### 2013-09-22 article-models-defn.tex
adding the source for the articles

\section{Introduction}
In this paper we define a notion of internal model which is adapted to
the univalent setting in which we work. Thus, we only describe what it means
to be an internal model when the ambient type theory is univalent. Our
presentation is derived from the definition of~\cite{Dybjer1996}. There,
an internal model is defined to be a category with families with interpretations
for the basic type constructors $\Pi$, $\Sigma$ and $\idtypevar{}$. However,
a category in~\cite{Dybjer1996} has a set of objects and for every two objects
a setoid of morphisms. This is too restrictive for our purpose. Ideally, we
would start with an $\infty$-category of contexts and an $\infty$-presheaf
of families over it. This presents us with the problem that a notion of
$\infty$-category has not been formulated \emph{in} type theory. Therefore,
we make a deviation from the approach in~\cite{Dybjer1996}. We shall
start with a type $\tfctx$ of contexts, a family $\mftypfunc:\tfctx\to\type$
of families over it and a family $\terms{\blank}:\prd*{\Gamma:\tfctx}\mftyp{\mfM}{\Gamma}
\to\type$ sending a type $A$ in context $\Gamma$ to its type $\terms{A}$ of
terms; then we shall directly interpret $\Pi$-types, declaring
$\terms{A\to B}$ to be $\ctxhom[\Gamma]{A}{B}$. The idea is that by letting
$\Pi$ play a fundamental role in the definition of a category with families
$\mathcal{C}$,
the morphisms get their properties directly from the internal type theory
of $\mathcal{C}$.

\begin{comment}
We briefly list the data of which an internal model of type theory consits. We
consider internal models in the style of~\cite{Dybjer1996}. Thus we will
describe a model $\mfM$ as a category with families. However, since our setting is
univalent type theory we will deviate from~\cite{Dybjer1996} in the following
respects: first of all, we make use of typical ambiguity and hence omit
reference to a thing called \emph{sort}.

We organize the definition of an internal model $\mfM$ as follows: in
\autoref{internal-model-contexts} we describe the category of contexts itself;
in \autoref{internal-model-families} we describe the families over a given context
and the related operations; in \autoref{internal-model-constructors} we
describe the basic constructors in the internal model.
\end{comment}

\section{Ideas in the definition}
An internal model of type theory is like a category with families, but we want
to avoid having to state higher coherences. In fact, we don't even start our
definition with a category of contexts; instead we just take a \emph{type} of contexts. 
The morphisms will come from the terms, evaluation of a function at a given
term will come from substitution. We recognize three basic ingredients to models:
first there is a type of contexts; second, for every context there is a model of types in
that context and third, for every type in a given context there is a type of its
terms. Then there are three basic attributes: context extension, weakening and
substitution. Context extension provides us with families over types as well as
with an interpretation of dependent pair types. We need weakening 
so that families can depend on the same type multiple times (the way the
identity type of a type depends two times on that type) and to be able
to talk about non-dependent function types,
the morphims of our category. Substitution will give us a way
to work with fibers of families as well as composition of functions and evaluation
of functions at terms.

Because we require a \emph{model} of types in a context, all the structure
which we require at the bottom level will be required to exist higher up as well.
Thus, the model of types in a given context $\Gamma$ will have a type of contexts
itself, which can be seen as the type of types in $\Gamma$; it will have its
own notion of types in a context, its own notion of terms, context extension,
weakening and substitution together with all the structure require for it. For
instance, when $A$ is a type in context $\Gamma$ in a model $\mfM$, then there
is the model of types in context $A$, which is the model of families over $A$. 
This model is required to be \emph{definitionally equal to} the model of types
in the context $\ctxext{\Gamma}A$, the context extension of $\Gamma$ and $A$.
In this way we protect ourselves from the need to dig an infinitely deep structure
of models when we want to consider examples.

To give the definition of a model we shall also need to consider certain morphisms
of internal models. Those should preserve all the structure: contexts are mapped
to contexts; for every context a morphism of models mapping the model of types
in that context to the model of types in the image of that context; there should
be a mapping of terms and context extension, weakening and substitution should be
preserved. We need to consider those morphisms because we require context extension,
weakening and substitution to be of that kind, thereby respecting each other
in all possible ways.

When we have this framework set up, we can interpret the basic type constructors
such as $\Pi$, $\Sigma$ and $\idtypevar{}$.
The higher categorical structure then comes from the
result that we have an interpretation of type theory.

\begingroup
\color{red}
\begin{rmk}
Currently, the definition seems to be circular. To define a model we need that
context extension, weakening and substitution be morphisms of models. A morphism
of models needs to preserve context extension, weakening and substitution.
Moreover, its definition requires the notion of composition of morphisms.

I've done it this way because it guides me to what the rules should be, trusting
that I can work my way back to give a (possibly less transparent) definition
which contains no circularities without doubt.
\end{rmk}
\endgroup

\section{Univalent universes as internal models}

Before we give the definition we illustrate the concepts that go into it in
the case of a univalent universe $\UU$.
Regardless of the definition of an internal model, a univalent universe should
be an example of one.

The type of contexts is $\UU$ itself and for every $\Gamma:\UU$, a type in
context $\Gamma$ is simply a family $A:\Gamma\to\UU$. The type $\terms{A}$ of terms of a type $A$
in context $\Gamma$ is defined to be $\prd{i:\Gamma}A(i)$. Note that $\Gamma\to
\UU$ itself also interprets type theory where a type in context $A:\Gamma\to\UU$
is a family $P:(\sm{i:\Gamma}A(i))\to\UU$. We may denote this model by
$\mftyp{\UU}{\Gamma}$. The terms of a type $P$ in context $A$ in the model
$\Gamma\to\UU$ are the terms of $\prd{i:\Gamma}{x:A(i)}P(i,x)$.

When $A$ is a type in context
$\Gamma$ we define the context extension $\ctxext{\Gamma}A$ to be
$\sm{i:\Gamma}A(i)$. Note that context extension $\ctxext{\Gamma}\blank$
can be seen acting not only as a function from the context of the model
$\Gamma\to\UU$ to the context of the model $\UU$, but it also acts on the types 
and terms: when $A$ is a type in context $\Gamma$ we may take the identity map
from $\mftyp{\Gamma\to\UU}{A}\to\mftyp{\UU}{\ctxext{\Gamma}A}$ because
$\mftyp{\Gamma\to\UU}{A}$ is taken to be $\mftyp{\UU}{\ctxext{\Gamma}A}$;
thus, $\ctxext{\Gamma}P\defeq P$ for every family $P$ over $A$ in context 
$\Gamma$. Likewise, when $P$ is a type in context $A$ in the model 
$\Gamma\to\UU$ then context extension should act on the terms of $P$ via a
function $\terms{P}\to\terms{\ctxext{\Gamma}P}$, which we take to be
the identity map once more.

Since $\mftyp{\UU}{\Gamma}$ is a model of type theory it
has it's own notion of context extension: when $P$ is a type in context $A$ in
the model $\mftyp{\UU}{\Gamma}$ then $\ctxext{A}P$ is the family
$\lam{i}\sm{x:A(i)}P(i,x):\Gamma\to\UU$. Also the context extension of
$\Gamma\to\UU$ acts trivially on the types and terms. Context extension is
analoguous to the Grothendieck construction that associates the category of
elements to a presheaf and it gives us $\Sigma$-types.

When $A$ and $B$ are types in context
$\Gamma$, the weakening $\ctxwk{A}{B}$ of $B$ along $A$ is defined to be
$\lam{\pairr{i,x}}B(i):(\sm{i:\Gamma}A(i))\to\UU$. Weakening along a type $A$ 
in context $\Gamma$ also acts on types and terms. When $Q$ is a type in context
$B$ in the model $\Gamma\to\UU$, we define $\ctxwk{A}{Q}$ to be 
$\lam{\pairr{i,x}}{y}Q(i,y)$. When $g:\terms{Q}$ we define $\ctxwk{A}{g}$ to
be $\lam{\pairr{i,x}}{y}g(i,y)$.

Note that for two types $A$ and $B$ in context $\Gamma$, the terms of
$\ctxwk{A}{B}$ are the terms of $\prd{i:\Gamma}A(i)\to B(i)$, i.e.~they are
the fiberwise maps from $A$ to $B$. We shall take the type $\terms{\ctxwk{A}{B}}$
of terms of $\ctxwk{A}{B}$ to be the type of morphisms from $A$ to $B$. Also,
any context of $\UU$ may be seen as a type in the empty context $\unit$. Thus
the type of terms of a context $\Gamma$ is $\unit\to\Gamma$, which is
equivalent to $\Gamma$. A context morphism from $\Delta$ to $\Gamma$ is a term
of $\ctxwk{\Gamma}{\Delta}$, i.e.~a function from $\Gamma$ to $\Delta$. We denote
the type $\terms{\ctxwk{\Gamma}{\Delta}}$ by $\ctxhom{\Delta}{\Gamma}$.

When $P$ is a family over
$A$ in context $\Gamma$ and $x$ is a term of $A$ we define the type $\subst{x}{P}$
in context $\Gamma$ to be $\lam{i}P(i,x(i))$. Like extension

We will interpret the dependent product $\mprd{A}{P}$ of a family $P$ over
$A$ in context $\Gamma$ by
\begin{equation*}
\mprd{A}{P}(i)\defeq \prd{x:A(i)}P(i,x)
\end{equation*}
With this interpretation there is an equivalence 
$\lambda:\eqv{\terms{P}}{\terms{\mprd{A}{P}}}$. This is lambda-abstraction; its
inverse being evaluation. We should note however, that the rule for evaluation
we interpret here is
\begin{equation*}
\inference{\Gamma\vdash f:\mprd{A}{P}}{\ctxext{\Gamma}A\vdash\tfev(f):P}
\end{equation*}
which is different than the usual rule
\begin{equation*}
\inference{\Gamma\vdash f:\mprd{A}{P} \qquad \Gamma\vdash a:A}{\Gamma \vdash \tfev(f,a):\subst{a}{P}}
\end{equation*}
The reason for this is that the interpretation of the usual rule would give a
function of type $\terms{\mprd{A}{P}}\to\prd{x:\terms{A}}\terms{\subst{x}{P}}$,
but this does not describe the terms of $\mprd{A}{P}$ by any means. Moreover,
since we have implemented substitution, we obtain from $\ctxext{\Gamma}A\vdash\tfev(f):P$ 
and $\Gamma\vdash x:A$ a term $\Gamma\vdash\subst{x}{\tfev(f)}:\subst{x}{P}$ and
therefore we do not loose anything with this approach.

{\color{red} point to example}

\section{A variable-free presentation of type theory}
We have seen in the example of $\UU$ that we will not need to refer to variables
when we are manipulating contexts. In fact, it is more natural not to.
This suggests that there is a presentation
of type theory where contexts do not mention variables. The first small gain is that such a presentation
would not be burdened with comments about variables being bounded or not, or fresh or
not occuring at all.

\subsection{The basic judgments}
The judgments we can make are the standard six basic judgments involving contexts, types and terms and
judgmental equality in each of these three cases.
\begin{align*}
\jctx*{\Gamma} & \jctxeq*{\Gamma}{\Gamma'}\\
\jtype*{\Gamma}{A} & \jtypeeq*{\Gamma}{A}{B}\\
\jterm*{\Gamma}{A}{x} & \jtermeq*{\Gamma}{A}{x}{y}.
\end{align*}

In contrast with standard practise, we don't assume that there is an empty context. The idea is that
if we interpret only the syntax of contexts, types and terms with extension, weakening and
substitution, we do not have a full model of type theory but we still have a 
structure resembling a category and categories need not have terminal objects. Therefore,
the rules that follow will describe only how to manipulate contexts, types and terms.
Of most of the inferences that we give there are two versions: one introducing an operation, the other asserting that the operation in question preserves judgmental equality.

\subsection{The basic rules for judgmental equality}
The rules for judgmental equality establish that it is an equivalence relation
in all three cases (contexts, types and terms).
%\begingroup
%\renewcommand*{\arraystretch}{3}
%\begin{equation*}
%\begin{array}{ccc}
\begin{infarray}{ccc}
\inference{\jctx{\Gamma}}{\jctxeq{\Gamma}{\Gamma}} & \inference{\jctxeq{\Gamma}{\Delta}}{\jctxeq{\Delta}{\Gamma}} & \inference{\jctxeq{\Gamma}{\Delta}\qquad\jctxeq{\Delta}{\Theta}}{\jctxeq{\Gamma}{\Theta}}\\
\inference{\jtype{\Gamma}{A}}{\jtypeeq{\Gamma}{A}{A}} &
\inference{\jtypeeq{\Gamma}{A}{B}}{\jtypeeq{\Gamma}{B}{A}} & 
\inference{\jtypeeq{\Gamma}{A}{B}\qquad\jtypeeq{\Gamma}{B}{C}}{\jtypeeq{\Gamma}{A}{C}}\\
\inference{\jterm{\Gamma}{A}{x}}{\jtermeq{\Gamma}{A}{x}{x}} & 
\inference{\jtermeq{\Gamma}{A}{x}{y}}{\jtermeq{\Gamma}{A}{y}{x}} &
\inference{\jtermeq{\Gamma}{A}{x}{y}\qquad\jtermeq{\Gamma}{A}{y}{z}}{\jtermeq{\Gamma}{A}{x}{z}}
\end{infarray}
%\end{array}
%\end{equation*}
\begin{infarray}{cc}
\inference{\jctxeq{\Gamma}{\Delta}\qquad\jtype{\Gamma}{A}}{\jtype{\Delta}{A}}
& \inference{\jctxeq{\Gamma}{\Delta}\qquad\jtypeeq{\Gamma}{A}{B}}{\jtypeeq{\Delta}{A}{B}}\\
\inference{\jctxeq{\Gamma}{\Delta}\qquad\jterm{\Gamma}{A}{x}}{\jterm{\Delta}{A}{x}}
& \inference{\jctxeq{\Gamma}{\Delta}\qquad\jtermeq{\Gamma}{A}{x}{y}}{\jtermeq{\Delta}{A}{x}{y}}\\
\inference{\jtypeeq{\Gamma}{A}{B}\qquad \jterm{\Gamma}{A}{x}}{\jterm{\Gamma}{B}{x}}
& \inference{\jtypeeq{\Gamma}{A}{B}\qquad\jtermeq{\Gamma}{A}{x}{y}}{\jtermeq{\Gamma}{B}{x}{y}}
\end{infarray}

\subsection{Extension}
We introduce extension which not only extends a context $\Gamma$ and a type
$A$ over it to a context $\ctxext{\Gamma}A$, but which also extends a type $A$
in context $\Gamma$ and a family $P$ over it to a type $\ctxext{A}P$ in context
$\Gamma$. We do this to ensure that all of type theory can be done in a context.
For instance, we could say (1) that a context in context $\Gamma$ is the same thing
as a type in context $\Gamma$; (2) When $A$ is a context in this sense, a type in
context $A$ is the same thing as a family $P$ over $A$ and (3) when $P$ is a type
in context $A$ in this sense, a term of $P$ keeps its original meaning.

Note that by introducing extension on the level of types and families,
we introduce $\Sigma$-types at a very early stage.
However, we need substitution to make this precise.
\begin{infarray}{cc}
\inference{\jtype{\Gamma}{A}}{\jctx{\ctxext{\Gamma}A}}
& \inference{\jctxeq{\Gamma}{\Delta}\qquad\jtypeeq{\Gamma}{A}{B}}{\jctxeq{\ctxext{\Gamma}A}{\ctxext{\Delta}B}}\\
\inference{\jtype{\ctxext{\Gamma}A}{P}}{\jtype{\Gamma}{\ctxext{A}P}}
& \inference{\jtypeeq{\Gamma}{A}{B}\qquad\jtypeeq{\ctxext{\Gamma}A}{P}{Q}}{\jtypeeq{\Gamma}{\ctxext{A}P}{\ctxext{B}Q}}
\end{infarray}

\subsection{Weakening}
We first define weakening by a context $\Gamma$. Since weakening by $\Gamma$
should in principle be a functor, it acts on contexts, types and terms alike.
Note that it is because of the weakening $\ctxwk{\Gamma}{\Delta}$ of a context
$\Delta$ by $\Gamma$ that we can speak of context morphisms from $\Gamma$ to $\Delta$: they are the terms
of $\ctxwk{\Gamma}{\Delta}$.
\begin{infarray}{cc}
\inference{\jctx{\Gamma}\qquad\jctx{\Delta}}{\jtype{\Gamma}{\ctxwk{\Gamma}{\Delta}}} 
& \inference{\jctxeq{\Gamma}{\Gamma'}\qquad\jctxeq{\Delta}{\Delta'}}{\jtypeeq{\Gamma}{\ctxwk{\Gamma}{\Delta}}{\ctxwk{\Gamma'}{\Delta'}}}\\
\inference{\jctx{\Gamma}\qquad\jtype{\Delta}{B}}{\jtype{\ctxext{\Gamma}\ctxwk{\Gamma}{\Delta}}{\ctxwk{\Gamma}{B}}} & \inference{\jctxeq{\Gamma}{\Gamma'}\qquad\jtypeeq{\Delta}{B}{B'}}{\jtypeeq{\ctxext{\Gamma}\ctxwk{\Gamma}{\Delta}}{\ctxwk{\Gamma}{B}}{\ctxwk{\Gamma'}{B'}}}\\
\inference{\jctx{\Gamma}\qquad\jterm{\Delta}{B}{y}}{\jterm{\ctxext{\Gamma}\ctxwk{\Gamma}{\Delta}}{\ctxwk{\Gamma}{B}}{\ctxwk{\Gamma}{y}}} 
& \inference{\jctxeq{\Gamma}{\Gamma'}\qquad\jtermeq{\Delta}{B}{y}{y'}}{\jtermeq{\ctxext{\Gamma}\ctxwk{\Gamma}{\Delta}}{\ctxwk{\Gamma}{B}}{\ctxwk{\Gamma}{y}}{\ctxwk{\Gamma'}{y'}}} 
\end{infarray}

A weakening operation is also defined for types. When $A$ and $B$ are both types
 in context $\Gamma$, the weakened type $\ctxwk{A}{B}$ in context $\ctxext{\Gamma}A$
 is the family which `doesn't really depend on $A$'. The terms of $\ctxwk{A}{B}$
 are the functions from $A$ to $B$. Likewise, the terms $\ctxwk{A}y$ are the
 constant maps at $y$ from $A$ to $B$, for $y:B$.
\begin{infarray}{cc}
\inference{\jtype{\Gamma}{A}\qquad\jtype{\Gamma}{B}}{\jtype{\ctxext{\Gamma}A}{\ctxwk{A}{B}}}
& \inference{\jtypeeq{\Gamma}{A}{A'}\qquad\jtypeeq{\Gamma}{B}{B'}}{\jtypeeq{\ctxext{\Gamma}A}{\ctxwk{A}{B}}{\ctxwk{A'}{B'}}}\\
\inference{\jtype{\Gamma}{A}\qquad\jtype{\ctxext{\Gamma}B}{Q}}
{\jtype{\ctxext({\Gamma}{A})\ctxwk{A}{B}}{\ctxwk{A}{Q}}}
& \inference{\jtypeeq{\Gamma}{A}{A'}\qquad\jtypeeq{\ctxext{\Gamma}B}{Q}{Q'}}
{\jtypeeq{\ctxext({\Gamma}{A})\ctxwk{A}{B}}{\ctxwk{A}{Q}}{\ctxwk{A'}{Q'}}}\\
\inference{\jtype{\Gamma}{A}\qquad\jterm{\Gamma}{B}{y}}{\jterm{\ctxext{\Gamma}A}{\ctxwk{A}{B}}{\ctxwk{A}{y}}}
& \inference{\jtypeeq{\Gamma}{A}{A'}\qquad\jtermeq{\Gamma}{B}{y}{y'}}{\jtermeq{\ctxext{\Gamma}A}{\ctxwk{A}{B}}{\ctxwk{A}{y}}{\ctxwk{A'}{y'}}}
\end{infarray}

\subsection{Substitution}
\begin{infarray}{cc}
\inference{\jterm{\Gamma}{A}{x}\qquad\jtype{\ctxext{\Gamma}A}{P}}{\jtype{\Gamma}{\subst{x}{P}}}
& \inference{\jtermeq{\Gamma}{A}{x}{x'}\qquad\jtypeeq{\ctxext{\Gamma}A}{P}{P'}}{\jtypeeq{\Gamma}{\subst{x}{P}}{\subst{x'}{P'}}}\\
\inference{\jterm{\Gamma}{A}{x}\qquad\jtype{\ctxext({\Gamma}{A})P}{Q}}{\jtype{\ctxext{\Gamma}\subst{x}{P}}{\subst{x}{Q}}}
& \inference{\jtermeq{\Gamma}{A}{x}{x'}\qquad\jtypeeq{\ctxext({\Gamma}{A})P}{Q}{Q'}}{\jtypeeq{\ctxext{\Gamma}\subst{x}{P}}{\subst{x}{Q}}{\subst{x'}{Q'}}}\\
\inference{\jterm{\Gamma}{A}{x}\qquad\jterm{\ctxext{\Gamma}A}{P}{f}}{\jterm{\Gamma}{\subst{x}{P}}{\subst{x}{f}}}
& \inference{\jtermeq{\Gamma}{A}{x}{x'}\qquad\jtermeq{\ctxext{\Gamma}A}{P}{f}{f'}}{\jtermeq{\Gamma}{\subst{x}{P}}{\subst{x}{f}}{\subst{x'}{f'}}}\\
\inference{\jterm{\Gamma}{A}{x}\qquad\jterm{\ctxext({\Gamma}{A}){P}}{Q}{g}}{\jterm{\ctxext{\Gamma}\subst{x}{P}}{\subst{x}{Q}}{\subst{x}{g}}}
& \inference{\jtermeq{\Gamma}{A}{x}{x'}\qquad\jtermeq{\ctxext({\Gamma}{A})P}{Q}{g}{g'}}{\jtermeq{\ctxext{\Gamma}\subst{x}{P}}{\subst{x}{Q}}{\subst{x}{g}}{\subst{x'}{g'}}}
\end{infarray}

\subsection{Compatibility of extension, weakening and substitution}
We describe the compatibility of extension, weakening and substitution. These
compatibility rules are formulated in a weak sense.

Extension is compatible with extension:
\begin{infarray}{c}
\inference
{ \jtype{\ctxext{\Gamma}A}{P}
}{\jterm
  {\ctxext{\Gamma}({A}P)}
  {\ctxwk{\ctxext{\Gamma}({A}P)}{(\ctxext({\Gamma}{A})P)}}{\typefont{extext}_0(A,P)}
}\\
\inference
{ \jtype{\ctxext({\Gamma}{A})P}{Q}
}{\jterm
  {\ctxext{\Gamma}({A}({P}Q))}
  {\ctxwk{\ctxext{A}({P}Q)}{(\ctxext({A}{P})Q)}}
  {\typefont{extext}_1(A,P,Q)}
}
\end{infarray}

Weakening is compatible with extension:
\begin{infarray}{c}
\inference{
  \jctx{\Gamma}
  \qquad
  \jtype{\Delta}{B}}
  {\jterm
    {\ctxext{\Gamma}\ctxwk{\Gamma}{(\ctxext{\Delta}B)}}
    {\ctxwk{\ctxwk{\Gamma}{(\ctxext{\Delta}B)}}{\ctxext{(\ctxwk{\Gamma}{\Delta})}(\ctxwk{\Gamma}{B})}}
    {\typefont{wkext}_0(\Gamma,B)}}
\end{infarray}

\subsection{More stuff in the type theory}
\begin{infarray}{c}
\inference{\jtype{\Gamma}{A}}{\jterm{\ctxext{\Gamma}A}{\ctxwk{A}{A}}{\idfunc[A]}}
\end{infarray}

\subsection{Leibniz equality on terms}
Occasionally we shall need a weaker version of equality than judgmental equality.
Our choice here is Leibniz equality, since it implies propositional equality when
identity types are present.

Let $P$ be a family over $A$ in context $\Gamma$. Then we can weaken
$\ctxwk{A}{P}$, which is a family over $\ctxwk{A}{A}$ in context $\ctxext{\Gamma}A$.
We also have the family $\ctxwk{\ctxwk{A}{A}}{P}$ over $\ctxwk{A}{A}$ in context
$\ctxext{\Gamma}A$ and thus we have the inference
\begin{equation*}
\inference{\jtype{\ctxext{\Gamma}A}{P}}{\jtype{\ctxext(({\Gamma}{A}){\ctxwk{A}{A}})\ctxwk{A}{P}}{\ctxwk{\ctxwk{A}{P}}{(\ctxwk{\ctxwk{A}{A}}{P})}}}
\end{equation*}
This family plays the role of the family $\jtype{\Gamma,x,y:A}{P(x)\to P(y)}$
in the type theory of \cite{TheBook}.

\subsection{Trivial cofibrations and weak equivalences of types}
We describe a relation between types that expresses when they are weakly equivalent.
Weak equivalence is introduced because we need a weaker notion of judgmental 
equality which also makes sense when identity types are not present, since that
would allow us to state that context extension, weakening and substitution
commute with each other.

A term $f:\ctxwk{A}{B}$ is a trivial cofibration if it has the
property that for any fibration $Q$ over $B$,
to find a section of $Q$ it suffices to find a section of the fibration
$f^\ast Q$ over $A$. In our type theoretical setting, the rôle of fibrations
is played by families, the rôle of the function type $A\to B$ is played by
$\ctxwk{A}{B}$ and our version of the pullback $f^\ast Q$ is $\subst{f}{(\ctxwk{A}{Q})}$.

\begin{defn}
Let $f$ be a term of $\ctxwk{A}{B}$ in context $\Gamma$.
\begin{enumerate}
\item For a family $Q$ over $B$ in context $\Gamma$ we define $f^\ast Q\jdeq\subst{f}{(\ctxwk{A}{Q})}$.
\item For a term $g$ of $Q$ in context $\ctxext{\Gamma}B$ we define $f^\ast g\jdeq\subst{f}{(\ctxwk{A}{g})}$.
\end{enumerate} 
\end{defn}
\begin{rmk}
The type $f^\ast(\ctxwk{B}{C})$ in context $\ctxext{\Gamma}A$ is can be viewed as the
type of functions from $A$ to $C$ which factor through $f$. Thus there should be
a function from $f^\ast(\ctxwk{B}{C})$ to $\ctxwk{A}{C}$.
\end{rmk}

\begin{rmk}
Every term $\jterm{\ctxext{\Gamma}A}{\ctxwk{A}{B}}{f}$ allows us to infer the following:
\begin{equation*}
\inference{\jtype{\ctxext{\Gamma}B}{Q}}{\jtype{\ctxext{\Gamma}A}{f^\ast Q}}
\qquad
\inference{\jterm{\ctxext{\Gamma}B}{Q}{g}}{\jterm{\ctxext{\Gamma}A}{f^\ast Q}{f^\ast g}}
\end{equation*}
\end{rmk}

\begin{defn}
A term $f:\ctxwk{A}{B}$ in context $\ctxext{\Gamma}A$ is said to be a trivial
cofibration if we can infer
\begin{equation*}
\inference{\jterm{\ctxext{\Gamma}A}{f^\ast Q}{t}}{\jterm{\ctxext{\Gamma}B}{Q}{\tilde{t}}}\qquad
\inference{\jterm{\ctxext{\Gamma}A}{f^\ast Q}{t}}{\jtermeq{\ctxext{\Gamma}A}{f^\ast Q}{f^\ast \tilde{t}}{t}}
\end{equation*}
{\color{red}This statement should be reformulated so that it only involves a single judgment...
but I don't see directly how to do that.}
\end{defn}

We have the following theorem in the type theory of \cite{TheBook}, which supports
our claim that we may indeed speak of a trivial cofibration.

\begin{thm}
Suppose $f:A\to B$ is a function. Then $f$ is an equivalence if and only if
for every $Q:B\to\type$ and every $g:\prd{x:A}Q(f(x))$ there is a section
$h:\prd{y:B}Q(b)$ with the property that $h\circ f\htpy g$. 
\end{thm}

\begin{proof}
We can first take $Q$ to be the constant family $\lam{y}A$. Furthermore, we may
take $g\defeq\idfunc[A]$. Then we get a term of type
\begin{equation*}
\sm{h:B\to A}h\circ f\htpy \idfunc[A],
\end{equation*}
i.e.~we get a left inverse $h$ for $f$. To show that $h$ is also a right inverse
of $f$, let $Q$ be the family $\lam{y}\id{f(h(y))}{y}$. To find a section of
$Q$, which is the homotopy we aim for, it suffices to find a section of
$Q\circ f$. In other words, we have to show that $\id{f(h(f(x)))}{f(x)}$ for
every $x:A$. This follows from the fact that $h$ is a left inverse for $f$.

The reverse direction is immediate.
\end{proof}

Another approach would be to define $f:\ctxwk{A}{B}$ to be left invertible
if there is a term $\ctxext({\Gamma}{A})P$

\subsection{The interactions of extension, weakening and substitution}

\subsection{The basic constructors of type theory}
\subsubsection{Dependent function types}
\begin{infarray}{cc}
\inference{\jtype{\ctxext{\Gamma}A}{P}}{\jtype{\Gamma}{\mprd{A}{P}}}
& \inference{\jtypeeq{\ctxext{\Gamma}A}{P}{P'}}{\jtypeeq{\Gamma}{\mprd{A}{P}}{\mprd{A}{P'}}}\\
\inference{\jterm{\ctxext{\Gamma}A}{P}{f}}{\jterm{\Gamma}{\mprd{A}{P}}{\lambda(f)}}
& \inference{\jtermeq{\ctxext{\Gamma}A}{P}{f}{f'}}{\jtermeq{\Gamma}{\mprd{A}{P}}{\lambda(f)}{\lambda(f')}}\\
\inference{\jterm{\Gamma}{\mprd{A}{P}}{g}}{\jterm{\ctxext{\Gamma}A}{P}{\tfev(g)}}
& \inference{\jtermeq{\Gamma}{\mprd{A}{P}}{g}{g'}}{\jtermeq{\ctxext{\Gamma}A}{P}{\tfev(g)}{\tfev(g')}}\\
\inference{\jterm{\ctxext{\Gamma}A}{P}{f}}{\jtermeq{\Gamma}{A}{\tfev(\lambda(f))}{f}}
\end{infarray}

With these rules we will not get the weak $\eta$-rule when identity types are present.
So it might be better to state that $\lambda$ is a trivial cofibration

\subsection{Identity types}
We need to define the diagonal function $\jterm{\ctxext{\Gamma}A}{\ctxwk{A}{(\ctxext{A}\ctxwk{A}{A})}}{\delta}$
\begin{infarray}{c}
\inference{\jtype{\Gamma}{A}}{\jtype{\ctxext({\Gamma}{A})\ctxwk{A}{A}}{\idtypevar{A}}}\\
\inference{\jtype{\Gamma}{A}}{\jterm{\ctxext{\Gamma}A}{\delta^\ast\idtypevar{A}}{\reflsym}}\\
\inference{\jtype{\ctxext(({\Gamma}{A}){\ctxwk{A}{A}})\idtypevar{A}}{P}\qquad
\jterm{\ctxext{\Gamma}A}{\pairr{\delta,\reflsym}^\ast P}{d}}
{\jterm{\ctxext(({\Gamma}{A}){\ctxwk{A}{A}})\idtypevar{A}}{P}{J}}
\end{infarray}

\section{The category with families $\mfM$}\label{internal-model-contexts}
\begin{defn}\label{defn:premodel}
An internal model $\mfM$ of type theory consists of the following data. 
\begin{enumerate}
\item A type $\tfctx(\mfM)$ of \emph{contexts}.
\item A function $\mftypfunc{\mfM}$ assigning to each context $\Gamma$ of $\mfM$ an internal model
$\mftyp{\mfM}{\Gamma}$ of type theory.
\begin{defn}
The type $\tfctx(\mftyp{\mfM}{\Gamma})$ is denoted by $\mftyp{\mfM}{\Gamma}$. A
term of $\mftyp{\mfM}{\Gamma}$ is called a \emph{type in context $\Gamma$}. To indicate
that $A$ is a type in context $\Gamma$ we also write $\Gamma\vdash A:\mfM$. 
When $P:\mftyp{\mftyp{\mfM}{\Gamma}}{A}$ for some type $A$ in context $\Gamma$, we
also speak of \emph{a family $P$ over $A$ in context $\Gamma$.}
\end{defn}
\item A family $\terms{\blank}:\prd*{\Gamma:\ctx(\mfM)}\mftyp{\mfM}{\Gamma}\to\type$
assigning to every type $A$ in context $\Gamma$ the type $\terms{A}$ of its
terms.
\begin{defn}
When $A$ is a type in context $\Gamma$, we define $\Gamma\vdash x:A$ 
to mean $x:\terms{A}$.
\end{defn}
\item Context extension: a morphism
\begin{equation*}
\tfext^\Gamma:\mftyp{\mfM}{\Gamma}\to\mfM
\end{equation*}
of internal models for every context $\Gamma$.
\begin{defn}
When $A$ is a type in context $\Gamma$, we also denote
the context $\tfext^\Gamma_0(A)$ of $\mfM$ by $\ctxext{\Gamma}A$.
We will write
$\ctxext({\Gamma}{A})P$ for $\ctxext({\Gamma}{A})P$.
\end{defn}
\item The judgmental equalities:
\begin{align*}
\mftyp{\mftyp{\mfM}{\Gamma}}{A} & \jdeq\mftyp{\mfM}{\ctxext{\Gamma}A}\\
\mftyp{\tfext^\Gamma}{A} & \jdeq \modelfont{id}_{\mftyp{\mfM}{\ctxext{\Gamma}A}}\\
\terms[{\tfext^\Gamma}]{P} & \jdeq \idfunc[\terms{P}].
\end{align*}
\begin{rmk}
In particular, we will have the judgmental equalities:
\begin{enumerate}
\item When $A$ is a type in context $\Gamma$ we have
\begin{equation*}
\mftyp{\mftyp{\mfM}{\Gamma}}{A}\jdeq\mftyp{\mfM}{\ctxext{\Gamma}A},
\end{equation*}
ensuring that a context in the model $\mftyp{\mftyp{\mfM}{\Gamma}}{A}$ is the same thing as a
context in the model $\mftyp{\mfM}{\ctxext{\Gamma}A}$.
\item If $Q$ is a family over $P$ where $P$ is a family over $A$ in context $\Gamma$, then
\begin{equation*}
\tfext^{\protect{\mftyp{\mftyp{\mfM}{\Gamma}}{A}}}(P,Q)
\jdeq
\tfext^{\protect{\mftyp{\mfM}{\ctxext{\Gamma}A}}}(P,Q)
\end{equation*}
ensuring that twe two possible notion of context extension are the same.
\end{enumerate}
Other judgmental equalities will be required with the ingredients that follow.
We will not list them all.
\end{rmk}
\item Weakening: a morphism
\begin{equation*}
\tfwk^A:\mftyp{\mfM}{\Gamma}\to\mftyp{\mfM}{\ctxext{\Gamma}A}
\end{equation*}
of internal models for every type $A$ in context $\Gamma$. When $B$ is a type
in context $\Gamma$, we denote $\tfwk^A(B)$ by $\ctxwk{A}{B}$. 
\item Substitution: a morphism
\begin{equation*}
\tfsubst^x:\mftyp{\mfM}{\ctxext{\Gamma}A}\to\mftyp{\mfM}{\Gamma}
\end{equation*}
of internal models for any $x:A$ in context $\Gamma$. When $P$ is a family over $A$ in context
$\Gamma$, we denote $\tfsubst^x(P)$ also by $\subst{x}{P}$. 
\item A judgmental equality
\begin{equation*}
\tfsubst^x\circ\tfwk^A\jdeq\modelfont{id}_{\mftyp{\mfM}{\Gamma}}
\end{equation*}
for any $x:A$ and $B$ in context $\Gamma$.
\item A context $\unit^\mfM:\tfctx(\mfM)$ and a term of type $\isequiv(\ctxext{\unit^\mfM}\blank)$. We will denote
this equivalence by $e_\unit$. The context $\unit^\mfM$ is also
called the \emph{empty context}.
\begin{defn}
For any context $\Gamma$, type $\terms{\Gamma}$ is defined to mean
$\terms{e_\unit^{-1}(\Gamma)}$. 
\end{defn}
\item A section $\unit^{\blank}:\prd{\Gamma:\ctx(\mfM)}\mftyp{\mfM}{\Gamma}$ assigning
a type $\unit^\Gamma$ in context $\Gamma$ to every context $\Gamma$ and
an identification $\alpha_\unit(\Gamma):\id{\ctxext{\Gamma}\unit^\Gamma}{\Gamma}$
for every context $\Gamma$. We also require that there is an identification
$\id{\trans{\alpha_\unit(\Gamma)}{\ctxwk{\unit^\Gamma}{A}}}{A}$ for every
type $A$ in context $\Gamma$.
\item For any type $A$ in context $\Gamma$, a term $\idfunc[A]:\terms{\ctxwk{A}{A}}$
\end{enumerate}
\begin{flushright}
\textsl{End of \autoref{defn:premodel}.}
\end{flushright}
\end{defn}

\begin{defn}
In an internal model $\mfM$ we define
\begin{equation*}
\ctxhom{\Delta}{\Gamma}\defeq \terms{\ctxwk{e_\unit^{-1}(\Delta)}{e_\unit^{-1}(\Gamma)}}
\end{equation*}
\end{defn}

\section{Morphisms of internal models}
\begin{defn}\label{defn:premodel-morphism}
A morphism $f:\mfM\to \mfN$ of internal models consists of
\begin{enumerate}
\item a function $\ctx(f):\ctx(\mfM)\to\ctx(\mfN)$. The function $\ctx(f)$ is also
denoted by $f_0$.
\item a morphism $\mftyp{f}{\Gamma}:\mftyp{\mfM}{\Gamma}\to\mftyp{\mfN}{f_0(\Gamma)}$ of internal models for every
$\Gamma:\ctx(M)$.
\item the judgmental equality
\begin{equation*}
\mftyp{\mftyp{f}{\Gamma}}{A}\jdeq\mftyp{f}{\ctxext{\Gamma}A}
\end{equation*}
\item a function $\terms[f]{A}:\terms[\mfM]{A}\to\terms[\mfN]{\mftyp{f}{\Gamma}_0(A)}$ for
every type $A$ in context $\Gamma$. 
\item preservation of context extension: 
\begin{align*}
\alpha^f_0 & :\id{f\circ\tfext^\Gamma}{\tfext^{f_0(\Gamma)}\circ\mftyp{f}{\Gamma}}\\
\alpha^f_1 & :\id{\mftyp{f}{\Gamma}_0\circ\tfext^A}{\tfext^{\mftyp{f}{\Gamma}_0(A)}\circ\mftyp{f}{\ctxext{\Gamma}A}}.
\end{align*}
\begin{comment}
the judgmental equality
\begin{equation*}
f_0(\ctxext{\Gamma}A)\jdeq\ctxext{f_0(\Gamma)}\mftyp{f}{\Gamma}_0(A)
\end{equation*}
for every type $A$ in context $\Gamma$.
\end{comment}
\item preservation of weakening: 
\begin{align*}
\beta^f_0 & :\id{\mftyp{f}{\Gamma}\circ\tfwk^\Gamma}{\tfwk^{f_0(\Gamma)}\circ f}\\
\beta^f_1 & :\id{\mftyp{f}{\ctxext{\Gamma}A}\circ\tfwk^A}{\tfwk^\protect{\mftyp{f}{\Gamma}_0(A)}\circ\mftyp{f}{\Gamma}}.
\end{align*}
\begin{comment}
This gives the following three judgmental equalities:
\begin{enumerate}
\item the judgmental equality
\begin{equation*}
\mftyp{f}{\ctxext{\Gamma}A}_0(\ctxwk{A}{B})\jdeq\ctxwk{\mftyp{f}{\Gamma}_0(A)}{\mftyp{f}{\Gamma}_0(B)}
\end{equation*}
for every two types $A$ and $B$ in context $\Gamma$.
\item the judgmental equality
\begin{equation*}
\mftyp{f}{\ctxext({\Gamma}{A})\ctxwk{A}{B}}\circ\tfwk^A\jdeq\tfwk^{\protect{\mftyp{f}{\Gamma}_0(A)}}\circ\mftyp{f}{\ctxext{\Gamma}B}
\end{equation*}
\item the judgmental equality
\begin{equation*}
\terms[f]{\ctxwk{A}{B}}(\ctxwk{A}{y})\jdeq\ctxwk{\mftyp{f}\Gamma_0(A)}{(\terms[f]{B}(y))}
\end{equation*}
\end{enumerate}
\end{comment}
\item preservation of substitution: 
\begin{equation*}
\gamma^f:\id{\mftyp{f}\Gamma_0\circ\tfsubst^x}{\tfsubst^\protect{\terms[f]{A}(x)}\circ\mftyp{f}{\ctxext{\Gamma}A}}.
\end{equation*}
\begin{comment}
\begin{enumerate}
\item the judgmental equality
\begin{equation*}
\mftyp{f}\Gamma_0(\subst{x}{P})\jdeq\subst{\terms[f]{A}(x)}{\mftyp{f}{\ctxext{\Gamma}A}_0(P)}
\end{equation*}
for every family $P$ over $A$ in context $\Gamma$ and every term $x:A$.
\item the judgmental equality
\begin{equation*}
\mftyp{f}\Gamma_0(\subst{x}{Q})\jdeq\subst{\terms[f]{A}(x)}{\mftyp{f}{\ctxext({\Gamma}{A})P}_0(Q)}
\end{equation*}
for every family $Q$ over $P$ over $A$ in context $\Gamma$ and every term $x:A$.
\end{enumerate}
\end{comment}
\end{enumerate}
\begin{flushright}
\textsl{End of \autoref{defn:premodel-morphism}.}
\end{flushright}
\end{defn}

\begin{defn}
Suppose that $f:\mfM\to\mfN$ and $g:\mfN\to\mfN'$ are morphisms of internal models.
We define the composition $g\circ f:\mfM\to\mfN'$ to be the morphism given by
\begin{enumerate}
\item $(g\circ f)_0\defeq g_0\circ f_0$
\item $\mftypfunc{g\circ f}(\Gamma)\defeq\mftypfunc{g}(f_0(\Gamma))\circ\mftypfunc{f}(\Gamma)$
\item $\terms[g\circ f]{A}\defeq\terms[g]{\mftyp{f}{\Gamma}_0(A)}\circ\terms[f]{A}$.
\item a definition of $\alpha^{g\circ f}_0$ and $\alpha^{g\circ f}_1$.
\item a definition of $\beta^{g\circ f}_0$ and $\beta^{f\circ f}_1$.
\item a definition of $\gamma^{g\circ f}$. 
\end{enumerate}
\end{defn}

\begin{rmk}
The requirement that a morphism of models acts on terms is reminiscent of the requirement
that a functor acts on morphisms. In this fasion, the requirement that a morphism of
models preserves substitution is the counterpart of a functor preserving composition.
\end{rmk}
Context extension, weakening and substitution are required to be such morphims
of models. Thus they must be extended for this purpose.

\begin{description}
\item[Context extension] We will extend context extension to a morphism 
$\tfext(\Gamma):\mftyp{\mfM}{\Gamma}\to M$. Thus, we already have
$\tfext(\Gamma)_0\defeq\ctxext{\Gamma}\blank$. We also require 
\begin{align*}
\mftyp{\tfext(\Gamma)}{A} & : \mftyp{\mftyp{\mfM}{\Gamma}}{A}\to \mftyp{\mfM}{\ctxext{\Gamma}A}\\
\terms[\tfext(\Gamma)]{P} & : \terms[\mftyp{\mfM}{\Gamma}]{P}\to\terms[\mfM]{\mftyp{\tfext(\Gamma)}{A}_0(P)}
\end{align*}
For both of these we take the identity function.
We require furthermore that context extension preserves context extension,
weakening and substitution:
\begin{enumerate}
\item Extension preserves extension: for every family $P$ over $A$ in context
$\Gamma$ an identification $\id{\ctxext({\Gamma}{A})P}{\ctxext{\Gamma}({A}P)}$.
\item Extension preserves weakening:
\item Extension preserves substitution:
Because context extension is the identity on the levels of types and terms,
the rules are easier.
\end{enumerate}
\item[Weakening] is a morphism $\tfwk:\mftyp{\mfM}{\Gamma}\to\mftyp{\mfM}{\ctxext{\Gamma}A}$, so we must have
\begin{align*}
\tfwk^A_0 & : \mftyp{\mfM}{\Gamma}\to\mftyp{\mfM}{\ctxext{\Gamma}A}
\intertext{which is denoted by $\ctxwk{A}{\blank}$,}
\mftypfunc{\tfwk^A} & : \mftyp{\mfM}{\ctxext{\Gamma}B}\to\mftyp{\mfM}{\ctxext({\Gamma}{A})\ctxwk{A}{B}}\\
\terms[\tfwk^A]{B} & : \terms[\mftyp{\mfM}{\Gamma}]{B}\to\terms[\mftyp{\mfM}{\ctxext{\Gamma}A}]{\ctxwk{A}{B}}
\end{align*}
for a term $y:B$, the term $\terms[{\tfwk^A}]{B}(y)$ is the constant map from
$A$ to $B$, assigning $y$ to every term of $A$.
\item[Substitution] is a morphism $\tfsubst^x:\mftyp{\mfM}{\ctxext{\Gamma}A}
\to\mftyp{\mfM}{\Gamma}$ for every $x:A$ in context $\Gamma$, so we must have
\begin{align*}
\tfsubst^x_0 & : \mftyp{\mfM}{\ctxext{\Gamma}A}\to\mftyp{\mfM}{\Gamma}
\intertext{which is denoted by $\subst{x}{\blank}$,}
\mftyp{\tfsubst^x}{P} & : \mftyp{\mfM}{\ctxext({\Gamma}{A})P}\to\mftyp{\mfM}{\ctxext{\Gamma}\subst{x}{P}}\\
\terms[\tfsubst^x]{P} & : \terms{P}\to\terms{\subst{x}{P}}
\end{align*}
The function $\terms[\tfsubst^x]{P}$ is usually denoted by $\tfev(\blank,x)$. 
\end{description}

\section{The basic type constructors in internal models}

\begin{defn}
\begin{enumerate}
\item A type $\mprd{A}{P}$ for every type $A$ in context $\Gamma$ and every type
$P$ in context $\ctxext{\Gamma}A$.
\begin{defn}
Suppose $A$ and $B$ are types in context $\Gamma$. We define
\begin{align*}
A\to B & \defeq\mprd{A}{\ctxwk{A}{B}}.
\intertext{For any two contexts $\Delta$ and $\Gamma$, we define}
\Delta\to\Gamma & \defeq e_\unit^{-1}(\Delta)\to e_\unit^{-1}(\Gamma)
\end{align*}
and furthermore we define $\ctxhom{\Delta}{\Gamma}\defeq\terms{\Delta\to\Gamma}$.
\end{defn}
\item An equivalence $\lambda:\eqv{\terms{P}}{\terms{\mprd{A}{P}}}$ for every
family $P$ over $A$ in context $\Gamma$. When $\ctxext{\Gamma}A\vdash 
u:P$, we call $\lambda(u)$ the \emph{$\lambda$-abstraction of $u$.}
\begin{rmk}
Note that we get an equivalence $\eqv{\terms{\unit^\Gamma\to A}}{\terms{A}}$ 
for every type $A$ in context $\Gamma$.
\end{rmk}
\begin{rmk}
Thus we see that $\eqv{\terms{A\to B}}{\terms{\ctxwk{A}{B}}}$
by $\lambda$-abstraction. 
\end{rmk}
%\item A function $\tfev:\terms{\mprd{A}{P}}\to\prd{x:\terms{A}}\terms{\subst{x}{P}}$.
\item A term $\pi^A:\ctxext{\Gamma}A\to \Gamma$ for every type $A$ in context
$\Gamma$.
\item A term $\iota^x:\ctxext{\Gamma}\subst{x}{P}\to\ctxext({\Gamma}{A})P$
for every family $P$ of types over $A$ in context $\Gamma$ and every
term $x:A$.
\item A type $A[f]$ in context $\Delta$ for every $f:\ctxhom{\Delta}{\Gamma}$ 
and every type $A$ in context $\Gamma$. {\color{blue}Could this be defined
in terms of the substitution $\subst{x}{P}$ we already have? It seems so. If this is
indeed the case we need either some identifications or we could just omit
this part of the definition.}
\item an identification $\id{B[\pi^A]}{\ctxwk{A}{B}}$ for any two types $A$
and $B$ in context $\Gamma$.
\item A type $\msm{A}{P}$ in context $\Gamma$ for every family $P$ over $A$
in context $\Gamma$, with an equivalence $\pairr{\blank,\blank}:\eqv{\sm{x:\terms{A}}\terms{\subst{x}{P}}}
{\terms{\msm{A}{P}}}$.
\item A family $\idtypevar{A}$ over $\ctxwk{A}{A}$ in context $\ctxext{\Gamma}A$ for
every type $A$ in context $\Gamma$ and a term of type.
\begin{rmk}
If $A$ and $B$ are types in context $\Gamma$, we may denote the type 
$\msm{A}{\ctxwk{A}{B}}$ by $A\times B$. By the end of the current definition
there will be an identification $\id{\ctxext({\Gamma}{A})P}
{\ctxext{\Gamma}\msm{A}{P}}$.

There is a term $\delta:\ctxhom{\ctxext{\Gamma}A}{\ctxext({\Gamma}{A})\ctxwk{A}{A}}$
defined by...
\end{rmk}
\item For any family $Q$ over $\idtypevar{A}$ in context $\ctxext({\Gamma}{A})\ctxwk{A}{A}$ an equivalence
$J:\eqv{\terms{Q}}{}$
\end{enumerate}
\end{defn}

\begingroup
\color{blue}
\subsubsection*{More desiderata}
Some of which hopefully follow from more elegant or general rules, but
this list is to keep them in mind:
\begin{enumerate}
\item $\id{e_\unit^{-1}(\Gamma.A)}{\msm{e_\unit^{-1}(\Gamma)}{\trans{(\eta_\Gamma)}{A}}}$, where
$\eta_\Gamma:\id{\Gamma}{e_\unit(e_\unit^{-1}(\Gamma))}$ is the unit of the
equivalence $e_\unit$.  
\item $\id{(\msm{A}{P}\to B)}{(P\to \ctxwk{A}{B})}$, expressing that $\Sigma$ is
left adjoint to weakening.
\item $\id{(\ctxwk{A}{B}\to P)}{(B\to\mprd{A}{P})}$ expressing that $\Pi$ is
right adjoint to weakening.
\item a term $\Gamma\vdash\idfunc[A]:A\to A$ for every type $A$ in context $\Gamma$.
\item a context $\UU$ for the universe.
\item an identification $\id{\pi^A\circ \pi^P\circ\iota^x}{\pi^{\subst{x}{P}}}$ for every
family $P$ of types over $A$ in context $\Gamma$ and every term $x:A$. In other
words, the diagram
\begin{equation*}
\begin{tikzcd}
\ctxext({\Gamma}{A})P \ar{r}{\pi^P} & \ctxext{\Gamma}A \ar{d}{\pi^A}\\
\ctxext{\Gamma}\subst{x}{P} \ar{u}{\iota^x} \ar{r}[swap]{\pi^{\subst{x}{P}}} & \Gamma
\end{tikzcd}
\end{equation*}
should commute.
\begin{comment}
\item For $\Gamma:\ctx$, $\Gamma\vdash A:\type$, $\ctxext{\Gamma}A\vdash P
:\type$ and $\ctxext({\Gamma}{A})P\vdash Q:\type$, $x:A$ and $u:P$
we desire that $\typefont{nat{-}ev{-}subst}(Q,u,x):\id{Q[\iota^x][\tfev(u,x)]}{Q[u][x]}$.
\item A function
\begin{equation*}
\tfcomp:\terms{\mprd{P}{Q}}\to\prd{x:\terms{A}}\terms{\mprd{\subst{x}{P}}{Q[\iota^x]}}
\end{equation*}
which could possibly replace $\tfev$ because it could be that $\tfev$ is
definable in terms of $\tfcomp$. It should satisfy
\begin{equation*}
\id{\trans{\typefont{nat{-}ev{-}subst}(Q,u,x)}{\tfev(\tfcomp(f,x),\tfev(u,x))}}{\tfev(\lambda(\tfev(f,u)),x)},
\end{equation*}
which is a naturality condition. There should be an equivalence between
$\terms{\mprd{A}{P}}$ and a type consisting of those terms of $\prd{x:\terms{A}}\terms{\subst{x}{P}}$ which satisfy
a certain naturality condition. One could hope that by formulating this dependently
enough, it is possible to write down such a naturality in one line.
\end{comment}
\end{enumerate}
\endgroup

\begin{comment}
\begingroup
\color{darkgreen}
\subsubsection*{The problem with evaluation.}
...or why I'm doing such strange things. It seems that evaluation is of a
different nature. At first sight, one might expect that evaluation compares
to $\lambda$-abstraction the same way that first and second projection compare
to the pairing function $\pairr{\blank,\blank}$. However, this is not the case
because evaluation cannot be seen as an inverse to $\lambda$-abstraction. The
$\lambda$-abstraction is of type $\terms{P}\to\terms{\mprd{A}{P}}$ whereas
evaluation is of type $\terms{\mprd{A}{P}}\to\prd{x:\terms{A}}\terms{\subst{x}{P}}$. There
are a few things to note about $\tfev$:
\begin{enumerate}
\item There's no function back so in particular $\tfev$ is not an equivalence.
My guess is that there should be a naturality condition playing a role.
\item Evaluation tells how functions interact with substitution of terms.
\end{enumerate}
So it seems that evaluation is like a variable binder in the same way
that quantifiers are binding variables. If I remember right, Voevodsky mentioned
something like this at the Formal Topology conference in Ljubljana, though I
forgot what exactly he said.
\endgroup
\end{comment}

\begin{comment}
\begingroup
\color{gray}
\subsubsection*{Old stuff below:}
\begin{enumerate}
\item a context morphism $\catid{\Gamma}:\ctxhom{\Gamma}{\Gamma}$ for every
context $\Gamma$.
\item a context morphism $f\circ g:\ctxhom{E}{\Gamma}$ for every two context
morphisms $f:\ctxhom{\Delta}{\Gamma}$ and $g:\ctxhom{E}{\Delta}$. 
\item paths $\id{f\circ\catid{\Delta}}{f}$ and $\id{\catid{\Gamma}\circ f}{f}$
for every $f:\ctxhom{\Delta}{\Gamma}$.
\item a path $\id{f\circ (g\circ h)}{(f\circ g)\circ h}$ for every
$f:\ctxhom{\Delta}{\Gamma}$, $g:\ctxhom{E}{\Delta}$ and $h:\ctxhom{Z}{E}$. 
\item a context morphism $\Gamma_!:\ctxhom{\Gamma}{\unit^\mfM}$ 
for every context $\Gamma$.
\item a path $\id{(\unit^\mfM)_!}{\catid{\unit^\mfM}}$.
\item a path $\id{\Gamma_!\circ f}{\Delta_!}$ for every $f:\ctxhom{\Delta}{\Gamma}$. 
\item a path $\id{A[\catid{\Gamma}]}{A}$ for every object $A$ in context $\Gamma$.
\item a path $\id{A[f][g]}{A[f\circ g]}$ for every object $A$ in context $\Gamma$
and every $f:\ctxhom{\Delta}{\Gamma}$ and $g:\ctxhom{E}{\Delta}$. 
\end{enumerate}
\endgroup
\end{comment}

\begin{desiderata}
We should have a theorem stating that our internal model indeed interprets
the rules of type theory.
\end{desiderata}

\begin{comment}
\section{Continuity of functions}
We wish to find a condition under which we have
\begin{equation*}
\eqv{\terms{\mprd{A}{P}}}{\sm{f:\prd{x:\terms{A}}\terms{\subst{x}{P}}}\text{``$f$ is continuous''}}.
\end{equation*}
for all families $P$ over $A$. 
Needless to say, we first need to formulate what we mean by ``$f$ is continuous''.
Thus, we need to find the family $\pairr{p,u,v}\mapsto \id{\trans{p}{u}}{v}$,
which also requires us to define transportation.

Let $P$ be a family over $A$ in context $\Gamma$. Then we can weaken
$\ctxwk{A}{P}$, which is a family over $\ctxwk{A}{A}$ in context $\ctxext{\Gamma}A$.
We also have the family $\ctxwk{\ctxwk{A}{A}}{P}$ over $\ctxwk{A}{A}$ in context
$\ctx{\Gamma}{A}$ and thus we can form the family
\begin{equation*}
\ctxwk{\ctxwk{A}{A}}{P}\to\ctxwk{A}{P}.
\end{equation*}
This family plays the role of the family $x,y:A\vdash P(x)\to P(y):\type$.
{\color{red}Explain this by performing the relevant substitution.}
Transportation is a term of the family
\begin{equation*}
\idtypevar{A}\to\ctxwk{\ctxwk{A}{A}}{P}\to\ctxwk{A}{P}
\end{equation*}
obtained by identity elimination, as usual. Thus we obtain a term
\begin{equation*}
\mathcal{T}rans_{x,y}:\terms{\id{x}{y}\to\subst{x}{P}\to\subst{y}{P}}.
\end{equation*}
for every $x,y:A$.

Now let $x,y:A$. Then we have the type $\subst{y}{P}$
in context $\Gamma$ and the family $\ctxwk{\subst{y}{P}}{\subst{x}{P}}$
over $\subst{y}{P}$ in context $\Gamma$. We could weaken $\subst{y}{P}$ subsequently by $A$, $\ctxwk{A}{A}$
and $\idtypevar{A}$ to obtain the family $\ctxwk{\idtypevar{A}}{\ctxwk{\ctxwk{A}{A}}{\ctxwk{A}{\subst{y}{P}}}}$ over $\idtypevar{A}$ in context $\ctxext({\Gamma}{A})\ctxwk{A}{A}$. We will
denote this family simply by $\ctxwk{\idtypevar{A}}{\subst{y}{P}}$. Likewise,
we obtain the family $\ctxwk{\idtypevar{A}}{\ctxwk{\subst{y}{P}}{\subst{x}{P}}}$ over
$\ctxwk{\idtypevar{A}}{\subst{y}{P}}$ in context $\ctxext(({\Gamma}{A}){\ctxwk{A}{A}})\idtypevar{A}$ and we obtain the family $\ctxwk{\idtypevar{A}}{\idtypevar{\subst{y}{P}}}$
over $\ctxwk{\idtypevar{A}}{\ctxwk{\subst{y}{P}}{\subst{y}{P}}}$. Now we
have the family $\subst{\ctxwk{A}{\mathcal{T}rans_{x,y}}}{\ctxwk{\idtypevar{A}}{\idtypevar{\subst{y}{P}}}}$ over
$\ctxwk{\idtypevar{A}}{\ctxwk{\subst{y}{P}}{\subst{x}{P}}}$, which is
the desired family to state the continuity condition:

\begin{defn}
A function $f:\prd{x:\terms{A}}\terms{\subst{x}{P}}$ is said to be \emph{continuous}
if there is a term of type
\begin{equation*}
\prd{x,y:\terms{A}}\terms{\subst{\ctxwk{A}{\mathcal{T}rans_{x,y}}}{\ctxwk{\idtypevar{A}}{\idtypevar{\subst{y}{P}}}}}.
\end{equation*}
\end{defn}

\begin{lem}
If all functions are ``continuous mappings of terms'' then types without terms
are empty, i.e.\ we have
\begin{equation*}
(\terms{A}\to\emptyt)\to\terms{A\to\emptyt^\mfGraph}
\end{equation*}
\end{lem}

\begin{proof}
Any function from a type without terms is vacuously continuous.
\end{proof}

We will see that in the graph model functions are not just continuous mappings
of terms, since there is a graph $\tilde{\emptyt}$ which has no terms but which
differs from the empty graph nevertheless.

\begin{desiderata}
The following may relate to functions being continuous mappings of terms (it
seems to be a very strong condition):
\begin{enumerate}
\item We have the adjunctions $\tfcolim\dashv\Delta\dashv\tflim$ in which the
two monads are idempotent (we don't even have the adjunctions for $\mfGraph$).
\item $\eqv{\terms{\Omega}}{\ctx}$
\end{enumerate}
We also need to check whether all functions in the reflexive graphs have this
property (if not, ditch this section).
\end{desiderata}
\end{comment}

\section{The graph model of type theory}
In this section we define the graph model of type theory, denoted by
$\mfGraph$.  In our presentation, we follow that of the definition internal
models. After we have established the graph model, we show how the graph
model can be seen as a presheaf model, namely over the category $\cdot
{\rightrightarrows}\cdot$.

\begin{defn}
A \emph{(directed) graph} $\Gamma$ is a pair $\pairr{\Gamma_0,\Gamma_1}$ 
consisting of a type $\Gamma_0$ of vertices and a family 
$\Gamma_1:\Gamma_0\to\Gamma_0\to\type$ of edges. The type $\ctx(\mfGraph)$
is defined to be the type of all graphs; we will usually denote it by
$\tfGraph$. Explicitly, we have
\begin{equation*}
\tfGraph\defeq\sm{\Gamma_0:\type}\Gamma_0\to\Gamma_0\to\type.
\end{equation*}
\end{defn}

\begin{eg}\label{ex:pb}
The underlying graph of the diagram
\begin{equation*}
\begin{tikzcd}
{} & A \ar{d}{f} \\
B \ar{r}[swap]{g} & C
\end{tikzcd}
\begin{comment}
\begin{tikzpicture}
\matrix (m) [std] { & A \\ B & C \\};
\draw[ar] (m-1-2) -- node[right] {$f$} (m-2-2);
\draw[ar] (m-2-1) -- node[below] {$g$} (m-2-2);
\end{tikzpicture}
\end{comment}
\end{equation*}
is $I\defeq\mathbf{3}$ and has $J(1,3)\defeq J(2,3)\defeq \unit $ 
and $J(x,y)\defeq\emptyt$ otherwise (it is defined using the induction 
principle of $\mathbf{3}$ and a universe).
\end{eg}

\begin{defn}
A \emph{graph $A$ in the context $\Gamma$} is a pair $\pairr{A_0,A_1}$ consisting
of 
\begin{align*}
A_0 & :\Gamma_0\to\type\\
A_1 & :\prd*{i,j:\Gamma_0}\Gamma_1(i,j)\to A_0(i)\to A_0(j)\to\type.
\end{align*}
Thus, for a graph $\Gamma$, the type $\mftyp{\mfGraph}{\Gamma}$ is the type
\begin{equation*}
\sm{A_0:\Gamma_0\to\type}\prd*{i,j:\Gamma_0}\Gamma_1(i,j)\to A_0(i)\to A_0(j)\to\type.
\end{equation*}
We will also write $\Gamma\vdash A:\mfGraph$ when $A$ is a graph in context
$\Gamma$.
\end{defn}

\begin{defn}
Suppose that $\Gamma\vdash A:\mfGraph$. Then we define the graph $\ctxext{\Gamma}A$
by
\begin{align*}
\ctxext{\Gamma}A_0 & \defeq \sm{i:\Gamma_0}A_0(i)\\
\ctxext{\Gamma}A_1(\pairr{i,x},\pairr{j,y}) & \defeq \sm{q:\Gamma_1(i,j)}A_1(q,x,y).
\end{align*}
\end{defn}

\begin{defn}
Suppose that $\Gamma\vdash A:\mfGraph$ and $\Gamma\vdash B:\mfGraph$. Then we
define the graph $\ctxwk{A}{B}$ in context $\ctxext{\Gamma}A$ by
\begin{align*}
(\ctxwk{A}{B})_0(\pairr{i,x}) & \defeq B_0(i)\\
(\ctxwk{A}{B})_1(\pairr{q,e},u,v) & \defeq B_1(q,u,v).
\end{align*}
\end{defn}

\begin{defn}
Suppose that $\Gamma\vdash A:\mfGraph$. A term $x$ of $A$ consists of a pair
$\pairr{x_0,x_1}$ where
\begin{align*}
x_0 & : \prd{i:\Gamma_0}A_0(i)\\
x_1 & : \prd*{i,j:\Gamma_0}{q:\Gamma_1}A_1(q,x_0(i),x_0(j)).
\end{align*}
Thus we define
\begin{equation*}
\terms{A}\defeq\sm{x_0:\prd{i:\Gamma_0}A_0(i)}\prd*{i,j:\Gamma_0}{q:\Gamma_1}A_1(q,x_0(i),x_0(j)).
\end{equation*}
We also write $\Gamma\vdash x:A$ when $x$ is a term of the graph $A$ in context
$\Gamma$.
\end{defn}

\begin{defn}
Suppose that $P$ is a family of graphs over $A$ in context $\Gamma$ and let
$x$ be a term of $A$. Then we define the graph $\subst{x}{P}$ 
in context $\Gamma$ by
\begin{align*}
\subst{x}{P}_0(i) & \defeq P_0(\pairr{i,x_0(i)})\\
\subst{x}{P}_1(q,u,v) & \defeq P_1(\pairr{q,x_1(q)},u,v)
\end{align*}
\end{defn}

\begin{rmk}
Note that we have the judgmental equality $\subst{x}{\ctxwk{A}{B}}\jdeq B$
for every two graphs $A$ and $B$ in context $\Gamma$ and every term $x$ of $A$.
\end{rmk}

\begin{defn}
The terminal graph $\unit^\mfGraph$, which we shall often denote simply 
by $\unit$, is defined by
\begin{align*}
{\unit^\mfGraph}_0 & \defeq \unit\\
{\unit^\mfGraph}_1(x,y) & \defeq \unit.
\end{align*}
\end{defn}

\begin{rmk}
We note that the function $\ctxext{\unit}\blank$ is an equivalence
from $\mftyp{\mfGraph}{\unit}$ to $\tfGraph$. It's inverse is the function which maps
a graph $\Gamma$ to the pair $\pairr{A_0,A_1}$ where $A_0$ is defined by
$A_0(\ttt)\defeq\Gamma_0$ and where $A_1$ is defined by $A_1(\ttt)\defeq
\Gamma_1$.

Note that we only have an equivalence here, not a judgmental equality.
\end{rmk}

\subsubsection{The contexts of $\mfGraph$}

The category $\psh(\cdot{\rightrightarrows}\cdot)$ of presheaves over
$\cdot{\rightrightarrows}\cdot$ is given by
\begin{equation*}
\sm{A_0,A_1:\type}(A_1\to A_0)\times(A_1\to A_0).
\end{equation*}
To see that $\psh(\cdot{\rightrightarrows}\cdot)$ is indeed equivalent to
the type $\tfGraph$ of all graphs, note that we have the equivalences
\begin{align*}
\psh(\cdot{\rightrightarrows}\cdot) & \eqvsym \sm{A_0,A_1:\type}A_1\to A_0\times A_0\\
& \eqvsym \sm{A_0:\type} A_0\times A_0\to\type\\
& \eqvsym \tfGraph.
\end{align*}

\begin{defn}
Suppose that $\Delta\defeq \pairr{\Delta_0,\Delta_1}$ and $\Gamma\defeq \pairr{\Gamma_0,
\Gamma_1}$ are graphs. A morphism $f$ from $\Delta$ to $\Gamma$ is a pair
$\pairr{f_0,f_1}$ consisting of
\begin{align*}
f_0 & : \Delta_0\to\Gamma_0\\
f_1 & : \prd*{u,v:\Delta_0}\Delta_1(u,v)\to\Gamma_1(f_0(u),f_0(v))
\end{align*}
We write $\ctxhom{\Delta}{\Gamma}$ for the type of morphisms of graphs.
\end{defn}

The terminal graph $\unit^\mfGraph$ plays the role of the empty context.

\begin{defn}
The terminal graph $\unit^\mfGraph$, which we shall often denote simply 
by $\unit$, is defined by
\begin{align*}
{\unit^\mfGraph}_0 & \defeq \unit\\
{\unit^\mfGraph}_1(x,y) & \defeq \unit.
\end{align*}
\end{defn}

It is immediate that the type of graph morphisms $\Gamma\to\unit^\mfGraph$ is
contractible for every graph $\Gamma$. The terms of a graph $\Gamma$ are the
global elements of $\Gamma$, i.e.~they are the morphisms of type 
$\ctxhom{\unit^\mfGraph}{\Gamma}$.

\begin{defn}
Let $\Gamma$ be a graph. The type $\terms[\mfGraph]{\Gamma}$ of terms of $\Gamma$ is
defined to be
\begin{equation*}
\sm{x_0:\Gamma_0}\Gamma_1(x_0,x_0).
\end{equation*}
\end{defn}

We now turn to the description of the universe $\gobjclass$ of graphs. Note
that for the category $I\defeq 0{\rightrightarrows}1$, we have
$I/0\defeq\catid{0}$ and $I/1\defeq s\rightarrow\catid{1}\leftarrow t$, where
$s$ and $t$ are the morphisms $0\to 1$. Thus we have 
\begin{align*}
\psh(I/0) & \eqvsym \type\\
\psh(I/1) & \eqvsym \sm{X,Y:\type}X\to Y\to\type.
\end{align*}
The functors $\psh(\Sigma_s)$ and $\psh(\Sigma_t)$ map a presheaf 
$\pairr{X,Y,R}$ to $X$ and $Y$, respectively. 
Thus we obtain the following graph $\gobjclass$.

\begin{defn}
The universe $\gobjclass$ of graphs is defined to be
\begin{align*}
{\gobjclass}_0 & \defeq \type\\
{\gobjclass}_1(X,Y) & \defeq X\to Y\to\type.
\end{align*}
\end{defn}

Note that the type $\terms{\gobjclass}$ of terms of $\gobjclass$ is 
exactly $\tfGraph$.

\subsubsection{Families of graphs}
Using $\gobjclass$, we obtain a description of what
it means to be a graph in a context $\Gamma$, where $\Gamma$ is itself a
graph. The notion of ``a graph in a context $\Gamma$'' is the interpretation
of ``a type in a context $\Gamma$''.

\begin{defn}
A \emph{graph $A$ in the context $\Gamma$} is a pair $\pairr{A_0,A_1}$ consisting
of 
\begin{align*}
A_0 & :\Gamma_0\to\type\\
A_1 & :\prd*{i,j:\Gamma_0}\Gamma_1(i,j)\to A_0(i)\to A_0(j)\to\type.
\end{align*}
We will also write $\Gamma\vdash A:\mfGraph$ when $A$ is a graph in context
$\Gamma$.
\end{defn}

\begin{rmk}
Note that every graph can be seen as a graph in context $\unit^\mfGraph$.
\end{rmk}

\begin{defn}
Let $A$ be a graph in context $\Gamma$ and let $f:\ctxhom{\Delta}{\Gamma}$
be a graph morphism. We define the graph $A[f]$ in context $\Delta$ by
\begin{align*}
A[f]_0(u) & \defeq A_0(f_0(u))\\
A[f]_1(p,x,y) & \defeq A_1(f_1(p),x,y).
\end{align*}
\end{defn}

\begin{defn}
Let $\Gamma\vdash A:\mfGraph$ and $\Gamma\vdash B:\mfGraph$ be two graphs in
the context $\Gamma$. A morphism $f$ from $A$ to $B$ is a pair $\pairr{f_0,f_1}$
consisting of
\begin{align*}
f_0 & : \prd*{i:\Gamma_0} A_0(i)\to B_0(i)\\
f_1 & : \prd*{i,j:\Gamma_0}{q:\Gamma_0}*{x:A_0(i)}*{y:A_0{j}}A_1(q,x,y)\to B_1(q,f_0(x),f_0(y)).
\end{align*}
We write $\ctxhom[\Gamma]{A}{B}$ for the type of morphisms between graphs in
context $\Gamma$.
\end{defn}

\begin{defn}
The unit graph $\unit^\Gamma$ in context $\Gamma$ consists of
\begin{align*}
(\unit^\Gamma)_0 & \defeq \lam{i}\unit\\
(\unit^\Gamma)_1(q) & \defeq \lam{x}{y}\unit.
\end{align*}
\end{defn}

As was the case with graphs, terms of graphs in a context are described by
morphisms from $\unit$. Explicating this in more detail, we get

\begin{defn}
Let $A$ be a graph in context $\Gamma$. A term $x$ of $A$ is a pair
$\pairr{x_0,x_1}$ consisting of
\begin{align*}
x_0 & : \prd{i:\Gamma_0}A_0(i)\\
x_1 & : \prd*{i,j:\Gamma_0}{q:\Gamma_1(i,j)}A_1(q,x_0(i),x_0(j)).
\end{align*}
We define $\terms[\mfGraph]{A}$ to be the type of terms of $A$.
\end{defn}

\begin{rmk}
Note that the type of terms of a graph $\Gamma$ is equivalent to the type
of terms of the graph $\Gamma$ in context $\unit^\mfGraph$.
\end{rmk}

\begin{defn}
Suppose that $\Gamma\vdash A:\mfGraph$. Then we define the graph $\ctxext{\Gamma}A$
by
\begin{align*}
(\ctxext{\Gamma}A)_0 & \defeq \sm{i:\Gamma_0}A_0(i)\\
(\ctxext{\Gamma}A)_1(\pairr{i,x},\pairr{j,y}) & \defeq \sm{q:\Gamma_1(i,j)}A_1(q,x,y).
\end{align*}
There is a morphism $\pi^A:\ctxhom{\ctxext{\Gamma}A,\Gamma}$ given by
\begin{align*}
\pi^A_0 & \defeq \lam{\pairr{i,x}}i && : (\sm{i:\Gamma_0}A_0(i))\to \Gamma_0\\
\pi^A_1(\pairr{i,x},\pairr{j,y}) & \defeq \lam{\pairr{q,e}}q && : (\sm{q:\Gamma_1(i,j)}A_1(q,x,y))\to\Gamma_1(i,j).
\end{align*}
\end{defn}

\begin{rmk}
The construction of $\ctxext{\Gamma}A$ is similar to the Grothendieck
construction of the category of elements for a presheaf.
\end{rmk}

\begin{defn}
Let $\Gamma\vdash A:\mfGraph$. A family $P$ of graphs over $A$ is by
definition a graph $P$ in the context $\ctxext{\Gamma}A$. More explicitly, a family
$P$ of graphs over $A$ in context $\Gamma$ consists of
\begin{align*}
P_0 & : \prd*{i:\Gamma_0}A_0(i)\to\type\\
P_1 & : \prd*{i,j:\Gamma_0}{q:\Gamma_1(i,j)}*{x:A_0(i)}*{y:A_0(j)} A_1(q,x,y)\to P_0(x)\to P_0(y)\to\type
%P_1 & : \prd{\pairr{i,x},\pairr{j,y}:(\ctxext{\Gamma}A)_0}(\ctxext{\Gamma}A)_1(\pairr{i,x},\pairr{j,y})\to P_0(i,x)\to P_0(j,y)\to\type.
\end{align*}
\end{defn}

\begin{defn}
When $P$ is a family of graphs over $A$ in context $\Gamma$ and $x:A$, we
get a graph $\subst{x}{P}$ in context $\Gamma$ given by
\begin{align*}
\subst{x}{P}_0(i) & \defeq P_0(x_0(i))\\
\subst{x}{P}_1(q,u,v) & \defeq P_1(\pairr{q,x_1(q)},u,v)
\end{align*}
\end{defn}

\subsubsection{The basic type constructors for graphs}
With families of graphs being available, we can give the interpretations of
dependent products, dependent sums and identity types. In the following, we
shall introduce the graph interpretations of dependent products, dependent
sums and identity types and describe the terms of the resulting graphs.\note{These definitions
should be connected to Mike's article but I don't really know how to do this}

\begin{defn}
Let $P$ be a family of graphs over $A$, where $\Gamma\vdash A:\mfGraph$. 
The dependent function graph $\mprd{A}{P}$ in context $\Gamma$ consists of
\begin{align*}
\mprd{A}{P}_0(i) & \defeq \prd{x:A_0(i)}P_0(x)\\
\mprd{A}{P}_1(q,f,g) & \defeq \prd*{x:A_0(i)}*{y:A_0(j)}{e:A_1(q,x,y)}P_1(\pairr{q,e},f(x),g(y)).
\end{align*}
\end{defn}

\begin{rmk}
A term $f:\mprd{A}{P}$ consists of
\begin{align*}
f_0 & : \prd*{i:\Gamma_0}{x:A_0(i)}P_0(x)\\
f_1 & : \prd*{i,j:\Gamma_0}{q:\Gamma_1(i,j)}*{x:A_0(i)}*{y:A_0(j)}{e:A_1(q,x,y)}P_1(\pairr{q,e},f_0(x),f_0(y))
\end{align*}
Therefore, we see that $\eqv{\terms{\mprd{A}{P}}}{\terms{P}}$. \emph{Warning:} it
is by no means the case that $\eqv{\terms{\mprd{A}{P}}}{\prd{x:\terms{A}}
\terms{\subst{x}{P}}}$ for all families $P$ over $A$. For instance, the graph
$\tilde{\emptyt}$ defined by $\tilde{\emptyt}_0\defeq\unit$ and 
$\tilde{\emptyt}_1(\ttt,\ttt)\defeq\emptyt$ has no terms and neither does
$\tilde{\emptyt}+\tilde{\emptyt}$. Nevertheless, there are two graph morphisms
from $\tilde{\emptyt}$ to $\tilde{\emptyt}+\tilde{\emptyt}$. More generally,
when $\Gamma$ is a graph such that $\Gamma_1(i,j)\jdeq\emptyt$ for all $i,j:\Gamma_0$,
then $\eqv{{\terms{\Gamma\to\Gamma'}}}{\Gamma_0\to\Gamma^\prime_0}$.
\end{rmk}

\begin{defn}
If $P$ is a family of graphs over $\Gamma$, the dependent pair graph
$\msm{A}{P}$ consists of
\begin{align*}
\msm{A}{P}_0(i) & \defeq \sm{x:A_0(i)}P_0(x)\\
\msm{A}{P}_1(q,\pairr{x,u},\pairr{y,v}) & \defeq \sm{e:A_1(q,x,y)}P_1(\pairr{q,e},u,v).
\end{align*}
\end{defn}

\begin{rmk}
A term $w:\msm{A}{P}$ consists of
\begin{align*}
w_0 & : \prd{i:\Gamma_0}\sm{x:A_0(i)}P_0(x)
\intertext{and, writing $\lam{i}\proj1(w_0(i))$ and $\lam{i}\proj2(w_0(i))$ as
$w_{00}$ and $w_{01}$ respectively,}
w_1 & : \prd*{i,j:\Gamma_0}{q:\Gamma_1(i,j)}\sm{e:A_1(q,w_{00}(i),w_{00}(j)}P_1(\pairr{q,e},w_{01}(i),w_{01}(j)).
\end{align*}
By $\choice{\infty}$ it follows that
\begin{equation*}
\eqv{\terms{\msm{A}{P}}}{\sm{x:\terms{A}}\terms{\subst{x}{P}}}.
\end{equation*}
\end{rmk}

\begin{rmk}
For any two graphs $A$ and $B$ in a context $\Gamma$, we have the graph
$B[\pi^A]$ in the context $\ctxext{\Gamma}A$, which we may just write
as $B$. Therefore, we can
consider the graphs $\mprd{A}{B}$ and $\msm{A}{B}$. As usually those shall be
denoted  by $A\to B$ and $A\times B$, respectively.

A term $f$ of $A\to B$ consists of
\begin{align*}
f_0 & : \prd*{i:\Gamma_0}A_0(i)\to B_0(i)\\
f_1 & : \prd*{i,j:\Gamma_0}{q:\Gamma_1(i,j)}*{x:A_0(i)}*{y:A_0(j)}A_1(q,x,y)\to B_1(q,f_0(x),f_0(y)),
\end{align*}
so we see that $\terms{A\to B}\jdeq\ctxhom[\Gamma]{A}{B}\jdeq\terms{B[\pi^A]}$.

A term $\pairr{x,y}$ of $A\times B$ consists of
\begin{align*}
\pairr{x,y}_0 & : \prd*{i:\Gamma_0} A_0(i)\times B_0(i)\\
\pairr{x,y}_1 & : \prd*{i,j:\Gamma_0}{q:\Gamma_1(i,j)} A_1(q,x_0(i),x_0(j))\times B_1(q,y_0(i),y_0(j))
\end{align*}
so we see that $\eqv{\terms{A\times B}}{\terms{A}\times\terms{B}}$.

Note that we also obtain
a \emph{graph} $\Delta\to\Gamma$ this way, for every two graphs $\Delta$ and
$\Gamma$ seen as graphs in the context $\unit^\mfGraph$. 
The type $\terms{\Delta\to\Gamma}$ of terms of $\Delta\to\Gamma$ is just
$\ctxhom{\Delta}{\Gamma}$.
\end{rmk}

\begin{defn}
For any graph $\Gamma$ there is a graph $\tffam{\Gamma}$ in context 
$\gobjclass$ defined by
\begin{align*}
(\tffam{\Gamma})_0 & \defeq \lam{X}X\to\Gamma_0\\
(\tffam{\Gamma})_1(X,Y,R) & \defeq \lam{f}{g}\prd*{x:X}*{y:Y}R(x,y)\to\Gamma_1(f(x),g(y)).
\end{align*}
\end{defn}

\begin{rmk}
A term $D$ of $\msm{\gobjclass}{\tffam{\Gamma}}$ consists of a term
$\pairr{\Delta_0,f_0}$ of type
\begin{equation*}
\msm{\gobjclass}{\tffam{\Gamma}}_0\jdeq \sm{\Delta_0:\type}\Delta_0\to\Gamma_0
\end{equation*}
and a term $\pairr{\Delta_1,f_1}$ of type
\begin{equation*}
\msm{\gobjclass}{\tffam{\Gamma}}_1(q)\jdeq \sm{\Delta_1}
\end{equation*}
\end{rmk}

\begin{defn}
There is a graph morphism
\begin{equation*}
\graphcharmapfunc{\Gamma} : \sm{\gobjclass}
\end{equation*}
\end{defn}

\begin{defn}
Let $A$ be a graph in context $\Gamma$. We define the family $\idtypevar{A}$ over $\ctxwk{A}{A}$ in
context $\ctxext{\Gamma}A$ by
\begin{align*}
(\idtypevar{A})_0(\pairr{i,x},y) & \defeq \id{x}{y}\\
(\idtypevar{A}){}_1(\pairr{q,e},d,\alpha,\alpha') & \defeq \id{\trans{\pairr{\alpha,\alpha'}}{e}}{d}
\end{align*}
where $q:\Gamma_1(i,j)$, $e:A_1(q,x,x')$, $d:A_1(q,y,y')$, $\alpha:\id{x}{y}$
and $\alpha':\id{x'}{y'}$. The transportation along the path 
$\pairr{\alpha,\alpha'}:\id{\pairr{x,x'}}{\pairr{y,y'}}$ in $A_0(i)\times A_0(j)$
is taken with respect to the family $\lam{x}{x'}A_1(q,x,x')$.

We define the term $\refl{A}$ of the family 
$\subst{\idfunc[A]}{\idtypevar{A}}$ over $A$ in context $\Gamma$ by
\begin{align*}
(\reflf{A})_0(i) & \defeq \lam{x}\refl{x}\\
(\reflf{A})_1(q) & \defeq \lam{e}\refl{e}
\end{align*}
\end{defn}

\begin{defn}
Let $D$ be a family over $\idtypevar{A}$ in context 
$\ctxext({\Gamma}{A})\ctxwk{A}{A}$. Then we have the family
$\subst{\idfunc[A]}{D}$ over $\subst{\idfunc[A]}{\idtypevar{A}}$ in
context $\ctxext{\Gamma}A$ given by
\begin{align*}
\subst{\idfunc[A]}{D}_0(\pairr{i,x},\alpha) & \defeq D_0(\pairr{i,x,x},\alpha)\\
\subst{\idfunc[A]}{D}_1(\pairr{q,e},\gamma) & \defeq D_1(\pairr{q,e,e},\gamma).
\end{align*}
The family $\subst{\reflf{A}}{\subst{\idfunc[A]}{D}}$ over $A$ in
context $\Gamma$ is given by
\begin{align*}
\subst{\reflf{A}}{\subst{\idfunc[A]}{D}}_0(i,x) & \defeq D_0(i,x,x,\refl{x})\\
\subst{\reflf{A}}{\subst{\idfunc[A]}{D}}_1(q,e) & \defeq D_1(q,e,e,\refl{e}).
\end{align*}
To show that the identity graphs correctly interpret the identity elimination
rule, we must give a function
\begin{equation*}
\tfJ : \terms{\subst{\reflf{A}}{\subst{\idfunc[A]}{D}}}\to\terms{D}.
\end{equation*}
Note that a term $d$ of $\subst{\reflf{A}}{\subst{\idfunc[A]}{D}}$
consists of
\begin{align*}
d_0 & : \prd*{i:\Gamma_0}{x:A_0(i)}D_0(i,x,x,\refl{x})\\
d_1 & : \prd*{i,j:\Gamma_0}{q:\Gamma_1(i,j)}*{x:A_0(i)}*{y:A_0(j)}{e:A_1(q,x,y)}D_1(q,e,e,\refl{e})
\end{align*}
A simple argument using path induction reveals that terms of
$\subst{\reflf{A}}{\subst{\idfunc[A]}{D}}$ indeed yield terms of $D$. 
\end{defn}

\begin{rmk}
Using the identity graph $\idtypevar{A}$ we can describe the identity
graph $\id[A]{x}{y}$ in context $\Gamma$ for any two terms $x,y:A$. 
The graph $\id[A]{x}{y}$ in context $\Gamma$ consists of
\begin{align*}
(\id[A]{x}{y})_0(i) & \defeq \id{x_0(i)}{y_0(i)}\\
(\id[A]{x}{y})_1(q,\alpha,\beta) & \defeq \id{\trans{\pairr{\alpha,\beta}}{x_1(q)}}{y_1(q)}
\end{align*}
From this, we see that a term of $p:\id[A]{x}{y}$ consists of
\begin{align*}
p_0 & : \prd{i:\Gamma_0}\id{x_0(i)}{y_0(i)}\\
p_1 & : \prd*{i,j:\Gamma_0}{q:\Gamma_1(i,j)}\id{\trans{\pairr{p_0(i),p_0(j)}}{x_1(q)}}{y_1(q)}.
\end{align*}
\end{rmk}

\subsection{Properties of graphs}

\subsubsection{Contractibility and equivalences of graphs}

\begin{defn}
A graph $A$ in context $\Gamma$ is said to be \emph{contractible} if there
is a term of the graph
\begin{equation*}
\msm{A}{\mprd{\ctxwk{A}{A}}{\idtypevar{A}}}
\end{equation*}
in context $\Gamma$.
\end{defn}

\begin{lem}\label{lem:contractible-graphs}
Let $A$ be a graph in context $\Gamma$. The following are equivalent:
\begin{enumerate}
\item $A$ is a contractible graph.
\item Both $A_0(i)$ and $A_1(q,x,y)$ are always contractible.
\end{enumerate}
\end{lem}

\begin{proof}
Let $H:\msm{A}{\mprd{\ctxwk{A}{A}}{\idtypevar{A}}}$. Unfolding the definitions, we have
an element $H_0(i)$ of type
\begin{align*}
\msm{A}{\mprd{\ctxwk{A}{A}}{\idtypevar{A}}}_0(i) & \jdeq \sm{x:A_0(i)}\mprd{\ctxwk{A}{A}}{\idtypevar{A}}_0(i,x)\\
& \jdeq \sm{x:A_0(i)}\prd{y:A_0(i)}(\idtypevar{A})_0(\pairr{i,x},y)\\
& \jdeq \sm{x:A_0(i)}\prd{y:A_0(i)}\id{x}{y}
\end{align*}
for all $i:\Gamma_0$ and, writing $H_{00}(i)$ for $\proj1 H_0(i)$ and
$H_{01}(i)$ for $\lam{y}(\proj2 H_0(i))(y)$, we have $H_1(q)$ of type
\begin{align*}
& \msm{A}{\mprd{\ctxwk{A}{A}}{\idtypevar{A}}}_1(q,H_0(i),H_0(j)) \\
& \jdeq \sm{e:A_1(q,H_{00}(i),H_{00}(j))}\mprd{\ctxwk{A}{A}}{\idtypevar{A}}_1(\pairr{q,e},H_{01}(i),H_{01}(j))\\
& \jdeq \sm{e:A_1(q,H_{00}(i),H_{00}(j))}\prd*{x:A_0(i)}*{y:A_0(j)}{d:A_1(q,x,y)}(\idtypevar{A})_1(\pairr{q,e},d,H_{01}(x),H_{01}(y))\\
& \jdeq \sm{e:A_1(q,H_{00}(i),H_{00}(j))}\prd*{x:A_0(i)}*{y:A_0(j)}{d:A_1(q,x,y)}\id{\trans{\pairr{H_{01}(x),H_{01}(y)}}{e}}{d}.
\end{align*}
By $H_0$, it follows that each $A_0(i)$ is contractible. By the contractibility
of each $A_0(i)$, it follows that the type of $H_1(q)$ is equivalent to
\begin{equation*}
\sm{e:A_1(q,H_{00}(i),H_{00}(j))}\prd{d:A_1(q,H_{00}(i),H_{00}(j))}\id{e}{d}
\end{equation*}
which asserts that $A_1(q,H_{00}(i),H_{00}(j))$ is contractible. By 
the contractibility of each $A_0(i)$, is is equivalent to the assertion
that each $A_1(q,x,y)$ is contractible.
\end{proof}

\begin{rmk}
We address the question whether it is the case that a graph $A$ in context
$\Gamma$ is contractible if and only if $\terms{A}$ is contractible. As a
consequence of \autoref{lem:contractible-graphs}, it is indeed the case
that $\terms{A}$ is contractible whenever $A$ is. However, the converse
does not hold.

To see this, we first construct a counter example to the converse of the
weak function extensionality principle, which states that there is a function
of type
\begin{equation}
\iscontr(\prd{x:X}P(x))\to\prd{x:X}\iscontr(P(x))\label{eq:wfe-converse}
\end{equation}
for any type family $P:X\to\type$. In the proof of \autoref{thm:wfe-converse}, 
we will find a family
$P:X\to\type$ with the property that $\prd{x:X}P(x)$ is contractible and for
which there is a term $x:A$ with $P(x)$ not contractible. Disproving the
converse of the weak function extensionality principle suffices for our
purposes, because if $P:X\to\type$ is a counter example to \autoref{eq:wfe-converse},
then we can take $\Gamma\defeq\pairr{\Gamma_0,\Gamma_1}$ to be given by
$\Gamma_0\defeq X$ and $\Gamma_1(i,j)\defeq\emptyt$ and we take
$A\defeq\pairr{A_0,A_1}$ to be given by $A_0\defeq P$ and $A_1(q,u,v)
\defeq\emptyt$.
\end{rmk}

We define the family $\mathcal{T}:\Sn^1\to\UU$ by
We define $\mathcal{T}(\base)\defeq\mathbf{3}$. To define $\mathcal{T}(\lloop):
\id{\mathbf{3}}{\mathbf{3}}$ we apply the univalence axiom. Hence it suffices to find an
equivalence $\eqv{\mathbf{3}}{\mathbf{3}}$, for which we take the function
$e$ defined by
\begin{equation*}
e(x)\defeq\begin{cases}
0_\mathbf{3} & \text{if }x\jdeq 0_\mathbf{3}\\
2_\mathbf{3} & \text{if }x\jdeq 1_\mathbf{3}\\
1_\mathbf{3} & \text{if }x\jdeq 2_\mathbf{3}.
\end{cases}
\end{equation*}

\begin{lem}
The type $\terms{\mathcal{T}}\defeq\prd{x:\Sn^1}\mathcal{T}(x)$ is contractible.
\end{lem}

\begin{proof}
The type of sections of $\mathcal{T}$ is equivalent to $\sm{u:\mathcal{T}(\base)}\id{e(u)}{u}$.
If we have a term $\pairr{u,\alpha}$ of the latter type, it follows by induction
on $\mathbf{3}$ that $\id{\pairr{u,\alpha}}
{\pairr{0_\mathbf{3},\refl{0_\mathbf{3}}}}$ for all $\pairr{u,\alpha}:
\sm{x:\mathcal{T}(\base)}\id{e(u)}{u}$,
which shows that the type of sections of $\mathcal{T}$ is contractible.
\end{proof}

\begin{thm}\label{thm:wfe-converse}
There is a type family $P:A\to\type$ for which
\begin{equation*}
\neg\Big(\iscontr(\prd{x:A}P(x))\to\prd{x:A}\iscontr(P(x))\Big).
\end{equation*}
\end{thm}

\begin{proof}
The type family of our counter example is $\mathcal{T}$: the fiber $\mathcal{T}(\base)$ isn't contractible.
\end{proof}

\begin{defn}
A graph morphism $f:\ctxhom{\Delta}{\Gamma}$ is an equivalence of graphs when
$\graphcharmap[\Gamma]{f}$ is a contractible graph in the context $\Gamma$.
\end{defn}

\begin{lem}
Let $A$ and $B$ be graphs in a context $\Gamma$ and let $f:A\to B$. The following are equivalent:
\begin{enumerate}
\item $f[i]:A[i]\to B[i]$ is an equivalence of graphs for every term $i:\Gamma$.
\item $\ctxext{\Gamma}f:\ctxext{\Gamma}A\to\ctxext{\Gamma}B$ is an equivalence of graphs.
\item Both $f_0(i)$ and $f_1(q,x,y)$ are always equivalences.
\item $\terms{f}:\terms{\Delta}\to\terms{\Gamma}$ is an equivalence.
\end{enumerate}
\end{lem}

\begin{rmk}
It follows that there is an equivalence
\begin{equation*}
\ctxext{\Gamma}\msm{A}{P}\simeq\ctxext({\Gamma}{A})P
\end{equation*}
for every family $P$ of graphs over a graph $A$ in context $\Gamma$.
\end{rmk}

\subsubsection{Homotopy levels}
\begin{itemize}
\item A graph $A$ in context $\Gamma$ is of homotopy level $n$ precisely when each
$A_0(i)$ and each $A_1(q,x,y)$ are of homotopy level $n$. 
\item We can name at least three different propositions in the empty context:
\begin{enumerate}
\item $\Gamma_0\defeq\emptyt$.
\item $\Gamma_0\defeq\unit$ and $\Gamma_1(\ttt,\ttt)\defeq\emptyt$.
\item $\Gamma_0\defeq\unit$ and $\Gamma_1(\ttt,\ttt)\defeq\unit$.
\end{enumerate}
Therefore $\mfGraph$ does not satisfy the law of excluded middle.
\end{itemize}

\subsubsection{Univalence for the graph model}
In the other direction, we also obtain a graph morphism $\graphcharmap[\Gamma]{f}:
\ctxhom{\Gamma}{\gobjclass}$ for every graph morphism $f:\ctxhom{\Delta}{\Gamma}$.
In \autoref{graph-object-classifier} we will prove that the maps
$\int_\Gamma:\ctxhom{\Gamma}{\gobjclass}\to\sm{\Delta:\tfGraph}\ctxhom{\Delta}{\Gamma}$
is an equivalence with iverse $\graphcharmapfunc{\Gamma}$.

\begin{defn}
Let $f:\Delta\to\Gamma$ be a graph morphism. We define the graph morphism
$\graphcharmap[\Gamma]{f}$ by
\begin{align*}
\graphcharmap[\Gamma]{f}_0 & \defeq \lam{i}\hfib{f_0}{i}\\
\graphcharmap[\Gamma]{f}_1(i,j) & \defeq \lam{q}\hfib{f_1(i,j)}{q}.
\end{align*}
\end{defn}

\begin{thm}\label{graph-object-classifier}
Main theorem here.
\end{thm}

\begin{comment}
To describe the object classifier for graphs, we will follow Streicher. Thus
we have to look at presheaves over $I/i$ for each object $i$ of the category
$I\defeq 0{\rightrightarrows}1$ with the morphisms named $s$ and $t$ for source
and target. The category $I/0$ is the terminal category;
the category $I/1$ looks like $\cdot{\rightarrow}\cdot{\leftarrow}\cdot$.
Therefore, we have
\begin{align*}
\type^{\op{(I/0)}} & \eqv{}{\type},\\
\type^{\op{(I/1)}} & \eqv{}{\sm{X,Y,A:\type}(A\to X)\times(A\to Y)}\\
& \eqv{}{\sm{X,Y:\type}X\to Y\to\type}.
\end{align*}
The functors $\type^\op{\Sigma_s}$ and $\type^\op{\Sigma_t}$ are given by
$\pi_1$ and $\pi_2$ respectively. This leads to our following definition
of the object classifier $\gobjclass$:

\begin{defn}
Define $\gobjclass$ to be the graph consisting of
\begin{align*}
\gobjclass_0 & \defeq  \type\\
\gobjclass_1(X,Y) & \defeq  X\to Y\to\type
\end{align*}
and define $\pointed{\gobjclass}$ by
\begin{align*}
(\pointed{\gobjclass})_0 & \defeq  \pointed{\type}\\
(\pointed{\gobjclass})_1(\pairr{X,x},\pairr{Y,y}) & \defeq  \sm{R:X\to Y\to\type}R(x,y)
\end{align*}
There is the obvious forgetful graph morphism $t:\pointed{\gobjclass}\to\gobjclass$,
given by projection on the first coordinate.

For any morphism $f:\Delta\to\Gamma$ of graphs we define a morphism
$\graphcharmap(f):\Gamma\to\gobjclass$ of graphs by
\begin{align*}
\graphcharmap(f)_0(i) & \defeq  \hfiber{f_0}{i}\\
\graphcharmap(f)_1(q,\pairr{u,\alpha},\pairr{v,\beta}) & \defeq  \sm{p:\Delta_1(u,v)}
\id{\trans{\pairr{\alpha,\beta}}{f_1(p)}}{q}
\end{align*}
where $\pairr{u,\alpha}:\graphcharmap(f)_0(i)$ and $\pairr{v,\beta}:\graphcharmap(f)_0(j)$. The
morphism $\graphcharmap(f)$ is called the \emph{characteristic map of $f$}. We obtain a function
\begin{equation*}
\graphcharmap : \big(\sm{\Delta:\tfGraph }\Delta\to\Gamma\big)\to\big(\Gamma\to\gobjclass\big)
\end{equation*}
for every graph $\Gamma$.
\end{defn}

\begin{thm}\label{thm:graph-classifier1}
The function $\graphcharmap$ is an equivalence for any graph $\Gamma$.
\end{thm}

\begin{proof}
We have to find a quasi-inverse
\begin{align*}
\Sigma : (\Gamma\to\gobjclass)\to\big(\sm{\Delta:\tfGraph}\Delta\to\Gamma\big)
\end{align*}
of $\graphcharmap$. Thus, we have to define $\Sigma_0:(\Gamma\to\gobjclass)\to\tfGraph$ and
$\Sigma_1:\prd{P:\Gamma\to\gobjclass}\Sigma_0(P)\to\Gamma$. For $P:\Gamma\to\gobjclass$ we define
\begin{align*}
\Sigma_0(P)_0 & \defeq \sm{i:\Gamma_0}P_0(i)\\
\Sigma_0(P)_1(\pairr{i,u},\pairr{j,v}) & \defeq \sm{q:\Gamma_1(i,j)}P_1(q,u,v)\\
\Sigma_1(P)_0 & \defeq \proj1\\
\Sigma_1(P)_1(\pairr{i,u},\pairr{j,v}) & \defeq \proj1.
\end{align*}
\end{proof}

\begin{thm}\label{conj:graph_classifier2}
For any graph morphism $f:\Delta\to\Gamma$, the diagram
\begin{equation*}
\begin{tikzcd}
\Delta \ar{r}{} \ar{d}[swap]{f} & \pointed{\gobjclass} \ar{d}{t} \\ 
\Gamma \ar{r}[swap]{\graphcharmap(f)} & \gobjclass 
\end{tikzcd}
\end{equation*}
is a pullback square.
\end{thm}
\note{We would like it to be a pb \emph{in} the graph model}
\end{comment}

\begingroup\color{blue}
\subsubsection{The adjunctions $\tfcolim\dashv\Delta\dashv\tflim$}
We define $\Delta:\type\to\tfGraph$ by
\begin{equation*}
\Delta(X)\defeq\pairr{X,\lam{x}{x'}\id{x}{x'}}
\end{equation*}
for $X:\type$. For $A:X\to\type$ we define $\Delta(A):\mftyp(\Delta(X))$ by
\begin{align*}
\Delta(A)_0(x) & \defeq A(x)\\
\Delta(A)_1(p,a,b) & \defeq \id{\trans{p}{a}}{b}.
\end{align*}

\begin{lem}
For any type $X$ and any graph $\Gamma$ there is an equivalence
\begin{equation*}
\eqv{(X\to\terms{\Gamma})}{\terms{\Delta(X)\to\Gamma}}.
\end{equation*}
\end{lem}

\begin{proof}
We have to find functions
\begin{align*}
\varphi & : (X\to\terms{\Gamma})\to\terms{\Delta(X)\to\Gamma}\\
\psi & : \terms{\Delta(X)\to\Gamma}\to X\to\terms{\Gamma}
\end{align*}
which are each others homotopy inverse. To define $\varphi$, let
\end{proof}
\endgroup

### 2013-09-22 article-models-main.tex
adding the source for the articles

\documentclass{article}

\input{preamble-articles}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title{Internal models of type theory}
\author{Egbert Rijke}
\date{\today}

\begin{document}

\maketitle

\tableofcontents

\begin{comment}
\section*{Introduction\phantomsection\addcontentsline{toc}{section}{Introduction}}
This paper contributes to the univalent homotopy type theory~\cite{kapulkin2012univalence,awodey2012type,pelayo2012homotopy}
program which connects and extends a number of topics such as type theory, (elementary) topos
theory~\cite{MacLaneMoerdijk,johnstone:elephant}, homotopy theory, higher category theory and higher topos
theory~\cite{rezk2010toposes,lurie2009higher}.
A key missing ingredient is a first-order (`elementary') definition of a higher topos. It has been
conjectured~\cite{awodey2012type,shulman2012univalence} that homotopy type theory can be used as the internal language, and thus may
provide an elementary definition of a higher topos; see also~\cite{LumsdaineWarren}.

A main ingredient in the theory of higher toposes is the descent property, which
is described in~\cite{rezk2010toposes,lurie2009higher}. The (weak) descent property
of an ordinary topos $\mathcal{E}$ is the property that
\begin{enumerate}
\item For any family $\{X_i\}_{i\in I}$ of objects of $\mathcal{E}$ with coproduct
$X:=\coprod_{i\in I}X_i$ and any function $f:Y\to X$ we get an isomorphism
\begin{equation*}
Y\cong\coprod_{i\in I} Y\times_X X_i
\end{equation*}
given by the canonical map.
\item Conversely, for any family $\{f_i:X_i\to Y_i\}_{i\in I}$ of maps inducing
$f:\coprod_{i\in I}X_i\to\coprod_{i\in I}Y_i$, the canonical map determines an
isomporphism
\begin{equation*}
Y_i\cong Y\times_X X_i.
\end{equation*}
\item
\item
\end{enumerate}
In model toposes, the descent property can be strengthened considerably, but
ordinary limits and colimits are to be replaced by \emph{homotopy} limits and
colimits. Since the only notion of limits and colimits in type theory is that
of homotopy limit and colimit, this is no restriction for us. The descent
property of model toposes as formulated by Rezk~\cite{rezk2010toposes} is: Let
$X:\mathbf{I}\to\mathcal{E}$ be a functor from a small category into a
model topos $\mathcal{E}$, with homotopy colimit $\bar{X}$. Then we have
\begin{enumerate}
\item Let $\bar{Y}$ be an object and let $f:\bar{Y}\to\bar{X}$ 
be any map in $\mathcal{E}$. Define
the functor $Y:\mathbf{I}\to\mathcal{E}$ by $Y(i):=Y\times_{\bar{X}} X_i$. Then
the canonical map $\typefont{hocolim}(Y)\to\bar{Y}$ is a weak equivalence.
\item Let $Y:\mathbf{I}\to\mathcal{E}$ be a functor and let $f:Y\to X$ be an
equifibered transformation from $Y$ to $X$. Define $\bar{f}:\bar{Y}\to\bar{X}$
be the homotopy colimit of $f$. Then the canonical map 
$Y_i\to \bar{Y}\times^h_{\bar{X}} X_i$ is a weak equivalence.
\end{enumerate}
These two properties are two sides of the same coin and in our type theoretical
setting it is natural to stress this. Given a diagram $D$, we will be able
to summarize the entire descent property in an equivalence between the type
$\typefont{equiFib}(D)$ of equifibered diagrams over $D$ and the type of functions
into $\typefont{colim}(D)$. We prove the descent property in 
theorem~\ref{thm:descent}.

We use universes ala Russell, universe polymorphism and typical ambiguity and hence create the illusion that $\type:\type$;
see~\cite{TheBook}. There are only a few places where we need to be explicit about universe levels.
As is common in HoTT, we will freely use the axiom of function extensionality.
Using this axiom, all limits can be constructed in type theory~\cite{Avigad:limits}.
We use higher inductive types (HITs) to construct co-limits; see~\cite{LumsdaineShulman,Shulman:HIT}.
Using the univalence axiom one can show that $\type$ is an object classifier~\cite{RijkeSpitters:Sets}.
An important feature property of $\infty$-toposes is the descent property. We prove this property in theorem~\ref{thm:descent}.

In the previous paragraph we have used informal categorical terminology such as `limits' to refer to what would be interpreted as homotopy
limits in the $\infty$-category of $\infty$-groupoids. We will continue to use this naive $\infty$-category style without any rigorous
claims to their interpretation in the model.

The paper is organized as follows. We start by defining limits and colimits of over graph. This is the most general case, as we
do not include the possibility of including higher coherence conditions.
We introduce diagrams and (co)limits in section~\ref{sec:diagram}. In section~\ref{sec:suspensions} we study suspensions.
Functions with contractible mapping cones are the same as epimorphisms, as we will see in section~\ref{sec:mapping}.
Finally, we prove the descent property.

\subsection{Notation}
A typical way we write a type in this article is:
\begin{equation}\label{eq:notation_example}
\prd{x,y:A}{p:B(x,y)}C(x,y,p)\to D(x,y,p).
\end{equation}
Assuming that $A$, $B$, $C$ and $D$ have been introduced, this type denotes a
dependent product, where the product ranges over $x,y:A$ and $p:B(x,y)$ and
the factors are the (non-dependent) function types $C(x,y,p)\to D(x,y,p)$. A
more common notation for the same type in mathematics would be
\begin{equation*}
\prod_{x,y:A}\prod_{p:B(x,y)}C(x,y,p)\to D(x,y,p),
\end{equation*}
like a big product of function spaces. Nevertheless we have chosen to use this
less common notation, because the types $A$ or $B$ can become quite complex.
% and formulas would become messy as as complex information starts `dancing
% around'.

Another thing we notice in formula \eqref{eq:notation_example} is the use of
curly braces around $x,y:A$. This indicates that whenever we have a term
\begin{equation*}
f:\prd{x,y:A}{p:B(x,y)}C(x,y,p)\to D(x,y,p),
\end{equation*}
we will omit reference to $x$ and $y$ whenever $f$ is applied to $x,y:A$ and
$p:B(x,y)$. In other words, we will write down $f(p):C(x,y,p)\to D(x,y,p)$
instead of $f(x,y,p)$. We will do this whenever it is possible to retrieve
the terms $x$ and $y$ from the term $p$, and in many similar situations like
this. Back to formula \autoref{eq:notation_example}, it is quite possible 
that we would have used the curly braces in the introduction
of $C$ and $D$ as well, and if we had done that we would have written $C(p)$
and $D(p)$.
\end{comment}

\input{article-models-defn}

\bibliographystyle{hplain}
%\phantomsection\addcontentsline{toc}{section}{References}
\bibliography{refs}

\end{document}

### 2013-09-27 article-models-defn.tex
resolve issue #7

@@
-\section{Introduction}
-In this paper we define a notion of internal model which is adapted to
-the univalent setting in which we work. Thus, we only describe what it means
-to be an internal model when the ambient type theory is univalent. Our
-presentation is derived from the definition of~\cite{Dybjer1996}. There,
-an internal model is defined to be a category with families with interpretations
-for the basic type constructors $\Pi$, $\Sigma$ and $\idtypevar{}$. However,
-a category in~\cite{Dybjer1996} has a set of objects and for every two objects
-a setoid of morphisms. This is too restrictive for our purpose. Ideally, we
-would start with an $\infty$-category of contexts and an $\infty$-presheaf
-of families over it. This presents us with the problem that a notion of
-$\infty$-category has not been formulated \emph{in} type theory. Therefore,
-we make a deviation from the approach in~\cite{Dybjer1996}. We shall
-start with a type $\tfctx$ of contexts, a family $\mftypfunc:\tfctx\to\type$
-of families over it and a family $\terms{\blank}:\prd*{\Gamma:\tfctx}\mftyp{\mfM}{\Gamma}
-\to\type$ sending a type $A$ in context $\Gamma$ to its type $\terms{A}$ of
-terms; then we shall directly interpret $\Pi$-types, declaring
-$\terms{A\to B}$ to be $\ctxhom[\Gamma]{A}{B}$. The idea is that by letting
-$\Pi$ play a fundamental role in the definition of a category with families
-$\mathcal{C}$,
-the morphisms get their properties directly from the internal type theory
-of $\mathcal{C}$.
-
-\begin{comment}
-We briefly list the data of which an internal model of type theory consits. We
-consider internal models in the style of~\cite{Dybjer1996}. Thus we will
-describe a model $\mfM$ as a category with families. However, since our setting is
-univalent type theory we will deviate from~\cite{Dybjer1996} in the following
-respects: first of all, we make use of typical ambiguity and hence omit
-reference to a thing called \emph{sort}.
-
-We organize the definition of an internal model $\mfM$ as follows: in
-\autoref{internal-model-contexts} we describe the category of contexts itself;
-in \autoref{internal-model-families} we describe the families over a given context
-and the related operations; in \autoref{internal-model-constructors} we
-describe the basic constructors in the internal model.
-\end{comment}
-
-\section{Ideas in the definition}
-An internal model of type theory is like a category with families, but we want
-to avoid having to state higher coherences. In fact, we don't even start our
-definition with a category of contexts; instead we just take a \emph{type} of contexts. 
-The morphisms will come from the terms, evaluation of a function at a given
-term will come from substitution. We recognize three basic ingredients to models:
-first there is a type of contexts; second, for every context there is a model of types in
-that context and third, for every type in a given context there is a type of its
-terms. Then there are three basic attributes: context extension, weakening and
-substitution. Context extension provides us with families over types as well as
-with an interpretation of dependent pair types. We need weakening 
-so that families can depend on the same type multiple times (the way the
-identity type of a type depends two times on that type) and to be able
-to talk about non-dependent function types,
-the morphims of our category. Substitution will give us a way
-to work with fibers of families as well as composition of functions and evaluation
-of functions at terms.
-
-Because we require a \emph{model} of types in a context, all the structure
-which we require at the bottom level will be required to exist higher up as well.
-Thus, the model of types in a given context $\Gamma$ will have a type of contexts
-itself, which can be seen as the type of types in $\Gamma$; it will have its
-own notion of types in a context, its own notion of terms, context extension,
-weakening and substitution together with all the structure require for it. For
-instance, when $A$ is a type in context $\Gamma$ in a model $\mfM$, then there
-is the model of types in context $A$, which is the model of families over $A$. 
-This model is required to be \emph{definitionally equal to} the model of types
-in the context $\ctxext{\Gamma}A$, the context extension of $\Gamma$ and $A$.
-In this way we protect ourselves from the need to dig an infinitely deep structure
-of models when we want to consider examples.
-
-To give the definition of a model we shall also need to consider certain morphisms
-of internal models. Those should preserve all the structure: contexts are mapped
-to contexts; for every context a morphism of models mapping the model of types
-in that context to the model of types in the image of that context; there should
-be a mapping of terms and context extension, weakening and substitution should be
-preserved. We need to consider those morphisms because we require context extension,
-weakening and substitution to be of that kind, thereby respecting each other
-in all possible ways.
-
-When we have this framework set up, we can interpret the basic type constructors
-such as $\Pi$, $\Sigma$ and $\idtypevar{}$.
-The higher categorical structure then comes from the
-result that we have an interpretation of type theory.
-
-\begingroup
-\color{red}
-\begin{rmk}
-Currently, the definition seems to be circular. To define a model we need that
-context extension, weakening and substitution be morphisms of models. A morphism
-of models needs to preserve context extension, weakening and substitution.
-Moreover, its definition requires the notion of composition of morphisms.
-
-I've done it this way because it guides me to what the rules should be, trusting
-that I can work my way back to give a (possibly less transparent) definition
-which contains no circularities without doubt.
-\end{rmk}
-\endgroup
-
-\section{Univalent universes as internal models}
-
-Before we give the definition we illustrate the concepts that go into it in
-the case of a univalent universe $\UU$.
-Regardless of the definition of an internal model, a univalent universe should
-be an example of one.
-
-The type of contexts is $\UU$ itself and for every $\Gamma:\UU$, a type in
-context $\Gamma$ is simply a family $A:\Gamma\to\UU$. The type $\terms{A}$ of terms of a type $A$
-in context $\Gamma$ is defined to be $\prd{i:\Gamma}A(i)$. Note that $\Gamma\to
-\UU$ itself also interprets type theory where a type in context $A:\Gamma\to\UU$
-is a family $P:(\sm{i:\Gamma}A(i))\to\UU$. We may denote this model by
-$\mftyp{\UU}{\Gamma}$. The terms of a type $P$ in context $A$ in the model
-$\Gamma\to\UU$ are the terms of $\prd{i:\Gamma}{x:A(i)}P(i,x)$.
-
-When $A$ is a type in context
-$\Gamma$ we define the context extension $\ctxext{\Gamma}A$ to be
-$\sm{i:\Gamma}A(i)$. Note that context extension $\ctxext{\Gamma}\blank$
-can be seen acting not only as a function from the context of the model
-$\Gamma\to\UU$ to the context of the model $\UU$, but it also acts on the types 
-and terms: when $A$ is a type in context $\Gamma$ we may take the identity map
-from $\mftyp{\Gamma\to\UU}{A}\to\mftyp{\UU}{\ctxext{\Gamma}A}$ because
-$\mftyp{\Gamma\to\UU}{A}$ is taken to be $\mftyp{\UU}{\ctxext{\Gamma}A}$;
-thus, $\ctxext{\Gamma}P\defeq P$ for every family $P$ over $A$ in context 
-$\Gamma$. Likewise, when $P$ is a type in context $A$ in the model 
-$\Gamma\to\UU$ then context extension should act on the terms of $P$ via a
-function $\terms{P}\to\terms{\ctxext{\Gamma}P}$, which we take to be
-the identity map once more.
-
-Since $\mftyp{\UU}{\Gamma}$ is a model of type theory it
-has it's own notion of context extension: when $P$ is a type in context $A$ in
-the model $\mftyp{\UU}{\Gamma}$ then $\ctxext{A}P$ is the family
-$\lam{i}\sm{x:A(i)}P(i,x):\Gamma\to\UU$. Also the context extension of
-$\Gamma\to\UU$ acts trivially on the types and terms. Context extension is
-analoguous to the Grothendieck construction that associates the category of
-elements to a presheaf and it gives us $\Sigma$-types.
-
-When $A$ and $B$ are types in context
-$\Gamma$, the weakening $\ctxwk{A}{B}$ of $B$ along $A$ is defined to be
-$\lam{\pairr{i,x}}B(i):(\sm{i:\Gamma}A(i))\to\UU$. Weakening along a type $A$ 
-in context $\Gamma$ also acts on types and terms. When $Q$ is a type in context
-$B$ in the model $\Gamma\to\UU$, we define $\ctxwk{A}{Q}$ to be 
-$\lam{\pairr{i,x}}{y}Q(i,y)$. When $g:\terms{Q}$ we define $\ctxwk{A}{g}$ to
-be $\lam{\pairr{i,x}}{y}g(i,y)$.
-
-Note that for two types $A$ and $B$ in context $\Gamma$, the terms of
-$\ctxwk{A}{B}$ are the terms of $\prd{i:\Gamma}A(i)\to B(i)$, i.e.~they are
-the fiberwise maps from $A$ to $B$. We shall take the type $\terms{\ctxwk{A}{B}}$
-of terms of $\ctxwk{A}{B}$ to be the type of morphisms from $A$ to $B$. Also,
-any context of $\UU$ may be seen as a type in the empty context $\unit$. Thus
-the type of terms of a context $\Gamma$ is $\unit\to\Gamma$, which is
-equivalent to $\Gamma$. A context morphism from $\Delta$ to $\Gamma$ is a term
-of $\ctxwk{\Gamma}{\Delta}$, i.e.~a function from $\Gamma$ to $\Delta$. We denote
-the type $\terms{\ctxwk{\Gamma}{\Delta}}$ by $\ctxhom{\Delta}{\Gamma}$.
-
-When $P$ is a family over
-$A$ in context $\Gamma$ and $x$ is a term of $A$ we define the type $\subst{x}{P}$
-in context $\Gamma$ to be $\lam{i}P(i,x(i))$. Like extension
-
-We will interpret the dependent product $\mprd{A}{P}$ of a family $P$ over
-$A$ in context $\Gamma$ by
-\begin{equation*}
-\mprd{A}{P}(i)\defeq \prd{x:A(i)}P(i,x)
-\end{equation*}
-With this interpretation there is an equivalence 
-$\lambda:\eqv{\terms{P}}{\terms{\mprd{A}{P}}}$. This is lambda-abstraction; its
-inverse being evaluation. We should note however, that the rule for evaluation
-we interpret here is
-\begin{equation*}
-\inference{\Gamma\vdash f:\mprd{A}{P}}{\ctxext{\Gamma}A\vdash\tfev(f):P}
-\end{equation*}
-which is different than the usual rule
-\begin{equation*}
-\inference{\Gamma\vdash f:\mprd{A}{P} \qquad \Gamma\vdash a:A}{\Gamma \vdash \tfev(f,a):\subst{a}{P}}
-\end{equation*}
-The reason for this is that the interpretation of the usual rule would give a
-function of type $\terms{\mprd{A}{P}}\to\prd{x:\terms{A}}\terms{\subst{x}{P}}$,
-but this does not describe the terms of $\mprd{A}{P}$ by any means. Moreover,
-since we have implemented substitution, we obtain from $\ctxext{\Gamma}A\vdash\tfev(f):P$ 
-and $\Gamma\vdash x:A$ a term $\Gamma\vdash\subst{x}{\tfev(f)}:\subst{x}{P}$ and
-therefore we do not loose anything with this approach.
-
-{\color{red} point to example}
-
 \section{A variable-free presentation of type theory}
@@
 
-\section{The category with families $\mfM$}\label{internal-model-contexts}
+\section{Internal models of type theory}
+\subsection{Models of type theory without basic constructors}\label{internal-model-contexts}
 \begin{defn}\label{defn:premodel}
@@
 
-\section{Morphisms of internal models}
+\subsection{Morphisms of internal models}
 \begin{defn}\label{defn:premodel-morphism}
@@
 
-\section{The basic type constructors in internal models}
+\subsection{The basic type constructors in internal models}
 
@@
 \end{comment}
-
-\section{The graph model of type theory}
-In this section we define the graph model of type theory, denoted by
-$\mfGraph$.  In our presentation, we follow that of the definition internal
-models. After we have established the graph model, we show how the graph
-model can be seen as a presheaf model, namely over the category $\cdot
-{\rightrightarrows}\cdot$.
-
-\begin{defn}
-A \emph{(directed) graph} $\Gamma$ is a pair $\pairr{\Gamma_0,\Gamma_1}$ 
-consisting of a type $\Gamma_0$ of vertices and a family 
-$\Gamma_1:\Gamma_0\to\Gamma_0\to\type$ of edges. The type $\ctx(\mfGraph)$
-is defined to be the type of all graphs; we will usually denote it by
-$\tfGraph$. Explicitly, we have
-\begin{equation*}
-\tfGraph\defeq\sm{\Gamma_0:\type}\Gamma_0\to\Gamma_0\to\type.
-\end{equation*}
-\end{defn}
-
-\begin{eg}\label{ex:pb}
-The underlying graph of the diagram
-\begin{equation*}
-\begin{tikzcd}
-{} & A \ar{d}{f} \\
-B \ar{r}[swap]{g} & C
-\end{tikzcd}
-\begin{comment}
-\begin{tikzpicture}
-\matrix (m) [std] { & A \\ B & C \\};
-\draw[ar] (m-1-2) -- node[right] {$f$} (m-2-2);
-\draw[ar] (m-2-1) -- node[below] {$g$} (m-2-2);
-\end{tikzpicture}
-\end{comment}
-\end{equation*}
-is $I\defeq\mathbf{3}$ and has $J(1,3)\defeq J(2,3)\defeq \unit $ 
-and $J(x,y)\defeq\emptyt$ otherwise (it is defined using the induction 
-principle of $\mathbf{3}$ and a universe).
-\end{eg}
-
-\begin{defn}
-A \emph{graph $A$ in the context $\Gamma$} is a pair $\pairr{A_0,A_1}$ consisting
-of 
-\begin{align*}
-A_0 & :\Gamma_0\to\type\\
-A_1 & :\prd*{i,j:\Gamma_0}\Gamma_1(i,j)\to A_0(i)\to A_0(j)\to\type.
-\end{align*}
-Thus, for a graph $\Gamma$, the type $\mftyp{\mfGraph}{\Gamma}$ is the type
-\begin{equation*}
-\sm{A_0:\Gamma_0\to\type}\prd*{i,j:\Gamma_0}\Gamma_1(i,j)\to A_0(i)\to A_0(j)\to\type.
-\end{equation*}
-We will also write $\Gamma\vdash A:\mfGraph$ when $A$ is a graph in context
-$\Gamma$.
-\end{defn}
-
-\begin{defn}
-Suppose that $\Gamma\vdash A:\mfGraph$. Then we define the graph $\ctxext{\Gamma}A$
-by
-\begin{align*}
-\ctxext{\Gamma}A_0 & \defeq \sm{i:\Gamma_0}A_0(i)\\
-\ctxext{\Gamma}A_1(\pairr{i,x},\pairr{j,y}) & \defeq \sm{q:\Gamma_1(i,j)}A_1(q,x,y).
-\end{align*}
-\end{defn}
-
-\begin{defn}
-Suppose that $\Gamma\vdash A:\mfGraph$ and $\Gamma\vdash B:\mfGraph$. Then we
-define the graph $\ctxwk{A}{B}$ in context $\ctxext{\Gamma}A$ by
-\begin{align*}
-(\ctxwk{A}{B})_0(\pairr{i,x}) & \defeq B_0(i)\\
-(\ctxwk{A}{B})_1(\pairr{q,e},u,v) & \defeq B_1(q,u,v).
-\end{align*}
-\end{defn}
-
-\begin{defn}
-Suppose that $\Gamma\vdash A:\mfGraph$. A term $x$ of $A$ consists of a pair
-$\pairr{x_0,x_1}$ where
-\begin{align*}
-x_0 & : \prd{i:\Gamma_0}A_0(i)\\
-x_1 & : \prd*{i,j:\Gamma_0}{q:\Gamma_1}A_1(q,x_0(i),x_0(j)).
-\end{align*}
-Thus we define
-\begin{equation*}
-\terms{A}\defeq\sm{x_0:\prd{i:\Gamma_0}A_0(i)}\prd*{i,j:\Gamma_0}{q:\Gamma_1}A_1(q,x_0(i),x_0(j)).
-\end{equation*}
-We also write $\Gamma\vdash x:A$ when $x$ is a term of the graph $A$ in context
-$\Gamma$.
-\end{defn}
-
-\begin{defn}
-Suppose that $P$ is a family of graphs over $A$ in context $\Gamma$ and let
-$x$ be a term of $A$. Then we define the graph $\subst{x}{P}$ 
-in context $\Gamma$ by
-\begin{align*}
-\subst{x}{P}_0(i) & \defeq P_0(\pairr{i,x_0(i)})\\
-\subst{x}{P}_1(q,u,v) & \defeq P_1(\pairr{q,x_1(q)},u,v)
-\end{align*}
-\end{defn}
-
-\begin{rmk}
-Note that we have the judgmental equality $\subst{x}{\ctxwk{A}{B}}\jdeq B$
-for every two graphs $A$ and $B$ in context $\Gamma$ and every term $x$ of $A$.
-\end{rmk}
-
-\begin{defn}
-The terminal graph $\unit^\mfGraph$, which we shall often denote simply 
-by $\unit$, is defined by
-\begin{align*}
-{\unit^\mfGraph}_0 & \defeq \unit\\
-{\unit^\mfGraph}_1(x,y) & \defeq \unit.
-\end{align*}
-\end{defn}
-
-\begin{rmk}
-We note that the function $\ctxext{\unit}\blank$ is an equivalence
-from $\mftyp{\mfGraph}{\unit}$ to $\tfGraph$. It's inverse is the function which maps
-a graph $\Gamma$ to the pair $\pairr{A_0,A_1}$ where $A_0$ is defined by
-$A_0(\ttt)\defeq\Gamma_0$ and where $A_1$ is defined by $A_1(\ttt)\defeq
-\Gamma_1$.
-
-Note that we only have an equivalence here, not a judgmental equality.
-\end{rmk}
-
-\subsubsection{The contexts of $\mfGraph$}
-
-The category $\psh(\cdot{\rightrightarrows}\cdot)$ of presheaves over
-$\cdot{\rightrightarrows}\cdot$ is given by
-\begin{equation*}
-\sm{A_0,A_1:\type}(A_1\to A_0)\times(A_1\to A_0).
-\end{equation*}
-To see that $\psh(\cdot{\rightrightarrows}\cdot)$ is indeed equivalent to
-the type $\tfGraph$ of all graphs, note that we have the equivalences
-\begin{align*}
-\psh(\cdot{\rightrightarrows}\cdot) & \eqvsym \sm{A_0,A_1:\type}A_1\to A_0\times A_0\\
-& \eqvsym \sm{A_0:\type} A_0\times A_0\to\type\\
-& \eqvsym \tfGraph.
-\end{align*}
-
-\begin{defn}
-Suppose that $\Delta\defeq \pairr{\Delta_0,\Delta_1}$ and $\Gamma\defeq \pairr{\Gamma_0,
-\Gamma_1}$ are graphs. A morphism $f$ from $\Delta$ to $\Gamma$ is a pair
-$\pairr{f_0,f_1}$ consisting of
-\begin{align*}
-f_0 & : \Delta_0\to\Gamma_0\\
-f_1 & : \prd*{u,v:\Delta_0}\Delta_1(u,v)\to\Gamma_1(f_0(u),f_0(v))
-\end{align*}
-We write $\ctxhom{\Delta}{\Gamma}$ for the type of morphisms of graphs.
-\end{defn}
-
-The terminal graph $\unit^\mfGraph$ plays the role of the empty context.
-
-\begin{defn}
-The terminal graph $\unit^\mfGraph$, which we shall often denote simply 
-by $\unit$, is defined by
-\begin{align*}
-{\unit^\mfGraph}_0 & \defeq \unit\\
-{\unit^\mfGraph}_1(x,y) & \defeq \unit.
-\end{align*}
-\end{defn}
-
-It is immediate that the type of graph morphisms $\Gamma\to\unit^\mfGraph$ is
-contractible for every graph $\Gamma$. The terms of a graph $\Gamma$ are the
-global elements of $\Gamma$, i.e.~they are the morphisms of type 
-$\ctxhom{\unit^\mfGraph}{\Gamma}$.
-
-\begin{defn}
-Let $\Gamma$ be a graph. The type $\terms[\mfGraph]{\Gamma}$ of terms of $\Gamma$ is
-defined to be
-\begin{equation*}
-\sm{x_0:\Gamma_0}\Gamma_1(x_0,x_0).
-\end{equation*}
-\end{defn}
-
-We now turn to the description of the universe $\gobjclass$ of graphs. Note
-that for the category $I\defeq 0{\rightrightarrows}1$, we have
-$I/0\defeq\catid{0}$ and $I/1\defeq s\rightarrow\catid{1}\leftarrow t$, where
-$s$ and $t$ are the morphisms $0\to 1$. Thus we have 
-\begin{align*}
-\psh(I/0) & \eqvsym \type\\
-\psh(I/1) & \eqvsym \sm{X,Y:\type}X\to Y\to\type.
-\end{align*}
-The functors $\psh(\Sigma_s)$ and $\psh(\Sigma_t)$ map a presheaf 
-$\pairr{X,Y,R}$ to $X$ and $Y$, respectively. 
-Thus we obtain the following graph $\gobjclass$.
-
-\begin{defn}
-The universe $\gobjclass$ of graphs is defined to be
-\begin{align*}
-{\gobjclass}_0 & \defeq \type\\
-{\gobjclass}_1(X,Y) & \defeq X\to Y\to\type.
-\end{align*}
-\end{defn}
-
-Note that the type $\terms{\gobjclass}$ of terms of $\gobjclass$ is 
-exactly $\tfGraph$.
-
-\subsubsection{Families of graphs}
-Using $\gobjclass$, we obtain a description of what
-it means to be a graph in a context $\Gamma$, where $\Gamma$ is itself a
-graph. The notion of ``a graph in a context $\Gamma$'' is the interpretation
-of ``a type in a context $\Gamma$''.
-
-\begin{defn}
-A \emph{graph $A$ in the context $\Gamma$} is a pair $\pairr{A_0,A_1}$ consisting
-of 
-\begin{align*}
-A_0 & :\Gamma_0\to\type\\
-A_1 & :\prd*{i,j:\Gamma_0}\Gamma_1(i,j)\to A_0(i)\to A_0(j)\to\type.
-\end{align*}
-We will also write $\Gamma\vdash A:\mfGraph$ when $A$ is a graph in context
-$\Gamma$.
-\end{defn}
-
-\begin{rmk}
-Note that every graph can be seen as a graph in context $\unit^\mfGraph$.
-\end{rmk}
-
-\begin{defn}
-Let $A$ be a graph in context $\Gamma$ and let $f:\ctxhom{\Delta}{\Gamma}$
-be a graph morphism. We define the graph $A[f]$ in context $\Delta$ by
-\begin{align*}
-A[f]_0(u) & \defeq A_0(f_0(u))\\
-A[f]_1(p,x,y) & \defeq A_1(f_1(p),x,y).
-\end{align*}
-\end{defn}
-
-\begin{defn}
-Let $\Gamma\vdash A:\mfGraph$ and $\Gamma\vdash B:\mfGraph$ be two graphs in
-the context $\Gamma$. A morphism $f$ from $A$ to $B$ is a pair $\pairr{f_0,f_1}$
-consisting of
-\begin{align*}
-f_0 & : \prd*{i:\Gamma_0} A_0(i)\to B_0(i)\\
-f_1 & : \prd*{i,j:\Gamma_0}{q:\Gamma_0}*{x:A_0(i)}*{y:A_0{j}}A_1(q,x,y)\to B_1(q,f_0(x),f_0(y)).
-\end{align*}
-We write $\ctxhom[\Gamma]{A}{B}$ for the type of morphisms between graphs in
-context $\Gamma$.
-\end{defn}
-
-\begin{defn}
-The unit graph $\unit^\Gamma$ in context $\Gamma$ consists of
-\begin{align*}
-(\unit^\Gamma)_0 & \defeq \lam{i}\unit\\
-(\unit^\Gamma)_1(q) & \defeq \lam{x}{y}\unit.
-\end{align*}
-\end{defn}
-
-As was the case with graphs, terms of graphs in a context are described by
-morphisms from $\unit$. Explicating this in more detail, we get
-
-\begin{defn}
-Let $A$ be a graph in context $\Gamma$. A term $x$ of $A$ is a pair
-$\pairr{x_0,x_1}$ consisting of
-\begin{align*}
-x_0 & : \prd{i:\Gamma_0}A_0(i)\\
-x_1 & : \prd*{i,j:\Gamma_0}{q:\Gamma_1(i,j)}A_1(q,x_0(i),x_0(j)).
-\end{align*}
-We define $\terms[\mfGraph]{A}$ to be the type of terms of $A$.
-\end{defn}
-
-\begin{rmk}
-Note that the type of terms of a graph $\Gamma$ is equivalent to the type
-of terms of the graph $\Gamma$ in context $\unit^\mfGraph$.
-\end{rmk}
-
-\begin{defn}
-Suppose that $\Gamma\vdash A:\mfGraph$. Then we define the graph $\ctxext{\Gamma}A$
-by
-\begin{align*}
-(\ctxext{\Gamma}A)_0 & \defeq \sm{i:\Gamma_0}A_0(i)\\
-(\ctxext{\Gamma}A)_1(\pairr{i,x},\pairr{j,y}) & \defeq \sm{q:\Gamma_1(i,j)}A_1(q,x,y).
-\end{align*}
-There is a morphism $\pi^A:\ctxhom{\ctxext{\Gamma}A,\Gamma}$ given by
-\begin{align*}
-\pi^A_0 & \defeq \lam{\pairr{i,x}}i && : (\sm{i:\Gamma_0}A_0(i))\to \Gamma_0\\
-\pi^A_1(\pairr{i,x},\pairr{j,y}) & \defeq \lam{\pairr{q,e}}q && : (\sm{q:\Gamma_1(i,j)}A_1(q,x,y))\to\Gamma_1(i,j).
-\end{align*}
-\end{defn}
-
-\begin{rmk}
-The construction of $\ctxext{\Gamma}A$ is similar to the Grothendieck
-construction of the category of elements for a presheaf.
-\end{rmk}
-
-\begin{defn}
-Let $\Gamma\vdash A:\mfGraph$. A family $P$ of graphs over $A$ is by
-definition a graph $P$ in the context $\ctxext{\Gamma}A$. More explicitly, a family
-$P$ of graphs over $A$ in context $\Gamma$ consists of
-\begin{align*}
-P_0 & : \prd*{i:\Gamma_0}A_0(i)\to\type\\
-P_1 & : \prd*{i,j:\Gamma_0}{q:\Gamma_1(i,j)}*{x:A_0(i)}*{y:A_0(j)} A_1(q,x,y)\to P_0(x)\to P_0(y)\to\type
-%P_1 & : \prd{\pairr{i,x},\pairr{j,y}:(\ctxext{\Gamma}A)_0}(\ctxext{\Gamma}A)_1(\pairr{i,x},\pairr{j,y})\to P_0(i,x)\to P_0(j,y)\to\type.
-\end{align*}
-\end{defn}
-
-\begin{defn}
-When $P$ is a family of graphs over $A$ in context $\Gamma$ and $x:A$, we
-get a graph $\subst{x}{P}$ in context $\Gamma$ given by
-\begin{align*}
-\subst{x}{P}_0(i) & \defeq P_0(x_0(i))\\
-\subst{x}{P}_1(q,u,v) & \defeq P_1(\pairr{q,x_1(q)},u,v)
-\end{align*}
-\end{defn}
-
-\subsubsection{The basic type constructors for graphs}
-With families of graphs being available, we can give the interpretations of
-dependent products, dependent sums and identity types. In the following, we
-shall introduce the graph interpretations of dependent products, dependent
-sums and identity types and describe the terms of the resulting graphs.\note{These definitions
-should be connected to Mike's article but I don't really know how to do this}
-
-\begin{defn}
-Let $P$ be a family of graphs over $A$, where $\Gamma\vdash A:\mfGraph$. 
-The dependent function graph $\mprd{A}{P}$ in context $\Gamma$ consists of
-\begin{align*}
-\mprd{A}{P}_0(i) & \defeq \prd{x:A_0(i)}P_0(x)\\
-\mprd{A}{P}_1(q,f,g) & \defeq \prd*{x:A_0(i)}*{y:A_0(j)}{e:A_1(q,x,y)}P_1(\pairr{q,e},f(x),g(y)).
-\end{align*}
-\end{defn}
-
-\begin{rmk}
-A term $f:\mprd{A}{P}$ consists of
-\begin{align*}
-f_0 & : \prd*{i:\Gamma_0}{x:A_0(i)}P_0(x)\\
-f_1 & : \prd*{i,j:\Gamma_0}{q:\Gamma_1(i,j)}*{x:A_0(i)}*{y:A_0(j)}{e:A_1(q,x,y)}P_1(\pairr{q,e},f_0(x),f_0(y))
-\end{align*}
-Therefore, we see that $\eqv{\terms{\mprd{A}{P}}}{\terms{P}}$. \emph{Warning:} it
-is by no means the case that $\eqv{\terms{\mprd{A}{P}}}{\prd{x:\terms{A}}
-\terms{\subst{x}{P}}}$ for all families $P$ over $A$. For instance, the graph
-$\tilde{\emptyt}$ defined by $\tilde{\emptyt}_0\defeq\unit$ and 
-$\tilde{\emptyt}_1(\ttt,\ttt)\defeq\emptyt$ has no terms and neither does
-$\tilde{\emptyt}+\tilde{\emptyt}$. Nevertheless, there are two graph morphisms
-from $\tilde{\emptyt}$ to $\tilde{\emptyt}+\tilde{\emptyt}$. More generally,
-when $\Gamma$ is a graph such that $\Gamma_1(i,j)\jdeq\emptyt$ for all $i,j:\Gamma_0$,
-then $\eqv{{\terms{\Gamma\to\Gamma'}}}{\Gamma_0\to\Gamma^\prime_0}$.
-\end{rmk}
-
-\begin{defn}
-If $P$ is a family of graphs over $\Gamma$, the dependent pair graph
-$\msm{A}{P}$ consists of
-\begin{align*}
-\msm{A}{P}_0(i) & \defeq \sm{x:A_0(i)}P_0(x)\\
-\msm{A}{P}_1(q,\pairr{x,u},\pairr{y,v}) & \defeq \sm{e:A_1(q,x,y)}P_1(\pairr{q,e},u,v).
-\end{align*}
-\end{defn}
-
-\begin{rmk}
-A term $w:\msm{A}{P}$ consists of
-\begin{align*}
-w_0 & : \prd{i:\Gamma_0}\sm{x:A_0(i)}P_0(x)
-\intertext{and, writing $\lam{i}\proj1(w_0(i))$ and $\lam{i}\proj2(w_0(i))$ as
-$w_{00}$ and $w_{01}$ respectively,}
-w_1 & : \prd*{i,j:\Gamma_0}{q:\Gamma_1(i,j)}\sm{e:A_1(q,w_{00}(i),w_{00}(j)}P_1(\pairr{q,e},w_{01}(i),w_{01}(j)).
-\end{align*}
-By $\choice{\infty}$ it follows that
-\begin{equation*}
-\eqv{\terms{\msm{A}{P}}}{\sm{x:\terms{A}}\terms{\subst{x}{P}}}.
-\end{equation*}
-\end{rmk}
-
-\begin{rmk}
-For any two graphs $A$ and $B$ in a context $\Gamma$, we have the graph
-$B[\pi^A]$ in the context $\ctxext{\Gamma}A$, which we may just write
-as $B$. Therefore, we can
-consider the graphs $\mprd{A}{B}$ and $\msm{A}{B}$. As usually those shall be
-denoted  by $A\to B$ and $A\times B$, respectively.
-
-A term $f$ of $A\to B$ consists of
-\begin{align*}
-f_0 & : \prd*{i:\Gamma_0}A_0(i)\to B_0(i)\\
-f_1 & : \prd*{i,j:\Gamma_0}{q:\Gamma_1(i,j)}*{x:A_0(i)}*{y:A_0(j)}A_1(q,x,y)\to B_1(q,f_0(x),f_0(y)),
-\end{align*}
-so we see that $\terms{A\to B}\jdeq\ctxhom[\Gamma]{A}{B}\jdeq\terms{B[\pi^A]}$.
-
-A term $\pairr{x,y}$ of $A\times B$ consists of
-\begin{align*}
-\pairr{x,y}_0 & : \prd*{i:\Gamma_0} A_0(i)\times B_0(i)\\
-\pairr{x,y}_1 & : \prd*{i,j:\Gamma_0}{q:\Gamma_1(i,j)} A_1(q,x_0(i),x_0(j))\times B_1(q,y_0(i),y_0(j))
-\end{align*}
-so we see that $\eqv{\terms{A\times B}}{\terms{A}\times\terms{B}}$.
-
-Note that we also obtain
-a \emph{graph} $\Delta\to\Gamma$ this way, for every two graphs $\Delta$ and
-$\Gamma$ seen as graphs in the context $\unit^\mfGraph$. 
-The type $\terms{\Delta\to\Gamma}$ of terms of $\Delta\to\Gamma$ is just
-$\ctxhom{\Delta}{\Gamma}$.
-\end{rmk}
-
-\begin{defn}
-For any graph $\Gamma$ there is a graph $\tffam{\Gamma}$ in context 
-$\gobjclass$ defined by
-\begin{align*}
-(\tffam{\Gamma})_0 & \defeq \lam{X}X\to\Gamma_0\\
-(\tffam{\Gamma})_1(X,Y,R) & \defeq \lam{f}{g}\prd*{x:X}*{y:Y}R(x,y)\to\Gamma_1(f(x),g(y)).
-\end{align*}
-\end{defn}
-
-\begin{rmk}
-A term $D$ of $\msm{\gobjclass}{\tffam{\Gamma}}$ consists of a term
-$\pairr{\Delta_0,f_0}$ of type
-\begin{equation*}
-\msm{\gobjclass}{\tffam{\Gamma}}_0\jdeq \sm{\Delta_0:\type}\Delta_0\to\Gamma_0
-\end{equation*}
-and a term $\pairr{\Delta_1,f_1}$ of type
-\begin{equation*}
-\msm{\gobjclass}{\tffam{\Gamma}}_1(q)\jdeq \sm{\Delta_1}
-\end{equation*}
-\end{rmk}
-
-\begin{defn}
-There is a graph morphism
-\begin{equation*}
-\graphcharmapfunc{\Gamma} : \sm{\gobjclass}
-\end{equation*}
-\end{defn}
-
-\begin{defn}
-Let $A$ be a graph in context $\Gamma$. We define the family $\idtypevar{A}$ over $\ctxwk{A}{A}$ in
-context $\ctxext{\Gamma}A$ by
-\begin{align*}
-(\idtypevar{A})_0(\pairr{i,x},y) & \defeq \id{x}{y}\\
-(\idtypevar{A}){}_1(\pairr{q,e},d,\alpha,\alpha') & \defeq \id{\trans{\pairr{\alpha,\alpha'}}{e}}{d}
-\end{align*}
-where $q:\Gamma_1(i,j)$, $e:A_1(q,x,x')$, $d:A_1(q,y,y')$, $\alpha:\id{x}{y}$
-and $\alpha':\id{x'}{y'}$. The transportation along the path 
-$\pairr{\alpha,\alpha'}:\id{\pairr{x,x'}}{\pairr{y,y'}}$ in $A_0(i)\times A_0(j)$
-is taken with respect to the family $\lam{x}{x'}A_1(q,x,x')$.
-
-We define the term $\refl{A}$ of the family 
-$\subst{\idfunc[A]}{\idtypevar{A}}$ over $A$ in context $\Gamma$ by
-\begin{align*}
-(\reflf{A})_0(i) & \defeq \lam{x}\refl{x}\\
-(\reflf{A})_1(q) & \defeq \lam{e}\refl{e}
-\end{align*}
-\end{defn}
-
-\begin{defn}
-Let $D$ be a family over $\idtypevar{A}$ in context 
-$\ctxext({\Gamma}{A})\ctxwk{A}{A}$. Then we have the family
-$\subst{\idfunc[A]}{D}$ over $\subst{\idfunc[A]}{\idtypevar{A}}$ in
-context $\ctxext{\Gamma}A$ given by
-\begin{align*}
-\subst{\idfunc[A]}{D}_0(\pairr{i,x},\alpha) & \defeq D_0(\pairr{i,x,x},\alpha)\\
-\subst{\idfunc[A]}{D}_1(\pairr{q,e},\gamma) & \defeq D_1(\pairr{q,e,e},\gamma).
-\end{align*}
-The family $\subst{\reflf{A}}{\subst{\idfunc[A]}{D}}$ over $A$ in
-context $\Gamma$ is given by
-\begin{align*}
-\subst{\reflf{A}}{\subst{\idfunc[A]}{D}}_0(i,x) & \defeq D_0(i,x,x,\refl{x})\\
-\subst{\reflf{A}}{\subst{\idfunc[A]}{D}}_1(q,e) & \defeq D_1(q,e,e,\refl{e}).
-\end{align*}
-To show that the identity graphs correctly interpret the identity elimination
-rule, we must give a function
-\begin{equation*}
-\tfJ : \terms{\subst{\reflf{A}}{\subst{\idfunc[A]}{D}}}\to\terms{D}.
-\end{equation*}
-Note that a term $d$ of $\subst{\reflf{A}}{\subst{\idfunc[A]}{D}}$
-consists of
-\begin{align*}
-d_0 & : \prd*{i:\Gamma_0}{x:A_0(i)}D_0(i,x,x,\refl{x})\\
-d_1 & : \prd*{i,j:\Gamma_0}{q:\Gamma_1(i,j)}*{x:A_0(i)}*{y:A_0(j)}{e:A_1(q,x,y)}D_1(q,e,e,\refl{e})
-\end{align*}
-A simple argument using path induction reveals that terms of
-$\subst{\reflf{A}}{\subst{\idfunc[A]}{D}}$ indeed yield terms of $D$. 
-\end{defn}
-
-\begin{rmk}
-Using the identity graph $\idtypevar{A}$ we can describe the identity
-graph $\id[A]{x}{y}$ in context $\Gamma$ for any two terms $x,y:A$. 
-The graph $\id[A]{x}{y}$ in context $\Gamma$ consists of
-\begin{align*}
-(\id[A]{x}{y})_0(i) & \defeq \id{x_0(i)}{y_0(i)}\\
-(\id[A]{x}{y})_1(q,\alpha,\beta) & \defeq \id{\trans{\pairr{\alpha,\beta}}{x_1(q)}}{y_1(q)}
-\end{align*}
-From this, we see that a term of $p:\id[A]{x}{y}$ consists of
-\begin{align*}
-p_0 & : \prd{i:\Gamma_0}\id{x_0(i)}{y_0(i)}\\
-p_1 & : \prd*{i,j:\Gamma_0}{q:\Gamma_1(i,j)}\id{\trans{\pairr{p_0(i),p_0(j)}}{x_1(q)}}{y_1(q)}.
-\end{align*}
-\end{rmk}
-
-\subsection{Properties of graphs}
-
-\subsubsection{Contractibility and equivalences of graphs}
-
-\begin{defn}
-A graph $A$ in context $\Gamma$ is said to be \emph{contractible} if there
-is a term of the graph
-\begin{equation*}
-\msm{A}{\mprd{\ctxwk{A}{A}}{\idtypevar{A}}}
-\end{equation*}
-in context $\Gamma$.
-\end{defn}
-
-\begin{lem}\label{lem:contractible-graphs}
-Let $A$ be a graph in context $\Gamma$. The following are equivalent:
-\begin{enumerate}
-\item $A$ is a contractible graph.
-\item Both $A_0(i)$ and $A_1(q,x,y)$ are always contractible.
-\end{enumerate}
-\end{lem}
-
-\begin{proof}
-Let $H:\msm{A}{\mprd{\ctxwk{A}{A}}{\idtypevar{A}}}$. Unfolding the definitions, we have
-an element $H_0(i)$ of type
-\begin{align*}
-\msm{A}{\mprd{\ctxwk{A}{A}}{\idtypevar{A}}}_0(i) & \jdeq \sm{x:A_0(i)}\mprd{\ctxwk{A}{A}}{\idtypevar{A}}_0(i,x)\\
-& \jdeq \sm{x:A_0(i)}\prd{y:A_0(i)}(\idtypevar{A})_0(\pairr{i,x},y)\\
-& \jdeq \sm{x:A_0(i)}\prd{y:A_0(i)}\id{x}{y}
-\end{align*}
-for all $i:\Gamma_0$ and, writing $H_{00}(i)$ for $\proj1 H_0(i)$ and
-$H_{01}(i)$ for $\lam{y}(\proj2 H_0(i))(y)$, we have $H_1(q)$ of type
-\begin{align*}
-& \msm{A}{\mprd{\ctxwk{A}{A}}{\idtypevar{A}}}_1(q,H_0(i),H_0(j)) \\
-& \jdeq \sm{e:A_1(q,H_{00}(i),H_{00}(j))}\mprd{\ctxwk{A}{A}}{\idtypevar{A}}_1(\pairr{q,e},H_{01}(i),H_{01}(j))\\
-& \jdeq \sm{e:A_1(q,H_{00}(i),H_{00}(j))}\prd*{x:A_0(i)}*{y:A_0(j)}{d:A_1(q,x,y)}(\idtypevar{A})_1(\pairr{q,e},d,H_{01}(x),H_{01}(y))\\
-& \jdeq \sm{e:A_1(q,H_{00}(i),H_{00}(j))}\prd*{x:A_0(i)}*{y:A_0(j)}{d:A_1(q,x,y)}\id{\trans{\pairr{H_{01}(x),H_{01}(y)}}{e}}{d}.
-\end{align*}
-By $H_0$, it follows that each $A_0(i)$ is contractible. By the contractibility
-of each $A_0(i)$, it follows that the type of $H_1(q)$ is equivalent to
-\begin{equation*}
-\sm{e:A_1(q,H_{00}(i),H_{00}(j))}\prd{d:A_1(q,H_{00}(i),H_{00}(j))}\id{e}{d}
-\end{equation*}
-which asserts that $A_1(q,H_{00}(i),H_{00}(j))$ is contractible. By 
-the contractibility of each $A_0(i)$, is is equivalent to the assertion
-that each $A_1(q,x,y)$ is contractible.
-\end{proof}
-
-\begin{rmk}
-We address the question whether it is the case that a graph $A$ in context
-$\Gamma$ is contractible if and only if $\terms{A}$ is contractible. As a
-consequence of \autoref{lem:contractible-graphs}, it is indeed the case
-that $\terms{A}$ is contractible whenever $A$ is. However, the converse
-does not hold.
-
-To see this, we first construct a counter example to the converse of the
-weak function extensionality principle, which states that there is a function
-of type
-\begin{equation}
-\iscontr(\prd{x:X}P(x))\to\prd{x:X}\iscontr(P(x))\label{eq:wfe-converse}
-\end{equation}
-for any type family $P:X\to\type$. In the proof of \autoref{thm:wfe-converse}, 
-we will find a family
-$P:X\to\type$ with the property that $\prd{x:X}P(x)$ is contractible and for
-which there is a term $x:A$ with $P(x)$ not contractible. Disproving the
-converse of the weak function extensionality principle suffices for our
-purposes, because if $P:X\to\type$ is a counter example to \autoref{eq:wfe-converse},
-then we can take $\Gamma\defeq\pairr{\Gamma_0,\Gamma_1}$ to be given by
-$\Gamma_0\defeq X$ and $\Gamma_1(i,j)\defeq\emptyt$ and we take
-$A\defeq\pairr{A_0,A_1}$ to be given by $A_0\defeq P$ and $A_1(q,u,v)
-\defeq\emptyt$.
-\end{rmk}
-
-We define the family $\mathcal{T}:\Sn^1\to\UU$ by
-We define $\mathcal{T}(\base)\defeq\mathbf{3}$. To define $\mathcal{T}(\lloop):
-\id{\mathbf{3}}{\mathbf{3}}$ we apply the univalence axiom. Hence it suffices to find an
-equivalence $\eqv{\mathbf{3}}{\mathbf{3}}$, for which we take the function
-$e$ defined by
-\begin{equation*}
-e(x)\defeq\begin{cases}
-0_\mathbf{3} & \text{if }x\jdeq 0_\mathbf{3}\\
-2_\mathbf{3} & \text{if }x\jdeq 1_\mathbf{3}\\
-1_\mathbf{3} & \text{if }x\jdeq 2_\mathbf{3}.
-\end{cases}
-\end{equation*}
-
-\begin{lem}
-The type $\terms{\mathcal{T}}\defeq\prd{x:\Sn^1}\mathcal{T}(x)$ is contractible.
-\end{lem}
-
-\begin{proof}
-The type of sections of $\mathcal{T}$ is equivalent to $\sm{u:\mathcal{T}(\base)}\id{e(u)}{u}$.
-If we have a term $\pairr{u,\alpha}$ of the latter type, it follows by induction
-on $\mathbf{3}$ that $\id{\pairr{u,\alpha}}
-{\pairr{0_\mathbf{3},\refl{0_\mathbf{3}}}}$ for all $\pairr{u,\alpha}:
-\sm{x:\mathcal{T}(\base)}\id{e(u)}{u}$,
-which shows that the type of sections of $\mathcal{T}$ is contractible.
-\end{proof}
-
-\begin{thm}\label{thm:wfe-converse}
-There is a type family $P:A\to\type$ for which
-\begin{equation*}
-\neg\Big(\iscontr(\prd{x:A}P(x))\to\prd{x:A}\iscontr(P(x))\Big).
-\end{equation*}
-\end{thm}
-
-\begin{proof}
-The type family of our counter example is $\mathcal{T}$: the fiber $\mathcal{T}(\base)$ isn't contractible.
-\end{proof}
-
-\begin{defn}
-A graph morphism $f:\ctxhom{\Delta}{\Gamma}$ is an equivalence of graphs when
-$\graphcharmap[\Gamma]{f}$ is a contractible graph in the context $\Gamma$.
-\end{defn}
-
-\begin{lem}
-Let $A$ and $B$ be graphs in a context $\Gamma$ and let $f:A\to B$. The following are equivalent:
-\begin{enumerate}
-\item $f[i]:A[i]\to B[i]$ is an equivalence of graphs for every term $i:\Gamma$.
-\item $\ctxext{\Gamma}f:\ctxext{\Gamma}A\to\ctxext{\Gamma}B$ is an equivalence of graphs.
-\item Both $f_0(i)$ and $f_1(q,x,y)$ are always equivalences.
-\item $\terms{f}:\terms{\Delta}\to\terms{\Gamma}$ is an equivalence.
-\end{enumerate}
-\end{lem}
-
-\begin{rmk}
-It follows that there is an equivalence
-\begin{equation*}
-\ctxext{\Gamma}\msm{A}{P}\simeq\ctxext({\Gamma}{A})P
-\end{equation*}
-for every family $P$ of graphs over a graph $A$ in context $\Gamma$.
-\end{rmk}
-
-\subsubsection{Homotopy levels}
-\begin{itemize}
-\item A graph $A$ in context $\Gamma$ is of homotopy level $n$ precisely when each
-$A_0(i)$ and each $A_1(q,x,y)$ are of homotopy level $n$. 
-\item We can name at least three different propositions in the empty context:
-\begin{enumerate}
-\item $\Gamma_0\defeq\emptyt$.
-\item $\Gamma_0\defeq\unit$ and $\Gamma_1(\ttt,\ttt)\defeq\emptyt$.
-\item $\Gamma_0\defeq\unit$ and $\Gamma_1(\ttt,\ttt)\defeq\unit$.
-\end{enumerate}
-Therefore $\mfGraph$ does not satisfy the law of excluded middle.
-\end{itemize}
-
-\subsubsection{Univalence for the graph model}
-In the other direction, we also obtain a graph morphism $\graphcharmap[\Gamma]{f}:
-\ctxhom{\Gamma}{\gobjclass}$ for every graph morphism $f:\ctxhom{\Delta}{\Gamma}$.
-In \autoref{graph-object-classifier} we will prove that the maps
-$\int_\Gamma:\ctxhom{\Gamma}{\gobjclass}\to\sm{\Delta:\tfGraph}\ctxhom{\Delta}{\Gamma}$
-is an equivalence with iverse $\graphcharmapfunc{\Gamma}$.
-
-\begin{defn}
-Let $f:\Delta\to\Gamma$ be a graph morphism. We define the graph morphism
-$\graphcharmap[\Gamma]{f}$ by
-\begin{align*}
-\graphcharmap[\Gamma]{f}_0 & \defeq \lam{i}\hfib{f_0}{i}\\
-\graphcharmap[\Gamma]{f}_1(i,j) & \defeq \lam{q}\hfib{f_1(i,j)}{q}.
-\end{align*}
-\end{defn}
-
-\begin{thm}\label{graph-object-classifier}
-Main theorem here.
-\end{thm}
-
-\begin{comment}
-To describe the object classifier for graphs, we will follow Streicher. Thus
-we have to look at presheaves over $I/i$ for each object $i$ of the category
-$I\defeq 0{\rightrightarrows}1$ with the morphisms named $s$ and $t$ for source
-and target. The category $I/0$ is the terminal category;
-the category $I/1$ looks like $\cdot{\rightarrow}\cdot{\leftarrow}\cdot$.
-Therefore, we have
-\begin{align*}
-\type^{\op{(I/0)}} & \eqv{}{\type},\\
-\type^{\op{(I/1)}} & \eqv{}{\sm{X,Y,A:\type}(A\to X)\times(A\to Y)}\\
-& \eqv{}{\sm{X,Y:\type}X\to Y\to\type}.
-\end{align*}
-The functors $\type^\op{\Sigma_s}$ and $\type^\op{\Sigma_t}$ are given by
-$\pi_1$ and $\pi_2$ respectively. This leads to our following definition
-of the object classifier $\gobjclass$:
-
-\begin{defn}
-Define $\gobjclass$ to be the graph consisting of
-\begin{align*}
-\gobjclass_0 & \defeq  \type\\
-\gobjclass_1(X,Y) & \defeq  X\to Y\to\type
-\end{align*}
-and define $\pointed{\gobjclass}$ by
-\begin{align*}
-(\pointed{\gobjclass})_0 & \defeq  \pointed{\type}\\
-(\pointed{\gobjclass})_1(\pairr{X,x},\pairr{Y,y}) & \defeq  \sm{R:X\to Y\to\type}R(x,y)
-\end{align*}
-There is the obvious forgetful graph morphism $t:\pointed{\gobjclass}\to\gobjclass$,
-given by projection on the first coordinate.
-
-For any morphism $f:\Delta\to\Gamma$ of graphs we define a morphism
-$\graphcharmap(f):\Gamma\to\gobjclass$ of graphs by
-\begin{align*}
-\graphcharmap(f)_0(i) & \defeq  \hfiber{f_0}{i}\\
-\graphcharmap(f)_1(q,\pairr{u,\alpha},\pairr{v,\beta}) & \defeq  \sm{p:\Delta_1(u,v)}
-\id{\trans{\pairr{\alpha,\beta}}{f_1(p)}}{q}
-\end{align*}
-where $\pairr{u,\alpha}:\graphcharmap(f)_0(i)$ and $\pairr{v,\beta}:\graphcharmap(f)_0(j)$. The
-morphism $\graphcharmap(f)$ is called the \emph{characteristic map of $f$}. We obtain a function
-\begin{equation*}
-\graphcharmap : \big(\sm{\Delta:\tfGraph }\Delta\to\Gamma\big)\to\big(\Gamma\to\gobjclass\big)
-\end{equation*}
-for every graph $\Gamma$.
-\end{defn}
-
-\begin{thm}\label{thm:graph-classifier1}
-The function $\graphcharmap$ is an equivalence for any graph $\Gamma$.
-\end{thm}
-
-\begin{proof}
-We have to find a quasi-inverse
-\begin{align*}
-\Sigma : (\Gamma\to\gobjclass)\to\big(\sm{\Delta:\tfGraph}\Delta\to\Gamma\big)
-\end{align*}
-of $\graphcharmap$. Thus, we have to define $\Sigma_0:(\Gamma\to\gobjclass)\to\tfGraph$ and
-$\Sigma_1:\prd{P:\Gamma\to\gobjclass}\Sigma_0(P)\to\Gamma$. For $P:\Gamma\to\gobjclass$ we define
-\begin{align*}
-\Sigma_0(P)_0 & \defeq \sm{i:\Gamma_0}P_0(i)\\
-\Sigma_0(P)_1(\pairr{i,u},\pairr{j,v}) & \defeq \sm{q:\Gamma_1(i,j)}P_1(q,u,v)\\
-\Sigma_1(P)_0 & \defeq \proj1\\
-\Sigma_1(P)_1(\pairr{i,u},\pairr{j,v}) & \defeq \proj1.
-\end{align*}
-\end{proof}
-
-\begin{thm}\label{conj:graph_classifier2}
-For any graph morphism $f:\Delta\to\Gamma$, the diagram
-\begin{equation*}
-\begin{tikzcd}
-\Delta \ar{r}{} \ar{d}[swap]{f} & \pointed{\gobjclass} \ar{d}{t} \\ 
-\Gamma \ar{r}[swap]{\graphcharmap(f)} & \gobjclass 
-\end{tikzcd}
-\end{equation*}
-is a pullback square.
-\end{thm}
-\note{We would like it to be a pb \emph{in} the graph model}
-\end{comment}
-
-\begingroup\color{blue}
-\subsubsection{The adjunctions $\tfcolim\dashv\Delta\dashv\tflim$}
-We define $\Delta:\type\to\tfGraph$ by
-\begin{equation*}
-\Delta(X)\defeq\pairr{X,\lam{x}{x'}\id{x}{x'}}
-\end{equation*}
-for $X:\type$. For $A:X\to\type$ we define $\Delta(A):\mftyp(\Delta(X))$ by
-\begin{align*}
-\Delta(A)_0(x) & \defeq A(x)\\
-\Delta(A)_1(p,a,b) & \defeq \id{\trans{p}{a}}{b}.
-\end{align*}
-
-\begin{lem}
-For any type $X$ and any graph $\Gamma$ there is an equivalence
-\begin{equation*}
-\eqv{(X\to\terms{\Gamma})}{\terms{\Delta(X)\to\Gamma}}.
-\end{equation*}
-\end{lem}
-
-\begin{proof}
-We have to find functions
-\begin{align*}
-\varphi & : (X\to\terms{\Gamma})\to\terms{\Delta(X)\to\Gamma}\\
-\psi & : \terms{\Delta(X)\to\Gamma}\to X\to\terms{\Gamma}
-\end{align*}
-which are each others homotopy inverse. To define $\varphi$, let
-\end{proof}
-\endgroup

### 2013-09-27 article-models-eg.tex
resolve issue #7

\section{Empirical evidence of what an internal model should be}

\subsection{Univalent universes as internal models}

Before we give the definition we illustrate the concepts that go into it in
the case of a univalent universe $\UU$.
Regardless of the definition of an internal model, a univalent universe should
be an example of one.

The type of contexts is $\UU$ itself and for every $\Gamma:\UU$, a type in
context $\Gamma$ is simply a family $A:\Gamma\to\UU$. The type $\terms{A}$ of terms of a type $A$
in context $\Gamma$ is defined to be $\prd{i:\Gamma}A(i)$. Note that $\Gamma\to
\UU$ itself also interprets type theory where a type in context $A:\Gamma\to\UU$
is a family $P:(\sm{i:\Gamma}A(i))\to\UU$. We may denote this model by
$\mftyp{\UU}{\Gamma}$. The terms of a type $P$ in context $A$ in the model
$\Gamma\to\UU$ are the terms of $\prd{i:\Gamma}{x:A(i)}P(i,x)$.

When $A$ is a type in context
$\Gamma$ we define the context extension $\ctxext{\Gamma}A$ to be
$\sm{i:\Gamma}A(i)$. Note that context extension $\ctxext{\Gamma}\blank$
can be seen acting not only as a function from the context of the model
$\Gamma\to\UU$ to the context of the model $\UU$, but it also acts on the types 
and terms: when $A$ is a type in context $\Gamma$ we may take the identity map
from $\mftyp{\Gamma\to\UU}{A}\to\mftyp{\UU}{\ctxext{\Gamma}A}$ because
$\mftyp{\Gamma\to\UU}{A}$ is taken to be $\mftyp{\UU}{\ctxext{\Gamma}A}$;
thus, $\ctxext{\Gamma}P\defeq P$ for every family $P$ over $A$ in context 
$\Gamma$. Likewise, when $P$ is a type in context $A$ in the model 
$\Gamma\to\UU$ then context extension should act on the terms of $P$ via a
function $\terms{P}\to\terms{\ctxext{\Gamma}P}$, which we take to be
the identity map once more.

Since $\mftyp{\UU}{\Gamma}$ is a model of type theory it
has it's own notion of context extension: when $P$ is a type in context $A$ in
the model $\mftyp{\UU}{\Gamma}$ then $\ctxext{A}P$ is the family
$\lam{i}\sm{x:A(i)}P(i,x):\Gamma\to\UU$. Also the context extension of
$\Gamma\to\UU$ acts trivially on the types and terms. Context extension is
analoguous to the Grothendieck construction that associates the category of
elements to a presheaf and it gives us $\Sigma$-types.

When $A$ and $B$ are types in context
$\Gamma$, the weakening $\ctxwk{A}{B}$ of $B$ along $A$ is defined to be
$\lam{\pairr{i,x}}B(i):(\sm{i:\Gamma}A(i))\to\UU$. Weakening along a type $A$ 
in context $\Gamma$ also acts on types and terms. When $Q$ is a type in context
$B$ in the model $\Gamma\to\UU$, we define $\ctxwk{A}{Q}$ to be 
$\lam{\pairr{i,x}}{y}Q(i,y)$. When $g:\terms{Q}$ we define $\ctxwk{A}{g}$ to
be $\lam{\pairr{i,x}}{y}g(i,y)$.

Note that for two types $A$ and $B$ in context $\Gamma$, the terms of
$\ctxwk{A}{B}$ are the terms of $\prd{i:\Gamma}A(i)\to B(i)$, i.e.~they are
the fiberwise maps from $A$ to $B$. We shall take the type $\terms{\ctxwk{A}{B}}$
of terms of $\ctxwk{A}{B}$ to be the type of morphisms from $A$ to $B$. Also,
any context of $\UU$ may be seen as a type in the empty context $\unit$. Thus
the type of terms of a context $\Gamma$ is $\unit\to\Gamma$, which is
equivalent to $\Gamma$. A context morphism from $\Delta$ to $\Gamma$ is a term
of $\ctxwk{\Gamma}{\Delta}$, i.e.~a function from $\Gamma$ to $\Delta$. We denote
the type $\terms{\ctxwk{\Gamma}{\Delta}}$ by $\ctxhom{\Delta}{\Gamma}$.

When $P$ is a family over
$A$ in context $\Gamma$ and $x$ is a term of $A$ we define the type $\subst{x}{P}$
in context $\Gamma$ to be $\lam{i}P(i,x(i))$. Like extension

We will interpret the dependent product $\mprd{A}{P}$ of a family $P$ over
$A$ in context $\Gamma$ by
\begin{equation*}
\mprd{A}{P}(i)\defeq \prd{x:A(i)}P(i,x)
\end{equation*}
With this interpretation there is an equivalence 
$\lambda:\eqv{\terms{P}}{\terms{\mprd{A}{P}}}$. This is lambda-abstraction; its
inverse being evaluation. We should note however, that the rule for evaluation
we interpret here is
\begin{equation*}
\inference{\Gamma\vdash f:\mprd{A}{P}}{\ctxext{\Gamma}A\vdash\tfev(f):P}
\end{equation*}
which is different than the usual rule
\begin{equation*}
\inference{\Gamma\vdash f:\mprd{A}{P} \qquad \Gamma\vdash a:A}{\Gamma \vdash \tfev(f,a):\subst{a}{P}}
\end{equation*}
The reason for this is that the interpretation of the usual rule would give a
function of type $\terms{\mprd{A}{P}}\to\prd{x:\terms{A}}\terms{\subst{x}{P}}$,
but this does not describe the terms of $\mprd{A}{P}$ by any means. Moreover,
since we have implemented substitution, we obtain from $\ctxext{\Gamma}A\vdash\tfev(f):P$ 
and $\Gamma\vdash x:A$ a term $\Gamma\vdash\subst{x}{\tfev(f)}:\subst{x}{P}$ and
therefore we do not loose anything with this approach.

{\color{red} point to example}

\subsection{The Sierpinsky model}.
In the Sierpinsky model, contexts are interpreted by dependent types $A\to\type$.

\subsection{The graph model of type theory}
In this section we define the graph model of type theory, denoted by
$\mfGraph$.  In our presentation, we follow that of the definition internal
models. After we have established the graph model, we show how the graph
model can be seen as a presheaf model, namely over the category $\cdot
{\rightrightarrows}\cdot$.

\begin{defn}
A \emph{(directed) graph} $\Gamma$ is a pair $\pairr{\Gamma_0,\Gamma_1}$ 
consisting of a type $\Gamma_0$ of vertices and a family 
$\Gamma_1:\Gamma_0\to\Gamma_0\to\type$ of edges. The type $\ctx(\mfGraph)$
is defined to be the type of all graphs; we will usually denote it by
$\tfGraph$. Explicitly, we have
\begin{equation*}
\tfGraph\defeq\sm{\Gamma_0:\type}\Gamma_0\to\Gamma_0\to\type.
\end{equation*}
\end{defn}

\begin{eg}\label{ex:pb}
The underlying graph of the diagram
\begin{equation*}
\begin{tikzcd}
{} & A \ar{d}{f} \\
B \ar{r}[swap]{g} & C
\end{tikzcd}
\begin{comment}
\begin{tikzpicture}
\matrix (m) [std] { & A \\ B & C \\};
\draw[ar] (m-1-2) -- node[right] {$f$} (m-2-2);
\draw[ar] (m-2-1) -- node[below] {$g$} (m-2-2);
\end{tikzpicture}
\end{comment}
\end{equation*}
is $I\defeq\mathbf{3}$ and has $J(1,3)\defeq J(2,3)\defeq \unit $ 
and $J(x,y)\defeq\emptyt$ otherwise (it is defined using the induction 
principle of $\mathbf{3}$ and a universe).
\end{eg}

\begin{defn}
A \emph{graph $A$ in the context $\Gamma$} is a pair $\pairr{A_0,A_1}$ consisting
of 
\begin{align*}
A_0 & :\Gamma_0\to\type\\
A_1 & :\prd*{i,j:\Gamma_0}\Gamma_1(i,j)\to A_0(i)\to A_0(j)\to\type.
\end{align*}
Thus, for a graph $\Gamma$, the type $\mftyp{\mfGraph}{\Gamma}$ is the type
\begin{equation*}
\sm{A_0:\Gamma_0\to\type}\prd*{i,j:\Gamma_0}\Gamma_1(i,j)\to A_0(i)\to A_0(j)\to\type.
\end{equation*}
We will also write $\Gamma\vdash A:\mfGraph$ when $A$ is a graph in context
$\Gamma$.
\end{defn}

\begin{defn}
Suppose that $\Gamma\vdash A:\mfGraph$. Then we define the graph $\ctxext{\Gamma}A$
by
\begin{align*}
\ctxext{\Gamma}A_0 & \defeq \sm{i:\Gamma_0}A_0(i)\\
\ctxext{\Gamma}A_1(\pairr{i,x},\pairr{j,y}) & \defeq \sm{q:\Gamma_1(i,j)}A_1(q,x,y).
\end{align*}
\end{defn}

\begin{defn}
Suppose that $\Gamma\vdash A:\mfGraph$ and $\Gamma\vdash B:\mfGraph$. Then we
define the graph $\ctxwk{A}{B}$ in context $\ctxext{\Gamma}A$ by
\begin{align*}
(\ctxwk{A}{B})_0(\pairr{i,x}) & \defeq B_0(i)\\
(\ctxwk{A}{B})_1(\pairr{q,e},u,v) & \defeq B_1(q,u,v).
\end{align*}
\end{defn}

\begin{defn}
Suppose that $\Gamma\vdash A:\mfGraph$. A term $x$ of $A$ consists of a pair
$\pairr{x_0,x_1}$ where
\begin{align*}
x_0 & : \prd{i:\Gamma_0}A_0(i)\\
x_1 & : \prd*{i,j:\Gamma_0}{q:\Gamma_1}A_1(q,x_0(i),x_0(j)).
\end{align*}
Thus we define
\begin{equation*}
\terms{A}\defeq\sm{x_0:\prd{i:\Gamma_0}A_0(i)}\prd*{i,j:\Gamma_0}{q:\Gamma_1}A_1(q,x_0(i),x_0(j)).
\end{equation*}
We also write $\Gamma\vdash x:A$ when $x$ is a term of the graph $A$ in context
$\Gamma$.
\end{defn}

\begin{defn}
Suppose that $P$ is a family of graphs over $A$ in context $\Gamma$ and let
$x$ be a term of $A$. Then we define the graph $\subst{x}{P}$ 
in context $\Gamma$ by
\begin{align*}
\subst{x}{P}_0(i) & \defeq P_0(\pairr{i,x_0(i)})\\
\subst{x}{P}_1(q,u,v) & \defeq P_1(\pairr{q,x_1(q)},u,v)
\end{align*}
\end{defn}

\begin{rmk}
Note that we have the judgmental equality $\subst{x}{\ctxwk{A}{B}}\jdeq B$
for every two graphs $A$ and $B$ in context $\Gamma$ and every term $x$ of $A$.
\end{rmk}

\begin{defn}
The terminal graph $\unit^\mfGraph$, which we shall often denote simply 
by $\unit$, is defined by
\begin{align*}
{\unit^\mfGraph}_0 & \defeq \unit\\
{\unit^\mfGraph}_1(x,y) & \defeq \unit.
\end{align*}
\end{defn}

\begin{rmk}
We note that the function $\ctxext{\unit}\blank$ is an equivalence
from $\mftyp{\mfGraph}{\unit}$ to $\tfGraph$. It's inverse is the function which maps
a graph $\Gamma$ to the pair $\pairr{A_0,A_1}$ where $A_0$ is defined by
$A_0(\ttt)\defeq\Gamma_0$ and where $A_1$ is defined by $A_1(\ttt)\defeq
\Gamma_1$.

Note that we only have an equivalence here, not a judgmental equality.
\end{rmk}

\subsubsection{The contexts of $\mfGraph$}

The category $\psh(\cdot{\rightrightarrows}\cdot)$ of presheaves over
$\cdot{\rightrightarrows}\cdot$ is given by
\begin{equation*}
\sm{A_0,A_1:\type}(A_1\to A_0)\times(A_1\to A_0).
\end{equation*}
To see that $\psh(\cdot{\rightrightarrows}\cdot)$ is indeed equivalent to
the type $\tfGraph$ of all graphs, note that we have the equivalences
\begin{align*}
\psh(\cdot{\rightrightarrows}\cdot) & \eqvsym \sm{A_0,A_1:\type}A_1\to A_0\times A_0\\
& \eqvsym \sm{A_0:\type} A_0\times A_0\to\type\\
& \eqvsym \tfGraph.
\end{align*}

\begin{defn}
Suppose that $\Delta\defeq \pairr{\Delta_0,\Delta_1}$ and $\Gamma\defeq \pairr{\Gamma_0,
\Gamma_1}$ are graphs. A morphism $f$ from $\Delta$ to $\Gamma$ is a pair
$\pairr{f_0,f_1}$ consisting of
\begin{align*}
f_0 & : \Delta_0\to\Gamma_0\\
f_1 & : \prd*{u,v:\Delta_0}\Delta_1(u,v)\to\Gamma_1(f_0(u),f_0(v))
\end{align*}
We write $\ctxhom{\Delta}{\Gamma}$ for the type of morphisms of graphs.
\end{defn}

The terminal graph $\unit^\mfGraph$ plays the role of the empty context.

\begin{defn}
The terminal graph $\unit^\mfGraph$, which we shall often denote simply 
by $\unit$, is defined by
\begin{align*}
{\unit^\mfGraph}_0 & \defeq \unit\\
{\unit^\mfGraph}_1(x,y) & \defeq \unit.
\end{align*}
\end{defn}

It is immediate that the type of graph morphisms $\Gamma\to\unit^\mfGraph$ is
contractible for every graph $\Gamma$. The terms of a graph $\Gamma$ are the
global elements of $\Gamma$, i.e.~they are the morphisms of type 
$\ctxhom{\unit^\mfGraph}{\Gamma}$.

\begin{defn}
Let $\Gamma$ be a graph. The type $\terms[\mfGraph]{\Gamma}$ of terms of $\Gamma$ is
defined to be
\begin{equation*}
\sm{x_0:\Gamma_0}\Gamma_1(x_0,x_0).
\end{equation*}
\end{defn}

We now turn to the description of the universe $\gobjclass$ of graphs. Note
that for the category $I\defeq 0{\rightrightarrows}1$, we have
$I/0\defeq\catid{0}$ and $I/1\defeq s\rightarrow\catid{1}\leftarrow t$, where
$s$ and $t$ are the morphisms $0\to 1$. Thus we have 
\begin{align*}
\psh(I/0) & \eqvsym \type\\
\psh(I/1) & \eqvsym \sm{X,Y:\type}X\to Y\to\type.
\end{align*}
The functors $\psh(\Sigma_s)$ and $\psh(\Sigma_t)$ map a presheaf 
$\pairr{X,Y,R}$ to $X$ and $Y$, respectively. 
Thus we obtain the following graph $\gobjclass$.

\begin{defn}
The universe $\gobjclass$ of graphs is defined to be
\begin{align*}
{\gobjclass}_0 & \defeq \type\\
{\gobjclass}_1(X,Y) & \defeq X\to Y\to\type.
\end{align*}
\end{defn}

Note that the type $\terms{\gobjclass}$ of terms of $\gobjclass$ is 
exactly $\tfGraph$.

\subsubsection{Families of graphs}
Using $\gobjclass$, we obtain a description of what
it means to be a graph in a context $\Gamma$, where $\Gamma$ is itself a
graph. The notion of ``a graph in a context $\Gamma$'' is the interpretation
of ``a type in a context $\Gamma$''.

\begin{defn}
A \emph{graph $A$ in the context $\Gamma$} is a pair $\pairr{A_0,A_1}$ consisting
of 
\begin{align*}
A_0 & :\Gamma_0\to\type\\
A_1 & :\prd*{i,j:\Gamma_0}\Gamma_1(i,j)\to A_0(i)\to A_0(j)\to\type.
\end{align*}
We will also write $\Gamma\vdash A:\mfGraph$ when $A$ is a graph in context
$\Gamma$.
\end{defn}

\begin{rmk}
Note that every graph can be seen as a graph in context $\unit^\mfGraph$.
\end{rmk}

\begin{defn}
Let $A$ be a graph in context $\Gamma$ and let $f:\ctxhom{\Delta}{\Gamma}$
be a graph morphism. We define the graph $A[f]$ in context $\Delta$ by
\begin{align*}
A[f]_0(u) & \defeq A_0(f_0(u))\\
A[f]_1(p,x,y) & \defeq A_1(f_1(p),x,y).
\end{align*}
\end{defn}

\begin{defn}
Let $\Gamma\vdash A:\mfGraph$ and $\Gamma\vdash B:\mfGraph$ be two graphs in
the context $\Gamma$. A morphism $f$ from $A$ to $B$ is a pair $\pairr{f_0,f_1}$
consisting of
\begin{align*}
f_0 & : \prd*{i:\Gamma_0} A_0(i)\to B_0(i)\\
f_1 & : \prd*{i,j:\Gamma_0}{q:\Gamma_0}*{x:A_0(i)}*{y:A_0{j}}A_1(q,x,y)\to B_1(q,f_0(x),f_0(y)).
\end{align*}
We write $\ctxhom[\Gamma]{A}{B}$ for the type of morphisms between graphs in
context $\Gamma$.
\end{defn}

\begin{defn}
The unit graph $\unit^\Gamma$ in context $\Gamma$ consists of
\begin{align*}
(\unit^\Gamma)_0 & \defeq \lam{i}\unit\\
(\unit^\Gamma)_1(q) & \defeq \lam{x}{y}\unit.
\end{align*}
\end{defn}

As was the case with graphs, terms of graphs in a context are described by
morphisms from $\unit$. Explicating this in more detail, we get

\begin{defn}
Let $A$ be a graph in context $\Gamma$. A term $x$ of $A$ is a pair
$\pairr{x_0,x_1}$ consisting of
\begin{align*}
x_0 & : \prd{i:\Gamma_0}A_0(i)\\
x_1 & : \prd*{i,j:\Gamma_0}{q:\Gamma_1(i,j)}A_1(q,x_0(i),x_0(j)).
\end{align*}
We define $\terms[\mfGraph]{A}$ to be the type of terms of $A$.
\end{defn}

\begin{rmk}
Note that the type of terms of a graph $\Gamma$ is equivalent to the type
of terms of the graph $\Gamma$ in context $\unit^\mfGraph$.
\end{rmk}

\begin{defn}
Suppose that $\Gamma\vdash A:\mfGraph$. Then we define the graph $\ctxext{\Gamma}A$
by
\begin{align*}
(\ctxext{\Gamma}A)_0 & \defeq \sm{i:\Gamma_0}A_0(i)\\
(\ctxext{\Gamma}A)_1(\pairr{i,x},\pairr{j,y}) & \defeq \sm{q:\Gamma_1(i,j)}A_1(q,x,y).
\end{align*}
There is a morphism $\pi^A:\ctxhom{\ctxext{\Gamma}A,\Gamma}$ given by
\begin{align*}
\pi^A_0 & \defeq \lam{\pairr{i,x}}i && : (\sm{i:\Gamma_0}A_0(i))\to \Gamma_0\\
\pi^A_1(\pairr{i,x},\pairr{j,y}) & \defeq \lam{\pairr{q,e}}q && : (\sm{q:\Gamma_1(i,j)}A_1(q,x,y))\to\Gamma_1(i,j).
\end{align*}
\end{defn}

\begin{rmk}
The construction of $\ctxext{\Gamma}A$ is similar to the Grothendieck
construction of the category of elements for a presheaf.
\end{rmk}

\begin{defn}
Let $\Gamma\vdash A:\mfGraph$. A family $P$ of graphs over $A$ is by
definition a graph $P$ in the context $\ctxext{\Gamma}A$. More explicitly, a family
$P$ of graphs over $A$ in context $\Gamma$ consists of
\begin{align*}
P_0 & : \prd*{i:\Gamma_0}A_0(i)\to\type\\
P_1 & : \prd*{i,j:\Gamma_0}{q:\Gamma_1(i,j)}*{x:A_0(i)}*{y:A_0(j)} A_1(q,x,y)\to P_0(x)\to P_0(y)\to\type
%P_1 & : \prd{\pairr{i,x},\pairr{j,y}:(\ctxext{\Gamma}A)_0}(\ctxext{\Gamma}A)_1(\pairr{i,x},\pairr{j,y})\to P_0(i,x)\to P_0(j,y)\to\type.
\end{align*}
\end{defn}

\begin{defn}
When $P$ is a family of graphs over $A$ in context $\Gamma$ and $x:A$, we
get a graph $\subst{x}{P}$ in context $\Gamma$ given by
\begin{align*}
\subst{x}{P}_0(i) & \defeq P_0(x_0(i))\\
\subst{x}{P}_1(q,u,v) & \defeq P_1(\pairr{q,x_1(q)},u,v)
\end{align*}
\end{defn}

\subsubsection{The basic type constructors for graphs}
With families of graphs being available, we can give the interpretations of
dependent products, dependent sums and identity types. In the following, we
shall introduce the graph interpretations of dependent products, dependent
sums and identity types and describe the terms of the resulting graphs.\note{These definitions
should be connected to Mike's article but I don't really know how to do this}

\begin{defn}
Let $P$ be a family of graphs over $A$, where $\Gamma\vdash A:\mfGraph$. 
The dependent function graph $\mprd{A}{P}$ in context $\Gamma$ consists of
\begin{align*}
\mprd{A}{P}_0(i) & \defeq \prd{x:A_0(i)}P_0(x)\\
\mprd{A}{P}_1(q,f,g) & \defeq \prd*{x:A_0(i)}*{y:A_0(j)}{e:A_1(q,x,y)}P_1(\pairr{q,e},f(x),g(y)).
\end{align*}
\end{defn}

\begin{rmk}
A term $f:\mprd{A}{P}$ consists of
\begin{align*}
f_0 & : \prd*{i:\Gamma_0}{x:A_0(i)}P_0(x)\\
f_1 & : \prd*{i,j:\Gamma_0}{q:\Gamma_1(i,j)}*{x:A_0(i)}*{y:A_0(j)}{e:A_1(q,x,y)}P_1(\pairr{q,e},f_0(x),f_0(y))
\end{align*}
Therefore, we see that $\eqv{\terms{\mprd{A}{P}}}{\terms{P}}$. \emph{Warning:} it
is by no means the case that $\eqv{\terms{\mprd{A}{P}}}{\prd{x:\terms{A}}
\terms{\subst{x}{P}}}$ for all families $P$ over $A$. For instance, the graph
$\tilde{\emptyt}$ defined by $\tilde{\emptyt}_0\defeq\unit$ and 
$\tilde{\emptyt}_1(\ttt,\ttt)\defeq\emptyt$ has no terms and neither does
$\tilde{\emptyt}+\tilde{\emptyt}$. Nevertheless, there are two graph morphisms
from $\tilde{\emptyt}$ to $\tilde{\emptyt}+\tilde{\emptyt}$. More generally,
when $\Gamma$ is a graph such that $\Gamma_1(i,j)\jdeq\emptyt$ for all $i,j:\Gamma_0$,
then $\eqv{{\terms{\Gamma\to\Gamma'}}}{\Gamma_0\to\Gamma^\prime_0}$.
\end{rmk}

\begin{defn}
If $P$ is a family of graphs over $\Gamma$, the dependent pair graph
$\msm{A}{P}$ consists of
\begin{align*}
\msm{A}{P}_0(i) & \defeq \sm{x:A_0(i)}P_0(x)\\
\msm{A}{P}_1(q,\pairr{x,u},\pairr{y,v}) & \defeq \sm{e:A_1(q,x,y)}P_1(\pairr{q,e},u,v).
\end{align*}
\end{defn}

\begin{rmk}
A term $w:\msm{A}{P}$ consists of
\begin{align*}
w_0 & : \prd{i:\Gamma_0}\sm{x:A_0(i)}P_0(x)
\intertext{and, writing $\lam{i}\proj1(w_0(i))$ and $\lam{i}\proj2(w_0(i))$ as
$w_{00}$ and $w_{01}$ respectively,}
w_1 & : \prd*{i,j:\Gamma_0}{q:\Gamma_1(i,j)}\sm{e:A_1(q,w_{00}(i),w_{00}(j)}P_1(\pairr{q,e},w_{01}(i),w_{01}(j)).
\end{align*}
By $\choice{\infty}$ it follows that
\begin{equation*}
\eqv{\terms{\msm{A}{P}}}{\sm{x:\terms{A}}\terms{\subst{x}{P}}}.
\end{equation*}
\end{rmk}

\begin{rmk}
For any two graphs $A$ and $B$ in a context $\Gamma$, we have the graph
$B[\pi^A]$ in the context $\ctxext{\Gamma}A$, which we may just write
as $B$. Therefore, we can
consider the graphs $\mprd{A}{B}$ and $\msm{A}{B}$. As usually those shall be
denoted  by $A\to B$ and $A\times B$, respectively.

A term $f$ of $A\to B$ consists of
\begin{align*}
f_0 & : \prd*{i:\Gamma_0}A_0(i)\to B_0(i)\\
f_1 & : \prd*{i,j:\Gamma_0}{q:\Gamma_1(i,j)}*{x:A_0(i)}*{y:A_0(j)}A_1(q,x,y)\to B_1(q,f_0(x),f_0(y)),
\end{align*}
so we see that $\terms{A\to B}\jdeq\ctxhom[\Gamma]{A}{B}\jdeq\terms{B[\pi^A]}$.

A term $\pairr{x,y}$ of $A\times B$ consists of
\begin{align*}
\pairr{x,y}_0 & : \prd*{i:\Gamma_0} A_0(i)\times B_0(i)\\
\pairr{x,y}_1 & : \prd*{i,j:\Gamma_0}{q:\Gamma_1(i,j)} A_1(q,x_0(i),x_0(j))\times B_1(q,y_0(i),y_0(j))
\end{align*}
so we see that $\eqv{\terms{A\times B}}{\terms{A}\times\terms{B}}$.

Note that we also obtain
a \emph{graph} $\Delta\to\Gamma$ this way, for every two graphs $\Delta$ and
$\Gamma$ seen as graphs in the context $\unit^\mfGraph$. 
The type $\terms{\Delta\to\Gamma}$ of terms of $\Delta\to\Gamma$ is just
$\ctxhom{\Delta}{\Gamma}$.
\end{rmk}

\begin{defn}
For any graph $\Gamma$ there is a graph $\tffam{\Gamma}$ in context 
$\gobjclass$ defined by
\begin{align*}
(\tffam{\Gamma})_0 & \defeq \lam{X}X\to\Gamma_0\\
(\tffam{\Gamma})_1(X,Y,R) & \defeq \lam{f}{g}\prd*{x:X}*{y:Y}R(x,y)\to\Gamma_1(f(x),g(y)).
\end{align*}
\end{defn}

\begin{rmk}
A term $D$ of $\msm{\gobjclass}{\tffam{\Gamma}}$ consists of a term
$\pairr{\Delta_0,f_0}$ of type
\begin{equation*}
\msm{\gobjclass}{\tffam{\Gamma}}_0\jdeq \sm{\Delta_0:\type}\Delta_0\to\Gamma_0
\end{equation*}
and a term $\pairr{\Delta_1,f_1}$ of type
\begin{equation*}
\msm{\gobjclass}{\tffam{\Gamma}}_1(q)\jdeq \sm{\Delta_1}
\end{equation*}
\end{rmk}

\begin{defn}
There is a graph morphism
\begin{equation*}
\graphcharmapfunc{\Gamma} : \sm{\gobjclass}
\end{equation*}
\end{defn}

\begin{defn}
Let $A$ be a graph in context $\Gamma$. We define the family $\idtypevar{A}$ over $\ctxwk{A}{A}$ in
context $\ctxext{\Gamma}A$ by
\begin{align*}
(\idtypevar{A})_0(\pairr{i,x},y) & \defeq \id{x}{y}\\
(\idtypevar{A}){}_1(\pairr{q,e},d,\alpha,\alpha') & \defeq \id{\trans{\pairr{\alpha,\alpha'}}{e}}{d}
\end{align*}
where $q:\Gamma_1(i,j)$, $e:A_1(q,x,x')$, $d:A_1(q,y,y')$, $\alpha:\id{x}{y}$
and $\alpha':\id{x'}{y'}$. The transportation along the path 
$\pairr{\alpha,\alpha'}:\id{\pairr{x,x'}}{\pairr{y,y'}}$ in $A_0(i)\times A_0(j)$
is taken with respect to the family $\lam{x}{x'}A_1(q,x,x')$.

We define the term $\refl{A}$ of the family 
$\subst{\idfunc[A]}{\idtypevar{A}}$ over $A$ in context $\Gamma$ by
\begin{align*}
(\reflf{A})_0(i) & \defeq \lam{x}\refl{x}\\
(\reflf{A})_1(q) & \defeq \lam{e}\refl{e}
\end{align*}
\end{defn}

\begin{defn}
Let $D$ be a family over $\idtypevar{A}$ in context 
$\ctxext({\Gamma}{A})\ctxwk{A}{A}$. Then we have the family
$\subst{\idfunc[A]}{D}$ over $\subst{\idfunc[A]}{\idtypevar{A}}$ in
context $\ctxext{\Gamma}A$ given by
\begin{align*}
\subst{\idfunc[A]}{D}_0(\pairr{i,x},\alpha) & \defeq D_0(\pairr{i,x,x},\alpha)\\
\subst{\idfunc[A]}{D}_1(\pairr{q,e},\gamma) & \defeq D_1(\pairr{q,e,e},\gamma).
\end{align*}
The family $\subst{\reflf{A}}{\subst{\idfunc[A]}{D}}$ over $A$ in
context $\Gamma$ is given by
\begin{align*}
\subst{\reflf{A}}{\subst{\idfunc[A]}{D}}_0(i,x) & \defeq D_0(i,x,x,\refl{x})\\
\subst{\reflf{A}}{\subst{\idfunc[A]}{D}}_1(q,e) & \defeq D_1(q,e,e,\refl{e}).
\end{align*}
To show that the identity graphs correctly interpret the identity elimination
rule, we must give a function
\begin{equation*}
\tfJ : \terms{\subst{\reflf{A}}{\subst{\idfunc[A]}{D}}}\to\terms{D}.
\end{equation*}
Note that a term $d$ of $\subst{\reflf{A}}{\subst{\idfunc[A]}{D}}$
consists of
\begin{align*}
d_0 & : \prd*{i:\Gamma_0}{x:A_0(i)}D_0(i,x,x,\refl{x})\\
d_1 & : \prd*{i,j:\Gamma_0}{q:\Gamma_1(i,j)}*{x:A_0(i)}*{y:A_0(j)}{e:A_1(q,x,y)}D_1(q,e,e,\refl{e})
\end{align*}
A simple argument using path induction reveals that terms of
$\subst{\reflf{A}}{\subst{\idfunc[A]}{D}}$ indeed yield terms of $D$. 
\end{defn}

\begin{rmk}
Using the identity graph $\idtypevar{A}$ we can describe the identity
graph $\id[A]{x}{y}$ in context $\Gamma$ for any two terms $x,y:A$. 
The graph $\id[A]{x}{y}$ in context $\Gamma$ consists of
\begin{align*}
(\id[A]{x}{y})_0(i) & \defeq \id{x_0(i)}{y_0(i)}\\
(\id[A]{x}{y})_1(q,\alpha,\beta) & \defeq \id{\trans{\pairr{\alpha,\beta}}{x_1(q)}}{y_1(q)}
\end{align*}
From this, we see that a term of $p:\id[A]{x}{y}$ consists of
\begin{align*}
p_0 & : \prd{i:\Gamma_0}\id{x_0(i)}{y_0(i)}\\
p_1 & : \prd*{i,j:\Gamma_0}{q:\Gamma_1(i,j)}\id{\trans{\pairr{p_0(i),p_0(j)}}{x_1(q)}}{y_1(q)}.
\end{align*}
\end{rmk}

\subsubsection{Properties of graphs}

\subsubsection{Contractibility and equivalences of graphs}

\begin{defn}
A graph $A$ in context $\Gamma$ is said to be \emph{contractible} if there
is a term of the graph
\begin{equation*}
\msm{A}{\mprd{\ctxwk{A}{A}}{\idtypevar{A}}}
\end{equation*}
in context $\Gamma$.
\end{defn}

\begin{lem}\label{lem:contractible-graphs}
Let $A$ be a graph in context $\Gamma$. The following are equivalent:
\begin{enumerate}
\item $A$ is a contractible graph.
\item Both $A_0(i)$ and $A_1(q,x,y)$ are always contractible.
\end{enumerate}
\end{lem}

\begin{proof}
Let $H:\msm{A}{\mprd{\ctxwk{A}{A}}{\idtypevar{A}}}$. Unfolding the definitions, we have
an element $H_0(i)$ of type
\begin{align*}
\msm{A}{\mprd{\ctxwk{A}{A}}{\idtypevar{A}}}_0(i) & \jdeq \sm{x:A_0(i)}\mprd{\ctxwk{A}{A}}{\idtypevar{A}}_0(i,x)\\
& \jdeq \sm{x:A_0(i)}\prd{y:A_0(i)}(\idtypevar{A})_0(\pairr{i,x},y)\\
& \jdeq \sm{x:A_0(i)}\prd{y:A_0(i)}\id{x}{y}
\end{align*}
for all $i:\Gamma_0$ and, writing $H_{00}(i)$ for $\proj1 H_0(i)$ and
$H_{01}(i)$ for $\lam{y}(\proj2 H_0(i))(y)$, we have $H_1(q)$ of type
\begin{align*}
& \msm{A}{\mprd{\ctxwk{A}{A}}{\idtypevar{A}}}_1(q,H_0(i),H_0(j)) \\
& \jdeq \sm{e:A_1(q,H_{00}(i),H_{00}(j))}\mprd{\ctxwk{A}{A}}{\idtypevar{A}}_1(\pairr{q,e},H_{01}(i),H_{01}(j))\\
& \jdeq \sm{e:A_1(q,H_{00}(i),H_{00}(j))}\prd*{x:A_0(i)}*{y:A_0(j)}{d:A_1(q,x,y)}(\idtypevar{A})_1(\pairr{q,e},d,H_{01}(x),H_{01}(y))\\
& \jdeq \sm{e:A_1(q,H_{00}(i),H_{00}(j))}\prd*{x:A_0(i)}*{y:A_0(j)}{d:A_1(q,x,y)}\id{\trans{\pairr{H_{01}(x),H_{01}(y)}}{e}}{d}.
\end{align*}
By $H_0$, it follows that each $A_0(i)$ is contractible. By the contractibility
of each $A_0(i)$, it follows that the type of $H_1(q)$ is equivalent to
\begin{equation*}
\sm{e:A_1(q,H_{00}(i),H_{00}(j))}\prd{d:A_1(q,H_{00}(i),H_{00}(j))}\id{e}{d}
\end{equation*}
which asserts that $A_1(q,H_{00}(i),H_{00}(j))$ is contractible. By 
the contractibility of each $A_0(i)$, is is equivalent to the assertion
that each $A_1(q,x,y)$ is contractible.
\end{proof}

\begin{rmk}
We address the question whether it is the case that a graph $A$ in context
$\Gamma$ is contractible if and only if $\terms{A}$ is contractible. As a
consequence of \autoref{lem:contractible-graphs}, it is indeed the case
that $\terms{A}$ is contractible whenever $A$ is. However, the converse
does not hold.

To see this, we first construct a counter example to the converse of the
weak function extensionality principle, which states that there is a function
of type
\begin{equation}
\iscontr(\prd{x:X}P(x))\to\prd{x:X}\iscontr(P(x))\label{eq:wfe-converse}
\end{equation}
for any type family $P:X\to\type$. In the proof of \autoref{thm:wfe-converse}, 
we will find a family
$P:X\to\type$ with the property that $\prd{x:X}P(x)$ is contractible and for
which there is a term $x:A$ with $P(x)$ not contractible. Disproving the
converse of the weak function extensionality principle suffices for our
purposes, because if $P:X\to\type$ is a counter example to \autoref{eq:wfe-converse},
then we can take $\Gamma\defeq\pairr{\Gamma_0,\Gamma_1}$ to be given by
$\Gamma_0\defeq X$ and $\Gamma_1(i,j)\defeq\emptyt$ and we take
$A\defeq\pairr{A_0,A_1}$ to be given by $A_0\defeq P$ and $A_1(q,u,v)
\defeq\emptyt$.
\end{rmk}

We define the family $\mathcal{T}:\Sn^1\to\UU$ by
We define $\mathcal{T}(\base)\defeq\mathbf{3}$. To define $\mathcal{T}(\lloop):
\id{\mathbf{3}}{\mathbf{3}}$ we apply the univalence axiom. Hence it suffices to find an
equivalence $\eqv{\mathbf{3}}{\mathbf{3}}$, for which we take the function
$e$ defined by
\begin{equation*}
e(x)\defeq\begin{cases}
0_\mathbf{3} & \text{if }x\jdeq 0_\mathbf{3}\\
2_\mathbf{3} & \text{if }x\jdeq 1_\mathbf{3}\\
1_\mathbf{3} & \text{if }x\jdeq 2_\mathbf{3}.
\end{cases}
\end{equation*}

\begin{lem}
The type $\terms{\mathcal{T}}\defeq\prd{x:\Sn^1}\mathcal{T}(x)$ is contractible.
\end{lem}

\begin{proof}
The type of sections of $\mathcal{T}$ is equivalent to $\sm{u:\mathcal{T}(\base)}\id{e(u)}{u}$.
If we have a term $\pairr{u,\alpha}$ of the latter type, it follows by induction
on $\mathbf{3}$ that $\id{\pairr{u,\alpha}}
{\pairr{0_\mathbf{3},\refl{0_\mathbf{3}}}}$ for all $\pairr{u,\alpha}:
\sm{x:\mathcal{T}(\base)}\id{e(u)}{u}$,
which shows that the type of sections of $\mathcal{T}$ is contractible.
\end{proof}

\begin{thm}\label{thm:wfe-converse}
There is a type family $P:A\to\type$ for which
\begin{equation*}
\neg\Big(\iscontr(\prd{x:A}P(x))\to\prd{x:A}\iscontr(P(x))\Big).
\end{equation*}
\end{thm}

\begin{proof}
The type family of our counter example is $\mathcal{T}$: the fiber $\mathcal{T}(\base)$ isn't contractible.
\end{proof}

\begin{defn}
A graph morphism $f:\ctxhom{\Delta}{\Gamma}$ is an equivalence of graphs when
$\graphcharmap[\Gamma]{f}$ is a contractible graph in the context $\Gamma$.
\end{defn}

\begin{lem}
Let $A$ and $B$ be graphs in a context $\Gamma$ and let $f:A\to B$. The following are equivalent:
\begin{enumerate}
\item $f[i]:A[i]\to B[i]$ is an equivalence of graphs for every term $i:\Gamma$.
\item $\ctxext{\Gamma}f:\ctxext{\Gamma}A\to\ctxext{\Gamma}B$ is an equivalence of graphs.
\item Both $f_0(i)$ and $f_1(q,x,y)$ are always equivalences.
\item $\terms{f}:\terms{\Delta}\to\terms{\Gamma}$ is an equivalence.
\end{enumerate}
\end{lem}

\begin{rmk}
It follows that there is an equivalence
\begin{equation*}
\ctxext{\Gamma}\msm{A}{P}\simeq\ctxext({\Gamma}{A})P
\end{equation*}
for every family $P$ of graphs over a graph $A$ in context $\Gamma$.
\end{rmk}

\subsubsection{Homotopy levels}
\begin{itemize}
\item A graph $A$ in context $\Gamma$ is of homotopy level $n$ precisely when each
$A_0(i)$ and each $A_1(q,x,y)$ are of homotopy level $n$. 
\item We can name at least three different propositions in the empty context:
\begin{enumerate}
\item $\Gamma_0\defeq\emptyt$.
\item $\Gamma_0\defeq\unit$ and $\Gamma_1(\ttt,\ttt)\defeq\emptyt$.
\item $\Gamma_0\defeq\unit$ and $\Gamma_1(\ttt,\ttt)\defeq\unit$.
\end{enumerate}
Therefore $\mfGraph$ does not satisfy the law of excluded middle.
\end{itemize}

\subsubsection{Univalence for the graph model}
In the other direction, we also obtain a graph morphism $\graphcharmap[\Gamma]{f}:
\ctxhom{\Gamma}{\gobjclass}$ for every graph morphism $f:\ctxhom{\Delta}{\Gamma}$.
In \autoref{graph-object-classifier} we will prove that the maps
$\int_\Gamma:\ctxhom{\Gamma}{\gobjclass}\to\sm{\Delta:\tfGraph}\ctxhom{\Delta}{\Gamma}$
is an equivalence with iverse $\graphcharmapfunc{\Gamma}$.

\begin{defn}
Let $f:\Delta\to\Gamma$ be a graph morphism. We define the graph morphism
$\graphcharmap[\Gamma]{f}$ by
\begin{align*}
\graphcharmap[\Gamma]{f}_0 & \defeq \lam{i}\hfib{f_0}{i}\\
\graphcharmap[\Gamma]{f}_1(i,j) & \defeq \lam{q}\hfib{f_1(i,j)}{q}.
\end{align*}
\end{defn}

\begin{thm}\label{graph-object-classifier}
Main theorem here.
\end{thm}

\begin{comment}
To describe the object classifier for graphs, we will follow Streicher. Thus
we have to look at presheaves over $I/i$ for each object $i$ of the category
$I\defeq 0{\rightrightarrows}1$ with the morphisms named $s$ and $t$ for source
and target. The category $I/0$ is the terminal category;
the category $I/1$ looks like $\cdot{\rightarrow}\cdot{\leftarrow}\cdot$.
Therefore, we have
\begin{align*}
\type^{\op{(I/0)}} & \eqv{}{\type},\\
\type^{\op{(I/1)}} & \eqv{}{\sm{X,Y,A:\type}(A\to X)\times(A\to Y)}\\
& \eqv{}{\sm{X,Y:\type}X\to Y\to\type}.
\end{align*}
The functors $\type^\op{\Sigma_s}$ and $\type^\op{\Sigma_t}$ are given by
$\pi_1$ and $\pi_2$ respectively. This leads to our following definition
of the object classifier $\gobjclass$:

\begin{defn}
Define $\gobjclass$ to be the graph consisting of
\begin{align*}
\gobjclass_0 & \defeq  \type\\
\gobjclass_1(X,Y) & \defeq  X\to Y\to\type
\end{align*}
and define $\pointed{\gobjclass}$ by
\begin{align*}
(\pointed{\gobjclass})_0 & \defeq  \pointed{\type}\\
(\pointed{\gobjclass})_1(\pairr{X,x},\pairr{Y,y}) & \defeq  \sm{R:X\to Y\to\type}R(x,y)
\end{align*}
There is the obvious forgetful graph morphism $t:\pointed{\gobjclass}\to\gobjclass$,
given by projection on the first coordinate.

For any morphism $f:\Delta\to\Gamma$ of graphs we define a morphism
$\graphcharmap(f):\Gamma\to\gobjclass$ of graphs by
\begin{align*}
\graphcharmap(f)_0(i) & \defeq  \hfiber{f_0}{i}\\
\graphcharmap(f)_1(q,\pairr{u,\alpha},\pairr{v,\beta}) & \defeq  \sm{p:\Delta_1(u,v)}
\id{\trans{\pairr{\alpha,\beta}}{f_1(p)}}{q}
\end{align*}
where $\pairr{u,\alpha}:\graphcharmap(f)_0(i)$ and $\pairr{v,\beta}:\graphcharmap(f)_0(j)$. The
morphism $\graphcharmap(f)$ is called the \emph{characteristic map of $f$}. We obtain a function
\begin{equation*}
\graphcharmap : \big(\sm{\Delta:\tfGraph }\Delta\to\Gamma\big)\to\big(\Gamma\to\gobjclass\big)
\end{equation*}
for every graph $\Gamma$.
\end{defn}

\begin{thm}\label{thm:graph-classifier1}
The function $\graphcharmap$ is an equivalence for any graph $\Gamma$.
\end{thm}

\begin{proof}
We have to find a quasi-inverse
\begin{align*}
\Sigma : (\Gamma\to\gobjclass)\to\big(\sm{\Delta:\tfGraph}\Delta\to\Gamma\big)
\end{align*}
of $\graphcharmap$. Thus, we have to define $\Sigma_0:(\Gamma\to\gobjclass)\to\tfGraph$ and
$\Sigma_1:\prd{P:\Gamma\to\gobjclass}\Sigma_0(P)\to\Gamma$. For $P:\Gamma\to\gobjclass$ we define
\begin{align*}
\Sigma_0(P)_0 & \defeq \sm{i:\Gamma_0}P_0(i)\\
\Sigma_0(P)_1(\pairr{i,u},\pairr{j,v}) & \defeq \sm{q:\Gamma_1(i,j)}P_1(q,u,v)\\
\Sigma_1(P)_0 & \defeq \proj1\\
\Sigma_1(P)_1(\pairr{i,u},\pairr{j,v}) & \defeq \proj1.
\end{align*}
\end{proof}

\begin{thm}\label{conj:graph_classifier2}
For any graph morphism $f:\Delta\to\Gamma$, the diagram
\begin{equation*}
\begin{tikzcd}
\Delta \ar{r}{} \ar{d}[swap]{f} & \pointed{\gobjclass} \ar{d}{t} \\ 
\Gamma \ar{r}[swap]{\graphcharmap(f)} & \gobjclass 
\end{tikzcd}
\end{equation*}
is a pullback square.
\end{thm}
\note{We would like it to be a pb \emph{in} the graph model}
\end{comment}

\begingroup\color{blue}
\subsubsection{The adjunctions $\tfcolim\dashv\Delta\dashv\tflim$}
We define $\Delta:\type\to\tfGraph$ by
\begin{equation*}
\Delta(X)\defeq\pairr{X,\lam{x}{x'}\id{x}{x'}}
\end{equation*}
for $X:\type$. For $A:X\to\type$ we define $\Delta(A):\mftyp(\Delta(X))$ by
\begin{align*}
\Delta(A)_0(x) & \defeq A(x)\\
\Delta(A)_1(p,a,b) & \defeq \id{\trans{p}{a}}{b}.
\end{align*}

\begin{lem}
For any type $X$ and any graph $\Gamma$ there is an equivalence
\begin{equation*}
\eqv{(X\to\terms{\Gamma})}{\terms{\Delta(X)\to\Gamma}}.
\end{equation*}
\end{lem}

\begin{proof}
We have to find functions
\begin{align*}
\varphi & : (X\to\terms{\Gamma})\to\terms{\Delta(X)\to\Gamma}\\
\psi & : \terms{\Delta(X)\to\Gamma}\to X\to\terms{\Gamma}
\end{align*}
which are each others homotopy inverse. To define $\varphi$, let
\end{proof}
\endgroup

### 2013-09-27 article-models-intro.tex
resolve issue #7

\section{Introduction}
In this paper we define a notion of internal model which is adapted to
the univalent setting in which we work. Thus, we only describe what it means
to be an internal model when the ambient type theory is univalent. Our
presentation is derived from the definition of~\cite{Dybjer1996}. There,
an internal model is defined to be a category with families with interpretations
for the basic type constructors $\Pi$, $\Sigma$ and $\idtypevar{}$. However,
a category in~\cite{Dybjer1996} has a set of objects and for every two objects
a setoid of morphisms. This is too restrictive for our purpose. Ideally, we
would start with an $\infty$-category of contexts and an $\infty$-presheaf
of families over it. This presents us with the problem that a notion of
$\infty$-category has not been formulated \emph{in} type theory. Therefore,
we make a deviation from the approach in~\cite{Dybjer1996}. We shall
start with a type $\tfctx$ of contexts, a family $\mftypfunc:\tfctx\to\type$
of families over it and a family $\terms{\blank}:\prd*{\Gamma:\tfctx}\mftyp{\mfM}{\Gamma}
\to\type$ sending a type $A$ in context $\Gamma$ to its type $\terms{A}$ of
terms; then we shall directly interpret $\Pi$-types, declaring
$\terms{A\to B}$ to be $\ctxhom[\Gamma]{A}{B}$. The idea is that by letting
$\Pi$ play a fundamental role in the definition of a category with families
$\mathcal{C}$,
the morphisms get their properties directly from the internal type theory
of $\mathcal{C}$.

\begin{comment}
We briefly list the data of which an internal model of type theory consits. We
consider internal models in the style of~\cite{Dybjer1996}. Thus we will
describe a model $\mfM$ as a category with families. However, since our setting is
univalent type theory we will deviate from~\cite{Dybjer1996} in the following
respects: first of all, we make use of typical ambiguity and hence omit
reference to a thing called \emph{sort}.

We organize the definition of an internal model $\mfM$ as follows: in
\autoref{internal-model-contexts} we describe the category of contexts itself;
in \autoref{internal-model-families} we describe the families over a given context
and the related operations; in \autoref{internal-model-constructors} we
describe the basic constructors in the internal model.
\end{comment}

\subsection{Ideas in the definition}
An internal model of type theory is like a category with families, but we want
to avoid having to state higher coherences. In fact, we don't even start our
definition with a category of contexts; instead we just take a \emph{type} of contexts. 
The morphisms will come from the terms, evaluation of a function at a given
term will come from substitution. We recognize three basic ingredients to models:
first there is a type of contexts; second, for every context there is a model of types in
that context and third, for every type in a given context there is a type of its
terms. Then there are three basic attributes: context extension, weakening and
substitution. Context extension provides us with families over types as well as
with an interpretation of dependent pair types. We need weakening 
so that families can depend on the same type multiple times (the way the
identity type of a type depends two times on that type) and to be able
to talk about non-dependent function types,
the morphims of our category. Substitution will give us a way
to work with fibers of families as well as composition of functions and evaluation
of functions at terms.

Because we require a \emph{model} of types in a context, all the structure
which we require at the bottom level will be required to exist higher up as well.
Thus, the model of types in a given context $\Gamma$ will have a type of contexts
itself, which can be seen as the type of types in $\Gamma$; it will have its
own notion of types in a context, its own notion of terms, context extension,
weakening and substitution together with all the structure require for it. For
instance, when $A$ is a type in context $\Gamma$ in a model $\mfM$, then there
is the model of types in context $A$, which is the model of families over $A$. 
This model is required to be \emph{definitionally equal to} the model of types
in the context $\ctxext{\Gamma}A$, the context extension of $\Gamma$ and $A$.
In this way we protect ourselves from the need to dig an infinitely deep structure
of models when we want to consider examples.

To give the definition of a model we shall also need to consider certain morphisms
of internal models. Those should preserve all the structure: contexts are mapped
to contexts; for every context a morphism of models mapping the model of types
in that context to the model of types in the image of that context; there should
be a mapping of terms and context extension, weakening and substitution should be
preserved. We need to consider those morphisms because we require context extension,
weakening and substitution to be of that kind, thereby respecting each other
in all possible ways.

When we have this framework set up, we can interpret the basic type constructors
such as $\Pi$, $\Sigma$ and $\idtypevar{}$.
The higher categorical structure then comes from the
result that we have an interpretation of type theory.

\begingroup
\color{red}
\begin{rmk}
Currently, the definition seems to be circular. To define a model we need that
context extension, weakening and substitution be morphisms of models. A morphism
of models needs to preserve context extension, weakening and substitution.
Moreover, its definition requires the notion of composition of morphisms.

I've done it this way because it guides me to what the rules should be, trusting
that I can work my way back to give a (possibly less transparent) definition
which contains no circularities without doubt.
\end{rmk}
\endgroup

### 2013-10-24 article-models-defn.tex
introducing strict compatibility properties for substitution and weakening
putting the new macros \ctxext, \ctxwk and \subst into use

@@
 We introduce extension which not only extends a context $\Gamma$ and a type
-$A$ over it to a context $\ctxext{\Gamma}A$, but which also extends a type $A$
-in context $\Gamma$ and a family $P$ over it to a type $\ctxext{A}P$ in context
+$A$ over it to a context $\ctxext{\Gamma}{A}$, but which also extends a type $A$
+in context $\Gamma$ and a family $P$ over it to a type $\ctxext{A}{P}$ in context
 $\Gamma$. We do this to ensure that all of type theory can be done in a context.
@@
 \begin{infarray}{cc}
-\inference{\jtype{\Gamma}{A}}{\jctx{\ctxext{\Gamma}A}}
-& \inference{\jctxeq{\Gamma}{\Delta}\qquad\jtypeeq{\Gamma}{A}{B}}{\jctxeq{\ctxext{\Gamma}A}{\ctxext{\Delta}B}}\\
-\inference{\jtype{\ctxext{\Gamma}A}{P}}{\jtype{\Gamma}{\ctxext{A}P}}
-& \inference{\jtypeeq{\Gamma}{A}{B}\qquad\jtypeeq{\ctxext{\Gamma}A}{P}{Q}}{\jtypeeq{\Gamma}{\ctxext{A}P}{\ctxext{B}Q}}
+\inference{\jtype{\Gamma}{A}}{\jctx{\ctxext{\Gamma}{A}}}
+& \inference{\jctxeq{\Gamma}{\Delta}\qquad\jtypeeq{\Gamma}{A}{B}}{\jctxeq{\ctxext{\Gamma}{A}}{\ctxext{\Delta}{B}}}\\
+\inference{\jtype{\ctxext{\Gamma}{A}}{P}}{\jtype{\Gamma}{\ctxext{A}{P}}}
+& \inference{\jtypeeq{\Gamma}{A}{B}\qquad\jtypeeq{\ctxext{\Gamma}{A}}{P}{Q}}{\jtypeeq{\Gamma}{\ctxext{A}{P}}{\ctxext{B}{Q}}}
 \end{infarray}
@@
 & \inference{\jctxeq{\Gamma}{\Gamma'}\qquad\jctxeq{\Delta}{\Delta'}}{\jtypeeq{\Gamma}{\ctxwk{\Gamma}{\Delta}}{\ctxwk{\Gamma'}{\Delta'}}}\\
-\inference{\jctx{\Gamma}\qquad\jtype{\Delta}{B}}{\jtype{\ctxext{\Gamma}\ctxwk{\Gamma}{\Delta}}{\ctxwk{\Gamma}{B}}} & \inference{\jctxeq{\Gamma}{\Gamma'}\qquad\jtypeeq{\Delta}{B}{B'}}{\jtypeeq{\ctxext{\Gamma}\ctxwk{\Gamma}{\Delta}}{\ctxwk{\Gamma}{B}}{\ctxwk{\Gamma'}{B'}}}\\
-\inference{\jctx{\Gamma}\qquad\jterm{\Delta}{B}{y}}{\jterm{\ctxext{\Gamma}\ctxwk{\Gamma}{\Delta}}{\ctxwk{\Gamma}{B}}{\ctxwk{\Gamma}{y}}} 
-& \inference{\jctxeq{\Gamma}{\Gamma'}\qquad\jtermeq{\Delta}{B}{y}{y'}}{\jtermeq{\ctxext{\Gamma}\ctxwk{\Gamma}{\Delta}}{\ctxwk{\Gamma}{B}}{\ctxwk{\Gamma}{y}}{\ctxwk{\Gamma'}{y'}}} 
+\inference{\jctx{\Gamma}\qquad\jtype{\Delta}{B}}{\jtype{\ctxext{\Gamma}{\ctxwk{\Gamma}{\Delta}}}{\ctxwk{\Gamma}{B}}} & \inference{\jctxeq{\Gamma}{\Gamma'}\qquad\jtypeeq{\Delta}{B}{B'}}{\jtypeeq{\ctxext{\Gamma}{\ctxwk{\Gamma}{\Delta}}}{\ctxwk{\Gamma}{B}}{\ctxwk{\Gamma'}{B'}}}\\
+\inference{\jctx{\Gamma}\qquad\jterm{\Delta}{B}{y}}{\jterm{\ctxext{\Gamma}{\ctxwk{\Gamma}{\Delta}}}{\ctxwk{\Gamma}{B}}{\ctxwk{\Gamma}{y}}} 
+& \inference{\jctxeq{\Gamma}{\Gamma'}\qquad\jtermeq{\Delta}{B}{y}{y'}}{\jtermeq{\ctxext{\Gamma}{\ctxwk{\Gamma}{\Delta}}}{\ctxwk{\Gamma}{B}}{\ctxwk{\Gamma}{y}}{\ctxwk{\Gamma'}{y'}}} 
 \end{infarray}
@@
 A weakening operation is also defined for types. When $A$ and $B$ are both types
- in context $\Gamma$, the weakened type $\ctxwk{A}{B}$ in context $\ctxext{\Gamma}A$
+ in context $\Gamma$, the weakened type $\ctxwk{A}{B}$ in context $\ctxext{\Gamma}{A}$
  is the family which `doesn't really depend on $A$'. The terms of $\ctxwk{A}{B}$
- are the functions from $A$ to $B$. Likewise, the terms $\ctxwk{A}y$ are the
+ are the functions from $A$ to $B$. Likewise, the terms $\ctxwk{A}{y}$ are the
  constant maps at $y$ from $A$ to $B$, for $y:B$.
 \begin{infarray}{cc}
-\inference{\jtype{\Gamma}{A}\qquad\jtype{\Gamma}{B}}{\jtype{\ctxext{\Gamma}A}{\ctxwk{A}{B}}}
-& \inference{\jtypeeq{\Gamma}{A}{A'}\qquad\jtypeeq{\Gamma}{B}{B'}}{\jtypeeq{\ctxext{\Gamma}A}{\ctxwk{A}{B}}{\ctxwk{A'}{B'}}}\\
-\inference{\jtype{\Gamma}{A}\qquad\jtype{\ctxext{\Gamma}B}{Q}}
+\inference{\jtype{\Gamma}{A}\qquad\jtype{\Gamma}{B}}{\jtype{\ctxext{\Gamma}{A}}{\ctxwk{A}{B}}}
+& \inference{\jtypeeq{\Gamma}{A}{A'}\qquad\jtypeeq{\Gamma}{B}{B'}}{\jtypeeq{\ctxext{\Gamma}{A}}{\ctxwk{A}{B}}{\ctxwk{A'}{B'}}}\\
+\inference{\jtype{\Gamma}{A}\qquad\jtype{\ctxext{\Gamma}{B}}{Q}}
 {\jtype{\ctxext({\Gamma}{A})\ctxwk{A}{B}}{\ctxwk{A}{Q}}}
-& \inference{\jtypeeq{\Gamma}{A}{A'}\qquad\jtypeeq{\ctxext{\Gamma}B}{Q}{Q'}}
+& \inference{\jtypeeq{\Gamma}{A}{A'}\qquad\jtypeeq{\ctxext{\Gamma}{B}}{Q}{Q'}}
 {\jtypeeq{\ctxext({\Gamma}{A})\ctxwk{A}{B}}{\ctxwk{A}{Q}}{\ctxwk{A'}{Q'}}}\\
-\inference{\jtype{\Gamma}{A}\qquad\jterm{\Gamma}{B}{y}}{\jterm{\ctxext{\Gamma}A}{\ctxwk{A}{B}}{\ctxwk{A}{y}}}
-& \inference{\jtypeeq{\Gamma}{A}{A'}\qquad\jtermeq{\Gamma}{B}{y}{y'}}{\jtermeq{\ctxext{\Gamma}A}{\ctxwk{A}{B}}{\ctxwk{A}{y}}{\ctxwk{A'}{y'}}}
+\inference{\jtype{\Gamma}{A}\qquad\jterm{\Gamma}{B}{y}}{\jterm{\ctxext{\Gamma}{A}}{\ctxwk{A}{B}}{\ctxwk{A}{y}}}
+& \inference{\jtypeeq{\Gamma}{A}{A'}\qquad\jtermeq{\Gamma}{B}{y}{y'}}{\jtermeq{\ctxext{\Gamma}{A}}{\ctxwk{A}{B}}{\ctxwk{A}{y}}{\ctxwk{A'}{y'}}}
 \end{infarray}
@@
 \begin{infarray}{cc}
-\inference{\jterm{\Gamma}{A}{x}\qquad\jtype{\ctxext{\Gamma}A}{P}}{\jtype{\Gamma}{\subst{x}{P}}}
-& \inference{\jtermeq{\Gamma}{A}{x}{x'}\qquad\jtypeeq{\ctxext{\Gamma}A}{P}{P'}}{\jtypeeq{\Gamma}{\subst{x}{P}}{\subst{x'}{P'}}}\\
-\inference{\jterm{\Gamma}{A}{x}\qquad\jtype{\ctxext({\Gamma}{A})P}{Q}}{\jtype{\ctxext{\Gamma}\subst{x}{P}}{\subst{x}{Q}}}
-& \inference{\jtermeq{\Gamma}{A}{x}{x'}\qquad\jtypeeq{\ctxext({\Gamma}{A})P}{Q}{Q'}}{\jtypeeq{\ctxext{\Gamma}\subst{x}{P}}{\subst{x}{Q}}{\subst{x'}{Q'}}}\\
-\inference{\jterm{\Gamma}{A}{x}\qquad\jterm{\ctxext{\Gamma}A}{P}{f}}{\jterm{\Gamma}{\subst{x}{P}}{\subst{x}{f}}}
-& \inference{\jtermeq{\Gamma}{A}{x}{x'}\qquad\jtermeq{\ctxext{\Gamma}A}{P}{f}{f'}}{\jtermeq{\Gamma}{\subst{x}{P}}{\subst{x}{f}}{\subst{x'}{f'}}}\\
-\inference{\jterm{\Gamma}{A}{x}\qquad\jterm{\ctxext({\Gamma}{A}){P}}{Q}{g}}{\jterm{\ctxext{\Gamma}\subst{x}{P}}{\subst{x}{Q}}{\subst{x}{g}}}
-& \inference{\jtermeq{\Gamma}{A}{x}{x'}\qquad\jtermeq{\ctxext({\Gamma}{A})P}{Q}{g}{g'}}{\jtermeq{\ctxext{\Gamma}\subst{x}{P}}{\subst{x}{Q}}{\subst{x}{g}}{\subst{x'}{g'}}}
+\inference{\jterm{\Gamma}{A}{x}\qquad\jtype{\ctxext{\Gamma}{A}}{P}}{\jtype{\Gamma}{\subst{x}{P}}}
+& \inference{\jtermeq{\Gamma}{A}{x}{x'}\qquad\jtypeeq{\ctxext{\Gamma}{A}}{P}{P'}}{\jtypeeq{\Gamma}{\subst{x}{P}}{\subst{x'}{P'}}}\\
+\inference{\jterm{\Gamma}{A}{x}\qquad\jtype{\ctxext{{\Gamma}{A}}{P}}{Q}}{\jtype{\ctxext{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}}
+& \inference{\jtermeq{\Gamma}{A}{x}{x'}\qquad\jtypeeq{\ctxext{{\Gamma}{A}}{P}}{Q}{Q'}}{\jtypeeq{\ctxext{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}{\subst{x'}{Q'}}}\\
+\inference{\jterm{\Gamma}{A}{x}\qquad\jterm{\ctxext{\Gamma}{A}}{P}{f}}{\jterm{\Gamma}{\subst{x}{P}}{\subst{x}{f}}}
+& \inference{\jtermeq{\Gamma}{A}{x}{x'}\qquad\jtermeq{\ctxext{\Gamma}{A}}{P}{f}{f'}}{\jtermeq{\Gamma}{\subst{x}{P}}{\subst{x}{f}}{\subst{x'}{f'}}}\\
+\inference{\jterm{\Gamma}{A}{x}\qquad\jterm{\ctxext{{\Gamma}{A}}{P}}{Q}{g}}{\jterm{\ctxext{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}{\subst{x}{g}}}
+& \inference{\jtermeq{\Gamma}{A}{x}{x'}\qquad\jtermeq{\ctxext{{\Gamma}{A}}{P}}{Q}{g}{g'}}{\jtermeq{\ctxext{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}{\subst{x}{g}}{\subst{x'}{g'}}}
 \end{infarray}
@@
 \inference
-{ \jtype{\ctxext{\Gamma}A}{P}
+{ \jtype{\ctxext{\Gamma}{A}}{P}
 }{\jterm
-  {\ctxext{\Gamma}({A}P)}
-  {\ctxwk{\ctxext{\Gamma}({A}P)}{(\ctxext({\Gamma}{A})P)}}{\typefont{extext}_0(A,P)}
+  {\ctxext{\Gamma}{{A}{P}}}
+  {\ctxwk{\ctxext{\Gamma}{{A}{P}}}{(\ctxext{{\Gamma}{A}}{P})}}{\typefont{extext}_0(A,P)}
 }\\
 \inference
-{ \jtype{\ctxext({\Gamma}{A})P}{Q}
+{ \jtype{\ctxext{{\Gamma}{A}}{P}}{Q}
 }{\jterm
-  {\ctxext{\Gamma}({A}({P}Q))}
-  {\ctxwk{\ctxext{A}({P}Q)}{(\ctxext({A}{P})Q)}}
+  {\ctxext{\Gamma}{{A}{{P}{Q}}}}
+  {\ctxwk{\ctxext{A}{{P}{Q}}}{(\ctxext{{A}{P}}{Q})}}
   {\typefont{extext}_1(A,P,Q)}
@@
   {\jterm
-    {\ctxext{\Gamma}\ctxwk{\Gamma}{(\ctxext{\Delta}B)}}
-    {\ctxwk{\ctxwk{\Gamma}{(\ctxext{\Delta}B)}}{\ctxext{(\ctxwk{\Gamma}{\Delta})}(\ctxwk{\Gamma}{B})}}
+    {\ctxext{\Gamma}{\ctxwk{\Gamma}{(\ctxext{\Delta}{B})}}}
+    {\ctxwk{{\Gamma}{(\ctxext{\Delta}{B})}}{\ctxext{(\ctxwk{\Gamma}{\Delta})}{(\ctxwk{\Gamma}{B})}}}
     {\typefont{wkext}_0(\Gamma,B)}}
+\end{infarray}
+
+Substitution is compatible with substitution:
+\begin{infarray}{c}
+\inference{\jtype{\ctxext{{\Gamma}{A}}{P}}{Q}\qquad\jterm{\Gamma}{A}{x}\qquad\jterm{\ctxext{\Gamma}{A}}{P}{f}}
+{\jtypeeq{\Gamma}{\subst{x}{\subst{f}{Q}}}{\subst{{x}{f}}{\subst{x}{Q}}}}
 \end{infarray}
@@
 \begin{infarray}{c}
-\inference{\jtype{\Gamma}{A}}{\jterm{\ctxext{\Gamma}A}{\ctxwk{A}{A}}{\idfunc[A]}}
+\inference{\jtype{\Gamma}{A}}{\jterm{\ctxext{\Gamma}{A}}{\ctxwk{A}{A}}{\idfunc[A]}}
 \end{infarray}
@@
 Let $P$ be a family over $A$ in context $\Gamma$. Then we can weaken
-$\ctxwk{A}{P}$, which is a family over $\ctxwk{A}{A}$ in context $\ctxext{\Gamma}A$.
-We also have the family $\ctxwk{\ctxwk{A}{A}}{P}$ over $\ctxwk{A}{A}$ in context
-$\ctxext{\Gamma}A$ and thus we have the inference
+$\ctxwk{A}{P}$, which is a family over $\ctxwk{A}{A}$ in context $\ctxext{\Gamma}{A}$.
+We also have the family $\ctxwk{{A}{A}}{P}$ over $\ctxwk{A}{A}$ in context
+$\ctxext{\Gamma}{A}$ and thus we have the inference
 \begin{equation*}
-\inference{\jtype{\ctxext{\Gamma}A}{P}}{\jtype{\ctxext(({\Gamma}{A}){\ctxwk{A}{A}})\ctxwk{A}{P}}{\ctxwk{\ctxwk{A}{P}}{(\ctxwk{\ctxwk{A}{A}}{P})}}}
+\inference{\jtype{\ctxext{\Gamma}{A}}{P}}{\jtype{\ctxext(({\Gamma}{A}){\ctxwk{A}{A}})\ctxwk{A}{P}}{\ctxwk{{A}{P}}{(\ctxwk{{A}{A}}{P})}}}
 \end{equation*}
@@
 \item For a family $Q$ over $B$ in context $\Gamma$ we define $f^\ast Q\jdeq\subst{f}{(\ctxwk{A}{Q})}$.
-\item For a term $g$ of $Q$ in context $\ctxext{\Gamma}B$ we define $f^\ast g\jdeq\subst{f}{(\ctxwk{A}{g})}$.
+\item For a term $g$ of $Q$ in context $\ctxext{\Gamma}{B}$ we define $f^\ast g\jdeq\subst{f}{(\ctxwk{A}{g})}$.
 \end{enumerate} 
@@
 \begin{rmk}
-The type $f^\ast(\ctxwk{B}{C})$ in context $\ctxext{\Gamma}A$ is can be viewed as the
+The type $f^\ast(\ctxwk{B}{C})$ in context $\ctxext{\Gamma}{A}$ is can be viewed as the
 type of functions from $A$ to $C$ which factor through $f$. Thus there should be
@@
 \begin{rmk}
-Every term $\jterm{\ctxext{\Gamma}A}{\ctxwk{A}{B}}{f}$ allows us to infer the following:
+Every term $\jterm{\ctxext{\Gamma}{A}}{\ctxwk{A}{B}}{f}$ allows us to infer the following:
 \begin{equation*}
-\inference{\jtype{\ctxext{\Gamma}B}{Q}}{\jtype{\ctxext{\Gamma}A}{f^\ast Q}}
+\inference{\jtype{\ctxext{\Gamma}{B}}{Q}}{\jtype{\ctxext{\Gamma}{A}}{f^\ast Q}}
 \qquad
-\inference{\jterm{\ctxext{\Gamma}B}{Q}{g}}{\jterm{\ctxext{\Gamma}A}{f^\ast Q}{f^\ast g}}
+\inference{\jterm{\ctxext{\Gamma}{B}}{Q}{g}}{\jterm{\ctxext{\Gamma}{A}}{f^\ast Q}{f^\ast g}}
 \end{equation*}
@@
 \begin{defn}
-A term $f:\ctxwk{A}{B}$ in context $\ctxext{\Gamma}A$ is said to be a trivial
+A term $f:\ctxwk{A}{B}$ in context $\ctxext{\Gamma}{A}$ is said to be a trivial
 cofibration if we can infer
 \begin{equation*}
-\inference{\jterm{\ctxext{\Gamma}A}{f^\ast Q}{t}}{\jterm{\ctxext{\Gamma}B}{Q}{\tilde{t}}}\qquad
-\inference{\jterm{\ctxext{\Gamma}A}{f^\ast Q}{t}}{\jtermeq{\ctxext{\Gamma}A}{f^\ast Q}{f^\ast \tilde{t}}{t}}
+\inference{\jterm{\ctxext{\Gamma}{A}}{f^\ast Q}{t}}{\jterm{\ctxext{\Gamma}{B}}{Q}{\tilde{t}}}\qquad
+\inference{\jterm{\ctxext{\Gamma}{A}}{f^\ast Q}{t}}{\jtermeq{\ctxext{\Gamma}{A}}{f^\ast Q}{f^\ast \tilde{t}}{t}}
 \end{equation*}
@@
 Another approach would be to define $f:\ctxwk{A}{B}$ to be left invertible
-if there is a term $\ctxext({\Gamma}{A})P$
+if there is a term $\ctxext{{\Gamma}{A}}{P}$
 
@@
 \begin{infarray}{cc}
-\inference{\jtype{\ctxext{\Gamma}A}{P}}{\jtype{\Gamma}{\mprd{A}{P}}}
-& \inference{\jtypeeq{\ctxext{\Gamma}A}{P}{P'}}{\jtypeeq{\Gamma}{\mprd{A}{P}}{\mprd{A}{P'}}}\\
-\inference{\jterm{\ctxext{\Gamma}A}{P}{f}}{\jterm{\Gamma}{\mprd{A}{P}}{\lambda(f)}}
-& \inference{\jtermeq{\ctxext{\Gamma}A}{P}{f}{f'}}{\jtermeq{\Gamma}{\mprd{A}{P}}{\lambda(f)}{\lambda(f')}}\\
-\inference{\jterm{\Gamma}{\mprd{A}{P}}{g}}{\jterm{\ctxext{\Gamma}A}{P}{\tfev(g)}}
-& \inference{\jtermeq{\Gamma}{\mprd{A}{P}}{g}{g'}}{\jtermeq{\ctxext{\Gamma}A}{P}{\tfev(g)}{\tfev(g')}}\\
-\inference{\jterm{\ctxext{\Gamma}A}{P}{f}}{\jtermeq{\Gamma}{A}{\tfev(\lambda(f))}{f}}
+\inference{\jtype{\ctxext{\Gamma}{A}}{P}}{\jtype{\Gamma}{\mprd{A}{P}}}
+& \inference{\jtypeeq{\ctxext{\Gamma}{A}}{P}{P'}}{\jtypeeq{\Gamma}{\mprd{A}{P}}{\mprd{A}{P'}}}\\
+\inference{\jterm{\ctxext{\Gamma}{A}}{P}{f}}{\jterm{\Gamma}{\mprd{A}{P}}{\lambda(f)}}
+& \inference{\jtermeq{\ctxext{\Gamma}{A}}{P}{f}{f'}}{\jtermeq{\Gamma}{\mprd{A}{P}}{\lambda(f)}{\lambda(f')}}\\
+\inference{\jterm{\Gamma}{\mprd{A}{P}}{g}}{\jterm{\ctxext{\Gamma}{A}}{P}{\tfev(g)}}
+& \inference{\jtermeq{\Gamma}{\mprd{A}{P}}{g}{g'}}{\jtermeq{\ctxext{\Gamma}{A}}{P}{\tfev(g)}{\tfev(g')}}\\
+\inference{\jterm{\ctxext{\Gamma}{A}}{P}{f}}{\jtermeq{\Gamma}{A}{\tfev(\lambda(f))}{f}}
 \end{infarray}
@@
 \subsection{Identity types}
-We need to define the diagonal function $\jterm{\ctxext{\Gamma}A}{\ctxwk{A}{(\ctxext{A}\ctxwk{A}{A})}}{\delta}$
+We need to define the diagonal function $\jterm{\ctxext{\Gamma}{A}}{\ctxwk{A}{(\ctxext{A}{\ctxwk{A}{A}})}}{\delta}$
 \begin{infarray}{c}
-\inference{\jtype{\Gamma}{A}}{\jtype{\ctxext({\Gamma}{A})\ctxwk{A}{A}}{\idtypevar{A}}}\\
-\inference{\jtype{\Gamma}{A}}{\jterm{\ctxext{\Gamma}A}{\delta^\ast\idtypevar{A}}{\reflsym}}\\
-\inference{\jtype{\ctxext(({\Gamma}{A}){\ctxwk{A}{A}})\idtypevar{A}}{P}\qquad
-\jterm{\ctxext{\Gamma}A}{\pairr{\delta,\reflsym}^\ast P}{d}}
-{\jterm{\ctxext(({\Gamma}{A}){\ctxwk{A}{A}})\idtypevar{A}}{P}{J}}
+\inference{\jtype{\Gamma}{A}}{\jtype{\ctxext{{\Gamma}{A}}{\ctxwk{A}{A}}}{\idtypevar{A}}}\\
+\inference{\jtype{\Gamma}{A}}{\jterm{\ctxext{\Gamma}{A}}{\delta^\ast\idtypevar{A}}{\reflsym}}\\
+\inference{\jtype{\ctxext{{{\Gamma}{A}}{\ctxwk{A}{A}}}{\idtypevar{A}}}{P}\qquad
+\jterm{\ctxext{\Gamma}{A}}{\pairr{\delta,\reflsym}^\ast P}{d}}
+{\jterm{\ctxext{{{\Gamma}{A}}{\ctxwk{A}{A}}}{\idtypevar{A}}}{P}{J}}
 \end{infarray}
@@
 When $A$ is a type in context $\Gamma$, we also denote
-the context $\tfext^\Gamma_0(A)$ of $\mfM$ by $\ctxext{\Gamma}A$.
+the context $\tfext^\Gamma_0(A)$ of $\mfM$ by $\ctxext{\Gamma}{A}$.
 We will write
-$\ctxext({\Gamma}{A})P$ for $\ctxext({\Gamma}{A})P$.
+$\ctxext{{\Gamma}{A}}{P}$ for $\ctxext{{\Gamma}{A}}{P}$.
 \end{defn}
@@
 \begin{align*}
-\mftyp{\mftyp{\mfM}{\Gamma}}{A} & \jdeq\mftyp{\mfM}{\ctxext{\Gamma}A}\\
-\mftyp{\tfext^\Gamma}{A} & \jdeq \modelfont{id}_{\mftyp{\mfM}{\ctxext{\Gamma}A}}\\
+\mftyp{\mftyp{\mfM}{\Gamma}}{A} & \jdeq\mftyp{\mfM}{\ctxext{\Gamma}{A}}\\
+\mftyp{\tfext^\Gamma}{A} & \jdeq \modelfont{id}_{\mftyp{\mfM}{\ctxext{\Gamma}{A}}}\\
 \terms[{\tfext^\Gamma}]{P} & \jdeq \idfunc[\terms{P}].
@@
 \begin{equation*}
-\mftyp{\mftyp{\mfM}{\Gamma}}{A}\jdeq\mftyp{\mfM}{\ctxext{\Gamma}A},
+\mftyp{\mftyp{\mfM}{\Gamma}}{A}\jdeq\mftyp{\mfM}{\ctxext{\Gamma}{A}},
 \end{equation*}
 ensuring that a context in the model $\mftyp{\mftyp{\mfM}{\Gamma}}{A}$ is the same thing as a
-context in the model $\mftyp{\mfM}{\ctxext{\Gamma}A}$.
+context in the model $\mftyp{\mfM}{\ctxext{\Gamma}{A}}$.
 \item If $Q$ is a family over $P$ where $P$ is a family over $A$ in context $\Gamma$, then
@@
 \jdeq
-\tfext^{\protect{\mftyp{\mfM}{\ctxext{\Gamma}A}}}(P,Q)
+\tfext^{\protect{\mftyp{\mfM}{\ctxext{\Gamma}{A}}}}(P,Q)
 \end{equation*}
@@
 \begin{equation*}
-\tfwk^A:\mftyp{\mfM}{\Gamma}\to\mftyp{\mfM}{\ctxext{\Gamma}A}
+\tfwk^A:\mftyp{\mfM}{\Gamma}\to\mftyp{\mfM}{\ctxext{\Gamma}{A}}
 \end{equation*}
@@
 \begin{equation*}
-\tfsubst^x:\mftyp{\mfM}{\ctxext{\Gamma}A}\to\mftyp{\mfM}{\Gamma}
+\tfsubst^x:\mftyp{\mfM}{\ctxext{\Gamma}{A}}\to\mftyp{\mfM}{\Gamma}
 \end{equation*}
@@
 for any $x:A$ and $B$ in context $\Gamma$.
-\item A context $\unit^\mfM:\tfctx(\mfM)$ and a term of type $\isequiv(\ctxext{\unit^\mfM}\blank)$. We will denote
+\item A context $\unit^\mfM:\tfctx(\mfM)$ and a term of type $\isequiv(\ctxext{\unit^\mfM}{\blank})$. We will denote
 this equivalence by $e_\unit$. The context $\unit^\mfM$ is also
@@
 a type $\unit^\Gamma$ in context $\Gamma$ to every context $\Gamma$ and
-an identification $\alpha_\unit(\Gamma):\id{\ctxext{\Gamma}\unit^\Gamma}{\Gamma}$
+an identification $\alpha_\unit(\Gamma):\id{\ctxext{\Gamma}{\unit^\Gamma}}{\Gamma}$
 for every context $\Gamma$. We also require that there is an identification
@@
 \begin{equation*}
-\mftyp{\mftyp{f}{\Gamma}}{A}\jdeq\mftyp{f}{\ctxext{\Gamma}A}
+\mftyp{\mftyp{f}{\Gamma}}{A}\jdeq\mftyp{f}{\ctxext{\Gamma}{A}}
 \end{equation*}
@@
 \alpha^f_0 & :\id{f\circ\tfext^\Gamma}{\tfext^{f_0(\Gamma)}\circ\mftyp{f}{\Gamma}}\\
-\alpha^f_1 & :\id{\mftyp{f}{\Gamma}_0\circ\tfext^A}{\tfext^{\mftyp{f}{\Gamma}_0(A)}\circ\mftyp{f}{\ctxext{\Gamma}A}}.
+\alpha^f_1 & :\id{\mftyp{f}{\Gamma}_0\circ\tfext^A}{\tfext^{\mftyp{f}{\Gamma}_0(A)}\circ\mftyp{f}{\ctxext{\Gamma}{A}}}.
 \end{align*}
@@
 \begin{equation*}
-f_0(\ctxext{\Gamma}A)\jdeq\ctxext{f_0(\Gamma)}\mftyp{f}{\Gamma}_0(A)
+f_0(\ctxext{\Gamma}{A})\jdeq\ctxext{f_0(\Gamma)}\mftyp{f}{\Gamma}_0(A)
 \end{equation*}
@@
 \beta^f_0 & :\id{\mftyp{f}{\Gamma}\circ\tfwk^\Gamma}{\tfwk^{f_0(\Gamma)}\circ f}\\
-\beta^f_1 & :\id{\mftyp{f}{\ctxext{\Gamma}A}\circ\tfwk^A}{\tfwk^\protect{\mftyp{f}{\Gamma}_0(A)}\circ\mftyp{f}{\Gamma}}.
+\beta^f_1 & :\id{\mftyp{f}{\ctxext{\Gamma}{A}}\circ\tfwk^A}{\tfwk^\protect{\mftyp{f}{\Gamma}_0(A)}\circ\mftyp{f}{\Gamma}}.
 \end{align*}
@@
 \begin{equation*}
-\mftyp{f}{\ctxext{\Gamma}A}_0(\ctxwk{A}{B})\jdeq\ctxwk{\mftyp{f}{\Gamma}_0(A)}{\mftyp{f}{\Gamma}_0(B)}
+\mftyp{f}{\ctxext{\Gamma}{A}}_0(\ctxwk{A}{B})\jdeq\ctxwk{\mftyp{f}{\Gamma}_0(A)}{\mftyp{f}{\Gamma}_0(B)}
 \end{equation*}
@@
 \begin{equation*}
-\mftyp{f}{\ctxext({\Gamma}{A})\ctxwk{A}{B}}\circ\tfwk^A\jdeq\tfwk^{\protect{\mftyp{f}{\Gamma}_0(A)}}\circ\mftyp{f}{\ctxext{\Gamma}B}
+\mftyp{f}{\ctxext({\Gamma}{A})\ctxwk{A}{B}}\circ\tfwk^A\jdeq\tfwk^{\protect{\mftyp{f}{\Gamma}_0(A)}}\circ\mftyp{f}{\ctxext{\Gamma}{B}}
 \end{equation*}
@@
 \begin{equation*}
-\gamma^f:\id{\mftyp{f}\Gamma_0\circ\tfsubst^x}{\tfsubst^\protect{\terms[f]{A}(x)}\circ\mftyp{f}{\ctxext{\Gamma}A}}.
+\gamma^f:\id{\mftyp{f}\Gamma_0\circ\tfsubst^x}{\tfsubst^\protect{\terms[f]{A}(x)}\circ\mftyp{f}{\ctxext{\Gamma}{A}}}.
 \end{equation*}
@@
 \begin{equation*}
-\mftyp{f}\Gamma_0(\subst{x}{P})\jdeq\subst{\terms[f]{A}(x)}{\mftyp{f}{\ctxext{\Gamma}A}_0(P)}
+\mftyp{f}\Gamma_0(\subst{x}{P})\jdeq\subst{\terms[f]{A}(x)}{\mftyp{f}{\ctxext{\Gamma}{A}}_0(P)}
 \end{equation*}
@@
 \begin{equation*}
-\mftyp{f}\Gamma_0(\subst{x}{Q})\jdeq\subst{\terms[f]{A}(x)}{\mftyp{f}{\ctxext({\Gamma}{A})P}_0(Q)}
+\mftyp{f}\Gamma_0(\subst{x}{Q})\jdeq\subst{\terms[f]{A}(x)}{\mftyp{f}{\ctxext{{\Gamma}{A}}{P}}_0(Q)}
 \end{equation*}
@@
 $\tfext(\Gamma):\mftyp{\mfM}{\Gamma}\to M$. Thus, we already have
-$\tfext(\Gamma)_0\defeq\ctxext{\Gamma}\blank$. We also require 
+$\tfext(\Gamma)_0\defeq\ctxext{\Gamma}{\blank}$. We also require 
 \begin{align*}
-\mftyp{\tfext(\Gamma)}{A} & : \mftyp{\mftyp{\mfM}{\Gamma}}{A}\to \mftyp{\mfM}{\ctxext{\Gamma}A}\\
+\mftyp{\tfext(\Gamma)}{A} & : \mftyp{\mftyp{\mfM}{\Gamma}}{A}\to \mftyp{\mfM}{\ctxext{\Gamma}{A}}\\
 \terms[\tfext(\Gamma)]{P} & : \terms[\mftyp{\mfM}{\Gamma}]{P}\to\terms[\mfM]{\mftyp{\tfext(\Gamma)}{A}_0(P)}
@@
 \item Extension preserves extension: for every family $P$ over $A$ in context
-$\Gamma$ an identification $\id{\ctxext({\Gamma}{A})P}{\ctxext{\Gamma}({A}P)}$.
+$\Gamma$ an identification $\id{\ctxext{{\Gamma}{A}}{P}}{\ctxext{\Gamma}{{A}{P}}}$.
 \item Extension preserves weakening:
@@
 \end{enumerate}
-\item[Weakening] is a morphism $\tfwk:\mftyp{\mfM}{\Gamma}\to\mftyp{\mfM}{\ctxext{\Gamma}A}$, so we must have
+\item[Weakening] is a morphism $\tfwk:\mftyp{\mfM}{\Gamma}\to\mftyp{\mfM}{\ctxext{\Gamma}{A}}$, so we must have
 \begin{align*}
-\tfwk^A_0 & : \mftyp{\mfM}{\Gamma}\to\mftyp{\mfM}{\ctxext{\Gamma}A}
+\tfwk^A_0 & : \mftyp{\mfM}{\Gamma}\to\mftyp{\mfM}{\ctxext{\Gamma}{A}}
 \intertext{which is denoted by $\ctxwk{A}{\blank}$,}
-\mftypfunc{\tfwk^A} & : \mftyp{\mfM}{\ctxext{\Gamma}B}\to\mftyp{\mfM}{\ctxext({\Gamma}{A})\ctxwk{A}{B}}\\
-\terms[\tfwk^A]{B} & : \terms[\mftyp{\mfM}{\Gamma}]{B}\to\terms[\mftyp{\mfM}{\ctxext{\Gamma}A}]{\ctxwk{A}{B}}
+\mftypfunc{\tfwk^A} & : \mftyp{\mfM}{\ctxext{\Gamma}{B}}\to\mftyp{\mfM}{\ctxext{{\Gamma}{A}}{\ctxwk{A}{B}}}\\
+\terms[\tfwk^A]{B} & : \terms[\mftyp{\mfM}{\Gamma}]{B}\to\terms[\mftyp{\mfM}{\ctxext{\Gamma}{A}}]{\ctxwk{A}{B}}
 \end{align*}
@@
 $A$ to $B$, assigning $y$ to every term of $A$.
-\item[Substitution] is a morphism $\tfsubst^x:\mftyp{\mfM}{\ctxext{\Gamma}A}
+\item[Substitution] is a morphism $\tfsubst^x:\mftyp{\mfM}{\ctxext{\Gamma}{A}}
 \to\mftyp{\mfM}{\Gamma}$ for every $x:A$ in context $\Gamma$, so we must have
 \begin{align*}
-\tfsubst^x_0 & : \mftyp{\mfM}{\ctxext{\Gamma}A}\to\mftyp{\mfM}{\Gamma}
+\tfsubst^x_0 & : \mftyp{\mfM}{\ctxext{\Gamma}{A}}\to\mftyp{\mfM}{\Gamma}
 \intertext{which is denoted by $\subst{x}{\blank}$,}
-\mftyp{\tfsubst^x}{P} & : \mftyp{\mfM}{\ctxext({\Gamma}{A})P}\to\mftyp{\mfM}{\ctxext{\Gamma}\subst{x}{P}}\\
+\mftyp{\tfsubst^x}{P} & : \mftyp{\mfM}{\ctxext{{\Gamma}{A}}{P}}\to\mftyp{\mfM}{\ctxext{\Gamma}{\subst{x}{P}}}\\
 \terms[\tfsubst^x]{P} & : \terms{P}\to\terms{\subst{x}{P}}
@@
 \item A type $\mprd{A}{P}$ for every type $A$ in context $\Gamma$ and every type
-$P$ in context $\ctxext{\Gamma}A$.
+$P$ in context $\ctxext{\Gamma}{A}$.
 \begin{defn}
@@
 \item An equivalence $\lambda:\eqv{\terms{P}}{\terms{\mprd{A}{P}}}$ for every
-family $P$ over $A$ in context $\Gamma$. When $\ctxext{\Gamma}A\vdash 
+family $P$ over $A$ in context $\Gamma$. When $\ctxext{\Gamma}{A}\vdash 
 u:P$, we call $\lambda(u)$ the \emph{$\lambda$-abstraction of $u$.}
@@
 %\item A function $\tfev:\terms{\mprd{A}{P}}\to\prd{x:\terms{A}}\terms{\subst{x}{P}}$.
-\item A term $\pi^A:\ctxext{\Gamma}A\to \Gamma$ for every type $A$ in context
+\item A term $\pi^A:\ctxext{\Gamma}{A}\to \Gamma$ for every type $A$ in context
 $\Gamma$.
-\item A term $\iota^x:\ctxext{\Gamma}\subst{x}{P}\to\ctxext({\Gamma}{A})P$
+\item A term $\iota^x:\ctxext{\Gamma}{\subst{x}{P}}\to\ctxext{{\Gamma}{A}}{P}$
 for every family $P$ of types over $A$ in context $\Gamma$ and every
@@
 {\terms{\msm{A}{P}}}$.
-\item A family $\idtypevar{A}$ over $\ctxwk{A}{A}$ in context $\ctxext{\Gamma}A$ for
+\item A family $\idtypevar{A}$ over $\ctxwk{A}{A}$ in context $\ctxext{\Gamma}{A}$ for
 every type $A$ in context $\Gamma$ and a term of type.
@@
 $\msm{A}{\ctxwk{A}{B}}$ by $A\times B$. By the end of the current definition
-there will be an identification $\id{\ctxext({\Gamma}{A})P}
-{\ctxext{\Gamma}\msm{A}{P}}$.
+there will be an identification $\id{\ctxext{{\Gamma}{A}}{P}}
+{\ctxext{\Gamma}{\msm{A}{P}}}$.
 
-There is a term $\delta:\ctxhom{\ctxext{\Gamma}A}{\ctxext({\Gamma}{A})\ctxwk{A}{A}}$
+There is a term $\delta:\ctxhom{\ctxext{\Gamma}{A}}{\ctxext{{\Gamma}{A}}{\ctxwk{A}{A}}}$
 defined by...
 \end{rmk}
-\item For any family $Q$ over $\idtypevar{A}$ in context $\ctxext({\Gamma}{A})\ctxwk{A}{A}$ an equivalence
+\item For any family $Q$ over $\idtypevar{A}$ in context $\ctxext{{\Gamma}{A}}{\ctxwk{A}{A}}$ an equivalence
 $J:\eqv{\terms{Q}}{}$
@@
 \begin{tikzcd}
-\ctxext({\Gamma}{A})P \ar{r}{\pi^P} & \ctxext{\Gamma}A \ar{d}{\pi^A}\\
-\ctxext{\Gamma}\subst{x}{P} \ar{u}{\iota^x} \ar{r}[swap]{\pi^{\subst{x}{P}}} & \Gamma
+\ctxext{{\Gamma}{A}}{P} \ar{r}{\pi^P} & \ctxext{\Gamma}{A} \ar{d}{\pi^A}\\
+\ctxext{\Gamma}{\subst{x}{P}} \ar{u}{\iota^x} \ar{r}[swap]{\pi^{\subst{x}{P}}} & \Gamma
 \end{tikzcd}
@@
 \begin{comment}
-\item For $\Gamma:\ctx$, $\Gamma\vdash A:\type$, $\ctxext{\Gamma}A\vdash P
-:\type$ and $\ctxext({\Gamma}{A})P\vdash Q:\type$, $x:A$ and $u:P$
+\item For $\Gamma:\ctx$, $\Gamma\vdash A:\type$, $\ctxext{\Gamma}{A}\vdash P
+:\type$ and $\ctxext{{\Gamma}{A}}{P}\vdash Q:\type$, $x:A$ and $u:P$
 we desire that $\typefont{nat{-}ev{-}subst}(Q,u,x):\id{Q[\iota^x][\tfev(u,x)]}{Q[u][x]}$.
@@
 Let $P$ be a family over $A$ in context $\Gamma$. Then we can weaken
-$\ctxwk{A}{P}$, which is a family over $\ctxwk{A}{A}$ in context $\ctxext{\Gamma}A$.
+$\ctxwk{A}{P}$, which is a family over $\ctxwk{A}{A}$ in context $\ctxext{\Gamma}{A}$.
 We also have the family $\ctxwk{\ctxwk{A}{A}}{P}$ over $\ctxwk{A}{A}$ in context

### 2013-10-24 article-models-eg.tex
putting the new macros \ctxext, \ctxwk and \subst into use

@@
 When $A$ is a type in context
-$\Gamma$ we define the context extension $\ctxext{\Gamma}A$ to be
-$\sm{i:\Gamma}A(i)$. Note that context extension $\ctxext{\Gamma}\blank$
+$\Gamma$ we define the context extension $\ctxext{\Gamma}{A}$ to be
+$\sm{i:\Gamma}A(i)$. Note that context extension $\ctxext{\Gamma}{\blank}$
 can be seen acting not only as a function from the context of the model
@@
 and terms: when $A$ is a type in context $\Gamma$ we may take the identity map
-from $\mftyp{\Gamma\to\UU}{A}\to\mftyp{\UU}{\ctxext{\Gamma}A}$ because
-$\mftyp{\Gamma\to\UU}{A}$ is taken to be $\mftyp{\UU}{\ctxext{\Gamma}A}$;
-thus, $\ctxext{\Gamma}P\defeq P$ for every family $P$ over $A$ in context 
+from $\mftyp{\Gamma\to\UU}{A}\to\mftyp{\UU}{\ctxext{\Gamma}{A}}$ because
+$\mftyp{\Gamma\to\UU}{A}$ is taken to be $\mftyp{\UU}{\ctxext{\Gamma}{A}}$;
+thus, $\ctxext{\Gamma}{P}\defeq P$ for every family $P$ over $A$ in context 
 $\Gamma$. Likewise, when $P$ is a type in context $A$ in the model 
 $\Gamma\to\UU$ then context extension should act on the terms of $P$ via a
-function $\terms{P}\to\terms{\ctxext{\Gamma}P}$, which we take to be
+function $\terms{P}\to\terms{\ctxext{\Gamma}{P}}$, which we take to be
 the identity map once more.
@@
 has it's own notion of context extension: when $P$ is a type in context $A$ in
-the model $\mftyp{\UU}{\Gamma}$ then $\ctxext{A}P$ is the family
+the model $\mftyp{\UU}{\Gamma}$ then $\ctxext{A}{P}$ is the family
 $\lam{i}\sm{x:A(i)}P(i,x):\Gamma\to\UU$. Also the context extension of
@@
 \begin{equation*}
-\inference{\Gamma\vdash f:\mprd{A}{P}}{\ctxext{\Gamma}A\vdash\tfev(f):P}
+\inference{\Gamma\vdash f:\mprd{A}{P}}{\ctxext{\Gamma}{A}\vdash\tfev(f):P}
 \end{equation*}
@@
 but this does not describe the terms of $\mprd{A}{P}$ by any means. Moreover,
-since we have implemented substitution, we obtain from $\ctxext{\Gamma}A\vdash\tfev(f):P$ 
+since we have implemented substitution, we obtain from $\ctxext{\Gamma}{A}\vdash\tfev(f):P$ 
 and $\Gamma\vdash x:A$ a term $\Gamma\vdash\subst{x}{\tfev(f)}:\subst{x}{P}$ and
@@
 \begin{defn}
-Suppose that $\Gamma\vdash A:\mfGraph$. Then we define the graph $\ctxext{\Gamma}A$
+Suppose that $\Gamma\vdash A:\mfGraph$. Then we define the graph $\ctxext{\Gamma}{A}$
 by
 \begin{align*}
-\ctxext{\Gamma}A_0 & \defeq \sm{i:\Gamma_0}A_0(i)\\
-\ctxext{\Gamma}A_1(\pairr{i,x},\pairr{j,y}) & \defeq \sm{q:\Gamma_1(i,j)}A_1(q,x,y).
+\ctxext{\Gamma}{A}_0 & \defeq \sm{i:\Gamma_0}A_0(i)\\
+\ctxext{\Gamma}{A}_1(\pairr{i,x},\pairr{j,y}) & \defeq \sm{q:\Gamma_1(i,j)}A_1(q,x,y).
 \end{align*}
@@
 Suppose that $\Gamma\vdash A:\mfGraph$ and $\Gamma\vdash B:\mfGraph$. Then we
-define the graph $\ctxwk{A}{B}$ in context $\ctxext{\Gamma}A$ by
+define the graph $\ctxwk{A}{B}$ in context $\ctxext{\Gamma}{A}$ by
 \begin{align*}
@@
 \begin{rmk}
-We note that the function $\ctxext{\unit}\blank$ is an equivalence
+We note that the function $\ctxext{\unit}{\blank}$ is an equivalence
 from $\mftyp{\mfGraph}{\unit}$ to $\tfGraph$. It's inverse is the function which maps
@@
 \begin{defn}
-Suppose that $\Gamma\vdash A:\mfGraph$. Then we define the graph $\ctxext{\Gamma}A$
+Suppose that $\Gamma\vdash A:\mfGraph$. Then we define the graph $\ctxext{\Gamma}{A}$
 by
 \begin{align*}
-(\ctxext{\Gamma}A)_0 & \defeq \sm{i:\Gamma_0}A_0(i)\\
-(\ctxext{\Gamma}A)_1(\pairr{i,x},\pairr{j,y}) & \defeq \sm{q:\Gamma_1(i,j)}A_1(q,x,y).
+(\ctxext{\Gamma}{A})_0 & \defeq \sm{i:\Gamma_0}A_0(i)\\
+(\ctxext{\Gamma}{A})_1(\pairr{i,x},\pairr{j,y}) & \defeq \sm{q:\Gamma_1(i,j)}A_1(q,x,y).
 \end{align*}
-There is a morphism $\pi^A:\ctxhom{\ctxext{\Gamma}A,\Gamma}$ given by
+There is a morphism $\pi^A:\ctxhom{\ctxext{\Gamma}{A},\Gamma}$ given by
 \begin{align*}
@@
 \begin{rmk}
-The construction of $\ctxext{\Gamma}A$ is similar to the Grothendieck
+The construction of $\ctxext{\Gamma}{A}$ is similar to the Grothendieck
 construction of the category of elements for a presheaf.
@@
 Let $\Gamma\vdash A:\mfGraph$. A family $P$ of graphs over $A$ is by
-definition a graph $P$ in the context $\ctxext{\Gamma}A$. More explicitly, a family
+definition a graph $P$ in the context $\ctxext{\Gamma}{A}$. More explicitly, a family
 $P$ of graphs over $A$ in context $\Gamma$ consists of
@@
 P_1 & : \prd*{i,j:\Gamma_0}{q:\Gamma_1(i,j)}*{x:A_0(i)}*{y:A_0(j)} A_1(q,x,y)\to P_0(x)\to P_0(y)\to\type
-%P_1 & : \prd{\pairr{i,x},\pairr{j,y}:(\ctxext{\Gamma}A)_0}(\ctxext{\Gamma}A)_1(\pairr{i,x},\pairr{j,y})\to P_0(i,x)\to P_0(j,y)\to\type.
+%P_1 & : \prd{\pairr{i,x},\pairr{j,y}:(\ctxext{\Gamma}{A})_0}(\ctxext{\Gamma}{A})_1(\pairr{i,x},\pairr{j,y})\to P_0(i,x)\to P_0(j,y)\to\type.
 \end{align*}
@@
 For any two graphs $A$ and $B$ in a context $\Gamma$, we have the graph
-$B[\pi^A]$ in the context $\ctxext{\Gamma}A$, which we may just write
+$B[\pi^A]$ in the context $\ctxext{\Gamma}{A}$, which we may just write
 as $B$. Therefore, we can
@@
 Let $A$ be a graph in context $\Gamma$. We define the family $\idtypevar{A}$ over $\ctxwk{A}{A}$ in
-context $\ctxext{\Gamma}A$ by
+context $\ctxext{\Gamma}{A}$ by
 \begin{align*}
@@
 Let $D$ be a family over $\idtypevar{A}$ in context 
-$\ctxext({\Gamma}{A})\ctxwk{A}{A}$. Then we have the family
+$\ctxext{{\Gamma}{A}}{\ctxwk{A}{A}}$. Then we have the family
 $\subst{\idfunc[A]}{D}$ over $\subst{\idfunc[A]}{\idtypevar{A}}$ in
-context $\ctxext{\Gamma}A$ given by
+context $\ctxext{\Gamma}{A}$ given by
 \begin{align*}
@@
 \item $f[i]:A[i]\to B[i]$ is an equivalence of graphs for every term $i:\Gamma$.
-\item $\ctxext{\Gamma}f:\ctxext{\Gamma}A\to\ctxext{\Gamma}B$ is an equivalence of graphs.
+\item $\ctxext{\Gamma}{f}:\ctxext{\Gamma}{A}\to\ctxext{\Gamma}{B}$ is an equivalence of graphs.
 \item Both $f_0(i)$ and $f_1(q,x,y)$ are always equivalences.
@@
 \begin{equation*}
-\ctxext{\Gamma}\msm{A}{P}\simeq\ctxext({\Gamma}{A})P
+\ctxext{\Gamma}\msm{A}{P}\simeq\ctxext{{\Gamma}{A}}{P}
 \end{equation*}

### 2013-10-24 article-models-intro.tex
putting the new macros \ctxext, \ctxwk and \subst into use

@@
 This model is required to be \emph{definitionally equal to} the model of types
-in the context $\ctxext{\Gamma}A$, the context extension of $\Gamma$ and $A$.
+in the context $\ctxext{\Gamma}{A}$, the context extension of $\Gamma$ and $A$.
 In this way we protect ourselves from the need to dig an infinitely deep structure

### 2013-10-24 article-models-main.tex
resolve issue #7
putting the new macros \ctxext, \ctxwk and \subst into use

@@
 
-\begin{comment}
-\section*{Introduction\phantomsection\addcontentsline{toc}{section}{Introduction}}
-This paper contributes to the univalent homotopy type theory~\cite{kapulkin2012univalence,awodey2012type,pelayo2012homotopy}
-program which connects and extends a number of topics such as type theory, (elementary) topos
-theory~\cite{MacLaneMoerdijk,johnstone:elephant}, homotopy theory, higher category theory and higher topos
-theory~\cite{rezk2010toposes,lurie2009higher}.
-A key missing ingredient is a first-order (`elementary') definition of a higher topos. It has been
-conjectured~\cite{awodey2012type,shulman2012univalence} that homotopy type theory can be used as the internal language, and thus may
-provide an elementary definition of a higher topos; see also~\cite{LumsdaineWarren}.
+\input{article-models-intro}
 
-A main ingredient in the theory of higher toposes is the descent property, which
-is described in~\cite{rezk2010toposes,lurie2009higher}. The (weak) descent property
-of an ordinary topos $\mathcal{E}$ is the property that
-\begin{enumerate}
-\item For any family $\{X_i\}_{i\in I}$ of objects of $\mathcal{E}$ with coproduct
-$X:=\coprod_{i\in I}X_i$ and any function $f:Y\to X$ we get an isomorphism
-\begin{equation*}
-Y\cong\coprod_{i\in I} Y\times_X X_i
-\end{equation*}
-given by the canonical map.
-\item Conversely, for any family $\{f_i:X_i\to Y_i\}_{i\in I}$ of maps inducing
-$f:\coprod_{i\in I}X_i\to\coprod_{i\in I}Y_i$, the canonical map determines an
-isomporphism
-\begin{equation*}
-Y_i\cong Y\times_X X_i.
-\end{equation*}
-\item
-\item
-\end{enumerate}
-In model toposes, the descent property can be strengthened considerably, but
-ordinary limits and colimits are to be replaced by \emph{homotopy} limits and
-colimits. Since the only notion of limits and colimits in type theory is that
-of homotopy limit and colimit, this is no restriction for us. The descent
-property of model toposes as formulated by Rezk~\cite{rezk2010toposes} is: Let
-$X:\mathbf{I}\to\mathcal{E}$ be a functor from a small category into a
-model topos $\mathcal{E}$, with homotopy colimit $\bar{X}$. Then we have
-\begin{enumerate}
-\item Let $\bar{Y}$ be an object and let $f:\bar{Y}\to\bar{X}$ 
-be any map in $\mathcal{E}$. Define
-the functor $Y:\mathbf{I}\to\mathcal{E}$ by $Y(i):=Y\times_{\bar{X}} X_i$. Then
-the canonical map $\typefont{hocolim}(Y)\to\bar{Y}$ is a weak equivalence.
-\item Let $Y:\mathbf{I}\to\mathcal{E}$ be a functor and let $f:Y\to X$ be an
-equifibered transformation from $Y$ to $X$. Define $\bar{f}:\bar{Y}\to\bar{X}$
-be the homotopy colimit of $f$. Then the canonical map 
-$Y_i\to \bar{Y}\times^h_{\bar{X}} X_i$ is a weak equivalence.
-\end{enumerate}
-These two properties are two sides of the same coin and in our type theoretical
-setting it is natural to stress this. Given a diagram $D$, we will be able
-to summarize the entire descent property in an equivalence between the type
-$\typefont{equiFib}(D)$ of equifibered diagrams over $D$ and the type of functions
-into $\typefont{colim}(D)$. We prove the descent property in 
-theorem~\ref{thm:descent}.
-
-We use universes ala Russell, universe polymorphism and typical ambiguity and hence create the illusion that $\type:\type$;
-see~\cite{TheBook}. There are only a few places where we need to be explicit about universe levels.
-As is common in HoTT, we will freely use the axiom of function extensionality.
-Using this axiom, all limits can be constructed in type theory~\cite{Avigad:limits}.
-We use higher inductive types (HITs) to construct co-limits; see~\cite{LumsdaineShulman,Shulman:HIT}.
-Using the univalence axiom one can show that $\type$ is an object classifier~\cite{RijkeSpitters:Sets}.
-An important feature property of $\infty$-toposes is the descent property. We prove this property in theorem~\ref{thm:descent}.
-
-In the previous paragraph we have used informal categorical terminology such as `limits' to refer to what would be interpreted as homotopy
-limits in the $\infty$-category of $\infty$-groupoids. We will continue to use this naive $\infty$-category style without any rigorous
-claims to their interpretation in the model.
-
-The paper is organized as follows. We start by defining limits and colimits of over graph. This is the most general case, as we
-do not include the possibility of including higher coherence conditions.
-We introduce diagrams and (co)limits in section~\ref{sec:diagram}. In section~\ref{sec:suspensions} we study suspensions.
-Functions with contractible mapping cones are the same as epimorphisms, as we will see in section~\ref{sec:mapping}.
-Finally, we prove the descent property.
-
-\subsection{Notation}
-A typical way we write a type in this article is:
-\begin{equation}\label{eq:notation_example}
-\prd{x,y:A}{p:B(x,y)}C(x,y,p)\to D(x,y,p).
-\end{equation}
-Assuming that $A$, $B$, $C$ and $D$ have been introduced, this type denotes a
-dependent product, where the product ranges over $x,y:A$ and $p:B(x,y)$ and
-the factors are the (non-dependent) function types $C(x,y,p)\to D(x,y,p)$. A
-more common notation for the same type in mathematics would be
-\begin{equation*}
-\prod_{x,y:A}\prod_{p:B(x,y)}C(x,y,p)\to D(x,y,p),
-\end{equation*}
-like a big product of function spaces. Nevertheless we have chosen to use this
-less common notation, because the types $A$ or $B$ can become quite complex.
-% and formulas would become messy as as complex information starts `dancing
-% around'.
-
-Another thing we notice in formula \eqref{eq:notation_example} is the use of
-curly braces around $x,y:A$. This indicates that whenever we have a term
-\begin{equation*}
-f:\prd{x,y:A}{p:B(x,y)}C(x,y,p)\to D(x,y,p),
-\end{equation*}
-we will omit reference to $x$ and $y$ whenever $f$ is applied to $x,y:A$ and
-$p:B(x,y)$. In other words, we will write down $f(p):C(x,y,p)\to D(x,y,p)$
-instead of $f(x,y,p)$. We will do this whenever it is possible to retrieve
-the terms $x$ and $y$ from the term $p$, and in many similar situations like
-this. Back to formula \autoref{eq:notation_example}, it is quite possible 
-that we would have used the curly braces in the introduction
-of $C$ and $D$ as well, and if we had done that we would have written $C(p)$
-and $D(p)$.
-\end{comment}
+\input{article-models-eg}
 

### 2013-10-25 article-models-defn.tex
moved content to the new file

@@
-\section{A variable-free presentation of type theory}
-We have seen in the example of $\UU$ that we will not need to refer to variables
-when we are manipulating contexts. In fact, it is more natural not to.
-This suggests that there is a presentation
-of type theory where contexts do not mention variables. The first small gain is that such a presentation
-would not be burdened with comments about variables being bounded or not, or fresh or
-not occuring at all.
-
-\subsection{The basic judgments}
-The judgments we can make are the standard six basic judgments involving contexts, types and terms and
-judgmental equality in each of these three cases.
-\begin{align*}
-\jctx*{\Gamma} & \jctxeq*{\Gamma}{\Gamma'}\\
-\jtype*{\Gamma}{A} & \jtypeeq*{\Gamma}{A}{B}\\
-\jterm*{\Gamma}{A}{x} & \jtermeq*{\Gamma}{A}{x}{y}.
-\end{align*}
-
-In contrast with standard practise, we don't assume that there is an empty context. The idea is that
-if we interpret only the syntax of contexts, types and terms with extension, weakening and
-substitution, we do not have a full model of type theory but we still have a 
-structure resembling a category and categories need not have terminal objects. Therefore,
-the rules that follow will describe only how to manipulate contexts, types and terms.
-Of most of the inferences that we give there are two versions: one introducing an operation, the other asserting that the operation in question preserves judgmental equality.
-
-\subsection{The basic rules for judgmental equality}
-The rules for judgmental equality establish that it is an equivalence relation
-in all three cases (contexts, types and terms).
-%\begingroup
-%\renewcommand*{\arraystretch}{3}
-%\begin{equation*}
-%\begin{array}{ccc}
-\begin{infarray}{ccc}
-\inference{\jctx{\Gamma}}{\jctxeq{\Gamma}{\Gamma}} & \inference{\jctxeq{\Gamma}{\Delta}}{\jctxeq{\Delta}{\Gamma}} & \inference{\jctxeq{\Gamma}{\Delta}\qquad\jctxeq{\Delta}{\Theta}}{\jctxeq{\Gamma}{\Theta}}\\
-\inference{\jtype{\Gamma}{A}}{\jtypeeq{\Gamma}{A}{A}} &
-\inference{\jtypeeq{\Gamma}{A}{B}}{\jtypeeq{\Gamma}{B}{A}} & 
-\inference{\jtypeeq{\Gamma}{A}{B}\qquad\jtypeeq{\Gamma}{B}{C}}{\jtypeeq{\Gamma}{A}{C}}\\
-\inference{\jterm{\Gamma}{A}{x}}{\jtermeq{\Gamma}{A}{x}{x}} & 
-\inference{\jtermeq{\Gamma}{A}{x}{y}}{\jtermeq{\Gamma}{A}{y}{x}} &
-\inference{\jtermeq{\Gamma}{A}{x}{y}\qquad\jtermeq{\Gamma}{A}{y}{z}}{\jtermeq{\Gamma}{A}{x}{z}}
-\end{infarray}
-%\end{array}
-%\end{equation*}
-\begin{infarray}{cc}
-\inference{\jctxeq{\Gamma}{\Delta}\qquad\jtype{\Gamma}{A}}{\jtype{\Delta}{A}}
-& \inference{\jctxeq{\Gamma}{\Delta}\qquad\jtypeeq{\Gamma}{A}{B}}{\jtypeeq{\Delta}{A}{B}}\\
-\inference{\jctxeq{\Gamma}{\Delta}\qquad\jterm{\Gamma}{A}{x}}{\jterm{\Delta}{A}{x}}
-& \inference{\jctxeq{\Gamma}{\Delta}\qquad\jtermeq{\Gamma}{A}{x}{y}}{\jtermeq{\Delta}{A}{x}{y}}\\
-\inference{\jtypeeq{\Gamma}{A}{B}\qquad \jterm{\Gamma}{A}{x}}{\jterm{\Gamma}{B}{x}}
-& \inference{\jtypeeq{\Gamma}{A}{B}\qquad\jtermeq{\Gamma}{A}{x}{y}}{\jtermeq{\Gamma}{B}{x}{y}}
-\end{infarray}
-
-\subsection{Extension}
-We introduce extension which not only extends a context $\Gamma$ and a type
-$A$ over it to a context $\ctxext{\Gamma}{A}$, but which also extends a type $A$
-in context $\Gamma$ and a family $P$ over it to a type $\ctxext{A}{P}$ in context
-$\Gamma$. We do this to ensure that all of type theory can be done in a context.
-For instance, we could say (1) that a context in context $\Gamma$ is the same thing
-as a type in context $\Gamma$; (2) When $A$ is a context in this sense, a type in
-context $A$ is the same thing as a family $P$ over $A$ and (3) when $P$ is a type
-in context $A$ in this sense, a term of $P$ keeps its original meaning.
-
-Note that by introducing extension on the level of types and families,
-we introduce $\Sigma$-types at a very early stage.
-However, we need substitution to make this precise.
-\begin{infarray}{cc}
-\inference{\jtype{\Gamma}{A}}{\jctx{\ctxext{\Gamma}{A}}}
-& \inference{\jctxeq{\Gamma}{\Delta}\qquad\jtypeeq{\Gamma}{A}{B}}{\jctxeq{\ctxext{\Gamma}{A}}{\ctxext{\Delta}{B}}}\\
-\inference{\jtype{\ctxext{\Gamma}{A}}{P}}{\jtype{\Gamma}{\ctxext{A}{P}}}
-& \inference{\jtypeeq{\Gamma}{A}{B}\qquad\jtypeeq{\ctxext{\Gamma}{A}}{P}{Q}}{\jtypeeq{\Gamma}{\ctxext{A}{P}}{\ctxext{B}{Q}}}
-\end{infarray}
-
-\subsection{Weakening}
-We first define weakening by a context $\Gamma$. Since weakening by $\Gamma$
-should in principle be a functor, it acts on contexts, types and terms alike.
-Note that it is because of the weakening $\ctxwk{\Gamma}{\Delta}$ of a context
-$\Delta$ by $\Gamma$ that we can speak of context morphisms from $\Gamma$ to $\Delta$: they are the terms
-of $\ctxwk{\Gamma}{\Delta}$.
-\begin{infarray}{cc}
-\inference{\jctx{\Gamma}\qquad\jctx{\Delta}}{\jtype{\Gamma}{\ctxwk{\Gamma}{\Delta}}} 
-& \inference{\jctxeq{\Gamma}{\Gamma'}\qquad\jctxeq{\Delta}{\Delta'}}{\jtypeeq{\Gamma}{\ctxwk{\Gamma}{\Delta}}{\ctxwk{\Gamma'}{\Delta'}}}\\
-\inference{\jctx{\Gamma}\qquad\jtype{\Delta}{B}}{\jtype{\ctxext{\Gamma}{\ctxwk{\Gamma}{\Delta}}}{\ctxwk{\Gamma}{B}}} & \inference{\jctxeq{\Gamma}{\Gamma'}\qquad\jtypeeq{\Delta}{B}{B'}}{\jtypeeq{\ctxext{\Gamma}{\ctxwk{\Gamma}{\Delta}}}{\ctxwk{\Gamma}{B}}{\ctxwk{\Gamma'}{B'}}}\\
-\inference{\jctx{\Gamma}\qquad\jterm{\Delta}{B}{y}}{\jterm{\ctxext{\Gamma}{\ctxwk{\Gamma}{\Delta}}}{\ctxwk{\Gamma}{B}}{\ctxwk{\Gamma}{y}}} 
-& \inference{\jctxeq{\Gamma}{\Gamma'}\qquad\jtermeq{\Delta}{B}{y}{y'}}{\jtermeq{\ctxext{\Gamma}{\ctxwk{\Gamma}{\Delta}}}{\ctxwk{\Gamma}{B}}{\ctxwk{\Gamma}{y}}{\ctxwk{\Gamma'}{y'}}} 
-\end{infarray}
-
-A weakening operation is also defined for types. When $A$ and $B$ are both types
- in context $\Gamma$, the weakened type $\ctxwk{A}{B}$ in context $\ctxext{\Gamma}{A}$
- is the family which `doesn't really depend on $A$'. The terms of $\ctxwk{A}{B}$
- are the functions from $A$ to $B$. Likewise, the terms $\ctxwk{A}{y}$ are the
- constant maps at $y$ from $A$ to $B$, for $y:B$.
-\begin{infarray}{cc}
-\inference{\jtype{\Gamma}{A}\qquad\jtype{\Gamma}{B}}{\jtype{\ctxext{\Gamma}{A}}{\ctxwk{A}{B}}}
-& \inference{\jtypeeq{\Gamma}{A}{A'}\qquad\jtypeeq{\Gamma}{B}{B'}}{\jtypeeq{\ctxext{\Gamma}{A}}{\ctxwk{A}{B}}{\ctxwk{A'}{B'}}}\\
-\inference{\jtype{\Gamma}{A}\qquad\jtype{\ctxext{\Gamma}{B}}{Q}}
-{\jtype{\ctxext({\Gamma}{A})\ctxwk{A}{B}}{\ctxwk{A}{Q}}}
-& \inference{\jtypeeq{\Gamma}{A}{A'}\qquad\jtypeeq{\ctxext{\Gamma}{B}}{Q}{Q'}}
-{\jtypeeq{\ctxext({\Gamma}{A})\ctxwk{A}{B}}{\ctxwk{A}{Q}}{\ctxwk{A'}{Q'}}}\\
-\inference{\jtype{\Gamma}{A}\qquad\jterm{\Gamma}{B}{y}}{\jterm{\ctxext{\Gamma}{A}}{\ctxwk{A}{B}}{\ctxwk{A}{y}}}
-& \inference{\jtypeeq{\Gamma}{A}{A'}\qquad\jtermeq{\Gamma}{B}{y}{y'}}{\jtermeq{\ctxext{\Gamma}{A}}{\ctxwk{A}{B}}{\ctxwk{A}{y}}{\ctxwk{A'}{y'}}}
-\end{infarray}
-
-\subsection{Substitution}
-\begin{infarray}{cc}
-\inference{\jterm{\Gamma}{A}{x}\qquad\jtype{\ctxext{\Gamma}{A}}{P}}{\jtype{\Gamma}{\subst{x}{P}}}
-& \inference{\jtermeq{\Gamma}{A}{x}{x'}\qquad\jtypeeq{\ctxext{\Gamma}{A}}{P}{P'}}{\jtypeeq{\Gamma}{\subst{x}{P}}{\subst{x'}{P'}}}\\
-\inference{\jterm{\Gamma}{A}{x}\qquad\jtype{\ctxext{{\Gamma}{A}}{P}}{Q}}{\jtype{\ctxext{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}}
-& \inference{\jtermeq{\Gamma}{A}{x}{x'}\qquad\jtypeeq{\ctxext{{\Gamma}{A}}{P}}{Q}{Q'}}{\jtypeeq{\ctxext{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}{\subst{x'}{Q'}}}\\
-\inference{\jterm{\Gamma}{A}{x}\qquad\jterm{\ctxext{\Gamma}{A}}{P}{f}}{\jterm{\Gamma}{\subst{x}{P}}{\subst{x}{f}}}
-& \inference{\jtermeq{\Gamma}{A}{x}{x'}\qquad\jtermeq{\ctxext{\Gamma}{A}}{P}{f}{f'}}{\jtermeq{\Gamma}{\subst{x}{P}}{\subst{x}{f}}{\subst{x'}{f'}}}\\
-\inference{\jterm{\Gamma}{A}{x}\qquad\jterm{\ctxext{{\Gamma}{A}}{P}}{Q}{g}}{\jterm{\ctxext{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}{\subst{x}{g}}}
-& \inference{\jtermeq{\Gamma}{A}{x}{x'}\qquad\jtermeq{\ctxext{{\Gamma}{A}}{P}}{Q}{g}{g'}}{\jtermeq{\ctxext{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}{\subst{x}{g}}{\subst{x'}{g'}}}
-\end{infarray}
-
-\subsection{Compatibility of extension, weakening and substitution}
-We describe the compatibility of extension, weakening and substitution. These
-compatibility rules are formulated in a weak sense.
-
-Extension is compatible with extension:
-\begin{infarray}{c}
-\inference
-{ \jtype{\ctxext{\Gamma}{A}}{P}
-}{\jterm
-  {\ctxext{\Gamma}{{A}{P}}}
-  {\ctxwk{\ctxext{\Gamma}{{A}{P}}}{(\ctxext{{\Gamma}{A}}{P})}}{\typefont{extext}_0(A,P)}
-}\\
-\inference
-{ \jtype{\ctxext{{\Gamma}{A}}{P}}{Q}
-}{\jterm
-  {\ctxext{\Gamma}{{A}{{P}{Q}}}}
-  {\ctxwk{\ctxext{A}{{P}{Q}}}{(\ctxext{{A}{P}}{Q})}}
-  {\typefont{extext}_1(A,P,Q)}
-}
-\end{infarray}
-
-Weakening is compatible with extension:
-\begin{infarray}{c}
-\inference{
-  \jctx{\Gamma}
-  \qquad
-  \jtype{\Delta}{B}}
-  {\jterm
-    {\ctxext{\Gamma}{\ctxwk{\Gamma}{(\ctxext{\Delta}{B})}}}
-    {\ctxwk{{\Gamma}{(\ctxext{\Delta}{B})}}{\ctxext{(\ctxwk{\Gamma}{\Delta})}{(\ctxwk{\Gamma}{B})}}}
-    {\typefont{wkext}_0(\Gamma,B)}}
-\end{infarray}
-
-Substitution is compatible with substitution:
-\begin{infarray}{c}
-\inference{\jtype{\ctxext{{\Gamma}{A}}{P}}{Q}\qquad\jterm{\Gamma}{A}{x}\qquad\jterm{\ctxext{\Gamma}{A}}{P}{f}}
-{\jtypeeq{\Gamma}{\subst{x}{\subst{f}{Q}}}{\subst{{x}{f}}{\subst{x}{Q}}}}
-\end{infarray}
-
-\subsection{More stuff in the type theory}
-\begin{infarray}{c}
-\inference{\jtype{\Gamma}{A}}{\jterm{\ctxext{\Gamma}{A}}{\ctxwk{A}{A}}{\idfunc[A]}}
-\end{infarray}
-
-\subsection{Leibniz equality on terms}
-Occasionally we shall need a weaker version of equality than judgmental equality.
-Our choice here is Leibniz equality, since it implies propositional equality when
-identity types are present.
-
-Let $P$ be a family over $A$ in context $\Gamma$. Then we can weaken
-$\ctxwk{A}{P}$, which is a family over $\ctxwk{A}{A}$ in context $\ctxext{\Gamma}{A}$.
-We also have the family $\ctxwk{{A}{A}}{P}$ over $\ctxwk{A}{A}$ in context
-$\ctxext{\Gamma}{A}$ and thus we have the inference
-\begin{equation*}
-\inference{\jtype{\ctxext{\Gamma}{A}}{P}}{\jtype{\ctxext(({\Gamma}{A}){\ctxwk{A}{A}})\ctxwk{A}{P}}{\ctxwk{{A}{P}}{(\ctxwk{{A}{A}}{P})}}}
-\end{equation*}
-This family plays the role of the family $\jtype{\Gamma,x,y:A}{P(x)\to P(y)}$
-in the type theory of \cite{TheBook}.
-
-\subsection{Trivial cofibrations and weak equivalences of types}
-We describe a relation between types that expresses when they are weakly equivalent.
-Weak equivalence is introduced because we need a weaker notion of judgmental 
-equality which also makes sense when identity types are not present, since that
-would allow us to state that context extension, weakening and substitution
-commute with each other.
-
-A term $f:\ctxwk{A}{B}$ is a trivial cofibration if it has the
-property that for any fibration $Q$ over $B$,
-to find a section of $Q$ it suffices to find a section of the fibration
-$f^\ast Q$ over $A$. In our type theoretical setting, the rôle of fibrations
-is played by families, the rôle of the function type $A\to B$ is played by
-$\ctxwk{A}{B}$ and our version of the pullback $f^\ast Q$ is $\subst{f}{(\ctxwk{A}{Q})}$.
-
-\begin{defn}
-Let $f$ be a term of $\ctxwk{A}{B}$ in context $\Gamma$.
-\begin{enumerate}
-\item For a family $Q$ over $B$ in context $\Gamma$ we define $f^\ast Q\jdeq\subst{f}{(\ctxwk{A}{Q})}$.
-\item For a term $g$ of $Q$ in context $\ctxext{\Gamma}{B}$ we define $f^\ast g\jdeq\subst{f}{(\ctxwk{A}{g})}$.
-\end{enumerate} 
-\end{defn}
-\begin{rmk}
-The type $f^\ast(\ctxwk{B}{C})$ in context $\ctxext{\Gamma}{A}$ is can be viewed as the
-type of functions from $A$ to $C$ which factor through $f$. Thus there should be
-a function from $f^\ast(\ctxwk{B}{C})$ to $\ctxwk{A}{C}$.
-\end{rmk}
-
-\begin{rmk}
-Every term $\jterm{\ctxext{\Gamma}{A}}{\ctxwk{A}{B}}{f}$ allows us to infer the following:
-\begin{equation*}
-\inference{\jtype{\ctxext{\Gamma}{B}}{Q}}{\jtype{\ctxext{\Gamma}{A}}{f^\ast Q}}
-\qquad
-\inference{\jterm{\ctxext{\Gamma}{B}}{Q}{g}}{\jterm{\ctxext{\Gamma}{A}}{f^\ast Q}{f^\ast g}}
-\end{equation*}
-\end{rmk}
-
-\begin{defn}
-A term $f:\ctxwk{A}{B}$ in context $\ctxext{\Gamma}{A}$ is said to be a trivial
-cofibration if we can infer
-\begin{equation*}
-\inference{\jterm{\ctxext{\Gamma}{A}}{f^\ast Q}{t}}{\jterm{\ctxext{\Gamma}{B}}{Q}{\tilde{t}}}\qquad
-\inference{\jterm{\ctxext{\Gamma}{A}}{f^\ast Q}{t}}{\jtermeq{\ctxext{\Gamma}{A}}{f^\ast Q}{f^\ast \tilde{t}}{t}}
-\end{equation*}
-{\color{red}This statement should be reformulated so that it only involves a single judgment...
-but I don't see directly how to do that.}
-\end{defn}
-
-We have the following theorem in the type theory of \cite{TheBook}, which supports
-our claim that we may indeed speak of a trivial cofibration.
-
-\begin{thm}
-Suppose $f:A\to B$ is a function. Then $f$ is an equivalence if and only if
-for every $Q:B\to\type$ and every $g:\prd{x:A}Q(f(x))$ there is a section
-$h:\prd{y:B}Q(b)$ with the property that $h\circ f\htpy g$. 
-\end{thm}
-
-\begin{proof}
-We can first take $Q$ to be the constant family $\lam{y}A$. Furthermore, we may
-take $g\defeq\idfunc[A]$. Then we get a term of type
-\begin{equation*}
-\sm{h:B\to A}h\circ f\htpy \idfunc[A],
-\end{equation*}
-i.e.~we get a left inverse $h$ for $f$. To show that $h$ is also a right inverse
-of $f$, let $Q$ be the family $\lam{y}\id{f(h(y))}{y}$. To find a section of
-$Q$, which is the homotopy we aim for, it suffices to find a section of
-$Q\circ f$. In other words, we have to show that $\id{f(h(f(x)))}{f(x)}$ for
-every $x:A$. This follows from the fact that $h$ is a left inverse for $f$.
-
-The reverse direction is immediate.
-\end{proof}
-
-Another approach would be to define $f:\ctxwk{A}{B}$ to be left invertible
-if there is a term $\ctxext{{\Gamma}{A}}{P}$
-
-\subsection{The interactions of extension, weakening and substitution}
-
-\subsection{The basic constructors of type theory}
-\subsubsection{Dependent function types}
-\begin{infarray}{cc}
-\inference{\jtype{\ctxext{\Gamma}{A}}{P}}{\jtype{\Gamma}{\mprd{A}{P}}}
-& \inference{\jtypeeq{\ctxext{\Gamma}{A}}{P}{P'}}{\jtypeeq{\Gamma}{\mprd{A}{P}}{\mprd{A}{P'}}}\\
-\inference{\jterm{\ctxext{\Gamma}{A}}{P}{f}}{\jterm{\Gamma}{\mprd{A}{P}}{\lambda(f)}}
-& \inference{\jtermeq{\ctxext{\Gamma}{A}}{P}{f}{f'}}{\jtermeq{\Gamma}{\mprd{A}{P}}{\lambda(f)}{\lambda(f')}}\\
-\inference{\jterm{\Gamma}{\mprd{A}{P}}{g}}{\jterm{\ctxext{\Gamma}{A}}{P}{\tfev(g)}}
-& \inference{\jtermeq{\Gamma}{\mprd{A}{P}}{g}{g'}}{\jtermeq{\ctxext{\Gamma}{A}}{P}{\tfev(g)}{\tfev(g')}}\\
-\inference{\jterm{\ctxext{\Gamma}{A}}{P}{f}}{\jtermeq{\Gamma}{A}{\tfev(\lambda(f))}{f}}
-\end{infarray}
-
-With these rules we will not get the weak $\eta$-rule when identity types are present.
-So it might be better to state that $\lambda$ is a trivial cofibration
-
-\subsection{Identity types}
-We need to define the diagonal function $\jterm{\ctxext{\Gamma}{A}}{\ctxwk{A}{(\ctxext{A}{\ctxwk{A}{A}})}}{\delta}$
-\begin{infarray}{c}
-\inference{\jtype{\Gamma}{A}}{\jtype{\ctxext{{\Gamma}{A}}{\ctxwk{A}{A}}}{\idtypevar{A}}}\\
-\inference{\jtype{\Gamma}{A}}{\jterm{\ctxext{\Gamma}{A}}{\delta^\ast\idtypevar{A}}{\reflsym}}\\
-\inference{\jtype{\ctxext{{{\Gamma}{A}}{\ctxwk{A}{A}}}{\idtypevar{A}}}{P}\qquad
-\jterm{\ctxext{\Gamma}{A}}{\pairr{\delta,\reflsym}^\ast P}{d}}
-{\jterm{\ctxext{{{\Gamma}{A}}{\ctxwk{A}{A}}}{\idtypevar{A}}}{P}{J}}
-\end{infarray}
-
 \section{Internal models of type theory}
@@
 \begin{equation*}
-\mftyp{f}{\ctxext({\Gamma}{A})\ctxwk{A}{B}}\circ\tfwk^A\jdeq\tfwk^{\protect{\mftyp{f}{\Gamma}_0(A)}}\circ\mftyp{f}{\ctxext{\Gamma}{B}}
+\mftyp{f}{\ctxext{{\Gamma}{A}}{\ctxwk{A}{B}}}\circ\tfwk^A\jdeq\tfwk^{\protect{\mftyp{f}{\Gamma}_0(A)}}\circ\mftyp{f}{\ctxext{\Gamma}{B}}
 \end{equation*}
@@
 \begin{equation*}
-\terms[f]{\ctxwk{A}{B}}(\ctxwk{A}{y})\jdeq\ctxwk{\mftyp{f}\Gamma_0(A)}{(\terms[f]{B}(y))}
+\terms[f]{\ctxwk{A}{B}}{{A}{y}}\jdeq\ctxwk{\mftyp{f}\Gamma_0(A)}{(\terms[f]{B}(y))}
 \end{equation*}
@@
 over $\subst{y}{P}$ in context $\Gamma$. We could weaken $\subst{y}{P}$ subsequently by $A$, $\ctxwk{A}{A}$
-and $\idtypevar{A}$ to obtain the family $\ctxwk{\idtypevar{A}}{\ctxwk{\ctxwk{A}{A}}{\ctxwk{A}{\subst{y}{P}}}}$ over $\idtypevar{A}$ in context $\ctxext({\Gamma}{A})\ctxwk{A}{A}$. We will
+and $\idtypevar{A}$ to obtain the family $\ctxwk{\idtypevar{A}}{\ctxwk{\ctxwk{A}{A}}{\ctxwk{A}{\subst{y}{P}}}}$ over $\idtypevar{A}$ in context $\ctxext{{\Gamma}{A}}{\ctxwk{A}{A}}$. We will
 denote this family simply by $\ctxwk{\idtypevar{A}}{\subst{y}{P}}$. Likewise,
 we obtain the family $\ctxwk{\idtypevar{A}}{\ctxwk{\subst{y}{P}}{\subst{x}{P}}}$ over
-$\ctxwk{\idtypevar{A}}{\subst{y}{P}}$ in context $\ctxext(({\Gamma}{A}){\ctxwk{A}{A}})\idtypevar{A}$ and we obtain the family $\ctxwk{\idtypevar{A}}{\idtypevar{\subst{y}{P}}}$
+$\ctxwk{\idtypevar{A}}{\subst{y}{P}}$ in context $\ctxext{{{\Gamma}{A}}{\ctxwk{A}{A}}}{\idtypevar{A}}$ and we obtain the family $\ctxwk{\idtypevar{A}}{\idtypevar{\subst{y}{P}}}$
 over $\ctxwk{\idtypevar{A}}{\ctxwk{\subst{y}{P}}{\subst{y}{P}}}$. Now we

### 2013-10-25 article-models-tt.tex
moved content to the new file

\section{A variable-free presentation of type theory}
We have seen in the example of $\UU$ that we will not need to refer to variables
when we are manipulating contexts. In fact, it is more natural not to.
This suggests that there is a presentation
of type theory where contexts do not mention variables. The first small gain is that such a presentation
would not be burdened with comments about variables being bounded or not, or fresh or
not occuring at all.

\subsection{The basic judgments}
The judgments we can make are the standard six basic judgments involving contexts, types and terms and
judgmental equality in each of these three cases.
\begin{align*}
\jctx*{\Gamma} & \jctxeq*{\Gamma}{\Gamma'}\\
\jtype*{\Gamma}{A} & \jtypeeq*{\Gamma}{A}{B}\\
\jterm*{\Gamma}{A}{x} & \jtermeq*{\Gamma}{A}{x}{y}.
\end{align*}

In contrast with standard practise, we don't assume that there is an empty context. The idea is that
if we interpret only the syntax of contexts, types and terms with extension, weakening and
substitution, we do not have a full model of type theory but we still have a 
structure resembling a category and categories need not have terminal objects. Therefore,
the rules that follow will describe only how to manipulate contexts, types and terms.
Of most of the inferences that we give there are two versions: one introducing an operation, the other asserting that the operation in question preserves judgmental equality.

\subsection{The basic rules for judgmental equality}
The rules for judgmental equality establish that it is an equivalence relation
in all three cases (contexts, types and terms).
%\begingroup
%\renewcommand*{\arraystretch}{3}
%\begin{equation*}
%\begin{array}{ccc}
\begin{infarray}{ccc}
\inference{\jctx{\Gamma}}{\jctxeq{\Gamma}{\Gamma}} & \inference{\jctxeq{\Gamma}{\Delta}}{\jctxeq{\Delta}{\Gamma}} & \inference{\jctxeq{\Gamma}{\Delta}\qquad\jctxeq{\Delta}{\Theta}}{\jctxeq{\Gamma}{\Theta}}\\
\inference{\jtype{\Gamma}{A}}{\jtypeeq{\Gamma}{A}{A}} &
\inference{\jtypeeq{\Gamma}{A}{B}}{\jtypeeq{\Gamma}{B}{A}} & 
\inference{\jtypeeq{\Gamma}{A}{B}\qquad\jtypeeq{\Gamma}{B}{C}}{\jtypeeq{\Gamma}{A}{C}}\\
\inference{\jterm{\Gamma}{A}{x}}{\jtermeq{\Gamma}{A}{x}{x}} & 
\inference{\jtermeq{\Gamma}{A}{x}{y}}{\jtermeq{\Gamma}{A}{y}{x}} &
\inference{\jtermeq{\Gamma}{A}{x}{y}\qquad\jtermeq{\Gamma}{A}{y}{z}}{\jtermeq{\Gamma}{A}{x}{z}}
\end{infarray}
%\end{array}
%\end{equation*}
\begin{infarray}{cc}
\inference{\jctxeq{\Gamma}{\Delta}\qquad\jtype{\Gamma}{A}}{\jtype{\Delta}{A}}
& \inference{\jctxeq{\Gamma}{\Delta}\qquad\jtypeeq{\Gamma}{A}{B}}{\jtypeeq{\Delta}{A}{B}}\\
\inference{\jctxeq{\Gamma}{\Delta}\qquad\jterm{\Gamma}{A}{x}}{\jterm{\Delta}{A}{x}}
& \inference{\jctxeq{\Gamma}{\Delta}\qquad\jtermeq{\Gamma}{A}{x}{y}}{\jtermeq{\Delta}{A}{x}{y}}\\
\inference{\jtypeeq{\Gamma}{A}{B}\qquad \jterm{\Gamma}{A}{x}}{\jterm{\Gamma}{B}{x}}
& \inference{\jtypeeq{\Gamma}{A}{B}\qquad\jtermeq{\Gamma}{A}{x}{y}}{\jtermeq{\Gamma}{B}{x}{y}}
\end{infarray}

\subsection{Extension}
We introduce extension which not only extends a context $\Gamma$ and a type
$A$ over it to a context $\ctxext{\Gamma}{A}$, but which also extends a type $A$
in context $\Gamma$ and a family $P$ over it to a type $\ctxext{A}{P}$ in context
$\Gamma$. We do this to ensure that all of type theory can be done in a context.
For instance, we could say (1) that a context in context $\Gamma$ is the same thing
as a type in context $\Gamma$; (2) When $A$ is a context in this sense, a type in
context $A$ is the same thing as a family $P$ over $A$ and (3) when $P$ is a type
in context $A$ in this sense, a term of $P$ keeps its original meaning.

Note that by introducing extension on the level of types and families,
we introduce $\Sigma$-types at a very early stage.
However, we need substitution to make this precise.
\begin{infarray}{cc}
\inference{\jtype{\Gamma}{A}}{\jctx{\ctxext{\Gamma}{A}}}
& \inference{\jctxeq{\Gamma}{\Delta}\qquad\jtypeeq{\Gamma}{A}{B}}{\jctxeq{\ctxext{\Gamma}{A}}{\ctxext{\Delta}{B}}}\\
\inference{\jtype{\ctxext{\Gamma}{A}}{P}}{\jtype{\Gamma}{\ctxext{A}{P}}}
& \inference{\jtypeeq{\Gamma}{A}{B}\qquad\jtypeeq{\ctxext{\Gamma}{A}}{P}{Q}}{\jtypeeq{\Gamma}{\ctxext{A}{P}}{\ctxext{B}{Q}}}
\end{infarray}

\subsection{Weakening}
We first define weakening by a context $\Gamma$. Since weakening by $\Gamma$
should in principle be a functor, it acts on contexts, types and terms alike.
Note that it is because of the weakening $\ctxwk{\Gamma}{\Delta}$ of a context
$\Delta$ by $\Gamma$ that we can speak of context morphisms from $\Gamma$ to $\Delta$: they are the terms
of $\ctxwk{\Gamma}{\Delta}$.
\begin{infarray}{cc}
\inference{\jctx{\Gamma}\qquad\jctx{\Delta}}{\jtype{\Gamma}{\ctxwk{\Gamma}{\Delta}}} 
& \inference{\jctxeq{\Gamma}{\Gamma'}\qquad\jctxeq{\Delta}{\Delta'}}{\jtypeeq{\Gamma}{\ctxwk{\Gamma}{\Delta}}{\ctxwk{\Gamma'}{\Delta'}}}\\
\inference{\jctx{\Gamma}\qquad\jtype{\Delta}{B}}{\jtype{\ctxext{\Gamma}{\ctxwk{\Gamma}{\Delta}}}{\ctxwk{\Gamma}{B}}} & \inference{\jctxeq{\Gamma}{\Gamma'}\qquad\jtypeeq{\Delta}{B}{B'}}{\jtypeeq{\ctxext{\Gamma}{\ctxwk{\Gamma}{\Delta}}}{\ctxwk{\Gamma}{B}}{\ctxwk{\Gamma'}{B'}}}\\
\inference{\jctx{\Gamma}\qquad\jterm{\Delta}{B}{y}}{\jterm{\ctxext{\Gamma}{\ctxwk{\Gamma}{\Delta}}}{\ctxwk{\Gamma}{B}}{\ctxwk{\Gamma}{y}}} 
& \inference{\jctxeq{\Gamma}{\Gamma'}\qquad\jtermeq{\Delta}{B}{y}{y'}}{\jtermeq{\ctxext{\Gamma}{\ctxwk{\Gamma}{\Delta}}}{\ctxwk{\Gamma}{B}}{\ctxwk{\Gamma}{y}}{\ctxwk{\Gamma'}{y'}}} 
\end{infarray}

A weakening operation is also defined for types. When $A$ and $B$ are both types
 in context $\Gamma$, the weakened type $\ctxwk{A}{B}$ in context $\ctxext{\Gamma}{A}$
 is the family which `doesn't really depend on $A$'. The terms of $\ctxwk{A}{B}$
 are the functions from $A$ to $B$. Likewise, the terms $\ctxwk{A}{y}$ are the
 constant maps at $y$ from $A$ to $B$, for $y:B$.
\begin{infarray}{cc}
\inference{\jtype{\Gamma}{A}\qquad\jtype{\Gamma}{B}}{\jtype{\ctxext{\Gamma}{A}}{\ctxwk{A}{B}}}
& \inference{\jtypeeq{\Gamma}{A}{A'}\qquad\jtypeeq{\Gamma}{B}{B'}}{\jtypeeq{\ctxext{\Gamma}{A}}{\ctxwk{A}{B}}{\ctxwk{A'}{B'}}}\\
\inference{\jtype{\Gamma}{A}\qquad\jtype{\ctxext{\Gamma}{B}}{Q}}
{\jtype{\ctxext{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{Q}}}
& \inference{\jtypeeq{\Gamma}{A}{A'}\qquad\jtypeeq{\ctxext{\Gamma}{B}}{Q}{Q'}}
{\jtypeeq{\ctxext{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{Q}}{\ctxwk{A'}{Q'}}}\\
\inference{\jtype{\Gamma}{A}\qquad\jterm{\Gamma}{B}{y}}{\jterm{\ctxext{\Gamma}{A}}{\ctxwk{A}{B}}{\ctxwk{A}{y}}}
& \inference{\jtypeeq{\Gamma}{A}{A'}\qquad\jtermeq{\Gamma}{B}{y}{y'}}{\jtermeq{\ctxext{\Gamma}{A}}{\ctxwk{A}{B}}{\ctxwk{A}{y}}{\ctxwk{A'}{y'}}}
\end{infarray}

\subsection{Substitution}
\begin{infarray}{cc}
\inference{\jterm{\Gamma}{A}{x}\qquad\jtype{\ctxext{\Gamma}{A}}{P}}{\jtype{\Gamma}{\subst{x}{P}}}
& \inference{\jtermeq{\Gamma}{A}{x}{x'}\qquad\jtypeeq{\ctxext{\Gamma}{A}}{P}{P'}}{\jtypeeq{\Gamma}{\subst{x}{P}}{\subst{x'}{P'}}}\\
\inference{\jterm{\Gamma}{A}{x}\qquad\jtype{\ctxext{{\Gamma}{A}}{P}}{Q}}{\jtype{\ctxext{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}}
& \inference{\jtermeq{\Gamma}{A}{x}{x'}\qquad\jtypeeq{\ctxext{{\Gamma}{A}}{P}}{Q}{Q'}}{\jtypeeq{\ctxext{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}{\subst{x'}{Q'}}}\\
\inference{\jterm{\Gamma}{A}{x}\qquad\jterm{\ctxext{\Gamma}{A}}{P}{f}}{\jterm{\Gamma}{\subst{x}{P}}{\subst{x}{f}}}
& \inference{\jtermeq{\Gamma}{A}{x}{x'}\qquad\jtermeq{\ctxext{\Gamma}{A}}{P}{f}{f'}}{\jtermeq{\Gamma}{\subst{x}{P}}{\subst{x}{f}}{\subst{x'}{f'}}}\\
\inference{\jterm{\Gamma}{A}{x}\qquad\jterm{\ctxext{{\Gamma}{A}}{P}}{Q}{g}}{\jterm{\ctxext{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}{\subst{x}{g}}}
& \inference{\jtermeq{\Gamma}{A}{x}{x'}\qquad\jtermeq{\ctxext{{\Gamma}{A}}{P}}{Q}{g}{g'}}{\jtermeq{\ctxext{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}{\subst{x}{g}}{\subst{x'}{g'}}}
\end{infarray}

\subsection{Compatibility of extension, weakening and substitution}
We describe the compatibility of extension, weakening and substitution. These
compatibility rules are formulated in a weak sense.

Extension is compatible with extension:
\begin{infarray}{c}
\inference
{ \jtype{\ctxext{\Gamma}{A}}{P}
}{\jterm
  {\ctxext{\Gamma}{{A}{P}}}
  {\ctxwk{\ctxext{\Gamma}{{A}{P}}}{\ctxext{{\Gamma}{A}}{P}}}{\typefont{extext}_0(A,P)}
}\\
\inference
{ \jtype{\ctxext{{\Gamma}{A}}{P}}{Q}
}{\jterm
  {\ctxext{\Gamma}{{A}{{P}{Q}}}}
  {\ctxwk{\ctxext{A}{{P}{Q}}}{\ctxext{{A}{P}}{Q}}}
  {\typefont{extext}_1(A,P,Q)}
}
\end{infarray}

Weakening is compatible with extension:
\begin{infarray}{c}
\inference{
  \jctx{\Gamma}
  \qquad
  \jtype{\Delta}{B}}
  {\jterm
    {\ctxext{\Gamma}{\ctxwk{\Gamma}{\ctxext{\Delta}{B}}}}
    {\ctxwk{{\Gamma}{\ctxext{\Delta}{B}}}{\ctxext{\ctxwk{\Gamma}{\Delta}}{\ctxwk{\Gamma}{B}}}}
    {\typefont{wkext}_0(\Gamma,B)}}
\end{infarray}

Substitution is compatible with substitution:
\begin{infarray}{c}
\inference{\jtype{\ctxext{{\Gamma}{A}}{P}}{Q}\qquad\jterm{\Gamma}{A}{x}\qquad\jterm{\ctxext{\Gamma}{A}}{P}{f}}
{\jtypeeq{\Gamma}{\subst{x}{\subst{f}{Q}}}{\subst{{x}{f}}{\subst{x}{Q}}}}
\end{infarray}

\subsection{More stuff in the type theory}
\begin{infarray}{c}
\inference{\jtype{\Gamma}{A}}{\jterm{\ctxext{\Gamma}{A}}{\ctxwk{A}{A}}{\idfunc[A]}}
\end{infarray}

\subsection{Leibniz equality on terms}
Occasionally we shall need a weaker version of equality than judgmental equality.
Our choice here is Leibniz equality, since it implies propositional equality when
identity types are present.

Let $P$ be a family over $A$ in context $\Gamma$. Then we can weaken
$\ctxwk{A}{P}$, which is a family over $\ctxwk{A}{A}$ in context $\ctxext{\Gamma}{A}$.
We also have the family $\ctxwk{{A}{A}}{P}$ over $\ctxwk{A}{A}$ in context
$\ctxext{\Gamma}{A}$ and thus we have the inference
\begin{equation*}
\inference{\jtype{\ctxext{\Gamma}{A}}{P}}{\jtype{\ctxext{{{\Gamma}{A}}{\ctxwk{A}{A}}}{\ctxwk{A}{P}}}{\ctxwk{{A}{P}}{\ctxwk{{A}{A}}{P}}}}
\end{equation*}
This family plays the role of the family $\jtype{\Gamma,x,y:A}{P(x)\to P(y)}$
in the type theory of \cite{TheBook}.

\subsection{Trivial cofibrations and weak equivalences of types}
We describe a relation between types that expresses when they are weakly equivalent.
Weak equivalence is introduced because we need a weaker notion of judgmental 
equality which also makes sense when identity types are not present, since that
would allow us to state that context extension, weakening and substitution
commute with each other.

A term $f:\ctxwk{A}{B}$ is a trivial cofibration if it has the
property that for any fibration $Q$ over $B$,
to find a section of $Q$ it suffices to find a section of the fibration
$f^\ast Q$ over $A$. In our type theoretical setting, the rôle of fibrations
is played by families, the rôle of the function type $A\to B$ is played by
$\ctxwk{A}{B}$ and our version of the pullback $f^\ast Q$ is $\subst{f}{\ctxwk{A}{Q}}$.

\begin{defn}
Let $f$ be a term of $\ctxwk{A}{B}$ in context $\Gamma$.
\begin{enumerate}
\item For a family $Q$ over $B$ in context $\Gamma$ we define $f^\ast Q\jdeq\subst{f}{\ctxwk{A}{Q}}$.
\item For a term $g$ of $Q$ in context $\ctxext{\Gamma}{B}$ we define $f^\ast g\jdeq\subst{f}{\ctxwk{A}{g}}$.
\end{enumerate} 
\end{defn}
\begin{rmk}
The type $f^\ast(\ctxwk{B}{C})$ in context $\ctxext{\Gamma}{A}$ is can be viewed as the
type of functions from $A$ to $C$ which factor through $f$. Thus there should be
a function from $f^\ast(\ctxwk{B}{C})$ to $\ctxwk{A}{C}$.
\end{rmk}

\begin{rmk}
Every term $\jterm{\ctxext{\Gamma}{A}}{\ctxwk{A}{B}}{f}$ allows us to infer the following:
\begin{equation*}
\inference{\jtype{\ctxext{\Gamma}{B}}{Q}}{\jtype{\ctxext{\Gamma}{A}}{f^\ast Q}}
\qquad
\inference{\jterm{\ctxext{\Gamma}{B}}{Q}{g}}{\jterm{\ctxext{\Gamma}{A}}{f^\ast Q}{f^\ast g}}
\end{equation*}
\end{rmk}

\begin{defn}
A term $f:\ctxwk{A}{B}$ in context $\ctxext{\Gamma}{A}$ is said to be a trivial
cofibration if we can infer
\begin{equation*}
\inference{\jterm{\ctxext{\Gamma}{A}}{f^\ast Q}{t}}{\jterm{\ctxext{\Gamma}{B}}{Q}{\tilde{t}}}\qquad
\inference{\jterm{\ctxext{\Gamma}{A}}{f^\ast Q}{t}}{\jtermeq{\ctxext{\Gamma}{A}}{f^\ast Q}{f^\ast \tilde{t}}{t}}
\end{equation*}
{\color{red}This statement should be reformulated so that it only involves a single judgment...
but I don't see directly how to do that.}
\end{defn}

We have the following theorem in the type theory of \cite{TheBook}, which supports
our claim that we may indeed speak of a trivial cofibration.

\begin{thm}
Suppose $f:A\to B$ is a function. Then $f$ is an equivalence if and only if
for every $Q:B\to\type$ and every $g:\prd{x:A}Q(f(x))$ there is a section
$h:\prd{y:B}Q(b)$ with the property that $h\circ f\htpy g$. 
\end{thm}

\begin{proof}
We can first take $Q$ to be the constant family $\lam{y}A$. Furthermore, we may
take $g\defeq\idfunc[A]$. Then we get a term of type
\begin{equation*}
\sm{h:B\to A}h\circ f\htpy \idfunc[A],
\end{equation*}
i.e.~we get a left inverse $h$ for $f$. To show that $h$ is also a right inverse
of $f$, let $Q$ be the family $\lam{y}\id{f(h(y))}{y}$. To find a section of
$Q$, which is the homotopy we aim for, it suffices to find a section of
$Q\circ f$. In other words, we have to show that $\id{f(h(f(x)))}{f(x)}$ for
every $x:A$. This follows from the fact that $h$ is a left inverse for $f$.

The reverse direction is immediate.
\end{proof}

Another approach would be to define $f:\ctxwk{A}{B}$ to be left invertible
if there is a term $\ctxext{{\Gamma}{A}}{P}$

\subsection{The interactions of extension, weakening and substitution}

\subsection{The basic constructors of type theory}
\subsubsection{Dependent function types}
\begin{infarray}{cc}
\inference{\jtype{\ctxext{\Gamma}{A}}{P}}{\jtype{\Gamma}{\mprd{A}{P}}}
& \inference{\jtypeeq{\ctxext{\Gamma}{A}}{P}{P'}}{\jtypeeq{\Gamma}{\mprd{A}{P}}{\mprd{A}{P'}}}\\
\inference{\jterm{\ctxext{\Gamma}{A}}{P}{f}}{\jterm{\Gamma}{\mprd{A}{P}}{\lambda(f)}}
& \inference{\jtermeq{\ctxext{\Gamma}{A}}{P}{f}{f'}}{\jtermeq{\Gamma}{\mprd{A}{P}}{\lambda(f)}{\lambda(f')}}\\
\inference{\jterm{\Gamma}{\mprd{A}{P}}{g}}{\jterm{\ctxext{\Gamma}{A}}{P}{\tfev(g)}}
& \inference{\jtermeq{\Gamma}{\mprd{A}{P}}{g}{g'}}{\jtermeq{\ctxext{\Gamma}{A}}{P}{\tfev(g)}{\tfev(g')}}\\
\inference{\jterm{\ctxext{\Gamma}{A}}{P}{f}}{\jtermeq{\Gamma}{A}{\tfev(\lambda(f))}{f}}
\end{infarray}

With these rules we will not get the weak $\eta$-rule when identity types are present.
So it might be better to state that $\lambda$ is a trivial cofibration

\subsection{Identity types}
We need to define the diagonal function $\jterm{\ctxext{\Gamma}{A}}{\ctxwk{A}{\ctxext{A}{\ctxwk{A}{A}}}}{\delta}$
\begin{infarray}{c}
\inference{\jtype{\Gamma}{A}}{\jtype{\ctxext{{\Gamma}{A}}{\ctxwk{A}{A}}}{\idtypevar{A}}}\\
\inference{\jtype{\Gamma}{A}}{\jterm{\ctxext{\Gamma}{A}}{\delta^\ast\idtypevar{A}}{\reflsym}}\\
\inference{\jtype{\ctxext{{{\Gamma}{A}}{\ctxwk{A}{A}}}{\idtypevar{A}}}{P}\qquad
\jterm{\ctxext{\Gamma}{A}}{\pairr{\delta,\reflsym}^\ast P}{d}}
{\jterm{\ctxext{{{\Gamma}{A}}{\ctxwk{A}{A}}}{\idtypevar{A}}}{P}{J}}
\end{infarray}

### 2013-10-25 article-models-tt.tex
added basic judgment for i is a term of context Gamma

@@
 \subsection{The basic judgments}
-The judgments we can make are the standard six basic judgments involving contexts, types and terms and
-judgmental equality in each of these three cases.
+The judgments we wish to make are not the standad six judgments for ``$\Gamma$ is a context'', ``$A$ is
+a type in context $\Gamma$'' and ``$x$ is a term of $A$ in context $\Gamma$'' and the three equality
+judgments coming with them. We have those, but we also add a judgment for ``$i$ is a term of context $\Gamma$''
+and ``$i$ and $j$ are equal terms of context $\Gamma$''. Thus, we have eight basic judgments we can make:
+
 \begin{align*}
 \jctx*{\Gamma} & \jctxeq*{\Gamma}{\Gamma'}\\
+\jtermc*{\Gamma}{i} & \jtermc*{\Gamma}{i}{j}\\
 \jtype*{\Gamma}{A} & \jtypeeq*{\Gamma}{A}{B}\\
-\jterm*{\Gamma}{A}{x} & \jtermeq*{\Gamma}{A}{x}{y}.
+\jtermt*{\Gamma}{A}{x} & \jtermteq*{\Gamma}{A}{x}{y}.
 \end{align*}
@@
 \inference{\jtypeeq{\Gamma}{A}{B}\qquad\jtypeeq{\Gamma}{B}{C}}{\jtypeeq{\Gamma}{A}{C}}\\
-\inference{\jterm{\Gamma}{A}{x}}{\jtermeq{\Gamma}{A}{x}{x}} & 
-\inference{\jtermeq{\Gamma}{A}{x}{y}}{\jtermeq{\Gamma}{A}{y}{x}} &
-\inference{\jtermeq{\Gamma}{A}{x}{y}\qquad\jtermeq{\Gamma}{A}{y}{z}}{\jtermeq{\Gamma}{A}{x}{z}}
+\inference{\jtermt{\Gamma}{A}{x}}{\jtermteq{\Gamma}{A}{x}{x}} & 
+\inference{\jtermteq{\Gamma}{A}{x}{y}}{\jtermteq{\Gamma}{A}{y}{x}} &
+\inference{\jtermteq{\Gamma}{A}{x}{y}\qquad\jtermteq{\Gamma}{A}{y}{z}}{\jtermteq{\Gamma}{A}{x}{z}}
 \end{infarray}
@@
 & \inference{\jctxeq{\Gamma}{\Delta}\qquad\jtypeeq{\Gamma}{A}{B}}{\jtypeeq{\Delta}{A}{B}}\\
-\inference{\jctxeq{\Gamma}{\Delta}\qquad\jterm{\Gamma}{A}{x}}{\jterm{\Delta}{A}{x}}
-& \inference{\jctxeq{\Gamma}{\Delta}\qquad\jtermeq{\Gamma}{A}{x}{y}}{\jtermeq{\Delta}{A}{x}{y}}\\
-\inference{\jtypeeq{\Gamma}{A}{B}\qquad \jterm{\Gamma}{A}{x}}{\jterm{\Gamma}{B}{x}}
-& \inference{\jtypeeq{\Gamma}{A}{B}\qquad\jtermeq{\Gamma}{A}{x}{y}}{\jtermeq{\Gamma}{B}{x}{y}}
+\inference{\jctxeq{\Gamma}{\Delta}\qquad\jtermt{\Gamma}{A}{x}}{\jtermt{\Delta}{A}{x}}
+& \inference{\jctxeq{\Gamma}{\Delta}\qquad\jtermteq{\Gamma}{A}{x}{y}}{\jtermteq{\Delta}{A}{x}{y}}\\
+\inference{\jtypeeq{\Gamma}{A}{B}\qquad \jtermt{\Gamma}{A}{x}}{\jtermt{\Gamma}{B}{x}}
+& \inference{\jtypeeq{\Gamma}{A}{B}\qquad\jtermteq{\Gamma}{A}{x}{y}}{\jtermteq{\Gamma}{B}{x}{y}}
 \end{infarray}
@@
 \inference{\jctx{\Gamma}\qquad\jtype{\Delta}{B}}{\jtype{\ctxext{\Gamma}{\ctxwk{\Gamma}{\Delta}}}{\ctxwk{\Gamma}{B}}} & \inference{\jctxeq{\Gamma}{\Gamma'}\qquad\jtypeeq{\Delta}{B}{B'}}{\jtypeeq{\ctxext{\Gamma}{\ctxwk{\Gamma}{\Delta}}}{\ctxwk{\Gamma}{B}}{\ctxwk{\Gamma'}{B'}}}\\
-\inference{\jctx{\Gamma}\qquad\jterm{\Delta}{B}{y}}{\jterm{\ctxext{\Gamma}{\ctxwk{\Gamma}{\Delta}}}{\ctxwk{\Gamma}{B}}{\ctxwk{\Gamma}{y}}} 
-& \inference{\jctxeq{\Gamma}{\Gamma'}\qquad\jtermeq{\Delta}{B}{y}{y'}}{\jtermeq{\ctxext{\Gamma}{\ctxwk{\Gamma}{\Delta}}}{\ctxwk{\Gamma}{B}}{\ctxwk{\Gamma}{y}}{\ctxwk{\Gamma'}{y'}}} 
+\inference{\jctx{\Gamma}\qquad\jtermt{\Delta}{B}{y}}{\jtermt{\ctxext{\Gamma}{\ctxwk{\Gamma}{\Delta}}}{\ctxwk{\Gamma}{B}}{\ctxwk{\Gamma}{y}}} 
+& \inference{\jctxeq{\Gamma}{\Gamma'}\qquad\jtermteq{\Delta}{B}{y}{y'}}{\jtermteq{\ctxext{\Gamma}{\ctxwk{\Gamma}{\Delta}}}{\ctxwk{\Gamma}{B}}{\ctxwk{\Gamma}{y}}{\ctxwk{\Gamma'}{y'}}} 
 \end{infarray}
@@
 {\jtypeeq{\ctxext{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{Q}}{\ctxwk{A'}{Q'}}}\\
-\inference{\jtype{\Gamma}{A}\qquad\jterm{\Gamma}{B}{y}}{\jterm{\ctxext{\Gamma}{A}}{\ctxwk{A}{B}}{\ctxwk{A}{y}}}
-& \inference{\jtypeeq{\Gamma}{A}{A'}\qquad\jtermeq{\Gamma}{B}{y}{y'}}{\jtermeq{\ctxext{\Gamma}{A}}{\ctxwk{A}{B}}{\ctxwk{A}{y}}{\ctxwk{A'}{y'}}}
+\inference{\jtype{\Gamma}{A}\qquad\jtermt{\Gamma}{B}{y}}{\jtermt{\ctxext{\Gamma}{A}}{\ctxwk{A}{B}}{\ctxwk{A}{y}}}
+& \inference{\jtypeeq{\Gamma}{A}{A'}\qquad\jtermteq{\Gamma}{B}{y}{y'}}{\jtermteq{\ctxext{\Gamma}{A}}{\ctxwk{A}{B}}{\ctxwk{A}{y}}{\ctxwk{A'}{y'}}}
 \end{infarray}
@@
 \begin{infarray}{cc}
-\inference{\jterm{\Gamma}{A}{x}\qquad\jtype{\ctxext{\Gamma}{A}}{P}}{\jtype{\Gamma}{\subst{x}{P}}}
-& \inference{\jtermeq{\Gamma}{A}{x}{x'}\qquad\jtypeeq{\ctxext{\Gamma}{A}}{P}{P'}}{\jtypeeq{\Gamma}{\subst{x}{P}}{\subst{x'}{P'}}}\\
-\inference{\jterm{\Gamma}{A}{x}\qquad\jtype{\ctxext{{\Gamma}{A}}{P}}{Q}}{\jtype{\ctxext{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}}
-& \inference{\jtermeq{\Gamma}{A}{x}{x'}\qquad\jtypeeq{\ctxext{{\Gamma}{A}}{P}}{Q}{Q'}}{\jtypeeq{\ctxext{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}{\subst{x'}{Q'}}}\\
-\inference{\jterm{\Gamma}{A}{x}\qquad\jterm{\ctxext{\Gamma}{A}}{P}{f}}{\jterm{\Gamma}{\subst{x}{P}}{\subst{x}{f}}}
-& \inference{\jtermeq{\Gamma}{A}{x}{x'}\qquad\jtermeq{\ctxext{\Gamma}{A}}{P}{f}{f'}}{\jtermeq{\Gamma}{\subst{x}{P}}{\subst{x}{f}}{\subst{x'}{f'}}}\\
-\inference{\jterm{\Gamma}{A}{x}\qquad\jterm{\ctxext{{\Gamma}{A}}{P}}{Q}{g}}{\jterm{\ctxext{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}{\subst{x}{g}}}
-& \inference{\jtermeq{\Gamma}{A}{x}{x'}\qquad\jtermeq{\ctxext{{\Gamma}{A}}{P}}{Q}{g}{g'}}{\jtermeq{\ctxext{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}{\subst{x}{g}}{\subst{x'}{g'}}}
+\inference{\jtermt{\Gamma}{A}{x}\qquad\jtype{\ctxext{\Gamma}{A}}{P}}{\jtype{\Gamma}{\subst{x}{P}}}
+& \inference{\jtermteq{\Gamma}{A}{x}{x'}\qquad\jtypeeq{\ctxext{\Gamma}{A}}{P}{P'}}{\jtypeeq{\Gamma}{\subst{x}{P}}{\subst{x'}{P'}}}\\
+\inference{\jtermt{\Gamma}{A}{x}\qquad\jtype{\ctxext{{\Gamma}{A}}{P}}{Q}}{\jtype{\ctxext{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}}
+& \inference{\jtermteq{\Gamma}{A}{x}{x'}\qquad\jtypeeq{\ctxext{{\Gamma}{A}}{P}}{Q}{Q'}}{\jtypeeq{\ctxext{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}{\subst{x'}{Q'}}}\\
+\inference{\jtermt{\Gamma}{A}{x}\qquad\jtermt{\ctxext{\Gamma}{A}}{P}{f}}{\jtermt{\Gamma}{\subst{x}{P}}{\subst{x}{f}}}
+& \inference{\jtermteq{\Gamma}{A}{x}{x'}\qquad\jtermteq{\ctxext{\Gamma}{A}}{P}{f}{f'}}{\jtermteq{\Gamma}{\subst{x}{P}}{\subst{x}{f}}{\subst{x'}{f'}}}\\
+\inference{\jtermt{\Gamma}{A}{x}\qquad\jtermt{\ctxext{{\Gamma}{A}}{P}}{Q}{g}}{\jtermt{\ctxext{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}{\subst{x}{g}}}
+& \inference{\jtermteq{\Gamma}{A}{x}{x'}\qquad\jtermteq{\ctxext{{\Gamma}{A}}{P}}{Q}{g}{g'}}{\jtermteq{\ctxext{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}{\subst{x}{g}}{\subst{x'}{g'}}}
 \end{infarray}
+
+\subsection{Definitional equality rules for weakening and substitution}
 
@@
 { \jtype{\ctxext{\Gamma}{A}}{P}
-}{\jterm
+}{\jtermt
   {\ctxext{\Gamma}{{A}{P}}}
@@
 { \jtype{\ctxext{{\Gamma}{A}}{P}}{Q}
-}{\jterm
+}{\jtermt
   {\ctxext{\Gamma}{{A}{{P}{Q}}}}
@@
   \jtype{\Delta}{B}}
-  {\jterm
+  {\jtermt
     {\ctxext{\Gamma}{\ctxwk{\Gamma}{\ctxext{\Delta}{B}}}}
@@
 \begin{infarray}{c}
-\inference{\jtype{\ctxext{{\Gamma}{A}}{P}}{Q}\qquad\jterm{\Gamma}{A}{x}\qquad\jterm{\ctxext{\Gamma}{A}}{P}{f}}
+\inference{\jtype{\ctxext{{\Gamma}{A}}{P}}{Q}\qquad\jtermt{\Gamma}{A}{x}\qquad\jtermt{\ctxext{\Gamma}{A}}{P}{f}}
 {\jtypeeq{\Gamma}{\subst{x}{\subst{f}{Q}}}{\subst{{x}{f}}{\subst{x}{Q}}}}
@@
 \begin{infarray}{c}
-\inference{\jtype{\Gamma}{A}}{\jterm{\ctxext{\Gamma}{A}}{\ctxwk{A}{A}}{\idfunc[A]}}
+\inference{\jtype{\Gamma}{A}}{\jtermt{\ctxext{\Gamma}{A}}{\ctxwk{A}{A}}{\idfunc[A]}}
 \end{infarray}
@@
 \begin{rmk}
-Every term $\jterm{\ctxext{\Gamma}{A}}{\ctxwk{A}{B}}{f}$ allows us to infer the following:
+Every term $\jtermt{\ctxext{\Gamma}{A}}{\ctxwk{A}{B}}{f}$ allows us to infer the following:
 \begin{equation*}
@@
 \qquad
-\inference{\jterm{\ctxext{\Gamma}{B}}{Q}{g}}{\jterm{\ctxext{\Gamma}{A}}{f^\ast Q}{f^\ast g}}
+\inference{\jtermt{\ctxext{\Gamma}{B}}{Q}{g}}{\jtermt{\ctxext{\Gamma}{A}}{f^\ast Q}{f^\ast g}}
 \end{equation*}
@@
 \begin{equation*}
-\inference{\jterm{\ctxext{\Gamma}{A}}{f^\ast Q}{t}}{\jterm{\ctxext{\Gamma}{B}}{Q}{\tilde{t}}}\qquad
-\inference{\jterm{\ctxext{\Gamma}{A}}{f^\ast Q}{t}}{\jtermeq{\ctxext{\Gamma}{A}}{f^\ast Q}{f^\ast \tilde{t}}{t}}
+\inference{\jtermt{\ctxext{\Gamma}{A}}{f^\ast Q}{t}}{\jtermt{\ctxext{\Gamma}{B}}{Q}{\tilde{t}}}\qquad
+\inference{\jtermt{\ctxext{\Gamma}{A}}{f^\ast Q}{t}}{\jtermteq{\ctxext{\Gamma}{A}}{f^\ast Q}{f^\ast \tilde{t}}{t}}
 \end{equation*}
@@
 & \inference{\jtypeeq{\ctxext{\Gamma}{A}}{P}{P'}}{\jtypeeq{\Gamma}{\mprd{A}{P}}{\mprd{A}{P'}}}\\
-\inference{\jterm{\ctxext{\Gamma}{A}}{P}{f}}{\jterm{\Gamma}{\mprd{A}{P}}{\lambda(f)}}
-& \inference{\jtermeq{\ctxext{\Gamma}{A}}{P}{f}{f'}}{\jtermeq{\Gamma}{\mprd{A}{P}}{\lambda(f)}{\lambda(f')}}\\
-\inference{\jterm{\Gamma}{\mprd{A}{P}}{g}}{\jterm{\ctxext{\Gamma}{A}}{P}{\tfev(g)}}
-& \inference{\jtermeq{\Gamma}{\mprd{A}{P}}{g}{g'}}{\jtermeq{\ctxext{\Gamma}{A}}{P}{\tfev(g)}{\tfev(g')}}\\
-\inference{\jterm{\ctxext{\Gamma}{A}}{P}{f}}{\jtermeq{\Gamma}{A}{\tfev(\lambda(f))}{f}}
+\inference{\jtermt{\ctxext{\Gamma}{A}}{P}{f}}{\jtermt{\Gamma}{\mprd{A}{P}}{\lambda(f)}}
+& \inference{\jtermteq{\ctxext{\Gamma}{A}}{P}{f}{f'}}{\jtermteq{\Gamma}{\mprd{A}{P}}{\lambda(f)}{\lambda(f')}}\\
+\inference{\jtermt{\Gamma}{\mprd{A}{P}}{g}}{\jtermt{\ctxext{\Gamma}{A}}{P}{\tfev(g)}}
+& \inference{\jtermteq{\Gamma}{\mprd{A}{P}}{g}{g'}}{\jtermteq{\ctxext{\Gamma}{A}}{P}{\tfev(g)}{\tfev(g')}}\\
+\inference{\jtermt{\ctxext{\Gamma}{A}}{P}{f}}{\jtermteq{\Gamma}{A}{\tfev(\lambda(f))}{f}}
 \end{infarray}
@@
 \subsection{Identity types}
-We need to define the diagonal function $\jterm{\ctxext{\Gamma}{A}}{\ctxwk{A}{\ctxext{A}{\ctxwk{A}{A}}}}{\delta}$
+We need to define the diagonal function $\jtermt{\ctxext{\Gamma}{A}}{\ctxwk{A}{\ctxext{A}{\ctxwk{A}{A}}}}{\delta}$
 \begin{infarray}{c}
 \inference{\jtype{\Gamma}{A}}{\jtype{\ctxext{{\Gamma}{A}}{\ctxwk{A}{A}}}{\idtypevar{A}}}\\
-\inference{\jtype{\Gamma}{A}}{\jterm{\ctxext{\Gamma}{A}}{\delta^\ast\idtypevar{A}}{\reflsym}}\\
+\inference{\jtype{\Gamma}{A}}{\jtermt{\ctxext{\Gamma}{A}}{\delta^\ast\idtypevar{A}}{\reflsym}}\\
 \inference{\jtype{\ctxext{{{\Gamma}{A}}{\ctxwk{A}{A}}}{\idtypevar{A}}}{P}\qquad
-\jterm{\ctxext{\Gamma}{A}}{\pairr{\delta,\reflsym}^\ast P}{d}}
-{\jterm{\ctxext{{{\Gamma}{A}}{\ctxwk{A}{A}}}{\idtypevar{A}}}{P}{J}}
+\jtermt{\ctxext{\Gamma}{A}}{\pairr{\delta,\reflsym}^\ast P}{d}}
+{\jtermt{\ctxext{{{\Gamma}{A}}{\ctxwk{A}{A}}}{\idtypevar{A}}}{P}{J}}
 \end{infarray}

### 2013-10-26 article-models-eg.tex
Started description of the model of equifibered diagrams

@@
 {\color{red} point to example}
-
-\subsection{The Sierpinsky model}.
-In the Sierpinsky model, contexts are interpreted by dependent types $A\to\type$.
 
@@
 \endgroup
+
+\subsection{The model of equifibered families of graphs}
+We describe the model $\equifib$ of equifibered families of graphs. The contexts
+are still the graphs. However, the types of $\equifib$ are not just the
+families of graphs, but the equifibered families of graphs. This will have
+far-reaching consequences for the rest of the structure of the model. As an
+example, a morphism of graphs will no longer be a pair consisting of a map of
+vertices and a map of edges.
+
+\begin{defn}
+Let $\Gamma$ be a graph. A type in context $\Gamma$ is a triple 
+$\pairr{A_0,A_1,A_2}$ consisting of
+\begin{align*}
+A_0 & : \Gamma_0\to\type\\
+A_1 & : \prd*{i,j:\Gamma_0}{q:\Gamma_1(i,j)}A_0(i)\to A_0(j)\\
+A_2 & : \prd*{i,j:\Gamma_0}{q:\Gamma_1(i,j)}\isequiv{A_1(q)}
+\end{align*}
+A type in context $\Gamma$ is also called an equifibered family over $\Gamma$.
+\end{defn}
+
+\begin{defn}
+Let $\Gamma$ be a graph. A term of $\Gamma$ is a term of $\Gamma_0$.
+\end{defn}
+
+\begin{defn}
+Let $A$ be an equifibered family over $\Gamma$. A term $x$ of $A$ is a pair
+$\pairr{x_0,x_1}$ consisting of
+\begin{align*}
+x_0 & : \prd{i:\Gamma_0}A_0(i)\\
+x_1 & : \prd*{i,j:\Gamma_0}{q:\Gamma_1(i,j)}\id{A_1(q,x_0(i))}{x_0(j)}
+\end{align*}
+\end{defn}
+
+\begin{defn}
+Suppose that $A$ is an equifibered family over $\Gamma$. We define the graph
+$\ctxext{\Gamma}{A}$ to consist of
+\begin{align*}
+\ctxext{\Gamma}{A}_0 & \defeq\sm{i:\Gamma_0}A_0(i)\\
+\ctxext{\Gamma}{A}_1(\pairr{i,x},\pairr{j,y}) & \defeq \sm{q:\Gamma_1(i,j)}\id{A_1(q,x)}{y}.
+\end{align*}
+\end{defn}
+
+\begin{defn}
+Suppose $\Gamma$ and $\Delta$ are graphs. We define the equifibered family
+$\ctxwk{\Gamma}{\Delta}$ over $\Gamma$ to consist of
+\begin{align*}
+\ctxwk{\Gamma}{\Delta}_0(i) & \defeq \Delta\\
+\ctxwk{\Gamma}{\Delta}_1(q) & \defeq \idfunc[\Delta]\\
+\ctxwk{\Gamma}{\Delta}_2(q) & \defeq \nameless.
+\end{align*}
+The proof $\ctxwk{\Gamma}{\Delta}_2(q)$ that $\ctxwk{\Gamma}{\Delta}_1(q)$ is an
+equivalence is the canonical proof that $\idfunc[\Delta]$ is an equivalence, for
+which we don't have a name.
+\end{defn}
+
+\begin{defn}
+Suppose $A$ and $B$ are equifibered families over $\Gamma$. We define the equifibered
+diagram $\ctxwk{A}{B}$ over $\ctxext{\Gamma}{A}$ to consist of
+\begin{align*}
+\ctxwk{A}{B}_0(\pairr{i,x}) & \defeq B_0(i)\\
+\ctxwk{A}{B}_1(\pairr{q,\alpha}) & \defeq B_1(q)\\
+\ctxwk{A}{B}_2(\pairr{q,\alpha}) & \defeq B_2(q).
+\end{align*}
+\end{defn}
+
+\begin{defn}
+Suppose $A$ is an equifibered family over $\Gamma$ with $x:A$ and that $P$ is an equifibered
+family over $\ctxext{\Gamma}{A}$. Then we define the equifibered family $\subst{x}{P}$
+over $\Gamma$ to consist of
+\begin{align*}
+\pts{\subst{x}{P}}(i) & \defeq \pts{P}(\pts{x}(i))\\
+\edg{\subst{x}{P}}(q,u) & \defeq \edg{P}(\pairr{q,\edg{x}(q)},u)\\
+\subst{x}{P}_2(q) & \defeq P_2(\pairr{q,\edg{x}(q)}).
+\end{align*}
+\end{defn}
+
+\begin{defn}
+Suppose that $A$ is an equifibered family over $\Gamma$ and that $P$ is an
+equifibered family over $\ctxext{\Gamma}{A}$. Then we define the equifibered
+family $\msm{A}{P}$ over $\Gamma$ by
+\begin{align*}
+\pts{\msm{A}{P}}(i) & \defeq \sm{x:\pts{A}(i)}\pts{P}(x)\\
+\edg{\msm{A}{P}}(q,\pairr{x,u}) & \defeq \pairr{\edg{A}(q,x),\edg{P}(\pairr{q,\refl{\edg{A}(q,x)}},u)}\\
+\msm{A}{P}_2(q,\pairr{x,u}) & \defeq \nameless.
+\end{align*}
+\end{defn}

### 2013-10-28 article-models-tt.tex
Started description of the model of equifibered diagrams
introducing judgmental equalities for weakening followed by substitution; introducing identity functions

@@
 \jctx*{\Gamma} & \jctxeq*{\Gamma}{\Gamma'}\\
-\jtermc*{\Gamma}{i} & \jtermc*{\Gamma}{i}{j}\\
+\jtermc*{\Gamma}{i} & \jtermceq*{\Gamma}{i}{j}\\
 \jtype*{\Gamma}{A} & \jtypeeq*{\Gamma}{A}{B}\\
@@
 \subsection{Definitional equality rules for weakening and substitution}
+The judgmental equalities we're about to describe assert that weakening followed
+by substitution leaves everything untouched. Thus, we get that each fiber
+$\subst{x}{\ctxwk{A}{B}}$ is just $B$, that $\ctxwk{A}{y}$ is the constant function
+mapping everything to $y:B$ and similar properties for families and terms thereof.
+
+\begin{infarray}{c}
+\inference{\jtype{\Gamma}{A}\qquad\jtype{\Gamma}{B}\qquad\jtermt{\Gamma}{A}{x}}{\jtypeeq{\Gamma}{\subst{x}{\ctxwk{A}{B}}}{B}}\\
+\inference{\jtype{\Gamma}{A}\qquad\jtype{\ctxext{\Gamma}{B}}{Q}\qquad\jtermt{A}{x}}{\jtypeeq{\ctxext{\Gamma}{B}}{\subst{x}{\ctxwk{A}{Q}}}{Q}}\\
+\inference{\jtermt{\Gamma}{A}{x}\qquad\jtermt{\Gamma}{B}{y}}{\jtermteq{\Gamma}{B}{\subst{x}{\ctxwk{A}{y}}}{y}}\\
+\inference{\jtermt{\Gamma}{A}{x}\qquad\jtermt{\ctxext{\Gamma}{B}}{Q}{g}}{\jtermteq{\ctxext{\Gamma}{B}}{Q}{\subst{x}{\ctxwk{A}{g}}}{g}}
+\end{infarray}
+
+\subsection{The variable rule -- identity functions}
+In the usual presentation of type theory, the variable rule asserts that ``in
+the context $x_1:A_1,...,x_n:A_n$ we have a term $x_i:A_i$ when $1\leq i\leq n$''.
+
+In our current system, we don't need this generality. Instead, it suffices for
+us to assert that there are identity morphisms. We get arbitrary projections
+from weakening.
+
+\begin{infarray}{c}
+\inference{\jtype{\Gamma}{A}}{\jtermt{\ctxext{\Gamma}{A}}{\ctxwk{A}{A}}{\idfunc[A]}}\\
+\inference{\jtermt{\Gamma}{A}{x}}{\jtermteq{\Gamma}{A}{\subst{x}{\idfunc[A]}}{x}}\\
+\inference{\jtermt{\ctxext{\Gamma}{A}}{P}{f}}{\jtermteq{\ctxext{\Gamma}{A}}{P}{\subst{\idfunc[A]}{\ctxwk{A}{f}}}{f}}\\
+\inference{\jtermt{\ctxext{\Gamma}{B}}{\ctxwk{B}{A}}{g}}{\jtermteq{\ctxext{\Gamma}{B}}{\ctxwk{B}{A}}{\subst{g}{\ctxwk{B}{\idfunc[A]}}}{g}}
+\end{infarray}
+
+Using these rules we can derive that there is a judgmental equality $\jtypeeq{\ctxext{\Gamma}{A}}{\subst{x}{{\idfunc[A]}{B}}}
+{\subst{x}{{x}{B}}}$.
+
+%\begin{infarray}{c}
+%\inference{\jtype{\ctxext{{\Gamma}{A}}{P}}{Q}\qquad\jtermt{\Gamma}{A}{x}\qquad\jtermt{\ctxext{A
+%\end{infarray}{c}
 

### 2013-10-29 article-models-tt.tex
compatibility rules and universal property of extension

@@
 
-\subsection{Substitution}
-\begin{infarray}{cc}
-\inference{\jtermt{\Gamma}{A}{x}\qquad\jtype{\ctxext{\Gamma}{A}}{P}}{\jtype{\Gamma}{\subst{x}{P}}}
-& \inference{\jtermteq{\Gamma}{A}{x}{x'}\qquad\jtypeeq{\ctxext{\Gamma}{A}}{P}{P'}}{\jtypeeq{\Gamma}{\subst{x}{P}}{\subst{x'}{P'}}}\\
-\inference{\jtermt{\Gamma}{A}{x}\qquad\jtype{\ctxext{{\Gamma}{A}}{P}}{Q}}{\jtype{\ctxext{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}}
-& \inference{\jtermteq{\Gamma}{A}{x}{x'}\qquad\jtypeeq{\ctxext{{\Gamma}{A}}{P}}{Q}{Q'}}{\jtypeeq{\ctxext{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}{\subst{x'}{Q'}}}\\
-\inference{\jtermt{\Gamma}{A}{x}\qquad\jtermt{\ctxext{\Gamma}{A}}{P}{f}}{\jtermt{\Gamma}{\subst{x}{P}}{\subst{x}{f}}}
-& \inference{\jtermteq{\Gamma}{A}{x}{x'}\qquad\jtermteq{\ctxext{\Gamma}{A}}{P}{f}{f'}}{\jtermteq{\Gamma}{\subst{x}{P}}{\subst{x}{f}}{\subst{x'}{f'}}}\\
-\inference{\jtermt{\Gamma}{A}{x}\qquad\jtermt{\ctxext{{\Gamma}{A}}{P}}{Q}{g}}{\jtermt{\ctxext{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}{\subst{x}{g}}}
-& \inference{\jtermteq{\Gamma}{A}{x}{x'}\qquad\jtermteq{\ctxext{{\Gamma}{A}}{P}}{Q}{g}{g'}}{\jtermteq{\ctxext{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}{\subst{x}{g}}{\subst{x'}{g'}}}
+\subsubsection{Weakening is compatible with itself}
+Suppose $\jtype{\ctxext{\Gamma}{A}}{P}$ and $\jtype{\Gamma}{B}$. Then we can
+consider the weakening $\ctxwk{A}{B}$ over $A$ and consecutively the weakening
+$\ctxwk{P}{{A}{B}}$. This should be isomorphic to $\ctxwk{\ctxext{A}{P}}{B}$.
+
+\subsection{The universal property of extension}
+
+Using weakening and substitution we are able to state the universal property
+for extension. It looks a bit more involved, since we cannot directly refer
+to the variables in the contexts. On the other hand, we can now plainly see
+were there were secretly weakenings going on.
+
+In all of the following inference rules we assume that $\jtype{\ctxext{\Gamma}{A}}{P}$
+is among the hypothesis. The induction principle for extension consists of three
+inference rules:
+
+\begin{infarray}{c}
+\inference{}
+{\jtermt{\ctxext{{\Gamma}{A}}{P}}{\ctxwk{P}{{A}{\ctxext{A}{P}}}}{\pair^P}}\\
+\inference{
+  \jtype{\ctxext{\Gamma}{{A}{P}}}{Q}
+  \qquad
+  \jtermt{\ctxext{{\Gamma}{A}}{P}}{\subst{\pair^P}{\ctxwk{P}{{A}{Q}}}}{f}}
+  {\jtermt{\ctxext{\Gamma}{{A}{P}}}{Q}{\ind{ext}(f)}}\\
+\inference{
+  \jtype{\ctxext{\Gamma}{{A}{P}}}{Q}
+  \qquad
+  \jtermt{\ctxext{{\Gamma}{A}}{P}}{\subst{\pair^P}{\ctxwk{P}{{A}{Q}}}}{f}}
+  {\jtermteq{\ctxext{{\Gamma}{A}}{P}}{\subst{\pair^P}{\ctxwk{P}{{A}{Q}}}}{\subst{\pair^P}{\ctxwk{P}{{A}{\ind{ext}(f)}}}}{f}}
 \end{infarray}
 
-\subsection{Definitional equality rules for weakening and substitution}
+\subsubsection{Weakening is compatible with extension}
+Suppose $\jtype{\Gamma}{A}$ and $\jtype{\ctxext{\Gamma}{B}}{Q}$. Then
+\begin{equation*}
+s:\ctxwk{{A}{\ctxext{B}{Q}}}{\ctxext{\ctxwk{A}{B}}{\ctxwk{A}{Q}}}
+\end{equation*}
+
+\subsection{Substitution}
+Given a family $P$ over $A$ and a term $x$ of $A$, substitution gives a way to
+consider the fiber $\subst{x}{P}$ of $P$ at $x$. Also, we get a way to evaluate
+terms $f$ of $P$ at $x$. This will give us ways to compose functions too. In
+this section, we shall first introduce the operations `substitution of a term $x$'
+for types, terms and families. Then we shall explain how substitution interacts
+with itself, extension and weakening.
+
+In the rules introducing the various substitutions we assume $\jtermt{\Gamma}{A}{x}$;
+in the rules introducing the definitional equalities we assume $\jtermteq{\Gamma}{A}{x}{x'}$.
+
+\begin{infarray}{cc}
+\inference{\jtype{\ctxext{\Gamma}{A}}{P}}{\jtype{\Gamma}{\subst{x}{P}}}
+& \inference{\jtypeeq{\ctxext{\Gamma}{A}}{P}{P'}}{\jtypeeq{\Gamma}{\subst{x}{P}}{\subst{x'}{P'}}}\\
+\inference{\jtype{\ctxext{{\Gamma}{A}}{P}}{Q}}{\jtype{\ctxext{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}}
+& \inference{\jtypeeq{\ctxext{{\Gamma}{A}}{P}}{Q}{Q'}}{\jtypeeq{\ctxext{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}{\subst{x'}{Q'}}}\\
+\inference{\jtermt{\ctxext{\Gamma}{A}}{P}{f}}{\jtermt{\Gamma}{\subst{x}{P}}{\subst{x}{f}}}
+& \inference{\jtermteq{\ctxext{\Gamma}{A}}{P}{f}{f'}}{\jtermteq{\Gamma}{\subst{x}{P}}{\subst{x}{f}}{\subst{x'}{f'}}}\\
+\inference{\jtermt{\ctxext{{\Gamma}{A}}{P}}{Q}{g}}{\jtermt{\ctxext{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}{\subst{x}{g}}}
+& \inference{\jtermteq{\ctxext{{\Gamma}{A}}{P}}{Q}{g}{g'}}{\jtermteq{\ctxext{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}{\subst{x}{g}}{\subst{x'}{g'}}}
+\end{infarray}
+
+\subsubsection{Substitution is compatible with substitution}
+
+We require that substitution is compatible with itself, which is roughly the
+assertion that substitution is associative. However, we cannot just state that
+$\subst{x}{{f}{g}}\jdeq\subst{{x}{f}}{g}$ since the expression $\subst{{x}{f}}{g}$
+is not well-formed. The term $\subst{x}{f}$ can be substituted in (terms of) families over
+$\subst{x}{P}$; the term $\subst{x}{g}$ is such. Therefore, associativity of
+substitution takes the form $\subst{x}{{f}{g}}\jdeq\subst{{x}{f}}{{x}{g}}$.
+
+In the following inference rules we assume
+$\jtermt{\Gamma}{A}{x}$ and $\jtermt{\ctxext{\Gamma}{A}}{P}{f}$.
+
+\begin{infarray}{c}
+\inference{\jtype{\ctxext{{\Gamma}{A}}{P}}{Q}}
+{\jtypeeq{\Gamma}{\subst{x}{{f}{Q}}}{\subst{{x}{f}}{{x}{Q}}}}\\
+\inference{\jtype{\ctxext{{{\Gamma}{A}}{P}}{Q}}{R}}
+{\jtypeeq{\ctxext{\Gamma}{\subst{x}{{f}{Q}}}}{\subst{x}{{f}{R}}}{\subst{{x}{f}}{{x}{R}}}}\\
+\inference{\jtermt{\ctxext{{\Gamma}{A}}{P}}{Q}{g}}
+{\jtermteq{\Gamma}{\subst{x}{{f}{Q}}}{\subst{x}{{f}{g}}}{\subst{{x}{f}}{{x}{g}}}}\\
+\inference{\jtermt{\ctxext{{{\Gamma}{A}}{P}}{Q}}{R}{h}}
+{\jtermteq{\ctxext{\Gamma}{\subst{x}{{f}{Q}}}}{\subst{x}{{f}{R}}}{\subst{x}{{f}{h}}}{\subst{{x}{f}}{{x}{h}}}}
+\end{infarray}
+
+\subsubsection{Substitution is compatible with extension}
+Suppose $\jtermt{\Gamma}{A}{x}$ in all of the following inference rules.
+\begin{infarray}{c}
+\inference{\jtype{\ctxext{{\Gamma}{A}}{P}}{Q}}{\jtypeeq{\Gamma}{\subst{x}{\ctxext{P}{Q}}}{\ctxext{\subst{x}{P}}{\subst{x}{Q}}}}
+\end{infarray}
+
+\subsubsection{Substitution is compatible with weakening}
 The judgmental equalities we're about to describe assert that weakening followed
@@
 \inference{\jtype{\Gamma}{A}\qquad\jtype{\Gamma}{B}\qquad\jtermt{\Gamma}{A}{x}}{\jtypeeq{\Gamma}{\subst{x}{\ctxwk{A}{B}}}{B}}\\
-\inference{\jtype{\Gamma}{A}\qquad\jtype{\ctxext{\Gamma}{B}}{Q}\qquad\jtermt{A}{x}}{\jtypeeq{\ctxext{\Gamma}{B}}{\subst{x}{\ctxwk{A}{Q}}}{Q}}\\
+\inference{\jtype{\Gamma}{A}\qquad\jtype{\ctxext{\Gamma}{B}}{Q}\qquad\jtermt{\Gamma}{A}{x}}{\jtypeeq{\ctxext{\Gamma}{B}}{\subst{x}{\ctxwk{A}{Q}}}{Q}}\\
 \inference{\jtermt{\Gamma}{A}{x}\qquad\jtermt{\Gamma}{B}{y}}{\jtermteq{\Gamma}{B}{\subst{x}{\ctxwk{A}{y}}}{y}}\\

### 2013-10-29 article-models-defn.tex
definition of weak omega groupoids
conjectured an equivalent definition of groupoids

@@
 \end{comment}
+
+\subsection{Weak $\omega$-groupoids}
+
+A weak omega groupoid is a model of type theory with $\Sigma$ and $\idtypevar{}$
+in which there is a term of $\terms{A}$ for every type $A$ in context $\Gamma$.
+
+We could also state that for every term $f:\terms{\ctxwk{A}{B}}$ and every
+family $Q:\tftyp{\mfM}{\ctxext{\Gamma}{B}}$ there is a function
+$\terms{\subst{f}{\ctxwk{A}{Q}}}\to\terms{Q}$ (asserting that $f$ is a trival
+cofibration). This should be equivalent to the previous condition.

### 2013-10-29 article-models-tt.tex
universal property for extension
wrapping up the description of type theory

@@
 
-\subsection{The universal property of extension}
-
-Using weakening and substitution we are able to state the universal property
-for extension. It looks a bit more involved, since we cannot directly refer
-to the variables in the contexts. On the other hand, we can now plainly see
-were there were secretly weakenings going on.
-
-In all of the following inference rules we assume that $\jtype{\ctxext{\Gamma}{A}}{P}$
-is among the hypothesis. The induction principle for extension consists of three
-inference rules:
-
-\begin{infarray}{c}
-\inference{}
-{\jtermt{\ctxext{{\Gamma}{A}}{P}}{\ctxwk{P}{{A}{\ctxext{A}{P}}}}{\pair^P}}\\
-\inference{
-  \jtype{\ctxext{\Gamma}{{A}{P}}}{Q}
-  \qquad
-  \jtermt{\ctxext{{\Gamma}{A}}{P}}{\subst{\pair^P}{\ctxwk{P}{{A}{Q}}}}{f}}
-  {\jtermt{\ctxext{\Gamma}{{A}{P}}}{Q}{\ind{ext}(f)}}\\
-\inference{
-  \jtype{\ctxext{\Gamma}{{A}{P}}}{Q}
-  \qquad
-  \jtermt{\ctxext{{\Gamma}{A}}{P}}{\subst{\pair^P}{\ctxwk{P}{{A}{Q}}}}{f}}
-  {\jtermteq{\ctxext{{\Gamma}{A}}{P}}{\subst{\pair^P}{\ctxwk{P}{{A}{Q}}}}{\subst{\pair^P}{\ctxwk{P}{{A}{\ind{ext}(f)}}}}{f}}
-\end{infarray}
-
 \subsubsection{Weakening is compatible with extension}
@@
 
+\subsection{The universal property of extension}
+
+Using weakening and substitution we are able to state the universal property
+for extension. It looks a bit more involved, since we cannot directly refer
+to the variables in the contexts. On the other hand, we can now plainly see
+were there were secretly weakenings going on.
+
+We begin with stating the universal property of the extension $\ctxext{\Gamma}{A}$.
+In these rules we assume we have $\jtype{\Gamma}{A}$ in the hypotheses.
+
+\begin{infarray}{c}
+\inference{}
+{\jtermt{\ctxext{\Gamma}{A}}{\ctxwk{A}{{\Gamma}{\ctxext{\Gamma}{A}}}}{\pair_A}}\\
+\inference{
+  \jtype{\ctxext{\Gamma}{A}}{P}
+  \qquad
+  \jtermt{\ctxext{\Gamma}{A}}{\subst{\pair_A}{\ctxwk{A}{{\Gamma}{P}}}}{f}}
+  {\jtermt{\ctxext{\Gamma}{A}}{P}{\ind{\tfext_\Gamma(A)}(f)}}\\
+\inference{
+  \jtype{\ctxext{\Gamma}{A}}{P}
+  \qquad
+  \jtermt{\ctxext{\Gamma}{A}}{\subst{\pair_A}{\ctxwk{A}{{\Gamma}{P}}}}{f}}
+  {\jtermteq{\ctxext{\Gamma}{A}}{\subst{\pair_A}{\ctxwk{A}{{\Gamma}{P}}}}{\subst{\pair_A}{\ctxwk{A}{{\Gamma}{\ind{\tfext_\Gamma(A)}(f)}}}}{f}}
+\end{infarray}
+
+Under the hypothesis that $\jtypeeq{\Gamma}{A}{A'}$
+we will also have the rules
+
+\begin{infarray}{c}
+\inference{}{\jtermteq{\ctxext{\Gamma}{A}}{\ctxwk{A}{{\Gamma}{\ctxext{\Gamma}{A}}}}{\pair_A}{\pair_{A'}}}\\
+\inference{\jtype{\ctxext{\Gamma}{A}}{P}\qquad\jtermt{\ctxext{\Gamma}{A}}{\subst{\pair_A}{\ctxwk{A}{{\Gamma}{P}}}}{f}}
+{\jtermteq{\ctxext{\Gamma}{A}}{P}{\ind{\tfext_\Gamma(A)}(f)}{\ind{\tfext_\Gamma(A')}(f)}}
+\end{infarray}
+
+Note that we don't need the notion of terms for contexts to state the universal
+property of context extension (which is a good thing, for we don't assume to have them).
+
+Next, we give the universal property of the extension $\ctxext{A}{P}$ in context
+$\Gamma$.
+In all of the following inference rules we assume that $\jtype{\ctxext{\Gamma}{A}}{P}$
+is among the hypothesis. The induction principle for extension consists of three
+inference rules:
+
+\begin{infarray}{c}
+\inference{}
+{\jtermt{\ctxext{{\Gamma}{A}}{P}}{\ctxwk{P}{{A}{\ctxext{A}{P}}}}{\pair_P}}\\
+\inference{
+  \jtype{\ctxext{\Gamma}{{A}{P}}}{Q}
+  \qquad
+  \jtermt{\ctxext{{\Gamma}{A}}{P}}{\subst{\pair_P}{\ctxwk{P}{{A}{Q}}}}{f}}
+  {\jtermt{\ctxext{\Gamma}{{A}{P}}}{Q}{\ind{\tfext_A(P)}(f)}}\\
+\inference{
+  \jtype{\ctxext{\Gamma}{{A}{P}}}{Q}
+  \qquad
+  \jtermt{\ctxext{{\Gamma}{A}}{P}}{\subst{\pair_P}{\ctxwk{P}{{A}{Q}}}}{f}}
+  {\jtermteq{\ctxext{{\Gamma}{A}}{P}}{\subst{\pair_P}{\ctxwk{P}{{A}{Q}}}}{\subst{\pair_P}{\ctxwk{P}{{A}{\ind{\tfext_A(P)}(f)}}}}{f}}
+\end{infarray}
+
+As with context extension, we shall require two more inference rules stating that
+$\pair_P$ and $\ind{\tfext_A(P)}$ are invariant under judgmental equality.
+
 \subsection{The variable rule -- identity functions}
-In the usual presentation of type theory, the variable rule asserts that ``in
-the context $x_1:A_1,...,x_n:A_n$ we have a term $x_i:A_i$ when $1\leq i\leq n$''.
+The variable rule for contexts:
+\begin{infarray}{c}
+\inference{\jctx{\Gamma}}{\jtermt{\Gamma}{\ctxwk{\Gamma}{\Gamma}}{\idfunc[\Gamma]}}\\
+\inference{\jtermt{\Gamma}{A}{x}}{\jtermteq{\Gamma}{A}{\subst{\idfunc[\Gamma]}{\ctxwk{\Gamma}{x}}}{x}}\\
+\inference{\jtermt{\Delta}{\ctxwk{\Delta}{\Gamma}}{g}}{\jtermteq{\Delta}{\ctxwk{\Delta}{\Gamma}}{\subst{g}{\ctxwk{\Delta}{\idfunc[\Gamma]}}}{g}}
+\end{infarray}
 
-In our current system, we don't need this generality. Instead, it suffices for
-us to assert that there are identity morphisms. We get arbitrary projections
-from weakening.
-
+The variable rule for types:
 \begin{infarray}{c}
@@
 
-%\begin{infarray}{c}
-%\inference{\jtype{\ctxext{{\Gamma}{A}}{P}}{Q}\qquad\jtermt{\Gamma}{A}{x}\qquad\jtermt{\ctxext{A
-%\end{infarray}{c}
-
-\subsection{Compatibility of extension, weakening and substitution}
-We describe the compatibility of extension, weakening and substitution. These
-compatibility rules are formulated in a weak sense.
-
-Extension is compatible with extension:
+\subsection{Identity types}
+We need to define the diagonal function $\jtermt{\ctxext{\Gamma}{A}}{\ctxwk{A}{\ctxext{A}{\ctxwk{A}{A}}}}{\delta}$
 \begin{infarray}{c}
-\inference
-{ \jtype{\ctxext{\Gamma}{A}}{P}
-}{\jtermt
-  {\ctxext{\Gamma}{{A}{P}}}
-  {\ctxwk{\ctxext{\Gamma}{{A}{P}}}{\ctxext{{\Gamma}{A}}{P}}}{\typefont{extext}_0(A,P)}
-}\\
-\inference
-{ \jtype{\ctxext{{\Gamma}{A}}{P}}{Q}
-}{\jtermt
-  {\ctxext{\Gamma}{{A}{{P}{Q}}}}
-  {\ctxwk{\ctxext{A}{{P}{Q}}}{\ctxext{{A}{P}}{Q}}}
-  {\typefont{extext}_1(A,P,Q)}
-}
+\inference{\jtype{\Gamma}{A}}{\jtype{\ctxext{{\Gamma}{A}}{\ctxwk{A}{A}}}{\idtypevar{A}}}\\
+\inference{\jtype{\Gamma}{A}}{\jtermt{\ctxext{\Gamma}{A}}{\delta^\ast\idtypevar{A}}{\reflsym}}\\
+\inference{\jtype{\ctxext{{{\Gamma}{A}}{\ctxwk{A}{A}}}{\idtypevar{A}}}{P}\qquad
+\jtermt{\ctxext{\Gamma}{A}}{\pairr{\delta,\reflsym}^\ast P}{d}}
+{\jtermt{\ctxext{{{\Gamma}{A}}{\ctxwk{A}{A}}}{\idtypevar{A}}}{P}{J}}
 \end{infarray}
 
-Weakening is compatible with extension:
-\begin{infarray}{c}
-\inference{
-  \jctx{\Gamma}
-  \qquad
-  \jtype{\Delta}{B}}
-  {\jtermt
-    {\ctxext{\Gamma}{\ctxwk{\Gamma}{\ctxext{\Delta}{B}}}}
-    {\ctxwk{{\Gamma}{\ctxext{\Delta}{B}}}{\ctxext{\ctxwk{\Gamma}{\Delta}}{\ctxwk{\Gamma}{B}}}}
-    {\typefont{wkext}_0(\Gamma,B)}}
+\subsection{Other type constructors}
+We will now describe the rules for the type constructors that we don't assume
+in our type theory by default. The first of these are the dependent function
+types. We also discuss $\tfW$-types, suspensions, truncations. More general higher inductive
+types will have to wait until we have introduced models, for the models of the
+basic type theory because we will use them as index categories of the diagrams.
+
+\subsubsection{Dependent function types}
+\begin{infarray}{cc}
+\inference{\jtype{\ctxext{\Gamma}{A}}{P}}{\jtype{\Gamma}{\mprd{A}{P}}}
+& \inference{\jtypeeq{\ctxext{\Gamma}{A}}{P}{P'}}{\jtypeeq{\Gamma}{\mprd{A}{P}}{\mprd{A}{P'}}}\\
+\inference{\jtermt{\ctxext{\Gamma}{A}}{P}{f}}{\jtermt{\Gamma}{\mprd{A}{P}}{\lambda(f)}}
+& \inference{\jtermteq{\ctxext{\Gamma}{A}}{P}{f}{f'}}{\jtermteq{\Gamma}{\mprd{A}{P}}{\lambda(f)}{\lambda(f')}}\\
+\inference{\jtermt{\Gamma}{\mprd{A}{P}}{g}}{\jtermt{\ctxext{\Gamma}{A}}{P}{\tfev(g)}}
+& \inference{\jtermteq{\Gamma}{\mprd{A}{P}}{g}{g'}}{\jtermteq{\ctxext{\Gamma}{A}}{P}{\tfev(g)}{\tfev(g')}}\\
+\inference{\jtermt{\ctxext{\Gamma}{A}}{P}{f}}{\jtermteq{\Gamma}{A}{\tfev(\lambda(f))}{f}}
 \end{infarray}
 
-Substitution is compatible with substitution:
-\begin{infarray}{c}
-\inference{\jtype{\ctxext{{\Gamma}{A}}{P}}{Q}\qquad\jtermt{\Gamma}{A}{x}\qquad\jtermt{\ctxext{\Gamma}{A}}{P}{f}}
-{\jtypeeq{\Gamma}{\subst{x}{\subst{f}{Q}}}{\subst{{x}{f}}{\subst{x}{Q}}}}
-\end{infarray}
-
-\subsection{More stuff in the type theory}
-\begin{infarray}{c}
-\inference{\jtype{\Gamma}{A}}{\jtermt{\ctxext{\Gamma}{A}}{\ctxwk{A}{A}}{\idfunc[A]}}
-\end{infarray}
+With these rules we will not get the weak $\eta$-rule when identity types are present.
+So it might be better to state that $\lambda$ is a trivial cofibration
 
@@
 if there is a term $\ctxext{{\Gamma}{A}}{P}$
-
-\subsection{The interactions of extension, weakening and substitution}
-
-\subsection{The basic constructors of type theory}
-\subsubsection{Dependent function types}
-\begin{infarray}{cc}
-\inference{\jtype{\ctxext{\Gamma}{A}}{P}}{\jtype{\Gamma}{\mprd{A}{P}}}
-& \inference{\jtypeeq{\ctxext{\Gamma}{A}}{P}{P'}}{\jtypeeq{\Gamma}{\mprd{A}{P}}{\mprd{A}{P'}}}\\
-\inference{\jtermt{\ctxext{\Gamma}{A}}{P}{f}}{\jtermt{\Gamma}{\mprd{A}{P}}{\lambda(f)}}
-& \inference{\jtermteq{\ctxext{\Gamma}{A}}{P}{f}{f'}}{\jtermteq{\Gamma}{\mprd{A}{P}}{\lambda(f)}{\lambda(f')}}\\
-\inference{\jtermt{\Gamma}{\mprd{A}{P}}{g}}{\jtermt{\ctxext{\Gamma}{A}}{P}{\tfev(g)}}
-& \inference{\jtermteq{\Gamma}{\mprd{A}{P}}{g}{g'}}{\jtermteq{\ctxext{\Gamma}{A}}{P}{\tfev(g)}{\tfev(g')}}\\
-\inference{\jtermt{\ctxext{\Gamma}{A}}{P}{f}}{\jtermteq{\Gamma}{A}{\tfev(\lambda(f))}{f}}
-\end{infarray}
-
-With these rules we will not get the weak $\eta$-rule when identity types are present.
-So it might be better to state that $\lambda$ is a trivial cofibration
-
-\subsection{Identity types}
-We need to define the diagonal function $\jtermt{\ctxext{\Gamma}{A}}{\ctxwk{A}{\ctxext{A}{\ctxwk{A}{A}}}}{\delta}$
-\begin{infarray}{c}
-\inference{\jtype{\Gamma}{A}}{\jtype{\ctxext{{\Gamma}{A}}{\ctxwk{A}{A}}}{\idtypevar{A}}}\\
-\inference{\jtype{\Gamma}{A}}{\jtermt{\ctxext{\Gamma}{A}}{\delta^\ast\idtypevar{A}}{\reflsym}}\\
-\inference{\jtype{\ctxext{{{\Gamma}{A}}{\ctxwk{A}{A}}}{\idtypevar{A}}}{P}\qquad
-\jtermt{\ctxext{\Gamma}{A}}{\pairr{\delta,\reflsym}^\ast P}{d}}
-{\jtermt{\ctxext{{{\Gamma}{A}}{\ctxwk{A}{A}}}{\idtypevar{A}}}{P}{J}}
-\end{infarray}

### 2013-10-29 article-models-tt.tex
Deleted the bit about Leibniz equality and trivial cofibrations since it seems we don't need them

@@
 With these rules we will not get the weak $\eta$-rule when identity types are present.
-So it might be better to state that $\lambda$ is a trivial cofibration
-
-\subsection{Leibniz equality on terms}
-Occasionally we shall need a weaker version of equality than judgmental equality.
-Our choice here is Leibniz equality, since it implies propositional equality when
-identity types are present.
-
-Let $P$ be a family over $A$ in context $\Gamma$. Then we can weaken
-$\ctxwk{A}{P}$, which is a family over $\ctxwk{A}{A}$ in context $\ctxext{\Gamma}{A}$.
-We also have the family $\ctxwk{{A}{A}}{P}$ over $\ctxwk{A}{A}$ in context
-$\ctxext{\Gamma}{A}$ and thus we have the inference
-\begin{equation*}
-\inference{\jtype{\ctxext{\Gamma}{A}}{P}}{\jtype{\ctxext{{{\Gamma}{A}}{\ctxwk{A}{A}}}{\ctxwk{A}{P}}}{\ctxwk{{A}{P}}{\ctxwk{{A}{A}}{P}}}}
-\end{equation*}
-This family plays the role of the family $\jtype{\Gamma,x,y:A}{P(x)\to P(y)}$
-in the type theory of \cite{TheBook}.
-
-\subsection{Trivial cofibrations and weak equivalences of types}
-We describe a relation between types that expresses when they are weakly equivalent.
-Weak equivalence is introduced because we need a weaker notion of judgmental 
-equality which also makes sense when identity types are not present, since that
-would allow us to state that context extension, weakening and substitution
-commute with each other.
-
-A term $f:\ctxwk{A}{B}$ is a trivial cofibration if it has the
-property that for any fibration $Q$ over $B$,
-to find a section of $Q$ it suffices to find a section of the fibration
-$f^\ast Q$ over $A$. In our type theoretical setting, the rôle of fibrations
-is played by families, the rôle of the function type $A\to B$ is played by
-$\ctxwk{A}{B}$ and our version of the pullback $f^\ast Q$ is $\subst{f}{\ctxwk{A}{Q}}$.
-
-\begin{defn}
-Let $f$ be a term of $\ctxwk{A}{B}$ in context $\Gamma$.
-\begin{enumerate}
-\item For a family $Q$ over $B$ in context $\Gamma$ we define $f^\ast Q\jdeq\subst{f}{\ctxwk{A}{Q}}$.
-\item For a term $g$ of $Q$ in context $\ctxext{\Gamma}{B}$ we define $f^\ast g\jdeq\subst{f}{\ctxwk{A}{g}}$.
-\end{enumerate} 
-\end{defn}
-\begin{rmk}
-The type $f^\ast(\ctxwk{B}{C})$ in context $\ctxext{\Gamma}{A}$ is can be viewed as the
-type of functions from $A$ to $C$ which factor through $f$. Thus there should be
-a function from $f^\ast(\ctxwk{B}{C})$ to $\ctxwk{A}{C}$.
-\end{rmk}
-
-\begin{rmk}
-Every term $\jtermt{\ctxext{\Gamma}{A}}{\ctxwk{A}{B}}{f}$ allows us to infer the following:
-\begin{equation*}
-\inference{\jtype{\ctxext{\Gamma}{B}}{Q}}{\jtype{\ctxext{\Gamma}{A}}{f^\ast Q}}
-\qquad
-\inference{\jtermt{\ctxext{\Gamma}{B}}{Q}{g}}{\jtermt{\ctxext{\Gamma}{A}}{f^\ast Q}{f^\ast g}}
-\end{equation*}
-\end{rmk}
-
-\begin{defn}
-A term $f:\ctxwk{A}{B}$ in context $\ctxext{\Gamma}{A}$ is said to be a trivial
-cofibration if we can infer
-\begin{equation*}
-\inference{\jtermt{\ctxext{\Gamma}{A}}{f^\ast Q}{t}}{\jtermt{\ctxext{\Gamma}{B}}{Q}{\tilde{t}}}\qquad
-\inference{\jtermt{\ctxext{\Gamma}{A}}{f^\ast Q}{t}}{\jtermteq{\ctxext{\Gamma}{A}}{f^\ast Q}{f^\ast \tilde{t}}{t}}
-\end{equation*}
-{\color{red}This statement should be reformulated so that it only involves a single judgment...
-but I don't see directly how to do that.}
-\end{defn}
-
-We have the following theorem in the type theory of \cite{TheBook}, which supports
-our claim that we may indeed speak of a trivial cofibration.
-
-\begin{thm}
-Suppose $f:A\to B$ is a function. Then $f$ is an equivalence if and only if
-for every $Q:B\to\type$ and every $g:\prd{x:A}Q(f(x))$ there is a section
-$h:\prd{y:B}Q(b)$ with the property that $h\circ f\htpy g$. 
-\end{thm}
-
-\begin{proof}
-We can first take $Q$ to be the constant family $\lam{y}A$. Furthermore, we may
-take $g\defeq\idfunc[A]$. Then we get a term of type
-\begin{equation*}
-\sm{h:B\to A}h\circ f\htpy \idfunc[A],
-\end{equation*}
-i.e.~we get a left inverse $h$ for $f$. To show that $h$ is also a right inverse
-of $f$, let $Q$ be the family $\lam{y}\id{f(h(y))}{y}$. To find a section of
-$Q$, which is the homotopy we aim for, it suffices to find a section of
-$Q\circ f$. In other words, we have to show that $\id{f(h(f(x)))}{f(x)}$ for
-every $x:A$. This follows from the fact that $h$ is a left inverse for $f$.
-
-The reverse direction is immediate.
-\end{proof}
-
-Another approach would be to define $f:\ctxwk{A}{B}$ to be left invertible
-if there is a term $\ctxext{{\Gamma}{A}}{P}$
+So it might be better to state that $\lambda$ is a trivial cofibration.

### 2013-10-30 article-models-defn.tex
funcy identity types

@@
 in which there is a term of $\terms{A}$ for every type $A$ in context $\Gamma$.
+This should work when univalence is around (so that functions can be replaced
+by families).
 
@@
 $\terms{\subst{f}{\ctxwk{A}{Q}}}\to\terms{Q}$ (asserting that $f$ is a trival
-cofibration). This should be equivalent to the previous condition.
+cofibration). This should always give that each $f$ is an equivalence and
+this should be equivalent to the previous condition when univalence is around.
+
+Another option arises when we see the identity type as an operation on a certain class of
+terms. Usually, $\idtypevar{}$ is defined using the class of identity functions.
+Let's take the class of all terms instead:
+
+\begin{infarray}{c}
+\inference{\jtermt{\Gamma}{A}{x}}{\jtype{\ctxext{\Gamma}{A}}{\eqtype{x}}}\\
+\inference{\jtermt{\Gamma}{A}{x}}{\jtermt{\Gamma}{\subst{x}{\eqtype{x}}}{\refl{x}}}\\
+\inference{\jtype{\ctxext{\Gamma}{A}}{P}\qquad\jtermt{\Gamma}{\subst{\refl{x}}{{x}{P}}}{d}}
+          {\jtermt{\ctxext{{\Gamma}{A}}{\eqtype{x}}}{P}{\ind{\eqtype{x}}(d)}}\\
+\inference{\jtype{\ctxext{\Gamma}{A}}{P}\qquad\jtermt{\Gamma}{\subst{\refl{x}}{{x}{P}}}{d}}
+          {\jtermteq{\ctxext{{\Gamma}{A}}{\eqtype{x}}}{\subst{\refl{x}}{{x}{P}}}{\subst{\refl{x}}{{x}{\ind{\eqtype{x}}(d)}}}{d}}
+\end{infarray}
+We get something that looks like Paulin-Mohring equality. But we can use it to
+show that every $\jtermt{\Gamma}{\ctxwk{\Gamma}{\Delta}}{f}$ is a trivial cofibration.
+
+Suppose that $\jtype{\Delta}{Q}$. Then we have $\jtype{\Gamma}{\subst{f}{\ctxwk{\Gamma}{Q}}}$. 
+For any term $\jtermt{\Gamma}{\subst{f}{\ctxwk{\Gamma}{Q}}}{d}$ we get a term 
+$\jtermt{\ctxext{{\Gamma}{\ctxwk{\Gamma}{\Delta}}}{\eqtype{f}}}{\ctxwk{\Gamma}{Q}}{\ind{\eqtype{f}}(d)}$.
+
+However, we could take other classes of terms, such as the projections
+$\proj1:\ctxwk{\ctxext{\Gamma}{A}}{\Gamma}$. Let's see how that goes:
+
+\begin{infarray}{c}
+\inference{\jtype{\Gamma}{A}}{\jtype{\ctxext{{\Gamma}{A}}{\ctxwk{\ctxext{\Gamma}{A}}{\Gamma}}}{\eqtype{\proj1^A}}}\\
+\inference{\jtype{\Gamma}{A}}{\jtermt{\ctxext{\Gamma}{A}}{\subst{\proj1^A}{\eqtype{\proj1^A}}}{\refl{\proj1^A}}}
+\end{infarray}

### 2013-10-30 article-models-tt.tex
funcy identity types

@@
 \subsection{Identity types}
-We need to define the diagonal function $\jtermt{\ctxext{\Gamma}{A}}{\ctxwk{A}{\ctxext{A}{\ctxwk{A}{A}}}}{\delta}$
+The identity types may now be formulated just as in ordinary Martin\nobreakdash-L\"of
+type theory. However, since all model categories have path objects, we will also
+introduce identity types for contexts.
+
+We will use the symbol $\reflsym$ slightly different than the book does. For us,
+$\refl{A}$ is a term of $\subst{\idfunc[A]}{\idtypevar{A}}$ in context
+$\ctxext{\Gamma}{A}$ and we will write $\subst{x}{\refl{A}}$ for the reflexivity
+path at $x$ (which would have been denoted by $\refl{x}$ in the book).
+
+\begin{infarray}{c}
+\inference{\jctx{\Gamma}}{\jtype{\ctxext{\Gamma}{\ctxwk{\Gamma}{\Gamma}}}{\idtypevar{\Gamma}}}\\
+\inference{\jctx{\Gamma}}{\jtermt{\Gamma}{\subst{\idfunc[\Gamma]}{\idtypevar{\Gamma}}}{\refl{\Gamma}}}\\
+\inference{\jtype{\ctxext{{\Gamma}{\ctxwk{\Gamma}{\Gamma}}}{\idtypevar{\Gamma}}}{P}
+           \qquad
+           \jtermt{\Gamma}{\subst{\refl{\Gamma}}{{\idfunc[\Gamma]}{P}}}{d}}
+           {\jtermt{\ctxext{{\Gamma}{\ctxwk{\Gamma}{\Gamma}}}{\idtypevar{\Gamma}}}{P}{\tfJ(d)}}\\
+\inference{\jtype{\ctxext{{\Gamma}{\ctxwk{\Gamma}{\Gamma}}}{\idtypevar{\Gamma}}}{P}\qquad
+\jtermt{\Gamma}{\subst{\refl{\Gamma}}{{\idfunc[\Gamma]}{P}}}{d}}
+{\jtermteq
+  {\ctxext{{\Gamma}{\ctxwk{\Gamma}{\Gamma}}}{\idtypevar{\Gamma}}}
+  {\subst{\refl{\Gamma}}{{\idfunc[\Gamma]}{P}}}
+  {\subst{\refl{\Gamma}}{{\idfunc[\Gamma]}{\tfJ(d)}}}
+  {d}}
+\end{infarray}
+
 \begin{infarray}{c}
 \inference{\jtype{\Gamma}{A}}{\jtype{\ctxext{{\Gamma}{A}}{\ctxwk{A}{A}}}{\idtypevar{A}}}\\
-\inference{\jtype{\Gamma}{A}}{\jtermt{\ctxext{\Gamma}{A}}{\delta^\ast\idtypevar{A}}{\reflsym}}\\
+\inference{\jtype{\Gamma}{A}}{\jtermt{\ctxext{\Gamma}{A}}{\subst{\idfunc[A]}{\idtypevar{A}}}{\refl{A}}}\\
 \inference{\jtype{\ctxext{{{\Gamma}{A}}{\ctxwk{A}{A}}}{\idtypevar{A}}}{P}\qquad
-\jtermt{\ctxext{\Gamma}{A}}{\pairr{\delta,\reflsym}^\ast P}{d}}
-{\jtermt{\ctxext{{{\Gamma}{A}}{\ctxwk{A}{A}}}{\idtypevar{A}}}{P}{J}}
+\jtermt{\ctxext{\Gamma}{A}}{\subst{\refl{A}}{{\idfunc[A]}{P}}}{d}}
+{\jtermt{\ctxext{{{\Gamma}{A}}{\ctxwk{A}{A}}}{\idtypevar{A}}}{P}{\tfJ(d)}}\\
+\inference{\jtype{\ctxext{{{\Gamma}{A}}{\ctxwk{A}{A}}}{\idtypevar{A}}}{P}\qquad
+\jtermt{\ctxext{\Gamma}{A}}{\subst{\refl{A}}{{\idfunc[A]}{P}}}{d}}
+{\jtermteq
+  {\ctxext{{{\Gamma}{A}}{\ctxwk{A}{A}}}{\idtypevar{A}}}
+  {\subst{\refl{A}}{{\idfunc[A]}{P}}}
+  {\subst{\refl{A}}{{\idfunc[A]}{\tfJ(d)}}}
+  {d}}
 \end{infarray}

### 2013-10-31 article-models-tt.tex
compatibility rules for weakening

@@
 \subsubsection{Weakening is compatible with itself}
-Suppose $\jtype{\ctxext{\Gamma}{A}}{P}$ and $\jtype{\Gamma}{B}$. Then we can
-consider the weakening $\ctxwk{A}{B}$ over $A$ and consecutively the weakening
-$\ctxwk{P}{{A}{B}}$. This should be isomorphic to $\ctxwk{\ctxext{A}{P}}{B}$.
+To conclude the rules for weakening, we state judgmental equality rules expressing
+that weakening is compatible with itself. These rules state that the following
+diagram commutes given any two contexts $\Gamma$ and $\Delta$:
+\begin{equation*}
+\begin{tikzcd}[column sep=huge]
+\jctx{\blank} \ar{r}{\greek{E}\mapsto\ctxwk{\Delta}{E}} \ar{d}[swap]{\greek{E}\mapsto\ctxwk{\Gamma}{E}} & \jtype{\Delta}{\blank} \ar{d}{B\mapsto\ctxwk{\Gamma}{B}}\\
+\jtype{\Gamma}{\blank} \ar{r}[swap]{A\mapsto\ctxwk{{\Gamma}{\Delta}}{A}} & \jtype{\ctxext{\Gamma}{\ctxwk{\Gamma}{\Delta}}}{\blank}
+\end{tikzcd}
+\end{equation*}
+There is also a version of this diagram in which all happens in a context. Thus,
+we get two sets of inference rules. For weakening by contexts we get:
 
-\subsubsection{Weakening is compatible with extension}
-Suppose $\jtype{\Gamma}{A}$ and $\jtype{\ctxext{\Gamma}{B}}{Q}$. Then
-\begin{equation*}
-s:\ctxwk{{A}{\ctxext{B}{Q}}}{\ctxext{\ctxwk{A}{B}}{\ctxwk{A}{Q}}}
-\end{equation*}
+\begin{infarray}{c}
+\inference{\jctx{\Gamma}\qquad\jctx{\Delta}\qquad\jctx{\greek{E}}}
+          {\jtypeeq{\ctxext{\Gamma}{\ctxwk{\Gamma}{\Delta}}}{\ctxwk{\Gamma}{{\Delta}{\greek{E}}}}
+            {\ctxwk{{\Gamma}{\Delta}}{{\Gamma}{\greek{E}}}}}\\
+\inference{\jctx{\Gamma}\qquad\jctx{\Delta}\qquad\jtype{\greek{E}}{C}}
+          {\jtypeeq{\ctxext{{\Gamma}{\ctxwk{\Gamma}{\Delta}}}{\ctxwk{{\Gamma}{\Delta}}{{\Gamma}{\greek{E}}}}}
+            {\ctxwk{\Gamma}{{\Delta}{C}}}
+            {\ctxwk{{\Gamma}{\Delta}}{{\Gamma}{C}}}}\\
+\inference{\jctx{\Gamma}\qquad\jctx{\Delta}\qquad\jtermt{\greek{E}}{C}{t}}
+          {\jtermteq{\ctxext{{\Gamma}{\ctxwk{\Gamma}{\Delta}}}{\ctxwk{{\Gamma}{\Delta}}{{\Gamma}{\greek{E}}}}}
+            {\ctxwk{{\Gamma}{\Delta}}{{\Gamma}{C}}}{\ctxwk{\Gamma}{{\Delta}{t}}}{\ctxwk{{\Gamma}{\Delta}}{{\Gamma}{t}}}}
+\end{infarray}
+
+For weakening by types we get:
+
+\begin{infarray}{c}
+\inference{\jtype{\Gamma}{A}\qquad\jtype{\Gamma}{B}\qquad\jtype{\Gamma}{C}}
+          {\jtypeeq{\ctxext{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{{B}{C}}}
+            {\ctxwk{{A}{B}}{{A}{C}}}}\\
+\inference{\jtype{\Gamma}{A}\qquad\jtype{\Gamma}{B}\qquad\jtype{\ctxext{\Gamma}{C}}{R}}
+          {\jtypeeq{\ctxext{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{{A}{B}}{{A}{C}}}}
+            {\ctxwk{A}{{B}{R}}}
+            {\ctxwk{{A}{B}}{{A}{R}}}}\\
+\inference{\jtype{\Gamma}{A}\qquad\jtype{\Gamma}{B}\qquad\jtermt{\Gamma}{R}{t}}
+          {\jtermteq{\ctxext{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{{A}{B}}{{A}{C}}}}
+            {\ctxwk{{A}{B}}{{A}{R}}}{\ctxwk{A}{{B}{t}}}{\ctxwk{{A}{B}}{{A}{t}}}}
+\end{infarray}
 

### 2013-10-31 article-models-defn.tex
equality types

@@
 
-\subsection{Weak $\omega$-groupoids}
+\section{Weak $\omega$-groupoids}
+
+\subsection{Trivial cofibrations and weak equivalences of types}
+We describe a relation between types that expresses when they are weakly equivalent.
+Weak equivalence is introduced because we need a weaker notion of judgmental 
+equality which also makes sense when identity types are not present, since that
+would allow us to state that context extension, weakening and substitution
+commute with each other.
+
+A term $f:\ctxwk{A}{B}$ is a trivial cofibration if it has the
+property that for any fibration $Q$ over $B$,
+to find a section of $Q$ it suffices to find a section of the fibration
+$f^\ast Q$ over $A$. In our type theoretical setting, the rôle of fibrations
+is played by families, the rôle of the function type $A\to B$ is played by
+$\ctxwk{A}{B}$ and our version of the pullback $f^\ast Q$ is $\subst{f}{\ctxwk{A}{Q}}$.
+
+\begin{defn}
+Let $f$ be a term of $\ctxwk{A}{B}$ in context $\Gamma$.
+\begin{enumerate}
+\item For a family $Q$ over $B$ in context $\Gamma$ we define $f^\ast Q\jdeq\subst{f}{\ctxwk{A}{Q}}$.
+\item For a term $g$ of $Q$ in context $\ctxext{\Gamma}{B}$ we define $f^\ast g\jdeq\subst{f}{\ctxwk{A}{g}}$.
+\end{enumerate} 
+\end{defn}
+\begin{rmk}
+The type $f^\ast(\ctxwk{B}{C})$ in context $\ctxext{\Gamma}{A}$ is can be viewed as the
+type of functions from $A$ to $C$ which factor through $f$. Thus there should be
+a function from $f^\ast(\ctxwk{B}{C})$ to $\ctxwk{A}{C}$.
+\end{rmk}
+
+\begin{rmk}
+Every term $\jtermt{\ctxext{\Gamma}{A}}{\ctxwk{A}{B}}{f}$ allows us to infer the following:
+\begin{equation*}
+\inference{\jtype{\ctxext{\Gamma}{B}}{Q}}{\jtype{\ctxext{\Gamma}{A}}{f^\ast Q}}
+\qquad
+\inference{\jtermt{\ctxext{\Gamma}{B}}{Q}{g}}{\jtermt{\ctxext{\Gamma}{A}}{f^\ast Q}{f^\ast g}}
+\end{equation*}
+\end{rmk}
+
+\begin{defn}
+A term $f:\ctxwk{A}{B}$ in context $\ctxext{\Gamma}{A}$ is said to be a trivial
+cofibration if we can infer
+\begin{equation*}
+\inference{\jtermt{\ctxext{\Gamma}{A}}{f^\ast Q}{t}}{\jtermt{\ctxext{\Gamma}{B}}{Q}{\tilde{t}}}\qquad
+\inference{\jtermt{\ctxext{\Gamma}{A}}{f^\ast Q}{t}}{\jtermteq{\ctxext{\Gamma}{A}}{f^\ast Q}{f^\ast \tilde{t}}{t}}
+\end{equation*}
+{\color{red}This statement should be reformulated so that it only involves a single judgment...
+but I don't see directly how to do that.}
+\end{defn}
+
+We have the following theorem in the type theory of \cite{TheBook}, which supports
+our claim that we may indeed speak of a trivial cofibration.
+
+\begin{thm}
+Suppose $f:A\to B$ is a function. Then $f$ is an equivalence if and only if
+for every $Q:B\to\type$ and every $g:\prd{x:A}Q(f(x))$ there is a section
+$h:\prd{y:B}Q(b)$ with the property that $h\circ f\htpy g$. 
+\end{thm}
+
+\begin{proof}
+We can first take $Q$ to be the constant family $\lam{y}A$. Furthermore, we may
+take $g\defeq\idfunc[A]$. Then we get a term of type
+\begin{equation*}
+\sm{h:B\to A}h\circ f\htpy \idfunc[A],
+\end{equation*}
+i.e.~we get a left inverse $h$ for $f$. To show that $h$ is also a right inverse
+of $f$, let $Q$ be the family $\lam{y}\id{f(h(y))}{y}$. To find a section of
+$Q$, which is the homotopy we aim for, it suffices to find a section of
+$Q\circ f$. In other words, we have to show that $\id{f(h(f(x)))}{f(x)}$ for
+every $x:A$. This follows from the fact that $h$ is a left inverse for $f$.
+
+The reverse direction is immediate.
+\end{proof}
+
+Another approach would be to define $f:\ctxwk{A}{B}$ to be left invertible
+if there is a term $\ctxext{{\Gamma}{A}}{P}$
+
+\subsection{Possible definitions of weak $\omega$-groupoids}
 
@@
 \inference{\jtermt{\Gamma}{A}{x}}{\jtermt{\Gamma}{\subst{x}{\eqtype{x}}}{\refl{x}}}\\
-\inference{\jtype{\ctxext{\Gamma}{A}}{P}\qquad\jtermt{\Gamma}{\subst{\refl{x}}{{x}{P}}}{d}}
+\inference{\jtype{\ctxext{{\Gamma}{A}}{\eqtype{x}}}{P}\qquad\jtermt{\Gamma}{\subst{\refl{x}}{{x}{P}}}{d}}
           {\jtermt{\ctxext{{\Gamma}{A}}{\eqtype{x}}}{P}{\ind{\eqtype{x}}(d)}}\\
-\inference{\jtype{\ctxext{\Gamma}{A}}{P}\qquad\jtermt{\Gamma}{\subst{\refl{x}}{{x}{P}}}{d}}
+\inference{\jtype{\ctxext{{\Gamma}{A}}{\eqtype{x}}}{P}\qquad\jtermt{\Gamma}{\subst{\refl{x}}{{x}{P}}}{d}}
           {\jtermteq{\ctxext{{\Gamma}{A}}{\eqtype{x}}}{\subst{\refl{x}}{{x}{P}}}{\subst{\refl{x}}{{x}{\ind{\eqtype{x}}(d)}}}{d}}
 \end{infarray}
-We get something that looks like Paulin-Mohring equality. But we can use it to
+
+We get something that looks like Paulin-Mohring equality. But we should be able to use it to
 show that every $\jtermt{\Gamma}{\ctxwk{\Gamma}{\Delta}}{f}$ is a trivial cofibration.
 
-Suppose that $\jtype{\Delta}{Q}$. Then we have $\jtype{\Gamma}{\subst{f}{\ctxwk{\Gamma}{Q}}}$. 
-For any term $\jtermt{\Gamma}{\subst{f}{\ctxwk{\Gamma}{Q}}}{d}$ we get a term 
+For the following conjecture, note that if we have $\jtermt{\Gamma}{\ctxwk{\Gamma}{\Delta}}{f}$
+Then we may consider the type $\jtype{\ctxext{{\Gamma}{\ctxwk{\Gamma}{\Delta}}}{\ctxwk{\Gamma}{{\Delta}{\Delta}}}}
+{\ctxwk{\Gamma}{\idtypevar{\Delta}}}$ and we may substitute $f$ to obtain the type
+\begin{equation*}
+\jtype{\ctxext{\Gamma}{\subst{f}{\ctxwk{\Gamma}{{\Delta}{\Delta}}}}}
+{\subst{f}{\ctxwk{\Gamma}{\idtypevar{\Delta}}}}
+\end{equation*}
+Note that we have the judgmental equality
+\begin{equation*}
+\jtypeeq{\Gamma}{\subst{f}{\ctxwk{\Gamma}{{\Delta}{\Delta}}}}{\ctxwk{\Gamma}{\Delta}}
+\end{equation*}
+so we obtain the type
+\begin{equation*}
+\jtype{\ctxext{\Gamma}{\ctxwk{\Gamma}{\Delta}}}
+{\subst{f}{\ctxwk{\Gamma}{\idtypevar{\Delta}}}}
+\end{equation*}
+We also have $\jtype{\ctxext{\Gamma}{\ctxwk{\Gamma}{\Delta}}}{\eqtype{f}}$, so we
+may ask for a term of type $\ctxwk{\eqtype{f}}{\subst{f}{\ctxwk{\Gamma}{\idtypevar{\Delta}}}}$
+in context $\ctxext{\Gamma}{\ctxwk{\Gamma}{\Delta}}$ and we can ask ourselves the question
+whether this is a trivial cofibration. This is the content of the following conjecture.
+
+\begin{conj}
+There is a term
+\begin{equation*}
+\jtermt{\ctxext{{\Gamma}{\ctxwk{\Gamma}{\Delta}}}{\eqtype{f}}}{\ctxwk{\eqtype{f}}{\subst{f}{\ctxwk{\Gamma}{\idtypevar{\Delta}}}}}{\eqtoid{f}}
+\end{equation*}
+Moreover, $\eqtoid{f}$ is a trivial cofibration.
+\end{conj}
+
+\begin{proof}
+We use the induction principle of $\eqtype{f}$ to construct $\eqtoid{f}$. Thus, we
+have to find a term
+\begin{equation*}
+\jtermt
+  {\Gamma}
+  {\subst{\refl{f}}{{f}{\ctxwk{\eqtype{f}}{\subst{f}{\ctxwk{\Gamma}{\idtypevar{\Delta}}}}}}}
+  {\subst{\refl{f}}{{f}{\eqtoid{f}}}}
+\end{equation*}
+Note that we have the judgmental equalities
+\begin{align*}
+\subst{\refl{f}}{{f}{\ctxwk{\eqtype{f}}{\subst{f}{\ctxwk{\Gamma}{\idtypevar{\Delta}}}}}}
+& \jdeq \subst{f}{{\ctxwk{{\Gamma}{\Delta}}{\refl{f}}}{\ctxwk{\eqtype{f}}{\subst{f}{\ctxwk{\Gamma}{\idtypevar{\Delta}}}}}} \\
+& \jdeq \subst{f}{\ctxwk{\Gamma}{\subst{f}{\ctxwk{\Gamma}{\idtypevar{\Delta}}}}}
+\end{align*}
+where we find the term $\subst{f}{\ctxwk{\Gamma}{\subst{f}{\ctxwk{\Gamma}{\refl{\Delta}}}}}$.
+\end{proof}
+
+\begin{conj}
+Suppose we have a type $\eqtype{x}$ for every term $\jtermt{\Gamma}{A}{x}$. Then
+every term $\jtermt{\Gamma}{\ctxwk{\Gamma}{\Delta}}{f}$ is a trivial cofibration.
+\end{conj}
+
+\begin{proof}
+Suppose that $Q$ is a type in context $\Delta$. We want to show that
+\begin{equation*}
+\inference{\jtermt{\Gamma}{\subst{f}{\ctxwk{\Gamma}{Q}}}{g}}{\jtermt{\Delta}{Q}{\tilde{g}}}
+\end{equation*}
+Let $g$ be a term of $\subst{f}{\ctxwk{\Gamma}{Q}}$ in context $\Gamma$. We get
+the term
+\begin{equation*}
+\jtermt{\Gamma}{\subst{\refl{f}}{{f}{\ctxwk{\eqtype{f}}{{\Gamma}{Q}}}}}{\subst{\refl{f}}{\ctxwk{\eqtype{f}}{g}}}
+\end{equation*}
+which is judgmentally equal to $g$; the types
+
+For any term $\jtermt{\Gamma}{\subst{f}{\ctxwk{\Gamma}{Q}}}{g}$ we get a term 
+$\jtermt{\Gamma}{\subst{\refl{x}}{{x}{P}}}{d}$.
+
 $\jtermt{\ctxext{{\Gamma}{\ctxwk{\Gamma}{\Delta}}}{\eqtype{f}}}{\ctxwk{\Gamma}{Q}}{\ind{\eqtype{f}}(d)}$.
+\end{proof}
 

### 2013-10-31 article-models-defn.tex
back to \jterm for terms

@@
 \begin{rmk}
-Every term $\jtermt{\ctxext{\Gamma}{A}}{\ctxwk{A}{B}}{f}$ allows us to infer the following:
+Every term $\jterm{\ctxext{\Gamma}{A}}{\ctxwk{A}{B}}{f}$ allows us to infer the following:
 \begin{equation*}
@@
 \qquad
-\inference{\jtermt{\ctxext{\Gamma}{B}}{Q}{g}}{\jtermt{\ctxext{\Gamma}{A}}{f^\ast Q}{f^\ast g}}
+\inference{\jterm{\ctxext{\Gamma}{B}}{Q}{g}}{\jterm{\ctxext{\Gamma}{A}}{f^\ast Q}{f^\ast g}}
 \end{equation*}
@@
 \begin{equation*}
-\inference{\jtermt{\ctxext{\Gamma}{A}}{f^\ast Q}{t}}{\jtermt{\ctxext{\Gamma}{B}}{Q}{\tilde{t}}}\qquad
-\inference{\jtermt{\ctxext{\Gamma}{A}}{f^\ast Q}{t}}{\jtermteq{\ctxext{\Gamma}{A}}{f^\ast Q}{f^\ast \tilde{t}}{t}}
+\inference{\jterm{\ctxext{\Gamma}{A}}{f^\ast Q}{t}}{\jterm{\ctxext{\Gamma}{B}}{Q}{\tilde{t}}}\qquad
+\inference{\jterm{\ctxext{\Gamma}{A}}{f^\ast Q}{t}}{\jtermeq{\ctxext{\Gamma}{A}}{f^\ast Q}{f^\ast \tilde{t}}{t}}
 \end{equation*}
@@
 \begin{infarray}{c}
-\inference{\jtermt{\Gamma}{A}{x}}{\jtype{\ctxext{\Gamma}{A}}{\eqtype{x}}}\\
-\inference{\jtermt{\Gamma}{A}{x}}{\jtermt{\Gamma}{\subst{x}{\eqtype{x}}}{\refl{x}}}\\
-\inference{\jtype{\ctxext{{\Gamma}{A}}{\eqtype{x}}}{P}\qquad\jtermt{\Gamma}{\subst{\refl{x}}{{x}{P}}}{d}}
-          {\jtermt{\ctxext{{\Gamma}{A}}{\eqtype{x}}}{P}{\ind{\eqtype{x}}(d)}}\\
-\inference{\jtype{\ctxext{{\Gamma}{A}}{\eqtype{x}}}{P}\qquad\jtermt{\Gamma}{\subst{\refl{x}}{{x}{P}}}{d}}
-          {\jtermteq{\ctxext{{\Gamma}{A}}{\eqtype{x}}}{\subst{\refl{x}}{{x}{P}}}{\subst{\refl{x}}{{x}{\ind{\eqtype{x}}(d)}}}{d}}
+\inference{\jterm{\Gamma}{A}{x}}{\jtype{\ctxext{\Gamma}{A}}{\eqtype{x}}}\\
+\inference{\jterm{\Gamma}{A}{x}}{\jterm{\Gamma}{\subst{x}{\eqtype{x}}}{\refl{x}}}\\
+\inference{\jtype{\ctxext{{\Gamma}{A}}{\eqtype{x}}}{P}\qquad\jterm{\Gamma}{\subst{\refl{x}}{{x}{P}}}{d}}
+          {\jterm{\ctxext{{\Gamma}{A}}{\eqtype{x}}}{P}{\ind{\eqtype{x}}(d)}}\\
+\inference{\jtype{\ctxext{{\Gamma}{A}}{\eqtype{x}}}{P}\qquad\jterm{\Gamma}{\subst{\refl{x}}{{x}{P}}}{d}}
+          {\jtermeq{\ctxext{{\Gamma}{A}}{\eqtype{x}}}{\subst{\refl{x}}{{x}{P}}}{\subst{\refl{x}}{{x}{\ind{\eqtype{x}}(d)}}}{d}}
 \end{infarray}
@@
 We get something that looks like Paulin-Mohring equality. But we should be able to use it to
-show that every $\jtermt{\Gamma}{\ctxwk{\Gamma}{\Delta}}{f}$ is a trivial cofibration.
+show that every $\jterm{\Gamma}{\ctxwk{\Gamma}{\Delta}}{f}$ is a trivial cofibration.
 
-For the following conjecture, note that if we have $\jtermt{\Gamma}{\ctxwk{\Gamma}{\Delta}}{f}$
+For the following conjecture, note that if we have $\jterm{\Gamma}{\ctxwk{\Gamma}{\Delta}}{f}$
 Then we may consider the type $\jtype{\ctxext{{\Gamma}{\ctxwk{\Gamma}{\Delta}}}{\ctxwk{\Gamma}{{\Delta}{\Delta}}}}
@@
 \begin{equation*}
-\jtermt{\ctxext{{\Gamma}{\ctxwk{\Gamma}{\Delta}}}{\eqtype{f}}}{\ctxwk{\eqtype{f}}{\subst{f}{\ctxwk{\Gamma}{\idtypevar{\Delta}}}}}{\eqtoid{f}}
+\jterm{\ctxext{{\Gamma}{\ctxwk{\Gamma}{\Delta}}}{\eqtype{f}}}{\ctxwk{\eqtype{f}}{\subst{f}{\ctxwk{\Gamma}{\idtypevar{\Delta}}}}}{\eqtoid{f}}
 \end{equation*}
@@
 \begin{equation*}
-\jtermt
+\jterm
   {\Gamma}
@@
 \begin{conj}
-Suppose we have a type $\eqtype{x}$ for every term $\jtermt{\Gamma}{A}{x}$. Then
-every term $\jtermt{\Gamma}{\ctxwk{\Gamma}{\Delta}}{f}$ is a trivial cofibration.
+Suppose we have a type $\eqtype{x}$ for every term $\jterm{\Gamma}{A}{x}$. Then
+every term $\jterm{\Gamma}{\ctxwk{\Gamma}{\Delta}}{f}$ is a trivial cofibration.
 \end{conj}
@@
 \begin{equation*}
-\inference{\jtermt{\Gamma}{\subst{f}{\ctxwk{\Gamma}{Q}}}{g}}{\jtermt{\Delta}{Q}{\tilde{g}}}
+\inference{\jterm{\Gamma}{\subst{f}{\ctxwk{\Gamma}{Q}}}{g}}{\jterm{\Delta}{Q}{\tilde{g}}}
 \end{equation*}
@@
 \begin{equation*}
-\jtermt{\Gamma}{\subst{\refl{f}}{{f}{\ctxwk{\eqtype{f}}{{\Gamma}{Q}}}}}{\subst{\refl{f}}{\ctxwk{\eqtype{f}}{g}}}
+\jterm{\Gamma}{\subst{\refl{f}}{{f}{\ctxwk{\eqtype{f}}{{\Gamma}{Q}}}}}{\subst{\refl{f}}{\ctxwk{\eqtype{f}}{g}}}
 \end{equation*}
@@
 
-For any term $\jtermt{\Gamma}{\subst{f}{\ctxwk{\Gamma}{Q}}}{g}$ we get a term 
-$\jtermt{\Gamma}{\subst{\refl{x}}{{x}{P}}}{d}$.
+For any term $\jterm{\Gamma}{\subst{f}{\ctxwk{\Gamma}{Q}}}{g}$ we get a term 
+$\jterm{\Gamma}{\subst{\refl{x}}{{x}{P}}}{d}$.
 
-$\jtermt{\ctxext{{\Gamma}{\ctxwk{\Gamma}{\Delta}}}{\eqtype{f}}}{\ctxwk{\Gamma}{Q}}{\ind{\eqtype{f}}(d)}$.
+$\jterm{\ctxext{{\Gamma}{\ctxwk{\Gamma}{\Delta}}}{\eqtype{f}}}{\ctxwk{\Gamma}{Q}}{\ind{\eqtype{f}}(d)}$.
 \end{proof}
@@
 \inference{\jtype{\Gamma}{A}}{\jtype{\ctxext{{\Gamma}{A}}{\ctxwk{\ctxext{\Gamma}{A}}{\Gamma}}}{\eqtype{\proj1^A}}}\\
-\inference{\jtype{\Gamma}{A}}{\jtermt{\ctxext{\Gamma}{A}}{\subst{\proj1^A}{\eqtype{\proj1^A}}}{\refl{\proj1^A}}}
+\inference{\jtype{\Gamma}{A}}{\jterm{\ctxext{\Gamma}{A}}{\subst{\proj1^A}{\eqtype{\proj1^A}}}{\refl{\proj1^A}}}
 \end{infarray}

### 2013-10-31 article-models-tt.tex
back to \jterm for terms

@@
 \jtype*{\Gamma}{A} & \jtypeeq*{\Gamma}{A}{B}\\
-\jtermt*{\Gamma}{A}{x} & \jtermteq*{\Gamma}{A}{x}{y}.
+\jterm*{\Gamma}{A}{x} & \jtermeq*{\Gamma}{A}{x}{y}.
 \end{align*}
@@
 \inference{\jtypeeq{\Gamma}{A}{B}\qquad\jtypeeq{\Gamma}{B}{C}}{\jtypeeq{\Gamma}{A}{C}}\\
-\inference{\jtermt{\Gamma}{A}{x}}{\jtermteq{\Gamma}{A}{x}{x}} & 
-\inference{\jtermteq{\Gamma}{A}{x}{y}}{\jtermteq{\Gamma}{A}{y}{x}} &
-\inference{\jtermteq{\Gamma}{A}{x}{y}\qquad\jtermteq{\Gamma}{A}{y}{z}}{\jtermteq{\Gamma}{A}{x}{z}}
+\inference{\jterm{\Gamma}{A}{x}}{\jtermeq{\Gamma}{A}{x}{x}} & 
+\inference{\jtermeq{\Gamma}{A}{x}{y}}{\jtermeq{\Gamma}{A}{y}{x}} &
+\inference{\jtermeq{\Gamma}{A}{x}{y}\qquad\jtermeq{\Gamma}{A}{y}{z}}{\jtermeq{\Gamma}{A}{x}{z}}
 \end{infarray}
@@
 & \inference{\jctxeq{\Gamma}{\Delta}\qquad\jtypeeq{\Gamma}{A}{B}}{\jtypeeq{\Delta}{A}{B}}\\
-\inference{\jctxeq{\Gamma}{\Delta}\qquad\jtermt{\Gamma}{A}{x}}{\jtermt{\Delta}{A}{x}}
-& \inference{\jctxeq{\Gamma}{\Delta}\qquad\jtermteq{\Gamma}{A}{x}{y}}{\jtermteq{\Delta}{A}{x}{y}}\\
-\inference{\jtypeeq{\Gamma}{A}{B}\qquad \jtermt{\Gamma}{A}{x}}{\jtermt{\Gamma}{B}{x}}
-& \inference{\jtypeeq{\Gamma}{A}{B}\qquad\jtermteq{\Gamma}{A}{x}{y}}{\jtermteq{\Gamma}{B}{x}{y}}
+\inference{\jctxeq{\Gamma}{\Delta}\qquad\jterm{\Gamma}{A}{x}}{\jterm{\Delta}{A}{x}}
+& \inference{\jctxeq{\Gamma}{\Delta}\qquad\jtermeq{\Gamma}{A}{x}{y}}{\jtermeq{\Delta}{A}{x}{y}}\\
+\inference{\jtypeeq{\Gamma}{A}{B}\qquad \jterm{\Gamma}{A}{x}}{\jterm{\Gamma}{B}{x}}
+& \inference{\jtypeeq{\Gamma}{A}{B}\qquad\jtermeq{\Gamma}{A}{x}{y}}{\jtermeq{\Gamma}{B}{x}{y}}
 \end{infarray}
@@
 \inference{\jctx{\Gamma}\qquad\jtype{\Delta}{B}}{\jtype{\ctxext{\Gamma}{\ctxwk{\Gamma}{\Delta}}}{\ctxwk{\Gamma}{B}}} & \inference{\jctxeq{\Gamma}{\Gamma'}\qquad\jtypeeq{\Delta}{B}{B'}}{\jtypeeq{\ctxext{\Gamma}{\ctxwk{\Gamma}{\Delta}}}{\ctxwk{\Gamma}{B}}{\ctxwk{\Gamma'}{B'}}}\\
-\inference{\jctx{\Gamma}\qquad\jtermt{\Delta}{B}{y}}{\jtermt{\ctxext{\Gamma}{\ctxwk{\Gamma}{\Delta}}}{\ctxwk{\Gamma}{B}}{\ctxwk{\Gamma}{y}}} 
-& \inference{\jctxeq{\Gamma}{\Gamma'}\qquad\jtermteq{\Delta}{B}{y}{y'}}{\jtermteq{\ctxext{\Gamma}{\ctxwk{\Gamma}{\Delta}}}{\ctxwk{\Gamma}{B}}{\ctxwk{\Gamma}{y}}{\ctxwk{\Gamma'}{y'}}} 
+\inference{\jctx{\Gamma}\qquad\jterm{\Delta}{B}{y}}{\jterm{\ctxext{\Gamma}{\ctxwk{\Gamma}{\Delta}}}{\ctxwk{\Gamma}{B}}{\ctxwk{\Gamma}{y}}} 
+& \inference{\jctxeq{\Gamma}{\Gamma'}\qquad\jtermeq{\Delta}{B}{y}{y'}}{\jtermeq{\ctxext{\Gamma}{\ctxwk{\Gamma}{\Delta}}}{\ctxwk{\Gamma}{B}}{\ctxwk{\Gamma}{y}}{\ctxwk{\Gamma'}{y'}}} 
 \end{infarray}
@@
 {\jtypeeq{\ctxext{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{Q}}{\ctxwk{A'}{Q'}}}\\
-\inference{\jtype{\Gamma}{A}\qquad\jtermt{\Gamma}{B}{y}}{\jtermt{\ctxext{\Gamma}{A}}{\ctxwk{A}{B}}{\ctxwk{A}{y}}}
-& \inference{\jtypeeq{\Gamma}{A}{A'}\qquad\jtermteq{\Gamma}{B}{y}{y'}}{\jtermteq{\ctxext{\Gamma}{A}}{\ctxwk{A}{B}}{\ctxwk{A}{y}}{\ctxwk{A'}{y'}}}
+\inference{\jtype{\Gamma}{A}\qquad\jterm{\Gamma}{B}{y}}{\jterm{\ctxext{\Gamma}{A}}{\ctxwk{A}{B}}{\ctxwk{A}{y}}}
+& \inference{\jtypeeq{\Gamma}{A}{A'}\qquad\jtermeq{\Gamma}{B}{y}{y'}}{\jtermeq{\ctxext{\Gamma}{A}}{\ctxwk{A}{B}}{\ctxwk{A}{y}}{\ctxwk{A'}{y'}}}
 \end{infarray}
@@
             {\ctxwk{{\Gamma}{\Delta}}{{\Gamma}{C}}}}\\
-\inference{\jctx{\Gamma}\qquad\jctx{\Delta}\qquad\jtermt{\greek{E}}{C}{t}}
-          {\jtermteq{\ctxext{{\Gamma}{\ctxwk{\Gamma}{\Delta}}}{\ctxwk{{\Gamma}{\Delta}}{{\Gamma}{\greek{E}}}}}
+\inference{\jctx{\Gamma}\qquad\jctx{\Delta}\qquad\jterm{\greek{E}}{C}{t}}
+          {\jtermeq{\ctxext{{\Gamma}{\ctxwk{\Gamma}{\Delta}}}{\ctxwk{{\Gamma}{\Delta}}{{\Gamma}{\greek{E}}}}}
             {\ctxwk{{\Gamma}{\Delta}}{{\Gamma}{C}}}{\ctxwk{\Gamma}{{\Delta}{t}}}{\ctxwk{{\Gamma}{\Delta}}{{\Gamma}{t}}}}
@@
             {\ctxwk{{A}{B}}{{A}{R}}}}\\
-\inference{\jtype{\Gamma}{A}\qquad\jtype{\Gamma}{B}\qquad\jtermt{\Gamma}{R}{t}}
-          {\jtermteq{\ctxext{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{{A}{B}}{{A}{C}}}}
+\inference{\jtype{\Gamma}{A}\qquad\jtype{\Gamma}{B}\qquad\jterm{\Gamma}{R}{t}}
+          {\jtermeq{\ctxext{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{{A}{B}}{{A}{C}}}}
             {\ctxwk{{A}{B}}{{A}{R}}}{\ctxwk{A}{{B}{t}}}{\ctxwk{{A}{B}}{{A}{t}}}}
@@
 
-In the rules introducing the various substitutions we assume $\jtermt{\Gamma}{A}{x}$;
-in the rules introducing the definitional equalities we assume $\jtermteq{\Gamma}{A}{x}{x'}$.
+In the rules introducing the various substitutions we assume $\jterm{\Gamma}{A}{x}$;
+in the rules introducing the definitional equalities we assume $\jtermeq{\Gamma}{A}{x}{x'}$.
 
@@
 & \inference{\jtypeeq{\ctxext{{\Gamma}{A}}{P}}{Q}{Q'}}{\jtypeeq{\ctxext{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}{\subst{x'}{Q'}}}\\
-\inference{\jtermt{\ctxext{\Gamma}{A}}{P}{f}}{\jtermt{\Gamma}{\subst{x}{P}}{\subst{x}{f}}}
-& \inference{\jtermteq{\ctxext{\Gamma}{A}}{P}{f}{f'}}{\jtermteq{\Gamma}{\subst{x}{P}}{\subst{x}{f}}{\subst{x'}{f'}}}\\
-\inference{\jtermt{\ctxext{{\Gamma}{A}}{P}}{Q}{g}}{\jtermt{\ctxext{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}{\subst{x}{g}}}
-& \inference{\jtermteq{\ctxext{{\Gamma}{A}}{P}}{Q}{g}{g'}}{\jtermteq{\ctxext{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}{\subst{x}{g}}{\subst{x'}{g'}}}
+\inference{\jterm{\ctxext{\Gamma}{A}}{P}{f}}{\jterm{\Gamma}{\subst{x}{P}}{\subst{x}{f}}}
+& \inference{\jtermeq{\ctxext{\Gamma}{A}}{P}{f}{f'}}{\jtermeq{\Gamma}{\subst{x}{P}}{\subst{x}{f}}{\subst{x'}{f'}}}\\
+\inference{\jterm{\ctxext{{\Gamma}{A}}{P}}{Q}{g}}{\jterm{\ctxext{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}{\subst{x}{g}}}
+& \inference{\jtermeq{\ctxext{{\Gamma}{A}}{P}}{Q}{g}{g'}}{\jtermeq{\ctxext{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}{\subst{x}{g}}{\subst{x'}{g'}}}
 \end{infarray}
@@
 In the following inference rules we assume
-$\jtermt{\Gamma}{A}{x}$ and $\jtermt{\ctxext{\Gamma}{A}}{P}{f}$.
+$\jterm{\Gamma}{A}{x}$ and $\jterm{\ctxext{\Gamma}{A}}{P}{f}$.
 
@@
 {\jtypeeq{\ctxext{\Gamma}{\subst{x}{{f}{Q}}}}{\subst{x}{{f}{R}}}{\subst{{x}{f}}{{x}{R}}}}\\
-\inference{\jtermt{\ctxext{{\Gamma}{A}}{P}}{Q}{g}}
-{\jtermteq{\Gamma}{\subst{x}{{f}{Q}}}{\subst{x}{{f}{g}}}{\subst{{x}{f}}{{x}{g}}}}\\
-\inference{\jtermt{\ctxext{{{\Gamma}{A}}{P}}{Q}}{R}{h}}
-{\jtermteq{\ctxext{\Gamma}{\subst{x}{{f}{Q}}}}{\subst{x}{{f}{R}}}{\subst{x}{{f}{h}}}{\subst{{x}{f}}{{x}{h}}}}
+\inference{\jterm{\ctxext{{\Gamma}{A}}{P}}{Q}{g}}
+{\jtermeq{\Gamma}{\subst{x}{{f}{Q}}}{\subst{x}{{f}{g}}}{\subst{{x}{f}}{{x}{g}}}}\\
+\inference{\jterm{\ctxext{{{\Gamma}{A}}{P}}{Q}}{R}{h}}
+{\jtermeq{\ctxext{\Gamma}{\subst{x}{{f}{Q}}}}{\subst{x}{{f}{R}}}{\subst{x}{{f}{h}}}{\subst{{x}{f}}{{x}{h}}}}
 \end{infarray}
@@
 \subsubsection{Substitution is compatible with extension}
-Suppose $\jtermt{\Gamma}{A}{x}$ in all of the following inference rules.
+Suppose $\jterm{\Gamma}{A}{x}$ in all of the following inference rules.
 \begin{infarray}{c}
@@
 \begin{infarray}{c}
-\inference{\jtype{\Gamma}{A}\qquad\jtype{\Gamma}{B}\qquad\jtermt{\Gamma}{A}{x}}{\jtypeeq{\Gamma}{\subst{x}{\ctxwk{A}{B}}}{B}}\\
-\inference{\jtype{\Gamma}{A}\qquad\jtype{\ctxext{\Gamma}{B}}{Q}\qquad\jtermt{\Gamma}{A}{x}}{\jtypeeq{\ctxext{\Gamma}{B}}{\subst{x}{\ctxwk{A}{Q}}}{Q}}\\
-\inference{\jtermt{\Gamma}{A}{x}\qquad\jtermt{\Gamma}{B}{y}}{\jtermteq{\Gamma}{B}{\subst{x}{\ctxwk{A}{y}}}{y}}\\
-\inference{\jtermt{\Gamma}{A}{x}\qquad\jtermt{\ctxext{\Gamma}{B}}{Q}{g}}{\jtermteq{\ctxext{\Gamma}{B}}{Q}{\subst{x}{\ctxwk{A}{g}}}{g}}
+\inference{\jtype{\Gamma}{A}\qquad\jtype{\Gamma}{B}\qquad\jterm{\Gamma}{A}{x}}{\jtypeeq{\Gamma}{\subst{x}{\ctxwk{A}{B}}}{B}}\\
+\inference{\jtype{\Gamma}{A}\qquad\jtype{\ctxext{\Gamma}{B}}{Q}\qquad\jterm{\Gamma}{A}{x}}{\jtypeeq{\ctxext{\Gamma}{B}}{\subst{x}{\ctxwk{A}{Q}}}{Q}}\\
+\inference{\jterm{\Gamma}{A}{x}\qquad\jterm{\Gamma}{B}{y}}{\jtermeq{\Gamma}{B}{\subst{x}{\ctxwk{A}{y}}}{y}}\\
+\inference{\jterm{\Gamma}{A}{x}\qquad\jterm{\ctxext{\Gamma}{B}}{Q}{g}}{\jtermeq{\ctxext{\Gamma}{B}}{Q}{\subst{x}{\ctxwk{A}{g}}}{g}}
 \end{infarray}
@@
 \inference{}
-{\jtermt{\ctxext{\Gamma}{A}}{\ctxwk{A}{{\Gamma}{\ctxext{\Gamma}{A}}}}{\pair_A}}\\
+{\jterm{\ctxext{\Gamma}{A}}{\ctxwk{A}{{\Gamma}{\ctxext{\Gamma}{A}}}}{\pair_A}}\\
 \inference{
@@
   \qquad
-  \jtermt{\ctxext{\Gamma}{A}}{\subst{\pair_A}{\ctxwk{A}{{\Gamma}{P}}}}{f}}
-  {\jtermt{\ctxext{\Gamma}{A}}{P}{\ind{\tfext_\Gamma(A)}(f)}}\\
+  \jterm{\ctxext{\Gamma}{A}}{\subst{\pair_A}{\ctxwk{A}{{\Gamma}{P}}}}{f}}
+  {\jterm{\ctxext{\Gamma}{A}}{P}{\ind{\tfext_\Gamma(A)}(f)}}\\
 \inference{
@@
   \qquad
-  \jtermt{\ctxext{\Gamma}{A}}{\subst{\pair_A}{\ctxwk{A}{{\Gamma}{P}}}}{f}}
-  {\jtermteq{\ctxext{\Gamma}{A}}{\subst{\pair_A}{\ctxwk{A}{{\Gamma}{P}}}}{\subst{\pair_A}{\ctxwk{A}{{\Gamma}{\ind{\tfext_\Gamma(A)}(f)}}}}{f}}
+  \jterm{\ctxext{\Gamma}{A}}{\subst{\pair_A}{\ctxwk{A}{{\Gamma}{P}}}}{f}}
+  {\jtermeq{\ctxext{\Gamma}{A}}{\subst{\pair_A}{\ctxwk{A}{{\Gamma}{P}}}}{\subst{\pair_A}{\ctxwk{A}{{\Gamma}{\ind{\tfext_\Gamma(A)}(f)}}}}{f}}
 \end{infarray}
@@
 \begin{infarray}{c}
-\inference{}{\jtermteq{\ctxext{\Gamma}{A}}{\ctxwk{A}{{\Gamma}{\ctxext{\Gamma}{A}}}}{\pair_A}{\pair_{A'}}}\\
-\inference{\jtype{\ctxext{\Gamma}{A}}{P}\qquad\jtermt{\ctxext{\Gamma}{A}}{\subst{\pair_A}{\ctxwk{A}{{\Gamma}{P}}}}{f}}
-{\jtermteq{\ctxext{\Gamma}{A}}{P}{\ind{\tfext_\Gamma(A)}(f)}{\ind{\tfext_\Gamma(A')}(f)}}
+\inference{}{\jtermeq{\ctxext{\Gamma}{A}}{\ctxwk{A}{{\Gamma}{\ctxext{\Gamma}{A}}}}{\pair_A}{\pair_{A'}}}\\
+\inference{\jtype{\ctxext{\Gamma}{A}}{P}\qquad\jterm{\ctxext{\Gamma}{A}}{\subst{\pair_A}{\ctxwk{A}{{\Gamma}{P}}}}{f}}
+{\jtermeq{\ctxext{\Gamma}{A}}{P}{\ind{\tfext_\Gamma(A)}(f)}{\ind{\tfext_\Gamma(A')}(f)}}
 \end{infarray}
@@
 \inference{}
-{\jtermt{\ctxext{{\Gamma}{A}}{P}}{\ctxwk{P}{{A}{\ctxext{A}{P}}}}{\pair_P}}\\
+{\jterm{\ctxext{{\Gamma}{A}}{P}}{\ctxwk{P}{{A}{\ctxext{A}{P}}}}{\pair_P}}\\
 \inference{
@@
   \qquad
-  \jtermt{\ctxext{{\Gamma}{A}}{P}}{\subst{\pair_P}{\ctxwk{P}{{A}{Q}}}}{f}}
-  {\jtermt{\ctxext{\Gamma}{{A}{P}}}{Q}{\ind{\tfext_A(P)}(f)}}\\
+  \jterm{\ctxext{{\Gamma}{A}}{P}}{\subst{\pair_P}{\ctxwk{P}{{A}{Q}}}}{f}}
+  {\jterm{\ctxext{\Gamma}{{A}{P}}}{Q}{\ind{\tfext_A(P)}(f)}}\\
 \inference{
@@
   \qquad
-  \jtermt{\ctxext{{\Gamma}{A}}{P}}{\subst{\pair_P}{\ctxwk{P}{{A}{Q}}}}{f}}
-  {\jtermteq{\ctxext{{\Gamma}{A}}{P}}{\subst{\pair_P}{\ctxwk{P}{{A}{Q}}}}{\subst{\pair_P}{\ctxwk{P}{{A}{\ind{\tfext_A(P)}(f)}}}}{f}}
+  \jterm{\ctxext{{\Gamma}{A}}{P}}{\subst{\pair_P}{\ctxwk{P}{{A}{Q}}}}{f}}
+  {\jtermeq{\ctxext{{\Gamma}{A}}{P}}{\subst{\pair_P}{\ctxwk{P}{{A}{Q}}}}{\subst{\pair_P}{\ctxwk{P}{{A}{\ind{\tfext_A(P)}(f)}}}}{f}}
 \end{infarray}
@@
 \begin{infarray}{c}
-\inference{\jctx{\Gamma}}{\jtermt{\Gamma}{\ctxwk{\Gamma}{\Gamma}}{\idfunc[\Gamma]}}\\
-\inference{\jtermt{\Gamma}{A}{x}}{\jtermteq{\Gamma}{A}{\subst{\idfunc[\Gamma]}{\ctxwk{\Gamma}{x}}}{x}}\\
-\inference{\jtermt{\Delta}{\ctxwk{\Delta}{\Gamma}}{g}}{\jtermteq{\Delta}{\ctxwk{\Delta}{\Gamma}}{\subst{g}{\ctxwk{\Delta}{\idfunc[\Gamma]}}}{g}}
+\inference{\jctx{\Gamma}}{\jterm{\Gamma}{\ctxwk{\Gamma}{\Gamma}}{\idfunc[\Gamma]}}\\
+\inference{\jterm{\Gamma}{A}{x}}{\jtermeq{\Gamma}{A}{\subst{\idfunc[\Gamma]}{\ctxwk{\Gamma}{x}}}{x}}\\
+\inference{\jterm{\Delta}{\ctxwk{\Delta}{\Gamma}}{g}}{\jtermeq{\Delta}{\ctxwk{\Delta}{\Gamma}}{\subst{g}{\ctxwk{\Delta}{\idfunc[\Gamma]}}}{g}}
 \end{infarray}
@@
 \begin{infarray}{c}
-\inference{\jtype{\Gamma}{A}}{\jtermt{\ctxext{\Gamma}{A}}{\ctxwk{A}{A}}{\idfunc[A]}}\\
-\inference{\jtermt{\Gamma}{A}{x}}{\jtermteq{\Gamma}{A}{\subst{x}{\idfunc[A]}}{x}}\\
-\inference{\jtermt{\ctxext{\Gamma}{A}}{P}{f}}{\jtermteq{\ctxext{\Gamma}{A}}{P}{\subst{\idfunc[A]}{\ctxwk{A}{f}}}{f}}\\
-\inference{\jtermt{\ctxext{\Gamma}{B}}{\ctxwk{B}{A}}{g}}{\jtermteq{\ctxext{\Gamma}{B}}{\ctxwk{B}{A}}{\subst{g}{\ctxwk{B}{\idfunc[A]}}}{g}}
+\inference{\jtype{\Gamma}{A}}{\jterm{\ctxext{\Gamma}{A}}{\ctxwk{A}{A}}{\idfunc[A]}}\\
+\inference{\jterm{\Gamma}{A}{x}}{\jtermeq{\Gamma}{A}{\subst{x}{\idfunc[A]}}{x}}\\
+\inference{\jterm{\ctxext{\Gamma}{A}}{P}{f}}{\jtermeq{\ctxext{\Gamma}{A}}{P}{\subst{\idfunc[A]}{\ctxwk{A}{f}}}{f}}\\
+\inference{\jterm{\ctxext{\Gamma}{B}}{\ctxwk{B}{A}}{g}}{\jtermeq{\ctxext{\Gamma}{B}}{\ctxwk{B}{A}}{\subst{g}{\ctxwk{B}{\idfunc[A]}}}{g}}
 \end{infarray}
@@
 \inference{\jctx{\Gamma}}{\jtype{\ctxext{\Gamma}{\ctxwk{\Gamma}{\Gamma}}}{\idtypevar{\Gamma}}}\\
-\inference{\jctx{\Gamma}}{\jtermt{\Gamma}{\subst{\idfunc[\Gamma]}{\idtypevar{\Gamma}}}{\refl{\Gamma}}}\\
+\inference{\jctx{\Gamma}}{\jterm{\Gamma}{\subst{\idfunc[\Gamma]}{\idtypevar{\Gamma}}}{\refl{\Gamma}}}\\
 \inference{\jtype{\ctxext{{\Gamma}{\ctxwk{\Gamma}{\Gamma}}}{\idtypevar{\Gamma}}}{P}
            \qquad
-           \jtermt{\Gamma}{\subst{\refl{\Gamma}}{{\idfunc[\Gamma]}{P}}}{d}}
-           {\jtermt{\ctxext{{\Gamma}{\ctxwk{\Gamma}{\Gamma}}}{\idtypevar{\Gamma}}}{P}{\tfJ(d)}}\\
+           \jterm{\Gamma}{\subst{\refl{\Gamma}}{{\idfunc[\Gamma]}{P}}}{d}}
+           {\jterm{\ctxext{{\Gamma}{\ctxwk{\Gamma}{\Gamma}}}{\idtypevar{\Gamma}}}{P}{\tfJ(d)}}\\
 \inference{\jtype{\ctxext{{\Gamma}{\ctxwk{\Gamma}{\Gamma}}}{\idtypevar{\Gamma}}}{P}\qquad
-\jtermt{\Gamma}{\subst{\refl{\Gamma}}{{\idfunc[\Gamma]}{P}}}{d}}
-{\jtermteq
+\jterm{\Gamma}{\subst{\refl{\Gamma}}{{\idfunc[\Gamma]}{P}}}{d}}
+{\jtermeq
   {\ctxext{{\Gamma}{\ctxwk{\Gamma}{\Gamma}}}{\idtypevar{\Gamma}}}
@@
 \inference{\jtype{\Gamma}{A}}{\jtype{\ctxext{{\Gamma}{A}}{\ctxwk{A}{A}}}{\idtypevar{A}}}\\
-\inference{\jtype{\Gamma}{A}}{\jtermt{\ctxext{\Gamma}{A}}{\subst{\idfunc[A]}{\idtypevar{A}}}{\refl{A}}}\\
+\inference{\jtype{\Gamma}{A}}{\jterm{\ctxext{\Gamma}{A}}{\subst{\idfunc[A]}{\idtypevar{A}}}{\refl{A}}}\\
 \inference{\jtype{\ctxext{{{\Gamma}{A}}{\ctxwk{A}{A}}}{\idtypevar{A}}}{P}\qquad
-\jtermt{\ctxext{\Gamma}{A}}{\subst{\refl{A}}{{\idfunc[A]}{P}}}{d}}
-{\jtermt{\ctxext{{{\Gamma}{A}}{\ctxwk{A}{A}}}{\idtypevar{A}}}{P}{\tfJ(d)}}\\
+\jterm{\ctxext{\Gamma}{A}}{\subst{\refl{A}}{{\idfunc[A]}{P}}}{d}}
+{\jterm{\ctxext{{{\Gamma}{A}}{\ctxwk{A}{A}}}{\idtypevar{A}}}{P}{\tfJ(d)}}\\
 \inference{\jtype{\ctxext{{{\Gamma}{A}}{\ctxwk{A}{A}}}{\idtypevar{A}}}{P}\qquad
-\jtermt{\ctxext{\Gamma}{A}}{\subst{\refl{A}}{{\idfunc[A]}{P}}}{d}}
-{\jtermteq
+\jterm{\ctxext{\Gamma}{A}}{\subst{\refl{A}}{{\idfunc[A]}{P}}}{d}}
+{\jtermeq
   {\ctxext{{{\Gamma}{A}}{\ctxwk{A}{A}}}{\idtypevar{A}}}
@@
 & \inference{\jtypeeq{\ctxext{\Gamma}{A}}{P}{P'}}{\jtypeeq{\Gamma}{\mprd{A}{P}}{\mprd{A}{P'}}}\\
-\inference{\jtermt{\ctxext{\Gamma}{A}}{P}{f}}{\jtermt{\Gamma}{\mprd{A}{P}}{\lambda(f)}}
-& \inference{\jtermteq{\ctxext{\Gamma}{A}}{P}{f}{f'}}{\jtermteq{\Gamma}{\mprd{A}{P}}{\lambda(f)}{\lambda(f')}}\\
-\inference{\jtermt{\Gamma}{\mprd{A}{P}}{g}}{\jtermt{\ctxext{\Gamma}{A}}{P}{\tfev(g)}}
-& \inference{\jtermteq{\Gamma}{\mprd{A}{P}}{g}{g'}}{\jtermteq{\ctxext{\Gamma}{A}}{P}{\tfev(g)}{\tfev(g')}}\\
-\inference{\jtermt{\ctxext{\Gamma}{A}}{P}{f}}{\jtermteq{\Gamma}{A}{\tfev(\lambda(f))}{f}}
+\inference{\jterm{\ctxext{\Gamma}{A}}{P}{f}}{\jterm{\Gamma}{\mprd{A}{P}}{\lambda(f)}}
+& \inference{\jtermeq{\ctxext{\Gamma}{A}}{P}{f}{f'}}{\jtermeq{\Gamma}{\mprd{A}{P}}{\lambda(f)}{\lambda(f')}}\\
+\inference{\jterm{\Gamma}{\mprd{A}{P}}{g}}{\jterm{\ctxext{\Gamma}{A}}{P}{\tfev(g)}}
+& \inference{\jtermeq{\Gamma}{\mprd{A}{P}}{g}{g'}}{\jtermeq{\ctxext{\Gamma}{A}}{P}{\tfev(g)}{\tfev(g')}}\\
+\inference{\jterm{\ctxext{\Gamma}{A}}{P}{f}}{\jtermeq{\Gamma}{A}{\tfev(\lambda(f))}{f}}
 \end{infarray}

### 2013-11-01 article-models-main.tex
moved content to the new file
funcy identity types
compatibility rules for weakening
developing category theory inside type theory; other constructors

@@
 \input{preamble-articles}
+
+\newcommand{\eqtype}[1]{\typefont{Eq}_{#1}}
+\newcommand{\eqtoid}[1]{\typefont{eqtoid}(#1)}
+\newcommand{\greek}{\mathrm}
+\newcommand{\product}[2]{{#1}\times{#2}}
+\newcommand{\pairp}[1]{(#1)}
+\newcommand{\ctxwkop}[2]{\ctxwk{#2}{#1}}
+\makeatletter
+\newcommand{\jhom}[3]{\jterm{\jhom@dom@ctx #1}{\jhom@dom@wk{#1}{#2}}{#3}}
+\newcommand{\jhom@dom@ctx}{\@ifnextchar\bgroup{\@jhom@dom@ctx}{}}
+\newcommand{\@jhom@dom@ctx}[1]{\@ifnextchar\bgroup{\@@jhom@dom@ctx{#1}}{#1}}
+\newcommand{\@@jhom@dom@ctx}[2]{\ctxext{\jhom@dom@ctx #1}{\jhom@dom@wk{#1}{#2}}}
+\newcommand{\jhom@dom@wk}[2]{{\@ifnextchar\bgroup{\@jhom@dom@wk{#2}}{\ctxwkop{#2}\bgroup\aftergroup{\aftergroup{\egroup}}}#1}}
+\newcommand{\@jhom@dom@wk}[2]{\@ifnextchar\bgroup{\@@jhom@dom@wk{#2}{#1}}{\ctxwk{#1}{#2}}}
+\newcommand{\@@jhom@dom@wk}[3]{\jhom@dom@wk{#2}{\ctxwk{\jhom@dom@ctx #3}{#1}}}
+\makeatother
+\newcommand{\jhomeq}[4]{\jtermeq{#1}{\ctxwk{#1}{#2}}{#3}{#4}}
+\newcommand{\jcomp}[3]{\subst{#2}{\ctxwk{#1}{#3}}}
+\newcommand{\jequalizer}[3]{\{#1|#2\jdeq #3\}}
+\newcommand{\jequalizerin}[2]{\iota_{#1,#2}}
+\newcommand{\tounit}[1]{{!_{#1}}}
+\newcommand{\trwk}{\typefont{trwk}}
+\newcommand{\trext}{\typefont{trext}}
 
@@
 
+\input{article-models-tt}
+
 \input{article-models-defn}

### 2013-11-01 article-models-tt.tex
developing category theory inside type theory; other constructors

@@
-\section{A variable-free presentation of type theory}
+\section{Variable-free type theory -- the basic theory}
 We have seen in the example of $\UU$ that we will not need to refer to variables
@@
 
+\subsubsection{A game of trees}
+We prove a little meta theorem about the type system we have so far. It is not
+really of importance to the development of the theory.
+
+Let's say
+that a binary tree of contexts is a pair $\pair{T,f}$ consisting of a binary 
+tree $T$ together with a function $f$ assigning to each leaf a context. The set
+of all binary trees of contexts is denoted by $B$. Such
+binary trees may be defined inductively: $(\unit,f):B$ for any $f:\unit\to ctx$
+and given any $(T_1,f_1)$ and $(T_2,f_2)$ in $B$ we have $(T_1,f_1)*(T_2,f_2)$
+in $B$.
+
+We simultaneously define
+the following two functions:
+\begin{align*}
+\trext & : B\to ctx\\
+\trwk_0 & : \prd{X:B} ctx\to typ(\trext(X))\\
+\trwk_1 & : \prd{X:B}{\Gamma~ctx} typ(\Gamma)\to typ(\ctxext{\trext(X)}{\trwk_0(X,\Gamma)})
+\end{align*}
+Both these functions are defined by induction on binary trees. We set
+\begin{align*}
+\trext((\unit,f)) & \defeq f(\unit)\\
+\trext((T_1,f_1)*(T_2,f_2)) & \defeq \ctxext{\trext((T_1,f_1))}
+{\trwk_0((T_1,f_1),\trext((T_2,f_2)))}
+\end{align*}
+and for any context $\Gamma$
+\begin{align*}
+\trwk_0((\unit,f),\Gamma) & \defeq \ctxwk{f(\unit)}{\Gamma}\\
+\trwk_0((T_1,f_1)*(T_2,f_2),\Gamma) 
+  & \defeq \trwk_1((T_1,f_1),\trwk_0((T_2,f_2)),\Gamma))
+\end{align*}
+and for any type $A$ in context $\Gamma$
+\begin{align*}
+\trwk_1((\unit,f),A) & \defeq \ctxwk{f(\unit)}{A}\\
+\trwk_1((T_1,f_1)*(T_2,f_2),A) & \defeq
+  \trwk_1((T_1,f_1),\trwk_1((T_2,f_2)),A))
+\end{align*}
+
+\begin{lem}
+$\trwk_0((T,f),\Gamma)$ is a type in context $\trext((T,f))$ and
+$\trwk_1((T,f),A)$ is a type in context $\ctxext{\trext((T,f))}{\trwk_0((T,f),\Gamma)}$
+for any type $A$ in context $\Gamma$.
+\end{lem}
+
+\begin{proof}
+It is immediate that $\trwk_0((\unit,f),\Gamma)$ is a type in context $\trext((\unit,f))$
+and that $\trwk_1((\unit,f),A)$ is a type in context $\ctxext{\trext((\unit,f))}
+{\trwk_0((\unit,f),\Gamma)}$ for any type $A$ in context $\Gamma$.
+
+Suppose that $(T_1,f_1)$ and $(T_2,f_2)$ are binary trees of contexts such that
+$\trwk_0((T_i,f_i),\Gamma)$ is a type in context
+$\trext((T_i,f_i))$ for $i\jdeq 1$ and $i\jdeq 2$, and such that
+$\trwk_1((T_i,f_1),A)$ is a type in context $\ctxext{\trext((T_i,f_i))}{\trwk_0((T_i,f_i),\Gamma)}$
+for $i\jdeq 1$ and $i\jdeq 2$. Then
+\begin{equation*}
+\trwk_0((T_1,f_1)*(T_2,f_2),\Gamma) 
+\jdeq \trwk_1((T_1,f_1),\trwk_0((T_2,f_2),\Gamma))
+\end{equation*}
+is a type in context
+\begin{equation*}
+\ctxext{\trext((T_1,f_1))}{\trwk_0((T_1,f_1),\trext((T_2,f_2)))}
+\jdeq \trext((T_1,f_1)*(T_2,f_2)).
+\end{equation*}
+Also,
+\begin{equation*}
+\trwk_1((T_1,f_1)*(T_2,f_2),A) \jdeq \trwk_1((T_1,f_1),\trwk_1((T_2,f_2),A))
+\end{equation*}
+is a type in context
+\begin{equation*}
+\ctxext{\trext((T_1,f_1))}{\trwk_0((T_1,f_1),\ctxext{\trext((T_2,f_2))}{\trwk_0((T_2,f_2),\Gamma))}}
+\end{equation*}
+\end{proof}
+
 \subsection{Substitution}
@@
 \end{infarray}
+
+Using the rules of the compatibility of substitution with weakening and of the
+compatibility of weakening with itself, we see that we can show
+\begin{equation*}
+\jtypeeq{\Gamma}{\subst{f}{\ctxwk{\Gamma}{{\Delta}{\greek{E}}}}}{\ctxwk{\Gamma}{E}}
+\end{equation*}
+for any three contexts $\Gamma$, $\Delta$ and $\greek{E}$ and any $\jhom{\Gamma}{\Delta}{f}$.
+It follows that for $\jhom{\Delta}{\greek{E}}{g}$ we have
+\begin{equation*}
+\jhom{\Gamma}{\greek{E}}{\subst{f}{\ctxwk{\Gamma}{g}}}
+\end{equation*}
+The term $\jcomp{\Gamma}{f}{g}$ is the composition of $g$ with $f$.
 
@@
 
-\subsection{Other type constructors}
+\section{Variable-free type theory -- other type constructors}
 We will now describe the rules for the type constructors that we don't assume
@@
 
-\subsubsection{Dependent function types}
+\subsection{The unit type}
+Since we don't have a notion of terms of a context, we just say that the context
+$\unit$ is the terminal context.
+
+\begin{infarray}{c}
+\inference{}{\jctx{\unit}}\\
+\inference{\jctx{\Gamma}}{\jhom{\Gamma}{\unit}{\tounit{\Gamma}}}\\
+\inference{\jctx{\Gamma}\qquad\jhom{\Gamma}{\unit}{f}}{\jhomeq{\Gamma}{\unit}{f}{\tounit{\Gamma}}}
+\end{infarray}
+
+Note that we don't have to require that $\jhomeq{\Gamma}{\unit}{\tounit{\Gamma}}{\tounit{\Delta}}$
+whenever we have a judgmental equality $\jctxeq{\Gamma}{\Delta}$, since this already follows from the third rule.
+
+When the context $\unit$ is present, we may use the expression $\jtype{}{\Gamma}$
+as a shorthand for the judgment $\jtype{\unit}{\ctxwk{\unit}{\Gamma}}$. Likewise,
+we may use the expression $\jtermc{\Gamma}{i}$ as a shorthand
+for the judgment $\jhom{\unit}{\Gamma}{i}$. If we have a type $A$ in context
+$\Gamma$, we may use the expression $\jtype{}{\subst{i}{A}}$ to stand for
+the judgment $\jtype{\unit}{\subst{i}{\ctxwk{\unit}{A}}}$. We see that in every
+respect, contexts are types in the empty context.
+
+We have created a strictly terminal object $\unit$. This is not necessary when
+we're working in a context. We introduce the unit type $\unit_\Gamma$ in context
+$\Gamma$ in the familiar type theoretical way.
+
+\begin{infarray}{c}
+\inference{\jctx{\Gamma}}{\jtype{\Gamma}{\unit_\Gamma}}\\
+\inference{\jctx{\Gamma}}{\jterm{\Gamma}{\unit_\Gamma}{\ttt_\Gamma}}\\
+\inference{\jtype{\ctxext{\Gamma}{\unit_\Gamma}}{P}\qquad\jterm{\Gamma}{\subst{\ttt_\Gamma}{P}}{u}}
+          {\jterm{\ctxext{\Gamma}{\unit_\Gamma}}{P}{\ind{\unit_\Gamma}(u)}}\\
+\inference{\jtype{\ctxext{\Gamma}{\unit_\Gamma}}{P}\qquad\jterm{\Gamma}{\subst{\ttt_\Gamma}{P}}{u}}
+          {\jtermeq{\Gamma}{\subst{\ttt_\Gamma}{P}}{\subst{\ttt_\Gamma}{\ind{\unit_\Gamma}(u)}}{u}}
+\end{infarray}
+
+\subsection{Product types}
+
+\subsubsection{Products}
+\begin{infarray}{c}
+\inference{\jctx{\Gamma}\qquad\jctx{\Delta}}{\jctx{\product{\Gamma}{\Delta}}}\\
+\inference{\jctx{\Gamma}\qquad\jctx{\Delta}}{\jhom{{\Gamma}{\Delta}}{\product{\Gamma}{\Delta}}{\pair}}
+\end{infarray}
+
+\subsubsection{Strict products}
+As with the unit type, we may use the categorical description of the product
+for our type theoretical definition. If we do that, we get strict products.
+
+\begin{infarray}{cc}
+\inference{\jctx{\Gamma}\qquad\jctx{\Delta}}{\jctx{\product{\Gamma}{\Delta}}}
+ & \inference{\jhom{\greek{E}}{\Gamma}{f}\qquad\jhom{\greek{E}}{\Delta}{g}}
+          {\jhom{\greek{E}}{\product{\Gamma}{\Delta}}{\pairp{f,g}}}\\
+\inference{\jctx{\Gamma}\qquad\jctx{\Delta}}{\jhom{\product{\Gamma}{\Delta}}{\Gamma}{\proj1}}
+ & \inference{\jhom{\greek{E}}{\Gamma}{f}\qquad\jhom{\greek{E}}{\Delta}{g}}
+          {\jhomeq{\greek{E}}{\Gamma}{\jcomp{\greek{E}}{\pairp{f,g}}{\proj1}}{f}}\\
+\inference{\jctx{\Gamma}\qquad\jctx{\Delta}}{\jhom{\product{\Gamma}{\Delta}}{\Delta}{\proj2}} 
+ & \inference{\jhom{\greek{E}}{\Gamma}{f}\qquad\jhom{\greek{E}}{\Delta}{g}}
+          {\jhomeq{\greek{E}}{\Delta}{\jcomp{\greek{E}}{\pairp{f,g}}{\proj2}}{g}}
+\end{infarray}
+\begin{equation*}
+\inference{\jhom{\greek{E}}{\product{\Gamma}{\Delta}}{h}
+           \qquad
+           \jhomeq{\greek{E}}{\Gamma}{\jcomp{E}{h}{\proj1}}{f}
+           \qquad
+           \jhomeq{\greek{E}}{\Delta}{\jcomp{E}{h}{\proj2}}{g}}
+          {\jhomeq{\greek{E}}{\product{\Gamma}{\Delta}}{h}{\pairp{f,g}}}
+\end{equation*}
+
+\subsection{Equalizer types}
+Now that we have introduced a terminal object and products the categorical way,
+we may just continue and present (strict) equalizers and pullbacks too, just to
+see where we get. These notions are probably just not very useful in a univalent
+setting until we got a good computational interpretation.
+
+\begin{infarray}{c}
+\inference{\jhom{\Gamma}{\Delta}{f}\qquad\jhom{\Gamma}{\Delta}{g}}{\jctx{\jequalizer{\Gamma}{f}{g}}}\\
+\inference{\jhom{\Gamma}{\Delta}{f}\qquad\jhom{\Gamma}{\Delta}{g}}{\jhom{\jequalizer{\Gamma}{f}{g}}{\Gamma}{\jequalizerin{f}{g}}}\\
+\inference{\jhom{\greek{E}}{\Gamma}{h}\qquad\jhomeq{\greek{E}}{\Delta}{\jcomp{\greek{E}}{h}{f}}{\jcomp{\greek{E}}{h}{g}}}{\jhom{\greek{E}}{\jequalizer{\Gamma}{f}{g}}{\jequalizer{h}{f}{g}}}\\
+\inference{
+  {\begin{array}{l}
+    \jhom{\greek{E}}{\Gamma}{h}\\
+    \jhomeq{\greek{E}}{\Delta}{\jcomp{\greek{E}}{h}{f}}{\jcomp{\greek{E}}{h}{g}}
+  \end{array}}
+  \qquad
+  {\begin{array}{l}
+    \jhom{\greek{E}}{\jequalizer{\Gamma}{f}{g}}{k}\\
+    \jhomeq{\greek{E}}{\Gamma}{\jcomp{\greek{E}}{k}{\jequalizerin{f}{g}}}{h}
+  \end{array}}
+}
+  {\jhom{\greek{E}}{\jequalizer{\Gamma}{f}{g}}{\jequalizer{h}{f}{g}}}
+\end{infarray}
+
+\subsection{Pullback types}
+
+\subsection{Dependent function types}
 \begin{infarray}{cc}
@@
 So it might be better to state that $\lambda$ is a trivial cofibration.
+
+\subsection{The empty type}
+
+\subsection{Coproduct types}
+
+\subsection{The natural numbers}

### 2013-11-07 article-models-tt.tex
worked on identity types

@@
 \trext & : B\to ctx\\
-\trwk_0 & : \prd{X:B} ctx\to typ(\trext(X))\\
+\trwk_0 & : \prd{X,Y:B} typ(\trext(X))\\
 \trwk_1 & : \prd{X:B}{\Gamma~ctx} typ(\Gamma)\to typ(\ctxext{\trext(X)}{\trwk_0(X,\Gamma)})
@@
 \begin{align*}
-\trwk_0((\unit,f),\Gamma) & \defeq \ctxwk{f(\unit)}{\Gamma}\\
+\trwk_0((\unit,f),(\unit,g)) & \defeq \ctxwk{f(\unit)}{g(\unit)}\\
 \trwk_0((T_1,f_1)*(T_2,f_2),\Gamma) 
@@
 \subsubsection{Substitution is compatible with weakening}
+\begin{equation*}
+\inference{\jterm{\Gamma}{A}{x}\qquad\jtype{\ctxext{\Gamma}{A}}{P}\qquad\jtype{\ctxext{\Gamma}{A}}{Q}}
+          {\jtypeeq{\ctxext{\Gamma}{\subst{x}{P}}}{\subst{x}{\ctxwk{P}{Q}}}{\ctxwk{\subst{x}{P}}{\subst{x}{Q}}}}
+\end{equation*}
+
 The judgmental equalities we're about to describe assert that weakening followed
@@
 \inference{\jterm{\Gamma}{A}{x}}{\jtermeq{\Gamma}{A}{\subst{x}{\idfunc[A]}}{x}}\\
+\inference{\jtype{\ctxext{\Gamma}{A}}{P}}{\jtypeeq{\ctxext{\Gamma}{A}}{\subst{\idfunc[A]}{\ctxwk{A}{P}}}{P}}\\
 \inference{\jterm{\ctxext{\Gamma}{A}}{P}{f}}{\jtermeq{\ctxext{\Gamma}{A}}{P}{\subst{\idfunc[A]}{\ctxwk{A}{f}}}{f}}\\
@@
 
+A notable difference in the formulation of identity types is that in our current
+setting we must state the elimination rule not only for families $P$ in context
+$\ctxext{{\Gamma}{\ctxwk{\Gamma}{\Gamma}}}{\idtypevar{\Gamma}}$,
+but also for families $Q$ in context 
+$\ctxext{{{\Gamma}{\ctxwk{\Gamma}{\Gamma}}}{\idtypevar{\Gamma}}}{P}$. The reason
+is that all operations have to be closed under slicing: everything may happen
+in a context. Secretly, a reason is that we don't have dependent function types.
+We wouldn't even be able to find the transport maps if we didn't state the
+identity elimination in an extended context.
+
 \begin{infarray}{c}
@@
   {\subst{\refl{\Gamma}}{{\idfunc[\Gamma]}{\tfJ(d)}}}
+  {d}}\\
+\inference{\jtype{\ctxext{{{\Gamma}{\ctxwk{\Gamma}{\Gamma}}}{\idtypevar{\Gamma}}}{P}}{Q}
+           \qquad
+           \jterm{\ctxext{\Gamma}{\subst{\refl{\Gamma}}{{\idfunc[\Gamma]}{P}}}}{\subst{\refl{\Gamma}}{{\idfunc[\Gamma]}{Q}}}{d}}
+           {\jterm{\ctxext{{{\Gamma}{\ctxwk{\Gamma}{\Gamma}}}{\idtypevar{\Gamma}}}{P}}{Q}{\tfJ(d)}}\\
+\inference
+  {\jtype
+    {\ctxext{{{\Gamma}{\ctxwk{\Gamma}{\Gamma}}}{\idtypevar{\Gamma}}}{P}}
+    {Q}
+  \qquad
+  \jterm
+    {\ctxext{\Gamma}{\subst{\refl{\Gamma}}{{\idfunc[\Gamma]}{P}}}}
+    {\subst{\refl{\Gamma}}{{\idfunc[\Gamma]}{Q}}}
+    {d}}
+  {\jtermeq
+    {\ctxext{{{\Gamma}{\ctxwk{\Gamma}{\Gamma}}}{\idtypevar{\Gamma}}}{P}}
+    {\subst{\refl{\Gamma}}{{\idfunc[\Gamma]}{Q}}}
+    {\subst{\refl{\Gamma}}{{\idfunc[\Gamma]}{\tfJ(d)}}}
     {d}}
 \end{infarray}
+
+Likewise, we introduce identity types in a context.
 
@@
 \end{infarray}
+
+Suppose we have terms $\jterm{\Gamma}{A}{x}$ and $\jterm{\Gamma}{A}{y}$. Then
+we may define $\id[A]{x}{y}\defeq\subst{y}{{x}{\idtypevar{A}}}$. A term
+$\jterm{\Gamma}{\id[A]{x}{y}}{p}$ is called an identification of $x$ and $y$.
+
+\subsubsection{Basic properties of identity types}
+In this subsubsection we prove some basic properties of identity types, just to
+know whether we got the type theory right.
+
+Suppose we have a family $\jtype{\ctxext{\Gamma}{A}}{P}$. Then we can consider
+the family $\jtype{\ctxext{{\Gamma}{A}}{\ctxwk{A}{A}}}{\ctxwk{A}{P}}$, which has the role of the family $\jtype{\Gamma,\,x,y:A}{P(y)}$
+of ordinary Martin-L\"of type theory. We may also consider the family
+$\jtype{\ctxext{{\Gamma}{A}}{\ctxwk{A}{A}}}{\ctxwk{{A}{A}}{P}}$; this one has the
+role of the family $\jtype{\Gamma,\,x,y:A}{P(x)}$. Those are families in the
+same context, so we have
+\begin{equation*}
+\jtype{\ctxext{{{\Gamma}{A}}{\ctxwk{A}{A}}}{\ctxwk{{A}{A}}{P}}}{\ctxwk{\ctxwk{{A}{A}}{P}}{{A}{P}}}
+\end{equation*}
+
+\begin{lem}
+There is a term
+\begin{equation*}
+\jterm{\ctxext{{{{\Gamma}{A}}{\ctxwk{A}{A}}}{\idtypevar{A}}}{\ctxwk{\idtypevar{A}}{{{A}{A}}{P}}}}{\ctxwk{\idtypevar{A}}{{{{A}{A}}{P}}{{A}{P}}}}{\transfibf{P}}
+\end{equation*}
+\end{lem}
+
+\begin{proof}
+By identity elimination it suffices to find a term
+\begin{equation*}
+\jterm{\ctxext{\Gamma}{A}}{\subst{\refl{A}}{{\idfunc[A]}{\ctxwk{\idtypevar{A}}{{{{A}{A}}{P}}{{A}{P}}}}}}{t}
+\end{equation*}
+By the judgmental equality $\jtypeeq{\Gamma}{\subst{{x}{f}}{{x}{Q}}}{\subst{x}{{f}{Q}}}$
+it follows that we have the judgmental equalities
+\begin{align*}
+& \subst{\refl{A}}{{\idfunc[A]}{\ctxwk{\idtypevar{A}}{{{{A}{A}}{P}}{{A}{P}}}}}\\
+& \qquad \jdeq \subst{\idfunc[A]}{{\ctxwk{{A}{A}}{\refl{A}}}{\ctxwk{\idtypevar{A}}{{{{A}{A}}{P}}{{A}{P}}}}}\\
+& \qquad \jdeq \subst{\idfunc[A]}{\ctxwk{{{A}{A}}{P}}{{A}{P}}}\\
+& \qquad \jdeq \ctxwk{\subst{\idfunc[A]}{\ctxwk{{A}{A}}{P}}}{\subst{\idfunc[A]}{\ctxwk{A}{P}}}\\
+& \qquad \jdeq \ctxwk{P}{P}
+\end{align*}
+in context $\ctxext{{\Gamma}{A}}{P}$. We have the term $\jterm{\ctxext{{\Gamma}{A}}{P}}{\ctxwk{P}{P}}{\idfunc[P]}$.
+\end{proof}
+
+Suppose we have a term $\jterm{\ctxext{\Gamma}{A}}{P}{f}$.
+
+Using identity types, we can assert that a function $\jhom{\Gamma}{\Delta}{f}$ has
+a left inverse $\jhom{\Delta}{\Gamma}{g}$ by asserting that there is an identification
+\begin{equation*}
+..
+\end{equation*}
 
@@
 \end{infarray}
+
+\subsection{Subterminal types}
+The subterminal types we're about to present are strict, so they're more like \verb+Prop+
+in \Coq. We can define subterminal types in two ways: the first equalizes all elements
+of the subject type; the second is universal with the property that for every constant
+map factors through it.
+
+\subsubsection{Equalizing all terms}
+
+\subsubsection{Factorizing constant maps}
+Let $\jhom{\Gamma}{\Delta}{f}$. We can weaken $f$ by $\Gamma$ to obtain a term
+$\jterm{\ctxext{\Gamma}{\ctxwk{\Gamma}{\Gamma}}}{\ctxwk{\Gamma}{{\Gamma}{\Delta}}}{\ctxwk{\Gamma}{f}}$.
+This term is `like the function $\lam{x}{y}f(y)$'. Likewise, we can weaken $f$
+by $\ctxwk{{\Gamma}{\Gamma}}$ to obtain a term
+$\jterm{\ctxext{\Gamma}{\ctxwk{\Gamma}{\Gamma}}}{\ctxwk{\Gamma}{{\Gamma}{\Delta}}}{\ctxwk{{\Gamma}{\Gamma}}{f}}$,
+which is `like the function $\lam{x}{y}f(x)$'. Since we have
+the judgmental equality 
+$\jtypeeq{\ctxext{\Gamma}{\ctxwk{\Gamma}{\Gamma}}}{\ctxwk{{\Gamma}{\Gamma}}{{\Gamma}{\Delta}}}{\ctxwk{\Gamma}{{\Gamma}{\Delta}}}$ we can consider the judgmental equality between $\ctxwk{\Gamma}{f}$
+and $\ctxwk{{\Gamma}{\Gamma}}{f}$, which is exactly what we'll do in the definition
+of judgmentally constant.
+
+\begin{defn}
+A term $\jhom{\Gamma}{\Delta}{f}$ is said to be \emph{judgmentally constant} if
+the judgment
+\begin{equation*}
+\jtermeq{\ctxext{\Gamma}{\ctxwk{\Gamma}{\Gamma}}}{\ctxwk{{\Gamma}{\Gamma}}{{\Gamma}{\Delta}}}{\ctxwk{\Gamma}{f}}{\ctxwk{{\Gamma}{\Gamma}}{f}}
+\end{equation*}
+can be derived.
+\end{defn}
+
+\begin{defn}
+\begin{infarray}{c}
+\inference{\jctx{\Gamma}}{\jctx{\tau\Gamma}}\\
+\inference{\jctx{\Gamma}}{\jhom{\Gamma}{\tau\Gamma}{t}}\\
+\inference{\jctx{\Gamma}}{\jtermeq{\ctxext{\Gamma}{\ctxwk{\Gamma}{\Gamma}}}{\ctxwk{{\Gamma}{\Gamma}}{{\Gamma}{\Delta}}}{\ctxwk{\Gamma}{t}}{\ctxwk{{\Gamma}{\Gamma}}{t}}}\\
+\inference{\jhom{\Gamma}{\Delta}{f}\qquad\jtermeq{\ctxext{\Gamma}{\ctxwk{\Gamma}{\Gamma}}}{\ctxwk{{\Gamma}{\Gamma}}{{\Gamma}{\Delta}}}{\ctxwk{\Gamma}{f}}{\ctxwk{{\Gamma}{\Gamma}}{f}}}{\jhom{\tau\Gamma}{\Delta}{\tilde{f}}}\\
+\inference{\jhom{\Gamma}{\Delta}{f}\qquad\jtermeq{\ctxext{\Gamma}{\ctxwk{\Gamma}{\Gamma}}}{\ctxwk{{\Gamma}{\Gamma}}{{\Gamma}{\Delta}}}{\ctxwk{\Gamma}{f}}{\ctxwk{{\Gamma}{\Gamma}}{f}}}
+{\jtermeq{\Gamma}{\ctxwk{\Gamma}{\Delta}}{\jcomp{\Gamma}{t}{\tilde{f}}}{f}}
+\end{infarray}
+\end{defn}
+
+\begin{lem}
+Any term $\jhom{\unit}{\Gamma}{i}$ is judgmentally constant.
+\end{lem}
+
+\begin{proof}
+
+\end{proof}
 
@@
 
+\subsection{Subterminal types}
+Let $\jhom{\Gamma}{\Delta}{f}$. Then we may also consider the type $\ctxwk{{\Gamma}{\Gamma}}{{\Gamma}{\Delta}}$
+in context $\ctxwk{\Gamma}{\Gamma}$ and we have the terms
+\begin{align*}
+& \jhom{\ctxext{\Gamma}{\ctxwk{\Gamma}{\Gamma}}}{\ctxwk{\Gamma}{A}}{\ctxwk{\Gamma}{f}}\\
+& \jhom{\ctxext{\Gamma}{\ctxwk{\Gamma}{\Gamma}}}{\ctxwk{\Gamma}{A}}{\ctxwk{{\Gamma}{\Gamma}}{f}}
+\end{align*}
+
 \subsection{The empty type}

### 2013-11-08 article-models-tt.tex
start leaving infarray environment; compatibility of typeconstructors with idtypes

@@
 
+Perhaps a bigger gain is that we obtain an algebraic theory for categories with
+fibrations and sections.
+
 \subsection{The basic judgments}
-The judgments we wish to make are not the standad six judgments for ``$\Gamma$ is a context'', ``$A$ is
+The judgments we wish to make are the standad six judgments for ``$\Gamma$ is a context'', ``$A$ is
 a type in context $\Gamma$'' and ``$x$ is a term of $A$ in context $\Gamma$'' and the three equality
-judgments coming with them. We have those, but we also add a judgment for ``$i$ is a term of context $\Gamma$''
-and ``$i$ and $j$ are equal terms of context $\Gamma$''. Thus, we have eight basic judgments we can make:
+judgments coming with them.
 
@@
 \jctx*{\Gamma} & \jctxeq*{\Gamma}{\Gamma'}\\
-\jtermc*{\Gamma}{i} & \jtermceq*{\Gamma}{i}{j}\\
 \jtype*{\Gamma}{A} & \jtypeeq*{\Gamma}{A}{B}\\
@@
 in all three cases (contexts, types and terms).
-%\begingroup
-%\renewcommand*{\arraystretch}{3}
-%\begin{equation*}
-%\begin{array}{ccc}
-\begin{infarray}{ccc}
-\inference{\jctx{\Gamma}}{\jctxeq{\Gamma}{\Gamma}} & \inference{\jctxeq{\Gamma}{\Delta}}{\jctxeq{\Delta}{\Gamma}} & \inference{\jctxeq{\Gamma}{\Delta}\qquad\jctxeq{\Delta}{\Theta}}{\jctxeq{\Gamma}{\Theta}}\\
-\inference{\jtype{\Gamma}{A}}{\jtypeeq{\Gamma}{A}{A}} &
-\inference{\jtypeeq{\Gamma}{A}{B}}{\jtypeeq{\Gamma}{B}{A}} & 
-\inference{\jtypeeq{\Gamma}{A}{B}\qquad\jtypeeq{\Gamma}{B}{C}}{\jtypeeq{\Gamma}{A}{C}}\\
-\inference{\jterm{\Gamma}{A}{x}}{\jtermeq{\Gamma}{A}{x}{x}} & 
-\inference{\jtermeq{\Gamma}{A}{x}{y}}{\jtermeq{\Gamma}{A}{y}{x}} &
-\inference{\jtermeq{\Gamma}{A}{x}{y}\qquad\jtermeq{\Gamma}{A}{y}{z}}{\jtermeq{\Gamma}{A}{x}{z}}
-\end{infarray}
-%\end{array}
-%\end{equation*}
-\begin{infarray}{cc}
-\inference{\jctxeq{\Gamma}{\Delta}\qquad\jtype{\Gamma}{A}}{\jtype{\Delta}{A}}
-& \inference{\jctxeq{\Gamma}{\Delta}\qquad\jtypeeq{\Gamma}{A}{B}}{\jtypeeq{\Delta}{A}{B}}\\
-\inference{\jctxeq{\Gamma}{\Delta}\qquad\jterm{\Gamma}{A}{x}}{\jterm{\Delta}{A}{x}}
-& \inference{\jctxeq{\Gamma}{\Delta}\qquad\jtermeq{\Gamma}{A}{x}{y}}{\jtermeq{\Delta}{A}{x}{y}}\\
-\inference{\jtypeeq{\Gamma}{A}{B}\qquad \jterm{\Gamma}{A}{x}}{\jterm{\Gamma}{B}{x}}
-& \inference{\jtypeeq{\Gamma}{A}{B}\qquad\jtermeq{\Gamma}{A}{x}{y}}{\jtermeq{\Gamma}{B}{x}{y}}
-\end{infarray}
+\begin{align}
+& \inference{\jctx{\Gamma}}{\jctxeq{\Gamma}{\Gamma}} 
+& & \inference{\jctxeq{\Gamma}{\Delta}}{\jctxeq{\Delta}{\Gamma}} 
+& & \inference{\jctxeq{\Gamma}{\Delta}\qquad\jctxeq{\Delta}{\Theta}}{\jctxeq{\Gamma}{\Theta}}\\
+& \inference{\jtype{\Gamma}{A}}{\jtypeeq{\Gamma}{A}{A}} 
+& & \inference{\jtypeeq{\Gamma}{A}{B}}{\jtypeeq{\Gamma}{B}{A}}
+& & \inference{\jtypeeq{\Gamma}{A}{B}\qquad\jtypeeq{\Gamma}{B}{C}}{\jtypeeq{\Gamma}{A}{C}}\\
+& \inference{\jterm{\Gamma}{A}{x}}{\jtermeq{\Gamma}{A}{x}{x}}
+& & \inference{\jtermeq{\Gamma}{A}{x}{y}}{\jtermeq{\Gamma}{A}{y}{x}}
+& & \inference{\jtermeq{\Gamma}{A}{x}{y}\qquad\jtermeq{\Gamma}{A}{y}{z}}{\jtermeq{\Gamma}{A}{x}{z}}
+\end{align}
+
+\begin{align}
+& \inference{\jctxeq{\Gamma}{\Delta}\qquad\jtype{\Gamma}{A}}{\jtype{\Delta}{A}}
+& & \inference{\jctxeq{\Gamma}{\Delta}\qquad\jtypeeq{\Gamma}{A}{B}}{\jtypeeq{\Delta}{A}{B}}\\
+& \inference{\jctxeq{\Gamma}{\Delta}\qquad\jterm{\Gamma}{A}{x}}{\jterm{\Delta}{A}{x}}
+& & \inference{\jctxeq{\Gamma}{\Delta}\qquad\jtermeq{\Gamma}{A}{x}{y}}{\jtermeq{\Delta}{A}{x}{y}}\\
+& \inference{\jtypeeq{\Gamma}{A}{B}\qquad \jterm{\Gamma}{A}{x}}{\jterm{\Gamma}{B}{x}}
+& & \inference{\jtypeeq{\Gamma}{A}{B}\qquad\jtermeq{\Gamma}{A}{x}{y}}{\jtermeq{\Gamma}{B}{x}{y}}
+\end{align}
 
@@
 \begin{infarray}{c}
-\inference{\jtype{\ctxext{{\Gamma}{A}}{P}}{Q}}{\jtypeeq{\Gamma}{\subst{x}{\ctxext{P}{Q}}}{\ctxext{\subst{x}{P}}{\subst{x}{Q}}}}
+\inference{\jtype{\ctxext{{\Gamma}{A}}{P}}{Q}}
+  {\jtypeeq{\Gamma}{\subst{x}{\ctxext{P}{Q}}}{\ctxext{\subst{x}{P}}{\subst{x}{Q}}}}\\
+\inference{\jtype{\ctxext{{{\Gamma}{A}}{P}}{Q}}{R}}
+  {\jtypeeq{\ctxext{\Gamma}{\subst{x}{P}}}{\subst{x}{\ctxext{Q}{R}}}{\ctxext{\subst{x}{Q}}{\subst{x}{R}}}}
 \end{infarray}
@@
 
-\subsection{The variable rule -- identity functions}
-The variable rule for contexts:
-\begin{infarray}{c}
-\inference{\jctx{\Gamma}}{\jterm{\Gamma}{\ctxwk{\Gamma}{\Gamma}}{\idfunc[\Gamma]}}\\
-\inference{\jterm{\Gamma}{A}{x}}{\jtermeq{\Gamma}{A}{\subst{\idfunc[\Gamma]}{\ctxwk{\Gamma}{x}}}{x}}\\
-\inference{\jterm{\Delta}{\ctxwk{\Delta}{\Gamma}}{g}}{\jtermeq{\Delta}{\ctxwk{\Delta}{\Gamma}}{\subst{g}{\ctxwk{\Delta}{\idfunc[\Gamma]}}}{g}}
-\end{infarray}
+\subsubsection{Extension is compatible with weakening}
 
-The variable rule for types:
-\begin{infarray}{c}
-\inference{\jtype{\Gamma}{A}}{\jterm{\ctxext{\Gamma}{A}}{\ctxwk{A}{A}}{\idfunc[A]}}\\
-\inference{\jterm{\Gamma}{A}{x}}{\jtermeq{\Gamma}{A}{\subst{x}{\idfunc[A]}}{x}}\\
-\inference{\jtype{\ctxext{\Gamma}{A}}{P}}{\jtypeeq{\ctxext{\Gamma}{A}}{\subst{\idfunc[A]}{\ctxwk{A}{P}}}{P}}\\
-\inference{\jterm{\ctxext{\Gamma}{A}}{P}{f}}{\jtermeq{\ctxext{\Gamma}{A}}{P}{\subst{\idfunc[A]}{\ctxwk{A}{f}}}{f}}\\
-\inference{\jterm{\ctxext{\Gamma}{B}}{\ctxwk{B}{A}}{g}}{\jtermeq{\ctxext{\Gamma}{B}}{\ctxwk{B}{A}}{\subst{g}{\ctxwk{B}{\idfunc[A]}}}{g}}
-\end{infarray}
+\begin{equation*}
+\jterm{\ctxext{\Gamma}{\ctxwk{\Gamma}{\ctxext{\Delta}{B}}}}{\ctxwk{{\Gamma}{\ctxext{\Delta}{B}}}{\ctxext{\ctxwk{\Gamma}{\Delta}}{\ctxwk{\Gamma}{B}}}}{ac}
+\end{equation*}
+
+\subsection{The identity functions}
+Without a rule explicitly asserting the existence of an identity morphism we don't
+get one, hence we do that here. The identity morphism is a term which introduced
+in ordinary type theory via the variable rule. The variable rule is a bit more
+general: it asserts that
+\begin{equation*}
+\jterm{\Gamma,\,x_1:A_1,\ldots,\,x_n:A_n}{A_i}{x_i}
+\end{equation*}
+for every $1\leq i\leq n$. Thus, it establishes the projections. In our setting,
+we get the projections from the identity morphisms together with weakening. We
+already have weakening, so here it suffices to introduce the identity morphisms.
+They are determined by their behavior with respect to substitution combined with
+weakening.
+
+\begin{align}
+&\inference{\jctx{\Gamma}}{\jterm{\Gamma}{\ctxwk{\Gamma}{\Gamma}}{\idfunc[\Gamma]}}\label{ctxIdFuncForm}\\
+&\inference{\jtype{\Gamma}{A}}{\jtypeeq{\Gamma}{\subst{\idfunc[\Gamma]}{\ctxwk{\Gamma}{A}}}{A}}\label{ctxIdFuncPbInvType}\\
+&\inference{\jterm{\Gamma}{A}{x}}{\jtermeq{\Gamma}{A}{\subst{\idfunc[\Gamma]}{\ctxwk{\Gamma}{x}}}{x}}\label{ctxIdFuncPbInvTerm}\\
+&\inference{\jterm{\Delta}{\ctxwk{\Delta}{\Gamma}}{g}}{\jtermeq{\Delta}{\ctxwk{\Delta}{\Gamma}}{\subst{g}{\ctxwk{\Delta}{\idfunc[\Gamma]}}}{g}}\label{ctxIdFuncSubstInvTerm}
+\end{align}
+
+The identity morphisms for types are introduced analogously to the identity
+morphisms of contexts, the only difference being that we have a rule asserting
+that $\subst{x}{\idfunc[A]}\jdeq x$ for every term $x:A$ in context $\Gamma$.
+
+\begin{align}
+& \inference{\jtype{\Gamma}{A}}{\jterm{\ctxext{\Gamma}{A}}{\ctxwk{A}{A}}{\idfunc[A]}}\\
+& \inference{\jterm{\Gamma}{A}{x}}{\jtermeq{\Gamma}{A}{\subst{x}{\idfunc[A]}}{x}}\\
+& \inference{\jtype{\ctxext{\Gamma}{A}}{P}}{\jtypeeq{\ctxext{\Gamma}{A}}{\subst{\idfunc[A]}{\ctxwk{A}{P}}}{P}}\\
+& \inference{\jterm{\ctxext{\Gamma}{A}}{P}{f}}{\jtermeq{\ctxext{\Gamma}{A}}{P}{\subst{\idfunc[A]}{\ctxwk{A}{f}}}{f}} \\
+& \inference{\jterm{\ctxext{\Gamma}{B}}{\ctxwk{B}{A}}{g}}{\jtermeq{\ctxext{\Gamma}{B}}{\ctxwk{B}{A}}{\subst{g}{\ctxwk{B}{\idfunc[A]}}}{g}}
+\end{align}
 
@@
 
-\subsection{Identity types}
+\section{Identity types}
 The identity types may now be formulated just as in ordinary Martin\nobreakdash-L\"of
@@
 
-\begin{infarray}{c}
-\inference{\jctx{\Gamma}}{\jtype{\ctxext{\Gamma}{\ctxwk{\Gamma}{\Gamma}}}{\idtypevar{\Gamma}}}\\
-\inference{\jctx{\Gamma}}{\jterm{\Gamma}{\subst{\idfunc[\Gamma]}{\idtypevar{\Gamma}}}{\refl{\Gamma}}}\\
-\inference{\jtype{\ctxext{{\Gamma}{\ctxwk{\Gamma}{\Gamma}}}{\idtypevar{\Gamma}}}{P}
+\begin{align}
+& \inference{\jctx{\Gamma}}{\jtype{\ctxext{\Gamma}{\ctxwk{\Gamma}{\Gamma}}}{\idtypevar{\Gamma}}}\\
+& \inference{\jctx{\Gamma}}{\jterm{\Gamma}{\subst{\idfunc[\Gamma]}{\idtypevar{\Gamma}}}{\refl{\Gamma}}}\\
+& \inference{\jtype{\ctxext{{\Gamma}{\ctxwk{\Gamma}{\Gamma}}}{\idtypevar{\Gamma}}}{P}
            \qquad
@@
            {\jterm{\ctxext{{\Gamma}{\ctxwk{\Gamma}{\Gamma}}}{\idtypevar{\Gamma}}}{P}{\tfJ(d)}}\\
-\inference{\jtype{\ctxext{{\Gamma}{\ctxwk{\Gamma}{\Gamma}}}{\idtypevar{\Gamma}}}{P}\qquad
+& \inference{\jtype{\ctxext{{\Gamma}{\ctxwk{\Gamma}{\Gamma}}}{\idtypevar{\Gamma}}}{P}\qquad
 \jterm{\Gamma}{\subst{\refl{\Gamma}}{{\idfunc[\Gamma]}{P}}}{d}}
@@
   {d}}\\
-\inference{\jtype{\ctxext{{{\Gamma}{\ctxwk{\Gamma}{\Gamma}}}{\idtypevar{\Gamma}}}{P}}{Q}
+& \inference{\jtype{\ctxext{{{\Gamma}{\ctxwk{\Gamma}{\Gamma}}}{\idtypevar{\Gamma}}}{P}}{Q}
            \qquad
@@
            {\jterm{\ctxext{{{\Gamma}{\ctxwk{\Gamma}{\Gamma}}}{\idtypevar{\Gamma}}}{P}}{Q}{\tfJ(d)}}\\
-\inference
+& \inference
   {\jtype
@@
     {d}}
-\end{infarray}
+\end{align}
 
@@
 
-\begin{infarray}{c}
-\inference{\jtype{\Gamma}{A}}{\jtype{\ctxext{{\Gamma}{A}}{\ctxwk{A}{A}}}{\idtypevar{A}}}\\
-\inference{\jtype{\Gamma}{A}}{\jterm{\ctxext{\Gamma}{A}}{\subst{\idfunc[A]}{\idtypevar{A}}}{\refl{A}}}\\
-\inference{\jtype{\ctxext{{{\Gamma}{A}}{\ctxwk{A}{A}}}{\idtypevar{A}}}{P}\qquad
+\begin{align}
+& \inference{\jtype{\Gamma}{A}}{\jtype{\ctxext{{\Gamma}{A}}{\ctxwk{A}{A}}}{\idtypevar{A}}}\\
+& \inference{\jtype{\Gamma}{A}}{\jterm{\ctxext{\Gamma}{A}}{\subst{\idfunc[A]}{\idtypevar{A}}}{\refl{A}}}\\
+& \inference{\jtype{\ctxext{{{\Gamma}{A}}{\ctxwk{A}{A}}}{\idtypevar{A}}}{P}\qquad
 \jterm{\ctxext{\Gamma}{A}}{\subst{\refl{A}}{{\idfunc[A]}{P}}}{d}}
 {\jterm{\ctxext{{{\Gamma}{A}}{\ctxwk{A}{A}}}{\idtypevar{A}}}{P}{\tfJ(d)}}\\
-\inference{\jtype{\ctxext{{{\Gamma}{A}}{\ctxwk{A}{A}}}{\idtypevar{A}}}{P}\qquad
+& \inference{\jtype{\ctxext{{{\Gamma}{A}}{\ctxwk{A}{A}}}{\idtypevar{A}}}{P}\qquad
 \jterm{\ctxext{\Gamma}{A}}{\subst{\refl{A}}{{\idfunc[A]}{P}}}{d}}
@@
   {d}}
-\end{infarray}
+\end{align}
 
@@
 
+\subsubsection{Compatibility of identity types with extension}
+
+\subsubsection{Compatibility of identity types with weakening}
+Suppose $A$ and $B$ are types in context $\Gamma$. Then we can consider the types
+\begin{align*}
+\jtype*{\ctxext{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{{B}{B}}}}{\ctxwk{A}{\idtypevar{B}}}\\
+\jtype*{\ctxext{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{{A}{B}}{{A}{B}}}}{\idtypevar{\ctxwk{A}{B}}}
+\end{align*}
+Note that we have the judgmental equality
+\begin{equation*}
+\jtypeeq
+  {\ctxext{{\Gamma}{A}}{\ctxwk{A}{B}}}
+  {\ctxwk{A}{{B}{B}}}
+  {\ctxwk{{A}{B}}{{A}{B}}}
+\end{equation*}
+so $\ctxwk{A}{\idtypevar{B}}$ and $\idtypevar{\ctxwk{A}{B}}$ are types in the
+same context.
+
+\begin{lem}
+There is a term of type
+\begin{equation*}
+\jterm
+  {\ctxext{{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{{A}{B}}{{A}{B}}}}{\idtypevar{\ctxwk{A}{B}}}}
+  {\ctxwk{\idtypevar{\ctxwk{A}{B}}}{\ctxwk{A}{\idtypevar{B}}}}
+  {\typefont{idwktowkid}}
+\end{equation*}
+\end{lem}
+
+\begin{proof}
+By the identity elimination rule it suffices to find a term
+\begin{equation*}
+\jterm
+  {\ctxext{{\Gamma}{A}}{\ctxwk{A}{B}}}
+  {\subst{\refl{\ctxwk{A}{B}}}{{\idfunc[\ctxwk{A}{B}]}{\ctxwk{\idtypevar{\ctxwk{A}{B}}}{\ctxwk{A}{\idtypevar{B}}}}}}
+  {i}
+\end{equation*}
+We may simplify the type $\subst{\refl{\ctxwk{A}{B}}}{{\idfunc[\ctxwk{A}{B}]}{\ctxwk{\idtypevar{\ctxwk{A}{B}}}{\ctxwk{A}{\idtypevar{B}}}}}$ as follows:
+\begin{align*}
+& \subst{\refl{\ctxwk{A}{B}}}{{\idfunc[\ctxwk{A}{B}]}{\ctxwk{\idtypevar{\ctxwk{A}{B}}}{\ctxwk{A}{\idtypevar{B}}}}}\\
+& \qquad\jdeq \subst{\idfunc[\ctxwk{A}{B}]}{{\ctxwk{..}{\refl{\ctxwk{A}{B}}}}{\ctxwk{\idtypevar{\ctxwk{A}{B}}}{\ctxwk{A}{\idtypevar{B}}}}}
+\end{align*}
+\end{proof}
+
+\subsubsection{Compatibility of identity types with substitution}
+
 \section{Variable-free type theory -- other type constructors}
@@
 
-\begin{infarray}{c}
-\inference{}{\jctx{\unit}}\\
-\inference{\jctx{\Gamma}}{\jhom{\Gamma}{\unit}{\tounit{\Gamma}}}\\
-\inference{\jctx{\Gamma}\qquad\jhom{\Gamma}{\unit}{f}}{\jhomeq{\Gamma}{\unit}{f}{\tounit{\Gamma}}}
-\end{infarray}
+\begin{align}
+& \inference{}{\jctx{\unit}}\\
+& \inference{\jctx{\Gamma}}{\jhom{\Gamma}{\unit}{\tounit{\Gamma}}}\\
+& \inference{\jctx{\Gamma}\qquad\jhom{\Gamma}{\unit}{f}}{\jhomeq{\Gamma}{\unit}{f}{\tounit{\Gamma}}}
+\end{align}
 
@@
 
-\begin{infarray}{c}
-\inference{\jctx{\Gamma}}{\jtype{\Gamma}{\unit_\Gamma}}\\
-\inference{\jctx{\Gamma}}{\jterm{\Gamma}{\unit_\Gamma}{\ttt_\Gamma}}\\
-\inference{\jtype{\ctxext{\Gamma}{\unit_\Gamma}}{P}\qquad\jterm{\Gamma}{\subst{\ttt_\Gamma}{P}}{u}}
+\begin{align}
+& \inference{\jctx{\Gamma}}{\jtype{\Gamma}{\unit_\Gamma}}\\
+& \inference{\jctx{\Gamma}}{\jterm{\Gamma}{\unit_\Gamma}{\ttt_\Gamma}}\\
+& \inference{\jtype{\ctxext{\Gamma}{\unit_\Gamma}}{P}\qquad\jterm{\Gamma}{\subst{\ttt_\Gamma}{P}}{u}}
           {\jterm{\ctxext{\Gamma}{\unit_\Gamma}}{P}{\ind{\unit_\Gamma}(u)}}\\
-\inference{\jtype{\ctxext{\Gamma}{\unit_\Gamma}}{P}\qquad\jterm{\Gamma}{\subst{\ttt_\Gamma}{P}}{u}}
+& \inference{\jtype{\ctxext{\Gamma}{\unit_\Gamma}}{P}\qquad\jterm{\Gamma}{\subst{\ttt_\Gamma}{P}}{u}}
           {\jtermeq{\Gamma}{\subst{\ttt_\Gamma}{P}}{\subst{\ttt_\Gamma}{\ind{\unit_\Gamma}(u)}}{u}}
-\end{infarray}
+\end{align}
 
@@
 \begin{defn}
-\begin{infarray}{c}
-\inference{\jctx{\Gamma}}{\jctx{\tau\Gamma}}\\
-\inference{\jctx{\Gamma}}{\jhom{\Gamma}{\tau\Gamma}{t}}\\
-\inference{\jctx{\Gamma}}{\jtermeq{\ctxext{\Gamma}{\ctxwk{\Gamma}{\Gamma}}}{\ctxwk{{\Gamma}{\Gamma}}{{\Gamma}{\Delta}}}{\ctxwk{\Gamma}{t}}{\ctxwk{{\Gamma}{\Gamma}}{t}}}\\
-\inference{\jhom{\Gamma}{\Delta}{f}\qquad\jtermeq{\ctxext{\Gamma}{\ctxwk{\Gamma}{\Gamma}}}{\ctxwk{{\Gamma}{\Gamma}}{{\Gamma}{\Delta}}}{\ctxwk{\Gamma}{f}}{\ctxwk{{\Gamma}{\Gamma}}{f}}}{\jhom{\tau\Gamma}{\Delta}{\tilde{f}}}\\
-\inference{\jhom{\Gamma}{\Delta}{f}\qquad\jtermeq{\ctxext{\Gamma}{\ctxwk{\Gamma}{\Gamma}}}{\ctxwk{{\Gamma}{\Gamma}}{{\Gamma}{\Delta}}}{\ctxwk{\Gamma}{f}}{\ctxwk{{\Gamma}{\Gamma}}{f}}}
+\begin{align}
+& \inference{\jctx{\Gamma}}{\jctx{\tau\Gamma}}\\
+& \inference{\jctx{\Gamma}}{\jhom{\Gamma}{\tau\Gamma}{t}}\\
+& \inference{\jctx{\Gamma}}{\jtermeq{\ctxext{\Gamma}{\ctxwk{\Gamma}{\Gamma}}}{\ctxwk{{\Gamma}{\Gamma}}{{\Gamma}{\Delta}}}{\ctxwk{\Gamma}{t}}{\ctxwk{{\Gamma}{\Gamma}}{t}}}\\
+& \inference{\jhom{\Gamma}{\Delta}{f}\qquad\jtermeq{\ctxext{\Gamma}{\ctxwk{\Gamma}{\Gamma}}}{\ctxwk{{\Gamma}{\Gamma}}{{\Gamma}{\Delta}}}{\ctxwk{\Gamma}{f}}{\ctxwk{{\Gamma}{\Gamma}}{f}}}{\jhom{\tau\Gamma}{\Delta}{\tilde{f}}}\\
+& \inference{\jhom{\Gamma}{\Delta}{f}\qquad\jtermeq{\ctxext{\Gamma}{\ctxwk{\Gamma}{\Gamma}}}{\ctxwk{{\Gamma}{\Gamma}}{{\Gamma}{\Delta}}}{\ctxwk{\Gamma}{f}}{\ctxwk{{\Gamma}{\Gamma}}{f}}}
 {\jtermeq{\Gamma}{\ctxwk{\Gamma}{\Delta}}{\jcomp{\Gamma}{t}{\tilde{f}}}{f}}
-\end{infarray}
+\end{align}
 \end{defn}
@@
 \subsubsection{Products}
-\begin{infarray}{c}
-\inference{\jctx{\Gamma}\qquad\jctx{\Delta}}{\jctx{\product{\Gamma}{\Delta}}}\\
-\inference{\jctx{\Gamma}\qquad\jctx{\Delta}}{\jhom{{\Gamma}{\Delta}}{\product{\Gamma}{\Delta}}{\pair}}
-\end{infarray}
+\begin{align}
+& \inference{\jctx{\Gamma}\qquad\jctx{\Delta}}{\jctx{\product{\Gamma}{\Delta}}}\\
+& \inference{\jctx{\Gamma}\qquad\jctx{\Delta}}{\jhom{{\Gamma}{\Delta}}{\product{\Gamma}{\Delta}}{\pair}}
+\end{align}
 
@@
 
-\begin{infarray}{cc}
-\inference{\jctx{\Gamma}\qquad\jctx{\Delta}}{\jctx{\product{\Gamma}{\Delta}}}
+\begin{align}
+& \inference{\jctx{\Gamma}\qquad\jctx{\Delta}}{\jctx{\product{\Gamma}{\Delta}}}\\
+& \inference{\jctx{\Gamma}\qquad\jctx{\Delta}}{\jhom{\product{\Gamma}{\Delta}}{\Gamma}{\proj1}}\\
+& \inference{\jctx{\Gamma}\qquad\jctx{\Delta}}{\jhom{\product{\Gamma}{\Delta}}{\Delta}{\proj2}}\\
 & \inference{\jhom{\greek{E}}{\Gamma}{f}\qquad\jhom{\greek{E}}{\Delta}{g}}
           {\jhom{\greek{E}}{\product{\Gamma}{\Delta}}{\pairp{f,g}}}\\
-\inference{\jctx{\Gamma}\qquad\jctx{\Delta}}{\jhom{\product{\Gamma}{\Delta}}{\Gamma}{\proj1}}
 & \inference{\jhom{\greek{E}}{\Gamma}{f}\qquad\jhom{\greek{E}}{\Delta}{g}}
           {\jhomeq{\greek{E}}{\Gamma}{\jcomp{\greek{E}}{\pairp{f,g}}{\proj1}}{f}}\\
-\inference{\jctx{\Gamma}\qquad\jctx{\Delta}}{\jhom{\product{\Gamma}{\Delta}}{\Delta}{\proj2}} 
 & \inference{\jhom{\greek{E}}{\Gamma}{f}\qquad\jhom{\greek{E}}{\Delta}{g}}
           {\jhomeq{\greek{E}}{\Delta}{\jcomp{\greek{E}}{\pairp{f,g}}{\proj2}}{g}}
-\end{infarray}
-\begin{equation*}
+\end{align}
+
+\begin{equation}
 \inference{\jhom{\greek{E}}{\product{\Gamma}{\Delta}}{h}
@@
           {\jhomeq{\greek{E}}{\product{\Gamma}{\Delta}}{h}{\pairp{f,g}}}
-\end{equation*}
+\end{equation}
 
@@
 
-\begin{infarray}{c}
-\inference{\jhom{\Gamma}{\Delta}{f}\qquad\jhom{\Gamma}{\Delta}{g}}{\jctx{\jequalizer{\Gamma}{f}{g}}}\\
-\inference{\jhom{\Gamma}{\Delta}{f}\qquad\jhom{\Gamma}{\Delta}{g}}{\jhom{\jequalizer{\Gamma}{f}{g}}{\Gamma}{\jequalizerin{f}{g}}}\\
-\inference{\jhom{\greek{E}}{\Gamma}{h}\qquad\jhomeq{\greek{E}}{\Delta}{\jcomp{\greek{E}}{h}{f}}{\jcomp{\greek{E}}{h}{g}}}{\jhom{\greek{E}}{\jequalizer{\Gamma}{f}{g}}{\jequalizer{h}{f}{g}}}\\
-\inference{
+\begin{align}
+& \inference{\jhom{\Gamma}{\Delta}{f}\qquad\jhom{\Gamma}{\Delta}{g}}{\jctx{\jequalizer{\Gamma}{f}{g}}}\\
+& \inference{\jhom{\Gamma}{\Delta}{f}\qquad\jhom{\Gamma}{\Delta}{g}}{\jhom{\jequalizer{\Gamma}{f}{g}}{\Gamma}{\jequalizerin{f}{g}}}\\
+& \inference{\jhom{\greek{E}}{\Gamma}{h}\qquad\jhomeq{\greek{E}}{\Delta}{\jcomp{\greek{E}}{h}{f}}{\jcomp{\greek{E}}{h}{g}}}{\jhom{\greek{E}}{\jequalizer{\Gamma}{f}{g}}{\jequalizer{h}{f}{g}}}\\
+& \inference{
   {\begin{array}{l}
@@
   {\jhom{\greek{E}}{\jequalizer{\Gamma}{f}{g}}{\jequalizer{h}{f}{g}}}
-\end{infarray}
+\end{align}
 
@@
 \subsection{Dependent function types}
-\begin{infarray}{cc}
-\inference{\jtype{\ctxext{\Gamma}{A}}{P}}{\jtype{\Gamma}{\mprd{A}{P}}}
+\begin{align}
+& \inference{\jtype{\ctxext{\Gamma}{A}}{P}}{\jtype{\Gamma}{\mprd{A}{P}}}\\
+& \inference{\jterm{\ctxext{\Gamma}{A}}{P}{f}}{\jterm{\Gamma}{\mprd{A}{P}}{\lambda(f)}}\\
+& \inference{\jterm{\Gamma}{\mprd{A}{P}}{g}}{\jterm{\ctxext{\Gamma}{A}}{P}{\tfev(g)}}\\
+& \inference{\jterm{\ctxext{\Gamma}{A}}{P}{f}}{\jtermeq{\Gamma}{A}{\tfev(\lambda(f))}{f}}\\
+& \inference{\jtermeq{\ctxext{\Gamma}{A}}{P}{f}{f'}}{\jtermeq{\Gamma}{\mprd{A}{P}}{\lambda(f)}{\lambda(f')}}\\
+& \inference{\jtermeq{\Gamma}{\mprd{A}{P}}{g}{g'}}{\jtermeq{\ctxext{\Gamma}{A}}{P}{\tfev(g)}{\tfev(g')}}\\
 & \inference{\jtypeeq{\ctxext{\Gamma}{A}}{P}{P'}}{\jtypeeq{\Gamma}{\mprd{A}{P}}{\mprd{A}{P'}}}\\
-\inference{\jterm{\ctxext{\Gamma}{A}}{P}{f}}{\jterm{\Gamma}{\mprd{A}{P}}{\lambda(f)}}
-& \inference{\jtermeq{\ctxext{\Gamma}{A}}{P}{f}{f'}}{\jtermeq{\Gamma}{\mprd{A}{P}}{\lambda(f)}{\lambda(f')}}\\
-\inference{\jterm{\Gamma}{\mprd{A}{P}}{g}}{\jterm{\ctxext{\Gamma}{A}}{P}{\tfev(g)}}
-& \inference{\jtermeq{\Gamma}{\mprd{A}{P}}{g}{g'}}{\jtermeq{\ctxext{\Gamma}{A}}{P}{\tfev(g)}{\tfev(g')}}\\
-\inference{\jterm{\ctxext{\Gamma}{A}}{P}{f}}{\jtermeq{\Gamma}{A}{\tfev(\lambda(f))}{f}}
-\end{infarray}
+\end{align}
 
@@
 \subsection{The natural numbers}
+
+\section{A categorical explanation of the variable-free type theory}
+
+The categorical interpretation of type theory is a category $\catfont{C}$ of
+contexts, for each object $\Gamma$ of $\catfont{C}$ a class $\catfont{type}(\Gamma)$
+of morphisms into $\Gamma$ and for each $A\in\catfont{type}(\Gamma)$ a category
+$\catfont{term}(A)$ of sections of $A$.
+
+\begin{itemize}
+\item The context extension $\ctxext{\Gamma}{A}$ is the domain of $A$.
+\item There is a choice of $\Gamma\times\Delta$ for each $\Gamma$ and $\Delta$
+      for which the first projection $\pi_1:\Gamma\times\Delta\to\Gamma$ is in $\catfont{type}(\Gamma)$;
+      this is the weakening $\ctxwk{\Gamma}{\Delta}$.
+\item For each $A\in\catfont{type}(\Gamma)$ and $x:A$ there is a choice of
+      $\pi_1:\Gamma\times_A P\to\Gamma$ which is in $\catfont{type}(\Gamma)$.
+\end{itemize}

### 2013-11-09 article-models-intro.tex
rewritten the introduction to give an overview of what the project is about, issue #15

@@
 \section{Introduction}
-In this paper we define a notion of internal model which is adapted to
-the univalent setting in which we work. Thus, we only describe what it means
-to be an internal model when the ambient type theory is univalent. Our
-presentation is derived from the definition of~\cite{Dybjer1996}. There,
-an internal model is defined to be a category with families with interpretations
-for the basic type constructors $\Pi$, $\Sigma$ and $\idtypevar{}$. However,
-a category in~\cite{Dybjer1996} has a set of objects and for every two objects
-a setoid of morphisms. This is too restrictive for our purpose. Ideally, we
-would start with an $\infty$-category of contexts and an $\infty$-presheaf
-of families over it. This presents us with the problem that a notion of
-$\infty$-category has not been formulated \emph{in} type theory. Therefore,
-we make a deviation from the approach in~\cite{Dybjer1996}. We shall
-start with a type $\tfctx$ of contexts, a family $\mftypfunc:\tfctx\to\type$
-of families over it and a family $\terms{\blank}:\prd*{\Gamma:\tfctx}\mftyp{\mfM}{\Gamma}
-\to\type$ sending a type $A$ in context $\Gamma$ to its type $\terms{A}$ of
-terms; then we shall directly interpret $\Pi$-types, declaring
-$\terms{A\to B}$ to be $\ctxhom[\Gamma]{A}{B}$. The idea is that by letting
-$\Pi$ play a fundamental role in the definition of a category with families
-$\mathcal{C}$,
-the morphisms get their properties directly from the internal type theory
-of $\mathcal{C}$.
+This project has its origins in the beginning of 2013, when I proved a version
+of the descent theorem for homotopy colimits in type theory while I was working
+with Bas Spitters to develop notions from higher category theory in the
+univalent foundations. To arrive at a
+notion of diagram general enough to capture the higher inductive types described
+in chapter 6 of \cite{TheBook} excluding the truncations we needed type
+theoretical graphs. The graphs form a model of type theory and indeed we needed
+several of the basic type theoretical operations to give an efficient approach
+to the descent theorem. However it was not an issue to describe the graph model
+and the sense in which it was a model, the presentation of it was unsystematic
+and in that sense unsatisfactory too.
 
-\begin{comment}
-We briefly list the data of which an internal model of type theory consits. We
-consider internal models in the style of~\cite{Dybjer1996}. Thus we will
-describe a model $\mfM$ as a category with families. However, since our setting is
-univalent type theory we will deviate from~\cite{Dybjer1996} in the following
-respects: first of all, we make use of typical ambiguity and hence omit
-reference to a thing called \emph{sort}.
+From the experience of working with the graph model it became appearent that the
+answer we had of the question of what it models, i.e.~Martin L\"of type theory
+with the univalence axiom, could have been different if we were to describe a
+type theory based on the mechanism on which the model was built. The particular
+featueres of the graph model were that there was a type of graphs resembling a category
+of contexts, for each context a notion of what it means to be a family of graphs
+and a term thereof; there are the operations of context extension, weakening and
+substitution. In other words, the graph model is an instance of type theretical
+categories with families, an invention of Peter Dybjer.
 
-We organize the definition of an internal model $\mfM$ as follows: in
-\autoref{internal-model-contexts} we describe the category of contexts itself;
-in \autoref{internal-model-families} we describe the families over a given context
-and the related operations; in \autoref{internal-model-constructors} we
-describe the basic constructors in the internal model.
-\end{comment}
+As we know however, the type theoretical description of a category as given in
+chapter 11 of \cite{TheBook} is restricted to categories which have a $1$-type
+as the type of objects and for every two objects a $0$-type of morphisms.
+Although categories with families are like that too, these are not very natural
+restrictions when one is concerned about internal models of type theory. It
+seems we're in the need of a higher categorical formulation of what an internal
+model, or a category with family, is. But type theory has the tendency to take
+care of it's own higher categorical structure and one could hope that type
+theory is an elementary theory of the category with families of category with
+families.
 
-\subsection{Ideas in the definition}
+\subsection{Ideas in the definition of internal models}
 An internal model of type theory is like a category with families, but we want
@@
 
-\begingroup
-\color{red}
-\begin{rmk}
-Currently, the definition seems to be circular. To define a model we need that
-context extension, weakening and substitution be morphisms of models. A morphism
-of models needs to preserve context extension, weakening and substitution.
-Moreover, its definition requires the notion of composition of morphisms.
+\subsection{The elementary theory of the category with families of categories
+with families}
 
-I've done it this way because it guides me to what the rules should be, trusting
-that I can work my way back to give a (possibly less transparent) definition
-which contains no circularities without doubt.
-\end{rmk}
-\endgroup
+Thus the idea was born to give a new description of Martin-L\"of type theory
+which is more faithful to the idea that type theory is an algebraic description
+of higher category theory. The features we have in mind for this description
+is that:
+\begin{enumerate}
+\item As in Martin-L\"of type theory there are the three basic judgments
+      \begin{align*}
+        & \Gamma\text{ is a context:} & \jctx*{\Gamma}\\
+        & A\text{ is a type in context }\Gamma\text{:} & \jtype*{\Gamma}{A}\\
+        & x\text{ is a term of $A$ in context }\Gamma\text{:} & \jterm*{\Gamma}{A}{x}
+      \end{align*}
+      together with the three accompanying judgments for judgmental equality.
+      Our new description of type theory should be such that we can faithfully
+      make the translation to the three assertions
+      \begin{enumerate}
+      \item $\Gamma$ is an object\\
+      \item $A$ is a fibration over $\Gamma$\\
+      \item $x$ is a section of the fibration $A$. 
+      \end{enumerate}
+\item There is a basic type theory which has just the operations
+      \begin{description}
+      \item[extension] taking the domain of the fibration $A$ over $\Gamma$.
+      \item[weakening] given an object $\Gamma$, weakening assigns to an object
+      $\Delta$ trivial fibration with $\Delta$ as its fibers.
+      \item[substitution] taking the fiber at a point. Fibers should always
+      be equivalent to homotopy fibers.
+      \end{description}
+      Thus we aim for a closer connection between the syntactic operations and
+      the operations of models.
+\item Models of this basic type theory should be weak $\omega$-categories. Thus,
+      it should not be necessary to start the definition of a model with a bunch
+      of contexts and substitution morphisms between them: these and the
+      operations on them, such as composition and associativity thereof, should
+      be a consequence of the interpretation of the basic type theory.
+\item There shouldn't be an empty context and dependent function types right
+      from the start. These correspond to having a terminal object and the
+      models of type theory with dependent function types are the locally
+      cartesian closed models.
+\item Contexts should not be lists of variable declarations and all the
+      operations should be explicitly invariant under context extension.
+\item There should be a large zoo of extensions of the basic type theory. First
+      off, we can extend the basic type theory with identity types, dependent
+      function types and inductive types. We want a general description of
+      homotopy colimits in type theory, which should become available once we
+      know what categories (models of the basic type theory) are. Other types
+      such as the truncations could be added as well.
+
+      An extension of a different kind is the addition of the univalence axiom.
+\end{enumerate}
+
+\subsection{The further goals of the project}
+The first aim of this project is thus to give this new description of type
+theory with these properties. The second aim is to define a general notion of
+an internal model and provide several examples thereof. Among the examples
+should be
+\begin{enumerate}
+\item The setoid model. This is a classical one so it should be there. It might
+      be a bit different in our case. We'll want to construct a setoid model
+      of the basic type theory to interpret identity types without necessarily
+      interpreting dependent function types (but that should remain possible).
+\item The graph model. This is the model that leads to the first version of the
+      descent theorem.
+\item A univalent universe should be a model.
+\item Equifibered diagrams over a graph should be the fibrations in some model.
+\item The model of all models. Very likely, we will even define this simultaneously
+      with the notion of model.
+\item The model of weak $\omega$-groupoids. Since we think of models of the
+      basic type theory as weak $\omega$-categories, it is not too hard to
+      provide a condition on those models which enables us to talk about
+      weak $\omega$-categories. In fact, we have several options for such a
+      condition. We investigate those; we also investigate how they relate to
+      Brunerie's weak $\omega$-groupoids. Presumably, they form an internal
+      model without too much fuss (because we already have an internal model
+      of internal models at this stage) and we can ask whether this models
+      the univalence axiom. In fact, the model of internal models might already
+      have modeled the univalence axiom. 
+\end{enumerate}
+
+We also wish to extend the descent property to a more general class of homotopy
+colimits. The general recipe has already been implicit in the first descent
+property where we have used diagrams over graphs. We have the following in mind:
+\begin{enumerate}
+\item Give an internal description of a diagram. Probably this is the same thing
+      as an internal model of the basic type theory with or without identity types.
+\item Define the notion of homotopy colimit thereof.
+\item Define the notion if equifibered diagrams.
+\item Formulate and prove descent.
+\end{enumerate}
+
+Using the descent property we can view $\tfcolim$ as an operation of type theory
+It should preserve all the basic operations, identity types, dependent function
+types, itself, etc.

### 2013-11-10 article-models-defn.tex
moved some stuff around

@@
 
-\subsection{Possible definitions of weak $\omega$-groupoids}
+\subsection{Some possibilities}
 
@@
 Let's take the class of all terms instead:
-
-\begin{infarray}{c}
-\inference{\jterm{\Gamma}{A}{x}}{\jtype{\ctxext{\Gamma}{A}}{\eqtype{x}}}\\
-\inference{\jterm{\Gamma}{A}{x}}{\jterm{\Gamma}{\subst{x}{\eqtype{x}}}{\refl{x}}}\\
-\inference{\jtype{\ctxext{{\Gamma}{A}}{\eqtype{x}}}{P}\qquad\jterm{\Gamma}{\subst{\refl{x}}{{x}{P}}}{d}}
-          {\jterm{\ctxext{{\Gamma}{A}}{\eqtype{x}}}{P}{\ind{\eqtype{x}}(d)}}\\
-\inference{\jtype{\ctxext{{\Gamma}{A}}{\eqtype{x}}}{P}\qquad\jterm{\Gamma}{\subst{\refl{x}}{{x}{P}}}{d}}
-          {\jtermeq{\ctxext{{\Gamma}{A}}{\eqtype{x}}}{\subst{\refl{x}}{{x}{P}}}{\subst{\refl{x}}{{x}{\ind{\eqtype{x}}(d)}}}{d}}
-\end{infarray}
-
-We get something that looks like Paulin-Mohring equality. But we should be able to use it to
-show that every $\jterm{\Gamma}{\ctxwk{\Gamma}{\Delta}}{f}$ is a trivial cofibration.
-
-For the following conjecture, note that if we have $\jterm{\Gamma}{\ctxwk{\Gamma}{\Delta}}{f}$
-Then we may consider the type $\jtype{\ctxext{{\Gamma}{\ctxwk{\Gamma}{\Delta}}}{\ctxwk{\Gamma}{{\Delta}{\Delta}}}}
-{\ctxwk{\Gamma}{\idtypevar{\Delta}}}$ and we may substitute $f$ to obtain the type
-\begin{equation*}
-\jtype{\ctxext{\Gamma}{\subst{f}{\ctxwk{\Gamma}{{\Delta}{\Delta}}}}}
-{\subst{f}{\ctxwk{\Gamma}{\idtypevar{\Delta}}}}
-\end{equation*}
-Note that we have the judgmental equality
-\begin{equation*}
-\jtypeeq{\Gamma}{\subst{f}{\ctxwk{\Gamma}{{\Delta}{\Delta}}}}{\ctxwk{\Gamma}{\Delta}}
-\end{equation*}
-so we obtain the type
-\begin{equation*}
-\jtype{\ctxext{\Gamma}{\ctxwk{\Gamma}{\Delta}}}
-{\subst{f}{\ctxwk{\Gamma}{\idtypevar{\Delta}}}}
-\end{equation*}
-We also have $\jtype{\ctxext{\Gamma}{\ctxwk{\Gamma}{\Delta}}}{\eqtype{f}}$, so we
-may ask for a term of type $\ctxwk{\eqtype{f}}{\subst{f}{\ctxwk{\Gamma}{\idtypevar{\Delta}}}}$
-in context $\ctxext{\Gamma}{\ctxwk{\Gamma}{\Delta}}$ and we can ask ourselves the question
-whether this is a trivial cofibration. This is the content of the following conjecture.
-
-\begin{conj}
-There is a term
-\begin{equation*}
-\jterm{\ctxext{{\Gamma}{\ctxwk{\Gamma}{\Delta}}}{\eqtype{f}}}{\ctxwk{\eqtype{f}}{\subst{f}{\ctxwk{\Gamma}{\idtypevar{\Delta}}}}}{\eqtoid{f}}
-\end{equation*}
-Moreover, $\eqtoid{f}$ is a trivial cofibration.
-\end{conj}
-
-\begin{proof}
-We use the induction principle of $\eqtype{f}$ to construct $\eqtoid{f}$. Thus, we
-have to find a term
-\begin{equation*}
-\jterm
-  {\Gamma}
-  {\subst{\refl{f}}{{f}{\ctxwk{\eqtype{f}}{\subst{f}{\ctxwk{\Gamma}{\idtypevar{\Delta}}}}}}}
-  {\subst{\refl{f}}{{f}{\eqtoid{f}}}}
-\end{equation*}
-Note that we have the judgmental equalities
-\begin{align*}
-\subst{\refl{f}}{{f}{\ctxwk{\eqtype{f}}{\subst{f}{\ctxwk{\Gamma}{\idtypevar{\Delta}}}}}}
-& \jdeq \subst{f}{{\ctxwk{{\Gamma}{\Delta}}{\refl{f}}}{\ctxwk{\eqtype{f}}{\subst{f}{\ctxwk{\Gamma}{\idtypevar{\Delta}}}}}} \\
-& \jdeq \subst{f}{\ctxwk{\Gamma}{\subst{f}{\ctxwk{\Gamma}{\idtypevar{\Delta}}}}}
-\end{align*}
-where we find the term $\subst{f}{\ctxwk{\Gamma}{\subst{f}{\ctxwk{\Gamma}{\refl{\Delta}}}}}$.
-\end{proof}
-
-\begin{conj}
-Suppose we have a type $\eqtype{x}$ for every term $\jterm{\Gamma}{A}{x}$. Then
-every term $\jterm{\Gamma}{\ctxwk{\Gamma}{\Delta}}{f}$ is a trivial cofibration.
-\end{conj}
-
-\begin{proof}
-Suppose that $Q$ is a type in context $\Delta$. We want to show that
-\begin{equation*}
-\inference{\jterm{\Gamma}{\subst{f}{\ctxwk{\Gamma}{Q}}}{g}}{\jterm{\Delta}{Q}{\tilde{g}}}
-\end{equation*}
-Let $g$ be a term of $\subst{f}{\ctxwk{\Gamma}{Q}}$ in context $\Gamma$. We get
-the term
-\begin{equation*}
-\jterm{\Gamma}{\subst{\refl{f}}{{f}{\ctxwk{\eqtype{f}}{{\Gamma}{Q}}}}}{\subst{\refl{f}}{\ctxwk{\eqtype{f}}{g}}}
-\end{equation*}
-which is judgmentally equal to $g$; the types
-
-For any term $\jterm{\Gamma}{\subst{f}{\ctxwk{\Gamma}{Q}}}{g}$ we get a term 
-$\jterm{\Gamma}{\subst{\refl{x}}{{x}{P}}}{d}$.
-
-$\jterm{\ctxext{{\Gamma}{\ctxwk{\Gamma}{\Delta}}}{\eqtype{f}}}{\ctxwk{\Gamma}{Q}}{\ind{\eqtype{f}}(d)}$.
-\end{proof}
-
-However, we could take other classes of terms, such as the projections
-$\proj1:\ctxwk{\ctxext{\Gamma}{A}}{\Gamma}$. Let's see how that goes:
-
-\begin{infarray}{c}
-\inference{\jtype{\Gamma}{A}}{\jtype{\ctxext{{\Gamma}{A}}{\ctxwk{\ctxext{\Gamma}{A}}{\Gamma}}}{\eqtype{\proj1^A}}}\\
-\inference{\jtype{\Gamma}{A}}{\jterm{\ctxext{\Gamma}{A}}{\subst{\proj1^A}{\eqtype{\proj1^A}}}{\refl{\proj1^A}}}
-\end{infarray}

### 2013-11-10 article-models-intro.tex
moved some stuff around

@@
 theory with these properties. The second aim is to define a general notion of
-an internal model and provide several examples thereof. Among the examples
-should be
-\begin{enumerate}
-\item The setoid model. This is a classical one so it should be there. It might
+an internal model and provide several examples thereof. In the definition of
+internal models we envision that all the operations of type theory become
+morphisms of type theory: they are compatible with each and every morphism
+of type theory. The compatibility of extension, weakening and substitution is
+already part of the basic type theory. The compatibility of identity types
+with weakening for instance, is a form of function extensionality. The 
+compatibility of dependent function types with identity types is the usual
+funchtion extensionality principle. The
+compatibility of a universe operation with identity types should be univalence.
+The descent theorem implies that the colimit operation is also compatible with
+all the type theoretic operations (sigma and id are checked).
+
+Among the examples of internal models we should have:
+\begin{description}
+\item[The setoid model] This is a classical one so it should be there. It might
       be a bit different in our case. We'll want to construct a setoid model
@@
       interpreting dependent function types (but that should remain possible).
-\item The graph model. This is the model that leads to the first version of the
+\item[The graph model] This is the model that leads to the first version of the
       descent theorem.
-\item A univalent universe should be a model.
-\item Equifibered diagrams over a graph should be the fibrations in some model.
-\item The model of all models. Very likely, we will even define this simultaneously
+\item[Univalent unvirses] A univalent universe should be a model.
+\item[Equifibered diagrams over a graph] should be the fibrations in some model.
+\item[The model of all models] Very likely, we will even define this simultaneously
       with the notion of model.
-\item The model of weak $\omega$-groupoids. Since we think of models of the
+\item[The model of weak $\omega$-groupoids] Since we think of models of the
       basic type theory as weak $\omega$-categories, it is not too hard to
@@
       have modeled the univalence axiom. 
-\end{enumerate}
+\end{description}
 
@@
 \item Define the notion if equifibered diagrams.
-\item Formulate and prove descent.
+\item Formulate and prove descent: the map defined by substitution
+ from families over the colimit to the equifibered diagrams is an equivalence.
 \end{enumerate}
@@
 types, itself, etc.
+
+Another thing we must keep in mind is that the compatibility rules in the basic
+type theory are strict. We should look to the possibility to weaken them.
+
+\subsection{Overview of the document}
+In \autoref{eg} we begin the project with exploring internal models of type theory
+in type theoretical setting of \cite{TheBook}.

### 2013-11-10 article-models-tt.tex
moved some stuff around

@@
 In these rules we assume we have $\jtype{\Gamma}{A}$ in the hypotheses.
-
-\begin{infarray}{c}
-\inference{}
+\begin{align}
+& \inference{}
 {\jterm{\ctxext{\Gamma}{A}}{\ctxwk{A}{{\Gamma}{\ctxext{\Gamma}{A}}}}{\pair_A}}\\
-\inference{
+& \inference{
   \jtype{\ctxext{\Gamma}{A}}{P}
@@
   {\jterm{\ctxext{\Gamma}{A}}{P}{\ind{\tfext_\Gamma(A)}(f)}}\\
-\inference{
+& \inference{
   \jtype{\ctxext{\Gamma}{A}}{P}
@@
   {\jtermeq{\ctxext{\Gamma}{A}}{\subst{\pair_A}{\ctxwk{A}{{\Gamma}{P}}}}{\subst{\pair_A}{\ctxwk{A}{{\Gamma}{\ind{\tfext_\Gamma(A)}(f)}}}}{f}}
-\end{infarray}
+\end{align}
 
@@
 
-\begin{infarray}{c}
-\inference{}{\jtermeq{\ctxext{\Gamma}{A}}{\ctxwk{A}{{\Gamma}{\ctxext{\Gamma}{A}}}}{\pair_A}{\pair_{A'}}}\\
-\inference{\jtype{\ctxext{\Gamma}{A}}{P}\qquad\jterm{\ctxext{\Gamma}{A}}{\subst{\pair_A}{\ctxwk{A}{{\Gamma}{P}}}}{f}}
+\begin{align}
+& \inference{}{\jtermeq{\ctxext{\Gamma}{A}}{\ctxwk{A}{{\Gamma}{\ctxext{\Gamma}{A}}}}{\pair_A}{\pair_{A'}}}\\
+& \inference{\jtype{\ctxext{\Gamma}{A}}{P}\qquad\jterm{\ctxext{\Gamma}{A}}{\subst{\pair_A}{\ctxwk{A}{{\Gamma}{P}}}}{f}}
 {\jtermeq{\ctxext{\Gamma}{A}}{P}{\ind{\tfext_\Gamma(A)}(f)}{\ind{\tfext_\Gamma(A')}(f)}}
-\end{infarray}
+\end{align}
 
@@
 
-\begin{infarray}{c}
-\inference{}
+\begin{align}
+& \inference{}
 {\jterm{\ctxext{{\Gamma}{A}}{P}}{\ctxwk{P}{{A}{\ctxext{A}{P}}}}{\pair_P}}\\
-\inference{
+& \inference{
   \jtype{\ctxext{\Gamma}{{A}{P}}}{Q}
@@
   {\jterm{\ctxext{\Gamma}{{A}{P}}}{Q}{\ind{\tfext_A(P)}(f)}}\\
-\inference{
+& \inference{
   \jtype{\ctxext{\Gamma}{{A}{P}}}{Q}
@@
   {\jtermeq{\ctxext{{\Gamma}{A}}{P}}{\subst{\pair_P}{\ctxwk{P}{{A}{Q}}}}{\subst{\pair_P}{\ctxwk{P}{{A}{\ind{\tfext_A(P)}(f)}}}}{f}}
-\end{infarray}
+\end{align}
 
@@
 
-\section{Identity types}
-The identity types may now be formulated just as in ordinary Martin\nobreakdash-L\"of
-type theory. However, since all model categories have path objects, we will also
-introduce identity types for contexts.
-
-We will use the symbol $\reflsym$ slightly different than the book does. For us,
-$\refl{A}$ is a term of $\subst{\idfunc[A]}{\idtypevar{A}}$ in context
-$\ctxext{\Gamma}{A}$ and we will write $\subst{x}{\refl{A}}$ for the reflexivity
-path at $x$ (which would have been denoted by $\refl{x}$ in the book).
-
-A notable difference in the formulation of identity types is that in our current
-setting we must state the elimination rule not only for families $P$ in context
-$\ctxext{{\Gamma}{\ctxwk{\Gamma}{\Gamma}}}{\idtypevar{\Gamma}}$,
-but also for families $Q$ in context 
-$\ctxext{{{\Gamma}{\ctxwk{\Gamma}{\Gamma}}}{\idtypevar{\Gamma}}}{P}$. The reason
-is that all operations have to be closed under slicing: everything may happen
-in a context. Secretly, a reason is that we don't have dependent function types.
-We wouldn't even be able to find the transport maps if we didn't state the
-identity elimination in an extended context.
-
-\begin{align}
-& \inference{\jctx{\Gamma}}{\jtype{\ctxext{\Gamma}{\ctxwk{\Gamma}{\Gamma}}}{\idtypevar{\Gamma}}}\\
-& \inference{\jctx{\Gamma}}{\jterm{\Gamma}{\subst{\idfunc[\Gamma]}{\idtypevar{\Gamma}}}{\refl{\Gamma}}}\\
-& \inference{\jtype{\ctxext{{\Gamma}{\ctxwk{\Gamma}{\Gamma}}}{\idtypevar{\Gamma}}}{P}
-           \qquad
-           \jterm{\Gamma}{\subst{\refl{\Gamma}}{{\idfunc[\Gamma]}{P}}}{d}}
-           {\jterm{\ctxext{{\Gamma}{\ctxwk{\Gamma}{\Gamma}}}{\idtypevar{\Gamma}}}{P}{\tfJ(d)}}\\
-& \inference{\jtype{\ctxext{{\Gamma}{\ctxwk{\Gamma}{\Gamma}}}{\idtypevar{\Gamma}}}{P}\qquad
-\jterm{\Gamma}{\subst{\refl{\Gamma}}{{\idfunc[\Gamma]}{P}}}{d}}
-{\jtermeq
-  {\ctxext{{\Gamma}{\ctxwk{\Gamma}{\Gamma}}}{\idtypevar{\Gamma}}}
-  {\subst{\refl{\Gamma}}{{\idfunc[\Gamma]}{P}}}
-  {\subst{\refl{\Gamma}}{{\idfunc[\Gamma]}{\tfJ(d)}}}
-  {d}}\\
-& \inference{\jtype{\ctxext{{{\Gamma}{\ctxwk{\Gamma}{\Gamma}}}{\idtypevar{\Gamma}}}{P}}{Q}
-           \qquad
-           \jterm{\ctxext{\Gamma}{\subst{\refl{\Gamma}}{{\idfunc[\Gamma]}{P}}}}{\subst{\refl{\Gamma}}{{\idfunc[\Gamma]}{Q}}}{d}}
-           {\jterm{\ctxext{{{\Gamma}{\ctxwk{\Gamma}{\Gamma}}}{\idtypevar{\Gamma}}}{P}}{Q}{\tfJ(d)}}\\
-& \inference
-  {\jtype
-    {\ctxext{{{\Gamma}{\ctxwk{\Gamma}{\Gamma}}}{\idtypevar{\Gamma}}}{P}}
-    {Q}
-  \qquad
-  \jterm
-    {\ctxext{\Gamma}{\subst{\refl{\Gamma}}{{\idfunc[\Gamma]}{P}}}}
-    {\subst{\refl{\Gamma}}{{\idfunc[\Gamma]}{Q}}}
-    {d}}
-  {\jtermeq
-    {\ctxext{{{\Gamma}{\ctxwk{\Gamma}{\Gamma}}}{\idtypevar{\Gamma}}}{P}}
-    {\subst{\refl{\Gamma}}{{\idfunc[\Gamma]}{Q}}}
-    {\subst{\refl{\Gamma}}{{\idfunc[\Gamma]}{\tfJ(d)}}}
-    {d}}
-\end{align}
-
-Likewise, we introduce identity types in a context.
-
-\begin{align}
-& \inference{\jtype{\Gamma}{A}}{\jtype{\ctxext{{\Gamma}{A}}{\ctxwk{A}{A}}}{\idtypevar{A}}}\\
-& \inference{\jtype{\Gamma}{A}}{\jterm{\ctxext{\Gamma}{A}}{\subst{\idfunc[A]}{\idtypevar{A}}}{\refl{A}}}\\
-& \inference{\jtype{\ctxext{{{\Gamma}{A}}{\ctxwk{A}{A}}}{\idtypevar{A}}}{P}\qquad
-\jterm{\ctxext{\Gamma}{A}}{\subst{\refl{A}}{{\idfunc[A]}{P}}}{d}}
-{\jterm{\ctxext{{{\Gamma}{A}}{\ctxwk{A}{A}}}{\idtypevar{A}}}{P}{\tfJ(d)}}\\
-& \inference{\jtype{\ctxext{{{\Gamma}{A}}{\ctxwk{A}{A}}}{\idtypevar{A}}}{P}\qquad
-\jterm{\ctxext{\Gamma}{A}}{\subst{\refl{A}}{{\idfunc[A]}{P}}}{d}}
-{\jtermeq
-  {\ctxext{{{\Gamma}{A}}{\ctxwk{A}{A}}}{\idtypevar{A}}}
-  {\subst{\refl{A}}{{\idfunc[A]}{P}}}
-  {\subst{\refl{A}}{{\idfunc[A]}{\tfJ(d)}}}
-  {d}}
-\end{align}
-
-Suppose we have terms $\jterm{\Gamma}{A}{x}$ and $\jterm{\Gamma}{A}{y}$. Then
-we may define $\id[A]{x}{y}\defeq\subst{y}{{x}{\idtypevar{A}}}$. A term
-$\jterm{\Gamma}{\id[A]{x}{y}}{p}$ is called an identification of $x$ and $y$.
-
-\subsubsection{Basic properties of identity types}
-In this subsubsection we prove some basic properties of identity types, just to
-know whether we got the type theory right.
-
-Suppose we have a family $\jtype{\ctxext{\Gamma}{A}}{P}$. Then we can consider
-the family $\jtype{\ctxext{{\Gamma}{A}}{\ctxwk{A}{A}}}{\ctxwk{A}{P}}$, which has the role of the family $\jtype{\Gamma,\,x,y:A}{P(y)}$
-of ordinary Martin-L\"of type theory. We may also consider the family
-$\jtype{\ctxext{{\Gamma}{A}}{\ctxwk{A}{A}}}{\ctxwk{{A}{A}}{P}}$; this one has the
-role of the family $\jtype{\Gamma,\,x,y:A}{P(x)}$. Those are families in the
-same context, so we have
-\begin{equation*}
-\jtype{\ctxext{{{\Gamma}{A}}{\ctxwk{A}{A}}}{\ctxwk{{A}{A}}{P}}}{\ctxwk{\ctxwk{{A}{A}}{P}}{{A}{P}}}
-\end{equation*}
-
-\begin{lem}
-There is a term
-\begin{equation*}
-\jterm{\ctxext{{{{\Gamma}{A}}{\ctxwk{A}{A}}}{\idtypevar{A}}}{\ctxwk{\idtypevar{A}}{{{A}{A}}{P}}}}{\ctxwk{\idtypevar{A}}{{{{A}{A}}{P}}{{A}{P}}}}{\transfibf{P}}
-\end{equation*}
-\end{lem}
-
-\begin{proof}
-By identity elimination it suffices to find a term
-\begin{equation*}
-\jterm{\ctxext{\Gamma}{A}}{\subst{\refl{A}}{{\idfunc[A]}{\ctxwk{\idtypevar{A}}{{{{A}{A}}{P}}{{A}{P}}}}}}{t}
-\end{equation*}
-By the judgmental equality $\jtypeeq{\Gamma}{\subst{{x}{f}}{{x}{Q}}}{\subst{x}{{f}{Q}}}$
-it follows that we have the judgmental equalities
-\begin{align*}
-& \subst{\refl{A}}{{\idfunc[A]}{\ctxwk{\idtypevar{A}}{{{{A}{A}}{P}}{{A}{P}}}}}\\
-& \qquad \jdeq \subst{\idfunc[A]}{{\ctxwk{{A}{A}}{\refl{A}}}{\ctxwk{\idtypevar{A}}{{{{A}{A}}{P}}{{A}{P}}}}}\\
-& \qquad \jdeq \subst{\idfunc[A]}{\ctxwk{{{A}{A}}{P}}{{A}{P}}}\\
-& \qquad \jdeq \ctxwk{\subst{\idfunc[A]}{\ctxwk{{A}{A}}{P}}}{\subst{\idfunc[A]}{\ctxwk{A}{P}}}\\
-& \qquad \jdeq \ctxwk{P}{P}
-\end{align*}
-in context $\ctxext{{\Gamma}{A}}{P}$. We have the term $\jterm{\ctxext{{\Gamma}{A}}{P}}{\ctxwk{P}{P}}{\idfunc[P]}$.
-\end{proof}
-
-Suppose we have a term $\jterm{\ctxext{\Gamma}{A}}{P}{f}$.
-
-Using identity types, we can assert that a function $\jhom{\Gamma}{\Delta}{f}$ has
-a left inverse $\jhom{\Delta}{\Gamma}{g}$ by asserting that there is an identification
-\begin{equation*}
-..
-\end{equation*}
-
-\subsubsection{Compatibility of identity types with extension}
-
-\subsubsection{Compatibility of identity types with weakening}
-Suppose $A$ and $B$ are types in context $\Gamma$. Then we can consider the types
-\begin{align*}
-\jtype*{\ctxext{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{{B}{B}}}}{\ctxwk{A}{\idtypevar{B}}}\\
-\jtype*{\ctxext{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{{A}{B}}{{A}{B}}}}{\idtypevar{\ctxwk{A}{B}}}
-\end{align*}
-Note that we have the judgmental equality
-\begin{equation*}
-\jtypeeq
-  {\ctxext{{\Gamma}{A}}{\ctxwk{A}{B}}}
-  {\ctxwk{A}{{B}{B}}}
-  {\ctxwk{{A}{B}}{{A}{B}}}
-\end{equation*}
-so $\ctxwk{A}{\idtypevar{B}}$ and $\idtypevar{\ctxwk{A}{B}}$ are types in the
-same context.
-
-\begin{lem}
-There is a term of type
-\begin{equation*}
-\jterm
-  {\ctxext{{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{{A}{B}}{{A}{B}}}}{\idtypevar{\ctxwk{A}{B}}}}
-  {\ctxwk{\idtypevar{\ctxwk{A}{B}}}{\ctxwk{A}{\idtypevar{B}}}}
-  {\typefont{idwktowkid}}
-\end{equation*}
-\end{lem}
-
-\begin{proof}
-By the identity elimination rule it suffices to find a term
-\begin{equation*}
-\jterm
-  {\ctxext{{\Gamma}{A}}{\ctxwk{A}{B}}}
-  {\subst{\refl{\ctxwk{A}{B}}}{{\idfunc[\ctxwk{A}{B}]}{\ctxwk{\idtypevar{\ctxwk{A}{B}}}{\ctxwk{A}{\idtypevar{B}}}}}}
-  {i}
-\end{equation*}
-We may simplify the type $\subst{\refl{\ctxwk{A}{B}}}{{\idfunc[\ctxwk{A}{B}]}{\ctxwk{\idtypevar{\ctxwk{A}{B}}}{\ctxwk{A}{\idtypevar{B}}}}}$ as follows:
-\begin{align*}
-& \subst{\refl{\ctxwk{A}{B}}}{{\idfunc[\ctxwk{A}{B}]}{\ctxwk{\idtypevar{\ctxwk{A}{B}}}{\ctxwk{A}{\idtypevar{B}}}}}\\
-& \qquad\jdeq \subst{\idfunc[\ctxwk{A}{B}]}{{\ctxwk{..}{\refl{\ctxwk{A}{B}}}}{\ctxwk{\idtypevar{\ctxwk{A}{B}}}{\ctxwk{A}{\idtypevar{B}}}}}
-\end{align*}
-\end{proof}
-
-\subsubsection{Compatibility of identity types with substitution}
+\input{article-models-idtypes}
 

### 2013-11-10 article-models-idtypes.tex
forgot to add the idtypes file

\section{Identity types}
The identity types may now be formulated just as in ordinary Martin\nobreakdash-L\"of
type theory. However, since all model categories have path objects, we will also
introduce identity types for contexts.

We will use the symbol $\reflsym$ slightly different than the book does. For us,
$\refl{A}$ is a term of $\subst{\idfunc[A]}{\idtypevar{A}}$ in context
$\ctxext{\Gamma}{A}$ and we will write $\subst{x}{\refl{A}}$ for the reflexivity
path at $x$ (which would have been denoted by $\refl{x}$ in the book).

A notable difference in the formulation of identity types is that in our current
setting we must state the elimination rule not only for families $P$ in context
$\ctxext{{\Gamma}{\ctxwk{\Gamma}{\Gamma}}}{\idtypevar{\Gamma}}$,
but also for families $Q$ in context 
$\ctxext{{{\Gamma}{\ctxwk{\Gamma}{\Gamma}}}{\idtypevar{\Gamma}}}{P}$. The reason
is that all operations have to be closed under slicing: everything may happen
in a context. Secretly, a reason is that we don't have dependent function types.
We wouldn't even be able to find the transport maps if we didn't state the
identity elimination in an extended context.

\begin{align}
& \inference{\jctx{\Gamma}}{\jtype{\ctxext{\Gamma}{\ctxwk{\Gamma}{\Gamma}}}{\idtypevar{\Gamma}}}\\
& \inference{\jctx{\Gamma}}{\jterm{\Gamma}{\subst{\idfunc[\Gamma]}{\idtypevar{\Gamma}}}{\refl{\Gamma}}}\\
& \inference{\jtype{\ctxext{{\Gamma}{\ctxwk{\Gamma}{\Gamma}}}{\idtypevar{\Gamma}}}{P}
           \qquad
           \jterm{\Gamma}{\subst{\refl{\Gamma}}{{\idfunc[\Gamma]}{P}}}{d}}
           {\jterm{\ctxext{{\Gamma}{\ctxwk{\Gamma}{\Gamma}}}{\idtypevar{\Gamma}}}{P}{\tfJ(d)}}\\
& \inference{\jtype{\ctxext{{\Gamma}{\ctxwk{\Gamma}{\Gamma}}}{\idtypevar{\Gamma}}}{P}\qquad
\jterm{\Gamma}{\subst{\refl{\Gamma}}{{\idfunc[\Gamma]}{P}}}{d}}
{\jtermeq
  {\ctxext{{\Gamma}{\ctxwk{\Gamma}{\Gamma}}}{\idtypevar{\Gamma}}}
  {\subst{\refl{\Gamma}}{{\idfunc[\Gamma]}{P}}}
  {\subst{\refl{\Gamma}}{{\idfunc[\Gamma]}{\tfJ(d)}}}
  {d}}\\
& \inference{\jtype{\ctxext{{{\Gamma}{\ctxwk{\Gamma}{\Gamma}}}{\idtypevar{\Gamma}}}{P}}{Q}
           \qquad
           \jterm{\ctxext{\Gamma}{\subst{\refl{\Gamma}}{{\idfunc[\Gamma]}{P}}}}{\subst{\refl{\Gamma}}{{\idfunc[\Gamma]}{Q}}}{d}}
           {\jterm{\ctxext{{{\Gamma}{\ctxwk{\Gamma}{\Gamma}}}{\idtypevar{\Gamma}}}{P}}{Q}{\tfJ(d)}}\\
& \inference
  {\jtype
    {\ctxext{{{\Gamma}{\ctxwk{\Gamma}{\Gamma}}}{\idtypevar{\Gamma}}}{P}}
    {Q}
  \qquad
  \jterm
    {\ctxext{\Gamma}{\subst{\refl{\Gamma}}{{\idfunc[\Gamma]}{P}}}}
    {\subst{\refl{\Gamma}}{{\idfunc[\Gamma]}{Q}}}
    {d}}
  {\jtermeq
    {\ctxext{{{\Gamma}{\ctxwk{\Gamma}{\Gamma}}}{\idtypevar{\Gamma}}}{P}}
    {\subst{\refl{\Gamma}}{{\idfunc[\Gamma]}{Q}}}
    {\subst{\refl{\Gamma}}{{\idfunc[\Gamma]}{\tfJ(d)}}}
    {d}}
\end{align}

Likewise, we introduce identity types in a context.

\begin{align}
& \inference{\jtype{\Gamma}{A}}{\jtype{\ctxext{{\Gamma}{A}}{\ctxwk{A}{A}}}{\idtypevar{A}}}\\
& \inference{\jtype{\Gamma}{A}}{\jterm{\ctxext{\Gamma}{A}}{\subst{\idfunc[A]}{\idtypevar{A}}}{\refl{A}}}\\
& \inference{\jtype{\ctxext{{{\Gamma}{A}}{\ctxwk{A}{A}}}{\idtypevar{A}}}{P}\qquad
\jterm{\ctxext{\Gamma}{A}}{\subst{\refl{A}}{{\idfunc[A]}{P}}}{d}}
{\jterm{\ctxext{{{\Gamma}{A}}{\ctxwk{A}{A}}}{\idtypevar{A}}}{P}{\tfJ(d)}}\\
& \inference{\jtype{\ctxext{{{\Gamma}{A}}{\ctxwk{A}{A}}}{\idtypevar{A}}}{P}\qquad
\jterm{\ctxext{\Gamma}{A}}{\subst{\refl{A}}{{\idfunc[A]}{P}}}{d}}
{\jtermeq
  {\ctxext{{{\Gamma}{A}}{\ctxwk{A}{A}}}{\idtypevar{A}}}
  {\subst{\refl{A}}{{\idfunc[A]}{P}}}
  {\subst{\refl{A}}{{\idfunc[A]}{\tfJ(d)}}}
  {d}}
\end{align}

Suppose we have terms $\jterm{\Gamma}{A}{x}$ and $\jterm{\Gamma}{A}{y}$. Then
we may define $\id[A]{x}{y}\defeq\subst{y}{{x}{\idtypevar{A}}}$. A term
$\jterm{\Gamma}{\id[A]{x}{y}}{p}$ is called an identification of $x$ and $y$.

\subsubsection{Basic properties of identity types}
In this subsubsection we prove some basic properties of identity types, just to
know whether we got the type theory right.

Suppose we have a family $\jtype{\ctxext{\Gamma}{A}}{P}$. Then we can consider
the family $\jtype{\ctxext{{\Gamma}{A}}{\ctxwk{A}{A}}}{\ctxwk{A}{P}}$, which has the role of the family $\jtype{\Gamma,\,x,y:A}{P(y)}$
of ordinary Martin-L\"of type theory. We may also consider the family
$\jtype{\ctxext{{\Gamma}{A}}{\ctxwk{A}{A}}}{\ctxwk{{A}{A}}{P}}$; this one has the
role of the family $\jtype{\Gamma,\,x,y:A}{P(x)}$. Those are families in the
same context, so we have
\begin{equation*}
\jtype{\ctxext{{{\Gamma}{A}}{\ctxwk{A}{A}}}{\ctxwk{{A}{A}}{P}}}{\ctxwk{\ctxwk{{A}{A}}{P}}{{A}{P}}}
\end{equation*}

\begin{lem}
There is a term
\begin{equation*}
\jterm{\ctxext{{{{\Gamma}{A}}{\ctxwk{A}{A}}}{\idtypevar{A}}}{\ctxwk{\idtypevar{A}}{{{A}{A}}{P}}}}{\ctxwk{\idtypevar{A}}{{{{A}{A}}{P}}{{A}{P}}}}{\transfibf{P}}
\end{equation*}
\end{lem}

\begin{proof}
By identity elimination it suffices to find a term
\begin{equation*}
\jterm{\ctxext{\Gamma}{A}}{\subst{\refl{A}}{{\idfunc[A]}{\ctxwk{\idtypevar{A}}{{{{A}{A}}{P}}{{A}{P}}}}}}{t}
\end{equation*}
By the judgmental equality $\jtypeeq{\Gamma}{\subst{{x}{f}}{{x}{Q}}}{\subst{x}{{f}{Q}}}$
it follows that we have the judgmental equalities
\begin{align*}
& \subst{\refl{A}}{{\idfunc[A]}{\ctxwk{\idtypevar{A}}{{{{A}{A}}{P}}{{A}{P}}}}}\\
& \qquad \jdeq \subst{\idfunc[A]}{{\ctxwk{{A}{A}}{\refl{A}}}{\ctxwk{\idtypevar{A}}{{{{A}{A}}{P}}{{A}{P}}}}}\\
& \qquad \jdeq \subst{\idfunc[A]}{\ctxwk{{{A}{A}}{P}}{{A}{P}}}\\
& \qquad \jdeq \ctxwk{\subst{\idfunc[A]}{\ctxwk{{A}{A}}{P}}}{\subst{\idfunc[A]}{\ctxwk{A}{P}}}\\
& \qquad \jdeq \ctxwk{P}{P}
\end{align*}
in context $\ctxext{{\Gamma}{A}}{P}$. We have the term $\jterm{\ctxext{{\Gamma}{A}}{P}}{\ctxwk{P}{P}}{\idfunc[P]}$.
\end{proof}

Suppose we have a term $\jterm{\ctxext{\Gamma}{A}}{P}{f}$.

Using identity types, we can assert that a function $\jhom{\Gamma}{\Delta}{f}$ has
a left inverse $\jhom{\Delta}{\Gamma}{g}$ by asserting that there is an identification
\begin{equation*}
..
\end{equation*}

\subsubsection{Compatibility of identity types with extension}

\subsubsection{Compatibility of identity types with weakening}
Suppose $A$ and $B$ are types in context $\Gamma$. Then we can consider the types
\begin{align*}
\jtype*{\ctxext{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{{B}{B}}}}{\ctxwk{A}{\idtypevar{B}}}\\
\jtype*{\ctxext{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{{A}{B}}{{A}{B}}}}{\idtypevar{\ctxwk{A}{B}}}
\end{align*}
Note that we have the judgmental equality
\begin{equation*}
\jtypeeq
  {\ctxext{{\Gamma}{A}}{\ctxwk{A}{B}}}
  {\ctxwk{A}{{B}{B}}}
  {\ctxwk{{A}{B}}{{A}{B}}}
\end{equation*}
so $\ctxwk{A}{\idtypevar{B}}$ and $\idtypevar{\ctxwk{A}{B}}$ are types in the
same context.

\begin{lem}
There is a term of type
\begin{equation*}
\jterm
  {\ctxext{{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{{A}{B}}{{A}{B}}}}{\idtypevar{\ctxwk{A}{B}}}}
  {\ctxwk{\idtypevar{\ctxwk{A}{B}}}{\ctxwk{A}{\idtypevar{B}}}}
  {\typefont{idwktowkid}}
\end{equation*}
\end{lem}

\begin{proof}
By the identity elimination rule it suffices to find a term
\begin{equation*}
\jterm
  {\ctxext{{\Gamma}{A}}{\ctxwk{A}{B}}}
  {\subst{\refl{\ctxwk{A}{B}}}{{\idfunc[\ctxwk{A}{B}]}{\ctxwk{\idtypevar{\ctxwk{A}{B}}}{\ctxwk{A}{\idtypevar{B}}}}}}
  {i}
\end{equation*}
We may simplify the type $\subst{\refl{\ctxwk{A}{B}}}{{\idfunc[\ctxwk{A}{B}]}{\ctxwk{\idtypevar{\ctxwk{A}{B}}}{\ctxwk{A}{\idtypevar{B}}}}}$ as follows:
\begin{align*}
& \subst{\refl{\ctxwk{A}{B}}}{{\idfunc[\ctxwk{A}{B}]}{\ctxwk{\idtypevar{\ctxwk{A}{B}}}{\ctxwk{A}{\idtypevar{B}}}}}\\
& \qquad\jdeq \subst{\idfunc[\ctxwk{A}{B}]}{{\ctxwk{..}{\refl{\ctxwk{A}{B}}}}{\ctxwk{\idtypevar{\ctxwk{A}{B}}}{\ctxwk{A}{\idtypevar{B}}}}}
\end{align*}
\end{proof}

\subsubsection{Compatibility of identity types with substitution}

\subsection{Generalized identity types}

\begin{infarray}{c}
\inference{\jterm{\Gamma}{A}{x}}{\jtype{\ctxext{\Gamma}{A}}{\eqtype{x}}}\\
\inference{\jterm{\Gamma}{A}{x}}{\jterm{\Gamma}{\subst{x}{\eqtype{x}}}{\refl{x}}}\\
\inference{\jtype{\ctxext{{\Gamma}{A}}{\eqtype{x}}}{P}\qquad\jterm{\Gamma}{\subst{\refl{x}}{{x}{P}}}{d}}
          {\jterm{\ctxext{{\Gamma}{A}}{\eqtype{x}}}{P}{\ind{\eqtype{x}}(d)}}\\
\inference{\jtype{\ctxext{{\Gamma}{A}}{\eqtype{x}}}{P}\qquad\jterm{\Gamma}{\subst{\refl{x}}{{x}{P}}}{d}}
          {\jtermeq{\ctxext{{\Gamma}{A}}{\eqtype{x}}}{\subst{\refl{x}}{{x}{P}}}{\subst{\refl{x}}{{x}{\ind{\eqtype{x}}(d)}}}{d}}
\end{infarray}

We get something that looks like Paulin-Mohring equality. But we should be able to use it to
show that every $\jterm{\Gamma}{\ctxwk{\Gamma}{\Delta}}{f}$ is a trivial cofibration.

For the following conjecture, note that if we have $\jterm{\Gamma}{\ctxwk{\Gamma}{\Delta}}{f}$
Then we may consider the type $\jtype{\ctxext{{\Gamma}{\ctxwk{\Gamma}{\Delta}}}{\ctxwk{\Gamma}{{\Delta}{\Delta}}}}
{\ctxwk{\Gamma}{\idtypevar{\Delta}}}$ and we may substitute $f$ to obtain the type
\begin{equation*}
\jtype{\ctxext{\Gamma}{\subst{f}{\ctxwk{\Gamma}{{\Delta}{\Delta}}}}}
{\subst{f}{\ctxwk{\Gamma}{\idtypevar{\Delta}}}}
\end{equation*}
Note that we have the judgmental equality
\begin{equation*}
\jtypeeq{\Gamma}{\subst{f}{\ctxwk{\Gamma}{{\Delta}{\Delta}}}}{\ctxwk{\Gamma}{\Delta}}
\end{equation*}
so we obtain the type
\begin{equation*}
\jtype{\ctxext{\Gamma}{\ctxwk{\Gamma}{\Delta}}}
{\subst{f}{\ctxwk{\Gamma}{\idtypevar{\Delta}}}}
\end{equation*}
We also have $\jtype{\ctxext{\Gamma}{\ctxwk{\Gamma}{\Delta}}}{\eqtype{f}}$, so we
may ask for a term of type $\ctxwk{\eqtype{f}}{\subst{f}{\ctxwk{\Gamma}{\idtypevar{\Delta}}}}$
in context $\ctxext{\Gamma}{\ctxwk{\Gamma}{\Delta}}$ and we can ask ourselves the question
whether this is a trivial cofibration. This is the content of the following conjecture.

\begin{conj}
There is a term
\begin{equation*}
\jterm{\ctxext{{\Gamma}{\ctxwk{\Gamma}{\Delta}}}{\eqtype{f}}}{\ctxwk{\eqtype{f}}{\subst{f}{\ctxwk{\Gamma}{\idtypevar{\Delta}}}}}{\eqtoid{f}}
\end{equation*}
Moreover, $\eqtoid{f}$ is a trivial cofibration.
\end{conj}

\begin{proof}
We use the induction principle of $\eqtype{f}$ to construct $\eqtoid{f}$. Thus, we
have to find a term
\begin{equation*}
\jterm
  {\Gamma}
  {\subst{\refl{f}}{{f}{\ctxwk{\eqtype{f}}{\subst{f}{\ctxwk{\Gamma}{\idtypevar{\Delta}}}}}}}
  {\subst{\refl{f}}{{f}{\eqtoid{f}}}}
\end{equation*}
Note that we have the judgmental equalities
\begin{align*}
\subst{\refl{f}}{{f}{\ctxwk{\eqtype{f}}{\subst{f}{\ctxwk{\Gamma}{\idtypevar{\Delta}}}}}}
& \jdeq \subst{f}{{\ctxwk{{\Gamma}{\Delta}}{\refl{f}}}{\ctxwk{\eqtype{f}}{\subst{f}{\ctxwk{\Gamma}{\idtypevar{\Delta}}}}}} \\
& \jdeq \subst{f}{\ctxwk{\Gamma}{\subst{f}{\ctxwk{\Gamma}{\idtypevar{\Delta}}}}}
\end{align*}
where we find the term $\subst{f}{\ctxwk{\Gamma}{\subst{f}{\ctxwk{\Gamma}{\refl{\Delta}}}}}$.
\end{proof}

\begin{conj}
Suppose we have a type $\eqtype{x}$ for every term $\jterm{\Gamma}{A}{x}$. Then
every term $\jterm{\Gamma}{\ctxwk{\Gamma}{\Delta}}{f}$ is a trivial cofibration.
\end{conj}

\begin{proof}
Suppose that $Q$ is a type in context $\Delta$. We want to show that
\begin{equation*}
\inference{\jterm{\Gamma}{\subst{f}{\ctxwk{\Gamma}{Q}}}{g}}{\jterm{\Delta}{Q}{\tilde{g}}}
\end{equation*}
Let $g$ be a term of $\subst{f}{\ctxwk{\Gamma}{Q}}$ in context $\Gamma$. We get
the term
\begin{equation*}
\jterm{\Gamma}{\subst{\refl{f}}{{f}{\ctxwk{\eqtype{f}}{{\Gamma}{Q}}}}}{\subst{\refl{f}}{\ctxwk{\eqtype{f}}{g}}}
\end{equation*}
which is judgmentally equal to $g$; the types

For any term $\jterm{\Gamma}{\subst{f}{\ctxwk{\Gamma}{Q}}}{g}$ we get a term 
$\jterm{\Gamma}{\subst{\refl{x}}{{x}{P}}}{d}$.

$\jterm{\ctxext{{\Gamma}{\ctxwk{\Gamma}{\Delta}}}{\eqtype{f}}}{\ctxwk{\Gamma}{Q}}{\ind{\eqtype{f}}(d)}$.
\end{proof}

However, we could take other classes of terms, such as the projections
$\proj1:\ctxwk{\ctxext{\Gamma}{A}}{\Gamma}$. Let's see how that goes:

\begin{infarray}{c}
\inference{\jtype{\Gamma}{A}}{\jtype{\ctxext{{\Gamma}{A}}{\ctxwk{\ctxext{\Gamma}{A}}{\Gamma}}}{\eqtype{\proj1^A}}}\\
\inference{\jtype{\Gamma}{A}}{\jterm{\ctxext{\Gamma}{A}}{\subst{\proj1^A}{\eqtype{\proj1^A}}}{\refl{\proj1^A}}}
\end{infarray}

### 2013-11-10 article-models-tt.tex
added remark on stability under extension

@@
 
+\subsection{Stability under extension}
+We need to come up with a suitable notion of stability under extension, which
+is to be the type theoretical analogue of category theoretical stability under
+slicing. When we have done this, we have to show that the constructions we have
+so far are indeed stable under extension.
+
 \input{article-models-idtypes}

### 2013-11-10 article-models-tt.tex
Complete basic type theory except for stability under extension

@@
 in all three cases (contexts, types and terms).
-\begin{align}
+\bgroup\small
+\begin{align*}
 & \inference{\jctx{\Gamma}}{\jctxeq{\Gamma}{\Gamma}} 
 & & \inference{\jctxeq{\Gamma}{\Delta}}{\jctxeq{\Delta}{\Gamma}} 
-& & \inference{\jctxeq{\Gamma}{\Delta}\qquad\jctxeq{\Delta}{\Theta}}{\jctxeq{\Gamma}{\Theta}}\\
+& & \inference{\jctxeq{\Gamma}{\Delta}\qquad\jctxeq{\Delta}{\greek{E}}}{\jctxeq{\Gamma}{\greek{E}}}\\
 & \inference{\jtype{\Gamma}{A}}{\jtypeeq{\Gamma}{A}{A}} 
@@
 & & \inference{\jtermeq{\Gamma}{A}{x}{y}\qquad\jtermeq{\Gamma}{A}{y}{z}}{\jtermeq{\Gamma}{A}{x}{z}}
-\end{align}
+\end{align*}
+\egroup
 
-\begin{align}
+The following convertibility rules are responsible for the strictness
+of judgmental equality, which sets it apart from equivalences or identifications:
+
+\begin{align*}
 & \inference{\jctxeq{\Gamma}{\Delta}\qquad\jtype{\Gamma}{A}}{\jtype{\Delta}{A}}
@@
 & & \inference{\jtypeeq{\Gamma}{A}{B}\qquad\jtermeq{\Gamma}{A}{x}{y}}{\jtermeq{\Gamma}{B}{x}{y}}
-\end{align}
+\end{align*}
 
@@
 However, we need substitution to make this precise.
-\begin{infarray}{cc}
-\inference{\jtype{\Gamma}{A}}{\jctx{\ctxext{\Gamma}{A}}}
-& \inference{\jctxeq{\Gamma}{\Delta}\qquad\jtypeeq{\Gamma}{A}{B}}{\jctxeq{\ctxext{\Gamma}{A}}{\ctxext{\Delta}{B}}}\\
-\inference{\jtype{\ctxext{\Gamma}{A}}{P}}{\jtype{\Gamma}{\ctxext{A}{P}}}
-& \inference{\jtypeeq{\Gamma}{A}{B}\qquad\jtypeeq{\ctxext{\Gamma}{A}}{P}{Q}}{\jtypeeq{\Gamma}{\ctxext{A}{P}}{\ctxext{B}{Q}}}
-\end{infarray}
+\begin{align}
+& \inference{\jtype{\Gamma}{A}}{\jctx{\ctxext{\Gamma}{A}}}
+& & \inference{\jctxeq{\Gamma}{\Delta}\qquad\jtypeeq{\Gamma}{A}{B}}{\jctxeq{\ctxext{\Gamma}{A}}{\ctxext{\Delta}{B}}}\\
+& \inference{\jtype{\ctxext{\Gamma}{A}}{P}}{\jtype{\Gamma}{\ctxext{A}{P}}}
+& & \inference{\jtypeeq{\Gamma}{A}{B}\qquad\jtypeeq{\ctxext{\Gamma}{A}}{P}{Q}}{\jtypeeq{\Gamma}{\ctxext{A}{P}}{\ctxext{B}{Q}}}
+\end{align}
 
@@
 of $\ctxwk{\Gamma}{\Delta}$.
-\begin{infarray}{cc}
-\inference{\jctx{\Gamma}\qquad\jctx{\Delta}}{\jtype{\Gamma}{\ctxwk{\Gamma}{\Delta}}} 
-& \inference{\jctxeq{\Gamma}{\Gamma'}\qquad\jctxeq{\Delta}{\Delta'}}{\jtypeeq{\Gamma}{\ctxwk{\Gamma}{\Delta}}{\ctxwk{\Gamma'}{\Delta'}}}\\
-\inference{\jctx{\Gamma}\qquad\jtype{\Delta}{B}}{\jtype{\ctxext{\Gamma}{\ctxwk{\Gamma}{\Delta}}}{\ctxwk{\Gamma}{B}}} & \inference{\jctxeq{\Gamma}{\Gamma'}\qquad\jtypeeq{\Delta}{B}{B'}}{\jtypeeq{\ctxext{\Gamma}{\ctxwk{\Gamma}{\Delta}}}{\ctxwk{\Gamma}{B}}{\ctxwk{\Gamma'}{B'}}}\\
-\inference{\jctx{\Gamma}\qquad\jterm{\Delta}{B}{y}}{\jterm{\ctxext{\Gamma}{\ctxwk{\Gamma}{\Delta}}}{\ctxwk{\Gamma}{B}}{\ctxwk{\Gamma}{y}}} 
-& \inference{\jctxeq{\Gamma}{\Gamma'}\qquad\jtermeq{\Delta}{B}{y}{y'}}{\jtermeq{\ctxext{\Gamma}{\ctxwk{\Gamma}{\Delta}}}{\ctxwk{\Gamma}{B}}{\ctxwk{\Gamma}{y}}{\ctxwk{\Gamma'}{y'}}} 
-\end{infarray}
+
+In the following inference rules we assume $\jctx{\Gamma}$ and in the rules
+asserting a judgmental equality we furthermore assume $\jctxeq{\Gamma}{\Gamma'}$.
+\begin{align}
+& \inference{\jctx{\Delta}}{\jtype{\Gamma}{\ctxwk{\Gamma}{\Delta}}} 
+& & \inference{\jctxeq{\Delta}{\Delta'}}{\jtypeeq{\Gamma}{\ctxwk{\Gamma}{\Delta}}{\ctxwk{\Gamma'}{\Delta'}}}\\
+& \inference{\jtype{\Delta}{B}}{\jtype{\ctxext{\Gamma}{\ctxwk{\Gamma}{\Delta}}}{\ctxwk{\Gamma}{B}}}
+& & \inference{\jtypeeq{\Delta}{B}{B'}}{\jtypeeq{\ctxext{\Gamma}{\ctxwk{\Gamma}{\Delta}}}{\ctxwk{\Gamma}{B}}{\ctxwk{\Gamma'}{B'}}}\\
+& \inference{\jterm{\Delta}{B}{y}}{\jterm{\ctxext{\Gamma}{\ctxwk{\Gamma}{\Delta}}}{\ctxwk{\Gamma}{B}}{\ctxwk{\Gamma}{y}}} 
+& & \inference{\jtermeq{\Delta}{B}{y}{y'}}{\jtermeq{\ctxext{\Gamma}{\ctxwk{\Gamma}{\Delta}}}{\ctxwk{\Gamma}{B}}{\ctxwk{\Gamma}{y}}{\ctxwk{\Gamma'}{y'}}} 
+\end{align}
 
@@
  constant maps at $y$ from $A$ to $B$, for $y:B$.
-\begin{infarray}{cc}
-\inference{\jtype{\Gamma}{A}\qquad\jtype{\Gamma}{B}}{\jtype{\ctxext{\Gamma}{A}}{\ctxwk{A}{B}}}
-& \inference{\jtypeeq{\Gamma}{A}{A'}\qquad\jtypeeq{\Gamma}{B}{B'}}{\jtypeeq{\ctxext{\Gamma}{A}}{\ctxwk{A}{B}}{\ctxwk{A'}{B'}}}\\
-\inference{\jtype{\Gamma}{A}\qquad\jtype{\ctxext{\Gamma}{B}}{Q}}
+
+ In the following inference rules we assume that $\jtype{\Gamma}{A}$ and in the
+ rules asserting a judgmental equality we assume furthermore that 
+ $\jtypeeq{\Gamma}{A}{A'}$.
+\begin{align}
+& \inference{\jtype{\Gamma}{B}}{\jtype{\ctxext{\Gamma}{A}}{\ctxwk{A}{B}}}
+& & \inference{\jtypeeq{\Gamma}{B}{B'}}{\jtypeeq{\ctxext{\Gamma}{A}}{\ctxwk{A}{B}}{\ctxwk{A'}{B'}}}\\
+& \inference{\jtype{\ctxext{\Gamma}{B}}{Q}}
 {\jtype{\ctxext{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{Q}}}
-& \inference{\jtypeeq{\Gamma}{A}{A'}\qquad\jtypeeq{\ctxext{\Gamma}{B}}{Q}{Q'}}
+& & \inference{\jtypeeq{\ctxext{\Gamma}{B}}{Q}{Q'}}
 {\jtypeeq{\ctxext{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{Q}}{\ctxwk{A'}{Q'}}}\\
-\inference{\jtype{\Gamma}{A}\qquad\jterm{\Gamma}{B}{y}}{\jterm{\ctxext{\Gamma}{A}}{\ctxwk{A}{B}}{\ctxwk{A}{y}}}
-& \inference{\jtypeeq{\Gamma}{A}{A'}\qquad\jtermeq{\Gamma}{B}{y}{y'}}{\jtermeq{\ctxext{\Gamma}{A}}{\ctxwk{A}{B}}{\ctxwk{A}{y}}{\ctxwk{A'}{y'}}}
-\end{infarray}
+& \inference{\jterm{\Gamma}{B}{y}}{\jterm{\ctxext{\Gamma}{A}}{\ctxwk{A}{B}}{\ctxwk{A}{y}}}
+& & \inference{\jtermeq{\Gamma}{B}{y}{y'}}{\jtermeq{\ctxext{\Gamma}{A}}{\ctxwk{A}{B}}{\ctxwk{A}{y}}{\ctxwk{A'}{y'}}}\\
+& \inference{\jterm{\ctxext{\Gamma}{B}}{Q}{g}}{\jterm{\ctxext{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{Q}}{\ctxwk{A}{g}}}
+& & \inference{\jtermeq{\ctxext{\Gamma}{B}}{Q}{g}{g'}}
+{\jtermeq{\ctxext{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{Q}}{\ctxwk{A}{g}}{\ctxwk{A'}{g'}}}
+\end{align}
+
+\subsubsection{Weakening is compatible with extension}
+
+The following rules assert the compatibility of extension with weakening: for
+every context $\Gamma$ and every type $B$ in context $\Delta$ there is a
+judgmental equality $\ctxwk{\Gamma}{\ctxext{\Delta}{B}}\jdeq\ctxext{\ctxwk{\Gamma}{\Delta}}
+{\ctxwk{\Gamma}{B}}$. When thinking of terms of $\ctxwk{\Gamma}{\Delta}$ as
+functions from $\Gamma$ to $\Delta$, this looks awfully lot like the type
+theoretical axiom of choice, of which it seems absurd to ask to hold judgmentally.
+But the rule should be there because in ordinary Martin-L\"of type theory
+there is no syntax for weakening, i.e.~weakening is invisible. 
+And thus both $\ctxwk{\Gamma}{\ctxext{\Delta}{B}}$ 
+and $\ctxext{\ctxwk{\Gamma}{\Delta}}{\ctxwk{\Gamma}{B}}$ is the type
+$\ctxext{\Delta}{B}$ in context $\Gamma$.
+
+In the following inference rules we assume that $\jctx{\Gamma}$.
+
+\begin{align}
+& \inference
+  {\jtype{\Delta}{B}}
+  {\jtypeeq{\Gamma}{\ctxwk{\Gamma}{\ctxext{\Delta}{B}}}{\ctxext{\ctxwk{\Gamma}{\Delta}}{\ctxwk{\Gamma}{B}}}}\\
+& \inference
+  {\jtype{\Delta}{B}\qquad\jtype{\ctxext{\Delta}{B}}{Q}}
+  {\jtypeeq
+    {\ctxext{\Gamma}{\ctxwk{\Gamma}{\Delta}}}
+    {\ctxwk{\Gamma}{\ctxext{B}{Q}}}
+    {\ctxext{\ctxwk{\Gamma}{B}}{\ctxwk{\Gamma}{Q}}}}
+\end{align}
+
+In the following inference rules we assume that $\jtype{\Gamma}{A}$.
+
+\begin{align}
+& \inference
+  {\jtype{\ctxext{\Gamma}{B}}{Q}}
+  {\jtypeeq{\ctxext{\Gamma}{A}}{\ctxwk{A}{\ctxext{B}{Q}}}{\ctxext{\ctxwk{A}{B}}{\ctxwk{A}{Q}}}}\\
+& \inference
+  {\jtype{\ctxext{{\Gamma}{B}}{Q}}{R}}
+  {\jtypeeq
+    {\ctxext{{\Gamma}{A}}{\ctxwk{A}{B}}}
+    {\ctxwk{A}{\ctxext{Q}{R}}}
+    {\ctxext{\ctxwk{A}{Q}}{\ctxwk{A}{R}}}}
+\end{align}
 
@@
 we get two sets of inference rules. For weakening by contexts we get:
-
-\begin{infarray}{c}
-\inference{\jctx{\Gamma}\qquad\jctx{\Delta}\qquad\jctx{\greek{E}}}
+\begin{align}
+& \inference{\jctx{\Gamma}\qquad\jctx{\Delta}\qquad\jctx{\greek{E}}}
           {\jtypeeq{\ctxext{\Gamma}{\ctxwk{\Gamma}{\Delta}}}{\ctxwk{\Gamma}{{\Delta}{\greek{E}}}}
             {\ctxwk{{\Gamma}{\Delta}}{{\Gamma}{\greek{E}}}}}\\
-\inference{\jctx{\Gamma}\qquad\jctx{\Delta}\qquad\jtype{\greek{E}}{C}}
+& \inference{\jctx{\Gamma}\qquad\jctx{\Delta}\qquad\jtype{\greek{E}}{C}}
           {\jtypeeq{\ctxext{{\Gamma}{\ctxwk{\Gamma}{\Delta}}}{\ctxwk{{\Gamma}{\Delta}}{{\Gamma}{\greek{E}}}}}
@@
             {\ctxwk{{\Gamma}{\Delta}}{{\Gamma}{C}}}}\\
-\inference{\jctx{\Gamma}\qquad\jctx{\Delta}\qquad\jterm{\greek{E}}{C}{t}}
+& \inference{\jctx{\Gamma}\qquad\jctx{\Delta}\qquad\jterm{\greek{E}}{C}{t}}
           {\jtermeq{\ctxext{{\Gamma}{\ctxwk{\Gamma}{\Delta}}}{\ctxwk{{\Gamma}{\Delta}}{{\Gamma}{\greek{E}}}}}
             {\ctxwk{{\Gamma}{\Delta}}{{\Gamma}{C}}}{\ctxwk{\Gamma}{{\Delta}{t}}}{\ctxwk{{\Gamma}{\Delta}}{{\Gamma}{t}}}}
-\end{infarray}
-
+\end{align}
 For weakening by types we get:
-
-\begin{infarray}{c}
-\inference{\jtype{\Gamma}{A}\qquad\jtype{\Gamma}{B}\qquad\jtype{\Gamma}{C}}
+\begin{align}
+& \inference{\jtype{\Gamma}{A}\qquad\jtype{\Gamma}{B}\qquad\jtype{\Gamma}{C}}
           {\jtypeeq{\ctxext{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{{B}{C}}}
             {\ctxwk{{A}{B}}{{A}{C}}}}\\
-\inference{\jtype{\Gamma}{A}\qquad\jtype{\Gamma}{B}\qquad\jtype{\ctxext{\Gamma}{C}}{R}}
+& \inference{\jtype{\Gamma}{A}\qquad\jtype{\Gamma}{B}\qquad\jtype{\ctxext{\Gamma}{C}}{R}}
           {\jtypeeq{\ctxext{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{{A}{B}}{{A}{C}}}}
@@
             {\ctxwk{{A}{B}}{{A}{R}}}}\\
-\inference{\jtype{\Gamma}{A}\qquad\jtype{\Gamma}{B}\qquad\jterm{\Gamma}{R}{t}}
+& \inference{\jtype{\Gamma}{A}\qquad\jtype{\Gamma}{B}\qquad\jterm{\Gamma}{R}{t}}
           {\jtermeq{\ctxext{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{{A}{B}}{{A}{C}}}}
             {\ctxwk{{A}{B}}{{A}{R}}}{\ctxwk{A}{{B}{t}}}{\ctxwk{{A}{B}}{{A}{t}}}}
-\end{infarray}
-
-\subsubsection{A game of trees}
-We prove a little meta theorem about the type system we have so far. It is not
-really of importance to the development of the theory.
-
-Let's say
-that a binary tree of contexts is a pair $\pair{T,f}$ consisting of a binary 
-tree $T$ together with a function $f$ assigning to each leaf a context. The set
-of all binary trees of contexts is denoted by $B$. Such
-binary trees may be defined inductively: $(\unit,f):B$ for any $f:\unit\to ctx$
-and given any $(T_1,f_1)$ and $(T_2,f_2)$ in $B$ we have $(T_1,f_1)*(T_2,f_2)$
-in $B$.
-
-We simultaneously define
-the following two functions:
-\begin{align*}
-\trext & : B\to ctx\\
-\trwk_0 & : \prd{X,Y:B} typ(\trext(X))\\
-\trwk_1 & : \prd{X:B}{\Gamma~ctx} typ(\Gamma)\to typ(\ctxext{\trext(X)}{\trwk_0(X,\Gamma)})
-\end{align*}
-Both these functions are defined by induction on binary trees. We set
-\begin{align*}
-\trext((\unit,f)) & \defeq f(\unit)\\
-\trext((T_1,f_1)*(T_2,f_2)) & \defeq \ctxext{\trext((T_1,f_1))}
-{\trwk_0((T_1,f_1),\trext((T_2,f_2)))}
-\end{align*}
-and for any context $\Gamma$
-\begin{align*}
-\trwk_0((\unit,f),(\unit,g)) & \defeq \ctxwk{f(\unit)}{g(\unit)}\\
-\trwk_0((T_1,f_1)*(T_2,f_2),\Gamma) 
-  & \defeq \trwk_1((T_1,f_1),\trwk_0((T_2,f_2)),\Gamma))
-\end{align*}
-and for any type $A$ in context $\Gamma$
-\begin{align*}
-\trwk_1((\unit,f),A) & \defeq \ctxwk{f(\unit)}{A}\\
-\trwk_1((T_1,f_1)*(T_2,f_2),A) & \defeq
-  \trwk_1((T_1,f_1),\trwk_1((T_2,f_2)),A))
-\end{align*}
-
-\begin{lem}
-$\trwk_0((T,f),\Gamma)$ is a type in context $\trext((T,f))$ and
-$\trwk_1((T,f),A)$ is a type in context $\ctxext{\trext((T,f))}{\trwk_0((T,f),\Gamma)}$
-for any type $A$ in context $\Gamma$.
-\end{lem}
-
-\begin{proof}
-It is immediate that $\trwk_0((\unit,f),\Gamma)$ is a type in context $\trext((\unit,f))$
-and that $\trwk_1((\unit,f),A)$ is a type in context $\ctxext{\trext((\unit,f))}
-{\trwk_0((\unit,f),\Gamma)}$ for any type $A$ in context $\Gamma$.
-
-Suppose that $(T_1,f_1)$ and $(T_2,f_2)$ are binary trees of contexts such that
-$\trwk_0((T_i,f_i),\Gamma)$ is a type in context
-$\trext((T_i,f_i))$ for $i\jdeq 1$ and $i\jdeq 2$, and such that
-$\trwk_1((T_i,f_1),A)$ is a type in context $\ctxext{\trext((T_i,f_i))}{\trwk_0((T_i,f_i),\Gamma)}$
-for $i\jdeq 1$ and $i\jdeq 2$. Then
-\begin{equation*}
-\trwk_0((T_1,f_1)*(T_2,f_2),\Gamma) 
-\jdeq \trwk_1((T_1,f_1),\trwk_0((T_2,f_2),\Gamma))
-\end{equation*}
-is a type in context
-\begin{equation*}
-\ctxext{\trext((T_1,f_1))}{\trwk_0((T_1,f_1),\trext((T_2,f_2)))}
-\jdeq \trext((T_1,f_1)*(T_2,f_2)).
-\end{equation*}
-Also,
-\begin{equation*}
-\trwk_1((T_1,f_1)*(T_2,f_2),A) \jdeq \trwk_1((T_1,f_1),\trwk_1((T_2,f_2),A))
-\end{equation*}
-is a type in context
-\begin{equation*}
-\ctxext{\trext((T_1,f_1))}{\trwk_0((T_1,f_1),\ctxext{\trext((T_2,f_2))}{\trwk_0((T_2,f_2),\Gamma))}}
-\end{equation*}
-\end{proof}
+\end{align}
 
@@
 
-\begin{infarray}{cc}
-\inference{\jtype{\ctxext{\Gamma}{A}}{P}}{\jtype{\Gamma}{\subst{x}{P}}}
-& \inference{\jtypeeq{\ctxext{\Gamma}{A}}{P}{P'}}{\jtypeeq{\Gamma}{\subst{x}{P}}{\subst{x'}{P'}}}\\
-\inference{\jtype{\ctxext{{\Gamma}{A}}{P}}{Q}}{\jtype{\ctxext{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}}
-& \inference{\jtypeeq{\ctxext{{\Gamma}{A}}{P}}{Q}{Q'}}{\jtypeeq{\ctxext{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}{\subst{x'}{Q'}}}\\
-\inference{\jterm{\ctxext{\Gamma}{A}}{P}{f}}{\jterm{\Gamma}{\subst{x}{P}}{\subst{x}{f}}}
-& \inference{\jtermeq{\ctxext{\Gamma}{A}}{P}{f}{f'}}{\jtermeq{\Gamma}{\subst{x}{P}}{\subst{x}{f}}{\subst{x'}{f'}}}\\
-\inference{\jterm{\ctxext{{\Gamma}{A}}{P}}{Q}{g}}{\jterm{\ctxext{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}{\subst{x}{g}}}
-& \inference{\jtermeq{\ctxext{{\Gamma}{A}}{P}}{Q}{g}{g'}}{\jtermeq{\ctxext{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}{\subst{x}{g}}{\subst{x'}{g'}}}
-\end{infarray}
+\begin{align}
+& \inference{\jtype{\ctxext{\Gamma}{A}}{P}}{\jtype{\Gamma}{\subst{x}{P}}}
+& & \inference{\jtypeeq{\ctxext{\Gamma}{A}}{P}{P'}}{\jtypeeq{\Gamma}{\subst{x}{P}}{\subst{x'}{P'}}}\\
+& \inference{\jtype{\ctxext{{\Gamma}{A}}{P}}{Q}}{\jtype{\ctxext{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}}
+& & \inference{\jtypeeq{\ctxext{{\Gamma}{A}}{P}}{Q}{Q'}}{\jtypeeq{\ctxext{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}{\subst{x'}{Q'}}}\\
+& \inference{\jterm{\ctxext{\Gamma}{A}}{P}{f}}{\jterm{\Gamma}{\subst{x}{P}}{\subst{x}{f}}}
+& & \inference{\jtermeq{\ctxext{\Gamma}{A}}{P}{f}{f'}}{\jtermeq{\Gamma}{\subst{x}{P}}{\subst{x}{f}}{\subst{x'}{f'}}}\\
+& \inference{\jterm{\ctxext{{\Gamma}{A}}{P}}{Q}{g}}{\jterm{\ctxext{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}{\subst{x}{g}}}
+& &\inference{\jtermeq{\ctxext{{\Gamma}{A}}{P}}{Q}{g}{g'}}{\jtermeq{\ctxext{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}{\subst{x}{g}}{\subst{x'}{g'}}}
+\end{align}
+
+\subsubsection{Substitution is compatible with extension}
+Suppose $\jterm{\Gamma}{A}{x}$ in all of the following inference rules.
+\begin{align}
+& \inference{\jtype{\ctxext{{\Gamma}{A}}{P}}{Q}}
+  {\jtypeeq{\Gamma}{\subst{x}{\ctxext{P}{Q}}}{\ctxext{\subst{x}{P}}{\subst{x}{Q}}}}\\
+& \inference{\jtype{\ctxext{{{\Gamma}{A}}{P}}{Q}}{R}}
+  {\jtypeeq{\ctxext{\Gamma}{\subst{x}{P}}}{\subst{x}{\ctxext{Q}{R}}}{\ctxext{\subst{x}{Q}}{\subst{x}{R}}}}
+\end{align}
+
+\subsubsection{Substitution is compatible with weakening}
+The rules asserting the compatibility of substitution with weakening assert
+that the following diagram commutes for any $\jterm{\Gamma}{A}{x}$ and any
+$\jtype{\ctxext{\Gamma}{A}}{P}$:
+\begin{equation*}
+\begin{tikzcd}[column sep=huge]
+\jtype{\ctxext{\Gamma}{A}}{\blank} \ar{d}[swap]{Q\mapsto\subst{x}{Q}} \ar{r}{Q\mapsto\ctxwk{P}{Q}} & \jtype{\ctxext{{\Gamma}{A}}{P}}{\blank} \ar{d}{R\mapsto\subst{x}{R}}\\ 
+\jtype{\Gamma}{\blank} \ar{r}[swap]{B\mapsto\ctxwk{\subst{x}{P}}{B}} & \jtype{\ctxext{\Gamma}{\subst{x}{P}}}{\blank}
+\end{tikzcd}
+\end{equation*}
+In the following inference rules we assume that $\jterm{\Gamma}{A}{x}$ and
+$\jtype{\ctxext{\Gamma}{A}}{P}$:
+\begin{align}
+& \inference
+    {\jtype{\ctxext{\Gamma}{A}}{Q}}
+    {\jtypeeq{\ctxext{\Gamma}{\subst{x}{P}}}{\subst{x}{\ctxwk{P}{Q}}}{\ctxwk{\subst{x}{P}}{\subst{x}{Q}}}}\\
+& \inference
+    {\jtype{\ctxext{{\Gamma}{A}}{Q}}{R}}
+    {\jtypeeq{\ctxext{{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}}{\subst{x}{\ctxwk{P}{R}}}{\ctxwk{\subst{x}{P}}{\subst{x}{R}}}}\\
+& \inference
+    {\jterm{\ctxext{\Gamma}{A}}{Q}{g}}
+    {\jtermeq
+      {\ctxext{\Gamma}{\subst{x}{P}}}
+      {\subst{x}{\ctxwk{P}{Q}}}
+      {\subst{x}{\ctxwk{P}{g}}}
+      {\ctxwk{\subst{x}{P}}{\subst{x}{g}}}}\\
+& \inference
+    {\jterm{\ctxext{{\Gamma}{A}}{Q}}{R}{h}}
+    {\jtermeq
+      {\ctxext{{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}}
+      {\subst{x}{\ctxwk{P}{R}}}
+      {\subst{x}{\ctxwk{P}{h}}}
+      {\ctxwk{\subst{x}{P}}{\subst{x}{h}}}}
+\end{align}
 
@@
 
-\begin{infarray}{c}
-\inference{\jtype{\ctxext{{\Gamma}{A}}{P}}{Q}}
+\begin{align}
+&\inference{\jtype{\ctxext{{\Gamma}{A}}{P}}{Q}}
 {\jtypeeq{\Gamma}{\subst{x}{{f}{Q}}}{\subst{{x}{f}}{{x}{Q}}}}\\
-\inference{\jtype{\ctxext{{{\Gamma}{A}}{P}}{Q}}{R}}
+&\inference{\jtype{\ctxext{{{\Gamma}{A}}{P}}{Q}}{R}}
 {\jtypeeq{\ctxext{\Gamma}{\subst{x}{{f}{Q}}}}{\subst{x}{{f}{R}}}{\subst{{x}{f}}{{x}{R}}}}\\
-\inference{\jterm{\ctxext{{\Gamma}{A}}{P}}{Q}{g}}
+&\inference{\jterm{\ctxext{{\Gamma}{A}}{P}}{Q}{g}}
 {\jtermeq{\Gamma}{\subst{x}{{f}{Q}}}{\subst{x}{{f}{g}}}{\subst{{x}{f}}{{x}{g}}}}\\
-\inference{\jterm{\ctxext{{{\Gamma}{A}}{P}}{Q}}{R}{h}}
+&\inference{\jterm{\ctxext{{{\Gamma}{A}}{P}}{Q}}{R}{h}}
 {\jtermeq{\ctxext{\Gamma}{\subst{x}{{f}{Q}}}}{\subst{x}{{f}{R}}}{\subst{x}{{f}{h}}}{\subst{{x}{f}}{{x}{h}}}}
-\end{infarray}
+\end{align}
 
-\subsubsection{Substitution is compatible with extension}
-Suppose $\jterm{\Gamma}{A}{x}$ in all of the following inference rules.
-\begin{infarray}{c}
-\inference{\jtype{\ctxext{{\Gamma}{A}}{P}}{Q}}
-  {\jtypeeq{\Gamma}{\subst{x}{\ctxext{P}{Q}}}{\ctxext{\subst{x}{P}}{\subst{x}{Q}}}}\\
-\inference{\jtype{\ctxext{{{\Gamma}{A}}{P}}{Q}}{R}}
-  {\jtypeeq{\ctxext{\Gamma}{\subst{x}{P}}}{\subst{x}{\ctxext{Q}{R}}}{\ctxext{\subst{x}{Q}}{\subst{x}{R}}}}
-\end{infarray}
+\subsubsection{Weakening is compatible with substitution}
+We already have rules for the compatibility of substitution with weakening, but
+we still need the rules the other way around, asserting that there is a 
+judgmental equality $\ctxwk{A}{\subst{y}{Q}}\jdeq\subst{\ctxwk{A}{y}}{\ctxwk{A}{Q}}$
+together with all its variants.
 
-\subsubsection{Substitution is compatible with weakening}
-\begin{equation*}
-\inference{\jterm{\Gamma}{A}{x}\qquad\jtype{\ctxext{\Gamma}{A}}{P}\qquad\jtype{\ctxext{\Gamma}{A}}{Q}}
-          {\jtypeeq{\ctxext{\Gamma}{\subst{x}{P}}}{\subst{x}{\ctxwk{P}{Q}}}{\ctxwk{\subst{x}{P}}{\subst{x}{Q}}}}
-\end{equation*}
+In the following inference rules we assume that $\jtype{\Gamma}{A}$ and that
+$\jterm{\Gamma}{B}{y}$.
 
+\begin{align}
+& \inference{\jtype{\ctxext{\Gamma}{B}}{Q}}{\jtypeeq{\ctxext{\Gamma}{A}}{\ctxwk{A}{\subst{y}{Q}}}{\subst{\ctxwk{A}{y}}{\ctxwk{A}{Q}}}}\\
+& \inference{\jtype{\ctxext{{\Gamma}{B}}{Q}}{R}}{\jtypeeq{\ctxext{{\Gamma}{A}}{\ctxwk{A}{Q}}}{\ctxwk{A}{\subst{y}{R}}}{\subst{\ctxwk{A}{y}}{\ctxwk{A}{R}}}}\\
+& \inference{\jterm{\ctxext{\Gamma}{B}}{Q}{g}}{\jtermeq{\ctxext{\Gamma}{A}}{\ctxwk{A}{\subst{y}{Q}}}{\ctxwk{A}{\subst{y}{g}}}{\subst{\ctxwk{A}{y}}{\ctxwk{A}{g}}}}\\
+& \inference{\jterm{\ctxext{{\Gamma}{B}}{Q}}{R}{h}}{\jtermeq{\ctxext{{\Gamma}{A}}{\ctxwk{A}{\subst{y}{Q}}}}{\ctxwk{A}{\subst{y}{R}}}{\ctxwk{A}{\subst{y}{h}}}{\subst{\ctxwk{A}{y}}{\ctxwk{A}{h}}}}
+\end{align}
+
+\subsection{The defining property of weakening}
 The judgmental equalities we're about to describe assert that weakening followed
@@
 
-\begin{infarray}{c}
-\inference{\jtype{\Gamma}{A}\qquad\jtype{\Gamma}{B}\qquad\jterm{\Gamma}{A}{x}}{\jtypeeq{\Gamma}{\subst{x}{\ctxwk{A}{B}}}{B}}\\
-\inference{\jtype{\Gamma}{A}\qquad\jtype{\ctxext{\Gamma}{B}}{Q}\qquad\jterm{\Gamma}{A}{x}}{\jtypeeq{\ctxext{\Gamma}{B}}{\subst{x}{\ctxwk{A}{Q}}}{Q}}\\
-\inference{\jterm{\Gamma}{A}{x}\qquad\jterm{\Gamma}{B}{y}}{\jtermeq{\Gamma}{B}{\subst{x}{\ctxwk{A}{y}}}{y}}\\
-\inference{\jterm{\Gamma}{A}{x}\qquad\jterm{\ctxext{\Gamma}{B}}{Q}{g}}{\jtermeq{\ctxext{\Gamma}{B}}{Q}{\subst{x}{\ctxwk{A}{g}}}{g}}
-\end{infarray}
+\begin{align}
+& \inference{\jtype{\Gamma}{A}\qquad\jtype{\Gamma}{B}\qquad\jterm{\Gamma}{A}{x}}{\jtypeeq{\Gamma}{\subst{x}{\ctxwk{A}{B}}}{B}}\\
+& \inference{\jtype{\Gamma}{A}\qquad\jtype{\ctxext{\Gamma}{B}}{Q}\qquad\jterm{\Gamma}{A}{x}}{\jtypeeq{\ctxext{\Gamma}{B}}{\subst{x}{\ctxwk{A}{Q}}}{Q}}\\
+& \inference{\jterm{\Gamma}{A}{x}\qquad\jterm{\Gamma}{B}{y}}{\jtermeq{\Gamma}{B}{\subst{x}{\ctxwk{A}{y}}}{y}}\\
+& \inference{\jterm{\Gamma}{A}{x}\qquad\jterm{\ctxext{\Gamma}{B}}{Q}{g}}{\jtermeq{\ctxext{\Gamma}{B}}{Q}{\subst{x}{\ctxwk{A}{g}}}{g}}
+\end{align}
 
@@
 \end{equation*}
-The term $\jcomp{\Gamma}{f}{g}$ is the composition of $g$ with $f$.
+The term $\jcomp{\Gamma}{f}{g}$ is the composition of $g$ with $f$. We may call
+a term $\jhom{\Gamma}{\Delta}{f}$ a morphism from $\Gamma$ to $\Delta$.
 
@@
 
-\subsubsection{Extension is compatible with weakening}
-
-\begin{equation*}
-\jterm{\ctxext{\Gamma}{\ctxwk{\Gamma}{\ctxext{\Delta}{B}}}}{\ctxwk{{\Gamma}{\ctxext{\Delta}{B}}}{\ctxext{\ctxwk{\Gamma}{\Delta}}{\ctxwk{\Gamma}{B}}}}{ac}
-\end{equation*}
-
 \subsection{The identity functions}
@@
 so far are indeed stable under extension.
+
+We can make weakening stable under extension by the following
+process. Suppose we have a context $\Gamma$. For a context $\Delta$ we add the
+judgment
+\begin{equation*}
+\vdash \Delta~\tfwk(\Gamma)-admissible
+\end{equation*}
+for the assertion that $\Delta$ is allowed as a subject for the action of
+$\tfwk(\Gamma)$. If $\Delta$ is a $\tfwk(\Gamma)$ admissible context and
 
@@
 \end{itemize}
+
+\section{Binary trees}
+We prove a little meta theorem about the type system we have so far. It is not
+really of importance to the development of the theory.
+
+Let's say
+that a binary tree of contexts is a pair $\pair{T,f}$ consisting of a binary 
+tree $T$ together with a function $f$ assigning to each leaf a context. The set
+of all binary trees of contexts is denoted by $B$. Such
+binary trees may be defined inductively: $(\unit,f):B$ for any $f:\unit\to ctx$
+and given any $(T_1,f_1)$ and $(T_2,f_2)$ in $B$ we have $(T_1,f_1)*(T_2,f_2)$
+in $B$.
+
+We simultaneously define
+the following two functions:
+\begin{align*}
+\trext & : B\to ctx\\
+\trwk_0 & : \prd{X,Y:B} typ(\trext(X))\\
+\trwk_1 & : \prd{X:B}{\Gamma~ctx} typ(\Gamma)\to typ(\ctxext{\trext(X)}{\trwk_0(X,\Gamma)})
+\end{align*}
+Both these functions are defined by induction on binary trees. We set
+\begin{align*}
+\trext((\unit,f)) & \defeq f(\unit)\\
+\trext((T_1,f_1)*(T_2,f_2)) & \defeq \ctxext{\trext((T_1,f_1))}
+{\trwk_0((T_1,f_1),\trext((T_2,f_2)))}
+\end{align*}
+and for any context $\Gamma$
+\begin{align*}
+\trwk_0((\unit,f),(\unit,g)) & \defeq \ctxwk{f(\unit)}{g(\unit)}\\
+\trwk_0((T_1,f_1)*(T_2,f_2),\Gamma) 
+  & \defeq \trwk_1((T_1,f_1),\trwk_0((T_2,f_2)),\Gamma))
+\end{align*}
+and for any type $A$ in context $\Gamma$
+\begin{align*}
+\trwk_1((\unit,f),A) & \defeq \ctxwk{f(\unit)}{A}\\
+\trwk_1((T_1,f_1)*(T_2,f_2),A) & \defeq
+  \trwk_1((T_1,f_1),\trwk_1((T_2,f_2)),A))
+\end{align*}
+
+\begin{lem}
+$\trwk_0((T,f),\Gamma)$ is a type in context $\trext((T,f))$ and
+$\trwk_1((T,f),A)$ is a type in context $\ctxext{\trext((T,f))}{\trwk_0((T,f),\Gamma)}$
+for any type $A$ in context $\Gamma$.
+\end{lem}
+
+\begin{proof}
+It is immediate that $\trwk_0((\unit,f),\Gamma)$ is a type in context $\trext((\unit,f))$
+and that $\trwk_1((\unit,f),A)$ is a type in context $\ctxext{\trext((\unit,f))}
+{\trwk_0((\unit,f),\Gamma)}$ for any type $A$ in context $\Gamma$.
+
+Suppose that $(T_1,f_1)$ and $(T_2,f_2)$ are binary trees of contexts such that
+$\trwk_0((T_i,f_i),\Gamma)$ is a type in context
+$\trext((T_i,f_i))$ for $i\jdeq 1$ and $i\jdeq 2$, and such that
+$\trwk_1((T_i,f_1),A)$ is a type in context $\ctxext{\trext((T_i,f_i))}{\trwk_0((T_i,f_i),\Gamma)}$
+for $i\jdeq 1$ and $i\jdeq 2$. Then
+\begin{equation*}
+\trwk_0((T_1,f_1)*(T_2,f_2),\Gamma) 
+\jdeq \trwk_1((T_1,f_1),\trwk_0((T_2,f_2),\Gamma))
+\end{equation*}
+is a type in context
+\begin{equation*}
+\ctxext{\trext((T_1,f_1))}{\trwk_0((T_1,f_1),\trext((T_2,f_2)))}
+\jdeq \trext((T_1,f_1)*(T_2,f_2)).
+\end{equation*}
+Also,
+\begin{equation*}
+\trwk_1((T_1,f_1)*(T_2,f_2),A) \jdeq \trwk_1((T_1,f_1),\trwk_1((T_2,f_2),A))
+\end{equation*}
+is a type in context
+\begin{equation*}
+\ctxext{\trext((T_1,f_1))}{\trwk_0((T_1,f_1),\ctxext{\trext((T_2,f_2))}{\trwk_0((T_2,f_2),\Gamma))}}
+\end{equation*}
+\end{proof}

### 2013-11-11 article-models-defn.tex
removed commented parts

@@
 \end{align*}
-\begin{comment}
-the judgmental equality
-\begin{equation*}
-f_0(\ctxext{\Gamma}{A})\jdeq\ctxext{f_0(\Gamma)}\mftyp{f}{\Gamma}_0(A)
-\end{equation*}
-for every type $A$ in context $\Gamma$.
-\end{comment}
 \item preservation of weakening: 
@@
 \end{align*}
-\begin{comment}
-This gives the following three judgmental equalities:
-\begin{enumerate}
-\item the judgmental equality
-\begin{equation*}
-\mftyp{f}{\ctxext{\Gamma}{A}}_0(\ctxwk{A}{B})\jdeq\ctxwk{\mftyp{f}{\Gamma}_0(A)}{\mftyp{f}{\Gamma}_0(B)}
-\end{equation*}
-for every two types $A$ and $B$ in context $\Gamma$.
-\item the judgmental equality
-\begin{equation*}
-\mftyp{f}{\ctxext{{\Gamma}{A}}{\ctxwk{A}{B}}}\circ\tfwk^A\jdeq\tfwk^{\protect{\mftyp{f}{\Gamma}_0(A)}}\circ\mftyp{f}{\ctxext{\Gamma}{B}}
-\end{equation*}
-\item the judgmental equality
-\begin{equation*}
-\terms[f]{\ctxwk{A}{B}}{{A}{y}}\jdeq\ctxwk{\mftyp{f}\Gamma_0(A)}{(\terms[f]{B}(y))}
-\end{equation*}
-\end{enumerate}
-\end{comment}
 \item preservation of substitution: 
@@
 \end{equation*}
-\begin{comment}
-\begin{enumerate}
-\item the judgmental equality
-\begin{equation*}
-\mftyp{f}\Gamma_0(\subst{x}{P})\jdeq\subst{\terms[f]{A}(x)}{\mftyp{f}{\ctxext{\Gamma}{A}}_0(P)}
-\end{equation*}
-for every family $P$ over $A$ in context $\Gamma$ and every term $x:A$.
-\item the judgmental equality
-\begin{equation*}
-\mftyp{f}\Gamma_0(\subst{x}{Q})\jdeq\subst{\terms[f]{A}(x)}{\mftyp{f}{\ctxext{{\Gamma}{A}}{P}}_0(Q)}
-\end{equation*}
-for every family $Q$ over $P$ over $A$ in context $\Gamma$ and every term $x:A$.
-\end{enumerate}
-\end{comment}
 \end{enumerate}
@@
 should commute.
-\begin{comment}
-\item For $\Gamma:\ctx$, $\Gamma\vdash A:\type$, $\ctxext{\Gamma}{A}\vdash P
-:\type$ and $\ctxext{{\Gamma}{A}}{P}\vdash Q:\type$, $x:A$ and $u:P$
-we desire that $\typefont{nat{-}ev{-}subst}(Q,u,x):\id{Q[\iota^x][\tfev(u,x)]}{Q[u][x]}$.
-\item A function
-\begin{equation*}
-\tfcomp:\terms{\mprd{P}{Q}}\to\prd{x:\terms{A}}\terms{\mprd{\subst{x}{P}}{Q[\iota^x]}}
-\end{equation*}
-which could possibly replace $\tfev$ because it could be that $\tfev$ is
-definable in terms of $\tfcomp$. It should satisfy
-\begin{equation*}
-\id{\trans{\typefont{nat{-}ev{-}subst}(Q,u,x)}{\tfev(\tfcomp(f,x),\tfev(u,x))}}{\tfev(\lambda(\tfev(f,u)),x)},
-\end{equation*}
-which is a naturality condition. There should be an equivalence between
-$\terms{\mprd{A}{P}}$ and a type consisting of those terms of $\prd{x:\terms{A}}\terms{\subst{x}{P}}$ which satisfy
-a certain naturality condition. One could hope that by formulating this dependently
-enough, it is possible to write down such a naturality in one line.
-\end{comment}
 \end{enumerate}
 \endgroup
-
-\begin{comment}
-\begingroup
-\color{darkgreen}
-\subsubsection*{The problem with evaluation.}
-...or why I'm doing such strange things. It seems that evaluation is of a
-different nature. At first sight, one might expect that evaluation compares
-to $\lambda$-abstraction the same way that first and second projection compare
-to the pairing function $\pairr{\blank,\blank}$. However, this is not the case
-because evaluation cannot be seen as an inverse to $\lambda$-abstraction. The
-$\lambda$-abstraction is of type $\terms{P}\to\terms{\mprd{A}{P}}$ whereas
-evaluation is of type $\terms{\mprd{A}{P}}\to\prd{x:\terms{A}}\terms{\subst{x}{P}}$. There
-are a few things to note about $\tfev$:
-\begin{enumerate}
-\item There's no function back so in particular $\tfev$ is not an equivalence.
-My guess is that there should be a naturality condition playing a role.
-\item Evaluation tells how functions interact with substitution of terms.
-\end{enumerate}
-So it seems that evaluation is like a variable binder in the same way
-that quantifiers are binding variables. If I remember right, Voevodsky mentioned
-something like this at the Formal Topology conference in Ljubljana, though I
-forgot what exactly he said.
-\endgroup
-\end{comment}
-
-\begin{comment}
-\begingroup
-\color{gray}
-\subsubsection*{Old stuff below:}
-\begin{enumerate}
-\item a context morphism $\catid{\Gamma}:\ctxhom{\Gamma}{\Gamma}$ for every
-context $\Gamma$.
-\item a context morphism $f\circ g:\ctxhom{E}{\Gamma}$ for every two context
-morphisms $f:\ctxhom{\Delta}{\Gamma}$ and $g:\ctxhom{E}{\Delta}$. 
-\item paths $\id{f\circ\catid{\Delta}}{f}$ and $\id{\catid{\Gamma}\circ f}{f}$
-for every $f:\ctxhom{\Delta}{\Gamma}$.
-\item a path $\id{f\circ (g\circ h)}{(f\circ g)\circ h}$ for every
-$f:\ctxhom{\Delta}{\Gamma}$, $g:\ctxhom{E}{\Delta}$ and $h:\ctxhom{Z}{E}$. 
-\item a context morphism $\Gamma_!:\ctxhom{\Gamma}{\unit^\mfM}$ 
-for every context $\Gamma$.
-\item a path $\id{(\unit^\mfM)_!}{\catid{\unit^\mfM}}$.
-\item a path $\id{\Gamma_!\circ f}{\Delta_!}$ for every $f:\ctxhom{\Delta}{\Gamma}$. 
-\item a path $\id{A[\catid{\Gamma}]}{A}$ for every object $A$ in context $\Gamma$.
-\item a path $\id{A[f][g]}{A[f\circ g]}$ for every object $A$ in context $\Gamma$
-and every $f:\ctxhom{\Delta}{\Gamma}$ and $g:\ctxhom{E}{\Delta}$. 
-\end{enumerate}
-\endgroup
-\end{comment}
 
@@
 \end{desiderata}
-
-\begin{comment}
-\section{Continuity of functions}
-We wish to find a condition under which we have
-\begin{equation*}
-\eqv{\terms{\mprd{A}{P}}}{\sm{f:\prd{x:\terms{A}}\terms{\subst{x}{P}}}\text{``$f$ is continuous''}}.
-\end{equation*}
-for all families $P$ over $A$. 
-Needless to say, we first need to formulate what we mean by ``$f$ is continuous''.
-Thus, we need to find the family $\pairr{p,u,v}\mapsto \id{\trans{p}{u}}{v}$,
-which also requires us to define transportation.
-
-Let $P$ be a family over $A$ in context $\Gamma$. Then we can weaken
-$\ctxwk{A}{P}$, which is a family over $\ctxwk{A}{A}$ in context $\ctxext{\Gamma}{A}$.
-We also have the family $\ctxwk{\ctxwk{A}{A}}{P}$ over $\ctxwk{A}{A}$ in context
-$\ctx{\Gamma}{A}$ and thus we can form the family
-\begin{equation*}
-\ctxwk{\ctxwk{A}{A}}{P}\to\ctxwk{A}{P}.
-\end{equation*}
-This family plays the role of the family $x,y:A\vdash P(x)\to P(y):\type$.
-{\color{red}Explain this by performing the relevant substitution.}
-Transportation is a term of the family
-\begin{equation*}
-\idtypevar{A}\to\ctxwk{\ctxwk{A}{A}}{P}\to\ctxwk{A}{P}
-\end{equation*}
-obtained by identity elimination, as usual. Thus we obtain a term
-\begin{equation*}
-\mathcal{T}rans_{x,y}:\terms{\id{x}{y}\to\subst{x}{P}\to\subst{y}{P}}.
-\end{equation*}
-for every $x,y:A$.
-
-Now let $x,y:A$. Then we have the type $\subst{y}{P}$
-in context $\Gamma$ and the family $\ctxwk{\subst{y}{P}}{\subst{x}{P}}$
-over $\subst{y}{P}$ in context $\Gamma$. We could weaken $\subst{y}{P}$ subsequently by $A$, $\ctxwk{A}{A}$
-and $\idtypevar{A}$ to obtain the family $\ctxwk{\idtypevar{A}}{\ctxwk{\ctxwk{A}{A}}{\ctxwk{A}{\subst{y}{P}}}}$ over $\idtypevar{A}$ in context $\ctxext{{\Gamma}{A}}{\ctxwk{A}{A}}$. We will
-denote this family simply by $\ctxwk{\idtypevar{A}}{\subst{y}{P}}$. Likewise,
-we obtain the family $\ctxwk{\idtypevar{A}}{\ctxwk{\subst{y}{P}}{\subst{x}{P}}}$ over
-$\ctxwk{\idtypevar{A}}{\subst{y}{P}}$ in context $\ctxext{{{\Gamma}{A}}{\ctxwk{A}{A}}}{\idtypevar{A}}$ and we obtain the family $\ctxwk{\idtypevar{A}}{\idtypevar{\subst{y}{P}}}$
-over $\ctxwk{\idtypevar{A}}{\ctxwk{\subst{y}{P}}{\subst{y}{P}}}$. Now we
-have the family $\subst{\ctxwk{A}{\mathcal{T}rans_{x,y}}}{\ctxwk{\idtypevar{A}}{\idtypevar{\subst{y}{P}}}}$ over
-$\ctxwk{\idtypevar{A}}{\ctxwk{\subst{y}{P}}{\subst{x}{P}}}$, which is
-the desired family to state the continuity condition:
-
-\begin{defn}
-A function $f:\prd{x:\terms{A}}\terms{\subst{x}{P}}$ is said to be \emph{continuous}
-if there is a term of type
-\begin{equation*}
-\prd{x,y:\terms{A}}\terms{\subst{\ctxwk{A}{\mathcal{T}rans_{x,y}}}{\ctxwk{\idtypevar{A}}{\idtypevar{\subst{y}{P}}}}}.
-\end{equation*}
-\end{defn}
-
-\begin{lem}
-If all functions are ``continuous mappings of terms'' then types without terms
-are empty, i.e.\ we have
-\begin{equation*}
-(\terms{A}\to\emptyt)\to\terms{A\to\emptyt^\mfGraph}
-\end{equation*}
-\end{lem}
-
-\begin{proof}
-Any function from a type without terms is vacuously continuous.
-\end{proof}
-
-We will see that in the graph model functions are not just continuous mappings
-of terms, since there is a graph $\tilde{\emptyt}$ which has no terms but which
-differs from the empty graph nevertheless.
-
-\begin{desiderata}
-The following may relate to functions being continuous mappings of terms (it
-seems to be a very strong condition):
-\begin{enumerate}
-\item We have the adjunctions $\tfcolim\dashv\Delta\dashv\tflim$ in which the
-two monads are idempotent (we don't even have the adjunctions for $\mfGraph$).
-\item $\eqv{\terms{\Omega}}{\ctx}$
-\end{enumerate}
-We also need to check whether all functions in the reflexive graphs have this
-property (if not, ditch this section).
-\end{desiderata}
-\end{comment}
 

### 2013-11-11 article-models-constructors.tex
Moved universal property of extension away since it is inclear whether we want it

--- 2013-11-10 article-models-tt.tex
@@
-\section{Variable-free type theory -- the basic theory}
-We have seen in the example of $\UU$ that we will not need to refer to variables
-when we are manipulating contexts. In fact, it is more natural not to.
-This suggests that there is a presentation
-of type theory where contexts do not mention variables. The first small gain is that such a presentation
-would not be burdened with comments about variables being bounded or not, or fresh or
-not occuring at all.
+\section{Variable-free type theory -- other type constructors}
+We will now describe the rules for the type constructors that we don't assume
+in our type theory by default. The first of these are the dependent function
+types. We also discuss $\tfW$-types, suspensions, truncations. More general higher inductive
+types will have to wait until we have introduced models, for the models of the
+basic type theory because we will use them as index categories of the diagrams.
 
-Perhaps a bigger gain is that we obtain an algebraic theory for categories with
-fibrations and sections.
-
-\subsection{The basic judgments}
-The judgments we wish to make are the standad six judgments for ``$\Gamma$ is a context'', ``$A$ is
-a type in context $\Gamma$'' and ``$x$ is a term of $A$ in context $\Gamma$'' and the three equality
-judgments coming with them.
-
-\begin{align*}
-\jctx*{\Gamma} & \jctxeq*{\Gamma}{\Gamma'}\\
-\jtype*{\Gamma}{A} & \jtypeeq*{\Gamma}{A}{B}\\
-\jterm*{\Gamma}{A}{x} & \jtermeq*{\Gamma}{A}{x}{y}.
-\end{align*}
-
-In contrast with standard practise, we don't assume that there is an empty context. The idea is that
-if we interpret only the syntax of contexts, types and terms with extension, weakening and
-substitution, we do not have a full model of type theory but we still have a 
-structure resembling a category and categories need not have terminal objects. Therefore,
-the rules that follow will describe only how to manipulate contexts, types and terms.
-Of most of the inferences that we give there are two versions: one introducing an operation, the other asserting that the operation in question preserves judgmental equality.
-
-\subsection{The basic rules for judgmental equality}
-The rules for judgmental equality establish that it is an equivalence relation
-in all three cases (contexts, types and terms).
-\begin{align}
-& \inference{\jctx{\Gamma}}{\jctxeq{\Gamma}{\Gamma}} 
-& & \inference{\jctxeq{\Gamma}{\Delta}}{\jctxeq{\Delta}{\Gamma}} 
-& & \inference{\jctxeq{\Gamma}{\Delta}\qquad\jctxeq{\Delta}{\Theta}}{\jctxeq{\Gamma}{\Theta}}\\
-& \inference{\jtype{\Gamma}{A}}{\jtypeeq{\Gamma}{A}{A}} 
-& & \inference{\jtypeeq{\Gamma}{A}{B}}{\jtypeeq{\Gamma}{B}{A}}
-& & \inference{\jtypeeq{\Gamma}{A}{B}\qquad\jtypeeq{\Gamma}{B}{C}}{\jtypeeq{\Gamma}{A}{C}}\\
-& \inference{\jterm{\Gamma}{A}{x}}{\jtermeq{\Gamma}{A}{x}{x}}
-& & \inference{\jtermeq{\Gamma}{A}{x}{y}}{\jtermeq{\Gamma}{A}{y}{x}}
-& & \inference{\jtermeq{\Gamma}{A}{x}{y}\qquad\jtermeq{\Gamma}{A}{y}{z}}{\jtermeq{\Gamma}{A}{x}{z}}
-\end{align}
-
-\begin{align}
-& \inference{\jctxeq{\Gamma}{\Delta}\qquad\jtype{\Gamma}{A}}{\jtype{\Delta}{A}}
-& & \inference{\jctxeq{\Gamma}{\Delta}\qquad\jtypeeq{\Gamma}{A}{B}}{\jtypeeq{\Delta}{A}{B}}\\
-& \inference{\jctxeq{\Gamma}{\Delta}\qquad\jterm{\Gamma}{A}{x}}{\jterm{\Delta}{A}{x}}
-& & \inference{\jctxeq{\Gamma}{\Delta}\qquad\jtermeq{\Gamma}{A}{x}{y}}{\jtermeq{\Delta}{A}{x}{y}}\\
-& \inference{\jtypeeq{\Gamma}{A}{B}\qquad \jterm{\Gamma}{A}{x}}{\jterm{\Gamma}{B}{x}}
-& & \inference{\jtypeeq{\Gamma}{A}{B}\qquad\jtermeq{\Gamma}{A}{x}{y}}{\jtermeq{\Gamma}{B}{x}{y}}
-\end{align}
-
-\subsection{Extension}
-We introduce extension which not only extends a context $\Gamma$ and a type
-$A$ over it to a context $\ctxext{\Gamma}{A}$, but which also extends a type $A$
-in context $\Gamma$ and a family $P$ over it to a type $\ctxext{A}{P}$ in context
-$\Gamma$. We do this to ensure that all of type theory can be done in a context.
-For instance, we could say (1) that a context in context $\Gamma$ is the same thing
-as a type in context $\Gamma$; (2) When $A$ is a context in this sense, a type in
-context $A$ is the same thing as a family $P$ over $A$ and (3) when $P$ is a type
-in context $A$ in this sense, a term of $P$ keeps its original meaning.
-
-Note that by introducing extension on the level of types and families,
-we introduce $\Sigma$-types at a very early stage.
-However, we need substitution to make this precise.
-\begin{infarray}{cc}
-\inference{\jtype{\Gamma}{A}}{\jctx{\ctxext{\Gamma}{A}}}
-& \inference{\jctxeq{\Gamma}{\Delta}\qquad\jtypeeq{\Gamma}{A}{B}}{\jctxeq{\ctxext{\Gamma}{A}}{\ctxext{\Delta}{B}}}\\
-\inference{\jtype{\ctxext{\Gamma}{A}}{P}}{\jtype{\Gamma}{\ctxext{A}{P}}}
-& \inference{\jtypeeq{\Gamma}{A}{B}\qquad\jtypeeq{\ctxext{\Gamma}{A}}{P}{Q}}{\jtypeeq{\Gamma}{\ctxext{A}{P}}{\ctxext{B}{Q}}}
-\end{infarray}
-
-\subsection{Weakening}
-We first define weakening by a context $\Gamma$. Since weakening by $\Gamma$
-should in principle be a functor, it acts on contexts, types and terms alike.
-Note that it is because of the weakening $\ctxwk{\Gamma}{\Delta}$ of a context
-$\Delta$ by $\Gamma$ that we can speak of context morphisms from $\Gamma$ to $\Delta$: they are the terms
-of $\ctxwk{\Gamma}{\Delta}$.
-\begin{infarray}{cc}
-\inference{\jctx{\Gamma}\qquad\jctx{\Delta}}{\jtype{\Gamma}{\ctxwk{\Gamma}{\Delta}}} 
-& \inference{\jctxeq{\Gamma}{\Gamma'}\qquad\jctxeq{\Delta}{\Delta'}}{\jtypeeq{\Gamma}{\ctxwk{\Gamma}{\Delta}}{\ctxwk{\Gamma'}{\Delta'}}}\\
-\inference{\jctx{\Gamma}\qquad\jtype{\Delta}{B}}{\jtype{\ctxext{\Gamma}{\ctxwk{\Gamma}{\Delta}}}{\ctxwk{\Gamma}{B}}} & \inference{\jctxeq{\Gamma}{\Gamma'}\qquad\jtypeeq{\Delta}{B}{B'}}{\jtypeeq{\ctxext{\Gamma}{\ctxwk{\Gamma}{\Delta}}}{\ctxwk{\Gamma}{B}}{\ctxwk{\Gamma'}{B'}}}\\
-\inference{\jctx{\Gamma}\qquad\jterm{\Delta}{B}{y}}{\jterm{\ctxext{\Gamma}{\ctxwk{\Gamma}{\Delta}}}{\ctxwk{\Gamma}{B}}{\ctxwk{\Gamma}{y}}} 
-& \inference{\jctxeq{\Gamma}{\Gamma'}\qquad\jtermeq{\Delta}{B}{y}{y'}}{\jtermeq{\ctxext{\Gamma}{\ctxwk{\Gamma}{\Delta}}}{\ctxwk{\Gamma}{B}}{\ctxwk{\Gamma}{y}}{\ctxwk{\Gamma'}{y'}}} 
-\end{infarray}
-
-A weakening operation is also defined for types. When $A$ and $B$ are both types
- in context $\Gamma$, the weakened type $\ctxwk{A}{B}$ in context $\ctxext{\Gamma}{A}$
- is the family which `doesn't really depend on $A$'. The terms of $\ctxwk{A}{B}$
- are the functions from $A$ to $B$. Likewise, the terms $\ctxwk{A}{y}$ are the
- constant maps at $y$ from $A$ to $B$, for $y:B$.
-\begin{infarray}{cc}
-\inference{\jtype{\Gamma}{A}\qquad\jtype{\Gamma}{B}}{\jtype{\ctxext{\Gamma}{A}}{\ctxwk{A}{B}}}
-& \inference{\jtypeeq{\Gamma}{A}{A'}\qquad\jtypeeq{\Gamma}{B}{B'}}{\jtypeeq{\ctxext{\Gamma}{A}}{\ctxwk{A}{B}}{\ctxwk{A'}{B'}}}\\
-\inference{\jtype{\Gamma}{A}\qquad\jtype{\ctxext{\Gamma}{B}}{Q}}
-{\jtype{\ctxext{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{Q}}}
-& \inference{\jtypeeq{\Gamma}{A}{A'}\qquad\jtypeeq{\ctxext{\Gamma}{B}}{Q}{Q'}}
-{\jtypeeq{\ctxext{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{Q}}{\ctxwk{A'}{Q'}}}\\
-\inference{\jtype{\Gamma}{A}\qquad\jterm{\Gamma}{B}{y}}{\jterm{\ctxext{\Gamma}{A}}{\ctxwk{A}{B}}{\ctxwk{A}{y}}}
-& \inference{\jtypeeq{\Gamma}{A}{A'}\qquad\jtermeq{\Gamma}{B}{y}{y'}}{\jtermeq{\ctxext{\Gamma}{A}}{\ctxwk{A}{B}}{\ctxwk{A}{y}}{\ctxwk{A'}{y'}}}
-\end{infarray}
-
-\subsubsection{Weakening is compatible with itself}
-To conclude the rules for weakening, we state judgmental equality rules expressing
-that weakening is compatible with itself. These rules state that the following
-diagram commutes given any two contexts $\Gamma$ and $\Delta$:
-\begin{equation*}
-\begin{tikzcd}[column sep=huge]
-\jctx{\blank} \ar{r}{\greek{E}\mapsto\ctxwk{\Delta}{E}} \ar{d}[swap]{\greek{E}\mapsto\ctxwk{\Gamma}{E}} & \jtype{\Delta}{\blank} \ar{d}{B\mapsto\ctxwk{\Gamma}{B}}\\
-\jtype{\Gamma}{\blank} \ar{r}[swap]{A\mapsto\ctxwk{{\Gamma}{\Delta}}{A}} & \jtype{\ctxext{\Gamma}{\ctxwk{\Gamma}{\Delta}}}{\blank}
-\end{tikzcd}
-\end{equation*}
-There is also a version of this diagram in which all happens in a context. Thus,
-we get two sets of inference rules. For weakening by contexts we get:
-
-\begin{infarray}{c}
-\inference{\jctx{\Gamma}\qquad\jctx{\Delta}\qquad\jctx{\greek{E}}}
-          {\jtypeeq{\ctxext{\Gamma}{\ctxwk{\Gamma}{\Delta}}}{\ctxwk{\Gamma}{{\Delta}{\greek{E}}}}
-            {\ctxwk{{\Gamma}{\Delta}}{{\Gamma}{\greek{E}}}}}\\
-\inference{\jctx{\Gamma}\qquad\jctx{\Delta}\qquad\jtype{\greek{E}}{C}}
-          {\jtypeeq{\ctxext{{\Gamma}{\ctxwk{\Gamma}{\Delta}}}{\ctxwk{{\Gamma}{\Delta}}{{\Gamma}{\greek{E}}}}}
-            {\ctxwk{\Gamma}{{\Delta}{C}}}
-            {\ctxwk{{\Gamma}{\Delta}}{{\Gamma}{C}}}}\\
-\inference{\jctx{\Gamma}\qquad\jctx{\Delta}\qquad\jterm{\greek{E}}{C}{t}}
-          {\jtermeq{\ctxext{{\Gamma}{\ctxwk{\Gamma}{\Delta}}}{\ctxwk{{\Gamma}{\Delta}}{{\Gamma}{\greek{E}}}}}
-            {\ctxwk{{\Gamma}{\Delta}}{{\Gamma}{C}}}{\ctxwk{\Gamma}{{\Delta}{t}}}{\ctxwk{{\Gamma}{\Delta}}{{\Gamma}{t}}}}
-\end{infarray}
-
-For weakening by types we get:
-
-\begin{infarray}{c}
-\inference{\jtype{\Gamma}{A}\qquad\jtype{\Gamma}{B}\qquad\jtype{\Gamma}{C}}
-          {\jtypeeq{\ctxext{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{{B}{C}}}
-            {\ctxwk{{A}{B}}{{A}{C}}}}\\
-\inference{\jtype{\Gamma}{A}\qquad\jtype{\Gamma}{B}\qquad\jtype{\ctxext{\Gamma}{C}}{R}}
-          {\jtypeeq{\ctxext{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{{A}{B}}{{A}{C}}}}
-            {\ctxwk{A}{{B}{R}}}
-            {\ctxwk{{A}{B}}{{A}{R}}}}\\
-\inference{\jtype{\Gamma}{A}\qquad\jtype{\Gamma}{B}\qquad\jterm{\Gamma}{R}{t}}
-          {\jtermeq{\ctxext{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{{A}{B}}{{A}{C}}}}
-            {\ctxwk{{A}{B}}{{A}{R}}}{\ctxwk{A}{{B}{t}}}{\ctxwk{{A}{B}}{{A}{t}}}}
-\end{infarray}
-
-\subsubsection{A game of trees}
-We prove a little meta theorem about the type system we have so far. It is not
-really of importance to the development of the theory.
-
-Let's say
-that a binary tree of contexts is a pair $\pair{T,f}$ consisting of a binary 
-tree $T$ together with a function $f$ assigning to each leaf a context. The set
-of all binary trees of contexts is denoted by $B$. Such
-binary trees may be defined inductively: $(\unit,f):B$ for any $f:\unit\to ctx$
-and given any $(T_1,f_1)$ and $(T_2,f_2)$ in $B$ we have $(T_1,f_1)*(T_2,f_2)$
-in $B$.
-
-We simultaneously define
-the following two functions:
-\begin{align*}
-\trext & : B\to ctx\\
-\trwk_0 & : \prd{X,Y:B} typ(\trext(X))\\
-\trwk_1 & : \prd{X:B}{\Gamma~ctx} typ(\Gamma)\to typ(\ctxext{\trext(X)}{\trwk_0(X,\Gamma)})
-\end{align*}
-Both these functions are defined by induction on binary trees. We set
-\begin{align*}
-\trext((\unit,f)) & \defeq f(\unit)\\
-\trext((T_1,f_1)*(T_2,f_2)) & \defeq \ctxext{\trext((T_1,f_1))}
-{\trwk_0((T_1,f_1),\trext((T_2,f_2)))}
-\end{align*}
-and for any context $\Gamma$
-\begin{align*}
-\trwk_0((\unit,f),(\unit,g)) & \defeq \ctxwk{f(\unit)}{g(\unit)}\\
-\trwk_0((T_1,f_1)*(T_2,f_2),\Gamma) 
-  & \defeq \trwk_1((T_1,f_1),\trwk_0((T_2,f_2)),\Gamma))
-\end{align*}
-and for any type $A$ in context $\Gamma$
-\begin{align*}
-\trwk_1((\unit,f),A) & \defeq \ctxwk{f(\unit)}{A}\\
-\trwk_1((T_1,f_1)*(T_2,f_2),A) & \defeq
-  \trwk_1((T_1,f_1),\trwk_1((T_2,f_2)),A))
-\end{align*}
-
-\begin{lem}
-$\trwk_0((T,f),\Gamma)$ is a type in context $\trext((T,f))$ and
-$\trwk_1((T,f),A)$ is a type in context $\ctxext{\trext((T,f))}{\trwk_0((T,f),\Gamma)}$
-for any type $A$ in context $\Gamma$.
-\end{lem}
-
-\begin{proof}
-It is immediate that $\trwk_0((\unit,f),\Gamma)$ is a type in context $\trext((\unit,f))$
-and that $\trwk_1((\unit,f),A)$ is a type in context $\ctxext{\trext((\unit,f))}
-{\trwk_0((\unit,f),\Gamma)}$ for any type $A$ in context $\Gamma$.
-
-Suppose that $(T_1,f_1)$ and $(T_2,f_2)$ are binary trees of contexts such that
-$\trwk_0((T_i,f_i),\Gamma)$ is a type in context
-$\trext((T_i,f_i))$ for $i\jdeq 1$ and $i\jdeq 2$, and such that
-$\trwk_1((T_i,f_1),A)$ is a type in context $\ctxext{\trext((T_i,f_i))}{\trwk_0((T_i,f_i),\Gamma)}$
-for $i\jdeq 1$ and $i\jdeq 2$. Then
-\begin{equation*}
-\trwk_0((T_1,f_1)*(T_2,f_2),\Gamma) 
-\jdeq \trwk_1((T_1,f_1),\trwk_0((T_2,f_2),\Gamma))
-\end{equation*}
-is a type in context
-\begin{equation*}
-\ctxext{\trext((T_1,f_1))}{\trwk_0((T_1,f_1),\trext((T_2,f_2)))}
-\jdeq \trext((T_1,f_1)*(T_2,f_2)).
-\end{equation*}
-Also,
-\begin{equation*}
-\trwk_1((T_1,f_1)*(T_2,f_2),A) \jdeq \trwk_1((T_1,f_1),\trwk_1((T_2,f_2),A))
-\end{equation*}
-is a type in context
-\begin{equation*}
-\ctxext{\trext((T_1,f_1))}{\trwk_0((T_1,f_1),\ctxext{\trext((T_2,f_2))}{\trwk_0((T_2,f_2),\Gamma))}}
-\end{equation*}
-\end{proof}
-
-\subsection{Substitution}
-Given a family $P$ over $A$ and a term $x$ of $A$, substitution gives a way to
-consider the fiber $\subst{x}{P}$ of $P$ at $x$. Also, we get a way to evaluate
-terms $f$ of $P$ at $x$. This will give us ways to compose functions too. In
-this section, we shall first introduce the operations `substitution of a term $x$'
-for types, terms and families. Then we shall explain how substitution interacts
-with itself, extension and weakening.
-
-In the rules introducing the various substitutions we assume $\jterm{\Gamma}{A}{x}$;
-in the rules introducing the definitional equalities we assume $\jtermeq{\Gamma}{A}{x}{x'}$.
-
-\begin{infarray}{cc}
-\inference{\jtype{\ctxext{\Gamma}{A}}{P}}{\jtype{\Gamma}{\subst{x}{P}}}
-& \inference{\jtypeeq{\ctxext{\Gamma}{A}}{P}{P'}}{\jtypeeq{\Gamma}{\subst{x}{P}}{\subst{x'}{P'}}}\\
-\inference{\jtype{\ctxext{{\Gamma}{A}}{P}}{Q}}{\jtype{\ctxext{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}}
-& \inference{\jtypeeq{\ctxext{{\Gamma}{A}}{P}}{Q}{Q'}}{\jtypeeq{\ctxext{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}{\subst{x'}{Q'}}}\\
-\inference{\jterm{\ctxext{\Gamma}{A}}{P}{f}}{\jterm{\Gamma}{\subst{x}{P}}{\subst{x}{f}}}
-& \inference{\jtermeq{\ctxext{\Gamma}{A}}{P}{f}{f'}}{\jtermeq{\Gamma}{\subst{x}{P}}{\subst{x}{f}}{\subst{x'}{f'}}}\\
-\inference{\jterm{\ctxext{{\Gamma}{A}}{P}}{Q}{g}}{\jterm{\ctxext{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}{\subst{x}{g}}}
-& \inference{\jtermeq{\ctxext{{\Gamma}{A}}{P}}{Q}{g}{g'}}{\jtermeq{\ctxext{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}{\subst{x}{g}}{\subst{x'}{g'}}}
-\end{infarray}
-
-\subsubsection{Substitution is compatible with substitution}
-
-We require that substitution is compatible with itself, which is roughly the
-assertion that substitution is associative. However, we cannot just state that
-$\subst{x}{{f}{g}}\jdeq\subst{{x}{f}}{g}$ since the expression $\subst{{x}{f}}{g}$
-is not well-formed. The term $\subst{x}{f}$ can be substituted in (terms of) families over
-$\subst{x}{P}$; the term $\subst{x}{g}$ is such. Therefore, associativity of
-substitution takes the form $\subst{x}{{f}{g}}\jdeq\subst{{x}{f}}{{x}{g}}$.
-
-In the following inference rules we assume
-$\jterm{\Gamma}{A}{x}$ and $\jterm{\ctxext{\Gamma}{A}}{P}{f}$.
-
-\begin{infarray}{c}
-\inference{\jtype{\ctxext{{\Gamma}{A}}{P}}{Q}}
-{\jtypeeq{\Gamma}{\subst{x}{{f}{Q}}}{\subst{{x}{f}}{{x}{Q}}}}\\
-\inference{\jtype{\ctxext{{{\Gamma}{A}}{P}}{Q}}{R}}
-{\jtypeeq{\ctxext{\Gamma}{\subst{x}{{f}{Q}}}}{\subst{x}{{f}{R}}}{\subst{{x}{f}}{{x}{R}}}}\\
-\inference{\jterm{\ctxext{{\Gamma}{A}}{P}}{Q}{g}}
-{\jtermeq{\Gamma}{\subst{x}{{f}{Q}}}{\subst{x}{{f}{g}}}{\subst{{x}{f}}{{x}{g}}}}\\
-\inference{\jterm{\ctxext{{{\Gamma}{A}}{P}}{Q}}{R}{h}}
-{\jtermeq{\ctxext{\Gamma}{\subst{x}{{f}{Q}}}}{\subst{x}{{f}{R}}}{\subst{x}{{f}{h}}}{\subst{{x}{f}}{{x}{h}}}}
-\end{infarray}
-
-\subsubsection{Substitution is compatible with extension}
-Suppose $\jterm{\Gamma}{A}{x}$ in all of the following inference rules.
-\begin{infarray}{c}
-\inference{\jtype{\ctxext{{\Gamma}{A}}{P}}{Q}}
-  {\jtypeeq{\Gamma}{\subst{x}{\ctxext{P}{Q}}}{\ctxext{\subst{x}{P}}{\subst{x}{Q}}}}\\
-\inference{\jtype{\ctxext{{{\Gamma}{A}}{P}}{Q}}{R}}
-  {\jtypeeq{\ctxext{\Gamma}{\subst{x}{P}}}{\subst{x}{\ctxext{Q}{R}}}{\ctxext{\subst{x}{Q}}{\subst{x}{R}}}}
-\end{infarray}
-
-\subsubsection{Substitution is compatible with weakening}
-\begin{equation*}
-\inference{\jterm{\Gamma}{A}{x}\qquad\jtype{\ctxext{\Gamma}{A}}{P}\qquad\jtype{\ctxext{\Gamma}{A}}{Q}}
-          {\jtypeeq{\ctxext{\Gamma}{\subst{x}{P}}}{\subst{x}{\ctxwk{P}{Q}}}{\ctxwk{\subst{x}{P}}{\subst{x}{Q}}}}
-\end{equation*}
-
-The judgmental equalities we're about to describe assert that weakening followed
-by substitution leaves everything untouched. Thus, we get that each fiber
-$\subst{x}{\ctxwk{A}{B}}$ is just $B$, that $\ctxwk{A}{y}$ is the constant function
-mapping everything to $y:B$ and similar properties for families and terms thereof.
-
-\begin{infarray}{c}
-\inference{\jtype{\Gamma}{A}\qquad\jtype{\Gamma}{B}\qquad\jterm{\Gamma}{A}{x}}{\jtypeeq{\Gamma}{\subst{x}{\ctxwk{A}{B}}}{B}}\\
-\inference{\jtype{\Gamma}{A}\qquad\jtype{\ctxext{\Gamma}{B}}{Q}\qquad\jterm{\Gamma}{A}{x}}{\jtypeeq{\ctxext{\Gamma}{B}}{\subst{x}{\ctxwk{A}{Q}}}{Q}}\\
-\inference{\jterm{\Gamma}{A}{x}\qquad\jterm{\Gamma}{B}{y}}{\jtermeq{\Gamma}{B}{\subst{x}{\ctxwk{A}{y}}}{y}}\\
-\inference{\jterm{\Gamma}{A}{x}\qquad\jterm{\ctxext{\Gamma}{B}}{Q}{g}}{\jtermeq{\ctxext{\Gamma}{B}}{Q}{\subst{x}{\ctxwk{A}{g}}}{g}}
-\end{infarray}
-
-Using the rules of the compatibility of substitution with weakening and of the
-compatibility of weakening with itself, we see that we can show
-\begin{equation*}
-\jtypeeq{\Gamma}{\subst{f}{\ctxwk{\Gamma}{{\Delta}{\greek{E}}}}}{\ctxwk{\Gamma}{E}}
-\end{equation*}
-for any three contexts $\Gamma$, $\Delta$ and $\greek{E}$ and any $\jhom{\Gamma}{\Delta}{f}$.
-It follows that for $\jhom{\Delta}{\greek{E}}{g}$ we have
-\begin{equation*}
-\jhom{\Gamma}{\greek{E}}{\subst{f}{\ctxwk{\Gamma}{g}}}
-\end{equation*}
-The term $\jcomp{\Gamma}{f}{g}$ is the composition of $g$ with $f$.
-
-\subsection{The universal property of extension}
+\subsection{The universal property of dependent pair types}
+This section shouldn't be about extension anymore.
 
@@
 $\pair_P$ and $\ind{\tfext_A(P)}$ are invariant under judgmental equality.
-
-\subsubsection{Extension is compatible with weakening}
-
-\begin{equation*}
-\jterm{\ctxext{\Gamma}{\ctxwk{\Gamma}{\ctxext{\Delta}{B}}}}{\ctxwk{{\Gamma}{\ctxext{\Delta}{B}}}{\ctxext{\ctxwk{\Gamma}{\Delta}}{\ctxwk{\Gamma}{B}}}}{ac}
-\end{equation*}
-
-\subsection{The identity functions}
-Without a rule explicitly asserting the existence of an identity morphism we don't
-get one, hence we do that here. The identity morphism is a term which introduced
-in ordinary type theory via the variable rule. The variable rule is a bit more
-general: it asserts that
-\begin{equation*}
-\jterm{\Gamma,\,x_1:A_1,\ldots,\,x_n:A_n}{A_i}{x_i}
-\end{equation*}
-for every $1\leq i\leq n$. Thus, it establishes the projections. In our setting,
-we get the projections from the identity morphisms together with weakening. We
-already have weakening, so here it suffices to introduce the identity morphisms.
-They are determined by their behavior with respect to substitution combined with
-weakening.
-
-\begin{align}
-&\inference{\jctx{\Gamma}}{\jterm{\Gamma}{\ctxwk{\Gamma}{\Gamma}}{\idfunc[\Gamma]}}\label{ctxIdFuncForm}\\
-&\inference{\jtype{\Gamma}{A}}{\jtypeeq{\Gamma}{\subst{\idfunc[\Gamma]}{\ctxwk{\Gamma}{A}}}{A}}\label{ctxIdFuncPbInvType}\\
-&\inference{\jterm{\Gamma}{A}{x}}{\jtermeq{\Gamma}{A}{\subst{\idfunc[\Gamma]}{\ctxwk{\Gamma}{x}}}{x}}\label{ctxIdFuncPbInvTerm}\\
-&\inference{\jterm{\Delta}{\ctxwk{\Delta}{\Gamma}}{g}}{\jtermeq{\Delta}{\ctxwk{\Delta}{\Gamma}}{\subst{g}{\ctxwk{\Delta}{\idfunc[\Gamma]}}}{g}}\label{ctxIdFuncSubstInvTerm}
-\end{align}
-
-The identity morphisms for types are introduced analogously to the identity
-morphisms of contexts, the only difference being that we have a rule asserting
-that $\subst{x}{\idfunc[A]}\jdeq x$ for every term $x:A$ in context $\Gamma$.
-
-\begin{align}
-& \inference{\jtype{\Gamma}{A}}{\jterm{\ctxext{\Gamma}{A}}{\ctxwk{A}{A}}{\idfunc[A]}}\\
-& \inference{\jterm{\Gamma}{A}{x}}{\jtermeq{\Gamma}{A}{\subst{x}{\idfunc[A]}}{x}}\\
-& \inference{\jtype{\ctxext{\Gamma}{A}}{P}}{\jtypeeq{\ctxext{\Gamma}{A}}{\subst{\idfunc[A]}{\ctxwk{A}{P}}}{P}}\\
-& \inference{\jterm{\ctxext{\Gamma}{A}}{P}{f}}{\jtermeq{\ctxext{\Gamma}{A}}{P}{\subst{\idfunc[A]}{\ctxwk{A}{f}}}{f}} \\
-& \inference{\jterm{\ctxext{\Gamma}{B}}{\ctxwk{B}{A}}{g}}{\jtermeq{\ctxext{\Gamma}{B}}{\ctxwk{B}{A}}{\subst{g}{\ctxwk{B}{\idfunc[A]}}}{g}}
-\end{align}
-
-Using these rules we can derive that there is a judgmental equality $\jtypeeq{\ctxext{\Gamma}{A}}{\subst{x}{{\idfunc[A]}{B}}}
-{\subst{x}{{x}{B}}}$.
-
-\input{article-models-idtypes}
-
-\section{Variable-free type theory -- other type constructors}
-We will now describe the rules for the type constructors that we don't assume
-in our type theory by default. The first of these are the dependent function
-types. We also discuss $\tfW$-types, suspensions, truncations. More general higher inductive
-types will have to wait until we have introduced models, for the models of the
-basic type theory because we will use them as index categories of the diagrams.
 

### 2013-11-11 article-models-tt.tex
Moved universal property of extension away since it is inclear whether we want it

@@
 
-\subsection{The universal property of extension}
-
-Using weakening and substitution we are able to state the universal property
-for extension. It looks a bit more involved, since we cannot directly refer
-to the variables in the contexts. On the other hand, we can now plainly see
-were there were secretly weakenings going on.
-
-We begin with stating the universal property of the extension $\ctxext{\Gamma}{A}$.
-In these rules we assume we have $\jtype{\Gamma}{A}$ in the hypotheses.
-\begin{align}
-& \inference{}
-{\jterm{\ctxext{\Gamma}{A}}{\ctxwk{A}{{\Gamma}{\ctxext{\Gamma}{A}}}}{\pair_A}}\\
-& \inference{
-  \jtype{\ctxext{\Gamma}{A}}{P}
-  \qquad
-  \jterm{\ctxext{\Gamma}{A}}{\subst{\pair_A}{\ctxwk{A}{{\Gamma}{P}}}}{f}}
-  {\jterm{\ctxext{\Gamma}{A}}{P}{\ind{\tfext_\Gamma(A)}(f)}}\\
-& \inference{
-  \jtype{\ctxext{\Gamma}{A}}{P}
-  \qquad
-  \jterm{\ctxext{\Gamma}{A}}{\subst{\pair_A}{\ctxwk{A}{{\Gamma}{P}}}}{f}}
-  {\jtermeq{\ctxext{\Gamma}{A}}{\subst{\pair_A}{\ctxwk{A}{{\Gamma}{P}}}}{\subst{\pair_A}{\ctxwk{A}{{\Gamma}{\ind{\tfext_\Gamma(A)}(f)}}}}{f}}
-\end{align}
-
-Under the hypothesis that $\jtypeeq{\Gamma}{A}{A'}$
-we will also have the rules
-
-\begin{align}
-& \inference{}{\jtermeq{\ctxext{\Gamma}{A}}{\ctxwk{A}{{\Gamma}{\ctxext{\Gamma}{A}}}}{\pair_A}{\pair_{A'}}}\\
-& \inference{\jtype{\ctxext{\Gamma}{A}}{P}\qquad\jterm{\ctxext{\Gamma}{A}}{\subst{\pair_A}{\ctxwk{A}{{\Gamma}{P}}}}{f}}
-{\jtermeq{\ctxext{\Gamma}{A}}{P}{\ind{\tfext_\Gamma(A)}(f)}{\ind{\tfext_\Gamma(A')}(f)}}
-\end{align}
-
-Note that we don't need the notion of terms for contexts to state the universal
-property of context extension (which is a good thing, for we don't assume to have them).
-
-Next, we give the universal property of the extension $\ctxext{A}{P}$ in context
-$\Gamma$.
-In all of the following inference rules we assume that $\jtype{\ctxext{\Gamma}{A}}{P}$
-is among the hypothesis. The induction principle for extension consists of three
-inference rules:
-
-\begin{align}
-& \inference{}
-{\jterm{\ctxext{{\Gamma}{A}}{P}}{\ctxwk{P}{{A}{\ctxext{A}{P}}}}{\pair_P}}\\
-& \inference{
-  \jtype{\ctxext{\Gamma}{{A}{P}}}{Q}
-  \qquad
-  \jterm{\ctxext{{\Gamma}{A}}{P}}{\subst{\pair_P}{\ctxwk{P}{{A}{Q}}}}{f}}
-  {\jterm{\ctxext{\Gamma}{{A}{P}}}{Q}{\ind{\tfext_A(P)}(f)}}\\
-& \inference{
-  \jtype{\ctxext{\Gamma}{{A}{P}}}{Q}
-  \qquad
-  \jterm{\ctxext{{\Gamma}{A}}{P}}{\subst{\pair_P}{\ctxwk{P}{{A}{Q}}}}{f}}
-  {\jtermeq{\ctxext{{\Gamma}{A}}{P}}{\subst{\pair_P}{\ctxwk{P}{{A}{Q}}}}{\subst{\pair_P}{\ctxwk{P}{{A}{\ind{\tfext_A(P)}(f)}}}}{f}}
-\end{align}
-
-As with context extension, we shall require two more inference rules stating that
-$\pair_P$ and $\ind{\tfext_A(P)}$ are invariant under judgmental equality.
-
 \subsection{The identity functions}
@@
 $\tfwk(\Gamma)$. If $\Delta$ is a $\tfwk(\Gamma)$ admissible context and
-
-\input{article-models-idtypes}
-
-\section{Variable-free type theory -- other type constructors}
-We will now describe the rules for the type constructors that we don't assume
-in our type theory by default. The first of these are the dependent function
-types. We also discuss $\tfW$-types, suspensions, truncations. More general higher inductive
-types will have to wait until we have introduced models, for the models of the
-basic type theory because we will use them as index categories of the diagrams.
-
-\subsection{The unit type}
-Since we don't have a notion of terms of a context, we just say that the context
-$\unit$ is the terminal context.
-
-\begin{align}
-& \inference{}{\jctx{\unit}}\\
-& \inference{\jctx{\Gamma}}{\jhom{\Gamma}{\unit}{\tounit{\Gamma}}}\\
-& \inference{\jctx{\Gamma}\qquad\jhom{\Gamma}{\unit}{f}}{\jhomeq{\Gamma}{\unit}{f}{\tounit{\Gamma}}}
-\end{align}
-
-Note that we don't have to require that $\jhomeq{\Gamma}{\unit}{\tounit{\Gamma}}{\tounit{\Delta}}$
-whenever we have a judgmental equality $\jctxeq{\Gamma}{\Delta}$, since this already follows from the third rule.
-
-When the context $\unit$ is present, we may use the expression $\jtype{}{\Gamma}$
-as a shorthand for the judgment $\jtype{\unit}{\ctxwk{\unit}{\Gamma}}$. Likewise,
-we may use the expression $\jtermc{\Gamma}{i}$ as a shorthand
-for the judgment $\jhom{\unit}{\Gamma}{i}$. If we have a type $A$ in context
-$\Gamma$, we may use the expression $\jtype{}{\subst{i}{A}}$ to stand for
-the judgment $\jtype{\unit}{\subst{i}{\ctxwk{\unit}{A}}}$. We see that in every
-respect, contexts are types in the empty context.
-
-We have created a strictly terminal object $\unit$. This is not necessary when
-we're working in a context. We introduce the unit type $\unit_\Gamma$ in context
-$\Gamma$ in the familiar type theoretical way.
-
-\begin{align}
-& \inference{\jctx{\Gamma}}{\jtype{\Gamma}{\unit_\Gamma}}\\
-& \inference{\jctx{\Gamma}}{\jterm{\Gamma}{\unit_\Gamma}{\ttt_\Gamma}}\\
-& \inference{\jtype{\ctxext{\Gamma}{\unit_\Gamma}}{P}\qquad\jterm{\Gamma}{\subst{\ttt_\Gamma}{P}}{u}}
-          {\jterm{\ctxext{\Gamma}{\unit_\Gamma}}{P}{\ind{\unit_\Gamma}(u)}}\\
-& \inference{\jtype{\ctxext{\Gamma}{\unit_\Gamma}}{P}\qquad\jterm{\Gamma}{\subst{\ttt_\Gamma}{P}}{u}}
-          {\jtermeq{\Gamma}{\subst{\ttt_\Gamma}{P}}{\subst{\ttt_\Gamma}{\ind{\unit_\Gamma}(u)}}{u}}
-\end{align}
-
-\subsection{Subterminal types}
-The subterminal types we're about to present are strict, so they're more like \verb+Prop+
-in \Coq. We can define subterminal types in two ways: the first equalizes all elements
-of the subject type; the second is universal with the property that for every constant
-map factors through it.
-
-\subsubsection{Equalizing all terms}
-
-\subsubsection{Factorizing constant maps}
-Let $\jhom{\Gamma}{\Delta}{f}$. We can weaken $f$ by $\Gamma$ to obtain a term
-$\jterm{\ctxext{\Gamma}{\ctxwk{\Gamma}{\Gamma}}}{\ctxwk{\Gamma}{{\Gamma}{\Delta}}}{\ctxwk{\Gamma}{f}}$.
-This term is `like the function $\lam{x}{y}f(y)$'. Likewise, we can weaken $f$
-by $\ctxwk{{\Gamma}{\Gamma}}$ to obtain a term
-$\jterm{\ctxext{\Gamma}{\ctxwk{\Gamma}{\Gamma}}}{\ctxwk{\Gamma}{{\Gamma}{\Delta}}}{\ctxwk{{\Gamma}{\Gamma}}{f}}$,
-which is `like the function $\lam{x}{y}f(x)$'. Since we have
-the judgmental equality 
-$\jtypeeq{\ctxext{\Gamma}{\ctxwk{\Gamma}{\Gamma}}}{\ctxwk{{\Gamma}{\Gamma}}{{\Gamma}{\Delta}}}{\ctxwk{\Gamma}{{\Gamma}{\Delta}}}$ we can consider the judgmental equality between $\ctxwk{\Gamma}{f}$
-and $\ctxwk{{\Gamma}{\Gamma}}{f}$, which is exactly what we'll do in the definition
-of judgmentally constant.
-
-\begin{defn}
-A term $\jhom{\Gamma}{\Delta}{f}$ is said to be \emph{judgmentally constant} if
-the judgment
-\begin{equation*}
-\jtermeq{\ctxext{\Gamma}{\ctxwk{\Gamma}{\Gamma}}}{\ctxwk{{\Gamma}{\Gamma}}{{\Gamma}{\Delta}}}{\ctxwk{\Gamma}{f}}{\ctxwk{{\Gamma}{\Gamma}}{f}}
-\end{equation*}
-can be derived.
-\end{defn}
-
-\begin{defn}
-\begin{align}
-& \inference{\jctx{\Gamma}}{\jctx{\tau\Gamma}}\\
-& \inference{\jctx{\Gamma}}{\jhom{\Gamma}{\tau\Gamma}{t}}\\
-& \inference{\jctx{\Gamma}}{\jtermeq{\ctxext{\Gamma}{\ctxwk{\Gamma}{\Gamma}}}{\ctxwk{{\Gamma}{\Gamma}}{{\Gamma}{\Delta}}}{\ctxwk{\Gamma}{t}}{\ctxwk{{\Gamma}{\Gamma}}{t}}}\\
-& \inference{\jhom{\Gamma}{\Delta}{f}\qquad\jtermeq{\ctxext{\Gamma}{\ctxwk{\Gamma}{\Gamma}}}{\ctxwk{{\Gamma}{\Gamma}}{{\Gamma}{\Delta}}}{\ctxwk{\Gamma}{f}}{\ctxwk{{\Gamma}{\Gamma}}{f}}}{\jhom{\tau\Gamma}{\Delta}{\tilde{f}}}\\
-& \inference{\jhom{\Gamma}{\Delta}{f}\qquad\jtermeq{\ctxext{\Gamma}{\ctxwk{\Gamma}{\Gamma}}}{\ctxwk{{\Gamma}{\Gamma}}{{\Gamma}{\Delta}}}{\ctxwk{\Gamma}{f}}{\ctxwk{{\Gamma}{\Gamma}}{f}}}
-{\jtermeq{\Gamma}{\ctxwk{\Gamma}{\Delta}}{\jcomp{\Gamma}{t}{\tilde{f}}}{f}}
-\end{align}
-\end{defn}
-
-\begin{lem}
-Any term $\jhom{\unit}{\Gamma}{i}$ is judgmentally constant.
-\end{lem}
-
-\begin{proof}
-
-\end{proof}
-
-\subsection{Product types}
-
-\subsubsection{Products}
-\begin{align}
-& \inference{\jctx{\Gamma}\qquad\jctx{\Delta}}{\jctx{\product{\Gamma}{\Delta}}}\\
-& \inference{\jctx{\Gamma}\qquad\jctx{\Delta}}{\jhom{{\Gamma}{\Delta}}{\product{\Gamma}{\Delta}}{\pair}}
-\end{align}
-
-\subsubsection{Strict products}
-As with the unit type, we may use the categorical description of the product
-for our type theoretical definition. If we do that, we get strict products.
-
-\begin{align}
-& \inference{\jctx{\Gamma}\qquad\jctx{\Delta}}{\jctx{\product{\Gamma}{\Delta}}}\\
-& \inference{\jctx{\Gamma}\qquad\jctx{\Delta}}{\jhom{\product{\Gamma}{\Delta}}{\Gamma}{\proj1}}\\
-& \inference{\jctx{\Gamma}\qquad\jctx{\Delta}}{\jhom{\product{\Gamma}{\Delta}}{\Delta}{\proj2}}\\
-& \inference{\jhom{\greek{E}}{\Gamma}{f}\qquad\jhom{\greek{E}}{\Delta}{g}}
-          {\jhom{\greek{E}}{\product{\Gamma}{\Delta}}{\pairp{f,g}}}\\
-& \inference{\jhom{\greek{E}}{\Gamma}{f}\qquad\jhom{\greek{E}}{\Delta}{g}}
-          {\jhomeq{\greek{E}}{\Gamma}{\jcomp{\greek{E}}{\pairp{f,g}}{\proj1}}{f}}\\
-& \inference{\jhom{\greek{E}}{\Gamma}{f}\qquad\jhom{\greek{E}}{\Delta}{g}}
-          {\jhomeq{\greek{E}}{\Delta}{\jcomp{\greek{E}}{\pairp{f,g}}{\proj2}}{g}}
-\end{align}
-
-\begin{equation}
-\inference{\jhom{\greek{E}}{\product{\Gamma}{\Delta}}{h}
-           \qquad
-           \jhomeq{\greek{E}}{\Gamma}{\jcomp{E}{h}{\proj1}}{f}
-           \qquad
-           \jhomeq{\greek{E}}{\Delta}{\jcomp{E}{h}{\proj2}}{g}}
-          {\jhomeq{\greek{E}}{\product{\Gamma}{\Delta}}{h}{\pairp{f,g}}}
-\end{equation}
-
-\subsection{Equalizer types}
-Now that we have introduced a terminal object and products the categorical way,
-we may just continue and present (strict) equalizers and pullbacks too, just to
-see where we get. These notions are probably just not very useful in a univalent
-setting until we got a good computational interpretation.
-
-\begin{align}
-& \inference{\jhom{\Gamma}{\Delta}{f}\qquad\jhom{\Gamma}{\Delta}{g}}{\jctx{\jequalizer{\Gamma}{f}{g}}}\\
-& \inference{\jhom{\Gamma}{\Delta}{f}\qquad\jhom{\Gamma}{\Delta}{g}}{\jhom{\jequalizer{\Gamma}{f}{g}}{\Gamma}{\jequalizerin{f}{g}}}\\
-& \inference{\jhom{\greek{E}}{\Gamma}{h}\qquad\jhomeq{\greek{E}}{\Delta}{\jcomp{\greek{E}}{h}{f}}{\jcomp{\greek{E}}{h}{g}}}{\jhom{\greek{E}}{\jequalizer{\Gamma}{f}{g}}{\jequalizer{h}{f}{g}}}\\
-& \inference{
-  {\begin{array}{l}
-    \jhom{\greek{E}}{\Gamma}{h}\\
-    \jhomeq{\greek{E}}{\Delta}{\jcomp{\greek{E}}{h}{f}}{\jcomp{\greek{E}}{h}{g}}
-  \end{array}}
-  \qquad
-  {\begin{array}{l}
-    \jhom{\greek{E}}{\jequalizer{\Gamma}{f}{g}}{k}\\
-    \jhomeq{\greek{E}}{\Gamma}{\jcomp{\greek{E}}{k}{\jequalizerin{f}{g}}}{h}
-  \end{array}}
-}
-  {\jhom{\greek{E}}{\jequalizer{\Gamma}{f}{g}}{\jequalizer{h}{f}{g}}}
-\end{align}
-
-\subsection{Pullback types}
-
-\subsection{Dependent function types}
-\begin{align}
-& \inference{\jtype{\ctxext{\Gamma}{A}}{P}}{\jtype{\Gamma}{\mprd{A}{P}}}\\
-& \inference{\jterm{\ctxext{\Gamma}{A}}{P}{f}}{\jterm{\Gamma}{\mprd{A}{P}}{\lambda(f)}}\\
-& \inference{\jterm{\Gamma}{\mprd{A}{P}}{g}}{\jterm{\ctxext{\Gamma}{A}}{P}{\tfev(g)}}\\
-& \inference{\jterm{\ctxext{\Gamma}{A}}{P}{f}}{\jtermeq{\Gamma}{A}{\tfev(\lambda(f))}{f}}\\
-& \inference{\jtermeq{\ctxext{\Gamma}{A}}{P}{f}{f'}}{\jtermeq{\Gamma}{\mprd{A}{P}}{\lambda(f)}{\lambda(f')}}\\
-& \inference{\jtermeq{\Gamma}{\mprd{A}{P}}{g}{g'}}{\jtermeq{\ctxext{\Gamma}{A}}{P}{\tfev(g)}{\tfev(g')}}\\
-& \inference{\jtypeeq{\ctxext{\Gamma}{A}}{P}{P'}}{\jtypeeq{\Gamma}{\mprd{A}{P}}{\mprd{A}{P'}}}\\
-\end{align}
-
-With these rules we will not get the weak $\eta$-rule when identity types are present.
-So it might be better to state that $\lambda$ is a trivial cofibration.
-
-\subsection{Subterminal types}
-Let $\jhom{\Gamma}{\Delta}{f}$. Then we may also consider the type $\ctxwk{{\Gamma}{\Gamma}}{{\Gamma}{\Delta}}$
-in context $\ctxwk{\Gamma}{\Gamma}$ and we have the terms
-\begin{align*}
-& \jhom{\ctxext{\Gamma}{\ctxwk{\Gamma}{\Gamma}}}{\ctxwk{\Gamma}{A}}{\ctxwk{\Gamma}{f}}\\
-& \jhom{\ctxext{\Gamma}{\ctxwk{\Gamma}{\Gamma}}}{\ctxwk{\Gamma}{A}}{\ctxwk{{\Gamma}{\Gamma}}{f}}
-\end{align*}
-
-\subsection{The empty type}
-
-\subsection{Coproduct types}
-
-\subsection{The natural numbers}
-
-\section{A categorical explanation of the variable-free type theory}
-
-The categorical interpretation of type theory is a category $\catfont{C}$ of
-contexts, for each object $\Gamma$ of $\catfont{C}$ a class $\catfont{type}(\Gamma)$
-of morphisms into $\Gamma$ and for each $A\in\catfont{type}(\Gamma)$ a category
-$\catfont{term}(A)$ of sections of $A$.
-
-\begin{itemize}
-\item The context extension $\ctxext{\Gamma}{A}$ is the domain of $A$.
-\item There is a choice of $\Gamma\times\Delta$ for each $\Gamma$ and $\Delta$
-      for which the first projection $\pi_1:\Gamma\times\Delta\to\Gamma$ is in $\catfont{type}(\Gamma)$;
-      this is the weakening $\ctxwk{\Gamma}{\Delta}$.
-\item For each $A\in\catfont{type}(\Gamma)$ and $x:A$ there is a choice of
-      $\pi_1:\Gamma\times_A P\to\Gamma$ which is in $\catfont{type}(\Gamma)$.
-\end{itemize}
-
-\section{Binary trees}
-We prove a little meta theorem about the type system we have so far. It is not
-really of importance to the development of the theory.
-
-Let's say
-that a binary tree of contexts is a pair $\pair{T,f}$ consisting of a binary 
-tree $T$ together with a function $f$ assigning to each leaf a context. The set
-of all binary trees of contexts is denoted by $B$. Such
-binary trees may be defined inductively: $(\unit,f):B$ for any $f:\unit\to ctx$
-and given any $(T_1,f_1)$ and $(T_2,f_2)$ in $B$ we have $(T_1,f_1)*(T_2,f_2)$
-in $B$.
-
-We simultaneously define
-the following two functions:
-\begin{align*}
-\trext & : B\to ctx\\
-\trwk_0 & : \prd{X,Y:B} typ(\trext(X))\\
-\trwk_1 & : \prd{X:B}{\Gamma~ctx} typ(\Gamma)\to typ(\ctxext{\trext(X)}{\trwk_0(X,\Gamma)})
-\end{align*}
-Both these functions are defined by induction on binary trees. We set
-\begin{align*}
-\trext((\unit,f)) & \defeq f(\unit)\\
-\trext((T_1,f_1)*(T_2,f_2)) & \defeq \ctxext{\trext((T_1,f_1))}
-{\trwk_0((T_1,f_1),\trext((T_2,f_2)))}
-\end{align*}
-and for any context $\Gamma$
-\begin{align*}
-\trwk_0((\unit,f),(\unit,g)) & \defeq \ctxwk{f(\unit)}{g(\unit)}\\
-\trwk_0((T_1,f_1)*(T_2,f_2),\Gamma) 
-  & \defeq \trwk_1((T_1,f_1),\trwk_0((T_2,f_2)),\Gamma))
-\end{align*}
-and for any type $A$ in context $\Gamma$
-\begin{align*}
-\trwk_1((\unit,f),A) & \defeq \ctxwk{f(\unit)}{A}\\
-\trwk_1((T_1,f_1)*(T_2,f_2),A) & \defeq
-  \trwk_1((T_1,f_1),\trwk_1((T_2,f_2)),A))
-\end{align*}
-
-\begin{lem}
-$\trwk_0((T,f),\Gamma)$ is a type in context $\trext((T,f))$ and
-$\trwk_1((T,f),A)$ is a type in context $\ctxext{\trext((T,f))}{\trwk_0((T,f),\Gamma)}$
-for any type $A$ in context $\Gamma$.
-\end{lem}
-
-\begin{proof}
-It is immediate that $\trwk_0((\unit,f),\Gamma)$ is a type in context $\trext((\unit,f))$
-and that $\trwk_1((\unit,f),A)$ is a type in context $\ctxext{\trext((\unit,f))}
-{\trwk_0((\unit,f),\Gamma)}$ for any type $A$ in context $\Gamma$.
-
-Suppose that $(T_1,f_1)$ and $(T_2,f_2)$ are binary trees of contexts such that
-$\trwk_0((T_i,f_i),\Gamma)$ is a type in context
-$\trext((T_i,f_i))$ for $i\jdeq 1$ and $i\jdeq 2$, and such that
-$\trwk_1((T_i,f_1),A)$ is a type in context $\ctxext{\trext((T_i,f_i))}{\trwk_0((T_i,f_i),\Gamma)}$
-for $i\jdeq 1$ and $i\jdeq 2$. Then
-\begin{equation*}
-\trwk_0((T_1,f_1)*(T_2,f_2),\Gamma) 
-\jdeq \trwk_1((T_1,f_1),\trwk_0((T_2,f_2),\Gamma))
-\end{equation*}
-is a type in context
-\begin{equation*}
-\ctxext{\trext((T_1,f_1))}{\trwk_0((T_1,f_1),\trext((T_2,f_2)))}
-\jdeq \trext((T_1,f_1)*(T_2,f_2)).
-\end{equation*}
-Also,
-\begin{equation*}
-\trwk_1((T_1,f_1)*(T_2,f_2),A) \jdeq \trwk_1((T_1,f_1),\trwk_1((T_2,f_2),A))
-\end{equation*}
-is a type in context
-\begin{equation*}
-\ctxext{\trext((T_1,f_1))}{\trwk_0((T_1,f_1),\ctxext{\trext((T_2,f_2))}{\trwk_0((T_2,f_2),\Gamma))}}
-\end{equation*}
-\end{proof}

### 2013-11-11 article-models-tt.tex
ok I did the evil thing; extension is now judgmentally associative

@@
 & & \inference{\jtypeeq{\Gamma}{A}{B}\qquad\jtypeeq{\ctxext{\Gamma}{A}}{P}{Q}}{\jtypeeq{\Gamma}{\ctxext{A}{P}}{\ctxext{B}{Q}}}
+\end{align}
+
+\subsubsection{Extension is compatible with itself}
+The inference rules asserting that extension is compatible with itself assert
+that contexts are unstructured lists of type declarations. This rule is
+unavoidable if we want that for a type $A$ in context $\Gamma$, a type
+in context $A$ is the same thing as a type in context $\ctxext{\Gamma}{A}$.
+
+\begin{align}
+& \inference{\jtype{\Gamma}{A}\qquad\jtype{\ctxext{\Gamma}{A}}{P}}
+  {\jctxeq{\ctxext{{\Gamma}{A}}{P}}{\ctxext{\Gamma}{{A}{P}}}}\\
+& \inference{\jtype{\ctxext{\Gamma}{A}}{P}\qquad\jtype{\ctxext{{\Gamma}{A}}{P}}{Q}}
+  {\jtypeeq{\Gamma}{\ctxext{{A}{P}}{Q}}{\ctxext{A}{{P}{Q}}}}
 \end{align}

### 2013-12-09 article-models-intro.tex
working on the introduction

@@
 \section{Introduction}
-This project has its origins in the beginning of 2013, when I proved a version
+The project I propose here has its origins in the beginning of 2013, when I proved a version
 of the descent theorem for homotopy colimits in type theory while I was working
@@
 several of the basic type theoretical operations to give an efficient approach
-to the descent theorem. However it was not an issue to describe the graph model
-and the sense in which it was a model, the presentation of it was unsystematic
-and in that sense unsatisfactory too.
+to the descent theorem. Although it was not an issue to describe the graph model
+and the sense in which it models the type constructors, 
+not all of type theory is interpreted very well. To start with, context
+extension isn't strictly associative if dependent pair types are used for
+its interpretation. Also, if $A$ and $B$ are families of graphs
+over a graph $\Gamma$, then $B$ isn't also a family of graphs over the extended
+graph $\ctxext{\Gamma}{A}$ (interfering with a good interpretation of weakening).
+There are no issues with the interpretation of substitution though. Actually
+it is not so straightforward to describe what it models!
 
-From the experience of working with the graph model it became appearent that the
-answer we had of the question of what it models, i.e.~Martin L\"of type theory
-with the univalence axiom, could have been different if we were to describe a
-type theory based on the mechanism on which the model was built. The particular
-featueres of the graph model were that there was a type of graphs resembling a category
-of contexts, for each context a notion of what it means to be a family of graphs
-and a term thereof; there are the operations of context extension, weakening and
-substitution. In other words, the graph model is an instance of type theretical
-categories with families, an invention of Peter Dybjer.
-
-As we know however, the type theoretical description of a category as given in
-chapter 11 of \cite{TheBook} is restricted to categories which have a $1$-type
-as the type of objects and for every two objects a $0$-type of morphisms.
-Although categories with families are like that too, these are not very natural
-restrictions when one is concerned about internal models of type theory. It
-seems we're in the need of a higher categorical formulation of what an internal
-model, or a category with family, is. But type theory has the tendency to take
-care of it's own higher categorical structure and one could hope that type
-theory is an elementary theory of the category with families of category with
-families.
+The graph model
+shows three operations: extension, weakening and substitution. All of them
+can be defined such that they act not only on contexts, but also on families
+and terms. Moreover, they are all compatible with each other. In fact, along
+with asserting the existence of identity morphisms (which is done in type
+theory), it seems
+that this structure is enough to establish some of the structure of a category.
+We might even go so far as to boldly assert that
+\begin{quote}
+\emph{Category theory is dependent type theory without type constructors.}
+\end{quote}
+It is a part of this proposal to test this hypothesis, with the note that by
+`category theory' we mean a theory of weak higher categories which does not
+have any of the truncatedness restrictions -- in particular not those appearing
+in the theory of AKS-categories as
+described in \cite{TheBook} in chapter 11. We note that even Dybjers theory
+of internal categories with families has such restrictions: he relies on setoids.
+It needs not much arguing that these are very unnatural from a univalent point
+of view.
 

### 2013-12-09 article-models-tt.tex
working on the introduction

@@
 
-\subsection{The identity functions}
+\subsection{Terms in the basic type theory}
+
+\subsubsection{Identity functions}
 Without a rule explicitly asserting the existence of an identity morphism we don't
@@
 
-\subsection{Stability under extension}
-We need to come up with a suitable notion of stability under extension, which
-is to be the type theoretical analogue of category theoretical stability under
-slicing. When we have done this, we have to show that the constructions we have
-so far are indeed stable under extension.
+The identity functions in combination with weakening give rise to the projections.
+We define
 
-We can make weakening stable under extension by the following
-process. Suppose we have a context $\Gamma$. For a context $\Delta$ we add the
-judgment
 \begin{equation*}
-\vdash \Delta~\tfwk(\Gamma)-admissible
+\proj A \defeq \ctxwk{A}{\idfunc[\Gamma]}:\ctxwk{A}{{\Gamma}{\Gamma}}.
 \end{equation*}
-for the assertion that $\Delta$ is allowed as a subject for the action of
-$\tfwk(\Gamma)$. If $\Delta$ is a $\tfwk(\Gamma)$ admissible context and
+
+\subsubsection{Inclusions}
+In ordinary type theory, there is a rule
+\begin{equation*}
+\inference{\jterm{\Gamma}{A}{x}\qquad\jtype{\Gamma,\,a:A}{P(a)}}
+  {\Gamma,\,u:P(x/a)\vdash x:A, u:P(x/a)}
+\end{equation*}
+This rule gives the inclusion of the fiber $\subst{x}{P}$ into $\ctxext{A}{P}$.
+
+Thus, we require
+\begin{equation}
+\inference{\jterm{\Gamma}{A}{x}\qquad\jtype{\ctxext{\Gamma}{A}}{P}}
+  {\jterm{\ctxext{\Gamma}{\subst{x}{P}}}{\ctxwk{\subst{x}{P}}{\ctxext{A}{P}}}{\typefont{in}^P_x}}
+\end{equation}
+
+The terms $\typefont{in}^P_x$ have the following behavior:
+
+\begin{align}
+& \inference
+    {\jtype{\ctxext{{\Gamma}{A}}{P}}{Q}}
+    {\jtypeeq{\ctxext{\Gamma}{\subst{x}{P}}}{\subst{\typefont{in}^P_x}{\ctxwk{\subst{x}{P}}{Q}}}{\subst{x}{Q}}}\\
+& \inference
+    {\jterm{\ctxext{{\Gamma}{A}}{P}}{Q}{g}}
+    {\jtermeq{\ctxext{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}{\subst{\typefont{in}^P_x}{\ctxwk{\subst{x}{P}}{g}}}{\subst{x}{g}}}
+\end{align}

### 2014-01-08 article-models-eg.tex
moved some stuff around
working on proposal

@@
-\section{Empirical evidence of what an internal model should be}
+\section{Empirical evidence of what an internal model should be}\label{eg}
 
@@
 \begin{defn}
-A \emph{graph $A$ in the context $\Gamma$} is a pair $\pairr{A_0,A_1}$ consisting
+A \emph{family $A$ of graphs in the context $\Gamma$} is a pair $\pairr{A_0,A_1}$ consisting
 of 
@@
 
+\begin{defn}
+Suppose that $\Gamma\vdash A:\mfGraph$. A term $x$ of $A$ consists of a pair
+$\pairr{x_0,x_1}$ where
+\begin{align*}
+x_0 & : \prd{i:\Gamma_0}A_0(i)\\
+x_1 & : \prd*{i,j:\Gamma_0}{q:\Gamma_1}A_1(q,x_0(i),x_0(j)).
+\end{align*}
+Thus we define
+\begin{equation*}
+\terms{A}\defeq\sm{x_0:\prd{i:\Gamma_0}A_0(i)}\prd*{i,j:\Gamma_0}{q:\Gamma_1}A_1(q,x_0(i),x_0(j)).
+\end{equation*}
+We also write $\Gamma\vdash x:A$ when $x$ is a term of the graph $A$ in context
+$\Gamma$.
+\end{defn}
+
+\subsubsection{The interpretations of extension, weakening and substitution}
 \begin{defn}
@@
 \begin{defn}
-Suppose that $\Gamma\vdash A:\mfGraph$. A term $x$ of $A$ consists of a pair
-$\pairr{x_0,x_1}$ where
-\begin{align*}
-x_0 & : \prd{i:\Gamma_0}A_0(i)\\
-x_1 & : \prd*{i,j:\Gamma_0}{q:\Gamma_1}A_1(q,x_0(i),x_0(j)).
-\end{align*}
-Thus we define
-\begin{equation*}
-\terms{A}\defeq\sm{x_0:\prd{i:\Gamma_0}A_0(i)}\prd*{i,j:\Gamma_0}{q:\Gamma_1}A_1(q,x_0(i),x_0(j)).
-\end{equation*}
-We also write $\Gamma\vdash x:A$ when $x$ is a term of the graph $A$ in context
-$\Gamma$.
-\end{defn}
-
-\begin{defn}
 Suppose that $P$ is a family of graphs over $A$ in context $\Gamma$ and let
@@
 \end{rmk}
+
+\begin{rmk}
+Using weakening we can describe the graph morphisms. Suppose that $\Delta$ and
+$\Gamma$ are graphs. A graph morphism from $\Gamma$ to $\Delta$ is a term of 
+the family $\ctxwk{\Gamma}{\Delta}$ of graphs over $\Gamma$. More explicitly,
+a graph morphism $f$ is a pair
+\begin{align*}
+\pts{f} & : \pts{\Gamma}\pts{\Delta}\\
+\edg{f} & : \prd*{i,j:\pts{\Gamma}} \edg{\Gamma}(i,j)\to\edg{\Delta}(\pts{f}(i),\pts{f}(j))
+\end{align*}
+just as expected.
+
+We may also describe morphisms of families of graphs this way.
+Let $A$ and $B$ be two families of graphs over a graph $\Gamma$. 
+A morphism from $A$ to $B$ is a term of $\ctxwk{A}{B}$. More explicitly,
+a morphism $f$ from $A$ to $B$ is a pair $\pairr{f_0,f_1}$
+consisting of
+\begin{align*}
+\pts{f} & : \prd*{i:\Gamma_0} A_0(i)\to B_0(i)\\
+\edg{f} & : \prd{\pairr{i,x},\pairr{j,y}:\pts{\ctxext{\Gamma}{A}}}{\pairr{q,e}:\edg{\ctxext{\Gamma}{A}}(\pairr{i,x},\pairr{j,y})} \edg{B}(q,f_0(x),f_0(y)).
+\end{align*}
+Here we have suppressed the notation for the projections.
+\end{rmk}
+
+\subsubsection{The interpretations of the type constructors}
 
@@
 
-\subsubsection{The contexts of $\mfGraph$}
-
-The category $\psh(\cdot{\rightrightarrows}\cdot)$ of presheaves over
-$\cdot{\rightrightarrows}\cdot$ is given by
-\begin{equation*}
-\sm{A_0,A_1:\type}(A_1\to A_0)\times(A_1\to A_0).
-\end{equation*}
-To see that $\psh(\cdot{\rightrightarrows}\cdot)$ is indeed equivalent to
-the type $\tfGraph$ of all graphs, note that we have the equivalences
-\begin{align*}
-\psh(\cdot{\rightrightarrows}\cdot) & \eqvsym \sm{A_0,A_1:\type}A_1\to A_0\times A_0\\
-& \eqvsym \sm{A_0:\type} A_0\times A_0\to\type\\
-& \eqvsym \tfGraph.
-\end{align*}
-
-\begin{defn}
-Suppose that $\Delta\defeq \pairr{\Delta_0,\Delta_1}$ and $\Gamma\defeq \pairr{\Gamma_0,
-\Gamma_1}$ are graphs. A morphism $f$ from $\Delta$ to $\Gamma$ is a pair
-$\pairr{f_0,f_1}$ consisting of
-\begin{align*}
-f_0 & : \Delta_0\to\Gamma_0\\
-f_1 & : \prd*{u,v:\Delta_0}\Delta_1(u,v)\to\Gamma_1(f_0(u),f_0(v))
-\end{align*}
-We write $\ctxhom{\Delta}{\Gamma}$ for the type of morphisms of graphs.
-\end{defn}
-
-The terminal graph $\unit^\mfGraph$ plays the role of the empty context.
-
-\begin{defn}
-The terminal graph $\unit^\mfGraph$, which we shall often denote simply 
-by $\unit$, is defined by
-\begin{align*}
-{\unit^\mfGraph}_0 & \defeq \unit\\
-{\unit^\mfGraph}_1(x,y) & \defeq \unit.
-\end{align*}
-\end{defn}
-
-It is immediate that the type of graph morphisms $\Gamma\to\unit^\mfGraph$ is
-contractible for every graph $\Gamma$. The terms of a graph $\Gamma$ are the
-global elements of $\Gamma$, i.e.~they are the morphisms of type 
-$\ctxhom{\unit^\mfGraph}{\Gamma}$.
-
-\begin{defn}
-Let $\Gamma$ be a graph. The type $\terms[\mfGraph]{\Gamma}$ of terms of $\Gamma$ is
-defined to be
-\begin{equation*}
-\sm{x_0:\Gamma_0}\Gamma_1(x_0,x_0).
-\end{equation*}
-\end{defn}
-
-We now turn to the description of the universe $\gobjclass$ of graphs. Note
-that for the category $I\defeq 0{\rightrightarrows}1$, we have
-$I/0\defeq\catid{0}$ and $I/1\defeq s\rightarrow\catid{1}\leftarrow t$, where
-$s$ and $t$ are the morphisms $0\to 1$. Thus we have 
-\begin{align*}
-\psh(I/0) & \eqvsym \type\\
-\psh(I/1) & \eqvsym \sm{X,Y:\type}X\to Y\to\type.
-\end{align*}
-The functors $\psh(\Sigma_s)$ and $\psh(\Sigma_t)$ map a presheaf 
-$\pairr{X,Y,R}$ to $X$ and $Y$, respectively. 
-Thus we obtain the following graph $\gobjclass$.
-
-\begin{defn}
-The universe $\gobjclass$ of graphs is defined to be
-\begin{align*}
-{\gobjclass}_0 & \defeq \type\\
-{\gobjclass}_1(X,Y) & \defeq X\to Y\to\type.
-\end{align*}
-\end{defn}
-
-Note that the type $\terms{\gobjclass}$ of terms of $\gobjclass$ is 
-exactly $\tfGraph$.
-
-\subsubsection{Families of graphs}
-Using $\gobjclass$, we obtain a description of what
-it means to be a graph in a context $\Gamma$, where $\Gamma$ is itself a
-graph. The notion of ``a graph in a context $\Gamma$'' is the interpretation
-of ``a type in a context $\Gamma$''.
-
-\begin{defn}
-A \emph{graph $A$ in the context $\Gamma$} is a pair $\pairr{A_0,A_1}$ consisting
-of 
-\begin{align*}
-A_0 & :\Gamma_0\to\type\\
-A_1 & :\prd*{i,j:\Gamma_0}\Gamma_1(i,j)\to A_0(i)\to A_0(j)\to\type.
-\end{align*}
-We will also write $\Gamma\vdash A:\mfGraph$ when $A$ is a graph in context
-$\Gamma$.
-\end{defn}
-
-\begin{rmk}
-Note that every graph can be seen as a graph in context $\unit^\mfGraph$.
-\end{rmk}
-
-\begin{defn}
-Let $A$ be a graph in context $\Gamma$ and let $f:\ctxhom{\Delta}{\Gamma}$
-be a graph morphism. We define the graph $A[f]$ in context $\Delta$ by
-\begin{align*}
-A[f]_0(u) & \defeq A_0(f_0(u))\\
-A[f]_1(p,x,y) & \defeq A_1(f_1(p),x,y).
-\end{align*}
-\end{defn}
-
-\begin{defn}
-Let $\Gamma\vdash A:\mfGraph$ and $\Gamma\vdash B:\mfGraph$ be two graphs in
-the context $\Gamma$. A morphism $f$ from $A$ to $B$ is a pair $\pairr{f_0,f_1}$
-consisting of
-\begin{align*}
-f_0 & : \prd*{i:\Gamma_0} A_0(i)\to B_0(i)\\
-f_1 & : \prd*{i,j:\Gamma_0}{q:\Gamma_0}*{x:A_0(i)}*{y:A_0{j}}A_1(q,x,y)\to B_1(q,f_0(x),f_0(y)).
-\end{align*}
-We write $\ctxhom[\Gamma]{A}{B}$ for the type of morphisms between graphs in
-context $\Gamma$.
-\end{defn}
-
-\begin{defn}
-The unit graph $\unit^\Gamma$ in context $\Gamma$ consists of
-\begin{align*}
-(\unit^\Gamma)_0 & \defeq \lam{i}\unit\\
-(\unit^\Gamma)_1(q) & \defeq \lam{x}{y}\unit.
-\end{align*}
-\end{defn}
-
-As was the case with graphs, terms of graphs in a context are described by
-morphisms from $\unit$. Explicating this in more detail, we get
-
-\begin{defn}
-Let $A$ be a graph in context $\Gamma$. A term $x$ of $A$ is a pair
-$\pairr{x_0,x_1}$ consisting of
-\begin{align*}
-x_0 & : \prd{i:\Gamma_0}A_0(i)\\
-x_1 & : \prd*{i,j:\Gamma_0}{q:\Gamma_1(i,j)}A_1(q,x_0(i),x_0(j)).
-\end{align*}
-We define $\terms[\mfGraph]{A}$ to be the type of terms of $A$.
-\end{defn}
-
-\begin{rmk}
-Note that the type of terms of a graph $\Gamma$ is equivalent to the type
-of terms of the graph $\Gamma$ in context $\unit^\mfGraph$.
-\end{rmk}
-
-\begin{defn}
-Suppose that $\Gamma\vdash A:\mfGraph$. Then we define the graph $\ctxext{\Gamma}{A}$
-by
-\begin{align*}
-(\ctxext{\Gamma}{A})_0 & \defeq \sm{i:\Gamma_0}A_0(i)\\
-(\ctxext{\Gamma}{A})_1(\pairr{i,x},\pairr{j,y}) & \defeq \sm{q:\Gamma_1(i,j)}A_1(q,x,y).
-\end{align*}
-There is a morphism $\pi^A:\ctxhom{\ctxext{\Gamma}{A},\Gamma}$ given by
-\begin{align*}
-\pi^A_0 & \defeq \lam{\pairr{i,x}}i && : (\sm{i:\Gamma_0}A_0(i))\to \Gamma_0\\
-\pi^A_1(\pairr{i,x},\pairr{j,y}) & \defeq \lam{\pairr{q,e}}q && : (\sm{q:\Gamma_1(i,j)}A_1(q,x,y))\to\Gamma_1(i,j).
-\end{align*}
-\end{defn}
-
-\begin{rmk}
-The construction of $\ctxext{\Gamma}{A}$ is similar to the Grothendieck
-construction of the category of elements for a presheaf.
-\end{rmk}
-
-\begin{defn}
-Let $\Gamma\vdash A:\mfGraph$. A family $P$ of graphs over $A$ is by
-definition a graph $P$ in the context $\ctxext{\Gamma}{A}$. More explicitly, a family
-$P$ of graphs over $A$ in context $\Gamma$ consists of
-\begin{align*}
-P_0 & : \prd*{i:\Gamma_0}A_0(i)\to\type\\
-P_1 & : \prd*{i,j:\Gamma_0}{q:\Gamma_1(i,j)}*{x:A_0(i)}*{y:A_0(j)} A_1(q,x,y)\to P_0(x)\to P_0(y)\to\type
-%P_1 & : \prd{\pairr{i,x},\pairr{j,y}:(\ctxext{\Gamma}{A})_0}(\ctxext{\Gamma}{A})_1(\pairr{i,x},\pairr{j,y})\to P_0(i,x)\to P_0(j,y)\to\type.
-\end{align*}
-\end{defn}
-
-\begin{defn}
-When $P$ is a family of graphs over $A$ in context $\Gamma$ and $x:A$, we
-get a graph $\subst{x}{P}$ in context $\Gamma$ given by
-\begin{align*}
-\subst{x}{P}_0(i) & \defeq P_0(x_0(i))\\
-\subst{x}{P}_1(q,u,v) & \defeq P_1(\pairr{q,x_1(q)},u,v)
-\end{align*}
-\end{defn}
-
-\subsubsection{The basic type constructors for graphs}
 With families of graphs being available, we can give the interpretations of
@@
 \end{rmk}
-
-\begin{rmk}
-For any two graphs $A$ and $B$ in a context $\Gamma$, we have the graph
-$B[\pi^A]$ in the context $\ctxext{\Gamma}{A}$, which we may just write
-as $B$. Therefore, we can
-consider the graphs $\mprd{A}{B}$ and $\msm{A}{B}$. As usually those shall be
-denoted  by $A\to B$ and $A\times B$, respectively.
-
-A term $f$ of $A\to B$ consists of
-\begin{align*}
-f_0 & : \prd*{i:\Gamma_0}A_0(i)\to B_0(i)\\
-f_1 & : \prd*{i,j:\Gamma_0}{q:\Gamma_1(i,j)}*{x:A_0(i)}*{y:A_0(j)}A_1(q,x,y)\to B_1(q,f_0(x),f_0(y)),
-\end{align*}
-so we see that $\terms{A\to B}\jdeq\ctxhom[\Gamma]{A}{B}\jdeq\terms{B[\pi^A]}$.
-
-A term $\pairr{x,y}$ of $A\times B$ consists of
-\begin{align*}
-\pairr{x,y}_0 & : \prd*{i:\Gamma_0} A_0(i)\times B_0(i)\\
-\pairr{x,y}_1 & : \prd*{i,j:\Gamma_0}{q:\Gamma_1(i,j)} A_1(q,x_0(i),x_0(j))\times B_1(q,y_0(i),y_0(j))
-\end{align*}
-so we see that $\eqv{\terms{A\times B}}{\terms{A}\times\terms{B}}$.
-
-Note that we also obtain
-a \emph{graph} $\Delta\to\Gamma$ this way, for every two graphs $\Delta$ and
-$\Gamma$ seen as graphs in the context $\unit^\mfGraph$. 
-The type $\terms{\Delta\to\Gamma}$ of terms of $\Delta\to\Gamma$ is just
-$\ctxhom{\Delta}{\Gamma}$.
-\end{rmk}
-
-\begin{defn}
-For any graph $\Gamma$ there is a graph $\tffam{\Gamma}$ in context 
-$\gobjclass$ defined by
-\begin{align*}
-(\tffam{\Gamma})_0 & \defeq \lam{X}X\to\Gamma_0\\
-(\tffam{\Gamma})_1(X,Y,R) & \defeq \lam{f}{g}\prd*{x:X}*{y:Y}R(x,y)\to\Gamma_1(f(x),g(y)).
-\end{align*}
-\end{defn}
-
-\begin{rmk}
-A term $D$ of $\msm{\gobjclass}{\tffam{\Gamma}}$ consists of a term
-$\pairr{\Delta_0,f_0}$ of type
-\begin{equation*}
-\msm{\gobjclass}{\tffam{\Gamma}}_0\jdeq \sm{\Delta_0:\type}\Delta_0\to\Gamma_0
-\end{equation*}
-and a term $\pairr{\Delta_1,f_1}$ of type
-\begin{equation*}
-\msm{\gobjclass}{\tffam{\Gamma}}_1(q)\jdeq \sm{\Delta_1}
-\end{equation*}
-\end{rmk}
-
-\begin{defn}
-There is a graph morphism
-\begin{equation*}
-\graphcharmapfunc{\Gamma} : \sm{\gobjclass}
-\end{equation*}
-\end{defn}
 
@@
 
-\subsubsection{Properties of graphs}
+\begin{comment}
+\subsubsection{The contexts of $\mfGraph$}
+
+The category $\psh(\cdot{\rightrightarrows}\cdot)$ of presheaves over
+$\cdot{\rightrightarrows}\cdot$ is given by
+\begin{equation*}
+\sm{A_0,A_1:\type}(A_1\to A_0)\times(A_1\to A_0).
+\end{equation*}
+To see that $\psh(\cdot{\rightrightarrows}\cdot)$ is indeed equivalent to
+the type $\tfGraph$ of all graphs, note that we have the equivalences
+\begin{align*}
+\psh(\cdot{\rightrightarrows}\cdot) & \eqvsym \sm{A_0,A_1:\type}A_1\to A_0\times A_0\\
+& \eqvsym \sm{A_0:\type} A_0\times A_0\to\type\\
+& \eqvsym \tfGraph.
+\end{align*}
+
+We now turn to the description of the universe $\gobjclass$ of graphs. Note
+that for the category $I\defeq 0{\rightrightarrows}1$, we have
+$I/0\defeq\catid{0}$ and $I/1\defeq s\rightarrow\catid{1}\leftarrow t$, where
+$s$ and $t$ are the morphisms $0\to 1$. Thus we have 
+\begin{align*}
+\psh(I/0) & \eqvsym \type\\
+\psh(I/1) & \eqvsym \sm{X,Y:\type}X\to Y\to\type.
+\end{align*}
+The functors $\psh(\Sigma_s)$ and $\psh(\Sigma_t)$ map a presheaf 
+$\pairr{X,Y,R}$ to $X$ and $Y$, respectively. 
+Thus we obtain the following graph $\gobjclass$.
+
+\begin{defn}
+The universe $\gobjclass$ of graphs is defined to be
+\begin{align*}
+{\gobjclass}_0 & \defeq \type\\
+{\gobjclass}_1(X,Y) & \defeq X\to Y\to\type.
+\end{align*}
+\end{defn}
+
+Note that the type $\terms{\gobjclass}$ of terms of $\gobjclass$ is 
+exactly $\tfGraph$.
+
+\subsubsection{The basic type constructors for graphs}
+
+\begin{defn}
+For any graph $\Gamma$ there is a graph $\tffam{\Gamma}$ in context 
+$\gobjclass$ defined by
+\begin{align*}
+(\tffam{\Gamma})_0 & \defeq \lam{X}X\to\Gamma_0\\
+(\tffam{\Gamma})_1(X,Y,R) & \defeq \lam{f}{g}\prd*{x:X}*{y:Y}R(x,y)\to\Gamma_1(f(x),g(y)).
+\end{align*}
+\end{defn}
+
+\begin{rmk}
+A term $D$ of $\msm{\gobjclass}{\tffam{\Gamma}}$ consists of a term
+$\pairr{\Delta_0,f_0}$ of type
+\begin{equation*}
+\msm{\gobjclass}{\tffam{\Gamma}}_0\jdeq \sm{\Delta_0:\type}\Delta_0\to\Gamma_0
+\end{equation*}
+and a term $\pairr{\Delta_1,f_1}$ of type
+\begin{equation*}
+\msm{\gobjclass}{\tffam{\Gamma}}_1(q)\jdeq \sm{\Delta_1}
+\end{equation*}
+\end{rmk}
+
+\begin{defn}
+There is a graph morphism
+\begin{equation*}
+\graphcharmapfunc{\Gamma} : \sm{\gobjclass}
+\end{equation*}
+\end{defn}
+\end{comment}
 

### 2014-01-08 article-models-intro.tex
working on proposal

@@
-\section{Introduction}
+\section{Stage one: establishing internal higher categories and internal models}
 The project I propose here has its origins in the beginning of 2013, when I proved a version
@@
 univalent foundations. To arrive at a
-notion of diagram general enough to capture the higher inductive types described
+notion of diagram general enough to capture all the higher inductive types described
 in chapter 6 of \cite{TheBook} excluding the truncations we needed type
 theoretical graphs. The graphs form a model of type theory and indeed we needed
-several of the basic type theoretical operations to give an efficient approach
-to the descent theorem. Although it was not an issue to describe the graph model
+the interpretations of several of the basic type constructors to give an efficient approach
+to the descent property and its proof. Although it was not an issue to describe the graph model
 and the sense in which it models the type constructors, 
-not all of type theory is interpreted very well. To start with, context
-extension isn't strictly associative if dependent pair types are used for
-its interpretation. Also, if $A$ and $B$ are families of graphs
-over a graph $\Gamma$, then $B$ isn't also a family of graphs over the extended
-graph $\ctxext{\Gamma}{A}$ (interfering with a good interpretation of weakening).
-There are no issues with the interpretation of substitution though. Actually
-it is not so straightforward to describe what it models!
+not all of type theory is interpreted very well. To start with, if $A$ and $B$ 
+are families of graphs over a graph $\Gamma$, then $B$ isn't also a family of 
+graphs over the extended graph $\ctxext{\Gamma}{A}$ (interfering with a good 
+interpretation of weakening). Since there are such issues, it is actually
+not so straightforward to describe what it models!
 
-The graph model
-shows three operations: extension, weakening and substitution. All of them
-can be defined such that they act not only on contexts, but also on families
-and terms. Moreover, they are all compatible with each other. In fact, along
-with asserting the existence of identity morphisms (which is done in type
-theory), it seems
-that this structure is enough to establish some of the structure of a category.
-We might even go so far as to boldly assert that
+Apart from all the type constructors like dependent function and pair types and
+identity types, the graph model shows three basic operations that should come
+before those type constructors. They are extension, weakening and substitution.
+All three of them can be defined such that they act not only on contexts (the graphs), 
+but also on families and terms. Moreover, they are all compatible with each 
+other. In fact, any structure which has a notion of contexts, families and terms
+and which models extension, weakening and substitution and which has identity
+functions is in a weak sense a category, hypothesizing that
 \begin{quote}
@@
 \end{quote}
-It is a part of this proposal to test this hypothesis, with the note that by
-`category theory' we mean a theory of weak higher categories which does not
-have any of the truncatedness restrictions -- in particular not those appearing
-in the theory of AKS-categories as
-described in \cite{TheBook} in chapter 11. We note that even Dybjers theory
-of internal categories with families has such restrictions: he relies on setoids.
-It needs not much arguing that these are very unnatural from a univalent point
-of view.
+In \autoref{tt} we sketch how type theory without type constructors but with
+\emph{explicit} extension, weakening and substitution might look like. This
+theory is written down keeping the guiding principle in mind that type theory
+without basic constructors should be a generalized algebraic theory of higher
+categories.
 
+Thus, models of this theory should be higher categories with some strictness
+built in, such as the strictness of associativity of substitution. But the
+theory really only talks about contexts, families and sections. The interpretation
+of this as a category works as follows: The objects are going to be the contexts.
+Then we may consider the terms of the weakening $\ctxwk{\Gamma}{\Delta}$ of a
+context $\Delta$ by a context $\Gamma$: they are the morphisms of the category.
+The composition of morphisms follows from the action of substitution and
+weakening on terms.
+
+In the proposed approach, a model of type theory isn't a category by assumption. Rather,
+a model of type theory \emph{becomes} a category because it interprets type
+theory. In other words, type theory does all the work for us concerning the
+categorical structure. In particular, we have neither started our notion of models
+with AKS-categories as described in \cite{1categories} nor with Dybjer's notion
+of internal categories with families. We should gain that we get rid of
+truncatedness assumptions that are present in both AKS-categories, where the
+type of objects has to be a $1$-type, and internal categories with families,
+where the type of morphisms has to be a setoid. It needs not much arguing that any way to impose
+conditions on internal models beyond the interpretations of the rules of
+type theory stands in the way of a clear understanding of what an internal
+model should be and it is a main goal of the project to design
+an internal type theory without the described deficiencies.
+
+\begin{comment}
 \subsection{Ideas in the definition of internal models}
@@
 result that we have an interpretation of type theory.
+\end{comment}
 
-\subsection{The elementary theory of the category with families of categories
-with families}
+Of course there is no way of studying internal models or categories without
+also studying the morphisms between them. In fact, we already have three of them
+right from the start: the interpretations of extension, weakening and substitution
+act on contexts, families and terms and preserve each of extension, weakening and
+substitution. These have to be ingredients of morphisms of models as well. A
+morphism of models should act on contexts because those are the contexts; it
+should act on terms because that is how it acts on the morphisms; it should
+preserve extension, weakening and substitution because that is exactly what
+makes it functorial (in particular: preserve composition).
 
-Thus the idea was born to give a new description of Martin-L\"of type theory
-which is more faithful to the idea that type theory is an algebraic description
-of higher category theory. The features we have in mind for this description
-is that:
-\begin{enumerate}
-\item As in Martin-L\"of type theory there are the three basic judgments
-      \begin{align*}
-        & \Gamma\text{ is a context:} & \jctx*{\Gamma}\\
-        & A\text{ is a type in context }\Gamma\text{:} & \jtype*{\Gamma}{A}\\
-        & x\text{ is a term of $A$ in context }\Gamma\text{:} & \jterm*{\Gamma}{A}{x}
-      \end{align*}
-      together with the three accompanying judgments for judgmental equality.
-      Our new description of type theory should be such that we can faithfully
-      make the translation to the three assertions
-      \begin{enumerate}
-      \item $\Gamma$ is an object\\
-      \item $A$ is a fibration over $\Gamma$\\
-      \item $x$ is a section of the fibration $A$. 
-      \end{enumerate}
-\item There is a basic type theory which has just the operations
-      \begin{description}
-      \item[extension] taking the domain of the fibration $A$ over $\Gamma$.
-      \item[weakening] given an object $\Gamma$, weakening assigns to an object
-      $\Delta$ trivial fibration with $\Delta$ as its fibers.
-      \item[substitution] taking the fiber at a point. Fibers should always
-      be equivalent to homotopy fibers.
-      \end{description}
-      Thus we aim for a closer connection between the syntactic operations and
-      the operations of models.
-\item Models of this basic type theory should be weak $\omega$-categories. Thus,
-      it should not be necessary to start the definition of a model with a bunch
-      of contexts and substitution morphisms between them: these and the
-      operations on them, such as composition and associativity thereof, should
-      be a consequence of the interpretation of the basic type theory.
-\item There shouldn't be an empty context and dependent function types right
-      from the start. These correspond to having a terminal object and the
-      models of type theory with dependent function types are the locally
-      cartesian closed models.
-\item Contexts should not be lists of variable declarations and all the
-      operations should be explicitly invariant under context extension.
-\item There should be a large zoo of extensions of the basic type theory. First
-      off, we can extend the basic type theory with identity types, dependent
-      function types and inductive types. We want a general description of
-      homotopy colimits in type theory, which should become available once we
-      know what categories (models of the basic type theory) are. Other types
-      such as the truncations could be added as well.
+In fact, now that we have arrived at the idea that categorical structure should
+come from the interpretation of type theory, the prettier (and necessary) way to go is to
+describe a model of models of type theory. Thus, we should find a way to express
+the notions of dependent model and terms thereof and explain what extension,
+weakening and substitution mean. Then we should be able to derive that a morphism
+of models coincides with a term of a weakened object.
 
-      An extension of a different kind is the addition of the univalence axiom.
-\end{enumerate}
-
-\subsection{The further goals of the project}
-The first aim of this project is thus to give this new description of type
-theory with these properties. The second aim is to define a general notion of
-an internal model and provide several examples thereof. In the definition of
-internal models we envision that all the operations of type theory become
-morphisms of type theory: they are compatible with each and every morphism
-of type theory. The compatibility of extension, weakening and substitution is
-already part of the basic type theory. The compatibility of identity types
-with weakening for instance, is a form of function extensionality. The 
-compatibility of dependent function types with identity types is the usual
-funchtion extensionality principle. The
-compatibility of a universe operation with identity types should be univalence.
-The descent theorem implies that the colimit operation is also compatible with
-all the type theoretic operations (sigma and id are checked).
+The idea that morphisms of models preserve the basic type theoretical structure
+can be carried further. Ideally, we would have the interpretations of the type
+constructors as morphisms of models. Thus, when interpreting the operations 
+$\Pi$, $\Sigma$, $\idtypevar{}$ or the universe operator 
+(such as described in \cite{Palmgren1998}) we will require that
+they act not only on objects, but also on families, terms and preserve extension,
+weakening and substitution. And when two or more type constructors are interpreted,
+they should furthermore be compatible with each other. In the case of
+dependent pair types and identity types, this means function extensionality. It
+should be investigated whether the compatibility of identity types and the
+universe operator with each other indeed means univalence. We should note,
+however, that we intend the compatibility with extension, weakening and substitution
+should be strict, whereas the compatibility with the other type constructors will
+not be strict. To give an exact meaning to these ideas is part of the proposal.
 
@@
       have modeled the univalence axiom. 
+\item[The set model] See \cite{RijkeSpitters:Sets}.
+\item[Polynomial functors]
 \end{description}
+
+\section{Stage two: extending the basic theory}
+The first aim of this project is thus to give this new description of type
+theory with these properties. The second aim is to define a general notion of
+an internal model and provide several examples thereof. In the definition of
+internal models we envision that all the operations of type theory become
+morphisms of type theory: they are compatible with each and every morphism
+of type theory. The compatibility of extension, weakening and substitution is
+already part of the basic type theory. The compatibility of identity types
+with weakening for instance, is a form of function extensionality. The 
+compatibility of dependent function types with identity types is the usual
+funchtion extensionality principle. The
+compatibility of a universe operation with identity types should be univalence.
+The descent theorem implies that the colimit operation is also compatible with
+all the type theoretic operations (sigma and id are checked).
 
@@
 type theory are strict. We should look to the possibility to weaken them.
-
-\subsection{Overview of the document}
-In \autoref{eg} we begin the project with exploring internal models of type theory
-in type theoretical setting of \cite{TheBook}.

### 2014-01-08 article-models-main.tex
moved some stuff around
Moved universal property of extension away since it is inclear whether we want it
working on the introduction
working on proposal

--- 2013-10-24 article-models-main.tex
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\title{Internal models of type theory}
+\title{Internal models of type theory and internal higher categories for the
+Univalent Foundations}
 \author{Egbert Rijke}
@@
 
+\begin{abstract}
+A project is proposed where we investigate notions of internal models of type
+theory and internal higher categories. We take the point of view that an
+internal model is an internal higher category with extra structure interpreting
+the type constructors. We propose to (1) explore possible definitions of both
+notions, (2) find a zoo of examples and (3) find extensions of the underlying
+theory and models thereof to include higher inductive types.
+\end{abstract}
+
 \input{article-models-intro}
 
-\input{article-models-eg}
+%\input{article-models-eg}
+
+\appendix
+\input{article-models-tt}
+
+\input{article-models-idtypes}
+
+%\input{article-models-constructors}
 
@@
 
-\bibliographystyle{hplain}
+\bibliographystyle{plain}
 %\phantomsection\addcontentsline{toc}{section}{References}

### 2014-01-08 article-models-tt.tex
working on proposal

@@
-\section{Variable-free type theory -- the basic theory}
+\section{Type theory without type constructors}\label{tt}
 We have seen in the example of $\UU$ that we will not need to refer to variables

### 2014-01-08 article-models-stage1.tex
working on proposal

--- 2014-01-08 article-models-intro.tex
@@
-\section{Stage one: establishing internal higher categories and internal models}
+\section{Stage one: establishing internal higher categories and internal models}\label{stage1}
 The project I propose here has its origins in the beginning of 2013, when I proved a version
@@
 \item[Equifibered diagrams over a graph] should be the fibrations in some model.
-\item[The model of all models] Very likely, we will even define this simultaneously
-      with the notion of model.
+\item[The model of all models] The objects in the model of models should be
+      the models of type theory itself. There should be various flavors of the
+      model of models. In the case of type theory without type constructors,
+      this would be the model of all (small) higher categories. The model of
+      models of type theory with several type constructors should interpret
+      these type constructors as well. (This might get interesting in the case
+      of type theory with a universe operator.)
 \item[The model of weak $\omega$-groupoids] Since we think of models of the
@@
       provide a condition on those models which enables us to talk about
-      weak $\omega$-categories. In fact, we have several options for such a
+      weak $\omega$-groupoids. In fact, we might have several options for such a
       condition. We investigate those; we also investigate how they relate to
@@
 \item[The set model] See \cite{RijkeSpitters:Sets}.
-\item[Polynomial functors]
+\item[Polynomial functors] The theory of polynomial functors is essential to the
+      theory of $\tfW$-types, see 
 \end{description}
-
-\section{Stage two: extending the basic theory}
-The first aim of this project is thus to give this new description of type
-theory with these properties. The second aim is to define a general notion of
-an internal model and provide several examples thereof. In the definition of
-internal models we envision that all the operations of type theory become
-morphisms of type theory: they are compatible with each and every morphism
-of type theory. The compatibility of extension, weakening and substitution is
-already part of the basic type theory. The compatibility of identity types
-with weakening for instance, is a form of function extensionality. The 
-compatibility of dependent function types with identity types is the usual
-funchtion extensionality principle. The
-compatibility of a universe operation with identity types should be univalence.
-The descent theorem implies that the colimit operation is also compatible with
-all the type theoretic operations (sigma and id are checked).
-
-We also wish to extend the descent property to a more general class of homotopy
-colimits. The general recipe has already been implicit in the first descent
-property where we have used diagrams over graphs. We have the following in mind:
-\begin{enumerate}
-\item Give an internal description of a diagram. Probably this is the same thing
-      as an internal model of the basic type theory with or without identity types.
-\item Define the notion of homotopy colimit thereof.
-\item Define the notion if equifibered diagrams.
-\item Formulate and prove descent: the map defined by substitution
- from families over the colimit to the equifibered diagrams is an equivalence.
-\end{enumerate}
-
-Using the descent property we can view $\tfcolim$ as an operation of type theory
-It should preserve all the basic operations, identity types, dependent function
-types, itself, etc.
-
-Another thing we must keep in mind is that the compatibility rules in the basic
-type theory are strict. We should look to the possibility to weaken them.

### 2014-01-08 article-models-stage2.tex
working on proposal

--- 2013-11-10 article-models-intro.tex
@@
-\section{Introduction}
-This project has its origins in the beginning of 2013, when I proved a version
-of the descent theorem for homotopy colimits in type theory while I was working
-with Bas Spitters to develop notions from higher category theory in the
-univalent foundations. To arrive at a
-notion of diagram general enough to capture the higher inductive types described
-in chapter 6 of \cite{TheBook} excluding the truncations we needed type
-theoretical graphs. The graphs form a model of type theory and indeed we needed
-several of the basic type theoretical operations to give an efficient approach
-to the descent theorem. However it was not an issue to describe the graph model
-and the sense in which it was a model, the presentation of it was unsystematic
-and in that sense unsatisfactory too.
-
-From the experience of working with the graph model it became appearent that the
-answer we had of the question of what it models, i.e.~Martin L\"of type theory
-with the univalence axiom, could have been different if we were to describe a
-type theory based on the mechanism on which the model was built. The particular
-featueres of the graph model were that there was a type of graphs resembling a category
-of contexts, for each context a notion of what it means to be a family of graphs
-and a term thereof; there are the operations of context extension, weakening and
-substitution. In other words, the graph model is an instance of type theretical
-categories with families, an invention of Peter Dybjer.
-
-As we know however, the type theoretical description of a category as given in
-chapter 11 of \cite{TheBook} is restricted to categories which have a $1$-type
-as the type of objects and for every two objects a $0$-type of morphisms.
-Although categories with families are like that too, these are not very natural
-restrictions when one is concerned about internal models of type theory. It
-seems we're in the need of a higher categorical formulation of what an internal
-model, or a category with family, is. But type theory has the tendency to take
-care of it's own higher categorical structure and one could hope that type
-theory is an elementary theory of the category with families of category with
-families.
-
-\subsection{Ideas in the definition of internal models}
-An internal model of type theory is like a category with families, but we want
-to avoid having to state higher coherences. In fact, we don't even start our
-definition with a category of contexts; instead we just take a \emph{type} of contexts. 
-The morphisms will come from the terms, evaluation of a function at a given
-term will come from substitution. We recognize three basic ingredients to models:
-first there is a type of contexts; second, for every context there is a model of types in
-that context and third, for every type in a given context there is a type of its
-terms. Then there are three basic attributes: context extension, weakening and
-substitution. Context extension provides us with families over types as well as
-with an interpretation of dependent pair types. We need weakening 
-so that families can depend on the same type multiple times (the way the
-identity type of a type depends two times on that type) and to be able
-to talk about non-dependent function types,
-the morphims of our category. Substitution will give us a way
-to work with fibers of families as well as composition of functions and evaluation
-of functions at terms.
-
-Because we require a \emph{model} of types in a context, all the structure
-which we require at the bottom level will be required to exist higher up as well.
-Thus, the model of types in a given context $\Gamma$ will have a type of contexts
-itself, which can be seen as the type of types in $\Gamma$; it will have its
-own notion of types in a context, its own notion of terms, context extension,
-weakening and substitution together with all the structure require for it. For
-instance, when $A$ is a type in context $\Gamma$ in a model $\mfM$, then there
-is the model of types in context $A$, which is the model of families over $A$. 
-This model is required to be \emph{definitionally equal to} the model of types
-in the context $\ctxext{\Gamma}{A}$, the context extension of $\Gamma$ and $A$.
-In this way we protect ourselves from the need to dig an infinitely deep structure
-of models when we want to consider examples.
-
-To give the definition of a model we shall also need to consider certain morphisms
-of internal models. Those should preserve all the structure: contexts are mapped
-to contexts; for every context a morphism of models mapping the model of types
-in that context to the model of types in the image of that context; there should
-be a mapping of terms and context extension, weakening and substitution should be
-preserved. We need to consider those morphisms because we require context extension,
-weakening and substitution to be of that kind, thereby respecting each other
-in all possible ways.
-
-When we have this framework set up, we can interpret the basic type constructors
-such as $\Pi$, $\Sigma$ and $\idtypevar{}$.
-The higher categorical structure then comes from the
-result that we have an interpretation of type theory.
-
-\subsection{The elementary theory of the category with families of categories
-with families}
-
-Thus the idea was born to give a new description of Martin-L\"of type theory
-which is more faithful to the idea that type theory is an algebraic description
-of higher category theory. The features we have in mind for this description
-is that:
-\begin{enumerate}
-\item As in Martin-L\"of type theory there are the three basic judgments
-      \begin{align*}
-        & \Gamma\text{ is a context:} & \jctx*{\Gamma}\\
-        & A\text{ is a type in context }\Gamma\text{:} & \jtype*{\Gamma}{A}\\
-        & x\text{ is a term of $A$ in context }\Gamma\text{:} & \jterm*{\Gamma}{A}{x}
-      \end{align*}
-      together with the three accompanying judgments for judgmental equality.
-      Our new description of type theory should be such that we can faithfully
-      make the translation to the three assertions
-      \begin{enumerate}
-      \item $\Gamma$ is an object\\
-      \item $A$ is a fibration over $\Gamma$\\
-      \item $x$ is a section of the fibration $A$. 
-      \end{enumerate}
-\item There is a basic type theory which has just the operations
-      \begin{description}
-      \item[extension] taking the domain of the fibration $A$ over $\Gamma$.
-      \item[weakening] given an object $\Gamma$, weakening assigns to an object
-      $\Delta$ trivial fibration with $\Delta$ as its fibers.
-      \item[substitution] taking the fiber at a point. Fibers should always
-      be equivalent to homotopy fibers.
-      \end{description}
-      Thus we aim for a closer connection between the syntactic operations and
-      the operations of models.
-\item Models of this basic type theory should be weak $\omega$-categories. Thus,
-      it should not be necessary to start the definition of a model with a bunch
-      of contexts and substitution morphisms between them: these and the
-      operations on them, such as composition and associativity thereof, should
-      be a consequence of the interpretation of the basic type theory.
-\item There shouldn't be an empty context and dependent function types right
-      from the start. These correspond to having a terminal object and the
-      models of type theory with dependent function types are the locally
-      cartesian closed models.
-\item Contexts should not be lists of variable declarations and all the
-      operations should be explicitly invariant under context extension.
-\item There should be a large zoo of extensions of the basic type theory. First
-      off, we can extend the basic type theory with identity types, dependent
-      function types and inductive types. We want a general description of
-      homotopy colimits in type theory, which should become available once we
-      know what categories (models of the basic type theory) are. Other types
-      such as the truncations could be added as well.
-
-      An extension of a different kind is the addition of the univalence axiom.
-\end{enumerate}
-
-\subsection{The further goals of the project}
+\section{Stage two: extending the basic theory}\label{stage2}
 The first aim of this project is thus to give this new description of type
@@
 all the type theoretic operations (sigma and id are checked).
-
-Among the examples of internal models we should have:
-\begin{description}
-\item[The setoid model] This is a classical one so it should be there. It might
-      be a bit different in our case. We'll want to construct a setoid model
-      of the basic type theory to interpret identity types without necessarily
-      interpreting dependent function types (but that should remain possible).
-\item[The graph model] This is the model that leads to the first version of the
-      descent theorem.
-\item[Univalent unvirses] A univalent universe should be a model.
-\item[Equifibered diagrams over a graph] should be the fibrations in some model.
-\item[The model of all models] Very likely, we will even define this simultaneously
-      with the notion of model.
-\item[The model of weak $\omega$-groupoids] Since we think of models of the
-      basic type theory as weak $\omega$-categories, it is not too hard to
-      provide a condition on those models which enables us to talk about
-      weak $\omega$-categories. In fact, we have several options for such a
-      condition. We investigate those; we also investigate how they relate to
-      Brunerie's weak $\omega$-groupoids. Presumably, they form an internal
-      model without too much fuss (because we already have an internal model
-      of internal models at this stage) and we can ask whether this models
-      the univalence axiom. In fact, the model of internal models might already
-      have modeled the univalence axiom. 
-\end{description}
 

### 2014-01-08 article-models-tt.tex
working on proposal

--- 2013-12-09 article-models-tt.tex
@@
-\section{Variable-free type theory -- the basic theory}
-We have seen in the example of $\UU$ that we will not need to refer to variables
-when we are manipulating contexts. In fact, it is more natural not to.
-This suggests that there is a presentation
-of type theory where contexts do not mention variables. The first small gain is that such a presentation
-would not be burdened with comments about variables being bounded or not, or fresh or
+\section{Type theory without type constructors}\label{tt}
+In this appendix we give a possible description of type theory without basic
+constructors and with explicit extension, weakening and substitution. We will
+formulate the type theory in such a way that contexts aren't lists of variable
+declarations. The reasons for this are twofold: (1) in the models of type theory
+such as a univalent universe or the graph model, contexts are types and graphs
+respectively -- they are objects rather than lists of variable declarations; we
+intend to give a general algebraic description of higher categories. Besides,
+we will not have to be burdened with comments about variables being bounded or not, or fresh or
 not occuring at all.
 
-Perhaps a bigger gain is that we obtain an algebraic theory for categories with
-fibrations and sections.
+It should be noted that this is just a prototype description. It may be the case
+that various other elements have to be added yet, such as terms
+$\jterm{\ctxext{\Gamma}{\subst{x}{P}}}{\ctxwk{\subst{x}{P}}{\ctxext{A}{P}}}{i}$
+including the fiber $\subst{x}{P}$ into the extended family $\ctxext{A}{P}$ over
+$\Gamma$. Moreover, it hasn't been thoroughly tested whether the theory presented
+here lives up to its promise.
 
@@
 \subsection{The defining property of weakening}
-The judgmental equalities we're about to describe assert that weakening followed
-by substitution leaves everything untouched. Thus, we get that each fiber
-$\subst{x}{\ctxwk{A}{B}}$ is just $B$, that $\ctxwk{A}{y}$ is the constant function
-mapping everything to $y:B$ and similar properties for families and terms thereof.
+The judgmental equalities we're about to describe assert that substituting a term
+in the weakening a thing gives you the thing back. In the case of contexts we get that each fiber
+$\subst{x}{\ctxwk{A}{B}}$ is just $B$ and in the case of terms we get 
+that $\ctxwk{A}{y}$ is the constant function
+mapping everything to $y:B$. Thus, these rules actually establish the weakening
+as the weakening. After stating the rules we will describe what it means to
+compose context morphisms (terms of weakened contexts).
 
@@
 
-\subsection{Terms in the basic type theory}
-
-\subsubsection{Identity functions}
+\subsection{Identity functions}
 Without a rule explicitly asserting the existence of an identity morphism we don't
@@
 Using these rules we can derive that there is a judgmental equality $\jtypeeq{\ctxext{\Gamma}{A}}{\subst{x}{{\idfunc[A]}{B}}}
-{\subst{x}{{x}{B}}}$.
+{\subst{x}{{x}{B}}}$. Note that $\subst{x}{{x}{B}}$ may just be denoted by
+$\subst{x,x}{B}$.
 
@@
 
+\begin{comment}
 \subsubsection{Inclusions}
@@
 \end{align}
+\end{comment}

### 2014-02-24 article-models-idtypes.tex
added empty context and empty families

@@
 \section{Identity types}
-The identity types may now be formulated just as in ordinary Martin\nobreakdash-L\"of
-type theory. However, since all model categories have path objects, we will also
-introduce identity types for contexts.
+In this appendix we introduce identity types in for the type theory without type
+constructors of \autoref{tt}. Since we do not use variable declarations in our
+contexts, the notational appearance is somewhat different. The purpose here is
+not only to demonstrate how identity types should be introduced, but also to
+show that the new presentation of type theory is workable.
 
+\subsection{The rules for identity types}
 We will use the symbol $\reflsym$ slightly different than the book does. For us,
@@
 
-Likewise, we introduce identity types in a context.
+Likewise, we introduce identity types in a context. The rules are slightly simpler.
 
@@
 
+\begin{comment}
 Suppose we have a term $\jterm{\ctxext{\Gamma}{A}}{P}{f}$.
@@
 \end{equation*}
+\end{comment}
 
+\begin{comment}
 \subsubsection{Compatibility of identity types with extension}
@@
 \subsubsection{Compatibility of identity types with substitution}
+\end{comment}
 
 \subsection{Generalized identity types}
+From the way identity types are treated and denoted, we might take the point of
+view that the identity types are instances of an operation which acts on the
+identity functions. In the following we introduce a new type constructor
+$\eqtype{x}$ for a given term $\jterm{\Gamma}{A}{x}$ which behaves very similarly
+to identity types. In the case of $\jterm{\Gamma}{\ctxwk{\Gamma}{\Gamma}}{\idfunc[\Gamma]}$
+the type $\eqtype{\idfunc[\Gamma]}$ is just the type $\idtypevar{\Gamma}$. In
+the case of a context morphism $\jterm{\Gamma}{\ctxwk{\Gamma}{\Delta}}{f}$, the
+type $\eqtype{f}$ can be thought of as the relation $\id{f(x)}{y}$ for terms $x:\Gamma$
+and $y:\Delta$. We prove that $\eqtype{f}$ is indeed equivalent to the type
+$\subst{f}{\ctxwk{\Gamma}{\idtypevar{\Delta}}}$ when both $\eqtype{f}$ and identity
+types are present. Since the relation $\id{f(x)}{y}$ is inductively generated
+by the reflexivity term if and only if $f$ is an equivalence, asserting the existence
+of $\eqtype{f}$ for all context morphisms $\jterm{\Gamma}{\ctxwk{\Gamma}{\Delta}}{f}$
+puts a groupoid condition on the type theory.
 

### 2014-02-24 article-models-main.tex
working on proposal
added empty context and empty families

@@
 
-\tableofcontents
-
 \begin{abstract}
@@
 
-\input{article-models-intro}
+\tableofcontents
 
-%\input{article-models-eg}
+\vspace{2\baselineskip}
+The project proposal is described in \autoref{stage1} and \autoref{stage2}. In
+the appendices we give elaborations on the topics discussed in the proposal itself.
+By including these we intend to clarify the proposal and present some of the 
+(unfinished) work that is already done.
 
-\appendix
-\input{article-models-tt}
+\input{article-models-stage1}
 
-\input{article-models-idtypes}
-
-%\input{article-models-constructors}
-
-\input{article-models-defn}
+\input{article-models-stage2}
 
@@
 
+%\input{article-models-eg}
+
+\newpage
+\appendix
+\input{article-models-tt}
+
+%\input{article-models-idtypes}
+
+%\input{article-models-constructors}
+
+%\input{article-models-defn}
+
 \end{document}

### 2014-02-24 article-models-stage1.tex
added empty context and empty families

@@
 with AKS-categories as described in \cite{1categories} nor with Dybjer's notion
-of internal categories with families. We should gain that we get rid of
+of internal categories with families \cite{Dybjer1996}. We should gain that we get rid of
 truncatedness assumptions that are present in both AKS-categories, where the
@@
 \item[The graph model] This is the model that leads to the first version of the
-      descent theorem.
+      descent theorem. There are at least four variations of the graph model.
+      The canonical graph model has the usual dependent graphs as families. We
+      could also take left and right fibrations of graphs, although we loose
+      interpretations of some type constructors by doing that. However, if we
+      take the equifibered families as families we seem to get a model of type
+      theory with the usual type constructors.
 \item[Univalent unvirses] A univalent universe should be a model.
@@
 \item[Polynomial functors] The theory of polynomial functors is essential to the
-      theory of $\tfW$-types, see 
+      theory of $\tfW$-types, see \cite{MoerdijkPalmgren2000}. The theory of
+      polynomial functors has been extensively developed by Kock \cite{Kock2011}.
+      Polynomial functors are generalizations of spans and graphs are endospans,
+      so a model of polynomial functors should generalize the graph model. 
 \end{description}

### 2014-02-24 article-models-stage2.tex
added empty context and empty families

@@
 \section{Stage two: extending the basic theory}\label{stage2}
-The first aim of this project is thus to give this new description of type
-theory with these properties. The second aim is to define a general notion of
-an internal model and provide several examples thereof. In the definition of
-internal models we envision that all the operations of type theory become
-morphisms of type theory: they are compatible with each and every morphism
-of type theory. The compatibility of extension, weakening and substitution is
-already part of the basic type theory. The compatibility of identity types
-with weakening for instance, is a form of function extensionality. The 
-compatibility of dependent function types with identity types is the usual
-funchtion extensionality principle. The
-compatibility of a universe operation with identity types should be univalence.
-The descent theorem implies that the colimit operation is also compatible with
-all the type theoretic operations (sigma and id are checked).
+With a firm notion of internal models of type theory and several extensions of
+type theory come within reach. The first of these extensions, colimits for
+diagrams over graphs, we have already started to explore. The operation 
+assigning the colimit to a graph is compatible with $\Sigma$ and $\idtypevar{}$
+types. The first was compatibility result was originally part of the descent
+theorem. We take the point of view that the descent theorem -- which asserts that
+for any graph $\Gamma$, the type of equifibered families over $\Gamma$ is
+equivalent in a canonical way to the type of families over $\tfcolim(\Gamma)$ --
+is just one of the aspects of many compatibility results for colimits. In this
+formulation, the descent theorem tells that the colimit operation acts not only
+on graphs (objects) but also on families (the equifibered diagrams over graphs)
+of the model of equifibered families.
+The colimit operation acts furthermore on terms and is compatible with extension,
+weakening and substitution. The fact that it is compatible with identity types is a new result
+which essentially expresses that the initial pointed equifibered diagram over
+a pointed graph (which is just another way of considering the notion of universal
+cover) describes
+the identity type of the colimit with respect to the base point. The methods
+in this proof actually describe all the higher identity types of the colimit
+as a certain higher inductive type; in particular it gives a description of the
+loop spaces of the higher inductive types appearing in \cite{TheBook}, so we
+should figure out which algebraic data we can extract from this description.
+In summary, this group of results related to the descent theorem may be viewed as the
+key to a good description of the colimit operation.
 
-We also wish to extend the descent property to a more general class of homotopy
-colimits. The general recipe has already been implicit in the first descent
-property where we have used diagrams over graphs. We have the following in mind:
-\begin{enumerate}
-\item Give an internal description of a diagram. Probably this is the same thing
-      as an internal model of the basic type theory with or without identity types.
-\item Define the notion of homotopy colimit thereof.
-\item Define the notion if equifibered diagrams.
-\item Formulate and prove descent: the map defined by substitution
- from families over the colimit to the equifibered diagrams is an equivalence.
-\end{enumerate}
+These results need to be extended. In the
+first place we could ask ourselves whether the colimit operation is compatible
+with $\Pi$ and the universe operation. Perhaps more importantly, the descent
+theorem and the related group of results stating that the colimit operation
+is compatible with various type constructors needs to be generalized to (diagrams
+over) arbitrary categories. Since we have a proposal for what a category is,
+namely a model of type theory without basic constructors, we seem to have the
+right tools at hand.  We envision that the colimit operation becomes a type theoretical operation
+(together with compatibility results) assigning a type to a diagram over a category.
+In the presence of a universe operator we might also consider to precompose
+the colimit operation with the universe operator, so that it acts on all families.
 
-Using the descent property we can view $\tfcolim$ as an operation of type theory
-It should preserve all the basic operations, identity types, dependent function
-types, itself, etc.
-
-Another thing we must keep in mind is that the compatibility rules in the basic
-type theory are strict. We should look to the possibility to weaken them.
+The result should be a fairly general theory of colimits and one of the first
+applications we must seek for is the establishment of $\im(f)$ as the colimit
+of the appropriate category, given a function $f:A\to B$ where $A$ and $B$
+are arbitrary types. In the case where $A$ and $B$ are sets, $\im(f)$ is
+the set-colimit of $\sm{x,y:A}\id{f(x)}{f(y)}\rightrightarrows A$; in general
+this diagram should be more complicated, see \S 6.1 of \cite{lurie2009higher}.
+It should be possible to describe, given $f:A\to B$, a diagram $K_f$ such that
+its colimit is $\im(f)$ generally. In particular we shouldn't need to take the
+set-colimit in the case that $f$ is a function between sets.  When we have 
+come this far we should also look for a possible definition of a
+(weak) higher predicative topos; the ingredients seem to be there. We hope that
+this gives us access to a description of sheaf models of type theory and we
+can possibly begin exploring some internal independence proofs.

### 2014-02-24 article-models-tt.tex
added empty context and empty families

@@
 
+Much of the rules we state are just compatibility rules of extension, weakening
+and substitution with each other. To get the overview, we list the sections
+where these compatibility rules are described in the following table (in this
+table, the
+subsection mentioned in row $X$ and column $Y$ consideres the rules of the
+operation $Y\circ X$):
+
+\begin{center}
+\begin{tabular}{r|ccc}
+& extension & weakening & substitution\\
+\hline
+extension & \autoref{comp-ee} & \autoref{comp-ew} & \\
+weakening & \autoref{comp-we} & \autoref{comp-ww} & \autoref{comp-ws}\\
+substitution & \autoref{comp-se} & \autoref{comp-sw} & \autoref{comp-ss}
+\end{tabular}
+\end{center}
+
+Since extension by a context acts trivially on families and terms there are
+no rules added for extension followed by substitution.
+
 \subsection{The basic judgments}
-The judgments we wish to make are the standad six judgments for ``$\Gamma$ is a context'', ``$A$ is
-a type in context $\Gamma$'' and ``$x$ is a term of $A$ in context $\Gamma$'' and the three equality
-judgments coming with them.
+The type theory we describe here is a theory of contexts, families of
+contexts and terms thereof. The families of contexts are by some authors called
+dependent contexts, but they are handled a bit differently here because they
+become the primary object of study. Dependent contexts can be types; they could
+be seen as atomic or indecomposable dependent contexts.
+
+Thus we make eight types of judgments in our type theory: ``$\Gamma$ is a context'',
+``$A$ is a family of contexts over $\Gamma$'', ``$A$ is a type in context $\Gamma$''
+and ``$x$ is a term of the family $A$ of contexts over $\Gamma$''. The other four
+judgments are for judgmental equality.
 
@@
 \jctx*{\Gamma} & \jctxeq*{\Gamma}{\Gamma'}\\
+\jfam*{\Gamma}{A} & \jfameq*{\Gamma}{A}{B}\\
 \jtype*{\Gamma}{A} & \jtypeeq*{\Gamma}{A}{B}\\
@@
 
-In contrast with standard practise, we don't assume that there is an empty context. The idea is that
-if we interpret only the syntax of contexts, types and terms with extension, weakening and
-substitution, we do not have a full model of type theory but we still have a 
-structure resembling a category and categories need not have terminal objects. Therefore,
-the rules that follow will describe only how to manipulate contexts, types and terms.
-Of most of the inferences that we give there are two versions: one introducing an operation, the other asserting that the operation in question preserves judgmental equality.
+If $A$ is a type
+in context $\Gamma$, then $A$ is also a family of contexts over $\Gamma$. Being
+a term of type $A$ means the same thing as being a term of the family $A$ of contexts.
+Two types in context $\Gamma$ are judgmentally equal precisely when they are equal
+as context families. Moreover, if a family $B$ of contexts over $\Gamma$ is
+judgmentally equal to a type $A$ in context $\Gamma$, then $B$ is a type in
+context $\Gamma$. This is expressed by the following four inference rules:
+
+\begin{align*}
+& \inference{\jfam{\Gamma}{A}}{\jfam{\Gamma}{A}} & & \inference{\jfameq{\Gamma}{A}{B}}{\jfameq{\Gamma}{A}{B}}\\
+& \inference{\jfam{\Gamma}{A}\quad\jfameq{\Gamma}{A}{B}}{\jfam{\Gamma}{B}}
+& & \inference{\jfam{\Gamma}{A}\quad\jfameq{\Gamma}{A}{B}}{\jfameq{\Gamma}{A}{B}}
+\end{align*}
 
@@
 & & \inference{\jctxeq{\Gamma}{\Delta}\qquad\jctxeq{\Delta}{\greek{E}}}{\jctxeq{\Gamma}{\greek{E}}}\\
-& \inference{\jtype{\Gamma}{A}}{\jtypeeq{\Gamma}{A}{A}} 
-& & \inference{\jtypeeq{\Gamma}{A}{B}}{\jtypeeq{\Gamma}{B}{A}}
-& & \inference{\jtypeeq{\Gamma}{A}{B}\qquad\jtypeeq{\Gamma}{B}{C}}{\jtypeeq{\Gamma}{A}{C}}\\
+& \inference{\jfam{\Gamma}{A}}{\jfameq{\Gamma}{A}{A}} 
+& & \inference{\jfameq{\Gamma}{A}{B}}{\jfameq{\Gamma}{B}{A}}
+& & \inference{\jfameq{\Gamma}{A}{B}\qquad\jfameq{\Gamma}{B}{C}}{\jfameq{\Gamma}{A}{C}}\\
 & \inference{\jterm{\Gamma}{A}{x}}{\jtermeq{\Gamma}{A}{x}{x}}
@@
 \begin{align*}
-& \inference{\jctxeq{\Gamma}{\Delta}\qquad\jtype{\Gamma}{A}}{\jtype{\Delta}{A}}
-& & \inference{\jctxeq{\Gamma}{\Delta}\qquad\jtypeeq{\Gamma}{A}{B}}{\jtypeeq{\Delta}{A}{B}}\\
+& \inference{\jctxeq{\Gamma}{\Delta}\qquad\jfam{\Gamma}{A}}{\jfam{\Delta}{A}}
+& & \inference{\jctxeq{\Gamma}{\Delta}\qquad\jfameq{\Gamma}{A}{B}}{\jfameq{\Delta}{A}{B}}\\
 & \inference{\jctxeq{\Gamma}{\Delta}\qquad\jterm{\Gamma}{A}{x}}{\jterm{\Delta}{A}{x}}
 & & \inference{\jctxeq{\Gamma}{\Delta}\qquad\jtermeq{\Gamma}{A}{x}{y}}{\jtermeq{\Delta}{A}{x}{y}}\\
-& \inference{\jtypeeq{\Gamma}{A}{B}\qquad \jterm{\Gamma}{A}{x}}{\jterm{\Gamma}{B}{x}}
-& & \inference{\jtypeeq{\Gamma}{A}{B}\qquad\jtermeq{\Gamma}{A}{x}{y}}{\jtermeq{\Gamma}{B}{x}{y}}
+& \inference{\jfameq{\Gamma}{A}{B}\qquad \jterm{\Gamma}{A}{x}}{\jterm{\Gamma}{B}{x}}
+& & \inference{\jfameq{\Gamma}{A}{B}\qquad\jtermeq{\Gamma}{A}{x}{y}}{\jtermeq{\Gamma}{B}{x}{y}}
 \end{align*}
+
+\subsection{The empty context}
+There is an empty context and over any context there is an empty family of
+contexts. We do not assume that the empty context is a type, that would be like
+assuming that the multiplicative unit of a ring is prime. The empty family
+always has a unique term.
+
+\begin{align}
+& \inference{}{\jctx{\emptyc}}\\
+& \inference{\jctx{\Gamma}}{\jfam{\Gamma}{\emptyf\Gamma}}\\
+& \inference{\jctx{\Gamma}}{\jterm{\Gamma}{\emptyf[\Gamma]}{\emptytm[\Gamma]}}\\
+& \inference{\jterm{\Gamma}{\emptyf[\Gamma]}{x}}{\jtermeq{\Gamma}{\emptyf[\Gamma]}{x}{\emptytm[\Gamma]}}
+\end{align}
+
+Moreover, if $\Gamma$ is a context family over the
+empty context, then $\Gamma$ is a context and every context is a context
+family over the empty context. Note that this allows us to speak
+of terms of contexts too.
+
+\begin{align}
+& \inference{\jctx{\Gamma}}{\jfam{\emptyc}{\Gamma}} 
+& & \inference{\jfam{\emptyc}{\Gamma}}{\jctx{\Gamma}}\\
+& \inference{\jctxeq{\Gamma}{\Delta}}{\jfameq{\emptyc}{\Gamma}{\Delta}}
+& & \inference{\jfameq{\emptyc}{\Gamma}{\Delta}}{\jctxeq{\Gamma}{\Delta}}
+\end{align}
+
+\subsubsection{The empty context is compatible with itslef}
+The empty context $\emptyc$ may be considered as a family of contexts over the empty
+context. When we do this, we get $\emptyf[\emptyc]$.
+\begin{equation}
+\inference{}{\jfameq{\emptyc}{\emptyc}{\emptyf[\emptyc]}}
+\end{equation}
+In the future, we shall denote $\emptyf[\Gamma]$ by $\emptyf$. The above rule
+guarantees that this will not cause confusion. Likewise, we shall denote
+$\emptytm[\Gamma]$ by $\emptytm$.
 
@@
 \begin{align}
-& \inference{\jtype{\Gamma}{A}}{\jctx{\ctxext{\Gamma}{A}}}
-& & \inference{\jctxeq{\Gamma}{\Delta}\qquad\jtypeeq{\Gamma}{A}{B}}{\jctxeq{\ctxext{\Gamma}{A}}{\ctxext{\Delta}{B}}}\\
-& \inference{\jtype{\ctxext{\Gamma}{A}}{P}}{\jtype{\Gamma}{\ctxext{A}{P}}}
-& & \inference{\jtypeeq{\Gamma}{A}{B}\qquad\jtypeeq{\ctxext{\Gamma}{A}}{P}{Q}}{\jtypeeq{\Gamma}{\ctxext{A}{P}}{\ctxext{B}{Q}}}
+& \inference{\jfam{\Gamma}{A}}{\jctx{\ctxext{\Gamma}{A}}}
+& & \inference{\jctxeq{\Gamma}{\Delta}\qquad\jfameq{\Gamma}{A}{B}}{\jctxeq{\ctxext{\Gamma}{A}}{\ctxext{\Delta}{B}}}\\
+& \inference{\jfam{\ctxext{\Gamma}{A}}{P}}{\jfam{\Gamma}{\ctxext{A}{P}}}
+& & \inference{\jfameq{\Gamma}{A}{B}\qquad\jfameq{\ctxext{\Gamma}{A}}{P}{Q}}{\jfameq{\Gamma}{\ctxext{A}{P}}{\ctxext{B}{Q}}}
 \end{align}
 
-\subsubsection{Extension is compatible with itself}
+\subsubsection{Extension is compatible with the empty context}
+The following rule asserts that extension by $\emptyc$ leaves the contexts unchanged.
+\begin{align}
+& \inference{\jctx{\Gamma}}{\jctxeq{\ctxext{\emptyc}{\Gamma}}{\Gamma}}\\
+& \inference{\jctx{\Gamma}}{\jctxeq{\ctxext{\Gamma}{\emptyf}}{\Gamma}}\\
+& \inference{\jfam{\Gamma}{A}}{\jfameq{\Gamma}{\ctxext{\emptyf}{A}}{A}}
+\end{align}
+
+\subsubsection{Extension is compatible with itself}\label{comp-ee}
 The inference rules asserting that extension is compatible with itself assert
@@
 \begin{align}
-& \inference{\jtype{\Gamma}{A}\qquad\jtype{\ctxext{\Gamma}{A}}{P}}
+& \inference{\jfam{\Gamma}{A}\qquad\jfam{\ctxext{\Gamma}{A}}{P}}
   {\jctxeq{\ctxext{{\Gamma}{A}}{P}}{\ctxext{\Gamma}{{A}{P}}}}\\
-& \inference{\jtype{\ctxext{\Gamma}{A}}{P}\qquad\jtype{\ctxext{{\Gamma}{A}}{P}}{Q}}
-  {\jtypeeq{\Gamma}{\ctxext{{A}{P}}{Q}}{\ctxext{A}{{P}{Q}}}}
+& \inference{\jfam{\ctxext{\Gamma}{A}}{P}\qquad\jfam{\ctxext{{\Gamma}{A}}{P}}{Q}}
+  {\jfameq{\Gamma}{\ctxext{{A}{P}}{Q}}{\ctxext{A}{{P}{Q}}}}
 \end{align}
@@
 \begin{align}
-& \inference{\jctx{\Delta}}{\jtype{\Gamma}{\ctxwk{\Gamma}{\Delta}}} 
-& & \inference{\jctxeq{\Delta}{\Delta'}}{\jtypeeq{\Gamma}{\ctxwk{\Gamma}{\Delta}}{\ctxwk{\Gamma'}{\Delta'}}}\\
-& \inference{\jtype{\Delta}{B}}{\jtype{\ctxext{\Gamma}{\ctxwk{\Gamma}{\Delta}}}{\ctxwk{\Gamma}{B}}}
-& & \inference{\jtypeeq{\Delta}{B}{B'}}{\jtypeeq{\ctxext{\Gamma}{\ctxwk{\Gamma}{\Delta}}}{\ctxwk{\Gamma}{B}}{\ctxwk{\Gamma'}{B'}}}\\
+& \inference{\jctx{\Delta}}{\jfam{\Gamma}{\ctxwk{\Gamma}{\Delta}}} 
+& & \inference{\jctxeq{\Delta}{\Delta'}}{\jfameq{\Gamma}{\ctxwk{\Gamma}{\Delta}}{\ctxwk{\Gamma'}{\Delta'}}}\\
+& \inference{\jfam{\Delta}{B}}{\jfam{\ctxext{\Gamma}{\ctxwk{\Gamma}{\Delta}}}{\ctxwk{\Gamma}{B}}}
+& & \inference{\jfameq{\Delta}{B}{B'}}{\jfameq{\ctxext{\Gamma}{\ctxwk{\Gamma}{\Delta}}}{\ctxwk{\Gamma}{B}}{\ctxwk{\Gamma'}{B'}}}\\
 & \inference{\jterm{\Delta}{B}{y}}{\jterm{\ctxext{\Gamma}{\ctxwk{\Gamma}{\Delta}}}{\ctxwk{\Gamma}{B}}{\ctxwk{\Gamma}{y}}} 
@@
 
- In the following inference rules we assume that $\jtype{\Gamma}{A}$ and in the
+ In the following inference rules we assume that $\jfam{\Gamma}{A}$ and in the
  rules asserting a judgmental equality we assume furthermore that 
- $\jtypeeq{\Gamma}{A}{A'}$.
+ $\jfameq{\Gamma}{A}{A'}$.
 \begin{align}
-& \inference{\jtype{\Gamma}{B}}{\jtype{\ctxext{\Gamma}{A}}{\ctxwk{A}{B}}}
-& & \inference{\jtypeeq{\Gamma}{B}{B'}}{\jtypeeq{\ctxext{\Gamma}{A}}{\ctxwk{A}{B}}{\ctxwk{A'}{B'}}}\\
-& \inference{\jtype{\ctxext{\Gamma}{B}}{Q}}
-{\jtype{\ctxext{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{Q}}}
-& & \inference{\jtypeeq{\ctxext{\Gamma}{B}}{Q}{Q'}}
-{\jtypeeq{\ctxext{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{Q}}{\ctxwk{A'}{Q'}}}\\
+& \inference{\jfam{\Gamma}{B}}{\jfam{\ctxext{\Gamma}{A}}{\ctxwk{A}{B}}}
+& & \inference{\jfameq{\Gamma}{B}{B'}}{\jfameq{\ctxext{\Gamma}{A}}{\ctxwk{A}{B}}{\ctxwk{A'}{B'}}}\\
+& \inference{\jfam{\ctxext{\Gamma}{B}}{Q}}
+{\jfam{\ctxext{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{Q}}}
+& & \inference{\jfameq{\ctxext{\Gamma}{B}}{Q}{Q'}}
+{\jfameq{\ctxext{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{Q}}{\ctxwk{A'}{Q'}}}\\
 & \inference{\jterm{\Gamma}{B}{y}}{\jterm{\ctxext{\Gamma}{A}}{\ctxwk{A}{B}}{\ctxwk{A}{y}}}
@@
 
-\subsubsection{Weakening is compatible with extension}
+\subsubsection{Weakening is compatible with the empty context}
+
+Weakening by the empty context $\emptyc$ leaves contexts, context families and terms
+unchanged:
+\begin{align}
+& \inference{\jctx{\Gamma}}{\jctxeq{\ctxwk{\emptyc}{\Gamma}}{\Gamma}}\\
+& \inference{\jfam{\Gamma}{A}}{\jfameq{\Gamma}{\ctxwk{\emptyc}{A}}{A}}\\
+& \inference{\jterm{\Gamma}{A}{x}}{\jtermeq{\Gamma}{A}{\ctxwk{\emptyc}{x}}{x}}
+\end{align}
+
+Likewise, weakening by the empty family $\emptyf$ over a context $\Gamma$ leaves families, families over that and
+terms of those unchanged:
+\begin{align}
+& \inference{\jfam{\Gamma}{A}}{\jfameq{\Gamma}{\ctxwk{\emptyf}{A}}{A}}\\
+& \inference{\jfam{\ctxext{\Gamma}{A}}{P}}{\jfameq{\ctxext{\Gamma}{A}}{\ctxwk{\emptyf}{P}}{P}}\\
+& \inference{\jterm{\ctxext{\Gamma}{A}}{P}{f}}{\jtermeq{\ctxext{\Gamma}{A}}{P}{\ctxwk{\emptyf}{f}}{f}}
+\end{align}
+
+\subsubsection{Weakening is compatible with extension}\label{comp-we}
 
@@
 & \inference
-  {\jtype{\Delta}{B}}
-  {\jtypeeq{\Gamma}{\ctxwk{\Gamma}{\ctxext{\Delta}{B}}}{\ctxext{\ctxwk{\Gamma}{\Delta}}{\ctxwk{\Gamma}{B}}}}\\
+  {\jfam{\Delta}{B}}
+  {\jfameq{\Gamma}{\ctxwk{\Gamma}{\ctxext{\Delta}{B}}}{\ctxext{\ctxwk{\Gamma}{\Delta}}{\ctxwk{\Gamma}{B}}}}\\
 & \inference
-  {\jtype{\Delta}{B}\qquad\jtype{\ctxext{\Delta}{B}}{Q}}
-  {\jtypeeq
+  {\jfam{\Delta}{B}\qquad\jfam{\ctxext{\Delta}{B}}{Q}}
+  {\jfameq
     {\ctxext{\Gamma}{\ctxwk{\Gamma}{\Delta}}}
@@
 
-In the following inference rules we assume that $\jtype{\Gamma}{A}$.
+In the following inference rules we assume that $\jfam{\Gamma}{A}$.
 
@@
 & \inference
-  {\jtype{\ctxext{\Gamma}{B}}{Q}}
-  {\jtypeeq{\ctxext{\Gamma}{A}}{\ctxwk{A}{\ctxext{B}{Q}}}{\ctxext{\ctxwk{A}{B}}{\ctxwk{A}{Q}}}}\\
+  {\jfam{\ctxext{\Gamma}{B}}{Q}}
+  {\jfameq{\ctxext{\Gamma}{A}}{\ctxwk{A}{\ctxext{B}{Q}}}{\ctxext{\ctxwk{A}{B}}{\ctxwk{A}{Q}}}}\\
 & \inference
-  {\jtype{\ctxext{{\Gamma}{B}}{Q}}{R}}
-  {\jtypeeq
+  {\jfam{\ctxext{{\Gamma}{B}}{Q}}{R}}
+  {\jfameq
     {\ctxext{{\Gamma}{A}}{\ctxwk{A}{B}}}
@@
 
-\subsubsection{Weakening is compatible with itself}
-To conclude the rules for weakening, we state judgmental equality rules expressing
+\subsubsection{Weakening is compatible with itself}\label{comp-ww}
+We state judgmental equality rules expressing
 that weakening is compatible with itself. These rules state that the following
@@
 \begin{tikzcd}[column sep=huge]
-\jctx{\blank} \ar{r}{\greek{E}\mapsto\ctxwk{\Delta}{E}} \ar{d}[swap]{\greek{E}\mapsto\ctxwk{\Gamma}{E}} & \jtype{\Delta}{\blank} \ar{d}{B\mapsto\ctxwk{\Gamma}{B}}\\
-\jtype{\Gamma}{\blank} \ar{r}[swap]{A\mapsto\ctxwk{{\Gamma}{\Delta}}{A}} & \jtype{\ctxext{\Gamma}{\ctxwk{\Gamma}{\Delta}}}{\blank}
+\jctx{\blank} \ar{r}{\greek{E}\mapsto\ctxwk{\Delta}{E}} \ar{d}[swap]{\greek{E}\mapsto\ctxwk{\Gamma}{E}} & \jfam{\Delta}{\blank} \ar{d}{B\mapsto\ctxwk{\Gamma}{B}}\\
+\jfam{\Gamma}{\blank} \ar{r}[swap]{A\mapsto\ctxwk{{\Gamma}{\Delta}}{A}} & \jfam{\ctxext{\Gamma}{\ctxwk{\Gamma}{\Delta}}}{\blank}
 \end{tikzcd}
@@
 & \inference{\jctx{\Gamma}\qquad\jctx{\Delta}\qquad\jctx{\greek{E}}}
-          {\jtypeeq{\ctxext{\Gamma}{\ctxwk{\Gamma}{\Delta}}}{\ctxwk{\Gamma}{{\Delta}{\greek{E}}}}
+          {\jfameq{\ctxext{\Gamma}{\ctxwk{\Gamma}{\Delta}}}{\ctxwk{\Gamma}{{\Delta}{\greek{E}}}}
             {\ctxwk{{\Gamma}{\Delta}}{{\Gamma}{\greek{E}}}}}\\
-& \inference{\jctx{\Gamma}\qquad\jctx{\Delta}\qquad\jtype{\greek{E}}{C}}
-          {\jtypeeq{\ctxext{{\Gamma}{\ctxwk{\Gamma}{\Delta}}}{\ctxwk{{\Gamma}{\Delta}}{{\Gamma}{\greek{E}}}}}
+& \inference{\jctx{\Gamma}\qquad\jctx{\Delta}\qquad\jfam{\greek{E}}{C}}
+          {\jfameq{\ctxext{{\Gamma}{\ctxwk{\Gamma}{\Delta}}}{\ctxwk{{\Gamma}{\Delta}}{{\Gamma}{\greek{E}}}}}
             {\ctxwk{\Gamma}{{\Delta}{C}}}
@@
 \begin{align}
-& \inference{\jtype{\Gamma}{A}\qquad\jtype{\Gamma}{B}\qquad\jtype{\Gamma}{C}}
-          {\jtypeeq{\ctxext{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{{B}{C}}}
+& \inference{\jfam{\Gamma}{A}\qquad\jfam{\Gamma}{B}\qquad\jfam{\Gamma}{C}}
+          {\jfameq{\ctxext{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{{B}{C}}}
             {\ctxwk{{A}{B}}{{A}{C}}}}\\
-& \inference{\jtype{\Gamma}{A}\qquad\jtype{\Gamma}{B}\qquad\jtype{\ctxext{\Gamma}{C}}{R}}
-          {\jtypeeq{\ctxext{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{{A}{B}}{{A}{C}}}}
+& \inference{\jfam{\Gamma}{A}\qquad\jfam{\Gamma}{B}\qquad\jfam{\ctxext{\Gamma}{C}}{R}}
+          {\jfameq{\ctxext{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{{A}{B}}{{A}{C}}}}
             {\ctxwk{A}{{B}{R}}}
             {\ctxwk{{A}{B}}{{A}{R}}}}\\
-& \inference{\jtype{\Gamma}{A}\qquad\jtype{\Gamma}{B}\qquad\jterm{\Gamma}{R}{t}}
+& \inference{\jfam{\Gamma}{A}\qquad\jfam{\Gamma}{B}\qquad\jterm{\Gamma}{R}{t}}
           {\jtermeq{\ctxext{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{{A}{B}}{{A}{C}}}}
             {\ctxwk{{A}{B}}{{A}{R}}}{\ctxwk{A}{{B}{t}}}{\ctxwk{{A}{B}}{{A}{t}}}}
+\end{align}
+
+\subsubsection{The empty context is compatible with weakening}
+The following rules express that when the empty context or context family is
+weakened, the result is the empty context family.
+
+\begin{align}
+& \inference{\jctx{\Gamma}}{\jfameq{\Gamma}{\ctxwk{\Gamma}{\emptyc}}{\emptyf}}\\
+& \inference{\jfam{\Gamma}{A}}{\jfameq{\ctxext{\Gamma}{A}}{\ctxwk{A}{\emptyf}}{\emptyf}}
+\end{align}
+
+\subsubsection{Extension is compatible with weakening}\label{comp-ew}
+The rules expressing that extension is compatible with weakening assert that
+weakening by an extension is the same thing as weakening twice in the
+appropriate way.
+
+In the following, we assume that $\jctx{\Gamma}$ and $\jfam{\Gamma}{A}$.
+
+\begin{align}
+& \inference{\jctx{\Delta}}
+    {\jfameq{\ctxext{\Gamma}{A}}{\ctxwk{\ctxext{\Gamma}{A}}{\Delta}}{\ctxwk{A}{{\Gamma}{\Delta}}}}\\
+& \inference{\jfam{\Delta}{B}}
+    {\jfameq{\ctxext{{\Gamma}{A}}{\ctxwk{A}{{\Gamma}{\Delta}}}}{\ctxwk{\ctxext{\Gamma}{A}}{B}}{\ctxwk{A}{{\Gamma}{B}}}}\\
+& \inference{\jterm{\Delta}{B}{g}}
+    {\jtermeq{\ctxext{{\Gamma}{A}}{\ctxwk{A}{{\Gamma}{\Delta}}}}{\ctxwk{A}{{\Gamma}{B}}}{\ctxwk{\ctxext{\Gamma}{A}}{g}}{\ctxwk{A}{{\Gamma}{g}}}}
+\end{align}
+
+The rules for weakening by a type and a family over it are similar. We assume that
+$\jfam{\Gamma}{A}$ and $\jfam{\ctxext{\Gamma}{A}}{P}$.
+
+\begin{align}
+& \inference{\jfam{\Gamma}{B}}
+    {\jfameq{\ctxext{{\Gamma}{A}}{P}}{\ctxwk{\ctxext{A}{P}}{B}}{\ctxwk{P}{{A}{B}}}}\\
+& \inference{\jfam{\ctxext{\Gamma}{B}}{Q}}
+    {\jfameq{\ctxext{{{\Gamma}{A}}{P}}{\ctxwk{P}{{A}{B}}}}{\ctxwk{\ctxext{A}{P}}{Q}}{\ctxwk{P}{{A}{Q}}}}\\
+& \inference{\jterm{\Gamma}{B}{g}}
+    {\jtermeq{\ctxext{{\Gamma}{A}}{P}}{\ctxwk{P}{{A}{B}}}{\ctxwk{\ctxext{A}{P}}{g}}{\ctxwk{P}{{A}{g}}}}\\
+& \inference{\jterm{\ctxext{\Gamma}{B}}{Q}{s}}
+    {\jtermeq{\ctxext{{{\Gamma}{A}}{P}}{\ctxwk{P}{{A}{B}}}}{\ctxwk{P}{{A}{Q}}}{\ctxwk{\ctxext{A}{P}}{s}}{\ctxwk{P}{{A}{s}}}}
 \end{align}
@@
 \begin{align}
-& \inference{\jtype{\ctxext{\Gamma}{A}}{P}}{\jtype{\Gamma}{\subst{x}{P}}}
-& & \inference{\jtypeeq{\ctxext{\Gamma}{A}}{P}{P'}}{\jtypeeq{\Gamma}{\subst{x}{P}}{\subst{x'}{P'}}}\\
-& \inference{\jtype{\ctxext{{\Gamma}{A}}{P}}{Q}}{\jtype{\ctxext{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}}
-& & \inference{\jtypeeq{\ctxext{{\Gamma}{A}}{P}}{Q}{Q'}}{\jtypeeq{\ctxext{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}{\subst{x'}{Q'}}}\\
+& \inference{\jfam{\ctxext{\Gamma}{A}}{P}}{\jfam{\Gamma}{\subst{x}{P}}}
+& & \inference{\jfameq{\ctxext{\Gamma}{A}}{P}{P'}}{\jfameq{\Gamma}{\subst{x}{P}}{\subst{x'}{P'}}}\\
+& \inference{\jfam{\ctxext{{\Gamma}{A}}{P}}{Q}}{\jfam{\ctxext{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}}
+& & \inference{\jfameq{\ctxext{{\Gamma}{A}}{P}}{Q}{Q'}}{\jfameq{\ctxext{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}{\subst{x'}{Q'}}}\\
 & \inference{\jterm{\ctxext{\Gamma}{A}}{P}{f}}{\jterm{\Gamma}{\subst{x}{P}}{\subst{x}{f}}}
@@
 
-\subsubsection{Substitution is compatible with extension}
+\subsubsection{Substitution is compatible with the empty context}
+
+The following rules assert that substituting by the term $\jterm{\Gamma}{\emptyf}{\emptytm}$
+leaves everything unchanged.
+\begin{align}
+& \inference{\jfam{\Gamma}{A}}{\jfameq{\Gamma}{\subst{\emptytm}{A}}{A}}\\
+& \inference{\jfam{\ctxext{\Gamma}{A}}{P}}{\jfameq{\ctxext{\Gamma}{A}}{\subst{\emptytm}{P}}{P}}\\
+& \inference{\jterm{\ctxext{\Gamma}{A}}{P}{f}}{\jtermeq{\ctxext{\Gamma}{A}}{P}{\subst{\emptytm}{f}}{f}}.
+\end{align}
+
+\subsubsection{Substitution is compatible with extension}\label{comp-se}
 Suppose $\jterm{\Gamma}{A}{x}$ in all of the following inference rules.
 \begin{align}
-& \inference{\jtype{\ctxext{{\Gamma}{A}}{P}}{Q}}
-  {\jtypeeq{\Gamma}{\subst{x}{\ctxext{P}{Q}}}{\ctxext{\subst{x}{P}}{\subst{x}{Q}}}}\\
-& \inference{\jtype{\ctxext{{{\Gamma}{A}}{P}}{Q}}{R}}
-  {\jtypeeq{\ctxext{\Gamma}{\subst{x}{P}}}{\subst{x}{\ctxext{Q}{R}}}{\ctxext{\subst{x}{Q}}{\subst{x}{R}}}}
+& \inference{\jfam{\ctxext{{\Gamma}{A}}{P}}{Q}}
+  {\jfameq{\Gamma}{\subst{x}{\ctxext{P}{Q}}}{\ctxext{\subst{x}{P}}{\subst{x}{Q}}}}\\
+& \inference{\jfam{\ctxext{{{\Gamma}{A}}{P}}{Q}}{R}}
+  {\jfameq{\ctxext{\Gamma}{\subst{x}{P}}}{\subst{x}{\ctxext{Q}{R}}}{\ctxext{\subst{x}{Q}}{\subst{x}{R}}}}
 \end{align}
 
-\subsubsection{Substitution is compatible with weakening}
+\subsubsection{Substitution is compatible with weakening}\label{comp-sw}
 The rules asserting the compatibility of substitution with weakening assert
 that the following diagram commutes for any $\jterm{\Gamma}{A}{x}$ and any
-$\jtype{\ctxext{\Gamma}{A}}{P}$:
+$\jfam{\ctxext{\Gamma}{A}}{P}$:
 \begin{equation*}
 \begin{tikzcd}[column sep=huge]
-\jtype{\ctxext{\Gamma}{A}}{\blank} \ar{d}[swap]{Q\mapsto\subst{x}{Q}} \ar{r}{Q\mapsto\ctxwk{P}{Q}} & \jtype{\ctxext{{\Gamma}{A}}{P}}{\blank} \ar{d}{R\mapsto\subst{x}{R}}\\ 
-\jtype{\Gamma}{\blank} \ar{r}[swap]{B\mapsto\ctxwk{\subst{x}{P}}{B}} & \jtype{\ctxext{\Gamma}{\subst{x}{P}}}{\blank}
+\jfam{\ctxext{\Gamma}{A}}{\blank} \ar{d}[swap]{Q\mapsto\subst{x}{Q}} \ar{r}{Q\mapsto\ctxwk{P}{Q}} & \jfam{\ctxext{{\Gamma}{A}}{P}}{\blank} \ar{d}{R\mapsto\subst{x}{R}}\\ 
+\jfam{\Gamma}{\blank} \ar{r}[swap]{B\mapsto\ctxwk{\subst{x}{P}}{B}} & \jfam{\ctxext{\Gamma}{\subst{x}{P}}}{\blank}
 \end{tikzcd}
@@
 In the following inference rules we assume that $\jterm{\Gamma}{A}{x}$ and
-$\jtype{\ctxext{\Gamma}{A}}{P}$:
+$\jfam{\ctxext{\Gamma}{A}}{P}$:
 \begin{align}
 & \inference
-    {\jtype{\ctxext{\Gamma}{A}}{Q}}
-    {\jtypeeq{\ctxext{\Gamma}{\subst{x}{P}}}{\subst{x}{\ctxwk{P}{Q}}}{\ctxwk{\subst{x}{P}}{\subst{x}{Q}}}}\\
+    {\jfam{\ctxext{\Gamma}{A}}{Q}}
+    {\jfameq{\ctxext{\Gamma}{\subst{x}{P}}}{\subst{x}{\ctxwk{P}{Q}}}{\ctxwk{\subst{x}{P}}{\subst{x}{Q}}}}\\
 & \inference
-    {\jtype{\ctxext{{\Gamma}{A}}{Q}}{R}}
-    {\jtypeeq{\ctxext{{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}}{\subst{x}{\ctxwk{P}{R}}}{\ctxwk{\subst{x}{P}}{\subst{x}{R}}}}\\
+    {\jfam{\ctxext{{\Gamma}{A}}{Q}}{R}}
+    {\jfameq{\ctxext{{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}}{\subst{x}{\ctxwk{P}{R}}}{\ctxwk{\subst{x}{P}}{\subst{x}{R}}}}\\
 & \inference
@@
 
-\subsubsection{Substitution is compatible with substitution}
+\subsubsection{Substitution is compatible with substitution}\label{comp-ss}
 
@@
 substitution takes the form $\subst{x}{{f}{g}}\jdeq\subst{{x}{f}}{{x}{g}}$.
+Note that the term $\subst{{x}{f}}{{x}{g}}$ may be written down more conveniently
+as $\subst{x,\subst{x}{f}}{g}$, although we will not do that here.
 
@@
 \begin{align}
-&\inference{\jtype{\ctxext{{\Gamma}{A}}{P}}{Q}}
-{\jtypeeq{\Gamma}{\subst{x}{{f}{Q}}}{\subst{{x}{f}}{{x}{Q}}}}\\
-&\inference{\jtype{\ctxext{{{\Gamma}{A}}{P}}{Q}}{R}}
-{\jtypeeq{\ctxext{\Gamma}{\subst{x}{{f}{Q}}}}{\subst{x}{{f}{R}}}{\subst{{x}{f}}{{x}{R}}}}\\
+&\inference{\jfam{\ctxext{{\Gamma}{A}}{P}}{Q}}
+{\jfameq{\Gamma}{\subst{x}{{f}{Q}}}{\subst{{x}{f}}{{x}{Q}}}}\\
+&\inference{\jfam{\ctxext{{{\Gamma}{A}}{P}}{Q}}{R}}
+{\jfameq{\ctxext{\Gamma}{\subst{x}{{f}{Q}}}}{\subst{x}{{f}{R}}}{\subst{{x}{f}}{{x}{R}}}}\\
 &\inference{\jterm{\ctxext{{\Gamma}{A}}{P}}{Q}{g}}
@@
 
-\subsubsection{Weakening is compatible with substitution}
+\subsubsection{The empty type is compatible with substitution}
+The fibers of the empty family are the empty families:
+\begin{equation}
+\inference{\jterm{\Gamma}{A}{x}}{\jfameq{\Gamma}{\subst{x}{\emptyf}}{\emptyf}}
+\end{equation}
+
+\subsubsection{Weakening is compatible with substitution}\label{comp-ws}
 We already have rules for the compatibility of substitution with weakening, but
@@
 
-In the following inference rules we assume that $\jtype{\Gamma}{A}$ and that
+In the following inference rules we assume that $\jfam{\Gamma}{A}$ and that
 $\jterm{\Gamma}{B}{y}$.
@@
 \begin{align}
-& \inference{\jtype{\ctxext{\Gamma}{B}}{Q}}{\jtypeeq{\ctxext{\Gamma}{A}}{\ctxwk{A}{\subst{y}{Q}}}{\subst{\ctxwk{A}{y}}{\ctxwk{A}{Q}}}}\\
-& \inference{\jtype{\ctxext{{\Gamma}{B}}{Q}}{R}}{\jtypeeq{\ctxext{{\Gamma}{A}}{\ctxwk{A}{Q}}}{\ctxwk{A}{\subst{y}{R}}}{\subst{\ctxwk{A}{y}}{\ctxwk{A}{R}}}}\\
+& \inference{\jfam{\ctxext{\Gamma}{B}}{Q}}{\jfameq{\ctxext{\Gamma}{A}}{\ctxwk{A}{\subst{y}{Q}}}{\subst{\ctxwk{A}{y}}{\ctxwk{A}{Q}}}}\\
+& \inference{\jfam{\ctxext{{\Gamma}{B}}{Q}}{R}}{\jfameq{\ctxext{{\Gamma}{A}}{\ctxwk{A}{Q}}}{\ctxwk{A}{\subst{y}{R}}}{\subst{\ctxwk{A}{y}}{\ctxwk{A}{R}}}}\\
 & \inference{\jterm{\ctxext{\Gamma}{B}}{Q}{g}}{\jtermeq{\ctxext{\Gamma}{A}}{\ctxwk{A}{\subst{y}{Q}}}{\ctxwk{A}{\subst{y}{g}}}{\subst{\ctxwk{A}{y}}{\ctxwk{A}{g}}}}\\
@@
 
-\subsection{The defining property of weakening}
+\subsection{The rules establishing the categorical properties of type theory}
+\subsubsection{The defining property of weakening}
 The judgmental equalities we're about to describe assert that substituting a term
@@
 \begin{align}
-& \inference{\jtype{\Gamma}{A}\qquad\jtype{\Gamma}{B}\qquad\jterm{\Gamma}{A}{x}}{\jtypeeq{\Gamma}{\subst{x}{\ctxwk{A}{B}}}{B}}\\
-& \inference{\jtype{\Gamma}{A}\qquad\jtype{\ctxext{\Gamma}{B}}{Q}\qquad\jterm{\Gamma}{A}{x}}{\jtypeeq{\ctxext{\Gamma}{B}}{\subst{x}{\ctxwk{A}{Q}}}{Q}}\\
+& \inference{\jfam{\Gamma}{A}\qquad\jfam{\Gamma}{B}\qquad\jterm{\Gamma}{A}{x}}{\jfameq{\Gamma}{\subst{x}{\ctxwk{A}{B}}}{B}}\\
+& \inference{\jfam{\Gamma}{A}\qquad\jfam{\ctxext{\Gamma}{B}}{Q}\qquad\jterm{\Gamma}{A}{x}}{\jfameq{\ctxext{\Gamma}{B}}{\subst{x}{\ctxwk{A}{Q}}}{Q}}\\
 & \inference{\jterm{\Gamma}{A}{x}\qquad\jterm{\Gamma}{B}{y}}{\jtermeq{\Gamma}{B}{\subst{x}{\ctxwk{A}{y}}}{y}}\\
@@
 \begin{equation*}
-\jtypeeq{\Gamma}{\subst{f}{\ctxwk{\Gamma}{{\Delta}{\greek{E}}}}}{\ctxwk{\Gamma}{E}}
+\jfameq{\Gamma}{\subst{f}{\ctxwk{\Gamma}{{\Delta}{\greek{E}}}}}{\ctxwk{\Gamma}{E}}
 \end{equation*}
@@
 
-\subsection{Identity functions}
+\subsubsection{Identity functions}
 Without a rule explicitly asserting the existence of an identity morphism we don't
@@
 &\inference{\jctx{\Gamma}}{\jterm{\Gamma}{\ctxwk{\Gamma}{\Gamma}}{\idfunc[\Gamma]}}\label{ctxIdFuncForm}\\
-&\inference{\jtype{\Gamma}{A}}{\jtypeeq{\Gamma}{\subst{\idfunc[\Gamma]}{\ctxwk{\Gamma}{A}}}{A}}\label{ctxIdFuncPbInvType}\\
+&\inference{\jfam{\Gamma}{A}}{\jfameq{\Gamma}{\subst{\idfunc[\Gamma]}{\ctxwk{\Gamma}{A}}}{A}}\label{ctxIdFuncPbInvType}\\
 &\inference{\jterm{\Gamma}{A}{x}}{\jtermeq{\Gamma}{A}{\subst{\idfunc[\Gamma]}{\ctxwk{\Gamma}{x}}}{x}}\label{ctxIdFuncPbInvTerm}\\
@@
 \begin{align}
-& \inference{\jtype{\Gamma}{A}}{\jterm{\ctxext{\Gamma}{A}}{\ctxwk{A}{A}}{\idfunc[A]}}\\
+& \inference{\jfam{\Gamma}{A}}{\jterm{\ctxext{\Gamma}{A}}{\ctxwk{A}{A}}{\idfunc[A]}}\\
 & \inference{\jterm{\Gamma}{A}{x}}{\jtermeq{\Gamma}{A}{\subst{x}{\idfunc[A]}}{x}}\\
-& \inference{\jtype{\ctxext{\Gamma}{A}}{P}}{\jtypeeq{\ctxext{\Gamma}{A}}{\subst{\idfunc[A]}{\ctxwk{A}{P}}}{P}}\\
+& \inference{\jfam{\ctxext{\Gamma}{A}}{P}}{\jfameq{\ctxext{\Gamma}{A}}{\subst{\idfunc[A]}{\ctxwk{A}{P}}}{P}}\\
 & \inference{\jterm{\ctxext{\Gamma}{A}}{P}{f}}{\jtermeq{\ctxext{\Gamma}{A}}{P}{\subst{\idfunc[A]}{\ctxwk{A}{f}}}{f}} \\
@@
 
-Using these rules we can derive that there is a judgmental equality $\jtypeeq{\ctxext{\Gamma}{A}}{\subst{x}{{\idfunc[A]}{B}}}
+Using these rules we can derive that there is a judgmental equality $\jfameq{\ctxext{\Gamma}{A}}{\subst{x}{{\idfunc[A]}{B}}}
 {\subst{x}{{x}{B}}}$. Note that $\subst{x}{{x}{B}}$ may just be denoted by
@@
 \begin{equation*}
-\inference{\jterm{\Gamma}{A}{x}\qquad\jtype{\Gamma,\,a:A}{P(a)}}
+\inference{\jterm{\Gamma}{A}{x}\qquad\jfam{\Gamma,\,a:A}{P(a)}}
   {\Gamma,\,u:P(x/a)\vdash x:A, u:P(x/a)}
@@
 \begin{equation}
-\inference{\jterm{\Gamma}{A}{x}\qquad\jtype{\ctxext{\Gamma}{A}}{P}}
+\inference{\jterm{\Gamma}{A}{x}\qquad\jfam{\ctxext{\Gamma}{A}}{P}}
   {\jterm{\ctxext{\Gamma}{\subst{x}{P}}}{\ctxwk{\subst{x}{P}}{\ctxext{A}{P}}}{\typefont{in}^P_x}}
@@
 & \inference
-    {\jtype{\ctxext{{\Gamma}{A}}{P}}{Q}}
-    {\jtypeeq{\ctxext{\Gamma}{\subst{x}{P}}}{\subst{\typefont{in}^P_x}{\ctxwk{\subst{x}{P}}{Q}}}{\subst{x}{Q}}}\\
+    {\jfam{\ctxext{{\Gamma}{A}}{P}}{Q}}
+    {\jfameq{\ctxext{\Gamma}{\subst{x}{P}}}{\subst{\typefont{in}^P_x}{\ctxwk{\subst{x}{P}}{Q}}}{\subst{x}{Q}}}\\
 & \inference

### 2014-02-25 article-models-tt.tex
started removing some of the redundant rules; added extension on terms

@@
 
-Note that by introducing extension on the level of types and families,
-we introduce $\Sigma$-types at a very early stage.
-However, we need substitution to make this precise.
 \begin{align}
@@
 
-\subsection{Weakening}
-We first define weakening by a context $\Gamma$. Since weakening by $\Gamma$
-should in principle be a functor, it acts on contexts, types and terms alike.
-Note that it is because of the weakening $\ctxwk{\Gamma}{\Delta}$ of a context
-$\Delta$ by $\Gamma$ that we can speak of context morphisms from $\Gamma$ to $\Delta$: they are the terms
-of $\ctxwk{\Gamma}{\Delta}$.
-
-In the following inference rules we assume $\jctx{\Gamma}$ and in the rules
-asserting a judgmental equality we furthermore assume $\jctxeq{\Gamma}{\Gamma'}$.
-\begin{align}
-& \inference{\jctx{\Delta}}{\jfam{\Gamma}{\ctxwk{\Gamma}{\Delta}}} 
-& & \inference{\jctxeq{\Delta}{\Delta'}}{\jfameq{\Gamma}{\ctxwk{\Gamma}{\Delta}}{\ctxwk{\Gamma'}{\Delta'}}}\\
-& \inference{\jfam{\Delta}{B}}{\jfam{\ctxext{\Gamma}{\ctxwk{\Gamma}{\Delta}}}{\ctxwk{\Gamma}{B}}}
-& & \inference{\jfameq{\Delta}{B}{B'}}{\jfameq{\ctxext{\Gamma}{\ctxwk{\Gamma}{\Delta}}}{\ctxwk{\Gamma}{B}}{\ctxwk{\Gamma'}{B'}}}\\
-& \inference{\jterm{\Delta}{B}{y}}{\jterm{\ctxext{\Gamma}{\ctxwk{\Gamma}{\Delta}}}{\ctxwk{\Gamma}{B}}{\ctxwk{\Gamma}{y}}} 
-& & \inference{\jtermeq{\Delta}{B}{y}{y'}}{\jtermeq{\ctxext{\Gamma}{\ctxwk{\Gamma}{\Delta}}}{\ctxwk{\Gamma}{B}}{\ctxwk{\Gamma}{y}}{\ctxwk{\Gamma'}{y'}}} 
-\end{align}
-
-A weakening operation is also defined for types. When $A$ and $B$ are both types
- in context $\Gamma$, the weakened type $\ctxwk{A}{B}$ in context $\ctxext{\Gamma}{A}$
- is the family which `doesn't really depend on $A$'. The terms of $\ctxwk{A}{B}$
- are the functions from $A$ to $B$. Likewise, the terms $\ctxwk{A}{y}$ are the
- constant maps at $y$ from $A$ to $B$, for $y:B$.
+\subsection{The type theoretic operation of weakening}
+When $A$ is a context family over a context $\Gamma$, we wish to define a weakening
+operation $\ctxwk{A}{}$. The weakening operation acts on context families $B$ 
+over $\Gamma$, terms thereof, context families over $B$ and terms thereof.
+It weakens those, which means that it ``adds $A$ to the context''. The context
+family $\ctxwk{A}{B}$ can be seen as the constant family $B$ over $\ctxext{\Gamma}{A}$.
+Likewise, when $y$ is a term of $B$, the term $\ctxwk{A}{y}$ of $\ctxwk{A}{B}$
+can be seen as the constant term with value $y$.
 
@@
 
+We add rules asserting that a weakened type is again a type:
+
+\begin{align}
+& \inference{\jtype{\Gamma}{B}}{\jtype{\ctxext{\Gamma}{A}}{\ctxwk{A}{B}}}\\
+& \inference{\jfam{\Gamma}{B}\quad\jtype{\ctxext{\Gamma}{B}}{Q}}{\jtype{\ctxext{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{Q}}}
+\end{align}
+
 \subsubsection{Weakening is compatible with the empty context}
 
-Weakening by the empty context $\emptyc$ leaves contexts, context families and terms
-unchanged:
+Weakening by the empty family $\emptyf$ over a context $\Gamma$ leaves families, 
+their terms, families over those families and
+terms of those unchanged:
 \begin{align}
-& \inference{\jctx{\Gamma}}{\jctxeq{\ctxwk{\emptyc}{\Gamma}}{\Gamma}}\\
-& \inference{\jfam{\Gamma}{A}}{\jfameq{\Gamma}{\ctxwk{\emptyc}{A}}{A}}\\
-& \inference{\jterm{\Gamma}{A}{x}}{\jtermeq{\Gamma}{A}{\ctxwk{\emptyc}{x}}{x}}
+& \inference{\jfam{\Gamma}{B}}{\jfameq{\Gamma}{\ctxwk{\emptyf}{B}}{B}}\\
+& \inference{\jterm{\Gamma}{B}{y}}{\jtermeq{\Gamma}{B}{\ctxwk{\emptyf}{y}}{y}}\\
+& \inference{\jfam{\ctxext{\Gamma}{B}}{Q}}{\jfameq{\ctxext{\Gamma}{B}}{\ctxwk{\emptyf}{Q}}{Q}}\\
+& \inference{\jterm{\ctxext{\Gamma}{B}}{Q}{g}}{\jtermeq{\ctxext{\Gamma}{B}}{Q}{\ctxwk{\emptyf}{g}}{g}}
 \end{align}
 
-Likewise, weakening by the empty family $\emptyf$ over a context $\Gamma$ leaves families, families over that and
-terms of those unchanged:
+\subsubsection{The empty context is compatible with weakening}
+The following rules express that when the empty context or context family is
+weakened, the result is the empty context family.
+
 \begin{align}
-& \inference{\jfam{\Gamma}{A}}{\jfameq{\Gamma}{\ctxwk{\emptyf}{A}}{A}}\\
-& \inference{\jfam{\ctxext{\Gamma}{A}}{P}}{\jfameq{\ctxext{\Gamma}{A}}{\ctxwk{\emptyf}{P}}{P}}\\
-& \inference{\jterm{\ctxext{\Gamma}{A}}{P}{f}}{\jtermeq{\ctxext{\Gamma}{A}}{P}{\ctxwk{\emptyf}{f}}{f}}
+& \inference{\jctx{\Gamma}}{\jfameq{\Gamma}{\ctxwk{\Gamma}{\emptyc}}{\emptyf}}\\
+& \inference{\jfam{\Gamma}{A}}{\jfameq{\ctxext{\Gamma}{A}}{\ctxwk{A}{\emptyf}}{\emptyf}}
 \end{align}
@@
 The following rules assert the compatibility of extension with weakening: for
-every context $\Gamma$ and every type $B$ in context $\Delta$ there is a
-judgmental equality $\ctxwk{\Gamma}{\ctxext{\Delta}{B}}\jdeq\ctxext{\ctxwk{\Gamma}{\Delta}}
-{\ctxwk{\Gamma}{B}}$. When thinking of terms of $\ctxwk{\Gamma}{\Delta}$ as
-functions from $\Gamma$ to $\Delta$, this looks awfully lot like the type
-theoretical axiom of choice, of which it seems absurd to ask to hold judgmentally.
-But the rule should be there because in ordinary Martin-L\"of type theory
-there is no syntax for weakening, i.e.~weakening is invisible. 
-And thus both $\ctxwk{\Gamma}{\ctxext{\Delta}{B}}$ 
-and $\ctxext{\ctxwk{\Gamma}{\Delta}}{\ctxwk{\Gamma}{B}}$ is the type
-$\ctxext{\Delta}{B}$ in context $\Gamma$.
+every family $A$ over $\Gamma$ and every family $Q$ over $\ctxext{\Gamma}{B}$
+there is a
+judgmental equality $\ctxwk{A}{\ctxext{B}{Q}}\jdeq\ctxext{\ctxwk{A}{B}}
+{\ctxwk{A}{Q}}$.
 
-In the following inference rules we assume that $\jctx{\Gamma}$.
-
-\begin{align}
-& \inference
-  {\jfam{\Delta}{B}}
-  {\jfameq{\Gamma}{\ctxwk{\Gamma}{\ctxext{\Delta}{B}}}{\ctxext{\ctxwk{\Gamma}{\Delta}}{\ctxwk{\Gamma}{B}}}}\\
-& \inference
-  {\jfam{\Delta}{B}\qquad\jfam{\ctxext{\Delta}{B}}{Q}}
-  {\jfameq
-    {\ctxext{\Gamma}{\ctxwk{\Gamma}{\Delta}}}
-    {\ctxwk{\Gamma}{\ctxext{B}{Q}}}
-    {\ctxext{\ctxwk{\Gamma}{B}}{\ctxwk{\Gamma}{Q}}}}
-\end{align}
+When thinking of terms of $\ctxwk{A}{B}$ as morphisms of families from $A$ to
+$B$, this looks already like form of type theoretic choice. It is weaker in that
+it is not stated with function types, yet it is stronger in that it states a
+judgmental equality between two families. When one makes the weakening operation
+notationally invisible -- as is in fact the usual practice in type theory -- the
+following compatibility rules become completely obvious.
 
 In the following inference rules we assume that $\jfam{\Gamma}{A}$.
-
 \begin{align}
@@
 that weakening is compatible with itself. These rules state that the following
-diagram commutes given any two contexts $\Gamma$ and $\Delta$:
+diagram commutes given any two families $A$ and $B$ in context $\Gamma$:
 \begin{equation*}
 \begin{tikzcd}[column sep=huge]
-\jctx{\blank} \ar{r}{\greek{E}\mapsto\ctxwk{\Delta}{E}} \ar{d}[swap]{\greek{E}\mapsto\ctxwk{\Gamma}{E}} & \jfam{\Delta}{\blank} \ar{d}{B\mapsto\ctxwk{\Gamma}{B}}\\
-\jfam{\Gamma}{\blank} \ar{r}[swap]{A\mapsto\ctxwk{{\Gamma}{\Delta}}{A}} & \jfam{\ctxext{\Gamma}{\ctxwk{\Gamma}{\Delta}}}{\blank}
+\jfam{\Gamma}{\blank} \ar{r}{C\mapsto\ctxwk{B}{C}} \ar{d}[swap]{C\mapsto\ctxwk{A}{C}} & \jfam{\ctxext{\Gamma}{B}}{\blank} \ar{d}{Q\mapsto\ctxwk{A}{Q}}\\
+\jfam{\ctxext{\Gamma}{A}}{\blank} \ar{r}[swap]{P\mapsto\ctxwk{{A}{B}}{P}} & \jfam{\ctxext{{\Gamma}{A}}{\ctxwk{A}{B}}}{\blank}
 \end{tikzcd}
 \end{equation*}
-There is also a version of this diagram in which all happens in a context. Thus,
-we get two sets of inference rules. For weakening by contexts we get:
-\begin{align}
-& \inference{\jctx{\Gamma}\qquad\jctx{\Delta}\qquad\jctx{\greek{E}}}
-          {\jfameq{\ctxext{\Gamma}{\ctxwk{\Gamma}{\Delta}}}{\ctxwk{\Gamma}{{\Delta}{\greek{E}}}}
-            {\ctxwk{{\Gamma}{\Delta}}{{\Gamma}{\greek{E}}}}}\\
-& \inference{\jctx{\Gamma}\qquad\jctx{\Delta}\qquad\jfam{\greek{E}}{C}}
-          {\jfameq{\ctxext{{\Gamma}{\ctxwk{\Gamma}{\Delta}}}{\ctxwk{{\Gamma}{\Delta}}{{\Gamma}{\greek{E}}}}}
-            {\ctxwk{\Gamma}{{\Delta}{C}}}
-            {\ctxwk{{\Gamma}{\Delta}}{{\Gamma}{C}}}}\\
-& \inference{\jctx{\Gamma}\qquad\jctx{\Delta}\qquad\jterm{\greek{E}}{C}{t}}
-          {\jtermeq{\ctxext{{\Gamma}{\ctxwk{\Gamma}{\Delta}}}{\ctxwk{{\Gamma}{\Delta}}{{\Gamma}{\greek{E}}}}}
-            {\ctxwk{{\Gamma}{\Delta}}{{\Gamma}{C}}}{\ctxwk{\Gamma}{{\Delta}{t}}}{\ctxwk{{\Gamma}{\Delta}}{{\Gamma}{t}}}}
-\end{align}
-For weakening by types we get:
+Thus, we get the following set of inference rules:
 \begin{align}
@@
             {\ctxwk{{A}{B}}{{A}{C}}}}\\
+& \inference{\jfam{\Gamma}{A}\qquad\jfam{\Gamma}{B}\qquad\jterm{\Gamma}{C}{z}}
+          {\jtermeq{\ctxext{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{{A}{B}}{{A}{C}}}
+            {\ctxwk{A}{{B}{z}}}
+            {\ctxwk{{A}{B}}{{A}{z}}}}\label{comp-ww-inf2}\\
 & \inference{\jfam{\Gamma}{A}\qquad\jfam{\Gamma}{B}\qquad\jfam{\ctxext{\Gamma}{C}}{R}}
@@
             {\ctxwk{{A}{B}}{{A}{R}}}{\ctxwk{A}{{B}{t}}}{\ctxwk{{A}{B}}{{A}{t}}}}
+            \label{comp-ww-inf4}
 \end{align}
-
-\subsubsection{The empty context is compatible with weakening}
-The following rules express that when the empty context or context family is
-weakened, the result is the empty context family.
-
-\begin{align}
-& \inference{\jctx{\Gamma}}{\jfameq{\Gamma}{\ctxwk{\Gamma}{\emptyc}}{\emptyf}}\\
-& \inference{\jfam{\Gamma}{A}}{\jfameq{\ctxext{\Gamma}{A}}{\ctxwk{A}{\emptyf}}{\emptyf}}
-\end{align}
+Note that strictly speaking, we did not need to require the inference in
+\autoref{comp-ww-inf2} since this follows from the inference in \autoref{comp-ww-inf4}
+and the fact that the weakening of the empty family is again the empty family.
 
@@
 
-In the following, we assume that $\jctx{\Gamma}$ and $\jfam{\Gamma}{A}$.
-
-\begin{align}
-& \inference{\jctx{\Delta}}
-    {\jfameq{\ctxext{\Gamma}{A}}{\ctxwk{\ctxext{\Gamma}{A}}{\Delta}}{\ctxwk{A}{{\Gamma}{\Delta}}}}\\
-& \inference{\jfam{\Delta}{B}}
-    {\jfameq{\ctxext{{\Gamma}{A}}{\ctxwk{A}{{\Gamma}{\Delta}}}}{\ctxwk{\ctxext{\Gamma}{A}}{B}}{\ctxwk{A}{{\Gamma}{B}}}}\\
-& \inference{\jterm{\Delta}{B}{g}}
-    {\jtermeq{\ctxext{{\Gamma}{A}}{\ctxwk{A}{{\Gamma}{\Delta}}}}{\ctxwk{A}{{\Gamma}{B}}}{\ctxwk{\ctxext{\Gamma}{A}}{g}}{\ctxwk{A}{{\Gamma}{g}}}}
-\end{align}
-
-The rules for weakening by a type and a family over it are similar. We assume that
+In the following inference rules we assume that
 $\jfam{\Gamma}{A}$ and $\jfam{\ctxext{\Gamma}{A}}{P}$. 
-
 \begin{align}
@@
     {\jfameq{\ctxext{{\Gamma}{A}}{P}}{\ctxwk{\ctxext{A}{P}}{B}}{\ctxwk{P}{{A}{B}}}}\\
+& \inference{\jterm{\Gamma}{B}{y}}
+    {\jtermeq{\ctxext{{\Gamma}{A}}{P}}{\ctxwk{P}{{A}{B}}}{\ctxwk{\ctxext{A}{P}}{y}}{\ctxwk{P}{{A}{y}}}}\\
 & \inference{\jfam{\ctxext{\Gamma}{B}}{Q}}
     {\jfameq{\ctxext{{{\Gamma}{A}}{P}}{\ctxwk{P}{{A}{B}}}}{\ctxwk{\ctxext{A}{P}}{Q}}{\ctxwk{P}{{A}{Q}}}}\\
-& \inference{\jterm{\Gamma}{B}{g}}
-    {\jtermeq{\ctxext{{\Gamma}{A}}{P}}{\ctxwk{P}{{A}{B}}}{\ctxwk{\ctxext{A}{P}}{g}}{\ctxwk{P}{{A}{g}}}}\\
-& \inference{\jterm{\ctxext{\Gamma}{B}}{Q}{s}}
-    {\jtermeq{\ctxext{{{\Gamma}{A}}{P}}{\ctxwk{P}{{A}{B}}}}{\ctxwk{P}{{A}{Q}}}{\ctxwk{\ctxext{A}{P}}{s}}{\ctxwk{P}{{A}{s}}}}
+& \inference{\jterm{\ctxext{\Gamma}{B}}{Q}{g}}
+    {\jtermeq{\ctxext{{{\Gamma}{A}}{P}}{\ctxwk{P}{{A}{B}}}}{\ctxwk{P}{{A}{Q}}}{\ctxwk{\ctxext{A}{P}}{g}}{\ctxwk{P}{{A}{g}}}}
 \end{align}
 
-\subsection{Substitution}
+\subsection{The type theoretic operation of substitution}
 Given a family $P$ over $A$ and a term $x$ of $A$, substitution gives a way to
@@
 
+\subsubsection{Extension on terms}
+When $\jterm{\Gamma}{A}{x}$ and $\jterm{\Gamma}{\subst{x}{P}}{u}$ are terms,
+we obtain a term $\jterm{\Gamma}{\ctxext{A}{P}}{\ctxext{x}{u}}$. This operation
+of weakening on terms is like the pairing operation. However, it is handled
+much stricter for it has a judgmental inverse and it is subject to various compatibility
+rules.
+\begin{align}
+& \inference{\jterm{\Gamma}{A}{x}\quad\jterm{\Gamma}{\subst{x}{P}}{u}}{\jterm{\Gamma}{\ctxext{A}{P}}{\ctxext{x}{u}}}\\
+& \inference{\jterm{\Gamma}{\ctxext{A}{P}}{w}}{\jterm{\Gamma}{A}{\pts{w}}}\\
+& \inference{\jterm{\Gamma}{\ctxext{A}{P}}{w}}{\jterm{\Gamma}{\subst{\pts{w}}{P}}{\edg{w}}}\\
+& \inference{\jterm{\Gamma}{\ctxext{A}{P}}{w}}{\jtermeq{\Gamma}{\ctxext{A}{P}}{\ctxext{\pts{w}}{\edg{w}}}{w}}\\
+& \inference{\jterm{\Gamma}{A}{x}\quad\jterm{\Gamma}{\subst{x}{P}}{u}}{\jtermeq{\Gamma}{A}{\pts{(\ctxext{x}{u})}}{x}}\\
+& \inference{\jterm{\Gamma}{A}{x}\quad\jterm{\Gamma}{\subst{x}{P}}{u}}{\jtermeq{\Gamma}{\subst{x}{P}}{\edg{(\ctxext{x}{u})}}{u}}
+\end{align}
+
+We add the following compatibility rule for two consecutive term extensions:
+\begin{equation}
+\inference{\jterm{\Gamma}{A}{x}\quad\jterm{\Gamma}{\subst{x}{P}}{u}\quad\jterm{\Gamma}{\subst{\ctxext{x}{u}}{Q}}{v}}
+{\jtermeq{\Gamma}{\ctxext{{A}{P}}{Q}}{\ctxext{x}{{u}{v}}}{\ctxext{{x}{u}}{v}}}
+\end{equation}
+
+When we weaken a term $\ctxext{y}{v}$ of $\ctxext{B}{Q}$ in context $\Gamma$ by
+a type $A$, the term that we get is $\ctxext{\ctxwk{A}{y}}{\ctxwk{A}{v}}$.
+\begin{align}
+& \inference{\jterm{\Gamma}{B}{y}\quad\jterm{\Gamma}{\subst{y}{Q}}{v}}{\jtermeq{\ctxext{\Gamma}{A}}{\ctxwk{A}{\ctxext{B}{Q}}}{\ctxwk{A}{\ctxext{y}{v}}}{\ctxext{\ctxwk{A}{y}}{\ctxwk{A}{v}}}}\\
+& \inference{\jterm{\ctxext{\Gamma}{B}}{Q}{g}\quad\jterm{\ctxext{\Gamma}{B}}{\subst{g}{R}}{t}}{\jtermeq{\ctxext{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{\ctxext{Q}{R}}}{\ctxwk{A}{\ctxext{g}{t}}}{\ctxext{\ctxwk{A}{g}}{\ctxwk{A}{t}}}}
+\end{align}
+
+When we substitute an extended term $\ctxext{f}{g}$ of $\ctxext{P}{Q}$ by a term
+$x$ of $A$, the term that we get is $\ctxext{\subst{x}{f}}{\subst{x}{g}}$.
+\begin{align}
+& \inference{\jterm{\ctxext{\Gamma}{A}}{P}{f}\quad\jterm{\ctxext{\Gamma}{A}}{\subst{f}{Q}}{g}}{\jtermeq{\Gamma}{\ctxext{\subst{x}{P}}{\subst{x}{Q}}}{\subst{x}{\ctxext{f}{g}}}{\ctxext{\subst{x}{f}}{\subst{x}{g}}}}\\
+& \inference{\jterm{\ctxext{{\Gamma}{A}}{P}}{Q}{g}\quad\jterm{\ctxext{{\Gamma}{A}}{P}}{\subst{g}{R}}{t}}{\jtermeq{\ctxext{\Gamma}{\subst{x}{P}}}{\ctxext{\subst{x}{Q}}{\subst{x}{R}}}{\subst{x}{\ctxext{g}{t}}}{\ctxext{\subst{x}{g}}{\subst{x}{t}}}}
+\end{align}
+
+Finally, when we substitute by an extended term we get an equal result as when we
+substitute two consecutive times. This looks like the Curry-Howard isomorphism.
+\begin{align}
+& \inference{\jterm{\Gamma}{A}{x}\quad\jterm{\Gamma}{\subst{x}{P}}{u}\quad\jfam{\ctxext{{\Gamma}{A}}{P}}{Q}}
+  {\jfameq{\Gamma}{\subst{\ctxext{x}{u}}{Q}}{\subst{u}{{x}{Q}}}}\\
+& \inference{\jterm{\Gamma}{A}{x}\quad\jterm{\Gamma}{\subst{x}{P}}{u}\quad\jterm{\ctxext{{\Gamma}{A}}{P}}{Q}{g}}
+  {\jtermeq{\Gamma}{\subst{u}{{x}{Q}}}{\subst{\ctxext{x}{u}}{g}}{\subst{u}{{x}{g}}}}\\
+& \inference{\jterm{\Gamma}{A}{x}\quad\jterm{\Gamma}{\subst{x}{P}}{u}\quad\jfam{\ctxext{{{\Gamma}{A}}{P}}{Q}}{R}}
+  {\jfameq{\ctxext{\Gamma}{\subst{u}{{x}{Q}}}}{\subst{\ctxext{x}{u}}{R}}{\subst{u}{{x}{R}}}}\\
+& \inference{\jterm{\Gamma}{A}{x}\quad\jterm{\Gamma}{\subst{x}{P}}{u}\quad\jterm{\ctxext{{{\Gamma}{A}}{P}}{Q}}{R}{t}}
+  {\jtermeq{\ctxext{\Gamma}{\subst{u}{{x}{Q}}}}{\subst{u}{{x}{R}}}{\subst{\ctxext{x}{u}}{t}}{\subst{u}{{x}{t}}}}
+\end{align}
+
 \begin{comment}

### 2014-02-25 article-models-tt.tex
reorganizing the notes

@@
 we obtain a term $\jterm{\Gamma}{\ctxext{A}{P}}{\ctxext{x}{u}}$. This operation
-of weakening on terms is like the pairing operation. However, it is handled
-much stricter for it has a judgmental inverse and it is subject to various compatibility
+of weakening on terms is like the pairing operation and like with the pairing
+operation, extension on terms tells what the terms of an extended family are. 
+However, it is handled
+much stricter than $\smsym$-types, 
+for it has a judgmental inverse and it is subject to various compatibility
 rules.

### 2014-02-25 article-models-tt.tex
extended intro to tt

@@
 no rules added for extension followed by substitution.
+
+In \autoref{extension-on-terms} we come back to the operation of extension, but
+now for terms. The rules in this section explain what it means to be a term of
+an extended family, namely to be an extended term. The compatibility rules for
+extended terms are then similar as the compatibility rules that have appeared
+earlier: there is a compatibility rule for the extension of an extension, for
+the weakening of an extended term, for the substutution of an extended term by
+an arbitrary term and for substitution by an extended term. The last set of
+compatibility rules is a form of the Curry-Howard isomorphism -- this famous
+correspondence naturally comes out of our scheme of compatibility rules.
 
@@
 
-\subsubsection{Extension on terms}
+\subsubsection{Extension on terms}\label{extension-on-terms}
 When $\jterm{\Gamma}{A}{x}$ and $\jterm{\Gamma}{\subst{x}{P}}{u}$ are terms,

### 2014-02-25 article-models-tt.tex
expanding intro of section 3

@@
-\section{Type theory without type constructors}\label{tt}
-In this appendix we give a possible description of type theory without basic
+\section{Type theory before type constructors}\label{tt}
+In this section we give a description of type theory before basic
 constructors and with explicit extension, weakening and substitution. We will
@@
 
+We will develop a theory which is in fact a theory of contexts and families of
+contexts where a family of context can be a type, behind which we have the
+intuition that it is an atomic or irreducible family of contexts. We get immediately
+that every construction can be made in an arbitrary context.
+
 It should be noted that this is just a prototype description. It may be the case
@@
 Much of the rules we state are just compatibility rules of extension, weakening
-and substitution with each other. To get the overview, we list the sections
+and substitution with each other. In a way, these rules assert that our contexts
+are just structureless lists of contexts and that likewise terms are structureless
+lists of terms. They are structureless in the sense that the order in which
+they are formed by pairing up is irrelevant. We note that this causes complications
+in the traditional way that categorical sematics of type theory is implemented,
+where contexts become objects of the category which is supposed to model type
+theory. The reason for this is that context extension will not satisfy all the
+compatibility rules we're about to state. The first step to resolving this is taking
+the types in the empty context as the objects.
+
+To get the overview of the compatibility rules, we list the sections
 where these compatibility rules are described in the following table (in this
@@
 operation $Y\circ X$):
-
 \begin{center}
@@
 \end{center}
-
 Since extension by a context acts trivially on families and terms there are

### 2014-02-25 article-models-tt.tex
extended the rules for identity functions

@@
 \subsubsection{Substitution is compatible with the empty context}
+The fibers of the empty family are the empty families:
+\begin{equation}
+\inference{\jterm{\Gamma}{A}{x}}{\jfameq{\Gamma}{\subst{x}{\emptyf}}{\emptyf}}
+\end{equation}
 
@@
 & \inference{\jfam{\Gamma}{A}}{\jfameq{\Gamma}{\subst{\emptytm}{A}}{A}}\\
+& \inference{\jterm{\Gamma}{A}{x}}{\jtermeq{\Gamma}{A}{\subst{\emptytm}{x}}{x}}\\
 & \inference{\jfam{\ctxext{\Gamma}{A}}{P}}{\jfameq{\ctxext{\Gamma}{A}}{\subst{\emptytm}{P}}{P}}\\
@@
 
-\subsubsection{The empty type is compatible with substitution}
-The fibers of the empty family are the empty families:
-\begin{equation}
-\inference{\jterm{\Gamma}{A}{x}}{\jfameq{\Gamma}{\subst{x}{\emptyf}}{\emptyf}}
-\end{equation}
-
 \subsubsection{Weakening is compatible with substitution}\label{comp-ws}
@@
 
-\subsection{The rules establishing the categorical properties of type theory}
-\subsubsection{The defining property of weakening}
-The judgmental equalities we're about to describe assert that substituting a term
-in the weakening a thing gives you the thing back. In the case of contexts we get that each fiber
-$\subst{x}{\ctxwk{A}{B}}$ is just $B$ and in the case of terms we get 
-that $\ctxwk{A}{y}$ is the constant function
-mapping everything to $y:B$. Thus, these rules actually establish the weakening
-as the weakening. After stating the rules we will describe what it means to
-compose context morphisms (terms of weakened contexts).
-
-\begin{align}
-& \inference{\jfam{\Gamma}{A}\qquad\jfam{\Gamma}{B}\qquad\jterm{\Gamma}{A}{x}}{\jfameq{\Gamma}{\subst{x}{\ctxwk{A}{B}}}{B}}\\
-& \inference{\jfam{\Gamma}{A}\qquad\jfam{\ctxext{\Gamma}{B}}{Q}\qquad\jterm{\Gamma}{A}{x}}{\jfameq{\ctxext{\Gamma}{B}}{\subst{x}{\ctxwk{A}{Q}}}{Q}}\\
-& \inference{\jterm{\Gamma}{A}{x}\qquad\jterm{\Gamma}{B}{y}}{\jtermeq{\Gamma}{B}{\subst{x}{\ctxwk{A}{y}}}{y}}\\
-& \inference{\jterm{\Gamma}{A}{x}\qquad\jterm{\ctxext{\Gamma}{B}}{Q}{g}}{\jtermeq{\ctxext{\Gamma}{B}}{Q}{\subst{x}{\ctxwk{A}{g}}}{g}}
-\end{align}
-
-Using the rules of the compatibility of substitution with weakening and of the
-compatibility of weakening with itself, we see that we can show
-\begin{equation*}
-\jfameq{\Gamma}{\subst{f}{\ctxwk{\Gamma}{{\Delta}{\greek{E}}}}}{\ctxwk{\Gamma}{E}}
-\end{equation*}
-for any three contexts $\Gamma$, $\Delta$ and $\greek{E}$ and any $\jhom{\Gamma}{\Delta}{f}$.
-It follows that for $\jhom{\Delta}{\greek{E}}{g}$ we have
-\begin{equation*}
-\jhom{\Gamma}{\greek{E}}{\subst{f}{\ctxwk{\Gamma}{g}}}
-\end{equation*}
-The term $\jcomp{\Gamma}{f}{g}$ is the composition of $g$ with $f$. We may call
-a term $\jhom{\Gamma}{\Delta}{f}$ a morphism from $\Gamma$ to $\Delta$.
-
-\subsubsection{Identity functions}
-Without a rule explicitly asserting the existence of an identity morphism we don't
-get one, hence we do that here. The identity morphism is a term which introduced
-in ordinary type theory via the variable rule. The variable rule is a bit more
-general: it asserts that
-\begin{equation*}
-\jterm{\Gamma,\,x_1:A_1,\ldots,\,x_n:A_n}{A_i}{x_i}
-\end{equation*}
-for every $1\leq i\leq n$. Thus, it establishes the projections. In our setting,
-we get the projections from the identity morphisms together with weakening. We
-already have weakening, so here it suffices to introduce the identity morphisms.
-They are determined by their behavior with respect to substitution combined with
-weakening.
-
-\begin{align}
-&\inference{\jctx{\Gamma}}{\jterm{\Gamma}{\ctxwk{\Gamma}{\Gamma}}{\idfunc[\Gamma]}}\label{ctxIdFuncForm}\\
-&\inference{\jfam{\Gamma}{A}}{\jfameq{\Gamma}{\subst{\idfunc[\Gamma]}{\ctxwk{\Gamma}{A}}}{A}}\label{ctxIdFuncPbInvType}\\
-&\inference{\jterm{\Gamma}{A}{x}}{\jtermeq{\Gamma}{A}{\subst{\idfunc[\Gamma]}{\ctxwk{\Gamma}{x}}}{x}}\label{ctxIdFuncPbInvTerm}\\
-&\inference{\jterm{\Delta}{\ctxwk{\Delta}{\Gamma}}{g}}{\jtermeq{\Delta}{\ctxwk{\Delta}{\Gamma}}{\subst{g}{\ctxwk{\Delta}{\idfunc[\Gamma]}}}{g}}\label{ctxIdFuncSubstInvTerm}
-\end{align}
-
-The identity morphisms for types are introduced analogously to the identity
-morphisms of contexts, the only difference being that we have a rule asserting
-that $\subst{x}{\idfunc[A]}\jdeq x$ for every term $x:A$ in context $\Gamma$.
-
-\begin{align}
-& \inference{\jfam{\Gamma}{A}}{\jterm{\ctxext{\Gamma}{A}}{\ctxwk{A}{A}}{\idfunc[A]}}\\
-& \inference{\jterm{\Gamma}{A}{x}}{\jtermeq{\Gamma}{A}{\subst{x}{\idfunc[A]}}{x}}\\
-& \inference{\jfam{\ctxext{\Gamma}{A}}{P}}{\jfameq{\ctxext{\Gamma}{A}}{\subst{\idfunc[A]}{\ctxwk{A}{P}}}{P}}\\
-& \inference{\jterm{\ctxext{\Gamma}{A}}{P}{f}}{\jtermeq{\ctxext{\Gamma}{A}}{P}{\subst{\idfunc[A]}{\ctxwk{A}{f}}}{f}} \\
-& \inference{\jterm{\ctxext{\Gamma}{B}}{\ctxwk{B}{A}}{g}}{\jtermeq{\ctxext{\Gamma}{B}}{\ctxwk{B}{A}}{\subst{g}{\ctxwk{B}{\idfunc[A]}}}{g}}
-\end{align}
-
-Using these rules we can derive that there is a judgmental equality $\jfameq{\ctxext{\Gamma}{A}}{\subst{x}{{\idfunc[A]}{B}}}
-{\subst{x}{{x}{B}}}$. Note that $\subst{x}{{x}{B}}$ may just be denoted by
-$\subst{x,x}{B}$.
-
-The identity functions in combination with weakening give rise to the projections.
-We define
-
-\begin{equation*}
-\proj A \defeq \ctxwk{A}{\idfunc[\Gamma]}:\ctxwk{A}{{\Gamma}{\Gamma}}.
-\end{equation*}
-
-\subsubsection{Extension on terms}\label{extension-on-terms}
+\subsection{Extension on terms}\label{extension-on-terms}
 When $\jterm{\Gamma}{A}{x}$ and $\jterm{\Gamma}{\subst{x}{P}}{u}$ are terms,
@@
 
-\begin{comment}
-\subsubsection{Inclusions}
-In ordinary type theory, there is a rule
+\subsection{Composition and identity functions}
+\subsubsection{The defining property of weakening}
+The judgmental equalities we're about to describe assert that substituting a term
+in the weakening a thing gives you the thing back. In the case of contexts we get that each fiber
+$\subst{x}{\ctxwk{A}{B}}$ is just $B$ and in the case of terms we get 
+that $\ctxwk{A}{y}$ is the constant function
+mapping everything to $y:B$. Thus, these rules actually establish the weakening
+as the weakening. After stating the rules we will describe what it means to
+compose context morphisms (terms of weakened contexts).
+
+\begin{align}
+& \inference{\jfam{\Gamma}{A}\qquad\jfam{\Gamma}{B}\qquad\jterm{\Gamma}{A}{x}}{\jfameq{\Gamma}{\subst{x}{\ctxwk{A}{B}}}{B}}\\
+& \inference{\jfam{\Gamma}{A}\qquad\jfam{\ctxext{\Gamma}{B}}{Q}\qquad\jterm{\Gamma}{A}{x}}{\jfameq{\ctxext{\Gamma}{B}}{\subst{x}{\ctxwk{A}{Q}}}{Q}}\\
+& \inference{\jterm{\Gamma}{A}{x}\qquad\jterm{\Gamma}{B}{y}}{\jtermeq{\Gamma}{B}{\subst{x}{\ctxwk{A}{y}}}{y}}\\
+& \inference{\jterm{\Gamma}{A}{x}\qquad\jterm{\ctxext{\Gamma}{B}}{Q}{g}}{\jtermeq{\ctxext{\Gamma}{B}}{Q}{\subst{x}{\ctxwk{A}{g}}}{g}}
+\end{align}
+
+Using the rules of the compatibility of substitution with weakening and of the
+compatibility of weakening with itself, we see that we can show
 \begin{equation*}
-\inference{\jterm{\Gamma}{A}{x}\qquad\jfam{\Gamma,\,a:A}{P(a)}}
-  {\Gamma,\,u:P(x/a)\vdash x:A, u:P(x/a)}
+\jfameq{\Gamma}{\subst{f}{\ctxwk{\Gamma}{{\Delta}{\greek{E}}}}}{\ctxwk{\Gamma}{E}}
 \end{equation*}
-This rule gives the inclusion of the fiber $\subst{x}{P}$ into $\ctxext{A}{P}$.
+for any three contexts $\Gamma$, $\Delta$ and $\greek{E}$ and any $\jhom{\Gamma}{\Delta}{f}$.
+It follows that for $\jhom{\Delta}{\greek{E}}{g}$ we have
+\begin{equation*}
+\jhom{\Gamma}{\greek{E}}{\subst{f}{\ctxwk{\Gamma}{g}}}
+\end{equation*}
+The term $\jcomp{\Gamma}{f}{g}$ is the composition of $g$ with $f$. We may call
+a term $\jhom{\Gamma}{\Delta}{f}$ a morphism from $\Gamma$ to $\Delta$.
 
-Thus, we require
+\subsubsection{Identity functions}
+Without a rule explicitly asserting the existence of an identity morphism we don't
+get one, hence we do that here. The identity morphism is a term which introduced
+in ordinary type theory via the variable rule. The variable rule is a bit more
+general: it asserts that
+\begin{equation*}
+\jterm{\Gamma,\,x_1:A_1,\ldots,\,x_n:A_n}{A_i}{x_i}
+\end{equation*}
+for every $1\leq i\leq n$. Thus, it establishes the projections. In our setting,
+we get the projections from the identity morphisms together with weakening. We
+already have weakening, so here it suffices to introduce the identity morphisms.
+They are determined by their behavior with respect to substitution combined with
+weakening. The identity functions will also be subject to compatibility rules.
+\begin{align}
+& \inference{\jfam{\Gamma}{A}}{\jterm{\ctxext{\Gamma}{A}}{\ctxwk{A}{A}}{\idfunc[A]}}\\
+& \inference{\jterm{\Gamma}{A}{x}}{\jtermeq{\Gamma}{A}{\subst{x}{\idfunc[A]}}{x}}\\
+& \inference{\jfam{\ctxext{\Gamma}{A}}{P}}{\jfameq{\ctxext{\Gamma}{A}}{\subst{\idfunc[A]}{\ctxwk{A}{P}}}{P}}\\
+& \inference{\jterm{\ctxext{\Gamma}{A}}{P}{f}}{\jtermeq{\ctxext{\Gamma}{A}}{P}{\subst{\idfunc[A]}{\ctxwk{A}{f}}}{f}} \\
+& \inference{\jterm{\ctxext{\Gamma}{B}}{\ctxwk{B}{A}}{g}}{\jtermeq{\ctxext{\Gamma}{B}}{\ctxwk{B}{A}}{\subst{g}{\ctxwk{B}{\idfunc[A]}}}{g}}
+\end{align}
+
+Using these rules we can derive that there is a judgmental equality 
+$\jfameq{\Gamma}{\subst{x}{{\idfunc[A]}{R}}}
+{\subst{x}{{x}{R}}}$ for any $\jfam{\ctxext{{\Gamma}{A}}{\ctxwk{A}{A}}}{R}$. 
+Note that $\subst{x}{{x}{R}}$ is judgmentatlly the same as
+$\subst{\ctxext{x}{x}}{R}$.
+
+The identity functions in combination with weakening give rise to the projections.
+We define the term
+\begin{equation*}
+\proj P \defeq \ctxwk{P}{\idfunc[A]}:\ctxwk{P}{{A}{A}}
+\end{equation*}
+in context $\ctxext{{\Gamma}{A}}{P}$.
+
+The identity function of an extended family is the extension of the identity
+functions on the components in the apropriate way. To find out what the
+apropriate way is, note that
+\begin{align*}
+\ctxwk{\ctxext{A}{P}}{\ctxext{A}{P}} & \jdeq \ctxwk{P}{{A}{\ctxext{A}{P}}}\\
+& \jdeq \ctxext{\ctxwk{P}{{A}{A}}}{\ctxwk{P}{{A}{P}}}
+\end{align*}
+We have the term $\jterm{\ctxext{\Gamma}{{A}{P}}}{\ctxwk{P}{A}}{\ctxwk{P}{\idfunc[A]}}$.
+Thus we need to find out what $\subst{\ctxwk{P}{\idfunc[A]}}{\ctxwk{P}{{A}{P}}}$ is:
+\begin{align*}
+\subst{\ctxwk{P}{\idfunc[A]}}{\ctxwk{P}{{A}{P}}} & \jdeq \ctxwk{P}{\subst{\idfunc[A]}{\ctxwk{A}{P}}}\\
+& \jdeq \ctxwk{P}{P},
+\end{align*}
+where we find the term $\idfunc[P]$. Therefore, we have the following inference rule:
 \begin{equation}
-\inference{\jterm{\Gamma}{A}{x}\qquad\jfam{\ctxext{\Gamma}{A}}{P}}
-  {\jterm{\ctxext{\Gamma}{\subst{x}{P}}}{\ctxwk{\subst{x}{P}}{\ctxext{A}{P}}}{\typefont{in}^P_x}}
+\inference{\jfam{\Gamma}{A}\quad\jfam{\ctxext{\Gamma}{A}}{P}}{\jtermeq{\ctxext{\Gamma}{{A}{P}}}{\ctxwk{\ctxext{A}{P}}{\ctxext{A}{P}}}{\idfunc[\ctxext{A}{P}]}{\ctxext{\ctxwk{P}{\idfunc[A]}}{\idfunc[P]}}}
 \end{equation}
 
-The terms $\typefont{in}^P_x$ have the following behavior:
+The identity function of a weakened family is the weakened identity function:
+\begin{equation}
+\inference{\jfam{\Gamma}{A}\quad\jfam{\Gamma}{B}}{\jtermeq{\ctxext{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{{A}{B}}{{A}{B}}}{\idfunc[\ctxwk{A}{B}]}{\ctxwk{A}{\idfunc[B]}}}
+\end{equation}
 
-\begin{align}
-& \inference
-    {\jfam{\ctxext{{\Gamma}{A}}{P}}{Q}}
-    {\jfameq{\ctxext{\Gamma}{\subst{x}{P}}}{\subst{\typefont{in}^P_x}{\ctxwk{\subst{x}{P}}{Q}}}{\subst{x}{Q}}}\\
-& \inference
-    {\jterm{\ctxext{{\Gamma}{A}}{P}}{Q}{g}}
-    {\jtermeq{\ctxext{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}{\subst{\typefont{in}^P_x}{\ctxwk{\subst{x}{P}}{g}}}{\subst{x}{g}}}
-\end{align}
-\end{comment}
+The identity function of a substituted family is the substitution of the identity function
+\begin{equation}
+\inference{\jterm{\Gamma}{A}{x}\quad\jfam{\ctxext{\Gamma}{A}}{P}}{\jtermeq{\ctxext{\Gamma}{\subst{x}{P}}}{\ctxwk{\subst{x}{P}}{\subst{x}{P}}}{\idfunc[\subst{x}{P}]}{\subst{x}{\idfunc[P]}}}
+\end{equation}

### 2014-02-25 article-models-tt.tex
small improvement

@@
 & \inference{\jfam{\Gamma}{A}\qquad\jfam{\Gamma}{B}\qquad\jterm{\Gamma}{A}{x}}{\jfameq{\Gamma}{\subst{x}{\ctxwk{A}{B}}}{B}}\\
+& \inference{\jterm{\Gamma}{A}{x}\qquad\jterm{\Gamma}{B}{y}}{\jtermeq{\Gamma}{B}{\subst{x}{\ctxwk{A}{y}}}{y}}\\
 & \inference{\jfam{\Gamma}{A}\qquad\jfam{\ctxext{\Gamma}{B}}{Q}\qquad\jterm{\Gamma}{A}{x}}{\jfameq{\ctxext{\Gamma}{B}}{\subst{x}{\ctxwk{A}{Q}}}{Q}}\\
-& \inference{\jterm{\Gamma}{A}{x}\qquad\jterm{\Gamma}{B}{y}}{\jtermeq{\Gamma}{B}{\subst{x}{\ctxwk{A}{y}}}{y}}\\
 & \inference{\jterm{\Gamma}{A}{x}\qquad\jterm{\ctxext{\Gamma}{B}}{Q}{g}}{\jtermeq{\ctxext{\Gamma}{B}}{Q}{\subst{x}{\ctxwk{A}{g}}}{g}}
@@
 \begin{equation*}
-\jfameq{\Gamma}{\subst{f}{\ctxwk{\Gamma}{{\Delta}{\greek{E}}}}}{\ctxwk{\Gamma}{E}}
+\jfameq{\ctxext{\Gamma}{A}}{\subst{f}{\ctxwk{A}{{B}{C}}}}{\ctxwk{A}{C}}
 \end{equation*}
-for any three contexts $\Gamma$, $\Delta$ and $\greek{E}$ and any $\jhom{\Gamma}{\Delta}{f}$.
-It follows that for $\jhom{\Delta}{\greek{E}}{g}$ we have
+for any three families $A$, $B$ and $C$ in context $\Gamma$ and any $\jterm{\ctxext{\Gamma}{A}}{\ctxwk{A}{B}}{f}$.
+It follows that for $\jterm{\ctxext{\Gamma}{B}}{\ctxwk{B}{C}}{g}$ we have
 \begin{equation*}
-\jhom{\Gamma}{\greek{E}}{\subst{f}{\ctxwk{\Gamma}{g}}}
+\jterm{\ctxext{\Gamma}{A}}{\ctxwk{A}{C}}{\subst{f}{\ctxwk{A}{g}}}
 \end{equation*}
-The term $\jcomp{\Gamma}{f}{g}$ is the composition of $g$ with $f$. We may call
-a term $\jhom{\Gamma}{\Delta}{f}$ a morphism from $\Gamma$ to $\Delta$.
+The term $\jcomp{A}{f}{g}$ is the composition of $g$ with $f$. We may call
+a term $\jterm{\ctxext{\Gamma}{A}}{\ctxwk{A}{B}}{f}$ a morphism from $A$ to $B$.
 

### 2014-02-25 article-models-tt.tex
substitution preserves types
small impro

@@
 \begin{align}
-& \inference{\jtype{\Gamma}{B}}{\jtype{\ctxext{\Gamma}{A}}{\ctxwk{A}{B}}}\\
 & \inference{\jfam{\Gamma}{B}\quad\jtype{\ctxext{\Gamma}{B}}{Q}}{\jtype{\ctxext{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{Q}}}
@@
 \subsubsection{Weakening is compatible with the empty context}
-
+The following rules express that when the empty context or context family is
+weakened, the result is the empty context family.
+\begin{align}
+& \inference{\jctx{\Gamma}}{\jfameq{\Gamma}{\ctxwk{\Gamma}{\emptyc}}{\emptyf}}\\
+& \inference{\jfam{\Gamma}{A}}{\jfameq{\ctxext{\Gamma}{A}}{\ctxwk{A}{\emptyf}}{\emptyf}}
+\end{align}
 Weakening by the empty family $\emptyf$ over a context $\Gamma$ leaves families, 
@@
 & \inference{\jterm{\ctxext{\Gamma}{B}}{Q}{g}}{\jtermeq{\ctxext{\Gamma}{B}}{Q}{\ctxwk{\emptyf}{g}}{g}}
-\end{align}
-
-\subsubsection{The empty context is compatible with weakening}
-The following rules express that when the empty context or context family is
-weakened, the result is the empty context family.
-
-\begin{align}
-& \inference{\jctx{\Gamma}}{\jfameq{\Gamma}{\ctxwk{\Gamma}{\emptyc}}{\emptyf}}\\
-& \inference{\jfam{\Gamma}{A}}{\jfameq{\ctxext{\Gamma}{A}}{\ctxwk{A}{\emptyf}}{\emptyf}}
 \end{align}
@@
 & &\inference{\jtermeq{\ctxext{{\Gamma}{A}}{P}}{Q}{g}{g'}}{\jtermeq{\ctxext{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}{\subst{x}{g}}{\subst{x'}{g'}}}
+\end{align}
+
+We also add the following rule asserting that substitution preserves the
+property of being a type:
+\begin{align}
+& \inference{\jterm{\Gamma}{A}{x}\quad\jtype{\ctxext{{\Gamma}{A}}{P}}{Q}}{\jtype{\ctxext{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}}
 \end{align}

### 2014-02-26 article-models-tt.tex
small impro
resolving issue #23

@@
 of weakening on terms is like the pairing operation and like with the pairing
-operation, extension on terms tells what the terms of an extended family are. 
-However, it is handled
+operation, extension on terms tells what the terms of an extended family are:
+every term of an extended family decomposes into the components of a pair. 
+\begin{align}
+& \inference{\jterm{\Gamma}{A}{x}\quad\jterm{\Gamma}{\subst{x}{P}}{u}}{\jterm{\Gamma}{\ctxext{A}{P}}{\ctxext{x}{u}}}
+& & \inference{\jtermeq{\Gamma}{A}{x}{x'}\quad\jtermeq{\Gamma}{\subst{x}{P}}{u}{u'}}{\jtermeq{\Gamma}{\ctxext{A}{P}}{\ctxext{x}{u}}{\ctxext{x'}{u'}}}\\
+& \inference{\jterm{\Gamma}{\ctxext{A}{P}}{w}}{\jterm{\Gamma}{A}{\pts{w}}}
+& & \inference{\jtermeq{\Gamma}{\ctxext{A}{P}}{w}{w'}}{\jtermeq{\Gamma}{A}{\pts{w}}{\pts{w'}}}\\
+& \inference{\jterm{\Gamma}{\ctxext{A}{P}}{w}}{\jterm{\Gamma}{\subst{\pts{w}}{P}}{\edg{w}}}
+& & \inference{\jtermeq{\Gamma}{\ctxext{A}{P}}{w}{w'}}{\jtermeq{\Gamma}{\subst{\pts{w}}{P}}{\edg{w}}{\edg{w'}}}
+\end{align}
+However, this pairing is handled
 much stricter than $\smsym$-types, 
-for it has a judgmental inverse and it is subject to various compatibility
-rules.
+because the decomposition is a judgmental inverse (and it will be subject to various compatibility
+rules).
 \begin{align}
-& \inference{\jterm{\Gamma}{A}{x}\quad\jterm{\Gamma}{\subst{x}{P}}{u}}{\jterm{\Gamma}{\ctxext{A}{P}}{\ctxext{x}{u}}}\\
-& \inference{\jterm{\Gamma}{\ctxext{A}{P}}{w}}{\jterm{\Gamma}{A}{\pts{w}}}\\
-& \inference{\jterm{\Gamma}{\ctxext{A}{P}}{w}}{\jterm{\Gamma}{\subst{\pts{w}}{P}}{\edg{w}}}\\
 & \inference{\jterm{\Gamma}{\ctxext{A}{P}}{w}}{\jtermeq{\Gamma}{\ctxext{A}{P}}{\ctxext{\pts{w}}{\edg{w}}}{w}}\\
@@
 already have weakening, so here it suffices to introduce the identity morphisms.
-They are determined by their behavior with respect to substitution combined with
+\begin{align}
+& \inference{\jfam{\Gamma}{A}}{\jterm{\ctxext{\Gamma}{A}}{\ctxwk{A}{A}}{\idfunc[A]}}
+& & \inference{\jfameq{\Gamma}{A}{A'}}{\jtermeq{\ctxext{\Gamma}{A}}{\ctxwk{A}{A}}{\idfunc[A]}{\idfunc[A']}}
+\end{align}
+Identity functions are determined by their behavior with respect to substitution combined with
 weakening. The identity functions will also be subject to compatibility rules.
 \begin{align}
-& \inference{\jfam{\Gamma}{A}}{\jterm{\ctxext{\Gamma}{A}}{\ctxwk{A}{A}}{\idfunc[A]}}\\
 & \inference{\jterm{\Gamma}{A}{x}}{\jtermeq{\Gamma}{A}{\subst{x}{\idfunc[A]}}{x}}\\

### 2014-02-26 article-models-tt.tex
lots of work on extension on terms

@@
 &\inference{\jfam{\ctxext{{\Gamma}{A}}{P}}{Q}}
-{\jfameq{\Gamma}{\subst{x}{{f}{Q}}}{\subst{{x}{f}}{{x}{Q}}}}\\
+{\jfameq{\Gamma}{\subst{x}{{f}{Q}}}{\subst{{x}{f}}{{x}{Q}}}}\label{comp-ss-1}\\
+&\inference{\jterm{\ctxext{{\Gamma}{A}}{P}}{Q}{g}}
+{\jtermeq{\Gamma}{\subst{x}{{f}{Q}}}{\subst{x}{{f}{g}}}{\subst{{x}{f}}{{x}{g}}}}\label{comp-ss-2}\\
 &\inference{\jfam{\ctxext{{{\Gamma}{A}}{P}}{Q}}{R}}
-{\jfameq{\ctxext{\Gamma}{\subst{x}{{f}{Q}}}}{\subst{x}{{f}{R}}}{\subst{{x}{f}}{{x}{R}}}}\\
-&\inference{\jterm{\ctxext{{\Gamma}{A}}{P}}{Q}{g}}
-{\jtermeq{\Gamma}{\subst{x}{{f}{Q}}}{\subst{x}{{f}{g}}}{\subst{{x}{f}}{{x}{g}}}}\\
+{\jfameq{\ctxext{\Gamma}{\subst{x}{{f}{Q}}}}{\subst{x}{{f}{R}}}{\subst{{x}{f}}{{x}{R}}}}\label{comp-ss-3}\\
 &\inference{\jterm{\ctxext{{{\Gamma}{A}}{P}}{Q}}{R}{h}}
-{\jtermeq{\ctxext{\Gamma}{\subst{x}{{f}{Q}}}}{\subst{x}{{f}{R}}}{\subst{x}{{f}{h}}}{\subst{{x}{f}}{{x}{h}}}}
+{\jtermeq{\ctxext{\Gamma}{\subst{x}{{f}{Q}}}}{\subst{x}{{f}{R}}}{\subst{x}{{f}{h}}}{\subst{{x}{f}}{{x}{h}}}}\label{comp-ss-4}
 \end{align}
@@
 
-\subsection{Extension on terms}\label{extension-on-terms}
-When $\jterm{\Gamma}{A}{x}$ and $\jterm{\Gamma}{\subst{x}{P}}{u}$ are terms,
-we obtain a term $\jterm{\Gamma}{\ctxext{A}{P}}{\ctxext{x}{u}}$. This operation
-of weakening on terms is like the pairing operation and like with the pairing
-operation, extension on terms tells what the terms of an extended family are:
-every term of an extended family decomposes into the components of a pair. 
-\begin{align}
-& \inference{\jterm{\Gamma}{A}{x}\quad\jterm{\Gamma}{\subst{x}{P}}{u}}{\jterm{\Gamma}{\ctxext{A}{P}}{\ctxext{x}{u}}}
-& & \inference{\jtermeq{\Gamma}{A}{x}{x'}\quad\jtermeq{\Gamma}{\subst{x}{P}}{u}{u'}}{\jtermeq{\Gamma}{\ctxext{A}{P}}{\ctxext{x}{u}}{\ctxext{x'}{u'}}}\\
-& \inference{\jterm{\Gamma}{\ctxext{A}{P}}{w}}{\jterm{\Gamma}{A}{\pts{w}}}
-& & \inference{\jtermeq{\Gamma}{\ctxext{A}{P}}{w}{w'}}{\jtermeq{\Gamma}{A}{\pts{w}}{\pts{w'}}}\\
-& \inference{\jterm{\Gamma}{\ctxext{A}{P}}{w}}{\jterm{\Gamma}{\subst{\pts{w}}{P}}{\edg{w}}}
-& & \inference{\jtermeq{\Gamma}{\ctxext{A}{P}}{w}{w'}}{\jtermeq{\Gamma}{\subst{\pts{w}}{P}}{\edg{w}}{\edg{w'}}}
-\end{align}
-However, this pairing is handled
-much stricter than $\smsym$-types, 
-because the decomposition is a judgmental inverse (and it will be subject to various compatibility
-rules).
-\begin{align}
-& \inference{\jterm{\Gamma}{\ctxext{A}{P}}{w}}{\jtermeq{\Gamma}{\ctxext{A}{P}}{\ctxext{\pts{w}}{\edg{w}}}{w}}\\
-& \inference{\jterm{\Gamma}{A}{x}\quad\jterm{\Gamma}{\subst{x}{P}}{u}}{\jtermeq{\Gamma}{A}{\pts{(\ctxext{x}{u})}}{x}}\\
-& \inference{\jterm{\Gamma}{A}{x}\quad\jterm{\Gamma}{\subst{x}{P}}{u}}{\jtermeq{\Gamma}{\subst{x}{P}}{\edg{(\ctxext{x}{u})}}{u}}
-\end{align}
-
-We add the following compatibility rule for two consecutive term extensions:
-\begin{equation}
-\inference{\jterm{\Gamma}{A}{x}\quad\jterm{\Gamma}{\subst{x}{P}}{u}\quad\jterm{\Gamma}{\subst{\ctxext{x}{u}}{Q}}{v}}
-{\jtermeq{\Gamma}{\ctxext{{A}{P}}{Q}}{\ctxext{x}{{u}{v}}}{\ctxext{{x}{u}}{v}}}
-\end{equation}
-
-When we weaken a term $\ctxext{y}{v}$ of $\ctxext{B}{Q}$ in context $\Gamma$ by
-a type $A$, the term that we get is $\ctxext{\ctxwk{A}{y}}{\ctxwk{A}{v}}$.
-\begin{align}
-& \inference{\jterm{\Gamma}{B}{y}\quad\jterm{\Gamma}{\subst{y}{Q}}{v}}{\jtermeq{\ctxext{\Gamma}{A}}{\ctxwk{A}{\ctxext{B}{Q}}}{\ctxwk{A}{\ctxext{y}{v}}}{\ctxext{\ctxwk{A}{y}}{\ctxwk{A}{v}}}}\\
-& \inference{\jterm{\ctxext{\Gamma}{B}}{Q}{g}\quad\jterm{\ctxext{\Gamma}{B}}{\subst{g}{R}}{t}}{\jtermeq{\ctxext{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{\ctxext{Q}{R}}}{\ctxwk{A}{\ctxext{g}{t}}}{\ctxext{\ctxwk{A}{g}}{\ctxwk{A}{t}}}}
-\end{align}
-
-When we substitute an extended term $\ctxext{f}{g}$ of $\ctxext{P}{Q}$ by a term
-$x$ of $A$, the term that we get is $\ctxext{\subst{x}{f}}{\subst{x}{g}}$.
-\begin{align}
-& \inference{\jterm{\ctxext{\Gamma}{A}}{P}{f}\quad\jterm{\ctxext{\Gamma}{A}}{\subst{f}{Q}}{g}}{\jtermeq{\Gamma}{\ctxext{\subst{x}{P}}{\subst{x}{Q}}}{\subst{x}{\ctxext{f}{g}}}{\ctxext{\subst{x}{f}}{\subst{x}{g}}}}\\
-& \inference{\jterm{\ctxext{{\Gamma}{A}}{P}}{Q}{g}\quad\jterm{\ctxext{{\Gamma}{A}}{P}}{\subst{g}{R}}{t}}{\jtermeq{\ctxext{\Gamma}{\subst{x}{P}}}{\ctxext{\subst{x}{Q}}{\subst{x}{R}}}{\subst{x}{\ctxext{g}{t}}}{\ctxext{\subst{x}{g}}{\subst{x}{t}}}}
-\end{align}
-
-Finally, when we substitute by an extended term we get an equal result as when we
-substitute two consecutive times. This looks like the Curry-Howard isomorphism.
-\begin{align}
-& \inference{\jterm{\Gamma}{A}{x}\quad\jterm{\Gamma}{\subst{x}{P}}{u}\quad\jfam{\ctxext{{\Gamma}{A}}{P}}{Q}}
-  {\jfameq{\Gamma}{\subst{\ctxext{x}{u}}{Q}}{\subst{u}{{x}{Q}}}}\\
-& \inference{\jterm{\Gamma}{A}{x}\quad\jterm{\Gamma}{\subst{x}{P}}{u}\quad\jterm{\ctxext{{\Gamma}{A}}{P}}{Q}{g}}
-  {\jtermeq{\Gamma}{\subst{u}{{x}{Q}}}{\subst{\ctxext{x}{u}}{g}}{\subst{u}{{x}{g}}}}\\
-& \inference{\jterm{\Gamma}{A}{x}\quad\jterm{\Gamma}{\subst{x}{P}}{u}\quad\jfam{\ctxext{{{\Gamma}{A}}{P}}{Q}}{R}}
-  {\jfameq{\ctxext{\Gamma}{\subst{u}{{x}{Q}}}}{\subst{\ctxext{x}{u}}{R}}{\subst{u}{{x}{R}}}}\\
-& \inference{\jterm{\Gamma}{A}{x}\quad\jterm{\Gamma}{\subst{x}{P}}{u}\quad\jterm{\ctxext{{{\Gamma}{A}}{P}}{Q}}{R}{t}}
-  {\jtermeq{\ctxext{\Gamma}{\subst{u}{{x}{Q}}}}{\subst{u}{{x}{R}}}{\subst{\ctxext{x}{u}}{t}}{\subst{u}{{x}{t}}}}
-\end{align}
-
 \subsection{Composition and identity functions}
@@
 \begin{align}
-& \inference{\jfam{\Gamma}{A}\qquad\jfam{\Gamma}{B}\qquad\jterm{\Gamma}{A}{x}}{\jfameq{\Gamma}{\subst{x}{\ctxwk{A}{B}}}{B}}\\
-& \inference{\jterm{\Gamma}{A}{x}\qquad\jterm{\Gamma}{B}{y}}{\jtermeq{\Gamma}{B}{\subst{x}{\ctxwk{A}{y}}}{y}}\\
-& \inference{\jfam{\Gamma}{A}\qquad\jfam{\ctxext{\Gamma}{B}}{Q}\qquad\jterm{\Gamma}{A}{x}}{\jfameq{\ctxext{\Gamma}{B}}{\subst{x}{\ctxwk{A}{Q}}}{Q}}\\
-& \inference{\jterm{\Gamma}{A}{x}\qquad\jterm{\ctxext{\Gamma}{B}}{Q}{g}}{\jtermeq{\ctxext{\Gamma}{B}}{Q}{\subst{x}{\ctxwk{A}{g}}}{g}}
+& \inference{\jfam{\Gamma}{A}\qquad\jfam{\Gamma}{B}\qquad\jterm{\Gamma}{A}{x}}{\jfameq{\Gamma}{\subst{x}{\ctxwk{A}{B}}}{B}}\label{defn-ws-1}\\
+& \inference{\jterm{\Gamma}{A}{x}\qquad\jterm{\Gamma}{B}{y}}{\jtermeq{\Gamma}{B}{\subst{x}{\ctxwk{A}{y}}}{y}}\label{defn-ws-2}\\
+& \inference{\jfam{\Gamma}{A}\qquad\jfam{\ctxext{\Gamma}{B}}{Q}\qquad\jterm{\Gamma}{A}{x}}{\jfameq{\ctxext{\Gamma}{B}}{\subst{x}{\ctxwk{A}{Q}}}{Q}}\label{defn-ws-3}\\
+& \inference{\jterm{\Gamma}{A}{x}\qquad\jterm{\ctxext{\Gamma}{B}}{Q}{g}}{\jtermeq{\ctxext{\Gamma}{B}}{Q}{\subst{x}{\ctxwk{A}{g}}}{g}}\label{defn-ws-4}
 \end{align}
@@
 \begin{align}
-& \inference{\jterm{\Gamma}{A}{x}}{\jtermeq{\Gamma}{A}{\subst{x}{\idfunc[A]}}{x}}\\
-& \inference{\jfam{\ctxext{\Gamma}{A}}{P}}{\jfameq{\ctxext{\Gamma}{A}}{\subst{\idfunc[A]}{\ctxwk{A}{P}}}{P}}\\
-& \inference{\jterm{\ctxext{\Gamma}{A}}{P}{f}}{\jtermeq{\ctxext{\Gamma}{A}}{P}{\subst{\idfunc[A]}{\ctxwk{A}{f}}}{f}} \\
-& \inference{\jterm{\ctxext{\Gamma}{B}}{\ctxwk{B}{A}}{g}}{\jtermeq{\ctxext{\Gamma}{B}}{\ctxwk{B}{A}}{\subst{g}{\ctxwk{B}{\idfunc[A]}}}{g}}
+& \inference{\jterm{\Gamma}{A}{x}}{\jtermeq{\Gamma}{A}{\subst{x}{\idfunc[A]}}{x}}\label{idfunc-subst-defn}\\
+& \inference{\jfam{\ctxext{\Gamma}{A}}{P}}{\jfameq{\ctxext{\Gamma}{A}}{\subst{\idfunc[A]}{\ctxwk{A}{P}}}{P}}\label{idfunc-wk-defn}\\
+& \inference{\jterm{\ctxext{\Gamma}{A}}{P}{f}}{\jtermeq{\ctxext{\Gamma}{A}}{P}{\subst{\idfunc[A]}{\ctxwk{A}{f}}}{f}}\label{idfunc-precomp} \\
+& \inference{\jterm{\ctxext{\Gamma}{B}}{\ctxwk{B}{A}}{g}}{\jtermeq{\ctxext{\Gamma}{B}}{\ctxwk{B}{A}}{\subst{g}{\ctxwk{B}{\idfunc[A]}}}{g}}\label{idfunc-postcomp}
 \end{align}
-
-Using these rules we can derive that there is a judgmental equality 
-$\jfameq{\Gamma}{\subst{x}{{\idfunc[A]}{R}}}
-{\subst{x}{{x}{R}}}$ for any $\jfam{\ctxext{{\Gamma}{A}}{\ctxwk{A}{A}}}{R}$. 
-Note that $\subst{x}{{x}{R}}$ is judgmentatlly the same as
-$\subst{\ctxext{x}{x}}{R}$.
-
-The identity functions in combination with weakening give rise to the projections.
-We define the term
-\begin{equation*}
-\proj P \defeq \ctxwk{P}{\idfunc[A]}:\ctxwk{P}{{A}{A}}
-\end{equation*}
-in context $\ctxext{{\Gamma}{A}}{P}$.
 
@@
 where we find the term $\idfunc[P]$. Therefore, we have the following inference rule:
-\begin{equation}
-\inference{\jfam{\Gamma}{A}\quad\jfam{\ctxext{\Gamma}{A}}{P}}{\jtermeq{\ctxext{\Gamma}{{A}{P}}}{\ctxwk{\ctxext{A}{P}}{\ctxext{A}{P}}}{\idfunc[\ctxext{A}{P}]}{\ctxext{\ctxwk{P}{\idfunc[A]}}{\idfunc[P]}}}
+\begin{equation}\label{idfunc-ext-comp}
+\inference{\jfam{\Gamma}{A}\quad\jfam{\ctxext{\Gamma}{A}}{P}}{\jtermeq{\ctxext{\Gamma}{{A}{P}}}{\ctxwk{\ctxext{A}{P}}{\ctxext{A}{P}}}{\idfunc[\ctxext{A}{P}]}{\subst{\idfunc[P]}{{\ctxwk{P}{\idfunc[A]}}{\idfunc[\ctxwk{\ctxext{A}{P}}{\ctxext{A}{P}}]}}}}
 \end{equation}
@@
 The identity function of a weakened family is the weakened identity function:
-\begin{equation}
+\begin{equation}\label{idfunc-wk-comp}
 \inference{\jfam{\Gamma}{A}\quad\jfam{\Gamma}{B}}{\jtermeq{\ctxext{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{{A}{B}}{{A}{B}}}{\idfunc[\ctxwk{A}{B}]}{\ctxwk{A}{\idfunc[B]}}}
@@
 The identity function of a substituted family is the substitution of the identity function
-\begin{equation}
+\begin{equation}\label{idfunc-subst-comp}
 \inference{\jterm{\Gamma}{A}{x}\quad\jfam{\ctxext{\Gamma}{A}}{P}}{\jtermeq{\ctxext{\Gamma}{\subst{x}{P}}}{\ctxwk{\subst{x}{P}}{\subst{x}{P}}}{\idfunc[\subst{x}{P}]}{\subst{x}{\idfunc[P]}}}
 \end{equation}
+
+\subsection{Extension on terms}\label{extension-on-terms}
+When $\jterm{\Gamma}{A}{x}$ and $\jterm{\Gamma}{\subst{x}{P}}{u}$ are terms,
+we obtain the term $\jterm{\Gamma}{\ctxext{A}{P}}{\subst{u}{{x}{\idfunc[\ctxext{A}{P}]}}}$. 
+We define 
+\begin{equation*}
+\ctxext{x}{u}\defeq\subst{u}{{x}{\idfunc[\ctxext{A}{P}]}}.
+\end{equation*}
+Thus
+$\ctxext{x}{u}$ is the pairing of $x$ and $u$. Note that because
+$\ctxwk{P}{{A}{\ctxext{A}{P}}}\jdeq\ctxwk{\ctxext{A}{P}}{\ctxext{A}{P}}$, the
+pairing function is just $\idfunc[\ctxext{A}{P}]$.
+
+There is also a decomposition of terms $\jterm{\Gamma}{\ctxext{A}{P}}{w}$ into
+components: the projections. We define
+\begin{align*}
+\cprojfstf{A}{P} & \defeq \ctxwk{P}{\idfunc[A]}\\
+\cprojsndf{A}{P} & \defeq \idfunc[P]
+\end{align*}
+
+The constructions of the terms $\ctxext{x}{u}$ and $\cprojfst{A}{P}{w}$ and
+$\cprojsnd{A}{P}{w}$ are subject to various rules, with all of them being
+consequences of earlier introduced inference rules.
+
+\begin{lem} The following inference rules expressing that pairing is a strict
+inverse to the combination of decompositions, are valid:
+\begin{align*}
+& \inference{\jterm{\Gamma}{\ctxext{A}{P}}{w}}{\jtermeq{\Gamma}{\ctxext{A}{P}}{\ctxext{\cprojfst{A}{P}{w}}{\cprojsnd{A}{P}{w}}}{w}}\\
+& \inference{\jterm{\Gamma}{A}{x}\quad\jterm{\Gamma}{\subst{x}{P}}{u}}{\jtermeq{\Gamma}{A}{\cprojfst{A}{P}{\ctxext{x}{u}}}{x}}\\
+& \inference{\jterm{\Gamma}{A}{x}\quad\jterm{\Gamma}{\subst{x}{P}}{u}}{\jtermeq{\Gamma}{\subst{x}{P}}{\cprojsnd{A}{P}{\ctxext{x}{u}}}{u}}
+\end{align*}
+\end{lem}
+
+\begin{proof}
+To prove the first judgmental equality, note that
+\begin{align*}
+w & \jdeq \subst{w}{\idfunc[\ctxext{A}{P}]} \tag{by \autoref{idfunc-subst-defn}}\\
+& \jdeq \subst{w}{\subst{\idfunc[P]}{{\ctxwk{P}{\idfunc[A]}}{\idfunc[\ctxwk{\ctxext{A}{P}}{\ctxext{A}{P}}]}}}
+  \tag{by \autoref{idfunc-ext-comp}}\\
+& \jdeq \subst{{w}{\idfunc[P]}}{\subst{w}{{\ctxwk{P}{\idfunc[A]}}{\idfunc[\ctxwk{\ctxext{A}{P}}{\ctxext{A}{P}}]}}}\tag{by \autoref{comp-ss-4}}\\
+& \jdeq \subst{{w}{\idfunc[P]}}{{\subst{w}{\ctxwk{P}{\idfunc[A]}}}{\subst{w}{\idfunc[\ctxwk{\ctxext{A}{P}}{\ctxext{A}{P}}]}}}\tag{by \autoref{comp-ss-2}}\\
+& \jdeq \subst{{w}{\idfunc[P]}}{{\subst{w}{\ctxwk{P}{\idfunc[A]}}}{\subst{w}{\ctxwk{\ctxext{A}{P}}{\idfunc[\ctxext{A}{P}]}}}}\tag{by \autoref{idfunc-wk-comp}}\\
+& \jdeq \subst{{w}{\idfunc[P]}}{{\subst{w}{\ctxwk{P}{\idfunc[A]}}}{\idfunc[\ctxext{A}{P}]}}\tag{by \autoref{defn-ws-2}}\\
+& \jdeq \ctxext{\cprojfst{A}{P}{w}}{\cprojsnd{A}{P}{w}}\tag{by definition}
+\end{align*}
+\end{proof}
+
+\begin{lem}
+The following compatibility rules for two consecutive term extensions is valid:
+\begin{align*}
+& \inference{\jterm{\Gamma}{A}{x}\quad\jterm{\Gamma}{\subst{x}{P}}{u}\quad\jterm{\Gamma}{\subst{\ctxext{x}{u}}{Q}}{v}}
+{\jtermeq{\Gamma}{\ctxext{{A}{P}}{Q}}{\ctxext{x}{{u}{v}}}{\ctxext{{x}{u}}{v}}}\\
+& \inference{\jterm{\Gamma}{\ctxext{{A}{P}}{Q}}{w}}
+  {\jtermeq{\Gamma}{A}{\cprojfst{A}{P}{\cprojfst{\ctxext{A}{P}}{Q}{w}}}{\cprojfst{A}{\ctxext{P}{Q}}{w}}}\\
+& \inference{\jterm{\Gamma}{\ctxext{{A}{P}}{Q}}{w}}
+  {\jtermeq{\Gamma}{\subst{\cprojfst{A}{\ctxext{P}{Q}}{w}}{P}}
+  {\cprojsnd{A}{P}{\cprojfst{\ctxext{A}{P}}{Q}{w}}}
+  {\cprojfst{P}{Q}{\cprojsnd{A}{\ctxext{P}{Q}}{w}}}}\\
+& \inference{\jterm{\Gamma}{\ctxext{{A}{P}}{Q}}{w}}
+  {\jtermeq{\Gamma}{\subst{\cprojfst{\ctxext{A}{P}}{Q}{w}}{Q}}
+  {\cprojsnd{P}{Q}{\cprojsnd{A}{\ctxext{P}{Q}}{w}}}
+  {\cprojsnd{\ctxext{A}{P}}{Q}{w}}}
+\end{align*}
+\end{lem}
+
+\begin{lem}
+When we weaken a term $\ctxext{y}{v}$ of $\ctxext{B}{Q}$ in context $\Gamma$ by
+a type $A$, the term that we get is $\ctxext{\ctxwk{A}{y}}{\ctxwk{A}{v}}$. More
+precisely, the following inference rules are valid:
+\begin{align*}
+& \inference{\jterm{\ctxext{\Gamma}{B}}{Q}{g}\quad\jterm{\ctxext{\Gamma}{B}}{\subst{g}{R}}{t}}{\jtermeq{\ctxext{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{\ctxext{Q}{R}}}{\ctxwk{A}{\ctxext{g}{t}}}{\ctxext{\ctxwk{A}{g}}{\ctxwk{A}{t}}}}
+% \\
+%& \inference{\jterm{\Gamma}{\ctxwk{A}{\ctxext{B}{Q}}}{\ctxwk{A}{w}}}{\jtermeq{\ctxext{\Gamma}{A}}{\ctxwk{A}{B}}{\cprojfst{\ctxwk{A}{B}}{\ctxwk{A}{Q}}{\ctxwk{A}{w}}}{\ctxwk{A}{(\cprojfst{B}{Q}{w})}}}\\
+\end{align*}
+\end{lem}
+
+\begin{lem}
+When we substitute an extended term $\ctxext{f}{g}$ of $\ctxext{P}{Q}$ by a term
+$x$ of $A$, the term that we get is $\ctxext{\subst{x}{f}}{\subst{x}{g}}$.
+More precisely, the following inference rules are valid:
+\begin{align*}
+& \inference{\jterm{\ctxext{\Gamma}{A}}{P}{f}\quad\jterm{\ctxext{\Gamma}{A}}{\subst{f}{Q}}{g}}{\jtermeq{\Gamma}{\ctxext{\subst{x}{P}}{\subst{x}{Q}}}{\subst{x}{\ctxext{f}{g}}}{\ctxext{\subst{x}{f}}{\subst{x}{g}}}}\\
+& \inference{\jterm{\ctxext{{\Gamma}{A}}{P}}{Q}{g}\quad\jterm{\ctxext{{\Gamma}{A}}{P}}{\subst{g}{R}}{t}}{\jtermeq{\ctxext{\Gamma}{\subst{x}{P}}}{\ctxext{\subst{x}{Q}}{\subst{x}{R}}}{\subst{x}{\ctxext{g}{t}}}{\ctxext{\subst{x}{g}}{\subst{x}{t}}}}
+\end{align*}
+\end{lem}
+Finally, when we substitute by an extended term we get an equal result as when we
+substitute two consecutive times. This looks like the Curry-Howard isomorphism.
+
+\begin{lem}
+The following inference rules are valid:
+\begin{align*}
+& \inference{\jterm{\Gamma}{A}{x}\quad\jterm{\Gamma}{\subst{x}{P}}{u}\quad\jfam{\ctxext{{\Gamma}{A}}{P}}{Q}}
+  {\jfameq{\Gamma}{\subst{\ctxext{x}{u}}{Q}}{\subst{u}{{x}{Q}}}}\\
+& \inference{\jterm{\Gamma}{A}{x}\quad\jterm{\Gamma}{\subst{x}{P}}{u}\quad\jterm{\ctxext{{\Gamma}{A}}{P}}{Q}{g}}
+  {\jtermeq{\Gamma}{\subst{u}{{x}{Q}}}{\subst{\ctxext{x}{u}}{g}}{\subst{u}{{x}{g}}}}\\
+& \inference{\jterm{\Gamma}{A}{x}\quad\jterm{\Gamma}{\subst{x}{P}}{u}\quad\jfam{\ctxext{{{\Gamma}{A}}{P}}{Q}}{R}}
+  {\jfameq{\ctxext{\Gamma}{\subst{u}{{x}{Q}}}}{\subst{\ctxext{x}{u}}{R}}{\subst{u}{{x}{R}}}}\\
+& \inference{\jterm{\Gamma}{A}{x}\quad\jterm{\Gamma}{\subst{x}{P}}{u}\quad\jterm{\ctxext{{{\Gamma}{A}}{P}}{Q}}{R}{t}}
+  {\jtermeq{\ctxext{\Gamma}{\subst{u}{{x}{Q}}}}{\subst{u}{{x}{R}}}{\subst{\ctxext{x}{u}}{t}}{\subst{u}{{x}{t}}}}
+\end{align*}
+\end{lem}

### 2014-02-28 article-models-tt.tex
rewritten part of intro

@@
 formulate the type theory in such a way that contexts aren't lists of variable
-declarations. The reasons for this are twofold: (1) in the models of type theory
-such as a univalent universe or the graph model, contexts are types and graphs
-respectively -- they are objects rather than lists of variable declarations; we
-intend to give a general algebraic description of higher categories. Besides,
-we will not have to be burdened with comments about variables being bounded or not, or fresh or
-not occuring at all.
+declarations. The main reason is that we don't see variable declarations in the
+models either. This way we also set out to a more algebraic approach of type
+theory and higher category theory. Thirdly, we will not have to be burdened with
+comments about variables being bounded or not, or fresh or not occuring at all.
 
@@
 When we weaken a term $\ctxext{y}{v}$ of $\ctxext{B}{Q}$ in context $\Gamma$ by
-a type $A$, the term that we get is $\ctxext{\ctxwk{A}{y}}{\ctxwk{A}{v}}$. More
+a family $A$, the term that we get is $\ctxext{\ctxwk{A}{y}}{\ctxwk{A}{v}}$. More
 precisely, the following inference rules are valid:

### 2014-02-28 article-models-tt.tex
fixed misuses of the word type

@@
 
-Thus we make eight types of judgments in our type theory: ``$\Gamma$ is a context'',
+Thus we make eight kinds of judgments in our type theory: ``$\Gamma$ is a context'',
 ``$A$ is a family of contexts over $\Gamma$'', ``$A$ is a type in context $\Gamma$''
@@
 \subsection{Extension}
-We introduce extension which not only extends a context $\Gamma$ and a type
-$A$ over it to a context $\ctxext{\Gamma}{A}$, but which also extends a type $A$
-in context $\Gamma$ and a family $P$ over it to a type $\ctxext{A}{P}$ in context
+We introduce extension which not only extends a context $\Gamma$ and a family
+$A$ over it to a context $\ctxext{\Gamma}{A}$, but which also extends a family $A$
+in context $\Gamma$ and a family $P$ over it to a family $\ctxext{A}{P}$ over context
 $\Gamma$. We do this to ensure that all of type theory can be done in a context.
 For instance, we could say (1) that a context in context $\Gamma$ is the same thing
-as a type in context $\Gamma$; (2) When $A$ is a context in this sense, a type in
-context $A$ is the same thing as a family $P$ over $A$ and (3) when $P$ is a type
-in context $A$ in this sense, a term of $P$ keeps its original meaning.
+as a family over $\Gamma$; (2) When $A$ is a context in this sense, a family over
+$A$ is the same thing as a family $P$ over $\ctxext{\Gamma}{A}$ and 
+(3) when $P$ is a family over $A$ in this sense, a term of $P$ keeps its original meaning.
 
@@
 that contexts are unstructured lists of type declarations. This rule is
-unavoidable if we want that for a type $A$ in context $\Gamma$, a type
-in context $A$ is the same thing as a type in context $\ctxext{\Gamma}{A}$.
+unavoidable if we want that for a family $A$ in context $\Gamma$, a family over
+$A$ is the same thing as a family over $\ctxext{\Gamma}{A}$.
 
@@
 this section, we shall first introduce the operations `substitution of a term $x$'
-for types, terms and families. Then we shall explain how substitution interacts
+for families $P$ over $\ctxext{\Gamma}{A}$, terms $f$ of those, families $Q$ over
+$\ctxext{{\Gamma}{A}}{P}$ and terms $g$ of those. 
+Then we shall explain how substitution interacts
 with itself, extension and weakening.

### 2014-02-28 article-models-tt.tex
typos
getting the theory down to 60 rules

@@
 \begin{align*}
-& \inference{\jfam{\Gamma}{A}}{\jfam{\Gamma}{A}} & & \inference{\jfameq{\Gamma}{A}{B}}{\jfameq{\Gamma}{A}{B}}\\
-& \inference{\jfam{\Gamma}{A}\quad\jfameq{\Gamma}{A}{B}}{\jfam{\Gamma}{B}}
-& & \inference{\jfam{\Gamma}{A}\quad\jfameq{\Gamma}{A}{B}}{\jfameq{\Gamma}{A}{B}}
+& \inference{\jtype{\Gamma}{A}}{\jfam{\Gamma}{A}} & & \inference{\jtypeeq{\Gamma}{A}{B}}{\jfameq{\Gamma}{A}{B}}\\
+& \inference{\jtype{\Gamma}{A}\quad\jfameq{\Gamma}{A}{B}}{\jtype{\Gamma}{B}}
+& & \inference{\jtype{\Gamma}{A}\quad\jfameq{\Gamma}{A}{B}}{\jtypeeq{\Gamma}{A}{B}}
 \end{align*}
@@
 & \inference{}{\jctx{\emptyc}}\\
-& \inference{\jctx{\Gamma}}{\jfam{\Gamma}{\emptyf\Gamma}}\\
+& \inference{\jctx{\Gamma}}{\jfam{\Gamma}{\emptyf[\Gamma]}}\\
 & \inference{\jctx{\Gamma}}{\jterm{\Gamma}{\emptyf[\Gamma]}{\emptytm[\Gamma]}}\\
@@
 & \inference
-  {\jfam{\ctxext{\Gamma}{B}}{Q}}
-  {\jfameq{\ctxext{\Gamma}{A}}{\ctxwk{A}{\ctxext{B}{Q}}}{\ctxext{\ctxwk{A}{B}}{\ctxwk{A}{Q}}}}\\
-& \inference
   {\jfam{\ctxext{{\Gamma}{B}}{Q}}{R}}
@@
 \begin{align}
-& \inference{\jfam{\Gamma}{A}\qquad\jfam{\Gamma}{B}\qquad\jfam{\Gamma}{C}}
-          {\jfameq{\ctxext{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{{B}{C}}}
-            {\ctxwk{{A}{B}}{{A}{C}}}}\\
-& \inference{\jfam{\Gamma}{A}\qquad\jfam{\Gamma}{B}\qquad\jterm{\Gamma}{C}{z}}
-          {\jtermeq{\ctxext{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{{A}{B}}{{A}{C}}}
-            {\ctxwk{A}{{B}{z}}}
-            {\ctxwk{{A}{B}}{{A}{z}}}}\label{comp-ww-inf2}\\
 & \inference{\jfam{\Gamma}{A}\qquad\jfam{\Gamma}{B}\qquad\jfam{\ctxext{\Gamma}{C}}{R}}
@@
             {\ctxwk{{A}{B}}{{A}{R}}}}\\
-& \inference{\jfam{\Gamma}{A}\qquad\jfam{\Gamma}{B}\qquad\jterm{\Gamma}{R}{t}}
+& \inference{\jfam{\Gamma}{A}\qquad\jfam{\Gamma}{B}\qquad\jterm{\ctxext{\Gamma}{C}}{R}{t}}
           {\jtermeq{\ctxext{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{{A}{B}}{{A}{C}}}}
@@
 \end{align}
-Note that strictly speaking, we did not need to require the inference in
-\autoref{comp-ww-inf2} since this follows from the inference in \autoref{comp-ww-inf4}
-and the fact that the weakening of the empty family is again the empty family.
 
@@
 \begin{align}
-& \inference{\jfam{\Gamma}{B}}
-    {\jfameq{\ctxext{{\Gamma}{A}}{P}}{\ctxwk{\ctxext{A}{P}}{B}}{\ctxwk{P}{{A}{B}}}}\\
-& \inference{\jterm{\Gamma}{B}{y}}
-    {\jtermeq{\ctxext{{\Gamma}{A}}{P}}{\ctxwk{P}{{A}{B}}}{\ctxwk{\ctxext{A}{P}}{y}}{\ctxwk{P}{{A}{y}}}}\\
 & \inference{\jfam{\ctxext{\Gamma}{B}}{Q}}
@@
 The fibers of the empty family are the empty families:
-\begin{equation}
-\inference{\jterm{\Gamma}{A}{x}}{\jfameq{\Gamma}{\subst{x}{\emptyf}}{\emptyf}}
-\end{equation}
+\begin{align}
+& \inference{\jterm{\Gamma}{A}{x}}{\jfameq{\Gamma}{\subst{x}{\emptyf}}{\emptyf}}\\
+& \inference{\jterm{\Gamma}{A}{x}\quad\jfam{\ctxext{\Gamma}{A}}{P}}{\jfameq{\ctxext{\Gamma}{\subst{x}{P}}}{\subst{x}{\emptyf}}{\emptyf}}
+\end{align}
 
@@
 \subsubsection{Substitution is compatible with extension}\label{comp-se}
-Suppose $\jterm{\Gamma}{A}{x}$ in all of the following inference rules.
+Suppose $\jterm{\Gamma}{A}{x}$ in all of the following inference rule.
 \begin{align}
-& \inference{\jfam{\ctxext{{\Gamma}{A}}{P}}{Q}}
-  {\jfameq{\Gamma}{\subst{x}{\ctxext{P}{Q}}}{\ctxext{\subst{x}{P}}{\subst{x}{Q}}}}\\
 & \inference{\jfam{\ctxext{{{\Gamma}{A}}{P}}{Q}}{R}}
@@
 \end{equation*}
+We plug in an extra layer of families to cover the most general case at once.
 In the following inference rules we assume that $\jterm{\Gamma}{A}{x}$ and
@@
 & \inference
-    {\jfam{\ctxext{\Gamma}{A}}{Q}}
-    {\jfameq{\ctxext{\Gamma}{\subst{x}{P}}}{\subst{x}{\ctxwk{P}{Q}}}{\ctxwk{\subst{x}{P}}{\subst{x}{Q}}}}\\
-& \inference
   {\jfam{\ctxext{{\Gamma}{A}}{Q}}{R}}
-    {\jfameq{\ctxext{{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}}{\subst{x}{\ctxwk{P}{R}}}{\ctxwk{\subst{x}{P}}{\subst{x}{R}}}}\\
-& \inference
-    {\jterm{\ctxext{\Gamma}{A}}{Q}{g}}
-    {\jtermeq
-      {\ctxext{\Gamma}{\subst{x}{P}}}
-      {\subst{x}{\ctxwk{P}{Q}}}
-      {\subst{x}{\ctxwk{P}{g}}}
-      {\ctxwk{\subst{x}{P}}{\subst{x}{g}}}}\\
+  {\jfameq{\ctxext{{\Gamma}{\subst{x}{P}}}{\subst{x}{\ctxwk{P}{Q}}}}{\subst{x}{\ctxwk{P}{R}}}{\ctxwk{\subst{x}{P}}{\subst{x}{R}}}}\\
 & \inference
@@
     {\jtermeq
-      {\ctxext{{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}}
+      {\ctxext{{\Gamma}{\subst{x}{P}}}{\subst{x}{\ctxwk{P}{Q}}}}
       {\subst{x}{\ctxwk{P}{R}}}
@@
 \begin{align}
-&\inference{\jfam{\ctxext{{\Gamma}{A}}{P}}{Q}}
-{\jfameq{\Gamma}{\subst{x}{{f}{Q}}}{\subst{{x}{f}}{{x}{Q}}}}\label{comp-ss-1}\\
-&\inference{\jterm{\ctxext{{\Gamma}{A}}{P}}{Q}{g}}
-{\jtermeq{\Gamma}{\subst{x}{{f}{Q}}}{\subst{x}{{f}{g}}}{\subst{{x}{f}}{{x}{g}}}}\label{comp-ss-2}\\
 &\inference{\jfam{\ctxext{{{\Gamma}{A}}{P}}{Q}}{R}}
@@
 \begin{align}
-& \inference{\jfam{\ctxext{\Gamma}{B}}{Q}}{\jfameq{\ctxext{\Gamma}{A}}{\ctxwk{A}{\subst{y}{Q}}}{\subst{\ctxwk{A}{y}}{\ctxwk{A}{Q}}}}\\
-& \inference{\jfam{\ctxext{{\Gamma}{B}}{Q}}{R}}{\jfameq{\ctxext{{\Gamma}{A}}{\ctxwk{A}{Q}}}{\ctxwk{A}{\subst{y}{R}}}{\subst{\ctxwk{A}{y}}{\ctxwk{A}{R}}}}\\
-& \inference{\jterm{\ctxext{\Gamma}{B}}{Q}{g}}{\jtermeq{\ctxext{\Gamma}{A}}{\ctxwk{A}{\subst{y}{Q}}}{\ctxwk{A}{\subst{y}{g}}}{\subst{\ctxwk{A}{y}}{\ctxwk{A}{g}}}}\\
+& \inference{\jfam{\ctxext{{\Gamma}{B}}{Q}}{R}}{\jfameq{\ctxext{{\Gamma}{A}}{\ctxwk{A}{\subst{y}{Q}}}}{\ctxwk{A}{\subst{y}{R}}}{\subst{\ctxwk{A}{y}}{\ctxwk{A}{R}}}}\\
 & \inference{\jterm{\ctxext{{\Gamma}{B}}{Q}}{R}{h}}{\jtermeq{\ctxext{{\Gamma}{A}}{\ctxwk{A}{\subst{y}{Q}}}}{\ctxwk{A}{\subst{y}{R}}}{\ctxwk{A}{\subst{y}{h}}}{\subst{\ctxwk{A}{y}}{\ctxwk{A}{h}}}}

### 2014-02-28 article-models-tt.tex
fixing a removed label

@@
 & \jdeq \subst{{w}{\idfunc[P]}}{\subst{w}{{\ctxwk{P}{\idfunc[A]}}{\idfunc[\ctxwk{\ctxext{A}{P}}{\ctxext{A}{P}}]}}}\tag{by \autoref{comp-ss-4}}\\
-& \jdeq \subst{{w}{\idfunc[P]}}{{\subst{w}{\ctxwk{P}{\idfunc[A]}}}{\subst{w}{\idfunc[\ctxwk{\ctxext{A}{P}}{\ctxext{A}{P}}]}}}\tag{by \autoref{comp-ss-2}}\\
+& \jdeq \subst{{w}{\idfunc[P]}}{{\subst{w}{\ctxwk{P}{\idfunc[A]}}}{\subst{w}{\idfunc[\ctxwk{\ctxext{A}{P}}{\ctxext{A}{P}}]}}}\tag{by \autoref{comp-ss-4}}\\
 & \jdeq \subst{{w}{\idfunc[P]}}{{\subst{w}{\ctxwk{P}{\idfunc[A]}}}{\subst{w}{\ctxwk{\ctxext{A}{P}}{\idfunc[\ctxext{A}{P}]}}}}\tag{by \autoref{idfunc-wk-comp}}\\

### 2014-02-28 article-models-tt.tex
proof of currying complete

@@
 an arbitrary term and for substitution by an extended term. The last set of
-compatibility rules is a form of the Curry-Howard isomorphism -- this famous
+compatibility rules is a form of currying-uncurrying -- this famous
 correspondence naturally comes out of our scheme of compatibility rules.
@@
             {\ctxwk{A}{{B}{R}}}
-            {\ctxwk{{A}{B}}{{A}{R}}}}\\
+            {\ctxwk{{A}{B}}{{A}{R}}}}\label{comp-ww-f}\\
 & \inference{\jfam{\Gamma}{A}\qquad\jfam{\Gamma}{B}\qquad\jterm{\ctxext{\Gamma}{C}}{R}{t}}
@@
             {\ctxwk{{A}{B}}{{A}{R}}}{\ctxwk{A}{{B}{t}}}{\ctxwk{{A}{B}}{{A}{t}}}}
-            \label{comp-ww-inf4}
+            \label{comp-ww-t}
 \end{align}
@@
 & \inference{\jfam{\ctxext{\Gamma}{B}}{Q}}
-    {\jfameq{\ctxext{{{\Gamma}{A}}{P}}{\ctxwk{P}{{A}{B}}}}{\ctxwk{\ctxext{A}{P}}{Q}}{\ctxwk{P}{{A}{Q}}}}\\
+    {\jfameq{\ctxext{{{\Gamma}{A}}{P}}{\ctxwk{P}{{A}{B}}}}{\ctxwk{\ctxext{A}{P}}{Q}}{\ctxwk{P}{{A}{Q}}}}\label{comp-ew-f}\\
 & \inference{\jterm{\ctxext{\Gamma}{B}}{Q}{g}}
-    {\jtermeq{\ctxext{{{\Gamma}{A}}{P}}{\ctxwk{P}{{A}{B}}}}{\ctxwk{P}{{A}{Q}}}{\ctxwk{\ctxext{A}{P}}{g}}{\ctxwk{P}{{A}{g}}}}
+    {\jtermeq{\ctxext{{{\Gamma}{A}}{P}}{\ctxwk{P}{{A}{B}}}}{\ctxwk{P}{{A}{Q}}}{\ctxwk{\ctxext{A}{P}}{g}}{\ctxwk{P}{{A}{g}}}}\label{comp-ew-t}
 \end{align}
@@
   {\jfam{\ctxext{{\Gamma}{A}}{Q}}{R}}
-  {\jfameq{\ctxext{{\Gamma}{\subst{x}{P}}}{\subst{x}{\ctxwk{P}{Q}}}}{\subst{x}{\ctxwk{P}{R}}}{\ctxwk{\subst{x}{P}}{\subst{x}{R}}}}\\
+  {\jfameq{\ctxext{{\Gamma}{\subst{x}{P}}}{\subst{x}{\ctxwk{P}{Q}}}}{\subst{x}{\ctxwk{P}{R}}}{\ctxwk{\subst{x}{P}}{\subst{x}{R}}}}\label{comp-sw-f}\\
 & \inference
@@
       {\subst{x}{\ctxwk{P}{h}}}
-      {\ctxwk{\subst{x}{P}}{\subst{x}{h}}}}
+      {\ctxwk{\subst{x}{P}}{\subst{x}{h}}}}\label{comp-sw-t}
 \end{align}
@@
 &\inference{\jfam{\ctxext{{{\Gamma}{A}}{P}}{Q}}{R}}
-{\jfameq{\ctxext{\Gamma}{\subst{x}{{f}{Q}}}}{\subst{x}{{f}{R}}}{\subst{{x}{f}}{{x}{R}}}}\label{comp-ss-3}\\
+{\jfameq{\ctxext{\Gamma}{\subst{x}{{f}{Q}}}}{\subst{x}{{f}{R}}}{\subst{{x}{f}}{{x}{R}}}}\label{comp-ss-f}\\
 &\inference{\jterm{\ctxext{{{\Gamma}{A}}{P}}{Q}}{R}{h}}
-{\jtermeq{\ctxext{\Gamma}{\subst{x}{{f}{Q}}}}{\subst{x}{{f}{R}}}{\subst{x}{{f}{h}}}{\subst{{x}{f}}{{x}{h}}}}\label{comp-ss-4}
+{\jtermeq{\ctxext{\Gamma}{\subst{x}{{f}{Q}}}}{\subst{x}{{f}{R}}}{\subst{x}{{f}{h}}}{\subst{{x}{f}}{{x}{h}}}}\label{comp-ss-t}
 \end{align}
@@
 \begin{align}
-& \inference{\jfam{\ctxext{{\Gamma}{B}}{Q}}{R}}{\jfameq{\ctxext{{\Gamma}{A}}{\ctxwk{A}{\subst{y}{Q}}}}{\ctxwk{A}{\subst{y}{R}}}{\subst{\ctxwk{A}{y}}{\ctxwk{A}{R}}}}\\
-& \inference{\jterm{\ctxext{{\Gamma}{B}}{Q}}{R}{h}}{\jtermeq{\ctxext{{\Gamma}{A}}{\ctxwk{A}{\subst{y}{Q}}}}{\ctxwk{A}{\subst{y}{R}}}{\ctxwk{A}{\subst{y}{h}}}{\subst{\ctxwk{A}{y}}{\ctxwk{A}{h}}}}
+& \inference{\jfam{\ctxext{{\Gamma}{B}}{Q}}{R}}{\jfameq{\ctxext{{\Gamma}{A}}{\ctxwk{A}{\subst{y}{Q}}}}{\ctxwk{A}{\subst{y}{R}}}{\subst{\ctxwk{A}{y}}{\ctxwk{A}{R}}}}\label{comp-ws-f}\\
+& \inference{\jterm{\ctxext{{\Gamma}{B}}{Q}}{R}{h}}{\jtermeq{\ctxext{{\Gamma}{A}}{\ctxwk{A}{\subst{y}{Q}}}}{\ctxwk{A}{\subst{y}{R}}}{\ctxwk{A}{\subst{y}{h}}}{\subst{\ctxwk{A}{y}}{\ctxwk{A}{h}}}}\label{comp-ws-t}
 \end{align}
@@
 \begin{align}
-& \inference{\jfam{\Gamma}{A}\qquad\jfam{\Gamma}{B}\qquad\jterm{\Gamma}{A}{x}}{\jfameq{\Gamma}{\subst{x}{\ctxwk{A}{B}}}{B}}\label{defn-ws-1}\\
-& \inference{\jterm{\Gamma}{A}{x}\qquad\jterm{\Gamma}{B}{y}}{\jtermeq{\Gamma}{B}{\subst{x}{\ctxwk{A}{y}}}{y}}\label{defn-ws-2}\\
 & \inference{\jfam{\Gamma}{A}\qquad\jfam{\ctxext{\Gamma}{B}}{Q}\qquad\jterm{\Gamma}{A}{x}}{\jfameq{\ctxext{\Gamma}{B}}{\subst{x}{\ctxwk{A}{Q}}}{Q}}\label{defn-ws-3}\\
@@
 & \inference{\jfam{\ctxext{\Gamma}{A}}{P}}{\jfameq{\ctxext{\Gamma}{A}}{\subst{\idfunc[A]}{\ctxwk{A}{P}}}{P}}\label{idfunc-wk-defn}\\
+& \inference{\jfam{\ctxext{\Gamma}{A}}{P}}{\jfameq{\ctxext{\Gamma}{A}}{\subst{\idfunc[A]}{\ctxwk{{A}{A}}{P}}}{P}}\label{idfunc-wk-defn2}\\
 & \inference{\jterm{\ctxext{\Gamma}{A}}{P}{f}}{\jtermeq{\ctxext{\Gamma}{A}}{P}{\subst{\idfunc[A]}{\ctxwk{A}{f}}}{f}}\label{idfunc-precomp} \\
+& \inference{\jterm{\ctxext{\Gamma}{A}}{P}{f}}{\jtermeq{\ctxext{\Gamma}{A}}{P}{\subst{\idfunc[A]}{\ctxwk{{A}{A}}{f}}}{f}}\label{idfunc-precomp} \\
 & \inference{\jterm{\ctxext{\Gamma}{B}}{\ctxwk{B}{A}}{g}}{\jtermeq{\ctxext{\Gamma}{B}}{\ctxwk{B}{A}}{\subst{g}{\ctxwk{B}{\idfunc[A]}}}{g}}\label{idfunc-postcomp}
@@
 
+\begin{lem}
+The following inference is valid:
+\begin{equation*}
+\inference{\jfam{\ctxext{\Gamma}{A}}{P}\quad\jfam{\ctxext{\Gamma}{A}}{Q}}
+{\jfameq{\ctxext{\Gamma}{{A}{P}}}{\subst{\cprojfstf{A}{P}}{\ctxwk{\ctxext{A}{P}}{Q}}}{\ctxwk{P}{Q}}}
+\end{equation*}
+\end{lem}
+
+\begin{proof}
+Let $\jfam{\ctxext{\Gamma}{A}}{P}$ and $\jfam{\ctxext{\Gamma}{A}}{Q}$ be
+families. Then we have
+\begin{align*}
+\subst{\cprojfstf{A}{P}}{\ctxwk{\ctxext{A}{P}}{Q}} 
+& \jdeq \subst{\ctxwk{P}{\idfunc[A]}}{\ctxwk{\ctxext{A}{P}}{Q}} \tag{by definition}\\
+& \jdeq \subst{\ctxwk{P}{\idfunc[A]}}{\ctxwk{P}{{A}{Q}}} \tag{by \autoref{comp-ww-f}}\\
+& \jdeq \ctxwk{P}{\subst{\idfunc[A]}{\ctxwk{A}{Q}}} \tag{by \autoref{comp-ws-f}}\\
+& \jdeq \ctxwk{P}{Q} \tag{by \autoref{idfunc-wk-defn}}
+\end{align*}
+\end{proof}
+
 The constructions of the terms $\ctxext{x}{u}$ and $\cprojfst{A}{P}{w}$ and
@@
   \tag{by \autoref{idfunc-ext-comp}}\\
-& \jdeq \subst{{w}{\idfunc[P]}}{\subst{w}{{\ctxwk{P}{\idfunc[A]}}{\idfunc[\ctxwk{\ctxext{A}{P}}{\ctxext{A}{P}}]}}}\tag{by \autoref{comp-ss-4}}\\
-& \jdeq \subst{{w}{\idfunc[P]}}{{\subst{w}{\ctxwk{P}{\idfunc[A]}}}{\subst{w}{\idfunc[\ctxwk{\ctxext{A}{P}}{\ctxext{A}{P}}]}}}\tag{by \autoref{comp-ss-4}}\\
+& \jdeq \subst{{w}{\idfunc[P]}}{\subst{w}{{\ctxwk{P}{\idfunc[A]}}{\idfunc[\ctxwk{\ctxext{A}{P}}{\ctxext{A}{P}}]}}}\tag{by \autoref{comp-ss-t}}\\
+& \jdeq \subst{{w}{\idfunc[P]}}{{\subst{w}{\ctxwk{P}{\idfunc[A]}}}{\subst{w}{\idfunc[\ctxwk{\ctxext{A}{P}}{\ctxext{A}{P}}]}}}\tag{by \autoref{comp-ss-t}}\\
 & \jdeq \subst{{w}{\idfunc[P]}}{{\subst{w}{\ctxwk{P}{\idfunc[A]}}}{\subst{w}{\ctxwk{\ctxext{A}{P}}{\idfunc[\ctxext{A}{P}]}}}}\tag{by \autoref{idfunc-wk-comp}}\\
-& \jdeq \subst{{w}{\idfunc[P]}}{{\subst{w}{\ctxwk{P}{\idfunc[A]}}}{\idfunc[\ctxext{A}{P}]}}\tag{by \autoref{defn-ws-2}}\\
+& \jdeq \subst{{w}{\idfunc[P]}}{{\subst{w}{\ctxwk{P}{\idfunc[A]}}}{\idfunc[\ctxext{A}{P}]}}\tag{by \autoref{defn-ws-4}}\\
 & \jdeq \ctxext{\cprojfst{A}{P}{w}}{\cprojsnd{A}{P}{w}}\tag{by definition}
@@
 Finally, when we substitute by an extended term we get an equal result as when we
-substitute two consecutive times. This looks like the Curry-Howard isomorphism.
+substitute two consecutive times. This looks like currying.
 
@@
 \end{lem}
+
+\begin{proof}
+We prove only the first judgmental equality. All the others are similar.
+Let $\jterm{\Gamma}{A}{x}$ and $\jterm{\Gamma}{\subst{x}{P}}{u}$
+be terms and let $\jfam{\ctxext{{\Gamma}{A}}{P}}{Q}$ be a family. Then we have
+\begin{align*}
+\subst{\ctxext{x}{u}}{Q} 
+& \jdeq \subst{{u}{{x}{\idfunc[\ctxext{A}{P}]}}}{Q} \tag{by definition}\\
+& \jdeq \subst{{u}{{x}{\idfunc[\ctxext{A}{P}]}}}{{x}{\ctxwk{A}{Q}}}\tag{by \autoref{defn-ws-3}}\\
+& \jdeq \subst{{u}{{x}{\idfunc[\ctxext{A}{P}]}}}{{u}{\ctxwk{\subst{x}{P}}{\subst{x}{\ctxwk{A}{Q}}}}}\tag{by \autoref{defn-ws-3}}\\
+& \jdeq \subst{{u}{{x}{\idfunc[\ctxext{A}{P}]}}}{{u}{{x}{\ctxwk{P}{{A}{Q}}}}}\tag{by \autoref{comp-sw-f}}\\
+& \jdeq \subst{u}{{{x}{\idfunc[\ctxext{A}{P}]}}{{x}{\ctxwk{{P}{{A}{Q}}}}}}\tag{by \autoref{comp-ss-f}}\\
+& \jdeq \subst{u}{{x}{{\idfunc[\ctxext{A}{P}]}{\ctxwk{P}{{A}{Q}}}}}\tag{by \autoref{comp-ss-f}}\\
+& \jdeq \subst{u}{{x}{{\idfunc[\ctxext{A}{P}]}{\ctxwk{\ctxext{A}{P}}{Q}}}}\tag{by \autoref{comp-ew-f}}\\
+& \jdeq \subst{u}{{x}{Q}}\tag{by \autoref{idfunc-wk-defn}}
+\end{align*}
+\end{proof}

### 2014-02-28 article-models-tt.tex
moved the currying lemma
first proj of x.u is x

@@
 
+When we substitute by an extended term we get an equal result as when we
+substitute two consecutive times, like the way currying works. Note however,
+that we can now \emph{prove} this property using the inference rules we have
+already established. It is not a matter of convention, the way it usually is
+in type theory by agreeing that $f(x)(y)$ is written as $f(x,y)$.
+
+\begin{lem}
+The following inference rules are valid:
+\begin{align*}
+& \inference{\jterm{\Gamma}{A}{x}\quad\jterm{\Gamma}{\subst{x}{P}}{u}\quad\jfam{\ctxext{{\Gamma}{A}}{P}}{Q}}
+  {\jfameq{\Gamma}{\subst{\ctxext{x}{u}}{Q}}{\subst{u}{{x}{Q}}}}\\
+& \inference{\jterm{\Gamma}{A}{x}\quad\jterm{\Gamma}{\subst{x}{P}}{u}\quad\jterm{\ctxext{{\Gamma}{A}}{P}}{Q}{g}}
+  {\jtermeq{\Gamma}{\subst{u}{{x}{Q}}}{\subst{\ctxext{x}{u}}{g}}{\subst{u}{{x}{g}}}}\\
+& \inference{\jterm{\Gamma}{A}{x}\quad\jterm{\Gamma}{\subst{x}{P}}{u}\quad\jfam{\ctxext{{{\Gamma}{A}}{P}}{Q}}{R}}
+  {\jfameq{\ctxext{\Gamma}{\subst{u}{{x}{Q}}}}{\subst{\ctxext{x}{u}}{R}}{\subst{u}{{x}{R}}}}\\
+& \inference{\jterm{\Gamma}{A}{x}\quad\jterm{\Gamma}{\subst{x}{P}}{u}\quad\jterm{\ctxext{{{\Gamma}{A}}{P}}{Q}}{R}{t}}
+  {\jtermeq{\ctxext{\Gamma}{\subst{u}{{x}{Q}}}}{\subst{u}{{x}{R}}}{\subst{\ctxext{x}{u}}{t}}{\subst{u}{{x}{t}}}}
+\end{align*}
+\end{lem}
+
+\begin{proof}
+We prove only the first judgmental equality. All the others are similar.
+Let $\jterm{\Gamma}{A}{x}$ and $\jterm{\Gamma}{\subst{x}{P}}{u}$
+be terms and let $\jfam{\ctxext{{\Gamma}{A}}{P}}{Q}$ be a family. Then we have
+\begin{align*}
+\subst{\ctxext{x}{u}}{Q} 
+& \jdeq \subst{{u}{{x}{\idfunc[\ctxext{A}{P}]}}}{Q} \tag{by definition}\\
+& \jdeq \subst{{u}{{x}{\idfunc[\ctxext{A}{P}]}}}{{x}{\ctxwk{A}{Q}}}\tag{by \autoref{defn-ws-3}}\\
+& \jdeq \subst{{u}{{x}{\idfunc[\ctxext{A}{P}]}}}{{u}{\ctxwk{\subst{x}{P}}{\subst{x}{\ctxwk{A}{Q}}}}}\tag{by \autoref{defn-ws-3}}\\
+& \jdeq \subst{{u}{{x}{\idfunc[\ctxext{A}{P}]}}}{{u}{{x}{\ctxwk{P}{{A}{Q}}}}}\tag{by \autoref{comp-sw-f}}\\
+& \jdeq \subst{u}{{{x}{\idfunc[\ctxext{A}{P}]}}{{x}{\ctxwk{{P}{{A}{Q}}}}}}\tag{by \autoref{comp-ss-f}}\\
+& \jdeq \subst{u}{{x}{{\idfunc[\ctxext{A}{P}]}{\ctxwk{P}{{A}{Q}}}}}\tag{by \autoref{comp-ss-f}}\\
+& \jdeq \subst{u}{{x}{{\idfunc[\ctxext{A}{P}]}{\ctxwk{\ctxext{A}{P}}{Q}}}}\tag{by \autoref{comp-ew-f}}\\
+& \jdeq \subst{u}{{x}{Q}}\tag{by \autoref{idfunc-wk-defn}}
+\end{align*}
+\end{proof}
+
 There is also a decomposition of terms $\jterm{\Gamma}{\ctxext{A}{P}}{w}$ into
@@
 & \jdeq \ctxwk{P}{Q} \tag{by \autoref{idfunc-wk-defn}}
+\end{align*}
+To prove the second judgmental equality, let $\jterm{\Gamma}{A}{x}$ and
+$\jterm{\Gamma}{\subst{x}{P}}{u}$. Then we have
+\begin{align*}
+\cprojfst{A}{P}{\ctxext{x}{u}}
+& \jdeq \subst{\ctxext{x}{u}}{\ctxwk{P}{\idfunc[A]}}\\
+& \jdeq \subst{u}{{x}{\ctxwk{P}{\idfunc[A]}}} \\
+& \jdeq \subst{u}{\ctxwk{\subst{x}{P}}{\subst{x}{\idfunc[A]}}}\\
+& \jdeq \subst{x}{\idfunc[A]}\\
+& \jdeq x.
 \end{align*}
@@
 \end{lem}
-Finally, when we substitute by an extended term we get an equal result as when we
-substitute two consecutive times. This looks like currying.
-
-\begin{lem}
-The following inference rules are valid:
-\begin{align*}
-& \inference{\jterm{\Gamma}{A}{x}\quad\jterm{\Gamma}{\subst{x}{P}}{u}\quad\jfam{\ctxext{{\Gamma}{A}}{P}}{Q}}
-  {\jfameq{\Gamma}{\subst{\ctxext{x}{u}}{Q}}{\subst{u}{{x}{Q}}}}\\
-& \inference{\jterm{\Gamma}{A}{x}\quad\jterm{\Gamma}{\subst{x}{P}}{u}\quad\jterm{\ctxext{{\Gamma}{A}}{P}}{Q}{g}}
-  {\jtermeq{\Gamma}{\subst{u}{{x}{Q}}}{\subst{\ctxext{x}{u}}{g}}{\subst{u}{{x}{g}}}}\\
-& \inference{\jterm{\Gamma}{A}{x}\quad\jterm{\Gamma}{\subst{x}{P}}{u}\quad\jfam{\ctxext{{{\Gamma}{A}}{P}}{Q}}{R}}
-  {\jfameq{\ctxext{\Gamma}{\subst{u}{{x}{Q}}}}{\subst{\ctxext{x}{u}}{R}}{\subst{u}{{x}{R}}}}\\
-& \inference{\jterm{\Gamma}{A}{x}\quad\jterm{\Gamma}{\subst{x}{P}}{u}\quad\jterm{\ctxext{{{\Gamma}{A}}{P}}{Q}}{R}{t}}
-  {\jtermeq{\ctxext{\Gamma}{\subst{u}{{x}{Q}}}}{\subst{u}{{x}{R}}}{\subst{\ctxext{x}{u}}{t}}{\subst{u}{{x}{t}}}}
-\end{align*}
-\end{lem}
-
-\begin{proof}
-We prove only the first judgmental equality. All the others are similar.
-Let $\jterm{\Gamma}{A}{x}$ and $\jterm{\Gamma}{\subst{x}{P}}{u}$
-be terms and let $\jfam{\ctxext{{\Gamma}{A}}{P}}{Q}$ be a family. Then we have
-\begin{align*}
-\subst{\ctxext{x}{u}}{Q} 
-& \jdeq \subst{{u}{{x}{\idfunc[\ctxext{A}{P}]}}}{Q} \tag{by definition}\\
-& \jdeq \subst{{u}{{x}{\idfunc[\ctxext{A}{P}]}}}{{x}{\ctxwk{A}{Q}}}\tag{by \autoref{defn-ws-3}}\\
-& \jdeq \subst{{u}{{x}{\idfunc[\ctxext{A}{P}]}}}{{u}{\ctxwk{\subst{x}{P}}{\subst{x}{\ctxwk{A}{Q}}}}}\tag{by \autoref{defn-ws-3}}\\
-& \jdeq \subst{{u}{{x}{\idfunc[\ctxext{A}{P}]}}}{{u}{{x}{\ctxwk{P}{{A}{Q}}}}}\tag{by \autoref{comp-sw-f}}\\
-& \jdeq \subst{u}{{{x}{\idfunc[\ctxext{A}{P}]}}{{x}{\ctxwk{{P}{{A}{Q}}}}}}\tag{by \autoref{comp-ss-f}}\\
-& \jdeq \subst{u}{{x}{{\idfunc[\ctxext{A}{P}]}{\ctxwk{P}{{A}{Q}}}}}\tag{by \autoref{comp-ss-f}}\\
-& \jdeq \subst{u}{{x}{{\idfunc[\ctxext{A}{P}]}{\ctxwk{\ctxext{A}{P}}{Q}}}}\tag{by \autoref{comp-ew-f}}\\
-& \jdeq \subst{u}{{x}{Q}}\tag{by \autoref{idfunc-wk-defn}}
-\end{align*}
-\end{proof}

### 2014-03-01 article-models-tt.tex
second proj of x.u is u
whoops, I proved the last two things in the wrong place. Hehe
worked on extension on terms, initiated two new sections

@@
 \subsection{Extension on terms}\label{extension-on-terms}
+In this subsection we consider the notion of extension on terms, which has now
+become definable inside our theory. Moreover, every compatibility rule one may
+dream of is provable as well, using the compatibility rules we have introduced
+earlier. \emph{This subsection contains no new assumptions.}
+
+\begin{defn}
 When $\jterm{\Gamma}{A}{x}$ and $\jterm{\Gamma}{\subst{x}{P}}{u}$ are terms,
-we obtain the term $\jterm{\Gamma}{\ctxext{A}{P}}{\subst{u}{{x}{\idfunc[\ctxext{A}{P}]}}}$. 
-We define 
+we define 
 \begin{equation*}
-\ctxext{x}{u}\defeq\subst{u}{{x}{\idfunc[\ctxext{A}{P}]}}.
+\jtermdefn{\Gamma}{\ctxext{A}{P}}{\ctxext{x}{u}}{\subst{u}{{x}{\idfunc[\ctxext{A}{P}]}}}.
 \end{equation*} 
-Thus
-$\ctxext{x}{u}$ is the pairing of $x$ and $u$. Note that because
-$\ctxwk{P}{{A}{\ctxext{A}{P}}}\jdeq\ctxwk{\ctxext{A}{P}}{\ctxext{A}{P}}$, the
-pairing function is just $\idfunc[\ctxext{A}{P}]$.
+\end{defn}
+
+Thus, the term $\ctxext{x}{u}$ is the pairing of $x$ and $u$. Note that because
+we have the judgmental equality 
+$\ctxwk{P}{{A}{\ctxext{A}{P}}}\jdeq\ctxwk{\ctxext{A}{P}}{\ctxext{A}{P}}$ in the
+context $\ctxext{{\Gamma}{A}}{P}$, the
+pairing function could just be defined as $\idfunc[\ctxext{A}{P}]$.
 
 When we substitute by an extended term we get an equal result as when we
-substitute two consecutive times, like the way currying works. Note however,
-that we can now \emph{prove} this property using the inference rules we have
-already established. It is not a matter of convention, the way it usually is
-in type theory by agreeing that $f(x)(y)$ is written as $f(x,y)$.
+substitute two consecutive times, like the way currying works.
 
@@
 There is also a decomposition of terms $\jterm{\Gamma}{\ctxext{A}{P}}{w}$ into
-components: the projections. We define
+components: the projections.
+
+\begin{defn}
+Let $\jfam{\Gamma}{A}$ and $\jfam{\ctxext{\Gamma}{A}}{P}$ be families. We define
 \begin{align*}
-\cprojfstf{A}{P} & \defeq \ctxwk{P}{\idfunc[A]}\\
-\cprojsndf{A}{P} & \defeq \idfunc[P]
+\jtermdefn*{\ctxext{\Gamma}{{A}{P}}}{\ctxwk{P}{{A}{A}}}{\cprojfstf{A}{P}}{\ctxwk{P}{\idfunc[A]}}\\
+\jtermdefn*{\ctxext{\Gamma}{{A}{P}}}{\ctxwk{P}{P}}{\cprojsndf{A}{P}}{\idfunc[P]}
 \end{align*}
+\end{defn}
 
@@
 & \jdeq \ctxwk{P}{Q} \tag{by \autoref{idfunc-wk-defn}}
-\end{align*}
-To prove the second judgmental equality, let $\jterm{\Gamma}{A}{x}$ and
-$\jterm{\Gamma}{\subst{x}{P}}{u}$. Then we have
-\begin{align*}
-\cprojfst{A}{P}{\ctxext{x}{u}}
-& \jdeq \subst{\ctxext{x}{u}}{\ctxwk{P}{\idfunc[A]}}\\
-& \jdeq \subst{u}{{x}{\ctxwk{P}{\idfunc[A]}}} \\
-& \jdeq \subst{u}{\ctxwk{\subst{x}{P}}{\subst{x}{\idfunc[A]}}}\\
-& \jdeq \subst{x}{\idfunc[A]}\\
-& \jdeq x.
 \end{align*}
@@
 \end{align*}
+To prove the second judgmental equality, let $\jterm{\Gamma}{A}{x}$ and
+$\jterm{\Gamma}{\subst{x}{P}}{u}$. Then we have
+\begin{align*}
+\cprojfst{A}{P}{\ctxext{x}{u}}
+& \jdeq \subst{\ctxext{x}{u}}{\ctxwk{P}{\idfunc[A]}}\\
+& \jdeq \subst{u}{{x}{\ctxwk{P}{\idfunc[A]}}} \\
+& \jdeq \subst{u}{\ctxwk{\subst{x}{P}}{\subst{x}{\idfunc[A]}}}\\
+& \jdeq \subst{x}{\idfunc[A]}\\
+& \jdeq x.
+\end{align*}
+To prove the third judgmental equality, note that
+\begin{align*}
+\cprojsnd{A}{P}{\ctxext{x}{u}}
+& \jdeq \subst{\ctxext{x}{u}}{\idfunc[P]}\\
+& \jdeq \subst{u}{{x}{\idfunc[P]}}\\
+& \jdeq \subst{u}{\idfunc[\subst{x}{P}]}\\
+& \jdeq u.
+\end{align*}
 \end{proof}
@@
 \end{lem}
+
+\begin{proof}
+Consider terms $\jterm{\Gamma}{A}{x}$, $\jterm{\Gamma}{\subst{x}{P}}{u}$ and
+$\jterm{\Gamma}{\subst{u}{{x}{Q}}}{v}$. Then we have
+\begin{align*}
+\ctxext{x}{{u}{v}} & \jdeq \subst{\ctxext{u}{v}}{{x}{\idfunc[\ctxext{A}{{P}{Q}}]}}\\
+& \jdeq \subst{v}{{u}{{x}{\idfunc[\ctxext{A}{{P}{Q}}]}}}\\
+& \jdeq \subst{v}{{u}{{x}{\idfunc[\ctxext{{A}{P}}{Q}]}}}\\
+& \jdeq \subst{v}{{\ctxext{x}{u}}{\idfunc[\ctxext{{A}{P}}{Q}]}}\\
+& \jdeq \ctxext{{x}{u}}{v}.
+\end{align*}
+Now consider a term $\jterm{\Gamma}{\ctxext{A}{{P}{Q}}}{w}$. Then we have
+\begin{align*}
+w 
+& \jdeq \ctxext{\cprojfst{A}{\ctxext{P}{Q}}{w}}{\cprojsnd{A}{\ctxext{P}{Q}}{w}}\\
+& \jdeq \ctxext{\cprojfst{A}{\ctxext{P}{Q}}{w}}{{\cprojfst{P}{Q}{\cprojsnd{A}{\ctxext{P}{Q}}{w}}}{\cprojsnd{P}{Q}{\cprojsnd{A}{\ctxext{P}{Q}}{w}}}}\\
+& \jdeq \ctxext{{\cprojfst{A}{\ctxext{P}{Q}}{w}}{\cprojfst{P}{Q}{\cprojsnd{A}{\ctxext{P}{Q}}{w}}}}{\cprojsnd{P}{Q}{\cprojsnd{A}{\ctxext{P}{Q}}{w}}}
+\end{align*}
+Thus we see that 
+\begin{align*}
+\cprojfst{\ctxext{A}{P}}{Q}{w} & \jdeq \ctxext{\cprojfst{A}{\ctxext{P}{Q}}{w}}{\cprojfst{P}{Q}{\cprojsnd{A}{\ctxext{P}{Q}}{w}}}\\ 
+\cprojsnd{\ctxext{A}{P}}{Q}{w} & \jdeq \cprojsnd{P}{Q}{\cprojsnd{A}{\ctxext{P}{Q}}{w}},
+\end{align*}
+proving the fourth judgmental equality, and therefore also that
+\begin{align*}
+\cprojfst{A}{P}{\cprojfst{\ctxext{A}{P}}{Q}{w}} & \jdeq \cprojfst{A}{\ctxext{P}{Q}}{w}\\
+\cprojsnd{A}{P}{\cprojfst{\ctxext{A}{P}}{Q}{w}} & \jdeq \cprojfst{P}{Q}{\cprojsnd{A}{\ctxext{P}{Q}}{w}},
+\end{align*}
+proving the second and the third judgmental equalities.
+\end{proof}
 
@@
 \end{lem}
+
+\section{Pretty type theory}
+In this section we will do three things. First we explain basic type theory with
+variable names, which we simply call \emph{Pretty Type Theory}. 
+Then we will show how every formula in basic type theory without
+variable names (shall we call it \emph{Structural Type Theory}?) 
+can be interpreted in pretty type theory. Finally, we will show how every formula
+in pretty type theory can be interpreted in structural type theory.
+
+\section{Chalmers type theory}
+In this section we will present the type theory that Coquand and Dybjer are using.
+It is a weak type theory (I think), with not so many operations and judgmental equalities.
+We will show how every formula of Chalmers type theory can be interpreted in
+structural type theory.

### 2014-03-01 article-models-tt.tex
did all the proofs of the extension on terms subsection

@@
 
+\begin{proof}
+Consider $\jterm{\ctxext{\Gamma}{B}}{Q}{g}$ and $\jterm{\ctxext{\Gamma}{B}}{\subst{g}{R}}{t}$.
+Then we have the judgmental equalities
+\begin{align*}
+\ctxwk{A}{\ctxext{g}{t}}
+& \jdeq \ctxwk{A}{\subst{t}{{g}{\idfunc[\ctxext{Q}{R}]}}}\\
+& \jdeq \subst{\ctxwk{A}{t}}{\ctxwk{A}{\subst{g}{\idfunc[\ctxext{Q}{R}]}}}\\
+& \jdeq \subst{\ctxwk{A}{t}}{{\ctxwk{A}{g}}{\ctxwk{A}{\idfunc[\ctxext{Q}{R}]}}}\\
+& \jdeq \subst{\ctxwk{A}{t}}{{\ctxwk{A}{g}}{\idfunc[\ctxwk{A}{\ctxext{Q}{R}}]}}\\
+& \jdeq \subst{\ctxwk{A}{t}}{{\ctxwk{A}{g}}{\idfunc[\ctxext{\ctxwk{A}{Q}}{\ctxwk{A}{R}}]}}\\
+& \jdeq \ctxext{\ctxwk{A}{g}}{\ctxwk{A}{t}}
+\end{align*}
+\end{proof}
+
 \begin{lem}
@@
 \begin{align*}
-& \inference{\jterm{\ctxext{\Gamma}{A}}{P}{f}\quad\jterm{\ctxext{\Gamma}{A}}{\subst{f}{Q}}{g}}{\jtermeq{\Gamma}{\ctxext{\subst{x}{P}}{\subst{x}{Q}}}{\subst{x}{\ctxext{f}{g}}}{\ctxext{\subst{x}{f}}{\subst{x}{g}}}}\\
 & \inference{\jterm{\ctxext{{\Gamma}{A}}{P}}{Q}{g}\quad\jterm{\ctxext{{\Gamma}{A}}{P}}{\subst{g}{R}}{t}}{\jtermeq{\ctxext{\Gamma}{\subst{x}{P}}}{\ctxext{\subst{x}{Q}}{\subst{x}{R}}}{\subst{x}{\ctxext{g}{t}}}{\ctxext{\subst{x}{g}}{\subst{x}{t}}}}
@@
 \end{lem}
+
+\begin{proof}
+Consider $\jterm{\ctxext{\Gamma}{B}}{Q}{g}$ and $\jterm{\ctxext{\Gamma}{B}}{\subst{g}{R}}{t}$.
+Then we have the judgmental equalities
+\begin{align*}
+\subst{x}{\ctxext{g}{t}}
+& \jdeq \subst{x}{{t}{{g}{\idfunc[\ctxext{Q}{R}]}}}\\
+& \jdeq \subst{{x}{t}}{{x}{{g}{\idfunc[\ctxext{Q}{R}]}}}\\
+& \jdeq \subst{{x}{t}}{{{x}{g}}{{x}{\idfunc[\ctxext{Q}{R}]}}}\\
+& \jdeq \subst{{x}{t}}{{{x}{g}}{\idfunc[\subst{x}{\ctxext{Q}{R}}]}}\\
+& \jdeq \subst{{x}{t}}{{{x}{g}}{\idfunc[\ctxext{\subst{x}{Q}}{\subst{x}{R}}]}}\\
+& \jdeq \ctxext{\subst{x}{g}}{\subst{x}{t}}.
+\end{align*}
+\end{proof}
 
@@
 
+Pretty type theory is pretty as an ugly little duckling. At first, it is actually uglier than
+structural type theory because one has to keep track of a variable nobody is
+actually interested in. The situation improves when we introduce the operators
+of extension and weakening, because from that point onwards one can draw notational
+advantages from having the variable around.
+
+\subsection{The basic judgments}
+The type theory we describe here is a theory of contexts, families of
+contexts and terms thereof. The families of contexts are by some authors called
+dependent contexts, but they are handled a bit differently here because they
+become the primary object of study. Dependent contexts can be types; they could
+be seen as atomic or indecomposable dependent contexts.
+
+Thus we make eight kinds of judgments in our type theory: ``$\Gamma$ is a context'',
+``$A(i)$ is a family of contexts over $\Gamma$'', ``$A$ is a type in context $\Gamma$''
+and ``$x$ is a term of the family $A$ of contexts over $\Gamma$''. The other four
+judgments are for judgmental equality.
+
+\begin{align*}
+\jvctx*{\Gamma} & \jvctxeq*{\Gamma}{\Gamma'}\\
+\jvfam*{i}{\Gamma}{A} & \jvfameq*{i}{\Gamma}{A}{B}\\
+\jvtype*{i}{\Gamma}{A} & \jvtypeeq*{i}{\Gamma}{A}{B}\\
+\jvterm*{i}{\Gamma}{A}{x} & \jvtermeq*{i}{\Gamma}{A}{x}{y}.
+\end{align*}
+
+If $A$ is a type
+in context $\Gamma$, then $A$ is also a family of contexts over $\Gamma$. Being
+a term of type $A$ means the same thing as being a term of the family $A$ of contexts.
+Two types in context $\Gamma$ are judgmentally equal precisely when they are equal
+as context families. Moreover, if a family $B$ of contexts over $\Gamma$ is
+judgmentally equal to a type $A$ in context $\Gamma$, then $B$ is a type in
+context $\Gamma$. This is expressed by the following four inference rules:
+
+\begin{small}
+\begin{align*}
+& \inference{\jvtype{i}{\Gamma}{A}}{\jvfam{i}{\Gamma}{A}} & & \inference{\jvtypeeq{i}{\Gamma}{A}{B}}{\jvfameq{i}{\Gamma}{A}{B}}\\
+& \inference{\jvtype{i}{\Gamma}{A}\quad\jvfameq{i}{\Gamma}{A}{B}}{\jvtype{i}{\Gamma}{B}}
+& & \inference{\jvtype{i}{\Gamma}{A}\quad\jvfameq{i}{\Gamma}{A}{B}}{\jvtypeeq{i}{\Gamma}{A}{B}}
+\end{align*}
+\end{small}
+
+\subsection{The basic rules for judgmental equality}
+The rules for judgmental equality establish that it is an equivalence relation
+in all three cases (contexts, types and terms).
+\bgroup\small
+\begin{align*}
+& \inference{\jvctx{\Gamma}}{\jvctxeq{\Gamma}{\Gamma}} 
+& & \inference{\jvctxeq{\Gamma}{\Delta}}{\jvctxeq{\Delta}{\Gamma}} 
+& & \inference{\jvctxeq{\Gamma}{\Delta}\qquad\jvctxeq{\Delta}{\greek{E}}}{\jvctxeq{\Gamma}{\greek{E}}}\\
+& \inference{\jvfam{i}{\Gamma}{A}}{\jvfameq{i}{\Gamma}{A}{A}} 
+& & \inference{\jvfameq{i}{\Gamma}{A}{B}}{\jvfameq{i}{\Gamma}{B}{A}}
+& & \inference{\jvfameq{i}{\Gamma}{A}{B}\qquad\jvfameq{i}{\Gamma}{B}{C}}{\jvfameq{i}{\Gamma}{A}{C}}\\
+& \inference{\jvterm{i}{\Gamma}{A}{x}}{\jvtermeq{i}{\Gamma}{A}{x}{x}}
+& & \inference{\jvtermeq{i}{\Gamma}{A}{x}{y}}{\jvtermeq{i}{\Gamma}{A}{y}{x}}
+& & \inference{\jvtermeq{i}{\Gamma}{A}{x}{y}\qquad\jvtermeq{i}{\Gamma}{A}{y}{z}}{\jvtermeq{i}{\Gamma}{A}{x}{z}}
+\end{align*}
+\egroup
+
+The following convertibility rules are responsible for the strictness
+of judgmental equality, which sets it apart from equivalences or identifications:
+
+\begin{align*}
+& \inference{\jvctxeq{\Gamma}{\Delta}\qquad\jvfam{i}{\Gamma}{A}}{\jvfam{i}{\Delta}{A}}
+& & \inference{\jvctxeq{\Gamma}{\Delta}\qquad\jvfameq{i}{\Gamma}{A}{B}}{\jvfameq{i}{\Delta}{A}{B}}\\
+& \inference{\jvctxeq{\Gamma}{\Delta}\qquad\jvterm{i}{\Gamma}{A}{x}}{\jvterm{i}{\Delta}{A}{x}}
+& & \inference{\jvctxeq{\Gamma}{\Delta}\qquad\jvtermeq{i}{\Gamma}{A}{x}{y}}{\jvtermeq{i}{\Delta}{A}{x}{y}}\\
+& \inference{\jvfameq{i}{\Gamma}{A}{B}\qquad \jvterm{i}{\Gamma}{A}{x}}{\jvterm{i}{\Gamma}{B}{x}}
+& & \inference{\jvfameq{i}{\Gamma}{A}{B}\qquad\jvtermeq{i}{\Gamma}{A}{x}{y}}{\jvtermeq{i}{\Gamma}{B}{x}{y}}
+\end{align*}
+
+\subsection{The empty context}
+There is an empty context and over any context there is an empty family of
+contexts. We do not assume that the empty context is a type, that would be like
+assuming that the multiplicative unit of a ring is prime. The empty family
+always has a unique term.
+
+\begin{align}
+& \inference{}{\jctx{\emptyc}}\\
+& \inference{\jctx{\Gamma}}{\jvfam{i}{\Gamma}{\emptyf[\Gamma]}}\\
+& \inference{\jctx{\Gamma}}{\jvterm{i}{\Gamma}{\emptyf[\Gamma]}{\emptytm[\Gamma]}}\\
+& \inference{\jvterm{i}{\Gamma}{\emptyf[\Gamma]}{x}}{\jvtermeq{i}{\Gamma}{\emptyf[\Gamma]}{x}{\emptytm[\Gamma]}}
+\end{align}
+
+Moreover, if $\Gamma$ is a context family over the
+empty context, then $\Gamma$ is a context and every context is a context
+family over the empty context. Note that this allows us to speak
+of terms of contexts too.
+
+\begin{align}
+& \inference{\jctx{\Gamma}}{\jvfam{\nameless}{\emptyc}{\Gamma}} 
+& & \inference{\jvfam{\nameless}{\emptyc}{\Gamma}}{\jctx{\Gamma}}\\
+& \inference{\jctxeq{\Gamma}{\Delta}}{\jvfameq{\nameless}{\emptyc}{\Gamma}{\Delta}}
+& & \inference{\jvfameq{\nameless}{\emptyc}{\Gamma}{\Delta}}{\jctxeq{\Gamma}{\Delta}}
+\end{align}
+
+\subsubsection{The empty context is compatible with itslef}
+The empty context $\emptyc$ may be considered as a family of contexts over the empty
+context. When we do this, we get $\emptyf[\emptyc]$.
+\begin{equation}
+\inference{}{\jvfameq{\nameless}{\emptyc}{\emptyc}{\emptyf[\emptyc]}}
+\end{equation}
+In the future, we shall denote $\emptyf[\Gamma]$ by $\emptyf$. The above rule
+guarantees that this will not cause confusion. Likewise, we shall denote
+$\emptytm[\Gamma]$ by $\emptytm$.
+
 \section{Chalmers type theory}

### 2014-03-01 article-models-tt.tex
starting to include the extension subsection to pretty, cool new commands

@@
 
-The identity function of an extended family is the extension of the identity
-functions on the components in the apropriate way. To find out what the
-apropriate way is, note that
-\begin{align*}
-\ctxwk{\ctxext{A}{P}}{\ctxext{A}{P}} & \jdeq \ctxwk{P}{{A}{\ctxext{A}{P}}}\\
-& \jdeq \ctxext{\ctxwk{P}{{A}{A}}}{\ctxwk{P}{{A}{P}}}
-\end{align*}
-We have the term $\jterm{\ctxext{\Gamma}{{A}{P}}}{\ctxwk{P}{A}}{\ctxwk{P}{\idfunc[A]}}$.
-Thus we need to find out what $\subst{\ctxwk{P}{\idfunc[A]}}{\ctxwk{P}{{A}{P}}}$ is:
-\begin{align*}
-\subst{\ctxwk{P}{\idfunc[A]}}{\ctxwk{P}{{A}{P}}} & \jdeq \ctxwk{P}{\subst{\idfunc[A]}{\ctxwk{A}{P}}}\\
-& \jdeq \ctxwk{P}{P},
-\end{align*}
-where we find the term $\idfunc[P]$. Therefore, we have the following inference rule:
-\begin{equation}\label{idfunc-ext-comp}
-\inference{\jfam{\Gamma}{A}\quad\jfam{\ctxext{\Gamma}{A}}{P}}{\jtermeq{\ctxext{\Gamma}{{A}{P}}}{\ctxwk{\ctxext{A}{P}}{\ctxext{A}{P}}}{\idfunc[\ctxext{A}{P}]}{\subst{\idfunc[P]}{{\ctxwk{P}{\idfunc[A]}}{\idfunc[\ctxwk{\ctxext{A}{P}}{\ctxext{A}{P}}]}}}}
-\end{equation}
+We won't state a compatibility rule stating that the identity function is
+compatible with extension because we will be able to prove that. Instead, we
+will just state the compatibility rules for the identity function combined with
+weakening and with substitution.
 
@@
 
-There is also a decomposition of terms $\jterm{\Gamma}{\ctxext{A}{P}}{w}$ into
-components: the projections.
+We have seen above that the pairing function into $\ctxext{A}{P}$ is just the identity function on
+$\ctxext{A}{P}$. To analyze the pairing functin a little further, we will also
+need the projection maps from $\ctxext{A}{P}$ to $A$ and from $\ctxext{A}{P}$
+to $P$. We will now define these and see that the identity function of an
+extended family is the extension (or pairing) of the identity
+functions on the components in the apropriate way.
+
+To find out what the
+apropriate way is, note that
+\begin{align*}
+\ctxwk{\ctxext{A}{P}}{\ctxext{A}{P}} & \jdeq \ctxwk{P}{{A}{\ctxext{A}{P}}}\\
+& \jdeq \ctxext{\ctxwk{P}{{A}{A}}}{\ctxwk{P}{{A}{P}}}
+\end{align*}
+We have the term $\jterm{\ctxext{\Gamma}{{A}{P}}}{\ctxwk{P}{A}}{\ctxwk{P}{\idfunc[A]}}$.
+Thus we need to find out what $\subst{\ctxwk{P}{\idfunc[A]}}{\ctxwk{P}{{A}{P}}}$ is:
+\begin{align*}
+\subst{\ctxwk{P}{\idfunc[A]}}{\ctxwk{P}{{A}{P}}} & \jdeq \ctxwk{P}{\subst{\idfunc[A]}{\ctxwk{A}{P}}}\\
+& \jdeq \ctxwk{P}{P},
+\end{align*}
+where we find the term $\idfunc[P]$. Therefore we define:
 
@@
 \end{defn}
+
+We find the following inference rule, which expresses that the identity function
+is compatible with extension:
+
+\begin{lem}
+For any $\jfam{\Gamma}{A}$ and $\jfam{\ctxext{\Gamma}{A}}{P}$ we have
+\begin{equation}\label{idfunc-ext-comp}
+\inference{\jfam{\Gamma}{A}\quad\jfam{\ctxext{\Gamma}{A}}{P}}{\jtermeq{\ctxext{\Gamma}{{A}{P}}}{\ctxwk{\ctxext{A}{P}}{\ctxext{A}{P}}}{\idfunc[\ctxext{A}{P}]}{\subst{\cprojsndf{A}{P}}{{\cprojfstf{A}{P}}{\idfunc[\ctxwk{\ctxext{A}{P}}{\ctxext{A}{P}}]}}}}
+\end{equation}
+\end{lem}
+
+\begin{proof}
+Consider the families $\jfam{\Gamma}{A}$ and $\jfam{\ctxext{\Gamma}{A}}{P}$. Then
+we have the judgmental equalities
+\begin{align*}
+\subst{\cprojsndf{A}{P}}{{\cprojfstf{A}{P}}{\idfunc[\ctxwk{\ctxext{A}{P}}{\ctxext{A}{P}}]}}
+& \jdeq \subst{\idfunc[P]}{{\ctxwk{P}{\idfunc[A]}}{\idfunc[\ctxwk{\ctxext{A}{P}}{\ctxext{A}{P}}]}}\\
+& \jdeq \subst{\idfunc[P]}{{\ctxwk{P}{\idfunc[A]}}{\ctxwk{\ctxext{A}{P}}{\idfunc[\ctxext{A}{P}]}}}\\
+& \jdeq \subst{\idfunc[P]}{{\ctxwk{P}{\idfunc[A]}}{\ctxwk{P}{{A}{\idfunc[\ctxext{A}{P}]}}}}\\
+& \jdeq \subst{\idfunc[P]}{\ctxwk{P}{\subst{\idfunc[A]}{\ctxwk{A}{\idfunc[\ctxext{A}{P}]}}}}\\
+& \jdeq \subst{\idfunc[A]}{\ctxwk{A}{\idfunc[\ctxext{A}{P}]}}\\
+& \jdeq \idfunc[\ctxext{A}{P}].
+\end{align*}
+\end{proof}
 
@@
 
-Pretty type theory is pretty as an ugly little duckling. At first, it is actually uglier than
+Pretty type theory is pretty as the ugly little duckling that grew up. At first, it is actually uglier than
 structural type theory because one has to keep track of a variable nobody is
@@
 
+\subsection{Extension}
+We introduce extension which not only extends a context $\Gamma$ and a family
+$A$ over it to a context $\ctxext{\Gamma}{A}$, but which also extends a family $A$
+in context $\Gamma$ and a family $P$ over it to a family $\ctxext{A}{P}$ over context
+$\Gamma$. We do this to ensure that all of type theory can be done in a context.
+For instance, we could say (1) that a context in context $\Gamma$ is the same thing
+as a family over $\Gamma$; (2) When $A$ is a context in this sense, a family over
+$A$ is the same thing as a family $P$ over $\ctxext{\Gamma}{A}$ and 
+(3) when $P$ is a family over $A$ in this sense, a term of $P$ keeps its original meaning.
+
+\begin{align}
+& \inference{\jvfam{i}{\Gamma}{A}}{\jvfamcombi{{i}{x}}{{\Gamma}{A}}{P}}
+& & \inference{\jctxeq{\Gamma}{\Delta}\qquad\jfameq{\Gamma}{A}{B}}{\jctxeq{\ctxext{\Gamma}{A}}{\ctxext{\Delta}{B}}}\\
+& \inference{\jfam{\ctxext{\Gamma}{A}}{P}}{\jfam{\Gamma}{\ctxext{A}{P}}}
+& & \inference{\jfameq{\Gamma}{A}{B}\qquad\jfameq{\ctxext{\Gamma}{A}}{P}{Q}}{\jfameq{\Gamma}{\ctxext{A}{P}}{\ctxext{B}{Q}}}
+\end{align}
+
+\subsubsection{Extension is compatible with the empty context}
+The following rule asserts that extension by $\emptyc$ leaves the contexts unchanged.
+\begin{align}
+& \inference{\jctx{\Gamma}}{\jctxeq{\ctxext{\emptyc}{\Gamma}}{\Gamma}}\\
+& \inference{\jctx{\Gamma}}{\jctxeq{\ctxext{\Gamma}{\emptyf}}{\Gamma}}\\
+& \inference{\jfam{\Gamma}{A}}{\jfameq{\Gamma}{\ctxext{\emptyf}{A}}{A}}
+\end{align}
+
+\subsubsection{Extension is compatible with itself}\label{comp-ee}
+The inference rules asserting that extension is compatible with itself assert
+that contexts are unstructured lists of type declarations. This rule is
+unavoidable if we want that for a family $A$ in context $\Gamma$, a family over
+$A$ is the same thing as a family over $\ctxext{\Gamma}{A}$.
+
+\begin{align}
+& \inference{\jfam{\Gamma}{A}\qquad\jfam{\ctxext{\Gamma}{A}}{P}}
+  {\jctxeq{\ctxext{{\Gamma}{A}}{P}}{\ctxext{\Gamma}{{A}{P}}}}\\
+& \inference{\jfam{\ctxext{\Gamma}{A}}{P}\qquad\jfam{\ctxext{{\Gamma}{A}}{P}}{Q}}
+  {\jfameq{\Gamma}{\ctxext{{A}{P}}{Q}}{\ctxext{A}{{P}{Q}}}}
+\end{align}
+
 \section{Chalmers type theory}

### 2014-03-02 article-models-tt.tex
small impro
using the new judgment macro

@@
 that various other elements have to be added yet, such as terms
-$\jterm{\ctxext{\Gamma}{\subst{x}{P}}}{\ctxwk{\subst{x}{P}}{\ctxext{A}{P}}}{i}$
+$\jterm{{\Gamma}{\subst{x}{P}}}{\ctxwk{\subst{x}{P}}{\ctxext{A}{P}}}{i}$
 including the fiber $\subst{x}{P}$ into the extended family $\ctxext{A}{P}$ over
@@
 & & \inference{\jctxeq{\Gamma}{\Delta}\qquad\jfameq{\Gamma}{A}{B}}{\jctxeq{\ctxext{\Gamma}{A}}{\ctxext{\Delta}{B}}}\\
-& \inference{\jfam{\ctxext{\Gamma}{A}}{P}}{\jfam{\Gamma}{\ctxext{A}{P}}}
-& & \inference{\jfameq{\Gamma}{A}{B}\qquad\jfameq{\ctxext{\Gamma}{A}}{P}{Q}}{\jfameq{\Gamma}{\ctxext{A}{P}}{\ctxext{B}{Q}}}
+& \inference{\jfam{{\Gamma}{A}}{P}}{\jfam{\Gamma}{\ctxext{A}{P}}}
+& & \inference{\jfameq{\Gamma}{A}{B}\qquad\jfameq{{\Gamma}{A}}{P}{Q}}{\jfameq{\Gamma}{\ctxext{A}{P}}{\ctxext{B}{Q}}}
 \end{align}
@@
 \begin{align}
-& \inference{\jfam{\Gamma}{A}\qquad\jfam{\ctxext{\Gamma}{A}}{P}}
+& \inference{\jfam{\Gamma}{A}\qquad\jfam{{\Gamma}{A}}{P}}
   {\jctxeq{\ctxext{{\Gamma}{A}}{P}}{\ctxext{\Gamma}{{A}{P}}}}\\
-& \inference{\jfam{\ctxext{\Gamma}{A}}{P}\qquad\jfam{\ctxext{{\Gamma}{A}}{P}}{Q}}
+& \inference{\jfam{{\Gamma}{A}}{P}\qquad\jfam{{{\Gamma}{A}}{P}}{Q}}
   {\jfameq{\Gamma}{\ctxext{{A}{P}}{Q}}{\ctxext{A}{{P}{Q}}}}
@@
 \begin{align}
-& \inference{\jfam{\Gamma}{B}}{\jfam{\ctxext{\Gamma}{A}}{\ctxwk{A}{B}}}
-& & \inference{\jfameq{\Gamma}{B}{B'}}{\jfameq{\ctxext{\Gamma}{A}}{\ctxwk{A}{B}}{\ctxwk{A'}{B'}}}\\
-& \inference{\jfam{\ctxext{\Gamma}{B}}{Q}}
-{\jfam{\ctxext{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{Q}}}
-& & \inference{\jfameq{\ctxext{\Gamma}{B}}{Q}{Q'}}
-{\jfameq{\ctxext{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{Q}}{\ctxwk{A'}{Q'}}}\\
-& \inference{\jterm{\Gamma}{B}{y}}{\jterm{\ctxext{\Gamma}{A}}{\ctxwk{A}{B}}{\ctxwk{A}{y}}}
-& & \inference{\jtermeq{\Gamma}{B}{y}{y'}}{\jtermeq{\ctxext{\Gamma}{A}}{\ctxwk{A}{B}}{\ctxwk{A}{y}}{\ctxwk{A'}{y'}}}\\
-& \inference{\jterm{\ctxext{\Gamma}{B}}{Q}{g}}{\jterm{\ctxext{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{Q}}{\ctxwk{A}{g}}}
-& & \inference{\jtermeq{\ctxext{\Gamma}{B}}{Q}{g}{g'}}
-{\jtermeq{\ctxext{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{Q}}{\ctxwk{A}{g}}{\ctxwk{A'}{g'}}}
+& \inference{\jfam{\Gamma}{B}}{\jfam{{\Gamma}{A}}{\ctxwk{A}{B}}}
+& & \inference{\jfameq{\Gamma}{B}{B'}}{\jfameq{{\Gamma}{A}}{\ctxwk{A}{B}}{\ctxwk{A'}{B'}}}\\
+& \inference{\jfam{{\Gamma}{B}}{Q}}
+{\jfam{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{Q}}}
+& & \inference{\jfameq{{\Gamma}{B}}{Q}{Q'}}
+{\jfameq{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{Q}}{\ctxwk{A'}{Q'}}}\\
+& \inference{\jterm{\Gamma}{B}{y}}{\jterm{{\Gamma}{A}}{\ctxwk{A}{B}}{\ctxwk{A}{y}}}
+& & \inference{\jtermeq{\Gamma}{B}{y}{y'}}{\jtermeq{{\Gamma}{A}}{\ctxwk{A}{B}}{\ctxwk{A}{y}}{\ctxwk{A'}{y'}}}\\
+& \inference{\jterm{{\Gamma}{B}}{Q}{g}}{\jterm{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{Q}}{\ctxwk{A}{g}}}
+& & \inference{\jtermeq{{\Gamma}{B}}{Q}{g}{g'}}
+{\jtermeq{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{Q}}{\ctxwk{A}{g}}{\ctxwk{A'}{g'}}}
 \end{align}
@@
 & \inference{\jctx{\Gamma}}{\jfameq{\Gamma}{\ctxwk{\Gamma}{\emptyc}}{\emptyf}}\\
-& \inference{\jfam{\Gamma}{A}}{\jfameq{\ctxext{\Gamma}{A}}{\ctxwk{A}{\emptyf}}{\emptyf}}
+& \inference{\jfam{\Gamma}{A}}{\jfameq{{\Gamma}{A}}{\ctxwk{A}{\emptyf}}{\emptyf}}
 \end{align}
@@
 & \inference{\jterm{\Gamma}{B}{y}}{\jtermeq{\Gamma}{B}{\ctxwk{\emptyf}{y}}{y}}\\
-& \inference{\jfam{\ctxext{\Gamma}{B}}{Q}}{\jfameq{\ctxext{\Gamma}{B}}{\ctxwk{\emptyf}{Q}}{Q}}\\
-& \inference{\jterm{\ctxext{\Gamma}{B}}{Q}{g}}{\jtermeq{\ctxext{\Gamma}{B}}{Q}{\ctxwk{\emptyf}{g}}{g}}
+& \inference{\jfam{{\Gamma}{B}}{Q}}{\jfameq{{\Gamma}{B}}{\ctxwk{\emptyf}{Q}}{Q}}\\
+& \inference{\jterm{{\Gamma}{B}}{Q}{g}}{\jtermeq{{\Gamma}{B}}{Q}{\ctxwk{\emptyf}{g}}{g}}
 \end{align}
@@
 & \inference
-  {\jfam{\ctxext{{\Gamma}{B}}{Q}}{R}}
+  {\jfam{{{\Gamma}{B}}{Q}}{R}}
   {\jfameq
@@
 \begin{tikzcd}[column sep=huge]
-\jfam{\Gamma}{\blank} \ar{r}{C\mapsto\ctxwk{B}{C}} \ar{d}[swap]{C\mapsto\ctxwk{A}{C}} & \jfam{\ctxext{\Gamma}{B}}{\blank} \ar{d}{Q\mapsto\ctxwk{A}{Q}}\\
-\jfam{\ctxext{\Gamma}{A}}{\blank} \ar{r}[swap]{P\mapsto\ctxwk{{A}{B}}{P}} & \jfam{\ctxext{{\Gamma}{A}}{\ctxwk{A}{B}}}{\blank}
+\jfam{\Gamma}{\blank} \ar{r}{C\mapsto\ctxwk{B}{C}} \ar{d}[swap]{C\mapsto\ctxwk{A}{C}} & \jfam{{\Gamma}{B}}{\blank} \ar{d}{Q\mapsto\ctxwk{A}{Q}}\\
+\jfam{{\Gamma}{A}}{\blank} \ar{r}[swap]{P\mapsto\ctxwk{{A}{B}}{P}} & \jfam{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\blank}
 \end{tikzcd}
@@
 \begin{align}
-& \inference{\jfam{\Gamma}{A}\qquad\jfam{\Gamma}{B}\qquad\jfam{\ctxext{\Gamma}{C}}{R}}
-          {\jfameq{\ctxext{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{{A}{B}}{{A}{C}}}}
+& \inference{\jfam{\Gamma}{A}\qquad\jfam{\Gamma}{B}\qquad\jfam{{\Gamma}{C}}{R}}
+          {\jfameq{{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{{A}{B}}{{A}{C}}}}
             {\ctxwk{A}{{B}{R}}}
             {\ctxwk{{A}{B}}{{A}{R}}}}\label{comp-ww-f}\\
-& \inference{\jfam{\Gamma}{A}\qquad\jfam{\Gamma}{B}\qquad\jterm{\ctxext{\Gamma}{C}}{R}{t}}
-          {\jtermeq{\ctxext{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{{A}{B}}{{A}{C}}}}
+& \inference{\jfam{\Gamma}{A}\qquad\jfam{\Gamma}{B}\qquad\jterm{{\Gamma}{C}}{R}{t}}
+          {\jtermeq{{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{{A}{B}}{{A}{C}}}}
             {\ctxwk{{A}{B}}{{A}{R}}}{\ctxwk{A}{{B}{t}}}{\ctxwk{{A}{B}}{{A}{t}}}}
@@
 In the following inference rules we assume that
-$\jfam{\Gamma}{A}$ and $\jfam{\ctxext{\Gamma}{A}}{P}$. 
+$\jfam{\Gamma}{A}$ and $\jfam{{\Gamma}{A}}{P}$. 
 \begin{align}
-& \inference{\jfam{\ctxext{\Gamma}{B}}{Q}}
-    {\jfameq{\ctxext{{{\Gamma}{A}}{P}}{\ctxwk{P}{{A}{B}}}}{\ctxwk{\ctxext{A}{P}}{Q}}{\ctxwk{P}{{A}{Q}}}}\label{comp-ew-f}\\
-& \inference{\jterm{\ctxext{\Gamma}{B}}{Q}{g}}
-    {\jtermeq{\ctxext{{{\Gamma}{A}}{P}}{\ctxwk{P}{{A}{B}}}}{\ctxwk{P}{{A}{Q}}}{\ctxwk{\ctxext{A}{P}}{g}}{\ctxwk{P}{{A}{g}}}}\label{comp-ew-t}
+& \inference{\jfam{{\Gamma}{B}}{Q}}
+    {\jfameq{{{{\Gamma}{A}}{P}}{\ctxwk{P}{{A}{B}}}}{\ctxwk{\ctxext{A}{P}}{Q}}{\ctxwk{P}{{A}{Q}}}}\label{comp-ew-f}\\
+& \inference{\jterm{{\Gamma}{B}}{Q}{g}}
+    {\jtermeq{{{{\Gamma}{A}}{P}}{\ctxwk{P}{{A}{B}}}}{\ctxwk{P}{{A}{Q}}}{\ctxwk{\ctxext{A}{P}}{g}}{\ctxwk{P}{{A}{g}}}}\label{comp-ew-t}
 \end{align}
@@
 \begin{align}
-& \inference{\jfam{\ctxext{\Gamma}{A}}{P}}{\jfam{\Gamma}{\subst{x}{P}}}
-& & \inference{\jfameq{\ctxext{\Gamma}{A}}{P}{P'}}{\jfameq{\Gamma}{\subst{x}{P}}{\subst{x'}{P'}}}\\
-& \inference{\jfam{\ctxext{{\Gamma}{A}}{P}}{Q}}{\jfam{\ctxext{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}}
-& & \inference{\jfameq{\ctxext{{\Gamma}{A}}{P}}{Q}{Q'}}{\jfameq{\ctxext{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}{\subst{x'}{Q'}}}\\
-& \inference{\jterm{\ctxext{\Gamma}{A}}{P}{f}}{\jterm{\Gamma}{\subst{x}{P}}{\subst{x}{f}}}
-& & \inference{\jtermeq{\ctxext{\Gamma}{A}}{P}{f}{f'}}{\jtermeq{\Gamma}{\subst{x}{P}}{\subst{x}{f}}{\subst{x'}{f'}}}\\
-& \inference{\jterm{\ctxext{{\Gamma}{A}}{P}}{Q}{g}}{\jterm{\ctxext{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}{\subst{x}{g}}}
-& &\inference{\jtermeq{\ctxext{{\Gamma}{A}}{P}}{Q}{g}{g'}}{\jtermeq{\ctxext{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}{\subst{x}{g}}{\subst{x'}{g'}}}
+& \inference{\jfam{{\Gamma}{A}}{P}}{\jfam{\Gamma}{\subst{x}{P}}}
+& & \inference{\jfameq{{\Gamma}{A}}{P}{P'}}{\jfameq{\Gamma}{\subst{x}{P}}{\subst{x'}{P'}}}\\
+& \inference{\jfam{{{\Gamma}{A}}{P}}{Q}}{\jfam{{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}}
+& & \inference{\jfameq{{{\Gamma}{A}}{P}}{Q}{Q'}}{\jfameq{{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}{\subst{x'}{Q'}}}\\
+& \inference{\jterm{{\Gamma}{A}}{P}{f}}{\jterm{\Gamma}{\subst{x}{P}}{\subst{x}{f}}}
+& & \inference{\jtermeq{{\Gamma}{A}}{P}{f}{f'}}{\jtermeq{\Gamma}{\subst{x}{P}}{\subst{x}{f}}{\subst{x'}{f'}}}\\
+& \inference{\jterm{{{\Gamma}{A}}{P}}{Q}{g}}{\jterm{{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}{\subst{x}{g}}}
+& &\inference{\jtermeq{{{\Gamma}{A}}{P}}{Q}{g}{g'}}{\jtermeq{{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}{\subst{x}{g}}{\subst{x'}{g'}}}
 \end{align}
@@
 & \inference{\jterm{\Gamma}{A}{x}}{\jfameq{\Gamma}{\subst{x}{\emptyf}}{\emptyf}}\\
-& \inference{\jterm{\Gamma}{A}{x}\quad\jfam{\ctxext{\Gamma}{A}}{P}}{\jfameq{\ctxext{\Gamma}{\subst{x}{P}}}{\subst{x}{\emptyf}}{\emptyf}}
+& \inference{\jterm{\Gamma}{A}{x}\quad\jfam{{\Gamma}{A}}{P}}{\jfameq{{\Gamma}{\subst{x}{P}}}{\subst{x}{\emptyf}}{\emptyf}}
 \end{align}
@@
 & \inference{\jterm{\Gamma}{A}{x}}{\jtermeq{\Gamma}{A}{\subst{\emptytm}{x}}{x}}\\
-& \inference{\jfam{\ctxext{\Gamma}{A}}{P}}{\jfameq{\ctxext{\Gamma}{A}}{\subst{\emptytm}{P}}{P}}\\
-& \inference{\jterm{\ctxext{\Gamma}{A}}{P}{f}}{\jtermeq{\ctxext{\Gamma}{A}}{P}{\subst{\emptytm}{f}}{f}}.
+& \inference{\jfam{{\Gamma}{A}}{P}}{\jfameq{{\Gamma}{A}}{\subst{\emptytm}{P}}{P}}\\
+& \inference{\jterm{{\Gamma}{A}}{P}{f}}{\jtermeq{{\Gamma}{A}}{P}{\subst{\emptytm}{f}}{f}}.
 \end{align}
@@
 \begin{align}
-& \inference{\jfam{\ctxext{{{\Gamma}{A}}{P}}{Q}}{R}}
-  {\jfameq{\ctxext{\Gamma}{\subst{x}{P}}}{\subst{x}{\ctxext{Q}{R}}}{\ctxext{\subst{x}{Q}}{\subst{x}{R}}}}
+& \inference{\jfam{{{{\Gamma}{A}}{P}}{Q}}{R}}
+  {\jfameq{{\Gamma}{\subst{x}{P}}}{\subst{x}{\ctxext{Q}{R}}}{\ctxext{\subst{x}{Q}}{\subst{x}{R}}}}
 \end{align}
@@
 that the following diagram commutes for any $\jterm{\Gamma}{A}{x}$ and any
-$\jfam{\ctxext{\Gamma}{A}}{P}$:
+$\jfam{{\Gamma}{A}}{P}$:
 \begin{equation*}
 \begin{tikzcd}[column sep=huge]
-\jfam{\ctxext{\Gamma}{A}}{\blank} \ar{d}[swap]{Q\mapsto\subst{x}{Q}} \ar{r}{Q\mapsto\ctxwk{P}{Q}} & \jfam{\ctxext{{\Gamma}{A}}{P}}{\blank} \ar{d}{R\mapsto\subst{x}{R}}\\ 
-\jfam{\Gamma}{\blank} \ar{r}[swap]{B\mapsto\ctxwk{\subst{x}{P}}{B}} & \jfam{\ctxext{\Gamma}{\subst{x}{P}}}{\blank}
+\jfam{{\Gamma}{A}}{\blank} \ar{d}[swap]{Q\mapsto\subst{x}{Q}} \ar{r}{Q\mapsto\ctxwk{P}{Q}} & \jfam{{{\Gamma}{A}}{P}}{\blank} \ar{d}{R\mapsto\subst{x}{R}}\\ 
+\jfam{\Gamma}{\blank} \ar{r}[swap]{B\mapsto\ctxwk{\subst{x}{P}}{B}} & \jfam{{\Gamma}{\subst{x}{P}}}{\blank}
 \end{tikzcd}
@@
 In the following inference rules we assume that $\jterm{\Gamma}{A}{x}$ and
-$\jfam{\ctxext{\Gamma}{A}}{P}$:
+$\jfam{{\Gamma}{A}}{P}$:
 \begin{align}
 & \inference
-  {\jfam{\ctxext{{\Gamma}{A}}{Q}}{R}}
-  {\jfameq{\ctxext{{\Gamma}{\subst{x}{P}}}{\subst{x}{\ctxwk{P}{Q}}}}{\subst{x}{\ctxwk{P}{R}}}{\ctxwk{\subst{x}{P}}{\subst{x}{R}}}}\label{comp-sw-f}\\
+  {\jfam{{{\Gamma}{A}}{Q}}{R}}
+  {\jfameq{{{\Gamma}{\subst{x}{P}}}{\subst{x}{\ctxwk{P}{Q}}}}{\subst{x}{\ctxwk{P}{R}}}{\ctxwk{\subst{x}{P}}{\subst{x}{R}}}}\label{comp-sw-f}\\
 & \inference
-    {\jterm{\ctxext{{\Gamma}{A}}{Q}}{R}{h}}
+    {\jterm{{{\Gamma}{A}}{Q}}{R}{h}}
     {\jtermeq
@@
 In the following inference rules we assume
-$\jterm{\Gamma}{A}{x}$ and $\jterm{\ctxext{\Gamma}{A}}{P}{f}$.
+$\jterm{\Gamma}{A}{x}$ and $\jterm{{\Gamma}{A}}{P}{f}$.
 
 \begin{align}
-&\inference{\jfam{\ctxext{{{\Gamma}{A}}{P}}{Q}}{R}}
-{\jfameq{\ctxext{\Gamma}{\subst{x}{{f}{Q}}}}{\subst{x}{{f}{R}}}{\subst{{x}{f}}{{x}{R}}}}\label{comp-ss-f}\\
-&\inference{\jterm{\ctxext{{{\Gamma}{A}}{P}}{Q}}{R}{h}}
-{\jtermeq{\ctxext{\Gamma}{\subst{x}{{f}{Q}}}}{\subst{x}{{f}{R}}}{\subst{x}{{f}{h}}}{\subst{{x}{f}}{{x}{h}}}}\label{comp-ss-t}
+&\inference{\jfam{{{{\Gamma}{A}}{P}}{Q}}{R}}
+{\jfameq{{\Gamma}{\subst{x}{{f}{Q}}}}{\subst{x}{{f}{R}}}{\subst{{x}{f}}{{x}{R}}}}\label{comp-ss-f}\\
+&\inference{\jterm{{{{\Gamma}{A}}{P}}{Q}}{R}{h}}
+{\jtermeq{{\Gamma}{\subst{x}{{f}{Q}}}}{\subst{x}{{f}{R}}}{\subst{x}{{f}{h}}}{\subst{{x}{f}}{{x}{h}}}}\label{comp-ss-t}
 \end{align}
@@
 \begin{align}
-& \inference{\jfam{\ctxext{{\Gamma}{B}}{Q}}{R}}{\jfameq{\ctxext{{\Gamma}{A}}{\ctxwk{A}{\subst{y}{Q}}}}{\ctxwk{A}{\subst{y}{R}}}{\subst{\ctxwk{A}{y}}{\ctxwk{A}{R}}}}\label{comp-ws-f}\\
-& \inference{\jterm{\ctxext{{\Gamma}{B}}{Q}}{R}{h}}{\jtermeq{\ctxext{{\Gamma}{A}}{\ctxwk{A}{\subst{y}{Q}}}}{\ctxwk{A}{\subst{y}{R}}}{\ctxwk{A}{\subst{y}{h}}}{\subst{\ctxwk{A}{y}}{\ctxwk{A}{h}}}}\label{comp-ws-t}
+& \inference{\jfam{{{\Gamma}{B}}{Q}}{R}}{\jfameq{{{\Gamma}{A}}{\ctxwk{A}{\subst{y}{Q}}}}{\ctxwk{A}{\subst{y}{R}}}{\subst{\ctxwk{A}{y}}{\ctxwk{A}{R}}}}\label{comp-ws-f}\\
+& \inference{\jterm{{{\Gamma}{B}}{Q}}{R}{h}}{\jtermeq{{{\Gamma}{A}}{\ctxwk{A}{\subst{y}{Q}}}}{\ctxwk{A}{\subst{y}{R}}}{\ctxwk{A}{\subst{y}{h}}}{\subst{\ctxwk{A}{y}}{\ctxwk{A}{h}}}}\label{comp-ws-t}
 \end{align}
@@
 \begin{align}
-& \inference{\jfam{\Gamma}{A}\qquad\jfam{\ctxext{\Gamma}{B}}{Q}\qquad\jterm{\Gamma}{A}{x}}{\jfameq{\ctxext{\Gamma}{B}}{\subst{x}{\ctxwk{A}{Q}}}{Q}}\label{defn-ws-3}\\
-& \inference{\jterm{\Gamma}{A}{x}\qquad\jterm{\ctxext{\Gamma}{B}}{Q}{g}}{\jtermeq{\ctxext{\Gamma}{B}}{Q}{\subst{x}{\ctxwk{A}{g}}}{g}}\label{defn-ws-4}
+& \inference{\jfam{\Gamma}{A}\qquad\jfam{{\Gamma}{B}}{Q}\qquad\jterm{\Gamma}{A}{x}}{\jfameq{{\Gamma}{B}}{\subst{x}{\ctxwk{A}{Q}}}{Q}}\label{defn-ws-3}\\
+& \inference{\jterm{\Gamma}{A}{x}\qquad\jterm{{\Gamma}{B}}{Q}{g}}{\jtermeq{{\Gamma}{B}}{Q}{\subst{x}{\ctxwk{A}{g}}}{g}}\label{defn-ws-4}
 \end{align}
@@
 \begin{equation*}
-\jfameq{\ctxext{\Gamma}{A}}{\subst{f}{\ctxwk{A}{{B}{C}}}}{\ctxwk{A}{C}}
+\jfameq{{\Gamma}{A}}{\subst{f}{\ctxwk{A}{{B}{C}}}}{\ctxwk{A}{C}}
 \end{equation*}
-for any three families $A$, $B$ and $C$ in context $\Gamma$ and any $\jterm{\ctxext{\Gamma}{A}}{\ctxwk{A}{B}}{f}$.
-It follows that for $\jterm{\ctxext{\Gamma}{B}}{\ctxwk{B}{C}}{g}$ we have
+for any three families $A$, $B$ and $C$ in context $\Gamma$ and any $\jterm{{\Gamma}{A}}{\ctxwk{A}{B}}{f}$.
+It follows that for $\jterm{{\Gamma}{B}}{\ctxwk{B}{C}}{g}$ we have
 \begin{equation*}
-\jterm{\ctxext{\Gamma}{A}}{\ctxwk{A}{C}}{\subst{f}{\ctxwk{A}{g}}}
+\jterm{{\Gamma}{A}}{\ctxwk{A}{C}}{\subst{f}{\ctxwk{A}{g}}}
 \end{equation*}
 The term $\jcomp{A}{f}{g}$ is the composition of $g$ with $f$. We may call
-a term $\jterm{\ctxext{\Gamma}{A}}{\ctxwk{A}{B}}{f}$ a morphism from $A$ to $B$.
+a term $\jterm{{\Gamma}{A}}{\ctxwk{A}{B}}{f}$ a morphism from $A$ to $B$.
 
@@
 \begin{align}
-& \inference{\jfam{\Gamma}{A}}{\jterm{\ctxext{\Gamma}{A}}{\ctxwk{A}{A}}{\idfunc[A]}}
-& & \inference{\jfameq{\Gamma}{A}{A'}}{\jtermeq{\ctxext{\Gamma}{A}}{\ctxwk{A}{A}}{\idfunc[A]}{\idfunc[A']}}
+& \inference{\jfam{\Gamma}{A}}{\jterm{{\Gamma}{A}}{\ctxwk{A}{A}}{\idfunc[A]}}
+& & \inference{\jfameq{\Gamma}{A}{A'}}{\jtermeq{{\Gamma}{A}}{\ctxwk{A}{A}}{\idfunc[A]}{\idfunc[A']}}
 \end{align}
@@
 & \inference{\jterm{\Gamma}{A}{x}}{\jtermeq{\Gamma}{A}{\subst{x}{\idfunc[A]}}{x}}\label{idfunc-subst-defn}\\
-& \inference{\jfam{\ctxext{\Gamma}{A}}{P}}{\jfameq{\ctxext{\Gamma}{A}}{\subst{\idfunc[A]}{\ctxwk{A}{P}}}{P}}\label{idfunc-wk-defn}\\
-& \inference{\jfam{\ctxext{\Gamma}{A}}{P}}{\jfameq{\ctxext{\Gamma}{A}}{\subst{\idfunc[A]}{\ctxwk{{A}{A}}{P}}}{P}}\label{idfunc-wk-defn2}\\
-& \inference{\jterm{\ctxext{\Gamma}{A}}{P}{f}}{\jtermeq{\ctxext{\Gamma}{A}}{P}{\subst{\idfunc[A]}{\ctxwk{A}{f}}}{f}}\label{idfunc-precomp} \\
-& \inference{\jterm{\ctxext{\Gamma}{A}}{P}{f}}{\jtermeq{\ctxext{\Gamma}{A}}{P}{\subst{\idfunc[A]}{\ctxwk{{A}{A}}{f}}}{f}}\label{idfunc-precomp} \\
-& \inference{\jterm{\ctxext{\Gamma}{B}}{\ctxwk{B}{A}}{g}}{\jtermeq{\ctxext{\Gamma}{B}}{\ctxwk{B}{A}}{\subst{g}{\ctxwk{B}{\idfunc[A]}}}{g}}\label{idfunc-postcomp}
+& \inference{\jfam{{\Gamma}{A}}{P}}{\jfameq{{\Gamma}{A}}{\subst{\idfunc[A]}{\ctxwk{A}{P}}}{P}}\label{idfunc-wk-defn}\\
+& \inference{\jfam{{\Gamma}{A}}{P}}{\jfameq{{\Gamma}{A}}{\subst{\idfunc[A]}{\ctxwk{{A}{A}}{P}}}{P}}\label{idfunc-wk-defn2}\\
+& \inference{\jterm{{\Gamma}{A}}{P}{f}}{\jtermeq{{\Gamma}{A}}{P}{\subst{\idfunc[A]}{\ctxwk{A}{f}}}{f}}\label{idfunc-precomp} \\
+& \inference{\jterm{{\Gamma}{A}}{P}{f}}{\jtermeq{{\Gamma}{A}}{P}{\subst{\idfunc[A]}{\ctxwk{{A}{A}}{f}}}{f}}\label{idfunc-precomp} \\
+& \inference{\jterm{{\Gamma}{B}}{\ctxwk{B}{A}}{g}}{\jtermeq{{\Gamma}{B}}{\ctxwk{B}{A}}{\subst{g}{\ctxwk{B}{\idfunc[A]}}}{g}}\label{idfunc-postcomp}
 \end{align}
@@
 \begin{equation}\label{idfunc-wk-comp}
-\inference{\jfam{\Gamma}{A}\quad\jfam{\Gamma}{B}}{\jtermeq{\ctxext{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{{A}{B}}{{A}{B}}}{\idfunc[\ctxwk{A}{B}]}{\ctxwk{A}{\idfunc[B]}}}
+\inference{\jfam{\Gamma}{A}\quad\jfam{\Gamma}{B}}{\jtermeq{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{{A}{B}}{{A}{B}}}{\idfunc[\ctxwk{A}{B}]}{\ctxwk{A}{\idfunc[B]}}}
 \end{equation}
@@
 \begin{equation}\label{idfunc-subst-comp}
-\inference{\jterm{\Gamma}{A}{x}\quad\jfam{\ctxext{\Gamma}{A}}{P}}{\jtermeq{\ctxext{\Gamma}{\subst{x}{P}}}{\ctxwk{\subst{x}{P}}{\subst{x}{P}}}{\idfunc[\subst{x}{P}]}{\subst{x}{\idfunc[P]}}}
+\inference{\jterm{\Gamma}{A}{x}\quad\jfam{{\Gamma}{A}}{P}}{\jtermeq{{\Gamma}{\subst{x}{P}}}{\ctxwk{\subst{x}{P}}{\subst{x}{P}}}{\idfunc[\subst{x}{P}]}{\subst{x}{\idfunc[P]}}}
 \end{equation}
@@
 \begin{align*}
-& \inference{\jterm{\Gamma}{A}{x}\quad\jterm{\Gamma}{\subst{x}{P}}{u}\quad\jfam{\ctxext{{\Gamma}{A}}{P}}{Q}}
+& \inference{\jterm{\Gamma}{A}{x}\quad\jterm{\Gamma}{\subst{x}{P}}{u}\quad\jfam{{{\Gamma}{A}}{P}}{Q}}
   {\jfameq{\Gamma}{\subst{\ctxext{x}{u}}{Q}}{\subst{u}{{x}{Q}}}}\\
-& \inference{\jterm{\Gamma}{A}{x}\quad\jterm{\Gamma}{\subst{x}{P}}{u}\quad\jterm{\ctxext{{\Gamma}{A}}{P}}{Q}{g}}
+& \inference{\jterm{\Gamma}{A}{x}\quad\jterm{\Gamma}{\subst{x}{P}}{u}\quad\jterm{{{\Gamma}{A}}{P}}{Q}{g}}
   {\jtermeq{\Gamma}{\subst{u}{{x}{Q}}}{\subst{\ctxext{x}{u}}{g}}{\subst{u}{{x}{g}}}}\\
-& \inference{\jterm{\Gamma}{A}{x}\quad\jterm{\Gamma}{\subst{x}{P}}{u}\quad\jfam{\ctxext{{{\Gamma}{A}}{P}}{Q}}{R}}
-  {\jfameq{\ctxext{\Gamma}{\subst{u}{{x}{Q}}}}{\subst{\ctxext{x}{u}}{R}}{\subst{u}{{x}{R}}}}\\
-& \inference{\jterm{\Gamma}{A}{x}\quad\jterm{\Gamma}{\subst{x}{P}}{u}\quad\jterm{\ctxext{{{\Gamma}{A}}{P}}{Q}}{R}{t}}
-  {\jtermeq{\ctxext{\Gamma}{\subst{u}{{x}{Q}}}}{\subst{u}{{x}{R}}}{\subst{\ctxext{x}{u}}{t}}{\subst{u}{{x}{t}}}}
+& \inference{\jterm{\Gamma}{A}{x}\quad\jterm{\Gamma}{\subst{x}{P}}{u}\quad\jfam{{{{\Gamma}{A}}{P}}{Q}}{R}}
+  {\jfameq{{\Gamma}{\subst{u}{{x}{Q}}}}{\subst{\ctxext{x}{u}}{R}}{\subst{u}{{x}{R}}}}\\
+& \inference{\jterm{\Gamma}{A}{x}\quad\jterm{\Gamma}{\subst{x}{P}}{u}\quad\jterm{{{{\Gamma}{A}}{P}}{Q}}{R}{t}}
+  {\jtermeq{{\Gamma}{\subst{u}{{x}{Q}}}}{\subst{u}{{x}{R}}}{\subst{\ctxext{x}{u}}{t}}{\subst{u}{{x}{t}}}}
 \end{align*}
@@
 Let $\jterm{\Gamma}{A}{x}$ and $\jterm{\Gamma}{\subst{x}{P}}{u}$
-be terms and let $\jfam{\ctxext{{\Gamma}{A}}{P}}{Q}$ be a family. Then we have
+be terms and let $\jfam{{{\Gamma}{A}}{P}}{Q}$ be a family. Then we have
 \begin{align*}
@@
 \end{align*}
-We have the term $\jterm{\ctxext{\Gamma}{{A}{P}}}{\ctxwk{P}{A}}{\ctxwk{P}{\idfunc[A]}}$.
+We have the term $\jterm{{\Gamma}{{A}{P}}}{\ctxwk{P}{A}}{\ctxwk{P}{\idfunc[A]}}$.
 Thus we need to find out what $\subst{\ctxwk{P}{\idfunc[A]}}{\ctxwk{P}{{A}{P}}}$ is:
@@
 \begin{defn}
-Let $\jfam{\Gamma}{A}$ and $\jfam{\ctxext{\Gamma}{A}}{P}$ be families. We define
+Let $\jfam{\Gamma}{A}$ and $\jfam{{\Gamma}{A}}{P}$ be families. We define
 \begin{align*}
@@
 \begin{lem}
-For any $\jfam{\Gamma}{A}$ and $\jfam{\ctxext{\Gamma}{A}}{P}$ we have
+For any $\jfam{\Gamma}{A}$ and $\jfam{{\Gamma}{A}}{P}$ we have
 \begin{equation}\label{idfunc-ext-comp}
-\inference{\jfam{\Gamma}{A}\quad\jfam{\ctxext{\Gamma}{A}}{P}}{\jtermeq{\ctxext{\Gamma}{{A}{P}}}{\ctxwk{\ctxext{A}{P}}{\ctxext{A}{P}}}{\idfunc[\ctxext{A}{P}]}{\subst{\cprojsndf{A}{P}}{{\cprojfstf{A}{P}}{\idfunc[\ctxwk{\ctxext{A}{P}}{\ctxext{A}{P}}]}}}}
+\inference{\jfam{\Gamma}{A}\quad\jfam{{\Gamma}{A}}{P}}{\jtermeq{{\Gamma}{{A}{P}}}{\ctxwk{\ctxext{A}{P}}{\ctxext{A}{P}}}{\idfunc[\ctxext{A}{P}]}{\ctxext{\cprojfstf{A}{P}}{\cprojsndf{A}{P}}}}
 \end{equation}
@@
 \begin{proof}
-Consider the families $\jfam{\Gamma}{A}$ and $\jfam{\ctxext{\Gamma}{A}}{P}$. Then
+Consider the families $\jfam{\Gamma}{A}$ and $\jfam{{\Gamma}{A}}{P}$. Then
 we have the judgmental equalities
 \begin{align*}
-\subst{\cprojsndf{A}{P}}{{\cprojfstf{A}{P}}{\idfunc[\ctxwk{\ctxext{A}{P}}{\ctxext{A}{P}}]}}
+\ctxext{\cprojfstf{A}{P}}{\cprojsndf{A}{P}}
+& \jdeq \subst{\cprojsndf{A}{P}}{{\cprojfstf{A}{P}}{\idfunc[\ctxext{\ctxwk{\ctxext{A}{P}}{A}}{\ctxwk{\ctxext{A}{P}}{P}}]}}\\
+& \jdeq \subst{\cprojsndf{A}{P}}{{\cprojfstf{A}{P}}{\idfunc[\ctxwk{\ctxext{A}{P}}{\ctxext{A}{P}}]}}\\
 & \jdeq \subst{\idfunc[P]}{{\ctxwk{P}{\idfunc[A]}}{\idfunc[\ctxwk{\ctxext{A}{P}}{\ctxext{A}{P}}]}}\\
@@
 \begin{equation*}
-\inference{\jfam{\ctxext{\Gamma}{A}}{P}\quad\jfam{\ctxext{\Gamma}{A}}{Q}}
-{\jfameq{\ctxext{\Gamma}{{A}{P}}}{\subst{\cprojfstf{A}{P}}{\ctxwk{\ctxext{A}{P}}{Q}}}{\ctxwk{P}{Q}}}
+\inference{\jfam{{\Gamma}{A}}{P}\quad\jfam{{\Gamma}{A}}{Q}}
+{\jfameq{{\Gamma}{{A}{P}}}{\subst{\cprojfstf{A}{P}}{\ctxwk{\ctxext{A}{P}}{Q}}}{\ctxwk{P}{Q}}}
 \end{equation*}
@@
 \begin{proof}
-Let $\jfam{\ctxext{\Gamma}{A}}{P}$ and $\jfam{\ctxext{\Gamma}{A}}{Q}$ be
+Let $\jfam{{\Gamma}{A}}{P}$ and $\jfam{{\Gamma}{A}}{Q}$ be
 families. Then we have
@@
 \begin{align*}
-& \inference{\jterm{\ctxext{\Gamma}{B}}{Q}{g}\quad\jterm{\ctxext{\Gamma}{B}}{\subst{g}{R}}{t}}{\jtermeq{\ctxext{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{\ctxext{Q}{R}}}{\ctxwk{A}{\ctxext{g}{t}}}{\ctxext{\ctxwk{A}{g}}{\ctxwk{A}{t}}}}
+& \inference{\jterm{{\Gamma}{B}}{Q}{g}\quad\jterm{{\Gamma}{B}}{\subst{g}{R}}{t}}{\jtermeq{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{\ctxext{Q}{R}}}{\ctxwk{A}{\ctxext{g}{t}}}{\ctxext{\ctxwk{A}{g}}{\ctxwk{A}{t}}}}
 % \\
-%& \inference{\jterm{\Gamma}{\ctxwk{A}{\ctxext{B}{Q}}}{\ctxwk{A}{w}}}{\jtermeq{\ctxext{\Gamma}{A}}{\ctxwk{A}{B}}{\cprojfst{\ctxwk{A}{B}}{\ctxwk{A}{Q}}{\ctxwk{A}{w}}}{\ctxwk{A}{(\cprojfst{B}{Q}{w})}}}\\
+%& \inference{\jterm{\Gamma}{\ctxwk{A}{\ctxext{B}{Q}}}{\ctxwk{A}{w}}}{\jtermeq{{\Gamma}{A}}{\ctxwk{A}{B}}{\cprojfst{\ctxwk{A}{B}}{\ctxwk{A}{Q}}{\ctxwk{A}{w}}}{\ctxwk{A}{(\cprojfst{B}{Q}{w})}}}\\
 \end{align*}
@@
 \begin{proof}
-Consider $\jterm{\ctxext{\Gamma}{B}}{Q}{g}$ and $\jterm{\ctxext{\Gamma}{B}}{\subst{g}{R}}{t}$.
+Consider $\jterm{{\Gamma}{B}}{Q}{g}$ and $\jterm{{\Gamma}{B}}{\subst{g}{R}}{t}$.
 Then we have the judgmental equalities
@@
 \begin{align*}
-& \inference{\jterm{\ctxext{{\Gamma}{A}}{P}}{Q}{g}\quad\jterm{\ctxext{{\Gamma}{A}}{P}}{\subst{g}{R}}{t}}{\jtermeq{\ctxext{\Gamma}{\subst{x}{P}}}{\ctxext{\subst{x}{Q}}{\subst{x}{R}}}{\subst{x}{\ctxext{g}{t}}}{\ctxext{\subst{x}{g}}{\subst{x}{t}}}}
+& \inference{\jterm{{{\Gamma}{A}}{P}}{Q}{g}\quad\jterm{{{\Gamma}{A}}{P}}{\subst{g}{R}}{t}}{\jtermeq{{\Gamma}{\subst{x}{P}}}{\ctxext{\subst{x}{Q}}{\subst{x}{R}}}{\subst{x}{\ctxext{g}{t}}}{\ctxext{\subst{x}{g}}{\subst{x}{t}}}}
 \end{align*}

### 2014-03-02 article-models-tt.tex
cleaning up the commets in pretty type theory

@@
 \subsection{The basic judgments}
-The type theory we describe here is a theory of contexts, families of
-contexts and terms thereof. The families of contexts are by some authors called
-dependent contexts, but they are handled a bit differently here because they
-become the primary object of study. Dependent contexts can be types; they could
-be seen as atomic or indecomposable dependent contexts.
-
-Thus we make eight kinds of judgments in our type theory: ``$\Gamma$ is a context'',
-``$A(i)$ is a family of contexts over $\Gamma$'', ``$A$ is a type in context $\Gamma$''
-and ``$x$ is a term of the family $A$ of contexts over $\Gamma$''. The other four
+The basic judgments of pretty type theory are the same as for structural type
+theory. There are judgments for: ``$\Gamma$ is a context'',
+``$A(i)$ over $i:\Gamma$ is a family over $\Gamma$'', ``$A(i)$ over $i:\Gamma$ 
+is a type in context $\Gamma$''
+and ``$x(i)$ is a term of $A(i)$ above $i:\Gamma$''. The other four
 judgments are for judgmental equality.
@@
 
-If $A$ is a type
-in context $\Gamma$, then $A$ is also a family of contexts over $\Gamma$. Being
-a term of type $A$ means the same thing as being a term of the family $A$ of contexts.
-Two types in context $\Gamma$ are judgmentally equal precisely when they are equal
-as context families. Moreover, if a family $B$ of contexts over $\Gamma$ is
-judgmentally equal to a type $A$ in context $\Gamma$, then $B$ is a type in
-context $\Gamma$. This is expressed by the following four inference rules:
+We have the following basic inference rules that relate types and families:
 
@@
 \subsection{The basic rules for judgmental equality}
-The rules for judgmental equality establish that it is an equivalence relation
-in all three cases (contexts, types and terms).
+The rules for judgmental equality establish that it is an equivalence relation.
 \bgroup\small
@@
 \subsection{The empty context}
-There is an empty context and over any context there is an empty family of
-contexts. We do not assume that the empty context is a type, that would be like
-assuming that the multiplicative unit of a ring is prime. The empty family
-always has a unique term.
+The empty context looks a bit strange when we explicitly denote the terms. But
+we will not do so anymore after this subsection.
 

### 2014-03-02 article-models-tt.tex
renamed section 3

@@
-\section{Type theory before type constructors}\label{tt}
+\section{Structural type theory}\label{tt}
 In this section we give a description of type theory before basic
@@
 
+Pretty type theory should be such that if you write down two contexts, families
+or terms in exactly the same way, then they are the same, and there should be
+notational shortcuts for extension, weakening and substitution which make
+this interesting.
+
 \subsection{The basic judgments}

### 2014-03-02 article-models-tt.tex
introducing lemmas about composition; big code cleanup

@@
 \begin{align*}
-& \inference{\jtype{\Gamma}{A}}{\jfam{\Gamma}{A}} & & \inference{\jtypeeq{\Gamma}{A}{B}}{\jfameq{\Gamma}{A}{B}}\\
-& \inference{\jtype{\Gamma}{A}\quad\jfameq{\Gamma}{A}{B}}{\jtype{\Gamma}{B}}
-& & \inference{\jtype{\Gamma}{A}\quad\jfameq{\Gamma}{A}{B}}{\jtypeeq{\Gamma}{A}{B}}
+& \inference
+  {\jtype{\Gamma}{A}}
+  {\jfam{\Gamma}{A}}
+& & \inference
+    {\jtypeeq{\Gamma}{A}{B}}
+    {\jfameq{\Gamma}{A}{B}}\\
+& \inference
+  {\jtype{\Gamma}{A}
+   \jfameq{\Gamma}{A}{B}}
+  {\jtype{\Gamma}{B}}
+& & \inference
+    {\jtype{\Gamma}{A}
+     \jfameq{\Gamma}{A}{B}}
+    {\jtypeeq{\Gamma}{A}{B}}
 \end{align*}
@@
 \begin{align*}
-& \inference{\jctx{\Gamma}}{\jctxeq{\Gamma}{\Gamma}} 
-& & \inference{\jctxeq{\Gamma}{\Delta}}{\jctxeq{\Delta}{\Gamma}} 
-& & \inference{\jctxeq{\Gamma}{\Delta}\qquad\jctxeq{\Delta}{\greek{E}}}{\jctxeq{\Gamma}{\greek{E}}}\\
-& \inference{\jfam{\Gamma}{A}}{\jfameq{\Gamma}{A}{A}} 
-& & \inference{\jfameq{\Gamma}{A}{B}}{\jfameq{\Gamma}{B}{A}}
-& & \inference{\jfameq{\Gamma}{A}{B}\qquad\jfameq{\Gamma}{B}{C}}{\jfameq{\Gamma}{A}{C}}\\
-& \inference{\jterm{\Gamma}{A}{x}}{\jtermeq{\Gamma}{A}{x}{x}}
-& & \inference{\jtermeq{\Gamma}{A}{x}{y}}{\jtermeq{\Gamma}{A}{y}{x}}
-& & \inference{\jtermeq{\Gamma}{A}{x}{y}\qquad\jtermeq{\Gamma}{A}{y}{z}}{\jtermeq{\Gamma}{A}{x}{z}}
+& \inference
+  {\jctx{\Gamma}}
+  {\jctxeq{\Gamma}{\Gamma}} 
+& & \inference
+    {\jctxeq{\Gamma}{\Delta}}
+    {\jctxeq{\Delta}{\Gamma}} 
+& & \inference
+    {\jctxeq{\Gamma}{\Delta}
+     \jctxeq{\Delta}{\greek{E}}}
+    {\jctxeq{\Gamma}{\greek{E}}}\\
+& \inference
+  {\jfam{\Gamma}{A}}
+  {\jfameq{\Gamma}{A}{A}} 
+& & \inference
+    {\jfameq{\Gamma}{A}{B}}
+    {\jfameq{\Gamma}{B}{A}}
+& & \inference
+    {\jfameq{\Gamma}{A}{B}
+     \jfameq{\Gamma}{B}{C}}
+    {\jfameq{\Gamma}{A}{C}}\\
+& \inference
+  {\jterm{\Gamma}{A}{x}}
+  {\jtermeq{\Gamma}{A}{x}{x}}
+& & \inference
+    {\jtermeq{\Gamma}{A}{x}{y}}
+    {\jtermeq{\Gamma}{A}{y}{x}}
+& & \inference
+    {\jtermeq{\Gamma}{A}{x}{y}
+     \jtermeq{\Gamma}{A}{y}{z}}
+    {\jtermeq{\Gamma}{A}{x}{z}}
 \end{align*}
@@
 \begin{align*}
-& \inference{\jctxeq{\Gamma}{\Delta}\qquad\jfam{\Gamma}{A}}{\jfam{\Delta}{A}}
-& & \inference{\jctxeq{\Gamma}{\Delta}\qquad\jfameq{\Gamma}{A}{B}}{\jfameq{\Delta}{A}{B}}\\
-& \inference{\jctxeq{\Gamma}{\Delta}\qquad\jterm{\Gamma}{A}{x}}{\jterm{\Delta}{A}{x}}
-& & \inference{\jctxeq{\Gamma}{\Delta}\qquad\jtermeq{\Gamma}{A}{x}{y}}{\jtermeq{\Delta}{A}{x}{y}}\\
-& \inference{\jfameq{\Gamma}{A}{B}\qquad \jterm{\Gamma}{A}{x}}{\jterm{\Gamma}{B}{x}}
-& & \inference{\jfameq{\Gamma}{A}{B}\qquad\jtermeq{\Gamma}{A}{x}{y}}{\jtermeq{\Gamma}{B}{x}{y}}
+& \inference
+  {\jctxeq{\Gamma}{\Delta}
+   \jfam{\Gamma}{A}}
+  {\jfam{\Delta}{A}}
+& & \inference
+    {\jctxeq{\Gamma}{\Delta}
+     \jfameq{\Gamma}{A}{B}}
+    {\jfameq{\Delta}{A}{B}}\\
+& \inference
+  {\jctxeq{\Gamma}{\Delta}
+   \jterm{\Gamma}{A}{x}}
+  {\jterm{\Delta}{A}{x}}
+& & \inference
+    {\jctxeq{\Gamma}{\Delta}
+     \jtermeq{\Gamma}{A}{x}{y}}
+    {\jtermeq{\Delta}{A}{x}{y}}\\
+& \inference
+  {\jfameq{\Gamma}{A}{B}
+   \jterm{\Gamma}{A}{x}}
+  {\jterm{\Gamma}{B}{x}}
+& & \inference
+    {\jfameq{\Gamma}{A}{B}
+     \jtermeq{\Gamma}{A}{x}{y}}
+    {\jtermeq{\Gamma}{B}{x}{y}}
 \end{align*}
@@
 \begin{align}
-& \inference{}{\jctx{\emptyc}}\\
-& \inference{\jctx{\Gamma}}{\jfam{\Gamma}{\emptyf[\Gamma]}}\\
-& \inference{\jctx{\Gamma}}{\jterm{\Gamma}{\emptyf[\Gamma]}{\emptytm[\Gamma]}}\\
-& \inference{\jterm{\Gamma}{\emptyf[\Gamma]}{x}}{\jtermeq{\Gamma}{\emptyf[\Gamma]}{x}{\emptytm[\Gamma]}}
+& \inference
+  {}
+  {\jctx{\emptyc}}\\
+& \inference
+  {\jctx{\Gamma}}
+  {\jfam{\Gamma}{\emptyf[\Gamma]}}\\
+& \inference
+  {\jctx{\Gamma}}
+  {\jterm{\Gamma}{\emptyf[\Gamma]}{\emptytm[\Gamma]}}\\
+& \inference
+  {\jterm{\Gamma}{\emptyf[\Gamma]}{x}}
+  {\jtermeq{\Gamma}{\emptyf[\Gamma]}{x}{\emptytm[\Gamma]}}
 \end{align}
@@
 \begin{align}
-& \inference{\jctx{\Gamma}}{\jfam{\emptyc}{\Gamma}} 
-& & \inference{\jfam{\emptyc}{\Gamma}}{\jctx{\Gamma}}\\
-& \inference{\jctxeq{\Gamma}{\Delta}}{\jfameq{\emptyc}{\Gamma}{\Delta}}
-& & \inference{\jfameq{\emptyc}{\Gamma}{\Delta}}{\jctxeq{\Gamma}{\Delta}}
+& \inference
+  {\jctx{\Gamma}}
+  {\jfam{\emptyc}{\Gamma}} 
+& & \inference
+    {\jfam{\emptyc}{\Gamma}}
+    {\jctx{\Gamma}}\\
+& \inference
+  {\jctxeq{\Gamma}{\Delta}}
+  {\jfameq{\emptyc}{\Gamma}{\Delta}}
+& & \inference
+    {\jfameq{\emptyc}{\Gamma}{\Delta}}
+    {\jctxeq{\Gamma}{\Delta}}
 \end{align}
@@
 \begin{equation}
-\inference{}{\jfameq{\emptyc}{\emptyc}{\emptyf[\emptyc]}}
+\inference
+{}
+{\jfameq{\emptyc}{\emptyc}{\emptyf[\emptyc]}}
 \end{equation}
@@
 \begin{align}
-& \inference{\jfam{\Gamma}{A}}{\jctx{\ctxext{\Gamma}{A}}}
-& & \inference{\jctxeq{\Gamma}{\Delta}\qquad\jfameq{\Gamma}{A}{B}}{\jctxeq{\ctxext{\Gamma}{A}}{\ctxext{\Delta}{B}}}\\
-& \inference{\jfam{{\Gamma}{A}}{P}}{\jfam{\Gamma}{\ctxext{A}{P}}}
-& & \inference{\jfameq{\Gamma}{A}{B}\qquad\jfameq{{\Gamma}{A}}{P}{Q}}{\jfameq{\Gamma}{\ctxext{A}{P}}{\ctxext{B}{Q}}}
+& \inference
+  {\jfam{\Gamma}{A}}
+  {\jctx{\ctxext{\Gamma}{A}}}
+& & \inference
+    {\jctxeq{\Gamma}{\Delta}
+     \jfameq{\Gamma}{A}{B}}
+    {\jctxeq{\ctxext{\Gamma}{A}}{\ctxext{\Delta}{B}}}\\
+& \inference
+  {\jfam{{\Gamma}{A}}{P}}
+  {\jfam{\Gamma}{\ctxext{A}{P}}}
+& & \inference
+    {\jfameq{\Gamma}{A}{B} 
+     \jfameq{{\Gamma}{A}}{P}{Q}}
+    {\jfameq{\Gamma}{\ctxext{A}{P}}{\ctxext{B}{Q}}}
 \end{align}
@@
 \begin{align}
-& \inference{\jctx{\Gamma}}{\jctxeq{\ctxext{\emptyc}{\Gamma}}{\Gamma}}\\
-& \inference{\jctx{\Gamma}}{\jctxeq{\ctxext{\Gamma}{\emptyf}}{\Gamma}}\\
-& \inference{\jfam{\Gamma}{A}}{\jfameq{\Gamma}{\ctxext{\emptyf}{A}}{A}}
+& \inference
+  {\jctx{\Gamma}}
+  {\jctxeq{\ctxext{\emptyc}{\Gamma}}{\Gamma}}\\
+& \inference
+  {\jctx{\Gamma}}
+  {\jctxeq{\ctxext{\Gamma}{\emptyf}}{\Gamma}}\\
+& \inference
+  {\jfam{\Gamma}{A}}
+  {\jfameq{\Gamma}{\ctxext{\emptyf}{A}}{A}}
 \end{align}
@@
 \begin{align}
-& \inference{\jfam{\Gamma}{A}\qquad\jfam{{\Gamma}{A}}{P}}
+& \inference
+  {\jfam{\Gamma}{A}
+   \jfam{{\Gamma}{A}}{P}}
   {\jctxeq{\ctxext{{\Gamma}{A}}{P}}{\ctxext{\Gamma}{{A}{P}}}}\\
-& \inference{\jfam{{\Gamma}{A}}{P}\qquad\jfam{{{\Gamma}{A}}{P}}{Q}}
+& \inference
+  {\jfam{{\Gamma}{A}}{P}
+   \jfam{{{\Gamma}{A}}{P}}{Q}}
   {\jfameq{\Gamma}{\ctxext{{A}{P}}{Q}}{\ctxext{A}{{P}{Q}}}}
@@
 \begin{align}
-& \inference{\jfam{\Gamma}{B}}{\jfam{{\Gamma}{A}}{\ctxwk{A}{B}}}
-& & \inference{\jfameq{\Gamma}{B}{B'}}{\jfameq{{\Gamma}{A}}{\ctxwk{A}{B}}{\ctxwk{A'}{B'}}}\\
-& \inference{\jfam{{\Gamma}{B}}{Q}}
+& \inference
+  {\jfam{\Gamma}{B}}
+  {\jfam{{\Gamma}{A}}{\ctxwk{A}{B}}}
+& & \inference
+    {\jfameq{\Gamma}{B}{B'}}
+    {\jfameq{{\Gamma}{A}}{\ctxwk{A}{B}}{\ctxwk{A'}{B'}}}\\
+& \inference
+  {\jfam{{\Gamma}{B}}{Q}}
   {\jfam{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{Q}}}
-& & \inference{\jfameq{{\Gamma}{B}}{Q}{Q'}}
+& & \inference
+    {\jfameq{{\Gamma}{B}}{Q}{Q'}}
     {\jfameq{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{Q}}{\ctxwk{A'}{Q'}}}\\
-& \inference{\jterm{\Gamma}{B}{y}}{\jterm{{\Gamma}{A}}{\ctxwk{A}{B}}{\ctxwk{A}{y}}}
-& & \inference{\jtermeq{\Gamma}{B}{y}{y'}}{\jtermeq{{\Gamma}{A}}{\ctxwk{A}{B}}{\ctxwk{A}{y}}{\ctxwk{A'}{y'}}}\\
-& \inference{\jterm{{\Gamma}{B}}{Q}{g}}{\jterm{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{Q}}{\ctxwk{A}{g}}}
-& & \inference{\jtermeq{{\Gamma}{B}}{Q}{g}{g'}}
+& \inference
+  {\jterm{\Gamma}{B}{y}}
+  {\jterm{{\Gamma}{A}}{\ctxwk{A}{B}}{\ctxwk{A}{y}}}
+& & \inference
+    {\jtermeq{\Gamma}{B}{y}{y'}}
+    {\jtermeq{{\Gamma}{A}}{\ctxwk{A}{B}}{\ctxwk{A}{y}}{\ctxwk{A'}{y'}}}\\
+& \inference
+  {\jterm{{\Gamma}{B}}{Q}{g}}
+  {\jterm{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{Q}}{\ctxwk{A}{g}}}
+& & \inference
+    {\jtermeq{{\Gamma}{B}}{Q}{g}{g'}}
     {\jtermeq{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{Q}}{\ctxwk{A}{g}}{\ctxwk{A'}{g'}}}
@@
 \begin{align}
-& \inference{\jfam{\Gamma}{B}\quad\jtype{\ctxext{\Gamma}{B}}{Q}}{\jtype{\ctxext{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{Q}}}
+& \inference
+  {\jfam{\Gamma}{B}
+   \jtype{\ctxext{\Gamma}{B}}{Q}}
+  {\jtype{\ctxext{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{Q}}}
 \end{align}
@@
 \begin{align}
-& \inference{\jctx{\Gamma}}{\jfameq{\Gamma}{\ctxwk{\Gamma}{\emptyc}}{\emptyf}}\\
-& \inference{\jfam{\Gamma}{A}}{\jfameq{{\Gamma}{A}}{\ctxwk{A}{\emptyf}}{\emptyf}}
+& \inference
+  {\jctx{\Gamma}}
+  {\jfameq{\Gamma}{\ctxwk{\Gamma}{\emptyc}}{\emptyf}}\\
+& \inference
+  {\jfam{\Gamma}{A}}
+  {\jfameq{{\Gamma}{A}}{\ctxwk{A}{\emptyf}}{\emptyf}}
 \end{align}
@@
 \begin{align}
-& \inference{\jfam{\Gamma}{B}}{\jfameq{\Gamma}{\ctxwk{\emptyf}{B}}{B}}\\
-& \inference{\jterm{\Gamma}{B}{y}}{\jtermeq{\Gamma}{B}{\ctxwk{\emptyf}{y}}{y}}\\
-& \inference{\jfam{{\Gamma}{B}}{Q}}{\jfameq{{\Gamma}{B}}{\ctxwk{\emptyf}{Q}}{Q}}\\
-& \inference{\jterm{{\Gamma}{B}}{Q}{g}}{\jtermeq{{\Gamma}{B}}{Q}{\ctxwk{\emptyf}{g}}{g}}
+& \inference
+  {\jfam{\Gamma}{B}}
+  {\jfameq{\Gamma}{\ctxwk{\emptyf}{B}}{B}}\\
+& \inference
+  {\jterm{\Gamma}{B}{y}}
+  {\jtermeq{\Gamma}{B}{\ctxwk{\emptyf}{y}}{y}}\\
+& \inference
+  {\jfam{{\Gamma}{B}}{Q}}
+  {\jfameq{{\Gamma}{B}}{\ctxwk{\emptyf}{Q}}{Q}}\\
+& \inference
+  {\jterm{{\Gamma}{B}}{Q}{g}}
+  {\jtermeq{{\Gamma}{B}}{Q}{\ctxwk{\emptyf}{g}}{g}}
 \end{align}
@@
 \begin{align}
-& \inference{\jfam{\Gamma}{A}\qquad\jfam{\Gamma}{B}\qquad\jfam{{\Gamma}{C}}{R}}
+& \inference
+  {\jfam{\Gamma}{A}
+   \jfam{\Gamma}{B}
+   \jfam{{\Gamma}{C}}{R}}
   {\jfameq{{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{{A}{B}}{{A}{C}}}}
@@
     {\ctxwk{{A}{B}}{{A}{R}}}}\label{comp-ww-f}\\
-& \inference{\jfam{\Gamma}{A}\qquad\jfam{\Gamma}{B}\qquad\jterm{{\Gamma}{C}}{R}{t}}
+& \inference
+  {\jfam{\Gamma}{A}
+   \jfam{\Gamma}{B}
+   \jterm{{\Gamma}{C}}{R}{t}}
   {\jtermeq{{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{{A}{B}}{{A}{C}}}}
@@
 \begin{align}
-& \inference{\jfam{{\Gamma}{B}}{Q}}
-    {\jfameq{{{{\Gamma}{A}}{P}}{\ctxwk{P}{{A}{B}}}}{\ctxwk{\ctxext{A}{P}}{Q}}{\ctxwk{P}{{A}{Q}}}}\label{comp-ew-f}\\
-& \inference{\jterm{{\Gamma}{B}}{Q}{g}}
-    {\jtermeq{{{{\Gamma}{A}}{P}}{\ctxwk{P}{{A}{B}}}}{\ctxwk{P}{{A}{Q}}}{\ctxwk{\ctxext{A}{P}}{g}}{\ctxwk{P}{{A}{g}}}}\label{comp-ew-t}
+& \inference
+  {\jfam{{\Gamma}{B}}{Q}}
+  {\jfameq{{{{\Gamma}{A}}{P}}{\ctxwk{P}{{A}{B}}}}{\ctxwk{\ctxext{A}{P}}{Q}}{\ctxwk{P}{{A}{Q}}}}
+  \label{comp-ew-f}\\
+& \inference
+  {\jterm{{\Gamma}{B}}{Q}{g}}
+  {\jtermeq{{{{\Gamma}{A}}{P}}{\ctxwk{P}{{A}{B}}}}{\ctxwk{P}{{A}{Q}}}{\ctxwk{\ctxext{A}{P}}{g}}{\ctxwk{P}{{A}{g}}}} 
+  \label{comp-ew-t}
 \end{align}
@@
 \begin{align}
-& \inference{\jfam{{\Gamma}{A}}{P}}{\jfam{\Gamma}{\subst{x}{P}}}
-& & \inference{\jfameq{{\Gamma}{A}}{P}{P'}}{\jfameq{\Gamma}{\subst{x}{P}}{\subst{x'}{P'}}}\\
-& \inference{\jfam{{{\Gamma}{A}}{P}}{Q}}{\jfam{{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}}
-& & \inference{\jfameq{{{\Gamma}{A}}{P}}{Q}{Q'}}{\jfameq{{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}{\subst{x'}{Q'}}}\\
-& \inference{\jterm{{\Gamma}{A}}{P}{f}}{\jterm{\Gamma}{\subst{x}{P}}{\subst{x}{f}}}
-& & \inference{\jtermeq{{\Gamma}{A}}{P}{f}{f'}}{\jtermeq{\Gamma}{\subst{x}{P}}{\subst{x}{f}}{\subst{x'}{f'}}}\\
-& \inference{\jterm{{{\Gamma}{A}}{P}}{Q}{g}}{\jterm{{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}{\subst{x}{g}}}
-& &\inference{\jtermeq{{{\Gamma}{A}}{P}}{Q}{g}{g'}}{\jtermeq{{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}{\subst{x}{g}}{\subst{x'}{g'}}}
+& \inference
+  {\jfam{{\Gamma}{A}}{P}}
+  {\jfam{\Gamma}{\subst{x}{P}}}
+& & \inference
+    {\jfameq{{\Gamma}{A}}{P}{P'}}
+    {\jfameq{\Gamma}{\subst{x}{P}}{\subst{x'}{P'}}}\\
+& \inference
+  {\jfam{{{\Gamma}{A}}{P}}{Q}}
+  {\jfam{{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}}
+& & \inference
+    {\jfameq{{{\Gamma}{A}}{P}}{Q}{Q'}}
+    {\jfameq{{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}{\subst{x'}{Q'}}}\\
+& \inference
+  {\jterm{{\Gamma}{A}}{P}{f}}
+  {\jterm{\Gamma}{\subst{x}{P}}{\subst{x}{f}}}
+& & \inference
+    {\jtermeq{{\Gamma}{A}}{P}{f}{f'}}
+    {\jtermeq{\Gamma}{\subst{x}{P}}{\subst{x}{f}}{\subst{x'}{f'}}}\\
+& \inference
+  {\jterm{{{\Gamma}{A}}{P}}{Q}{g}}
+  {\jterm{{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}{\subst{x}{g}}}
+& & \inference
+    {\jtermeq{{{\Gamma}{A}}{P}}{Q}{g}{g'}}
+    {\jtermeq{{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}{\subst{x}{g}}{\subst{x'}{g'}}}
 \end{align}
@@
 \begin{align}
-& \inference{\jterm{\Gamma}{A}{x}\quad\jtype{\ctxext{{\Gamma}{A}}{P}}{Q}}{\jtype{\ctxext{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}}
+& \inference
+  {\jterm{\Gamma}{A}{x}
+   \jtype{\ctxext{{\Gamma}{A}}{P}}{Q}}
+  {\jtype{\ctxext{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}}
 \end{align}
@@
 \begin{align}
-& \inference{\jterm{\Gamma}{A}{x}}{\jfameq{\Gamma}{\subst{x}{\emptyf}}{\emptyf}}\\
-& \inference{\jterm{\Gamma}{A}{x}\quad\jfam{{\Gamma}{A}}{P}}{\jfameq{{\Gamma}{\subst{x}{P}}}{\subst{x}{\emptyf}}{\emptyf}}
+& \inference
+  {\jterm{\Gamma}{A}{x}}
+  {\jfameq{\Gamma}{\subst{x}{\emptyf}}{\emptyf}}\\
+& \inference
+  {\jterm{\Gamma}{A}{x}
+   \jfam{{\Gamma}{A}}{P}}
+  {\jfameq{{\Gamma}{\subst{x}{P}}}{\subst{x}{\emptyf}}{\emptyf}}
 \end{align}
@@
 \begin{align}
-& \inference{\jfam{\Gamma}{A}}{\jfameq{\Gamma}{\subst{\emptytm}{A}}{A}}\\
-& \inference{\jterm{\Gamma}{A}{x}}{\jtermeq{\Gamma}{A}{\subst{\emptytm}{x}}{x}}\\
-& \inference{\jfam{{\Gamma}{A}}{P}}{\jfameq{{\Gamma}{A}}{\subst{\emptytm}{P}}{P}}\\
-& \inference{\jterm{{\Gamma}{A}}{P}{f}}{\jtermeq{{\Gamma}{A}}{P}{\subst{\emptytm}{f}}{f}}.
+& \inference
+  {\jfam{\Gamma}{A}}
+  {\jfameq{\Gamma}{\subst{\emptytm}{A}}{A}}\\
+& \inference
+  {\jterm{\Gamma}{A}{x}}
+  {\jtermeq{\Gamma}{A}{\subst{\emptytm}{x}}{x}}\\
+& \inference
+  {\jfam{{\Gamma}{A}}{P}}
+  {\jfameq{{\Gamma}{A}}{\subst{\emptytm}{P}}{P}}\\
+& \inference
+  {\jterm{{\Gamma}{A}}{P}{f}}
+  {\jtermeq{{\Gamma}{A}}{P}{\subst{\emptytm}{f}}{f}}.
 \end{align}
@@
 \begin{align}
-& \inference{\jfam{{{{\Gamma}{A}}{P}}{Q}}{R}}
+& \inference
+  {\jfam{{{{\Gamma}{A}}{P}}{Q}}{R}}
   {\jfameq{{\Gamma}{\subst{x}{P}}}{\subst{x}{\ctxext{Q}{R}}}{\ctxext{\subst{x}{Q}}{\subst{x}{R}}}}
@@
   {\jfam{{{\Gamma}{A}}{Q}}{R}}
-  {\jfameq{{{\Gamma}{\subst{x}{P}}}{\subst{x}{\ctxwk{P}{Q}}}}{\subst{x}{\ctxwk{P}{R}}}{\ctxwk{\subst{x}{P}}{\subst{x}{R}}}}\label{comp-sw-f}\\
+  {\jfameq{{{\Gamma}{\subst{x}{P}}}{\subst{x}{\ctxwk{P}{Q}}}}{\subst{x}{\ctxwk{P}{R}}}{\ctxwk{\subst{x}{P}}{\subst{x}{R}}}}
+  \label{comp-sw-f}\\
 & \inference
@@
       {\subst{x}{\ctxwk{P}{h}}}
-      {\ctxwk{\subst{x}{P}}{\subst{x}{h}}}}\label{comp-sw-t}
+      {\ctxwk{\subst{x}{P}}{\subst{x}{h}}}}
+  \label{comp-sw-t}
 \end{align}
@@
 \begin{align}
-&\inference{\jfam{{{{\Gamma}{A}}{P}}{Q}}{R}}
-{\jfameq{{\Gamma}{\subst{x}{{f}{Q}}}}{\subst{x}{{f}{R}}}{\subst{{x}{f}}{{x}{R}}}}\label{comp-ss-f}\\
-&\inference{\jterm{{{{\Gamma}{A}}{P}}{Q}}{R}{h}}
-{\jtermeq{{\Gamma}{\subst{x}{{f}{Q}}}}{\subst{x}{{f}{R}}}{\subst{x}{{f}{h}}}{\subst{{x}{f}}{{x}{h}}}}\label{comp-ss-t}
+& \inference
+  {\jfam{{{{\Gamma}{A}}{P}}{Q}}{R}}
+  {\jfameq{{\Gamma}{\subst{x}{{f}{Q}}}}{\subst{x}{{f}{R}}}{\subst{{x}{f}}{{x}{R}}}}
+  \label{comp-ss-f}\\
+& \inference
+  {\jterm{{{{\Gamma}{A}}{P}}{Q}}{R}{h}}
+  {\jtermeq{{\Gamma}{\subst{x}{{f}{Q}}}}{\subst{x}{{f}{R}}}{\subst{x}{{f}{h}}}{\subst{{x}{f}}{{x}{h}}}}
+  \label{comp-ss-t}
 \end{align}
@@
 \begin{align}
-& \inference{\jfam{{{\Gamma}{B}}{Q}}{R}}{\jfameq{{{\Gamma}{A}}{\ctxwk{A}{\subst{y}{Q}}}}{\ctxwk{A}{\subst{y}{R}}}{\subst{\ctxwk{A}{y}}{\ctxwk{A}{R}}}}\label{comp-ws-f}\\
-& \inference{\jterm{{{\Gamma}{B}}{Q}}{R}{h}}{\jtermeq{{{\Gamma}{A}}{\ctxwk{A}{\subst{y}{Q}}}}{\ctxwk{A}{\subst{y}{R}}}{\ctxwk{A}{\subst{y}{h}}}{\subst{\ctxwk{A}{y}}{\ctxwk{A}{h}}}}\label{comp-ws-t}
+& \inference
+  {\jfam{{{\Gamma}{B}}{Q}}{R}}
+  {\jfameq{{{\Gamma}{A}}{\ctxwk{A}{\subst{y}{Q}}}}{\ctxwk{A}{\subst{y}{R}}}{\subst{\ctxwk{A}{y}}{\ctxwk{A}{R}}}}
+  \label{comp-ws-f}\\
+& \inference
+  {\jterm{{{\Gamma}{B}}{Q}}{R}{h}}
+  {\jtermeq{{{\Gamma}{A}}{\ctxwk{A}{\subst{y}{Q}}}}{\ctxwk{A}{\subst{y}{R}}}{\ctxwk{A}{\subst{y}{h}}}{\subst{\ctxwk{A}{y}}{\ctxwk{A}{h}}}}
+  \label{comp-ws-t}
 \end{align}
@@
 \begin{align}
-& \inference{\jfam{\Gamma}{A}\qquad\jfam{{\Gamma}{B}}{Q}\qquad\jterm{\Gamma}{A}{x}}{\jfameq{{\Gamma}{B}}{\subst{x}{\ctxwk{A}{Q}}}{Q}}\label{defn-ws-3}\\
-& \inference{\jterm{\Gamma}{A}{x}\qquad\jterm{{\Gamma}{B}}{Q}{g}}{\jtermeq{{\Gamma}{B}}{Q}{\subst{x}{\ctxwk{A}{g}}}{g}}\label{defn-ws-4}
+& \inference
+  {\jfam{\Gamma}{A}
+   \jfam{{\Gamma}{B}}{Q}
+   \jterm{\Gamma}{A}{x}}
+  {\jfameq{{\Gamma}{B}}{\subst{x}{\ctxwk{A}{Q}}}{Q}}
+  \label{defn-ws-3}\\
+& \inference
+  {\jterm{\Gamma}{A}{x}
+   \jterm{{\Gamma}{B}}{Q}{g}}
+  {\jtermeq{{\Gamma}{B}}{Q}{\subst{x}{\ctxwk{A}{g}}}{g}}
+  \label{defn-ws-4}
 \end{align}
@@
 for any three families $A$, $B$ and $C$ in context $\Gamma$ and any $\jterm{{\Gamma}{A}}{\ctxwk{A}{B}}{f}$.
-It follows that for $\jterm{{\Gamma}{B}}{\ctxwk{B}{C}}{g}$ we have
+It follows that for $\jterm{{\Gamma}{B}}{\ctxwk{B}{C}}{g}$ we can compose $f$
+with $g$ to obtain a term of $\ctxwk{A}{C}$ in context $\ctxext{\Gamma}{A}$.
+In the following definition, we work with in a slightly greater generality.
+
+\begin{defn}
+For $\jhom{\Gamma}{A}{B}{f}$ and consider a term $\jterm{{\Gamma}{B}}{Q}{g}$.
+We define
+\begin{align*}
+\jfamdefn*{{\Gamma}{A}}{\jhomcomp{f}{Q}}{\subst{f}{\ctxwk{A}{Q}}}\\
+\jtermdefn*{{\Gamma}{A}}{\jhomcomp{f}{Q}}{\jhomcomp{f}{g}}{\subst{f}{\ctxwk{A}{g}}}.
+\end{align*}
+Likewise, when we have a family $\jfam{{{\Gamma}{B}}{Q}}{R}$ and a term
+$\jterm{{{\Gamma}{B}}{Q}}{R}{h}$, we define
+\begin{align*}
+\jfamdefn*{{{\Gamma}{A}}{\jhomcomp{f}{Q}}}{\jhomcomp{f}{R}}{\subst{f}{\ctxwk{A}{R}}}\\
+\jtermdefn*{{{\Gamma}{A}}{\jhomcomp{f}{Q}}}{\jhomcomp{f}{R}}{\jhomcomp{f}{h}}{\subst{f}{\ctxwk{A}{h}}}.
+\end{align*}
+\end{defn}
+
+We have lots of compatibility properties for composition:
+
+\begin{lem}
+We have the following inference rules about the situation where something is
+substituted by a composition:
+\begin{align*}
+& \inference
+  {\jhom{\Gamma}{A}{B}{f}
+   \jhom{\Gamma}{B}{C}{g}
+   \jfam{{{\Gamma}{A}}{\ctxwk{A}{C}}}{R}}
+  {\jfameq{{\Gamma}{A}}{\subst{\jhomcomp{f}{g}}{R}}{\subst{f}{{\ctxwk{A}{g}}{\ctxwk{{A}{B}}{R}}}}}\\
+& \inference
+  {\jhom{\Gamma}{A}{B}{f}
+   \jhom{\Gamma}{B}{C}{g}
+   \jterm{{{\Gamma}{A}}{\ctxwk{A}{C}}}{R}{h}}
+  {\jfameq{{\Gamma}{A}}{\subst{\jhomcomp{f}{g}}{h}}{\subst{f}{{\ctxwk{A}{g}}{\ctxwk{{A}{B}}{h}}}}}
+\end{align*}
+In particular, when $\jhom{\Gamma}{C}{D}{h}$, we have
 \begin{equation*}
-\jterm{{\Gamma}{A}}{\ctxwk{A}{C}}{\subst{f}{\ctxwk{A}{g}}}
+\jhomeq{\Gamma}{A}{D}{\jhomcomp{\jhomcomp{f}{g}}{h}}{\jhomcomp{f}{\jhomcomp{g}{h}}}.
 \end{equation*}
-The term $\jcomp{A}{f}{g}$ is the composition of $g$ with $f$. We may call
-a term $\jterm{{\Gamma}{A}}{\ctxwk{A}{B}}{f}$ a morphism from $A$ to $B$.
+\end{lem}
+
+\begin{proof}
+Consider family morphisms $\jhom{\Gamma}{A}{B}{f}$ and $\jhom{\Gamma}{B}{C}{g}$
+and a family $\jfam{{{\Gamma}{A}}{\ctxwk{A}{C}}}{R}$. Then we have the judgmental
+equalities
+\begin{align*}
+\subst{\jhomcomp{f}{g}}{R} & \jdeq \subst{{f}{\ctxwk{A}{g}}}{R}\\
+& \jdeq \subst{{f}{\ctxwk{A}{g}}}{\subst{f}{\ctxwk{{A}{B}}{R}}}\\
+& \jdeq \subst{f}{{\ctxwk{A}{g}}{\ctxwk{{A}{B}}{R}}}
+\end{align*}
+\end{proof}
+
+\begin{lem}
+We have the following inference rules about the compatibility of composition with
+weakening:
+\begin{align*}
+& \inference
+  {\jhom{\Gamma}{A}{B}{f}
+   \jhom{\Gamma}{B}{C}{g}
+   \jfam{{\Gamma}{A}}{P}}
+  {\jhomeq{\Gamma}{{A}{P}}{C}{\ctxwk{P}{\jhomcomp{f}{g}}}{\jhomcomp{\ctxwk{P}{f}}{g}}}\\
+& \inference
+  {\jterm{\Gamma}{B}{y}
+   \jhom{\Gamma}{B}{C}{g}}
+  {\jhomeq{\Gamma}{A}{C}{\jhomcomp{\ctxwk{A}{y}}{g}}{\ctxwk{A}{\subst{y}{g}}}}\\
+& \inference
+  {\jhom{\Gamma}{A}{B}{f}
+   \jterm{\Gamma}{C}{z}}
+  {\jhomeq{\Gamma}{A}{C}{\jhomcomp{f}{\ctxwk{B}{z}}}{\ctxwk{A}{z}}}
+\end{align*}
+\end{lem}
+
+\begin{lem}
+We have the following inference rules about the compatibility of composition with
+substitution:
+\begin{align*}
+& \inference
+  {\jhom{{\Gamma}{A}}{P}{Q}{f}
+   \jhom{{\Gamma}{A}}{Q}{R}{g}
+   \jterm{\Gamma}{A}{x}}
+  {\jhomeq{\Gamma}{\subst{x}{P}}{\subst{x}{R}}{\subst{x}{\jhomcomp{f}{g}}}{\jhomcomp{\subst{x}{f}}{\subst{x}{g}}}}\\
+& \inference
+  {\jhom{\Gamma}{A}{B}{f}
+   \jhom{\Gamma}{B}{C}{g}
+   \jterm{\Gamma}{A}{x}}
+  {\jtermeq{\Gamma}{C}{\subst{x}{\jhomcomp{f}{g}}}{\subst{{x}{f}}{g}}}
+\end{align*}
+\end{lem}
 
@@
 \begin{align}
-& \inference{\jfam{\Gamma}{A}}{\jterm{{\Gamma}{A}}{\ctxwk{A}{A}}{\idfunc[A]}}
-& & \inference{\jfameq{\Gamma}{A}{A'}}{\jtermeq{{\Gamma}{A}}{\ctxwk{A}{A}}{\idfunc[A]}{\idfunc[A']}}
+& \inference
+  {\jfam{\Gamma}{A}}
+  {\jterm{{\Gamma}{A}}{\ctxwk{A}{A}}{\idfunc[A]}}
+& & \inference
+    {\jfameq{\Gamma}{A}{A'}}
+    {\jtermeq{{\Gamma}{A}}{\ctxwk{A}{A}}{\idfunc[A]}{\idfunc[A']}}
 \end{align}
@@
 \begin{align}
-& \inference{\jterm{\Gamma}{A}{x}}{\jtermeq{\Gamma}{A}{\subst{x}{\idfunc[A]}}{x}}\label{idfunc-subst-defn}\\
-& \inference{\jfam{{\Gamma}{A}}{P}}{\jfameq{{\Gamma}{A}}{\subst{\idfunc[A]}{\ctxwk{A}{P}}}{P}}\label{idfunc-wk-defn}\\
-& \inference{\jfam{{\Gamma}{A}}{P}}{\jfameq{{\Gamma}{A}}{\subst{\idfunc[A]}{\ctxwk{{A}{A}}{P}}}{P}}\label{idfunc-wk-defn2}\\
-& \inference{\jterm{{\Gamma}{A}}{P}{f}}{\jtermeq{{\Gamma}{A}}{P}{\subst{\idfunc[A]}{\ctxwk{A}{f}}}{f}}\label{idfunc-precomp} \\
-& \inference{\jterm{{\Gamma}{A}}{P}{f}}{\jtermeq{{\Gamma}{A}}{P}{\subst{\idfunc[A]}{\ctxwk{{A}{A}}{f}}}{f}}\label{idfunc-precomp} \\
-& \inference{\jterm{{\Gamma}{B}}{\ctxwk{B}{A}}{g}}{\jtermeq{{\Gamma}{B}}{\ctxwk{B}{A}}{\subst{g}{\ctxwk{B}{\idfunc[A]}}}{g}}\label{idfunc-postcomp}
+& \inference
+  {\jterm{\Gamma}{A}{x}}
+  {\jtermeq{\Gamma}{A}{\subst{x}{\idfunc[A]}}{x}}
+  \label{idfunc-subst-defn}\\
+& \inference
+  {\jfam{{\Gamma}{A}}{P}}
+  {\jfameq{{\Gamma}{A}}{\subst{\idfunc[A]}{\ctxwk{A}{P}}}{P}}
+  \label{idfunc-wk-defn}\\
+& \inference
+  {\jfam{{\Gamma}{A}}{P}}
+  {\jfameq{{\Gamma}{A}}{\subst{\idfunc[A]}{\ctxwk{{A}{A}}{P}}}{P}}
+  \label{idfunc-wk-defn2}\\
+& \inference
+  {\jterm{{\Gamma}{A}}{P}{f}}
+  {\jtermeq{{\Gamma}{A}}{P}{\subst{\idfunc[A]}{\ctxwk{A}{f}}}{f}}
+  \label{idfunc-precomp}\\
+& \inference
+  {\jterm{{\Gamma}{A}}{P}{f}}
+  {\jtermeq{{\Gamma}{A}}{P}{\subst{\idfunc[A]}{\ctxwk{{A}{A}}{f}}}{f}}
+  \label{idfunc-precomp}\\
+& \inference
+  {\jterm{{\Gamma}{B}}{\ctxwk{B}{A}}{g}}
+  {\jtermeq{{\Gamma}{B}}{\ctxwk{B}{A}}{\subst{g}{\ctxwk{B}{\idfunc[A]}}}{g}}
+  \label{idfunc-postcomp}
 \end{align}
@@
 \begin{equation}\label{idfunc-wk-comp}
-\inference{\jfam{\Gamma}{A}\quad\jfam{\Gamma}{B}}{\jtermeq{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{{A}{B}}{{A}{B}}}{\idfunc[\ctxwk{A}{B}]}{\ctxwk{A}{\idfunc[B]}}}
+\inference
+  {\jfam{\Gamma}{A}
+   \jfam{\Gamma}{B}}
+  {\jtermeq{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{{A}{B}}{{A}{B}}}{\idfunc[\ctxwk{A}{B}]}{\ctxwk{A}{\idfunc[B]}}}
 \end{equation}
@@
 \begin{equation}\label{idfunc-subst-comp}
-\inference{\jterm{\Gamma}{A}{x}\quad\jfam{{\Gamma}{A}}{P}}{\jtermeq{{\Gamma}{\subst{x}{P}}}{\ctxwk{\subst{x}{P}}{\subst{x}{P}}}{\idfunc[\subst{x}{P}]}{\subst{x}{\idfunc[P]}}}
+\inference
+  {\jterm{\Gamma}{A}{x}
+   \jfam{{\Gamma}{A}}{P}}
+  {\jtermeq{{\Gamma}{\subst{x}{P}}}{\ctxwk{\subst{x}{P}}{\subst{x}{P}}}{\idfunc[\subst{x}{P}]}{\subst{x}{\idfunc[P]}}}
 \end{equation}
@@
 \begin{align*}
-& \inference{\jterm{\Gamma}{A}{x}\quad\jterm{\Gamma}{\subst{x}{P}}{u}\quad\jfam{{{\Gamma}{A}}{P}}{Q}}
+& \inference
+  {\jterm{\Gamma}{A}{x}
+   \jterm{\Gamma}{\subst{x}{P}}{u}
+   \jfam{{{\Gamma}{A}}{P}}{Q}}
   {\jfameq{\Gamma}{\subst{\ctxext{x}{u}}{Q}}{\subst{u}{{x}{Q}}}}\\
-& \inference{\jterm{\Gamma}{A}{x}\quad\jterm{\Gamma}{\subst{x}{P}}{u}\quad\jterm{{{\Gamma}{A}}{P}}{Q}{g}}
+& \inference
+  {\jterm{\Gamma}{A}{x}
+   \jterm{\Gamma}{\subst{x}{P}}{u}
+   \jterm{{{\Gamma}{A}}{P}}{Q}{g}}
   {\jtermeq{\Gamma}{\subst{u}{{x}{Q}}}{\subst{\ctxext{x}{u}}{g}}{\subst{u}{{x}{g}}}}\\
-& \inference{\jterm{\Gamma}{A}{x}\quad\jterm{\Gamma}{\subst{x}{P}}{u}\quad\jfam{{{{\Gamma}{A}}{P}}{Q}}{R}}
+& \inference
+  {\jterm{\Gamma}{A}{x}
+   \jterm{\Gamma}{\subst{x}{P}}{u}
+   \jfam{{{{\Gamma}{A}}{P}}{Q}}{R}}
   {\jfameq{{\Gamma}{\subst{u}{{x}{Q}}}}{\subst{\ctxext{x}{u}}{R}}{\subst{u}{{x}{R}}}}\\
-& \inference{\jterm{\Gamma}{A}{x}\quad\jterm{\Gamma}{\subst{x}{P}}{u}\quad\jterm{{{{\Gamma}{A}}{P}}{Q}}{R}{t}}
+& \inference
+  {\jterm{\Gamma}{A}{x}
+   \jterm{\Gamma}{\subst{x}{P}}{u}
+   \jterm{{{{\Gamma}{A}}{P}}{Q}}{R}{t}}
   {\jtermeq{{\Gamma}{\subst{u}{{x}{Q}}}}{\subst{u}{{x}{R}}}{\subst{\ctxext{x}{u}}{t}}{\subst{u}{{x}{t}}}}
@@
 \begin{equation}\label{idfunc-ext-comp}
-\inference{\jfam{\Gamma}{A}\quad\jfam{{\Gamma}{A}}{P}}{\jtermeq{{\Gamma}{{A}{P}}}{\ctxwk{\ctxext{A}{P}}{\ctxext{A}{P}}}{\idfunc[\ctxext{A}{P}]}{\ctxext{\cprojfstf{A}{P}}{\cprojsndf{A}{P}}}}
+\inference
+  {\jfam{\Gamma}{A}
+   \jfam{{\Gamma}{A}}{P}}
+  {\jtermeq{{\Gamma}{{A}{P}}}{\ctxwk{\ctxext{A}{P}}{\ctxext{A}{P}}}{\idfunc[\ctxext{A}{P}]}{\ctxext{\cprojfstf{A}{P}}{\cprojsndf{A}{P}}}}
 \end{equation}
@@
 \begin{equation*}
-\inference{\jfam{{\Gamma}{A}}{P}\quad\jfam{{\Gamma}{A}}{Q}}
+\inference
+  {\jfam{{\Gamma}{A}}{P}
+   \jfam{{\Gamma}{A}}{Q}}
   {\jfameq{{\Gamma}{{A}{P}}}{\subst{\cprojfstf{A}{P}}{\ctxwk{\ctxext{A}{P}}{Q}}}{\ctxwk{P}{Q}}}
@@
 \begin{align*}
-& \inference{\jterm{\Gamma}{\ctxext{A}{P}}{w}}{\jtermeq{\Gamma}{\ctxext{A}{P}}{\ctxext{\cprojfst{A}{P}{w}}{\cprojsnd{A}{P}{w}}}{w}}\\
-& \inference{\jterm{\Gamma}{A}{x}\quad\jterm{\Gamma}{\subst{x}{P}}{u}}{\jtermeq{\Gamma}{A}{\cprojfst{A}{P}{\ctxext{x}{u}}}{x}}\\
-& \inference{\jterm{\Gamma}{A}{x}\quad\jterm{\Gamma}{\subst{x}{P}}{u}}{\jtermeq{\Gamma}{\subst{x}{P}}{\cprojsnd{A}{P}{\ctxext{x}{u}}}{u}}
+& \inference
+  {\jterm{\Gamma}{\ctxext{A}{P}}{w}}
+  {\jtermeq{\Gamma}{\ctxext{A}{P}}{\ctxext{\cprojfst{A}{P}{w}}{\cprojsnd{A}{P}{w}}}{w}}\\
+& \inference
+  {\jterm{\Gamma}{A}{x}
+   \jterm{\Gamma}{\subst{x}{P}}{u}}
+  {\jtermeq{\Gamma}{A}{\cprojfst{A}{P}{\ctxext{x}{u}}}{x}}\\
+& \inference
+  {\jterm{\Gamma}{A}{x}
+   \jterm{\Gamma}{\subst{x}{P}}{u}}
+  {\jtermeq{\Gamma}{\subst{x}{P}}{\cprojsnd{A}{P}{\ctxext{x}{u}}}{u}}
 \end{align*}
@@
 \begin{align*}
-& \inference{\jterm{\Gamma}{A}{x}\quad\jterm{\Gamma}{\subst{x}{P}}{u}\quad\jterm{\Gamma}{\subst{\ctxext{x}{u}}{Q}}{v}}
+& \inference
+  {\jterm{\Gamma}{A}{x}
+   \jterm{\Gamma}{\subst{x}{P}}{u}
+   \jterm{\Gamma}{\subst{\ctxext{x}{u}}{Q}}{v}}
   {\jtermeq{\Gamma}{\ctxext{{A}{P}}{Q}}{\ctxext{x}{{u}{v}}}{\ctxext{{x}{u}}{v}}}\\
-& \inference{\jterm{\Gamma}{\ctxext{{A}{P}}{Q}}{w}}
+& \inference
+  {\jterm{\Gamma}{\ctxext{{A}{P}}{Q}}{w}}
   {\jtermeq{\Gamma}{A}{\cprojfst{A}{P}{\cprojfst{\ctxext{A}{P}}{Q}{w}}}{\cprojfst{A}{\ctxext{P}{Q}}{w}}}\\
-& \inference{\jterm{\Gamma}{\ctxext{{A}{P}}{Q}}{w}}
+& \inference
+  {\jterm{\Gamma}{\ctxext{{A}{P}}{Q}}{w}}
   {\jtermeq{\Gamma}{\subst{\cprojfst{A}{\ctxext{P}{Q}}{w}}{P}}
@@
     {\cprojfst{P}{Q}{\cprojsnd{A}{\ctxext{P}{Q}}{w}}}}\\
-& \inference{\jterm{\Gamma}{\ctxext{{A}{P}}{Q}}{w}}
+& \inference
+  {\jterm{\Gamma}{\ctxext{{A}{P}}{Q}}{w}}
   {\jtermeq{\Gamma}{\subst{\cprojfst{\ctxext{A}{P}}{Q}{w}}{Q}}
@@
 \begin{align*}
-& \inference{\jterm{{\Gamma}{B}}{Q}{g}\quad\jterm{{\Gamma}{B}}{\subst{g}{R}}{t}}{\jtermeq{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{\ctxext{Q}{R}}}{\ctxwk{A}{\ctxext{g}{t}}}{\ctxext{\ctxwk{A}{g}}{\ctxwk{A}{t}}}}
+& \inference
+  {\jterm{{\Gamma}{B}}{Q}{g}
+   \jterm{{\Gamma}{B}}{\subst{g}{R}}{t}}
+  {\jtermeq
+    {{{\Gamma}{A}}{\ctxwk{A}{B}}}
+    {\ctxwk{A}{\ctxext{Q}{R}}}
+    {\ctxwk{A}{\ctxext{g}{t}}}
+    {\ctxext{\ctxwk{A}{g}}{\ctxwk{A}{t}}}}
 % \\
@@
 \begin{align*}
-& \inference{\jterm{{{\Gamma}{A}}{P}}{Q}{g}\quad\jterm{{{\Gamma}{A}}{P}}{\subst{g}{R}}{t}}{\jtermeq{{\Gamma}{\subst{x}{P}}}{\ctxext{\subst{x}{Q}}{\subst{x}{R}}}{\subst{x}{\ctxext{g}{t}}}{\ctxext{\subst{x}{g}}{\subst{x}{t}}}}
+& \inference
+  {\jterm{{{\Gamma}{A}}{P}}{Q}{g}
+   \jterm{{{\Gamma}{A}}{P}}{\subst{g}{R}}{t}}
+  {\jtermeq
+    {{\Gamma}{\subst{x}{P}}}
+    {\ctxext{\subst{x}{Q}}{\subst{x}{R}}}
+    {\subst{x}{\ctxext{g}{t}}}
+    {\ctxext{\subst{x}{g}}{\subst{x}{t}}}}
 \end{align*}
@@
 \begin{align*}
-& \inference{\jvtype{i}{\Gamma}{A}}{\jvfam{i}{\Gamma}{A}} & & \inference{\jvtypeeq{i}{\Gamma}{A}{B}}{\jvfameq{i}{\Gamma}{A}{B}}\\
-& \inference{\jvtype{i}{\Gamma}{A}\quad\jvfameq{i}{\Gamma}{A}{B}}{\jvtype{i}{\Gamma}{B}}
-& & \inference{\jvtype{i}{\Gamma}{A}\quad\jvfameq{i}{\Gamma}{A}{B}}{\jvtypeeq{i}{\Gamma}{A}{B}}
+& \inference
+  {\jvtype{i}{\Gamma}{A}}
+  {\jvfam{i}{\Gamma}{A}}
+& & \inference
+    {\jvtypeeq{i}{\Gamma}{A}{B}}
+    {\jvfameq{i}{\Gamma}{A}{B}}\\
+& \inference
+  {\jvtype{i}{\Gamma}{A}
+   \jvfameq{i}{\Gamma}{A}{B}}
+  {\jvtype{i}{\Gamma}{B}}
+& & \inference
+    {\jvtype{i}{\Gamma}{A}
+     \jvfameq{i}{\Gamma}{A}{B}}
+    {\jvtypeeq{i}{\Gamma}{A}{B}}
 \end{align*}
@@
 \begin{align*}
-& \inference{\jvctx{\Gamma}}{\jvctxeq{\Gamma}{\Gamma}} 
-& & \inference{\jvctxeq{\Gamma}{\Delta}}{\jvctxeq{\Delta}{\Gamma}} 
-& & \inference{\jvctxeq{\Gamma}{\Delta}\qquad\jvctxeq{\Delta}{\greek{E}}}{\jvctxeq{\Gamma}{\greek{E}}}\\
-& \inference{\jvfam{i}{\Gamma}{A}}{\jvfameq{i}{\Gamma}{A}{A}} 
-& & \inference{\jvfameq{i}{\Gamma}{A}{B}}{\jvfameq{i}{\Gamma}{B}{A}}
-& & \inference{\jvfameq{i}{\Gamma}{A}{B}\qquad\jvfameq{i}{\Gamma}{B}{C}}{\jvfameq{i}{\Gamma}{A}{C}}\\
-& \inference{\jvterm{i}{\Gamma}{A}{x}}{\jvtermeq{i}{\Gamma}{A}{x}{x}}
-& & \inference{\jvtermeq{i}{\Gamma}{A}{x}{y}}{\jvtermeq{i}{\Gamma}{A}{y}{x}}
-& & \inference{\jvtermeq{i}{\Gamma}{A}{x}{y}\qquad\jvtermeq{i}{\Gamma}{A}{y}{z}}{\jvtermeq{i}{\Gamma}{A}{x}{z}}
+& \inference
+  {\jvctx{\Gamma}}
+  {\jvctxeq{\Gamma}{\Gamma}} 
+& & \inference
+    {\jvctxeq{\Gamma}{\Delta}}
+    {\jvctxeq{\Delta}{\Gamma}} 
+& & \inference
+    {\jvctxeq{\Gamma}{\Delta}
+     \jvctxeq{\Delta}{\greek{E}}}
+    {\jvctxeq{\Gamma}{\greek{E}}}\\
+& \inference
+  {\jvfam{i}{\Gamma}{A}}
+  {\jvfameq{i}{\Gamma}{A}{A}} 
+& & \inference
+    {\jvfameq{i}{\Gamma}{A}{B}}
+    {\jvfameq{i}{\Gamma}{B}{A}}
+& & \inference
+    {\jvfameq{i}{\Gamma}{A}{B}
+     \jvfameq{i}{\Gamma}{B}{C}}
+    {\jvfameq{i}{\Gamma}{A}{C}}\\
+& \inference
+  {\jvterm{i}{\Gamma}{A}{x}}
+  {\jvtermeq{i}{\Gamma}{A}{x}{x}}
+& & \inference
+    {\jvtermeq{i}{\Gamma}{A}{x}{y}}
+    {\jvtermeq{i}{\Gamma}{A}{y}{x}}
+& & \inference
+    {\jvtermeq{i}{\Gamma}{A}{x}{y}
+     \jvtermeq{i}{\Gamma}{A}{y}{z}}
+    {\jvtermeq{i}{\Gamma}{A}{x}{z}}
 \end{align*}
@@
 \begin{align*}
-& \inference{\jvctxeq{\Gamma}{\Delta}\qquad\jvfam{i}{\Gamma}{A}}{\jvfam{i}{\Delta}{A}}
-& & \inference{\jvctxeq{\Gamma}{\Delta}\qquad\jvfameq{i}{\Gamma}{A}{B}}{\jvfameq{i}{\Delta}{A}{B}}\\
-& \inference{\jvctxeq{\Gamma}{\Delta}\qquad\jvterm{i}{\Gamma}{A}{x}}{\jvterm{i}{\Delta}{A}{x}}
-& & \inference{\jvctxeq{\Gamma}{\Delta}\qquad\jvtermeq{i}{\Gamma}{A}{x}{y}}{\jvtermeq{i}{\Delta}{A}{x}{y}}\\
-& \inference{\jvfameq{i}{\Gamma}{A}{B}\qquad \jvterm{i}{\Gamma}{A}{x}}{\jvterm{i}{\Gamma}{B}{x}}
-& & \inference{\jvfameq{i}{\Gamma}{A}{B}\qquad\jvtermeq{i}{\Gamma}{A}{x}{y}}{\jvtermeq{i}{\Gamma}{B}{x}{y}}
+& \inference
+  {\jvctxeq{\Gamma}{\Delta}
+   \jvfam{i}{\Gamma}{A}}
+  {\jvfam{i}{\Delta}{A}}
+& & \inference
+    {\jvctxeq{\Gamma}{\Delta}
+     \jvfameq{i}{\Gamma}{A}{B}}
+    {\jvfameq{i}{\Delta}{A}{B}}\\
+& \inference
+  {\jvctxeq{\Gamma}{\Delta}
+   \jvterm{i}{\Gamma}{A}{x}}
+  {\jvterm{i}{\Delta}{A}{x}}
+& & \inference
+    {\jvctxeq{\Gamma}{\Delta}
+     \jvtermeq{i}{\Gamma}{A}{x}{y}}
+    {\jvtermeq{i}{\Delta}{A}{x}{y}}\\
+& \inference
+  {\jvfameq{i}{\Gamma}{A}{B}
+   \jvterm{i}{\Gamma}{A}{x}}
+  {\jvterm{i}{\Gamma}{B}{x}}
+& & \inference
+    {\jvfameq{i}{\Gamma}{A}{B}
+     \jvtermeq{i}{\Gamma}{A}{x}{y}}
+    {\jvtermeq{i}{\Gamma}{B}{x}{y}}
 \end{align*}
@@
 \begin{align}
-& \inference{}{\jctx{\emptyc}}\\
-& \inference{\jctx{\Gamma}}{\jvfam{i}{\Gamma}{\emptyf[\Gamma]}}\\
-& \inference{\jctx{\Gamma}}{\jvterm{i}{\Gamma}{\emptyf[\Gamma]}{\emptytm[\Gamma]}}\\
-& \inference{\jvterm{i}{\Gamma}{\emptyf[\Gamma]}{x}}{\jvtermeq{i}{\Gamma}{\emptyf[\Gamma]}{x}{\emptytm[\Gamma]}}
+& \inference
+  {}
+  {\jctx{\emptyc}}\\
+& \inference
+  {\jctx{\Gamma}}
+  {\jvfam{i}{\Gamma}{\emptyf[\Gamma]}}\\
+& \inference
+  {\jctx{\Gamma}}
+  {\jvterm{i}{\Gamma}{\emptyf[\Gamma]}{\emptytm[\Gamma]}}\\
+& \inference
+  {\jvterm{i}{\Gamma}{\emptyf[\Gamma]}{x}}
+  {\jvtermeq{i}{\Gamma}{\emptyf[\Gamma]}{x}{\emptytm[\Gamma]}}
 \end{align}
@@
 \begin{align}
-& \inference{\jctx{\Gamma}}{\jvfam{\nameless}{\emptyc}{\Gamma}} 
-& & \inference{\jvfam{\nameless}{\emptyc}{\Gamma}}{\jctx{\Gamma}}\\
-& \inference{\jctxeq{\Gamma}{\Delta}}{\jvfameq{\nameless}{\emptyc}{\Gamma}{\Delta}}
-& & \inference{\jvfameq{\nameless}{\emptyc}{\Gamma}{\Delta}}{\jctxeq{\Gamma}{\Delta}}
+& \inference
+  {\jctx{\Gamma}}
+  {\jvfam{\nameless}{\emptyc}{\Gamma}} 
+& & \inference
+    {\jvfam{\nameless}{\emptyc}{\Gamma}}
+    {\jctx{\Gamma}}\\
+& \inference
+  {\jctxeq{\Gamma}{\Delta}}
+  {\jvfameq{\nameless}{\emptyc}{\Gamma}{\Delta}}
+& & \inference
+    {\jvfameq{\nameless}{\emptyc}{\Gamma}{\Delta}}
+    {\jctxeq{\Gamma}{\Delta}}
 \end{align}
@@
 \begin{equation}
-\inference{}{\jvfameq{\nameless}{\emptyc}{\emptyc}{\emptyf[\emptyc]}}
+\inference
+  {}
+  {\jvfameq{\nameless}{\emptyc}{\emptyc}{\emptyf[\emptyc]}}
 \end{equation}
@@
 \begin{align}
-& \inference{\jvfam{i}{\Gamma}{A}}{\jvfamcombi{{i}{x}}{{\Gamma}{A}}{P}}
-& & \inference{\jctxeq{\Gamma}{\Delta}\qquad\jfameq{\Gamma}{A}{B}}{\jctxeq{\ctxext{\Gamma}{A}}{\ctxext{\Delta}{B}}}\\
-& \inference{\jfam{{\Gamma}{A}}{P}}{\jfam{\Gamma}{\ctxext{A}{P}}}
-& & \inference{\jfameq{\Gamma}{A}{B}\qquad\jfameq{{\Gamma}{A}}{P}{Q}}{\jfameq{\Gamma}{\ctxext{A}{P}}{\ctxext{B}{Q}}}
+& \inference
+  {\jvfam{i}{\Gamma}{A}}
+  {\jvfamcombi{{i}{x}}{{\Gamma}{A}}{P}}
+& & \inference
+    {\jctxeq{\Gamma}{\Delta}
+     \jfameq{\Gamma}{A}{B}}
+    {\jctxeq{\ctxext{\Gamma}{A}}{\ctxext{\Delta}{B}}}\\
+& \inference
+  {\jfam{{\Gamma}{A}}{P}}
+  {\jfam{\Gamma}{\ctxext{A}{P}}}
+& & \inference
+    {\jfameq{\Gamma}{A}{B}
+     \jfameq{{\Gamma}{A}}{P}{Q}}
+    {\jfameq{\Gamma}{\ctxext{A}{P}}{\ctxext{B}{Q}}}
 \end{align}

### 2014-03-02 article-models-tt.tex
filled in all the proofs

@@
 \begin{align*}
-\jfamdefn*{{\Gamma}{A}}{\jhomcomp{f}{Q}}{\subst{f}{\ctxwk{A}{Q}}}\\
-\jtermdefn*{{\Gamma}{A}}{\jhomcomp{f}{Q}}{\jhomcomp{f}{g}}{\subst{f}{\ctxwk{A}{g}}}.
+\jfamdefn*{{\Gamma}{A}}{\jcomp{f}{Q}}{\subst{f}{\ctxwk{A}{Q}}}\\
+\jtermdefn*{{\Gamma}{A}}{\jcomp{f}{Q}}{\jcomp{f}{g}}{\subst{f}{\ctxwk{A}{g}}}.
 \end{align*}
@@
 \begin{align*}
-\jfamdefn*{{{\Gamma}{A}}{\jhomcomp{f}{Q}}}{\jhomcomp{f}{R}}{\subst{f}{\ctxwk{A}{R}}}\\
-\jtermdefn*{{{\Gamma}{A}}{\jhomcomp{f}{Q}}}{\jhomcomp{f}{R}}{\jhomcomp{f}{h}}{\subst{f}{\ctxwk{A}{h}}}.
+\jfamdefn*{{{\Gamma}{A}}{\jcomp{f}{Q}}}{\jcomp{f}{R}}{\subst{f}{\ctxwk{A}{R}}}\\
+\jtermdefn*{{{\Gamma}{A}}{\jcomp{f}{Q}}}{\jcomp{f}{R}}{\jcomp{f}{h}}{\subst{f}{\ctxwk{A}{h}}}.
 \end{align*}
@@
    \jfam{{{\Gamma}{A}}{\ctxwk{A}{C}}}{R}}
-  {\jfameq{{\Gamma}{A}}{\subst{\jhomcomp{f}{g}}{R}}{\subst{f}{{\ctxwk{A}{g}}{\ctxwk{{A}{B}}{R}}}}}\\
+  {\jfameq
+    {{\Gamma}{A}}
+    {\subst{\jcomp{f}{g}}{R}}
+    {\subst{f}{{\ctxwk{A}{g}}{\ctxwk{{A}{B}}{R}}}}}\\
 & \inference
@@
    \jterm{{{\Gamma}{A}}{\ctxwk{A}{C}}}{R}{h}}
-  {\jfameq{{\Gamma}{A}}{\subst{\jhomcomp{f}{g}}{h}}{\subst{f}{{\ctxwk{A}{g}}{\ctxwk{{A}{B}}{h}}}}}
-\end{align*}
-In particular, when $\jhom{\Gamma}{C}{D}{h}$, we have
-\begin{equation*}
-\jhomeq{\Gamma}{A}{D}{\jhomcomp{\jhomcomp{f}{g}}{h}}{\jhomcomp{f}{\jhomcomp{g}{h}}}.
-\end{equation*}
+  {\jfameq
+    {{\Gamma}{A}}
+    {\subst{\jcomp{f}{g}}{h}}
+    {\subst{f}{{\ctxwk{A}{g}}{\ctxwk{{A}{B}}{h}}}}}
+\end{align*}
+We also have the following related inference rules, asserting that composition
+is strictly associative:
+\begin{align*}
+& \inference
+  {\jhom{\Gamma}{A}{B}{f}
+   \jhom{\Gamma}{B}{C}{g}
+   \jfam{{\Gamma}{C}}{R}}
+  {\jfameq
+    {{\Gamma}{A}}
+    {\jcomp{{f}{g}}{R}}
+    {\jcomp{f}{{g}{R}}}}\\
+& \inference
+  {\jhom{\Gamma}{A}{B}{f}
+   \jhom{\Gamma}{B}{C}{g}
+   \jterm{{\Gamma}{C}}{R}{h}}
+  {\jtermeq
+    {{\Gamma}{A}}
+    {\jcomp{{f}{g}}{R}}
+    {\jcomp{{f}{g}}{h}}
+    {\jcomp{f}{{g}{h}}}}
+\end{align*}
 \end{lem}
@@
 \begin{align*}
-\subst{\jhomcomp{f}{g}}{R} & \jdeq \subst{{f}{\ctxwk{A}{g}}}{R}\\
+\subst{\jcomp{f}{g}}{R} & \jdeq \subst{{f}{\ctxwk{A}{g}}}{R}\\
 & \jdeq \subst{{f}{\ctxwk{A}{g}}}{\subst{f}{\ctxwk{{A}{B}}{R}}}\\
@@
 \end{align*}
+The proof that 
+$\subst{\jcomp{f}{g}}{h}\jdeq\subst{f}{{\ctxwk{A}{g}}{\ctxwk{{A}{B}}{h}}}$
+is similar.
+
+Now suppose that $\jfam{{\Gamma}{C}}{R}$ instead. Then we have
+\begin{align*}
+\jcomp{{f}{g}}{R} & \jdeq \subst{\jcomp{f}{g}}{\ctxwk{A}{R}}\\
+& \jdeq \subst{{f}{\ctxwk{A}{g}}}{\ctxwk{A}{R}}\\
+& \jdeq \subst{f}{{\ctxwk{A}{g}}{\ctxwk{{A}{B}}{{A}{R}}}}\\
+& \jdeq \subst{f}{{\ctxwk{A}{g}}{\ctxwk{A}{{B}{R}}}}\\
+& \jdeq \subst{f}{\ctxwk{A}{\subst{g}{\ctxwk{B}{R}}}}\\
+& \jdeq \subst{f}{\ctxwk{A}{\jcomp{g}{R}}}\\
+& \jdeq \jcomp{f}{{g}{R}}.
+\end{align*}
+Again, the proof is similar for terms $h$ of $R$ in context $\ctxext{\Gamma}{C}$.
 \end{proof}
@@
    \jfam{{\Gamma}{A}}{P}}
-  {\jhomeq{\Gamma}{{A}{P}}{C}{\ctxwk{P}{\jhomcomp{f}{g}}}{\jhomcomp{\ctxwk{P}{f}}{g}}}\\
+  {\jhomeq{\Gamma}{{A}{P}}{C}
+    {\ctxwk{P}{\jcomp{f}{g}}}
+    {\jcomp{\ctxwk{P}{f}}{g}}}\\
 & \inference
@@
    \jhom{\Gamma}{B}{C}{g}}
-  {\jhomeq{\Gamma}{A}{C}{\jhomcomp{\ctxwk{A}{y}}{g}}{\ctxwk{A}{\subst{y}{g}}}}\\
+  {\jhomeq{\Gamma}{A}{C}
+    {\jcomp{\ctxwk{A}{y}}{g}}
+    {\ctxwk{A}{\subst{y}{g}}}}\\
 & \inference
@@
    \jterm{\Gamma}{C}{z}}
-  {\jhomeq{\Gamma}{A}{C}{\jhomcomp{f}{\ctxwk{B}{z}}}{\ctxwk{A}{z}}}
+  {\jhomeq{\Gamma}{A}{C}
+    {\jcomp{f}{\ctxwk{B}{z}}}
+    {\ctxwk{A}{z}}}
 \end{align*}
 \end{lem}
+
+\begin{proof}
+Let $\jhom{\Gamma}{A}{B}{f}$, $\jhom{\Gamma}{B}{C}{g}$ and $\jfam{{\Gamma}{A}}{P}$.
+Then we have the judgmental equalities
+\begin{align*}
+\ctxwk{P}{\jcomp{f}{g}} 
+& \jdeq \ctxwk{P}{\subst{f}{\ctxwk{A}{g}}}\\
+& \jdeq \subst{\ctxwk{P}{f}}{\ctxwk{P}{{A}{g}}}\\
+& \jdeq \subst{\ctxwk{P}{f}}{\ctxwk{\ctxext{A}{P}}{g}}\\
+& \jdeq \jcomp{\ctxwk{P}{f}}{g}.
+\end{align*}
+Now let $\jterm{\Gamma}{B}{y}$ and $\jhom{\Gamma}{B}{C}{g}$. Then we have the
+judgmental equalities
+\begin{align*}
+\jcomp{\ctxwk{A}{y}}{g}
+& \jdeq \subst{\ctxwk{A}{y}}{\ctxwk{A}{g}}\\
+& \jdeq \ctxwk{A}{\subst{y}{g}}.
+\end{align*}
+For the third assertion, let $\jhom{\Gamma}{A}{B}{f}$ and $\jterm{\Gamma}{C}{z}$.
+Then we have the judgmental equalities
+\begin{align*}
+\jcomp{f}{\ctxwk{B}{z}} 
+& \jdeq \subst{f}{\ctxwk{A}{{B}{z}}}\\
+& \jdeq \subst{f}{\ctxwk{{A}{B}}{{A}{z}}}\\
+& \jdeq \ctxwk{A}{z}.
+\end{align*}
+\end{proof}
 
@@
    \jterm{\Gamma}{A}{x}}
-  {\jhomeq{\Gamma}{\subst{x}{P}}{\subst{x}{R}}{\subst{x}{\jhomcomp{f}{g}}}{\jhomcomp{\subst{x}{f}}{\subst{x}{g}}}}\\
+  {\jhomeq{\Gamma}{\subst{x}{P}}{\subst{x}{R}}
+    {\subst{x}{\jcomp{f}{g}}}
+    {\jcomp{\subst{x}{f}}{\subst{x}{g}}}}\\
 & \inference
@@
    \jterm{\Gamma}{A}{x}}
-  {\jtermeq{\Gamma}{C}{\subst{x}{\jhomcomp{f}{g}}}{\subst{{x}{f}}{g}}}
+  {\jtermeq{\Gamma}{C}
+    {\subst{x}{\jcomp{f}{g}}}
+    {\subst{{x}{f}}{g}}}
 \end{align*}
 \end{lem}
+
+\begin{proof}
+Let $\jhom{{\Gamma}{A}}{P}{Q}{f}$, $\jhom{{\Gamma}{A}}{Q}{R}{g}$ and $\jterm{\Gamma}{A}{x}$.
+Then we have the judgmental equalities
+\begin{align*}
+\subst{x}{\jcomp{f}{g}}
+& \jdeq \subst{x}{{f}{\ctxwk{P}{g}}}\\
+& \jdeq \subst{{x}{f}}{{x}{\ctxwk{P}{g}}}\\
+& \jdeq \subst{{x}{f}}{\ctxwk{\subst{x}{P}}{\subst{x}{g}}}\\
+& \jdeq \jcomp{\subst{x}{f}}{\subst{x}{g}}.
+\end{align*}
+Now let $\jhom{\Gamma}{A}{B}{f}$, $\jhom{\Gamma}{B}{C}{g}$ and $\jterm{\Gamma}{A}{x}$.
+Then we have the judgmental equalities
+\begin{align*}
+\subst{x}{\jcomp{f}{g}}
+& \jdeq \subst{x}{{f}{\ctxwk{A}{g}}}\\
+& \jdeq \subst{{x}{f}}{{x}{\ctxwk{A}{g}}}\\
+& \jdeq \subst{{x}{f}}{g}.
+\end{align*}
+\end{proof}
 
@@
    \jfam{{\Gamma}{A}}{P}}
-  {\jtermeq{{\Gamma}{{A}{P}}}{\ctxwk{\ctxext{A}{P}}{\ctxext{A}{P}}}{\idfunc[\ctxext{A}{P}]}{\ctxext{\cprojfstf{A}{P}}{\cprojsndf{A}{P}}}}
+  {\jtermeq
+    {{\Gamma}{{A}{P}}}
+    {\ctxwk{\ctxext{A}{P}}{\ctxext{A}{P}}}
+    {\idfunc[\ctxext{A}{P}]}
+    {\ctxext{\cprojfstf{A}{P}}{\cprojsndf{A}{P}}}}
 \end{equation}
@@
 & \jdeq \subst{\cprojsndf{A}{P}}{{\cprojfstf{A}{P}}{\idfunc[\ctxext{\ctxwk{\ctxext{A}{P}}{A}}{\ctxwk{\ctxext{A}{P}}{P}}]}}\\
-& \jdeq \subst{\cprojsndf{A}{P}}{{\cprojfstf{A}{P}}{\idfunc[\ctxwk{\ctxext{A}{P}}{\ctxext{A}{P}}]}}\\
-& \jdeq \subst{\idfunc[P]}{{\ctxwk{P}{\idfunc[A]}}{\idfunc[\ctxwk{\ctxext{A}{P}}{\ctxext{A}{P}}]}}\\
-& \jdeq \subst{\idfunc[P]}{{\ctxwk{P}{\idfunc[A]}}{\ctxwk{\ctxext{A}{P}}{\idfunc[\ctxext{A}{P}]}}}\\
-& \jdeq \subst{\idfunc[P]}{{\ctxwk{P}{\idfunc[A]}}{\ctxwk{P}{{A}{\idfunc[\ctxext{A}{P}]}}}}\\
-& \jdeq \subst{\idfunc[P]}{\ctxwk{P}{\subst{\idfunc[A]}{\ctxwk{A}{\idfunc[\ctxext{A}{P}]}}}}\\
-& \jdeq \subst{\idfunc[A]}{\ctxwk{A}{\idfunc[\ctxext{A}{P}]}}\\
+& \jdeq \subst
+          {\cprojsndf{A}{P}}
+          { {\cprojfstf{A}{P}}
+            {\idfunc[\ctxwk{\ctxext{A}{P}}{\ctxext{A}{P}}]}}\\
+& \jdeq \subst
+          {\idfunc[P]}
+          { {\ctxwk{P}{\idfunc[A]}}
+            {\idfunc[\ctxwk{\ctxext{A}{P}}{\ctxext{A}{P}}]}}\\
+& \jdeq \subst
+          {\idfunc[P]}
+          { {\ctxwk{P}{\idfunc[A]}}
+            {\ctxwk{\ctxext{A}{P}}{\idfunc[\ctxext{A}{P}]}}}\\
+& \jdeq \subst
+          {\idfunc[P]}
+          { {\ctxwk{P}{\idfunc[A]}}
+            {\ctxwk{P}{{A}{\idfunc[\ctxext{A}{P}]}}}}\\
+& \jdeq \subst
+          {\idfunc[P]}
+          {\ctxwk
+            {P}
+            {\subst
+              {\idfunc[A]}
+              {\ctxwk{A}{\idfunc[\ctxext{A}{P}]}}}}\\
+& \jdeq \subst
+          {\idfunc[A]}
+          {\ctxwk{A}{\idfunc[\ctxext{A}{P}]}}\\
 & \jdeq \idfunc[\ctxext{A}{P}].

### 2014-03-03 article-models-tt.tex
motivating property made into lemma
compatibility of comp with ctxext
a start for issue #27

@@
 compatibility of weakening with itself, we see that we can show
+
+\begin{lem}
+The inference rule
 \begin{equation*}
-\jfameq{{\Gamma}{A}}{\subst{f}{\ctxwk{A}{{B}{C}}}}{\ctxwk{A}{C}}
+\inference
+  {\jfam{\Gamma}{A}
+   \jfam{\Gamma}{B}
+   \jfam{\Gamma}{C}
+   \jhom{\Gamma}{A}{B}{f}}
+  {\jfameq{{\Gamma}{A}}
+    {\subst{f}{\ctxwk{A}{{B}{C}}}}
+    {\ctxwk{A}{C}}}
 \end{equation*}
-for any three families $A$, $B$ and $C$ in context $\Gamma$ and any $\jterm{{\Gamma}{A}}{\ctxwk{A}{B}}{f}$.
+is valid.
+\end{lem}
+
+\begin{proof}
+Let $\jfam{\Gamma}{A}$, $\jfam{\Gamma}{B}$, $\jfam{\Gamma}{C}$ and $\jhom{\Gamma}{A}{B}{f}$.
+Then we have the judgmental equalities
+\begin{align*}
+\subst{f}{\ctxwk{A}{{B}{C}}}
+& \jdeq \subst{f}{\ctxwk{{A}{B}}{{A}{C}}}\\
+& \jdeq \ctxwk{A}{C}.
+\end{align*}
+\end{proof}
+
 It follows that for $\jterm{{\Gamma}{B}}{\ctxwk{B}{C}}{g}$ we can compose $f$
@@
 
+We also have the following lemma about the compatibility of pairing and composition:
+
+\begin{lem}
+The following inference rule is valid
+\begin{align*}
+& \inference
+  {\jhom{\Gamma}{A}{B}{f}
+   \jhom{\Gamma}{B}{C}{g}
+   \jfam{{\Gamma}{C}}{R}
+   \jterm{{\Gamma}{B}}{\subst{g}{\ctxwk{B}{R}}}{w}}
+  {\jhomeq{\Gamma}{A}{{C}{R}}
+    {\jcomp{f}{\ctxext{g}{w}}}
+    {\ctxext{\jcomp{f}{g}}{\jcomp{f}{w}}}}
+\end{align*}
+\end{lem}
+
+\begin{proof}
+Let $\jhom{\Gamma}{A}{B}{f}$, $\jhom{\Gamma}{B}{C}{g}$, $\jfam{{\Gamma}{C}}{R}$
+and $\jterm{{\Gamma}{B}}{\subst{g}{\ctxwk{B}{R}}}{w}$. Then we have the
+judgmental equalities
+\begin{align*}
+\jcomp{f}{\ctxext{g}{w}}
+& \jdeq \subst{f}{\ctxwk{A}{\ctxext{g}{w}}}\\
+& \jdeq \subst{f}{\ctxext{\ctxwk{A}{g}}{\ctxwk{A}{w}}}\\
+& \jdeq \ctxext{\subst{f}{\ctxwk{A}{g}}}{\subst{f}{\ctxwk{A}{w}}}\\
+& \jdeq \ctxext{\jcomp{f}{g}}{\jcomp{f}{w}}.
+\end{align*}
+\end{proof}
+
+\subsection{The unit type in structural type theory}
+In this section we explore what we get if we pose compatibility conditions on
+an inductively defined unit type. We will assume that not only the unit type
+is compatible with extension, weakening, substitution and identity functions,
+but also its induction principle should be compatible with those.
+
 \section{Pretty type theory}

### 2014-03-04 article-models-tt.tex
outlined the structure of the type constructors section
starting to work on models

@@
-\section{Structural type theory}\label{tt}
+\section{Type theory before type constructors}\label{tt}
 In this section we give a description of type theory before basic
@@
 \end{proof}
+
+There is also a notion of morphism \emph{over} a morphism. We will develop this
+notion because it will be needed in the theory of models later on.
+
+\begin{defn}
+Let $\jhom{\Gamma}{A}{B}{f}$ be a morphism from $A$ to $B$ in context $\Gamma$
+and consider $\jfam{{\Gamma}{A}}{P}$ and $\jfam{{\Gamma}{B}}{Q}$. A morphism
+from $P$ to $Q$ over $f$ in context $\Gamma$ is a term
+\begin{equation*}
+\jfhom{\Gamma}{A}{B}{f}{P}{Q}{F}.
+\end{equation*}
+\end{defn}
+
+We can also define composition for this version of morphism and formulate
+compatibility rules for them.
 
@@
 \end{proof}
+%\end{document}
+
+\section{Type constructors in structural type theory}
+In this section we will introduce the usual type constructors to structural
+type theory. We do so with the point of view that each type constructor is a
+(class of) operations on type theory that should be compatible with extension,
+weakening, substitution and identity morphisms. Likewise, each induction
+principle of an inductively defined type constructor is going to be such an
+operator, and it is therefore also required to be compatible with extension,
+weakening, substitution and identity functions.
+
+We will first state all the rules of the individual type constructors, where
+only the $\wtypesym$-type constructor depends on the presence of dependent
+function types (or universes too?), and hence comes with the notion that $\wtypesym$
+is compatible with dependent product types and vice versa. All the other
+compatibility properties are stated in \autoref{compatibility-of-type-constructors}.
+
+\subsubsection{Issues to keep in mind}
+\begin{enumerate}
+\item The rule asserting that unit types are compatible with context extension makes
+no sense from the point of view that contexts are lists of types and types
+are contexts of length one; in other words, that types are indecomposable (i.e.~non-extended)
+families. Remidies:
+\begin{enumerate}
+\item This point of view is wrong and should be abandoned.
+\item We have, as Vladimir proposed, two kinds of judgmental equalities. One
+      could be used for the very strict equalities, the other could be used
+      to state the compatibility rules with. In other words, the other equality
+      is a compatibility relation. If we do that, we shouldn't require that
+      if a family is compatible with a type, then the family is a type. The
+      compatibility relation could relate things by uniqueness up to unique
+      isomorphism.
+\item Don't state such compatibility rules for the type constructors.
+\item Don't even give the unit type a dependent action.
+\end{enumerate}
+\end{enumerate}
 
@@
 but also its induction principle should be compatible with those.
+
+\begin{align*}
+& \inference
+  {\jfam{\Gamma}{A}}
+  {\jtype{{\Gamma}{A}}{\unitc{A}}}
+& & \inference
+    {\jfameq{\Gamma}{A}{A'}}
+    {\jtypeeq{{\Gamma}{A}}
+      {\unitc{A}}
+      {\unitc{A'}}}\\
+& \inference
+  {\jfam{\Gamma}{A}}
+  {\jterm{{\Gamma}{A}}{\unitc{A}}{\unitct{A}}}
+& & \inference
+    {\jfameq{\Gamma}{A}{A'}}
+    {\jtermeq{{\Gamma}{A}}{\unitc{A}}{\unitct{A}}{\unitct{A'}}}\\
+& \inference
+  {\jfam{{\Gamma}{A}}{P}}
+  {\jtype{{{{\Gamma}{A}}{P}}{\ctxwk{P}{\unitc{A}}}}{\unitf{A}{P}}}
+& & \inference
+    {\jfameq{{\Gamma}{A}}{P}{P'}}
+    {\jtypeeq{{{{\Gamma}{A}}{P}}{\ctxwk{P}{\unitc{A}}}}
+      {\unitf{A}{P}}
+      {\unitf{A}{P'}}}\\
+& \inference
+  {\jfam{{\Gamma}{A}}{P}}
+  {\jterm{{{{\Gamma}{A}}{P}}{\ctxwk{P}{\unitc{A}}}}{\unitf{A}{P}}{\unitft{A}{P}}}
+& & \inference
+    {\jfameq{{\Gamma}{A}}{P}{P'}}
+    {\jtermeq{{{{\Gamma}{A}}{P}}{\ctxwk{P}{\unitc{A}}}}{\unitf{A}{P}}
+      {\unitft{A}{P}}
+      {\unitft{A}{P'}}}
+\end{align*}
+
+We impose the following compatibility rules for the unit type:
+
+\begin{align*}
+& \inference
+  {\jfam{{\Gamma}{A}}{P}}
+  {\jtypeeq{{{\Gamma}{A}}{P}}{\ctxext{\ctxwk{P}{\unitc{A}}}{\unitf{A}{P}}}{\unitc{{\Gamma}{A}}}}\\
+& \inference
+  {\jfam{\Gamma}{A}
+   \jfam{\Gamma}{B}}
+  {\jtypeeq{{{\Gamma}{A}}{\ctxwk{A}{B}}}{}{}}  
+\end{align*}
+
+\subsection{The empty type}
+
+\subsection{The natural numbers}
+
+\subsection{Dependent pair types}
+
+\subsection{Identity types}
+
+\subsection{Dependent function types}
+
+\subsection{$\wtypesym$-types}
+
+\subsection{Universes}
+
+\subsection{The compatibility of the type constructors with each other}
+\label{compatibility-of-type-constructors}
+
+\section{Models of type theory}
+In this section we pursue the idea of what a general model of type theory is by
+axiomatizing what you can do with them. We have the following ideas:
+\begin{enumerate}
+\item There are dependent models and sections thereof. Particular instances
+of sections: extension, weakening and substitution (and something for identity
+function?). And like the original
+extension, weakening and substitution, they're going to be compatible with each
+other. Therefore we state our theory of models as an extension of type theory
+without constructors.
+\item Type theory without constructors is a model of itself, the canonical model $\mctx$.
+\item for any (family of) model(s) $A$, there is the family model $\mfam{A}$ which
+is a family of models over $A$.
+\item The terms of $\mctx$ should be precisely contexts. Terms of $\subst{\Gamma}{\mfam{\mctx}}$
+should be families over $\Gamma$.
+\item if a model and a family of models over it are given, there is an extended model.
+If we extend $A$ by $\mfam{A}$, we get the Sierpinski model of $A$.
+\item Likewise, models can be weakened and substituted and there are identity
+functions.
+
+\end{enumerate}
+So the theory of models is going to be an extension of this theory with this
+data.
+
+\subsection{The basic ingredients of the theory of models}
+We first introduce the basic ingredients of our abstract theory of models.
+\begin{align*}
+& \inference
+  {}
+  {\jctx{\mctx}}
+  \tag{the canonical model}\\
+& \inference
+  {\jfam{\Gamma}{A}}
+  {\jfam{{\Gamma}{A}}{\mfam{A}}}
+  \tag{families}\\
+& \inference
+  {\jterm{\Gamma}{A}{x}}
+  {\jterm{{\Gamma}{A}}{\mfam{A}}{\mfam{x}}}
+  \tag{term action of $\mfam{}$}\\
+& \inference
+  {\jctx{\Gamma}}
+  {\jfam{{\Gamma}{\mfam{\Gamma}}}{\mtm{\Gamma}}}
+  \tag{terms}
+\intertext{The empty context of a model $A$ is simply a term of $A$:}
+& \inference
+  {\jfam{\Gamma}{A}}
+  {\jterm{\Gamma}{A}{\tfemp{A}}}
+  \tag{empty context}
+\intertext{Extension is going to be a morphism from $\ctxext{A}{\mfam{A}}$ to
+$A$ in context $\Gamma$:}
+& \inference
+  {\jfam{\Gamma}{A}}
+  {\jhom{\Gamma}{{A}{\mfam{A}}}{A}{\tfext{A}}}
+  \tag{extension}
+\intertext{The action of weakening on families
+is a morphism from $\ctxwk{\mfam{A}}{\mfam{A}}$ to
+$\mfam{\mfam{A}}$ in context $\ctxext{{\Gamma}{A}}{\mfam{A}}$:}
+& \inference
+  {\jfam{\Gamma}{A}}
+  {\jhom{{{\Gamma}{A}}{\mfam{A}}}{\ctxwk{\mfam{A}}{\mfam{A}}}
+    {\mfam{\mfam{A}}}
+    {\tfwk{A}^0}}
+  \tag{weakening}
+\intertext{The action of weakening on terms
+is a morphism from $\ctxwk{\mfam{A}}{\mtm{A}}$ to $\mtm{\mfam{\mfam{A}}}$ over
+$\tfwk{A}^0$ in context $\ctxext{{\Gamma}{A}}{\mfam{A}}$:}
+& \inference
+  {\jfam{\Gamma}{A}}
+  {\jfhom
+    {{{\Gamma}{A}}{\mfam{A}}}
+    {\ctxwk{\mfam{A}}{\mfam{A}}}
+    {\mfam{\mfam{A}}}
+    {\tfwk{A}^0}
+    {\ctxwk{\mfam{A}}{\mtm{A}}}
+    {\mtm{\mfam{\mfam{A}}}}
+    {\tfwk{A}^1}}
+  \tag{weakening}
+\intertext{The action of substitution by a term on families
+is going to be a morphism from $\ctxwk{\mtm{A}}{\mfam{\mfam{A}}}$ to 
+$\ctxwk{\mtm{A}}{{\mfam{A}}{\mfam{A}}}$
+in context $\ctxext{{{\Gamma}{A}}{\mfam{A}}}{\mtm{A}}$:}
+& \inference
+  {\jfam{\Gamma}{A}}
+  {\jhom
+  {{{{\Gamma}{A}}{\mfam{A}}}{\mtm{A}}}
+  {\ctxwk{\mtm{A}}{\mfam{\mfam{A}}}}
+  {\ctxwk{\mtm{A}}{{\mfam{A}}{\mfam{A}}}}
+  {\tfsubst{A}^0}}
+\intertext{Likewise, the action of substitution by a term on terms of those
+families is going to be a morphism from $\ctxwk{\mtm{A}}{\mtm{\mfam{\mfam{A}}}}$
+to $\ctxwk{\mtm{A}}{{\mfam{A}}{\mtm{A}}}$ over $\tfsubst{A}^0$ in context
+$\ctxext{{{\Gamma}{A}}{\mfam{A}}}{\mtm{A}}$:}
+& \inference
+  {\jfam{\Gamma}{A}}
+  {\jfhom
+  {{{{\Gamma}{A}}{\mfam{A}}}{\mtm{A}}}
+  {\ctxwk{\mtm{A}}{\mfam{\mfam{A}}}}
+  {\ctxwk{\mtm{A}}{{\mfam{A}}{\mfam{A}}}}
+  {\tfsubst{A}^0}
+  {\ctxwk{\mtm{A}}{\mtm{\mfam{\mfam{A}}}}}
+  {\ctxwk{\mtm{A}}{{\mfam{A}}{\mtm{A}}}}
+  {\tfsubst{A}^1}}
+\intertext{The identity functions are coded by}
+& \inference
+  {\jfam{\Gamma}{A}}
+  {\jterm
+   {{{\Gamma}{A}}{\mfam{A}}}
+   {\subst{{\idfunc[\mfam{A}]}{\tfwk{A}^0}}{\mtm{\mfam{\mfam{A}}}}}
+   {\tfid{A}}}
+\end{align*}
+We will call the terms that we introduced here the model constructors.
+
+\subsection{The compatibility rules}
+We need to do several things in this section:
+\begin{enumerate}
+\item postulate that ordinary extension, weakening, substitution and identity
+functions are compatible with the model constructors.
+\item postulate that each of the model constructors is compatible with ordinary
+extension, weakening, substitution and identity functions.
+\item postulate that the model constructors are compatible with each other,
+so that they come to model ordinary extension, weakening, substitution and
+identity functions respectively.
+\end{enumerate}
+Note that the rules for compatibility with extension are going to explain which
+model $\ctxext{A}{P}$ is by telling what the families, the terms, extension,
+weakening, substitution and identity functions are. Likewise the rules for
+compatibility with weakening and substitution do this for their respective
+cases.
+
+\subsubsection{Families of extensions}
+\begin{equation*}
+\inference
+  {\jfam{{\Gamma}{A}}{P}}
+  {\jfameq{{{\Gamma}{A}}{P}}
+    {\mfam{{A}{P}}}
+    {{\ctxwk{P}{\mfam{A}}}{\ctxwk{{P}{\mfam{A}}}{\mfam{P}}}}}
+\end{equation*}
+
+\subsubsection{Terms of families of extensions}
+\begin{equation*}
+\inference
+  {\jterm{\Gamma}{A}{x}
+   \jterm{\Gamma}{\subst{x}{P}}{u}}
+  {\jtermeq
+    {{{\Gamma}{A}}{P}}
+    {\mfam{{A}{P}}}
+    {\mfam{{x}{u}}}
+    {{\ctxwk{P}{\mfam{x}}}{\mfam{u}}}}
+\end{equation*}
+
+\subsubsection{Terms of extensions}
+\begin{equation*}
+\inference
+  {\jfam{{\Gamma}{A}}{P}}
+  {\jfameq
+    {{{{\Gamma}{A}}{P}}{\mfam{{A}{P}}}}
+    {\mtm{{A}{P}}}
+    {{\ctxwk{{P}{\mfam{A}}}{{P}{\mtm{A}}}}{\ctxwk{{P}{\mfam{A}}}{\mtm{P}}}}}
+\end{equation*}
 

### 2014-03-04 article-models-tt.tex
removed the rubbish rule, there may be more
small changes

@@
 
-\section{Models of type theory}
+\section{The type theory of models of type theory}
 In this section we pursue the idea of what a general model of type theory is by
@@
 functions.
-
 \end{enumerate}
 So the theory of models is going to be an extension of this theory with this
-data.
-
-\subsection{The basic ingredients of the theory of models}
+data. The theory we are about to describe can be seen as an elementary theory
+of the category (with families (and terms)) of categories (with families (and terms)).
+
+\subsection{The basic ingredients of the type theory of models}
 We first introduce the basic ingredients of our abstract theory of models.
@@
 & \inference
-  {\jterm{\Gamma}{A}{x}}
-  {\jterm{{\Gamma}{A}}{\mfam{A}}{\mfam{x}}}
-  \tag{term action of $\mfam{}$}\\
-& \inference
   {\jctx{\Gamma}}
@@
   \tag{terms}
+\intertext{The following two rules essentially make a start with saying that every term is functorial
+in the apropriate sense (there will be more rules contributing to this vies):}
+& \inference
+  {\jterm{{\Gamma}{A}}{P}{f}}
+  {\jterm{{{\Gamma}{A}}{\mfam{A}}}{\subst{f}{\mfam{P}}}{\mfam{f}}}\\
+& \inference
+  {\jterm{{\Gamma}{A}}{P}{f}}
+  {\jterm{{{{\Gamma}{A}}{\mfam{A}}}{\mtm{A}}}{\subst{\mfam{f}}{{f}{\mtm{P}}}}{\mtm{f}}}
 \intertext{The empty context of a model $A$ is simply a term of $A$:}
@@
 \subsubsection{Terms of families of extensions}
+
+\subsubsection{Terms of extensions}
+
+\subsubsection{Families over the empty context}
 \begin{equation*}
 \inference
-  {\jterm{\Gamma}{A}{x}
-   \jterm{\Gamma}{\subst{x}{P}}{u}}
-  {\jtermeq
-    {{{\Gamma}{A}}{P}}
-    {\mfam{{A}{P}}}
-    {\mfam{{x}{u}}}
-    {{\ctxwk{P}{\mfam{x}}}{\mfam{u}}}}
+  {\jfam{\Gamma}{A}}
+  {\jfameq{\Gamma}{\subst{\tfemp{A}}{\mfam{A}}}{A}}
 \end{equation*}
 
-\subsubsection{Terms of extensions}
+\subsubsection{Terms of the empty family}
 \begin{equation*}
 \inference
-  {\jfam{{\Gamma}{A}}{P}}
-  {\jfameq
-    {{{{\Gamma}{A}}{P}}{\mfam{{A}{P}}}}
-    {\mtm{{A}{P}}}
-    {{\ctxwk{{P}{\mfam{A}}}{{P}{\mtm{A}}}}{\ctxwk{{P}{\mfam{A}}}{\mtm{P}}}}}
+  {\jfam{\Gamma}{A}}
+  {\jfameq{{\Gamma}{A}}{\subst{\tfemp{\mfam{A}}}{\mtm{A}}}{\emptyf[A]}}
 \end{equation*}

### 2014-03-04 article-models-tt.tex
removed a false rule
typo
working on unfold
the unfold command works nicely now

@@
 \begin{defn}
-For $\jhom{\Gamma}{A}{B}{f}$ and consider a term $\jterm{{\Gamma}{B}}{Q}{g}$.
+We define the judgment
+\begin{equation*}
+\jhom{\Gamma}{A}{B}{f},
+\end{equation*}
+which is pronounced as `$f$ is a morphism from $A$ to $B$ in context $\Gamma$',
+to be the judgment
+\begin{equation*}
+\unfold{\jhom{\Gamma}{A}{B}{f}}.
+\end{equation*}
+Likewise, we define the judgment
+\begin{equation*}
+\jhomeq{\Gamma}{A}{B}{f}{f'}
+\end{equation*}
+to be the judgment
+\begin{equation*}
+\unfold{\jhomeq{\Gamma}{A}{B}{f}{f'}}.
+\end{equation*}
+\end{defn}
+
+\begin{defn}
+Let $\jhom{\Gamma}{A}{B}{f}$ and consider a term $\jterm{{\Gamma}{B}}{Q}{g}$.
 We define
 \begin{align*}
-\jfamdefn*{{\Gamma}{A}}{\jcomp{f}{Q}}{\subst{f}{\ctxwk{A}{Q}}}\\
-\jtermdefn*{{\Gamma}{A}}{\jcomp{f}{Q}}{\jcomp{f}{g}}{\subst{f}{\ctxwk{A}{g}}}.
+\jfamdefn*
+  {{\Gamma}{A}}
+  {\jcomp{A}{f}{Q}}
+  {\unfold{\jcomp{A}{f}{Q}}}\\
+\jtermdefn*
+  {{\Gamma}{A}}
+  {\jcomp{A}{f}{Q}}
+  {\jcomp{A}{f}{g}}
+  {\unfold{\jcomp{A}{f}{g}}}.
 \end{align*}
@@
 \begin{align*}
-\jfamdefn*{{{\Gamma}{A}}{\jcomp{f}{Q}}}{\jcomp{f}{R}}{\subst{f}{\ctxwk{A}{R}}}\\
-\jtermdefn*{{{\Gamma}{A}}{\jcomp{f}{Q}}}{\jcomp{f}{R}}{\jcomp{f}{h}}{\subst{f}{\ctxwk{A}{h}}}.
+\jfamdefn*{{{\Gamma}{A}}{\jcomp{A}{f}{Q}}}{\jcomp{A}{f}{R}}{\unfold{\jcomp{A}{f}{R}}}\\
+\jtermdefn*{{{\Gamma}{A}}{\jcomp{A}{f}{Q}}}{\jcomp{A}{f}{R}}{\jcomp{A}{f}{h}}{\unfold{\jcomp{A}{f}{h}}}.
 \end{align*}
@@
     {{\Gamma}{A}}
-    {\subst{\jcomp{f}{g}}{R}}
+    {\subst{\jcomp{A}{f}{g}}{R}}
     {\subst{f}{{\ctxwk{A}{g}}{\ctxwk{{A}{B}}{R}}}}}\\
@@
     {{\Gamma}{A}}
-    {\subst{\jcomp{f}{g}}{h}}
+    {\subst{\jcomp{A}{f}{g}}{h}}
     {\subst{f}{{\ctxwk{A}{g}}{\ctxwk{{A}{B}}{h}}}}}
@@
       {{\Gamma}{A}}
-    {\jcomp{{f}{g}}{R}}
-    {\jcomp{f}{{g}{R}}}}\\
+      {\jcomp{A}{{A}{f}{g}}{R}}
+      {\jcomp{A}{f}{{B}{g}{R}}}}\\
 & \inference
@@
       {{\Gamma}{A}}
-    {\jcomp{{f}{g}}{R}}
-    {\jcomp{{f}{g}}{h}}
-    {\jcomp{f}{{g}{h}}}}
+      {\jcomp{A}{{A}{f}{g}}{R}}
+      {\jcomp{A}{{A}{f}{g}}{h}}
+      {\jcomp{A}{f}{{B}{g}{h}}}}
 \end{align*}
@@
 \begin{align*}
-\subst{\jcomp{f}{g}}{R} & \jdeq \subst{{f}{\ctxwk{A}{g}}}{R}\\
+\subst{\jcomp{A}{f}{g}}{R} & \jdeq \subst{{f}{\ctxwk{A}{g}}}{R}\\
 & \jdeq \subst{{f}{\ctxwk{A}{g}}}{\subst{f}{\ctxwk{{A}{B}}{R}}}\\
@@
 The proof that 
-$\subst{\jcomp{f}{g}}{h}\jdeq\subst{f}{{\ctxwk{A}{g}}{\ctxwk{{A}{B}}{h}}}$
+$\subst{\jcomp{A}{f}{g}}{h}\jdeq\subst{f}{{\ctxwk{A}{g}}{\ctxwk{{A}{B}}{h}}}$
 is similar.
@@
 \begin{align*}
-\jcomp{{f}{g}}{R} & \jdeq \subst{\jcomp{f}{g}}{\ctxwk{A}{R}}\\
+\jcomp{A}{{A}{f}{g}}{R} & \jdeq \subst{\jcomp{A}{f}{g}}{\ctxwk{A}{R}}\\
 & \jdeq \subst{{f}{\ctxwk{A}{g}}}{\ctxwk{A}{R}}\\
@@
 & \jdeq \subst{f}{\ctxwk{A}{\subst{g}{\ctxwk{B}{R}}}}\\
-& \jdeq \subst{f}{\ctxwk{A}{\jcomp{g}{R}}}\\
-& \jdeq \jcomp{f}{{g}{R}}.
+& \jdeq \subst{f}{\ctxwk{A}{\jcomp{B}{g}{R}}}\\
+& \jdeq \jcomp{A}{f}{{B}{g}{R}}.
 \end{align*}
@@
   {\jhomeq{\Gamma}{{A}{P}}{C}
-    {\ctxwk{P}{\jcomp{f}{g}}}
-    {\jcomp{\ctxwk{P}{f}}{g}}}\\
+    {\ctxwk{P}{\jcomp{A}{f}{g}}}
+    {\jcomp{{A}{P}}{\ctxwk{P}{f}}{g}}}\\
 & \inference
@@
   {\jhomeq{\Gamma}{A}{C}
-    {\jcomp{\ctxwk{A}{y}}{g}}
+    {\jcomp{A}{\ctxwk{A}{y}}{g}}
     {\ctxwk{A}{\subst{y}{g}}}}\\
@@
   {\jhomeq{\Gamma}{A}{C}
-    {\jcomp{f}{\ctxwk{B}{z}}}
+    {\jcomp{A}{f}{\ctxwk{B}{z}}}
     {\ctxwk{A}{z}}}
@@
 \begin{align*}
-\ctxwk{P}{\jcomp{f}{g}} 
+\ctxwk{P}{\jcomp{A}{f}{g}} 
 & \jdeq \ctxwk{P}{\subst{f}{\ctxwk{A}{g}}}\\
@@
 & \jdeq \subst{\ctxwk{P}{f}}{\ctxwk{\ctxext{A}{P}}{g}}\\
-& \jdeq \jcomp{\ctxwk{P}{f}}{g}.
+& \jdeq \jcomp{{A}{P}}{\ctxwk{P}{f}}{g}.
 \end{align*}
@@
 \begin{align*}
-\jcomp{\ctxwk{A}{y}}{g}
+\jcomp{A}{\ctxwk{A}{y}}{g}
 & \jdeq \subst{\ctxwk{A}{y}}{\ctxwk{A}{g}}\\
@@
 \begin{align*}
-\jcomp{f}{\ctxwk{B}{z}} 
+\jcomp{A}{f}{\ctxwk{B}{z}} 
 & \jdeq \subst{f}{\ctxwk{A}{{B}{z}}}\\
@@
   {\jhomeq{\Gamma}{\subst{x}{P}}{\subst{x}{R}}
-    {\subst{x}{\jcomp{f}{g}}}
-    {\jcomp{\subst{x}{f}}{\subst{x}{g}}}}\\
+    {\subst{x}{\jcomp{P}{f}{g}}}
+    {\jcomp{\subst{x}{P}}{\subst{x}{f}}{\subst{x}{g}}}}\\
 & \inference
@@
   {\jtermeq{\Gamma}{C}
-    {\subst{x}{\jcomp{f}{g}}}
+    {\subst{x}{\jcomp{A}{f}{g}}}
     {\subst{{x}{f}}{g}}}
@@
 \begin{align*}
-\subst{x}{\jcomp{f}{g}}
+\subst{x}{\jcomp{A}{f}{g}}
 & \jdeq \subst{x}{{f}{\ctxwk{P}{g}}}\\
@@
 & \jdeq \subst{{x}{f}}{\ctxwk{\subst{x}{P}}{\subst{x}{g}}}\\
-& \jdeq \jcomp{\subst{x}{f}}{\subst{x}{g}}.
+& \jdeq \jcomp{\subst{x}{P}}{\subst{x}{f}}{\subst{x}{g}}.
 \end{align*}
@@
 \begin{align*}
-\subst{x}{\jcomp{f}{g}}
+\subst{x}{\jcomp{A}{f}{g}}
 & \jdeq \subst{x}{{f}{\ctxwk{A}{g}}}\\
@@
 Let $\jhom{\Gamma}{A}{B}{f}$ be a morphism from $A$ to $B$ in context $\Gamma$
-and consider $\jfam{{\Gamma}{A}}{P}$ and $\jfam{{\Gamma}{B}}{Q}$. A morphism
-from $P$ to $Q$ over $f$ in context $\Gamma$ is a term
+and consider $\jfam{{\Gamma}{A}}{P}$ and $\jfam{{\Gamma}{B}}{Q}$. We define the
+judgment
 \begin{equation*}
-\jfhom{\Gamma}{A}{B}{f}{P}{Q}{F}.
+\jfhom{\Gamma}{A}{B}{f}{P}{Q}{F},
+\end{equation*}
+which is pronounced as `$F$ is a morphism from $P$ to $Q$ over $f$ in context
+$\Gamma$, to be the judgment
+\begin{equation*}
+\unfold{\jfhom{\Gamma}{A}{B}{f}{P}{Q}{F}}.
 \end{equation*}
@@
   {\jfam{\Gamma}{A}}
-  {\jterm{{\Gamma}{A}}{\ctxwk{A}{A}}{\idfunc[A]}}
+  {\jhom{\Gamma}{A}{A}{\idfunc[A]}}
 & & \inference
     {\jfameq{\Gamma}{A}{A'}}
-    {\jtermeq{{\Gamma}{A}}{\ctxwk{A}{A}}{\idfunc[A]}{\idfunc[A']}}
+    {\jhomeq{\Gamma}{A}{A}{\idfunc[A]}{\idfunc[A']}}
 \end{align}
@@
 & \inference
-  {\jterm{{\Gamma}{B}}{\ctxwk{B}{A}}{g}}
-  {\jtermeq{{\Gamma}{B}}{\ctxwk{B}{A}}{\subst{g}{\ctxwk{B}{\idfunc[A]}}}{g}}
+  {\jhom{\Gamma}{A}{B}{f}}
+  {\jhomeq{\Gamma}{A}{B}{\jcomp{A}{f}{\idfunc[B]}}{f}}
   \label{idfunc-postcomp}
@@
    \jfam{\Gamma}{B}}
-  {\jtermeq{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{{A}{B}}{{A}{B}}}{\idfunc[\ctxwk{A}{B}]}{\ctxwk{A}{\idfunc[B]}}}
+  {\jhomeq{{\Gamma}{A}}{\ctxwk{A}{B}}{\ctxwk{A}{B}}{\ctxwk{A}{\idfunc[B]}}{\idfunc[\ctxwk{A}{B}]}}
 \end{equation}
@@
    \jfam{{\Gamma}{A}}{P}}
-  {\jtermeq{{\Gamma}{\subst{x}{P}}}{\ctxwk{\subst{x}{P}}{\subst{x}{P}}}{\idfunc[\subst{x}{P}]}{\subst{x}{\idfunc[P]}}}
+  {\jhomeq{\Gamma}{\subst{x}{P}}{\subst{x}{P}}{\subst{x}{\idfunc[P]}}{\idfunc[\subst{x}{P}]}}
 \end{equation}
@@
   {\jhomeq{\Gamma}{A}{{C}{R}}
-    {\jcomp{f}{\ctxext{g}{w}}}
-    {\ctxext{\jcomp{f}{g}}{\jcomp{f}{w}}}}
+    {\jcomp{A}{f}{\ctxext{g}{w}}}
+    {\ctxext{\jcomp{A}{f}{g}}{\jcomp{A}{f}{w}}}}
 \end{align*}
@@
 \begin{align*}
-\jcomp{f}{\ctxext{g}{w}}
+\jcomp{A}{f}{\ctxext{g}{w}}
 & \jdeq \subst{f}{\ctxwk{A}{\ctxext{g}{w}}}\\
@@
 & \jdeq \ctxext{\subst{f}{\ctxwk{A}{g}}}{\subst{f}{\ctxwk{A}{w}}}\\
-& \jdeq \ctxext{\jcomp{f}{g}}{\jcomp{f}{w}}.
+& \jdeq \ctxext{\jcomp{A}{f}{g}}{\jcomp{A}{f}{w}}.
 \end{align*}
@@
 \subsubsection{Families of extensions}
-\begin{equation*}
-\inference
-  {\jfam{{\Gamma}{A}}{P}}
-  {\jfameq{{{\Gamma}{A}}{P}}
-    {\mfam{{A}{P}}}
-    {{\ctxwk{P}{\mfam{A}}}{\ctxwk{{P}{\mfam{A}}}{\mfam{P}}}}}
-\end{equation*}
 

### 2014-03-04 article-models-tt.tex
remark on internalizing valid inference rules

@@
 
+We would also like to remark explicitly that the valid judgments of the original type
+theory become contexts, families or terms, depending on what kind of judgments
+it was. Then, valid inferences of the original type theory become valid
+judgments here. In some sense, the theory we present here is therefore a second
+order theory over the basic theory of types.
+
 \subsection{The basic ingredients of the type theory of models}

### 2014-03-04 article-models-tt.tex
code cleanup, plus fixes of small bugs

@@
 \begin{equation*}
-\jtermdefn{\Gamma}{\ctxext{A}{P}}{\ctxext{x}{u}}{\subst{u}{{x}{\idfunc[\ctxext{A}{P}]}}}.
+\jtermdefn{\Gamma}{\ctxext{A}{P}}{\tmext{A}{P}{x}{u}}{\unfold{\tmext{A}{P}{x}{u}}}.
 \end{equation*} 
@@
 
-Thus, the term $\ctxext{x}{u}$ is the pairing of $x$ and $u$. Note that because
+Thus, the term $\tmext{A}{P}{x}{u}$ is the pairing of $x$ and $u$. Note that because
 we have the judgmental equality 
@@
    \jfam{{{\Gamma}{A}}{P}}{Q}}
-  {\jfameq{\Gamma}{\subst{\ctxext{x}{u}}{Q}}{\subst{u}{{x}{Q}}}}\\
+  {\jfameq{\Gamma}{\subst{\tmext{A}{P}{x}{u}}{Q}}{\subst{u}{{x}{Q}}}}\\
 & \inference
@@
    \jterm{{{\Gamma}{A}}{P}}{Q}{g}}
-  {\jtermeq{\Gamma}{\subst{u}{{x}{Q}}}{\subst{\ctxext{x}{u}}{g}}{\subst{u}{{x}{g}}}}\\
+  {\jtermeq{\Gamma}{\subst{u}{{x}{Q}}}{\subst{\tmext{A}{P}{x}{u}}{g}}{\subst{u}{{x}{g}}}}\\
 & \inference
@@
    \jfam{{{{\Gamma}{A}}{P}}{Q}}{R}}
-  {\jfameq{{\Gamma}{\subst{u}{{x}{Q}}}}{\subst{\ctxext{x}{u}}{R}}{\subst{u}{{x}{R}}}}\\
+  {\jfameq{{\Gamma}{\subst{u}{{x}{Q}}}}{\subst{\tmext{A}{P}{x}{u}}{R}}{\subst{u}{{x}{R}}}}\\
 & \inference
@@
    \jterm{{{{\Gamma}{A}}{P}}{Q}}{R}{t}}
-  {\jtermeq{{\Gamma}{\subst{u}{{x}{Q}}}}{\subst{u}{{x}{R}}}{\subst{\ctxext{x}{u}}{t}}{\subst{u}{{x}{t}}}}
+  {\jtermeq{{\Gamma}{\subst{u}{{x}{Q}}}}{\subst{u}{{x}{R}}}{\subst{\tmext{A}{P}{x}{u}}{t}}{\subst{u}{{x}{t}}}}
 \end{align*}
@@
 \begin{align*}
-\subst{\ctxext{x}{u}}{Q} 
+\subst{\tmext{A}{P}{x}{u}}{Q} 
 & \jdeq \subst{{u}{{x}{\idfunc[\ctxext{A}{P}]}}}{Q} \tag{by definition}\\
@@
 \begin{align*}
-\jtermdefn*{\ctxext{\Gamma}{{A}{P}}}{\ctxwk{P}{{A}{A}}}{\cprojfstf{A}{P}}{\ctxwk{P}{\idfunc[A]}}\\
+\jhomdefn*{\Gamma}{{A}{P}}{A}{\cprojfstf{A}{P}}{\unfold{\cprojfstf{A}{P}}}\\
 \jtermdefn*{\ctxext{\Gamma}{{A}{P}}}{\ctxwk{P}{P}}{\cprojsndf{A}{P}}{\idfunc[P]}
@@
    \jfam{{\Gamma}{A}}{P}}
-  {\jtermeq
-    {{\Gamma}{{A}{P}}}
-    {\ctxwk{\ctxext{A}{P}}{\ctxext{A}{P}}}
+  {\jhomeq{\Gamma}{{A}{P}}{{A}{P}}
     {\idfunc[\ctxext{A}{P}]}
-    {\ctxext{\cprojfstf{A}{P}}{\cprojsndf{A}{P}}}}
+    {\tmext{\ctxwk{\ctxext{A}{P}}{A}}{\ctxwk{\ctxext{A}{P}}{P}}{\cprojfstf{A}{P}}{\cprojsndf{A}{P}}}}
 \end{equation}
@@
 \begin{align*}
-\ctxext{\cprojfstf{A}{P}}{\cprojsndf{A}{P}}
-& \jdeq \subst{\cprojsndf{A}{P}}{{\cprojfstf{A}{P}}{\idfunc[\ctxext{\ctxwk{\ctxext{A}{P}}{A}}{\ctxwk{\ctxext{A}{P}}{P}}]}}\\
-& \jdeq \subst
-          {\cprojsndf{A}{P}}
-          { {\cprojfstf{A}{P}}
-            {\idfunc[\ctxwk{\ctxext{A}{P}}{\ctxext{A}{P}}]}}\\
+\tmext{\ctxwk{\ctxext{A}{P}}{A}}{\ctxwk{\ctxext{A}{P}}{P}}{\cprojfstf{A}{P}}{\cprojsndf{A}{P}}
+& \jdeq \unfold{\tmext{\ctxwk{\ctxext{A}{P}}{A}}{\ctxwk{\ctxext{A}{P}}{P}}{\cprojfstf{A}{P}}{\cprojsndf{A}{P}}}\\
 & \jdeq \subst
@@
 
-The constructions of the terms $\ctxext{x}{u}$ and $\cprojfst{A}{P}{w}$ and
+The constructions of the terms $\tmext{A}{P}{x}{u}$ and $\cprojfst{A}{P}{w}$ and
 $\cprojsnd{A}{P}{w}$ are subject to various rules, with all of them being
@@
     {\jterm{\Gamma}{\ctxext{A}{P}}{w}}
-  {\jtermeq{\Gamma}{\ctxext{A}{P}}{\ctxext{\cprojfst{A}{P}{w}}{\cprojsnd{A}{P}{w}}}{w}}\\
+    {\jtermeq
+      {\Gamma}
+      {\ctxext{A}{P}}
+      {\tmext{A}{P}{\cprojfst{A}{P}{w}}{\cprojsnd{A}{P}{w}}}
+      {w}}\\
 & \inference
@@
    \jterm{\Gamma}{\subst{x}{P}}{u}}
-  {\jtermeq{\Gamma}{A}{\cprojfst{A}{P}{\ctxext{x}{u}}}{x}}\\
+  {\jtermeq{\Gamma}{A}{\cprojfst{A}{P}{\tmext{A}{P}{x}{u}}}{x}}\\
 & \inference
@@
    \jterm{\Gamma}{\subst{x}{P}}{u}}
-  {\jtermeq{\Gamma}{\subst{x}{P}}{\cprojsnd{A}{P}{\ctxext{x}{u}}}{u}}
+  {\jtermeq{\Gamma}{\subst{x}{P}}{\cprojsnd{A}{P}{\tmext{A}{P}{x}{u}}}{u}}
 \end{align*}
@@
 & \jdeq \subst{{w}{\idfunc[P]}}{{\subst{w}{\ctxwk{P}{\idfunc[A]}}}{\idfunc[\ctxext{A}{P}]}}\tag{by \autoref{defn-ws-4}}\\
-& \jdeq \ctxext{\cprojfst{A}{P}{w}}{\cprojsnd{A}{P}{w}}\tag{by definition}
+& \jdeq \tmext{A}{P}{\cprojfst{A}{P}{w}}{\cprojsnd{A}{P}{w}}\tag{by definition}
 \end{align*}
@@
 \begin{align*}
-\cprojfst{A}{P}{\ctxext{x}{u}}
-& \jdeq \subst{\ctxext{x}{u}}{\ctxwk{P}{\idfunc[A]}}\\
+\cprojfst{A}{P}{\tmext{A}{P}{x}{u}}
+& \jdeq \subst{\tmext{A}{P}{x}{u}}{\ctxwk{P}{\idfunc[A]}}\\
 & \jdeq \subst{u}{{x}{\ctxwk{P}{\idfunc[A]}}} \\
@@
 \begin{align*}
-\cprojsnd{A}{P}{\ctxext{x}{u}}
-& \jdeq \subst{\ctxext{x}{u}}{\idfunc[P]}\\
+\cprojsnd{A}{P}{\tmext{A}{P}{x}{u}}
+& \jdeq \subst{\tmext{A}{P}{x}{u}}{\idfunc[P]}\\
 & \jdeq \subst{u}{{x}{\idfunc[P]}}\\
@@
    \jterm{\Gamma}{\subst{x}{P}}{u}
-   \jterm{\Gamma}{\subst{\ctxext{x}{u}}{Q}}{v}}
-  {\jtermeq{\Gamma}{\ctxext{{A}{P}}{Q}}{\ctxext{x}{{u}{v}}}{\ctxext{{x}{u}}{v}}}\\
+   \jterm{\Gamma}{\subst{\tmext{A}{P}{x}{u}}{Q}}{v}}
+  {\jtermeq{\Gamma}{\ctxext{{A}{P}}{Q}}{\tmext{A}{{P}{Q}}{x}{{u}{v}}}{\tmext{{A}{P}}{Q}{{x}{u}}{v}}}\\
 & \inference
@@
 \begin{align*}
-\ctxext{x}{{u}{v}} & \jdeq \subst{\ctxext{u}{v}}{{x}{\idfunc[\ctxext{A}{{P}{Q}}]}}\\
+\tmext{A}{{P}{Q}}{x}{{u}{v}} & \jdeq \subst{\tmext{P}{Q}{u}{v}}{{x}{\idfunc[\ctxext{A}{{P}{Q}}]}}\\
 & \jdeq \subst{v}{{u}{{x}{\idfunc[\ctxext{A}{{P}{Q}}]}}}\\
 & \jdeq \subst{v}{{u}{{x}{\idfunc[\ctxext{{A}{P}}{Q}]}}}\\
-& \jdeq \subst{v}{{\ctxext{x}{u}}{\idfunc[\ctxext{{A}{P}}{Q}]}}\\
-& \jdeq \ctxext{{x}{u}}{v}.
+& \jdeq \subst{v}{{\tmext{A}{P}{x}{u}}{\idfunc[\ctxext{{A}{P}}{Q}]}}\\
+& \jdeq \tmext{{A}{P}}{Q}{{x}{u}}{v}.
 \end{align*}
@@
 w 
-& \jdeq \ctxext{\cprojfst{A}{\ctxext{P}{Q}}{w}}{\cprojsnd{A}{\ctxext{P}{Q}}{w}}\\
-& \jdeq \ctxext{\cprojfst{A}{\ctxext{P}{Q}}{w}}{{\cprojfst{P}{Q}{\cprojsnd{A}{\ctxext{P}{Q}}{w}}}{\cprojsnd{P}{Q}{\cprojsnd{A}{\ctxext{P}{Q}}{w}}}}\\
-& \jdeq \ctxext{{\cprojfst{A}{\ctxext{P}{Q}}{w}}{\cprojfst{P}{Q}{\cprojsnd{A}{\ctxext{P}{Q}}{w}}}}{\cprojsnd{P}{Q}{\cprojsnd{A}{\ctxext{P}{Q}}{w}}}
+& \jdeq \tmext{A}{{P}{Q}}{\cprojfst{A}{\ctxext{P}{Q}}{w}}{\cprojsnd{A}{\ctxext{P}{Q}}{w}}\\
+& \jdeq \tmext{A}{{P}{Q}}{\cprojfst{A}{\ctxext{P}{Q}}{w}}{{\cprojfst{P}{Q}{\cprojsnd{A}{\ctxext{P}{Q}}{w}}}{\cprojsnd{P}{Q}{\cprojsnd{A}{\ctxext{P}{Q}}{w}}}}\\
+& \jdeq \tmext{{A}{P}}{Q}{{\cprojfst{A}{\ctxext{P}{Q}}{w}}{\cprojfst{P}{Q}{\cprojsnd{A}{\ctxext{P}{Q}}{w}}}}{\cprojsnd{P}{Q}{\cprojsnd{A}{\ctxext{P}{Q}}{w}}}
 \end{align*}
@@
 \begin{align*}
-\cprojfst{\ctxext{A}{P}}{Q}{w} & \jdeq \ctxext{\cprojfst{A}{\ctxext{P}{Q}}{w}}{\cprojfst{P}{Q}{\cprojsnd{A}{\ctxext{P}{Q}}{w}}}\\ 
+\cprojfst{\ctxext{A}{P}}{Q}{w} & \jdeq \tmext{A}{P}{\cprojfst{A}{\ctxext{P}{Q}}{w}}{\cprojfst{P}{Q}{\cprojsnd{A}{\ctxext{P}{Q}}{w}}}\\ 
 \cprojsnd{\ctxext{A}{P}}{Q}{w} & \jdeq \cprojsnd{P}{Q}{\cprojsnd{A}{\ctxext{P}{Q}}{w}},
@@
 \begin{lem}
-When we weaken a term $\ctxext{y}{v}$ of $\ctxext{B}{Q}$ in context $\Gamma$ by
-a family $A$, the term that we get is $\ctxext{\ctxwk{A}{y}}{\ctxwk{A}{v}}$. More
+When we weaken a term $\tmext{B}{Q}{y}{v}$ of $\ctxext{B}{Q}$ in context $\Gamma$ by
+a family $A$, the term that we get is $\tmext{\ctxwk{A}{B}}{\ctxwk{A}{Q}}{\ctxwk{A}{y}}{\ctxwk{A}{v}}$. More
 precisely, the following inference rules are valid:
@@
     {\ctxwk{A}{\ctxext{Q}{R}}}
-    {\ctxwk{A}{\ctxext{g}{t}}}
-    {\ctxext{\ctxwk{A}{g}}{\ctxwk{A}{t}}}}
+    {\ctxwk{A}{\tmext{Q}{R}{g}{t}}}
+    {\tmext{\ctxwk{A}{Q}}{\ctxwk{A}{R}}{\ctxwk{A}{g}}{\ctxwk{A}{t}}}}
 % \\
@@
 \begin{align*}
-\ctxwk{A}{\ctxext{g}{t}}
+\ctxwk{A}{\ctxext{Q}{R}{g}{t}}
 & \jdeq \ctxwk{A}{\subst{t}{{g}{\idfunc[\ctxext{Q}{R}]}}}\\
@@
 & \jdeq \subst{\ctxwk{A}{t}}{{\ctxwk{A}{g}}{\idfunc[\ctxext{\ctxwk{A}{Q}}{\ctxwk{A}{R}}]}}\\
-& \jdeq \ctxext{\ctxwk{A}{g}}{\ctxwk{A}{t}}
+& \jdeq \tmext{\ctxwk{A}{Q}}{\ctxwk{A}{R}}{\ctxwk{A}{g}}{\ctxwk{A}{t}}
 \end{align*}
@@
 \begin{lem}
-When we substitute an extended term $\ctxext{f}{g}$ of $\ctxext{P}{Q}$ by a term
-$x$ of $A$, the term that we get is $\ctxext{\subst{x}{f}}{\subst{x}{g}}$.
+When we substitute an extended term $\tmext{P}{Q}{f}{g}$ of $\ctxext{P}{Q}$ by a term
+$x$ of $A$, the term that we get is $\tmext{\subst{x}{P}}{\subst{x}{Q}}{\subst{x}{f}}{\subst{x}{g}}$.
 More precisely, the following inference rules are valid:
@@
     {\ctxext{\subst{x}{Q}}{\subst{x}{R}}}
-    {\subst{x}{\ctxext{g}{t}}}
-    {\ctxext{\subst{x}{g}}{\subst{x}{t}}}}
+    {\subst{x}{\tmext{Q}{R}{g}{t}}}
+    {\tmext{\subst{x}{Q}}{\subst{x}{R}}{\subst{x}{g}}{\subst{x}{t}}}}
 \end{align*}
@@
 \begin{align*}
-\subst{x}{\ctxext{g}{t}}
+\subst{x}{\tmext{Q}{R}{g}{t}}
 & \jdeq \subst{x}{{t}{{g}{\idfunc[\ctxext{Q}{R}]}}}\\
@@
 & \jdeq \subst{{x}{t}}{{{x}{g}}{\idfunc[\ctxext{\subst{x}{Q}}{\subst{x}{R}}]}}\\
-& \jdeq \ctxext{\subst{x}{g}}{\subst{x}{t}}.
+& \jdeq \tmext{\subst{x}{Q}}{\subst{x}{R}}{\subst{x}{g}}{\subst{x}{t}}.
 \end{align*}
@@
   {\jhomeq{\Gamma}{A}{{C}{R}}
-    {\jcomp{A}{f}{\ctxext{g}{w}}}
-    {\ctxext{\jcomp{A}{f}{g}}{\jcomp{A}{f}{w}}}}
+    {\jcomp{A}{f}{\tmext{\ctxwk{B}{C}}{\ctxwk{B}{R}}{g}{w}}}
+    {\tmext{\ctxwk{A}{C}}{\ctxwk{A}{R}}{\jcomp{A}{f}{g}}{\jcomp{A}{f}{w}}}}
 \end{align*}
@@
 \begin{align*}
-\jcomp{A}{f}{\ctxext{g}{w}}
-& \jdeq \subst{f}{\ctxwk{A}{\ctxext{g}{w}}}\\
-& \jdeq \subst{f}{\ctxext{\ctxwk{A}{g}}{\ctxwk{A}{w}}}\\
-& \jdeq \ctxext{\subst{f}{\ctxwk{A}{g}}}{\subst{f}{\ctxwk{A}{w}}}\\
-& \jdeq \ctxext{\jcomp{A}{f}{g}}{\jcomp{A}{f}{w}}.
+\jcomp{A}{f}{\tmext{\ctxwk{B}{C}}{\ctxwk{B}{R}}{g}{w}}
+& \jdeq \subst{f}{\ctxwk{A}{\tmext{\ctxwk{B}{C}}{\ctxwk{B}{R}}{g}{w}}}\\
+& \jdeq \subst{f}{\tmext{\ctxwk{A}{{B}{C}}}{\ctxwk{A}{{B}{R}}}{\ctxwk{A}{g}}{\ctxwk{A}{w}}}\\
+& \jdeq \tmext{\ctxwk{A}{C}}{\ctxwk{A}{R}}{\subst{f}{\ctxwk{A}{g}}}{\subst{f}{\ctxwk{A}{w}}}\\
+& \jdeq \tmext{\ctxwk{A}{C}}{\ctxwk{A}{R}}{\jcomp{A}{f}{g}}{\jcomp{A}{f}{w}}.
 \end{align*}

### 2014-03-04 article-models-tt.tex
added tags to model constructor rules, they look easier to parse now
working on the compatibility rules for the constructor of model type theory

@@
 & \inference
+  {\jfam{\mctx}{P}}
+  {\jterm{\mctx}{P}{i}}
+  \tag{initiality of $\mctx$}\\
+& \inference
   {\jfam{\Gamma}{A}}
@@
   {\jterm{\Gamma}{A}{\tfemp{A}}}
-  \tag{empty context}
+  \tag{empty context}\\
+& \inference
+  {\jfam{\Gamma}{A}}
+  {\jterm{{{\Gamma}{A}}{\mfam{A}}}{\subst{\tfemp{\mfam{A}}}{\mtm{A}}}{\tft{A}}}
 \intertext{Extension is going to be a morphism from $\ctxext{A}{\mfam{A}}$ to
@@
   {\tfsubst{A}^0}}
+  \tag{substitution}
 \intertext{Likewise, the action of substitution by a term on terms of those
@@
   {\tfsubst{A}^1}}
+  \tag{substitution}
 \intertext{The identity functions are coded by}
@@
    {\tfid{A}}}
+   \tag{identity functions}
 \end{align*}
@@
 
+The easiest set of compatibility rules comes with weakening. The compatibility
+rules that deal with substitution are likely going to have to do with the
+Yoneda lemma, which we should be able to implement at some point.
+
+\subsubsection{Families of weakenings}
+\begin{equation*}
+\inference
+  { \jfam{\Gamma}{A}
+    \jfam{\Gamma}{B}}
+  { \jfameq
+    {{{\Gamma}{A}}{\ctxwk{A}{B}}}
+    {\mfam{\ctxwk{A}{B}}}
+    {\ctxwk{A}{\mfam{B}}}}.
+\end{equation*}
+
+\subsubsection{Terms of weakenings}
+\begin{equation*}
+\inference
+  { \jfam{\Gamma}{A}
+    \jfam{\Gamma}{B}}
+  { \jfameq
+    {{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\mfam{\ctxwk{A}{B}}}}
+    {\mtm{\ctxwk{A}{B}}}
+    {\ctxwk{A}{\mtm{B}}}}
+\end{equation*}
+
 \subsubsection{Families of extensions}
@@
 \end{equation*}
+
+\subsubsection{The empty family over the empty family}
+\begin{equation*}
+\inference
+  {\jfam{\Gamma}{A}}
+  {\jtermeq{\Gamma}{A}{\subst{\tfemp{A}}{\tfemp{\mfam{A}}}}{\tfemp{A}}}
+\end{equation*}
+
+\subsubsection{Extension by the empty family}
+To understand the following inference rule, recall that for $\jfam{\Gamma}{A}$
+we have the judgmental equalities
+\begin{equation*}
+\subst{\tfemp{\mfam{A}}}{\ctxwk{\ctxext{A}{\mfam{A}}}{A}}\jdeq
+\subst{\tfemp{\mfam{A}}}{\ctxwk{\mfam{A}}{{A}{A}}}\jdeq
+\ctxwk{A}{A}.
+\end{equation*}
+Therefore, we can postulate:
+\begin{equation*}
+\inference
+  {\jfam{\Gamma}{A}}
+  {\jhomeq{\Gamma}{A}{A}{\subst{\tfemp{\mfam{A}}}{\tfext{A}}}{\idfunc[A]}}
+\end{equation*}
+
+\subsubsection{Extensions of the empty context}
+To understand the following inference rule, recall that for $\jfam{\Gamma}{A}$
+we have the judgmental equalities
+\begin{align*}
+\subst{\tfemp{A}}{\ctxwk{\ctxext{A}{\mfam{A}}}{A}}
+& \jdeq \subst{\tfemp{A}}{\ctxwk{\mfam{A}}{{A}{A}}}\\
+& \jdeq \ctxwk{\subst{\tfemp{A}}{\mfam{A}}}{\subst{\tfemp{A}}{\ctxwk{A}{A}}}\\
+& \jdeq \ctxwk{A}{\subst{\tfemp{A}}{\ctxwk{A}{A}}}\\
+& \jdeq \ctxwk{A}{A}
+\end{align*}
+Therefore, we can postulate
+\begin{equation*}
+\inference
+  {\jfam{\Gamma}{A}}
+  {\jhomeq{\Gamma}{A}{A}{\subst{\tfemp{A}}{\tfext{A}}}{\idfunc[A]}}
+\end{equation*}
+
+\subsubsection{Weakening by the empty family}
+
+\subsubsection{Weakenings of the empty family}
+
+\subsubsection{Substitutions of the empty family}
+
+\subsubsection{Substitution by the term of the empty family}
+
+\subsubsection{Extensions of extensions}
+
+\subsubsection{Extensions of weakenings}
+
+\subsubsection{Weakenings by extensions}
+
+\subsubsection{Substitutions of extensions}
 

### 2014-03-05 article-models-tt.tex
added rule for families of families

@@
   \tag{terms}
-\intertext{The following two rules essentially make a start with saying that every term is functorial
+\intertext{%
+  The following two rules essentially make a start with saying that every term is functorial
   in the apropriate sense (there will be more rules contributing to this vies):}
@@
   {\jterm{{\Gamma}{A}}{P}{f}}
-  {\jterm{{{\Gamma}{A}}{\mfam{A}}}{\subst{f}{\mfam{P}}}{\mfam{f}}}\\
+  {\jhom
+    {{\Gamma}{A}}
+    {\mfam{A}}
+    {\subst{f}{\mfam{P}}}
+    {\mfam{f}}}\\
 & \inference
   {\jterm{{\Gamma}{A}}{P}{f}}
-  {\jterm{{{{\Gamma}{A}}{\mfam{A}}}{\mtm{A}}}{\subst{\mfam{f}}{{f}{\mtm{P}}}}{\mtm{f}}}
-\intertext{The empty context of a model $A$ is simply a term of $A$:}
+  {\jfhom
+    {{\Gamma}{A}}
+    {\mfam{A}}
+    {\subst{f}{\mfam{P}}}
+    {\mfam{f}}
+    {\mtm{A}}
+    {\subst{f}{\mtm{P}}}
+    {\mtm{f}}}
+\intertext{%
+  When the situation requires clarity --  for instance when extensions
+  are involved -- we will write $\mfam{f}^\Gamma$ and $\mtm{f}^\Gamma$ to make
+  what we regard as the context explicit.}
+\intertext{%
+The empty context of a model $A$ is simply a term of $A$:}
 & \inference
@@
   {\jterm{{{\Gamma}{A}}{\mfam{A}}}{\subst{\tfemp{\mfam{A}}}{\mtm{A}}}{\tft{A}}}
-\intertext{Extension is going to be a morphism from $\ctxext{A}{\mfam{A}}$ to
+\intertext{%
+  Extension is going to be a morphism from $\ctxext{A}{\mfam{A}}$ to
   $A$ in context $\Gamma$:}
@@
   \tag{extension}
-\intertext{The action of weakening on families
+\intertext{%
+  The action of weakening on families
   is a morphism from $\ctxwk{\mfam{A}}{\mfam{A}}$ to
@@
   \tag{weakening}
-\intertext{The action of weakening on terms
+\intertext{%
+  The action of weakening on terms
   is a morphism from $\ctxwk{\mfam{A}}{\mtm{A}}$ to $\mtm{\mfam{\mfam{A}}}$ over
@@
   \tag{weakening}
-\intertext{The action of substitution by a term on families
+\intertext{%
+  The action of substitution by a term on families
   is going to be a morphism from $\ctxwk{\mtm{A}}{\mfam{\mfam{A}}}$ to 
@@
   \tag{substitution}
-\intertext{Likewise, the action of substitution by a term on terms of those
+\intertext{%
+  Likewise, the action of substitution by a term on terms of those
   families is going to be a morphism from $\ctxwk{\mtm{A}}{\mtm{\mfam{\mfam{A}}}}$
@@
   \tag{substitution}
-\intertext{The identity functions are coded by}
+\intertext{%
+  The identity functions are coded by}
 & \inference
@@
 \item postulate that each of the model constructors is compatible with ordinary
-extension, weakening, substitution and identity functions.
+extension, weakening, substitution and identity functions. Actually, we want
+that every term of this type theory is compatible with those. I.e.~every term
+is a functor/morphism of models. Also the sections.
 \item postulate that the model constructors are compatible with each other,
@@
 
+\subsubsection{Families over families are families over extensions}
+\begin{equation*}
+\inference
+  {\jfam{\Gamma}{A}}
+  {\jfameq
+    {{{\Gamma}{A}}{\mfam{A}}}
+    {\mfam{\mfam{A}}}
+    {\jcomp{{A}{\mfam{A}}}{\tfext{A}}{\mfam{A}}}}
+\end{equation*}
+\emph{(Note: this rule might be a consequence of an explanation of what $\mfam{{A}{P}}$
+is in general, but at the moment I don't see how to do this)}
+
+\subsubsection{Extension acts as the identity on families of families}
+The family of which $\tfext{A}$ is a term, is unfolded as
+\begin{equation*}
+\unfold{\jhom{\Gamma}{{A}{\mfam{A}}}{A}{\tfext{A}}}
+\end{equation*}
+Therefore, there is a term
+\begin{equation*}
+\jhom
+  {{\Gamma}{A}}
+  {\mfam{\mfam{A}}}
+  {\subst{\tfext{A}}{\mfam{\ctxwk{\ctxext{A}{\mfam{A}}}{A}}}}
+  {\mfam{\tfext{A}}^{\ctxext{\Gamma}{A}}}
+\end{equation*}
+By the rule asserting that families of families are families of extensions, we
+have $\mfam{\mfam{A}}\jdeq\jcomp{}{\tfext{A}}{\mfam{A}}$. By the rule that
+families over weakenings are weakenings of families, we have the judgmental
+equalities
+\begin{equation*}
+\subst{\tfext{A}}{\mfam{\ctxwk{\ctxext{A}{\mfam{A}}}{A}}}
+\jdeq \unfold{\jcomp{{A}{P}}{\tfext{A}}{\mfam{A}}}
+\jdeq \jcomp{{A}{P}}{\tfext{A}}{\mfam{A}}
+\end{equation*}
+Therefore, we can compare the term $\mfam{\tfext{A}}^{\ctxext{\Gamma}{A}}$
+to the identity function $\idfunc[\jcomp{}{\tfext{A}}{\mfam{A}}]$. 
+The compatibility
+rule for the action of extension on families of families asserts that
+\begin{equation*}
+\inference
+  {\jfam{\Gamma}{A}}
+  {\jhomeq
+    {{\Gamma}{A}}
+    {\jcomp{}{\tfext{A}}{\mfam{A}}}
+    {\jcomp{}{\tfext{A}}{\mfam{A}}}
+    {\mfam{\tfext{A}}^{\ctxext{\Gamma}{A}}}
+    {\idfunc[\jcomp{}{\tfext{A}}{\mfam{A}}]}}
+\end{equation*}
+\emph{(Note: this rule might be a consequence of an explanation of what $\mfam{{A}{P}}$
+is in general, but at the moment I don't see how to do this)}
+
 \subsubsection{Families of extensions}
@@
 \subsubsection{Substitution by the term of the empty family}
+
+\subsubsection{Extension acts as the identity functor on families}
+\begin{equation*}
+df
+\end{equation*}
+
+\subsubsection{Extension acts as the identity functor on terms}
 

### 2014-03-05 article-models-tt.tex
big code cleanup

@@
 \begin{align*}
-\jctx*{\Gamma} & \jctxeq*{\Gamma}{\Gamma'}\\
-\jfam*{\Gamma}{A} & \jfameq*{\Gamma}{A}{B}\\
-\jtype*{\Gamma}{A} & \jtypeeq*{\Gamma}{A}{B}\\
-\jterm*{\Gamma}{A}{x} & \jtermeq*{\Gamma}{A}{x}{y}.
+\jctx*{\Gamma} 
+& \jctxeq*{\Gamma}{\Gamma'}
+  \\
+\jfam*{\Gamma}{A} 
+& \jfameq*{\Gamma}{A}{B}
+  \\
+\jtype*{\Gamma}{A} 
+& \jtypeeq*{\Gamma}{A}{B}
+  \\
+\jterm*{\Gamma}{A}{x} 
+& \jtermeq*{\Gamma}{A}{x}{y}.
 \end{align*}
@@
 & \inference
-  {\jtype{\Gamma}{A}}
-  {\jfam{\Gamma}{A}}
-& & \inference
-    {\jtypeeq{\Gamma}{A}{B}}
-    {\jfameq{\Gamma}{A}{B}}\\
-& \inference
-  {\jtype{\Gamma}{A}
-   \jfameq{\Gamma}{A}{B}}
-  {\jtype{\Gamma}{B}}
-& & \inference
-    {\jtype{\Gamma}{A}
-     \jfameq{\Gamma}{A}{B}}
-    {\jtypeeq{\Gamma}{A}{B}}
+  { \jtype{\Gamma}{A}
+    }
+  { \jfam{\Gamma}{A}
+    }
+& & \inference
+    { \jtypeeq{\Gamma}{A}{B}
+      }
+    { \jfameq{\Gamma}{A}{B}
+      }
+    \\
+& \inference
+  { \jtype{\Gamma}{A}
+    \jfameq{\Gamma}{A}{B}
+    }
+  { \jtype{\Gamma}{B}
+    }
+& & \inference
+    { \jtype{\Gamma}{A}
+      \jfameq{\Gamma}{A}{B}
+      }
+    { \jtypeeq{\Gamma}{A}{B}
+      }
 \end{align*}
@@
 & \inference
-  {\jctx{\Gamma}}
-  {\jctxeq{\Gamma}{\Gamma}} 
-& & \inference
-    {\jctxeq{\Gamma}{\Delta}}
-    {\jctxeq{\Delta}{\Gamma}} 
-& & \inference
-    {\jctxeq{\Gamma}{\Delta}
-     \jctxeq{\Delta}{\greek{E}}}
-    {\jctxeq{\Gamma}{\greek{E}}}\\
-& \inference
-  {\jfam{\Gamma}{A}}
-  {\jfameq{\Gamma}{A}{A}} 
-& & \inference
-    {\jfameq{\Gamma}{A}{B}}
-    {\jfameq{\Gamma}{B}{A}}
-& & \inference
-    {\jfameq{\Gamma}{A}{B}
-     \jfameq{\Gamma}{B}{C}}
-    {\jfameq{\Gamma}{A}{C}}\\
-& \inference
-  {\jterm{\Gamma}{A}{x}}
-  {\jtermeq{\Gamma}{A}{x}{x}}
-& & \inference
-    {\jtermeq{\Gamma}{A}{x}{y}}
-    {\jtermeq{\Gamma}{A}{y}{x}}
-& & \inference
-    {\jtermeq{\Gamma}{A}{x}{y}
-     \jtermeq{\Gamma}{A}{y}{z}}
-    {\jtermeq{\Gamma}{A}{x}{z}}
+  { \jctx{\Gamma}
+    }
+  { \jctxeq{\Gamma}{\Gamma}
+    } 
+& & \inference
+    { \jctxeq{\Gamma}{\Delta}
+      }
+    { \jctxeq{\Delta}{\Gamma}
+      } 
+& & \inference
+    { \jctxeq{\Gamma}{\Delta}
+      \jctxeq{\Delta}{\greek{E}}
+      }
+    { \jctxeq{\Gamma}{\greek{E}}
+      }
+    \\
+& \inference
+  { \jfam{\Gamma}{A}
+    }
+  { \jfameq{\Gamma}{A}{A}
+    } 
+& & \inference
+    { \jfameq{\Gamma}{A}{B}
+      }
+    { \jfameq{\Gamma}{B}{A}
+      }
+& & \inference
+    { \jfameq{\Gamma}{A}{B}
+      \jfameq{\Gamma}{B}{C}
+      }
+    { \jfameq{\Gamma}{A}{C}
+      }
+    \\
+& \inference
+  { \jterm{\Gamma}{A}{x}
+    }
+  { \jtermeq{\Gamma}{A}{x}{x}
+    }
+& & \inference
+    { \jtermeq{\Gamma}{A}{x}{y}
+      }
+    { \jtermeq{\Gamma}{A}{y}{x}
+      }
+& & \inference
+    { \jtermeq{\Gamma}{A}{x}{y}
+      \jtermeq{\Gamma}{A}{y}{z}
+      }
+    { \jtermeq{\Gamma}{A}{x}{z}
+      }
 \end{align*}
@@
 & \inference
-  {\jctxeq{\Gamma}{\Delta}
-   \jfam{\Gamma}{A}}
-  {\jfam{\Delta}{A}}
-& & \inference
-    {\jctxeq{\Gamma}{\Delta}
-     \jfameq{\Gamma}{A}{B}}
-    {\jfameq{\Delta}{A}{B}}\\
-& \inference
-  {\jctxeq{\Gamma}{\Delta}
-   \jterm{\Gamma}{A}{x}}
-  {\jterm{\Delta}{A}{x}}
-& & \inference
-    {\jctxeq{\Gamma}{\Delta}
-     \jtermeq{\Gamma}{A}{x}{y}}
-    {\jtermeq{\Delta}{A}{x}{y}}\\
-& \inference
-  {\jfameq{\Gamma}{A}{B}
-   \jterm{\Gamma}{A}{x}}
-  {\jterm{\Gamma}{B}{x}}
-& & \inference
-    {\jfameq{\Gamma}{A}{B}
-     \jtermeq{\Gamma}{A}{x}{y}}
-    {\jtermeq{\Gamma}{B}{x}{y}}
+  { \jctxeq{\Gamma}{\Delta}
+    \jfam{\Gamma}{A}
+    }
+  { \jfam{\Delta}{A}
+    }
+& & \inference
+    { \jctxeq{\Gamma}{\Delta}
+      \jfameq{\Gamma}{A}{B}
+      }
+    { \jfameq{\Delta}{A}{B}
+      }
+    \\
+& \inference
+  { \jctxeq{\Gamma}{\Delta}
+    \jterm{\Gamma}{A}{x}
+    }
+  { \jterm{\Delta}{A}{x}
+    }
+& & \inference
+    { \jctxeq{\Gamma}{\Delta}
+      \jtermeq{\Gamma}{A}{x}{y}
+      }
+    { \jtermeq{\Delta}{A}{x}{y}
+      }
+    \\
+& \inference
+  { \jfameq{\Gamma}{A}{B}
+    \jterm{\Gamma}{A}{x}
+    }
+  { \jterm{\Gamma}{B}{x}
+    }
+& & \inference
+    { \jfameq{\Gamma}{A}{B}
+      \jtermeq{\Gamma}{A}{x}{y}
+      }
+    { \jtermeq{\Gamma}{B}{x}{y}
+      }
 \end{align*}
@@
 & \inference
-  {}
-  {\jctx{\emptyc}}\\
-& \inference
-  {\jctx{\Gamma}}
-  {\jfam{\Gamma}{\emptyf[\Gamma]}}\\
-& \inference
-  {\jctx{\Gamma}}
-  {\jterm{\Gamma}{\emptyf[\Gamma]}{\emptytm[\Gamma]}}\\
-& \inference
-  {\jterm{\Gamma}{\emptyf[\Gamma]}{x}}
-  {\jtermeq{\Gamma}{\emptyf[\Gamma]}{x}{\emptytm[\Gamma]}}
+  { }
+  { \jctx{\emptyc}
+    }
+  \\
+& \inference
+  { \jctx{\Gamma}
+    }
+  { \jfam{\Gamma}{\emptyf[\Gamma]}
+    }
+  \\
+& \inference
+  { \jctx{\Gamma}
+    }
+  { \jterm{\Gamma}{\emptyf[\Gamma]}{\emptytm[\Gamma]}
+    }
+  \\
+& \inference
+  { \jterm{\Gamma}{\emptyf[\Gamma]}{x}
+    }
+  { \jtermeq
+      {\Gamma}
+      {\emptyf[\Gamma]}
+      {x}
+      {\emptytm[\Gamma]}
+    }
 \end{align}
@@
 & \inference
-  {\jctx{\Gamma}}
-  {\jfam{\emptyc}{\Gamma}} 
-& & \inference
-    {\jfam{\emptyc}{\Gamma}}
-    {\jctx{\Gamma}}\\
-& \inference
-  {\jctxeq{\Gamma}{\Delta}}
-  {\jfameq{\emptyc}{\Gamma}{\Delta}}
-& & \inference
-    {\jfameq{\emptyc}{\Gamma}{\Delta}}
-    {\jctxeq{\Gamma}{\Delta}}
+  { \jctx{\Gamma}
+    }
+  { \jfam{\emptyc}{\Gamma}
+    } 
+& & \inference
+    { \jfam{\emptyc}{\Gamma}
+      }
+    { \jctx{\Gamma}
+      }
+    \\
+& \inference
+  { \jctxeq{\Gamma}{\Delta}
+    }
+  { \jfameq{\emptyc}{\Gamma}{\Delta}
+    }
+& & \inference
+    { \jfameq{\emptyc}{\Gamma}{\Delta}
+      }
+    { \jctxeq{\Gamma}{\Delta}
+      }
 \end{align}
@@
 \inference
-{}
-{\jfameq{\emptyc}{\emptyc}{\emptyf[\emptyc]}}
+  { }
+  { \jfameq
+      {\emptyc}
+      {\emptyc}
+      {\emptyf[\emptyc]}
+    }
 \end{equation}
@@
 & \inference
-  {\jfam{\Gamma}{A}}
-  {\jctx{\ctxext{\Gamma}{A}}}
-& & \inference
-    {\jctxeq{\Gamma}{\Delta}
-     \jfameq{\Gamma}{A}{B}}
-    {\jctxeq{\ctxext{\Gamma}{A}}{\ctxext{\Delta}{B}}}\\
-& \inference
-  {\jfam{{\Gamma}{A}}{P}}
-  {\jfam{\Gamma}{\ctxext{A}{P}}}
-& & \inference
-    {\jfameq{\Gamma}{A}{B} 
-     \jfameq{{\Gamma}{A}}{P}{Q}}
-    {\jfameq{\Gamma}{\ctxext{A}{P}}{\ctxext{B}{Q}}}
+  { \jfam{\Gamma}{A}
+    }
+  { \jctx{\ctxext{\Gamma}{A}}
+    }
+& & \inference
+    { \jctxeq{\Gamma}{\Delta}
+      \jfameq{\Gamma}{A}{B}
+      }
+    { \jctxeq{\ctxext{\Gamma}{A}}{\ctxext{\Delta}{B}}
+      }
+    \\
+& \inference
+  { \jfam{{\Gamma}{A}}{P}
+    }
+  { \jfam{\Gamma}{\ctxext{A}{P}}
+    }
+& & \inference
+    { \jfameq{\Gamma}{A}{B} 
+      \jfameq{{\Gamma}{A}}{P}{Q}
+      }
+    { \jfameq{\Gamma}{\ctxext{A}{P}}{\ctxext{B}{Q}}
+      }
 \end{align}
@@
 & \inference
-  {\jctx{\Gamma}}
-  {\jctxeq{\ctxext{\emptyc}{\Gamma}}{\Gamma}}\\
-& \inference
-  {\jctx{\Gamma}}
-  {\jctxeq{\ctxext{\Gamma}{\emptyf}}{\Gamma}}\\
-& \inference
-  {\jfam{\Gamma}{A}}
-  {\jfameq{\Gamma}{\ctxext{\emptyf}{A}}{A}}
+  { \jctx{\Gamma}
+    }
+  { \jctxeq{\ctxext{\emptyc}{\Gamma}}{\Gamma}
+    }
+  \\
+& \inference
+  { \jctx{\Gamma}
+    }
+  { \jctxeq{\ctxext{\Gamma}{\emptyf}}{\Gamma}
+    }
+  \\
+& \inference
+  { \jfam{\Gamma}{A}
+    }
+  { \jfameq{\Gamma}{\ctxext{\emptyf}{A}}{A}
+    }
 \end{align}
@@
 & \inference
-  {\jfam{\Gamma}{A}
-   \jfam{{\Gamma}{A}}{P}}
-  {\jctxeq{\ctxext{{\Gamma}{A}}{P}}{\ctxext{\Gamma}{{A}{P}}}}\\
-& \inference
-  {\jfam{{\Gamma}{A}}{P}
-   \jfam{{{\Gamma}{A}}{P}}{Q}}
-  {\jfameq{\Gamma}{\ctxext{{A}{P}}{Q}}{\ctxext{A}{{P}{Q}}}}
+  { \jfam{\Gamma}{A}
+    \jfam{{\Gamma}{A}}{P}
+    }
+  { \jctxeq{\ctxext{{\Gamma}{A}}{P}}{\ctxext{\Gamma}{{A}{P}}}
+    }
+  \\
+& \inference
+  { \jfam{{\Gamma}{A}}{P}
+    \jfam{{{\Gamma}{A}}{P}}{Q}
+    }
+  { \jfameq{\Gamma}{\ctxext{{A}{P}}{Q}}{\ctxext{A}{{P}{Q}}}
+    }
 \end{align}
@@
 & \inference
-  {\jfam{\Gamma}{B}}
-  {\jfam{{\Gamma}{A}}{\ctxwk{A}{B}}}
-& & \inference
-    {\jfameq{\Gamma}{B}{B'}}
-    {\jfameq{{\Gamma}{A}}{\ctxwk{A}{B}}{\ctxwk{A'}{B'}}}\\
-& \inference
-  {\jfam{{\Gamma}{B}}{Q}}
-  {\jfam{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{Q}}}
-& & \inference
-    {\jfameq{{\Gamma}{B}}{Q}{Q'}}
-    {\jfameq{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{Q}}{\ctxwk{A'}{Q'}}}\\
-& \inference
-  {\jterm{\Gamma}{B}{y}}
-  {\jterm{{\Gamma}{A}}{\ctxwk{A}{B}}{\ctxwk{A}{y}}}
-& & \inference
-    {\jtermeq{\Gamma}{B}{y}{y'}}
-    {\jtermeq{{\Gamma}{A}}{\ctxwk{A}{B}}{\ctxwk{A}{y}}{\ctxwk{A'}{y'}}}\\
-& \inference
-  {\jterm{{\Gamma}{B}}{Q}{g}}
-  {\jterm{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{Q}}{\ctxwk{A}{g}}}
-& & \inference
-    {\jtermeq{{\Gamma}{B}}{Q}{g}{g'}}
-    {\jtermeq{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{Q}}{\ctxwk{A}{g}}{\ctxwk{A'}{g'}}}
+  { \jfam{\Gamma}{B}
+    }
+  { \jfam{{\Gamma}{A}}{\ctxwk{A}{B}}
+    }
+& & \inference
+    { \jfameq{\Gamma}{B}{B'}
+      }
+    { \jfameq{{\Gamma}{A}}{\ctxwk{A}{B}}{\ctxwk{A'}{B'}}
+      }
+    \\
+& \inference
+  { \jfam{{\Gamma}{B}}{Q}
+    }
+  { \jfam{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{Q}}
+    }
+& & \inference
+    { \jfameq{{\Gamma}{B}}{Q}{Q'}
+      }
+    { \jfameq
+        {{{\Gamma}{A}}{\ctxwk{A}{B}}}
+        {\ctxwk{A}{Q}}
+        {\ctxwk{A'}{Q'}}
+      }
+    \\
+& \inference
+  { \jterm{\Gamma}{B}{y}
+    }
+  { \jterm{{\Gamma}{A}}{\ctxwk{A}{B}}{\ctxwk{A}{y}}
+    }
+& & \inference
+    { \jtermeq{\Gamma}{B}{y}{y'}
+      }
+    { \jtermeq
+        {{\Gamma}{A}}
+        {\ctxwk{A}{B}}
+        {\ctxwk{A}{y}}
+        {\ctxwk{A'}{y'}}
+      }
+    \\
+& \inference
+  { \jterm{{\Gamma}{B}}{Q}{g}
+    }
+  { \jterm{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{Q}}{\ctxwk{A}{g}}
+    }
+& & \inference
+    { \jtermeq{{\Gamma}{B}}{Q}{g}{g'}
+      }
+    { \jtermeq
+        {{{\Gamma}{A}}{\ctxwk{A}{B}}}
+        {\ctxwk{A}{Q}}
+        {\ctxwk{A}{g}}
+        {\ctxwk{A'}{g'}}
+      }
 \end{align}
@@
 & \inference
-  {\jfam{\Gamma}{B}
-   \jtype{\ctxext{\Gamma}{B}}{Q}}
-  {\jtype{\ctxext{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{Q}}}
+  { \jfam{\Gamma}{B}
+    \jtype{\ctxext{\Gamma}{B}}{Q}
+    }
+  { \jtype{\ctxext{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{Q}}
+    }
 \end{align}
@@
 & \inference
-  {\jctx{\Gamma}}
-  {\jfameq{\Gamma}{\ctxwk{\Gamma}{\emptyc}}{\emptyf}}\\
-& \inference
-  {\jfam{\Gamma}{A}}
-  {\jfameq{{\Gamma}{A}}{\ctxwk{A}{\emptyf}}{\emptyf}}
+  { \jctx{\Gamma}
+    }
+  { \jfameq{\Gamma}{\ctxwk{\Gamma}{\emptyc}}{\emptyf}
+    }
+  \\
+& \inference
+  { \jfam{\Gamma}{A}
+    }
+  { \jfameq{{\Gamma}{A}}{\ctxwk{A}{\emptyf}}{\emptyf}
+    }
 \end{align}
@@
 & \inference
-  {\jfam{\Gamma}{B}}
-  {\jfameq{\Gamma}{\ctxwk{\emptyf}{B}}{B}}\\
-& \inference
-  {\jterm{\Gamma}{B}{y}}
-  {\jtermeq{\Gamma}{B}{\ctxwk{\emptyf}{y}}{y}}\\
-& \inference
-  {\jfam{{\Gamma}{B}}{Q}}
-  {\jfameq{{\Gamma}{B}}{\ctxwk{\emptyf}{Q}}{Q}}\\
-& \inference
-  {\jterm{{\Gamma}{B}}{Q}{g}}
-  {\jtermeq{{\Gamma}{B}}{Q}{\ctxwk{\emptyf}{g}}{g}}
+  { \jfam{\Gamma}{B}
+    }
+  { \jfameq{\Gamma}{\ctxwk{\emptyf}{B}}{B}
+    }
+  \\
+& \inference
+  { \jterm{\Gamma}{B}{y}
+    }
+  { \jtermeq{\Gamma}{B}{\ctxwk{\emptyf}{y}}{y}
+    }
+  \\
+& \inference
+  { \jfam{{\Gamma}{B}}{Q}
+    }
+  { \jfameq{{\Gamma}{B}}{\ctxwk{\emptyf}{Q}}{Q}
+    }
+  \\
+& \inference
+  { \jterm{{\Gamma}{B}}{Q}{g}
+    }
+  { \jtermeq{{\Gamma}{B}}{Q}{\ctxwk{\emptyf}{g}}{g}
+    }
 \end{align}
@@
 & \inference
-  {\jfam{{{\Gamma}{B}}{Q}}{R}}
-  {\jfameq
+  { \jfam{{{\Gamma}{B}}{Q}}{R}
+    }
+  { \jfameq
       {\ctxext{{\Gamma}{A}}{\ctxwk{A}{B}}}
       {\ctxwk{A}{\ctxext{Q}{R}}}
-    {\ctxext{\ctxwk{A}{Q}}{\ctxwk{A}{R}}}}
+      {\ctxext{\ctxwk{A}{Q}}{\ctxwk{A}{R}}}
+    }
 \end{align}
@@
 \begin{tikzcd}[column sep=huge]
-\jfam{\Gamma}{\blank} \ar{r}{C\mapsto\ctxwk{B}{C}} \ar{d}[swap]{C\mapsto\ctxwk{A}{C}} & \jfam{{\Gamma}{B}}{\blank} \ar{d}{Q\mapsto\ctxwk{A}{Q}}\\
-\jfam{{\Gamma}{A}}{\blank} \ar{r}[swap]{P\mapsto\ctxwk{{A}{B}}{P}} & \jfam{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\blank}
+\jfam{\Gamma}{\blank} 
+  \ar{r}{C\mapsto\ctxwk{B}{C}} 
+  \ar{d}[swap]{C\mapsto\ctxwk{A}{C}} 
+& \jfam{{\Gamma}{B}}{\blank} 
+  \ar{d}{Q\mapsto\ctxwk{A}{Q}}
+  \\
+\jfam{{\Gamma}{A}}{\blank} 
+  \ar{r}[swap]{P\mapsto\ctxwk{{A}{B}}{P}} 
+& \jfam{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\blank}
 \end{tikzcd}
@@
 & \inference
-  {\jfam{\Gamma}{A}
+  { \jfam{\Gamma}{A}
     \jfam{\Gamma}{B}
-   \jfam{{\Gamma}{C}}{R}}
-  {\jfameq{{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{{A}{B}}{{A}{C}}}}
+    \jfam{{\Gamma}{C}}{R}
+    }
+  { \jfameq
+      {{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{{A}{B}}{{A}{C}}}}
       {\ctxwk{A}{{B}{R}}}
-    {\ctxwk{{A}{B}}{{A}{R}}}}\label{comp-ww-f}\\
-& \inference
-  {\jfam{\Gamma}{A}
+      {\ctxwk{{A}{B}}{{A}{R}}}
+    }
+  \label{comp-ww-f}\\
+& \inference
+  { \jfam{\Gamma}{A}
     \jfam{\Gamma}{B}
-   \jterm{{\Gamma}{C}}{R}{t}}
-  {\jtermeq{{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{{A}{B}}{{A}{C}}}}
-    {\ctxwk{{A}{B}}{{A}{R}}}{\ctxwk{A}{{B}{t}}}{\ctxwk{{A}{B}}{{A}{t}}}}
+    \jterm{{\Gamma}{C}}{R}{t}
+    }
+  { \jtermeq
+      {{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{{A}{B}}{{A}{C}}}}
+      {\ctxwk{{A}{B}}{{A}{R}}}
+      {\ctxwk{A}{{B}{t}}}
+      {\ctxwk{{A}{B}}{{A}{t}}}
+    }
   \label{comp-ww-t}
@@
 & \inference
-  {\jfam{{\Gamma}{B}}{Q}}
-  {\jfameq{{{{\Gamma}{A}}{P}}{\ctxwk{P}{{A}{B}}}}{\ctxwk{\ctxext{A}{P}}{Q}}{\ctxwk{P}{{A}{Q}}}}
+  { \jfam{{\Gamma}{B}}{Q}
+    }
+  { \jfameq
+      {{{{\Gamma}{A}}{P}}{\ctxwk{P}{{A}{B}}}}
+      {\ctxwk{\ctxext{A}{P}}{Q}}
+      {\ctxwk{P}{{A}{Q}}}
+    }
   \label{comp-ew-f}\\
 & \inference
-  {\jterm{{\Gamma}{B}}{Q}{g}}
-  {\jtermeq{{{{\Gamma}{A}}{P}}{\ctxwk{P}{{A}{B}}}}{\ctxwk{P}{{A}{Q}}}{\ctxwk{\ctxext{A}{P}}{g}}{\ctxwk{P}{{A}{g}}}} 
+  { \jterm{{\Gamma}{B}}{Q}{g}
+    }
+  { \jtermeq
+      {{{{\Gamma}{A}}{P}}{\ctxwk{P}{{A}{B}}}}
+      {\ctxwk{P}{{A}{Q}}}
+      {\ctxwk{\ctxext{A}{P}}{g}}
+      {\ctxwk{P}{{A}{g}}}
+    } 
   \label{comp-ew-t}
@@
 & \inference
-  {\jfam{{\Gamma}{A}}{P}}
-  {\jfam{\Gamma}{\subst{x}{P}}}
-& & \inference
-    {\jfameq{{\Gamma}{A}}{P}{P'}}
-    {\jfameq{\Gamma}{\subst{x}{P}}{\subst{x'}{P'}}}\\
-& \inference
-  {\jfam{{{\Gamma}{A}}{P}}{Q}}
-  {\jfam{{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}}
-& & \inference
-    {\jfameq{{{\Gamma}{A}}{P}}{Q}{Q'}}
-    {\jfameq{{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}{\subst{x'}{Q'}}}\\
-& \inference
-  {\jterm{{\Gamma}{A}}{P}{f}}
-  {\jterm{\Gamma}{\subst{x}{P}}{\subst{x}{f}}}
-& & \inference
-    {\jtermeq{{\Gamma}{A}}{P}{f}{f'}}
-    {\jtermeq{\Gamma}{\subst{x}{P}}{\subst{x}{f}}{\subst{x'}{f'}}}\\
-& \inference
-  {\jterm{{{\Gamma}{A}}{P}}{Q}{g}}
-  {\jterm{{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}{\subst{x}{g}}}
-& & \inference
-    {\jtermeq{{{\Gamma}{A}}{P}}{Q}{g}{g'}}
-    {\jtermeq{{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}{\subst{x}{g}}{\subst{x'}{g'}}}
+  { \jfam{{\Gamma}{A}}{P}
+    }
+  { \jfam{\Gamma}{\subst{x}{P}}
+    }
+& & \inference
+    { \jfameq{{\Gamma}{A}}{P}{P'}
+      }
+    { \jfameq{\Gamma}{\subst{x}{P}}{\subst{x'}{P'}}
+      }
+    \\
+& \inference
+  { \jfam{{{\Gamma}{A}}{P}}{Q}
+    }
+  { \jfam{{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}
+    }
+& & \inference
+    { \jfameq{{{\Gamma}{A}}{P}}{Q}{Q'}
+      }
+    { \jfameq{{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}{\subst{x'}{Q'}}
+      }
+    \\
+& \inference
+  { \jterm{{\Gamma}{A}}{P}{f}
+    }
+  { \jterm{\Gamma}{\subst{x}{P}}{\subst{x}{f}}
+    }
+& & \inference
+    { \jtermeq{{\Gamma}{A}}{P}{f}{f'}
+      }
+    { \jtermeq{\Gamma}{\subst{x}{P}}{\subst{x}{f}}{\subst{x'}{f'}}
+      }
+    \\
+& \inference
+  { \jterm{{{\Gamma}{A}}{P}}{Q}{g}
+    }
+  { \jterm{{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}{\subst{x}{g}}
+    }
+& & \inference
+    { \jtermeq{{{\Gamma}{A}}{P}}{Q}{g}{g'}
+      }
+    { \jtermeq
+        {{\Gamma}{\subst{x}{P}}}
+        {\subst{x}{Q}}
+        {\subst{x}{g}}
+        {\subst{x'}{g'}}
+      }
 \end{align}
@@
 & \inference
-  {\jterm{\Gamma}{A}{x}
-   \jtype{\ctxext{{\Gamma}{A}}{P}}{Q}}
-  {\jtype{\ctxext{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}}
+  { \jterm{\Gamma}{A}{x}
+    \jtype{\ctxext{{\Gamma}{A}}{P}}{Q}
+    }
+  { \jtype{\ctxext{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}
+    }
 \end{align}
@@
 & \inference
-  {\jterm{\Gamma}{A}{x}}
-  {\jfameq{\Gamma}{\subst{x}{\emptyf}}{\emptyf}}\\
-& \inference
-  {\jterm{\Gamma}{A}{x}
-   \jfam{{\Gamma}{A}}{P}}
-  {\jfameq{{\Gamma}{\subst{x}{P}}}{\subst{x}{\emptyf}}{\emptyf}}
+  { \jterm{\Gamma}{A}{x}
+    }
+  { \jfameq{\Gamma}{\subst{x}{\emptyf}}{\emptyf}
+    }
+  \\
+& \inference
+  { \jterm{\Gamma}{A}{x}
+    \jfam{{\Gamma}{A}}{P}
+    }
+  { \jfameq
+      {{\Gamma}{\subst{x}{P}}}
+      {\subst{x}{\emptyf}}
+      {\emptyf}
+    }
 \end{align}
@@
 & \inference
-  {\jfam{\Gamma}{A}}
-  {\jfameq{\Gamma}{\subst{\emptytm}{A}}{A}}\\
-& \inference
-  {\jterm{\Gamma}{A}{x}}
-  {\jtermeq{\Gamma}{A}{\subst{\emptytm}{x}}{x}}\\
-& \inference
-  {\jfam{{\Gamma}{A}}{P}}
-  {\jfameq{{\Gamma}{A}}{\subst{\emptytm}{P}}{P}}\\
-& \inference
-  {\jterm{{\Gamma}{A}}{P}{f}}
-  {\jtermeq{{\Gamma}{A}}{P}{\subst{\emptytm}{f}}{f}}.
+  { \jfam{\Gamma}{A}
+    }
+  { \jfameq{\Gamma}{\subst{\emptytm}{A}}{A}
+    }
+  \\
+& \inference
+  { \jterm{\Gamma}{A}{x}
+    }
+  { \jtermeq{\Gamma}{A}{\subst{\emptytm}{x}}{x}
+    }
+  \\
+& \inference
+  { \jfam{{\Gamma}{A}}{P}
+    }
+  { \jfameq{{\Gamma}{A}}{\subst{\emptytm}{P}}{P}
+    }
+  \\
+& \inference
+  { \jterm{{\Gamma}{A}}{P}{f}
+    }
+  { \jtermeq{{\Gamma}{A}}{P}{\subst{\emptytm}{f}}{f}
+    }.
 \end{align}
@@
 & \inference
-  {\jfam{{{{\Gamma}{A}}{P}}{Q}}{R}}
-  {\jfameq{{\Gamma}{\subst{x}{P}}}{\subst{x}{\ctxext{Q}{R}}}{\ctxext{\subst{x}{Q}}{\subst{x}{R}}}}
+  { \jfam{{{{\Gamma}{A}}{P}}{Q}}{R}
+    }
+  { \jfameq
+      {{\Gamma}{\subst{x}{P}}}
+      {\subst{x}{\ctxext{Q}{R}}}
+      {\ctxext{\subst{x}{Q}}{\subst{x}{R}}}
+    }
 \end{align}
@@
 \begin{tikzcd}[column sep=huge]
-\jfam{{\Gamma}{A}}{\blank} \ar{d}[swap]{Q\mapsto\subst{x}{Q}} \ar{r}{Q\mapsto\ctxwk{P}{Q}} & \jfam{{{\Gamma}{A}}{P}}{\blank} \ar{d}{R\mapsto\subst{x}{R}}\\ 
-\jfam{\Gamma}{\blank} \ar{r}[swap]{B\mapsto\ctxwk{\subst{x}{P}}{B}} & \jfam{{\Gamma}{\subst{x}{P}}}{\blank}
+\jfam{{\Gamma}{A}}{\blank} 
+  \ar{d}[swap]{Q\mapsto\subst{x}{Q}} 
+  \ar{r}{Q\mapsto\ctxwk{P}{Q}} 
+& \jfam{{{\Gamma}{A}}{P}}{\blank} 
+    \ar{d}{R\mapsto\subst{x}{R}}
+  \\ 
+\jfam{\Gamma}{\blank} 
+  \ar{r}[swap]{B\mapsto\ctxwk{\subst{x}{P}}{B}} 
+& \jfam{{\Gamma}{\subst{x}{P}}}{\blank}
 \end{tikzcd}
@@
 & \inference
-  {\jfam{{{\Gamma}{A}}{Q}}{R}}
-  {\jfameq{{{\Gamma}{\subst{x}{P}}}{\subst{x}{\ctxwk{P}{Q}}}}{\subst{x}{\ctxwk{P}{R}}}{\ctxwk{\subst{x}{P}}{\subst{x}{R}}}}
+  { \jfam{{{\Gamma}{A}}{Q}}{R}
+    }
+  { \jfameq
+      {{{\Gamma}{\subst{x}{P}}}{\subst{x}{\ctxwk{P}{Q}}}}
+      {\subst{x}{\ctxwk{P}{R}}}
+      {\ctxwk{\subst{x}{P}}{\subst{x}{R}}}
+    }
   \label{comp-sw-f}\\
 & \inference
-  {\jterm{{{\Gamma}{A}}{Q}}{R}{h}}
-  {\jtermeq
+  { \jterm{{{\Gamma}{A}}{Q}}{R}{h}
+    }
+  { \jtermeq
       {\ctxext{{\Gamma}{\subst{x}{P}}}{\subst{x}{\ctxwk{P}{Q}}}}
@@
       {\subst{x}{\ctxwk{P}{h}}}
-      {\ctxwk{\subst{x}{P}}{\subst{x}{h}}}}
+      {\ctxwk{\subst{x}{P}}{\subst{x}{h}}}
+    }
   \label{comp-sw-t}
@@
 & \inference
-  {\jfam{{{{\Gamma}{A}}{P}}{Q}}{R}}
-  {\jfameq{{\Gamma}{\subst{x}{{f}{Q}}}}{\subst{x}{{f}{R}}}{\subst{{x}{f}}{{x}{R}}}}
+  { \jfam{{{{\Gamma}{A}}{P}}{Q}}{R}
+    }
+  { \jfameq
+      {{\Gamma}{\subst{x}{{f}{Q}}}}
+      {\subst{x}{{f}{R}}}
+      {\subst{{x}{f}}{{x}{R}}}
+    }
   \label{comp-ss-f}\\
 & \inference
-  {\jterm{{{{\Gamma}{A}}{P}}{Q}}{R}{h}}
-  {\jtermeq{{\Gamma}{\subst{x}{{f}{Q}}}}{\subst{x}{{f}{R}}}{\subst{x}{{f}{h}}}{\subst{{x}{f}}{{x}{h}}}}
+  { \jterm{{{{\Gamma}{A}}{P}}{Q}}{R}{h}
+    }
+  { \jtermeq
+      {{\Gamma}{\subst{x}{{f}{Q}}}}
+      {\subst{x}{{f}{R}}}
+      {\subst{x}{{f}{h}}}
+      {\subst{{x}{f}}{{x}{h}}}
+    }
   \label{comp-ss-t}
@@
 & \inference
-  {\jfam{{{\Gamma}{B}}{Q}}{R}}
-  {\jfameq{{{\Gamma}{A}}{\ctxwk{A}{\subst{y}{Q}}}}{\ctxwk{A}{\subst{y}{R}}}{\subst{\ctxwk{A}{y}}{\ctxwk{A}{R}}}}
+  { \jfam{{{\Gamma}{B}}{Q}}{R}
+    }
+  { \jfameq
+      {{{\Gamma}{A}}{\ctxwk{A}{\subst{y}{Q}}}}
+      {\ctxwk{A}{\subst{y}{R}}}
+      {\subst{\ctxwk{A}{y}}{\ctxwk{A}{R}}}
+    }
   \label{comp-ws-f}\\
 & \inference
-  {\jterm{{{\Gamma}{B}}{Q}}{R}{h}}
-  {\jtermeq{{{\Gamma}{A}}{\ctxwk{A}{\subst{y}{Q}}}}{\ctxwk{A}{\subst{y}{R}}}{\ctxwk{A}{\subst{y}{h}}}{\subst{\ctxwk{A}{y}}{\ctxwk{A}{h}}}}
+  { \jterm{{{\Gamma}{B}}{Q}}{R}{h}
+    }
+  { \jtermeq
+      {{{\Gamma}{A}}{\ctxwk{A}{\subst{y}{Q}}}}
+      {\ctxwk{A}{\subst{y}{R}}}
+      {\ctxwk{A}{\subst{y}{h}}}
+      {\subst{\ctxwk{A}{y}}{\ctxwk{A}{h}}}
+    }
   \label{comp-ws-t}
@@
 & \inference
-  {\jfam{\Gamma}{A}
+  { \jfam{\Gamma}{A}
     \jfam{{\Gamma}{B}}{Q}
-   \jterm{\Gamma}{A}{x}}
-  {\jfameq{{\Gamma}{B}}{\subst{x}{\ctxwk{A}{Q}}}{Q}}
+    \jterm{\Gamma}{A}{x}
+    }
+  { \jfameq{{\Gamma}{B}}{\subst{x}{\ctxwk{A}{Q}}}{Q}
+    }
   \label{defn-ws-3}\\
 & \inference
-  {\jterm{\Gamma}{A}{x}
-   \jterm{{\Gamma}{B}}{Q}{g}}
-  {\jtermeq{{\Gamma}{B}}{Q}{\subst{x}{\ctxwk{A}{g}}}{g}}
+  { \jterm{\Gamma}{A}{x}
+    \jterm{{\Gamma}{B}}{Q}{g}
+    }
+  { \jtermeq{{\Gamma}{B}}{Q}{\subst{x}{\ctxwk{A}{g}}}{g}
+    }
   \label{defn-ws-4}
@@
 \inference
-  {\jfam{\Gamma}{A}
+  { \jfam{\Gamma}{A}
     \jfam{\Gamma}{B}
     \jfam{\Gamma}{C}
-   \jhom{\Gamma}{A}{B}{f}}
-  {\jfameq{{\Gamma}{A}}
+    \jhom{\Gamma}{A}{B}{f}
+    }
+  { \jfameq
+    {{\Gamma}{A}}
     {\subst{f}{\ctxwk{A}{{B}{C}}}}
-    {\ctxwk{A}{C}}}
+    {\ctxwk{A}{C}}
+    }
 \end{equation*}
@@
 \subst{f}{\ctxwk{A}{{B}{C}}}
-& \jdeq \subst{f}{\ctxwk{{A}{B}}{{A}{C}}}\\
-& \jdeq \ctxwk{A}{C}.
+& \jdeq 
+  \subst{f}{\ctxwk{{A}{B}}{{A}{C}}}
+  \\
+& \jdeq 
+  \ctxwk{A}{C}.
 \end{align*}
@@
 \begin{align*}
-\jfamdefn*{{{\Gamma}{A}}{\jcomp{A}{f}{Q}}}{\jcomp{A}{f}{R}}{\unfold{\jcomp{A}{f}{R}}}\\
-\jtermdefn*{{{\Gamma}{A}}{\jcomp{A}{f}{Q}}}{\jcomp{A}{f}{R}}{\jcomp{A}{f}{h}}{\unfold{\jcomp{A}{f}{h}}}.
+\jfamdefn*
+  {{{\Gamma}{A}}{\jcomp{A}{f}{Q}}}
+  {\jcomp{A}{f}{R}}
+  {\unfold{\jcomp{A}{f}{R}}}
+  \\
+\jtermdefn*
+  {{{\Gamma}{A}}{\jcomp{A}{f}{Q}}}
+  {\jcomp{A}{f}{R}}
+  {\jcomp{A}{f}{h}}
+  {\unfold{\jcomp{A}{f}{h}}}.
 \end{align*}
@@
 & \inference
-  {\jhom{\Gamma}{A}{B}{f}
+  { \jhom{\Gamma}{A}{B}{f}
     \jhom{\Gamma}{B}{C}{g}
-    \jfam{{{\Gamma}{A}}{\ctxwk{A}{C}}}{R}}
-  {\jfameq
+    \jfam{{{\Gamma}{A}}{\ctxwk{A}{C}}}{R}
+    }
+  { \jfameq
       {{\Gamma}{A}}
       {\subst{\jcomp{A}{f}{g}}{R}}
-    {\subst{f}{{\ctxwk{A}{g}}{\ctxwk{{A}{B}}{R}}}}}\\
-& \inference
-  {\jhom{\Gamma}{A}{B}{f}
+      {\subst{f}{{\ctxwk{A}{g}}{\ctxwk{{A}{B}}{R}}}}
+    }
+  \\
+& \inference
+  { \jhom{\Gamma}{A}{B}{f}
     \jhom{\Gamma}{B}{C}{g}
-    \jterm{{{\Gamma}{A}}{\ctxwk{A}{C}}}{R}{h}}
-  {\jfameq
+    \jterm{{{\Gamma}{A}}{\ctxwk{A}{C}}}{R}{h}
+    }
+  { \jfameq
     {{\Gamma}{A}}
     {\subst{\jcomp{A}{f}{g}}{h}}
-    {\subst{f}{{\ctxwk{A}{g}}{\ctxwk{{A}{B}}{h}}}}}
+    {\subst{f}{{\ctxwk{A}{g}}{\ctxwk{{A}{B}}{h}}}}
+    }
 \end{align*}
@@
 & \inference
-    {\jhom{\Gamma}{A}{B}{f}
+  { \jhom{\Gamma}{A}{B}{f}
     \jhom{\Gamma}{B}{C}{g}
-      \jfam{{\Gamma}{C}}{R}}
-    {\jfameq
+    \jfam{{\Gamma}{C}}{R}
+    }
+  { \jfameq
       {{\Gamma}{A}}
       {\jcomp{A}{{A}{f}{g}}{R}}
-      {\jcomp{A}{f}{{B}{g}{R}}}}\\
-& \inference
-    {\jhom{\Gamma}{A}{B}{f}
+      {\jcomp{A}{f}{{B}{g}{R}}}
+    }
+  \\
+& \inference
+    { \jhom{\Gamma}{A}{B}{f}
       \jhom{\Gamma}{B}{C}{g}
-      \jterm{{\Gamma}{C}}{R}{h}}
-    {\jtermeq
+      \jterm{{\Gamma}{C}}{R}{h}
+      }
+    { \jtermeq
         {{\Gamma}{A}}
@@
         {\jcomp{A}{{A}{f}{g}}{h}}
-      {\jcomp{A}{f}{{B}{g}{h}}}}
+        {\jcomp{A}{f}{{B}{g}{h}}}
+      }
 \end{align*}
@@
 \begin{align*}
-\subst{\jcomp{A}{f}{g}}{R} & \jdeq \subst{{f}{\ctxwk{A}{g}}}{R}\\
-& \jdeq \subst{{f}{\ctxwk{A}{g}}}{\subst{f}{\ctxwk{{A}{B}}{R}}}\\
-& \jdeq \subst{f}{{\ctxwk{A}{g}}{\ctxwk{{A}{B}}{R}}}
+\subst{\jcomp{A}{f}{g}}{R} 
+& \jdeq 
+  \subst{{f}{\ctxwk{A}{g}}}{R}
+  \\
+& \jdeq 
+  \subst{{f}{\ctxwk{A}{g}}}{\subst{f}{\ctxwk{{A}{B}}{R}}}
+  \\
+& \jdeq 
+  \subst{f}{{\ctxwk{A}{g}}{\ctxwk{{A}{B}}{R}}}
 \end{align*}
@@
 \begin{align*}
-\jcomp{A}{{A}{f}{g}}{R} & \jdeq \subst{\jcomp{A}{f}{g}}{\ctxwk{A}{R}}\\
-& \jdeq \subst{{f}{\ctxwk{A}{g}}}{\ctxwk{A}{R}}\\
-& \jdeq \subst{f}{{\ctxwk{A}{g}}{\ctxwk{{A}{B}}{{A}{R}}}}\\
-& \jdeq \subst{f}{{\ctxwk{A}{g}}{\ctxwk{A}{{B}{R}}}}\\
-& \jdeq \subst{f}{\ctxwk{A}{\subst{g}{\ctxwk{B}{R}}}}\\
-& \jdeq \subst{f}{\ctxwk{A}{\jcomp{B}{g}{R}}}\\
-& \jdeq \jcomp{A}{f}{{B}{g}{R}}.
+\jcomp{A}{{A}{f}{g}}{R} 
+& \jdeq 
+  \subst{\jcomp{A}{f}{g}}{\ctxwk{A}{R}}
+  \\
+& \jdeq 
+  \subst{{f}{\ctxwk{A}{g}}}{\ctxwk{A}{R}}
+  \\
+& \jdeq 
+  \subst{f}{{\ctxwk{A}{g}}{\ctxwk{{A}{B}}{{A}{R}}}}
+  \\
+& \jdeq 
+  \subst{f}{{\ctxwk{A}{g}}{\ctxwk{A}{{B}{R}}}}
+  \\
+& \jdeq 
+  \subst{f}{\ctxwk{A}{\subst{g}{\ctxwk{B}{R}}}}
+  \\
+& \jdeq 
+  \subst{f}{\ctxwk{A}{\jcomp{B}{g}{R}}}
+  \\
+& \jdeq 
+  \jcomp{A}{f}{{B}{g}{R}}.
 \end{align*}
@@
 & \inference
-  {\jhom{\Gamma}{A}{B}{f}
+  { \jhom{\Gamma}{A}{B}{f}
     \jhom{\Gamma}{B}{C}{g}
-   \jfam{{\Gamma}{A}}{P}}
-  {\jhomeq{\Gamma}{{A}{P}}{C}
+    \jfam{{\Gamma}{A}}{P}
+    }
+  { \jhomeq
+      {\Gamma}
+      {{A}{P}}
+      {C}
       {\ctxwk{P}{\jcomp{A}{f}{g}}}
-    {\jcomp{{A}{P}}{\ctxwk{P}{f}}{g}}}\\
-& \inference
-  {\jterm{\Gamma}{B}{y}
-   \jhom{\Gamma}{B}{C}{g}}
-  {\jhomeq{\Gamma}{A}{C}
+      {\jcomp{{A}{P}}{\ctxwk{P}{f}}{g}}
+    }
+  \\
+& \inference
+  { \jterm{\Gamma}{B}{y}
+    \jhom{\Gamma}{B}{C}{g}
+    }
+  { \jhomeq
+      {\Gamma}
+      {A}
+      {C}
       {\jcomp{A}{\ctxwk{A}{y}}{g}}
-    {\ctxwk{A}{\subst{y}{g}}}}\\
-& \inference
-  {\jhom{\Gamma}{A}{B}{f}
-   \jterm{\Gamma}{C}{z}}
-  {\jhomeq{\Gamma}{A}{C}
+      {\ctxwk{A}{\subst{y}{g}}}
+    }
+  \\
+& \inference
+  { \jhom{\Gamma}{A}{B}{f}
+    \jterm{\Gamma}{C}{z}
+    }
+  { \jhomeq
+      {\Gamma}
+      {A}
+      {C}
       {\jcomp{A}{f}{\ctxwk{B}{z}}}
-    {\ctxwk{A}{z}}}
+      {\ctxwk{A}{z}}
+    }
 \end{align*}
@@
 \ctxwk{P}{\jcomp{A}{f}{g}} 
-& \jdeq \ctxwk{P}{\subst{f}{\ctxwk{A}{g}}}\\
-& \jdeq \subst{\ctxwk{P}{f}}{\ctxwk{P}{{A}{g}}}\\
-& \jdeq \subst{\ctxwk{P}{f}}{\ctxwk{\ctxext{A}{P}}{g}}\\
-& \jdeq \jcomp{{A}{P}}{\ctxwk{P}{f}}{g}.
+& \jdeq 
+  \ctxwk{P}{\subst{f}{\ctxwk{A}{g}}}
+  \\
+& \jdeq 
+  \subst{\ctxwk{P}{f}}{\ctxwk{P}{{A}{g}}}
+  \\
+& \jdeq 
+  \subst{\ctxwk{P}{f}}{\ctxwk{\ctxext{A}{P}}{g}}
+  \\
+& \jdeq 
+  \jcomp{{A}{P}}{\ctxwk{P}{f}}{g}.
 \end{align*}
@@
 \jcomp{A}{\ctxwk{A}{y}}{g}
-& \jdeq \subst{\ctxwk{A}{y}}{\ctxwk{A}{g}}\\
-& \jdeq \ctxwk{A}{\subst{y}{g}}.
+& \jdeq 
+  \subst{\ctxwk{A}{y}}{\ctxwk{A}{g}}
+  \\
+& \jdeq 
+  \ctxwk{A}{\subst{y}{g}}.
 \end{align*}
@@
 \jcomp{A}{f}{\ctxwk{B}{z}} 
-& \jdeq \subst{f}{\ctxwk{A}{{B}{z}}}\\
-& \jdeq \subst{f}{\ctxwk{{A}{B}}{{A}{z}}}\\
-& \jdeq \ctxwk{A}{z}.
+& \jdeq 
+  \subst{f}{\ctxwk{A}{{B}{z}}}
+  \\
+& \jdeq 
+  \subst{f}{\ctxwk{{A}{B}}{{A}{z}}}
+  \\
+& \jdeq 
+  \ctxwk{A}{z}.
 \end{align*}
@@
 & \inference
-  {\jhom{{\Gamma}{A}}{P}{Q}{f}
+  { \jhom{{\Gamma}{A}}{P}{Q}{f}
     \jhom{{\Gamma}{A}}{Q}{R}{g}
-   \jterm{\Gamma}{A}{x}}
-  {\jhomeq{\Gamma}{\subst{x}{P}}{\subst{x}{R}}
+    \jterm{\Gamma}{A}{x}
+    }
+  { \jhomeq
+      {\Gamma}
+      {\subst{x}{P}}
+      {\subst{x}{R}}
       {\subst{x}{\jcomp{P}{f}{g}}}
-    {\jcomp{\subst{x}{P}}{\subst{x}{f}}{\subst{x}{g}}}}\\
-& \inference
-  {\jhom{\Gamma}{A}{B}{f}
+      {\jcomp{\subst{x}{P}}{\subst{x}{f}}{\subst{x}{g}}}
+    }
+  \\
+& \inference
+  { \jhom{\Gamma}{A}{B}{f}
     \jhom{\Gamma}{B}{C}{g}
-   \jterm{\Gamma}{A}{x}}
-  {\jtermeq{\Gamma}{C}
+    \jterm{\Gamma}{A}{x}
+    }
+  { \jtermeq
+      {\Gamma}
+      {C}
       {\subst{x}{\jcomp{A}{f}{g}}}
-    {\subst{{x}{f}}{g}}}
+      {\subst{{x}{f}}{g}}
+    }
 \end{align*}
@@
 \begin{proof}
-Let $\jhom{{\Gamma}{A}}{P}{Q}{f}$, $\jhom{{\Gamma}{A}}{Q}{R}{g}$ and $\jterm{\Gamma}{A}{x}$.
+Let $\jhom{{\Gamma}{A}}{P}{Q}{f}$, $\jhom{{\Gamma}{A}}{Q}{R}{g}$ and 
+$\jterm{\Gamma}{A}{x}$.
 Then we have the judgmental equalities
@@
 \subst{x}{\jcomp{A}{f}{g}}
-& \jdeq \subst{x}{{f}{\ctxwk{P}{g}}}\\
-& \jdeq \subst{{x}{f}}{{x}{\ctxwk{P}{g}}}\\
-& \jdeq \subst{{x}{f}}{\ctxwk{\subst{x}{P}}{\subst{x}{g}}}\\
-& \jdeq \jcomp{\subst{x}{P}}{\subst{x}{f}}{\subst{x}{g}}.
+& \jdeq 
+  \subst{x}{{f}{\ctxwk{P}{g}}}
+  \\
+& \jdeq 
+  \subst{{x}{f}}{{x}{\ctxwk{P}{g}}}
+  \\
+& \jdeq 
+  \subst{{x}{f}}{\ctxwk{\subst{x}{P}}{\subst{x}{g}}}
+  \\
+& \jdeq 
+  \jcomp{\subst{x}{P}}{\subst{x}{f}}{\subst{x}{g}}.
 \end{align*}
@@
 \subst{x}{\jcomp{A}{f}{g}}
-& \jdeq \subst{x}{{f}{\ctxwk{A}{g}}}\\
-& \jdeq \subst{{x}{f}}{{x}{\ctxwk{A}{g}}}\\
-& \jdeq \subst{{x}{f}}{g}.
+& \jdeq 
+  \subst{x}{{f}{\ctxwk{A}{g}}}
+  \\
+& \jdeq 
+  \subst{{x}{f}}{{x}{\ctxwk{A}{g}}}
+  \\
+& \jdeq 
+  \subst{{x}{f}}{g}.
 \end{align*}
@@
 & \inference
-  {\jfam{\Gamma}{A}}
-  {\jhom{\Gamma}{A}{A}{\idfunc[A]}}
-& & \inference
-    {\jfameq{\Gamma}{A}{A'}}
-    {\jhomeq{\Gamma}{A}{A}{\idfunc[A]}{\idfunc[A']}}
+  { \jfam{\Gamma}{A}
+    }
+  { \jhom{\Gamma}{A}{A}{\idfunc[A]}
+    }
+& & \inference
+    { \jfameq{\Gamma}{A}{A'}
+      }
+    { \jhomeq{\Gamma}{A}{A}{\idfunc[A]}{\idfunc[A']}
+      }
 \end{align}
@@
 & \inference
-  {\jterm{\Gamma}{A}{x}}
-  {\jtermeq{\Gamma}{A}{\subst{x}{\idfunc[A]}}{x}}
+  { \jterm{\Gamma}{A}{x}
+    }
+  { \jtermeq{\Gamma}{A}{\subst{x}{\idfunc[A]}}{x}
+    }
   \label{idfunc-subst-defn}\\
 & \inference
-  {\jfam{{\Gamma}{A}}{P}}
-  {\jfameq{{\Gamma}{A}}{\subst{\idfunc[A]}{\ctxwk{A}{P}}}{P}}
+  { \jfam{{\Gamma}{A}}{P}
+    }
+  { \jfameq{{\Gamma}{A}}{\subst{\idfunc[A]}{\ctxwk{A}{P}}}{P}
+    }
   \label{idfunc-wk-defn}\\
 & \inference
-  {\jfam{{\Gamma}{A}}{P}}
-  {\jfameq{{\Gamma}{A}}{\subst{\idfunc[A]}{\ctxwk{{A}{A}}{P}}}{P}}
+  { \jfam{{\Gamma}{A}}{P}
+    }
+  { \jfameq
+      {{\Gamma}{A}}
+      {\subst{\idfunc[A]}{\ctxwk{{A}{A}}{P}}}
+      {P}
+    }
   \label{idfunc-wk-defn2}\\
 & \inference
-  {\jterm{{\Gamma}{A}}{P}{f}}
-  {\jtermeq{{\Gamma}{A}}{P}{\subst{\idfunc[A]}{\ctxwk{A}{f}}}{f}}
+  { \jterm{{\Gamma}{A}}{P}{f}
+    }
+  { \jtermeq
+      {{\Gamma}{A}}
+      {P}
+      {\subst{\idfunc[A]}{\ctxwk{A}{f}}}
+      {f}
+    }
   \label{idfunc-precomp}\\
 & \inference
-  {\jterm{{\Gamma}{A}}{P}{f}}
-  {\jtermeq{{\Gamma}{A}}{P}{\subst{\idfunc[A]}{\ctxwk{{A}{A}}{f}}}{f}}
+  { \jterm{{\Gamma}{A}}{P}{f}
+    }
+  { \jtermeq
+      {{\Gamma}{A}}
+      {P}
+      {\subst{\idfunc[A]}{\ctxwk{{A}{A}}{f}}}
+      {f}
+    }
   \label{idfunc-precomp}\\
 & \inference
-  {\jhom{\Gamma}{A}{B}{f}}
-  {\jhomeq{\Gamma}{A}{B}{\jcomp{A}{f}{\idfunc[B]}}{f}}
+  { \jhom{\Gamma}{A}{B}{f}
+    }
+  { \jhomeq{\Gamma}{A}{B}{\jcomp{A}{f}{\idfunc[B]}}{f}
+    }
   \label{idfunc-postcomp}
@@
 \inference
-  {\jfam{\Gamma}{A}
-   \jfam{\Gamma}{B}}
-  {\jhomeq{{\Gamma}{A}}{\ctxwk{A}{B}}{\ctxwk{A}{B}}{\ctxwk{A}{\idfunc[B]}}{\idfunc[\ctxwk{A}{B}]}}
+  { \jfam{\Gamma}{A}
+    \jfam{\Gamma}{B}
+    }
+  { \jhomeq
+      {{\Gamma}{A}}
+      {\ctxwk{A}{B}}
+      {\ctxwk{A}{B}}
+      {\ctxwk{A}{\idfunc[B]}}
+      {\idfunc[\ctxwk{A}{B}]}
+    }
 \end{equation}
@@
 \inference
-  {\jterm{\Gamma}{A}{x}
-   \jfam{{\Gamma}{A}}{P}}
-  {\jhomeq{\Gamma}{\subst{x}{P}}{\subst{x}{P}}{\subst{x}{\idfunc[P]}}{\idfunc[\subst{x}{P}]}}
+  { \jterm{\Gamma}{A}{x}
+    \jfam{{\Gamma}{A}}{P}
+    }
+  { \jhomeq
+      {\Gamma}
+      {\subst{x}{P}}
+      {\subst{x}{P}}
+      {\subst{x}{\idfunc[P]}}
+      {\idfunc[\subst{x}{P}]}
+    }
 \end{equation}
@@
 \begin{equation*}
-\jtermdefn{\Gamma}{\ctxext{A}{P}}{\tmext{A}{P}{x}{u}}{\unfold{\tmext{A}{P}{x}{u}}}.
+\jtermdefn
+  {\Gamma}
+  {\ctxext{A}{P}}
+  {\tmext{A}{P}{x}{u}}
+  {\unfold{\tmext{A}{P}{x}{u}}}.
 \end{equation*} 
@@
 & \inference
-  {\jterm{\Gamma}{A}{x}
+  { \jterm{\Gamma}{A}{x}
     \jterm{\Gamma}{\subst{x}{P}}{u}
-   \jfam{{{\Gamma}{A}}{P}}{Q}}
-  {\jfameq{\Gamma}{\subst{\tmext{A}{P}{x}{u}}{Q}}{\subst{u}{{x}{Q}}}}\\
-& \inference
-  {\jterm{\Gamma}{A}{x}
+    \jfam{{{\Gamma}{A}}{P}}{Q}
+    }
+  { \jfameq
+      {\Gamma}
+      {\subst{\tmext{A}{P}{x}{u}}{Q}}
+      {\subst{u}{{x}{Q}}}
+    }
+  \\
+& \inference
+  { \jterm{\Gamma}{A}{x}
     \jterm{\Gamma}{\subst{x}{P}}{u}
-   \jterm{{{\Gamma}{A}}{P}}{Q}{g}}
-  {\jtermeq{\Gamma}{\subst{u}{{x}{Q}}}{\subst{\tmext{A}{P}{x}{u}}{g}}{\subst{u}{{x}{g}}}}\\
-& \inference
-  {\jterm{\Gamma}{A}{x}
+    \jterm{{{\Gamma}{A}}{P}}{Q}{g}
+    }
+  { \jtermeq
+      {\Gamma}
+      {\subst{u}{{x}{Q}}}
+      {\subst{\tmext{A}{P}{x}{u}}{g}}
+      {\subst{u}{{x}{g}}}
+    }
+  \\
+& \inference
+  { \jterm{\Gamma}{A}{x}
     \jterm{\Gamma}{\subst{x}{P}}{u}
-   \jfam{{{{\Gamma}{A}}{P}}{Q}}{R}}
-  {\jfameq{{\Gamma}{\subst{u}{{x}{Q}}}}{\subst{\tmext{A}{P}{x}{u}}{R}}{\subst{u}{{x}{R}}}}\\
-& \inference
-  {\jterm{\Gamma}{A}{x}
+    \jfam{{{{\Gamma}{A}}{P}}{Q}}{R}
+    }
+  { \jfameq
+      {{\Gamma}{\subst{u}{{x}{Q}}}}
+      {\subst{\tmext{A}{P}{x}{u}}{R}}
+      {\subst{u}{{x}{R}}}
+    }
+  \\
+& \inference
+  { \jterm{\Gamma}{A}{x}
     \jterm{\Gamma}{\subst{x}{P}}{u}
-   \jterm{{{{\Gamma}{A}}{P}}{Q}}{R}{t}}
-  {\jtermeq{{\Gamma}{\subst{u}{{x}{Q}}}}{\subst{u}{{x}{R}}}{\subst{\tmext{A}{P}{x}{u}}{t}}{\subst{u}{{x}{t}}}}
+    \jterm{{{{\Gamma}{A}}{P}}{Q}}{R}{t}
+    }
+  { \jtermeq
+      {{\Gamma}{\subst{u}{{x}{Q}}}}
+      {\subst{u}{{x}{R}}}
+      {\subst{\tmext{A}{P}{x}{u}}{t}}
+      {\subst{u}{{x}{t}}}
+    }
 \end{align*}
@@
 \begin{align*}
-\subst{\tmext{A}{P}{x}{u}}{Q} 
-& \jdeq \subst{{u}{{x}{\idfunc[\ctxext{A}{P}]}}}{Q} \tag{by definition}\\
-& \jdeq \subst{{u}{{x}{\idfunc[\ctxext{A}{P}]}}}{{x}{\ctxwk{A}{Q}}}\tag{by \autoref{defn-ws-3}}\\
-& \jdeq \subst{{u}{{x}{\idfunc[\ctxext{A}{P}]}}}{{u}{\ctxwk{\subst{x}{P}}{\subst{x}{\ctxwk{A}{Q}}}}}\tag{by \autoref{defn-ws-3}}\\
-& \jdeq \subst{{u}{{x}{\idfunc[\ctxext{A}{P}]}}}{{u}{{x}{\ctxwk{P}{{A}{Q}}}}}\tag{by \autoref{comp-sw-f}}\\
-& \jdeq \subst{u}{{{x}{\idfunc[\ctxext{A}{P}]}}{{x}{\ctxwk{{P}{{A}{Q}}}}}}\tag{by \autoref{comp-ss-f}}\\
-& \jdeq \subst{u}{{x}{{\idfunc[\ctxext{A}{P}]}{\ctxwk{P}{{A}{Q}}}}}\tag{by \autoref{comp-ss-f}}\\
-& \jdeq \subst{u}{{x}{{\idfunc[\ctxext{A}{P}]}{\ctxwk{\ctxext{A}{P}}{Q}}}}\tag{by \autoref{comp-ew-f}}\\
-& \jdeq \subst{u}{{x}{Q}}\tag{by \autoref{idfunc-wk-defn}}
+\subst
+  {\tmext{A}{P}{x}{u}}
+  {Q} 
+& \jdeq 
+  \subst
+    {{u}{{x}{\idfunc[\ctxext{A}{P}]}}}
+    {Q}
+  \tag{by definition}\\
+& \jdeq 
+  \subst
+    {{u}{{x}{\idfunc[\ctxext{A}{P}]}}}
+    {{x}{\ctxwk{A}{Q}}}
+  \tag{by \autoref{defn-ws-3}}\\
+& \jdeq 
+  \subst
+    {{u}{{x}{\idfunc[\ctxext{A}{P}]}}}
+    {{u}{\ctxwk{\subst{x}{P}}{\subst{x}{\ctxwk{A}{Q}}}}}
+  \tag{by \autoref{defn-ws-3}}\\
+& \jdeq 
+  \subst
+    {{u}{{x}{\idfunc[\ctxext{A}{P}]}}}
+    {{u}{{x}{\ctxwk{P}{{A}{Q}}}}}
+  \tag{by \autoref{comp-sw-f}}\\
+& \jdeq 
+  \subst
+    {u}
+    {{{x}{\idfunc[\ctxext{A}{P}]}}{{x}{\ctxwk{{P}{{A}{Q}}}}}}
+  \tag{by \autoref{comp-ss-f}}\\
+& \jdeq 
+  \subst
+    {u}
+    {{x}{{\idfunc[\ctxext{A}{P}]}{\ctxwk{P}{{A}{Q}}}}}
+  \tag{by \autoref{comp-ss-f}}\\
+& \jdeq 
+  \subst
+    {u}
+    {{x}{{\idfunc[\ctxext{A}{P}]}{\ctxwk{\ctxext{A}{P}}{Q}}}}
+  \tag{by \autoref{comp-ew-f}}\\
+& \jdeq 
+  \subst
+    {u}
+    {{x}{Q}}
+  \tag{by \autoref{idfunc-wk-defn}}
 \end{align*}
@@
 \begin{align*}
-\ctxwk{\ctxext{A}{P}}{\ctxext{A}{P}} & \jdeq \ctxwk{P}{{A}{\ctxext{A}{P}}}\\
-& \jdeq \ctxext{\ctxwk{P}{{A}{A}}}{\ctxwk{P}{{A}{P}}}
+\ctxwk{\ctxext{A}{P}}{\ctxext{A}{P}} 
+& \jdeq 
+  \ctxwk{P}{{A}{\ctxext{A}{P}}}
+  \\
+& \jdeq 
+  \ctxext{\ctxwk{P}{{A}{A}}}{\ctxwk{P}{{A}{P}}}
 \end{align*}
@@
 \begin{align*}
-\subst{\ctxwk{P}{\idfunc[A]}}{\ctxwk{P}{{A}{P}}} & \jdeq \ctxwk{P}{\subst{\idfunc[A]}{\ctxwk{A}{P}}}\\
-& \jdeq \ctxwk{P}{P},
+\subst{\ctxwk{P}{\idfunc[A]}}{\ctxwk{P}{{A}{P}}} 
+& \jdeq 
+  \ctxwk{P}{\subst{\idfunc[A]}{\ctxwk{A}{P}}}
+  \\
+& \jdeq 
+  \ctxwk{P}{P},
 \end{align*}
@@
 \begin{align*}
-\jhomdefn*{\Gamma}{{A}{P}}{A}{\cprojfstf{A}{P}}{\unfold{\cprojfstf{A}{P}}}\\
-\jtermdefn*{\ctxext{\Gamma}{{A}{P}}}{\ctxwk{P}{P}}{\cprojsndf{A}{P}}{\idfunc[P]}
+\jhomdefn*
+  {\Gamma}
+  {{A}{P}}
+  {A}
+  {\cprojfstf{A}{P}}
+  {\unfold{\cprojfstf{A}{P}}}
+  \\
+\jtermdefn*
+  {\ctxext{\Gamma}{{A}{P}}}
+  {\ctxwk{P}{P}}
+  {\cprojsndf{A}{P}}
+  {\unfold{\cprojsndf{A}{P}}}
 \end{align*}
@@
 \inference
-  {\jfam{\Gamma}{A}
-   \jfam{{\Gamma}{A}}{P}}
-  {\jhomeq{\Gamma}{{A}{P}}{{A}{P}}
+  { \jfam{\Gamma}{A}
+    \jfam{{\Gamma}{A}}{P}
+    }
+  { \jhomeq
+      {\Gamma}
+      {{A}{P}}{{A}{P}}
       {\idfunc[\ctxext{A}{P}]}
-    {\tmext{\ctxwk{\ctxext{A}{P}}{A}}{\ctxwk{\ctxext{A}{P}}{P}}{\cprojfstf{A}{P}}{\cprojsndf{A}{P}}}}
+      { \tmext
+          {\ctxwk{\ctxext{A}{P}}{A}}
+          {\ctxwk{\ctxext{A}{P}}{P}}
+          {\cprojfstf{A}{P}}
+          {\cprojsndf{A}{P}}
+        }
+    }
 \end{equation}
@@
 \begin{align*}
-\tmext{\ctxwk{\ctxext{A}{P}}{A}}{\ctxwk{\ctxext{A}{P}}{P}}{\cprojfstf{A}{P}}{\cprojsndf{A}{P}}
-& \jdeq \unfold{\tmext{\ctxwk{\ctxext{A}{P}}{A}}{\ctxwk{\ctxext{A}{P}}{P}}{\cprojfstf{A}{P}}{\cprojsndf{A}{P}}}\\
-& \jdeq \subst
-          {\idfunc[P]}
+\tmext
+  {\ctxwk{\ctxext{A}{P}}{A}}
+  {\ctxwk{\ctxext{A}{P}}{P}}
+  {\cprojfstf{A}{P}}
+  {\cprojsndf{A}{P}}
+& \jdeq 
+  \unfold
+  { \tmext
+      {\ctxwk{\ctxext{A}{P}}{A}}
+      {\ctxwk{\ctxext{A}{P}}{P}}
+      {\cprojfstf{A}{P}}
+      {\cprojsndf{A}{P}}
+    }
+  \\
+& \jdeq 
+  \subst
+    { \idfunc[P]
+      }
     { {\ctxwk{P}{\idfunc[A]}}
-            {\idfunc[\ctxwk{\ctxext{A}{P}}{\ctxext{A}{P}}]}}\\
-& \jdeq \subst
-          {\idfunc[P]}
+      {\idfunc[\ctxwk{\ctxext{A}{P}}{\ctxext{A}{P}}]}
+      }
+  \\
+& \jdeq 
+  \subst
+    { \idfunc[P]
+      }
     { {\ctxwk{P}{\idfunc[A]}}
-            {\ctxwk{\ctxext{A}{P}}{\idfunc[\ctxext{A}{P}]}}}\\
-& \jdeq \subst
-          {\idfunc[P]}
+      {\ctxwk{\ctxext{A}{P}}{\idfunc[\ctxext{A}{P}]}}
+      }
+  \\
+& \jdeq 
+  \subst
+    { \idfunc[P]
+      }
     { {\ctxwk{P}{\idfunc[A]}}
-            {\ctxwk{P}{{A}{\idfunc[\ctxext{A}{P}]}}}}\\
-& \jdeq \subst
-          {\idfunc[P]}
+      {\ctxwk{P}{{A}{\idfunc[\ctxext{A}{P}]}}}
+      }
+  \\
+& \jdeq 
+  \subst
+    { \idfunc[P]
+      }
     {\ctxwk
       {P}
-            {\subst
+      { \subst
         {\idfunc[A]}
-              {\ctxwk{A}{\idfunc[\ctxext{A}{P}]}}}}\\
-& \jdeq \subst
-          {\idfunc[A]}
-          {\ctxwk{A}{\idfunc[\ctxext{A}{P}]}}\\
-& \jdeq \idfunc[\ctxext{A}{P}].
+        {\ctxwk{A}{\idfunc[\ctxext{A}{P}]}}
+        }
+      }
+  \\
+& \jdeq 
+  \subst
+    { \idfunc[A]
+      }
+    { \ctxwk{A}{\idfunc[\ctxext{A}{P}]}
+      }
+  \\
+& \jdeq 
+  \idfunc[\ctxext{A}{P}].
 \end{align*}
@@
 \inference
-  {\jfam{{\Gamma}{A}}{P}
-   \jfam{{\Gamma}{A}}{Q}}
-  {\jfameq{{\Gamma}{{A}{P}}}{\subst{\cprojfstf{A}{P}}{\ctxwk{\ctxext{A}{P}}{Q}}}{\ctxwk{P}{Q}}}
+  { \jfam{{\Gamma}{A}}{P}
+    \jfam{{\Gamma}{A}}{Q}
+    }
+  { \jfameq
+      {{\Gamma}{{A}{P}}}
+      {\subst{\cprojfstf{A}{P}}{\ctxwk{\ctxext{A}{P}}{Q}}}
+      {\ctxwk{P}{Q}}
+    }
 \end{equation*}
@@
 \subst{\cprojfstf{A}{P}}{\ctxwk{\ctxext{A}{P}}{Q}} 
-& \jdeq \subst{\ctxwk{P}{\idfunc[A]}}{\ctxwk{\ctxext{A}{P}}{Q}} \tag{by definition}\\
-& \jdeq \subst{\ctxwk{P}{\idfunc[A]}}{\ctxwk{P}{{A}{Q}}} \tag{by \autoref{comp-ww-f}}\\
-& \jdeq \ctxwk{P}{\subst{\idfunc[A]}{\ctxwk{A}{Q}}} \tag{by \autoref{comp-ws-f}}\\
-& \jdeq \ctxwk{P}{Q} \tag{by \autoref{idfunc-wk-defn}}
+& \jdeq 
+  \subst{\ctxwk{P}{\idfunc[A]}}{\ctxwk{\ctxext{A}{P}}{Q}} 
+  \tag{by definition}\\
+& \jdeq 
+  \subst{\ctxwk{P}{\idfunc[A]}}{\ctxwk{P}{{A}{Q}}} 
+  \tag{by \autoref{comp-ww-f}}\\
+& \jdeq 
+  \ctxwk{P}{\subst{\idfunc[A]}{\ctxwk{A}{Q}}} 
+  \tag{by \autoref{comp-ws-f}}\\
+& \jdeq 
+  \ctxwk{P}{Q} 
+  \tag{by \autoref{idfunc-wk-defn}}
 \end{align*}
@@
 & \inference
-    {\jterm{\Gamma}{\ctxext{A}{P}}{w}}
-    {\jtermeq
+  { \jterm{\Gamma}{\ctxext{A}{P}}{w}
+    }
+  { \jtermeq
       {\Gamma}
@@
       {\tmext{A}{P}{\cprojfst{A}{P}{w}}{\cprojsnd{A}{P}{w}}}
-      {w}}\\
-& \inference
-  {\jterm{\Gamma}{A}{x}
-   \jterm{\Gamma}{\subst{x}{P}}{u}}
-  {\jtermeq{\Gamma}{A}{\cprojfst{A}{P}{\tmext{A}{P}{x}{u}}}{x}}\\
-& \inference
-  {\jterm{\Gamma}{A}{x}
-   \jterm{\Gamma}{\subst{x}{P}}{u}}
-  {\jtermeq{\Gamma}{\subst{x}{P}}{\cprojsnd{A}{P}{\tmext{A}{P}{x}{u}}}{u}}
+      {w}
+    }
+  \\
+& \inference
+  { \jterm{\Gamma}{A}{x}
+    \jterm{\Gamma}{\subst{x}{P}}{u}
+    }
+  { \jtermeq
+      {\Gamma}
+      {A}
+      {\cprojfst{A}{P}{\tmext{A}{P}{x}{u}}}
+      {x}
+    }
+  \\
+& \inference
+  { \jterm{\Gamma}{A}{x}
+    \jterm{\Gamma}{\subst{x}{P}}{u}
+    }
+  { \jtermeq
+      {\Gamma}
+      {\subst{x}{P}}
+      {\cprojsnd{A}{P}{\tmext{A}{P}{x}{u}}}
+      {u}
+    }
 \end{align*}
@@
 \begin{align*}
-w & \jdeq \subst{w}{\idfunc[\ctxext{A}{P}]} \tag{by \autoref{idfunc-subst-defn}}\\
-& \jdeq \subst{w}{\subst{\idfunc[P]}{{\ctxwk{P}{\idfunc[A]}}{\idfunc[\ctxwk{\ctxext{A}{P}}{\ctxext{A}{P}}]}}}
+w 
+& \jdeq 
+  \subst{w}{\idfunc[\ctxext{A}{P}]} 
+  \tag{by \autoref{idfunc-subst-defn}}\\
+& \jdeq 
+  \subst
+    { w}
+    { { \idfunc[P]}
+      { { \ctxwk{P}{\idfunc[A]}}
+        { \idfunc[\ctxwk{\ctxext{A}{P}}{\ctxext{A}{P}}]}
+        }
+      }
   \tag{by \autoref{idfunc-ext-comp}}\\
-& \jdeq \subst{{w}{\idfunc[P]}}{\subst{w}{{\ctxwk{P}{\idfunc[A]}}{\idfunc[\ctxwk{\ctxext{A}{P}}{\ctxext{A}{P}}]}}}\tag{by \autoref{comp-ss-t}}\\
-& \jdeq \subst{{w}{\idfunc[P]}}{{\subst{w}{\ctxwk{P}{\idfunc[A]}}}{\subst{w}{\idfunc[\ctxwk{\ctxext{A}{P}}{\ctxext{A}{P}}]}}}\tag{by \autoref{comp-ss-t}}\\
-& \jdeq \subst{{w}{\idfunc[P]}}{{\subst{w}{\ctxwk{P}{\idfunc[A]}}}{\subst{w}{\ctxwk{\ctxext{A}{P}}{\idfunc[\ctxext{A}{P}]}}}}\tag{by \autoref{idfunc-wk-comp}}\\
-& \jdeq \subst{{w}{\idfunc[P]}}{{\subst{w}{\ctxwk{P}{\idfunc[A]}}}{\idfunc[\ctxext{A}{P}]}}\tag{by \autoref{defn-ws-4}}\\
-& \jdeq \tmext{A}{P}{\cprojfst{A}{P}{w}}{\cprojsnd{A}{P}{w}}\tag{by definition}
+& \jdeq 
+  \subst
+    { {w}
+      {\idfunc[P]}
+      }
+    { {w}
+      { { \ctxwk{P}{\idfunc[A]}
+          }
+        { \idfunc[\ctxwk{\ctxext{A}{P}}{\ctxext{A}{P}}]
+          }
+        }
+      }
+  \tag{by \autoref{comp-ss-t}}\\
+& \jdeq 
+  \subst
+    { {w}
+      {\idfunc[P]}
+      }
+    { { {w}
+        {\ctxwk{P}{\idfunc[A]}}
+        }
+      { {w}
+        {\idfunc[\ctxwk{\ctxext{A}{P}}{\ctxext{A}{P}}]}
+        }
+      }
+  \tag{by \autoref{comp-ss-t}}\\
+& \jdeq 
+  \subst
+    { {w}
+      {\idfunc[P]}
+      }
+    { { {w}
+        {\ctxwk{P}{\idfunc[A]}}
+        }
+      { {w}
+        {\ctxwk{\ctxext{A}{P}}{\idfunc[\ctxext{A}{P}]}}
+        }
+      }
+  \tag{by \autoref{idfunc-wk-comp}}\\
+& \jdeq 
+  \subst
+    { {w}
+      {\idfunc[P]}
+      }
+    { { {w}
+        {\ctxwk{P}{\idfunc[A]}}
+        }
+      { \idfunc[\ctxext{A}{P}]
+        }
+      }
+  \tag{by \autoref{defn-ws-4}}\\
+& \jdeq 
+  \tmext{A}{P}{\cprojfst{A}{P}{w}}{\cprojsnd{A}{P}{w}}
+  \tag{by definition}
 \end{align*}
@@
 \cprojfst{A}{P}{\tmext{A}{P}{x}{u}}
-& \jdeq \subst{\tmext{A}{P}{x}{u}}{\ctxwk{P}{\idfunc[A]}}\\
-& \jdeq \subst{u}{{x}{\ctxwk{P}{\idfunc[A]}}} \\
-& \jdeq \subst{u}{\ctxwk{\subst{x}{P}}{\subst{x}{\idfunc[A]}}}\\
-& \jdeq \subst{x}{\idfunc[A]}\\
-& \jdeq x.
+& \jdeq 
+  \subst{\tmext{A}{P}{x}{u}}{\ctxwk{P}{\idfunc[A]}}
+  \\
+& \jdeq 
+  \subst{u}{{x}{\ctxwk{P}{\idfunc[A]}}} 
+  \\
+& \jdeq 
+  \subst{u}{\ctxwk{\subst{x}{P}}{\subst{x}{\idfunc[A]}}}
+  \\
+& \jdeq 
+  \subst{x}{\idfunc[A]}
+  \\
+& \jdeq 
+  x.
 \end{align*}
@@
 \cprojsnd{A}{P}{\tmext{A}{P}{x}{u}}
-& \jdeq \subst{\tmext{A}{P}{x}{u}}{\idfunc[P]}\\
-& \jdeq \subst{u}{{x}{\idfunc[P]}}\\
-& \jdeq \subst{u}{\idfunc[\subst{x}{P}]}\\
-& \jdeq u.
+& \jdeq 
+  \subst{\tmext{A}{P}{x}{u}}{\idfunc[P]}
+  \\
+& \jdeq 
+  \subst{u}{{x}{\idfunc[P]}}
+  \\
+& \jdeq 
+  \subst{u}{\idfunc[\subst{x}{P}]}
+  \\
+& \jdeq 
+  u.
 \end{align*}
@@
 & \inference
-  {\jterm{\Gamma}{A}{x}
+  { \jterm{\Gamma}{A}{x}
     \jterm{\Gamma}{\subst{x}{P}}{u}
-   \jterm{\Gamma}{\subst{\tmext{A}{P}{x}{u}}{Q}}{v}}
-  {\jtermeq{\Gamma}{\ctxext{{A}{P}}{Q}}{\tmext{A}{{P}{Q}}{x}{{u}{v}}}{\tmext{{A}{P}}{Q}{{x}{u}}{v}}}\\
-& \inference
-  {\jterm{\Gamma}{\ctxext{{A}{P}}{Q}}{w}}
-  {\jtermeq{\Gamma}{A}{\cprojfst{A}{P}{\cprojfst{\ctxext{A}{P}}{Q}{w}}}{\cprojfst{A}{\ctxext{P}{Q}}{w}}}\\
-& \inference
-  {\jterm{\Gamma}{\ctxext{{A}{P}}{Q}}{w}}
-  {\jtermeq{\Gamma}{\subst{\cprojfst{A}{\ctxext{P}{Q}}{w}}{P}}
+    \jterm{\Gamma}{\subst{\tmext{A}{P}{x}{u}}{Q}}{v}
+    }
+  { \jtermeq
+      {\Gamma}
+      {\ctxext{{A}{P}}{Q}}
+      {\tmext{A}{{P}{Q}}{x}{{u}{v}}}
+      {\tmext{{A}{P}}{Q}{{x}{u}}{v}}
+    }
+  \\
+& \inference
+  { \jterm{\Gamma}{\ctxext{{A}{P}}{Q}}{w}
+    }
+  { \jtermeq
+      {\Gamma}
+      {A}
+      {\cprojfst{A}{P}{\cprojfst{\ctxext{A}{P}}{Q}{w}}}
+      {\cprojfst{A}{\ctxext{P}{Q}}{w}}
+    }
+  \\
+& \inference
+  { \jterm{\Gamma}{\ctxext{{A}{P}}{Q}}{w}
+    }
+  { \jtermeq
+      {\Gamma}
+      {\subst{\cprojfst{A}{\ctxext{P}{Q}}{w}}{P}}
       {\cprojsnd{A}{P}{\cprojfst{\ctxext{A}{P}}{Q}{w}}}
-    {\cprojfst{P}{Q}{\cprojsnd{A}{\ctxext{P}{Q}}{w}}}}\\
-& \inference
-  {\jterm{\Gamma}{\ctxext{{A}{P}}{Q}}{w}}
-  {\jtermeq{\Gamma}{\subst{\cprojfst{\ctxext{A}{P}}{Q}{w}}{Q}}
+      {\cprojfst{P}{Q}{\cprojsnd{A}{\ctxext{P}{Q}}{w}}}
+    }
+  \\
+& \inference
+    { \jterm{\Gamma}{\ctxext{{A}{P}}{Q}}{w}
+      }
+    { \jtermeq
+        {\Gamma}
+        {\subst{\cprojfst{\ctxext{A}{P}}{Q}{w}}{Q}}
         {\cprojsnd{P}{Q}{\cprojsnd{A}{\ctxext{P}{Q}}{w}}}
-    {\cprojsnd{\ctxext{A}{P}}{Q}{w}}}
+        {\cprojsnd{\ctxext{A}{P}}{Q}{w}}
+      }
 \end{align*}
@@
 \begin{align*}
-\tmext{A}{{P}{Q}}{x}{{u}{v}} & \jdeq \subst{\tmext{P}{Q}{u}{v}}{{x}{\idfunc[\ctxext{A}{{P}{Q}}]}}\\
-& \jdeq \subst{v}{{u}{{x}{\idfunc[\ctxext{A}{{P}{Q}}]}}}\\
-& \jdeq \subst{v}{{u}{{x}{\idfunc[\ctxext{{A}{P}}{Q}]}}}\\
-& \jdeq \subst{v}{{\tmext{A}{P}{x}{u}}{\idfunc[\ctxext{{A}{P}}{Q}]}}\\
-& \jdeq \tmext{{A}{P}}{Q}{{x}{u}}{v}.
+\tmext{A}{{P}{Q}}{x}{{\subst{x}{P}}{\subst{x}{Q}}{u}{v}} 
+& \jdeq 
+  \subst
+    {\tmext{\subst{x}{P}}{\subst{x}{Q}}{u}{v}}
+    {{x}{\idfunc[\ctxext{A}{{P}{Q}}]}}
+  \\
+& \jdeq 
+  \subst{v}{{u}{{x}{\idfunc[\ctxext{A}{{P}{Q}}]}}}
+  \\
+& \jdeq 
+  \subst{v}{{u}{{x}{\idfunc[\ctxext{{A}{P}}{Q}]}}}
+  \\
+& \jdeq 
+  \subst{v}{{\tmext{A}{P}{x}{u}}{\idfunc[\ctxext{{A}{P}}{Q}]}}
+  \\
+& \jdeq 
+  \tmext{{A}{P}}{Q}{{A}{P}{x}{u}}{v}.
 \end{align*}
@@
 w 
-& \jdeq \tmext{A}{{P}{Q}}{\cprojfst{A}{\ctxext{P}{Q}}{w}}{\cprojsnd{A}{\ctxext{P}{Q}}{w}}\\
-& \jdeq \tmext{A}{{P}{Q}}{\cprojfst{A}{\ctxext{P}{Q}}{w}}{{\cprojfst{P}{Q}{\cprojsnd{A}{\ctxext{P}{Q}}{w}}}{\cprojsnd{P}{Q}{\cprojsnd{A}{\ctxext{P}{Q}}{w}}}}\\
-& \jdeq \tmext{{A}{P}}{Q}{{\cprojfst{A}{\ctxext{P}{Q}}{w}}{\cprojfst{P}{Q}{\cprojsnd{A}{\ctxext{P}{Q}}{w}}}}{\cprojsnd{P}{Q}{\cprojsnd{A}{\ctxext{P}{Q}}{w}}}
+& \jdeq 
+  \tmext
+    {A}
+    {{P}{Q}}
+    {\cprojfst{A}{\ctxext{P}{Q}}{w}}
+    {\cprojsnd{A}{\ctxext{P}{Q}}{w}}
+  \\
+& \jdeq 
+  \tmext
+    {A}
+    {{P}{Q}}
+    {\cprojfst{A}{\ctxext{P}{Q}}{w}}
+    { {P}
+      {Q}
+      {\cprojfst{P}{Q}{\cprojsnd{A}{\ctxext{P}{Q}}{w}}}
+      {\cprojsnd{P}{Q}{\cprojsnd{A}{\ctxext{P}{Q}}{w}}}
+      }
+  \\
+& \jdeq 
+  \tmext
+    {{A}{P}}
+    {Q}
+    { {} % need to provide base and family, but there's no unfold.
+      {}
+      {\cprojfst{A}{\ctxext{P}{Q}}{w}}
+      {\cprojfst{P}{Q}{\cprojsnd{A}{\ctxext{P}{Q}}{w}}}
+      }
+    { \cprojsnd{P}{Q}{\cprojsnd{A}{\ctxext{P}{Q}}{w}}
+      }
 \end{align*}
@@
 \begin{align*}
-\cprojfst{\ctxext{A}{P}}{Q}{w} & \jdeq \tmext{A}{P}{\cprojfst{A}{\ctxext{P}{Q}}{w}}{\cprojfst{P}{Q}{\cprojsnd{A}{\ctxext{P}{Q}}{w}}}\\ 
-\cprojsnd{\ctxext{A}{P}}{Q}{w} & \jdeq \cprojsnd{P}{Q}{\cprojsnd{A}{\ctxext{P}{Q}}{w}},
+\cprojfst{\ctxext{A}{P}}{Q}{w} 
+& \jdeq 
+  \tmext
+    {A}
+    {P}
+    {\cprojfst{A}{\ctxext{P}{Q}}{w}}
+    {\cprojfst{P}{Q}{\cprojsnd{A}{\ctxext{P}{Q}}{w}}}
+  \\ 
+\cprojsnd{\ctxext{A}{P}}{Q}{w} 
+& \jdeq 
+  \cprojsnd{P}{Q}{\cprojsnd{A}{\ctxext{P}{Q}}{w}},
 \end{align*}
@@
 \begin{align*}
-\cprojfst{A}{P}{\cprojfst{\ctxext{A}{P}}{Q}{w}} & \jdeq \cprojfst{A}{\ctxext{P}{Q}}{w}\\
-\cprojsnd{A}{P}{\cprojfst{\ctxext{A}{P}}{Q}{w}} & \jdeq \cprojfst{P}{Q}{\cprojsnd{A}{\ctxext{P}{Q}}{w}},
+\cprojfst{A}{P}{\cprojfst{\ctxext{A}{P}}{Q}{w}} 
+& \jdeq 
+  \cprojfst{A}{\ctxext{P}{Q}}{w}
+  \\
+\cprojsnd{A}{P}{\cprojfst{\ctxext{A}{P}}{Q}{w}} 
+& \jdeq 
+  \cprojfst{P}{Q}{\cprojsnd{A}{\ctxext{P}{Q}}{w}},
 \end{align*}
@@
 & \inference
-  {\jterm{{\Gamma}{B}}{Q}{g}
-   \jterm{{\Gamma}{B}}{\subst{g}{R}}{t}}
-  {\jtermeq
+  { \jterm{{\Gamma}{B}}{Q}{g}
+    \jterm{{\Gamma}{B}}{\subst{g}{R}}{t}
+    }
+  { \jtermeq
       {{{\Gamma}{A}}{\ctxwk{A}{B}}}
@@
       {\ctxwk{A}{\tmext{Q}{R}{g}{t}}}
-    {\tmext{\ctxwk{A}{Q}}{\ctxwk{A}{R}}{\ctxwk{A}{g}}{\ctxwk{A}{t}}}}
+      {\tmext{\ctxwk{A}{Q}}{\ctxwk{A}{R}}{\ctxwk{A}{g}}{\ctxwk{A}{t}}}
+    }
 % \\
@@
 \ctxwk{A}{\ctxext{Q}{R}{g}{t}}
-& \jdeq \ctxwk{A}{\subst{t}{{g}{\idfunc[\ctxext{Q}{R}]}}}\\
-& \jdeq \subst{\ctxwk{A}{t}}{\ctxwk{A}{\subst{g}{\idfunc[\ctxext{Q}{R}]}}}\\
-& \jdeq \subst{\ctxwk{A}{t}}{{\ctxwk{A}{g}}{\ctxwk{A}{\idfunc[\ctxext{Q}{R}]}}}\\
-& \jdeq \subst{\ctxwk{A}{t}}{{\ctxwk{A}{g}}{\idfunc[\ctxwk{A}{\ctxext{Q}{R}}]}}\\
-& \jdeq \subst{\ctxwk{A}{t}}{{\ctxwk{A}{g}}{\idfunc[\ctxext{\ctxwk{A}{Q}}{\ctxwk{A}{R}}]}}\\
-& \jdeq \tmext{\ctxwk{A}{Q}}{\ctxwk{A}{R}}{\ctxwk{A}{g}}{\ctxwk{A}{t}}
+& \jdeq 
+  \ctxwk{A}{\subst{t}{{g}{\idfunc[\ctxext{Q}{R}]}}}
+  \\
+& \jdeq 
+  \subst{\ctxwk{A}{t}}{\ctxwk{A}{\subst{g}{\idfunc[\ctxext{Q}{R}]}}}
+  \\
+& \jdeq 
+  \subst{\ctxwk{A}{t}}{{\ctxwk{A}{g}}{\ctxwk{A}{\idfunc[\ctxext{Q}{R}]}}}
+  \\
+& \jdeq 
+  \subst{\ctxwk{A}{t}}{{\ctxwk{A}{g}}{\idfunc[\ctxwk{A}{\ctxext{Q}{R}}]}}
+  \\
+& \jdeq 
+  \subst{\ctxwk{A}{t}}{{\ctxwk{A}{g}}{\idfunc[\ctxext{\ctxwk{A}{Q}}{\ctxwk{A}{R}}]}}
+  \\
+& \jdeq 
+  \tmext{\ctxwk{A}{Q}}{\ctxwk{A}{R}}{\ctxwk{A}{g}}{\ctxwk{A}{t}}
 \end{align*}
@@
 & \inference
-  {\jterm{{{\Gamma}{A}}{P}}{Q}{g}
-   \jterm{{{\Gamma}{A}}{P}}{\subst{g}{R}}{t}}
-  {\jtermeq
+  { \jterm{{{\Gamma}{A}}{P}}{Q}{g}
+    \jterm{{{\Gamma}{A}}{P}}{\subst{g}{R}}{t}
+    }
+  { \jtermeq
       {{\Gamma}{\subst{x}{P}}}
@@
       {\subst{x}{\tmext{Q}{R}{g}{t}}}
-    {\tmext{\subst{x}{Q}}{\subst{x}{R}}{\subst{x}{g}}{\subst{x}{t}}}}
+      {\tmext{\subst{x}{Q}}{\subst{x}{R}}{\subst{x}{g}}{\subst{x}{t}}}
+    }
 \end{align*}
@@
 \subst{x}{\tmext{Q}{R}{g}{t}}
-& \jdeq \subst{x}{{t}{{g}{\idfunc[\ctxext{Q}{R}]}}}\\
-& \jdeq \subst{{x}{t}}{{x}{{g}{\idfunc[\ctxext{Q}{R}]}}}\\
-& \jdeq \subst{{x}{t}}{{{x}{g}}{{x}{\idfunc[\ctxext{Q}{R}]}}}\\
-& \jdeq \subst{{x}{t}}{{{x}{g}}{\idfunc[\subst{x}{\ctxext{Q}{R}}]}}\\
-& \jdeq \subst{{x}{t}}{{{x}{g}}{\idfunc[\ctxext{\subst{x}{Q}}{\subst{x}{R}}]}}\\
-& \jdeq \tmext{\subst{x}{Q}}{\subst{x}{R}}{\subst{x}{g}}{\subst{x}{t}}.
+& \jdeq 
+  \subst{x}{{t}{{g}{\idfunc[\ctxext{Q}{R}]}}}
+  \\
+& \jdeq 
+  \subst{{x}{t}}{{x}{{g}{\idfunc[\ctxext{Q}{R}]}}}
+  \\
+& \jdeq 
+  \subst{{x}{t}}{{{x}{g}}{{x}{\idfunc[\ctxext{Q}{R}]}}}
+  \\
+& \jdeq 
+  \subst{{x}{t}}{{{x}{g}}{\idfunc[\subst{x}{\ctxext{Q}{R}}]}}
+  \\
+& \jdeq 
+  \subst{{x}{t}}{{{x}{g}}{\idfunc[\ctxext{\subst{x}{Q}}{\subst{x}{R}}]}}
+  \\
+& \jdeq 
+  \tmext{\subst{x}{Q}}{\subst{x}{R}}{\subst{x}{g}}{\subst{x}{t}}.
 \end{align*}
@@
 & \inference
-  {\jhom{\Gamma}{A}{B}{f}
+  { \jhom{\Gamma}{A}{B}{f}
     \jhom{\Gamma}{B}{C}{g}
     \jfam{{\Gamma}{C}}{R}
-   \jterm{{\Gamma}{B}}{\subst{g}{\ctxwk{B}{R}}}{w}}
-  {\jhomeq{\Gamma}{A}{{C}{R}}
+    \jterm{{\Gamma}{B}}{\subst{g}{\ctxwk{B}{R}}}{w}
+    }
+  { \jhomeq
+      {\Gamma}
+      {A}
+      {{C}{R}}
       {\jcomp{A}{f}{\tmext{\ctxwk{B}{C}}{\ctxwk{B}{R}}{g}{w}}}
-    {\tmext{\ctxwk{A}{C}}{\ctxwk{A}{R}}{\jcomp{A}{f}{g}}{\jcomp{A}{f}{w}}}}
+      {\tmext{\ctxwk{A}{C}}{\ctxwk{A}{R}}{\jcomp{A}{f}{g}}{\jcomp{A}{f}{w}}}
+    }
 \end{align*}
@@
 \jcomp{A}{f}{\tmext{\ctxwk{B}{C}}{\ctxwk{B}{R}}{g}{w}}
-& \jdeq \subst{f}{\ctxwk{A}{\tmext{\ctxwk{B}{C}}{\ctxwk{B}{R}}{g}{w}}}\\
-& \jdeq \subst{f}{\tmext{\ctxwk{A}{{B}{C}}}{\ctxwk{A}{{B}{R}}}{\ctxwk{A}{g}}{\ctxwk{A}{w}}}\\
-& \jdeq \tmext{\ctxwk{A}{C}}{\ctxwk{A}{R}}{\subst{f}{\ctxwk{A}{g}}}{\subst{f}{\ctxwk{A}{w}}}\\
-& \jdeq \tmext{\ctxwk{A}{C}}{\ctxwk{A}{R}}{\jcomp{A}{f}{g}}{\jcomp{A}{f}{w}}.
+& \jdeq 
+  \subst{f}{\ctxwk{A}{\tmext{\ctxwk{B}{C}}{\ctxwk{B}{R}}{g}{w}}}
+  \\
+& \jdeq 
+  \subst
+    {f}
+    {\tmext{\ctxwk{A}{{B}{C}}}{\ctxwk{A}{{B}{R}}}{\ctxwk{A}{g}}{\ctxwk{A}{w}}}
+  \\
+& \jdeq 
+  \tmext
+    {\ctxwk{A}{C}}
+    {\ctxwk{A}{R}}
+    {\subst{f}{\ctxwk{A}{g}}}
+    {\subst{f}{\ctxwk{A}{w}}}
+  \\
+& \jdeq 
+  \tmext{\ctxwk{A}{C}}{\ctxwk{A}{R}}{\jcomp{A}{f}{g}}{\jcomp{A}{f}{w}}.
 \end{align*}
@@
 & \inference
-  {\jfam{\Gamma}{A}}
-  {\jtype{{\Gamma}{A}}{\unitc{A}}}
-& & \inference
-    {\jfameq{\Gamma}{A}{A'}}
-    {\jtypeeq{{\Gamma}{A}}
+  { \jfam{\Gamma}{A}
+    }
+  { \jtype{{\Gamma}{A}}{\unitc{A}}
+    }
+& & \inference
+    { \jfameq{\Gamma}{A}{A'}
+      }
+    { \jtypeeq
+        {{\Gamma}{A}}
         {\unitc{A}}
-      {\unitc{A'}}}\\
-& \inference
-  {\jfam{\Gamma}{A}}
-  {\jterm{{\Gamma}{A}}{\unitc{A}}{\unitct{A}}}
-& & \inference
-    {\jfameq{\Gamma}{A}{A'}}
-    {\jtermeq{{\Gamma}{A}}{\unitc{A}}{\unitct{A}}{\unitct{A'}}}\\
-& \inference
-  {\jfam{{\Gamma}{A}}{P}}
-  {\jtype{{{{\Gamma}{A}}{P}}{\ctxwk{P}{\unitc{A}}}}{\unitf{A}{P}}}
-& & \inference
-    {\jfameq{{\Gamma}{A}}{P}{P'}}
-    {\jtypeeq{{{{\Gamma}{A}}{P}}{\ctxwk{P}{\unitc{A}}}}
+        {\unitc{A'}}
+      }
+  \\
+& \inference
+    { \jfam{\Gamma}{A}
+      }
+    { \jterm{{\Gamma}{A}}{\unitc{A}}{\unitct{A}}
+      }
+& & \inference
+    { \jfameq{\Gamma}{A}{A'}
+      }
+    { \jtermeq{{\Gamma}{A}}{\unitc{A}}{\unitct{A}}{\unitct{A'}}
+      }
+  \\
+& \inference
+  { \jfam{{\Gamma}{A}}{P}
+    }
+  { \jtype{{{{\Gamma}{A}}{P}}{\ctxwk{P}{\unitc{A}}}}{\unitf{A}{P}}
+    }
+& & \inference
+    { \jfameq{{\Gamma}{A}}{P}{P'}
+      }
+    { \jtypeeq
+        {{{{\Gamma}{A}}{P}}{\ctxwk{P}{\unitc{A}}}}
         {\unitf{A}{P}}
-      {\unitf{A}{P'}}}\\
-& \inference
-  {\jfam{{\Gamma}{A}}{P}}
-  {\jterm{{{{\Gamma}{A}}{P}}{\ctxwk{P}{\unitc{A}}}}{\unitf{A}{P}}{\unitft{A}{P}}}
-& & \inference
-    {\jfameq{{\Gamma}{A}}{P}{P'}}
-    {\jtermeq{{{{\Gamma}{A}}{P}}{\ctxwk{P}{\unitc{A}}}}{\unitf{A}{P}}
+        {\unitf{A}{P'}}
+      }
+  \\
+& \inference
+  { \jfam{{\Gamma}{A}}{P}
+    }
+  { \jterm
+      {{{{\Gamma}{A}}{P}}{\ctxwk{P}{\unitc{A}}}}
+      {\unitf{A}{P}}
       {\unitft{A}{P}}
-      {\unitft{A}{P'}}}
+    }
+& & \inference
+    { \jfameq{{\Gamma}{A}}{P}{P'}
+      }
+    { \jtermeq
+        {{{{\Gamma}{A}}{P}}{\ctxwk{P}{\unitc{A}}}}
+        {\unitf{A}{P}}
+        {\unitft{A}{P}}
+        {\unitft{A}{P'}}
+      }
 \end{align*}
@@
 & \inference
-  {\jfam{{\Gamma}{A}}{P}}
-  {\jtypeeq{{{\Gamma}{A}}{P}}{\ctxext{\ctxwk{P}{\unitc{A}}}{\unitf{A}{P}}}{\unitc{{\Gamma}{A}}}}\\
-& \inference
-  {\jfam{\Gamma}{A}
-   \jfam{\Gamma}{B}}
-  {\jtypeeq{{{\Gamma}{A}}{\ctxwk{A}{B}}}{}{}}  
+  { \jfam{{\Gamma}{A}}{P}
+    }
+  { \jtypeeq
+      {{{\Gamma}{A}}{P}}
+      {\ctxext{\ctxwk{P}{\unitc{A}}}{\unitf{A}{P}}}
+      {\unitc{{\Gamma}{A}}}
+    }\\
+& \inference
+  { \jfam{\Gamma}{A}
+    \jfam{\Gamma}{B}
+    }
+  { \jtypeeq{{{\Gamma}{A}}{\ctxwk{A}{B}}}{}{}
+    }  
 \end{align*}
@@
 & \inference
-  {}
-  {\jctx{\mctx}}
+  { }
+  { \jctx{\mctx}
+    }
   \tag{the canonical model}\\
 & \inference
-  {\jfam{\mctx}{P}}
-  {\jterm{\mctx}{P}{i}}
+  { \jfam{\mctx}{P}
+    }
+  { \jterm{\mctx}{P}{i}
+    }
   \tag{initiality of $\mctx$}\\
 & \inference
-  {\jfam{\Gamma}{A}}
-  {\jfam{{\Gamma}{A}}{\mfam{A}}}
+  { \jfam{\Gamma}{A}
+    }
+  { \jfam{{\Gamma}{A}}{\mfam{A}}
+    }
   \tag{families}\\
 & \inference
-  {\jctx{\Gamma}}
-  {\jfam{{\Gamma}{\mfam{\Gamma}}}{\mtm{\Gamma}}}
+  { \jctx{\Gamma}
+    }
+  { \jfam{{\Gamma}{\mfam{\Gamma}}}{\mtm{\Gamma}}
+    }
   \tag{terms}
@@
 & \inference
-  {\jterm{{\Gamma}{A}}{P}{f}}
-  {\jhom
+  { \jterm{{\Gamma}{A}}{P}{f}
+    }
+  { \jhom
       {{\Gamma}{A}}
@@
       {\subst{f}{\mfam{P}}}
-    {\mfam{f}}}\\
-& \inference
-  {\jterm{{\Gamma}{A}}{P}{f}}
-  {\jfhom
+      {\mfam{f}}
+    }
+  \\
+& \inference
+  { \jterm{{\Gamma}{A}}{P}{f}
+    }
+  { \jfhom
       {{\Gamma}{A}}
@@
       {\subst{f}{\mtm{P}}}
-    {\mtm{f}}}
+      {\mtm{f}}
+    }
 \intertext{%
@@
 & \inference
-  {\jfam{\Gamma}{A}}
-  {\jterm{\Gamma}{A}{\tfemp{A}}}
+  { \jfam{\Gamma}{A}
+    }
+  { \jterm{\Gamma}{A}{\tfemp{A}}
+    }
   \tag{empty context}\\
 & \inference
-  {\jfam{\Gamma}{A}}
-  {\jterm{{{\Gamma}{A}}{\mfam{A}}}{\subst{\tfemp{\mfam{A}}}{\mtm{A}}}{\tft{A}}}
+  { \jfam{\Gamma}{A}
+    }
+  { \jterm{{{\Gamma}{A}}{\mfam{A}}}{\subst{\tfemp{\mfam{A}}}{\mtm{A}}}{\tft{A}}
+    }
 \intertext{%
@@
 & \inference
-  {\jfam{\Gamma}{A}}
-  {\jhom{\Gamma}{{A}{\mfam{A}}}{A}{\tfext{A}}}
+  { \jfam{\Gamma}{A}
+    }
+  { \jhom{\Gamma}{{A}{\mfam{A}}}{A}{\tfext{A}}
+    }
   \tag{extension}
@@
 & \inference
-  {\jfam{\Gamma}{A}}
-  {\jhom{{{\Gamma}{A}}{\mfam{A}}}{\ctxwk{\mfam{A}}{\mfam{A}}}
+  { \jfam{\Gamma}{A}
+    }
+  { \jhom
+      {{{\Gamma}{A}}{\mfam{A}}}
+      {\ctxwk{\mfam{A}}{\mfam{A}}}
       {\mfam{\mfam{A}}}
-    {\tfwk{A}^0}}
+      {\tfwk{A}^0}
+    }
   \tag{weakening}
@@
 & \inference
-  {\jfam{\Gamma}{A}}
-  {\jfhom
+  { \jfam{\Gamma}{A}
+    }
+  { \jfhom
       {{{\Gamma}{A}}{\mfam{A}}}
@@
       {\mtm{\mfam{\mfam{A}}}}
-    {\tfwk{A}^1}}
+      {\tfwk{A}^1}
+    }
   \tag{weakening}
@@
 & \inference
-  {\jfam{\Gamma}{A}}
-  {\jhom
+  { \jfam{\Gamma}{A}
+    }
+  { \jhom
       {{{{\Gamma}{A}}{\mfam{A}}}{\mtm{A}}}
@@
       {\ctxwk{\mtm{A}}{{\mfam{A}}{\mfam{A}}}}
-  {\tfsubst{A}^0}}
+      {\tfsubst{A}^0}
+    }
   \tag{substitution}
@@
 & \inference
-  {\jfam{\Gamma}{A}}
-  {\jfhom
+  { \jfam{\Gamma}{A}
+    }
+  { \jfhom
       {{{{\Gamma}{A}}{\mfam{A}}}{\mtm{A}}}
@@
       {\ctxwk{\mtm{A}}{{\mfam{A}}{\mtm{A}}}}
-  {\tfsubst{A}^1}}
+      {\tfsubst{A}^1}
+    }
   \tag{substitution}
@@
 & \inference
-  {\jfam{\Gamma}{A}}
-  {\jterm
+  { \jfam{\Gamma}{A}
+    }
+  { \jterm
       {{{\Gamma}{A}}{\mfam{A}}}
       {\subst{{\idfunc[\mfam{A}]}{\tfwk{A}^0}}{\mtm{\mfam{\mfam{A}}}}}
-   {\tfid{A}}}
+      {\tfid{A}}
+    }
   \tag{identity functions}
@@
   { \jfam{\Gamma}{A}
-    \jfam{\Gamma}{B}}
+    \jfam{\Gamma}{B}
+    }
   { \jfameq
@@
       {\mfam{\ctxwk{A}{B}}}
-    {\ctxwk{A}{\mfam{B}}}}.
+      {\ctxwk{A}{\mfam{B}}}
+    }.
 \end{equation*}
@@
     {\mtm{\ctxwk{A}{B}}}
-    {\ctxwk{A}{\mtm{B}}}}
+    {\ctxwk{A}{\mtm{B}}}
+    }
 \end{equation*}
@@
 \inference
-  {\jfam{\Gamma}{A}}
-  {\jfameq
+  { \jfam{\Gamma}{A}
+    }
+  { \jfameq
       {{{\Gamma}{A}}{\mfam{A}}}
       {\mfam{\mfam{A}}}
-    {\jcomp{{A}{\mfam{A}}}{\tfext{A}}{\mfam{A}}}}
+      {\jcomp{{A}{\mfam{A}}}{\tfext{A}}{\mfam{A}}}
+    }
 \end{equation*}
@@
 \subst{\tfext{A}}{\mfam{\ctxwk{\ctxext{A}{\mfam{A}}}{A}}}
-\jdeq \unfold{\jcomp{{A}{P}}{\tfext{A}}{\mfam{A}}}
-\jdeq \jcomp{{A}{P}}{\tfext{A}}{\mfam{A}}
+\jdeq 
+  \unfold{\jcomp{{A}{P}}{\tfext{A}}{\mfam{A}}}
+\jdeq 
+  \jcomp{{A}{P}}{\tfext{A}}{\mfam{A}}
 \end{equation*}
@@
 \inference
-  {\jfam{\Gamma}{A}}
-  {\jhomeq
+  { \jfam{\Gamma}{A}
+    }
+  { \jhomeq
       {{\Gamma}{A}}
@@
       {\mfam{\tfext{A}}^{\ctxext{\Gamma}{A}}}
-    {\idfunc[\jcomp{}{\tfext{A}}{\mfam{A}}]}}
+      {\idfunc[\jcomp{}{\tfext{A}}{\mfam{A}}]}
+    }
 \end{equation*}
@@
 \inference
-  {\jfam{\Gamma}{A}}
-  {\jfameq{\Gamma}{\subst{\tfemp{A}}{\mfam{A}}}{A}}
+  { \jfam{\Gamma}{A}
+    }
+  { \jfameq{\Gamma}{\subst{\tfemp{A}}{\mfam{A}}}{A}
+    }
 \end{equation*}
@@
 \inference
-  {\jfam{\Gamma}{A}}
-  {\jfameq{{\Gamma}{A}}{\subst{\tfemp{\mfam{A}}}{\mtm{A}}}{\emptyf[A]}}
+  { \jfam{\Gamma}{A}
+    }
+  { \jfameq{{\Gamma}{A}}{\subst{\tfemp{\mfam{A}}}{\mtm{A}}}{\emptyf[A]}
+    }
 \end{equation*}
@@
 \inference
-  {\jfam{\Gamma}{A}}
-  {\jtermeq{\Gamma}{A}{\subst{\tfemp{A}}{\tfemp{\mfam{A}}}}{\tfemp{A}}}
+  { \jfam{\Gamma}{A}
+    }
+  { \jtermeq{\Gamma}{A}{\subst{\tfemp{A}}{\tfemp{\mfam{A}}}}{\tfemp{A}}
+    }
 \end{equation*}
@@
 \begin{equation*}
-\subst{\tfemp{\mfam{A}}}{\ctxwk{\ctxext{A}{\mfam{A}}}{A}}\jdeq
-\subst{\tfemp{\mfam{A}}}{\ctxwk{\mfam{A}}{{A}{A}}}\jdeq
+\subst{\tfemp{\mfam{A}}}{\ctxwk{\ctxext{A}{\mfam{A}}}{A}}
+\jdeq
+  \subst{\tfemp{\mfam{A}}}{\ctxwk{\mfam{A}}{{A}{A}}}
+\jdeq
   \ctxwk{A}{A}.
@@
 \inference
-  {\jfam{\Gamma}{A}}
-  {\jhomeq{\Gamma}{A}{A}{\subst{\tfemp{\mfam{A}}}{\tfext{A}}}{\idfunc[A]}}
+  { \jfam{\Gamma}{A}
+    }
+  { \jhomeq{\Gamma}{A}{A}{\subst{\tfemp{\mfam{A}}}{\tfext{A}}}{\idfunc[A]}
+    }
 \end{equation*}
@@
 \subst{\tfemp{A}}{\ctxwk{\ctxext{A}{\mfam{A}}}{A}}
-& \jdeq \subst{\tfemp{A}}{\ctxwk{\mfam{A}}{{A}{A}}}\\
-& \jdeq \ctxwk{\subst{\tfemp{A}}{\mfam{A}}}{\subst{\tfemp{A}}{\ctxwk{A}{A}}}\\
-& \jdeq \ctxwk{A}{\subst{\tfemp{A}}{\ctxwk{A}{A}}}\\
-& \jdeq \ctxwk{A}{A}
+& \jdeq 
+  \subst{\tfemp{A}}{\ctxwk{\mfam{A}}{{A}{A}}}
+  \\
+& \jdeq 
+  \ctxwk{\subst{\tfemp{A}}{\mfam{A}}}{\subst{\tfemp{A}}{\ctxwk{A}{A}}}
+  \\
+& \jdeq 
+  \ctxwk{A}{\subst{\tfemp{A}}{\ctxwk{A}{A}}}
+  \\
+& \jdeq 
+  \ctxwk{A}{A}
 \end{align*}
@@
 \inference
-  {\jfam{\Gamma}{A}}
-  {\jhomeq{\Gamma}{A}{A}{\subst{\tfemp{A}}{\tfext{A}}}{\idfunc[A]}}
+  { \jfam{\Gamma}{A}
+    }
+  { \jhomeq{\Gamma}{A}{A}{\subst{\tfemp{A}}{\tfext{A}}}{\idfunc[A]}
+    }
 \end{equation*}

### 2014-03-05 article-models-stage2.tex
written rule for weakening followed by substitution

@@
 can possibly begin exploring some internal independence proofs.
+
+\section{Stage three: Beyond}
+We also pose the following questions which, I'm afraid, are beyond the scope
+of the proposed research program.
+
+\begin{enumerate}
+\item In this document we have developed a custom type theory without basic
+constructors. It is not a question whether someone will want to propose
+a slightly different type theory without basic constructors, someone will. So
+it would be nice if there were a type theory of type theories, formalizing what
+a type theory actually is. Then one could define his or her new type theory
+as an object of that theory. Moreover, it could then be possible to write a
+program that automatically generates a theorem prover for user defined type
+theories in this language, so that a person who has come up with a new type
+theory (such as Guillaume who came up with his type theory of higher groupoids)
+can immediately start giving formally verified proofs in his new type theory.
+One could also envision that there will be morphisms in this type theory which
+automatically transform a formalized development of a theory from one type
+theory to another, so that not every simple thing has to be done over and over
+again. So the question is, could something like that actually be possible? We
+may actually use our type theory before type constructors for that. Or, if we
+manage to get hold on the internal logic of a model
+(which should be a type theory such that the model in question is its initial
+model?)
+for each object of our type theory of models of type theory, we might actually
+already have a `type theory of type theories'.
+\end{enumerate}

### 2014-03-05 article-models-tt.tex
written rule for weakening followed by substitution

@@
 \end{equation}
+
+Now let $\jfam{{\Gamma}{A}}{P}$ and $\jfam{{\Gamma}{A}}{Q}$ be families. A
+morphism from $P$ to $Q$ over the identity function $\idfunc[A]$ in context
+$\Gamma$ is the same thing as a morphism from $P$ to $Q$ in context
+$\ctxext{\Gamma}{A}$:
+
+\begin{lem}\label{hom-over-id-is-hom}
+We have the following valid inference rules:
+\begin{align*}
+& \inference
+  { \jfam{{\Gamma}{A}}{P}
+    \jfam{{\Gamma}{A}}{Q}
+    \jfhom{\Gamma}{A}{A}{\idfunc[A]}{P}{Q}{f}
+    }
+  { \jhom{{\Gamma}{A}}{P}{Q}{f}
+    }
+  \\
+& \inference
+  { \jfam{{\Gamma}{A}}{P}
+    \jfam{{\Gamma}{A}}{Q}
+    \jhom{{\Gamma}{A}}{P}{Q}{f}
+    }
+  { \jfhom{\Gamma}{A}{A}{\idfunc[A]}{P}{Q}{f}
+    }
+\end{align*}
+\end{lem}
+
+\begin{proof}
+If we unfold the judgments $\jhom{{\Gamma}{A}}{P}{Q}{f}$ and
+$\jfhom{\Gamma}{A}{A}{\idfunc[A]}{P}{Q}{f}$, we get the judgments
+\begin{align*}
+& \unfold{\jhom{{\Gamma}{A}}{P}{Q}{f}}
+  \\
+& \unfold{\jfhom{\Gamma}{A}{A}{\idfunc[A]}{P}{Q}{f}}
+\end{align*}
+respectively. Therefore, we only need to verify that
+$\ctxwk{P}{\subst{\idfunc[A]}{\ctxwk{A}{Q}}}\jdeq\ctxwk{P}{Q}$, which is indeed
+the case by \autoref{idfunc-wk-defn}.
+\end{proof}
 
@@
 
+\subsubsection{Terms of families over families are terms of families over extensions}
+\begin{equation*}
+\inference
+  { \jfam{\Gamma}{A}
+    }
+  { \jfameq
+      {{{{\Gamma}{A}}{\mfam{A}}}{\mfam{\mfam{A}}}}
+      {\mtm{\mfam{A}}}
+      {\jcomp{}{\tfext{A}}{\mtm{A}}}
+      }
+\end{equation*}
+
 \subsubsection{Extension acts as the identity on families of families}
@@
 \end{equation*}
-Therefore, there is a term
+Because we have the judgmental equality $\ctxext{\Gamma}{{A}{\mfam{A}}}
+\jdeq \ctxext{{\Gamma}{A}}{\mfam{A}}$, there is a term
 \begin{equation*}
 \jhom
-  {{\Gamma}{A}}
+  {{{\Gamma}{A}}{\mfam{A}}}
   {\mfam{\mfam{A}}}
@@
   { \jhomeq
-      {{\Gamma}{A}}
+      {{{\Gamma}{A}}{\mfam{A}}}
       {\jcomp{}{\tfext{A}}{\mfam{A}}}
@@
 is in general, but at the moment I don't see how to do this)}
+
+\subsubsection{Extension acts as the identity on terms of families of families}
+We now investigate the nature of the morphism.
+\begin{equation*}
+\jfhom
+  {{{\Gamma}{A}}{\mfam{A}}}
+  {\mfam{\mfam{A}}}
+  {\subst{\tfext{A}}{\mfam{\ctxwk{\ctxext{A}{\mfam{A}}}{A}}}}
+  {\mfam{\tfext{A}}^{\ctxext{\Gamma}{A}}}
+  {\mtm{\mfam{A}}}
+  {\subst{\tfext{A}}{\mtm{\ctxwk{\ctxext{A}{\mfam{A}}}{A}}}}
+  {\mtm{\tfext{A}}^{\ctxext{\Gamma}{A}}}
+\end{equation*}
+Note that we have the judgmental equality 
+$\mtm{\mfam{A}}\jdeq\jcomp{}{\tfext{A}}{\mtm{A}}$. Likewise, we have the
+judgmental equality $\subst{\tfext{A}}{\mtm{\ctxwk{\ctxext{A}{\mfam{A}}}{A}}}
+\jdeq\jcomp{}{\tfext{A}}{\mtm{A}}$. Thirdly, we have the judgmental equality
+$\mfam{\tfext{A}}^{\ctxext{\Gamma}{A}}\jdeq\idfunc[\jcomp{}{\tfext{A}}{\mfam{A}}]$.
+We may combine these three facts with \autoref{hom-over-id-is-hom} to see that
+\begin{equation*}
+\jhom
+  {{{{\Gamma}{A}}{\mfam{A}}}{\jcomp{}{\tfext{A}}}{\mfam{A}}}
+  {\jcomp{}{\tfext{A}}{\mtm{A}}}
+  {\jcomp{}{\tfext{A}}{\mtm{A}}}
+  {\mtm{\tfext{A}}^{\ctxext{\Gamma}{A}}}
+\end{equation*}
+Now we see that we can require that $\mtm{\tfext{A}}^{\ctxext{\Gamma}{A}}$ is
+the identity function on $\jcomp{}{\tfext{A}}{\mtm{A}}$:
+\begin{equation*}
+\inference
+  { \jfam{\Gamma}{A}
+    }
+  { \jhomeq
+    {{{{\Gamma}{A}}{\mfam{A}}}{\jcomp{}{\tfext{A}}}{\mfam{A}}}
+    {\jcomp{}{\tfext{A}}{\mtm{A}}}
+    {\jcomp{}{\tfext{A}}{\mtm{A}}}
+    {\mtm{\tfext{A}}^{\ctxext{\Gamma}{A}}}
+    {\idfunc[\jcomp{}{\tfext{A}}{\mtm{A}}]}
+    }
+\end{equation*}
+
+\subsubsection{Weakening followed by substitution is the identity}
+The weakening morphism $\tfwk{A}^0$ gives a morphism
+\begin{equation*}
+\jhom
+  {{{{\Gamma}{A}}{\mfam{A}}}{\mtm{A}}}
+  {\ctxwk{\mtm{A}}{{\mfam{A}}{\mfam{A}}}}
+  {\ctxwk{\mtm{A}}{\mfam{\mfam{A}}}}
+  {\ctxwk{\mtm{A}}{\tfwk{A}^0}}
+\end{equation*}
+which can be composed with the morphism $\tfsubst{A}^0$. We require that this
+composition is the identity function on $\ctxwk{\mtm{A}}{{\mfam{A}}{\mfam{A}}}$:
+\begin{equation*}
+\inference
+  { \jfam{\Gamma}{A}
+    }
+  { \jhomeq
+      {{{{\Gamma}{A}}{\mfam{A}}}{\mtm{A}}}
+      {\ctxwk{\mtm{A}}{{\mfam{A}}{\mfam{A}}}}
+      {\ctxwk{\mtm{A}}{{\mfam{A}}{\mfam{A}}}}
+      {\ctxwk{\mtm{A}}{\tfwk{A}^0}}
+      {\idfunc[\ctxwk{\mtm{A}}{{\mfam{A}}{\mfam{A}}}]}
+    }
+\end{equation*}
 

### 2014-03-05 article-models-tt.tex
placed the section on the model type theory in a separate file

--- 2014-03-05 article-models-tt.tex
@@
 
+Now let $\jfam{{\Gamma}{A}}{P}$ and $\jfam{{\Gamma}{A}}{Q}$ be families. A
+morphism from $P$ to $Q$ over the identity function $\idfunc[A]$ in context
+$\Gamma$ is the same thing as a morphism from $P$ to $Q$ in context
+$\ctxext{\Gamma}{A}$:
+
+\begin{lem}\label{hom-over-id-is-hom}
+We have the following valid inference rules:
+\begin{align*}
+& \inference
+  { \jfam{{\Gamma}{A}}{P}
+    \jfam{{\Gamma}{A}}{Q}
+    \jfhom{\Gamma}{A}{A}{\idfunc[A]}{P}{Q}{f}
+    }
+  { \jhom{{\Gamma}{A}}{P}{Q}{f}
+    }
+  \\
+& \inference
+  { \jfam{{\Gamma}{A}}{P}
+    \jfam{{\Gamma}{A}}{Q}
+    \jhom{{\Gamma}{A}}{P}{Q}{f}
+    }
+  { \jfhom{\Gamma}{A}{A}{\idfunc[A]}{P}{Q}{f}
+    }
+\end{align*}
+\end{lem}
+
+\begin{proof}
+If we unfold the judgments $\jhom{{\Gamma}{A}}{P}{Q}{f}$ and
+$\jfhom{\Gamma}{A}{A}{\idfunc[A]}{P}{Q}{f}$, we get the judgments
+\begin{align*}
+& \unfold{\jhom{{\Gamma}{A}}{P}{Q}{f}}
+  \\
+& \unfold{\jfhom{\Gamma}{A}{A}{\idfunc[A]}{P}{Q}{f}}
+\end{align*}
+respectively. Therefore, we only need to verify that
+$\ctxwk{P}{\subst{\idfunc[A]}{\ctxwk{A}{Q}}}\jdeq\ctxwk{P}{Q}$, which is indeed
+the case by \autoref{idfunc-wk-defn}.
+\end{proof}
+
 \subsection{Extension on terms}\label{extension-on-terms}
@@
 \label{compatibility-of-type-constructors}
-
-\section{The type theory of models of type theory}
-In this section we pursue the idea of what a general model of type theory is by
-axiomatizing what you can do with them. We have the following ideas:
-\begin{enumerate}
-\item There are dependent models and sections thereof. Particular instances
-  of sections: extension, weakening and substitution (and something for identity
-  function?). And like the original
-  extension, weakening and substitution, they're going to be compatible with each
-  other. Therefore we state our theory of models as an extension of type theory
-  without constructors.
-\item Type theory without constructors is a model of itself, the canonical model $\mctx$.
-\item for any (family of) model(s) $A$, there is the family model $\mfam{A}$ which
-  is a family of models over $A$.
-\item The terms of $\mctx$ should be precisely contexts. Terms of $\subst{\Gamma}{\mfam{\mctx}}$
-  should be families over $\Gamma$.
-\item if a model and a family of models over it are given, there is an extended model.
-  If we extend $A$ by $\mfam{A}$, we get the Sierpinski model of $A$.
-\item Likewise, models can be weakened and substituted and there are identity
-  functions.
-\end{enumerate}
-So the theory of models is going to be an extension of this theory with this
-data. The theory we are about to describe can be seen as an elementary theory
-of the category (with families (and terms)) of categories (with families (and terms)).
-
-We would also like to remark explicitly that the valid judgments of the original type
-theory become contexts, families or terms, depending on what kind of judgments
-it was. Then, valid inferences of the original type theory become valid
-judgments here. In some sense, the theory we present here is therefore a second
-order theory over the basic theory of types.
-
-\subsection{The basic ingredients of the type theory of models}
-We first introduce the basic ingredients of our abstract theory of models.
-\begin{align*}
-& \inference
-  { }
-  { \jctx{\mctx}
-    }
-  \tag{the canonical model}\\
-& \inference
-  { \jfam{\mctx}{P}
-    }
-  { \jterm{\mctx}{P}{i}
-    }
-  \tag{initiality of $\mctx$}\\
-& \inference
-  { \jfam{\Gamma}{A}
-    }
-  { \jfam{{\Gamma}{A}}{\mfam{A}}
-    }
-  \tag{families}\\
-& \inference
-  { \jctx{\Gamma}
-    }
-  { \jfam{{\Gamma}{\mfam{\Gamma}}}{\mtm{\Gamma}}
-    }
-  \tag{terms}
-\intertext{%
-  The following two rules essentially make a start with saying that every term is functorial
-  in the apropriate sense (there will be more rules contributing to this vies):}
-& \inference
-  { \jterm{{\Gamma}{A}}{P}{f}
-    }
-  { \jhom
-      {{\Gamma}{A}}
-      {\mfam{A}}
-      {\subst{f}{\mfam{P}}}
-      {\mfam{f}}
-    }
-  \\
-& \inference
-  { \jterm{{\Gamma}{A}}{P}{f}
-    }
-  { \jfhom
-      {{\Gamma}{A}}
-      {\mfam{A}}
-      {\subst{f}{\mfam{P}}}
-      {\mfam{f}}
-      {\mtm{A}}
-      {\subst{f}{\mtm{P}}}
-      {\mtm{f}}
-    }
-\intertext{%
-  When the situation requires clarity --  for instance when extensions
-  are involved -- we will write $\mfam{f}^\Gamma$ and $\mtm{f}^\Gamma$ to make
-  what we regard as the context explicit.}
-\intertext{%
-The empty context of a model $A$ is simply a term of $A$:}
-& \inference
-  { \jfam{\Gamma}{A}
-    }
-  { \jterm{\Gamma}{A}{\tfemp{A}}
-    }
-  \tag{empty context}\\
-& \inference
-  { \jfam{\Gamma}{A}
-    }
-  { \jterm{{{\Gamma}{A}}{\mfam{A}}}{\subst{\tfemp{\mfam{A}}}{\mtm{A}}}{\tft{A}}
-    }
-\intertext{%
-  Extension is going to be a morphism from $\ctxext{A}{\mfam{A}}$ to
-  $A$ in context $\Gamma$:}
-& \inference
-  { \jfam{\Gamma}{A}
-    }
-  { \jhom{\Gamma}{{A}{\mfam{A}}}{A}{\tfext{A}}
-    }
-  \tag{extension}
-\intertext{%
-  The action of weakening on families
-  is a morphism from $\ctxwk{\mfam{A}}{\mfam{A}}$ to
-  $\mfam{\mfam{A}}$ in context $\ctxext{{\Gamma}{A}}{\mfam{A}}$:}
-& \inference
-  { \jfam{\Gamma}{A}
-    }
-  { \jhom
-      {{{\Gamma}{A}}{\mfam{A}}}
-      {\ctxwk{\mfam{A}}{\mfam{A}}}
-      {\mfam{\mfam{A}}}
-      {\tfwk{A}^0}
-    }
-  \tag{weakening}
-\intertext{%
-  The action of weakening on terms
-  is a morphism from $\ctxwk{\mfam{A}}{\mtm{A}}$ to $\mtm{\mfam{\mfam{A}}}$ over
-  $\tfwk{A}^0$ in context $\ctxext{{\Gamma}{A}}{\mfam{A}}$:}
-& \inference
-  { \jfam{\Gamma}{A}
-    }
-  { \jfhom
-      {{{\Gamma}{A}}{\mfam{A}}}
-      {\ctxwk{\mfam{A}}{\mfam{A}}}
-      {\mfam{\mfam{A}}}
-      {\tfwk{A}^0}
-      {\ctxwk{\mfam{A}}{\mtm{A}}}
-      {\mtm{\mfam{\mfam{A}}}}
-      {\tfwk{A}^1}
-    }
-  \tag{weakening}
-\intertext{%
-  The action of substitution by a term on families
-  is going to be a morphism from $\ctxwk{\mtm{A}}{\mfam{\mfam{A}}}$ to 
-  $\ctxwk{\mtm{A}}{{\mfam{A}}{\mfam{A}}}$
-  in context $\ctxext{{{\Gamma}{A}}{\mfam{A}}}{\mtm{A}}$:}
-& \inference
-  { \jfam{\Gamma}{A}
-    }
-  { \jhom
-      {{{{\Gamma}{A}}{\mfam{A}}}{\mtm{A}}}
-      {\ctxwk{\mtm{A}}{\mfam{\mfam{A}}}}
-      {\ctxwk{\mtm{A}}{{\mfam{A}}{\mfam{A}}}}
-      {\tfsubst{A}^0}
-    }
-  \tag{substitution}
-\intertext{%
-  Likewise, the action of substitution by a term on terms of those
-  families is going to be a morphism from $\ctxwk{\mtm{A}}{\mtm{\mfam{\mfam{A}}}}$
-  to $\ctxwk{\mtm{A}}{{\mfam{A}}{\mtm{A}}}$ over $\tfsubst{A}^0$ in context
-  $\ctxext{{{\Gamma}{A}}{\mfam{A}}}{\mtm{A}}$:}
-& \inference
-  { \jfam{\Gamma}{A}
-    }
-  { \jfhom
-      {{{{\Gamma}{A}}{\mfam{A}}}{\mtm{A}}}
-      {\ctxwk{\mtm{A}}{\mfam{\mfam{A}}}}
-      {\ctxwk{\mtm{A}}{{\mfam{A}}{\mfam{A}}}}
-      {\tfsubst{A}^0}
-      {\ctxwk{\mtm{A}}{\mtm{\mfam{\mfam{A}}}}}
-      {\ctxwk{\mtm{A}}{{\mfam{A}}{\mtm{A}}}}
-      {\tfsubst{A}^1}
-    }
-  \tag{substitution}
-\intertext{%
-  The identity functions are coded by}
-& \inference
-  { \jfam{\Gamma}{A}
-    }
-  { \jterm
-      {{{\Gamma}{A}}{\mfam{A}}}
-      {\subst{{\idfunc[\mfam{A}]}{\tfwk{A}^0}}{\mtm{\mfam{\mfam{A}}}}}
-      {\tfid{A}}
-    }
-  \tag{identity functions}
-\end{align*}
-We will call the terms that we introduced here the model constructors.
-
-\subsection{The compatibility rules}
-We need to do several things in this section:
-\begin{enumerate}
-\item postulate that ordinary extension, weakening, substitution and identity
-functions are compatible with the model constructors.
-\item postulate that each of the model constructors is compatible with ordinary
-extension, weakening, substitution and identity functions. Actually, we want
-that every term of this type theory is compatible with those. I.e.~every term
-is a functor/morphism of models. Also the sections.
-\item postulate that the model constructors are compatible with each other,
-so that they come to model ordinary extension, weakening, substitution and
-identity functions respectively.
-\end{enumerate}
-Note that the rules for compatibility with extension are going to explain which
-model $\ctxext{A}{P}$ is by telling what the families, the terms, extension,
-weakening, substitution and identity functions are. Likewise the rules for
-compatibility with weakening and substitution do this for their respective
-cases.
-
-The easiest set of compatibility rules comes with weakening. The compatibility
-rules that deal with substitution are likely going to have to do with the
-Yoneda lemma, which we should be able to implement at some point.
-
-\subsubsection{Families of weakenings}
-\begin{equation*}
-\inference
-  { \jfam{\Gamma}{A}
-    \jfam{\Gamma}{B}
-    }
-  { \jfameq
-      {{{\Gamma}{A}}{\ctxwk{A}{B}}}
-      {\mfam{\ctxwk{A}{B}}}
-      {\ctxwk{A}{\mfam{B}}}
-    }.
-\end{equation*}
-
-\subsubsection{Terms of weakenings}
-\begin{equation*}
-\inference
-  { \jfam{\Gamma}{A}
-    \jfam{\Gamma}{B}
-    }
-  { \jfameq
-    {{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\mfam{\ctxwk{A}{B}}}}
-    {\mtm{\ctxwk{A}{B}}}
-    {\ctxwk{A}{\mtm{B}}}
-    }
-\end{equation*}
-
-\subsubsection{Families over families are families over extensions}
-\begin{equation*}
-\inference
-  { \jfam{\Gamma}{A}
-    }
-  { \jfameq
-      {{{\Gamma}{A}}{\mfam{A}}}
-      {\mfam{\mfam{A}}}
-      {\jcomp{{A}{\mfam{A}}}{\tfext{A}}{\mfam{A}}}
-    }
-\end{equation*}
-\emph{(Note: this rule might be a consequence of an explanation of what $\mfam{{A}{P}}$
-is in general, but at the moment I don't see how to do this)}
-
-\subsubsection{Extension acts as the identity on families of families}
-The family of which $\tfext{A}$ is a term, is unfolded as
-\begin{equation*}
-\unfold{\jhom{\Gamma}{{A}{\mfam{A}}}{A}{\tfext{A}}}
-\end{equation*}
-Therefore, there is a term
-\begin{equation*}
-\jhom
-  {{\Gamma}{A}}
-  {\mfam{\mfam{A}}}
-  {\subst{\tfext{A}}{\mfam{\ctxwk{\ctxext{A}{\mfam{A}}}{A}}}}
-  {\mfam{\tfext{A}}^{\ctxext{\Gamma}{A}}}
-\end{equation*}
-By the rule asserting that families of families are families of extensions, we
-have $\mfam{\mfam{A}}\jdeq\jcomp{}{\tfext{A}}{\mfam{A}}$. By the rule that
-families over weakenings are weakenings of families, we have the judgmental
-equalities
-\begin{equation*}
-\subst{\tfext{A}}{\mfam{\ctxwk{\ctxext{A}{\mfam{A}}}{A}}}
-\jdeq 
-  \unfold{\jcomp{{A}{P}}{\tfext{A}}{\mfam{A}}}
-\jdeq 
-  \jcomp{{A}{P}}{\tfext{A}}{\mfam{A}}
-\end{equation*}
-Therefore, we can compare the term $\mfam{\tfext{A}}^{\ctxext{\Gamma}{A}}$
-to the identity function $\idfunc[\jcomp{}{\tfext{A}}{\mfam{A}}]$. 
-The compatibility
-rule for the action of extension on families of families asserts that
-\begin{equation*}
-\inference
-  { \jfam{\Gamma}{A}
-    }
-  { \jhomeq
-      {{\Gamma}{A}}
-      {\jcomp{}{\tfext{A}}{\mfam{A}}}
-      {\jcomp{}{\tfext{A}}{\mfam{A}}}
-      {\mfam{\tfext{A}}^{\ctxext{\Gamma}{A}}}
-      {\idfunc[\jcomp{}{\tfext{A}}{\mfam{A}}]}
-    }
-\end{equation*}
-\emph{(Note: this rule might be a consequence of an explanation of what $\mfam{{A}{P}}$
-is in general, but at the moment I don't see how to do this)}
-
-\subsubsection{Families of extensions}
-
-\subsubsection{Terms of families of extensions}
-
-\subsubsection{Terms of extensions}
-
-\subsubsection{Families over the empty context}
-\begin{equation*}
-\inference
-  { \jfam{\Gamma}{A}
-    }
-  { \jfameq{\Gamma}{\subst{\tfemp{A}}{\mfam{A}}}{A}
-    }
-\end{equation*}
-
-\subsubsection{Terms of the empty family}
-\begin{equation*}
-\inference
-  { \jfam{\Gamma}{A}
-    }
-  { \jfameq{{\Gamma}{A}}{\subst{\tfemp{\mfam{A}}}{\mtm{A}}}{\emptyf[A]}
-    }
-\end{equation*}
-
-\subsubsection{The empty family over the empty family}
-\begin{equation*}
-\inference
-  { \jfam{\Gamma}{A}
-    }
-  { \jtermeq{\Gamma}{A}{\subst{\tfemp{A}}{\tfemp{\mfam{A}}}}{\tfemp{A}}
-    }
-\end{equation*}
-
-\subsubsection{Extension by the empty family}
-To understand the following inference rule, recall that for $\jfam{\Gamma}{A}$
-we have the judgmental equalities
-\begin{equation*}
-\subst{\tfemp{\mfam{A}}}{\ctxwk{\ctxext{A}{\mfam{A}}}{A}}
-\jdeq
-  \subst{\tfemp{\mfam{A}}}{\ctxwk{\mfam{A}}{{A}{A}}}
-\jdeq
-  \ctxwk{A}{A}.
-\end{equation*}
-Therefore, we can postulate:
-\begin{equation*}
-\inference
-  { \jfam{\Gamma}{A}
-    }
-  { \jhomeq{\Gamma}{A}{A}{\subst{\tfemp{\mfam{A}}}{\tfext{A}}}{\idfunc[A]}
-    }
-\end{equation*}
-
-\subsubsection{Extensions of the empty context}
-To understand the following inference rule, recall that for $\jfam{\Gamma}{A}$
-we have the judgmental equalities
-\begin{align*}
-\subst{\tfemp{A}}{\ctxwk{\ctxext{A}{\mfam{A}}}{A}}
-& \jdeq 
-  \subst{\tfemp{A}}{\ctxwk{\mfam{A}}{{A}{A}}}
-  \\
-& \jdeq 
-  \ctxwk{\subst{\tfemp{A}}{\mfam{A}}}{\subst{\tfemp{A}}{\ctxwk{A}{A}}}
-  \\
-& \jdeq 
-  \ctxwk{A}{\subst{\tfemp{A}}{\ctxwk{A}{A}}}
-  \\
-& \jdeq 
-  \ctxwk{A}{A}
-\end{align*}
-Therefore, we can postulate
-\begin{equation*}
-\inference
-  { \jfam{\Gamma}{A}
-    }
-  { \jhomeq{\Gamma}{A}{A}{\subst{\tfemp{A}}{\tfext{A}}}{\idfunc[A]}
-    }
-\end{equation*}
-
-\subsubsection{Weakening by the empty family}
-
-\subsubsection{Weakenings of the empty family}
-
-\subsubsection{Substitutions of the empty family}
-
-\subsubsection{Substitution by the term of the empty family}
-
-\subsubsection{Extension acts as the identity functor on families}
-\begin{equation*}
-df
-\end{equation*}
-
-\subsubsection{Extension acts as the identity functor on terms}
-
-\subsubsection{Extensions of extensions}
-
-\subsubsection{Extensions of weakenings}
-
-\subsubsection{Weakenings by extensions}
-
-\subsubsection{Substitutions of extensions}
-
-\section{Pretty type theory}
-In this section we will do three things. First we explain basic type theory with
-variable names, which we simply call \emph{Pretty Type Theory}. 
-Then we will show how every formula in basic type theory without
-variable names (shall we call it \emph{Structural Type Theory}?) 
-can be interpreted in pretty type theory. Finally, we will show how every formula
-in pretty type theory can be interpreted in structural type theory.
-
-Pretty type theory is pretty as the ugly little duckling that grew up. At first, it is actually uglier than
-structural type theory because one has to keep track of a variable nobody is
-actually interested in. The situation improves when we introduce the operators
-of extension and weakening, because from that point onwards one can draw notational
-advantages from having the variable around.
-
-Pretty type theory should be such that if you write down two contexts, families
-or terms in exactly the same way, then they are the same, and there should be
-notational shortcuts for extension, weakening and substitution which make
-this interesting.
-
-\subsection{The basic judgments}
-The basic judgments of pretty type theory are the same as for structural type
-theory. There are judgments for: ``$\Gamma$ is a context'',
-``$A(i)$ over $i:\Gamma$ is a family over $\Gamma$'', ``$A(i)$ over $i:\Gamma$ 
-is a type in context $\Gamma$''
-and ``$x(i)$ is a term of $A(i)$ above $i:\Gamma$''. The other four
-judgments are for judgmental equality.
-
-\begin{align*}
-\jvctx*{\Gamma} & \jvctxeq*{\Gamma}{\Gamma'}\\
-\jvfam*{i}{\Gamma}{A} & \jvfameq*{i}{\Gamma}{A}{B}\\
-\jvtype*{i}{\Gamma}{A} & \jvtypeeq*{i}{\Gamma}{A}{B}\\
-\jvterm*{i}{\Gamma}{A}{x} & \jvtermeq*{i}{\Gamma}{A}{x}{y}.
-\end{align*}
-
-We have the following basic inference rules that relate types and families:
-
-\begin{small}
-\begin{align*}
-& \inference
-  {\jvtype{i}{\Gamma}{A}}
-  {\jvfam{i}{\Gamma}{A}}
-& & \inference
-    {\jvtypeeq{i}{\Gamma}{A}{B}}
-    {\jvfameq{i}{\Gamma}{A}{B}}\\
-& \inference
-  {\jvtype{i}{\Gamma}{A}
-   \jvfameq{i}{\Gamma}{A}{B}}
-  {\jvtype{i}{\Gamma}{B}}
-& & \inference
-    {\jvtype{i}{\Gamma}{A}
-     \jvfameq{i}{\Gamma}{A}{B}}
-    {\jvtypeeq{i}{\Gamma}{A}{B}}
-\end{align*}
-\end{small}
-
-\subsection{The basic rules for judgmental equality}
-The rules for judgmental equality establish that it is an equivalence relation.
-\bgroup\small
-\begin{align*}
-& \inference
-  {\jvctx{\Gamma}}
-  {\jvctxeq{\Gamma}{\Gamma}} 
-& & \inference
-    {\jvctxeq{\Gamma}{\Delta}}
-    {\jvctxeq{\Delta}{\Gamma}} 
-& & \inference
-    {\jvctxeq{\Gamma}{\Delta}
-     \jvctxeq{\Delta}{\greek{E}}}
-    {\jvctxeq{\Gamma}{\greek{E}}}\\
-& \inference
-  {\jvfam{i}{\Gamma}{A}}
-  {\jvfameq{i}{\Gamma}{A}{A}} 
-& & \inference
-    {\jvfameq{i}{\Gamma}{A}{B}}
-    {\jvfameq{i}{\Gamma}{B}{A}}
-& & \inference
-    {\jvfameq{i}{\Gamma}{A}{B}
-     \jvfameq{i}{\Gamma}{B}{C}}
-    {\jvfameq{i}{\Gamma}{A}{C}}\\
-& \inference
-  {\jvterm{i}{\Gamma}{A}{x}}
-  {\jvtermeq{i}{\Gamma}{A}{x}{x}}
-& & \inference
-    {\jvtermeq{i}{\Gamma}{A}{x}{y}}
-    {\jvtermeq{i}{\Gamma}{A}{y}{x}}
-& & \inference
-    {\jvtermeq{i}{\Gamma}{A}{x}{y}
-     \jvtermeq{i}{\Gamma}{A}{y}{z}}
-    {\jvtermeq{i}{\Gamma}{A}{x}{z}}
-\end{align*}
-\egroup
-
-The following convertibility rules are responsible for the strictness
-of judgmental equality, which sets it apart from equivalences or identifications:
-
-\begin{align*}
-& \inference
-  {\jvctxeq{\Gamma}{\Delta}
-   \jvfam{i}{\Gamma}{A}}
-  {\jvfam{i}{\Delta}{A}}
-& & \inference
-    {\jvctxeq{\Gamma}{\Delta}
-     \jvfameq{i}{\Gamma}{A}{B}}
-    {\jvfameq{i}{\Delta}{A}{B}}\\
-& \inference
-  {\jvctxeq{\Gamma}{\Delta}
-   \jvterm{i}{\Gamma}{A}{x}}
-  {\jvterm{i}{\Delta}{A}{x}}
-& & \inference
-    {\jvctxeq{\Gamma}{\Delta}
-     \jvtermeq{i}{\Gamma}{A}{x}{y}}
-    {\jvtermeq{i}{\Delta}{A}{x}{y}}\\
-& \inference
-  {\jvfameq{i}{\Gamma}{A}{B}
-   \jvterm{i}{\Gamma}{A}{x}}
-  {\jvterm{i}{\Gamma}{B}{x}}
-& & \inference
-    {\jvfameq{i}{\Gamma}{A}{B}
-     \jvtermeq{i}{\Gamma}{A}{x}{y}}
-    {\jvtermeq{i}{\Gamma}{B}{x}{y}}
-\end{align*}
-
-\subsection{The empty context}
-The empty context looks a bit strange when we explicitly denote the terms. But
-we will not do so anymore after this subsection.
-
-\begin{align}
-& \inference
-  {}
-  {\jctx{\emptyc}}\\
-& \inference
-  {\jctx{\Gamma}}
-  {\jvfam{i}{\Gamma}{\emptyf[\Gamma]}}\\
-& \inference
-  {\jctx{\Gamma}}
-  {\jvterm{i}{\Gamma}{\emptyf[\Gamma]}{\emptytm[\Gamma]}}\\
-& \inference
-  {\jvterm{i}{\Gamma}{\emptyf[\Gamma]}{x}}
-  {\jvtermeq{i}{\Gamma}{\emptyf[\Gamma]}{x}{\emptytm[\Gamma]}}
-\end{align}
-
-Moreover, if $\Gamma$ is a context family over the
-empty context, then $\Gamma$ is a context and every context is a context
-family over the empty context. Note that this allows us to speak
-of terms of contexts too.
-
-\begin{align}
-& \inference
-  {\jctx{\Gamma}}
-  {\jvfam{\nameless}{\emptyc}{\Gamma}} 
-& & \inference
-    {\jvfam{\nameless}{\emptyc}{\Gamma}}
-    {\jctx{\Gamma}}\\
-& \inference
-  {\jctxeq{\Gamma}{\Delta}}
-  {\jvfameq{\nameless}{\emptyc}{\Gamma}{\Delta}}
-& & \inference
-    {\jvfameq{\nameless}{\emptyc}{\Gamma}{\Delta}}
-    {\jctxeq{\Gamma}{\Delta}}
-\end{align}
-
-\subsubsection{The empty context is compatible with itslef}
-The empty context $\emptyc$ may be considered as a family of contexts over the empty
-context. When we do this, we get $\emptyf[\emptyc]$.
-\begin{equation}
-\inference
-  {}
-  {\jvfameq{\nameless}{\emptyc}{\emptyc}{\emptyf[\emptyc]}}
-\end{equation}
-In the future, we shall denote $\emptyf[\Gamma]$ by $\emptyf$. The above rule
-guarantees that this will not cause confusion. Likewise, we shall denote
-$\emptytm[\Gamma]$ by $\emptytm$.
-
-\subsection{Extension}
-We introduce extension which not only extends a context $\Gamma$ and a family
-$A$ over it to a context $\ctxext{\Gamma}{A}$, but which also extends a family $A$
-in context $\Gamma$ and a family $P$ over it to a family $\ctxext{A}{P}$ over context
-$\Gamma$. We do this to ensure that all of type theory can be done in a context.
-For instance, we could say (1) that a context in context $\Gamma$ is the same thing
-as a family over $\Gamma$; (2) When $A$ is a context in this sense, a family over
-$A$ is the same thing as a family $P$ over $\ctxext{\Gamma}{A}$ and 
-(3) when $P$ is a family over $A$ in this sense, a term of $P$ keeps its original meaning.
-
-\begin{align}
-& \inference
-  {\jvfam{i}{\Gamma}{A}}
-  {\jvfamcombi{{i}{x}}{{\Gamma}{A}}{P}}
-& & \inference
-    {\jctxeq{\Gamma}{\Delta}
-     \jfameq{\Gamma}{A}{B}}
-    {\jctxeq{\ctxext{\Gamma}{A}}{\ctxext{\Delta}{B}}}\\
-& \inference
-  {\jfam{{\Gamma}{A}}{P}}
-  {\jfam{\Gamma}{\ctxext{A}{P}}}
-& & \inference
-    {\jfameq{\Gamma}{A}{B}
-     \jfameq{{\Gamma}{A}}{P}{Q}}
-    {\jfameq{\Gamma}{\ctxext{A}{P}}{\ctxext{B}{Q}}}
-\end{align}
-
-\subsubsection{Extension is compatible with the empty context}
-The following rule asserts that extension by $\emptyc$ leaves the contexts unchanged.
-\begin{align}
-& \inference
-  {\jctx{\Gamma}}
-  {\jctxeq{\ctxext{\emptyc}{\Gamma}}{\Gamma}}\\
-& \inference
-  {\jctx{\Gamma}}
-  {\jctxeq{\ctxext{\Gamma}{\emptyf}}{\Gamma}}\\
-& \inference
-  {\jfam{\Gamma}{A}}
-  {\jfameq{\Gamma}{\ctxext{\emptyf}{A}}{A}}
-\end{align}
-
-\subsubsection{Extension is compatible with itself}\label{comp-ee}
-The inference rules asserting that extension is compatible with itself assert
-that contexts are unstructured lists of type declarations. This rule is
-unavoidable if we want that for a family $A$ in context $\Gamma$, a family over
-$A$ is the same thing as a family over $\ctxext{\Gamma}{A}$.
-
-\begin{align}
-& \inference
-  {\jfam{\Gamma}{A}
-   \jfam{{\Gamma}{A}}{P}}
-  {\jctxeq{\ctxext{{\Gamma}{A}}{P}}{\ctxext{\Gamma}{{A}{P}}}}\\
-& \inference
-  {\jfam{{\Gamma}{A}}{P}
-   \jfam{{{\Gamma}{A}}{P}}{Q}}
-  {\jfameq{\Gamma}{\ctxext{{A}{P}}{Q}}{\ctxext{A}{{P}{Q}}}}
-\end{align}
-
-\section{Chalmers type theory}
-In this section we will present the type theory that Coquand and Dybjer are using.
-It is a weak type theory (I think), with not so many operations and judgmental equalities.
-We will show how every formula of Chalmers type theory can be interpreted in
-structural type theory.

### 2014-03-05 article-models-ttmodels.tex
placed the section on the model type theory in a separate file

\section{The type theory of models of type theory}
In this section we pursue the idea of what a general model of type theory is by
axiomatizing what you can do with them. We have the following ideas:
\begin{enumerate}
\item There are dependent models and sections thereof. Particular instances
  of sections: extension, weakening and substitution (and something for identity
  function?). And like the original
  extension, weakening and substitution, they're going to be compatible with each
  other. Therefore we state our theory of models as an extension of type theory
  without constructors.
\item Type theory without constructors is a model of itself, the canonical model $\mctx$.
\item for any (family of) model(s) $A$, there is the family model $\mfam{A}$ which
  is a family of models over $A$.
\item The terms of $\mctx$ should be precisely contexts. Terms of $\subst{\Gamma}{\mfam{\mctx}}$
  should be families over $\Gamma$.
\item if a model and a family of models over it are given, there is an extended model.
  If we extend $A$ by $\mfam{A}$, we get the Sierpinski model of $A$.
\item Likewise, models can be weakened and substituted and there are identity
  functions.
\end{enumerate}
So the theory of models is going to be an extension of this theory with this
data. The theory we are about to describe can be seen as an elementary theory
of the category (with families (and terms)) of categories (with families (and terms)).

We would also like to remark explicitly that the valid judgments of the original type
theory become contexts, families or terms, depending on what kind of judgments
it was. Then, valid inferences of the original type theory become valid
judgments here. In some sense, the theory we present here is therefore a second
order theory over the basic theory of types.

\subsection{The basic ingredients of the type theory of models}
We first introduce the basic ingredients of our abstract theory of models.
\begin{align*}
& \inference
  { }
  { \jctx{\mctx}
    }
  \tag{the canonical model}\\
& \inference
  { \jfam{\mctx}{P}
    }
  { \jterm{\mctx}{P}{i}
    }
  \tag{initiality of $\mctx$}\\
& \inference
  { \jfam{\Gamma}{A}
    }
  { \jfam{{\Gamma}{A}}{\mfam{A}}
    }
  \tag{families}\\
& \inference
  { \jctx{\Gamma}
    }
  { \jfam{{\Gamma}{\mfam{\Gamma}}}{\mtm{\Gamma}}
    }
  \tag{terms}
\intertext{%
  The following two rules essentially make a start with saying that every term is functorial
  in the apropriate sense (there will be more rules contributing to this vies):}
& \inference
  { \jterm{{\Gamma}{A}}{P}{f}
    }
  { \jhom
      {{\Gamma}{A}}
      {\mfam{A}}
      {\subst{f}{\mfam{P}}}
      {\mfam{f}}
    }
  \\
& \inference
  { \jterm{{\Gamma}{A}}{P}{f}
    }
  { \jfhom
      {{\Gamma}{A}}
      {\mfam{A}}
      {\subst{f}{\mfam{P}}}
      {\mfam{f}}
      {\mtm{A}}
      {\subst{f}{\mtm{P}}}
      {\mtm{f}}
    }
\intertext{%
  When the situation requires clarity --  for instance when extensions
  are involved -- we will write $\mfam{f}^\Gamma$ and $\mtm{f}^\Gamma$ to make
  what we regard as the context explicit.}
\intertext{%
The empty context of a model $A$ is simply a term of $A$:}
& \inference
  { \jfam{\Gamma}{A}
    }
  { \jterm{\Gamma}{A}{\tfemp{A}}
    }
  \tag{empty context}\\
& \inference
  { \jfam{\Gamma}{A}
    }
  { \jterm{{{\Gamma}{A}}{\mfam{A}}}{\subst{\tfemp{\mfam{A}}}{\mtm{A}}}{\tft{A}}
    }
\intertext{%
  Extension is going to be a morphism from $\ctxext{A}{\mfam{A}}$ to
  $A$ in context $\Gamma$:}
& \inference
  { \jfam{\Gamma}{A}
    }
  { \jhom{\Gamma}{{A}{\mfam{A}}}{A}{\tfext{A}}
    }
  \tag{extension}
\intertext{%
  The action of weakening on families
  is a morphism from $\ctxwk{\mfam{A}}{\mfam{A}}$ to
  $\mfam{\mfam{A}}$ in context $\ctxext{{\Gamma}{A}}{\mfam{A}}$:}
& \inference
  { \jfam{\Gamma}{A}
    }
  { \jhom
      {{{\Gamma}{A}}{\mfam{A}}}
      {\ctxwk{\mfam{A}}{\mfam{A}}}
      {\mfam{\mfam{A}}}
      {\tfwk{A}^0}
    }
  \tag{weakening}
\intertext{%
  The action of weakening on terms
  is a morphism from $\ctxwk{\mfam{A}}{\mtm{A}}$ to $\mtm{\mfam{\mfam{A}}}$ over
  $\tfwk{A}^0$ in context $\ctxext{{\Gamma}{A}}{\mfam{A}}$:}
& \inference
  { \jfam{\Gamma}{A}
    }
  { \jfhom
      {{{\Gamma}{A}}{\mfam{A}}}
      {\ctxwk{\mfam{A}}{\mfam{A}}}
      {\mfam{\mfam{A}}}
      {\tfwk{A}^0}
      {\ctxwk{\mfam{A}}{\mtm{A}}}
      {\mtm{\mfam{\mfam{A}}}}
      {\tfwk{A}^1}
    }
  \tag{weakening}
\intertext{%
  The action of substitution by a term on families
  is going to be a morphism from $\ctxwk{\mtm{A}}{\mfam{\mfam{A}}}$ to 
  $\ctxwk{\mtm{A}}{{\mfam{A}}{\mfam{A}}}$
  in context $\ctxext{{{\Gamma}{A}}{\mfam{A}}}{\mtm{A}}$:}
& \inference
  { \jfam{\Gamma}{A}
    }
  { \jhom
      {{{{\Gamma}{A}}{\mfam{A}}}{\mtm{A}}}
      {\ctxwk{\mtm{A}}{\mfam{\mfam{A}}}}
      {\ctxwk{\mtm{A}}{{\mfam{A}}{\mfam{A}}}}
      {\tfsubst{A}^0}
    }
  \tag{substitution}
\intertext{%
  Likewise, the action of substitution by a term on terms of those
  families is going to be a morphism from $\ctxwk{\mtm{A}}{\mtm{\mfam{\mfam{A}}}}$
  to $\ctxwk{\mtm{A}}{{\mfam{A}}{\mtm{A}}}$ over $\tfsubst{A}^0$ in context
  $\ctxext{{{\Gamma}{A}}{\mfam{A}}}{\mtm{A}}$:}
& \inference
  { \jfam{\Gamma}{A}
    }
  { \jfhom
      {{{{\Gamma}{A}}{\mfam{A}}}{\mtm{A}}}
      {\ctxwk{\mtm{A}}{\mfam{\mfam{A}}}}
      {\ctxwk{\mtm{A}}{{\mfam{A}}{\mfam{A}}}}
      {\tfsubst{A}^0}
      {\ctxwk{\mtm{A}}{\mtm{\mfam{\mfam{A}}}}}
      {\ctxwk{\mtm{A}}{{\mfam{A}}{\mtm{A}}}}
      {\tfsubst{A}^1}
    }
  \tag{substitution}
\intertext{%
  The identity functions are coded by}
& \inference
  { \jfam{\Gamma}{A}
    }
  { \jterm
      {{{\Gamma}{A}}{\mfam{A}}}
      {\subst{{\idfunc[\mfam{A}]}{\tfwk{A}^0}}{\mtm{\mfam{\mfam{A}}}}}
      {\tfid{A}}
    }
  \tag{identity functions}
\end{align*}
We will call the terms that we introduced here the model constructors.

\subsection{The compatibility rules}
We need to do several things in this section:
\begin{enumerate}
\item postulate that ordinary extension, weakening, substitution and identity
functions are compatible with the model constructors.
\item postulate that each of the model constructors is compatible with ordinary
extension, weakening, substitution and identity functions. Actually, we want
that every term of this type theory is compatible with those. I.e.~every term
is a functor/morphism of models. Also the sections.
\item postulate that the model constructors are compatible with each other,
so that they come to model ordinary extension, weakening, substitution and
identity functions respectively.
\end{enumerate}
Note that the rules for compatibility with extension are going to explain which
model $\ctxext{A}{P}$ is by telling what the families, the terms, extension,
weakening, substitution and identity functions are. Likewise the rules for
compatibility with weakening and substitution do this for their respective
cases.

The easiest set of compatibility rules comes with weakening. The compatibility
rules that deal with substitution are likely going to have to do with the
Yoneda lemma, which we should be able to implement at some point.

\subsubsection{Families of weakenings}
\begin{equation*}
\inference
  { \jfam{\Gamma}{A}
    \jfam{\Gamma}{B}
    }
  { \jfameq
      {{{\Gamma}{A}}{\ctxwk{A}{B}}}
      {\mfam{\ctxwk{A}{B}}}
      {\ctxwk{A}{\mfam{B}}}
    }.
\end{equation*}

\subsubsection{Terms of weakenings}
\begin{equation*}
\inference
  { \jfam{\Gamma}{A}
    \jfam{\Gamma}{B}
    }
  { \jfameq
    {{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\mfam{\ctxwk{A}{B}}}}
    {\mtm{\ctxwk{A}{B}}}
    {\ctxwk{A}{\mtm{B}}}
    }
\end{equation*}

\subsubsection{Families over families are families over extensions}
\begin{equation*}
\inference
  { \jfam{\Gamma}{A}
    }
  { \jfameq
      {{{\Gamma}{A}}{\mfam{A}}}
      {\mfam{\mfam{A}}}
      {\jcomp{{A}{\mfam{A}}}{\tfext{A}}{\mfam{A}}}
    }
\end{equation*}
\emph{(Note: this rule might be a consequence of an explanation of what $\mfam{{A}{P}}$
is in general, but at the moment I don't see how to do this)}

\subsubsection{Terms of families over families are terms of families over extensions}
\begin{equation*}
\inference
  { \jfam{\Gamma}{A}
    }
  { \jfameq
      {{{{\Gamma}{A}}{\mfam{A}}}{\mfam{\mfam{A}}}}
      {\mtm{\mfam{A}}}
      {\jcomp{}{\tfext{A}}{\mtm{A}}}
      }
\end{equation*}

\subsubsection{Extension acts as the identity on families of families}
The family of which $\tfext{A}$ is a term, is unfolded as
\begin{equation*}
\unfold{\jhom{\Gamma}{{A}{\mfam{A}}}{A}{\tfext{A}}}
\end{equation*}
Because we have the judgmental equality $\ctxext{\Gamma}{{A}{\mfam{A}}}
\jdeq \ctxext{{\Gamma}{A}}{\mfam{A}}$, there is a term
\begin{equation*}
\jhom
  {{{\Gamma}{A}}{\mfam{A}}}
  {\mfam{\mfam{A}}}
  {\subst{\tfext{A}}{\mfam{\ctxwk{\ctxext{A}{\mfam{A}}}{A}}}}
  {\mfam{\tfext{A}}^{\ctxext{\Gamma}{A}}}
\end{equation*}
By the rule asserting that families of families are families of extensions, we
have $\mfam{\mfam{A}}\jdeq\jcomp{}{\tfext{A}}{\mfam{A}}$. By the rule that
families over weakenings are weakenings of families, we have the judgmental
equalities
\begin{equation*}
\subst{\tfext{A}}{\mfam{\ctxwk{\ctxext{A}{\mfam{A}}}{A}}}
\jdeq 
  \unfold{\jcomp{{A}{P}}{\tfext{A}}{\mfam{A}}}
\jdeq 
  \jcomp{{A}{P}}{\tfext{A}}{\mfam{A}}
\end{equation*}
Therefore, we can compare the term $\mfam{\tfext{A}}^{\ctxext{\Gamma}{A}}$
to the identity function $\idfunc[\jcomp{}{\tfext{A}}{\mfam{A}}]$. 
The compatibility
rule for the action of extension on families of families asserts that
\begin{equation*}
\inference
  { \jfam{\Gamma}{A}
    }
  { \jhomeq
      {{{\Gamma}{A}}{\mfam{A}}}
      {\jcomp{}{\tfext{A}}{\mfam{A}}}
      {\jcomp{}{\tfext{A}}{\mfam{A}}}
      {\mfam{\tfext{A}}^{\ctxext{\Gamma}{A}}}
      {\idfunc[\jcomp{}{\tfext{A}}{\mfam{A}}]}
    }
\end{equation*}
\emph{(Note: this rule might be a consequence of an explanation of what $\mfam{{A}{P}}$
is in general, but at the moment I don't see how to do this)}

\subsubsection{Extension acts as the identity on terms of families of families}
We now investigate the nature of the morphism.
\begin{equation*}
\jfhom
  {{{\Gamma}{A}}{\mfam{A}}}
  {\mfam{\mfam{A}}}
  {\subst{\tfext{A}}{\mfam{\ctxwk{\ctxext{A}{\mfam{A}}}{A}}}}
  {\mfam{\tfext{A}}^{\ctxext{\Gamma}{A}}}
  {\mtm{\mfam{A}}}
  {\subst{\tfext{A}}{\mtm{\ctxwk{\ctxext{A}{\mfam{A}}}{A}}}}
  {\mtm{\tfext{A}}^{\ctxext{\Gamma}{A}}}
\end{equation*}
Note that we have the judgmental equality 
$\mtm{\mfam{A}}\jdeq\jcomp{}{\tfext{A}}{\mtm{A}}$. Likewise, we have the
judgmental equality $\subst{\tfext{A}}{\mtm{\ctxwk{\ctxext{A}{\mfam{A}}}{A}}}
\jdeq\jcomp{}{\tfext{A}}{\mtm{A}}$. Thirdly, we have the judgmental equality
$\mfam{\tfext{A}}^{\ctxext{\Gamma}{A}}\jdeq\idfunc[\jcomp{}{\tfext{A}}{\mfam{A}}]$.
We may combine these three facts with \autoref{hom-over-id-is-hom} to see that
\begin{equation*}
\jhom
  {{{{\Gamma}{A}}{\mfam{A}}}{\jcomp{}{\tfext{A}}}{\mfam{A}}}
  {\jcomp{}{\tfext{A}}{\mtm{A}}}
  {\jcomp{}{\tfext{A}}{\mtm{A}}}
  {\mtm{\tfext{A}}^{\ctxext{\Gamma}{A}}}
\end{equation*}
Now we see that we can require that $\mtm{\tfext{A}}^{\ctxext{\Gamma}{A}}$ is
the identity function on $\jcomp{}{\tfext{A}}{\mtm{A}}$:
\begin{equation*}
\inference
  { \jfam{\Gamma}{A}
    }
  { \jhomeq
    {{{{\Gamma}{A}}{\mfam{A}}}{\jcomp{}{\tfext{A}}}{\mfam{A}}}
    {\jcomp{}{\tfext{A}}{\mtm{A}}}
    {\jcomp{}{\tfext{A}}{\mtm{A}}}
    {\mtm{\tfext{A}}^{\ctxext{\Gamma}{A}}}
    {\idfunc[\jcomp{}{\tfext{A}}{\mtm{A}}]}
    }
\end{equation*}

\subsubsection{Weakening followed by substitution is the identity}
The weakening morphism $\tfwk{A}^0$ gives a morphism
\begin{equation*}
\jhom
  {{{{\Gamma}{A}}{\mfam{A}}}{\mtm{A}}}
  {\ctxwk{\mtm{A}}{{\mfam{A}}{\mfam{A}}}}
  {\ctxwk{\mtm{A}}{\mfam{\mfam{A}}}}
  {\ctxwk{\mtm{A}}{\tfwk{A}^0}}
\end{equation*}
which can be composed with the morphism $\tfsubst{A}^0$. We require that this
composition is the identity function on $\ctxwk{\mtm{A}}{{\mfam{A}}{\mfam{A}}}$:
\begin{equation*}
\inference
  { \jfam{\Gamma}{A}
    }
  { \jhomeq
      {{{{\Gamma}{A}}{\mfam{A}}}{\mtm{A}}}
      {\ctxwk{\mtm{A}}{{\mfam{A}}{\mfam{A}}}}
      {\ctxwk{\mtm{A}}{{\mfam{A}}{\mfam{A}}}}
      {\ctxwk{\mtm{A}}{\tfwk{A}^0}}
      {\idfunc[\ctxwk{\mtm{A}}{{\mfam{A}}{\mfam{A}}}]}
    }
\end{equation*}

\subsubsection{Families of extensions}

\subsubsection{Terms of families of extensions}

\subsubsection{Terms of extensions}

\subsubsection{Families over the empty context}
\begin{equation*}
\inference
  { \jfam{\Gamma}{A}
    }
  { \jfameq{\Gamma}{\subst{\tfemp{A}}{\mfam{A}}}{A}
    }
\end{equation*}

\subsubsection{Terms of the empty family}
\begin{equation*}
\inference
  { \jfam{\Gamma}{A}
    }
  { \jfameq{{\Gamma}{A}}{\subst{\tfemp{\mfam{A}}}{\mtm{A}}}{\emptyf[A]}
    }
\end{equation*}

\subsubsection{The empty family over the empty family}
\begin{equation*}
\inference
  { \jfam{\Gamma}{A}
    }
  { \jtermeq{\Gamma}{A}{\subst{\tfemp{A}}{\tfemp{\mfam{A}}}}{\tfemp{A}}
    }
\end{equation*}

\subsubsection{Extension by the empty family}
To understand the following inference rule, recall that for $\jfam{\Gamma}{A}$
we have the judgmental equalities
\begin{equation*}
\subst{\tfemp{\mfam{A}}}{\ctxwk{\ctxext{A}{\mfam{A}}}{A}}
\jdeq
  \subst{\tfemp{\mfam{A}}}{\ctxwk{\mfam{A}}{{A}{A}}}
\jdeq
  \ctxwk{A}{A}.
\end{equation*}
Therefore, we can postulate:
\begin{equation*}
\inference
  { \jfam{\Gamma}{A}
    }
  { \jhomeq{\Gamma}{A}{A}{\subst{\tfemp{\mfam{A}}}{\tfext{A}}}{\idfunc[A]}
    }
\end{equation*}

\subsubsection{Extensions of the empty context}
To understand the following inference rule, recall that for $\jfam{\Gamma}{A}$
we have the judgmental equalities
\begin{align*}
\subst{\tfemp{A}}{\ctxwk{\ctxext{A}{\mfam{A}}}{A}}
& \jdeq 
  \subst{\tfemp{A}}{\ctxwk{\mfam{A}}{{A}{A}}}
  \\
& \jdeq 
  \ctxwk{\subst{\tfemp{A}}{\mfam{A}}}{\subst{\tfemp{A}}{\ctxwk{A}{A}}}
  \\
& \jdeq 
  \ctxwk{A}{\subst{\tfemp{A}}{\ctxwk{A}{A}}}
  \\
& \jdeq 
  \ctxwk{A}{A}
\end{align*}
Therefore, we can postulate
\begin{equation*}
\inference
  { \jfam{\Gamma}{A}
    }
  { \jhomeq{\Gamma}{A}{A}{\subst{\tfemp{A}}{\tfext{A}}}{\idfunc[A]}
    }
\end{equation*}

\subsubsection{Weakening by the empty family}

\subsubsection{Weakenings of the empty family}

\subsubsection{Substitutions of the empty family}

\subsubsection{Substitution by the term of the empty family}

\subsubsection{Extension acts as the identity functor on families}
\begin{equation*}
df
\end{equation*}

\subsubsection{Extension acts as the identity functor on terms}

\subsubsection{Extensions of extensions}

\subsubsection{Extensions of weakenings}

\subsubsection{Weakenings by extensions}

\subsubsection{Substitutions of extensions}

\section{Pretty type theory}
In this section we will do three things. First we explain basic type theory with
variable names, which we simply call \emph{Pretty Type Theory}. 
Then we will show how every formula in basic type theory without
variable names (shall we call it \emph{Structural Type Theory}?) 
can be interpreted in pretty type theory. Finally, we will show how every formula
in pretty type theory can be interpreted in structural type theory.

Pretty type theory is pretty as the ugly little duckling that grew up. At first, it is actually uglier than
structural type theory because one has to keep track of a variable nobody is
actually interested in. The situation improves when we introduce the operators
of extension and weakening, because from that point onwards one can draw notational
advantages from having the variable around.

Pretty type theory should be such that if you write down two contexts, families
or terms in exactly the same way, then they are the same, and there should be
notational shortcuts for extension, weakening and substitution which make
this interesting.

\subsection{The basic judgments}
The basic judgments of pretty type theory are the same as for structural type
theory. There are judgments for: ``$\Gamma$ is a context'',
``$A(i)$ over $i:\Gamma$ is a family over $\Gamma$'', ``$A(i)$ over $i:\Gamma$ 
is a type in context $\Gamma$''
and ``$x(i)$ is a term of $A(i)$ above $i:\Gamma$''. The other four
judgments are for judgmental equality.

\begin{align*}
\jvctx*{\Gamma} & \jvctxeq*{\Gamma}{\Gamma'}\\
\jvfam*{i}{\Gamma}{A} & \jvfameq*{i}{\Gamma}{A}{B}\\
\jvtype*{i}{\Gamma}{A} & \jvtypeeq*{i}{\Gamma}{A}{B}\\
\jvterm*{i}{\Gamma}{A}{x} & \jvtermeq*{i}{\Gamma}{A}{x}{y}.
\end{align*}

We have the following basic inference rules that relate types and families:

\begin{small}
\begin{align*}
& \inference
  {\jvtype{i}{\Gamma}{A}}
  {\jvfam{i}{\Gamma}{A}}
& & \inference
    {\jvtypeeq{i}{\Gamma}{A}{B}}
    {\jvfameq{i}{\Gamma}{A}{B}}\\
& \inference
  {\jvtype{i}{\Gamma}{A}
   \jvfameq{i}{\Gamma}{A}{B}}
  {\jvtype{i}{\Gamma}{B}}
& & \inference
    {\jvtype{i}{\Gamma}{A}
     \jvfameq{i}{\Gamma}{A}{B}}
    {\jvtypeeq{i}{\Gamma}{A}{B}}
\end{align*}
\end{small}

\subsection{The basic rules for judgmental equality}
The rules for judgmental equality establish that it is an equivalence relation.
\bgroup\small
\begin{align*}
& \inference
  {\jvctx{\Gamma}}
  {\jvctxeq{\Gamma}{\Gamma}} 
& & \inference
    {\jvctxeq{\Gamma}{\Delta}}
    {\jvctxeq{\Delta}{\Gamma}} 
& & \inference
    {\jvctxeq{\Gamma}{\Delta}
     \jvctxeq{\Delta}{\greek{E}}}
    {\jvctxeq{\Gamma}{\greek{E}}}\\
& \inference
  {\jvfam{i}{\Gamma}{A}}
  {\jvfameq{i}{\Gamma}{A}{A}} 
& & \inference
    {\jvfameq{i}{\Gamma}{A}{B}}
    {\jvfameq{i}{\Gamma}{B}{A}}
& & \inference
    {\jvfameq{i}{\Gamma}{A}{B}
     \jvfameq{i}{\Gamma}{B}{C}}
    {\jvfameq{i}{\Gamma}{A}{C}}\\
& \inference
  {\jvterm{i}{\Gamma}{A}{x}}
  {\jvtermeq{i}{\Gamma}{A}{x}{x}}
& & \inference
    {\jvtermeq{i}{\Gamma}{A}{x}{y}}
    {\jvtermeq{i}{\Gamma}{A}{y}{x}}
& & \inference
    {\jvtermeq{i}{\Gamma}{A}{x}{y}
     \jvtermeq{i}{\Gamma}{A}{y}{z}}
    {\jvtermeq{i}{\Gamma}{A}{x}{z}}
\end{align*}
\egroup

The following convertibility rules are responsible for the strictness
of judgmental equality, which sets it apart from equivalences or identifications:

\begin{align*}
& \inference
  {\jvctxeq{\Gamma}{\Delta}
   \jvfam{i}{\Gamma}{A}}
  {\jvfam{i}{\Delta}{A}}
& & \inference
    {\jvctxeq{\Gamma}{\Delta}
     \jvfameq{i}{\Gamma}{A}{B}}
    {\jvfameq{i}{\Delta}{A}{B}}\\
& \inference
  {\jvctxeq{\Gamma}{\Delta}
   \jvterm{i}{\Gamma}{A}{x}}
  {\jvterm{i}{\Delta}{A}{x}}
& & \inference
    {\jvctxeq{\Gamma}{\Delta}
     \jvtermeq{i}{\Gamma}{A}{x}{y}}
    {\jvtermeq{i}{\Delta}{A}{x}{y}}\\
& \inference
  {\jvfameq{i}{\Gamma}{A}{B}
   \jvterm{i}{\Gamma}{A}{x}}
  {\jvterm{i}{\Gamma}{B}{x}}
& & \inference
    {\jvfameq{i}{\Gamma}{A}{B}
     \jvtermeq{i}{\Gamma}{A}{x}{y}}
    {\jvtermeq{i}{\Gamma}{B}{x}{y}}
\end{align*}

\subsection{The empty context}
The empty context looks a bit strange when we explicitly denote the terms. But
we will not do so anymore after this subsection.

\begin{align}
& \inference
  {}
  {\jctx{\emptyc}}\\
& \inference
  {\jctx{\Gamma}}
  {\jvfam{i}{\Gamma}{\emptyf[\Gamma]}}\\
& \inference
  {\jctx{\Gamma}}
  {\jvterm{i}{\Gamma}{\emptyf[\Gamma]}{\emptytm[\Gamma]}}\\
& \inference
  {\jvterm{i}{\Gamma}{\emptyf[\Gamma]}{x}}
  {\jvtermeq{i}{\Gamma}{\emptyf[\Gamma]}{x}{\emptytm[\Gamma]}}
\end{align}

Moreover, if $\Gamma$ is a context family over the
empty context, then $\Gamma$ is a context and every context is a context
family over the empty context. Note that this allows us to speak
of terms of contexts too.

\begin{align}
& \inference
  {\jctx{\Gamma}}
  {\jvfam{\nameless}{\emptyc}{\Gamma}} 
& & \inference
    {\jvfam{\nameless}{\emptyc}{\Gamma}}
    {\jctx{\Gamma}}\\
& \inference
  {\jctxeq{\Gamma}{\Delta}}
  {\jvfameq{\nameless}{\emptyc}{\Gamma}{\Delta}}
& & \inference
    {\jvfameq{\nameless}{\emptyc}{\Gamma}{\Delta}}
    {\jctxeq{\Gamma}{\Delta}}
\end{align}

\subsubsection{The empty context is compatible with itslef}
The empty context $\emptyc$ may be considered as a family of contexts over the empty
context. When we do this, we get $\emptyf[\emptyc]$.
\begin{equation}
\inference
  {}
  {\jvfameq{\nameless}{\emptyc}{\emptyc}{\emptyf[\emptyc]}}
\end{equation}
In the future, we shall denote $\emptyf[\Gamma]$ by $\emptyf$. The above rule
guarantees that this will not cause confusion. Likewise, we shall denote
$\emptytm[\Gamma]$ by $\emptytm$.

\subsection{Extension}
We introduce extension which not only extends a context $\Gamma$ and a family
$A$ over it to a context $\ctxext{\Gamma}{A}$, but which also extends a family $A$
in context $\Gamma$ and a family $P$ over it to a family $\ctxext{A}{P}$ over context
$\Gamma$. We do this to ensure that all of type theory can be done in a context.
For instance, we could say (1) that a context in context $\Gamma$ is the same thing
as a family over $\Gamma$; (2) When $A$ is a context in this sense, a family over
$A$ is the same thing as a family $P$ over $\ctxext{\Gamma}{A}$ and 
(3) when $P$ is a family over $A$ in this sense, a term of $P$ keeps its original meaning.

\begin{align}
& \inference
  {\jvfam{i}{\Gamma}{A}}
  {\jvfamcombi{{i}{x}}{{\Gamma}{A}}{P}}
& & \inference
    {\jctxeq{\Gamma}{\Delta}
     \jfameq{\Gamma}{A}{B}}
    {\jctxeq{\ctxext{\Gamma}{A}}{\ctxext{\Delta}{B}}}\\
& \inference
  {\jfam{{\Gamma}{A}}{P}}
  {\jfam{\Gamma}{\ctxext{A}{P}}}
& & \inference
    {\jfameq{\Gamma}{A}{B}
     \jfameq{{\Gamma}{A}}{P}{Q}}
    {\jfameq{\Gamma}{\ctxext{A}{P}}{\ctxext{B}{Q}}}
\end{align}

\subsubsection{Extension is compatible with the empty context}
The following rule asserts that extension by $\emptyc$ leaves the contexts unchanged.
\begin{align}
& \inference
  {\jctx{\Gamma}}
  {\jctxeq{\ctxext{\emptyc}{\Gamma}}{\Gamma}}\\
& \inference
  {\jctx{\Gamma}}
  {\jctxeq{\ctxext{\Gamma}{\emptyf}}{\Gamma}}\\
& \inference
  {\jfam{\Gamma}{A}}
  {\jfameq{\Gamma}{\ctxext{\emptyf}{A}}{A}}
\end{align}

\subsubsection{Extension is compatible with itself}\label{comp-ee}
The inference rules asserting that extension is compatible with itself assert
that contexts are unstructured lists of type declarations. This rule is
unavoidable if we want that for a family $A$ in context $\Gamma$, a family over
$A$ is the same thing as a family over $\ctxext{\Gamma}{A}$.

\begin{align}
& \inference
  {\jfam{\Gamma}{A}
   \jfam{{\Gamma}{A}}{P}}
  {\jctxeq{\ctxext{{\Gamma}{A}}{P}}{\ctxext{\Gamma}{{A}{P}}}}\\
& \inference
  {\jfam{{\Gamma}{A}}{P}
   \jfam{{{\Gamma}{A}}{P}}{Q}}
  {\jfameq{\Gamma}{\ctxext{{A}{P}}{Q}}{\ctxext{A}{{P}{Q}}}}
\end{align}

\section{Chalmers type theory}
In this section we will present the type theory that Coquand and Dybjer are using.
It is a weak type theory (I think), with not so many operations and judgmental equalities.
We will show how every formula of Chalmers type theory can be interpreted in
structural type theory.

### 2014-03-06 article-models-tt.tex
worked on morphisms over morphisms

@@
 which is pronounced as `$F$ is a morphism from $P$ to $Q$ over $f$ in context
-$\Gamma$, to be the judgment
+$\Gamma$', to be the judgment
 \begin{equation*}
@@
 
-We can also define composition for this version of morphism and formulate
-compatibility rules for them.
+\begin{rmk}
+The judgment $\jfhom{\Gamma}{A}{B}{f}{P}{Q}{F}$ means the same thing as
+\begin{equation*}
+\jhom{{\Gamma}{A}}{P}{\jcomp{A}{f}{Q}}{F}.
+\end{equation*}
+\end{rmk}
+
+Suppose we have morphisms $\jhom{\Gamma}{A}{B}{f}$ and $\jhom{\Gamma}{B}{C}{g}$
+and that we have the morphisms $\jfhom{\Gamma}{A}{B}{f}{P}{Q}{F}$ and
+$\jfhom{\Gamma}{B}{C}{g}{Q}{R}{G}$ over them. Then we have
+\begin{equation*}
+\jhom
+  {{\Gamma}{A}}
+  {\jcomp{A}{f}{Q}}
+  {\jcomp{A}{f}{{B}{g}{R}}}
+  {\unfold{\jcomp{A}{f}{G}}}
+\end{equation*}
+Because we also have $\jhom{{\Gamma}{A}}{P}{\jcomp{A}{f}{Q}}{F}$, we have the
+composition
+\begin{equation*}
+\jhom
+  {{\Gamma}{A}}
+  {P}
+  {\jcomp{A}{f}{{B}{g}{R}}}
+  {\jcomp{P}{F}{\unfold{\jcomp{A}{f}{G}}}}.
+\end{equation*}
+Because of 
+the judgmental equality $\jcomp{A}{f}{{B}{g}{R}}\jdeq
+\jcomp{A}{{A}{f}{g}}{R}$, it follows that 
+$\jcomp{P}{F}{\unfold{\jcomp{A}{f}{G}}}$ is a morphism from $P$ to $R$ over
+$\jcomp{A}{f}{g}$. We make the following definition:
+
+\begin{defn}
+Let $\jhom{\Gamma}{A}{B}{f}$ and $\jhom{\Gamma}{B}{C}{g}$
+be morphisms and let $\jfhom{\Gamma}{A}{B}{f}{P}{Q}{F}$ and
+$\jfhom{\Gamma}{B}{C}{g}{Q}{R}{G}$ be morphisms over them. Then we define
+\begin{equation*}
+\jfhomdefn
+  {\Gamma}
+  {A}
+  {C}
+  {\jcomp{A}{f}{g}}
+  {P}
+  {R}
+  {\jfcomp{A}{f}{P}{F}{G}}
+  {\unfold{\jfcomp{A}{f}{P}{F}{G}}}.
+\end{equation*}
+\end{defn}
+
+This composition is also judgmentally associative.
 

### 2014-03-06 article-models-ttmodels.tex
worked on morphisms over morphisms

@@
 
+The model $\subst{\tfemp{A}}{\mtm{A}}$ is going to be initial in the category of
+all models over $A$. That means: whenever $P$ is a family over $\subst{\tfemp{A}}
+{\mtm{A}}$ there will be a term of $P$. This means in particular that
+$\subst{\tfemp{\emptyf}}{\mtm{\emptyf}}$ is going to be $\mctx$.
+
 \subsubsection{Families of weakenings}
@@
 \subsubsection{Weakening followed by substitution is the identity}
-The weakening morphism $\tfwk{A}^0$ gives a morphism
+The weakening morphism $\tfwk{A}^0$ gives a weakened morphism
 \begin{equation*}
@@
       {\ctxwk{\mtm{A}}{{\mfam{A}}{\mfam{A}}}}
-      {\ctxwk{\mtm{A}}{\tfwk{A}^0}}
+      {\jcomp{}{\ctxwk{\mtm{A}}{\tfwk{A}^0}}{\tfsubst{A}^0}}
       {\idfunc[\ctxwk{\mtm{A}}{{\mfam{A}}{\mfam{A}}}]}
@@
 \end{equation*}
+Likewise, the weakening morphism $\tfwk{A}^1$ gives a weakened morphism
+\begin{equation*}
+\jfhom
+  {{{{\Gamma}{A}}{\mfam{A}}}{\mtm{A}}}
+  {\ctxwk{\mtm{A}}{{\mfam{A}}{\mfam{A}}}}
+  {\ctxwk{\mtm{A}}{\mfam{\mfam{A}}}}
+  {\ctxwk{\mtm{A}}{\tfwk{A}^0}}
+  {\ctxwk{\mtm{A}}{\ctxwk{\mfam{A}}{\mtm{A}}}}
+  {\ctxwk{\mtm{A}}{\mtm{\mfam{\mfam{A}}}}}
+  {\ctxwk{\mtm{A}}{\tfwk{A}^1}}
+\end{equation*}
+which can be composed with the the morphism $\tfsubst{A}^1$. We require that
+this composition is the identity function on 
+$\ctxwk{\mtm{A}}{\ctxwk{\mfam{A}}{\mtm{A}}}$. Note that this composition gives
+a morphism over $\jcomp{}{\ctxwk{\mtm{A}}{\tfwk{A}^0}}{\tfsubst{A}^0}$, which
+is the identity function on $\ctxwk{\mtm{A}}{{\mfam{A}}{\mfam{A}}}$; so in fact
+it is an ordinary morphism in the context
+$\ctxext{{{{\Gamma}{A}}{\mfam{A}}}{\mtm{A}}}{\ctxwk{\mtm{A}}{{\mfam{A}}{\mfam{A}}}}$.
+Thus the inference rule we require to become valid is:
+\begin{equation*}
+\inference
+  { \jfam{\Gamma}{A}
+    }
+  { \jhomeq
+      {{{{{\Gamma}{A}}{\mfam{A}}}{\mtm{A}}}{\ctxwk{\mtm{A}}{{\mfam{A}}{\mfam{A}}}}}
+      {\ctxwk{\mtm{A}}{\ctxwk{\mfam{A}}{\mtm{A}}}}
+      {\ctxwk{\mtm{A}}{\ctxwk{\mfam{A}}{\mtm{A}}}}
+      {\jcomp{}{\ctxwk{\mtm{A}}{\tfwk{A}^1}}{\tfsubst{A}^1}}
+      {\idfunc[\ctxwk{\mtm{A}}{\ctxwk{\mfam{A}}{\mtm{A}}}]}
+    }
+\end{equation*}
+
+\subsubsection{Weakening by a family is precomposing with extension}
+We seem to need the following rule:
+\begin{equation*}
+\inference
+  { \jfam{{\Gamma}{A}}{P}
+    }
+  { \jfameq
+      {{{\Gamma}{A}}{\mfam{A}}}
+      {\ctxwk{\mfam{A}}{P}}
+      {\jcomp{}{\tfext{A}}{P}}
+    }
+\end{equation*}
+I don't know very well how to motivate requiring this rule. Maybe it's provable?
 
@@
 
-\subsubsection{Extension acts as the identity functor on families}
+\subsection{Compatibility properties of arbitrary terms}
+
+\subsubsection{Every term is compatible with the empty context}
 \begin{equation*}
-df
+\inference
+  { \jterm{{\Gamma}{A}}{P}{f}
+    }
+  { \jtermeq
+      {\Gamma}
+      {\subst{\tfemp{A}}{P}}
+      {\subst{\tfemp{A}}{f}}
+      {\tfemp{\subst{\tfemp{A}}{P}}}
+    }
 \end{equation*}
 
-\subsubsection{Extension acts as the identity functor on terms}
+\subsubsection{Every term is compatible with extension}
+Let $\jterm{{\Gamma}{A}}{P}{f}$. The rule we're about to explain is that $f$
+commutes with extension.
 
-\subsubsection{Extensions of extensions}
+We have the morphism
+$\jhom{{\Gamma}{A}}{\mfam{A}}{\subst{f}{\mfam{P}}}{\mfam{f}}$. Let's unfold
+to remind ourselves what this means:
+\begin{equation*}
+\unfold{\jhom{{\Gamma}{A}}{\mfam{A}}{\subst{f}{\mfam{P}}}{\mfam{f}}}
+\end{equation*}
+Now note that we have the judgmental equality
+\begin{equation*}
+  \ctxwk{\mfam{A}}{\subst{f}{\mfam{P}}}
+  \jdeq
+  \subst{\ctxwk{\mfam{A}}{f}}{\ctxwk{\mfam{A}}{\mfam{P}}}
+\end{equation*}
+and we see that we get the term
+$ \tmext
+    {\ctxwk{\mfam{A}}{P}}
+    {\ctxwk{\mfam{A}}{\mfam{P}}}
+    {\ctxwk{\mfam{A}}{f}}
+    {\mfam{f}}
+  $
+of the family $\ctxwk{\mfam{A}}{\ctxext{P}{\mfam{P}}}$. We can substitute this
+term in the morphism
+\begin{equation*}
+\jhom
+  {{{\Gamma}{A}}{\mfam{A}}}
+  {\ctxwk{\mfam{A}}{\ctxext{P}{\mfam{P}}}}
+  {\ctxwk{\mfam{A}}{P}}
+  {\ctxwk{\mfam{A}}{\tfext{P}}}
+\end{equation*}
+to obtain the term 
+$ \jterm
+    {{{\Gamma}{A}}{\mfam{A}}}
+    {\ctxwk{\mfam{A}}{P}}
+    { \subst
+        { \tmext
+            {\ctxwk{\mfam{A}}{P}}
+            {\ctxwk{\mfam{A}}{\mfam{P}}}
+            {\ctxwk{\mfam{A}}{f}}
+            {\mfam{f}}
+          }
+        { \ctxwk{\mfam{A}}{\tfext{P}}
+          }
+      }.
+  $
+We can also compose $\tfext{A}$ with $f$ to obtain the term
+\begin{equation*}
+\jterm{{{\Gamma}{A}}{\mfam{f}}}{\jcomp{}{\tfext{A}}{P}}{\jcomp{}{\tfext{A}}{f}}
+\end{equation*}
+Note that we have (?) the judgmental equality $\jcomp{}{\tfext{A}}{P}
+\jdeq\ctxwk{\mfam{A}}{P}$ and therefore
+we can require:
+\begin{equation*}
+\inference
+  { \jterm{{\Gamma}{A}}{P}{f}
+    }
+  { \jtermeq
+       {{{\Gamma}{A}}{\mfam{A}}}
+       {\ctxwk{\mfam{A}}{P}}
+       { \subst
+           { \tmext
+               {\ctxwk{\mfam{A}}{P}}
+               {\ctxwk{\mfam{A}}{\mfam{P}}}
+               {\ctxwk{\mfam{A}}{f}}
+               {\mfam{f}}
+             }
+           { \ctxwk{\mfam{A}}{\tfext{P}}
+             }
+         }
+       {\jcomp{}{\tfext{A}}{f}}
+    }
+\end{equation*}
 
-\subsubsection{Extensions of weakenings}
+\subsubsection{Every term is compatible with weakening}
 
-\subsubsection{Weakenings by extensions}
+\subsubsection{Every term is compatible with substitution}
 
-\subsubsection{Substitutions of extensions}
+\subsubsection{Every term preserves identity}
 

### 2014-03-06 article-models-ttmodels.tex
organizing the introduction of section 6

@@
 axiomatizing what you can do with them. We have the following ideas:
-\begin{enumerate}
+\begin{itemize}
 \item There are dependent models and sections thereof. Particular instances
@@
   functions.
-\end{enumerate}
-So the theory of models is going to be an extension of this theory with this
+\item So the theory of models is going to be an extension of this theory with this
   data. The theory we are about to describe can be seen as an elementary theory
   of the category (with families (and terms)) of categories (with families (and terms)).
-
-We would also like to remark explicitly that the valid judgments of the original type
+\item We would also like to remark explicitly that the valid judgments of the original type
   theory become contexts, families or terms, depending on what kind of judgments
@@
   order theory over the basic theory of types. 
+\end{itemize}
 
@@
 We need to do several things in this section:
-\begin{enumerate}
+\begin{itemize}
 \item postulate that ordinary extension, weakening, substitution and identity
@@
 identity functions respectively.
-\end{enumerate}
-Note that the rules for compatibility with extension are going to explain which
+\item Note that the rules for compatibility with extension are going to explain which
 model $\ctxext{A}{P}$ is by telling what the families, the terms, extension,
@@
 cases.
-
-The easiest set of compatibility rules comes with weakening. The compatibility
+\item The easiest set of compatibility rules comes with weakening. The compatibility
 rules that deal with substitution are likely going to have to do with the
 Yoneda lemma, which we should be able to implement at some point.
-
-The model $\subst{\tfemp{A}}{\mtm{A}}$ is going to be initial in the category of
+\item The model $\subst{\tfemp{A}}{\mtm{A}}$ is going to be initial in the category of
 all models over $A$. That means: whenever $P$ is a family over $\subst{\tfemp{A}}
@@
 $\subst{\tfemp{\emptyf}}{\mtm{\emptyf}}$ is going to be $\mctx$.
+\end{itemize}
 

### 2014-03-06 article-models-tt.tex
separating the constructors section into a separate file

@@
 \end{proof}
-%\end{document}
-
-\section{Type constructors in structural type theory}
-In this section we will introduce the usual type constructors to structural
-type theory. We do so with the point of view that each type constructor is a
-(class of) operations on type theory that should be compatible with extension,
-weakening, substitution and identity morphisms. Likewise, each induction
-principle of an inductively defined type constructor is going to be such an
-operator, and it is therefore also required to be compatible with extension,
-weakening, substitution and identity functions.
-
-We will first state all the rules of the individual type constructors, where
-only the $\wtypesym$-type constructor depends on the presence of dependent
-function types (or universes too?), and hence comes with the notion that $\wtypesym$
-is compatible with dependent product types and vice versa. All the other
-compatibility properties are stated in \autoref{compatibility-of-type-constructors}.
-
-\subsubsection{Issues to keep in mind}
-\begin{enumerate}
-\item The rule asserting that unit types are compatible with context extension makes
-no sense from the point of view that contexts are lists of types and types
-are contexts of length one; in other words, that types are indecomposable (i.e.~non-extended)
-families. Remidies:
-\begin{enumerate}
-\item This point of view is wrong and should be abandoned.
-\item We have, as Vladimir proposed, two kinds of judgmental equalities. One
-      could be used for the very strict equalities, the other could be used
-      to state the compatibility rules with. In other words, the other equality
-      is a compatibility relation. If we do that, we shouldn't require that
-      if a family is compatible with a type, then the family is a type. The
-      compatibility relation could relate things by uniqueness up to unique
-      isomorphism.
-\item Don't state such compatibility rules for the type constructors.
-\item Don't even give the unit type a dependent action.
-\end{enumerate}
-\end{enumerate}
-
-\subsection{The unit type in structural type theory}
-In this section we explore what we get if we pose compatibility conditions on
-an inductively defined unit type. We will assume that not only the unit type
-is compatible with extension, weakening, substitution and identity functions,
-but also its induction principle should be compatible with those.
-
-\begin{align*}
-& \inference
-  { \jfam{\Gamma}{A}
-    }
-  { \jtype{{\Gamma}{A}}{\unitc{A}}
-    }
-& & \inference
-    { \jfameq{\Gamma}{A}{A'}
-      }
-    { \jtypeeq
-        {{\Gamma}{A}}
-        {\unitc{A}}
-        {\unitc{A'}}
-      }
-  \\
-& \inference
-    { \jfam{\Gamma}{A}
-      }
-    { \jterm{{\Gamma}{A}}{\unitc{A}}{\unitct{A}}
-      }
-& & \inference
-    { \jfameq{\Gamma}{A}{A'}
-      }
-    { \jtermeq{{\Gamma}{A}}{\unitc{A}}{\unitct{A}}{\unitct{A'}}
-      }
-  \\
-& \inference
-  { \jfam{{\Gamma}{A}}{P}
-    }
-  { \jtype{{{{\Gamma}{A}}{P}}{\ctxwk{P}{\unitc{A}}}}{\unitf{A}{P}}
-    }
-& & \inference
-    { \jfameq{{\Gamma}{A}}{P}{P'}
-      }
-    { \jtypeeq
-        {{{{\Gamma}{A}}{P}}{\ctxwk{P}{\unitc{A}}}}
-        {\unitf{A}{P}}
-        {\unitf{A}{P'}}
-      }
-  \\
-& \inference
-  { \jfam{{\Gamma}{A}}{P}
-    }
-  { \jterm
-      {{{{\Gamma}{A}}{P}}{\ctxwk{P}{\unitc{A}}}}
-      {\unitf{A}{P}}
-      {\unitft{A}{P}}
-    }
-& & \inference
-    { \jfameq{{\Gamma}{A}}{P}{P'}
-      }
-    { \jtermeq
-        {{{{\Gamma}{A}}{P}}{\ctxwk{P}{\unitc{A}}}}
-        {\unitf{A}{P}}
-        {\unitft{A}{P}}
-        {\unitft{A}{P'}}
-      }
-\end{align*}
-
-We impose the following compatibility rules for the unit type:
-
-\begin{align*}
-& \inference
-  { \jfam{{\Gamma}{A}}{P}
-    }
-  { \jtypeeq
-      {{{\Gamma}{A}}{P}}
-      {\ctxext{\ctxwk{P}{\unitc{A}}}{\unitf{A}{P}}}
-      {\unitc{{\Gamma}{A}}}
-    }\\
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{\Gamma}{B}
-    }
-  { \jtypeeq{{{\Gamma}{A}}{\ctxwk{A}{B}}}{}{}
-    }  
-\end{align*}
-
-\subsection{The empty type}
-
-\subsection{The natural numbers}
-
-\subsection{Dependent pair types}
-
-\subsection{Identity types}
-
-\subsection{Dependent function types}
-
-\subsection{$\wtypesym$-types}
-
-\subsection{Universes}
-
-\subsection{The compatibility of the type constructors with each other}
-\label{compatibility-of-type-constructors}

### 2014-03-06 article-models-ttconstr.tex
separating the constructors section into a separate file

\section{Type constructors in structural type theory}
In this section we will introduce the usual type constructors to structural
type theory. We do so with the point of view that each type constructor is a
(class of) operations on type theory that should be compatible with extension,
weakening, substitution and identity morphisms. Likewise, each induction
principle of an inductively defined type constructor is going to be such an
operator, and it is therefore also required to be compatible with extension,
weakening, substitution and identity functions.

We will first state all the rules of the individual type constructors, where
only the $\wtypesym$-type constructor depends on the presence of dependent
function types (or universes too?), and hence comes with the notion that $\wtypesym$
is compatible with dependent product types and vice versa. All the other
compatibility properties are stated in \autoref{compatibility-of-type-constructors}.

\subsubsection{Issues to keep in mind}
\begin{enumerate}
\item The rule asserting that unit types are compatible with context extension makes
no sense from the point of view that contexts are lists of types and types
are contexts of length one; in other words, that types are indecomposable (i.e.~non-extended)
families. Remidies:
\begin{enumerate}
\item This point of view is wrong and should be abandoned.
\item We have, as Vladimir proposed, two kinds of judgmental equalities. One
      could be used for the very strict equalities, the other could be used
      to state the compatibility rules with. In other words, the other equality
      is a compatibility relation. If we do that, we shouldn't require that
      if a family is compatible with a type, then the family is a type. The
      compatibility relation could relate things by uniqueness up to unique
      isomorphism.
\item Don't state such compatibility rules for the type constructors.
\item Don't even give the unit type a dependent action.
\end{enumerate}
\end{enumerate}

\subsection{The unit type in structural type theory}
In this section we explore what we get if we pose compatibility conditions on
an inductively defined unit type. We will assume that not only the unit type
is compatible with extension, weakening, substitution and identity functions,
but also its induction principle should be compatible with those.

\begin{align*}
& \inference
  { \jfam{\Gamma}{A}
    }
  { \jtype{{\Gamma}{A}}{\unitc{A}}
    }
& & \inference
    { \jfameq{\Gamma}{A}{A'}
      }
    { \jtypeeq
        {{\Gamma}{A}}
        {\unitc{A}}
        {\unitc{A'}}
      }
  \\
& \inference
    { \jfam{\Gamma}{A}
      }
    { \jterm{{\Gamma}{A}}{\unitc{A}}{\unitct{A}}
      }
& & \inference
    { \jfameq{\Gamma}{A}{A'}
      }
    { \jtermeq{{\Gamma}{A}}{\unitc{A}}{\unitct{A}}{\unitct{A'}}
      }
  \\
& \inference
  { \jfam{{\Gamma}{A}}{P}
    }
  { \jtype{{{{\Gamma}{A}}{P}}{\ctxwk{P}{\unitc{A}}}}{\unitf{A}{P}}
    }
& & \inference
    { \jfameq{{\Gamma}{A}}{P}{P'}
      }
    { \jtypeeq
        {{{{\Gamma}{A}}{P}}{\ctxwk{P}{\unitc{A}}}}
        {\unitf{A}{P}}
        {\unitf{A}{P'}}
      }
  \\
& \inference
  { \jfam{{\Gamma}{A}}{P}
    }
  { \jterm
      {{{{\Gamma}{A}}{P}}{\ctxwk{P}{\unitc{A}}}}
      {\unitf{A}{P}}
      {\unitft{A}{P}}
    }
& & \inference
    { \jfameq{{\Gamma}{A}}{P}{P'}
      }
    { \jtermeq
        {{{{\Gamma}{A}}{P}}{\ctxwk{P}{\unitc{A}}}}
        {\unitf{A}{P}}
        {\unitft{A}{P}}
        {\unitft{A}{P'}}
      }
\end{align*}

We impose the following compatibility rules for the unit type:

\begin{align*}
& \inference
  { \jfam{{\Gamma}{A}}{P}
    }
  { \jtypeeq
      {{{\Gamma}{A}}{P}}
      {\ctxext{\ctxwk{P}{\unitc{A}}}{\unitf{A}{P}}}
      {\unitc{{\Gamma}{A}}}
    }\\
& \inference
  { \jfam{\Gamma}{A}
    \jfam{\Gamma}{B}
    }
  { \jtypeeq{{{\Gamma}{A}}{\ctxwk{A}{B}}}{}{}
    }  
\end{align*}

\subsection{The empty type}

\subsection{The natural numbers}

\subsection{Dependent pair types}

\subsection{Identity types}

\subsection{Dependent function types}

\subsection{$\wtypesym$-types}

\subsection{Universes}

\subsection{The compatibility of the type constructors with each other}
\label{compatibility-of-type-constructors}

### 2014-03-07 article-models-ttmodels.tex
reorganizing compatibility rules

@@
 \subsection{Compatibility properties of arbitrary terms}
+In this subsection we will state the inference rules that assert that every
+term $\jterm{{\Gamma}{A}}{P}{f}$ acts functorially. That means, every term
+preserves the empty context, extension, weakening, substitution and the
+identity functions. Moreover, there will be several inference rules
+involving the behavior of $\mfam{f}$ and $\mtm{f}$.
 
@@
 \end{equation*}
+
+\subsubsection{The action of the action on families of a term is the action
+on contexts of that term}
+Consider a term $\jterm{{\Gamma}{A}}{P}{f}$. Then we have the term
+\begin{equation*}
+\jterm
+  {{\Gamma}{\subst{\tfemp{A}}{\mfam{A}}}}
+  {\subst{\tfemp{A}}{\ctxwk{\mfam{A}}{\subst{f}{\mfam{P}}}}}
+  {\subst{\tfemp{A}}{\mfam{f}}}
+\end{equation*}
+We have required that $\subst{\tfemp{A}}{\mfam{A}}\jdeq A$. Also, we have the
+judgmental equalities
+\begin{align*}
+\subst{\tfemp{A}}{\ctxwk{\mfam{A}}{\subst{f}{\mfam{P}}}}
+& \jdeq
+  \ctxwk{\subst{\tfemp{A}}{\mfam{A}}}{\subst{\tfemp{A}}{{f}{\mfam{P}}}}
+  \\
+& \jdeq
+  \ctxwk{A}{\subst{{\tfemp{A}}{f}}{{\tfemp{A}}{\mfam{P}}}}
+  \\
+& \jdeq
+  \ctxwk{A}{\subst{\tfemp{\subst{\tfemp{A}}{P}}}{{\tfemp{A}}{\mfam{P}}}}
+  \\
+& \jdeq
+  \ctxwk{A}{\subst{{\tfemp{A}}{\tfemp{P}}}{{\tfemp{A}}{\mfam{P}}}}
+  \\
+& \jdeq
+  \ctxwk{A}{\subst{\tfemp{A}}{{\tfemp{P}}{\mfam{P}}}}
+  \\
+& \jdeq
+  \ctxwk{A}{\subst{\tfemp{A}}{P}}
+\end{align*}
+If we require also the following inference rule, we would get $\mfam{A}\jdeq
+\ctxwk{A}{A}$.
+\begin{equation*}
+\inference
+  { \jfam{{\Gamma}{A}}{P}
+    }
+  { \jfameq{{\Gamma}{A}}{P}{\ctxwk{A}{\subst{\tfemp{A}}{P}}}
+    }
+\end{equation*}
+This rule would say that a family $P$ of models over $A$ is determined by the
+model $\subst{\tfemp{A}}{P}$ and that the only possible families of models
+are the constant families.
 

### 2014-03-12 article-models-ttmodels.tex
added item to intro

@@
   order theory over the basic theory of types.
+\item Martin Escardo and Mike Shulman have been promoting the use of 
+  inductive-inductive definitions for internal models. What we're doing here looks
+  different, because we're writing down a type theory for internal models, but it
+  might not be that different at all. The type theory can be seen as the theory
+  of (dependent) algebras over the inductive-inductively defined model and the
+  terms of this type theory are the (dependent) algebra homomorphisms. The
+  asserted initial object of our type theory is the inductive-inductively defined
+  type. 
 \end{itemize}
@@
 
+\subsubsection{The action on families of families of a term is the action on
+families over extensions}
+Let $\jterm{{\Gamma}{A}}{P}{f}$. Then we have the term
+\begin{equation*}
+\jterm{{{\Gamma}{A}}{\mfam{A}}}{\ctxwk{\mfam{A}}{\subst{f}{\mfam{P}}}}{\mfam{f}}
+\end{equation*}
+and we have the term
+\begin{equation*}
+\jterm
+  { {{{\Gamma}{A}}{\mfam{A}}}{\mfam{\mfam{A}}}
+    }
+  { \ctxwk
+      {\mfam{\mfam{A}}}
+      {\subst{\mfam{f}}{\mfam{\ctxwk{\mfam{A}}{\subst{f}{\mfam{P}}}}}}
+    }
+  { \mfam{\mfam{f}}
+    }.
+\end{equation*}
+We have the judgmental equality $\mfam{\mfam{A}}\jdeq
+\jcomp{}{\tfext{A}}{\mfam{A}}$; in this subsubsection we wish to establish a
+similar judgmental equality explaining what $\mfam{\mfam{f}}$ is. Note that
+we have the judgmental equalities
+\begin{align*}
+\ctxwk
+  {\mfam{\mfam{A}}}
+  {\subst{\mfam{f}}{\mfam{\ctxwk{\mfam{A}}{\subst{f}{\mfam{P}}}}}}
+& \jdeq
+  \ctxwk
+    {\jcomp{}{\tfext{A}}{\mfam{A}}}
+    {\subst{\mfam{f}}{\mfam{\ctxwk{\mfam{A}}{\subst{f}{\mfam{P}}}}}}
+  \\
+& \jdeq
+  \ctxwk
+    {\unfold{\jcomp{\ctxext{A}{\mfam{A}}}{\tfext{A}}{\mfam{A}}}}
+    {\subst{\mfam{f}}{\mfam{\ctxwk{\mfam{A}}{\subst{f}{\mfam{P}}}}}}
+  \\
+& \jdeq
+  \ctxwk
+    {\subst{\tfext{A}}{\ctxwk{\mfam{A}}{{A}{\mfam{A}}}}}
+    {\subst{\mfam{f}}{\mfam{\ctxwk{\mfam{A}}{\subst{f}{\mfam{P}}}}}}
+  \\
+& \jdeq
+  \ctxwk
+    { \subst{\tfext{A}}{\ctxwk{\mfam{A}}{{A}{\mfam{A}}}}
+      }
+    { \subst
+        { \tfext{A}
+          }
+        { \ctxwk
+            {A}
+            {\subst{\mfam{f}}{\mfam{\ctxwk{\mfam{A}}{\subst{f}{\mfam{P}}}}}}}
+      }
+  \\
+& \jdeq
+  \subst
+    { \tfext{A}
+      }
+    { \ctxwk
+        { {\mfam{A}}{{A}{\mfam{A}}}
+          }
+        { {A}
+          {\subst{\mfam{f}}{\mfam{\ctxwk{\mfam{A}}{\subst{f}{\mfam{P}}}}}}
+          }
+      }
+  \\
+& \jdeq \unfold{\jcomp{{A}{\mfam{A}}}{\tfext{A}}{\mfam{\ctxwk{\mfam{A}}{\subst{f}{\mfam{P}}}}}}
+\end{align*}
+
 \subsubsection{Every term is compatible with the empty context}
@@
 on contexts of that term}
+\emph{In this subsubsection we attempted to establish a rule asserting that
+$\subst{\tfemp{A}}{\mfam{f}}\jdeq f$ for any $\jterm{{\Gamma}{A}}{P}{f}$. This
+is however not true.}
+
 Consider a term $\jterm{{\Gamma}{A}}{P}{f}$. Then we have the term
@@
 \subsubsection{Every term is compatible with weakening}
+Let $\jterm{{\Gamma}{A}}{P}{f}$ be a term. We will establish an inference rule
+asserting that $f$ commutes with weakening.
 

### 2014-03-17 article-models-tt.tex
adding ideas to the introduction of section 4

@@
 \section{Type theory before type constructors}\label{tt}
-In this section we give a description of type theory before basic
-constructors and with explicit extension, weakening and substitution. We will
-formulate the type theory in such a way that contexts aren't lists of variable
-declarations. The main reason is that we don't see variable declarations in the
+In this section we give a description of dependent type theory before type
+constructors with explicit operations of extension, weakening and substitution. We will
+formulate the type theory in such a way that contexts aren't defined
+to be lists of variable
+declarations. We rather take a syntactic approach in our treatment of contexts,
+axiomatizing only what you can do with them and allowing for the freedom to
+interpret the axiom in any desired way. 
+The main reason is that we don't see variable declarations in the internal
 models either. This way we also set out to a more algebraic approach of type
 theory and higher category theory. Thirdly, we will not have to be burdened with
-comments about variables being bounded or not, or fresh or not occuring at all.
-
-We will develop a theory which is in fact a theory of contexts and families of
-contexts where a family of context can be a type, behind which we have the
-intuition that it is an atomic or irreducible family of contexts. We get immediately
-that every construction can be made in an arbitrary context.
-
-It should be noted that this is just a prototype description. It may be the case
-that various other elements have to be added yet, such as terms
-$\jterm{{\Gamma}{\subst{x}{P}}}{\ctxwk{\subst{x}{P}}{\ctxext{A}{P}}}{i}$
-including the fiber $\subst{x}{P}$ into the extended family $\ctxext{A}{P}$ over
-$\Gamma$. Moreover, it hasn't been thoroughly tested whether the theory presented
-here lives up to its promise.
+superficial comments about variables being bounded or not, or fresh or free or 
+not occuring at all.
+
+The theory we will develop here is a theory of contexts, families of contexts
+and terms thereof. Families of contexts could have been named dependent contexts
+or telescopes.
+
+In our treatment, a context can be the empty context or it can be a binary
+planar tree of which `the leaves are (families of) contexts'. 
+The judgmental equality relation on contexts is an equivalence relation which 
+expresses that binary planar
+trees of contexts are judgmentally equal if their leaves are, taking only the
+order of the leaves into account \emph{and not the actual shape of the three}.
+The intuition behind this equivalence relation is indeed that unstructured
+(i.e.~unbracketed) lists such as the lists of variable declarations which
+usually appear in type theoretical syntax, may be regarded as contexts.
+
+Besides contexts, families and terms there will also be a notion of `type in
+a context'. Only a family of contexts over a context $\Gamma$ is eligible to
+be a type in the context $\Gamma$ and we have the intuition that `being a type'
+expresses the property of `being an atomic or irreducible family of contexts'.
+
+We will formulate fairly strict rules governing the judgmental equalities,
+expressing that extension, weakening and substitution are combatible with
+each other in a judgmental manner. This does not, however, diminish the role
+of isomorphisms or of homotopies could play in the theory once identity types
+are added. Indeed, types could still have non-trivial identity relations and
+the category of types in a certain context could genuinely display higher
+categorical structure, or so we conjecture.
 

### 2014-03-17 article-models-tt.tex
small improvement

@@
 expresses that binary planar
-trees of contexts are judgmentally equal if their leaves are, taking only the
-order of the leaves into account \emph{and not the actual shape of the three}.
+trees of contexts are judgmentally equal if their leaves are, taking only
+(the isomorphism class of) 
+the order of the leaves into account \emph{and not the actual shape of the three}.
 The intuition behind this equivalence relation is indeed that unstructured

### 2014-03-19 article-models-tt.tex
added a sentence
new section title

@@
-\section{Type theory before type constructors}\label{tt}
+\section{Operational syntactics for type theory before type constructors}\label{tt}
 In this section we give a description of dependent type theory before type
@@
 expresses the property of `being an atomic or irreducible family of contexts'.
+We will not axiomatize that every context is either judgmentally equal to the
+empty context or judgmentally equal to the extension of a context by a type,
+although it is certainly worth investigating the class of contexts which fall
+under this category.
 

### 2014-03-19 article-models-ttmodels.tex
adding ideas to the introduction of section 4
deleting a duplicated label

@@
 
-\subsubsection{The action of the action on families of a term is the action
+\subsubsection{The action on families of the action on families of a term is the action
 on contexts of that term}
@@
 
-\subsubsection{Extension is compatible with itself}\label{comp-ee}
+\subsubsection{Extension is compatible with itself}
 The inference rules asserting that extension is compatible with itself assert

### 2014-03-22 article-models-tt.tex
tweaking the positions of the QED signs
small changes

@@
   \ctxwk{A}{C}.
+  \qedhere
 \end{align*}
@@
   \ctxwk{A}{z}.
+  \qedhere
 \end{align*}
@@
   \subst{{x}{f}}{g}.
+  \qedhere
 \end{align*}
@@
   \idfunc[\ctxext{A}{P}].
+  \qedhere
 \end{align*}
@@
   u.
+  \qedhere
 \end{align*}
@@
   \tmext{\ctxwk{A}{Q}}{\ctxwk{A}{R}}{\ctxwk{A}{g}}{\ctxwk{A}{t}}
+  \qedhere
 \end{align*}
@@
   \tmext{\subst{x}{Q}}{\subst{x}{R}}{\subst{x}{g}}{\subst{x}{t}}.
+  \qedhere
 \end{align*}
@@
   \tmext{\ctxwk{A}{C}}{\ctxwk{A}{R}}{\jcomp{A}{f}{g}}{\jcomp{A}{f}{w}}.
+  \qedhere
 \end{align*}
 \end{proof}
+
+\subsection{Judgmental trivial cofibrations}
+\begin{defn}
+Let $\jhom{\Gamma}{A}{B}{f}$ be a context morphism. We say that $f$ is a 
+\emph{judgmental trivial cofibration} if the following inference rules are valid:
+\begin{align*}
+& \inference
+  { \jfam{{\Gamma}{B}}{Q}
+    \jterm{{\Gamma}{A}}{\jcomp{A}{f}{Q}}{g}
+    }
+  { \jterm{{\Gamma}{B}}{Q}{\jtcext{g}}
+    }
+& & \inference
+  { \jfam{{\Gamma}{B}}{Q}
+    \jterm{{\Gamma}{A}}{\jcomp{A}{f}{Q}}{g}
+    }
+  { \jtermeq{{\Gamma}{A}}{\jcomp{A}{f}{Q}}{\jcomp{A}{f}{\jtcext{g}}}{g}
+    }
+\end{align*}
+\end{defn}
+
+When $f$ is a context morphism from $A$ to $B$ in context $\Gamma$, 
+finding a term of a family $Q$ over $\ctxext{\Gamma}{B}$ can be accomplished
+by finding a term of the family $\jcomp{A}{f}{Q}$ over $\ctxext{\Gamma}{A}$.
+Judgmental trivial cofibrations come with a computation rule. Thus, term
+constructors of inductively defined types are going to be the major source of
+examples of judgmental trivial cofibrations. Since identity functions behave
+so nicely, they are judgmental trivial cofibrations too.
+
+\begin{lem}
+The identity function
+$\jhom{\Gamma}{A}{A}{\idfunc[A]}$ is a judgmental trivial cofibration
+for each family $A$ of contexts over $\Gamma$
+\end{lem}

### 2014-03-22 article-models-ttconstr.tex
adding ideas to the introduction of section 4
small changes

@@
 \subsection{Dependent function types}
+The dependent function type constructor bundles up the terms of a family $P$
+of contexts over $\ctxext{\Gamma}{A}$ and produces a family $\mprd{A}{P}$ of
+contexts over $\Gamma$. Its terms are copies of those of $P$. The operation
+$\mprd{A}{\blank}$ also acts on terms of families -- this is called
+$\lambda$-abstraction.
+
+The dependent function type constructor not only acts on families $P$ of
+contexts over $\ctxext{\Gamma}{A}$, but it also sends families $Q$ of contexts
+over $\ctxext{{\Gamma}{A}}{P}$ to families $\mprd{A}{Q}$ over $\ctxext{\Gamma}
+{\mprd{A}{P}}$. Moreover, we will also have a version of $\lambda$-abstraction
+for terms of such families $Q$.
+
+\begin{align*}
+& \inference
+    { \jfam{{\Gamma}{A}}{P}
+      }
+    { \jfam{\Gamma}{\mprd{A}{P}}
+      }
+& & \inference
+      { \jfameq{\Gamma}{A}{A'}
+        \jfameq{{\Gamma}{A}}{P}{P'}
+        }
+      { \jfameq{\Gamma}{\mprd{A}{P}}{\mprd{A'}{P'}}
+        }
+  \\
+& \inference
+    { \jterm{{\Gamma}{A}}{P}{f}
+      }
+    { \jterm{\Gamma}{\mprd{A}{P}}{\slam{A}{P}{f}}
+      }
+& & \inference
+      { \jtermeq{{\Gamma}{A}}{P}{f}{f'}
+        }
+      { \jtermeq{\Gamma}{\mprd{A}{P}}{\slam{A}{P}{f}}{\slam{A}{P}{f'}}
+        }
+  \\
+& \inference
+    { \jterm{\Gamma}{\mprd{A}{P}}{g}
+      }
+    { \jterm{{\Gamma}{A}}{P}{\sev{g}}
+      }
+& & \inference
+      { \jtermeq{\Gamma}{\mprd{A}{P}}{g}{g'}
+        }
+      { \jtermeq{{\Gamma}{A}}{P}{\sev{g}}{\sev{g'}}
+        }
+\end{align*}
+
+\begin{defn}
+Let $A$ and $B$ be families of contexts over $\Gamma$. Then we define
+\begin{equation*}
+\jfamdefn{\Gamma}{\jfun{A}{B}}{\sprd{A}{\ctxwk{A}{B}}}.
+\end{equation*}
+\end{defn}
+
+\subsubsection{Dependent products are compatible with the empty context}
+The empty context can appear in the domain and in the codomain of the dependent
+function type constructor. We have the following inference rules explaining
+what happens when the empty context appears in the domain:
+\begin{align}
+& \inference
+    { \jfam{\Gamma}{P}
+      }
+    { \jfameq{\Gamma}{\mprd{\emptyf}{P}}{P}
+      }
+  \\
+& \inference
+    { \jterm{\Gamma}{P}{f}
+      }
+    { \jtermeq{\Gamma}{P}{\slam{A}{P}{f}}{f}
+      }
+  \\
+& \inference
+    { \jterm{\Gamma}{P}{g}
+      }
+    { \jtermeq{\Gamma}{P}{\sev{g}}{g}
+      }
+\end{align}
+We have the following infernece rules explaining what happens when the empty
+context appears in the codomain:
+\begin{align}
+& \inference
+    { \jfam{\Gamma}{A}
+      }
+    { \jfameq{\Gamma}{\mprd{A}{\emptyf}}{\emptyf}
+      }
+\end{align}
+
+\subsubsection{Dependent products are compatible with extension}
+\begin{align}
+& \inference
+  { \jfam{{{{\Gamma}{A}}{B}}{P}}{Q}
+    }
+  { \jfameq
+      {\Gamma}
+      {\sprd{{A}{B}}{{P}{Q}}}
+      {\ctxext{\sprd{A}{\sprd{B}{P}}}{\sprd{A}{\sprd{B}{Q}}}}
+    }
+  \\
+& \inference
+  { \jterm{{{\Gamma}{A}}{B}}{P}{f}
+    }
+  { \jtermeq
+      {\Gamma}
+      {\sprd{A}{\sprd{B}{P}}}
+      {\slam{{A}{B}}{P}{f}}
+      {\slam{A}{\sprd{B}{P}}{\slam{B}{P}{f}}}
+    }
+\end{align}
+\subsubsection{Dependent products are compatible with weakening}
+\subsubsection{Dependent products are compatible with substitution}?
+\subsubsection{Dependent products are compatible with the identity functions}?
 

### 2014-03-26 article-models-tt.tex
changes in notation, most notably the change from identity functions to identity terms

@@
 
-\subsection{Composition and identity functions}
+\subsection{Composition and identity terms}
 \subsubsection{The defining property of weakening}
@@
 
-\subsubsection{Identity functions}
+\subsubsection{identity terms}
 Without a rule explicitly asserting the existence of an identity morphism we don't
@@
     }
-  { \jhom{\Gamma}{A}{A}{\idfunc[A]}
+  { \jhom{\Gamma}{A}{A}{\idtm{A}}
     }
@@
       }
-    { \jhomeq{\Gamma}{A}{A}{\idfunc[A]}{\idfunc[A']}
+    { \jhomeq{\Gamma}{A}{A}{\idtm{A}}{\idtm{A'}}
       }
 \end{align}
-Identity functions are determined by their behavior with respect to substitution combined with
-weakening. The identity functions will also be subject to compatibility rules.
+Identity terms are determined by their behavior with respect to substitution combined with
+weakening. The identity terms will also be subject to compatibility rules.
 \begin{align}
@@
     }
-  { \jtermeq{\Gamma}{A}{\subst{x}{\idfunc[A]}}{x}
+  { \jtermeq{\Gamma}{A}{\subst{x}{\idtm{A}}}{x}
     }
@@
     }
-  { \jfameq{{\Gamma}{A}}{\subst{\idfunc[A]}{\ctxwk{A}{P}}}{P}
+  { \jfameq{{\Gamma}{A}}{\subst{\idtm{A}}{\ctxwk{A}{P}}}{P}
     }
@@
       {{\Gamma}{A}}
-      {\subst{\idfunc[A]}{\ctxwk{{A}{A}}{P}}}
+      {\subst{\idtm{A}}{\ctxwk{{A}{A}}{P}}}
       {P}
@@
       {P}
-      {\subst{\idfunc[A]}{\ctxwk{A}{f}}}
+      {\subst{\idtm{A}}{\ctxwk{A}{f}}}
       {f}
@@
       {P}
-      {\subst{\idfunc[A]}{\ctxwk{{A}{A}}{f}}}
+      {\subst{\idtm{A}}{\ctxwk{{A}{A}}{f}}}
       {f}
@@
     }
-  { \jhomeq{\Gamma}{A}{B}{\jcomp{A}{f}{\idfunc[B]}}{f}
+  { \jhomeq{\Gamma}{A}{B}{\jcomp{A}{f}{\idtm{B}}}{f}
     }
@@
 
-We won't state a compatibility rule stating that the identity function is
+We won't state a compatibility rule stating that the identity term is
 compatible with extension because we will be able to prove that. Instead, we
-will just state the compatibility rules for the identity function combined with
+will just state the compatibility rules for the identity term combined with
 weakening and with substitution.
 
-The identity function of a weakened family is the weakened identity function:
+The identity term of a weakened family is the weakened identity term:
 \begin{equation}\label{idfunc-wk-comp}
@@
       {\ctxwk{A}{B}}
-      {\ctxwk{A}{\idfunc[B]}}
-      {\idfunc[\ctxwk{A}{B}]}
+      {\ctxwk{A}{\idtm{B}}}
+      {\idtm{\ctxwk{A}{B}}}
     }
@@
 
-The identity function of a substituted family is the substitution of the identity function
+The identity term of a substituted family is the substitution of the identity term
 \begin{equation}\label{idfunc-subst-comp}
@@
       {\subst{x}{P}}
-      {\subst{x}{\idfunc[P]}}
-      {\idfunc[\subst{x}{P}]}
+      {\subst{x}{\idtm{P}}}
+      {\idtm{\subst{x}{P}}}
     }
@@
 Now let $\jfam{{\Gamma}{A}}{P}$ and $\jfam{{\Gamma}{A}}{Q}$ be families. A
-morphism from $P$ to $Q$ over the identity function $\idfunc[A]$ in context
+morphism from $P$ to $Q$ over the identity term $\idtm{A}$ in context
 $\Gamma$ is the same thing as a morphism from $P$ to $Q$ in context
@@
     \jfam{{\Gamma}{A}}{Q}
-    \jfhom{\Gamma}{A}{A}{\idfunc[A]}{P}{Q}{f}
+    \jfhom{\Gamma}{A}{A}{\idtm{A}}{P}{Q}{f}
     }
@@
     }
-  { \jfhom{\Gamma}{A}{A}{\idfunc[A]}{P}{Q}{f}
+  { \jfhom{\Gamma}{A}{A}{\idtm{A}}{P}{Q}{f}
     }
@@
 If we unfold the judgments $\jhom{{\Gamma}{A}}{P}{Q}{f}$ and
-$\jfhom{\Gamma}{A}{A}{\idfunc[A]}{P}{Q}{f}$, we get the judgments
+$\jfhom{\Gamma}{A}{A}{\idtm{A}}{P}{Q}{f}$, we get the judgments
 \begin{align*}
@@
   \\
-& \unfold{\jfhom{\Gamma}{A}{A}{\idfunc[A]}{P}{Q}{f}}
+& \unfold{\jfhom{\Gamma}{A}{A}{\idtm{A}}{P}{Q}{f}}
 \end{align*}
 respectively. Therefore, we only need to verify that
-$\ctxwk{P}{\subst{\idfunc[A]}{\ctxwk{A}{Q}}}\jdeq\ctxwk{P}{Q}$, which is indeed
+$\ctxwk{P}{\subst{\idtm{A}}{\ctxwk{A}{Q}}}\jdeq\ctxwk{P}{Q}$, which is indeed
 the case by \autoref{idfunc-wk-defn}.
@@
 context $\ctxext{{\Gamma}{A}}{P}$, the
-pairing function could just be defined as $\idfunc[\ctxext{A}{P}]$.
+pairing function could just be defined as $\idtm{\ctxext{A}{P}}$.
 
@@
   \subst
-    {{u}{{x}{\idfunc[\ctxext{A}{P}]}}}
+    {{u}{{x}{\idtm{\ctxext{A}{P}}}}}
     {Q}
@@
   \subst
-    {{u}{{x}{\idfunc[\ctxext{A}{P}]}}}
+    {{u}{{x}{\idtm{\ctxext{A}{P}}}}}
     {{x}{\ctxwk{A}{Q}}}
@@
   \subst
-    {{u}{{x}{\idfunc[\ctxext{A}{P}]}}}
+    {{u}{{x}{\idtm{\ctxext{A}{P}}}}}
     {{u}{\ctxwk{\subst{x}{P}}{\subst{x}{\ctxwk{A}{Q}}}}}
@@
   \subst
-    {{u}{{x}{\idfunc[\ctxext{A}{P}]}}}
+    {{u}{{x}{\idtm{\ctxext{A}{P}}}}}
     {{u}{{x}{\ctxwk{P}{{A}{Q}}}}}
@@
     {u}
-    {{{x}{\idfunc[\ctxext{A}{P}]}}{{x}{\ctxwk{{P}{{A}{Q}}}}}}
+    {{{x}{\idtm{\ctxext{A}{P}}}}{{x}{\ctxwk{{P}{{A}{Q}}}}}}
   \tag{by \autoref{comp-ss-f}}\\
@@
     {u}
-    {{x}{{\idfunc[\ctxext{A}{P}]}{\ctxwk{P}{{A}{Q}}}}}
+    {{x}{{\idtm{\ctxext{A}{P}}}{\ctxwk{P}{{A}{Q}}}}}
   \tag{by \autoref{comp-ss-f}}\\
@@
     {u}
-    {{x}{{\idfunc[\ctxext{A}{P}]}{\ctxwk{\ctxext{A}{P}}{Q}}}}
+    {{x}{{\idtm{\ctxext{A}{P}}}{\ctxwk{\ctxext{A}{P}}{Q}}}}
   \tag{by \autoref{comp-ew-f}}\\
@@
 
-We have seen above that the pairing function into $\ctxext{A}{P}$ is just the identity function on
+We have seen above that the pairing function into $\ctxext{A}{P}$ is just the identity term on
 $\ctxext{A}{P}$. To analyze the pairing functin a little further, we will also
 need the projection maps from $\ctxext{A}{P}$ to $A$ and from $\ctxext{A}{P}$
-to $P$. We will now define these and see that the identity function of an
+to $P$. We will now define these and see that the identity term of an
 extended family is the extension (or pairing) of the identity
@@
 \end{align*}
-We have the term $\jterm{{\Gamma}{{A}{P}}}{\ctxwk{P}{A}}{\ctxwk{P}{\idfunc[A]}}$.
-Thus we need to find out what $\subst{\ctxwk{P}{\idfunc[A]}}{\ctxwk{P}{{A}{P}}}$ is:
-\begin{align*}
-\subst{\ctxwk{P}{\idfunc[A]}}{\ctxwk{P}{{A}{P}}} 
-& \jdeq 
-  \ctxwk{P}{\subst{\idfunc[A]}{\ctxwk{A}{P}}}
+We have the term $\jterm{{\Gamma}{{A}{P}}}{\ctxwk{P}{A}}{\ctxwk{P}{\idtm{A}}}$.
+Thus we need to find out what $\subst{\ctxwk{P}{\idtm{A}}}{\ctxwk{P}{{A}{P}}}$ is:
+\begin{align*}
+\subst{\ctxwk{P}{\idtm{A}}}{\ctxwk{P}{{A}{P}}} 
+& \jdeq 
+  \ctxwk{P}{\subst{\idtm{A}}{\ctxwk{A}{P}}}
   \\
@@
 \end{align*}
-where we find the term $\idfunc[P]$. Therefore we define:
+where we find the term $\idtm{P}$. Therefore we define:
 
@@
 
-We find the following inference rule, which expresses that the identity function
+We find the following inference rule, which expresses that the identity term
 is compatible with extension:
@@
       {{A}{P}}{{A}{P}}
-      {\idfunc[\ctxext{A}{P}]}
+      {\idtm{\ctxext{A}{P}}}
       { \tmext
@@
   \subst
-    { \idfunc[P]
-      }
-    { {\ctxwk{P}{\idfunc[A]}}
-      {\idfunc[\ctxwk{\ctxext{A}{P}}{\ctxext{A}{P}}]}
+    { \idtm{P}
+      }
+    { {\ctxwk{P}{\idtm{A}}}
+      {\idtm{\ctxwk{\ctxext{A}{P}}{\ctxext{A}{P}}}}
       }
@@
   \subst
-    { \idfunc[P]
-      }
-    { {\ctxwk{P}{\idfunc[A]}}
-      {\ctxwk{\ctxext{A}{P}}{\idfunc[\ctxext{A}{P}]}}
+    { \idtm{P}
+      }
+    { {\ctxwk{P}{\idtm{A}}}
+      {\ctxwk{\ctxext{A}{P}}{\idtm{\ctxext{A}{P}}}}
       }
@@
   \subst
-    { \idfunc[P]
-      }
-    { {\ctxwk{P}{\idfunc[A]}}
-      {\ctxwk{P}{{A}{\idfunc[\ctxext{A}{P}]}}}
+    { \idtm{P}
+      }
+    { {\ctxwk{P}{\idtm{A}}}
+      {\ctxwk{P}{{A}{\idtm{\ctxext{A}{P}}}}}
       }
@@
   \subst
-    { \idfunc[P]
+    { \idtm{P}
       }
@@
       { \subst
-        {\idfunc[A]}
-        {\ctxwk{A}{\idfunc[\ctxext{A}{P}]}}
+        {\idtm{A}}
+        {\ctxwk{A}{\idtm{\ctxext{A}{P}}}}
         }
@@
   \subst
-    { \idfunc[A]
-      }
-    { \ctxwk{A}{\idfunc[\ctxext{A}{P}]}
-      }
-  \\
-& \jdeq 
-  \idfunc[\ctxext{A}{P}].
+    { \idtm{A}
+      }
+    { \ctxwk{A}{\idtm{\ctxext{A}{P}}}
+      }
+  \\
+& \jdeq 
+  \idtm{\ctxext{A}{P}}.
   \qedhere
@@
 & \jdeq 
-  \subst{\ctxwk{P}{\idfunc[A]}}{\ctxwk{\ctxext{A}{P}}{Q}} 
+  \subst{\ctxwk{P}{\idtm{A}}}{\ctxwk{\ctxext{A}{P}}{Q}} 
   \tag{by definition}\\
 & \jdeq 
-  \subst{\ctxwk{P}{\idfunc[A]}}{\ctxwk{P}{{A}{Q}}} 
+  \subst{\ctxwk{P}{\idtm{A}}}{\ctxwk{P}{{A}{Q}}} 
   \tag{by \autoref{comp-ww-f}}\\
 & \jdeq 
-  \ctxwk{P}{\subst{\idfunc[A]}{\ctxwk{A}{Q}}} 
+  \ctxwk{P}{\subst{\idtm{A}}{\ctxwk{A}{Q}}} 
   \tag{by \autoref{comp-ws-f}}\\
@@
 & \jdeq 
-  \subst{w}{\idfunc[\ctxext{A}{P}]} 
+  \subst{w}{\idtm{\ctxext{A}{P}}} 
   \tag{by \autoref{idfunc-subst-defn}}\\
@@
     { w}
-    { { \idfunc[P]}
-      { { \ctxwk{P}{\idfunc[A]}}
-        { \idfunc[\ctxwk{\ctxext{A}{P}}{\ctxext{A}{P}}]}
+    { { \idtm{P}}
+      { { \ctxwk{P}{\idtm{A}}}
+        { \idtm{\ctxwk{\ctxext{A}{P}}{\ctxext{A}{P}}}}
         }
@@
     { {w}
-      {\idfunc[P]}
+      {\idtm{P}}
       }
     { {w}
-      { { \ctxwk{P}{\idfunc[A]}
-          }
-        { \idfunc[\ctxwk{\ctxext{A}{P}}{\ctxext{A}{P}}]
+      { { \ctxwk{P}{\idtm{A}}
+          }
+        { \idtm{\ctxwk{\ctxext{A}{P}}{\ctxext{A}{P}}}
           }
@@
     { {w}
-      {\idfunc[P]}
+      {\idtm{P}}
       }
     { { {w}
-        {\ctxwk{P}{\idfunc[A]}}
+        {\ctxwk{P}{\idtm{A}}}
         }
       { {w}
-        {\idfunc[\ctxwk{\ctxext{A}{P}}{\ctxext{A}{P}}]}
+        {\idtm{\ctxwk{\ctxext{A}{P}}{\ctxext{A}{P}}}}
         }
@@
     { {w}
-      {\idfunc[P]}
+      {\idtm{P}}
       }
     { { {w}
-        {\ctxwk{P}{\idfunc[A]}}
+        {\ctxwk{P}{\idtm{A}}}
         }
       { {w}
-        {\ctxwk{\ctxext{A}{P}}{\idfunc[\ctxext{A}{P}]}}
+        {\ctxwk{\ctxext{A}{P}}{\idtm{\ctxext{A}{P}}}}
         }
@@
     { {w}
-      {\idfunc[P]}
+      {\idtm{P}}
       }
     { { {w}
-        {\ctxwk{P}{\idfunc[A]}}
-        }
-      { \idfunc[\ctxext{A}{P}]
+        {\ctxwk{P}{\idtm{A}}}
+        }
+      { \idtm{\ctxext{A}{P}}
         }
@@
 & \jdeq 
-  \subst{\tmext{A}{P}{x}{u}}{\ctxwk{P}{\idfunc[A]}}
-  \\
-& \jdeq 
-  \subst{u}{{x}{\ctxwk{P}{\idfunc[A]}}} 
-  \\
-& \jdeq 
-  \subst{u}{\ctxwk{\subst{x}{P}}{\subst{x}{\idfunc[A]}}}
-  \\
-& \jdeq 
-  \subst{x}{\idfunc[A]}
+  \subst{\tmext{A}{P}{x}{u}}{\ctxwk{P}{\idtm{A}}}
+  \\
+& \jdeq 
+  \subst{u}{{x}{\ctxwk{P}{\idtm{A}}}} 
+  \\
+& \jdeq 
+  \subst{u}{\ctxwk{\subst{x}{P}}{\subst{x}{\idtm{A}}}}
+  \\
+& \jdeq 
+  \subst{x}{\idtm{A}}
   \\
@@
 & \jdeq 
-  \subst{\tmext{A}{P}{x}{u}}{\idfunc[P]}
-  \\
-& \jdeq 
-  \subst{u}{{x}{\idfunc[P]}}
-  \\
-& \jdeq 
-  \subst{u}{\idfunc[\subst{x}{P}]}
+  \subst{\tmext{A}{P}{x}{u}}{\idtm{P}}
+  \\
+& \jdeq 
+  \subst{u}{{x}{\idtm{P}}}
+  \\
+& \jdeq 
+  \subst{u}{\idtm{\subst{x}{P}}}
   \\
@@
     {\tmext{\subst{x}{P}}{\subst{x}{Q}}{u}{v}}
-    {{x}{\idfunc[\ctxext{A}{{P}{Q}}]}}
-  \\
-& \jdeq 
-  \subst{v}{{u}{{x}{\idfunc[\ctxext{A}{{P}{Q}}]}}}
-  \\
-& \jdeq 
-  \subst{v}{{u}{{x}{\idfunc[\ctxext{{A}{P}}{Q}]}}}
-  \\
-& \jdeq 
-  \subst{v}{{\tmext{A}{P}{x}{u}}{\idfunc[\ctxext{{A}{P}}{Q}]}}
+    {{x}{\idtm{\ctxext{A}{{P}{Q}}}}}
+  \\
+& \jdeq 
+  \subst{v}{{u}{{x}{\idtm{\ctxext{A}{{P}{Q}}}}}}
+  \\
+& \jdeq 
+  \subst{v}{{u}{{x}{\idtm{\ctxext{{A}{P}}{Q}}}}}
+  \\
+& \jdeq 
+  \subst{v}{{\tmext{A}{P}{x}{u}}{\idtm{\ctxext{{A}{P}}{Q}}}}
   \\
@@
 & \jdeq 
-  \ctxwk{A}{\subst{t}{{g}{\idfunc[\ctxext{Q}{R}]}}}
-  \\
-& \jdeq 
-  \subst{\ctxwk{A}{t}}{\ctxwk{A}{\subst{g}{\idfunc[\ctxext{Q}{R}]}}}
-  \\
-& \jdeq 
-  \subst{\ctxwk{A}{t}}{{\ctxwk{A}{g}}{\ctxwk{A}{\idfunc[\ctxext{Q}{R}]}}}
-  \\
-& \jdeq 
-  \subst{\ctxwk{A}{t}}{{\ctxwk{A}{g}}{\idfunc[\ctxwk{A}{\ctxext{Q}{R}}]}}
-  \\
-& \jdeq 
-  \subst{\ctxwk{A}{t}}{{\ctxwk{A}{g}}{\idfunc[\ctxext{\ctxwk{A}{Q}}{\ctxwk{A}{R}}]}}
+  \ctxwk{A}{\subst{t}{{g}{\idtm{\ctxext{Q}{R}}}}}
+  \\
+& \jdeq 
+  \subst{\ctxwk{A}{t}}{\ctxwk{A}{\subst{g}{\idtm{\ctxext{Q}{R}}}}}
+  \\
+& \jdeq 
+  \subst{\ctxwk{A}{t}}{{\ctxwk{A}{g}}{\ctxwk{A}{\idtm{\ctxext{Q}{R}}}}}
+  \\
+& \jdeq 
+  \subst{\ctxwk{A}{t}}{{\ctxwk{A}{g}}{\idtm{\ctxwk{A}{\ctxext{Q}{R}}}}}
+  \\
+& \jdeq 
+  \subst{\ctxwk{A}{t}}{{\ctxwk{A}{g}}{\idtm{\ctxext{\ctxwk{A}{Q}}{\ctxwk{A}{R}}}}}
   \\
@@
 & \jdeq 
-  \subst{x}{{t}{{g}{\idfunc[\ctxext{Q}{R}]}}}
-  \\
-& \jdeq 
-  \subst{{x}{t}}{{x}{{g}{\idfunc[\ctxext{Q}{R}]}}}
-  \\
-& \jdeq 
-  \subst{{x}{t}}{{{x}{g}}{{x}{\idfunc[\ctxext{Q}{R}]}}}
-  \\
-& \jdeq 
-  \subst{{x}{t}}{{{x}{g}}{\idfunc[\subst{x}{\ctxext{Q}{R}}]}}
-  \\
-& \jdeq 
-  \subst{{x}{t}}{{{x}{g}}{\idfunc[\ctxext{\subst{x}{Q}}{\subst{x}{R}}]}}
+  \subst{x}{{t}{{g}{\idtm{\ctxext{Q}{R}}}}}
+  \\
+& \jdeq 
+  \subst{{x}{t}}{{x}{{g}{\idtm{\ctxext{Q}{R}}}}}
+  \\
+& \jdeq 
+  \subst{{x}{t}}{{{x}{g}}{{x}{\idtm{\ctxext{Q}{R}}}}}
+  \\
+& \jdeq 
+  \subst{{x}{t}}{{{x}{g}}{\idtm{\subst{x}{\ctxext{Q}{R}}}}}
+  \\
+& \jdeq 
+  \subst{{x}{t}}{{{x}{g}}{\idtm{\ctxext{\subst{x}{Q}}{\subst{x}{R}}}}}
   \\
@@
 constructors of inductively defined types are going to be the major source of
-examples of judgmental trivial cofibrations. Since identity functions behave
+examples of judgmental trivial cofibrations. Since identity terms behave
 so nicely, they are judgmental trivial cofibrations too.
@@
 \begin{lem}
-The identity function
-$\jhom{\Gamma}{A}{A}{\idfunc[A]}$ is a judgmental trivial cofibration
+The identity term
+$\jhom{\Gamma}{A}{A}{\idtm{A}}$ is a judgmental trivial cofibration
 for each family $A$ of contexts over $\Gamma$

### 2014-03-26 article-models-ttmodels.tex
changes in notation, most notably the change from identity functions to identity terms

@@
   When the situation requires clarity --  for instance when extensions
-  are involved -- we will write $\mfam{f}^\Gamma$ and $\mtm{f}^\Gamma$ to make
+  are involved -- we will write $\mfam[\Gamma]{f}$ and $\mtm[\Gamma]{f}$ to make
   what we regard as the context explicit.}
@@
 \intertext{%
-  The identity functions are coded by}
+  The identity terms are coded by}
 & \inference
@@
     }
-  \tag{identity functions}
+  \tag{identity terms}
 \end{align*}
@@
 \item postulate that each of the model constructors is compatible with ordinary
-extension, weakening, substitution and identity functions. Actually, we want
+extension, weakening, substitution and identity terms. Actually, we want
 that every term of this type theory is compatible with those. I.e.~every term
@@
 so that they come to model ordinary extension, weakening, substitution and
-identity functions respectively.
+identity terms respectively.
 \item Note that the rules for compatibility with extension are going to explain which
 model $\ctxext{A}{P}$ is by telling what the families, the terms, extension,
-weakening, substitution and identity functions are. Likewise the rules for
+weakening, substitution and identity terms are. Likewise the rules for
 compatibility with weakening and substitution do this for their respective
@@
   {\subst{\tfext{A}}{\mfam{\ctxwk{\ctxext{A}{\mfam{A}}}{A}}}}
-  {\mfam{\tfext{A}}^{\ctxext{\Gamma}{A}}}
+  {\mfam[\ctxext{\Gamma}{A}]{\tfext{A}}}
 \end{equation*}
@@
 \end{equation*}
-Therefore, we can compare the term $\mfam{\tfext{A}}^{\ctxext{\Gamma}{A}}$
-to the identity function $\idfunc[\jcomp{}{\tfext{A}}{\mfam{A}}]$. 
+Therefore, we can compare the term $\mfam[\ctxext{\Gamma}{A}]{\tfext{A}}$
+to the identity term $\idfunc[\jcomp{}{\tfext{A}}{\mfam{A}}]$. 
 The compatibility
@@
       {\jcomp{}{\tfext{A}}{\mfam{A}}}
-      {\mfam{\tfext{A}}^{\ctxext{\Gamma}{A}}}
+      {\mfam[\ctxext{\Gamma}{A}]{\tfext{A}}}
       {\idfunc[\jcomp{}{\tfext{A}}{\mfam{A}}]}
@@
   {\subst{\tfext{A}}{\mfam{\ctxwk{\ctxext{A}{\mfam{A}}}{A}}}}
-  {\mfam{\tfext{A}}^{\ctxext{\Gamma}{A}}}
+  {\mfam[\ctxext{\Gamma}{A}]{\tfext{A}}}
   {\mtm{\mfam{A}}}
   {\subst{\tfext{A}}{\mtm{\ctxwk{\ctxext{A}{\mfam{A}}}{A}}}}
-  {\mtm{\tfext{A}}^{\ctxext{\Gamma}{A}}}
+  {\mtm[\ctxext{\Gamma}{A}]{\tfext{A}}}
 \end{equation*}
@@
 \jdeq\jcomp{}{\tfext{A}}{\mtm{A}}$. Thirdly, we have the judgmental equality
-$\mfam{\tfext{A}}^{\ctxext{\Gamma}{A}}\jdeq\idfunc[\jcomp{}{\tfext{A}}{\mfam{A}}]$.
+$\mfam[\ctxext{\Gamma}{A}]{\tfext{A}}\jdeq\idfunc[\jcomp{}{\tfext{A}}{\mfam{A}}]$.
 We may combine these three facts with \autoref{hom-over-id-is-hom} to see that
@@
   {\jcomp{}{\tfext{A}}{\mtm{A}}}
-  {\mtm{\tfext{A}}^{\ctxext{\Gamma}{A}}}
+  {\mtm[\ctxext{\Gamma}{A}]{\tfext{A}}}
 \end{equation*}
-Now we see that we can require that $\mtm{\tfext{A}}^{\ctxext{\Gamma}{A}}$ is
-the identity function on $\jcomp{}{\tfext{A}}{\mtm{A}}$:
+Now we see that we can require that $\mtm[\ctxext{\Gamma}{A}]{\tfext{A}}$ is
+the identity term on $\jcomp{}{\tfext{A}}{\mtm{A}}$:
 \begin{equation*}
@@
     {\jcomp{}{\tfext{A}}{\mtm{A}}}
-    {\mtm{\tfext{A}}^{\ctxext{\Gamma}{A}}}
+    {\mtm[\ctxext{\Gamma}{A}]{\tfext{A}}}
     {\idfunc[\jcomp{}{\tfext{A}}{\mtm{A}}]}
@@
 which can be composed with the morphism $\tfsubst{A}^0$. We require that this
-composition is the identity function on $\ctxwk{\mtm{A}}{{\mfam{A}}{\mfam{A}}}$:
+composition is the identity term on $\ctxwk{\mtm{A}}{{\mfam{A}}{\mfam{A}}}$:
 \begin{equation*}
@@
 which can be composed with the the morphism $\tfsubst{A}^1$. We require that
-this composition is the identity function on 
+this composition is the identity term on 
 $\ctxwk{\mtm{A}}{\ctxwk{\mfam{A}}{\mtm{A}}}$. Note that this composition gives
 a morphism over $\jcomp{}{\ctxwk{\mtm{A}}{\tfwk{A}^0}}{\tfsubst{A}^0}$, which
-is the identity function on $\ctxwk{\mtm{A}}{{\mfam{A}}{\mfam{A}}}$; so in fact
+is the identity term on $\ctxwk{\mtm{A}}{{\mfam{A}}{\mfam{A}}}$; so in fact
 it is an ordinary morphism in the context
@@
 preserves the empty context, extension, weakening, substitution and the
-identity functions. Moreover, there will be several inference rules
+identity terms. Moreover, there will be several inference rules
 involving the behavior of $\mfam{f}$ and $\mtm{f}$.

### 2014-03-26 article-models-ttmodels.tex
new compatibility rules

@@
 
+\subsubsection{Families of extended models}
+Suppose we have a model $\ctxext{A}{P}$ in context $\Gamma$. In this section we
+will figure out what the families of this model are. This is a bit tricky, so
+we first treat the case where $P\jdeq\mfam{A}$. A context $\gamma$ in this
+model is decomposed as a pair $\tmext{A}{\mfam{A}}{\pts{\gamma}}{\edg{\gamma}}$
+which consists of a context $\pts{\gamma}$ of the model $A$ and a context
+$\edg{\gamma}$ of the model $\mfam{A}$ in the fiber above $\pts{\Gamma}$. In
+other words, $\edg{\gamma}$ is a family over $\pts{\gamma}$. A family $\beta$
+over $\gamma$ is then a pair $\tmext{}{}{\pts{\beta}}{\edg{\beta}}$ consisting
+of a family $\pts{\beta}$ over $\pts{\gamma}$ and a family $\edg{\beta}$ over
+the extended term $\tmext{}{}{\pts{\beta}}{\edg{\gamma}}$. 
+It is this extension which appears in the type of $\edg{\beta}$ which requires
+a workaround for general $P$.
+
+We consider the following diagram helpful:
+\begin{equation*}
+\begin{tikzcd}
+\jcomp{}{\bar{\typefont{pr}}}{{}{\tfext{A}}{\mfam{P}}}
+  \rar{}
+  \ar[fib]{d}
+  &
+  \jcomp{}{\tfext{A}}{\mfam{P}}
+    \rar{}
+    \ar[fib]{d}{}
+  &
+  \mfam{P}
+    \ar[fib]{d}{}
+  \\
+\jcomp{}{\bar{\typefont{pr}}}{{}{\tfext{A}}{P}}
+  \rar{}
+  \ar[fib]{d}{}
+  &
+  \jcomp{}{\tfext{A}}{P}
+  \rar{}
+  \ar[fib]{d}{}
+  &
+  P
+  \ar[fib]{d}{}\\
+\ctxext{{A}{P}}{\mfam{A}}
+  \ar{r}[swap]{\bar{\typefont{pr}}}
+  &
+  \ctxext{A}{\mfam{A}}
+  \ar{r}[swap]{\tfext{A}}
+  &
+  A
+\end{tikzcd}
+\end{equation*}
+Here $\bar{\typefont{pr}}\jdeq\tmext{}{}{\cprojfstf{A}{P}}{\idtm{\mfam{A}}}$,
+the unmarked horizontal arrows indicate identity terms and the vertical arrows
+merely indicate that what's above is a family depending on what's below.
+
+\subsubsection{Families of substituted modes}
+The following inference rule explains that the fiber at a term $f$ of $P$
+of the model of families of a family of models $Q$ over $\ctxext{{\Gamma}{A}}{P}$
+is judgmentally the same as the model of families of the fiber at the term $f$
+of the family $Q$. In more practical terms, suppose we have a context $\gamma$
+in $Q$ which happens to be in the fiber at $f$ of $Q$ over $P$, then a family
+over $\gamma$ according to $Q$ is the same thing as a family over $\Gamma$
+according to $\subst{f}{Q}$. The following inference rule captures this idea
+precisely:
+
+\begin{equation}
+\inference
+  { \jfam{{{\Gamma}{A}}{P}}{Q}
+    \jterm{{\Gamma}{A}}{P}{f}
+    }
+  { \jfameq
+      {{\Gamma}{A}}
+      {\subst{f}{\mfam{Q}}}
+      {\mfam{\subst{f}{Q}}}
+    }
+\end{equation}
+
+\subsubsection{Families of weakened models}
+\begin{equation*}
+\inference
+  { \jfam{\Gamma}{A}
+    \jfam{\Gamma}{B}
+    }
+  { \jfameq
+      {{{\Gamma}{A}}{\ctxwk{A}{B}}}
+      {\ctxwk{A}{\mfam{B}}}
+      {\mfam{\ctxwk{A}{B}}}
+    }
+\end{equation*}
+
+\subsubsection{The action on families of the identity term}
+Let $A$ be a model in context $\Gamma$. Then we have the identity term
+$\jhom{\Gamma}{A}{A}{\idtm{A}}$, which also has an action on families with the
+following specification:
+\begin{equation*}
+\jhom{{\Gamma}{A}}{\mfam{A}}{\subst{\idtm{A}}{\mfam{\ctxwk{A}{A}}}}{\mfam{\idtm{A}}}.
+\end{equation*}
+Note, however that we have the judgmental equalities
+\begin{equation*}
+\subst{\idtm{A}}{\mfam{\ctxwk{A}{A}}}\jdeq\subst{\idtm{A}}{\ctxwk{A}{\mfam{A}}}
+\jdeq\mfam{A},
+\end{equation*}
+so we may require that the action on families of the identity
+term $\idtm{A}$ is the identity term $\idtm{\mfam{A}}$. We get the following
+inference rule:
+\begin{equation*}
+\inference
+  { \jfam{\Gamma}{A}
+    }
+  { \jhomeq
+      {{\Gamma}{A}}
+      {\mfam{A}}
+      {\mfam{A}}
+      {\mfam{\idtm{A}}}
+      {\idtm{\mfam{A}}}
+    }
+\end{equation*}
+
 \subsubsection{Weakening by a family is precomposing with extension}

### 2014-03-27 article-models-ttmodels.tex
figuring out what families of extended models are

@@
 
-We consider the following diagram helpful:
+We consider the following dependency diagram helpful:
 \begin{equation*}
@@
 \jcomp{}{\bar{\typefont{pr}}}{{}{\tfext{A}}{\mfam{P}}}
-  \rar{}
   \ar[fib]{d}
@@
   \jcomp{}{\tfext{A}}{\mfam{P}}
-    \rar{}
     \ar[fib]{d}{}
@@
 \jcomp{}{\bar{\typefont{pr}}}{{}{\tfext{A}}{P}}
-  \rar{}
   \ar[fib]{d}{}
@@
   \jcomp{}{\tfext{A}}{P}
-  \rar{}
   \ar[fib]{d}{}
@@
   \ar[fib]{d}{}\\
-\ctxext{{A}{P}}{\mfam{A}}
+\ctxext{{A}{P}}{\ctxwk{P}{\mfam{A}}}
   \ar{r}[swap]{\bar{\typefont{pr}}}
@@
 \end{equation*}
-Here $\bar{\typefont{pr}}\jdeq\tmext{}{}{\cprojfstf{A}{P}}{\idtm{\mfam{A}}}$,
-the unmarked horizontal arrows indicate identity terms and the vertical arrows
-merely indicate that what's above is a family depending on what's below.
+Here the morphism 
+$\bar{\typefont{pr}}$ should be explained (skip ahead to 
+\autoref{barproj} for the definition). Note that we have the judgmental
+equalities
+\begin{align*}
+\ctxwk{\ctxext{{A}{P}}{\ctxwk{P}{\mfam{A}}}}{\ctxext{A}{\mfam{A}}}
+& \jdeq 
+  \ctxext
+    {\ctxwk{\ctxext{{A}{P}}{\ctxwk{P}{\mfam{A}}}}{A}}
+    {\ctxwk{\ctxext{{A}{P}}{\ctxwk{P}{\mfam{A}}}}{\mfam{A}}}
+  \\
+& \jdeq
+  \ctxext
+    {\ctxwk{{P}{\mfam{A}}}{{\ctxext{A}{P}}{A}}}
+    {\ctxwk{\ctxext{{A}{P}}{\ctxwk{P}{\mfam{A}}}}{\mfam{A}}}
+\end{align*}
+Note that we have the term $\ctxwk{{P}{\mfam{A}}}{\cprojfstf{A}{P}}$ of the
+family $\ctxwk{{P}{\mfam{A}}}{{\ctxext{A}{P}}{A}}$. Therefore, we need to
+find a term of type $\subst{\ctxwk{{P}{\mfam{A}}}{\cprojfstf{A}{P}}}
+{\ctxwk{\ctxext{{A}{P}}{\ctxwk{P}{\mfam{A}}}}{\mfam{A}}}$. Note that we have
+the judgmental equalities:
+\begin{align*}
+\subst
+  {\ctxwk{{P}{\mfam{A}}}{\cprojfstf{A}{P}}}
+  {\ctxwk{\ctxext{{A}{P}}{\ctxwk{P}{\mfam{A}}}}{\mfam{A}}}
+& \jdeq
+  \subst
+    {\ctxwk{{P}{\mfam{A}}}{\cprojfstf{A}{P}}}
+    {\ctxwk{{P}{\mfam{A}}}{{\ctxext{A}{P}}{\mfam{A}}}}
+  \\
+& \jdeq
+  \ctxwk
+    { {P}{\mfam{A}}
+      }
+    { \subst
+        {\cprojfstf{A}{P}}
+        {\ctxwk{\ctxext{A}{P}}{\mfam{A}}}
+      }
+  \\
+& \jdeq
+  \ctxwk
+    { {P}{\mfam{A}}
+      }
+    { {P}{\mfam{A}}
+      }
+  \\
+& \jdeq
+  \ctxwk{P}{{\mfam{A}}{\mfam{A}}}
+\end{align*}
+We find the term $\ctxwk{P}{\idtm{\mfam{A}}}$ here. Thus we can now define
+$\bar{\typefont{pr}}$ by:
+\begin{equation}\label{barproj}
+\jhomdefn
+  {\Gamma}
+  {{{A}{P}}{\mfam{A}}}
+  {{A}{\mfam{A}}}
+  {\bar{\typefont{pr}}}
+  {\tmext{}{}{\ctxwk{{P}{\mfam{A}}}{\cprojfstf{A}{P}}}{\ctxwk{P}{\idtm{\mfam{A}}}}}
+\end{equation}
+
+\begin{lem}
+We have the judgmental equality
+\begin{equation*}
+\jfameq
+  {{{{\Gamma}{A}}{P}}{\ctxwk{P}{\mfam{A}}}}
+  {\jcomp{}{\bar{\typefont{pr}}}{Q}}
+  {\ctxwk{P}{Q}}
+\end{equation*}
+for any family $Q$ of contexts over $\ctxext{{\Gamma}{A}}{\mfam{A}}$ 
+\end{lem}
 

### 2014-03-27 article-models-tt.tex
fixing bugs of tmext
commutative dependence diagrams

@@
       {\ctxext{{A}{P}}{Q}}
-      {\tmext{A}{{P}{Q}}{x}{{u}{v}}}
-      {\tmext{{A}{P}}{Q}{{x}{u}}{v}}
+      {\tmext{A}{{P}{Q}}{x}{{\subst{x}{P}}{\subst{x}{Q}}{u}{v}}}
+      {\tmext{{A}{P}}{Q}{{A}{P}{x}{u}}{v}}
     }
@@
 
+\subsubsection{Commuting dependence diagrams}
+Now that we have introduced the notions of morphisms, composition and identity
+terms, we can develop a diagramatic style of of displaying type dependencies
+combined with morphisms. We give an informal, metatheoretical definition of
+such diagrams by indicating what the various components mean. The definition
+is informal because we will only use such diagrams occasionally to provide a
+graphical indication of the situation in which we're working.
+
+\begin{defn}
+A diagram is said to be a \emph{dependency diagram in context $\Gamma$}
+if it is built up according to the following steps:
+\begin{itemize}
+\item An ordinary arrow 
+\begin{equation*}
+\begin{tikzcd}
+A \ar{r}{f} & B
+\end{tikzcd}
+\end{equation*}
+between two families $A$ and $B$ of contexts over $\Gamma$ indicates that
+$f$ is a morphism from $A$ to $B$ in context $\Gamma$, i.e.~that we have the
+judgment $\jhom{\Gamma}{A}{B}{f}$.
+\item The set of double-headed arrows must form a forest where the root of
+each tree is a family of contexts over $\Gamma$. 
+\item A sequence of double-headed 
+arrows
+\begin{equation*}
+\begin{tikzcd}
+P_{n+1} \ar[fib]{r} & \cdots \ar[fib]{r} & P_1 \ar[fib]{r} & A
+\end{tikzcd}
+\end{equation*}
+indicates that $P_1$ is a family of contexts over $\ctxext{\Gamma}{A}$, that
+$P_2$ is a family of contexts over $\ctxext{{\Gamma}{A}}{P_1}$, etcetera.
+\item A ladder of double-headed arrows
+\begin{equation*}
+\begin{tikzcd}
+P_{n+1} \ar{r}{F_{n+1}} \ar[fib]{d} & Q_{n+1} \ar[fib]{d}\\
+\vdots \ar[fib]{d} & \vdots \ar[fib]{d}\\
+P_1 \ar{r}{F_1} \ar[fib]{d} & Q_1 \ar[fib]{d}\\
+A \ar{r}{f} & B
+\end{tikzcd}
+\end{equation*}
+indicates that $F_1$ is a morphism from $P_1$ to $Q_1$ \emph{over} $f$,
+i.e.~that the judgment $\jfhom{\Gamma}{A}{B}{f}{P_1}{Q_1}{F_1}$ holds, that
+$F_2$ is a morphism from $P_2$ to $Q_2$ over
+the morphism $\tmext{}{}{\ctxwk{P_1}{f}}{F_1}$ from $\ctxext{A}{P_1}$ to
+$\ctxext{B}{Q_1}$, etcetera.
+\end{itemize}
+Such a diagram is said to be commutative if the subdiagram consisting of only
+the normal headed arrows is commutative in the usual sense (using judgmental
+equality). Note that the ladders are inherently commutative.
+\end{defn}
+
+\subsubsection{Another special case of projections}
+In this subsubsection we investigate the special case of a projection which
+appears as a morphism from $\ctxext{{A}{P}}{\ctxwk{P}{Q}}$ to $\ctxext{A}{Q}$
+in context $\Gamma$, where we assume to have the families 
+$\jfam{{\Gamma}{A}}{P}$ and $\jfam{{\Gamma}{A}}{Q}$.
+
+Note that we have the judgmental
+equalities
+\begin{align*}
+\ctxwk{\ctxext{{A}{P}}{\ctxwk{P}{\mfam{A}}}}{\ctxext{A}{\mfam{A}}}
+& \jdeq 
+  \ctxext
+    {\ctxwk{\ctxext{{A}{P}}{\ctxwk{P}{\mfam{A}}}}{A}}
+    {\ctxwk{\ctxext{{A}{P}}{\ctxwk{P}{\mfam{A}}}}{\mfam{A}}}
+  \\
+& \jdeq
+  \ctxext
+    {\ctxwk{{P}{\mfam{A}}}{{\ctxext{A}{P}}{A}}}
+    {\ctxwk{\ctxext{{A}{P}}{\ctxwk{P}{\mfam{A}}}}{\mfam{A}}}
+\end{align*}
+Note that we have the term $\ctxwk{{P}{\mfam{A}}}{\cprojfstf{A}{P}}$ of the
+family $\ctxwk{{P}{\mfam{A}}}{{\ctxext{A}{P}}{A}}$. Therefore, we need to
+find a term of type $\subst{\ctxwk{{P}{\mfam{A}}}{\cprojfstf{A}{P}}}
+{\ctxwk{\ctxext{{A}{P}}{\ctxwk{P}{\mfam{A}}}}{\mfam{A}}}$. Note that we have
+the judgmental equalities:
+\begin{align*}
+\subst
+  {\ctxwk{{P}{\mfam{A}}}{\cprojfstf{A}{P}}}
+  {\ctxwk{\ctxext{{A}{P}}{\ctxwk{P}{\mfam{A}}}}{\mfam{A}}}
+& \jdeq
+  \subst
+    {\ctxwk{{P}{\mfam{A}}}{\cprojfstf{A}{P}}}
+    {\ctxwk{{P}{\mfam{A}}}{{\ctxext{A}{P}}{\mfam{A}}}}
+  \\
+& \jdeq
+  \ctxwk
+    { {P}{\mfam{A}}
+      }
+    { \subst
+        {\cprojfstf{A}{P}}
+        {\ctxwk{\ctxext{A}{P}}{\mfam{A}}}
+      }
+  \\
+& \jdeq
+  \ctxwk
+    { {P}{\mfam{A}}
+      }
+    { {P}{\mfam{A}}
+      }
+  \\
+& \jdeq
+  \ctxwk{P}{{\mfam{A}}{\mfam{A}}}
+\end{align*}
+We find the term $\ctxwk{P}{\idtm{\mfam{A}}}$ here. Thus we can now define
+$\bar{\typefont{pr}}$ by:
+\begin{equation}\label{barproj}
+\jhomdefn
+  {\Gamma}
+  {{{A}{P}}{\mfam{A}}}
+  {{A}{\mfam{A}}}
+  {\bar{\typefont{pr}}}
+  {\tmext{}{}{\ctxwk{{P}{\mfam{A}}}{\cprojfstf{A}{P}}}{\ctxwk{P}{\idtm{\mfam{A}}}}}
+\end{equation}
+
+\begin{lem}
+We have the judgmental equality
+\begin{equation*}
+\jfameq
+  {{{{\Gamma}{A}}{P}}{\ctxwk{P}{\mfam{A}}}}
+  {\jcomp{}{\bar{\typefont{pr}}}{Q}}
+  {\ctxwk{P}{Q}}
+\end{equation*}
+for any family $Q$ of contexts over $\ctxext{{\Gamma}{A}}{\mfam{A}}$ 
+\end{lem}
+
 \subsection{Judgmental trivial cofibrations}

### 2014-03-27 article-models-ttmodels.tex
commutative dependence diagrams

--- 2014-03-26 article-models-ttmodels.tex
@@
 \end{equation*}
-Here $\bar{\typefont{pr}}\jdeq\tmext{}{}{\cprojfstf{A}{P}}{\idtm{\mfam{A}}}$,
-the unmarked horizontal arrows indicate identity terms and the vertical arrows
-merely indicate that what's above is a family depending on what's below.
+Here the morphism 
+$\bar{\typefont{pr}}$ should be explained (skip ahead to 
+\autoref{barproj} for the definition).
 

### 2014-03-27 article-models-tt.tex
judgmental pullbacks

@@
 
-\subsection{Extension on terms}\label{extension-on-terms}
-In this subsection we consider the notion of extension on terms, which has now
+\subsection{Digging a little deeper in the theory of contexts, families and terms}
+
+There is a good amount of things we can already state and prove about in the
+theory we have developed so far. We will develop a bit more theory that will
+come in handy later on.
+
+\emph{This subsection contains no new assumptions.}
+
+\subsubsection{Extension on terms}\label{extension-on-terms}
+In this subsubsection we consider the notion of extension on terms, which has now
 become definable inside our theory. Moreover, every compatibility rule one may
 dream of is provable as well, using the compatibility rules we have introduced
-earlier. \emph{This subsection contains no new assumptions.}
+earlier.
 
@@
 is informal because we will only use such diagrams occasionally to provide a
-graphical indication of the situation in which we're working.
+graphical indication of the situation in which we're working. In particular,
+we will not shy away from using natural numbers and trust that the reader can
+figure out what we mean.
 
@@
 \begin{itemize}
+\item The arrows appearing in a dependency diagram are either ordinary, like the
+arrow%
+$\begin{tikzcd}[ampersand replacement = \&]
+X \ar{r} \& Y,
+\end{tikzcd}$
+or double-headed, like
+$\begin{tikzcd}[ampersand replacement = \&]
+X \ar[fib]{r} \& Y.
+\end{tikzcd}$
 \item An ordinary arrow 
@@
 judgment $\jhom{\Gamma}{A}{B}{f}$.
-\item The set of double-headed arrows must form a forest where the root of
-each tree is a family of contexts over $\Gamma$. 
+\item The set of double-headed arrows must form a forest and the root of
+each maximal tree of double-headed arrows is a family of contexts over $\Gamma$.
+In particular, if an object is not the domain of a double-headed arrow it must
+be a family of contexts over $\Gamma$.
 \item A sequence of double-headed 
@@
 \begin{tikzcd}
-P_{n+1} \ar[fib]{r} & \cdots \ar[fib]{r} & P_1 \ar[fib]{r} & A
+P_{n} \ar[fib]{r} & \cdots \ar[fib]{r} & P_1 \ar[fib]{r} & A
 \end{tikzcd}
@@
 $P_2$ is a family of contexts over $\ctxext{{\Gamma}{A}}{P_1}$, etcetera.
-\item A ladder of double-headed arrows
+\item There can be two kinds of ladders of double-headed arrows:
 \begin{equation*}
 \begin{tikzcd}
-P_{n+1} \ar{r}{F_{n+1}} \ar[fib]{d} & Q_{n+1} \ar[fib]{d}\\
+P_{n} \ar{r}{F_{n}} \ar[fib]{d} & Q_{n} \ar[fib]{d}\\
 \vdots \ar[fib]{d} & \vdots \ar[fib]{d}\\
@@
 \end{tikzcd}
+\qquad
+\begin{tikzcd}[column sep = tiny]
+P_{n+m} \ar{rr}{F_{n+m}} \ar[fib]{d} & & Q_{n+m} \ar[fib]{d}\\
+\vdots \ar[fib]{d} & & \vdots \ar[fib]{d}\\
+P_{n+1} \ar{rr}{F_{n+1}} \ar[fib]{dr} & & Q_{n+1} \ar[fib]{dl}\\
+& P_n \ar[fib]{d}\\
+& \vdots \ar[fib]{d}\\
+& P_1 \ar[fib]{d}\\
+& A
+\end{tikzcd}
 \end{equation*}
+The ladder on the left 
 indicates that $F_1$ is a morphism from $P_1$ to $Q_1$ \emph{over} $f$,
@@
 $\ctxext{B}{Q_1}$, etcetera.
+
+The ladder on the right indicates that $F_{n+1}$ is a morphism from $P_{n+1}$ to
+$Q_{n+1}$ in the appropriate context, that $F_{n+2}$ is a morphism from
+$P_{n+2}$ to $Q_{n+2}$ over $F_{n+1}$, etcetera.
+
+Note that the object(s) at the bottom of a ladder are always families of contexts
+over $\Gamma$, so that the typing of the various ingredients makes sense.
 \end{itemize}
@@
 \end{defn}
+
+The most basic illustrative example of a commutative dependency diagram is
+the diagram
+\begin{equation*}
+\begin{tikzcd}
+P \ar[fib]{d} \ar{r}{F} & Q \ar[fib]{d} \\
+A \ar{r}{f} & B
+\end{tikzcd}
+\end{equation*}
+indicating a morphism $F$ from $P$ to $Q$ over the morphism $f$ from $A$ to
+$B$ in a context $\Gamma$.
+
+\begin{defn}
+We say that a commutative dependency diagram of the form
+\begin{equation*}
+\begin{tikzcd}
+P \ar[fib]{d} \ar{r}{F} & Q \ar[fib]{d} \\
+A \ar{r}{f} & B
+\end{tikzcd}
+\end{equation*}
+is a judgmental pullback if the following inference rules are valid:
+\begin{align*}
+& \inference
+  { \jfam{{\Gamma}{A}}{P'}
+    \jfhom{\Gamma}{A}{B}{f}{P'}{Q}{F'}
+    }
+  { \jhom{{\Gamma}{A}}{P'}{P}{u}
+    }
+  \\
+& \inference
+  { \jfam{{\Gamma}{A}}{P'}
+    \jfhom{\Gamma}{A}{B}{f}{P'}{Q}{F'}
+    }
+  { \jfhomeq{\Gamma}{A}{B}{f}{P'}{Q}{\jcomp{}{u}{F}}{F'}
+    }
+  \\
+& \inference
+  { \jhom{{\Gamma}{A}}{P'}{P}{v}
+    \jfhomeq{\Gamma}{A}{B}{f}{P'}{Q}{\jcomp{}{v}{F}}{F'}
+    }
+  { \jhomeq{{\Gamma}{A}}{P'}{P}{v}{u}
+    }
+\end{align*}
+\end{defn}
+
+\begin{lem}
+The diagram
+\begin{equation*}
+\begin{tikzcd}
+\jcomp{}{f}{Q} \ar[fib]{d} \ar{r}{\idtm{\jcomp{}{f}{Q}}} & Q \ar[fib]{d} \\
+A \ar{r}{f} & B
+\end{tikzcd}
+\end{equation*}
+is a judgmental pullback diagram.
+\end{lem}
 
@@
 
-\subsection{Judgmental trivial cofibrations}
+\subsubsection{Judgmental trivial cofibrations}
 \begin{defn}

### 2014-03-27 article-models-ttconstr.tex
judgmental pullbacks

@@
 \subsection{Universes}
+The type operator $\UU$ has as input a pair $(A,P)$ consisting of a family
+$A$ in context $\Gamma$ and a family $P$ in context $\ctxext{\Gamma}{A}$. Its
+output is a pair $(U(A,P),\tilde{U}(A,P))$ consisting of a family
+$U(A,P)$ in context $\Gamma$ and a family $\tilde{U}(A,P)$ over it for which
+there are terms representing the empty context, the operations of extension,
+weakening and substitution and the identity terms. Moreover there is a morphism
+$i$ from $A$ to $U(A,P)$ with the property that 
+$\jcomp{}{i}{\tilde{U}(A,P)}\jdeq P$.
 

### 2014-03-28 article-models-tt.tex
judgmental pullbacks

@@
 
+\begin{proof}
+The proof is a triviality because $\jhom{{\Gamma}{A}}{P'}{\jcomp{}{f}{Q}}{F'}$
+is the same judgment as $\jfhom{\Gamma}{A}{B}{f}{P'}{Q}{F'}$ and
+$\jcomp{}{\idtm{\jcomp{}{f}{Q}}}{F'}\jdeq F'$.
+\end{proof}
+
+\begin{lem}
+A square
+\begin{equation*}
+\begin{tikzcd}
+P \ar[fib]{d} \ar{r}{F} & Q \ar[fib]{d} \\
+A \ar{r}{f} & B
+\end{tikzcd}
+\end{equation*}
+is a pullback precisely when the square
+\begin{equation*}
+\begin{tikzcd}[column sep = large]
+\ctxext{A}{P} \ar{d}[swap]{\cprojfstf{A}{P}} \ar{r}{\tmext{}{}{\ctxwk{P}{f}}{F}} & \ctxext{B}{Q} \ar{d}{\cprojfstf{B}{Q}} \\
+A \ar{r}{f} & B
+\end{tikzcd}
+\end{equation*}
+is a pullback
+\end{lem}
+
+\begin{lem}
+Suppose we have the diagram
+\begin{equation*}
+\begin{tikzcd}
+P \ar{r}{F} \ar[fib]{d} & Q \ar{r}{G} \ar[fib]{d} & R \ar[fib]{d}\\
+A \ar{r}{f} & B \ar{r}{g} & C
+\end{tikzcd}
+\end{equation*}
+where the square on the right and the outer rectangle are judgmental pullback 
+diagrams. Then the square on the left is a judgmental pullback diagram.
+\end{lem}
+
+\begin{proof}
+Let $\jfam{{\Gamma}{A}}{P'}$ be a family and let $\jfhom{\Gamma}{A}{B}{f}
+{P'}{Q}{F}$ be a morphism over $f$.
+\begin{itemize}
+\item Then we compose $F'$ with $G$ to obtain a morphism over $\jcomp{}{f}{g}$.
+\item Then we get $\jhom{{\Gamma}{A}}{P'}{P}{u}$ with a uniqueness property.
+      The property that $\jcomp{}{u}{F}\jdeq F'$ follows from the assumption
+      that the right square is a pullback.
+\item Now assume that we have another such $v$. Compose it with $F$ and $G$.
+      By the assumed properties this is the same as $u$ composed with $F$ and
+      $G$. By the pullback condition we now get $u\jdeq v$. 
+\end{itemize}
+\end{proof}
+
 \subsubsection{Another special case of projections}

### 2014-03-31 article-models-tt.tex
renaming section 4.1

@@
 
-\subsection{The basic judgments}
+\subsection{Judgments and inference rules}
 The type theory we describe here is a theory of contexts, families of
@@
 \end{equation*}
-is a judgmental pullback if the following inference rules are valid:
+is a family pullback if the following inference rules are valid:
 \begin{align*}
@@
 \end{equation*}
-is a judgmental pullback diagram.
+is a family pullback diagram.
 \end{lem}
@@
 A square
-\begin{equation*}
+\begin{equation}\label{eq:fpb_to_pb_eqv_fpb}
 \begin{tikzcd}
@@
 \end{tikzcd}
-\end{equation*}
-is a pullback precisely when the square
-\begin{equation*}
+\end{equation}
+is a family pullback square whenever the square
+\begin{equation}\label{eq:fpb_to_pb_eqv_pb}
 \begin{tikzcd}[column sep = large]
@@
 \end{tikzcd}
-\end{equation*}
-is a pullback
+\end{equation}
+is a pullback square.
 \end{lem}
+
+\begin{proof}
+First assume that the diagram in \autoref{eq:fpb_to_pb_eqv_fpb} is a family 
+pullback square. Consider a family $\jfam{\Gamma}{X}$, the morphisms
+$\jhom{\Gamma}{X}{A}{g}$ and $\jhom{\Gamma}{X}{{A}{P}}{h}$ and the judgmental
+equality $\jhomeq{\Gamma}{X}{B}{\jcomp{}{g}{f}}{\jcomp{}{h}{\cprojfstf{B}{Q}}}$.
+
+We have the family $\jfam{{\Gamma}{A}}{\ctxwk{A}{X}}$.
+\end{proof}
 
@@
 \end{equation*}
-where the square on the right and the outer rectangle are judgmental pullback 
-diagrams. Then the square on the left is a judgmental pullback diagram.
+where the square on the right and the outer rectangle are family pullback 
+diagrams. Then the square on the left is a family pullback diagram.
 \end{lem}

### 2014-03-31 article-models-tt.tex
added stuff to the intro of section 4

@@
-\section{Operational syntactics for type theory before type constructors}\label{tt}
+\section{The theory of contexts, families and terms}\label{tt}
 In this section we give a description of dependent type theory before type
@@
 \end{center}
-Since extension by a context acts trivially on families and terms there are
-no rules added for extension followed by substitution.
-
-In \autoref{extension-on-terms} we come back to the operation of extension, but
-now for terms. The rules in this section explain what it means to be a term of
-an extended family, namely to be an extended term. The compatibility rules for
-extended terms are then similar as the compatibility rules that have appeared
-earlier: there is a compatibility rule for the extension of an extension, for
-the weakening of an extended term, for the substutution of an extended term by
-an arbitrary term and for substitution by an extended term. The last set of
-compatibility rules is a form of currying-uncurrying -- this famous
-correspondence naturally comes out of our scheme of compatibility rules.
-
-\subsection{Judgments and inference rules}
+
+In \autoref{digging_deeper} we use the framework we have developed in this
+section to derive new notions and their properties. In particular, we will
+develop the notion of \emph{extension on terms} together with the projection
+maps from an extension to the `base' context of family, the 
+\emph{family pullback} which is a version of pullbacks for families and thirdly
+the \emph{inductive morphisms} which are morphisms of type theory that allow
+to find terms of families over the codomain context by pulling them back to
+the domain context and finding a term there.
+
+In \autoref{extension-on-terms} on the extension operation on terms we will
+derive all the compatibility rules that one would expect to hold for extension
+on terms. The key to most of these results is the currying operation, which
+could be seen as the missing feature in the table above. The extension on terms
+operation depends in an essential way on the substitution operation, on the
+identity terms and therefore indirectly also on the weakening operation. Thus,
+we will see here all of the features of the theory we develop in
+\autoref{judgments} through \autoref{categorical_properties} come to the 
+surface.
+
+It should be kept in mind though that in the current formulation there is no
+sealed deal establishing a relationship between families over a context
+and any kind of morphisms -- neither with morphisms having the `base' of the
+family as its codomain nor with families into a universe (universes will be
+introduced in \autoref{universes}). The only thing we know here is
+that a family $P$ over $\ctxext{\Gamma}{A}$ determines a context morphism
+from $\ctxext{A}{P}$ to $A$ in context $\Gamma$, the projection. 
+We do not see this as
+a shortcoming of the theory of contexts families and terms. Rather, such a
+correspondence is a feature of a theory which does incorporate universes. The
+fact that we're lacking a clear connection between families and (a specified
+class of) morphisms, however, does show up in our treatment of the notion we
+called familie pullbacks. For instance, we can't show that a square of families
+is a pullback precisely when the corresponding square of projections is a
+pullback: only the backwards direction holds. Neither can we show that the
+composition of two family pullback squares is again a family pullback square.
+We feel that pointing out what we can and can't do in the current setting is
+an important aspect of developing an intuition with the system and therefore
+we include this subsubsection even though the theory might feel `incomplete'
+compared to the usual theory of pullbacks.
+
+In the last subsubsection we give a treatment of inductive morphisms. These
+morphisms appear also in the introduction of the many inductive type
+constructors in \autoref{tt_constructors} and therefore a general treatment of
+the subject is insightful.
+
+\subsection{Judgments and inference rules}\label{judgments}
 The type theory we describe here is a theory of contexts, families of
@@
 
-\subsection{Composition and identity terms}
+\subsection{Composition and identity terms}\label{categorical_properties}
 \subsubsection{The defining property of weakening}
@@
 \subsection{Digging a little deeper in the theory of contexts, families and terms}
+\label{digging_deeper}
 

### 2014-03-31 article-models-ttconstr.tex
added stuff to the intro of section 4

@@
 \section{Type constructors in structural type theory}
+\label{tt_constructors}
 In this section we will introduce the usual type constructors to structural
@@
 
-\subsection{Universes}
+\subsection{Universes}\label{universes}
 The type operator $\UU$ has as input a pair $(A,P)$ consisting of a family

### 2014-03-31 article-models-tt.tex
added stuff to the intro

@@
 
+Next, we introduce the inclusion of the fibers $\subst{x}{P}$ into the extension
+$\ctxext{A}{P}$ as a morphism in context $\Gamma$. As was the case with
+extension on terms and with projections, there will be a ton of compatibility
+properties which we will prove about these inclusions.
+
 It should be kept in mind though that in the current formulation there is no
@@
 is a pullback precisely when the corresponding square of projections is a
-pullback: only the backwards direction holds. Neither can we show that the
-composition of two family pullback squares is again a family pullback square.
+pullback: only the backwards direction holds. 
+The discrepancies continue: ordinary pullbacks do not always exist
+whereas family pullbacks do but the composition of two family pullback squares
+need not be a family pullback square again whereas the pasting lemma of
+ordinary pullbacks holds as usually.
 We feel that pointing out what we can and can't do in the current setting is
 an important aspect of developing an intuition with the system and therefore
-we include this subsubsection even though the theory might feel `incomplete'
-compared to the usual theory of pullbacks.
+we include this subsubsection even though the theory of family pullbacks
+might feel a bit different than the usual theory of pullbacks.
 

### 2014-04-01 article-models-tt.tex
working on fiber inclusions and pullbacks

@@
 
-\subsubsection{Extension on terms}\label{extension-on-terms}
+\subsubsection{Projections and extension on terms}\label{extension-on-terms}
 In this subsubsection we consider the notion of extension on terms, which has now
@@
 & \inference
-  { \jterm{\Gamma}{\ctxext{{A}{P}}{Q}}{w}
-    }
-  { \jtermeq
+  { \jfam{{{\Gamma}{A}}{P}}{Q}
+    }
+  { \jhomeq
       {\Gamma}
+      {\ctxext{{A}{P}}{Q}}
       {A}
-      {\cprojfst{A}{P}{\cprojfst{\ctxext{A}{P}}{Q}{w}}}
-      {\cprojfst{A}{\ctxext{P}{Q}}{w}}
-    }
-  \\
-& \inference
-  { \jterm{\Gamma}{\ctxext{{A}{P}}{Q}}{w}
-    }
-  { \jtermeq
-      {\Gamma}
-      {\subst{\cprojfst{A}{\ctxext{P}{Q}}{w}}{P}}
-      {\cprojsnd{A}{P}{\cprojfst{\ctxext{A}{P}}{Q}{w}}}
-      {\cprojfst{P}{Q}{\cprojsnd{A}{\ctxext{P}{Q}}{w}}}
-    }
-  \\
-& \inference
-    { \jterm{\Gamma}{\ctxext{{A}{P}}{Q}}{w}
-      }
-    { \jtermeq
-        {\Gamma}
-        {\subst{\cprojfst{\ctxext{A}{P}}{Q}{w}}{Q}}
-        {\cprojsnd{P}{Q}{\cprojsnd{A}{\ctxext{P}{Q}}{w}}}
-        {\cprojsnd{\ctxext{A}{P}}{Q}{w}}
+      {\jcomp{}{\cprojfstf{{A}{P}}{Q}}{\cprojfstf{A}{P}}}
+      {\cprojfstf{A}{{P}{Q}}}
+    }
+  \\
+& \inference
+  { \jfam{{{\Gamma}{A}}{P}}{Q}
+    }
+  { \jhomeq
+      {{\Gamma}{A}}
+      {{P}{Q}}
+      {P}
+      {\jcomp{}{\cprojfstf{{A}{P}}{Q}}{\cprojsndf{A}{P}}}
+      {\jcomp{}{\cprojsndf{A}{{P}{Q}}}{\cprojfstf{P}{Q}}}
+    }
+  \\
+& \inference
+    { \jfam{{{\Gamma}{A}}{P}}{Q}
+      }
+    { \jhomeq
+        {{{\Gamma}{A}}{P}}
+        {Q}
+        {Q}
+        {\cprojsndf{{A}{P}}{Q}}
+        {\jcomp{}{\cprojsndf{A}{{P}{Q}}}{\cprojsndf{P}{Q}}}
       }
@@
     }
-% \\
-%& \inference{\jterm{\Gamma}{\ctxwk{A}{\ctxext{B}{Q}}}{\ctxwk{A}{w}}}{\jtermeq{{\Gamma}{A}}{\ctxwk{A}{B}}{\cprojfst{\ctxwk{A}{B}}{\ctxwk{A}{Q}}{\ctxwk{A}{w}}}{\ctxwk{A}{(\cprojfst{B}{Q}{w})}}}\\
+  \\
+& \inference
+  { \jfam{{{\Gamma}{B}}{Q}}{R}
+    }
+  { \jhomeq
+      {{{\Gamma}{A}}{\ctxwk{A}{B}}}
+      {{\ctxwk{A}{Q}}{\ctxwk{A}{R}}}
+      {\ctxwk{A}{Q}}
+      {\cprojfstf{\ctxwk{A}{Q}}{\ctxwk{A}{R}}}
+      {\ctxwk{A}{\cprojfstf{Q}{R}}}
+    }
+  \\
+& \inference
+  { \jfam{{{\Gamma}{B}}{Q}}{R}
+    }
+  { \jhomeq
+      {{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{Q}}}
+      {\ctxwk{A}{R}}
+      {\ctxwk{A}{R}}
+      {\cprojsndf{\ctxwk{A}{Q}}{\ctxwk{A}{R}}}
+      {\ctxwk{A}{\cprojsndf{Q}{R}}}
+    }
 \end{align*}
@@
 & \inference
-  { \jterm{{{\Gamma}{A}}{P}}{Q}{g}
+  { \jterm{\Gamma}{A}{x}
+    \jterm{{{\Gamma}{A}}{P}}{Q}{g}
     \jterm{{{\Gamma}{A}}{P}}{\subst{g}{R}}{t}
@@
     }
+  \\
+& \inference
+  { \jterm{\Gamma}{A}{x}
+    \jfam{{{\Gamma}{A}}{P}}{Q}
+    }
+  { \jhomeq
+      {\Gamma}
+      {{\subst{x}{P}}{\subst{x}{Q}}}
+      {\subst{x}{P}}
+      {\cprojfstf{\subst{x}{P}}{\subst{x}{Q}}}
+      {\subst{x}{\cprojfstf{P}{Q}}}
+    }
+  \\
+& \inference
+  { \jterm{\Gamma}{A}{x}
+    \jfam{{{\Gamma}{A}}{P}}{Q}
+    }
+  { \jhomeq
+      {{\Gamma}{\subst{x}{P}}}
+      {\subst{x}{Q}}
+      {\subst{x}{Q}}
+      {\cprojsndf{\subst{x}{P}}{\subst{x}{Q}}}
+      {\subst{x}{\cprojsndf{P}{Q}}}
+    }
 \end{align*}
@@
 \end{align*}
+\end{proof}
+
+\subsubsection{Fiber inclusions}
+We will use the insights of \autoref{extension-on-terms} to define and study
+\emph{fiber inclusions}. The fiber inclusion of the \emph{fiber}
+$\subst{x}{P}$ into the extension $\ctxext{A}{P}$ is a morphism
+$\jhom{\Gamma}{\subst{x}{P}}{{A}{P}}{\finc{x}{P}}$, for any family
+$\jfam{{\Gamma}{A}}{P}$ and any term $\jterm{\Gamma}{A}{x}$. Then we will determine
+the ways in which it is compatible with the other operators. Note that in this
+subsubsection we will focus on the compatibility properties; the fact that
+the fiber inclusions also appear in a pullback diagram will be established in
+\autoref{pullback}. Because 
+$\finc{x}{P}$ is a morphism into an extension, we will first study morphisms
+into extensions more generally.
+
+For the following lemma, recall that the judgment $\jhom{\Gamma}{A}{{B}{Q}}{f}$
+unfolds as
+\begin{equation*}
+\unfold{\jhom{\Gamma}{A}{{B}{Q}}{f}}
+\end{equation*}
+and that we have the judgmental equality 
+$ \jfameq
+    {{\Gamma}{A}}
+    {\ctxwk{A}{\ctxext{B}{Q}}}
+    {\ctxext{\ctxwk{A}{B}}{\ctxwk{A}{Q}}}.
+  $
+Therefore, each morphism into an extended family can itself be described as
+an extended term. The following lemma explains how this goes.
+
+\begin{lem}
+Let $\jhom{\Gamma}{A}{{B}{Q}}{f}$ be a morphism from $A$ to $\ctxext{B}{Q}$
+in a context $\Gamma$. Then we have
+\begin{equation*}
+\jhomeq
+  {\Gamma}
+  {A}
+  {{B}{Q}}
+  {f}
+  {\tmext{}{}{\jcomp{}{f}{\cprojfstf{B}{Q}}}{\jcomp{}{f}{\cprojsndf{B}{Q}}}}.
+\end{equation*}
+Alternatively, when $\jhom{\Gamma}{A}{B}{f_0}$ and 
+$\jterm{{\Gamma}{A}}{\jcomp{}{f_0}{Q}}{f_1}$ we obtain a morphism
+$\jhom{\Gamma}{A}{{B}{Q}}{\tmext{}{}{f_0}{f_1}}$ with the property that
+\begin{align*}
+\jhomeq*{\Gamma}{A}{B}{\jcomp{}{\tmext{}{}{f_0}{f_1}}{\cprojfstf{B}{Q}}}{f_0}\\
+\jtermeq*{{\Gamma}{A}}{\jcomp{}{f_0}{Q}}{\jcomp{}{\tmext{}{}{f_0}{f_1}}{\cprojsndf{B}{Q}}}{f_1}.
+\end{align*}
+\end{lem}
+
+\begin{proof}
+Let $\jhom{\Gamma}{A}{{B}{Q}}{f}$ be a morphism in context $\Gamma$. Then we
+have the judgmental equalities
+\begin{align*}
+\cprojfst{\ctxwk{A}{B}}{\ctxwk{A}{Q}}{f}
+& \jdeq
+  \subst{f}{\ctxwk{A}{\cprojfstf{B}{Q}}}
+  \\
+& \jdeq
+  \jcomp{}{f}{\cprojfstf{B}{Q}}
+  \\
+\cprojsnd{\ctxwk{A}{B}}{\ctxwk{A}{Q}}{f}
+& \jdeq
+  \subst{f}{\ctxwk{A}{\cprojsndf{B}{Q}}}
+  \\
+& \jdeq
+  \jcomp{}{f}{\cprojsndf{B}{Q}}
+\end{align*}
+The alternative formulation of the statement is a direct corollary.
 \end{proof}
@@
 
-\subsubsection{Commuting dependence diagrams}
+\begin{defn}
+Let $\jterm{\Gamma}{A}{x}$ be a term and let $\jfam{{\Gamma}{A}}{P}$ be a
+family. Then we define
+\begin{equation*}
+\jhomdefn{\Gamma}{\subst{x}{P}}{{A}{P}}{\finc{x}{P}}{\unfoldnext{\finc{x}{P}}}
+\end{equation*}
+\end{defn}
+
+\subsubsection{Pullback squares and family pullback squares}
+\label{pullback}
 Now that we have introduced the notions of morphisms, composition and identity
@@
 
+\begin{lem}
+For any $\jterm{\Gamma}{A}{x}$ and any $\jfam{{\Gamma}{A}}{P}$, the square
+\begin{equation*}
+\begin{tikzcd}
+\subst{x}{P} \ar{d} \ar{r}{\finc{x}{P}} & \ctxext{A}{P} \ar{d}{\cprojfstf{A}{P}}\\
+\emptyf \ar{r}{\ctxwk{\emptyf}{x}} & A
+\end{tikzcd}
+\end{equation*}
+is a pullback square.
+\end{lem}
+
+\begin{lem}
+We have the following inference rule:
+\begin{equation*}
+\inference
+  { \jfam{{{\Gamma}{A}}{P}}{Q}
+    \jterm{\Gamma}{A}{x}
+    }
+  { \jfameq
+      {{\Gamma}{\subst{x}{P}}}
+      {\jcomp{}{\finc{x}{P}}{Q}}
+      {\subst{x}{Q}}
+    }
+\end{equation*}
+\end{lem}
+
+\begin{proof}
+We have the judgmental equalities:
+\begin{align*}
+\jcomp{\subst{x}{P}}{\finc{x}{P}}{Q}
+& \jdeq
+  \unfold{\jcomp{\subst{x}{P}}{\finc{x}{P}}{Q}}
+  \\
+& \jdeq
+  \subst{\idtm{\subst{x}{P}}}{{\ctxwk{\subst{x}{P}}{x}}{\ctxwk{\subst{x}{P}}{Q}}}
+  \\
+& \jdeq
+  \subst{\idtm{\subst{x}{P}}}{\ctxwk{\subst{x}{P}}{\subst{x}{Q}}}
+  \\
+& \jdeq
+  \subst{x}{Q}.
+\end{align*}
+\end{proof}
+
 \subsubsection{Another special case of projections}

### 2014-04-01 article-models-tt.tex
small impro
added comments in the pullback section and changed judgmental trivial cofibration to inductive morphism

@@
 
+We can just copy the usual categorical definition of a pullback square to our
+current situation, but we have to require that each arrow in the pullback square
+is an ordinary arrow. When families (i.e. double-headed arrows) are involved
+in the diagram, we make the following definition of a family pullback:
+
 \begin{defn}
@@
 \end{equation*}
-is a family pullback if the following inference rules are valid:
+is a \emph{family pullback} if the following inference rules are valid:
 \begin{align*}
@@
 
-\begin{lem}
-The diagram
-\begin{equation*}
-\begin{tikzcd}
-\jcomp{}{f}{Q} \ar[fib]{d} \ar{r}{\idtm{\jcomp{}{f}{Q}}} & Q \ar[fib]{d} \\
-A \ar{r}{f} & B
-\end{tikzcd}
-\end{equation*}
-is a family pullback diagram.
-\end{lem}
-
-\begin{proof}
-The proof is a triviality because $\jhom{{\Gamma}{A}}{P'}{\jcomp{}{f}{Q}}{F'}$
-is the same judgment as $\jfhom{\Gamma}{A}{B}{f}{P'}{Q}{F'}$ and
-$\jcomp{}{\idtm{\jcomp{}{f}{Q}}}{F'}\jdeq F'$.
-\end{proof}
+The following lemma explains that when a square involving families is a
+family pullback square whenever the corresponding square involving projections is a
+pullback square. There is no proof in the the opposite direction.
 
@@
 
+The family pullback of a family along any morphism always exists. It is simply given
+by the precomposition of the family with the morphism. Note that this fact does
+not carry over to arbitrary pullbacks.
+
+\begin{lem}
+The diagram
+\begin{equation*}
+\begin{tikzcd}
+\jcomp{}{f}{Q} \ar[fib]{d} \ar{r}{\idtm{\jcomp{}{f}{Q}}} & Q \ar[fib]{d} \\
+A \ar{r}{f} & B
+\end{tikzcd}
+\end{equation*}
+is a family pullback diagram.
+\end{lem}
+
 \begin{proof}
-First assume that the diagram in \autoref{eq:fpb_to_pb_eqv_fpb} is a family 
-pullback square. Consider a family $\jfam{\Gamma}{X}$, the morphisms
-$\jhom{\Gamma}{X}{A}{g}$ and $\jhom{\Gamma}{X}{{A}{P}}{h}$ and the judgmental
-equality $\jhomeq{\Gamma}{X}{B}{\jcomp{}{g}{f}}{\jcomp{}{h}{\cprojfstf{B}{Q}}}$.
-
-We have the family $\jfam{{\Gamma}{A}}{\ctxwk{A}{X}}$.
+The proof is a triviality because $\jhom{{\Gamma}{A}}{P'}{\jcomp{}{f}{Q}}{F'}$
+is the same judgment as $\jfhom{\Gamma}{A}{B}{f}{P'}{Q}{F'}$ and
+$\jcomp{}{\idtm{\jcomp{}{f}{Q}}}{F'}\jdeq F'$.
 \end{proof}
+
+For arbitrary pullbacks we have the pasting lemma as usual, but for family
+pullbacks we can only derive one of the parts of the pasting lemma.
 
@@
 
+In the following lemma we assert that pulling back a family $Q$ 
+over $\ctxext{{\Gamma}{A}}{P}$ along a fiber
+inclusion $\finc{x}{P}$ gives the family $\subst{x}{Q}$ over $\ctxext{\Gamma}{\subst{x}{P}}$.
+
 \begin{lem}
@@
 & \jdeq
-  \unfold{\jcomp{\subst{x}{P}}{\finc{x}{P}}{Q}}
+  \subst{\tmext{}{}{\ctxwk{\subst{x}{P}}{x}}{\idtm{\subst{x}{P}}}}{\ctxwk{\subst{x}{P}}{Q}}
   \\
@@
 
-\subsubsection{Judgmental trivial cofibrations}
+\subsubsection{Inductive morphisms}
 \begin{defn}
-Let $\jhom{\Gamma}{A}{B}{f}$ be a context morphism. We say that $f$ is a 
-\emph{judgmental trivial cofibration} if the following inference rules are valid:
+Let $\jhom{\Gamma}{A}{B}{f}$ be a context morphism. We say that $f$ is an 
+\emph{inductive morphism} if the following inference rules are valid:
 \begin{align*}
@@
 by finding a term of the family $\jcomp{A}{f}{Q}$ over $\ctxext{\Gamma}{A}$.
-Judgmental trivial cofibrations come with a computation rule. Thus, term
+Inductive morphisms come with a computation rule. Thus, term
 constructors of inductively defined types are going to be the major source of
-examples of judgmental trivial cofibrations. Since identity terms behave
-so nicely, they are judgmental trivial cofibrations too.
+examples of inductive morphisms. Since identity terms behave
+so nicely, they are inductive morphisms too.
 
@@
 The identity term
-$\jhom{\Gamma}{A}{A}{\idtm{A}}$ is a judgmental trivial cofibration
+$\jhom{\Gamma}{A}{A}{\idtm{A}}$ is an inductive morphism
 for each family $A$ of contexts over $\Gamma$

### 2014-04-01 article-models-main.tex
reorganizing the notes
placed the section on the model type theory in a separate file
separating the constructors section into a separate file
added intro to part II

@@
 
-\vspace{2\baselineskip}
+\part{Introduction}
 The project proposal is described in \autoref{stage1} and \autoref{stage2}. In
@@
 
-\bibliographystyle{plain}
-%\phantomsection\addcontentsline{toc}{section}{References}
-\bibliography{refs}
-
 %\input{article-models-eg}
 
-\newpage
-\appendix
+\part{Type theory}
+
+In this part we develop type theory from the ground up. We start with a type
+theory without any of the basic constructors. This is the theory of contexts
+families and terms which has the basic operations of extension, weakening,
+substitution and identity terms. Type theory before type constructors has not
+been studied very much. Dependent product types or even universes tend to make
+an early appearance in just about any presentation of type theory. While one
+could argue that in type theories like the one in \cite{Dybjer1996} the type
+constructors could be left out and a study of the remnants could \emph{in
+principle} be carried out, this is not the usual practice. In the type theory
+of \cite{TheBook}, which seems to have won the race of introducing universes
+as early as possible hands down, this seems entirely unfeasible.
+
+Nevertheless, type theory without type constructors has received some attention
+recently. In \cite{Garner2014}, Garner describes the combinatorial structure
+of the type operations of the weakening, substitution and projection monads
+(their projections are our identity terms) and suggests lots of further research
+that can be done on type theory without constructors. Also, Joyal has been
+lecturing about tribes lately, which are a categorical semantics for type theory
+without type constructors. Although his slides have been going around on
+mailing lists, there is as of yet no easily accessible written accound on tribes.
+
 \input{article-models-tt}
+
+\input{article-models-ttconstr}
+
+\input{article-models-ttmodels}
 
@@
 
+\bibliographystyle{plain}
+%\phantomsection\addcontentsline{toc}{section}{References}
+\bibliography{refs}
+
 \end{document}

### 2014-04-01 article-models-main.tex
added lots of references

@@
 constructors could be left out and a study of the remnants could \emph{in
-principle} be carried out, this is not the usual practice. In the type theory
+principle} be carried out, this is not the usual practice. A noteable exception
+is the unpublished work \cite{Awodey2013} on natural models of type theory,
+which makes the connection between categories with families and representable
+transformations of presheaves. In the type theory
 of \cite{TheBook}, which seems to have won the race of introducing universes
-as early as possible hands down, this seems entirely unfeasible.
+as early as possible hands down, this seems entirely unfeasible. Also, the 
+proof-assistant \Coq\ {\color{red}(and \Agda too?)} has universes and dependent product and pair types 
+built-in, making it impossible to study type theory without type constructors in 
+that environment.
 
 Nevertheless, type theory without type constructors has received some attention
-recently. In \cite{Garner2014}, Garner describes the combinatorial structure
+contributors to the Univalent Foundations Program recently. We name two
+further investigations on this topic, other than the mentioned work by Awodey.
+In \cite{Garner2014}, Garner describes the combinatorial structure
 of the type operations of the weakening, substitution and projection monads
@@
 mailing lists, there is as of yet no easily accessible written accound on tribes.
+
+The theory that is layed out in this part is built up modularly. As announced,
+we begin by presenting type theory without type constructors. This type theory
+can be extended by adding type constructors. (Probably) There are three flavors
+of type constructors. Dependent function types and universes are each of their
+own flavor; the inductive types form the third flavor.
+
+However the type theory can also be extended with local versions of extenison,
+weakening, substitution and identity terms. This comprises a type theory of
+(internal) models of the theory of contexts families and terms presented in
+\autoref{tt} and it can be further extended to also model any combination of
+type constructors.
 

### 2014-04-01 article-models-tt.tex
added intro to part II
added lots of references

@@
 In this section we give a description of dependent type theory before type
-constructors with explicit operations of extension, weakening and substitution. We will
+constructors. Apart from contexts, families and terms -- which provide for the
+core of the language of dependent logic -- the basic ingredients
+of this theory will be the operations of extension, weakening and substitution
+and the identity terms. The resulting theory can be seen as the structure of
+dependent type theory. We note up front that what we call families over contexts
+here, could have been named dependent contexts or telescopes, see
+\cite{deBruijn1991,hofmann1995extensional}.
+
+We will
 formulate the type theory in such a way that contexts aren't defined
-to be lists of variable
-declarations. We rather take a syntactic approach in our treatment of contexts,
-axiomatizing only what you can do with them and allowing for the freedom to
-interpret the axiom in any desired way. 
-The main reason is that we don't see variable declarations in the internal
-models either. This way we also set out to a more algebraic approach of type
-theory and higher category theory. Thirdly, we will not have to be burdened with
-superficial comments about variables being bounded or not, or fresh or free or 
-not occuring at all.
-
-The theory we will develop here is a theory of contexts, families of contexts
-and terms thereof. Families of contexts could have been named dependent contexts
-or telescopes.
+to be lists of variable declarations. The variable-free (or name-free) approach 
+we take here is rather different than those appearing in 
+\cite{hofmann1995extensional,TheBook} but it has appeared in the work of Coquand
+and in \cite{Dybjer1996}.
+The main reason we don't let variable declarations in is that we don't see them 
+in the internal models either. This way we also set out to a more algebraic 
+approach of type theory and higher category theory. Thirdly, we will not have to
+be burdened with superficial comments about variables being bounded or not, or 
+fresh or free or not occuring at all.
 
@@
 \end{lem}
+
+\begin{itemize}
+\item Extensions of inductive morphisms are inductive
+\item Weakenings of inductive morphisms are inductive
+\item Substitutions of inductive morphisms are inductive
+\end{itemize}

### 2014-04-01 article-models-tt.tex
small impros

@@
 of this theory will be the operations of extension, weakening and substitution
-and the identity terms. The resulting theory can be seen as the structure of
-dependent type theory. We note up front that what we call families over contexts
-here, could have been named dependent contexts or telescopes, see
-\cite{deBruijn1991,hofmann1995extensional}.
+and the identity terms. An empty context (and empty families) are also included
+in the theory. The resulting theory can be seen as a manifestation of the 
+structure underlying dependent type theory.
 
@@
 
+We note that what we call families over contexts
+here could also have been named dependent contexts or telescopes, see
+\cite{deBruijn1991,hofmann1995extensional}. The term family is in agreement
+with the terminology scheme of \cite{TheBook}, though the reader should be
+warned that the notion of familie means something slightly different there than
+it does here.
+
 If $A$ is a type

### 2014-04-01 article-models-tt.tex
created an extra subsubsection about morphisms
added vertical composition to the preamble
fixing tmext

@@
 
+\subsubsection{More on morphisms}
+
+For the following lemma, recall that the judgment $\jhom{\Gamma}{A}{{B}{Q}}{f}$
+unfolds as
+\begin{equation*}
+\unfold{\jhom{\Gamma}{A}{{B}{Q}}{f}}
+\end{equation*}
+and that we have the judgmental equality 
+$ \jfameq
+    {{\Gamma}{A}}
+    {\ctxwk{A}{\ctxext{B}{Q}}}
+    {\ctxext{\ctxwk{A}{B}}{\ctxwk{A}{Q}}}.
+  $
+Therefore, each morphism into an extended family can itself be described as
+an extended term. The following lemma explains how this goes.
+
+\begin{lem}
+Let $\jhom{\Gamma}{A}{{B}{Q}}{f}$ be a morphism from $A$ to $\ctxext{B}{Q}$
+in a context $\Gamma$. Then we have
+\begin{equation*}
+\jhomeq
+  {\Gamma}
+  {A}
+  {{B}{Q}}
+  {f}
+  {\tmext{\jcomp{}{f}{\cprojfstf{B}{Q}}}{\jcomp{}{f}{\cprojsndf{B}{Q}}}}.
+\end{equation*}
+Alternatively, when $\jhom{\Gamma}{A}{B}{f_0}$ and 
+$\jterm{{\Gamma}{A}}{\jcomp{}{f_0}{Q}}{f_1}$ we obtain a morphism
+$\jhom{\Gamma}{A}{{B}{Q}}{\tmext{f_0}{f_1}}$ with the property that
+\begin{align*}
+\jhomeq*{\Gamma}{A}{B}{\jcomp{}{\tmext{f_0}{f_1}}{\cprojfstf{B}{Q}}}{f_0}\\
+\jtermeq*{{\Gamma}{A}}{\jcomp{}{f_0}{Q}}{\jcomp{}{\tmext{f_0}{f_1}}{\cprojsndf{B}{Q}}}{f_1}.
+\end{align*}
+\end{lem}
+
+\begin{proof}
+Let $\jhom{\Gamma}{A}{{B}{Q}}{f}$ be a morphism in context $\Gamma$. Then we
+have the judgmental equalities
+\begin{align*}
+\cprojfst{\ctxwk{A}{B}}{\ctxwk{A}{Q}}{f}
+& \jdeq
+  \subst{f}{\ctxwk{A}{\cprojfstf{B}{Q}}}
+  \\
+& \jdeq
+  \jcomp{}{f}{\cprojfstf{B}{Q}}
+  \\
+\cprojsnd{\ctxwk{A}{B}}{\ctxwk{A}{Q}}{f}
+& \jdeq
+  \subst{f}{\ctxwk{A}{\cprojsndf{B}{Q}}}
+  \\
+& \jdeq
+  \jcomp{}{f}{\cprojsndf{B}{Q}}
+\end{align*}
+The alternative formulation of the statement is a direct corollary.
+\end{proof}
+
+We also have the following lemma about the compatibility of pairing and composition:
+
+\begin{lem}
+The following inference rule is valid
+\begin{align*}
+& \inference
+  { \jhom{\Gamma}{A}{B}{f}
+    \jhom{\Gamma}{B}{C}{g}
+    \jfam{{\Gamma}{C}}{R}
+    \jterm{{\Gamma}{B}}{\subst{g}{\ctxwk{B}{R}}}{w}
+    }
+  { \jhomeq
+      {\Gamma}
+      {A}
+      {{C}{R}}
+      {\jcomp{A}{f}{\tmext{\ctxwk{B}{C}}{\ctxwk{B}{R}}{g}{w}}}
+      {\tmext{\ctxwk{A}{C}}{\ctxwk{A}{R}}{\jcomp{A}{f}{g}}{\jcomp{A}{f}{w}}}
+    }
+\end{align*}
+\end{lem}
+
+\begin{proof}
+Let $\jhom{\Gamma}{A}{B}{f}$, $\jhom{\Gamma}{B}{C}{g}$, $\jfam{{\Gamma}{C}}{R}$
+and $\jterm{{\Gamma}{B}}{\subst{g}{\ctxwk{B}{R}}}{w}$. Then we have the
+judgmental equalities
+\begin{align*}
+\jcomp{A}{f}{\tmext{\ctxwk{B}{C}}{\ctxwk{B}{R}}{g}{w}}
+& \jdeq 
+  \subst{f}{\ctxwk{A}{\tmext{\ctxwk{B}{C}}{\ctxwk{B}{R}}{g}{w}}}
+  \\
+& \jdeq 
+  \subst
+    {f}
+    {\tmext{\ctxwk{A}{{B}{C}}}{\ctxwk{A}{{B}{R}}}{\ctxwk{A}{g}}{\ctxwk{A}{w}}}
+  \\
+& \jdeq 
+  \tmext
+    {\ctxwk{A}{C}}
+    {\ctxwk{A}{R}}
+    {\subst{f}{\ctxwk{A}{g}}}
+    {\subst{f}{\ctxwk{A}{w}}}
+  \\
+& \jdeq 
+  \tmext{\ctxwk{A}{C}}{\ctxwk{A}{R}}{\jcomp{A}{f}{g}}{\jcomp{A}{f}{w}}.
+  \qedhere
+\end{align*}
+\end{proof}
+
+\begin{defn}
+Let $\jhom{\Gamma}{A}{B}{f}$ be a morphism from $A$ to $B$ in context $\Gamma$
+and let $\jfhom{\Gamma}{A}{B}{f}{P}{Q}{F}$ be a morphism over $f$ in context 
+$\Gamma$. We define
+\begin{equation*}
+\jhomdefn{\Gamma}{{A}{P}}{{B}{Q}}{\jvcomp{P}{f}{F}}{\unfold{\jvcomp{P}{f}{F}}}
+\end{equation*}
+\end{defn}
+
 \subsubsection{Fiber inclusions}
@@
 the fiber inclusions also appear in a pullback diagram will be established in
-\autoref{pullback}. Because 
-$\finc{x}{P}$ is a morphism into an extension, we will first study morphisms
-into extensions more generally.
-
-For the following lemma, recall that the judgment $\jhom{\Gamma}{A}{{B}{Q}}{f}$
-unfolds as
-\begin{equation*}
-\unfold{\jhom{\Gamma}{A}{{B}{Q}}{f}}
-\end{equation*}
-and that we have the judgmental equality 
-$ \jfameq
-    {{\Gamma}{A}}
-    {\ctxwk{A}{\ctxext{B}{Q}}}
-    {\ctxext{\ctxwk{A}{B}}{\ctxwk{A}{Q}}}.
-  $
-Therefore, each morphism into an extended family can itself be described as
-an extended term. The following lemma explains how this goes.
-
-\begin{lem}
-Let $\jhom{\Gamma}{A}{{B}{Q}}{f}$ be a morphism from $A$ to $\ctxext{B}{Q}$
-in a context $\Gamma$. Then we have
-\begin{equation*}
-\jhomeq
-  {\Gamma}
-  {A}
-  {{B}{Q}}
-  {f}
-  {\tmext{}{}{\jcomp{}{f}{\cprojfstf{B}{Q}}}{\jcomp{}{f}{\cprojsndf{B}{Q}}}}.
-\end{equation*}
-Alternatively, when $\jhom{\Gamma}{A}{B}{f_0}$ and 
-$\jterm{{\Gamma}{A}}{\jcomp{}{f_0}{Q}}{f_1}$ we obtain a morphism
-$\jhom{\Gamma}{A}{{B}{Q}}{\tmext{}{}{f_0}{f_1}}$ with the property that
-\begin{align*}
-\jhomeq*{\Gamma}{A}{B}{\jcomp{}{\tmext{}{}{f_0}{f_1}}{\cprojfstf{B}{Q}}}{f_0}\\
-\jtermeq*{{\Gamma}{A}}{\jcomp{}{f_0}{Q}}{\jcomp{}{\tmext{}{}{f_0}{f_1}}{\cprojsndf{B}{Q}}}{f_1}.
-\end{align*}
-\end{lem}
-
-\begin{proof}
-Let $\jhom{\Gamma}{A}{{B}{Q}}{f}$ be a morphism in context $\Gamma$. Then we
-have the judgmental equalities
-\begin{align*}
-\cprojfst{\ctxwk{A}{B}}{\ctxwk{A}{Q}}{f}
-& \jdeq
-  \subst{f}{\ctxwk{A}{\cprojfstf{B}{Q}}}
-  \\
-& \jdeq
-  \jcomp{}{f}{\cprojfstf{B}{Q}}
-  \\
-\cprojsnd{\ctxwk{A}{B}}{\ctxwk{A}{Q}}{f}
-& \jdeq
-  \subst{f}{\ctxwk{A}{\cprojsndf{B}{Q}}}
-  \\
-& \jdeq
-  \jcomp{}{f}{\cprojsndf{B}{Q}}
-\end{align*}
-The alternative formulation of the statement is a direct corollary.
-\end{proof}
-
-We also have the following lemma about the compatibility of pairing and composition:
-
-\begin{lem}
-The following inference rule is valid
-\begin{align*}
-& \inference
-  { \jhom{\Gamma}{A}{B}{f}
-    \jhom{\Gamma}{B}{C}{g}
-    \jfam{{\Gamma}{C}}{R}
-    \jterm{{\Gamma}{B}}{\subst{g}{\ctxwk{B}{R}}}{w}
-    }
-  { \jhomeq
-      {\Gamma}
-      {A}
-      {{C}{R}}
-      {\jcomp{A}{f}{\tmext{\ctxwk{B}{C}}{\ctxwk{B}{R}}{g}{w}}}
-      {\tmext{\ctxwk{A}{C}}{\ctxwk{A}{R}}{\jcomp{A}{f}{g}}{\jcomp{A}{f}{w}}}
-    }
-\end{align*}
-\end{lem}
-
-\begin{proof}
-Let $\jhom{\Gamma}{A}{B}{f}$, $\jhom{\Gamma}{B}{C}{g}$, $\jfam{{\Gamma}{C}}{R}$
-and $\jterm{{\Gamma}{B}}{\subst{g}{\ctxwk{B}{R}}}{w}$. Then we have the
-judgmental equalities
-\begin{align*}
-\jcomp{A}{f}{\tmext{\ctxwk{B}{C}}{\ctxwk{B}{R}}{g}{w}}
-& \jdeq 
-  \subst{f}{\ctxwk{A}{\tmext{\ctxwk{B}{C}}{\ctxwk{B}{R}}{g}{w}}}
-  \\
-& \jdeq 
-  \subst
-    {f}
-    {\tmext{\ctxwk{A}{{B}{C}}}{\ctxwk{A}{{B}{R}}}{\ctxwk{A}{g}}{\ctxwk{A}{w}}}
-  \\
-& \jdeq 
-  \tmext
-    {\ctxwk{A}{C}}
-    {\ctxwk{A}{R}}
-    {\subst{f}{\ctxwk{A}{g}}}
-    {\subst{f}{\ctxwk{A}{w}}}
-  \\
-& \jdeq 
-  \tmext{\ctxwk{A}{C}}{\ctxwk{A}{R}}{\jcomp{A}{f}{g}}{\jcomp{A}{f}{w}}.
-  \qedhere
-\end{align*}
-\end{proof}
+\autoref{pullback}.
 
@@
 $F_2$ is a morphism from $P_2$ to $Q_2$ over
-the morphism $\tmext{}{}{\ctxwk{P_1}{f}}{F_1}$ from $\ctxext{A}{P_1}$ to
+the morphism $\tmext{\ctxwk{P_1}{f}}{F_1}$ from $\ctxext{A}{P_1}$ to
 $\ctxext{B}{Q_1}$, etcetera.
@@
 \begin{tikzcd}[column sep = large]
-\ctxext{A}{P} \ar{d}[swap]{\cprojfstf{A}{P}} \ar{r}{\tmext{}{}{\ctxwk{P}{f}}{F}} & \ctxext{B}{Q} \ar{d}{\cprojfstf{B}{Q}} \\
+\ctxext{A}{P} \ar{d}[swap]{\cprojfstf{A}{P}} \ar{r}{\tmext{\ctxwk{P}{f}}{F}} & \ctxext{B}{Q} \ar{d}{\cprojfstf{B}{Q}} \\
 A \ar{r}{f} & B
@@
 & \jdeq
-  \subst{\tmext{}{}{\ctxwk{\subst{x}{P}}{x}}{\idtm{\subst{x}{P}}}}{\ctxwk{\subst{x}{P}}{Q}}
+  \subst{\tmext{\ctxwk{\subst{x}{P}}{x}}{\idtm{\subst{x}{P}}}}{\ctxwk{\subst{x}{P}}{Q}}
   \\
@@
   {\bar{\typefont{pr}}}
-  {\tmext{}{}{\ctxwk{{P}{\mfam{A}}}{\cprojfstf{A}{P}}}{\ctxwk{P}{\idtm{\mfam{A}}}}}
+  {\tmext{\ctxwk{{P}{\mfam{A}}}{\cprojfstf{A}{P}}}{\ctxwk{P}{\idtm{\mfam{A}}}}}
 \end{equation}

### 2014-04-01 article-models-ttmodels.tex
fixing tmext

@@
 other words, $\edg{\gamma}$ is a family over $\pts{\gamma}$. A family $\beta$
-over $\gamma$ is then a pair $\tmext{}{}{\pts{\beta}}{\edg{\beta}}$ consisting
+over $\gamma$ is then a pair $\tmext{\pts{\beta}}{\edg{\beta}}$ consisting
 of a family $\pts{\beta}$ over $\pts{\gamma}$ and a family $\edg{\beta}$ over
-the extended term $\tmext{}{}{\pts{\beta}}{\edg{\gamma}}$. 
+the extended term $\tmext{\pts{\beta}}{\edg{\gamma}}$. 
 It is this extension which appears in the type of $\edg{\beta}$ which requires

### 2014-04-02 article-models-tt.tex
rewriting the proof of consecutive term extensions
completed proof of consecutive term extension lemma
finished section 4.8.1

@@
 \begin{align*}
-\tmext{A}{{P}{Q}}{x}{{\subst{x}{P}}{\subst{x}{Q}}{u}{v}} 
+\tmext{x}{{u}{v}}
 & \jdeq 
   \subst
-    {\tmext{\subst{x}{P}}{\subst{x}{Q}}{u}{v}}
-    {{x}{\idtm{\ctxext{A}{{P}{Q}}}}}
+    {\tmext{u}{v}}{{x}{\idtm{\ctxext{A}{{P}{Q}}}}}
   \\
@@
 & \jdeq 
-  \subst{v}{{\tmext{A}{P}{x}{u}}{\idtm{\ctxext{{A}{P}}{Q}}}}
-  \\
-& \jdeq 
-  \tmext{{A}{P}}{Q}{{A}{P}{x}{u}}{v}.
-\end{align*}
+  \subst{v}{{\tmext{x}{u}}{\idtm{\ctxext{{A}{P}}{Q}}}}
+  \\
+& \jdeq 
+  \tmext{{x}{u}}{v}.
+\end{align*}
+To prove the judmental equality
+\begin{equation*}
+\jhomeq
+  {\Gamma}
+  {\ctxext{{A}{P}}{Q}}
+  {A}
+  {\jcomp{}{\cprojfstf{{A}{P}}{Q}}{\cprojfstf{A}{P}}}
+  {\cprojfstf{A}{{P}{Q}}}
+\end{equation*}
+note that we have the judgmental equalities
+\begin{align*}
+\jcomp{{{A}{P}}{Q}}{\cprojfstf{{A}{P}}{Q}}{\cprojfstf{A}{P}}
+& \jdeq 
+  \unfoldall{\jcomp{{{A}{P}}{Q}}{\cprojfstf{{A}{P}}{Q}}{\cprojfstf{A}{P}}}
+  \\
+& \jdeq 
+  \subst
+    {\ctxwk{Q}{\idtm{{A}{P}}}}
+    {\ctxwk{Q}{{\ctxext{A}{P}}{{P}{\idtm{A}}}}}
+  \\
+& \jdeq
+  \ctxwk{Q}{\subst{\idtm{{A}{P}}}{\ctxwk{\ctxext{A}{P}}{{P}{\idtm{A}}}}}
+  \\
+& \jdeq
+  \ctxwk{Q}{{P}{\idtm{A}}}
+  \\
+& \jdeq
+  \ctxwk{\ctxext{P}{Q}}{\idtm{A}}
+  \\
+& \jdeq
+  \cprojfstf{A}{{P}{Q}}
+\end{align*}
+To prove the judgmental equality
+\begin{equation*}
+\jhomeq
+  {{\Gamma}{A}}
+  {{P}{Q}}
+  {P}
+  {\jcomp{}{\cprojfstf{{A}{P}}{Q}}{\cprojsndf{A}{P}}}
+  {\jcomp{}{\cprojsndf{A}{{P}{Q}}}{\cprojfstf{P}{Q}}}
+\end{equation*}
+note that we have the judgmental equalities
+\begin{align*}
+\jcomp{{{A}{P}}{Q}}{\cprojfstf{{A}{P}}{Q}}{\cprojsndf{A}{P}}
+& \jdeq 
+  \unfoldall{\jcomp{{{A}{P}}{Q}}{\cprojfstf{{A}{P}}{Q}}{\cprojsndf{A}{P}}}
+  \\
+& \jdeq
+  \subst{\ctxwk{Q}{\idtm{{A}{P}}}}{\ctxwk{Q}{{\ctxext{A}{P}}{\idtm{P}}}}
+  \\
+& \jdeq
+  \ctxwk{Q}{\subst{\idtm{{A}{P}}}{\ctxwk{\ctxext{A}{P}}{\idtm{P}}}}
+  \\
+& \jdeq
+  \ctxwk{Q}{\idtm{P}}
+  \\
+& \jdeq
+  \unfoldall{\jcomp{{P}{Q}}{\cprojsndf{A}{{P}{Q}}}{\cprojfstf{P}{Q}}}
+  \\
+& \jdeq
+  \jcomp{{P}{Q}}{\cprojsndf{A}{{P}{Q}}}{\cprojfstf{P}{Q}}
+\end{align*}
+To prove the judgmental equality
+\begin{equation*}
+\jhomeq
+  {{{\Gamma}{A}}{P}}
+  {Q}
+  {Q}
+  {\cprojsndf{{A}{P}}{Q}}
+  {\jcomp{}{\cprojsndf{A}{{P}{Q}}}{\cprojsndf{P}{Q}}}
+\end{equation*}
+note that we have the judgmental equalities
+\begin{align*}
+\cprojsndf{{A}{P}}{Q}
+& \jdeq
+  \unfoldall{\cprojsndf{{A}{P}}{Q}}
+  \\
+& \jdeq 
+  \unfoldall{\jcomp{{P}{Q}}{\cprojsndf{A}{{P}{Q}}}{\cprojsndf{P}{Q}}}
+  \\
+& \jdeq
+  \jcomp{}{\cprojsndf{A}{{P}{Q}}}{\cprojsndf{P}{Q}}\qedhere
+\end{align*}
+\begin{comment}
+%%%% This was a proof of an old version of the statement
 Now consider a term $\jterm{\Gamma}{\ctxext{A}{{P}{Q}}}{w}$. Then we have
@@
 proving the second and the third judgmental equalities.
+\end{comment}
 \end{proof}
@@
   \tmext{\ctxwk{A}{Q}}{\ctxwk{A}{R}}{\ctxwk{A}{g}}{\ctxwk{A}{t}}
+\end{align*}
+Next, we want to prove the judgmental equality
+\begin{equation*}
+\jhomeq
+  {{{\Gamma}{A}}{\ctxwk{A}{B}}}
+  {{\ctxwk{A}{Q}}{\ctxwk{A}{R}}}
+  {\ctxwk{A}{Q}}
+  {\cprojfstf{\ctxwk{A}{Q}}{\ctxwk{A}{R}}}
+  {\ctxwk{A}{\cprojfstf{Q}{R}}}
+\end{equation*}
+Note that we have the judgmental equalities
+\begin{align*}
+\cprojfstf{\ctxwk{A}{Q}}{\ctxwk{A}{R}}
+& \jdeq
+  \unfoldall{\cprojfstf{\ctxwk{A}{Q}}{\ctxwk{A}{R}}}
+  \\
+& \jdeq
+  \ctxwk{{A}{R}}{{A}{\idtm{Q}}}
+  \\
+& \jdeq
+  \unfoldall{\ctxwk{A}{\cprojfstf{Q}{R}}}
+  \\
+& \jdeq
+  \ctxwk{A}{\cprojfstf{Q}{R}}.
+\end{align*}
+Finally, we want to prove the judgmental equality
+\begin{equation*}
+\jhomeq
+  {{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{Q}}}
+  {\ctxwk{A}{R}}
+  {\ctxwk{A}{R}}
+  {\cprojsndf{\ctxwk{A}{Q}}{\ctxwk{A}{R}}}
+  {\ctxwk{A}{\cprojsndf{Q}{R}}}
+\end{equation*}
+Note that we have the judgmental equalities
+\begin{align*}
+\cprojsndf{\ctxwk{A}{Q}}{\ctxwk{A}{R}}
+& \jdeq
+  \unfoldall{\cprojsndf{\ctxwk{A}{Q}}{\ctxwk{A}{R}}}
+  \\
+& \jdeq
+  \unfoldall{\ctxwk{A}{\cprojsndf{Q}{R}}}
+  \\
+& \jdeq
+  \ctxwk{A}{\cprojsndf{Q}{R}}.
   \qedhere
@@
   { \jterm{\Gamma}{A}{x}
-    \jfam{{{\Gamma}{A}}{P}}{Q}
-    }
-  { \jhomeq
-      {\Gamma}
-      {{\subst{x}{P}}{\subst{x}{Q}}}
-      {\subst{x}{P}}
-      {\cprojfstf{\subst{x}{P}}{\subst{x}{Q}}}
-      {\subst{x}{\cprojfstf{P}{Q}}}
-    }
-  \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jfam{{{\Gamma}{A}}{P}}{Q}
+    \jfam{{{{\Gamma}{A}}{P}}{Q}}{R}
     }
@@
       {{\Gamma}{\subst{x}{P}}}
+      {{\subst{x}{Q}}{\subst{x}{R}}}
       {\subst{x}{Q}}
-      {\subst{x}{Q}}
-      {\cprojsndf{\subst{x}{P}}{\subst{x}{Q}}}
-      {\subst{x}{\cprojsndf{P}{Q}}}
+      {\cprojfstf{\subst{x}{Q}}{\subst{x}{R}}}
+      {\subst{x}{\cprojfstf{Q}{R}}}
+    }
+  \\
+& \inference
+  { \jterm{\Gamma}{A}{x}
+    \jfam{{{{\Gamma}{A}}{P}}{Q}}{R}
+    }
+  { \jhomeq
+      {{{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}}
+      {\subst{x}{R}}
+      {\subst{x}{R}}
+      {\cprojsndf{\subst{x}{Q}}{\subst{x}{R}}}
+      {\subst{x}{\cprojsndf{Q}{R}}}
     }
@@
   \tmext{\subst{x}{Q}}{\subst{x}{R}}{\subst{x}{g}}{\subst{x}{t}}.
+\end{align*}
+Next, we want to prove the judgmental equality
+\begin{equation*}
+\jhomeq
+  {{\Gamma}{\subst{x}{P}}}
+  {{\subst{x}{Q}}{\subst{x}{R}}}
+  {\subst{x}{Q}}
+  {\cprojfstf{\subst{x}{Q}}{\subst{x}{R}}}
+  {\subst{x}{\cprojfstf{Q}{R}}}
+\end{equation*}
+Note that we have the judgmental equalities
+\begin{align*}
+\cprojfstf{\subst{x}{Q}}{\subst{x}{R}}
+& \jdeq
+  \unfoldall{\cprojfstf{\subst{x}{Q}}{\subst{x}{R}}}
+  \\
+& \jdeq
+  \ctxwk{\subst{x}{R}}{\subst{x}{\idtm{Q}}}
+  \\
+& \jdeq
+  \unfoldall{\subst{x}{\cprojfstf{Q}{R}}}
+  \\
+& \jdeq
+  \subst{x}{\cprojfstf{Q}{R}}.
+\end{align*}
+And finally we want to prove the judgmental equality
+\begin{equation*}
+\jhomeq
+  {{{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}}
+  {\subst{x}{R}}
+  {\subst{x}{R}}
+  {\cprojsndf{\subst{x}{Q}}{\subst{x}{R}}}
+  {\subst{x}{\cprojsndf{Q}{R}}}
+\end{equation*}
+Note that we have the judgmental equalities
+\begin{align*}
+\cprojsndf{\subst{x}{Q}}{\subst{x}{R}}
+& \jdeq
+  \unfoldall{\cprojsndf{\subst{x}{Q}}{\subst{x}{R}}}
+  \\
+& \jdeq
+  \unfoldall{\subst{x}{\cprojsndf{Q}{R}}}
+  \\
+& \jdeq
+  \subst{x}{\cprojsndf{Q}{R}}.
   \qedhere

### 2014-04-02 article-models-tt.tex
new section for derived notions

--- 2014-04-01 article-models-tt.tex
@@
 \end{center}
-
-In \autoref{digging_deeper} we use the framework we have developed in this
-section to derive new notions and their properties. In particular, we will
-develop the notion of \emph{extension on terms} together with the projection
-maps from an extension to the `base' context of family, the 
-\emph{family pullback} which is a version of pullbacks for families and thirdly
-the \emph{inductive morphisms} which are morphisms of type theory that allow
-to find terms of families over the codomain context by pulling them back to
-the domain context and finding a term there.
-
-In \autoref{extension-on-terms} on the extension operation on terms we will
-derive all the compatibility rules that one would expect to hold for extension
-on terms. The key to most of these results is the currying operation, which
-could be seen as the missing feature in the table above. The extension on terms
-operation depends in an essential way on the substitution operation, on the
-identity terms and therefore indirectly also on the weakening operation. Thus,
-we will see here all of the features of the theory we develop in
-\autoref{judgments} through \autoref{categorical_properties} come to the 
-surface.
-
-Next, we introduce the inclusion of the fibers $\subst{x}{P}$ into the extension
-$\ctxext{A}{P}$ as a morphism in context $\Gamma$. As was the case with
-extension on terms and with projections, there will be a ton of compatibility
-properties which we will prove about these inclusions.
-
-It should be kept in mind though that in the current formulation there is no
-sealed deal establishing a relationship between families over a context
-and any kind of morphisms -- neither with morphisms having the `base' of the
-family as its codomain nor with families into a universe (universes will be
-introduced in \autoref{universes}). The only thing we know here is
-that a family $P$ over $\ctxext{\Gamma}{A}$ determines a context morphism
-from $\ctxext{A}{P}$ to $A$ in context $\Gamma$, the projection. 
-We do not see this as
-a shortcoming of the theory of contexts families and terms. Rather, such a
-correspondence is a feature of a theory which does incorporate universes. The
-fact that we're lacking a clear connection between families and (a specified
-class of) morphisms, however, does show up in our treatment of the notion we
-called familie pullbacks. For instance, we can't show that a square of families
-is a pullback precisely when the corresponding square of projections is a
-pullback: only the backwards direction holds. 
-The discrepancies continue: ordinary pullbacks do not always exist
-whereas family pullbacks do but the composition of two family pullback squares
-need not be a family pullback square again whereas the pasting lemma of
-ordinary pullbacks holds as usually.
-We feel that pointing out what we can and can't do in the current setting is
-an important aspect of developing an intuition with the system and therefore
-we include this subsubsection even though the theory of family pullbacks
-might feel a bit different than the usual theory of pullbacks.
-
-In the last subsubsection we give a treatment of inductive morphisms. These
-morphisms appear also in the introduction of the many inductive type
-constructors in \autoref{tt_constructors} and therefore a general treatment of
-the subject is insightful.
 
@@
 \end{proof}
-
-\subsection{Digging a little deeper in the theory of contexts, families and terms}
-\label{digging_deeper}
-
-There is a good amount of things we can already state and prove about in the
-theory we have developed so far. We will develop a bit more theory that will
-come in handy later on.
-
-\emph{This subsection contains no new assumptions.}
-
-\subsubsection{Projections and extension on terms}\label{extension-on-terms}
-In this subsubsection we consider the notion of extension on terms, which has now
-become definable inside our theory. Moreover, every compatibility rule one may
-dream of is provable as well, using the compatibility rules we have introduced
-earlier.
-
-\begin{defn}
-When $\jterm{\Gamma}{A}{x}$ and $\jterm{\Gamma}{\subst{x}{P}}{u}$ are terms,
-we define 
-\begin{equation*}
-\jtermdefn
-  {\Gamma}
-  {\ctxext{A}{P}}
-  {\tmext{A}{P}{x}{u}}
-  {\unfold{\tmext{A}{P}{x}{u}}}.
-\end{equation*} 
-\end{defn}
-
-Thus, the term $\tmext{A}{P}{x}{u}$ is the pairing of $x$ and $u$. Note that because
-we have the judgmental equality 
-$\ctxwk{P}{{A}{\ctxext{A}{P}}}\jdeq\ctxwk{\ctxext{A}{P}}{\ctxext{A}{P}}$ in the
-context $\ctxext{{\Gamma}{A}}{P}$, the
-pairing function could just be defined as $\idtm{\ctxext{A}{P}}$.
-
-When we substitute by an extended term we get an equal result as when we
-substitute two consecutive times, like the way currying works.
-
-\begin{lem}
-The following inference rules are valid:
-\begin{align*}
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jterm{\Gamma}{\subst{x}{P}}{u}
-    \jfam{{{\Gamma}{A}}{P}}{Q}
-    }
-  { \jfameq
-      {\Gamma}
-      {\subst{\tmext{A}{P}{x}{u}}{Q}}
-      {\subst{u}{{x}{Q}}}
-    }
-  \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jterm{\Gamma}{\subst{x}{P}}{u}
-    \jterm{{{\Gamma}{A}}{P}}{Q}{g}
-    }
-  { \jtermeq
-      {\Gamma}
-      {\subst{u}{{x}{Q}}}
-      {\subst{\tmext{A}{P}{x}{u}}{g}}
-      {\subst{u}{{x}{g}}}
-    }
-  \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jterm{\Gamma}{\subst{x}{P}}{u}
-    \jfam{{{{\Gamma}{A}}{P}}{Q}}{R}
-    }
-  { \jfameq
-      {{\Gamma}{\subst{u}{{x}{Q}}}}
-      {\subst{\tmext{A}{P}{x}{u}}{R}}
-      {\subst{u}{{x}{R}}}
-    }
-  \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jterm{\Gamma}{\subst{x}{P}}{u}
-    \jterm{{{{\Gamma}{A}}{P}}{Q}}{R}{t}
-    }
-  { \jtermeq
-      {{\Gamma}{\subst{u}{{x}{Q}}}}
-      {\subst{u}{{x}{R}}}
-      {\subst{\tmext{A}{P}{x}{u}}{t}}
-      {\subst{u}{{x}{t}}}
-    }
-\end{align*}
-\end{lem}
-
-\begin{proof}
-We prove only the first judgmental equality. All the others are similar.
-Let $\jterm{\Gamma}{A}{x}$ and $\jterm{\Gamma}{\subst{x}{P}}{u}$
-be terms and let $\jfam{{{\Gamma}{A}}{P}}{Q}$ be a family. Then we have
-\begin{align*}
-\subst
-  {\tmext{A}{P}{x}{u}}
-  {Q} 
-& \jdeq 
-  \subst
-    {{u}{{x}{\idtm{\ctxext{A}{P}}}}}
-    {Q}
-  \tag{by definition}\\
-& \jdeq 
-  \subst
-    {{u}{{x}{\idtm{\ctxext{A}{P}}}}}
-    {{x}{\ctxwk{A}{Q}}}
-  \tag{by \autoref{defn-ws-3}}\\
-& \jdeq 
-  \subst
-    {{u}{{x}{\idtm{\ctxext{A}{P}}}}}
-    {{u}{\ctxwk{\subst{x}{P}}{\subst{x}{\ctxwk{A}{Q}}}}}
-  \tag{by \autoref{defn-ws-3}}\\
-& \jdeq 
-  \subst
-    {{u}{{x}{\idtm{\ctxext{A}{P}}}}}
-    {{u}{{x}{\ctxwk{P}{{A}{Q}}}}}
-  \tag{by \autoref{comp-sw-f}}\\
-& \jdeq 
-  \subst
-    {u}
-    {{{x}{\idtm{\ctxext{A}{P}}}}{{x}{\ctxwk{{P}{{A}{Q}}}}}}
-  \tag{by \autoref{comp-ss-f}}\\
-& \jdeq 
-  \subst
-    {u}
-    {{x}{{\idtm{\ctxext{A}{P}}}{\ctxwk{P}{{A}{Q}}}}}
-  \tag{by \autoref{comp-ss-f}}\\
-& \jdeq 
-  \subst
-    {u}
-    {{x}{{\idtm{\ctxext{A}{P}}}{\ctxwk{\ctxext{A}{P}}{Q}}}}
-  \tag{by \autoref{comp-ew-f}}\\
-& \jdeq 
-  \subst
-    {u}
-    {{x}{Q}}
-  \tag{by \autoref{idfunc-wk-defn}}
-\end{align*}
-\end{proof}
-
-We have seen above that the pairing function into $\ctxext{A}{P}$ is just the identity term on
-$\ctxext{A}{P}$. To analyze the pairing functin a little further, we will also
-need the projection maps from $\ctxext{A}{P}$ to $A$ and from $\ctxext{A}{P}$
-to $P$. We will now define these and see that the identity term of an
-extended family is the extension (or pairing) of the identity
-functions on the components in the apropriate way.
-
-To find out what the
-apropriate way is, note that
-\begin{align*}
-\ctxwk{\ctxext{A}{P}}{\ctxext{A}{P}} 
-& \jdeq 
-  \ctxwk{P}{{A}{\ctxext{A}{P}}}
-  \\
-& \jdeq 
-  \ctxext{\ctxwk{P}{{A}{A}}}{\ctxwk{P}{{A}{P}}}
-\end{align*}
-We have the term $\jterm{{\Gamma}{{A}{P}}}{\ctxwk{P}{A}}{\ctxwk{P}{\idtm{A}}}$.
-Thus we need to find out what $\subst{\ctxwk{P}{\idtm{A}}}{\ctxwk{P}{{A}{P}}}$ is:
-\begin{align*}
-\subst{\ctxwk{P}{\idtm{A}}}{\ctxwk{P}{{A}{P}}} 
-& \jdeq 
-  \ctxwk{P}{\subst{\idtm{A}}{\ctxwk{A}{P}}}
-  \\
-& \jdeq 
-  \ctxwk{P}{P},
-\end{align*}
-where we find the term $\idtm{P}$. Therefore we define:
-
-\begin{defn}
-Let $\jfam{\Gamma}{A}$ and $\jfam{{\Gamma}{A}}{P}$ be families. We define
-\begin{align*}
-\jhomdefn*
-  {\Gamma}
-  {{A}{P}}
-  {A}
-  {\cprojfstf{A}{P}}
-  {\unfold{\cprojfstf{A}{P}}}
-  \\
-\jtermdefn*
-  {\ctxext{\Gamma}{{A}{P}}}
-  {\ctxwk{P}{P}}
-  {\cprojsndf{A}{P}}
-  {\unfold{\cprojsndf{A}{P}}}
-\end{align*}
-\end{defn}
-
-We find the following inference rule, which expresses that the identity term
-is compatible with extension:
-
-\begin{lem}
-For any $\jfam{\Gamma}{A}$ and $\jfam{{\Gamma}{A}}{P}$ we have
-\begin{equation}\label{idfunc-ext-comp}
-\inference
-  { \jfam{\Gamma}{A}
-    \jfam{{\Gamma}{A}}{P}
-    }
-  { \jhomeq
-      {\Gamma}
-      {{A}{P}}{{A}{P}}
-      {\idtm{\ctxext{A}{P}}}
-      { \tmext
-          {\ctxwk{\ctxext{A}{P}}{A}}
-          {\ctxwk{\ctxext{A}{P}}{P}}
-          {\cprojfstf{A}{P}}
-          {\cprojsndf{A}{P}}
-        }
-    }
-\end{equation}
-\end{lem}
-
-\begin{proof}
-Consider the families $\jfam{\Gamma}{A}$ and $\jfam{{\Gamma}{A}}{P}$. Then
-we have the judgmental equalities
-\begin{align*}
-\tmext
-  {\ctxwk{\ctxext{A}{P}}{A}}
-  {\ctxwk{\ctxext{A}{P}}{P}}
-  {\cprojfstf{A}{P}}
-  {\cprojsndf{A}{P}}
-& \jdeq 
-  \unfold
-  { \tmext
-      {\ctxwk{\ctxext{A}{P}}{A}}
-      {\ctxwk{\ctxext{A}{P}}{P}}
-      {\cprojfstf{A}{P}}
-      {\cprojsndf{A}{P}}
-    }
-  \\
-& \jdeq 
-  \subst
-    { \idtm{P}
-      }
-    { {\ctxwk{P}{\idtm{A}}}
-      {\idtm{\ctxwk{\ctxext{A}{P}}{\ctxext{A}{P}}}}
-      }
-  \\
-& \jdeq 
-  \subst
-    { \idtm{P}
-      }
-    { {\ctxwk{P}{\idtm{A}}}
-      {\ctxwk{\ctxext{A}{P}}{\idtm{\ctxext{A}{P}}}}
-      }
-  \\
-& \jdeq 
-  \subst
-    { \idtm{P}
-      }
-    { {\ctxwk{P}{\idtm{A}}}
-      {\ctxwk{P}{{A}{\idtm{\ctxext{A}{P}}}}}
-      }
-  \\
-& \jdeq 
-  \subst
-    { \idtm{P}
-      }
-    {\ctxwk
-      {P}
-      { \subst
-        {\idtm{A}}
-        {\ctxwk{A}{\idtm{\ctxext{A}{P}}}}
-        }
-      }
-  \\
-& \jdeq 
-  \subst
-    { \idtm{A}
-      }
-    { \ctxwk{A}{\idtm{\ctxext{A}{P}}}
-      }
-  \\
-& \jdeq 
-  \idtm{\ctxext{A}{P}}.
-  \qedhere
-\end{align*}
-\end{proof}
-
-\begin{lem}
-The following inference is valid:
-\begin{equation*}
-\inference
-  { \jfam{{\Gamma}{A}}{P}
-    \jfam{{\Gamma}{A}}{Q}
-    }
-  { \jfameq
-      {{\Gamma}{{A}{P}}}
-      {\subst{\cprojfstf{A}{P}}{\ctxwk{\ctxext{A}{P}}{Q}}}
-      {\ctxwk{P}{Q}}
-    }
-\end{equation*}
-\end{lem}
-
-\begin{proof}
-Let $\jfam{{\Gamma}{A}}{P}$ and $\jfam{{\Gamma}{A}}{Q}$ be
-families. Then we have
-\begin{align*}
-\subst{\cprojfstf{A}{P}}{\ctxwk{\ctxext{A}{P}}{Q}} 
-& \jdeq 
-  \subst{\ctxwk{P}{\idtm{A}}}{\ctxwk{\ctxext{A}{P}}{Q}} 
-  \tag{by definition}\\
-& \jdeq 
-  \subst{\ctxwk{P}{\idtm{A}}}{\ctxwk{P}{{A}{Q}}} 
-  \tag{by \autoref{comp-ww-f}}\\
-& \jdeq 
-  \ctxwk{P}{\subst{\idtm{A}}{\ctxwk{A}{Q}}} 
-  \tag{by \autoref{comp-ws-f}}\\
-& \jdeq 
-  \ctxwk{P}{Q} 
-  \tag{by \autoref{idfunc-wk-defn}}
-\end{align*}
-\end{proof}
-
-The constructions of the terms $\tmext{A}{P}{x}{u}$ and $\cprojfst{A}{P}{w}$ and
-$\cprojsnd{A}{P}{w}$ are subject to various rules, with all of them being
-consequences of earlier introduced inference rules.
-
-\begin{lem} The following inference rules expressing that pairing is a strict
-inverse to the combination of decompositions, are valid:
-\begin{align*}
-& \inference
-  { \jterm{\Gamma}{\ctxext{A}{P}}{w}
-    }
-  { \jtermeq
-      {\Gamma}
-      {\ctxext{A}{P}}
-      {\tmext{A}{P}{\cprojfst{A}{P}{w}}{\cprojsnd{A}{P}{w}}}
-      {w}
-    }
-  \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jterm{\Gamma}{\subst{x}{P}}{u}
-    }
-  { \jtermeq
-      {\Gamma}
-      {A}
-      {\cprojfst{A}{P}{\tmext{A}{P}{x}{u}}}
-      {x}
-    }
-  \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jterm{\Gamma}{\subst{x}{P}}{u}
-    }
-  { \jtermeq
-      {\Gamma}
-      {\subst{x}{P}}
-      {\cprojsnd{A}{P}{\tmext{A}{P}{x}{u}}}
-      {u}
-    }
-\end{align*}
-\end{lem}
-
-\begin{proof}
-To prove the first judgmental equality, note that
-\begin{align*}
-w 
-& \jdeq 
-  \subst{w}{\idtm{\ctxext{A}{P}}} 
-  \tag{by \autoref{idfunc-subst-defn}}\\
-& \jdeq 
-  \subst
-    { w}
-    { { \idtm{P}}
-      { { \ctxwk{P}{\idtm{A}}}
-        { \idtm{\ctxwk{\ctxext{A}{P}}{\ctxext{A}{P}}}}
-        }
-      }
-  \tag{by \autoref{idfunc-ext-comp}}\\
-& \jdeq 
-  \subst
-    { {w}
-      {\idtm{P}}
-      }
-    { {w}
-      { { \ctxwk{P}{\idtm{A}}
-          }
-        { \idtm{\ctxwk{\ctxext{A}{P}}{\ctxext{A}{P}}}
-          }
-        }
-      }
-  \tag{by \autoref{comp-ss-t}}\\
-& \jdeq 
-  \subst
-    { {w}
-      {\idtm{P}}
-      }
-    { { {w}
-        {\ctxwk{P}{\idtm{A}}}
-        }
-      { {w}
-        {\idtm{\ctxwk{\ctxext{A}{P}}{\ctxext{A}{P}}}}
-        }
-      }
-  \tag{by \autoref{comp-ss-t}}\\
-& \jdeq 
-  \subst
-    { {w}
-      {\idtm{P}}
-      }
-    { { {w}
-        {\ctxwk{P}{\idtm{A}}}
-        }
-      { {w}
-        {\ctxwk{\ctxext{A}{P}}{\idtm{\ctxext{A}{P}}}}
-        }
-      }
-  \tag{by \autoref{idfunc-wk-comp}}\\
-& \jdeq 
-  \subst
-    { {w}
-      {\idtm{P}}
-      }
-    { { {w}
-        {\ctxwk{P}{\idtm{A}}}
-        }
-      { \idtm{\ctxext{A}{P}}
-        }
-      }
-  \tag{by \autoref{defn-ws-4}}\\
-& \jdeq 
-  \tmext{A}{P}{\cprojfst{A}{P}{w}}{\cprojsnd{A}{P}{w}}
-  \tag{by definition}
-\end{align*}
-To prove the second judgmental equality, let $\jterm{\Gamma}{A}{x}$ and
-$\jterm{\Gamma}{\subst{x}{P}}{u}$. Then we have
-\begin{align*}
-\cprojfst{A}{P}{\tmext{A}{P}{x}{u}}
-& \jdeq 
-  \subst{\tmext{A}{P}{x}{u}}{\ctxwk{P}{\idtm{A}}}
-  \\
-& \jdeq 
-  \subst{u}{{x}{\ctxwk{P}{\idtm{A}}}} 
-  \\
-& \jdeq 
-  \subst{u}{\ctxwk{\subst{x}{P}}{\subst{x}{\idtm{A}}}}
-  \\
-& \jdeq 
-  \subst{x}{\idtm{A}}
-  \\
-& \jdeq 
-  x.
-\end{align*}
-To prove the third judgmental equality, note that
-\begin{align*}
-\cprojsnd{A}{P}{\tmext{A}{P}{x}{u}}
-& \jdeq 
-  \subst{\tmext{A}{P}{x}{u}}{\idtm{P}}
-  \\
-& \jdeq 
-  \subst{u}{{x}{\idtm{P}}}
-  \\
-& \jdeq 
-  \subst{u}{\idtm{\subst{x}{P}}}
-  \\
-& \jdeq 
-  u.
-  \qedhere
-\end{align*}
-\end{proof}
-
-\begin{lem}
-The following compatibility rules for two consecutive term extensions is valid:
-\begin{align*}
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jterm{\Gamma}{\subst{x}{P}}{u}
-    \jterm{\Gamma}{\subst{\tmext{A}{P}{x}{u}}{Q}}{v}
-    }
-  { \jtermeq
-      {\Gamma}
-      {\ctxext{{A}{P}}{Q}}
-      {\tmext{A}{{P}{Q}}{x}{{\subst{x}{P}}{\subst{x}{Q}}{u}{v}}}
-      {\tmext{{A}{P}}{Q}{{A}{P}{x}{u}}{v}}
-    }
-  \\
-& \inference
-  { \jfam{{{\Gamma}{A}}{P}}{Q}
-    }
-  { \jhomeq
-      {\Gamma}
-      {\ctxext{{A}{P}}{Q}}
-      {A}
-      {\jcomp{}{\cprojfstf{{A}{P}}{Q}}{\cprojfstf{A}{P}}}
-      {\cprojfstf{A}{{P}{Q}}}
-    }
-  \\
-& \inference
-  { \jfam{{{\Gamma}{A}}{P}}{Q}
-    }
-  { \jhomeq
-      {{\Gamma}{A}}
-      {{P}{Q}}
-      {P}
-      {\jcomp{}{\cprojfstf{{A}{P}}{Q}}{\cprojsndf{A}{P}}}
-      {\jcomp{}{\cprojsndf{A}{{P}{Q}}}{\cprojfstf{P}{Q}}}
-    }
-  \\
-& \inference
-    { \jfam{{{\Gamma}{A}}{P}}{Q}
-      }
-    { \jhomeq
-        {{{\Gamma}{A}}{P}}
-        {Q}
-        {Q}
-        {\cprojsndf{{A}{P}}{Q}}
-        {\jcomp{}{\cprojsndf{A}{{P}{Q}}}{\cprojsndf{P}{Q}}}
-      }
-\end{align*}
-\end{lem}
-
-\begin{proof}
-Consider terms $\jterm{\Gamma}{A}{x}$, $\jterm{\Gamma}{\subst{x}{P}}{u}$ and
-$\jterm{\Gamma}{\subst{u}{{x}{Q}}}{v}$. Then we have
-\begin{align*}
-\tmext{A}{{P}{Q}}{x}{{\subst{x}{P}}{\subst{x}{Q}}{u}{v}} 
-& \jdeq 
-  \subst
-    {\tmext{\subst{x}{P}}{\subst{x}{Q}}{u}{v}}
-    {{x}{\idtm{\ctxext{A}{{P}{Q}}}}}
-  \\
-& \jdeq 
-  \subst{v}{{u}{{x}{\idtm{\ctxext{A}{{P}{Q}}}}}}
-  \\
-& \jdeq 
-  \subst{v}{{u}{{x}{\idtm{\ctxext{{A}{P}}{Q}}}}}
-  \\
-& \jdeq 
-  \subst{v}{{\tmext{A}{P}{x}{u}}{\idtm{\ctxext{{A}{P}}{Q}}}}
-  \\
-& \jdeq 
-  \tmext{{A}{P}}{Q}{{A}{P}{x}{u}}{v}.
-\end{align*}
-Now consider a term $\jterm{\Gamma}{\ctxext{A}{{P}{Q}}}{w}$. Then we have
-\begin{align*}
-w 
-& \jdeq 
-  \tmext
-    {A}
-    {{P}{Q}}
-    {\cprojfst{A}{\ctxext{P}{Q}}{w}}
-    {\cprojsnd{A}{\ctxext{P}{Q}}{w}}
-  \\
-& \jdeq 
-  \tmext
-    {A}
-    {{P}{Q}}
-    {\cprojfst{A}{\ctxext{P}{Q}}{w}}
-    { {P}
-      {Q}
-      {\cprojfst{P}{Q}{\cprojsnd{A}{\ctxext{P}{Q}}{w}}}
-      {\cprojsnd{P}{Q}{\cprojsnd{A}{\ctxext{P}{Q}}{w}}}
-      }
-  \\
-& \jdeq 
-  \tmext
-    {{A}{P}}
-    {Q}
-    { {} % need to provide base and family, but there's no unfold.
-      {}
-      {\cprojfst{A}{\ctxext{P}{Q}}{w}}
-      {\cprojfst{P}{Q}{\cprojsnd{A}{\ctxext{P}{Q}}{w}}}
-      }
-    { \cprojsnd{P}{Q}{\cprojsnd{A}{\ctxext{P}{Q}}{w}}
-      }
-\end{align*}
-Thus we see that 
-\begin{align*}
-\cprojfst{\ctxext{A}{P}}{Q}{w} 
-& \jdeq 
-  \tmext
-    {A}
-    {P}
-    {\cprojfst{A}{\ctxext{P}{Q}}{w}}
-    {\cprojfst{P}{Q}{\cprojsnd{A}{\ctxext{P}{Q}}{w}}}
-  \\ 
-\cprojsnd{\ctxext{A}{P}}{Q}{w} 
-& \jdeq 
-  \cprojsnd{P}{Q}{\cprojsnd{A}{\ctxext{P}{Q}}{w}},
-\end{align*}
-proving the fourth judgmental equality, and therefore also that
-\begin{align*}
-\cprojfst{A}{P}{\cprojfst{\ctxext{A}{P}}{Q}{w}} 
-& \jdeq 
-  \cprojfst{A}{\ctxext{P}{Q}}{w}
-  \\
-\cprojsnd{A}{P}{\cprojfst{\ctxext{A}{P}}{Q}{w}} 
-& \jdeq 
-  \cprojfst{P}{Q}{\cprojsnd{A}{\ctxext{P}{Q}}{w}},
-\end{align*}
-proving the second and the third judgmental equalities.
-\end{proof}
-
-\begin{lem}
-When we weaken a term $\tmext{B}{Q}{y}{v}$ of $\ctxext{B}{Q}$ in context $\Gamma$ by
-a family $A$, the term that we get is $\tmext{\ctxwk{A}{B}}{\ctxwk{A}{Q}}{\ctxwk{A}{y}}{\ctxwk{A}{v}}$. More
-precisely, the following inference rules are valid:
-\begin{align*}
-& \inference
-  { \jterm{{\Gamma}{B}}{Q}{g}
-    \jterm{{\Gamma}{B}}{\subst{g}{R}}{t}
-    }
-  { \jtermeq
-      {{{\Gamma}{A}}{\ctxwk{A}{B}}}
-      {\ctxwk{A}{\ctxext{Q}{R}}}
-      {\ctxwk{A}{\tmext{Q}{R}{g}{t}}}
-      {\tmext{\ctxwk{A}{Q}}{\ctxwk{A}{R}}{\ctxwk{A}{g}}{\ctxwk{A}{t}}}
-    }
-  \\
-& \inference
-  { \jfam{{{\Gamma}{B}}{Q}}{R}
-    }
-  { \jhomeq
-      {{{\Gamma}{A}}{\ctxwk{A}{B}}}
-      {{\ctxwk{A}{Q}}{\ctxwk{A}{R}}}
-      {\ctxwk{A}{Q}}
-      {\cprojfstf{\ctxwk{A}{Q}}{\ctxwk{A}{R}}}
-      {\ctxwk{A}{\cprojfstf{Q}{R}}}
-    }
-  \\
-& \inference
-  { \jfam{{{\Gamma}{B}}{Q}}{R}
-    }
-  { \jhomeq
-      {{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{Q}}}
-      {\ctxwk{A}{R}}
-      {\ctxwk{A}{R}}
-      {\cprojsndf{\ctxwk{A}{Q}}{\ctxwk{A}{R}}}
-      {\ctxwk{A}{\cprojsndf{Q}{R}}}
-    }
-\end{align*}
-\end{lem}
-
-\begin{proof}
-Consider $\jterm{{\Gamma}{B}}{Q}{g}$ and $\jterm{{\Gamma}{B}}{\subst{g}{R}}{t}$.
-Then we have the judgmental equalities
-\begin{align*}
-\ctxwk{A}{\ctxext{Q}{R}{g}{t}}
-& \jdeq 
-  \ctxwk{A}{\subst{t}{{g}{\idtm{\ctxext{Q}{R}}}}}
-  \\
-& \jdeq 
-  \subst{\ctxwk{A}{t}}{\ctxwk{A}{\subst{g}{\idtm{\ctxext{Q}{R}}}}}
-  \\
-& \jdeq 
-  \subst{\ctxwk{A}{t}}{{\ctxwk{A}{g}}{\ctxwk{A}{\idtm{\ctxext{Q}{R}}}}}
-  \\
-& \jdeq 
-  \subst{\ctxwk{A}{t}}{{\ctxwk{A}{g}}{\idtm{\ctxwk{A}{\ctxext{Q}{R}}}}}
-  \\
-& \jdeq 
-  \subst{\ctxwk{A}{t}}{{\ctxwk{A}{g}}{\idtm{\ctxext{\ctxwk{A}{Q}}{\ctxwk{A}{R}}}}}
-  \\
-& \jdeq 
-  \tmext{\ctxwk{A}{Q}}{\ctxwk{A}{R}}{\ctxwk{A}{g}}{\ctxwk{A}{t}}
-  \qedhere
-\end{align*}
-\end{proof}
-
-\begin{lem}
-When we substitute an extended term $\tmext{P}{Q}{f}{g}$ of $\ctxext{P}{Q}$ by a term
-$x$ of $A$, the term that we get is $\tmext{\subst{x}{P}}{\subst{x}{Q}}{\subst{x}{f}}{\subst{x}{g}}$.
-More precisely, the following inference rules are valid:
-\begin{align*}
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jterm{{{\Gamma}{A}}{P}}{Q}{g}
-    \jterm{{{\Gamma}{A}}{P}}{\subst{g}{R}}{t}
-    }
-  { \jtermeq
-      {{\Gamma}{\subst{x}{P}}}
-      {\ctxext{\subst{x}{Q}}{\subst{x}{R}}}
-      {\subst{x}{\tmext{Q}{R}{g}{t}}}
-      {\tmext{\subst{x}{Q}}{\subst{x}{R}}{\subst{x}{g}}{\subst{x}{t}}}
-    }
-  \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jfam{{{\Gamma}{A}}{P}}{Q}
-    }
-  { \jhomeq
-      {\Gamma}
-      {{\subst{x}{P}}{\subst{x}{Q}}}
-      {\subst{x}{P}}
-      {\cprojfstf{\subst{x}{P}}{\subst{x}{Q}}}
-      {\subst{x}{\cprojfstf{P}{Q}}}
-    }
-  \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jfam{{{\Gamma}{A}}{P}}{Q}
-    }
-  { \jhomeq
-      {{\Gamma}{\subst{x}{P}}}
-      {\subst{x}{Q}}
-      {\subst{x}{Q}}
-      {\cprojsndf{\subst{x}{P}}{\subst{x}{Q}}}
-      {\subst{x}{\cprojsndf{P}{Q}}}
-    }
-\end{align*}
-\end{lem}
-
-\begin{proof}
-Consider $\jterm{{\Gamma}{B}}{Q}{g}$ and $\jterm{{\Gamma}{B}}{\subst{g}{R}}{t}$.
-Then we have the judgmental equalities
-\begin{align*}
-\subst{x}{\tmext{Q}{R}{g}{t}}
-& \jdeq 
-  \subst{x}{{t}{{g}{\idtm{\ctxext{Q}{R}}}}}
-  \\
-& \jdeq 
-  \subst{{x}{t}}{{x}{{g}{\idtm{\ctxext{Q}{R}}}}}
-  \\
-& \jdeq 
-  \subst{{x}{t}}{{{x}{g}}{{x}{\idtm{\ctxext{Q}{R}}}}}
-  \\
-& \jdeq 
-  \subst{{x}{t}}{{{x}{g}}{\idtm{\subst{x}{\ctxext{Q}{R}}}}}
-  \\
-& \jdeq 
-  \subst{{x}{t}}{{{x}{g}}{\idtm{\ctxext{\subst{x}{Q}}{\subst{x}{R}}}}}
-  \\
-& \jdeq 
-  \tmext{\subst{x}{Q}}{\subst{x}{R}}{\subst{x}{g}}{\subst{x}{t}}.
-  \qedhere
-\end{align*}
-\end{proof}
-
-\subsubsection{Fiber inclusions}
-We will use the insights of \autoref{extension-on-terms} to define and study
-\emph{fiber inclusions}. The fiber inclusion of the \emph{fiber}
-$\subst{x}{P}$ into the extension $\ctxext{A}{P}$ is a morphism
-$\jhom{\Gamma}{\subst{x}{P}}{{A}{P}}{\finc{x}{P}}$, for any family
-$\jfam{{\Gamma}{A}}{P}$ and any term $\jterm{\Gamma}{A}{x}$. Then we will determine
-the ways in which it is compatible with the other operators. Note that in this
-subsubsection we will focus on the compatibility properties; the fact that
-the fiber inclusions also appear in a pullback diagram will be established in
-\autoref{pullback}. Because 
-$\finc{x}{P}$ is a morphism into an extension, we will first study morphisms
-into extensions more generally.
-
-For the following lemma, recall that the judgment $\jhom{\Gamma}{A}{{B}{Q}}{f}$
-unfolds as
-\begin{equation*}
-\unfold{\jhom{\Gamma}{A}{{B}{Q}}{f}}
-\end{equation*}
-and that we have the judgmental equality 
-$ \jfameq
-    {{\Gamma}{A}}
-    {\ctxwk{A}{\ctxext{B}{Q}}}
-    {\ctxext{\ctxwk{A}{B}}{\ctxwk{A}{Q}}}.
-  $
-Therefore, each morphism into an extended family can itself be described as
-an extended term. The following lemma explains how this goes.
-
-\begin{lem}
-Let $\jhom{\Gamma}{A}{{B}{Q}}{f}$ be a morphism from $A$ to $\ctxext{B}{Q}$
-in a context $\Gamma$. Then we have
-\begin{equation*}
-\jhomeq
-  {\Gamma}
-  {A}
-  {{B}{Q}}
-  {f}
-  {\tmext{}{}{\jcomp{}{f}{\cprojfstf{B}{Q}}}{\jcomp{}{f}{\cprojsndf{B}{Q}}}}.
-\end{equation*}
-Alternatively, when $\jhom{\Gamma}{A}{B}{f_0}$ and 
-$\jterm{{\Gamma}{A}}{\jcomp{}{f_0}{Q}}{f_1}$ we obtain a morphism
-$\jhom{\Gamma}{A}{{B}{Q}}{\tmext{}{}{f_0}{f_1}}$ with the property that
-\begin{align*}
-\jhomeq*{\Gamma}{A}{B}{\jcomp{}{\tmext{}{}{f_0}{f_1}}{\cprojfstf{B}{Q}}}{f_0}\\
-\jtermeq*{{\Gamma}{A}}{\jcomp{}{f_0}{Q}}{\jcomp{}{\tmext{}{}{f_0}{f_1}}{\cprojsndf{B}{Q}}}{f_1}.
-\end{align*}
-\end{lem}
-
-\begin{proof}
-Let $\jhom{\Gamma}{A}{{B}{Q}}{f}$ be a morphism in context $\Gamma$. Then we
-have the judgmental equalities
-\begin{align*}
-\cprojfst{\ctxwk{A}{B}}{\ctxwk{A}{Q}}{f}
-& \jdeq
-  \subst{f}{\ctxwk{A}{\cprojfstf{B}{Q}}}
-  \\
-& \jdeq
-  \jcomp{}{f}{\cprojfstf{B}{Q}}
-  \\
-\cprojsnd{\ctxwk{A}{B}}{\ctxwk{A}{Q}}{f}
-& \jdeq
-  \subst{f}{\ctxwk{A}{\cprojsndf{B}{Q}}}
-  \\
-& \jdeq
-  \jcomp{}{f}{\cprojsndf{B}{Q}}
-\end{align*}
-The alternative formulation of the statement is a direct corollary.
-\end{proof}
-
-We also have the following lemma about the compatibility of pairing and composition:
-
-\begin{lem}
-The following inference rule is valid
-\begin{align*}
-& \inference
-  { \jhom{\Gamma}{A}{B}{f}
-    \jhom{\Gamma}{B}{C}{g}
-    \jfam{{\Gamma}{C}}{R}
-    \jterm{{\Gamma}{B}}{\subst{g}{\ctxwk{B}{R}}}{w}
-    }
-  { \jhomeq
-      {\Gamma}
-      {A}
-      {{C}{R}}
-      {\jcomp{A}{f}{\tmext{\ctxwk{B}{C}}{\ctxwk{B}{R}}{g}{w}}}
-      {\tmext{\ctxwk{A}{C}}{\ctxwk{A}{R}}{\jcomp{A}{f}{g}}{\jcomp{A}{f}{w}}}
-    }
-\end{align*}
-\end{lem}
-
-\begin{proof}
-Let $\jhom{\Gamma}{A}{B}{f}$, $\jhom{\Gamma}{B}{C}{g}$, $\jfam{{\Gamma}{C}}{R}$
-and $\jterm{{\Gamma}{B}}{\subst{g}{\ctxwk{B}{R}}}{w}$. Then we have the
-judgmental equalities
-\begin{align*}
-\jcomp{A}{f}{\tmext{\ctxwk{B}{C}}{\ctxwk{B}{R}}{g}{w}}
-& \jdeq 
-  \subst{f}{\ctxwk{A}{\tmext{\ctxwk{B}{C}}{\ctxwk{B}{R}}{g}{w}}}
-  \\
-& \jdeq 
-  \subst
-    {f}
-    {\tmext{\ctxwk{A}{{B}{C}}}{\ctxwk{A}{{B}{R}}}{\ctxwk{A}{g}}{\ctxwk{A}{w}}}
-  \\
-& \jdeq 
-  \tmext
-    {\ctxwk{A}{C}}
-    {\ctxwk{A}{R}}
-    {\subst{f}{\ctxwk{A}{g}}}
-    {\subst{f}{\ctxwk{A}{w}}}
-  \\
-& \jdeq 
-  \tmext{\ctxwk{A}{C}}{\ctxwk{A}{R}}{\jcomp{A}{f}{g}}{\jcomp{A}{f}{w}}.
-  \qedhere
-\end{align*}
-\end{proof}
-
-\begin{defn}
-Let $\jterm{\Gamma}{A}{x}$ be a term and let $\jfam{{\Gamma}{A}}{P}$ be a
-family. Then we define
-\begin{equation*}
-\jhomdefn{\Gamma}{\subst{x}{P}}{{A}{P}}{\finc{x}{P}}{\unfoldnext{\finc{x}{P}}}
-\end{equation*}
-\end{defn}
-
-\subsubsection{Pullback squares and family pullback squares}
-\label{pullback}
-Now that we have introduced the notions of morphisms, composition and identity
-terms, we can develop a diagramatic style of of displaying type dependencies
-combined with morphisms. We give an informal, metatheoretical definition of
-such diagrams by indicating what the various components mean. The definition
-is informal because we will only use such diagrams occasionally to provide a
-graphical indication of the situation in which we're working. In particular,
-we will not shy away from using natural numbers and trust that the reader can
-figure out what we mean.
-
-\begin{defn}
-A diagram is said to be a \emph{dependency diagram in context $\Gamma$}
-if it is built up according to the following steps:
-\begin{itemize}
-\item The arrows appearing in a dependency diagram are either ordinary, like the
-arrow%
-$\begin{tikzcd}[ampersand replacement = \&]
-X \ar{r} \& Y,
-\end{tikzcd}$
-or double-headed, like
-$\begin{tikzcd}[ampersand replacement = \&]
-X \ar[fib]{r} \& Y.
-\end{tikzcd}$
-\item An ordinary arrow 
-\begin{equation*}
-\begin{tikzcd}
-A \ar{r}{f} & B
-\end{tikzcd}
-\end{equation*}
-between two families $A$ and $B$ of contexts over $\Gamma$ indicates that
-$f$ is a morphism from $A$ to $B$ in context $\Gamma$, i.e.~that we have the
-judgment $\jhom{\Gamma}{A}{B}{f}$.
-\item The set of double-headed arrows must form a forest and the root of
-each maximal tree of double-headed arrows is a family of contexts over $\Gamma$.
-In particular, if an object is not the domain of a double-headed arrow it must
-be a family of contexts over $\Gamma$.
-\item A sequence of double-headed 
-arrows
-\begin{equation*}
-\begin{tikzcd}
-P_{n} \ar[fib]{r} & \cdots \ar[fib]{r} & P_1 \ar[fib]{r} & A
-\end{tikzcd}
-\end{equation*}
-indicates that $P_1$ is a family of contexts over $\ctxext{\Gamma}{A}$, that
-$P_2$ is a family of contexts over $\ctxext{{\Gamma}{A}}{P_1}$, etcetera.
-\item There can be two kinds of ladders of double-headed arrows:
-\begin{equation*}
-\begin{tikzcd}
-P_{n} \ar{r}{F_{n}} \ar[fib]{d} & Q_{n} \ar[fib]{d}\\
-\vdots \ar[fib]{d} & \vdots \ar[fib]{d}\\
-P_1 \ar{r}{F_1} \ar[fib]{d} & Q_1 \ar[fib]{d}\\
-A \ar{r}{f} & B
-\end{tikzcd}
-\qquad
-\begin{tikzcd}[column sep = tiny]
-P_{n+m} \ar{rr}{F_{n+m}} \ar[fib]{d} & & Q_{n+m} \ar[fib]{d}\\
-\vdots \ar[fib]{d} & & \vdots \ar[fib]{d}\\
-P_{n+1} \ar{rr}{F_{n+1}} \ar[fib]{dr} & & Q_{n+1} \ar[fib]{dl}\\
-& P_n \ar[fib]{d}\\
-& \vdots \ar[fib]{d}\\
-& P_1 \ar[fib]{d}\\
-& A
-\end{tikzcd}
-\end{equation*}
-The ladder on the left 
-indicates that $F_1$ is a morphism from $P_1$ to $Q_1$ \emph{over} $f$,
-i.e.~that the judgment $\jfhom{\Gamma}{A}{B}{f}{P_1}{Q_1}{F_1}$ holds, that
-$F_2$ is a morphism from $P_2$ to $Q_2$ over
-the morphism $\tmext{}{}{\ctxwk{P_1}{f}}{F_1}$ from $\ctxext{A}{P_1}$ to
-$\ctxext{B}{Q_1}$, etcetera.
-
-The ladder on the right indicates that $F_{n+1}$ is a morphism from $P_{n+1}$ to
-$Q_{n+1}$ in the appropriate context, that $F_{n+2}$ is a morphism from
-$P_{n+2}$ to $Q_{n+2}$ over $F_{n+1}$, etcetera.
-
-Note that the object(s) at the bottom of a ladder are always families of contexts
-over $\Gamma$, so that the typing of the various ingredients makes sense.
-\end{itemize}
-Such a diagram is said to be commutative if the subdiagram consisting of only
-the normal headed arrows is commutative in the usual sense (using judgmental
-equality). Note that the ladders are inherently commutative.
-\end{defn}
-
-The most basic illustrative example of a commutative dependency diagram is
-the diagram
-\begin{equation*}
-\begin{tikzcd}
-P \ar[fib]{d} \ar{r}{F} & Q \ar[fib]{d} \\
-A \ar{r}{f} & B
-\end{tikzcd}
-\end{equation*}
-indicating a morphism $F$ from $P$ to $Q$ over the morphism $f$ from $A$ to
-$B$ in a context $\Gamma$.
-
-We can just copy the usual categorical definition of a pullback square to our
-current situation, but we have to require that each arrow in the pullback square
-is an ordinary arrow. When families (i.e. double-headed arrows) are involved
-in the diagram, we make the following definition of a family pullback:
-
-\begin{defn}
-We say that a commutative dependency diagram of the form
-\begin{equation*}
-\begin{tikzcd}
-P \ar[fib]{d} \ar{r}{F} & Q \ar[fib]{d} \\
-A \ar{r}{f} & B
-\end{tikzcd}
-\end{equation*}
-is a \emph{family pullback} if the following inference rules are valid:
-\begin{align*}
-& \inference
-  { \jfam{{\Gamma}{A}}{P'}
-    \jfhom{\Gamma}{A}{B}{f}{P'}{Q}{F'}
-    }
-  { \jhom{{\Gamma}{A}}{P'}{P}{u}
-    }
-  \\
-& \inference
-  { \jfam{{\Gamma}{A}}{P'}
-    \jfhom{\Gamma}{A}{B}{f}{P'}{Q}{F'}
-    }
-  { \jfhomeq{\Gamma}{A}{B}{f}{P'}{Q}{\jcomp{}{u}{F}}{F'}
-    }
-  \\
-& \inference
-  { \jhom{{\Gamma}{A}}{P'}{P}{v}
-    \jfhomeq{\Gamma}{A}{B}{f}{P'}{Q}{\jcomp{}{v}{F}}{F'}
-    }
-  { \jhomeq{{\Gamma}{A}}{P'}{P}{v}{u}
-    }
-\end{align*}
-\end{defn}
-
-The following lemma explains that when a square involving families is a
-family pullback square whenever the corresponding square involving projections is a
-pullback square. There is no proof in the the opposite direction.
-
-\begin{lem}
-A square
-\begin{equation}\label{eq:fpb_to_pb_eqv_fpb}
-\begin{tikzcd}
-P \ar[fib]{d} \ar{r}{F} & Q \ar[fib]{d} \\
-A \ar{r}{f} & B
-\end{tikzcd}
-\end{equation}
-is a family pullback square whenever the square
-\begin{equation}\label{eq:fpb_to_pb_eqv_pb}
-\begin{tikzcd}[column sep = large]
-\ctxext{A}{P} \ar{d}[swap]{\cprojfstf{A}{P}} \ar{r}{\tmext{}{}{\ctxwk{P}{f}}{F}} & \ctxext{B}{Q} \ar{d}{\cprojfstf{B}{Q}} \\
-A \ar{r}{f} & B
-\end{tikzcd}
-\end{equation}
-is a pullback square.
-\end{lem}
-
-The family pullback of a family along any morphism always exists. It is simply given
-by the precomposition of the family with the morphism. Note that this fact does
-not carry over to arbitrary pullbacks.
-
-\begin{lem}
-The diagram
-\begin{equation*}
-\begin{tikzcd}
-\jcomp{}{f}{Q} \ar[fib]{d} \ar{r}{\idtm{\jcomp{}{f}{Q}}} & Q \ar[fib]{d} \\
-A \ar{r}{f} & B
-\end{tikzcd}
-\end{equation*}
-is a family pullback diagram.
-\end{lem}
-
-\begin{proof}
-The proof is a triviality because $\jhom{{\Gamma}{A}}{P'}{\jcomp{}{f}{Q}}{F'}$
-is the same judgment as $\jfhom{\Gamma}{A}{B}{f}{P'}{Q}{F'}$ and
-$\jcomp{}{\idtm{\jcomp{}{f}{Q}}}{F'}\jdeq F'$.
-\end{proof}
-
-For arbitrary pullbacks we have the pasting lemma as usual, but for family
-pullbacks we can only derive one of the parts of the pasting lemma.
-
-\begin{lem}
-Suppose we have the diagram
-\begin{equation*}
-\begin{tikzcd}
-P \ar{r}{F} \ar[fib]{d} & Q \ar{r}{G} \ar[fib]{d} & R \ar[fib]{d}\\
-A \ar{r}{f} & B \ar{r}{g} & C
-\end{tikzcd}
-\end{equation*}
-where the square on the right and the outer rectangle are family pullback 
-diagrams. Then the square on the left is a family pullback diagram.
-\end{lem}
-
-\begin{proof}
-Let $\jfam{{\Gamma}{A}}{P'}$ be a family and let $\jfhom{\Gamma}{A}{B}{f}
-{P'}{Q}{F}$ be a morphism over $f$.
-\begin{itemize}
-\item Then we compose $F'$ with $G$ to obtain a morphism over $\jcomp{}{f}{g}$.
-\item Then we get $\jhom{{\Gamma}{A}}{P'}{P}{u}$ with a uniqueness property.
-      The property that $\jcomp{}{u}{F}\jdeq F'$ follows from the assumption
-      that the right square is a pullback.
-\item Now assume that we have another such $v$. Compose it with $F$ and $G$.
-      By the assumed properties this is the same as $u$ composed with $F$ and
-      $G$. By the pullback condition we now get $u\jdeq v$. 
-\end{itemize}
-\end{proof}
-
-\begin{lem}
-For any $\jterm{\Gamma}{A}{x}$ and any $\jfam{{\Gamma}{A}}{P}$, the square
-\begin{equation*}
-\begin{tikzcd}
-\subst{x}{P} \ar{d} \ar{r}{\finc{x}{P}} & \ctxext{A}{P} \ar{d}{\cprojfstf{A}{P}}\\
-\emptyf \ar{r}{\ctxwk{\emptyf}{x}} & A
-\end{tikzcd}
-\end{equation*}
-is a pullback square.
-\end{lem}
-
-In the following lemma we assert that pulling back a family $Q$ 
-over $\ctxext{{\Gamma}{A}}{P}$ along a fiber
-inclusion $\finc{x}{P}$ gives the family $\subst{x}{Q}$ over $\ctxext{\Gamma}{\subst{x}{P}}$.
-
-\begin{lem}
-We have the following inference rule:
-\begin{equation*}
-\inference
-  { \jfam{{{\Gamma}{A}}{P}}{Q}
-    \jterm{\Gamma}{A}{x}
-    }
-  { \jfameq
-      {{\Gamma}{\subst{x}{P}}}
-      {\jcomp{}{\finc{x}{P}}{Q}}
-      {\subst{x}{Q}}
-    }
-\end{equation*}
-\end{lem}
-
-\begin{proof}
-We have the judgmental equalities:
-\begin{align*}
-\jcomp{\subst{x}{P}}{\finc{x}{P}}{Q}
-& \jdeq
-  \subst{\tmext{}{}{\ctxwk{\subst{x}{P}}{x}}{\idtm{\subst{x}{P}}}}{\ctxwk{\subst{x}{P}}{Q}}
-  \\
-& \jdeq
-  \subst{\idtm{\subst{x}{P}}}{{\ctxwk{\subst{x}{P}}{x}}{\ctxwk{\subst{x}{P}}{Q}}}
-  \\
-& \jdeq
-  \subst{\idtm{\subst{x}{P}}}{\ctxwk{\subst{x}{P}}{\subst{x}{Q}}}
-  \\
-& \jdeq
-  \subst{x}{Q}.
-\end{align*}
-\end{proof}
-
-\subsubsection{Another special case of projections}
-In this subsubsection we investigate the special case of a projection which
-appears as a morphism from $\ctxext{{A}{P}}{\ctxwk{P}{Q}}$ to $\ctxext{A}{Q}$
-in context $\Gamma$, where we assume to have the families 
-$\jfam{{\Gamma}{A}}{P}$ and $\jfam{{\Gamma}{A}}{Q}$.
-
-Note that we have the judgmental
-equalities
-\begin{align*}
-\ctxwk{\ctxext{{A}{P}}{\ctxwk{P}{\mfam{A}}}}{\ctxext{A}{\mfam{A}}}
-& \jdeq 
-  \ctxext
-    {\ctxwk{\ctxext{{A}{P}}{\ctxwk{P}{\mfam{A}}}}{A}}
-    {\ctxwk{\ctxext{{A}{P}}{\ctxwk{P}{\mfam{A}}}}{\mfam{A}}}
-  \\
-& \jdeq
-  \ctxext
-    {\ctxwk{{P}{\mfam{A}}}{{\ctxext{A}{P}}{A}}}
-    {\ctxwk{\ctxext{{A}{P}}{\ctxwk{P}{\mfam{A}}}}{\mfam{A}}}
-\end{align*}
-Note that we have the term $\ctxwk{{P}{\mfam{A}}}{\cprojfstf{A}{P}}$ of the
-family $\ctxwk{{P}{\mfam{A}}}{{\ctxext{A}{P}}{A}}$. Therefore, we need to
-find a term of type $\subst{\ctxwk{{P}{\mfam{A}}}{\cprojfstf{A}{P}}}
-{\ctxwk{\ctxext{{A}{P}}{\ctxwk{P}{\mfam{A}}}}{\mfam{A}}}$. Note that we have
-the judgmental equalities:
-\begin{align*}
-\subst
-  {\ctxwk{{P}{\mfam{A}}}{\cprojfstf{A}{P}}}
-  {\ctxwk{\ctxext{{A}{P}}{\ctxwk{P}{\mfam{A}}}}{\mfam{A}}}
-& \jdeq
-  \subst
-    {\ctxwk{{P}{\mfam{A}}}{\cprojfstf{A}{P}}}
-    {\ctxwk{{P}{\mfam{A}}}{{\ctxext{A}{P}}{\mfam{A}}}}
-  \\
-& \jdeq
-  \ctxwk
-    { {P}{\mfam{A}}
-      }
-    { \subst
-        {\cprojfstf{A}{P}}
-        {\ctxwk{\ctxext{A}{P}}{\mfam{A}}}
-      }
-  \\
-& \jdeq
-  \ctxwk
-    { {P}{\mfam{A}}
-      }
-    { {P}{\mfam{A}}
-      }
-  \\
-& \jdeq
-  \ctxwk{P}{{\mfam{A}}{\mfam{A}}}
-\end{align*}
-We find the term $\ctxwk{P}{\idtm{\mfam{A}}}$ here. Thus we can now define
-$\bar{\typefont{pr}}$ by:
-\begin{equation}\label{barproj}
-\jhomdefn
-  {\Gamma}
-  {{{A}{P}}{\mfam{A}}}
-  {{A}{\mfam{A}}}
-  {\bar{\typefont{pr}}}
-  {\tmext{}{}{\ctxwk{{P}{\mfam{A}}}{\cprojfstf{A}{P}}}{\ctxwk{P}{\idtm{\mfam{A}}}}}
-\end{equation}
-
-\begin{lem}
-We have the judgmental equality
-\begin{equation*}
-\jfameq
-  {{{{\Gamma}{A}}{P}}{\ctxwk{P}{\mfam{A}}}}
-  {\jcomp{}{\bar{\typefont{pr}}}{Q}}
-  {\ctxwk{P}{Q}}
-\end{equation*}
-for any family $Q$ of contexts over $\ctxext{{\Gamma}{A}}{\mfam{A}}$ 
-\end{lem}
-
-\subsubsection{Inductive morphisms}
-\begin{defn}
-Let $\jhom{\Gamma}{A}{B}{f}$ be a context morphism. We say that $f$ is an 
-\emph{inductive morphism} if the following inference rules are valid:
-\begin{align*}
-& \inference
-  { \jfam{{\Gamma}{B}}{Q}
-    \jterm{{\Gamma}{A}}{\jcomp{A}{f}{Q}}{g}
-    }
-  { \jterm{{\Gamma}{B}}{Q}{\jtcext{g}}
-    }
-& & \inference
-  { \jfam{{\Gamma}{B}}{Q}
-    \jterm{{\Gamma}{A}}{\jcomp{A}{f}{Q}}{g}
-    }
-  { \jtermeq{{\Gamma}{A}}{\jcomp{A}{f}{Q}}{\jcomp{A}{f}{\jtcext{g}}}{g}
-    }
-\end{align*}
-\end{defn}
-
-When $f$ is a context morphism from $A$ to $B$ in context $\Gamma$, 
-finding a term of a family $Q$ over $\ctxext{\Gamma}{B}$ can be accomplished
-by finding a term of the family $\jcomp{A}{f}{Q}$ over $\ctxext{\Gamma}{A}$.
-Inductive morphisms come with a computation rule. Thus, term
-constructors of inductively defined types are going to be the major source of
-examples of inductive morphisms. Since identity terms behave
-so nicely, they are inductive morphisms too.
-
-\begin{lem}
-The identity term
-$\jhom{\Gamma}{A}{A}{\idtm{A}}$ is an inductive morphism
-for each family $A$ of contexts over $\Gamma$
-\end{lem}
-
-\begin{itemize}
-\item Extensions of inductive morphisms are inductive
-\item Weakenings of inductive morphisms are inductive
-\item Substitutions of inductive morphisms are inductive
-\end{itemize}

### 2014-04-02 article-models-ttderived.tex
there it is :)

\section{Derived notions of the theory of contexts, families and terms}
\label{digging_deeper}

There is a good amount of things we can already state and prove about in the
theory we have developed so far. We will develop a bit more theory that will
come in handy later on.

In this section we use the framework we have developed in this
section to derive new notions and their properties. In particular, we will
develop the notion of \emph{extension on terms} together with the projection
maps from an extension to the `base' context of family, the 
\emph{family pullback} which is a version of pullbacks for families and thirdly
the \emph{inductive morphisms} which are morphisms of type theory that allow
to find terms of families over the codomain context by pulling them back to
the domain context and finding a term there.

In \autoref{extension-on-terms} on the extension operation on terms we will
derive all the compatibility rules that one would expect to hold for extension
on terms. The key to most of these results is the currying operation, which
could be seen as the missing feature in the table above. The extension on terms
operation depends in an essential way on the substitution operation, on the
identity terms and therefore indirectly also on the weakening operation. Thus,
we will see here all of the features of the theory we develop in
\autoref{judgments} through \autoref{categorical_properties} come to the 
surface.

Next, we introduce the inclusion of the fibers $\subst{x}{P}$ into the extension
$\ctxext{A}{P}$ as a morphism in context $\Gamma$. As was the case with
extension on terms and with projections, there will be a ton of compatibility
properties which we will prove about these inclusions.

It should be kept in mind though that in the current formulation there is no
sealed deal establishing a relationship between families over a context
and any kind of morphisms -- neither with morphisms having the `base' of the
family as its codomain nor with families into a universe (universes will be
introduced in \autoref{universes}). The only thing we know here is
that a family $P$ over $\ctxext{\Gamma}{A}$ determines a context morphism
from $\ctxext{A}{P}$ to $A$ in context $\Gamma$, the projection. 
We do not see this as
a shortcoming of the theory of contexts families and terms. Rather, such a
correspondence is a feature of a theory which does incorporate universes. The
fact that we're lacking a clear connection between families and (a specified
class of) morphisms, however, does show up in our treatment of the notion we
called familie pullbacks. For instance, we can't show that a square of families
is a pullback precisely when the corresponding square of projections is a
pullback: only the backwards direction holds. 
The discrepancies continue: ordinary pullbacks do not always exist
whereas family pullbacks do but the composition of two family pullback squares
need not be a family pullback square again whereas the pasting lemma of
ordinary pullbacks holds as usually.
We feel that pointing out what we can and can't do in the current setting is
an important aspect of developing an intuition with the system and therefore
we include this subsection even though the theory of family pullbacks
might feel a bit different than the usual theory of pullbacks.

In the last subsection we give a treatment of inductive morphisms. These
morphisms appear also in the introduction of the many inductive type
constructors in \autoref{tt_constructors} and therefore a general treatment of
the subject is insightful.

\emph{This subsection contains no new assumptions.}

\subsection{Projections and extension on terms}\label{extension-on-terms}
In this subsection we consider the notion of extension on terms, which has now
become definable inside our theory. Moreover, every compatibility rule one may
dream of is provable as well, using the compatibility rules we have introduced
earlier.

\begin{defn}
When $\jterm{\Gamma}{A}{x}$ and $\jterm{\Gamma}{\subst{x}{P}}{u}$ are terms,
we define 
\begin{equation*}
\jtermdefn
  {\Gamma}
  {\ctxext{A}{P}}
  {\tmext{A}{P}{x}{u}}
  {\unfold{\tmext{A}{P}{x}{u}}}.
\end{equation*} 
\end{defn}

Thus, the term $\tmext{A}{P}{x}{u}$ is the pairing of $x$ and $u$. Note that because
we have the judgmental equality 
$\ctxwk{P}{{A}{\ctxext{A}{P}}}\jdeq\ctxwk{\ctxext{A}{P}}{\ctxext{A}{P}}$ in the
context $\ctxext{{\Gamma}{A}}{P}$, the
pairing function could just be defined as $\idtm{\ctxext{A}{P}}$.

When we substitute by an extended term we get an equal result as when we
substitute two consecutive times, like the way currying works.

\begin{lem}
The following inference rules are valid:
\begin{align*}
& \inference
  { \jterm{\Gamma}{A}{x}
    \jterm{\Gamma}{\subst{x}{P}}{u}
    \jfam{{{\Gamma}{A}}{P}}{Q}
    }
  { \jfameq
      {\Gamma}
      {\subst{\tmext{A}{P}{x}{u}}{Q}}
      {\subst{u}{{x}{Q}}}
    }
  \\
& \inference
  { \jterm{\Gamma}{A}{x}
    \jterm{\Gamma}{\subst{x}{P}}{u}
    \jterm{{{\Gamma}{A}}{P}}{Q}{g}
    }
  { \jtermeq
      {\Gamma}
      {\subst{u}{{x}{Q}}}
      {\subst{\tmext{A}{P}{x}{u}}{g}}
      {\subst{u}{{x}{g}}}
    }
  \\
& \inference
  { \jterm{\Gamma}{A}{x}
    \jterm{\Gamma}{\subst{x}{P}}{u}
    \jfam{{{{\Gamma}{A}}{P}}{Q}}{R}
    }
  { \jfameq
      {{\Gamma}{\subst{u}{{x}{Q}}}}
      {\subst{\tmext{A}{P}{x}{u}}{R}}
      {\subst{u}{{x}{R}}}
    }
  \\
& \inference
  { \jterm{\Gamma}{A}{x}
    \jterm{\Gamma}{\subst{x}{P}}{u}
    \jterm{{{{\Gamma}{A}}{P}}{Q}}{R}{t}
    }
  { \jtermeq
      {{\Gamma}{\subst{u}{{x}{Q}}}}
      {\subst{u}{{x}{R}}}
      {\subst{\tmext{A}{P}{x}{u}}{t}}
      {\subst{u}{{x}{t}}}
    }
\end{align*}
\end{lem}

\begin{proof}
We prove only the first judgmental equality. All the others are similar.
Let $\jterm{\Gamma}{A}{x}$ and $\jterm{\Gamma}{\subst{x}{P}}{u}$
be terms and let $\jfam{{{\Gamma}{A}}{P}}{Q}$ be a family. Then we have
\begin{align*}
\subst
  {\tmext{A}{P}{x}{u}}
  {Q} 
& \jdeq 
  \subst
    {{u}{{x}{\idtm{\ctxext{A}{P}}}}}
    {Q}
  \tag{by definition}\\
& \jdeq 
  \subst
    {{u}{{x}{\idtm{\ctxext{A}{P}}}}}
    {{x}{\ctxwk{A}{Q}}}
  \tag{by \autoref{defn-ws-3}}\\
& \jdeq 
  \subst
    {{u}{{x}{\idtm{\ctxext{A}{P}}}}}
    {{u}{\ctxwk{\subst{x}{P}}{\subst{x}{\ctxwk{A}{Q}}}}}
  \tag{by \autoref{defn-ws-3}}\\
& \jdeq 
  \subst
    {{u}{{x}{\idtm{\ctxext{A}{P}}}}}
    {{u}{{x}{\ctxwk{P}{{A}{Q}}}}}
  \tag{by \autoref{comp-sw-f}}\\
& \jdeq 
  \subst
    {u}
    {{{x}{\idtm{\ctxext{A}{P}}}}{{x}{\ctxwk{{P}{{A}{Q}}}}}}
  \tag{by \autoref{comp-ss-f}}\\
& \jdeq 
  \subst
    {u}
    {{x}{{\idtm{\ctxext{A}{P}}}{\ctxwk{P}{{A}{Q}}}}}
  \tag{by \autoref{comp-ss-f}}\\
& \jdeq 
  \subst
    {u}
    {{x}{{\idtm{\ctxext{A}{P}}}{\ctxwk{\ctxext{A}{P}}{Q}}}}
  \tag{by \autoref{comp-ew-f}}\\
& \jdeq 
  \subst
    {u}
    {{x}{Q}}
  \tag{by \autoref{idfunc-wk-defn}}
\end{align*}
\end{proof}

We have seen above that the pairing function into $\ctxext{A}{P}$ is just the identity term on
$\ctxext{A}{P}$. To analyze the pairing functin a little further, we will also
need the projection maps from $\ctxext{A}{P}$ to $A$ and from $\ctxext{A}{P}$
to $P$. We will now define these and see that the identity term of an
extended family is the extension (or pairing) of the identity
functions on the components in the apropriate way.

To find out what the
apropriate way is, note that
\begin{align*}
\ctxwk{\ctxext{A}{P}}{\ctxext{A}{P}} 
& \jdeq 
  \ctxwk{P}{{A}{\ctxext{A}{P}}}
  \\
& \jdeq 
  \ctxext{\ctxwk{P}{{A}{A}}}{\ctxwk{P}{{A}{P}}}
\end{align*}
We have the term $\jterm{{\Gamma}{{A}{P}}}{\ctxwk{P}{A}}{\ctxwk{P}{\idtm{A}}}$.
Thus we need to find out what $\subst{\ctxwk{P}{\idtm{A}}}{\ctxwk{P}{{A}{P}}}$ is:
\begin{align*}
\subst{\ctxwk{P}{\idtm{A}}}{\ctxwk{P}{{A}{P}}} 
& \jdeq 
  \ctxwk{P}{\subst{\idtm{A}}{\ctxwk{A}{P}}}
  \\
& \jdeq 
  \ctxwk{P}{P},
\end{align*}
where we find the term $\idtm{P}$. Therefore we define:

\begin{defn}
Let $\jfam{\Gamma}{A}$ and $\jfam{{\Gamma}{A}}{P}$ be families. We define
\begin{align*}
\jhomdefn*
  {\Gamma}
  {{A}{P}}
  {A}
  {\cprojfstf{A}{P}}
  {\unfold{\cprojfstf{A}{P}}}
  \\
\jtermdefn*
  {\ctxext{\Gamma}{{A}{P}}}
  {\ctxwk{P}{P}}
  {\cprojsndf{A}{P}}
  {\unfold{\cprojsndf{A}{P}}}
\end{align*}
\end{defn}

We find the following inference rule, which expresses that the identity term
is compatible with extension:

\begin{lem}
For any $\jfam{\Gamma}{A}$ and $\jfam{{\Gamma}{A}}{P}$ we have
\begin{equation}\label{idfunc-ext-comp}
\inference
  { \jfam{\Gamma}{A}
    \jfam{{\Gamma}{A}}{P}
    }
  { \jhomeq
      {\Gamma}
      {{A}{P}}{{A}{P}}
      {\idtm{\ctxext{A}{P}}}
      { \tmext
          {\ctxwk{\ctxext{A}{P}}{A}}
          {\ctxwk{\ctxext{A}{P}}{P}}
          {\cprojfstf{A}{P}}
          {\cprojsndf{A}{P}}
        }
    }
\end{equation}
\end{lem}

\begin{proof}
Consider the families $\jfam{\Gamma}{A}$ and $\jfam{{\Gamma}{A}}{P}$. Then
we have the judgmental equalities
\begin{align*}
\tmext
  {\ctxwk{\ctxext{A}{P}}{A}}
  {\ctxwk{\ctxext{A}{P}}{P}}
  {\cprojfstf{A}{P}}
  {\cprojsndf{A}{P}}
& \jdeq 
  \unfold
  { \tmext
      {\ctxwk{\ctxext{A}{P}}{A}}
      {\ctxwk{\ctxext{A}{P}}{P}}
      {\cprojfstf{A}{P}}
      {\cprojsndf{A}{P}}
    }
  \\
& \jdeq 
  \subst
    { \idtm{P}
      }
    { {\ctxwk{P}{\idtm{A}}}
      {\idtm{\ctxwk{\ctxext{A}{P}}{\ctxext{A}{P}}}}
      }
  \\
& \jdeq 
  \subst
    { \idtm{P}
      }
    { {\ctxwk{P}{\idtm{A}}}
      {\ctxwk{\ctxext{A}{P}}{\idtm{\ctxext{A}{P}}}}
      }
  \\
& \jdeq 
  \subst
    { \idtm{P}
      }
    { {\ctxwk{P}{\idtm{A}}}
      {\ctxwk{P}{{A}{\idtm{\ctxext{A}{P}}}}}
      }
  \\
& \jdeq 
  \subst
    { \idtm{P}
      }
    {\ctxwk
      {P}
      { \subst
        {\idtm{A}}
        {\ctxwk{A}{\idtm{\ctxext{A}{P}}}}
        }
      }
  \\
& \jdeq 
  \subst
    { \idtm{A}
      }
    { \ctxwk{A}{\idtm{\ctxext{A}{P}}}
      }
  \\
& \jdeq 
  \idtm{\ctxext{A}{P}}.
  \qedhere
\end{align*}
\end{proof}

\begin{lem}
The following inference is valid:
\begin{equation*}
\inference
  { \jfam{{\Gamma}{A}}{P}
    \jfam{{\Gamma}{A}}{Q}
    }
  { \jfameq
      {{\Gamma}{{A}{P}}}
      {\subst{\cprojfstf{A}{P}}{\ctxwk{\ctxext{A}{P}}{Q}}}
      {\ctxwk{P}{Q}}
    }
\end{equation*}
\end{lem}

\begin{proof}
Let $\jfam{{\Gamma}{A}}{P}$ and $\jfam{{\Gamma}{A}}{Q}$ be
families. Then we have
\begin{align*}
\subst{\cprojfstf{A}{P}}{\ctxwk{\ctxext{A}{P}}{Q}} 
& \jdeq 
  \subst{\ctxwk{P}{\idtm{A}}}{\ctxwk{\ctxext{A}{P}}{Q}} 
  \tag{by definition}\\
& \jdeq 
  \subst{\ctxwk{P}{\idtm{A}}}{\ctxwk{P}{{A}{Q}}} 
  \tag{by \autoref{comp-ww-f}}\\
& \jdeq 
  \ctxwk{P}{\subst{\idtm{A}}{\ctxwk{A}{Q}}} 
  \tag{by \autoref{comp-ws-f}}\\
& \jdeq 
  \ctxwk{P}{Q} 
  \tag{by \autoref{idfunc-wk-defn}}
\end{align*}
\end{proof}

The constructions of the terms $\tmext{A}{P}{x}{u}$ and $\cprojfst{A}{P}{w}$ and
$\cprojsnd{A}{P}{w}$ are subject to various rules, with all of them being
consequences of earlier introduced inference rules.

\begin{lem} The following inference rules expressing that pairing is a strict
inverse to the combination of decompositions, are valid:
\begin{align*}
& \inference
  { \jterm{\Gamma}{\ctxext{A}{P}}{w}
    }
  { \jtermeq
      {\Gamma}
      {\ctxext{A}{P}}
      {\tmext{A}{P}{\cprojfst{A}{P}{w}}{\cprojsnd{A}{P}{w}}}
      {w}
    }
  \\
& \inference
  { \jterm{\Gamma}{A}{x}
    \jterm{\Gamma}{\subst{x}{P}}{u}
    }
  { \jtermeq
      {\Gamma}
      {A}
      {\cprojfst{A}{P}{\tmext{A}{P}{x}{u}}}
      {x}
    }
  \\
& \inference
  { \jterm{\Gamma}{A}{x}
    \jterm{\Gamma}{\subst{x}{P}}{u}
    }
  { \jtermeq
      {\Gamma}
      {\subst{x}{P}}
      {\cprojsnd{A}{P}{\tmext{A}{P}{x}{u}}}
      {u}
    }
\end{align*}
\end{lem}

\begin{proof}
To prove the first judgmental equality, note that
\begin{align*}
w 
& \jdeq 
  \subst{w}{\idtm{\ctxext{A}{P}}} 
  \tag{by \autoref{idfunc-subst-defn}}\\
& \jdeq 
  \subst
    { w}
    { { \idtm{P}}
      { { \ctxwk{P}{\idtm{A}}}
        { \idtm{\ctxwk{\ctxext{A}{P}}{\ctxext{A}{P}}}}
        }
      }
  \tag{by \autoref{idfunc-ext-comp}}\\
& \jdeq 
  \subst
    { {w}
      {\idtm{P}}
      }
    { {w}
      { { \ctxwk{P}{\idtm{A}}
          }
        { \idtm{\ctxwk{\ctxext{A}{P}}{\ctxext{A}{P}}}
          }
        }
      }
  \tag{by \autoref{comp-ss-t}}\\
& \jdeq 
  \subst
    { {w}
      {\idtm{P}}
      }
    { { {w}
        {\ctxwk{P}{\idtm{A}}}
        }
      { {w}
        {\idtm{\ctxwk{\ctxext{A}{P}}{\ctxext{A}{P}}}}
        }
      }
  \tag{by \autoref{comp-ss-t}}\\
& \jdeq 
  \subst
    { {w}
      {\idtm{P}}
      }
    { { {w}
        {\ctxwk{P}{\idtm{A}}}
        }
      { {w}
        {\ctxwk{\ctxext{A}{P}}{\idtm{\ctxext{A}{P}}}}
        }
      }
  \tag{by \autoref{idfunc-wk-comp}}\\
& \jdeq 
  \subst
    { {w}
      {\idtm{P}}
      }
    { { {w}
        {\ctxwk{P}{\idtm{A}}}
        }
      { \idtm{\ctxext{A}{P}}
        }
      }
  \tag{by \autoref{defn-ws-4}}\\
& \jdeq 
  \tmext{A}{P}{\cprojfst{A}{P}{w}}{\cprojsnd{A}{P}{w}}
  \tag{by definition}
\end{align*}
To prove the second judgmental equality, let $\jterm{\Gamma}{A}{x}$ and
$\jterm{\Gamma}{\subst{x}{P}}{u}$. Then we have
\begin{align*}
\cprojfst{A}{P}{\tmext{A}{P}{x}{u}}
& \jdeq 
  \subst{\tmext{A}{P}{x}{u}}{\ctxwk{P}{\idtm{A}}}
  \\
& \jdeq 
  \subst{u}{{x}{\ctxwk{P}{\idtm{A}}}} 
  \\
& \jdeq 
  \subst{u}{\ctxwk{\subst{x}{P}}{\subst{x}{\idtm{A}}}}
  \\
& \jdeq 
  \subst{x}{\idtm{A}}
  \\
& \jdeq 
  x.
\end{align*}
To prove the third judgmental equality, note that
\begin{align*}
\cprojsnd{A}{P}{\tmext{A}{P}{x}{u}}
& \jdeq 
  \subst{\tmext{A}{P}{x}{u}}{\idtm{P}}
  \\
& \jdeq 
  \subst{u}{{x}{\idtm{P}}}
  \\
& \jdeq 
  \subst{u}{\idtm{\subst{x}{P}}}
  \\
& \jdeq 
  u.
  \qedhere
\end{align*}
\end{proof}

\begin{lem}
The following compatibility rules for two consecutive term extensions is valid:
\begin{align*}
& \inference
  { \jterm{\Gamma}{A}{x}
    \jterm{\Gamma}{\subst{x}{P}}{u}
    \jterm{\Gamma}{\subst{\tmext{A}{P}{x}{u}}{Q}}{v}
    }
  { \jtermeq
      {\Gamma}
      {\ctxext{{A}{P}}{Q}}
      {\tmext{A}{{P}{Q}}{x}{{\subst{x}{P}}{\subst{x}{Q}}{u}{v}}}
      {\tmext{{A}{P}}{Q}{{A}{P}{x}{u}}{v}}
    }
  \\
& \inference
  { \jfam{{{\Gamma}{A}}{P}}{Q}
    }
  { \jhomeq
      {\Gamma}
      {\ctxext{{A}{P}}{Q}}
      {A}
      {\jcomp{}{\cprojfstf{{A}{P}}{Q}}{\cprojfstf{A}{P}}}
      {\cprojfstf{A}{{P}{Q}}}
    }
  \\
& \inference
  { \jfam{{{\Gamma}{A}}{P}}{Q}
    }
  { \jhomeq
      {{\Gamma}{A}}
      {{P}{Q}}
      {P}
      {\jcomp{}{\cprojfstf{{A}{P}}{Q}}{\cprojsndf{A}{P}}}
      {\jcomp{}{\cprojsndf{A}{{P}{Q}}}{\cprojfstf{P}{Q}}}
    }
  \\
& \inference
    { \jfam{{{\Gamma}{A}}{P}}{Q}
      }
    { \jhomeq
        {{{\Gamma}{A}}{P}}
        {Q}
        {Q}
        {\cprojsndf{{A}{P}}{Q}}
        {\jcomp{}{\cprojsndf{A}{{P}{Q}}}{\cprojsndf{P}{Q}}}
      }
\end{align*}
\end{lem}

\begin{proof}
Consider terms $\jterm{\Gamma}{A}{x}$, $\jterm{\Gamma}{\subst{x}{P}}{u}$ and
$\jterm{\Gamma}{\subst{u}{{x}{Q}}}{v}$. Then we have
\begin{align*}
\tmext{x}{{u}{v}}
& \jdeq 
  \subst
    {\tmext{u}{v}}{{x}{\idtm{\ctxext{A}{{P}{Q}}}}}
  \\
& \jdeq 
  \subst{v}{{u}{{x}{\idtm{\ctxext{A}{{P}{Q}}}}}}
  \\
& \jdeq 
  \subst{v}{{u}{{x}{\idtm{\ctxext{{A}{P}}{Q}}}}}
  \\
& \jdeq 
  \subst{v}{{\tmext{x}{u}}{\idtm{\ctxext{{A}{P}}{Q}}}}
  \\
& \jdeq 
  \tmext{{x}{u}}{v}.
\end{align*}
To prove the judmental equality
\begin{equation*}
\jhomeq
  {\Gamma}
  {\ctxext{{A}{P}}{Q}}
  {A}
  {\jcomp{}{\cprojfstf{{A}{P}}{Q}}{\cprojfstf{A}{P}}}
  {\cprojfstf{A}{{P}{Q}}}
\end{equation*}
note that we have the judgmental equalities
\begin{align*}
\jcomp{{{A}{P}}{Q}}{\cprojfstf{{A}{P}}{Q}}{\cprojfstf{A}{P}}
& \jdeq 
  \unfoldall{\jcomp{{{A}{P}}{Q}}{\cprojfstf{{A}{P}}{Q}}{\cprojfstf{A}{P}}}
  \\
& \jdeq 
  \subst
    {\ctxwk{Q}{\idtm{{A}{P}}}}
    {\ctxwk{Q}{{\ctxext{A}{P}}{{P}{\idtm{A}}}}}
  \\
& \jdeq
  \ctxwk{Q}{\subst{\idtm{{A}{P}}}{\ctxwk{\ctxext{A}{P}}{{P}{\idtm{A}}}}}
  \\
& \jdeq
  \ctxwk{Q}{{P}{\idtm{A}}}
  \\
& \jdeq
  \ctxwk{\ctxext{P}{Q}}{\idtm{A}}
  \\
& \jdeq
  \cprojfstf{A}{{P}{Q}}
\end{align*}
To prove the judgmental equality
\begin{equation*}
\jhomeq
  {{\Gamma}{A}}
  {{P}{Q}}
  {P}
  {\jcomp{}{\cprojfstf{{A}{P}}{Q}}{\cprojsndf{A}{P}}}
  {\jcomp{}{\cprojsndf{A}{{P}{Q}}}{\cprojfstf{P}{Q}}}
\end{equation*}
note that we have the judgmental equalities
\begin{align*}
\jcomp{{{A}{P}}{Q}}{\cprojfstf{{A}{P}}{Q}}{\cprojsndf{A}{P}}
& \jdeq 
  \unfoldall{\jcomp{{{A}{P}}{Q}}{\cprojfstf{{A}{P}}{Q}}{\cprojsndf{A}{P}}}
  \\
& \jdeq
  \subst{\ctxwk{Q}{\idtm{{A}{P}}}}{\ctxwk{Q}{{\ctxext{A}{P}}{\idtm{P}}}}
  \\
& \jdeq
  \ctxwk{Q}{\subst{\idtm{{A}{P}}}{\ctxwk{\ctxext{A}{P}}{\idtm{P}}}}
  \\
& \jdeq
  \ctxwk{Q}{\idtm{P}}
  \\
& \jdeq
  \unfoldall{\jcomp{{P}{Q}}{\cprojsndf{A}{{P}{Q}}}{\cprojfstf{P}{Q}}}
  \\
& \jdeq
  \jcomp{{P}{Q}}{\cprojsndf{A}{{P}{Q}}}{\cprojfstf{P}{Q}}
\end{align*}
To prove the judgmental equality
\begin{equation*}
\jhomeq
  {{{\Gamma}{A}}{P}}
  {Q}
  {Q}
  {\cprojsndf{{A}{P}}{Q}}
  {\jcomp{}{\cprojsndf{A}{{P}{Q}}}{\cprojsndf{P}{Q}}}
\end{equation*}
note that we have the judgmental equalities
\begin{align*}
\cprojsndf{{A}{P}}{Q}
& \jdeq
  \unfoldall{\cprojsndf{{A}{P}}{Q}}
  \\
& \jdeq 
  \unfoldall{\jcomp{{P}{Q}}{\cprojsndf{A}{{P}{Q}}}{\cprojsndf{P}{Q}}}
  \\
& \jdeq
  \jcomp{}{\cprojsndf{A}{{P}{Q}}}{\cprojsndf{P}{Q}}\qedhere
\end{align*}
\begin{comment}
%%%% This was a proof of an old version of the statement
Now consider a term $\jterm{\Gamma}{\ctxext{A}{{P}{Q}}}{w}$. Then we have
\begin{align*}
w 
& \jdeq 
  \tmext
    {A}
    {{P}{Q}}
    {\cprojfst{A}{\ctxext{P}{Q}}{w}}
    {\cprojsnd{A}{\ctxext{P}{Q}}{w}}
  \\
& \jdeq 
  \tmext
    {A}
    {{P}{Q}}
    {\cprojfst{A}{\ctxext{P}{Q}}{w}}
    { {P}
      {Q}
      {\cprojfst{P}{Q}{\cprojsnd{A}{\ctxext{P}{Q}}{w}}}
      {\cprojsnd{P}{Q}{\cprojsnd{A}{\ctxext{P}{Q}}{w}}}
      }
  \\
& \jdeq 
  \tmext
    {{A}{P}}
    {Q}
    { {} % need to provide base and family, but there's no unfold.
      {}
      {\cprojfst{A}{\ctxext{P}{Q}}{w}}
      {\cprojfst{P}{Q}{\cprojsnd{A}{\ctxext{P}{Q}}{w}}}
      }
    { \cprojsnd{P}{Q}{\cprojsnd{A}{\ctxext{P}{Q}}{w}}
      }
\end{align*}
Thus we see that 
\begin{align*}
\cprojfst{\ctxext{A}{P}}{Q}{w} 
& \jdeq 
  \tmext
    {A}
    {P}
    {\cprojfst{A}{\ctxext{P}{Q}}{w}}
    {\cprojfst{P}{Q}{\cprojsnd{A}{\ctxext{P}{Q}}{w}}}
  \\ 
\cprojsnd{\ctxext{A}{P}}{Q}{w} 
& \jdeq 
  \cprojsnd{P}{Q}{\cprojsnd{A}{\ctxext{P}{Q}}{w}},
\end{align*}
proving the fourth judgmental equality, and therefore also that
\begin{align*}
\cprojfst{A}{P}{\cprojfst{\ctxext{A}{P}}{Q}{w}} 
& \jdeq 
  \cprojfst{A}{\ctxext{P}{Q}}{w}
  \\
\cprojsnd{A}{P}{\cprojfst{\ctxext{A}{P}}{Q}{w}} 
& \jdeq 
  \cprojfst{P}{Q}{\cprojsnd{A}{\ctxext{P}{Q}}{w}},
\end{align*}
proving the second and the third judgmental equalities.
\end{comment}
\end{proof}

\begin{lem}
When we weaken a term $\tmext{B}{Q}{y}{v}$ of $\ctxext{B}{Q}$ in context $\Gamma$ by
a family $A$, the term that we get is $\tmext{\ctxwk{A}{B}}{\ctxwk{A}{Q}}{\ctxwk{A}{y}}{\ctxwk{A}{v}}$. More
precisely, the following inference rules are valid:
\begin{align*}
& \inference
  { \jterm{{\Gamma}{B}}{Q}{g}
    \jterm{{\Gamma}{B}}{\subst{g}{R}}{t}
    }
  { \jtermeq
      {{{\Gamma}{A}}{\ctxwk{A}{B}}}
      {\ctxwk{A}{\ctxext{Q}{R}}}
      {\ctxwk{A}{\tmext{Q}{R}{g}{t}}}
      {\tmext{\ctxwk{A}{Q}}{\ctxwk{A}{R}}{\ctxwk{A}{g}}{\ctxwk{A}{t}}}
    }
  \\
& \inference
  { \jfam{{{\Gamma}{B}}{Q}}{R}
    }
  { \jhomeq
      {{{\Gamma}{A}}{\ctxwk{A}{B}}}
      {{\ctxwk{A}{Q}}{\ctxwk{A}{R}}}
      {\ctxwk{A}{Q}}
      {\cprojfstf{\ctxwk{A}{Q}}{\ctxwk{A}{R}}}
      {\ctxwk{A}{\cprojfstf{Q}{R}}}
    }
  \\
& \inference
  { \jfam{{{\Gamma}{B}}{Q}}{R}
    }
  { \jhomeq
      {{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{Q}}}
      {\ctxwk{A}{R}}
      {\ctxwk{A}{R}}
      {\cprojsndf{\ctxwk{A}{Q}}{\ctxwk{A}{R}}}
      {\ctxwk{A}{\cprojsndf{Q}{R}}}
    }
\end{align*}
\end{lem}

\begin{proof}
Consider $\jterm{{\Gamma}{B}}{Q}{g}$ and $\jterm{{\Gamma}{B}}{\subst{g}{R}}{t}$.
Then we have the judgmental equalities
\begin{align*}
\ctxwk{A}{\ctxext{Q}{R}{g}{t}}
& \jdeq 
  \ctxwk{A}{\subst{t}{{g}{\idtm{\ctxext{Q}{R}}}}}
  \\
& \jdeq 
  \subst{\ctxwk{A}{t}}{\ctxwk{A}{\subst{g}{\idtm{\ctxext{Q}{R}}}}}
  \\
& \jdeq 
  \subst{\ctxwk{A}{t}}{{\ctxwk{A}{g}}{\ctxwk{A}{\idtm{\ctxext{Q}{R}}}}}
  \\
& \jdeq 
  \subst{\ctxwk{A}{t}}{{\ctxwk{A}{g}}{\idtm{\ctxwk{A}{\ctxext{Q}{R}}}}}
  \\
& \jdeq 
  \subst{\ctxwk{A}{t}}{{\ctxwk{A}{g}}{\idtm{\ctxext{\ctxwk{A}{Q}}{\ctxwk{A}{R}}}}}
  \\
& \jdeq 
  \tmext{\ctxwk{A}{Q}}{\ctxwk{A}{R}}{\ctxwk{A}{g}}{\ctxwk{A}{t}}
\end{align*}
Next, we want to prove the judgmental equality
\begin{equation*}
\jhomeq
  {{{\Gamma}{A}}{\ctxwk{A}{B}}}
  {{\ctxwk{A}{Q}}{\ctxwk{A}{R}}}
  {\ctxwk{A}{Q}}
  {\cprojfstf{\ctxwk{A}{Q}}{\ctxwk{A}{R}}}
  {\ctxwk{A}{\cprojfstf{Q}{R}}}
\end{equation*}
Note that we have the judgmental equalities
\begin{align*}
\cprojfstf{\ctxwk{A}{Q}}{\ctxwk{A}{R}}
& \jdeq
  \unfoldall{\cprojfstf{\ctxwk{A}{Q}}{\ctxwk{A}{R}}}
  \\
& \jdeq
  \ctxwk{{A}{R}}{{A}{\idtm{Q}}}
  \\
& \jdeq
  \unfoldall{\ctxwk{A}{\cprojfstf{Q}{R}}}
  \\
& \jdeq
  \ctxwk{A}{\cprojfstf{Q}{R}}.
\end{align*}
Finally, we want to prove the judgmental equality
\begin{equation*}
\jhomeq
  {{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{Q}}}
  {\ctxwk{A}{R}}
  {\ctxwk{A}{R}}
  {\cprojsndf{\ctxwk{A}{Q}}{\ctxwk{A}{R}}}
  {\ctxwk{A}{\cprojsndf{Q}{R}}}
\end{equation*}
Note that we have the judgmental equalities
\begin{align*}
\cprojsndf{\ctxwk{A}{Q}}{\ctxwk{A}{R}}
& \jdeq
  \unfoldall{\cprojsndf{\ctxwk{A}{Q}}{\ctxwk{A}{R}}}
  \\
& \jdeq
  \unfoldall{\ctxwk{A}{\cprojsndf{Q}{R}}}
  \\
& \jdeq
  \ctxwk{A}{\cprojsndf{Q}{R}}.
  \qedhere
\end{align*}
\end{proof}

\begin{lem}
When we substitute an extended term $\tmext{P}{Q}{f}{g}$ of $\ctxext{P}{Q}$ by a term
$x$ of $A$, the term that we get is $\tmext{\subst{x}{P}}{\subst{x}{Q}}{\subst{x}{f}}{\subst{x}{g}}$.
More precisely, the following inference rules are valid:
\begin{align*}
& \inference
  { \jterm{\Gamma}{A}{x}
    \jterm{{{\Gamma}{A}}{P}}{Q}{g}
    \jterm{{{\Gamma}{A}}{P}}{\subst{g}{R}}{t}
    }
  { \jtermeq
      {{\Gamma}{\subst{x}{P}}}
      {\ctxext{\subst{x}{Q}}{\subst{x}{R}}}
      {\subst{x}{\tmext{Q}{R}{g}{t}}}
      {\tmext{\subst{x}{Q}}{\subst{x}{R}}{\subst{x}{g}}{\subst{x}{t}}}
    }
  \\
& \inference
  { \jterm{\Gamma}{A}{x}
    \jfam{{{{\Gamma}{A}}{P}}{Q}}{R}
    }
  { \jhomeq
      {{\Gamma}{\subst{x}{P}}}
      {{\subst{x}{Q}}{\subst{x}{R}}}
      {\subst{x}{Q}}
      {\cprojfstf{\subst{x}{Q}}{\subst{x}{R}}}
      {\subst{x}{\cprojfstf{Q}{R}}}
    }
  \\
& \inference
  { \jterm{\Gamma}{A}{x}
    \jfam{{{{\Gamma}{A}}{P}}{Q}}{R}
    }
  { \jhomeq
      {{{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}}
      {\subst{x}{R}}
      {\subst{x}{R}}
      {\cprojsndf{\subst{x}{Q}}{\subst{x}{R}}}
      {\subst{x}{\cprojsndf{Q}{R}}}
    }
\end{align*}
\end{lem}

\begin{proof}
Consider $\jterm{{\Gamma}{B}}{Q}{g}$ and $\jterm{{\Gamma}{B}}{\subst{g}{R}}{t}$.
Then we have the judgmental equalities
\begin{align*}
\subst{x}{\tmext{Q}{R}{g}{t}}
& \jdeq 
  \subst{x}{{t}{{g}{\idtm{\ctxext{Q}{R}}}}}
  \\
& \jdeq 
  \subst{{x}{t}}{{x}{{g}{\idtm{\ctxext{Q}{R}}}}}
  \\
& \jdeq 
  \subst{{x}{t}}{{{x}{g}}{{x}{\idtm{\ctxext{Q}{R}}}}}
  \\
& \jdeq 
  \subst{{x}{t}}{{{x}{g}}{\idtm{\subst{x}{\ctxext{Q}{R}}}}}
  \\
& \jdeq 
  \subst{{x}{t}}{{{x}{g}}{\idtm{\ctxext{\subst{x}{Q}}{\subst{x}{R}}}}}
  \\
& \jdeq 
  \tmext{\subst{x}{Q}}{\subst{x}{R}}{\subst{x}{g}}{\subst{x}{t}}.
\end{align*}
Next, we want to prove the judgmental equality
\begin{equation*}
\jhomeq
  {{\Gamma}{\subst{x}{P}}}
  {{\subst{x}{Q}}{\subst{x}{R}}}
  {\subst{x}{Q}}
  {\cprojfstf{\subst{x}{Q}}{\subst{x}{R}}}
  {\subst{x}{\cprojfstf{Q}{R}}}
\end{equation*}
Note that we have the judgmental equalities
\begin{align*}
\cprojfstf{\subst{x}{Q}}{\subst{x}{R}}
& \jdeq
  \unfoldall{\cprojfstf{\subst{x}{Q}}{\subst{x}{R}}}
  \\
& \jdeq
  \ctxwk{\subst{x}{R}}{\subst{x}{\idtm{Q}}}
  \\
& \jdeq
  \unfoldall{\subst{x}{\cprojfstf{Q}{R}}}
  \\
& \jdeq
  \subst{x}{\cprojfstf{Q}{R}}.
\end{align*}
And finally we want to prove the judgmental equality
\begin{equation*}
\jhomeq
  {{{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}}
  {\subst{x}{R}}
  {\subst{x}{R}}
  {\cprojsndf{\subst{x}{Q}}{\subst{x}{R}}}
  {\subst{x}{\cprojsndf{Q}{R}}}
\end{equation*}
Note that we have the judgmental equalities
\begin{align*}
\cprojsndf{\subst{x}{Q}}{\subst{x}{R}}
& \jdeq
  \unfoldall{\cprojsndf{\subst{x}{Q}}{\subst{x}{R}}}
  \\
& \jdeq
  \unfoldall{\subst{x}{\cprojsndf{Q}{R}}}
  \\
& \jdeq
  \subst{x}{\cprojsndf{Q}{R}}.
  \qedhere
\end{align*}
\end{proof}

\subsection{More on morphisms}

For the following lemma, recall that the judgment $\jhom{\Gamma}{A}{{B}{Q}}{f}$
unfolds as
\begin{equation*}
\unfold{\jhom{\Gamma}{A}{{B}{Q}}{f}}
\end{equation*}
and that we have the judgmental equality 
$ \jfameq
    {{\Gamma}{A}}
    {\ctxwk{A}{\ctxext{B}{Q}}}
    {\ctxext{\ctxwk{A}{B}}{\ctxwk{A}{Q}}}.
  $
Therefore, each morphism into an extended family can itself be described as
an extended term. The following lemma explains how this goes.

\begin{lem}
Let $\jhom{\Gamma}{A}{{B}{Q}}{f}$ be a morphism from $A$ to $\ctxext{B}{Q}$
in a context $\Gamma$. Then we have
\begin{equation*}
\jhomeq
  {\Gamma}
  {A}
  {{B}{Q}}
  {f}
  {\tmext{\jcomp{}{f}{\cprojfstf{B}{Q}}}{\jcomp{}{f}{\cprojsndf{B}{Q}}}}.
\end{equation*}
Alternatively, when $\jhom{\Gamma}{A}{B}{f_0}$ and 
$\jterm{{\Gamma}{A}}{\jcomp{}{f_0}{Q}}{f_1}$ we obtain a morphism
$\jhom{\Gamma}{A}{{B}{Q}}{\tmext{f_0}{f_1}}$ with the property that
\begin{align*}
\jhomeq*{\Gamma}{A}{B}{\jcomp{}{\tmext{f_0}{f_1}}{\cprojfstf{B}{Q}}}{f_0}\\
\jtermeq*{{\Gamma}{A}}{\jcomp{}{f_0}{Q}}{\jcomp{}{\tmext{f_0}{f_1}}{\cprojsndf{B}{Q}}}{f_1}.
\end{align*}
\end{lem}

\begin{proof}
Let $\jhom{\Gamma}{A}{{B}{Q}}{f}$ be a morphism in context $\Gamma$. Then we
have the judgmental equalities
\begin{align*}
\cprojfst{\ctxwk{A}{B}}{\ctxwk{A}{Q}}{f}
& \jdeq
  \subst{f}{\ctxwk{A}{\cprojfstf{B}{Q}}}
  \\
& \jdeq
  \jcomp{}{f}{\cprojfstf{B}{Q}}
  \\
\cprojsnd{\ctxwk{A}{B}}{\ctxwk{A}{Q}}{f}
& \jdeq
  \subst{f}{\ctxwk{A}{\cprojsndf{B}{Q}}}
  \\
& \jdeq
  \jcomp{}{f}{\cprojsndf{B}{Q}}
\end{align*}
The alternative formulation of the statement is a direct corollary.
\end{proof}

We also have the following lemma about the compatibility of pairing and composition:

\begin{lem}
The following inference rule is valid
\begin{align*}
& \inference
  { \jhom{\Gamma}{A}{B}{f}
    \jhom{\Gamma}{B}{C}{g}
    \jfam{{\Gamma}{C}}{R}
    \jterm{{\Gamma}{B}}{\subst{g}{\ctxwk{B}{R}}}{w}
    }
  { \jhomeq
      {\Gamma}
      {A}
      {{C}{R}}
      {\jcomp{A}{f}{\tmext{\ctxwk{B}{C}}{\ctxwk{B}{R}}{g}{w}}}
      {\tmext{\ctxwk{A}{C}}{\ctxwk{A}{R}}{\jcomp{A}{f}{g}}{\jcomp{A}{f}{w}}}
    }
\end{align*}
\end{lem}

\begin{proof}
Let $\jhom{\Gamma}{A}{B}{f}$, $\jhom{\Gamma}{B}{C}{g}$, $\jfam{{\Gamma}{C}}{R}$
and $\jterm{{\Gamma}{B}}{\subst{g}{\ctxwk{B}{R}}}{w}$. Then we have the
judgmental equalities
\begin{align*}
\jcomp{A}{f}{\tmext{\ctxwk{B}{C}}{\ctxwk{B}{R}}{g}{w}}
& \jdeq 
  \subst{f}{\ctxwk{A}{\tmext{\ctxwk{B}{C}}{\ctxwk{B}{R}}{g}{w}}}
  \\
& \jdeq 
  \subst
    {f}
    {\tmext{\ctxwk{A}{{B}{C}}}{\ctxwk{A}{{B}{R}}}{\ctxwk{A}{g}}{\ctxwk{A}{w}}}
  \\
& \jdeq 
  \tmext
    {\ctxwk{A}{C}}
    {\ctxwk{A}{R}}
    {\subst{f}{\ctxwk{A}{g}}}
    {\subst{f}{\ctxwk{A}{w}}}
  \\
& \jdeq 
  \tmext{\ctxwk{A}{C}}{\ctxwk{A}{R}}{\jcomp{A}{f}{g}}{\jcomp{A}{f}{w}}.
  \qedhere
\end{align*}
\end{proof}

\begin{defn}
Let $\jhom{\Gamma}{A}{B}{f}$ be a morphism from $A$ to $B$ in context $\Gamma$
and let $\jfhom{\Gamma}{A}{B}{f}{P}{Q}{F}$ be a morphism over $f$ in context 
$\Gamma$. We define
\begin{equation*}
\jhomdefn{\Gamma}{{A}{P}}{{B}{Q}}{\jvcomp{P}{f}{F}}{\unfold{\jvcomp{P}{f}{F}}}
\end{equation*}
\end{defn}

\subsection{Fiber inclusions}
We will use the insights of \autoref{extension-on-terms} to define and study
\emph{fiber inclusions}. The fiber inclusion of the \emph{fiber}
$\subst{x}{P}$ into the extension $\ctxext{A}{P}$ is a morphism
$\jhom{\Gamma}{\subst{x}{P}}{{A}{P}}{\finc{x}{P}}$, for any family
$\jfam{{\Gamma}{A}}{P}$ and any term $\jterm{\Gamma}{A}{x}$. Then we will determine
the ways in which it is compatible with the other operators. Note that in this
subsection we will focus on the compatibility properties; the fact that
the fiber inclusions also appear in a pullback diagram will be established in
\autoref{pullback}.

\begin{defn}
Let $\jterm{\Gamma}{A}{x}$ be a term and let $\jfam{{\Gamma}{A}}{P}$ be a
family. Then we define
\begin{equation*}
\jhomdefn{\Gamma}{\subst{x}{P}}{{A}{P}}{\finc{x}{P}}{\unfoldnext{\finc{x}{P}}}
\end{equation*}
\end{defn}

\subsection{Pullback squares and family pullback squares}
\label{pullback}
Now that we have introduced the notions of morphisms, composition and identity
terms, we can develop a diagramatic style of of displaying type dependencies
combined with morphisms. We give an informal, metatheoretical definition of
such diagrams by indicating what the various components mean. The definition
is informal because we will only use such diagrams occasionally to provide a
graphical indication of the situation in which we're working. In particular,
we will not shy away from using natural numbers and trust that the reader can
figure out what we mean.

\begin{defn}
A diagram is said to be a \emph{dependency diagram in context $\Gamma$}
if it is built up according to the following steps:
\begin{itemize}
\item The arrows appearing in a dependency diagram are either ordinary, like the
arrow%
$\begin{tikzcd}[ampersand replacement = \&]
X \ar{r} \& Y,
\end{tikzcd}$
or double-headed, like
$\begin{tikzcd}[ampersand replacement = \&]
X \ar[fib]{r} \& Y.
\end{tikzcd}$
\item An ordinary arrow 
\begin{equation*}
\begin{tikzcd}
A \ar{r}{f} & B
\end{tikzcd}
\end{equation*}
between two families $A$ and $B$ of contexts over $\Gamma$ indicates that
$f$ is a morphism from $A$ to $B$ in context $\Gamma$, i.e.~that we have the
judgment $\jhom{\Gamma}{A}{B}{f}$.
\item The set of double-headed arrows must form a forest and the root of
each maximal tree of double-headed arrows is a family of contexts over $\Gamma$.
In particular, if an object is not the domain of a double-headed arrow it must
be a family of contexts over $\Gamma$.
\item A sequence of double-headed 
arrows
\begin{equation*}
\begin{tikzcd}
P_{n} \ar[fib]{r} & \cdots \ar[fib]{r} & P_1 \ar[fib]{r} & A
\end{tikzcd}
\end{equation*}
indicates that $P_1$ is a family of contexts over $\ctxext{\Gamma}{A}$, that
$P_2$ is a family of contexts over $\ctxext{{\Gamma}{A}}{P_1}$, etcetera.
\item There can be two kinds of ladders of double-headed arrows:
\begin{equation*}
\begin{tikzcd}
P_{n} \ar{r}{F_{n}} \ar[fib]{d} & Q_{n} \ar[fib]{d}\\
\vdots \ar[fib]{d} & \vdots \ar[fib]{d}\\
P_1 \ar{r}{F_1} \ar[fib]{d} & Q_1 \ar[fib]{d}\\
A \ar{r}{f} & B
\end{tikzcd}
\qquad
\begin{tikzcd}[column sep = tiny]
P_{n+m} \ar{rr}{F_{n+m}} \ar[fib]{d} & & Q_{n+m} \ar[fib]{d}\\
\vdots \ar[fib]{d} & & \vdots \ar[fib]{d}\\
P_{n+1} \ar{rr}{F_{n+1}} \ar[fib]{dr} & & Q_{n+1} \ar[fib]{dl}\\
& P_n \ar[fib]{d}\\
& \vdots \ar[fib]{d}\\
& P_1 \ar[fib]{d}\\
& A
\end{tikzcd}
\end{equation*}
The ladder on the left 
indicates that $F_1$ is a morphism from $P_1$ to $Q_1$ \emph{over} $f$,
i.e.~that the judgment $\jfhom{\Gamma}{A}{B}{f}{P_1}{Q_1}{F_1}$ holds, that
$F_2$ is a morphism from $P_2$ to $Q_2$ over
the morphism $\tmext{\ctxwk{P_1}{f}}{F_1}$ from $\ctxext{A}{P_1}$ to
$\ctxext{B}{Q_1}$, etcetera.

The ladder on the right indicates that $F_{n+1}$ is a morphism from $P_{n+1}$ to
$Q_{n+1}$ in the appropriate context, that $F_{n+2}$ is a morphism from
$P_{n+2}$ to $Q_{n+2}$ over $F_{n+1}$, etcetera.

Note that the object(s) at the bottom of a ladder are always families of contexts
over $\Gamma$, so that the typing of the various ingredients makes sense.
\end{itemize}
Such a diagram is said to be commutative if the subdiagram consisting of only
the normal headed arrows is commutative in the usual sense (using judgmental
equality). Note that the ladders are inherently commutative.
\end{defn}

The most basic illustrative example of a commutative dependency diagram is
the diagram
\begin{equation*}
\begin{tikzcd}
P \ar[fib]{d} \ar{r}{F} & Q \ar[fib]{d} \\
A \ar{r}{f} & B
\end{tikzcd}
\end{equation*}
indicating a morphism $F$ from $P$ to $Q$ over the morphism $f$ from $A$ to
$B$ in a context $\Gamma$.

We can just copy the usual categorical definition of a pullback square to our
current situation, but we have to require that each arrow in the pullback square
is an ordinary arrow. When families (i.e. double-headed arrows) are involved
in the diagram, we make the following definition of a family pullback:

\begin{defn}
We say that a commutative dependency diagram of the form
\begin{equation*}
\begin{tikzcd}
P \ar[fib]{d} \ar{r}{F} & Q \ar[fib]{d} \\
A \ar{r}{f} & B
\end{tikzcd}
\end{equation*}
is a \emph{family pullback} if the following inference rules are valid:
\begin{align*}
& \inference
  { \jfam{{\Gamma}{A}}{P'}
    \jfhom{\Gamma}{A}{B}{f}{P'}{Q}{F'}
    }
  { \jhom{{\Gamma}{A}}{P'}{P}{u}
    }
  \\
& \inference
  { \jfam{{\Gamma}{A}}{P'}
    \jfhom{\Gamma}{A}{B}{f}{P'}{Q}{F'}
    }
  { \jfhomeq{\Gamma}{A}{B}{f}{P'}{Q}{\jcomp{}{u}{F}}{F'}
    }
  \\
& \inference
  { \jhom{{\Gamma}{A}}{P'}{P}{v}
    \jfhomeq{\Gamma}{A}{B}{f}{P'}{Q}{\jcomp{}{v}{F}}{F'}
    }
  { \jhomeq{{\Gamma}{A}}{P'}{P}{v}{u}
    }
\end{align*}
\end{defn}

The following lemma explains that when a square involving families is a
family pullback square whenever the corresponding square involving projections is a
pullback square. There is no proof in the the opposite direction.

\begin{lem}
A square
\begin{equation}\label{eq:fpb_to_pb_eqv_fpb}
\begin{tikzcd}
P \ar[fib]{d} \ar{r}{F} & Q \ar[fib]{d} \\
A \ar{r}{f} & B
\end{tikzcd}
\end{equation}
is a family pullback square whenever the square
\begin{equation}\label{eq:fpb_to_pb_eqv_pb}
\begin{tikzcd}[column sep = large]
\ctxext{A}{P} \ar{d}[swap]{\cprojfstf{A}{P}} \ar{r}{\tmext{\ctxwk{P}{f}}{F}} & \ctxext{B}{Q} \ar{d}{\cprojfstf{B}{Q}} \\
A \ar{r}{f} & B
\end{tikzcd}
\end{equation}
is a pullback square.
\end{lem}

The family pullback of a family along any morphism always exists. It is simply given
by the precomposition of the family with the morphism. Note that this fact does
not carry over to arbitrary pullbacks.

\begin{lem}
The diagram
\begin{equation*}
\begin{tikzcd}
\jcomp{}{f}{Q} \ar[fib]{d} \ar{r}{\idtm{\jcomp{}{f}{Q}}} & Q \ar[fib]{d} \\
A \ar{r}{f} & B
\end{tikzcd}
\end{equation*}
is a family pullback diagram.
\end{lem}

\begin{proof}
The proof is a triviality because $\jhom{{\Gamma}{A}}{P'}{\jcomp{}{f}{Q}}{F'}$
is the same judgment as $\jfhom{\Gamma}{A}{B}{f}{P'}{Q}{F'}$ and
$\jcomp{}{\idtm{\jcomp{}{f}{Q}}}{F'}\jdeq F'$.
\end{proof}

For arbitrary pullbacks we have the pasting lemma as usual, but for family
pullbacks we can only derive one of the parts of the pasting lemma.

\begin{lem}
Suppose we have the diagram
\begin{equation*}
\begin{tikzcd}
P \ar{r}{F} \ar[fib]{d} & Q \ar{r}{G} \ar[fib]{d} & R \ar[fib]{d}\\
A \ar{r}{f} & B \ar{r}{g} & C
\end{tikzcd}
\end{equation*}
where the square on the right and the outer rectangle are family pullback 
diagrams. Then the square on the left is a family pullback diagram.
\end{lem}

\begin{proof}
Let $\jfam{{\Gamma}{A}}{P'}$ be a family and let $\jfhom{\Gamma}{A}{B}{f}
{P'}{Q}{F}$ be a morphism over $f$.
\begin{itemize}
\item Then we compose $F'$ with $G$ to obtain a morphism over $\jcomp{}{f}{g}$.
\item Then we get $\jhom{{\Gamma}{A}}{P'}{P}{u}$ with a uniqueness property.
      The property that $\jcomp{}{u}{F}\jdeq F'$ follows from the assumption
      that the right square is a pullback.
\item Now assume that we have another such $v$. Compose it with $F$ and $G$.
      By the assumed properties this is the same as $u$ composed with $F$ and
      $G$. By the pullback condition we now get $u\jdeq v$. 
\end{itemize}
\end{proof}

\begin{lem}
For any $\jterm{\Gamma}{A}{x}$ and any $\jfam{{\Gamma}{A}}{P}$, the square
\begin{equation*}
\begin{tikzcd}
\subst{x}{P} \ar{d} \ar{r}{\finc{x}{P}} & \ctxext{A}{P} \ar{d}{\cprojfstf{A}{P}}\\
\emptyf \ar{r}{\ctxwk{\emptyf}{x}} & A
\end{tikzcd}
\end{equation*}
is a pullback square.
\end{lem}

In the following lemma we assert that pulling back a family $Q$ 
over $\ctxext{{\Gamma}{A}}{P}$ along a fiber
inclusion $\finc{x}{P}$ gives the family $\subst{x}{Q}$ over $\ctxext{\Gamma}{\subst{x}{P}}$.

\begin{lem}
We have the following inference rule:
\begin{equation*}
\inference
  { \jfam{{{\Gamma}{A}}{P}}{Q}
    \jterm{\Gamma}{A}{x}
    }
  { \jfameq
      {{\Gamma}{\subst{x}{P}}}
      {\jcomp{}{\finc{x}{P}}{Q}}
      {\subst{x}{Q}}
    }
\end{equation*}
\end{lem}

\begin{proof}
We have the judgmental equalities:
\begin{align*}
\jcomp{\subst{x}{P}}{\finc{x}{P}}{Q}
& \jdeq
  \subst{\tmext{\ctxwk{\subst{x}{P}}{x}}{\idtm{\subst{x}{P}}}}{\ctxwk{\subst{x}{P}}{Q}}
  \\
& \jdeq
  \subst{\idtm{\subst{x}{P}}}{{\ctxwk{\subst{x}{P}}{x}}{\ctxwk{\subst{x}{P}}{Q}}}
  \\
& \jdeq
  \subst{\idtm{\subst{x}{P}}}{\ctxwk{\subst{x}{P}}{\subst{x}{Q}}}
  \\
& \jdeq
  \subst{x}{Q}.
\end{align*}
\end{proof}

\subsection{Another special case of projections}
In this subsection we investigate the special case of a projection which
appears as a morphism from $\ctxext{{A}{P}}{\ctxwk{P}{Q}}$ to $\ctxext{A}{Q}$
in context $\Gamma$, where we assume to have the families 
$\jfam{{\Gamma}{A}}{P}$ and $\jfam{{\Gamma}{A}}{Q}$.

Note that we have the judgmental
equalities
\begin{align*}
\ctxwk{\ctxext{{A}{P}}{\ctxwk{P}{\mfam{A}}}}{\ctxext{A}{\mfam{A}}}
& \jdeq 
  \ctxext
    {\ctxwk{\ctxext{{A}{P}}{\ctxwk{P}{\mfam{A}}}}{A}}
    {\ctxwk{\ctxext{{A}{P}}{\ctxwk{P}{\mfam{A}}}}{\mfam{A}}}
  \\
& \jdeq
  \ctxext
    {\ctxwk{{P}{\mfam{A}}}{{\ctxext{A}{P}}{A}}}
    {\ctxwk{\ctxext{{A}{P}}{\ctxwk{P}{\mfam{A}}}}{\mfam{A}}}
\end{align*}
Note that we have the term $\ctxwk{{P}{\mfam{A}}}{\cprojfstf{A}{P}}$ of the
family $\ctxwk{{P}{\mfam{A}}}{{\ctxext{A}{P}}{A}}$. Therefore, we need to
find a term of type $\subst{\ctxwk{{P}{\mfam{A}}}{\cprojfstf{A}{P}}}
{\ctxwk{\ctxext{{A}{P}}{\ctxwk{P}{\mfam{A}}}}{\mfam{A}}}$. Note that we have
the judgmental equalities:
\begin{align*}
\subst
  {\ctxwk{{P}{\mfam{A}}}{\cprojfstf{A}{P}}}
  {\ctxwk{\ctxext{{A}{P}}{\ctxwk{P}{\mfam{A}}}}{\mfam{A}}}
& \jdeq
  \subst
    {\ctxwk{{P}{\mfam{A}}}{\cprojfstf{A}{P}}}
    {\ctxwk{{P}{\mfam{A}}}{{\ctxext{A}{P}}{\mfam{A}}}}
  \\
& \jdeq
  \ctxwk
    { {P}{\mfam{A}}
      }
    { \subst
        {\cprojfstf{A}{P}}
        {\ctxwk{\ctxext{A}{P}}{\mfam{A}}}
      }
  \\
& \jdeq
  \ctxwk
    { {P}{\mfam{A}}
      }
    { {P}{\mfam{A}}
      }
  \\
& \jdeq
  \ctxwk{P}{{\mfam{A}}{\mfam{A}}}
\end{align*}
We find the term $\ctxwk{P}{\idtm{\mfam{A}}}$ here. Thus we can now define
$\bar{\typefont{pr}}$ by:
\begin{equation}\label{barproj}
\jhomdefn
  {\Gamma}
  {{{A}{P}}{\mfam{A}}}
  {{A}{\mfam{A}}}
  {\bar{\typefont{pr}}}
  {\tmext{\ctxwk{{P}{\mfam{A}}}{\cprojfstf{A}{P}}}{\ctxwk{P}{\idtm{\mfam{A}}}}}
\end{equation}

\begin{lem}
We have the judgmental equality
\begin{equation*}
\jfameq
  {{{{\Gamma}{A}}{P}}{\ctxwk{P}{\mfam{A}}}}
  {\jcomp{}{\bar{\typefont{pr}}}{Q}}
  {\ctxwk{P}{Q}}
\end{equation*}
for any family $Q$ of contexts over $\ctxext{{\Gamma}{A}}{\mfam{A}}$ 
\end{lem}

\subsection{Inductive morphisms}
\begin{defn}
Let $\jhom{\Gamma}{A}{B}{f}$ be a context morphism. We say that $f$ is an 
\emph{inductive morphism} if the following inference rules are valid:
\begin{align*}
& \inference
  { \jfam{{\Gamma}{B}}{Q}
    \jterm{{\Gamma}{A}}{\jcomp{A}{f}{Q}}{g}
    }
  { \jterm{{\Gamma}{B}}{Q}{\jtcext{g}}
    }
& & \inference
  { \jfam{{\Gamma}{B}}{Q}
    \jterm{{\Gamma}{A}}{\jcomp{A}{f}{Q}}{g}
    }
  { \jtermeq{{\Gamma}{A}}{\jcomp{A}{f}{Q}}{\jcomp{A}{f}{\jtcext{g}}}{g}
    }
\end{align*}
\end{defn}

When $f$ is a context morphism from $A$ to $B$ in context $\Gamma$, 
finding a term of a family $Q$ over $\ctxext{\Gamma}{B}$ can be accomplished
by finding a term of the family $\jcomp{A}{f}{Q}$ over $\ctxext{\Gamma}{A}$.
Inductive morphisms come with a computation rule. Thus, term
constructors of inductively defined types are going to be the major source of
examples of inductive morphisms. Since identity terms behave
so nicely, they are inductive morphisms too.

\begin{lem}
The identity term
$\jhom{\Gamma}{A}{A}{\idtm{A}}$ is an inductive morphism
for each family $A$ of contexts over $\Gamma$
\end{lem}

\begin{itemize}
\item Extensions of inductive morphisms are inductive
\item Weakenings of inductive morphisms are inductive
\item Substitutions of inductive morphisms are inductive
\end{itemize}

### 2014-04-03 article-models-ttderived.tex
added compatibility properties of fiber inclusions

@@
 
-There is a good amount of things we can already state and prove about in the
-theory we have developed so far. We will develop a bit more theory that will
-come in handy later on.
-
-In this section we use the framework we have developed in this
-section to derive new notions and their properties. In particular, we will
+In this section we use the framework we have developed in \autoref{tt}
+to derive new notions and their properties. In particular, we will
 develop the notion of \emph{extension on terms} together with the projection
@@
 the domain context and finding a term there.
+\emph{This section contains no new assumptions.}
 
@@
 
-\emph{This subsection contains no new assumptions.}
-
 \subsection{Projections and extension on terms}\label{extension-on-terms}
@@
 Let $\jterm{\Gamma}{A}{x}$ be a term and let $\jfam{{\Gamma}{A}}{P}$ be a
-family. Then we define
-\begin{equation*}
-\jhomdefn{\Gamma}{\subst{x}{P}}{{A}{P}}{\finc{x}{P}}{\unfoldnext{\finc{x}{P}}}
+family. Then we define the \emph{fiber inclusion} of $\subst{x}{P}$ into
+$\ctxext{A}{P}$ in context $\Gamma$ to be the morphism
+\begin{equation*}
+\jhomdefn{\Gamma}{\subst{x}{P}}{{A}{P}}{\finc{x}{P}}{\unfoldnext{\finc{x}{P}}}.
 \end{equation*}
 \end{defn}
+
+We have the following lemmas expressing the compatibility of the fiber
+inclusions with the empty context, extension, weakening and substitution.
+
+\begin{lem}
+The fiber inclusions are compatible with the empty families; i.e.~the following
+inference rules are valid
+\begin{align*}
+& \inference
+  { \jterm{\Gamma}{A}{x}
+    }
+  { \jtermeq
+      {\Gamma}
+      {A}
+      {\finc{x}{\emptyf}}
+      {x}
+    }
+  \\
+& \inference
+  { \jfam{\Gamma}{A}
+    }
+  { \jhomeq
+      {\Gamma}
+      {A}
+      {A}
+      {\finc{\emptytm}{A}}
+      {\idtm{A}}
+    }
+\end{align*}
+\end{lem}
+
+\begin{lem}
+The fiber inclusions are compatible with extension; i.e.~the following inference
+rule is valid
+\begin{equation*}
+\inference
+  { \jterm{\Gamma}{A}{x}
+    \jterm{\Gamma}{\subst{x}{P}}{u}
+    \jfam{{{\Gamma}{A}}{P}}{Q}
+    }
+  { \jhomeq
+      {\Gamma}
+      {\subst{\tmext{x}{u}}{Q}}
+      {{{A}{P}}{Q}}
+      {\finc{\tmext{x}{u}}{Q}}
+      {\jcomp{}{\finc{u}{\subst{x}{Q}}}{\finc{x}{\ctxext{P}{Q}}}}
+    }
+\end{equation*}
+\end{lem}
+
+\begin{lem}
+The fiber inclusions are compatible with weakening; i.e.~the following inference
+rule is valid
+\begin{equation*}
+\inference
+  { \jfam{\Gamma}{A}
+    \jterm{\Gamma}{B}{y}
+    \jfam{{\Gamma}{B}}{Q}
+    }
+  { \jhomeq
+      {{\Gamma}{A}}
+      {\subst{\ctxwk{A}{y}}{\ctxwk{A}{Q}}}
+      {{\ctxwk{A}{B}}{\ctxwk{A}{Q}}}
+      {\finc{\ctxwk{A}{y}}{\ctxwk{A}{Q}}}
+      {\ctxwk{A}{\finc{y}{Q}}}
+    }
+\end{equation*}
+\end{lem}
+
+\begin{lem}
+The fiber inclusions are compatible with substitution; i.e.~the following
+inference rule is valid
+\begin{equation*}
+\inference
+  { \jterm{\Gamma}{A}{x}
+    \jfam{{{\Gamma}{A}}{P}}{Q}
+    \jterm{{\Gamma}{A}}{P}{f}
+    }
+  { \jhomeq
+      {\Gamma}
+      {\subst{{x}{f}}{{x}{Q}}}
+      {{\subst{x}{P}}{\subst{x}{Q}}}
+      {\finc{\subst{x}{f}}{\subst{x}{Q}}}
+      {\subst{x}{\finc{f}{Q}}}
+    }
+\end{equation*}
+\end{lem}
+
+\begin{lem}
+The fiber inclusions are compatible with identity terms; i.e.~the following
+inference rule is valid
+\begin{equation*}
+\inference
+  { \jfam{{\Gamma}{A}}{P}
+    }
+  { \jhomeq
+      {{\Gamma}{A}}
+      {P}
+      {\ctxwk{A}{\ctxext{A}{P}}}
+      {\finc{\idtm{A}}{\ctxwk{A}{P}}}
+      {\idtm{{A}{P}}}
+    }
+\end{equation*}
+\end{lem}
 

### 2014-04-03 article-models-ttderived.tex
working on section 5

@@
 
-We find the following inference rule, which expresses that the identity term
-is compatible with extension:
-
-\begin{lem}
-For any $\jfam{\Gamma}{A}$ and $\jfam{{\Gamma}{A}}{P}$ we have
-\begin{equation}\label{idfunc-ext-comp}
-\inference
-  { \jfam{\Gamma}{A}
-    \jfam{{\Gamma}{A}}{P}
-    }
-  { \jhomeq
-      {\Gamma}
-      {{A}{P}}{{A}{P}}
-      {\idtm{\ctxext{A}{P}}}
-      { \tmext
-          {\ctxwk{\ctxext{A}{P}}{A}}
-          {\ctxwk{\ctxext{A}{P}}{P}}
-          {\cprojfstf{A}{P}}
-          {\cprojsndf{A}{P}}
-        }
-    }
-\end{equation}
-\end{lem}
-
-\begin{proof}
-Consider the families $\jfam{\Gamma}{A}$ and $\jfam{{\Gamma}{A}}{P}$. Then
-we have the judgmental equalities
-\begin{align*}
-\tmext
-  {\ctxwk{\ctxext{A}{P}}{A}}
-  {\ctxwk{\ctxext{A}{P}}{P}}
-  {\cprojfstf{A}{P}}
-  {\cprojsndf{A}{P}}
-& \jdeq 
-  \unfold
-  { \tmext
-      {\ctxwk{\ctxext{A}{P}}{A}}
-      {\ctxwk{\ctxext{A}{P}}{P}}
-      {\cprojfstf{A}{P}}
-      {\cprojsndf{A}{P}}
-    }
-  \\
-& \jdeq 
-  \subst
-    { \idtm{P}
-      }
-    { {\ctxwk{P}{\idtm{A}}}
-      {\idtm{\ctxwk{\ctxext{A}{P}}{\ctxext{A}{P}}}}
-      }
-  \\
-& \jdeq 
-  \subst
-    { \idtm{P}
-      }
-    { {\ctxwk{P}{\idtm{A}}}
-      {\ctxwk{\ctxext{A}{P}}{\idtm{\ctxext{A}{P}}}}
-      }
-  \\
-& \jdeq 
-  \subst
-    { \idtm{P}
-      }
-    { {\ctxwk{P}{\idtm{A}}}
-      {\ctxwk{P}{{A}{\idtm{\ctxext{A}{P}}}}}
-      }
-  \\
-& \jdeq 
-  \subst
-    { \idtm{P}
-      }
-    {\ctxwk
-      {P}
-      { \subst
-        {\idtm{A}}
-        {\ctxwk{A}{\idtm{\ctxext{A}{P}}}}
-        }
-      }
-  \\
-& \jdeq 
-  \subst
-    { \idtm{A}
-      }
-    { \ctxwk{A}{\idtm{\ctxext{A}{P}}}
-      }
-  \\
-& \jdeq 
-  \idtm{\ctxext{A}{P}}.
-  \qedhere
-\end{align*}
-\end{proof}
-
-\begin{lem}
-The following inference is valid:
-\begin{equation*}
-\inference
-  { \jfam{{\Gamma}{A}}{P}
-    \jfam{{\Gamma}{A}}{Q}
-    }
-  { \jfameq
-      {{\Gamma}{{A}{P}}}
-      {\subst{\cprojfstf{A}{P}}{\ctxwk{\ctxext{A}{P}}{Q}}}
-      {\ctxwk{P}{Q}}
-    }
-\end{equation*}
-\end{lem}
-
-\begin{proof}
-Let $\jfam{{\Gamma}{A}}{P}$ and $\jfam{{\Gamma}{A}}{Q}$ be
-families. Then we have
-\begin{align*}
-\subst{\cprojfstf{A}{P}}{\ctxwk{\ctxext{A}{P}}{Q}} 
-& \jdeq 
-  \subst{\ctxwk{P}{\idtm{A}}}{\ctxwk{\ctxext{A}{P}}{Q}} 
-  \tag{by definition}\\
-& \jdeq 
-  \subst{\ctxwk{P}{\idtm{A}}}{\ctxwk{P}{{A}{Q}}} 
-  \tag{by \autoref{comp-ww-f}}\\
-& \jdeq 
-  \ctxwk{P}{\subst{\idtm{A}}{\ctxwk{A}{Q}}} 
-  \tag{by \autoref{comp-ws-f}}\\
-& \jdeq 
-  \ctxwk{P}{Q} 
-  \tag{by \autoref{idfunc-wk-defn}}
-\end{align*}
-\end{proof}
-
 The constructions of the terms $\tmext{A}{P}{x}{u}$ and $\cprojfst{A}{P}{w}$ and
@@
 
-\begin{lem} The following inference rules expressing that pairing is a strict
+\begin{lem}\label{lem:tmext-basic}
+The following inference rules expressing that pairing is a strict
 inverse to the combination of decompositions, are valid:
@@
 
-\begin{lem}
-The following compatibility rules for two consecutive term extensions is valid:
+In \autoref{lem:tmext-emp,lem:tmext-ext,lem:tmext-wk,lem:tmext-subst,lem:tmext-id}
+we show that term extension and the projections are
+compatible with the empty families, extension,
+weakening, substitution and the identity terms (in that order). \autoref{lem:tmext-id}
+is in fact a generalization of the above \autoref{lem:tmext-basic}.
+
+\begin{lem}\label{lem:tmext-emp}
+The following compatibility rules for extensions of the term of the empty family
+are valid:
+\begin{align*}
+& \inference
+  { \jterm{\Gamma}{A}{x}
+    }
+  { \jtermeq{\Gamma}{A}{\tmext{\emptytm}{x}}{x}
+    }
+& & \inference
+  { \jterm{\Gamma}{A}{x}
+    }
+  { \jtermeq{\Gamma}{A}{\tmext{x}{\emptytm}}{x}
+    }
+  \\
+& \inference
+  { \jfam{\Gamma}{A}
+    }
+  { \jtermeq
+      {{\Gamma}{A}}
+      {\emptyf}
+      {\cprojfstf{\emptyf}{A}}
+      {\emptytm}
+    }
+& & \inference
+  { \jfam{\Gamma}{A}
+    }
+  { \jhomeq
+      {\Gamma}
+      {A}
+      {A}
+      {\cprojfstf{A}{\emptyf}}
+      {\idtm{A}}
+    }
+  \\
+& \inference
+  { \jfam{\Gamma}{A}
+    }
+  { \jhomeq
+      {\Gamma}
+      {A}
+      {A}
+      {\cprojsndf{\emptyf}{A}}
+      {\idtm{A}}
+    }
+& & \inference
+  { \jfam{\Gamma}{A}
+    }
+  { \jtermeq
+      {{\Gamma}{A}}
+      {\emptyf}
+      {\cprojsndf{A}{\emptyf}}
+      {\emptytm}
+    }
+\end{align*}
+\end{lem}
+
+\begin{proof}
+These equalities are very easy to verify. We only display a proof of the first:
+\begin{equation*}
+\tmext{\emptytm}{x}
+\jdeq \unfold{\tmext{\emptyf}{A}{\emptytm}{x}}
+\jdeq \subst{x}{\idtm{{\emptyf}{A}}}
+\jdeq \subst{x}{\idtm{A}}
+\jdeq x.\qedhere
+\end{equation*}
+\end{proof}
+
+\begin{lem}\label{lem:tmext-ext}
+The following compatibility rules for two consecutive term extensions are valid:
 \begin{align*}
@@
 
-\begin{lem}
+\begin{lem}\label{lem:tmext-wk}
 When we weaken a term $\tmext{B}{Q}{y}{v}$ of $\ctxext{B}{Q}$ in context $\Gamma$ by
@@
 
-\begin{lem}
+\begin{lem}\label{lem:tmext-subst}
 When we substitute an extended term $\tmext{P}{Q}{f}{g}$ of $\ctxext{P}{Q}$ by a term
@@
   \subst{x}{\cprojsndf{Q}{R}}.
+  \qedhere
+\end{align*}
+\end{proof}
+
+We find the following inference rule, which expresses that the identity term
+is compatible with extension:
+
+\begin{lem}\label{lem:tmext-id}
+For any $\jfam{\Gamma}{A}$ and $\jfam{{\Gamma}{A}}{P}$ we have
+\begin{equation}\label{idfunc-ext-comp}
+\inference
+  { \jfam{\Gamma}{A}
+    \jfam{{\Gamma}{A}}{P}
+    }
+  { \jhomeq
+      {\Gamma}
+      {{A}{P}}{{A}{P}}
+      {\idtm{\ctxext{A}{P}}}
+      { \tmext
+          {\ctxwk{\ctxext{A}{P}}{A}}
+          {\ctxwk{\ctxext{A}{P}}{P}}
+          {\cprojfstf{A}{P}}
+          {\cprojsndf{A}{P}}
+        }
+    }
+\end{equation}
+\end{lem}
+
+\begin{proof}
+Consider the families $\jfam{\Gamma}{A}$ and $\jfam{{\Gamma}{A}}{P}$. Then
+we have the judgmental equalities
+\begin{align*}
+\tmext
+  {\ctxwk{\ctxext{A}{P}}{A}}
+  {\ctxwk{\ctxext{A}{P}}{P}}
+  {\cprojfstf{A}{P}}
+  {\cprojsndf{A}{P}}
+& \jdeq 
+  \unfold
+  { \tmext
+      {\ctxwk{\ctxext{A}{P}}{A}}
+      {\ctxwk{\ctxext{A}{P}}{P}}
+      {\cprojfstf{A}{P}}
+      {\cprojsndf{A}{P}}
+    }
+  \\
+& \jdeq 
+  \subst
+    { \idtm{P}
+      }
+    { {\ctxwk{P}{\idtm{A}}}
+      {\idtm{\ctxwk{\ctxext{A}{P}}{\ctxext{A}{P}}}}
+      }
+  \\
+& \jdeq 
+  \subst
+    { \idtm{P}
+      }
+    { {\ctxwk{P}{\idtm{A}}}
+      {\ctxwk{\ctxext{A}{P}}{\idtm{\ctxext{A}{P}}}}
+      }
+  \\
+& \jdeq 
+  \subst
+    { \idtm{P}
+      }
+    { {\ctxwk{P}{\idtm{A}}}
+      {\ctxwk{P}{{A}{\idtm{\ctxext{A}{P}}}}}
+      }
+  \\
+& \jdeq 
+  \subst
+    { \idtm{P}
+      }
+    {\ctxwk
+      {P}
+      { \subst
+        {\idtm{A}}
+        {\ctxwk{A}{\idtm{\ctxext{A}{P}}}}
+        }
+      }
+  \\
+& \jdeq 
+  \subst
+    { \idtm{A}
+      }
+    { \ctxwk{A}{\idtm{\ctxext{A}{P}}}
+      }
+  \\
+& \jdeq 
+  \idtm{\ctxext{A}{P}}.
   \qedhere
@@
 
+\begin{lem}
+The following inference rule is valid:
+\begin{equation*}
+\inference
+  { \jfam{{\Gamma}{A}}{P}
+    \jfam{{\Gamma}{A}}{Q}
+    }
+  { \jfameq
+      {{\Gamma}{{A}{P}}}
+      {\jcomp{{A}{P}}{\cprojfstf{A}{P}}{Q}}
+      {\ctxwk{P}{Q}}
+    }
+\end{equation*}
+\end{lem}
+
+\begin{proof}
+Let $\jfam{{\Gamma}{A}}{P}$ and $\jfam{{\Gamma}{A}}{Q}$ be
+families. Then we have
+\begin{align*}
+\jcomp{{A}{P}}{\cprojfstf{A}{P}}{Q}
+& \jdeq
+  \unfoldall{\jcomp{{A}{P}}{\cprojfstf{A}{P}}{Q}}
+  \tag{by definition}\\
+& \jdeq 
+  \subst{\ctxwk{P}{\idtm{A}}}{\ctxwk{P}{{A}{Q}}} 
+  \tag{by \autoref{comp-ww-f}}\\
+& \jdeq 
+  \ctxwk{P}{\subst{\idtm{A}}{\ctxwk{A}{Q}}} 
+  \tag{by \autoref{comp-ws-f}}\\
+& \jdeq 
+  \ctxwk{P}{Q} 
+  \tag{by \autoref{idfunc-wk-defn}}
+\end{align*}
+\end{proof}
+
 \subsection{Another special case of projections}

### 2014-04-08 article-models-tt.tex
rewritten the section on the empty context, moved all the type-stuff to the end

@@
 We will
-formulate the type theory in such a way that contexts aren't defined
+formulate the theory of contexts, families and terms in such a way that contexts aren't defined
 to be lists of variable declarations. The variable-free (or name-free) approach 
@@
 \subsection{Judgments and inference rules}\label{judgments}
-The type theory we describe here is a theory of contexts, families of
+The theory we describe here is a theory of contexts, families of
 contexts and terms thereof. The families of contexts are by some authors called
@@
 
-Thus we make eight kinds of judgments in our type theory: ``$\Gamma$ is a context'',
+Thus we make eight kinds of judgments: ``$\Gamma$ is a context'',
 ``$A$ is a family of contexts over $\Gamma$'', ``$A$ is a type in context $\Gamma$''
@@
   \\
-\jtype*{\Gamma}{A} 
-& \jtypeeq*{\Gamma}{A}{B}
-  \\
 \jterm*{\Gamma}{A}{x} 
@@
 \end{align*}
+
+Strictly speaking, we have three different judgmental equalities in play and one
+could request for a notational difference to signify that fact. For instance,
+we could denote the judgmental equalities of contexts, families and terms by
+$\jdeq_c$, $\jdeq_f$ and $\jdeq_t$ respectively. It will, however, always be
+clear which of the three kinds of judgmental equality is meant when we assert
+a judgmental equality and therefore we shall not bother to make this notational
+distinction.
 
@@
 
-If $A$ is a type
-in context $\Gamma$, then $A$ is also a family of contexts over $\Gamma$. Being
-a term of type $A$ means the same thing as being a term of the family $A$ of contexts.
-Two types in context $\Gamma$ are judgmentally equal precisely when they are equal
-as context families. Moreover, if a family $B$ of contexts over $\Gamma$ is
-judgmentally equal to a type $A$ in context $\Gamma$, then $B$ is a type in
-context $\Gamma$. This is expressed by the following four inference rules:
-
-\begin{align*}
-& \inference
-  { \jtype{\Gamma}{A}
-    }
-  { \jfam{\Gamma}{A}
-    }
-& & \inference
-    { \jtypeeq{\Gamma}{A}{B}
-      }
-    { \jfameq{\Gamma}{A}{B}
-      }
-    \\
-& \inference
-  { \jtype{\Gamma}{A}
-    \jfameq{\Gamma}{A}{B}
-    }
-  { \jtype{\Gamma}{B}
-    }
-& & \inference
-    { \jtype{\Gamma}{A}
-      \jfameq{\Gamma}{A}{B}
-      }
-    { \jtypeeq{\Gamma}{A}{B}
-      }
-\end{align*}
-
 \subsection{The basic rules for judgmental equality}
@@
 \subsection{The empty context}
-There is an empty context and over any context there is an empty family of
-contexts. We do not assume that the empty context is a type, that would be like
-assuming that the multiplicative unit of a ring is prime. The empty family
-always has a unique term.
+In the theory of contexts, families and terms we introduce an empty context and,
+for every context $\Gamma$, an empty family over $\Gamma$. 
+When contexts are viewed as statements, the empty context is the statement that
+asserts nothing. Likewise, in any context one can assert nothing and this is
+achieved by the empty family. 
+Since somebody who states nothing asserts it all right, there is a term of the empty
+family over any context. Moreover, any two terms of the empty family are always
+judgmentally equal.
 
@@
 
-Moreover, if $\Gamma$ is a context family over the
-empty context, then $\Gamma$ is a context and every context is a context
-family over the empty context. Note that this allows us to speak
-of terms of contexts too.
+The empty context and the empty families together with the operation of
+extension will provide several shortcuts for the rest
+of the theory of families and terms. This is (in part) obtained by requiring
+that a context is exactly the same thing as a family of contexts over the
+empty context and that judgmental equality of contexts is exactly the same
+as judgmental equality of families of contexts over the empty context.
+
+By regarding contexts as families of contexts over the empty context, we
+enable ourselves also to speak of terms of contexts. A term of a context
+$\Gamma$ is a term of the family $\Gamma$ over the empty context.
 
@@
 
-\subsubsection{The empty context is compatible with itslef}
+\subsubsection{The empty context is compatible with itself}
 The empty context $\emptyc$ may be considered as a family of contexts over the empty
@@
 in context $\Gamma$ and a family $P$ over it to a family $\ctxext{A}{P}$ over context
-$\Gamma$. We do this to ensure that all of type theory can be done in a context.
+$\Gamma$. We do this to ensure that all of the theory of contexts, families and
+terms can be done in a context.
 For instance, we could say (1) that a context in context $\Gamma$ is the same thing
@@
   { \jfameq{\Gamma}{\ctxext{\emptyf}{A}}{A}
+    }
+  \\
+& \inference
+  { \jfam{\Gamma}{A}
+    }
+  { \jfameq{\Gamma}{{A}{\emptyf}}{A}
     }
@@
 
-\subsection{The type theoretic operation of weakening}
+\subsection{Weakening}
 When $A$ is a context family over a context $\Gamma$, we wish to define a weakening
@@
         {\ctxwk{A'}{g'}}
-      }
-\end{align}
-
-We add rules asserting that a weakened type is again a type:
-
-\begin{align}
-& \inference
-  { \jfam{\Gamma}{B}
-    \jtype{\ctxext{\Gamma}{B}}{Q}
-    }
-  { \jtype{\ctxext{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{Q}}
       }
@@
 
-\subsection{The type theoretic operation of substitution}
+\subsection{Substitution}
 Given a family $P$ over $A$ and a term $x$ of $A$, substitution gives a way to
@@
         {\subst{x'}{g'}}
-      }
-\end{align}
-
-We also add the following rule asserting that substitution preserves the
-property of being a type:
-\begin{align}
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jtype{\ctxext{{\Gamma}{A}}{P}}{Q}
-    }
-  { \jtype{\ctxext{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}
       }
@@
 \end{equation*}
-for every $1\leq i\leq n$. Thus, it establishes the projections. In our setting,
+for every $1\leq i\leq n$. Thus, it establishes the projections. In 
+the current setting,
 we get the projections from the identity morphisms together with weakening. We
@@
 \end{proof}
+
+\subsection{The possiblity of types in the theory of contexts, families and
+terms}
+We have deliberately not spoken of types so far because we have taken the point
+of view that a type in a context is nothing but a family in that context which
+belongs to the class of types. We think of types as \emph{irreducible} families,
+i.e.\ families which are neither the empty context nor the extension of two
+families which are both not the empty context (in algebraic terminology: which
+are both non-trivial). To allow ourselves to speak of types we introduce two
+new judgments: the judgment that something is a type and the judgment that two
+types are equal.
+\begin{align*}
+\jtype*{\Gamma}{A} 
+& \jtypeeq*{\Gamma}{A}{B}
+\end{align*}
+But only families of contexts are eligible to be types. If $A$ is a type
+in context $\Gamma$, then $A$ is also a family of contexts over $\Gamma$. 
+Moreover, two types in context $\Gamma$ are judgmentally equal precisely when they are equal
+as context families and if a family $B$ of contexts over $\Gamma$ is
+judgmentally equal to a type $A$ in context $\Gamma$, then $B$ is a type in
+context $\Gamma$. This is expressed by the following four inference rules:
+\begin{align*}
+& \inference
+  { \jtype{\Gamma}{A}
+    }
+  { \jfam{\Gamma}{A}
+    }
+& & \inference
+    { \jtypeeq{\Gamma}{A}{B}
+      }
+    { \jfameq{\Gamma}{A}{B}
+      }
+    \\
+& \inference
+  { \jtype{\Gamma}{A}
+    \jfameq{\Gamma}{A}{B}
+    }
+  { \jtype{\Gamma}{B}
+    }
+& & \inference
+    { \jtype{\Gamma}{A}
+      \jfameq{\Gamma}{A}{B}
+      }
+    { \jtypeeq{\Gamma}{A}{B}
+      }
+\end{align*}
+As pointed out at the beginning of this subsection, 
+we do not assume that the empty context is a type, that would be like
+assuming that the multiplicative unit of a ring is prime. We envision a
+categorical interpretation of the theory of contexts, families and terms where 
+contexts are interpreted as finite paths of arrows belonging to a predetermined
+class. The empty context would be such a path of length $0$ whereas the types
+would be such paths of length $1$.
+
+We add rules asserting that a weakened type is again a type that 
+substitution preserves the property of being a type:
+\begin{align}
+& \inference
+  { \jfam{\Gamma}{B}
+    \jtype{\ctxext{\Gamma}{B}}{Q}
+    }
+  { \jtype{\ctxext{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{Q}}
+    }
+  \\
+& \inference
+  { \jterm{\Gamma}{A}{x}
+    \jtype{\ctxext{{\Gamma}{A}}{P}}{Q}
+    }
+  { \jtype{\ctxext{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}
+    }
+\end{align}
+With only the current rules, the possibility of making the judgment that
+something is a type does not add much to the theory of contexts, families and
+terms. Nevertheless, when studying models, having an interpretation for the
+judgment that something is a type will allow for the possibility to study
+conditions such as the one asserting that every family `factorizes' uniquely
+as multiple applications of extension to types.

### 2014-04-08 article-models-tt.tex
small impro

@@
 conditions such as the one asserting that every family `factorizes' uniquely
-as multiple applications of extension to types.
+as multiple applications of extension to types, analogous to the condition on
+unique factorization domains in ring theory. All the tradional models of type
+theory should translate to models with such a condition.

### 2014-04-08 article-models-tt.tex
small impros

@@
-\section{The theory of contexts, families and terms}\label{tt}
+\section{The algebraic theory of contexts, families and terms}\label{tt}
 In this section we give a description of dependent type theory before type
@@
 and ``$x$ is a term of the family $A$ of contexts over $\Gamma$''. The other four
-judgments are for judgmental equality.
-
-\begin{align*}
+judgments are for judgmental equality. \begin{align*}
 \jctx*{\Gamma} 
@@
 in all three cases (contexts, types and terms).
-\bgroup\small
-\begin{align*}
+\bgroup\small\begin{align*}
 & \inference
@@
 The following convertibility rules are responsible for the strictness
-of judgmental equality, which sets it apart from equivalences or identifications:
-
-\begin{align*}
+of judgmental equality, which sets it apart from equivalences or identifications:\begin{align*}
 & \inference
@@
 family over any context. Moreover, any two terms of the empty family are always
-judgmentally equal.
-
-\begin{align}
+judgmentally equal.\begin{align}
 & \inference
@@
 $\Gamma$ is a term of the family $\Gamma$ over the empty context. 
-
 \begin{align}
@@
 The empty context $\emptyc$ may be considered as a family of contexts over the empty
-context. When we do this, we get $\emptyf[\emptyc]$.
-\begin{equation}
+context. When we do this, we get $\emptyf[\emptyc]$.\begin{equation}
 \inference
@@
 (3) when $P$ is a family over $A$ in this sense, a term of $P$ keeps its original meaning.
-
 \begin{align}
@@
 \subsubsection{Extension is compatible with the empty context}
-The following rule asserts that extension by $\emptyc$ leaves the contexts unchanged.
-\begin{align}
+The following rule asserts that extension by $\emptyc$ leaves the contexts unchanged.\begin{align}
 & \inference
@@
 $A$ is the same thing as a family over $\ctxext{\Gamma}{A}$. 
-
 \begin{align}
@@
  rules asserting a judgmental equality we assume furthermore that 
- $\jfameq{\Gamma}{A}{A'}$.
-\begin{align}
+ $\jfameq{\Gamma}{A}{A'}$.\begin{align}
 & \inference
@@
 The following rules express that when the empty context or context family is
-weakened, the result is the empty context family.
-\begin{align}
+weakened, the result is the empty context family.\begin{align}
 & \inference
@@
 their terms, families over those families and
-terms of those unchanged:
-\begin{align}
+terms of those unchanged:\begin{align}
 & \inference
@@
 
-In the following inference rules we assume that $\jfam{\Gamma}{A}$.
-\begin{align}
+In the following inference rules we assume that $\jfam{\Gamma}{A}$.\begin{align}
 & \inference
@@
 that weakening is compatible with itself. These rules state that the following
-diagram commutes given any two families $A$ and $B$ in context $\Gamma$:
-\begin{equation*}
+diagram commutes given any two families $A$ and $B$ in context $\Gamma$:\begin{equation*}
 \begin{tikzcd}[column sep=huge]
@@
 \end{equation*}
-Thus, we get the following set of inference rules:
-\begin{align}
+Thus, we get the following set of inference rules:\begin{align}
 & \inference
@@
 In the following inference rules we assume that
-$\jfam{\Gamma}{A}$ and $\jfam{{\Gamma}{A}}{P}$. 
-\begin{align}
+$\jfam{\Gamma}{A}$ and $\jfam{{\Gamma}{A}}{P}$. \begin{align}
 & \inference
@@
 in the rules introducing the definitional equalities we assume $\jtermeq{\Gamma}{A}{x}{x'}$.
-
 \begin{align}
@@
 \subsubsection{Substitution is compatible with the empty context}
-The fibers of the empty family are the empty families:
-\begin{align}
+The fibers of the empty family are the empty families:\begin{align}
 & \inference
@@
 The following rules assert that substituting by the term $\jterm{\Gamma}{\emptyf}{\emptytm}$
-leaves everything unchanged.
-\begin{align}
+leaves everything unchanged.\begin{align}
 & \inference
@@
 \subsubsection{Substitution is compatible with extension}\label{comp-se}
-Suppose $\jterm{\Gamma}{A}{x}$ in all of the following inference rule.
-\begin{align}
+Suppose $\jterm{\Gamma}{A}{x}$ in all of the following inference rule.\begin{align}
 & \inference
@@
 that the following diagram commutes for any $\jterm{\Gamma}{A}{x}$ and any
-$\jfam{{\Gamma}{A}}{P}$:
-\begin{equation*}
+$\jfam{{\Gamma}{A}}{P}$:\begin{equation*}
 \begin{tikzcd}[column sep=huge]
@@
 In the following inference rules we assume that $\jterm{\Gamma}{A}{x}$ and
-$\jfam{{\Gamma}{A}}{P}$:
-\begin{align}
+$\jfam{{\Gamma}{A}}{P}$:\begin{align}
 & \inference
@@
 $\jterm{\Gamma}{A}{x}$ and $\jterm{{\Gamma}{A}}{P}{f}$.
-
 \begin{align}
@@
 $\jterm{\Gamma}{B}{y}$.
-
 \begin{align}
@@
 compose context morphisms (terms of weakened contexts).
-
 \begin{align}
@@
 \begin{lem}
-The inference rule
-\begin{equation*}
+The inference rule\begin{equation*}
 \inference
@@
 Let $\jfam{\Gamma}{A}$, $\jfam{\Gamma}{B}$, $\jfam{\Gamma}{C}$ and $\jhom{\Gamma}{A}{B}{f}$.
-Then we have the judgmental equalities
-\begin{align*}
+Then we have the judgmental equalities\begin{align*}
 \subst{f}{\ctxwk{A}{{B}{C}}}
@@
 \begin{defn}
-We define the judgment
-\begin{equation*}
+We define the judgment\begin{equation*}
 \jhom{\Gamma}{A}{B}{f},
@@
 which is pronounced as `$f$ is a morphism from $A$ to $B$ in context $\Gamma$',
-to be the judgment
-\begin{equation*}
+to be the judgment\begin{equation*}
 \unfold{\jhom{\Gamma}{A}{B}{f}}.
 \end{equation*}
-Likewise, we define the judgment
-\begin{equation*}
+Likewise, we define the judgment\begin{equation*}
 \jhomeq{\Gamma}{A}{B}{f}{f'}
 \end{equation*}
-to be the judgment
-\begin{equation*}
+to be the judgment\begin{equation*}
 \unfold{\jhomeq{\Gamma}{A}{B}{f}{f'}}.
@@
 Let $\jhom{\Gamma}{A}{B}{f}$ and consider a term $\jterm{{\Gamma}{B}}{Q}{g}$.
-We define
-\begin{align*}
+We define\begin{align*}
 \jfamdefn*
@@
 Likewise, when we have a family $\jfam{{{\Gamma}{B}}{Q}}{R}$ and a term
-$\jterm{{{\Gamma}{B}}{Q}}{R}{h}$, we define
-\begin{align*}
+$\jterm{{{\Gamma}{B}}{Q}}{R}{h}$, we define\begin{align*}
 \jfamdefn*
@@
 We have the following inference rules about the situation where something is
-substituted by a composition:
-\begin{align*}
+substituted by a composition:\begin{align*}
 & \inference
@@
 We also have the following related inference rules, asserting that composition
-is strictly associative:
-\begin{align*}
+is strictly associative:\begin{align*}
 & \inference
@@
 and a family $\jfam{{{\Gamma}{A}}{\ctxwk{A}{C}}}{R}$. Then we have the judgmental
-equalities
-\begin{align*}
+equalities\begin{align*}
 \subst{\jcomp{A}{f}{g}}{R} 
@@
 
-Now suppose that $\jfam{{\Gamma}{C}}{R}$ instead. Then we have
-\begin{align*}
+Now suppose that $\jfam{{\Gamma}{C}}{R}$ instead. Then we have\begin{align*}
 \jcomp{A}{{A}{f}{g}}{R} 
@@
 We have the following inference rules about the compatibility of composition with
-weakening:
-\begin{align*}
+weakening:\begin{align*}
 & \inference
@@
 Let $\jhom{\Gamma}{A}{B}{f}$, $\jhom{\Gamma}{B}{C}{g}$ and $\jfam{{\Gamma}{A}}{P}$.
-Then we have the judgmental equalities
-\begin{align*}
+Then we have the judgmental equalities\begin{align*}
 \ctxwk{P}{\jcomp{A}{f}{g}} 
@@
 Now let $\jterm{\Gamma}{B}{y}$ and $\jhom{\Gamma}{B}{C}{g}$. Then we have the
-judgmental equalities
-\begin{align*}
+judgmental equalities\begin{align*}
 \jcomp{A}{\ctxwk{A}{y}}{g}
@@
 For the third assertion, let $\jhom{\Gamma}{A}{B}{f}$ and $\jterm{\Gamma}{C}{z}$.
-Then we have the judgmental equalities
-\begin{align*}
+Then we have the judgmental equalities\begin{align*}
 \jcomp{A}{f}{\ctxwk{B}{z}} 
@@
 We have the following inference rules about the compatibility of composition with
-substitution:
-\begin{align*}
+substitution:\begin{align*}
 & \inference
@@
 $\jterm{\Gamma}{A}{x}$.
-Then we have the judgmental equalities
-\begin{align*}
+Then we have the judgmental equalities\begin{align*}
 \subst{x}{\jcomp{A}{f}{g}}
@@
 Now let $\jhom{\Gamma}{A}{B}{f}$, $\jhom{\Gamma}{B}{C}{g}$ and $\jterm{\Gamma}{A}{x}$.
-Then we have the judgmental equalities
-\begin{align*}
+Then we have the judgmental equalities\begin{align*}
 \subst{x}{\jcomp{A}{f}{g}}
@@
 and consider $\jfam{{\Gamma}{A}}{P}$ and $\jfam{{\Gamma}{B}}{Q}$. We define the
-judgment
-\begin{equation*}
+judgment\begin{equation*}
 \jfhom{\Gamma}{A}{B}{f}{P}{Q}{F},
@@
 which is pronounced as `$F$ is a morphism from $P$ to $Q$ over $f$ in context
-$\Gamma$', to be the judgment
-\begin{equation*}
+$\Gamma$', to be the judgment\begin{equation*}
 \unfold{\jfhom{\Gamma}{A}{B}{f}{P}{Q}{F}}.
@@
 \begin{rmk}
-The judgment $\jfhom{\Gamma}{A}{B}{f}{P}{Q}{F}$ means the same thing as
-\begin{equation*}
+The judgment $\jfhom{\Gamma}{A}{B}{f}{P}{Q}{F}$ means the same thing as\begin{equation*}
 \jhom{{\Gamma}{A}}{P}{\jcomp{A}{f}{Q}}{F}.
@@
 and that we have the morphisms $\jfhom{\Gamma}{A}{B}{f}{P}{Q}{F}$ and
-$\jfhom{\Gamma}{B}{C}{g}{Q}{R}{G}$ over them. Then we have
-\begin{equation*}
+$\jfhom{\Gamma}{B}{C}{g}{Q}{R}{G}$ over them. Then we have\begin{equation*}
 \jhom
@@
 Because we also have $\jhom{{\Gamma}{A}}{P}{\jcomp{A}{f}{Q}}{F}$, we have the
-composition
-\begin{equation*}
+composition\begin{equation*}
 \jhom
@@
 be morphisms and let $\jfhom{\Gamma}{A}{B}{f}{P}{Q}{F}$ and
-$\jfhom{\Gamma}{B}{C}{g}{Q}{R}{G}$ be morphisms over them. Then we define
-\begin{equation*}
+$\jfhom{\Gamma}{B}{C}{g}{Q}{R}{G}$ be morphisms over them. Then we define\begin{equation*}
 \jfhomdefn
@@
 
-\subsubsection{identity terms}
+\subsubsection{Identity terms}
 Without a rule explicitly asserting the existence of an identity morphism we don't
@@
 in ordinary type theory via the variable rule. The variable rule is a bit more
-general: it asserts that
-\begin{equation*}
+general: it asserts that\begin{equation*}
 \jterm{\Gamma,\,x_1:A_1,\ldots,\,x_n:A_n}{A_i}{x_i}
@@
 we get the projections from the identity morphisms together with weakening. We
-already have weakening, so here it suffices to introduce the identity morphisms.
-\begin{align}
+already have weakening, so here it suffices to introduce the identity morphisms.\begin{align}
 & \inference
@@
 Identity terms are determined by their behavior with respect to substitution combined with
-weakening. The identity terms will also be subject to compatibility rules.
-\begin{align}
+weakening. The identity terms will also be subject to compatibility rules.\begin{align}
 & \inference
@@
 
-The identity term of a weakened family is the weakened identity term:
-\begin{equation}\label{idfunc-wk-comp}
+The identity term of a weakened family is the weakened identity term:\begin{equation}\label{idfunc-wk-comp}
 \inference
@@
 
-The identity term of a substituted family is the substitution of the identity term
-\begin{equation}\label{idfunc-subst-comp}
+The identity term of a substituted family is the substitution of the identity term\begin{equation}\label{idfunc-subst-comp}
 \inference
@@
 \begin{lem}\label{hom-over-id-is-hom}
-We have the following valid inference rules:
-\begin{align*}
+We have the following valid inference rules:\begin{align*}
 & \inference
@@
 If we unfold the judgments $\jhom{{\Gamma}{A}}{P}{Q}{f}$ and
-$\jfhom{\Gamma}{A}{A}{\idtm{A}}{P}{Q}{f}$, we get the judgments
-\begin{align*}
+$\jfhom{\Gamma}{A}{A}{\idtm{A}}{P}{Q}{f}$, we get the judgments\begin{align*}
 & \unfold{\jhom{{\Gamma}{A}}{P}{Q}{f}}
@@
 new judgments: the judgment that something is a type and the judgment that two
-types are equal.
-\begin{align*}
+types are equal.\begin{align*}
 \jtype*{\Gamma}{A} 
@@
 judgmentally equal to a type $A$ in context $\Gamma$, then $B$ is a type in
-context $\Gamma$. This is expressed by the following four inference rules:
-\begin{align*}
+context $\Gamma$. This is expressed by the following four inference rules:\begin{align*}
 & \inference
@@
 We add rules asserting that a weakened type is again a type that 
-substitution preserves the property of being a type:
-\begin{align}
+substitution preserves the property of being a type:\begin{align}
 & \inference
@@
 unique factorization domains in ring theory. All the tradional models of type
-theory should translate to models with such a condition.
+theory should translate to models with such a condition, simply because
+traditionally contexts are viewed as lists of variable (and type) declarations.

### 2014-04-08 article-models-tt.tex
reorganizing section 4

@@
 
-\subsection{The empty context}
+\subsection{The operations of the theory of contexts, families and terms}
+
+\subsubsection{The empty context}
 In the theory of contexts, families and terms we introduce an empty context and,
@@
 
-\subsubsection{The empty context is compatible with itself}
-The empty context $\emptyc$ may be considered as a family of contexts over the empty
-context. When we do this, we get $\emptyf[\emptyc]$.\begin{equation}
-\inference
-  { }
-  { \jfameq
-      {\emptyc}
-      {\emptyc}
-      {\emptyf[\emptyc]}
-    }
-\end{equation}
-In the future, we shall denote $\emptyf[\Gamma]$ by $\emptyf$. The above rule
-guarantees that this will not cause confusion. Likewise, we shall denote
-$\emptytm[\Gamma]$ by $\emptytm$.
-
-\subsection{Extension}
+\subsubsection{Extension}
 We introduce extension which not only extends a context $\Gamma$ and a family
@@
 
-\subsubsection{Extension is compatible with the empty context}
-The following rule asserts that extension by $\emptyc$ leaves the contexts unchanged.\begin{align}
-& \inference
-  { \jctx{\Gamma}
-    }
-  { \jctxeq{\ctxext{\emptyc}{\Gamma}}{\Gamma}
-    }
-  \\
-& \inference
-  { \jctx{\Gamma}
-    }
-  { \jctxeq{\ctxext{\Gamma}{\emptyf}}{\Gamma}
-    }
-  \\
-& \inference
-  { \jfam{\Gamma}{A}
-    }
-  { \jfameq{\Gamma}{\ctxext{\emptyf}{A}}{A}
-    }
-  \\
-& \inference
-  { \jfam{\Gamma}{A}
-    }
-  { \jfameq{\Gamma}{{A}{\emptyf}}{A}
-    }
-\end{align}
-
-\subsubsection{Extension is compatible with itself}\label{comp-ee}
-The inference rules asserting that extension is compatible with itself assert
-that contexts are unstructured lists of type declarations. This rule is
-unavoidable if we want that for a family $A$ in context $\Gamma$, a family over
-$A$ is the same thing as a family over $\ctxext{\Gamma}{A}$. 
-\begin{align}
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{{\Gamma}{A}}{P}
-    }
-  { \jctxeq{\ctxext{{\Gamma}{A}}{P}}{\ctxext{\Gamma}{{A}{P}}}
-    }
-  \\
-& \inference
-  { \jfam{{\Gamma}{A}}{P}
-    \jfam{{{\Gamma}{A}}{P}}{Q}
-    }
-  { \jfameq{\Gamma}{\ctxext{{A}{P}}{Q}}{\ctxext{A}{{P}{Q}}}
-    }
-\end{align}
-
-\subsection{Weakening}
+\subsubsection{Weakening}
 When $A$ is a context family over a context $\Gamma$, we wish to define a weakening
@@
 
-\subsubsection{Weakening is compatible with the empty context}
-The following rules express that when the empty context or context family is
-weakened, the result is the empty context family.\begin{align}
-& \inference
-  { \jctx{\Gamma}
-    }
-  { \jfameq{\Gamma}{\ctxwk{\Gamma}{\emptyc}}{\emptyf}
-    }
-  \\
-& \inference
-  { \jfam{\Gamma}{A}
-    }
-  { \jfameq{{\Gamma}{A}}{\ctxwk{A}{\emptyf}}{\emptyf}
-    }
-\end{align}
-Weakening by the empty family $\emptyf$ over a context $\Gamma$ leaves families, 
-their terms, families over those families and
-terms of those unchanged:\begin{align}
-& \inference
-  { \jfam{\Gamma}{B}
-    }
-  { \jfameq{\Gamma}{\ctxwk{\emptyf}{B}}{B}
-    }
-  \\
-& \inference
-  { \jterm{\Gamma}{B}{y}
-    }
-  { \jtermeq{\Gamma}{B}{\ctxwk{\emptyf}{y}}{y}
-    }
-  \\
-& \inference
-  { \jfam{{\Gamma}{B}}{Q}
-    }
-  { \jfameq{{\Gamma}{B}}{\ctxwk{\emptyf}{Q}}{Q}
-    }
-  \\
-& \inference
-  { \jterm{{\Gamma}{B}}{Q}{g}
-    }
-  { \jtermeq{{\Gamma}{B}}{Q}{\ctxwk{\emptyf}{g}}{g}
-    }
-\end{align}
-
-\subsubsection{Weakening is compatible with extension}\label{comp-we}
-
-The following rules assert the compatibility of extension with weakening: for
-every family $A$ over $\Gamma$ and every family $Q$ over $\ctxext{\Gamma}{B}$
-there is a
-judgmental equality $\ctxwk{A}{\ctxext{B}{Q}}\jdeq\ctxext{\ctxwk{A}{B}}
-{\ctxwk{A}{Q}}$.
-
-When thinking of terms of $\ctxwk{A}{B}$ as morphisms of families from $A$ to
-$B$, this looks already like form of type theoretic choice. It is weaker in that
-it is not stated with function types, yet it is stronger in that it states a
-judgmental equality between two families. When one makes the weakening operation
-notationally invisible -- as is in fact the usual practice in type theory -- the
-following compatibility rules become completely obvious.
-
-In the following inference rules we assume that $\jfam{\Gamma}{A}$.\begin{align}
-& \inference
-  { \jfam{{{\Gamma}{B}}{Q}}{R}
-    }
-  { \jfameq
-      {\ctxext{{\Gamma}{A}}{\ctxwk{A}{B}}}
-      {\ctxwk{A}{\ctxext{Q}{R}}}
-      {\ctxext{\ctxwk{A}{Q}}{\ctxwk{A}{R}}}
-    }
-\end{align}
-
-\subsubsection{Weakening is compatible with itself}\label{comp-ww}
-We state judgmental equality rules expressing
-that weakening is compatible with itself. These rules state that the following
-diagram commutes given any two families $A$ and $B$ in context $\Gamma$:\begin{equation*}
-\begin{tikzcd}[column sep=huge]
-\jfam{\Gamma}{\blank} 
-  \ar{r}{C\mapsto\ctxwk{B}{C}} 
-  \ar{d}[swap]{C\mapsto\ctxwk{A}{C}} 
-& \jfam{{\Gamma}{B}}{\blank} 
-  \ar{d}{Q\mapsto\ctxwk{A}{Q}}
-  \\
-\jfam{{\Gamma}{A}}{\blank} 
-  \ar{r}[swap]{P\mapsto\ctxwk{{A}{B}}{P}} 
-& \jfam{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\blank}
-\end{tikzcd}
-\end{equation*}
-Thus, we get the following set of inference rules:\begin{align}
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{\Gamma}{B}
-    \jfam{{\Gamma}{C}}{R}
-    }
-  { \jfameq
-      {{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{{A}{B}}{{A}{C}}}}
-      {\ctxwk{A}{{B}{R}}}
-      {\ctxwk{{A}{B}}{{A}{R}}}
-    }
-  \label{comp-ww-f}\\
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{\Gamma}{B}
-    \jterm{{\Gamma}{C}}{R}{t}
-    }
-  { \jtermeq
-      {{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{{A}{B}}{{A}{C}}}}
-      {\ctxwk{{A}{B}}{{A}{R}}}
-      {\ctxwk{A}{{B}{t}}}
-      {\ctxwk{{A}{B}}{{A}{t}}}
-    }
-  \label{comp-ww-t}
-\end{align}
-
-\subsubsection{Extension is compatible with weakening}\label{comp-ew}
-The rules expressing that extension is compatible with weakening assert that
-weakening by an extension is the same thing as weakening twice in the
-appropriate way.
-
-In the following inference rules we assume that
-$\jfam{\Gamma}{A}$ and $\jfam{{\Gamma}{A}}{P}$. \begin{align}
-& \inference
-  { \jfam{{\Gamma}{B}}{Q}
-    }
-  { \jfameq
-      {{{{\Gamma}{A}}{P}}{\ctxwk{P}{{A}{B}}}}
-      {\ctxwk{\ctxext{A}{P}}{Q}}
-      {\ctxwk{P}{{A}{Q}}}
-    }
-  \label{comp-ew-f}\\
-& \inference
-  { \jterm{{\Gamma}{B}}{Q}{g}
-    }
-  { \jtermeq
-      {{{{\Gamma}{A}}{P}}{\ctxwk{P}{{A}{B}}}}
-      {\ctxwk{P}{{A}{Q}}}
-      {\ctxwk{\ctxext{A}{P}}{g}}
-      {\ctxwk{P}{{A}{g}}}
-    } 
-  \label{comp-ew-t}
-\end{align}
-
-\subsection{Substitution}
+\subsubsection{Substitution}
 Given a family $P$ over $A$ and a term $x$ of $A$, substitution gives a way to
@@
 \end{align}
-
-\subsubsection{Substitution is compatible with the empty context}
-The fibers of the empty family are the empty families:\begin{align}
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    }
-  { \jfameq{\Gamma}{\subst{x}{\emptyf}}{\emptyf}
-    }
-  \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jfam{{\Gamma}{A}}{P}
-    }
-  { \jfameq
-      {{\Gamma}{\subst{x}{P}}}
-      {\subst{x}{\emptyf}}
-      {\emptyf}
-    }
-\end{align}
-
-The following rules assert that substituting by the term $\jterm{\Gamma}{\emptyf}{\emptytm}$
-leaves everything unchanged.\begin{align}
-& \inference
-  { \jfam{\Gamma}{A}
-    }
-  { \jfameq{\Gamma}{\subst{\emptytm}{A}}{A}
-    }
-  \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    }
-  { \jtermeq{\Gamma}{A}{\subst{\emptytm}{x}}{x}
-    }
-  \\
-& \inference
-  { \jfam{{\Gamma}{A}}{P}
-    }
-  { \jfameq{{\Gamma}{A}}{\subst{\emptytm}{P}}{P}
-    }
-  \\
-& \inference
-  { \jterm{{\Gamma}{A}}{P}{f}
-    }
-  { \jtermeq{{\Gamma}{A}}{P}{\subst{\emptytm}{f}}{f}
-    }.
-\end{align}
-
-\subsubsection{Substitution is compatible with extension}\label{comp-se}
-Suppose $\jterm{\Gamma}{A}{x}$ in all of the following inference rule.\begin{align}
-& \inference
-  { \jfam{{{{\Gamma}{A}}{P}}{Q}}{R}
-    }
-  { \jfameq
-      {{\Gamma}{\subst{x}{P}}}
-      {\subst{x}{\ctxext{Q}{R}}}
-      {\ctxext{\subst{x}{Q}}{\subst{x}{R}}}
-    }
-\end{align}
-
-\subsubsection{Substitution is compatible with weakening}\label{comp-sw}
-The rules asserting the compatibility of substitution with weakening assert
-that the following diagram commutes for any $\jterm{\Gamma}{A}{x}$ and any
-$\jfam{{\Gamma}{A}}{P}$:\begin{equation*}
-\begin{tikzcd}[column sep=huge]
-\jfam{{\Gamma}{A}}{\blank} 
-  \ar{d}[swap]{Q\mapsto\subst{x}{Q}} 
-  \ar{r}{Q\mapsto\ctxwk{P}{Q}} 
-& \jfam{{{\Gamma}{A}}{P}}{\blank} 
-    \ar{d}{R\mapsto\subst{x}{R}}
-  \\ 
-\jfam{\Gamma}{\blank} 
-  \ar{r}[swap]{B\mapsto\ctxwk{\subst{x}{P}}{B}} 
-& \jfam{{\Gamma}{\subst{x}{P}}}{\blank}
-\end{tikzcd}
-\end{equation*}
-We plug in an extra layer of families to cover the most general case at once.
-In the following inference rules we assume that $\jterm{\Gamma}{A}{x}$ and
-$\jfam{{\Gamma}{A}}{P}$:\begin{align}
-& \inference
-  { \jfam{{{\Gamma}{A}}{Q}}{R}
-    }
-  { \jfameq
-      {{{\Gamma}{\subst{x}{P}}}{\subst{x}{\ctxwk{P}{Q}}}}
-      {\subst{x}{\ctxwk{P}{R}}}
-      {\ctxwk{\subst{x}{P}}{\subst{x}{R}}}
-    }
-  \label{comp-sw-f}\\
-& \inference
-  { \jterm{{{\Gamma}{A}}{Q}}{R}{h}
-    }
-  { \jtermeq
-      {\ctxext{{\Gamma}{\subst{x}{P}}}{\subst{x}{\ctxwk{P}{Q}}}}
-      {\subst{x}{\ctxwk{P}{R}}}
-      {\subst{x}{\ctxwk{P}{h}}}
-      {\ctxwk{\subst{x}{P}}{\subst{x}{h}}}
-    }
-  \label{comp-sw-t}
-\end{align}
-
-\subsubsection{Substitution is compatible with substitution}\label{comp-ss}
-
-We require that substitution is compatible with itself, which is roughly the
-assertion that substitution is associative. However, we cannot just state that
-$\subst{x}{{f}{g}}\jdeq\subst{{x}{f}}{g}$ since the expression $\subst{{x}{f}}{g}$
-is not well-formed. The term $\subst{x}{f}$ can be substituted in (terms of) families over
-$\subst{x}{P}$; the term $\subst{x}{g}$ is such. Therefore, associativity of
-substitution takes the form $\subst{x}{{f}{g}}\jdeq\subst{{x}{f}}{{x}{g}}$.
-Note that the term $\subst{{x}{f}}{{x}{g}}$ may be written down more conveniently
-as $\subst{x,\subst{x}{f}}{g}$, although we will not do that here.
-
-In the following inference rules we assume
-$\jterm{\Gamma}{A}{x}$ and $\jterm{{\Gamma}{A}}{P}{f}$.
-\begin{align}
-& \inference
-  { \jfam{{{{\Gamma}{A}}{P}}{Q}}{R}
-    }
-  { \jfameq
-      {{\Gamma}{\subst{x}{{f}{Q}}}}
-      {\subst{x}{{f}{R}}}
-      {\subst{{x}{f}}{{x}{R}}}
-    }
-  \label{comp-ss-f}\\
-& \inference
-  { \jterm{{{{\Gamma}{A}}{P}}{Q}}{R}{h}
-    }
-  { \jtermeq
-      {{\Gamma}{\subst{x}{{f}{Q}}}}
-      {\subst{x}{{f}{R}}}
-      {\subst{x}{{f}{h}}}
-      {\subst{{x}{f}}{{x}{h}}}
-    }
-  \label{comp-ss-t}
-\end{align}
-
-\subsubsection{Weakening is compatible with substitution}\label{comp-ws}
-We already have rules for the compatibility of substitution with weakening, but
-we still need the rules the other way around, asserting that there is a 
-judgmental equality $\ctxwk{A}{\subst{y}{Q}}\jdeq\subst{\ctxwk{A}{y}}{\ctxwk{A}{Q}}$
-together with all its variants.
-
-In the following inference rules we assume that $\jfam{\Gamma}{A}$ and that
-$\jterm{\Gamma}{B}{y}$.
-\begin{align}
-& \inference
-  { \jfam{{{\Gamma}{B}}{Q}}{R}
-    }
-  { \jfameq
-      {{{\Gamma}{A}}{\ctxwk{A}{\subst{y}{Q}}}}
-      {\ctxwk{A}{\subst{y}{R}}}
-      {\subst{\ctxwk{A}{y}}{\ctxwk{A}{R}}}
-    }
-  \label{comp-ws-f}\\
-& \inference
-  { \jterm{{{\Gamma}{B}}{Q}}{R}{h}
-    }
-  { \jtermeq
-      {{{\Gamma}{A}}{\ctxwk{A}{\subst{y}{Q}}}}
-      {\ctxwk{A}{\subst{y}{R}}}
-      {\ctxwk{A}{\subst{y}{h}}}
-      {\subst{\ctxwk{A}{y}}{\ctxwk{A}{h}}}
-    }
-  \label{comp-ws-t}
-\end{align}
-
-\subsection{Composition and identity terms}\label{categorical_properties}
-\subsubsection{The defining property of weakening}
-The judgmental equalities we're about to describe assert that substituting a term
-in the weakening a thing gives you the thing back. In the case of contexts we get that each fiber
-$\subst{x}{\ctxwk{A}{B}}$ is just $B$ and in the case of terms we get 
-that $\ctxwk{A}{y}$ is the constant function
-mapping everything to $y:B$. Thus, these rules actually establish the weakening
-as the weakening. After stating the rules we will describe what it means to
-compose context morphisms (terms of weakened contexts).
-\begin{align}
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{{\Gamma}{B}}{Q}
-    \jterm{\Gamma}{A}{x}
-    }
-  { \jfameq{{\Gamma}{B}}{\subst{x}{\ctxwk{A}{Q}}}{Q}
-    }
-  \label{defn-ws-3}\\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jterm{{\Gamma}{B}}{Q}{g}
-    }
-  { \jtermeq{{\Gamma}{B}}{Q}{\subst{x}{\ctxwk{A}{g}}}{g}
-    }
-  \label{defn-ws-4}
-\end{align}
-
-Using the rules of the compatibility of substitution with weakening and of the
-compatibility of weakening with itself, we see that we can show
-
-\begin{lem}
-The inference rule\begin{equation*}
-\inference
-  { \jfam{\Gamma}{A}
-    \jfam{\Gamma}{B}
-    \jfam{\Gamma}{C}
-    \jhom{\Gamma}{A}{B}{f}
-    }
-  { \jfameq
-    {{\Gamma}{A}}
-    {\subst{f}{\ctxwk{A}{{B}{C}}}}
-    {\ctxwk{A}{C}}
-    }
-\end{equation*}
-is valid.
-\end{lem}
-
-\begin{proof}
-Let $\jfam{\Gamma}{A}$, $\jfam{\Gamma}{B}$, $\jfam{\Gamma}{C}$ and $\jhom{\Gamma}{A}{B}{f}$.
-Then we have the judgmental equalities\begin{align*}
-\subst{f}{\ctxwk{A}{{B}{C}}}
-& \jdeq 
-  \subst{f}{\ctxwk{{A}{B}}{{A}{C}}}
-  \\
-& \jdeq 
-  \ctxwk{A}{C}.
-  \qedhere
-\end{align*}
-\end{proof}
-
-It follows that for $\jterm{{\Gamma}{B}}{\ctxwk{B}{C}}{g}$ we can compose $f$
-with $g$ to obtain a term of $\ctxwk{A}{C}$ in context $\ctxext{\Gamma}{A}$.
-In the following definition, we work with in a slightly greater generality.
-
-\begin{defn}
-We define the judgment\begin{equation*}
-\jhom{\Gamma}{A}{B}{f},
-\end{equation*}
-which is pronounced as `$f$ is a morphism from $A$ to $B$ in context $\Gamma$',
-to be the judgment\begin{equation*}
-\unfold{\jhom{\Gamma}{A}{B}{f}}.
-\end{equation*}
-Likewise, we define the judgment\begin{equation*}
-\jhomeq{\Gamma}{A}{B}{f}{f'}
-\end{equation*}
-to be the judgment\begin{equation*}
-\unfold{\jhomeq{\Gamma}{A}{B}{f}{f'}}.
-\end{equation*}
-\end{defn}
-
-\begin{defn}
-Let $\jhom{\Gamma}{A}{B}{f}$ and consider a term $\jterm{{\Gamma}{B}}{Q}{g}$.
-We define\begin{align*}
-\jfamdefn*
-  {{\Gamma}{A}}
-  {\jcomp{A}{f}{Q}}
-  {\unfold{\jcomp{A}{f}{Q}}}\\
-\jtermdefn*
-  {{\Gamma}{A}}
-  {\jcomp{A}{f}{Q}}
-  {\jcomp{A}{f}{g}}
-  {\unfold{\jcomp{A}{f}{g}}}.
-\end{align*}
-Likewise, when we have a family $\jfam{{{\Gamma}{B}}{Q}}{R}$ and a term
-$\jterm{{{\Gamma}{B}}{Q}}{R}{h}$, we define\begin{align*}
-\jfamdefn*
-  {{{\Gamma}{A}}{\jcomp{A}{f}{Q}}}
-  {\jcomp{A}{f}{R}}
-  {\unfold{\jcomp{A}{f}{R}}}
-  \\
-\jtermdefn*
-  {{{\Gamma}{A}}{\jcomp{A}{f}{Q}}}
-  {\jcomp{A}{f}{R}}
-  {\jcomp{A}{f}{h}}
-  {\unfold{\jcomp{A}{f}{h}}}.
-\end{align*}
-\end{defn}
-
-We have lots of compatibility properties for composition:
-
-\begin{lem}
-We have the following inference rules about the situation where something is
-substituted by a composition:\begin{align*}
-& \inference
-  { \jhom{\Gamma}{A}{B}{f}
-    \jhom{\Gamma}{B}{C}{g}
-    \jfam{{{\Gamma}{A}}{\ctxwk{A}{C}}}{R}
-    }
-  { \jfameq
-      {{\Gamma}{A}}
-      {\subst{\jcomp{A}{f}{g}}{R}}
-      {\subst{f}{{\ctxwk{A}{g}}{\ctxwk{{A}{B}}{R}}}}
-    }
-  \\
-& \inference
-  { \jhom{\Gamma}{A}{B}{f}
-    \jhom{\Gamma}{B}{C}{g}
-    \jterm{{{\Gamma}{A}}{\ctxwk{A}{C}}}{R}{h}
-    }
-  { \jfameq
-    {{\Gamma}{A}}
-    {\subst{\jcomp{A}{f}{g}}{h}}
-    {\subst{f}{{\ctxwk{A}{g}}{\ctxwk{{A}{B}}{h}}}}
-    }
-\end{align*}
-We also have the following related inference rules, asserting that composition
-is strictly associative:\begin{align*}
-& \inference
-  { \jhom{\Gamma}{A}{B}{f}
-    \jhom{\Gamma}{B}{C}{g}
-    \jfam{{\Gamma}{C}}{R}
-    }
-  { \jfameq
-      {{\Gamma}{A}}
-      {\jcomp{A}{{A}{f}{g}}{R}}
-      {\jcomp{A}{f}{{B}{g}{R}}}
-    }
-  \\
-& \inference
-    { \jhom{\Gamma}{A}{B}{f}
-      \jhom{\Gamma}{B}{C}{g}
-      \jterm{{\Gamma}{C}}{R}{h}
-      }
-    { \jtermeq
-        {{\Gamma}{A}}
-        {\jcomp{A}{{A}{f}{g}}{R}}
-        {\jcomp{A}{{A}{f}{g}}{h}}
-        {\jcomp{A}{f}{{B}{g}{h}}}
-      }
-\end{align*}
-\end{lem}
-
-\begin{proof}
-Consider family morphisms $\jhom{\Gamma}{A}{B}{f}$ and $\jhom{\Gamma}{B}{C}{g}$
-and a family $\jfam{{{\Gamma}{A}}{\ctxwk{A}{C}}}{R}$. Then we have the judgmental
-equalities\begin{align*}
-\subst{\jcomp{A}{f}{g}}{R} 
-& \jdeq 
-  \subst{{f}{\ctxwk{A}{g}}}{R}
-  \\
-& \jdeq 
-  \subst{{f}{\ctxwk{A}{g}}}{\subst{f}{\ctxwk{{A}{B}}{R}}}
-  \\
-& \jdeq 
-  \subst{f}{{\ctxwk{A}{g}}{\ctxwk{{A}{B}}{R}}}
-\end{align*}
-The proof that 
-$\subst{\jcomp{A}{f}{g}}{h}\jdeq\subst{f}{{\ctxwk{A}{g}}{\ctxwk{{A}{B}}{h}}}$
-is similar.
-
-Now suppose that $\jfam{{\Gamma}{C}}{R}$ instead. Then we have\begin{align*}
-\jcomp{A}{{A}{f}{g}}{R} 
-& \jdeq 
-  \subst{\jcomp{A}{f}{g}}{\ctxwk{A}{R}}
-  \\
-& \jdeq 
-  \subst{{f}{\ctxwk{A}{g}}}{\ctxwk{A}{R}}
-  \\
-& \jdeq 
-  \subst{f}{{\ctxwk{A}{g}}{\ctxwk{{A}{B}}{{A}{R}}}}
-  \\
-& \jdeq 
-  \subst{f}{{\ctxwk{A}{g}}{\ctxwk{A}{{B}{R}}}}
-  \\
-& \jdeq 
-  \subst{f}{\ctxwk{A}{\subst{g}{\ctxwk{B}{R}}}}
-  \\
-& \jdeq 
-  \subst{f}{\ctxwk{A}{\jcomp{B}{g}{R}}}
-  \\
-& \jdeq 
-  \jcomp{A}{f}{{B}{g}{R}}.
-\end{align*}
-Again, the proof is similar for terms $h$ of $R$ in context $\ctxext{\Gamma}{C}$.
-\end{proof}
-
-\begin{lem}
-We have the following inference rules about the compatibility of composition with
-weakening:\begin{align*}
-& \inference
-  { \jhom{\Gamma}{A}{B}{f}
-    \jhom{\Gamma}{B}{C}{g}
-    \jfam{{\Gamma}{A}}{P}
-    }
-  { \jhomeq
-      {\Gamma}
-      {{A}{P}}
-      {C}
-      {\ctxwk{P}{\jcomp{A}{f}{g}}}
-      {\jcomp{{A}{P}}{\ctxwk{P}{f}}{g}}
-    }
-  \\
-& \inference
-  { \jterm{\Gamma}{B}{y}
-    \jhom{\Gamma}{B}{C}{g}
-    }
-  { \jhomeq
-      {\Gamma}
-      {A}
-      {C}
-      {\jcomp{A}{\ctxwk{A}{y}}{g}}
-      {\ctxwk{A}{\subst{y}{g}}}
-    }
-  \\
-& \inference
-  { \jhom{\Gamma}{A}{B}{f}
-    \jterm{\Gamma}{C}{z}
-    }
-  { \jhomeq
-      {\Gamma}
-      {A}
-      {C}
-      {\jcomp{A}{f}{\ctxwk{B}{z}}}
-      {\ctxwk{A}{z}}
-    }
-\end{align*}
-\end{lem}
-
-\begin{proof}
-Let $\jhom{\Gamma}{A}{B}{f}$, $\jhom{\Gamma}{B}{C}{g}$ and $\jfam{{\Gamma}{A}}{P}$.
-Then we have the judgmental equalities\begin{align*}
-\ctxwk{P}{\jcomp{A}{f}{g}} 
-& \jdeq 
-  \ctxwk{P}{\subst{f}{\ctxwk{A}{g}}}
-  \\
-& \jdeq 
-  \subst{\ctxwk{P}{f}}{\ctxwk{P}{{A}{g}}}
-  \\
-& \jdeq 
-  \subst{\ctxwk{P}{f}}{\ctxwk{\ctxext{A}{P}}{g}}
-  \\
-& \jdeq 
-  \jcomp{{A}{P}}{\ctxwk{P}{f}}{g}.
-\end{align*}
-Now let $\jterm{\Gamma}{B}{y}$ and $\jhom{\Gamma}{B}{C}{g}$. Then we have the
-judgmental equalities\begin{align*}
-\jcomp{A}{\ctxwk{A}{y}}{g}
-& \jdeq 
-  \subst{\ctxwk{A}{y}}{\ctxwk{A}{g}}
-  \\
-& \jdeq 
-  \ctxwk{A}{\subst{y}{g}}.
-\end{align*}
-For the third assertion, let $\jhom{\Gamma}{A}{B}{f}$ and $\jterm{\Gamma}{C}{z}$.
-Then we have the judgmental equalities\begin{align*}
-\jcomp{A}{f}{\ctxwk{B}{z}} 
-& \jdeq 
-  \subst{f}{\ctxwk{A}{{B}{z}}}
-  \\
-& \jdeq 
-  \subst{f}{\ctxwk{{A}{B}}{{A}{z}}}
-  \\
-& \jdeq 
-  \ctxwk{A}{z}.
-  \qedhere
-\end{align*}
-\end{proof}
-
-\begin{lem}
-We have the following inference rules about the compatibility of composition with
-substitution:\begin{align*}
-& \inference
-  { \jhom{{\Gamma}{A}}{P}{Q}{f}
-    \jhom{{\Gamma}{A}}{Q}{R}{g}
-    \jterm{\Gamma}{A}{x}
-    }
-  { \jhomeq
-      {\Gamma}
-      {\subst{x}{P}}
-      {\subst{x}{R}}
-      {\subst{x}{\jcomp{P}{f}{g}}}
-      {\jcomp{\subst{x}{P}}{\subst{x}{f}}{\subst{x}{g}}}
-    }
-  \\
-& \inference
-  { \jhom{\Gamma}{A}{B}{f}
-    \jhom{\Gamma}{B}{C}{g}
-    \jterm{\Gamma}{A}{x}
-    }
-  { \jtermeq
-      {\Gamma}
-      {C}
-      {\subst{x}{\jcomp{A}{f}{g}}}
-      {\subst{{x}{f}}{g}}
-    }
-\end{align*}
-\end{lem}
-
-\begin{proof}
-Let $\jhom{{\Gamma}{A}}{P}{Q}{f}$, $\jhom{{\Gamma}{A}}{Q}{R}{g}$ and 
-$\jterm{\Gamma}{A}{x}$.
-Then we have the judgmental equalities\begin{align*}
-\subst{x}{\jcomp{A}{f}{g}}
-& \jdeq 
-  \subst{x}{{f}{\ctxwk{P}{g}}}
-  \\
-& \jdeq 
-  \subst{{x}{f}}{{x}{\ctxwk{P}{g}}}
-  \\
-& \jdeq 
-  \subst{{x}{f}}{\ctxwk{\subst{x}{P}}{\subst{x}{g}}}
-  \\
-& \jdeq 
-  \jcomp{\subst{x}{P}}{\subst{x}{f}}{\subst{x}{g}}.
-\end{align*}
-Now let $\jhom{\Gamma}{A}{B}{f}$, $\jhom{\Gamma}{B}{C}{g}$ and $\jterm{\Gamma}{A}{x}$.
-Then we have the judgmental equalities\begin{align*}
-\subst{x}{\jcomp{A}{f}{g}}
-& \jdeq 
-  \subst{x}{{f}{\ctxwk{A}{g}}}
-  \\
-& \jdeq 
-  \subst{{x}{f}}{{x}{\ctxwk{A}{g}}}
-  \\
-& \jdeq 
-  \subst{{x}{f}}{g}.
-  \qedhere
-\end{align*}
-\end{proof}
-
-There is also a notion of morphism \emph{over} a morphism. We will develop this
-notion because it will be needed in the theory of models later on.
-
-\begin{defn}
-Let $\jhom{\Gamma}{A}{B}{f}$ be a morphism from $A$ to $B$ in context $\Gamma$
-and consider $\jfam{{\Gamma}{A}}{P}$ and $\jfam{{\Gamma}{B}}{Q}$. We define the
-judgment\begin{equation*}
-\jfhom{\Gamma}{A}{B}{f}{P}{Q}{F},
-\end{equation*}
-which is pronounced as `$F$ is a morphism from $P$ to $Q$ over $f$ in context
-$\Gamma$', to be the judgment\begin{equation*}
-\unfold{\jfhom{\Gamma}{A}{B}{f}{P}{Q}{F}}.
-\end{equation*}
-\end{defn}
-
-\begin{rmk}
-The judgment $\jfhom{\Gamma}{A}{B}{f}{P}{Q}{F}$ means the same thing as\begin{equation*}
-\jhom{{\Gamma}{A}}{P}{\jcomp{A}{f}{Q}}{F}.
-\end{equation*}
-\end{rmk}
-
-Suppose we have morphisms $\jhom{\Gamma}{A}{B}{f}$ and $\jhom{\Gamma}{B}{C}{g}$
-and that we have the morphisms $\jfhom{\Gamma}{A}{B}{f}{P}{Q}{F}$ and
-$\jfhom{\Gamma}{B}{C}{g}{Q}{R}{G}$ over them. Then we have\begin{equation*}
-\jhom
-  {{\Gamma}{A}}
-  {\jcomp{A}{f}{Q}}
-  {\jcomp{A}{f}{{B}{g}{R}}}
-  {\unfold{\jcomp{A}{f}{G}}}
-\end{equation*}
-Because we also have $\jhom{{\Gamma}{A}}{P}{\jcomp{A}{f}{Q}}{F}$, we have the
-composition\begin{equation*}
-\jhom
-  {{\Gamma}{A}}
-  {P}
-  {\jcomp{A}{f}{{B}{g}{R}}}
-  {\jcomp{P}{F}{\unfold{\jcomp{A}{f}{G}}}}.
-\end{equation*}
-Because of 
-the judgmental equality $\jcomp{A}{f}{{B}{g}{R}}\jdeq
-\jcomp{A}{{A}{f}{g}}{R}$, it follows that 
-$\jcomp{P}{F}{\unfold{\jcomp{A}{f}{G}}}$ is a morphism from $P$ to $R$ over
-$\jcomp{A}{f}{g}$. We make the following definition:
-
-\begin{defn}
-Let $\jhom{\Gamma}{A}{B}{f}$ and $\jhom{\Gamma}{B}{C}{g}$
-be morphisms and let $\jfhom{\Gamma}{A}{B}{f}{P}{Q}{F}$ and
-$\jfhom{\Gamma}{B}{C}{g}{Q}{R}{G}$ be morphisms over them. Then we define\begin{equation*}
-\jfhomdefn
-  {\Gamma}
-  {A}
-  {C}
-  {\jcomp{A}{f}{g}}
-  {P}
-  {R}
-  {\jfcomp{A}{f}{P}{F}{G}}
-  {\unfold{\jfcomp{A}{f}{P}{F}{G}}}.
-\end{equation*}
-\end{defn}
-
-This composition is also judgmentally associative.
 
@@
     }
-  { \jhom{\Gamma}{A}{A}{\idtm{A}}
+  { \unfoldall{\jhom{\Gamma}{A}{A}{\idtm{A}}}
     }
@@
       }
-    { \jhomeq{\Gamma}{A}{A}{\idtm{A}}{\idtm{A'}}
-      }
-\end{align}
+    { \unfoldall{\jhomeq{\Gamma}{A}{A}{\idtm{A}}{\idtm{A'}}}
+      }
+\end{align}
+
+\subsection{The judgmental equalities of the theory of contexts, families and
+terms}
+
+\subsubsection{The empty context is compatible with itself}
+The empty context $\emptyc$ may be considered as a family of contexts over the empty
+context. When we do this, we get $\emptyf[\emptyc]$.\begin{equation}
+\inference
+  { }
+  { \jfameq
+      {\emptyc}
+      {\emptyc}
+      {\emptyf[\emptyc]}
+    }
+\end{equation}
+In the future, we shall denote $\emptyf[\Gamma]$ by $\emptyf$. The above rule
+guarantees that this will not cause confusion. Likewise, we shall denote
+$\emptytm[\Gamma]$ by $\emptytm$.
+
+\subsubsection{Extension is compatible with the empty context}
+The following rule asserts that extension by $\emptyc$ leaves the contexts unchanged.\begin{align}
+& \inference
+  { \jctx{\Gamma}
+    }
+  { \jctxeq{\ctxext{\emptyc}{\Gamma}}{\Gamma}
+    }
+  \\
+& \inference
+  { \jctx{\Gamma}
+    }
+  { \jctxeq{\ctxext{\Gamma}{\emptyf}}{\Gamma}
+    }
+  \\
+& \inference
+  { \jfam{\Gamma}{A}
+    }
+  { \jfameq{\Gamma}{\ctxext{\emptyf}{A}}{A}
+    }
+  \\
+& \inference
+  { \jfam{\Gamma}{A}
+    }
+  { \jfameq{\Gamma}{{A}{\emptyf}}{A}
+    }
+\end{align}
+
+\subsubsection{Extension is compatible with itself}\label{comp-ee}
+The inference rules asserting that extension is compatible with itself assert
+that contexts are unstructured lists of type declarations. This rule is
+unavoidable if we want that for a family $A$ in context $\Gamma$, a family over
+$A$ is the same thing as a family over $\ctxext{\Gamma}{A}$. 
+\begin{align}
+& \inference
+  { \jfam{\Gamma}{A}
+    \jfam{{\Gamma}{A}}{P}
+    }
+  { \jctxeq{\ctxext{{\Gamma}{A}}{P}}{\ctxext{\Gamma}{{A}{P}}}
+    }
+  \\
+& \inference
+  { \jfam{{\Gamma}{A}}{P}
+    \jfam{{{\Gamma}{A}}{P}}{Q}
+    }
+  { \jfameq{\Gamma}{\ctxext{{A}{P}}{Q}}{\ctxext{A}{{P}{Q}}}
+    }
+\end{align}
+
+\subsubsection{Weakening is compatible with the empty context}
+The following rules express that when the empty context or context family is
+weakened, the result is the empty context family.\begin{align}
+& \inference
+  { \jctx{\Gamma}
+    }
+  { \jfameq{\Gamma}{\ctxwk{\Gamma}{\emptyc}}{\emptyf}
+    }
+  \\
+& \inference
+  { \jfam{\Gamma}{A}
+    }
+  { \jfameq{{\Gamma}{A}}{\ctxwk{A}{\emptyf}}{\emptyf}
+    }
+\end{align}
+Weakening by the empty family $\emptyf$ over a context $\Gamma$ leaves families, 
+their terms, families over those families and
+terms of those unchanged:\begin{align}
+& \inference
+  { \jfam{\Gamma}{B}
+    }
+  { \jfameq{\Gamma}{\ctxwk{\emptyf}{B}}{B}
+    }
+  \\
+& \inference
+  { \jterm{\Gamma}{B}{y}
+    }
+  { \jtermeq{\Gamma}{B}{\ctxwk{\emptyf}{y}}{y}
+    }
+  \\
+& \inference
+  { \jfam{{\Gamma}{B}}{Q}
+    }
+  { \jfameq{{\Gamma}{B}}{\ctxwk{\emptyf}{Q}}{Q}
+    }
+  \\
+& \inference
+  { \jterm{{\Gamma}{B}}{Q}{g}
+    }
+  { \jtermeq{{\Gamma}{B}}{Q}{\ctxwk{\emptyf}{g}}{g}
+    }
+\end{align}
+
+\subsubsection{Weakening is compatible with extension}\label{comp-we}
+
+The following rules assert the compatibility of extension with weakening: for
+every family $A$ over $\Gamma$ and every family $Q$ over $\ctxext{\Gamma}{B}$
+there is a
+judgmental equality $\ctxwk{A}{\ctxext{B}{Q}}\jdeq\ctxext{\ctxwk{A}{B}}
+{\ctxwk{A}{Q}}$.
+
+When thinking of terms of $\ctxwk{A}{B}$ as morphisms of families from $A$ to
+$B$, this looks already like form of type theoretic choice. It is weaker in that
+it is not stated with function types, yet it is stronger in that it states a
+judgmental equality between two families. When one makes the weakening operation
+notationally invisible -- as is in fact the usual practice in type theory -- the
+following compatibility rules become completely obvious.
+
+In the following inference rules we assume that $\jfam{\Gamma}{A}$.\begin{align}
+& \inference
+  { \jfam{{{\Gamma}{B}}{Q}}{R}
+    }
+  { \jfameq
+      {\ctxext{{\Gamma}{A}}{\ctxwk{A}{B}}}
+      {\ctxwk{A}{\ctxext{Q}{R}}}
+      {\ctxext{\ctxwk{A}{Q}}{\ctxwk{A}{R}}}
+    }
+\end{align}
+
+\subsubsection{Weakening is compatible with itself}\label{comp-ww}
+We state judgmental equality rules expressing
+that weakening is compatible with itself. These rules state that the following
+diagram commutes given any two families $A$ and $B$ in context $\Gamma$:\begin{equation*}
+\begin{tikzcd}[column sep=huge]
+\jfam{\Gamma}{\blank} 
+  \ar{r}{C\mapsto\ctxwk{B}{C}} 
+  \ar{d}[swap]{C\mapsto\ctxwk{A}{C}} 
+& \jfam{{\Gamma}{B}}{\blank} 
+  \ar{d}{Q\mapsto\ctxwk{A}{Q}}
+  \\
+\jfam{{\Gamma}{A}}{\blank} 
+  \ar{r}[swap]{P\mapsto\ctxwk{{A}{B}}{P}} 
+& \jfam{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\blank}
+\end{tikzcd}
+\end{equation*}
+Thus, we get the following set of inference rules:\begin{align}
+& \inference
+  { \jfam{\Gamma}{A}
+    \jfam{\Gamma}{B}
+    \jfam{{\Gamma}{C}}{R}
+    }
+  { \jfameq
+      {{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{{A}{B}}{{A}{C}}}}
+      {\ctxwk{A}{{B}{R}}}
+      {\ctxwk{{A}{B}}{{A}{R}}}
+    }
+  \label{comp-ww-f}\\
+& \inference
+  { \jfam{\Gamma}{A}
+    \jfam{\Gamma}{B}
+    \jterm{{\Gamma}{C}}{R}{t}
+    }
+  { \jtermeq
+      {{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{{A}{B}}{{A}{C}}}}
+      {\ctxwk{{A}{B}}{{A}{R}}}
+      {\ctxwk{A}{{B}{t}}}
+      {\ctxwk{{A}{B}}{{A}{t}}}
+    }
+  \label{comp-ww-t}
+\end{align}
+
+\subsubsection{Extension is compatible with weakening}\label{comp-ew}
+The rules expressing that extension is compatible with weakening assert that
+weakening by an extension is the same thing as weakening twice in the
+appropriate way.
+
+In the following inference rules we assume that
+$\jfam{\Gamma}{A}$ and $\jfam{{\Gamma}{A}}{P}$. \begin{align}
+& \inference
+  { \jfam{{\Gamma}{B}}{Q}
+    }
+  { \jfameq
+      {{{{\Gamma}{A}}{P}}{\ctxwk{P}{{A}{B}}}}
+      {\ctxwk{\ctxext{A}{P}}{Q}}
+      {\ctxwk{P}{{A}{Q}}}
+    }
+  \label{comp-ew-f}\\
+& \inference
+  { \jterm{{\Gamma}{B}}{Q}{g}
+    }
+  { \jtermeq
+      {{{{\Gamma}{A}}{P}}{\ctxwk{P}{{A}{B}}}}
+      {\ctxwk{P}{{A}{Q}}}
+      {\ctxwk{\ctxext{A}{P}}{g}}
+      {\ctxwk{P}{{A}{g}}}
+    } 
+  \label{comp-ew-t}
+\end{align}
+
+\subsubsection{Substitution is compatible with the empty context}
+The fibers of the empty family are the empty families:\begin{align}
+& \inference
+  { \jterm{\Gamma}{A}{x}
+    }
+  { \jfameq{\Gamma}{\subst{x}{\emptyf}}{\emptyf}
+    }
+  \\
+& \inference
+  { \jterm{\Gamma}{A}{x}
+    \jfam{{\Gamma}{A}}{P}
+    }
+  { \jfameq
+      {{\Gamma}{\subst{x}{P}}}
+      {\subst{x}{\emptyf}}
+      {\emptyf}
+    }
+\end{align}
+
+The following rules assert that substituting by the term $\jterm{\Gamma}{\emptyf}{\emptytm}$
+leaves everything unchanged.\begin{align}
+& \inference
+  { \jfam{\Gamma}{A}
+    }
+  { \jfameq{\Gamma}{\subst{\emptytm}{A}}{A}
+    }
+  \\
+& \inference
+  { \jterm{\Gamma}{A}{x}
+    }
+  { \jtermeq{\Gamma}{A}{\subst{\emptytm}{x}}{x}
+    }
+  \\
+& \inference
+  { \jfam{{\Gamma}{A}}{P}
+    }
+  { \jfameq{{\Gamma}{A}}{\subst{\emptytm}{P}}{P}
+    }
+  \\
+& \inference
+  { \jterm{{\Gamma}{A}}{P}{f}
+    }
+  { \jtermeq{{\Gamma}{A}}{P}{\subst{\emptytm}{f}}{f}
+    }.
+\end{align}
+
+\subsubsection{Substitution is compatible with extension}\label{comp-se}
+Suppose $\jterm{\Gamma}{A}{x}$ in all of the following inference rule.\begin{align}
+& \inference
+  { \jfam{{{{\Gamma}{A}}{P}}{Q}}{R}
+    }
+  { \jfameq
+      {{\Gamma}{\subst{x}{P}}}
+      {\subst{x}{\ctxext{Q}{R}}}
+      {\ctxext{\subst{x}{Q}}{\subst{x}{R}}}
+    }
+\end{align}
+
+\subsubsection{Substitution is compatible with weakening}\label{comp-sw}
+The rules asserting the compatibility of substitution with weakening assert
+that the following diagram commutes for any $\jterm{\Gamma}{A}{x}$ and any
+$\jfam{{\Gamma}{A}}{P}$:\begin{equation*}
+\begin{tikzcd}[column sep=huge]
+\jfam{{\Gamma}{A}}{\blank} 
+  \ar{d}[swap]{Q\mapsto\subst{x}{Q}} 
+  \ar{r}{Q\mapsto\ctxwk{P}{Q}} 
+& \jfam{{{\Gamma}{A}}{P}}{\blank} 
+    \ar{d}{R\mapsto\subst{x}{R}}
+  \\ 
+\jfam{\Gamma}{\blank} 
+  \ar{r}[swap]{B\mapsto\ctxwk{\subst{x}{P}}{B}} 
+& \jfam{{\Gamma}{\subst{x}{P}}}{\blank}
+\end{tikzcd}
+\end{equation*}
+We plug in an extra layer of families to cover the most general case at once.
+In the following inference rules we assume that $\jterm{\Gamma}{A}{x}$ and
+$\jfam{{\Gamma}{A}}{P}$:\begin{align}
+& \inference
+  { \jfam{{{\Gamma}{A}}{Q}}{R}
+    }
+  { \jfameq
+      {{{\Gamma}{\subst{x}{P}}}{\subst{x}{\ctxwk{P}{Q}}}}
+      {\subst{x}{\ctxwk{P}{R}}}
+      {\ctxwk{\subst{x}{P}}{\subst{x}{R}}}
+    }
+  \label{comp-sw-f}\\
+& \inference
+  { \jterm{{{\Gamma}{A}}{Q}}{R}{h}
+    }
+  { \jtermeq
+      {\ctxext{{\Gamma}{\subst{x}{P}}}{\subst{x}{\ctxwk{P}{Q}}}}
+      {\subst{x}{\ctxwk{P}{R}}}
+      {\subst{x}{\ctxwk{P}{h}}}
+      {\ctxwk{\subst{x}{P}}{\subst{x}{h}}}
+    }
+  \label{comp-sw-t}
+\end{align}
+
+\subsubsection{Substitution is compatible with substitution}\label{comp-ss}
+
+We require that substitution is compatible with itself, which is roughly the
+assertion that substitution is associative. However, we cannot just state that
+$\subst{x}{{f}{g}}\jdeq\subst{{x}{f}}{g}$ since the expression $\subst{{x}{f}}{g}$
+is not well-formed. The term $\subst{x}{f}$ can be substituted in (terms of) families over
+$\subst{x}{P}$; the term $\subst{x}{g}$ is such. Therefore, associativity of
+substitution takes the form $\subst{x}{{f}{g}}\jdeq\subst{{x}{f}}{{x}{g}}$.
+Note that the term $\subst{{x}{f}}{{x}{g}}$ may be written down more conveniently
+as $\subst{x,\subst{x}{f}}{g}$, although we will not do that here.
+
+In the following inference rules we assume
+$\jterm{\Gamma}{A}{x}$ and $\jterm{{\Gamma}{A}}{P}{f}$.
+\begin{align}
+& \inference
+  { \jfam{{{{\Gamma}{A}}{P}}{Q}}{R}
+    }
+  { \jfameq
+      {{\Gamma}{\subst{x}{{f}{Q}}}}
+      {\subst{x}{{f}{R}}}
+      {\subst{{x}{f}}{{x}{R}}}
+    }
+  \label{comp-ss-f}\\
+& \inference
+  { \jterm{{{{\Gamma}{A}}{P}}{Q}}{R}{h}
+    }
+  { \jtermeq
+      {{\Gamma}{\subst{x}{{f}{Q}}}}
+      {\subst{x}{{f}{R}}}
+      {\subst{x}{{f}{h}}}
+      {\subst{{x}{f}}{{x}{h}}}
+    }
+  \label{comp-ss-t}
+\end{align}
+
+\subsubsection{Weakening is compatible with substitution}\label{comp-ws}
+We already have rules for the compatibility of substitution with weakening, but
+we still need the rules the other way around, asserting that there is a 
+judgmental equality $\ctxwk{A}{\subst{y}{Q}}\jdeq\subst{\ctxwk{A}{y}}{\ctxwk{A}{Q}}$
+together with all its variants.
+
+In the following inference rules we assume that $\jfam{\Gamma}{A}$ and that
+$\jterm{\Gamma}{B}{y}$.
+\begin{align}
+& \inference
+  { \jfam{{{\Gamma}{B}}{Q}}{R}
+    }
+  { \jfameq
+      {{{\Gamma}{A}}{\ctxwk{A}{\subst{y}{Q}}}}
+      {\ctxwk{A}{\subst{y}{R}}}
+      {\subst{\ctxwk{A}{y}}{\ctxwk{A}{R}}}
+    }
+  \label{comp-ws-f}\\
+& \inference
+  { \jterm{{{\Gamma}{B}}{Q}}{R}{h}
+    }
+  { \jtermeq
+      {{{\Gamma}{A}}{\ctxwk{A}{\subst{y}{Q}}}}
+      {\ctxwk{A}{\subst{y}{R}}}
+      {\ctxwk{A}{\subst{y}{h}}}
+      {\subst{\ctxwk{A}{y}}{\ctxwk{A}{h}}}
+    }
+  \label{comp-ws-t}
+\end{align}
+
+\subsubsection{The cancellation properties of identity terms}
 Identity terms are determined by their behavior with respect to substitution combined with
@@
 & \inference
-  { \jhom{\Gamma}{A}{B}{f}
-    }
-  { \jhomeq{\Gamma}{A}{B}{\jcomp{A}{f}{\idtm{B}}}{f}
+  { \unfoldall{\jhom{\Gamma}{A}{B}{f}}
+    }
+  { \unfoldall{\jhomeq{\Gamma}{A}{B}{\jcomp{A}{f}{\idtm{B}}}{f}}
     }
@@
     }
-  { \jhomeq
+  { \unfoldall{\jhomeq
       {{\Gamma}{A}}
@@
       {\idtm{\ctxwk{A}{B}}}
-    }
+    }}
 \end{equation}
@@
     }
-  { \jhomeq
+  { \unfoldall{\jhomeq
       {\Gamma}
@@
       {\idtm{\subst{x}{P}}}
-    }
+    }}
 \end{equation}
 
-Now let $\jfam{{\Gamma}{A}}{P}$ and $\jfam{{\Gamma}{A}}{Q}$ be families. A
-morphism from $P$ to $Q$ over the identity term $\idtm{A}$ in context
-$\Gamma$ is the same thing as a morphism from $P$ to $Q$ in context
-$\ctxext{\Gamma}{A}$:
-
-\begin{lem}\label{hom-over-id-is-hom}
-We have the following valid inference rules:\begin{align*}
-& \inference
-  { \jfam{{\Gamma}{A}}{P}
-    \jfam{{\Gamma}{A}}{Q}
-    \jfhom{\Gamma}{A}{A}{\idtm{A}}{P}{Q}{f}
-    }
-  { \jhom{{\Gamma}{A}}{P}{Q}{f}
-    }
-  \\
-& \inference
-  { \jfam{{\Gamma}{A}}{P}
-    \jfam{{\Gamma}{A}}{Q}
-    \jhom{{\Gamma}{A}}{P}{Q}{f}
-    }
-  { \jfhom{\Gamma}{A}{A}{\idtm{A}}{P}{Q}{f}
-    }
-\end{align*}
-\end{lem}
-
-\begin{proof}
-If we unfold the judgments $\jhom{{\Gamma}{A}}{P}{Q}{f}$ and
-$\jfhom{\Gamma}{A}{A}{\idtm{A}}{P}{Q}{f}$, we get the judgments\begin{align*}
-& \unfold{\jhom{{\Gamma}{A}}{P}{Q}{f}}
-  \\
-& \unfold{\jfhom{\Gamma}{A}{A}{\idtm{A}}{P}{Q}{f}}
-\end{align*}
-respectively. Therefore, we only need to verify that
-$\ctxwk{P}{\subst{\idtm{A}}{\ctxwk{A}{Q}}}\jdeq\ctxwk{P}{Q}$, which is indeed
-the case by \autoref{idfunc-wk-defn}.
-\end{proof}
+\subsubsection{The cancellation property of weakening and substitution}
+The judgmental equalities we're about to describe assert that substituting a term
+in the weakening a thing gives you the thing back. In the case of contexts we get that each fiber
+$\subst{x}{\ctxwk{A}{B}}$ is just $B$ and in the case of terms we get 
+that $\ctxwk{A}{y}$ is the constant function
+mapping everything to $y:B$. Thus, these rules actually establish the weakening
+as the weakening. After stating the rules we will describe what it means to
+compose context morphisms (terms of weakened contexts).
+\begin{align}
+& \inference
+  { \jfam{\Gamma}{A}
+    \jfam{{\Gamma}{B}}{Q}
+    \jterm{\Gamma}{A}{x}
+    }
+  { \jfameq{{\Gamma}{B}}{\subst{x}{\ctxwk{A}{Q}}}{Q}
+    }
+  \label{defn-ws-3}\\
+& \inference
+  { \jterm{\Gamma}{A}{x}
+    \jterm{{\Gamma}{B}}{Q}{g}
+    }
+  { \jtermeq{{\Gamma}{B}}{Q}{\subst{x}{\ctxwk{A}{g}}}{g}
+    }
+  \label{defn-ws-4}
+\end{align}
 

### 2014-04-08 article-models-ttderived.tex
reorganizing section 4

@@
 
+\subsection{Morphisms}\label{categorical_properties}
+Using the rules of the compatibility of substitution with weakening and of the
+compatibility of weakening with itself, we see that we can show
+
+\begin{lem}
+The inference rule\begin{equation*}
+\inference
+  { \jfam{\Gamma}{A}
+    \jfam{\Gamma}{B}
+    \jfam{\Gamma}{C}
+    \jhom{\Gamma}{A}{B}{f}
+    }
+  { \jfameq
+    {{\Gamma}{A}}
+    {\subst{f}{\ctxwk{A}{{B}{C}}}}
+    {\ctxwk{A}{C}}
+    }
+\end{equation*}
+is valid.
+\end{lem}
+
+\begin{proof}
+Let $\jfam{\Gamma}{A}$, $\jfam{\Gamma}{B}$, $\jfam{\Gamma}{C}$ and $\jhom{\Gamma}{A}{B}{f}$.
+Then we have the judgmental equalities\begin{align*}
+\subst{f}{\ctxwk{A}{{B}{C}}}
+& \jdeq 
+  \subst{f}{\ctxwk{{A}{B}}{{A}{C}}}
+  \\
+& \jdeq 
+  \ctxwk{A}{C}.
+  \qedhere
+\end{align*}
+\end{proof}
+
+It follows that for $\jterm{{\Gamma}{B}}{\ctxwk{B}{C}}{g}$ we can compose $f$
+with $g$ to obtain a term of $\ctxwk{A}{C}$ in context $\ctxext{\Gamma}{A}$.
+In the following definition, we work with in a slightly greater generality.
+
+\begin{defn}
+We define the judgment\begin{equation*}
+\jhom{\Gamma}{A}{B}{f},
+\end{equation*}
+which is pronounced as `$f$ is a morphism from $A$ to $B$ in context $\Gamma$',
+to be the judgment\begin{equation*}
+\unfold{\jhom{\Gamma}{A}{B}{f}}.
+\end{equation*}
+Likewise, we define the judgment\begin{equation*}
+\jhomeq{\Gamma}{A}{B}{f}{f'}
+\end{equation*}
+to be the judgment\begin{equation*}
+\unfold{\jhomeq{\Gamma}{A}{B}{f}{f'}}.
+\end{equation*}
+\end{defn}
+
+\begin{defn}
+Let $\jhom{\Gamma}{A}{B}{f}$ and consider a term $\jterm{{\Gamma}{B}}{Q}{g}$.
+We define\begin{align*}
+\jfamdefn*
+  {{\Gamma}{A}}
+  {\jcomp{A}{f}{Q}}
+  {\unfold{\jcomp{A}{f}{Q}}}\\
+\jtermdefn*
+  {{\Gamma}{A}}
+  {\jcomp{A}{f}{Q}}
+  {\jcomp{A}{f}{g}}
+  {\unfold{\jcomp{A}{f}{g}}}.
+\end{align*}
+Likewise, when we have a family $\jfam{{{\Gamma}{B}}{Q}}{R}$ and a term
+$\jterm{{{\Gamma}{B}}{Q}}{R}{h}$, we define\begin{align*}
+\jfamdefn*
+  {{{\Gamma}{A}}{\jcomp{A}{f}{Q}}}
+  {\jcomp{A}{f}{R}}
+  {\unfold{\jcomp{A}{f}{R}}}
+  \\
+\jtermdefn*
+  {{{\Gamma}{A}}{\jcomp{A}{f}{Q}}}
+  {\jcomp{A}{f}{R}}
+  {\jcomp{A}{f}{h}}
+  {\unfold{\jcomp{A}{f}{h}}}.
+\end{align*}
+\end{defn}
+
+We have lots of compatibility properties for composition:
+
+\begin{lem}
+We have the following inference rules about the situation where something is
+substituted by a composition:\begin{align*}
+& \inference
+  { \jhom{\Gamma}{A}{B}{f}
+    \jhom{\Gamma}{B}{C}{g}
+    \jfam{{{\Gamma}{A}}{\ctxwk{A}{C}}}{R}
+    }
+  { \jfameq
+      {{\Gamma}{A}}
+      {\subst{\jcomp{A}{f}{g}}{R}}
+      {\subst{f}{{\ctxwk{A}{g}}{\ctxwk{{A}{B}}{R}}}}
+    }
+  \\
+& \inference
+  { \jhom{\Gamma}{A}{B}{f}
+    \jhom{\Gamma}{B}{C}{g}
+    \jterm{{{\Gamma}{A}}{\ctxwk{A}{C}}}{R}{h}
+    }
+  { \jfameq
+    {{\Gamma}{A}}
+    {\subst{\jcomp{A}{f}{g}}{h}}
+    {\subst{f}{{\ctxwk{A}{g}}{\ctxwk{{A}{B}}{h}}}}
+    }
+\end{align*}
+We also have the following related inference rules, asserting that composition
+is strictly associative:\begin{align*}
+& \inference
+  { \jhom{\Gamma}{A}{B}{f}
+    \jhom{\Gamma}{B}{C}{g}
+    \jfam{{\Gamma}{C}}{R}
+    }
+  { \jfameq
+      {{\Gamma}{A}}
+      {\jcomp{A}{{A}{f}{g}}{R}}
+      {\jcomp{A}{f}{{B}{g}{R}}}
+    }
+  \\
+& \inference
+    { \jhom{\Gamma}{A}{B}{f}
+      \jhom{\Gamma}{B}{C}{g}
+      \jterm{{\Gamma}{C}}{R}{h}
+      }
+    { \jtermeq
+        {{\Gamma}{A}}
+        {\jcomp{A}{{A}{f}{g}}{R}}
+        {\jcomp{A}{{A}{f}{g}}{h}}
+        {\jcomp{A}{f}{{B}{g}{h}}}
+      }
+\end{align*}
+\end{lem}
+
+\begin{proof}
+Consider family morphisms $\jhom{\Gamma}{A}{B}{f}$ and $\jhom{\Gamma}{B}{C}{g}$
+and a family $\jfam{{{\Gamma}{A}}{\ctxwk{A}{C}}}{R}$. Then we have the judgmental
+equalities\begin{align*}
+\subst{\jcomp{A}{f}{g}}{R} 
+& \jdeq 
+  \subst{{f}{\ctxwk{A}{g}}}{R}
+  \\
+& \jdeq 
+  \subst{{f}{\ctxwk{A}{g}}}{\subst{f}{\ctxwk{{A}{B}}{R}}}
+  \\
+& \jdeq 
+  \subst{f}{{\ctxwk{A}{g}}{\ctxwk{{A}{B}}{R}}}
+\end{align*}
+The proof that 
+$\subst{\jcomp{A}{f}{g}}{h}\jdeq\subst{f}{{\ctxwk{A}{g}}{\ctxwk{{A}{B}}{h}}}$
+is similar.
+
+Now suppose that $\jfam{{\Gamma}{C}}{R}$ instead. Then we have\begin{align*}
+\jcomp{A}{{A}{f}{g}}{R} 
+& \jdeq 
+  \subst{\jcomp{A}{f}{g}}{\ctxwk{A}{R}}
+  \\
+& \jdeq 
+  \subst{{f}{\ctxwk{A}{g}}}{\ctxwk{A}{R}}
+  \\
+& \jdeq 
+  \subst{f}{{\ctxwk{A}{g}}{\ctxwk{{A}{B}}{{A}{R}}}}
+  \\
+& \jdeq 
+  \subst{f}{{\ctxwk{A}{g}}{\ctxwk{A}{{B}{R}}}}
+  \\
+& \jdeq 
+  \subst{f}{\ctxwk{A}{\subst{g}{\ctxwk{B}{R}}}}
+  \\
+& \jdeq 
+  \subst{f}{\ctxwk{A}{\jcomp{B}{g}{R}}}
+  \\
+& \jdeq 
+  \jcomp{A}{f}{{B}{g}{R}}.
+\end{align*}
+Again, the proof is similar for terms $h$ of $R$ in context $\ctxext{\Gamma}{C}$.
+\end{proof}
+
+\begin{lem}
+We have the following inference rules about the compatibility of composition with
+weakening:\begin{align*}
+& \inference
+  { \jhom{\Gamma}{A}{B}{f}
+    \jhom{\Gamma}{B}{C}{g}
+    \jfam{{\Gamma}{A}}{P}
+    }
+  { \jhomeq
+      {\Gamma}
+      {{A}{P}}
+      {C}
+      {\ctxwk{P}{\jcomp{A}{f}{g}}}
+      {\jcomp{{A}{P}}{\ctxwk{P}{f}}{g}}
+    }
+  \\
+& \inference
+  { \jterm{\Gamma}{B}{y}
+    \jhom{\Gamma}{B}{C}{g}
+    }
+  { \jhomeq
+      {\Gamma}
+      {A}
+      {C}
+      {\jcomp{A}{\ctxwk{A}{y}}{g}}
+      {\ctxwk{A}{\subst{y}{g}}}
+    }
+  \\
+& \inference
+  { \jhom{\Gamma}{A}{B}{f}
+    \jterm{\Gamma}{C}{z}
+    }
+  { \jhomeq
+      {\Gamma}
+      {A}
+      {C}
+      {\jcomp{A}{f}{\ctxwk{B}{z}}}
+      {\ctxwk{A}{z}}
+    }
+\end{align*}
+\end{lem}
+
+\begin{proof}
+Let $\jhom{\Gamma}{A}{B}{f}$, $\jhom{\Gamma}{B}{C}{g}$ and $\jfam{{\Gamma}{A}}{P}$.
+Then we have the judgmental equalities\begin{align*}
+\ctxwk{P}{\jcomp{A}{f}{g}} 
+& \jdeq 
+  \ctxwk{P}{\subst{f}{\ctxwk{A}{g}}}
+  \\
+& \jdeq 
+  \subst{\ctxwk{P}{f}}{\ctxwk{P}{{A}{g}}}
+  \\
+& \jdeq 
+  \subst{\ctxwk{P}{f}}{\ctxwk{\ctxext{A}{P}}{g}}
+  \\
+& \jdeq 
+  \jcomp{{A}{P}}{\ctxwk{P}{f}}{g}.
+\end{align*}
+Now let $\jterm{\Gamma}{B}{y}$ and $\jhom{\Gamma}{B}{C}{g}$. Then we have the
+judgmental equalities\begin{align*}
+\jcomp{A}{\ctxwk{A}{y}}{g}
+& \jdeq 
+  \subst{\ctxwk{A}{y}}{\ctxwk{A}{g}}
+  \\
+& \jdeq 
+  \ctxwk{A}{\subst{y}{g}}.
+\end{align*}
+For the third assertion, let $\jhom{\Gamma}{A}{B}{f}$ and $\jterm{\Gamma}{C}{z}$.
+Then we have the judgmental equalities\begin{align*}
+\jcomp{A}{f}{\ctxwk{B}{z}} 
+& \jdeq 
+  \subst{f}{\ctxwk{A}{{B}{z}}}
+  \\
+& \jdeq 
+  \subst{f}{\ctxwk{{A}{B}}{{A}{z}}}
+  \\
+& \jdeq 
+  \ctxwk{A}{z}.
+  \qedhere
+\end{align*}
+\end{proof}
+
+\begin{lem}
+We have the following inference rules about the compatibility of composition with
+substitution:\begin{align*}
+& \inference
+  { \jhom{{\Gamma}{A}}{P}{Q}{f}
+    \jhom{{\Gamma}{A}}{Q}{R}{g}
+    \jterm{\Gamma}{A}{x}
+    }
+  { \jhomeq
+      {\Gamma}
+      {\subst{x}{P}}
+      {\subst{x}{R}}
+      {\subst{x}{\jcomp{P}{f}{g}}}
+      {\jcomp{\subst{x}{P}}{\subst{x}{f}}{\subst{x}{g}}}
+    }
+  \\
+& \inference
+  { \jhom{\Gamma}{A}{B}{f}
+    \jhom{\Gamma}{B}{C}{g}
+    \jterm{\Gamma}{A}{x}
+    }
+  { \jtermeq
+      {\Gamma}
+      {C}
+      {\subst{x}{\jcomp{A}{f}{g}}}
+      {\subst{{x}{f}}{g}}
+    }
+\end{align*}
+\end{lem}
+
+\begin{proof}
+Let $\jhom{{\Gamma}{A}}{P}{Q}{f}$, $\jhom{{\Gamma}{A}}{Q}{R}{g}$ and 
+$\jterm{\Gamma}{A}{x}$.
+Then we have the judgmental equalities\begin{align*}
+\subst{x}{\jcomp{A}{f}{g}}
+& \jdeq 
+  \subst{x}{{f}{\ctxwk{P}{g}}}
+  \\
+& \jdeq 
+  \subst{{x}{f}}{{x}{\ctxwk{P}{g}}}
+  \\
+& \jdeq 
+  \subst{{x}{f}}{\ctxwk{\subst{x}{P}}{\subst{x}{g}}}
+  \\
+& \jdeq 
+  \jcomp{\subst{x}{P}}{\subst{x}{f}}{\subst{x}{g}}.
+\end{align*}
+Now let $\jhom{\Gamma}{A}{B}{f}$, $\jhom{\Gamma}{B}{C}{g}$ and $\jterm{\Gamma}{A}{x}$.
+Then we have the judgmental equalities\begin{align*}
+\subst{x}{\jcomp{A}{f}{g}}
+& \jdeq 
+  \subst{x}{{f}{\ctxwk{A}{g}}}
+  \\
+& \jdeq 
+  \subst{{x}{f}}{{x}{\ctxwk{A}{g}}}
+  \\
+& \jdeq 
+  \subst{{x}{f}}{g}.
+  \qedhere
+\end{align*}
+\end{proof}
+
+There is also a notion of morphism \emph{over} a morphism. We will develop this
+notion because it will be needed in the theory of models later on.
+
+\begin{defn}
+Let $\jhom{\Gamma}{A}{B}{f}$ be a morphism from $A$ to $B$ in context $\Gamma$
+and consider $\jfam{{\Gamma}{A}}{P}$ and $\jfam{{\Gamma}{B}}{Q}$. We define the
+judgment\begin{equation*}
+\jfhom{\Gamma}{A}{B}{f}{P}{Q}{F},
+\end{equation*}
+which is pronounced as `$F$ is a morphism from $P$ to $Q$ over $f$ in context
+$\Gamma$', to be the judgment\begin{equation*}
+\unfold{\jfhom{\Gamma}{A}{B}{f}{P}{Q}{F}}.
+\end{equation*}
+\end{defn}
+
+\begin{rmk}
+The judgment $\jfhom{\Gamma}{A}{B}{f}{P}{Q}{F}$ means the same thing as\begin{equation*}
+\jhom{{\Gamma}{A}}{P}{\jcomp{A}{f}{Q}}{F}.
+\end{equation*}
+\end{rmk}
+
+Suppose we have morphisms $\jhom{\Gamma}{A}{B}{f}$ and $\jhom{\Gamma}{B}{C}{g}$
+and that we have the morphisms $\jfhom{\Gamma}{A}{B}{f}{P}{Q}{F}$ and
+$\jfhom{\Gamma}{B}{C}{g}{Q}{R}{G}$ over them. Then we have\begin{equation*}
+\jhom
+  {{\Gamma}{A}}
+  {\jcomp{A}{f}{Q}}
+  {\jcomp{A}{f}{{B}{g}{R}}}
+  {\unfold{\jcomp{A}{f}{G}}}
+\end{equation*}
+Because we also have $\jhom{{\Gamma}{A}}{P}{\jcomp{A}{f}{Q}}{F}$, we have the
+composition\begin{equation*}
+\jhom
+  {{\Gamma}{A}}
+  {P}
+  {\jcomp{A}{f}{{B}{g}{R}}}
+  {\jcomp{P}{F}{\unfold{\jcomp{A}{f}{G}}}}.
+\end{equation*}
+Because of 
+the judgmental equality $\jcomp{A}{f}{{B}{g}{R}}\jdeq
+\jcomp{A}{{A}{f}{g}}{R}$, it follows that 
+$\jcomp{P}{F}{\unfold{\jcomp{A}{f}{G}}}$ is a morphism from $P$ to $R$ over
+$\jcomp{A}{f}{g}$. We make the following definition:
+
+\begin{defn}
+Let $\jhom{\Gamma}{A}{B}{f}$ and $\jhom{\Gamma}{B}{C}{g}$
+be morphisms and let $\jfhom{\Gamma}{A}{B}{f}{P}{Q}{F}$ and
+$\jfhom{\Gamma}{B}{C}{g}{Q}{R}{G}$ be morphisms over them. Then we define\begin{equation*}
+\jfhomdefn
+  {\Gamma}
+  {A}
+  {C}
+  {\jcomp{A}{f}{g}}
+  {P}
+  {R}
+  {\jfcomp{A}{f}{P}{F}{G}}
+  {\unfold{\jfcomp{A}{f}{P}{F}{G}}}.
+\end{equation*}
+\end{defn}
+
+This composition is also judgmentally associative.
+
+Now let $\jfam{{\Gamma}{A}}{P}$ and $\jfam{{\Gamma}{A}}{Q}$ be families. A
+morphism from $P$ to $Q$ over the identity term $\idtm{A}$ in context
+$\Gamma$ is the same thing as a morphism from $P$ to $Q$ in context
+$\ctxext{\Gamma}{A}$:
+
+\begin{lem}\label{hom-over-id-is-hom}
+We have the following valid inference rules:\begin{align*}
+& \inference
+  { \jfam{{\Gamma}{A}}{P}
+    \jfam{{\Gamma}{A}}{Q}
+    \jfhom{\Gamma}{A}{A}{\idtm{A}}{P}{Q}{f}
+    }
+  { \jhom{{\Gamma}{A}}{P}{Q}{f}
+    }
+  \\
+& \inference
+  { \jfam{{\Gamma}{A}}{P}
+    \jfam{{\Gamma}{A}}{Q}
+    \jhom{{\Gamma}{A}}{P}{Q}{f}
+    }
+  { \jfhom{\Gamma}{A}{A}{\idtm{A}}{P}{Q}{f}
+    }
+\end{align*}
+\end{lem}
+
+\begin{proof}
+If we unfold the judgments $\jhom{{\Gamma}{A}}{P}{Q}{f}$ and
+$\jfhom{\Gamma}{A}{A}{\idtm{A}}{P}{Q}{f}$, we get the judgments\begin{align*}
+& \unfold{\jhom{{\Gamma}{A}}{P}{Q}{f}}
+  \\
+& \unfold{\jfhom{\Gamma}{A}{A}{\idtm{A}}{P}{Q}{f}}
+\end{align*}
+respectively. Therefore, we only need to verify that
+$\ctxwk{P}{\subst{\idtm{A}}{\ctxwk{A}{Q}}}\jdeq\ctxwk{P}{Q}$, which is indeed
+the case by \autoref{idfunc-wk-defn}.
+\end{proof}
+
 \subsection{Projections and extension on terms}\label{extension-on-terms}

### 2014-04-08 article-models-tt.tex
improvements

@@
 
-\subsection{Judgments and inference rules}\label{judgments}
+\subsection{Judgments and valid inference rules}\label{judgments}
 The theory we describe here is a theory of contexts, families of
@@
 and ``$x$ is a term of the family $A$ of contexts over $\Gamma$''. The other four
-judgments are for judgmental equality. \begin{align*}
+judgments are for judgmental equality.
+\begin{align*}
 \jctx*{\Gamma} 
@@
 
+The theory of contexts, families and terms shall be described by means of a
+declaration of valid inference rules. An inference rule is an expression of the
+form
+\begin{equation*}
+\inference{\mathcal{H}_1\quad\cdots\quad\mathcal{H}_n}{\mathcal{J}}
+\end{equation*}
+where each of $\mathcal{H}_1$, ..., $\mathcal{H}_n$ and $\mathcal{J}$ is 
+judgment. The judgments $\mathcal{H}_1$, ..., $\mathcal{H}_n$ are called the
+hypotheses and the judgment $\mathcal{J}$ is the conclusion of the inference
+rule. The class of valid inference rules is closed under composition. That
+means that if
+\begin{equation*}
+\inference{\mathcal{I}_{1k}\quad\cdots\quad\mathcal{I}_{m_kk}}{\mathcal{H}_k}
+\end{equation*}
+and
+\begin{equation*}
+\inference{\mathcal{H}_1\quad\cdots\quad\mathcal{H}_n}{\mathcal{J}}
+\end{equation*}
+are valid inference rules for $1\leq k\leq n$, then
+\begin{equation*}
+\inference{\mathcal{I}_{11}\quad\cdots\quad\mathcal{I}_{m_11}\quad\cdots\quad\mathcal{I}_{1n}\quad\cdots\quad\mathcal{I}_{m_nn}}{\mathcal{J}}
+\end{equation*}
+is also a valid inference rule. Also, in a valid inference rule 
+duplicated hypotheses may always be contracted and hypotheses may always be 
+reordered; the result is still considered a valid inference rule.
+
 \subsection{The basic rules for judgmental equality}
 The rules for judgmental equality establish that it is an equivalence relation
-in all three cases (contexts, types and terms).
-\bgroup\small\begin{align*}
+in all three cases (contexts, families and terms). Thus, the following inference
+rules shall be required to be valid:
+\bgroup\small
+\begin{align*}
 & \inference
@@
 The following convertibility rules are responsible for the strictness
-of judgmental equality, which sets it apart from equivalences or identifications:\begin{align*}
+of judgmental equality, which sets it apart from equivalences or identifications:
+\begin{align*}
 & \inference
@@
 family over any context. Moreover, any two terms of the empty family are always
-judgmentally equal.\begin{align}
+judgmentally equal.
+\begin{align}
 & \inference
@@
  rules asserting a judgmental equality we assume furthermore that 
- $\jfameq{\Gamma}{A}{A'}$.\begin{align}
+ $\jfameq{\Gamma}{A}{A'}$.
+ \begin{align}
 & \inference
@@
 \subsubsection{Identity terms}
-Without a rule explicitly asserting the existence of an identity morphism we don't
-get one, hence we do that here. The identity morphism is a term which introduced
-in ordinary type theory via the variable rule. The variable rule is a bit more
-general: it asserts that\begin{equation*}
-\jterm{\Gamma,\,x_1:A_1,\ldots,\,x_n:A_n}{A_i}{x_i}
-\end{equation*}
-for every $1\leq i\leq n$. Thus, it establishes the projections. In 
-the current setting,
-we get the projections from the identity morphisms together with weakening. We
-already have weakening, so here it suffices to introduce the identity morphisms.\begin{align}
+Last, but not least, we introduce the identity terms. In combination with
+weakening and substitution, identity terms will allow us to speak of projections,
+of a variant of extension for terms and of fiber inclusions.
+
+\begin{align}
 & \inference
@@
 terms}
-
-\subsubsection{The empty context is compatible with itself}
-The empty context $\emptyc$ may be considered as a family of contexts over the empty
-context. When we do this, we get $\emptyf[\emptyc]$.\begin{equation}
+In this section we extend the judgmental equality relation in such a way that
+the empty context, extension, weakening, substitution and the identity terms
+all become compatible with each other. Since we have five operations, there
+are going to be $25$ finite sets of rules explaining how each operation is
+compatible with each of the operations. In addition to these $25$ finite sets
+of compatibility rules, there are also two sets of rules explaining the
+cancellation properties of weakening and substitution and of the identity terms.
+
+\subsubsection{The empty context is compatible with itself}\label{comp-00}
+Because every context is a family of contexts over the empty context, 
+the empty context $\emptyc$ may be considered as a family of contexts over the empty
+context. When we do this, we get $\emptyf[\emptyc]$.
+\begin{equation}\label{eq:comp-00}
 \inference
@@
 
-\subsubsection{Extension is compatible with the empty context}
-The following rule asserts that extension by $\emptyc$ leaves the contexts unchanged.\begin{align}
+\subsubsection{Extension is compatible with the empty context}\label{comp-e0}
+The following rule asserts that extension by $\emptyc$ leaves the contexts unchanged.
+\begin{align}
 & \inference
@@
     }
-  \\
+  \label{eq:comp-e0}\\
 & \inference
@@
     }
+\end{align}
+
+\subsubsection{Weakening is compatible with the empty context}
+The following rules express that when the empty context or context family is
+weakened, the result is the empty context family.
+\begin{align}
+& \inference
+  { \jctx{\Gamma}
+    }
+  { \jfameq{\Gamma}{\ctxwk{\Gamma}{\emptyc}}{\emptyf}
+    }
+  \\
+& \inference
+  { \jfam{\Gamma}{A}
+    }
+  { \jfameq{{\Gamma}{A}}{\ctxwk{A}{\emptyf}}{\emptyf}
+    }
+\end{align}
+Weakening by the empty family $\emptyf$ over a context $\Gamma$ leaves families, 
+their terms, families over those families and
+terms of those unchanged:
+\begin{align}
+& \inference
+  { \jfam{\Gamma}{B}
+    }
+  { \jfameq{\Gamma}{\ctxwk{\emptyf}{B}}{B}
+    }
+  \\
+& \inference
+  { \jterm{\Gamma}{B}{y}
+    }
+  { \jtermeq{\Gamma}{B}{\ctxwk{\emptyf}{y}}{y}
+    }
+  \\
+& \inference
+  { \jfam{{\Gamma}{B}}{Q}
+    }
+  { \jfameq{{\Gamma}{B}}{\ctxwk{\emptyf}{Q}}{Q}
+    }
+  \\
+& \inference
+  { \jterm{{\Gamma}{B}}{Q}{g}
+    }
+  { \jtermeq{{\Gamma}{B}}{Q}{\ctxwk{\emptyf}{g}}{g}
+    }
+\end{align}
+
+\subsubsection{Substitution is compatible with the empty context}
+The fibers of the empty family are the empty families:
+\begin{align}
+& \inference
+  { \jterm{\Gamma}{A}{x}
+    }
+  { \jfameq{\Gamma}{\subst{x}{\emptyf}}{\emptyf}
+    }
+  \\
+& \inference
+  { \jterm{\Gamma}{A}{x}
+    \jfam{{\Gamma}{A}}{P}
+    }
+  { \jfameq
+      {{\Gamma}{\subst{x}{P}}}
+      {\subst{x}{\emptyf}}
+      {\emptyf}
+    }
+\end{align}
+
+The following rules assert that substituting by the term $\jterm{\Gamma}{\emptyf}{\emptytm}$
+leaves everything unchanged.
+\begin{align}
+& \inference
+  { \jfam{\Gamma}{A}
+    }
+  { \jfameq{\Gamma}{\subst{\emptytm}{A}}{A}
+    }
+  \\
+& \inference
+  { \jterm{\Gamma}{A}{x}
+    }
+  { \jtermeq{\Gamma}{A}{\subst{\emptytm}{x}}{x}
+    }
+  \\
+& \inference
+  { \jfam{{\Gamma}{A}}{P}
+    }
+  { \jfameq{{\Gamma}{A}}{\subst{\emptytm}{P}}{P}
+    }
+  \\
+& \inference
+  { \jterm{{\Gamma}{A}}{P}{f}
+    }
+  { \jtermeq{{\Gamma}{A}}{P}{\subst{\emptytm}{f}}{f}
+    }.
 \end{align}
@@
   { \jfameq{\Gamma}{\ctxext{{A}{P}}{Q}}{\ctxext{A}{{P}{Q}}}
-    }
-\end{align}
-
-\subsubsection{Weakening is compatible with the empty context}
-The following rules express that when the empty context or context family is
-weakened, the result is the empty context family.\begin{align}
-& \inference
-  { \jctx{\Gamma}
-    }
-  { \jfameq{\Gamma}{\ctxwk{\Gamma}{\emptyc}}{\emptyf}
-    }
-  \\
-& \inference
-  { \jfam{\Gamma}{A}
-    }
-  { \jfameq{{\Gamma}{A}}{\ctxwk{A}{\emptyf}}{\emptyf}
-    }
-\end{align}
-Weakening by the empty family $\emptyf$ over a context $\Gamma$ leaves families, 
-their terms, families over those families and
-terms of those unchanged:\begin{align}
-& \inference
-  { \jfam{\Gamma}{B}
-    }
-  { \jfameq{\Gamma}{\ctxwk{\emptyf}{B}}{B}
-    }
-  \\
-& \inference
-  { \jterm{\Gamma}{B}{y}
-    }
-  { \jtermeq{\Gamma}{B}{\ctxwk{\emptyf}{y}}{y}
-    }
-  \\
-& \inference
-  { \jfam{{\Gamma}{B}}{Q}
-    }
-  { \jfameq{{\Gamma}{B}}{\ctxwk{\emptyf}{Q}}{Q}
-    }
-  \\
-& \inference
-  { \jterm{{\Gamma}{B}}{Q}{g}
-    }
-  { \jtermeq{{\Gamma}{B}}{Q}{\ctxwk{\emptyf}{g}}{g}
     }
@@
 
-In the following inference rules we assume that $\jfam{\Gamma}{A}$.\begin{align}
+In the following inference rules we assume that $\jfam{\Gamma}{A}$.
+\begin{align}
 & \inference
@@
 that weakening is compatible with itself. These rules state that the following
-diagram commutes given any two families $A$ and $B$ in context $\Gamma$:\begin{equation*}
+diagram commutes given any two families $A$ and $B$ in context $\Gamma$:
+\begin{equation*}
 \begin{tikzcd}[column sep=huge]
@@
 \end{equation*}
-Thus, we get the following set of inference rules:\begin{align}
+Thus, we get the following set of inference rules:
+\begin{align}
 & \inference
@@
 In the following inference rules we assume that
-$\jfam{\Gamma}{A}$ and $\jfam{{\Gamma}{A}}{P}$. \begin{align}
+$\jfam{\Gamma}{A}$ and $\jfam{{\Gamma}{A}}{P}$.
+\begin{align}
 & \inference
@@
 
-\subsubsection{Substitution is compatible with the empty context}
-The fibers of the empty family are the empty families:\begin{align}
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    }
-  { \jfameq{\Gamma}{\subst{x}{\emptyf}}{\emptyf}
-    }
-  \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jfam{{\Gamma}{A}}{P}
-    }
-  { \jfameq
-      {{\Gamma}{\subst{x}{P}}}
-      {\subst{x}{\emptyf}}
-      {\emptyf}
-    }
-\end{align}
-
-The following rules assert that substituting by the term $\jterm{\Gamma}{\emptyf}{\emptytm}$
-leaves everything unchanged.\begin{align}
-& \inference
-  { \jfam{\Gamma}{A}
-    }
-  { \jfameq{\Gamma}{\subst{\emptytm}{A}}{A}
-    }
-  \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    }
-  { \jtermeq{\Gamma}{A}{\subst{\emptytm}{x}}{x}
-    }
-  \\
-& \inference
-  { \jfam{{\Gamma}{A}}{P}
-    }
-  { \jfameq{{\Gamma}{A}}{\subst{\emptytm}{P}}{P}
-    }
-  \\
-& \inference
-  { \jterm{{\Gamma}{A}}{P}{f}
-    }
-  { \jtermeq{{\Gamma}{A}}{P}{\subst{\emptytm}{f}}{f}
-    }.
-\end{align}
-
 \subsubsection{Substitution is compatible with extension}\label{comp-se}
-Suppose $\jterm{\Gamma}{A}{x}$ in all of the following inference rule.\begin{align}
+Suppose $\jterm{\Gamma}{A}{x}$ in all of the following inference rule.
+\begin{align}
 & \inference
@@
 that the following diagram commutes for any $\jterm{\Gamma}{A}{x}$ and any
-$\jfam{{\Gamma}{A}}{P}$:\begin{equation*}
+$\jfam{{\Gamma}{A}}{P}$:
+\begin{equation*}
 \begin{tikzcd}[column sep=huge]
@@
 In the following inference rules we assume that $\jterm{\Gamma}{A}{x}$ and
-$\jfam{{\Gamma}{A}}{P}$:\begin{align}
+$\jfam{{\Gamma}{A}}{P}$:
+\begin{align}
 & \inference
@@
 Identity terms are determined by their behavior with respect to substitution combined with
-weakening. The identity terms will also be subject to compatibility rules.\begin{align}
+weakening. The identity terms will also be subject to compatibility rules.
+\begin{align}
 & \inference
@@
 
-The identity term of a weakened family is the weakened identity term:\begin{equation}\label{idfunc-wk-comp}
+The identity term of a weakened family is the weakened identity term:
+\begin{equation}\label{idfunc-wk-comp}
 \inference
@@
 
-The identity term of a substituted family is the substitution of the identity term\begin{equation}\label{idfunc-subst-comp}
+The identity term of a substituted family is the substitution of the identity term
+\begin{equation}\label{idfunc-subst-comp}
 \inference
@@
 new judgments: the judgment that something is a type and the judgment that two
-types are equal.\begin{align*}
+types are equal.
+\begin{align*}
 \jtype*{\Gamma}{A} 
@@
 judgmentally equal to a type $A$ in context $\Gamma$, then $B$ is a type in
-context $\Gamma$. This is expressed by the following four inference rules:\begin{align*}
+context $\Gamma$. This is expressed by the following four inference rules:
+\begin{align*}
 & \inference
@@
 We add rules asserting that a weakened type is again a type that 
-substitution preserves the property of being a type:\begin{align}
+substitution preserves the property of being a type:
+\begin{align}
 & \inference

### 2014-04-08 article-models-tt.tex
added explanations

@@
 \subsection{The operations of the theory of contexts, families and terms}
+In this subsection we introduce the notions of empty context,
+extension, weakening, substitution and identity terms. These are all viewed as
+operations of the theory of contexts, families and terms.
+
+Generally, 
+each operation should
+have an action on contexts, on families and on terms at the `appropriate level'
+and should be compatible with each of the operations we introduce here. In
+that will become more practical after the theory of contexts, families and terms
+is formulated, this means that when we will write down a type theory that
+interprets itself in \autoref{ttmodels}, each of the notions of empty context,
+extension, weakening, substitution and identity terms are going to be internalized
+as terms of the theory and all terms are going to be compatible with these.
+That vague idea will be our guiding light.
+
+So in principle, the introductions of each of the empty context, extension,
+weakening, substitution and identity terms should come in three parts. In 
+practice, they do not all come in three parts.
+
+Apart from the empty context, there will indeed be an empty family and a term
+of the empty family (which will be deemed unique). But context extension will
+not have an explicit action `at the level of families' since extension
+is used to define higher families and the action would have to be trivial
+anyway. The operations of weakening and substitution are actually introduced
+by four inference rules and the identity terms by one.
 
@@
 consider the fiber $\subst{x}{P}$ of $P$ at $x$. Also, we get a way to evaluate
-terms $f$ of $P$ at $x$. This will give us ways to compose functions too. In
-this section, we shall first introduce the operations `substitution of a term $x$'
-for families $P$ over $\ctxext{\Gamma}{A}$, terms $f$ of those, families $Q$ over
-$\ctxext{{\Gamma}{A}}{P}$ and terms $g$ of those. 
-Then we shall explain how substitution interacts
-with itself, extension and weakening.
-
+terms $f$ of $P$ at $x$. This will give us ways to compose morphisms too.
 In the rules introducing the various substitutions we assume $\jterm{\Gamma}{A}{x}$;
@@
 weakening and substitution, identity terms will allow us to speak of projections,
-of a variant of extension for terms and of fiber inclusions.
-
+of a variant of extension for terms and of fiber inclusions. The following
+inference rules are required to be valid:
 \begin{align}

### 2014-04-08 article-models-ttmodels.tex
added explanations

@@
-\section{The type theory of models of type theory}
+\section{The type theory of models of type theory}\label{ttmodels}
 In this section we pursue the idea of what a general model of type theory is by

### 2014-04-09 article-models-tt.tex
rewriting some of the section on the operators of CFT theory

@@
 hypotheses and the judgment $\mathcal{J}$ is the conclusion of the inference
-rule. The class of valid inference rules is closed under composition. That
+rule. When a judgment is the conclusion of a valid inference rule, we also say
+that the judgment is valid or well-formed. In fact, the only way to construct
+new judgments is via the valid inference rules we describe in this section.
+
+The class of valid inference rules is closed under composition and under
+rearranging the hypotheses. That
 means that if
@@
 enable ourselves also to speak of terms of contexts. A term of a context
-$\Gamma$ is a term of the family $\Gamma$ over the empty context. 
+$\Gamma$ is a term of the family $\Gamma$ over the empty context. These ideas
+are captured in the following convertibility rules:
 \begin{align}
@@
 
+Because a context is the same thing as a family of contexts over the empty
+context, we may introduce other type theoretic operations by only describing
+operations on families of contexts. These will then automatically translate
+to opearations on contexts as well. This and the notion of terms of a context
+give the empty context a fundamental role in the theory of contexts, families
+and terms and we feel that the purpose of the empty context should not be
+diminished to mere esthetical reasons, as might be the case in the setting
+of models of the theory \cite{Awodey2013}.
+
 \subsubsection{Extension}
-We introduce extension which not only extends a context $\Gamma$ and a family
-$A$ over it to a context $\ctxext{\Gamma}{A}$, but which also extends a family $A$
-in context $\Gamma$ and a family $P$ over it to a family $\ctxext{A}{P}$ over context
-$\Gamma$. We do this to ensure that all of the theory of contexts, families and
-terms can be done in a context.
-For instance, we could say (1) that a context in context $\Gamma$ is the same thing
-as a family over $\Gamma$; (2) When $A$ is a context in this sense, a family over
-$A$ is the same thing as a family $P$ over $\ctxext{\Gamma}{A}$ and 
-(3) when $P$ is a family over $A$ in this sense, a term of $P$ keeps its original meaning.
+The operation of extension has its place right at the core of the theory of
+contexts, families and terms since it enables one to consider families over
+families and higher level families alike.
+
+We will need the extension operation at the level of contexts and families 
+over them and at the level of families and families over them. 
 \begin{align}
@@
 \subsubsection{Weakening}
-When $A$ is a context family over a context $\Gamma$, we wish to define a weakening
-operation $\ctxwk{A}{}$. The weakening operation acts on context families $B$ 
-over $\Gamma$, terms thereof, context families over $B$ and terms thereof.
-It weakens those, which means that it ``adds $A$ to the context''. The context
-family $\ctxwk{A}{B}$ can be seen as the constant family $B$ over $\ctxext{\Gamma}{A}$.
-Likewise, when $y$ is a term of $B$, the term $\ctxwk{A}{y}$ of $\ctxwk{A}{B}$
-can be seen as the constant term with value $y$.
-
- In the following inference rules we assume that $\jfam{\Gamma}{A}$ and in the
- rules asserting a judgmental equality we assume furthermore that 
- $\jfameq{\Gamma}{A}{A'}$.
- \begin{align}
-& \inference
-  { \jfam{\Gamma}{B}
+When $A$ is a family in context $\Gamma$, the operation of weakening by $A$
+takes a family $B$ in context $\Gamma$ and provides a family $\ctxwk{A}{B}$
+in context $\ctxext{\Gamma}{A}$. The context family $\ctxwk{A}{B}$ can be seen
+as the constant family over $\ctxext{\Gamma}{A}$ with value $B$. This idea will
+be axiomatized in the cancellation property of weakening and substitution in
+\autoref{cancellation-ws}.
+There is also an action of weakening by $A$
+on families $Q$ over $\ctxext{\Gamma}{B}$ -- the weakened family $\ctxwk{A}{Q}$
+will be a family over $\ctxext{{\Gamma}{A}}{\ctxwk{A}{B}}$ -- and likewise
+there is an action on terms of those families.
+
+We will not need to postulate an action of weakening on terms $y$ of families 
+$B$ to provide for the `constant terms' $\ctxwk{A}{y}$ of type
+$\ctxwk{A}{B}$. These
+will be obtained from the action of weakening on terms described above once
+we have established the judgmental equality $\ctxwk{A}{\emptyf}\jdeq\emptyf$
+of families over $\ctxext{\Gamma}{A}$ in \autoref{comp-w0}.
+\begin{align}
+& \inference
+  { \jfam{\Gamma}{A}
+    \jfam{\Gamma}{B}
     }
@@
 & & \inference
-    { \jfameq{\Gamma}{B}{B'}
+    { \jfameq{\Gamma}{A}{A'}
+      \jfameq{\Gamma}{B}{B'}
       }
@@
 & \inference
-  { \jfam{{\Gamma}{B}}{Q}
+  { \jfam{\Gamma}{A}
+    \jfam{{\Gamma}{B}}{Q}
     }
@@
 & & \inference
-    { \jfameq{{\Gamma}{B}}{Q}{Q'}
+    { \jfameq{\Gamma}{A}{A'}
+      \jfameq{{\Gamma}{B}}{Q}{Q'}
       }
@@
     \\
-& \inference
-  { \jterm{\Gamma}{B}{y}
-    }
-  { \jterm{{\Gamma}{A}}{\ctxwk{A}{B}}{\ctxwk{A}{y}}
-    }
-& & \inference
-    { \jtermeq{\Gamma}{B}{y}{y'}
-      }
-    { \jtermeq
-        {{\Gamma}{A}}
-        {\ctxwk{A}{B}}
-        {\ctxwk{A}{y}}
-        {\ctxwk{A'}{y'}}
-      }
-    \\
-& \inference
-  { \jterm{{\Gamma}{B}}{Q}{g}
+%& \inference
+%  { \jterm{\Gamma}{B}{y}
+%    }
+%  { \jterm{{\Gamma}{A}}{\ctxwk{A}{B}}{\ctxwk{A}{y}}
+%    }
+%& & \inference
+%    { \jtermeq{\Gamma}{B}{y}{y'}
+%      }
+%    { \jtermeq
+%        {{\Gamma}{A}}
+%        {\ctxwk{A}{B}}
+%        {\ctxwk{A}{y}}
+%        {\ctxwk{A'}{y'}}
+%      }
+%    \\
+& \inference
+  { \jfam{\Gamma}{A}
+    \jterm{{\Gamma}{B}}{Q}{g}
     }
@@
 & & \inference
-    { \jtermeq{{\Gamma}{B}}{Q}{g}{g'}
+    { \jfameq{\Gamma}{A}{A'}
+      \jtermeq{{\Gamma}{B}}{Q}{g}{g'}
       }
@@
 \end{align}
+In \autoref{morphisms} we will take the terms $\unfold{\jhom{\Gamma}{A}{B}{f}}$
+to be the morphisms of families from $A$ to $B$. These will be at the heart of
+the categorical structure of the theory.
 
@@
 
-\subsubsection{Weakening is compatible with the empty context}
+\subsubsection{Weakening is compatible with the empty context}\label{comp-w0}
 The following rules express that when the empty context or context family is
@@
 \subsubsection{The cancellation property of weakening and substitution}
+\label{cancellation-ws}
 The judgmental equalities we're about to describe assert that substituting a term

### 2014-04-09 article-models-ttderived.tex
added explanations
rewriting some of the section on the operators of CFT theory

@@
 we will see here all of the features of the theory we develop in
-\autoref{judgments} through \autoref{categorical_properties} come to the 
+\autoref{tt} come to the 
 surface.
@@
 
-\subsection{Morphisms}\label{categorical_properties}
+\subsection{Morphisms}\label{morphisms}
 Using the rules of the compatibility of substitution with weakening and of the

### 2014-04-10 article-models-tt.tex
working on section 4.4

@@
 the types in the empty context as the objects.
-
-To get the overview of the compatibility rules, we list the sections
-where these compatibility rules are described in the following table (in this
-table, the
-subsection mentioned in row $X$ and column $Y$ consideres the rules of the
-operation $Y\circ X$):
-\begin{center}
-\begin{tabular}{r|ccc}
-& extension & weakening & substitution\\
-\hline
-extension & \autoref{comp-ee} & \autoref{comp-ew} & \\
-weakening & \autoref{comp-we} & \autoref{comp-ww} & \autoref{comp-ws}\\
-substitution & \autoref{comp-se} & \autoref{comp-sw} & \autoref{comp-ss}
-\end{tabular}
-\end{center}
 
@@
 \subsubsection{The empty context}
+\label{empty}
 In the theory of contexts, families and terms we introduce an empty context and,
@@
 \subsubsection{Extension}
+\label{extension}
 The operation of extension has its place right at the core of the theory of
@@
 \subsubsection{Weakening}
+\label{weakening}
 When $A$ is a family in context $\Gamma$, the operation of weakening by $A$
@@
 \subsubsection{Substitution}
+\label{substitution}
 Given a family $P$ over $A$ and a term $x$ of $A$, substitution gives a way to
-consider the fiber $\subst{x}{P}$ of $P$ at $x$. Also, we get a way to evaluate
-terms $f$ of $P$ at $x$. This will give us ways to compose morphisms too.
-In the rules introducing the various substitutions we assume $\jterm{\Gamma}{A}{x}$;
-in the rules introducing the definitional equalities we assume $\jtermeq{\Gamma}{A}{x}{x'}$.
-\begin{align}
-& \inference
-  { \jfam{{\Gamma}{A}}{P}
+consider the \emph{fiber $\subst{x}{P}$ of $P$ at $x$}. The substitution
+operation furthermore restricts families $Q$ over $\ctxext{{\Gamma}{A}}{P}$ to
+families $\subst{x}{Q}$ over $\ctxext{\Gamma}{\subst{x}{P}}$ and likewise it
+restricts terms $g$ of such $Q$ to terms $\subst{x}{g}$ of $\subst{x}{Q}$.
+
+As was the case with weakening, we will get to apply the substitution operation
+to terms $f$ of $P$ as well, but only after we have established the judgmental
+equality $\subst{x}{\emptyf}\jdeq\emptyf$ of families over the context $\Gamma$.
+We will state this inference rule in \autoref{comp-s0}.
+\begin{align}
+& \inference
+  { \jterm{\Gamma}{A}{x}
+    \jfam{{\Gamma}{A}}{P}
     }
@@
 & & \inference
-    { \jfameq{{\Gamma}{A}}{P}{P'}
+    { \jtermeq{\Gamma}{A}{x}{x'}
+      \jfameq{{\Gamma}{A}}{P}{P'}
       }
@@
 & \inference
-  { \jfam{{{\Gamma}{A}}{P}}{Q}
+  { \jterm{\Gamma}{A}{x}
+    \jfam{{{\Gamma}{A}}{P}}{Q}
     }
@@
 & & \inference
-    { \jfameq{{{\Gamma}{A}}{P}}{Q}{Q'}
+    { \jtermeq{\Gamma}{A}{x}{x'}
+      \jfameq{{{\Gamma}{A}}{P}}{Q}{Q'}
       }
@@
     \\
-& \inference
-  { \jterm{{\Gamma}{A}}{P}{f}
-    }
-  { \jterm{\Gamma}{\subst{x}{P}}{\subst{x}{f}}
-    }
-& & \inference
-    { \jtermeq{{\Gamma}{A}}{P}{f}{f'}
-      }
-    { \jtermeq{\Gamma}{\subst{x}{P}}{\subst{x}{f}}{\subst{x'}{f'}}
-      }
-    \\
-& \inference
-  { \jterm{{{\Gamma}{A}}{P}}{Q}{g}
+%& \inference
+%  { \jterm{{\Gamma}{A}}{P}{f}
+%    }
+%  { \jterm{\Gamma}{\subst{x}{P}}{\subst{x}{f}}
+%    }
+%& & \inference
+%    { \jtermeq{{\Gamma}{A}}{P}{f}{f'}
+%      }
+%    { \jtermeq{\Gamma}{\subst{x}{P}}{\subst{x}{f}}{\subst{x'}{f'}}
+%      }
+%    \\
+& \inference
+  { \jterm{\Gamma}{A}{x}
+    \jterm{{{\Gamma}{A}}{P}}{Q}{g}
     }
@@
 & & \inference
-    { \jtermeq{{{\Gamma}{A}}{P}}{Q}{g}{g'}
+    { \jtermeq{\Gamma}{A}{x}{x'}
+      \jtermeq{{{\Gamma}{A}}{P}}{Q}{g}{g'}
       }
@@
 \end{align}
+In \autoref{morphisms} we will use a combination of weakening and substitution
+to define composition of morphisms of families. However, we have to rely
+on the cancellation rule stated in \autoref{cancellation-ws} before we can
+meaningfully state the definition of composition.
 
 \subsubsection{Identity terms}
+\label{identityterms}
 Last, but not least, we introduce the identity terms. In combination with
@@
 are going to be $25$ finite sets of rules explaining how each operation is
-compatible with each of the operations. In addition to these $25$ finite sets
+compatible with each of the operations.
+
+In addition to these $25$ finite sets
 of compatibility rules, there are also two sets of rules explaining the
@@
 
-\subsubsection{The empty context is compatible with itself}\label{comp-00}
+A typical compatibility axiom captures the idea that a diagram like
+\begin{equation*}
+\begin{tikzcd}[column sep=huge]
+\jfam{\Gamma}{\blank} 
+  \ar{r}{C\mapsto\ctxwk{B}{C}} 
+  \ar{d}[swap]{C\mapsto\ctxwk{A}{C}} 
+& \jfam{{\Gamma}{B}}{\blank} 
+  \ar{d}{Q\mapsto\ctxwk{A}{Q}}
+  \\
+\jfam{{\Gamma}{A}}{\blank} 
+  \ar{r}[swap]{P\mapsto\ctxwk{{A}{B}}{P}} 
+& \jfam{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\blank}
+\end{tikzcd}
+\end{equation*}
+in the case of two consecutive weakenings, or
+\begin{equation*}
+\begin{tikzcd}[column sep=huge]
+\jfam{{\Gamma}{A}}{\blank} 
+  \ar{d}[swap]{Q\mapsto\subst{x}{Q}} 
+  \ar{r}{Q\mapsto\ctxwk{P}{Q}} 
+& \jfam{{{\Gamma}{A}}{P}}{\blank} 
+    \ar{d}{R\mapsto\subst{x}{R}}
+  \\ 
+\jfam{\Gamma}{\blank} 
+  \ar{r}[swap]{B\mapsto\ctxwk{\subst{x}{P}}{B}} 
+& \jfam{{\Gamma}{\subst{x}{P}}}{\blank}
+\end{tikzcd}
+\end{equation*}
+in the case of a combination of weakening and substitution, commutes. We won't
+actually draw any of these diagrams in this section, but readers are encouraged
+to do so if they think it can help them understanding the axioms. Another thing
+one could keep in mind with the axioms that involve weakening is that all of
+them become completely obvious when weakening is made notationally invisible.
+Of course, we have an explicit notation for weakening on purpuse, but it is the
+usual practice of type theorists to make weakening invisible.
+
+The table below indicates where
+we can find the axioms of each of the combinations of operators. The reference
+at the row of operation $X$ and the column of operation $Y$ refers to the
+axiom which explains $Y$ distributes over the arguments of $X$ when $Y$ is
+applied after $X$. The empty context has no arguments, whereas extension,
+and weakening have two arguments which are both of the family-kind, substitution
+has two arguments of which one is a term and the other is a family and
+idenitity terms have one argument, a family.
+\begin{center}
+\begin{tabular}{r|rrrrr}
+& emp & ext & wk & subst & id\\
+\hline
+empty context & \autoref{comp-00} & \autoref{comp-0e}* & \autoref{comp-0w} & \autoref{comp-0s} & \xmark \\
+extension & \autoref{comp-e0}* & \autoref{comp-ee}* & \autoref{comp-ew} & \cmark & \cmark \\
+weakening & \autoref{comp-w0} & \autoref{comp-we}* & \autoref{comp-ww} & \autoref{comp-sw} & \autoref{comp-wi} \\
+substitution & \autoref{comp-s0} & \autoref{comp-se}* & \autoref{comp-ws} & \autoref{comp-ss} & \autoref{comp-si} \\
+identity terms & \cmark & \cmark & \autoref{comp-wi} & \autoref{comp-si} & \xmark
+\end{tabular}
+\end{center}
+Apart from references to sections where axioms are given, we also see checkmarks,
+crosses, marked references and duplicated references in this table. 
+The checkmarks indicate that the desired property
+can be proved and need not be axiomatized. To clarify those:
+\begin{itemize}
+\item The identity term of the empty context is the unique term of the empty
+family. Recall that we already have axiomatized in \autoref{empty} that every
+term of the empty family is judgmentally equal to $\emptytm$.
+\item The identity term of an extension can be explained by means of the notion
+of \emph{extension on terms}. This notion is defined in \autoref{extension-on-terms}
+using only identity terms and substitution. With extension on terms we can
+combine a term $\jterm{\Gamma}{A}{x}$ and a term $\jterm{\Gamma}{\subst{x}{P}}{u}$
+into a term $\jterm{\Gamma}{{A}{P}}{\tmext{x}{u}}$, which may be seen as the
+pair consisting of $x$ and $u$. Since it is a defined notion, we will
+be able to derive lots of properties for extension of terms. One of them explains
+how we can see the identity term of an extension as the extension of two certain
+terms, see \autoref{comp-ie}.
+\item Using extension on terms we can also make sense of `fibers at an extension'
+because that would exactly be the fiber at an extended term. We will see in
+\autoref{comp-es} that, for instance, $\subst{\tmext{x}{u}}{Q}\jdeq\subst{u}{{x}{Q}}$
+for any family $\jfam{{{\Gamma}{A}}{P}}{Q}$ and any two terms $\jterm{\Gamma}{A}{x}$
+and $\jterm{\Gamma}{\subst{x}{P}}{u}$. That's a form of currying.
+\end{itemize}
+The spots with a \xmark simply make no sense. The marked references point to
+axioms which \emph{do} have a variant for terms, but those require the notion
+of extension on terms, can be proved and therefore need not be included in our
+list of compatibility axioms. In each of these sections, we give a forward
+reference to the lemma in which this variant is stated and proved.
+
+\subsubsection{The empty context as a family over the empty context}
+\label{comp-00}
 Because every context is a family of contexts over the empty context, 
@@
 
-\subsubsection{Extension is compatible with the empty context}\label{comp-e0}
-The following rule asserts that extension by $\emptyc$ leaves the contexts unchanged.
-\begin{align}
-& \inference
-  { \jctx{\Gamma}
-    }
-  { \jctxeq{\ctxext{\emptyc}{\Gamma}}{\Gamma}
-    }
-  \label{eq:comp-e0}\\
-& \inference
-  { \jctx{\Gamma}
-    }
-  { \jctxeq{\ctxext{\Gamma}{\emptyf}}{\Gamma}
-    }
-  \\
+\subsubsection{Extensions of the empty context}
+\label{comp-e0}
+\begin{align}
+%& \inference
+%  { \jctx{\Gamma}
+%    }
+%  { \jctxeq{\ctxext{\Gamma}{\emptyf}}{\Gamma}
+%    }
+%  \label{comp-e0-c}\\
+& \inference
+  { \jfam{\Gamma}{A}
+    }
+  { \jfameq{\Gamma}{{A}{\emptyf}}{A}
+    }
+  \label{comp-e0-f}
+\end{align}
+
+\subsubsection{Extension by the empty context}
+\label{comp-0e}
+\begin{align}
+%& \inference
+%  { \jctx{\Gamma}
+%    }
+%  { \jctxeq{\ctxext{\emptyc}{\Gamma}}{\Gamma}
+%    }
+%  \label{comp-0e-c}\\
 & \inference
@@
     }
-  \\
-& \inference
-  { \jfam{\Gamma}{A}
-    }
-  { \jfameq{\Gamma}{{A}{\emptyf}}{A}
-    }
-\end{align}
-
-\subsubsection{Weakening is compatible with the empty context}\label{comp-w0}
+  \label{comp-0e-f}
+\end{align}
+
+\subsubsection{Weakenings of the empty context}
+\label{comp-w0}
 The following rules express that when the empty context or context family is
@@
 \begin{align}
-& \inference
-  { \jctx{\Gamma}
-    }
-  { \jfameq{\Gamma}{\ctxwk{\Gamma}{\emptyc}}{\emptyf}
-    }
-  \\
+%& \inference
+%  { \jctx{\Gamma}
+%    }
+%  { \jfameq{\Gamma}{\ctxwk{\Gamma}{\emptyc}}{\emptyf}
+%    }
+%  \\
 & \inference
@@
 \end{align}
+Note that we do not have to axiomatize the action of weakening by $A$ on the term
+of the empty family, because the result will be in the empty family and every
+term in the empty family is judgmentally equal to the term $\emptytm$.
+
+\subsubsection{Weakening by the empty context}
+\label{comp-0w}
 Weakening by the empty family $\emptyf$ over a context $\Gamma$ leaves families, 
@@
 & \inference
-  { \jterm{\Gamma}{B}{y}
-    }
-  { \jtermeq{\Gamma}{B}{\ctxwk{\emptyf}{y}}{y}
-    }
-  \\
-& \inference
   { \jfam{{\Gamma}{B}}{Q}
@@
 
-\subsubsection{Substitution is compatible with the empty context}
+\subsubsection{Substitutions of the empty context}
+\label{comp-s0}
 The fibers of the empty family are the empty families:
@@
 \end{align}
-
+Note again that we do not have to axiomatize the action of substitution by $x$ on the term
+of the empty family, because the result will be in the empty family and every
+term in the empty family is judgmentally equal to the term $\emptytm$.
+
+\subsubsection{Substitution by the term of the empty context}
+\label{comp-0s}
 The following rules assert that substituting by the term $\jterm{\Gamma}{\emptyf}{\emptytm}$
@@
   \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    }
-  { \jtermeq{\Gamma}{A}{\subst{\emptytm}{x}}{x}
-    }
-  \\
+%& \inference
+%  { \jterm{\Gamma}{A}{x}
+%    }
+%  { \jtermeq{\Gamma}{A}{\subst{\emptytm}{x}}{x}
+%    }
+%  \\
 & \inference
@@
 
-\subsubsection{Extension is compatible with itself}\label{comp-ee}
+\subsubsection{Extensions of extensions are extensions by extensions}
+\label{comp-ee}
 The inference rules asserting that extension is compatible with itself assert
@@
 \begin{align}
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{{\Gamma}{A}}{P}
-    }
-  { \jctxeq{\ctxext{{\Gamma}{A}}{P}}{\ctxext{\Gamma}{{A}{P}}}
-    }
-  \\
+%& \inference
+%  { \jfam{\Gamma}{A}
+%    \jfam{{\Gamma}{A}}{P}
+%    }
+%  { \jctxeq{\ctxext{{\Gamma}{A}}{P}}{\ctxext{\Gamma}{{A}{P}}}
+%    }
+%  \\
 & \inference
@@
     }
-  { \jfameq{\Gamma}{\ctxext{{A}{P}}{Q}}{\ctxext{A}{{P}{Q}}}
-    }
-\end{align}
-
-\subsubsection{Weakening is compatible with extension}\label{comp-we}
-
-The following rules assert the compatibility of extension with weakening: for
-every family $A$ over $\Gamma$ and every family $Q$ over $\ctxext{\Gamma}{B}$
-there is a
-judgmental equality $\ctxwk{A}{\ctxext{B}{Q}}\jdeq\ctxext{\ctxwk{A}{B}}
-{\ctxwk{A}{Q}}$.
-
-When thinking of terms of $\ctxwk{A}{B}$ as morphisms of families from $A$ to
-$B$, this looks already like form of type theoretic choice. It is weaker in that
-it is not stated with function types, yet it is stronger in that it states a
-judgmental equality between two families. When one makes the weakening operation
-notationally invisible -- as is in fact the usual practice in type theory -- the
-following compatibility rules become completely obvious.
-
-In the following inference rules we assume that $\jfam{\Gamma}{A}$.
-\begin{align}
-& \inference
-  { \jfam{{{\Gamma}{B}}{Q}}{R}
+  { \jfameq{\Gamma}{\ctxext{A}{{P}{Q}}}{\ctxext{{A}{P}}{Q}}
+    }
+  \label{comp-ee-c}
+\end{align}
+Note that we may take the point of view that extension distributes over
+extensions. If we want to extend an extended family $\ctxext{P}{Q}$ over
+$\ctxext{\Gamma}{A}$ by the family $A$, the result is the extension of (1)
+the extension of $P$ by $A$ and (2) the extension of $Q$ by $A$. The latter is
+simply $Q$ since extension acts trivially at the level of families (and terms).
+
+\subsubsection{Weakenings of extensions}
+\label{comp-we}
+The following rules assert that when an extended family is weakened, the
+weakening distributes over the extension factors.
+\begin{align}\label{comp-we-f}
+& \inference
+  { \jfam{\Gamma}{A}
+    \jfam{{{\Gamma}{B}}{Q}}{R}
     }
@@
 \end{align}
-
-\subsubsection{Weakening is compatible with itself}\label{comp-ww}
-We state judgmental equality rules expressing
-that weakening is compatible with itself. These rules state that the following
-diagram commutes given any two families $A$ and $B$ in context $\Gamma$:
-\begin{equation*}
-\begin{tikzcd}[column sep=huge]
-\jfam{\Gamma}{\blank} 
-  \ar{r}{C\mapsto\ctxwk{B}{C}} 
-  \ar{d}[swap]{C\mapsto\ctxwk{A}{C}} 
-& \jfam{{\Gamma}{B}}{\blank} 
-  \ar{d}{Q\mapsto\ctxwk{A}{Q}}
-  \\
-\jfam{{\Gamma}{A}}{\blank} 
-  \ar{r}[swap]{P\mapsto\ctxwk{{A}{B}}{P}} 
-& \jfam{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\blank}
-\end{tikzcd}
-\end{equation*}
-Thus, we get the following set of inference rules:
+When thinking of terms of $\ctxwk{A}{B}$ as morphisms of families from $A$ to
+$B$, this looks already like form of type theoretic choice. It is weaker in that
+it is not stated with function types, yet it is stronger in that it states a
+judgmental equality between two families.
+
+There is also a version of this property where an extended term is weakened.
+This variant is stated and proved in \autoref{comp-we-t}.
+
+\subsubsection{Weakening of weakenings}
+When a weakened family is weakened even further, the outer weakening distributes
+over the `weakening factors'.
+\label{comp-ww}
 \begin{align}
@@
 
-\subsubsection{Extension is compatible with weakening}\label{comp-ew}
+\subsubsection{Weakening by extensions}
+\label{comp-ew}
 The rules expressing that extension is compatible with weakening assert that
@@
 appropriate way.
-
-In the following inference rules we assume that
-$\jfam{\Gamma}{A}$ and $\jfam{{\Gamma}{A}}{P}$.
-\begin{align}
-& \inference
-  { \jfam{{\Gamma}{B}}{Q}
+\begin{align}
+& \inference
+  { \jfam{\Gamma}{A}
+    \jfam{{\Gamma}{A}}{P}
+    \jfam{{\Gamma}{B}}{Q}
     }
@@
 & \inference
-  { \jterm{{\Gamma}{B}}{Q}{g}
+  { \jfam{\Gamma}{A}
+    \jfam{{\Gamma}{A}}{P}
+    \jterm{{\Gamma}{B}}{Q}{g}
     }
@@
 
-\subsubsection{Substitution is compatible with extension}\label{comp-se}
-Suppose $\jterm{\Gamma}{A}{x}$ in all of the following inference rule.
-\begin{align}
-& \inference
-  { \jfam{{{{\Gamma}{A}}{P}}{Q}}{R}
+\subsubsection{Fibers of extensions}
+\label{comp-se}
+The following inference rule asserts that if we take the fiber of an extended
+family at a term $x$ of $A$ in context $\Gamma$, the substitution by $x$
+distributes over the factors of the extension.
+\begin{align}\label{comp-se-f}
+& \inference
+  { \jterm{\Gamma}{A}{x}
+    \jfam{{{{\Gamma}{A}}{P}}{Q}}{R}
     }
@@
 \end{align}
-
-\subsubsection{Substitution is compatible with weakening}\label{comp-sw}
-The rules asserting the compatibility of substitution with weakening assert
-that the following diagram commutes for any $\jterm{\Gamma}{A}{x}$ and any
-$\jfam{{\Gamma}{A}}{P}$:
-\begin{equation*}
-\begin{tikzcd}[column sep=huge]
-\jfam{{\Gamma}{A}}{\blank} 
-  \ar{d}[swap]{Q\mapsto\subst{x}{Q}} 
-  \ar{r}{Q\mapsto\ctxwk{P}{Q}} 
-& \jfam{{{\Gamma}{A}}{P}}{\blank} 
-    \ar{d}{R\mapsto\subst{x}{R}}
-  \\ 
-\jfam{\Gamma}{\blank} 
-  \ar{r}[swap]{B\mapsto\ctxwk{\subst{x}{P}}{B}} 
-& \jfam{{\Gamma}{\subst{x}{P}}}{\blank}
-\end{tikzcd}
-\end{equation*}
-We plug in an extra layer of families to cover the most general case at once.
-In the following inference rules we assume that $\jterm{\Gamma}{A}{x}$ and
-$\jfam{{\Gamma}{A}}{P}$:
-\begin{align}
-& \inference
-  { \jfam{{{\Gamma}{A}}{Q}}{R}
+There is also a version of this statement in which extended terms are considered.
+This variant is stated and proved in \autoref{comp-se-t}.
+
+\subsubsection{Fibers of weakenings}\label{comp-sw}
+Substitutions distribute over weakenings:
+\begin{align}
+& \inference
+  { \jterm{\Gamma}{A}{x}
+    \jfam{{\Gamma}{A}}{P}
+    \jfam{{{\Gamma}{A}}{Q}}{R}
     }
@@
 & \inference
-  { \jterm{{{\Gamma}{A}}{Q}}{R}{h}
+  { \jterm{\Gamma}{A}{x}
+    \jfam{{\Gamma}{A}}{P}
+    \jterm{{{\Gamma}{A}}{Q}}{R}{h}
     }
@@
 
-\subsubsection{Substitution is compatible with substitution}\label{comp-ss}
+\subsubsection{Fibers of fibers}
+\label{comp-ss}
 
@@
 
-\subsubsection{Weakening is compatible with substitution}\label{comp-ws}
+\subsubsection{Weakenings of fibers}
+\label{comp-ws}
 We already have rules for the compatibility of substitution with weakening, but
@@
 
-\subsubsection{The cancellation properties of identity terms}
-Identity terms are determined by their behavior with respect to substitution combined with
-weakening. The identity terms will also be subject to compatibility rules.
-\begin{align}
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    }
-  { \jtermeq{\Gamma}{A}{\subst{x}{\idtm{A}}}{x}
-    }
-  \label{idfunc-subst-defn}\\
-& \inference
-  { \jfam{{\Gamma}{A}}{P}
-    }
-  { \jfameq{{\Gamma}{A}}{\subst{\idtm{A}}{\ctxwk{A}{P}}}{P}
-    }
-  \label{idfunc-wk-defn}\\
-& \inference
-  { \jfam{{\Gamma}{A}}{P}
-    }
-  { \jfameq
-      {{\Gamma}{A}}
-      {\subst{\idtm{A}}{\ctxwk{{A}{A}}{P}}}
-      {P}
-    }
-  \label{idfunc-wk-defn2}\\
-& \inference
-  { \jterm{{\Gamma}{A}}{P}{f}
-    }
-  { \jtermeq
-      {{\Gamma}{A}}
-      {P}
-      {\subst{\idtm{A}}{\ctxwk{A}{f}}}
-      {f}
-    }
-  \label{idfunc-precomp}\\
-& \inference
-  { \jterm{{\Gamma}{A}}{P}{f}
-    }
-  { \jtermeq
-      {{\Gamma}{A}}
-      {P}
-      {\subst{\idtm{A}}{\ctxwk{{A}{A}}{f}}}
-      {f}
-    }
-  \label{idfunc-precomp}\\
-& \inference
-  { \unfoldall{\jhom{\Gamma}{A}{B}{f}}
-    }
-  { \unfoldall{\jhomeq{\Gamma}{A}{B}{\jcomp{A}{f}{\idtm{B}}}{f}}
-    }
-  \label{idfunc-postcomp}
-\end{align}
-
-We won't state a compatibility rule stating that the identity term is
-compatible with extension because we will be able to prove that. Instead, we
-will just state the compatibility rules for the identity term combined with
-weakening and with substitution.
-
+\subsubsection{The identity term of a weakened family}
+\label{comp-wi}
 The identity term of a weakened family is the weakened identity term:
@@
 
+\subsubsection{The identity term of a substituted family}
+\label{comp-si}
 The identity term of a substituted family is the substitution of the identity term
@@
 
+\subsubsection{The cancellation properties of identity terms}
+\label{cancellation-i}
+Identity terms are determined by their behavior with respect to substitution combined with
+weakening. The identity terms will also be subject to compatibility rules.
+\begin{align}
+& \inference
+  { \jterm{\Gamma}{A}{x}
+    }
+  { \jtermeq{\Gamma}{A}{\subst{x}{\idtm{A}}}{x}
+    }
+  \label{idfunc-subst-defn}\\
+& \inference
+  { \jfam{{\Gamma}{A}}{P}
+    }
+  { \jfameq{{\Gamma}{A}}{\subst{\idtm{A}}{\ctxwk{A}{P}}}{P}
+    }
+  \label{idfunc-wk-defn}\\
+%An unnecessary rule:
+%& \inference
+%  { \jfam{{\Gamma}{A}}{P}
+%    }
+%  { \jfameq
+%      {{\Gamma}{A}}
+%      {\subst{\idtm{A}}{\ctxwk{{A}{A}}{P}}}
+%      {P}
+%    }
+%  \label{idfunc-wk-defn2}\\
+& \inference
+  { \jterm{{\Gamma}{A}}{P}{f}
+    }
+  { \jtermeq
+      {{\Gamma}{A}}
+      {P}
+      {\subst{\idtm{A}}{\ctxwk{A}{f}}}
+      {f}
+    }
+  \label{idfunc-precomp}\\
+%An unnecessary rule:
+%& \inference
+%  { \jterm{{\Gamma}{A}}{P}{f}
+%    }
+%  { \jtermeq
+%      {{\Gamma}{A}}
+%      {P}
+%      {\subst{\idtm{A}}{\ctxwk{{A}{A}}{f}}}
+%      {f}
+%    }
+%  \label{idfunc-precomp}\\
+& \inference
+  { \unfoldall{\jhom{\Gamma}{A}{B}{f}}
+    }
+  { \unfoldall{\jhomeq{\Gamma}{A}{B}{\jcomp{A}{f}{\idtm{B}}}{f}}
+    }
+  \label{idfunc-postcomp}
+\end{align}
+
 \subsection{The possiblity of types in the theory of contexts, families and
 terms}
+\label{types}
 We have deliberately not spoken of types so far because we have taken the point

### 2014-04-10 article-models-ttderived.tex
working on section 4.4

@@
 
-\begin{lem}
+\begin{lem}\label{comp-es}
 The following inference rules are valid:
@@
 
-\begin{lem}\label{lem:tmext-wk}
+\begin{lem}\label{lem:tmext-wk}\label{comp-we-t}
 When we weaken a term $\tmext{B}{Q}{y}{v}$ of $\ctxext{B}{Q}$ in context $\Gamma$ by
@@
 
-\begin{lem}\label{lem:tmext-subst}
+\begin{lem}\label{lem:tmext-subst}\label{comp-se-t}
 When we substitute an extended term $\tmext{P}{Q}{f}{g}$ of $\ctxext{P}{Q}$ by a term
@@
 
-\begin{lem}\label{lem:tmext-id}
+\begin{lem}\label{lem:tmext-id}\label{comp-ie}
 For any $\jfam{\Gamma}{A}$ and $\jfam{{\Gamma}{A}}{P}$ we have

### 2014-04-12 article-models-tt.tex
small improvisations

@@
 \end{align}
+The extension operation may also be defined to act on families over families
+and terms thereof. For instance, When $\jfam{{\Gamma}{A}}{P}$ is a family,
+we would get a family $\jfam{{\Gamma}{A}}{{\Gamma}{P}}$ and when
+$\jterm{{\Gamma}{A}}{P}{f}$ is a term, we would get a term
+$\jterm{{\Gamma}{A}}{{\Gamma}{P}}{\ctxext{\Gamma}{f}}$. Then there would be
+axioms stating the judgmental equalities $\ctxext{\Gamma}{P}\jdeq P$ and
+$\ctxext{\Gamma}{f}\jdeq f$. Thus, it is of no actual use to introduce the
+action of context extension on families and terms.
+
+Nevertheless, the fact that extension \emph{does} act on families
+and terms (albeit trivially) reflects in the internal extension term in the type theory of
+models of type theory, which is developed in \autoref{ttmodels}. Also, in the
+categorical semantics, the extension operation would be the concatenation of
+lists of arrows belonging to the class of arrows representing families. Since
+families over families are interpreted by lists of those family arrows extending
+concateneted lists of family arrows, the action of concatenation of such lists
+may be taken to be trivial. The $\Sigma$-functor, which is the version of this
+operation on those lists which have length one, acts trivially on morphisms;
+another manifestation of the same phenomenon.
 
@@
 \end{itemize}
-The spots with a \xmark simply make no sense. The marked references point to
+The spots with a \xmark simply make no sense 
+{\color{red}(or do they, when the term model
+is better understood (by me)?)}. The marked references point to
 axioms which \emph{do} have a variant for terms, but those require the notion

### 2014-04-30 article-models-tt.tex
added stuff to intro of 4.3

@@
 Apart from the empty context, there will indeed be an empty family and a term
-of the empty family (which will be deemed unique). But context extension will
+of the empty family (which will be deemed unique). But extension will
 not have an explicit action `at the level of families' since extension
 is used to define higher families and the action would have to be trivial
-anyway. The operations of weakening and substitution are actually introduced
-by four inference rules and the identity terms by one.
+anyway. The empty context and extension provide the essentials of the rest of
+the theory in the sense that higher families are families over extensions,
+contexts are families over the empty context and the `action on families' of
+any operation applied to families over the empty context is the same as the
+original operation. Thus, when we have introduced the operations of weakening
+and substitution we will see that they also have an `action on the terms of 
+contexts'. We will state the relevant judgmental equality rules for the empty
+context in this section rather than in the next, since they are required to
+state that the action on families of an operation is compatible with the
+original operation.
 

### 2014-04-30 article-models-tt.tex
edited the weakening subsubsection

@@
 
+\label{comp-00}
+Because every context is a family of contexts over the empty context, 
+the empty context $\emptyc$ may be considered as a family of contexts over the empty
+context. When we do this, we get $\emptyf[\emptyc]$.
+\begin{equation}\label{eq:comp-00}
+\inference
+  { }
+  { \jfameq
+      {\emptyc}
+      {\emptyc}
+      {\emptyf[\emptyc]}
+    }
+\end{equation}
+In the future, we shall denote $\emptyf[\Gamma]$ by $\emptyf$. The above rule
+guarantees that this will not cause confusion. Likewise, we shall denote
+$\emptytm[\Gamma]$ by $\emptytm$.
+
 \subsubsection{Extension}
@@
 
+In \autoref{extension-on-terms} we will also define an extension operation
+on terms, which defines a term $\jterm{\Gamma}{{A}{P}}{\tmext{x}{u}}$ for
+every $\jterm{\Gamma}{A}{x}$ and $\jterm{\Gamma}{\subst{x}{P}}{u}$. This
+operation is defined by using the substitution operation and identity terms.
+
+In the following set of inference rules we state that the empty context and
+the empty family are neutral objects when extension is concerned.
+\label{comp-e0}\label{comp-0e}
+\begin{align}
+%& \inference
+%  { \jctx{\Gamma}
+%    }
+%  { \jctxeq{\ctxext{\Gamma}{\emptyf}}{\Gamma}
+%    }
+%  \label{comp-e0-c}\\
+& \inference
+  { \jfam{\Gamma}{A}
+    }
+  { \jfameq{\Gamma}{{A}{\emptyf}}{A}
+    }
+  \label{comp-e0-f}
+  \\
+%& \inference
+%  { \jctx{\Gamma}
+%    }
+%  { \jctxeq{\ctxext{\emptyc}{\Gamma}}{\Gamma}
+%    }
+%  \label{comp-0e-c}\\
+& \inference
+  { \jfam{\Gamma}{A}
+    }
+  { \jfameq{\Gamma}{\ctxext{\emptyf}{A}}{A}
+    }
+  \label{comp-0e-f}
+\end{align}
+Note the special case where we take $\Gamma$ to be the empty context.
+
 \subsubsection{Weakening}
@@
 be axiomatized in the cancellation property of weakening and substitution in
-\autoref{cancellation-ws}.
-There is also an action of weakening by $A$
-on families $Q$ over $\ctxext{\Gamma}{B}$ -- the weakened family $\ctxwk{A}{Q}$
-will be a family over $\ctxext{{\Gamma}{A}}{\ctxwk{A}{B}}$ -- and likewise
-there is an action on terms of those families.
-
-We will not need to postulate an action of weakening on terms $y$ of families 
-$B$ to provide for the `constant terms' $\ctxwk{A}{y}$ of type
-$\ctxwk{A}{B}$. These
-will be obtained from the action of weakening on terms described above once
-we have established the judgmental equality $\ctxwk{A}{\emptyf}\jdeq\emptyf$
-of families over $\ctxext{\Gamma}{A}$ in \autoref{comp-w0}.
+\autoref{cancellation-ws}. In \autoref{morphisms} we will take the terms 
+$\unfold{\jhom{\Gamma}{A}{B}{f}}$ to be the morphisms of families from $A$ to 
+$B$. These will be at the heart of the categorical structure of the theory.
+
+The weakening operation acts on three levels: on contexts, on families and
+on terms. The `action on contexts' of weakening is the action we described
+above: it takes a family $B$ over $\Gamma$ to a family $\ctxwk{A}{B}$ over
+$\ctxext{\Gamma}{A}$; the `action on families' of weakening takes a family
+$Q$ over $\ctxext{\Gamma}{B}$ to a family $\ctxwk[\famsym]{A}{Q}$ over
+$\ctxext{{\Gamma}{A}}{\ctxwk{A}{B}}$; the `action on terms' of weakening takes
+a term $g$ of $Q$ to a term $\ctxwk[\tmsym]{A}{g}$ of $\ctxwk[\famsym]{A}{Q}$.
 \begin{align}
@@
     }
-  { \jfam{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{Q}}
+  { \jfam{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk[\famsym]{A}{Q}}
     }
@@
         {{{\Gamma}{A}}{\ctxwk{A}{B}}}
-        {\ctxwk{A}{Q}}
-        {\ctxwk{A'}{Q'}}
+        {\ctxwk[\famsym]{A}{Q}}
+        {\ctxwk[\famsym]{A'}{Q'}}
       }
@@
     }
-  { \jterm{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{Q}}{\ctxwk{A}{g}}
+  { \jterm{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk[\famsym]{A}{Q}}{\ctxwk[\tmsym]{A}{g}}
     }
@@
         {{{\Gamma}{A}}{\ctxwk{A}{B}}}
-        {\ctxwk{A}{Q}}
-        {\ctxwk{A}{g}}
-        {\ctxwk{A'}{g'}}
-      }
-\end{align}
-In \autoref{morphisms} we will take the terms $\unfold{\jhom{\Gamma}{A}{B}{f}}$
-to be the morphisms of families from $A$ to $B$. These will be at the heart of
-the categorical structure of the theory.
+        {\ctxwk[\famsym]{A}{Q}}
+        {\ctxwk[\tmsym]{A}{g}}
+        {\ctxwk[\tmsym]{A'}{g'}}
+      }
+\end{align}
+The following rule expresses that when the empty context or context family is
+weakened, the result is the empty context family.
+\label{comp-w0}
+\begin{align}
+%& \inference
+%  { \jctx{\Gamma}
+%    }
+%  { \jfameq{\Gamma}{\ctxwk{\Gamma}{\emptyc}}{\emptyf}
+%    }
+%  \\
+& \inference
+  { \jfam{\Gamma}{A}
+    }
+  { \jfameq{{\Gamma}{A}}{\ctxwk{A}{\emptyf}}{\emptyf}
+    }
+\end{align}
+The special case where $\Gamma$ is taken to be the empty context is again
+noteworthy. Also, note that we do not have to axiomatize the action of weakening
+by $A$ on the term of the empty family, because the result will be in the empty 
+family and every term in the empty family is judgmentally equal to the term 
+$\emptytm$.
+
+Because a family over $\Gamma$ is the same as a family over 
+$\ctxext{\Gamma}{\emptyf}$, we can apply both the action on contexts and the
+action on families of weakening to a family $B$ over $\Gamma$. When we apply
+the action on families, we obtain a family $\ctxwk[\famsym]{A}{B}$ over the
+context $\ctxext{{\Gamma}{A}}{\ctxwk{A}{\emptyf}}$. However, since we have
+postulated the judgmental equalities $\ctxwk{A}{\emptyf}\jdeq\emptyf$ and
+$\ctxext{{\Gamma}{A}}{\emptyf}\jdeq\ctxext{\Gamma}{A}$, we see that we can
+compare $\ctxwk[\famsym]{A}{B}$ with $\ctxwk{A}{B}$. The following inference
+rule postulates that these two are judgmentally equal:
+\begin{equation}
+\inference
+{ \jfam{\Gamma}{A}
+  \jfam{\Gamma}{B}
+  }
+{ \jfameq{{\Gamma}{A}}{\ctxwk[\famsym]{A}{B}}{\ctxwk{A}{B}}
+  }
+\end{equation}
+Due to this rule, the action on contexts and the action on families of weakening
+are compatible with each other and consequently there can be no possible
+confusion when we omit the annotations $\famsym$ and $\tmsym$ alltogether. In
+the future, the weakening of a family $Q$ over $\ctxext{\Gamma}{B}$ shall
+be denoted just by $\ctxwk{A}{Q}$ and likewise the weakening of a term $g$ of
+$Q$ shall be denoted by $\ctxwk{A}{g}$.
+
+Because a family $B$ over $\Gamma$ can be treated as a family by the operation
+of weakening, weakening also acts on the terms of $B$. The weakening of a term
+$y$ of $B$ by $A$ can be seen as the constant term (or function) of the
+family $\ctxwk{A}{B}$.
+
+Note that we can also weaken by the empty family over $\Gamma$ itself.
+Weakening by the empty family $\emptyf$ over a context $\Gamma$ leaves families, 
+their terms, families over those families and terms of those unchanged:
+\label{comp-0w}\begin{align}
+& \inference
+  { \jfam{\Gamma}{B}
+    }
+  { \jfameq{\Gamma}{\ctxwk{\emptyf}{B}}{B}
+    }
+  \\
+& \inference
+  { \jfam{{\Gamma}{B}}{Q}
+    }
+  { \jfameq{{\Gamma}{B}}{\ctxwk{\emptyf}{Q}}{Q}
+    }
+  \\
+& \inference
+  { \jterm{{\Gamma}{B}}{Q}{g}
+    }
+  { \jtermeq{{\Gamma}{B}}{Q}{\ctxwk{\emptyf}{g}}{g}
+    }
+\end{align}
 
@@
 
-\subsubsection{The empty context as a family over the empty context}
-\label{comp-00}
-Because every context is a family of contexts over the empty context, 
-the empty context $\emptyc$ may be considered as a family of contexts over the empty
-context. When we do this, we get $\emptyf[\emptyc]$.
-\begin{equation}\label{eq:comp-00}
-\inference
-  { }
-  { \jfameq
-      {\emptyc}
-      {\emptyc}
-      {\emptyf[\emptyc]}
-    }
-\end{equation}
-In the future, we shall denote $\emptyf[\Gamma]$ by $\emptyf$. The above rule
-guarantees that this will not cause confusion. Likewise, we shall denote
-$\emptytm[\Gamma]$ by $\emptytm$.
-
-\subsubsection{Extensions of the empty context}
-\label{comp-e0}
-\begin{align}
-%& \inference
-%  { \jctx{\Gamma}
-%    }
-%  { \jctxeq{\ctxext{\Gamma}{\emptyf}}{\Gamma}
-%    }
-%  \label{comp-e0-c}\\
-& \inference
-  { \jfam{\Gamma}{A}
-    }
-  { \jfameq{\Gamma}{{A}{\emptyf}}{A}
-    }
-  \label{comp-e0-f}
-\end{align}
-
-\subsubsection{Extension by the empty context}
-\label{comp-0e}
-\begin{align}
-%& \inference
-%  { \jctx{\Gamma}
-%    }
-%  { \jctxeq{\ctxext{\emptyc}{\Gamma}}{\Gamma}
-%    }
-%  \label{comp-0e-c}\\
-& \inference
-  { \jfam{\Gamma}{A}
-    }
-  { \jfameq{\Gamma}{\ctxext{\emptyf}{A}}{A}
-    }
-  \label{comp-0e-f}
-\end{align}
-
-\subsubsection{Weakenings of the empty context}
-\label{comp-w0}
-The following rules express that when the empty context or context family is
-weakened, the result is the empty context family.
-\begin{align}
-%& \inference
-%  { \jctx{\Gamma}
-%    }
-%  { \jfameq{\Gamma}{\ctxwk{\Gamma}{\emptyc}}{\emptyf}
-%    }
-%  \\
-& \inference
-  { \jfam{\Gamma}{A}
-    }
-  { \jfameq{{\Gamma}{A}}{\ctxwk{A}{\emptyf}}{\emptyf}
-    }
-\end{align}
-Note that we do not have to axiomatize the action of weakening by $A$ on the term
-of the empty family, because the result will be in the empty family and every
-term in the empty family is judgmentally equal to the term $\emptytm$.
-
-\subsubsection{Weakening by the empty context}
-\label{comp-0w}
-Weakening by the empty family $\emptyf$ over a context $\Gamma$ leaves families, 
-their terms, families over those families and
-terms of those unchanged:
-\begin{align}
-& \inference
-  { \jfam{\Gamma}{B}
-    }
-  { \jfameq{\Gamma}{\ctxwk{\emptyf}{B}}{B}
-    }
-  \\
-& \inference
-  { \jfam{{\Gamma}{B}}{Q}
-    }
-  { \jfameq{{\Gamma}{B}}{\ctxwk{\emptyf}{Q}}{Q}
-    }
-  \\
-& \inference
-  { \jterm{{\Gamma}{B}}{Q}{g}
-    }
-  { \jtermeq{{\Gamma}{B}}{Q}{\ctxwk{\emptyf}{g}}{g}
-    }
-\end{align}
-
 \subsubsection{Substitutions of the empty context}

### 2014-04-30 article-models-tt.tex
edited the subsubsection of substitution

@@
 \end{align}
-The following rule expresses that when the empty context or context family is
-weakened, the result is the empty context family.
+The following inference rules express that when the empty context or family is
+weakened, the result is the empty family.
 \label{comp-w0}
@@
     }
-\end{align}
-The special case where $\Gamma$ is taken to be the empty context is again
-noteworthy. Also, note that we do not have to axiomatize the action of weakening
+  \\
+& \inference
+  { \jfam{\Gamma}{A}
+    \jfam{\Gamma}{B}
+    }
+  { \jfameq
+    {{{\Gamma}{A}}{\ctxwk{A}{B}}}
+    {\ctxwk[\famsym]{A}{\emptyf}}
+    {\emptyf}
+    }
+\end{align}
+Note that we do not have to axiomatize the action of weakening
 by $A$ on the term of the empty family, because the result will be in the empty 
@@
 Given a family $P$ over $A$ and a term $x$ of $A$, substitution gives a way to
-consider the \emph{fiber $\subst{x}{P}$ of $P$ at $x$}. The substitution
-operation furthermore restricts families $Q$ over $\ctxext{{\Gamma}{A}}{P}$ to
-families $\subst{x}{Q}$ over $\ctxext{\Gamma}{\subst{x}{P}}$ and likewise it
-restricts terms $g$ of such $Q$ to terms $\subst{x}{g}$ of $\subst{x}{Q}$.
-
-As was the case with weakening, we will get to apply the substitution operation
-to terms $f$ of $P$ as well, but only after we have established the judgmental
-equality $\subst{x}{\emptyf}\jdeq\emptyf$ of families over the context $\Gamma$.
-We will state this inference rule in \autoref{comp-s0}.
+consider the \emph{fiber $\subst{x}{P}$ of $P$ at $x$}. As was the case with
+weakening, the substitution operation comes in three parts: the `action on
+contexts' of substitution is the operation just described; the `action on
+families' of substitution takes a family $Q$ over $\ctxext{{\Gamma}{A}}{P}$
+to a family $\subst[\famsym]{x}{Q}$ over $\ctxext{\Gamma}{\subst{x}{P}}$; the
+`action on terms' of substitution takes a term $g$ of $Q$ over
+$\ctxext{{\Gamma}{A}}{P}$ to a term $\subst[\tmsym]{x}{g}$ of 
+$\subst[\famsym]{x}{Q}$.
 \begin{align}
@@
     }
-  { \jfam{{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}
+  { \jfam{{\Gamma}{\subst{x}{P}}}{\subst[\famsym]{x}{Q}}
     }
@@
       }
-    { \jfameq{{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}{\subst{x'}{Q'}}
+    { \jfameq{{\Gamma}{\subst{x}{P}}}{\subst[\famsym]{x}{Q}}{\subst[\famsym]{x'}{Q'}}
       }
@@
     }
-  { \jterm{{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}{\subst{x}{g}}
+  { \jterm{{\Gamma}{\subst{x}{P}}}{\subst[\famsym]{x}{Q}}{\subst[\tmsym]{x}{g}}
     }
@@
         {{\Gamma}{\subst{x}{P}}}
-        {\subst{x}{Q}}
-        {\subst{x}{g}}
-        {\subst{x'}{g'}}
-      }
-\end{align}
+        {\subst[\famsym]{x}{Q}}
+        {\subst[\tmsym]{x}{g}}
+        {\subst[\tmsym]{x'}{g'}}
+      }
+\end{align}
+The following inference rules establish that the fibers of the empty family are 
+the empty families:
+\label{comp-s0}
+\begin{align}
+& \inference
+  { \jterm{\Gamma}{A}{x}
+    }
+  { \jfameq{\Gamma}{\subst{x}{\emptyf}}{\emptyf}
+    }
+  \\
+& \inference
+  { \jterm{\Gamma}{A}{x}
+    \jfam{{\Gamma}{A}}{P}
+    }
+  { \jfameq
+      {{\Gamma}{\subst{x}{P}}}
+      {\subst{x}{\emptyf}}
+      {\emptyf}
+    }
+\end{align}
+Note again that we do not have to axiomatize the action of substitution by $x$ on the term
+of the empty family, because the result will be in the empty family and every
+term in the empty family is judgmentally equal to the term $\emptytm$.
+
+We use the above rule to state the compatibility of the action on families of
+substitution with the action on contexts of substitution. Note that a family
+$P$ over $\ctxext{\Gamma}{A}$ may be regarded as a family over
+$\ctxext{{\Gamma}{A}}{\emptyf}$. Thus, we may consider the family
+$\subst[\famsym]{x}{P}$ over $\ctxext{\Gamma}{\subst{x}{\emptyf}}$. Since
+$\subst{x}{\emptyf}$ is judgmentally equal to the empty family, we may compare
+$\subst[\famsym]{x}{P}$ with $\subst{x}{P}$:
+\begin{equation}
+\inference
+{ \jfam{{\Gamma}{A}}{P}
+  }
+{ \jfameq{\Gamma}{\subst[\famsym]{x}{P}}{\subst{x}{P}}
+  }
+\end{equation}
+Due to this rule we need not make the annotations $\famsym$ and $\tmsym$ in
+the notation for substitution anymore and thus we shall omit them from now on.
+Note that because a family $P$ over $\ctxext{\Gamma}{A}$ is eligible for
+application of the action on families of substitution, we may also substitute
+terms of $P$. Thus, given terms $\jterm{\Gamma}{A}{x}$ and $\jterm{{\Gamma}{A}}{P}{f}$,
+we get a term $\jterm{\Gamma}{\subst{x}{P}}{\subst{x}{f}}$, the \emph{value of
+$f$ at $x$}.
+
+\label{comp-0s}
+The following rules assert that substituting by the term $\jterm{\Gamma}{\emptyf}{\emptytm}$
+leaves everything unchanged.
+\begin{align}
+& \inference
+  { \jfam{\Gamma}{A}
+    }
+  { \jfameq{\Gamma}{\subst{\emptytm}{A}}{A}
+    }
+  \\
+%& \inference
+%  { \jterm{\Gamma}{A}{x}
+%    }
+%  { \jtermeq{\Gamma}{A}{\subst{\emptytm}{x}}{x}
+%    }
+%  \\
+& \inference
+  { \jfam{{\Gamma}{A}}{P}
+    }
+  { \jfameq{{\Gamma}{A}}{\subst{\emptytm}{P}}{P}
+    }
+  \\
+& \inference
+  { \jterm{{\Gamma}{A}}{P}{f}
+    }
+  { \jtermeq{{\Gamma}{A}}{P}{\subst{\emptytm}{f}}{f}
+    }.
+\end{align}
+
 In \autoref{morphisms} we will use a combination of weakening and substitution
@@
 
-\subsubsection{Substitutions of the empty context}
-\label{comp-s0}
-The fibers of the empty family are the empty families:
-\begin{align}
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    }
-  { \jfameq{\Gamma}{\subst{x}{\emptyf}}{\emptyf}
-    }
-  \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jfam{{\Gamma}{A}}{P}
-    }
-  { \jfameq
-      {{\Gamma}{\subst{x}{P}}}
-      {\subst{x}{\emptyf}}
-      {\emptyf}
-    }
-\end{align}
-Note again that we do not have to axiomatize the action of substitution by $x$ on the term
-of the empty family, because the result will be in the empty family and every
-term in the empty family is judgmentally equal to the term $\emptytm$.
-
-\subsubsection{Substitution by the term of the empty context}
-\label{comp-0s}
-The following rules assert that substituting by the term $\jterm{\Gamma}{\emptyf}{\emptytm}$
-leaves everything unchanged.
-\begin{align}
-& \inference
-  { \jfam{\Gamma}{A}
-    }
-  { \jfameq{\Gamma}{\subst{\emptytm}{A}}{A}
-    }
-  \\
-%& \inference
-%  { \jterm{\Gamma}{A}{x}
-%    }
-%  { \jtermeq{\Gamma}{A}{\subst{\emptytm}{x}}{x}
-%    }
-%  \\
-& \inference
-  { \jfam{{\Gamma}{A}}{P}
-    }
-  { \jfameq{{\Gamma}{A}}{\subst{\emptytm}{P}}{P}
-    }
-  \\
-& \inference
-  { \jterm{{\Gamma}{A}}{P}{f}
-    }
-  { \jtermeq{{\Gamma}{A}}{P}{\subst{\emptytm}{f}}{f}
-    }.
-\end{align}
-
 \subsubsection{Extensions of extensions are extensions by extensions}

### 2014-04-30 article-models-tt.tex
edited the subsubsection of substitution

@@
 \hline
-empty context & \autoref{comp-00} & \autoref{comp-0e}* & \autoref{comp-0w} & \autoref{comp-0s} & \xmark \\
-extension & \autoref{comp-e0}* & \autoref{comp-ee}* & \autoref{comp-ew} & \cmark & \cmark \\
-weakening & \autoref{comp-w0} & \autoref{comp-we}* & \autoref{comp-ww} & \autoref{comp-sw} & \autoref{comp-wi} \\
-substitution & \autoref{comp-s0} & \autoref{comp-se}* & \autoref{comp-ws} & \autoref{comp-ss} & \autoref{comp-si} \\
-identity terms & \cmark & \cmark & \autoref{comp-wi} & \autoref{comp-si} & \xmark
+extension & \autoref{comp-ee}* & \autoref{comp-ew} & \cmark & \cmark \\
+weakening & \autoref{comp-we}* & \autoref{comp-ww} & \autoref{comp-sw} & \autoref{comp-wi} \\
+substitution & \autoref{comp-se}* & \autoref{comp-ws} & \autoref{comp-ss} & \autoref{comp-si} \\
+identity terms & \cmark & \autoref{comp-wi} & \autoref{comp-si} & \xmark
 \end{tabular}

### 2014-04-30 article-models-tt.tex
editing 4.4

@@
 \begin{center}
-\begin{tabular}{r|rrrrr}
-& emp & ext & wk & subst & id\\
+\begin{tabular}{r|rrrr}
+& ext & wk & subst & id\\
 \hline
-extension & \autoref{comp-ee}* & \autoref{comp-ew} & \cmark & \cmark \\
+extension & \autoref{comp-ee}* &  & \cmark & \cmark \\
 weakening & \autoref{comp-we}* & \autoref{comp-ww} & \autoref{comp-sw} & \autoref{comp-wi} \\
@@
 
-\subsubsection{Weakening of weakenings}
-When a weakened family is weakened even further, the outer weakening distributes
-over the `weakening factors'.
-\label{comp-ww}
-\begin{align}
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{\Gamma}{B}
-    \jfam{{\Gamma}{C}}{R}
-    }
-  { \jfameq
-      {{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{{A}{B}}{{A}{C}}}}
-      {\ctxwk{A}{{B}{R}}}
-      {\ctxwk{{A}{B}}{{A}{R}}}
-    }
-  \label{comp-ww-f}\\
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{\Gamma}{B}
-    \jterm{{\Gamma}{C}}{R}{t}
-    }
-  { \jtermeq
-      {{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{{A}{B}}{{A}{C}}}}
-      {\ctxwk{{A}{B}}{{A}{R}}}
-      {\ctxwk{A}{{B}{t}}}
-      {\ctxwk{{A}{B}}{{A}{t}}}
-    }
-  \label{comp-ww-t}
-\end{align}
-
-\subsubsection{Weakening by extensions}
-\label{comp-ew}
-The rules expressing that extension is compatible with weakening assert that
-weakening by an extension is the same thing as weakening twice in the
-appropriate way.
-\begin{align}
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{{\Gamma}{A}}{P}
-    \jfam{{\Gamma}{B}}{Q}
-    }
-  { \jfameq
-      {{{{\Gamma}{A}}{P}}{\ctxwk{P}{{A}{B}}}}
-      {\ctxwk{\ctxext{A}{P}}{Q}}
-      {\ctxwk{P}{{A}{Q}}}
-    }
-  \label{comp-ew-f}\\
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{{\Gamma}{A}}{P}
-    \jterm{{\Gamma}{B}}{Q}{g}
-    }
-  { \jtermeq
-      {{{{\Gamma}{A}}{P}}{\ctxwk{P}{{A}{B}}}}
-      {\ctxwk{P}{{A}{Q}}}
-      {\ctxwk{\ctxext{A}{P}}{g}}
-      {\ctxwk{P}{{A}{g}}}
-    } 
-  \label{comp-ew-t}
-\end{align}
-
 \subsubsection{Fibers of extensions}
@@
 This variant is stated and proved in \autoref{comp-se-t}.
+
+\subsubsection{Weakening of weakenings}
+When a weakened family is weakened even further, the outer weakening distributes
+over the `weakening factors'.
+\label{comp-ww}
+\begin{align}
+& \inference
+  { \jfam{\Gamma}{A}
+    \jfam{\Gamma}{B}
+    \jfam{{\Gamma}{C}}{R}
+    }
+  { \jfameq
+      {{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{{A}{B}}{{A}{C}}}}
+      {\ctxwk{A}{{B}{R}}}
+      {\ctxwk{{A}{B}}{{A}{R}}}
+    }
+  \label{comp-ww-f}\\
+& \inference
+  { \jfam{\Gamma}{A}
+    \jfam{\Gamma}{B}
+    \jterm{{\Gamma}{C}}{R}{t}
+    }
+  { \jtermeq
+      {{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{{A}{B}}{{A}{C}}}}
+      {\ctxwk{{A}{B}}{{A}{R}}}
+      {\ctxwk{A}{{B}{t}}}
+      {\ctxwk{{A}{B}}{{A}{t}}}
+    }
+  \label{comp-ww-t}
+\end{align}
 
@@
 \end{equation}
+
+\subsubsection{Currying for weakening}
+\label{comp-ew}
+The rules expressing that extension is compatible with weakening assert that
+weakening by an extension is the same thing as weakening twice in the
+appropriate way.
+\begin{align}
+& \inference
+  { \jfam{\Gamma}{A}
+    \jfam{{\Gamma}{A}}{P}
+    \jfam{{\Gamma}{B}}{Q}
+    }
+  { \jfameq
+      {{{{\Gamma}{A}}{P}}{\ctxwk{P}{{A}{B}}}}
+      {\ctxwk{\ctxext{A}{P}}{Q}}
+      {\ctxwk{P}{{A}{Q}}}
+    }
+  \label{comp-ew-f}\\
+& \inference
+  { \jfam{\Gamma}{A}
+    \jfam{{\Gamma}{A}}{P}
+    \jterm{{\Gamma}{B}}{Q}{g}
+    }
+  { \jtermeq
+      {{{{\Gamma}{A}}{P}}{\ctxwk{P}{{A}{B}}}}
+      {\ctxwk{P}{{A}{Q}}}
+      {\ctxwk{\ctxext{A}{P}}{g}}
+      {\ctxwk{P}{{A}{g}}}
+    } 
+  \label{comp-ew-t}
+\end{align}
 

### 2014-05-01 article-models-tt.tex
edited fibers of fibers

@@
 Note that the term $\subst{{x}{f}}{{x}{g}}$ may be written down more conveniently
-as $\subst{x,\subst{x}{f}}{g}$, although we will not do that here.
+as $\subst{x,\subst{x}{f}}{g}$, although we will not do that. In fact, after
+we have introduced extension on terms in \autoref{extension-on-terms}, we will 
+see that $\subst{{x}{f}}{{x}{g}}\jdeq\subst{\tmext{x}{\subst{x}{f}}}{g}$.
 

### 2014-05-09 article-models-ttconstr.tex
making a start with extension algebras

@@
 
-\subsection{Universes}\label{universes}
+\subsection{Pre-universes}
+We would like to investigate the universe operator, which assigns to a family
+$A$ over $\Gamma$ and a family $P$ over $\ctxext{\Gamma}{A}$ a universe
+$\UU(A,P)$, which is a family over $\Gamma$, and a decoding $\tilde{\UU}(A,P)$,
+which is a family over $\ctxext{\Gamma}{\UU(A,P)}$. However, we find it useful
+to first look at a pre-universe operator which extends the pair $(A,P)$ to a
+quadruple $(V_0(A,P),\tilde{V}_0(A,P),V_1(A,P),\tilde{V}_1(A,P))$ consisting of
+a family $V_0(A,P)$ in context $\Gamma$, a family $\tilde{V}_0(A,P)$ over
+$\ctxext{\Gamma}{V_0(A,P)}$, a family $V_1(A,P)$ over 
+$\ctxext{\Gamma}{V_0(A,P)}$ and a family $\tilde{V}_1(A,P)$ over.... and only
+later blur the distinction between $\subst{x}{V_1(A,P)}$ and family morphisms
+from $\subst{x}{\tilde{V}_0(A,P)}$ to $V_0(A,P)$ for terms $x:V_0(A,P)$. The
+pair $(V_0(A,P),V_1(A,P))$ can be seen as a version of $(A,P)$ which is
+expanded to incorporate extension, the empty context, weakening, substitution
+and identity terms.
+
+Our initial setup for pre-universes consists just of the two inverence rules
+calling for $V_0$ and $V_1$:
+\begin{align*}
+& \inference
+  { \jfam{{\Gamma}{A}}{P}
+    }
+  { \jfam{\Gamma}{V_0(A,P)}
+    }
+  \\
+& \inference
+  { \jfam{{\Gamma}{A}}{P}
+    }
+  { \jfam{{\Gamma}{V_0(A,P)}}{V_1(A,P)}
+    }
+\end{align*}
+
+\subsection{Universes}
+\label{universes}
+Now we investigate universes. When $A$ is a family over context $\Gamma$ and
+$P$ is a family over $\ctxext{\Gamma}{A}$, the universe at $(A,P)$ consists,
+among other things, of a family $\UU_0(A,P)$ over context $\Gamma$ and a family
+$\UU_1(A,P)$
+
 The type operator $\UU$ has as input a pair $(A,P)$ consisting of a family

### 2014-05-09 article-models-ttderived.tex
making a start with extension algebras

@@
 \end{itemize}
+
+\subsection{Extension algebras}
+Let $P$ be a family over an extended context $\ctxext{\Gamma}{A}$. We could
+mimic extension by requiring to have terms
+\begin{align*}
+\jhom*{\Gamma}{{A}{P}}{A}{e}\\
+\jhom*{{\Gamma}{A}}{{P}{\jcomp{}{e}{P}}}{P}{f}.
+\end{align*}
+We say that a quadruple $(A,P,e,f)$ as above is an extension algebra if the
+following diagrams commute judgmentally:
+\begin{equation*}
+\begin{tikzcd}[column sep=huge]
+\ctxext{A}{{P}{\jcomp{}{e}{P}}} 
+  \ar{d}[swap]{\tmext{\ctxwk{\jcomp{}{e}{P}}{e}}{\idtm{\jcomp{}{e}{P}}}} 
+  \ar{r}{\tmext{\ctxwk{\ctxext{P}{\jcomp{}{e}{P}}}{\idtm{A}}}{f}} 
+  & \ctxext{A}{P} \ar{d}{e}\\
+\ctxext{A}{P} \ar{r}[swap]{e} & A
+\end{tikzcd}
+\end{equation*}
+\begin{equation*}
+\begin{tikzcd}[column sep=huge]
+\ctxext{P}{{\jcomp{}{e}{P}}{\jcomp{}{\bar{e}}{{}{e}{P}}}} 
+  \ar{r} 
+  \ar{d}[swap]{
+    \tmext
+      {\ctxwk{\jcomp{}{\bar{e}}{{}{e}{P}}}{f}}
+      {\idtm{\jcomp{}{\bar{e}}{{}{e}{P}}}}
+    }
+& \ctxext{P}{\jcomp{}{e}{P}} \ar{d}{f}\\
+\ctxext{P}{\jcomp{}{e}{P}} \ar{r}[swap]{f} & P
+\end{tikzcd}
+\end{equation*}
+where we have taken 
+$\bar{e}\defeq\tmext{\ctxwk{\jcomp{}{e}{P}}{e}}{\idtm{\jcomp{}{e}{P}}}$, which
+is the morphism on the left of the first diagram (taking the upper morphism
+would yield a judgmentally equal result).

### 2014-05-09 article-models-ttmodels.tex
making a start with extension algebras

@@
 \subsection{The basic ingredients of the type theory of models}
-We first introduce the basic ingredients of our abstract theory of models.
+\subsubsection{The family and term models and the initial model}
+For every model $A$ there is a dependent model $\mfam{A}$ of families over $A$,
+called the \emph{family model of $A$}.
+The terms of a model $A$ are to be thought of as the contexts of $A$. Thus, when
+we have a term $x:A$, the families over $x$ are the terms of the substituted
+model $\subst{x}{\mfam{A}}$.
+\begin{equation*}
+\inference
+  { \jfam{\Gamma}{A}
+    }
+  { \jfam{{\Gamma}{A}}{\mfam{A}}
+    }
+  \tag{families}
+\end{equation*}
+
+Likewise, there is a \emph{term model $\mtm{A}$ of $A$}, 
+which is a family of models over the family model $\mfam{A}$ of $A$. When
+$x_0:A$ and $x_1:\mfam{A}$, then a term of $\subst{x_1}{{x_0}{\mtm{A}}}$ is
+to be thought of as a term of $x_1$ in context $x_0$.
+\begin{equation*}
+\inference
+  { \jctx{\Gamma}
+    }
+  { \jfam{{\Gamma}{\mfam{\Gamma}}}{\mtm{\Gamma}}
+    }
+  \tag{terms}
+\end{equation*}
+We shall furthermore assume that there is an initial model $\mctx$ in the theory
+of models.
 \begin{align*}
@@
     }
-  \tag{initiality of $\mctx$}\\
-& \inference
-  { \jfam{\Gamma}{A}
-    }
-  { \jfam{{\Gamma}{A}}{\mfam{A}}
-    }
-  \tag{families}\\
-& \inference
-  { \jctx{\Gamma}
-    }
-  { \jfam{{\Gamma}{\mfam{\Gamma}}}{\mtm{\Gamma}}
-    }
-  \tag{terms}
-\intertext{%
-  The following two rules essentially make a start with saying that every term is functorial
-  in the apropriate sense (there will be more rules contributing to this vies):}
+  \tag{initiality of $\mctx$}
+\end{align*}
+
+\subsubsection{The action of terms on families and on terms}
+We will require that every term $\jterm{{\Gamma}{A}}{P}{f}$ has an action on
+families and on terms. It tells that when $x_1$ is a family over $x_0$ in $A$
+with a term $y$, then we obtain a family $\subst{x_1}{{x_0}{\mfam{f}}}$ over
+$\subst{x_0}{f}$ with a term $\subst{y}{{x_1}{{x_0}{\mtm{f}}}}$.
+However, there is a choice in the precise formulation of this idea. One option
+is to require that $\mfam{f}$ is a term of the family $\ctxwk{\mfam{A}}
+{\subst{f}{\mfam{P}}}$; the other option, first investigated over a pastis
+and peanuts before having breakfast, is that $\mfam{f}$ is a term of
+the family $\jcomp{}{\tfext{A}}{\subst{f}{\mfam{P}}}$. We explore these two
+options in more detail separately.
+
+\paragraph{Option one: pullback along the projection}
+Suppose we require the inference rules
+\begin{align*}
 & \inference
@@
     }
-\intertext{%
+\end{align*}
 When the situation requires clarity --  for instance when extensions
 are involved -- we will write $\mfam[\Gamma]{f}$ and $\mtm[\Gamma]{f}$ to make
-  what we regard as the context explicit.}
-\intertext{%
-The empty context of a model $A$ is simply a term of $A$:}
+what we regard as the context explicit.
+
+With these two inference rules we see that
+\begin{align*}
+\subst{\tfemp{A}}{\ctxwk{\mfam{A}}{\subst{f}{\mfam{P}}}}
+& \jdeq
+  \ctxwk{\subst{\tfemp{A}}{\mfam{A}}}{\subst{\tfemp{A}}{{f}{\mfam{P}}}}
+  \\
+& \jdeq
+  \ctxwk{A}{\subst{\tfemp{A}}{{f}{\mfam{P}}}}
+  \\
+& \jdeq
+  \ctxwk{A}{\subst{{\tfemp{A}}{f}}{{\tfemp{A}}{\mfam{P}}}}
+  \\
+& \jdeq
+  \ctxwk{A}{\subst{\tfemp{\subst{\tfemp{A}}{P}}}{{\tfemp{A}}{\mfam{P}}}}
+  \\
+& \jdeq
+  \ctxwk{A}{\subst{\tfemp{\subst{\tfemp{A}}{P}}}}{\mfam{\subst{\tfemp{A}}{P}}}
+  \\
+& \jdeq
+  \ctxwk{A}{\subst{\tfemp{A}}{P}}
+\end{align*}
+
+\paragraph{Option two: pullback along extension}
+
+\subsubsection{The rest}
+The empty context of a model $A$ is simply a term of $A$:
+\begin{align*}
 & \inference

### 2014-05-09 article-models-ttderived.tex
extension algebras finished (initial version)

--- 2014-04-10 article-models-ttderived.tex
@@
 \end{itemize}
+
+\subsection{Extension algebras}
+Let $P$ be a family over an extended context $\ctxext{\Gamma}{A}$. We could
+mimic extension by requiring to have terms
+\begin{align*}
+\jhom*{\Gamma}{{A}{P}}{A}{e}\\
+\jhom*{{\Gamma}{A}}{{P}{\jcomp{}{e}{P}}}{P}{f}.
+\end{align*}
+We say that a quadruple $(A,P,e,f)$ as above is an extension algebra if the
+following diagrams commute judgmentally:
+\begin{equation*}
+\begin{tikzcd}[column sep=huge]
+\ctxext{A}{{P}{\jcomp{}{e}{P}}} 
+  \ar{d}[swap]{\jvcomp{}{e}{\idtm{\jcomp{}{e}{P}}}
+    %\tmext{\ctxwk{\jcomp{}{e}{P}}{e}}{\idtm{\jcomp{}{e}{P}}}
+    } 
+  \ar{r}{\jvcomp{}{\idtm{A}}{f}
+    %\tmext{\ctxwk{\ctxext{P}{\jcomp{}{e}{P}}}{\idtm{A}}}{f}
+    } 
+  & \ctxext{A}{P} \ar{d}{e}\\
+\ctxext{A}{P} \ar{r}[swap]{e} & A
+\end{tikzcd}
+\end{equation*}
+\begin{equation*}
+\begin{tikzcd}[column sep=huge]
+\ctxext{P}{{\jcomp{}{e}{P}}{\jcomp{}{\bar{e}}{{}{e}{P}}}} 
+  \ar{r}{\jvcomp{}{\idtm{P}}{\jcomp{}{e}{f}}}
+  \ar{d}[swap]{
+    \jvcomp{}{f}{\idtm{\jcomp{}{\bar{e}}{{}{e}{P}}}}
+%    \tmext
+%      {\ctxwk{\jcomp{}{\bar{e}}{{}{e}{P}}}{f}}
+%      {\idtm{\jcomp{}{\bar{e}}{{}{e}{P}}}}
+    }
+& \ctxext{P}{\jcomp{}{e}{P}} \ar{d}{f}\\
+\ctxext{P}{\jcomp{}{e}{P}} \ar{r}[swap]{f} & P
+\end{tikzcd}
+\end{equation*}
+where we have taken 
+$\bar{e}\defeq\jvcomp{}{e}{\idtm{\jcomp{}{e}{P}}}$, which
+is the morphism on the left of the first diagram (taking the upper morphism
+would yield a judgmentally equal result). To see that the above diagram makes
+sense, note that we can compute the upper left family in two different ways.
+First, we have the judgmental equalities
+\begin{align*}
+\jcomp{}{\bar{e}}{\jcomp{}{e}{P}}
+& \jdeq 
+  \jcomp{}{\jcomp{}{\bar{e}}{e}}{P}
+  \\
+& \jdeq
+  \jcomp{}{
+    \jcomp{}{\jvcomp{}{f}{\idtm{\jcomp{}{\bar{e}}{{}{e}{P}}}}}{e}}{P}
+  \\
+& \jdeq
+  \jcomp{}{\jvcomp{}{f}{\idtm{\jcomp{}{\bar{e}}{{}{e}{P}}}}}{
+    \jcomp{}{e}{P}}
+  \\
+& \jdeq
+  \subst
+    {\tmext{\ctxwk{\ctxext{P}{\jcomp{}{e}{P}}}{\idtm{A}}}{f}}
+    {\ctxwk{\ctxext{A}{{P}{\jcomp{}{e}{P}}}}{\jcomp{}{e}{P}}}
+  \\
+& \jdeq
+  \subst
+    {f}
+    { {\ctxwk{\ctxext{P}{\jcomp{}{e}{P}}}{\idtm{A}}}
+      {\ctxwk{\ctxext{A}{{P}{\jcomp{}{e}{P}}}}{\jcomp{}{e}{P}}}
+      }
+  \\
+& \jdeq
+  \subst
+    {f}
+    {\ctxwk{\ctxext{P}{\jcomp{}{e}{P}}}{\jcomp{}{e}{P}}}
+  \\
+& \jdeq
+  \jcomp{}{f}{\jcomp{}{e}{P}}
+\end{align*}
+so the downwards arrow on the left indeed makes sense. Second, we have the
+judgmental equalities
+\begin{align*}
+\ctxext{\jcomp{}{e}{P}}{\jcomp{}{\bar{e}}{\jcomp{}{e}{P}}}
+& \jdeq
+  \ctxext{\jcomp{}{e}{P}}{\subst{\jvcomp{}{e}{\idtm{\jcomp{}{e}{P}}}}
+    {\ctxwk{\ctxext{A}{{P}{\jcomp{}{e}{P}}}}{\jcomp{}{e}{P}}}}
+  \\
+& \jdeq
+  \ctxext{\jcomp{}{e}{P}}{\subst
+    {\idtm{\jcomp{}{e}{P}}}
+    {{\ctxwk{\jcomp{}{e}{P}}{e}}
+    {\ctxwk{\ctxext{A}{{P}{\jcomp{}{e}{P}}}}{\jcomp{}{e}{P}}}}}
+  \\
+& \jdeq
+  \ctxext{\jcomp{}{e}{P}}{\subst
+    {\idtm{\jcomp{}{e}{P}}}
+    {{\ctxwk{\jcomp{}{e}{P}}{e}}
+    {\ctxwk{\ctxext{{A}{P}}{\jcomp{}{e}{P}}}{\jcomp{}{e}{P}}}}}
+  \\
+& \jdeq
+  \ctxext{\jcomp{}{e}{P}}{\subst
+    {\idtm{\jcomp{}{e}{P}}}
+    {{\ctxwk{\jcomp{}{e}{P}}{e}}
+    {\ctxwk{\jcomp{}{e}{P}}{{\ctxext{A}{P}}{\jcomp{}{e}{P}}}}}}
+  \\
+& \jdeq
+  \ctxext{\jcomp{}{e}{P}}{\subst
+    {\idtm{\jcomp{}{e}{P}}}
+    {\ctxwk{\jcomp{}{e}{P}}{\subst{e}
+    {\ctxwk{\ctxext{A}{P}}{\jcomp{}{e}{P}}}}}}
+  \\
+& \jdeq
+  \ctxext{\jcomp{}{e}{P}}{\subst{e}
+    {\ctxwk{\ctxext{A}{P}}{\jcomp{}{e}{P}}}}
+  \\
+& \jdeq
+  \ctxext{\jcomp{}{e}{P}}{\jcomp{}{e}{\jcomp{}{e}{P}}}
+  \\
+& \jdeq
+  \jcomp{}{e}{\ctxext{P}{\jcomp{}{e}{P}}}
+\end{align*}
+and we indeed have the morphism $\jcomp{}{e}{f}$ from
+$\jcomp{}{e}{\ctxext{P}{\jcomp{}{e}{P}}}$ to $\jcomp{}{e}{P}$.
+\subsection{Extension-empty algebras}
+\begin{defn}
+Let $P$ be a family over the extended context $\ctxext{\Gamma}{A}$, let
+$\jterm{\Gamma}{A}{x}$ and $\jterm{{\Gamma}{A}}{P}{f}$ be terms. Then the
+quadruple $(A,P,x,f)$ is said to be an \emph{empty algebra in context $\Gamma$}
+if the following judgmental equalities hold:
+\begin{align*}
+\jfameq*{\Gamma}{\subst{x}{P}}{A}\\
+\jtermeq*{\Gamma}{A}{\subst{x}{f}}{x}.
+\end{align*}
+\end{defn}

### 2014-05-09 article-models-ttderived.tex
rewritten the part about the second judgmental equality of extension algebras

--- 2014-04-10 article-models-ttderived.tex
@@
 \end{itemize}
+
+\subsection{Extension algebras}
+In this subsection our goal is to define the notion of extension algebras,
+which are internal versions of the extension operation of the theory of
+contexts, families and terms. In this article, their use will be mainly in
+universes.
+
+Let $P$ be a family over an extended context $\ctxext{\Gamma}{A}$. We could
+mimic extension by requiring to have terms
+\begin{align*}
+\jhom*{\Gamma}{{A}{P}}{A}{e}\\
+\jhom*{{\Gamma}{A}}{{P}{\jcomp{}{e}{P}}}{P}{f}.
+\end{align*}
+Extension also satisfies the properties explained in \autoref{comp-ee}, so we
+must find the two judgmental equalities for $e$ and $f$ mimicing those. 
+The first of these judgmental equalities is easy to give: it says that the
+following diagram commutes:
+\begin{equation}\label{eq:ealg-eq1}
+\begin{tikzcd}[column sep=huge]
+\ctxext{A}{{P}{\jcomp{}{e}{P}}} 
+  \ar{d}[swap]{\jvcomp{}{e}{\idtm{\jcomp{}{e}{P}}}
+    %\tmext{\ctxwk{\jcomp{}{e}{P}}{e}}{\idtm{\jcomp{}{e}{P}}}
+    } 
+  \ar{r}{\jvcomp{}{\idtm{A}}{f}
+    %\tmext{\ctxwk{\ctxext{P}{\jcomp{}{e}{P}}}{\idtm{A}}}{f}
+    } 
+  & \ctxext{A}{P} \ar{d}{e}\\
+\ctxext{A}{P} \ar{r}[swap]{e} & A
+\end{tikzcd}
+\end{equation}
+To get a feel for this judgmental equality we include the following lemma.
+
+\begin{lem}
+Let $A$, $P$, $e$ and $f$ be as above, satisfying \autoref{eq:ealg-eq1} and
+let $x_0:A$,
+$x_1:\subst{x_0}{P}$ and $x_2:\subst{x_1}{{x_0}{\jcomp{}{e}{P}}}$.
+Then we have the judgmental equality
+\begin{equation*}
+\subst{{x_2}{{x_1}{f}}}{{x_0}{e}}
+\jdeq
+\subst{x_2}{{{x_1}{{x_0}{e}}}{e}}.
+\end{equation*}
+\end{lem}
+
+\begin{proof}
+The proof is a simple computation:
+\begin{align*}
+\subst{{x_2}{{x_1}{f}}}{{x_0}{e}}
+& \jdeq
+  \subst{\tmext{x_0}{\subst{x_2}{{x_1}{f}}}}{e}
+  \\
+& \jdeq 
+  \subst{{x_2}{{x_1}{{x_0}{\jvcomp{}{\idtm{A}}{f}}}}}{e}
+  \\
+& \jdeq
+  \subst{x_2}{{x_1}{{x_0}{\jcomp{}{\jvcomp{}{\idtm{A}}{f}}{e}}}}
+  \\
+& \jdeq
+  \subst{x_2}{{x_1}{{x_0}{\jcomp{}{\jvcomp{}{e}{\idtm{\jcomp{}{e}{P}}}}{e}}}}
+  \\
+& \jdeq 
+  \subst{{x_2}{{x_1}{{x_0}{\jvcomp{}{e}{\idtm{\jcomp{}{e}{P}}}}}}}{e}
+  \\
+& \jdeq
+  \subst{\tmext{\subst{x_1}{{x_0}{e}}}{x_2}}{e}
+  \\
+& \jdeq
+  \subst{x_2}{{{x_1}{{x_0}{e}}}{e}}.\qedhere
+\end{align*}
+\end{proof}
+
+The second of the judgmental equalities is harder to describe, however. We need
+to consider `higher' families, i.e.~families over families over families, and
+thus we need to look at the family $\jcomp{}{e}{\jcomp{}{e}{P}}$ and find the
+two dotted morphisms in the diagram
+\begin{equation*}
+\begin{tikzcd}
+\ctxext{P}{{\jcomp{}{e}{P}}{\jcomp{}{e}{\jcomp{}{e}{P}}}}
+  \ar[densely dotted]{d}
+  \ar[densely dotted]{r}
+& \ctxext{P}{\jcomp{}{e}{P}} \ar{d}{f}\\
+\ctxext{P}{\jcomp{}{e}{P}} \ar{r}{f} & P
+\end{tikzcd}
+\end{equation*}
+The first is easy to find. Note that we have the judgmental equality
+\begin{equation*}
+\ctxext{\jcomp{}{e}{P}}{\jcomp{}{e}{\jcomp{}{e}{P}}}
+  \jdeq
+  \jcomp{}{e}{\ctxext{P}{\jcomp{}{e}{P}}}
+\end{equation*}
+and therefore we may just take the morphism
+\begin{equation*}
+\jhom
+  {{\Gamma}{A}}
+  {{P}{{\jcomp{}{e}{P}}{\jcomp{}{e}{\jcomp{}{e}{P}}}}}
+  {{P}{\jcomp{}{e}{P}}}
+  {\jvcomp{}{\idtm{P}}{\jcomp{}{e}{f}}}.
+\end{equation*}
+For the other morphism we need to look at the family
+$\ctxext{P}{{\jcomp{}{e}{P}}{\jcomp{}{e}{\jcomp{}{e}{P}}}}$ differently. We do
+that in the following lemma.
+
+\begin{lem}
+Suppose we have $A$, $P$, $e$ and $f$ satisfying \autoref{eq:ealg-eq1}. Then
+we have the judgmental equality
+\begin{equation*}
+\jcomp{}{e}{\jcomp{}{e}{P}}
+  \jdeq
+  \jcomp{}{f}{\jcomp{}{e}{P}}.
+\end{equation*}
+\end{lem}
+
+\begin{proof}
+The proof is a rather long computation:
+\begin{align*}
+\jcomp{}{e}{\jcomp{}{e}{P}}
+& \jdeq
+  \subst{e}
+    {\ctxwk{\ctxext{A}{P}}{\jcomp{}{e}{P}}}
+  \\
+& \jdeq
+  \subst
+    {\idtm{\jcomp{}{e}{P}}}
+    { \ctxwk
+        {\jcomp{}{e}{P}}{\subst{e}
+        {\ctxwk{\ctxext{A}{P}}{\jcomp{}{e}{P}}}}}
+  \\
+& \jdeq
+  \subst
+    {\idtm{\jcomp{}{e}{P}}}
+    { {\ctxwk{\jcomp{}{e}{P}}{e}}
+      {\ctxwk{\jcomp{}{e}{P}}{{\ctxext{A}{P}}{\jcomp{}{e}{P}}}}}
+  \\
+& \jdeq
+  \subst
+    {\idtm{\jcomp{}{e}{P}}}
+    { {\ctxwk{\jcomp{}{e}{P}}{e}}
+      {\ctxwk{\ctxext{{A}{P}}{\jcomp{}{e}{P}}}{\jcomp{}{e}{P}}}}
+  \\
+& \jdeq
+  \subst
+    {\idtm{\jcomp{}{e}{P}}}
+    { {\ctxwk{\jcomp{}{e}{P}}{e}}
+      {\ctxwk{\ctxext{A}{{P}{\jcomp{}{e}{P}}}}{\jcomp{}{e}{P}}}}
+  \\
+& \jdeq
+  \subst
+    {\unfold{\jvcomp{\jcomp{}{e}{P}}{e}{\idtm{\jcomp{}{e}{P}}}}}
+    {\ctxwk{\ctxext{A}{{P}{\jcomp{}{e}{P}}}}{\jcomp{}{e}{P}}}
+  \\
+& \jdeq
+  \subst
+    {\jvcomp{}{e}{\idtm{\jcomp{}{e}{P}}}}
+    {\ctxwk{\ctxext{A}{{P}{\jcomp{}{e}{P}}}}{\jcomp{}{e}{P}}}
+  \\
+& \jdeq
+  \jcomp{}{\jvcomp{}{e}{\idtm{\jcomp{}{e}{P}}}}{\jcomp{}{e}{P}}
+  \\
+& \jdeq 
+  \jcomp{}{\jcomp{}{\jvcomp{}{e}{\idtm{\jcomp{}{e}{P}}}}{e}}{P}
+  \\
+& \jdeq
+  \jcomp{}{
+    \jcomp{}{\jvcomp{}{\idtm{A}}{f}}{e}}{P}
+  \\
+& \jdeq
+  \jcomp{}{\jvcomp{}{\idtm{A}}{f}}{
+    \jcomp{}{e}{P}}
+  \\
+& \jdeq
+  \subst
+    {\tmext{\ctxwk{\ctxext{P}{\jcomp{}{e}{P}}}{\idtm{A}}}{f}}
+    {\ctxwk{\ctxext{A}{{P}{\jcomp{}{e}{P}}}}{\jcomp{}{e}{P}}}
+  \\
+& \jdeq
+  \subst
+    {f}
+    { {\ctxwk{\ctxext{P}{\jcomp{}{e}{P}}}{\idtm{A}}}
+      {\ctxwk{\ctxext{A}{{P}{\jcomp{}{e}{P}}}}{\jcomp{}{e}{P}}}
+      }
+  \\
+& \jdeq
+  \subst
+    {f}
+    { {\cprojfstf{A}{\ctxext{P}{\jcomp{}{e}{P}}}}
+      {\ctxwk{\ctxext{A}{{P}{\jcomp{}{e}{P}}}}{\jcomp{}{e}{P}}}
+      }
+  \\
+& \jdeq
+  \subst
+    {f}
+    {\jcomp{}{\cprojfstf{A}{\ctxext{P}{\jcomp{}{e}{P}}}}{\jcomp{}{e}{P}}}
+  \\
+& \jdeq
+  \subst
+    {f}
+    {\ctxwk{\ctxext{P}{\jcomp{}{e}{P}}}{\jcomp{}{e}{P}}}
+  \\
+& \jdeq
+  \jcomp{}{f}{\jcomp{}{e}{P}}.
+  \qedhere
+\end{align*}
+\end{proof}
+
+Now we see that we can use the morphism
+\begin{equation*}
+\jhom
+  {{\Gamma}{A}}
+  {{P}{{\jcomp{}{e}{P}}{\jcomp{}{f}{{}{e}{P}}}}}
+  {{P}{\jcomp{}{e}{P}}}
+  {\jvcomp{}{f}{\idtm{\jcomp{}{f}{{}{e}{P}}}}}.
+\end{equation*}
+Thus, the second judgmental equality we will need is that the diagram
+\begin{equation}\label{eq:ealg-eq2}
+\begin{tikzcd}[column sep=huge]
+\ctxext{P}{{\jcomp{}{e}{P}}{\jcomp{}{e}{{}{e}{P}}}} 
+  \ar{r}{\jvcomp{}{\idtm{P}}{\jcomp{}{e}{f}}}
+  \ar{d}[swap]{
+    \jvcomp{}{f}{\idtm{\jcomp{}{f}{{}{e}{P}}}}
+%    \tmext
+%      {\ctxwk{\jcomp{}{\bar{e}}{{}{e}{P}}}{f}}
+%      {\idtm{\jcomp{}{\bar{e}}{{}{e}{P}}}}
+    }
+& \ctxext{P}{\jcomp{}{e}{P}} \ar{d}{f}\\
+\ctxext{P}{\jcomp{}{e}{P}} \ar{r}[swap]{f} & P
+\end{tikzcd}
+\end{equation}
+commutes judgmentally. Now we can confidently formulate the definition of
+extension algebras.
+
+\begin{defn}
+An \emph{extension algebra in context $\Gamma$} is a quadruple $(A,P,e,f)$
+consisting of a family $A$ over context $\Gamma$, a family $P$ over the context
+$\ctxext{\Gamma}{A}$, a morphism $e$ from $\ctxext{A}{P}$ to $A$ in context
+$\Gamma$ and a morphism $f$ from $\ctxext{P}{\jcomp{}{e}{P}}$ to $P$ in context
+$\ctxext{\Gamma}{A}$, satisfying the judgmental equalities of
+\autoref{eq:ealg-eq1,eq:ealg-eq2}.
+\end{defn}
+
+We also give a bit of intuition to the requirement of \autoref{eq:ealg-eq2} by
+means of the following lemma.
+
+\begin{lem}
+Let $(A,P,e,f)$ be an extension algebra in context $\Gamma$ and let
+$y_0:P$, $y_1:\subst{y_0}{\jcomp{}{e}{P}}$ and 
+$y_2:\subst{y_1}{{y_0}{\jcomp{}{e}{\jcomp{}{e}{P}}}}$. Then we have the
+judgmental equality
+\begin{equation*}
+\subst{{y_2}{{y_1}{\jcomp{}{e}{f}}}}{{y_0}{f}}
+  \jdeq
+  \subst{y_2}{{{y_1}{{y_0}{f}}}{f}}
+\end{equation*}
+\end{lem}
+
+\subsection{Extension-empty algebras}
+\begin{defn}
+Let $P$ be a family over the extended context $\ctxext{\Gamma}{A}$, let
+$\jterm{\Gamma}{A}{x}$ and $\jterm{{\Gamma}{A}}{P}{f}$ be terms. Then the
+quadruple $(A,P,x,f)$ is said to be an \emph{empty algebra in context $\Gamma$}
+if the following judgmental equalities hold:
+\begin{align*}
+\jfameq*{\Gamma}{\subst{x}{P}}{A}\\
+\jtermeq*{\Gamma}{A}{\subst{x}{f}}{x}.
+\end{align*}
+\end{defn}

### 2014-05-09 article-models-ttderived.tex
small remark added
adding material to the morphisms subsection

@@
 
-\begin{lem}
+\begin{lem}\label{lem:prehom}
 The inference rule\begin{equation*}
@@
 \begin{defn}
-Let $\jhom{\Gamma}{A}{B}{f}$ and consider a term $\jterm{{\Gamma}{B}}{Q}{g}$.
+Let $\jhom{\Gamma}{A}{B}{f}$ and consider a family $\jfam{{\Gamma}{B}}{Q}$,
+a family $\jfam{{{\Gamma}{B}}{Q}}{R}$ and a term $\jterm{{{\Gamma}{B}}{Q}}{R}{h}$.
 We define\begin{align*}
@@
   {\jcomp{A}{f}{Q}}
-  {\unfold{\jcomp{A}{f}{Q}}}\\
-\jtermdefn*
-  {{\Gamma}{A}}
-  {\jcomp{A}{f}{Q}}
-  {\jcomp{A}{f}{g}}
-  {\unfold{\jcomp{A}{f}{g}}}.
-\end{align*}
-Likewise, when we have a family $\jfam{{{\Gamma}{B}}{Q}}{R}$ and a term
-$\jterm{{{\Gamma}{B}}{Q}}{R}{h}$, we define\begin{align*}
+  {\unfold{\jcomp{A}{f}{Q}}}
+  \\
 \jfamdefn*
@@
 
-We have lots of compatibility properties for composition:
+\begin{lem}
+Let $\jhom{\Gamma}{A}{B}{f}$. Then the inference rules
+\begin{align*}
+& \inference
+  { %
+    }
+  { \jfameq{{\Gamma}{A}}{\jcomp{A}{f}{\emptyf}}{\emptyf}
+    }
+  \\
+& \inference
+  { \jfam{{\Gamma}{B}}{Q}
+    }
+  { \jfameq{{{\Gamma}{B}}{\jcomp{A}{f}{Q}}}{\jcomp{A}{f}{\emptyf}}{\emptyf}
+    }
+\end{align*}
+are valid.
+\end{lem}
+
+\begin{proof}
+For the first, note that we have the judgmental equalities
+\begin{equation*}
+\jcomp{A}{f}{\emptyf}
+  \jdeq
+  \unfold{\jcomp{A}{f}{\emptyf}}
+  \jdeq
+  \subst{f}{\emptyf}
+  \jdeq
+  \emptyf.
+\end{equation*}
+The second judgmental equality is proven similarly.
+\end{proof}
+
+\begin{rmk}
+Recall that we can treat a family $\jfam{{\Gamma}{B}}{Q}$ as a family
+$\jfam{{{\Gamma}{B}}{\emptyf}}{Q}$ and that $\jcomp{A}{f}{\emptyf}\jdeq
+\emptyf$. Thus we can apply composition with $f$ to terms 
+$\jterm{{\Gamma}{B}}{Q}{g}$. We get
+\begin{equation*}
+\jtermeq
+  {{\Gamma}{B}}
+  {\jcomp{A}{f}{Q}}
+  {\jcomp{A}{f}{g}}
+  {\unfold{\jcomp{A}{f}{g}}}.
+\end{equation*}
+In the particular situation where we take $Q$ to be a weakened family
+$\ctxwk{B}{C}$, we see that we can apply composition with $f$ to morphisms
+from $B$ to $C$ and we can use \autoref{lem:prehom} to see that we get
+\begin{equation*}
+\jhomeq{\Gamma}{A}{C}{\jcomp{A}{f}{g}}{\unfold{\jcomp{A}{f}{g}}}
+\end{equation*}
+for $\jhom{\Gamma}{B}{C}{g}$.
+
+One might argue that the notation for composition should be reserved to only
+this special case, to not confuse with common intuition of composition. It is
+however very convenient to see composition as an operation of the theory of
+contexts, families and terms. This allows us to follow the scheme of
+compatibility rules which are provable for this form of composition. 
+\end{rmk}
+
+We have lots of compatibility properties for composition. On the one hand we
+have the properties that composition with a morphism $f$ is compatible with
+the empty family, extension, weakening, substitution and identity terms. On
+the other hand, there are compatibility properties saying what happens when
+we substitute by a composition, or when we weaken or substitute a composition.
+Proving all these compatibility properties is the content of the rest of this
+subsection. None of it is very difficult.
+
+\begin{lem}
+We have the following inference rule expressing that composition with $f$ is
+compatible with extension:
+\begin{equation*}
+\inference
+  { \jfam{{{{\Gamma}{B}}{Q}}{R}}{S}
+    }
+  { \jfameq
+      {{{\Gamma}{A}}{\jcomp{A}{f}{Q}}}
+      {\jcomp{A}{f}{\ctxext{R}{S}}}
+      {\ctxext{\jcomp{A}{f}{R}}{\jcomp{A}{f}{S}}}
+    }
+\end{equation*}
+\end{lem}
+
+\begin{proof}
+Let $\jfam{{{{\Gamma}{B}}{Q}}{R}}{S}$. Then we have the judgmental equalities
+\begin{align*}
+\jcomp{A}{f}{\ctxext{R}{S}}
+& \jdeq
+  \unfold{\jcomp{A}{f}{\ctxext{R}{S}}}
+  \\
+& \jdeq
+  \subst{f}{\ctxext{\ctxwk{A}{R}}{\ctxwk{A}{S}}}
+  \\
+& \jdeq
+  \unfoldall{\ctxext{\jcomp{A}{f}{R}}{\jcomp{A}{f}{S}}}
+  \\
+& \jdeq
+  \ctxext{\jcomp{A}{f}{R}}{\jcomp{A}{f}{S}}.\qedhere
+\end{align*}
+\end{proof}
+
+\begin{lem}
+We have the following inference rules expressing that composition with $f$ is
+compatible with weakening:
+\begin{align*}
+& \inference
+  { \jfam{{{\Gamma}{B}}{Q}}{R}
+    \jfam{{{\Gamma}{B}}{Q}}{S}
+    }
+  { \jfameq
+      {{{{\Gamma}{A}}{\jcomp{A}{f}{Q}}}{\jcomp{A}{f}{R}}}
+      {\jcomp{A}{f}{\ctxwk{R}{S}}}
+      {\ctxwk{\jcomp{A}{f}{R}}{\jcomp{A}{f}{S}}}
+    }
+  \\
+& \inference
+  { \jfam{{{\Gamma}{B}}{Q}}{R}
+    \jterm{{{\Gamma}{B}}{Q}}{S}{k}
+    }
+  { \jtermeq
+      {{{{\Gamma}{A}}{\jcomp{A}{f}{Q}}}{\jcomp{A}{f}{R}}}
+      {\jcomp{A}{f}{\ctxwk{R}{S}}}
+      {\jcomp{A}{f}{\ctxwk{R}{k}}}
+      {\ctxwk{\jcomp{A}{f}{R}}{\jcomp{A}{f}{k}}}
+    }
+\end{align*}
+\end{lem}
+
+\begin{proof}
+Consider the families $\jfam{{{\Gamma}{B}}{Q}}{R}$ and 
+$\jfam{{{\Gamma}{B}}{Q}}{S}$. Then we have the judgmental equalities
+\begin{align*}
+\jcomp{A}{f}{\ctxwk{R}{S}}
+& \jdeq
+  \unfold{\jcomp{A}{f}{\ctxwk{R}{S}}}
+  \\
+& \jdeq
+  \subst{f}{\ctxwk{{A}{R}}{{A}{S}}}
+  \\
+& \jdeq
+  \unfoldall{\ctxwk{\jcomp{A}{f}{R}}{\jcomp{A}{f}{S}}}
+  \\
+& \jdeq
+  \ctxwk{\jcomp{A}{f}{R}}{\jcomp{A}{f}{S}}.
+\end{align*}
+The proof of the second property is similar.
+\end{proof}
+
+\begin{lem}
+We have the following inference rules expressing that composition with $f$ is
+compatible with substitution:
+\begin{align*}
+& \inference
+  { \jterm{{{\Gamma}{B}}{Q}}{R}{h}
+    \jfam{{{{\Gamma}{B}}{Q}}{R}}{S}
+    }
+  { \jfameq
+      {{{\Gamma}{A}}{\jcomp{A}{f}{Q}}}
+      {\jcomp{A}{f}{\subst{h}{S}}}
+      {\subst{\jcomp{A}{f}{h}}{\jcomp{A}{f}{S}}}
+    }
+  \\
+& \inference
+  { \jterm{{{\Gamma}{B}}{Q}}{R}{h}
+    \jterm{{{{\Gamma}{B}}{Q}}{R}}{S}{k}
+    }
+  { \jtermeq
+      {{{\Gamma}{A}}{\jcomp{A}{f}{Q}}}
+      {\jcomp{A}{f}{\subst{h}{S}}}
+      {\jcomp{A}{f}{\subst{h}{k}}}
+      {\subst{\jcomp{A}{f}{h}}{\jcomp{A}{f}{k}}}
+    }
+\end{align*}
+\end{lem}
+
+\begin{proof}
+Let $\jterm{{{\Gamma}{B}}{Q}}{R}{h}$ and $\jfam{{{{\Gamma}{B}}{Q}}{R}}{S}$.
+Then we have the judgmental equalities
+\begin{align*}
+\jcomp{A}{f}{\subst{h}{S}}
+& \jdeq
+  \unfold{\jcomp{A}{f}{\subst{h}{S}}}
+  \\
+& \jdeq
+  \subst{f}{{\ctxwk{A}{h}}{\ctxwk{A}{S}}}
+  \\
+& \jdeq
+  \unfoldall{\subst{\jcomp{A}{f}{h}}{\jcomp{A}{f}{S}}}
+  \\
+& \jdeq
+  \subst{\jcomp{A}{f}{h}}{\jcomp{A}{f}{S}}.
+\end{align*}
+The proof of the second inference rule is similar.
+\end{proof}
+
+\begin{lem}
+We have the following inference rule expressing that composition with $f$ is
+compatible with identity terms:
+\begin{equation*}
+\inference
+  { \jfam{{{\Gamma}{B}}{Q}}{R}
+    }
+  { \jtermeq
+      {{{{\Gamma}{A}}{\jcomp{A}{f}{Q}}}{\jcomp{A}{f}{R}}}
+      {\ctxwk{\jcomp{A}{f}{R}}{\jcomp{A}{f}{R}}}
+      {\jcomp{A}{f}{\idtm{R}}}
+      {\idtm{\jcomp{A}{f}{R}}}
+    }
+\end{equation*}
+\end{lem}
+
+\begin{proof}
+Let $\jfam{{{\Gamma}{B}}{Q}}{R}$. We have the following judgmental equalities:
+\begin{align*}
+\jcomp{A}{f}{\idtm{R}}
+& \jdeq
+  \unfold{\jcomp{A}{f}{\idtm{R}}}
+  \\
+& \jdeq
+  \subst{f}{\idtm{\ctxwk{A}{R}}}
+  \\
+& \jdeq
+  \unfoldall{\idtm{\jcomp{A}{f}{R}}}
+  \\
+& \jdeq
+  \idtm{\jcomp{A}{f}{R}}.\qedhere
+\end{align*}
+\end{proof}
 
@@
 contexts, families and terms. In this article, their use will be mainly in
-universes.
+universes. The theory of extension algebras requires the full power (i.e.~all
+of the ingredients) of the theory of contexts, families and terms in its
+formulation.
 

### 2014-05-09 article-models-ttderived.tex
reorganization: move morphisms over morphisms to the later subsection
initiated a investigation of a stronger version of inductive morphism

@@
 
+\subsection{Projections and extension on terms}\label{extension-on-terms}
+In this subsection we consider the notion of extension on terms, which has now
+become definable inside our theory. Moreover, every compatibility rule one may
+dream of is provable as well, using the compatibility rules we have introduced
+earlier.
+
+\begin{defn}
+When $\jterm{\Gamma}{A}{x}$ and $\jterm{\Gamma}{\subst{x}{P}}{u}$ are terms,
+we define 
+\begin{equation*}
+\jtermdefn
+  {\Gamma}
+  {\ctxext{A}{P}}
+  {\tmext{A}{P}{x}{u}}
+  {\unfold{\tmext{A}{P}{x}{u}}}.
+\end{equation*} 
+\end{defn}
+
+Thus, the term $\tmext{A}{P}{x}{u}$ is the pairing of $x$ and $u$. Note that because
+we have the judgmental equality 
+$\ctxwk{P}{{A}{\ctxext{A}{P}}}\jdeq\ctxwk{\ctxext{A}{P}}{\ctxext{A}{P}}$ in the
+context $\ctxext{{\Gamma}{A}}{P}$, the
+pairing function could just be defined as $\idtm{\ctxext{A}{P}}$.
+
+When we substitute by an extended term we get an equal result as when we
+substitute two consecutive times, like the way currying works.
+
+\begin{lem}\label{comp-es}
+The following inference rules are valid:
+\begin{align*}
+& \inference
+  { \jterm{\Gamma}{A}{x}
+    \jterm{\Gamma}{\subst{x}{P}}{u}
+    \jfam{{{\Gamma}{A}}{P}}{Q}
+    }
+  { \jfameq
+      {\Gamma}
+      {\subst{\tmext{A}{P}{x}{u}}{Q}}
+      {\subst{u}{{x}{Q}}}
+    }
+  \\
+& \inference
+  { \jterm{\Gamma}{A}{x}
+    \jterm{\Gamma}{\subst{x}{P}}{u}
+    \jfam{{{{\Gamma}{A}}{P}}{Q}}{R}
+    }
+  { \jfameq
+      {{\Gamma}{\subst{u}{{x}{Q}}}}
+      {\subst{\tmext{A}{P}{x}{u}}{R}}
+      {\subst{u}{{x}{R}}}
+    }
+  \\
+& \inference
+  { \jterm{\Gamma}{A}{x}
+    \jterm{\Gamma}{\subst{x}{P}}{u}
+    \jterm{{{{\Gamma}{A}}{P}}{Q}}{R}{t}
+    }
+  { \jtermeq
+      {{\Gamma}{\subst{u}{{x}{Q}}}}
+      {\subst{u}{{x}{R}}}
+      {\subst{\tmext{A}{P}{x}{u}}{h}}
+      {\subst{u}{{x}{h}}}
+    }
+\end{align*}
+\end{lem}
+
+\begin{proof}
+We prove only the first judgmental equality. All the others are similar.
+Let $\jterm{\Gamma}{A}{x}$ and $\jterm{\Gamma}{\subst{x}{P}}{u}$
+be terms and let $\jfam{{{\Gamma}{A}}{P}}{Q}$ be a family. Then we have
+\begin{align*}
+\subst
+  {\tmext{A}{P}{x}{u}}
+  {Q} 
+& \jdeq 
+  \subst
+    {{u}{{x}{\idtm{\ctxext{A}{P}}}}}
+    {Q}
+  \tag{by definition}\\
+& \jdeq 
+  \subst
+    {{u}{{x}{\idtm{\ctxext{A}{P}}}}}
+    {{x}{\ctxwk{A}{Q}}}
+  \tag{by \autoref{defn-ws-3}}\\
+& \jdeq 
+  \subst
+    {{u}{{x}{\idtm{\ctxext{A}{P}}}}}
+    {{u}{\ctxwk{\subst{x}{P}}{\subst{x}{\ctxwk{A}{Q}}}}}
+  \tag{by \autoref{defn-ws-3}}\\
+& \jdeq 
+  \subst
+    {{u}{{x}{\idtm{\ctxext{A}{P}}}}}
+    {{u}{{x}{\ctxwk{P}{{A}{Q}}}}}
+  \tag{by \autoref{comp-sw-f}}\\
+& \jdeq 
+  \subst
+    {u}
+    {{{x}{\idtm{\ctxext{A}{P}}}}{{x}{\ctxwk{{P}{{A}{Q}}}}}}
+  \tag{by \autoref{comp-ss-f}}\\
+& \jdeq 
+  \subst
+    {u}
+    {{x}{{\idtm{\ctxext{A}{P}}}{\ctxwk{P}{{A}{Q}}}}}
+  \tag{by \autoref{comp-ss-f}}\\
+& \jdeq 
+  \subst
+    {u}
+    {{x}{{\idtm{\ctxext{A}{P}}}{\ctxwk{\ctxext{A}{P}}{Q}}}}
+  \tag{by \autoref{comp-ew-f}}\\
+& \jdeq 
+  \subst
+    {u}
+    {{x}{Q}}
+  \tag{by \autoref{idfunc-wk-defn}}
+\end{align*}
+\end{proof}
+
+We have seen above that the pairing function into $\ctxext{A}{P}$ is just the identity term on
+$\ctxext{A}{P}$. To analyze the pairing functin a little further, we will also
+need the projection maps from $\ctxext{A}{P}$ to $A$ and from $\ctxext{A}{P}$
+to $P$. We will now define these and see that the identity term of an
+extended family is the extension (or pairing) of the identity
+functions on the components in the apropriate way.
+
+To find out what the
+apropriate way is, note that
+\begin{align*}
+\ctxwk{\ctxext{A}{P}}{\ctxext{A}{P}} 
+& \jdeq 
+  \ctxwk{P}{{A}{\ctxext{A}{P}}}
+  \\
+& \jdeq 
+  \ctxext{\ctxwk{P}{{A}{A}}}{\ctxwk{P}{{A}{P}}}
+\end{align*}
+We have the term $\jterm{{\Gamma}{{A}{P}}}{\ctxwk{P}{A}}{\ctxwk{P}{\idtm{A}}}$.
+Thus we need to find out what $\subst{\ctxwk{P}{\idtm{A}}}{\ctxwk{P}{{A}{P}}}$ is:
+\begin{align*}
+\subst{\ctxwk{P}{\idtm{A}}}{\ctxwk{P}{{A}{P}}} 
+& \jdeq 
+  \ctxwk{P}{\subst{\idtm{A}}{\ctxwk{A}{P}}}
+  \\
+& \jdeq 
+  \ctxwk{P}{P},
+\end{align*}
+where we find the term $\idtm{P}$. Therefore we define:
+
+\begin{defn}
+Let $\jfam{\Gamma}{A}$ and $\jfam{{\Gamma}{A}}{P}$ be families. We define
+\begin{align*}
+\jhomdefn*
+  {\Gamma}
+  {{A}{P}}
+  {A}
+  {\cprojfstf{A}{P}}
+  {\unfold{\cprojfstf{A}{P}}}
+  \\
+\jtermdefn*
+  {\ctxext{\Gamma}{{A}{P}}}
+  {\ctxwk{P}{P}}
+  {\cprojsndf{A}{P}}
+  {\unfold{\cprojsndf{A}{P}}}
+\end{align*}
+\end{defn}
+
+The constructions of the terms $\tmext{A}{P}{x}{u}$ and $\cprojfst{A}{P}{w}$ and
+$\cprojsnd{A}{P}{w}$ are subject to various rules, with all of them being
+consequences of earlier introduced inference rules.
+
+\begin{lem}\label{lem:tmext-basic}
+The following inference rules expressing that pairing is a strict
+inverse to the combination of decompositions, are valid:
+\begin{align*}
+& \inference
+  { \jterm{\Gamma}{\ctxext{A}{P}}{w}
+    }
+  { \jtermeq
+      {\Gamma}
+      {\ctxext{A}{P}}
+      {\tmext{A}{P}{\cprojfst{A}{P}{w}}{\cprojsnd{A}{P}{w}}}
+      {w}
+    }
+  \\
+& \inference
+  { \jterm{\Gamma}{A}{x}
+    \jterm{\Gamma}{\subst{x}{P}}{u}
+    }
+  { \jtermeq
+      {\Gamma}
+      {A}
+      {\cprojfst{A}{P}{\tmext{A}{P}{x}{u}}}
+      {x}
+    }
+  \\
+& \inference
+  { \jterm{\Gamma}{A}{x}
+    \jterm{\Gamma}{\subst{x}{P}}{u}
+    }
+  { \jtermeq
+      {\Gamma}
+      {\subst{x}{P}}
+      {\cprojsnd{A}{P}{\tmext{A}{P}{x}{u}}}
+      {u}
+    }
+\end{align*}
+\end{lem}
+
+\begin{proof}
+To prove the first judgmental equality, note that
+\begin{align*}
+w 
+& \jdeq 
+  \subst{w}{\idtm{\ctxext{A}{P}}} 
+  \tag{by \autoref{idfunc-subst-defn}}\\
+& \jdeq 
+  \subst
+    { w}
+    { { \idtm{P}}
+      { { \ctxwk{P}{\idtm{A}}}
+        { \idtm{\ctxwk{\ctxext{A}{P}}{\ctxext{A}{P}}}}
+        }
+      }
+  \tag{by \autoref{idfunc-ext-comp}}\\
+& \jdeq 
+  \subst
+    { {w}
+      {\idtm{P}}
+      }
+    { {w}
+      { { \ctxwk{P}{\idtm{A}}
+          }
+        { \idtm{\ctxwk{\ctxext{A}{P}}{\ctxext{A}{P}}}
+          }
+        }
+      }
+  \tag{by \autoref{comp-ss-t}}\\
+& \jdeq 
+  \subst
+    { {w}
+      {\idtm{P}}
+      }
+    { { {w}
+        {\ctxwk{P}{\idtm{A}}}
+        }
+      { {w}
+        {\idtm{\ctxwk{\ctxext{A}{P}}{\ctxext{A}{P}}}}
+        }
+      }
+  \tag{by \autoref{comp-ss-t}}\\
+& \jdeq 
+  \subst
+    { {w}
+      {\idtm{P}}
+      }
+    { { {w}
+        {\ctxwk{P}{\idtm{A}}}
+        }
+      { {w}
+        {\ctxwk{\ctxext{A}{P}}{\idtm{\ctxext{A}{P}}}}
+        }
+      }
+  \tag{by \autoref{idfunc-wk-comp}}\\
+& \jdeq 
+  \subst
+    { {w}
+      {\idtm{P}}
+      }
+    { { {w}
+        {\ctxwk{P}{\idtm{A}}}
+        }
+      { \idtm{\ctxext{A}{P}}
+        }
+      }
+  \tag{by \autoref{defn-ws-4}}\\
+& \jdeq 
+  \tmext{A}{P}{\cprojfst{A}{P}{w}}{\cprojsnd{A}{P}{w}}
+  \tag{by definition}
+\end{align*}
+To prove the second judgmental equality, let $\jterm{\Gamma}{A}{x}$ and
+$\jterm{\Gamma}{\subst{x}{P}}{u}$. Then we have
+\begin{align*}
+\cprojfst{A}{P}{\tmext{A}{P}{x}{u}}
+& \jdeq 
+  \subst{\tmext{A}{P}{x}{u}}{\ctxwk{P}{\idtm{A}}}
+  \\
+& \jdeq 
+  \subst{u}{{x}{\ctxwk{P}{\idtm{A}}}} 
+  \\
+& \jdeq 
+  \subst{u}{\ctxwk{\subst{x}{P}}{\subst{x}{\idtm{A}}}}
+  \\
+& \jdeq 
+  \subst{x}{\idtm{A}}
+  \\
+& \jdeq 
+  x.
+\end{align*}
+To prove the third judgmental equality, note that
+\begin{align*}
+\cprojsnd{A}{P}{\tmext{A}{P}{x}{u}}
+& \jdeq 
+  \subst{\tmext{A}{P}{x}{u}}{\idtm{P}}
+  \\
+& \jdeq 
+  \subst{u}{{x}{\idtm{P}}}
+  \\
+& \jdeq 
+  \subst{u}{\idtm{\subst{x}{P}}}
+  \\
+& \jdeq 
+  u.
+  \qedhere
+\end{align*}
+\end{proof}
+
+In \autoref{lem:tmext-emp,lem:tmext-ext,lem:tmext-wk,lem:tmext-subst,lem:tmext-id}
+we show that term extension and the projections are
+compatible with the empty families, extension,
+weakening, substitution and the identity terms (in that order). \autoref{lem:tmext-id}
+is in fact a generalization of the above \autoref{lem:tmext-basic}.
+
+\begin{lem}\label{lem:tmext-emp}
+The following compatibility rules for extensions of the term of the empty family
+are valid:
+\begin{align*}
+& \inference
+  { \jterm{\Gamma}{A}{x}
+    }
+  { \jtermeq{\Gamma}{A}{\tmext{\emptytm}{x}}{x}
+    }
+& & \inference
+  { \jterm{\Gamma}{A}{x}
+    }
+  { \jtermeq{\Gamma}{A}{\tmext{x}{\emptytm}}{x}
+    }
+  \\
+& \inference
+  { \jfam{\Gamma}{A}
+    }
+  { \jtermeq
+      {{\Gamma}{A}}
+      {\emptyf}
+      {\cprojfstf{\emptyf}{A}}
+      {\emptytm}
+    }
+& & \inference
+  { \jfam{\Gamma}{A}
+    }
+  { \jhomeq
+      {\Gamma}
+      {A}
+      {A}
+      {\cprojfstf{A}{\emptyf}}
+      {\idtm{A}}
+    }
+  \\
+& \inference
+  { \jfam{\Gamma}{A}
+    }
+  { \jhomeq
+      {\Gamma}
+      {A}
+      {A}
+      {\cprojsndf{\emptyf}{A}}
+      {\idtm{A}}
+    }
+& & \inference
+  { \jfam{\Gamma}{A}
+    }
+  { \jtermeq
+      {{\Gamma}{A}}
+      {\emptyf}
+      {\cprojsndf{A}{\emptyf}}
+      {\emptytm}
+    }
+\end{align*}
+\end{lem}
+
+\begin{proof}
+These equalities are very easy to verify. We only display a proof of the first:
+\begin{equation*}
+\tmext{\emptytm}{x}
+\jdeq \unfold{\tmext{\emptyf}{A}{\emptytm}{x}}
+\jdeq \subst{x}{\idtm{{\emptyf}{A}}}
+\jdeq \subst{x}{\idtm{A}}
+\jdeq x.\qedhere
+\end{equation*}
+\end{proof}
+
+\begin{lem}\label{lem:tmext-ext}
+The following compatibility rules for two consecutive term extensions are valid:
+\begin{align*}
+& \inference
+  { \jterm{\Gamma}{A}{x}
+    \jterm{\Gamma}{\subst{x}{P}}{u}
+    \jterm{\Gamma}{\subst{\tmext{A}{P}{x}{u}}{Q}}{v}
+    }
+  { \jtermeq
+      {\Gamma}
+      {\ctxext{{A}{P}}{Q}}
+      {\tmext{A}{{P}{Q}}{x}{{\subst{x}{P}}{\subst{x}{Q}}{u}{v}}}
+      {\tmext{{A}{P}}{Q}{{A}{P}{x}{u}}{v}}
+    }
+  \\
+& \inference
+  { \jfam{{{\Gamma}{A}}{P}}{Q}
+    }
+  { \jhomeq
+      {\Gamma}
+      {\ctxext{{A}{P}}{Q}}
+      {A}
+      {\jcomp{}{\cprojfstf{{A}{P}}{Q}}{\cprojfstf{A}{P}}}
+      {\cprojfstf{A}{{P}{Q}}}
+    }
+  \\
+& \inference
+  { \jfam{{{\Gamma}{A}}{P}}{Q}
+    }
+  { \jhomeq
+      {{\Gamma}{A}}
+      {{P}{Q}}
+      {P}
+      {\jcomp{}{\cprojfstf{{A}{P}}{Q}}{\cprojsndf{A}{P}}}
+      {\jcomp{}{\cprojsndf{A}{{P}{Q}}}{\cprojfstf{P}{Q}}}
+    }
+  \\
+& \inference
+    { \jfam{{{\Gamma}{A}}{P}}{Q}
+      }
+    { \jhomeq
+        {{{\Gamma}{A}}{P}}
+        {Q}
+        {Q}
+        {\cprojsndf{{A}{P}}{Q}}
+        {\jcomp{}{\cprojsndf{A}{{P}{Q}}}{\cprojsndf{P}{Q}}}
+      }
+\end{align*}
+\end{lem}
+
+\begin{proof}
+Consider terms $\jterm{\Gamma}{A}{x}$, $\jterm{\Gamma}{\subst{x}{P}}{u}$ and
+$\jterm{\Gamma}{\subst{u}{{x}{Q}}}{v}$. Then we have
+\begin{align*}
+\tmext{x}{{u}{v}}
+& \jdeq 
+  \subst
+    {\tmext{u}{v}}{{x}{\idtm{\ctxext{A}{{P}{Q}}}}}
+  \\
+& \jdeq 
+  \subst{v}{{u}{{x}{\idtm{\ctxext{A}{{P}{Q}}}}}}
+  \\
+& \jdeq 
+  \subst{v}{{u}{{x}{\idtm{\ctxext{{A}{P}}{Q}}}}}
+  \\
+& \jdeq 
+  \subst{v}{{\tmext{x}{u}}{\idtm{\ctxext{{A}{P}}{Q}}}}
+  \\
+& \jdeq 
+  \tmext{{x}{u}}{v}.
+\end{align*}
+To prove the judmental equality
+\begin{equation*}
+\jhomeq
+  {\Gamma}
+  {\ctxext{{A}{P}}{Q}}
+  {A}
+  {\jcomp{}{\cprojfstf{{A}{P}}{Q}}{\cprojfstf{A}{P}}}
+  {\cprojfstf{A}{{P}{Q}}}
+\end{equation*}
+note that we have the judgmental equalities
+\begin{align*}
+\jcomp{{{A}{P}}{Q}}{\cprojfstf{{A}{P}}{Q}}{\cprojfstf{A}{P}}
+& \jdeq 
+  \unfoldall{\jcomp{{{A}{P}}{Q}}{\cprojfstf{{A}{P}}{Q}}{\cprojfstf{A}{P}}}
+  \\
+& \jdeq 
+  \subst
+    {\ctxwk{Q}{\idtm{{A}{P}}}}
+    {\ctxwk{Q}{{\ctxext{A}{P}}{{P}{\idtm{A}}}}}
+  \\
+& \jdeq
+  \ctxwk{Q}{\subst{\idtm{{A}{P}}}{\ctxwk{\ctxext{A}{P}}{{P}{\idtm{A}}}}}
+  \\
+& \jdeq
+  \ctxwk{Q}{{P}{\idtm{A}}}
+  \\
+& \jdeq
+  \ctxwk{\ctxext{P}{Q}}{\idtm{A}}
+  \\
+& \jdeq
+  \cprojfstf{A}{{P}{Q}}
+\end{align*}
+To prove the judgmental equality
+\begin{equation*}
+\jhomeq
+  {{\Gamma}{A}}
+  {{P}{Q}}
+  {P}
+  {\jcomp{}{\cprojfstf{{A}{P}}{Q}}{\cprojsndf{A}{P}}}
+  {\jcomp{}{\cprojsndf{A}{{P}{Q}}}{\cprojfstf{P}{Q}}}
+\end{equation*}
+note that we have the judgmental equalities
+\begin{align*}
+\jcomp{{{A}{P}}{Q}}{\cprojfstf{{A}{P}}{Q}}{\cprojsndf{A}{P}}
+& \jdeq 
+  \unfoldall{\jcomp{{{A}{P}}{Q}}{\cprojfstf{{A}{P}}{Q}}{\cprojsndf{A}{P}}}
+  \\
+& \jdeq
+  \subst{\ctxwk{Q}{\idtm{{A}{P}}}}{\ctxwk{Q}{{\ctxext{A}{P}}{\idtm{P}}}}
+  \\
+& \jdeq
+  \ctxwk{Q}{\subst{\idtm{{A}{P}}}{\ctxwk{\ctxext{A}{P}}{\idtm{P}}}}
+  \\
+& \jdeq
+  \ctxwk{Q}{\idtm{P}}
+  \\
+& \jdeq
+  \unfoldall{\jcomp{{P}{Q}}{\cprojsndf{A}{{P}{Q}}}{\cprojfstf{P}{Q}}}
+  \\
+& \jdeq
+  \jcomp{{P}{Q}}{\cprojsndf{A}{{P}{Q}}}{\cprojfstf{P}{Q}}
+\end{align*}
+To prove the judgmental equality
+\begin{equation*}
+\jhomeq
+  {{{\Gamma}{A}}{P}}
+  {Q}
+  {Q}
+  {\cprojsndf{{A}{P}}{Q}}
+  {\jcomp{}{\cprojsndf{A}{{P}{Q}}}{\cprojsndf{P}{Q}}}
+\end{equation*}
+note that we have the judgmental equalities
+\begin{align*}
+\cprojsndf{{A}{P}}{Q}
+& \jdeq
+  \unfoldall{\cprojsndf{{A}{P}}{Q}}
+  \\
+& \jdeq 
+  \unfoldall{\jcomp{{P}{Q}}{\cprojsndf{A}{{P}{Q}}}{\cprojsndf{P}{Q}}}
+  \\
+& \jdeq
+  \jcomp{}{\cprojsndf{A}{{P}{Q}}}{\cprojsndf{P}{Q}}\qedhere
+\end{align*}
+\begin{comment}
+%%%% This was a proof of an old version of the statement
+Now consider a term $\jterm{\Gamma}{\ctxext{A}{{P}{Q}}}{w}$. Then we have
+\begin{align*}
+w 
+& \jdeq 
+  \tmext
+    {A}
+    {{P}{Q}}
+    {\cprojfst{A}{\ctxext{P}{Q}}{w}}
+    {\cprojsnd{A}{\ctxext{P}{Q}}{w}}
+  \\
+& \jdeq 
+  \tmext
+    {A}
+    {{P}{Q}}
+    {\cprojfst{A}{\ctxext{P}{Q}}{w}}
+    { {P}
+      {Q}
+      {\cprojfst{P}{Q}{\cprojsnd{A}{\ctxext{P}{Q}}{w}}}
+      {\cprojsnd{P}{Q}{\cprojsnd{A}{\ctxext{P}{Q}}{w}}}
+      }
+  \\
+& \jdeq 
+  \tmext
+    {{A}{P}}
+    {Q}
+    { {} % need to provide base and family, but there's no unfold.
+      {}
+      {\cprojfst{A}{\ctxext{P}{Q}}{w}}
+      {\cprojfst{P}{Q}{\cprojsnd{A}{\ctxext{P}{Q}}{w}}}
+      }
+    { \cprojsnd{P}{Q}{\cprojsnd{A}{\ctxext{P}{Q}}{w}}
+      }
+\end{align*}
+Thus we see that 
+\begin{align*}
+\cprojfst{\ctxext{A}{P}}{Q}{w} 
+& \jdeq 
+  \tmext
+    {A}
+    {P}
+    {\cprojfst{A}{\ctxext{P}{Q}}{w}}
+    {\cprojfst{P}{Q}{\cprojsnd{A}{\ctxext{P}{Q}}{w}}}
+  \\ 
+\cprojsnd{\ctxext{A}{P}}{Q}{w} 
+& \jdeq 
+  \cprojsnd{P}{Q}{\cprojsnd{A}{\ctxext{P}{Q}}{w}},
+\end{align*}
+proving the fourth judgmental equality, and therefore also that
+\begin{align*}
+\cprojfst{A}{P}{\cprojfst{\ctxext{A}{P}}{Q}{w}} 
+& \jdeq 
+  \cprojfst{A}{\ctxext{P}{Q}}{w}
+  \\
+\cprojsnd{A}{P}{\cprojfst{\ctxext{A}{P}}{Q}{w}} 
+& \jdeq 
+  \cprojfst{P}{Q}{\cprojsnd{A}{\ctxext{P}{Q}}{w}},
+\end{align*}
+proving the second and the third judgmental equalities.
+\end{comment}
+\end{proof}
+
+\begin{lem}\label{lem:tmext-wk}\label{comp-we-t}
+When we weaken a term $\tmext{B}{Q}{y}{v}$ of $\ctxext{B}{Q}$ in context $\Gamma$ by
+a family $A$, the term that we get is $\tmext{\ctxwk{A}{B}}{\ctxwk{A}{Q}}{\ctxwk{A}{y}}{\ctxwk{A}{v}}$. More
+precisely, the following inference rules are valid:
+\begin{align*}
+& \inference
+  { \jterm{{\Gamma}{B}}{Q}{g}
+    \jterm{{\Gamma}{B}}{\subst{g}{R}}{t}
+    }
+  { \jtermeq
+      {{{\Gamma}{A}}{\ctxwk{A}{B}}}
+      {\ctxwk{A}{\ctxext{Q}{R}}}
+      {\ctxwk{A}{\tmext{Q}{R}{g}{t}}}
+      {\tmext{\ctxwk{A}{Q}}{\ctxwk{A}{R}}{\ctxwk{A}{g}}{\ctxwk{A}{t}}}
+    }
+  \\
+& \inference
+  { \jfam{{{\Gamma}{B}}{Q}}{R}
+    }
+  { \jhomeq
+      {{{\Gamma}{A}}{\ctxwk{A}{B}}}
+      {{\ctxwk{A}{Q}}{\ctxwk{A}{R}}}
+      {\ctxwk{A}{Q}}
+      {\cprojfstf{\ctxwk{A}{Q}}{\ctxwk{A}{R}}}
+      {\ctxwk{A}{\cprojfstf{Q}{R}}}
+    }
+  \\
+& \inference
+  { \jfam{{{\Gamma}{B}}{Q}}{R}
+    }
+  { \jhomeq
+      {{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{Q}}}
+      {\ctxwk{A}{R}}
+      {\ctxwk{A}{R}}
+      {\cprojsndf{\ctxwk{A}{Q}}{\ctxwk{A}{R}}}
+      {\ctxwk{A}{\cprojsndf{Q}{R}}}
+    }
+\end{align*}
+\end{lem}
+
+\begin{proof}
+Consider $\jterm{{\Gamma}{B}}{Q}{g}$ and $\jterm{{\Gamma}{B}}{\subst{g}{R}}{t}$.
+Then we have the judgmental equalities
+\begin{align*}
+\ctxwk{A}{\ctxext{Q}{R}{g}{t}}
+& \jdeq 
+  \ctxwk{A}{\subst{t}{{g}{\idtm{\ctxext{Q}{R}}}}}
+  \\
+& \jdeq 
+  \subst{\ctxwk{A}{t}}{\ctxwk{A}{\subst{g}{\idtm{\ctxext{Q}{R}}}}}
+  \\
+& \jdeq 
+  \subst{\ctxwk{A}{t}}{{\ctxwk{A}{g}}{\ctxwk{A}{\idtm{\ctxext{Q}{R}}}}}
+  \\
+& \jdeq 
+  \subst{\ctxwk{A}{t}}{{\ctxwk{A}{g}}{\idtm{\ctxwk{A}{\ctxext{Q}{R}}}}}
+  \\
+& \jdeq 
+  \subst{\ctxwk{A}{t}}{{\ctxwk{A}{g}}{\idtm{\ctxext{\ctxwk{A}{Q}}{\ctxwk{A}{R}}}}}
+  \\
+& \jdeq 
+  \tmext{\ctxwk{A}{Q}}{\ctxwk{A}{R}}{\ctxwk{A}{g}}{\ctxwk{A}{t}}
+\end{align*}
+Next, we want to prove the judgmental equality
+\begin{equation*}
+\jhomeq
+  {{{\Gamma}{A}}{\ctxwk{A}{B}}}
+  {{\ctxwk{A}{Q}}{\ctxwk{A}{R}}}
+  {\ctxwk{A}{Q}}
+  {\cprojfstf{\ctxwk{A}{Q}}{\ctxwk{A}{R}}}
+  {\ctxwk{A}{\cprojfstf{Q}{R}}}
+\end{equation*}
+Note that we have the judgmental equalities
+\begin{align*}
+\cprojfstf{\ctxwk{A}{Q}}{\ctxwk{A}{R}}
+& \jdeq
+  \unfoldall{\cprojfstf{\ctxwk{A}{Q}}{\ctxwk{A}{R}}}
+  \\
+& \jdeq
+  \ctxwk{{A}{R}}{{A}{\idtm{Q}}}
+  \\
+& \jdeq
+  \unfoldall{\ctxwk{A}{\cprojfstf{Q}{R}}}
+  \\
+& \jdeq
+  \ctxwk{A}{\cprojfstf{Q}{R}}.
+\end{align*}
+Finally, we want to prove the judgmental equality
+\begin{equation*}
+\jhomeq
+  {{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{Q}}}
+  {\ctxwk{A}{R}}
+  {\ctxwk{A}{R}}
+  {\cprojsndf{\ctxwk{A}{Q}}{\ctxwk{A}{R}}}
+  {\ctxwk{A}{\cprojsndf{Q}{R}}}
+\end{equation*}
+Note that we have the judgmental equalities
+\begin{align*}
+\cprojsndf{\ctxwk{A}{Q}}{\ctxwk{A}{R}}
+& \jdeq
+  \unfoldall{\cprojsndf{\ctxwk{A}{Q}}{\ctxwk{A}{R}}}
+  \\
+& \jdeq
+  \unfoldall{\ctxwk{A}{\cprojsndf{Q}{R}}}
+  \\
+& \jdeq
+  \ctxwk{A}{\cprojsndf{Q}{R}}.
+  \qedhere
+\end{align*}
+\end{proof}
+
+\begin{lem}\label{lem:tmext-subst}\label{comp-se-t}
+When we substitute an extended term $\tmext{P}{Q}{f}{g}$ of $\ctxext{P}{Q}$ by a term
+$x$ of $A$, the term that we get is $\tmext{\subst{x}{P}}{\subst{x}{Q}}{\subst{x}{f}}{\subst{x}{g}}$.
+More precisely, the following inference rules are valid:
+\begin{align*}
+& \inference
+  { \jterm{\Gamma}{A}{x}
+    \jterm{{{\Gamma}{A}}{P}}{Q}{g}
+    \jterm{{{\Gamma}{A}}{P}}{\subst{g}{R}}{t}
+    }
+  { \jtermeq
+      {{\Gamma}{\subst{x}{P}}}
+      {\ctxext{\subst{x}{Q}}{\subst{x}{R}}}
+      {\subst{x}{\tmext{Q}{R}{g}{t}}}
+      {\tmext{\subst{x}{Q}}{\subst{x}{R}}{\subst{x}{g}}{\subst{x}{t}}}
+    }
+  \\
+& \inference
+  { \jterm{\Gamma}{A}{x}
+    \jfam{{{{\Gamma}{A}}{P}}{Q}}{R}
+    }
+  { \jhomeq
+      {{\Gamma}{\subst{x}{P}}}
+      {{\subst{x}{Q}}{\subst{x}{R}}}
+      {\subst{x}{Q}}
+      {\cprojfstf{\subst{x}{Q}}{\subst{x}{R}}}
+      {\subst{x}{\cprojfstf{Q}{R}}}
+    }
+  \\
+& \inference
+  { \jterm{\Gamma}{A}{x}
+    \jfam{{{{\Gamma}{A}}{P}}{Q}}{R}
+    }
+  { \jhomeq
+      {{{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}}
+      {\subst{x}{R}}
+      {\subst{x}{R}}
+      {\cprojsndf{\subst{x}{Q}}{\subst{x}{R}}}
+      {\subst{x}{\cprojsndf{Q}{R}}}
+    }
+\end{align*}
+\end{lem}
+
+\begin{proof}
+Consider $\jterm{{\Gamma}{B}}{Q}{g}$ and $\jterm{{\Gamma}{B}}{\subst{g}{R}}{t}$.
+Then we have the judgmental equalities
+\begin{align*}
+\subst{x}{\tmext{Q}{R}{g}{t}}
+& \jdeq 
+  \subst{x}{{t}{{g}{\idtm{\ctxext{Q}{R}}}}}
+  \\
+& \jdeq 
+  \subst{{x}{t}}{{x}{{g}{\idtm{\ctxext{Q}{R}}}}}
+  \\
+& \jdeq 
+  \subst{{x}{t}}{{{x}{g}}{{x}{\idtm{\ctxext{Q}{R}}}}}
+  \\
+& \jdeq 
+  \subst{{x}{t}}{{{x}{g}}{\idtm{\subst{x}{\ctxext{Q}{R}}}}}
+  \\
+& \jdeq 
+  \subst{{x}{t}}{{{x}{g}}{\idtm{\ctxext{\subst{x}{Q}}{\subst{x}{R}}}}}
+  \\
+& \jdeq 
+  \tmext{\subst{x}{Q}}{\subst{x}{R}}{\subst{x}{g}}{\subst{x}{t}}.
+\end{align*}
+Next, we want to prove the judgmental equality
+\begin{equation*}
+\jhomeq
+  {{\Gamma}{\subst{x}{P}}}
+  {{\subst{x}{Q}}{\subst{x}{R}}}
+  {\subst{x}{Q}}
+  {\cprojfstf{\subst{x}{Q}}{\subst{x}{R}}}
+  {\subst{x}{\cprojfstf{Q}{R}}}
+\end{equation*}
+Note that we have the judgmental equalities
+\begin{align*}
+\cprojfstf{\subst{x}{Q}}{\subst{x}{R}}
+& \jdeq
+  \unfoldall{\cprojfstf{\subst{x}{Q}}{\subst{x}{R}}}
+  \\
+& \jdeq
+  \ctxwk{\subst{x}{R}}{\subst{x}{\idtm{Q}}}
+  \\
+& \jdeq
+  \unfoldall{\subst{x}{\cprojfstf{Q}{R}}}
+  \\
+& \jdeq
+  \subst{x}{\cprojfstf{Q}{R}}.
+\end{align*}
+And finally we want to prove the judgmental equality
+\begin{equation*}
+\jhomeq
+  {{{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}}
+  {\subst{x}{R}}
+  {\subst{x}{R}}
+  {\cprojsndf{\subst{x}{Q}}{\subst{x}{R}}}
+  {\subst{x}{\cprojsndf{Q}{R}}}
+\end{equation*}
+Note that we have the judgmental equalities
+\begin{align*}
+\cprojsndf{\subst{x}{Q}}{\subst{x}{R}}
+& \jdeq
+  \unfoldall{\cprojsndf{\subst{x}{Q}}{\subst{x}{R}}}
+  \\
+& \jdeq
+  \unfoldall{\subst{x}{\cprojsndf{Q}{R}}}
+  \\
+& \jdeq
+  \subst{x}{\cprojsndf{Q}{R}}.
+  \qedhere
+\end{align*}
+\end{proof}
+
+We find the following inference rule, which expresses that the identity term
+is compatible with extension:
+
+\begin{lem}\label{lem:tmext-id}\label{comp-ie}
+For any $\jfam{\Gamma}{A}$ and $\jfam{{\Gamma}{A}}{P}$ we have
+\begin{equation}\label{idfunc-ext-comp}
+\inference
+  { \jfam{\Gamma}{A}
+    \jfam{{\Gamma}{A}}{P}
+    }
+  { \jhomeq
+      {\Gamma}
+      {{A}{P}}{{A}{P}}
+      {\idtm{\ctxext{A}{P}}}
+      { \tmext
+          {\ctxwk{\ctxext{A}{P}}{A}}
+          {\ctxwk{\ctxext{A}{P}}{P}}
+          {\cprojfstf{A}{P}}
+          {\cprojsndf{A}{P}}
+        }
+    }
+\end{equation}
+\end{lem}
+
+\begin{proof}
+Consider the families $\jfam{\Gamma}{A}$ and $\jfam{{\Gamma}{A}}{P}$. Then
+we have the judgmental equalities
+\begin{align*}
+\tmext
+  {\ctxwk{\ctxext{A}{P}}{A}}
+  {\ctxwk{\ctxext{A}{P}}{P}}
+  {\cprojfstf{A}{P}}
+  {\cprojsndf{A}{P}}
+& \jdeq 
+  \unfold
+  { \tmext
+      {\ctxwk{\ctxext{A}{P}}{A}}
+      {\ctxwk{\ctxext{A}{P}}{P}}
+      {\cprojfstf{A}{P}}
+      {\cprojsndf{A}{P}}
+    }
+  \\
+& \jdeq 
+  \subst
+    { \idtm{P}
+      }
+    { {\ctxwk{P}{\idtm{A}}}
+      {\idtm{\ctxwk{\ctxext{A}{P}}{\ctxext{A}{P}}}}
+      }
+  \\
+& \jdeq 
+  \subst
+    { \idtm{P}
+      }
+    { {\ctxwk{P}{\idtm{A}}}
+      {\ctxwk{\ctxext{A}{P}}{\idtm{\ctxext{A}{P}}}}
+      }
+  \\
+& \jdeq 
+  \subst
+    { \idtm{P}
+      }
+    { {\ctxwk{P}{\idtm{A}}}
+      {\ctxwk{P}{{A}{\idtm{\ctxext{A}{P}}}}}
+      }
+  \\
+& \jdeq 
+  \subst
+    { \idtm{P}
+      }
+    {\ctxwk
+      {P}
+      { \subst
+        {\idtm{A}}
+        {\ctxwk{A}{\idtm{\ctxext{A}{P}}}}
+        }
+      }
+  \\
+& \jdeq 
+  \subst
+    { \idtm{A}
+      }
+    { \ctxwk{A}{\idtm{\ctxext{A}{P}}}
+      }
+  \\
+& \jdeq 
+  \idtm{\ctxext{A}{P}}.
+  \qedhere
+\end{align*}
+\end{proof}
+
+\subsection{More on morphisms}
+
 There is also a notion of morphism \emph{over} a morphism. We will develop this
@@
 \end{proof}
-
-\subsection{Projections and extension on terms}\label{extension-on-terms}
-In this subsection we consider the notion of extension on terms, which has now
-become definable inside our theory. Moreover, every compatibility rule one may
-dream of is provable as well, using the compatibility rules we have introduced
-earlier.
-
-\begin{defn}
-When $\jterm{\Gamma}{A}{x}$ and $\jterm{\Gamma}{\subst{x}{P}}{u}$ are terms,
-we define 
-\begin{equation*}
-\jtermdefn
-  {\Gamma}
-  {\ctxext{A}{P}}
-  {\tmext{A}{P}{x}{u}}
-  {\unfold{\tmext{A}{P}{x}{u}}}.
-\end{equation*} 
-\end{defn}
-
-Thus, the term $\tmext{A}{P}{x}{u}$ is the pairing of $x$ and $u$. Note that because
-we have the judgmental equality 
-$\ctxwk{P}{{A}{\ctxext{A}{P}}}\jdeq\ctxwk{\ctxext{A}{P}}{\ctxext{A}{P}}$ in the
-context $\ctxext{{\Gamma}{A}}{P}$, the
-pairing function could just be defined as $\idtm{\ctxext{A}{P}}$.
-
-When we substitute by an extended term we get an equal result as when we
-substitute two consecutive times, like the way currying works.
-
-\begin{lem}\label{comp-es}
-The following inference rules are valid:
-\begin{align*}
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jterm{\Gamma}{\subst{x}{P}}{u}
-    \jfam{{{\Gamma}{A}}{P}}{Q}
-    }
-  { \jfameq
-      {\Gamma}
-      {\subst{\tmext{A}{P}{x}{u}}{Q}}
-      {\subst{u}{{x}{Q}}}
-    }
-  \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jterm{\Gamma}{\subst{x}{P}}{u}
-    \jterm{{{\Gamma}{A}}{P}}{Q}{g}
-    }
-  { \jtermeq
-      {\Gamma}
-      {\subst{u}{{x}{Q}}}
-      {\subst{\tmext{A}{P}{x}{u}}{g}}
-      {\subst{u}{{x}{g}}}
-    }
-  \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jterm{\Gamma}{\subst{x}{P}}{u}
-    \jfam{{{{\Gamma}{A}}{P}}{Q}}{R}
-    }
-  { \jfameq
-      {{\Gamma}{\subst{u}{{x}{Q}}}}
-      {\subst{\tmext{A}{P}{x}{u}}{R}}
-      {\subst{u}{{x}{R}}}
-    }
-  \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jterm{\Gamma}{\subst{x}{P}}{u}
-    \jterm{{{{\Gamma}{A}}{P}}{Q}}{R}{t}
-    }
-  { \jtermeq
-      {{\Gamma}{\subst{u}{{x}{Q}}}}
-      {\subst{u}{{x}{R}}}
-      {\subst{\tmext{A}{P}{x}{u}}{t}}
-      {\subst{u}{{x}{t}}}
-    }
-\end{align*}
-\end{lem}
-
-\begin{proof}
-We prove only the first judgmental equality. All the others are similar.
-Let $\jterm{\Gamma}{A}{x}$ and $\jterm{\Gamma}{\subst{x}{P}}{u}$
-be terms and let $\jfam{{{\Gamma}{A}}{P}}{Q}$ be a family. Then we have
-\begin{align*}
-\subst
-  {\tmext{A}{P}{x}{u}}
-  {Q} 
-& \jdeq 
-  \subst
-    {{u}{{x}{\idtm{\ctxext{A}{P}}}}}
-    {Q}
-  \tag{by definition}\\
-& \jdeq 
-  \subst
-    {{u}{{x}{\idtm{\ctxext{A}{P}}}}}
-    {{x}{\ctxwk{A}{Q}}}
-  \tag{by \autoref{defn-ws-3}}\\
-& \jdeq 
-  \subst
-    {{u}{{x}{\idtm{\ctxext{A}{P}}}}}
-    {{u}{\ctxwk{\subst{x}{P}}{\subst{x}{\ctxwk{A}{Q}}}}}
-  \tag{by \autoref{defn-ws-3}}\\
-& \jdeq 
-  \subst
-    {{u}{{x}{\idtm{\ctxext{A}{P}}}}}
-    {{u}{{x}{\ctxwk{P}{{A}{Q}}}}}
-  \tag{by \autoref{comp-sw-f}}\\
-& \jdeq 
-  \subst
-    {u}
-    {{{x}{\idtm{\ctxext{A}{P}}}}{{x}{\ctxwk{{P}{{A}{Q}}}}}}
-  \tag{by \autoref{comp-ss-f}}\\
-& \jdeq 
-  \subst
-    {u}
-    {{x}{{\idtm{\ctxext{A}{P}}}{\ctxwk{P}{{A}{Q}}}}}
-  \tag{by \autoref{comp-ss-f}}\\
-& \jdeq 
-  \subst
-    {u}
-    {{x}{{\idtm{\ctxext{A}{P}}}{\ctxwk{\ctxext{A}{P}}{Q}}}}
-  \tag{by \autoref{comp-ew-f}}\\
-& \jdeq 
-  \subst
-    {u}
-    {{x}{Q}}
-  \tag{by \autoref{idfunc-wk-defn}}
-\end{align*}
-\end{proof}
-
-We have seen above that the pairing function into $\ctxext{A}{P}$ is just the identity term on
-$\ctxext{A}{P}$. To analyze the pairing functin a little further, we will also
-need the projection maps from $\ctxext{A}{P}$ to $A$ and from $\ctxext{A}{P}$
-to $P$. We will now define these and see that the identity term of an
-extended family is the extension (or pairing) of the identity
-functions on the components in the apropriate way.
-
-To find out what the
-apropriate way is, note that
-\begin{align*}
-\ctxwk{\ctxext{A}{P}}{\ctxext{A}{P}} 
-& \jdeq 
-  \ctxwk{P}{{A}{\ctxext{A}{P}}}
-  \\
-& \jdeq 
-  \ctxext{\ctxwk{P}{{A}{A}}}{\ctxwk{P}{{A}{P}}}
-\end{align*}
-We have the term $\jterm{{\Gamma}{{A}{P}}}{\ctxwk{P}{A}}{\ctxwk{P}{\idtm{A}}}$.
-Thus we need to find out what $\subst{\ctxwk{P}{\idtm{A}}}{\ctxwk{P}{{A}{P}}}$ is:
-\begin{align*}
-\subst{\ctxwk{P}{\idtm{A}}}{\ctxwk{P}{{A}{P}}} 
-& \jdeq 
-  \ctxwk{P}{\subst{\idtm{A}}{\ctxwk{A}{P}}}
-  \\
-& \jdeq 
-  \ctxwk{P}{P},
-\end{align*}
-where we find the term $\idtm{P}$. Therefore we define:
-
-\begin{defn}
-Let $\jfam{\Gamma}{A}$ and $\jfam{{\Gamma}{A}}{P}$ be families. We define
-\begin{align*}
-\jhomdefn*
-  {\Gamma}
-  {{A}{P}}
-  {A}
-  {\cprojfstf{A}{P}}
-  {\unfold{\cprojfstf{A}{P}}}
-  \\
-\jtermdefn*
-  {\ctxext{\Gamma}{{A}{P}}}
-  {\ctxwk{P}{P}}
-  {\cprojsndf{A}{P}}
-  {\unfold{\cprojsndf{A}{P}}}
-\end{align*}
-\end{defn}
-
-The constructions of the terms $\tmext{A}{P}{x}{u}$ and $\cprojfst{A}{P}{w}$ and
-$\cprojsnd{A}{P}{w}$ are subject to various rules, with all of them being
-consequences of earlier introduced inference rules.
-
-\begin{lem}\label{lem:tmext-basic}
-The following inference rules expressing that pairing is a strict
-inverse to the combination of decompositions, are valid:
-\begin{align*}
-& \inference
-  { \jterm{\Gamma}{\ctxext{A}{P}}{w}
-    }
-  { \jtermeq
-      {\Gamma}
-      {\ctxext{A}{P}}
-      {\tmext{A}{P}{\cprojfst{A}{P}{w}}{\cprojsnd{A}{P}{w}}}
-      {w}
-    }
-  \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jterm{\Gamma}{\subst{x}{P}}{u}
-    }
-  { \jtermeq
-      {\Gamma}
-      {A}
-      {\cprojfst{A}{P}{\tmext{A}{P}{x}{u}}}
-      {x}
-    }
-  \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jterm{\Gamma}{\subst{x}{P}}{u}
-    }
-  { \jtermeq
-      {\Gamma}
-      {\subst{x}{P}}
-      {\cprojsnd{A}{P}{\tmext{A}{P}{x}{u}}}
-      {u}
-    }
-\end{align*}
-\end{lem}
-
-\begin{proof}
-To prove the first judgmental equality, note that
-\begin{align*}
-w 
-& \jdeq 
-  \subst{w}{\idtm{\ctxext{A}{P}}} 
-  \tag{by \autoref{idfunc-subst-defn}}\\
-& \jdeq 
-  \subst
-    { w}
-    { { \idtm{P}}
-      { { \ctxwk{P}{\idtm{A}}}
-        { \idtm{\ctxwk{\ctxext{A}{P}}{\ctxext{A}{P}}}}
-        }
-      }
-  \tag{by \autoref{idfunc-ext-comp}}\\
-& \jdeq 
-  \subst
-    { {w}
-      {\idtm{P}}
-      }
-    { {w}
-      { { \ctxwk{P}{\idtm{A}}
-          }
-        { \idtm{\ctxwk{\ctxext{A}{P}}{\ctxext{A}{P}}}
-          }
-        }
-      }
-  \tag{by \autoref{comp-ss-t}}\\
-& \jdeq 
-  \subst
-    { {w}
-      {\idtm{P}}
-      }
-    { { {w}
-        {\ctxwk{P}{\idtm{A}}}
-        }
-      { {w}
-        {\idtm{\ctxwk{\ctxext{A}{P}}{\ctxext{A}{P}}}}
-        }
-      }
-  \tag{by \autoref{comp-ss-t}}\\
-& \jdeq 
-  \subst
-    { {w}
-      {\idtm{P}}
-      }
-    { { {w}
-        {\ctxwk{P}{\idtm{A}}}
-        }
-      { {w}
-        {\ctxwk{\ctxext{A}{P}}{\idtm{\ctxext{A}{P}}}}
-        }
-      }
-  \tag{by \autoref{idfunc-wk-comp}}\\
-& \jdeq 
-  \subst
-    { {w}
-      {\idtm{P}}
-      }
-    { { {w}
-        {\ctxwk{P}{\idtm{A}}}
-        }
-      { \idtm{\ctxext{A}{P}}
-        }
-      }
-  \tag{by \autoref{defn-ws-4}}\\
-& \jdeq 
-  \tmext{A}{P}{\cprojfst{A}{P}{w}}{\cprojsnd{A}{P}{w}}
-  \tag{by definition}
-\end{align*}
-To prove the second judgmental equality, let $\jterm{\Gamma}{A}{x}$ and
-$\jterm{\Gamma}{\subst{x}{P}}{u}$. Then we have
-\begin{align*}
-\cprojfst{A}{P}{\tmext{A}{P}{x}{u}}
-& \jdeq 
-  \subst{\tmext{A}{P}{x}{u}}{\ctxwk{P}{\idtm{A}}}
-  \\
-& \jdeq 
-  \subst{u}{{x}{\ctxwk{P}{\idtm{A}}}} 
-  \\
-& \jdeq 
-  \subst{u}{\ctxwk{\subst{x}{P}}{\subst{x}{\idtm{A}}}}
-  \\
-& \jdeq 
-  \subst{x}{\idtm{A}}
-  \\
-& \jdeq 
-  x.
-\end{align*}
-To prove the third judgmental equality, note that
-\begin{align*}
-\cprojsnd{A}{P}{\tmext{A}{P}{x}{u}}
-& \jdeq 
-  \subst{\tmext{A}{P}{x}{u}}{\idtm{P}}
-  \\
-& \jdeq 
-  \subst{u}{{x}{\idtm{P}}}
-  \\
-& \jdeq 
-  \subst{u}{\idtm{\subst{x}{P}}}
-  \\
-& \jdeq 
-  u.
-  \qedhere
-\end{align*}
-\end{proof}
-
-In \autoref{lem:tmext-emp,lem:tmext-ext,lem:tmext-wk,lem:tmext-subst,lem:tmext-id}
-we show that term extension and the projections are
-compatible with the empty families, extension,
-weakening, substitution and the identity terms (in that order). \autoref{lem:tmext-id}
-is in fact a generalization of the above \autoref{lem:tmext-basic}.
-
-\begin{lem}\label{lem:tmext-emp}
-The following compatibility rules for extensions of the term of the empty family
-are valid:
-\begin{align*}
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    }
-  { \jtermeq{\Gamma}{A}{\tmext{\emptytm}{x}}{x}
-    }
-& & \inference
-  { \jterm{\Gamma}{A}{x}
-    }
-  { \jtermeq{\Gamma}{A}{\tmext{x}{\emptytm}}{x}
-    }
-  \\
-& \inference
-  { \jfam{\Gamma}{A}
-    }
-  { \jtermeq
-      {{\Gamma}{A}}
-      {\emptyf}
-      {\cprojfstf{\emptyf}{A}}
-      {\emptytm}
-    }
-& & \inference
-  { \jfam{\Gamma}{A}
-    }
-  { \jhomeq
-      {\Gamma}
-      {A}
-      {A}
-      {\cprojfstf{A}{\emptyf}}
-      {\idtm{A}}
-    }
-  \\
-& \inference
-  { \jfam{\Gamma}{A}
-    }
-  { \jhomeq
-      {\Gamma}
-      {A}
-      {A}
-      {\cprojsndf{\emptyf}{A}}
-      {\idtm{A}}
-    }
-& & \inference
-  { \jfam{\Gamma}{A}
-    }
-  { \jtermeq
-      {{\Gamma}{A}}
-      {\emptyf}
-      {\cprojsndf{A}{\emptyf}}
-      {\emptytm}
-    }
-\end{align*}
-\end{lem}
-
-\begin{proof}
-These equalities are very easy to verify. We only display a proof of the first:
-\begin{equation*}
-\tmext{\emptytm}{x}
-\jdeq \unfold{\tmext{\emptyf}{A}{\emptytm}{x}}
-\jdeq \subst{x}{\idtm{{\emptyf}{A}}}
-\jdeq \subst{x}{\idtm{A}}
-\jdeq x.\qedhere
-\end{equation*}
-\end{proof}
-
-\begin{lem}\label{lem:tmext-ext}
-The following compatibility rules for two consecutive term extensions are valid:
-\begin{align*}
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jterm{\Gamma}{\subst{x}{P}}{u}
-    \jterm{\Gamma}{\subst{\tmext{A}{P}{x}{u}}{Q}}{v}
-    }
-  { \jtermeq
-      {\Gamma}
-      {\ctxext{{A}{P}}{Q}}
-      {\tmext{A}{{P}{Q}}{x}{{\subst{x}{P}}{\subst{x}{Q}}{u}{v}}}
-      {\tmext{{A}{P}}{Q}{{A}{P}{x}{u}}{v}}
-    }
-  \\
-& \inference
-  { \jfam{{{\Gamma}{A}}{P}}{Q}
-    }
-  { \jhomeq
-      {\Gamma}
-      {\ctxext{{A}{P}}{Q}}
-      {A}
-      {\jcomp{}{\cprojfstf{{A}{P}}{Q}}{\cprojfstf{A}{P}}}
-      {\cprojfstf{A}{{P}{Q}}}
-    }
-  \\
-& \inference
-  { \jfam{{{\Gamma}{A}}{P}}{Q}
-    }
-  { \jhomeq
-      {{\Gamma}{A}}
-      {{P}{Q}}
-      {P}
-      {\jcomp{}{\cprojfstf{{A}{P}}{Q}}{\cprojsndf{A}{P}}}
-      {\jcomp{}{\cprojsndf{A}{{P}{Q}}}{\cprojfstf{P}{Q}}}
-    }
-  \\
-& \inference
-    { \jfam{{{\Gamma}{A}}{P}}{Q}
-      }
-    { \jhomeq
-        {{{\Gamma}{A}}{P}}
-        {Q}
-        {Q}
-        {\cprojsndf{{A}{P}}{Q}}
-        {\jcomp{}{\cprojsndf{A}{{P}{Q}}}{\cprojsndf{P}{Q}}}
-      }
-\end{align*}
-\end{lem}
-
-\begin{proof}
-Consider terms $\jterm{\Gamma}{A}{x}$, $\jterm{\Gamma}{\subst{x}{P}}{u}$ and
-$\jterm{\Gamma}{\subst{u}{{x}{Q}}}{v}$. Then we have
-\begin{align*}
-\tmext{x}{{u}{v}}
-& \jdeq 
-  \subst
-    {\tmext{u}{v}}{{x}{\idtm{\ctxext{A}{{P}{Q}}}}}
-  \\
-& \jdeq 
-  \subst{v}{{u}{{x}{\idtm{\ctxext{A}{{P}{Q}}}}}}
-  \\
-& \jdeq 
-  \subst{v}{{u}{{x}{\idtm{\ctxext{{A}{P}}{Q}}}}}
-  \\
-& \jdeq 
-  \subst{v}{{\tmext{x}{u}}{\idtm{\ctxext{{A}{P}}{Q}}}}
-  \\
-& \jdeq 
-  \tmext{{x}{u}}{v}.
-\end{align*}
-To prove the judmental equality
-\begin{equation*}
-\jhomeq
-  {\Gamma}
-  {\ctxext{{A}{P}}{Q}}
-  {A}
-  {\jcomp{}{\cprojfstf{{A}{P}}{Q}}{\cprojfstf{A}{P}}}
-  {\cprojfstf{A}{{P}{Q}}}
-\end{equation*}
-note that we have the judgmental equalities
-\begin{align*}
-\jcomp{{{A}{P}}{Q}}{\cprojfstf{{A}{P}}{Q}}{\cprojfstf{A}{P}}
-& \jdeq 
-  \unfoldall{\jcomp{{{A}{P}}{Q}}{\cprojfstf{{A}{P}}{Q}}{\cprojfstf{A}{P}}}
-  \\
-& \jdeq 
-  \subst
-    {\ctxwk{Q}{\idtm{{A}{P}}}}
-    {\ctxwk{Q}{{\ctxext{A}{P}}{{P}{\idtm{A}}}}}
-  \\
-& \jdeq
-  \ctxwk{Q}{\subst{\idtm{{A}{P}}}{\ctxwk{\ctxext{A}{P}}{{P}{\idtm{A}}}}}
-  \\
-& \jdeq
-  \ctxwk{Q}{{P}{\idtm{A}}}
-  \\
-& \jdeq
-  \ctxwk{\ctxext{P}{Q}}{\idtm{A}}
-  \\
-& \jdeq
-  \cprojfstf{A}{{P}{Q}}
-\end{align*}
-To prove the judgmental equality
-\begin{equation*}
-\jhomeq
-  {{\Gamma}{A}}
-  {{P}{Q}}
-  {P}
-  {\jcomp{}{\cprojfstf{{A}{P}}{Q}}{\cprojsndf{A}{P}}}
-  {\jcomp{}{\cprojsndf{A}{{P}{Q}}}{\cprojfstf{P}{Q}}}
-\end{equation*}
-note that we have the judgmental equalities
-\begin{align*}
-\jcomp{{{A}{P}}{Q}}{\cprojfstf{{A}{P}}{Q}}{\cprojsndf{A}{P}}
-& \jdeq 
-  \unfoldall{\jcomp{{{A}{P}}{Q}}{\cprojfstf{{A}{P}}{Q}}{\cprojsndf{A}{P}}}
-  \\
-& \jdeq
-  \subst{\ctxwk{Q}{\idtm{{A}{P}}}}{\ctxwk{Q}{{\ctxext{A}{P}}{\idtm{P}}}}
-  \\
-& \jdeq
-  \ctxwk{Q}{\subst{\idtm{{A}{P}}}{\ctxwk{\ctxext{A}{P}}{\idtm{P}}}}
-  \\
-& \jdeq
-  \ctxwk{Q}{\idtm{P}}
-  \\
-& \jdeq
-  \unfoldall{\jcomp{{P}{Q}}{\cprojsndf{A}{{P}{Q}}}{\cprojfstf{P}{Q}}}
-  \\
-& \jdeq
-  \jcomp{{P}{Q}}{\cprojsndf{A}{{P}{Q}}}{\cprojfstf{P}{Q}}
-\end{align*}
-To prove the judgmental equality
-\begin{equation*}
-\jhomeq
-  {{{\Gamma}{A}}{P}}
-  {Q}
-  {Q}
-  {\cprojsndf{{A}{P}}{Q}}
-  {\jcomp{}{\cprojsndf{A}{{P}{Q}}}{\cprojsndf{P}{Q}}}
-\end{equation*}
-note that we have the judgmental equalities
-\begin{align*}
-\cprojsndf{{A}{P}}{Q}
-& \jdeq
-  \unfoldall{\cprojsndf{{A}{P}}{Q}}
-  \\
-& \jdeq 
-  \unfoldall{\jcomp{{P}{Q}}{\cprojsndf{A}{{P}{Q}}}{\cprojsndf{P}{Q}}}
-  \\
-& \jdeq
-  \jcomp{}{\cprojsndf{A}{{P}{Q}}}{\cprojsndf{P}{Q}}\qedhere
-\end{align*}
-\begin{comment}
-%%%% This was a proof of an old version of the statement
-Now consider a term $\jterm{\Gamma}{\ctxext{A}{{P}{Q}}}{w}$. Then we have
-\begin{align*}
-w 
-& \jdeq 
-  \tmext
-    {A}
-    {{P}{Q}}
-    {\cprojfst{A}{\ctxext{P}{Q}}{w}}
-    {\cprojsnd{A}{\ctxext{P}{Q}}{w}}
-  \\
-& \jdeq 
-  \tmext
-    {A}
-    {{P}{Q}}
-    {\cprojfst{A}{\ctxext{P}{Q}}{w}}
-    { {P}
-      {Q}
-      {\cprojfst{P}{Q}{\cprojsnd{A}{\ctxext{P}{Q}}{w}}}
-      {\cprojsnd{P}{Q}{\cprojsnd{A}{\ctxext{P}{Q}}{w}}}
-      }
-  \\
-& \jdeq 
-  \tmext
-    {{A}{P}}
-    {Q}
-    { {} % need to provide base and family, but there's no unfold.
-      {}
-      {\cprojfst{A}{\ctxext{P}{Q}}{w}}
-      {\cprojfst{P}{Q}{\cprojsnd{A}{\ctxext{P}{Q}}{w}}}
-      }
-    { \cprojsnd{P}{Q}{\cprojsnd{A}{\ctxext{P}{Q}}{w}}
-      }
-\end{align*}
-Thus we see that 
-\begin{align*}
-\cprojfst{\ctxext{A}{P}}{Q}{w} 
-& \jdeq 
-  \tmext
-    {A}
-    {P}
-    {\cprojfst{A}{\ctxext{P}{Q}}{w}}
-    {\cprojfst{P}{Q}{\cprojsnd{A}{\ctxext{P}{Q}}{w}}}
-  \\ 
-\cprojsnd{\ctxext{A}{P}}{Q}{w} 
-& \jdeq 
-  \cprojsnd{P}{Q}{\cprojsnd{A}{\ctxext{P}{Q}}{w}},
-\end{align*}
-proving the fourth judgmental equality, and therefore also that
-\begin{align*}
-\cprojfst{A}{P}{\cprojfst{\ctxext{A}{P}}{Q}{w}} 
-& \jdeq 
-  \cprojfst{A}{\ctxext{P}{Q}}{w}
-  \\
-\cprojsnd{A}{P}{\cprojfst{\ctxext{A}{P}}{Q}{w}} 
-& \jdeq 
-  \cprojfst{P}{Q}{\cprojsnd{A}{\ctxext{P}{Q}}{w}},
-\end{align*}
-proving the second and the third judgmental equalities.
-\end{comment}
-\end{proof}
-
-\begin{lem}\label{lem:tmext-wk}\label{comp-we-t}
-When we weaken a term $\tmext{B}{Q}{y}{v}$ of $\ctxext{B}{Q}$ in context $\Gamma$ by
-a family $A$, the term that we get is $\tmext{\ctxwk{A}{B}}{\ctxwk{A}{Q}}{\ctxwk{A}{y}}{\ctxwk{A}{v}}$. More
-precisely, the following inference rules are valid:
-\begin{align*}
-& \inference
-  { \jterm{{\Gamma}{B}}{Q}{g}
-    \jterm{{\Gamma}{B}}{\subst{g}{R}}{t}
-    }
-  { \jtermeq
-      {{{\Gamma}{A}}{\ctxwk{A}{B}}}
-      {\ctxwk{A}{\ctxext{Q}{R}}}
-      {\ctxwk{A}{\tmext{Q}{R}{g}{t}}}
-      {\tmext{\ctxwk{A}{Q}}{\ctxwk{A}{R}}{\ctxwk{A}{g}}{\ctxwk{A}{t}}}
-    }
-  \\
-& \inference
-  { \jfam{{{\Gamma}{B}}{Q}}{R}
-    }
-  { \jhomeq
-      {{{\Gamma}{A}}{\ctxwk{A}{B}}}
-      {{\ctxwk{A}{Q}}{\ctxwk{A}{R}}}
-      {\ctxwk{A}{Q}}
-      {\cprojfstf{\ctxwk{A}{Q}}{\ctxwk{A}{R}}}
-      {\ctxwk{A}{\cprojfstf{Q}{R}}}
-    }
-  \\
-& \inference
-  { \jfam{{{\Gamma}{B}}{Q}}{R}
-    }
-  { \jhomeq
-      {{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{Q}}}
-      {\ctxwk{A}{R}}
-      {\ctxwk{A}{R}}
-      {\cprojsndf{\ctxwk{A}{Q}}{\ctxwk{A}{R}}}
-      {\ctxwk{A}{\cprojsndf{Q}{R}}}
-    }
-\end{align*}
-\end{lem}
-
-\begin{proof}
-Consider $\jterm{{\Gamma}{B}}{Q}{g}$ and $\jterm{{\Gamma}{B}}{\subst{g}{R}}{t}$.
-Then we have the judgmental equalities
-\begin{align*}
-\ctxwk{A}{\ctxext{Q}{R}{g}{t}}
-& \jdeq 
-  \ctxwk{A}{\subst{t}{{g}{\idtm{\ctxext{Q}{R}}}}}
-  \\
-& \jdeq 
-  \subst{\ctxwk{A}{t}}{\ctxwk{A}{\subst{g}{\idtm{\ctxext{Q}{R}}}}}
-  \\
-& \jdeq 
-  \subst{\ctxwk{A}{t}}{{\ctxwk{A}{g}}{\ctxwk{A}{\idtm{\ctxext{Q}{R}}}}}
-  \\
-& \jdeq 
-  \subst{\ctxwk{A}{t}}{{\ctxwk{A}{g}}{\idtm{\ctxwk{A}{\ctxext{Q}{R}}}}}
-  \\
-& \jdeq 
-  \subst{\ctxwk{A}{t}}{{\ctxwk{A}{g}}{\idtm{\ctxext{\ctxwk{A}{Q}}{\ctxwk{A}{R}}}}}
-  \\
-& \jdeq 
-  \tmext{\ctxwk{A}{Q}}{\ctxwk{A}{R}}{\ctxwk{A}{g}}{\ctxwk{A}{t}}
-\end{align*}
-Next, we want to prove the judgmental equality
-\begin{equation*}
-\jhomeq
-  {{{\Gamma}{A}}{\ctxwk{A}{B}}}
-  {{\ctxwk{A}{Q}}{\ctxwk{A}{R}}}
-  {\ctxwk{A}{Q}}
-  {\cprojfstf{\ctxwk{A}{Q}}{\ctxwk{A}{R}}}
-  {\ctxwk{A}{\cprojfstf{Q}{R}}}
-\end{equation*}
-Note that we have the judgmental equalities
-\begin{align*}
-\cprojfstf{\ctxwk{A}{Q}}{\ctxwk{A}{R}}
-& \jdeq
-  \unfoldall{\cprojfstf{\ctxwk{A}{Q}}{\ctxwk{A}{R}}}
-  \\
-& \jdeq
-  \ctxwk{{A}{R}}{{A}{\idtm{Q}}}
-  \\
-& \jdeq
-  \unfoldall{\ctxwk{A}{\cprojfstf{Q}{R}}}
-  \\
-& \jdeq
-  \ctxwk{A}{\cprojfstf{Q}{R}}.
-\end{align*}
-Finally, we want to prove the judgmental equality
-\begin{equation*}
-\jhomeq
-  {{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{Q}}}
-  {\ctxwk{A}{R}}
-  {\ctxwk{A}{R}}
-  {\cprojsndf{\ctxwk{A}{Q}}{\ctxwk{A}{R}}}
-  {\ctxwk{A}{\cprojsndf{Q}{R}}}
-\end{equation*}
-Note that we have the judgmental equalities
-\begin{align*}
-\cprojsndf{\ctxwk{A}{Q}}{\ctxwk{A}{R}}
-& \jdeq
-  \unfoldall{\cprojsndf{\ctxwk{A}{Q}}{\ctxwk{A}{R}}}
-  \\
-& \jdeq
-  \unfoldall{\ctxwk{A}{\cprojsndf{Q}{R}}}
-  \\
-& \jdeq
-  \ctxwk{A}{\cprojsndf{Q}{R}}.
-  \qedhere
-\end{align*}
-\end{proof}
-
-\begin{lem}\label{lem:tmext-subst}\label{comp-se-t}
-When we substitute an extended term $\tmext{P}{Q}{f}{g}$ of $\ctxext{P}{Q}$ by a term
-$x$ of $A$, the term that we get is $\tmext{\subst{x}{P}}{\subst{x}{Q}}{\subst{x}{f}}{\subst{x}{g}}$.
-More precisely, the following inference rules are valid:
-\begin{align*}
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jterm{{{\Gamma}{A}}{P}}{Q}{g}
-    \jterm{{{\Gamma}{A}}{P}}{\subst{g}{R}}{t}
-    }
-  { \jtermeq
-      {{\Gamma}{\subst{x}{P}}}
-      {\ctxext{\subst{x}{Q}}{\subst{x}{R}}}
-      {\subst{x}{\tmext{Q}{R}{g}{t}}}
-      {\tmext{\subst{x}{Q}}{\subst{x}{R}}{\subst{x}{g}}{\subst{x}{t}}}
-    }
-  \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jfam{{{{\Gamma}{A}}{P}}{Q}}{R}
-    }
-  { \jhomeq
-      {{\Gamma}{\subst{x}{P}}}
-      {{\subst{x}{Q}}{\subst{x}{R}}}
-      {\subst{x}{Q}}
-      {\cprojfstf{\subst{x}{Q}}{\subst{x}{R}}}
-      {\subst{x}{\cprojfstf{Q}{R}}}
-    }
-  \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jfam{{{{\Gamma}{A}}{P}}{Q}}{R}
-    }
-  { \jhomeq
-      {{{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}}
-      {\subst{x}{R}}
-      {\subst{x}{R}}
-      {\cprojsndf{\subst{x}{Q}}{\subst{x}{R}}}
-      {\subst{x}{\cprojsndf{Q}{R}}}
-    }
-\end{align*}
-\end{lem}
-
-\begin{proof}
-Consider $\jterm{{\Gamma}{B}}{Q}{g}$ and $\jterm{{\Gamma}{B}}{\subst{g}{R}}{t}$.
-Then we have the judgmental equalities
-\begin{align*}
-\subst{x}{\tmext{Q}{R}{g}{t}}
-& \jdeq 
-  \subst{x}{{t}{{g}{\idtm{\ctxext{Q}{R}}}}}
-  \\
-& \jdeq 
-  \subst{{x}{t}}{{x}{{g}{\idtm{\ctxext{Q}{R}}}}}
-  \\
-& \jdeq 
-  \subst{{x}{t}}{{{x}{g}}{{x}{\idtm{\ctxext{Q}{R}}}}}
-  \\
-& \jdeq 
-  \subst{{x}{t}}{{{x}{g}}{\idtm{\subst{x}{\ctxext{Q}{R}}}}}
-  \\
-& \jdeq 
-  \subst{{x}{t}}{{{x}{g}}{\idtm{\ctxext{\subst{x}{Q}}{\subst{x}{R}}}}}
-  \\
-& \jdeq 
-  \tmext{\subst{x}{Q}}{\subst{x}{R}}{\subst{x}{g}}{\subst{x}{t}}.
-\end{align*}
-Next, we want to prove the judgmental equality
-\begin{equation*}
-\jhomeq
-  {{\Gamma}{\subst{x}{P}}}
-  {{\subst{x}{Q}}{\subst{x}{R}}}
-  {\subst{x}{Q}}
-  {\cprojfstf{\subst{x}{Q}}{\subst{x}{R}}}
-  {\subst{x}{\cprojfstf{Q}{R}}}
-\end{equation*}
-Note that we have the judgmental equalities
-\begin{align*}
-\cprojfstf{\subst{x}{Q}}{\subst{x}{R}}
-& \jdeq
-  \unfoldall{\cprojfstf{\subst{x}{Q}}{\subst{x}{R}}}
-  \\
-& \jdeq
-  \ctxwk{\subst{x}{R}}{\subst{x}{\idtm{Q}}}
-  \\
-& \jdeq
-  \unfoldall{\subst{x}{\cprojfstf{Q}{R}}}
-  \\
-& \jdeq
-  \subst{x}{\cprojfstf{Q}{R}}.
-\end{align*}
-And finally we want to prove the judgmental equality
-\begin{equation*}
-\jhomeq
-  {{{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}}
-  {\subst{x}{R}}
-  {\subst{x}{R}}
-  {\cprojsndf{\subst{x}{Q}}{\subst{x}{R}}}
-  {\subst{x}{\cprojsndf{Q}{R}}}
-\end{equation*}
-Note that we have the judgmental equalities
-\begin{align*}
-\cprojsndf{\subst{x}{Q}}{\subst{x}{R}}
-& \jdeq
-  \unfoldall{\cprojsndf{\subst{x}{Q}}{\subst{x}{R}}}
-  \\
-& \jdeq
-  \unfoldall{\subst{x}{\cprojsndf{Q}{R}}}
-  \\
-& \jdeq
-  \subst{x}{\cprojsndf{Q}{R}}.
-  \qedhere
-\end{align*}
-\end{proof}
-
-We find the following inference rule, which expresses that the identity term
-is compatible with extension:
-
-\begin{lem}\label{lem:tmext-id}\label{comp-ie}
-For any $\jfam{\Gamma}{A}$ and $\jfam{{\Gamma}{A}}{P}$ we have
-\begin{equation}\label{idfunc-ext-comp}
-\inference
-  { \jfam{\Gamma}{A}
-    \jfam{{\Gamma}{A}}{P}
-    }
-  { \jhomeq
-      {\Gamma}
-      {{A}{P}}{{A}{P}}
-      {\idtm{\ctxext{A}{P}}}
-      { \tmext
-          {\ctxwk{\ctxext{A}{P}}{A}}
-          {\ctxwk{\ctxext{A}{P}}{P}}
-          {\cprojfstf{A}{P}}
-          {\cprojsndf{A}{P}}
-        }
-    }
-\end{equation}
-\end{lem}
-
-\begin{proof}
-Consider the families $\jfam{\Gamma}{A}$ and $\jfam{{\Gamma}{A}}{P}$. Then
-we have the judgmental equalities
-\begin{align*}
-\tmext
-  {\ctxwk{\ctxext{A}{P}}{A}}
-  {\ctxwk{\ctxext{A}{P}}{P}}
-  {\cprojfstf{A}{P}}
-  {\cprojsndf{A}{P}}
-& \jdeq 
-  \unfold
-  { \tmext
-      {\ctxwk{\ctxext{A}{P}}{A}}
-      {\ctxwk{\ctxext{A}{P}}{P}}
-      {\cprojfstf{A}{P}}
-      {\cprojsndf{A}{P}}
-    }
-  \\
-& \jdeq 
-  \subst
-    { \idtm{P}
-      }
-    { {\ctxwk{P}{\idtm{A}}}
-      {\idtm{\ctxwk{\ctxext{A}{P}}{\ctxext{A}{P}}}}
-      }
-  \\
-& \jdeq 
-  \subst
-    { \idtm{P}
-      }
-    { {\ctxwk{P}{\idtm{A}}}
-      {\ctxwk{\ctxext{A}{P}}{\idtm{\ctxext{A}{P}}}}
-      }
-  \\
-& \jdeq 
-  \subst
-    { \idtm{P}
-      }
-    { {\ctxwk{P}{\idtm{A}}}
-      {\ctxwk{P}{{A}{\idtm{\ctxext{A}{P}}}}}
-      }
-  \\
-& \jdeq 
-  \subst
-    { \idtm{P}
-      }
-    {\ctxwk
-      {P}
-      { \subst
-        {\idtm{A}}
-        {\ctxwk{A}{\idtm{\ctxext{A}{P}}}}
-        }
-      }
-  \\
-& \jdeq 
-  \subst
-    { \idtm{A}
-      }
-    { \ctxwk{A}{\idtm{\ctxext{A}{P}}}
-      }
-  \\
-& \jdeq 
-  \idtm{\ctxext{A}{P}}.
-  \qedhere
-\end{align*}
-\end{proof}
-
-\subsection{More on morphisms}
 
@@
 \subsection{Inductive morphisms}
+In this subsection we investigate the notion of inductive morphisms. An 
+inductive morphism $f$ from $A$ to $B$ in context $\Gamma$ is a morphism which
+induces an operation which is judgmentally the right inverse of composition
+with $f$. We explore a notion of inductive morphism which is much stronger
+than the usual notion: it pushes families over $A$ to families over $B$,
+families over families over $A$ to families over families over $B$ and terms
+thereof to terms of the output families over families in a manner compatible
+with the empty family, extension, weakening, substitution and identity terms.
+As a result, inductive morphisms will be stable under extension, weakening,
+substitution and the identity term is an inductive morphism.
+
 \begin{defn}
@@
 \begin{proof}
-The proof is a rather long computation:
+The proof is a rather long computation. It was found by starting with the
+application of \autoref{eq:ealg-eq1} and making the computation in both ways
+from there. The judgmental equality below where \autoref{eq:ealg-eq1} is applied,
+is marked with an $*$. The computation goes as follows:
 \begin{align*}
@@
   \\
-& \jdeq
+& \stackrel{*}{\jdeq}
   \jcomp{}{
@@
 
+\begin{proof}
+The proof is a straightforward calculation (start in the middle):
+\begin{align*}
+\subst{{y_2}{{y_1}{\jcomp{}{e}{f}}}}{{y_0}{f}}
+& \jdeq
+  \subst{\tmext{y_0}{\subst{y_2}{{y_1}{\jcomp{}{e}{f}}}}}{f}
+  \\
+& \jdeq
+  \subst{{y_2}{{y_1}{{y_0}{\jvcomp{}{\idtm{P}}{\jcomp{}{e}{f}}}}}}{f}
+  \\
+& \jdeq
+  \subst
+    {y_2}
+    { {y_1}
+      { {y_0}
+        {\jcomp{}{\jvcomp{}{\idtm{P}}{\jcomp{}{e}{f}}}{f}}
+        }
+      }
+  \\
+& \jdeq
+  \subst
+    {y_2}
+    { {y_1}
+      { {y_0}
+        {\jcomp{}{\jvcomp{}{f}{\idtm{\jcomp{}{f}{{}{e}{P}}}}}{f}}
+        }
+      }
+  \\
+& \jdeq
+  \subst{{y_2}{{y_1}{{y_0}{\jvcomp{}{f}{\idtm{\jcomp{}{f}{{}{e}{P}}}}}}}}{f}
+  \\
+& \jdeq
+  \subst{\tmext{\subst{y_1}{{y_0}{f}}}{y_2}}{f}
+  \\
+& \jdeq
+  \subst{y_2}{{{y_1}{{y_0}{f}}}{f}}.\qedhere
+\end{align*}
+\end{proof}
+
 \subsection{Extension-empty algebras}
@@
 \end{defn}
+
+\begin{defn}
+Let $(A_0,P_0,x_0,f_0)$ be an empty algebra in context $\Gamma$ and consider
+\begin{enumerate}
+\item a family $\jfam{{\Gamma}{A_0}}{A_1}$.
+\item a family $\jfam{{{{\Gamma}{A_0}}{P_0}}{\ctxwk{P_0}{A_1}}}{P_1}$.
+\end{enumerate}
+\end{defn}

### 2014-05-10 article-models-ttderived.tex
starting to write the dependent logic of extension algebras

@@
 \begin{proof}
-The proof is a straightforward calculation (start in the middle):
+The proof is a straightforward calculation:
 \begin{align*}
@@
 \end{proof}
+
+Extension algebras don't come in isolation. There are also extension algebra
+families and extension algebra terms. We now aim to define these and to
+establish various constructions of new extension algebras out of old ones:
+the empty extension algebra, and extensions, weakenings and substitutions
+of extension algebras and of course the identity term as an extension algebra
+term. We start with extension algebra families.
+
+Consider an extension algebra $\mathcal{A}\defeq(A,P,e,f)$. 
+An extension algebra family over $\mathcal{A}$ is likewise a quadruple
+$\mathcal{B}\defeq(B,Q,g,h)$. Here we have a family $\jfam{{\Gamma}{A}}{B}$, a
+family $\jfam{{{{\Gamma}{A}}{P}}{\jcomp{}{e}{B}}}{Q}$ and
+\begin{align*}
+\jhom*{{{\Gamma}{A}}{P}}{\ctxext{\jcomp{}{e}{B}}{Q}}{\jcomp{}{e}{B}}{g}\\
+\jhom*{{{{\Gamma}{A}}{P}}{\jcomp{}{e}{B}}}{\ctxext{Q}{\jcomp{}{g}{Q}}}{Q}{h}.
+\end{align*}
+The quadruple $(\jcomp{}{e}{B},Q,g,h)$ is required to be an extension algebra
+in context $\ctxext{{\Gamma}{A}}{P}$.
+
+Suppose $\mathcal{A}$ is an extension algebra and $\mathcal{B}$ is an extension
+algebra family over $\mathcal{A}$. A term of $\mathcal{B}$ is a pair $(x,y)$
+consisting of
+\begin{align*}
+\jterm*{{\Gamma}{A}}{B}{x}\\
+\jterm*{{{\Gamma}{A}}{P}}{\subst{\jcomp{}{e}{x}}{Q}}{y}
+\end{align*}
+such that... (compatibility of the extensions)
 

### 2014-05-10 article-models-ttderived.tex
strong inductive morphisms

@@
 
-\subsection{Inductive morphisms}
-In this subsection we investigate the notion of inductive morphisms. An 
-inductive morphism $f$ from $A$ to $B$ in context $\Gamma$ is a morphism which
-induces an operation which is judgmentally the right inverse of composition
-with $f$. We explore a notion of inductive morphism which is much stronger
-than the usual notion: it pushes families over $A$ to families over $B$,
-families over families over $A$ to families over families over $B$ and terms
-thereof to terms of the output families over families in a manner compatible
-with the empty family, extension, weakening, substitution and identity terms.
-As a result, inductive morphisms will be stable under extension, weakening,
-substitution and the identity term is an inductive morphism.
-
-\begin{defn}
-Let $\jhom{\Gamma}{A}{B}{f}$ be a context morphism. We say that $f$ is an 
-\emph{inductive morphism} if the following inference rules are valid:
-\begin{align*}
-& \inference
-  { \jfam{{\Gamma}{B}}{Q}
-    \jterm{{\Gamma}{A}}{\jcomp{A}{f}{Q}}{g}
-    }
-  { \jterm{{\Gamma}{B}}{Q}{\jtcext{g}}
-    }
-& & \inference
-  { \jfam{{\Gamma}{B}}{Q}
-    \jterm{{\Gamma}{A}}{\jcomp{A}{f}{Q}}{g}
-    }
-  { \jtermeq{{\Gamma}{A}}{\jcomp{A}{f}{Q}}{\jcomp{A}{f}{\jtcext{g}}}{g}
-    }
-\end{align*}
-\end{defn}
-
-When $f$ is a context morphism from $A$ to $B$ in context $\Gamma$, 
-finding a term of a family $Q$ over $\ctxext{\Gamma}{B}$ can be accomplished
-by finding a term of the family $\jcomp{A}{f}{Q}$ over $\ctxext{\Gamma}{A}$.
-Inductive morphisms come with a computation rule. Thus, term
-constructors of inductively defined types are going to be the major source of
-examples of inductive morphisms. Since identity terms behave
-so nicely, they are inductive morphisms too.
-
-\begin{lem}
-The identity term
-$\jhom{\Gamma}{A}{A}{\idtm{A}}$ is an inductive morphism
-for each family $A$ of contexts over $\Gamma$
-\end{lem}
-
-\begin{itemize}
-\item Extensions of inductive morphisms are inductive
-\item Weakenings of inductive morphisms are inductive
-\item Substitutions of inductive morphisms are inductive
-\end{itemize}
-
 \subsection{Extension algebras}
@@
 
+\begin{defn}
 Consider an extension algebra $\mathcal{A}\defeq(A,P,e,f)$. 
@@
 in context $\ctxext{{\Gamma}{A}}{P}$.
-
+\end{defn}
+
+\begin{defn}
 Suppose $\mathcal{A}$ is an extension algebra and $\mathcal{B}$ is an extension
@@
 such that... (compatibility of the extensions)
+\end{defn}
 

### 2014-05-10 article-models-ttderived.tex
pretty diagram :)

@@
 of the ingredients) of the theory of contexts, families and terms in its
-formulation.
+formulation and it is (perhaps surprisingly) quite involved to formulate it.
+It is however a necessary task, because extension algebras stand at the basis
+of implementing any theory with dependency in the theory of contexts, families
+and terms (i.e.~you don't stand a chance without extension algebras).
 
@@
 \end{align*}
-such that... (compatibility of the extensions)
+such that the diagrams
+\begin{equation*}
+\begin{tikzcd}
+\ctxext{\jcomp{}{e}{B}}{Q} 
+  \ar{r}{g} 
+  \ar[shift right=.7ex,fib]{d}
+& B 
+  \ar[shift right=.7ex,fib]{d} 
+  \\
+\ctxext{A}{P} 
+  \ar[shift right=.7ex,dotted]{u}[swap]{\tmext{\jcomp{}{e}{x}}{y}}
+  \ar{r}{e}
+& A
+  \ar[shift right=.7ex,dotted]{u}[swap]{x}
+\end{tikzcd}
+\end{equation*}
+commute
+\end{defn}
+
+\begin{rmk}
+The compatibility 
+\end{rmk}
+
+\begin{defn}
+Suppose $\mathcal{A}$ and $\mathcal{B}$ are extension algebras in context
+$\Gamma$. We define the extension algebra $\ctxwk{\mathcal{A}}{\mathcal{B}}$
+to be the quadruple
+\begin{equation*}
+(\ctxwk{A}{B},\ctxwk{\ctxext{A}{P}}{Q},\ctxwk{\ctxext{A}{P}}{g},\ctxwk{\ctxext{A}{P}}{h}).
+\end{equation*}
+Note that $\ctxwk{\ctxext{A}{P}}{Q}$ is a family over $\ctxwk{\ctxext{A}{P}}{B}$,
+whereas it should be a family over $\jcomp{}{e}{\ctxwk{A}{B}}$. These are the
+same by \autoref{lem:prehom}.
 \end{defn}

### 2014-05-10 article-models-ttderived.tex
another fancy diagram

@@
 \end{equation*}
-commute
+and
+\begin{equation*}
+\begin{tikzcd}
+\jcomp{}{f}{\ctxext{Q}{\jcomp{}{g}{Q}}}
+  \ar{r}{\jcomp{}{f}{h}}
+  \ar[shift right=.7ex,fib]{d}
+& Q
+  \ar[shift right=.7ex,fib]{d}
+  \\
+\jcomp{}{f}{\jcomp{}{e}{B}}
+  \ar{r}{\idtm{\jcomp{}{f}{\jcomp{}{e}{B}}}}
+  \ar[shift right=.7ex,fib]{d}
+  \ar[shift right=.7ex,dotted,mapsto]{u}[swap]{\jcomp{}{f}{y}}
+& \jcomp{}{e}{B}
+  \ar[shift right=.7ex,fib]{d}
+  \ar[shift right=.7ex,dotted,mapsto]{u}[swap]{y}
+  \\
+\ctxext{P}{\jcomp{}{e}{P}}
+  \ar{r}[swap]{f}
+  \ar[shift right=.7ex,dotted]{u}[swap]{\jcomp{}{f}{\jcomp{}{e}{x}}}
+& P
+  \ar[shift right=.7ex,dotted]{u}[swap]{\jcomp{}{e}{x}}
+\end{tikzcd}
+\end{equation*}
+commute.
 \end{defn}
@@
 \end{defn}
+
+\subsection{Extension-weakening algebras}
+
+\subsection{Extension-substitution algebras}
+
+\subsection{Pre-universes}
+Pre-universes are internal versions of the theory of contexts, families and
+terms. They interpret extension, the empty context, weakening, substitution
+and identity terms all at once in a compatible way.

### 2014-05-11 article-models-ttderived.tex
tagging equations

@@
 Let $\jfam{\Gamma}{A}$, $\jfam{\Gamma}{B}$, $\jfam{\Gamma}{C}$ and $\jhom{\Gamma}{A}{B}{f}$.
-Then we have the judgmental equalities\begin{align*}
+Then we have the judgmental equalities
+\begin{align*}
 \subst{f}{\ctxwk{A}{{B}{C}}}
@@
   \subst{f}{\ctxwk{{A}{B}}{{A}{C}}}
-  \\
+  \tag{by \autoref{comp-ww-f}}\\
 & \jdeq 
   \ctxwk{A}{C}.
-  \qedhere
+  \tag{by \autoref{cancellation-ws-t}}
 \end{align*}
@@
 For the first, note that we have the judgmental equalities
-\begin{equation*}
+\begin{align*}
 \jcomp{A}{f}{\emptyf}
-  \jdeq
+& \jdeq
   \unfold{\jcomp{A}{f}{\emptyf}}
-  \jdeq
+  \tag{by definition}
+  \\
+& \jdeq
   \subst{f}{\emptyf}
-  \jdeq
+  \tag{by \autoref{comp-w0-c}}
+  \\
+& \jdeq
   \emptyf.
-\end{equation*}
+  \tag{by \autoref{comp-s0-c}}
+\end{align*}
 The second judgmental equality is proven similarly.
@@
   \unfold{\jcomp{A}{f}{\ctxext{R}{S}}}
+  \tag{by definition}
   \\
@@
   \subst{f}{\ctxext{\ctxwk{A}{R}}{\ctxwk{A}{S}}}
+  \tag{by \autoref{comp-we-f}}
   \\
@@
   \unfoldall{\ctxext{\jcomp{A}{f}{R}}{\jcomp{A}{f}{S}}}
-  \\
-& \jdeq
-  \ctxext{\jcomp{A}{f}{R}}{\jcomp{A}{f}{S}}.\qedhere
+  \tag{by \autoref{comp-se-f}}
+  \\
+& \jdeq
+  \ctxext{\jcomp{A}{f}{R}}{\jcomp{A}{f}{S}}.
+  \tag{by definition}
 \end{align*}
@@
   \unfold{\jcomp{A}{f}{\ctxwk{R}{S}}}
+  \tag{by definition}
   \\
@@
   \subst{f}{\ctxwk{{A}{R}}{{A}{S}}}
+  \tag{by \autoref{comp-ww-f}}
   \\
@@
   \unfoldall{\ctxwk{\jcomp{A}{f}{R}}{\jcomp{A}{f}{S}}}
+  \tag{by \autoref{comp-sw-f}}
   \\
@@
   \ctxwk{\jcomp{A}{f}{R}}{\jcomp{A}{f}{S}}.
+  \tag{by definition}
 \end{align*}
@@
   \unfold{\jcomp{A}{f}{\subst{h}{S}}}
+  \tag{by definition}
   \\
@@
   \subst{f}{{\ctxwk{A}{h}}{\ctxwk{A}{S}}}
+  \tag{by \autoref{comp-ws-f}}
   \\
@@
   \unfoldall{\subst{\jcomp{A}{f}{h}}{\jcomp{A}{f}{S}}}
+  \tag{by \autoref{comp-ss-f}}
   \\
@@
   \subst{\jcomp{A}{f}{h}}{\jcomp{A}{f}{S}}.
+  \tag{by definition}
 \end{align*}
@@
   \unfold{\jcomp{A}{f}{\idtm{R}}}
+  \tag{by definition}
   \\
@@
   \subst{f}{\idtm{\ctxwk{A}{R}}}
+  \tag{by \autoref{comp-wi-t}}
   \\
@@
   \unfoldall{\idtm{\jcomp{A}{f}{R}}}
-  \\
-& \jdeq
-  \idtm{\jcomp{A}{f}{R}}.\qedhere
+  \tag{by \autoref{comp-si-t}}
+  \\
+& \jdeq
+  \idtm{\jcomp{A}{f}{R}}.
+  \tag{by definition}
 \end{align*}
@@
   \subst{{f}{\ctxwk{A}{g}}}{R}
+  \tag{by definition}
   \\
@@
   \subst{{f}{\ctxwk{A}{g}}}{\subst{f}{\ctxwk{{A}{B}}{R}}}
-  \\
-& \jdeq 
-  \subst{f}{{\ctxwk{A}{g}}{\ctxwk{{A}{B}}{R}}}
+  \tag{by \autoref{cancellation-ws-f}}
+  \\
+& \jdeq 
+  \subst{f}{{\ctxwk{A}{g}}{\ctxwk{{A}{B}}{R}}}.
+  \tag{by \autoref{comp-ss-f}}
 \end{align*}
@@
   \subst{\jcomp{A}{f}{g}}{\ctxwk{A}{R}}
+  \tag{by definition}
   \\
@@
   \subst{{f}{\ctxwk{A}{g}}}{\ctxwk{A}{R}}
+  \tag{by definition}
+  \\
+& \jdeq
+  \subst{{f}{\ctxwk{A}{g}}}{{f}{\ctxwk{{A}{B}}{{A}{R}}}}
+  \tag{by \autoref{cancellation-ws-t}}
   \\
@@
   \subst{f}{{\ctxwk{A}{g}}{\ctxwk{{A}{B}}{{A}{R}}}}
+  \tag{by \autoref{comp-ss-f}}
   \\
@@
   \subst{f}{{\ctxwk{A}{g}}{\ctxwk{A}{{B}{R}}}}
+  \tag{by \autoref{comp-ww-f}}
   \\
@@
   \subst{f}{\ctxwk{A}{\subst{g}{\ctxwk{B}{R}}}}
+  \tag{by \autoref{comp-ws-f}}
   \\
@@
   \subst{f}{\ctxwk{A}{\jcomp{B}{g}{R}}}
+  \tag{by definition}
   \\
@@
   \jcomp{A}{f}{{B}{g}{R}}.
+  \tag{by definition}
 \end{align*}
@@
   \ctxwk{P}{\subst{f}{\ctxwk{A}{g}}}
+  \tag{by definition}
   \\
@@
   \subst{\ctxwk{P}{f}}{\ctxwk{P}{{A}{g}}}
+  \tag{by \autoref{comp-ws-t}}
   \\
@@
   \subst{\ctxwk{P}{f}}{\ctxwk{\ctxext{A}{P}}{g}}
+  \tag{by \autoref{comp-ew-t}}
   \\
@@
   \jcomp{{A}{P}}{\ctxwk{P}{f}}{g}.
+  \tag{by definition}
 \end{align*}
@@
   \subst{\ctxwk{A}{y}}{\ctxwk{A}{g}}
+  \tag{by definition}
   \\
@@
   \ctxwk{A}{\subst{y}{g}}.
+  \tag{by \autoref{comp-ws-t}}
 \end{align*}
@@
   \subst{f}{\ctxwk{A}{{B}{z}}}
+  \tag{by definition}
   \\
@@
   \subst{f}{\ctxwk{{A}{B}}{{A}{z}}}
+  \tag{by \autoref{comp-ww-t}}
   \\
@@
   \ctxwk{A}{z}.
-  \qedhere
+  \tag{by \autoref{cancellation-ws-t}}
 \end{align*}
@@
   \subst{x}{{f}{\ctxwk{P}{g}}}
+  \tag{by definition}
   \\
@@
   \subst{{x}{f}}{{x}{\ctxwk{P}{g}}}
+  \tag{by \autoref{comp-ss-t}}
   \\
@@
   \subst{{x}{f}}{\ctxwk{\subst{x}{P}}{\subst{x}{g}}}
+  \tag{by \autoref{comp-sw-t}}
   \\
@@
   \jcomp{\subst{x}{P}}{\subst{x}{f}}{\subst{x}{g}}.
+  \tag{by definition}
 \end{align*}
@@
   \subst{x}{{f}{\ctxwk{A}{g}}}
+  \tag{by definition}
   \\
@@
   \subst{{x}{f}}{{x}{\ctxwk{A}{g}}}
+  \tag{by \autoref{comp-ss-t}}
   \\
@@
   \subst{{x}{f}}{g}.
-  \qedhere
+  \tag{by \autoref{cancellation-ws-t}}
 \end{align*}
@@
     {{x}{\ctxwk{A}{Q}}}
-  \tag{by \autoref{defn-ws-3}}\\
+  \tag{by \autoref{cancellation-ws-f}}\\
 & \jdeq 
@@
     {{u}{\ctxwk{\subst{x}{P}}{\subst{x}{\ctxwk{A}{Q}}}}}
-  \tag{by \autoref{defn-ws-3}}\\
+  \tag{by \autoref{cancellation-ws-f}}\\
 & \jdeq 
@@
 & \jdeq
-  \ctxwk{P}{{A}{\ctxext{A}{P}}}
+  \ctxext{\ctxwk{\ctxext{A}{P}}{A}}{\ctxwk{\ctxext{A}{P}}{P}}
+  \tag{by \autoref{comp-we-f}}
   \\
@@
   \ctxext{\ctxwk{P}{{A}{A}}}{\ctxwk{P}{{A}{P}}}
+  \tag{by \autoref{comp-ew-f}}
 \end{align*}
@@
       }
-  \tag{by \autoref{idfunc-wk-comp}}\\
+  \tag{by \autoref{comp-wi-t}}\\
 & \jdeq 
@@
       }
-  \tag{by \autoref{defn-ws-4}}\\
+  \tag{by \autoref{cancellation-ws-t}}\\
 & \jdeq 
@@
 \end{align*}
+
+\begin{rmk}
+Instead of looking at $f$ as a context morphism from $\ctxext{P}{\jcomp{}{e}{P}}$
+to $P$, one could also look at $f$ as a morphism \emph{over $\cprojfstf{A}{P}$},
+as indicated in the following diagram:
+\begin{equation*}
+\begin{tikzcd}
+P
+  \ar[fib]{d}
+& \jcomp{}{e}{P}
+  \ar[fib]{d}
+  \ar{l}[swap]{f}
+  \ar{r}
+& P
+  \ar[fib]{d}
+  \\
+A
+& \ctxext{A}{P}
+  \ar{l}{\cprojfstf{A}{P}}
+  \ar{r}[swap]{e}
+& A
+\end{tikzcd}
+\end{equation*}
+This makes it clear that $f$ takes a family over an extended context as an
+argument. The extended context consists of a `base part' and a `family part'. 
+The output of $f$ is a new (extended) family over that base part. Forgetting 
+the family part is what the projection takes care of.
+\end{rmk}
+
 Extension also satisfies the properties explained in \autoref{comp-ee}, so we
@@
 $\mathcal{B}\defeq(B,Q,g,h)$. Here we have a family $\jfam{{\Gamma}{A}}{B}$, a
-family $\jfam{{{{\Gamma}{A}}{P}}{\jcomp{}{e}{B}}}{Q}$ and
-\begin{align*}
-\jhom*{{{\Gamma}{A}}{P}}{\ctxext{\jcomp{}{e}{B}}{Q}}{\jcomp{}{e}{B}}{g}\\
-\jhom*{{{{\Gamma}{A}}{P}}{\jcomp{}{e}{B}}}{\ctxext{Q}{\jcomp{}{g}{Q}}}{Q}{h}.
+family $\jfam{{{{\Gamma}{A}}{P}}{\ctxwk{P}{B}}}{Q}$ and
+\begin{align*}
+\jhom*{{{\Gamma}{A}}{P}}{\ctxext{\ctxwk{P}{B}}{Q}}{\jcomp{}{e}{B}}{g}\\
+\jhom*{{{{\Gamma}{A}}{P}}{\ctxwk{P}{B}}}{\ctxext{Q}{\jcomp{}{g}{Q}}}{Q}{h}.
 \end{align*}
@@
 
-\begin{rmk}
-The compatibility 
-\end{rmk}
-
 \begin{defn}
@@
 \end{defn}
+
+\begin{rmk}
+Before we continue, let us explore what it means to be an extension algebra
+term of the extension algebra $\ctxwk{\mathcal{A}}{\mathcal{B}}$. Such an
+extension algebra term $(x,y)$ would consist of
+\begin{align*}
+\jterm*{{\Gamma}{A}}{\ctxwk{A}{B}}{x}\\
+\jterm*{{{\Gamma}{A}}{P}}{\subst{\jcomp{}{e}{x}}{\ctxwk{\ctxext{A}{P}}{Q}}}{y}.
+\end{align*}
+Thus, $x$ is a context morphism from $A$ to $B$ and $y$ is nothing but a term
+of $\jcomp{}{\jcomp{}{e}{x}}{Q}$. For $x$, we see that the diagram
+\begin{equation*}
+\begin{tikzcd}
+\ctxext{B}{Q} 
+  \ar{r}{g} 
+& B 
+  \\
+\ctxext{A}{P} 
+  \ar{u}{\jvcomp{}{x}{y}}
+  \ar{r}{e}
+& A
+  \ar{u}[swap]{x}
+\end{tikzcd}
+\end{equation*}
+commutes.
+\end{rmk}
 

### 2014-05-13 article-models-ttderived.tex
a few more tags added
commenting stuff so that the article becomes shorter
nearly finished with the section about fiber inclusions

@@
   \ctxwk{P}{\subst{\idtm{A}}{\ctxwk{A}{P}}}
+  \tag{by \autoref{comp-ws-f}}
   \\
@@
   \ctxwk{P}{P},
+  \tag{by \autoref{idfunc-wk-defn}}
 \end{align*}
@@
   \subst{w}{\idtm{\ctxext{A}{P}}} 
-  \tag{by \autoref{idfunc-subst-defn}}\\
+  \tag{by \autoref{cancellation-si}}\\
 & \jdeq 
@@
   \subst{\tmext{A}{P}{x}{u}}{\ctxwk{P}{\idtm{A}}}
+  \tag{by definition}
   \\
@@
   \subst{u}{{x}{\ctxwk{P}{\idtm{A}}}} 
+  \tag{by \autoref{comp-es}}
   \\
@@
   \subst{u}{\ctxwk{\subst{x}{P}}{\subst{x}{\idtm{A}}}}
+  \tag{by \autoref{comp-sw-t}}
   \\
@@
   \subst{x}{\idtm{A}}
+  \tag{by \autoref{cancellation-ws-t}}
   \\
@@
   x.
+  \tag{by \autoref{cancellation-si}}
 \end{align*}
@@
   \subst{\tmext{A}{P}{x}{u}}{\idtm{P}}
+  \tag{by definition}
   \\
@@
   \subst{u}{{x}{\idtm{P}}}
+  \tag{by \autoref{comp-es}}
   \\
@@
   \subst{u}{\idtm{\subst{x}{P}}}
+  \tag{by \autoref{comp-si-t}}
   \\
@@
   u.
-  \qedhere
+  \tag{by \autoref{cancellation-si}}
 \end{align*}
@@
 
+We can immediately show that composing with a fiber inclusion is substitution.
+Thus, the following lemma asserts that we have a family pullback square
+\begin{equation*}
+\begin{tikzcd}
+\subst{x}{Q} \ar[fib]{d} \ar{r} & Q \ar[fib]{d} \\
+\subst{x}{P} \ar{r}[swap]{\finc{x}{P}} & \ctxext{A}{P}
+\end{tikzcd}
+\end{equation*}
+in context $\Gamma$.
+The upper arrow is just the identity morphism over $\finc{x}{P}$.
+
+\begin{lem}\label{lem:finc-precomp}
+Composition with $\finc{x}{P}$ is (the action on families of)
+substitution with $x$:
+\begin{align*}
+& \inference
+  { \jterm{\Gamma}{A}{x}
+    \jfam{{{\Gamma}{A}}{P}}{Q}
+    }
+  { \jfameq
+      {{\Gamma}{\subst{x}{P}}}
+      {\jcomp{\subst{x}{P}}{\finc{x}{P}}{Q}}
+      {\subst{x}{Q}}
+    }
+  \\
+& \inference
+  { \jterm{\Gamma}{A}{x}
+    \jfam{{{{\Gamma}{A}}{P}}{Q}}{R}
+    }
+  { \jfameq
+      {{{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}}
+      {\jcomp{\subst{x}{P}}{\finc{x}{P}}{R}}
+      {\subst{x}{R}}
+    }
+  \\
+& \inference
+  { \jterm{\Gamma}{A}{x}
+    \jterm{{{{\Gamma}{A}}{P}}{Q}}{R}{h}
+    }
+  { \jtermeq
+      {{{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}}
+      {\subst{x}{R}}
+      {\jcomp{\subst{x}{P}}{\finc{x}{P}}{h}}
+      {\subst{x}{h}}
+    }
+\end{align*}
+\end{lem}
+
+\begin{proof}
+We only prove the validity of the first inference rule. Let $\jterm{\Gamma}{A}{x}$
+be a term and let $\jfam{{{\Gamma}{A}}{P}}{Q}$ be a family. Then we have the
+judgmental equalities
+\begin{align*}
+\jcomp{\subst{x}{P}}{\finc{x}{P}}{Q}
+& \jdeq
+  \unfold{\jcomp{\subst{x}{P}}{\unfold{\finc{x}{P}}}{Q}}
+  \tag{by definition}
+  \\
+& \jdeq
+  \subst
+    {\idtm{\subst{x}{P}}}
+    {{\ctxwk{\subst{x}{P}}{x}}{\ctxwk{\subst{x}{P}}{Q}}}
+  \tag{by \autoref{comp-es}}
+  \\
+& \jdeq
+  \subst
+    {\idtm{\subst{x}{P}}}
+    {\ctxwk{\subst{x}{P}}{\subst{x}{Q}}}
+  \tag{by \autoref{comp-ws-f}}
+  \\
+& \jdeq
+  \subst{x}{Q}.
+  \tag{by \autoref{idfunc-precomp}}
+\end{align*}
+\end{proof}
+
+We can give a second characterization of fiber inclusions:
+
+\begin{lem}\label{lem:finc-char2}
+We have the following inference rule
+\begin{equation*}
+\inference
+  { \jterm{\Gamma}{A}{x}
+    \jfam{{\Gamma}{A}}{P}
+    }
+  { \jhomeq{\Gamma}{\subst{x}{P}}{{A}{P}}{\finc{x}{P}}{\subst{x}{\idtm{{A}{P}}}}
+    } 
+\end{equation*}
+\end{lem}
+
+\begin{proof}
+The proof is the following calculation:
+\begin{align*}
+\finc{x}{P}
+& \jdeq 
+  \tmext{\ctxwk{\subst{x}{P}}{x}}{\idtm{\subst{x}{P}}}
+  \tag{by definition}
+  \\
+& \jdeq
+  \subst
+    { \idtm{\subst{x}{P}}
+      }
+    { { \ctxwk{\subst{x}{P}}{x}
+        }
+      { \idtm{\ctxext{\ctxwk{\subst{x}{P}}{A}}{\ctxwk{\subst{x}{P}}{P}}}
+        }
+      }
+  \tag{by definition}
+  \\
+& \jdeq
+  \subst
+    { \idtm{\subst{x}{P}}
+      }
+    { { \ctxwk{\subst{x}{P}}{x}
+        }
+      { \idtm{\ctxwk{\subst{x}{P}}{\ctxext{A}{P}}}
+        }
+      }
+  \tag{by \autoref{comp-we-f}}
+  \\
+& \jdeq
+  \subst
+    { \idtm{\subst{x}{P}}
+      }
+    { { \ctxwk{\subst{x}{P}}{x}
+        }
+      { \ctxwk{\subst{x}{P}}{\idtm{\ctxext{A}{P}}}
+        }
+      }
+  \tag{by \autoref{comp-wi-t}}
+  \\
+& \jdeq
+  \subst
+    { \idtm{\subst{x}{P}}
+      }
+    { \ctxwk
+        { \subst{x}{P}
+          }
+        { \subst{x}{\idtm{\ctxext{A}{P}}}
+          }
+      }
+  \tag{by \autoref{comp-ws-t}}
+  \\
+& \jdeq
+  \subst{x}{\idtm{\ctxext{A}{P}}}.
+  \tag{by \autoref{idfunc-precomp}}
+\end{align*}
+\end{proof}
+
+\begin{cor}
+The following inference rule is valid:
+\begin{equation*}
+\inference
+  { \jterm{\Gamma}{A}{x}
+    \jterm{\Gamma}{\subst{x}{P}}{u}
+    }
+  { \jtermeq{\Gamma}{{A}{P}}{\subst{u}{\finc{x}{P}}}{\tmext{x}{u}}
+    }
+\end{equation*}
+\end{cor}
+
 We have the following lemmas expressing the compatibility of the fiber
@@
 
+\begin{proof}
+Let $\jterm{\Gamma}{A}{x}$. We have the judgmental equalities
+\begin{align*}
+\finc{x}{\emptyf}
+& \jdeq
+  \unfold{\finc{x}{\emptyf}}
+  \tag{by definition}
+  \\
+& \jdeq
+  \tmext{\ctxwk{\emptyf}{x}}{\idtm{\emptyf}}
+  \tag{by \autoref{comp-s0-c}}
+  \\
+& \jdeq
+  \tmext{x}{\idtm{\emptyf}}
+  \tag{by \autoref{comp-0w-t}}
+  \\
+& \jdeq
+  \tmext{x}{\emptytm}
+  \tag{by \autoref{comp-00-t}}
+  \\
+& \jdeq
+  x
+  \tag{by \autoref{lem:tmext-emp}}
+\end{align*}
+to prove the first judgmental equality. For the second, we have
+\begin{align*}
+\finc{\emptytm}{A}
+& \jdeq
+  \unfold{\finc{\emptytm}{A}}
+  \tag{by definition}
+  \\
+& \jdeq
+  \tmext{\emptytm}{\idtm{\subst{\emptytm}{A}}}
+  \tag{by \autoref{comp-00-t}}
+  \\
+& \jdeq
+  \idtm{\subst{\emptytm}{A}}
+  \tag{by \autoref{lem:tmext-emp}}
+  \\
+& \jdeq
+  \idtm{A}
+  \tag{by \autoref{comp-0s-f}}.
+\end{align*}
+\end{proof}
+
 \begin{lem}
@@
 \end{lem}
+
+\begin{proof}
+Let $\jterm{\Gamma}{A}{x}$, $\jterm{\Gamma}{\subst{x}{P}}{u}$ and
+$\jfam{{{\Gamma}{A}}{P}}{Q}$. Then we have the judgmental equalities
+\begin{align*}
+\finc{\tmext{x}{u}}{Q}
+& \jdeq
+  \subst{\tmext{}{}{x}{u}}{\idtm{{{A}{P}}{Q}}}
+  \tag{by \autoref{lem:finc-char2}}
+  \\
+& \jdeq
+  \subst{u}{{x}{\idtm{{{A}{P}}{Q}}}}
+  \tag{by \autoref{comp-es}}
+  \\
+& \jdeq
+  \subst{u}{{x}{\idtm{{A}{{P}{Q}}}}}
+  \tag{by \autoref{comp-ee-c}}
+  \\
+& \jdeq
+  \subst{u}{\finc{x}{\ctxext{P}{Q}}}
+  \tag{by \autoref{lem:finc-char2}}
+  \\
+& \jdeq
+  \jcomp{}{\finc{u}{\subst{x}{Q}}}{\finc{x}{\ctxext{P}{Q}}}
+  \tag{by \autoref{lem:finc-precomp}}
+\end{align*}
+\end{proof}
 
@@
 
+\begin{comment}
 \subsection{Another special case of projections}
@@
 \end{lem}
+\end{comment}
 

### 2014-05-13 article-models-ttderived.tex
section on fiber inclusions finished!
working on dependent morphisms

@@
   {\jcomp{A}{f}{{B}{g}{R}}}
-  {\unfold{\jcomp{A}{f}{G}}}
+  {\jcomp{A}{f}{G}}
 \end{equation*}
@@
   {\jcomp{A}{f}{{B}{g}{R}}}
-  {\jcomp{P}{F}{\unfold{\jcomp{A}{f}{G}}}}.
+  {\jcomp{P}{F}{\jcomp{A}{f}{G}}}.
 \end{equation*}
@@
 \jcomp{A}{{A}{f}{g}}{R}$, it follows that 
-$\jcomp{P}{F}{\unfold{\jcomp{A}{f}{G}}}$ is a morphism from $P$ to $R$ over
-$\jcomp{A}{f}{g}$. We make the following definition:
+$\jcomp{P}{F}{\jcomp{A}{f}{G}}$ is a morphism from $P$ to $R$ over
+$\jcomp{A}{f}{g}$. This could be considered as the composition of $G$ with $F$.
+In the following definition, we formulate this more generally.
 
 \begin{defn}
-Let $\jhom{\Gamma}{A}{B}{f}$ and $\jhom{\Gamma}{B}{C}{g}$
-be morphisms and let $\jfhom{\Gamma}{A}{B}{f}{P}{Q}{F}$ and
-$\jfhom{\Gamma}{B}{C}{g}{Q}{R}{G}$ be morphisms over them. Then we define\begin{equation*}
-\jfhomdefn
+\emph{Horizontal composition of morphisms over morphisms} is defined by
+\begin{align*}
+& \inference
+  { \jfhom{\Gamma}{A}{B}{f}{P}{Q}{F}
+    \jfam{{{\Gamma}{B}}{Q}}{R}
+    }
+  { \jfamdefn
+      {{{\Gamma}{A}}{P}}
+      {\jfcomp{A}{f}{P}{F}{R}}
+      {\unfold{\jfcomp{A}{f}{P}{F}{R}}}
+    }
+  \\
+& \inference
+  { \jfhom{\Gamma}{A}{B}{f}{P}{Q}{F}
+    \jfam{{{{\Gamma}{B}}{Q}}{R}}{S}
+    }
+  { \jfamdefn
+      {{{{\Gamma}{A}}{P}}{\jfcomp{A}{f}{P}{F}{R}}}
+      {\jfcomp{A}{f}{P}{F}{S}}
+      {\unfold{\jfcomp{A}{f}{P}{F}{S}}}
+    }
+  \\
+& \inference
+  { \jfhom{\Gamma}{A}{B}{f}{P}{Q}{F}
+    \jterm{{{{\Gamma}{B}}{Q}}{R}}{S}{k}
+    }
+  { \jtermdefn
+      {{{{\Gamma}{A}}{P}}{\jfcomp{A}{f}{P}{F}{R}}}
+      {\jfcomp{A}{f}{P}{F}{S}}
+      {\jfcomp{A}{f}{P}{F}{k}}
+      {\unfold{\jfcomp{A}{f}{P}{F}{S}}}
+    }
+\end{align*}
+\end{defn}
+
+There is also a notion of vertical composition, although this is not an
+operation the way horizontal composition is. Nevertheless, to analyze the
+properties of horizontal composition it is useful to
+consider vertical composition.
+
+\begin{defn}
+Let $\jhom{\Gamma}{A}{B}{f}$ be a morphism from $A$ to $B$ in context $\Gamma$
+and let $\jfhom{\Gamma}{A}{B}{f}{P}{Q}{F}$ be a morphism over $f$ in context 
+$\Gamma$. We define
+\begin{equation*}
+\jhomdefn{\Gamma}{{A}{P}}{{B}{Q}}{\jvcomp{P}{f}{F}}{\unfold{\jvcomp{P}{f}{F}}}
+\end{equation*}
+\end{defn}
+
+\begin{lem}
+The following inference rule is valid:
+\begin{equation*}
+\inference
+  { \jfhom{\Gamma}{A}{B}{f}{P}{Q}{F}
+    \jfam{{{\Gamma}{B}}{Q}}{R}
+    }
+  { \jfameq
+      {{{\Gamma}{A}}{P}}
+      {\jfcomp{A}{f}{P}{F}{R}}
+      {\jcomp{{A}{P}}{\jvcomp{P}{f}{F}}{R}}
+    }
+\end{equation*}
+\end{lem}
+
+\begin{proof}
+Let $F$ be a morphism from $P$ to $Q$ over $f$ in context $\Gamma$ and let
+$R$ be a family over $\ctxext{{\Gamma}{B}}{Q}$. Then we have
+\begin{align*}
+\jfcomp{A}{f}{P}{F}{R}
+& \jdeq
+  \unfold{\jfcomp{A}{f}{P}{F}{R}}
+  \tag{by definition}
+  \\
+& \jdeq
+  \subst{F}{{\ctxwk{P}{f}}{\ctxwk{P}{{A}{R}}}}
+  \tag{by \autoref{comp-ws-f}}
+  \\
+& \jdeq
+  \subst{F}{{\ctxwk{P}{f}}{\ctxwk{\ctxext{A}{P}}{R}}}
+  \tag{by \autoref{comp-ew-f}}
+  \\
+& \jdeq
+  \subst{\tmext{\ctxwk{P}{f}}{F}}{\ctxwk{\ctxext{A}{P}}{R}}
+  \tag{by \autoref{comp-es}}
+  \\
+& \jdeq
+  \jcomp{{A}{P}}{\jvcomp{P}{f}{F}}{R}.
+  \tag{by definition}
+\end{align*}
+\end{proof}
+
+\begin{lem}
+We have the following composition threesome:
+\begin{equation*}
+\inference
+  { \jfhom{\Gamma}{A}{B}{f}{P}{Q}{F}
+    \jfhom{\Gamma}{B}{C}{g}{Q}{R}{G}
+    }
+  { \jhomeq
       {\Gamma}
-  {A}
-  {C}
-  {\jcomp{A}{f}{g}}
-  {P}
-  {R}
-  {\jfcomp{A}{f}{P}{F}{G}}
-  {\unfold{\jfcomp{A}{f}{P}{F}{G}}}.
-\end{equation*}
-\end{defn}
-
-This composition is also judgmentally associative.
+      {{A}{P}}
+      {{C}{R}}
+      {\jcomp{{A}{P}}{\jvcomp{P}{f}{F}}{\jvcomp{Q}{g}{G}}}
+      {\jvcomp{P}{\jcomp{A}{f}{g}}{\jfcomp{A}{f}{P}{F}{G}}}
+    }
+\end{equation*}
+\end{lem}
+
+\begin{proof}
+Consider $\jfhom{\Gamma}{A}{B}{f}{P}{Q}{F}$ and 
+$\jfhom{\Gamma}{B}{C}{g}{Q}{R}{G}$. Then we have
+\begin{align*}
+\jcomp{{A}{P}}{\jvcomp{P}{f}{F}}{\jvcomp{Q}{g}{G}}
+& \jdeq
+  \unfold{\jcomp{{A}{P}}{\unfold{\jvcomp{P}{f}{F}}}{\unfold{\jvcomp{Q}{g}{G}}}}
+  \tag{by definition}
+  \\
+& \jdeq
+  \subst
+    {F}
+    { {\ctxwk{P}{f}}
+      {\ctxwk{\ctxext{A}{P}}{\tmext{\ctxwk{Q}{g}}{G}}}
+      }
+  \tag{by \autoref{comp-es}}
+  \\
+& \jdeq
+  \subst
+    {F}
+    { {\ctxwk{P}{f}}
+      {\ctxwk{P}{{A}{\tmext{\ctxwk{Q}{g}}{G}}}}
+      }
+  \tag{by \autoref{comp-ew-t}}
+  \\
+& \jdeq
+  \subst
+    {F}
+    { \ctxwk{P}{\subst{f}{\ctxwk{A}{\tmext{\ctxwk{Q}{g}}{G}}}}
+      }
+  \tag{by \autoref{comp-ws-t}}
+  \\
+& \jdeq
+  \jcomp{P}{F}{\jcomp{A}{f}{\tmext{}{}{\ctxwk{Q}{g}}{G}}}
+  \tag{by definition}
+  \\
+& \jdeq
+  \subst
+    {F}
+    {\ctxwk{P}{\tmext{}{}{\jcomp{A}{f}{\ctxwk{Q}{g}}}{\jcomp{A}{f}{G}}}}
+\end{align*}
+\end{proof}
+
+\begin{lem}
+Horizontal composition of morphisms over morphisms is associative, i.e.~the
+following inference rules are valid:
+\begin{align*}
+& \inference
+  { \jfhom{\Gamma}{A}{B}{f}{P}{Q}{F}
+    \jfhom{\Gamma}{B}{C}{g}{Q}{R}{G}
+    \jfam{{{\Gamma}{C}}{R}}{S}
+    }
+  { \jfameq
+      {{{\Gamma}{A}}{P}}
+      {\jfcomp{A}{f}{P}{F}{\jfcomp{B}{g}{Q}{G}{S}}}
+      {\jfcomp{A}{\jcomp{A}{f}{g}}{P}{\jfcomp{A}{f}{P}{F}{G}}{S}}
+    }
+  \\
+& \inference
+  { \jfhom{\Gamma}{A}{B}{f}{P}{Q}{F}
+    \jfhom{\Gamma}{B}{C}{g}{Q}{R}{G}
+    \jfam{{{{\Gamma}{C}}{R}}{S}}{T}
+    }
+  { \jfameq
+      {{{{\Gamma}{A}}{P}}{\jfcomp{A}{f}{P}{F}{\jfcomp{B}{g}{Q}{G}{S}}}}
+      {\jfcomp{A}{f}{P}{F}{\jfcomp{B}{g}{Q}{G}{T}}}
+      {\jfcomp{A}{\jcomp{A}{f}{g}}{P}{\jfcomp{A}{f}{P}{F}{G}}{T}}
+    }
+  \\
+& \inference
+  { \jfhom{\Gamma}{A}{B}{f}{P}{Q}{F}
+    \jfhom{\Gamma}{B}{C}{g}{Q}{R}{G}
+    \jterm{{{{\Gamma}{C}}{R}}{S}}{T}{l}
+    }
+  { \jtermeq
+      {{{{\Gamma}{A}}{P}}{\jfcomp{A}{f}{P}{F}{\jfcomp{B}{g}{Q}{G}{S}}}}
+      {\jfcomp{A}{f}{P}{F}{\jfcomp{B}{g}{Q}{G}{T}}}
+      {\jfcomp{A}{f}{P}{F}{\jfcomp{B}{g}{Q}{G}{l}}}
+      {\jfcomp{A}{\jcomp{A}{f}{g}}{P}{\jfcomp{A}{f}{P}{F}{G}}{l}}
+    }
+\end{align*}
+\end{lem}
+
+\begin{proof}
+We only prove the first inference rule. Let
+$\jfhom{\Gamma}{A}{B}{f}{P}{Q}{F}$, $\jfhom{\Gamma}{B}{C}{g}{Q}{R}{G}$ and
+$\jfam{{{\Gamma}{C}}{R}}{S}$. Then we have the judgmental equalities
+\begin{align*}
+\jfcomp{A}{f}{P}{F}{\jfcomp{B}{g}{Q}{G}{S}}
+& \jdeq
+  \jcomp{P}{F}{\jcomp{A}{f}{\jcomp{Q}{G}{\jcomp{B}{g}{S}}}}
+  \tag{by definition}
+  \\
+& \jdeq 
+  \unfoldall{\jcomp{P}{F}{\jcomp{A}{f}{\jcomp{Q}{G}{\jcomp{B}{g}{S}}}}}
+  \tag{by definition}
+  \\
+& \jdeq 
+\end{align*}
+\end{proof}
 
@@
 \end{proof}
-
-\begin{defn}
-Let $\jhom{\Gamma}{A}{B}{f}$ be a morphism from $A$ to $B$ in context $\Gamma$
-and let $\jfhom{\Gamma}{A}{B}{f}{P}{Q}{F}$ be a morphism over $f$ in context 
-$\Gamma$. We define
-\begin{equation*}
-\jhomdefn{\Gamma}{{A}{P}}{{B}{Q}}{\jvcomp{P}{f}{F}}{\unfold{\jvcomp{P}{f}{F}}}
-\end{equation*}
-\end{defn}
 
@@
 
+\begin{proof}
+Let $\jterm{\Gamma}{B}{y}$ and $\jfam{{\Gamma}{B}}{Q}$. Then we have the
+judgmental equalities
+\begin{align*}
+\finc{\ctxwk{A}{y}}{\ctxwk{A}{Q}}
+& \jdeq
+  \subst{\ctxwk{A}{y}}{\idtm{\ctxext{\ctxwk{A}{B}}{\ctxwk{A}{Q}}}}
+  \tag{by \autoref{lem:finc-char2}}
+  \\
+& \jdeq
+  \subst{\ctxwk{A}{y}}{\idtm{\ctxwk{A}{\ctxext{B}{Q}}}}
+  \tag{by \autoref{comp-we-f}}
+  \\
+& \jdeq
+  \subst{\ctxwk{A}{y}}{\ctxwk{A}{\idtm{\ctxext{B}{Q}}}}
+  \tag{by \autoref{comp-wi-t}}
+  \\
+& \jdeq
+  \ctxwk{A}{\subst{y}{\idtm{{B}{Q}}}}
+  \tag{by \autoref{comp-ws-t}}
+  \\
+& \jdeq
+  \ctxwk{A}{\finc{y}{Q}}.
+  \tag{by \autoref{lem:finc-char2}}
+\end{align*}
+\end{proof}
+
 \begin{lem}
@@
 
+\begin{proof}
+Let $\jterm{\Gamma}{A}{x}$, $\jterm{{\Gamma}{A}}{P}{f}$ and 
+$\jfam{{{\Gamma}{A}}{P}}{Q}$. Then we have the judgmental equalities
+\begin{align*}
+\finc{\subst{x}{f}}{\subst{x}{Q}}
+& \jdeq
+  \subst{x}{{f}{\idtm{{P}{Q}}}}
+  \tag{by \autoref{lem:finc-char2}}
+  \\
+& \jdeq
+  \subst{{x}{f}}{{x}{\idtm{{P}{Q}}}}
+  \tag{by \autoref{comp-ss-t}}
+  \\
+& \jdeq
+  \subst{{x}{f}}{\idtm{\subst{x}{\ctxext{P}{Q}}}}
+  \tag{by \autoref{comp-si-t}}
+  \\
+& \jdeq
+  \subst{{x}{f}}{\idtm{\ctxext{\subst{x}{P}}{\subst{x}{Q}}}}
+  \tag{by \autoref{comp-se-f}}
+  \\
+& \jdeq
+  \finc{\subst{x}{f}}{\subst{x}{Q}}.
+  \tag{by \autoref{lem:finc-char2}}
+\end{align*}
+\end{proof}
+
 \begin{lem}
@@
 \end{lem}
+
+\begin{proof}
+Let $\jfam{\Gamma}{A}$. Then we have the judgmental equalities
+\begin{align*}
+\finc{\idtm{A}}{\ctxwk{A}{P}}
+& \jdeq
+  \subst{\idtm{A}}{\idtm{\ctxext{\ctxwk{A}{A}}{\ctxwk{A}{P}}}}
+  \tag{by \autoref{lem:finc-char2}}
+  \\
+& \jdeq
+  \subst{\idtm{A}}{\idtm{\ctxwk{A}{\ctxext{A}{P}}}}
+  \tag{by \autoref{comp-we-f}}
+  \\
+& \jdeq
+  \subst{\idtm{A}}{\ctxwk{A}{\idtm{{A}{P}}}}
+  \tag{by \autoref{comp-wi-t}}
+  \\
+& \jdeq
+  \idtm{{A}{P}}
+  \tag{by \autoref{idfunc-precomp}}
+\end{align*}
+\end{proof}
 

### 2014-05-14 article-models-ttderived.tex
finished the associativity of horizontal composition

@@
 
-\begin{lem}
+\begin{lem}\label{lem:jcomp-emp}
 Let $\jhom{\Gamma}{A}{B}{f}$. Then the inference rules
@@
 
-\begin{lem}
+\begin{lem}\label{lem:jcomp-ext}
 We have the following inference rule expressing that composition with $f$ is
@@
 
-\begin{lem}
+\begin{lem}\label{lem:jcomp-wk}
 We have the following inference rules expressing that composition with $f$ is
@@
 
-\begin{lem}
+\begin{lem}\label{lem:jcomp-subst}
 We have the following inference rules expressing that composition with $f$ is
@@
 
-\begin{lem}
+\begin{lem}\label{lem:jcomp-idtm}
 We have the following inference rule expressing that composition with $f$ is
@@
 
-\begin{lem}
+\begin{lem}\label{lem:jcomp-jcomp}
 We have the following inference rules about the situation where something is
@@
 
-\begin{lem}
-We have the following inference rules about the compatibility of composition with
-weakening:\begin{align*}
+We established already in \autoref{lem:jcomp-wk} that composition is compatible
+with weakening. However, in that lemma we did not consider the possibility
+of weakening a composition and neither did we consider the possibility of
+composition by or with a constant morphism. We do this in the following lemma.
+
+\begin{lem}\label{lem:jcomp-const}
+The inference
+\begin{align*}
 & \inference
@@
     }
-  \\
+\end{align*}
+tells what happens when we weaken a composition, which is a term of 
+$\ctxwk{A}{C}$, by a family $P$ over $\ctxext{\Gamma}{P}$. We have the following
+inference rules expressing that compositions by or with constant morphisms are
+again constant morphisms.
+\begin{align*}
 & \inference
@@
 
-\begin{lem}
-We have the following inference rules about the compatibility of composition with
-substitution:\begin{align*}
+Likewise, we can substitute a term $\jterm{\Gamma}{A}{x}$ in a composition
+$\jhom{\Gamma}{A}{C}{\jcomp{A}{f}{g}}$. What we get is the \emph{value}
+$\subst{{x}{f}}{g}$. There is a second case where we can substitute: when
+we consider morphisms $\jhom{{\Gamma}{A}}{P}{Q}{f}$ and 
+$\jhom{{\Gamma}{A}}{Q}{R}{g}$ we can restrict them to the fibers. We do this
+in the following lemma:
+
+\begin{lem}\label{lem:jcomp-fiber}
+The following inference rules are valid:
+\begin{align*}
 & \inference
@@
 \begin{lem}\label{lem:tmext-wk}\label{comp-we-t}
-When we weaken a term $\tmext{B}{Q}{y}{v}$ of $\ctxext{B}{Q}$ in context $\Gamma$ by
-a family $A$, the term that we get is $\tmext{\ctxwk{A}{B}}{\ctxwk{A}{Q}}{\ctxwk{A}{y}}{\ctxwk{A}{v}}$. More
+When we weaken a term $\tmext{B}{Q}{y}{v}$ of $\ctxext{B}{Q}$ in context 
+$\Gamma$ by a family $A$, the term that we get is 
+$\tmext{\ctxwk{A}{B}}{\ctxwk{A}{Q}}{\ctxwk{A}{y}}{\ctxwk{A}{v}}$. More
 precisely, the following inference rules are valid:
@@
 \subsection{More on morphisms}
-
-There is also a notion of morphism \emph{over} a morphism. We will develop this
+For the following lemma, recall that the judgment $\jhom{\Gamma}{A}{{B}{Q}}{f}$
+unfolds as
+\begin{equation*}
+\unfold{\jhom{\Gamma}{A}{{B}{Q}}{f}}
+\end{equation*}
+and that we have the judgmental equality 
+$ \jfameq
+    {{\Gamma}{A}}
+    {\ctxwk{A}{\ctxext{B}{Q}}}
+    {\ctxext{\ctxwk{A}{B}}{\ctxwk{A}{Q}}}.
+  $
+Therefore, each morphism into an extended family can itself be described as
+an extended term. The following lemma explains how this goes.
+
+\begin{lem}
+Let $\jhom{\Gamma}{A}{{B}{Q}}{f}$ be a morphism from $A$ to $\ctxext{B}{Q}$
+in a context $\Gamma$. Then we have
+\begin{equation*}
+\jhomeq
+  {\Gamma}
+  {A}
+  {{B}{Q}}
+  {f}
+  {\tmext{\jcomp{}{f}{\cprojfstf{B}{Q}}}{\jcomp{}{f}{\cprojsndf{B}{Q}}}}.
+\end{equation*}
+Alternatively, when $\jhom{\Gamma}{A}{B}{f_0}$ and 
+$\jterm{{\Gamma}{A}}{\jcomp{}{f_0}{Q}}{f_1}$ we obtain a morphism
+$\jhom{\Gamma}{A}{{B}{Q}}{\tmext{f_0}{f_1}}$ with the property that
+\begin{align*}
+\jhomeq*{\Gamma}{A}{B}{\jcomp{}{\tmext{f_0}{f_1}}{\cprojfstf{B}{Q}}}{f_0}\\
+\jtermeq*{{\Gamma}{A}}{\jcomp{}{f_0}{Q}}{\jcomp{}{\tmext{f_0}{f_1}}{\cprojsndf{B}{Q}}}{f_1}.
+\end{align*}
+\end{lem}
+
+\begin{proof}
+Let $\jhom{\Gamma}{A}{{B}{Q}}{f}$ be a morphism in context $\Gamma$. Then we
+have the judgmental equalities
+\begin{align*}
+\cprojfst{\ctxwk{A}{B}}{\ctxwk{A}{Q}}{f}
+& \jdeq
+  \subst{f}{\ctxwk{A}{\cprojfstf{B}{Q}}}
+  \\
+& \jdeq
+  \jcomp{}{f}{\cprojfstf{B}{Q}}
+  \\
+\cprojsnd{\ctxwk{A}{B}}{\ctxwk{A}{Q}}{f}
+& \jdeq
+  \subst{f}{\ctxwk{A}{\cprojsndf{B}{Q}}}
+  \\
+& \jdeq
+  \jcomp{}{f}{\cprojsndf{B}{Q}}
+\end{align*}
+The alternative formulation of the statement is a direct corollary.
+\end{proof}
+
+We also have the following lemma about the compatibility of pairing and composition:
+
+\begin{lem}\label{lem:tmext-jcomp}
+The following inference rule is valid
+\begin{align*}
+& \inference
+  { \jhom{\Gamma}{A}{B}{f}
+    \jhom{\Gamma}{B}{C}{g}
+    \jfam{{\Gamma}{C}}{R}
+    \jterm{{\Gamma}{B}}{\subst{g}{\ctxwk{B}{R}}}{w}
+    }
+  { \jhomeq
+      {\Gamma}
+      {A}
+      {{C}{R}}
+      {\jcomp{A}{f}{\tmext{\ctxwk{B}{C}}{\ctxwk{B}{R}}{g}{w}}}
+      {\tmext{\ctxwk{A}{C}}{\ctxwk{A}{R}}{\jcomp{A}{f}{g}}{\jcomp{A}{f}{w}}}
+    }
+\end{align*}
+\end{lem}
+
+\begin{proof}
+Let $\jhom{\Gamma}{A}{B}{f}$, $\jhom{\Gamma}{B}{C}{g}$, $\jfam{{\Gamma}{C}}{R}$
+and $\jterm{{\Gamma}{B}}{\subst{g}{\ctxwk{B}{R}}}{w}$. Then we have the
+judgmental equalities
+\begin{align*}
+\jcomp{A}{f}{\tmext{\ctxwk{B}{C}}{\ctxwk{B}{R}}{g}{w}}
+& \jdeq 
+  \subst{f}{\ctxwk{A}{\tmext{\ctxwk{B}{C}}{\ctxwk{B}{R}}{g}{w}}}
+  \\
+& \jdeq 
+  \subst
+    {f}
+    {\tmext{\ctxwk{A}{{B}{C}}}{\ctxwk{A}{{B}{R}}}{\ctxwk{A}{g}}{\ctxwk{A}{w}}}
+  \\
+& \jdeq 
+  \tmext
+    {\ctxwk{A}{C}}
+    {\ctxwk{A}{R}}
+    {\subst{f}{\ctxwk{A}{g}}}
+    {\subst{f}{\ctxwk{A}{w}}}
+  \\
+& \jdeq 
+  \tmext{\ctxwk{A}{C}}{\ctxwk{A}{R}}{\jcomp{A}{f}{g}}{\jcomp{A}{f}{w}}.
+  \qedhere
+\end{align*}
+\end{proof}
+
+There is a notion of morphism \emph{over} a morphism. We will develop this
 notion because it will be needed in the theory of models later on.
@@
 \begin{rmk}
-The judgment $\jfhom{\Gamma}{A}{B}{f}{P}{Q}{F}$ means the same thing as\begin{equation*}
+The judgment $\jfhom{\Gamma}{A}{B}{f}{P}{Q}{F}$ means the same thing as
+\begin{equation*}
 \jhom{{\Gamma}{A}}{P}{\jcomp{A}{f}{Q}}{F}.
 \end{equation*}
+Thus we see that a morphism from $P$ to $Q$ over the identity term $\idtm{A}$ in
+context $\Gamma$ is the same thing as a morphism from $P$ to $Q$ in context
+$\ctxext{\Gamma}{A}$, i.e.~the following inference rules are valid:
+\begin{align*}
+& \inference
+  { \jfam{{\Gamma}{A}}{P}
+    \jfam{{\Gamma}{A}}{Q}
+    \jfhom{\Gamma}{A}{A}{\idtm{A}}{P}{Q}{f}
+    }
+  { \jhom{{\Gamma}{A}}{P}{Q}{f}
+    }
+  \\
+& \inference
+  { \jfam{{\Gamma}{A}}{P}
+    \jfam{{\Gamma}{A}}{Q}
+    \jhom{{\Gamma}{A}}{P}{Q}{f}
+    }
+  { \jfhom{\Gamma}{A}{A}{\idtm{A}}{P}{Q}{f}
+    }
+\end{align*}
+To see this, we only have to note that
+$\ctxwk{P}{\subst{\idtm{A}}{\ctxwk{A}{Q}}}\jdeq\ctxwk{P}{Q}$, which
+holds by \autoref{idfunc-wk-defn}.
 \end{rmk}
@@
 
-\begin{lem}
-The following inference rule is valid:
-\begin{equation*}
-\inference
+\begin{lem}\label{lem:composition-threesome}
+We have the following composition threesome:
+\begin{align*}
+& \inference
   { \jfhom{\Gamma}{A}{B}{f}{P}{Q}{F}
@@
     }
-\end{equation*}
+  \\
+& \inference
+  { \jfhom{\Gamma}{A}{B}{f}{P}{Q}{F}
+    \jfam{{{{\Gamma}{B}}{Q}}{R}}{S}
+    }
+  { \jfameq
+      {{{{\Gamma}{A}}{P}}{\jfcomp{A}{f}{P}{F}{R}}}
+      {\jfcomp{A}{f}{P}{F}{S}}
+      {\jcomp{{A}{P}}{\jvcomp{P}{f}{F}}{S}}
+    }
+  \\
+& \inference
+  { \jfhom{\Gamma}{A}{B}{f}{P}{Q}{F}
+    \jterm{{{{\Gamma}{B}}{Q}}{R}}{S}{k}
+    }
+  { \jtermeq
+      {{{{\Gamma}{A}}{P}}{\jfcomp{A}{f}{P}{F}{R}}}
+      {\jfcomp{A}{f}{P}{F}{S}}
+      {\jfcomp{A}{f}{P}{F}{k}}
+      {\jcomp{{A}{P}}{\jvcomp{P}{f}{F}}{k}}
+    }
+\end{align*}
 \end{lem}
@@
 
-\begin{lem}
-We have the following composition threesome:
+\begin{lem}\label{lem:composition-interchange}
+We have the following interchange law for composition:
 \begin{equation*}
@@
 & \jdeq
-  \unfold{\jcomp{{A}{P}}{\unfold{\jvcomp{P}{f}{F}}}{\unfold{\jvcomp{Q}{g}{G}}}}
-  \tag{by definition}
-  \\
-& \jdeq
-  \subst
-    {F}
-    { {\ctxwk{P}{f}}
-      {\ctxwk{\ctxext{A}{P}}{\tmext{\ctxwk{Q}{g}}{G}}}
-      }
-  \tag{by \autoref{comp-es}}
-  \\
-& \jdeq
-  \subst
-    {F}
-    { {\ctxwk{P}{f}}
-      {\ctxwk{P}{{A}{\tmext{\ctxwk{Q}{g}}{G}}}}
-      }
-  \tag{by \autoref{comp-ew-t}}
-  \\
-& \jdeq
-  \subst
-    {F}
-    { \ctxwk{P}{\subst{f}{\ctxwk{A}{\tmext{\ctxwk{Q}{g}}{G}}}}
-      }
-  \tag{by \autoref{comp-ws-t}}
-  \\
-& \jdeq
   \jcomp{P}{F}{\jcomp{A}{f}{\tmext{}{}{\ctxwk{Q}{g}}{G}}}
-  \tag{by definition}
-  \\
-& \jdeq
-  \subst
-    {F}
-    {\ctxwk{P}{\tmext{}{}{\jcomp{A}{f}{\ctxwk{Q}{g}}}{\jcomp{A}{f}{G}}}}
+  \tag{by \autoref{lem:composition-threesome}}
+  \\
+& \jdeq
+  \tmext{}{}
+    {\jcomp{P}{F}{\jcomp{A}{f}{\ctxwk{Q}{g}}}}
+    {\jcomp{P}{F}{\jcomp{A}{f}{G}}}
+  \tag{by \autoref{lem:tmext-jcomp}}
+  \\
+& \jdeq
+  \tmext{}{}
+    {\jcomp{P}{F}{\jcomp{A}{f}{\ctxwk{Q}{g}}}}
+    {\jfcomp{A}{f}{P}{F}{G}}
+  \tag{by definition}
+  \\
+& \jdeq
+  \tmext{}{}
+    {\jcomp{P}{F}{\ctxwk{\jcomp{A}{f}{Q}}{\jcomp{A}{f}{g}}}}
+    {\jfcomp{A}{f}{P}{F}{G}}
+  \tag{by \autoref{lem:jcomp-wk}}
+  \\
+& \jdeq
+  \tmext{}{}
+    {\subst{F}{\ctxwk{P}{\ctxwk{\jcomp{A}{f}{Q}}{\jcomp{A}{f}{g}}}}}
+    {\jfcomp{A}{f}{P}{F}{G}}
+  \tag{by definition}
+  \\
+& \jdeq
+  \tmext{}{}
+    {\subst{F}{\ctxwk{{P}{\jcomp{A}{f}{Q}}}{{P}{\jcomp{A}{f}{g}}}}}
+    {\jfcomp{A}{f}{P}{F}{G}}
+  \tag{by \autoref{comp-ww-t}}
+  \\
+& \jdeq
+  \tmext{}{}
+    {\ctxwk{P}{\jcomp{A}{f}{g}}}
+    {\jfcomp{A}{f}{P}{F}{G}}
+  \tag{by \autoref{cancellation-ws-t}}
+  \\
+& \jdeq
+  \jvcomp{P}{\jcomp{A}{f}{g}}{\jfcomp{A}{f}{P}{F}{G}}
+  \tag{by definition}
 \end{align*}
@@
 & \jdeq
-  \jcomp{P}{F}{\jcomp{A}{f}{\jcomp{Q}{G}{\jcomp{B}{g}{S}}}}
-  \tag{by definition}
-  \\
-& \jdeq 
-  \unfoldall{\jcomp{P}{F}{\jcomp{A}{f}{\jcomp{Q}{G}{\jcomp{B}{g}{S}}}}}
-  \tag{by definition}
-  \\
-& \jdeq 
-\end{align*}
-\end{proof}
-
-Now let $\jfam{{\Gamma}{A}}{P}$ and $\jfam{{\Gamma}{A}}{Q}$ be families. A
-morphism from $P$ to $Q$ over the identity term $\idtm{A}$ in context
-$\Gamma$ is the same thing as a morphism from $P$ to $Q$ in context
-$\ctxext{\Gamma}{A}$:
-
-\begin{lem}\label{hom-over-id-is-hom}
-We have the following valid inference rules:\begin{align*}
-& \inference
-  { \jfam{{\Gamma}{A}}{P}
-    \jfam{{\Gamma}{A}}{Q}
-    \jfhom{\Gamma}{A}{A}{\idtm{A}}{P}{Q}{f}
-    }
-  { \jhom{{\Gamma}{A}}{P}{Q}{f}
-    }
-  \\
-& \inference
-  { \jfam{{\Gamma}{A}}{P}
-    \jfam{{\Gamma}{A}}{Q}
-    \jhom{{\Gamma}{A}}{P}{Q}{f}
-    }
-  { \jfhom{\Gamma}{A}{A}{\idtm{A}}{P}{Q}{f}
-    }
-\end{align*}
-\end{lem}
-
-\begin{proof}
-If we unfold the judgments $\jhom{{\Gamma}{A}}{P}{Q}{f}$ and
-$\jfhom{\Gamma}{A}{A}{\idtm{A}}{P}{Q}{f}$, we get the judgments\begin{align*}
-& \unfold{\jhom{{\Gamma}{A}}{P}{Q}{f}}
-  \\
-& \unfold{\jfhom{\Gamma}{A}{A}{\idtm{A}}{P}{Q}{f}}
-\end{align*}
-respectively. Therefore, we only need to verify that
-$\ctxwk{P}{\subst{\idtm{A}}{\ctxwk{A}{Q}}}\jdeq\ctxwk{P}{Q}$, which is indeed
-the case by \autoref{idfunc-wk-defn}.
-\end{proof}
-
-For the following lemma, recall that the judgment $\jhom{\Gamma}{A}{{B}{Q}}{f}$
-unfolds as
-\begin{equation*}
-\unfold{\jhom{\Gamma}{A}{{B}{Q}}{f}}
-\end{equation*}
-and that we have the judgmental equality 
-$ \jfameq
-    {{\Gamma}{A}}
-    {\ctxwk{A}{\ctxext{B}{Q}}}
-    {\ctxext{\ctxwk{A}{B}}{\ctxwk{A}{Q}}}.
-  $
-Therefore, each morphism into an extended family can itself be described as
-an extended term. The following lemma explains how this goes.
-
-\begin{lem}
-Let $\jhom{\Gamma}{A}{{B}{Q}}{f}$ be a morphism from $A$ to $\ctxext{B}{Q}$
-in a context $\Gamma$. Then we have
-\begin{equation*}
-\jhomeq
-  {\Gamma}
-  {A}
-  {{B}{Q}}
-  {f}
-  {\tmext{\jcomp{}{f}{\cprojfstf{B}{Q}}}{\jcomp{}{f}{\cprojsndf{B}{Q}}}}.
-\end{equation*}
-Alternatively, when $\jhom{\Gamma}{A}{B}{f_0}$ and 
-$\jterm{{\Gamma}{A}}{\jcomp{}{f_0}{Q}}{f_1}$ we obtain a morphism
-$\jhom{\Gamma}{A}{{B}{Q}}{\tmext{f_0}{f_1}}$ with the property that
-\begin{align*}
-\jhomeq*{\Gamma}{A}{B}{\jcomp{}{\tmext{f_0}{f_1}}{\cprojfstf{B}{Q}}}{f_0}\\
-\jtermeq*{{\Gamma}{A}}{\jcomp{}{f_0}{Q}}{\jcomp{}{\tmext{f_0}{f_1}}{\cprojsndf{B}{Q}}}{f_1}.
-\end{align*}
-\end{lem}
-
-\begin{proof}
-Let $\jhom{\Gamma}{A}{{B}{Q}}{f}$ be a morphism in context $\Gamma$. Then we
-have the judgmental equalities
-\begin{align*}
-\cprojfst{\ctxwk{A}{B}}{\ctxwk{A}{Q}}{f}
-& \jdeq
-  \subst{f}{\ctxwk{A}{\cprojfstf{B}{Q}}}
-  \\
-& \jdeq
-  \jcomp{}{f}{\cprojfstf{B}{Q}}
-  \\
-\cprojsnd{\ctxwk{A}{B}}{\ctxwk{A}{Q}}{f}
-& \jdeq
-  \subst{f}{\ctxwk{A}{\cprojsndf{B}{Q}}}
-  \\
-& \jdeq
-  \jcomp{}{f}{\cprojsndf{B}{Q}}
-\end{align*}
-The alternative formulation of the statement is a direct corollary.
-\end{proof}
-
-We also have the following lemma about the compatibility of pairing and composition:
-
-\begin{lem}
-The following inference rule is valid
-\begin{align*}
-& \inference
-  { \jhom{\Gamma}{A}{B}{f}
-    \jhom{\Gamma}{B}{C}{g}
-    \jfam{{\Gamma}{C}}{R}
-    \jterm{{\Gamma}{B}}{\subst{g}{\ctxwk{B}{R}}}{w}
-    }
-  { \jhomeq
-      {\Gamma}
-      {A}
-      {{C}{R}}
-      {\jcomp{A}{f}{\tmext{\ctxwk{B}{C}}{\ctxwk{B}{R}}{g}{w}}}
-      {\tmext{\ctxwk{A}{C}}{\ctxwk{A}{R}}{\jcomp{A}{f}{g}}{\jcomp{A}{f}{w}}}
-    }
-\end{align*}
-\end{lem}
-
-\begin{proof}
-Let $\jhom{\Gamma}{A}{B}{f}$, $\jhom{\Gamma}{B}{C}{g}$, $\jfam{{\Gamma}{C}}{R}$
-and $\jterm{{\Gamma}{B}}{\subst{g}{\ctxwk{B}{R}}}{w}$. Then we have the
-judgmental equalities
-\begin{align*}
-\jcomp{A}{f}{\tmext{\ctxwk{B}{C}}{\ctxwk{B}{R}}{g}{w}}
-& \jdeq 
-  \subst{f}{\ctxwk{A}{\tmext{\ctxwk{B}{C}}{\ctxwk{B}{R}}{g}{w}}}
-  \\
-& \jdeq 
-  \subst
-    {f}
-    {\tmext{\ctxwk{A}{{B}{C}}}{\ctxwk{A}{{B}{R}}}{\ctxwk{A}{g}}{\ctxwk{A}{w}}}
-  \\
-& \jdeq 
-  \tmext
-    {\ctxwk{A}{C}}
-    {\ctxwk{A}{R}}
-    {\subst{f}{\ctxwk{A}{g}}}
-    {\subst{f}{\ctxwk{A}{w}}}
-  \\
-& \jdeq 
-  \tmext{\ctxwk{A}{C}}{\ctxwk{A}{R}}{\jcomp{A}{f}{g}}{\jcomp{A}{f}{w}}.
-  \qedhere
+  \jcomp{{A}{P}}{\jvcomp{P}{f}{F}}{\jcomp{{B}{Q}}{\jvcomp{Q}{g}{G}}{S}}
+  \tag{by \autoref{lem:composition-threesome}}
+  \\
+& \jdeq
+  \jcomp{{A}{P}}{\jcomp{{A}{P}}{\jvcomp{P}{f}{F}}{\jvcomp{Q}{g}{G}}}{S}
+  \tag{by \autoref{lem:jcomp-jcomp}}
+  \\
+& \jdeq
+  \jcomp{{A}{P}}{\jvcomp{P}{\jcomp{A}{f}{g}}{\jfcomp{A}{f}{P}{F}{G}}}{S}
+  \tag{by \autoref{lem:composition-interchange}}
+  \\
+& \jdeq
+  \jfcomp{A}{\jcomp{A}{f}{g}}{P}{\jfcomp{A}{f}{P}{F}{G}}{S}.
+  \tag{by \autoref{lem:composition-threesome}}
 \end{align*}

### 2014-05-14 article-models-ttderived.tex
various small changes

@@
 
-\subsection{More on morphisms}
+\subsection{Family morphisms}
+In this section we are going to develop dependent morphisms and analyze various
+sorts of composition. The notion of family morphism is the notion of `a 
+morphism over a context morphism'. When we have established all the associativity
+and interchange laws for all the compositions that come with the theory of
+family morphisms, we will also be able to consider \emph{family diagrams}. Those
+are graphical displays of situations in the theory of contexts, families and
+terms, but unlike diagrams in category theory they can contain families (which
+are strictly speaking not morphisms) and terms (which can be seen as morphisms,
+but not in the way we're about to display them).
+
 For the following lemma, recall that the judgment $\jhom{\Gamma}{A}{{B}{Q}}{f}$
@@
 
+Since horizontal composition is just consecutive composition, we will get the
+compatibility with the empty context, extension, weakening, substitution and
+the identity terms for free. We will state these compatibility properities in
+\autoref{lem:jfcomp-emp,lem:jfcomp-ext,lem:jfcomp-wk,lem:jfcomp-subst,lem:jfcomp-idtm},
+but the proofs are left to the reader.
+
+\begin{lem}\label{lem:jfcomp-emp}
+Horizontal composition is compatible with the empty context:
+\begin{align*}
+& \inference
+  { \jfhom{\Gamma}{A}{B}{f}{P}{Q}{F}
+    }
+  { \jfameq
+      {{{\Gamma}{A}}{P}}
+      {\jfcomp{A}{f}{P}{F}{\emptyf}}
+      {\emptyf}
+    }
+  \\
+& \inference
+  { \jfhom{\Gamma}{A}{B}{f}{P}{Q}{F}
+    \jfam{{{\Gamma}{B}}{Q}}{R}
+    }
+  { \jfameq
+      {{{{\Gamma}{A}}{P}}{\jfcomp{A}{f}{P}{F}{R}}}
+      {\jfcomp{A}{f}{P}{F}{\emptyf}}
+      {\emptyf}
+    }
+\end{align*}
+\end{lem}
+
+\begin{lem}\label{lem:jfcomp-ext}
+Horizontal composition is compatible with extension:
+\begin{align*}
+& \inference
+  { \jfhom{\Gamma}{A}{B}{f}{P}{Q}{F}
+    \jfam{{{{\Gamma}{B}}{Q}}{R}}{S}
+    }
+  { \jfameq
+      {{{\Gamma}{A}}{P}}
+      {\jfcomp{A}{f}{P}{F}{\ctxext{R}{S}}}
+      {\ctxext{\jfcomp{A}{f}{P}{F}{R}}{\jfcomp{A}{f}{P}{F}{S}}}
+    }
+  \\
+& \inference
+  { \jfhom{\Gamma}{A}{B}{f}{P}{Q}{F}
+    \jfam{{{{{\Gamma}{B}}{Q}}{R}}{S}}{T}
+    }
+  { \jfameq
+      {{{{\Gamma}{A}}{P}}{\jfcomp{A}{f}{P}{F}{R}}}
+      {\jfcomp{A}{f}{P}{F}{\ctxext{S}{T}}}
+      {\ctxext{\jfcomp{A}{f}{P}{F}{S}}{\jfcomp{A}{f}{P}{F}{T}}}
+    }
+  \\
+& \inference
+  { \jfhom{\Gamma}{A}{B}{f}{P}{Q}{F}
+    \jterm{{{{\Gamma}{B}}{Q}}{R}}{{S}{T}}{\tmext{k}{l}}
+    }
+  { \jtermdefn
+      {{{{\Gamma}{A}}{P}}{\jfcomp{A}{f}{P}{F}{R}}}
+      {\ctxext{\jfcomp{A}{f}{P}{F}{S}}{\jfcomp{A}{f}{P}{F}{T}}}
+      {\jfcomp{A}{f}{P}{F}{\tmext{k}{l}}}
+      {\tmext{\jfcomp{A}{f}{P}{F}{k}}{\jfcomp{A}{f}{P}{F}{l}}}
+    }
+\end{align*}
+\end{lem}
+
+\begin{lem}\label{lem:jfcomp-wk}
+Horizontal composition is compatible with weakening:
+\begin{align*}
+& \inference
+  { \jfhom{\Gamma}{A}{B}{f}{P}{Q}{F}
+    \jfam{{{{\Gamma}{B}}{Q}}{R}}{S}
+    \jfam{{{{\Gamma}{B}}{Q}}{R}}{T}
+    }
+  { \jfameq
+      {{{{{\Gamma}{A}}{P}}{\jfcomp{A}{f}{P}{F}{R}}}{\jfcomp{A}{f}{P}{F}{S}}}
+      {\jfcomp{A}{f}{P}{F}{\ctxwk{S}{T}}}
+      {\ctxwk{\jfcomp{A}{f}{P}{F}{S}}{\jfcomp{A}{f}{P}{F}{T}}}
+    }
+  \\
+& \inference
+  { \jfhom{\Gamma}{A}{B}{f}{P}{Q}{F}
+    \jfam{{{{\Gamma}{B}}{Q}}{R}}{S}
+    \jfam{{{{{\Gamma}{B}}{Q}}{R}}{T}}{U}
+    }
+  { \jfameq
+      {{{{{{\Gamma}{A}}{P}}
+        {\jfcomp{A}{f}{P}{F}{R}}}
+        {\jfcomp{A}{f}{P}{F}{S}}}
+        {\ctxwk{\jfcomp{A}{f}{P}{F}{S}}{\jfcomp{A}{f}{P}{F}{T}}}}
+      {\jfcomp{A}{f}{P}{F}{\ctxwk{S}{U}}}
+      {\ctxwk{\jfcomp{A}{f}{P}{F}{S}}{\jfcomp{A}{f}{P}{F}{U}}}
+    }
+  \\
+& \inference
+  { \jfhom{\Gamma}{A}{B}{f}{P}{Q}{F}
+    \jfam{{{{\Gamma}{B}}{Q}}{R}}{S}
+    \jterm{{{{{\Gamma}{B}}{Q}}{R}}{T}}{U}{m}
+    }
+  { \jtermeq
+      {{{{{{\Gamma}{A}}{P}}
+        {\jfcomp{A}{f}{P}{F}{R}}}
+        {\jfcomp{A}{f}{P}{F}{S}}}
+        {\ctxwk{\jfcomp{A}{f}{P}{F}{S}}{\jfcomp{A}{f}{P}{F}{T}}}}
+      {\jfcomp{A}{f}{P}{F}{\ctxwk{S}{U}}}
+      {\jfcomp{A}{f}{P}{F}{\ctxwk{S}{m}}}
+      {\ctxwk{\jfcomp{A}{f}{P}{F}{S}}{\jfcomp{A}{f}{P}{F}{m}}}
+    }
+\end{align*}
+\end{lem}
+
+\begin{lem}\label{lem:jfcomp-subst}
+Horizontal composition is compatible with substitution:
+\begin{align*}
+& \inference
+  { \jfhom{\Gamma}{A}{B}{f}{P}{Q}{F}
+    \jterm{{{{\Gamma}{B}}{Q}}{R}}{S}{k}
+    \jfam{{{{{\Gamma}{B}}{Q}}{R}}{S}}{T}
+    }
+  { \jfameq
+      {{{{\Gamma}{A}}{P}}{\jfcomp{A}{f}{P}{F}{R}}}
+      {\jfcomp{A}{f}{P}{F}{\subst{k}{T}}}
+      {\subst{\jfcomp{A}{f}{P}{F}{k}}{\jfcomp{A}{f}{P}{F}{T}}}
+    }
+  \\
+& \inference
+  { \jfhom{\Gamma}{A}{B}{f}{P}{Q}{F}
+    \jterm{{{{\Gamma}{B}}{Q}}{R}}{S}{k}
+    \jfam{{{{{{\Gamma}{B}}{Q}}{R}}{S}}{T}}{U}
+    }
+  { \jfameq
+      {{{{{\Gamma}{A}}{P}}{\jfcomp{A}{f}{P}{F}{R}}}{\jfcomp{A}{f}{P}{F}{\subst{k}{T}}}}
+      {\jfcomp{A}{f}{P}{F}{\subst{k}{U}}}
+      {\subst{\jfcomp{A}{f}{P}{F}{k}}{\jfcomp{A}{f}{P}{F}{U}}}
+    }
+  \\
+& \inference
+  { \jfhom{\Gamma}{A}{B}{f}{P}{Q}{F}
+    \jterm{{{{\Gamma}{B}}{Q}}{R}}{S}{k}
+    \jterm{{{{{{\Gamma}{B}}{Q}}{R}}{S}}{T}}{U}{m}
+    }
+  { \jfameq
+      {{{{{\Gamma}{A}}{P}}{\jfcomp{A}{f}{P}{F}{R}}}{\jfcomp{A}{f}{P}{F}{\subst{k}{T}}}}
+      {\jfcomp{A}{f}{P}{F}{\subst{k}{U}}}
+      {\jfcomp{A}{f}{P}{F}{\subst{k}{m}}}
+      {\subst{\jfcomp{A}{f}{P}{F}{k}}{\jfcomp{A}{f}{P}{F}{m}}}
+    }
+\end{align*}
+\end{lem}
+
+\begin{lem}\label{lem:jfcomp-idtm}
+Horizontal composition is compatible with identity terms:
+\begin{equation*}
+\inference
+  { \jfhom{\Gamma}{A}{B}{f}{P}{Q}{F}
+    \jfam{{{{\Gamma}{B}}{Q}}{R}}{S}
+    }
+  { \jtermeq
+      {{{{{\Gamma}{A}}{P}}{\jfcomp{A}{f}{P}{F}{R}}}{\jfcomp{A}{f}{P}{F}{S}}}
+      {\ctxwk{{\jfcomp{A}{f}{P}{F}{S}}}{\jfcomp{A}{f}{P}{F}{S}}}
+      {\jfcomp{A}{f}{P}{F}{\idtm{S}}}
+      {\idtm{\jfcomp{A}{f}{P}{F}{S}}}
+    }
+\end{equation*}
+\end{lem}
+
 There is also a notion of vertical composition, although this is not an
@@
 
-\subsection{Fiber inclusions}
-We will use the insights of \autoref{extension-on-terms} to define and study
-\emph{fiber inclusions}. The fiber inclusion of the \emph{fiber}
-$\subst{x}{P}$ into the extension $\ctxext{A}{P}$ is a morphism
-$\jhom{\Gamma}{\subst{x}{P}}{{A}{P}}{\finc{x}{P}}$, for any family
-$\jfam{{\Gamma}{A}}{P}$ and any term $\jterm{\Gamma}{A}{x}$. Then we will determine
-the ways in which it is compatible with the other operators. Note that in this
-subsection we will focus on the compatibility properties; the fact that
-the fiber inclusions also appear in a pullback diagram will be established in
-\autoref{pullback}.
-
-\begin{defn}
-Let $\jterm{\Gamma}{A}{x}$ be a term and let $\jfam{{\Gamma}{A}}{P}$ be a
-family. Then we define the \emph{fiber inclusion} of $\subst{x}{P}$ into
-$\ctxext{A}{P}$ in context $\Gamma$ to be the morphism
-\begin{equation*}
-\jhomdefn{\Gamma}{\subst{x}{P}}{{A}{P}}{\finc{x}{P}}{\unfoldnext{\finc{x}{P}}}.
-\end{equation*}
-\end{defn}
-
-We can immediately show that composing with a fiber inclusion is substitution.
-Thus, the following lemma asserts that we have a family pullback square
-\begin{equation*}
-\begin{tikzcd}
-\subst{x}{Q} \ar[fib]{d} \ar{r} & Q \ar[fib]{d} \\
-\subst{x}{P} \ar{r}[swap]{\finc{x}{P}} & \ctxext{A}{P}
-\end{tikzcd}
-\end{equation*}
-in context $\Gamma$.
-The upper arrow is just the identity morphism over $\finc{x}{P}$.
-
-\begin{lem}\label{lem:finc-precomp}
-Composition with $\finc{x}{P}$ is (the action on families of)
-substitution with $x$:
-\begin{align*}
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jfam{{{\Gamma}{A}}{P}}{Q}
-    }
-  { \jfameq
-      {{\Gamma}{\subst{x}{P}}}
-      {\jcomp{\subst{x}{P}}{\finc{x}{P}}{Q}}
-      {\subst{x}{Q}}
-    }
-  \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jfam{{{{\Gamma}{A}}{P}}{Q}}{R}
-    }
-  { \jfameq
-      {{{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}}
-      {\jcomp{\subst{x}{P}}{\finc{x}{P}}{R}}
-      {\subst{x}{R}}
-    }
-  \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jterm{{{{\Gamma}{A}}{P}}{Q}}{R}{h}
-    }
-  { \jtermeq
-      {{{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}}
-      {\subst{x}{R}}
-      {\jcomp{\subst{x}{P}}{\finc{x}{P}}{h}}
-      {\subst{x}{h}}
-    }
-\end{align*}
-\end{lem}
-
-\begin{proof}
-We only prove the validity of the first inference rule. Let $\jterm{\Gamma}{A}{x}$
-be a term and let $\jfam{{{\Gamma}{A}}{P}}{Q}$ be a family. Then we have the
-judgmental equalities
-\begin{align*}
-\jcomp{\subst{x}{P}}{\finc{x}{P}}{Q}
-& \jdeq
-  \unfold{\jcomp{\subst{x}{P}}{\unfold{\finc{x}{P}}}{Q}}
-  \tag{by definition}
-  \\
-& \jdeq
-  \subst
-    {\idtm{\subst{x}{P}}}
-    {{\ctxwk{\subst{x}{P}}{x}}{\ctxwk{\subst{x}{P}}{Q}}}
-  \tag{by \autoref{comp-es}}
-  \\
-& \jdeq
-  \subst
-    {\idtm{\subst{x}{P}}}
-    {\ctxwk{\subst{x}{P}}{\subst{x}{Q}}}
-  \tag{by \autoref{comp-ws-f}}
-  \\
-& \jdeq
-  \subst{x}{Q}.
-  \tag{by \autoref{idfunc-precomp}}
-\end{align*}
-\end{proof}
-
-We can give a second characterization of fiber inclusions:
-
-\begin{lem}\label{lem:finc-char2}
-We have the following inference rule
-\begin{equation*}
-\inference
-  { \jterm{\Gamma}{A}{x}
-    \jfam{{\Gamma}{A}}{P}
-    }
-  { \jhomeq{\Gamma}{\subst{x}{P}}{{A}{P}}{\finc{x}{P}}{\subst{x}{\idtm{{A}{P}}}}
-    } 
-\end{equation*}
-\end{lem}
-
-\begin{proof}
-The proof is the following calculation:
-\begin{align*}
-\finc{x}{P}
-& \jdeq 
-  \tmext{\ctxwk{\subst{x}{P}}{x}}{\idtm{\subst{x}{P}}}
-  \tag{by definition}
-  \\
-& \jdeq
-  \subst
-    { \idtm{\subst{x}{P}}
-      }
-    { { \ctxwk{\subst{x}{P}}{x}
-        }
-      { \idtm{\ctxext{\ctxwk{\subst{x}{P}}{A}}{\ctxwk{\subst{x}{P}}{P}}}
-        }
-      }
-  \tag{by definition}
-  \\
-& \jdeq
-  \subst
-    { \idtm{\subst{x}{P}}
-      }
-    { { \ctxwk{\subst{x}{P}}{x}
-        }
-      { \idtm{\ctxwk{\subst{x}{P}}{\ctxext{A}{P}}}
-        }
-      }
-  \tag{by \autoref{comp-we-f}}
-  \\
-& \jdeq
-  \subst
-    { \idtm{\subst{x}{P}}
-      }
-    { { \ctxwk{\subst{x}{P}}{x}
-        }
-      { \ctxwk{\subst{x}{P}}{\idtm{\ctxext{A}{P}}}
-        }
-      }
-  \tag{by \autoref{comp-wi-t}}
-  \\
-& \jdeq
-  \subst
-    { \idtm{\subst{x}{P}}
-      }
-    { \ctxwk
-        { \subst{x}{P}
-          }
-        { \subst{x}{\idtm{\ctxext{A}{P}}}
-          }
-      }
-  \tag{by \autoref{comp-ws-t}}
-  \\
-& \jdeq
-  \subst{x}{\idtm{\ctxext{A}{P}}}.
-  \tag{by \autoref{idfunc-precomp}}
-\end{align*}
-\end{proof}
-
-\begin{cor}
-The following inference rule is valid:
-\begin{equation*}
-\inference
-  { \jterm{\Gamma}{A}{x}
+\begin{lem}
+The following inference rules explain how vertically and horizontally composed 
+morphisms are evaluated:
+\begin{align*}
+& \inference
+  { \jfhom{\Gamma}{A}{B}{f}{P}{Q}{F}
+    \jterm{\Gamma}{A}{x}
     \jterm{\Gamma}{\subst{x}{P}}{u}
-    }
-  { \jtermeq{\Gamma}{{A}{P}}{\subst{u}{\finc{x}{P}}}{\tmext{x}{u}}
-    }
-\end{equation*}
-\end{cor}
-
-We have the following lemmas expressing the compatibility of the fiber
-inclusions with the empty context, extension, weakening and substitution.
-
-\begin{lem}
-The fiber inclusions are compatible with the empty families; i.e.~the following
-inference rules are valid
-\begin{align*}
-& \inference
-  { \jterm{\Gamma}{A}{x}
     }
@@
       {\Gamma}
-      {A}
-      {\finc{x}{\emptyf}}
-      {x}
-    }
-  \\
-& \inference
-  { \jfam{\Gamma}{A}
-    }
-  { \jhomeq
-      {\Gamma}
-      {A}
-      {A}
-      {\finc{\emptytm}{A}}
-      {\idtm{A}}
-    }
-\end{align*}
-\end{lem}
-
-\begin{proof}
-Let $\jterm{\Gamma}{A}{x}$. We have the judgmental equalities
-\begin{align*}
-\finc{x}{\emptyf}
-& \jdeq
-  \unfold{\finc{x}{\emptyf}}
+      {{B}{Q}}
+      {\subst{\tmext{}{}{x}{u}}{\jvcomp{P}{f}{F}}}
+      {\tmext{}{}{\subst{x}{f}}{\subst{u}{{x}{F}}}}
+    }
+\end{align*}
+\end{lem}
+
+\begin{proof}
+Let $\jfhom{\Gamma}{A}{B}{f}{P}{Q}{F}$, $\jterm{\Gamma}{A}{x}$ and
+$\jterm{\Gamma}{\subst{x}{P}}{u}$. Then we have the judgmental equalities
+\begin{align*}
+\subst{\tmext{}{}{x}{u}}{\jvcomp{P}{f}{F}}
+& \jdeq
+  \subst{\tmext{}{}{x}{u}}{\tmext{}{}{\ctxwk{P}{f}}{F}}
   \tag{by definition}
@@
 & \jdeq
-  \tmext{\ctxwk{\emptyf}{x}}{\idtm{\emptyf}}
-  \tag{by \autoref{comp-s0-c}}
-  \\
-& \jdeq
-  \tmext{x}{\idtm{\emptyf}}
-  \tag{by \autoref{comp-0w-t}}
-  \\
-& \jdeq
-  \tmext{x}{\emptytm}
-  \tag{by \autoref{comp-00-t}}
-  \\
-& \jdeq
-  x
-  \tag{by \autoref{lem:tmext-emp}}
-\end{align*}
-to prove the first judgmental equality. For the second, we have
-\begin{align*}
-\finc{\emptytm}{A}
-& \jdeq
-  \unfold{\finc{\emptytm}{A}}
-  \tag{by definition}
-  \\
-& \jdeq
-  \tmext{\emptytm}{\idtm{\subst{\emptytm}{A}}}
-  \tag{by \autoref{comp-00-t}}
-  \\
-& \jdeq
-  \idtm{\subst{\emptytm}{A}}
-  \tag{by \autoref{lem:tmext-emp}}
-  \\
-& \jdeq
-  \idtm{A}
-  \tag{by \autoref{comp-0s-f}}.
-\end{align*}
-\end{proof}
-
-\begin{lem}
-The fiber inclusions are compatible with extension; i.e.~the following inference
-rule is valid
-\begin{equation*}
-\inference
-  { \jterm{\Gamma}{A}{x}
-    \jterm{\Gamma}{\subst{x}{P}}{u}
-    \jfam{{{\Gamma}{A}}{P}}{Q}
-    }
-  { \jhomeq
-      {\Gamma}
-      {\subst{\tmext{x}{u}}{Q}}
-      {{{A}{P}}{Q}}
-      {\finc{\tmext{x}{u}}{Q}}
-      {\jcomp{}{\finc{u}{\subst{x}{Q}}}{\finc{x}{\ctxext{P}{Q}}}}
-    }
-\end{equation*}
-\end{lem}
-
-\begin{proof}
-Let $\jterm{\Gamma}{A}{x}$, $\jterm{\Gamma}{\subst{x}{P}}{u}$ and
-$\jfam{{{\Gamma}{A}}{P}}{Q}$. Then we have the judgmental equalities
-\begin{align*}
-\finc{\tmext{x}{u}}{Q}
-& \jdeq
-  \subst{\tmext{}{}{x}{u}}{\idtm{{{A}{P}}{Q}}}
-  \tag{by \autoref{lem:finc-char2}}
-  \\
-& \jdeq
-  \subst{u}{{x}{\idtm{{{A}{P}}{Q}}}}
+  \tmext{\subst{\tmext{x}{u}}{\ctxwk{P}{f}}}{\subst{\tmext{x}{u}}{F}}
+  \tag{by \autoref{lem:tmext-subst}}
+  \\
+& \jdeq
+  \tmext{\subst{u}{{x}{\ctxwk{P}{f}}}}{\subst{u}{{x}{F}}}
   \tag{by \autoref{comp-es}}
@@
 & \jdeq
-  \subst{u}{{x}{\idtm{{A}{{P}{Q}}}}}
-  \tag{by \autoref{comp-ee-c}}
-  \\
-& \jdeq
-  \subst{u}{\finc{x}{\ctxext{P}{Q}}}
-  \tag{by \autoref{lem:finc-char2}}
-  \\
-& \jdeq
-  \jcomp{}{\finc{u}{\subst{x}{Q}}}{\finc{x}{\ctxext{P}{Q}}}
-  \tag{by \autoref{lem:finc-precomp}}
-\end{align*}
-\end{proof}
-
-\begin{lem}
-The fiber inclusions are compatible with weakening; i.e.~the following inference
-rule is valid
-\begin{equation*}
-\inference
-  { \jfam{\Gamma}{A}
-    \jterm{\Gamma}{B}{y}
-    \jfam{{\Gamma}{B}}{Q}
-    }
-  { \jhomeq
-      {{\Gamma}{A}}
-      {\subst{\ctxwk{A}{y}}{\ctxwk{A}{Q}}}
-      {{\ctxwk{A}{B}}{\ctxwk{A}{Q}}}
-      {\finc{\ctxwk{A}{y}}{\ctxwk{A}{Q}}}
-      {\ctxwk{A}{\finc{y}{Q}}}
-    }
-\end{equation*}
-\end{lem}
-
-\begin{proof}
-Let $\jterm{\Gamma}{B}{y}$ and $\jfam{{\Gamma}{B}}{Q}$. Then we have the
-judgmental equalities
-\begin{align*}
-\finc{\ctxwk{A}{y}}{\ctxwk{A}{Q}}
-& \jdeq
-  \subst{\ctxwk{A}{y}}{\idtm{\ctxext{\ctxwk{A}{B}}{\ctxwk{A}{Q}}}}
-  \tag{by \autoref{lem:finc-char2}}
-  \\
-& \jdeq
-  \subst{\ctxwk{A}{y}}{\idtm{\ctxwk{A}{\ctxext{B}{Q}}}}
-  \tag{by \autoref{comp-we-f}}
-  \\
-& \jdeq
-  \subst{\ctxwk{A}{y}}{\ctxwk{A}{\idtm{\ctxext{B}{Q}}}}
-  \tag{by \autoref{comp-wi-t}}
-  \\
-& \jdeq
-  \ctxwk{A}{\subst{y}{\idtm{{B}{Q}}}}
-  \tag{by \autoref{comp-ws-t}}
-  \\
-& \jdeq
-  \ctxwk{A}{\finc{y}{Q}}.
-  \tag{by \autoref{lem:finc-char2}}
-\end{align*}
-\end{proof}
-
-\begin{lem}
-The fiber inclusions are compatible with substitution; i.e.~the following
-inference rule is valid
-\begin{equation*}
-\inference
-  { \jterm{\Gamma}{A}{x}
-    \jfam{{{\Gamma}{A}}{P}}{Q}
-    \jterm{{\Gamma}{A}}{P}{f}
-    }
-  { \jhomeq
-      {\Gamma}
-      {\subst{{x}{f}}{{x}{Q}}}
-      {{\subst{x}{P}}{\subst{x}{Q}}}
-      {\finc{\subst{x}{f}}{\subst{x}{Q}}}
-      {\subst{x}{\finc{f}{Q}}}
-    }
-\end{equation*}
-\end{lem}
-
-\begin{proof}
-Let $\jterm{\Gamma}{A}{x}$, $\jterm{{\Gamma}{A}}{P}{f}$ and 
-$\jfam{{{\Gamma}{A}}{P}}{Q}$. Then we have the judgmental equalities
-\begin{align*}
-\finc{\subst{x}{f}}{\subst{x}{Q}}
-& \jdeq
-  \subst{x}{{f}{\idtm{{P}{Q}}}}
-  \tag{by \autoref{lem:finc-char2}}
-  \\
-& \jdeq
-  \subst{{x}{f}}{{x}{\idtm{{P}{Q}}}}
-  \tag{by \autoref{comp-ss-t}}
-  \\
-& \jdeq
-  \subst{{x}{f}}{\idtm{\subst{x}{\ctxext{P}{Q}}}}
-  \tag{by \autoref{comp-si-t}}
-  \\
-& \jdeq
-  \subst{{x}{f}}{\idtm{\ctxext{\subst{x}{P}}{\subst{x}{Q}}}}
-  \tag{by \autoref{comp-se-f}}
-  \\
-& \jdeq
-  \finc{\subst{x}{f}}{\subst{x}{Q}}.
-  \tag{by \autoref{lem:finc-char2}}
-\end{align*}
-\end{proof}
-
-\begin{lem}
-The fiber inclusions are compatible with identity terms; i.e.~the following
-inference rule is valid
-\begin{equation*}
-\inference
-  { \jfam{{\Gamma}{A}}{P}
-    }
-  { \jhomeq
-      {{\Gamma}{A}}
-      {P}
-      {\ctxwk{A}{\ctxext{A}{P}}}
-      {\finc{\idtm{A}}{\ctxwk{A}{P}}}
-      {\idtm{{A}{P}}}
-    }
-\end{equation*}
-\end{lem}
-
-\begin{proof}
-Let $\jfam{\Gamma}{A}$. Then we have the judgmental equalities
-\begin{align*}
-\finc{\idtm{A}}{\ctxwk{A}{P}}
-& \jdeq
-  \subst{\idtm{A}}{\idtm{\ctxext{\ctxwk{A}{A}}{\ctxwk{A}{P}}}}
-  \tag{by \autoref{lem:finc-char2}}
-  \\
-& \jdeq
-  \subst{\idtm{A}}{\idtm{\ctxwk{A}{\ctxext{A}{P}}}}
-  \tag{by \autoref{comp-we-f}}
-  \\
-& \jdeq
-  \subst{\idtm{A}}{\ctxwk{A}{\idtm{{A}{P}}}}
-  \tag{by \autoref{comp-wi-t}}
-  \\
-& \jdeq
-  \idtm{{A}{P}}
-  \tag{by \autoref{idfunc-precomp}}
-\end{align*}
-\end{proof}
-
-\subsection{Pullback squares and family pullback squares}
+  \tmext{\subst{u}{\ctxwk{\subst{x}{P}}{\subst{x}{f}}}}{\subst{u}{{x}{F}}}
+  \tag{by \autoref{comp-sw-t}}
+  \\
+& \jdeq
+  \tmext{\subst{x}{f}}{\subst{u}{{x}{F}}}.
+  \tag{by \autoref{cancellation-ws-t}}
+\end{align*}
+\end{proof}
+
 \label{pullback}
-Now that we have introduced the notions of morphisms, composition and identity
-terms, we can develop a diagramatic style of of displaying type dependencies
+Now that we have introduced the notions of morphisms and composition,
+we can develop a diagramatic style of of displaying type dependencies
 combined with morphisms. We give an informal, metatheoretical definition of
@@
 
+\subsection{Fiber inclusions}
+We will use the insights of \autoref{extension-on-terms} to define and study
+\emph{fiber inclusions}. The fiber inclusion of the \emph{fiber}
+$\subst{x}{P}$ into the extension $\ctxext{A}{P}$ is a morphism
+$\jhom{\Gamma}{\subst{x}{P}}{{A}{P}}{\finc{x}{P}}$, for any family
+$\jfam{{\Gamma}{A}}{P}$ and any term $\jterm{\Gamma}{A}{x}$. Then we will determine
+the ways in which it is compatible with the other operators. Note that in this
+subsection we will focus on the compatibility properties; the fact that
+the fiber inclusions also appear in a pullback diagram will be established in
+\autoref{pullback}.
+
+\begin{defn}
+Let $\jterm{\Gamma}{A}{x}$ be a term and let $\jfam{{\Gamma}{A}}{P}$ be a
+family. Then we define the \emph{fiber inclusion} of $\subst{x}{P}$ into
+$\ctxext{A}{P}$ in context $\Gamma$ to be the morphism
+\begin{equation*}
+\jhomdefn{\Gamma}{\subst{x}{P}}{{A}{P}}{\finc{x}{P}}{\unfoldnext{\finc{x}{P}}}.
+\end{equation*}
+\end{defn}
+
+We can immediately show that composing with a fiber inclusion is substitution.
+Thus, the following lemma asserts that we have a family pullback square
+\begin{equation*}
+\begin{tikzcd}
+\subst{x}{Q} \ar[fib]{d} \ar{r} & Q \ar[fib]{d} \\
+\subst{x}{P} \ar{r}[swap]{\finc{x}{P}} & \ctxext{A}{P}
+\end{tikzcd}
+\end{equation*}
+in context $\Gamma$.
+The upper arrow is just the identity morphism over $\finc{x}{P}$.
+
+\begin{lem}\label{lem:finc-precomp}
+Composition with $\finc{x}{P}$ is (the action on families of)
+substitution with $x$:
+\begin{align*}
+& \inference
+  { \jterm{\Gamma}{A}{x}
+    \jfam{{{\Gamma}{A}}{P}}{Q}
+    }
+  { \jfameq
+      {{\Gamma}{\subst{x}{P}}}
+      {\jcomp{\subst{x}{P}}{\finc{x}{P}}{Q}}
+      {\subst{x}{Q}}
+    }
+  \\
+& \inference
+  { \jterm{\Gamma}{A}{x}
+    \jfam{{{{\Gamma}{A}}{P}}{Q}}{R}
+    }
+  { \jfameq
+      {{{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}}
+      {\jcomp{\subst{x}{P}}{\finc{x}{P}}{R}}
+      {\subst{x}{R}}
+    }
+  \\
+& \inference
+  { \jterm{\Gamma}{A}{x}
+    \jterm{{{{\Gamma}{A}}{P}}{Q}}{R}{h}
+    }
+  { \jtermeq
+      {{{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}}
+      {\subst{x}{R}}
+      {\jcomp{\subst{x}{P}}{\finc{x}{P}}{h}}
+      {\subst{x}{h}}
+    }
+\end{align*}
+\end{lem}
+
+\begin{proof}
+We only prove the validity of the first inference rule. Let $\jterm{\Gamma}{A}{x}$
+be a term and let $\jfam{{{\Gamma}{A}}{P}}{Q}$ be a family. Then we have the
+judgmental equalities
+\begin{align*}
+\jcomp{\subst{x}{P}}{\finc{x}{P}}{Q}
+& \jdeq
+  \unfold{\jcomp{\subst{x}{P}}{\unfold{\finc{x}{P}}}{Q}}
+  \tag{by definition}
+  \\
+& \jdeq
+  \subst
+    {\idtm{\subst{x}{P}}}
+    {{\ctxwk{\subst{x}{P}}{x}}{\ctxwk{\subst{x}{P}}{Q}}}
+  \tag{by \autoref{comp-es}}
+  \\
+& \jdeq
+  \subst
+    {\idtm{\subst{x}{P}}}
+    {\ctxwk{\subst{x}{P}}{\subst{x}{Q}}}
+  \tag{by \autoref{comp-ws-f}}
+  \\
+& \jdeq
+  \subst{x}{Q}.
+  \tag{by \autoref{idfunc-precomp}}
+\end{align*}
+\end{proof}
+
+We can give a second characterization of fiber inclusions:
+
+\begin{lem}\label{lem:finc-char2}
+We have the following inference rule
+\begin{equation*}
+\inference
+  { \jterm{\Gamma}{A}{x}
+    \jfam{{\Gamma}{A}}{P}
+    }
+  { \jhomeq{\Gamma}{\subst{x}{P}}{{A}{P}}{\finc{x}{P}}{\subst{x}{\idtm{{A}{P}}}}
+    } 
+\end{equation*}
+\end{lem}
+
+\begin{proof}
+The proof is the following calculation:
+\begin{align*}
+\finc{x}{P}
+& \jdeq 
+  \tmext{\ctxwk{\subst{x}{P}}{x}}{\idtm{\subst{x}{P}}}
+  \tag{by definition}
+  \\
+& \jdeq
+  \subst
+    { \idtm{\subst{x}{P}}
+      }
+    { { \ctxwk{\subst{x}{P}}{x}
+        }
+      { \idtm{\ctxext{\ctxwk{\subst{x}{P}}{A}}{\ctxwk{\subst{x}{P}}{P}}}
+        }
+      }
+  \tag{by definition}
+  \\
+& \jdeq
+  \subst
+    { \idtm{\subst{x}{P}}
+      }
+    { { \ctxwk{\subst{x}{P}}{x}
+        }
+      { \idtm{\ctxwk{\subst{x}{P}}{\ctxext{A}{P}}}
+        }
+      }
+  \tag{by \autoref{comp-we-f}}
+  \\
+& \jdeq
+  \subst
+    { \idtm{\subst{x}{P}}
+      }
+    { { \ctxwk{\subst{x}{P}}{x}
+        }
+      { \ctxwk{\subst{x}{P}}{\idtm{\ctxext{A}{P}}}
+        }
+      }
+  \tag{by \autoref{comp-wi-t}}
+  \\
+& \jdeq
+  \subst
+    { \idtm{\subst{x}{P}}
+      }
+    { \ctxwk
+        { \subst{x}{P}
+          }
+        { \subst{x}{\idtm{\ctxext{A}{P}}}
+          }
+      }
+  \tag{by \autoref{comp-ws-t}}
+  \\
+& \jdeq
+  \subst{x}{\idtm{\ctxext{A}{P}}}.
+  \tag{by \autoref{idfunc-precomp}}
+\end{align*}
+\end{proof}
+
+\begin{cor}
+The following inference rule is valid:
+\begin{equation*}
+\inference
+  { \jterm{\Gamma}{A}{x}
+    \jterm{\Gamma}{\subst{x}{P}}{u}
+    }
+  { \jtermeq{\Gamma}{{A}{P}}{\subst{u}{\finc{x}{P}}}{\tmext{x}{u}}
+    }
+\end{equation*}
+\end{cor}
+
+We have the following lemmas expressing the compatibility of the fiber
+inclusions with the empty context, extension, weakening and substitution.
+
+\begin{lem}
+The fiber inclusions are compatible with the empty families; i.e.~the following
+inference rules are valid
+\begin{align*}
+& \inference
+  { \jterm{\Gamma}{A}{x}
+    }
+  { \jtermeq
+      {\Gamma}
+      {A}
+      {\finc{x}{\emptyf}}
+      {x}
+    }
+  \\
+& \inference
+  { \jfam{\Gamma}{A}
+    }
+  { \jhomeq
+      {\Gamma}
+      {A}
+      {A}
+      {\finc{\emptytm}{A}}
+      {\idtm{A}}
+    }
+\end{align*}
+\end{lem}
+
+\begin{proof}
+Let $\jterm{\Gamma}{A}{x}$. We have the judgmental equalities
+\begin{align*}
+\finc{x}{\emptyf}
+& \jdeq
+  \unfold{\finc{x}{\emptyf}}
+  \tag{by definition}
+  \\
+& \jdeq
+  \tmext{\ctxwk{\emptyf}{x}}{\idtm{\emptyf}}
+  \tag{by \autoref{comp-s0-c}}
+  \\
+& \jdeq
+  \tmext{x}{\idtm{\emptyf}}
+  \tag{by \autoref{comp-0w-t}}
+  \\
+& \jdeq
+  \tmext{x}{\emptytm}
+  \tag{by \autoref{comp-00-t}}
+  \\
+& \jdeq
+  x
+  \tag{by \autoref{lem:tmext-emp}}
+\end{align*}
+to prove the first judgmental equality. For the second, we have
+\begin{align*}
+\finc{\emptytm}{A}
+& \jdeq
+  \unfold{\finc{\emptytm}{A}}
+  \tag{by definition}
+  \\
+& \jdeq
+  \tmext{\emptytm}{\idtm{\subst{\emptytm}{A}}}
+  \tag{by \autoref{comp-00-t}}
+  \\
+& \jdeq
+  \idtm{\subst{\emptytm}{A}}
+  \tag{by \autoref{lem:tmext-emp}}
+  \\
+& \jdeq
+  \idtm{A}
+  \tag{by \autoref{comp-0s-f}}.
+\end{align*}
+\end{proof}
+
+\begin{lem}
+The fiber inclusions are compatible with extension; i.e.~the following inference
+rule is valid
+\begin{equation*}
+\inference
+  { \jterm{\Gamma}{A}{x}
+    \jterm{\Gamma}{\subst{x}{P}}{u}
+    \jfam{{{\Gamma}{A}}{P}}{Q}
+    }
+  { \jhomeq
+      {\Gamma}
+      {\subst{\tmext{x}{u}}{Q}}
+      {{{A}{P}}{Q}}
+      {\finc{\tmext{x}{u}}{Q}}
+      {\jcomp{}{\finc{u}{\subst{x}{Q}}}{\finc{x}{\ctxext{P}{Q}}}}
+    }
+\end{equation*}
+\end{lem}
+
+\begin{proof}
+Let $\jterm{\Gamma}{A}{x}$, $\jterm{\Gamma}{\subst{x}{P}}{u}$ and
+$\jfam{{{\Gamma}{A}}{P}}{Q}$. Then we have the judgmental equalities
+\begin{align*}
+\finc{\tmext{x}{u}}{Q}
+& \jdeq
+  \subst{\tmext{}{}{x}{u}}{\idtm{{{A}{P}}{Q}}}
+  \tag{by \autoref{lem:finc-char2}}
+  \\
+& \jdeq
+  \subst{u}{{x}{\idtm{{{A}{P}}{Q}}}}
+  \tag{by \autoref{comp-es}}
+  \\
+& \jdeq
+  \subst{u}{{x}{\idtm{{A}{{P}{Q}}}}}
+  \tag{by \autoref{comp-ee-c}}
+  \\
+& \jdeq
+  \subst{u}{\finc{x}{\ctxext{P}{Q}}}
+  \tag{by \autoref{lem:finc-char2}}
+  \\
+& \jdeq
+  \jcomp{}{\finc{u}{\subst{x}{Q}}}{\finc{x}{\ctxext{P}{Q}}}
+  \tag{by \autoref{lem:finc-precomp}}
+\end{align*}
+\end{proof}
+
+\begin{lem}
+The fiber inclusions are compatible with weakening; i.e.~the following inference
+rule is valid
+\begin{equation*}
+\inference
+  { \jfam{\Gamma}{A}
+    \jterm{\Gamma}{B}{y}
+    \jfam{{\Gamma}{B}}{Q}
+    }
+  { \jhomeq
+      {{\Gamma}{A}}
+      {\subst{\ctxwk{A}{y}}{\ctxwk{A}{Q}}}
+      {{\ctxwk{A}{B}}{\ctxwk{A}{Q}}}
+      {\finc{\ctxwk{A}{y}}{\ctxwk{A}{Q}}}
+      {\ctxwk{A}{\finc{y}{Q}}}
+    }
+\end{equation*}
+\end{lem}
+
+\begin{proof}
+Let $\jterm{\Gamma}{B}{y}$ and $\jfam{{\Gamma}{B}}{Q}$. Then we have the
+judgmental equalities
+\begin{align*}
+\finc{\ctxwk{A}{y}}{\ctxwk{A}{Q}}
+& \jdeq
+  \subst{\ctxwk{A}{y}}{\idtm{\ctxext{\ctxwk{A}{B}}{\ctxwk{A}{Q}}}}
+  \tag{by \autoref{lem:finc-char2}}
+  \\
+& \jdeq
+  \subst{\ctxwk{A}{y}}{\idtm{\ctxwk{A}{\ctxext{B}{Q}}}}
+  \tag{by \autoref{comp-we-f}}
+  \\
+& \jdeq
+  \subst{\ctxwk{A}{y}}{\ctxwk{A}{\idtm{\ctxext{B}{Q}}}}
+  \tag{by \autoref{comp-wi-t}}
+  \\
+& \jdeq
+  \ctxwk{A}{\subst{y}{\idtm{{B}{Q}}}}
+  \tag{by \autoref{comp-ws-t}}
+  \\
+& \jdeq
+  \ctxwk{A}{\finc{y}{Q}}.
+  \tag{by \autoref{lem:finc-char2}}
+\end{align*}
+\end{proof}
+
+\begin{lem}
+The fiber inclusions are compatible with substitution; i.e.~the following
+inference rule is valid
+\begin{equation*}
+\inference
+  { \jterm{\Gamma}{A}{x}
+    \jfam{{{\Gamma}{A}}{P}}{Q}
+    \jterm{{\Gamma}{A}}{P}{f}
+    }
+  { \jhomeq
+      {\Gamma}
+      {\subst{{x}{f}}{{x}{Q}}}
+      {{\subst{x}{P}}{\subst{x}{Q}}}
+      {\finc{\subst{x}{f}}{\subst{x}{Q}}}
+      {\subst{x}{\finc{f}{Q}}}
+    }
+\end{equation*}
+\end{lem}
+
+\begin{proof}
+Let $\jterm{\Gamma}{A}{x}$, $\jterm{{\Gamma}{A}}{P}{f}$ and 
+$\jfam{{{\Gamma}{A}}{P}}{Q}$. Then we have the judgmental equalities
+\begin{align*}
+\finc{\subst{x}{f}}{\subst{x}{Q}}
+& \jdeq
+  \subst{x}{{f}{\idtm{{P}{Q}}}}
+  \tag{by \autoref{lem:finc-char2}}
+  \\
+& \jdeq
+  \subst{{x}{f}}{{x}{\idtm{{P}{Q}}}}
+  \tag{by \autoref{comp-ss-t}}
+  \\
+& \jdeq
+  \subst{{x}{f}}{\idtm{\subst{x}{\ctxext{P}{Q}}}}
+  \tag{by \autoref{comp-si-t}}
+  \\
+& \jdeq
+  \subst{{x}{f}}{\idtm{\ctxext{\subst{x}{P}}{\subst{x}{Q}}}}
+  \tag{by \autoref{comp-se-f}}
+  \\
+& \jdeq
+  \finc{\subst{x}{f}}{\subst{x}{Q}}.
+  \tag{by \autoref{lem:finc-char2}}
+\end{align*}
+\end{proof}
+
+\begin{lem}
+The fiber inclusions are compatible with identity terms; i.e.~the following
+inference rule is valid
+\begin{equation*}
+\inference
+  { \jfam{{\Gamma}{A}}{P}
+    }
+  { \jhomeq
+      {{\Gamma}{A}}
+      {P}
+      {\ctxwk{A}{\ctxext{A}{P}}}
+      {\finc{\idtm{A}}{\ctxwk{A}{P}}}
+      {\idtm{{A}{P}}}
+    }
+\end{equation*}
+\end{lem}
+
+\begin{proof}
+Let $\jfam{\Gamma}{A}$. Then we have the judgmental equalities
+\begin{align*}
+\finc{\idtm{A}}{\ctxwk{A}{P}}
+& \jdeq
+  \subst{\idtm{A}}{\idtm{\ctxext{\ctxwk{A}{A}}{\ctxwk{A}{P}}}}
+  \tag{by \autoref{lem:finc-char2}}
+  \\
+& \jdeq
+  \subst{\idtm{A}}{\idtm{\ctxwk{A}{\ctxext{A}{P}}}}
+  \tag{by \autoref{comp-we-f}}
+  \\
+& \jdeq
+  \subst{\idtm{A}}{\ctxwk{A}{\idtm{{A}{P}}}}
+  \tag{by \autoref{comp-wi-t}}
+  \\
+& \jdeq
+  \idtm{{A}{P}}
+  \tag{by \autoref{idfunc-precomp}}
+\end{align*}
+\end{proof}
+
 \begin{comment}
@@
 formulation and it is (perhaps surprisingly) quite involved to formulate it.
-It is however a necessary task, because extension algebras stand at the basis
-of implementing any theory with dependency in the theory of contexts, families
-and terms (i.e.~you don't stand a chance without extension algebras).
 
@@
 \begin{align*}
-\jhom*{\Gamma}{{A}{P}}{A}{e}\\
-\jhom*{{\Gamma}{A}}{{P}{\jcomp{}{e}{P}}}{P}{f}.
+\jhom*{\Gamma}{{A}{P}}{A}{\epsilon_0}\\
+\jhom*{{\Gamma}{A}}{{P}{\jcomp{}{\epsilon_0}{P}}}{P}{\epsilon_1}.
 \end{align*}
@@
 \begin{rmk}
-Instead of looking at $f$ as a context morphism from $\ctxext{P}{\jcomp{}{e}{P}}$
-to $P$, one could also look at $f$ as a morphism \emph{over $\cprojfstf{A}{P}$},
+Instead of looking at $\epsilon_1$ as a context morphism from $\ctxext{P}{\jcomp{}{\epsilon_0}{P}}$
+to $P$, one could also look at $\epsilon_1$ as a morphism \emph{over $\cprojfstf{A}{P}$},
 as indicated in the following diagram:
@@
   \ar[fib]{d}
-& \jcomp{}{e}{P}
+& \jcomp{}{\epsilon_0}{P}
   \ar[fib]{d}
-  \ar{l}[swap]{f}
+  \ar{l}[swap]{\epsilon_1}
   \ar{r}
@@
   \ar{l}{\cprojfstf{A}{P}}
-  \ar{r}[swap]{e}
+  \ar{r}[swap]{\epsilon_0}
 & A
@@
 \end{equation*}
-This makes it clear that $f$ takes a family over an extended context as an
+This makes it clear that $\epsilon_1$ takes a family over an extended context as an
 argument. The extended context consists of a `base part' and a `family part'. 
-The output of $f$ is a new (extended) family over that base part. Forgetting 
+The output of $\epsilon_1$ is a new (extended) family over that base part. Forgetting 
 the family part is what the projection takes care of.
@@
 Extension also satisfies the properties explained in \autoref{comp-ee}, so we
-must find the two judgmental equalities for $e$ and $f$ mimicing those. 
+must find the two judgmental equalities for $\epsilon_0$ and $\epsilon_1$ mimicing those. 
 The first of these judgmental equalities is easy to give: it says that the
@@
 \begin{tikzcd}[column sep=huge]
-\ctxext{A}{{P}{\jcomp{}{e}{P}}} 
-  \ar{d}[swap]{\jvcomp{}{e}{\idtm{\jcomp{}{e}{P}}}
-    %\tmext{\ctxwk{\jcomp{}{e}{P}}{e}}{\idtm{\jcomp{}{e}{P}}}
-    } 
-  \ar{r}{\jvcomp{}{\idtm{A}}{f}
-    %\tmext{\ctxwk{\ctxext{P}{\jcomp{}{e}{P}}}{\idtm{A}}}{f}
-    } 
-  & \ctxext{A}{P} \ar{d}{e}\\
-\ctxext{A}{P} \ar{r}[swap]{e} & A
+\ctxext{A}{{P}{\jcomp{}{\epsilon_0}{P}}} 
+  \ar{d}[swap]{\jvcomp{}{\epsilon_0}{\idtm{\jcomp{}{\epsilon_0}{P}}}
+    } 
+  \ar{r}{\jvcomp{}{\idtm{A}}{\epsilon_1}
+    } 
+  & \ctxext{A}{P} \ar{d}{\epsilon_0}\\
+\ctxext{A}{P} \ar{r}[swap]{\epsilon_0} & A
 \end{tikzcd}
@@
 \begin{lem}
-Let $A$, $P$, $e$ and $f$ be as above, satisfying \autoref{eq:ealg-eq1} and
+Let $A$, $P$, $\epsilon_0$ and $\epsilon_1$ be as above, satisfying \autoref{eq:ealg-eq1} and
 let $x_0:A$,
-$x_1:\subst{x_0}{P}$ and $x_2:\subst{x_1}{{x_0}{\jcomp{}{e}{P}}}$.
+$x_1:\subst{x_0}{P}$ and $x_2:\subst{x_1}{{x_0}{\jcomp{}{\epsilon_0}{P}}}$.
 Then we have the judgmental equality
 \begin{equation*}
-\subst{{x_2}{{x_1}{f}}}{{x_0}{e}}
+\subst{{x_2}{{x_1}{\epsilon_1}}}{{x_0}{\epsilon_0}}
 \jdeq
-\subst{x_2}{{{x_1}{{x_0}{e}}}{e}}.
+\subst{x_2}{{{x_1}{{x_0}{\epsilon_0}}}{\epsilon_0}}.
 \end{equation*}
@@
 \begin{align*}
-\subst{{x_2}{{x_1}{f}}}{{x_0}{e}}
-& \jdeq
-  \subst{\tmext{x_0}{\subst{x_2}{{x_1}{f}}}}{e}
-  \\
-& \jdeq 
-  \subst{{x_2}{{x_1}{{x_0}{\jvcomp{}{\idtm{A}}{f}}}}}{e}
-  \\
-& \jdeq
-  \subst{x_2}{{x_1}{{x_0}{\jcomp{}{\jvcomp{}{\idtm{A}}{f}}{e}}}}
-  \\
-& \jdeq
-  \subst{x_2}{{x_1}{{x_0}{\jcomp{}{\jvcomp{}{e}{\idtm{\jcomp{}{e}{P}}}}{e}}}}
-  \\
-& \jdeq 
-  \subst{{x_2}{{x_1}{{x_0}{\jvcomp{}{e}{\idtm{\jcomp{}{e}{P}}}}}}}{e}
-  \\
-& \jdeq
-  \subst{\tmext{\subst{x_1}{{x_0}{e}}}{x_2}}{e}
-  \\
-& \jdeq
-  \subst{x_2}{{{x_1}{{x_0}{e}}}{e}}.\qedhere
+\subst{{x_2}{{x_1}{\epsilon_1}}}{{x_0}{\epsilon_0}}
+& \jdeq
+  \subst{\tmext{x_0}{\subst{x_2}{{x_1}{\epsilon_1}}}}{\epsilon_0}
+  \\
+& \jdeq 
+  \subst{{x_2}{{x_1}{{x_0}{\jvcomp{}{\idtm{A}}{\epsilon_1}}}}}{\epsilon_0}
+  \\
+& \jdeq
+  \subst{x_2}{{x_1}{{x_0}{\jcomp{}{\jvcomp{}{\idtm{A}}{\epsilon_1}}{\epsilon_0}}}}
+  \\
+& \jdeq
+  \subst{x_2}{{x_1}{{x_0}{\jcomp{}{\jvcomp{}{\epsilon_0}{\idtm{\jcomp{}{\epsilon_0}{P}}}}{\epsilon_0}}}}
+  \\
+& \jdeq 
+  \subst{{x_2}{{x_1}{{x_0}{\jvcomp{}{\epsilon_0}{\idtm{\jcomp{}{\epsilon_0}{P}}}}}}}{\epsilon_0}
+  \\
+& \jdeq
+  \subst{\tmext{\subst{x_1}{{x_0}{\epsilon_0}}}{x_2}}{\epsilon_0}
+  \\
+& \jdeq
+  \subst{x_2}{{{x_1}{{x_0}{\epsilon_0}}}{\epsilon_0}}.\qedhere
 \end{align*}
@@
 to consider `higher' families, i.e.~families over families over families, and
-thus we need to look at the family $\jcomp{}{e}{\jcomp{}{e}{P}}$ and find the
+thus we need to look at the family $\jcomp{}{\epsilon_0}{\jcomp{}{\epsilon_0}{P}}$ and find the
 two dotted morphisms in the diagram
@@
 \begin{tikzcd}
-\ctxext{P}{{\jcomp{}{e}{P}}{\jcomp{}{e}{\jcomp{}{e}{P}}}}
+\ctxext{P}{{\jcomp{}{\epsilon_0}{P}}{\jcomp{}{\epsilon_0}{\jcomp{}{\epsilon_0}{P}}}}
   \ar[densely dotted]{d}
   \ar[densely dotted]{r}
-& \ctxext{P}{\jcomp{}{e}{P}} \ar{d}{f}\\
-\ctxext{P}{\jcomp{}{e}{P}} \ar{r}{f} & P
+& \ctxext{P}{\jcomp{}{\epsilon_0}{P}} \ar{d}{\epsilon_1}\\
+\ctxext{P}{\jcomp{}{\epsilon_0}{P}} \ar{r}{\epsilon_1} & P
 \end{tikzcd}
@@
 \begin{equation*}
-\ctxext{\jcomp{}{e}{P}}{\jcomp{}{e}{\jcomp{}{e}{P}}}
+\ctxext{\jcomp{}{\epsilon_0}{P}}{\jcomp{}{\epsilon_0}{\jcomp{}{\epsilon_0}{P}}}
   \jdeq
-  \jcomp{}{e}{\ctxext{P}{\jcomp{}{e}{P}}}
+  \jcomp{}{\epsilon_0}{\ctxext{P}{\jcomp{}{\epsilon_0}{P}}}
 \end{equation*}
@@
   {{\Gamma}{A}}
-  {{P}{{\jcomp{}{e}{P}}{\jcomp{}{e}{\jcomp{}{e}{P}}}}}
-  {{P}{\jcomp{}{e}{P}}}
-  {\jvcomp{}{\idtm{P}}{\jcomp{}{e}{f}}}.
+  {{P}{{\jcomp{}{\epsilon_0}{P}}{\jcomp{}{\epsilon_0}{\jcomp{}{\epsilon_0}{P}}}}}
+  {{P}{\jcomp{}{\epsilon_0}{P}}}
+  {\jvcomp{}{\idtm{P}}{\jcomp{}{\epsilon_0}{\epsilon_1}}}.
 \end{equation*}
 For the other morphism we need to look at the family
-$\ctxext{P}{{\jcomp{}{e}{P}}{\jcomp{}{e}{\jcomp{}{e}{P}}}}$ differently. We do
+$\ctxext{P}{{\jcomp{}{\epsilon_0}{P}}{\jcomp{}{\epsilon_0}{\jcomp{}{\epsilon_0}{P}}}}$ differently. We do
 that in the following lemma.
@@
 \begin{lem}
-Suppose we have $A$, $P$, $e$ and $f$ satisfying \autoref{eq:ealg-eq1}. Then
+Suppose we have $A$, $P$, $\epsilon_0$ and $\epsilon_1$ satisfying \autoref{eq:ealg-eq1}. Then
 we have the judgmental equality
 \begin{equation*}
-\jcomp{}{e}{\jcomp{}{e}{P}}
+\jcomp{}{\epsilon_0}{\jcomp{}{\epsilon_0}{P}}
   \jdeq
-  \jcomp{}{f}{\jcomp{}{e}{P}}.
+  \jcomp{}{\epsilon_1}{\jcomp{}{\epsilon_0}{P}}.
 \end{equation*}
@@
 \begin{align*}
-\jcomp{}{e}{\jcomp{}{e}{P}}
-& \jdeq
-  \subst{e}
-    {\ctxwk{\ctxext{A}{P}}{\jcomp{}{e}{P}}}
-  \\
-& \jdeq
-  \subst
-    {\idtm{\jcomp{}{e}{P}}}
+\jcomp{}{\epsilon_0}{\jcomp{}{\epsilon_0}{P}}
+& \jdeq
+  \subst{\epsilon_0}
+    {\ctxwk{\ctxext{A}{P}}{\jcomp{}{\epsilon_0}{P}}}
+  \\
+& \jdeq
+  \subst
+    {\idtm{\jcomp{}{\epsilon_0}{P}}}
     { \ctxwk
-        {\jcomp{}{e}{P}}{\subst{e}
-        {\ctxwk{\ctxext{A}{P}}{\jcomp{}{e}{P}}}}}
-  \\
-& \jdeq
-  \subst
-    {\idtm{\jcomp{}{e}{P}}}
-    { {\ctxwk{\jcomp{}{e}{P}}{e}}
-      {\ctxwk{\jcomp{}{e}{P}}{{\ctxext{A}{P}}{\jcomp{}{e}{P}}}}}
-  \\
-& \jdeq
-  \subst
-    {\idtm{\jcomp{}{e}{P}}}
-    { {\ctxwk{\jcomp{}{e}{P}}{e}}
-      {\ctxwk{\ctxext{{A}{P}}{\jcomp{}{e}{P}}}{\jcomp{}{e}{P}}}}
-  \\
-& \jdeq
-  \subst
-    {\idtm{\jcomp{}{e}{P}}}
-    { {\ctxwk{\jcomp{}{e}{P}}{e}}
-      {\ctxwk{\ctxext{A}{{P}{\jcomp{}{e}{P}}}}{\jcomp{}{e}{P}}}}
-  \\
-& \jdeq
-  \subst
-    {\unfold{\jvcomp{\jcomp{}{e}{P}}{e}{\idtm{\jcomp{}{e}{P}}}}}
-    {\ctxwk{\ctxext{A}{{P}{\jcomp{}{e}{P}}}}{\jcomp{}{e}{P}}}
-  \\
-& \jdeq
-  \subst
-    {\jvcomp{}{e}{\idtm{\jcomp{}{e}{P}}}}
-    {\ctxwk{\ctxext{A}{{P}{\jcomp{}{e}{P}}}}{\jcomp{}{e}{P}}}
-  \\
-& \jdeq
-  \jcomp{}{\jvcomp{}{e}{\idtm{\jcomp{}{e}{P}}}}{\jcomp{}{e}{P}}
-  \\
-& \jdeq
-  \jcomp{}{\jcomp{}{\jvcomp{}{e}{\idtm{\jcomp{}{e}{P}}}}{e}}{P}
+        {\jcomp{}{\epsilon_0}{P}}{\subst{\epsilon_0}
+        {\ctxwk{\ctxext{A}{P}}{\jcomp{}{\epsilon_0}{P}}}}}
+  \\
+& \jdeq
+  \subst
+    {\idtm{\jcomp{}{\epsilon_0}{P}}}
+    { {\ctxwk{\jcomp{}{\epsilon_0}{P}}{\epsilon_0}}
+      {\ctxwk{\jcomp{}{\epsilon_0}{P}}{{\ctxext{A}{P}}{\jcomp{}{\epsilon_0}{P}}}}}
+  \\
+& \jdeq
+  \subst
+    {\idtm{\jcomp{}{\epsilon_0}{P}}}
+    { {\ctxwk{\jcomp{}{\epsilon_0}{P}}{\epsilon_0}}
+      {\ctxwk{\ctxext{{A}{P}}{\jcomp{}{\epsilon_0}{P}}}{\jcomp{}{\epsilon_0}{P}}}}
+  \\
+& \jdeq
+  \subst
+    {\idtm{\jcomp{}{\epsilon_0}{P}}}
+    { {\ctxwk{\jcomp{}{\epsilon_0}{P}}{\epsilon_0}}
+      {\ctxwk{\ctxext{A}{{P}{\jcomp{}{\epsilon_0}{P}}}}{\jcomp{}{\epsilon_0}{P}}}}
+  \\
+& \jdeq
+  \subst
+    {\unfold{\jvcomp{\jcomp{}{\epsilon_0}{P}}{\epsilon_0}{\idtm{\jcomp{}{\epsilon_0}{P}}}}}
+    {\ctxwk{\ctxext{A}{{P}{\jcomp{}{\epsilon_0}{P}}}}{\jcomp{}{\epsilon_0}{P}}}
+  \\
+& \jdeq
+  \subst
+    {\jvcomp{}{\epsilon_0}{\idtm{\jcomp{}{\epsilon_0}{P}}}}
+    {\ctxwk{\ctxext{A}{{P}{\jcomp{}{\epsilon_0}{P}}}}{\jcomp{}{\epsilon_0}{P}}}
+  \\
+& \jdeq
+  \jcomp{}{\jvcomp{}{\epsilon_0}{\idtm{\jcomp{}{\epsilon_0}{P}}}}{\jcomp{}{\epsilon_0}{P}}
+  \\
+& \jdeq
+  \jcomp{}{\jcomp{}{\jvcomp{}{\epsilon_0}{\idtm{\jcomp{}{\epsilon_0}{P}}}}{\epsilon_0}}{P}
   \\
@@
   \jcomp{}{
-    \jcomp{}{\jvcomp{}{\idtm{A}}{f}}{e}}{P}
-  \\
-& \jdeq
-  \jcomp{}{\jvcomp{}{\idtm{A}}{f}}{
-    \jcomp{}{e}{P}}
-  \\
-& \jdeq
-  \subst
-    {\tmext{\ctxwk{\ctxext{P}{\jcomp{}{e}{P}}}{\idtm{A}}}{f}}
-    {\ctxwk{\ctxext{A}{{P}{\jcomp{}{e}{P}}}}{\jcomp{}{e}{P}}}
-  \\
-& \jdeq
-  \subst
-    {f}
-    { {\ctxwk{\ctxext{P}{\jcomp{}{e}{P}}}{\idtm{A}}}
-      {\ctxwk{\ctxext{A}{{P}{\jcomp{}{e}{P}}}}{\jcomp{}{e}{P}}}
-      }
-  \\
-& \jdeq
-  \subst
-    {f}
-    { {\cprojfstf{A}{\ctxext{P}{\jcomp{}{e}{P}}}}
-      {\ctxwk{\ctxext{A}{{P}{\jcomp{}{e}{P}}}}{\jcomp{}{e}{P}}}
-      }
-  \\
-& \jdeq
-  \subst
-    {f}
-    {\jcomp{}{\cprojfstf{A}{\ctxext{P}{\jcomp{}{e}{P}}}}{\jcomp{}{e}{P}}}
-  \\
-& \jdeq
-  \subst
-    {f}
-    {\ctxwk{\ctxext{P}{\jcomp{}{e}{P}}}{\jcomp{}{e}{P}}}
-  \\
-& \jdeq
-  \jcomp{}{f}{\jcomp{}{e}{P}}.
+    \jcomp{}{\jvcomp{}{\idtm{A}}{\epsilon_1}}{\epsilon_0}}{P}
+  \\
+& \jdeq
+  \jcomp{}{\jvcomp{}{\idtm{A}}{\epsilon_1}}{
+    \jcomp{}{\epsilon_0}{P}}
+  \\
+& \jdeq
+  \subst
+    {\tmext{\ctxwk{\ctxext{P}{\jcomp{}{\epsilon_0}{P}}}{\idtm{A}}}{\epsilon_1}}
+    {\ctxwk{\ctxext{A}{{P}{\jcomp{}{\epsilon_0}{P}}}}{\jcomp{}{\epsilon_0}{P}}}
+  \\
+& \jdeq
+  \subst
+    {\epsilon_1}
+    { {\ctxwk{\ctxext{P}{\jcomp{}{\epsilon_0}{P}}}{\idtm{A}}}
+      {\ctxwk{\ctxext{A}{{P}{\jcomp{}{\epsilon_0}{P}}}}{\jcomp{}{\epsilon_0}{P}}}
+      }
+  \\
+& \jdeq
+  \subst
+    {\epsilon_1}
+    { {\cprojfstf{A}{\ctxext{P}{\jcomp{}{\epsilon_0}{P}}}}
+      {\ctxwk{\ctxext{A}{{P}{\jcomp{}{\epsilon_0}{P}}}}{\jcomp{}{\epsilon_0}{P}}}
+      }
+  \\
+& \jdeq
+  \subst
+    {\epsilon_1}
+    {\jcomp{}{\cprojfstf{A}{\ctxext{P}{\jcomp{}{\epsilon_0}{P}}}}{\jcomp{}{\epsilon_0}{P}}}
+  \\
+& \jdeq
+  \subst
+    {\epsilon_1}
+    {\ctxwk{\ctxext{P}{\jcomp{}{\epsilon_0}{P}}}{\jcomp{}{\epsilon_0}{P}}}
+  \\
+& \jdeq
+  \jcomp{}{\epsilon_1}{\jcomp{}{\epsilon_0}{P}}.
   \qedhere
@@
   {{\Gamma}{A}}
-  {{P}{{\jcomp{}{e}{P}}{\jcomp{}{f}{{}{e}{P}}}}}
-  {{P}{\jcomp{}{e}{P}}}
-  {\jvcomp{}{f}{\idtm{\jcomp{}{f}{{}{e}{P}}}}}.
+  {{P}{{\jcomp{}{\epsilon_0}{P}}{\jcomp{}{\epsilon_1}{{}{\epsilon_0}{P}}}}}
+  {{P}{\jcomp{}{\epsilon_0}{P}}}
+  {\jvcomp{}{\epsilon_1}{\idtm{\jcomp{}{\epsilon_1}{{}{\epsilon_0}{P}}}}}.
 \end{equation*}
@@
 \begin{tikzcd}[column sep=huge]
-\ctxext{P}{{\jcomp{}{e}{P}}{\jcomp{}{e}{{}{e}{P}}}} 
-  \ar{r}{\jvcomp{}{\idtm{P}}{\jcomp{}{e}{f}}}
+\ctxext{P}{{\jcomp{}{\epsilon_0}{P}}{\jcomp{}{\epsilon_0}{{}{\epsilon_0}{P}}}} 
+  \ar{r}{\jvcomp{}{\idtm{P}}{\jcomp{}{\epsilon_0}{\epsilon_1}}}
   \ar{d}[swap]{
-    \jvcomp{}{f}{\idtm{\jcomp{}{f}{{}{e}{P}}}}
+    \jvcomp{}{\epsilon_1}{\idtm{\jcomp{}{\epsilon_1}{{}{\epsilon_0}{P}}}}
 %    \tmext
@@
     }
-& \ctxext{P}{\jcomp{}{e}{P}} \ar{d}{f}\\
-\ctxext{P}{\jcomp{}{e}{P}} \ar{r}[swap]{f} & P
+& \ctxext{P}{\jcomp{}{\epsilon_0}{P}} \ar{d}{\epsilon_1}\\
+\ctxext{P}{\jcomp{}{\epsilon_0}{P}} \ar{r}[swap]{\epsilon_1} & P
 \end{tikzcd}
@@
 \begin{defn}
-An \emph{extension algebra in context $\Gamma$} is a quadruple $(A,P,e,f)$
+An \emph{extension algebra in context $\Gamma$} is a quadruple $(A,P,\epsilon_0,\epsilon_1)$
 consisting of a family $A$ over context $\Gamma$, a family $P$ over the context
-$\ctxext{\Gamma}{A}$, a morphism $e$ from $\ctxext{A}{P}$ to $A$ in context
-$\Gamma$ and a morphism $f$ from $\ctxext{P}{\jcomp{}{e}{P}}$ to $P$ in context
+$\ctxext{\Gamma}{A}$, a morphism $\epsilon_0$ from $\ctxext{A}{P}$ to $A$ in context
+$\Gamma$ and a morphism $\epsilon_1$ from $\ctxext{P}{\jcomp{}{\epsilon_0}{P}}$ to $P$ in context
 $\ctxext{\Gamma}{A}$, satisfying the judgmental equalities of
@@
 \begin{lem}
-Let $(A,P,e,f)$ be an extension algebra in context $\Gamma$ and let
-$y_0:P$, $y_1:\subst{y_0}{\jcomp{}{e}{P}}$ and 
-$y_2:\subst{y_1}{{y_0}{\jcomp{}{e}{\jcomp{}{e}{P}}}}$. Then we have the
+Let $(A,P,\epsilon_0,\epsilon_1)$ be an extension algebra in context $\Gamma$ and let
+$y_0:P$, $y_1:\subst{y_0}{\jcomp{}{\epsilon_0}{P}}$ and 
+$y_2:\subst{y_1}{{y_0}{\jcomp{}{\epsilon_0}{\jcomp{}{\epsilon_0}{P}}}}$. Then we have the
 judgmental equality
 \begin{equation*}
-\subst{{y_2}{{y_1}{\jcomp{}{e}{f}}}}{{y_0}{f}}
+\subst{{y_2}{{y_1}{\jcomp{}{\epsilon_0}{\epsilon_1}}}}{{y_0}{\epsilon_1}}
   \jdeq
-  \subst{y_2}{{{y_1}{{y_0}{f}}}{f}}
+  \subst{y_2}{{{y_1}{{y_0}{\epsilon_1}}}{\epsilon_1}}
 \end{equation*}
@@
 \begin{align*}
-\subst{{y_2}{{y_1}{\jcomp{}{e}{f}}}}{{y_0}{f}}
-& \jdeq
-  \subst{\tmext{y_0}{\subst{y_2}{{y_1}{\jcomp{}{e}{f}}}}}{f}
-  \\
-& \jdeq
-  \subst{{y_2}{{y_1}{{y_0}{\jvcomp{}{\idtm{P}}{\jcomp{}{e}{f}}}}}}{f}
+\subst{{y_2}{{y_1}{\jcomp{}{\epsilon_0}{\epsilon_1}}}}{{y_0}{\epsilon_1}}
+& \jdeq
+  \subst{\tmext{y_0}{\subst{y_2}{{y_1}{\jcomp{}{\epsilon_0}{\epsilon_1}}}}}{\epsilon_1}
+  \\
+& \jdeq
+  \subst{{y_2}{{y_1}{{y_0}{\jvcomp{}{\idtm{P}}{\jcomp{}{\epsilon_0}{\epsilon_1}}}}}}{\epsilon_1}
   \\
@@
       { {y_0}
-        {\jcomp{}{\jvcomp{}{\idtm{P}}{\jcomp{}{e}{f}}}{f}}
+        {\jcomp{}{\jvcomp{}{\idtm{P}}{\jcomp{}{\epsilon_0}{\epsilon_1}}}{\epsilon_1}}
         }
@@
       { {y_0}
-        {\jcomp{}{\jvcomp{}{f}{\idtm{\jcomp{}{f}{{}{e}{P}}}}}{f}}
-        }
-      }
-  \\
-& \jdeq
-  \subst{{y_2}{{y_1}{{y_0}{\jvcomp{}{f}{\idtm{\jcomp{}{f}{{}{e}{P}}}}}}}}{f}
-  \\
-& \jdeq
-  \subst{\tmext{\subst{y_1}{{y_0}{f}}}{y_2}}{f}
-  \\
-& \jdeq
-  \subst{y_2}{{{y_1}{{y_0}{f}}}{f}}.\qedhere
-\end{align*}
-\end{proof}
-
+        {\jcomp{}{\jvcomp{}{\epsilon_1}{\idtm{\jcomp{}{\epsilon_1}{{}{\epsilon_0}{P}}}}}{\epsilon_1}}
+        }
+      }
+  \\
+& \jdeq
+  \subst{{y_2}{{y_1}{{y_0}{\jvcomp{}{\epsilon_1}{\idtm{\jcomp{}{\epsilon_1}{{}{\epsilon_0}{P}}}}}}}}{\epsilon_1}
+  \\
+& \jdeq
+  \subst{\tmext{\subst{y_1}{{y_0}{\epsilon_1}}}{y_2}}{\epsilon_1}
+  \\
+& \jdeq
+  \subst{y_2}{{{y_1}{{y_0}{\epsilon_1}}}{\epsilon_1}}.\qedhere
+\end{align*}
+\end{proof}
+
+\begin{comment}
 Extension algebras don't come in isolation. There are also extension algebra
@@
 \begin{align*}
-\jhom*{{{\Gamma}{A}}{P}}{\ctxext{\ctxwk{P}{B}}{Q}}{\jcomp{}{e}{B}}{g}\\
+\jhom*{{{\Gamma}{A}}{P}}{\ctxext{\ctxwk{P}{B}}{Q}}{\jcomp{}{\epsilon_0}{B}}{g}\\
 \jhom*{{{{\Gamma}{A}}{P}}{\ctxwk{P}{B}}}{\ctxext{Q}{\jcomp{}{g}{Q}}}{Q}{h}.
 \end{align*}
-The quadruple $(\jcomp{}{e}{B},Q,g,h)$ is required to be an extension algebra
+The quadruple $(\jcomp{}{\epsilon_0}{B},Q,g,h)$ is required to be an extension algebra
 in context $\ctxext{{\Gamma}{A}}{P}$.
@@
 \jterm*{{\Gamma}{A}}{B}{x}\\
-\jterm*{{{\Gamma}{A}}{P}}{\subst{\jcomp{}{e}{x}}{Q}}{y}
+\jterm*{{{\Gamma}{A}}{P}}{\subst{\jcomp{}{\epsilon_0}{x}}{Q}}{y}
 \end{align*}
@@
 \begin{tikzcd}
-\ctxext{\jcomp{}{e}{B}}{Q} 
+\ctxext{\jcomp{}{\epsilon_0}{B}}{Q} 
   \ar{r}{g} 
@@
 \ctxext{A}{P} 
-  \ar[shift right=.7ex,dotted]{u}[swap]{\tmext{\jcomp{}{e}{x}}{y}}
+  \ar[shift right=.7ex,dotted]{u}[swap]{\tmext{\jcomp{}{\epsilon_0}{x}}{y}}
   \ar{r}{e}
@@
   \\
-\jcomp{}{f}{\jcomp{}{e}{B}}
-  \ar{r}{\idtm{\jcomp{}{f}{\jcomp{}{e}{B}}}}
+\jcomp{}{f}{\jcomp{}{\epsilon_0}{B}}
+  \ar{r}{\idtm{\jcomp{}{f}{\jcomp{}{\epsilon_0}{B}}}}
   \ar[shift right=.7ex,fib]{d}
   \ar[shift right=.7ex,dotted,mapsto]{u}[swap]{\jcomp{}{f}{y}}
-& \jcomp{}{e}{B}
+& \jcomp{}{\epsilon_0}{B}
   \ar[shift right=.7ex,fib]{d}
@@
   \\
-\ctxext{P}{\jcomp{}{e}{P}}
+\ctxext{P}{\jcomp{}{\epsilon_0}{P}}
   \ar{r}[swap]{f}
-  \ar[shift right=.7ex,dotted]{u}[swap]{\jcomp{}{f}{\jcomp{}{e}{x}}}
+  \ar[shift right=.7ex,dotted]{u}[swap]{\jcomp{}{f}{\jcomp{}{\epsilon_0}{x}}}
 & P
-  \ar[shift right=.7ex,dotted]{u}[swap]{\jcomp{}{e}{x}}
+  \ar[shift right=.7ex,dotted]{u}[swap]{\jcomp{}{\epsilon_0}{x}}
 \end{tikzcd}
@@
 Note that $\ctxwk{\ctxext{A}{P}}{Q}$ is a family over $\ctxwk{\ctxext{A}{P}}{B}$,
-whereas it should be a family over $\jcomp{}{e}{\ctxwk{A}{B}}$. These are the
+whereas it should be a family over $\jcomp{}{\epsilon_0}{\ctxwk{A}{B}}$. These are the
 same by \autoref{lem:prehom}.
@@
 \jterm*{{\Gamma}{A}}{\ctxwk{A}{B}}{x}\\
-\jterm*{{{\Gamma}{A}}{P}}{\subst{\jcomp{}{e}{x}}{\ctxwk{\ctxext{A}{P}}{Q}}}{y}.
+\jterm*{{{\Gamma}{A}}{P}}{\subst{\jcomp{}{\epsilon_0}{x}}{\ctxwk{\ctxext{A}{P}}{Q}}}{y}.
 \end{align*}
 Thus, $x$ is a context morphism from $A$ to $B$ and $y$ is nothing but a term
-of $\jcomp{}{\jcomp{}{e}{x}}{Q}$. For $x$, we see that the diagram
+of $\jcomp{}{\jcomp{}{\epsilon_0}{x}}{Q}$. For $x$, we see that the diagram
 \begin{equation*}
@@
 \end{rmk}
+\end{comment}
 
@@
 Let $P$ be a family over the extended context $\ctxext{\Gamma}{A}$, let
-$\jterm{\Gamma}{A}{x}$ and $\jterm{{\Gamma}{A}}{P}{f}$ be terms. Then the
-quadruple $(A,P,x,f)$ is said to be an \emph{empty algebra in context $\Gamma$}
+$\jterm{\Gamma}{A}{\phi_0}$ and $\jterm{{\Gamma}{A}}{P}{\phi_1}$ be terms. Then the
+quadruple $(A,P,\phi_0,\phi_1)$ is said to be an \emph{empty algebra in context $\Gamma$}
 if the following judgmental equalities hold:
 \begin{align*}
-\jfameq*{\Gamma}{\subst{x}{P}}{A}\\
-\jtermeq*{\Gamma}{A}{\subst{x}{f}}{x}.
-\end{align*}
-\end{defn}
-
-\begin{defn}
-Let $(A_0,P_0,x_0,f_0)$ be an empty algebra in context $\Gamma$ and consider
-\begin{enumerate}
-\item a family $\jfam{{\Gamma}{A_0}}{A_1}$.
-\item a family $\jfam{{{{\Gamma}{A_0}}{P_0}}{\ctxwk{P_0}{A_1}}}{P_1}$.
-\end{enumerate}
+\jfameq*{\Gamma}{\subst{\phi_0}{P}}{A}\\
+\jtermeq*{\Gamma}{A}{\subst{\phi_0}{\phi_1}}{\phi_0}.
+\end{align*}
 \end{defn}

### 2014-05-14 article-models-ttderived.tex
working on algebras

@@
 tells what happens when we weaken a composition, which is a term of 
-$\ctxwk{A}{C}$, by a family $P$ over $\ctxext{\Gamma}{P}$. We have the following
+$\ctxwk{A}{C}$, by a family $P$ over $\ctxext{\Gamma}{P}$. We also have the following
 inference rules expressing that compositions by or with constant morphisms are
@@
 
-Likewise, we can substitute a term $\jterm{\Gamma}{A}{x}$ in a composition
+Likewise, we can substitute a term $\jterm{\Gamma}{A}{x}$ in a composed morphism
 $\jhom{\Gamma}{A}{C}{\jcomp{A}{f}{g}}$. What we get is the \emph{value}
-$\subst{{x}{f}}{g}$. There is a second case where we can substitute: when
+$\subst{{x}{f}}{g}$. This is the content of one part of the following lemma.
+
+There is a second, related case where we can substitute: when
 we consider morphisms $\jhom{{\Gamma}{A}}{P}{Q}{f}$ and 
-$\jhom{{\Gamma}{A}}{Q}{R}{g}$ we can restrict them to the fibers. We do this
-in the following lemma:
-
-\begin{lem}\label{lem:jcomp-fiber}
-The following inference rules are valid:
-\begin{align*}
-& \inference
+$\jhom{{\Gamma}{A}}{Q}{R}{g}$ we can restrict them to the fibers, obtaining
+the inference rule
+\begin{equation*}
+\inference
   { \jhom{{\Gamma}{A}}{P}{Q}{f}
@@
     }
-  \\
+\end{equation*}
+In the following lemma we prove the validity of a more general version of this
+inference rule.
+
+\begin{lem}\label{lem:jcomp-fiber}
+The following inference rules are valid:
+\begin{align*}
+& \inference
+  { \jhom{{\Gamma}{A}}{P}{Q}{f}
+    \jfam{{{\Gamma}{A}}{Q}}{R}
+    \jterm{\Gamma}{A}{x}
+    }
+  { \jfameq
+      {{\Gamma}{\subst{x}{P}}}
+      {\subst{x}{\jcomp{P}{f}{R}}}
+      {\jcomp{\subst{x}{P}}{\subst{x}{f}}{\subst{x}{R}}}
+    }
+  \\
+& \inference
+  { \jhom{{\Gamma}{A}}{P}{Q}{f}
+    \jfam{{{{\Gamma}{A}}{Q}}{R}}{S}
+    \jterm{\Gamma}{A}{x}
+    }
+  { \jfameq
+      {{{\Gamma}{\subst{x}{P}}}{\subst{x}{\jcomp{P}{f}{R}}}}
+      {\subst{x}{\jcomp{P}{f}{S}}}
+      {\jcomp{\subst{x}{P}}{\subst{x}{f}}{\subst{x}{S}}}
+    }
+  \\
+& \inference
+  { \jhom{{\Gamma}{A}}{P}{Q}{f}
+    \jterm{{{{\Gamma}{A}}{Q}}{R}}{S}{k}
+    \jterm{\Gamma}{A}{x}
+    }
+  { \jtermeq
+      {{{\Gamma}{\subst{x}{P}}}{\subst{x}{\jcomp{P}{f}{R}}}}
+      {\subst{x}{\jcomp{P}{f}{S}}}
+      {\subst{x}{\jcomp{P}{f}{k}}}
+      {\jcomp{\subst{x}{P}}{\subst{x}{f}}{\subst{x}{k}}}
+    }
+\end{align*}
+Also, the following inference rule computing the value at $x$ of a composed
+morphism is valid:
+\begin{align*}
 & \inference
@@
 \begin{proof}
-Let $\jhom{{\Gamma}{A}}{P}{Q}{f}$, $\jhom{{\Gamma}{A}}{Q}{R}{g}$ and 
+Of the first three inference rules, we only prove the first.
+Let $\jhom{{\Gamma}{A}}{P}{Q}{f}$, $\jfam{{{\Gamma}{A}}{Q}}{R}$ and 
 $\jterm{\Gamma}{A}{x}$.
 Then we have the judgmental equalities\begin{align*}
-\subst{x}{\jcomp{A}{f}{g}}
-& \jdeq 
-  \subst{x}{{f}{\ctxwk{P}{g}}}
+\subst{x}{\jcomp{A}{f}{R}}
+& \jdeq 
+  \subst{x}{{f}{\ctxwk{P}{R}}}
   \tag{by definition}
@@
 & \jdeq 
-  \subst{{x}{f}}{{x}{\ctxwk{P}{g}}}
-  \tag{by \autoref{comp-ss-t}}
-  \\
-& \jdeq 
-  \subst{{x}{f}}{\ctxwk{\subst{x}{P}}{\subst{x}{g}}}
-  \tag{by \autoref{comp-sw-t}}
-  \\
-& \jdeq 
-  \jcomp{\subst{x}{P}}{\subst{x}{f}}{\subst{x}{g}}.
+  \subst{{x}{f}}{{x}{\ctxwk{P}{R}}}
+  \tag{by \autoref{comp-ss-f}}
+  \\
+& \jdeq 
+  \subst{{x}{f}}{\ctxwk{\subst{x}{P}}{\subst{x}{R}}}
+  \tag{by \autoref{comp-sw-f}}
+  \\
+& \jdeq 
+  \jcomp{\subst{x}{P}}{\subst{x}{f}}{\subst{x}{R}}.
   \tag{by definition}
@@
 if the following judgmental equalities hold:
-\begin{align*}
-\jfameq*{\Gamma}{\subst{\phi_0}{P}}{A}\\
+\begin{align}
+\jfameq*{\Gamma}{\subst{\phi_0}{P}}{A}
+  \label{empalg-eq1}
+  \\
 \jtermeq*{\Gamma}{A}{\subst{\phi_0}{\phi_1}}{\phi_0}.
-\end{align*}
+  \label{empalg-eq2}
+\end{align}
 \end{defn}
 
+Thus, the empty algebras are the kind of algebras that require that families
+are compatible with contexts, just as our motivation in \autoref{empty}. We
+now combine the notion of extension algebras and empty algebras.
+
+An extension-empty algebra in context $\Gamma$ is going to be a sextuple
+$(A,P,\epsilon_0,\epsilon_1,\phi_0,\phi_1)$ for which 
+the quadruple $(A,P,\epsilon_0,\epsilon_1)$ is an extension algebra in context 
+$\Gamma$, the quadruple $(A,P,\phi_0,\phi_1)$ is an empty algebra in context
+$\Gamma$, satisfying additional judgmental equalities expressing the 
+compatibility of $\epsilon_0$ and $\epsilon_1$ with $\phi_0$ and $\phi_1$.
+There will be four such judgmental equalities.
+
+We can immediately state the first two:
+\begin{align}
+\jtermeq*{{\Gamma}{A}}{\ctxwk{A}{A}}{\subst{\phi_0}{\epsilon_0}}{\idtm{A}}
+  \label{extempalg-eq1}
+  \\
+\jtermeq*{{\Gamma}{A}}{\ctxwk{A}{A}}{\subst{\phi_1}{\epsilon_0}}{\idtm{A}}
+  \label{extempalg-eq2}
+\end{align}
+To see what $\subst{\phi_0}{\epsilon_1}$ can be, we must know its type first.
+It is a morphism from $\subst{\phi_0}{\ctxext{P}{\jcomp{}{\epsilon_0}{P}}}$ to
+$\subst{\phi_0}{P}$. We already know that $\subst{\phi_0}{P}\jdeq A$ by
+\autoref{empalg-eq1} and to compute $\subst{\phi_0}{\jcomp{}{\epsilon_0}{P}}$
+we use the following lemma.
+
+\begin{lem}
+Consider a sextuple $(A,P,\epsilon_0,\epsilon_1,\phi_0,\phi_1)$ where
+the quadruple $(A,P,\epsilon_0,\epsilon_1)$ is an extension algebra in context 
+$\Gamma$, the quadruple $(A,P,\phi_0,\phi_1)$ is an empty algebra in context
+$\Gamma$, satisfying the judgmental equality \autoref{extempalg-eq1}. Then
+the following inference rules are valid:
+\begin{align*}
+& \inference
+  { \jfam{{\Gamma}{A}}{Q}
+    }
+  { \jfameq
+      {{\Gamma}{A}}
+      {\subst{\phi_0}{\jcomp{}{\epsilon_0}{Q}}}
+      {Q}
+    }
+  \\
+& \inference
+  { \jfam{{{\Gamma}{A}}{Q}}{R}
+    }
+  { \jfameq
+      {{{\Gamma}{A}}{\subst{\phi_0}{Q}}}
+      {\subst{\phi_0}{\jcomp{}{\epsilon_0}{R}}}
+      {R}
+    }
+  \\
+& \inference
+  { \jterm{{{\Gamma}{A}}{Q}}{R}{h}
+    }
+  { \jtermeq
+      {{{\Gamma}{A}}{\subst{\phi_0}{Q}}}
+      {\subst{\phi_0}{R}}
+      {\subst{\phi_0}{\jcomp{}{\epsilon_0}{h}}}
+      {h}
+    }
+\end{align*}
+\end{lem}
+
+\begin{proof}
+We only prove the first inference rule.
+Let $Q$ be a family over $\ctxext{\Gamma}{A}$. Then we have the judgmental
+equalities
+\begin{align*}
+\subst{\phi_0}{\jcomp{}{\epsilon_0}{Q}}
+& \jdeq
+  \subst{\phi_0}{{\epsilon_0}{\ctxwk{\ctxext{A}{P}}{Q}}}
+  \\
+& \jdeq
+  \subst{{\phi_0}{\epsilon_0}}{{\phi_0}{\ctxwk{\ctxext{A}{P}}{Q}}}
+  \\
+& \jdeq
+  \subst{\idtm{A}}{{\phi_0}{\ctxwk{\ctxext{A}{P}}{Q}}}
+  \\
+& \jdeq
+  \subst{\idtm{A}}{{\phi_0}{\ctxwk{P}{{A}{Q}}}}
+  \\
+& \jdeq
+  \subst{\idtm{A}}{\ctxwk{\subst{\phi_0}{P}}{\subst{\phi_0}{\ctxwk{A}{Q}}}}
+  \\
+& \jdeq
+  \subst{\idtm{A}}{\ctxwk{\subst{\phi_0}{P}}{Q}}
+  \\
+& \jdeq
+  \subst{\idtm{A}}{\ctxwk{A}{Q}}
+  \\
+& \jdeq
+  Q.
+\end{align*}
+\end{proof}
+
+\begin{defn}
+An \emph{extension-empty algebra in context $\Gamma$} 
+is a sextuple $(A,P,\epsilon_0,\epsilon_1,\phi_0,\phi_1)$ for which 
+the quadruple $(A,P,\epsilon_0,\epsilon_1)$ is an extension algebra in context 
+$\Gamma$, the quadruple $(A,P,\phi_0,\phi_1)$ is an empty algebra in context
+$\Gamma$, and the following additional judgmental equalities hold:
+\begin{align*}
+\jtermeq*{{\Gamma}{A}}{\ctxwk{A}{A}}{\subst{\phi_0}{\epsilon_0}}{\idtm{A}}
+  \\
+\jtermeq*{{\Gamma}{A}}{\ctxwk{A}{A}}{\subst{\phi_1}{\epsilon_0}}{\idtm{A}}
+  \\
+\jhomeq*{\Gamma}{{A}{P}}{A}{\subst{\phi_0}{\epsilon_1}}{\epsilon_0}
+  \\
+\jtermeq*{{{\Gamma}{A}}{P}}{\ctxwk{P}{P}}{\subst{\phi_1}{\epsilon_1}}{\idtm{P}}
+\end{align*}
+To see that the last two of these equalities make sense, one has to
+use the first two.
+\end{defn}
+
 \subsection{Extension-weakening algebras}
+The notion of (extension-)weakening algebra is dependent on the notion of extension algebra.
+Although it is strictly speaking not dependent on the notion of empty-algebra,
+we shall only formulate a weakening algebras in the setting of an
+extension-empty algebras. When one wants to have a weakening operation which
+also acts on the level of contexts in an extension algebra without empty
+context, extra work has to be done to introduce these separately.
+
+\begin{defn}
+An extension-weakening algebra in context $\Gamma$ is an octuple
+\begin{equation*}
+(A,P,\epsilon_0,\epsilon_1,\phi_0,\phi_1,\omega_0,\omega_1)
+\end{equation*}
+where $(A,P,\epsilon_0,\epsilon_1,\phi_0,\phi_1)$ is an extension-empty algebra
+in context $\Gamma$ and where
+\begin{align*}
+\jhom*
+  {{{\Gamma}{A}}{P}}
+  {\ctxwk{P}{P}}
+  {\jcomp{}{\epsilon_0}{P}}
+  {\omega_0}
+  \\
+\jfhom*
+  {{{\Gamma}{A}}{P}}
+  {\ctxwk{P}{P}}
+  {\jcomp{}{\epsilon_0}{P}}
+  {\omega_0}
+  {\ctxwk{P}{\jcomp{}{\epsilon_0}{P}}}
+  {\jcomp{}{\epsilon_0}{\jcomp{}{\epsilon_0}{P}}}
+  {\omega_1}
+\end{align*}
+satisfying the judgmental equalities
+\end{defn}
 

### 2014-05-15 article-models-ttderived.tex
section on extension-empty algebras finished

@@
 \begin{lem}
-Consider a sextuple $(A,P,\epsilon_0,\epsilon_1,\phi_0,\phi_1)$ where
-the quadruple $(A,P,\epsilon_0,\epsilon_1)$ is an extension algebra in context 
-$\Gamma$, the quadruple $(A,P,\phi_0,\phi_1)$ is an empty algebra in context
-$\Gamma$, satisfying the judgmental equality \autoref{extempalg-eq1}. Then
-the following inference rules are valid:
+Consider an empty algebra $(A,P,\phi_0,\phi_1)$ in context $\Gamma$
+and a morphism $\jhom{\Gamma}{{A}{P}}{B}{f}$.
+Then $\subst{\phi_i}{f}$ is a morphism from $A$ to $B$ in context $\Gamma$ and
+the following inference rules are valid for $i$ being $0$ or $1$:
 \begin{align*}
@@
       {{\Gamma}{A}}
-      {\subst{\phi_0}{\jcomp{}{\epsilon_0}{Q}}}
-      {Q}
+      {\subst{\phi_i}{\jcomp{}{f}{Q}}}
+      {\jcomp{}{\subst{\phi_i}{f}}{Q}}
     }
@@
   { \jfameq
-      {{{\Gamma}{A}}{\subst{\phi_0}{Q}}}
-      {\subst{\phi_0}{\jcomp{}{\epsilon_0}{R}}}
-      {R}
+      {{{\Gamma}{A}}{\jcomp{}{\subst{\phi_i}{f}}{Q}}}
+      {\subst{\phi_i}{\jcomp{}{f}{R}}}
+      {\jcomp{}{\subst{\phi_i}{f}}{R}}
     }
@@
   { \jtermeq
-      {{{\Gamma}{A}}{\subst{\phi_0}{Q}}}
-      {\subst{\phi_0}{R}}
-      {\subst{\phi_0}{\jcomp{}{\epsilon_0}{h}}}
-      {h}
+      {{{\Gamma}{A}}{\jcomp{}{\subst{\phi_i}{f}}{Q}}}
+      {\jcomp{}{\subst{\phi_i}{f}}{R}}
+      {\subst{\phi_i}{\jcomp{}{f}{h}}}
+      {\jcomp{}{\subst{\phi_i}{f}}{h}}
     }
@@
 \begin{proof}
-We only prove the first inference rule.
-Let $Q$ be a family over $\ctxext{\Gamma}{A}$. Then we have the judgmental
-equalities
-\begin{align*}
-\subst{\phi_0}{\jcomp{}{\epsilon_0}{Q}}
-& \jdeq
-  \subst{\phi_0}{{\epsilon_0}{\ctxwk{\ctxext{A}{P}}{Q}}}
-  \\
-& \jdeq
-  \subst{{\phi_0}{\epsilon_0}}{{\phi_0}{\ctxwk{\ctxext{A}{P}}{Q}}}
-  \\
-& \jdeq
-  \subst{\idtm{A}}{{\phi_0}{\ctxwk{\ctxext{A}{P}}{Q}}}
-  \\
-& \jdeq
-  \subst{\idtm{A}}{{\phi_0}{\ctxwk{P}{{A}{Q}}}}
-  \\
-& \jdeq
-  \subst{\idtm{A}}{\ctxwk{\subst{\phi_0}{P}}{\subst{\phi_0}{\ctxwk{A}{Q}}}}
-  \\
-& \jdeq
-  \subst{\idtm{A}}{\ctxwk{\subst{\phi_0}{P}}{Q}}
-  \\
-& \jdeq
-  \subst{\idtm{A}}{\ctxwk{A}{Q}}
-  \\
-& \jdeq
-  Q.
+We only prove the first inference rule in both cases.
+Let $Q$ be a family over $\ctxext{\Gamma}{A}$. In the case $i=0$
+ we have the judgmental equalities
+\begin{align*}
+\subst{\phi_0}{\jcomp{}{f}{Q}}
+& \jdeq
+  \subst{\phi_0}{{f}{\ctxwk{\ctxext{A}{P}}{Q}}}
+  \\
+& \jdeq
+  \subst{{\phi_0}{f}}{{\phi_0}{\ctxwk{\ctxext{A}{P}}{Q}}}
+  \\
+& \jdeq
+  \subst{{\phi_0}{f}}{{\phi_0}{\ctxwk{P}{{A}{Q}}}}
+  \\
+& \jdeq
+  \subst{{\phi_0}{f}}{\ctxwk{\subst{\phi_0}{P}}{\subst{\phi_0}{\ctxwk{A}{Q}}}}
+  \\
+& \jdeq
+  \subst{{\phi_0}{f}}{\ctxwk{\subst{\phi_0}{P}}{Q}}
+  \\
+& \jdeq
+  \subst{{\phi_0}{f}}{\ctxwk{A}{Q}}
+  \\
+& \jdeq
+  \jcomp{}{\subst{\phi_0}{f}}{Q}.
+\end{align*}
+In the case $i=1$ we have the judgmental equalities
+\begin{align*}
+\subst{\phi_1}{\jcomp{}{f}{Q}}
+& \jdeq
+  \subst{\phi_1}{{f}{\ctxwk{\ctxext{A}{P}}{Q}}}
+  \\
+& \jdeq
+  \subst{{\phi_1}{f}}{{\phi_1}{\ctxwk{\ctxext{A}{P}}{Q}}}
+  \\
+& \jdeq
+  \subst{{\phi_1}{f}}{{\phi_1}{\ctxwk{P}{{A}{Q}}}}
+  \\
+& \jdeq
+  \subst{{\phi_1}{f}}{\ctxwk{A}{Q}}
+  \\
+& \jdeq
+  \jcomp{}{\subst{\phi_1}{f}}{Q}.
 \end{align*}
 \end{proof}
+
+As an immediate corollary, if we assume the judgmental equalities
+\autoref{extempalg-eq1,extempalg-eq2} we get that 
+\begin{equation*}
+\jfameq{{\Gamma}{A}}{\subst{\phi_i}{\jcomp{}{\epsilon_0}{P}}}{P}
+\end{equation*}
+and hence that $\subst{\phi_0}{\epsilon_1}$ is a 
+morphism from $\ctxext{A}{P}$ to $A$. Thus, we can require
+\begin{equation}\label{extempalg-eq3}
+\jhomeq{\Gamma}{{A}{P}}{A}{\subst{\phi_0}{\epsilon_1}}{\epsilon_0}.
+\end{equation}
+For the final judgmental equality we need to explain the term
+\begin{equation*}
+\jterm
+  {{{\Gamma}{A}}{\subst{\phi_1}{\jcomp{}{\epsilon_0}{P}}}}
+  {\subst{\phi_1}{\ctxwk{\ctxext{P}{\jcomp{}{\epsilon_0}{P}}}{P}}}
+  {\subst{\phi_1}{\epsilon_1}}.
+\end{equation*}
+We have already established that $\subst{\phi_1}{\jcomp{}{\epsilon_0}{P}}\jdeq
+P$. We also see that 
+\begin{align*}
+\subst{\phi_1}{\ctxwk{\ctxext{P}{\jcomp{}{\epsilon_0}{P}}}{P}}
+& \jdeq
+  \subst{\phi_1}{\ctxwk{\jcomp{}{\epsilon_0}{P}}{{P}{P}}}
+  \\
+& \jdeq
+  \ctxwk{\subst{\phi_1}{\jcomp{}{\epsilon_0}{P}}}{\subst{\phi_1}{\ctxwk{P}{P}}}
+  \\
+& \jdeq
+  \ctxwk{P}{\subst{\phi_1}{\ctxwk{P}{P}}}
+  \\
+& \jdeq
+  \ctxwk{P}{P}
+\end{align*}
+and we will therefore require that
+\begin{equation}\label{extempalg-eq4}
+\jtermeq{{{\Gamma}{A}}{P}}{\ctxwk{P}{P}}{\subst{\phi_1}{\epsilon_1}}{\idtm{P}}.
+\end{equation}
+We bring all this together in the definition of extension-empty algebras:
 
@@
 $\Gamma$, the quadruple $(A,P,\phi_0,\phi_1)$ is an empty algebra in context
-$\Gamma$, and the following additional judgmental equalities hold:
-\begin{align*}
-\jtermeq*{{\Gamma}{A}}{\ctxwk{A}{A}}{\subst{\phi_0}{\epsilon_0}}{\idtm{A}}
-  \\
-\jtermeq*{{\Gamma}{A}}{\ctxwk{A}{A}}{\subst{\phi_1}{\epsilon_0}}{\idtm{A}}
-  \\
-\jhomeq*{\Gamma}{{A}{P}}{A}{\subst{\phi_0}{\epsilon_1}}{\epsilon_0}
-  \\
-\jtermeq*{{{\Gamma}{A}}{P}}{\ctxwk{P}{P}}{\subst{\phi_1}{\epsilon_1}}{\idtm{P}}
-\end{align*}
-To see that the last two of these equalities make sense, one has to
-use the first two.
+$\Gamma$, satisfying the judgmental equalities 
+\autoref{extempalg-eq1,extempalg-eq2,extempalg-eq3,extempalg-eq4}.
 \end{defn}

### 2014-05-15 article-models-ttderived.tex
section on extension-empty algebras tagged

@@
 
-\begin{lem}
+\begin{lem}\label{lem:empalg-mor}
 Consider an empty algebra $(A,P,\phi_0,\phi_1)$ in context $\Gamma$
@@
   \subst{\phi_0}{{f}{\ctxwk{\ctxext{A}{P}}{Q}}}
+  \tag{by definition}
   \\
@@
   \subst{{\phi_0}{f}}{{\phi_0}{\ctxwk{\ctxext{A}{P}}{Q}}}
+  \tag{by \autoref{comp-ss-f}}
   \\
@@
   \subst{{\phi_0}{f}}{{\phi_0}{\ctxwk{P}{{A}{Q}}}}
+  \tag{by \autoref{comp-ew-f}}
   \\
@@
   \subst{{\phi_0}{f}}{\ctxwk{\subst{\phi_0}{P}}{\subst{\phi_0}{\ctxwk{A}{Q}}}}
+  \tag{by \autoref{comp-sw-f}}
   \\
@@
   \subst{{\phi_0}{f}}{\ctxwk{\subst{\phi_0}{P}}{Q}}
+  \tag{by \autoref{cancellation-ws-f}}
   \\
@@
   \subst{{\phi_0}{f}}{\ctxwk{A}{Q}}
+  \tag{by \autoref{empalg-eq1}}
   \\
@@
   \jcomp{}{\subst{\phi_0}{f}}{Q}.
+  \tag{by definition}
 \end{align*}
@@
   \subst{\phi_1}{{f}{\ctxwk{\ctxext{A}{P}}{Q}}}
+  \tag{by definition}
   \\
@@
   \subst{{\phi_1}{f}}{{\phi_1}{\ctxwk{\ctxext{A}{P}}{Q}}}
+  \tag{by \autoref{comp-ss-f}}
   \\
@@
   \subst{{\phi_1}{f}}{{\phi_1}{\ctxwk{P}{{A}{Q}}}}
+  \tag{by \autoref{comp-ew-f}}
   \\
@@
   \subst{{\phi_1}{f}}{\ctxwk{A}{Q}}
+  \tag{by \autoref{cancellation-ws-f}}
   \\
@@
   \jcomp{}{\subst{\phi_1}{f}}{Q}.
+  \tag{by definition}
 \end{align*}
@@
 \autoref{extempalg-eq1,extempalg-eq2} we get that 
-\begin{equation*}
+\begin{equation}\label{cor:empalg-mor}
 \jfameq{{\Gamma}{A}}{\subst{\phi_i}{\jcomp{}{\epsilon_0}{P}}}{P}
-\end{equation*}
+\end{equation}
 and hence that $\subst{\phi_0}{\epsilon_1}$ is a 
@@
   \subst{\phi_1}{\ctxwk{\jcomp{}{\epsilon_0}{P}}{{P}{P}}}
+  \tag{by \autoref{comp-ew-f}}
   \\
@@
   \ctxwk{\subst{\phi_1}{\jcomp{}{\epsilon_0}{P}}}{\subst{\phi_1}{\ctxwk{P}{P}}}
+  \tag{by \autoref{comp-sw-f}}
   \\
@@
   \ctxwk{P}{\subst{\phi_1}{\ctxwk{P}{P}}}
+  \tag{by \autoref{cor:empalg-mor}}
   \\
@@
   \ctxwk{P}{P}
+  \tag{by \autoref{cancellation-ws-f}}
 \end{align*}

### 2014-05-15 article-models-ttderived.tex
developed intuition for internal weakening

@@
 
-\begin{defn}
-An extension-weakening algebra in context $\Gamma$ is an octuple
+An extension-weakening algebra in context $\Gamma$ will be an octuple
 \begin{equation*}
@@
 \end{align*}
-satisfying the judgmental equalities
-\end{defn}
+satisfying nine additional judgmental equalities expressing that $\omega_0$
+and $\omega_1$ are compatible with $\epsilon_1$, $\phi_1$
+and with each other. Our current goal is to figure out what these are.
+
+Before we go into that, we develop a bit of intuition by explaining how
+$\omega_0$ and $\omega_1$ act when applied to a family $x_1:\subst{x_0}{P}$
+over $x_0:A$. Note that
+\begin{equation*}
+\jterm{{\Gamma}{\subst{x_0}{P}}}{\subst{{x_1}{{x_0}{\epsilon_0}}}{P}}{\subst{x_1}{{x_0}{\omega_0}}},
+\end{equation*}
+so $\subst{x_1}{{x_0}{\omega_0}}$ takes families over $x_0$ to families over
+the extended $\subst{x_1}{{x_0}{\epsilon_0}}$.
+
+To compute the type of $\subst{x_2}{{x_1}{{x_0}{\omega_1}}}$ for families
+$x_1,x_2:\subst{x_0}{P}$ over $x_0:A$ we have to do a
+bit more work. Note that
+\begin{align*}
+& \subst{x_2}{{x_1}{{x_0}{\jcomp{}{\omega_0}{\jcomp{}{\epsilon_0}{\jcomp{}{\epsilon_0}{P}}}}}}
+  \\
+& \jdeq
+  \subst{x_2}{{x_1}{{x_0}{\unfold{\jcomp{\ctxwk{P}{P}}{\omega_0}{\jcomp{}{\epsilon_0}{\jcomp{}{\epsilon_0}{P}}}}}}}
+  \tag{by definition}
+  \\
+& \jdeq
+  \subst
+    { {x_2}
+      { {x_1}
+        { {x_0}
+          {\omega_0}
+          }
+        }
+      }
+    { {x_2}
+      { {x_1}
+        { {x_0}
+          {\ctxwk{{P}{P}}{\jcomp{}{\epsilon_0}{\jcomp{}{\epsilon_0}{P}}}}
+          }
+        }
+      }
+  \tag{by \autoref{comp-ss-f}}
+  \\
+& \jdeq
+  \subst
+    { {x_2}
+      { {x_1}
+        { {x_0}
+          {\omega_0}
+          }
+        }
+      }
+    { {x_2}
+      { \ctxwk
+          {\subst{x_1}{\ctxwk{\subst{x_0}{P}}{\subst{x_0}{P}}}}
+          {\subst{x_1}{{x_0}{\jcomp{}{\epsilon_0}{\jcomp{}{\epsilon_0}{P}}}}}
+        }
+      }
+  \tag{by \autoref{comp-sw-f}}
+  \\
+& \jdeq
+  \subst
+    { {x_2}
+      { {x_1}
+        { {x_0}
+          {\omega_0}
+          }
+        }
+      }
+    { {x_1}
+      { {x_0}
+        {\jcomp{}{\epsilon_0}{\jcomp{}{\epsilon_0}{P}}}
+        }
+      }
+  \tag{by \autoref{cancellation-ws-f}}
+  \\
+& \jdeq
+  \subst
+    { {{x_2}{{x_1}{{x_0}{\omega_0}}}}
+      {{{x_1}{{x_0}{\epsilon_0}}}{\epsilon_0}}
+      }
+    { P
+      }.
+  \tag{by \autoref{comp-ss-f}}
+\end{align*}
+Therefore, we see that
+\begin{equation*}
+\jterm
+  {{\Gamma}{\subst{{x_2}{{x_0}{\epsilon_0}}}{P}}}
+  { \subst
+      { {{x_2}{{x_1}{{x_0}{\omega_0}}}}
+        {{{x_1}{{x_0}{\epsilon_0}}}{\epsilon_0}}
+        }
+      { P
+        }
+    }
+  { \subst{x_2}{{x_1}{{x_0}{\omega_1}}}
+    }
+\end{equation*}
+Thus, the term $\subst{x_2}{{x_1}{{x_0}{\omega_1}}}$ takes a family over the
+extended context $\subst{x_2}{{x_0}{\epsilon_0}}$ to a family over the
+weakening $\subst{x_2}{{x_1}{{x_0}{\omega_0}}}$, which is itself a family
+over $\subst{{x_1}{{x_0}{\epsilon_0}}}{\epsilon_0}$. In other words, $\omega_1$
+is precisely the internalization of the action on families of weakening, as
+intended.
+\subsubsection{The compatibility of weakening with extension}
+
+\subsubsection{The compatibility of weakening with the empty context and family}
+The first two judgmental equalities expressing that $\omega_0$ and $\omega_1$
+are compatible with $\phi_1$ are easy to state:
+\begin{align}
+\jhomeq*{{\Gamma}{A}}{P}{P}{\subst{\phi_1}{\omega_0}}{\idtm{P}}\\
+\jhomeq*{{\Gamma}{A}}{P}{P}{\subst{\phi_1}{\omega_1}}{\idtm{P}}
+\end{align}
+
+\subsubsection{The compatibility of weakening with itself}
+
+\subsubsection{The definition of weakening algebras}
 
@@
 terms. They interpret extension, the empty context, weakening, substitution
-and identity terms all at once in a compatible way.
+and identity terms all at once in a compatible way. Besides the compatibility
+properties there will be judgmental equalities analoguous to the cancellation
+properties of \autoref{cancellation-ws,cancellation-i}. Pre-universes are to
+the theory of contexts, families and terms what internal categories to a
+category.

### 2014-05-15 article-models-ttderived.tex
typo

@@
 weakening $\subst{x_2}{{x_1}{{x_0}{\omega_0}}}$, which is itself a family
-over $\subst{{x_1}{{x_0}{\epsilon_0}}}{\epsilon_0}$. In other words, $\omega_1$
+over $\subst{x_1}{{x_0}{\epsilon_0}}$. In other words, $\omega_1$
 is precisely the internalization of the action on families of weakening, as

### 2014-05-15 article-models-ttderived.tex
first rule for compatibility of weakening with extension
split the derived notions section in two

--- 2014-05-15 article-models-ttderived.tex
@@
 \end{comment}
-
-\subsection{Extension algebras}
-In this subsection our goal is to define the notion of extension algebras,
-which are internal versions of the extension operation of the theory of
-contexts, families and terms. In this article, their use will be mainly in
-universes. The theory of extension algebras requires the full power (i.e.~all
-of the ingredients) of the theory of contexts, families and terms in its
-formulation and it is (perhaps surprisingly) quite involved to formulate it.
-
-Let $P$ be a family over an extended context $\ctxext{\Gamma}{A}$. We could
-mimic extension by requiring to have terms
-\begin{align*}
-\jhom*{\Gamma}{{A}{P}}{A}{\epsilon_0}\\
-\jhom*{{\Gamma}{A}}{{P}{\jcomp{}{\epsilon_0}{P}}}{P}{\epsilon_1}.
-\end{align*}
-
-\begin{rmk}
-Instead of looking at $\epsilon_1$ as a context morphism from $\ctxext{P}{\jcomp{}{\epsilon_0}{P}}$
-to $P$, one could also look at $\epsilon_1$ as a morphism \emph{over $\cprojfstf{A}{P}$},
-as indicated in the following diagram:
-\begin{equation*}
-\begin{tikzcd}
-P
-  \ar[fib]{d}
-& \jcomp{}{\epsilon_0}{P}
-  \ar[fib]{d}
-  \ar{l}[swap]{\epsilon_1}
-  \ar{r}
-& P
-  \ar[fib]{d}
-  \\
-A
-& \ctxext{A}{P}
-  \ar{l}{\cprojfstf{A}{P}}
-  \ar{r}[swap]{\epsilon_0}
-& A
-\end{tikzcd}
-\end{equation*}
-This makes it clear that $\epsilon_1$ takes a family over an extended context as an
-argument. The extended context consists of a `base part' and a `family part'. 
-The output of $\epsilon_1$ is a new (extended) family over that base part. Forgetting 
-the family part is what the projection takes care of.
-\end{rmk}
-
-Extension also satisfies the properties explained in \autoref{comp-ee}, so we
-must find the two judgmental equalities for $\epsilon_0$ and $\epsilon_1$ mimicing those. 
-The first of these judgmental equalities is easy to give: it says that the
-following diagram commutes:
-\begin{equation}\label{eq:ealg-eq1}
-\begin{tikzcd}[column sep=huge]
-\ctxext{A}{{P}{\jcomp{}{\epsilon_0}{P}}} 
-  \ar{d}[swap]{\jvcomp{}{\epsilon_0}{\idtm{\jcomp{}{\epsilon_0}{P}}}
-    } 
-  \ar{r}{\jvcomp{}{\idtm{A}}{\epsilon_1}
-    } 
-  & \ctxext{A}{P} \ar{d}{\epsilon_0}\\
-\ctxext{A}{P} \ar{r}[swap]{\epsilon_0} & A
-\end{tikzcd}
-\end{equation}
-To get a feel for this judgmental equality we include the following lemma.
-
-\begin{lem}
-Let $A$, $P$, $\epsilon_0$ and $\epsilon_1$ be as above, satisfying \autoref{eq:ealg-eq1} and
-let $x_0:A$,
-$x_1:\subst{x_0}{P}$ and $x_2:\subst{x_1}{{x_0}{\jcomp{}{\epsilon_0}{P}}}$.
-Then we have the judgmental equality
-\begin{equation*}
-\subst{{x_2}{{x_1}{\epsilon_1}}}{{x_0}{\epsilon_0}}
-\jdeq
-\subst{x_2}{{{x_1}{{x_0}{\epsilon_0}}}{\epsilon_0}}.
-\end{equation*}
-\end{lem}
-
-\begin{proof}
-The proof is a simple computation:
-\begin{align*}
-\subst{{x_2}{{x_1}{\epsilon_1}}}{{x_0}{\epsilon_0}}
-& \jdeq
-  \subst{\tmext{x_0}{\subst{x_2}{{x_1}{\epsilon_1}}}}{\epsilon_0}
-  \\
-& \jdeq 
-  \subst{{x_2}{{x_1}{{x_0}{\jvcomp{}{\idtm{A}}{\epsilon_1}}}}}{\epsilon_0}
-  \\
-& \jdeq
-  \subst{x_2}{{x_1}{{x_0}{\jcomp{}{\jvcomp{}{\idtm{A}}{\epsilon_1}}{\epsilon_0}}}}
-  \\
-& \jdeq
-  \subst{x_2}{{x_1}{{x_0}{\jcomp{}{\jvcomp{}{\epsilon_0}{\idtm{\jcomp{}{\epsilon_0}{P}}}}{\epsilon_0}}}}
-  \\
-& \jdeq 
-  \subst{{x_2}{{x_1}{{x_0}{\jvcomp{}{\epsilon_0}{\idtm{\jcomp{}{\epsilon_0}{P}}}}}}}{\epsilon_0}
-  \\
-& \jdeq
-  \subst{\tmext{\subst{x_1}{{x_0}{\epsilon_0}}}{x_2}}{\epsilon_0}
-  \\
-& \jdeq
-  \subst{x_2}{{{x_1}{{x_0}{\epsilon_0}}}{\epsilon_0}}.\qedhere
-\end{align*}
-\end{proof}
-
-The second of the judgmental equalities is harder to describe, however. We need
-to consider `higher' families, i.e.~families over families over families, and
-thus we need to look at the family $\jcomp{}{\epsilon_0}{\jcomp{}{\epsilon_0}{P}}$ and find the
-two dotted morphisms in the diagram
-\begin{equation*}
-\begin{tikzcd}
-\ctxext{P}{{\jcomp{}{\epsilon_0}{P}}{\jcomp{}{\epsilon_0}{\jcomp{}{\epsilon_0}{P}}}}
-  \ar[densely dotted]{d}
-  \ar[densely dotted]{r}
-& \ctxext{P}{\jcomp{}{\epsilon_0}{P}} \ar{d}{\epsilon_1}\\
-\ctxext{P}{\jcomp{}{\epsilon_0}{P}} \ar{r}{\epsilon_1} & P
-\end{tikzcd}
-\end{equation*}
-The first is easy to find. Note that we have the judgmental equality
-\begin{equation*}
-\ctxext{\jcomp{}{\epsilon_0}{P}}{\jcomp{}{\epsilon_0}{\jcomp{}{\epsilon_0}{P}}}
-  \jdeq
-  \jcomp{}{\epsilon_0}{\ctxext{P}{\jcomp{}{\epsilon_0}{P}}}
-\end{equation*}
-and therefore we may just take the morphism
-\begin{equation*}
-\jhom
-  {{\Gamma}{A}}
-  {{P}{{\jcomp{}{\epsilon_0}{P}}{\jcomp{}{\epsilon_0}{\jcomp{}{\epsilon_0}{P}}}}}
-  {{P}{\jcomp{}{\epsilon_0}{P}}}
-  {\jvcomp{}{\idtm{P}}{\jcomp{}{\epsilon_0}{\epsilon_1}}}.
-\end{equation*}
-For the other morphism we need to look at the family
-$\ctxext{P}{{\jcomp{}{\epsilon_0}{P}}{\jcomp{}{\epsilon_0}{\jcomp{}{\epsilon_0}{P}}}}$ differently. We do
-that in the following lemma.
-
-\begin{lem}
-Suppose we have $A$, $P$, $\epsilon_0$ and $\epsilon_1$ satisfying \autoref{eq:ealg-eq1}. Then
-we have the judgmental equality
-\begin{equation*}
-\jcomp{}{\epsilon_0}{\jcomp{}{\epsilon_0}{P}}
-  \jdeq
-  \jcomp{}{\epsilon_1}{\jcomp{}{\epsilon_0}{P}}.
-\end{equation*}
-\end{lem}
-
-\begin{proof}
-The proof is a rather long computation. It was found by starting with the
-application of \autoref{eq:ealg-eq1} and making the computation in both ways
-from there. The judgmental equality below where \autoref{eq:ealg-eq1} is applied,
-is marked with an $*$. The computation goes as follows:
-\begin{align*}
-\jcomp{}{\epsilon_0}{\jcomp{}{\epsilon_0}{P}}
-& \jdeq
-  \subst{\epsilon_0}
-    {\ctxwk{\ctxext{A}{P}}{\jcomp{}{\epsilon_0}{P}}}
-  \\
-& \jdeq
-  \subst
-    {\idtm{\jcomp{}{\epsilon_0}{P}}}
-    { \ctxwk
-        {\jcomp{}{\epsilon_0}{P}}{\subst{\epsilon_0}
-        {\ctxwk{\ctxext{A}{P}}{\jcomp{}{\epsilon_0}{P}}}}}
-  \\
-& \jdeq
-  \subst
-    {\idtm{\jcomp{}{\epsilon_0}{P}}}
-    { {\ctxwk{\jcomp{}{\epsilon_0}{P}}{\epsilon_0}}
-      {\ctxwk{\jcomp{}{\epsilon_0}{P}}{{\ctxext{A}{P}}{\jcomp{}{\epsilon_0}{P}}}}}
-  \\
-& \jdeq
-  \subst
-    {\idtm{\jcomp{}{\epsilon_0}{P}}}
-    { {\ctxwk{\jcomp{}{\epsilon_0}{P}}{\epsilon_0}}
-      {\ctxwk{\ctxext{{A}{P}}{\jcomp{}{\epsilon_0}{P}}}{\jcomp{}{\epsilon_0}{P}}}}
-  \\
-& \jdeq
-  \subst
-    {\idtm{\jcomp{}{\epsilon_0}{P}}}
-    { {\ctxwk{\jcomp{}{\epsilon_0}{P}}{\epsilon_0}}
-      {\ctxwk{\ctxext{A}{{P}{\jcomp{}{\epsilon_0}{P}}}}{\jcomp{}{\epsilon_0}{P}}}}
-  \\
-& \jdeq
-  \subst
-    {\unfold{\jvcomp{\jcomp{}{\epsilon_0}{P}}{\epsilon_0}{\idtm{\jcomp{}{\epsilon_0}{P}}}}}
-    {\ctxwk{\ctxext{A}{{P}{\jcomp{}{\epsilon_0}{P}}}}{\jcomp{}{\epsilon_0}{P}}}
-  \\
-& \jdeq
-  \subst
-    {\jvcomp{}{\epsilon_0}{\idtm{\jcomp{}{\epsilon_0}{P}}}}
-    {\ctxwk{\ctxext{A}{{P}{\jcomp{}{\epsilon_0}{P}}}}{\jcomp{}{\epsilon_0}{P}}}
-  \\
-& \jdeq
-  \jcomp{}{\jvcomp{}{\epsilon_0}{\idtm{\jcomp{}{\epsilon_0}{P}}}}{\jcomp{}{\epsilon_0}{P}}
-  \\
-& \jdeq
-  \jcomp{}{\jcomp{}{\jvcomp{}{\epsilon_0}{\idtm{\jcomp{}{\epsilon_0}{P}}}}{\epsilon_0}}{P}
-  \\
-& \stackrel{*}{\jdeq}
-  \jcomp{}{
-    \jcomp{}{\jvcomp{}{\idtm{A}}{\epsilon_1}}{\epsilon_0}}{P}
-  \\
-& \jdeq
-  \jcomp{}{\jvcomp{}{\idtm{A}}{\epsilon_1}}{
-    \jcomp{}{\epsilon_0}{P}}
-  \\
-& \jdeq
-  \subst
-    {\tmext{\ctxwk{\ctxext{P}{\jcomp{}{\epsilon_0}{P}}}{\idtm{A}}}{\epsilon_1}}
-    {\ctxwk{\ctxext{A}{{P}{\jcomp{}{\epsilon_0}{P}}}}{\jcomp{}{\epsilon_0}{P}}}
-  \\
-& \jdeq
-  \subst
-    {\epsilon_1}
-    { {\ctxwk{\ctxext{P}{\jcomp{}{\epsilon_0}{P}}}{\idtm{A}}}
-      {\ctxwk{\ctxext{A}{{P}{\jcomp{}{\epsilon_0}{P}}}}{\jcomp{}{\epsilon_0}{P}}}
-      }
-  \\
-& \jdeq
-  \subst
-    {\epsilon_1}
-    { {\cprojfstf{A}{\ctxext{P}{\jcomp{}{\epsilon_0}{P}}}}
-      {\ctxwk{\ctxext{A}{{P}{\jcomp{}{\epsilon_0}{P}}}}{\jcomp{}{\epsilon_0}{P}}}
-      }
-  \\
-& \jdeq
-  \subst
-    {\epsilon_1}
-    {\jcomp{}{\cprojfstf{A}{\ctxext{P}{\jcomp{}{\epsilon_0}{P}}}}{\jcomp{}{\epsilon_0}{P}}}
-  \\
-& \jdeq
-  \subst
-    {\epsilon_1}
-    {\ctxwk{\ctxext{P}{\jcomp{}{\epsilon_0}{P}}}{\jcomp{}{\epsilon_0}{P}}}
-  \\
-& \jdeq
-  \jcomp{}{\epsilon_1}{\jcomp{}{\epsilon_0}{P}}.
-  \qedhere
-\end{align*}
-\end{proof}
-
-Now we see that we can use the morphism
-\begin{equation*}
-\jhom
-  {{\Gamma}{A}}
-  {{P}{{\jcomp{}{\epsilon_0}{P}}{\jcomp{}{\epsilon_1}{{}{\epsilon_0}{P}}}}}
-  {{P}{\jcomp{}{\epsilon_0}{P}}}
-  {\jvcomp{}{\epsilon_1}{\idtm{\jcomp{}{\epsilon_1}{{}{\epsilon_0}{P}}}}}.
-\end{equation*}
-Thus, the second judgmental equality we will need is that the diagram
-\begin{equation}\label{eq:ealg-eq2}
-\begin{tikzcd}[column sep=huge]
-\ctxext{P}{{\jcomp{}{\epsilon_0}{P}}{\jcomp{}{\epsilon_0}{{}{\epsilon_0}{P}}}} 
-  \ar{r}{\jvcomp{}{\idtm{P}}{\jcomp{}{\epsilon_0}{\epsilon_1}}}
-  \ar{d}[swap]{
-    \jvcomp{}{\epsilon_1}{\idtm{\jcomp{}{\epsilon_1}{{}{\epsilon_0}{P}}}}
-%    \tmext
-%      {\ctxwk{\jcomp{}{\bar{e}}{{}{e}{P}}}{f}}
-%      {\idtm{\jcomp{}{\bar{e}}{{}{e}{P}}}}
-    }
-& \ctxext{P}{\jcomp{}{\epsilon_0}{P}} \ar{d}{\epsilon_1}\\
-\ctxext{P}{\jcomp{}{\epsilon_0}{P}} \ar{r}[swap]{\epsilon_1} & P
-\end{tikzcd}
-\end{equation}
-commutes judgmentally. Now we can confidently formulate the definition of
-extension algebras.
-
-\begin{defn}
-An \emph{extension algebra in context $\Gamma$} is a quadruple $(A,P,\epsilon_0,\epsilon_1)$
-consisting of a family $A$ over context $\Gamma$, a family $P$ over the context
-$\ctxext{\Gamma}{A}$, a morphism $\epsilon_0$ from $\ctxext{A}{P}$ to $A$ in context
-$\Gamma$ and a morphism $\epsilon_1$ from $\ctxext{P}{\jcomp{}{\epsilon_0}{P}}$ to $P$ in context
-$\ctxext{\Gamma}{A}$, satisfying the judgmental equalities of
-\autoref{eq:ealg-eq1,eq:ealg-eq2}.
-\end{defn}
-
-We also give a bit of intuition to the requirement of \autoref{eq:ealg-eq2} by
-means of the following lemma.
-
-\begin{lem}
-Let $(A,P,\epsilon_0,\epsilon_1)$ be an extension algebra in context $\Gamma$ and let
-$y_0:P$, $y_1:\subst{y_0}{\jcomp{}{\epsilon_0}{P}}$ and 
-$y_2:\subst{y_1}{{y_0}{\jcomp{}{\epsilon_0}{\jcomp{}{\epsilon_0}{P}}}}$. Then we have the
-judgmental equality
-\begin{equation*}
-\subst{{y_2}{{y_1}{\jcomp{}{\epsilon_0}{\epsilon_1}}}}{{y_0}{\epsilon_1}}
-  \jdeq
-  \subst{y_2}{{{y_1}{{y_0}{\epsilon_1}}}{\epsilon_1}}
-\end{equation*}
-\end{lem}
-
-\begin{proof}
-The proof is a straightforward calculation:
-\begin{align*}
-\subst{{y_2}{{y_1}{\jcomp{}{\epsilon_0}{\epsilon_1}}}}{{y_0}{\epsilon_1}}
-& \jdeq
-  \subst{\tmext{y_0}{\subst{y_2}{{y_1}{\jcomp{}{\epsilon_0}{\epsilon_1}}}}}{\epsilon_1}
-  \\
-& \jdeq
-  \subst{{y_2}{{y_1}{{y_0}{\jvcomp{}{\idtm{P}}{\jcomp{}{\epsilon_0}{\epsilon_1}}}}}}{\epsilon_1}
-  \\
-& \jdeq
-  \subst
-    {y_2}
-    { {y_1}
-      { {y_0}
-        {\jcomp{}{\jvcomp{}{\idtm{P}}{\jcomp{}{\epsilon_0}{\epsilon_1}}}{\epsilon_1}}
-        }
-      }
-  \\
-& \jdeq
-  \subst
-    {y_2}
-    { {y_1}
-      { {y_0}
-        {\jcomp{}{\jvcomp{}{\epsilon_1}{\idtm{\jcomp{}{\epsilon_1}{{}{\epsilon_0}{P}}}}}{\epsilon_1}}
-        }
-      }
-  \\
-& \jdeq
-  \subst{{y_2}{{y_1}{{y_0}{\jvcomp{}{\epsilon_1}{\idtm{\jcomp{}{\epsilon_1}{{}{\epsilon_0}{P}}}}}}}}{\epsilon_1}
-  \\
-& \jdeq
-  \subst{\tmext{\subst{y_1}{{y_0}{\epsilon_1}}}{y_2}}{\epsilon_1}
-  \\
-& \jdeq
-  \subst{y_2}{{{y_1}{{y_0}{\epsilon_1}}}{\epsilon_1}}.\qedhere
-\end{align*}
-\end{proof}
-
-\begin{comment}
-Extension algebras don't come in isolation. There are also extension algebra
-families and extension algebra terms. We now aim to define these and to
-establish various constructions of new extension algebras out of old ones:
-the empty extension algebra, and extensions, weakenings and substitutions
-of extension algebras and of course the identity term as an extension algebra
-term. We start with extension algebra families.
-
-\begin{defn}
-Consider an extension algebra $\mathcal{A}\defeq(A,P,e,f)$. 
-An extension algebra family over $\mathcal{A}$ is likewise a quadruple
-$\mathcal{B}\defeq(B,Q,g,h)$. Here we have a family $\jfam{{\Gamma}{A}}{B}$, a
-family $\jfam{{{{\Gamma}{A}}{P}}{\ctxwk{P}{B}}}{Q}$ and
-\begin{align*}
-\jhom*{{{\Gamma}{A}}{P}}{\ctxext{\ctxwk{P}{B}}{Q}}{\jcomp{}{\epsilon_0}{B}}{g}\\
-\jhom*{{{{\Gamma}{A}}{P}}{\ctxwk{P}{B}}}{\ctxext{Q}{\jcomp{}{g}{Q}}}{Q}{h}.
-\end{align*}
-The quadruple $(\jcomp{}{\epsilon_0}{B},Q,g,h)$ is required to be an extension algebra
-in context $\ctxext{{\Gamma}{A}}{P}$.
-\end{defn}
-
-\begin{defn}
-Suppose $\mathcal{A}$ is an extension algebra and $\mathcal{B}$ is an extension
-algebra family over $\mathcal{A}$. A term of $\mathcal{B}$ is a pair $(x,y)$
-consisting of
-\begin{align*}
-\jterm*{{\Gamma}{A}}{B}{x}\\
-\jterm*{{{\Gamma}{A}}{P}}{\subst{\jcomp{}{\epsilon_0}{x}}{Q}}{y}
-\end{align*}
-such that the diagrams
-\begin{equation*}
-\begin{tikzcd}
-\ctxext{\jcomp{}{\epsilon_0}{B}}{Q} 
-  \ar{r}{g} 
-  \ar[shift right=.7ex,fib]{d}
-& B 
-  \ar[shift right=.7ex,fib]{d} 
-  \\
-\ctxext{A}{P} 
-  \ar[shift right=.7ex,dotted]{u}[swap]{\tmext{\jcomp{}{\epsilon_0}{x}}{y}}
-  \ar{r}{e}
-& A
-  \ar[shift right=.7ex,dotted]{u}[swap]{x}
-\end{tikzcd}
-\end{equation*}
-and
-\begin{equation*}
-\begin{tikzcd}
-\jcomp{}{f}{\ctxext{Q}{\jcomp{}{g}{Q}}}
-  \ar{r}{\jcomp{}{f}{h}}
-  \ar[shift right=.7ex,fib]{d}
-& Q
-  \ar[shift right=.7ex,fib]{d}
-  \\
-\jcomp{}{f}{\jcomp{}{\epsilon_0}{B}}
-  \ar{r}{\idtm{\jcomp{}{f}{\jcomp{}{\epsilon_0}{B}}}}
-  \ar[shift right=.7ex,fib]{d}
-  \ar[shift right=.7ex,dotted,mapsto]{u}[swap]{\jcomp{}{f}{y}}
-& \jcomp{}{\epsilon_0}{B}
-  \ar[shift right=.7ex,fib]{d}
-  \ar[shift right=.7ex,dotted,mapsto]{u}[swap]{y}
-  \\
-\ctxext{P}{\jcomp{}{\epsilon_0}{P}}
-  \ar{r}[swap]{f}
-  \ar[shift right=.7ex,dotted]{u}[swap]{\jcomp{}{f}{\jcomp{}{\epsilon_0}{x}}}
-& P
-  \ar[shift right=.7ex,dotted]{u}[swap]{\jcomp{}{\epsilon_0}{x}}
-\end{tikzcd}
-\end{equation*}
-commute.
-\end{defn}
-
-\begin{defn}
-Suppose $\mathcal{A}$ and $\mathcal{B}$ are extension algebras in context
-$\Gamma$. We define the extension algebra $\ctxwk{\mathcal{A}}{\mathcal{B}}$
-to be the quadruple
-\begin{equation*}
-(\ctxwk{A}{B},\ctxwk{\ctxext{A}{P}}{Q},\ctxwk{\ctxext{A}{P}}{g},\ctxwk{\ctxext{A}{P}}{h}).
-\end{equation*}
-Note that $\ctxwk{\ctxext{A}{P}}{Q}$ is a family over $\ctxwk{\ctxext{A}{P}}{B}$,
-whereas it should be a family over $\jcomp{}{\epsilon_0}{\ctxwk{A}{B}}$. These are the
-same by \autoref{lem:prehom}.
-\end{defn}
-
-\begin{rmk}
-Before we continue, let us explore what it means to be an extension algebra
-term of the extension algebra $\ctxwk{\mathcal{A}}{\mathcal{B}}$. Such an
-extension algebra term $(x,y)$ would consist of
-\begin{align*}
-\jterm*{{\Gamma}{A}}{\ctxwk{A}{B}}{x}\\
-\jterm*{{{\Gamma}{A}}{P}}{\subst{\jcomp{}{\epsilon_0}{x}}{\ctxwk{\ctxext{A}{P}}{Q}}}{y}.
-\end{align*}
-Thus, $x$ is a context morphism from $A$ to $B$ and $y$ is nothing but a term
-of $\jcomp{}{\jcomp{}{\epsilon_0}{x}}{Q}$. For $x$, we see that the diagram
-\begin{equation*}
-\begin{tikzcd}
-\ctxext{B}{Q} 
-  \ar{r}{g} 
-& B 
-  \\
-\ctxext{A}{P} 
-  \ar{u}{\jvcomp{}{x}{y}}
-  \ar{r}{e}
-& A
-  \ar{u}[swap]{x}
-\end{tikzcd}
-\end{equation*}
-commutes.
-\end{rmk}
-\end{comment}
-
-\subsection{Extension-empty algebras}
-\begin{defn}
-Let $P$ be a family over the extended context $\ctxext{\Gamma}{A}$, let
-$\jterm{\Gamma}{A}{\phi_0}$ and $\jterm{{\Gamma}{A}}{P}{\phi_1}$ be terms. Then the
-quadruple $(A,P,\phi_0,\phi_1)$ is said to be an \emph{empty algebra in context $\Gamma$}
-if the following judgmental equalities hold:
-\begin{align}
-\jfameq*{\Gamma}{\subst{\phi_0}{P}}{A}
-  \label{empalg-eq1}
-  \\
-\jtermeq*{\Gamma}{A}{\subst{\phi_0}{\phi_1}}{\phi_0}.
-  \label{empalg-eq2}
-\end{align}
-\end{defn}
-
-Thus, the empty algebras are the kind of algebras that require that families
-are compatible with contexts, just as our motivation in \autoref{empty}. We
-now combine the notion of extension algebras and empty algebras.
-
-An extension-empty algebra in context $\Gamma$ is going to be a sextuple
-$(A,P,\epsilon_0,\epsilon_1,\phi_0,\phi_1)$ for which 
-the quadruple $(A,P,\epsilon_0,\epsilon_1)$ is an extension algebra in context 
-$\Gamma$, the quadruple $(A,P,\phi_0,\phi_1)$ is an empty algebra in context
-$\Gamma$, satisfying additional judgmental equalities expressing the 
-compatibility of $\epsilon_0$ and $\epsilon_1$ with $\phi_0$ and $\phi_1$.
-There will be four such judgmental equalities.
-
-We can immediately state the first two:
-\begin{align}
-\jtermeq*{{\Gamma}{A}}{\ctxwk{A}{A}}{\subst{\phi_0}{\epsilon_0}}{\idtm{A}}
-  \label{extempalg-eq1}
-  \\
-\jtermeq*{{\Gamma}{A}}{\ctxwk{A}{A}}{\subst{\phi_1}{\epsilon_0}}{\idtm{A}}
-  \label{extempalg-eq2}
-\end{align}
-To see what $\subst{\phi_0}{\epsilon_1}$ can be, we must know its type first.
-It is a morphism from $\subst{\phi_0}{\ctxext{P}{\jcomp{}{\epsilon_0}{P}}}$ to
-$\subst{\phi_0}{P}$. We already know that $\subst{\phi_0}{P}\jdeq A$ by
-\autoref{empalg-eq1} and to compute $\subst{\phi_0}{\jcomp{}{\epsilon_0}{P}}$
-we use the following lemma.
-
-\begin{lem}\label{lem:empalg-mor}
-Consider an empty algebra $(A,P,\phi_0,\phi_1)$ in context $\Gamma$
-and a morphism $\jhom{\Gamma}{{A}{P}}{B}{f}$.
-Then $\subst{\phi_i}{f}$ is a morphism from $A$ to $B$ in context $\Gamma$ and
-the following inference rules are valid for $i$ being $0$ or $1$:
-\begin{align*}
-& \inference
-  { \jfam{{\Gamma}{A}}{Q}
-    }
-  { \jfameq
-      {{\Gamma}{A}}
-      {\subst{\phi_i}{\jcomp{}{f}{Q}}}
-      {\jcomp{}{\subst{\phi_i}{f}}{Q}}
-    }
-  \\
-& \inference
-  { \jfam{{{\Gamma}{A}}{Q}}{R}
-    }
-  { \jfameq
-      {{{\Gamma}{A}}{\jcomp{}{\subst{\phi_i}{f}}{Q}}}
-      {\subst{\phi_i}{\jcomp{}{f}{R}}}
-      {\jcomp{}{\subst{\phi_i}{f}}{R}}
-    }
-  \\
-& \inference
-  { \jterm{{{\Gamma}{A}}{Q}}{R}{h}
-    }
-  { \jtermeq
-      {{{\Gamma}{A}}{\jcomp{}{\subst{\phi_i}{f}}{Q}}}
-      {\jcomp{}{\subst{\phi_i}{f}}{R}}
-      {\subst{\phi_i}{\jcomp{}{f}{h}}}
-      {\jcomp{}{\subst{\phi_i}{f}}{h}}
-    }
-\end{align*}
-\end{lem}
-
-\begin{proof}
-We only prove the first inference rule in both cases.
-Let $Q$ be a family over $\ctxext{\Gamma}{A}$. In the case $i=0$
- we have the judgmental equalities
-\begin{align*}
-\subst{\phi_0}{\jcomp{}{f}{Q}}
-& \jdeq
-  \subst{\phi_0}{{f}{\ctxwk{\ctxext{A}{P}}{Q}}}
-  \tag{by definition}
-  \\
-& \jdeq
-  \subst{{\phi_0}{f}}{{\phi_0}{\ctxwk{\ctxext{A}{P}}{Q}}}
-  \tag{by \autoref{comp-ss-f}}
-  \\
-& \jdeq
-  \subst{{\phi_0}{f}}{{\phi_0}{\ctxwk{P}{{A}{Q}}}}
-  \tag{by \autoref{comp-ew-f}}
-  \\
-& \jdeq
-  \subst{{\phi_0}{f}}{\ctxwk{\subst{\phi_0}{P}}{\subst{\phi_0}{\ctxwk{A}{Q}}}}
-  \tag{by \autoref{comp-sw-f}}
-  \\
-& \jdeq
-  \subst{{\phi_0}{f}}{\ctxwk{\subst{\phi_0}{P}}{Q}}
-  \tag{by \autoref{cancellation-ws-f}}
-  \\
-& \jdeq
-  \subst{{\phi_0}{f}}{\ctxwk{A}{Q}}
-  \tag{by \autoref{empalg-eq1}}
-  \\
-& \jdeq
-  \jcomp{}{\subst{\phi_0}{f}}{Q}.
-  \tag{by definition}
-\end{align*}
-In the case $i=1$ we have the judgmental equalities
-\begin{align*}
-\subst{\phi_1}{\jcomp{}{f}{Q}}
-& \jdeq
-  \subst{\phi_1}{{f}{\ctxwk{\ctxext{A}{P}}{Q}}}
-  \tag{by definition}
-  \\
-& \jdeq
-  \subst{{\phi_1}{f}}{{\phi_1}{\ctxwk{\ctxext{A}{P}}{Q}}}
-  \tag{by \autoref{comp-ss-f}}
-  \\
-& \jdeq
-  \subst{{\phi_1}{f}}{{\phi_1}{\ctxwk{P}{{A}{Q}}}}
-  \tag{by \autoref{comp-ew-f}}
-  \\
-& \jdeq
-  \subst{{\phi_1}{f}}{\ctxwk{A}{Q}}
-  \tag{by \autoref{cancellation-ws-f}}
-  \\
-& \jdeq
-  \jcomp{}{\subst{\phi_1}{f}}{Q}.
-  \tag{by definition}
-\end{align*}
-\end{proof}
-
-As an immediate corollary, if we assume the judgmental equalities
-\autoref{extempalg-eq1,extempalg-eq2} we get that 
-\begin{equation}\label{cor:empalg-mor}
-\jfameq{{\Gamma}{A}}{\subst{\phi_i}{\jcomp{}{\epsilon_0}{P}}}{P}
-\end{equation}
-and hence that $\subst{\phi_0}{\epsilon_1}$ is a 
-morphism from $\ctxext{A}{P}$ to $A$. Thus, we can require
-\begin{equation}\label{extempalg-eq3}
-\jhomeq{\Gamma}{{A}{P}}{A}{\subst{\phi_0}{\epsilon_1}}{\epsilon_0}.
-\end{equation}
-For the final judgmental equality we need to explain the term
-\begin{equation*}
-\jterm
-  {{{\Gamma}{A}}{\subst{\phi_1}{\jcomp{}{\epsilon_0}{P}}}}
-  {\subst{\phi_1}{\ctxwk{\ctxext{P}{\jcomp{}{\epsilon_0}{P}}}{P}}}
-  {\subst{\phi_1}{\epsilon_1}}.
-\end{equation*}
-We have already established that $\subst{\phi_1}{\jcomp{}{\epsilon_0}{P}}\jdeq
-P$. We also see that 
-\begin{align*}
-\subst{\phi_1}{\ctxwk{\ctxext{P}{\jcomp{}{\epsilon_0}{P}}}{P}}
-& \jdeq
-  \subst{\phi_1}{\ctxwk{\jcomp{}{\epsilon_0}{P}}{{P}{P}}}
-  \tag{by \autoref{comp-ew-f}}
-  \\
-& \jdeq
-  \ctxwk{\subst{\phi_1}{\jcomp{}{\epsilon_0}{P}}}{\subst{\phi_1}{\ctxwk{P}{P}}}
-  \tag{by \autoref{comp-sw-f}}
-  \\
-& \jdeq
-  \ctxwk{P}{\subst{\phi_1}{\ctxwk{P}{P}}}
-  \tag{by \autoref{cor:empalg-mor}}
-  \\
-& \jdeq
-  \ctxwk{P}{P}
-  \tag{by \autoref{cancellation-ws-f}}
-\end{align*}
-and we will therefore require that
-\begin{equation}\label{extempalg-eq4}
-\jtermeq{{{\Gamma}{A}}{P}}{\ctxwk{P}{P}}{\subst{\phi_1}{\epsilon_1}}{\idtm{P}}.
-\end{equation}
-We bring all this together in the definition of extension-empty algebras:
-
-\begin{defn}
-An \emph{extension-empty algebra in context $\Gamma$} 
-is a sextuple $(A,P,\epsilon_0,\epsilon_1,\phi_0,\phi_1)$ for which 
-the quadruple $(A,P,\epsilon_0,\epsilon_1)$ is an extension algebra in context 
-$\Gamma$, the quadruple $(A,P,\phi_0,\phi_1)$ is an empty algebra in context
-$\Gamma$, satisfying the judgmental equalities 
-\autoref{extempalg-eq1,extempalg-eq2,extempalg-eq3,extempalg-eq4}.
-\end{defn}
-
-\subsection{Extension-weakening algebras}
-The notion of (extension-)weakening algebra is dependent on the notion of extension algebra.
-Although it is strictly speaking not dependent on the notion of empty-algebra,
-we shall only formulate a weakening algebras in the setting of an
-extension-empty algebras. When one wants to have a weakening operation which
-also acts on the level of contexts in an extension algebra without empty
-context, extra work has to be done to introduce these separately.
-
-\begin{defn}
-An extension-weakening algebra in context $\Gamma$ is an octuple
-\begin{equation*}
-(A,P,\epsilon_0,\epsilon_1,\phi_0,\phi_1,\omega_0,\omega_1)
-\end{equation*}
-where $(A,P,\epsilon_0,\epsilon_1,\phi_0,\phi_1)$ is an extension-empty algebra
-in context $\Gamma$ and where
-\begin{align*}
-\jhom*
-  {{{\Gamma}{A}}{P}}
-  {\ctxwk{P}{P}}
-  {\jcomp{}{\epsilon_0}{P}}
-  {\omega_0}
-  \\
-\jfhom*
-  {{{\Gamma}{A}}{P}}
-  {\ctxwk{P}{P}}
-  {\jcomp{}{\epsilon_0}{P}}
-  {\omega_0}
-  {\ctxwk{P}{\jcomp{}{\epsilon_0}{P}}}
-  {\jcomp{}{\epsilon_0}{\jcomp{}{\epsilon_0}{P}}}
-  {\omega_1}
-\end{align*}
-satisfying the judgmental equalities
-\end{defn}
-
-\subsection{Extension-substitution algebras}
-
-\subsection{Pre-universes}
-Pre-universes are internal versions of the theory of contexts, families and
-terms. They interpret extension, the empty context, weakening, substitution
-and identity terms all at once in a compatible way.

### 2014-05-15 article-models-ttinternal.tex
added file

\section{Internalizing the theory of contexts families and terms}
One of the guiding ideas behind the design of the theory of contexts, families
and terms was that it would have to be possible to consider internal versions
of the theory. In this section we aim for this internalization.

\subsection{Extension algebras}
In this subsection our goal is to define the notion of extension algebras,
which are internal versions of the extension operation of the theory of
contexts, families and terms. In this article, their use will be mainly in
universes. The theory of extension algebras requires the full power (i.e.~all
of the ingredients) of the theory of contexts, families and terms in its
formulation and it is (perhaps surprisingly) quite involved to formulate it.

Let $P$ be a family over an extended context $\ctxext{\Gamma}{A}$. We could
mimic extension by requiring to have terms
\begin{align*}
\jhom*{\Gamma}{{A}{P}}{A}{\epsilon_0}\\
\jhom*{{\Gamma}{A}}{{P}{\jcomp{}{\epsilon_0}{P}}}{P}{\epsilon_1}.
\end{align*}

\begin{rmk}
Instead of looking at $\epsilon_1$ as a context morphism from $\ctxext{P}{\jcomp{}{\epsilon_0}{P}}$
to $P$, one could also look at $\epsilon_1$ as a morphism \emph{over $\cprojfstf{A}{P}$},
as indicated in the following diagram:
\begin{equation*}
\begin{tikzcd}
P
  \ar[fib]{d}
& \jcomp{}{\epsilon_0}{P}
  \ar[fib]{d}
  \ar{l}[swap]{\epsilon_1}
  \ar{r}
& P
  \ar[fib]{d}
  \\
A
& \ctxext{A}{P}
  \ar{l}{\cprojfstf{A}{P}}
  \ar{r}[swap]{\epsilon_0}
& A
\end{tikzcd}
\end{equation*}
This makes it clear that $\epsilon_1$ takes a family over an extended context as an
argument. The extended context consists of a `base part' and a `family part'. 
The output of $\epsilon_1$ is a new (extended) family over that base part. Forgetting 
the family part is what the projection takes care of.
\end{rmk}

Extension also satisfies the properties explained in \autoref{comp-ee}, so we
must find the two judgmental equalities for $\epsilon_0$ and $\epsilon_1$ mimicing those. 
The first of these judgmental equalities is easy to give: it says that the
following diagram commutes:
\begin{equation}\label{eq:ealg-eq1}
\begin{tikzcd}[column sep=huge]
\ctxext{A}{{P}{\jcomp{}{\epsilon_0}{P}}} 
  \ar{d}[swap]{\jvcomp{}{\epsilon_0}{\idtm{\jcomp{}{\epsilon_0}{P}}}
    } 
  \ar{r}{\jvcomp{}{\idtm{A}}{\epsilon_1}
    } 
  & \ctxext{A}{P} \ar{d}{\epsilon_0}\\
\ctxext{A}{P} \ar{r}[swap]{\epsilon_0} & A
\end{tikzcd}
\end{equation}
To get a feel for this judgmental equality we include the following lemma.

\begin{lem}
Let $A$, $P$, $\epsilon_0$ and $\epsilon_1$ be as above, satisfying \autoref{eq:ealg-eq1} and
let $x_0:A$,
$x_1:\subst{x_0}{P}$ and $x_2:\subst{x_1}{{x_0}{\jcomp{}{\epsilon_0}{P}}}$.
Then we have the judgmental equality
\begin{equation*}
\subst{{x_2}{{x_1}{\epsilon_1}}}{{x_0}{\epsilon_0}}
\jdeq
\subst{x_2}{{{x_1}{{x_0}{\epsilon_0}}}{\epsilon_0}}.
\end{equation*}
\end{lem}

\begin{proof}
The proof is a simple computation:
\begin{align*}
\subst{{x_2}{{x_1}{\epsilon_1}}}{{x_0}{\epsilon_0}}
& \jdeq
  \subst{\tmext{x_0}{\subst{x_2}{{x_1}{\epsilon_1}}}}{\epsilon_0}
  \\
& \jdeq 
  \subst{{x_2}{{x_1}{{x_0}{\jvcomp{}{\idtm{A}}{\epsilon_1}}}}}{\epsilon_0}
  \\
& \jdeq
  \subst{x_2}{{x_1}{{x_0}{\jcomp{}{\jvcomp{}{\idtm{A}}{\epsilon_1}}{\epsilon_0}}}}
  \\
& \jdeq
  \subst{x_2}{{x_1}{{x_0}{\jcomp{}{\jvcomp{}{\epsilon_0}{\idtm{\jcomp{}{\epsilon_0}{P}}}}{\epsilon_0}}}}
  \\
& \jdeq 
  \subst{{x_2}{{x_1}{{x_0}{\jvcomp{}{\epsilon_0}{\idtm{\jcomp{}{\epsilon_0}{P}}}}}}}{\epsilon_0}
  \\
& \jdeq
  \subst{\tmext{\subst{x_1}{{x_0}{\epsilon_0}}}{x_2}}{\epsilon_0}
  \\
& \jdeq
  \subst{x_2}{{{x_1}{{x_0}{\epsilon_0}}}{\epsilon_0}}.\qedhere
\end{align*}
\end{proof}

The second of the judgmental equalities is harder to describe, however. We need
to consider `higher' families, i.e.~families over families over families, and
thus we need to look at the family $\jcomp{}{\epsilon_0}{\jcomp{}{\epsilon_0}{P}}$ and find the
two dotted morphisms in the diagram
\begin{equation*}
\begin{tikzcd}
\ctxext{P}{{\jcomp{}{\epsilon_0}{P}}{\jcomp{}{\epsilon_0}{\jcomp{}{\epsilon_0}{P}}}}
  \ar[densely dotted]{d}
  \ar[densely dotted]{r}
& \ctxext{P}{\jcomp{}{\epsilon_0}{P}} \ar{d}{\epsilon_1}\\
\ctxext{P}{\jcomp{}{\epsilon_0}{P}} \ar{r}{\epsilon_1} & P
\end{tikzcd}
\end{equation*}
The first is easy to find. Note that we have the judgmental equality
\begin{equation*}
\ctxext{\jcomp{}{\epsilon_0}{P}}{\jcomp{}{\epsilon_0}{\jcomp{}{\epsilon_0}{P}}}
  \jdeq
  \jcomp{}{\epsilon_0}{\ctxext{P}{\jcomp{}{\epsilon_0}{P}}}
\end{equation*}
and therefore we may just take the morphism
\begin{equation*}
\jhom
  {{\Gamma}{A}}
  {{P}{{\jcomp{}{\epsilon_0}{P}}{\jcomp{}{\epsilon_0}{\jcomp{}{\epsilon_0}{P}}}}}
  {{P}{\jcomp{}{\epsilon_0}{P}}}
  {\jvcomp{}{\idtm{P}}{\jcomp{}{\epsilon_0}{\epsilon_1}}}.
\end{equation*}
For the other morphism we need to look at the family
$\ctxext{P}{{\jcomp{}{\epsilon_0}{P}}{\jcomp{}{\epsilon_0}{\jcomp{}{\epsilon_0}{P}}}}$ differently. We do
that in the following lemma.

\begin{lem}
Suppose we have $A$, $P$, $\epsilon_0$ and $\epsilon_1$ satisfying \autoref{eq:ealg-eq1}. Then
we have the judgmental equality
\begin{equation*}
\jcomp{}{\epsilon_0}{\jcomp{}{\epsilon_0}{P}}
  \jdeq
  \jcomp{}{\epsilon_1}{\jcomp{}{\epsilon_0}{P}}.
\end{equation*}
\end{lem}

\begin{proof}
The proof is a rather long computation. It was found by starting with the
application of \autoref{eq:ealg-eq1} and making the computation in both ways
from there. The judgmental equality below where \autoref{eq:ealg-eq1} is applied,
is marked with an $*$. The computation goes as follows:
\begin{align*}
\jcomp{}{\epsilon_0}{\jcomp{}{\epsilon_0}{P}}
& \jdeq
  \subst{\epsilon_0}
    {\ctxwk{\ctxext{A}{P}}{\jcomp{}{\epsilon_0}{P}}}
  \\
& \jdeq
  \subst
    {\idtm{\jcomp{}{\epsilon_0}{P}}}
    { \ctxwk
        {\jcomp{}{\epsilon_0}{P}}{\subst{\epsilon_0}
        {\ctxwk{\ctxext{A}{P}}{\jcomp{}{\epsilon_0}{P}}}}}
  \\
& \jdeq
  \subst
    {\idtm{\jcomp{}{\epsilon_0}{P}}}
    { {\ctxwk{\jcomp{}{\epsilon_0}{P}}{\epsilon_0}}
      {\ctxwk{\jcomp{}{\epsilon_0}{P}}{{\ctxext{A}{P}}{\jcomp{}{\epsilon_0}{P}}}}}
  \\
& \jdeq
  \subst
    {\idtm{\jcomp{}{\epsilon_0}{P}}}
    { {\ctxwk{\jcomp{}{\epsilon_0}{P}}{\epsilon_0}}
      {\ctxwk{\ctxext{{A}{P}}{\jcomp{}{\epsilon_0}{P}}}{\jcomp{}{\epsilon_0}{P}}}}
  \\
& \jdeq
  \subst
    {\idtm{\jcomp{}{\epsilon_0}{P}}}
    { {\ctxwk{\jcomp{}{\epsilon_0}{P}}{\epsilon_0}}
      {\ctxwk{\ctxext{A}{{P}{\jcomp{}{\epsilon_0}{P}}}}{\jcomp{}{\epsilon_0}{P}}}}
  \\
& \jdeq
  \subst
    {\unfold{\jvcomp{\jcomp{}{\epsilon_0}{P}}{\epsilon_0}{\idtm{\jcomp{}{\epsilon_0}{P}}}}}
    {\ctxwk{\ctxext{A}{{P}{\jcomp{}{\epsilon_0}{P}}}}{\jcomp{}{\epsilon_0}{P}}}
  \\
& \jdeq
  \subst
    {\jvcomp{}{\epsilon_0}{\idtm{\jcomp{}{\epsilon_0}{P}}}}
    {\ctxwk{\ctxext{A}{{P}{\jcomp{}{\epsilon_0}{P}}}}{\jcomp{}{\epsilon_0}{P}}}
  \\
& \jdeq
  \jcomp{}{\jvcomp{}{\epsilon_0}{\idtm{\jcomp{}{\epsilon_0}{P}}}}{\jcomp{}{\epsilon_0}{P}}
  \\
& \jdeq
  \jcomp{}{\jcomp{}{\jvcomp{}{\epsilon_0}{\idtm{\jcomp{}{\epsilon_0}{P}}}}{\epsilon_0}}{P}
  \\
& \stackrel{*}{\jdeq}
  \jcomp{}{
    \jcomp{}{\jvcomp{}{\idtm{A}}{\epsilon_1}}{\epsilon_0}}{P}
  \\
& \jdeq
  \jcomp{}{\jvcomp{}{\idtm{A}}{\epsilon_1}}{
    \jcomp{}{\epsilon_0}{P}}
  \\
& \jdeq
  \subst
    {\tmext{\ctxwk{\ctxext{P}{\jcomp{}{\epsilon_0}{P}}}{\idtm{A}}}{\epsilon_1}}
    {\ctxwk{\ctxext{A}{{P}{\jcomp{}{\epsilon_0}{P}}}}{\jcomp{}{\epsilon_0}{P}}}
  \\
& \jdeq
  \subst
    {\epsilon_1}
    { {\ctxwk{\ctxext{P}{\jcomp{}{\epsilon_0}{P}}}{\idtm{A}}}
      {\ctxwk{\ctxext{A}{{P}{\jcomp{}{\epsilon_0}{P}}}}{\jcomp{}{\epsilon_0}{P}}}
      }
  \\
& \jdeq
  \subst
    {\epsilon_1}
    { {\cprojfstf{A}{\ctxext{P}{\jcomp{}{\epsilon_0}{P}}}}
      {\ctxwk{\ctxext{A}{{P}{\jcomp{}{\epsilon_0}{P}}}}{\jcomp{}{\epsilon_0}{P}}}
      }
  \\
& \jdeq
  \subst
    {\epsilon_1}
    {\jcomp{}{\cprojfstf{A}{\ctxext{P}{\jcomp{}{\epsilon_0}{P}}}}{\jcomp{}{\epsilon_0}{P}}}
  \\
& \jdeq
  \subst
    {\epsilon_1}
    {\ctxwk{\ctxext{P}{\jcomp{}{\epsilon_0}{P}}}{\jcomp{}{\epsilon_0}{P}}}
  \\
& \jdeq
  \jcomp{}{\epsilon_1}{\jcomp{}{\epsilon_0}{P}}.
  \qedhere
\end{align*}
\end{proof}

Now we see that we can use the morphism
\begin{equation*}
\jhom
  {{\Gamma}{A}}
  {{P}{{\jcomp{}{\epsilon_0}{P}}{\jcomp{}{\epsilon_1}{{}{\epsilon_0}{P}}}}}
  {{P}{\jcomp{}{\epsilon_0}{P}}}
  {\jvcomp{}{\epsilon_1}{\idtm{\jcomp{}{\epsilon_1}{{}{\epsilon_0}{P}}}}}.
\end{equation*}
Thus, the second judgmental equality we will need is that the diagram
\begin{equation}\label{eq:ealg-eq2}
\begin{tikzcd}[column sep=huge]
\ctxext{P}{{\jcomp{}{\epsilon_0}{P}}{\jcomp{}{\epsilon_0}{{}{\epsilon_0}{P}}}} 
  \ar{r}{\jvcomp{}{\idtm{P}}{\jcomp{}{\epsilon_0}{\epsilon_1}}}
  \ar{d}[swap]{
    \jvcomp{}{\epsilon_1}{\idtm{\jcomp{}{\epsilon_1}{{}{\epsilon_0}{P}}}}
%    \tmext
%      {\ctxwk{\jcomp{}{\bar{e}}{{}{e}{P}}}{f}}
%      {\idtm{\jcomp{}{\bar{e}}{{}{e}{P}}}}
    }
& \ctxext{P}{\jcomp{}{\epsilon_0}{P}} \ar{d}{\epsilon_1}\\
\ctxext{P}{\jcomp{}{\epsilon_0}{P}} \ar{r}[swap]{\epsilon_1} & P
\end{tikzcd}
\end{equation}
commutes judgmentally. Now we can confidently formulate the definition of
extension algebras.

\begin{defn}
An \emph{extension algebra in context $\Gamma$} is a quadruple $(A,P,\epsilon_0,\epsilon_1)$
consisting of a family $A$ over context $\Gamma$, a family $P$ over the context
$\ctxext{\Gamma}{A}$, a morphism $\epsilon_0$ from $\ctxext{A}{P}$ to $A$ in context
$\Gamma$ and a morphism $\epsilon_1$ from $\ctxext{P}{\jcomp{}{\epsilon_0}{P}}$ to $P$ in context
$\ctxext{\Gamma}{A}$, satisfying the judgmental equalities of
\autoref{eq:ealg-eq1,eq:ealg-eq2}.
\end{defn}

We also give a bit of intuition to the requirement of \autoref{eq:ealg-eq2} by
means of the following lemma.

\begin{lem}
Let $(A,P,\epsilon_0,\epsilon_1)$ be an extension algebra in context $\Gamma$ and let
$y_0:P$, $y_1:\subst{y_0}{\jcomp{}{\epsilon_0}{P}}$ and 
$y_2:\subst{y_1}{{y_0}{\jcomp{}{\epsilon_0}{\jcomp{}{\epsilon_0}{P}}}}$. Then we have the
judgmental equality
\begin{equation*}
\subst{{y_2}{{y_1}{\jcomp{}{\epsilon_0}{\epsilon_1}}}}{{y_0}{\epsilon_1}}
  \jdeq
  \subst{y_2}{{{y_1}{{y_0}{\epsilon_1}}}{\epsilon_1}}
\end{equation*}
\end{lem}

\begin{proof}
The proof is a straightforward calculation:
\begin{align*}
\subst{{y_2}{{y_1}{\jcomp{}{\epsilon_0}{\epsilon_1}}}}{{y_0}{\epsilon_1}}
& \jdeq
  \subst{\tmext{y_0}{\subst{y_2}{{y_1}{\jcomp{}{\epsilon_0}{\epsilon_1}}}}}{\epsilon_1}
  \\
& \jdeq
  \subst{{y_2}{{y_1}{{y_0}{\jvcomp{}{\idtm{P}}{\jcomp{}{\epsilon_0}{\epsilon_1}}}}}}{\epsilon_1}
  \\
& \jdeq
  \subst
    {y_2}
    { {y_1}
      { {y_0}
        {\jcomp{}{\jvcomp{}{\idtm{P}}{\jcomp{}{\epsilon_0}{\epsilon_1}}}{\epsilon_1}}
        }
      }
  \\
& \jdeq
  \subst
    {y_2}
    { {y_1}
      { {y_0}
        {\jcomp{}{\jvcomp{}{\epsilon_1}{\idtm{\jcomp{}{\epsilon_1}{{}{\epsilon_0}{P}}}}}{\epsilon_1}}
        }
      }
  \\
& \jdeq
  \subst{{y_2}{{y_1}{{y_0}{\jvcomp{}{\epsilon_1}{\idtm{\jcomp{}{\epsilon_1}{{}{\epsilon_0}{P}}}}}}}}{\epsilon_1}
  \\
& \jdeq
  \subst{\tmext{\subst{y_1}{{y_0}{\epsilon_1}}}{y_2}}{\epsilon_1}
  \\
& \jdeq
  \subst{y_2}{{{y_1}{{y_0}{\epsilon_1}}}{\epsilon_1}}.\qedhere
\end{align*}
\end{proof}

\begin{comment}
Extension algebras don't come in isolation. There are also extension algebra
families and extension algebra terms. We now aim to define these and to
establish various constructions of new extension algebras out of old ones:
the empty extension algebra, and extensions, weakenings and substitutions
of extension algebras and of course the identity term as an extension algebra
term. We start with extension algebra families.

\begin{defn}
Consider an extension algebra $\mathcal{A}\defeq(A,P,e,f)$. 
An extension algebra family over $\mathcal{A}$ is likewise a quadruple
$\mathcal{B}\defeq(B,Q,g,h)$. Here we have a family $\jfam{{\Gamma}{A}}{B}$, a
family $\jfam{{{{\Gamma}{A}}{P}}{\ctxwk{P}{B}}}{Q}$ and
\begin{align*}
\jhom*{{{\Gamma}{A}}{P}}{\ctxext{\ctxwk{P}{B}}{Q}}{\jcomp{}{\epsilon_0}{B}}{g}\\
\jhom*{{{{\Gamma}{A}}{P}}{\ctxwk{P}{B}}}{\ctxext{Q}{\jcomp{}{g}{Q}}}{Q}{h}.
\end{align*}
The quadruple $(\jcomp{}{\epsilon_0}{B},Q,g,h)$ is required to be an extension algebra
in context $\ctxext{{\Gamma}{A}}{P}$.
\end{defn}

\begin{defn}
Suppose $\mathcal{A}$ is an extension algebra and $\mathcal{B}$ is an extension
algebra family over $\mathcal{A}$. A term of $\mathcal{B}$ is a pair $(x,y)$
consisting of
\begin{align*}
\jterm*{{\Gamma}{A}}{B}{x}\\
\jterm*{{{\Gamma}{A}}{P}}{\subst{\jcomp{}{\epsilon_0}{x}}{Q}}{y}
\end{align*}
such that the diagrams
\begin{equation*}
\begin{tikzcd}
\ctxext{\jcomp{}{\epsilon_0}{B}}{Q} 
  \ar{r}{g} 
  \ar[shift right=.7ex,fib]{d}
& B 
  \ar[shift right=.7ex,fib]{d} 
  \\
\ctxext{A}{P} 
  \ar[shift right=.7ex,dotted]{u}[swap]{\tmext{\jcomp{}{\epsilon_0}{x}}{y}}
  \ar{r}{e}
& A
  \ar[shift right=.7ex,dotted]{u}[swap]{x}
\end{tikzcd}
\end{equation*}
and
\begin{equation*}
\begin{tikzcd}
\jcomp{}{f}{\ctxext{Q}{\jcomp{}{g}{Q}}}
  \ar{r}{\jcomp{}{f}{h}}
  \ar[shift right=.7ex,fib]{d}
& Q
  \ar[shift right=.7ex,fib]{d}
  \\
\jcomp{}{f}{\jcomp{}{\epsilon_0}{B}}
  \ar{r}{\idtm{\jcomp{}{f}{\jcomp{}{\epsilon_0}{B}}}}
  \ar[shift right=.7ex,fib]{d}
  \ar[shift right=.7ex,dotted,mapsto]{u}[swap]{\jcomp{}{f}{y}}
& \jcomp{}{\epsilon_0}{B}
  \ar[shift right=.7ex,fib]{d}
  \ar[shift right=.7ex,dotted,mapsto]{u}[swap]{y}
  \\
\ctxext{P}{\jcomp{}{\epsilon_0}{P}}
  \ar{r}[swap]{f}
  \ar[shift right=.7ex,dotted]{u}[swap]{\jcomp{}{f}{\jcomp{}{\epsilon_0}{x}}}
& P
  \ar[shift right=.7ex,dotted]{u}[swap]{\jcomp{}{\epsilon_0}{x}}
\end{tikzcd}
\end{equation*}
commute.
\end{defn}

\begin{defn}
Suppose $\mathcal{A}$ and $\mathcal{B}$ are extension algebras in context
$\Gamma$. We define the extension algebra $\ctxwk{\mathcal{A}}{\mathcal{B}}$
to be the quadruple
\begin{equation*}
(\ctxwk{A}{B},\ctxwk{\ctxext{A}{P}}{Q},\ctxwk{\ctxext{A}{P}}{g},\ctxwk{\ctxext{A}{P}}{h}).
\end{equation*}
Note that $\ctxwk{\ctxext{A}{P}}{Q}$ is a family over $\ctxwk{\ctxext{A}{P}}{B}$,
whereas it should be a family over $\jcomp{}{\epsilon_0}{\ctxwk{A}{B}}$. These are the
same by \autoref{lem:prehom}.
\end{defn}

\begin{rmk}
Before we continue, let us explore what it means to be an extension algebra
term of the extension algebra $\ctxwk{\mathcal{A}}{\mathcal{B}}$. Such an
extension algebra term $(x,y)$ would consist of
\begin{align*}
\jterm*{{\Gamma}{A}}{\ctxwk{A}{B}}{x}\\
\jterm*{{{\Gamma}{A}}{P}}{\subst{\jcomp{}{\epsilon_0}{x}}{\ctxwk{\ctxext{A}{P}}{Q}}}{y}.
\end{align*}
Thus, $x$ is a context morphism from $A$ to $B$ and $y$ is nothing but a term
of $\jcomp{}{\jcomp{}{\epsilon_0}{x}}{Q}$. For $x$, we see that the diagram
\begin{equation*}
\begin{tikzcd}
\ctxext{B}{Q} 
  \ar{r}{g} 
& B 
  \\
\ctxext{A}{P} 
  \ar{u}{\jvcomp{}{x}{y}}
  \ar{r}{e}
& A
  \ar{u}[swap]{x}
\end{tikzcd}
\end{equation*}
commutes.
\end{rmk}
\end{comment}

\subsection{Extension-empty algebras}
\begin{defn}
Let $P$ be a family over the extended context $\ctxext{\Gamma}{A}$, let
$\jterm{\Gamma}{A}{\phi_0}$ and $\jterm{{\Gamma}{A}}{P}{\phi_1}$ be terms. Then the
quadruple $(A,P,\phi_0,\phi_1)$ is said to be an \emph{empty algebra in context $\Gamma$}
if the following judgmental equalities hold:
\begin{align}
\jfameq*{\Gamma}{\subst{\phi_0}{P}}{A}
  \label{empalg-eq1}
  \\
\jtermeq*{\Gamma}{A}{\subst{\phi_0}{\phi_1}}{\phi_0}.
  \label{empalg-eq2}
\end{align}
\end{defn}

Thus, the empty algebras are the kind of algebras that require that families
are compatible with contexts, just as our motivation in \autoref{empty}. We
now combine the notion of extension algebras and empty algebras.

An extension-empty algebra in context $\Gamma$ is going to be a sextuple
$(A,P,\epsilon_0,\epsilon_1,\phi_0,\phi_1)$ for which 
the quadruple $(A,P,\epsilon_0,\epsilon_1)$ is an extension algebra in context 
$\Gamma$, the quadruple $(A,P,\phi_0,\phi_1)$ is an empty algebra in context
$\Gamma$, satisfying additional judgmental equalities expressing the 
compatibility of $\epsilon_0$ and $\epsilon_1$ with $\phi_0$ and $\phi_1$.
There will be four such judgmental equalities.

We can immediately state the first two:
\begin{align}
\jtermeq*{{\Gamma}{A}}{\ctxwk{A}{A}}{\subst{\phi_0}{\epsilon_0}}{\idtm{A}}
  \label{extempalg-eq1}
  \\
\jtermeq*{{\Gamma}{A}}{\ctxwk{A}{A}}{\subst{\phi_1}{\epsilon_0}}{\idtm{A}}
  \label{extempalg-eq2}
\end{align}
To see what $\subst{\phi_0}{\epsilon_1}$ can be, we must know its type first.
It is a morphism from $\subst{\phi_0}{\ctxext{P}{\jcomp{}{\epsilon_0}{P}}}$ to
$\subst{\phi_0}{P}$. We already know that $\subst{\phi_0}{P}\jdeq A$ by
\autoref{empalg-eq1} and to compute $\subst{\phi_0}{\jcomp{}{\epsilon_0}{P}}$
we use the following lemma.

\begin{lem}\label{lem:empalg-mor}
Consider an empty algebra $(A,P,\phi_0,\phi_1)$ in context $\Gamma$
and a morphism $\jhom{\Gamma}{{A}{P}}{B}{f}$.
Then $\subst{\phi_i}{f}$ is a morphism from $A$ to $B$ in context $\Gamma$ and
the following inference rules are valid for $i$ being $0$ or $1$:
\begin{align*}
& \inference
  { \jfam{{\Gamma}{A}}{Q}
    }
  { \jfameq
      {{\Gamma}{A}}
      {\subst{\phi_i}{\jcomp{}{f}{Q}}}
      {\jcomp{}{\subst{\phi_i}{f}}{Q}}
    }
  \\
& \inference
  { \jfam{{{\Gamma}{A}}{Q}}{R}
    }
  { \jfameq
      {{{\Gamma}{A}}{\jcomp{}{\subst{\phi_i}{f}}{Q}}}
      {\subst{\phi_i}{\jcomp{}{f}{R}}}
      {\jcomp{}{\subst{\phi_i}{f}}{R}}
    }
  \\
& \inference
  { \jterm{{{\Gamma}{A}}{Q}}{R}{h}
    }
  { \jtermeq
      {{{\Gamma}{A}}{\jcomp{}{\subst{\phi_i}{f}}{Q}}}
      {\jcomp{}{\subst{\phi_i}{f}}{R}}
      {\subst{\phi_i}{\jcomp{}{f}{h}}}
      {\jcomp{}{\subst{\phi_i}{f}}{h}}
    }
\end{align*}
\end{lem}

\begin{proof}
We only prove the first inference rule in both cases.
Let $Q$ be a family over $\ctxext{\Gamma}{A}$. In the case $i=0$
 we have the judgmental equalities
\begin{align*}
\subst{\phi_0}{\jcomp{}{f}{Q}}
& \jdeq
  \subst{\phi_0}{{f}{\ctxwk{\ctxext{A}{P}}{Q}}}
  \tag{by definition}
  \\
& \jdeq
  \subst{{\phi_0}{f}}{{\phi_0}{\ctxwk{\ctxext{A}{P}}{Q}}}
  \tag{by \autoref{comp-ss-f}}
  \\
& \jdeq
  \subst{{\phi_0}{f}}{{\phi_0}{\ctxwk{P}{{A}{Q}}}}
  \tag{by \autoref{comp-ew-f}}
  \\
& \jdeq
  \subst{{\phi_0}{f}}{\ctxwk{\subst{\phi_0}{P}}{\subst{\phi_0}{\ctxwk{A}{Q}}}}
  \tag{by \autoref{comp-sw-f}}
  \\
& \jdeq
  \subst{{\phi_0}{f}}{\ctxwk{\subst{\phi_0}{P}}{Q}}
  \tag{by \autoref{cancellation-ws-f}}
  \\
& \jdeq
  \subst{{\phi_0}{f}}{\ctxwk{A}{Q}}
  \tag{by \autoref{empalg-eq1}}
  \\
& \jdeq
  \jcomp{}{\subst{\phi_0}{f}}{Q}.
  \tag{by definition}
\end{align*}
In the case $i=1$ we have the judgmental equalities
\begin{align*}
\subst{\phi_1}{\jcomp{}{f}{Q}}
& \jdeq
  \subst{\phi_1}{{f}{\ctxwk{\ctxext{A}{P}}{Q}}}
  \tag{by definition}
  \\
& \jdeq
  \subst{{\phi_1}{f}}{{\phi_1}{\ctxwk{\ctxext{A}{P}}{Q}}}
  \tag{by \autoref{comp-ss-f}}
  \\
& \jdeq
  \subst{{\phi_1}{f}}{{\phi_1}{\ctxwk{P}{{A}{Q}}}}
  \tag{by \autoref{comp-ew-f}}
  \\
& \jdeq
  \subst{{\phi_1}{f}}{\ctxwk{A}{Q}}
  \tag{by \autoref{cancellation-ws-f}}
  \\
& \jdeq
  \jcomp{}{\subst{\phi_1}{f}}{Q}.
  \tag{by definition}
\end{align*}
\end{proof}

As an immediate corollary, if we assume the judgmental equalities
\autoref{extempalg-eq1,extempalg-eq2} we get that 
\begin{equation}\label{cor:empalg-mor}
\jfameq{{\Gamma}{A}}{\subst{\phi_i}{\jcomp{}{\epsilon_0}{P}}}{P}
\end{equation}
and hence that $\subst{\phi_0}{\epsilon_1}$ is a 
morphism from $\ctxext{A}{P}$ to $A$. Thus, we can require
\begin{equation}\label{extempalg-eq3}
\jhomeq{\Gamma}{{A}{P}}{A}{\subst{\phi_0}{\epsilon_1}}{\epsilon_0}.
\end{equation}
For the final judgmental equality we need to explain the term
\begin{equation*}
\jterm
  {{{\Gamma}{A}}{\subst{\phi_1}{\jcomp{}{\epsilon_0}{P}}}}
  {\subst{\phi_1}{\ctxwk{\ctxext{P}{\jcomp{}{\epsilon_0}{P}}}{P}}}
  {\subst{\phi_1}{\epsilon_1}}.
\end{equation*}
We have already established that $\subst{\phi_1}{\jcomp{}{\epsilon_0}{P}}\jdeq
P$. We also see that 
\begin{align*}
\subst{\phi_1}{\ctxwk{\ctxext{P}{\jcomp{}{\epsilon_0}{P}}}{P}}
& \jdeq
  \subst{\phi_1}{\ctxwk{\jcomp{}{\epsilon_0}{P}}{{P}{P}}}
  \tag{by \autoref{comp-ew-f}}
  \\
& \jdeq
  \ctxwk{\subst{\phi_1}{\jcomp{}{\epsilon_0}{P}}}{\subst{\phi_1}{\ctxwk{P}{P}}}
  \tag{by \autoref{comp-sw-f}}
  \\
& \jdeq
  \ctxwk{P}{\subst{\phi_1}{\ctxwk{P}{P}}}
  \tag{by \autoref{cor:empalg-mor}}
  \\
& \jdeq
  \ctxwk{P}{P}
  \tag{by \autoref{cancellation-ws-f}}
\end{align*}
and we will therefore require that
\begin{equation}\label{extempalg-eq4}
\jtermeq{{{\Gamma}{A}}{P}}{\ctxwk{P}{P}}{\subst{\phi_1}{\epsilon_1}}{\idtm{P}}.
\end{equation}
We bring all this together in the definition of extension-empty algebras:

\begin{defn}
An \emph{extension-empty algebra in context $\Gamma$} 
is a sextuple $(A,P,\epsilon_0,\epsilon_1,\phi_0,\phi_1)$ for which 
the quadruple $(A,P,\epsilon_0,\epsilon_1)$ is an extension algebra in context 
$\Gamma$, the quadruple $(A,P,\phi_0,\phi_1)$ is an empty algebra in context
$\Gamma$, satisfying the judgmental equalities 
\autoref{extempalg-eq1,extempalg-eq2,extempalg-eq3,extempalg-eq4}.
\end{defn}

\subsection{Extension-weakening algebras}
The notion of (extension-)weakening algebra is dependent on the notion of extension algebra.
Although it is strictly speaking not dependent on the notion of empty-algebra,
we shall only formulate a weakening algebras in the setting of an
extension-empty algebras. When one wants to have a weakening operation which
also acts on the level of contexts in an extension algebra without empty
context, extra work has to be done to introduce these separately.

An extension-weakening algebra in context $\Gamma$ will be an octuple
\begin{equation*}
(A,P,\epsilon_0,\epsilon_1,\phi_0,\phi_1,\omega_0,\omega_1)
\end{equation*}
where $(A,P,\epsilon_0,\epsilon_1,\phi_0,\phi_1)$ is an extension-empty algebra
in context $\Gamma$ and where
\begin{align*}
\jhom*
  {{{\Gamma}{A}}{P}}
  {\ctxwk{P}{P}}
  {\jcomp{}{\epsilon_0}{P}}
  {\omega_0}
  \\
\jfhom*
  {{{\Gamma}{A}}{P}}
  {\ctxwk{P}{P}}
  {\jcomp{}{\epsilon_0}{P}}
  {\omega_0}
  {\ctxwk{P}{\jcomp{}{\epsilon_0}{P}}}
  {\jcomp{}{\epsilon_0}{\jcomp{}{\epsilon_0}{P}}}
  {\omega_1}
\end{align*}
satisfying nine additional judgmental equalities expressing that $\omega_0$
and $\omega_1$ are compatible with $\epsilon_1$, $\phi_1$
and with each other. Our current goal is to figure out what these are.

Before we go into that, we develop a bit of intuition by explaining how
$\omega_0$ and $\omega_1$ act when applied to a family $x_1:\subst{x_0}{P}$
over $x_0:A$. Note that
\begin{equation*}
\jterm{{\Gamma}{\subst{x_0}{P}}}{\subst{{x_1}{{x_0}{\epsilon_0}}}{P}}{\subst{x_1}{{x_0}{\omega_0}}},
\end{equation*}
so $\subst{x_1}{{x_0}{\omega_0}}$ takes families over $x_0$ to families over
the extended $\subst{x_1}{{x_0}{\epsilon_0}}$.

To compute the type of $\subst{x_2}{{x_1}{{x_0}{\omega_1}}}$ for families
$x_1,x_2:\subst{x_0}{P}$ over $x_0:A$ we have to do a
bit more work. Note that
\begin{align*}
& \subst{x_2}{{x_1}{{x_0}{\jcomp{}{\omega_0}{\jcomp{}{\epsilon_0}{\jcomp{}{\epsilon_0}{P}}}}}}
  \\
& \jdeq
  \subst{x_2}{{x_1}{{x_0}{\unfold{\jcomp{\ctxwk{P}{P}}{\omega_0}{\jcomp{}{\epsilon_0}{\jcomp{}{\epsilon_0}{P}}}}}}}
  \tag{by definition}
  \\
& \jdeq
  \subst
    { {x_2}
      { {x_1}
        { {x_0}
          {\omega_0}
          }
        }
      }
    { {x_2}
      { {x_1}
        { {x_0}
          {\ctxwk{{P}{P}}{\jcomp{}{\epsilon_0}{\jcomp{}{\epsilon_0}{P}}}}
          }
        }
      }
  \tag{by \autoref{comp-ss-f}}
  \\
& \jdeq
  \subst
    { {x_2}
      { {x_1}
        { {x_0}
          {\omega_0}
          }
        }
      }
    { {x_2}
      { \ctxwk
          {\subst{x_1}{\ctxwk{\subst{x_0}{P}}{\subst{x_0}{P}}}}
          {\subst{x_1}{{x_0}{\jcomp{}{\epsilon_0}{\jcomp{}{\epsilon_0}{P}}}}}
        }
      }
  \tag{by \autoref{comp-sw-f}}
  \\
& \jdeq
  \subst
    { {x_2}
      { {x_1}
        { {x_0}
          {\omega_0}
          }
        }
      }
    { {x_1}
      { {x_0}
        {\jcomp{}{\epsilon_0}{\jcomp{}{\epsilon_0}{P}}}
        }
      }
  \tag{by \autoref{cancellation-ws-f}}
  \\
& \jdeq
  \subst
    { {{x_2}{{x_1}{{x_0}{\omega_0}}}}
      {{{x_1}{{x_0}{\epsilon_0}}}{\epsilon_0}}
      }
    { P
      }.
  \tag{by \autoref{comp-ss-f}}
\end{align*}
Therefore, we see that
\begin{equation*}
\jterm
  {{\Gamma}{\subst{{x_2}{{x_0}{\epsilon_0}}}{P}}}
  { \subst
      { {{x_2}{{x_1}{{x_0}{\omega_0}}}}
        {{{x_1}{{x_0}{\epsilon_0}}}{\epsilon_0}}
        }
      { P
        }
    }
  { \subst{x_2}{{x_1}{{x_0}{\omega_1}}}
    }
\end{equation*}
Thus, the term $\subst{x_2}{{x_1}{{x_0}{\omega_1}}}$ takes a family over the
extended context $\subst{x_2}{{x_0}{\epsilon_0}}$ to a family over the
weakening $\subst{x_2}{{x_1}{{x_0}{\omega_0}}}$, which is itself a family
over $\subst{x_1}{{x_0}{\epsilon_0}}$. In other words, $\omega_1$
is precisely the internalization of the action on families of weakening, as
intended.

\subsubsection{The compatibility of weakening with extension}
The first rule expressing that weakening is compatible with extension is that
the diagram
\begin{equation}\label{wkalg-exteq1}
\begin{tikzcd}[column sep=large]
\ctxwk{P}{\ctxext{P}{\jcomp{}{\epsilon_0}{P}}}
  \ar{r}{\jvcomp{}{\omega_0}{\omega_1}}
  \ar{d}[swap]{\ctxwk{P}{\epsilon_1}}
& \ctxext
    {\jcomp{}{\epsilon_0}{P}}
    {\jcomp{}{\epsilon_0}{\jcomp{}{\epsilon_0}{P}}}
  \ar{d}{\jcomp{}{\epsilon_0}{\epsilon_1}}
  \\
\ctxwk{P}{P}
  \ar{r}[swap]{\omega_0}
& \jcomp{}{\epsilon_0}{P}
\end{tikzcd}
\end{equation}
commutes judgmentally.

\subsubsection{The compatibility of weakening with the empty context and family}
The first two judgmental equalities expressing that $\omega_0$ and $\omega_1$
are compatible with $\phi_1$ are easy to state:
\begin{align}
\jhomeq*{{\Gamma}{A}}{P}{P}{\subst{\phi_1}{\omega_0}}{\idtm{P}}\\
\jhomeq*{{\Gamma}{A}}{P}{P}{\subst{\phi_1}{\omega_1}}{\idtm{P}}
\end{align}

\subsubsection{The compatibility of weakening with itself}

\subsubsection{The definition of weakening algebras}

\subsection{Extension-substitution algebras}

\subsection{Pre-universes}
Pre-universes are internal versions of the theory of contexts, families and
terms. They interpret extension, the empty context, weakening, substitution
and identity terms all at once in a compatible way. Besides the compatibility
properties there will be judgmental equalities analoguous to the cancellation
properties of \autoref{cancellation-ws,cancellation-i}. Pre-universes are to
the theory of contexts, families and terms what internal categories to a
category.

### 2014-05-15 article-models-ttderived.tex
stated and proved lem:jcomp-higherjcomp

@@
 
+We end this subsection with a simple application of the above lemmas that is
+typical for the more general formulation of composition we have
+introduced. It is a special case of a consecutive composition where we consider
+a morphism $\jhom{\Gamma}{A}{B}{f}$ and a morphism
+$\jhom{{{\Gamma}{B}}{Q}}{R}{S}{g}$.
+
+\begin{lem}\label{lem:jcomp-higherjcomp}
+Let $\jhom{\Gamma}{A}{B}{f}$ and $\jhom{{{\Gamma}{B}}{Q}}{R}{S}{g}$ be context
+morphisms. Then the following inference rules are valid:
+\begin{align*}
+& \inference
+  { \jfam{{{{\Gamma}{B}}{Q}}{S}}{T}
+    }
+  { \jfameq
+      {{{{\Gamma}{A}}{\jcomp{A}{f}{Q}}}{\jcomp{A}{f}{R}}}
+      {\jcomp{A}{f}{\jcomp{R}{g}{T}}}
+      {\jcomp{\jcomp{A}{f}{R}}{\jcomp{A}{f}{g}}{\jcomp{A}{f}{T}}}
+    }  
+  \\
+& \inference
+  { \jfam{{{{{\Gamma}{B}}{Q}}{S}}{T}}{U}
+    }
+  { \jfameq
+      {{{{{\Gamma}{A}}{\jcomp{A}{f}{Q}}}{\jcomp{A}{f}{R}}}{\jcomp{A}{f}{\jcomp{R}{g}{T}}}}
+      {\jcomp{A}{f}{\jcomp{R}{g}{U}}}
+      {\jcomp{\jcomp{A}{f}{R}}{\jcomp{A}{f}{g}}{\jcomp{A}{f}{U}}}
+    }  
+  \\
+& \inference
+  { \jterm{{{{{\Gamma}{B}}{Q}}{S}}{T}}{U}{m}
+    }
+  { \jtermeq
+      {{{{{\Gamma}{A}}{\jcomp{A}{f}{Q}}}{\jcomp{A}{f}{R}}}{\jcomp{A}{f}{\jcomp{R}{g}{T}}}}
+      {\jcomp{A}{f}{\jcomp{R}{g}{U}}}
+      {\jcomp{A}{f}{\jcomp{R}{g}{m}}}
+      {\jcomp{\jcomp{A}{f}{R}}{\jcomp{A}{f}{g}}{\jcomp{A}{f}{m}}}
+    }  
+\end{align*}
+\end{lem}
+
+\begin{proof}
+We only prove the first inference rule. Let $\jfam{{{{\Gamma}{B}}{Q}}{S}}{T}$ be
+a family. Then we have the judgmental equalities
+\begin{align*}
+\jcomp{A}{f}{\jcomp{R}{g}{T}}
+& \jdeq
+  \jcomp{A}{f}{\subst{g}{\ctxwk{R}{T}}}
+  \tag{by definition}
+  \\
+& \jdeq
+  \subst{\jcomp{A}{f}{g}}{\jcomp{A}{f}{\ctxwk{R}{T}}}
+  \tag{by \autoref{lem:jcomp-subst}}
+  \\
+& \jdeq
+  \subst{\jcomp{A}{f}{g}}{\ctxwk{\jcomp{A}{f}{R}}{\jcomp{A}{f}{T}}}
+  \tag{by \autoref{lem:jcomp-wk}}
+  \\
+& \jdeq
+  \jcomp{\jcomp{A}{f}{R}}{\jcomp{A}{f}{g}}{\jcomp{A}{f}{T}}.
+  \tag{by definition}
+\end{align*}
+\end{proof}
+
 \subsection{Projections and extension on terms}\label{extension-on-terms}

### 2014-05-15 article-models-ttinternal.tex
stated and proved lem:jcomp-higherjcomp

@@
 \end{align*}
-satisfying nine additional judgmental equalities expressing that $\omega_0$
+satisfying five(?) additional judgmental equalities expressing that $\omega_0$
 and $\omega_1$ are compatible with $\epsilon_1$, $\phi_1$

### 2014-05-15 article-models-ttinternal.tex
stated and proved a fundamental lemma for wkalg-exteq2

@@
 
-\begin{lem}
+\begin{lem}\label{lem:extalg-twins}
 Suppose we have $A$, $P$, $\epsilon_0$ and $\epsilon_1$ satisfying \autoref{eq:ealg-eq1}. Then
@@
 \subsubsection{The compatibility of weakening with extension}
-The first rule expressing that weakening is compatible with extension is that
+The first rule, expressing that $\omega_0$ is compatible with $\epsilon_1$ is that
 the diagram
@@
 \end{equation}
-commutes judgmentally.
+commutes judgmentally. For the second rule, expressing that $\omega_1$ is 
+compatible with $\epsilon_1$, we wish to fill in a commutative diagram
+\begin{equation*}
+\begin{tikzcd}[column sep=huge]
+\ctxwk
+  {P}
+  { \ctxext
+      {\jcomp{}{\epsilon_0}{P}}
+      {\jcomp{}{\epsilon_0}{\jcomp{}{\epsilon_0}{P}}}
+    }
+  \ar[densely dotted]{r}{\jvcomp{}{\omega_1}{?}}
+  \ar{d}[swap]{\ctxwk{P}{\jcomp{}{\epsilon_0}{\epsilon_1}}}
+& \jcomp{}{\omega_0}{%
+    \jcomp{}{\epsilon_0}{%
+      \jcomp{}{\epsilon_0}{%
+        \ctxext{P}{\jcomp{}{\epsilon_0}{P}}
+        }
+      }
+    }
+  \ar{d}{ \jcomp{}{\omega_0}{%
+            \jcomp{}{\epsilon_0}{%
+              \jcomp{}{\epsilon_0}{\epsilon_1}
+              }
+            }
+          }
+  \\
+\ctxwk{P}{\jcomp{}{\epsilon_0}{P}}
+  \ar{r}[swap]{\omega_1}
+& \jcomp{}{\omega_0}{%
+    \jcomp{}{\epsilon_0}{%
+      \jcomp{}{\epsilon_0}{P}
+      }
+    }
+\end{tikzcd}
+\end{equation*}
+The morphism indicated by the question mark is a morphism from
+$\ctxwk{P}{\jcomp{}{\epsilon_0}{\jcomp{}{\epsilon_0}{P}}}$ to
+$%
+\jcomp{}{\omega_1}{%
+  \jcomp{}{\omega_0}{%
+    \jcomp{}{\epsilon_0}{%
+      \jcomp{}{\epsilon_0}{%
+        \ctxext{P}{\jcomp{}{\epsilon_0}{P}}
+        }
+      }
+    }
+  }
+$ in context $\ctxext{{\Gamma}{A}}{P}$. One would expect this morphism to be
+\begin{equation*}
+\begin{tikzcd}[column sep=large]
+\jcomp{}{\ctxwk{P}{\epsilon_0}}{\ctxwk{P}{\jcomp{}{\epsilon_0}{P}}}
+  \ar{r}{\jcomp{}{\ctxwk{P}{\epsilon_0}}{\omega_1}}
+& \jcomp{}{\ctxwk{P}{\epsilon_0}}{%
+    \jcomp{}{\omega_0}{%
+      \jcomp{}{\epsilon_0}{%
+        \jcomp{}{\epsilon_0}{%
+          P
+          }
+        }
+      }
+    }
+\end{tikzcd}
+\end{equation*}
+in the context $\ctxext{{\Gamma}{A}}{P}$, 
+so we must verify the judgmental equalities
+\begin{align*}
+\jcomp{}{\ctxwk{P}{\epsilon_0}}{\ctxwk{P}{\jcomp{}{\epsilon_0}{P}}}
+& \jdeq 
+  \ctxwk{P}{\jcomp{}{\epsilon_0}{\jcomp{}{\epsilon_0}{P}}}
+  \\
+\jcomp{}{\ctxwk{P}{\epsilon_0}}{%
+    \jcomp{}{\omega_0}{%
+      \jcomp{}{\epsilon_0}{%
+        \jcomp{}{\epsilon_0}{%
+          P
+          }
+        }
+      }
+    }
+& \jdeq
+  \jcomp{}{\omega_1}{%
+    \jcomp{}{\omega_0}{%
+      \jcomp{}{\epsilon_0}{%
+        \jcomp{}{\epsilon_0}{%
+          \ctxext{P}{\jcomp{}{\epsilon_0}{P}}
+          }
+        }
+      }
+    } 
+\end{align*}
+of families over the context $\ctxext{{\Gamma}{A}}{P}$. The first of these
+judgmental equalities is nothing special: it follows at straight from
+\autoref{comp-ws-f}. We prove the second equality in the following lemma.
+
+\begin{lem}
+Let $(A,P,\epsilon_0,\epsilon_1)$ be an extension algebra and let
+\begin{align*}
+\jhom*
+  {{{\Gamma}{A}}{P}}
+  {\ctxwk{P}{P}}
+  {\jcomp{}{\epsilon_0}{P}}
+  {\omega_0}
+  \\
+\jfhom*
+  {{{\Gamma}{A}}{P}}
+  {\ctxwk{P}{P}}
+  {\jcomp{}{\epsilon_0}{P}}
+  {\omega_0}
+  {\ctxwk{P}{\jcomp{}{\epsilon_0}{P}}}
+  {\jcomp{}{\epsilon_0}{\jcomp{}{\epsilon_0}{P}}}
+  {\omega_1}
+\end{align*}
+be morphisms satisfying \autoref{wkalg-exteq1}. Then the inference rules
+\begin{align*}
+& \inference
+  { \jfam{{\Gamma}{A}}{Q}
+    }
+  { \jfameq
+      {\blank}
+      { \jcomp{}{\omega_1}{%
+          \jcomp{}{\omega_0}{%
+            \jcomp{}{\epsilon_0}{%
+              \jcomp{}{\epsilon_0}{%
+                \jcomp{}{\epsilon_0}{Q}
+                }
+              }
+            }
+          } 
+        }
+      { \jcomp{}{\ctxwk{P}{\epsilon_0}}{%
+          \jcomp{}{\omega_0}{%
+            \jcomp{}{\epsilon_0}{%
+              \jcomp{}{\epsilon_0}{%
+                Q
+                }
+              }
+            }
+          }
+        }
+      }
+\end{align*}
+are valid.
+\end{lem}
+
+\begin{proof}
+We have the judgmental equalities
+\begin{align*}
+& \jcomp{}{\omega_1}{%
+          \jcomp{}{\omega_0}{%
+            \jcomp{}{\epsilon_0}{%
+              \jcomp{}{\epsilon_0}{%
+                \jcomp{}{\epsilon_0}{Q}
+                }
+              }
+            }
+          } 
+  \\
+& \jdeq
+  \jcomp{}{\jvcomp{}{\omega_0}{\omega_1}}{%
+            \jcomp{}{\epsilon_0}{%
+              \jcomp{}{\epsilon_0}{%
+                \jcomp{}{\epsilon_0}{Q}
+                }
+              }
+            }
+  \tag{by \autoref{lem:composition-threesome}}
+  \\
+& \jdeq
+  \jcomp{}{\jvcomp{}{\omega_0}{\omega_1}}{%
+            \jcomp{}{\epsilon_0}{%
+              \jcomp{}{\epsilon_1}{%
+                \jcomp{}{\epsilon_0}{Q}
+                }
+              }
+            }
+  \tag{by \autoref{lem:extalg-twins}}
+  \\
+& \jdeq
+  \jcomp{}{\jvcomp{}{\omega_0}{\omega_1}}{%
+            \jcomp{}{\jcomp{}{\epsilon_0}{\epsilon_1}}{%
+              \jcomp{}{\epsilon_0}{%
+                \jcomp{}{\epsilon_0}{Q}
+                }
+              }
+            }
+   \tag{by \autoref{lem:jcomp-higherjcomp}}
+   \\
+& \jdeq
+  \jcomp{}{\jcomp{}{%
+             \jvcomp{}{\omega_0}{\omega_1}}{%
+               \jcomp{}{\epsilon_0}{\epsilon_1}}}{%
+              \jcomp{}{\epsilon_0}{%
+                \jcomp{}{\epsilon_0}{Q}
+                }
+              }
+  \tag{by \autoref{lem:jcomp-jcomp}}
+  \\
+& \jdeq
+  \jcomp{}{\jcomp{}{\ctxwk{P}{\epsilon_0}}{\omega_0}}{%
+              \jcomp{}{\epsilon_0}{%
+                \jcomp{}{\epsilon_0}{Q}
+                }
+              }
+  \tag{by \autoref{wkalg-exteq1}}
+  \\
+& \jdeq
+  \jcomp{}{\ctxwk{P}{\epsilon_0}}{%
+    \jcomp{}{\omega_0}{%
+      \jcomp{}{\epsilon_0}{%
+        \jcomp{}{\epsilon_0}{%
+          Q
+          }
+        }
+      }
+    }
+  \tag{by \autoref{lem:jcomp-jcomp}}
+\end{align*}
+\end{proof}
 

### 2014-05-16 article-models-ttinternal.tex
formulated wkalg-exteq2
stated and proved thm:extalg-infty

@@
 following diagram commutes:
-\begin{equation}\label{eq:ealg-eq1}
+\begin{equation}\label{eq:extalg-eq1}
 \begin{tikzcd}[column sep=huge]
@@
 \begin{lem}
-Let $A$, $P$, $\epsilon_0$ and $\epsilon_1$ be as above, satisfying \autoref{eq:ealg-eq1} and
+Let $A$, $P$, $\epsilon_0$ and $\epsilon_1$ be as above, satisfying \autoref{eq:extalg-eq1} and
 let $x_0:A$,
@@
 \begin{lem}\label{lem:extalg-twins}
-Suppose we have $A$, $P$, $\epsilon_0$ and $\epsilon_1$ satisfying \autoref{eq:ealg-eq1}. Then
-we have the judgmental equality
-\begin{equation*}
-\jcomp{}{\epsilon_0}{\jcomp{}{\epsilon_0}{P}}
-  \jdeq
-  \jcomp{}{\epsilon_1}{\jcomp{}{\epsilon_0}{P}}.
-\end{equation*}
+Suppose we have $A$, $P$, $\epsilon_0$ and $\epsilon_1$ satisfying \autoref{eq:extalg-eq1}. Then
+the inference rules
+\begin{align*}
+& \inference
+  { \jfam{{\Gamma}{A}}{Q}
+    }
+  { \jfameq
+      {{{{\Gamma}{A}}{P}}{\jcomp{}{\epsilon_0}{P}}}
+      {\jcomp{}{\epsilon_0}{\jcomp{}{\epsilon_0}{Q}}}
+      {\jcomp{}{\epsilon_1}{\jcomp{}{\epsilon_0}{Q}}}
+    }
+  \\
+& \inference
+  { \jfam{{{\Gamma}{A}}{Q}}{R}
+    }
+  { \jfameq
+      {{{{{\Gamma}{A}}{P}}{\jcomp{}{\epsilon_0}{P}}}
+        {\jcomp{}{\epsilon_0}{\jcomp{}{\epsilon_0}{Q}}}}
+      {\jcomp{}{\epsilon_0}{\jcomp{}{\epsilon_0}{R}}}
+      {\jcomp{}{\epsilon_1}{\jcomp{}{\epsilon_0}{R}}}
+    }
+  \\
+& \inference
+  { \jterm{{{\Gamma}{A}}{Q}}{R}{h}
+    }
+  { \jtermeq
+      {{{{{\Gamma}{A}}{P}}{\jcomp{}{\epsilon_0}{P}}}
+        {\jcomp{}{\epsilon_0}{\jcomp{}{\epsilon_0}{Q}}}}
+      {\jcomp{}{\epsilon_0}{\jcomp{}{\epsilon_0}{R}}}
+      {\jcomp{}{\epsilon_0}{\jcomp{}{\epsilon_0}{h}}}
+      {\jcomp{}{\epsilon_1}{\jcomp{}{\epsilon_0}{h}}}
+    }
+\end{align*}
+are valid.
 \end{lem}
@@
 The proof is a rather long computation. It was found by starting with the
-application of \autoref{eq:ealg-eq1} and making the computation in both ways
-from there. The judgmental equality below where \autoref{eq:ealg-eq1} is applied,
+application of \autoref{eq:extalg-eq1} and making the computation in both ways
+from there. The judgmental equality below where \autoref{eq:extalg-eq1} is applied,
 is marked with an $*$. The computation goes as follows:
 \begin{align*}
-\jcomp{}{\epsilon_0}{\jcomp{}{\epsilon_0}{P}}
+\jcomp{}{\epsilon_0}{\jcomp{}{\epsilon_0}{Q}}
 & \jdeq
   \subst{\epsilon_0}
-    {\ctxwk{\ctxext{A}{P}}{\jcomp{}{\epsilon_0}{P}}}
+    {\ctxwk{\ctxext{A}{P}}{\jcomp{}{\epsilon_0}{Q}}}
   \\
@@
         {\jcomp{}{\epsilon_0}{P}}{\subst{\epsilon_0}
-        {\ctxwk{\ctxext{A}{P}}{\jcomp{}{\epsilon_0}{P}}}}}
+        {\ctxwk{\ctxext{A}{P}}{\jcomp{}{\epsilon_0}{Q}}}}}
   \\
@@
     { {\ctxwk{\jcomp{}{\epsilon_0}{P}}{\epsilon_0}}
-      {\ctxwk{\jcomp{}{\epsilon_0}{P}}{{\ctxext{A}{P}}{\jcomp{}{\epsilon_0}{P}}}}}
+      {\ctxwk{\jcomp{}{\epsilon_0}{P}}{{\ctxext{A}{P}}{\jcomp{}{\epsilon_0}{Q}}}}}
   \\
@@
     { {\ctxwk{\jcomp{}{\epsilon_0}{P}}{\epsilon_0}}
-      {\ctxwk{\ctxext{{A}{P}}{\jcomp{}{\epsilon_0}{P}}}{\jcomp{}{\epsilon_0}{P}}}}
+      {\ctxwk{\ctxext{{A}{P}}{\jcomp{}{\epsilon_0}{P}}}{\jcomp{}{\epsilon_0}{Q}}}}
   \\
@@
     { {\ctxwk{\jcomp{}{\epsilon_0}{P}}{\epsilon_0}}
-      {\ctxwk{\ctxext{A}{{P}{\jcomp{}{\epsilon_0}{P}}}}{\jcomp{}{\epsilon_0}{P}}}}
+      {\ctxwk{\ctxext{A}{{P}{\jcomp{}{\epsilon_0}{P}}}}{\jcomp{}{\epsilon_0}{Q}}}}
   \\
@@
     {\unfold{\jvcomp{\jcomp{}{\epsilon_0}{P}}{\epsilon_0}{\idtm{\jcomp{}{\epsilon_0}{P}}}}}
-    {\ctxwk{\ctxext{A}{{P}{\jcomp{}{\epsilon_0}{P}}}}{\jcomp{}{\epsilon_0}{P}}}
+    {\ctxwk{\ctxext{A}{{P}{\jcomp{}{\epsilon_0}{P}}}}{\jcomp{}{\epsilon_0}{Q}}}
   \\
@@
     {\jvcomp{}{\epsilon_0}{\idtm{\jcomp{}{\epsilon_0}{P}}}}
-    {\ctxwk{\ctxext{A}{{P}{\jcomp{}{\epsilon_0}{P}}}}{\jcomp{}{\epsilon_0}{P}}}
-  \\
-& \jdeq
-  \jcomp{}{\jvcomp{}{\epsilon_0}{\idtm{\jcomp{}{\epsilon_0}{P}}}}{\jcomp{}{\epsilon_0}{P}}
-  \\
-& \jdeq
-  \jcomp{}{\jcomp{}{\jvcomp{}{\epsilon_0}{\idtm{\jcomp{}{\epsilon_0}{P}}}}{\epsilon_0}}{P}
+    {\ctxwk{\ctxext{A}{{P}{\jcomp{}{\epsilon_0}{P}}}}{\jcomp{}{\epsilon_0}{Q}}}
+  \\
+& \jdeq
+  \jcomp{}{\jvcomp{}{\epsilon_0}{\idtm{\jcomp{}{\epsilon_0}{P}}}}{\jcomp{}{\epsilon_0}{Q}}
+  \\
+& \jdeq
+  \jcomp{}{\jcomp{}{\jvcomp{}{\epsilon_0}{\idtm{\jcomp{}{\epsilon_0}{P}}}}{\epsilon_0}}{Q}
   \\
 & \stackrel{*}{\jdeq}
-  \jcomp{}{
-    \jcomp{}{\jvcomp{}{\idtm{A}}{\epsilon_1}}{\epsilon_0}}{P}
+  \jcomp{}{\jcomp{}{\jvcomp{}{\idtm{A}}{\epsilon_1}}{\epsilon_0}}{Q}
   \\
@@
   \jcomp{}{\jvcomp{}{\idtm{A}}{\epsilon_1}}{
-    \jcomp{}{\epsilon_0}{P}}
+    \jcomp{}{\epsilon_0}{Q}}
   \\
@@
     {\tmext{\ctxwk{\ctxext{P}{\jcomp{}{\epsilon_0}{P}}}{\idtm{A}}}{\epsilon_1}}
-    {\ctxwk{\ctxext{A}{{P}{\jcomp{}{\epsilon_0}{P}}}}{\jcomp{}{\epsilon_0}{P}}}
+    {\ctxwk{\ctxext{A}{{P}{\jcomp{}{\epsilon_0}{P}}}}{\jcomp{}{\epsilon_0}{Q}}}
   \\
@@
     { {\ctxwk{\ctxext{P}{\jcomp{}{\epsilon_0}{P}}}{\idtm{A}}}
-      {\ctxwk{\ctxext{A}{{P}{\jcomp{}{\epsilon_0}{P}}}}{\jcomp{}{\epsilon_0}{P}}}
+      {\ctxwk{\ctxext{A}{{P}{\jcomp{}{\epsilon_0}{P}}}}{\jcomp{}{\epsilon_0}{Q}}}
       }
@@
     { {\cprojfstf{A}{\ctxext{P}{\jcomp{}{\epsilon_0}{P}}}}
-      {\ctxwk{\ctxext{A}{{P}{\jcomp{}{\epsilon_0}{P}}}}{\jcomp{}{\epsilon_0}{P}}}
+      {\ctxwk{\ctxext{A}{{P}{\jcomp{}{\epsilon_0}{P}}}}{\jcomp{}{\epsilon_0}{Q}}}
       }
@@
     {\epsilon_1}
-    {\jcomp{}{\cprojfstf{A}{\ctxext{P}{\jcomp{}{\epsilon_0}{P}}}}{\jcomp{}{\epsilon_0}{P}}}
+    {\jcomp{}{\cprojfstf{A}{\ctxext{P}{\jcomp{}{\epsilon_0}{P}}}}{\jcomp{}{\epsilon_0}{Q}}}
   \\
@@
     {\epsilon_1}
-    {\ctxwk{\ctxext{P}{\jcomp{}{\epsilon_0}{P}}}{\jcomp{}{\epsilon_0}{P}}}
-  \\
-& \jdeq
-  \jcomp{}{\epsilon_1}{\jcomp{}{\epsilon_0}{P}}.
+    {\ctxwk{\ctxext{P}{\jcomp{}{\epsilon_0}{P}}}{\jcomp{}{\epsilon_0}{Q}}}
+  \\
+& \jdeq
+  \jcomp{}{\epsilon_1}{\jcomp{}{\epsilon_0}{Q}}.
   \qedhere
@@
 Thus, the second judgmental equality we will need is that the diagram
-\begin{equation}\label{eq:ealg-eq2}
+\begin{equation}\label{eq:extalg-eq2}
 \begin{tikzcd}[column sep=huge]
@@
     \jvcomp{}{\epsilon_1}{\idtm{\jcomp{}{\epsilon_1}{{}{\epsilon_0}{P}}}}
-%    \tmext
-%      {\ctxwk{\jcomp{}{\bar{e}}{{}{e}{P}}}{f}}
-%      {\idtm{\jcomp{}{\bar{e}}{{}{e}{P}}}}
     }
@@
 $\ctxext{\Gamma}{A}$, satisfying the judgmental equalities of
-\autoref{eq:ealg-eq1,eq:ealg-eq2}.
+\autoref{eq:extalg-eq1,eq:extalg-eq2}.
 \end{defn}
 
-We also give a bit of intuition to the requirement of \autoref{eq:ealg-eq2} by
+We also give a bit of intuition to the requirement of \autoref{eq:extalg-eq2} by
 means of the following lemma.
@@
 \end{align*}
+\end{proof}
+
+The following lemma explains how each extension algebra gives rise to infinitely
+many extension algebras.
+
+\begin{thm}
+Suppose that $(A,P,\epsilon_0,\epsilon_1)$ is an extension algebra in context
+$\Gamma$. Then 
+$(P,\jcomp{}{\epsilon_0}{P},\epsilon_1,\jcomp{}{\epsilon_0}{\epsilon_1})$
+is an extension algebra in context $\ctxext{\Gamma}{A}$.
+\end{thm}
+
+\begin{proof}
+We first need to verify that $\jcomp{}{\epsilon_0}{\epsilon_1}$ is a morphism
+from $\ctxext{\jcomp{}{\epsilon_0}{P}}{\jcomp{}{\epsilon_1}{{}{\epsilon_0}{P}}}$
+to $\jcomp{}{\epsilon_0}{P}$. This follows from the judgmental equality
+$\jcomp{}{\epsilon_1}{{}{\epsilon_0}{P}}\jdeq
+\jcomp{}{\epsilon_0}{{}{\epsilon_0}{P}}$, which we have proved in
+\autoref{lem:extalg-twins}. Notice how the diagram in \autoref{eq:extalg-eq2} is
+of exactly the right sort, so the quadruple
+$(P,\jcomp{}{\epsilon_0}{P},\epsilon_1,\jcomp{}{\epsilon_0}{\epsilon_1})$
+satisfies its version of \autoref{eq:extalg-eq1}. It is left to verify that the diagram
+\begin{equation*}
+\begin{tikzcd}[column sep=huge]
+\ctxext
+  { \jcomp{}{\epsilon_0}{P}
+    }
+  { { \jcomp{}{\epsilon_1}{%
+        \jcomp{}{\epsilon_0}{P}
+        }
+      }
+    { \jcomp{}{\epsilon_1}{%
+        \jcomp{}{\epsilon_1}{%
+          \jcomp{}{\epsilon_0}{P}
+          }
+        }
+      }
+    } 
+  \ar{r}
+    { \jvcomp{}{\idtm{\jcomp{}{\epsilon_0}{P}}}{%
+        \jcomp{}{\epsilon_1}{%
+          \jcomp{}{\epsilon_0}{\epsilon_1}}}}
+  \ar{d}[swap]{
+    \jvcomp{}{\jcomp{}{\epsilon_0}{\epsilon_1}}{%
+      \idtm{
+        \jcomp{}{\jcomp{}{\epsilon_0}{\epsilon_1}}{%
+          \jcomp{}{\epsilon_1}{%
+            \jcomp{}{\epsilon_0}{P}
+            }
+          }
+        }
+      }
+    }
+& \ctxext
+    {\jcomp{}{\epsilon_0}{P}}
+    {\jcomp{}{\epsilon_1}{\jcomp{}{\epsilon_0}{P}}} 
+  \ar{d}{\jcomp{}{\epsilon_0}{\epsilon_1}}
+  \\
+\ctxext
+  {\jcomp{}{\epsilon_0}{P}}
+  {\jcomp{}{\epsilon_1}{\jcomp{}{\epsilon_0}{P}}} 
+  \ar{r}[swap]{\jcomp{}{\epsilon_0}{\epsilon_1}} 
+& \jcomp{}{\epsilon_0}{P}
+\end{tikzcd}
+\end{equation*}
+commutes judgmentally; this diagram is the version of \autoref{eq:extalg-eq2}
+for the quadruple
+$(P,\jcomp{}{\epsilon_0}{P},\epsilon_1,\jcomp{}{\epsilon_0}{\epsilon_1})$. Note
+that this follows from \autoref{eq:extalg-eq2} provided that we can show that
+\begin{align}
+\jvcomp{}{\idtm{\jcomp{}{\epsilon_0}{P}}}{%
+  \jcomp{}{\epsilon_1}{%
+    \jcomp{}{\epsilon_0}{\epsilon_1}}}
+& \jdeq
+  \jcomp{}{\epsilon_0}{%
+    \jvcomp{}{\idtm{P}}{\jcomp{}{\epsilon_0}{\epsilon_1}}
+    }
+  \label{eq:extalg-infty1}
+  \\
+\jvcomp{}{\jcomp{}{\epsilon_0}{\epsilon_1}}{%
+  \idtm{
+    \jcomp{}{\jcomp{}{\epsilon_0}{\epsilon_1}}{%
+      \jcomp{}{\epsilon_1}{%
+        \jcomp{}{\epsilon_0}{P}
+        }
+      }
+    }
+  }
+& \jdeq
+\jcomp{}{\epsilon_0}{%
+  \jvcomp{}{\epsilon_1}{%
+    \idtm{
+      \jcomp{}{\epsilon_1}{%
+        \jcomp{}{\epsilon_0}{P}
+        }
+      }
+    }
+  }
+  \label{eq:extalg-infty2}
+\end{align}
+Note that \autoref{eq:extalg-infty1} follows if we can show that
+\begin{equation*}
+\jcomp{}{\epsilon_1}{\jcomp{}{\epsilon_0}{\epsilon_1}}
+  \jdeq
+  \jcomp{}{\epsilon_0}{\jcomp{}{\epsilon_0}{\epsilon_1}}.
+\end{equation*}
+This is a special case of \autoref{lem:extalg-twins}. The second judgmental
+equality, \autoref{eq:extalg-infty2}, is trivial.
 \end{proof}
@@
 
+This enables us to require that the diagram
+\begin{equation}\label{wkalg-exteq2}
+\begin{tikzcd}[column sep=huge]
+\ctxwk
+  {P}
+  { \ctxext
+      {\jcomp{}{\epsilon_0}{P}}
+      {\jcomp{}{\epsilon_0}{\jcomp{}{\epsilon_0}{P}}}
+    }
+  \ar{r}{\jvcomp{}{\omega_1}{\jcomp{}{\ctxwk{P}{\epsilon_0}}{\omega_1}}}
+  \ar{d}[swap]{\ctxwk{P}{\jcomp{}{\epsilon_0}{\epsilon_1}}}
+& \jcomp{}{\omega_0}{%
+    \jcomp{}{\epsilon_0}{%
+      \jcomp{}{\epsilon_0}{%
+        \ctxext{P}{\jcomp{}{\epsilon_0}{P}}
+        }
+      }
+    }
+  \ar{d}{ \jcomp{}{\omega_0}{%
+            \jcomp{}{\epsilon_0}{%
+              \jcomp{}{\epsilon_0}{\epsilon_1}
+              }
+            }
+          }
+  \\
+\ctxwk{P}{\jcomp{}{\epsilon_0}{P}}
+  \ar{r}[swap]{\omega_1}
+& \jcomp{}{\omega_0}{%
+    \jcomp{}{\epsilon_0}{%
+      \jcomp{}{\epsilon_0}{P}
+      }
+    }
+\end{tikzcd}
+\end{equation}
+commutes judgmentally.
+
 \subsubsection{The compatibility of weakening with the empty context and family}

### 2014-05-16 article-models-ttinternal.tex
given examples, defined homomorphisms

@@
 \end{proof}
+
+There is a trivial class of examples of extension algebras we can give right
+away. More examples will be introduced by universes, later on.
+
+\begin{eg}
+Let $A$ be a family in context $\Gamma$. Then the quadruple
+\begin{equation*}
+(A,\emptyf,\idtm{A},\emptytm)
+\end{equation*}
+is an extension algebra in context $\Gamma$, as is the quadruple
+\begin{equation*}
+(\emptyf,A,\emptytm,\idtm{A}).
+\end{equation*}
+Also, the quadruple
+\begin{equation*}
+(A,\ctxwk{A}{A},\cprojfstf{A}{\ctxwk{A}{A}},\cprojfstf{\ctxwk{A}}{\ctxwk{A}{{A}{A}}})
+\end{equation*}
+is an extension algebra in context $\Gamma$.
+\end{eg}
 
@@
 \end{proof}
+
+\begin{defn}
+An extension algebra homomorphism from $(A,P,\epsilon_0,\epsilon_1)$ to
+$(B,Q,\eta_0,\eta_1)$ in context $\Gamma$ is a pair $(f_0,f_1)$ consisting of
+\begin{align*}
+\jhom*{\Gamma}{A}{B}{f_0}
+  \\
+\jfhom*{\Gamma}{A}{B}{f_0}{P}{Q}{f_1}
+\end{align*}
+for which the diagrams
+\begin{equation}\label{eq:extalg-hom1}
+\begin{tikzcd}
+\ctxext{A}{P}
+  \ar{r}{\jvcomp{}{f_0}{f_1}}
+  \ar{d}[swap]{\epsilon_0}
+& \ctxext{B}{Q}
+  \ar{d}{\eta_0}
+  \\
+A
+  \ar{r}[swap]{f_0}
+& B
+\end{tikzcd}
+\end{equation}
+and
+\begin{equation}\label{eq:extalg-hom2}
+\begin{tikzcd}[column sep=large]
+\ctxext{P}{\jcomp{}{\epsilon_0}{P}}
+  \ar{r}{\jvcomp{}{f_1}{\jcomp{}{\epsilon_0}{f_1}}}
+  \ar{d}[swap]{\epsilon_1}
+& \ctxext{Q}{\jcomp{}{\eta_0}{Q}}
+  \ar{d}{\eta_1}
+  \\
+P
+  \ar{r}[swap]{f_1}
+& Q
+\end{tikzcd}
+\end{equation}
+commute judgmentally.
+\end{defn}
+
+\begin{rmk}
+I suspect that if we copy this theory of extension algebras to Martin-L\"of
+type theory, with the judgmental equalities replaced by identifications, with
+dependent pair types rather than those strict extensions, etcetera, then
+the type of $f_1$ for which these two diagrams commute is a mere proposition.
+\end{rmk}
 

### 2014-05-16 article-models-ttinternal.tex
corrected the definition of extension algebra homomorphism

@@
 \end{proof}
+
+\begin{thm}
+Let $(Q,R,\eta_0,\eta_1)$ be an extension algebra in context $\ctxext{\Gamma}{B}$ and let
+$\jfam{\Gamma}{A}$. Then the quadruple
+\begin{equation*}
+(\ctxwk{A}{Q},\ctxwk{A}{R},\ctxwk{A}{\eta_0},\ctxwk{A}{\eta_1})
+\end{equation*}
+is an extension algebra in context $\ctxext{{\Gamma}{A}}{\ctxwk{A}{B}}$.
+\end{thm}
+
+\begin{thm}
+Let $(Q,R,\eta_0,\eta_1)$ be an extension algebra in context $\ctxext{{\Gamma}{A}}{P}$
+and let $\jterm{\Gamma}{A}{x}$. Then the quadruple
+\begin{equation*}
+(\subst{x}{Q},\subst{x}{R},\subst{x}{\eta_0},\subst{x}{\eta_1})
+\end{equation*}
+is an extension algebra in context $\ctxext{\Gamma}{\subst{x}{P}}$.
+\end{thm}
+
+\begin{cor}
+Let $(Q,R,\eta_0,\eta_1)$ be an extension algebra in context $\ctxext{\Gamma}{B}$
+and let $\jhom{\Gamma}{A}{B}{f}$. Then the quadruple
+\begin{equation*}
+(\jcomp{A}{f}{Q},\jcomp{A}{f}{R},\jcomp{A}{f}{\eta_0},\jcomp{A}{f}{\eta_1})
+\end{equation*}
+is an extension algebra in context $\ctxext{\Gamma}{A}$.
+\end{cor}
 
@@
   \ar{d}[swap]{\epsilon_1}
-& \ctxext{Q}{\jcomp{}{\eta_0}{Q}}
-  \ar{d}{\eta_1}
+& \jcomp{}{f_0}{\ctxext{Q}{\jcomp{}{\eta_0}{Q}}}
+  \ar{d}{\jcomp{}{f_0}{\eta_1}}
   \\
@@
   \ar{r}[swap]{f_1}
-& Q
+& \jcomp{}{f_0}{Q}
 \end{tikzcd}
@@
 \end{defn}
+
+\begin{rmk}
+To see that the upper morphism in the diagram of \autoref{eq:extalg-hom2} has
+indeed the indicated codomain provided that the diagram of \autoref{eq:extalg-hom1}
+commutes judgmentally, note that we have the judgmental equalities
+\begin{align*}
+\jcomp{}{f_1}{\jcomp{}{f_0}{\jcomp{}{\eta_0}{Q}}}
+& \jdeq 
+  \jcomp{}{\jvcomp{}{f_0}{f_1}}{\jcomp{}{\eta_0}{Q}}
+  \tag{by \autoref{lem:composition-threesome}}
+  \\
+& \jdeq
+  \jcomp{}{\jcomp{}{\jvcomp{}{f_0}{f_1}}{\eta_0}}{Q}
+  \tag{by \autoref{lem:jcomp-jcomp}}
+  \\
+& \jdeq
+  \jcomp{}{\jcomp{}{\epsilon_0}{f_0}}{Q}
+  \tag{by \autoref{eq:extalg-hom1}}
+  \\
+& \jdeq
+  \jcomp{}{\epsilon_0}{\jcomp{}{f_0}{Q}}.
+  \tag{by \autoref{lem:jcomp-jcomp}}
+\end{align*}
+and we indeed have the morphism $\jcomp{}{\epsilon_0}{f_1}$ from 
+$\jcomp{}{\epsilon_0}{P}$ to $\jcomp{}{\epsilon_0}{\jcomp{}{f_0}{Q}}$.
+\end{rmk}
 
@@
 the type of $f_1$ for which these two diagrams commute is a mere proposition.
+
+With this notion of morphism, a term of an extension
+algebra $(A,P,\epsilon_0,\epsilon_1)$ is a pair $(x_0,x_1)$ such that
+$\subst{x_1}{{x_0}{\epsilon_0}}\jdeq x_0$.
 \end{rmk}

### 2014-05-16 article-models-ttinternal.tex
small improvements

@@
 \begin{defn}
-An extension algebra homomorphism from $(A,P,\epsilon_0,\epsilon_1)$ to
+An \emph{extension homomorphism} from $(A,P,\epsilon_0,\epsilon_1)$ to
 $(B,Q,\eta_0,\eta_1)$ in context $\Gamma$ is a pair $(f_0,f_1)$ consisting of

### 2014-05-16 article-models-ttderived.tex
stated and proved a fundamental lemma for wkalg-exteq2
changed a word

@@
 \begin{lem}\label{lem:jcomp-const}
-The inference
+The inference rule
 \begin{align*}
@@
 We end this subsection with a simple application of the above lemmas that is
-typical for the more general formulation of composition we have
+illustrative for the more general formulation of composition we have
 introduced. It is a special case of a consecutive composition where we consider

### 2014-05-16 article-models-ttinternal.tex
shortened the long proof

@@
 \begin{proof}
-The proof is a rather long computation. It was found by starting with the
-application of \autoref{eq:extalg-eq1} and making the computation in both ways
-from there. The judgmental equality below where \autoref{eq:extalg-eq1} is applied,
-is marked with an $*$. The computation goes as follows:
+We only prove the first inference rule. Let $\jfam{{\Gamma}{A}}{Q}$ be a family.
+Then we have the judgmental equalities
 \begin{align*}
@@
 & \jdeq
-  \subst{\epsilon_0}
-    {\ctxwk{\ctxext{A}{P}}{\jcomp{}{\epsilon_0}{Q}}}
-  \\
-& \jdeq
-  \subst
-    {\idtm{\jcomp{}{\epsilon_0}{P}}}
-    { \ctxwk
-        {\jcomp{}{\epsilon_0}{P}}{\subst{\epsilon_0}
-        {\ctxwk{\ctxext{A}{P}}{\jcomp{}{\epsilon_0}{Q}}}}}
-  \\
-& \jdeq
-  \subst
-    {\idtm{\jcomp{}{\epsilon_0}{P}}}
-    { {\ctxwk{\jcomp{}{\epsilon_0}{P}}{\epsilon_0}}
-      {\ctxwk{\jcomp{}{\epsilon_0}{P}}{{\ctxext{A}{P}}{\jcomp{}{\epsilon_0}{Q}}}}}
-  \\
-& \jdeq
-  \subst
-    {\idtm{\jcomp{}{\epsilon_0}{P}}}
-    { {\ctxwk{\jcomp{}{\epsilon_0}{P}}{\epsilon_0}}
-      {\ctxwk{\ctxext{{A}{P}}{\jcomp{}{\epsilon_0}{P}}}{\jcomp{}{\epsilon_0}{Q}}}}
-  \\
-& \jdeq
-  \subst
-    {\idtm{\jcomp{}{\epsilon_0}{P}}}
-    { {\ctxwk{\jcomp{}{\epsilon_0}{P}}{\epsilon_0}}
-      {\ctxwk{\ctxext{A}{{P}{\jcomp{}{\epsilon_0}{P}}}}{\jcomp{}{\epsilon_0}{Q}}}}
-  \\
-& \jdeq
-  \subst
-    {\unfold{\jvcomp{\jcomp{}{\epsilon_0}{P}}{\epsilon_0}{\idtm{\jcomp{}{\epsilon_0}{P}}}}}
-    {\ctxwk{\ctxext{A}{{P}{\jcomp{}{\epsilon_0}{P}}}}{\jcomp{}{\epsilon_0}{Q}}}
-  \\
-& \jdeq
-  \subst
-    {\jvcomp{}{\epsilon_0}{\idtm{\jcomp{}{\epsilon_0}{P}}}}
-    {\ctxwk{\ctxext{A}{{P}{\jcomp{}{\epsilon_0}{P}}}}{\jcomp{}{\epsilon_0}{Q}}}
+  \jcomp{}{\idtm{\jcomp{}{\epsilon_0}{P}}}{\jcomp{}{\epsilon_0}{\jcomp{}{\epsilon_0}{Q}}}
+  \tag{by \autoref{idfunc-precomp}}
   \\
@@
   \jcomp{}{\jvcomp{}{\epsilon_0}{\idtm{\jcomp{}{\epsilon_0}{P}}}}{\jcomp{}{\epsilon_0}{Q}}
+  \tag{by \autoref{lem:composition-threesome}}
   \\
@@
   \jcomp{}{\jcomp{}{\jvcomp{}{\epsilon_0}{\idtm{\jcomp{}{\epsilon_0}{P}}}}{\epsilon_0}}{Q}
-  \\
-& \stackrel{*}{\jdeq}
+  \tag{by \autoref{lem:jcomp-jcomp}}
+  \\
+& \jdeq
   \jcomp{}{\jcomp{}{\jvcomp{}{\idtm{A}}{\epsilon_1}}{\epsilon_0}}{Q}
-  \\
-& \jdeq
-  \jcomp{}{\jvcomp{}{\idtm{A}}{\epsilon_1}}{
-    \jcomp{}{\epsilon_0}{Q}}
-  \\
-& \jdeq
-  \subst
-    {\tmext{\ctxwk{\ctxext{P}{\jcomp{}{\epsilon_0}{P}}}{\idtm{A}}}{\epsilon_1}}
-    {\ctxwk{\ctxext{A}{{P}{\jcomp{}{\epsilon_0}{P}}}}{\jcomp{}{\epsilon_0}{Q}}}
-  \\
-& \jdeq
-  \subst
-    {\epsilon_1}
-    { {\ctxwk{\ctxext{P}{\jcomp{}{\epsilon_0}{P}}}{\idtm{A}}}
-      {\ctxwk{\ctxext{A}{{P}{\jcomp{}{\epsilon_0}{P}}}}{\jcomp{}{\epsilon_0}{Q}}}
-      }
-  \\
-& \jdeq
-  \subst
-    {\epsilon_1}
-    { {\cprojfstf{A}{\ctxext{P}{\jcomp{}{\epsilon_0}{P}}}}
-      {\ctxwk{\ctxext{A}{{P}{\jcomp{}{\epsilon_0}{P}}}}{\jcomp{}{\epsilon_0}{Q}}}
-      }
-  \\
-& \jdeq
-  \subst
-    {\epsilon_1}
-    {\jcomp{}{\cprojfstf{A}{\ctxext{P}{\jcomp{}{\epsilon_0}{P}}}}{\jcomp{}{\epsilon_0}{Q}}}
-  \\
-& \jdeq
-  \subst
-    {\epsilon_1}
-    {\ctxwk{\ctxext{P}{\jcomp{}{\epsilon_0}{P}}}{\jcomp{}{\epsilon_0}{Q}}}
+  \tag{by \autoref{eq:extalg-eq1}}
+  \\
+& \jdeq
+  \jcomp{}{\jvcomp{}{\idtm{A}}{\epsilon_1}}{\jcomp{}{\epsilon_0}{Q}}
+  \tag{by \autoref{lem:jcomp-jcomp}}
+  \\
+& \jdeq
+  \jcomp{}{\epsilon_1}{\jcomp{}{\idtm{A}}{\jcomp{}{\epsilon_0}{Q}}}
+  \tag{by \autoref{lem:composition-threesome}}
   \\
@@
   \jcomp{}{\epsilon_1}{\jcomp{}{\epsilon_0}{Q}}.
-  \qedhere
+  \tag{by \autoref{idfunc-precomp}}
 \end{align*}

### 2014-05-16 article-models-ttinternal.tex
shortened the long proof

@@
 satisfies its version of \autoref{eq:extalg-eq1}. It is left to verify that the diagram
+\begin{small}
 \begin{equation*}
@@
 \end{equation*}
+\end{small}
 commutes judgmentally; this diagram is the version of \autoref{eq:extalg-eq2}

### 2014-05-16 article-models-ttinternal.tex
starting to work on compatibility of weakening with itself

@@
 
+\begin{proof}
+The proof follows from the fact that weakening by $A$ is compatible with all
+the involved operations.
+\end{proof}
+
 \begin{thm}
@@
 \end{thm}
+
+\begin{proof}
+The proof follows from the fact that substitution with $x$ is compatible with
+all the involved operations.
+\end{proof}
 
@@
 \end{align}
+\end{defn}
+
+\begin{defn}
+Let $(A,P,\phi_0,\phi_1)$ and $(B,Q,\psi_0,\psi_1)$ be empty algebras in context
+$\Gamma$. An empty homomorphism from $(A,P,\phi_0,\phi_1)$ to $(B,Q,\psi_0,\psi_1)$
+is a pair $(f_0,f_1)$ consisting of
+\begin{align*}
+\jhom*{\Gamma}{A}{B}{f_0}\\
+\jfhom*{\Gamma}{A}{B}{f}{P}{Q}{f_1}
+\end{align*}
+satisfying the judgmental equalities
+\begin{align*}
+\jtermeq*{\Gamma}{B}{\subst{\phi_0}{f_0}}{\psi_0}\\
+\jhomeq*{\Gamma}{A}{B}{\subst{\phi_0}{f_1}}{f_0}\\
+\jtermeq*{{\Gamma}{A}}{\jcomp{}{f_0}{Q}}{\subst{\phi_1}{f_1}}{\jcomp{}{f_0}{\psi_1}}.
+\end{align*}
 \end{defn}
@@
 \end{align*}
+such that $(\omega_0,\omega_1)$ is both an extension homomorphism
+\begin{equation*}
+\begin{tikzcd}
+( \ctxwk{P}{P},
+  \ctxwk{P}{\jcomp{}{\epsilon_0}{P}},
+  \ctxwk{P}{\epsilon_1},
+  \ctxwk{P}{\jcomp{}{\epsilon_0}{\epsilon_1}})
+\ar{r}{(\omega_0,\omega_1)}
+& ( \jcomp{}{\epsilon_0}{P},
+  \jcomp{}{\epsilon_0}{\jcomp{}{\epsilon_0}{P}},
+  \jcomp{}{\epsilon_0}{\epsilon_1},
+  \jcomp{}{\epsilon_0}{\jcomp{}{\epsilon_0}{\epsilon_1}})
+\end{tikzcd}
+\end{equation*}
+and an empty homomorphism
+
 satisfying five(?) additional judgmental equalities expressing that $\omega_0$
@@
 \subsubsection{The compatibility of weakening with itself}
+To express the compatibility of weakening with itself, we must fill in the
+following diagram:
+\begin{equation*}
+\begin{tikzcd}
+\ctxwk{P}{{P}{P}}
+  \ar{r}{\ctxwk{P}{\omega_0}}
+  \ar{d}[swap]{\ctxwk{{P}{P}}{\omega_0}}
+& \ctxwk{P}{\jcomp{}{\epsilon_0}{P}}
+  \ar{d}{\omega_1}
+  \\
+\ctxwk{{P}{P}}{\jcomp{}{\epsilon_0}{P}}
+  \ar[densely dotted]{r}
+& \jcomp{}{\omega_0}{\jcomp{}{\epsilon_0}{\jcomp{}{\epsilon_0}{P}}}
+\end{tikzcd}
+\end{equation*}
 

### 2014-05-17 article-models-ttderived.tex
proved associativity of vertical composition

@@
 
-\begin{lem}\label{lem:tmext-ext}
+\begin{lem}\label{lem:tmext-ext}\label{lem:tmext-tmext}
 The following compatibility rules for two consecutive term extensions are valid:
@@
 
-\begin{lem}
+\begin{lem}\label{lem:jfcomp-jfcomp}
 Horizontal composition of morphisms over morphisms is associative, i.e.~the
@@
 
+Likewise, vertical composition is associative. For the following lemma, keep in
+mind that the inference rules
+\begin{align*}
+& \inference
+  { \jfhom{\Gamma}{{A}{P}}{{B}{Q}}{\jvcomp{P}{f_0}{f_1}}{R}{S}{f_2}
+    }
+  { \jfhom{{\Gamma}{A}}{P}{\jcomp{A}{f_0}{Q}}{f_1}{R}{\jcomp{A}{f_0}{S}}{f_2}
+    }
+  \\
+& \inference
+  { \jfhom{{\Gamma}{A}}{P}{\jcomp{A}{f_0}{Q}}{f_1}{R}{\jcomp{A}{f_0}{S}}{f_2}
+    }
+  { \jfhom{\Gamma}{{A}{P}}{{B}{Q}}{\jvcomp{P}{f_0}{f_1}}{R}{S}{f_2}
+    }
+\end{align*}
+are valid. Thus, a morphism over the vertical composition of $f_1$ over $f_0$
+is the same thing as a morphism over $f_1$ in the appropriate sense. Thus, we
+have vertical composition also available to compose $f_2$ with $f_1$ vertically
+and we don't need to introduce a seperate `vertical composition' for this.
+
+\begin{lem}\label{lem:jvcomp-jvcomp}
+Vertical composition is associative: the inverence rule
+\begin{equation*}
+\inference
+  { \jhom{\Gamma}{A}{B}{f_0}
+    \jfhom{\Gamma}{A}{B}{f_0}{P}{Q}{f_1}
+    \jfhom{\Gamma}{{A}{P}}{{B}{Q}}{\jvcomp{P}{f_0}{f_1}}{R}{S}{f_2}
+    }
+  { \jhomeq
+      {\Gamma}
+      {{{A}{P}}{R}}
+      {{{B}{Q}}{S}}
+      {\jvcomp{R}{\jvcomp{P}{f_0}{f_1}}{f_2}}
+      {\jvcomp{{P}{R}}{f_0}{\jvcomp{R}{f_1}{f_2}}}
+    }
+\end{equation*}
+\end{lem}
+
+\begin{proof}
+Consider the morphisms $\jhom{\Gamma}{A}{B}{f_0}$, 
+$\jfhom{\Gamma}{A}{B}{f_0}{P}{Q}{f_1}$ and 
+$\jfhom{\Gamma}{{A}{P}}{{B}{Q}}{\jvcomp{P}{f_0}{f_1}}{R}{S}{f_2}$. Then we have
+the judgmental equalities
+\begin{align*}
+\jvcomp{R}{\jvcomp{P}{f_0}{f_1}}{f_2}
+& \jdeq 
+  \tmext{\ctxwk{R}{\tmext{\ctxwk{P}{f_0}}{f_1}}}{f_2}
+  \tag{by definition}
+  \\
+& \jdeq
+  \tmext{{\ctxwk{R}{{P}{f_0}}}{\ctxwk{R}{f_1}}}{f_2}
+  \tag{by \autoref{lem:tmext-wk}}
+  \\
+& \jdeq
+  \tmext{\ctxwk{R}{{P}{f_0}}}{{\ctxwk{R}{f_1}}{f_2}}
+  \tag{by \autoref{lem:tmext-tmext}}
+  \\
+& \jdeq
+  \tmext{\ctxwk{R}{{P}{f_0}}}{\jvcomp{R}{f_1}{f_2}}
+  \tag{by definition}
+  \\
+& \jdeq
+  \tmext{\ctxwk{\ctxext{P}{R}}{f_0}}{\jvcomp{R}{f_1}{f_2}}
+  \tag{by \autoref{comp-ew-t}}
+  \\
+& \jdeq
+  \jvcomp{{P}{R}}{f_0}{\jvcomp{R}{f_1}{f_2}}.
+  \tag{by definition}
+\end{align*}
+\end{proof}
+
 \begin{lem}
@@
 we can develop a diagramatic style of of displaying type dependencies
-combined with morphisms. We give an informal, metatheoretical definition of
+combined with morphisms. We have proved all the possible associativity and
+interchange properties concerning the various kinds of composition, which
+ensures that when a diagram is asserted to commute, the order in which
+morphisms are combined has no relevance.
+
+We give an informal, metatheoretical definition of
 such diagrams by indicating what the various components mean. The definition

### 2014-05-17 article-models-ttderived.tex
reorganizing section 2.3

@@
 
-\begin{lem}
+\begin{lem}\label{lem:mor-tmext-cproj}
 Let $\jhom{\Gamma}{A}{{B}{Q}}{f}$ be a morphism from $A$ to $\ctxext{B}{Q}$
@@
 
+In the same spirit, we have the following compatibility of composition with
+projections:
+
+\begin{lem}\label{lem:jcomp-cprojfstf}
+The following inference rule is valid:
+\begin{equation*}
+\inference
+  { \jfam{{\Gamma}{A}}{P}
+    \jfam{{\Gamma}{A}}{Q}
+    }
+  { \jfameq
+      {{\Gamma}{{A}{P}}}
+      {\jcomp{{A}{P}}{\cprojfstf{A}{P}}{Q}}
+      {\ctxwk{P}{Q}}
+    }
+\end{equation*}
+\end{lem}
+
+\begin{proof}
+Let $\jfam{{\Gamma}{A}}{P}$ and $\jfam{{\Gamma}{A}}{Q}$ be
+families. Then we have
+\begin{align*}
+\jcomp{{A}{P}}{\cprojfstf{A}{P}}{Q}
+& \jdeq
+  \unfoldall{\jcomp{{A}{P}}{\cprojfstf{A}{P}}{Q}}
+  \tag{by definition}\\
+& \jdeq 
+  \subst{\ctxwk{P}{\idtm{A}}}{\ctxwk{P}{{A}{Q}}} 
+  \tag{by \autoref{comp-ww-f}}\\
+& \jdeq 
+  \ctxwk{P}{\subst{\idtm{A}}{\ctxwk{A}{Q}}} 
+  \tag{by \autoref{comp-ws-f}}\\
+& \jdeq 
+  \ctxwk{P}{Q} 
+  \tag{by \autoref{idfunc-wk-defn}}
+\end{align*}
+\end{proof}
+
 We also have the following lemma about the compatibility of pairing and composition:
@@
 \end{rmk}
+
+\subsubsection{Horizontal and vertical composition}
 
@@
 
+\begin{lem}
+The following inference rule explains how vertically and horizontally composed 
+morphisms are evaluated:
+\begin{align*}
+& \inference
+  { \jfhom{\Gamma}{A}{B}{f}{P}{Q}{F}
+    \jterm{\Gamma}{A}{x}
+    \jterm{\Gamma}{\subst{x}{P}}{u}
+    }
+  { \jtermeq
+      {\Gamma}
+      {{B}{Q}}
+      {\subst{\tmext{}{}{x}{u}}{\jvcomp{P}{f}{F}}}
+      {\tmext{}{}{\subst{x}{f}}{\subst{u}{{x}{F}}}}
+    }
+\end{align*}
+\end{lem}
+
+\begin{proof}
+Let $\jfhom{\Gamma}{A}{B}{f}{P}{Q}{F}$, $\jterm{\Gamma}{A}{x}$ and
+$\jterm{\Gamma}{\subst{x}{P}}{u}$. Then we have the judgmental equalities
+\begin{align*}
+\subst{\tmext{}{}{x}{u}}{\jvcomp{P}{f}{F}}
+& \jdeq
+  \subst{\tmext{}{}{x}{u}}{\tmext{}{}{\ctxwk{P}{f}}{F}}
+  \tag{by definition}
+  \\
+& \jdeq
+  \tmext{\subst{\tmext{x}{u}}{\ctxwk{P}{f}}}{\subst{\tmext{x}{u}}{F}}
+  \tag{by \autoref{lem:tmext-subst}}
+  \\
+& \jdeq
+  \tmext{\subst{u}{{x}{\ctxwk{P}{f}}}}{\subst{u}{{x}{F}}}
+  \tag{by \autoref{comp-es}}
+  \\
+& \jdeq
+  \tmext{\subst{u}{\ctxwk{\subst{x}{P}}{\subst{x}{f}}}}{\subst{u}{{x}{F}}}
+  \tag{by \autoref{comp-sw-t}}
+  \\
+& \jdeq
+  \tmext{\subst{x}{f}}{\subst{u}{{x}{F}}}.
+  \tag{by \autoref{cancellation-ws-t}}
+\end{align*}
+\end{proof}
+
+\begin{lem}\label{lem:jvcomp-cprojfstf}
+Consider $\jhom{\Gamma}{A}{B}{f}$ and $\jfhom{\Gamma}{A}{B}{f}{P}{Q}{F}$. Then
+the diagram
+\begin{equation*}
+\begin{tikzcd}
+\ctxext{A}{P}
+  \ar{r}{\jvcomp{P}{f}{F}}
+  \ar{d}[swap]{\cprojfstf{A}{P}}
+& \ctxext{B}{Q}
+  \ar{d}{\cprojfstf{B}{Q}}
+  \\
+A \ar{r}[swap]{f}
+& B
+\end{tikzcd}
+\end{equation*}
+commutes judgmentally.
+\end{lem}
+
+\begin{proof}
+We have the judgmental equalities
+\begin{align*}
+\jcomp{{A}{P}}{\jvcomp{P}{f}{F}}{\cprojfstf{B}{Q}}
+& \jdeq
+  \jcomp{{A}{P}}{\tmext{\ctxwk{P}{f}}{F}}{\cprojfstf{B}{Q}}
+  \tag{by definition}
+  \\
+& \jdeq
+  \ctxwk{P}{f}
+  \tag{by \autoref{lem:mor-tmext-cproj}}
+  \\
+& \jdeq
+  \jcomp{{A}{P}}{\cprojfstf{A}{P}}{f}
+  \tag{by \autoref{lem:jcomp-cprojfstf}}
+\end{align*}
+\end{proof}
+
+\subsubsection{Interactions of the composition threesome}
+Our current goal is to show that the three sorts of composition are all
+compatbile with each other. This includes showing associativity and interchange
+properties. The following lemma of the composition threesome stands at the
+basis of the further compatibility results.
+
 \begin{lem}\label{lem:composition-threesome}
-We have the following composition threesome:
+The following inference rules are valid:
 \begin{align*}
@@
 
-\begin{lem}
-The following inference rules explain how vertically and horizontally composed 
-morphisms are evaluated:
-\begin{align*}
-& \inference
-  { \jfhom{\Gamma}{A}{B}{f}{P}{Q}{F}
-    \jterm{\Gamma}{A}{x}
-    \jterm{\Gamma}{\subst{x}{P}}{u}
-    }
-  { \jtermeq
-      {\Gamma}
-      {{B}{Q}}
-      {\subst{\tmext{}{}{x}{u}}{\jvcomp{P}{f}{F}}}
-      {\tmext{}{}{\subst{x}{f}}{\subst{u}{{x}{F}}}}
-    }
-\end{align*}
-\end{lem}
-
-\begin{proof}
-Let $\jfhom{\Gamma}{A}{B}{f}{P}{Q}{F}$, $\jterm{\Gamma}{A}{x}$ and
-$\jterm{\Gamma}{\subst{x}{P}}{u}$. Then we have the judgmental equalities
-\begin{align*}
-\subst{\tmext{}{}{x}{u}}{\jvcomp{P}{f}{F}}
-& \jdeq
-  \subst{\tmext{}{}{x}{u}}{\tmext{}{}{\ctxwk{P}{f}}{F}}
-  \tag{by definition}
-  \\
-& \jdeq
-  \tmext{\subst{\tmext{x}{u}}{\ctxwk{P}{f}}}{\subst{\tmext{x}{u}}{F}}
-  \tag{by \autoref{lem:tmext-subst}}
-  \\
-& \jdeq
-  \tmext{\subst{u}{{x}{\ctxwk{P}{f}}}}{\subst{u}{{x}{F}}}
-  \tag{by \autoref{comp-es}}
-  \\
-& \jdeq
-  \tmext{\subst{u}{\ctxwk{\subst{x}{P}}{\subst{x}{f}}}}{\subst{u}{{x}{F}}}
-  \tag{by \autoref{comp-sw-t}}
-  \\
-& \jdeq
-  \tmext{\subst{x}{f}}{\subst{u}{{x}{F}}}.
-  \tag{by \autoref{cancellation-ws-t}}
-\end{align*}
-\end{proof}
-
+\subsubsection{Diagrams in the theory of contexts, families and terms}
 \label{pullback}
@@
 
-\begin{lem}
-For any $\jterm{\Gamma}{A}{x}$ and any $\jfam{{\Gamma}{A}}{P}$, the square
-\begin{equation*}
-\begin{tikzcd}
-\subst{x}{P} \ar{d} \ar{r}{\finc{x}{P}} & \ctxext{A}{P} \ar{d}{\cprojfstf{A}{P}}\\
-\emptyf \ar{r}{\ctxwk{\emptyf}{x}} & A
-\end{tikzcd}
-\end{equation*}
-is a pullback square.
-\end{lem}
-
-In the following lemma we assert that pulling back a family $Q$ 
-over $\ctxext{{\Gamma}{A}}{P}$ along a fiber
-inclusion $\finc{x}{P}$ gives the family $\subst{x}{Q}$ over $\ctxext{\Gamma}{\subst{x}{P}}$.
-
-\begin{lem}
-We have the following inference rule:
-\begin{equation*}
-\inference
-  { \jfam{{{\Gamma}{A}}{P}}{Q}
-    \jterm{\Gamma}{A}{x}
-    }
-  { \jfameq
-      {{\Gamma}{\subst{x}{P}}}
-      {\jcomp{}{\finc{x}{P}}{Q}}
-      {\subst{x}{Q}}
-    }
-\end{equation*}
-\end{lem}
-
-\begin{proof}
-We have the judgmental equalities:
-\begin{align*}
-\jcomp{\subst{x}{P}}{\finc{x}{P}}{Q}
-& \jdeq
-  \subst{\tmext{\ctxwk{\subst{x}{P}}{x}}{\idtm{\subst{x}{P}}}}{\ctxwk{\subst{x}{P}}{Q}}
-  \\
-& \jdeq
-  \subst{\idtm{\subst{x}{P}}}{{\ctxwk{\subst{x}{P}}{x}}{\ctxwk{\subst{x}{P}}{Q}}}
-  \\
-& \jdeq
-  \subst{\idtm{\subst{x}{P}}}{\ctxwk{\subst{x}{P}}{\subst{x}{Q}}}
-  \\
-& \jdeq
-  \subst{x}{Q}.
-\end{align*}
-\end{proof}
-
-\begin{lem}
-The following inference rule is valid:
-\begin{equation*}
-\inference
-  { \jfam{{\Gamma}{A}}{P}
-    \jfam{{\Gamma}{A}}{Q}
-    }
-  { \jfameq
-      {{\Gamma}{{A}{P}}}
-      {\jcomp{{A}{P}}{\cprojfstf{A}{P}}{Q}}
-      {\ctxwk{P}{Q}}
-    }
-\end{equation*}
-\end{lem}
-
-\begin{proof}
-Let $\jfam{{\Gamma}{A}}{P}$ and $\jfam{{\Gamma}{A}}{Q}$ be
-families. Then we have
-\begin{align*}
-\jcomp{{A}{P}}{\cprojfstf{A}{P}}{Q}
-& \jdeq
-  \unfoldall{\jcomp{{A}{P}}{\cprojfstf{A}{P}}{Q}}
-  \tag{by definition}\\
-& \jdeq 
-  \subst{\ctxwk{P}{\idtm{A}}}{\ctxwk{P}{{A}{Q}}} 
-  \tag{by \autoref{comp-ww-f}}\\
-& \jdeq 
-  \ctxwk{P}{\subst{\idtm{A}}{\ctxwk{A}{Q}}} 
-  \tag{by \autoref{comp-ws-f}}\\
-& \jdeq 
-  \ctxwk{P}{Q} 
-  \tag{by \autoref{idfunc-wk-defn}}
-\end{align*}
-\end{proof}
-
 \subsection{Fiber inclusions}
@@
 
+\begin{lem}
+For any $\jterm{\Gamma}{A}{x}$ and any $\jfam{{\Gamma}{A}}{P}$, the square
+\begin{equation*}
+\begin{tikzcd}
+\subst{x}{P} \ar{d} \ar{r}{\finc{x}{P}} & \ctxext{A}{P} \ar{d}{\cprojfstf{A}{P}}\\
+\emptyf \ar{r}{\ctxwk{\emptyf}{x}} & A
+\end{tikzcd}
+\end{equation*}
+is a pullback square.
+\end{lem}
+
 We can give a second characterization of fiber inclusions:

### 2014-05-17 article-models-ttinternal.tex
reorganizing section 2.3

@@
 
-\subsection{Extension-weakening algebras}
+\subsection{Weakening algebras}
 The notion of (extension-)weakening algebra is dependent on the notion of extension algebra.
@@
 
+\subsection{Term-algebras}
+Until now we had no need to consider term-algebras. Extension is an operation
+letting contexts and families interact, the empty object is a term which
+requires that contexts and families are compatible and weakening is an operation
+which takes families to extended families. However, we need to consider
+the notion of term-algebras before we can go into the internalization of
+substitution. In this subsection we will define the notion of term-algebras
+with extension and an empty object (which are simply called term-algebras),
+term-homomorphisms between them and weakening term-algebras with their
+homomorphisms.
+
 \subsection{Extension-substitution algebras}

### 2014-05-17 article-models-ttinternal.tex
initiated the section on substitution algebras
formulated the groupoid conjecture

--- 2014-05-16 article-models-ttinternal.tex
@@
 of the theory. In this section we aim for this internalization.
+
+It would be interesting to write out a weak version of pre-universes, internal
+to Martin-L\"of type theory with the function extensionality principle. 
+To do this, the empty context needs to
+be replaced by a contractible type, extension by dependent pair types,
+judgmental equalities of terms by identifications and judgmental equalities
+of families by equivalences of types. We conjecture that it is possible to
+carry this out (in particular to make sure that all the constructions 
+type-check). Moreover, one could then extend the notion of `weak' pre-universes
+with the requirement that every internal morphism is weakly anodyne. This
+should give an internal theory of weak higher groupoids.
 
@@
 
-\subsection{Extension-weakening algebras}
-The notion of (extension-)weakening algebra is dependent on the notion of extension algebra.
-Although it is strictly speaking not dependent on the notion of empty-algebra,
-we shall only formulate a weakening algebras in the setting of an
-extension-empty algebras. When one wants to have a weakening operation which
-also acts on the level of contexts in an extension algebra without empty
-context, extra work has to be done to introduce these separately.
+\subsection{Term-algebras and term-operations}
+Until now we had no need to consider term-algebras. Extension is an operation
+letting contexts and families interact and the empty object is a term which
+requires that contexts and families are compatible. However, to model the
+theory of contexts, families and terms we do need to take terms into account.
+
+\begin{defn}
+A term-algebra in context $\Gamma$ is a septuple 
+$(A,P,T,\epsilon_0,\epsilon_1,\phi_0,\phi_1)$ such that the quadruple
+$(A,P,\epsilon_0,\epsilon_1)$ is an extension algebra in context $\Gamma$, 
+the quadruple $(A,P,\phi_0,\phi_1)$ is an empty algebra in context $\Gamma$,
+for which we additionally have the following judgmental equality:
+\begin{align*}
+\jfameq*{{\Gamma}{A}}{\subst{\phi_1}{T}}{\emptyf}.
+\end{align*}
+\end{defn}
+
+\begin{thm}
+Let $\mathcal{A}\defeq(A,P,T,\epsilon_0,\epsilon_1,\phi_0,\phi_1)$ be a term-algebra in context
+$\Gamma$. Then the septuple
+\begin{equation*}
+(P,\jcomp{}{\epsilon_0}{P},\jcomp{}{\epsilon_0}{T},\epsilon_1,\jcomp{}{\epsilon_0}{\epsilon_1},\phi_1,\jcomp{}{\epsilon_0}{\phi_1})
+\end{equation*}
+is a term-algebra in context $\ctxext{\Gamma}{A}$, called the family term-algebra
+of $\mathcal{A}$.
+\end{thm}
+
+\begin{thm}
+Let $\mathcal{A}\defeq(A,P,T,\epsilon_0,\epsilon_1,\phi_0,\phi_1)$ be a term-algebra in context
+$\Gamma$ and let $X$ be a family in context $\Gamma$. Then
+\begin{equation*}
+\ctxwk{X}{\mathcal{A}}\defeq(\ctxwk{X}{A},\ctxwk{X}{P},\ctxwk{X}{T},\ctxwk{X}{\epsilon_0},\ctxwk{X}{\epsilon_1},\ctxwk{X}{\phi_0},\ctxwk{X}{\phi_1})
+\end{equation*}
+is a term-algebra in context $\ctxext{\Gamma}{X}$.
+\end{thm}
+
+\begin{thm}
+Let $\mathcal{A}\defeq(A,P,T,\epsilon_0,\epsilon_1,\phi_0,\phi_1)$ be a term-algebra in context
+$\ctxext{\Gamma}{X}$ and let $\jterm{\Gamma}{X}{x}$. Then
+\begin{equation*}
+\subst{x}{\mathcal{A}}\defeq
+(\subst{x}{A},\subst{x}{P},\subst{x}{T},\subst{x}{\epsilon_0},\subst{x}{\epsilon_1},
+\subst{x}{\phi_0},\subst{x}{\phi_1})
+\end{equation*}
+is a term-algebra in context $\Gamma$.
+\end{thm}
+
+\begin{defn}
+Let $\mathcal{A}$ and $\mathcal{B}$ be term-algebras in context $\Gamma$. 
+A term-operation from $\mathcal{A}$ to $\mathcal{B}$ is a triple $(f_0,f_1,f_2)$
+consisting of
+\begin{align*}
+\jhom*{\Gamma}{A}{B}{f_0}
+  \\
+\jfhom*{\Gamma}{A}{B}{f_0}{P}{Q}{f_1}
+  \\
+\jfhom*{\Gamma}{{A}{P}}{{B}{Q}}{\jvcomp{P}{f_0}{f_1}}{T}{U}{f_2}
+\end{align*}
+such that the pair $(f_0,f_1)$ is both an extension homomorphism from
+$(A,P,\epsilon_0,\epsilon_1)$ to $(B,Q,\eta_0,\eta_1)$ in context $\Gamma$
+and an empty homomorphism from $(A,P,\phi_0,\phi_1)$ to $(B,Q,\psi_0,\psi_1)$
+in context $\Gamma$.
+\end{defn}
+
+\subsection{Weakening term-algebras}
+Weakening term-algebras will be term-algebras with certain added structure.
+Although strictly speaking one could formulate a notion of weakening algebra
+which only depends on extension algebras and which omits both empty objects
+and terms, we shall not do so here.
+
+Let $(A,P,T,\epsilon_0,\epsilon_1,\phi_0,\phi_1)$ be a term-algebra in context
+$\Gamma$.
 
@@
 \begin{equation*}
-\begin{tikzcd}
-( \ctxwk{P}{P},
-  \ctxwk{P}{\jcomp{}{\epsilon_0}{P}},
-  \ctxwk{P}{\epsilon_1},
-  \ctxwk{P}{\jcomp{}{\epsilon_0}{\epsilon_1}})
+\begin{tikzcd}[column sep=large]
+\ctxwk{P}{(P,
+  \jcomp{}{\epsilon_0}{P},
+  \epsilon_1,
+  \jcomp{}{\epsilon_0}{\epsilon_1})}
 \ar{r}{(\omega_0,\omega_1)}
-& ( \jcomp{}{\epsilon_0}{P},
-  \jcomp{}{\epsilon_0}{\jcomp{}{\epsilon_0}{P}},
-  \jcomp{}{\epsilon_0}{\epsilon_1},
-  \jcomp{}{\epsilon_0}{\jcomp{}{\epsilon_0}{\epsilon_1}})
+& \jcomp{}{\epsilon_0}{(P,
+  \jcomp{}{\epsilon_0}{P},
+  \epsilon_1,
+  \jcomp{}{\epsilon_0}{\epsilon_1})}
 \end{tikzcd}
@@
 
+Let $(A,P,T,\epsilon_0,\epsilon_1,\phi_0,\phi_1)$ be a term-algebra. To define
+substitution algebras, we will consider
+\begin{align*}
+\jhom*
+  {{{{\Gamma}{A}}{P}}{T}}
+  {\ctxwk{T}{\jcomp{}{\epsilon_0}{P}}}
+  {P}
+  {\sigma_0}
+  \\
+\jfhom*
+  {{{{\Gamma}{A}}{P}}{T}}
+  {\ctxwk{T}{\jcomp{}{\epsilon_0}{P}}}
+  {P}
+  {\sigma_0}
+  {\ctxwk{T}{\jcomp{}{\epsilon_0}{\jcomp{}{\epsilon_0}{P}}}}
+  {\jcomp{}{\epsilon_0}{P}}
+  {\sigma_1}
+  \\
+\jfhom*
+  {{{{\Gamma}{A}}{P}}{T}}
+  { \ctxext
+      {\ctxwk{T}{\jcomp{}{\epsilon_0}{P}}}
+      {\ctxwk{T}{\jcomp{}{\epsilon_0}{\jcomp{}{\epsilon_0}{P}}}}
+    }
+  {\ctxext{P}{\jcomp{}{\epsilon_0}{P}}}
+  {\jvcomp{}{\sigma_0}{\sigma_1}}
+  {\ctxwk{T}{\jcomp{}{\epsilon_0}{\jcomp{}{\epsilon_0}{T}}}}
+  {\jcomp{}{\epsilon_0}{T}}
+  {\sigma_2}
+\end{align*}
+for which the pair $(\sigma_0,\sigma_1)$ is required to be an extension
+homomorphism and the triple $(\sigma_0,\sigma_1,\sigma_2)$ is required to
+be an empty term-homomorphism.
+
 \subsection{Pre-universes}

### 2014-05-17 article-models-ttinternal.tex
reorganization

@@
 carry this out (in particular to make sure that all the constructions 
-type-check). Moreover, one could then extend the notion of `weak' pre-universes
+type-check). This could serve as a starting point for investigating internal
+models without truncatedness assumptions and for investigating internal higher
+categories. 
+Moreover, one could then extend the notion of `weak' pre-universes
 with the requirement that every internal morphism is weakly anodyne. This
-should give an internal theory of weak higher groupoids.
+could give an internal theory of weak higher groupoids.
 
@@
 \end{defn}
+
+\begin{thm}
+Let $(A,P,\epsilon_0,\epsilon_1,\phi_0,\phi_1)$ be an extension-empty algebra
+in context $\Gamma$. Then the sextuple
+\begin{equation*}
+(P,\jcomp{}{\epsilon_0}{P},\epsilon_1,\jcomp{}{\epsilon_0}{\epsilon_1},\phi_1,
+\jcomp{}{\epsilon_0}{\phi_1})
+\end{equation*}
+is an extension-empty algebra in context $\ctxext{\Gamma}{A}$. 
+\end{thm}
+
+\begin{proof}
+We first check whether the quadruple $(P,\jcomp{}{\epsilon_0}{P},\phi_1,\jcomp{}{\epsilon_0}{\phi_1})$
+is an empty algebra in context $\ctxext{\Gamma}{A}$. Thus we need to verify
+that $\subst{\phi_1}{\jcomp{}{\epsilon_0}{P}}\jdeq P$ and that
+$\subst{\phi_1}{\jcomp{}{\epsilon_0}{\phi_1}}\jdeq\phi_1$. Both these judgmental
+equalities follow from \autoref{lem:empalg-mor}.
+
+To complete the proof, we need to verify that
+\autoref{extempalg-eq1,extempalg-eq2,extempalg-eq3,extempalg-eq4} hold in our
+current setting.
+\end{proof}
 

### 2014-05-17 article-models-ttconstr.tex
reorganization

--- 2014-03-31 article-models-ttconstr.tex
@@
 
-\subsection{Universes}\label{universes}
-The type operator $\UU$ has as input a pair $(A,P)$ consisting of a family
-$A$ in context $\Gamma$ and a family $P$ in context $\ctxext{\Gamma}{A}$. Its
-output is a pair $(U(A,P),\tilde{U}(A,P))$ consisting of a family
-$U(A,P)$ in context $\Gamma$ and a family $\tilde{U}(A,P)$ over it for which
-there are terms representing the empty context, the operations of extension,
-weakening and substitution and the identity terms. Moreover there is a morphism
-$i$ from $A$ to $U(A,P)$ with the property that 
-$\jcomp{}{i}{\tilde{U}(A,P)}\jdeq P$.
-
 \subsection{The compatibility of the type constructors with each other}

### 2014-05-17 article-models-ttinternal.tex
reorganization

@@
 and terms was that it would have to be possible to consider internal versions
-of the theory. In this section we aim for this internalization.
+of the theory. In this section we aim for this internalization. We stress that
+we shall not make any further assumptions in this section, and thus that
+it is \emph{by default} possible to consider internal models of the theory
+of contexts, families and terms in itself. In particular, we do not assume that
+there are universes; this is the subject of a later section in this part.
 

### 2014-05-17 article-models-ttmodels.tex
reorganization

@@
 \subsubsection{Every term preserves identity}
-
-\section{Pretty type theory}
-In this section we will do three things. First we explain basic type theory with
-variable names, which we simply call \emph{Pretty Type Theory}. 
-Then we will show how every formula in basic type theory without
-variable names (shall we call it \emph{Structural Type Theory}?) 
-can be interpreted in pretty type theory. Finally, we will show how every formula
-in pretty type theory can be interpreted in structural type theory.
-
-Pretty type theory is pretty as the ugly little duckling that grew up. At first, it is actually uglier than
-structural type theory because one has to keep track of a variable nobody is
-actually interested in. The situation improves when we introduce the operators
-of extension and weakening, because from that point onwards one can draw notational
-advantages from having the variable around.
-
-Pretty type theory should be such that if you write down two contexts, families
-or terms in exactly the same way, then they are the same, and there should be
-notational shortcuts for extension, weakening and substitution which make
-this interesting.
-
-\subsection{The basic judgments}
-The basic judgments of pretty type theory are the same as for structural type
-theory. There are judgments for: ``$\Gamma$ is a context'',
-``$A(i)$ over $i:\Gamma$ is a family over $\Gamma$'', ``$A(i)$ over $i:\Gamma$ 
-is a type in context $\Gamma$''
-and ``$x(i)$ is a term of $A(i)$ above $i:\Gamma$''. The other four
-judgments are for judgmental equality.
-
-\begin{align*}
-\jvctx*{\Gamma} & \jvctxeq*{\Gamma}{\Gamma'}\\
-\jvfam*{i}{\Gamma}{A} & \jvfameq*{i}{\Gamma}{A}{B}\\
-\jvtype*{i}{\Gamma}{A} & \jvtypeeq*{i}{\Gamma}{A}{B}\\
-\jvterm*{i}{\Gamma}{A}{x} & \jvtermeq*{i}{\Gamma}{A}{x}{y}.
-\end{align*}
-
-We have the following basic inference rules that relate types and families:
-
-\begin{small}
-\begin{align*}
-& \inference
-  {\jvtype{i}{\Gamma}{A}}
-  {\jvfam{i}{\Gamma}{A}}
-& & \inference
-    {\jvtypeeq{i}{\Gamma}{A}{B}}
-    {\jvfameq{i}{\Gamma}{A}{B}}\\
-& \inference
-  {\jvtype{i}{\Gamma}{A}
-   \jvfameq{i}{\Gamma}{A}{B}}
-  {\jvtype{i}{\Gamma}{B}}
-& & \inference
-    {\jvtype{i}{\Gamma}{A}
-     \jvfameq{i}{\Gamma}{A}{B}}
-    {\jvtypeeq{i}{\Gamma}{A}{B}}
-\end{align*}
-\end{small}
-
-\subsection{The basic rules for judgmental equality}
-The rules for judgmental equality establish that it is an equivalence relation.
-\bgroup\small
-\begin{align*}
-& \inference
-  {\jvctx{\Gamma}}
-  {\jvctxeq{\Gamma}{\Gamma}} 
-& & \inference
-    {\jvctxeq{\Gamma}{\Delta}}
-    {\jvctxeq{\Delta}{\Gamma}} 
-& & \inference
-    {\jvctxeq{\Gamma}{\Delta}
-     \jvctxeq{\Delta}{\greek{E}}}
-    {\jvctxeq{\Gamma}{\greek{E}}}\\
-& \inference
-  {\jvfam{i}{\Gamma}{A}}
-  {\jvfameq{i}{\Gamma}{A}{A}} 
-& & \inference
-    {\jvfameq{i}{\Gamma}{A}{B}}
-    {\jvfameq{i}{\Gamma}{B}{A}}
-& & \inference
-    {\jvfameq{i}{\Gamma}{A}{B}
-     \jvfameq{i}{\Gamma}{B}{C}}
-    {\jvfameq{i}{\Gamma}{A}{C}}\\
-& \inference
-  {\jvterm{i}{\Gamma}{A}{x}}
-  {\jvtermeq{i}{\Gamma}{A}{x}{x}}
-& & \inference
-    {\jvtermeq{i}{\Gamma}{A}{x}{y}}
-    {\jvtermeq{i}{\Gamma}{A}{y}{x}}
-& & \inference
-    {\jvtermeq{i}{\Gamma}{A}{x}{y}
-     \jvtermeq{i}{\Gamma}{A}{y}{z}}
-    {\jvtermeq{i}{\Gamma}{A}{x}{z}}
-\end{align*}
-\egroup
-
-The following convertibility rules are responsible for the strictness
-of judgmental equality, which sets it apart from equivalences or identifications:
-
-\begin{align*}
-& \inference
-  {\jvctxeq{\Gamma}{\Delta}
-   \jvfam{i}{\Gamma}{A}}
-  {\jvfam{i}{\Delta}{A}}
-& & \inference
-    {\jvctxeq{\Gamma}{\Delta}
-     \jvfameq{i}{\Gamma}{A}{B}}
-    {\jvfameq{i}{\Delta}{A}{B}}\\
-& \inference
-  {\jvctxeq{\Gamma}{\Delta}
-   \jvterm{i}{\Gamma}{A}{x}}
-  {\jvterm{i}{\Delta}{A}{x}}
-& & \inference
-    {\jvctxeq{\Gamma}{\Delta}
-     \jvtermeq{i}{\Gamma}{A}{x}{y}}
-    {\jvtermeq{i}{\Delta}{A}{x}{y}}\\
-& \inference
-  {\jvfameq{i}{\Gamma}{A}{B}
-   \jvterm{i}{\Gamma}{A}{x}}
-  {\jvterm{i}{\Gamma}{B}{x}}
-& & \inference
-    {\jvfameq{i}{\Gamma}{A}{B}
-     \jvtermeq{i}{\Gamma}{A}{x}{y}}
-    {\jvtermeq{i}{\Gamma}{B}{x}{y}}
-\end{align*}
-
-\subsection{The empty context}
-The empty context looks a bit strange when we explicitly denote the terms. But
-we will not do so anymore after this subsection.
-
-\begin{align}
-& \inference
-  {}
-  {\jctx{\emptyc}}\\
-& \inference
-  {\jctx{\Gamma}}
-  {\jvfam{i}{\Gamma}{\emptyf[\Gamma]}}\\
-& \inference
-  {\jctx{\Gamma}}
-  {\jvterm{i}{\Gamma}{\emptyf[\Gamma]}{\emptytm[\Gamma]}}\\
-& \inference
-  {\jvterm{i}{\Gamma}{\emptyf[\Gamma]}{x}}
-  {\jvtermeq{i}{\Gamma}{\emptyf[\Gamma]}{x}{\emptytm[\Gamma]}}
-\end{align}
-
-Moreover, if $\Gamma$ is a context family over the
-empty context, then $\Gamma$ is a context and every context is a context
-family over the empty context. Note that this allows us to speak
-of terms of contexts too.
-
-\begin{align}
-& \inference
-  {\jctx{\Gamma}}
-  {\jvfam{\nameless}{\emptyc}{\Gamma}} 
-& & \inference
-    {\jvfam{\nameless}{\emptyc}{\Gamma}}
-    {\jctx{\Gamma}}\\
-& \inference
-  {\jctxeq{\Gamma}{\Delta}}
-  {\jvfameq{\nameless}{\emptyc}{\Gamma}{\Delta}}
-& & \inference
-    {\jvfameq{\nameless}{\emptyc}{\Gamma}{\Delta}}
-    {\jctxeq{\Gamma}{\Delta}}
-\end{align}
-
-\subsubsection{The empty context is compatible with itslef}
-The empty context $\emptyc$ may be considered as a family of contexts over the empty
-context. When we do this, we get $\emptyf[\emptyc]$.
-\begin{equation}
-\inference
-  {}
-  {\jvfameq{\nameless}{\emptyc}{\emptyc}{\emptyf[\emptyc]}}
-\end{equation}
-In the future, we shall denote $\emptyf[\Gamma]$ by $\emptyf$. The above rule
-guarantees that this will not cause confusion. Likewise, we shall denote
-$\emptytm[\Gamma]$ by $\emptytm$.
-
-\subsection{Extension}
-We introduce extension which not only extends a context $\Gamma$ and a family
-$A$ over it to a context $\ctxext{\Gamma}{A}$, but which also extends a family $A$
-in context $\Gamma$ and a family $P$ over it to a family $\ctxext{A}{P}$ over context
-$\Gamma$. We do this to ensure that all of type theory can be done in a context.
-For instance, we could say (1) that a context in context $\Gamma$ is the same thing
-as a family over $\Gamma$; (2) When $A$ is a context in this sense, a family over
-$A$ is the same thing as a family $P$ over $\ctxext{\Gamma}{A}$ and 
-(3) when $P$ is a family over $A$ in this sense, a term of $P$ keeps its original meaning.
-
-\begin{align}
-& \inference
-  {\jvfam{i}{\Gamma}{A}}
-  {\jvfamcombi{{i}{x}}{{\Gamma}{A}}{P}}
-& & \inference
-    {\jctxeq{\Gamma}{\Delta}
-     \jfameq{\Gamma}{A}{B}}
-    {\jctxeq{\ctxext{\Gamma}{A}}{\ctxext{\Delta}{B}}}\\
-& \inference
-  {\jfam{{\Gamma}{A}}{P}}
-  {\jfam{\Gamma}{\ctxext{A}{P}}}
-& & \inference
-    {\jfameq{\Gamma}{A}{B}
-     \jfameq{{\Gamma}{A}}{P}{Q}}
-    {\jfameq{\Gamma}{\ctxext{A}{P}}{\ctxext{B}{Q}}}
-\end{align}
-
-\subsubsection{Extension is compatible with the empty context}
-The following rule asserts that extension by $\emptyc$ leaves the contexts unchanged.
-\begin{align}
-& \inference
-  {\jctx{\Gamma}}
-  {\jctxeq{\ctxext{\emptyc}{\Gamma}}{\Gamma}}\\
-& \inference
-  {\jctx{\Gamma}}
-  {\jctxeq{\ctxext{\Gamma}{\emptyf}}{\Gamma}}\\
-& \inference
-  {\jfam{\Gamma}{A}}
-  {\jfameq{\Gamma}{\ctxext{\emptyf}{A}}{A}}
-\end{align}
-
-\subsubsection{Extension is compatible with itself}
-The inference rules asserting that extension is compatible with itself assert
-that contexts are unstructured lists of type declarations. This rule is
-unavoidable if we want that for a family $A$ in context $\Gamma$, a family over
-$A$ is the same thing as a family over $\ctxext{\Gamma}{A}$.
-
-\begin{align}
-& \inference
-  {\jfam{\Gamma}{A}
-   \jfam{{\Gamma}{A}}{P}}
-  {\jctxeq{\ctxext{{\Gamma}{A}}{P}}{\ctxext{\Gamma}{{A}{P}}}}\\
-& \inference
-  {\jfam{{\Gamma}{A}}{P}
-   \jfam{{{\Gamma}{A}}{P}}{Q}}
-  {\jfameq{\Gamma}{\ctxext{{A}{P}}{Q}}{\ctxext{A}{{P}{Q}}}}
-\end{align}
-
-\section{Chalmers type theory}
-In this section we will present the type theory that Coquand and Dybjer are using.
-It is a weak type theory (I think), with not so many operations and judgmental equalities.
-We will show how every formula of Chalmers type theory can be interpreted in
-structural type theory.

### 2014-05-17 article-models-ttuniverses.tex
reorganization

\section{Universes}\label{sec:universes}

\subsection{Pre-universes}
We would like to investigate the universe operator, which assigns to a family
$A$ over $\Gamma$ and a family $P$ over $\ctxext{\Gamma}{A}$ a universe
$\UU(A,P)$, which is a family over $\Gamma$, and a decoding $\tilde{\UU}(A,P)$,
which is a family over $\ctxext{\Gamma}{\UU(A,P)}$. However, we find it useful
to first look at a pre-universe operator which extends the pair $(A,P)$ to a
quadruple $(V_0(A,P),\tilde{V}_0(A,P),V_1(A,P),\tilde{V}_1(A,P))$ consisting of
a family $V_0(A,P)$ in context $\Gamma$, a family $\tilde{V}_0(A,P)$ over
$\ctxext{\Gamma}{V_0(A,P)}$, a family $V_1(A,P)$ over 
$\ctxext{\Gamma}{V_0(A,P)}$ and a family $\tilde{V}_1(A,P)$ over.... and only
later blur the distinction between $\subst{x}{V_1(A,P)}$ and family morphisms
from $\subst{x}{\tilde{V}_0(A,P)}$ to $V_0(A,P)$ for terms $x:V_0(A,P)$. The
pair $(V_0(A,P),V_1(A,P))$ can be seen as a version of $(A,P)$ which is
expanded to incorporate extension, the empty context, weakening, substitution
and identity terms.

Our initial setup for pre-universes consists just of the two inverence rules
calling for $V_0$ and $V_1$:
\begin{align*}
& \inference
  { \jfam{{\Gamma}{A}}{P}
    }
  { \jfam{\Gamma}{V_0(A,P)}
    }
  \\
& \inference
  { \jfam{{\Gamma}{A}}{P}
    }
  { \jfam{{\Gamma}{V_0(A,P)}}{V_1(A,P)}
    }
\end{align*}

\subsection{Universe operators}
\label{universes}
Now we investigate universes. When $A$ is a family over context $\Gamma$ and
$P$ is a family over $\ctxext{\Gamma}{A}$, the universe at $(A,P)$ consists,
among other things, of a family $\UU_0(A,P)$ over context $\Gamma$ and a family
$\UU_1(A,P)$

The type operator $\UU$ has as input a pair $(A,P)$ consisting of a family
$A$ in context $\Gamma$ and a family $P$ in context $\ctxext{\Gamma}{A}$. Its
output is a pair $(U(A,P),\tilde{U}(A,P))$ consisting of a family
$U(A,P)$ in context $\Gamma$ and a family $\tilde{U}(A,P)$ over it for which
there are terms representing the empty context, the operations of extension,
weakening and substitution and the identity terms. Moreover there is a morphism
$i$ from $A$ to $U(A,P)$ with the property that 
$\jcomp{}{i}{\tilde{U}(A,P)}\jdeq P$.

### 2014-05-19 article-models-ttinternal.tex
small addition
worked on the forgotten property of extempalgs

@@
 compatibility of $\epsilon_0$ and $\epsilon_1$ with $\phi_0$ and $\phi_1$.
-There will be four such judgmental equalities.
+There will be five such judgmental equalities. Four of them will express that
+extension acts as the identity term when applied to the empty object in one of
+its arguments. The other expresses that extension of families restricted to
+families above the empty object is the same as extension of contexts.
 
@@
 and hence that $\subst{\phi_0}{\epsilon_1}$ is a 
-morphism from $\ctxext{A}{P}$ to $A$. Thus, we can require
+morphism from $\ctxext{A}{P}$ to $A$. Thus, we can require that
 \begin{equation}\label{extempalg-eq3}
-\jhomeq{\Gamma}{{A}{P}}{A}{\subst{\phi_0}{\epsilon_1}}{\epsilon_0}.
+\jhomeq{\Gamma}{{A}{P}}{A}{\subst{\phi_0}{\epsilon_1}}{\epsilon_0},
 \end{equation}
-For the final judgmental equality we need to explain the term
-\begin{equation*}
-\jterm
+expressing that the action of extension at the level of families, restricted to the empty object
+$\phi_0$, is just extension at the level of contexts. We also have two judgmental
+equalities expressing that $\epsilon_1$ restricted to $\phi_1$ and
+$\jcomp{}{\epsilon_0}{\phi_1}$ are both the identity term on $P$. Note that
+the term $\jcomp{}{\epsilon_0}{\phi_1}$ serves as the empty family of families,
+being a term of $\jcomp{}{\epsilon_0}{P}$.
+
+Before we can require that $\subst{\phi_1}{\epsilon_1}\jdeq\idtm{P}$, we need
+to compute the family
+\begin{equation*}
+\jfam
   {{{\Gamma}{A}}{\subst{\phi_1}{\jcomp{}{\epsilon_0}{P}}}}
-  {\subst{\phi_1}{\ctxwk{\ctxext{P}{\jcomp{}{\epsilon_0}{P}}}{P}}}
-  {\subst{\phi_1}{\epsilon_1}}.
-\end{equation*}
+  {\subst{\phi_1}{\ctxwk{\ctxext{P}{\jcomp{}{\epsilon_0}{P}}}{P}}}.
+\end{equation*}
+of which $\subst{\phi_1}{\epsilon_1}$ is a term.
 We have already established that $\subst{\phi_1}{\jcomp{}{\epsilon_0}{P}}\jdeq
@@
 & \jdeq
-  \ctxwk{P}{P}
+  \ctxwk{P}{P},
   \tag{by \autoref{cancellation-ws-f}}
 \end{align*}
-and we will therefore require that
+so we see that it indeed makes sense to require that
 \begin{equation}\label{extempalg-eq4}
@@
 \end{equation}
+For the last judgmental equality, we need to compute the family
+\begin{equation*}
+\jfam
+  {{{\Gamma}{A}}{\subst{\phi_1}{\jcomp{}{\epsilon_0}{P}}}}
+  {\subst{\jcomp{}{\epsilon_0}{\phi_1}}{\ctxwk{\ctxext{P}{\jcomp{}{\epsilon_0}{P}}}{P}}}.
+\end{equation*}
+This is easy:
+\begin{align*}
+\subst{\jcomp{}{\epsilon_0}{\phi_1}}{\ctxwk{\ctxext{P}{\jcomp{}{\epsilon_0}{P}}}{P}}
+& \jdeq
+  \subst{\jcomp{}{\epsilon_0}{\phi_1}}{\ctxwk{\jcomp{}{\epsilon_0}{P}}{{P}{P}}}
+  \\
+& \jdeq
+  \ctxwk{P}{P}.
+\end{align*}
+Thus we are allowed to require that
+\begin{equation}\label{extempalg-eq5}
+\subst{\jcomp{}{\epsilon_0}{\phi_1}}{\epsilon_1}.
+\end{equation}
+
 We bring all this together in the definition of extension-empty algebras:
@@
 \autoref{extempalg-eq1,extempalg-eq2,extempalg-eq3,extempalg-eq4} hold in our
-current setting.
+current setting. More precisely, we have to verify that
+\begin{align*}
+\subst{\phi_1}{\epsilon_1}
+& \jdeq 
+  \idtm{P}
+  \tag{1}
+  \\
+\subst{\jcomp{}{\epsilon_0}{\phi_1}}{\epsilon_1}
+& \jdeq 
+  \idtm{P}
+  \tag{2}
+  \\
+\subst{\phi_1}{\jcomp{}{\epsilon_0}{\epsilon_1}}
+& \jdeq
+  \epsilon_1
+  \tag{3}
+  \\
+\subst{\jcomp{}{\epsilon_0}{\phi_1}}{\jcomp{}{\epsilon_0}{\epsilon_1}}
+& \jdeq
+  \idtm{\jcomp{}{\epsilon_0}{P}}
+  \tag{4}
+  \\
+\end{align*}
 \end{proof}

### 2014-05-19 article-models-ttinternal.tex
reorganizing the weakening section

@@
 $\Gamma$, satisfying the judgmental equalities 
-\autoref{extempalg-eq1,extempalg-eq2,extempalg-eq3,extempalg-eq4}.
+\autoref{extempalg-eq1,extempalg-eq2,extempalg-eq3,extempalg-eq4,extempalg-eq5}.
 \end{defn}
@@
 To complete the proof, we need to verify that
-\autoref{extempalg-eq1,extempalg-eq2,extempalg-eq3,extempalg-eq4} hold in our
+\autoref{extempalg-eq1,extempalg-eq2,extempalg-eq3,extempalg-eq4,extempalg-eq5} hold in our
 current setting. More precisely, we have to verify that
@@
   \idtm{P}
-  \tag{1}
   \\
@@
   \idtm{P}
-  \tag{2}
   \\
@@
   \epsilon_1
-  \tag{3}
   \\
@@
   \idtm{\jcomp{}{\epsilon_0}{P}}
-  \tag{4}
-  \\
-\end{align*}
+  \\
+\subst{\jcomp{}{\epsilon_0}{\jcomp{}{\epsilon_0}{\phi_1}}}{\jcomp{}{\epsilon_0}{\epsilon_1}}
+& \jdeq
+  \idtm{\jcomp{}{\epsilon_0}{P}}
+\end{align*}
+None of these present any difficulties.
 \end{proof}
@@
 \begin{equation*}
-(P,\jcomp{}{\epsilon_0}{P},\jcomp{}{\epsilon_0}{T},\epsilon_1,\jcomp{}{\epsilon_0}{\epsilon_1},\phi_1,\jcomp{}{\epsilon_0}{\phi_1})
+\mathcal{F}_{\mathcal{A}}\defeq(P,\jcomp{}{\epsilon_0}{P},\jcomp{}{\epsilon_0}{T},\epsilon_1,\jcomp{}{\epsilon_0}{\epsilon_1},\phi_1,\jcomp{}{\epsilon_0}{\phi_1})
 \end{equation*}
@@
 Let $(A,P,T,\epsilon_0,\epsilon_1,\phi_0,\phi_1)$ be a term-algebra in context
-$\Gamma$.
-
-An extension-weakening algebra in context $\Gamma$ will be an octuple
-\begin{equation*}
-(A,P,\epsilon_0,\epsilon_1,\phi_0,\phi_1,\omega_0,\omega_1)
-\end{equation*}
-where $(A,P,\epsilon_0,\epsilon_1,\phi_0,\phi_1)$ is an extension-empty algebra
-in context $\Gamma$ and where
+$\Gamma$. The weakening homomorphism $\omega$ will be a term-homomorphism from 
+$\ctxwk{P}{\mathcal{F}_\mathcal{A}}$ to 
+$\jcomp{}{\epsilon_0}{\mathcal{F}_\mathcal{A}}$ satisfying certain extra
+judgmental equalities. The goal of this subsection is to spell these out.
+
+Concretely, the weakening homomorphism is a triple $(\omega_0,\omega_1,\omega_2)$
+consisting of
 \begin{align*}
@@
   {\omega_1}
-\end{align*}
-such that $(\omega_0,\omega_1)$ is both an extension homomorphism
-\begin{equation*}
-\begin{tikzcd}[column sep=large]
-\ctxwk{P}{(P,
-  \jcomp{}{\epsilon_0}{P},
-  \epsilon_1,
-  \jcomp{}{\epsilon_0}{\epsilon_1})}
-\ar{r}{(\omega_0,\omega_1)}
-& \jcomp{}{\epsilon_0}{(P,
-  \jcomp{}{\epsilon_0}{P},
-  \epsilon_1,
-  \jcomp{}{\epsilon_0}{\epsilon_1})}
+  \\
+\jfhom*
+  {{{\Gamma}{A}}{P}}
+  {\ctxwk{P}{\ctxext{P}{\jcomp{}{\epsilon_0}{P}}}}
+  {\jcomp{}{\epsilon_0}{\ctxext{P}{\jcomp{}{\epsilon_0}{P}}}}
+  {\jvcomp{\ctxwk{P}{\jcomp{}{\epsilon_0}{P}}}{\omega_0}{\omega_1}}
+  {\ctxwk{P}{\jcomp{}{\epsilon_0}{T}}}
+  {\jcomp{}{\epsilon_0}{\jcomp{}{\epsilon_0}{T}}}
+  {\omega_2}
+\end{align*}
+We could also summarize the situation in a diagram:
+\begin{equation*}
+\begin{tikzcd}
+\ctxwk{P}{\jcomp{}{\epsilon_0}{T}}
+  \ar[fib]{d}
+  \ar{r}{\omega_2}
+& \jcomp{}{\epsilon_0}{\jcomp{}{\epsilon_0}{T}}
+  \ar[fib]{d}
+  \\
+\ctxwk{P}{\jcomp{}{\epsilon_0}{P}}
+  \ar[fib]{d}
+  \ar{r}{\omega_1}
+& \jcomp{}{\epsilon_0}{\jcomp{}{\epsilon_0}{P}}
+  \ar[fib]{d}
+  \\
+\ctxwk{P}{P}
+  \ar{r}[swap]{\omega_0}
+& \jcomp{}{\epsilon_0}{P}
 \end{tikzcd}
 \end{equation*}
-and an empty homomorphism
-
-satisfying five(?) additional judgmental equalities expressing that $\omega_0$
-and $\omega_1$ are compatible with $\epsilon_1$, $\phi_1$
-and with each other. Our current goal is to figure out what these are.
-
-Before we go into that, we develop a bit of intuition by explaining how
+
+Before we go into chasing the judgmental equalities that are to be imposed on
+weakening, we develop a bit of intuition by explaining how
 $\omega_0$ and $\omega_1$ act when applied to a family $x_1:\subst{x_0}{P}$
-over $x_0:A$. Note that
+over $x_0:A$, and we will provide the explicit requirements that make
+$\omega$ a term-homomorphism. Note that
 \begin{equation*}
@@
 
-\subsubsection{The compatibility of weakening with extension}
-The first rule, expressing that $\omega_0$ is compatible with $\epsilon_1$ is that
-the diagram
+Because the pair $(\omega_0,\omega_1)$ is required to be an extension homomorphism,
+the diagrams
 \begin{equation}\label{wkalg-exteq1}
@@
 \end{equation}
-commutes judgmentally. For the second rule, expressing that $\omega_1$ is 
-compatible with $\epsilon_1$, we wish to fill in a commutative diagram
-\begin{equation*}
+in context $\ctxext{{\Gamma}{A}}{P}$, and
+\begin{equation}\label{wkalg-exteq2}
 \begin{tikzcd}[column sep=huge]
@@
     }
-  \ar[densely dotted]{r}{\jvcomp{}{\omega_1}{?}}
+  \ar{r}{\jvcomp{}{\omega_1}{\jcomp{}{\ctxwk{P}{\epsilon_1}}{\omega_1}}}
   \ar{d}[swap]{\ctxwk{P}{\jcomp{}{\epsilon_0}{\epsilon_1}}}
@@
 \end{tikzcd}
-\end{equation*}
-The morphism indicated by the question mark is a morphism from
-$\ctxwk{P}{\jcomp{}{\epsilon_0}{\jcomp{}{\epsilon_0}{P}}}$ to
-$%
-\jcomp{}{\omega_1}{%
-  \jcomp{}{\omega_0}{%
-    \jcomp{}{\epsilon_0}{%
-      \jcomp{}{\epsilon_0}{%
-        \ctxext{P}{\jcomp{}{\epsilon_0}{P}}
-        }
-      }
-    }
-  }
-$ in context $\ctxext{{\Gamma}{A}}{P}$. One would expect this morphism to be
-\begin{equation*}
-\begin{tikzcd}[column sep=large]
-\jcomp{}{\ctxwk{P}{\epsilon_0}}{\ctxwk{P}{\jcomp{}{\epsilon_0}{P}}}
-  \ar{r}{\jcomp{}{\ctxwk{P}{\epsilon_0}}{\omega_1}}
-& \jcomp{}{\ctxwk{P}{\epsilon_0}}{%
-    \jcomp{}{\omega_0}{%
-      \jcomp{}{\epsilon_0}{%
-        \jcomp{}{\epsilon_0}{%
-          P
-          }
-        }
-      }
-    }
-\end{tikzcd}
-\end{equation*}
-in the context $\ctxext{{\Gamma}{A}}{P}$, 
-so we must verify the judgmental equalities
-\begin{align*}
-\jcomp{}{\ctxwk{P}{\epsilon_0}}{\ctxwk{P}{\jcomp{}{\epsilon_0}{P}}}
-& \jdeq 
-  \ctxwk{P}{\jcomp{}{\epsilon_0}{\jcomp{}{\epsilon_0}{P}}}
-  \\
-\jcomp{}{\ctxwk{P}{\epsilon_0}}{%
-    \jcomp{}{\omega_0}{%
-      \jcomp{}{\epsilon_0}{%
-        \jcomp{}{\epsilon_0}{%
-          P
-          }
-        }
-      }
-    }
-& \jdeq
-  \jcomp{}{\omega_1}{%
-    \jcomp{}{\omega_0}{%
-      \jcomp{}{\epsilon_0}{%
-        \jcomp{}{\epsilon_0}{%
-          \ctxext{P}{\jcomp{}{\epsilon_0}{P}}
-          }
-        }
-      }
-    } 
-\end{align*}
-of families over the context $\ctxext{{\Gamma}{A}}{P}$. The first of these
-judgmental equalities is nothing special: it follows at straight from
-\autoref{comp-ws-f}. We prove the second equality in the following lemma.
+\end{equation}
+in context $\ctxext{{{\Gamma}{A}}{P}}{\ctxwk{P}{P}}$, commute judgmentally.
+To see that the morphism $\jvcomp{}{\omega_1}{\jcomp{}{\ctxwk{P}{\epsilon_0}}{\omega_1}}$
+indeed has the suggested codomain, we have the following lemma:
 
@@
 
-This enables us to require that the diagram
-\begin{equation}\label{wkalg-exteq2}
-\begin{tikzcd}[column sep=huge]
-\ctxwk
-  {P}
-  { \ctxext
-      {\jcomp{}{\epsilon_0}{P}}
-      {\jcomp{}{\epsilon_0}{\jcomp{}{\epsilon_0}{P}}}
-    }
-  \ar{r}{\jvcomp{}{\omega_1}{\jcomp{}{\ctxwk{P}{\epsilon_0}}{\omega_1}}}
-  \ar{d}[swap]{\ctxwk{P}{\jcomp{}{\epsilon_0}{\epsilon_1}}}
-& \jcomp{}{\omega_0}{%
-    \jcomp{}{\epsilon_0}{%
-      \jcomp{}{\epsilon_0}{%
-        \ctxext{P}{\jcomp{}{\epsilon_0}{P}}
-        }
-      }
-    }
-  \ar{d}{ \jcomp{}{\omega_0}{%
-            \jcomp{}{\epsilon_0}{%
-              \jcomp{}{\epsilon_0}{\epsilon_1}
-              }
-            }
-          }
-  \\
-\ctxwk{P}{\jcomp{}{\epsilon_0}{P}}
-  \ar{r}[swap]{\omega_1}
-& \jcomp{}{\omega_0}{%
-    \jcomp{}{\epsilon_0}{%
-      \jcomp{}{\epsilon_0}{P}
-      }
-    }
-\end{tikzcd}
-\end{equation}
-commutes judgmentally.
-
 \subsubsection{The compatibility of weakening with the empty context and family}

### 2014-05-20 article-models-ttinternal.tex
starting to implement the new code in the text

@@
 \begin{defn}
-An \emph{extension algebra in context $\Gamma$} is a quadruple $(A,P,\epsilon_0,\epsilon_1)$
-consisting of a family $A$ over context $\Gamma$, a family $P$ over the context
-$\ctxext{\Gamma}{A}$, a morphism $\epsilon_0$ from $\ctxext{A}{P}$ to $A$ in context
-$\Gamma$ and a morphism $\epsilon_1$ from $\ctxext{P}{\jcomp{}{\epsilon_0}{P}}$ to $P$ in context
-$\ctxext{\Gamma}{A}$, satisfying the judgmental equalities of
+An \emph{extension algebra $\cftalg{A}$ in context $\Gamma$} is a quadruple 
+\begin{equation*}
+(\cftalgc{\cftalg{A}},\cftalgf{\cftalg{A}},\cftctxext{\cftalg{A}},\cftfamext{\cftalg{A}})
+\end{equation*}
+consisting of a family $\cftalgc{\cftalg{A}}$ over context $\Gamma$, 
+a family $\cftalgf{\cftalg{A}}$ over the context
+$\ctxext{\Gamma}{\cftalgc{\cftalg{A}}}$,
+a morphism $\cftctxext{\cftalg{A}}$ from $\ctxext{\cftalgc{\cftalg{A}}}{
+\cftalgf{\cftalg{A}}}$ to $\cftalgc{\cftalg{A}}$ in context
+$\Gamma$ and a morphism $\cftfamext{\cftalg{A}}$ from 
+$\ctxext{\cftalgf{\cftalg{A}}}{\jcomp{}{\cftctxext{\cftalg{A}}}{\cftalgf{\cftalg{A}}}}$ 
+to $\cftalgf{\cftalg{A}}$ in context
+$\ctxext{\Gamma}{\cftalgc{\cftalg{A}}}$, satisfying the judgmental equalities of
 \autoref{eq:extalg-eq1,eq:extalg-eq2}.
@@
 \begin{defn}
-Let $\mathcal{A}$ and $\mathcal{B}$ be term-algebras in context $\Gamma$. 
-A term-operation from $\mathcal{A}$ to $\mathcal{B}$ is a triple $(f_0,f_1,f_2)$
+Let $\cftalg{A}$ and $\cftalg{B}$ be term-algebras in context $\Gamma$. 
+A term-operation $\cfthom{f}$ from $\cftalg{A}$ to $\cftalg{B}$ in context
+$\Gamma$ is a triple $(\cfthomc{\cfthom{f}},\cfthomf{\cfthom{f}},\cfthomt{\cfthom{f}})$
 consisting of
 \begin{align*}
-\jhom*{\Gamma}{A}{B}{f_0}
-  \\
-\jfhom*{\Gamma}{A}{B}{f_0}{P}{Q}{f_1}
-  \\
-\jfhom*{\Gamma}{{A}{P}}{{B}{Q}}{\jvcomp{P}{f_0}{f_1}}{T}{U}{f_2}
-\end{align*}
-such that the pair $(f_0,f_1)$ is both an extension homomorphism from
-$(A,P,\epsilon_0,\epsilon_1)$ to $(B,Q,\eta_0,\eta_1)$ in context $\Gamma$
-and an empty homomorphism from $(A,P,\phi_0,\phi_1)$ to $(B,Q,\psi_0,\psi_1)$
+\jhom*{\Gamma}{\cftalgc{\cftalg{A}}}{\cftalgc{\cftalg{B}}}{\cfthomc{\cfthom{f}}}
+  \\
+\jfhom*
+  {\Gamma}
+  {\cftalgc{\cftalg{A}}}
+  {\cftalgc{\cftalg{B}}}
+  {\cfthomc{\cfthom{f}}}
+  {\cftalgf{\cftalg{A}}}
+  {\cftalgf{\cftalg{B}}}
+  {\cfthomf{\cfthom{f}}}
+  \\
+\jfhom*
+  {\Gamma}
+  {{\cftalgc{\cftalg{A}}}{\cftalgf{\cftalg{A}}}}
+  {{\cftalgc{\cftalg{B}}}{\cftalgf{\cftalg{B}}}}
+  {\jvcomp{\cftalgf{\cftalg{A}}}{\cfthomc{\cfthom{f}}}{\cfthomf{\cfthom{f}}}}
+  {\cftalgt{\cftalg{A}}}
+  {\cftalgt{\cftalg{B}}}
+  {\cfthomt{\cfthom{f}}}
+\end{align*}
+such that the pair $(\cfthomc{\cfthom{f}},\cfthomf{\cfthom{f}})$ is both an 
+extension homomorphism from
+$(\cftalgc{\cftalg{A}},\cftalgf{\cftalg{A}},\cftctxext{\cftalg{A}},\cftfamext{\cftalg{A}})$ 
+to 
+$(\cftalgc{\cftalg{B}},\cftalgf{\cftalg{B}},\cftctxext{\cftalg{B}},\cftfamext{\cftalg{B}})$ 
+in context $\Gamma$
+and an empty homomorphism from 
+$(\cftalgc{\cftalg{A}},\cftalgf{\cftalg{A}},\cftempc{\cftalg{A}},\cftempf{\cftalg{A}})$
+to
+$(\cftalgc{\cftalg{B}},\cftalgf{\cftalg{B}},\cftempc{\cftalg{B}},\cftempf{\cftalg{B}})$
 in context $\Gamma$.
 \end{defn}
+
+Thus, when $\cfthom{f}$ is a term-operation from $\cftalg{A}$ to $\cftalg{B}$, we are
+in the situation of the following judgmentally commutative diagram:
+\begin{equation*}
+\begin{tikzcd}
+\cftalgt{\cftalg{A}}
+  \ar[fib]{d}
+  \ar{r}{\cfthomt{\cfthom{f}}}
+& \cftalgt{\cftalg{B}}
+  \ar[fib]{d}
+  \\
+\cftalgf{\cftalg{A}}
+  \ar[fib]{d}
+  \ar{r}{\cfthomf{\cfthom{f}}}
+& \cftalgf{\cftalg{B}}
+  \ar[fib]{d}
+  \\
+\cftalgc{\cftalg{A}}
+  \ar{r}{\cfthomc{\cfthom{f}}}
+& \cftalgc{\cftalg{B}}
+\end{tikzcd}
+\end{equation*}
 

### 2014-05-20 article-models-ttinternal.tex
implementing code in text --- not working!

@@
 \begin{defn}
-A term-algebra in context $\Gamma$ is a septuple 
-$(A,P,T,\epsilon_0,\epsilon_1,\phi_0,\phi_1)$ such that the quadruple
-$(A,P,\epsilon_0,\epsilon_1)$ is an extension algebra in context $\Gamma$, 
-the quadruple $(A,P,\phi_0,\phi_1)$ is an empty algebra in context $\Gamma$,
+A term-algebra $\cftalg{A}$ in context $\Gamma$ is a septuple 
+\begin{equation*}
+\unfold{\cftalg{A}}
+\end{equation*}
+such that the quadruple
+$\unfold{\extalg{A}}$ is an extension algebra in context $\Gamma$, 
+the quadruple $\unfold{\empalg{A}}$ is an empty algebra in context $\Gamma$,
 for which we additionally have the following judgmental equality:
 \begin{align*}
-\jfameq*{{\Gamma}{A}}{\subst{\phi_1}{T}}{\emptyf}.
+\jfameq*{{\Gamma}{\cftalgc{\cftalg{A}}}}{\subst{\cftempf{\cftalg{A}}}{\cftalgt{\cftalg{A}}}}{\emptyf}.
 \end{align*}

### 2014-05-22 article-models-ttinternal.tex
rewriting empty-object algebras -- in progress

@@
 
-\subsection{Extension-empty algebras}
+\subsection{Term-algebras}
+In studying extension algebras we had no need to consider terms. However, we
+do need to consider terms if we want a full interpretation of the theory of
+contexts, families and terms. Therefore, we shall now consider algebras where
+we have a further family 
+$\jfam{{{\Gamma}{A}}{\cftalgf{\cftalg{A}}}}{\cftalgt{\cftalg{A}}}$, assigning
+terms to families of $\cftalg{A}$. The notion of term-algebras that we will
+study in this subsection will also have an empty object $(\cftempc{\cftalg{A}},
+\cftempf{\cftalg{A}})$ and guarantiees that families over the empty object
+are just the terms of the underlying family $\cftalgc{\cftalg{A}}$ of contexts
+of the term-algebra $\cftalg{A}$. As with the theory of contexts, families and
+terms, this enables us to speak of terms of contexts of $\cftalg{A}$, and this
+will provide the categorical structure for $\cftalg{A}$ when we also consider
+internal weakening, substitution and identity terms.
+
 \begin{defn}
-Let $P$ be a family over the extended context $\ctxext{\Gamma}{A}$, let
-$\jterm{\Gamma}{A}{\phi_0}$ and $\jterm{{\Gamma}{A}}{P}{\phi_1}$ be terms. Then the
-quadruple $(A,P,\phi_0,\phi_1)$ is said to be an \emph{empty algebra in context $\Gamma$}
-if the following judgmental equalities hold:
+A quintuple $(\cftalgc{\cftalg{A}},\cftalgf{\cftalg{A}},\cftalgt{\cftalg{A}},
+\cftempc{\cftalg{A}},\cftempf{\cftalg{A}})$ consisting of
+\begin{enumerate}
+\item A family $\cftalgc{\cftalg{A}}$ in context $\Gamma$,
+\item A family $\cftalgf{\cftalg{A}}$ in context $\ctxext{\Gamma}{\cftalgc{\cftalg{A}}}$,
+\item A family $\cftalgt{\cftalg{A}}$ in context $\ctxext{{\Gamma}{A}}{\cftalgf{\cftalg{A}}}$,
+\item A term $\cftempc{\cftalg{A}}$ of $\cftalgc{\cftalg{A}}$ in context $\Gamma$,
+\item A term $\cftempf{\cftalg{A}}$ of $\cftalgf{\cftalg{A}}$ in context $\ctxext{\Gamma}{\cftalgc{\cftalg{A}}}$,
+\end{enumerate}
+is said to be an \emph{empty-object algebra in context $\Gamma$} if additionally
+the following judgmental equalities hold:
 \begin{align}
-\jfameq*{\Gamma}{\subst{\phi_0}{P}}{A}
+\jfameq*{\Gamma}{\subst{\cftempc{\cftalg{A}}}{\cftalgf{\cftalg{A}}}}{\cftalgc{\cftalg{A}}}
   \label{empalg-eq1}
   \\
-\jtermeq*{\Gamma}{A}{\subst{\phi_0}{\phi_1}}{\phi_0}.
+\jtermeq*{\Gamma}{\cftalgc{\cftalg{A}}}{\subst{\cftempc{\cftalg{A}}}{\cftempf{\cftalg{A}}}}{\cftempc{\cftalg{A}}}.
   \label{empalg-eq2}
+  \\
+\jfameq*{{\Gamma}{\cftalgc{\cftalg{A}}}}{\subst{\cftempf{\cftalg{A}}}{\cftalgt{\cftalg{A}}}}{\emptyf}
 \end{align}
+When $(\cftalgc{\cftalg{A}},\cftalgf{\cftalg{A}},\cftalgt{\cftalg{A}},
+\cftempc{\cftalg{A}},\cftempf{\cftalg{A}})$ is an empty-object algebra
+in context $\Gamma$, we will
+also refer to the pair $(\cftempc{\cftalg{A}},\cftempf{\cftalg{A}})$ as the
+\emph{empty object of $\cftalg{A}$}.
 \end{defn}
@@
 \begin{defn}
-Let $(A,P,\phi_0,\phi_1)$ and $(B,Q,\psi_0,\psi_1)$ be empty algebras in context
-$\Gamma$. An empty homomorphism from $(A,P,\phi_0,\phi_1)$ to $(B,Q,\psi_0,\psi_1)$
-is a pair $(f_0,f_1)$ consisting of
-\begin{align*}
-\jhom*{\Gamma}{A}{B}{f_0}\\
-\jfhom*{\Gamma}{A}{B}{f}{P}{Q}{f_1}
-\end{align*}
-satisfying the judgmental equalities
-\begin{align*}
-\jtermeq*{\Gamma}{B}{\subst{\phi_0}{f_0}}{\psi_0}\\
-\jhomeq*{\Gamma}{A}{B}{\subst{\phi_0}{f_1}}{f_0}\\
-\jtermeq*{{\Gamma}{A}}{\jcomp{}{f_0}{Q}}{\subst{\phi_1}{f_1}}{\jcomp{}{f_0}{\psi_1}}.
+Let $\cftalg{A}$ and $\cftalg{B}$ be empty-object algebras in context $\Gamma$.
+An \emph{empty-object homomorphism $\cfthom{f}$ from $\cftalg{A}$ to $\cftalg{B}$
+in context $\Gamma$} is a triple $(\cfthomc{\cfthom{f}},\cfthomf{\cfthom{f}},\cfthomt{\cfthom{f}})$ consisting of
+\begin{align*}
+\jhom*
+  {\Gamma}
+  {\cftalgc{\cftalg{A}}}
+  {\cftalgc{\cftalg{B}}}
+  {\cfthomc{\cfthom{f}}}\\
+\jfhom*
+  {\Gamma}
+  {\cftalgc{\cftalg{A}}}
+  {\cftalgc{\cftalg{B}}}
+  {\cfthomc{\cfthom{f}}}
+  {\cftalgf{\cftalg{A}}}
+  {\cftalgf{\cftalg{B}}}
+  {\cfthomf{\cfthom{f}}}\\
+\jfhom*
+  {\Gamma}
+  {{\cftalgc{\cftalg{A}}}{\cftalgf{\cftalg{A}}}}
+  {{\cftalgc{\cftalg{B}}}{\cftalgf{\cftalg{B}}}}
+  {\jvcomp{\cftalgf{\cftalg{A}}}{\cfthomc{\cfthom{f}}}{\cfthomf{\cfthom{f}}}}
+  {\cftalgt{\cftalg{A}}}
+  {\cftalgt{\cftalg{B}}}
+  {\cfthomt{\cfthom{f}}}
+\end{align*}
+as displayed in \autoref{fig:empty-object-hom}, satisfying the judgmental equalities
+\begin{align*}
+\jtermeq*
+  {\Gamma}
+  {\cftalgc{\cftalg{B}}}
+  {\subst{\cftemp{\cftalg{A}}}{\cfthomc{\cfthom{f}}}}
+  {\cftempc{\cftalg{B}}}
+  \\
+\jhomeq*
+  {\Gamma}
+  {\cftalgc{\cftalg{A}}}
+  {\cftalgc{\cftalg{B}}}
+  {\subst{\cftempc{\cftalg{A}}}{\cfthomf{\cfthom{f}}}}
+  {\cfthomc{\cfthom{f}}}
+  \\
+\jtermeq*
+  {{\Gamma}{\cftalgc{\cftalg{A}}}}
+  {\jcomp{\cftalgc{\cftalg{A}}}{\cfthomc{\cfthom{f}}}{\cftalgf{\cftalg{B}}}}
+  {\subst{\cftempf{\cftalg{A}}}{\cfthomf{\cfthom{f}}}}
+  {\jcomp{\cftalgc{\cftalg{A}}}{\cfthomc{\cfthom{f}}}{\cftempf{\cftalg{B}}}}.
 \end{align*}
 \end{defn}
+
+\begin{figure}
+\begin{equation*}
+\begin{tikzcd}
+\cftalgt{\cftalg{A}}
+  \ar[fib]{d}
+  \ar{r}{\cfthomt{\cfthom{f}}}
+& \cftalgt{\cftalg{B}}
+  \ar[fib]{d}
+  \\
+\cftalgf{\cftalg{A}}
+  \ar[fib]{d}
+  \ar{r}{\cfthomf{\cfthom{f}}}
+& \cftalgf{\cftalg{B}}
+  \ar[fib]{d}
+  \\
+\cftalgc{\cftalg{A}}
+  \ar{r}[swap]{\cfthomc{\cfthom{f}}}
+& \cftalgc{\cftalg{B}}
+\end{tikzcd}
+\end{equation*}
+\caption{\label{fig:empty-object-hom}The ingredients of an empty-object homomorphism $\cfthom{f}$ from 
+$\cftalg{A}$ to $\cftalg{B}$ in context $\Gamma$.}
+\end{figure}
 
@@
 such that the quadruple
-$\unfold{\extalg{A}}$ is an extension algebra in context $\Gamma$, 
+$\unfold{\extalg}{A}$ is an extension algebra in context $\Gamma$, 
 the quadruple $\unfold{\empalg{A}}$ is an empty algebra in context $\Gamma$,

### 2014-05-22 article-models-ttinternal.tex
working on new notation for term-algebras

@@
 
-Thus, the empty algebras are the kind of algebras that require that families
-are compatible with contexts, just as our motivation in \autoref{empty}. We
-now combine the notion of extension algebras and empty algebras.
-
-An extension-empty algebra in context $\Gamma$ is going to be a sextuple
-$(A,P,\epsilon_0,\epsilon_1,\phi_0,\phi_1)$ for which 
-the quadruple $(A,P,\epsilon_0,\epsilon_1)$ is an extension algebra in context 
-$\Gamma$, the quadruple $(A,P,\phi_0,\phi_1)$ is an empty algebra in context
+The following lemma presents a property of empty-object algebras that will get
+a lot of use.
+
+\begin{lem}\label{lem:empalg-mor}
+Consider an empty-object algebra $\cftalg{A}$ in context $\Gamma$
+and a morphism 
+\begin{equation*}
+\jhom{\Gamma}{{\cftalgc{\cftalg{A}}}{\cftalgf{\cftalg{A}}}}{B}{f}.
+\end{equation*}
+Then $\subst{\cftemp[i]{\cftalg{A}}}{f}$ is a morphism from $\cftalgc{\cftalg{A}}$ to $B$ 
+in context $\Gamma$ and the following inference rules are valid for $i$ being 
+$\mathsf{c}$ or $\mathsf{f}$:
+\begin{align*}
+& \inference
+  { \jfam{{\Gamma}{\cftalgc{\cftalg{A}}}}{Q}
+    }
+  { \jfameq
+      {{\Gamma}{\cftalgc{\cftalg{A}}}}
+      {\subst{\cftemp[i]{\cftalg{A}}}{\jcomp{}{f}{Q}}}
+      {\jcomp{}{\subst{\cftemp[i]{\cftalg{A}}}{f}}{Q}}
+    }
+  \\
+& \inference
+  { \jfam{{{\Gamma}{\cftalgc{\cftalg{A}}}}{Q}}{R}
+    }
+  { \jfameq
+      {{{\Gamma}{\cftalgc{\cftalg{A}}}}{\jcomp{}{\subst{\cftemp[i]{\cftalg{A}}}{f}}{Q}}}
+      {\subst{\cftemp[i]{\cftalg{A}}}{\jcomp{}{f}{R}}}
+      {\jcomp{}{\subst{\cftemp[i]{\cftalg{A}}}{f}}{R}}
+    }
+  \\
+& \inference
+  { \jterm{{{\Gamma}{\cftalgc{\cftalg{A}}}}{Q}}{R}{h}
+    }
+  { \jtermeq
+      {{{\Gamma}{\cftalgc{\cftalg{A}}}}{\jcomp{}{\subst{\cftemp[i]{\cftalg{A}}}{f}}{Q}}}
+      {\jcomp{}{\subst{\cftemp[i]{\cftalg{A}}}{f}}{R}}
+      {\subst{\cftemp[i]{\cftalg{A}}}{\jcomp{}{f}{h}}}
+      {\jcomp{}{\subst{\cftemp[i]{\cftalg{A}}}{f}}{h}}
+    }
+\end{align*}
+\end{lem}
+
+\begin{proof}
+We only prove the first inference rule in both cases.
+Let $Q$ be a family over $\ctxext{\Gamma}{\cftalgc{\cftalg{A}}}$. In the case 
+$i=\mathsf{c}$ we have the judgmental equalities
+\begin{align*}
+\subst{\cftempc{\cftalg{A}}}{\jcomp{}{f}{Q}}
+& \jdeq
+  \subst
+    {\cftempc{\cftalg{A}}}
+    {{f}{\ctxwk{\ctxext{\cftalgc{\cftalg{A}}}{\cftalgf{\cftalg{A}}}}{Q}}}
+  \tag{by definition}
+  \\
+& \jdeq
+  \subst
+    { {\cftempc{\cftalg{A}}}{f}
+      }
+    { {\cftempc{\cftalg{A}}}
+      {\ctxwk{\ctxext{\cftalgc{\cftalg{A}}}{\cftalgf{\cftalg{A}}}}{Q}}
+      }
+  \tag{by \autoref{comp-ss-f}}
+  \\
+& \jdeq
+  \subst
+    { {\cftempc{\cftalg{A}}}{f}
+      }
+    { {\cftempc{\cftalg{A}}}
+      {\ctxwk{\cftalgf{\cftalg{A}}}{{\cftalgc{\cftalg{A}}}{Q}}}
+      }
+  \tag{by \autoref{comp-ew-f}}
+  \\
+& \jdeq
+  \subst
+    { {\cftempc{\cftalg{A}}}{f}
+      }
+    { \ctxwk
+        {\subst{\cftempc{\cftalg{A}}}{\cftalgf{\cftalg{A}}}}
+        {\subst{\cftempc{\cftalg{A}}}{\ctxwk{\cftalgc{\cftalg{A}}}{Q}}}
+      }
+  \tag{by \autoref{comp-sw-f}}
+  \\
+& \jdeq
+  \subst
+    { {\cftempc{\cftalg{A}}}{f}
+      }
+    { \ctxwk
+        {\subst{\cftempc{\cftalg{A}}}{\cftalgf{\cftalg{A}}}}
+        {Q}
+      }
+  \tag{by \autoref{cancellation-ws-f}}
+  \\
+& \jdeq
+  \subst{{\cftempc{\cftalg{A}}}{f}}{\ctxwk{\cftalgc{\cftalg{A}}}{Q}}
+  \tag{by \autoref{empalg-eq1}}
+  \\
+& \jdeq
+  \jcomp{}{\subst{\cftempc{\cftalg{A}}}{f}}{Q}.
+  \tag{by definition}
+\end{align*}
+In the case $i=\mathsf{f}$ we have the judgmental equalities
+\begin{align*}
+\subst{\cftempf{\cftalg{A}}}{\jcomp{}{f}{Q}}
+& \jdeq
+  \subst
+    { \cftempf{\cftalg{A}}
+      }
+    { {f}
+      {\ctxwk{\ctxext{\cftalgc{\cftalg{A}}}{\cftalgf{\cftalg{A}}}}{Q}}
+      }
+  \tag{by definition}
+  \\
+& \jdeq
+  \subst
+    { {\cftempf{\cftalg{A}}}{f}
+      }
+    { {\cftempf{\cftalg{A}}}
+      {\ctxwk{\ctxext{\cftalgc{\cftalg{A}}}{\cftalgf{\cftalg{A}}}}{Q}}
+      }
+  \tag{by \autoref{comp-ss-f}}
+  \\
+& \jdeq
+  \subst
+    { {\cftempf{\cftalg{A}}}{f}
+      }
+    { {\cftempf{\cftalg{A}}}
+      {\ctxwk{\cftalgf{\cftalg{A}}}{{\cftalgc{\cftalg{A}}}{Q}}}
+      }
+  \tag{by \autoref{comp-ew-f}}
+  \\
+& \jdeq
+  \subst
+    { {\cftempf{\cftalg{A}}}{f}
+      }
+    { \ctxwk{\cftalgc{\cftalg{A}}}{Q}
+      }
+  \tag{by \autoref{cancellation-ws-f}}
+  \\
+& \jdeq
+  \jcomp{}{\subst{\cftempf{\cftalg{A}}}{f}}{Q}.
+  \tag{by definition}
+\end{align*}
+\end{proof}
+
+A term-algebra $\cftalg{A}$ in context $\Gamma$ is going to be a septuple
+\begin{equation*}
+(\cftalgc{\cftalg{A}},\cftalgf{\cftalg{A}},\cftalgt{\cftalg{A}},
+\cftctxext{\cftalg{A}},\cftfamext{\cftalg{A}},\cftempc{\cftalg{A}},
+\cftempf{\cftalg{A}})
+\end{equation*}
+for which 
+the quadruple $\unfold{\extalg{A}}$ is an extension algebra in context 
+$\Gamma$, the quintuple $\unfold{\empalg{A}}$ is an empty algebra in context
 $\Gamma$, satisfying additional judgmental equalities expressing the 
-compatibility of $\epsilon_0$ and $\epsilon_1$ with $\phi_0$ and $\phi_1$.
+compatibility of $\cftctxext{\cftalg{A}}$ and $\cftfamext{\cftalg{A}}$ with 
+$\cftempc{\cftalg{A}}$ and $\cftempf{\cftalg{A}}$.
 There will be five such judgmental equalities. Four of them will express that
@@
 \begin{align}
-\jtermeq*{{\Gamma}{A}}{\ctxwk{A}{A}}{\subst{\phi_0}{\epsilon_0}}{\idtm{A}}
+\jtermeq*{{\Gamma}{\cftalgc{\cftalg{A}}}}{\ctxwk{\cftalgc{\cftalg{A}}}{\cftalgc{\cftalg{A}}}}{\subst{\cftempc{\cftalg{A}}}{\cftctxext{\cftalg{A}}}}{\idtm{\cftalgc{\cftalg{A}}}}
   \label{extempalg-eq1}
   \\
-\jtermeq*{{\Gamma}{A}}{\ctxwk{A}{A}}{\subst{\phi_1}{\epsilon_0}}{\idtm{A}}
+\jtermeq*{{\Gamma}{\cftalgc{\cftalg{A}}}}{\ctxwk{\cftalgc{\cftalg{A}}}{\cftalgc{\cftalg{A}}}}{\subst{\cftempf{\cftalg{A}}}{\cftctxext{\cftalg{A}}}}{\idtm{\cftalgc{\cftalg{A}}}}
   \label{extempalg-eq2}
@@
 we use the following lemma.
-
-\begin{lem}\label{lem:empalg-mor}
-Consider an empty algebra $(A,P,\phi_0,\phi_1)$ in context $\Gamma$
-and a morphism $\jhom{\Gamma}{{A}{P}}{B}{f}$.
-Then $\subst{\phi_i}{f}$ is a morphism from $A$ to $B$ in context $\Gamma$ and
-the following inference rules are valid for $i$ being $0$ or $1$:
-\begin{align*}
-& \inference
-  { \jfam{{\Gamma}{A}}{Q}
-    }
-  { \jfameq
-      {{\Gamma}{A}}
-      {\subst{\phi_i}{\jcomp{}{f}{Q}}}
-      {\jcomp{}{\subst{\phi_i}{f}}{Q}}
-    }
-  \\
-& \inference
-  { \jfam{{{\Gamma}{A}}{Q}}{R}
-    }
-  { \jfameq
-      {{{\Gamma}{A}}{\jcomp{}{\subst{\phi_i}{f}}{Q}}}
-      {\subst{\phi_i}{\jcomp{}{f}{R}}}
-      {\jcomp{}{\subst{\phi_i}{f}}{R}}
-    }
-  \\
-& \inference
-  { \jterm{{{\Gamma}{A}}{Q}}{R}{h}
-    }
-  { \jtermeq
-      {{{\Gamma}{A}}{\jcomp{}{\subst{\phi_i}{f}}{Q}}}
-      {\jcomp{}{\subst{\phi_i}{f}}{R}}
-      {\subst{\phi_i}{\jcomp{}{f}{h}}}
-      {\jcomp{}{\subst{\phi_i}{f}}{h}}
-    }
-\end{align*}
-\end{lem}
-
-\begin{proof}
-We only prove the first inference rule in both cases.
-Let $Q$ be a family over $\ctxext{\Gamma}{A}$. In the case $i=0$
- we have the judgmental equalities
-\begin{align*}
-\subst{\phi_0}{\jcomp{}{f}{Q}}
-& \jdeq
-  \subst{\phi_0}{{f}{\ctxwk{\ctxext{A}{P}}{Q}}}
-  \tag{by definition}
-  \\
-& \jdeq
-  \subst{{\phi_0}{f}}{{\phi_0}{\ctxwk{\ctxext{A}{P}}{Q}}}
-  \tag{by \autoref{comp-ss-f}}
-  \\
-& \jdeq
-  \subst{{\phi_0}{f}}{{\phi_0}{\ctxwk{P}{{A}{Q}}}}
-  \tag{by \autoref{comp-ew-f}}
-  \\
-& \jdeq
-  \subst{{\phi_0}{f}}{\ctxwk{\subst{\phi_0}{P}}{\subst{\phi_0}{\ctxwk{A}{Q}}}}
-  \tag{by \autoref{comp-sw-f}}
-  \\
-& \jdeq
-  \subst{{\phi_0}{f}}{\ctxwk{\subst{\phi_0}{P}}{Q}}
-  \tag{by \autoref{cancellation-ws-f}}
-  \\
-& \jdeq
-  \subst{{\phi_0}{f}}{\ctxwk{A}{Q}}
-  \tag{by \autoref{empalg-eq1}}
-  \\
-& \jdeq
-  \jcomp{}{\subst{\phi_0}{f}}{Q}.
-  \tag{by definition}
-\end{align*}
-In the case $i=1$ we have the judgmental equalities
-\begin{align*}
-\subst{\phi_1}{\jcomp{}{f}{Q}}
-& \jdeq
-  \subst{\phi_1}{{f}{\ctxwk{\ctxext{A}{P}}{Q}}}
-  \tag{by definition}
-  \\
-& \jdeq
-  \subst{{\phi_1}{f}}{{\phi_1}{\ctxwk{\ctxext{A}{P}}{Q}}}
-  \tag{by \autoref{comp-ss-f}}
-  \\
-& \jdeq
-  \subst{{\phi_1}{f}}{{\phi_1}{\ctxwk{P}{{A}{Q}}}}
-  \tag{by \autoref{comp-ew-f}}
-  \\
-& \jdeq
-  \subst{{\phi_1}{f}}{\ctxwk{A}{Q}}
-  \tag{by \autoref{cancellation-ws-f}}
-  \\
-& \jdeq
-  \jcomp{}{\subst{\phi_1}{f}}{Q}.
-  \tag{by definition}
-\end{align*}
-\end{proof}
 

### 2014-05-22 article-models-ttinternal.tex
implementing the new commands

@@
 \begin{align*}
-\jhom*{\Gamma}{{A}{P}}{A}{\epsilon_0}\\
-\jhom*{{\Gamma}{A}}{{P}{\jcomp{}{\epsilon_0}{P}}}{P}{\epsilon_1}.
+\jhom*{\Gamma}{{A}{P}}{A}{e_0}\\
+\jhom*{{\Gamma}{A}}{{P}{\jcomp{}{e_0}{P}}}{P}{e_1}.
 \end{align*}
@@
 \begin{rmk}
-Instead of looking at $\epsilon_1$ as a context morphism from $\ctxext{P}{\jcomp{}{\epsilon_0}{P}}$
-to $P$, one could also look at $\epsilon_1$ as a morphism \emph{over $\cprojfstf{A}{P}$},
+Instead of looking at $e_1$ as a context morphism from $\ctxext{P}{\jcomp{}{e_0}{P}}$
+to $P$, one could also look at $e_1$ as a morphism \emph{over $\cprojfstf{A}{P}$},
 as indicated in the following diagram:
@@
   \ar[fib]{d}
-& \jcomp{}{\epsilon_0}{P}
+& \jcomp{}{e_0}{P}
   \ar[fib]{d}
-  \ar{l}[swap]{\epsilon_1}
+  \ar{l}[swap]{e_1}
   \ar{r}
@@
   \ar{l}{\cprojfstf{A}{P}}
-  \ar{r}[swap]{\epsilon_0}
+  \ar{r}[swap]{e_0}
 & A
@@
 \end{equation*}
-This makes it clear that $\epsilon_1$ takes a family over an extended context as an
+This makes it clear that $e_1$ takes a family over an extended context as an
 argument. The extended context consists of a `base part' and a `family part'. 
-The output of $\epsilon_1$ is a new (extended) family over that base part. Forgetting 
+The output of $e_1$ is a new (extended) family over that base part. Forgetting 
 the family part is what the projection takes care of.
@@
 Extension also satisfies the properties explained in \autoref{comp-ee}, so we
-must find the two judgmental equalities for $\epsilon_0$ and $\epsilon_1$ mimicing those. 
+must find the two judgmental equalities for $e_0$ and $e_1$ mimicing those. 
 The first of these judgmental equalities is easy to give: it says that the
@@
 \begin{tikzcd}[column sep=huge]
-\ctxext{A}{{P}{\jcomp{}{\epsilon_0}{P}}} 
-  \ar{d}[swap]{\jvcomp{}{\epsilon_0}{\idtm{\jcomp{}{\epsilon_0}{P}}}
-    } 
-  \ar{r}{\jvcomp{}{\idtm{A}}{\epsilon_1}
-    } 
-  & \ctxext{A}{P} \ar{d}{\epsilon_0}\\
-\ctxext{A}{P} \ar{r}[swap]{\epsilon_0} & A
+\ctxext{A}{{P}{\jcomp{}{e_0}{P}}} 
+  \ar{d}[swap]{\jvcomp{}{e_0}{\idtm{\jcomp{}{e_0}{P}}}
+    } 
+  \ar{r}{\jvcomp{}{\idtm{A}}{e_1}
+    } 
+  & \ctxext{A}{P} \ar{d}{e_0}\\
+\ctxext{A}{P} \ar{r}[swap]{e_0} & A
 \end{tikzcd}
 \end{equation}
-To get a feel for this judgmental equality we include the following lemma.
-
-\begin{lem}
-Let $A$, $P$, $\epsilon_0$ and $\epsilon_1$ be as above, satisfying \autoref{eq:extalg-eq1} and
-let $x_0:A$,
-$x_1:\subst{x_0}{P}$ and $x_2:\subst{x_1}{{x_0}{\jcomp{}{\epsilon_0}{P}}}$.
-Then we have the judgmental equality
-\begin{equation*}
-\subst{{x_2}{{x_1}{\epsilon_1}}}{{x_0}{\epsilon_0}}
-\jdeq
-\subst{x_2}{{{x_1}{{x_0}{\epsilon_0}}}{\epsilon_0}}.
-\end{equation*}
-\end{lem}
-
-\begin{proof}
-The proof is a simple computation:
-\begin{align*}
-\subst{{x_2}{{x_1}{\epsilon_1}}}{{x_0}{\epsilon_0}}
-& \jdeq
-  \subst{\tmext{x_0}{\subst{x_2}{{x_1}{\epsilon_1}}}}{\epsilon_0}
-  \\
-& \jdeq 
-  \subst{{x_2}{{x_1}{{x_0}{\jvcomp{}{\idtm{A}}{\epsilon_1}}}}}{\epsilon_0}
-  \\
-& \jdeq
-  \subst{x_2}{{x_1}{{x_0}{\jcomp{}{\jvcomp{}{\idtm{A}}{\epsilon_1}}{\epsilon_0}}}}
-  \\
-& \jdeq
-  \subst{x_2}{{x_1}{{x_0}{\jcomp{}{\jvcomp{}{\epsilon_0}{\idtm{\jcomp{}{\epsilon_0}{P}}}}{\epsilon_0}}}}
-  \\
-& \jdeq 
-  \subst{{x_2}{{x_1}{{x_0}{\jvcomp{}{\epsilon_0}{\idtm{\jcomp{}{\epsilon_0}{P}}}}}}}{\epsilon_0}
-  \\
-& \jdeq
-  \subst{\tmext{\subst{x_1}{{x_0}{\epsilon_0}}}{x_2}}{\epsilon_0}
-  \\
-& \jdeq
-  \subst{x_2}{{{x_1}{{x_0}{\epsilon_0}}}{\epsilon_0}}.\qedhere
-\end{align*}
-\end{proof}
 
@@
 to consider `higher' families, i.e.~families over families over families, and
-thus we need to look at the family $\jcomp{}{\epsilon_0}{\jcomp{}{\epsilon_0}{P}}$ and find the
+thus we need to look at the family $\jcomp{}{e_0}{\jcomp{}{e_0}{P}}$ and find the
 two dotted morphisms in the diagram
@@
 \begin{tikzcd}
-\ctxext{P}{{\jcomp{}{\epsilon_0}{P}}{\jcomp{}{\epsilon_0}{\jcomp{}{\epsilon_0}{P}}}}
+\ctxext{P}{{\jcomp{}{e_0}{P}}{\jcomp{}{e_0}{\jcomp{}{e_0}{P}}}}
   \ar[densely dotted]{d}
   \ar[densely dotted]{r}
-& \ctxext{P}{\jcomp{}{\epsilon_0}{P}} \ar{d}{\epsilon_1}\\
-\ctxext{P}{\jcomp{}{\epsilon_0}{P}} \ar{r}{\epsilon_1} & P
+& \ctxext{P}{\jcomp{}{e_0}{P}} \ar{d}{e_1}\\
+\ctxext{P}{\jcomp{}{e_0}{P}} \ar{r}{e_1} & P
 \end{tikzcd}
@@
 \begin{equation*}
-\ctxext{\jcomp{}{\epsilon_0}{P}}{\jcomp{}{\epsilon_0}{\jcomp{}{\epsilon_0}{P}}}
+\ctxext{\jcomp{}{e_0}{P}}{\jcomp{}{e_0}{\jcomp{}{e_0}{P}}}
   \jdeq
-  \jcomp{}{\epsilon_0}{\ctxext{P}{\jcomp{}{\epsilon_0}{P}}}
+  \jcomp{}{e_0}{\ctxext{P}{\jcomp{}{e_0}{P}}}
 \end{equation*}
@@
   {{\Gamma}{A}}
-  {{P}{{\jcomp{}{\epsilon_0}{P}}{\jcomp{}{\epsilon_0}{\jcomp{}{\epsilon_0}{P}}}}}
-  {{P}{\jcomp{}{\epsilon_0}{P}}}
-  {\jvcomp{}{\idtm{P}}{\jcomp{}{\epsilon_0}{\epsilon_1}}}.
+  {{P}{{\jcomp{}{e_0}{P}}{\jcomp{}{e_0}{\jcomp{}{e_0}{P}}}}}
+  {{P}{\jcomp{}{e_0}{P}}}
+  {\jvcomp{}{\idtm{P}}{\jcomp{}{e_0}{e_1}}}.
 \end{equation*}
 For the other morphism we need to look at the family
-$\ctxext{P}{{\jcomp{}{\epsilon_0}{P}}{\jcomp{}{\epsilon_0}{\jcomp{}{\epsilon_0}{P}}}}$ differently. We do
+$\ctxext{P}{{\jcomp{}{e_0}{P}}{\jcomp{}{e_0}{\jcomp{}{e_0}{P}}}}$ differently. We do
 that in the following lemma.
@@
 \begin{lem}\label{lem:extalg-twins}
-Suppose we have $A$, $P$, $\epsilon_0$ and $\epsilon_1$ satisfying \autoref{eq:extalg-eq1}. Then
+Suppose we have $A$, $P$, $e_0$ and $e_1$ satisfying \autoref{eq:extalg-eq1}. Then
 the inference rules
@@
   { \jfameq
-      {{{{\Gamma}{A}}{P}}{\jcomp{}{\epsilon_0}{P}}}
-      {\jcomp{}{\epsilon_0}{\jcomp{}{\epsilon_0}{Q}}}
-      {\jcomp{}{\epsilon_1}{\jcomp{}{\epsilon_0}{Q}}}
+      {{{{\Gamma}{A}}{P}}{\jcomp{}{e_0}{P}}}
+      {\jcomp{}{e_0}{\jcomp{}{e_0}{Q}}}
+      {\jcomp{}{e_1}{\jcomp{}{e_0}{Q}}}
     }
@@
   { \jfameq
-      {{{{{\Gamma}{A}}{P}}{\jcomp{}{\epsilon_0}{P}}}
-        {\jcomp{}{\epsilon_0}{\jcomp{}{\epsilon_0}{Q}}}}
-      {\jcomp{}{\epsilon_0}{\jcomp{}{\epsilon_0}{R}}}
-      {\jcomp{}{\epsilon_1}{\jcomp{}{\epsilon_0}{R}}}
+      {{{{{\Gamma}{A}}{P}}{\jcomp{}{e_0}{P}}}
+        {\jcomp{}{e_0}{\jcomp{}{e_0}{Q}}}}
+      {\jcomp{}{e_0}{\jcomp{}{e_0}{R}}}
+      {\jcomp{}{e_1}{\jcomp{}{e_0}{R}}}
     }
@@
   { \jtermeq
-      {{{{{\Gamma}{A}}{P}}{\jcomp{}{\epsilon_0}{P}}}
-        {\jcomp{}{\epsilon_0}{\jcomp{}{\epsilon_0}{Q}}}}
-      {\jcomp{}{\epsilon_0}{\jcomp{}{\epsilon_0}{R}}}
-      {\jcomp{}{\epsilon_0}{\jcomp{}{\epsilon_0}{h}}}
-      {\jcomp{}{\epsilon_1}{\jcomp{}{\epsilon_0}{h}}}
+      {{{{{\Gamma}{A}}{P}}{\jcomp{}{e_0}{P}}}
+        {\jcomp{}{e_0}{\jcomp{}{e_0}{Q}}}}
+      {\jcomp{}{e_0}{\jcomp{}{e_0}{R}}}
+      {\jcomp{}{e_0}{\jcomp{}{e_0}{h}}}
+      {\jcomp{}{e_1}{\jcomp{}{e_0}{h}}}
     }
@@
 \begin{align*}
-\jcomp{}{\epsilon_0}{\jcomp{}{\epsilon_0}{Q}}
-& \jdeq
-  \jcomp{}{\idtm{\jcomp{}{\epsilon_0}{P}}}{\jcomp{}{\epsilon_0}{\jcomp{}{\epsilon_0}{Q}}}
+\jcomp{}{e_0}{\jcomp{}{e_0}{Q}}
+& \jdeq
+  \jcomp{}{\idtm{\jcomp{}{e_0}{P}}}{\jcomp{}{e_0}{\jcomp{}{e_0}{Q}}}
   \tag{by \autoref{idfunc-precomp}}
@@
 & \jdeq
-  \jcomp{}{\jvcomp{}{\epsilon_0}{\idtm{\jcomp{}{\epsilon_0}{P}}}}{\jcomp{}{\epsilon_0}{Q}}
+  \jcomp{}{\jvcomp{}{e_0}{\idtm{\jcomp{}{e_0}{P}}}}{\jcomp{}{e_0}{Q}}
   \tag{by \autoref{lem:composition-threesome}}
@@
 & \jdeq
-  \jcomp{}{\jcomp{}{\jvcomp{}{\epsilon_0}{\idtm{\jcomp{}{\epsilon_0}{P}}}}{\epsilon_0}}{Q}
+  \jcomp{}{\jcomp{}{\jvcomp{}{e_0}{\idtm{\jcomp{}{e_0}{P}}}}{e_0}}{Q}
   \tag{by \autoref{lem:jcomp-jcomp}}
@@
 & \jdeq
-  \jcomp{}{\jcomp{}{\jvcomp{}{\idtm{A}}{\epsilon_1}}{\epsilon_0}}{Q}
+  \jcomp{}{\jcomp{}{\jvcomp{}{\idtm{A}}{e_1}}{e_0}}{Q}
   \tag{by \autoref{eq:extalg-eq1}}
@@
 & \jdeq
-  \jcomp{}{\jvcomp{}{\idtm{A}}{\epsilon_1}}{\jcomp{}{\epsilon_0}{Q}}
+  \jcomp{}{\jvcomp{}{\idtm{A}}{e_1}}{\jcomp{}{e_0}{Q}}
   \tag{by \autoref{lem:jcomp-jcomp}}
@@
 & \jdeq
-  \jcomp{}{\epsilon_1}{\jcomp{}{\idtm{A}}{\jcomp{}{\epsilon_0}{Q}}}
+  \jcomp{}{e_1}{\jcomp{}{\idtm{A}}{\jcomp{}{e_0}{Q}}}
   \tag{by \autoref{lem:composition-threesome}}
@@
 & \jdeq
-  \jcomp{}{\epsilon_1}{\jcomp{}{\epsilon_0}{Q}}.
+  \jcomp{}{e_1}{\jcomp{}{e_0}{Q}}.
   \tag{by \autoref{idfunc-precomp}}
@@
   {{\Gamma}{A}}
-  {{P}{{\jcomp{}{\epsilon_0}{P}}{\jcomp{}{\epsilon_1}{{}{\epsilon_0}{P}}}}}
-  {{P}{\jcomp{}{\epsilon_0}{P}}}
-  {\jvcomp{}{\epsilon_1}{\idtm{\jcomp{}{\epsilon_1}{{}{\epsilon_0}{P}}}}}.
+  {{P}{{\jcomp{}{e_0}{P}}{\jcomp{}{e_1}{{}{e_0}{P}}}}}
+  {{P}{\jcomp{}{e_0}{P}}}
+  {\jvcomp{}{e_1}{\idtm{\jcomp{}{e_1}{{}{e_0}{P}}}}}.
 \end{equation*}
@@
 \begin{tikzcd}[column sep=huge]
-\ctxext{P}{{\jcomp{}{\epsilon_0}{P}}{\jcomp{}{\epsilon_0}{{}{\epsilon_0}{P}}}} 
-  \ar{r}{\jvcomp{}{\idtm{P}}{\jcomp{}{\epsilon_0}{\epsilon_1}}}
+\ctxext{P}{{\jcomp{}{e_0}{P}}{\jcomp{}{e_0}{{}{e_0}{P}}}} 
+  \ar{r}{\jvcomp{}{\idtm{P}}{\jcomp{}{e_0}{e_1}}}
   \ar{d}[swap]{
-    \jvcomp{}{\epsilon_1}{\idtm{\jcomp{}{\epsilon_1}{{}{\epsilon_0}{P}}}}
-    }
-& \ctxext{P}{\jcomp{}{\epsilon_0}{P}} \ar{d}{\epsilon_1}\\
-\ctxext{P}{\jcomp{}{\epsilon_0}{P}} \ar{r}[swap]{\epsilon_1} & P
+    \jvcomp{}{e_1}{\idtm{\jcomp{}{e_1}{{}{e_0}{P}}}}
+    }
+& \ctxext{P}{\jcomp{}{e_0}{P}} \ar{d}{e_1}\\
+\ctxext{P}{\jcomp{}{e_0}{P}} \ar{r}[swap]{e_1} & P
 \end{tikzcd}
@@
 \begin{defn}
-An \emph{extension algebra $\cftalg{A}$ in context $\Gamma$} is a quadruple 
-\begin{equation*}
-(\cftalgc{\cftalg{A}},\cftalgf{\cftalg{A}},\cftctxext{\cftalg{A}},\cftfamext{\cftalg{A}})
+An \emph{extension algebra $\extalg{A}$ in context $\Gamma$} is a quadruple 
+\begin{equation*}
+\unfold{\extalg{A}}
 \end{equation*}
@@
 
-We also give a bit of intuition to the requirement of \autoref{eq:extalg-eq2} by
-means of the following lemma.
+The following lemma both provides intuition behind the judgmental equalities
+we have required for context and family extension and it proves that context 
+and family extension in fact satisfy the compatibility rules stated in 
+\autoref{comp-ee}.
 
 \begin{lem}
-Let $(A,P,\epsilon_0,\epsilon_1)$ be an extension algebra in context $\Gamma$ and let
-$y_0:P$, $y_1:\subst{y_0}{\jcomp{}{\epsilon_0}{P}}$ and 
-$y_2:\subst{y_1}{{y_0}{\jcomp{}{\epsilon_0}{\jcomp{}{\epsilon_0}{P}}}}$. Then we have the
-judgmental equality
-\begin{equation*}
-\subst{{y_2}{{y_1}{\jcomp{}{\epsilon_0}{\epsilon_1}}}}{{y_0}{\epsilon_1}}
-  \jdeq
-  \subst{y_2}{{{y_1}{{y_0}{\epsilon_1}}}{\epsilon_1}}
-\end{equation*}
+Let $\extalg{A}$ be an extension algebra in context $\Gamma$.
+Then $\cftctxext{\extalg{A}}$ and $\cftfamext{\extalg{A}}$ 
+correctly interpret the rules for context
+extension and family extension.
+
+More precisely, for $x_0:\cftalgc{\cftalg{A}}$, $x_1:\subst{x_0}{\cftalgf{\cftalg{A}}}$ and 
+$x_2:\subst{{x_1}{{x_0}{\cftctxext{\extalg{A}}}}}{\cftalgf{\cftalg{A}}}$ we have the judgmental equality
+\begin{align*}
+\subst{{x_2}{{x_1}{\cftfamext{\extalg{A}}}}}{{x_0}{\cftctxext{\extalg{A}}}}
+& \jdeq
+\subst{x_2}{{{x_1}{{x_0}{\cftctxext{\extalg{A}}}}}{\cftctxext{\extalg{A}}}}.
+\intertext{and for $y_0:\cftalgf{\extalg{A}}$, $y_1:\subst{y_0}{\jcomp{}{\cftctxext{\extalg{A}}}{\cftalgf{\extalg{A}}}}$ and 
+$y_2:\subst{y_1}{{y_0}{\jcomp{}{\cftctxext{\extalg{A}}}{\jcomp{}{\cftctxext{\extalg{A}}}{\cftalgf{\extalg{A}}}}}}$
+we have the judgmental equality}
+\subst{{y_2}{{y_1}{\jcomp{}{\cftctxext{\extalg{A}}}{\cftfamext{\extalg{A}}}}}}{{y_0}{\cftfamext{\extalg{A}}}}
+& \jdeq
+  \subst{y_2}{{{y_1}{{y_0}{\cftfamext{\extalg{A}}}}}{\cftfamext{\extalg{A}}}}.
+\end{align*}
 \end{lem}
@@
 \begin{proof}
-The proof is a straightforward calculation:
+Both proofs are simple calculations. For the first judgmental equality we have
+\begin{align*}
+\subst{{x_2}{{x_1}{\epsilon_1}}}{{x_0}{\epsilon_0}}
+& \jdeq
+  \subst{\tmext{x_0}{\subst{x_2}{{x_1}{\epsilon_1}}}}{\epsilon_0}
+  \\
+& \jdeq 
+  \subst{{x_2}{{x_1}{{x_0}{\jvcomp{}{\idtm{A}}{\epsilon_1}}}}}{\epsilon_0}
+  \\
+& \jdeq
+  \subst{x_2}{{x_1}{{x_0}{\jcomp{}{\jvcomp{}{\idtm{A}}{\epsilon_1}}{\epsilon_0}}}}
+  \\
+& \jdeq
+  \subst{x_2}{{x_1}{{x_0}{\jcomp{}{\jvcomp{}{\epsilon_0}{\idtm{\jcomp{}{\epsilon_0}{P}}}}{\epsilon_0}}}}
+  \\
+& \jdeq 
+  \subst{{x_2}{{x_1}{{x_0}{\jvcomp{}{\epsilon_0}{\idtm{\jcomp{}{\epsilon_0}{P}}}}}}}{\epsilon_0}
+  \\
+& \jdeq
+  \subst{\tmext{\subst{x_1}{{x_0}{\epsilon_0}}}{x_2}}{\epsilon_0}
+  \\
+& \jdeq
+  \subst{x_2}{{{x_1}{{x_0}{\epsilon_0}}}{\epsilon_0}}.
+\end{align*}
+and to prove the second judgmental equality we calculate
 \begin{align*}

### 2014-05-23 article-models-tt.tex
tagging equations
a few more tags added
nearly finished with the section about fiber inclusions
initiated CwF section

@@
 Besides contexts, families and terms there will also be a notion of `type in
-a context'. Only a family of contexts over a context $\Gamma$ is eligible to
+a context', which one may assume.
+Only a family of contexts over a context $\Gamma$ is eligible to
 be a type in the context $\Gamma$ and we have the intuition that `being a type'
@@
 although it is certainly worth investigating the class of contexts which fall
-under this category.
+under this category. However, we will primarily investigate families of
+contexts and leave discussion on the possibility of a typing judgment until the
+end of this section. With families of contexts being the principal things of
+study, we have added a bit more generality to the theory, compared to the
+theory of categories with families presented in \cite{Dybjer1996}. The reason
+for this added generality is twofold:
+\begin{enumerate}
+\item We want to be very sure that we can interpret all the rules of type theory
+in a straightforward manner in any context.
+\item We do want do end up with a categorical structure on the contexts, derived
+from the theory by requiring that a morphism from $\Gamma$ to $\Delta$ is a
+term of the weakening of $\Delta$ by $\Gamma$ in context $\Gamma$. Thus, we
+must be able to lift the context $\Delta$ to become a family of contexts over
+$\Gamma$ and it is not reasonable to expect that this results in a type.
+\end{enumerate}
 
@@
     }
+  \label{comp-00-t}
 \end{align}
@@
     }
-  \\
+  \label{comp-w0-c}\\
 & \inference
@@
     }
+  \label{comp-w0-f}
 \end{align}
@@
     }
+  \label{comp-0w-c}
   \\
@@
     }
+  \label{comp-0w-f}
   \\
@@
     }
+  \label{comp-0w-t}
 \end{align}
@@
     }
+  \label{comp-s0-c}
   \\
@@
     }
+  \label{comp-s0-f}
 \end{align}
@@
     }
+  \label{comp-0s-c}
   \\
@@
     }
+  \label{comp-0s-f}
   \\
@@
     }.
+  \label{comp-0s-t}
 \end{align}
@@
 The identity term of a weakened family is the weakened identity term:
-\begin{equation}\label{idfunc-wk-comp}
+\begin{equation}
 \inference
@@
     }}
+  \label{comp-wi-t}
 \end{equation}
@@
 The identity term of a substituted family is the substitution of the identity term
-\begin{equation}\label{idfunc-subst-comp}
+\begin{equation}
 \inference
@@
     }}
+  \label{comp-si-t}
 \end{equation}
@@
     }
-  \label{defn-ws-3}\\
+  \label{cancellation-ws-f}\\
 & \inference
@@
     }
-  \label{defn-ws-4}
+  \label{cancellation-ws-t}
 \end{align}
@@
     }
-  \label{idfunc-subst-defn}\\
+  \label{cancellation-si}\\
 & \inference

### 2014-05-23 article-models-ttinternal.tex
initiated CwF section

@@
 $\ctxext{\Gamma}{\cftalgc{\cftalg{A}}}$,
-a morphism $\cftctxext{\cftalg{A}}$ from $\ctxext{\cftalgc{\cftalg{A}}}{
+a morphism $\cftctxext[\cftalg{A}]$ from $\ctxext{\cftalgc{\cftalg{A}}}{
 \cftalgf{\cftalg{A}}}$ to $\cftalgc{\cftalg{A}}$ in context
-$\Gamma$ and a morphism $\cftfamext{\cftalg{A}}$ from 
-$\ctxext{\cftalgf{\cftalg{A}}}{\jcomp{}{\cftctxext{\cftalg{A}}}{\cftalgf{\cftalg{A}}}}$ 
+$\Gamma$ and a morphism $\cftfamext[\cftalg{A}]$ from 
+$\ctxext{\cftalgf{\cftalg{A}}}{\jcomp{}{\cftctxext[\cftalg{A}]}{\cftalgf{\cftalg{A}}}}$ 
 to $\cftalgf{\cftalg{A}}$ in context
@@
 Let $\extalg{A}$ be an extension algebra in context $\Gamma$.
-Then $\cftctxext{\extalg{A}}$ and $\cftfamext{\extalg{A}}$ 
+Then $\cftctxext[\extalg{A}]$ and $\cftfamext[\extalg{A}]$ 
 correctly interpret the rules for context
@@
 
-More precisely, for $x_0:\cftalgc{\cftalg{A}}$, $x_1:\subst{x_0}{\cftalgf{\cftalg{A}}}$ and 
-$x_2:\subst{{x_1}{{x_0}{\cftctxext{\extalg{A}}}}}{\cftalgf{\cftalg{A}}}$ we have the judgmental equality
-\begin{align*}
-\subst{{x_2}{{x_1}{\cftfamext{\extalg{A}}}}}{{x_0}{\cftctxext{\extalg{A}}}}
-& \jdeq
-\subst{x_2}{{{x_1}{{x_0}{\cftctxext{\extalg{A}}}}}{\cftctxext{\extalg{A}}}}.
-\intertext{and for $y_0:\cftalgf{\extalg{A}}$, $y_1:\subst{y_0}{\jcomp{}{\cftctxext{\extalg{A}}}{\cftalgf{\extalg{A}}}}$ and 
-$y_2:\subst{y_1}{{y_0}{\jcomp{}{\cftctxext{\extalg{A}}}{\jcomp{}{\cftctxext{\extalg{A}}}{\cftalgf{\extalg{A}}}}}}$
+More precisely, for $x_0:\cftalgc{\cftalg{A}}$, 
+$x_1:\subst{x_0}{\cftalgf{\cftalg{A}}}$ and 
+$x_2:\subst{{x_1}{{x_0}{\cftctxext[\extalg{A}]}}}{\cftalgf{\cftalg{A}}}$ 
+we have the judgmental equality
+\begin{align*}
+\subst{{x_2}{{x_1}{\cftfamext[\extalg{A}]}}}{{x_0}{\cftfamext[\extalg{A}]}}
+& \jdeq
+  \subst
+    {x_2}
+    {{{x_1}{{x_0}{\cftfamext[\extalg{A}]}}}{\cftfamext[\extalg{A}]}}.
+  \intertext{%
+and for $y_0:\cftalgf{\extalg{A}}$, 
+$y_1:\subst{y_0}{\jcomp{}{\cftfamext[\extalg{A}]}{\cftalgf{\extalg{A}}}}$ and 
+$y_2:
+  \subst
+    {y_1}
+    { {y_0}
+      { \jcomp
+          {}
+          {\cftfamext[\extalg{A}]}
+          {\jcomp{}{\cftfamext[\extalg{A}]}{\cftalgf{\extalg{A}}}}
+        }
+      }
+$
 we have the judgmental equality}
-\subst{{y_2}{{y_1}{\jcomp{}{\cftctxext{\extalg{A}}}{\cftfamext{\extalg{A}}}}}}{{y_0}{\cftfamext{\extalg{A}}}}
-& \jdeq
-  \subst{y_2}{{{y_1}{{y_0}{\cftfamext{\extalg{A}}}}}{\cftfamext{\extalg{A}}}}.
+\subst
+  {{y_2}{{y_1}{\jcomp{}{\cftfamext[\extalg{A}]}{\cftfamext[\extalg{A}]}}}}
+  {{y_0}{\cftfamext[\extalg{A}]}}
+& \jdeq
+  \subst{y_2}{{{y_1}{{y_0}{\cftfamext[\extalg{A}]}}}{\cftfamext[\extalg{A}]}}.
 \end{align*}
@@
 \begin{align*}
-\subst{{x_2}{{x_1}{\epsilon_1}}}{{x_0}{\epsilon_0}}
-& \jdeq
-  \subst{\tmext{x_0}{\subst{x_2}{{x_1}{\epsilon_1}}}}{\epsilon_0}
-  \\
-& \jdeq 
-  \subst{{x_2}{{x_1}{{x_0}{\jvcomp{}{\idtm{A}}{\epsilon_1}}}}}{\epsilon_0}
-  \\
-& \jdeq
-  \subst{x_2}{{x_1}{{x_0}{\jcomp{}{\jvcomp{}{\idtm{A}}{\epsilon_1}}{\epsilon_0}}}}
-  \\
-& \jdeq
-  \subst{x_2}{{x_1}{{x_0}{\jcomp{}{\jvcomp{}{\epsilon_0}{\idtm{\jcomp{}{\epsilon_0}{P}}}}{\epsilon_0}}}}
-  \\
-& \jdeq 
-  \subst{{x_2}{{x_1}{{x_0}{\jvcomp{}{\epsilon_0}{\idtm{\jcomp{}{\epsilon_0}{P}}}}}}}{\epsilon_0}
-  \\
-& \jdeq
-  \subst{\tmext{\subst{x_1}{{x_0}{\epsilon_0}}}{x_2}}{\epsilon_0}
-  \\
-& \jdeq
-  \subst{x_2}{{{x_1}{{x_0}{\epsilon_0}}}{\epsilon_0}}.
+\subst{{x_2}{{x_1}{\cftfamext[\cftalg{A}]}}}{{x_0}{\cftctxext[\cftalg{A}]}}
+& \jdeq
+  \subst{\tmext{x_0}{\subst{x_2}{{x_1}{\cftfamext[\cftalg{A}]}}}}{\cftctxext[\cftalg{A}]}
+  \\
+& \jdeq 
+  \subst{{x_2}{{x_1}{{x_0}{\jvcomp{}{\idtm{\cftalgc{\cftalg{A}}}}{\cftfamext[\cftalg{A}]}}}}}{\cftctxext[\cftalg{A}]}
+  \\
+& \jdeq
+  \subst{x_2}{{x_1}{{x_0}{\jcomp{}{\jvcomp{}{\idtm{\cftalgc{\cftalg{A}}}}{\cftfamext[\cftalg{A}]}}{\cftctxext[\cftalg{A}]}}}}
+  \\
+& \jdeq
+  \subst{x_2}{{x_1}{{x_0}{\jcomp{}{\jvcomp{}{\cftctxext[\cftalg{A}]}{\idtm{\jcomp{}{\cftctxext[\cftalg{A}]}{\cftalgf{\cftalg{A}}}}}}{\cftctxext[\cftalg{A}]}}}}
+  \\
+& \jdeq 
+  \subst{{x_2}{{x_1}{{x_0}{\jvcomp{}{\cftctxext[\cftalg{A}]}{\idtm{\jcomp{}{\cftctxext[\cftalg{A}]}{\cftalgf{\cftalg{A}}}}}}}}}{\cftctxext[\cftalg{A}]}
+  \\
+& \jdeq
+  \subst{\tmext{\subst{x_1}{{x_0}{\cftctxext[\cftalg{A}]}}}{x_2}}{\cftctxext[\cftalg{A}]}
+  \\
+& \jdeq
+  \subst{x_2}{{{x_1}{{x_0}{\cftctxext[\cftalg{A}]}}}{\cftctxext[\cftalg{A}]}}.
 \end{align*}
@@
 \begin{align*}
-\subst{{y_2}{{y_1}{\jcomp{}{\epsilon_0}{\epsilon_1}}}}{{y_0}{\epsilon_1}}
-& \jdeq
-  \subst{\tmext{y_0}{\subst{y_2}{{y_1}{\jcomp{}{\epsilon_0}{\epsilon_1}}}}}{\epsilon_1}
-  \\
-& \jdeq
-  \subst{{y_2}{{y_1}{{y_0}{\jvcomp{}{\idtm{P}}{\jcomp{}{\epsilon_0}{\epsilon_1}}}}}}{\epsilon_1}
+\subst{{y_2}{{y_1}{\jcomp{}{\cftctxext[\cftalg{A}]}{\cftfamext[\cftalg{A}]}}}}{{y_0}{\cftfamext[\cftalg{A}]}}
+& \jdeq
+  \subst{\tmext{y_0}{\subst{y_2}{{y_1}{\jcomp{}{\cftctxext[\cftalg{A}]}{\cftfamext[\cftalg{A}]}}}}}{\cftfamext[\cftalg{A}]}
+  \\
+& \jdeq
+  \subst{{y_2}{{y_1}{{y_0}{\jvcomp{}{\idtm{P}}{\jcomp{}{\cftctxext[\cftalg{A}]}{\cftfamext[\cftalg{A}]}}}}}}{\cftfamext[\cftalg{A}]}
   \\
@@
       { {y_0}
-        {\jcomp{}{\jvcomp{}{\idtm{P}}{\jcomp{}{\epsilon_0}{\epsilon_1}}}{\epsilon_1}}
+        {\jcomp{}{\jvcomp{}{\idtm{P}}{\jcomp{}{\cftctxext[\cftalg{A}]}{\cftfamext[\cftalg{A}]}}}{\cftfamext[\cftalg{A}]}}
         }
@@
       { {y_0}
-        {\jcomp{}{\jvcomp{}{\epsilon_1}{\idtm{\jcomp{}{\epsilon_1}{{}{\epsilon_0}{P}}}}}{\epsilon_1}}
-        }
-      }
-  \\
-& \jdeq
-  \subst{{y_2}{{y_1}{{y_0}{\jvcomp{}{\epsilon_1}{\idtm{\jcomp{}{\epsilon_1}{{}{\epsilon_0}{P}}}}}}}}{\epsilon_1}
-  \\
-& \jdeq
-  \subst{\tmext{\subst{y_1}{{y_0}{\epsilon_1}}}{y_2}}{\epsilon_1}
-  \\
-& \jdeq
-  \subst{y_2}{{{y_1}{{y_0}{\epsilon_1}}}{\epsilon_1}}.\qedhere
+        {\jcomp{}{\jvcomp{}{\cftfamext[\cftalg{A}]}{\idtm{\jcomp{}{\cftfamext[\cftalg{A}]}{{}{\cftctxext[\cftalg{A}]}{\cftalgf{\cftalg{A}}}}}}}{\cftfamext[\cftalg{A}]}}
+        }
+      }
+  \\
+& \jdeq
+  \subst{{y_2}{{y_1}{{y_0}{\jvcomp{}{\cftfamext[\cftalg{A}]}{\idtm{\jcomp{}{\cftfamext[\cftalg{A}]}{{}{\cftctxext[\cftalg{A}]}{\cftalgf{\cftalg{A}}}}}}}}}}{\cftfamext[\cftalg{A}]}
+  \\
+& \jdeq
+  \subst{\tmext{\subst{y_1}{{y_0}{\cftfamext[\cftalg{A}]}}}{y_2}}{\cftfamext[\cftalg{A}]}
+  \\
+& \jdeq
+  \subst{y_2}{{{y_1}{{y_0}{\cftfamext[\cftalg{A}]}}}{\cftfamext[\cftalg{A}]}}.\qedhere
 \end{align*}
@@
 \begin{thm}
-Suppose that $(A,P,\epsilon_0,\epsilon_1)$ is an extension algebra in context
+Suppose that $\extalg{A}$ is an extension algebra in context
 $\Gamma$. Then 
-$(P,\jcomp{}{\epsilon_0}{P},\epsilon_1,\jcomp{}{\epsilon_0}{\epsilon_1})$
-is an extension algebra in context $\ctxext{\Gamma}{A}$.
+\begin{equation*}
+\cftalg{F_A}\defeq(\cftalgf{\cftalg{A}},\jcomp{}{\cftctxext}{\cftalgf{\cftalg{A}}},\cftfamext,\jcomp{}{\cftctxext}{\cftfamext})
+\end{equation*}
+is an extension algebra in context $\ctxext{\Gamma}{\cftalgc{\cftalg{A}}}$.
 \end{thm}
@@
 \begin{proof}
-We first need to verify that $\jcomp{}{\epsilon_0}{\epsilon_1}$ is a morphism
-from $\ctxext{\jcomp{}{\epsilon_0}{P}}{\jcomp{}{\epsilon_1}{{}{\epsilon_0}{P}}}$
-to $\jcomp{}{\epsilon_0}{P}$. This follows from the judgmental equality
-$\jcomp{}{\epsilon_1}{{}{\epsilon_0}{P}}\jdeq
-\jcomp{}{\epsilon_0}{{}{\epsilon_0}{P}}$, which we have proved in
+In this proof, we shall use the short-hand notations $\cftctxext$ and $\cftfamext$
+only to refer to $\cftctxext[\cftalg{A}]$ and $\cftfamext[\cftalg{A}]$, respectively,
+and not to $\cftctxext[\cftalg{F_A}]$ or $\cftfamext[\cftalg{F_A}]$.
+
+We first need to verify that the domain of the morphism 
+$\jcomp{}{\cftctxext}{\cftfamext}$ is indeed
+$\ctxext{\jcomp{}{\cftctxext}{\cftalgf{\cftalg{A}}}}{\jcomp{}{\cftfamext}{{}{\cftctxext}{\cftalgf{\cftalg{A}}}}}$. 
+This follows from the judgmental equality
+$\jcomp{}{\cftfamext}{{}{\cftctxext}{\cftalgf{\cftalg{A}}}}\jdeq
+\jcomp{}{\cftctxext}{{}{\cftctxext}{\cftalgf{\cftalg{A}}}}$, which we have proved in
 \autoref{lem:extalg-twins}. Notice how the diagram in \autoref{eq:extalg-eq2} is
 of exactly the right sort, so the quadruple
-$(P,\jcomp{}{\epsilon_0}{P},\epsilon_1,\jcomp{}{\epsilon_0}{\epsilon_1})$
+$(\cftalgf{\cftalg{A}},\jcomp{}{\cftctxext}{\cftalgf{\cftalg{A}}},\cftfamext,\jcomp{}{\cftctxext}{\cftfamext})$
 satisfies its version of \autoref{eq:extalg-eq1}. It is left to verify that the diagram
@@
 \ctxext
-  { \jcomp{}{\epsilon_0}{P}
-    }
-  { { \jcomp{}{\epsilon_1}{%
-        \jcomp{}{\epsilon_0}{P}
-        }
-      }
-    { \jcomp{}{\epsilon_1}{%
-        \jcomp{}{\epsilon_1}{%
-          \jcomp{}{\epsilon_0}{P}
+  { \jcomp{}{\cftctxext}{\cftalgf{\cftalg{A}}}
+    }
+  { { \jcomp{}{\cftfamext}{%
+        \jcomp{}{\cftctxext}{\cftalgf{\cftalg{A}}}
+        }
+      }
+    { \jcomp{}{\cftfamext}{%
+        \jcomp{}{\cftfamext}{%
+          \jcomp{}{\cftctxext}{\cftalgf{\cftalg{A}}}
           }
@@
   \ar{r}
-    { \jvcomp{}{\idtm{\jcomp{}{\epsilon_0}{P}}}{%
-        \jcomp{}{\epsilon_1}{%
-          \jcomp{}{\epsilon_0}{\epsilon_1}}}}
+    { \jvcomp{}{\idtm{\jcomp{}{\cftctxext}{\cftalgf{\cftalg{A}}}}}{%
+        \jcomp{}{\cftfamext}{%
+          \jcomp{}{\cftctxext}{\cftfamext}}}}
   \ar{d}[swap]{
-    \jvcomp{}{\jcomp{}{\epsilon_0}{\epsilon_1}}{%
+    \jvcomp{}{\jcomp{}{\cftctxext}{\cftfamext}}{%
       \idtm{
-        \jcomp{}{\jcomp{}{\epsilon_0}{\epsilon_1}}{%
-          \jcomp{}{\epsilon_1}{%
-            \jcomp{}{\epsilon_0}{P}
+        \jcomp{}{\jcomp{}{\cftctxext}{\cftfamext}}{%
+          \jcomp{}{\cftfamext}{%
+            \jcomp{}{\cftctxext}{\cftalgf{\cftalg{A}}}
             }
@@
 & \ctxext
-    {\jcomp{}{\epsilon_0}{P}}
-    {\jcomp{}{\epsilon_1}{\jcomp{}{\epsilon_0}{P}}} 
-  \ar{d}{\jcomp{}{\epsilon_0}{\epsilon_1}}
+    {\jcomp{}{\cftctxext}{\cftalgf{\cftalg{A}}}}
+    {\jcomp{}{\cftfamext}{\jcomp{}{\cftctxext}{\cftalgf{\cftalg{A}}}}} 
+  \ar{d}{\jcomp{}{\cftctxext}{\cftfamext}}
   \\
 \ctxext
-  {\jcomp{}{\epsilon_0}{P}}
-  {\jcomp{}{\epsilon_1}{\jcomp{}{\epsilon_0}{P}}} 
-  \ar{r}[swap]{\jcomp{}{\epsilon_0}{\epsilon_1}} 
-& \jcomp{}{\epsilon_0}{P}
+  {\jcomp{}{\cftctxext}{\cftalgf{\cftalg{A}}}}
+  {\jcomp{}{\cftfamext}{\jcomp{}{\cftctxext}{\cftalgf{\cftalg{A}}}}} 
+  \ar{r}[swap]{\jcomp{}{\cftctxext}{\cftfamext}} 
+& \jcomp{}{\cftctxext}{\cftalgf{\cftalg{A}}}
 \end{tikzcd}
 \end{equation*}
-\end{small}
+\end{small}%
 commutes judgmentally; this diagram is the version of \autoref{eq:extalg-eq2}
 for the quadruple
-$(P,\jcomp{}{\epsilon_0}{P},\epsilon_1,\jcomp{}{\epsilon_0}{\epsilon_1})$. Note
+$(\cftalgf{\cftalg{A}},\jcomp{}{\cftctxext}{\cftalgf{\cftalg{A}}},\cftfamext,\jcomp{}{\cftctxext}{\cftfamext})$. Note
 that this follows from \autoref{eq:extalg-eq2} provided that we can show that
 \begin{align}
-\jvcomp{}{\idtm{\jcomp{}{\epsilon_0}{P}}}{%
-  \jcomp{}{\epsilon_1}{%
-    \jcomp{}{\epsilon_0}{\epsilon_1}}}
-& \jdeq
-  \jcomp{}{\epsilon_0}{%
-    \jvcomp{}{\idtm{P}}{\jcomp{}{\epsilon_0}{\epsilon_1}}
+\jvcomp{}{\idtm{\jcomp{}{\cftctxext}{\cftalgf{\cftalg{A}}}}}{%
+  \jcomp{}{\cftfamext}{%
+    \jcomp{}{\cftctxext}{\cftfamext}}}
+& \jdeq
+  \jcomp{}{\cftctxext}{%
+    \jvcomp{}{\idtm{\cftalgf{\cftalg{A}}}}{\jcomp{}{\cftctxext}{\cftfamext}}
     }
@@
   \\
-\jvcomp{}{\jcomp{}{\epsilon_0}{\epsilon_1}}{%
+\jvcomp{}{\jcomp{}{\cftctxext}{\cftfamext}}{%
   \idtm{
-    \jcomp{}{\jcomp{}{\epsilon_0}{\epsilon_1}}{%
-      \jcomp{}{\epsilon_1}{%
-        \jcomp{}{\epsilon_0}{P}
-        }
-      }
-    }
-  }
-& \jdeq
-\jcomp{}{\epsilon_0}{%
-  \jvcomp{}{\epsilon_1}{%
+    \jcomp{}{\jcomp{}{\cftctxext}{\cftfamext}}{%
+      \jcomp{}{\cftfamext}{%
+        \jcomp{}{\cftctxext}{\cftalgf{\cftalg{A}}}
+        }
+      }
+    }
+  }
+& \jdeq
+\jcomp{}{\cftctxext}{%
+  \jvcomp{}{\cftfamext}{%
     \idtm{
-      \jcomp{}{\epsilon_1}{%
-        \jcomp{}{\epsilon_0}{P}
+      \jcomp{}{\cftfamext}{%
+        \jcomp{}{\cftctxext}{\cftalgf{\cftalg{A}}}
         }
@@
 \begin{equation*}
-\jcomp{}{\epsilon_1}{\jcomp{}{\epsilon_0}{\epsilon_1}}
+\jcomp{}{\cftfamext}{\jcomp{}{\cftctxext}{\cftfamext}}
   \jdeq
-  \jcomp{}{\epsilon_0}{\jcomp{}{\epsilon_0}{\epsilon_1}}.
+  \jcomp{}{\cftctxext}{\jcomp{}{\cftctxext}{\cftfamext}}.
 \end{equation*}
@@
 (\cftalgc{\cftalg{A}},\cftalgf{\cftalg{A}},\cftalgt{\cftalg{A}},
-\cftctxext{\cftalg{A}},\cftfamext{\cftalg{A}},\cftempc{\cftalg{A}},
+\cftctxext[\cftalg{A}],\cftfamext[\cftalg{A}],\cftempc{\cftalg{A}},
 \cftempf{\cftalg{A}})
@@
 $\Gamma$, satisfying additional judgmental equalities expressing the 
-compatibility of $\cftctxext{\cftalg{A}}$ and $\cftfamext{\cftalg{A}}$ with 
+compatibility of $\cftctxext[\cftalg{A}]$ and $\cftfamext[\cftalg{A}]$ with 
 $\cftempc{\cftalg{A}}$ and $\cftempf{\cftalg{A}}$.
@@
 \begin{align}
-\jtermeq*{{\Gamma}{\cftalgc{\cftalg{A}}}}{\ctxwk{\cftalgc{\cftalg{A}}}{\cftalgc{\cftalg{A}}}}{\subst{\cftempc{\cftalg{A}}}{\cftctxext{\cftalg{A}}}}{\idtm{\cftalgc{\cftalg{A}}}}
+\jtermeq*{{\Gamma}{\cftalgc{\cftalg{A}}}}{\ctxwk{\cftalgc{\cftalg{A}}}{\cftalgc{\cftalg{A}}}}{\subst{\cftempc{\cftalg{A}}}{\cftctxext[\cftalg{A}]}}{\idtm{\cftalgc{\cftalg{A}}}}
   \label{extempalg-eq1}
   \\
-\jtermeq*{{\Gamma}{\cftalgc{\cftalg{A}}}}{\ctxwk{\cftalgc{\cftalg{A}}}{\cftalgc{\cftalg{A}}}}{\subst{\cftempf{\cftalg{A}}}{\cftctxext{\cftalg{A}}}}{\idtm{\cftalgc{\cftalg{A}}}}
+\jtermeq*{{\Gamma}{\cftalgc{\cftalg{A}}}}{\ctxwk{\cftalgc{\cftalg{A}}}{\cftalgc{\cftalg{A}}}}{\subst{\cftempf{\cftalg{A}}}{\cftctxext[\cftalg{A}]}}{\idtm{\cftalgc{\cftalg{A}}}}
   \label{extempalg-eq2}
@@
 extension homomorphism from
-$(\cftalgc{\cftalg{A}},\cftalgf{\cftalg{A}},\cftctxext{\cftalg{A}},\cftfamext{\cftalg{A}})$ 
+$(\cftalgc{\cftalg{A}},\cftalgf{\cftalg{A}},\cftctxext[\cftalg{A}],\cftfamext[\cftalg{A}])$ 
 to 
-$(\cftalgc{\cftalg{B}},\cftalgf{\cftalg{B}},\cftctxext{\cftalg{B}},\cftfamext{\cftalg{B}})$ 
+$(\cftalgc{\cftalg{B}},\cftalgf{\cftalg{B}},\cftfamext[\cftalg{B}],\cftfamext[\cftalg{B}])$ 
 in context $\Gamma$

### 2014-05-26 article-models-ttinternal.tex
finished updating section 4.1

@@
 
-\subsection{Extension algebras}
+\subsection{Extension algebras}\label{sec:extension-algebras}
 In this subsection our goal is to define the notion of extension algebras,
@@
 formulation and it is (perhaps surprisingly) quite involved to formulate it.
+The definition of extension algebras is given in \autoref{defn:extension-algebras}.
 
@@
 
-\begin{defn}
+\begin{defn}\label{defn:extension-algebras}
 An \emph{extension algebra $\extalg{A}$ in context $\Gamma$} is a quadruple 
@@
 \begin{thm}
-Let $(Q,R,\eta_0,\eta_1)$ be an extension algebra in context $\ctxext{\Gamma}{B}$ and let
-$\jfam{\Gamma}{A}$. Then the quadruple
-\begin{equation*}
-(\ctxwk{A}{Q},\ctxwk{A}{R},\ctxwk{A}{\eta_0},\ctxwk{A}{\eta_1})
+Let $\extalg{Q}$ be an extension algebra in context $\ctxext{\Gamma}{B}$ and let
+$\jfam{\Gamma}{A}$ be a family of contexts. Then the quadruple
+\begin{equation*}
+\ctxwk{A}{\extalg{Q}}
+  \defeq
+  (\ctxwk{A}{\cftalgc{\cftalg{Q}}},\ctxwk{A}{\cftalgf{\cftalg{Q}}},\ctxwk{A}{\cftctxext[\cftalg{Q}]},\ctxwk{A}{\cftfamext[\cftalg{Q}]})
 \end{equation*}
@@
 \begin{thm}
-Let $(Q,R,\eta_0,\eta_1)$ be an extension algebra in context $\ctxext{{\Gamma}{A}}{P}$
-and let $\jterm{\Gamma}{A}{x}$. Then the quadruple
-\begin{equation*}
-(\subst{x}{Q},\subst{x}{R},\subst{x}{\eta_0},\subst{x}{\eta_1})
+Let $\extalg{Q}$ be an extension algebra in context $\ctxext{{\Gamma}{A}}{P}$
+and let $\jterm{\Gamma}{A}{x}$ be a term. Then the quadruple
+\begin{equation*}
+\subst{x}{\cftalg{Q}}
+  \defeq
+  (\subst{x}{\cftalgc{\cftalg{Q}}},\subst{x}{\cftalgf{\cftalg{Q}}},\subst{x}{\cftctxext[\cftalg{Q}]},\subst{x}{\cftfamext[\cftalg{Q}]})
 \end{equation*}
@@
 \begin{cor}
-Let $(Q,R,\eta_0,\eta_1)$ be an extension algebra in context $\ctxext{\Gamma}{B}$
+Let $\extalg{Q}$ be an extension algebra in context $\ctxext{\Gamma}{B}$
 and let $\jhom{\Gamma}{A}{B}{f}$. Then the quadruple
 \begin{equation*}
-(\jcomp{A}{f}{Q},\jcomp{A}{f}{R},\jcomp{A}{f}{\eta_0},\jcomp{A}{f}{\eta_1})
+\jcomp{A}{f}{\extalg{Q}}
+  \defeq
+  (\jcomp{A}{f}{\cftalgc{\cftalg{Q}}},\jcomp{A}{f}{\cftalgf{\cftalg{Q}}},\jcomp{A}{f}{\cftctxext[\cftalg{Q}]},\jcomp{A}{f}{\cftfamext[\cftalg{Q}]})
 \end{equation*}
@@
 \begin{defn}
-An \emph{extension homomorphism} from $(A,P,\epsilon_0,\epsilon_1)$ to
-$(B,Q,\eta_0,\eta_1)$ in context $\Gamma$ is a pair $(f_0,f_1)$ consisting of
-\begin{align*}
-\jhom*{\Gamma}{A}{B}{f_0}
-  \\
-\jfhom*{\Gamma}{A}{B}{f_0}{P}{Q}{f_1}
+An \emph{extension homomorphism $\exthom{f}$ from $\extalg{A}$ to
+$\extalg{B}$ in context $\Gamma$} is a pair $\unfold{\exthom{f}}$ consisting of
+\begin{align*}
+\jhom*{\Gamma}{\cftalgc{\cftalg{A}}}{\cftalgc{\cftalg{B}}}{\cfthomc{\cfthom{f}}}
+  \\
+\jfhom*{\Gamma}{\cftalgc{\cftalg{A}}}{\cftalgc{\cftalg{B}}}{\cfthomc{\cfthom{f}}}{\cftalgf{\cftalg{A}}}{\cftalgf{\cftalg{B}}}{\cfthomf{\cfthom{f}}}
 \end{align*}
@@
 \begin{tikzcd}
-\ctxext{A}{P}
-  \ar{r}{\jvcomp{}{f_0}{f_1}}
-  \ar{d}[swap]{\epsilon_0}
-& \ctxext{B}{Q}
-  \ar{d}{\eta_0}
-  \\
-A
-  \ar{r}[swap]{f_0}
-& B
+\ctxext{\cftalgc{\cftalg{A}}}{\cftalgf{\cftalg{A}}}
+  \ar{r}{\jvcomp{}{\cfthomc{\cfthom{f}}}{\cfthomf{\cfthom{f}}}}
+  \ar{d}[swap]{\cftctxext[\cftalg{A}]}
+& \ctxext{\cftalgc{\cftalg{B}}}{\cftalgf{\cftalg{B}}}
+  \ar{d}{\cftctxext[\cftalg{B}]}
+  \\
+\cftalgc{\cftalg{A}}
+  \ar{r}[swap]{\cfthomc{\cfthom{f}}}
+& \cftalgc{\cftalg{A}}
 \end{tikzcd}
@@
 \begin{tikzcd}[column sep=large]
-\ctxext{P}{\jcomp{}{\epsilon_0}{P}}
-  \ar{r}{\jvcomp{}{f_1}{\jcomp{}{\epsilon_0}{f_1}}}
-  \ar{d}[swap]{\epsilon_1}
-& \jcomp{}{f_0}{\ctxext{Q}{\jcomp{}{\eta_0}{Q}}}
-  \ar{d}{\jcomp{}{f_0}{\eta_1}}
-  \\
-P
-  \ar{r}[swap]{f_1}
-& \jcomp{}{f_0}{Q}
+\ctxext{\cftalgf{\cftalg{A}}}{\jcomp{}{\cftctxext[\cftalg{A}]}{\cftalgf{\cftalg{A}}}}
+  \ar{r}{\jvcomp{}{\cfthomf{\cfthom{f}}}{\jcomp{}{\cftctxext[\cftalg{A}]}{\cfthomf{\cfthom{f}}}}}
+  \ar{d}[swap]{\cftfamext[\cftalg{A}]}
+& \jcomp{}{\cfthomc{\cfthom{f}}}{\ctxext{\cftalgf{\cftalg{B}}}{\jcomp{}{\cftctxext[\cftalg{B}]}{\cftalgf{\cftalg{B}}}}}
+  \ar{d}{\jcomp{}{\cfthomc{\cfthom{f}}}{\cftfamext[\cftalg{B}]}}
+  \\
+\cftalgf{\cftalg{A}}
+  \ar{r}[swap]{\cfthomf{\cfthom{f}}}
+& \jcomp{}{\cfthomc{\cfthom{f}}}{\cftalgf{\cftalg{B}}}
 \end{tikzcd}
@@
 \begin{align*}
-\jcomp{}{f_1}{\jcomp{}{f_0}{\jcomp{}{\eta_0}{Q}}}
-& \jdeq 
-  \jcomp{}{\jvcomp{}{f_0}{f_1}}{\jcomp{}{\eta_0}{Q}}
+\jcomp{}{\cfthomf{\cfthom{f}}}{\jcomp{}{\cfthomc{\cfthom{f}}}{\jcomp{}{\cftctxext[\cftalg{B}]}{\cftalgf{\cftalg{B}}}}}
+& \jdeq 
+  \jcomp{}{\jvcomp{}{\cfthomc{\cfthom{f}}}{\cfthomf{\cfthom{f}}}}{\jcomp{}{\cftctxext[\cftalg{B}]}{\cftalgf{\cftalg{B}}}}
   \tag{by \autoref{lem:composition-threesome}}
@@
 & \jdeq
-  \jcomp{}{\jcomp{}{\jvcomp{}{f_0}{f_1}}{\eta_0}}{Q}
+  \jcomp{}{\jcomp{}{\jvcomp{}{\cfthomc{\cfthom{f}}}{\cfthomf{\cfthom{f}}}}{\cftctxext[\cftalg{B}]}}{\cftalgf{\cftalg{B}}}
   \tag{by \autoref{lem:jcomp-jcomp}}
@@
 & \jdeq
-  \jcomp{}{\jcomp{}{\epsilon_0}{f_0}}{Q}
+  \jcomp{}{\jcomp{}{\cftctxext[\cftalg{A}]}{\cfthomc{\cfthom{f}}}}{\cftalgf{\cftalg{B}}}
   \tag{by \autoref{eq:extalg-hom1}}
@@
 & \jdeq
-  \jcomp{}{\epsilon_0}{\jcomp{}{f_0}{Q}}.
+  \jcomp{}{\cftctxext[\cftalg{A}]}{\jcomp{}{\cfthomc{\cfthom{f}}}{\cftalgf{\cftalg{B}}}}.
   \tag{by \autoref{lem:jcomp-jcomp}}
 \end{align*}
-and we indeed have the morphism $\jcomp{}{\epsilon_0}{f_1}$ from 
-$\jcomp{}{\epsilon_0}{P}$ to $\jcomp{}{\epsilon_0}{\jcomp{}{f_0}{Q}}$.
+and we indeed have the morphism $\jcomp{}{\cftctxext[\cftalg{A}]}{\cfthomf{\cfthom{f}}}$ from 
+$\jcomp{}{\cftctxext[\cftalg{A}]}{\cftalgf{\cftalg{A}}}$ to $\jcomp{}{\cftctxext[\cftalg{A}]}{\jcomp{}{\cfthomc{\cfthom{f}}}{\cftalgf{\cftalg{B}}}}$.
 \end{rmk}
 
+\begin{comment}
 \begin{rmk}

### 2014-05-26 article-models-ttinternal.tex
initiating proof of thm about cft homomorphisms

@@
 \begin{equation}\label{eq:extalg-hom2}
-\begin{tikzcd}[column sep=large]
+\begin{tikzcd}[column sep=huge]
 \ctxext{\cftalgf{\cftalg{A}}}{\jcomp{}{\cftctxext[\cftalg{A}]}{\cftalgf{\cftalg{A}}}}
@@
 \end{rmk}
+
+\begin{thm}
+Let $\cfthom{f}$ be an extension homomorphism from $\cftalg{A}$ to $\cftalg{B}$
+in context $\Gamma$. Then
+\begin{equation*}
+\cfthom{F_f}
+  \defeq
+  (\cfthomf{\cfthom{f}},\jcomp{}{\cftctxext[\cftalg{A}]}{\cfthomf{\cfthom{f}}})
+\end{equation*}
+is an extension homomorphism from $\cftalg{F_A}$ to $\jcomp{}{\cfthomc{\cfthom{f}}}{\cftalg{F_B}}$
+in context $\ctxext{\Gamma}{\cftalgc{\cftalg{A}}}$. 
+\end{thm}
+
+\begin{proof}
+We have to verify that the diagrams in \autoref{eq:extalg-hom1,eq:extalg-hom2}
+commute for $\cfthom{F_f}$. Unfolding the ingredients of \autoref{eq:extalg-hom1}
+for $\cfthom{F_f}$ gives us quite directly \autoref{eq:extalg-hom2}.
+\end{proof}
 
@@
   {\cftalgc{\cftalg{B}}}
-  {\subst{\cftemp{\cftalg{A}}}{\cfthomc{\cfthom{f}}}}
+  {\subst{\cftempc{\cftalg{A}}}{\cfthomc{\cfthom{f}}}}
   {\cftempc{\cftalg{B}}}

### 2014-05-27 article-models-ttinternal.tex
working

@@
 \end{equation*}
-consisting of a family $\cftalgc{\cftalg{A}}$ over context $\Gamma$, 
-a family $\cftalgf{\cftalg{A}}$ over the context
+consisting of 
+\begin{enumerate}
+\item a family $\cftalgc{\cftalg{A}}$ over context $\Gamma$, 
+\item a family $\cftalgf{\cftalg{A}}$ over the context
 $\ctxext{\Gamma}{\cftalgc{\cftalg{A}}}$,
-a morphism $\cftctxext[\cftalg{A}]$ from $\ctxext{\cftalgc{\cftalg{A}}}{
+\item a morphism $\cftctxext[\cftalg{A}]$ from $\ctxext{\cftalgc{\cftalg{A}}}{
 \cftalgf{\cftalg{A}}}$ to $\cftalgc{\cftalg{A}}$ in context
-$\Gamma$ and a morphism $\cftfamext[\cftalg{A}]$ from 
+$\Gamma$
+\item a morphism $\cftfamext[\cftalg{A}]$ from 
 $\ctxext{\cftalgf{\cftalg{A}}}{\jcomp{}{\cftctxext[\cftalg{A}]}{\cftalgf{\cftalg{A}}}}$ 
 to $\cftalgf{\cftalg{A}}$ in context
-$\ctxext{\Gamma}{\cftalgc{\cftalg{A}}}$, satisfying the judgmental equalities of
+$\ctxext{\Gamma}{\cftalgc{\cftalg{A}}}$,
+\end{enumerate}
+satisfying the judgmental equalities of
 \autoref{eq:extalg-eq1,eq:extalg-eq2}.
+
+When $\extalg{A}$ is an extension algebra in context $\Gamma$, we also refer to
+$\cftctxext[\cftalg{A}]$ as the \emph{context extension of $\extalg{A}$} and to
+$\cftfamext[\cftalg{A}]$ as the \emph{family extension of $\extalg{A}$}.
 \end{defn}
@@
 
-The following lemma explains how each extension algebra gives rise to infinitely
-many extension algebras.
+As a consequence of the following theorem, every extension algebra gives rise
+to infinitely many extension algebras by constructing the extension algebra
+of families of $\cftalg{A}$ for each extension algebra $\cftalg{A}$. Notice
+that this also explains that the family $\jcomp{}{\cftctxext}{\cftalgf{\cftalg{A}}}$
+over $\ctxext{{\Gamma}{\cftalgc{\cftalg{A}}}}{\cftalgf{\cftalg{A}}}$ is the family
+of families over families. Likewise, higher families are obtained by pulling
+back more times along $\cftctxext[\cftalg{A}]$.
 
@@
 commute for $\cfthom{F_f}$. Unfolding the ingredients of \autoref{eq:extalg-hom1}
-for $\cfthom{F_f}$ gives us quite directly \autoref{eq:extalg-hom2}.
+for $\cfthom{F_f}$ gives us quite directly \autoref{eq:extalg-hom2}. The
+diagram in \autoref{eq:extalg-hom2} for $\cfthom{F_f}$ is judgmentally equal
+to the pullback of everything in the diagram in \autoref{eq:extalg-hom2} for
+$\cfthom{f}$ by $\epsilon_0$, and therefore it commutes too.
 \end{proof}
@@
 
+Below, we give the definition of empty-object homomorphisms. They have, like
+extension homomorphisms, an action on contexts and an action on families and
+in addition they have an action on terms. Thus, we can picture a term
+homomorphism as follows:
+\begin{equation*}
+\begin{tikzcd}
+\cftalgt{\cftalg{A}}
+  \ar[fib]{d}
+  \ar{r}{\cfthomt{\cfthom{f}}}
+& \cftalgt{\cftalg{B}}
+  \ar[fib]{d}
+  \\
+\cftalgf{\cftalg{A}}
+  \ar[fib]{d}
+  \ar{r}{\cfthomf{\cfthom{f}}}
+& \cftalgf{\cftalg{B}}
+  \ar[fib]{d}
+  \\
+\cftalgc{\cftalg{A}}
+  \ar{r}[swap]{\cfthomc{\cfthom{f}}}
+& \cftalgc{\cftalg{B}}
+\end{tikzcd}
+\end{equation*}
+
 \begin{defn}
-Let $\cftalg{A}$ and $\cftalg{B}$ be empty-object algebras in context $\Gamma$.
+Let $\cftalg{A}$ and $\cftalg{B}$ be term algebras in context $\Gamma$.
 An \emph{empty-object homomorphism $\cfthom{f}$ from $\cftalg{A}$ to $\cftalg{B}$
-in context $\Gamma$} is a triple $(\cfthomc{\cfthom{f}},\cfthomf{\cfthom{f}},\cfthomt{\cfthom{f}})$ consisting of
+in context $\Gamma$} is a triple 
+$(\cfthomc{\cfthom{f}},\cfthomf{\cfthom{f}},\cfthomt{\cfthom{f}})$ consisting of
 \begin{align*}
@@
 \end{align*}
-as displayed in \autoref{fig:empty-object-hom}, satisfying the judgmental equalities
+satisfying the judgmental equalities
 \begin{align*}
@@
 
-\begin{figure}
-\begin{equation*}
-\begin{tikzcd}
-\cftalgt{\cftalg{A}}
-  \ar[fib]{d}
-  \ar{r}{\cfthomt{\cfthom{f}}}
-& \cftalgt{\cftalg{B}}
-  \ar[fib]{d}
-  \\
-\cftalgf{\cftalg{A}}
-  \ar[fib]{d}
-  \ar{r}{\cfthomf{\cfthom{f}}}
-& \cftalgf{\cftalg{B}}
-  \ar[fib]{d}
-  \\
-\cftalgc{\cftalg{A}}
-  \ar{r}[swap]{\cfthomc{\cfthom{f}}}
-& \cftalgc{\cftalg{B}}
-\end{tikzcd}
-\end{equation*}
-\caption{\label{fig:empty-object-hom}The ingredients of an empty-object homomorphism $\cfthom{f}$ from 
-$\cftalg{A}$ to $\cftalg{B}$ in context $\Gamma$.}
-\end{figure}
-
 The following lemma presents a property of empty-object algebras that will get
@@
 the quadruple $\unfold{\extalg{A}}$ is an extension algebra in context 
-$\Gamma$, the quintuple $\unfold{\empalg{A}}$ is an empty algebra in context
+$\Gamma$, the quintuple $\unfold{\empalg{A}}$ is an empty-object algebra in context
 $\Gamma$, satisfying additional judgmental equalities expressing the 
@@
 extension acts as the identity term when applied to the empty object in one of
-its arguments. The other expresses that extension of families restricted to
-families above the empty object is the same as extension of contexts.
+its arguments. The other expresses that family extension restricted to
+families above the empty object is the same as context extension. To be able
+to state these, we need to know that $\unfold{\empalg{A}}$ is an empty-object
+algebra. In the following, we verify step-by-step that we can indeed make the
+suggested requirements.
+\begin{enumerate}
+\item The type of $\subst{\cftempc{\cftalg{A}}}{\cftctxext[\cftalg{A}]}$ is
+$\subst
+  { \cftempc{\cftalg{A}}
+    }
+  { \ctxwk
+      {\ctxext{\cftalgc{\cftalg{A}}}{\cftalgf{\cftalg{A}}}}
+      {\cftalgc{\cftalg{A}}}
+    }$.
+We have the judgmental equalities
+\begin{align*}
+\subst
+  { \cftempc{\cftalg{A}}
+    }
+  { \ctxwk
+      {\ctxext{\cftalgc{\cftalg{A}}}{\cftalgf{\cftalg{A}}}}
+      {\cftalgc{\cftalg{A}}}
+    }
+& \jdeq
+  \subst
+    { \cftempc{\cftalg{A}}
+      }
+    { \ctxwk
+        { \cftalgf{\cftalg{A}}
+          }
+        { {\cftalgc{\cftalg{A}}}
+          {\cftalgc{\cftalg{A}}}
+          }
+      }
+  \tag{by \autoref{comp-ew-f}}
+  \\
+& \jdeq
+  \ctxwk
+    { \subst{\cftempc{\cftalg{A}}}{\cftalgf{\cftalg{A}}}
+      }
+    { \subst{\cftempc{\cftalg{A}}}{\ctxwk{\cftalgc{\cftalg{A}}}{\cftalgc{\cftalg{A}}}}
+      }
+  \tag{by \autoref{comp-sw-f}}
+\end{align*}
+\end{enumerate}
 
@@
 \autoref{empalg-eq1} and to compute $\subst{\phi_0}{\jcomp{}{\epsilon_0}{P}}$
-we use the following lemma.
-
-As an immediate corollary, if we assume the judgmental equalities
-\autoref{extempalg-eq1,extempalg-eq2} we get that 
+we use \autoref{lem:empalg-mor}. Note that in combination with 
+\autoref{extempalg-eq1} and \autoref{extempalg-eq2} we get the judgmental
+equalities
 \begin{equation}\label{cor:empalg-mor}
@@
 \end{equation}
-and hence that $\subst{\phi_0}{\epsilon_1}$ is a 
-morphism from $\ctxext{A}{P}$ to $A$. Thus, we can require that
+for $i$ being $\mathsf{c}$ or $\mathsf{f}$. Consequently, 
+$\subst{\phi_0}{\epsilon_1}$ is a morphism from $\ctxext{A}{P}$ to $A$, so we 
+can require that
 \begin{equation}\label{extempalg-eq3}
-\jhomeq{\Gamma}{{A}{P}}{A}{\subst{\phi_0}{\epsilon_1}}{\epsilon_0},
+\jhomeq{\Gamma}{{A}{P}}{A}{\subst{\phi_0}{\epsilon_1}}{\epsilon_0}.
 \end{equation}
-expressing that the action of extension at the level of families, restricted to the empty object
-$\phi_0$, is just extension at the level of contexts. We also have two judgmental
+This judgmental equality expresses that family extension restricted to the empty object
+$\phi_0$ is context extension. We also have two judgmental
 equalities expressing that $\epsilon_1$ restricted to $\phi_1$ and

### 2014-06-03 article-models-ttderived.tex
added remark about diagrams of type theory

@@
 
+Because composition is proven to be associative, we can draw diagrams involving
+families and morphisms between them in a context $\Gamma$, just as is usual in
+category theory and the order in which the morphisms are composed has no
+relevance. We can also speak about pullback diagrams, with universality in a
+judgmental sense.
+
 \subsection{Projections and extension on terms}\label{extension-on-terms}

### 2014-06-03 article-models-ttderived.tex
added ref to sec:fhom

@@
 relevance. We can also speak about pullback diagrams, with universality in a
-judgmental sense.
+judgmental sense. In \autoref{sec:fhom} we will introduce a more general kind
+of diagrams which may also include families and `morphisms over morphisms'.
 
@@
 
-\subsection{Family morphisms}
+\subsection{Family morphisms}\label{sec:fhom}
 In this section we are going to develop dependent morphisms and analyze various

### 2014-06-04 article-models-ttconstr.tex
working on extension algebras

@@
 \label{compatibility-of-type-constructors}
+
+\begin{comment}
+\section{Intensional algebras for the theory of contexts, families and terms}
+In this section we investigate intensional algebras for the theory of contexts,
+families and terms. These are algebras in a weak sense: the judgmental
+equalities are replaced by propositional equalities and in some cases coherence
+properties are added. The goal writing intensional algebras is twofold: first,
+it allows for easy formalization in existing proof assistants such as \Coq;
+second, we are shooting at an internal definition of internal weak 
+$\infty$-groupoids.
+
+\subsection{Intensional extension algebras}
+
+\begin{defn}
+An intensional extension algebra in context $\Gamma$ consists of
+\begin{enumerate}
+\item A type $C$ in context $\Gamma$
+\item A type family $F:C\to\type$
+\item A function $\epsilon_0:(\sm{x:C}F(x))\to C$.
+\item A function $\epsilon_1:\prd{x:C}(\sm{y:F(x)}F(\epsilon_0(x,y)))\to F(x)$
+\end{enumerate}
+\end{defn}
+\end{comment}

### 2014-06-04 article-models-ttinternal.tex
added remark about diagrams of type theory
working on extension algebras

@@
 \begin{defn}\label{defn:extension-algebras}
-An \emph{extension algebra $\extalg{A}$ in context $\Gamma$} is a quadruple 
-\begin{equation*}
-\unfold{\extalg{A}}
-\end{equation*}
-consisting of 
-\begin{enumerate}
-\item a family $\cftalgc{\cftalg{A}}$ over context $\Gamma$, 
-\item a family $\cftalgf{\cftalg{A}}$ over the context
-$\ctxext{\Gamma}{\cftalgc{\cftalg{A}}}$,
-\item a morphism $\cftctxext[\cftalg{A}]$ from $\ctxext{\cftalgc{\cftalg{A}}}{
-\cftalgf{\cftalg{A}}}$ to $\cftalgc{\cftalg{A}}$ in context
-$\Gamma$
-\item a morphism $\cftfamext[\cftalg{A}]$ from 
-$\ctxext{\cftalgf{\cftalg{A}}}{\jcomp{}{\cftctxext[\cftalg{A}]}{\cftalgf{\cftalg{A}}}}$ 
-to $\cftalgf{\cftalg{A}}$ in context
-$\ctxext{\Gamma}{\cftalgc{\cftalg{A}}}$,
-\end{enumerate}
+We define the judgment
+\begin{equation*}
+\jextalg{\Gamma}{A}
+\end{equation*}
+asserting that $\extalg{A}$ is an \emph{extension algebra in context $\Gamma$}
+to be the conjunction of the following six judmgents:
+\begin{align*}
+\jextalgctx*{\Gamma}{A}
+  \\
+\jextalgfam*{\Gamma}{A}
+  \\
+\jextalgctxext*{\Gamma}{A}
+  \\
+\jextalgfamext*{\Gamma}{A}
+  \\
+\jhomeq*
+  { \Gamma}
+  { {\cftalgc{\cftalg{A}}}
+    { {\cftalgf{\cftalg{A}}}
+      {\jcomp{}{\cftctxext[\cftalg{A}]}{\cftalgf{\cftalg{A}}}}
+      }
+    }
+  { \cftalgc{\cftalg{A}}}
+  { \jcomp{}
+      { \jvcomp{}
+          {\cftctxext[\cftalg{A}]}
+          {\idtm{\jcomp{}{\cftctxext[\cftalg{A}]}{\cftalgf{\cftalg{A}}}}}
+        }
+      { \cftctxext[\cftalg{A}]}
+    }
+  { \jcomp{}
+      { \jvcomp{}
+          {\idtm{\cftalgc{\cftalg{A}}}}
+          {\cftfamext[\cftalg{A}]}
+        }
+      { \cftctxext[\cftalg{A}]}
+    }
+  \\
+\jhomeq*
+  { {\Gamma}{\cftalgc{\cftalg{A}}}}
+  { { \cftalgf{\cftalg{A}}}
+    { { \jcomp{}{\cftctxext[\cftalg{A}]}{\cftalgf{\cftalg{A}}}}
+      { \jcomp{}
+          {\cftctxext[\cftalg{A}]}
+          {{}{\cftctxext[\cftalg{A}]}{\cftalgf{\cftalg{A}}}}
+        }
+      }
+    }
+  { \cftalgf{\cftalg{A}}}
+  { \jcomp{}
+      { \cftfamext[\cftalg{A}]}
+      { \jvcomp{}
+          {\cftfamext[\cftalg{A}]}
+          {\idtm{\jcomp{}{\cftfamext[\cftalg{A}]}{{}{\cftctxext[\cftalg{A}]}{\cftalgf{\cftalg{A}}}}}}}
+    }
+  { \jcomp{}
+      { \cftfamext[\cftalg{A}]}
+      { \jvcomp{}
+          {\idtm{\cftalgf{\cftalg{A}}}}
+          {\jcomp{}{\cftctxext[\cftalg{A}]}{\cftfamext[\cftalg{A}]}}
+        }
+    }
+\end{align*}
+In other words, an extension algebra $\extalg{A}$ in context $\Gamma$
+is a quadruple $\unfold{\extalg{A}}$ 
 satisfying the judgmental equalities of
-\autoref{eq:extalg-eq1,eq:extalg-eq2}.
+\autoref{eq:extalg-eq1,eq:extalg-eq2}. Extension algebras are judgmentally
+equal if they are componentwise judgmentally equal.
 
@@
 
-\begin{thm}
+\begin{thm}\label{thm:extalg-fam}
 Suppose that $\extalg{A}$ is an extension algebra in context
@@
 
+\begin{cor}
+Let $\extalg{A}$ be an extension algebra in context $\Gamma$. Then
+$\extalg{F_{F_A}}$ and $\jcomp{}{\cftctxext[\extalg{A}]}{\extalg{F_A}}$ are
+judgmentally equal extension algebras in context $\ctxext{{\Gamma}{\cftalgc{\cftalg{A}}}}{\cftalgf{\cftalg{A}}}$.
+\end{cor}
+
 \begin{defn}
@@
 for which the diagrams
-\begin{equation}\label{eq:extalg-hom1}
+\begin{equation}\label{eq:exthom1}
 \begin{tikzcd}
@@
 and
-\begin{equation}\label{eq:extalg-hom2}
+\begin{equation}\label{eq:exthom2}
 \begin{tikzcd}[column sep=huge]
@@
 \begin{rmk}
-To see that the upper morphism in the diagram of \autoref{eq:extalg-hom2} has
-indeed the indicated codomain provided that the diagram of \autoref{eq:extalg-hom1}
+To see that the upper morphism in the diagram of \autoref{eq:exthom2} has
+indeed the indicated codomain provided that the diagram of \autoref{eq:exthom1}
 commutes judgmentally, note that we have the judgmental equalities
@@
   \jcomp{}{\jcomp{}{\cftctxext[\cftalg{A}]}{\cfthomc{\cfthom{f}}}}{\cftalgf{\cftalg{B}}}
-  \tag{by \autoref{eq:extalg-hom1}}
+  \tag{by \autoref{eq:exthom1}}
   \\
@@
 \begin{proof}
-We have to verify that the diagrams in \autoref{eq:extalg-hom1,eq:extalg-hom2}
-commute for $\cfthom{F_f}$. Unfolding the ingredients of \autoref{eq:extalg-hom1}
-for $\cfthom{F_f}$ gives us quite directly \autoref{eq:extalg-hom2}. The
-diagram in \autoref{eq:extalg-hom2} for $\cfthom{F_f}$ is judgmentally equal
-to the pullback of everything in the diagram in \autoref{eq:extalg-hom2} for
+We have to verify that the diagrams in \autoref{eq:exthom1,eq:exthom2}
+commute for $\cfthom{F_f}$. Unfolding the ingredients of \autoref{eq:exthom1}
+for $\cfthom{F_f}$ gives us quite directly \autoref{eq:exthom2}. The
+diagram in \autoref{eq:exthom2} for $\cfthom{F_f}$ is judgmentally equal
+to the pullback of everything in the diagram in \autoref{eq:exthom2} for
 $\cfthom{f}$ by $\epsilon_0$, and therefore it commutes too.
 \end{proof}
+
+\begin{defn}
+Let $\exthom{f}$ and $\exthom{g}$ be extension homomorphisms from
+$\extalg{A}$ to $\extalg{B}$ and from $\extalg{B}$ to $\extalg{C}$, respectively.
+We define the composition
+\begin{equation*}
+\exthomcomp{\exthom{f}}{\exthom{g}}
+  \defeq
+  \unfold{\exthomcomp{\exthom{f}}{\exthom{g}}}.
+\end{equation*}
+In other words, extension homomorphisms are composed by taking the outer rectangle
+in the diagram
+\begin{equation*}
+\begin{tikzcd}
+\extalgf{\extalg{A}} 
+  \ar[fib]{d}
+  \ar{r}{\exthomf{\exthom{f}}}
+& \extalgf{\extalg{B}} 
+  \ar[fib]{d}
+  \ar{r}{\exthomf{\exthom{g}}}
+& \extalgf{\extalg{C}}
+  \ar[fib]{d}
+  \\
+\extalgc{\extalg{A}}
+  \ar{r}{\exthomc{\exthom{f}}}
+& \extalgc{\extalg{B}}
+  \ar{r}{\exthomc{\exthom{g}}}
+& \extalgc{\extalg{C}}
+\end{tikzcd}
+\end{equation*}
+\end{defn}
+
+\begin{rmk}
+It follows from \autoref{lem:jcomp-jcomp,lem:jfcomp-jfcomp} that composition
+of extension homomorphisms is associative.
+\end{rmk}
+
+\begin{thm}
+Let $\exthom{f}$ and $\exthom{g}$ be extension homomorphisms from
+$\extalg{A}$ to $\extalg{B}$ and from $\extalg{B}$ to $\extalg{C}$, respectively.
+Then $\exthomcomp{\exthom{f}}{\exthom{g}}$ is an extension homomorphism from
+$\extalg{A}$ to $\extalg{C}$. 
+\end{thm}
+
+\begin{proof}
+Both judgmental equalities are applications of the interchange law for composition,
+\autoref{lem:composition-interchange}. 
+\end{proof}
+
+The following theorem explains how context extension can be seen as an extension
+homomorphism. Note that in combination with \autoref{thm:extalg-fam}, this also
+explains how family extension can be seen as an extension homomorphism.
+
+\begin{thm}
+Let $\extalg{A}$ be an extension algebra in context $\Gamma$. Then
+\begin{align*}
+& ( \ctxext{\extalgc{\extalg{A}}}{\extalgf{\extalg{A}}},
+    \jcomp{}{\cftctxext}{\extalgf{\extalg{A}}},
+    \jvcomp{}{\ctxwk{\extalgf{\extalg{A}}}{\idtm{\extalgc{\extalg{A}}}}}{\cftfamext},
+    \jcomp{}{\cftctxext}{\cftfamext}
+    )
+\intertext{and}
+& ( \ctxext{\extalgc{\extalg{A}}}{\extalgf{\extalg{A}}},
+    \jcomp{}{\cftctxext}{\extalgf{\extalg{A}}},
+    \jvcomp{}{\cftctxext}{\idtm{\jcomp{}{\cftctxext}{\extalgf{\extalg{A}}}}},
+    \jcomp{}{\cftctxext}{\cftfamext}
+    )
+\end{align*}
+are extension algebras in context $\Gamma$ and
+\begin{equation*}
+\boldsymbol{\cftctxext}\defeq ( \cftctxext,
+  \idtm{\jcomp{}{\cftctxext}{\extalgf{\extalg{A}}}}
+  )
+\end{equation*}
+is an extension homomorphism from both of them to $\extalg{A}$. 
+\end{thm}
 
@@
   \tag{by \autoref{comp-sw-f}}
-\end{align*}
+  \\
+& \jdeq
+  \ctxwk
+    { \subst{\cftempc{\cftalg{A}}}{\cftalgf{\cftalg{A}}}
+      }
+    { A
+      }
+  \tag{by \autoref{cancellation-ws-f}}
+  \\
+& \jdeq
+  \ctxwk{A}{A}
+  \tag{by \autoref{empalg-eq2}}
+\end{align*}
+Thus, we require that
+\begin{equation}
+\jtermeq{{\Gamma}{\cftalgc{\cftalg{A}}}}{\ctxwk{\cftalgc{\cftalg{A}}}{\cftalgc{\cftalg{A}}}}{\subst{\cftempc{\cftalg{A}}}{\cftctxext[\cftalg{A}]}}{\idtm{\cftalgc{\cftalg{A}}}}
+  \label{extempalg-eq1}
+\end{equation}
 \end{enumerate}

### 2014-06-04 article-models-tt.tex
reorganized first section

@@
 \label{extension}
-The operation of extension has its place right at the core of the theory of
-contexts, families and terms since it enables one to consider families over
-families and higher level families alike.
-
-We will need the extension operation at the level of contexts and families 
-over them and at the level of families and families over them. 
+The operations of extension enable one to consider families over
+families and higher level families alike. We will need both a context extension
+operation and a family extension operation.
+
+The inference rules introducing context and family extension are as follows:
 \begin{align}
@@
 Note the special case where we take $\Gamma$ to be the empty context.
+
+\paragraph{Associativity of extension}
+\label{comp-ee}
+The inference rules asserting that extension is compatible with itself assert
+that contexts are unstructured lists of type declarations. This rule is
+unavoidable if we want that for a family $A$ in context $\Gamma$, a family over
+$A$ is the same thing as a family over $\ctxext{\Gamma}{A}$. 
+\begin{align}
+%& \inference
+%  { \jfam{\Gamma}{A}
+%    \jfam{{\Gamma}{A}}{P}
+%    }
+%  { \jctxeq{\ctxext{{\Gamma}{A}}{P}}{\ctxext{\Gamma}{{A}{P}}}
+%    }
+%  \\
+& \inference
+  { \jfam{{\Gamma}{A}}{P}
+    \jfam{{{\Gamma}{A}}{P}}{Q}
+    }
+  { \jfameq{\Gamma}{\ctxext{A}{{P}{Q}}}{\ctxext{{A}{P}}{Q}}
+    }
+  \label{comp-ee-c}
+\end{align}
+Note that we may take the point of view that extension distributes over
+extensions. If we want to extend an extended family $\ctxext{P}{Q}$ over
+$\ctxext{\Gamma}{A}$ by the family $A$, the result is the extension of (1)
+the extension of $P$ by $A$ and (2) the extension of $Q$ by $A$. The latter is
+simply $Q$ since extension acts trivially at the level of families (and terms).
 
@@
 
-Note that we can also weaken by the empty family over $\Gamma$ itself.
-Weakening by the empty family $\emptyf$ over a context $\Gamma$ leaves families, 
-their terms, families over those families and terms of those unchanged:
-\label{comp-0w}\begin{align}
-& \inference
-  { \jfam{\Gamma}{B}
-    }
-  { \jfameq{\Gamma}{\ctxwk{\emptyf}{B}}{B}
-    }
-  \label{comp-0w-c}
-  \\
-& \inference
-  { \jfam{{\Gamma}{B}}{Q}
-    }
-  { \jfameq{{\Gamma}{B}}{\ctxwk{\emptyf}{Q}}{Q}
-    }
-  \label{comp-0w-f}
-  \\
-& \inference
-  { \jterm{{\Gamma}{B}}{Q}{g}
-    }
-  { \jtermeq{{\Gamma}{B}}{Q}{\ctxwk{\emptyf}{g}}{g}
-    }
-  \label{comp-0w-t}
+\paragraph{Weakenings of extensions}
+\label{comp-we}
+The following rules assert that when an extended family is weakened, the
+weakening distributes over the extension factors.
+\begin{align}\label{comp-we-f}
+& \inference
+  { \jfam{\Gamma}{A}
+    \jfam{{{\Gamma}{B}}{Q}}{R}
+    }
+  { \jfameq
+      {\ctxext{{\Gamma}{A}}{\ctxwk{A}{B}}}
+      {\ctxwk{A}{\ctxext{Q}{R}}}
+      {\ctxext{\ctxwk{A}{Q}}{\ctxwk{A}{R}}}
+    }
+\end{align}
+When thinking of terms of $\ctxwk{A}{B}$ as morphisms of families from $A$ to
+$B$, this looks already like form of type theoretic choice. It is weaker in that
+it is not stated with function types, yet it is stronger in that it states a
+judgmental equality between two families.
+
+There is also a version of this property where an extended term is weakened.
+This variant is stated and proved in \autoref{comp-we-t}.
+
+\paragraph{Weakening of weakenings}
+When a weakened family is weakened even further, the outer weakening distributes
+over the `weakening factors'.
+\label{comp-ww}
+\begin{align}
+& \inference
+  { \jfam{\Gamma}{A}
+    \jfam{\Gamma}{B}
+    \jfam{{\Gamma}{C}}{R}
+    }
+  { \jfameq
+      {{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{{A}{B}}{{A}{C}}}}
+      {\ctxwk{A}{{B}{R}}}
+      {\ctxwk{{A}{B}}{{A}{R}}}
+    }
+  \label{comp-ww-f}\\
+& \inference
+  { \jfam{\Gamma}{A}
+    \jfam{\Gamma}{B}
+    \jterm{{\Gamma}{C}}{R}{t}
+    }
+  { \jtermeq
+      {{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{{A}{B}}{{A}{C}}}}
+      {\ctxwk{{A}{B}}{{A}{R}}}
+      {\ctxwk{A}{{B}{t}}}
+      {\ctxwk{{A}{B}}{{A}{t}}}
+    }
+  \label{comp-ww-t}
 \end{align}
@@
 
-\label{comp-0s}
-The following rules assert that substituting by the term $\jterm{\Gamma}{\emptyf}{\emptytm}$
-leaves everything unchanged.
-\begin{align}
-& \inference
-  { \jfam{\Gamma}{A}
-    }
-  { \jfameq{\Gamma}{\subst{\emptytm}{A}}{A}
-    }
-  \label{comp-0s-c}
-  \\
-%& \inference
-%  { \jterm{\Gamma}{A}{x}
-%    }
-%  { \jtermeq{\Gamma}{A}{\subst{\emptytm}{x}}{x}
-%    }
-%  \\
-& \inference
-  { \jfam{{\Gamma}{A}}{P}
-    }
-  { \jfameq{{\Gamma}{A}}{\subst{\emptytm}{P}}{P}
-    }
-  \label{comp-0s-f}
-  \\
-& \inference
-  { \jterm{{\Gamma}{A}}{P}{f}
-    }
-  { \jtermeq{{\Gamma}{A}}{P}{\subst{\emptytm}{f}}{f}
-    }.
-  \label{comp-0s-t}
-\end{align}
-
 In \autoref{morphisms} we will use a combination of weakening and substitution
@@
 meaningfully state the definition of composition.
+
+\paragraph{Fibers of extensions}
+\label{comp-se}
+The following inference rule asserts that if we take the fiber of an extended
+family at a term $x$ of $A$ in context $\Gamma$, the substitution by $x$
+distributes over the factors of the extension.
+\begin{align}\label{comp-se-f}
+& \inference
+  { \jterm{\Gamma}{A}{x}
+    \jfam{{{{\Gamma}{A}}{P}}{Q}}{R}
+    }
+  { \jfameq
+      {{\Gamma}{\subst{x}{P}}}
+      {\subst{x}{\ctxext{Q}{R}}}
+      {\ctxext{\subst{x}{Q}}{\subst{x}{R}}}
+    }
+\end{align}
+There is also a version of this statement in which extended terms are considered.
+This variant is stated and proved in \autoref{comp-se-t}.
+
+\paragraph{Fibers of weakenings}\label{comp-sw}
+Substitutions distribute over weakenings:
+\begin{align}
+& \inference
+  { \jterm{\Gamma}{A}{x}
+    \jfam{{\Gamma}{A}}{P}
+    \jfam{{{\Gamma}{A}}{Q}}{R}
+    }
+  { \jfameq
+      {{{\Gamma}{\subst{x}{P}}}{\subst{x}{\ctxwk{P}{Q}}}}
+      {\subst{x}{\ctxwk{P}{R}}}
+      {\ctxwk{\subst{x}{P}}{\subst{x}{R}}}
+    }
+  \label{comp-sw-f}\\
+& \inference
+  { \jterm{\Gamma}{A}{x}
+    \jfam{{\Gamma}{A}}{P}
+    \jterm{{{\Gamma}{A}}{Q}}{R}{h}
+    }
+  { \jtermeq
+      {\ctxext{{\Gamma}{\subst{x}{P}}}{\subst{x}{\ctxwk{P}{Q}}}}
+      {\subst{x}{\ctxwk{P}{R}}}
+      {\subst{x}{\ctxwk{P}{h}}}
+      {\ctxwk{\subst{x}{P}}{\subst{x}{h}}}
+    }
+  \label{comp-sw-t}
+\end{align}
+
+\paragraph{Fibers of fibers}
+\label{comp-ss}
+
+We require that substitution is compatible with itself, which is roughly the
+assertion that substitution is associative. However, we cannot just state that
+$\subst{x}{{f}{g}}\jdeq\subst{{x}{f}}{g}$ since the expression $\subst{{x}{f}}{g}$
+is not well-formed. The term $\subst{x}{f}$ can be substituted in (terms of) families over
+$\subst{x}{P}$; the term $\subst{x}{g}$ is such. Therefore, associativity of
+substitution takes the form $\subst{x}{{f}{g}}\jdeq\subst{{x}{f}}{{x}{g}}$.
+Note that the term $\subst{{x}{f}}{{x}{g}}$ may be written down more conveniently
+as $\subst{x,\subst{x}{f}}{g}$, although we will not do that. In fact, after
+we have introduced extension on terms in \autoref{extension-on-terms}, we will 
+see that $\subst{{x}{f}}{{x}{g}}\jdeq\subst{\tmext{x}{\subst{x}{f}}}{g}$.
+
+In the following inference rules we assume
+$\jterm{\Gamma}{A}{x}$ and $\jterm{{\Gamma}{A}}{P}{f}$.
+\begin{align}
+& \inference
+  { \jfam{{{{\Gamma}{A}}{P}}{Q}}{R}
+    }
+  { \jfameq
+      {{\Gamma}{\subst{x}{{f}{Q}}}}
+      {\subst{x}{{f}{R}}}
+      {\subst{{x}{f}}{{x}{R}}}
+    }
+  \label{comp-ss-f}\\
+& \inference
+  { \jterm{{{{\Gamma}{A}}{P}}{Q}}{R}{h}
+    }
+  { \jtermeq
+      {{\Gamma}{\subst{x}{{f}{Q}}}}
+      {\subst{x}{{f}{R}}}
+      {\subst{x}{{f}{h}}}
+      {\subst{{x}{f}}{{x}{h}}}
+    }
+  \label{comp-ss-t}
+\end{align}
+
+\paragraph{Weakenings of fibers}
+\label{comp-ws}
+We already have rules for the compatibility of substitution with weakening, but
+we still need the rules the other way around, asserting that there is a 
+judgmental equality $\ctxwk{A}{\subst{y}{Q}}\jdeq\subst{\ctxwk{A}{y}}{\ctxwk{A}{Q}}$
+together with all its variants.
+
+In the following inference rules we assume that $\jfam{\Gamma}{A}$ and that
+$\jterm{\Gamma}{B}{y}$.
+\begin{align}
+& \inference
+  { \jfam{{{\Gamma}{B}}{Q}}{R}
+    }
+  { \jfameq
+      {{{\Gamma}{A}}{\ctxwk{A}{\subst{y}{Q}}}}
+      {\ctxwk{A}{\subst{y}{R}}}
+      {\subst{\ctxwk{A}{y}}{\ctxwk{A}{R}}}
+    }
+  \label{comp-ws-f}\\
+& \inference
+  { \jterm{{{\Gamma}{B}}{Q}}{R}{h}
+    }
+  { \jtermeq
+      {{{\Gamma}{A}}{\ctxwk{A}{\subst{y}{Q}}}}
+      {\ctxwk{A}{\subst{y}{R}}}
+      {\ctxwk{A}{\subst{y}{h}}}
+      {\subst{\ctxwk{A}{y}}{\ctxwk{A}{h}}}
+    }
+  \label{comp-ws-t}
+\end{align}
 
@@
 
-\subsection{The judgmental equalities of the theory of contexts, families and
+\paragraph{The identity term of an extended family}
+The fact that identity terms are compatible with extensions is derived in
+\autoref{comp-ie}.
+
+\paragraph{The identity term of a weakened family}
+\label{comp-wi}
+The identity term of a weakened family is the weakened identity term:
+\begin{equation}
+\inference
+  { \jfam{\Gamma}{A}
+    \jfam{\Gamma}{B}
+    }
+  { \unfoldall{\jhomeq
+      {{\Gamma}{A}}
+      {\ctxwk{A}{B}}
+      {\ctxwk{A}{B}}
+      {\ctxwk{A}{\idtm{B}}}
+      {\idtm{\ctxwk{A}{B}}}
+    }}
+  \label{comp-wi-t}
+\end{equation}
+
+\paragraph{The identity term of a substituted family}
+\label{comp-si}
+The identity term of a substituted family is the substitution of the identity term
+\begin{equation}
+\inference
+  { \jterm{\Gamma}{A}{x}
+    \jfam{{\Gamma}{A}}{P}
+    }
+  { \unfoldall{\jhomeq
+      {\Gamma}
+      {\subst{x}{P}}
+      {\subst{x}{P}}
+      {\subst{x}{\idtm{P}}}
+      {\idtm{\subst{x}{P}}}
+    }}
+  \label{comp-si-t}
+\end{equation}
+
+\subsection{Further judgmental equalities of the theory of contexts, families and
 terms}
+We have already seen that empty families, extension, weakening, substitution
+and the identity terms are compatible with each other. There are, however, a
+few more judgmental equalities explaining what happens when
+\begin{enumerate}
+\item something is weakend by the empty family. This is used to consider a term
+      $x$ of $A$ in context $\Gamma$ as a morphism from $\emptyf$ to $A$ in
+      context $\Gamma$, see \autoref{morphisms};
+\item something is substituted by the term of the empty family;
+\item something is weakened by an extended family. This is Currying for weakening;
+\item something is first weakened by a family $A$ in context $\Gamma$ and then
+      substituted by a term of $A$. This is used to show that composition is
+      compatible with itself (i.e. that composition is associative);
+\item the identity terms are joining the party. Identity terms are empty contexts
+      in certain models, so they should behave accordingly.
+\end{enumerate}
+In other words, the rules seem to not fit in the scheme that all operations are
+compatible with each other, but in fact each of these rules plays a role in
+completing the scheme (work out further).
+
+\begin{comment}
 In this section we extend the judgmental equality relation in such a way that
@@
 reference to the lemma in which this variant is stated and proved.
-
-\subsubsection{Extensions of extensions are extensions by extensions}
-\label{comp-ee}
-The inference rules asserting that extension is compatible with itself assert
-that contexts are unstructured lists of type declarations. This rule is
-unavoidable if we want that for a family $A$ in context $\Gamma$, a family over
-$A$ is the same thing as a family over $\ctxext{\Gamma}{A}$. 
-\begin{align}
+\end{comment}
+
+\subsubsection{Weakening by the empty family}
+Note that we can also weaken by the empty family over $\Gamma$.
+Weakening by the empty family $\emptyf$ over a context $\Gamma$ leaves families, 
+their terms, families over those families and terms of those unchanged:
+\label{comp-0w}
+\begin{align}
+& \inference
+  { \jfam{\Gamma}{B}
+    }
+  { \jfameq{\Gamma}{\ctxwk{\emptyf}{B}}{B}
+    }
+  \label{comp-0w-c}
+  \\
+& \inference
+  { \jfam{{\Gamma}{B}}{Q}
+    }
+  { \jfameq{{\Gamma}{B}}{\ctxwk{\emptyf}{Q}}{Q}
+    }
+  \label{comp-0w-f}
+  \\
+& \inference
+  { \jterm{{\Gamma}{B}}{Q}{g}
+    }
+  { \jtermeq{{\Gamma}{B}}{Q}{\ctxwk{\emptyf}{g}}{g}
+    }
+  \label{comp-0w-t}
+\end{align}
+
+\subsubsection{Substituting by the term of the empty family}
+\label{comp-0s}
+The following rules assert that substituting by the term $\jterm{\Gamma}{\emptyf}{\emptytm}$
+leaves everything unchanged.
+\begin{align}
+& \inference
+  { \jfam{\Gamma}{A}
+    }
+  { \jfameq{\Gamma}{\subst{\emptytm}{A}}{A}
+    }
+  \label{comp-0s-c}
+  \\
 %& \inference
-%  { \jfam{\Gamma}{A}
-%    \jfam{{\Gamma}{A}}{P}
-%    }
-%  { \jctxeq{\ctxext{{\Gamma}{A}}{P}}{\ctxext{\Gamma}{{A}{P}}}
+%  { \jterm{\Gamma}{A}{x}
+%    }
+%  { \jtermeq{\Gamma}{A}{\subst{\emptytm}{x}}{x}
 %    }
@@
   { \jfam{{\Gamma}{A}}{P}
-    \jfam{{{\Gamma}{A}}{P}}{Q}
-    }
-  { \jfameq{\Gamma}{\ctxext{A}{{P}{Q}}}{\ctxext{{A}{P}}{Q}}
-    }
-  \label{comp-ee-c}
-\end{align}
-Note that we may take the point of view that extension distributes over
-extensions. If we want to extend an extended family $\ctxext{P}{Q}$ over
-$\ctxext{\Gamma}{A}$ by the family $A$, the result is the extension of (1)
-the extension of $P$ by $A$ and (2) the extension of $Q$ by $A$. The latter is
-simply $Q$ since extension acts trivially at the level of families (and terms).
-
-\subsubsection{Weakenings of extensions}
-\label{comp-we}
-The following rules assert that when an extended family is weakened, the
-weakening distributes over the extension factors.
-\begin{align}\label{comp-we-f}
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{{{\Gamma}{B}}{Q}}{R}
-    }
-  { \jfameq
-      {\ctxext{{\Gamma}{A}}{\ctxwk{A}{B}}}
-      {\ctxwk{A}{\ctxext{Q}{R}}}
-      {\ctxext{\ctxwk{A}{Q}}{\ctxwk{A}{R}}}
-    }
-\end{align}
-When thinking of terms of $\ctxwk{A}{B}$ as morphisms of families from $A$ to
-$B$, this looks already like form of type theoretic choice. It is weaker in that
-it is not stated with function types, yet it is stronger in that it states a
-judgmental equality between two families.
-
-There is also a version of this property where an extended term is weakened.
-This variant is stated and proved in \autoref{comp-we-t}.
-
-\subsubsection{Fibers of extensions}
-\label{comp-se}
-The following inference rule asserts that if we take the fiber of an extended
-family at a term $x$ of $A$ in context $\Gamma$, the substitution by $x$
-distributes over the factors of the extension.
-\begin{align}\label{comp-se-f}
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jfam{{{{\Gamma}{A}}{P}}{Q}}{R}
-    }
-  { \jfameq
-      {{\Gamma}{\subst{x}{P}}}
-      {\subst{x}{\ctxext{Q}{R}}}
-      {\ctxext{\subst{x}{Q}}{\subst{x}{R}}}
-    }
-\end{align}
-There is also a version of this statement in which extended terms are considered.
-This variant is stated and proved in \autoref{comp-se-t}.
-
-\subsubsection{Weakening of weakenings}
-When a weakened family is weakened even further, the outer weakening distributes
-over the `weakening factors'.
-\label{comp-ww}
-\begin{align}
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{\Gamma}{B}
-    \jfam{{\Gamma}{C}}{R}
-    }
-  { \jfameq
-      {{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{{A}{B}}{{A}{C}}}}
-      {\ctxwk{A}{{B}{R}}}
-      {\ctxwk{{A}{B}}{{A}{R}}}
-    }
-  \label{comp-ww-f}\\
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{\Gamma}{B}
-    \jterm{{\Gamma}{C}}{R}{t}
-    }
-  { \jtermeq
-      {{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{{A}{B}}{{A}{C}}}}
-      {\ctxwk{{A}{B}}{{A}{R}}}
-      {\ctxwk{A}{{B}{t}}}
-      {\ctxwk{{A}{B}}{{A}{t}}}
-    }
-  \label{comp-ww-t}
-\end{align}
-
-\subsubsection{Fibers of weakenings}\label{comp-sw}
-Substitutions distribute over weakenings:
-\begin{align}
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jfam{{\Gamma}{A}}{P}
-    \jfam{{{\Gamma}{A}}{Q}}{R}
-    }
-  { \jfameq
-      {{{\Gamma}{\subst{x}{P}}}{\subst{x}{\ctxwk{P}{Q}}}}
-      {\subst{x}{\ctxwk{P}{R}}}
-      {\ctxwk{\subst{x}{P}}{\subst{x}{R}}}
-    }
-  \label{comp-sw-f}\\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jfam{{\Gamma}{A}}{P}
-    \jterm{{{\Gamma}{A}}{Q}}{R}{h}
-    }
-  { \jtermeq
-      {\ctxext{{\Gamma}{\subst{x}{P}}}{\subst{x}{\ctxwk{P}{Q}}}}
-      {\subst{x}{\ctxwk{P}{R}}}
-      {\subst{x}{\ctxwk{P}{h}}}
-      {\ctxwk{\subst{x}{P}}{\subst{x}{h}}}
-    }
-  \label{comp-sw-t}
-\end{align}
-
-\subsubsection{Fibers of fibers}
-\label{comp-ss}
-
-We require that substitution is compatible with itself, which is roughly the
-assertion that substitution is associative. However, we cannot just state that
-$\subst{x}{{f}{g}}\jdeq\subst{{x}{f}}{g}$ since the expression $\subst{{x}{f}}{g}$
-is not well-formed. The term $\subst{x}{f}$ can be substituted in (terms of) families over
-$\subst{x}{P}$; the term $\subst{x}{g}$ is such. Therefore, associativity of
-substitution takes the form $\subst{x}{{f}{g}}\jdeq\subst{{x}{f}}{{x}{g}}$.
-Note that the term $\subst{{x}{f}}{{x}{g}}$ may be written down more conveniently
-as $\subst{x,\subst{x}{f}}{g}$, although we will not do that. In fact, after
-we have introduced extension on terms in \autoref{extension-on-terms}, we will 
-see that $\subst{{x}{f}}{{x}{g}}\jdeq\subst{\tmext{x}{\subst{x}{f}}}{g}$.
-
-In the following inference rules we assume
-$\jterm{\Gamma}{A}{x}$ and $\jterm{{\Gamma}{A}}{P}{f}$.
-\begin{align}
-& \inference
-  { \jfam{{{{\Gamma}{A}}{P}}{Q}}{R}
-    }
-  { \jfameq
-      {{\Gamma}{\subst{x}{{f}{Q}}}}
-      {\subst{x}{{f}{R}}}
-      {\subst{{x}{f}}{{x}{R}}}
-    }
-  \label{comp-ss-f}\\
-& \inference
-  { \jterm{{{{\Gamma}{A}}{P}}{Q}}{R}{h}
-    }
-  { \jtermeq
-      {{\Gamma}{\subst{x}{{f}{Q}}}}
-      {\subst{x}{{f}{R}}}
-      {\subst{x}{{f}{h}}}
-      {\subst{{x}{f}}{{x}{h}}}
-    }
-  \label{comp-ss-t}
-\end{align}
-
-\subsubsection{Weakenings of fibers}
-\label{comp-ws}
-We already have rules for the compatibility of substitution with weakening, but
-we still need the rules the other way around, asserting that there is a 
-judgmental equality $\ctxwk{A}{\subst{y}{Q}}\jdeq\subst{\ctxwk{A}{y}}{\ctxwk{A}{Q}}$
-together with all its variants.
-
-In the following inference rules we assume that $\jfam{\Gamma}{A}$ and that
-$\jterm{\Gamma}{B}{y}$.
-\begin{align}
-& \inference
-  { \jfam{{{\Gamma}{B}}{Q}}{R}
-    }
-  { \jfameq
-      {{{\Gamma}{A}}{\ctxwk{A}{\subst{y}{Q}}}}
-      {\ctxwk{A}{\subst{y}{R}}}
-      {\subst{\ctxwk{A}{y}}{\ctxwk{A}{R}}}
-    }
-  \label{comp-ws-f}\\
-& \inference
-  { \jterm{{{\Gamma}{B}}{Q}}{R}{h}
-    }
-  { \jtermeq
-      {{{\Gamma}{A}}{\ctxwk{A}{\subst{y}{Q}}}}
-      {\ctxwk{A}{\subst{y}{R}}}
-      {\ctxwk{A}{\subst{y}{h}}}
-      {\subst{\ctxwk{A}{y}}{\ctxwk{A}{h}}}
-    }
-  \label{comp-ws-t}
-\end{align}
-
-\subsubsection{The identity term of a weakened family}
-\label{comp-wi}
-The identity term of a weakened family is the weakened identity term:
-\begin{equation}
-\inference
-  { \jfam{\Gamma}{A}
-    \jfam{\Gamma}{B}
-    }
-  { \unfoldall{\jhomeq
-      {{\Gamma}{A}}
-      {\ctxwk{A}{B}}
-      {\ctxwk{A}{B}}
-      {\ctxwk{A}{\idtm{B}}}
-      {\idtm{\ctxwk{A}{B}}}
-    }}
-  \label{comp-wi-t}
-\end{equation}
-
-\subsubsection{The identity term of a substituted family}
-\label{comp-si}
-The identity term of a substituted family is the substitution of the identity term
-\begin{equation}
-\inference
-  { \jterm{\Gamma}{A}{x}
-    \jfam{{\Gamma}{A}}{P}
-    }
-  { \unfoldall{\jhomeq
-      {\Gamma}
-      {\subst{x}{P}}
-      {\subst{x}{P}}
-      {\subst{x}{\idtm{P}}}
-      {\idtm{\subst{x}{P}}}
-    }}
-  \label{comp-si-t}
-\end{equation}
+    }
+  { \jfameq{{\Gamma}{A}}{\subst{\emptytm}{P}}{P}
+    }
+  \label{comp-0s-f}
+  \\
+& \inference
+  { \jterm{{\Gamma}{A}}{P}{f}
+    }
+  { \jtermeq{{\Gamma}{A}}{P}{\subst{\emptytm}{f}}{f}
+    }.
+  \label{comp-0s-t}
+\end{align}
 

### 2014-06-05 article-models-ttderived.tex
terms as sections

@@
 
+Note that we can directly view any term $x$ of $A$ in context $\Gamma$ as a
+morphism from the empty family $\emptyf$ to $A$ in context $\Gamma$.
+
+Now let $\jterm{{\Gamma}{A}}{P}{f}$ be a term of $P$. Then we can view $f$ as a
+morphism from $A$ to $\ctxext{A}{P}$ in the following way:
+
+\begin{defn}
+Let $\jterm{{\Gamma}{A}}{P}{f}$ be a term. Then we define
+\begin{equation*}
+\jhomdefn{\Gamma}{A}{{A}{P}}{\jtermashom{f}}{\unfold{\jtermashom{f}}}
+\end{equation*}
+\end{defn}
+
+In the following lemma we will see that $\jtermashom{f}$ is a section for
+$\cprojfstf{A}{P}$.
+
+\begin{lem}
+Consider a family $\jfam{{\Gamma}{A}}{P}$ and a term $\jterm{\Gamma}{A}{x}$.
+Then we have the judgmental equality
+\begin{equation*}
+\jhomeq{\Gamma}{A}{A}{\jcomp{A}{\jtermashom{f}}{\cprojfstf{A}{P}}}{\idtm{A}}.
+\end{equation*}
+\end{lem}
+
+\begin{proof}
+We have the judgmental equalities
+\begin{align*}
+\jtermashom{f}
+& \jdeq
+  \unfold{\jtermashom{f}}
+  \tag{by definition}
+  \\
+& \jdeq
+  \jcomp{}{f}{\jcomp{}{\idtm{A}}{\cprojfstf{A}{P}}}
+  \tag{by \autoref{lem:composition-threesome}}
+  \\
+& \jdeq
+  \jcomp{}{f}{\cprojfstf{A}{P}}
+  \tag{by \autoref{lem:jcomp-idtm}}
+  \\
+& \jdeq
+  \jcomp{}{f}{\ctxwk{P}{\idtm{A}}}
+\end{align*}
+\end{proof}
+
 In \autoref{lem:tmext-emp,lem:tmext-ext,lem:tmext-wk,lem:tmext-subst,lem:tmext-id}

### 2014-06-06 article-models-ttderived.tex
more on terms as sections
terms as sections

@@
 
-Note that we can directly view any term $x$ of $A$ in context $\Gamma$ as a
-morphism from the empty family $\emptyf$ to $A$ in context $\Gamma$.
-
-Now let $\jterm{{\Gamma}{A}}{P}{f}$ be a term of $P$. Then we can view $f$ as a
-morphism from $A$ to $\ctxext{A}{P}$ in the following way:
-
-\begin{defn}
-Let $\jterm{{\Gamma}{A}}{P}{f}$ be a term. Then we define
-\begin{equation*}
-\jhomdefn{\Gamma}{A}{{A}{P}}{\jtermashom{f}}{\unfold{\jtermashom{f}}}
-\end{equation*}
-\end{defn}
-
-In the following lemma we will see that $\jtermashom{f}$ is a section for
-$\cprojfstf{A}{P}$.
-
-\begin{lem}
-Consider a family $\jfam{{\Gamma}{A}}{P}$ and a term $\jterm{\Gamma}{A}{x}$.
-Then we have the judgmental equality
-\begin{equation*}
-\jhomeq{\Gamma}{A}{A}{\jcomp{A}{\jtermashom{f}}{\cprojfstf{A}{P}}}{\idtm{A}}.
-\end{equation*}
-\end{lem}
-
-\begin{proof}
-We have the judgmental equalities
-\begin{align*}
-\jtermashom{f}
-& \jdeq
-  \unfold{\jtermashom{f}}
-  \tag{by definition}
-  \\
-& \jdeq
-  \jcomp{}{f}{\jcomp{}{\idtm{A}}{\cprojfstf{A}{P}}}
-  \tag{by \autoref{lem:composition-threesome}}
-  \\
-& \jdeq
-  \jcomp{}{f}{\cprojfstf{A}{P}}
-  \tag{by \autoref{lem:jcomp-idtm}}
-  \\
-& \jdeq
-  \jcomp{}{f}{\ctxwk{P}{\idtm{A}}}
-\end{align*}
-\end{proof}
-
 In \autoref{lem:tmext-emp,lem:tmext-ext,lem:tmext-wk,lem:tmext-subst,lem:tmext-id}
@@
 \end{align*}
+\end{proof}
+
+As an application of these lemmas we prove that every term $\jterm{{\Gamma}{A}}{P}{f}$
+induces a section $\jtermashom{f}$ of $\cprojfstf{A}{P}$.
+Note that we can directly view any term $x$ of $A$ in context $\Gamma$ as a
+morphism from the empty family $\emptyf$ to $A$ in context $\Gamma$.
+
+Now let $\jterm{{\Gamma}{A}}{P}{f}$ be a term of $P$. Then we can view $f$ as a
+morphism from $A$ to $\ctxext{A}{P}$ in the following way:
+
+\begin{defn}\label{defn:terms-as-sections}
+Let $\jterm{{\Gamma}{A}}{P}{f}$ be a term. Then we define
+\begin{equation*}
+\jhomdefn{\Gamma}{A}{{A}{P}}{\jtermashom{f}}{\unfold{\jtermashom{f}}}
+\end{equation*}
+\end{defn}
+
+In the following lemma we will see that $\jtermashom{f}$ is a section for
+$\cprojfstf{A}{P}$.
+
+\begin{cor}
+Consider a family $\jfam{{\Gamma}{A}}{P}$ and a term $\jterm{\Gamma}{A}{x}$.
+Then we have the judgmental equality
+\begin{equation*}
+\jhomeq{\Gamma}{A}{A}{\jcomp{A}{\jtermashom{f}}{\cprojfstf{A}{P}}}{\idtm{A}}.
+\end{equation*}
+\end{cor}
+
+\begin{proof}
+The asserted judgmental equality follows immediately from \autoref{lem:mor-tmext-cproj}.
 \end{proof}

### 2014-06-15 article-models-tt.tex
terms as sections
starting to remove the empty context

@@
 
-\subsubsection{The empty context}
+\subsubsection{Extension}
+\label{extension}
+The operations of extension enable one to consider families over
+families and higher level families alike. We will need both a context extension
+operation and a family extension operation.
+
+The inference rules introducing context and family extension are as follows:
+\begin{align}
+& \inference
+  { \jfam{\Gamma}{A}
+    }
+  { \jctx{\ctxext{\Gamma}{A}}
+    }
+& & \inference
+    { \jctxeq{\Gamma}{\Delta}
+      \jfameq{\Gamma}{A}{B}
+      }
+    { \jctxeq{\ctxext{\Gamma}{A}}{\ctxext{\Delta}{B}}
+      }
+    \\
+& \inference
+  { \jfam{{\Gamma}{A}}{P}
+    }
+  { \jfam{\Gamma}{\ctxext{A}{P}}
+    }
+& & \inference
+    { \jfameq{\Gamma}{A}{B} 
+      \jfameq{{\Gamma}{A}}{P}{Q}
+      }
+    { \jfameq{\Gamma}{\ctxext{A}{P}}{\ctxext{B}{Q}}
+      }
+\end{align}
+The extension operation may also be defined to act on families over families
+and terms thereof. For instance, When $\jfam{{\Gamma}{A}}{P}$ is a family,
+we would get a family $\jfam{{\Gamma}{A}}{{\Gamma}{P}}$ and when
+$\jterm{{\Gamma}{A}}{P}{f}$ is a term, we would get a term
+$\jterm{{\Gamma}{A}}{{\Gamma}{P}}{\ctxext{\Gamma}{f}}$. Then there would be
+axioms stating the judgmental equalities $\ctxext{\Gamma}{P}\jdeq P$ and
+$\ctxext{\Gamma}{f}\jdeq f$. Thus, it is of no actual use to introduce the
+action of context extension on families and terms.
+
+Nevertheless, the fact that extension \emph{does} act on families
+and terms (albeit trivially) reflects in the internal extension term in the type theory of
+models of type theory, which is developed in \autoref{ttmodels}. Also, in the
+categorical semantics, the extension operation would be the concatenation of
+lists of arrows belonging to the class of arrows representing families. Since
+families over families are interpreted by lists of those family arrows extending
+concateneted lists of family arrows, the action of concatenation of such lists
+may be taken to be trivial. The $\Sigma$-functor, which is the version of this
+operation on those lists which have length one, acts trivially on morphisms;
+another manifestation of the same phenomenon.
+
+In \autoref{extension-on-terms} we will also define an extension operation
+on terms, which defines a term $\jterm{\Gamma}{{A}{P}}{\tmext{x}{u}}$ for
+every $\jterm{\Gamma}{A}{x}$ and $\jterm{\Gamma}{\subst{x}{P}}{u}$. This
+operation is defined by using the substitution operation and identity terms.
+
+\paragraph{Associativity of extension}
+\label{comp-ee}
+The inference rules asserting that extension is compatible with itself assert
+that contexts are unstructured lists of type declarations. This rule is
+unavoidable if we want that for a family $A$ in context $\Gamma$, a family over
+$A$ is the same thing as a family over $\ctxext{\Gamma}{A}$. 
+\begin{align}
+%& \inference
+%  { \jfam{\Gamma}{A}
+%    \jfam{{\Gamma}{A}}{P}
+%    }
+%  { \jctxeq{\ctxext{{\Gamma}{A}}{P}}{\ctxext{\Gamma}{{A}{P}}}
+%    }
+%  \\
+& \inference
+  { \jfam{{\Gamma}{A}}{P}
+    \jfam{{{\Gamma}{A}}{P}}{Q}
+    }
+  { \jfameq{\Gamma}{\ctxext{A}{{P}{Q}}}{\ctxext{{A}{P}}{Q}}
+    }
+  \label{comp-ee-c}
+\end{align}
+Note that we may take the point of view that extension distributes over
+extensions. If we want to extend an extended family $\ctxext{P}{Q}$ over
+$\ctxext{\Gamma}{A}$ by the family $A$, the result is the extension of (1)
+the extension of $P$ by $A$ and (2) the extension of $Q$ by $A$. The latter is
+simply $Q$ since extension acts trivially at the level of families (and terms).
+
+\subsubsection{The empty families}
 \label{empty}
-In the theory of contexts, families and terms we introduce an empty context and,
-for every context $\Gamma$, an empty family over $\Gamma$. 
+We introduce an empty family over $\Gamma$ for every context $\Gamma$. It was
+a suggestion of Voevodsky and Garner to only include empty families and not an
+empty context. Having an empty context requires also rules asserting that a
+context is the same thing as a family over the empty context and that operations
+act the same on families over the empty context as on contexts. Not including
+an empty context changes the theory a bit. One example of a difference is that
+if we want weakening by a context we shall have to explicitly state it as it 
+would not follow from weakening by a family. A second example is that without
+an empty context, some rule schemes (like the compatibility of extension with
+the empty families, stated in \autoref{comp-0e} below) become slightly
+assymetrical. Another example is that it would
+not be possible to consider closed types and terms without an empty context.
+Although we have opted to not include an empty context by default, we
+shall give the rules in \autoref{empty-context} so that one can look them up
+if they are desired.
+
+{\color{red}Explain why empty families are needed}
+
+The empty family over a context $\Gamma$ is introduced by the following rule
+inference rule:
+\begin{equation}
+\inference
+  { \jctx{\Gamma}
+    }
+  { \jfam{\Gamma}{\emptyf[\Gamma]}
+    }
+\end{equation}
+Since it is not really necessary to annotate the empty family with its context,
+we shall usually denote $\emptyf[\Gamma]$ by $\emptyf$.
+
+The reader may wonder whether the empty family $\emptyf$ always has a
+term. This shall follow from the rules stating the compatibility of extension
+with the empty families in \autoref{comp-0e} below and from
+identity terms (\autoref{identityterms}).
+\begin{comment}
 When contexts are viewed as statements, the empty context is the statement that
@@
 $\emptytm[\Gamma]$ by $\emptytm$.
-
-\subsubsection{Extension}
-\label{extension}
-The operations of extension enable one to consider families over
-families and higher level families alike. We will need both a context extension
-operation and a family extension operation.
-
-The inference rules introducing context and family extension are as follows:
-\begin{align}
-& \inference
-  { \jfam{\Gamma}{A}
-    }
-  { \jctx{\ctxext{\Gamma}{A}}
-    }
-& & \inference
-    { \jctxeq{\Gamma}{\Delta}
-      \jfameq{\Gamma}{A}{B}
-      }
-    { \jctxeq{\ctxext{\Gamma}{A}}{\ctxext{\Delta}{B}}
-      }
-    \\
-& \inference
-  { \jfam{{\Gamma}{A}}{P}
-    }
-  { \jfam{\Gamma}{\ctxext{A}{P}}
-    }
-& & \inference
-    { \jfameq{\Gamma}{A}{B} 
-      \jfameq{{\Gamma}{A}}{P}{Q}
-      }
-    { \jfameq{\Gamma}{\ctxext{A}{P}}{\ctxext{B}{Q}}
-      }
-\end{align}
-The extension operation may also be defined to act on families over families
-and terms thereof. For instance, When $\jfam{{\Gamma}{A}}{P}$ is a family,
-we would get a family $\jfam{{\Gamma}{A}}{{\Gamma}{P}}$ and when
-$\jterm{{\Gamma}{A}}{P}{f}$ is a term, we would get a term
-$\jterm{{\Gamma}{A}}{{\Gamma}{P}}{\ctxext{\Gamma}{f}}$. Then there would be
-axioms stating the judgmental equalities $\ctxext{\Gamma}{P}\jdeq P$ and
-$\ctxext{\Gamma}{f}\jdeq f$. Thus, it is of no actual use to introduce the
-action of context extension on families and terms.
-
-Nevertheless, the fact that extension \emph{does} act on families
-and terms (albeit trivially) reflects in the internal extension term in the type theory of
-models of type theory, which is developed in \autoref{ttmodels}. Also, in the
-categorical semantics, the extension operation would be the concatenation of
-lists of arrows belonging to the class of arrows representing families. Since
-families over families are interpreted by lists of those family arrows extending
-concateneted lists of family arrows, the action of concatenation of such lists
-may be taken to be trivial. The $\Sigma$-functor, which is the version of this
-operation on those lists which have length one, acts trivially on morphisms;
-another manifestation of the same phenomenon.
-
-In \autoref{extension-on-terms} we will also define an extension operation
-on terms, which defines a term $\jterm{\Gamma}{{A}{P}}{\tmext{x}{u}}$ for
-every $\jterm{\Gamma}{A}{x}$ and $\jterm{\Gamma}{\subst{x}{P}}{u}$. This
-operation is defined by using the substitution operation and identity terms.
+\end{comment}
 
@@
 \begin{align}
+& \inference
+  { \jctx{\Gamma}
+    }
+  { \jctxeq{\ctxext{\Gamma}{\emptyf}}{\Gamma}
+    }
+  \label{comp-e0-c}\\
+& \inference
+  { \jfam{\Gamma}{A}
+    }
+  { \jfameq{\Gamma}{\ctxext{\emptyf}{A}}{A}
+    }
+  \label{comp-0e-f}
+  \\
+& \inference
+  { \jfam{\Gamma}{A}
+    }
+  { \jfameq{\Gamma}{{A}{\emptyf}}{A}
+    }
+  \label{comp-e0-f}
 %& \inference
@@
 %    }
-%  { \jctxeq{\ctxext{\Gamma}{\emptyf}}{\Gamma}
-%    }
-%  \label{comp-e0-c}\\
-& \inference
-  { \jfam{\Gamma}{A}
-    }
-  { \jfameq{\Gamma}{{A}{\emptyf}}{A}
-    }
-  \label{comp-e0-f}
-  \\
-%& \inference
-%  { \jctx{\Gamma}
-%    }
 %  { \jctxeq{\ctxext{\emptyc}{\Gamma}}{\Gamma}
@@
 %  \label{comp-0e-c}\\
-& \inference
-  { \jfam{\Gamma}{A}
-    }
-  { \jfameq{\Gamma}{\ctxext{\emptyf}{A}}{A}
-    }
-  \label{comp-0e-f}
-\end{align}
-Note the special case where we take $\Gamma$ to be the empty context.
-
-\paragraph{Associativity of extension}
-\label{comp-ee}
-The inference rules asserting that extension is compatible with itself assert
-that contexts are unstructured lists of type declarations. This rule is
-unavoidable if we want that for a family $A$ in context $\Gamma$, a family over
-$A$ is the same thing as a family over $\ctxext{\Gamma}{A}$. 
-\begin{align}
-%& \inference
-%  { \jfam{\Gamma}{A}
-%    \jfam{{\Gamma}{A}}{P}
-%    }
-%  { \jctxeq{\ctxext{{\Gamma}{A}}{P}}{\ctxext{\Gamma}{{A}{P}}}
-%    }
-%  \\
-& \inference
-  { \jfam{{\Gamma}{A}}{P}
-    \jfam{{{\Gamma}{A}}{P}}{Q}
-    }
-  { \jfameq{\Gamma}{\ctxext{A}{{P}{Q}}}{\ctxext{{A}{P}}{Q}}
-    }
-  \label{comp-ee-c}
-\end{align}
-Note that we may take the point of view that extension distributes over
-extensions. If we want to extend an extended family $\ctxext{P}{Q}$ over
-$\ctxext{\Gamma}{A}$ by the family $A$, the result is the extension of (1)
-the extension of $P$ by $A$ and (2) the extension of $Q$ by $A$. The latter is
-simply $Q$ since extension acts trivially at the level of families (and terms).
+\end{align}
+%Note the special case where we take $\Gamma$ to be the empty context.
 
@@
       $x$ of $A$ in context $\Gamma$ as a morphism from $\emptyf$ to $A$ in
-      context $\Gamma$, see \autoref{morphisms};
+      context $\Gamma$;
 \item something is substituted by the term of the empty family;

### 2014-06-17 article-models-main.tex
new section for derived notions
strong inductive morphisms
commenting stuff so that the article becomes shorter
split the derived notions section in two
reorganization
version esystems sent to VV and RG

--- 2014-04-01 article-models-main.tex
@@
 
+\begin{comment}
 \begin{abstract}
@@
 \end{abstract}
+\end{comment}
 
@@
 
+\begin{comment}
 \part{Introduction}
@@
 %\input{article-models-eg}
+\end{comment}
 
@@
 been studied very much. Dependent product types or even universes tend to make
-an early appearance in just about any presentation of type theory. While one
-could argue that in type theories like the one in \cite{Dybjer1996} the type
-constructors could be left out and a study of the remnants could \emph{in
-principle} be carried out, this is not the usual practice. In the type theory
-of \cite{TheBook}, which seems to have won the race of introducing universes
-as early as possible hands down, this seems entirely unfeasible.
+an early appearance in just about any presentation of type theory.
+A noteable exception is the theory of categories with families of Dybjer
+in \cite{Dybjer1996}, which has been elaborated on further by Dybjer and
+Clairambault in \cite{DybjerClairambault2011} and in unpublished work by
+Awody \cite{Awodey2013} on natural models of type theory,
+which makes the connection between categories with families and representable
+transformations of presheaves. In the way type theory is presented by the
+Univalent Foundations Project in \cite{TheBook}, which seems to have won the race of introducing universes
+as early as possible hands down, it seems entirely unfeasible to study type
+theory without type constructors. Also, the 
+proof-assistant \Coq\ {\color{red}(and \Agda too?)} has universes and dependent product and pair types 
+built-in, making it impossible to study type theory without type constructors in 
+that environment.
 
 Nevertheless, type theory without type constructors has received some attention
-recently. In \cite{Garner2014}, Garner describes the combinatorial structure
+contributors to the Univalent Foundations Program recently. We name two
+further investigations on this topic, other than the mentioned work by Awodey.
+In \cite{Garner2014}, Garner describes the combinatorial structure
 of the type operations of the weakening, substitution and projection monads
 (their projections are our identity terms) and suggests lots of further research
-that can be done on type theory without constructors. Also, Joyal has been
-lecturing about tribes lately, which are a categorical semantics for type theory
-without type constructors. Although his slides have been going around on
-mailing lists, there is as of yet no easily accessible written accound on tribes.
+that can be done on type theory without constructors. Also, Joyal has presented
+his theory of tribes, which is a categorical explanation of type theory
+without type constructors.
+
+After we have described the E-system, which is the flavor of type theory without
+type constructors presented in \autoref{tt}, we will demonstrate in \autoref{ttderived}
+that the theory gives rise to a rich categorical structure. The notions and
+properties we derive here will be essential for the further work on internal
+models for E-systems, presented in \autoref{part:models}.
 
@@
 
+\input{article-models-ttderived}
+
+%\input{article-models-ttpretty}
+
+\part{Internal Models}\label{part:models}
+
+\input{article-models-ttinternal}
+
+\begin{comment}
+\input{article-models-ttmodels}
+
+\input{article-models-ttuniverses}
+
+\part{Inductive constructions}
+
+\input{article-models-ttind}
+
 \input{article-models-ttconstr}
-
-\input{article-models-ttmodels}
+\end{comment}
 

### 2014-06-17 article-models-tt.tex
version esystems sent to VV and RG

@@
-\section{The algebraic theory of contexts, families and terms}\label{tt}
+\section{The generalized algebraic theory of contexts, families and terms (the E-system)}\label{tt}
 In this section we give a description of dependent type theory before type
@@
 core of the language of dependent logic -- the basic ingredients
-of this theory will be the operations of extension, weakening and substitution
-and the identity terms. An empty context (and empty families) are also included
-in the theory. The resulting theory can be seen as a manifestation of the 
+of this theory will be the operations of extension, empty families, weakening and substitution
+and the identity terms. The resulting theory can be seen as a manifestation of the 
 structure underlying dependent type theory.
@@
 formulate the theory of contexts, families and terms in such a way that contexts aren't defined
-to be lists of variable declarations. The variable-free (or name-free) approach 
+to be lists of variable declarations. The variable-free (a.k.a.~\emph{name-free}) approach 
 we take here is rather different than those appearing in 
@@
 
-In our treatment, a context can be the empty context or it can be a binary
-planar tree of which `the leaves are (families of) contexts'. 
+In our treatment, a contexts can be seen as binary
+planar trees of which the leaves are (families of) contexts. 
 The judgmental equality relation on contexts is an equivalence relation which 
@@
 a context', which one may assume.
-Only a family of contexts over a context $\Gamma$ is eligible to
-be a type in the context $\Gamma$ and we have the intuition that `being a type'
-expresses the property of `being an atomic or irreducible family of contexts'.
-We will not axiomatize that every context is either judgmentally equal to the
-empty context or judgmentally equal to the extension of a context by a type,
-although it is certainly worth investigating the class of contexts which fall
-under this category. However, we will primarily investigate families of
-contexts and leave discussion on the possibility of a typing judgment until the
-end of this section. With families of contexts being the principal things of
+A context is eligible to be a \emph{closed type} and over a context, a family
+may be a type. We have the intuition that `being a type'
+expresses the property of `being atomic or irreducible'. With a notion of types,
+one has the means to require a stratification by which every context is a finite
+extension of types starting at a closed type, and every family is likewise a
+finite extension of types. The idea behind this is to recover the theory of
+\autoref{Garner2014} {\color{red}(and possibly B-systems?)}. We stress that
+we will not include axioms asserting such a stratification in the E-system
+and we leave discussion on the possibility of a typing judgment until the
+end of this section.
+
+With families of contexts being the principal things of
 study, we have added a bit more generality to the theory, compared to the
-theory of categories with families presented in \cite{Dybjer1996}. The reason
-for this added generality is twofold:
+theory of categories with families presented in \cite{Dybjer1996}. Here are
+a couple of reasons for this extra generality:
 \begin{enumerate}
@@
 $\Gamma$ and it is not reasonable to expect that this results in a type.
+\item It is very easy in this setting to assert that an operation has an action
+on very high levels of families, i.e.~when $F$ is an operation taking things in
+context $\Gamma$ to things in context $\Delta$, then it takes a family $A$ in
+context $\Gamma$ to $F(A)$ in context $\Delta$; a family $P$ in context
+$\ctxext{\Gamma}{A}$ to a family $F(P)$ in context $\ctxext{\Delta}{(F(A))}$,
+and so on. The reason it becomes simple is that we don't have to explain what
+`and so on' means.
 \end{enumerate}
@@
 \subsection{The operations of the theory of contexts, families and terms}
-In this subsection we introduce the notions of empty context,
-extension, weakening, substitution and identity terms. These are all viewed as
-operations of the theory of contexts, families and terms.
-
+In this subsection we introduce the notions of
+extension, empty families, 
+weakening by families, weakening by contexts, 
+substitution and identity terms. The weakenings and substitution are both
+operations of the theory of contexts, families and terms and therefore they will
+have three levels of actions: on contexts (at the appropriate level), families
+and terms.
+
+\begin{comment}
 Generally, 
@@
 is formulated, this means that when we will write down a type theory that
-interprets itself in \autoref{ttmodels}, each of the notions of empty context,
-extension, weakening, substitution and identity terms are going to be internalized
+interprets itself in \autoref{ttmodels}, each of the notions of
+extension, empty family, weakening, substitution and identity terms are going to be internalized
 as terms of the theory and all terms are going to be compatible with these.
@@
 original operation.
+\end{comment}
+
+We note upfront that both context extension and family extension can also be
+seen as operations of contexts, families and terms, even though we will not
+introduce them as such. The reason is that extension is used to define what it
+means to be a family over a family over a context (namely, a family over an
+extended context). As such, it will have a trivial action on families and terms
+which will not need to be introduced. We will come back to this point in
+\autoref{extension}.
 
@@
 \begin{align}
-%& \inference
-%  { \jfam{\Gamma}{A}
-%    \jfam{{\Gamma}{A}}{P}
-%    }
-%  { \jctxeq{\ctxext{{\Gamma}{A}}{P}}{\ctxext{\Gamma}{{A}{P}}}
-%    }
-%  \\
+& \inference
+  { \jfam{\Gamma}{A}
+    \jfam{{\Gamma}{A}}{P}
+    }
+  { \jctxeq{\ctxext{{\Gamma}{A}}{P}}{\ctxext{\Gamma}{{A}{P}}}
+    }
+  \\
 & \inference
@@
 
-\subsubsection{Weakening}
+\subsubsection{Weakening by a family}
 \label{weakening}
@@
 \end{align}
-The following inference rules express that when the empty context or family is
+The following inference rules express that when the empty family is
 weakened, the result is the empty family.
@@
 \end{align}
+\begin{comment}
 Note that we do not have to axiomatize the action of weakening
@@
 $\emptytm$.
+\end{comment}
 
@@
 
+\subsubsection{Weakening by a context}
+Since we have not required that there is an empty context with the property that
+a context is the same thing as a family in the empty context, we will have to
+explicitly introduce weakening by contexts. Since it will allways be clear
+whether something is weakened by a context or by a family, we shall use the
+same notation for weakening by contexts as for weakening by families. We shall
+have to repeat all the compatibility rules for weakening by families for
+weakening by contexts; there is no way around it. It is
+actually this sort of inconvenience that has originally led us to include an
+empty context and be done with it.
+
+In the following inference rules we assume that we have a context $\Gamma$ and
+that $\jctxeq{\Gamma}{\Gamma'}$ in the rules where $\Gamma'$ makes an appearence.
+\begin{align}
+& \inference
+  { \jctx{\Delta}
+    }
+  { \jfam{\Gamma}{\ctxwk{\Gamma}{\Delta}}
+    }
+& & \inference
+  { \jctxeq{\Delta}{\Delta'}
+    }
+  { \jfameq{\Gamma}{\ctxwk{\Gamma}{\Delta}}{\ctxwk{\Gamma'}{\Delta'}}
+    }
+  \\
+& \inference
+  { \jfam{\Delta}{B}
+    }
+  { \jfam{\ctxext{\Gamma}{\ctxwk{\Gamma}{\Delta}}}{\ctxwk{\Gamma}{B}}
+    }
+& & \inference
+  { \jfameq{\Delta}{B}{B'}
+    }
+  { \jfameq
+      {\ctxext{\Gamma}{\ctxwk{\Gamma}{\Delta}}}
+      {\ctxwk{\Gamma}{B}}
+      {\ctxwk{\Gamma'}{B'}}
+    }
+  \\
+& \inference
+  { \jterm{\Delta}{B}{y}
+    }
+  { \jterm
+      {\ctxext{\Gamma}{\ctxwk{\Gamma}{\Delta}}}
+      {\ctxwk{\Gamma}{B}}
+      {\ctxwk{\Gamma}{y}}
+    }
+& & \inference
+  { \jtermeq{\Delta}{B}{y}{y'}
+    }
+  { \jtermeq
+      {\ctxext{\Gamma}{\ctxwk{\Gamma}{\Delta}}}
+      {\ctxwk{\Gamma}{B}}
+      {\ctxwk{\Gamma}{y}}
+      {\ctxwk{\Gamma}{y'}}
+    }
+\end{align}
+Note that we do not employ three different notations for the three different
+actions of weakening by a context because there can be no confusion about which
+of the three actions is meant. If there were an empty context we should have
+distinguished the three as we did for weakening by a family.
+
+We shall have similar compatibility rules for weakening by a context as we did
+for weakening by a family.
+
+{\color{red}To do. It will look exactly the same as for weakening by families}
+
 \subsubsection{Substitution}
@@
 \end{align}
+\begin{comment}
 Note again that we do not have to axiomatize the action of substitution by $x$ on the term
@@
 term in the empty family is judgmentally equal to the term $\emptytm$.
+\end{comment}
 
@@
 weakening and substitution, identity terms will allow us to speak of projections,
-of a variant of extension for terms and of fiber inclusions. The following
+of a variant of extension for terms and of fiber inclusions.
+The following
 inference rules are required to be valid:
@@
 \end{align}
+We stress again
+that because we have not required there to be an empty context, we need to
+explicitly call for identity terms $\idtm{\Gamma}$ of $\ctxwk{\Gamma}{\Gamma}$
+in context $\Gamma$ for each context $\Gamma$. Thus, we will also require
+the inference rule
+\begin{align}
+& \inference
+  { \jctx{\Gamma}
+    }
+  { \jterm{\Gamma}{\ctxwk{\Gamma}{\Gamma}}{\idtm{\Gamma}}
+    }
+& &
+  \inference
+  { \jctxeq{\Gamma}{\Gamma'}
+    }
+  { \jtermeq{\Gamma}{\ctxwk{\Gamma}{\Gamma}}{\idtm{\Gamma}}{\idtm{\Gamma'}}
+    }
+\end{align}
+For the following inference rules, there should also be versions of identity
+terms at contexts.
 
@@
 
+\begin{comment}
 \subsubsection{Substituting by the term of the empty family}
@@
 \end{align}
+\end{comment}
 
@@
 belongs to the class of types. We think of types as \emph{irreducible} families,
-i.e.\ families which are neither the empty context nor the extension of two
-families which are both not the empty context (in algebraic terminology: which
+i.e.\ families which are neither empty nor the extension of two
+families which are both not empty (in algebraic terminology: which
 are both non-trivial). To allow ourselves to speak of types we introduce two
@@
 As pointed out at the beginning of this subsection, 
-we do not assume that the empty context is a type, that would be like
+we do not assume that the empty family is a type, that would be like
 assuming that the multiplicative unit of a ring is prime. We envision a
@@
 contexts are interpreted as finite paths of arrows belonging to a predetermined
-class. The empty context would be such a path of length $0$ whereas the types
+class. The empty family would be such a path of length $0$ whereas the types
 would be such paths of length $1$.

### 2014-06-18 article-models-tt.tex
reintroduced the empty context

@@
 core of the language of dependent logic -- the basic ingredients
-of this theory will be the operations of extension, empty families, weakening and substitution
+of this theory will be the operations of extension on context and on families, 
+and empty context and empty families, weakening, substitution
 and the identity terms. The resulting theory can be seen as a manifestation of the 
@@
 In this subsection we introduce the notions of
-extension, empty families, 
-weakening by families, weakening by contexts, 
-substitution and identity terms. The weakenings and substitution are both
+extension on context and on families, the empty context and the empty families, 
+weakening, 
+substitution and identity terms.
+
+Weakening and substitution are both
 operations of the theory of contexts, families and terms and therefore they will
@@
 
-\subsubsection{The empty families}
+\subsubsection{The empty context and the empty families}
 \label{empty}
-We introduce an empty family over $\Gamma$ for every context $\Gamma$. It was
-a suggestion of Voevodsky and Garner to only include empty families and not an
-empty context. Having an empty context requires also rules asserting that a
-context is the same thing as a family over the empty context and that operations
-act the same on families over the empty context as on contexts. Not including
-an empty context changes the theory a bit. One example of a difference is that
-if we want weakening by a context we shall have to explicitly state it as it 
-would not follow from weakening by a family. A second example is that without
-an empty context, some rule schemes (like the compatibility of extension with
-the empty families, stated in \autoref{comp-0e} below) become slightly
-assymetrical. Another example is that it would
-not be possible to consider closed types and terms without an empty context.
-Although we have opted to not include an empty context by default, we
-shall give the rules in \autoref{empty-context} so that one can look them up
-if they are desired.
-
-{\color{red}Explain why empty families are needed}
+We introduce an empty context and an family over $\Gamma$ for every context $\Gamma$. 
+It has been suggested to only include empty families and not an
+empty context because an empty context is not necessary, but we do have several
+reasons to include them. Having an empty context requires also rules asserting 
+that a context is the same thing as a family over the empty context and this
+gives the categorical structure on contexts for free once one has it for 
+families. The main ingredients that will be missing from the theory once an
+empty context is avoided are weakening by a context and identity morphisms from
+a context to itself. Including these by hand also requires to formulate all the
+compatibility rules involving weakening once more for the cases of weakening by
+a context and identity terms at contexts. We prefer to state these rules once
+and only once and including an empty context helps in this respect.
+
+We also prefer to have our scheme of compatibility rules as symmetrical as
+possible. The structure of type dependency should look exactly the same in the
+default case as in any context. In that respect, having an empty family but not
+an empty context seems a bit odd. Also, some sets of compatibility rules 
+(like the rules stating that extension is compatible with the empty families)
+will become assymetrical as a result of not including an empty context. Moreover,
+we would eventually like to include a stratification of the theory by means of
+a type judgment (asserting that a family in a context $\Gamma$ is a type) and
+study closed types (i.e.~types in the empty context). It would be possible to
+provide a notion of closed types without having an empty context, but this would
+have to be formulated separately and we would have to restate all the rules for
+types (if any) for closed types all over again.
+
+One of the main uses of the empty context and the empty families is that we
+get the property that the `action on contexts' of an operation is compatible
+with its `action on families'.
 
@@
 inference rule:
-\begin{equation}
-\inference
+\begin{align}
+& \inference
+  { }
+  { \jctx{\emptyc}
+    }
+  \\
+& \inference
   { \jctx{\Gamma}
@@
     }
-\end{equation}
-Since it is not really necessary to annotate the empty family with its context,
-we shall usually denote $\emptyf[\Gamma]$ by $\emptyf$.
+  \\
+& \inference
+  { \jctxeq{\Gamma}{\Gamma'}
+    }
+  { \jfameq{\Gamma}{\emptyf[\Gamma]}{\emptyf[\Gamma']}
+    }
+\end{align}
+
+By regarding contexts as families of contexts over the empty context, we
+enable ourselves also to speak of terms of contexts. A term of a context
+$\Gamma$ is a term of the family $\Gamma$ over the empty context. These ideas
+are captured in the following convertibility rules:
+\begin{align}
+& \inference
+  { \jctx{\Gamma}
+    }
+  { \jfam{\emptyc}{\Gamma}
+    } 
+& & \inference
+    { \jfam{\emptyc}{\Gamma}
+      }
+    { \jctx{\Gamma}
+      }
+    \\
+& \inference
+  { \jctxeq{\Gamma}{\Delta}
+    }
+  { \jfameq{\emptyc}{\Gamma}{\Delta}
+    }
+& & \inference
+    { \jfameq{\emptyc}{\Gamma}{\Delta}
+      }
+    { \jctxeq{\Gamma}{\Delta}
+      }
+\end{align}
 
@@
 
-By regarding contexts as families of contexts over the empty context, we
-enable ourselves also to speak of terms of contexts. A term of a context
-$\Gamma$ is a term of the family $\Gamma$ over the empty context. These ideas
-are captured in the following convertibility rules:
-\begin{align}
-& \inference
-  { \jctx{\Gamma}
-    }
-  { \jfam{\emptyc}{\Gamma}
-    } 
-& & \inference
-    { \jfam{\emptyc}{\Gamma}
-      }
-    { \jctx{\Gamma}
-      }
-    \\
-& \inference
-  { \jctxeq{\Gamma}{\Delta}
-    }
-  { \jfameq{\emptyc}{\Gamma}{\Delta}
-    }
-& & \inference
-    { \jfameq{\emptyc}{\Gamma}{\Delta}
-      }
-    { \jctxeq{\Gamma}{\Delta}
-      }
-\end{align}
-
-Because a context is the same thing as a family of contexts over the empty
-context, we may introduce other type theoretic operations by only describing
-operations on families of contexts. These will then automatically translate
-to opearations on contexts as well. This and the notion of terms of a context
-give the empty context a fundamental role in the theory of contexts, families
-and terms and we feel that the purpose of the empty context should not be
-diminished to mere esthetical reasons, as might be the case in the setting
-of models of the theory \cite{Awodey2013}.
-
 \label{comp-00}
@@
 
+\paragraph{Compatibility of extension with the empty context and families}
 In the following set of inference rules we state that the empty context and
-the empty family are neutral objects when extension is concerned.
+the empty family are neutral objects for both context extension (the first two
+rules) and family extension (the last two rules).
 \label{comp-e0}\label{comp-0e}
@@
     }
+  { \jctxeq{\ctxext{\emptyc}{\Gamma}}{\Gamma}
+    }
+  \label{comp-0e-c}
+  \\
+& \inference
+  { \jctx{\Gamma}
+    }
   { \jctxeq{\ctxext{\Gamma}{\emptyf}}{\Gamma}
@@
   \label{comp-e0-f}
-%& \inference
-%  { \jctx{\Gamma}
-%    }
-%  { \jctxeq{\ctxext{\emptyc}{\Gamma}}{\Gamma}
-%    }
-%  \label{comp-0e-c}\\
-\end{align}
-%Note the special case where we take $\Gamma$ to be the empty context.
-
-\subsubsection{Weakening by a family}
+\end{align}
+
+\paragraph{Family extension restricted to the empty context}
+Strictly speaking we should have used a different notation for context extension
+as for family extension, because the following rule asserting that family extension
+in the empty context is the same thing as context extension would look tautological
+without a difference. So let us denote, only for the moment, context extension
+of $\Gamma$ by $A$ by $(\ctxext{\Gamma}{A})^c$ and family extension of $A$ by
+$P$ in context $\Gamma$ by $(\ctxext{A}{P})^\famsym$.
+
+Note that we may consider a context $\Gamma$ as a family over $\emptyc$ and
+a family $\jfam{\Gamma}{A}$ as a family $\jfam{{\emptyc}{\Gamma}}{A}$. 
+Therefore we will require the following rule:
+\begin{align}
+& \inference
+  { \jfam{\Gamma}{A}
+    }
+  { \jctxeq{(\ctxext{\Gamma}{A})^c}{(\ctxext{\Gamma}{A})^\famsym}
+    }
+\end{align}
+Note that this rule actually justifies that we have not utilized two different
+notations for context extension and family extension.
+
+\subsubsection{Weakening}
 \label{weakening}
@@
 \end{align}
+
+\paragraph{Weakening of an empty family}
 The following inference rules express that when the empty family is
@@
 Because a family over $\Gamma$ is the same as a family over 
-$\ctxext{\Gamma}{\emptyf}$, we can apply both the action on contexts and the
+$\ctxext{\Gamma}{\emptyf}$ we can apply both the action on contexts and the
 action on families of weakening to a family $B$ over $\Gamma$. When we apply
@@
 
+\begin{comment}
 \subsubsection{Weakening by a context}
@@
 {\color{red}To do. It will look exactly the same as for weakening by families}
+\end{comment}
 
@@
 \end{align}
+
+\paragraph{Fibers of an empty family}
 The following inference rules establish that the fibers of the empty family are 
@@
 \end{align}
+
+\begin{comment}
 We stress again
@@
 terms at contexts.
+\end{comment}
 

### 2014-06-18 article-models-tt.tex
reintroduced the empty context

@@
 We introduce an empty context and an family over $\Gamma$ for every context $\Gamma$. 
-It has been suggested to only include empty families and not an
+It has been suggested by some to only include empty families and not an
 empty context because an empty context is not necessary, but we do have several

### 2014-06-18 article-models-main.tex
typo

@@
 Clairambault in \cite{DybjerClairambault2011} and in unpublished work by
-Awody \cite{Awodey2013} on natural models of type theory,
+Awodey \cite{Awodey2013} on natural models of type theory,
 which makes the connection between categories with families and representable

### 2014-06-18 article-models-ttderived.tex
version esystems sent to VV and RG
terms of the empty family

@@
 \section{Derived notions of the theory of contexts, families and terms}
-\label{digging_deeper}
+\label{ttderived}
 
@@
 we show that term extension and the projections are
-compatible with the empty families, extension,
-weakening, substitution and the identity terms (in that order). \autoref{lem:tmext-id}
+compatible with extension (\autoref{lem:tmext-tmext}),
+weakening (\autoref{lem:tmext-wk}), substitution (\autoref{lem:tmext-subst}) and the identity terms (\autoref{lem:tmext-id}). The latter property, \autoref{lem:tmext-id},
 is in fact a generalization of the above \autoref{lem:tmext-basic}.
-
-\begin{lem}\label{lem:tmext-emp}
-The following compatibility rules for extensions of the term of the empty family
-are valid:
-\begin{align*}
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    }
-  { \jtermeq{\Gamma}{A}{\tmext{\emptytm}{x}}{x}
-    }
-& & \inference
-  { \jterm{\Gamma}{A}{x}
-    }
-  { \jtermeq{\Gamma}{A}{\tmext{x}{\emptytm}}{x}
-    }
-  \\
-& \inference
-  { \jfam{\Gamma}{A}
-    }
-  { \jtermeq
-      {{\Gamma}{A}}
-      {\emptyf}
-      {\cprojfstf{\emptyf}{A}}
-      {\emptytm}
-    }
-& & \inference
-  { \jfam{\Gamma}{A}
-    }
-  { \jhomeq
-      {\Gamma}
-      {A}
-      {A}
-      {\cprojfstf{A}{\emptyf}}
-      {\idtm{A}}
-    }
-  \\
-& \inference
-  { \jfam{\Gamma}{A}
-    }
-  { \jhomeq
-      {\Gamma}
-      {A}
-      {A}
-      {\cprojsndf{\emptyf}{A}}
-      {\idtm{A}}
-    }
-& & \inference
-  { \jfam{\Gamma}{A}
-    }
-  { \jtermeq
-      {{\Gamma}{A}}
-      {\emptyf}
-      {\cprojsndf{A}{\emptyf}}
-      {\emptytm}
-    }
-\end{align*}
-\end{lem}
-
-\begin{proof}
-These equalities are very easy to verify. We only display a proof of the first:
-\begin{equation*}
-\tmext{\emptytm}{x}
-\jdeq \unfold{\tmext{\emptyf}{A}{\emptytm}{x}}
-\jdeq \subst{x}{\idtm{{\emptyf}{A}}}
-\jdeq \subst{x}{\idtm{A}}
-\jdeq x.\qedhere
-\end{equation*}
-\end{proof}
 
@@
 We have the following lemmas expressing the compatibility of the fiber
-inclusions with the empty context, extension, weakening and substitution.
-
-\begin{lem}
-The fiber inclusions are compatible with the empty families; i.e.~the following
-inference rules are valid
-\begin{align*}
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    }
-  { \jtermeq
-      {\Gamma}
-      {A}
-      {\finc{x}{\emptyf}}
-      {x}
-    }
-  \\
-& \inference
-  { \jfam{\Gamma}{A}
-    }
-  { \jhomeq
-      {\Gamma}
-      {A}
-      {A}
-      {\finc{\emptytm}{A}}
-      {\idtm{A}}
-    }
-\end{align*}
-\end{lem}
-
-\begin{proof}
-Let $\jterm{\Gamma}{A}{x}$. We have the judgmental equalities
-\begin{align*}
-\finc{x}{\emptyf}
-& \jdeq
-  \unfold{\finc{x}{\emptyf}}
-  \tag{by definition}
-  \\
-& \jdeq
-  \tmext{\ctxwk{\emptyf}{x}}{\idtm{\emptyf}}
-  \tag{by \autoref{comp-s0-c}}
-  \\
-& \jdeq
-  \tmext{x}{\idtm{\emptyf}}
-  \tag{by \autoref{comp-0w-t}}
-  \\
-& \jdeq
-  \tmext{x}{\emptytm}
-  \tag{by \autoref{comp-00-t}}
-  \\
-& \jdeq
-  x
-  \tag{by \autoref{lem:tmext-emp}}
-\end{align*}
-to prove the first judgmental equality. For the second, we have
-\begin{align*}
-\finc{\emptytm}{A}
-& \jdeq
-  \unfold{\finc{\emptytm}{A}}
-  \tag{by definition}
-  \\
-& \jdeq
-  \tmext{\emptytm}{\idtm{\subst{\emptytm}{A}}}
-  \tag{by \autoref{comp-00-t}}
-  \\
-& \jdeq
-  \idtm{\subst{\emptytm}{A}}
-  \tag{by \autoref{lem:tmext-emp}}
-  \\
-& \jdeq
-  \idtm{A}
-  \tag{by \autoref{comp-0s-f}}.
-\end{align*}
-\end{proof}
+inclusions with extension, weakening and substitution.
 
@@
 \end{comment}
+
+\subsection{The terms of the empty families}
+In this subsection we will define and derive properties of the term of the empty
+family in a context. We will leave most proofs to the reader.
+
+For the following definition, note that we have the judgmental equalities
+$\jctxeq{\Gamma}{{\Gamma}{\emptyf}}$ and 
+$\jfameq{\Gamma}{\ctxwk{\emptyf}{\emptyf}}{\emptyf}$.
+
+\begin{defn}
+Let $\Gamma$ be a context. We define
+\begin{equation*}
+\jtermdefn{\Gamma}{\emptyf}{\emptytm}{\idtm{\emptyf}}
+\end{equation*}
+\end{defn}
+
+\begin{rmk}
+There seems to be no way to derive that this term is
+judgmentally unique. Also, there seems to be no reason we need that. This has
+been a suggestion of Voevodsky and Garner.
+\end{rmk}
+
+We will first show that the terms of the empty families are compatible with
+weakening and substitution before we show that they are compatible with
+extension on terms.
+
+\begin{lem}\label{lem:emptytm-wk}
+The following inference rule expressing that the terms of the empty families are
+compatible with weakening is valid:
+\begin{equation*}
+\inference
+  { \jfam{\Gamma}{A}
+    }
+  { \jtermeq
+      {{\Gamma}{A}}
+      {\emptyf}
+      {\ctxwk{A}{\emptytm}}
+      {\emptytm}
+    }
+\end{equation*}
+\end{lem}
+
+\begin{lem}\label{lem:emptytm-subst}
+The following inference rule expressing that the terms of the empty families are
+compatible with substitutuion is valid:
+\begin{equation*}
+\inference
+  { \jterm{\Gamma}{A}{x}
+    }
+  { \jtermeq
+      {\Gamma}
+      {\emptyf}
+      {\subst{x}{\emptytm}}
+      {\emptytm}
+    }
+\end{equation*}
+\end{lem}
+
+\begin{lem}\label{comp-0s}
+The following rules asserting that substituting by the term $\jterm{\Gamma}{\emptyf}{\emptytm}$
+leaves everything unchanged are valid:
+\begin{align}
+& \inference
+  { \jfam{\Gamma}{A}
+    }
+  { \jfameq{\Gamma}{\subst{\emptytm}{A}}{A}
+    }
+  \\
+%& \inference
+%  { \jterm{\Gamma}{A}{x}
+%    }
+%  { \jtermeq{\Gamma}{A}{\subst{\emptytm}{x}}{x}
+%    }
+%  \\
+& \inference
+  { \jfam{{\Gamma}{A}}{P}
+    }
+  { \jfameq{{\Gamma}{A}}{\subst{\emptytm}{P}}{P}
+    }
+  \\
+& \inference
+  { \jterm{{\Gamma}{A}}{P}{f}
+    }
+  { \jtermeq{{\Gamma}{A}}{P}{\subst{\emptytm}{f}}{f}
+    }.
+\end{align}
+\end{lem}
+
+\begin{proof}
+This follows directly from the properties in \autoref{comp-0w,cancellation-i}.
+\end{proof}
+
+\begin{lem}\label{lem:tmext-emp}
+The following compatibility rules for extensions of the term of the empty family
+are valid:
+\begin{align*}
+& \inference
+  { \jterm{\Gamma}{A}{x}
+    }
+  { \jtermeq{\Gamma}{A}{\tmext{\emptytm}{x}}{x}
+    }
+& & \inference
+  { \jterm{\Gamma}{A}{x}
+    }
+  { \jtermeq{\Gamma}{A}{\tmext{x}{\emptytm}}{x}
+    }
+  \\
+& \inference
+  { \jfam{\Gamma}{A}
+    }
+  { \jtermeq
+      {{\Gamma}{A}}
+      {\emptyf}
+      {\cprojfstf{\emptyf}{A}}
+      {\emptytm}
+    }
+& & \inference
+  { \jfam{\Gamma}{A}
+    }
+  { \jhomeq
+      {\Gamma}
+      {A}
+      {A}
+      {\cprojfstf{A}{\emptyf}}
+      {\idtm{A}}
+    }
+  \\
+& \inference
+  { \jfam{\Gamma}{A}
+    }
+  { \jhomeq
+      {\Gamma}
+      {A}
+      {A}
+      {\cprojsndf{\emptyf}{A}}
+      {\idtm{A}}
+    }
+& & \inference
+  { \jfam{\Gamma}{A}
+    }
+  { \jtermeq
+      {{\Gamma}{A}}
+      {\emptyf}
+      {\cprojsndf{A}{\emptyf}}
+      {\emptytm}
+    }
+\end{align*}
+\end{lem}
+
+\begin{proof}
+We only display a proof of the first:
+\begin{equation*}
+\tmext{\emptytm}{x}
+\jdeq \unfold{\tmext{\emptyf}{A}{\emptytm}{x}}
+\jdeq \subst{x}{\idtm{{\emptyf}{A}}}
+\jdeq \subst{x}{\idtm{A}}
+\jdeq x.\qedhere
+\end{equation*}
+\end{proof}
+
+\begin{lem}\label{lem:finc-emp}
+The fiber inclusions are compatible with the empty families; i.e.~the following
+inference rules are valid
+\begin{align*}
+& \inference
+  { \jterm{\Gamma}{A}{x}
+    }
+  { \jtermeq
+      {\Gamma}
+      {A}
+      {\finc{x}{\emptyf}}
+      {x}
+    }
+  \\
+& \inference
+  { \jfam{\Gamma}{A}
+    }
+  { \jhomeq
+      {\Gamma}
+      {A}
+      {A}
+      {\finc{\emptytm}{A}}
+      {\idtm{A}}
+    }
+\end{align*}
+\end{lem}
+
+\begin{proof}
+Let $\jterm{\Gamma}{A}{x}$. We have the judgmental equalities
+\begin{align*}
+\finc{x}{\emptyf}
+& \jdeq
+  \unfold{\finc{x}{\emptyf}}
+  \tag{by definition}
+  \\
+& \jdeq
+  \tmext{\ctxwk{\emptyf}{x}}{\idtm{\emptyf}}
+  \tag{by \autoref{comp-s0-c}}
+  \\
+& \jdeq
+  \tmext{x}{\idtm{\emptyf}}
+  \tag{by \autoref{comp-0w-t}}
+  \\
+& \jdeq
+  \tmext{x}{\emptytm}
+  \tag{by definition}
+  \\
+& \jdeq
+  x
+  \tag{by \autoref{lem:tmext-emp}}
+\end{align*}
+to prove the first judgmental equality. For the second, we have
+\begin{align*}
+\finc{\emptytm}{A}
+& \jdeq
+  \unfold{\finc{\emptytm}{A}}
+  \tag{by definition}
+  \\
+& \jdeq
+  \tmext{\emptytm}{\idtm{\subst{\emptytm}{A}}}
+  \tag{by \autoref{lem:emptytm-wk}}
+  \\
+& \jdeq
+  \idtm{\subst{\emptytm}{A}}
+  \tag{by \autoref{lem:tmext-emp}}
+  \\
+& \jdeq
+  \idtm{A}
+  \tag{by \autoref{comp-0s}}.
+\end{align*}
+\end{proof}

### 2014-06-18 article-models-ttderived.tex
revised rules of identity terms

@@
 \end{equation*}
+As usual, when $f$ is a morphism from $A$ to $B$ in context $\Gamma$ we say that
+$A$ is the \emph{domain of $f$} and that $B$ is the \emph{codomain of $f$}. 
 \end{defn}
+
+In two special cases we have alternative explanations of what a morphism is.
+These special cases are when the domain is either empty or an extension.
+
+\begin{lem}
+We have the following valid inference rules expressing that a morphism from
+the empty family in context $\Gamma$ is the same thing as a term:
+\begin{align*}
+& \inference
+  { \jhom{\Gamma}{\emptyf}{A}{x}
+    }
+  { \jterm{\Gamma}{A}{x}
+    }
+  \\
+& \inference
+  { \jterm{\Gamma}{A}{x}
+    }
+  { \jhom{\Gamma}{\emptyf}{A}{x}
+    }
+\end{align*}
+\end{lem}
+
+\begin{lem}
+We have the following valid inference rules expressing that a morphism from
+an extension $\ctxext{A}{P}$ to $B$ is the same thing as a morphism from $P$ into
+$\ctxwk{A}{B}$:
+\begin{align*}
+& \inference
+  { \jhom{\Gamma}{{A}{P}}{B}{f}
+    }
+  { \jhom{{\Gamma}{A}}{P}{\ctxwk{A}{B}}{f}
+    }
+  \\
+& \inference
+  { \jhom{{\Gamma}{A}}{P}{\ctxwk{A}{B}}{f}
+    }
+  { \jhom{\Gamma}{{A}{P}}{B}{f}
+    }
+\end{align*}
+\end{lem}
+
+We devote the rest of this subsection to precomposition by a morphism. We treat
+precomposition by a morphism $f$ from $A$ to $B$ in context $\Gamma$ as an
+\emph{operation} from things in context $\ctxext{\Gamma}{B}$ to things in
+context $\ctxext{\Gamma}{A}$ rather than as something which only acts on
+morphisms with domain $B$. Since precomposition by a morphism is seen as an
+operation, we will give it an `action on contexts', an `action on families' and
+and `action on terms' and we will follow the usual scheme in verifying that
+it is compatible with the empty context and families, with context and family
+extension, weakening, substitution and identity terms. Moreover, we will show
+that precomposition is compatible with itself; this is a slight generalization
+of associativity of composition. We will be able to retrieve the usual
+composition of morphisms via the action on terms of the precomposition operation.
 
@@
 
-We have lots of compatibility properties for composition. On the one hand we
-have the properties that composition with a morphism $f$ is compatible with
-the empty family, extension, weakening, substitution and identity terms. On
-the other hand, there are compatibility properties saying what happens when
-we substitute by a composition, or when we weaken or substitute a composition.
-Proving all these compatibility properties is the content of the rest of this
-subsection. None of it is very difficult.
-
 \begin{lem}\label{lem:jcomp-ext}
@@
 
-Because composition is proven to be associative, we can draw diagrams involving
+\begin{lem}
+The following inference rule is valid
+\begin{equation*}
+\inference
+  { \jhom{\Gamma}{A}{B}{f}
+    }
+  { \jhomeq{\Gamma}{A}{B}{\jcomp{A}{f}{\idtm{B}}}{f}
+    }
+\end{equation*}
+\end{lem}
+
+\begin{proof}
+We have the judgmental equalities
+\begin{align*}
+\jcomp{A}{f}{\idtm{B}}
+& \jdeq
+  \subst{f}{\ctxwk{A}{\idtm{B}}}
+  \tag{by definition}
+  \\
+& \jdeq
+  \subst{f}{\idtm{\ctxwk{A}{B}}}
+  \tag{by \autoref{comp-wi}}
+  \\
+& \jdeq
+  f.
+  \tag{by \autoref{cancellation-si}}
+\end{align*}
+\end{proof}
+
+Because composition is proven to be associative and because identity morphisms
+are neutral for both pre- and postcomposition, we can draw diagrams involving
 families and morphisms between them in a context $\Gamma$, just as is usual in
@@
     {{x}{Q}}
-  \tag{by \autoref{idfunc-wk-defn}}
+  \tag{by \autoref{precomp-idtm-c}}
 \end{align*}
@@
   \ctxwk{P}{P},
-  \tag{by \autoref{idfunc-wk-defn}}
+  \tag{by \autoref{precomp-idtm-c}}
 \end{align*}
@@
     {\tmext{u}{v}}{{x}{\idtm{\ctxext{A}{{P}{Q}}}}}
+  \tag{by \autoref{comp-es}}
   \\
@@
   \subst{v}{{u}{{x}{\idtm{\ctxext{A}{{P}{Q}}}}}}
+  \tag{by \autoref{comp-es}}
   \\
@@
   \subst{v}{{u}{{x}{\idtm{\ctxext{{A}{P}}{Q}}}}}
+  \tag{by \autoref{comp-ee-c}}
   \\
@@
   \subst{v}{{\tmext{x}{u}}{\idtm{\ctxext{{A}{P}}{Q}}}}
+  \tag{by \autoref{comp-es}}
   \\
@@
   \tmext{{x}{u}}{v}.
+  \tag{by \autoref{comp-es}}
 \end{align*}
@@
   \unfoldall{\jcomp{{{A}{P}}{Q}}{\cprojfstf{{A}{P}}{Q}}{\cprojfstf{A}{P}}}
+  \tag{by definition}
   \\
@@
     {\ctxwk{Q}{{\ctxext{A}{P}}{{P}{\idtm{A}}}}}
+  \tag{by \autoref{comp-ew-t}}
   \\
@@
   \ctxwk{Q}{\subst{\idtm{{A}{P}}}{\ctxwk{\ctxext{A}{P}}{{P}{\idtm{A}}}}}
+  \tag{by \autoref{comp-ws-t}}
   \\
@@
   \ctxwk{P}{Q} 
-  \tag{by \autoref{idfunc-wk-defn}}
+  \tag{by \autoref{precomp-idtm-c}}
 \end{align*}
@@
 $\ctxwk{P}{\subst{\idtm{A}}{\ctxwk{A}{Q}}}\jdeq\ctxwk{P}{Q}$, which
-holds by \autoref{idfunc-wk-defn}.
+holds by \autoref{precomp-idtm-c}.
 \end{rmk}
@@
   \subst{x}{Q}.
-  \tag{by \autoref{idfunc-precomp}}
+  \tag{by \autoref{precomp-idtm-t}}
 \end{align*}
@@
   \subst{x}{\idtm{\ctxext{A}{P}}}.
-  \tag{by \autoref{idfunc-precomp}}
+  \tag{by \autoref{precomp-idtm-t}}
 \end{align*}
@@
   \idtm{{A}{P}}
-  \tag{by \autoref{idfunc-precomp}}
+  \tag{by \autoref{precomp-idtm-t}}
 \end{align*}

### 2014-06-18 article-models-tt.tex
terms of the empty family
revised rules of identity terms
the big generalization of Richard

@@
 weakening distributes over the extension factors.
-\begin{align}\label{comp-we-f}
+\begin{align}
+& \inference
+  { \jfam{\Gamma}{A}
+    \jfam{{\Gamma}{B}}{Q}
+    }
+  { \jfameq
+      {\ctxext{\Gamma}{A}}
+      {\ctxwk{A}{\ctxext{B}{Q}}}
+      {\ctxext{\ctxwk{A}{B}}{\ctxwk{A}{Q}}}
+    }
+  \label{comp-we-c}
+  \\
 & \inference
@@
     }
+  \label{comp-we-f}
 \end{align}
@@
 \paragraph{Weakening of weakenings}
-When a weakened family is weakened even further, the outer weakening distributes
-over the `weakening factors'.
+Weakening by a family $A$ in context $\Gamma$ brings things in context $\Gamma$
+to things in context $\ctxext{\Gamma}{A}$. Since we have all the ingredients of
+the theory of contexts, families and terms in the context $\Gamma$ as well it
+has in particular it's own weakening by families. Suppose we have a family
+$\jfam{{\Gamma}{B}}{Q}$ over $B$ in context $\Gamma$. Weakening by $Q$ brings
+things from context $\ctxext{\Gamma}{B}$ to $\ctxext{{\Gamma}{B}}{Q}$. Thus
+we can provide rules asserting what will happen when we first weaken by $Q$ and
+then (via the action on families) by $A$. We will require the following
+inference rules to be valid:
 \label{comp-ww}
 \begin{align}
+& \inference
+  { \jfam{\Gamma}{A}
+    \jfam{{\Gamma}{B}}{Q}
+    \jfam{{\Gamma}{B}}{R}
+    }
+  { \jfameq
+      {{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{Q}}}
+      {\ctxwk{A}{{Q}{R}}}
+      {\ctxwk{{A}{Q}}{{A}{R}}}
+    }
+  \label{comp-ww-c}\\
+& \inference
+  { \jfam{\Gamma}{A}
+    \jfam{{\Gamma}{B}}{Q}
+    \jfam{{{\Gamma}{B}}{R}}{S}
+    }
+  { \jfameq
+      {{{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{Q}}}{\ctxwk{A}{{Q}{R}}}}
+      {\ctxwk{A}{{Q}{S}}}
+      {\ctxwk{{A}{Q}}{{A}{S}}}
+    }
+  \label{comp-ww-f}\\
+& \inference
+  { \jfam{\Gamma}{A}
+    \jfam{{\Gamma}{B}}{Q}
+    \jterm{{{\Gamma}{B}}{R}}{S}{k}
+    }
+  { \jtermeq
+      {{{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{Q}}}{\ctxwk{A}{{Q}{R}}}}
+      {\ctxwk{A}{{Q}{S}}}
+      {\ctxwk{A}{{Q}{k}}}
+      {\ctxwk{{A}{Q}}{{A}{k}}}
+    }
+  \label{comp-ww-t}
+\end{align}
+
+\begin{rmk}
+As an important special case of these inference rules we have the following
+valid inference rules:
+\begin{align*}
+& \inference
+  { \jfam{\Gamma}{A}
+    \jfam{\Gamma}{B}
+    \jfam{\Gamma}{C}
+    }
+  { \jfameq
+      {{{\Gamma}{A}}{\ctxwk{A}{B}}}
+      {\ctxwk{A}{{B}{C}}}
+      {\ctxwk{{A}{B}}{{A}{C}}}
+    }
+  \\
 & \inference
@@
   { \jfameq
-      {{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{{A}{B}}{{A}{C}}}}
+      {{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{{B}{C}}}}
       {\ctxwk{A}{{B}{R}}}
@@
     }
-  \label{comp-ww-f}\\
+  \\
 & \inference
@@
     \jfam{\Gamma}{B}
-    \jterm{{\Gamma}{C}}{R}{t}
+    \jterm{{\Gamma}{C}}{R}{h}
     }
   { \jtermeq
-      {{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{{A}{B}}{{A}{C}}}}
-      {\ctxwk{{A}{B}}{{A}{R}}}
-      {\ctxwk{A}{{B}{t}}}
-      {\ctxwk{{A}{B}}{{A}{t}}}
-    }
-  \label{comp-ww-t}
-\end{align}
+      {{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{{B}{C}}}}
+      {\ctxwk{A}{{B}{R}}}
+      {\ctxwk{A}{{B}{h}}}
+      {\ctxwk{{A}{B}}{{A}{h}}}
+    }
+\end{align*}
+Moreover, we have
+\begin{equation*}
+\inference
+  { \jfam{\Gamma}{A}
+    \jfam{\Gamma}{B}
+    \jterm{\Gamma}{C}{z}
+    }
+  { \jtermeq
+      {{{\Gamma}{A}}{\ctxwk{A}{B}}}
+      {\ctxwk{A}{{B}{C}}}
+      {\ctxwk{A}{{B}{z}}}
+      {\ctxwk{{A}{B}}{{A}{z}}}
+    }
+\end{equation*}
+\end{rmk}
 
@@
 \label{comp-se}
-The following inference rule asserts that if we take the fiber of an extended
+The following inference rules assert that if we take the fiber of an extended
 family at a term $x$ of $A$ in context $\Gamma$, the substitution by $x$
 distributes over the factors of the extension.
-\begin{align}\label{comp-se-f}
+\begin{align}
+& \inference
+  { \jterm{\Gamma}{A}{x}
+    \jfam{{{\Gamma}{A}}{P}}{Q}
+    }
+  { \jfameq
+      {\Gamma}
+      {\subst{x}{\ctxext{P}{Q}}}
+      {\ctxext{\subst{x}{P}}{\subst{x}{Q}}}
+    }
+  \label{comp-se-c}
+  \\
 & \inference
@@
     }
+  \label{comp-se-f}
 \end{align}
@@
 \paragraph{Fibers of weakenings}\label{comp-sw}
-Substitutions distribute over weakenings:
-\begin{align}
+The following rules assert what happens when we first weaken by a family
+$Q$ over $P$ in context $\ctxext{\Gamma}{A}$ and then substitute by a term
+$x$ of $A$:
+\begin{align}
+& \inference
+  { \jterm{\Gamma}{A}{x}
+    \jfam{{{\Gamma}{A}}{P}}{Q}
+    \jfam{{{\Gamma}{A}}{P}}{R}
+    }
+  { \jfameq
+      {{{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}}
+      {\subst{x}{\ctxwk{Q}{R}}}
+      {\ctxwk{\subst{x}{Q}}{\subst{x}{R}}}
+    }
+  \label{comp-sw-c}
+  \\
+& \inference
+  { \jterm{\Gamma}{A}{x}
+    \jfam{{{\Gamma}{A}}{P}}{Q}
+    \jfam{{{{\Gamma}{A}}{P}}{R}}{S}
+    }
+  { \jfameq
+      {{{{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}}{\subst{x}{\ctxwk{Q}{R}}}}
+      {\subst{x}{\ctxwk{Q}{S}}}
+      {\ctxwk{\subst{x}{Q}}{\subst{x}{S}}}
+    }
+  \label{comp-sw-f}
+  \\
+& \inference
+  { \jterm{\Gamma}{A}{x}
+    \jfam{{{\Gamma}{A}}{P}}{Q}
+    \jterm{{{{\Gamma}{A}}{P}}{R}}{S}{k}
+    }
+  { \jtermeq
+      {{{{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}}{\subst{x}{\ctxwk{Q}{R}}}}
+      {\subst{x}{\ctxwk{Q}{S}}}
+      {\subst{x}{\ctxwk{Q}{k}}}
+      {\ctxwk{\subst{x}{Q}}{\subst{x}{k}}}
+    }
+  \label{comp-sw-t}
+\end{align}
+
+\begin{rmk}
+As an important special case of these inference rules we have the following
+valid inference rules:
+\begin{align*}
+& \inference
+  { \jterm{\Gamma}{A}{x}
+    \jfam{{\Gamma}{A}}{P}
+    \jfam{{\Gamma}{A}}{Q}
+    }
+  { \jfameq
+      {{\Gamma}{\subst{x}{P}}}
+      {\subst{x}{\ctxwk{P}{Q}}}
+      {\ctxwk{\subst{x}{P}}{\subst{x}{Q}}}
+    }
+  \\
 & \inference
@@
     }
-  \label{comp-sw-f}\\
+  \\
 & \inference
@@
   { \jtermeq
-      {\ctxext{{\Gamma}{\subst{x}{P}}}{\subst{x}{\ctxwk{P}{Q}}}}
+      {{{\Gamma}{\subst{x}{P}}}{\subst{x}{\ctxwk{P}{Q}}}}
       {\subst{x}{\ctxwk{P}{R}}}
@@
     }
-  \label{comp-sw-t}
-\end{align}
+\end{align*}
+Moreover, we get
+\begin{equation*}
+\inference
+  { \jterm{\Gamma}{A}{x}
+    \jfam{{\Gamma}{A}}{P}
+    \jterm{{\Gamma}{A}}{Q}{g}
+    }
+  { \jtermeq
+      {{\Gamma}{\subst{x}{P}}}
+      {\subst{x}{\ctxwk{P}{Q}}}
+      {\subst{x}{\ctxwk{P}{g}}}
+      {\ctxwk{\subst{x}{P}}{\subst{x}{g}}}
+    }
+\end{equation*}
+\end{rmk}
 
@@
 
-We require that substitution is compatible with itself, which is roughly the
-assertion that substitution is associative. However, we cannot just state that
-$\subst{x}{{f}{g}}\jdeq\subst{{x}{f}}{g}$ since the expression $\subst{{x}{f}}{g}$
-is not well-formed. The term $\subst{x}{f}$ can be substituted in (terms of) families over
-$\subst{x}{P}$; the term $\subst{x}{g}$ is such. Therefore, associativity of
-substitution takes the form $\subst{x}{{f}{g}}\jdeq\subst{{x}{f}}{{x}{g}}$.
-Note that the term $\subst{{x}{f}}{{x}{g}}$ may be written down more conveniently
-as $\subst{x,\subst{x}{f}}{g}$, although we will not do that. In fact, after
-we have introduced extension on terms in \autoref{extension-on-terms}, we will 
-see that $\subst{{x}{f}}{{x}{g}}\jdeq\subst{\tmext{x}{\subst{x}{f}}}{g}$.
-
-In the following inference rules we assume
-$\jterm{\Gamma}{A}{x}$ and $\jterm{{\Gamma}{A}}{P}{f}$.
-\begin{align}
-& \inference
-  { \jfam{{{{\Gamma}{A}}{P}}{Q}}{R}
+The following rules explain what happens when we first substitute by a term
+$g$ of $Q$ over $P$ in context $\ctxext{\Gamma}{A}$ and then by a term $x$ of
+$A$:
+\begin{align}
+& \inference
+  { \jterm{\Gamma}{A}{x}
+    \jterm{{{\Gamma}{A}}{P}}{Q}{g}
+    \jfam{{{{\Gamma}{A}}{P}}{Q}}{R}
+    }
+  { \jfameq
+      {{{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}}
+      {\subst{x}{{g}{R}}}
+      {\subst{{x}{g}}{{x}{R}}}
+    }
+  \label{comp-ss-c}
+  \\
+& \inference
+  { \jterm{\Gamma}{A}{x}
+    \jterm{{{\Gamma}{A}}{P}}{Q}{g}
+    \jfam{{{{{\Gamma}{A}}{P}}{Q}}{R}}{S}
+    }
+  { \jfameq
+      {{{{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}}{\subst{x}{{g}{R}}}}
+      {\subst{x}{{g}{S}}}
+      {\subst{{x}{g}}{{x}{S}}}
+    }
+  \label{comp-ss-f}
+  \\
+& \inference
+  { \jterm{\Gamma}{A}{x}
+    \jterm{{{\Gamma}{A}}{P}}{Q}{g}
+    \jterm{{{{{\Gamma}{A}}{P}}{Q}}{R}}{S}{k}
+    }
+  { \jtermeq
+      {{{{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}}{\subst{x}{{g}{R}}}}
+      {\subst{x}{{g}{S}}}
+      {\subst{x}{{g}{k}}}
+      {\subst{{x}{g}}{{x}{k}}}
+    }
+  \label{comp-ss-f}
+\end{align}
+
+\begin{rmk}
+As an important special case of these inference rules we have the following
+valid inference rules:
+\begin{align*}
+& \inference
+  { \jterm{\Gamma}{A}{x}
+    \jterm{{\Gamma}{A}}{P}{f}
+    \jfam{{{\Gamma}{A}}{P}}{Q}
+    }
+  { \jfameq
+      {\Gamma}
+      {\subst{x}{{f}{Q}}}
+      {\subst{{x}{f}}{{x}{Q}}}
+    }
+  \\
+& \inference
+  { \jterm{\Gamma}{A}{x}
+    \jterm{{\Gamma}{A}}{P}{f}
+    \jfam{{{{\Gamma}{A}}{P}}{Q}}{R}
     }
@@
     }
-  \label{comp-ss-f}\\
-& \inference
-  { \jterm{{{{\Gamma}{A}}{P}}{Q}}{R}{h}
+  \\
+& \inference
+  { \jterm{\Gamma}{A}{x}
+    \jterm{{\Gamma}{A}}{P}{f}
+    \jterm{{{{\Gamma}{A}}{P}}{Q}}{R}{h}
     }
@@
     }
-  \label{comp-ss-t}
-\end{align}
+\end{align*}
+Moreover, we get
+\begin{equation*}
+\inference
+  { \jterm{\Gamma}{A}{x}
+    \jterm{{\Gamma}{A}}{P}{f}
+    \jterm{{{\Gamma}{A}}{P}}{Q}{g}
+    }
+  { \jtermeq
+      {\Gamma}
+      {\subst{x}{{f}{Q}}}
+      {\subst{x}{{f}{g}}}
+      {\subst{{x}{f}}{{x}{g}}}
+    }
+\end{equation*}
+\end{rmk}
 
@@
 \label{comp-ws}
-We already have rules for the compatibility of substitution with weakening, but
-we still need the rules the other way around, asserting that there is a 
-judgmental equality $\ctxwk{A}{\subst{y}{Q}}\jdeq\subst{\ctxwk{A}{y}}{\ctxwk{A}{Q}}$
-together with all its variants.
-
-In the following inference rules we assume that $\jfam{\Gamma}{A}$ and that
-$\jterm{\Gamma}{B}{y}$.
-\begin{align}
-& \inference
-  { \jfam{{{\Gamma}{B}}{Q}}{R}
+The following inference rules explain what happens when we first weaken by a
+term $g$ of $Q$ in context $\ctxext{\Gamma}{B}$ and then weaken by a family
+$A$ over $\Gamma$.
+\begin{align}
+& \inference
+  { \jfam{\Gamma}{A}
+    \jterm{{\Gamma}{B}}{Q}{g}
+    \jfam{{{\Gamma}{B}}{Q}}{R}
+    }
+  { \jfameq
+      {{{\Gamma}{A}}{\ctxwk{A}{B}}}
+      {\ctxwk{A}{\subst{g}{R}}}
+      {\subst{\ctxwk{A}{g}}{\ctxwk{A}{R}}}
+    }
+  \label{comp-ws-c}
+  \\
+& \inference
+  { \jfam{\Gamma}{A}
+    \jterm{{\Gamma}{B}}{Q}{g}
+    \jfam{{{{\Gamma}{B}}{Q}}{R}}{S}
+    }
+  { \jfameq
+      {{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{\subst{g}{R}}}}
+      {\ctxwk{A}{\subst{g}{S}}}
+      {\subst{\ctxwk{A}{g}}{\ctxwk{A}{S}}}
+    }
+  \label{comp-ws-f}
+  \\
+& \inference
+  { \jfam{\Gamma}{A}
+    \jterm{{\Gamma}{B}}{Q}{g}
+    \jterm{{{{\Gamma}{B}}{Q}}{R}}{S}{k}
+    }
+  { \jtermeq
+      {{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{\subst{g}{R}}}}
+      {\ctxwk{A}{\subst{g}{S}}}
+      {\ctxwk{A}{\subst{g}{k}}}
+      {\subst{\ctxwk{A}{g}}{\ctxwk{A}{k}}}
+    }
+  \label{comp-ws-t}
+\end{align}
+
+\begin{rmk}
+As an important special case of these inference rules we have the following 
+valid inference rules:
+\begin{align*}
+& \inference
+  { \jfam{\Gamma}{A}
+    \jterm{\Gamma}{B}{y}
+    \jfam{{\Gamma}{B}}{Q}
+    }
+  { \jfameq
+      {{\Gamma}{A}}
+      {\ctxwk{A}{\subst{y}{Q}}}
+      {\subst{\ctxwk{A}{y}}{\ctxwk{A}{Q}}}
+    }
+  \\
+& \inference
+  { \jfam{\Gamma}{A}
+    \jterm{\Gamma}{B}{y}
+    \jfam{{{\Gamma}{B}}{Q}}{R}
     }
@@
     }
-  \label{comp-ws-f}\\
-& \inference
-  { \jterm{{{\Gamma}{B}}{Q}}{R}{h}
+  \\
+& \inference
+  { \jfam{\Gamma}{A}
+    \jterm{\Gamma}{B}{y}
+    \jterm{{{\Gamma}{B}}{Q}}{R}{h}
     }
@@
     }
-  \label{comp-ws-t}
-\end{align}
+\end{align*}
+Moreover, we get
+\begin{equation*}
+\inference
+  { \jfam{\Gamma}{A}
+    \jterm{\Gamma}{B}{y}
+    \jterm{{\Gamma}{B}}{Q}{g}
+    }
+  { \jtermeq
+      {{\Gamma}{A}}
+      {\ctxwk{A}{\subst{y}{Q}}}
+      {\ctxwk{A}{\subst{y}{g}}}
+      {\subst{\ctxwk{A}{y}}{\ctxwk{A}{g}}}
+    }
+\end{equation*}
+\end{rmk}
 
@@
   { \jfam{\Gamma}{A}
-    \jfam{\Gamma}{B}
+    \jfam{{\Gamma}{B}}{Q}
     }
   { \unfoldall{\jhomeq
+      {{{\Gamma}{A}}{\ctxwk{A}{B}}}
+      {\ctxwk{A}{Q}}
+      {\ctxwk{A}{Q}}
+      {\ctxwk{A}{\idtm{Q}}}
+      {\idtm{\ctxwk{A}{Q}}}
+      }
+    }
+  \label{comp-wi-t}
+\end{equation}
+
+\begin{rmk}
+An important special case is the judgmental equality
+\begin{equation*}
+\unfoldall{\jhomeq
       {{\Gamma}{A}}
@@
       {\idtm{\ctxwk{A}{B}}}
-    }}
-  \label{comp-wi-t}
-\end{equation}
+      }
+\end{equation*}
+for a family $\jfam{\Gamma}{B}$.
+\end{rmk}
 
@@
   { \jterm{\Gamma}{A}{x}
-    \jfam{{\Gamma}{A}}{P}
+    \jfam{{{\Gamma}{A}}{P}}{Q}
     }
   { \unfoldall{\jhomeq
+      {{\Gamma}{\subst{x}{P}}}
+      {\subst{x}{Q}}
+      {\subst{x}{Q}}
+      {\subst{x}{\idtm{Q}}}
+      {\idtm{\subst{x}{Q}}}
+    }}
+  \label{comp-si-t}
+\end{equation}
+
+\begin{rmk}
+An important special case is the judgmental equality
+\begin{equation*}
+\unfoldall{\jhomeq
       {\Gamma}
@@
       {\idtm{\subst{x}{P}}}
-    }}
-  \label{comp-si-t}
-\end{equation}
+    }
+\end{equation*}
+for a family $\jfam{{\Gamma}{A}}{P}$.
+\end{rmk}
 
@@
 
-\begin{comment}
-\subsubsection{Substituting by the term of the empty family}
-\label{comp-0s}
-The following rules assert that substituting by the term $\jterm{\Gamma}{\emptyf}{\emptytm}$
-leaves everything unchanged.
-\begin{align}
-& \inference
-  { \jfam{\Gamma}{A}
-    }
-  { \jfameq{\Gamma}{\subst{\emptytm}{A}}{A}
-    }
-  \label{comp-0s-c}
-  \\
-%& \inference
-%  { \jterm{\Gamma}{A}{x}
-%    }
-%  { \jtermeq{\Gamma}{A}{\subst{\emptytm}{x}}{x}
-%    }
-%  \\
-& \inference
-  { \jfam{{\Gamma}{A}}{P}
-    }
-  { \jfameq{{\Gamma}{A}}{\subst{\emptytm}{P}}{P}
-    }
-  \label{comp-0s-f}
-  \\
-& \inference
-  { \jterm{{\Gamma}{A}}{P}{f}
-    }
-  { \jtermeq{{\Gamma}{A}}{P}{\subst{\emptytm}{f}}{f}
-    }.
-  \label{comp-0s-t}
-\end{align}
-\end{comment}
-
 \subsubsection{Currying for weakening}
@@
 \begin{align}
+& \inference
+  { \jfam{\Gamma}{A}
+    \jfam{{\Gamma}{A}}{P}
+    \jfam{\Gamma}{B}
+    }
+  { \jfameq
+      {{{\Gamma}{A}}{P}}
+      {\ctxwk{\ctxext{A}{P}}{B}}
+      {\ctxwk{P}{{A}{B}}}
+    }
+  \label{comp-ew-c}\\
 & \inference
@@
   { \jfam{\Gamma}{A}
+    \jfam{\Gamma}{B}
+    \jterm{\Gamma}{A}{x}
+    }
+  { \jfameq
+      {\Gamma}
+      {\subst{x}{\ctxwk{A}{B}}}
+      {B}
+    }
+  \label{cancellation-ws-c}\\
+& \inference
+  { \jfam{\Gamma}{A}
     \jfam{{\Gamma}{B}}{Q}
@@
     }
-  \label{idfunc-wk-defn}\\
+  \label{precomp-idtm-c}
+  \\
+& \inference
+  { \jfam{{{\Gamma}{A}}{P}}{Q}
+    }
+  { \jfameq{{{\Gamma}{A}}{P}}{\subst{\idtm{A}}{\ctxwk{A}{Q}}}{Q}
+    }
+  \label{precomp-idtm-f}
+  \\
 %An unnecessary rule:
@@
 & \inference
-  { \jterm{{\Gamma}{A}}{P}{f}
+  { \jterm{{{\Gamma}{A}}{P}}{Q}{g}
     }
   { \jtermeq
-      {{\Gamma}{A}}
-      {P}
-      {\subst{\idtm{A}}{\ctxwk{A}{f}}}
-      {f}
-    }
-  \label{idfunc-precomp}\\
+      {{{\Gamma}{A}}{P}}
+      {Q}
+      {\subst{\idtm{A}}{\ctxwk{A}{g}}}
+      {g}
+    }
+  \label{precomp-idtm-t}
 %An unnecessary rule:
@@
 %    }
-%  \label{idfunc-precomp}\\
-& \inference
-  { \unfoldall{\jhom{\Gamma}{A}{B}{f}}
-    }
-  { \unfoldall{\jhomeq{\Gamma}{A}{B}{\jcomp{A}{f}{\idtm{B}}}{f}}
-    }
-  \label{idfunc-postcomp}
+%  \label{precomp-idtm-t}
 \end{align}

### 2014-06-19 article-models-ttconstr.tex
writing rules for the dependent function constructor

@@
 
-\subsection{Dependent function types}
-The dependent function type constructor bundles up the terms of a family $P$
-of contexts over $\ctxext{\Gamma}{A}$ and produces a family $\mprd{A}{P}$ of
-contexts over $\Gamma$. Its terms are copies of those of $P$. The operation
-$\mprd{A}{\blank}$ also acts on terms of families -- this is called
-$\lambda$-abstraction.
-
-The dependent function type constructor not only acts on families $P$ of
-contexts over $\ctxext{\Gamma}{A}$, but it also sends families $Q$ of contexts
-over $\ctxext{{\Gamma}{A}}{P}$ to families $\mprd{A}{Q}$ over $\ctxext{\Gamma}
-{\mprd{A}{P}}$. Moreover, we will also have a version of $\lambda$-abstraction
-for terms of such families $Q$.
-
-\begin{align*}
-& \inference
-    { \jfam{{\Gamma}{A}}{P}
-      }
-    { \jfam{\Gamma}{\mprd{A}{P}}
-      }
-& & \inference
-      { \jfameq{\Gamma}{A}{A'}
-        \jfameq{{\Gamma}{A}}{P}{P'}
-        }
-      { \jfameq{\Gamma}{\mprd{A}{P}}{\mprd{A'}{P'}}
-        }
-  \\
-& \inference
-    { \jterm{{\Gamma}{A}}{P}{f}
-      }
-    { \jterm{\Gamma}{\mprd{A}{P}}{\slam{A}{P}{f}}
-      }
-& & \inference
-      { \jtermeq{{\Gamma}{A}}{P}{f}{f'}
-        }
-      { \jtermeq{\Gamma}{\mprd{A}{P}}{\slam{A}{P}{f}}{\slam{A}{P}{f'}}
-        }
-  \\
-& \inference
-    { \jterm{\Gamma}{\mprd{A}{P}}{g}
-      }
-    { \jterm{{\Gamma}{A}}{P}{\sev{g}}
-      }
-& & \inference
-      { \jtermeq{\Gamma}{\mprd{A}{P}}{g}{g'}
-        }
-      { \jtermeq{{\Gamma}{A}}{P}{\sev{g}}{\sev{g'}}
-        }
-\end{align*}
-
-\begin{defn}
-Let $A$ and $B$ be families of contexts over $\Gamma$. Then we define
-\begin{equation*}
-\jfamdefn{\Gamma}{\jfun{A}{B}}{\sprd{A}{\ctxwk{A}{B}}}.
-\end{equation*}
-\end{defn}
-
-\subsubsection{Dependent products are compatible with the empty context}
-The empty context can appear in the domain and in the codomain of the dependent
-function type constructor. We have the following inference rules explaining
-what happens when the empty context appears in the domain:
-\begin{align}
-& \inference
-    { \jfam{\Gamma}{P}
-      }
-    { \jfameq{\Gamma}{\mprd{\emptyf}{P}}{P}
-      }
-  \\
-& \inference
-    { \jterm{\Gamma}{P}{f}
-      }
-    { \jtermeq{\Gamma}{P}{\slam{A}{P}{f}}{f}
-      }
-  \\
-& \inference
-    { \jterm{\Gamma}{P}{g}
-      }
-    { \jtermeq{\Gamma}{P}{\sev{g}}{g}
-      }
-\end{align}
-We have the following infernece rules explaining what happens when the empty
-context appears in the codomain:
-\begin{align}
-& \inference
-    { \jfam{\Gamma}{A}
-      }
-    { \jfameq{\Gamma}{\mprd{A}{\emptyf}}{\emptyf}
-      }
-\end{align}
-
-\subsubsection{Dependent products are compatible with extension}
-\begin{align}
-& \inference
-  { \jfam{{{{\Gamma}{A}}{B}}{P}}{Q}
-    }
-  { \jfameq
-      {\Gamma}
-      {\sprd{{A}{B}}{{P}{Q}}}
-      {\ctxext{\sprd{A}{\sprd{B}{P}}}{\sprd{A}{\sprd{B}{Q}}}}
-    }
-  \\
-& \inference
-  { \jterm{{{\Gamma}{A}}{B}}{P}{f}
-    }
-  { \jtermeq
-      {\Gamma}
-      {\sprd{A}{\sprd{B}{P}}}
-      {\slam{{A}{B}}{P}{f}}
-      {\slam{A}{\sprd{B}{P}}{\slam{B}{P}{f}}}
-    }
-\end{align}
-\subsubsection{Dependent products are compatible with weakening}
-\subsubsection{Dependent products are compatible with substitution}?
-\subsubsection{Dependent products are compatible with the identity functions}?
-
 \subsection{$\wtypesym$-types}

### 2014-06-19 article-models-ttdp.tex
rules for the dependent function constructor

\section{The dependent function constructor}
As we have done with most everything so far, we will take the point of view that
the dependent function constructor is an operation with an action on contexts,
on families and on terms. When $A$ is a family of contexts over $\Gamma$,
$\mprd{A}{\blank}$ takes things in context $\ctxext{\Gamma}{A}$ to the context
$\Gamma$. It's action on terms restricted to the empty context is usually 
denoted by $\lambda$, so we shall denote the entire action by $\lambda$. As
usual, $\lambda$ is reversed by evaluation.

The dependent function type constructor not only acts on families $P$ of
contexts over $\ctxext{\Gamma}{A}$, but it also sends families $Q$ of contexts
over $\ctxext{{\Gamma}{A}}{P}$ to families $\mprd[\famsym]{A}{Q}$ over $\ctxext{\Gamma}
{\mprd{A}{P}}$. Moreover, we will also have a version of $\lambda$-abstraction
for terms of such families $Q$.

\begin{align*}
& \inference
    { \jfam{{\Gamma}{A}}{P}
      }
    { \jfam{\Gamma}{\mprd{A}{P}}
      }
& & \inference
      { \jfameq{\Gamma}{A}{A'}
        \jfameq{{\Gamma}{A}}{P}{P'}
        }
      { \jfameq{\Gamma}{\mprd{A}{P}}{\mprd{A'}{P'}}
        }
  \\
& \inference
    { \jfam{{{\Gamma}{A}}{P}}{Q}
      }
    { \jfam{{\Gamma}{\mprd{A}{P}}}{\mprd[\famsym]{A}{Q}}
      }
& & \inference
      { \jfameq{\Gamma}{A}{A'}
        \jfameq{{{\Gamma}{A}}{P}}{Q}{Q'}
        }
      { \jfameq{{\Gamma}{\mprd{A}{P}}}{\mprd[\famsym]{A}{Q}}{\mprd[\famsym]{A'}{Q'}}
        }
  \\
& \inference
    { \jterm{{{\Gamma}{A}}{P}}{Q}{g}
      }
    { \jterm{{\Gamma}{\mprd{A}{P}}}{\mprd[\famsym]{A}{Q}}{\slam{A}{Q}{g}}
      }
& & \inference
      { \jfameq{\Gamma}{A}{A'}
        \jtermeq{{{\Gamma}{A}}{P}}{Q}{g}{g'}
        }
      { \jtermeq{{\Gamma}{\mprd{A}{P}}}{\mprd[\famsym]{A}{Q}}{\slam{A}{Q}{g}}{\slam{A'}{Q}{g'}}
        }
  \\
& \inference
    { \jterm{{\Gamma}{\mprd{A}{P}}}{\mprd[\famsym]{A}{Q}}{g}
      }
    { \jterm{{{\Gamma}{A}}{P}}{Q}{\sev{A}{Q}{g}}
      }
& & \inference
      { \jfameq{\Gamma}{A}{A'}
        \jtermeq{{\Gamma}{\mprd{A}{P}}}{\mprd[\famsym]{A}{Q}}{g}{g'}
        }
      { \jtermeq{{{\Gamma}{A}}{P}}{Q}{\sev{A}{Q}{g}}{\sev{A'}{Q}{g'}}
        }
\end{align*}

\begin{defn}
Let $A$ and $B$ be families of contexts over $\Gamma$. Then we define
\begin{equation*}
\jfamdefn{\Gamma}{\jfun{A}{B}}{\sprd{A}{\ctxwk{A}{B}}}.
\end{equation*}
\end{defn}

\subsection{The compatibility rules for the dependent product constructor}
In this subsection we lay out the compatibility rules which we will require
for the dependent product constructor.

\subsubsection{Dependent products are compatible with the empty context}
The empty context can appear in the domain and in the codomain of the dependent
function type constructor. We have the following inference rules explaining
what happens when the empty context appears in the domain:
\begin{align}
& \inference
    { \jfam{\Gamma}{P}
      }
    { \jfameq{\Gamma}{\mprd{\emptyf}{P}}{P}
      }
  \\
& \inference
    { \jfam{{\Gamma}{P}}{Q}
      }
    { \jfameq{{\Gamma}{P}}{\mprd[\famsym]{\emptyf}{Q}}{Q}
      }
  \\
& \inference
    { \jterm{{\Gamma}{P}}{Q}{g}
      }
    { \jtermeq{{\Gamma}{P}}{Q}{\slam{\emptyf}{Q}{g}}{g}
      }
  \\
& \inference
    { \jterm{{\Gamma}{P}}{Q}{g}
      }
    { \jtermeq{{\Gamma}{P}}{Q}{\sev{\emptyf}{Q}{g}}{g}
      }
\end{align}
We have the following infernece rules explaining what happens when the empty
context appears in the codomain:
\begin{align}
& \inference
    { \jfam{\Gamma}{A}
      }
    { \jfameq{\Gamma}{\mprd{A}{\emptyf}}{\emptyf}
      }
  \\
& \inference
    { \jfam{{\Gamma}{A}}{P}
      }
    { \jfameq{{\Gamma}{\mprd{A}{P}}}{\mprd[\famsym]{A}{\emptyf}}{\emptyf}
      }
\end{align}
Now we see that we can compare $\mprd{A}{P}$ with $\mprd[\famsym]{A}{P}$ by
seeing $P$ as a family over $\ctxext{{\Gamma}{A}}{\emptyf}$. We will require
the following inference rule to be valid:
\begin{equation}
\inference
  { \jfam{{\Gamma}{A}}{P}
    }
  { \jfameq{\Gamma}{\mprd{A}{P}}{\mprd[\famsym]{A}{P}}
    }
\end{equation}
This rule will allow us to omit the annotation indicating the action on families,
which we will do from now on. We also note that because the empty family over
$\ctxext{\Gamma}{A}$ is mapped to the empty family over $\Gamma$ by the
dependent function constructor, we obtain the important special cases of
lambda abstraction and evaluation that
\begin{align*}
& \inference
  { \jterm{{\Gamma}{A}}{P}{f}
    }
  { \jterm{\Gamma}{\mprd{A}{P}}{\slam{A}{P}{f}}
    }
  \\
& \inference
  { \jterm{\Gamma}{\mprd{A}{P}}{f}
    }
  { \jterm{{\Gamma}{A}}{P}{\sev{A}{P}{f}}
    }
\end{align*}
Thus, we retrieve what is originally meant by lambda abstraction.

\subsubsection{Dependent products are compatible with extension}
\begin{align}
& \inference
  { \jfam{{{{\Gamma}{A}}{B}}{P}}{Q}
    }
  { \jfameq
      {\Gamma}
      {\sprd{{A}{B}}{{P}{Q}}}
      {\ctxext{\sprd{A}{\sprd{B}{P}}}{\sprd{A}{\sprd{B}{Q}}}}
    }
  \\
& \inference
  { \jterm{{{\Gamma}{A}}{B}}{P}{f}
    }
  { \jtermeq
      {\Gamma}
      {\sprd{A}{\sprd{B}{P}}}
      {\slam{{A}{B}}{P}{f}}
      {\slam{A}{\sprd{B}{P}}{\slam{B}{P}{f}}}
    }
\end{align}

\subsubsection{Weakening is compatible with the dependent product constructor}

\begin{align*}
& \inference
  { \jfam{\Gamma}{A}
    \jfam{{{\Gamma}{B}}{Q}}{R}
    }
  { \jfameq
      {{{\Gamma}{A}}{\ctxwk{A}{B}}}
      {\ctxwk{A}{\mprd{Q}{R}}}
      {\mprd{\ctxwk{A}{Q}}{\ctxwk{A}{R}}}
    }
  \\
& \inference
  { \jfam{\Gamma}{A}
    \jfam{{{{\Gamma}{B}}{Q}}{R}}{S}
    }
  { \jfameq
      {{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{\mprd{Q}{R}}}}
      {\ctxwk{A}{\mprd{Q}{S}}}
      {\mprd{\ctxwk{A}{Q}}{\ctxwk{A}{S}}}
    }
  \\
& \inference
  { \jfam{\Gamma}{A}
    \jterm{{{{\Gamma}{B}}{Q}}{R}}{S}{k}
    }
  { \jtermeq
      {{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{\mprd{Q}{R}}}}
      {\ctxwk{A}{\mprd{Q}{S}}}
      {\ctxwk{A}{\slam{Q}{S}{k}}}
      {\slam{\ctxwk{A}{Q}}{\ctxwk{A}{S}}{\ctxwk{A}{k}}}
    }
  \\
& \inference
  { \jfam{\Gamma}{A}
    \jterm{{{\Gamma}{B}}{\mprd{Q}{R}}}{\mprd{Q}{S}}{k}
    }
  { \jtermeq
      {{{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{Q}}}{\ctxwk{A}{R}}}
      {\ctxwk{A}{S}}
      {\ctxwk{A}{\sev{Q}{S}{k}}}
      {\sev{\ctxwk{A}{Q}}{\ctxwk{A}{S}}{\ctxwk{A}{k}}}
    }
\end{align*}

\subsection{Dependent products are taken fiberwise}
The rules explaining that substitution is compatible with the dependent product
constructor assert that dependent products are taken fiberwise (as is usual).

\begin{align*}
& \inference
  { \jterm{\Gamma}{A}{x}
    \jfam{{{{\Gamma}{A}}{P}}{Q}}{R}
    }
  { \jfameq
      {{\Gamma}{\subst{x}{P}}}
      {\subst{x}{\mprd{Q}{R}}}
      {\mprd{\subst{x}{Q}}{\subst{x}{R}}}
    }
  \\
& \inference
  { \jterm{\Gamma}{A}{x}
    \jfam{{{{{\Gamma}{A}}{P}}{Q}}{R}}{S}
    }
  { \jfameq
      {{{\Gamma}{\subst{x}{P}}}{\mprd{\subst{x}{Q}}{\subst{x}{R}}}}
      {\subst{x}{\mprd{Q}{S}}}
      {\mprd{\subst{x}{Q}}{\subst{x}{S}}}
    }
  \\
& \inference
  { \jterm{\Gamma}{A}{x}
    \jterm{{{{{\Gamma}{A}}{P}}{Q}}{R}}{S}{k}
    }
  { \jtermeq
      {{{\Gamma}{\subst{x}{P}}}{\mprd{\subst{x}{Q}}{\subst{x}{R}}}}
      {\subst{x}{\mprd{Q}{S}}}
      {\subst{x}{\slam{Q}{S}{k}}}
      {\slam{\subst{x}{Q}}{\subst{x}{S}}{\subst{x}{k}}}
    }
  \\
& \inference
  { \jterm{\Gamma}{A}{x}
    \jterm{{{{\Gamma}{A}}{P}}{\mprd{Q}{R}}}{\mprd{Q}{S}}{k}
    }
  { \jtermeq
      {{{{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}}{\subst{x}{R}}}
      {\subst{x}{S}}
      {\subst{x}{\sev{Q}{S}{k}}}
      {\sev{\subst{x}{Q}}{\subst{x}{S}}{\subst{x}{k}}}
    } 
\end{align*}

\subsubsection{Dependent products are compatible with the identity functions}?

### 2014-06-19 article-models-ttdp.tex
small improvs

@@
 
-\subsection{Dependent products are taken fiberwise}
+\subsubsection{Dependent products are taken fiberwise}
 The rules explaining that substitution is compatible with the dependent product
@@
 
-\subsubsection{Dependent products are compatible with the identity functions}?
+\subsubsection{Dependent products are compatible with the identity functions}

### 2014-06-20 article-models-ttdp.tex
The first version of the set of rules for the dependent function constructor is finished

@@
 \subsubsection{Dependent products are compatible with extension}
-\begin{align}
+The following rules describe what happens when a dependent function constructor
+is taken over an extension
+
+\begin{align*}
 & \inference
-  { \jfam{{{{\Gamma}{A}}{B}}{P}}{Q}
+  { \jfam{{\Gamma}{A}}{P}
+    \jfam{{{\Gamma}{A}}{P}}{Q}
     }
@@
       {\Gamma}
-      {\sprd{{A}{B}}{{P}{Q}}}
-      {\ctxext{\sprd{A}{\sprd{B}{P}}}{\sprd{A}{\sprd{B}{Q}}}}
+      {\mprd{\ctxext{A}{P}}{Q}}
+      {\mprd{A}{\mprd{P}{Q}}}
     }
@@
 & \inference
-  { \jterm{{{\Gamma}{A}}{B}}{P}{f}
+  { \jfam{{\Gamma}{A}}{P}
+    \jfam{{{{\Gamma}{A}}{P}}{Q}}{R}
+    }
+  { \jfameq
+      {{\Gamma}{\mprd{\ctxext{A}{P}}{Q}}}
+      {\mprd{\ctxext{A}{P}}{R}}
+      {\mprd{A}{\mprd{P}{R}}}
+    }
+  \\
+& \inference
+  { \jfam{{\Gamma}{A}}{P}
+    \jterm{{{{\Gamma}{A}}{P}}{Q}}{R}{h}
     }
   { \jtermeq
+      {{\Gamma}{\mprd{\ctxext{A}{P}}{Q}}}
+      {\mprd{\ctxext{A}{P}}{R}}
+      {\slam{\ctxext{A}{P}}{R}{h}}
+      {\slam{A}{\mprd{P}{R}}{\slam{P}{R}{h}}}
+    }
+  \\
+& \inference
+  { \jfam{{\Gamma}{A}}{P}
+    \jterm{{\Gamma}{\mprd{\ctxext{A}{P}}{Q}}}{\mprd{\ctxext{A}{P}}{R}}{h}
+    }
+  { \jtermeq
+      {{{\Gamma}{{A}{P}}}{Q}}
+      {R}
+      {\sev{\ctxext{A}{P}}{R}{h}}
+      {\sev{P}{R}{\sev{A}{\mprd{P}{R}}{h}}}
+    }
+\end{align*}
+
+When the dependent product constructor is applied to an extension, we get the
+following:
+
+\begin{align*}
+& \inference
+  { \jfam{{{\Gamma}{A}}{P}}{Q}
+    }
+  { \jfameq
       {\Gamma}
-      {\sprd{A}{\sprd{B}{P}}}
-      {\slam{{A}{B}}{P}{f}}
-      {\slam{A}{\sprd{B}{P}}{\slam{B}{P}{f}}}
+      {\mprd{A}{\ctxext{P}{Q}}}
+      {\ctxext{\mprd{A}{P}}{\mprd{A}{Q}}}
     }
-\end{align}
+  \\
+& \inference
+  { \jfam{{{{\Gamma}{A}}{P}}{Q}}{R}
+    }
+  { \jfameq
+      {{\Gamma}{\mprd{A}{P}}}
+      {\mprd{A}{\ctxext{Q}{R}}}
+      {\ctxext{\mprd{A}{Q}}{\mprd{A}{R}}}
+    }
+\end{align*}
+We show in \autoref{lem:slam-ext} that the remaining two properties
+involving lambda abstraction and evaluation are derivable.
 
@@
 
-\subsubsection{Dependent products are compatible with the identity functions}
+\subsubsection{The dependent function constructor is compatible with weakening}
+\begin{align*}
+& \inference
+  { \jfam{{{\Gamma}{A}}{P}}{Q}
+    \jfam{{{\Gamma}{Q}}{P}}{R}
+    }
+  { \jfameq
+      {{{\Gamma}{\mprd{A}{P}}}{\mprd{A}{Q}}}
+      {\mprd{A}{\ctxwk{Q}{R}}}
+      {\ctxwk{\mprd{A}{Q}}{\mprd{A}{R}}}
+    }
+  \\
+& \inference
+  { \jfam{{{\Gamma}{A}}{P}}{Q}
+    \jfam{{{{\Gamma}{P}}{Q}}{R}}{S}
+    }
+  { \jfameq
+      {{{{\Gamma}{\mprd{A}{P}}}{\mprd{A}{Q}}}{\mprd{A}{\ctxwk{Q}{R}}}}
+      {\mprd{A}{\ctxwk{Q}{S}}}
+      {\ctxwk{\mprd{A}{Q}}{\mprd{A}{S}}}
+    }
+  \\
+& \inference
+  { \jfam{{{\Gamma}{A}}{P}}{Q}
+    \jterm{{{{\Gamma}{P}}{Q}}{R}}{S}{k}
+    }
+  { \jtermeq
+      {{{{\Gamma}{\mprd{A}{P}}}{\mprd{A}{Q}}}{\mprd{A}{\ctxwk{Q}{R}}}}
+      {\mprd{A}{\ctxwk{Q}{S}}}
+      {\slam{A}{\ctxwk{Q}{S}}{\ctxwk{Q}{k}}}
+      {\ctxwk{\mprd{A}{Q}}{\slam{A}{S}{k}}}
+    }
+  \\
+& \inference
+  { \jfam{{{\Gamma}{A}}{P}}{Q}
+    \jfam{{{{\Gamma}{P}}{Q}}{R}}{S}
+    \jterm{{{\Gamma}{\mprd{A}{P}}}{\mprd{A}{R}}}{\mprd{A}{S}}{k}
+    }
+  { \jtermeq
+      {{{{{\Gamma}{A}}{P}}{Q}}{\ctxwk{Q}{R}}}
+      {\ctxwk{Q}{S}}
+      {\sev{A}{\ctxwk{Q}{S}}{\ctxwk{\mprd{A}{Q}}{k}}}
+      {\ctxwk{Q}{\sev{A}{S}{k}}}
+    }
+\end{align*}
+
+\subsubsection{The dependent function constructor is compatible with substitution}
+\begin{align*}
+& \inference
+  { \jterm{{{\Gamma}{A}}{P}}{Q}{g}
+    \jfam{{{{\Gamma}{A}}{P}}{Q}}{R}
+    }
+  { \jfameq
+      {{\Gamma}{\mprd{A}{P}}}
+      {\mprd{A}{\subst{g}{R}}}
+      {\subst{\slam{A}{Q}{g}}{\mprd{A}{R}}}
+    }
+  \\
+& \inference
+  { \jterm{{{\Gamma}{A}}{P}}{Q}{g}
+    \jfam{{{{{\Gamma}{A}}{P}}{Q}}{R}}{S}
+    }
+  { \jfameq
+      {{{\Gamma}{\mprd{A}{P}}}{\mprd{A}{\subst{g}{R}}}}
+      {\mprd{A}{\subst{g}{S}}}
+      {\subst{\slam{A}{Q}{g}}{\mprd{A}{S}}}
+    }
+  \\
+& \inference
+  { \jterm{{{\Gamma}{A}}{P}}{Q}{g}
+    \jterm{{{{{\Gamma}{A}}{P}}{Q}}{R}}{S}{k}
+    }
+  { \jtermeq
+      {{{\Gamma}{\mprd{A}{P}}}{\mprd{A}{\subst{g}{R}}}}
+      {\mprd{A}{\subst{g}{S}}}
+      {\slam{A}{\subst{g}{S}}{\subst{g}{k}}}
+      {\subst{\slam{A}{Q}{g}}{\slam{A}{S}{k}}}
+    }
+  \\
+& \inference
+  { \jterm{{{\Gamma}{A}}{P}}{Q}{g}
+    \jterm{{\Gamma}{\mprd{A}{\ctxext{{P}{Q}}{R}}}}{\mprd{A}{S}}{k}
+    }
+  { \jtermeq
+      {{{{\Gamma}{A}}{P}}{\subst{g}{R}}}
+      {\subst{g}{S}}
+      {\sev{A}{\subst{g}{S}}{\subst{\slam{A}{Q}{g}}{k}}}
+      {\subst{g}{\sev{A}{S}{k}}}
+    }
+\end{align*}
+
+\subsubsection{The dependent function constructor is compatible with the identity terms}
+\begin{align*}
+& \inference
+  { \jfam{{{\Gamma}{A}}{P}}{Q}
+    }
+  { \jtermeq
+      {{{\Gamma}{\mprd{A}{P}}}{\mprd{A}{Q}}}
+      {\ctxwk{\mprd{A}{Q}}{\mprd{A}{Q}}}
+      {\slam{A}{\ctxwk{Q}{Q}}{\idtm{Q}}}
+      {\idtm{\mprd{A}{Q}}}
+    }
+  \\
+& \inference
+  { \jfam{{{\Gamma}{A}}{P}}{Q}
+    }
+  { \jtermeq
+      {{{{\Gamma}{A}}{P}}{Q}}
+      {\ctxwk{Q}{Q}}
+      {\sev{A}{\ctxwk{Q}{Q}}{\idtm{\mprd{A}{Q}}}}
+      {\idtm{Q}}
+    }
+\end{align*}
+
+\subsection{Consequences of the rules for the dependent function constructors}
+
+\begin{lem}\label{lem:slam-ext}
+The inference rules asserting that lambda abstraction is compatible with term
+extension
+\begin{align*}
+& \inference
+  { \jterm{{{\Gamma}{A}}{P}}{{Q}{R}}{h}
+    }
+  { \jtermeq
+      {{\Gamma}{\mprd{A}{P}}}
+      {\mprd{A}{\ctxext{Q}{R}}}
+      {\cprojfst{\mprd{A}{Q}}{\mprd{A}{R}}{\slam{A}{\ctxext{Q}{R}}{h}}}
+      {\slam{A}{Q}{\cprojfst{Q}{R}{h}}}
+    }
+  \\
+& \inference
+  { \jterm{{{\Gamma}{A}}{P}}{{Q}{R}}{h}
+    }
+  { \jtermeq
+      {{\Gamma}{\mprd{A}{P}}}
+      {\mprd{A}{\ctxext{Q}{R}}}
+      {\cprojfst{\mprd{A}{Q}}{\mprd{A}{R}}{\slam{A}{\ctxext{Q}{R}}{h}}}
+      {\slam{A}{R}{\cprojsnd{Q}{R}{h}}}
+    }
+\intertext{and the inference rules asserting that evaluation is compatible with
+term extension}
+& \inference
+  { \jterm{{\Gamma}{\mprd{A}{P}}}{\mprd{A}{\ctxext{Q}{R}}}{h}
+    }
+  { \jtermeq
+      {{{\Gamma}{A}}{P}}
+      {{Q}{R}}
+      {\cprojfst{Q}{R}{\sev{A}{\ctxext{Q}{R}}{h}}}
+      {\sev{A}{Q}{\cprojfst{\mprd{A}{Q}}{\mprd{A}{R}}{h}}}
+    }
+  \\
+& \inference
+  { \jterm{{\Gamma}{\mprd{A}{P}}}{\mprd{A}{\ctxext{Q}{R}}}{h}
+    }
+  { \jtermeq
+      {{{\Gamma}{A}}{P}}
+      {{Q}{R}}
+      {\cprojsnd{Q}{R}{\sev{A}{\ctxext{Q}{R}}{h}}}
+      {\sev{A}{R}{\cprojsnd{\mprd{A}{Q}}{\mprd{A}{R}}{h}}}
+    }
+\end{align*}
+are valid.
+\end{lem}
+
+\begin{proof}
+We have the judgmental equalities
+\begin{align*}
+\cprojfst{\mprd{A}{Q}}{\mprd{A}{R}}{\slam{A}{\ctxext{Q}{R}}{h}}
+& \jdeq
+  \subst{\slam{A}{\ctxext{Q}{R}}{h}}{\ctxwk{\mprd{A}{R}}{\idtm{\mprd{A}{Q}}}}
+  \\
+& \jdeq 
+  \subst{\slam{A}{\ctxext{Q}{R}}{h}}{\ctxwk{\mprd{A}{R}}{\slam{A}{\ctxwk{Q}{Q}}{\idtm{Q}}}}
+  \\
+& \jdeq
+  \subst{\slam{A}{\ctxext{Q}{R}}{h}}{\slam{A}{\ctxwk{R}{{Q}{Q}}}{\ctxwk{R}{\idtm{Q}}}}
+  \\
+& \jdeq
+  \subst{\slam{A}{\ctxext{Q}{R}}{h}}{\slam{A}{\ctxwk{\ctxext{Q}{R}}{Q}}{\ctxwk{R}{\idtm{Q}}}}
+  \\
+& \jdeq
+  \slam{A}{Q}{\subst{h}{\ctxwk{R}{\idtm{Q}}}}
+  \\
+& \jdeq
+  \slam{A}{Q}{\cprojfst{Q}{R}{h}}
+\end{align*}
+\end{proof}
+
+\begin{rmk}
+It follows from these rules that on the projections (i.e.~on the fibrations),
+lambda abstraction and evaluation are each other's inverses even though we
+have not even postulated that evaluation is a retraction of lambda 
+abstraction. Thus, in combination with univalence this should give that
+they are homotopically inverse to each other.
+\end{rmk}

### 2014-06-22 article-models-ttdp.tex
revising the rules for evaluation

@@
         }
-  \\
-& \inference
-    { \jterm{{\Gamma}{\mprd{A}{P}}}{\mprd[\famsym]{A}{Q}}{g}
-      }
-    { \jterm{{{\Gamma}{A}}{P}}{Q}{\sev{A}{Q}{g}}
-      }
-& & \inference
-      { \jfameq{\Gamma}{A}{A'}
-        \jtermeq{{\Gamma}{\mprd{A}{P}}}{\mprd[\famsym]{A}{Q}}{g}{g'}
-        }
-      { \jtermeq{{{\Gamma}{A}}{P}}{Q}{\sev{A}{Q}{g}}{\sev{A'}{Q}{g'}}
-        }
 \end{align*}
@@
     { \jtermeq{{\Gamma}{P}}{Q}{\slam{\emptyf}{Q}{g}}{g}
-      }
-  \\
-& \inference
-    { \jterm{{\Gamma}{P}}{Q}{g}
-      }
-    { \jtermeq{{\Gamma}{P}}{Q}{\sev{\emptyf}{Q}{g}}{g}
       }
@@
     }
-  \\
-& \inference
-  { \jterm{\Gamma}{\mprd{A}{P}}{f}
-    }
-  { \jterm{{\Gamma}{A}}{P}{\sev{A}{P}{f}}
-    }
 \end{align*}
@@
       {\slam{A}{\mprd{P}{R}}{\slam{P}{R}{h}}}
-    }
-  \\
-& \inference
-  { \jfam{{\Gamma}{A}}{P}
-    \jterm{{\Gamma}{\mprd{\ctxext{A}{P}}{Q}}}{\mprd{\ctxext{A}{P}}{R}}{h}
-    }
-  { \jtermeq
-      {{{\Gamma}{{A}{P}}}{Q}}
-      {R}
-      {\sev{\ctxext{A}{P}}{R}{h}}
-      {\sev{P}{R}{\sev{A}{\mprd{P}{R}}{h}}}
     }
@@
     }
-  \\
-& \inference
-  { \jfam{\Gamma}{A}
-    \jterm{{{\Gamma}{B}}{\mprd{Q}{R}}}{\mprd{Q}{S}}{k}
-    }
-  { \jtermeq
-      {{{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{Q}}}{\ctxwk{A}{R}}}
-      {\ctxwk{A}{S}}
-      {\ctxwk{A}{\sev{Q}{S}{k}}}
-      {\sev{\ctxwk{A}{Q}}{\ctxwk{A}{S}}{\ctxwk{A}{k}}}
-    }
 \end{align*}
@@
     }
-  \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jterm{{{{\Gamma}{A}}{P}}{\mprd{Q}{R}}}{\mprd{Q}{S}}{k}
-    }
-  { \jtermeq
-      {{{{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}}{\subst{x}{R}}}
-      {\subst{x}{S}}
-      {\subst{x}{\sev{Q}{S}{k}}}
-      {\sev{\subst{x}{Q}}{\subst{x}{S}}{\subst{x}{k}}}
-    } 
 \end{align*}
@@
       {\ctxwk{\mprd{A}{Q}}{\slam{A}{S}{k}}}
-    }
-  \\
-& \inference
-  { \jfam{{{\Gamma}{A}}{P}}{Q}
-    \jfam{{{{\Gamma}{P}}{Q}}{R}}{S}
-    \jterm{{{\Gamma}{\mprd{A}{P}}}{\mprd{A}{R}}}{\mprd{A}{S}}{k}
-    }
-  { \jtermeq
-      {{{{{\Gamma}{A}}{P}}{Q}}{\ctxwk{Q}{R}}}
-      {\ctxwk{Q}{S}}
-      {\sev{A}{\ctxwk{Q}{S}}{\ctxwk{\mprd{A}{Q}}{k}}}
-      {\ctxwk{Q}{\sev{A}{S}{k}}}
     }
@@
 \end{rmk}
+
+\subsection{The evaluation term}
+Until now we have just introduced the dependent function constructor, but we 
+have not treated one of its main characteristic, the evaluation map. Usually
+in type theory, evaluation is introduces by a rule like
+\begin{equation*}
+\inference
+    { \jterm{\Gamma}{\mprd{A}{P}}{f}
+      }
+    { \jterm{{\Gamma}{A}}{P}{\mathsf{ev}(f)}
+      }
+\end{equation*}
+or even a rule like
+\begin{equation*}
+\inference
+    { \jterm{\Gamma}{\mprd{A}{P}}{f}
+      \jterm{\Gamma}{A}{x}
+      }
+    { \jterm{\Gamma}{\subst{x}{P}}{\mathsf{ev}(f,x)}
+      }
+\end{equation*}
+involving a term of $A$. We immediately discard the latter as a viable option.
+In the current setting however, there is an even better option than the first. 
+We need a better option because given a family 
+$\jfam{{{\Gamma}{A}}{P}}{Q}$, we will need to have a family over
+$\ctxext{{\Gamma}{\mprd{A}{P}}}{\ctxwk{\mprd{A}{P}}{A}}$ with fibers
+$\subst{{x}{\mathsf{ev}(f)}}{{x}{Q}}$ at $f:\mprd{A}{P}$ and $x:A$. In a 
+name-free setting, such a family is impossible to obtain when using evaluation
+as introduced above. The solution is to introduce evaluation via the rule
+\begin{equation}\label{eq:evtm}
+\inference
+  { \jfam{{\Gamma}{A}}{P}
+    }
+  { \jterm
+      {{{\Gamma}{\mprd{A}{P}}}{\ctxwk{\mprd{A}{P}}{A}}}
+      {\ctxwk{\mprd{A}{P}}{P}}
+      {\evtm{A}{P}}
+    }
+\end{equation}
+The advantages are as follows:
+\begin{enumerate}
+\item It matches the categorical notion of locally cartesian closedness, where
+the unit and the counit consist of morphisms in the category.
+\item We now have the family 
+\begin{equation*}
+\jfam
+  {{{\Gamma}{\mprd{A}{P}}}{\ctxwk{\mprd{A}{P}}{A}}}
+  {\subst{\evtm{A}{P}}{\ctxwk{\mprd{A}{P}}{Q}}}
+\end{equation*}
+Using this family, we hope to find the judgmental equality
+\begin{equation*}
+\jfameq
+  {{\Gamma}{\mprd{A}{P}}}
+  {\mprd{A}{Q}}
+  {\mprd{\ctxwk{\mprd{A}{P}}{A}}{\subst{\evtm{A}{P}}{\ctxwk{\mprd{A}{P}}{Q}}}}
+\end{equation*}
+expressing the action on families of the dependent function constructor in terms
+of its action on contexts. This is in agreement with the usual approach of
+\cite{TheBook}, where no action on families of the dependent function type
+constructor is needed. We will also be able to use the family
+\begin{equation*}
+\jfam
+  {{\Gamma}{\mprd{A}{P}}}
+  {{\ctxwk{\mprd{A}{P}}{A}}{\subst{\evtm{A}{P}}{\ctxwk{\mprd{A}{P}}{Q}}}}
+\end{equation*}
+to formulate the compatibility of the dependent function constructor with itself.
+\end{enumerate}
+Of course, we will also need the following convertibility rule before we 
+continue:
+
+\begin{rmk}
+It is common practice, for example in \cite{TheBook}, to omit explicit notation
+for evaluation. In our setting that would mean that we would denote a term 
+$\subst{f}{\evtm{A}{P}}$ of $P$ in context $\ctxext{\Gamma}{A}$ by $f$.
+However, since the evaluation term is set to be an actual term of the theory,
+it seems to be harmful to the exposition of the theory to copy this practice.
+\end{rmk}
+
+\subsubsection{The compatibility rules of the evaluation terms}
+
+\paragraph{Compatibility with the empty context}
+The following rule explains that the evaluation term is judgmentally equal to
+the identity term when the domain of the dependent function constructor is the
+empty family over $\Gamma$.
+\begin{align*}
+& \inference
+    { \jfam{\Gamma}{P}
+      }
+    { \jtermeq
+        {{\Gamma}{P}}
+        {\ctxwk{P}{P}}
+        {\evtm{\emptyf}{P}}
+        {\idtm{P}}
+      }
+\end{align*}
+
+\paragraph{Currying for the evaluation term}
+We have asserted that there is a judgmental equality
+\begin{equation*}
+\jfameq
+      {\Gamma}
+      {\mprd{\ctxext{A}{P}}{Q}}
+      {\mprd{A}{\mprd{P}{Q}}}
+\end{equation*}
+for every family $Q$ over $\ctxext{{\Gamma}{A}}{P}$. It follows that we have
+the judgmental equalities
+\begin{align*}
+& \ctxwk{\mprd{A}{\mprd{P}{Q}}}{Q}
+  \\
+& \jdeq
+  \subst
+    { \evtm{A}{\mprd{P}{Q}}}
+    { \ctxwk
+        {{\mprd{A}{\mprd{P}{Q}}}{\mprd{P}{Q}}}
+        {{\mprd{A}{\mprd{P}{Q}}}{Q}}
+      }
+  \\
+& \jdeq
+  \subst
+    { \evtm{A}{\mprd{P}{Q}}}
+    { \ctxwk
+        {\mprd{A}{\mprd{P}{Q}}}
+        {{\mprd{P}{Q}}{Q}}
+      }
+\end{align*}
+Since $\evtm{P}{Q}$ is a term of $\ctxwk{\mprd{P}{Q}}{Q}$ in 
+context $\ctxext{{{\Gamma}{A}}{\mprd{P}{Q}}}{\ctxwk{\mprd{P}{Q}}{P}}$ we may
+require the following compatibility rule:
+\begin{align}
+& \inference
+  { \jfam{{{\Gamma}{A}}{P}}{Q}
+    }
+  { \begin{array}{l}
+    \ctxext
+      {{\Gamma}{\mprd{\ctxext{A}{P}}{Q}}}{\ctxwk{\mprd{\ctxext{A}{P}}{Q}}{\ctxext{A}{P}}}
+      \\
+    \jtermeq
+      {\qquad}
+      {\ctxwk{\mprd{\ctxext{A}{P}}{Q}}{Q}}
+      {\evtm{\ctxext{A}{P}}{Q}}
+      {\subst{\evtm{A}{\mprd{P}{Q}}}{\ctxwk{\mprd{A}{\mprd{P}{Q}}}{\evtm{P}{Q}}}}
+    \end{array}
+    }
+\end{align}
+To get a feel for what it says we substitute terms in both expressions involved
+in the following lemma.
+
+\begin{lem}
+Consider $f:\mprd{\ctxext{A}{P}}{Q}$, $x:A$ and $u:\subst{x}{P}$ be terms in
+context $\Gamma$. Then we have the judgmental equality
+\begin{equation*}
+\jtermeq
+  {\Gamma}
+  {\subst{u}{{x}{Q}}}
+  {\subst{\tmext{x}{u}}{{f}{\evtm{\ctxext{A}{P}}{Q}}}}
+  {\subst{u}{{{x}{{f}{\evtm{A}{\mprd{P}{Q}}}}}{\evtm{\subst{x}{P}}{\subst{x}{Q}}}}}
+\end{equation*}
+is valid.
+\end{lem}
+
+\paragraph{Weakening is compatible with the evaluation term}
+\begin{align*}
+& \inference
+  { \jfam{\Gamma}{A}
+    \jfam{{{\Gamma}{B}}{Q}}{R}
+    }
+  { \begin{array}{l}
+    \ctxext
+        { { { {\Gamma}
+              {A}
+              }
+            { \ctxwk{A}{B}
+              }
+            }
+          { \mprd{\ctxwk{A}{Q}}{\ctxwk{A}{R}}
+            }
+          }
+        { \ctxwk{\mprd{\ctxwk{A}{Q}}{\ctxwk{A}{R}}}{\ctxwk{A}{Q}}
+          }
+      \\
+    \jtermeq
+      {\qquad}
+      {\ctxwk{\mprd{\ctxwk{A}{Q}}{\ctxwk{A}{R}}}{\ctxwk{A}{R}}}
+      {\evtm{\ctxwk{A}{Q}}{\ctxwk{A}{R}}}
+      {\ctxwk{A}{\evtm{Q}{R}}}  
+    \end{array}
+    }
+\end{align*}
+
+\paragraph{Substitution is compatible with the evaluation term}
+\begin{align*}
+& \inference
+  { \jterm{\Gamma}{A}{x}
+    \jfam{{{{\Gamma}{A}}{P}}{Q}}{R}
+    }
+  { \begin{array}{l}
+    \ctxext
+      {{{\Gamma}{\subst{x}{P}}}{\mprd{\subst{x}{Q}}{\subst{x}{R}}}}
+      {\ctxwk{\mprd{\subst{x}{Q}}{\subst{x}{R}}}{\subst{x}{Q}}}
+      \\
+    \jtermeq
+      {\qquad}
+      {\ctxwk{\mprd{\subst{x}{Q}}{\subst{x}{R}}}{\subst{x}{R}}}
+      {\subst{x}{\evtm{Q}{R}}}
+      {\evtm{\subst{x}{Q}}{\subst{x}{R}}}
+    \end{array}
+    }  
+\end{align*}
+
+\paragraph{The evaluation term is compatible with weakening}
+\begin{align*}
+& \inference
+  { \jfam{{{\Gamma}{A}}{P}}{Q}
+    \jfam{{{{\Gamma}{P}}{Q}}{R}}{S}
+    \jterm{{{\Gamma}{\mprd{A}{P}}}{\mprd{A}{R}}}{\mprd{A}{S}}{k}
+    }
+  { \jtermeq
+      {{{{{\Gamma}{A}}{P}}{Q}}{\ctxwk{Q}{R}}}
+      {\ctxwk{Q}{S}}
+      {\sev{A}{\ctxwk{Q}{S}}{\ctxwk{\mprd{A}{Q}}{k}}}
+      {\ctxwk{Q}{\sev{A}{S}{k}}}
+    }
+\end{align*}
+
+\subsection{Composition of dependent functions}

### 2014-06-28 article-models-ttinternal.tex
fixed broken references
small improvements; starting to write the rules of weakening algebras

@@
   \jcomp{}{\idtm{\jcomp{}{e_0}{P}}}{\jcomp{}{e_0}{\jcomp{}{e_0}{Q}}}
-  \tag{by \autoref{idfunc-precomp}}
+  \tag{by \autoref{precomp-idtm-f}}
   \\
@@
   \jcomp{}{e_1}{\jcomp{}{e_0}{Q}}.
-  \tag{by \autoref{idfunc-precomp}}
+  \tag{by \autoref{precomp-idtm-f}}
 \end{align*}
@@
 \begin{align*}
-\subst{{x_2}{{x_1}{\cftfamext[\extalg{A}]}}}{{x_0}{\cftfamext[\extalg{A}]}}
+\subst{{x_2}{{x_1}{\cftfamext[\extalg{A}]}}}{{x_0}{\cftctxext[\extalg{A}]}}
 & \jdeq
@@
     {x_2}
-    {{{x_1}{{x_0}{\cftfamext[\extalg{A}]}}}{\cftfamext[\extalg{A}]}}.
+    {{{x_1}{{x_0}{\cftctxext[\extalg{A}]}}}{\cftctxext[\extalg{A}]}}.
   \intertext{%
@@
   \label{empalg-eq2}
-  \\
-\jfameq*{{\Gamma}{\cftalgc{\cftalg{A}}}}{\subst{\cftempf{\cftalg{A}}}{\cftalgt{\cftalg{A}}}}{\emptyf}
+%  \\
+%\jfameq*{{\Gamma}{\cftalgc{\cftalg{A}}}}{\subst{\cftempf{\cftalg{A}}}{\cftalgt{\cftalg{A}}}}{\emptyf}
 \end{align}
@@
 
-\subsection{Weakening term-algebras}
-Weakening term-algebras will be term-algebras with certain added structure.
+\subsection{Weakening algebras}
+Weakening algebras will be term-algebras with certain added structure.
 Although strictly speaking one could formulate a notion of weakening algebra
@@
 
-Let $(A,P,T,\epsilon_0,\epsilon_1,\phi_0,\phi_1)$ be a term-algebra in context
+Let $\cftalg{A}$ be a term-algebra in context $\Gamma$. Weakening will be a
+homomorphism
+\begin{equation*}
+\jhom
+  {{{\Gamma}{\cftalgc{\cftalg{A}}}}{\cftalgf{\cftalg{A}}}}
+  {\ctxwk{\cftalgf{\cftalg{A}}}{\cftalg{F_A}}}
+  {\jcomp{}{\cftctxext}{\cftalg{F_A}}}
+  {\cftwk{\cftalg{A}}}
+\end{equation*}
+satisfying judgmental equalities which express the compatibility with extension,
+the empty object and itself. Thus, the rules we interpret for weakening are
+the rules in sections \autoref{comp-we,comp-w0,comp-ww,comp-ew,comp-0w}. The
+definition of weakening algebras can be found in \autoref{sec:cftwkalg-defn}.
+
+\subsubsection{Weakening preserves itself}
+The following diagram displays the compatibility of weakening with itself:
+\begin{equation*}
+\begin{tikzcd}[column sep=huge]
+\ctxwk{\cftalgf{\cftalg{A}}}{{\cftalgf{\cftalg{F_A}}}{\cftalg{F_{F_A}}}}
+  \ar{d}
+    [swap]{ \ctxwk
+        {{\cftalgf{\cftalg{A}}}{\cftalgf{\cftalg{F_A}}}}
+        {\boldsymbol{\mathcal{F}}_{\cftwk{\cftalg{A}}}}
+      }
+  \ar{r}{\ctxwk{\cftalgf{\cftalg{A}}}{\jcomp{}{\cftctxext}{\cftwk{\cftalg{A}}}}}
+& \ctxwk{\cftalgf{\cftalg{A}}}{\cftalg{F_{F_{F_A}}}}
+  \ar{d}{\jcomp{}{\cftctxext}{\boldsymbol{\mathcal{F}}_{\cftwk{\cftalg{A}}}}}
+  \\
+\jcomp{}{\cftwkc{A}}{
+\ctxwk{{\cftalgf{\cftalg{A}}}{\cftalgf{\cftalg{F_A}}}}{\cftalg{F_{F_{F_A}}}}}
+  \ar{r}[swap]{\jcomp{}{\cftwkc{A}}{\jcomp{}{\cftctxext}{\cftwk{A}}}}
+& \jcomp{}{\cftwkc{A}}{\cftalg{F_{F_{F_{F_A}}}}}
+\end{tikzcd}
+\end{equation*}
+
+\subsubsection{Currying for weakening}
+\begin{equation*}
+\jhomeq
+  { {{{\Gamma}{\cftalgc{\cftalg{A}}}}{\cftalgf{\cftalg{A}}}}
+    {\cftalgf{\cftalg{F_A}}}
+    }
+  { \ctxwk
+      {\cftalgf{\cftalg{F_A}}}
+      {\cftalg{F_{F_A}}}
+    }
+  { \cftalg{F_{F_{F_A}}}
+    }
+  { \jcomp{}{\cftfamext}{\cftwk{\cftalg{A}}}
+    }
+  { \jcomp{}{\cftwk{\cftalg{A}}}{\jcomp{}{\cftctxext}{\cftwk{\cftalg{A}}}}
+    }
+\end{equation*}
+
+\subsubsection{Weakening by the empty family}
+\begin{equation*}
+\jhomeq
+  { {\Gamma}{\cftalgc{\cftalg{A}}}
+    }
+  { \cftalg{F_A}
+    }
+  { \cftalg{F_A}
+    }
+  { \subst{\cftempf{\cftalg{A}}}{\cftwk{\cftalg{A}}}
+    }
+  { id_{\cftalg{F_A}}
+    }
+\end{equation*}
+
+\subsubsection{The definition of weakening algebras}\label{sec:cftwkalg-defn}
+
+\subsubsection{Derivable properties of weakening algebras}
+We should theoremize the following statements:
+\begin{enumerate}
+\item If $\cftalg{A}$ is a weakening algebra, then so is $\cftalg{F_A}$ if we take
+$\jcomp{}{\cftctxext}{\cftwk{\cftalg{A}}}$ as its weakening.
+\end{enumerate}
+
+Let $\cftalg{A}$ be a term-algebra in context
 $\Gamma$. The weakening homomorphism $\omega$ will be a term-homomorphism from 
@@
 
-\subsubsection{The definition of weakening algebras}
-
 \subsection{Extension-substitution algebras}

### 2014-06-28 article-models-ttinternal.tex
lots of improvements

@@
 asserting that $\extalg{A}$ is an \emph{extension algebra in context $\Gamma$}
-to be the conjunction of the following six judmgents:
+to be the conjunction of the following seven judmgents:
 \begin{align*}
@@
 \jextalgfam*{\Gamma}{A}
+  \\
+\jextalgtm*{\Gamma}{A}
   \\
@@
 In other words, an extension algebra $\extalg{A}$ in context $\Gamma$
-is a quadruple $\unfold{\extalg{A}}$ 
-satisfying the judgmental equalities of
+is a quintuple $\unfold{\extalg{A}}$ 
+satisfying the judgmental equalities displayed in the diagrams in
 \autoref{eq:extalg-eq1,eq:extalg-eq2}. Extension algebras are judgmentally
@@
 \begin{equation*}
-(A,\emptyf,\idtm{A},\emptytm)
+(A,\emptyf,\emptyf,\idtm{A},\emptytm)
 \end{equation*}
@@
 \begin{equation*}
-(\emptyf,A,\emptytm,\idtm{A}).
+(\emptyf,A,\emptyf,\emptytm,\idtm{A}).
 \end{equation*}
@@
 \begin{equation*}
-(A,\ctxwk{A}{A},\cprojfstf{A}{\ctxwk{A}{A}},\cprojfstf{\ctxwk{A}}{\ctxwk{A}{{A}{A}}})
+(A,\ctxwk{A}{A},\emptyf,\cprojfstf{A}{\ctxwk{A}{A}},\cprojfstf{\ctxwk{A}}{\ctxwk{A}{{A}{A}}})
 \end{equation*}
@@
 \begin{equation*}
-\cftalg{F_A}\defeq(\cftalgf{\cftalg{A}},\jcomp{}{\cftctxext}{\cftalgf{\cftalg{A}}},\cftfamext,\jcomp{}{\cftctxext}{\cftfamext})
+\cftalg{F_A}
+\defeq
+( \cftalgf{\cftalg{A}},
+  \jcomp{}{\cftctxext[\cftalg{A}]}{\cftalgf{\cftalg{A}}},
+  \jcomp{}{\cftctxext[\cftalg{A}]}{\cftalgt{\cftalg{A}}},
+  \cftfamext,
+  \jcomp{}{\cftctxext[\cftalg{A}]}{\cftfamext})
 \end{equation*}
@@
 Let $\extalg{Q}$ be an extension algebra in context $\ctxext{\Gamma}{B}$ and let
-$\jfam{\Gamma}{A}$ be a family of contexts. Then the quadruple
+$\jfam{\Gamma}{A}$ be a family of contexts. Then the quintuple
 \begin{equation*}
@@
   \defeq
-  (\ctxwk{A}{\cftalgc{\cftalg{Q}}},\ctxwk{A}{\cftalgf{\cftalg{Q}}},\ctxwk{A}{\cftctxext[\cftalg{Q}]},\ctxwk{A}{\cftfamext[\cftalg{Q}]})
+  ( \ctxwk{A}{\cftalgc{\cftalg{Q}}},
+    \ctxwk{A}{\cftalgf{\cftalg{Q}}},
+    \ctxwk{A}{\cftalgt{\cftalg{Q}}},
+    \ctxwk{A}{\cftctxext[\cftalg{Q}]},
+    \ctxwk{A}{\cftfamext[\cftalg{Q}]})
 \end{equation*}
@@
 Let $\extalg{Q}$ be an extension algebra in context $\ctxext{{\Gamma}{A}}{P}$
-and let $\jterm{\Gamma}{A}{x}$ be a term. Then the quadruple
+and let $\jterm{\Gamma}{A}{x}$ be a term. Then the quintuple
 \begin{equation*}
@@
   \defeq
-  (\subst{x}{\cftalgc{\cftalg{Q}}},\subst{x}{\cftalgf{\cftalg{Q}}},\subst{x}{\cftctxext[\cftalg{Q}]},\subst{x}{\cftfamext[\cftalg{Q}]})
+  ( \subst{x}{\cftalgc{\cftalg{Q}}},
+    \subst{x}{\cftalgf{\cftalg{Q}}},
+    \subst{x}{\cftalgt{\cftalg{Q}}},
+    \subst{x}{\cftctxext[\cftalg{Q}]},
+    \subst{x}{\cftfamext[\cftalg{Q}]})
 \end{equation*}
@@
 Let $\extalg{Q}$ be an extension algebra in context $\ctxext{\Gamma}{B}$
-and let $\jhom{\Gamma}{A}{B}{f}$. Then the quadruple
+and let $\jhom{\Gamma}{A}{B}{f}$. Then the quintuple
 \begin{equation*}
@@
   \defeq
-  (\jcomp{A}{f}{\cftalgc{\cftalg{Q}}},\jcomp{A}{f}{\cftalgf{\cftalg{Q}}},\jcomp{A}{f}{\cftctxext[\cftalg{Q}]},\jcomp{A}{f}{\cftfamext[\cftalg{Q}]})
+  ( \jcomp{A}{f}{\cftalgc{\cftalg{Q}}},
+    \jcomp{A}{f}{\cftalgf{\cftalg{Q}}},
+    \jcomp{A}{f}{\cftalgt{\cftalg{Q}}},
+    \jcomp{A}{f}{\cftctxext[\cftalg{Q}]},
+    \jcomp{A}{f}{\cftfamext[\cftalg{Q}]})
 \end{equation*}
@@
 
+\subsection{Extension homomorphisms}
+
 \begin{defn}
-An \emph{extension homomorphism $\exthom{f}$ from $\extalg{A}$ to
-$\extalg{B}$ in context $\Gamma$} is a pair $\unfold{\exthom{f}}$ consisting of
-\begin{align*}
-\jhom*{\Gamma}{\cftalgc{\cftalg{A}}}{\cftalgc{\cftalg{B}}}{\cfthomc{\cfthom{f}}}
-  \\
-\jfhom*{\Gamma}{\cftalgc{\cftalg{A}}}{\cftalgc{\cftalg{B}}}{\cfthomc{\cfthom{f}}}{\cftalgf{\cftalg{A}}}{\cftalgf{\cftalg{B}}}{\cfthomf{\cfthom{f}}}
+An \emph{extension homomorphism $\cfthom{f}$ from $\extalg{A}$ to
+$\extalg{B}$ in context $\Gamma$} is a triple $\unfold{\cfthom{f}}$ consisting of
+\begin{align*}
+\jhom*
+  {\Gamma}
+  {\cftalgc{\cftalg{A}}}
+  {\cftalgc{\cftalg{B}}}
+  {\cfthomc{\cfthom{f}}}
+  \\
+\jfhom*
+  {\Gamma}
+  {\cftalgc{\cftalg{A}}}
+  {\cftalgc{\cftalg{B}}}
+  {\cfthomc{\cfthom{f}}}
+  {\cftalgf{\cftalg{A}}}
+  {\cftalgf{\cftalg{B}}}
+  {\cfthomf{\cfthom{f}}}
+  \\
+\jfhom*
+  {\Gamma}
+  {{\cftalgc{\cftalg{A}}}{\cftalgf{\cftalg{A}}}}
+  {{\cftalgc{\cftalg{B}}}{\cftalgf{\cftalg{B}}}}
+  {\jvcomp{}{\cfthomc{\cfthom{f}}}{\cfthomf{\cfthom{f}}}}
+  {\cftalgt{\cftalg{A}}}
+  {\cftalgt{\cftalg{B}}}
+  {\cfthomt{\cfthom{f}}}
 \end{align*}
@@
 \begin{equation*}
-\cfthom{F_f}
+\boldsymbol{\mathcal{F}}_\cfthom{f}
   \defeq
-  (\cfthomf{\cfthom{f}},\jcomp{}{\cftctxext[\cftalg{A}]}{\cfthomf{\cfthom{f}}})
+  ( \cfthomf{\cfthom{f}},
+    \jcomp{}{\cftctxext[\cftalg{A}]}{\cfthomf{\cfthom{f}}},
+    \jcomp{}{\cftctxext[\cftalg{A}]}{\cfthomt{\cfthom{f}}})
 \end{equation*}
@@
 \begin{defn}
-Let $\exthom{f}$ and $\exthom{g}$ be extension homomorphisms from
+Let $\cfthom{f}$ and $\cfthom{g}$ be extension homomorphisms from
 $\extalg{A}$ to $\extalg{B}$ and from $\extalg{B}$ to $\extalg{C}$, respectively.
@@
 \begin{equation*}
-\exthomcomp{\exthom{f}}{\exthom{g}}
+\cfthomcomp{\cfthom{f}}{\cfthom{g}}
   \defeq
-  \unfold{\exthomcomp{\exthom{f}}{\exthom{g}}}.
-\end{equation*}
-In other words, extension homomorphisms are composed by taking the outer rectangle
-in the diagram
+  \unfold{\cfthomcomp{\cfthom{f}}{\cfthom{g}}}.
+\end{equation*}
+In other words, extension homomorphisms are composed by taking the horizontal 
+rectangles in the diagram
 \begin{equation*}
 \begin{tikzcd}
+\extalgt{\extalg{A}} 
+  \ar[fib]{d}
+  \ar{r}{\cfthomt{\cfthom{f}}}
+& \extalgt{\extalg{B}} 
+  \ar[fib]{d}
+  \ar{r}{\cfthomt{\cfthom{g}}}
+& \extalgt{\extalg{C}}
+  \ar[fib]{d}
+  \\
 \extalgf{\extalg{A}} 
   \ar[fib]{d}
-  \ar{r}{\exthomf{\exthom{f}}}
+  \ar{r}{\cfthomf{\cfthom{f}}}
 & \extalgf{\extalg{B}} 
   \ar[fib]{d}
-  \ar{r}{\exthomf{\exthom{g}}}
+  \ar{r}{\cfthomf{\cfthom{g}}}
 & \extalgf{\extalg{C}}
@@
 \extalgc{\extalg{A}}
-  \ar{r}{\exthomc{\exthom{f}}}
+  \ar{r}{\cfthomc{\cfthom{f}}}
 & \extalgc{\extalg{B}}
-  \ar{r}{\exthomc{\exthom{g}}}
+  \ar{r}{\cfthomc{\cfthom{g}}}
 & \extalgc{\extalg{C}}
@@
 \begin{thm}
-Let $\exthom{f}$ and $\exthom{g}$ be extension homomorphisms from
+Let $\cfthom{f}$ and $\cfthom{g}$ be extension homomorphisms from
 $\extalg{A}$ to $\extalg{B}$ and from $\extalg{B}$ to $\extalg{C}$, respectively.
-Then $\exthomcomp{\exthom{f}}{\exthom{g}}$ is an extension homomorphism from
+Then $\cfthomcomp{\cfthom{f}}{\cfthom{g}}$ is an extension homomorphism from
 $\extalg{A}$ to $\extalg{C}$. 
@@
 \begin{lem}\label{lem:empalg-mor}
-Consider an empty-object algebra $\cftalg{A}$ in context $\Gamma$
-and a morphism 
-\begin{equation*}
-\jhom{\Gamma}{{\cftalgc{\cftalg{A}}}{\cftalgf{\cftalg{A}}}}{B}{f}.
-\end{equation*}
-Then $\subst{\cftemp[i]{\cftalg{A}}}{f}$ is a morphism from $\cftalgc{\cftalg{A}}$ to $B$ 
-in context $\Gamma$ and the following inference rules are valid for $i$ being 
-$\mathsf{c}$ or $\mathsf{f}$:
+Consider a family $\jfam{\Gamma}{A}$ with a term $\jterm{\Gamma}{A}{x}$ and 
+a family $\jfam{{\Gamma}{A}}{P}$ with a term $\jterm{{\Gamma}{A}}{P}{y}$
+satisfying the judgmental equalities 
+\begin{align}
+\jfameq*{\Gamma}{\subst{x}{P}}{A}
+  \label{eq1:empalg-mor}
+  \\
+\jtermeq*{\Gamma}{A}{\subst{x}{y}}{x}.
+  \label{eq2:empalg-mor}
+\end{align}
+Consider also a morphism
+\begin{equation*}
+\jhom{\Gamma}{{A}{P}}{B}{f}.
+\end{equation*}
+Then $\subst{\xi}{f}$ is a morphism from $A$ to $B$ 
+in context $\Gamma$ and the following inference rules are valid for $\xi$ being 
+$x$ or $y$:
 \begin{align*}
 & \inference
-  { \jfam{{\Gamma}{\cftalgc{\cftalg{A}}}}{Q}
+  { \jfam{{\Gamma}{B}}{Q}
     }
   { \jfameq
-      {{\Gamma}{\cftalgc{\cftalg{A}}}}
-      {\subst{\cftemp[i]{\cftalg{A}}}{\jcomp{}{f}{Q}}}
-      {\jcomp{}{\subst{\cftemp[i]{\cftalg{A}}}{f}}{Q}}
+      {{\Gamma}{A}}
+      {\subst{\xi}{\jcomp{}{f}{Q}}}
+      {\jcomp{}{\subst{\xi}{f}}{Q}}
     }
@@
 & \inference
-  { \jfam{{{\Gamma}{\cftalgc{\cftalg{A}}}}{Q}}{R}
+  { \jfam{{{\Gamma}{B}}{Q}}{R}
     }
   { \jfameq
-      {{{\Gamma}{\cftalgc{\cftalg{A}}}}{\jcomp{}{\subst{\cftemp[i]{\cftalg{A}}}{f}}{Q}}}
-      {\subst{\cftemp[i]{\cftalg{A}}}{\jcomp{}{f}{R}}}
-      {\jcomp{}{\subst{\cftemp[i]{\cftalg{A}}}{f}}{R}}
+      {{{\Gamma}{A}}{\jcomp{}{\subst{\xi}{f}}{Q}}}
+      {\subst{\xi}{\jcomp{}{f}{R}}}
+      {\jcomp{}{\subst{\xi}{f}}{R}}
     }
@@
 & \inference
-  { \jterm{{{\Gamma}{\cftalgc{\cftalg{A}}}}{Q}}{R}{h}
+  { \jterm{{{\Gamma}{B}}{Q}}{R}{h}
     }
   { \jtermeq
-      {{{\Gamma}{\cftalgc{\cftalg{A}}}}{\jcomp{}{\subst{\cftemp[i]{\cftalg{A}}}{f}}{Q}}}
-      {\jcomp{}{\subst{\cftemp[i]{\cftalg{A}}}{f}}{R}}
-      {\subst{\cftemp[i]{\cftalg{A}}}{\jcomp{}{f}{h}}}
-      {\jcomp{}{\subst{\cftemp[i]{\cftalg{A}}}{f}}{h}}
+      {{{\Gamma}{A}}{\jcomp{}{\subst{\xi}{f}}{Q}}}
+      {\jcomp{}{\subst{\xi}{f}}{R}}
+      {\subst{\xi}{\jcomp{}{f}{h}}}
+      {\jcomp{}{\subst{\xi}{f}}{h}}
     }
@@
 \begin{align*}
-\subst{\cftempc{\cftalg{A}}}{\jcomp{}{f}{Q}}
+\subst{x}{\jcomp{}{f}{Q}}
 & \jdeq
   \subst
-    {\cftempc{\cftalg{A}}}
+    {x}
     {{f}{\ctxwk{\ctxext{\cftalgc{\cftalg{A}}}{\cftalgf{\cftalg{A}}}}{Q}}}
@@
   \subst
-    { {\cftempc{\cftalg{A}}}{f}
-      }
-    { {\cftempc{\cftalg{A}}}
+    { {x}{f}
+      }
+    { {x}
       {\ctxwk{\ctxext{\cftalgc{\cftalg{A}}}{\cftalgf{\cftalg{A}}}}{Q}}
@@
   \subst
-    { {\cftempc{\cftalg{A}}}{f}
-      }
-    { {\cftempc{\cftalg{A}}}
+    { {x}{f}
+      }
+    { {x}
       {\ctxwk{\cftalgf{\cftalg{A}}}{{\cftalgc{\cftalg{A}}}{Q}}}
@@
   \subst
-    { {\cftempc{\cftalg{A}}}{f}
+    { {x}{f}
       }
     { \ctxwk
-        {\subst{\cftempc{\cftalg{A}}}{\cftalgf{\cftalg{A}}}}
-        {\subst{\cftempc{\cftalg{A}}}{\ctxwk{\cftalgc{\cftalg{A}}}{Q}}}
+        {\subst{x}{\cftalgf{\cftalg{A}}}}
+        {\subst{x}{\ctxwk{\cftalgc{\cftalg{A}}}{Q}}}
       }
@@
   \subst
-    { {\cftempc{\cftalg{A}}}{f}
+    { {x}{f}
       }
     { \ctxwk
-        {\subst{\cftempc{\cftalg{A}}}{\cftalgf{\cftalg{A}}}}
+        {\subst{x}{\cftalgf{\cftalg{A}}}}
         {Q}
@@
 & \jdeq
-  \subst{{\cftempc{\cftalg{A}}}{f}}{\ctxwk{\cftalgc{\cftalg{A}}}{Q}}
-  \tag{by \autoref{empalg-eq1}}
-  \\
-& \jdeq
-  \jcomp{}{\subst{\cftempc{\cftalg{A}}}{f}}{Q}.
+  \subst{{x}{f}}{\ctxwk{\cftalgc{\cftalg{A}}}{Q}}
+  \tag{by \autoref{eq1:empalg-mor}}
+  \\
+& \jdeq
+  \jcomp{}{\subst{x}{f}}{Q}.
   \tag{by definition}
@@
 \begin{align*}
-\subst{\cftempf{\cftalg{A}}}{\jcomp{}{f}{Q}}
+\subst{y}{\jcomp{}{f}{Q}}
 & \jdeq
   \subst
-    { \cftempf{\cftalg{A}}
+    { y
       }
@@
   \subst
-    { {\cftempf{\cftalg{A}}}{f}
-      }
-    { {\cftempf{\cftalg{A}}}
+    { {y}{f}
+      }
+    { {y}
       {\ctxwk{\ctxext{\cftalgc{\cftalg{A}}}{\cftalgf{\cftalg{A}}}}{Q}}
@@
   \subst
-    { {\cftempf{\cftalg{A}}}{f}
-      }
-    { {\cftempf{\cftalg{A}}}
+    { {y}{f}
+      }
+    { {y}
       {\ctxwk{\cftalgf{\cftalg{A}}}{{\cftalgc{\cftalg{A}}}{Q}}}
@@
   \subst
-    { {\cftempf{\cftalg{A}}}{f}
+    { {y}{f}
       }
@@
 & \jdeq
-  \jcomp{}{\subst{\cftempf{\cftalg{A}}}{f}}{Q}.
+  \jcomp{}{\subst{y}{f}}{Q}.
   \tag{by definition}
@@
 
+\subsection{Homorphisms of term-algebras}
+\begin{defn}
+A homomorphism of term algebras $\cfthom{f}$ from $\cftalg{A}$ to $\cftalg{B}$
+consists of a triple $(\cfthomc{\cfthom{f}},\cfthomf{\cfthom{f}},\cfthomt{\cfthom{f}})$
+such that $(\cfthomc{\cfthom{f}},\cfthomf{\cfthom{f}})$ is an extension homomorphism
+\end{defn}
+
 \subsection{Term-algebras and term-operations}
@@
 homomorphism
-\begin{equation*}
+\begin{equation}\label{cftwk}
 \jhom
@@
   {\ctxwk{\cftalgf{\cftalg{A}}}{\cftalg{F_A}}}
-  {\jcomp{}{\cftctxext}{\cftalg{F_A}}}
+  {\cftalg{F_{F_A}}}
   {\cftwk{\cftalg{A}}}
-\end{equation*}
+\end{equation}
 satisfying judgmental equalities which express the compatibility with extension,
@@
 
+We begin with some trivial remarks to familiarize ourselves with the situation.
+The morphism $\cftwk{\cftalg{A}}$ of \autoref{cftwk} is a triple
+$(\cftwkc{\cftalg{A}},\cftwkf{\cftalg{A}},\cftwkt{\cftalg{A}})$ consisting of
+the morphisms displayed in the following diagram:
+\begin{equation*}
+\begin{tikzcd}
+\ctxwk{\cftalgf{\cftalg{A}}}{\jcomp{}{\cftctxext}{\cftalgt{\cftalg{A}}}}
+  \ar[fib]{d}
+  \ar{r}{\cftwkt{\cftalg{A}}}
+& \jcomp{}{\cftctxext}{\jcomp{}{\cftctxext}{\cftalgt{\cftalg{A}}}}
+  \ar[fib]{d}
+  \\
+\ctxwk{\cftalgf{\cftalg{A}}}{\jcomp{}{\cftctxext}{\cftalgf{\cftalg{A}}}}
+  \ar[fib]{d}
+  \ar{r}{\cftwkf{\cftalg{A}}}
+& \jcomp{}{\cftctxext}{\jcomp{}{\cftctxext}{\cftalgf{\cftalg{A}}}}
+  \ar[fib]{d}
+  \\
+\ctxwk{\cftalgf{\cftalg{A}}}{\cftalgf{\cftalg{A}}}
+  \ar{r}[swap]{\cftwkc{\cftalg{A}}}
+& \jcomp{}{\cftctxext}{\cftalgf{\cftalg{A}}}
+\end{tikzcd}
+\end{equation*}
+
+\begin{rmk}
+Let $\cftalg{A}$ be a term-algebra and let $\cftwk{\cftalg{A}}$ be as in
+\autoref{cftwk}. Then we have
+\begin{equation*}
+\jterm
+  {\Gamma}
+  {\subst{{a}{{\gamma}{\cftctxext}}}{\cftalgf{\cftalg{A}}}}
+  {\subst{b}{{a}{{\gamma}{\cftwkc{\cftalg{A}}}}}}
+\end{equation*}
+for $\gamma:A$ and $a,b:\subst{\gamma}{\cftalgf{\cftalg{A}}}$. We have
+\begin{equation*}
+\jterm
+  {\Gamma}
+  {\subst{{{b}{{a}{{\gamma}{\cftwkc{\cftalg{A}}}}}}{{{a}{{\gamma}{\cftctxext}}}{\cftctxext}}}{\cftalgf{\cftalg{A}}}}
+  {\subst{q}{{b}{{a}{{\gamma}{\cftwkf{\cftalg{A}}}}}}}
+\end{equation*}
+for $\gamma:A$, $a,b:\subst{\gamma}{\cftalgf{\cftalg{A}}}$ and
+$q:\subst{{b}{{\gamma}{\cftctxext}}}{\cftalgf{\cftalg{A}}}$. Finally, we have
+\begin{equation*}
+\jterm
+  {\Gamma}
+  {\subst{{q}{{b}{{a}{{\gamma}{\cftwkf{\cftalg{A}}}}}}}{{{{b}{{a}{{\gamma}{\cftwkc{\cftalg{A}}}}}}{{{a}{{\gamma}{\cftctxext}}}{\cftctxext}}}{\cftalgt{\cftalg{A}}}}}
+  {\subst{g}{{q}{{b}{{a}{{\gamma}{\cftwkt{\cftalg{A}}}}}}}}
+\end{equation*}
+for $\gamma:A$, $a,b:\subst{\gamma}{\cftalgf{\cftalg{A}}}$,
+$q:\subst{{b}{{\gamma}{\cftctxext}}}{\cftalgf{\cftalg{A}}}$ and
+$g:\subst{q}{{{b}{{\gamma}{\cftctxext}}}{\cftalgt{\cftalg{A}}}}$.
+These observations show that the weakening morphism $\cftwk{\cftalg{A}}$ is
+indeed an abstraction of the introduction rules for the weakening operation.
+\end{rmk}
+
 \subsubsection{Weakening preserves itself}
 The following diagram displays the compatibility of weakening with itself:
-\begin{equation*}
+\begin{equation*}\label{diag:cftwk-cftwk}
 \begin{tikzcd}[column sep=huge]
@@
   \\
-\jcomp{}{\cftwkc{A}}{
-\ctxwk{{\cftalgf{\cftalg{A}}}{\cftalgf{\cftalg{F_A}}}}{\cftalg{F_{F_{F_A}}}}}
+\ctxwk{{\cftalgf{\cftalg{A}}}{\cftalgf{\cftalg{F_A}}}}{\jcomp{}{\cftwkc{A}}{\cftalg{F_{F_{F_A}}}}}
   \ar{r}[swap]{\jcomp{}{\cftwkc{A}}{\jcomp{}{\cftctxext}{\cftwk{A}}}}
@@
 \end{equation*}
+This is a complicated diagram and it is not immediately clear that all of it
+makes sense. Therefore we examine its ingredients in more detail.
+
+The morphisms on the top and on the right are fine as they are displayed
+
+\begin{lem}
+Let $\cftalg{A}$ be a term-algebra in context $\Gamma$ and let
+\begin{equation*}
+\jhom
+  {{{\Gamma}{\cftalgc{\cftalg{A}}}}{\cftalgf{\cftalg{A}}}}
+  {\ctxwk{\cftalgf{\cftalg{A}}}{\cftalg{F_A}}}
+  {\jcomp{}{\cftctxext}{\cftalg{F_A}}}
+  {\cftwk{\cftalg{A}}}
+\end{equation*}
+be a homomorphism of term-algebras satisfying \autoref{diag:cftwk-cftwk}. Then
+we have the judgmental equality
+\begin{equation*}
+\jtermeq
+  {}
+  {}
+  {}
+  {}
+\end{equation*}
+for $\gamma:A$, $x:\cftalgf{\cftalg{A}}$, $y_0:\cftalgf{\cftalg{A}}$ and
+$y_1:\cftalgf{\cftalg{F_A}}$.
+\end{lem}
 
@@
 
-Let $\cftalg{A}$ be a term-algebra in context
-$\Gamma$. The weakening homomorphism $\omega$ will be a term-homomorphism from 
-$\ctxwk{P}{\mathcal{F}_\mathcal{A}}$ to 
-$\jcomp{}{\epsilon_0}{\mathcal{F}_\mathcal{A}}$ satisfying certain extra
-judgmental equalities. The goal of this subsection is to spell these out.
-
-Concretely, the weakening homomorphism is a triple $(\omega_0,\omega_1,\omega_2)$
-consisting of
-\begin{align*}
-\jhom*
-  {{{\Gamma}{A}}{P}}
-  {\ctxwk{P}{P}}
-  {\jcomp{}{\epsilon_0}{P}}
-  {\omega_0}
-  \\
-\jfhom*
-  {{{\Gamma}{A}}{P}}
-  {\ctxwk{P}{P}}
-  {\jcomp{}{\epsilon_0}{P}}
-  {\omega_0}
-  {\ctxwk{P}{\jcomp{}{\epsilon_0}{P}}}
-  {\jcomp{}{\epsilon_0}{\jcomp{}{\epsilon_0}{P}}}
-  {\omega_1}
-  \\
-\jfhom*
-  {{{\Gamma}{A}}{P}}
-  {\ctxwk{P}{\ctxext{P}{\jcomp{}{\epsilon_0}{P}}}}
-  {\jcomp{}{\epsilon_0}{\ctxext{P}{\jcomp{}{\epsilon_0}{P}}}}
-  {\jvcomp{\ctxwk{P}{\jcomp{}{\epsilon_0}{P}}}{\omega_0}{\omega_1}}
-  {\ctxwk{P}{\jcomp{}{\epsilon_0}{T}}}
-  {\jcomp{}{\epsilon_0}{\jcomp{}{\epsilon_0}{T}}}
-  {\omega_2}
-\end{align*}
-We could also summarize the situation in a diagram:
-\begin{equation*}
-\begin{tikzcd}
-\ctxwk{P}{\jcomp{}{\epsilon_0}{T}}
-  \ar[fib]{d}
-  \ar{r}{\omega_2}
-& \jcomp{}{\epsilon_0}{\jcomp{}{\epsilon_0}{T}}
-  \ar[fib]{d}
-  \\
-\ctxwk{P}{\jcomp{}{\epsilon_0}{P}}
-  \ar[fib]{d}
-  \ar{r}{\omega_1}
-& \jcomp{}{\epsilon_0}{\jcomp{}{\epsilon_0}{P}}
-  \ar[fib]{d}
-  \\
-\ctxwk{P}{P}
-  \ar{r}[swap]{\omega_0}
-& \jcomp{}{\epsilon_0}{P}
-\end{tikzcd}
-\end{equation*}
-
-Before we go into chasing the judgmental equalities that are to be imposed on
-weakening, we develop a bit of intuition by explaining how
-$\omega_0$ and $\omega_1$ act when applied to a family $x_1:\subst{x_0}{P}$
-over $x_0:A$, and we will provide the explicit requirements that make
-$\omega$ a term-homomorphism. Note that
-\begin{equation*}
-\jterm{{\Gamma}{\subst{x_0}{P}}}{\subst{{x_1}{{x_0}{\epsilon_0}}}{P}}{\subst{x_1}{{x_0}{\omega_0}}},
-\end{equation*}
-so $\subst{x_1}{{x_0}{\omega_0}}$ takes families over $x_0$ to families over
-the extended $\subst{x_1}{{x_0}{\epsilon_0}}$.
-
-To compute the type of $\subst{x_2}{{x_1}{{x_0}{\omega_1}}}$ for families
-$x_1,x_2:\subst{x_0}{P}$ over $x_0:A$ we have to do a
-bit more work. Note that
-\begin{align*}
-& \subst{x_2}{{x_1}{{x_0}{\jcomp{}{\omega_0}{\jcomp{}{\epsilon_0}{\jcomp{}{\epsilon_0}{P}}}}}}
-  \\
-& \jdeq
-  \subst{x_2}{{x_1}{{x_0}{\unfold{\jcomp{\ctxwk{P}{P}}{\omega_0}{\jcomp{}{\epsilon_0}{\jcomp{}{\epsilon_0}{P}}}}}}}
-  \tag{by definition}
-  \\
-& \jdeq
-  \subst
-    { {x_2}
-      { {x_1}
-        { {x_0}
-          {\omega_0}
-          }
-        }
-      }
-    { {x_2}
-      { {x_1}
-        { {x_0}
-          {\ctxwk{{P}{P}}{\jcomp{}{\epsilon_0}{\jcomp{}{\epsilon_0}{P}}}}
-          }
-        }
-      }
-  \tag{by \autoref{comp-ss-f}}
-  \\
-& \jdeq
-  \subst
-    { {x_2}
-      { {x_1}
-        { {x_0}
-          {\omega_0}
-          }
-        }
-      }
-    { {x_2}
-      { \ctxwk
-          {\subst{x_1}{\ctxwk{\subst{x_0}{P}}{\subst{x_0}{P}}}}
-          {\subst{x_1}{{x_0}{\jcomp{}{\epsilon_0}{\jcomp{}{\epsilon_0}{P}}}}}
-        }
-      }
-  \tag{by \autoref{comp-sw-f}}
-  \\
-& \jdeq
-  \subst
-    { {x_2}
-      { {x_1}
-        { {x_0}
-          {\omega_0}
-          }
-        }
-      }
-    { {x_1}
-      { {x_0}
-        {\jcomp{}{\epsilon_0}{\jcomp{}{\epsilon_0}{P}}}
-        }
-      }
-  \tag{by \autoref{cancellation-ws-f}}
-  \\
-& \jdeq
-  \subst
-    { {{x_2}{{x_1}{{x_0}{\omega_0}}}}
-      {{{x_1}{{x_0}{\epsilon_0}}}{\epsilon_0}}
-      }
-    { P
-      }.
-  \tag{by \autoref{comp-ss-f}}
-\end{align*}
-Therefore, we see that
-\begin{equation*}
-\jterm
-  {{\Gamma}{\subst{{x_2}{{x_0}{\epsilon_0}}}{P}}}
-  { \subst
-      { {{x_2}{{x_1}{{x_0}{\omega_0}}}}
-        {{{x_1}{{x_0}{\epsilon_0}}}{\epsilon_0}}
-        }
-      { P
-        }
-    }
-  { \subst{x_2}{{x_1}{{x_0}{\omega_1}}}
-    }
-\end{equation*}
-Thus, the term $\subst{x_2}{{x_1}{{x_0}{\omega_1}}}$ takes a family over the
-extended context $\subst{x_2}{{x_0}{\epsilon_0}}$ to a family over the
-weakening $\subst{x_2}{{x_1}{{x_0}{\omega_0}}}$, which is itself a family
-over $\subst{x_1}{{x_0}{\epsilon_0}}$. In other words, $\omega_1$
-is precisely the internalization of the action on families of weakening, as
-intended.
-
+\begin{comment}
 Because the pair $(\omega_0,\omega_1)$ is required to be an extension homomorphism,
@@
 \end{equation*}
-
-\subsection{Extension-substitution algebras}
+\end{comment}
+
+\subsection{Projection algebras}
+
+\subsection{Substitution algebras}
 
@@
 
+\subsection{E-algebras}
+
 \subsection{Pre-universes}

### 2014-06-28 article-models-ttinternal.tex
working on cft-algebras

@@
 
-\begin{thm}
+\begin{thm}\label{thm:extalg-wk}
 Let $\extalg{Q}$ be an extension algebra in context $\ctxext{\Gamma}{B}$ and let
@@
 
-\begin{thm}
+\begin{thm}\label{thm:extalg-subst}
 Let $\extalg{Q}$ be an extension algebra in context $\ctxext{{\Gamma}{A}}{P}$
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Extension homomorphisms}
@@
 
-\subsection{Term-algebras}
-In studying extension algebras we had no need to consider terms. However, we
-do need to consider terms if we want a full interpretation of the theory of
-contexts, families and terms. Therefore, we shall now consider algebras where
-we have a further family 
-$\jfam{{{\Gamma}{A}}{\cftalgf{\cftalg{A}}}}{\cftalgt{\cftalg{A}}}$, assigning
-terms to families of $\cftalg{A}$. The notion of term-algebras that we will
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{CFT-algebras}
+The notion of CFT-algebras that we will
 study in this subsection will also have an empty object $(\cftempc{\cftalg{A}},
@@
 are just the terms of the underlying family $\cftalgc{\cftalg{A}}$ of contexts
-of the term-algebra $\cftalg{A}$. As with the theory of contexts, families and
-terms, this enables us to speak of terms of contexts of $\cftalg{A}$, and this
-will provide the categorical structure for $\cftalg{A}$ when we also consider
-internal weakening, substitution and identity terms.
-
+of the term-algebra $\cftalg{A}$. We will begin with an auxilary lemma which
+will help us dealing with the internal empty contexts:
+
+\begin{lem}\label{lem:empalg-mor}
+Consider a family $\jfam{\Gamma}{A}$ with a term $\jterm{\Gamma}{A}{x}$ and 
+a family $\jfam{{\Gamma}{A}}{P}$ % with a term $\jterm{{\Gamma}{A}}{P}{y}$
+satisfying the judgmental equality
+\begin{align}
+\jfameq*{\Gamma}{\subst{x}{P}}{A}
+  \label{eq1:empalg-mor}
+%  \\
+%\jtermeq*{\Gamma}{A}{\subst{x}{y}}{x}.
+%  \label{eq2:empalg-mor}
+\end{align}
+Consider also a morphism
+\begin{equation*}
+\jhom{\Gamma}{{A}{P}}{B}{f}.
+\end{equation*}
+Then $\subst{\xi}{f}$ is a morphism from $A$ to $B$ 
+in context $\Gamma$ and the following inference rules are valid for $\xi$ being 
+$x$ or $y$:
+\begin{align*}
+& \inference
+  { \jfam{{\Gamma}{B}}{Q}
+    }
+  { \jfameq
+      {{\Gamma}{A}}
+      {\subst{\xi}{\jcomp{}{f}{Q}}}
+      {\jcomp{}{\subst{\xi}{f}}{Q}}
+    }
+  \\
+& \inference
+  { \jfam{{{\Gamma}{B}}{Q}}{R}
+    }
+  { \jfameq
+      {{{\Gamma}{A}}{\jcomp{}{\subst{\xi}{f}}{Q}}}
+      {\subst{\xi}{\jcomp{}{f}{R}}}
+      {\jcomp{}{\subst{\xi}{f}}{R}}
+    }
+  \\
+& \inference
+  { \jterm{{{\Gamma}{B}}{Q}}{R}{h}
+    }
+  { \jtermeq
+      {{{\Gamma}{A}}{\jcomp{}{\subst{\xi}{f}}{Q}}}
+      {\jcomp{}{\subst{\xi}{f}}{R}}
+      {\subst{\xi}{\jcomp{}{f}{h}}}
+      {\jcomp{}{\subst{\xi}{f}}{h}}
+    }
+\end{align*}
+\end{lem}
+
+\begin{proof}
+We only prove the first inference rule in both cases.
+Let $Q$ be a family over $\ctxext{\Gamma}{\cftalgc{\cftalg{A}}}$. In the case 
+where $\xi$ is $x$ we have the judgmental equalities
+\begin{align*}
+\subst{x}{\jcomp{}{f}{Q}}
+& \jdeq
+  \subst
+    {x}
+    {{f}{\ctxwk{\ctxext{\cftalgc{\cftalg{A}}}{\cftalgf{\cftalg{A}}}}{Q}}}
+  \tag{by definition}
+  \\
+& \jdeq
+  \subst
+    { {x}{f}
+      }
+    { {x}
+      {\ctxwk{\ctxext{\cftalgc{\cftalg{A}}}{\cftalgf{\cftalg{A}}}}{Q}}
+      }
+  \tag{by \autoref{comp-ss-f}}
+  \\
+& \jdeq
+  \subst
+    { {x}{f}
+      }
+    { {x}
+      {\ctxwk{\cftalgf{\cftalg{A}}}{{\cftalgc{\cftalg{A}}}{Q}}}
+      }
+  \tag{by \autoref{comp-ew-f}}
+  \\
+& \jdeq
+  \subst
+    { {x}{f}
+      }
+    { \ctxwk
+        {\subst{x}{\cftalgf{\cftalg{A}}}}
+        {\subst{x}{\ctxwk{\cftalgc{\cftalg{A}}}{Q}}}
+      }
+  \tag{by \autoref{comp-sw-f}}
+  \\
+& \jdeq
+  \subst
+    { {x}{f}
+      }
+    { \ctxwk
+        {\subst{x}{\cftalgf{\cftalg{A}}}}
+        {Q}
+      }
+  \tag{by \autoref{cancellation-ws-f}}
+  \\
+& \jdeq
+  \subst{{x}{f}}{\ctxwk{\cftalgc{\cftalg{A}}}{Q}}
+  \tag{by \autoref{eq1:empalg-mor}}
+  \\
+& \jdeq
+  \jcomp{}{\subst{x}{f}}{Q}.
+  \tag{by definition}
+\end{align*}
+In the case where $\xi$ is $y$ we have the judgmental equalities
+\begin{align*}
+\subst{y}{\jcomp{}{f}{Q}}
+& \jdeq
+  \subst
+    { y
+      }
+    { {f}
+      {\ctxwk{\ctxext{\cftalgc{\cftalg{A}}}{\cftalgf{\cftalg{A}}}}{Q}}
+      }
+  \tag{by definition}
+  \\
+& \jdeq
+  \subst
+    { {y}{f}
+      }
+    { {y}
+      {\ctxwk{\ctxext{\cftalgc{\cftalg{A}}}{\cftalgf{\cftalg{A}}}}{Q}}
+      }
+  \tag{by \autoref{comp-ss-f}}
+  \\
+& \jdeq
+  \subst
+    { {y}{f}
+      }
+    { {y}
+      {\ctxwk{\cftalgf{\cftalg{A}}}{{\cftalgc{\cftalg{A}}}{Q}}}
+      }
+  \tag{by \autoref{comp-ew-f}}
+  \\
+& \jdeq
+  \subst
+    { {y}{f}
+      }
+    { \ctxwk{\cftalgc{\cftalg{A}}}{Q}
+      }
+  \tag{by \autoref{cancellation-ws-f}}
+  \\
+& \jdeq
+  \jcomp{}{\subst{y}{f}}{Q}.
+  \tag{by definition}
+\end{align*}
+\end{proof}
+
+\begin{defn}
+A \emph{CFT-algebra $\cftalg{A}$ in context $\Gamma$} is a septuple
+\begin{equation*}
+\unfold{\cftalg{A}}
+\end{equation*}
+where the quintuple $\unfold{\extalg{A}}$ is an extension
+algebra in context $\Gamma$, and where
+\begin{align*}
+\jterm*{\Gamma}{\cftalgc{\cftalg{A}}}{\cftempc{\cftalg{A}}}
+  \\
+\jterm*{{\Gamma}{\cftalgc{\cftalg{A}}}}{\cftalgf{\cftalg{A}}}{\cftempf{\cftalg{A}}}
+\end{align*}
+which satisfy the following judgmental equalities
+\begin{enumerate}
+\item Families over the empty context in $\cftalg{A}$ are contexts of $\cftalg{A}$:
+\begin{align}
+\jfameq*
+  {\Gamma}
+  {\subst{\cftempc{\cftalg{A}}}{\cftalgf{\cftalg{A}}}}
+  {\cftalgc{\cftalg{A}}}
+  \label{empalg-eq1}
+\end{align}
+\item The empty family over the empty context of $\cftalg{A}$ is the empty context of $\cftalg{A}$:
+\begin{align}
+\jtermeq*
+  {\Gamma}
+  {\cftalgc{\cftalg{A}}}
+  {\subst{\cftempc{\cftalg{A}}}{\cftempf{\cftalg{A}}}}
+  {\cftempc{\cftalg{A}}}.
+  \label{empalg-eq2}
+\end{align}
+\item Context extension of $\cftalg{A}$ is compatible with the empty context and
+family of $\cftalg{A}$:
+\begin{align}
+\jtermeq*
+  {{\Gamma}{\cftalgc{\cftalg{A}}}}
+  {\ctxwk{\cftalgc{\cftalg{A}}}{\cftalgc{\cftalg{A}}}}
+  {\subst{\cftempc{\cftalg{A}}}{\cftctxext[\cftalg{A}]}}
+  {\idtm{\cftalgc{\cftalg{A}}}}
+  \label{cftalg-eq1}
+  \\
+\jtermeq*
+  {{\Gamma}{\cftalgc{\cftalg{A}}}}
+  {\ctxwk{\cftalgc{\cftalg{A}}}{\cftalgc{\cftalg{A}}}}
+  {\subst{\cftempf{\cftalg{A}}}{\cftctxext[\cftalg{A}]}}
+  {\idtm{\cftalgc{\cftalg{A}}}}
+  \label{cftalg-eq2}
+\end{align}
+\item Family extension of $\cftalg{A}$ is compatible with the empty family and the empty family
+over families of $\cftalg{A}$:
+\begin{align}
+\jtermeq*
+  {{{\Gamma}{\cftalgc{\cftalg{A}}}}{\cftalgf{\cftalg{A}}}}
+  {\ctxwk{\cftalgf{\cftalg{A}}}{\cftalgf{\cftalg{A}}}}
+  {\subst{\cftempf{\cftalg{A}}}{\cftfamext}}
+  {\idtm{\cftalgf{\cftalg{A}}}}
+  \label{cftalg-eq3}
+  \\
+\jtermeq*
+  {{{\Gamma}{\cftalgc{\cftalg{A}}}}{\cftalgf{\cftalg{A}}}}
+  {\ctxwk{\cftalgf{\cftalg{A}}}{\cftalgf{\cftalg{A}}}}
+  {\subst{\jcomp{}{\cftctxext}{\cftempf{\cftalg{A}}}}{\cftfamext}}
+  {\idtm{\cftalgf{\cftalg{A}}}}
+  \label{cftalg-eq4}
+\end{align}
+\item Family extension over the empty context of $\cftalg{A}$ is context extension
+of $\cftalg{A}$:
+\begin{align}
+\jhomeq*
+  {\Gamma}
+  {{\cftalgc{\cftalg{A}}}{\cftalgf{\cftalg{A}}}}
+  {\cftalgc{\cftalg{A}}}
+  {\subst{\cftempc{\cftalg{A}}}{\cftfamext}}
+  {\cftctxext}
+  \label{cftalg-eq5}
+\end{align}
+\end{enumerate}
+\end{defn}
+
+\begin{rmk}\label{rmk:cftalg-defn}
+We need to verify that the judgmental equalities
+\autoref{cftalg-eq1,cftalg-eq2,cftalg-eq3,cftalg-eq4,cftalg-eq5}
+are indeed well-typed (i.e.~compare two terms of the same type).
+\begin{enumerate}
+\item \autoref{lem:empalg-mor} implies that the morphisms 
+$\subst{\cftempc{\cftalg{A}}}{\cftctxext}$ and
+$\subst{\cftempf{\cftalg{A}}}{\cftctxext}$ both go from $\cftalgc{\cftalg{A}}$ 
+to $\cftalgc{\cftalg{A}}$.
+\item \label{rmkenum:famfamempf-jdeq-fam}
+Now we have the judgmental equalities
+\begin{align*}
+\subst{\cftempf{\cftalg{A}}}{\jcomp{}{\cftctxext}{\cftalgf{\cftalg{A}}}}
+& \jdeq
+  \jcomp{}{\subst{\cftempf{\cftalg{A}}}{\cftctxext}}{\cftalgf{\cftalg{A}}}
+  \tag{by \autoref{lem:empalg-mor}}
+  \\
+& \jdeq
+  \jcomp{}{\idtm{\cftalgc{\cftalg{A}}}}{\cftalgf{\cftalg{A}}}
+  \tag{by \autoref{cftalg-eq1}}
+  \\
+& \jdeq
+  \cftalgf{\cftalg{A}}
+  \tag{by \autoref{precomp-idtm-c}}
+\end{align*}
+Therefore we can apply \autoref{lem:empalg-mor} with the triple
+$(\cftalgf{\cftalg{A}},\jcomp{}{\cftctxext}{\cftalgf{\cftalg{A}}},
+\cftempf{\cftalg{A}})$ and the morphism $\cftfamext$ to see that both 
+$\subst{\cftempf{\cftalg{A}}}{\cftfamext}$ and
+$\subst{\jcomp{}{\cftctxext}{\cftempf{\cftalg{A}}}}{\cftfamext}$
+are morphisms from $\cftalgf{\cftalg{A}}$ to $\cftalgf{\cftalg{A}}$.
+\item The morphism $\subst{\cftempc{\cftalg{A}}}{\cftfamext}$ goes from
+$ \subst
+    {\cftempc{\cftalg{A}}}
+    {\ctxext{\cftalgf{\cftalg{A}}}{\jcomp{}{\cftctxext}{\cftalgf{\cftalg{A}}}}}
+  $
+to $\subst{\cftempc{\cftalg{A}}}{\cftalgf{\cftalg{A}}}$. For the domain we
+have the judgmental equalities
+\begin{align*}
+  \subst
+    {\cftempc{\cftalg{A}}}
+    {\ctxext{\cftalgf{\cftalg{A}}}{\jcomp{}{\cftctxext}{\cftalgf{\cftalg{A}}}}}
+& \jdeq
+  \ctxext
+    {\subst{\cftempc{\cftalg{A}}}{\cftalgf{\cftalg{A}}}}
+    {\subst{\cftempc{\cftalg{A}}}{\jcomp{}{\cftctxext}{\cftalgf{\cftalg{A}}}}}
+  \tag{by \autoref{comp-se-c}}
+  \\
+& \jdeq
+  \ctxext
+    {\cftalgc{\cftalg{A}}}
+    {\subst{\cftempc{\cftalg{A}}}{\jcomp{}{\cftctxext}{\cftalgf{\cftalg{A}}}}}
+  \tag{by \autoref{empalg-eq1}}
+  \\
+& \jdeq
+  \ctxext
+    {\cftalgc{\cftalg{A}}}
+    {\cftalgf{\cftalg{A}}}
+  \tag{by \autoref{rmkenum:famfamempf-jdeq-fam} above}
+\end{align*}
+The codomain of $\subst{\cftempc{\cftalg{A}}}{\cftfamext}$
+is judgmentally equal to $\cftalgc{\cftalg{A}}$ by
+\autoref{empalg-eq1}.
+\end{enumerate}
+\end{rmk}
+
+\begin{comment}
 \begin{defn}
@@
 
-The following lemma presents a property of empty-object algebras that will get
-a lot of use.
-
-\begin{lem}\label{lem:empalg-mor}
-Consider a family $\jfam{\Gamma}{A}$ with a term $\jterm{\Gamma}{A}{x}$ and 
-a family $\jfam{{\Gamma}{A}}{P}$ with a term $\jterm{{\Gamma}{A}}{P}{y}$
-satisfying the judgmental equalities 
-\begin{align}
-\jfameq*{\Gamma}{\subst{x}{P}}{A}
-  \label{eq1:empalg-mor}
-  \\
-\jtermeq*{\Gamma}{A}{\subst{x}{y}}{x}.
-  \label{eq2:empalg-mor}
-\end{align}
-Consider also a morphism
-\begin{equation*}
-\jhom{\Gamma}{{A}{P}}{B}{f}.
-\end{equation*}
-Then $\subst{\xi}{f}$ is a morphism from $A$ to $B$ 
-in context $\Gamma$ and the following inference rules are valid for $\xi$ being 
-$x$ or $y$:
-\begin{align*}
-& \inference
-  { \jfam{{\Gamma}{B}}{Q}
-    }
-  { \jfameq
-      {{\Gamma}{A}}
-      {\subst{\xi}{\jcomp{}{f}{Q}}}
-      {\jcomp{}{\subst{\xi}{f}}{Q}}
-    }
-  \\
-& \inference
-  { \jfam{{{\Gamma}{B}}{Q}}{R}
-    }
-  { \jfameq
-      {{{\Gamma}{A}}{\jcomp{}{\subst{\xi}{f}}{Q}}}
-      {\subst{\xi}{\jcomp{}{f}{R}}}
-      {\jcomp{}{\subst{\xi}{f}}{R}}
-    }
-  \\
-& \inference
-  { \jterm{{{\Gamma}{B}}{Q}}{R}{h}
-    }
-  { \jtermeq
-      {{{\Gamma}{A}}{\jcomp{}{\subst{\xi}{f}}{Q}}}
-      {\jcomp{}{\subst{\xi}{f}}{R}}
-      {\subst{\xi}{\jcomp{}{f}{h}}}
-      {\jcomp{}{\subst{\xi}{f}}{h}}
-    }
-\end{align*}
-\end{lem}
-
-\begin{proof}
-We only prove the first inference rule in both cases.
-Let $Q$ be a family over $\ctxext{\Gamma}{\cftalgc{\cftalg{A}}}$. In the case 
-$i=\mathsf{c}$ we have the judgmental equalities
-\begin{align*}
-\subst{x}{\jcomp{}{f}{Q}}
-& \jdeq
-  \subst
-    {x}
-    {{f}{\ctxwk{\ctxext{\cftalgc{\cftalg{A}}}{\cftalgf{\cftalg{A}}}}{Q}}}
-  \tag{by definition}
-  \\
-& \jdeq
-  \subst
-    { {x}{f}
-      }
-    { {x}
-      {\ctxwk{\ctxext{\cftalgc{\cftalg{A}}}{\cftalgf{\cftalg{A}}}}{Q}}
-      }
-  \tag{by \autoref{comp-ss-f}}
-  \\
-& \jdeq
-  \subst
-    { {x}{f}
-      }
-    { {x}
-      {\ctxwk{\cftalgf{\cftalg{A}}}{{\cftalgc{\cftalg{A}}}{Q}}}
-      }
-  \tag{by \autoref{comp-ew-f}}
-  \\
-& \jdeq
-  \subst
-    { {x}{f}
-      }
-    { \ctxwk
-        {\subst{x}{\cftalgf{\cftalg{A}}}}
-        {\subst{x}{\ctxwk{\cftalgc{\cftalg{A}}}{Q}}}
-      }
-  \tag{by \autoref{comp-sw-f}}
-  \\
-& \jdeq
-  \subst
-    { {x}{f}
-      }
-    { \ctxwk
-        {\subst{x}{\cftalgf{\cftalg{A}}}}
-        {Q}
-      }
-  \tag{by \autoref{cancellation-ws-f}}
-  \\
-& \jdeq
-  \subst{{x}{f}}{\ctxwk{\cftalgc{\cftalg{A}}}{Q}}
-  \tag{by \autoref{eq1:empalg-mor}}
-  \\
-& \jdeq
-  \jcomp{}{\subst{x}{f}}{Q}.
-  \tag{by definition}
-\end{align*}
-In the case $i=\mathsf{f}$ we have the judgmental equalities
-\begin{align*}
-\subst{y}{\jcomp{}{f}{Q}}
-& \jdeq
-  \subst
-    { y
-      }
-    { {f}
-      {\ctxwk{\ctxext{\cftalgc{\cftalg{A}}}{\cftalgf{\cftalg{A}}}}{Q}}
-      }
-  \tag{by definition}
-  \\
-& \jdeq
-  \subst
-    { {y}{f}
-      }
-    { {y}
-      {\ctxwk{\ctxext{\cftalgc{\cftalg{A}}}{\cftalgf{\cftalg{A}}}}{Q}}
-      }
-  \tag{by \autoref{comp-ss-f}}
-  \\
-& \jdeq
-  \subst
-    { {y}{f}
-      }
-    { {y}
-      {\ctxwk{\cftalgf{\cftalg{A}}}{{\cftalgc{\cftalg{A}}}{Q}}}
-      }
-  \tag{by \autoref{comp-ew-f}}
-  \\
-& \jdeq
-  \subst
-    { {y}{f}
-      }
-    { \ctxwk{\cftalgc{\cftalg{A}}}{Q}
-      }
-  \tag{by \autoref{cancellation-ws-f}}
-  \\
-& \jdeq
-  \jcomp{}{\subst{y}{f}}{Q}.
-  \tag{by definition}
-\end{align*}
-\end{proof}
-
 A term-algebra $\cftalg{A}$ in context $\Gamma$ is going to be a septuple
@@
 \jtermeq{{\Gamma}{\cftalgc{\cftalg{A}}}}{\ctxwk{\cftalgc{\cftalg{A}}}{\cftalgc{\cftalg{A}}}}{\subst{\cftempc{\cftalg{A}}}{\cftctxext[\cftalg{A}]}}{\idtm{\cftalgc{\cftalg{A}}}}
-  \label{extempalg-eq1}
+  \label{cftalg-eq1}
 \end{equation}
@@
 \jtermeq*{{\Gamma}{\cftalgc{\cftalg{A}}}}{\ctxwk{\cftalgc{\cftalg{A}}}{\cftalgc{\cftalg{A}}}}{\subst{\cftempc{\cftalg{A}}}{\cftctxext[\cftalg{A}]}}{\idtm{\cftalgc{\cftalg{A}}}}
-  \label{extempalg-eq1}
+  \label{cftalg-eq1}
   \\
 \jtermeq*{{\Gamma}{\cftalgc{\cftalg{A}}}}{\ctxwk{\cftalgc{\cftalg{A}}}{\cftalgc{\cftalg{A}}}}{\subst{\cftempf{\cftalg{A}}}{\cftctxext[\cftalg{A}]}}{\idtm{\cftalgc{\cftalg{A}}}}
-  \label{extempalg-eq2}
+  \label{cftalg-eq2}
 \end{align}
@@
 we use \autoref{lem:empalg-mor}. Note that in combination with 
-\autoref{extempalg-eq1} and \autoref{extempalg-eq2} we get the judgmental
+\autoref{cftalg-eq1} and \autoref{cftalg-eq2} we get the judgmental
 equalities
@@
 can require that
-\begin{equation}\label{extempalg-eq3}
+\begin{equation}\label{cftalg-eq5}
 \jhomeq{\Gamma}{{A}{P}}{A}{\subst{\phi_0}{\epsilon_1}}{\epsilon_0}.
@@
 so we see that it indeed makes sense to require that
-\begin{equation}\label{extempalg-eq4}
+\begin{equation}\label{cftalg-eq3}
 \jtermeq{{{\Gamma}{A}}{P}}{\ctxwk{P}{P}}{\subst{\phi_1}{\epsilon_1}}{\idtm{P}}.
@@
 Thus we are allowed to require that
-\begin{equation}\label{extempalg-eq5}
+\begin{equation}\label{cftalg-eq4}
 \subst{\jcomp{}{\epsilon_0}{\phi_1}}{\epsilon_1}.
@@
 $\Gamma$, satisfying the judgmental equalities 
-\autoref{extempalg-eq1,extempalg-eq2,extempalg-eq3,extempalg-eq4,extempalg-eq5}.
+\autoref{cftalg-eq1,cftalg-eq2,cftalg-eq5,cftalg-eq3,cftalg-eq4}.
 \end{defn}
-
-\begin{thm}
-Let $(A,P,\epsilon_0,\epsilon_1,\phi_0,\phi_1)$ be an extension-empty algebra
-in context $\Gamma$. Then the sextuple
-\begin{equation*}
-(P,\jcomp{}{\epsilon_0}{P},\epsilon_1,\jcomp{}{\epsilon_0}{\epsilon_1},\phi_1,
-\jcomp{}{\epsilon_0}{\phi_1})
-\end{equation*}
-is an extension-empty algebra in context $\ctxext{\Gamma}{A}$. 
+\end{comment}
+
+We can extend the results of 
+\autoref{thm:extalg-fam,thm:extalg-wk,thm:extalg-subst} to CFT-algebras.
+
+\begin{thm}\label{thm:cftalg-fam}
+Let $\cftalg{A}$ be a CFT-algebra
+in context $\Gamma$. Then the septuple
+\begin{equation*}
+\cftalg{F_A}
+  \defeq
+  ( \cftalgf{\cftalg{A}},
+    \jcomp{}{\cftctxext}{\cftalgf{\cftalg{A}}},
+    \jcomp{}{\cftctxext}{\cftalgt{\cftalg{A}}},
+    \cftfamext,
+    \jcomp{}{\cftctxext}{\cftfamext},
+    \cftempf{\cftalg{A}},
+    \jcomp{}{\cftctxext}{\cftempf{\cftalg{A}}})
+\end{equation*}
+is a CFT-algebra in context $\ctxext{\Gamma}{\cftalgc{\cftalg{A}}}$. 
 \end{thm}
@@
 \begin{proof}
-We first check whether the quadruple $(P,\jcomp{}{\epsilon_0}{P},\phi_1,\jcomp{}{\epsilon_0}{\phi_1})$
-is an empty algebra in context $\ctxext{\Gamma}{A}$. Thus we need to verify
-that $\subst{\phi_1}{\jcomp{}{\epsilon_0}{P}}\jdeq P$ and that
-$\subst{\phi_1}{\jcomp{}{\epsilon_0}{\phi_1}}\jdeq\phi_1$. Both these judgmental
-equalities follow from \autoref{lem:empalg-mor}.
-
-To complete the proof, we need to verify that
-\autoref{extempalg-eq1,extempalg-eq2,extempalg-eq3,extempalg-eq4,extempalg-eq5} hold in our
-current setting. More precisely, we have to verify that
-\begin{align*}
-\subst{\phi_1}{\epsilon_1}
-& \jdeq 
-  \idtm{P}
-  \\
-\subst{\jcomp{}{\epsilon_0}{\phi_1}}{\epsilon_1}
-& \jdeq 
-  \idtm{P}
-  \\
-\subst{\phi_1}{\jcomp{}{\epsilon_0}{\epsilon_1}}
-& \jdeq
-  \epsilon_1
-  \\
-\subst{\jcomp{}{\epsilon_0}{\phi_1}}{\jcomp{}{\epsilon_0}{\epsilon_1}}
-& \jdeq
-  \idtm{\jcomp{}{\epsilon_0}{P}}
-  \\
-\subst{\jcomp{}{\epsilon_0}{\jcomp{}{\epsilon_0}{\phi_1}}}{\jcomp{}{\epsilon_0}{\epsilon_1}}
-& \jdeq
-  \idtm{\jcomp{}{\epsilon_0}{P}}
-\end{align*}
-None of these present any difficulties.
+\begin{enumerate}
+\item The judgmental equality
+\begin{equation*}
+\jfameq{{\Gamma}{\cftalgc{\cftalg{A}}}}{\subst{\cftempf{\cftalg{A}}}{\jcomp{}{\cftctxext}{\cftalgf{\cftalg{A}}}}}{\cftalgf{\cftalg{A}}}
+\end{equation*}
+was verified in \autoref{rmkenum:famfamempf-jdeq-fam} of \autoref{rmk:cftalg-defn}.
+\item Now we verify the judgmental equality
+\begin{equation*}
+\jtermeq
+  {{\Gamma}{\cftalgc{\cftalg{A}}}}
+  {\cftalgf{\cftalg{A}}}
+  {\subst{\cftempf{\cftalg{A}}}{\jcomp{}{\cftctxext}{\cftempf{\cftalg{A}}}}}
+  {\cftempf{\cftalg{A}}}.
+\end{equation*}
+Note that we can apply \autoref{lem:empalg-mor}, so we get the judgmental
+equalities
+\begin{align*}
+\subst{\cftempf{\cftalg{A}}}{\jcomp{}{\cftctxext}{\cftempf{\cftalg{A}}}}
+& \jdeq
+  \jcomp{}{\subst{\cftempf{\cftalg{A}}}{\cftctxext}}{\cftempf{\cftalg{A}}}
+  \tag{by \autoref{lem:empalg-mor}}
+  \\
+& \jdeq
+  \jcomp{}{\idtm{\cftalgf{\cftalg{A}}}}{\cftempf{\cftalg{A}}}
+  \tag{by \autoref{cftalg-eq3}}
+  \\
+& \jdeq
+  \cftempf{\cftalg{A}}.
+  \tag{by \autoref{precomp-idtm-t}}
+\end{align*}
+\item The judgmental equalities
+\begin{align*}
+\subst{\cftempf{\cftalg{A}}}{\cftfamext}
+& \jdeq 
+  \idtm{\cftalgf{\cftalg{A}}}
+  \\
+\subst{\jcomp{}{\cftctxext}{\cftempf{\cftalg{A}}}}{\cftfamext}
+& \jdeq 
+  \idtm{\cftalgf{\cftalg{A}}}
+\end{align*}
+are given by assumption.
+\item The judgmental equalities
+\begin{align*}
+\subst{\jcomp{}{\cftctxext}{\cftempf{\cftalg{A}}}}{\jcomp{}{\cftctxext}{\cftfamext}}
+& \jdeq
+  \idtm{\jcomp{}{\cftctxext}{\cftalgf{\cftalg{A}}}}
+  \\
+\subst{\jcomp{}{\cftfamext}{\jcomp{}{\cftctxext}{\cftempf{\cftalg{A}}}}}{\jcomp{}{\cftctxext}{\cftfamext}}
+& \jdeq
+  \idtm{\jcomp{}{\cftctxext}{\cftalgf{\cftalg{A}}}}
+\end{align*}
+follow from the judgmental equalities
+\begin{align*}
+\subst{\jcomp{}{\cftctxext}{\cftempf{\cftalg{A}}}}{\jcomp{}{\cftctxext}{\cftfamext}}
+& \jdeq
+  \jcomp{}{\cftctxext}{\subst{\cftempf{\cftalg{A}}}{\cftfamext}}
+  \tag{by \autoref{lem:jcomp-subst}}
+  \\
+& \jdeq
+  \jcomp{}{\cftctxext}{\idtm{\cftalgf{\cftalg{A}}}}
+  \tag{by \autoref{cftalg-eq3}}
+  \\
+& \jdeq
+  \idtm{\jcomp{}{\cftctxext}{\cftalgf{\cftalg{A}}}}
+  \tag{by \autoref{comp-wi-t}}
+\end{align*}
+and
+\begin{align*}
+\subst
+  {\jcomp{}{\cftfamext}{\jcomp{}{\cftctxext}{\cftempf{\cftalg{A}}}}}
+  {\jcomp{}{\cftctxext}{\cftfamext}}
+& \jdeq
+  \subst
+    {\jcomp{}{\cftctxext}{\jcomp{}{\cftctxext}{\cftempf{\cftalg{A}}}}}
+    {\jcomp{}{\cftctxext}{\cftfamext}}
+  \tag{by \autoref{lem:extalg-twins}}
+  \\
+& \jdeq
+  \jcomp
+    {}
+    {\cftctxext}
+    {\subst{\jcomp{}{\cftctxext}{\cftempf{\cftalg{A}}}}{\cftfamext}}
+  \tag{by \autoref{comp-sw-t}}
+  \\
+& \jdeq
+  \jcomp
+    {}
+    {\cftctxext}
+    {\idtm{\cftalgf{\cftalg{A}}}}
+  \tag{by \autoref{cftalg-eq4}}
+  \\
+& \jdeq
+  \idtm{\jcomp{}{\cftctxext}{\cftalgf{\cftalg{A}}}}.
+  \tag{by \autoref{comp-wi-t}}
+\end{align*}
+\item Finally, the judgmental equality
+\begin{equation*}
+\subst{\cftempf{\cftalg{A}}}{\jcomp{}{\cftctxext}{\cftfamext}}
+  \jdeq
+  \cftfamext
+\end{equation*}
+is verified as follows:
+\begin{align*}
+\subst{\cftempf{\cftalg{A}}}{\jcomp{}{\cftctxext}{\cftfamext}}
+& \jdeq
+  \jcomp{}{\subst{\cftempf{\cftalg{A}}}{\cftctxext}}{\cftfamext}
+  \tag{by \autoref{lem:empalg-mor}}
+  \\
+& \jdeq
+  \jcomp{}{\idtm{\cftalgc{\cftalg{A}}}}{\cftfamext}
+  \tag{by \autoref{cftalg-eq1}}
+  \\
+& \jdeq
+  \cftfamext.
+  \tag{by \autoref{precomp-idtm-t}}
+\end{align*}
+\end{enumerate}
 \end{proof}
@@
 \end{defn}
-
-\subsection{Term-algebras and term-operations}
-Until now we had no need to consider term-algebras. Extension is an operation
-letting contexts and families interact and the empty object is a term which
-requires that contexts and families are compatible. However, to model the
-theory of contexts, families and terms we do need to take terms into account.
-
-\begin{defn}
-A term-algebra $\cftalg{A}$ in context $\Gamma$ is a septuple 
-\begin{equation*}
-\unfold{\cftalg{A}}
-\end{equation*}
-such that the quadruple
-$\unfold{\extalg}{A}$ is an extension algebra in context $\Gamma$, 
-the quadruple $\unfold{\empalg{A}}$ is an empty algebra in context $\Gamma$,
-for which we additionally have the following judgmental equality:
-\begin{align*}
-\jfameq*{{\Gamma}{\cftalgc{\cftalg{A}}}}{\subst{\cftempf{\cftalg{A}}}{\cftalgt{\cftalg{A}}}}{\emptyf}.
-\end{align*}
-\end{defn}
-
-\begin{thm}
-Let $\mathcal{A}\defeq(A,P,T,\epsilon_0,\epsilon_1,\phi_0,\phi_1)$ be a term-algebra in context
-$\Gamma$. Then the septuple
-\begin{equation*}
-\mathcal{F}_{\mathcal{A}}\defeq(P,\jcomp{}{\epsilon_0}{P},\jcomp{}{\epsilon_0}{T},\epsilon_1,\jcomp{}{\epsilon_0}{\epsilon_1},\phi_1,\jcomp{}{\epsilon_0}{\phi_1})
-\end{equation*}
-is a term-algebra in context $\ctxext{\Gamma}{A}$, called the family term-algebra
-of $\mathcal{A}$.
-\end{thm}
-
-\begin{thm}
-Let $\mathcal{A}\defeq(A,P,T,\epsilon_0,\epsilon_1,\phi_0,\phi_1)$ be a term-algebra in context
-$\Gamma$ and let $X$ be a family in context $\Gamma$. Then
-\begin{equation*}
-\ctxwk{X}{\mathcal{A}}\defeq(\ctxwk{X}{A},\ctxwk{X}{P},\ctxwk{X}{T},\ctxwk{X}{\epsilon_0},\ctxwk{X}{\epsilon_1},\ctxwk{X}{\phi_0},\ctxwk{X}{\phi_1})
-\end{equation*}
-is a term-algebra in context $\ctxext{\Gamma}{X}$.
-\end{thm}
-
-\begin{thm}
-Let $\mathcal{A}\defeq(A,P,T,\epsilon_0,\epsilon_1,\phi_0,\phi_1)$ be a term-algebra in context
-$\ctxext{\Gamma}{X}$ and let $\jterm{\Gamma}{X}{x}$. Then
-\begin{equation*}
-\subst{x}{\mathcal{A}}\defeq
-(\subst{x}{A},\subst{x}{P},\subst{x}{T},\subst{x}{\epsilon_0},\subst{x}{\epsilon_1},
-\subst{x}{\phi_0},\subst{x}{\phi_1})
-\end{equation*}
-is a term-algebra in context $\Gamma$.
-\end{thm}
-
-\begin{defn}
-Let $\cftalg{A}$ and $\cftalg{B}$ be term-algebras in context $\Gamma$. 
-A term-operation $\cfthom{f}$ from $\cftalg{A}$ to $\cftalg{B}$ in context
-$\Gamma$ is a triple $(\cfthomc{\cfthom{f}},\cfthomf{\cfthom{f}},\cfthomt{\cfthom{f}})$
-consisting of
-\begin{align*}
-\jhom*{\Gamma}{\cftalgc{\cftalg{A}}}{\cftalgc{\cftalg{B}}}{\cfthomc{\cfthom{f}}}
-  \\
-\jfhom*
-  {\Gamma}
-  {\cftalgc{\cftalg{A}}}
-  {\cftalgc{\cftalg{B}}}
-  {\cfthomc{\cfthom{f}}}
-  {\cftalgf{\cftalg{A}}}
-  {\cftalgf{\cftalg{B}}}
-  {\cfthomf{\cfthom{f}}}
-  \\
-\jfhom*
-  {\Gamma}
-  {{\cftalgc{\cftalg{A}}}{\cftalgf{\cftalg{A}}}}
-  {{\cftalgc{\cftalg{B}}}{\cftalgf{\cftalg{B}}}}
-  {\jvcomp{\cftalgf{\cftalg{A}}}{\cfthomc{\cfthom{f}}}{\cfthomf{\cfthom{f}}}}
-  {\cftalgt{\cftalg{A}}}
-  {\cftalgt{\cftalg{B}}}
-  {\cfthomt{\cfthom{f}}}
-\end{align*}
-such that the pair $(\cfthomc{\cfthom{f}},\cfthomf{\cfthom{f}})$ is both an 
-extension homomorphism from
-$(\cftalgc{\cftalg{A}},\cftalgf{\cftalg{A}},\cftctxext[\cftalg{A}],\cftfamext[\cftalg{A}])$ 
-to 
-$(\cftalgc{\cftalg{B}},\cftalgf{\cftalg{B}},\cftfamext[\cftalg{B}],\cftfamext[\cftalg{B}])$ 
-in context $\Gamma$
-and an empty homomorphism from 
-$(\cftalgc{\cftalg{A}},\cftalgf{\cftalg{A}},\cftempc{\cftalg{A}},\cftempf{\cftalg{A}})$
-to
-$(\cftalgc{\cftalg{B}},\cftalgf{\cftalg{B}},\cftempc{\cftalg{B}},\cftempf{\cftalg{B}})$
-in context $\Gamma$.
-\end{defn}
-
-Thus, when $\cfthom{f}$ is a term-operation from $\cftalg{A}$ to $\cftalg{B}$, we are
-in the situation of the following judgmentally commutative diagram:
-\begin{equation*}
-\begin{tikzcd}
-\cftalgt{\cftalg{A}}
-  \ar[fib]{d}
-  \ar{r}{\cfthomt{\cfthom{f}}}
-& \cftalgt{\cftalg{B}}
-  \ar[fib]{d}
-  \\
-\cftalgf{\cftalg{A}}
-  \ar[fib]{d}
-  \ar{r}{\cfthomf{\cfthom{f}}}
-& \cftalgf{\cftalg{B}}
-  \ar[fib]{d}
-  \\
-\cftalgc{\cftalg{A}}
-  \ar{r}{\cfthomc{\cfthom{f}}}
-& \cftalgc{\cftalg{B}}
-\end{tikzcd}
-\end{equation*}
 

### 2014-06-29 article-models-ttinternal.tex
small improvements

@@
 
-\subsection{Homorphisms of term-algebras}
+\subsection{CFT-homomorphisms}
 \begin{defn}
@@
 Let $\cftalg{A}$ be a term-algebra in context $\Gamma$. Weakening will be a
-homomorphism
+CFT-homomorphism
 \begin{equation}\label{cftwk}
@@
 \end{equation}
-satisfying judgmental equalities which express the compatibility with extension,
-the empty object and itself. Thus, the rules we interpret for weakening are
-the rules in sections \autoref{comp-we,comp-w0,comp-ww,comp-ew,comp-0w}. The
+satisfying judgmental equalities which express abstracted analogues of
+the rules in sections \autoref{comp-ww,comp-ew,comp-0w}. The
 definition of weakening algebras can be found in \autoref{sec:cftwkalg-defn}.

### 2014-06-29 article-models-ttinternal.tex
initiated the sections on identity terms, substitution (revised) and e-algebras

@@
 \subsection{Projection algebras}
+Projection algebras will be weakening algebras with additional terms implementing
+the identity terms. We call these algebras projection algebras because the
+identity terms can only be formulated in weakening algebras and together with
+weakening, the identity terms provide for all the projections. In a weakening
+algebra $\cftalg{A}$ in context $\Gamma$, identity terms are implemented by a term
+\begin{equation*}
+\jterm
+  {context}
+  {\subst{{\idtm{\cftalgf{\cftalg{A}}}}{\cftwkc{\cftalg{A}}}}{\cftalgt{\cftalg{F_{F_A}}}}}
+  {\cftidtm{\cftalg{A}}}
+\end{equation*}
+We will require additionally that identity terms are compatible with weakening.
+Thus we will formulate an abstraction of the rule stated in \autoref{comp-wi}.
 
 \subsection{Substitution algebras}
-
-Let $(A,P,T,\epsilon_0,\epsilon_1,\phi_0,\phi_1)$ be a term-algebra. To define
-substitution algebras, we will consider
-\begin{align*}
-\jhom*
-  {{{{\Gamma}{A}}{P}}{T}}
-  {\ctxwk{T}{\jcomp{}{\epsilon_0}{P}}}
-  {P}
-  {\sigma_0}
-  \\
-\jfhom*
-  {{{{\Gamma}{A}}{P}}{T}}
-  {\ctxwk{T}{\jcomp{}{\epsilon_0}{P}}}
-  {P}
-  {\sigma_0}
-  {\ctxwk{T}{\jcomp{}{\epsilon_0}{\jcomp{}{\epsilon_0}{P}}}}
-  {\jcomp{}{\epsilon_0}{P}}
-  {\sigma_1}
-  \\
-\jfhom*
-  {{{{\Gamma}{A}}{P}}{T}}
-  { \ctxext
-      {\ctxwk{T}{\jcomp{}{\epsilon_0}{P}}}
-      {\ctxwk{T}{\jcomp{}{\epsilon_0}{\jcomp{}{\epsilon_0}{P}}}}
-    }
-  {\ctxext{P}{\jcomp{}{\epsilon_0}{P}}}
-  {\jvcomp{}{\sigma_0}{\sigma_1}}
-  {\ctxwk{T}{\jcomp{}{\epsilon_0}{\jcomp{}{\epsilon_0}{T}}}}
-  {\jcomp{}{\epsilon_0}{T}}
-  {\sigma_2}
-\end{align*}
-for which the pair $(\sigma_0,\sigma_1)$ is required to be an extension
-homomorphism and the triple $(\sigma_0,\sigma_1,\sigma_2)$ is required to
-be an empty term-homomorphism.
+Substitution algebras will be CFT-algebras with an additional substitution 
+operation. Given a CFT-algebra $\cftalg{A}$ in context $\Gamma$, substitution 
+will be a CFT-homomorphism
+\begin{equation*}
+\jhom
+  {{{{\Gamma}{\cftalgc{\cftalg{A}}}}{\cftalgf{\cftalg{A}}}}{\cftalgt{\cftalg{A}}}}
+  {\ctxwk{\cftalgt{\cftalg{A}}}{\cftalg{F_{F_A}}}}
+  {\cftalgf{\cftalg{A}}}
+  {\cftsubst{\cftalg{A}}}
+\end{equation*}
+We will require additionally that substitution preserves itself. Thus, we will
+formulate an abstraction of the rules stated in \autoref{comp-ss}.
 
 \subsection{E-algebras}
+E-algebras will be CFT-algebras with a weakening operation, identity terms and
+a substitution operation making it a projection algebra and a substitution 
+algebra at the same time. Additionally the weakening operation, identity terms
+and the substitution operation are all required to be compatible with each
+other. Thus, we need to require the abstractions of the rules stated in
+\autoref{comp-sw,comp-ws,cancellation-ws,cancellation-i}.
 

### 2014-06-30 article-models-ttinternal.tex
improvement identity terms
attempting to write the rule for cftidtm
finally stated the rule that weakening is compatible with itself

@@
 The following diagram displays the compatibility of weakening with itself:
-\begin{equation*}\label{diag:cftwk-cftwk}
-\begin{tikzcd}[column sep=huge]
+\begin{equation}\label{diag:cftwk-cftwk}
+\begin{tikzcd}[column sep=10em]
 \ctxwk{\cftalgf{\cftalg{A}}}{{\cftalgf{\cftalg{F_A}}}{\cftalg{F_{F_A}}}}
@@
 & \ctxwk{\cftalgf{\cftalg{A}}}{\cftalg{F_{F_{F_A}}}}
-  \ar{d}{\jcomp{}{\cftctxext}{\boldsymbol{\mathcal{F}}_{\cftwk{\cftalg{A}}}}}
+  \ar{d}{\jcomp{}{\ctxwk{\cftalgf{\cftalg{A}}}{\cftfamext}}{\boldsymbol{\mathcal{F}}_{\cftwk{\cftalg{A}}}}}
   \\
 \ctxwk{{\cftalgf{\cftalg{A}}}{\cftalgf{\cftalg{F_A}}}}{\jcomp{}{\cftwkc{A}}{\cftalg{F_{F_{F_A}}}}}
-  \ar{r}[swap]{\jcomp{}{\cftwkc{A}}{\jcomp{}{\cftctxext}{\cftwk{A}}}}
+  \ar{r}[swap]{%
+    \jcomp{}
+      { \cftwkf{\cftalg{A}}}
+      { \jcomp{}
+          {\cftwkc{\cftalg{A}}}
+          { \jcomp{}
+              {\cftctxext}
+              { \jcomp{}
+                  {\cftctxext}
+                  {\cftwk{\cftalg{A}}}
+                }
+            }
+        }
+    }
 & \jcomp{}{\cftwkc{A}}{\cftalg{F_{F_{F_{F_A}}}}}
 \end{tikzcd}
-\end{equation*}
+\end{equation}
 This is a complicated diagram and it is not immediately clear that all of it
@@
 \end{equation*}
-be a homomorphism of term-algebras satisfying \autoref{diag:cftwk-cftwk}. Then
+be a homomorphism of term-algebras for which the diagram in
+\autoref{diag:cftwk-cftwk} commutes judgmentally. Then
 we have the judgmental equality
-\begin{equation*}
-\jtermeq
-  {}
-  {}
-  {}
-  {}
-\end{equation*}
-for $\gamma:A$, $x:\cftalgf{\cftalg{A}}$, $y_0:\cftalgf{\cftalg{A}}$ and
-$y_1:\cftalgf{\cftalg{F_A}}$.
+\begin{align*}
+& \jcomp{}
+  { \subst{q}{{{b}{{\gamma}{\cftctxext}}}{\cftwk{\cftalg{A}}}}
+    }
+  { \subst
+      {{q}{{b}{\cftfamext}}}
+      {{a}{{\gamma}{\boldsymbol{\mathcal{F}}_{\cftwk{\cftalg{A}}}}}}
+    }
+  \\
+& \jdeq
+  \jcomp{}
+    { \subst{b}{{a}{{\gamma}{\boldsymbol{\mathcal{F}}_{\cftwk{\cftalg{A}}}}}}
+      }
+    { \subst
+        { {q}{{b}{{a}{{\gamma}{\cftwkf{\cftalg{A}}}}}}}
+        { {{{b}{{a}{{\gamma}{\cftwkc{\cftalg{A}}}}}}{{{a}{{\gamma}{\cftctxext}}}{\cftctxext}}}
+          {\cftwk{\cftalg{A}}}
+          }
+      }
+\end{align*}
+of CFT-homomorphisms from $\subst{b}{{\gamma}{\cftalg{F_{F_A}}}}$ to
+$ \subst
+      { {q}{{b}{{a}{{\gamma}{\cftwkf{\cftalg{A}}}}}}
+        }
+      { {{b}{{a}{{\gamma}{\cftwkc{\cftalg{A}}}}}}
+        {{a}{{\gamma}{\cftalg{F_{F_{F_{F_{A}}}}}}}}
+        }$
+in context $\Gamma$, for $\gamma:A$, $a,b:\subst{\gamma}{\cftalgf{\cftalg{A}}}$ and
+$q:\subst{b}{{\gamma}{\cftalgf{\cftalg{F_A}}}}$.
 \end{lem}
@@
 \jterm
-  {context}
-  {\subst{{\idtm{\cftalgf{\cftalg{A}}}}{\cftwkc{\cftalg{A}}}}{\cftalgt{\cftalg{F_{F_A}}}}}
+  { { {\Gamma}
+      {\cftalgc{\cftalg{A}}}
+      }
+    { \cftalgf{\cftalg{A}}
+      }
+    }
+  {\subst{{\idtm{\cftalgf{\cftalg{A}}}}{\cftwkc{\cftalg{A}}}}{\cftalgt{\cftalg{F_A}}}}
   {\cftidtm{\cftalg{A}}}
@@
 Thus we will formulate an abstraction of the rule stated in \autoref{comp-wi}.
+
+\begin{equation}
+\jtermeq
+  { \xi}
+  { \xi}
+  { \subst
+      { \jcomp
+          { }
+          { \cftctxext}
+          { \cftidtm{\cftalg{A}}
+            }
+        }
+      { { { \idtm
+              { \jcomp
+                  { }
+                  { \cftctxext}
+                  { \cftalgf{\cftalg{A}}
+                    }
+                }
+            }
+            { \jcomp
+                { }
+                { \cftctxext
+                  }
+                { \cftwkc{\cftalg{A}}
+                  }
+              }
+          }
+        { \jcomp
+            { }
+            { \jcomp{}{\cftctxext}{\cftfamext}}
+            { \cftwkt{\cftalg{A}}}
+          }
+        }
+    }
+  { \xi}
+\end{equation}
 
@@
   {\ctxwk{\cftalgt{\cftalg{A}}}{\cftalg{F_{F_A}}}}
-  {\cftalgf{\cftalg{A}}}
+  {\cftalg{F_A}}
   {\cftsubst{\cftalg{A}}}

### 2014-07-01 article-models-ttinternal.tex
fixed the currying for weakening rule in section 3

@@
 \begin{lem}
-Let $\cftalg{A}$ be a term-algebra in context $\Gamma$ and let
+Let $\cftalg{A}$ be a CFT-algebra in context $\Gamma$ and let
 \begin{equation*}
@@
 \subsubsection{Currying for weakening}
-\begin{equation*}
+\begin{equation}\label{eq:cftwk-currying}
 \jhomeq
@@
   { \ctxwk
-      {\cftalgf{\cftalg{F_A}}}
-      {\cftalg{F_{F_A}}}
+      {\ctxext{\cftalgf{\cftalg{A}}}{\cftalgf{\cftalg{F_A}}}}
+      {\cftalg{F_A}}
     }
@@
     }
-  { \jcomp{}{\cftwk{\cftalg{A}}}{\jcomp{}{\cftctxext}{\cftwk{\cftalg{A}}}}
-    }
-\end{equation*}
+  { \jcomp{}{\ctxwk{\cftalgf{\cftalg{F_A}}}{\cftwk{\cftalg{A}}}}{\jcomp{}{\cftctxext}{\cftwk{\cftalg{A}}}}
+    }
+\end{equation}
+
+\begin{lem}
+Let $\cftalg{A}$ be a CFT-algebra in context gamma with a CFT-homomorphism 
+$ \jhom
+  {{{\Gamma}{\cftalgc{\cftalg{A}}}}{\cftalgf{\cftalg{A}}}}
+  {\ctxwk{\cftalgf{\cftalg{A}}}{\cftalg{F_A}}}
+  {\jcomp{}{\cftctxext}{\cftalg{F_A}}}
+  {\cftwk{\cftalg{A}}}
+$
+satisfying the judgmental equality of \autoref{eq:cftwk-currying}. Then we
+have the judgmental equality
+\begin{equation*}
+\subst{{p}{{a}{\cftfamext}}}{{\gamma}{\cftwk{\cftalg{A}}}}
+\jdeq
+\jcomp{}
+    {\subst{a}{{\gamma}{\cftwk{\cftalg{A}}}}}
+    {\subst{p}{{{a}{{\gamma}{\cftctxext}}}{\cftwk{\cftalg{A}}}}}
+\end{equation*}
+of CFT-homomorphisms from $\subst{\gamma}{\cftalg{F_A}}$ to 
+$\subst{p}{{a}{{\gamma}{\cftalg{F_{F_{F_A}}}}}}$ in context $\Gamma$,
+for every $\gamma:A$, $a:\subst{\gamma}{\cftalgf{\cftalg{A}}}$ and
+$p:\subst{a}{{\gamma}{\cftalgf{\cftalg{F_A}}}}$. 
+\end{lem}
 

### 2014-07-02 article-models-ttinternal.tex
lots of small changes

--- 2014-06-30 article-models-ttinternal.tex
@@
 \begin{lem}
-Let $\extalg{A}$ be an extension algebra in context $\Gamma$.
-Then $\cftctxext[\extalg{A}]$ and $\cftfamext[\extalg{A}]$ 
-correctly interpret the rules for context
-extension and family extension.
-
-More precisely, for $x_0:\cftalgc{\cftalg{A}}$, 
-$x_1:\subst{x_0}{\cftalgf{\cftalg{A}}}$ and 
-$x_2:\subst{{x_1}{{x_0}{\cftctxext[\extalg{A}]}}}{\cftalgf{\cftalg{A}}}$ 
+For $\gamma:\cftalgc{\cftalg{A}}$, 
+$a:\subst{\gamma}{\cftalgf{\cftalg{A}}}$ and 
+$p:\subst{\apply\cftctxext[\extalg{A}]{\gamma}{a}}{\cftalgf{\cftalg{A}}}$ 
 we have the judgmental equality
 \begin{align*}
-\subst{{x_2}{{x_1}{\cftfamext[\extalg{A}]}}}{{x_0}{\cftctxext[\extalg{A}]}}
-& \jdeq
-  \subst
-    {x_2}
-    {{{x_1}{{x_0}{\cftctxext[\extalg{A}]}}}{\cftctxext[\extalg{A}]}}.
+\apply\cftctxext[\extalg{A}]{\gamma}{\apply\cftfamext[\extalg{A}]{\gamma}{a}{p}}
+& \jdeq
+  \apply\cftctxext[\extalg{A}]{\apply\cftctxext[\extalg{A}]{\gamma}{a}}{p}.
   \intertext{%
-and for $y_0:\cftalgf{\extalg{A}}$, 
-$y_1:\subst{y_0}{\jcomp{}{\cftfamext[\extalg{A}]}{\cftalgf{\extalg{A}}}}$ and 
-$y_2:
-  \subst
-    {y_1}
-    { {y_0}
-      { \jcomp
-          {}
-          {\cftfamext[\extalg{A}]}
-          {\jcomp{}{\cftfamext[\extalg{A}]}{\cftalgf{\extalg{A}}}}
-        }
-      }
-$
+and when we also have 
+$q: \subst
+      {\apply\cftctxext[\extalg{A}]{\apply\cftctxext[\extalg{A}]{\gamma}{a}}{p}}
+      {\cftalgf{\extalg{A}}}$, %
 we have the judgmental equality}
-\subst
-  {{y_2}{{y_1}{\jcomp{}{\cftfamext[\extalg{A}]}{\cftfamext[\extalg{A}]}}}}
-  {{y_0}{\cftfamext[\extalg{A}]}}
-& \jdeq
-  \subst{y_2}{{{y_1}{{y_0}{\cftfamext[\extalg{A}]}}}{\cftfamext[\extalg{A}]}}.
+\apply\cftfamext[\extalg{A}]{\gamma}{a}{%
+  \apply\cftfamext[\extalg{A}]{\apply\cftctxext[\extalg{A}]{\gamma}{a}}{p}{q}}
+& \jdeq
+  \apply\cftfamext[\extalg{A}]{\gamma}{%
+    \apply\cftfamext[\extalg{A}]{\gamma}{a}{p}}{q}.
 \end{align*}
@@
 \begin{align*}
-\subst{{x_2}{{x_1}{\cftfamext[\cftalg{A}]}}}{{x_0}{\cftctxext[\cftalg{A}]}}
-& \jdeq
-  \subst{\tmext{x_0}{\subst{x_2}{{x_1}{\cftfamext[\cftalg{A}]}}}}{\cftctxext[\cftalg{A}]}
-  \\
-& \jdeq 
-  \subst{{x_2}{{x_1}{{x_0}{\jvcomp{}{\idtm{\cftalgc{\cftalg{A}}}}{\cftfamext[\cftalg{A}]}}}}}{\cftctxext[\cftalg{A}]}
-  \\
-& \jdeq
-  \subst{x_2}{{x_1}{{x_0}{\jcomp{}{\jvcomp{}{\idtm{\cftalgc{\cftalg{A}}}}{\cftfamext[\cftalg{A}]}}{\cftctxext[\cftalg{A}]}}}}
-  \\
-& \jdeq
-  \subst{x_2}{{x_1}{{x_0}{\jcomp{}{\jvcomp{}{\cftctxext[\cftalg{A}]}{\idtm{\jcomp{}{\cftctxext[\cftalg{A}]}{\cftalgf{\cftalg{A}}}}}}{\cftctxext[\cftalg{A}]}}}}
-  \\
-& \jdeq 
-  \subst{{x_2}{{x_1}{{x_0}{\jvcomp{}{\cftctxext[\cftalg{A}]}{\idtm{\jcomp{}{\cftctxext[\cftalg{A}]}{\cftalgf{\cftalg{A}}}}}}}}}{\cftctxext[\cftalg{A}]}
-  \\
-& \jdeq
-  \subst{\tmext{\subst{x_1}{{x_0}{\cftctxext[\cftalg{A}]}}}{x_2}}{\cftctxext[\cftalg{A}]}
-  \\
-& \jdeq
-  \subst{x_2}{{{x_1}{{x_0}{\cftctxext[\cftalg{A}]}}}{\cftctxext[\cftalg{A}]}}.
+\apply\cftctxext[\extalg{A}]{\gamma}{\apply\cftfamext[\extalg{A}]{\gamma}{a}{p}}
+& \jdeq
+  \subst
+    {\tmext{\gamma}{\apply\cftfamext[\extalg{A}]{\gamma}{a}{p}}}
+    {\cftctxext[\cftalg{A}]}
+  \\
+& \jdeq 
+  \subst
+    { {p}
+      { {a}
+        { {\gamma}
+          {\jvcomp{}{\idtm{\cftalgc{\cftalg{A}}}}{\cftfamext[\cftalg{A}]}}
+          }
+        }
+      }
+    { \cftctxext[\cftalg{A}]}
+  \\
+& \jdeq
+  \subst{p}{{a}{{\gamma}{\jcomp{}{\jvcomp{}{\idtm{\cftalgc{\cftalg{A}}}}{\cftfamext[\cftalg{A}]}}{\cftctxext[\cftalg{A}]}}}}
+  \\
+& \jdeq
+  \subst{p}{{a}{{\gamma}{\jcomp{}{\jvcomp{}{\cftctxext[\cftalg{A}]}{\idtm{\jcomp{}{\cftctxext[\cftalg{A}]}{\cftalgf{\cftalg{A}}}}}}{\cftctxext[\cftalg{A}]}}}}
+  \\
+& \jdeq 
+  \subst{{p}{{a}{{\gamma}{\jvcomp{}{\cftctxext[\cftalg{A}]}{\idtm{\jcomp{}{\cftctxext[\cftalg{A}]}{\cftalgf{\cftalg{A}}}}}}}}}{\cftctxext[\cftalg{A}]}
+  \\
+& \jdeq
+  \subst
+    {\tmext{\apply\cftctxext[\extalg{A}]{\gamma}{a}}{p}}
+    {\cftctxext[\cftalg{A}]}
+  \\
+& \jdeq
+  \apply\cftctxext[\extalg{A}]{\apply\cftctxext[\extalg{A}]{\gamma}{a}}{p}.
 \end{align*}
@@
 \begin{align*}
-\subst{{y_2}{{y_1}{\jcomp{}{\cftctxext[\cftalg{A}]}{\cftfamext[\cftalg{A}]}}}}{{y_0}{\cftfamext[\cftalg{A}]}}
-& \jdeq
-  \subst{\tmext{y_0}{\subst{y_2}{{y_1}{\jcomp{}{\cftctxext[\cftalg{A}]}{\cftfamext[\cftalg{A}]}}}}}{\cftfamext[\cftalg{A}]}
-  \\
-& \jdeq
-  \subst{{y_2}{{y_1}{{y_0}{\jvcomp{}{\idtm{P}}{\jcomp{}{\cftctxext[\cftalg{A}]}{\cftfamext[\cftalg{A}]}}}}}}{\cftfamext[\cftalg{A}]}
-  \\
-& \jdeq
-  \subst
-    {y_2}
-    { {y_1}
-      { {y_0}
-        {\jcomp{}{\jvcomp{}{\idtm{P}}{\jcomp{}{\cftctxext[\cftalg{A}]}{\cftfamext[\cftalg{A}]}}}{\cftfamext[\cftalg{A}]}}
-        }
-      }
-  \\
-& \jdeq
-  \subst
-    {y_2}
-    { {y_1}
-      { {y_0}
-        {\jcomp{}{\jvcomp{}{\cftfamext[\cftalg{A}]}{\idtm{\jcomp{}{\cftfamext[\cftalg{A}]}{{}{\cftctxext[\cftalg{A}]}{\cftalgf{\cftalg{A}}}}}}}{\cftfamext[\cftalg{A}]}}
-        }
-      }
-  \\
-& \jdeq
-  \subst{{y_2}{{y_1}{{y_0}{\jvcomp{}{\cftfamext[\cftalg{A}]}{\idtm{\jcomp{}{\cftfamext[\cftalg{A}]}{{}{\cftctxext[\cftalg{A}]}{\cftalgf{\cftalg{A}}}}}}}}}}{\cftfamext[\cftalg{A}]}
-  \\
-& \jdeq
-  \subst{\tmext{\subst{y_1}{{y_0}{\cftfamext[\cftalg{A}]}}}{y_2}}{\cftfamext[\cftalg{A}]}
-  \\
-& \jdeq
-  \subst{y_2}{{{y_1}{{y_0}{\cftfamext[\cftalg{A}]}}}{\cftfamext[\cftalg{A}]}}.\qedhere
+\apply\cftfamext[\extalg{A}]{\gamma}{a}{%
+  \apply\cftfamext[\extalg{A}]{\apply\cftctxext[\extalg{A}]{\gamma}{a}}{p}{q}}
+& \jdeq
+  \apply\cftfamext[\extalg{A}]{\gamma}{%
+    \apply\cftfamext[\extalg{A}]{\gamma}{a}{p}}{q}.\qedhere
 \end{align*}
@@
 
-\begin{comment}
-\begin{rmk}
-I suspect that if we copy this theory of extension algebras to Martin-L\"of
-type theory, with the judgmental equalities replaced by identifications, with
-dependent pair types rather than those strict extensions, etcetera, then
-the type of $f_1$ for which these two diagrams commute is a mere proposition.
-
-With this notion of morphism, a term of an extension
-algebra $(A,P,\epsilon_0,\epsilon_1)$ is a pair $(x_0,x_1)$ such that
-$\subst{x_1}{{x_0}{\epsilon_0}}\jdeq x_0$.
-\end{rmk}
-\end{comment}
-
-\begin{comment}
-Extension algebras don't come in isolation. There are also extension algebra
-families and extension algebra terms. We now aim to define these and to
-establish various constructions of new extension algebras out of old ones:
-the empty extension algebra, and extensions, weakenings and substitutions
-of extension algebras and of course the identity term as an extension algebra
-term. We start with extension algebra families.
-
-\begin{defn}
-Consider an extension algebra $\mathcal{A}\defeq(A,P,e,f)$. 
-An extension algebra family over $\mathcal{A}$ is likewise a quadruple
-$\mathcal{B}\defeq(B,Q,g,h)$. Here we have a family $\jfam{{\Gamma}{A}}{B}$, a
-family $\jfam{{{{\Gamma}{A}}{P}}{\ctxwk{P}{B}}}{Q}$ and
-\begin{align*}
-\jhom*{{{\Gamma}{A}}{P}}{\ctxext{\ctxwk{P}{B}}{Q}}{\jcomp{}{\epsilon_0}{B}}{g}\\
-\jhom*{{{{\Gamma}{A}}{P}}{\ctxwk{P}{B}}}{\ctxext{Q}{\jcomp{}{g}{Q}}}{Q}{h}.
-\end{align*}
-The quadruple $(\jcomp{}{\epsilon_0}{B},Q,g,h)$ is required to be an extension algebra
-in context $\ctxext{{\Gamma}{A}}{P}$.
-\end{defn}
-
-\begin{defn}
-Suppose $\mathcal{A}$ is an extension algebra and $\mathcal{B}$ is an extension
-algebra family over $\mathcal{A}$. A term of $\mathcal{B}$ is a pair $(x,y)$
-consisting of
-\begin{align*}
-\jterm*{{\Gamma}{A}}{B}{x}\\
-\jterm*{{{\Gamma}{A}}{P}}{\subst{\jcomp{}{\epsilon_0}{x}}{Q}}{y}
-\end{align*}
-such that the diagrams
-\begin{equation*}
-\begin{tikzcd}
-\ctxext{\jcomp{}{\epsilon_0}{B}}{Q} 
-  \ar{r}{g} 
-  \ar[shift right=.7ex,fib]{d}
-& B 
-  \ar[shift right=.7ex,fib]{d} 
-  \\
-\ctxext{A}{P} 
-  \ar[shift right=.7ex,dotted]{u}[swap]{\tmext{\jcomp{}{\epsilon_0}{x}}{y}}
-  \ar{r}{e}
-& A
-  \ar[shift right=.7ex,dotted]{u}[swap]{x}
-\end{tikzcd}
-\end{equation*}
-and
-\begin{equation*}
-\begin{tikzcd}
-\jcomp{}{f}{\ctxext{Q}{\jcomp{}{g}{Q}}}
-  \ar{r}{\jcomp{}{f}{h}}
-  \ar[shift right=.7ex,fib]{d}
-& Q
-  \ar[shift right=.7ex,fib]{d}
-  \\
-\jcomp{}{f}{\jcomp{}{\epsilon_0}{B}}
-  \ar{r}{\idtm{\jcomp{}{f}{\jcomp{}{\epsilon_0}{B}}}}
-  \ar[shift right=.7ex,fib]{d}
-  \ar[shift right=.7ex,dotted,mapsto]{u}[swap]{\jcomp{}{f}{y}}
-& \jcomp{}{\epsilon_0}{B}
-  \ar[shift right=.7ex,fib]{d}
-  \ar[shift right=.7ex,dotted,mapsto]{u}[swap]{y}
-  \\
-\ctxext{P}{\jcomp{}{\epsilon_0}{P}}
-  \ar{r}[swap]{f}
-  \ar[shift right=.7ex,dotted]{u}[swap]{\jcomp{}{f}{\jcomp{}{\epsilon_0}{x}}}
-& P
-  \ar[shift right=.7ex,dotted]{u}[swap]{\jcomp{}{\epsilon_0}{x}}
-\end{tikzcd}
-\end{equation*}
-commute.
-\end{defn}
-
-\begin{defn}
-Suppose $\mathcal{A}$ and $\mathcal{B}$ are extension algebras in context
-$\Gamma$. We define the extension algebra $\ctxwk{\mathcal{A}}{\mathcal{B}}$
-to be the quadruple
-\begin{equation*}
-(\ctxwk{A}{B},\ctxwk{\ctxext{A}{P}}{Q},\ctxwk{\ctxext{A}{P}}{g},\ctxwk{\ctxext{A}{P}}{h}).
-\end{equation*}
-Note that $\ctxwk{\ctxext{A}{P}}{Q}$ is a family over $\ctxwk{\ctxext{A}{P}}{B}$,
-whereas it should be a family over $\jcomp{}{\epsilon_0}{\ctxwk{A}{B}}$. These are the
-same by \autoref{lem:prehom}.
-\end{defn}
-
-\begin{rmk}
-Before we continue, let us explore what it means to be an extension algebra
-term of the extension algebra $\ctxwk{\mathcal{A}}{\mathcal{B}}$. Such an
-extension algebra term $(x,y)$ would consist of
-\begin{align*}
-\jterm*{{\Gamma}{A}}{\ctxwk{A}{B}}{x}\\
-\jterm*{{{\Gamma}{A}}{P}}{\subst{\jcomp{}{\epsilon_0}{x}}{\ctxwk{\ctxext{A}{P}}{Q}}}{y}.
-\end{align*}
-Thus, $x$ is a context morphism from $A$ to $B$ and $y$ is nothing but a term
-of $\jcomp{}{\jcomp{}{\epsilon_0}{x}}{Q}$. For $x$, we see that the diagram
-\begin{equation*}
-\begin{tikzcd}
-\ctxext{B}{Q} 
-  \ar{r}{g} 
-& B 
-  \\
-\ctxext{A}{P} 
-  \ar{u}{\jvcomp{}{x}{y}}
-  \ar{r}{e}
-& A
-  \ar{u}[swap]{x}
-\end{tikzcd}
-\end{equation*}
-commutes.
-\end{rmk}
-\end{comment}
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 The notion of CFT-algebras that we will
-study in this subsection will also have an empty object $(\cftempc{\cftalg{A}},
-\cftempf{\cftalg{A}})$ and guarantiees that families over the empty object
-are just the terms of the underlying family $\cftalgc{\cftalg{A}}$ of contexts
-of the term-algebra $\cftalg{A}$. We will begin with an auxilary lemma which
-will help us dealing with the internal empty contexts:
+study in this subsection will be extension algebras which also have an empty 
+context $\cftempc{\cftalg{A}}$ and an empty family $\cftempf{\cftalg{A}}$ 
+satisfying (among other conditions) that families over the empty context in
+$\cftalg{A}$ are just the contexts of $\cftalg{A}$. We will begin with an 
+auxilary lemma which will help us dealing with the internal empty contexts:
 
@@
 
-\begin{comment}
-\begin{defn}
-A quintuple $(\cftalgc{\cftalg{A}},\cftalgf{\cftalg{A}},\cftalgt{\cftalg{A}},
-\cftempc{\cftalg{A}},\cftempf{\cftalg{A}})$ consisting of
-\begin{enumerate}
-\item A family $\cftalgc{\cftalg{A}}$ in context $\Gamma$,
-\item A family $\cftalgf{\cftalg{A}}$ in context $\ctxext{\Gamma}{\cftalgc{\cftalg{A}}}$,
-\item A family $\cftalgt{\cftalg{A}}$ in context $\ctxext{{\Gamma}{A}}{\cftalgf{\cftalg{A}}}$,
-\item A term $\cftempc{\cftalg{A}}$ of $\cftalgc{\cftalg{A}}$ in context $\Gamma$,
-\item A term $\cftempf{\cftalg{A}}$ of $\cftalgf{\cftalg{A}}$ in context $\ctxext{\Gamma}{\cftalgc{\cftalg{A}}}$,
-\end{enumerate}
-is said to be an \emph{empty-object algebra in context $\Gamma$} if additionally
-the following judgmental equalities hold:
-\begin{align}
-\jfameq*{\Gamma}{\subst{\cftempc{\cftalg{A}}}{\cftalgf{\cftalg{A}}}}{\cftalgc{\cftalg{A}}}
-  \label{empalg-eq1}
-  \\
-\jtermeq*{\Gamma}{\cftalgc{\cftalg{A}}}{\subst{\cftempc{\cftalg{A}}}{\cftempf{\cftalg{A}}}}{\cftempc{\cftalg{A}}}.
-  \label{empalg-eq2}
-%  \\
-%\jfameq*{{\Gamma}{\cftalgc{\cftalg{A}}}}{\subst{\cftempf{\cftalg{A}}}{\cftalgt{\cftalg{A}}}}{\emptyf}
-\end{align}
-When $(\cftalgc{\cftalg{A}},\cftalgf{\cftalg{A}},\cftalgt{\cftalg{A}},
-\cftempc{\cftalg{A}},\cftempf{\cftalg{A}})$ is an empty-object algebra
-in context $\Gamma$, we will
-also refer to the pair $(\cftempc{\cftalg{A}},\cftempf{\cftalg{A}})$ as the
-\emph{empty object of $\cftalg{A}$}.
-\end{defn}
-
-Below, we give the definition of empty-object homomorphisms. They have, like
-extension homomorphisms, an action on contexts and an action on families and
-in addition they have an action on terms. Thus, we can picture a term
-homomorphism as follows:
-\begin{equation*}
-\begin{tikzcd}
-\cftalgt{\cftalg{A}}
-  \ar[fib]{d}
-  \ar{r}{\cfthomt{\cfthom{f}}}
-& \cftalgt{\cftalg{B}}
-  \ar[fib]{d}
-  \\
-\cftalgf{\cftalg{A}}
-  \ar[fib]{d}
-  \ar{r}{\cfthomf{\cfthom{f}}}
-& \cftalgf{\cftalg{B}}
-  \ar[fib]{d}
-  \\
-\cftalgc{\cftalg{A}}
-  \ar{r}[swap]{\cfthomc{\cfthom{f}}}
-& \cftalgc{\cftalg{B}}
-\end{tikzcd}
-\end{equation*}
-
-\begin{defn}
-Let $\cftalg{A}$ and $\cftalg{B}$ be term algebras in context $\Gamma$.
-An \emph{empty-object homomorphism $\cfthom{f}$ from $\cftalg{A}$ to $\cftalg{B}$
-in context $\Gamma$} is a triple 
-$(\cfthomc{\cfthom{f}},\cfthomf{\cfthom{f}},\cfthomt{\cfthom{f}})$ consisting of
-\begin{align*}
-\jhom*
-  {\Gamma}
-  {\cftalgc{\cftalg{A}}}
-  {\cftalgc{\cftalg{B}}}
-  {\cfthomc{\cfthom{f}}}\\
-\jfhom*
-  {\Gamma}
-  {\cftalgc{\cftalg{A}}}
-  {\cftalgc{\cftalg{B}}}
-  {\cfthomc{\cfthom{f}}}
-  {\cftalgf{\cftalg{A}}}
-  {\cftalgf{\cftalg{B}}}
-  {\cfthomf{\cfthom{f}}}\\
-\jfhom*
-  {\Gamma}
-  {{\cftalgc{\cftalg{A}}}{\cftalgf{\cftalg{A}}}}
-  {{\cftalgc{\cftalg{B}}}{\cftalgf{\cftalg{B}}}}
-  {\jvcomp{\cftalgf{\cftalg{A}}}{\cfthomc{\cfthom{f}}}{\cfthomf{\cfthom{f}}}}
-  {\cftalgt{\cftalg{A}}}
-  {\cftalgt{\cftalg{B}}}
-  {\cfthomt{\cfthom{f}}}
-\end{align*}
-satisfying the judgmental equalities
-\begin{align*}
-\jtermeq*
-  {\Gamma}
-  {\cftalgc{\cftalg{B}}}
-  {\subst{\cftempc{\cftalg{A}}}{\cfthomc{\cfthom{f}}}}
-  {\cftempc{\cftalg{B}}}
-  \\
-\jhomeq*
-  {\Gamma}
-  {\cftalgc{\cftalg{A}}}
-  {\cftalgc{\cftalg{B}}}
-  {\subst{\cftempc{\cftalg{A}}}{\cfthomf{\cfthom{f}}}}
-  {\cfthomc{\cfthom{f}}}
-  \\
-\jtermeq*
-  {{\Gamma}{\cftalgc{\cftalg{A}}}}
-  {\jcomp{\cftalgc{\cftalg{A}}}{\cfthomc{\cfthom{f}}}{\cftalgf{\cftalg{B}}}}
-  {\subst{\cftempf{\cftalg{A}}}{\cfthomf{\cfthom{f}}}}
-  {\jcomp{\cftalgc{\cftalg{A}}}{\cfthomc{\cfthom{f}}}{\cftempf{\cftalg{B}}}}.
-\end{align*}
-\end{defn}
-
-A term-algebra $\cftalg{A}$ in context $\Gamma$ is going to be a septuple
-\begin{equation*}
-(\cftalgc{\cftalg{A}},\cftalgf{\cftalg{A}},\cftalgt{\cftalg{A}},
-\cftctxext[\cftalg{A}],\cftfamext[\cftalg{A}],\cftempc{\cftalg{A}},
-\cftempf{\cftalg{A}})
-\end{equation*}
-for which 
-the quadruple $\unfold{\extalg{A}}$ is an extension algebra in context 
-$\Gamma$, the quintuple $\unfold{\empalg{A}}$ is an empty-object algebra in context
-$\Gamma$, satisfying additional judgmental equalities expressing the 
-compatibility of $\cftctxext[\cftalg{A}]$ and $\cftfamext[\cftalg{A}]$ with 
-$\cftempc{\cftalg{A}}$ and $\cftempf{\cftalg{A}}$.
-There will be five such judgmental equalities. Four of them will express that
-extension acts as the identity term when applied to the empty object in one of
-its arguments. The other expresses that family extension restricted to
-families above the empty object is the same as context extension. To be able
-to state these, we need to know that $\unfold{\empalg{A}}$ is an empty-object
-algebra. In the following, we verify step-by-step that we can indeed make the
-suggested requirements.
-\begin{enumerate}
-\item The type of $\subst{\cftempc{\cftalg{A}}}{\cftctxext[\cftalg{A}]}$ is
-$\subst
-  { \cftempc{\cftalg{A}}
-    }
-  { \ctxwk
-      {\ctxext{\cftalgc{\cftalg{A}}}{\cftalgf{\cftalg{A}}}}
-      {\cftalgc{\cftalg{A}}}
-    }$.
-We have the judgmental equalities
-\begin{align*}
-\subst
-  { \cftempc{\cftalg{A}}
-    }
-  { \ctxwk
-      {\ctxext{\cftalgc{\cftalg{A}}}{\cftalgf{\cftalg{A}}}}
-      {\cftalgc{\cftalg{A}}}
-    }
-& \jdeq
-  \subst
-    { \cftempc{\cftalg{A}}
-      }
-    { \ctxwk
-        { \cftalgf{\cftalg{A}}
-          }
-        { {\cftalgc{\cftalg{A}}}
-          {\cftalgc{\cftalg{A}}}
-          }
-      }
-  \tag{by \autoref{comp-ew-f}}
-  \\
-& \jdeq
-  \ctxwk
-    { \subst{\cftempc{\cftalg{A}}}{\cftalgf{\cftalg{A}}}
-      }
-    { \subst{\cftempc{\cftalg{A}}}{\ctxwk{\cftalgc{\cftalg{A}}}{\cftalgc{\cftalg{A}}}}
-      }
-  \tag{by \autoref{comp-sw-f}}
-  \\
-& \jdeq
-  \ctxwk
-    { \subst{\cftempc{\cftalg{A}}}{\cftalgf{\cftalg{A}}}
-      }
-    { A
-      }
-  \tag{by \autoref{cancellation-ws-f}}
-  \\
-& \jdeq
-  \ctxwk{A}{A}
-  \tag{by \autoref{empalg-eq2}}
-\end{align*}
-Thus, we require that
-\begin{equation}
-\jtermeq{{\Gamma}{\cftalgc{\cftalg{A}}}}{\ctxwk{\cftalgc{\cftalg{A}}}{\cftalgc{\cftalg{A}}}}{\subst{\cftempc{\cftalg{A}}}{\cftctxext[\cftalg{A}]}}{\idtm{\cftalgc{\cftalg{A}}}}
-  \label{cftalg-eq1}
-\end{equation}
-\end{enumerate}
-
-We can immediately state the first two:
-\begin{align}
-\jtermeq*{{\Gamma}{\cftalgc{\cftalg{A}}}}{\ctxwk{\cftalgc{\cftalg{A}}}{\cftalgc{\cftalg{A}}}}{\subst{\cftempc{\cftalg{A}}}{\cftctxext[\cftalg{A}]}}{\idtm{\cftalgc{\cftalg{A}}}}
-  \label{cftalg-eq1}
-  \\
-\jtermeq*{{\Gamma}{\cftalgc{\cftalg{A}}}}{\ctxwk{\cftalgc{\cftalg{A}}}{\cftalgc{\cftalg{A}}}}{\subst{\cftempf{\cftalg{A}}}{\cftctxext[\cftalg{A}]}}{\idtm{\cftalgc{\cftalg{A}}}}
-  \label{cftalg-eq2}
-\end{align}
-To see what $\subst{\phi_0}{\epsilon_1}$ can be, we must know its type first.
-It is a morphism from $\subst{\phi_0}{\ctxext{P}{\jcomp{}{\epsilon_0}{P}}}$ to
-$\subst{\phi_0}{P}$. We already know that $\subst{\phi_0}{P}\jdeq A$ by
-\autoref{empalg-eq1} and to compute $\subst{\phi_0}{\jcomp{}{\epsilon_0}{P}}$
-we use \autoref{lem:empalg-mor}. Note that in combination with 
-\autoref{cftalg-eq1} and \autoref{cftalg-eq2} we get the judgmental
-equalities
-\begin{equation}\label{cor:empalg-mor}
-\jfameq{{\Gamma}{A}}{\subst{\phi_i}{\jcomp{}{\epsilon_0}{P}}}{P}
-\end{equation}
-for $i$ being $\mathsf{c}$ or $\mathsf{f}$. Consequently, 
-$\subst{\phi_0}{\epsilon_1}$ is a morphism from $\ctxext{A}{P}$ to $A$, so we 
-can require that
-\begin{equation}\label{cftalg-eq5}
-\jhomeq{\Gamma}{{A}{P}}{A}{\subst{\phi_0}{\epsilon_1}}{\epsilon_0}.
-\end{equation}
-This judgmental equality expresses that family extension restricted to the empty object
-$\phi_0$ is context extension. We also have two judgmental
-equalities expressing that $\epsilon_1$ restricted to $\phi_1$ and
-$\jcomp{}{\epsilon_0}{\phi_1}$ are both the identity term on $P$. Note that
-the term $\jcomp{}{\epsilon_0}{\phi_1}$ serves as the empty family of families,
-being a term of $\jcomp{}{\epsilon_0}{P}$.
-
-Before we can require that $\subst{\phi_1}{\epsilon_1}\jdeq\idtm{P}$, we need
-to compute the family
-\begin{equation*}
-\jfam
-  {{{\Gamma}{A}}{\subst{\phi_1}{\jcomp{}{\epsilon_0}{P}}}}
-  {\subst{\phi_1}{\ctxwk{\ctxext{P}{\jcomp{}{\epsilon_0}{P}}}{P}}}.
-\end{equation*}
-of which $\subst{\phi_1}{\epsilon_1}$ is a term.
-We have already established that $\subst{\phi_1}{\jcomp{}{\epsilon_0}{P}}\jdeq
-P$. We also see that 
-\begin{align*}
-\subst{\phi_1}{\ctxwk{\ctxext{P}{\jcomp{}{\epsilon_0}{P}}}{P}}
-& \jdeq
-  \subst{\phi_1}{\ctxwk{\jcomp{}{\epsilon_0}{P}}{{P}{P}}}
-  \tag{by \autoref{comp-ew-f}}
-  \\
-& \jdeq
-  \ctxwk{\subst{\phi_1}{\jcomp{}{\epsilon_0}{P}}}{\subst{\phi_1}{\ctxwk{P}{P}}}
-  \tag{by \autoref{comp-sw-f}}
-  \\
-& \jdeq
-  \ctxwk{P}{\subst{\phi_1}{\ctxwk{P}{P}}}
-  \tag{by \autoref{cor:empalg-mor}}
-  \\
-& \jdeq
-  \ctxwk{P}{P},
-  \tag{by \autoref{cancellation-ws-f}}
-\end{align*}
-so we see that it indeed makes sense to require that
-\begin{equation}\label{cftalg-eq3}
-\jtermeq{{{\Gamma}{A}}{P}}{\ctxwk{P}{P}}{\subst{\phi_1}{\epsilon_1}}{\idtm{P}}.
-\end{equation}
-For the last judgmental equality, we need to compute the family
-\begin{equation*}
-\jfam
-  {{{\Gamma}{A}}{\subst{\phi_1}{\jcomp{}{\epsilon_0}{P}}}}
-  {\subst{\jcomp{}{\epsilon_0}{\phi_1}}{\ctxwk{\ctxext{P}{\jcomp{}{\epsilon_0}{P}}}{P}}}.
-\end{equation*}
-This is easy:
-\begin{align*}
-\subst{\jcomp{}{\epsilon_0}{\phi_1}}{\ctxwk{\ctxext{P}{\jcomp{}{\epsilon_0}{P}}}{P}}
-& \jdeq
-  \subst{\jcomp{}{\epsilon_0}{\phi_1}}{\ctxwk{\jcomp{}{\epsilon_0}{P}}{{P}{P}}}
-  \\
-& \jdeq
-  \ctxwk{P}{P}.
-\end{align*}
-Thus we are allowed to require that
-\begin{equation}\label{cftalg-eq4}
-\subst{\jcomp{}{\epsilon_0}{\phi_1}}{\epsilon_1}.
-\end{equation}
-
-We bring all this together in the definition of extension-empty algebras:
-
-\begin{defn}
-An \emph{extension-empty algebra in context $\Gamma$} 
-is a sextuple $(A,P,\epsilon_0,\epsilon_1,\phi_0,\phi_1)$ for which 
-the quadruple $(A,P,\epsilon_0,\epsilon_1)$ is an extension algebra in context 
-$\Gamma$, the quadruple $(A,P,\phi_0,\phi_1)$ is an empty algebra in context
-$\Gamma$, satisfying the judgmental equalities 
-\autoref{cftalg-eq1,cftalg-eq2,cftalg-eq5,cftalg-eq3,cftalg-eq4}.
-\end{defn}
-\end{comment}
-
 We can extend the results of 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{CFT-homomorphisms}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Weakening algebras}
-Weakening algebras will be term-algebras with certain added structure.
+Weakening algebras will be CFT-algebras with certain added structure.
 Although strictly speaking one could formulate a notion of weakening algebra
@@
 
-Let $\cftalg{A}$ be a term-algebra in context $\Gamma$. Weakening will be a
+Let $\cftalg{A}$ be a CFT-algebra in context $\Gamma$. Weakening will be a
 CFT-homomorphism
@@
 \begin{rmk}
-Let $\cftalg{A}$ be a term-algebra and let $\cftwk{\cftalg{A}}$ be as in
+Let $\cftalg{A}$ be a CFT-algebra and let $\cftwk{\cftalg{A}}$ be as in
 \autoref{cftwk}. Then we have
@@
   {\Gamma}
-  {\subst{{a}{{\gamma}{\cftctxext}}}{\cftalgf{\cftalg{A}}}}
+  {\subst{\apply\cftctxext{\gamma}{a}}{\cftalgf{\cftalg{A}}}}
   {\subst{b}{{a}{{\gamma}{\cftwkc{\cftalg{A}}}}}}
@@
 \jterm
-  {\Gamma}
-  {\subst{{{b}{{a}{{\gamma}{\cftwkc{\cftalg{A}}}}}}{{{a}{{\gamma}{\cftctxext}}}{\cftctxext}}}{\cftalgf{\cftalg{A}}}}
-  {\subst{q}{{b}{{a}{{\gamma}{\cftwkf{\cftalg{A}}}}}}}
+  { \Gamma}
+  { \subst
+      { \apply\cftctxext
+          {\apply\cftctxext{\gamma}{a}}
+          {\subst{b}{{a}{{\gamma}{\cftwkc{\cftalg{A}}}}}}
+        }
+      { \cftalgf{\cftalg{A}}}}
+  { \subst{q}{{b}{{a}{{\gamma}{\cftwkf{\cftalg{A}}}}}}}
 \end{equation*}
 for $\gamma:A$, $a,b:\subst{\gamma}{\cftalgf{\cftalg{A}}}$ and
-$q:\subst{{b}{{\gamma}{\cftctxext}}}{\cftalgf{\cftalg{A}}}$. Finally, we have
+$q:\subst{\apply\cftctxext{\gamma}{b}}{\cftalgf{\cftalg{A}}}$. Finally, we have
 \begin{equation*}
 \jterm
-  {\Gamma}
-  {\subst{{q}{{b}{{a}{{\gamma}{\cftwkf{\cftalg{A}}}}}}}{{{{b}{{a}{{\gamma}{\cftwkc{\cftalg{A}}}}}}{{{a}{{\gamma}{\cftctxext}}}{\cftctxext}}}{\cftalgt{\cftalg{A}}}}}
+  { \Gamma}
+  { \subst
+      { {q}{{b}{{a}{{\gamma}{\cftwkf{\cftalg{A}}}}}}}
+      { { \apply\cftctxext
+            {\apply\cftctxext{\gamma}{a}}
+            {{b}{{a}{{\gamma}{\cftwkc{\cftalg{A}}}}}}
+          }
+        {\cftalgt{\cftalg{A}}}
+        }
+    }
   {\subst{g}{{q}{{b}{{a}{{\gamma}{\cftwkt{\cftalg{A}}}}}}}}
@@
 for $\gamma:A$, $a,b:\subst{\gamma}{\cftalgf{\cftalg{A}}}$,
-$q:\subst{{b}{{\gamma}{\cftctxext}}}{\cftalgf{\cftalg{A}}}$ and
-$g:\subst{q}{{{b}{{\gamma}{\cftctxext}}}{\cftalgt{\cftalg{A}}}}$.
+$q:\subst{\apply\cftctxext{\gamma}{b}}{\cftalgf{\cftalg{A}}}$ and
+$g:\subst{q}{{\apply\cftctxext{\gamma}{b}}{\cftalgt{\cftalg{A}}}}$.
 These observations show that the weakening morphism $\cftwk{\cftalg{A}}$ is
@@
 \subsubsection{Weakening preserves itself}
-The following diagram displays the compatibility of weakening with itself:
+Let $\cftalg{A}$ be a CFT-algebra in context $\Gamma$ and let
+\begin{equation*}
+\jhom
+  {{{\Gamma}{\cftalgc{\cftalg{A}}}}{\cftalgf{\cftalg{A}}}}
+  {\ctxwk{\cftalgf{\cftalg{A}}}{\cftalg{F_A}}}
+  {\jcomp{}{\cftctxext}{\cftalg{F_A}}}
+  {\cfthom{w}}
+\end{equation*}
+be a CFT-homomorphism. The abstraction of the rules in \autoref{comp-ww} is the
+condition that the diagram
 \begin{equation}\label{diag:cftwk-cftwk}
@@
         {{\cftalgf{\cftalg{A}}}{\cftalgf{\cftalg{F_A}}}}
-        {\boldsymbol{\mathcal{F}}_{\cftwk{\cftalg{A}}}}
-      }
-  \ar{r}{\ctxwk{\cftalgf{\cftalg{A}}}{\jcomp{}{\cftctxext}{\cftwk{\cftalg{A}}}}}
+        {\boldsymbol{\mathcal{F}}_{\cfthom{w}}}
+      }
+  \ar{r}{\ctxwk{\cftalgf{\cftalg{A}}}{\jcomp{}{\cftctxext}{\cfthom{w}}}}
 & \ctxwk{\cftalgf{\cftalg{A}}}{\cftalg{F_{F_{F_A}}}}
-  \ar{d}{\jcomp{}{\ctxwk{\cftalgf{\cftalg{A}}}{\cftfamext}}{\boldsymbol{\mathcal{F}}_{\cftwk{\cftalg{A}}}}}
-  \\
-\ctxwk{{\cftalgf{\cftalg{A}}}{\cftalgf{\cftalg{F_A}}}}{\jcomp{}{\cftwkc{A}}{\cftalg{F_{F_{F_A}}}}}
+  \ar{d}{\jcomp{}{\ctxwk{\cftalgf{\cftalg{A}}}{\cftfamext}}{\boldsymbol{\mathcal{F}}_{\cfthom{w}}}}
+  \\
+\ctxwk{{\cftalgf{\cftalg{A}}}{\cftalgf{\cftalg{F_A}}}}{\jcomp{}{\cfthomc{\cfthom{w}}}{\cftalg{F_{F_{F_A}}}}}
   \ar{r}[swap]{%
     \jcomp{}
-      { \cftwkf{\cftalg{A}}}
+      { \cfthomf{\cfthom{w}}}
       { \jcomp{}
-          {\cftwkc{\cftalg{A}}}
+          {\cfthomc{\cfthom{w}}}
           { \jcomp{}
@@
                   {\cftctxext}
-                  {\cftwk{\cftalg{A}}}
-                }
-            }
-        }
-    }
-& \jcomp{}{\cftwkc{A}}{\cftalg{F_{F_{F_{F_A}}}}}
+                  {\cfthom{w}}
+                }
+            }
+        }
+    }
+& \jcomp{}{\cfthomc{\cfthom{w}}}{\cftalg{F_{F_{F_{F_A}}}}}
 \end{tikzcd}
 \end{equation}
-This is a complicated diagram and it is not immediately clear that all of it
-makes sense. Therefore we examine its ingredients in more detail.
-
-The morphisms on the top and on the right are fine as they are displayed
+of CFT-algebras in context 
+$ \ctxext
+    {{{\Gamma}{\cftalgc{\cftalg{A}}}}{\cftalgf{\cftalg{A}}}}
+    {\cftalgf{\cftalg{F_A}}}$ %
+commutes. In this diagram, it is not immediately obvious that the domain and
+codomain of the CFT-homomorphism
+$ \jcomp{}
+      { \cfthomf{\cfthom{w}}}
+      { \jcomp{}
+          {\cfthomc{\cfthom{w}}}
+          { \jcomp{}
+              {\cftctxext}
+              { \jcomp{}
+                  {\cftctxext}
+                  {\cfthom{w}}
+                }
+            }
+        }$ %
+from
+$ \jcomp{}
+      { \cfthomf{\cfthom{w}}}
+      { \jcomp{}
+          {\cfthomc{\cfthom{w}}}
+          { \jcomp{}
+              {\cftctxext}
+              { \jcomp{}
+                  {\cftctxext}
+                  {\ctxwk{\cftalgf{\cftalg{A}}}{\cftalg{F_A}}}
+                }
+            }
+        }
+$ to
+$ \jcomp{}
+      { \cfthomf{\cfthom{w}}}
+      { \jcomp{}
+          {\cfthomc{\cfthom{w}}}
+          { \jcomp{}
+              {\cftctxext}
+              { \jcomp{}
+                  {\cftctxext}
+                  {\cftalg{F_{F_A}}}
+                }
+            }
+        }
+$ are indeed judgmentally equal to the indicated domain and codomain.
 
 \begin{lem}
-Let $\cftalg{A}$ be a term-algebra in context $\Gamma$ and let
+Let $\cftalg{A}$ be a CFT-algebra in context $\Gamma$ and let
 \begin{equation*}
@@
   {\jcomp{}{\cftctxext}{\cftalg{F_A}}}
-  {\cftwk{\cftalg{A}}}
-\end{equation*}
-be a homomorphism of term-algebras for which the diagram in
+  {\cfthom{w}}
+\end{equation*}
+be a homomorphism of CFT-algebras for which the diagram in
 \autoref{diag:cftwk-cftwk} commutes judgmentally. Then
@@
 & \jcomp{}
-  { \subst{q}{{{b}{{\gamma}{\cftctxext}}}{\cftwk{\cftalg{A}}}}
+  { \subst{q}{{\apply\cftctxext{\gamma}{b}}{\cfthom{w}}}
     }
   { \subst
-      {{q}{{b}{\cftfamext}}}
-      {{a}{{\gamma}{\boldsymbol{\mathcal{F}}_{\cftwk{\cftalg{A}}}}}}
+      {\apply\cftfamext{\gamma}{b}{q}}
+      {{a}{{\gamma}{\boldsymbol{\mathcal{F}}_{\cfthom{w}}}}}
     }
@@
   \jcomp{}
-    { \subst{b}{{a}{{\gamma}{\boldsymbol{\mathcal{F}}_{\cftwk{\cftalg{A}}}}}}
+    { \subst{b}{{a}{{\gamma}{\boldsymbol{\mathcal{F}}_{\cfthom{w}}}}}
       }
     { \subst
-        { {q}{{b}{{a}{{\gamma}{\cftwkf{\cftalg{A}}}}}}}
-        { {{{b}{{a}{{\gamma}{\cftwkc{\cftalg{A}}}}}}{{{a}{{\gamma}{\cftctxext}}}{\cftctxext}}}
-          {\cftwk{\cftalg{A}}}
+        { {q}{{b}{{a}{{\gamma}{\cfthomf{\cfthom{w}}}}}}}
+        { {\apply\cftctxext{\apply\cftctxext{\gamma}{a}}{\subst{b}{{a}{{\gamma}{\cfthomc{\cfthom{w}}}}}}}
+          {\cfthom{w}}
           }
@@
 $ \subst
-      { {q}{{b}{{a}{{\gamma}{\cftwkf{\cftalg{A}}}}}}
-        }
-      { {{b}{{a}{{\gamma}{\cftwkc{\cftalg{A}}}}}}
+      { {q}{{b}{{a}{{\gamma}{\cfthomf{\cfthom{w}}}}}}
+        }
+      { {{b}{{a}{{\gamma}{\cfthomc{\cfthom{w}}}}}}
         {{a}{{\gamma}{\cftalg{F_{F_{F_{F_{A}}}}}}}}
@@
 \subsubsection{Currying for weakening}
-\begin{equation*}
-\jhomeq
-  { {{{\Gamma}{\cftalgc{\cftalg{A}}}}{\cftalgf{\cftalg{A}}}}
+Let $\cftalg{A}$ be a CFT-algebra in context $\Gamma$ and let
+\begin{equation*}
+\jhom
+  {{{\Gamma}{\cftalgc{\cftalg{A}}}}{\cftalgf{\cftalg{A}}}}
+  {\ctxwk{\cftalgf{\cftalg{A}}}{\cftalg{F_A}}}
+  {\jcomp{}{\cftctxext}{\cftalg{F_A}}}
+  {\cfthom{w}}
+\end{equation*}
+be a CFT-homomorphism. The abstraction of the rules in \autoref{comp-ew} is the
+condition that the diagram
+\begin{equation}\label{eq:cftwk-cftextcurry}
+\begin{tikzcd}[column sep=large]
+\ctxwk
+  {\ctxext{\cftalgf{\cftalg{A}}}{\cftalgf{\cftalg{F_A}}}}
+  {\cftalg{F_A}}
+  \ar{r}{\ctxwk{\cftalgf{\cftalg{F_A}}}{\cfthom{w}}}
+  \ar{dr}[swap]{\jcomp{}{\cftfamext}{\cfthom{w}}}
+& \ctxwk{\cftalgf{\cftalg{F_A}}}{\cftalg{F_{F_A}}}
+  \ar{d}{\jcomp{}{\cftctxext}{\cfthom{w}}}
+  \\
+& \cftalg{F_{F_{F_A}}}
+\end{tikzcd}
+\end{equation}
+of CFT-algebras in context 
+$\ctxext
+    {{{\Gamma}{\cftalgc{\cftalg{A}}}}{\cftalgf{\cftalg{A}}}}
     {\cftalgf{\cftalg{F_A}}}
-    }
-  { \ctxwk
-      {\cftalgf{\cftalg{F_A}}}
-      {\cftalg{F_{F_A}}}
-    }
-  { \cftalg{F_{F_{F_A}}}
-    }
-  { \jcomp{}{\cftfamext}{\cftwk{\cftalg{A}}}
-    }
-  { \jcomp{}{\cftwk{\cftalg{A}}}{\jcomp{}{\cftctxext}{\cftwk{\cftalg{A}}}}
-    }
-\end{equation*}
+$ %
+commutes.
+
+\begin{lem}
+Let $\cftalg{A}$ be a CFT-algebra in context gamma with a CFT-homomorphism 
+$ \jhom
+  {{{\Gamma}{\cftalgc{\cftalg{A}}}}{\cftalgf{\cftalg{A}}}}
+  {\ctxwk{\cftalgf{\cftalg{A}}}{\cftalg{F_A}}}
+  {\jcomp{}{\cftctxext}{\cftalg{F_A}}}
+  {\cfthom{w}}
+$
+satisfying the judgmental equality of \autoref{eq:cftwk-cftextcurry}. Then we
+have the judgmental equality
+\begin{equation*}
+\subst{{p}{{a}{\cftfamext}}}{{\gamma}{\cfthom{w}}}
+\jdeq
+\jcomp{}
+    {\subst{a}{{\gamma}{\cfthom{w}}}}
+    {\subst{p}{{\apply\cftctxext{\gamma}{a}}{\cfthom{w}}}}
+\end{equation*}
+of CFT-homomorphisms from $\subst{\gamma}{\cftalg{F_A}}$ to 
+$\subst{p}{{a}{{\gamma}{\cftalg{F_{F_{F_A}}}}}}$ in context $\Gamma$,
+for every $\gamma:A$, $a:\subst{\gamma}{\cftalgf{\cftalg{A}}}$ and
+$p:\subst{a}{{\gamma}{\cftalgf{\cftalg{F_A}}}}$. 
+\end{lem}
 
 \subsubsection{Weakening by the empty family}
-\begin{equation*}
+\begin{equation}\label{eq:cftwk-cftempcurry}
 \jhomeq
@@
     }
-  { \subst{\cftempf{\cftalg{A}}}{\cftwk{\cftalg{A}}}
+  { \subst{\cftempf{\cftalg{A}}}{\cfthom{w}}
     }
@@
     }
-\end{equation*}
+\end{equation}
 
 \subsubsection{The definition of weakening algebras}\label{sec:cftwkalg-defn}
+
+\begin{defn}
+A weakening algebra in context $\Gamma$ is a CFT-algebra $\cftalg{A}$ in context
+$\Gamma$ for which there is a CFT-homomorphism
+\begin{equation*}
+\jhom
+  {{{\Gamma}{\cftalgc{\cftalg{A}}}}{\cftalgf{\cftalg{A}}}}
+  {\ctxwk{\cftalgf{\cftalg{A}}}{\cftalg{F_A}}}
+  {\cftalg{F_{F_A}}}
+  {\cftwk{\cftalg{A}}}
+\end{equation*}
+satisfying the judgmental equalities of %
+\autoref{diag:cftwk-cftwk,eq:cftwk-cftextcurry,eq:cftwk-cftempcurry}.
+\end{defn}
 
@@
 
-\begin{comment}
-Because the pair $(\omega_0,\omega_1)$ is required to be an extension homomorphism,
-the diagrams
-\begin{equation}\label{wkalg-exteq1}
-\begin{tikzcd}[column sep=large]
-\ctxwk{P}{\ctxext{P}{\jcomp{}{\epsilon_0}{P}}}
-  \ar{r}{\jvcomp{}{\omega_0}{\omega_1}}
-  \ar{d}[swap]{\ctxwk{P}{\epsilon_1}}
-& \ctxext
-    {\jcomp{}{\epsilon_0}{P}}
-    {\jcomp{}{\epsilon_0}{\jcomp{}{\epsilon_0}{P}}}
-  \ar{d}{\jcomp{}{\epsilon_0}{\epsilon_1}}
-  \\
-\ctxwk{P}{P}
-  \ar{r}[swap]{\omega_0}
-& \jcomp{}{\epsilon_0}{P}
-\end{tikzcd}
-\end{equation}
-in context $\ctxext{{\Gamma}{A}}{P}$, and
-\begin{equation}\label{wkalg-exteq2}
-\begin{tikzcd}[column sep=huge]
-\ctxwk
-  {P}
-  { \ctxext
-      {\jcomp{}{\epsilon_0}{P}}
-      {\jcomp{}{\epsilon_0}{\jcomp{}{\epsilon_0}{P}}}
-    }
-  \ar{r}{\jvcomp{}{\omega_1}{\jcomp{}{\ctxwk{P}{\epsilon_1}}{\omega_1}}}
-  \ar{d}[swap]{\ctxwk{P}{\jcomp{}{\epsilon_0}{\epsilon_1}}}
-& \jcomp{}{\omega_0}{%
-    \jcomp{}{\epsilon_0}{%
-      \jcomp{}{\epsilon_0}{%
-        \ctxext{P}{\jcomp{}{\epsilon_0}{P}}
-        }
-      }
-    }
-  \ar{d}{ \jcomp{}{\omega_0}{%
-            \jcomp{}{\epsilon_0}{%
-              \jcomp{}{\epsilon_0}{\epsilon_1}
-              }
-            }
-          }
-  \\
-\ctxwk{P}{\jcomp{}{\epsilon_0}{P}}
-  \ar{r}[swap]{\omega_1}
-& \jcomp{}{\omega_0}{%
-    \jcomp{}{\epsilon_0}{%
-      \jcomp{}{\epsilon_0}{P}
-      }
-    }
-\end{tikzcd}
-\end{equation}
-in context $\ctxext{{{\Gamma}{A}}{P}}{\ctxwk{P}{P}}$, commute judgmentally.
-To see that the morphism $\jvcomp{}{\omega_1}{\jcomp{}{\ctxwk{P}{\epsilon_0}}{\omega_1}}$
-indeed has the suggested codomain, we have the following lemma:
-
-\begin{lem}
-Let $(A,P,\epsilon_0,\epsilon_1)$ be an extension algebra and let
-\begin{align*}
-\jhom*
-  {{{\Gamma}{A}}{P}}
-  {\ctxwk{P}{P}}
-  {\jcomp{}{\epsilon_0}{P}}
-  {\omega_0}
-  \\
-\jfhom*
-  {{{\Gamma}{A}}{P}}
-  {\ctxwk{P}{P}}
-  {\jcomp{}{\epsilon_0}{P}}
-  {\omega_0}
-  {\ctxwk{P}{\jcomp{}{\epsilon_0}{P}}}
-  {\jcomp{}{\epsilon_0}{\jcomp{}{\epsilon_0}{P}}}
-  {\omega_1}
-\end{align*}
-be morphisms satisfying \autoref{wkalg-exteq1}. Then the inference rules
-\begin{align*}
-& \inference
-  { \jfam{{\Gamma}{A}}{Q}
-    }
-  { \jfameq
-      {\blank}
-      { \jcomp{}{\omega_1}{%
-          \jcomp{}{\omega_0}{%
-            \jcomp{}{\epsilon_0}{%
-              \jcomp{}{\epsilon_0}{%
-                \jcomp{}{\epsilon_0}{Q}
-                }
-              }
-            }
-          } 
-        }
-      { \jcomp{}{\ctxwk{P}{\epsilon_0}}{%
-          \jcomp{}{\omega_0}{%
-            \jcomp{}{\epsilon_0}{%
-              \jcomp{}{\epsilon_0}{%
-                Q
-                }
-              }
-            }
-          }
-        }
-      }
-\end{align*}
-are valid.
-\end{lem}
-
-\begin{proof}
-We have the judgmental equalities
-\begin{align*}
-& \jcomp{}{\omega_1}{%
-          \jcomp{}{\omega_0}{%
-            \jcomp{}{\epsilon_0}{%
-              \jcomp{}{\epsilon_0}{%
-                \jcomp{}{\epsilon_0}{Q}
-                }
-              }
-            }
-          } 
-  \\
-& \jdeq
-  \jcomp{}{\jvcomp{}{\omega_0}{\omega_1}}{%
-            \jcomp{}{\epsilon_0}{%
-              \jcomp{}{\epsilon_0}{%
-                \jcomp{}{\epsilon_0}{Q}
-                }
-              }
-            }
-  \tag{by \autoref{lem:composition-threesome}}
-  \\
-& \jdeq
-  \jcomp{}{\jvcomp{}{\omega_0}{\omega_1}}{%
-            \jcomp{}{\epsilon_0}{%
-              \jcomp{}{\epsilon_1}{%
-                \jcomp{}{\epsilon_0}{Q}
-                }
-              }
-            }
-  \tag{by \autoref{lem:extalg-twins}}
-  \\
-& \jdeq
-  \jcomp{}{\jvcomp{}{\omega_0}{\omega_1}}{%
-            \jcomp{}{\jcomp{}{\epsilon_0}{\epsilon_1}}{%
-              \jcomp{}{\epsilon_0}{%
-                \jcomp{}{\epsilon_0}{Q}
-                }
-              }
-            }
-   \tag{by \autoref{lem:jcomp-higherjcomp}}
-   \\
-& \jdeq
-  \jcomp{}{\jcomp{}{%
-             \jvcomp{}{\omega_0}{\omega_1}}{%
-               \jcomp{}{\epsilon_0}{\epsilon_1}}}{%
-              \jcomp{}{\epsilon_0}{%
-                \jcomp{}{\epsilon_0}{Q}
-                }
-              }
-  \tag{by \autoref{lem:jcomp-jcomp}}
-  \\
-& \jdeq
-  \jcomp{}{\jcomp{}{\ctxwk{P}{\epsilon_0}}{\omega_0}}{%
-              \jcomp{}{\epsilon_0}{%
-                \jcomp{}{\epsilon_0}{Q}
-                }
-              }
-  \tag{by \autoref{wkalg-exteq1}}
-  \\
-& \jdeq
-  \jcomp{}{\ctxwk{P}{\epsilon_0}}{%
-    \jcomp{}{\omega_0}{%
-      \jcomp{}{\epsilon_0}{%
-        \jcomp{}{\epsilon_0}{%
-          Q
-          }
-        }
-      }
-    }
-  \tag{by \autoref{lem:jcomp-jcomp}}
-\end{align*}
-\end{proof}
-
-\subsubsection{The compatibility of weakening with the empty context and family}
-The first two judgmental equalities expressing that $\omega_0$ and $\omega_1$
-are compatible with $\phi_1$ are easy to state:
-\begin{align}
-\jhomeq*{{\Gamma}{A}}{P}{P}{\subst{\phi_1}{\omega_0}}{\idtm{P}}\\
-\jhomeq*{{\Gamma}{A}}{P}{P}{\subst{\phi_1}{\omega_1}}{\idtm{P}}
-\end{align}
-
-\subsubsection{The compatibility of weakening with itself}
-To express the compatibility of weakening with itself, we must fill in the
-following diagram:
-\begin{equation*}
-\begin{tikzcd}
-\ctxwk{P}{{P}{P}}
-  \ar{r}{\ctxwk{P}{\omega_0}}
-  \ar{d}[swap]{\ctxwk{{P}{P}}{\omega_0}}
-& \ctxwk{P}{\jcomp{}{\epsilon_0}{P}}
-  \ar{d}{\omega_1}
-  \\
-\ctxwk{{P}{P}}{\jcomp{}{\epsilon_0}{P}}
-  \ar[densely dotted]{r}
-& \jcomp{}{\omega_0}{\jcomp{}{\epsilon_0}{\jcomp{}{\epsilon_0}{P}}}
-\end{tikzcd}
-\end{equation*}
-\end{comment}
-
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Projection algebras}

### 2014-07-03 article-models-ttinternal.tex
reorganized the preamble
source readability improvement

@@
 $ \ctxext
-    {{{\Gamma}{\cftalgc{\cftalg{A}}}}{\cftalgf{\cftalg{A}}}}
-    {\cftalgf{\cftalg{F_A}}}$ %
+    { { { {\Gamma}
+          {\cftalgc{\cftalg{A}}}
+          }
+        { \cftalgf{\cftalg{A}}}
+        }
+      { \ctxwk
+          {\cftalgf{\cftalg{A}}}
+          {\cftalgf{\cftalg{A}}}
+        }
+      }
+    { \ctxwk
+        {\cftalgf{\cftalg{A}}}
+        {\cftalgf{\cftalg{F_A}}}
+      }$ %
 commutes. In this diagram, it is not immediately obvious that the domain and
@@
 \begin{equation*}
-\subst{{p}{{a}{\cftfamext}}}{{\gamma}{\cfthom{w}}}
+\subst{\apply\cftfamext{\gamma}{a}{p}}{{\gamma}{\cfthom{w}}}
 \jdeq
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Substitution algebras}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{E-algebras}

### 2014-07-03 article-models-tt.tex
fixed broken references
source readability improvement
reorganized to match content of article-models-ttinternal.tex

@@
 finite extension of types. The idea behind this is to recover the theory of
-\autoref{Garner2014} {\color{red}(and possibly B-systems?)}. We stress that
+\cite{Garner2014} {\color{red}(and possibly B-systems?)}. We stress that
 we will not include axioms asserting such a stratification in the E-system
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Judgments and valid inference rules}\label{judgments}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{The basic rules for judgmental equality}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{The operations of the theory of contexts, families and terms}
@@
 
-\subsubsection{Extension}
+\subsection{Extension}
 \label{extension}
@@
 
-\paragraph{Associativity of extension}
+\subsubsection{Associativity of extension}
 \label{comp-ee}
@@
 
-\subsubsection{The empty context and the empty families}
+\subsection{The empty context and the empty families}
 \label{empty}
@@
 
-\paragraph{Compatibility of extension with the empty context and families}
+\subsubsection{Compatibility of extension with the empty context and families}
 In the following set of inference rules we state that the empty context and
@@
 
-\paragraph{Family extension restricted to the empty context}
+\subsubsection{Family extension restricted to the empty context}
 Strictly speaking we should have used a different notation for context extension
@@
 
-\subsubsection{Weakening}
+\subsection{The theory of weakening}
 \label{weakening}
@@
 
-\paragraph{Weakening of an empty family}
+\subsubsection{Weakening of an empty family}
 The following inference rules express that when the empty family is
@@
 
-\paragraph{Weakenings of extensions}
+\subsubsection{Weakenings of extensions}
 \label{comp-we}
@@
 
-\paragraph{Weakening of weakenings}
+\subsubsection{Weakening of weakenings}
 Weakening by a family $A$ in context $\Gamma$ brings things in context $\Gamma$
@@
 \end{rmk}
+
+\subsubsection{Weakening by the empty family}
+Note that we can also weaken by the empty family over $\Gamma$.
+Weakening by the empty family $\emptyf$ over a context $\Gamma$ leaves families, 
+their terms, families over those families and terms of those unchanged:
+\label{comp-0w}
+\begin{align}
+& \inference
+  { \jfam{\Gamma}{B}
+    }
+  { \jfameq{\Gamma}{\ctxwk{\emptyf}{B}}{B}
+    }
+  \label{comp-0w-c}
+  \\
+& \inference
+  { \jfam{{\Gamma}{B}}{Q}
+    }
+  { \jfameq{{\Gamma}{B}}{\ctxwk{\emptyf}{Q}}{Q}
+    }
+  \label{comp-0w-f}
+  \\
+& \inference
+  { \jterm{{\Gamma}{B}}{Q}{g}
+    }
+  { \jtermeq{{\Gamma}{B}}{Q}{\ctxwk{\emptyf}{g}}{g}
+    }
+  \label{comp-0w-t}
+\end{align}
+
+\subsubsection{Currying for weakening}
+\label{comp-ew}
+The rules expressing that extension is compatible with weakening assert that
+weakening by an extension is the same thing as weakening twice in the
+appropriate way.
+\begin{align}
+& \inference
+  { \jfam{\Gamma}{A}
+    \jfam{{\Gamma}{A}}{P}
+    \jfam{\Gamma}{B}
+    }
+  { \jfameq
+      {{{\Gamma}{A}}{P}}
+      {\ctxwk{\ctxext{A}{P}}{B}}
+      {\ctxwk{P}{{A}{B}}}
+    }
+  \label{comp-ew-c}\\
+& \inference
+  { \jfam{\Gamma}{A}
+    \jfam{{\Gamma}{A}}{P}
+    \jfam{{\Gamma}{B}}{Q}
+    }
+  { \jfameq
+      {{{{\Gamma}{A}}{P}}{\ctxwk{P}{{A}{B}}}}
+      {\ctxwk{\ctxext{A}{P}}{Q}}
+      {\ctxwk{P}{{A}{Q}}}
+    }
+  \label{comp-ew-f}\\
+& \inference
+  { \jfam{\Gamma}{A}
+    \jfam{{\Gamma}{A}}{P}
+    \jterm{{\Gamma}{B}}{Q}{g}
+    }
+  { \jtermeq
+      {{{{\Gamma}{A}}{P}}{\ctxwk{P}{{A}{B}}}}
+      {\ctxwk{P}{{A}{Q}}}
+      {\ctxwk{\ctxext{A}{P}}{g}}
+      {\ctxwk{P}{{A}{g}}}
+    } 
+  \label{comp-ew-t}
+\end{align}
 
@@
 
-\subsubsection{Substitution}
+\subsection{The theory of projections}
+\label{identityterms}
+We formulate the theory of projections on top of the theory of weakening
+described in \autoref{weakening}. In the theory of projections we introduce
+the identity terms. The identity term at a family $A$ in context $\Gamma$ is
+a term $\jterm{{\Gamma}{A}}{\ctxwk{A}{A}}{\idtm{A}}$, so weakening is necessary.
+Together with weakening, the identity terms will provide for the projections.
+The projection from $\cftext{A}{P}$ to $A$ will be the term
+$\jterm{{\Gamma}{{A}{P}}}{\ctxwk{\ctxext{A}{P}}{A}}{\ctxwk{P}{\idtm{A}}}$ and
+will be studied extensively in \autoref{extension-on-terms}.
+
+The following inference rules are required to be valid:
+\begin{align}
+& \inference
+  { \jfam{\Gamma}{A}
+    }
+  { \unfoldall{\jhom{\Gamma}{A}{A}{\idtm{A}}}
+    }
+& & \inference
+    { \jfameq{\Gamma}{A}{A'}
+      }
+    { \unfoldall{\jhomeq{\Gamma}{A}{A}{\idtm{A}}{\idtm{A'}}}
+      }
+\end{align}
+
+\begin{comment}
+We stress again
+that because we have not required there to be an empty context, we need to
+explicitly call for identity terms $\idtm{\Gamma}$ of $\ctxwk{\Gamma}{\Gamma}$
+in context $\Gamma$ for each context $\Gamma$. Thus, we will also require
+the inference rule
+\begin{align}
+& \inference
+  { \jctx{\Gamma}
+    }
+  { \jterm{\Gamma}{\ctxwk{\Gamma}{\Gamma}}{\idtm{\Gamma}}
+    }
+& &
+  \inference
+  { \jctxeq{\Gamma}{\Gamma'}
+    }
+  { \jtermeq{\Gamma}{\ctxwk{\Gamma}{\Gamma}}{\idtm{\Gamma}}{\idtm{\Gamma'}}
+    }
+\end{align}
+For the following inference rules, there should also be versions of identity
+terms at contexts.
+\end{comment}
+
+\subsubsection{The identity term of an extended family}
+The fact that identity terms are compatible with extensions is derived in
+\autoref{comp-ie}.
+
+\subsubsection{The identity term of a weakened family}
+\label{comp-wi}
+The identity term of a weakened family is the weakened identity term:
+\begin{equation}
+\inference
+  { \jfam{\Gamma}{A}
+    \jfam{{\Gamma}{B}}{Q}
+    }
+  { \unfoldall{\jhomeq
+      {{{\Gamma}{A}}{\ctxwk{A}{B}}}
+      {\ctxwk{A}{Q}}
+      {\ctxwk{A}{Q}}
+      {\ctxwk{A}{\idtm{Q}}}
+      {\idtm{\ctxwk{A}{Q}}}
+      }
+    }
+  \label{comp-wi-t}
+\end{equation}
+
+\begin{rmk}
+An important special case is the judgmental equality
+\begin{equation*}
+\unfoldall{\jhomeq
+      {{\Gamma}{A}}
+      {\ctxwk{A}{B}}
+      {\ctxwk{A}{B}}
+      {\ctxwk{A}{\idtm{B}}}
+      {\idtm{\ctxwk{A}{B}}}
+      }
+\end{equation*}
+for a family $\jfam{\Gamma}{B}$.
+\end{rmk}
+
+\subsection{The theory of substitution}
 \label{substitution}
+The theory of substitution can be formulated separately from the theory of
+weakening. Therefore, we will only require the theory of extension and the
+empty context and families. In \autoref{sec:esystem-equalities} we will
+provide the additional judgmental equality rules that will be required when
+combining the theories of substitution and of projections.
+
 Given a family $P$ over $A$ and a term $x$ of $A$, substitution gives a way to
@@
     \\
-%& \inference
-%  { \jterm{{\Gamma}{A}}{P}{f}
-%    }
-%  { \jterm{\Gamma}{\subst{x}{P}}{\subst{x}{f}}
-%    }
-%& & \inference
-%    { \jtermeq{{\Gamma}{A}}{P}{f}{f'}
-%      }
-%    { \jtermeq{\Gamma}{\subst{x}{P}}{\subst{x}{f}}{\subst{x'}{f'}}
-%      }
-%    \\
 & \inference
@@
 
-\paragraph{Fibers of an empty family}
+\subsubsection{Fibers of an empty family}
 The following inference rules establish that the fibers of the empty family are 
@@
 
-\paragraph{Fibers of extensions}
+\subsubsection{Fibers of extensions}
 \label{comp-se}
@@
 
-\paragraph{Fibers of weakenings}\label{comp-sw}
-The following rules assert what happens when we first weaken by a family
-$Q$ over $P$ in context $\ctxext{\Gamma}{A}$ and then substitute by a term
-$x$ of $A$:
-\begin{align}
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jfam{{{\Gamma}{A}}{P}}{Q}
-    \jfam{{{\Gamma}{A}}{P}}{R}
-    }
-  { \jfameq
-      {{{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}}
-      {\subst{x}{\ctxwk{Q}{R}}}
-      {\ctxwk{\subst{x}{Q}}{\subst{x}{R}}}
-    }
-  \label{comp-sw-c}
-  \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jfam{{{\Gamma}{A}}{P}}{Q}
-    \jfam{{{{\Gamma}{A}}{P}}{R}}{S}
-    }
-  { \jfameq
-      {{{{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}}{\subst{x}{\ctxwk{Q}{R}}}}
-      {\subst{x}{\ctxwk{Q}{S}}}
-      {\ctxwk{\subst{x}{Q}}{\subst{x}{S}}}
-    }
-  \label{comp-sw-f}
-  \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jfam{{{\Gamma}{A}}{P}}{Q}
-    \jterm{{{{\Gamma}{A}}{P}}{R}}{S}{k}
-    }
-  { \jtermeq
-      {{{{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}}{\subst{x}{\ctxwk{Q}{R}}}}
-      {\subst{x}{\ctxwk{Q}{S}}}
-      {\subst{x}{\ctxwk{Q}{k}}}
-      {\ctxwk{\subst{x}{Q}}{\subst{x}{k}}}
-    }
-  \label{comp-sw-t}
-\end{align}
-
-\begin{rmk}
-As an important special case of these inference rules we have the following
-valid inference rules:
-\begin{align*}
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jfam{{\Gamma}{A}}{P}
-    \jfam{{\Gamma}{A}}{Q}
-    }
-  { \jfameq
-      {{\Gamma}{\subst{x}{P}}}
-      {\subst{x}{\ctxwk{P}{Q}}}
-      {\ctxwk{\subst{x}{P}}{\subst{x}{Q}}}
-    }
-  \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jfam{{\Gamma}{A}}{P}
-    \jfam{{{\Gamma}{A}}{Q}}{R}
-    }
-  { \jfameq
-      {{{\Gamma}{\subst{x}{P}}}{\subst{x}{\ctxwk{P}{Q}}}}
-      {\subst{x}{\ctxwk{P}{R}}}
-      {\ctxwk{\subst{x}{P}}{\subst{x}{R}}}
-    }
-  \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jfam{{\Gamma}{A}}{P}
-    \jterm{{{\Gamma}{A}}{Q}}{R}{h}
-    }
-  { \jtermeq
-      {{{\Gamma}{\subst{x}{P}}}{\subst{x}{\ctxwk{P}{Q}}}}
-      {\subst{x}{\ctxwk{P}{R}}}
-      {\subst{x}{\ctxwk{P}{h}}}
-      {\ctxwk{\subst{x}{P}}{\subst{x}{h}}}
-    }
-\end{align*}
-Moreover, we get
-\begin{equation*}
-\inference
-  { \jterm{\Gamma}{A}{x}
-    \jfam{{\Gamma}{A}}{P}
-    \jterm{{\Gamma}{A}}{Q}{g}
-    }
-  { \jtermeq
-      {{\Gamma}{\subst{x}{P}}}
-      {\subst{x}{\ctxwk{P}{Q}}}
-      {\subst{x}{\ctxwk{P}{g}}}
-      {\ctxwk{\subst{x}{P}}{\subst{x}{g}}}
-    }
-\end{equation*}
-\end{rmk}
-
-\paragraph{Fibers of fibers}
+\subsubsection{Fibers of fibers}
 \label{comp-ss}
@@
     }
-  \label{comp-ss-f}
+  \label{comp-ss-t}
 \end{align}
@@
 
-\paragraph{Weakenings of fibers}
-\label{comp-ws}
-The following inference rules explain what happens when we first weaken by a
-term $g$ of $Q$ in context $\ctxext{\Gamma}{B}$ and then weaken by a family
-$A$ over $\Gamma$.
-\begin{align}
-& \inference
-  { \jfam{\Gamma}{A}
-    \jterm{{\Gamma}{B}}{Q}{g}
-    \jfam{{{\Gamma}{B}}{Q}}{R}
-    }
-  { \jfameq
-      {{{\Gamma}{A}}{\ctxwk{A}{B}}}
-      {\ctxwk{A}{\subst{g}{R}}}
-      {\subst{\ctxwk{A}{g}}{\ctxwk{A}{R}}}
-    }
-  \label{comp-ws-c}
-  \\
-& \inference
-  { \jfam{\Gamma}{A}
-    \jterm{{\Gamma}{B}}{Q}{g}
-    \jfam{{{{\Gamma}{B}}{Q}}{R}}{S}
-    }
-  { \jfameq
-      {{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{\subst{g}{R}}}}
-      {\ctxwk{A}{\subst{g}{S}}}
-      {\subst{\ctxwk{A}{g}}{\ctxwk{A}{S}}}
-    }
-  \label{comp-ws-f}
-  \\
-& \inference
-  { \jfam{\Gamma}{A}
-    \jterm{{\Gamma}{B}}{Q}{g}
-    \jterm{{{{\Gamma}{B}}{Q}}{R}}{S}{k}
-    }
-  { \jtermeq
-      {{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{\subst{g}{R}}}}
-      {\ctxwk{A}{\subst{g}{S}}}
-      {\ctxwk{A}{\subst{g}{k}}}
-      {\subst{\ctxwk{A}{g}}{\ctxwk{A}{k}}}
-    }
-  \label{comp-ws-t}
-\end{align}
-
-\begin{rmk}
-As an important special case of these inference rules we have the following 
-valid inference rules:
-\begin{align*}
-& \inference
-  { \jfam{\Gamma}{A}
-    \jterm{\Gamma}{B}{y}
-    \jfam{{\Gamma}{B}}{Q}
-    }
-  { \jfameq
-      {{\Gamma}{A}}
-      {\ctxwk{A}{\subst{y}{Q}}}
-      {\subst{\ctxwk{A}{y}}{\ctxwk{A}{Q}}}
-    }
-  \\
-& \inference
-  { \jfam{\Gamma}{A}
-    \jterm{\Gamma}{B}{y}
-    \jfam{{{\Gamma}{B}}{Q}}{R}
-    }
-  { \jfameq
-      {{{\Gamma}{A}}{\ctxwk{A}{\subst{y}{Q}}}}
-      {\ctxwk{A}{\subst{y}{R}}}
-      {\subst{\ctxwk{A}{y}}{\ctxwk{A}{R}}}
-    }
-  \\
-& \inference
-  { \jfam{\Gamma}{A}
-    \jterm{\Gamma}{B}{y}
-    \jterm{{{\Gamma}{B}}{Q}}{R}{h}
-    }
-  { \jtermeq
-      {{{\Gamma}{A}}{\ctxwk{A}{\subst{y}{Q}}}}
-      {\ctxwk{A}{\subst{y}{R}}}
-      {\ctxwk{A}{\subst{y}{h}}}
-      {\subst{\ctxwk{A}{y}}{\ctxwk{A}{h}}}
-    }
-\end{align*}
-Moreover, we get
-\begin{equation*}
-\inference
-  { \jfam{\Gamma}{A}
-    \jterm{\Gamma}{B}{y}
-    \jterm{{\Gamma}{B}}{Q}{g}
-    }
-  { \jtermeq
-      {{\Gamma}{A}}
-      {\ctxwk{A}{\subst{y}{Q}}}
-      {\ctxwk{A}{\subst{y}{g}}}
-      {\subst{\ctxwk{A}{y}}{\ctxwk{A}{g}}}
-    }
-\end{equation*}
-\end{rmk}
-
-\subsubsection{Identity terms}
-\label{identityterms}
-Last, but not least, we introduce the identity terms. In combination with
-weakening and substitution, identity terms will allow us to speak of projections,
-of a variant of extension for terms and of fiber inclusions.
-The following
-inference rules are required to be valid:
-\begin{align}
-& \inference
-  { \jfam{\Gamma}{A}
-    }
-  { \unfoldall{\jhom{\Gamma}{A}{A}{\idtm{A}}}
-    }
-& & \inference
-    { \jfameq{\Gamma}{A}{A'}
-      }
-    { \unfoldall{\jhomeq{\Gamma}{A}{A}{\idtm{A}}{\idtm{A'}}}
-      }
-\end{align}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Joining the theory of projections with the theory of substitution}
+\label{sec:esystem-equalities}
 
 \begin{comment}
-We stress again
-that because we have not required there to be an empty context, we need to
-explicitly call for identity terms $\idtm{\Gamma}$ of $\ctxwk{\Gamma}{\Gamma}$
-in context $\Gamma$ for each context $\Gamma$. Thus, we will also require
-the inference rule
-\begin{align}
-& \inference
-  { \jctx{\Gamma}
-    }
-  { \jterm{\Gamma}{\ctxwk{\Gamma}{\Gamma}}{\idtm{\Gamma}}
-    }
-& &
-  \inference
-  { \jctxeq{\Gamma}{\Gamma'}
-    }
-  { \jtermeq{\Gamma}{\ctxwk{\Gamma}{\Gamma}}{\idtm{\Gamma}}{\idtm{\Gamma'}}
-    }
-\end{align}
-For the following inference rules, there should also be versions of identity
-terms at contexts.
-\end{comment}
-
-\paragraph{The identity term of an extended family}
-The fact that identity terms are compatible with extensions is derived in
-\autoref{comp-ie}.
-
-\paragraph{The identity term of a weakened family}
-\label{comp-wi}
-The identity term of a weakened family is the weakened identity term:
-\begin{equation}
-\inference
-  { \jfam{\Gamma}{A}
-    \jfam{{\Gamma}{B}}{Q}
-    }
-  { \unfoldall{\jhomeq
-      {{{\Gamma}{A}}{\ctxwk{A}{B}}}
-      {\ctxwk{A}{Q}}
-      {\ctxwk{A}{Q}}
-      {\ctxwk{A}{\idtm{Q}}}
-      {\idtm{\ctxwk{A}{Q}}}
-      }
-    }
-  \label{comp-wi-t}
-\end{equation}
-
-\begin{rmk}
-An important special case is the judgmental equality
-\begin{equation*}
-\unfoldall{\jhomeq
-      {{\Gamma}{A}}
-      {\ctxwk{A}{B}}
-      {\ctxwk{A}{B}}
-      {\ctxwk{A}{\idtm{B}}}
-      {\idtm{\ctxwk{A}{B}}}
-      }
-\end{equation*}
-for a family $\jfam{\Gamma}{B}$.
-\end{rmk}
-
-\paragraph{The identity term of a substituted family}
-\label{comp-si}
-The identity term of a substituted family is the substitution of the identity term
-\begin{equation}
-\inference
-  { \jterm{\Gamma}{A}{x}
-    \jfam{{{\Gamma}{A}}{P}}{Q}
-    }
-  { \unfoldall{\jhomeq
-      {{\Gamma}{\subst{x}{P}}}
-      {\subst{x}{Q}}
-      {\subst{x}{Q}}
-      {\subst{x}{\idtm{Q}}}
-      {\idtm{\subst{x}{Q}}}
-    }}
-  \label{comp-si-t}
-\end{equation}
-
-\begin{rmk}
-An important special case is the judgmental equality
-\begin{equation*}
-\unfoldall{\jhomeq
-      {\Gamma}
-      {\subst{x}{P}}
-      {\subst{x}{P}}
-      {\subst{x}{\idtm{P}}}
-      {\idtm{\subst{x}{P}}}
-    }
-\end{equation*}
-for a family $\jfam{{\Gamma}{A}}{P}$.
-\end{rmk}
-
-\subsection{Further judgmental equalities of the theory of contexts, families and
-terms}
 We have already seen that empty families, extension, weakening, substitution
@@
       context $\Gamma$;
-\item something is substituted by the term of the empty family;
 \item something is weakened by an extended family. This is Currying for weakening;
@@
 completing the scheme (work out further).
+\end{comment}
 
@@
 
-\subsubsection{Weakening by the empty family}
-Note that we can also weaken by the empty family over $\Gamma$.
-Weakening by the empty family $\emptyf$ over a context $\Gamma$ leaves families, 
-their terms, families over those families and terms of those unchanged:
-\label{comp-0w}
-\begin{align}
-& \inference
-  { \jfam{\Gamma}{B}
-    }
-  { \jfameq{\Gamma}{\ctxwk{\emptyf}{B}}{B}
-    }
-  \label{comp-0w-c}
-  \\
-& \inference
-  { \jfam{{\Gamma}{B}}{Q}
-    }
-  { \jfameq{{\Gamma}{B}}{\ctxwk{\emptyf}{Q}}{Q}
-    }
-  \label{comp-0w-f}
-  \\
-& \inference
-  { \jterm{{\Gamma}{B}}{Q}{g}
-    }
-  { \jtermeq{{\Gamma}{B}}{Q}{\ctxwk{\emptyf}{g}}{g}
-    }
-  \label{comp-0w-t}
-\end{align}
-
-\subsubsection{Currying for weakening}
-\label{comp-ew}
-The rules expressing that extension is compatible with weakening assert that
-weakening by an extension is the same thing as weakening twice in the
-appropriate way.
-\begin{align}
-& \inference
-  { \jfam{\Gamma}{A}
+\subsubsection{Fibers of weakenings}\label{comp-sw}
+The following rules assert what happens when we first weaken by a family
+$Q$ over $P$ in context $\ctxext{\Gamma}{A}$ and then substitute by a term
+$x$ of $A$:
+\begin{align}
+& \inference
+  { \jterm{\Gamma}{A}{x}
+    \jfam{{{\Gamma}{A}}{P}}{Q}
+    \jfam{{{\Gamma}{A}}{P}}{R}
+    }
+  { \jfameq
+      {{{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}}
+      {\subst{x}{\ctxwk{Q}{R}}}
+      {\ctxwk{\subst{x}{Q}}{\subst{x}{R}}}
+    }
+  \label{comp-sw-c}
+  \\
+& \inference
+  { \jterm{\Gamma}{A}{x}
+    \jfam{{{\Gamma}{A}}{P}}{Q}
+    \jfam{{{{\Gamma}{A}}{P}}{R}}{S}
+    }
+  { \jfameq
+      {{{{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}}{\subst{x}{\ctxwk{Q}{R}}}}
+      {\subst{x}{\ctxwk{Q}{S}}}
+      {\ctxwk{\subst{x}{Q}}{\subst{x}{S}}}
+    }
+  \label{comp-sw-f}
+  \\
+& \inference
+  { \jterm{\Gamma}{A}{x}
+    \jfam{{{\Gamma}{A}}{P}}{Q}
+    \jterm{{{{\Gamma}{A}}{P}}{R}}{S}{k}
+    }
+  { \jtermeq
+      {{{{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}}{\subst{x}{\ctxwk{Q}{R}}}}
+      {\subst{x}{\ctxwk{Q}{S}}}
+      {\subst{x}{\ctxwk{Q}{k}}}
+      {\ctxwk{\subst{x}{Q}}{\subst{x}{k}}}
+    }
+  \label{comp-sw-t}
+\end{align}
+
+\begin{rmk}
+As an important special case of these inference rules we have the following
+valid inference rules:
+\begin{align*}
+& \inference
+  { \jterm{\Gamma}{A}{x}
     \jfam{{\Gamma}{A}}{P}
-    \jfam{\Gamma}{B}
-    }
-  { \jfameq
-      {{{\Gamma}{A}}{P}}
-      {\ctxwk{\ctxext{A}{P}}{B}}
-      {\ctxwk{P}{{A}{B}}}
-    }
-  \label{comp-ew-c}\\
-& \inference
-  { \jfam{\Gamma}{A}
+    \jfam{{\Gamma}{A}}{Q}
+    }
+  { \jfameq
+      {{\Gamma}{\subst{x}{P}}}
+      {\subst{x}{\ctxwk{P}{Q}}}
+      {\ctxwk{\subst{x}{P}}{\subst{x}{Q}}}
+    }
+  \\
+& \inference
+  { \jterm{\Gamma}{A}{x}
     \jfam{{\Gamma}{A}}{P}
+    \jfam{{{\Gamma}{A}}{Q}}{R}
+    }
+  { \jfameq
+      {{{\Gamma}{\subst{x}{P}}}{\subst{x}{\ctxwk{P}{Q}}}}
+      {\subst{x}{\ctxwk{P}{R}}}
+      {\ctxwk{\subst{x}{P}}{\subst{x}{R}}}
+    }
+  \\
+& \inference
+  { \jterm{\Gamma}{A}{x}
+    \jfam{{\Gamma}{A}}{P}
+    \jterm{{{\Gamma}{A}}{Q}}{R}{h}
+    }
+  { \jtermeq
+      {{{\Gamma}{\subst{x}{P}}}{\subst{x}{\ctxwk{P}{Q}}}}
+      {\subst{x}{\ctxwk{P}{R}}}
+      {\subst{x}{\ctxwk{P}{h}}}
+      {\ctxwk{\subst{x}{P}}{\subst{x}{h}}}
+    }
+\end{align*}
+Moreover, we get
+\begin{equation*}
+\inference
+  { \jterm{\Gamma}{A}{x}
+    \jfam{{\Gamma}{A}}{P}
+    \jterm{{\Gamma}{A}}{Q}{g}
+    }
+  { \jtermeq
+      {{\Gamma}{\subst{x}{P}}}
+      {\subst{x}{\ctxwk{P}{Q}}}
+      {\subst{x}{\ctxwk{P}{g}}}
+      {\ctxwk{\subst{x}{P}}{\subst{x}{g}}}
+    }
+\end{equation*}
+\end{rmk}
+
+\subsubsection{Weakenings of fibers}
+\label{comp-ws}
+The following inference rules explain what happens when we first weaken by a
+term $g$ of $Q$ in context $\ctxext{\Gamma}{B}$ and then weaken by a family
+$A$ over $\Gamma$.
+\begin{align}
+& \inference
+  { \jfam{\Gamma}{A}
+    \jterm{{\Gamma}{B}}{Q}{g}
+    \jfam{{{\Gamma}{B}}{Q}}{R}
+    }
+  { \jfameq
+      {{{\Gamma}{A}}{\ctxwk{A}{B}}}
+      {\ctxwk{A}{\subst{g}{R}}}
+      {\subst{\ctxwk{A}{g}}{\ctxwk{A}{R}}}
+    }
+  \label{comp-ws-c}
+  \\
+& \inference
+  { \jfam{\Gamma}{A}
+    \jterm{{\Gamma}{B}}{Q}{g}
+    \jfam{{{{\Gamma}{B}}{Q}}{R}}{S}
+    }
+  { \jfameq
+      {{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{\subst{g}{R}}}}
+      {\ctxwk{A}{\subst{g}{S}}}
+      {\subst{\ctxwk{A}{g}}{\ctxwk{A}{S}}}
+    }
+  \label{comp-ws-f}
+  \\
+& \inference
+  { \jfam{\Gamma}{A}
+    \jterm{{\Gamma}{B}}{Q}{g}
+    \jterm{{{{\Gamma}{B}}{Q}}{R}}{S}{k}
+    }
+  { \jtermeq
+      {{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{\subst{g}{R}}}}
+      {\ctxwk{A}{\subst{g}{S}}}
+      {\ctxwk{A}{\subst{g}{k}}}
+      {\subst{\ctxwk{A}{g}}{\ctxwk{A}{k}}}
+    }
+  \label{comp-ws-t}
+\end{align}
+
+\begin{rmk}
+As an important special case of these inference rules we have the following 
+valid inference rules:
+\begin{align*}
+& \inference
+  { \jfam{\Gamma}{A}
+    \jterm{\Gamma}{B}{y}
     \jfam{{\Gamma}{B}}{Q}
@@
   { \jfameq
-      {{{{\Gamma}{A}}{P}}{\ctxwk{P}{{A}{B}}}}
-      {\ctxwk{\ctxext{A}{P}}{Q}}
-      {\ctxwk{P}{{A}{Q}}}
-    }
-  \label{comp-ew-f}\\
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{{\Gamma}{A}}{P}
+      {{\Gamma}{A}}
+      {\ctxwk{A}{\subst{y}{Q}}}
+      {\subst{\ctxwk{A}{y}}{\ctxwk{A}{Q}}}
+    }
+  \\
+& \inference
+  { \jfam{\Gamma}{A}
+    \jterm{\Gamma}{B}{y}
+    \jfam{{{\Gamma}{B}}{Q}}{R}
+    }
+  { \jfameq
+      {{{\Gamma}{A}}{\ctxwk{A}{\subst{y}{Q}}}}
+      {\ctxwk{A}{\subst{y}{R}}}
+      {\subst{\ctxwk{A}{y}}{\ctxwk{A}{R}}}
+    }
+  \\
+& \inference
+  { \jfam{\Gamma}{A}
+    \jterm{\Gamma}{B}{y}
+    \jterm{{{\Gamma}{B}}{Q}}{R}{h}
+    }
+  { \jtermeq
+      {{{\Gamma}{A}}{\ctxwk{A}{\subst{y}{Q}}}}
+      {\ctxwk{A}{\subst{y}{R}}}
+      {\ctxwk{A}{\subst{y}{h}}}
+      {\subst{\ctxwk{A}{y}}{\ctxwk{A}{h}}}
+    }
+\end{align*}
+Moreover, we get
+\begin{equation*}
+\inference
+  { \jfam{\Gamma}{A}
+    \jterm{\Gamma}{B}{y}
     \jterm{{\Gamma}{B}}{Q}{g}
@@
   { \jtermeq
-      {{{{\Gamma}{A}}{P}}{\ctxwk{P}{{A}{B}}}}
-      {\ctxwk{P}{{A}{Q}}}
-      {\ctxwk{\ctxext{A}{P}}{g}}
-      {\ctxwk{P}{{A}{g}}}
-    } 
-  \label{comp-ew-t}
-\end{align}
+      {{\Gamma}{A}}
+      {\ctxwk{A}{\subst{y}{Q}}}
+      {\ctxwk{A}{\subst{y}{g}}}
+      {\subst{\ctxwk{A}{y}}{\ctxwk{A}{g}}}
+    }
+\end{equation*}
+\end{rmk}
 
@@
 
+\subsubsection{The identity term of a substituted family}
+\label{comp-si}
+The identity term of a substituted family is the substitution of the identity term
+\begin{equation}
+\inference
+  { \jterm{\Gamma}{A}{x}
+    \jfam{{{\Gamma}{A}}{P}}{Q}
+    }
+  { \unfoldall{\jhomeq
+      {{\Gamma}{\subst{x}{P}}}
+      {\subst{x}{Q}}
+      {\subst{x}{Q}}
+      {\subst{x}{\idtm{Q}}}
+      {\idtm{\subst{x}{Q}}}
+    }}
+  \label{comp-si-t}
+\end{equation}
+
+\begin{rmk}
+An important special case is the judgmental equality
+\begin{equation*}
+\unfoldall{\jhomeq
+      {\Gamma}
+      {\subst{x}{P}}
+      {\subst{x}{P}}
+      {\subst{x}{\idtm{P}}}
+      {\idtm{\subst{x}{P}}}
+    }
+\end{equation*}
+for a family $\jfam{{\Gamma}{A}}{P}$.
+\end{rmk}
+
 \subsubsection{The cancellation properties of identity terms}
@@
   \\
-%An unnecessary rule:
-%& \inference
-%  { \jfam{{\Gamma}{A}}{P}
-%    }
-%  { \jfameq
-%      {{\Gamma}{A}}
-%      {\subst{\idtm{A}}{\ctxwk{{A}{A}}{P}}}
-%      {P}
-%    }
-%  \label{idfunc-wk-defn2}\\
 & \inference
@@
   \label{precomp-idtm-t}
-%An unnecessary rule:
-%& \inference
-%  { \jterm{{\Gamma}{A}}{P}{f}
-%    }
-%  { \jtermeq
-%      {{\Gamma}{A}}
-%      {P}
-%      {\subst{\idtm{A}}{\ctxwk{{A}{A}}{f}}}
-%      {f}
-%    }
-%  \label{precomp-idtm-t}
-\end{align}
-
+\end{align}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{The possiblity of types in the theory of contexts, families and

### 2014-07-03 article-models-ttinternal.tex
reorganized to match content of article-models-ttinternal.tex

@@
   \ar{r}[swap]{\cfthomc{\cfthom{f}}}
-& \cftalgc{\cftalg{A}}
+& \cftalgc{\cftalg{B}}
 \end{tikzcd}
@@
 \end{thm}
+
+As with ordinary morphisms, extension homomorphism can be considered \emph{over} a
+morphism.
+
+\begin{defn}
+Let $\jhom{\Gamma}{A}{B}{f}$ be a morphism, let $\extalg{P}$ be an extension
+algebra in context $\ctxext{\Gamma}{A}$ and let $\extalg{Q}$ be an extension
+algebra in context $\ctxext{\Gamma}{B}$. An \emph{extension homomorphism from
+$\cftalg{P}$ to $\cftalg{Q}$ over $f$ in context $\Gamma$} is defined to be
+an extension homomorphism from $\extalg{P}$ to $\jcomp{A}{f}{\extalg{Q}}$ in
+context $\ctxext{\Gamma}{A}$. 
+\end{defn}
+
+\begin{eg}
+Let $\cfthom{f}$ be an extension homomorphism from $\extalg{A}$ to $\extalg{B}$
+in context $\Gamma$. Then $\boldsymbol{\mathcal{F}}_{\cfthom{f}}$ is an
+extension homomorphism from $\extalg{F_A}$ to $\extalg{F_B}$ over
+$\cfthomc{\cfthom{f}}$ in context $\Gamma$.
+\end{eg}
 
@@
 satisfying judgmental equalities which express abstracted analogues of
-the rules in sections \autoref{comp-ww,comp-ew,comp-0w}. The
+the rules in \autoref{comp-ww,comp-ew,comp-0w}. The
 definition of weakening algebras can be found in \autoref{sec:cftwkalg-defn}.
@@
 \end{equation*}
+In particular, we see that $\cftwk{\cftalg{A}}$ is a CFT-homomorphism from
+$\ctxwk{\cftalgf{\cftalg{A}}}{\cftalg{F_A}}$ to $\cftalg{F_A}$ over
+$\cftctxext$ in context $\Gamma$.
 
@@
     }
-& \jcomp{}{\cfthomc{\cfthom{w}}}{\cftalg{F_{F_{F_{F_A}}}}}
+& \jcomp{}
+    {\ctxwk{\cftalgf{\cftalg{A}}}{\cftfamext}}
+    {\jcomp{}{\cfthomc{\cfthom{w}}}{\cftalg{F_{F_{F_A}}}}}
 \end{tikzcd}
@@
       }$ %
-commutes. In this diagram, it is not immediately obvious that the domain and
-codomain of the CFT-homomorphism
-$ \jcomp{}
+commutes. In this diagram, it is not immediately obvious that 
+\begin{enumerate}
+\item the domain of the CFT-homomorphism
+$\jcomp{}{\ctxwk{\cftalgf{\cftalg{A}}}{\cftfamext}}{\boldsymbol{\mathcal{F}}_{\cfthom{w}}}$
+is the indicated domain. Note that we have the judgmental equalities
+\begin{align*}
+\jcomp{}
+  {\ctxwk{\cftalgf{\cftalg{A}}}{\cftfamext}}
+  {\ctxwk{\cftalgf{\cftalg{A}}}{\cftalgf{\cftalg{F_A}}}}
+& \jdeq
+\ctxwk
+  {\cftalgf{\cftalg{A}}}
+  {\jcomp{}{\cftfamext}{\cftalg{F_{F_A}}}}
+  \\
+& \jdeq
+\ctxwk
+  {\cftalgf{\cftalg{A}}}
+  {\jcomp{}{\jcomp{}{\cftctxext}{\cftfamext}}{\cftalg{F_{F_A}}}}
+  \\
+& \jdeq
+\ctxwk
+  {\cftalgf{\cftalg{A}}}
+  {\cftalg{F_{F_{F_A}}}}.
+\end{align*}
+\item the domain and codomain of the CFT-homomorphism
+\begin{equation*}
+\jcomp{}
       { \cfthomf{\cfthom{w}}}
@@
             }
-        }$ %
+        }
+\end{equation*}
 from
@@
         }
-$ are indeed judgmentally equal to the indicated domain and codomain.
+$ are indeed judgmentally equal to the indicated domain and codomain. To see
+this, we have the following lemma:
+\end{enumerate}
+
+\begin{lem}
+The following three inference rules are valid
+\end{lem}
 
@@
 
+\subsubsection{Substitution is compatible with itself}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2014-07-04 article-models-ttinternal.tex
preamble cleanup

@@
 \begin{align*}
-\jhom*{\Gamma}{{A}{P}}{A}{e_0}\\
-\jhom*{{\Gamma}{A}}{{P}{\jcomp{}{e_0}{P}}}{P}{e_1}.
+\jalign\jhom{\Gamma}{{A}{P}}{A}{e_0}\\
+\jalign\jhom{{\Gamma}{A}}{{P}{\jcomp{}{e_0}{P}}}{P}{e_1}.
 \end{align*}
@@
 \begin{align*}
-\jhom*
+\jalign\jhom
   {\Gamma}
@@
 \begin{align}
-\jfameq*{\Gamma}{\subst{x}{P}}{A}
+\jalign\jfameq{\Gamma}{\subst{x}{P}}{A}
   \label{eq1:empalg-mor}
 %  \\
-%\jtermeq*{\Gamma}{A}{\subst{x}{y}}{x}.
+%\jalign\jtermeq{\Gamma}{A}{\subst{x}{y}}{x}.
 %  \label{eq2:empalg-mor}
@@
 \begin{align*}
-\jterm*{\Gamma}{\cftalgc{\cftalg{A}}}{\cftempc{\cftalg{A}}}
-  \\
-\jterm*{{\Gamma}{\cftalgc{\cftalg{A}}}}{\cftalgf{\cftalg{A}}}{\cftempf{\cftalg{A}}}
+\jalign\jterm{\Gamma}{\cftalgc{\cftalg{A}}}{\cftempc{\cftalg{A}}}
+  \\
+\jalign\jterm{{\Gamma}{\cftalgc{\cftalg{A}}}}{\cftalgf{\cftalg{A}}}{\cftempf{\cftalg{A}}}
 \end{align*}
@@
 \begin{align}
-\jfameq*
+\jalign\jfameq
   {\Gamma}
@@
   {\jcomp{}{\cftctxext}{\cftalg{F_A}}}
-  {\cfthom{w}}
+  {\cftwk{\cftalg{A}}}
 \end{equation*}
@@
         {{\cftalgf{\cftalg{A}}}{\cftalgf{\cftalg{F_A}}}}
-        {\boldsymbol{\mathcal{F}}_{\cfthom{w}}}
-      }
-  \ar{r}{\ctxwk{\cftalgf{\cftalg{A}}}{\jcomp{}{\cftctxext}{\cfthom{w}}}}
+        {\boldsymbol{\mathcal{F}}_{\cftwk{\cftalg{A}}}}
+      }
+  \ar{r}{\ctxwk{\cftalgf{\cftalg{A}}}{\jcomp{}{\cftctxext}{\cftwk{\cftalg{A}}}}}
 & \ctxwk{\cftalgf{\cftalg{A}}}{\cftalg{F_{F_{F_A}}}}
-  \ar{d}{\jcomp{}{\ctxwk{\cftalgf{\cftalg{A}}}{\cftfamext}}{\boldsymbol{\mathcal{F}}_{\cfthom{w}}}}
-  \\
-\ctxwk{{\cftalgf{\cftalg{A}}}{\cftalgf{\cftalg{F_A}}}}{\jcomp{}{\cfthomc{\cfthom{w}}}{\cftalg{F_{F_{F_A}}}}}
+  \ar{d}{\jcomp{}{\ctxwk{\cftalgf{\cftalg{A}}}{\cftfamext}}{\boldsymbol{\mathcal{F}}_{\cftwk{\cftalg{A}}}}}
+  \\
+\ctxwk{{\cftalgf{\cftalg{A}}}{\cftalgf{\cftalg{F_A}}}}{\jcomp{}{\cftwkc{\cftalg{A}}}{\cftalg{F_{F_{F_A}}}}}
   \ar{r}[swap]{%
     \jcomp{}
-      { \cfthomf{\cfthom{w}}}
+      { \cftwkf{\cftalg{A}}}
       { \jcomp{}
-          {\cfthomc{\cfthom{w}}}
+          {\cftwkc{\cftalg{A}}}
           { \jcomp{}
@@
                   {\cftctxext}
-                  {\cfthom{w}}
+                  {\cftwk{\cftalg{A}}}
                 }
@@
     {\ctxwk{\cftalgf{\cftalg{A}}}{\cftfamext}}
-    {\jcomp{}{\cfthomc{\cfthom{w}}}{\cftalg{F_{F_{F_A}}}}}
+    {\jcomp{}{\cftwkc{\cftalg{A}}}{\cftalg{F_{F_{F_A}}}}}
 \end{tikzcd}
@@
 \item the domain of the CFT-homomorphism
-$\jcomp{}{\ctxwk{\cftalgf{\cftalg{A}}}{\cftfamext}}{\boldsymbol{\mathcal{F}}_{\cfthom{w}}}$
+$\jcomp{}{\ctxwk{\cftalgf{\cftalg{A}}}{\cftfamext}}{\boldsymbol{\mathcal{F}}_{\cftwk{\cftalg{A}}}}$
 is the indicated domain. Note that we have the judgmental equalities
@@
 \jcomp{}
-      { \cfthomf{\cfthom{w}}}
+      { \cftwkf{\cftalg{A}}}
       { \jcomp{}
-          {\cfthomc{\cfthom{w}}}
+          {\cftwkc{\cftalg{A}}}
           { \jcomp{}
@@
 $ \jcomp{}
-      { \cfthomf{\cfthom{w}}}
+      { \cftwkf{\cftalg{A}}}
       { \jcomp{}
-          {\cfthomc{\cfthom{w}}}
+          {\cftwkc{\cftalg{A}}}
           { \jcomp{}
@@
 & \jcomp{}
-  { \subst{q}{{\apply\cftctxext{\gamma}{b}}{\cfthom{w}}}
+  { \subst{q}{{\apply\cftctxext{\gamma}{b}}{\cftwk{\cftalg{A}}}}
     }
@@
       {\apply\cftfamext{\gamma}{b}{q}}
-      {{a}{{\gamma}{\boldsymbol{\mathcal{F}}_{\cfthom{w}}}}}
+      {{a}{{\gamma}{\boldsymbol{\mathcal{F}}_{\cftwk{\cftalg{A}}}}}}
     }
@@
   \jcomp{}
-    { \subst{b}{{a}{{\gamma}{\boldsymbol{\mathcal{F}}_{\cfthom{w}}}}}
+    { \subst{b}{{a}{{\gamma}{\boldsymbol{\mathcal{F}}_{\cftwk{\cftalg{A}}}}}}
       }
     { \subst
-        { {q}{{b}{{a}{{\gamma}{\cfthomf{\cfthom{w}}}}}}}
-        { {\apply\cftctxext{\apply\cftctxext{\gamma}{a}}{\subst{b}{{a}{{\gamma}{\cfthomc{\cfthom{w}}}}}}}
-          {\cfthom{w}}
+        { {q}{{b}{{a}{{\gamma}{\cftwkf{\cftalg{A}}}}}}}
+        { {\apply\cftctxext{\apply\cftctxext{\gamma}{a}}{\subst{b}{{a}{{\gamma}{\cftwkc{\cftalg{A}}}}}}}
+          {\cftwk{\cftalg{A}}}
           }
@@
 $ \subst
-      { {q}{{b}{{a}{{\gamma}{\cfthomf{\cfthom{w}}}}}}
-        }
-      { {{b}{{a}{{\gamma}{\cfthomc{\cfthom{w}}}}}}
+      { {q}{{b}{{a}{{\gamma}{\cftwkf{\cftalg{A}}}}}}
+        }
+      { {{b}{{a}{{\gamma}{\cftwkc{\cftalg{A}}}}}}
         {{a}{{\gamma}{\cftalg{F_{F_{F_{F_{A}}}}}}}}
@@
   {\cftalg{F_A}}
-  \ar{r}{\ctxwk{\cftalgf{\cftalg{F_A}}}{\cfthom{w}}}
-  \ar{dr}[swap]{\jcomp{}{\cftfamext}{\cfthom{w}}}
+  \ar{r}{\ctxwk{\cftalgf{\cftalg{F_A}}}{\cftwk{\cftalg{A}}}}
+  \ar{dr}[swap]{\jcomp{}{\cftfamext}{\cftwk{\cftalg{A}}}}
 & \ctxwk{\cftalgf{\cftalg{F_A}}}{\cftalg{F_{F_A}}}
-  \ar{d}{\jcomp{}{\cftctxext}{\cfthom{w}}}
+  \ar{d}{\jcomp{}{\cftctxext}{\cftwk{\cftalg{A}}}}
   \\
@@
   {\jcomp{}{\cftctxext}{\cftalg{F_A}}}
-  {\cfthom{w}}
+  {\cftwk{\cftalg{A}}}
 $
@@
 \begin{equation*}
-\subst{\apply\cftfamext{\gamma}{a}{p}}{{\gamma}{\cfthom{w}}}
+\apply\cftwk{\cftalg{A}}{\gamma}{\apply\cftfamext{\gamma}{a}{p}}
 \jdeq
 \jcomp{}
-    {\subst{a}{{\gamma}{\cfthom{w}}}}
-    {\subst{p}{{\apply\cftctxext{\gamma}{a}}{\cfthom{w}}}}
+    {\apply\cftwk{\cftalg{A}}{\gamma}{a}}
+    {\apply\cftwk{\cftalg{A}}{\apply\cftctxext{\gamma}{a}}{p}}
 \end{equation*}
@@
     }
-  { \subst{\cftempf{\cftalg{A}}}{\cfthom{w}}
+  { \subst{\cftempf{\cftalg{A}}}{\cftwk{\cftalg{A}}}
     }

### 2014-07-08 article-models-ttderived.tex
source readability improvement
preamble cleanup
small changes

@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Morphisms}\label{morphisms}
@@
 \begin{defn}
-We define the judgment\begin{equation*}
+We define the judgment
+\begin{equation*}
 \jhom{\Gamma}{A}{B}{f},
@@
 which is pronounced as `$f$ is a morphism from $A$ to $B$ in context $\Gamma$',
-to be the judgment\begin{equation*}
+to be the judgment
+\begin{equation*}
 \unfold{\jhom{\Gamma}{A}{B}{f}}.
 \end{equation*}
-Likewise, we define the judgment\begin{equation*}
+Likewise, we define the judgment
+\begin{equation*}
 \jhomeq{\Gamma}{A}{B}{f}{f'}
 \end{equation*}
-to be the judgment\begin{equation*}
+to be the judgment
+\begin{equation*}
 \unfold{\jhomeq{\Gamma}{A}{B}{f}{f'}}.
@@
 
-\begin{lem}
+\begin{lem}\label{lem:jhomdomext-jhomcodwk}
 We have the following valid inference rules expressing that a morphism from
@@
 a family $\jfam{{{\Gamma}{B}}{Q}}{R}$ and a term $\jterm{{{\Gamma}{B}}{Q}}{R}{h}$.
-We define\begin{align*}
-\jfamdefn*
+We define
+\begin{align*}
+\jalign\jfamdefn
   {{\Gamma}{A}}
@@
   \\
-\jfamdefn*
+\jalign\jfamdefn
   {{{\Gamma}{A}}{\jcomp{A}{f}{Q}}}
@@
   \\
-\jtermdefn*
+\jalign\jtermdefn
   {{{\Gamma}{A}}{\jcomp{A}{f}{Q}}}
@@
   \subst{f}{\idtm{\ctxwk{A}{B}}}
-  \tag{by \autoref{comp-wi}}
+  \tag{by \autoref{comp-wi-t}}
   \\
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Projections and extension on terms}\label{extension-on-terms}
@@
 \begin{align*}
-\jhomdefn*
+\jalign\jhomdefn
   {\Gamma}
@@
   \\
-\jtermdefn*
+\jalign\jtermdefn
   {\ctxext{\Gamma}{{A}{P}}}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Family morphisms}\label{sec:fhom}
@@
 \begin{align*}
-\jhomeq*{\Gamma}{A}{B}{\jcomp{}{\tmext{f_0}{f_1}}{\cprojfstf{B}{Q}}}{f_0}\\
-\jtermeq*{{\Gamma}{A}}{\jcomp{}{f_0}{Q}}{\jcomp{}{\tmext{f_0}{f_1}}{\cprojsndf{B}{Q}}}{f_1}.
+\jalign\jhomeq{\Gamma}{A}{B}{\jcomp{}{\tmext{f_0}{f_1}}{\cprojfstf{B}{Q}}}{f_0}\\
+\jalign\jtermeq{{\Gamma}{A}}{\jcomp{}{f_0}{Q}}{\jcomp{}{\tmext{f_0}{f_1}}{\cprojsndf{B}{Q}}}{f_1}.
 \end{align*}
@@
 
-\begin{rmk}
+\begin{rmk}\label{rmk:jhom-as-jfhom}
 The judgment $\jfhom{\Gamma}{A}{B}{f}{P}{Q}{F}$ means the same thing as
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Fiber inclusions}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{The terms of the empty families}
@@
 \end{proof}
+
+\subsection{Morphisms with extended domains}
+
+We cover some properties of morphisms with extended domains because we will
+need them later on.
+
+\begin{lem}
+Let $\jhom{\Gamma}{{A}{P}}{B}{f}$ be a morphism and let $\jterm{\Gamma}{A}{x}$
+be a term. The following three inference rules are valid
+\begin{align*}
+& \inference
+  { \jfam{{\Gamma}{B}}{Q}
+    }
+  { \jfameq
+      {{\Gamma}{\subst{x}{P}}}
+      {\subst{x}{\jcomp{}{f}{Q}}}
+      {\jcomp{}{\subst{x}{f}}{Q}}
+    }
+  \\
+& \inference
+  { \jfam{{{\Gamma}{B}}{Q}}{R}
+    }
+  { \jfameq
+      {{{\Gamma}{\subst{x}{P}}}{\jcomp{}{\subst{x}{f}}{Q}}}
+      {\subst{x}{\jcomp{}{f}{R}}}
+      {\jcomp{}{\subst{x}{f}}{R}}
+    }
+  \\
+& \inference
+  { \jterm{{{\Gamma}{B}}{Q}}{R}{h}
+    }
+  { \jfameq
+      {{{\Gamma}{\subst{x}{P}}}{\jcomp{}{\subst{x}{f}}{Q}}}
+      {\jcomp{}{\subst{x}{f}}{R}}
+      {\subst{x}{\jcomp{}{f}{h}}}
+      {\jcomp{}{\subst{x}{f}}{h}}
+    }
+\end{align*}
+\end{lem}
+
+\begin{proof}
+This follows from the observation in \autoref{lem:jhomdomext-jhomcodwk} 
+that a morphism from
+$\ctxext{A}{P}$ to $B$ in context $\Gamma$ is the same thing as a morphism
+from $P$ to $\ctxwk{A}{B}$ in context $\ctxext{\Gamma}{A}$.
+\end{proof}
+
+\begin{lem}
+Consider a morphism $\jhom{{\Gamma}{A}}{{P}{Q}}{R}{f}$ and a morphism
+$\jhom{\Gamma}{{A}{R}}{B}{g}$. Then the diagram
+\begin{equation*}
+\begin{tikzcd}
+\ctxext{A}{{P}{Q}}
+  \ar{r}{\jvcomp{}{\idtm{A}}{f}}
+  \ar{dr}[swap]{\jcomp{}{f}{g}}
+  &
+\ctxext{A}{R}
+  \ar{d}{g}
+  \\
+  &
+B
+\end{tikzcd}
+\end{equation*}
+in context $\Gamma$ commutes judgmentally.
+\end{lem}
+
+\begin{proof}
+It is just a triviality, as we have the judgmental equalities
+\begin{align*}
+\jcomp{}{f}{g}
+& \jdeq
+  \jcomp{}{f}{\jcomp{}{\idtm{A}}{g}}
+  \tag{by \autoref{precomp-idtm-t}}
+  \\
+& \jdeq
+  \jcomp{}{\jvcomp{}{\idtm{A}}{f}}{g}.
+  \tag{by \autoref{lem:composition-threesome}}
+\end{align*}
+Note that we may apply \autoref{lem:composition-threesome} because by
+\autoref{rmk:jhom-as-jfhom} every ordinary morphism is a morphism over the
+identity morphism.
+\end{proof}

### 2014-07-08 article-models-ttinternal.tex
small changes

@@
 \begin{defn}\label{defn:extension-algebras}
+An \emph{extension algebra $\extalg{A}$ in context $\Gamma$} is a quintuple
+\begin{equation*}
+\unfold{\extalg{A}}
+\end{equation*}
+consisting of a family $\jextalgctx{\Gamma}{A}$ of \emph{$\extalg{A}$-contexts}, 
+a family $\jextalgfam{\Gamma}{A}$ of \emph{$\extalg{A}$-families}, a family 
+$\jextalgtm{\Gamma}{A}$ of \emph{$\extalg{A}$-terms}, a term
+$\jextalgctxext{\Gamma}{A}$ called the \emph{$\extalg{A}$-context extension}
+and a term $\jextalgfamext{\Gamma}{A}$ called the \emph{$\extalg{A}$-family
+extension}, satisfying the judgmental equalities of
+\autoref{eq:extalg-eq1,eq:extalg-eq2}.
+\begin{comment}
 We define the judgment
@@
 \begin{align*}
-\jextalgctx*{\Gamma}{A}
-  \\
-\jextalgfam*{\Gamma}{A}
-  \\
-\jextalgtm*{\Gamma}{A}
-  \\
-\jextalgctxext*{\Gamma}{A}
-  \\
-\jextalgfamext*{\Gamma}{A}
-  \\
-\jhomeq*
+\jalign\jextalgctx{\Gamma}{A}
+  \\
+\jalign\jextalgfam{\Gamma}{A}
+  \\
+\jalign\jextalgtm{\Gamma}{A}
+  \\
+\jalign\jextalgctxext{\Gamma}{A}
+  \\
+\jalign\jextalgfamext{\Gamma}{A}
+  \\
+\jalign\jhomeq
   { \Gamma}
@@
   \\
-\jhomeq*
+\jalign\jhomeq
   { {\Gamma}{\cftalgc{\cftalg{A}}}}
@@
 satisfying the judgmental equalities displayed in the diagrams in
-\autoref{eq:extalg-eq1,eq:extalg-eq2}. Extension algebras are judgmentally
+\autoref{eq:extalg-eq1,eq:extalg-eq2}.
+\end{comment}
+Extension algebras are judgmentally
 equal if they are componentwise judgmentally equal.
@@
     {\cftctxext[\cftalg{A}]}
+  \tag{by \autoref{comp-es}}
   \\
@@
   {\subst{\apply\cftctxext{\gamma}{a}}{\cftalgf{\cftalg{A}}}}
-  {\subst{b}{{a}{{\gamma}{\cftwkc{\cftalg{A}}}}}}
+  {\apply\cftwkc{\cftalg{A}}{\gamma}{a}{b}}
 \end{equation*}
@@
           {\apply\cftctxext{\gamma}{a}}
-          {\subst{b}{{a}{{\gamma}{\cftwkc{\cftalg{A}}}}}}
+          {\apply\cftwkc{\cftalg{A}}{\gamma}{a}{b}}
         }
       { \cftalgf{\cftalg{A}}}}
-  { \subst{q}{{b}{{a}{{\gamma}{\cftwkf{\cftalg{A}}}}}}}
+  { \apply\cftwkf{\cftalg{A}}{\gamma}{a}{b}{q}}
 \end{equation*}
@@
   { \subst
-      { {q}{{b}{{a}{{\gamma}{\cftwkf{\cftalg{A}}}}}}}
+      { \apply\cftwkf{\cftalg{A}}{\gamma}{a}{b}{q}}
       { { \apply\cftctxext
             {\apply\cftctxext{\gamma}{a}}
-            {{b}{{a}{{\gamma}{\cftwkc{\cftalg{A}}}}}}
+            {\apply\cftwkc{\cftalg{A}}{\gamma}{a}{b}}
           }
@@
     { \subst
-        { {q}{{b}{{a}{{\gamma}{\cftwkf{\cftalg{A}}}}}}}
-        { {\apply\cftctxext{\apply\cftctxext{\gamma}{a}}{\subst{b}{{a}{{\gamma}{\cftwkc{\cftalg{A}}}}}}}
+        { \apply\cftwkf{\cftalg{A}}{\gamma}{a}{b}{q}}
+        { {\apply\cftctxext{\apply\cftctxext{\gamma}{a}}{\apply\cftwkc{\cftalg{A}}{\gamma}{a}{b}}}
           {\cftwk{\cftalg{A}}}
@@
 $ \subst
-      { {q}{{b}{{a}{{\gamma}{\cftwkf{\cftalg{A}}}}}}
-        }
-      { {{b}{{a}{{\gamma}{\cftwkc{\cftalg{A}}}}}}
+      { \apply\cftwkf{\cftalg{A}}{\gamma}{a}{b}{q}
+        }
+      { {\apply\cftwkc{\cftalg{A}}{\gamma}{a}{b}}
         {{a}{{\gamma}{\cftalg{F_{F_{F_{F_{A}}}}}}}}
@@
 
-\subsubsection{Substitution is compatible with itself}
+\begin{defn}
+Let $\cftalg{A}$ be a CFT-algebra in context $\Gamma$ and let
+\begin{equation*}
+\jhom
+  {{{{\Gamma}{\cftalgc{\cftalg{A}}}}{\cftalgf{\cftalg{A}}}}{\cftalgt{\cftalg{A}}}}
+  {\ctxwk{\cftalgt{\cftalg{A}}}{\cftalg{F_{F_A}}}}
+  {\cftalg{F_A}}
+  {\cftsubst{\cftalg{A}}}
+\end{equation*}
+be a CFT-homomorphism. We say that the pair $(\cftalg{A},\cftsubst{\cftalg{A}})$
+is a substitution algebra in context $\Gamma$ if the diagram
+\begin{equation}\label{diag:substalg}
+\begin{tikzcd}
+X
+  \ar{r}
+  \ar{d}
+& X
+  \ar{d}
+  \\
+X 
+  \ar{r}
+& X
+\end{tikzcd}
+\end{equation}
+of CFT-algebras in context ... commutes judgmentally.
+\end{defn}
+
+\begin{lem}
+Let $\cftalg{A}$ be a substitution algebra in context $\Gamma$. Then we have the
+judgmental equality
+\begin{align*}
+& \jcomp{}
+    { \apply\cftsubst{\cftalg{A}}%
+      { \apply\cftctxext%
+          { \apply\cftctxext{\gamma}{a}}{p}
+        }
+      { q}
+      { g}
+      }
+    { \subst
+        { q}
+        { { p}
+          { \boldsymbol{\mathcal{F}}_{\apply\cftsubst{\cftalg{A}}{\gamma}{a}{x}}}
+          }
+      }
+  \\
+& \jdeq
+\end{align*}
+of CFT-homomorphisms from ... to ... in context $\Gamma$,
+for $\gamma:A$, $a:\subst{\gamma}{\cftalgf{\cftalg{A}}}$, 
+$x:\subst{a}{{\gamma}{\cftalgt{\cftalg{A}}}}$,
+$p:\subst{\apply\cftctxext{\gamma}{a}}{\cftalgf{\cftalg{A}}}$
+$q:\subst{\apply\cftctxext{\apply\cftctxext{\gamma}{a}}{p}}{\cftalgf{\cftalg{A}}}$
+and $g:\subst{\apply\cftctxext{\apply\cftctxext{\apply\cftctxext{\gamma}{a}}{p}}{q}}{\cftalgt{\cftalg{A}}}$
+\end{lem}
 

### 2014-07-11 article-models-ttderived.tex
added properties to extension algebras and cft-algebras

@@
 
-\subsection{Morphisms with extended domains}
+\subsection{Morphisms with extended domains}\label{sec:ehom}
 
@@
 
-\begin{lem}
+\begin{rmk}
+Suppose $\jhom{\Gamma}{{A}{P}}{B}{f}$ is a morphism with an extended domain.
+Then the inference rules
+\begin{align*}
+& \inference
+  { \jterm{\Gamma}{A}{x}
+    }
+  { \jhom{\Gamma}{\subst{x}{P}}{B}{\subst{x}{f}}
+    }
+  \\
+& \inference
+  { \jterm{{\Gamma}{A}}{P}{y}
+    }
+  { \jhom{\Gamma}{A}{B}{\subst{y}{f}}
+    }
+\end{align*}
+are valid.
+This follows from the observation in \autoref{lem:jhomdomext-jhomcodwk} 
+that a morphism from
+$\ctxext{A}{P}$ to $B$ in context $\Gamma$ is the same thing as a morphism
+from $P$ to $\ctxwk{A}{B}$ in context $\ctxext{\Gamma}{A}$.
+\end{rmk}
+
+\begin{lem}\label{lem:ehom-subst}
 Let $\jhom{\Gamma}{{A}{P}}{B}{f}$ be a morphism and let $\jterm{\Gamma}{A}{x}$
@@
     }
-  { \jfameq
+  { \jtermeq
       {{{\Gamma}{\subst{x}{P}}}{\jcomp{}{\subst{x}{f}}{Q}}}
@@
 \begin{proof}
-This follows from the observation in \autoref{lem:jhomdomext-jhomcodwk} 
-that a morphism from
-$\ctxext{A}{P}$ to $B$ in context $\Gamma$ is the same thing as a morphism
-from $P$ to $\ctxwk{A}{B}$ in context $\ctxext{\Gamma}{A}$.
+Let $Q$ be a family over $\ctxext{\Gamma}{A}$.
+We have the judgmental equalities
+\begin{align*}
+\subst{x}{\jcomp{}{f}{Q}}
+& \jdeq
+  \subst
+    {x}
+    {{f}{\ctxwk{\ctxext{A}{P}}{Q}}}
+  \tag{by definition}
+  \\
+& \jdeq
+  \subst
+    { {x}{f}
+      }
+    { {x}
+      {\ctxwk{\ctxext{A}{P}}{Q}}
+      }
+  \tag{by \autoref{comp-ss-f}}
+  \\
+& \jdeq
+  \subst
+    { {x}{f}
+      }
+    { {x}
+      {\ctxwk{P}{{A}{Q}}}
+      }
+  \tag{by \autoref{comp-ew-f}}
+  \\
+& \jdeq
+  \subst
+    { {x}{f}
+      }
+    { \ctxwk
+        {\subst{x}{P}}
+        {\subst{x}{\ctxwk{A}{Q}}}
+      }
+  \tag{by \autoref{comp-sw-f}}
+  \\
+& \jdeq
+  \subst
+    { {x}{f}
+      }
+    { \ctxwk
+        {\subst{x}{P}}
+        {Q}
+      }
+  \tag{by \autoref{cancellation-ws-f}}
+  \\
+& \jdeq
+  \jcomp{}{\subst{x}{f}}{Q}.
+  \tag{by definition}
+\end{align*}
+\begin{comment}
+In the case where $\xi$ is $y$ we have the judgmental equalities
+\begin{align*}
+\subst{y}{\jcomp{}{f}{Q}}
+& \jdeq
+  \subst
+    { y
+      }
+    { {f}
+      {\ctxwk{\ctxext{\cftalgc{\cftalg{A}}}{\cftalgf{\cftalg{A}}}}{Q}}
+      }
+  \tag{by definition}
+  \\
+& \jdeq
+  \subst
+    { {y}{f}
+      }
+    { {y}
+      {\ctxwk{\ctxext{\cftalgc{\cftalg{A}}}{\cftalgf{\cftalg{A}}}}{Q}}
+      }
+  \tag{by \autoref{comp-ss-f}}
+  \\
+& \jdeq
+  \subst
+    { {y}{f}
+      }
+    { {y}
+      {\ctxwk{\cftalgf{\cftalg{A}}}{{\cftalgc{\cftalg{A}}}{Q}}}
+      }
+  \tag{by \autoref{comp-ew-f}}
+  \\
+& \jdeq
+  \subst
+    { {y}{f}
+      }
+    { \ctxwk{\cftalgc{\cftalg{A}}}{Q}
+      }
+  \tag{by \autoref{cancellation-ws-f}}
+  \\
+& \jdeq
+  \jcomp{}{\subst{y}{f}}{Q}.
+  \tag{by definition}
+\end{align*}
+\end{comment}
 \end{proof}

### 2014-07-11 article-models-ttinternal.tex
added properties to extension algebras and cft-algebras

@@
 
+\begin{lem}
+Let $\extalg{Q}$ be an extension algebra in context $\ctxext{\Gamma}{B}$ and let
+$\jfam{\Gamma}{A}$. Then we have the judgmental equality
+\begin{equation*}
+\ctxwk{A}{\extalg{F_Q}}\jdeq\extalg{F}_{\ctxwk{A}{\extalg{Q}}}
+\end{equation*}
+of extension algebras in context $\ctxext{{\Gamma}{A}}{\ctxwk{A}{B}}$.
+\end{lem}
+
+\begin{lem}
+Let $\extalg{Q}$ be an extension algebra in context $\ctxext{{\Gamma}{A}}{P}$ 
+and let $\jterm{\Gamma}{A}{x}$. Then we have the judgmental equality
+\begin{equation*}
+\subst{x}{\extalg{F_Q}}\jdeq\extalg{F}_{\subst{x}{\extalg{Q}}}
+\end{equation*}
+of extension algebras in context $\ctxext{\Gamma}{\subst{x}{P}}$.
+\end{lem}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
   \\
-\jfhom*
+\jalign\jfhom
   {\Gamma}
@@
 satisfying (among other conditions) that families over the empty context in
-$\cftalg{A}$ are just the contexts of $\cftalg{A}$. We will begin with an 
-auxilary lemma which will help us dealing with the internal empty contexts:
-
-\begin{lem}\label{lem:empalg-mor}
-Consider a family $\jfam{\Gamma}{A}$ with a term $\jterm{\Gamma}{A}{x}$ and 
-a family $\jfam{{\Gamma}{A}}{P}$ % with a term $\jterm{{\Gamma}{A}}{P}{y}$
-satisfying the judgmental equality
-\begin{align}
-\jalign\jfameq{\Gamma}{\subst{x}{P}}{A}
-  \label{eq1:empalg-mor}
-%  \\
-%\jalign\jtermeq{\Gamma}{A}{\subst{x}{y}}{x}.
-%  \label{eq2:empalg-mor}
-\end{align}
-Consider also a morphism
-\begin{equation*}
-\jhom{\Gamma}{{A}{P}}{B}{f}.
-\end{equation*}
-Then $\subst{\xi}{f}$ is a morphism from $A$ to $B$ 
-in context $\Gamma$ and the following inference rules are valid for $\xi$ being 
-$x$ or $y$:
-\begin{align*}
-& \inference
-  { \jfam{{\Gamma}{B}}{Q}
-    }
-  { \jfameq
-      {{\Gamma}{A}}
-      {\subst{\xi}{\jcomp{}{f}{Q}}}
-      {\jcomp{}{\subst{\xi}{f}}{Q}}
-    }
-  \\
-& \inference
-  { \jfam{{{\Gamma}{B}}{Q}}{R}
-    }
-  { \jfameq
-      {{{\Gamma}{A}}{\jcomp{}{\subst{\xi}{f}}{Q}}}
-      {\subst{\xi}{\jcomp{}{f}{R}}}
-      {\jcomp{}{\subst{\xi}{f}}{R}}
-    }
-  \\
-& \inference
-  { \jterm{{{\Gamma}{B}}{Q}}{R}{h}
-    }
-  { \jtermeq
-      {{{\Gamma}{A}}{\jcomp{}{\subst{\xi}{f}}{Q}}}
-      {\jcomp{}{\subst{\xi}{f}}{R}}
-      {\subst{\xi}{\jcomp{}{f}{h}}}
-      {\jcomp{}{\subst{\xi}{f}}{h}}
-    }
-\end{align*}
-\end{lem}
-
-\begin{proof}
-We only prove the first inference rule in both cases.
-Let $Q$ be a family over $\ctxext{\Gamma}{\cftalgc{\cftalg{A}}}$. In the case 
-where $\xi$ is $x$ we have the judgmental equalities
-\begin{align*}
-\subst{x}{\jcomp{}{f}{Q}}
-& \jdeq
-  \subst
-    {x}
-    {{f}{\ctxwk{\ctxext{\cftalgc{\cftalg{A}}}{\cftalgf{\cftalg{A}}}}{Q}}}
-  \tag{by definition}
-  \\
-& \jdeq
-  \subst
-    { {x}{f}
-      }
-    { {x}
-      {\ctxwk{\ctxext{\cftalgc{\cftalg{A}}}{\cftalgf{\cftalg{A}}}}{Q}}
-      }
-  \tag{by \autoref{comp-ss-f}}
-  \\
-& \jdeq
-  \subst
-    { {x}{f}
-      }
-    { {x}
-      {\ctxwk{\cftalgf{\cftalg{A}}}{{\cftalgc{\cftalg{A}}}{Q}}}
-      }
-  \tag{by \autoref{comp-ew-f}}
-  \\
-& \jdeq
-  \subst
-    { {x}{f}
-      }
-    { \ctxwk
-        {\subst{x}{\cftalgf{\cftalg{A}}}}
-        {\subst{x}{\ctxwk{\cftalgc{\cftalg{A}}}{Q}}}
-      }
-  \tag{by \autoref{comp-sw-f}}
-  \\
-& \jdeq
-  \subst
-    { {x}{f}
-      }
-    { \ctxwk
-        {\subst{x}{\cftalgf{\cftalg{A}}}}
-        {Q}
-      }
-  \tag{by \autoref{cancellation-ws-f}}
-  \\
-& \jdeq
-  \subst{{x}{f}}{\ctxwk{\cftalgc{\cftalg{A}}}{Q}}
-  \tag{by \autoref{eq1:empalg-mor}}
-  \\
-& \jdeq
-  \jcomp{}{\subst{x}{f}}{Q}.
-  \tag{by definition}
-\end{align*}
-In the case where $\xi$ is $y$ we have the judgmental equalities
-\begin{align*}
-\subst{y}{\jcomp{}{f}{Q}}
-& \jdeq
-  \subst
-    { y
-      }
-    { {f}
-      {\ctxwk{\ctxext{\cftalgc{\cftalg{A}}}{\cftalgf{\cftalg{A}}}}{Q}}
-      }
-  \tag{by definition}
-  \\
-& \jdeq
-  \subst
-    { {y}{f}
-      }
-    { {y}
-      {\ctxwk{\ctxext{\cftalgc{\cftalg{A}}}{\cftalgf{\cftalg{A}}}}{Q}}
-      }
-  \tag{by \autoref{comp-ss-f}}
-  \\
-& \jdeq
-  \subst
-    { {y}{f}
-      }
-    { {y}
-      {\ctxwk{\cftalgf{\cftalg{A}}}{{\cftalgc{\cftalg{A}}}{Q}}}
-      }
-  \tag{by \autoref{comp-ew-f}}
-  \\
-& \jdeq
-  \subst
-    { {y}{f}
-      }
-    { \ctxwk{\cftalgc{\cftalg{A}}}{Q}
-      }
-  \tag{by \autoref{cancellation-ws-f}}
-  \\
-& \jdeq
-  \jcomp{}{\subst{y}{f}}{Q}.
-  \tag{by definition}
-\end{align*}
-\end{proof}
+$\cftalg{A}$ are just the contexts of $\cftalg{A}$.
 
@@
 \begin{align}
-\jtermeq*
+\jalign\jtermeq
   {\Gamma}
@@
 \begin{align}
-\jtermeq*
+\jalign\jtermeq
   {{\Gamma}{\cftalgc{\cftalg{A}}}}
@@
   \\
-\jtermeq*
+\jalign\jtermeq
   {{\Gamma}{\cftalgc{\cftalg{A}}}}
@@
 \begin{align}
-\jtermeq*
+\jalign\jtermeq
   {{{\Gamma}{\cftalgc{\cftalg{A}}}}{\cftalgf{\cftalg{A}}}}
@@
   \\
-\jtermeq*
+\jalign\jtermeq
   {{{\Gamma}{\cftalgc{\cftalg{A}}}}{\cftalgf{\cftalg{A}}}}
@@
 \begin{align}
-\jhomeq*
+\jalign\jhomeq
   {\Gamma}
@@
 \begin{enumerate}
-\item \autoref{lem:empalg-mor} implies that the morphisms 
+\item \autoref{lem:ehom-subst} implies that the morphisms 
 $\subst{\cftempc{\cftalg{A}}}{\cftctxext}$ and
@@
   \jcomp{}{\subst{\cftempf{\cftalg{A}}}{\cftctxext}}{\cftalgf{\cftalg{A}}}
-  \tag{by \autoref{lem:empalg-mor}}
+  \tag{by \autoref{lem:ehom-subst}}
   \\
@@
 \end{align*}
-Therefore we can apply \autoref{lem:empalg-mor} with the triple
+Therefore we can apply \autoref{lem:ehom-subst} with the triple
 $(\cftalgf{\cftalg{A}},\jcomp{}{\cftctxext}{\cftalgf{\cftalg{A}}},
@@
 \end{equation*}
-Note that we can apply \autoref{lem:empalg-mor}, so we get the judgmental
+Note that we can apply \autoref{lem:ehom-subst}, so we get the judgmental
 equalities
@@
   \jcomp{}{\subst{\cftempf{\cftalg{A}}}{\cftctxext}}{\cftempf{\cftalg{A}}}
-  \tag{by \autoref{lem:empalg-mor}}
+  \tag{by \autoref{lem:ehom-subst}}
   \\
@@
   \jcomp{}{\subst{\cftempf{\cftalg{A}}}{\cftctxext}}{\cftfamext}
-  \tag{by \autoref{lem:empalg-mor}}
+  \tag{by \autoref{lem:ehom-subst}}
   \\
@@
 \end{proof}
+
+\begin{thm}\label{thm:cftalg-wk}
+Let $\cftalg{Q}$ be a CFT-algebra in context $\ctxext{\Gamma}{B}$ and let
+$\jfam{\Gamma}{A}$ be a family of contexts. Then the septuple
+\begin{equation*}
+\ctxwk{A}{\cftalg{Q}}
+  \defeq
+  ( \ctxwk{A}{\cftalgc{\cftalg{Q}}},
+    \ctxwk{A}{\cftalgf{\cftalg{Q}}},
+    \ctxwk{A}{\cftalgt{\cftalg{Q}}},
+    \ctxwk{A}{\cftctxext[\cftalg{Q}]},
+    \ctxwk{A}{\cftfamext[\cftalg{Q}]},
+    \ctxwk{A}{\cftempc{\cftalg{Q}}},
+    \ctxwk{A}{\cftempf{\cftalg{Q}}})
+\end{equation*}
+is a CFT-algebra in context $\ctxext{{\Gamma}{A}}{\ctxwk{A}{B}}$.
+\end{thm}
+
+\begin{proof}
+The proof follows from the fact that weakening by $A$ is compatible with all
+the involved operations.
+\end{proof}
+
+\begin{thm}\label{thm:cftalg-subst}
+Let $\extalg{Q}$ be a CFT-algebra in context $\ctxext{{\Gamma}{A}}{P}$
+and let $\jterm{\Gamma}{A}{x}$ be a term. Then the septuple
+\begin{equation*}
+\subst{x}{\cftalg{Q}}
+  \defeq
+  ( \subst{x}{\cftalgc{\cftalg{Q}}},
+    \subst{x}{\cftalgf{\cftalg{Q}}},
+    \subst{x}{\cftalgt{\cftalg{Q}}},
+    \subst{x}{\cftctxext[\cftalg{Q}]},
+    \subst{x}{\cftfamext[\cftalg{Q}]},
+    \subst{x}{\cftempc{\cftalg{Q}}},
+    \subst{x}{\cftempf{\cftalg{Q}}})
+\end{equation*}
+is a CFT-algebra in context $\ctxext{\Gamma}{\subst{x}{P}}$.
+\end{thm}
+
+\begin{proof}
+The proof follows from the fact that substitution with $x$ is compatible with
+all the involved operations.
+\end{proof}
+
+\begin{cor}
+Let $\cftalg{Q}$ be a CFT-algebra in context $\ctxext{\Gamma}{B}$
+and let $\jhom{\Gamma}{A}{B}{f}$. Then the septuple
+\begin{equation*}
+\jcomp{A}{f}{\cftalg{Q}}
+  \defeq
+  ( \jcomp{A}{f}{\cftalgc{\cftalg{Q}}},
+    \jcomp{A}{f}{\cftalgf{\cftalg{Q}}},
+    \jcomp{A}{f}{\cftalgt{\cftalg{Q}}},
+    \jcomp{A}{f}{\cftctxext[\cftalg{Q}]},
+    \jcomp{A}{f}{\cftfamext[\cftalg{Q}]},
+    \jcomp{A}{f}{\cftempc{\cftalg{Q}}},
+    \jcomp{A}{f}{\cftempf{\cftalg{Q}}})
+\end{equation*}
+is a CFT-algebra in context $\ctxext{\Gamma}{A}$.
+\end{cor}
+
+\begin{cor}
+Let $\cftalg{A}$ be a CFT-algebra in context $\Gamma$. Then
+$\cftalg{F_{F_A}}$ and $\jcomp{}{\cftctxext[\cftalg{A}]}{\cftalg{F_A}}$ are
+judgmentally equal CFT-algebras in context $\ctxext{{\Gamma}{\cftalgc{\cftalg{A}}}}{\cftalgf{\cftalg{A}}}$.
+\end{cor}
+
+\begin{lem}
+Let $\cftalg{Q}$ be a CFT-algebra in context $\ctxext{\Gamma}{B}$ and let
+$\jfam{\Gamma}{A}$. Then we have the judgmental equality
+\begin{equation*}
+\ctxwk{A}{\cftalg{F_Q}}\jdeq\cftalg{F}_{\ctxwk{A}{\cftalg{Q}}}
+\end{equation*}
+of CFT-algebras in context $\ctxext{{\Gamma}{A}}{\ctxwk{A}{B}}$.
+\end{lem}
+
+\begin{lem}
+Let $\cftalg{Q}$ be a CFT-algebra in context $\ctxext{{\Gamma}{A}}{P}$ 
+and let $\jterm{\Gamma}{A}{x}$. Then we have the judgmental equality
+\begin{equation*}
+\subst{x}{\cftalg{F_Q}}\jdeq\cftalg{F}_{\subst{x}{\cftalg{Q}}}
+\end{equation*}
+of CFT-algebras in context $\ctxext{\Gamma}{\subst{x}{P}}$.
+\end{lem}
 

### 2014-07-12 article-models-ttinternal.tex
victorypdflatex article-models-main.tex

@@
 \begin{equation*}
-\cftalg{F_A}
+\cftfamalg{\cftalg{A}}
 \defeq
@@
 only to refer to $\cftctxext[\cftalg{A}]$ and $\cftfamext[\cftalg{A}]$, respectively,
-and not to $\cftctxext[\cftalg{F_A}]$ or $\cftfamext[\cftalg{F_A}]$.
+and not to $\cftctxext[\cftfamalg{\cftalg{A}}]$ or 
+$\cftfamext[\cftfamalg{\cftalg{A}}]$.
 
@@
 Let $\extalg{A}$ be an extension algebra in context $\Gamma$. Then
-$\extalg{F_{F_A}}$ and $\jcomp{}{\cftctxext[\extalg{A}]}{\extalg{F_A}}$ are
+$\extfamalg{\extfamalg{\extalg{A}}}$ and $\jcomp{}{\cftctxext[\extalg{A}]}{\extfamalg{\extalg{A}}}$ are
 judgmentally equal extension algebras in context $\ctxext{{\Gamma}{\cftalgc{\cftalg{A}}}}{\cftalgf{\cftalg{A}}}$.
@@
 \begin{equation*}
-\ctxwk{A}{\extalg{F_Q}}\jdeq\extalg{F}_{\ctxwk{A}{\extalg{Q}}}
+\ctxwk{A}{\extfamalg{\extalg{Q}}}\jdeq\extfamalg{\ctxwk{A}{\extalg{Q}}}
 \end{equation*}
@@
 \begin{equation*}
-\subst{x}{\extalg{F_Q}}\jdeq\extalg{F}_{\subst{x}{\extalg{Q}}}
+\subst{x}{\extfamalg{\extalg{Q}}}\jdeq\extfamalg{\subst{x}{\extalg{Q}}}
 \end{equation*}
@@
 \begin{equation*}
-\boldsymbol{\mathcal{F}}_\cfthom{f}
+\cftfamhom{\cfthom{f}}
   \defeq
@@
 \end{equation*}
-is an extension homomorphism from $\cftalg{F_A}$ to $\jcomp{}{\cfthomc{\cfthom{f}}}{\cftalg{F_B}}$
+is an extension homomorphism from $\cftfamalg{\cftalg{A}}$ to 
+$\jcomp{}{\cfthomc{\cfthom{f}}}{\cftfamalg{\cftalg{B}}}$
 in context $\ctxext{\Gamma}{\cftalgc{\cftalg{A}}}$. 
@@
 We have to verify that the diagrams in \autoref{eq:exthom1,eq:exthom2}
-commute for $\cfthom{F_f}$. Unfolding the ingredients of \autoref{eq:exthom1}
-for $\cfthom{F_f}$ gives us quite directly \autoref{eq:exthom2}. The
-diagram in \autoref{eq:exthom2} for $\cfthom{F_f}$ is judgmentally equal
+commute for $\cftfamhom{\cfthom{f}}$. Unfolding the ingredients of \autoref{eq:exthom1}
+for $\cftfamhom{\cfthom{f}}$ gives us quite directly \autoref{eq:exthom2}. The
+diagram in \autoref{eq:exthom2} for $\cftfamhom{\cfthom{f}}$ is judgmentally equal
 to the pullback of everything in the diagram in \autoref{eq:exthom2} for
@@
 \end{proof}
+
+\begin{lem}
+Let $\cfthom{f}$ be an extension homomorphism from $\cftalg{A}$ to $\cftalg{B}$
+in context $\Gamma$. Then we have the judgmental equality
+\begin{equation*}
+\cftfamhom{\cftfamhom{\cfthom{f}}}
+  \jdeq
+  \jcomp{}{\cftctxext[\cftalg{A}]}{\cftfamhom{\cfthom{f}}}
+\end{equation*}
+of extension homomorphisms from $\cftfamalg{\cftfamalg{\cftalg{A}}}$ to
+$\jcomp{}{\cfthomf{\cfthom{f}}}
+  {\jcomp{}{\cfthomc{\cfthom{f}}}{\cftfamalg{\cftfamalg{\cftalg{B}}}}}$ in
+context $\ctxext{{\Gamma}{A}}{\cftalgf{\cftalg{A}}}$.
+\end{lem}
 
@@
 \end{proof}
+
+\begin{lem}
+Let $\exthom{g}$ be an extension homomorphism from $\extalg{Q}$ to $\extalg{R}$ in
+context $\ctxext{\Gamma}{B}$ and let $\jfam{\Gamma}{A}$. Then the triple
+\begin{equation*}
+\ctxwk{A}{\cfthom{g}}
+  \defeq
+  ( \ctxwk{A}{\cfthomc{\cfthom{g}}},
+    \ctxwk{A}{\cfthomf{\cfthom{g}}},
+    \ctxwk{A}{\cfthomt{\cfthom{g}}})
+\end{equation*}
+is an extension homomorphism from $\ctxwk{A}{\cftalg{Q}}$ to $\ctxwk{A}{\cftalg{R}}$
+in context $\ctxext{{\Gamma}{A}}{\ctxwk{A}{B}}$.
+\end{lem}
+
+\begin{lem}
+Let $\cfthom{g}$ be an extension homomorphism from $\cftalg{Q}$ to $\cftalg{R}$ in
+context $\ctxext{{\Gamma}{A}}{P}$ and let $\jterm{\Gamma}{A}{x}$. Then the
+triple
+\begin{equation*}
+\subst{x}{\cfthom{g}}
+  \defeq
+  ( \subst{x}{\cfthomc{\cfthom{g}}},
+    \subst{x}{\cfthomf{\cfthom{g}}},
+    \subst{x}{\cfthomt{\cfthom{g}}})
+\end{equation*}
+is an extension homomorphism from $\subst{x}{\cftalg{Q}}$ to $\subst{x}{\cftalg{R}}$
+in context $\ctxext{\Gamma}{\subst{x}{P}}$.
+\end{lem}
 
@@
 Let $\cfthom{f}$ be an extension homomorphism from $\extalg{A}$ to $\extalg{B}$
-in context $\Gamma$. Then $\boldsymbol{\mathcal{F}}_{\cfthom{f}}$ is an
-extension homomorphism from $\extalg{F_A}$ to $\extalg{F_B}$ over
+in context $\Gamma$. Then $\cftfamhom{\cfthom{f}}$ is an
+extension homomorphism from $\extfamalg{\extalg{A}}$ to $\extfamalg{\extalg{B}}$ over
 $\cfthomc{\cfthom{f}}$ in context $\Gamma$.
@@
 \begin{equation*}
-\cftalg{F_A}
+\cftfamalg{\cftalg{A}}
   \defeq
@@
 Let $\cftalg{A}$ be a CFT-algebra in context $\Gamma$. Then
-$\cftalg{F_{F_A}}$ and $\jcomp{}{\cftctxext[\cftalg{A}]}{\cftalg{F_A}}$ are
-judgmentally equal CFT-algebras in context $\ctxext{{\Gamma}{\cftalgc{\cftalg{A}}}}{\cftalgf{\cftalg{A}}}$.
+$\cftfamalg{\cftfamalg{\cftalg{A}}}$ and 
+$\jcomp{}{\cftctxext[\cftalg{A}]}{\cftfamalg{\cftalg{A}}}$ are
+judgmentally equal CFT-algebras in context 
+$\ctxext{{\Gamma}{\cftalgc{\cftalg{A}}}}{\cftalgf{\cftalg{A}}}$.
 \end{cor}
@@
 \begin{equation*}
-\ctxwk{A}{\cftalg{F_Q}}\jdeq\cftalg{F}_{\ctxwk{A}{\cftalg{Q}}}
+\ctxwk{A}{\cftfamalg{\cftalg{Q}}}\jdeq\cftalg{F}_{\ctxwk{A}{\cftalg{Q}}}
 \end{equation*}
@@
 \begin{equation*}
-\subst{x}{\cftalg{F_Q}}\jdeq\cftalg{F}_{\subst{x}{\cftalg{Q}}}
+\subst{x}{\cftfamalg{\cftalg{Q}}}\jdeq\cftalg{F}_{\subst{x}{\cftalg{Q}}}
 \end{equation*}
@@
 \begin{defn}
-A homomorphism of term algebras $\cfthom{f}$ from $\cftalg{A}$ to $\cftalg{B}$
-consists of a triple $(\cfthomc{\cfthom{f}},\cfthomf{\cfthom{f}},\cfthomt{\cfthom{f}})$
-such that $(\cfthomc{\cfthom{f}},\cfthomf{\cfthom{f}})$ is an extension homomorphism
+A CFT-homomorphism $\cfthom{f}$ from $\cftalg{A}$ to $\cftalg{B}$ in context
+$\Gamma$ is an extension homomorphism from $\unfold{\extalg{A}}$ to
+$\unfold{\extalg{B}}$ in context $\Gamma$ for which we have the judgmental
+equalities
+\begin{align*}
+\subst{\cftempc{\cftalg{A}}}{\cfthomc{\cfthom{f}}}
+& \jdeq
+  \cftempc{\cftalg{B}}
+  \\
+\subst{\cftempc{\cftalg{A}}}{\cfthomf{\cfthom{f}}}
+& \jdeq
+  \cfthomc{\cfthom{f}}
+  \\
+\subst{\cftempf{\cftalg{A}}}{\cfthomf{\cfthom{f}}}
+& \jdeq
+  \jcomp{}{\cfthomc{\cfthom{f}}}{\cftempf{\cftalg{B}}}
+\end{align*}
+
+Judgmental equality of CFT-homomorphisms from $\cftalg{A}$ to $\cftalg{B}$ in
+context $\Gamma$ is taken to be judgmental equality of extension homomorphisms
+from $\extalg{A}$ to $\extalg{B}$ in context $\Gamma$.
 \end{defn}
+
+\begin{lem}
+Suppose that $\cfthom{f}$ is a CFT-homomorphism from $\cftalg{A}$ to $\cftalg{B}$
+in context $\Gamma$ and that $\cfthom{g}$ is a CFT-homomorphism from $\cftalg{B}$
+to $\cftalg{C}$ in context $\Gamma$. Then $\jcomp{}{\cfthom{f}}{\cfthom{g}}$
+is a CFT-homomorphism from $\cftalg{A}$ to $\cftalg{C}$ in context $\Gamma$.
+\end{lem}
+
+\begin{lem}
+Let $\cfthom{f}$ be a CFT-homomorphism from $\cftalg{A}$ to $\cftalg{B}$ in
+context $\Gamma$. Then $\cftfamhom{\exthom{f}}$ is a CFT-homomorphism from
+$\cftfamalg{\cftalg{A}}$ to $\jcomp{}{\cfthomc{\cfthom{f}}}{\cftfamalg{\cftalg{B}}}$
+in context $\ctxext{\Gamma}{\cftalgc{\cftalg{A}}}$. 
+\end{lem}
+
+\begin{lem}
+Let $\cfthom{g}$ be a CFT-homomorphism from $\cftalg{Q}$ to $\cftalg{R}$ in
+context $\ctxext{\Gamma}{B}$ and let $\jfam{\Gamma}{A}$. Then 
+$\ctxwk{A}{\cfthom{g}}$
+is a CFT-homomorphism from $\ctxwk{A}{\cftalg{Q}}$ to $\ctxwk{A}{\cftalg{R}}$
+in context $\ctxext{{\Gamma}{A}}{\ctxwk{A}{B}}$.
+\end{lem}
+
+\begin{lem}
+Let $\cfthom{g}$ be a CFT-homomorphism from $\cftalg{Q}$ to $\cftalg{R}$ in
+context $\ctxext{{\Gamma}{A}}{P}$ and let $\jterm{\Gamma}{A}{x}$. Then 
+$\subst{x}{\cfthom{g}}$
+is a CFT-homomorphism from $\subst{x}{\cftalg{Q}}$ to $\subst{x}{\cftalg{R}}$
+in context $\ctxext{\Gamma}{\subst{x}{P}}$.
+\end{lem}
 
@@
   {{{\Gamma}{\cftalgc{\cftalg{A}}}}{\cftalgf{\cftalg{A}}}}
-  {\ctxwk{\cftalgf{\cftalg{A}}}{\cftalg{F_A}}}
-  {\cftalg{F_{F_A}}}
+  {\ctxwk{\cftalgf{\cftalg{A}}}{\cftfamalg{\cftalg{A}}}}
+  {\cftfamalg{\cftfamalg{\cftalg{A}}}}
   {\cftwk{\cftalg{A}}}
@@
 In particular, we see that $\cftwk{\cftalg{A}}$ is a CFT-homomorphism from
-$\ctxwk{\cftalgf{\cftalg{A}}}{\cftalg{F_A}}$ to $\cftalg{F_A}$ over
+$\ctxwk{\cftalgf{\cftalg{A}}}{\cftfamalg{\cftalg{A}}}$ to $\cftfamalg{\cftalg{A}}$ over
 $\cftctxext$ in context $\Gamma$.
@@
 Let $\cftalg{A}$ be a CFT-algebra and let $\cftwk{\cftalg{A}}$ be as in
-\autoref{cftwk}. Then we have
+\autoref{cftwk}. Then $\apply\cftwk{\cftalg{A}}{\gamma}{a}$ is a CFT-homomorphism
+from $\subst{\gamma}{\cftfamalg{\cftalg{A}}}$ to 
+$\subst{a}{{\gamma}{\cftfamalg{\cftfamalg{\cftalg{A}}}}}$. Furthermore, we have
 \begin{equation*}
@@
 \end{equation*}
-for $\gamma:A$ and $a,b:\subst{\gamma}{\cftalgf{\cftalg{A}}}$. We have
+for $\gamma:A$ and $a,b:\subst{\gamma}{\cftalgf{\cftalg{A}}}$; we have
 \begin{equation*}
@@
 for $\gamma:A$, $a,b:\subst{\gamma}{\cftalgf{\cftalg{A}}}$ and
-$q:\subst{\apply\cftctxext{\gamma}{b}}{\cftalgf{\cftalg{A}}}$. Finally, we have
+$q:\subst{\apply\cftctxext{\gamma}{b}}{\cftalgf{\cftalg{A}}}$; and finally, we have
 \begin{equation*}
@@
 
+\subsubsection{Currying for weakening}
+Let $\cftalg{A}$ be a CFT-algebra in context $\Gamma$ and let
+\begin{equation*}
+\jhom
+  {{{\Gamma}{\cftalgc{\cftalg{A}}}}{\cftalgf{\cftalg{A}}}}
+  {\ctxwk{\cftalgf{\cftalg{A}}}{\cftfamalg{\cftalg{A}}}}
+  {\cftfamalg{\cftfamalg{\cftalg{A}}}}
+  {\cftwk{\cftalg{A}}}
+\end{equation*}
+be a CFT-homomorphism. The abstraction of the rules in \autoref{comp-ew} is the
+condition that the diagram
+\begin{equation}\label{eq:cftwk-cftextcurry}
+\begin{tikzcd}[column sep=large]
+\ctxwk
+  {\ctxext{\cftalgf{\cftalg{A}}}{\cftalgf{\cftfamalg{\cftalg{A}}}}}
+  {\cftfamalg{\cftalg{A}}}
+  \ar{r}{\ctxwk{\cftalgf{\cftfamalg{\cftalg{A}}}}{\cftwk{\cftalg{A}}}}
+  \ar{dr}[swap]{\jcomp{}{\cftfamext}{\cftwk{\cftalg{A}}}}
+& \ctxwk{\cftalgf{\cftfamalg{\cftalg{A}}}}{\cftfamalg{\cftfamalg{\cftalg{A}}}}
+  \ar{d}{\jcomp{}{\cftctxext}{\cftwk{\cftalg{A}}}}
+  \\
+& \cftfamalg{\cftfamalg{\cftfamalg{\cftalg{A}}}}
+\end{tikzcd}
+\end{equation}
+of CFT-algebras in context 
+$\ctxext
+    {{{\Gamma}{\cftalgc{\cftalg{A}}}}{\cftalgf{\cftalg{A}}}}
+    {\cftalgf{\cftfamalg{\cftalg{A}}}}
+$ %
+commutes.
+
+\begin{lem}
+Let $\cftalg{A}$ be a CFT-algebra in context gamma with a CFT-homomorphism 
+$ \jhom
+  {{{\Gamma}{\cftalgc{\cftalg{A}}}}{\cftalgf{\cftalg{A}}}}
+  {\ctxwk{\cftalgf{\cftalg{A}}}{\cftfamalg{\cftalg{A}}}}
+  {\jcomp{}{\cftctxext}{\cftfamalg{\cftalg{A}}}}
+  {\cftwk{\cftalg{A}}}
+$
+satisfying the judgmental equality of \autoref{eq:cftwk-cftextcurry}. Then we
+have the judgmental equality
+\begin{equation*}
+\apply\cftwk{\cftalg{A}}{\gamma}{\apply\cftfamext{\gamma}{a}{p}}
+\jdeq
+\jcomp{}
+    {\apply\cftwk{\cftalg{A}}{\gamma}{a}}
+    {\apply\cftwk{\cftalg{A}}{\apply\cftctxext{\gamma}{a}}{p}}
+\end{equation*}
+of CFT-homomorphisms from $\subst{\gamma}{\cftfamalg{\cftalg{A}}}$ to 
+$\subst{p}{{a}{{\gamma}{\cftalg{F_{F_{F_A}}}}}}$ in context $\Gamma$,
+for every $\gamma:A$, $a:\subst{\gamma}{\cftalgf{\cftalg{A}}}$ and
+$p:\subst{a}{{\gamma}{\cftalgf{\cftfamalg{\cftalg{A}}}}}$. 
+\end{lem}
+
+\subsubsection{Weakening by the empty family}
+\begin{equation}\label{eq:cftwk-cftempcurry}
+\jhomeq
+  { {\Gamma}{\cftalgc{\cftalg{A}}}
+    }
+  { \cftfamalg{\cftalg{A}}
+    }
+  { \cftfamalg{\cftalg{A}}
+    }
+  { \subst{\cftempf{\cftalg{A}}}{\cftwk{\cftalg{A}}}
+    }
+  { id_{\cftfamalg{\cftalg{A}}}
+    }
+\end{equation}
+
 \subsubsection{Weakening preserves itself}
@@
   {{{\Gamma}{\cftalgc{\cftalg{A}}}}{\cftalgf{\cftalg{A}}}}
-  {\ctxwk{\cftalgf{\cftalg{A}}}{\cftalg{F_A}}}
-  {\jcomp{}{\cftctxext}{\cftalg{F_A}}}
+  {\ctxwk{\cftalgf{\cftalg{A}}}{\cftfamalg{\cftalg{A}}}}
+  {\cftfamalg{\cftfamalg{\cftalg{A}}}}
   {\cftwk{\cftalg{A}}}
@@
 \begin{tikzcd}[column sep=10em]
-\ctxwk{\cftalgf{\cftalg{A}}}{{\cftalgf{\cftalg{F_A}}}{\cftalg{F_{F_A}}}}
+\ctxwk{\cftalgf{\cftalg{A}}}{{\cftalgf{\cftfamalg{\cftalg{A}}}}{\cftfamalg{\cftfamalg{\cftalg{A}}}}}
   \ar{d}
     [swap]{ \ctxwk
-        {{\cftalgf{\cftalg{A}}}{\cftalgf{\cftalg{F_A}}}}
-        {\boldsymbol{\mathcal{F}}_{\cftwk{\cftalg{A}}}}
+        {{\cftalgf{\cftalg{A}}}{\cftalgf{\cftfamalg{\cftalg{A}}}}}
+        {\cftfamhom{\cftwk{\cftalg{A}}}}
       }
   \ar{r}{\ctxwk{\cftalgf{\cftalg{A}}}{\jcomp{}{\cftctxext}{\cftwk{\cftalg{A}}}}}
-& \ctxwk{\cftalgf{\cftalg{A}}}{\cftalg{F_{F_{F_A}}}}
-  \ar{d}{\jcomp{}{\ctxwk{\cftalgf{\cftalg{A}}}{\cftfamext}}{\boldsymbol{\mathcal{F}}_{\cftwk{\cftalg{A}}}}}
-  \\
-\ctxwk{{\cftalgf{\cftalg{A}}}{\cftalgf{\cftalg{F_A}}}}{\jcomp{}{\cftwkc{\cftalg{A}}}{\cftalg{F_{F_{F_A}}}}}
+& \ctxwk{\cftalgf{\cftalg{A}}}{\cftfamalg{\cftfamalg{\cftfamalg{\cftalg{A}}}}}
+  \ar{d}{\jcomp{}{\ctxwk{\cftalgf{\cftalg{A}}}{\cftfamext}}{\cftfamhom{\cftwk{\cftalg{A}}}}}
+  \\
+\ctxwk{{\cftalgf{\cftalg{A}}}{\cftalgf{\cftfamalg{\cftalg{A}}}}}{\jcomp{}{\cftwkc{\cftalg{A}}}{\cftfamalg{\cftfamalg{\cftfamalg{\cftalg{A}}}}}}
   \ar{r}[swap]{%
@@
     {\ctxwk{\cftalgf{\cftalg{A}}}{\cftfamext}}
-    {\jcomp{}{\cftwkc{\cftalg{A}}}{\cftalg{F_{F_{F_A}}}}}
+    {\jcomp{}{\cftwkc{\cftalg{A}}}{\cftfamalg{\cftfamalg{\cftfamalg{\cftalg{A}}}}}}
 \end{tikzcd}
@@
         {\cftalgf{\cftalg{A}}}
-        {\cftalgf{\cftalg{F_A}}}
+        {\cftalgf{\cftfamalg{\cftalg{A}}}}
       }$ %
-commutes. In this diagram, it is not immediately obvious that 
+commutes. In this diagram, it is not immediately obvious that\ldots
 \begin{enumerate}
-\item the domain of the CFT-homomorphism
-$\jcomp{}{\ctxwk{\cftalgf{\cftalg{A}}}{\cftfamext}}{\boldsymbol{\mathcal{F}}_{\cftwk{\cftalg{A}}}}$
+\item \ldots the domain of the CFT-homomorphism
+$\jcomp{}{\ctxwk{\cftalgf{\cftalg{A}}}{\cftfamext}}{\cftfamhom{\cftwk{\cftalg{A}}}}$
 is the indicated domain. Note that we have the judgmental equalities
@@
   {\ctxwk{\cftalgf{\cftalg{A}}}{\cftfamext}}
-  {\ctxwk{\cftalgf{\cftalg{A}}}{\cftalgf{\cftalg{F_A}}}}
+  {\ctxwk{\cftalgf{\cftalg{A}}}{\cftalgf{\cftfamalg{\cftalg{A}}}}}
 & \jdeq
@@
   {\cftalgf{\cftalg{A}}}
-  {\jcomp{}{\cftfamext}{\cftalg{F_{F_A}}}}
+  {\jcomp{}{\cftfamext}{\cftfamalg{\cftfamalg{\cftalg{A}}}}}
   \\
@@
   {\cftalgf{\cftalg{A}}}
-  {\jcomp{}{\jcomp{}{\cftctxext}{\cftfamext}}{\cftalg{F_{F_A}}}}
+  {\jcomp{}{\jcomp{}{\cftctxext}{\cftfamext}}{\cftfamalg{\cftfamalg{\cftalg{A}}}}}
   \\
@@
   {\cftalgf{\cftalg{A}}}
-  {\cftalg{F_{F_{F_A}}}}.
-\end{align*}
-\item the domain and codomain of the CFT-homomorphism
+  {\cftfamalg{\cftfamalg{\cftfamalg{\cftalg{A}}}}}.
+\end{align*}
+\item \ldots the domain and codomain of the CFT-homomorphism
 \begin{equation*}
@@
                   {\cftctxext}
-                  {\ctxwk{\cftalgf{\cftalg{A}}}{\cftalg{F_A}}}
+                  {\ctxwk{\cftalgf{\cftalg{A}}}{\cftfamalg{\cftalg{A}}}}
                 }
@@
                   {\cftctxext}
-                  {\cftalg{F_{F_A}}}
-                }
-            }
-        }
-$ are indeed judgmentally equal to the indicated domain and codomain. To see
-this, we have the following lemma:
+                  {\cftfamalg{\cftfamalg{\cftalg{A}}}}
+                }
+            }
+        }
+$ are indeed judgmentally equal to the indicated domain and codomain.
+
+To see
+that the two domains are equal, note that $\cftwkf{\cftalg{A}}$ may be seen as
+a morphism from $\ctxwk{\cftalgf{\cftalg{A}}}{\cftalgf{\cftfamalg{\cftalg{A}}}}$
+to $\cftalgf{\cftalg{A}}$ over 
+$\jcomp{}{\cftwkc{\cftalg{A}}}{\jcomp{}{\cftctxext}{\cftctxext}}$ in context
+$\Gamma$, i.e. we have the following commuting diagram in context $\Gamma$:
+\begin{equation*}
+\begin{tikzcd}[column sep=huge]
+\ctxwk{\cftalgf{\cftalg{A}}}{\cftalgf{\cftfamalg{\cftalg{A}}}}
+  \ar{r}
+    {\cftwkf{\cftalg{A}}}
+  \ar[fib]{d}
+  &
+\cftalgf{\cftalg{A}}
+  \ar[fib]{d}
+  \\
+\ctxext{ {\cftalgc{\cftalg{A}}}
+         {\cftalgf{\cftalg{A}}}
+         }
+       { \ctxwk{\cftalgf{\cftalg{A}}}{\cftalgf{\cftalg{A}}}
+         }
+  \ar{r}[swap]
+    {\jcomp{}{\cftwkc{\cftalg{A}}}{\jcomp{}{\cftctxext}{\cftctxext}}}
+  &
+\cftalgc{\cftalg{A}}
+\end{tikzcd}
+\end{equation*}
+Using the associativity of composition, we see that we have the judgmental 
+equality
+\begin{equation*}
+\jcomp{}
+  { \cftwkf{\cftalg{A}}}
+  { \jcomp{}
+      {\cftwkc{\cftalg{A}}}
+      { \jcomp{}
+          {\cftctxext}
+          { \jcomp{}
+              {\cftctxext}
+              {\ctxwk{\cftalgf{\cftalg{A}}}{\cftfamalg{\cftalg{A}}}}
+            }
+        }
+    }
+\jdeq
+\jcomp{}
+  { \jvcomp{}
+      { \jcomp{}
+          { \cftwkc{\cftalg{A}}}
+          { \jcomp{}
+              {\cftctxext}
+              {\cftctxext}
+            }
+        }
+      { \cftwkf{\cftalg{A}}
+        }
+    }
+  { \ctxwk{\cftalgf{\cftalg{A}}}{\cftfamalg{\cftalg{A}}}}  
+\end{equation*}
+Recall that weakening is the same as pulling
+back along a projection. Therefore, we see that
+it suffices to show that the diagram
+\begin{equation*}
+\begin{tikzcd}[column sep=10em]
+\ctxext{ { { \cftalgc{\cftalg{A}}}
+           { \cftalgf{\cftalg{A}}}
+           }
+         { \ctxwk{\cftalgf{\cftalg{A}}}{\cftalgf{\cftalg{A}}}}
+         }
+       { \ctxwk{\cftalgf{\cftalg{A}}}{\cftalgf{\cftfamalg{\cftalg{A}}}}
+         }
+  \ar{r}
+    { \jvcomp{}
+        { \jcomp{}
+            {\cftwkc{\cftalg{A}}}
+            { \jcomp{}
+                {\cftctxext}
+                {\cftctxext}
+              }
+          }
+        { \cftwkf{\cftalg{A}}}
+      }
+  \ar{d}[swap]
+    { \cprojfstf
+        { { { \cftalgc{\cftalg{A}}}
+            { \cftalgf{\cftalg{A}}}
+            }
+          { \ctxwk{\cftalgf{\cftalg{A}}}{\cftalgf{\cftalg{A}}}}
+          }
+        { \ctxwk{\cftalgf{\cftalg{A}}}{\cftalgf{\cftfamalg{\cftalg{A}}}}
+          }
+      }
+  &
+\ctxext{\cftalgc{\cftalg{A}}}{\cftalgf{\cftalg{A}}}
+  \ar{d}
+    {\cprojfstf{\cftalgc{\cftalg{A}}}{\cftalgf{\cftalg{A}}}}
+  \\
+\ctxext{ {\cftalgc{\cftalg{A}}}
+         {\cftalgf{\cftalg{A}}}
+         }
+       { \ctxwk{\cftalgf{\cftalg{A}}}{\cftalgf{\cftalg{A}}}
+         }
+  \ar{r}[swap]
+    {\jcomp{}{\cftwkc{\cftalg{A}}}{\jcomp{}{\cftctxext}{\cftctxext}}}
+  &
+\cftalgc{\cftalg{A}}
+\end{tikzcd}
+\end{equation*}
+commutes judgmentally. This follows from \autoref{lem:jvcomp-cprojfstf}.
+
+To see that the codomains are equal, use the fact that $\cftwk{\cftalg{A}}$ is
+an extension homomorphism.
 \end{enumerate}
-
-\begin{lem}
-The following three inference rules are valid
-\end{lem}
 
@@
 be a homomorphism of CFT-algebras for which the diagram in
-\autoref{diag:cftwk-cftwk} commutes judgmentally. Then
-we have the judgmental equality
-\begin{align*}
-& \jcomp{}
-  { \subst{q}{{\apply\cftctxext{\gamma}{b}}{\cftwk{\cftalg{A}}}}
-    }
+\autoref{diag:cftwk-cftwk} commutes judgmentally. 
+Then the diagram
+\begin{equation*}
+\begin{tikzcd}[column sep=large]
+\subst
+  {\apply\cftctxext{\gamma}{b}}
+  {\cftfamalg{\cftalg{A}}}
+  \ar{r}{\apply\cftwk{\cftalg{A}}{\apply\cftctxext{\gamma}{b}}{q}}
+  \ar{d}[swap]{\subst{b}{{a}{{\gamma}{\cftfamhom{\cftwk{\cftalg{A}}}}}}}
+& \subst
+    { q}
+    { {\apply\cftctxext{\gamma}{b}}
+      {\cftfamalg{\cftfamalg{\cftalg{A}}}}
+      }
+  \ar{d}
     { \subst
         {\apply\cftfamext{\gamma}{b}{q}}
-      {{a}{{\gamma}{\boldsymbol{\mathcal{F}}_{\cftwk{\cftalg{A}}}}}}
-    }
-  \\
-& \jdeq
-  \jcomp{}
-    { \subst{b}{{a}{{\gamma}{\boldsymbol{\mathcal{F}}_{\cftwk{\cftalg{A}}}}}}
-      }
+        {{a}{{\gamma}{\cftfamhom{\cftwk{\cftalg{A}}}}}}
+      }
+  \\
+\subst
+  { \apply\cftwkc{\cftalg{A}}{\gamma}{a}{b}
+    }
+  { { \apply\cftctxext{\gamma}{a}}
+    { \cftfamalg{\cftfamalg{\cftalg{A}}}}
+    }
+  \ar{r}[swap,yshift=-1em]
     { \subst
@@
       }
-\end{align*}
-of CFT-homomorphisms from $\subst{b}{{\gamma}{\cftalg{F_{F_A}}}}$ to
-$ \subst
+& \subst
     { \apply\cftwkf{\cftalg{A}}{\gamma}{a}{b}{q}
@@
     { {\apply\cftwkc{\cftalg{A}}{\gamma}{a}{b}}
-        {{a}{{\gamma}{\cftalg{F_{F_{F_{F_{A}}}}}}}}
-        }$
-in context $\Gamma$, for $\gamma:A$, $a,b:\subst{\gamma}{\cftalgf{\cftalg{A}}}$ and
-$q:\subst{b}{{\gamma}{\cftalgf{\cftalg{F_A}}}}$.
+      {{a}{{\gamma}{\cftfamalg{\cftfamalg{\cftfamalg{\cftfamalg{\cftalg{A}}}}}}}}
+      }
+\end{tikzcd}
+\end{equation*}
+of CFT-algebras in context $\Gamma$ commutes judgmentally for
+$\gamma:A$, $a,b:\subst{\gamma}{\cftalgf{\cftalg{A}}}$ and
+$q:\subst{b}{{\gamma}{\cftalgf{\cftfamalg{\cftalg{A}}}}}$.
 \end{lem}
 
-\subsubsection{Currying for weakening}
-Let $\cftalg{A}$ be a CFT-algebra in context $\Gamma$ and let
-\begin{equation*}
-\jhom
-  {{{\Gamma}{\cftalgc{\cftalg{A}}}}{\cftalgf{\cftalg{A}}}}
-  {\ctxwk{\cftalgf{\cftalg{A}}}{\cftalg{F_A}}}
-  {\jcomp{}{\cftctxext}{\cftalg{F_A}}}
-  {\cftwk{\cftalg{A}}}
-\end{equation*}
-be a CFT-homomorphism. The abstraction of the rules in \autoref{comp-ew} is the
-condition that the diagram
-\begin{equation}\label{eq:cftwk-cftextcurry}
-\begin{tikzcd}[column sep=large]
-\ctxwk
-  {\ctxext{\cftalgf{\cftalg{A}}}{\cftalgf{\cftalg{F_A}}}}
-  {\cftalg{F_A}}
-  \ar{r}{\ctxwk{\cftalgf{\cftalg{F_A}}}{\cftwk{\cftalg{A}}}}
-  \ar{dr}[swap]{\jcomp{}{\cftfamext}{\cftwk{\cftalg{A}}}}
-& \ctxwk{\cftalgf{\cftalg{F_A}}}{\cftalg{F_{F_A}}}
-  \ar{d}{\jcomp{}{\cftctxext}{\cftwk{\cftalg{A}}}}
-  \\
-& \cftalg{F_{F_{F_A}}}
-\end{tikzcd}
-\end{equation}
-of CFT-algebras in context 
-$\ctxext
-    {{{\Gamma}{\cftalgc{\cftalg{A}}}}{\cftalgf{\cftalg{A}}}}
-    {\cftalgf{\cftalg{F_A}}}
-$ %
-commutes.
-
-\begin{lem}
-Let $\cftalg{A}$ be a CFT-algebra in context gamma with a CFT-homomorphism 
-$ \jhom
-  {{{\Gamma}{\cftalgc{\cftalg{A}}}}{\cftalgf{\cftalg{A}}}}
-  {\ctxwk{\cftalgf{\cftalg{A}}}{\cftalg{F_A}}}
-  {\jcomp{}{\cftctxext}{\cftalg{F_A}}}
-  {\cftwk{\cftalg{A}}}
-$
-satisfying the judgmental equality of \autoref{eq:cftwk-cftextcurry}. Then we
-have the judgmental equality
-\begin{equation*}
-\apply\cftwk{\cftalg{A}}{\gamma}{\apply\cftfamext{\gamma}{a}{p}}
-\jdeq
-\jcomp{}
-    {\apply\cftwk{\cftalg{A}}{\gamma}{a}}
-    {\apply\cftwk{\cftalg{A}}{\apply\cftctxext{\gamma}{a}}{p}}
-\end{equation*}
-of CFT-homomorphisms from $\subst{\gamma}{\cftalg{F_A}}$ to 
-$\subst{p}{{a}{{\gamma}{\cftalg{F_{F_{F_A}}}}}}$ in context $\Gamma$,
-for every $\gamma:A$, $a:\subst{\gamma}{\cftalgf{\cftalg{A}}}$ and
-$p:\subst{a}{{\gamma}{\cftalgf{\cftalg{F_A}}}}$. 
-\end{lem}
-
-\subsubsection{Weakening by the empty family}
-\begin{equation}\label{eq:cftwk-cftempcurry}
-\jhomeq
-  { {\Gamma}{\cftalgc{\cftalg{A}}}
-    }
-  { \cftalg{F_A}
-    }
-  { \cftalg{F_A}
-    }
-  { \subst{\cftempf{\cftalg{A}}}{\cftwk{\cftalg{A}}}
-    }
-  { id_{\cftalg{F_A}}
-    }
-\end{equation}
+\begin{proof}
+All one has to do to prove this is to substitute all of the ingredients of
+the diagram in \autoref{diag:cftwk-cftwk} consecutively by
+$\gamma$, by $a$, by $b$ and finally by $q$.
+\end{proof}
 
@@
   {{{\Gamma}{\cftalgc{\cftalg{A}}}}{\cftalgf{\cftalg{A}}}}
-  {\ctxwk{\cftalgf{\cftalg{A}}}{\cftalg{F_A}}}
+  {\ctxwk{\cftalgf{\cftalg{A}}}{\cftfamalg{\cftalg{A}}}}
   {\cftalg{F_{F_A}}}
@@
 \subsubsection{Derivable properties of weakening algebras}
-We should theoremize the following statements:
-\begin{enumerate}
-\item If $\cftalg{A}$ is a weakening algebra, then so is $\cftalg{F_A}$ if we take
-$\jcomp{}{\cftctxext}{\cftwk{\cftalg{A}}}$ as its weakening.
-\end{enumerate}
+
+\begin{thm}
+Suppose $\cftalg{A}$ is a weakening algebra in context $\Gamma$. Then
+$\cftfamalg{\cftalg{A}}$ together with the CFT-homomorphism
+\begin{equation*}
+\cftwk{\cftfamalg{\cftalg{A}}}
+  \defeq \jcomp{}{\cftctxext}{\cftwk{\cftalg{A}}}
+\end{equation*}
+forms a weakening algebra in context $\ctxext{\Gamma}{\cftalgc{\cftalg{A}}}$.
+\end{thm}
+
+\begin{thm}
+Suppose $\cftalg{Q}$ is a weakening algebra in context $\ctxext{\Gamma}{B}$ and
+let $\jfam{\Gamma}{A}$. Then $\ctxwk{A}{\cftalg{Q}}$ together with the
+CFT-homomorphism
+\begin{equation*}
+\cftwk{\ctxwk{A}{\cftalg{Q}}}\defeq \ctxwk{A}\cftwk{\cftalg{Q}}
+\end{equation*}
+forms a weakening algebra.
+\end{thm}
+
+\begin{thm}
+Suppose $\cftalg{Q}$ is a weakening algebra in context $\ctxext{{\Gamma}{A}}{P}$
+and let $\jterm{\Gamma}{A}{x}$. Then $\subst{x}{\cftalg{Q}}$ together with the
+CFT-homomorphism
+\begin{equation*}
+\cftwk{\subst{x}{\cftalg{Q}}} \defeq \subst{x}{\cftwk{\cftalg{Q}}}
+\end{equation*}
+forms a weakening algebra.
+\end{thm}
 
@@
     }
-  {\subst{{\idtm{\cftalgf{\cftalg{A}}}}{\cftwkc{\cftalg{A}}}}{\cftalgt{\cftalg{F_A}}}}
+  {\subst{{\idtm{\cftalgf{\cftalg{A}}}}{\cftwkc{\cftalg{A}}}}{\cftalgt{\cftfamalg{\cftalg{A}}}}}
   {\cftidtm{\cftalg{A}}}
@@
   {\ctxwk{\cftalgt{\cftalg{A}}}{\cftalg{F_{F_A}}}}
-  {\cftalg{F_A}}
+  {\cftfamalg{\cftalg{A}}}
   {\cftsubst{\cftalg{A}}}
@@
         { { p}
-          { \boldsymbol{\mathcal{F}}_{\apply\cftsubst{\cftalg{A}}{\gamma}{a}{x}}}
+          { \cftfamhom{\apply\cftsubst{\cftalg{A}}{\gamma}{a}{x}}}
           }

### 2014-07-12 article-models-ttinternal.tex
small changes

@@
 \end{equation*}
-We will require additionally that identity terms are compatible with weakening.
-Thus we will formulate an abstraction of the rule stated in \autoref{comp-wi}.
-
-\begin{equation}
-\jtermeq
-  { \xi}
-  { \xi}
-  { \subst
+We will require additionally that identity terms are compatible with weakening,
+i.e.~an abstraction of the rule stated in \autoref{comp-wi}. Since this is the
+only additional rule for projection algebras, we immediately state the 
+definition.
+
+\begin{defn}
+A \emph{projection algebra $\cftalg{A}$ in context $\Gamma$} is a weakening algebra
+$\cftalg{A}$ in context $\Gamma$ together with a term
+\begin{equation*}
+\jterm
+  { { {\Gamma}
+      {\cftalgc{\cftalg{A}}}
+      }
+    { \cftalgf{\cftalg{A}}
+      }
+    }
+  {\subst{{\idtm{\cftalgf{\cftalg{A}}}}{\cftwkc{\cftalg{A}}}}{\cftalgt{\cftfamalg{\cftalg{A}}}}}
+  {\cftidtm{\cftalg{A}}}
+\end{equation*}
+satisfying the judgmental equality
+\begin{align*}
+& \subst
+  { \ctxwk
+      { \cftalgf{\cftalg{A}}}
       { \jcomp
@@
         }
-      { { { \idtm
+    }
+  { { \ctxwk
+        { \cftalgf{\cftalg{A}}}
+        { \subst { \idtm
             { \jcomp
@@
           }
+      }
     { \jcomp
         { }
-            { \jcomp{}{\cftctxext}{\cftfamext}}
+        { \ctxwk{\cftalgf{\cftalg{A}}}{\cftfamext}}
         { \cftwkt{\cftalg{A}}}
@@
     }
-    }
-  { \xi}
-\end{equation}
+  \\
+& \jdeq
+  \jcomp{}
+    { \cftwkf{\cftalg{A}}}
+    { \jcomp{}
+        { \cftwkc{\cftalg{A}}}
+        { \jcomp{}
+            { \cftctxext}
+            { \jcomp{}
+              { \cftctxext}
+              { \cftidtm{\cftalg{A}}}
+              }
+          }
+      }
+\end{align*}
+of terms of %
+$ \jcomp{}
+    { \cftwkf{\cftalg{A}}}
+    { \jcomp{}
+        { \cftwkc{\cftalg{A}}}
+        { \jcomp{}
+            { \cftctxext}
+            { \jcomp{}
+              { \cftctxext}
+              { \subst
+                  { {\idtm{\cftalgf{\cftalg{A}}}}
+                    {\cftwkc{\cftalg{A}}}
+                    }
+                  { \cftalgt{\cftfamalg{\cftalg{A}}}
+                    }
+                }
+              }
+          }
+      }$
+in context \ldots.
+
+The term $\cftidtm{\cftalg{A}}$ of a projection algebra $\cftalg{A}$ in context
+$\Gamma$ is called the \emph{internal identity term of $\cftalg{A}$}.
+\end{defn}
+
+\begin{rmk}
+We shall have to verify that the condition on the internal identity term 
+$\cftidtm{\cftalg{A}}$ of a projection algebra $\cftalg{A}$ in context $\Gamma$
+is indeed well-typed. More precisely, we have to demonstrate that the judgmental
+equality
+\begin{align*}
+& \jdeq
+  \jcomp{}
+    { \cftwkf{\cftalg{A}}}
+    { \jcomp{}
+        { \cftwkc{\cftalg{A}}}
+        { \jcomp{}
+            { \cftctxext}
+            { \jcomp{}
+              { \cftctxext}
+              { \subst
+                  { {\idtm{\cftalgf{\cftalg{A}}}}
+                    {\cftwkc{\cftalg{A}}}
+                    }
+                  { \cftalgt{\cftfamalg{\cftalg{A}}}
+                    }
+                }
+              }
+          }
+      }  
+\end{align*}
+of families in context \ldots holds.
+\end{rmk}
 
@@
 other. Thus, we need to require the abstractions of the rules stated in
-\autoref{comp-sw,comp-ws,cancellation-ws,cancellation-i}.
+\autoref{comp-sw,comp-ws,cancellation-ws,comp-si,cancellation-i}.
+
+\subsection{E-homomorphisms}
+\begin{defn}
+An \emph{E-homomorphism from $\cftalg{A}$ to $\cftalg{B}$ in context $\Gamma$}
+is a CFT-homomorphism which is both a projection homomorphism and a
+substitution homomorphism.
+\end{defn}
+
+\subsection{E-transformations}
+E-transformations are going to be the analogues of natural transformations.
+
+\subsection{E-adjunctions}
 

### 2014-07-12 article-models-ttinternal.tex
progress on weakening algebras

@@
 condition that the diagram
-\begin{equation}\label{eq:cftwk-cftextcurry}
+\begin{equation}\label{diag:cftwk-curry}
 \begin{tikzcd}[column sep=large]
@@
 $
-satisfying the judgmental equality of \autoref{eq:cftwk-cftextcurry}. Then we
+satisfying the judgmental equality of \autoref{diag:cftwk-curry}. Then we
 have the judgmental equality
@@
 
+\begin{proof}
+To show that $\cftwk{\cftfamalg{\cftalg{A}}}$ satisfies the Currying rule, we
+must verify that the diagram
+\begin{equation*}
+\begin{tikzcd}[column sep=large]
+\ctxwk
+  {\ctxext{\cftalgf{\cftfamalg{\cftalg{A}}}}{\cftalgf{\cftfamalg{\cftfamalg{\cftalg{A}}}}}}
+  {\cftfamalg{\cftfamalg{\cftalg{A}}}}
+  \ar{r}{\ctxwk{\cftalgf{\cftfamalg{\cftfamalg{\cftalg{A}}}}}{\cftwk{\cftfamalg{\cftalg{A}}}}}
+  \ar{dr}[swap]{\jcomp{}{\jcomp{}{\cftctxext}{\cftfamext}}{\cftwk{\cftfamalg{\cftalg{A}}}}}
+& \ctxwk{\cftalgf{\cftfamalg{\cftfamalg{\cftalg{A}}}}}{\cftfamalg{\cftfamalg{\cftfamalg{\cftalg{A}}}}}
+  \ar{d}{\jcomp{}{\cftfamext}{\cftwk{\cftfamalg{\cftalg{A}}}}}
+  \\
+& \cftfamalg{\cftfamalg{\cftfamalg{\cftfamalg{\cftalg{A}}}}}
+\end{tikzcd}
+\end{equation*}
+commutes judgmentally. We do this by showing that each of the ingredients of this
+diagram is judgmentally equal to the corresponding ingredient of the diagram in
+\autoref{diag:cftwk-curry} pulled back along $\cftctxext$. Thus, we are going
+to show that
+\begin{align*}
+\ctxwk{\cftalgf{\cftfamalg{\cftfamalg{\cftalg{A}}}}}{\cftwk{\cftfamalg{\cftalg{A}}}}
+& \jdeq
+\jcomp{}{\cftctxext}{\ctxwk{\cftalgf{\cftfamalg{\cftalg{A}}}}{\cftwk{\cftalg{A}}}}
+  \tag{1}
+  \\
+\jcomp{}{\cftfamext}{\cftwk{\cftfamalg{\cftalg{A}}}}
+& \jdeq
+\jcomp{}{\cftctxext}{\jcomp{}{\cftctxext}{\cftwk{\cftalg{A}}}}
+  \tag{2}
+  \\
+\jcomp{}{\jcomp{}{\cftctxext}{\cftfamext}}{\cftwk{\cftfamalg{\cftalg{A}}}}
+& \jdeq
+\jcomp{}{\cftctxext}{\jcomp{}{\cftfamext}{\cftwk{\cftalg{A}}}}
+  \tag{3}
+\end{align*}
+For the first, we have the judgmental equalities
+
+To show that $\cftwk{\cftfamalg{\cftalg{A}}}$ preserves itself, we must verify
+that the diagram
+\begin{equation*}
+\begin{tikzcd}[column sep=huge]
+\ctxwk{\cftalgf{\cftfamalg{\cftalg{A}}}}{{\cftalgf{\cftfamalg{\cftfamalg{\cftalg{A}}}}}{\cftfamalg{\cftfamalg{\cftfamalg{\cftalg{A}}}}}}
+  \ar{d}
+    [swap]{ \ctxwk
+        {{\cftalgf{\cftfamalg{\cftalg{A}}}}{\cftalgf{\cftfamalg{\cftfamalg{\cftalg{A}}}}}}
+        {\cftfamhom{\cftwk{\cftfamalg{\cftalg{A}}}}}
+      }
+  \ar{r}{\ctxwk{\cftalgf{\cftfamalg{\cftalg{A}}}}{\jcomp{}{\cftfamext}{\cftwk{\cftfamalg{\cftalg{A}}}}}}
+& \ctxwk{\cftalgf{\cftfamalg{\cftalg{A}}}}{\cftfamalg{\cftfamalg{\cftfamalg{\cftfamalg{\cftalg{A}}}}}}
+  \ar{d}{\jcomp{}{\ctxwk{\cftalgf{\cftfamalg{\cftalg{A}}}}{\jcomp{}{\cftctxext}{\cftfamext}}}{\cftfamhom{\cftwk{\cftfamalg{\cftalg{A}}}}}}
+  \\
+\ctxwk{{\cftalgf{\cftfamalg{\cftalg{A}}}}{\cftalgf{\cftfamalg{\cftfamalg{\cftalg{A}}}}}}{\jcomp{}{\cftwkc{\cftfamalg{\cftalg{A}}}}{\cftfamalg{\cftfamalg{\cftfamalg{\cftfamalg{\cftalg{A}}}}}}}
+  \ar{r}[swap,yshift=-1em]{%
+    \jcomp{}
+      { \cftwkf{\cftfamalg{\cftalg{A}}}}
+      { \jcomp{}
+          {\cftwkc{\cftfamalg{\cftalg{A}}}}
+          { \jcomp{}
+              {\cftfamext}
+              { \jcomp{}
+                  {\cftfamext}
+                  {\cftwk{\cftfamalg{\cftalg{A}}}}
+                }
+            }
+        }
+    }
+& \jcomp{}
+    {\ctxwk{\cftalgf{\cftfamalg{\cftalg{A}}}}{\jcomp{}{\cftctxext}{\cftfamext}}}
+    {\jcomp{}{\cftwkc{\cftfamalg{\cftalg{A}}}}{\cftfamalg{\cftfamalg{\cftfamalg{\cftfamalg{\cftalg{A}}}}}}}
+\end{tikzcd}
+\end{equation*}
+commutes judgmentally. We do this by showing that each of the ingredients of this
+diagram is judgmentally equal to the corresponding ingredient of the diagram in
+\autoref{diag:cftwk-cftwk} pulled back along $\cftctxext$. Thus, we are going
+to show that
+\begin{align*}
+\ctxwk
+  {\cftalgf{\cftfamalg{\cftalg{A}}}}
+  {\jcomp{}{\cftfamext}{\cftwk{\cftfamalg{\cftalg{A}}}}}
+& \jdeq
+  \jcomp{}
+    { \cftctxext}
+    { \ctxwk{\cftalgf{\cftalg{A}}}{\jcomp{}{\cftctxext}{\cftwk{\cftalg{A}}}}}
+  \tag{1}
+  \\
+\jcomp{}
+  {\ctxwk{\cftalgf{\cftfamalg{\cftalg{A}}}}{\jcomp{}{\cftctxext}{\cftfamext}}}
+  {\cftfamhom{\cftwk{\cftfamalg{\cftalg{A}}}}}
+& \jdeq
+  \jcomp{}
+    { \cftctxext}
+    { \jcomp{}
+        {\ctxwk{\cftalgf{\cftalg{A}}}{\cftfamext}}
+        {\cftfamhom{\cftwk{\cftalg{A}}}}
+      }
+  \tag{2}
+  \\
+\ctxwk
+  {{\cftalgf{\cftfamalg{\cftalg{A}}}}{\cftalgf{\cftfamalg{\cftfamalg{\cftalg{A}}}}}}
+  {\cftfamhom{\cftwk{\cftfamalg{\cftalg{A}}}}}
+& \jdeq
+  \jcomp{}
+    { \cftctxext}
+    { \ctxwk
+        {{\cftalgf{\cftalg{A}}}{\cftalgf{\cftfamalg{\cftalg{A}}}}}
+        {\cftfamhom{\cftwk{\cftalg{A}}}}
+      }
+  \tag{3}
+  \\
+\jcomp{}
+      { \cftwkf{\cftfamalg{\cftalg{A}}}}
+      { \jcomp{}
+          {\cftwkc{\cftfamalg{\cftalg{A}}}}
+          { \jcomp{}
+              {\cftfamext}
+              { \jcomp{}
+                  {\cftfamext}
+                  {\cftwk{\cftfamalg{\cftalg{A}}}}
+                }
+            }
+        }
+& \jdeq
+  \jcomp{}
+    { \cftctxext}
+    { \jcomp{}
+        { \cftwkf{\cftalg{A}}}
+        { \jcomp{}
+            {\cftwkc{\cftalg{A}}}
+            { \jcomp{}
+                {\cftctxext}
+                { \jcomp{}
+                    {\cftctxext}
+                    {\cftwk{\cftalg{A}}}
+                  }
+              }
+          }
+      }
+  \tag{4}
+\end{align*}
+For the first, note that we have the judgmental equalities
+\begin{align*}
+\ctxwk
+  {\cftalgf{\cftfamalg{\cftalg{A}}}}
+  {\jcomp{}{\cftfamext}{\cftwk{\cftfamalg{\cftalg{A}}}}}
+& \jdeq 
+  \ctxwk
+    {\jcomp{}{\cftctxext}{\cftalgf{\cftalg{A}}}}
+    {\jcomp{}{\cftfamext}{\jcomp{}{\cftctxext}{\cftwk{\cftalg{A}}}}}
+  \\
+& \jdeq
+  \ctxwk
+    {\jcomp{}{\cftctxext}{\cftalgf{\cftalg{A}}}}
+    {\jcomp{}{\cftctxext}{\jcomp{}{\cftctxext}{\cftwk{\cftalg{A}}}}}
+  \\
+& \jdeq
+  \jcomp{}
+    { \cftctxext}
+    { \ctxwk{\cftalgf{\cftalg{A}}}{\jcomp{}{\cftctxext}{\cftwk{\cftalg{A}}}}}
+\end{align*}
+For the second, note that we have the judgmental equalities
+\begin{align*}
+\jcomp{}
+  {\ctxwk{\cftalgf{\cftfamalg{\cftalg{A}}}}{\jcomp{}{\cftctxext}{\cftfamext}}}
+  {\cftfamhom{\cftwk{\cftfamalg{\cftalg{A}}}}}
+& \jdeq
+  \jcomp{}
+    {\ctxwk{\jcomp{}{\cftctxext}{\cftalgf{\cftalg{A}}}}{\jcomp{}{\cftctxext}{\cftfamext}}}
+    {\cftfamhom{\jcomp{}{\cftctxext}{\cftwk{\cftalg{A}}}}}
+  \\
+& \jdeq
+  \jcomp{}
+    {\jcomp{}{\cftctxext}{\ctxwk{\cftalgf{\cftalg{A}}}{\cftfamext}}}
+    {\cftfamhom{\jcomp{}{\cftctxext}{\cftwk{\cftalg{A}}}}}
+  \\
+& \jdeq
+  \jcomp{}
+    {\jcomp{}{\cftctxext}{\ctxwk{\cftalgf{\cftalg{A}}}{\cftfamext}}}
+    {\jcomp{}{\cftctxext}{\cftfamhom{\cftwk{\cftalg{A}}}}}
+  \\
+& \jdeq
+  \jcomp{}
+    { \cftctxext}
+    { \jcomp{}
+        {\ctxwk{\cftalgf{\cftalg{A}}}{\cftfamext}}
+        {\cftfamhom{\cftwk{\cftalg{A}}}}
+      }
+\end{align*}
+For the third, note that we have the judgmental equalities
+\begin{align*}
+\ctxwk
+  {{\cftalgf{\cftfamalg{\cftalg{A}}}}{\cftalgf{\cftfamalg{\cftfamalg{\cftalg{A}}}}}}
+  {\cftfamhom{\cftwk{\cftfamalg{\cftalg{A}}}}}
+& \jdeq
+  \ctxwk
+    {{\jcomp{}{\cftctxext}{\cftalgf{\cftalg{A}}}}{\jcomp{}{\cftctxext}{\cftalgf{\cftfamalg{\cftalg{A}}}}}}
+    {\cftfamhom{\cftwk{\cftfamalg{\cftalg{A}}}}}
+  \\
+& \jdeq
+  \ctxwk
+    {\jcomp{}{\cftctxext}{\ctxwk{\cftalgf{\cftalg{A}}}{\cftalgf{\cftfamalg{\cftalg{A}}}}}}
+    {\cftfamhom{\cftwk{\cftfamalg{\cftalg{A}}}}}
+  \\
+& \jdeq
+  \ctxwk
+    {\jcomp{}{\cftctxext}{\ctxwk{\cftalgf{\cftalg{A}}}{\cftalgf{\cftfamalg{\cftalg{A}}}}}}
+    {\cftfamhom{\jcomp{}{\cftctxext}{\cftwk{\cftalg{A}}}}}
+  \\
+& \jdeq
+  \ctxwk
+    {\jcomp{}{\cftctxext}{\ctxwk{\cftalgf{\cftalg{A}}}{\cftalgf{\cftfamalg{\cftalg{A}}}}}}
+    {\jcomp{}{\cftctxext}{\cftfamhom{\cftwk{\cftalg{A}}}}}
+  \\
+& \jdeq
+  \jcomp{}
+    { \cftctxext}
+    { \ctxwk
+        {{\cftalgf{\cftalg{A}}}{\cftalgf{\cftfamalg{\cftalg{A}}}}}
+        {\cftfamhom{\cftwk{\cftalg{A}}}}
+      }
+\end{align*}
+For the forth, note that we have the judgmental equalities
+\begin{align*}
+\jcomp{}
+      { \cftwkf{\cftfamalg{\cftalg{A}}}}
+      { \jcomp{}
+          {\cftwkc{\cftfamalg{\cftalg{A}}}}
+          { \jcomp{}
+              {\cftfamext}
+              { \jcomp{}
+                  {\cftfamext}
+                  {\cftwk{\cftfamalg{\cftalg{A}}}}
+                }
+            }
+        }
+& \jdeq
+  \jcomp{}
+      { \cftwkf{\cftfamalg{\cftalg{A}}}}
+      { \jcomp{}
+          { \jcomp{}{\cftctxext}{\cftwkc{\cftalg{A}}}}
+          { \jcomp{}
+              {\cftfamext}
+              { \jcomp{}
+                  {\cftfamext}
+                  {\jcomp{}{\cftctxext}{\cftwk{\cftalg{A}}}}
+                }
+            }
+        }
+  \\
+& \jdeq
+  \jcomp{}
+      { \cftwkf{\cftfamalg{\cftalg{A}}}}
+      { \jcomp{}
+          { \jcomp{}{\cftctxext}{\cftwkc{\cftalg{A}}}}
+          { \jcomp{}
+              {\cftctxext}
+              { \jcomp{}
+                  {\cftctxext}
+                  {\jcomp{}{\cftctxext}{\cftwk{\cftalg{A}}}}
+                }
+            }
+        }
+  \\
+& \jdeq
+  \jcomp{}
+      { \cftwkf{\cftfamalg{\cftalg{A}}}}
+      { \jcomp{}
+          { \cftctxext}
+          { \jcomp{}
+              { \cftwkc{\cftalg{A}}}
+              { \jcomp{}
+                  {\cftctxext}
+                  {\jcomp{}{\cftctxext}{\cftwk{\cftalg{A}}}}
+                }
+            }
+        }
+  \\
+& \jdeq
+  \jcomp{}
+      { \jcomp{}{\cftctxext}{\cftwkf{\cftalg{A}}}}
+      { \jcomp{}
+          { \cftctxext}
+          { \jcomp{}
+              { \cftwkc{\cftalg{A}}}
+              { \jcomp{}
+                  {\cftctxext}
+                  {\jcomp{}{\cftctxext}{\cftwk{\cftalg{A}}}}
+                }
+            }
+        }
+  \\
+& \jdeq
+  \jcomp{}
+    { \cftctxext}
+    { \jcomp{}
+        { \cftwkf{\cftalg{A}}}
+        { \jcomp{}
+            {\cftwkc{\cftalg{A}}}
+            { \jcomp{}
+                {\cftctxext}
+                { \jcomp{}
+                    {\cftctxext}
+                    {\cftwk{\cftalg{A}}}
+                  }
+              }
+          }
+      }
+\end{align*}
+Thus we may conclude that the diagram above of which we had the aim to show that
+it commutes judgmentally is a pullback of the diagram in \autoref{diag:cftwk-cftwk}
+by $\cftctxext$ and therefore it commutes as desired.
+\end{proof}
+
 \begin{thm}

### 2014-07-12 article-models-ttinternal.tex
a bit of progress on cftsubst

@@
 condition that the diagram
-\begin{equation}\label{diag:cftwk-curry}
+\begin{equation}\label{diag:cftwk-cftextcurry}
 \begin{tikzcd}[column sep=large]
@@
 $
-satisfying the judgmental equality of \autoref{diag:cftwk-curry}. Then we
+satisfying the judgmental equality of \autoref{diag:cftwk-cftextcurry}. Then we
 have the judgmental equality
@@
 satisfying the judgmental equalities of %
-\autoref{diag:cftwk-cftwk,eq:cftwk-cftextcurry,eq:cftwk-cftempcurry}.
+\autoref{diag:cftwk-cftwk,diag:cftwk-cftextcurry,eq:cftwk-cftempcurry}.
 \end{defn}
@@
 diagram is judgmentally equal to the corresponding ingredient of the diagram in
-\autoref{diag:cftwk-curry} pulled back along $\cftctxext$. Thus, we are going
+\autoref{diag:cftwk-cftextcurry} pulled back along $\cftctxext$. Thus, we are going
 to show that
@@
 For the first, we have the judgmental equalities
+\begin{align*}
+\ctxwk{\cftalgf{\cftfamalg{\cftfamalg{\cftalg{A}}}}}{\cftwk{\cftfamalg{\cftalg{A}}}}
+& \jdeq
+\ctxwk
+  {\jcomp{}{\cftctxext}{\cftalgf{\cftfamalg{\cftalg{A}}}}}
+  {\jcomp{}{\cftctxext}{\cftwk{\cftalg{A}}}}
+  \\
+& \jdeq
+\jcomp{}{\cftctxext}{\ctxwk{\cftalgf{\cftfamalg{\cftalg{A}}}}{\cftwk{\cftalg{A}}}}
+\end{align*}
+For the second, we have the judgmental equalities
+\begin{align*}
+\jcomp{}{\cftfamext}{\cftwk{\cftfamalg{\cftalg{A}}}}
+& \jdeq
+  \jcomp{}{\cftfamext}{\jcomp{}{\cftctxext}{\cftwk{\cftalg{A}}}}
+  \\
+& \jdeq
+  \jcomp{}{\cftctxext}{\jcomp{}{\cftctxext}{\cftwk{\cftalg{A}}}}
+\end{align*}
+For the third, note that we have the judgmental equalities
+\begin{align*}
+\jcomp{}{\jcomp{}{\cftctxext}{\cftfamext}}{\cftwk{\cftfamalg{\cftalg{A}}}}
+& \jdeq
+  \jcomp{}{\jcomp{}{\cftctxext}{\cftfamext}}{\jcomp{}{\cftctxext}{\cftwk{\cftalg{A}}}}
+  \\
+& \jdeq
+\jcomp{}{\cftctxext}{\jcomp{}{\cftfamext}{\cftwk{\cftalg{A}}}}
+\end{align*}
+Thus we may conclude that the diagram above a pullback of the diagram in 
+\autoref{diag:cftwk-cftextcurry} by $\cftctxext$ and therefore it commutes as desired.
 
@@
 \end{align*}
-For the forth, note that we have the judgmental equalities
+For the fourth, note that we have the judgmental equalities
 \begin{align*}
@@
 \end{align*}
-Thus we may conclude that the diagram above of which we had the aim to show that
-it commutes judgmentally is a pullback of the diagram in \autoref{diag:cftwk-cftwk}
+Thus we may conclude that the diagram above a pullback of the diagram in \autoref{diag:cftwk-cftwk}
 by $\cftctxext$ and therefore it commutes as desired.
@@
 
+\begin{thm}
+Suppose $\cftalg{A}$ is a projection algebra in context $\Gamma$. Then
+$\cftfamalg{\cftalg{A}}$ with the internal identity term
+\begin{equation*}
+\cftidtm{\cftfamalg{\cftalg{A}}}\defeq\jcomp{}{\cftctxext}{\cftidtm{\cftalg{A}}}
+\end{equation*}
+is a projection algebra in context $\ctxext{\Gamma}{\cftalgc{\cftalg{A}}}$.
+\end{thm}
+
+\begin{thm}
+Suppose $\cftalg{Q}$ is a projection algebra in context $\ctxext{\Gamma}{B}$ and
+let $\jfam{\Gamma}{A}$. Then $\ctxwk{A}{\cftalg{Q}}$ together with the
+internal identity term
+\begin{equation*}
+\cftidtm{\ctxwk{A}{\cftalg{Q}}}\defeq \ctxwk{A}\cftidtm{\cftalg{Q}}
+\end{equation*}
+is a projection algebra in context $\ctxext{{\Gamma}{A}}{B}$.
+\end{thm}
+
+\begin{thm}
+Suppose $\cftalg{Q}$ is a projection algebra in context $\ctxext{{\Gamma}{A}}{P}$
+and let $\jterm{\Gamma}{A}{x}$. Then $\subst{x}{\cftalg{Q}}$ together with the
+internal identity term
+\begin{equation*}
+\cftidtm{\subst{x}{\cftalg{Q}}} \defeq \subst{x}{\cftidtm{\cftalg{Q}}}
+\end{equation*}
+is a projection algebra in context $\ctxext{\Gamma}{\subst{x}{P}}$.
+\end{thm}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 is a substitution algebra in context $\Gamma$ if the diagram
-\begin{equation}\label{diag:substalg}
-\begin{tikzcd}
-X
+\begin{equation}\label{diag:cftsubst-cftsubst}
+\begin{tikzcd}[column sep=9em]
+\ctxwk
+  { \cftalgt{\cftalg{A}}}
+  { { \cftalgt{\cftfamalg{\cftfamalg{\cftalg{A}}}}}
+    { \cftfamalg{\cftfamalg{\cftfamalg{\cftfamalg{\cftalg{A}}}}}}
+    }
   \ar{r}
-  \ar{d}
-& X
-  \ar{d}
-  \\
-X 
+    { \ctxwk
+        { \cftalgt{\cftalg{A}}}
+        { \jcomp{}
+            { \cftctxext}
+            { \jcomp{}
+                { \cftctxext}
+                { \cftsubst{\cftalg{A}}}
+              }
+          }
+      }
+  \ar{d}[swap]
+    { \ctxwk
+        { { \cftalgt{\cftalg{A}}}
+          { \cftalgt{\cftfamalg{\cftfamalg{\cftalg{A}}}}}
+          }
+        { \jcomp{}{\ctxwk{\cftalgt{\cftalg{A}}}{\jcomp{}{\cftctxext}{\cftfamext}}}{\cftfamhom{\cftsubst{\cftalg{A}}}}}
+      }
+  & 
+\ctxwk
+  { \cftalgt{\cftalg{A}}}
+  { \cftfamalg{\cftfamalg{\cftfamalg{\cftalg{A}}}}}
+  \ar{d}{ \cftfamhom{\cftsubst{\cftalg{A}}}}
+  \\
+\ctxwk
+  { { \cftalgt{\cftalg{A}}}
+    { \cftalgt{\cftfamalg{\cftfamalg{\cftalg{A}}}}}
+    }
+  { \jcomp{}
+      { \ctxwk{\cftalgt{\cftalg{A}}}{\jcomp{}{\cftctxext}{\cftfamext}}}
+      { \jcomp{}
+          {\cftsubstc{\cftalg{A}}}
+          {\cftfamalg{\cftfamalg{\cftalg{A}}}}
+        }
+    }
   \ar{r}
-& X
+& \jcomp{}{\cftsubstc{\cftalg{A}}}{\cftfamalg{\cftfamalg{\cftalg{A}}}}
 \end{tikzcd}
 \end{equation}
-of CFT-algebras in context ... commutes judgmentally.
+of CFT-algebras in context 
+$ \ctxext
+    { { { { { {\Gamma}
+              {\cftalgc{\cftalg{A}}}
+              }
+            { \cftalgf{\cftalg{A}}}
+            }
+          { \cftalgt{\cftalg{A}}}
+          }
+        { \ctxwk{\cftalgt{\cftalg{A}}}{\cftalgf{\cftfamalg{\cftalg{A}}}}}
+        }
+      { \ctxwk{\cftalgt{\cftalg{A}}}{\cftalgf{\cftfamalg{\cftfamalg{\cftalg{A}}}}}}
+      }
+    { \ctxwk{\cftalgt{\cftalg{A}}}{\cftalgt{\cftfamalg{\cftfamalg{\cftalg{A}}}}}}$
+commutes judgmentally.
 \end{defn}
+
+\begin{rmk}
+We shall have to verify that the diagram in \autoref{diag:cftsubst-cftsubst} is
+well-typed.
+\end{rmk}
 

### 2014-07-13 article-models-ttinternal.tex
pre-weakening homomorphisms

@@
 \end{rmk}
+
+\begin{thm}
+Let $\extalg{A}$ be an extension algebra in context $\Gamma$. Then the triple
+\begin{equation*}
+\cftidhom{\extalg{A}}\defeq\unfold{\cftidhom{\extalg{A}}}
+\end{equation*}
+is an extension homomorphism from $\cftalg{A}$ to $\cftalg{A}$ in context
+$\Gamma$.
+\end{thm}
 
@@
 \begin{lem}
+For any CFT-algebra $\cftalg{A}$, the extension homomorphism $\cftidhom{\cftalg{A}}$
+is a CFT-homomorphism.
+\end{lem}
+
+\begin{lem}
 Suppose that $\cfthom{f}$ is a CFT-homomorphism from $\cftalg{A}$ to $\cftalg{B}$
@@
 
-Let $\cftalg{A}$ be a CFT-algebra in context $\Gamma$. Weakening will be a
-CFT-homomorphism
+\begin{defn}
+A pre-weakening algebra $\cftalg{A}$ in context $\Gamma$ is a CFT-algebra
+$\cftalg{A}$ in context $\Gamma$ together with a CFT-homomorphism
 \begin{equation}\label{cftwk}
@@
 \end{equation}
-satisfying judgmental equalities which express abstracted analogues of
+called the \emph{weakening of $\cftalg{A}$}.
+\end{defn}
+
+Weakening algebras will be pre-weakening algebras for which the weakening
+satisfies judgmental equalities which express abstracted analogues of
 the rules in \autoref{comp-ww,comp-ew,comp-0w}. The
@@
 
-We begin with some trivial remarks to familiarize ourselves with the situation.
-The morphism $\cftwk{\cftalg{A}}$ of \autoref{cftwk} is a triple
+We begin with some trivial remarks to familiarize ourselves with the situation
+of pre-weakening algebras.
+The CFT-homomorphism $\cftwk{\cftalg{A}}$ of \autoref{cftwk} is a triple
 $(\cftwkc{\cftalg{A}},\cftwkf{\cftalg{A}},\cftwkt{\cftalg{A}})$ consisting of
@@
 indeed an abstraction of the introduction rules for the weakening operation.
+\end{rmk}
+
+\begin{defn}
+Suppose $\cftalg{A}$ is a pre-weakening algebra in context $\Gamma$. Then we
+define the pre-weakening algebra $\cftfamalg{\cftalg{A}}$ in context 
+$\ctxext{\Gamma}{\cftalgc{\cftalg{A}}}$ by taking
+\begin{equation*}
+\cftwk{\cftfamalg{\cftalg{A}}}
+  \defeq \jcomp{}{\cftctxext}{\cftwk{\cftalg{A}}}
+\end{equation*}
+\end{defn}
+
+\begin{defn}
+Suppose $\cftalg{Q}$ is a pre-weakening algebra in context $\ctxext{\Gamma}{B}$ and
+let $\jfam{\Gamma}{A}$. Then we define the pre-weakening algebra 
+$\ctxwk{A}{\cftalg{Q}}$ in context $\ctxext{{\Gamma}{A}}{\ctxwk{A}{B}}$ by 
+taking
+\begin{equation*}
+\cftwk{\ctxwk{A}{\cftalg{Q}}}\defeq \ctxwk{A}\cftwk{\cftalg{Q}}.
+\end{equation*}
+\end{defn}
+
+\begin{defn}
+Suppose $\cftalg{Q}$ is a pre-weakening algebra in context $\ctxext{{\Gamma}{A}}{P}$
+and let $\jterm{\Gamma}{A}{x}$. Then we define the pre-weakening algebra 
+$\subst{x}{\cftalg{Q}}$ in context $\ctxext{\Gamma}{\subst{x}{P}}$ by taking
+\begin{equation*}
+\cftwk{\subst{x}{\cftalg{Q}}} \defeq \subst{x}{\cftwk{\cftalg{Q}}}.
+\end{equation*}
+\end{defn}
+
+\subsubsection{Pre-weakening homomorphisms}
+It is useful to start with the definition of pre-weakening homomorphisms. Those
+are CFT-homomorphisms which also preserve weakening. Before we give the
+definition, recall from \autoref{lem:jvcomp-cprojfstf} that we have a
+commutative square
+\begin{equation*}
+\begin{tikzcd}
+\ctxext{\cftalgc{\cftalg{A}}}{\cftalgf{\cftalg{A}}}
+  \ar{r}{\jvcomp{}{\cfthomc{\cfthom{f}}}{\cfthomf{\cfthom{f}}}}
+  \ar{d}[swap]{\cprojfstf{\cftalgc{\cftalg{A}}}{\cftalgf{\cftalg{A}}}}
+  &
+\ctxext{\cftalgc{\cftalg{A}}}{\cftalgf{\cftalg{B}}}
+  \ar{d}{\cprojfstf{\cftalgc{\cftalg{B}}}{\cftalgf{\cftalg{B}}}}
+  \\
+\cftalgc{\cftalg{A}}
+  \ar{r}[swap]{\cfthomc{\cfthom{f}}}
+  &
+\cftalgc{\cftalg{B}}
+\end{tikzcd}
+\end{equation*}
+of morphisms in context $\Gamma$ for every CFT-homomorphism $\cfthom{f}$ from
+$\cftalg{A}$ to $\cftalg{B}$ in context $\Gamma$. Because weakening is the
+same thing as pulling back along projection, we therefore see that the
+judgmental equality
+\begin{equation*}
+\jfameq
+  {{{\Gamma}{\cftalgc{\cftalg{A}}}}{\cftalgf{\cftalg{A}}}}
+  { \ctxwk
+      {\cftalgf{\cftalg{A}}}
+      {\jcomp{}{\cfthomc{\cfthom{f}}}{\cftfamalg{\cftalg{B}}}}
+    }
+  { \jcomp{}
+      { \cfthomf{\cfthom{f}}}
+      { \jcomp{}
+        {\cfthomc{\cfthom{f}}}
+        {\ctxwk{\cftalgf{\cftalg{B}}}{\cftfamalg{\cftalg{B}}}}
+        }
+    }
+\end{equation*}
+is valid for every CFT-homomorphism $\cfthom{f}$ from $\cftalg{A}$ to $\cftalg{B}$
+in context $\Gamma$. This judgmental equality helps us seeing that the condition
+on pre-weakening homomorphisms is well-typed.
+
+\begin{defn}
+Let $\cftalg{A}$ and $\cftalg{B}$ be pre-weakening algebras in context $\Gamma$.
+A \emph{pre-weakening homomorphism from $\cftalg{A}$ to $\cftalg{B}$ in context
+$\Gamma$} is a CFT-homomorphism $\cfthom{f}$ from $\cftalg{A}$ to $\cftalg{B}$
+in context $\Gamma$ for which the diagram
+\begin{equation*}
+\begin{tikzcd}[column sep=huge]
+\ctxwk{\cftalgf{\cftalg{A}}}{\cftfamalg{\cftalg{A}}}
+  \ar{r}{\ctxwk{\cftalgf{\cftalg{A}}}{\cftfamhom{\cfthom{f}}}}
+  \ar{d}[swap]{\cftwk{\cftalg{A}}}
+  &
+\jcomp{}
+  { \cfthomf{\cfthom{f}}}
+  { \jcomp{}
+    {\cfthomc{\cfthom{f}}}
+    {\ctxwk{\cftalgf{\cftalg{B}}}{\cftfamalg{\cftalg{B}}}}
+    }
+  \ar{d}
+    { \jcomp{}
+        {\cfthomf{\cfthom{f}}}
+        {\jcomp{}{\cfthomc{\cfthom{f}}}{\cftwk{\cftalg{B}}}}
+      }
+  \\
+\cftfamalg{\cftfamalg{\cftalg{A}}}
+  \ar{r}[swap]{\cftfamhom{\cftfamhom{\cfthom{f}}}}
+  &
+\jcomp{}
+  {\cfthomf{\cfthom{f}}}
+  {\jcomp{}{\cfthomc{\cfthom{f}}}{\cftfamalg{\cftfamalg{\cftalg{B}}}}}
+\end{tikzcd}
+\end{equation*}
+of CFT-algebras in context 
+$\ctxext{{\Gamma}{\cftalgc{\cftalg{A}}}}{\cftalgf{\cftalg{A}}}$ commutes
+judgmentally.
+\end{defn}
+
+\begin{rmk}
+In this remark we explain what it means that weakening is a pre-weakening
+homomorphism. Let $\cftalg{A}$ be a pre-weakening algebra. Then its weakening
+$\cftwk{\cftalg{A}}$ is a CFT-homomorphism from
+$\ctxwk{\cftalgf{\cftalg{A}}}{\cftfamalg{\cftalg{A}}}$ to
+$\cftfamalg{\cftfamalg{\cftalg{A}}}$. Therefore, $\cftwk{\cftalg{A}}$ is a
+pre-weakening homomorphism if the diagram
+\begin{equation*}
+\begin{tikzcd}[column sep=8em]
+\ctxwk
+  {\cftalgf{\ctxwk{\cftalgf{\cftalg{A}}}{\cftfamalg{\cftalg{A}}}}}
+  {\cftfamalg{\ctxwk{\cftalgf{\cftalg{A}}}{\cftfamalg{\cftalg{A}}}}}
+  \ar{r}
+    { \ctxwk
+        {\cftalgf{\ctxwk{\cftalgf{\cftalg{A}}}{\cftfamalg{\cftalg{A}}}}}
+        {\cftfamhom{\cftwk{\cftalg{A}}}}
+      }
+  \ar{d}[swap]{\cftwk{\ctxwk{\cftalgf{\cftalg{A}}}{\cftfamalg{\cftalg{A}}}}}
+  &
+\jcomp{}
+  { \cftwkf{\cftalg{A}}}
+  { \jcomp{}
+    {\cftwkc{\cftalg{A}}}
+    {\ctxwk{\cftalgf{\cftalg{B}}}{\cftfamalg{\cftfamalg{\cftfamalg{\cftalg{A}}}}}}
+    }
+  \ar{d}
+    { \jcomp{}
+        { \cftwkf{\cftalg{A}}}
+        { \jcomp{}
+            { \cftwkc{\cftalg{A}}}
+            { \cftwk{\cftfamalg{\cftfamalg{\cftalg{A}}}}}}
+      }
+  \\
+\cftfamalg{\cftfamalg{\ctxwk{\cftalgf{\cftalg{A}}}{\cftfamalg{\cftalg{A}}}}}
+  \ar{r}[swap]
+    { \cftfamhom
+        { \cftfamhom
+            { \cftwk{\cftalg{A}}}
+          }
+      }
+  &
+\jcomp{}
+  { \cftwkf{\cftalg{A}}}
+  { \jcomp{}
+      {\cftwkc{\cftalg{A}}}
+      {\cftfamalg{\cftfamalg{\cftfamalg{\cftfamalg{\cftalg{A}}}}}}
+    }
+\end{tikzcd}
+\end{equation*}
+Both the domain and the codomain are
+again pre-weakening algebras: the weakening of
+$\ctxwk{\cftalgf{\cftalg{A}}}{\cftfamalg{\cftalg{A}}}$ is
+$\ctxwk{\cftalgf{\cftalg{A}}}{\cftfamhom{\cftwk{\cftalg{A}}}}$; the weakening of
+$\cftfamalg{\cftfamalg{\cftalg{A}}}$ is
+$\cftfamhom{\cftfamhom{\cftwk{\cftalg{A}}}}$. Therefore, the condition that
+the weakening of $\cftalg{A}$ is a pre-weakening homomorphism becomes that
+the diagram
 \end{rmk}

### 2014-07-14 article-models-ttinternal.tex
progress on substitution algebras

@@
 
-\begin{rmk}
+\begin{rmk}\label{rmk:cftwk-prewk}
 In this remark we explain what it means that weakening is a pre-weakening
@@
     {\cftwkc{\cftalg{A}}}
-    {\ctxwk{\cftalgf{\cftalg{B}}}{\cftfamalg{\cftfamalg{\cftfamalg{\cftalg{A}}}}}}
+    {\ctxwk{\cftalgf{\cftfamalg{\cftfamalg{\cftalg{A}}}}}{\cftfamalg{\cftfamalg{\cftfamalg{\cftalg{A}}}}}}
     }
@@
 \end{equation*}
-Both the domain and the codomain are
-again pre-weakening algebras: the weakening of
-$\ctxwk{\cftalgf{\cftalg{A}}}{\cftfamalg{\cftalg{A}}}$ is
-$\ctxwk{\cftalgf{\cftalg{A}}}{\cftfamhom{\cftwk{\cftalg{A}}}}$; the weakening of
-$\cftfamalg{\cftfamalg{\cftalg{A}}}$ is
-$\cftfamhom{\cftfamhom{\cftwk{\cftalg{A}}}}$. Therefore, the condition that
-the weakening of $\cftalg{A}$ is a pre-weakening homomorphism becomes that
-the diagram
+of pre-weakening algebras in context
+$ \ctxext
+    { { { { \Gamma}
+          { \cftalgc{\cftalg{A}}}
+          }
+        { \cftalgf{\cftalg{A}}}
+        }
+      { \ctxwk{\cftalgf{\cftalg{A}}}{\cftalgf{\cftalg{A}}}}
+      }
+    { \cftalgf{\ctxwk{\cftalgf{\cftalg{A}}}{\cftfamalg{\cftalg{A}}}}
+      }$
+commutes judgmentally.
 \end{rmk}
 
+\begin{comment}
 \subsubsection{Currying for weakening}
@@
     }
-  { id_{\cftfamalg{\cftalg{A}}}
+  { \cftidhom{\cftfamalg{\cftalg{A}}}
     }
@@
 \end{proof}
+\end{comment}
 
@@
 \begin{defn}
-A weakening algebra in context $\Gamma$ is a CFT-algebra $\cftalg{A}$ in context
-$\Gamma$ for which there is a CFT-homomorphism
-\begin{equation*}
-\jhom
-  {{{\Gamma}{\cftalgc{\cftalg{A}}}}{\cftalgf{\cftalg{A}}}}
-  {\ctxwk{\cftalgf{\cftalg{A}}}{\cftfamalg{\cftalg{A}}}}
-  {\cftalg{F_{F_A}}}
-  {\cftwk{\cftalg{A}}}
-\end{equation*}
-satisfying the judgmental equalities of %
-\autoref{diag:cftwk-cftwk,diag:cftwk-cftextcurry,eq:cftwk-cftempcurry}.
+A weakening algebra in context $\Gamma$ is a pre-weakening algebra $\cftalg{A}$
+in context $\Gamma$ for which the weakening $\cftwk{\cftalg{A}}$ is a
+pre-weakening homomorphism as explained in \autoref{rmk:cftwk-prewk},
+satisfying the two additional judgmental equalities
+expressing that: (1) the diagram
+\begin{equation}\label{diag:cftwk-cftextcurry}
+\begin{tikzcd}[column sep=large]
+\ctxwk
+  {\ctxext{\cftalgf{\cftalg{A}}}{\cftalgf{\cftfamalg{\cftalg{A}}}}}
+  {\cftfamalg{\cftalg{A}}}
+  \ar{r}{\ctxwk{\cftalgf{\cftfamalg{\cftalg{A}}}}{\cftwk{\cftalg{A}}}}
+  \ar{dr}[swap]{\jcomp{}{\cftfamext}{\cftwk{\cftalg{A}}}}
+& \ctxwk{\cftalgf{\cftfamalg{\cftalg{A}}}}{\cftfamalg{\cftfamalg{\cftalg{A}}}}
+  \ar{d}{\jcomp{}{\cftctxext}{\cftwk{\cftalg{A}}}}
+  \\
+& \cftfamalg{\cftfamalg{\cftfamalg{\cftalg{A}}}}
+\end{tikzcd}
+\end{equation}
+and (2) that
+\begin{equation}\label{eq:cftwk-cftempcurry}
+\jhomeq
+  { {\Gamma}{\cftalgc{\cftalg{A}}}
+    }
+  { \cftfamalg{\cftalg{A}}
+    }
+  { \cftfamalg{\cftalg{A}}
+    }
+  { \subst{\cftempf{\cftalg{A}}}{\cftwk{\cftalg{A}}}
+    }
+  { \cftidhom{\cftfamalg{\cftalg{A}}}
+    }
+\end{equation}
+\end{defn}
+
+\begin{rmk}
+The condition in \autoref{diag:cftwk-cftextcurry} implements the rules of
+\autoref{comp-ew} and the condition in \autoref{eq:cftwk-cftempcurry} implements
+the rules of \autoref{comp-0w}.
+\end{rmk}
+
+\begin{defn}
+A \emph{weakening homomorphism} is a pre-weakening homomorphism between
+weakening algebras.
 \end{defn}
@@
 \subsection{Substitution algebras}
-Substitution algebras will be CFT-algebras with an additional substitution 
-operation. Given a CFT-algebra $\cftalg{A}$ in context $\Gamma$, substitution 
-will be a CFT-homomorphism
+\begin{defn}
+A \emph{pre-substitution algebra $\cftalg{A}$ in context $\Gamma$} consists of a
+CFT-algebra $\cftalg{A}$ in context $\Gamma$ and a CFT-homomorphism
 \begin{equation*}
@@
   {\ctxwk{\cftalgt{\cftalg{A}}}{\cftalg{F_{F_A}}}}
-  {\cftfamalg{\cftalg{A}}}
-  {\cftsubst{\cftalg{A}}}
-\end{equation*}
-We will require additionally that substitution preserves itself. Thus, we will
-formulate an abstraction of the rules stated in \autoref{comp-ss}.
+  {\ctxwk{\cftalgt{\cftalg{A}}}{{\cftalgf{\cftalg{A}}}{\cftfamalg{\cftalg{A}}}}}
+  {\cftsubst{\cftalg{A}}}.
+\end{equation*}
+\end{defn}
 
 \begin{defn}
-Let $\cftalg{A}$ be a CFT-algebra in context $\Gamma$ and let
-\begin{equation*}
-\jhom
-  {{{{\Gamma}{\cftalgc{\cftalg{A}}}}{\cftalgf{\cftalg{A}}}}{\cftalgt{\cftalg{A}}}}
-  {\ctxwk{\cftalgt{\cftalg{A}}}{\cftalg{F_{F_A}}}}
-  {\cftfamalg{\cftalg{A}}}
-  {\cftsubst{\cftalg{A}}}
-\end{equation*}
-be a CFT-homomorphism. We say that the pair $(\cftalg{A},\cftsubst{\cftalg{A}})$
-is a substitution algebra in context $\Gamma$ if the diagram
-\begin{equation}\label{diag:cftsubst-cftsubst}
-\begin{tikzcd}[column sep=9em]
+Suppose $\cftalg{A}$ is a pre-substitution algebra in context $\Gamma$. Then we
+define the pre-substitution algebra $\cftfamalg{\cftalg{A}}$ in context 
+$\ctxext{\Gamma}{\cftalgc{\cftalg{A}}}$ by taking
+\begin{equation*}
+\cftsubst{\cftfamalg{\cftalg{A}}}
+  \defeq \jcomp{}{\cftctxext}{\cftsubst{\cftalg{A}}}.
+\end{equation*}
+\end{defn}
+
+\begin{defn}
+Suppose $\cftalg{Q}$ is a pre-substitution algebra in context $\ctxext{\Gamma}{B}$ and
+let $\jfam{\Gamma}{A}$. Then we define the pre-substitution algebra 
+$\ctxwk{A}{\cftalg{Q}}$ in context $\ctxext{{\Gamma}{A}}{\ctxwk{A}{B}}$ by 
+taking
+\begin{equation*}
+\cftsubst{\ctxwk{A}{\cftalg{Q}}}\defeq \ctxwk{A}\cftsubst{\cftalg{Q}}.
+\end{equation*}
+\end{defn}
+
+\begin{defn}
+Suppose $\cftalg{Q}$ is a pre-substitution algebra in context $\ctxext{{\Gamma}{A}}{P}$
+and let $\jterm{\Gamma}{A}{x}$. Then we define the pre-substitution algebra 
+$\subst{x}{\cftalg{Q}}$ in context $\ctxext{\Gamma}{\subst{x}{P}}$ by taking
+\begin{equation*}
+\cftsubst{\subst{x}{\cftalg{Q}}} \defeq \subst{x}{\cftsubst{\cftalg{Q}}}.
+\end{equation*}
+\end{defn}
+
+\begin{defn}
+A \emph{pre-substitution homomorphism $\cfthom{f}$ from $\cftalg{A}$ to $\cftalg{B}$ in
+context $\Gamma$} is a CFT-homomorphism $\cfthom{f}$ for which the diagram
+\begin{equation*}
+\begin{tikzcd}[column sep=8em]
+\ctxwk{\cftalgt{\cftalg{A}}}{\cftfamalg{\cftfamalg{\cftalg{A}}}}
+  \ar{r}{\ctxwk{\cftalgt{\cftalg{A}}}{\cftfamhom{\cftfamhom{\cfthom{f}}}}}
+  \ar{d}[swap]{\cftsubst{\cftalg{A}}}
+  &
+\jcomp{}
+  { \cfthomt{\cfthom{f}}}
+  { \jcomp{}
+      { \cfthomf{\cfthom{f}}}
+      { \jcomp{}
+          { \cfthomc{\cfthom{f}}}
+          { \ctxwk{\cftalgt{\cftalg{B}}}{\cftfamalg{\cftfamalg{\cftalg{B}}}}}
+        }
+    }
+  \ar{d}
+    { \jcomp{}
+        { \cfthomt{\cfthom{f}}}
+        { \jcomp{}
+            { \cfthomf{\cfthom{f}}}
+            { \jcomp{}
+                { \cfthomc{\cfthom{f}}}
+                {\cftsubst{\cftalg{B}}}
+              }
+          }
+      }   
+  \\
+\ctxwk{\cftalgt{\cftalg{A}}}{{\cftalgf{\cftalg{A}}}{\cftfamalg{\cftalg{A}}}}
+  \ar{r}[swap]
+    {\ctxwk{\cftalgt{\cftalg{A}}}{{\cftalgf{\cftalg{A}}}{\cftfamhom{\cfthom{f}}}}}
+  &
+\jcomp{}
+  { \cfthomt{\cfthom{f}}}
+  { \jcomp{}
+      { \cfthomf{\cfthom{f}}}
+      { \jcomp{}
+          { \cfthomc{\cfthom{f}}}
+          { \ctxwk
+              {\cftalgt{\cftalg{B}}}
+              {{\cftalgf{\cftalg{B}}}{\cftfamalg{\cftalg{B}}}}
+            }
+        }
+    }
+\end{tikzcd}
+\end{equation*}
+of pre-substitution algebras in context
+$\ctxext{{{\Gamma}{\cftalgc{\cftalg{A}}}}{\cftalgf{\cftalg{A}}}}{\cftalgt{\cftalg{A}}}$
+commutes judgmentally.
+\end{defn}
+
+\begin{defn}
+A \emph{substitution algebra in context $\Gamma$} is a pre-substitution algebra
+in context $\Gamma$ for which substitution is a pre-substitution homomorphism.
+\end{defn}
+
+\begin{rmk}
+The condition that the substitution of a pre-substitution algebra is a
+pre-substitution homomorphism assert that the diagram
+\begin{small}
+\begin{equation*}
+\begin{tikzcd}[column sep=large]
 \ctxwk
-  { \cftalgt{\cftalg{A}}}
-  { { \cftalgt{\cftfamalg{\cftfamalg{\cftalg{A}}}}}
-    { \cftfamalg{\cftfamalg{\cftfamalg{\cftfamalg{\cftalg{A}}}}}}
-    }
+  {\cftalgt{\ctxwk{\cftalgt{\cftalg{A}}}{\cftalg{F_{F_A}}}}}
+  {\cftfamalg{\cftfamalg{\ctxwk{\cftalgt{\cftalg{A}}}{\cftalg{F_{F_A}}}}}}
   \ar{r}
     { \ctxwk
+        {\cftalgt{\ctxwk{\cftalgt{\cftalg{A}}}{\cftalg{F_{F_A}}}}}
+        {\cftfamhom{\cftfamhom{\cftsubst{\cftalg{A}}}}}
+      }
+  \ar{d}[swap]
+    {\cftsubst{\ctxwk{\cftalgt{\cftalg{A}}}{\cftalg{F_{F_A}}}}}
+  &
+\jcomp{}
+  { \cftsubstt{\cftalg{A}}}
+  { \jcomp{}
+      { \cftsubstf{\cftalg{A}}}
+      { \jcomp{}
+          { \cftsubstc{\cftalg{A}}}
+          { \ctxwk
+              { \cftalgt
+                  { \ctxwk
                       { \cftalgt{\cftalg{A}}}
-        { \jcomp{}
-            { \cftctxext}
-            { \jcomp{}
-                { \cftctxext}
-                { \cftsubst{\cftalg{A}}}
-              }
-          }
-      }
-  \ar{d}[swap]
+                      { {\cftalgf{\cftalg{A}}}{\cftfamalg{\cftalg{A}}}}
+                    }
+                }
+              { \cftfamalg
+                  { \cftfamalg
                       { \ctxwk
-        { { \cftalgt{\cftalg{A}}}
-          { \cftalgt{\cftfamalg{\cftfamalg{\cftalg{A}}}}}
-          }
-        { \jcomp{}{\ctxwk{\cftalgt{\cftalg{A}}}{\jcomp{}{\cftctxext}{\cftfamext}}}{\cftfamhom{\cftsubst{\cftalg{A}}}}}
+                          { \cftalgt{\cftalg{A}}}
+                          { {\cftalgf{\cftalg{A}}}{\cftfamalg{\cftalg{A}}}}
+                        }
+                    }
+                }
+            }
+        }
+    }
+  \ar{d}
+    { \jcomp{}
+        { \cftsubstt{\cftalg{A}}}
+        { \jcomp{}
+            { \cftsubstf{\cftalg{A}}}
+            { \jcomp{}
+                { \cftsubstc{\cftalg{A}}}
+                { \cftsubst
+                    { \ctxwk
+                        {\cftalgt{\cftalg{A}}}
+                        {{\cftalgf{\cftalg{A}}}{\cftfamalg{\cftalg{A}}}}
+                      }
+                  }
+              }
+          }
+      }
+  \\
+\ctxwk
+  { \cftalgt{\ctxwk{\cftalgt{\cftalg{A}}}{\cftalg{F_{F_A}}}}}
+  { {\cftalgf{\ctxwk{\cftalgt{\cftalg{A}}}{\cftalg{F_{F_A}}}}}
+    {\cftfamalg{\ctxwk{\cftalgt{\cftalg{A}}}{\cftalg{F_{F_A}}}}}
+    }
+  \ar{r}[swap,yshift=-1em]
+    { \ctxwk
+        { \cftalgt{\ctxwk{\cftalgt{\cftalg{A}}}{\cftalg{F_{F_A}}}}}
+        { {\cftalgf{\ctxwk{\cftalgt{\cftalg{A}}}{\cftalg{F_{F_A}}}}}
+          {\cftfamhom{\cftsubst{\cftalg{A}}}}}
           }
   &
-\ctxwk
+\jcomp{}
+  { \cftsubstt{\cftalg{A}}}
+  { \jcomp{}
+      { \cftsubstf{\cftalg{A}}}
+      { \jcomp{}
+          { \cftsubstc{\cftalg{A}}}
+          { \ctxwk
+              { \cftalgt
+                  { \ctxwk
+                      {\cftalgt{\cftalg{A}}}
+                      {{\cftalgf{\cftalg{A}}}{\cftfamalg{\cftalg{A}}}}
+                    }
+                }
+              { { \cftalgf
+                    { \ctxwk
                         { \cftalgt{\cftalg{A}}}
-  { \cftfamalg{\cftfamalg{\cftfamalg{\cftalg{A}}}}}
-  \ar{d}{ \cftfamhom{\cftsubst{\cftalg{A}}}}
-  \\
-\ctxwk
-  { { \cftalgt{\cftalg{A}}}
-    { \cftalgt{\cftfamalg{\cftfamalg{\cftalg{A}}}}}
-    }
-  { \jcomp{}
-      { \ctxwk{\cftalgt{\cftalg{A}}}{\jcomp{}{\cftctxext}{\cftfamext}}}
-      { \jcomp{}
-          {\cftsubstc{\cftalg{A}}}
-          {\cftfamalg{\cftfamalg{\cftalg{A}}}}
-        }
-    }
-  \ar{r}
-& \jcomp{}{\cftsubstc{\cftalg{A}}}{\cftfamalg{\cftfamalg{\cftalg{A}}}}
+                        {{\cftalgf{\cftalg{A}}}{\cftfamalg{\cftalg{A}}}}
+                      }
+                  }
+                { \cftfamalg
+                    { \ctxwk
+                        {\cftalgt{\cftalg{A}}}
+                        {{\cftalgf{\cftalg{A}}}{\cftfamalg{\cftalg{A}}}}
+                      }
+                  }
+                }
+            }
+        }
+    }
 \end{tikzcd}
-\end{equation}
-of CFT-algebras in context 
-$ \ctxext
-    { { { { { {\Gamma}
-              {\cftalgc{\cftalg{A}}}
-              }
-            { \cftalgf{\cftalg{A}}}
-            }
-          { \cftalgt{\cftalg{A}}}
-          }
-        { \ctxwk{\cftalgt{\cftalg{A}}}{\cftalgf{\cftfamalg{\cftalg{A}}}}}
-        }
-      { \ctxwk{\cftalgt{\cftalg{A}}}{\cftalgf{\cftfamalg{\cftfamalg{\cftalg{A}}}}}}
-      }
-    { \ctxwk{\cftalgt{\cftalg{A}}}{\cftalgt{\cftfamalg{\cftfamalg{\cftalg{A}}}}}}$
-commutes judgmentally.
+\end{equation*}
+\end{small}
+\end{rmk}
+
+\begin{defn}
+A substitution homomorphism is a pre-substitution homomorphism between
+substitution algebras.
 \end{defn}
-
-\begin{rmk}
-We shall have to verify that the diagram in \autoref{diag:cftsubst-cftsubst} is
-well-typed.
-\end{rmk}
 

### 2014-07-14 article-models-ttinternal.tex
small changes

@@
 \end{equation}
-and (2) that
+of pre-weakening algebras in context $\ctxext{{{\Gamma}{\cftalgc{\cftalg{A}}}}{\cftalgf{\cftalg{A}}}}{\cftalgf{\cftfamalg{\cftalg{A}}}}$ commutes judgmentally and (2) that
 \begin{equation}\label{eq:cftwk-cftempcurry}
@@
 identity terms can only be formulated in weakening algebras and together with
-weakening, the identity terms provide for all the projections. In a weakening
-algebra $\cftalg{A}$ in context $\Gamma$, identity terms are implemented by a term
+weakening, the identity terms provide for all the projections.
+
+\begin{defn}
+A \emph{pre-projection algebra $\cftalg{A}$ in context $\Gamma$} consists of a 
+weakening algebra $\cftalg{A}$ in context $\Gamma$ together with an 
+\emph{internal identity term}
 \begin{equation*}
@@
   {\subst{{\idtm{\cftalgf{\cftalg{A}}}}{\cftwkc{\cftalg{A}}}}{\cftalgt{\cftfamalg{\cftalg{A}}}}}
-  {\cftidtm{\cftalg{A}}}
-\end{equation*}
-We will require additionally that identity terms are compatible with weakening,
-i.e.~an abstraction of the rule stated in \autoref{comp-wi}. Since this is the
-only additional rule for projection algebras, we immediately state the 
-definition.
+  {\cftidtm{\cftalg{A}}}.
+\end{equation*}
+\end{defn}
 
 \begin{defn}
-A \emph{projection algebra $\cftalg{A}$ in context $\Gamma$} is a weakening algebra
-$\cftalg{A}$ in context $\Gamma$ together with a term
-\begin{equation*}
-\jterm
-  { { {\Gamma}
-      {\cftalgc{\cftalg{A}}}
-      }
-    { \cftalgf{\cftalg{A}}
-      }
-    }
-  {\subst{{\idtm{\cftalgf{\cftalg{A}}}}{\cftwkc{\cftalg{A}}}}{\cftalgt{\cftfamalg{\cftalg{A}}}}}
-  {\cftidtm{\cftalg{A}}}
-\end{equation*}
-satisfying the judgmental equality
-\begin{align*}
-& \subst
-  { \ctxwk
-      { \cftalgf{\cftalg{A}}}
-      { \jcomp
-        { }
-        { \cftctxext}
-        { \cftidtm{\cftalg{A}}
-          }
-        }
-    }
-  { { \ctxwk
-        { \cftalgf{\cftalg{A}}}
-        { \subst { \idtm
-            { \jcomp
-                { }
-                { \cftctxext}
-                { \cftalgf{\cftalg{A}}
-                  }
-              }
-          }
-          { \jcomp
-              { }
-              { \cftctxext
-                }
-              { \cftwkc{\cftalg{A}}
-                }
-            }
-          }
-      }
-    { \jcomp
-        { }
-        { \ctxwk{\cftalgf{\cftalg{A}}}{\cftfamext}}
-        { \cftwkt{\cftalg{A}}}
-      }
-    }
-  \\
-& \jdeq
-  \jcomp{}
-    { \cftwkf{\cftalg{A}}}
-    { \jcomp{}
-        { \cftwkc{\cftalg{A}}}
-        { \jcomp{}
-            { \cftctxext}
-            { \jcomp{}
-              { \cftctxext}
-              { \cftidtm{\cftalg{A}}}
-              }
-          }
-      }
-\end{align*}
-of terms of %
-$ \jcomp{}
-    { \cftwkf{\cftalg{A}}}
-    { \jcomp{}
-        { \cftwkc{\cftalg{A}}}
-        { \jcomp{}
-            { \cftctxext}
-            { \jcomp{}
-              { \cftctxext}
-              { \subst
-                  { {\idtm{\cftalgf{\cftalg{A}}}}
-                    {\cftwkc{\cftalg{A}}}
-                    }
-                  { \cftalgt{\cftfamalg{\cftalg{A}}}
-                    }
-                }
-              }
-          }
-      }$
-in context \ldots.
-
-The term $\cftidtm{\cftalg{A}}$ of a projection algebra $\cftalg{A}$ in context
-$\Gamma$ is called the \emph{internal identity term of $\cftalg{A}$}.
+Suppose $\cftalg{A}$ is a pre-projection algebra in context $\Gamma$. Then we
+define the pre-projection algebra $\cftfamalg{\cftalg{A}}$ in context 
+$\ctxext{\Gamma}{\cftalgc{\cftalg{A}}}$ by taking
+\begin{equation*}
+\cftidtm{\cftfamalg{\cftalg{A}}}
+  \defeq \jcomp{}{\cftctxext}{\cftidtm{\cftalg{A}}}.
+\end{equation*}
 \end{defn}
 
-\begin{rmk}
-We shall have to verify that the condition on the internal identity term 
-$\cftidtm{\cftalg{A}}$ of a projection algebra $\cftalg{A}$ in context $\Gamma$
-is indeed well-typed. More precisely, we have to demonstrate that the judgmental
-equality
-\begin{align*}
-& \jdeq
-  \jcomp{}
-    { \cftwkf{\cftalg{A}}}
-    { \jcomp{}
-        { \cftwkc{\cftalg{A}}}
-        { \jcomp{}
-            { \cftctxext}
-            { \jcomp{}
-              { \cftctxext}
-              { \subst
-                  { {\idtm{\cftalgf{\cftalg{A}}}}
-                    {\cftwkc{\cftalg{A}}}
-                    }
-                  { \cftalgt{\cftfamalg{\cftalg{A}}}
-                    }
-                }
-              }
-          }
-      }  
-\end{align*}
-of families in context \ldots holds.
-\end{rmk}
+\begin{defn}
+Suppose $\cftalg{Q}$ is a pre-projection algebra in context $\ctxext{\Gamma}{B}$ and
+let $\jfam{\Gamma}{A}$. Then we define the pre-projection algebra 
+$\ctxwk{A}{\cftalg{Q}}$ in context $\ctxext{{\Gamma}{A}}{\ctxwk{A}{B}}$ by 
+taking
+\begin{equation*}
+\cftidtm{\ctxwk{A}{\cftalg{Q}}}\defeq \ctxwk{A}\cftidtm{\cftalg{Q}}.
+\end{equation*}
+\end{defn}
+
+\begin{defn}
+Suppose $\cftalg{Q}$ is a pre-projection algebra in context $\ctxext{{\Gamma}{A}}{P}$
+and let $\jterm{\Gamma}{A}{x}$. Then we define the pre-projection algebra 
+$\subst{x}{\cftalg{Q}}$ in context $\ctxext{\Gamma}{\subst{x}{P}}$ by taking
+\begin{equation*}
+\cftidtm{\subst{x}{\cftalg{Q}}} \defeq \subst{x}{\cftidtm{\cftalg{Q}}}.
+\end{equation*}
+\end{defn}
+
+\begin{defn}
+A pre-projection homomorphism from $\cftalg{A}$ to $\cftalg{B}$ in context
+$\Gamma$ is a weakening homomorphism $\cfthom{f}$ from $\cftalg{A}$ to
+$\cftalg{B}$ in context $\Gamma$ for which the diagram
+\begin{equation*}
+\begin{tikzcd}[column sep=10em]
+\subst
+  {{\idtm{\cftalgf{\cftalg{A}}}}{\cftwkc{\cftalg{A}}}}
+  {\cftalgt{\cftfamalg{\cftalg{A}}}}
+  \ar{r}{\subst{{\idtm{\cftalgf{\cftalg{A}}}}{\cftwkc{\cftalg{A}}}}{\cfthomt{\cftfamhom{\cfthom{f}}}}}
+  \ar[fib,xshift=-.7ex]{d}
+  &
+\subst
+  {{\idtm{\cftalgf{\cftalg{B}}}}{\cftwkc{\cftalg{B}}}}
+  {\cftalgt{\cftfamalg{\cftalg{B}}}}
+  \ar[fib,xshift=-.7ex]{d}
+  \\
+\cftalgf{\cftalg{A}}
+  \ar{r}[swap]{\cfthomf{\cfthom{f}}}
+  \ar[fib]{d}
+  \ar[densely dotted,xshift=.7ex]{u}[swap]{\cftidtm{\cftalg{A}}}
+  &
+\cftalgf{\cftalg{B}}
+  \ar[fib]{d}
+  \ar[densely dotted,xshift=.7ex]{u}[swap]{\cftidtm{\cftalg{B}}}
+  \\
+\cftalgc{\cftalg{A}}
+  \ar{r}[swap]{\cfthomc{\cfthom{f}}}
+  &
+\cftalgc{\cftalg{B}}
+\end{tikzcd}
+\end{equation*}
+\end{defn}
+
+\begin{defn}
+A projection algebra in context $\Gamma$ is a pre-projection algebra in context
+$\Gamma$ for which weakening is a pre-projection homomorphism.
+\end{defn}
+
+\begin{defn}
+A projection homomorphism is a pre-projection homomorphism between projection
+algebras.
+\end{defn}
 
 \begin{thm}
-Suppose $\cftalg{A}$ is a projection algebra in context $\Gamma$. Then
-$\cftfamalg{\cftalg{A}}$ with the internal identity term
-\begin{equation*}
-\cftidtm{\cftfamalg{\cftalg{A}}}\defeq\jcomp{}{\cftctxext}{\cftidtm{\cftalg{A}}}
-\end{equation*}
-is a projection algebra in context $\ctxext{\Gamma}{\cftalgc{\cftalg{A}}}$.
+Suppose $\cftalg{A}$ is a projection algebra in context $\Gamma$. Then the 
+pre-projection algebra $\cftfamalg{\cftalg{A}}$ is a projection algebra in context 
+$\ctxext{\Gamma}{\cftalgc{\cftalg{A}}}$.
 \end{thm}
@@
 Suppose $\cftalg{Q}$ is a projection algebra in context $\ctxext{\Gamma}{B}$ and
-let $\jfam{\Gamma}{A}$. Then $\ctxwk{A}{\cftalg{Q}}$ together with the
-internal identity term
-\begin{equation*}
-\cftidtm{\ctxwk{A}{\cftalg{Q}}}\defeq \ctxwk{A}\cftidtm{\cftalg{Q}}
-\end{equation*}
-is a projection algebra in context $\ctxext{{\Gamma}{A}}{B}$.
+let $\jfam{\Gamma}{A}$. Then the pre-projection algebra 
+$\ctxwk{A}{\cftalg{Q}}$ is a
+projection algebra in context $\ctxext{{\Gamma}{A}}{\ctxwk{A}{B}}$.
 \end{thm}
@@
 Suppose $\cftalg{Q}$ is a projection algebra in context $\ctxext{{\Gamma}{A}}{P}$
-and let $\jterm{\Gamma}{A}{x}$. Then $\subst{x}{\cftalg{Q}}$ together with the
-internal identity term
-\begin{equation*}
-\cftidtm{\subst{x}{\cftalg{Q}}} \defeq \subst{x}{\cftidtm{\cftalg{Q}}}
-\end{equation*}
-is a projection algebra in context $\ctxext{\Gamma}{\subst{x}{P}}$.
+and let $\jterm{\Gamma}{A}{x}$. Then the pre-projection algebra 
+$\subst{x}{\cftalg{Q}}$ is a
+projection algebra in context $\ctxext{\Gamma}{\subst{x}{P}}$.
 \end{thm}
+
+\begin{cor}
+In a projection algebra $\cftalg{A}$ in context $\Gamma$, weakening is a
+projection homomorphism.
+\end{cor}
 
@@
 
-\begin{lem}
-Let $\cftalg{A}$ be a substitution algebra in context $\Gamma$. Then we have the
-judgmental equality
-\begin{align*}
-& \jcomp{}
-    { \apply\cftsubst{\cftalg{A}}%
-      { \apply\cftctxext%
-          { \apply\cftctxext{\gamma}{a}}{p}
-        }
-      { q}
-      { g}
-      }
-    { \subst
-        { q}
-        { { p}
-          { \cftfamhom{\apply\cftsubst{\cftalg{A}}{\gamma}{a}{x}}}
-          }
-      }
-  \\
-& \jdeq
-\end{align*}
-of CFT-homomorphisms from ... to ... in context $\Gamma$,
-for $\gamma:A$, $a:\subst{\gamma}{\cftalgf{\cftalg{A}}}$, 
-$x:\subst{a}{{\gamma}{\cftalgt{\cftalg{A}}}}$,
-$p:\subst{\apply\cftctxext{\gamma}{a}}{\cftalgf{\cftalg{A}}}$
-$q:\subst{\apply\cftctxext{\apply\cftctxext{\gamma}{a}}{p}}{\cftalgf{\cftalg{A}}}$
-and $g:\subst{\apply\cftctxext{\apply\cftctxext{\apply\cftctxext{\gamma}{a}}{p}}{q}}{\cftalgt{\cftalg{A}}}$
-\end{lem}
+\begin{thm}
+Suppose $\cftalg{A}$ is a substitution algebra in context $\Gamma$. Then
+the pre-substitution algebra $\cftfamalg{\cftalg{A}}$ is a substitution algebra
+in context 
+$\ctxext{\Gamma}{\cftalgc{\cftalg{A}}}$.
+\end{thm}
+
+\begin{thm}
+Suppose $\cftalg{Q}$ is a substitution algebra in context $\ctxext{\Gamma}{B}$ 
+and let $\jfam{\Gamma}{A}$. Then the pre-substitution algebra 
+$\ctxwk{A}{\cftalg{Q}}$ is a substitution algebra in context 
+$\ctxext{{\Gamma}{A}}{\ctxwk{A}{B}}$.
+\end{thm}
+
+\begin{thm}
+Suppose $\cftalg{Q}$ is a substitution algebra in context $\ctxext{{\Gamma}{A}}{P}$
+and let $\jterm{\Gamma}{A}{x}$. Then the pre-substitution algebra 
+$\subst{x}{\cftalg{Q}}$ is a substitution algebra in context 
+$\ctxext{\Gamma}{\subst{x}{P}}$.
+\end{thm}
+
+\begin{cor}
+In a substitution algebra $\cftalg{A}$ in context $\Gamma$, substitution is a
+substitution homomorphism.
+\end{cor}
 
@@
 \subsection{E-algebras}
-E-algebras will be CFT-algebras with a weakening operation, identity terms and
-a substitution operation making it a projection algebra and a substitution 
-algebra at the same time. Additionally the weakening operation, identity terms
-and the substitution operation are all required to be compatible with each
-other. Thus, we need to require the abstractions of the rules stated in
-\autoref{comp-sw,comp-ws,cancellation-ws,comp-si,cancellation-i}.
+\begin{defn}
+A CFT-algebra $\cftalg{A}$ in context $\Gamma$ with internal weakening, identity terms and
+substitution
+\begin{align*}
+\jalign\jhom
+  {{{\Gamma}{\cftalgc{\cftalg{A}}}}{\cftalgf{\cftalg{A}}}}
+  {\ctxwk{\cftalgf{\cftalg{A}}}{\cftfamalg{\cftalg{A}}}}
+  {\cftfamalg{\cftfamalg{\cftalg{A}}}}
+  {\cftwk{\cftalg{A}}}
+  \\
+\jalign\jterm
+  { { {\Gamma}
+      {\cftalgc{\cftalg{A}}}
+      }
+    { \cftalgf{\cftalg{A}}
+      }
+    }
+  {\subst{{\idtm{\cftalgf{\cftalg{A}}}}{\cftwkc{\cftalg{A}}}}{\cftalgt{\cftfamalg{\cftalg{A}}}}}
+  {\cftidtm{\cftalg{A}}}
+  \\
+\jalign\jhom
+  {{{{\Gamma}{\cftalgc{\cftalg{A}}}}{\cftalgf{\cftalg{A}}}}{\cftalgt{\cftalg{A}}}}
+  {\ctxwk{\cftalgt{\cftalg{A}}}{\cftalg{F_{F_A}}}}
+  {\ctxwk{\cftalgt{\cftalg{A}}}{{\cftalgf{\cftalg{A}}}{\cftfamalg{\cftalg{A}}}}}
+  {\cftsubst{\cftalg{A}}}
+\end{align*}
+is said to be an \emph{E-algebra in context $\Gamma$} if both weakening and substitution are both
+projection and substitution homomorphisms and if (1) the diagram
+\begin{equation*}
+\begin{tikzcd}[column sep=large]
+\ctxwk
+  { \cftalgt{\cftalg{A}}}
+  { { \cftalgf{\cftalg{A}}}
+    { \cftfamalg{\cftalg{A}}}
+    }
+  \ar{r}{\ctxwk{\cftalgt{\cftalg{A}}}{\cftwk{\cftalg{A}}}}
+  \ar{dr}[swap]{\cftidhom{}}
+  &
+\ctxwk
+  { \cftalgt{\cftalg{A}}}
+  { \cftfamalg{\cftfamalg{\cftalg{A}}}}
+  \ar{d}{\cftsubst{\cftalg{A}}}
+  \\
+  &
+\ctxwk
+  { \cftalgt{\cftalg{A}}}
+  { { \cftalgf{\cftalg{A}}}
+    { \cftfamalg{\cftalg{A}}}
+    }
+\end{tikzcd}
+\end{equation*}
+of CFT-algebras in context 
+$\ctxext{{{\Gamma}{\cftalgc{\cftalg{A}}}}{\cftalgf{\cftalg{A}}}}{\cftalgt{\cftalg{A}}}$
+commutes judgmentally
+and (2) the diagram
+\begin{equation*}
+\begin{tikzcd}[column sep=large]
+\cftfamalg{\cftfamalg{\cftalg{A}}}
+  \ar{r}{\subst{\idtm{\cftalgf{\cftalg{A}}}}{\cftfamhom{\cftwk{\cftalg{A}}}}}
+  \ar{dr}[swap]{\cftidhom{}}
+  &
+\subst
+  { \idtm{\cftalgf{\cftalg{A}}}}
+  { \jcomp{}
+      { \cftwk{\cftalg{A}}}
+      { \cftfamalg{\cftfamalg{\cftfamalg{\cftalg{A}}}}}
+    }
+  \ar{d}
+    { \subst
+        { \cftidtm{\cftalg{A}}}
+        { { { \idtm{\cftalgf{\cftalg{A}}}}
+            { \cftwkc{\cftalg{A}}}
+            }
+          { \cftsubst{\cftalgf{\cftalg{A}}}}
+          }
+      }
+  \\
+  &
+    { \subst
+        { \cftidtm{\cftalg{A}}}
+        { { { \idtm{\cftalgf{\cftalg{A}}}}
+            { \cftwkc{\cftalg{A}}}
+            }
+          { \ctxwk
+              { \cftalgt{\cftfamalg{\cftalg{A}}}}
+              { { \cftalgf{\cftfamalg{\cftalg{A}}}}
+                { \cftfamalg{\cftfamalg{\cftalg{A}}}}
+                }
+            }
+          }
+      }
+\end{tikzcd}
+\end{equation*}
+of CFT-algebras in context 
+$\ctxext{{\Gamma}{\cftalgc{\cftalg{A}}}}{\cftalgf{\cftalg{A}}}$ commutes
+judgmentally.
+
+E-algebras are considered judgmentally equal if their underlying CFT-algebras
+and their internal weakening, identity terms and substitution are judgmentally
+equal.
+\end{defn}
+
+\begin{thm}
+Suppose $\cftalg{A}$ is an E-algebra in context $\Gamma$. Then
+$\cftfamalg{\cftalg{A}}$ is an E-algebra in context $\ctxext{\Gamma}{\cftalgc{\cftalg{A}}}$.
+\end{thm}
+
+\begin{thm}
+Suppose $\cftalg{Q}$ is an E-algebra in context $\ctxext{\Gamma}{B}$ and that
+$\jfam{\Gamma}{A}$. Then $\ctxwk{A}{\cftalg{Q}}$ is an E-algebra in context
+$\ctxext{{\Gamma}{A}}{\ctxwk{A}{B}}$. 
+\end{thm}
+
+\begin{thm}
+Suppose $\cftalg{Q}$ is an E-algebra in context $\ctxext{{\Gamma}{A}}{P}$ and
+that $\jterm{\Gamma}{A}{x}$. Then $\subst{x}{\cftalg{Q}}$ is an E-algebra in
+context $\ctxext{\Gamma}{\subst{x}{P}}$.
+\end{thm}
+
+\begin{thm}
+Suppose $\cftalg{Q}$ is an E-algebra in context $\ctxext{\Gamma}{B}$ and that
+$\jfam{\Gamma}{A}$. Then we have the
+judgmental equality $\cftfamalg{\ctxwk{A}{\cftalg{Q}}}\jdeq\ctxwk{A}{\cftfamalg{\cftalg{Q}}}$
+of E-algebras in context $\ctxext{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{\cftalgc{\cftalg{Q}}}}$.
+\end{thm}
+
+\begin{thm}
+Suppose $\cftalg{Q}$ is an E-algebra in context $\ctxext{{\Gamma}{A}}{P}$ and
+that $\jterm{\Gamma}{A}{x}$. Then we have the judgmental equality
+$\cftfamalg{\subst{x}{\cftalg{Q}}}\jdeq\subst{x}{\cftfamalg{\cftalg{Q}}}$
+of E-algebras in context $\ctxext{{\Gamma}{\subst{x}{P}}}{\subst{x}{\cftalgc{\cftalg{Q}}}}$.
+\end{thm}
 

### 2014-09-19 article-models-ttderived.tex
added forgotten files

@@
 
-\subsubsection{Interactions of the composition threesome}
+\subsubsection{Interactions of horizontal and vertical composition}
 Our current goal is to show that the three sorts of composition are all
@@
 
-\begin{lem}
+\begin{lem}\label{lem:edom-edom-comp}
 Consider a morphism $\jhom{{\Gamma}{A}}{{P}{Q}}{R}{f}$ and a morphism

### 2014-09-19 article-models-ttind.tex
added forgotten files

\section{Inductive constructions}

\subsection{Strong inductive morphisms}
In this subsection we investigate the notion of inductive morphisms. An 
inductive morphism $f$ from $A$ to $B$ in context $\Gamma$ is a morphism which
induces an operation which is judgmentally the inverse of composition
with $f$. We explore a notion of inductive morphism which is much stronger
than the usual notion: it pushes families over $A$ to families over $B$,
families over families over $A$ to families over families over $B$ and terms
thereof to terms of the output families over families in a manner compatible
with the empty family, extension, weakening, substitution and identity terms.
As a result, inductive morphisms will be stable under extension, weakening,
substitution and the identity term is an inductive morphism.

Inductive morphisms are useful to for inductive types which are defined using
only one (ordinary morphism as) constructor, like the unit type and dependent 
pair types.
They can't be used to define the empty type, disjoint sums,
the natural numbers or identity types.

{\color{red} Maybe
we can solve this by writing down a type theory of inductive constructions.}

\begin{defn}
Let $\jhom{\gamma}{A}{B}{f}$ be a context morphism. We say that $f$ is an
\emph{inductive morphism} if there is an operation $\tfindf{f}$ given by
\begin{align*}
& \inference
  { \jfam{{\Gamma}{A}}{P}
    }
  { \jfam{{\Gamma}{B}}{\tfind{f}{P}}
    }
  \\
& \inference
  { \jfam{{{\Gamma}{A}}{P}}{Q}
    }
  { \jfam{{{\Gamma}{B}}{\tfind{f}{P}}}{\tfind[\famsym]{f}{Q}}
    }
  \\
& \inference
  { \jterm{{{\Gamma}{A}}{P}}{Q}{g}
    }
  { \jterm{{{\Gamma}{B}}{\tfind{f}{P}}}{\tfind[\famsym]{f}{Q}}{\tfind[\tmsym]{f}{g}}
    }
\end{align*}
for which the inference rules in the following list are valid:
\begin{enumerate}
\item The operation $\tfindf{f}$ is compatible with the empty context:
\begin{align*}
& \inference
  { 
    }
  { \jfameq{{\Gamma}{B}}{\tfind{f}{\emptyf}}{\emptyf}
    }
  \\
& \inference
  { \jfam{{\Gamma}{A}}{P}
    }
  { \jfameq{{{\Gamma}{B}}{\tfind{f}{P}}}{\tfind[\famsym]{f}{\emptyf}}{\emptyf}
    }
\end{align*}
\item The action on families $\tfindf[\famsym]{f}$ of $\tfindf{f}$ is compatible
with the action on contexts:
\begin{equation*}
\inference
  { \jfam{{\Gamma}{A}}{P}
    }
  { \jfameq{{\Gamma}{B}}{\tfind[\famsym]{f}{P}}{\tfind{f}{P}}
    }
\end{equation*}
Because of this inference rule we shall henceforth omit the annotations
$\famsym$ and $\tmsym$ from the operation $\tfindf{f}$ as usual.
\item The operation $\tfindf{f}$ is compatible with extension:
\begin{align*}
& \inference
  { \jfam{{{\Gamma}{A}}{P}}{Q}
    }
  { \jfameq
      {{\Gamma}{B}}
      {\tfind{f}{\ctxext{P}{Q}}}
      {\ctxext{\tfind{f}{P}}{\tfind{f}{Q}}}
    }
  \\
& \inference
  { \jfam{{{{\Gamma}{A}}{P}}{Q}}{R}
    }
  { \jfameq
      {{{\Gamma}{B}}{\tfind{f}{P}}}
      {\tfind{f}{\ctxext{Q}{R}}}
      {\ctxext{\tfind{f}{Q}}{\tfind{f}{R}}}
    }
\end{align*}
\item The operation $\tfindf{f}$ is compatible with weakening:
\begin{align*}
& \inference
  { \jfam{{\Gamma}{A}}{P}
    \jfam{{\Gamma}{A}}{Q}
    }
  { \jfameq
      {{{\Gamma}{B}}{\tfind{f}{P}}}
      {\tfind{f}{\ctxwk{P}{Q}}}
      {\ctxwk{\tfind{f}{P}}{\tfind{f}{Q}}}
    }
  \\
& \inference
  { \jfam{{\Gamma}{A}}{P}
    \jfam{{{\Gamma}{A}}{Q}}{R}
    }
  { \jfameq
      {{{{\Gamma}{B}}{\tfind{f}{P}}}{\ctxwk{\tfind{f}{P}}{\tfind{f}{Q}}}}
      {\tfind{f}{\ctxwk{P}{R}}}
      {\ctxwk{\tfind{f}{P}}{\tfind{f}{R}}}
    }
  \\
& \inference
  { \jfam{{\Gamma}{A}}{P}
    \jterm{{{\Gamma}{A}}{Q}}{R}{h}
    }
  { \jtermeq
      {{{{\Gamma}{B}}{\tfind{f}{P}}}{\ctxwk{\tfind{f}{P}}{\tfind{f}{Q}}}}
      {\ctxwk{\tfind{f}{P}}{\tfind{f}{R}}}
      {\tfind{f}{\ctxwk{P}{h}}}
      {\ctxwk{\tfind{f}{P}}{\tfind{f}{h}}}
    }
\end{align*}
\item We will also require that the operation $\tfindf{f}$ is compatible with
weakening by $A$:
\begin{align*}
& \inference
  { \jfam{\Gamma}{X}
    }
  { \jfameq
      {{\Gamma}{B}}
      {\tfind{f}{\ctxwk{A}{X}}}
      {\ctxwk{B}{X}}
    }
  \\
& \inference
  { \jfam{{\Gamma}{X}}{Y}
    }
  { \jfameq
      {{{\Gamma}{B}}{\ctxwk{B}{X}}}
      {\tfind{f}{\ctxwk{A}{Y}}}
      {\ctxwk{B}{Y}}
    }
  \\
& \inference
  { \jterm{{\Gamma}{X}}{Y}{y}
    }
  { \jtermeq
      {{{\Gamma}{B}}{\ctxwk{B}{X}}}
      {\ctxwk{B}{Y}}
      {\tfind{f}{\ctxwk{A}{y}}}
      {\ctxwk{B}{y}}
    }
\end{align*}
These rules assert that constant families and terms are mapped to constant
families and terms.
\item The operation $\tfindf{f}$ is compatible with substitution:
\begin{align*}
& \inference
  { \jterm{{{\Gamma}{A}}{P}}{Q}{g}
    \jfam{{{{\Gamma}{A}}{P}}{Q}}{R}
    }
  { \jfameq
      {{{\Gamma}{B}}{\tfind{f}{P}}}
      {\tfind{f}{\subst{g}{R}}}
      {\subst{\tfind{f}{g}}{\tfind{f}{R}}}
    }
  \\
& \inference
  { \jterm{{{\Gamma}{A}}{P}}{Q}{g}
    \jfam{{{{{\Gamma}{A}}{P}}{Q}}{R}}{S}
    }
  { \jfameq
      {{{{\Gamma}{B}}{\tfind{f}{P}}}{\subst{\tfind{f}{g}}{\tfind{f}{R}}}}
      {\tfind{f}{\subst{g}{S}}}
      {\subst{\tfind{f}{g}}{\tfind{f}{S}}}
    }
  \\
& \inference
  { \jterm{{{\Gamma}{A}}{P}}{Q}{g}
    \jterm{{{{{\Gamma}{A}}{P}}{Q}}{R}}{S}{k}
    }
  { \jtermeq
      {{{{\Gamma}{B}}{\tfind{f}{P}}}{\subst{\tfind{f}{g}}{\tfind{f}{R}}}}
      {\subst{\tfind{f}{g}}{\tfind{f}{S}}}
      {\tfind{f}{\subst{g}{k}}}
      {\subst{\tfind{f}{g}}{\tfind{f}{k}}}
    }
\end{align*}
\item The operation $\tfindf{f}$ is compatible with the identity terms:
\begin{equation*}
\inference
  { \jfam{{{\Gamma}{A}}{P}}{Q}
    }
  { \jtermeq
      {{{{\Gamma}{B}}{\tfind{f}{P}}}{\tfind{f}{Q}}}
      {\ctxwk{\tfind{f}{Q}}{\tfind{f}{Q}}}
      {\tfind{f}{\idtm{Q}}}
      {\idtm{\tfind{f}{Q}}}
    }
\end{equation*}
\item We will also require that $\tfindf{f}$ is compatible with $f$ itself:
\begin{equation*}
\inference
  {
    }
  { \jtermeq
      {{\Gamma}{B}}
      {\ctxwk{B}{B}}
      {\tfind{f}{f}}
      {\idtm{B}}
    }
\end{equation*}
\item Finally, we require that $\tfindf{f}$ is the right inverse of composition
with $f$:
\begin{align*}
& \inference
  { \jfam{{\Gamma}{A}}{P}
    }
  { \jfameq
      {{\Gamma}{A}}
      {\jcomp{}{f}{\tfind{f}{P}}}
      {P}
    }
  \\
& \inference
  { \jfam{{{\Gamma}{A}}{P}}{Q}
    }
  { \jfameq
      {{{\Gamma}{A}}{P}}
      {\jcomp{}{f}{\tfind{f}{Q}}}
      {Q}
    }
  \\
& \inference
  { \jterm{{{\Gamma}{A}}{P}}{Q}{g}
    }
  { \jtermeq
      {{{\Gamma}{A}}{P}}
      {Q}
      {\jcomp{}{f}{\tfind{f}{g}}}
      {g}
    }
\end{align*}
\end{enumerate}
\end{defn}

\begin{rmk}
The rules expressing that $\tfindf{f}$ is a right inverse to composition with
$f$ are usually called the `computation rules' of the induction principle.

Recall that we had announced that $\tfindf{f}$ would be an actual inverse of
composition with $f$, while we only have stated explicitly that $\tfindf{f}$
is a right inverse. We get the fact that it is also a left inverse from the
other compatibility rules. For example: given $\jfam{{\Gamma}{B}}{Q}$ we get
\begin{equation*}
\tfind{f}{\jcomp{}{f}{Q}}
  \jdeq
  \tfind{f}{\subst{f}{\ctxwk{A}{Q}}}
  \jdeq
  \subst{\tfind{f}{f}}{\tfind{f}{\ctxwk{A}{Q}}}
  \jdeq
  \subst{\idtm{B}}{\ctxwk{B}{Q}}
  \jdeq
  Q.
\end{equation*}
We thus recover the usual sort of induction principle. When $Q$ is a family
over $\ctxext{\Gamma}{B}$, all we have to do to find a term of $Q$ is to find
a term $g$ of $\jcomp{}{f}{Q}$. The result of applying $\tfindf{f}$ to $g$
will be a term of $Q$.
\end{rmk}

\begin{rmk}
These stronger rules also seem to imply that not every equivalence is going
to be an inductive morphism (when we add all the type constructors to the
theory). For instance, the interval is equivalent to the unit type. If the
unit type is defined via an inductive morphism $\emptyc\to\unit$ we get that
every family over $\unit$ is definitionally a constant type because
every family over $\emptyc$ is a weakening by the empty family. If the
equivalence from $\unit$ to the interval were inductive, this would in turn
imply that every type family over the interval is constant. However, this is
not the case because we have the family which has the unit type as a fiber
above one endpoint and the interval above the other.
\end{rmk}

\begin{lem}
The identity term
$\jhom{\Gamma}{A}{A}{\idtm{A}}$ is an inductive morphism
for each family $A$ of contexts over $\Gamma$
\end{lem}

\begin{proof}
Composition with the identity morphism is an identity operation.
\end{proof}

\begin{itemize}
\item Extensions of inductive morphisms are inductive
\item Weakenings of inductive morphisms are inductive
\item Substitutions of inductive morphisms are inductive
\end{itemize}

### 2014-09-19 article-models-ttinternal.tex
added forgotten files

@@
 Instead of looking at $e_1$ as a context morphism from $\ctxext{P}{\jcomp{}{e_0}{P}}$
-to $P$, one could also look at $e_1$ as a morphism \emph{over $\cprojfstf{A}{P}$},
-as indicated in the following diagram:
+to $P$ in context $\ctxext{\Gamma}{A}$, one could also look at $e_1$ as a 
+morphism \emph{over $\cprojfstf{A}{P}$} in context $\Gamma$, as indicated in the
+following diagram:
 \begin{equation*}
@@
 Extension also satisfies the properties explained in \autoref{comp-ee}, so we
-must find the two judgmental equalities for $e_0$ and $e_1$ mimicing those. 
+must find the two judgmental equalities for $e_0$ and $e_1$ interpreting those. 
 The first of these judgmental equalities is easy to give: it says that the
-following diagram commutes:
+following diagram in context $\Gamma$ commutes judgmentally:
 \begin{equation}\label{eq:extalg-eq1}
@@
 
-The second of the judgmental equalities is harder to describe, however. We need
-to consider `higher' families, i.e.~families over families over families, and
-thus we need to look at the family $\jcomp{}{e_0}{\jcomp{}{e_0}{P}}$ and find the
-two dotted morphisms in the diagram
-\begin{equation*}
-\begin{tikzcd}
-\ctxext{P}{{\jcomp{}{e_0}{P}}{\jcomp{}{e_0}{\jcomp{}{e_0}{P}}}}
-  \ar[densely dotted]{d}
-  \ar[densely dotted]{r}
-& \ctxext{P}{\jcomp{}{e_0}{P}} \ar{d}{e_1}\\
-\ctxext{P}{\jcomp{}{e_0}{P}} \ar{r}{e_1} & P
-\end{tikzcd}
-\end{equation*}
-The first is easy to find. Note that we have the judgmental equality
-\begin{equation*}
-\ctxext{\jcomp{}{e_0}{P}}{\jcomp{}{e_0}{\jcomp{}{e_0}{P}}}
-  \jdeq
-  \jcomp{}{e_0}{\ctxext{P}{\jcomp{}{e_0}{P}}}
-\end{equation*}
-and therefore we may just take the morphism
-\begin{equation*}
-\jhom
-  {{\Gamma}{A}}
-  {{P}{{\jcomp{}{e_0}{P}}{\jcomp{}{e_0}{\jcomp{}{e_0}{P}}}}}
-  {{P}{\jcomp{}{e_0}{P}}}
-  {\jvcomp{}{\idtm{P}}{\jcomp{}{e_0}{e_1}}}.
-\end{equation*}
-For the other morphism we need to look at the family
-$\ctxext{P}{{\jcomp{}{e_0}{P}}{\jcomp{}{e_0}{\jcomp{}{e_0}{P}}}}$ differently. We do
-that in the following lemma.
+In the following lemma we find an equivalent reformulation of the condition in 
+\autoref{eq:extalg-eq1}.
 
 \begin{lem}\label{lem:extalg-twins}
-Suppose we have $A$, $P$, $e_0$ and $e_1$ satisfying \autoref{eq:extalg-eq1}. Then
-the inference rules
+Suppose we have $\jfam{\Gamma}{A}$, $\jfam{{\Gamma}{A}}{P}$, 
+$\jhom{\Gamma}{{A}{P}}{A}{e_0}$ and 
+$\jhom{{\Gamma}{A}}{{P}{\jcomp{}{P}{e_0}}}{P}{e_1}$. 
+Then the inference rules
 \begin{align*}
 & \inference
-  { \jfam{{\Gamma}{A}}{Q}
-    }
-  { \jfameq
-      {{{{\Gamma}{A}}{P}}{\jcomp{}{e_0}{P}}}
-      {\jcomp{}{e_0}{\jcomp{}{e_0}{Q}}}
-      {\jcomp{}{e_1}{\jcomp{}{e_0}{Q}}}
-    }
-  \\
+  { }
+  { \jhomeq
+      {\Gamma}
+      {\ctxext{A}{{P}{\jcomp{}{e_0}{P}}}}
+      {A}
+      {\jcomp{}{\jvcomp{}{e_0}{\idtm{\jcomp{}{e_0}{P}}}}{e_0}}
+      {\jcomp{}{e_0}{e_0}}
+    }
+\intertext{and}
 & \inference
-  { \jfam{{{\Gamma}{A}}{Q}}{R}
-    }
-  { \jfameq
-      {{{{{\Gamma}{A}}{P}}{\jcomp{}{e_0}{P}}}
-        {\jcomp{}{e_0}{\jcomp{}{e_0}{Q}}}}
-      {\jcomp{}{e_0}{\jcomp{}{e_0}{R}}}
-      {\jcomp{}{e_1}{\jcomp{}{e_0}{R}}}
-    }
-  \\
+  { }
+  { \jhomeq
+      {\Gamma}
+      {\ctxext{A}{{P}{\jcomp{}{e_0}{P}}}}
+      {A}
+      {\jcomp{}{\jvcomp{}{\idtm{A}}{e_1}}{e_0}}
+      {\jcomp{}{e_1}{e_0}}
+    }
+\end{align*}
+As a consequence, the inference rules
+\begin{align*}
 & \inference
-  { \jterm{{{\Gamma}{A}}{Q}}{R}{h}
-    }
-  { \jtermeq
-      {{{{{\Gamma}{A}}{P}}{\jcomp{}{e_0}{P}}}
-        {\jcomp{}{e_0}{\jcomp{}{e_0}{Q}}}}
-      {\jcomp{}{e_0}{\jcomp{}{e_0}{R}}}
-      {\jcomp{}{e_0}{\jcomp{}{e_0}{h}}}
-      {\jcomp{}{e_1}{\jcomp{}{e_0}{h}}}
+  { \jhomeq
+      {\Gamma}
+      {\ctxext{A}{{P}{\jcomp{}{e_0}{P}}}}
+      {A}
+      {\jcomp{}{\jvcomp{}{e_0}{\idtm{\jcomp{}{e_0}{P}}}}{e_0}}
+      {\jcomp{}{\jvcomp{}{\idtm{A}}{e_1}}{e_0}}
+    }
+  { \jhomeq
+      {\Gamma}
+      {\ctxext{A}{{P}{\jcomp{}{e_0}{P}}}}
+      {A}
+      {\jcomp{}{e_0}{e_0}}
+      {\jcomp{}{e_1}{e_0}}
+    }
+\intertext{and}
+& \inference
+  { \jhomeq
+      {\Gamma}
+      {\ctxext{A}{{P}{\jcomp{}{e_0}{P}}}}
+      {A}
+      {\jcomp{}{e_0}{e_0}}
+      {\jcomp{}{e_1}{e_0}}
+    }
+  { \jhomeq
+      {\Gamma}
+      {\ctxext{A}{{P}{\jcomp{}{e_0}{P}}}}
+      {A}
+      {\jcomp{}{\jvcomp{}{e_0}{\idtm{\jcomp{}{e_0}{P}}}}{e_0}}
+      {\jcomp{}{\jvcomp{}{\idtm{A}}{e_1}}{e_0}}
     }
@@
 \begin{proof}
-We only prove the first inference rule. Let $\jfam{{\Gamma}{A}}{Q}$ be a family.
-Then we have the judgmental equalities
-\begin{align*}
-\jcomp{}{e_0}{\jcomp{}{e_0}{Q}}
-& \jdeq
-  \jcomp{}{\idtm{\jcomp{}{e_0}{P}}}{\jcomp{}{e_0}{\jcomp{}{e_0}{Q}}}
-  \tag{by \autoref{precomp-idtm-f}}
-  \\
-& \jdeq
-  \jcomp{}{\jvcomp{}{e_0}{\idtm{\jcomp{}{e_0}{P}}}}{\jcomp{}{e_0}{Q}}
+To prove the first inference rule, note that we have the judgmental equalities
+\begin{align*}
+\jcomp{}{\jvcomp{}{e_0}{\idtm{\jcomp{}{e_0}{P}}}}{e_0}
+& \jdeq
+  \jcomp{}{\idtm{\jcomp{}{e_0}{P}}}{\jcomp{}{e_0}{e_0}}
   \tag{by \autoref{lem:composition-threesome}}
@@
 & \jdeq
-  \jcomp{}{\jcomp{}{\jvcomp{}{e_0}{\idtm{\jcomp{}{e_0}{P}}}}{e_0}}{Q}
-  \tag{by \autoref{lem:jcomp-jcomp}}
-  \\
-& \jdeq
-  \jcomp{}{\jcomp{}{\jvcomp{}{\idtm{A}}{e_1}}{e_0}}{Q}
-  \tag{by \autoref{eq:extalg-eq1}}
-  \\
-& \jdeq
-  \jcomp{}{\jvcomp{}{\idtm{A}}{e_1}}{\jcomp{}{e_0}{Q}}
-  \tag{by \autoref{lem:jcomp-jcomp}}
-  \\
-& \jdeq
-  \jcomp{}{e_1}{\jcomp{}{\idtm{A}}{\jcomp{}{e_0}{Q}}}
-  \tag{by \autoref{lem:composition-threesome}}
-  \\
-& \jdeq
-  \jcomp{}{e_1}{\jcomp{}{e_0}{Q}}.
-  \tag{by \autoref{precomp-idtm-f}}
-\end{align*}
+  \jcomp{}{e_0}{e_0}
+  \tag{by \autoref{precomp-idtm-t}}
+\end{align*}
+The second inference rule follows immediately from \autoref{lem:edom-edom-comp}.
 \end{proof}
 
-Now we see that we can use the morphism
-\begin{equation*}
-\jhom
-  {{\Gamma}{A}}
-  {{P}{{\jcomp{}{e_0}{P}}{\jcomp{}{e_1}{{}{e_0}{P}}}}}
-  {{P}{\jcomp{}{e_0}{P}}}
-  {\jvcomp{}{e_1}{\idtm{\jcomp{}{e_1}{{}{e_0}{P}}}}}.
-\end{equation*}
-Thus, the second judgmental equality we will need is that the diagram
+The second condition we will put on context and family extension essentially 
+asserts the quadruple $(P,\jcomp{}{e_0}{P},e_1,\jcomp{}{e_0}{e_1})$ also satisfies 
+\autoref{eq:extalg-eq1}. Here we consider `higher' families, i.e.~families over
+families over families, which are described by the family 
+$\jcomp{}{e_0}{P}$ over $\ctxext{{\Gamma}{A}}{P}$. The family
+extension $e_1$ becomes context extension when we consider the quadruple
+$(P,\jcomp{}{e_0}{P},e_1,\jcomp{}{e_0}{e_1})$ and likewise, the morphism
+$\jcomp{}{e_0}{e_1}$ becomes family extension. Thus, the second requirement is 
+that the diagram
 \begin{equation}\label{eq:extalg-eq2}
 \begin{tikzcd}[column sep=huge]
-\ctxext{P}{{\jcomp{}{e_0}{P}}{\jcomp{}{e_0}{{}{e_0}{P}}}} 
-  \ar{r}{\jvcomp{}{\idtm{P}}{\jcomp{}{e_0}{e_1}}}
-  \ar{d}[swap]{
-    \jvcomp{}{e_1}{\idtm{\jcomp{}{e_1}{{}{e_0}{P}}}}
+\ctxext{P}{{\jcomp{}{e_0}{P}}{\jcomp{}{e_1}{\jcomp{}{e_0}{P}}}} 
+  \ar{d}[swap]{\jvcomp{}{e_1}{\idtm{\jcomp{}{e_1}{\jcomp{}{e_0}{P}}}}
+    } 
+  \ar{r}{\jvcomp{}{\idtm{P}}{\jcomp{}{e_0}{e_1}}
     } 
@@
 \end{equation}
-commutes judgmentally. Now we can confidently formulate the definition of
-extension algebras.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{CFT-algebras}
-The notion of CFT-algebras that we will
+\subsection{Empty family algebras}
+The notion of empty family algebras that we will
 study in this subsection will be extension algebras which also have an empty 
@@
 \begin{defn}
-A \emph{CFT-algebra $\cftalg{A}$ in context $\Gamma$} is a septuple
+An \emph{empty family algebra $\cftalg{A}$ in context $\Gamma$} is a septuple
 \begin{equation*}
@@
 We can extend the results of 
-\autoref{thm:extalg-fam,thm:extalg-wk,thm:extalg-subst} to CFT-algebras.
+\autoref{thm:extalg-fam,thm:extalg-wk,thm:extalg-subst} to empty family algebras.
 
 \begin{thm}\label{thm:cftalg-fam}
-Let $\cftalg{A}$ be a CFT-algebra
+Let $\cftalg{A}$ be an empty family algebra
 in context $\Gamma$. Then the septuple
@@
 \end{equation*}
-is a CFT-algebra in context $\ctxext{\Gamma}{\cftalgc{\cftalg{A}}}$. 
+is an empty family algebra in context $\ctxext{\Gamma}{\cftalgc{\cftalg{A}}}$. 
 \end{thm}
@@
 \begin{thm}\label{thm:cftalg-wk}
-Let $\cftalg{Q}$ be a CFT-algebra in context $\ctxext{\Gamma}{B}$ and let
+Let $\cftalg{Q}$ be an empty family algebra in context $\ctxext{\Gamma}{B}$ and let
 $\jfam{\Gamma}{A}$ be a family of contexts. Then the septuple
@@
 \end{equation*}
-is a CFT-algebra in context $\ctxext{{\Gamma}{A}}{\ctxwk{A}{B}}$.
+is an empty family algebra in context $\ctxext{{\Gamma}{A}}{\ctxwk{A}{B}}$.
 \end{thm}
@@
 \begin{thm}\label{thm:cftalg-subst}
-Let $\extalg{Q}$ be a CFT-algebra in context $\ctxext{{\Gamma}{A}}{P}$
+Let $\extalg{Q}$ be an empty family algebra in context $\ctxext{{\Gamma}{A}}{P}$
 and let $\jterm{\Gamma}{A}{x}$ be a term. Then the septuple
@@
 \end{equation*}
-is a CFT-algebra in context $\ctxext{\Gamma}{\subst{x}{P}}$.
+is an empty family algebra in context $\ctxext{\Gamma}{\subst{x}{P}}$.
 \end{thm}
@@
 \begin{cor}
-Let $\cftalg{Q}$ be a CFT-algebra in context $\ctxext{\Gamma}{B}$
+Let $\cftalg{Q}$ be an empty family algebra in context $\ctxext{\Gamma}{B}$
 and let $\jhom{\Gamma}{A}{B}{f}$. Then the septuple
@@
 \end{equation*}
-is a CFT-algebra in context $\ctxext{\Gamma}{A}$.
+is an empty family algebra in context $\ctxext{\Gamma}{A}$.
 \end{cor}
@@
 \begin{cor}
-Let $\cftalg{A}$ be a CFT-algebra in context $\Gamma$. Then
+Let $\cftalg{A}$ be an empty family algebra in context $\Gamma$. Then
 $\cftfamalg{\cftfamalg{\cftalg{A}}}$ and 
 $\jcomp{}{\cftctxext[\cftalg{A}]}{\cftfamalg{\cftalg{A}}}$ are
-judgmentally equal CFT-algebras in context 
+judgmentally equal empty family algebras in context 
 $\ctxext{{\Gamma}{\cftalgc{\cftalg{A}}}}{\cftalgf{\cftalg{A}}}$.
@@
 \begin{lem}
-Let $\cftalg{Q}$ be a CFT-algebra in context $\ctxext{\Gamma}{B}$ and let
+Let $\cftalg{Q}$ be an empty family algebra in context $\ctxext{\Gamma}{B}$ and let
 $\jfam{\Gamma}{A}$. Then we have the judgmental equality
@@
 \end{equation*}
-of CFT-algebras in context $\ctxext{{\Gamma}{A}}{\ctxwk{A}{B}}$.
+of empty family algebras in context $\ctxext{{\Gamma}{A}}{\ctxwk{A}{B}}$.
 \end{lem}
@@
 \begin{lem}
-Let $\cftalg{Q}$ be a CFT-algebra in context $\ctxext{{\Gamma}{A}}{P}$ 
+Let $\cftalg{Q}$ be an empty family algebra in context $\ctxext{{\Gamma}{A}}{P}$ 
 and let $\jterm{\Gamma}{A}{x}$. Then we have the judgmental equality
@@
 \end{equation*}
-of CFT-algebras in context $\ctxext{\Gamma}{\subst{x}{P}}$.
+of empty family algebras in context $\ctxext{\Gamma}{\subst{x}{P}}$.
 \end{lem}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{CFT-homomorphisms}
+\subsection{Empty family homomorphisms}
 \begin{defn}
-A CFT-homomorphism $\cfthom{f}$ from $\cftalg{A}$ to $\cftalg{B}$ in context
+An empty family homomorphism $\cfthom{f}$ from $\cftalg{A}$ to $\cftalg{B}$ in context
 $\Gamma$ is an extension homomorphism from $\unfold{\extalg{A}}$ to
@@
 
-Judgmental equality of CFT-homomorphisms from $\cftalg{A}$ to $\cftalg{B}$ in
+Judgmental equality of empty family homomorphisms from $\cftalg{A}$ to $\cftalg{B}$ in
 context $\Gamma$ is taken to be judgmental equality of extension homomorphisms
@@
 \begin{lem}
-For any CFT-algebra $\cftalg{A}$, the extension homomorphism $\cftidhom{\cftalg{A}}$
-is a CFT-homomorphism.
+For any empty family algebra $\cftalg{A}$, the extension homomorphism $\cftidhom{\cftalg{A}}$
+is an empty family homomorphism.
 \end{lem}
@@
 \begin{lem}
-Suppose that $\cfthom{f}$ is a CFT-homomorphism from $\cftalg{A}$ to $\cftalg{B}$
-in context $\Gamma$ and that $\cfthom{g}$ is a CFT-homomorphism from $\cftalg{B}$
+Suppose that $\cfthom{f}$ is an empty family homomorphism from $\cftalg{A}$ to $\cftalg{B}$
+in context $\Gamma$ and that $\cfthom{g}$ is an empty family homomorphism from $\cftalg{B}$
 to $\cftalg{C}$ in context $\Gamma$. Then $\jcomp{}{\cfthom{f}}{\cfthom{g}}$
-is a CFT-homomorphism from $\cftalg{A}$ to $\cftalg{C}$ in context $\Gamma$.
+is an empty family homomorphism from $\cftalg{A}$ to $\cftalg{C}$ in context $\Gamma$.
 \end{lem}
@@
 \begin{lem}
-Let $\cfthom{f}$ be a CFT-homomorphism from $\cftalg{A}$ to $\cftalg{B}$ in
-context $\Gamma$. Then $\cftfamhom{\exthom{f}}$ is a CFT-homomorphism from
+Let $\cfthom{f}$ be an empty family homomorphism from $\cftalg{A}$ to $\cftalg{B}$ in
+context $\Gamma$. Then $\cftfamhom{\exthom{f}}$ is an empty family homomorphism from
 $\cftfamalg{\cftalg{A}}$ to $\jcomp{}{\cfthomc{\cfthom{f}}}{\cftfamalg{\cftalg{B}}}$
@@
 \begin{lem}
-Let $\cfthom{g}$ be a CFT-homomorphism from $\cftalg{Q}$ to $\cftalg{R}$ in
+Let $\cfthom{g}$ be an empty family homomorphism from $\cftalg{Q}$ to $\cftalg{R}$ in
 context $\ctxext{\Gamma}{B}$ and let $\jfam{\Gamma}{A}$. Then 
 $\ctxwk{A}{\cfthom{g}}$
-is a CFT-homomorphism from $\ctxwk{A}{\cftalg{Q}}$ to $\ctxwk{A}{\cftalg{R}}$
+is an empty family homomorphism from $\ctxwk{A}{\cftalg{Q}}$ to $\ctxwk{A}{\cftalg{R}}$
 in context $\ctxext{{\Gamma}{A}}{\ctxwk{A}{B}}$.
@@
 \begin{lem}
-Let $\cfthom{g}$ be a CFT-homomorphism from $\cftalg{Q}$ to $\cftalg{R}$ in
+Let $\cfthom{g}$ be an empty family homomorphism from $\cftalg{Q}$ to $\cftalg{R}$ in
 context $\ctxext{{\Gamma}{A}}{P}$ and let $\jterm{\Gamma}{A}{x}$. Then 
 $\subst{x}{\cfthom{g}}$
-is a CFT-homomorphism from $\subst{x}{\cftalg{Q}}$ to $\subst{x}{\cftalg{R}}$
+is an empty family homomorphism from $\subst{x}{\cftalg{Q}}$ to $\subst{x}{\cftalg{R}}$
 in context $\ctxext{\Gamma}{\subst{x}{P}}$.
@@
 \subsection{Weakening algebras}
-Weakening algebras will be CFT-algebras with certain added structure.
-Although strictly speaking one could formulate a notion of weakening algebra
-which only depends on extension algebras and which omits both empty objects
-and terms, we shall not do so here.
+Weakening algebras will be extension algebras with certain added structure.
+We shall first formulate the notion of a pre-weakening algrebra, which is an
+extension algebra together with an extension homomorphism describing weakening,
+but without the additional judgmental equalities which will have to hold for
+weakening algebras. The advantage of this approach is that this allows us to
+consider pre-weakening homomorphisms and require that the weakening of the
+pre-weakening algebra is a pre-weakening homomorphism. This is the condition
+of \autoref{comp-ww} that weakening is compatible with itself. We could also
+formulate this condition directly, but it is complicated to state and it would
+not be very insightful if we did that.
 
 \begin{defn}
-A pre-weakening algebra $\cftalg{A}$ in context $\Gamma$ is a CFT-algebra
-$\cftalg{A}$ in context $\Gamma$ together with a CFT-homomorphism
+A pre-weakening algebra $\cftalg{A}$ in context $\Gamma$ is an extension algebra
+$\cftalg{A}$ in context $\Gamma$ together with an extension homomorphism
 \begin{equation}\label{cftwk}
@@
 of pre-weakening algebras.
-The CFT-homomorphism $\cftwk{\cftalg{A}}$ of \autoref{cftwk} is a triple
+The extension homomorphism $\cftwk{\cftalg{A}}$ of \autoref{cftwk} is a triple
 $(\cftwkc{\cftalg{A}},\cftwkf{\cftalg{A}},\cftwkt{\cftalg{A}})$ consisting of
@@
 \end{equation*}
-In particular, we see that $\cftwk{\cftalg{A}}$ is a CFT-homomorphism from
+In particular, we see that $\cftwk{\cftalg{A}}$ is an extension homomorphism from
 $\ctxwk{\cftalgf{\cftalg{A}}}{\cftfamalg{\cftalg{A}}}$ to $\cftfamalg{\cftalg{A}}$ over
@@
 \begin{rmk}
-Let $\cftalg{A}$ be a CFT-algebra and let $\cftwk{\cftalg{A}}$ be as in
-\autoref{cftwk}. Then $\apply\cftwk{\cftalg{A}}{\gamma}{a}$ is a CFT-homomorphism
+Let $\cftalg{A}$ be an extension algebra and let $\cftwk{\cftalg{A}}$ be as in
+\autoref{cftwk}. Then $\apply\cftwk{\cftalg{A}}{\gamma}{a}$ is an extension homomorphism
 from $\subst{\gamma}{\cftfamalg{\cftalg{A}}}$ to 
@@
 
+\begin{lem}
+Suppose $\cftalg{A}$ is a pre-weakening algebra in context $\Gamma$. Then
+$\jcomp{}{\cftctxext}{\cftfamalg{\cftalg{A}}}$ and $\cftfamalg{\cftfamalg{\cftalg{A}}}$
+are judgmentally equal pre-weakening algebras.
+\end{lem}
+
+\begin{proof}
+This follows immediately from the facts that
+$\jcomp{}{\cftctxext}{\cftfamalg{\cftalg{A}}}$ and $\cftfamalg{\cftfamalg{\cftalg{A}}}$
+are judgmentally equal extension algebras in context $\Gamma$ and that
+\begin{equation*}
+\jcomp{}{\cftctxext}{\cftwk{\cftfamalg{\cftalg{A}}}}
+  \jdeq
+\jcomp{}{\cftctxext}{\jcomp{}{\cftctxext}{\cftwk{\cftalg{A}}}}
+  \jdeq
+\jcomp{}{\cftfamext}{\jcomp{}{\cftctxext}{\cftwk{\cftalg{A}}}}
+  \jdeq
+\cftwk{\cftfamalg{\cftfamalg{\cftalg{A}}}}.
+  \qedhere
+\end{equation*}
+\end{proof}
+
+\begin{lem}
+Suppose $\cftalg{Q}$ is a pre-weakening algebra in context $\ctxext{\Gamma}{B}$
+and that $\jfam{\Gamma}{A}$. Then
+\end{lem}
+
 \subsubsection{Pre-weakening homomorphisms}
 It is useful to start with the definition of pre-weakening homomorphisms. Those
-are CFT-homomorphisms which also preserve weakening. Before we give the
+are extension homomorphisms which also preserve weakening. Before we give the
 definition, recall from \autoref{lem:jvcomp-cprojfstf} that we have a
@@
 \end{equation*}
-of morphisms in context $\Gamma$ for every CFT-homomorphism $\cfthom{f}$ from
+of morphisms in context $\Gamma$ for every extension homomorphism $\cfthom{f}$ from
 $\cftalg{A}$ to $\cftalg{B}$ in context $\Gamma$. Because weakening is the
@@
 \end{equation*}
-is valid for every CFT-homomorphism $\cfthom{f}$ from $\cftalg{A}$ to $\cftalg{B}$
+is valid for every extension homomorphism $\cfthom{f}$ from $\cftalg{A}$ to $\cftalg{B}$
 in context $\Gamma$. This judgmental equality helps us seeing that the condition
@@
 A \emph{pre-weakening homomorphism from $\cftalg{A}$ to $\cftalg{B}$ in context
-$\Gamma$} is a CFT-homomorphism $\cfthom{f}$ from $\cftalg{A}$ to $\cftalg{B}$
+$\Gamma$} is an extension homomorphism $\cfthom{f}$ from $\cftalg{A}$ to $\cftalg{B}$
 in context $\Gamma$ for which the diagram
@@
 \end{equation*}
-of CFT-algebras in context 
+of extension algebras in context 
 $\ctxext{{\Gamma}{\cftalgc{\cftalg{A}}}}{\cftalgf{\cftalg{A}}}$ commutes
@@
 homomorphism. Let $\cftalg{A}$ be a pre-weakening algebra. Then its weakening
-$\cftwk{\cftalg{A}}$ is a CFT-homomorphism from
+$\cftwk{\cftalg{A}}$ is an extension homomorphism from
 $\ctxwk{\cftalgf{\cftalg{A}}}{\cftfamalg{\cftalg{A}}}$ to
@@
 \subsubsection{Currying for weakening}
-Let $\cftalg{A}$ be a CFT-algebra in context $\Gamma$ and let
+Let $\cftalg{A}$ be an extension algebra in context $\Gamma$ and let
 \begin{equation*}
@@
 \end{equation*}
-be a CFT-homomorphism. The abstraction of the rules in \autoref{comp-ew} is the
+be an extension homomorphism. The abstraction of the rules in \autoref{comp-ew} is the
 condition that the diagram
@@
 \end{equation}
-of CFT-algebras in context 
+of extension algebras in context 
 $\ctxext
@@
 \begin{lem}
-Let $\cftalg{A}$ be a CFT-algebra in context gamma with a CFT-homomorphism 
+Let $\cftalg{A}$ be an extension algebra in context gamma with an extension homomorphism 
 $ \jhom
@@
 \end{equation*}
-of CFT-homomorphisms from $\subst{\gamma}{\cftfamalg{\cftalg{A}}}$ to 
+of extension homomorphisms from $\subst{\gamma}{\cftfamalg{\cftalg{A}}}$ to 
 $\subst{p}{{a}{{\gamma}{\cftalg{F_{F_{F_A}}}}}}$ in context $\Gamma$,
@@
 \subsubsection{Weakening preserves itself}
-Let $\cftalg{A}$ be a CFT-algebra in context $\Gamma$ and let
+Let $\cftalg{A}$ be an extension algebra in context $\Gamma$ and let
 \begin{equation*}
@@
 \end{equation*}
-be a CFT-homomorphism. The abstraction of the rules in \autoref{comp-ww} is the
+be an extension homomorphism. The abstraction of the rules in \autoref{comp-ww} is the
 condition that the diagram
@@
 \end{equation}
-of CFT-algebras in context 
+of extension algebras in context 
 $ \ctxext
@@
 \begin{enumerate}
-\item \ldots the domain of the CFT-homomorphism
+\item \ldots the domain of the extension homomorphism
 $\jcomp{}{\ctxwk{\cftalgf{\cftalg{A}}}{\cftfamext}}{\cftfamhom{\cftwk{\cftalg{A}}}}$
@@
 \end{align*}
-\item \ldots the domain and codomain of the CFT-homomorphism
+\item \ldots the domain and codomain of the extension homomorphism
 \begin{equation*}
@@
 \begin{lem}
-Let $\cftalg{A}$ be a CFT-algebra in context $\Gamma$ and let
+Let $\cftalg{A}$ be an extension algebra in context $\Gamma$ and let
 \begin{equation*}
@@
 \end{equation*}
-be a homomorphism of CFT-algebras for which the diagram in
+be a homomorphism of extension algebras for which the diagram in
 \autoref{diag:cftwk-cftwk} commutes judgmentally. 
@@
 \end{equation*}
-of CFT-algebras in context $\Gamma$ commutes judgmentally for
+of extension algebras in context $\Gamma$ commutes judgmentally for
 $\gamma:A$, $a,b:\subst{\gamma}{\cftalgf{\cftalg{A}}}$ and
@@
 Suppose $\cftalg{A}$ is a weakening algebra in context $\Gamma$. Then
-$\cftfamalg{\cftalg{A}}$ together with the CFT-homomorphism
+$\cftfamalg{\cftalg{A}}$ together with the extension homomorphism
 \begin{equation*}
@@
 let $\jfam{\Gamma}{A}$. Then $\ctxwk{A}{\cftalg{Q}}$ together with the
-CFT-homomorphism
+extension homomorphism
 \begin{equation*}
@@
 and let $\jterm{\Gamma}{A}{x}$. Then $\subst{x}{\cftalg{Q}}$ together with the
-CFT-homomorphism
+extension homomorphism
 \begin{equation*}
@@
 \begin{defn}
-A \emph{pre-substitution algebra $\cftalg{A}$ in context $\Gamma$} consists of a
-CFT-algebra $\cftalg{A}$ in context $\Gamma$ and a CFT-homomorphism
+A \emph{pre-substitution algebra $\cftalg{A}$ in context $\Gamma$} consists of an extension
+algebra $\cftalg{A}$ in context $\Gamma$ and an extension homomorphism
 \begin{equation*}
@@
 A \emph{pre-substitution homomorphism $\cfthom{f}$ from $\cftalg{A}$ to $\cftalg{B}$ in
-context $\Gamma$} is a CFT-homomorphism $\cfthom{f}$ for which the diagram
+context $\Gamma$} is an extension homomorphism $\cfthom{f}$ for which the diagram
 \begin{equation*}
@@
 \begin{defn}
-A CFT-algebra $\cftalg{A}$ in context $\Gamma$ with internal weakening, identity terms and
+An extension algebra $\cftalg{A}$ in context $\Gamma$ with internal empty context,
+empty family, weakening, identity terms and
 substitution
@@
 \end{equation*}
-of CFT-algebras in context 
+of extension algebras in context 
 $\ctxext{{{\Gamma}{\cftalgc{\cftalg{A}}}}{\cftalgf{\cftalg{A}}}}{\cftalgt{\cftalg{A}}}$
@@
 
-E-algebras are considered judgmentally equal if their underlying CFT-algebras
+E-algebras are considered judgmentally equal if their underlying extension algebras
 and their internal weakening, identity terms and substitution are judgmentally
@@
 An \emph{E-homomorphism from $\cftalg{A}$ to $\cftalg{B}$ in context $\Gamma$}
-is a CFT-homomorphism which is both a projection homomorphism and a
+is a extension homomorphism which is both a projection homomorphism and a
 substitution homomorphism.

### 2014-09-20 article-models-ttinternal.tex
spelling errors

@@
 asserting that $\extalg{A}$ is an \emph{extension algebra in context $\Gamma$}
-to be the conjunction of the following seven judmgents:
+to be the conjunction of the following seven judgments:
 \begin{align*}
@@
 Extension algebras are judgmentally
-equal if they are componentwise judgmentally equal.
+equal if they are component-wise judgmentally equal.
 
@@
 Weakening algebras will be extension algebras with certain added structure.
-We shall first formulate the notion of a pre-weakening algrebra, which is an
+We shall first formulate the notion of a pre-weakening algebra, which is an
 extension algebra together with an extension homomorphism describing weakening,
@@
 and identity terms all at once in a compatible way. Besides the compatibility
-properties there will be judgmental equalities analoguous to the cancellation
+properties there will be judgmental equalities analogous to the cancellation
 properties of \autoref{cancellation-ws,cancellation-i}. Pre-universes are to

### 2014-09-20 article-models-tt.tex
preamble cleanup
typos, universes

@@
 
-In our treatment, a contexts can be seen as binary
+In the current presentation, contexts can be seen as binary
 planar trees of which the leaves are (families of) contexts. 
@@
 \begin{align*}
-\jctx*{\Gamma} 
-& \jctxeq*{\Gamma}{\Gamma'}
-  \\
-\jfam*{\Gamma}{A} 
-& \jfameq*{\Gamma}{A}{B}
-  \\
-\jterm*{\Gamma}{A}{x} 
-& \jtermeq*{\Gamma}{A}{x}{y}.
+\jalign\jctx{\Gamma} 
+& \jalign\jctxeq{\Gamma}{\Gamma'}
+  \\
+\jalign\jfam{\Gamma}{A} 
+& \jalign\jfameq{\Gamma}{A}{B}
+  \\
+\jalign\jterm{\Gamma}{A}{x} 
+& \jalign\jtermeq{\Gamma}{A}{x}{y}.
 \end{align*}
@@
 action of context extension on families and terms.
-
-Nevertheless, the fact that extension \emph{does} act on families
-and terms (albeit trivially) reflects in the internal extension term in the type theory of
-models of type theory, which is developed in \autoref{ttmodels}. Also, in the
-categorical semantics, the extension operation would be the concatenation of
-lists of arrows belonging to the class of arrows representing families. Since
-families over families are interpreted by lists of those family arrows extending
-concateneted lists of family arrows, the action of concatenation of such lists
-may be taken to be trivial. The $\Sigma$-functor, which is the version of this
-operation on those lists which have length one, acts trivially on morphisms;
-another manifestation of the same phenomenon.
 
@@
 \begin{align*}
-\jtype*{\Gamma}{A} 
-& \jtypeeq*{\Gamma}{A}{B}
+\jalign\jtype{\Gamma}{A} 
+& \jalign\jtypeeq{\Gamma}{A}{B}
 \end{align*}
@@
 we do not assume that the empty family is a type, that would be like
-assuming that the multiplicative unit of a ring is prime. We envision a
+assuming that the multiplicative unit of a ring is prime.
+
+\begin{comment}
+We envision a
 categorical interpretation of the theory of contexts, families and terms where 
@@
 would be such paths of length $1$.
+\end{comment}
 

### 2014-09-21 article-models-ttdp.tex
evaluation operator(?)
editing derived.tex

@@
 
+\begin{comment}
+\subsubsection{Introducing the evaluation operator}
+The operation $\ctxev{A}{P}{\blank}$ brings things in context $\ctxext{{\Gamma}{A}}{P}$ to
+the context $\ctxext{{\Gamma}{\mprd{A}{P}}}{\ctxwk{\mprd{A}{P}}{A}}$.
+\begin{align*}
+& \inference
+  { \jfam{{{\Gamma}{A}}{P}}{Q}
+    }
+  { \jfamdefn
+      {{{\Gamma}{\mprd{A}{P}}}{\ctxwk{\mprd{A}{P}}{A}}}
+      {\ctxev{A}{P}{Q}}
+      {\subst{\evtm{A}{P}}{\ctxwk{\mprd{A}{P}}{Q}}}
+    }
+  \\
+& \inference
+  { \jfam{{{{\Gamma}{A}}{P}}{Q}}{R}
+    }
+  { \jfamdefn
+      {{{{\Gamma}{\mprd{A}{P}}}{\ctxwk{\mprd{A}{P}}{A}}}{\ctxev{A}{P}{Q}}}
+      {\ctxev[\famsym]{A}{P}{R}}
+      {\subst{\evtm{A}{P}}{\ctxwk{\mprd{A}{P}}{R}}}
+    }
+  \\
+& \inference
+  { \jterm{{{{\Gamma}{A}}{P}}{Q}}{R}{h}
+    }
+  { \jtermdefn
+      {{{{\Gamma}{\mprd{A}{P}}}{\ctxwk{\mprd{A}{P}}{A}}}{\ctxev{A}{P}{Q}}}
+      {\ctxev[\famsym]{A}{P}{R}}
+      {\ctxev[\tmsym]{A}{P}{h}}
+      {\subst{\evtm{A}{P}}{\ctxwk{\mprd{A}{P}}{h}}}
+    }
+\end{align*}
+\end{comment}
+
 \subsubsection{The compatibility rules of the evaluation terms}
@@
 \subsection{Composition of dependent functions}
+
+\subsection{\texorpdfstring{$\Pi$}{Π}E-algebras}
+
+\begin{defn}
+An E-algebra $\cftalg{A}$ is said to be a $\Pi$E-algebra if there is 
+an $E$-homomorphism
+\begin{equation*}
+\check\Pi:\cftfamalg{\cftfamalg{\cftalg{A}}}\to\ctxwk{\cftalgf{\cftalg{A}}}{\cftfamalg{\cftalg{A}}}
+\end{equation*} 
+in context $\ctxext{{\Gamma}{\cftalgc{\cftalg{A}}}}{\cftalgf{\cftalg{A}}}$, and a term
+\begin{equation*}
+\check{\mathsf{ev}}:
+\end{equation*}
+\end{defn}

### 2014-09-22 article-models-intro.tex
intro

--- 2013-11-10 article-models-intro.tex
@@
-\section{Introduction}
-This project has its origins in the beginning of 2013, when I proved a version
-of the descent theorem for homotopy colimits in type theory while I was working
-with Bas Spitters to develop notions from higher category theory in the
-univalent foundations. To arrive at a
-notion of diagram general enough to capture the higher inductive types described
-in chapter 6 of \cite{TheBook} excluding the truncations we needed type
-theoretical graphs. The graphs form a model of type theory and indeed we needed
-several of the basic type theoretical operations to give an efficient approach
-to the descent theorem. However it was not an issue to describe the graph model
-and the sense in which it was a model, the presentation of it was unsystematic
-and in that sense unsatisfactory too.
-
-From the experience of working with the graph model it became appearent that the
-answer we had of the question of what it models, i.e.~Martin L\"of type theory
-with the univalence axiom, could have been different if we were to describe a
-type theory based on the mechanism on which the model was built. The particular
-featueres of the graph model were that there was a type of graphs resembling a category
-of contexts, for each context a notion of what it means to be a family of graphs
-and a term thereof; there are the operations of context extension, weakening and
-substitution. In other words, the graph model is an instance of type theretical
-categories with families, an invention of Peter Dybjer.
-
-As we know however, the type theoretical description of a category as given in
-chapter 11 of \cite{TheBook} is restricted to categories which have a $1$-type
-as the type of objects and for every two objects a $0$-type of morphisms.
-Although categories with families are like that too, these are not very natural
-restrictions when one is concerned about internal models of type theory. It
-seems we're in the need of a higher categorical formulation of what an internal
-model, or a category with family, is. But type theory has the tendency to take
-care of it's own higher categorical structure and one could hope that type
-theory is an elementary theory of the category with families of category with
-families.
-
-\subsection{Ideas in the definition of internal models}
-An internal model of type theory is like a category with families, but we want
-to avoid having to state higher coherences. In fact, we don't even start our
-definition with a category of contexts; instead we just take a \emph{type} of contexts. 
-The morphisms will come from the terms, evaluation of a function at a given
-term will come from substitution. We recognize three basic ingredients to models:
-first there is a type of contexts; second, for every context there is a model of types in
-that context and third, for every type in a given context there is a type of its
-terms. Then there are three basic attributes: context extension, weakening and
-substitution. Context extension provides us with families over types as well as
-with an interpretation of dependent pair types. We need weakening 
-so that families can depend on the same type multiple times (the way the
-identity type of a type depends two times on that type) and to be able
-to talk about non-dependent function types,
-the morphims of our category. Substitution will give us a way
-to work with fibers of families as well as composition of functions and evaluation
-of functions at terms.
-
-Because we require a \emph{model} of types in a context, all the structure
-which we require at the bottom level will be required to exist higher up as well.
-Thus, the model of types in a given context $\Gamma$ will have a type of contexts
-itself, which can be seen as the type of types in $\Gamma$; it will have its
-own notion of types in a context, its own notion of terms, context extension,
-weakening and substitution together with all the structure require for it. For
-instance, when $A$ is a type in context $\Gamma$ in a model $\mfM$, then there
-is the model of types in context $A$, which is the model of families over $A$. 
-This model is required to be \emph{definitionally equal to} the model of types
-in the context $\ctxext{\Gamma}{A}$, the context extension of $\Gamma$ and $A$.
-In this way we protect ourselves from the need to dig an infinitely deep structure
-of models when we want to consider examples.
-
-To give the definition of a model we shall also need to consider certain morphisms
-of internal models. Those should preserve all the structure: contexts are mapped
-to contexts; for every context a morphism of models mapping the model of types
-in that context to the model of types in the image of that context; there should
-be a mapping of terms and context extension, weakening and substitution should be
-preserved. We need to consider those morphisms because we require context extension,
-weakening and substitution to be of that kind, thereby respecting each other
-in all possible ways.
-
-When we have this framework set up, we can interpret the basic type constructors
-such as $\Pi$, $\Sigma$ and $\idtypevar{}$.
-The higher categorical structure then comes from the
-result that we have an interpretation of type theory.
-
-\subsection{The elementary theory of the category with families of categories
-with families}
-
-Thus the idea was born to give a new description of Martin-L\"of type theory
-which is more faithful to the idea that type theory is an algebraic description
-of higher category theory. The features we have in mind for this description
-is that:
-\begin{enumerate}
-\item As in Martin-L\"of type theory there are the three basic judgments
-      \begin{align*}
-        & \Gamma\text{ is a context:} & \jctx*{\Gamma}\\
-        & A\text{ is a type in context }\Gamma\text{:} & \jtype*{\Gamma}{A}\\
-        & x\text{ is a term of $A$ in context }\Gamma\text{:} & \jterm*{\Gamma}{A}{x}
-      \end{align*}
-      together with the three accompanying judgments for judgmental equality.
-      Our new description of type theory should be such that we can faithfully
-      make the translation to the three assertions
-      \begin{enumerate}
-      \item $\Gamma$ is an object\\
-      \item $A$ is a fibration over $\Gamma$\\
-      \item $x$ is a section of the fibration $A$. 
-      \end{enumerate}
-\item There is a basic type theory which has just the operations
-      \begin{description}
-      \item[extension] taking the domain of the fibration $A$ over $\Gamma$.
-      \item[weakening] given an object $\Gamma$, weakening assigns to an object
-      $\Delta$ trivial fibration with $\Delta$ as its fibers.
-      \item[substitution] taking the fiber at a point. Fibers should always
-      be equivalent to homotopy fibers.
-      \end{description}
-      Thus we aim for a closer connection between the syntactic operations and
-      the operations of models.
-\item Models of this basic type theory should be weak $\omega$-categories. Thus,
-      it should not be necessary to start the definition of a model with a bunch
-      of contexts and substitution morphisms between them: these and the
-      operations on them, such as composition and associativity thereof, should
-      be a consequence of the interpretation of the basic type theory.
-\item There shouldn't be an empty context and dependent function types right
-      from the start. These correspond to having a terminal object and the
-      models of type theory with dependent function types are the locally
-      cartesian closed models.
-\item Contexts should not be lists of variable declarations and all the
-      operations should be explicitly invariant under context extension.
-\item There should be a large zoo of extensions of the basic type theory. First
-      off, we can extend the basic type theory with identity types, dependent
-      function types and inductive types. We want a general description of
-      homotopy colimits in type theory, which should become available once we
-      know what categories (models of the basic type theory) are. Other types
-      such as the truncations could be added as well.
-
-      An extension of a different kind is the addition of the univalence axiom.
-\end{enumerate}
-
-\subsection{The further goals of the project}
-The first aim of this project is thus to give this new description of type
-theory with these properties. The second aim is to define a general notion of
-an internal model and provide several examples thereof. In the definition of
-internal models we envision that all the operations of type theory become
-morphisms of type theory: they are compatible with each and every morphism
-of type theory. The compatibility of extension, weakening and substitution is
-already part of the basic type theory. The compatibility of identity types
-with weakening for instance, is a form of function extensionality. The 
-compatibility of dependent function types with identity types is the usual
-funchtion extensionality principle. The
-compatibility of a universe operation with identity types should be univalence.
-The descent theorem implies that the colimit operation is also compatible with
-all the type theoretic operations (sigma and id are checked).
-
-Among the examples of internal models we should have:
-\begin{description}
-\item[The setoid model] This is a classical one so it should be there. It might
-      be a bit different in our case. We'll want to construct a setoid model
-      of the basic type theory to interpret identity types without necessarily
-      interpreting dependent function types (but that should remain possible).
-\item[The graph model] This is the model that leads to the first version of the
-      descent theorem.
-\item[Univalent unvirses] A univalent universe should be a model.
-\item[Equifibered diagrams over a graph] should be the fibrations in some model.
-\item[The model of all models] Very likely, we will even define this simultaneously
-      with the notion of model.
-\item[The model of weak $\omega$-groupoids] Since we think of models of the
-      basic type theory as weak $\omega$-categories, it is not too hard to
-      provide a condition on those models which enables us to talk about
-      weak $\omega$-categories. In fact, we have several options for such a
-      condition. We investigate those; we also investigate how they relate to
-      Brunerie's weak $\omega$-groupoids. Presumably, they form an internal
-      model without too much fuss (because we already have an internal model
-      of internal models at this stage) and we can ask whether this models
-      the univalence axiom. In fact, the model of internal models might already
-      have modeled the univalence axiom. 
-\end{description}
-
-We also wish to extend the descent property to a more general class of homotopy
-colimits. The general recipe has already been implicit in the first descent
-property where we have used diagrams over graphs. We have the following in mind:
-\begin{enumerate}
-\item Give an internal description of a diagram. Probably this is the same thing
-      as an internal model of the basic type theory with or without identity types.
-\item Define the notion of homotopy colimit thereof.
-\item Define the notion if equifibered diagrams.
-\item Formulate and prove descent: the map defined by substitution
- from families over the colimit to the equifibered diagrams is an equivalence.
-\end{enumerate}
-
-Using the descent property we can view $\tfcolim$ as an operation of type theory
-It should preserve all the basic operations, identity types, dependent function
-types, itself, etc.
-
-Another thing we must keep in mind is that the compatibility rules in the basic
-type theory are strict. We should look to the possibility to weaken them.
-
-\subsection{Overview of the document}
-In \autoref{eg} we begin the project with exploring internal models of type theory
-in type theoretical setting of \cite{TheBook}.
+As of yet, there is no theory of internal models of type theory of sufficient
+generality. The problem of internalizing type theory has two facets. On the one
+hand one may choose to model the definitional equalities of type theory by
+identifications. While this approach would have the benefint of being completely
+internal (allowing one to use the internal logic of type theory to prove things
+about internal models), complications in this approach do arise when the 
+underlying types are not sets. This immediately leaves univalent universes beyond
+the scope of that program. On the other hand, one may want to interpret the
+judgmental equalities of type theory with new judgmental equalities. The problem
+here is that the theory of judgmental equality in many variations of dependent
+type theory is not developed systematically, leaving anyone essentially in the
+dark on this matter.

### 2014-10-11 article-models-fibsec.tex
starting standard model

\section{Categories with fibrations and sections}

\subsection{The  model for extension}

\begin{defn}
Let $\cat{C}$ be a category with a class $\cat{F}\subseteq\mathrm{mor}(\cat{C})$ of fibrations and for each fibration $f:X\to Y$ in $\cat{F}$ a set $\cat{T}$ of sections of $f$. We call a category with this structure a category with fibrations and sections. The class of fibrations must have the property that 
\begin{description}
\item{weakening} whenever $f:X\to Y$ and $g:U\to V$ are fibrations, then so are $\catid{X}\times g:X\times U\to X\times V$ and $\pi_1\circ(\catid{X}\times g):X\times U\to X$. Whenever $s:Y\to X$ is a section of $g$, then so is $\catid{X}\times s$ a section of $\catid{X}\times g$ and 
\item{substitution} 
\end{defn}

\begin{defn}
A context $\Gamma$ in a category with fibrations and sections is a finite list $f_1:X_1\to X_0$, \dots, $f_n:X_n\to X_{n-1}$ of composable fibrations. When $\Gamma$ is has positive length, we write $\Gamma_0$ for the ultimate codomain and $\Gamma_1$ for the ultimate domain, i.e.~the composition of all the morphisms in
$\Gamma$ is a morphism from $\Gamma_1$ to $\Gamma_0$. 
\end{defn}

\begin{defn}
When $\Gamma$ is a context in a category with fibrations and sections, a family $A$ over $\Gamma$ consists of a finite list $g_1:Y_1\to Y_0$, \dots, $g_m:Y_m\to Y_{m-1}$ such that $Y_0\defeq \Gamma_1$. When $A$ has positive length, we will write $A_1$ for the ultimate domain and sometimes we will write $A_0$ for the ultimate codomain.
\end{defn}

\begin{defn}
When $A$ is a family over $\Gamma$ of length $n$ in a category with fibrations and sections, a term $x$ of $A$ consists of a section $s_i:Y_{i-1}\to Y_i$ of $g_i$ for each $1\leq i\leq n$. 
\end{defn}

\begin{defn}
Suppose $A$ is a family over $\Gamma$. The context extension of $\Gamma$ with $A$ is the concatenation of $\Gamma$ with $A$. It is denoted by $\ctxext{\Gamma}{A}$ as usual.
\end{defn}

\begin{defn}
Suppose $P$ is a family over $\ctxext{\Gamma}{A}$. The family extension of $A$ with $P$ is the concatenation of $A$ with $P$. It is denoted by $\ctxext{\Gamma}{A}$. 
\end{defn}

Since list concatenation is taken to be associative, we automatically get that extension is associative.

\begin{defn}
The empty list of fibrations is taken to be the empty context. In each context $\Gamma$, the empty list of fibrations is taken to be the empty family over $\Gamma$. With this choice of empty context and empty family, we immediately see that the empty list and empty family are neutral with respect to context extension and family extension. We also immediately see that every context is a family over the empty context. 
\end{defn}

\section{The model for weakening}

\begin{defn}

\end{defn}

### 2014-10-16 Presentation/presentation.tex
initiated presentation

--- 2014-09-20 article-models-tt.tex
@@
-\section{The generalized algebraic theory of contexts, families and terms (the E-system)}\label{tt}
-In this section we give a description of dependent type theory before type
-constructors. Apart from contexts, families and terms -- which provide for the
-core of the language of dependent logic -- the basic ingredients
-of this theory will be the operations of extension on context and on families, 
-and empty context and empty families, weakening, substitution
-and the identity terms. The resulting theory can be seen as a manifestation of the 
-structure underlying dependent type theory.
-
-We will
-formulate the theory of contexts, families and terms in such a way that contexts aren't defined
-to be lists of variable declarations. The variable-free (a.k.a.~\emph{name-free}) approach 
-we take here is rather different than those appearing in 
-\cite{hofmann1995extensional,TheBook} but it has appeared in the work of Coquand
-and in \cite{Dybjer1996}.
-The main reason we don't let variable declarations in is that we don't see them 
-in the internal models either. This way we also set out to a more algebraic 
-approach of type theory and higher category theory. Thirdly, we will not have to
-be burdened with superficial comments about variables being bounded or not, or 
-fresh or free or not occuring at all.
-
-In the current presentation, contexts can be seen as binary
-planar trees of which the leaves are (families of) contexts. 
-The judgmental equality relation on contexts is an equivalence relation which 
-expresses that binary planar
-trees of contexts are judgmentally equal if their leaves are, taking only
-(the isomorphism class of) 
-the order of the leaves into account \emph{and not the actual shape of the three}.
-The intuition behind this equivalence relation is indeed that unstructured
-(i.e.~unbracketed) lists such as the lists of variable declarations which
-usually appear in type theoretical syntax, may be regarded as contexts.
-
-Besides contexts, families and terms there will also be a notion of `type in
-a context', which one may assume.
-A context is eligible to be a \emph{closed type} and over a context, a family
-may be a type. We have the intuition that `being a type'
-expresses the property of `being atomic or irreducible'. With a notion of types,
-one has the means to require a stratification by which every context is a finite
-extension of types starting at a closed type, and every family is likewise a
-finite extension of types. The idea behind this is to recover the theory of
-\cite{Garner2014} {\color{red}(and possibly B-systems?)}. We stress that
-we will not include axioms asserting such a stratification in the E-system
-and we leave discussion on the possibility of a typing judgment until the
-end of this section.
-
-With families of contexts being the principal things of
-study, we have added a bit more generality to the theory, compared to the
-theory of categories with families presented in \cite{Dybjer1996}. Here are
-a couple of reasons for this extra generality:
-\begin{enumerate}
-\item We want to be very sure that we can interpret all the rules of type theory
-in a straightforward manner in any context.
-\item We do want do end up with a categorical structure on the contexts, derived
-from the theory by requiring that a morphism from $\Gamma$ to $\Delta$ is a
-term of the weakening of $\Delta$ by $\Gamma$ in context $\Gamma$. Thus, we
-must be able to lift the context $\Delta$ to become a family of contexts over
-$\Gamma$ and it is not reasonable to expect that this results in a type.
-\item It is very easy in this setting to assert that an operation has an action
-on very high levels of families, i.e.~when $F$ is an operation taking things in
-context $\Gamma$ to things in context $\Delta$, then it takes a family $A$ in
-context $\Gamma$ to $F(A)$ in context $\Delta$; a family $P$ in context
-$\ctxext{\Gamma}{A}$ to a family $F(P)$ in context $\ctxext{\Delta}{(F(A))}$,
-and so on. The reason it becomes simple is that we don't have to explain what
-`and so on' means.
-\end{enumerate}
-
-We will formulate fairly strict rules governing the judgmental equalities,
-expressing that extension, weakening and substitution are combatible with
-each other in a judgmental manner. This does not, however, diminish the role
-of isomorphisms or of homotopies could play in the theory once identity types
-are added. Indeed, types could still have non-trivial identity relations and
-the category of types in a certain context could genuinely display higher
-categorical structure, or so we conjecture.
-
-Much of the rules we state are just compatibility rules of extension, weakening
-and substitution with each other. In a way, these rules assert that our contexts
-are just structureless lists of contexts and that likewise terms are structureless
-lists of terms. They are structureless in the sense that the order in which
-they are formed by pairing up is irrelevant. We note that this causes complications
-in the traditional way that categorical sematics of type theory is implemented,
-where contexts become objects of the category which is supposed to model type
-theory. The reason for this is that context extension will not satisfy all the
-compatibility rules we're about to state. The first step to resolving this is taking
-the types in the empty context as the objects.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Judgments and valid inference rules}\label{judgments}
-The theory we describe here is a theory of contexts, families of
-contexts and terms thereof. The families of contexts are by some authors called
-dependent contexts, but they are handled a bit differently here because they
-become the primary object of study. Dependent contexts can be types; they could
-be seen as atomic or indecomposable dependent contexts.
-
-Thus we make eight kinds of judgments: ``$\Gamma$ is a context'',
-``$A$ is a family of contexts over $\Gamma$'', ``$A$ is a type in context $\Gamma$''
-and ``$x$ is a term of the family $A$ of contexts over $\Gamma$''. The other four
-judgments are for judgmental equality.
+\documentclass[handout]{beamer}
+
+\input{preamble-articles}
+
+\title{An algebraic formulation of dependent type theory}
+\author{Egbert Rijke}
+\institute{Carnegie Mellon University}
+\date{November 7th -- 10th, 2014}
+
+\begin{document}
+
+\begin{frame}
+\titlepage
+\end{frame}
+
+\begin{frame}
+\frametitle{Desired properties of the algebraic theory}
+\begin{itemize}
+\item Finitely many sorts. The basic ingredients are contexts, families and
+terms. Thus there will be 3 sorts.
+\item The theory is algebraic and the operations on it are homomorphisms. The
+wish that operations are homomorphisms tells which judgmental equalities to
+require.
+\item Finite set of rules. The theory is invariant under slicing as a consequence, rather than by
+convention. There will be no rule schemes.
+\item The theory is capable of interpreting itself. Investigations to internal
+models of type theory originally motivated me to write type theory this way.
+\end{itemize}
+\end{frame}
+
+\begin{frame}
+\frametitle{Overview of the theories}
+\begin{equation*}
+\begin{tikzcd}[ampersand replacement=\&,column sep=tiny]
+{} \& \text{Extension} \ar{dr} \ar{d} \ar{dl}
+  \\
+\text{Weakening} \ar{d} \& \text{Substitution} \ar{dd} \& \text{Empty context} \ar{ddl}
+\\
+\text{Projection} \ar{dr}
+\\
+{} \& \text{Dependent Type Theory}
+\end{tikzcd}
+\end{equation*}
+\end{frame}
+
+\begin{frame}
+\frametitle{The basic judgments}
 \begin{align*}
@@
 \end{align*}
-
-Strictly speaking, we have three different judgmental equalities in play and one
-could request for a notational difference to signify that fact. For instance,
-we could denote the judgmental equalities of contexts, families and terms by
-$\jdeq_c$, $\jdeq_f$ and $\jdeq_t$ respectively. It will, however, always be
-clear which of the three kinds of judgmental equality is meant when we assert
-a judgmental equality and therefore we shall not bother to make this notational
-distinction.
-
-We note that what we call families over contexts
-here could also have been named dependent contexts or telescopes, see
-\cite{deBruijn1991,hofmann1995extensional}. The term family is in agreement
-with the terminology scheme of \cite{TheBook}, though the reader should be
-warned that the notion of familie means something slightly different there than
-it does here.
-
-The theory of contexts, families and terms shall be described by means of a
-declaration of valid inference rules. An inference rule is an expression of the
-form
-\begin{equation*}
-\inference{\mathcal{H}_1\quad\cdots\quad\mathcal{H}_n}{\mathcal{J}}
-\end{equation*}
-where each of $\mathcal{H}_1$, ..., $\mathcal{H}_n$ and $\mathcal{J}$ is 
-judgment. The judgments $\mathcal{H}_1$, ..., $\mathcal{H}_n$ are called the
-hypotheses and the judgment $\mathcal{J}$ is the conclusion of the inference
-rule. When a judgment is the conclusion of a valid inference rule, we also say
-that the judgment is valid or well-formed. In fact, the only way to construct
-new judgments is via the valid inference rules we describe in this section.
-
-The class of valid inference rules is closed under composition and under
-rearranging the hypotheses. That
-means that if
-\begin{equation*}
-\inference{\mathcal{I}_{1k}\quad\cdots\quad\mathcal{I}_{m_kk}}{\mathcal{H}_k}
-\end{equation*}
-and
-\begin{equation*}
-\inference{\mathcal{H}_1\quad\cdots\quad\mathcal{H}_n}{\mathcal{J}}
-\end{equation*}
-are valid inference rules for $1\leq k\leq n$, then
-\begin{equation*}
-\inference{\mathcal{I}_{11}\quad\cdots\quad\mathcal{I}_{m_11}\quad\cdots\quad\mathcal{I}_{1n}\quad\cdots\quad\mathcal{I}_{m_nn}}{\mathcal{J}}
-\end{equation*}
-is also a valid inference rule. Also, in a valid inference rule 
-duplicated hypotheses may always be contracted and hypotheses may always be 
-reordered; the result is still considered a valid inference rule.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{The basic rules for judgmental equality}
-The rules for judgmental equality establish that it is an equivalence relation
-in all three cases (contexts, families and terms). Thus, the following inference
-rules shall be required to be valid:
-\bgroup\small
+\end{frame}
+
+\begin{frame}
+\frametitle{Judgmental equality is an equivalence relation}
+\begin{footnotesize}
 \begin{align*}
@@
 \end{align*}
-\egroup
-
-The following convertibility rules are responsible for the strictness
-of judgmental equality, which sets it apart from equivalences or identifications:
 \begin{align*}
@@
 \end{align*}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{The operations of the theory of contexts, families and terms}
-In this subsection we introduce the notions of
-extension on context and on families, the empty context and the empty families, 
-weakening, 
-substitution and identity terms.
-
-Weakening and substitution are both
-operations of the theory of contexts, families and terms and therefore they will
-have three levels of actions: on contexts (at the appropriate level), families
-and terms.
-
-\begin{comment}
-Generally, 
-each operation should
-have an action on contexts, on families and on terms at the `appropriate level'
-and should be compatible with each of the operations we introduce here. In
-that will become more practical after the theory of contexts, families and terms
-is formulated, this means that when we will write down a type theory that
-interprets itself in \autoref{ttmodels}, each of the notions of
-extension, empty family, weakening, substitution and identity terms are going to be internalized
-as terms of the theory and all terms are going to be compatible with these.
-That vague idea will be our guiding light.
-
-So in principle, the introductions of each of the empty context, extension,
-weakening, substitution and identity terms should come in three parts. In 
-practice, they do not all come in three parts.
-
-Apart from the empty context, there will indeed be an empty family and a term
-of the empty family (which will be deemed unique). But extension will
-not have an explicit action `at the level of families' since extension
-is used to define higher families and the action would have to be trivial
-anyway. The empty context and extension provide the essentials of the rest of
-the theory in the sense that higher families are families over extensions,
-contexts are families over the empty context and the `action on families' of
-any operation applied to families over the empty context is the same as the
-original operation. Thus, when we have introduced the operations of weakening
-and substitution we will see that they also have an `action on the terms of 
-contexts'. We will state the relevant judgmental equality rules for the empty
-context in this section rather than in the next, since they are required to
-state that the action on families of an operation is compatible with the
-original operation.
-\end{comment}
-
-We note upfront that both context extension and family extension can also be
-seen as operations of contexts, families and terms, even though we will not
-introduce them as such. The reason is that extension is used to define what it
-means to be a family over a family over a context (namely, a family over an
-extended context). As such, it will have a trivial action on families and terms
-which will not need to be introduced. We will come back to this point in
-\autoref{extension}.
-
-\subsection{Extension}
-\label{extension}
-The operations of extension enable one to consider families over
-families and higher level families alike. We will need both a context extension
-operation and a family extension operation.
-
-The inference rules introducing context and family extension are as follows:
-\begin{align}
+\end{footnotesize}
+\end{frame}
+
+\begin{frame}
+\frametitle{Extension}
+Introduction rules for extension:
+\begin{align*}
 & \inference
@@
       }
-\end{align}
-The extension operation may also be defined to act on families over families
-and terms thereof. For instance, When $\jfam{{\Gamma}{A}}{P}$ is a family,
-we would get a family $\jfam{{\Gamma}{A}}{{\Gamma}{P}}$ and when
-$\jterm{{\Gamma}{A}}{P}{f}$ is a term, we would get a term
-$\jterm{{\Gamma}{A}}{{\Gamma}{P}}{\ctxext{\Gamma}{f}}$. Then there would be
-axioms stating the judgmental equalities $\ctxext{\Gamma}{P}\jdeq P$ and
-$\ctxext{\Gamma}{f}\jdeq f$. Thus, it is of no actual use to introduce the
-action of context extension on families and terms.
-
-In \autoref{extension-on-terms} we will also define an extension operation
-on terms, which defines a term $\jterm{\Gamma}{{A}{P}}{\tmext{x}{u}}$ for
-every $\jterm{\Gamma}{A}{x}$ and $\jterm{\Gamma}{\subst{x}{P}}{u}$. This
-operation is defined by using the substitution operation and identity terms.
-
-\subsubsection{Associativity of extension}
-\label{comp-ee}
-The inference rules asserting that extension is compatible with itself assert
-that contexts are unstructured lists of type declarations. This rule is
-unavoidable if we want that for a family $A$ in context $\Gamma$, a family over
-$A$ is the same thing as a family over $\ctxext{\Gamma}{A}$. 
-\begin{align}
+\end{align*}
+\pause
+Extension is associative:
+\begin{align*}
 & \inference
@@
     }
-  \label{comp-ee-c}
-\end{align}
-Note that we may take the point of view that extension distributes over
-extensions. If we want to extend an extended family $\ctxext{P}{Q}$ over
-$\ctxext{\Gamma}{A}$ by the family $A$, the result is the extension of (1)
-the extension of $P$ by $A$ and (2) the extension of $Q$ by $A$. The latter is
-simply $Q$ since extension acts trivially at the level of families (and terms).
-
-\subsection{The empty context and the empty families}
-\label{empty}
-We introduce an empty context and an family over $\Gamma$ for every context $\Gamma$. 
-It has been suggested by some to only include empty families and not an
-empty context because an empty context is not necessary, but we do have several
-reasons to include them. Having an empty context requires also rules asserting 
-that a context is the same thing as a family over the empty context and this
-gives the categorical structure on contexts for free once one has it for 
-families. The main ingredients that will be missing from the theory once an
-empty context is avoided are weakening by a context and identity morphisms from
-a context to itself. Including these by hand also requires to formulate all the
-compatibility rules involving weakening once more for the cases of weakening by
-a context and identity terms at contexts. We prefer to state these rules once
-and only once and including an empty context helps in this respect.
-
-We also prefer to have our scheme of compatibility rules as symmetrical as
-possible. The structure of type dependency should look exactly the same in the
-default case as in any context. In that respect, having an empty family but not
-an empty context seems a bit odd. Also, some sets of compatibility rules 
-(like the rules stating that extension is compatible with the empty families)
-will become assymetrical as a result of not including an empty context. Moreover,
-we would eventually like to include a stratification of the theory by means of
-a type judgment (asserting that a family in a context $\Gamma$ is a type) and
-study closed types (i.e.~types in the empty context). It would be possible to
-provide a notion of closed types without having an empty context, but this would
-have to be formulated separately and we would have to restate all the rules for
-types (if any) for closed types all over again.
-
-One of the main uses of the empty context and the empty families is that we
-get the property that the `action on contexts' of an operation is compatible
-with its `action on families'.
-
-The empty family over a context $\Gamma$ is introduced by the following rule
-inference rule:
-\begin{align}
+\end{align*}
+\end{frame}
+
+\begin{frame}
+\frametitle{The empty context and family}
+\begin{small}
+Introduction rules for the empty context and family:
+\begin{align*}
 & \inference
@@
     }
-  \\
-& \inference
+& & \inference
   { \jctx{\Gamma}
@@
     }
-  \\
-& \inference
+& & \inference
   { \jctxeq{\Gamma}{\Gamma'}
@@
     }
-\end{align}
-
-By regarding contexts as families of contexts over the empty context, we
-enable ourselves also to speak of terms of contexts. A term of a context
-$\Gamma$ is a term of the family $\Gamma$ over the empty context. These ideas
-are captured in the following convertibility rules:
-\begin{align}
+\end{align*}
+\pause
+Compatibility rules for families and contexts:
+\begin{align*}
 & \inference
@@
       }
-\end{align}
-
-The reader may wonder whether the empty family $\emptyf$ always has a
-term. This shall follow from the rules stating the compatibility of extension
-with the empty families in \autoref{comp-0e} below and from
-identity terms (\autoref{identityterms}).
-\begin{comment}
-When contexts are viewed as statements, the empty context is the statement that
-asserts nothing. Likewise, in any context one can assert nothing and this is
-achieved by the empty family. 
-Since somebody who states nothing asserts it all right, there is a term of the empty
-family over any context. Moreover, any two terms of the empty family are always
-judgmentally equal.
-\begin{align}
-& \inference
-  { }
-  { \jctx{\emptyc}
-    }
-  \\
+\end{align*}
+\pause
+Compatibility rules for extension and the empty families:
+\begin{align*}
 & \inference
@@
     }
-  { \jfam{\Gamma}{\emptyf[\Gamma]}
-    }
-  \\
-& \inference
+  { \jctxeq{\ctxext{\emptyc}{\Gamma}}{\Gamma}
+    }
+& & \inference
   { \jctx{\Gamma}
     }
-  { \jterm{\Gamma}{\emptyf[\Gamma]}{\emptytm[\Gamma]}
-    }
-  \\
-& \inference
-  { \jterm{\Gamma}{\emptyf[\Gamma]}{x}
-    }
-  { \jtermeq
-      {\Gamma}
-      {\emptyf[\Gamma]}
-      {x}
-      {\emptytm[\Gamma]}
-    }
-  \label{comp-00-t}
-\end{align}
-
-The empty context and the empty families together with the operation of
-extension will provide several shortcuts for the rest
-of the theory of families and terms. This is (in part) obtained by requiring
-that a context is exactly the same thing as a family of contexts over the
-empty context and that judgmental equality of contexts is exactly the same
-as judgmental equality of families of contexts over the empty context.
-
-\label{comp-00}
-Because every context is a family of contexts over the empty context, 
-the empty context $\emptyc$ may be considered as a family of contexts over the empty
-context. When we do this, we get $\emptyf[\emptyc]$.
-\begin{equation}\label{eq:comp-00}
-\inference
-  { }
-  { \jfameq
-      {\emptyc}
-      {\emptyc}
-      {\emptyf[\emptyc]}
-    }
-\end{equation}
-In the future, we shall denote $\emptyf[\Gamma]$ by $\emptyf$. The above rule
-guarantees that this will not cause confusion. Likewise, we shall denote
-$\emptytm[\Gamma]$ by $\emptytm$.
-\end{comment}
-
-\subsubsection{Compatibility of extension with the empty context and families}
-In the following set of inference rules we state that the empty context and
-the empty family are neutral objects for both context extension (the first two
-rules) and family extension (the last two rules).
-\label{comp-e0}\label{comp-0e}
-\begin{align}
-& \inference
-  { \jctx{\Gamma}
-    }
-  { \jctxeq{\ctxext{\emptyc}{\Gamma}}{\Gamma}
-    }
-  \label{comp-0e-c}
-  \\
-& \inference
-  { \jctx{\Gamma}
-    }
   { \jctxeq{\ctxext{\Gamma}{\emptyf}}{\Gamma}
     }
-  \label{comp-e0-c}\\
+  \\
 & \inference
@@
     }
-  \label{comp-0e-f}
-  \\
-& \inference
+& & \inference
   { \jfam{\Gamma}{A}
@@
     }
-  \label{comp-e0-f}
-\end{align}
-
-\subsubsection{Family extension restricted to the empty context}
-Strictly speaking we should have used a different notation for context extension
-as for family extension, because the following rule asserting that family extension
-in the empty context is the same thing as context extension would look tautological
-without a difference. So let us denote, only for the moment, context extension
-of $\Gamma$ by $A$ by $(\ctxext{\Gamma}{A})^c$ and family extension of $A$ by
-$P$ in context $\Gamma$ by $(\ctxext{A}{P})^\famsym$.
-
-Note that we may consider a context $\Gamma$ as a family over $\emptyc$ and
-a family $\jfam{\Gamma}{A}$ as a family $\jfam{{\emptyc}{\Gamma}}{A}$. 
-Therefore we will require the following rule:
-\begin{align}
+\end{align*}
+Compatibility of context extension and family extension:
+\begin{align*}
 & \inference
@@
     }
-\end{align}
-Note that this rule actually justifies that we have not utilized two different
-notations for context extension and family extension.
-
-\subsection{The theory of weakening}
-\label{weakening}
-When $A$ is a family in context $\Gamma$, the operation of weakening by $A$
-takes a family $B$ in context $\Gamma$ and provides a family $\ctxwk{A}{B}$
-in context $\ctxext{\Gamma}{A}$. The context family $\ctxwk{A}{B}$ can be seen
-as the constant family over $\ctxext{\Gamma}{A}$ with value $B$. This idea will
-be axiomatized in the cancellation property of weakening and substitution in
-\autoref{cancellation-ws}. In \autoref{morphisms} we will take the terms 
-$\unfold{\jhom{\Gamma}{A}{B}{f}}$ to be the morphisms of families from $A$ to 
-$B$. These will be at the heart of the categorical structure of the theory.
-
-The weakening operation acts on three levels: on contexts, on families and
-on terms. The `action on contexts' of weakening is the action we described
-above: it takes a family $B$ over $\Gamma$ to a family $\ctxwk{A}{B}$ over
-$\ctxext{\Gamma}{A}$; the `action on families' of weakening takes a family
-$Q$ over $\ctxext{\Gamma}{B}$ to a family $\ctxwk[\famsym]{A}{Q}$ over
-$\ctxext{{\Gamma}{A}}{\ctxwk{A}{B}}$; the `action on terms' of weakening takes
-a term $g$ of $Q$ to a term $\ctxwk[\tmsym]{A}{g}$ of $\ctxwk[\famsym]{A}{Q}$.
-\begin{align}
+\end{align*}
+\end{small}
+\end{frame}
+
+\begin{frame}
+\frametitle{The theory of weakening}
+\begin{footnotesize}
+The introduction rules for weakening:
+\begin{align*}
 & \inference
@@
     \\
-%& \inference
-%  { \jterm{\Gamma}{B}{y}
-%    }
-%  { \jterm{{\Gamma}{A}}{\ctxwk{A}{B}}{\ctxwk{A}{y}}
-%    }
-%& & \inference
-%    { \jtermeq{\Gamma}{B}{y}{y'}
-%      }
-%    { \jtermeq
-%        {{\Gamma}{A}}
-%        {\ctxwk{A}{B}}
-%        {\ctxwk{A}{y}}
-%        {\ctxwk{A'}{y'}}
-%      }
-%    \\
 & \inference
@@
       }
-\end{align}
-
-\subsubsection{Weakening of an empty family}
-The following inference rules express that when the empty family is
-weakened, the result is the empty family.
-\label{comp-w0}
-\begin{align}
-%& \inference
-%  { \jctx{\Gamma}
-%    }
-%  { \jfameq{\Gamma}{\ctxwk{\Gamma}{\emptyc}}{\emptyf}
-%    }
-%  \\
-& \inference
-  { \jfam{\Gamma}{A}
-    }
-  { \jfameq{{\Gamma}{A}}{\ctxwk{A}{\emptyf}}{\emptyf}
-    }
-  \label{comp-w0-c}\\
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{\Gamma}{B}
-    }
-  { \jfameq
-    {{{\Gamma}{A}}{\ctxwk{A}{B}}}
-    {\ctxwk[\famsym]{A}{\emptyf}}
-    {\emptyf}
-    }
-  \label{comp-w0-f}
-\end{align}
-\begin{comment}
-Note that we do not have to axiomatize the action of weakening
-by $A$ on the term of the empty family, because the result will be in the empty 
-family and every term in the empty family is judgmentally equal to the term 
-$\emptytm$.
-\end{comment}
-
-Because a family over $\Gamma$ is the same as a family over 
-$\ctxext{\Gamma}{\emptyf}$ we can apply both the action on contexts and the
-action on families of weakening to a family $B$ over $\Gamma$. When we apply
-the action on families, we obtain a family $\ctxwk[\famsym]{A}{B}$ over the
-context $\ctxext{{\Gamma}{A}}{\ctxwk{A}{\emptyf}}$. However, since we have
-postulated the judgmental equalities $\ctxwk{A}{\emptyf}\jdeq\emptyf$ and
-$\ctxext{{\Gamma}{A}}{\emptyf}\jdeq\ctxext{\Gamma}{A}$, we see that we can
-compare $\ctxwk[\famsym]{A}{B}$ with $\ctxwk{A}{B}$. The following inference
-rule postulates that these two are judgmentally equal:
-\begin{equation}
-\inference
-{ \jfam{\Gamma}{A}
-  \jfam{\Gamma}{B}
-  }
-{ \jfameq{{\Gamma}{A}}{\ctxwk[\famsym]{A}{B}}{\ctxwk{A}{B}}
-  }
-\end{equation}
-Due to this rule, the action on contexts and the action on families of weakening
-are compatible with each other and consequently there can be no possible
-confusion when we omit the annotations $\famsym$ and $\tmsym$ alltogether. In
-the future, the weakening of a family $Q$ over $\ctxext{\Gamma}{B}$ shall
-be denoted just by $\ctxwk{A}{Q}$ and likewise the weakening of a term $g$ of
-$Q$ shall be denoted by $\ctxwk{A}{g}$.
-
-Because a family $B$ over $\Gamma$ can be treated as a family by the operation
-of weakening, weakening also acts on the terms of $B$. The weakening of a term
-$y$ of $B$ by $A$ can be seen as the constant term (or function) of the
-family $\ctxwk{A}{B}$.
-
-\subsubsection{Weakenings of extensions}
-\label{comp-we}
-The following rules assert that when an extended family is weakened, the
-weakening distributes over the extension factors.
-\begin{align}
+\end{align*}
+\pause
+Weakening preserves extension:
+\begin{align*}
 & \inference
@@
     }
-  \label{comp-we-c}
   \\
@@
     }
-  \label{comp-we-f}
-\end{align}
-When thinking of terms of $\ctxwk{A}{B}$ as morphisms of families from $A$ to
-$B$, this looks already like form of type theoretic choice. It is weaker in that
-it is not stated with function types, yet it is stronger in that it states a
-judgmental equality between two families.
-
-There is also a version of this property where an extended term is weakened.
-This variant is stated and proved in \autoref{comp-we-t}.
-
-\subsubsection{Weakening of weakenings}
-Weakening by a family $A$ in context $\Gamma$ brings things in context $\Gamma$
-to things in context $\ctxext{\Gamma}{A}$. Since we have all the ingredients of
-the theory of contexts, families and terms in the context $\Gamma$ as well it
-has in particular it's own weakening by families. Suppose we have a family
-$\jfam{{\Gamma}{B}}{Q}$ over $B$ in context $\Gamma$. Weakening by $Q$ brings
-things from context $\ctxext{\Gamma}{B}$ to $\ctxext{{\Gamma}{B}}{Q}$. Thus
-we can provide rules asserting what will happen when we first weaken by $Q$ and
-then (via the action on families) by $A$. We will require the following
-inference rules to be valid:
-\label{comp-ww}
-\begin{align}
+\end{align*}
+\end{footnotesize}
+\end{frame}
+
+\begin{frame}
+\frametitle{The theory of weakening (continued)}
+\begin{small}
+Weakening preserves itself:
+\begin{align*}
 & \inference
@@
     }
-  \label{comp-ww-c}\\
+\\
 & \inference
@@
     }
-  \label{comp-ww-f}\\
+\\
 & \inference
@@
     }
-  \label{comp-ww-t}
-\end{align}
-
-\begin{rmk}
-As an important special case of these inference rules we have the following
-valid inference rules:
-\begin{align*}
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{\Gamma}{B}
-    \jfam{\Gamma}{C}
-    }
-  { \jfameq
-      {{{\Gamma}{A}}{\ctxwk{A}{B}}}
-      {\ctxwk{A}{{B}{C}}}
-      {\ctxwk{{A}{B}}{{A}{C}}}
-    }
-  \\
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{\Gamma}{B}
-    \jfam{{\Gamma}{C}}{R}
-    }
-  { \jfameq
-      {{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{{B}{C}}}}
-      {\ctxwk{A}{{B}{R}}}
-      {\ctxwk{{A}{B}}{{A}{R}}}
-    }
-  \\
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{\Gamma}{B}
-    \jterm{{\Gamma}{C}}{R}{h}
-    }
-  { \jtermeq
-      {{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{{B}{C}}}}
-      {\ctxwk{A}{{B}{R}}}
-      {\ctxwk{A}{{B}{h}}}
-      {\ctxwk{{A}{B}}{{A}{h}}}
-    }
-\end{align*}
-Moreover, we have
-\begin{equation*}
-\inference
-  { \jfam{\Gamma}{A}
-    \jfam{\Gamma}{B}
-    \jterm{\Gamma}{C}{z}
-    }
-  { \jtermeq
-      {{{\Gamma}{A}}{\ctxwk{A}{B}}}
-      {\ctxwk{A}{{B}{C}}}
-      {\ctxwk{A}{{B}{z}}}
-      {\ctxwk{{A}{B}}{{A}{z}}}
-    }
-\end{equation*}
-\end{rmk}
-
-\subsubsection{Weakening by the empty family}
-Note that we can also weaken by the empty family over $\Gamma$.
-Weakening by the empty family $\emptyf$ over a context $\Gamma$ leaves families, 
-their terms, families over those families and terms of those unchanged:
-\label{comp-0w}
-\begin{align}
-& \inference
-  { \jfam{\Gamma}{B}
-    }
-  { \jfameq{\Gamma}{\ctxwk{\emptyf}{B}}{B}
-    }
-  \label{comp-0w-c}
-  \\
-& \inference
-  { \jfam{{\Gamma}{B}}{Q}
-    }
-  { \jfameq{{\Gamma}{B}}{\ctxwk{\emptyf}{Q}}{Q}
-    }
-  \label{comp-0w-f}
-  \\
-& \inference
-  { \jterm{{\Gamma}{B}}{Q}{g}
-    }
-  { \jtermeq{{\Gamma}{B}}{Q}{\ctxwk{\emptyf}{g}}{g}
-    }
-  \label{comp-0w-t}
-\end{align}
-
-\subsubsection{Currying for weakening}
-\label{comp-ew}
-The rules expressing that extension is compatible with weakening assert that
-weakening by an extension is the same thing as weakening twice in the
-appropriate way.
-\begin{align}
+\end{align*}
+\pause
+Currying for weakening:
+\begin{align*}
 & \inference
@@
     }
-  \label{comp-ew-c}\\
+\\
 & \inference
@@
     }
-  \label{comp-ew-f}\\
+\\
 & \inference
@@
     } 
-  \label{comp-ew-t}
-\end{align}
-
-\begin{comment}
-\subsubsection{Weakening by a context}
-Since we have not required that there is an empty context with the property that
-a context is the same thing as a family in the empty context, we will have to
-explicitly introduce weakening by contexts. Since it will allways be clear
-whether something is weakened by a context or by a family, we shall use the
-same notation for weakening by contexts as for weakening by families. We shall
-have to repeat all the compatibility rules for weakening by families for
-weakening by contexts; there is no way around it. It is
-actually this sort of inconvenience that has originally led us to include an
-empty context and be done with it.
-
-In the following inference rules we assume that we have a context $\Gamma$ and
-that $\jctxeq{\Gamma}{\Gamma'}$ in the rules where $\Gamma'$ makes an appearence.
-\begin{align}
-& \inference
-  { \jctx{\Delta}
-    }
-  { \jfam{\Gamma}{\ctxwk{\Gamma}{\Delta}}
-    }
-& & \inference
-  { \jctxeq{\Delta}{\Delta'}
-    }
-  { \jfameq{\Gamma}{\ctxwk{\Gamma}{\Delta}}{\ctxwk{\Gamma'}{\Delta'}}
-    }
-  \\
-& \inference
-  { \jfam{\Delta}{B}
-    }
-  { \jfam{\ctxext{\Gamma}{\ctxwk{\Gamma}{\Delta}}}{\ctxwk{\Gamma}{B}}
-    }
-& & \inference
-  { \jfameq{\Delta}{B}{B'}
-    }
-  { \jfameq
-      {\ctxext{\Gamma}{\ctxwk{\Gamma}{\Delta}}}
-      {\ctxwk{\Gamma}{B}}
-      {\ctxwk{\Gamma'}{B'}}
-    }
-  \\
-& \inference
-  { \jterm{\Delta}{B}{y}
-    }
-  { \jterm
-      {\ctxext{\Gamma}{\ctxwk{\Gamma}{\Delta}}}
-      {\ctxwk{\Gamma}{B}}
-      {\ctxwk{\Gamma}{y}}
-    }
-& & \inference
-  { \jtermeq{\Delta}{B}{y}{y'}
-    }
-  { \jtermeq
-      {\ctxext{\Gamma}{\ctxwk{\Gamma}{\Delta}}}
-      {\ctxwk{\Gamma}{B}}
-      {\ctxwk{\Gamma}{y}}
-      {\ctxwk{\Gamma}{y'}}
-    }
-\end{align}
-Note that we do not employ three different notations for the three different
-actions of weakening by a context because there can be no confusion about which
-of the three actions is meant. If there were an empty context we should have
-distinguished the three as we did for weakening by a family.
-
-We shall have similar compatibility rules for weakening by a context as we did
-for weakening by a family.
-
-{\color{red}To do. It will look exactly the same as for weakening by families}
-\end{comment}
-
-\subsection{The theory of projections}
-\label{identityterms}
-We formulate the theory of projections on top of the theory of weakening
-described in \autoref{weakening}. In the theory of projections we introduce
-the identity terms. The identity term at a family $A$ in context $\Gamma$ is
-a term $\jterm{{\Gamma}{A}}{\ctxwk{A}{A}}{\idtm{A}}$, so weakening is necessary.
-Together with weakening, the identity terms will provide for the projections.
-The projection from $\cftext{A}{P}$ to $A$ will be the term
-$\jterm{{\Gamma}{{A}{P}}}{\ctxwk{\ctxext{A}{P}}{A}}{\ctxwk{P}{\idtm{A}}}$ and
-will be studied extensively in \autoref{extension-on-terms}.
-
-The following inference rules are required to be valid:
-\begin{align}
+\end{align*}
+\end{small}
+\end{frame}
+
+\begin{frame}
+\frametitle{The theory of projections}
+Introduction rules for identity morphisms:
+\begin{align*}
 & \inference
@@
       }
-\end{align}
-
-\begin{comment}
-We stress again
-that because we have not required there to be an empty context, we need to
-explicitly call for identity terms $\idtm{\Gamma}$ of $\ctxwk{\Gamma}{\Gamma}$
-in context $\Gamma$ for each context $\Gamma$. Thus, we will also require
-the inference rule
-\begin{align}
-& \inference
-  { \jctx{\Gamma}
-    }
-  { \jterm{\Gamma}{\ctxwk{\Gamma}{\Gamma}}{\idtm{\Gamma}}
-    }
-& &
-  \inference
-  { \jctxeq{\Gamma}{\Gamma'}
-    }
-  { \jtermeq{\Gamma}{\ctxwk{\Gamma}{\Gamma}}{\idtm{\Gamma}}{\idtm{\Gamma'}}
-    }
-\end{align}
-For the following inference rules, there should also be versions of identity
-terms at contexts.
-\end{comment}
-
-\subsubsection{The identity term of an extended family}
-The fact that identity terms are compatible with extensions is derived in
-\autoref{comp-ie}.
-
-\subsubsection{The identity term of a weakened family}
-\label{comp-wi}
-The identity term of a weakened family is the weakened identity term:
-\begin{equation}
+\end{align*}
+Weakening preserves the identity morphism:
+\begin{equation*}
 \inference
@@
     }
-  \label{comp-wi-t}
-\end{equation}
-
-\begin{rmk}
-An important special case is the judgmental equality
-\begin{equation*}
-\unfoldall{\jhomeq
-      {{\Gamma}{A}}
-      {\ctxwk{A}{B}}
-      {\ctxwk{A}{B}}
-      {\ctxwk{A}{\idtm{B}}}
-      {\idtm{\ctxwk{A}{B}}}
-      }
 \end{equation*}
-for a family $\jfam{\Gamma}{B}$.
-\end{rmk}
-
-\subsection{The theory of substitution}
-\label{substitution}
-The theory of substitution can be formulated separately from the theory of
-weakening. Therefore, we will only require the theory of extension and the
-empty context and families. In \autoref{sec:esystem-equalities} we will
-provide the additional judgmental equality rules that will be required when
-combining the theories of substitution and of projections.
-
-Given a family $P$ over $A$ and a term $x$ of $A$, substitution gives a way to
-consider the \emph{fiber $\subst{x}{P}$ of $P$ at $x$}. As was the case with
-weakening, the substitution operation comes in three parts: the `action on
-contexts' of substitution is the operation just described; the `action on
-families' of substitution takes a family $Q$ over $\ctxext{{\Gamma}{A}}{P}$
-to a family $\subst[\famsym]{x}{Q}$ over $\ctxext{\Gamma}{\subst{x}{P}}$; the
-`action on terms' of substitution takes a term $g$ of $Q$ over
-$\ctxext{{\Gamma}{A}}{P}$ to a term $\subst[\tmsym]{x}{g}$ of 
-$\subst[\famsym]{x}{Q}$.
-\begin{align}
+\end{frame}
+
+\begin{frame}
+\frametitle{The theory of substitution}
+\begin{footnotesize}
+Introduction rules for substitution:
+\begin{align*}
 & \inference
@@
       }
-\end{align}
-
-\subsubsection{Fibers of an empty family}
-The following inference rules establish that the fibers of the empty family are 
-the empty families:
-\label{comp-s0}
-\begin{align}
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    }
-  { \jfameq{\Gamma}{\subst{x}{\emptyf}}{\emptyf}
-    }
-  \label{comp-s0-c}
-  \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jfam{{\Gamma}{A}}{P}
-    }
-  { \jfameq
-      {{\Gamma}{\subst{x}{P}}}
-      {\subst{x}{\emptyf}}
-      {\emptyf}
-    }
-  \label{comp-s0-f}
-\end{align}
-\begin{comment}
-Note again that we do not have to axiomatize the action of substitution by $x$ on the term
-of the empty family, because the result will be in the empty family and every
-term in the empty family is judgmentally equal to the term $\emptytm$.
-\end{comment}
-
-We use the above rule to state the compatibility of the action on families of
-substitution with the action on contexts of substitution. Note that a family
-$P$ over $\ctxext{\Gamma}{A}$ may be regarded as a family over
-$\ctxext{{\Gamma}{A}}{\emptyf}$. Thus, we may consider the family
-$\subst[\famsym]{x}{P}$ over $\ctxext{\Gamma}{\subst{x}{\emptyf}}$. Since
-$\subst{x}{\emptyf}$ is judgmentally equal to the empty family, we may compare
-$\subst[\famsym]{x}{P}$ with $\subst{x}{P}$:
-\begin{equation}
-\inference
-{ \jfam{{\Gamma}{A}}{P}
-  }
-{ \jfameq{\Gamma}{\subst[\famsym]{x}{P}}{\subst{x}{P}}
-  }
-\end{equation}
-Due to this rule we need not make the annotations $\famsym$ and $\tmsym$ in
-the notation for substitution anymore and thus we shall omit them from now on.
-Note that because a family $P$ over $\ctxext{\Gamma}{A}$ is eligible for
-application of the action on families of substitution, we may also substitute
-terms of $P$. Thus, given terms $\jterm{\Gamma}{A}{x}$ and $\jterm{{\Gamma}{A}}{P}{f}$,
-we get a term $\jterm{\Gamma}{\subst{x}{P}}{\subst{x}{f}}$, the \emph{value of
-$f$ at $x$}.
-
-In \autoref{morphisms} we will use a combination of weakening and substitution
-to define composition of morphisms of families. However, we have to rely
-on the cancellation rule stated in \autoref{cancellation-ws} before we can
-meaningfully state the definition of composition.
-
-\subsubsection{Fibers of extensions}
-\label{comp-se}
-The following inference rules assert that if we take the fiber of an extended
-family at a term $x$ of $A$ in context $\Gamma$, the substitution by $x$
-distributes over the factors of the extension.
-\begin{align}
+\end{align*}
+\end{footnotesize}
+\end{frame}
+
+\begin{frame}
+\frametitle{The theory of substitution (continued)}
+Substitution is compatible with extension:
+\begin{align*}
 & \inference
@@
     }
-  \label{comp-se-c}
   \\
@@
     }
-  \label{comp-se-f}
-\end{align}
-There is also a version of this statement in which extended terms are considered.
-This variant is stated and proved in \autoref{comp-se-t}.
-
-\subsubsection{Fibers of fibers}
-\label{comp-ss}
-
-The following rules explain what happens when we first substitute by a term
-$g$ of $Q$ over $P$ in context $\ctxext{\Gamma}{A}$ and then by a term $x$ of
-$A$:
-\begin{align}
+\end{align*}
+\pause
+Substitution is compatible with itself:
+\begin{align*}
 & \inference
@@
     }
-  \label{comp-ss-c}
   \\
@@
     }
-  \label{comp-ss-f}
   \\
@@
     }
-  \label{comp-ss-t}
-\end{align}
-
-\begin{rmk}
-As an important special case of these inference rules we have the following
-valid inference rules:
-\begin{align*}
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jterm{{\Gamma}{A}}{P}{f}
-    \jfam{{{\Gamma}{A}}{P}}{Q}
-    }
-  { \jfameq
-      {\Gamma}
-      {\subst{x}{{f}{Q}}}
-      {\subst{{x}{f}}{{x}{Q}}}
-    }
-  \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jterm{{\Gamma}{A}}{P}{f}
-    \jfam{{{{\Gamma}{A}}{P}}{Q}}{R}
-    }
-  { \jfameq
-      {{\Gamma}{\subst{x}{{f}{Q}}}}
-      {\subst{x}{{f}{R}}}
-      {\subst{{x}{f}}{{x}{R}}}
-    }
-  \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jterm{{\Gamma}{A}}{P}{f}
-    \jterm{{{{\Gamma}{A}}{P}}{Q}}{R}{h}
-    }
-  { \jtermeq
-      {{\Gamma}{\subst{x}{{f}{Q}}}}
-      {\subst{x}{{f}{R}}}
-      {\subst{x}{{f}{h}}}
-      {\subst{{x}{f}}{{x}{h}}}
-    }
-\end{align*}
-Moreover, we get
+\end{align*}
+\end{frame}
+
+\begin{frame}
+\frametitle{Joining the theories of projections, substitution and the empty context}
+To join the three theories we have formulated on top of the theory of extension,
+we need to provide rules describing:
+\begin{itemize}
+\item what happens when something is weakened by the empty family;\pause
+\item that weakening preserves the empty context and family;\pause
+\item that substitution preserves the empty context and family;\pause
+\item that weakening preserves substitution;\pause
+\item that substitution preserves weakening;\pause
+\item that weakenings are constant families;\pause
+\item that substitution preserves the identity morphisms;\pause
+\item that identity morphisms act as identity functions;\pause
+\item that pullback along the identity morphism has no effect.
+\end{itemize}
+\end{frame}
+
+\begin{frame}
+\frametitle{Compatibility rules}
+Weakening by the empty family:
+\begin{align*}
+& \inference
+  { \jfam{\Gamma}{B}
+    }
+  { \jfameq{\Gamma}{\ctxwk{\emptyf}{B}}{B}
+    }
+  \\
+& \inference
+  { \jfam{{\Gamma}{B}}{Q}
+    }
+  { \jfameq{{\Gamma}{B}}{\ctxwk{\emptyf}{Q}}{Q}
+    }
+  \\
+& \inference
+  { \jterm{{\Gamma}{B}}{Q}{g}
+    }
+  { \jtermeq{{\Gamma}{B}}{Q}{\ctxwk{\emptyf}{g}}{g}
+    }
+\end{align*}
+\pause
+Weakening of the empty family:
+\begin{align*}
+& \inference
+  { \jfam{\Gamma}{A}
+    }
+  { \jfameq{{\Gamma}{A}}{\ctxwk{A}{\emptyf}}{\emptyf}
+    }
+  \\
+& \inference
+  { \jfam{\Gamma}{A}
+    \jfam{\Gamma}{B}
+    }
+  { \jfameq
+    {{{\Gamma}{A}}{\ctxwk{A}{B}}}
+    {\ctxwk[\famsym]{A}{\emptyf}}
+    {\emptyf}
+    }
+\end{align*}
+\pause
+Compatibility of weakening of a family with weakening of a context:
 \begin{equation*}
 \inference
-  { \jterm{\Gamma}{A}{x}
-    \jterm{{\Gamma}{A}}{P}{f}
-    \jterm{{{\Gamma}{A}}{P}}{Q}{g}
+{ \jfam{\Gamma}{A}
+  \jfam{\Gamma}{B}
+  }
+{ \jfameq{{\Gamma}{A}}{\ctxwk[\famsym]{A}{B}}{\ctxwk{A}{B}}
+  }
+\end{equation*}
+\end{frame}
+
+\begin{frame}
+\frametitle{Compatibility rules (continued)}
+Substitution of an empty family
+\begin{align*}
+& \inference
+  { \jterm{\Gamma}{A}{x}
+    }
+  { \jfameq{\Gamma}{\subst{x}{\emptyf}}{\emptyf}
+    }
+  \\
+& \inference
+  { \jterm{\Gamma}{A}{x}
+    \jfam{{\Gamma}{A}}{P}
+    }
+  { \jfameq
+      {{\Gamma}{\subst{x}{P}}}
+      {\subst{x}{\emptyf}}
+      {\emptyf}
+    }
+\end{align*}
+Compatibility of substitution of a family with substitution of a context:
+\begin{equation*}
+\inference
+{ \jfam{{\Gamma}{A}}{P}
+  }
+{ \jfameq{\Gamma}{\subst[\famsym]{x}{P}}{\subst{x}{P}}
+  }
+\end{equation*}
+\end{frame}
+
+\begin{frame}
+\frametitle{Compatibility rules (continued)}
+\begin{footnotesize}
+Weakening preserves substitution:
+\begin{align*}
+& \inference
+  { \jfam{\Gamma}{A}
+    \jterm{{\Gamma}{B}}{Q}{g}
+    \jfam{{{\Gamma}{B}}{Q}}{R}
+    }
+  { \jfameq
+      {{{\Gamma}{A}}{\ctxwk{A}{B}}}
+      {\ctxwk{A}{\subst{g}{R}}}
+      {\subst{\ctxwk{A}{g}}{\ctxwk{A}{R}}}
+    }
+  \\
+& \inference
+  { \jfam{\Gamma}{A}
+    \jterm{{\Gamma}{B}}{Q}{g}
+    \jfam{{{{\Gamma}{B}}{Q}}{R}}{S}
+    }
+  { \jfameq
+      {{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{\subst{g}{R}}}}
+      {\ctxwk{A}{\subst{g}{S}}}
+      {\subst{\ctxwk{A}{g}}{\ctxwk{A}{S}}}
+    }
+  \\
+& \inference
+  { \jfam{\Gamma}{A}
+    \jterm{{\Gamma}{B}}{Q}{g}
+    \jterm{{{{\Gamma}{B}}{Q}}{R}}{S}{k}
     }
   { \jtermeq
-      {\Gamma}
-      {\subst{x}{{f}{Q}}}
-      {\subst{x}{{f}{g}}}
-      {\subst{{x}{f}}{{x}{g}}}
-    }
-\end{equation*}
-\end{rmk}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Joining the theory of projections with the theory of substitution}
-\label{sec:esystem-equalities}
-
-\begin{comment}
-We have already seen that empty families, extension, weakening, substitution
-and the identity terms are compatible with each other. There are, however, a
-few more judgmental equalities explaining what happens when
-\begin{enumerate}
-\item something is weakend by the empty family. This is used to consider a term
-      $x$ of $A$ in context $\Gamma$ as a morphism from $\emptyf$ to $A$ in
-      context $\Gamma$;
-\item something is weakened by an extended family. This is Currying for weakening;
-\item something is first weakened by a family $A$ in context $\Gamma$ and then
-      substituted by a term of $A$. This is used to show that composition is
-      compatible with itself (i.e. that composition is associative);
-\item the identity terms are joining the party. Identity terms are empty contexts
-      in certain models, so they should behave accordingly.
-\end{enumerate}
-In other words, the rules seem to not fit in the scheme that all operations are
-compatible with each other, but in fact each of these rules plays a role in
-completing the scheme (work out further).
-\end{comment}
-
-\begin{comment}
-In this section we extend the judgmental equality relation in such a way that
-the empty context, extension, weakening, substitution and the identity terms
-all become compatible with each other. Since we have five operations, there
-are going to be $25$ finite sets of rules explaining how each operation is
-compatible with each of the operations.
-
-In addition to these $25$ finite sets
-of compatibility rules, there are also two sets of rules explaining the
-cancellation properties of weakening and substitution and of the identity terms.
-
-A typical compatibility axiom captures the idea that a diagram like
-\begin{equation*}
-\begin{tikzcd}[column sep=huge]
-\jfam{\Gamma}{\blank} 
-  \ar{r}{C\mapsto\ctxwk{B}{C}} 
-  \ar{d}[swap]{C\mapsto\ctxwk{A}{C}} 
-& \jfam{{\Gamma}{B}}{\blank} 
-  \ar{d}{Q\mapsto\ctxwk{A}{Q}}
-  \\
-\jfam{{\Gamma}{A}}{\blank} 
-  \ar{r}[swap]{P\mapsto\ctxwk{{A}{B}}{P}} 
-& \jfam{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\blank}
-\end{tikzcd}
-\end{equation*}
-in the case of two consecutive weakenings, or
-\begin{equation*}
-\begin{tikzcd}[column sep=huge]
-\jfam{{\Gamma}{A}}{\blank} 
-  \ar{d}[swap]{Q\mapsto\subst{x}{Q}} 
-  \ar{r}{Q\mapsto\ctxwk{P}{Q}} 
-& \jfam{{{\Gamma}{A}}{P}}{\blank} 
-    \ar{d}{R\mapsto\subst{x}{R}}
-  \\ 
-\jfam{\Gamma}{\blank} 
-  \ar{r}[swap]{B\mapsto\ctxwk{\subst{x}{P}}{B}} 
-& \jfam{{\Gamma}{\subst{x}{P}}}{\blank}
-\end{tikzcd}
-\end{equation*}
-in the case of a combination of weakening and substitution, commutes. We won't
-actually draw any of these diagrams in this section, but readers are encouraged
-to do so if they think it can help them understanding the axioms. Another thing
-one could keep in mind with the axioms that involve weakening is that all of
-them become completely obvious when weakening is made notationally invisible.
-Of course, we have an explicit notation for weakening on purpuse, but it is the
-usual practice of type theorists to make weakening invisible.
-
-The table below indicates where
-we can find the axioms of each of the combinations of operators. The reference
-at the row of operation $X$ and the column of operation $Y$ refers to the
-axiom which explains $Y$ distributes over the arguments of $X$ when $Y$ is
-applied after $X$. The empty context has no arguments, whereas extension,
-and weakening have two arguments which are both of the family-kind, substitution
-has two arguments of which one is a term and the other is a family and
-idenitity terms have one argument, a family.
-\begin{center}
-\begin{tabular}{r|rrrr}
-& ext & wk & subst & id\\
-\hline
-extension & \autoref{comp-ee}* &  & \cmark & \cmark \\
-weakening & \autoref{comp-we}* & \autoref{comp-ww} & \autoref{comp-sw} & \autoref{comp-wi} \\
-substitution & \autoref{comp-se}* & \autoref{comp-ws} & \autoref{comp-ss} & \autoref{comp-si} \\
-identity terms & \cmark & \autoref{comp-wi} & \autoref{comp-si} & \xmark
-\end{tabular}
-\end{center}
-Apart from references to sections where axioms are given, we also see checkmarks,
-crosses, marked references and duplicated references in this table. 
-The checkmarks indicate that the desired property
-can be proved and need not be axiomatized. To clarify those:
-\begin{itemize}
-\item The identity term of the empty context is the unique term of the empty
-family. Recall that we already have axiomatized in \autoref{empty} that every
-term of the empty family is judgmentally equal to $\emptytm$.
-\item The identity term of an extension can be explained by means of the notion
-of \emph{extension on terms}. This notion is defined in \autoref{extension-on-terms}
-using only identity terms and substitution. With extension on terms we can
-combine a term $\jterm{\Gamma}{A}{x}$ and a term $\jterm{\Gamma}{\subst{x}{P}}{u}$
-into a term $\jterm{\Gamma}{{A}{P}}{\tmext{x}{u}}$, which may be seen as the
-pair consisting of $x$ and $u$. Since it is a defined notion, we will
-be able to derive lots of properties for extension of terms. One of them explains
-how we can see the identity term of an extension as the extension of two certain
-terms, see \autoref{comp-ie}.
-\item Using extension on terms we can also make sense of `fibers at an extension'
-because that would exactly be the fiber at an extended term. We will see in
-\autoref{comp-es} that, for instance, $\subst{\tmext{x}{u}}{Q}\jdeq\subst{u}{{x}{Q}}$
-for any family $\jfam{{{\Gamma}{A}}{P}}{Q}$ and any two terms $\jterm{\Gamma}{A}{x}$
-and $\jterm{\Gamma}{\subst{x}{P}}{u}$. That's a form of currying.
-\end{itemize}
-The spots with a \xmark simply make no sense 
-{\color{red}(or do they, when the term model
-is better understood (by me)?)}. The marked references point to
-axioms which \emph{do} have a variant for terms, but those require the notion
-of extension on terms, can be proved and therefore need not be included in our
-list of compatibility axioms. In each of these sections, we give a forward
-reference to the lemma in which this variant is stated and proved.
-\end{comment}
-
-\subsubsection{Fibers of weakenings}\label{comp-sw}
-The following rules assert what happens when we first weaken by a family
-$Q$ over $P$ in context $\ctxext{\Gamma}{A}$ and then substitute by a term
-$x$ of $A$:
-\begin{align}
+      {{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{\subst{g}{R}}}}
+      {\ctxwk{A}{\subst{g}{S}}}
+      {\ctxwk{A}{\subst{g}{k}}}
+      {\subst{\ctxwk{A}{g}}{\ctxwk{A}{k}}}
+    }
+\end{align*}
+\pause
+Substitution preserves weakening:
+\begin{align*}
 & \inference
@@
     }
-  \label{comp-sw-c}
   \\
@@
     }
-  \label{comp-sw-f}
   \\
@@
     }
-  \label{comp-sw-t}
-\end{align}
-
-\begin{rmk}
-As an important special case of these inference rules we have the following
-valid inference rules:
-\begin{align*}
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jfam{{\Gamma}{A}}{P}
-    \jfam{{\Gamma}{A}}{Q}
-    }
-  { \jfameq
-      {{\Gamma}{\subst{x}{P}}}
-      {\subst{x}{\ctxwk{P}{Q}}}
-      {\ctxwk{\subst{x}{P}}{\subst{x}{Q}}}
-    }
-  \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jfam{{\Gamma}{A}}{P}
-    \jfam{{{\Gamma}{A}}{Q}}{R}
-    }
-  { \jfameq
-      {{{\Gamma}{\subst{x}{P}}}{\subst{x}{\ctxwk{P}{Q}}}}
-      {\subst{x}{\ctxwk{P}{R}}}
-      {\ctxwk{\subst{x}{P}}{\subst{x}{R}}}
-    }
-  \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jfam{{\Gamma}{A}}{P}
-    \jterm{{{\Gamma}{A}}{Q}}{R}{h}
-    }
-  { \jtermeq
-      {{{\Gamma}{\subst{x}{P}}}{\subst{x}{\ctxwk{P}{Q}}}}
-      {\subst{x}{\ctxwk{P}{R}}}
-      {\subst{x}{\ctxwk{P}{h}}}
-      {\ctxwk{\subst{x}{P}}{\subst{x}{h}}}
-    }
-\end{align*}
-Moreover, we get
-\begin{equation*}
-\inference
-  { \jterm{\Gamma}{A}{x}
-    \jfam{{\Gamma}{A}}{P}
-    \jterm{{\Gamma}{A}}{Q}{g}
-    }
-  { \jtermeq
-      {{\Gamma}{\subst{x}{P}}}
-      {\subst{x}{\ctxwk{P}{Q}}}
-      {\subst{x}{\ctxwk{P}{g}}}
-      {\ctxwk{\subst{x}{P}}{\subst{x}{g}}}
-    }
-\end{equation*}
-\end{rmk}
-
-\subsubsection{Weakenings of fibers}
-\label{comp-ws}
-The following inference rules explain what happens when we first weaken by a
-term $g$ of $Q$ in context $\ctxext{\Gamma}{B}$ and then weaken by a family
-$A$ over $\Gamma$.
-\begin{align}
-& \inference
-  { \jfam{\Gamma}{A}
-    \jterm{{\Gamma}{B}}{Q}{g}
-    \jfam{{{\Gamma}{B}}{Q}}{R}
-    }
-  { \jfameq
-      {{{\Gamma}{A}}{\ctxwk{A}{B}}}
-      {\ctxwk{A}{\subst{g}{R}}}
-      {\subst{\ctxwk{A}{g}}{\ctxwk{A}{R}}}
-    }
-  \label{comp-ws-c}
-  \\
-& \inference
-  { \jfam{\Gamma}{A}
-    \jterm{{\Gamma}{B}}{Q}{g}
-    \jfam{{{{\Gamma}{B}}{Q}}{R}}{S}
-    }
-  { \jfameq
-      {{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{\subst{g}{R}}}}
-      {\ctxwk{A}{\subst{g}{S}}}
-      {\subst{\ctxwk{A}{g}}{\ctxwk{A}{S}}}
-    }
-  \label{comp-ws-f}
-  \\
-& \inference
-  { \jfam{\Gamma}{A}
-    \jterm{{\Gamma}{B}}{Q}{g}
-    \jterm{{{{\Gamma}{B}}{Q}}{R}}{S}{k}
-    }
-  { \jtermeq
-      {{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{\subst{g}{R}}}}
-      {\ctxwk{A}{\subst{g}{S}}}
-      {\ctxwk{A}{\subst{g}{k}}}
-      {\subst{\ctxwk{A}{g}}{\ctxwk{A}{k}}}
-    }
-  \label{comp-ws-t}
-\end{align}
-
-\begin{rmk}
-As an important special case of these inference rules we have the following 
-valid inference rules:
-\begin{align*}
-& \inference
-  { \jfam{\Gamma}{A}
-    \jterm{\Gamma}{B}{y}
-    \jfam{{\Gamma}{B}}{Q}
-    }
-  { \jfameq
-      {{\Gamma}{A}}
-      {\ctxwk{A}{\subst{y}{Q}}}
-      {\subst{\ctxwk{A}{y}}{\ctxwk{A}{Q}}}
-    }
-  \\
-& \inference
-  { \jfam{\Gamma}{A}
-    \jterm{\Gamma}{B}{y}
-    \jfam{{{\Gamma}{B}}{Q}}{R}
-    }
-  { \jfameq
-      {{{\Gamma}{A}}{\ctxwk{A}{\subst{y}{Q}}}}
-      {\ctxwk{A}{\subst{y}{R}}}
-      {\subst{\ctxwk{A}{y}}{\ctxwk{A}{R}}}
-    }
-  \\
-& \inference
-  { \jfam{\Gamma}{A}
-    \jterm{\Gamma}{B}{y}
-    \jterm{{{\Gamma}{B}}{Q}}{R}{h}
-    }
-  { \jtermeq
-      {{{\Gamma}{A}}{\ctxwk{A}{\subst{y}{Q}}}}
-      {\ctxwk{A}{\subst{y}{R}}}
-      {\ctxwk{A}{\subst{y}{h}}}
-      {\subst{\ctxwk{A}{y}}{\ctxwk{A}{h}}}
-    }
-\end{align*}
-Moreover, we get
-\begin{equation*}
-\inference
-  { \jfam{\Gamma}{A}
-    \jterm{\Gamma}{B}{y}
-    \jterm{{\Gamma}{B}}{Q}{g}
-    }
-  { \jtermeq
-      {{\Gamma}{A}}
-      {\ctxwk{A}{\subst{y}{Q}}}
-      {\ctxwk{A}{\subst{y}{g}}}
-      {\subst{\ctxwk{A}{y}}{\ctxwk{A}{g}}}
-    }
-\end{equation*}
-\end{rmk}
-
-\subsubsection{The cancellation property of weakening and substitution}
-\label{cancellation-ws}
-The judgmental equalities we're about to describe assert that substituting a term
-in the weakening a thing gives you the thing back. In the case of contexts we get that each fiber
-$\subst{x}{\ctxwk{A}{B}}$ is just $B$ and in the case of terms we get 
-that $\ctxwk{A}{y}$ is the constant function
-mapping everything to $y:B$. Thus, these rules actually establish the weakening
-as the weakening. After stating the rules we will describe what it means to
-compose context morphisms (terms of weakened contexts).
-\begin{align}
+\end{align*}
+\end{footnotesize}
+\end{frame}
+
+\begin{frame}
+\frametitle{Compatibility rules (continued)}
+Weakened families are constant:
+\begin{align*}
 & \inference
@@
     }
-  \label{cancellation-ws-c}\\
+  \\
 & \inference
@@
     }
-  \label{cancellation-ws-f}\\
+  \\
 & \inference
@@
     }
-  \label{cancellation-ws-t}
-\end{align}
-
-\subsubsection{The identity term of a substituted family}
-\label{comp-si}
-The identity term of a substituted family is the substitution of the identity term
-\begin{equation}
+\end{align*}
+\pause
+Substitution preserves the identity morphisms
+\begin{equation*}
 \inference
@@
     }}
-  \label{comp-si-t}
-\end{equation}
-
-\begin{rmk}
-An important special case is the judgmental equality
+\end{equation*}
+\pause
+Identity morphisms act as identity functions:
 \begin{equation*}
-\unfoldall{\jhomeq
-      {\Gamma}
-      {\subst{x}{P}}
-      {\subst{x}{P}}
-      {\subst{x}{\idtm{P}}}
-      {\idtm{\subst{x}{P}}}
+\inference
+  { \jterm{\Gamma}{A}{x}
+    }
+  { \jtermeq{\Gamma}{A}{\subst{x}{\idtm{A}}}{x}
     }
 \end{equation*}
-for a family $\jfam{{\Gamma}{A}}{P}$.
-\end{rmk}
-
-\subsubsection{The cancellation properties of identity terms}
-\label{cancellation-i}
-Identity terms are determined by their behavior with respect to substitution combined with
-weakening. The identity terms will also be subject to compatibility rules.
-\begin{align}
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    }
-  { \jtermeq{\Gamma}{A}{\subst{x}{\idtm{A}}}{x}
-    }
-  \label{cancellation-si}\\
+\end{frame}
+
+\begin{frame}
+\frametitle{Compatibility rules (continued)}
+Pullback along identity morphisms has no effect:
+\begin{align*}
 & \inference
@@
     }
-  \label{precomp-idtm-c}
   \\
@@
     }
-  \label{precomp-idtm-f}
   \\
@@
     }
-  \label{precomp-idtm-t}
-\end{align}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{The possiblity of types in the theory of contexts, families and
-terms}
-\label{types}
-We have deliberately not spoken of types so far because we have taken the point
-of view that a type in a context is nothing but a family in that context which
-belongs to the class of types. We think of types as \emph{irreducible} families,
-i.e.\ families which are neither empty nor the extension of two
-families which are both not empty (in algebraic terminology: which
-are both non-trivial). To allow ourselves to speak of types we introduce two
-new judgments: the judgment that something is a type and the judgment that two
-types are equal.
-\begin{align*}
-\jalign\jtype{\Gamma}{A} 
-& \jalign\jtypeeq{\Gamma}{A}{B}
-\end{align*}
-But only families of contexts are eligible to be types. If $A$ is a type
-in context $\Gamma$, then $A$ is also a family of contexts over $\Gamma$. 
-Moreover, two types in context $\Gamma$ are judgmentally equal precisely when they are equal
-as context families and if a family $B$ of contexts over $\Gamma$ is
-judgmentally equal to a type $A$ in context $\Gamma$, then $B$ is a type in
-context $\Gamma$. This is expressed by the following four inference rules:
-\begin{align*}
-& \inference
-  { \jtype{\Gamma}{A}
-    }
-  { \jfam{\Gamma}{A}
-    }
-& & \inference
-    { \jtypeeq{\Gamma}{A}{B}
-      }
-    { \jfameq{\Gamma}{A}{B}
-      }
-    \\
-& \inference
-  { \jtype{\Gamma}{A}
-    \jfameq{\Gamma}{A}{B}
-    }
-  { \jtype{\Gamma}{B}
-    }
-& & \inference
-    { \jtype{\Gamma}{A}
-      \jfameq{\Gamma}{A}{B}
-      }
-    { \jtypeeq{\Gamma}{A}{B}
-      }
-\end{align*}
-As pointed out at the beginning of this subsection, 
-we do not assume that the empty family is a type, that would be like
-assuming that the multiplicative unit of a ring is prime.
-
-\begin{comment}
-We envision a
-categorical interpretation of the theory of contexts, families and terms where 
-contexts are interpreted as finite paths of arrows belonging to a predetermined
-class. The empty family would be such a path of length $0$ whereas the types
-would be such paths of length $1$.
-\end{comment}
-
-We add rules asserting that a weakened type is again a type that 
-substitution preserves the property of being a type:
-\begin{align}
-& \inference
-  { \jfam{\Gamma}{B}
-    \jtype{\ctxext{\Gamma}{B}}{Q}
-    }
-  { \jtype{\ctxext{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{Q}}
-    }
-  \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jtype{\ctxext{{\Gamma}{A}}{P}}{Q}
-    }
-  { \jtype{\ctxext{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}
-    }
-\end{align}
-With only the current rules, the possibility of making the judgment that
-something is a type does not add much to the theory of contexts, families and
-terms. Nevertheless, when studying models, having an interpretation for the
-judgment that something is a type will allow for the possibility to study
-conditions such as the one asserting that every family `factorizes' uniquely
-as multiple applications of extension to types, analogous to the condition on
-unique factorization domains in ring theory. All the tradional models of type
-theory should translate to models with such a condition, simply because
-traditionally contexts are viewed as lists of variable (and type) declarations.
+\end{align*}
+\end{frame}
+\end{document}

### 2014-10-28 Derivations/article-models-ttderived.tex
reorganization and slight change in the rules

--- 2014-09-19 article-models-ttderived.tex
@@
 
-In the last subsection we give a treatment of inductive morphisms. These
-morphisms appear also in the introduction of the many inductive type
-constructors in \autoref{tt_constructors} and therefore a general treatment of
-the subject is insightful.
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
       }
-  \tag{by \autoref{idfunc-ext-comp}}\\
+  \tag{by \autoref{comp-ie}}\\
 & \jdeq 
@@
 
-In \autoref{lem:tmext-emp,lem:tmext-ext,lem:tmext-wk,lem:tmext-subst,lem:tmext-id}
+In \autoref{lem:tmext-ext,lem:tmext-wk,lem:tmext-subst,comp-ie,lem:tmext-jcomp}
 we show that term extension and the projections are
 compatible with extension (\autoref{lem:tmext-tmext}),
-weakening (\autoref{lem:tmext-wk}), substitution (\autoref{lem:tmext-subst}) and the identity terms (\autoref{lem:tmext-id}). The latter property, \autoref{lem:tmext-id},
-is in fact a generalization of the above \autoref{lem:tmext-basic}.
+weakening (\autoref{lem:tmext-wk}), substitution (\autoref{lem:tmext-subst}), 
+the identity terms (\autoref{comp-ie}) and with precomposition (\autoref{lem:tmext-jcomp}). 
+The property that term extension and the projections are compatible with identity terms
+is in fact a generalization of the above \autoref{lem:tmext-basic}. In
+\autoref{lem:tmext-emp} we will also show that term extension and the projections
+are compatible with the empty object.
 
@@
   \ctxwk{Q}{{P}{\idtm{A}}}
+  \tag{by \autoref{precomp-idtm-t}}
   \\
@@
   \ctxwk{\ctxext{P}{Q}}{\idtm{A}}
+  \tag{by \autoref{comp-ew-t}}
   \\
@@
   \cprojfstf{A}{{P}{Q}}
+  \tag{by definition}
 \end{align*}
@@
   \unfoldall{\jcomp{{{A}{P}}{Q}}{\cprojfstf{{A}{P}}{Q}}{\cprojsndf{A}{P}}}
+  \tag{by definition}
   \\
@@
   \subst{\ctxwk{Q}{\idtm{{A}{P}}}}{\ctxwk{Q}{{\ctxext{A}{P}}{\idtm{P}}}}
+  \tag{by \autoref{comp-ew-t}}
   \\
@@
   \ctxwk{Q}{\subst{\idtm{{A}{P}}}{\ctxwk{\ctxext{A}{P}}{\idtm{P}}}}
+  \tag{by \autoref{comp-ws-t}}
   \\
@@
   \ctxwk{Q}{\idtm{P}}
+  \tag{by \autoref{precomp-idtm-t}}
   \\
@@
   \unfoldall{\jcomp{{P}{Q}}{\cprojsndf{A}{{P}{Q}}}{\cprojfstf{P}{Q}}}
+  \tag{by \autoref{precomp-idtm-t}}
   \\
@@
   \jcomp{{P}{Q}}{\cprojsndf{A}{{P}{Q}}}{\cprojfstf{P}{Q}}
+  \tag{by definition}
 \end{align*}
@@
   \unfoldall{\cprojsndf{{A}{P}}{Q}}
+  \tag{by definition}
   \\
@@
   \unfoldall{\jcomp{{P}{Q}}{\cprojsndf{A}{{P}{Q}}}{\cprojsndf{P}{Q}}}
-  \\
-& \jdeq
-  \jcomp{}{\cprojsndf{A}{{P}{Q}}}{\cprojsndf{P}{Q}}\qedhere
-\end{align*}
-\begin{comment}
-%%%% This was a proof of an old version of the statement
-Now consider a term $\jterm{\Gamma}{\ctxext{A}{{P}{Q}}}{w}$. Then we have
-\begin{align*}
-w 
-& \jdeq 
-  \tmext
-    {A}
-    {{P}{Q}}
-    {\cprojfst{A}{\ctxext{P}{Q}}{w}}
-    {\cprojsnd{A}{\ctxext{P}{Q}}{w}}
-  \\
-& \jdeq 
-  \tmext
-    {A}
-    {{P}{Q}}
-    {\cprojfst{A}{\ctxext{P}{Q}}{w}}
-    { {P}
-      {Q}
-      {\cprojfst{P}{Q}{\cprojsnd{A}{\ctxext{P}{Q}}{w}}}
-      {\cprojsnd{P}{Q}{\cprojsnd{A}{\ctxext{P}{Q}}{w}}}
-      }
-  \\
-& \jdeq 
-  \tmext
-    {{A}{P}}
-    {Q}
-    { {} % need to provide base and family, but there's no unfold.
-      {}
-      {\cprojfst{A}{\ctxext{P}{Q}}{w}}
-      {\cprojfst{P}{Q}{\cprojsnd{A}{\ctxext{P}{Q}}{w}}}
-      }
-    { \cprojsnd{P}{Q}{\cprojsnd{A}{\ctxext{P}{Q}}{w}}
-      }
-\end{align*}
-Thus we see that 
-\begin{align*}
-\cprojfst{\ctxext{A}{P}}{Q}{w} 
-& \jdeq 
-  \tmext
-    {A}
-    {P}
-    {\cprojfst{A}{\ctxext{P}{Q}}{w}}
-    {\cprojfst{P}{Q}{\cprojsnd{A}{\ctxext{P}{Q}}{w}}}
-  \\ 
-\cprojsnd{\ctxext{A}{P}}{Q}{w} 
-& \jdeq 
-  \cprojsnd{P}{Q}{\cprojsnd{A}{\ctxext{P}{Q}}{w}},
-\end{align*}
-proving the fourth judgmental equality, and therefore also that
-\begin{align*}
-\cprojfst{A}{P}{\cprojfst{\ctxext{A}{P}}{Q}{w}} 
-& \jdeq 
-  \cprojfst{A}{\ctxext{P}{Q}}{w}
-  \\
-\cprojsnd{A}{P}{\cprojfst{\ctxext{A}{P}}{Q}{w}} 
-& \jdeq 
-  \cprojfst{P}{Q}{\cprojsnd{A}{\ctxext{P}{Q}}{w}},
-\end{align*}
-proving the second and the third judgmental equalities.
-\end{comment}
+  \tag{by \autoref{precomp-idtm-t}}
+  \\
+& \jdeq
+  \jcomp{}{\cprojsndf{A}{{P}{Q}}}{\cprojsndf{P}{Q}}
+  \tag{by definition}
+\end{align*}
 \end{proof}
@@
 \begin{align*}
-\ctxwk{A}{\ctxext{Q}{R}{g}{t}}
+\ctxwk{A}{\tmext{Q}{R}{g}{t}}
 & \jdeq 
   \ctxwk{A}{\subst{t}{{g}{\idtm{\ctxext{Q}{R}}}}}
+  \tag{by definition}
   \\
@@
   \subst{\ctxwk{A}{t}}{\ctxwk{A}{\subst{g}{\idtm{\ctxext{Q}{R}}}}}
+  \tag{by \autoref{comp-ws-t}}
   \\
@@
   \subst{\ctxwk{A}{t}}{{\ctxwk{A}{g}}{\ctxwk{A}{\idtm{\ctxext{Q}{R}}}}}
+  \tag{by \autoref{comp-ws-t}}
   \\
@@
   \subst{\ctxwk{A}{t}}{{\ctxwk{A}{g}}{\idtm{\ctxwk{A}{\ctxext{Q}{R}}}}}
+  \tag{by \autoref{comp-wi-t}}
   \\
@@
   \subst{\ctxwk{A}{t}}{{\ctxwk{A}{g}}{\idtm{\ctxext{\ctxwk{A}{Q}}{\ctxwk{A}{R}}}}}
+  \tag{by \autoref{comp-we-f}}
   \\
@@
   \tmext{\ctxwk{A}{Q}}{\ctxwk{A}{R}}{\ctxwk{A}{g}}{\ctxwk{A}{t}}
+  \tag{by definition}
 \end{align*}
@@
   \unfoldall{\cprojfstf{\ctxwk{A}{Q}}{\ctxwk{A}{R}}}
+  \tag{by definition}
   \\
@@
   \ctxwk{{A}{R}}{{A}{\idtm{Q}}}
+  \tag{by \autoref{comp-wi-t}}
   \\
@@
   \unfoldall{\ctxwk{A}{\cprojfstf{Q}{R}}}
+  \tag{by \autoref{comp-ww-t}}
   \\
@@
   \ctxwk{A}{\cprojfstf{Q}{R}}.
+  \tag{by definition}
 \end{align*}
@@
   \unfoldall{\cprojsndf{\ctxwk{A}{Q}}{\ctxwk{A}{R}}}
+  \tag{by definition}
   \\
@@
   \unfoldall{\ctxwk{A}{\cprojsndf{Q}{R}}}
+  \tag{by \autoref{comp-wi-t}}
   \\
@@
   \ctxwk{A}{\cprojsndf{Q}{R}}.
-  \qedhere
+  \tag{by definition}
 \end{align*}
@@
   \subst{x}{{t}{{g}{\idtm{\ctxext{Q}{R}}}}}
+  \tag{by definition}
   \\
@@
   \subst{{x}{t}}{{x}{{g}{\idtm{\ctxext{Q}{R}}}}}
+  \tag{by \autoref{comp-ss-t}}
   \\
@@
   \subst{{x}{t}}{{{x}{g}}{{x}{\idtm{\ctxext{Q}{R}}}}}
+  \tag{by \autoref{comp-ss-t}}
   \\
@@
   \subst{{x}{t}}{{{x}{g}}{\idtm{\subst{x}{\ctxext{Q}{R}}}}}
+  \tag{by \autoref{comp-si-t}}
   \\
@@
   \subst{{x}{t}}{{{x}{g}}{\idtm{\ctxext{\subst{x}{Q}}{\subst{x}{R}}}}}
+  \tag{by \autoref{comp-se-f}}
   \\
@@
   \tmext{\subst{x}{Q}}{\subst{x}{R}}{\subst{x}{g}}{\subst{x}{t}}.
+  \tag{by definition}
 \end{align*}
@@
   \unfoldall{\cprojfstf{\subst{x}{Q}}{\subst{x}{R}}}
+  \tag{by definition}
   \\
@@
   \ctxwk{\subst{x}{R}}{\subst{x}{\idtm{Q}}}
+  \tag{by \autoref{comp-si-t}}
   \\
@@
   \unfoldall{\subst{x}{\cprojfstf{Q}{R}}}
+  \tag{by \autoref{comp-ss-t}}
   \\
@@
   \subst{x}{\cprojfstf{Q}{R}}.
+  \tag{by definition}
 \end{align*}
@@
   \unfoldall{\cprojsndf{\subst{x}{Q}}{\subst{x}{R}}}
+  \tag{by definition}
   \\
@@
   \unfoldall{\subst{x}{\cprojsndf{Q}{R}}}
+  \tag{by \autoref{comp-si-t}}
   \\
@@
   \subst{x}{\cprojsndf{Q}{R}}.
-  \qedhere
+  \tag{by definition}
 \end{align*}
@@
 
-\begin{lem}\label{lem:tmext-id}\label{comp-ie}
+\begin{lem}\label{comp-ie}
 For any $\jfam{\Gamma}{A}$ and $\jfam{{\Gamma}{A}}{P}$ we have
-\begin{equation}\label{idfunc-ext-comp}
+\begin{equation*}
 \inference
@@
     }
-\end{equation}
+\end{equation*}
 \end{lem}
@@
     }
+  \tag{by definition}
   \\
@@
       }
+  \tag{by definition}
   \\
@@
       }
+  \tag{by \autoref{comp-wi-t}}
   \\
@@
       }
+  \tag{by \autoref{comp-ew-t}}
   \\
@@
       }
+  \tag{by \autoref{comp-ws-t}}
   \\
@@
       }
+  \tag{by \autoref{precomp-idtm-t}}
   \\
@@
   \idtm{\ctxext{A}{P}}.
-  \qedhere
-\end{align*}
+  \tag{by \autoref{precomp-idtm-t}}
+\end{align*}
+\end{proof}
+
+We also have the following lemma about the compatibility of pairing and composition:
+
+\begin{lem}\label{lem:tmext-jcomp}
+The following inference rules are valid
+\begin{align*}
+& \inference
+  { \jhom{\Gamma}{A}{B}{f}
+    \jterm{{{\Gamma}{B}}{Q}}{R}{h}
+    \jterm{{{\Gamma}{B}}{Q}}{\subst{h}{S}}{k}
+    }
+  { \jtermeq
+      {{{\Gamma}{A}}{\jcomp{A}{f}{Q}}}
+      {\jcomp{A}{f}{\ctxext{R}{S}}}
+      {\jcomp{A}{f}{\tmext{h}{k}}}
+      {\tmext{\jcomp{A}{f}{h}}{\jcomp{A}{f}{k}}}
+    }
+  \\
+& \inference
+  { \jhom{\Gamma}{A}{B}{f}
+    \jfam{{{{\Gamma}{B}}{Q}}{R}}{S}
+    }
+  { \jhomeq
+      {{{\Gamma}{A}}{\jcomp{A}{f}{Q}}}
+      {\jcomp{A}{f}{\ctxext{R}{S}}}
+      {\jcomp{A}{f}{R}}
+      {\jcomp{A}{f}{\cprojfstf{R}{S}}}
+      {\cprojfstf{\jcomp{A}{f}{R}}{\jcomp{A}{f}{S}}}
+    }
+  \\
+& \inference
+  { \jhom{\Gamma}{A}{B}{f}
+    \jfam{{{{\Gamma}{B}}{Q}}{R}}{S}
+    }
+  { \jhomeq
+      {{{{\Gamma}{A}}{\jcomp{A}{f}{Q}}}{\jcomp{A}{f}{S}}}
+      {\jcomp{A}{f}{S}}
+      {\jcomp{A}{f}{S}}
+      {\jcomp{A}{f}{\cprojsndf{R}{S}}}
+      {\cprojsndf{\jcomp{A}{f}{R}}{\jcomp{A}{f}{S}}}
+    }
+\end{align*}
+\end{lem}
+
+\begin{proof}
+Let $\jhom{\Gamma}{A}{B}{f}$, $\jterm{{{\Gamma}{B}}{Q}}{R}{h}$
+and $\jterm{{{\Gamma}{B}}{Q}}{\subst{h}{S}}{k}$. Then we have the
+judgmental equalities
+\begin{align*}
+\jcomp{A}{f}{\tmext{\ctxwk{B}{C}}{\ctxwk{B}{R}}{g}{h}}
+& \jdeq 
+  \subst{f}{\ctxwk{A}{\tmext{\ctxwk{B}{C}}{\ctxwk{B}{R}}{g}{h}}}
+  \tag{by definition}
+  \\
+& \jdeq 
+  \subst
+    {f}
+    {\tmext{\ctxwk{A}{{B}{C}}}{\ctxwk{A}{{B}{R}}}{\ctxwk{A}{g}}{\ctxwk{A}{h}}}
+  \tag{by \autoref{lem:tmext-wk}}
+  \\
+& \jdeq 
+  \tmext
+    {\ctxwk{A}{C}}
+    {\ctxwk{A}{R}}
+    {\subst{f}{\ctxwk{A}{g}}}
+    {\subst{f}{\ctxwk{A}{h}}}
+  \tag{by \autoref{lem:tmext-subst}}
+  \\
+& \jdeq 
+  \tmext{\ctxwk{A}{C}}{\ctxwk{A}{R}}{\jcomp{A}{f}{g}}{\jcomp{A}{f}{h}}
+  \tag{by definition}
+\end{align*}
+validating the first inference rule. For the second and the third, note that
+we have the judgmental equalities
+\begin{align*}
+\jcomp{A}{f}{\cprojfstf{R}{S}}
+& \jdeq
+  \unfold\jcomp{A}{f}{\cprojfstf{R}{S}}
+  \tag{by definition}
+  \\
+& \jdeq
+  \subst{f}{\cprojfstf{\ctxwk{A}{R}}{\ctxwk{A}{S}}}
+  \tag{by \autoref{lem:tmext-wk}}
+  \\
+& \jdeq
+  \cprojfstf{\subst{f}{\ctxwk{A}{R}}}{\subst{f}{\ctxwk{A}{R}}}
+  \tag{by \autoref{lem:tmext-subst}}
+  \\
+& \jdeq
+  \cprojfstf{\jcomp{A}{f}{R}}{\jcomp{A}{f}{S}}
+  \tag{by definition}
+\end{align*}
+and
+\begin{align*}
+\jcomp{A}{f}{\cprojsndf{R}{S}}
+& \jdeq
+  \unfold\jcomp{A}{f}{\cprojsndf{R}{S}}
+  \tag{by definition}
+  \\
+& \jdeq
+  \subst{f}{\cprojsndf{\ctxwk{A}{R}}{\ctxwk{A}{S}}}
+  \tag{by \autoref{lem:tmext-wk}}
+  \\
+& \jdeq
+  \cprojsndf{\subst{f}{\ctxwk{A}{R}}}{\subst{f}{\ctxwk{A}{R}}}
+  \tag{by \autoref{lem:tmext-subst}}
+  \\
+& \jdeq
+  \cprojsndf{\jcomp{A}{f}{R}}{\jcomp{A}{f}{S}}
+  \tag{by definition}
+\end{align*}
+\end{proof}
+
+In \autoref{lem:mor-tmext-cproj} below we will explain what happens in the
+case where the codomain of the function $f$ of the previous lemma is itself
+extended.
+
+Recall that the judgment $\jhom{\Gamma}{A}{{B}{Q}}{f}$
+unfolds as
+\begin{equation*}
+\unfold{\jhom{\Gamma}{A}{{B}{Q}}{f}}
+\end{equation*}
+and that we have the judgmental equality 
+$ \jfameq
+    {{\Gamma}{A}}
+    {\ctxwk{A}{\ctxext{B}{Q}}}
+    {\ctxext{\ctxwk{A}{B}}{\ctxwk{A}{Q}}}.
+  $
+Therefore, each morphism into an extended family can itself be described as
+an extended term. The following lemma explains how this goes.
+
+\begin{lem}\label{lem:mor-tmext-cproj}
+Let $\jhom{\Gamma}{A}{{B}{Q}}{f}$ be a morphism from $A$ to $\ctxext{B}{Q}$
+in a context $\Gamma$. Then we have
+\begin{equation*}
+\jhomeq
+  {\Gamma}
+  {A}
+  {{B}{Q}}
+  {f}
+  {\tmext{\jcomp{}{f}{\cprojfstf{B}{Q}}}{\jcomp{}{f}{\cprojsndf{B}{Q}}}}.
+\end{equation*}
+Alternatively, when $\jhom{\Gamma}{A}{B}{f_0}$ and 
+$\jterm{{\Gamma}{A}}{\jcomp{}{f_0}{Q}}{f_1}$ we obtain a morphism
+$\jhom{\Gamma}{A}{{B}{Q}}{\tmext{f_0}{f_1}}$ with the property that
+\begin{align*}
+\jalign\jhomeq{\Gamma}{A}{B}{\jcomp{}{\tmext{f_0}{f_1}}{\cprojfstf{B}{Q}}}{f_0}\\
+\jalign\jtermeq{{\Gamma}{A}}{\jcomp{}{f_0}{Q}}{\jcomp{}{\tmext{f_0}{f_1}}{\cprojsndf{B}{Q}}}{f_1}.
+\end{align*}
+\end{lem}
+
+\begin{proof}
+Let $\jhom{\Gamma}{A}{{B}{Q}}{f}$ be a morphism in context $\Gamma$. Then we
+have the judgmental equalities
+\begin{align*}
+\cprojfst{\ctxwk{A}{B}}{\ctxwk{A}{Q}}{f}
+& \jdeq
+  \subst{f}{\ctxwk{A}{\cprojfstf{B}{Q}}}
+  \tag{by \autoref{lem:tmext-wk}}
+  \\
+& \jdeq
+  \jcomp{}{f}{\cprojfstf{B}{Q}}
+  \tag{by definition}
+  \\
+\cprojsnd{\ctxwk{A}{B}}{\ctxwk{A}{Q}}{f}
+& \jdeq
+  \subst{f}{\ctxwk{A}{\cprojsndf{B}{Q}}}
+  \tag{by \autoref{lem:tmext-wk}}
+  \\
+& \jdeq
+  \jcomp{}{f}{\cprojsndf{B}{Q}}
+  \tag{by definition}
+\end{align*}
+The alternative formulation of the statement is a direct corollary.
+\end{proof}
+
+Precomposing by the first projection $\cprojfstf{A}{P}:\ctxext{A}{P}\to A$
+is judgmentally equal to weakening by $P$:
+
+\begin{lem}\label{lem:jcomp-cprojfstf}
+The following inference rule is valid:
+\begin{align*}
+& \inference
+  { \jfam{{\Gamma}{A}}{P}
+    \jfam{{\Gamma}{A}}{Q}
+    }
+  { \jfameq
+      {{\Gamma}{{A}{P}}}
+      {\jcomp{{A}{P}}{\cprojfstf{A}{P}}{Q}}
+      {\ctxwk{P}{Q}}
+    }
+  \\
+& \inference
+  { \jfam{{\Gamma}{A}}{P}
+    \jfam{{{\Gamma}{A}}{Q}}{R}
+    }
+  { \jfameq
+      {{{\Gamma}{{A}{P}}}{\ctxwk{A}{Q}}}
+      {\jcomp{{A}{P}}{\cprojfstf{A}{P}}{R}}
+      {\ctxwk{P}{R}}
+    }
+  \\
+& \inference
+  { \jfam{{\Gamma}{A}}{P}
+    \jterm{{{\Gamma}{A}}{Q}}{R}{h}
+    }
+  { \jtermeq
+      {{{\Gamma}{{A}{P}}}{\ctxwk{A}{Q}}}
+      {\ctxwk{A}{R}}
+      {\jcomp{{A}{P}}{\cprojfstf{A}{P}}{h}}
+      {\ctxwk{P}{h}}
+    }
+\end{align*}
+\end{lem}
+
+\begin{proof}
+Let $\jfam{{\Gamma}{A}}{P}$ and $\jfam{{\Gamma}{A}}{Q}$ be
+families. Then we have
+\begin{align*}
+\jcomp{{A}{P}}{\cprojfstf{A}{P}}{Q}
+& \jdeq
+  \unfoldall{\jcomp{{A}{P}}{\cprojfstf{A}{P}}{Q}}
+  \tag{by definition}\\
+& \jdeq 
+  \subst{\ctxwk{P}{\idtm{A}}}{\ctxwk{P}{{A}{Q}}} 
+  \tag{by \autoref{comp-ww-f}}\\
+& \jdeq 
+  \ctxwk{P}{\subst{\idtm{A}}{\ctxwk{A}{Q}}} 
+  \tag{by \autoref{comp-ws-f}}\\
+& \jdeq 
+  \ctxwk{P}{Q} 
+  \tag{by \autoref{precomp-idtm-c}}
+\end{align*}
+The other proofs are similar.
+\end{proof}
+
+As an application of these lemmas we prove that every term $\jterm{{\Gamma}{A}}{P}{f}$
+induces a section $\jtermashom{f}$ of $\cprojfstf{A}{P}$.
+Note that we can directly view any term $x$ of $A$ in context $\Gamma$ as a
+morphism from the empty family $\emptyf$ to $A$ in context $\Gamma$.
+
+Now let $\jterm{{\Gamma}{A}}{P}{f}$ be a term of $P$. Then we can view $f$ as a
+morphism from $A$ to $\ctxext{A}{P}$ in the following way:
+
+\begin{defn}\label{defn:terms-as-sections}
+Let $\jterm{{\Gamma}{A}}{P}{f}$ be a term. Then we define
+\begin{equation*}
+\jhomdefn{\Gamma}{A}{{A}{P}}{\jtermashom{f}}{\unfold{\jtermashom{f}}}
+\end{equation*}
+\end{defn}
+
+In the following lemma we will see that $\jtermashom{f}$ is a section for
+$\cprojfstf{A}{P}$.
+
+\begin{cor}
+Consider a family $\jfam{{\Gamma}{A}}{P}$ and a term $\jterm{\Gamma}{A}{x}$.
+Then we have the judgmental equality
+\begin{equation*}
+\jhomeq{\Gamma}{A}{A}{\jcomp{A}{\jtermashom{f}}{\cprojfstf{A}{P}}}{\idtm{A}}.
+\end{equation*}
+\end{cor}
+
+\begin{proof}
+The asserted judgmental equality follows immediately from \autoref{lem:mor-tmext-cproj}.
 \end{proof}
@@
 but not in the way we're about to display them).
-
-For the following lemma, recall that the judgment $\jhom{\Gamma}{A}{{B}{Q}}{f}$
-unfolds as
-\begin{equation*}
-\unfold{\jhom{\Gamma}{A}{{B}{Q}}{f}}
-\end{equation*}
-and that we have the judgmental equality 
-$ \jfameq
-    {{\Gamma}{A}}
-    {\ctxwk{A}{\ctxext{B}{Q}}}
-    {\ctxext{\ctxwk{A}{B}}{\ctxwk{A}{Q}}}.
-  $
-Therefore, each morphism into an extended family can itself be described as
-an extended term. The following lemma explains how this goes.
-
-\begin{lem}\label{lem:mor-tmext-cproj}
-Let $\jhom{\Gamma}{A}{{B}{Q}}{f}$ be a morphism from $A$ to $\ctxext{B}{Q}$
-in a context $\Gamma$. Then we have
-\begin{equation*}
-\jhomeq
-  {\Gamma}
-  {A}
-  {{B}{Q}}
-  {f}
-  {\tmext{\jcomp{}{f}{\cprojfstf{B}{Q}}}{\jcomp{}{f}{\cprojsndf{B}{Q}}}}.
-\end{equation*}
-Alternatively, when $\jhom{\Gamma}{A}{B}{f_0}$ and 
-$\jterm{{\Gamma}{A}}{\jcomp{}{f_0}{Q}}{f_1}$ we obtain a morphism
-$\jhom{\Gamma}{A}{{B}{Q}}{\tmext{f_0}{f_1}}$ with the property that
-\begin{align*}
-\jalign\jhomeq{\Gamma}{A}{B}{\jcomp{}{\tmext{f_0}{f_1}}{\cprojfstf{B}{Q}}}{f_0}\\
-\jalign\jtermeq{{\Gamma}{A}}{\jcomp{}{f_0}{Q}}{\jcomp{}{\tmext{f_0}{f_1}}{\cprojsndf{B}{Q}}}{f_1}.
-\end{align*}
-\end{lem}
-
-\begin{proof}
-Let $\jhom{\Gamma}{A}{{B}{Q}}{f}$ be a morphism in context $\Gamma$. Then we
-have the judgmental equalities
-\begin{align*}
-\cprojfst{\ctxwk{A}{B}}{\ctxwk{A}{Q}}{f}
-& \jdeq
-  \subst{f}{\ctxwk{A}{\cprojfstf{B}{Q}}}
-  \\
-& \jdeq
-  \jcomp{}{f}{\cprojfstf{B}{Q}}
-  \\
-\cprojsnd{\ctxwk{A}{B}}{\ctxwk{A}{Q}}{f}
-& \jdeq
-  \subst{f}{\ctxwk{A}{\cprojsndf{B}{Q}}}
-  \\
-& \jdeq
-  \jcomp{}{f}{\cprojsndf{B}{Q}}
-\end{align*}
-The alternative formulation of the statement is a direct corollary.
-\end{proof}
-
-In the same spirit, we have the following compatibility of composition with
-projections:
-
-\begin{lem}\label{lem:jcomp-cprojfstf}
-The following inference rule is valid:
-\begin{equation*}
-\inference
-  { \jfam{{\Gamma}{A}}{P}
-    \jfam{{\Gamma}{A}}{Q}
-    }
-  { \jfameq
-      {{\Gamma}{{A}{P}}}
-      {\jcomp{{A}{P}}{\cprojfstf{A}{P}}{Q}}
-      {\ctxwk{P}{Q}}
-    }
-\end{equation*}
-\end{lem}
-
-\begin{proof}
-Let $\jfam{{\Gamma}{A}}{P}$ and $\jfam{{\Gamma}{A}}{Q}$ be
-families. Then we have
-\begin{align*}
-\jcomp{{A}{P}}{\cprojfstf{A}{P}}{Q}
-& \jdeq
-  \unfoldall{\jcomp{{A}{P}}{\cprojfstf{A}{P}}{Q}}
-  \tag{by definition}\\
-& \jdeq 
-  \subst{\ctxwk{P}{\idtm{A}}}{\ctxwk{P}{{A}{Q}}} 
-  \tag{by \autoref{comp-ww-f}}\\
-& \jdeq 
-  \ctxwk{P}{\subst{\idtm{A}}{\ctxwk{A}{Q}}} 
-  \tag{by \autoref{comp-ws-f}}\\
-& \jdeq 
-  \ctxwk{P}{Q} 
-  \tag{by \autoref{precomp-idtm-c}}
-\end{align*}
-\end{proof}
-
-We also have the following lemma about the compatibility of pairing and composition:
-
-\begin{lem}\label{lem:tmext-jcomp}
-The following inference rule is valid
-\begin{align*}
-& \inference
-  { \jhom{\Gamma}{A}{B}{f}
-    \jhom{\Gamma}{B}{C}{g}
-    \jfam{{\Gamma}{C}}{R}
-    \jterm{{\Gamma}{B}}{\subst{g}{\ctxwk{B}{R}}}{w}
-    }
-  { \jhomeq
-      {\Gamma}
-      {A}
-      {{C}{R}}
-      {\jcomp{A}{f}{\tmext{\ctxwk{B}{C}}{\ctxwk{B}{R}}{g}{w}}}
-      {\tmext{\ctxwk{A}{C}}{\ctxwk{A}{R}}{\jcomp{A}{f}{g}}{\jcomp{A}{f}{w}}}
-    }
-\end{align*}
-\end{lem}
-
-\begin{proof}
-Let $\jhom{\Gamma}{A}{B}{f}$, $\jhom{\Gamma}{B}{C}{g}$, $\jfam{{\Gamma}{C}}{R}$
-and $\jterm{{\Gamma}{B}}{\subst{g}{\ctxwk{B}{R}}}{w}$. Then we have the
-judgmental equalities
-\begin{align*}
-\jcomp{A}{f}{\tmext{\ctxwk{B}{C}}{\ctxwk{B}{R}}{g}{w}}
-& \jdeq 
-  \subst{f}{\ctxwk{A}{\tmext{\ctxwk{B}{C}}{\ctxwk{B}{R}}{g}{w}}}
-  \\
-& \jdeq 
-  \subst
-    {f}
-    {\tmext{\ctxwk{A}{{B}{C}}}{\ctxwk{A}{{B}{R}}}{\ctxwk{A}{g}}{\ctxwk{A}{w}}}
-  \\
-& \jdeq 
-  \tmext
-    {\ctxwk{A}{C}}
-    {\ctxwk{A}{R}}
-    {\subst{f}{\ctxwk{A}{g}}}
-    {\subst{f}{\ctxwk{A}{w}}}
-  \\
-& \jdeq 
-  \tmext{\ctxwk{A}{C}}{\ctxwk{A}{R}}{\jcomp{A}{f}{g}}{\jcomp{A}{f}{w}}.
-  \qedhere
-\end{align*}
-\end{proof}
-
-As an application of these lemmas we prove that every term $\jterm{{\Gamma}{A}}{P}{f}$
-induces a section $\jtermashom{f}$ of $\cprojfstf{A}{P}$.
-Note that we can directly view any term $x$ of $A$ in context $\Gamma$ as a
-morphism from the empty family $\emptyf$ to $A$ in context $\Gamma$.
-
-Now let $\jterm{{\Gamma}{A}}{P}{f}$ be a term of $P$. Then we can view $f$ as a
-morphism from $A$ to $\ctxext{A}{P}$ in the following way:
-
-\begin{defn}\label{defn:terms-as-sections}
-Let $\jterm{{\Gamma}{A}}{P}{f}$ be a term. Then we define
-\begin{equation*}
-\jhomdefn{\Gamma}{A}{{A}{P}}{\jtermashom{f}}{\unfold{\jtermashom{f}}}
-\end{equation*}
-\end{defn}
-
-In the following lemma we will see that $\jtermashom{f}$ is a section for
-$\cprojfstf{A}{P}$.
-
-\begin{cor}
-Consider a family $\jfam{{\Gamma}{A}}{P}$ and a term $\jterm{\Gamma}{A}{x}$.
-Then we have the judgmental equality
-\begin{equation*}
-\jhomeq{\Gamma}{A}{A}{\jcomp{A}{\jtermashom{f}}{\cprojfstf{A}{P}}}{\idtm{A}}.
-\end{equation*}
-\end{cor}
-
-\begin{proof}
-The asserted judgmental equality follows immediately from \autoref{lem:mor-tmext-cproj}.
-\end{proof}
-
-There is a notion of morphism \emph{over} a morphism. We will develop this
-notion because it will be needed in the theory of models later on.
 

### 2014-10-28 Rules/rules_aftermath.tex
reorganization

--- 2014-09-20 article-models-tt.tex
@@
-\section{The generalized algebraic theory of contexts, families and terms (the E-system)}\label{tt}
-In this section we give a description of dependent type theory before type
-constructors. Apart from contexts, families and terms -- which provide for the
-core of the language of dependent logic -- the basic ingredients
-of this theory will be the operations of extension on context and on families, 
-and empty context and empty families, weakening, substitution
-and the identity terms. The resulting theory can be seen as a manifestation of the 
-structure underlying dependent type theory.
+\subsection{%
+  The possiblity of types in the theory of contexts, families and terms}
+\label{types}
 
-We will
-formulate the theory of contexts, families and terms in such a way that contexts aren't defined
-to be lists of variable declarations. The variable-free (a.k.a.~\emph{name-free}) approach 
-we take here is rather different than those appearing in 
-\cite{hofmann1995extensional,TheBook} but it has appeared in the work of Coquand
-and in \cite{Dybjer1996}.
-The main reason we don't let variable declarations in is that we don't see them 
-in the internal models either. This way we also set out to a more algebraic 
-approach of type theory and higher category theory. Thirdly, we will not have to
-be burdened with superficial comments about variables being bounded or not, or 
-fresh or free or not occuring at all.
-
-In the current presentation, contexts can be seen as binary
-planar trees of which the leaves are (families of) contexts. 
-The judgmental equality relation on contexts is an equivalence relation which 
-expresses that binary planar
-trees of contexts are judgmentally equal if their leaves are, taking only
-(the isomorphism class of) 
-the order of the leaves into account \emph{and not the actual shape of the three}.
-The intuition behind this equivalence relation is indeed that unstructured
-(i.e.~unbracketed) lists such as the lists of variable declarations which
-usually appear in type theoretical syntax, may be regarded as contexts.
-
-Besides contexts, families and terms there will also be a notion of `type in
-a context', which one may assume.
-A context is eligible to be a \emph{closed type} and over a context, a family
-may be a type. We have the intuition that `being a type'
-expresses the property of `being atomic or irreducible'. With a notion of types,
-one has the means to require a stratification by which every context is a finite
-extension of types starting at a closed type, and every family is likewise a
-finite extension of types. The idea behind this is to recover the theory of
-\cite{Garner2014} {\color{red}(and possibly B-systems?)}. We stress that
-we will not include axioms asserting such a stratification in the E-system
-and we leave discussion on the possibility of a typing judgment until the
-end of this section.
-
-With families of contexts being the principal things of
-study, we have added a bit more generality to the theory, compared to the
-theory of categories with families presented in \cite{Dybjer1996}. Here are
-a couple of reasons for this extra generality:
-\begin{enumerate}
-\item We want to be very sure that we can interpret all the rules of type theory
-in a straightforward manner in any context.
-\item We do want do end up with a categorical structure on the contexts, derived
-from the theory by requiring that a morphism from $\Gamma$ to $\Delta$ is a
-term of the weakening of $\Delta$ by $\Gamma$ in context $\Gamma$. Thus, we
-must be able to lift the context $\Delta$ to become a family of contexts over
-$\Gamma$ and it is not reasonable to expect that this results in a type.
-\item It is very easy in this setting to assert that an operation has an action
-on very high levels of families, i.e.~when $F$ is an operation taking things in
-context $\Gamma$ to things in context $\Delta$, then it takes a family $A$ in
-context $\Gamma$ to $F(A)$ in context $\Delta$; a family $P$ in context
-$\ctxext{\Gamma}{A}$ to a family $F(P)$ in context $\ctxext{\Delta}{(F(A))}$,
-and so on. The reason it becomes simple is that we don't have to explain what
-`and so on' means.
-\end{enumerate}
-
-We will formulate fairly strict rules governing the judgmental equalities,
-expressing that extension, weakening and substitution are combatible with
-each other in a judgmental manner. This does not, however, diminish the role
-of isomorphisms or of homotopies could play in the theory once identity types
-are added. Indeed, types could still have non-trivial identity relations and
-the category of types in a certain context could genuinely display higher
-categorical structure, or so we conjecture.
-
-Much of the rules we state are just compatibility rules of extension, weakening
-and substitution with each other. In a way, these rules assert that our contexts
-are just structureless lists of contexts and that likewise terms are structureless
-lists of terms. They are structureless in the sense that the order in which
-they are formed by pairing up is irrelevant. We note that this causes complications
-in the traditional way that categorical sematics of type theory is implemented,
-where contexts become objects of the category which is supposed to model type
-theory. The reason for this is that context extension will not satisfy all the
-compatibility rules we're about to state. The first step to resolving this is taking
-the types in the empty context as the objects.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Judgments and valid inference rules}\label{judgments}
-The theory we describe here is a theory of contexts, families of
-contexts and terms thereof. The families of contexts are by some authors called
-dependent contexts, but they are handled a bit differently here because they
-become the primary object of study. Dependent contexts can be types; they could
-be seen as atomic or indecomposable dependent contexts.
-
-Thus we make eight kinds of judgments: ``$\Gamma$ is a context'',
-``$A$ is a family of contexts over $\Gamma$'', ``$A$ is a type in context $\Gamma$''
-and ``$x$ is a term of the family $A$ of contexts over $\Gamma$''. The other four
-judgments are for judgmental equality.
-\begin{align*}
-\jalign\jctx{\Gamma} 
-& \jalign\jctxeq{\Gamma}{\Gamma'}
-  \\
-\jalign\jfam{\Gamma}{A} 
-& \jalign\jfameq{\Gamma}{A}{B}
-  \\
-\jalign\jterm{\Gamma}{A}{x} 
-& \jalign\jtermeq{\Gamma}{A}{x}{y}.
-\end{align*}
-
-Strictly speaking, we have three different judgmental equalities in play and one
-could request for a notational difference to signify that fact. For instance,
-we could denote the judgmental equalities of contexts, families and terms by
-$\jdeq_c$, $\jdeq_f$ and $\jdeq_t$ respectively. It will, however, always be
-clear which of the three kinds of judgmental equality is meant when we assert
-a judgmental equality and therefore we shall not bother to make this notational
-distinction.
-
-We note that what we call families over contexts
-here could also have been named dependent contexts or telescopes, see
-\cite{deBruijn1991,hofmann1995extensional}. The term family is in agreement
-with the terminology scheme of \cite{TheBook}, though the reader should be
-warned that the notion of familie means something slightly different there than
-it does here.
-
-The theory of contexts, families and terms shall be described by means of a
-declaration of valid inference rules. An inference rule is an expression of the
-form
-\begin{equation*}
-\inference{\mathcal{H}_1\quad\cdots\quad\mathcal{H}_n}{\mathcal{J}}
-\end{equation*}
-where each of $\mathcal{H}_1$, ..., $\mathcal{H}_n$ and $\mathcal{J}$ is 
-judgment. The judgments $\mathcal{H}_1$, ..., $\mathcal{H}_n$ are called the
-hypotheses and the judgment $\mathcal{J}$ is the conclusion of the inference
-rule. When a judgment is the conclusion of a valid inference rule, we also say
-that the judgment is valid or well-formed. In fact, the only way to construct
-new judgments is via the valid inference rules we describe in this section.
-
-The class of valid inference rules is closed under composition and under
-rearranging the hypotheses. That
-means that if
-\begin{equation*}
-\inference{\mathcal{I}_{1k}\quad\cdots\quad\mathcal{I}_{m_kk}}{\mathcal{H}_k}
-\end{equation*}
-and
-\begin{equation*}
-\inference{\mathcal{H}_1\quad\cdots\quad\mathcal{H}_n}{\mathcal{J}}
-\end{equation*}
-are valid inference rules for $1\leq k\leq n$, then
-\begin{equation*}
-\inference{\mathcal{I}_{11}\quad\cdots\quad\mathcal{I}_{m_11}\quad\cdots\quad\mathcal{I}_{1n}\quad\cdots\quad\mathcal{I}_{m_nn}}{\mathcal{J}}
-\end{equation*}
-is also a valid inference rule. Also, in a valid inference rule 
-duplicated hypotheses may always be contracted and hypotheses may always be 
-reordered; the result is still considered a valid inference rule.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{The basic rules for judgmental equality}
-The rules for judgmental equality establish that it is an equivalence relation
-in all three cases (contexts, families and terms). Thus, the following inference
-rules shall be required to be valid:
-\bgroup\small
-\begin{align*}
-& \inference
-  { \jctx{\Gamma}
-    }
-  { \jctxeq{\Gamma}{\Gamma}
-    } 
-& & \inference
-    { \jctxeq{\Gamma}{\Delta}
-      }
-    { \jctxeq{\Delta}{\Gamma}
-      } 
-& & \inference
-    { \jctxeq{\Gamma}{\Delta}
-      \jctxeq{\Delta}{\greek{E}}
-      }
-    { \jctxeq{\Gamma}{\greek{E}}
-      }
-    \\
-& \inference
-  { \jfam{\Gamma}{A}
-    }
-  { \jfameq{\Gamma}{A}{A}
-    } 
-& & \inference
-    { \jfameq{\Gamma}{A}{B}
-      }
-    { \jfameq{\Gamma}{B}{A}
-      }
-& & \inference
-    { \jfameq{\Gamma}{A}{B}
-      \jfameq{\Gamma}{B}{C}
-      }
-    { \jfameq{\Gamma}{A}{C}
-      }
-    \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    }
-  { \jtermeq{\Gamma}{A}{x}{x}
-    }
-& & \inference
-    { \jtermeq{\Gamma}{A}{x}{y}
-      }
-    { \jtermeq{\Gamma}{A}{y}{x}
-      }
-& & \inference
-    { \jtermeq{\Gamma}{A}{x}{y}
-      \jtermeq{\Gamma}{A}{y}{z}
-      }
-    { \jtermeq{\Gamma}{A}{x}{z}
-      }
-\end{align*}
-\egroup
-
-The following convertibility rules are responsible for the strictness
-of judgmental equality, which sets it apart from equivalences or identifications:
-\begin{align*}
-& \inference
-  { \jctxeq{\Gamma}{\Delta}
-    \jfam{\Gamma}{A}
-    }
-  { \jfam{\Delta}{A}
-    }
-& & \inference
-    { \jctxeq{\Gamma}{\Delta}
-      \jfameq{\Gamma}{A}{B}
-      }
-    { \jfameq{\Delta}{A}{B}
-      }
-    \\
-& \inference
-  { \jctxeq{\Gamma}{\Delta}
-    \jterm{\Gamma}{A}{x}
-    }
-  { \jterm{\Delta}{A}{x}
-    }
-& & \inference
-    { \jctxeq{\Gamma}{\Delta}
-      \jtermeq{\Gamma}{A}{x}{y}
-      }
-    { \jtermeq{\Delta}{A}{x}{y}
-      }
-    \\
-& \inference
-  { \jfameq{\Gamma}{A}{B}
-    \jterm{\Gamma}{A}{x}
-    }
-  { \jterm{\Gamma}{B}{x}
-    }
-& & \inference
-    { \jfameq{\Gamma}{A}{B}
-      \jtermeq{\Gamma}{A}{x}{y}
-      }
-    { \jtermeq{\Gamma}{B}{x}{y}
-      }
-\end{align*}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{The operations of the theory of contexts, families and terms}
-In this subsection we introduce the notions of
-extension on context and on families, the empty context and the empty families, 
-weakening, 
-substitution and identity terms.
-
-Weakening and substitution are both
-operations of the theory of contexts, families and terms and therefore they will
-have three levels of actions: on contexts (at the appropriate level), families
-and terms.
-
-\begin{comment}
-Generally, 
-each operation should
-have an action on contexts, on families and on terms at the `appropriate level'
-and should be compatible with each of the operations we introduce here. In
-that will become more practical after the theory of contexts, families and terms
-is formulated, this means that when we will write down a type theory that
-interprets itself in \autoref{ttmodels}, each of the notions of
-extension, empty family, weakening, substitution and identity terms are going to be internalized
-as terms of the theory and all terms are going to be compatible with these.
-That vague idea will be our guiding light.
-
-So in principle, the introductions of each of the empty context, extension,
-weakening, substitution and identity terms should come in three parts. In 
-practice, they do not all come in three parts.
-
-Apart from the empty context, there will indeed be an empty family and a term
-of the empty family (which will be deemed unique). But extension will
-not have an explicit action `at the level of families' since extension
-is used to define higher families and the action would have to be trivial
-anyway. The empty context and extension provide the essentials of the rest of
-the theory in the sense that higher families are families over extensions,
-contexts are families over the empty context and the `action on families' of
-any operation applied to families over the empty context is the same as the
-original operation. Thus, when we have introduced the operations of weakening
-and substitution we will see that they also have an `action on the terms of 
-contexts'. We will state the relevant judgmental equality rules for the empty
-context in this section rather than in the next, since they are required to
-state that the action on families of an operation is compatible with the
-original operation.
-\end{comment}
-
-We note upfront that both context extension and family extension can also be
-seen as operations of contexts, families and terms, even though we will not
-introduce them as such. The reason is that extension is used to define what it
-means to be a family over a family over a context (namely, a family over an
-extended context). As such, it will have a trivial action on families and terms
-which will not need to be introduced. We will come back to this point in
-\autoref{extension}.
-
-\subsection{Extension}
-\label{extension}
-The operations of extension enable one to consider families over
-families and higher level families alike. We will need both a context extension
-operation and a family extension operation.
-
-The inference rules introducing context and family extension are as follows:
-\begin{align}
-& \inference
-  { \jfam{\Gamma}{A}
-    }
-  { \jctx{\ctxext{\Gamma}{A}}
-    }
-& & \inference
-    { \jctxeq{\Gamma}{\Delta}
-      \jfameq{\Gamma}{A}{B}
-      }
-    { \jctxeq{\ctxext{\Gamma}{A}}{\ctxext{\Delta}{B}}
-      }
-    \\
-& \inference
-  { \jfam{{\Gamma}{A}}{P}
-    }
-  { \jfam{\Gamma}{\ctxext{A}{P}}
-    }
-& & \inference
-    { \jfameq{\Gamma}{A}{B} 
-      \jfameq{{\Gamma}{A}}{P}{Q}
-      }
-    { \jfameq{\Gamma}{\ctxext{A}{P}}{\ctxext{B}{Q}}
-      }
-\end{align}
-The extension operation may also be defined to act on families over families
-and terms thereof. For instance, When $\jfam{{\Gamma}{A}}{P}$ is a family,
-we would get a family $\jfam{{\Gamma}{A}}{{\Gamma}{P}}$ and when
-$\jterm{{\Gamma}{A}}{P}{f}$ is a term, we would get a term
-$\jterm{{\Gamma}{A}}{{\Gamma}{P}}{\ctxext{\Gamma}{f}}$. Then there would be
-axioms stating the judgmental equalities $\ctxext{\Gamma}{P}\jdeq P$ and
-$\ctxext{\Gamma}{f}\jdeq f$. Thus, it is of no actual use to introduce the
-action of context extension on families and terms.
-
-In \autoref{extension-on-terms} we will also define an extension operation
-on terms, which defines a term $\jterm{\Gamma}{{A}{P}}{\tmext{x}{u}}$ for
-every $\jterm{\Gamma}{A}{x}$ and $\jterm{\Gamma}{\subst{x}{P}}{u}$. This
-operation is defined by using the substitution operation and identity terms.
-
-\subsubsection{Associativity of extension}
-\label{comp-ee}
-The inference rules asserting that extension is compatible with itself assert
-that contexts are unstructured lists of type declarations. This rule is
-unavoidable if we want that for a family $A$ in context $\Gamma$, a family over
-$A$ is the same thing as a family over $\ctxext{\Gamma}{A}$. 
-\begin{align}
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{{\Gamma}{A}}{P}
-    }
-  { \jctxeq{\ctxext{{\Gamma}{A}}{P}}{\ctxext{\Gamma}{{A}{P}}}
-    }
-  \\
-& \inference
-  { \jfam{{\Gamma}{A}}{P}
-    \jfam{{{\Gamma}{A}}{P}}{Q}
-    }
-  { \jfameq{\Gamma}{\ctxext{A}{{P}{Q}}}{\ctxext{{A}{P}}{Q}}
-    }
-  \label{comp-ee-c}
-\end{align}
-Note that we may take the point of view that extension distributes over
-extensions. If we want to extend an extended family $\ctxext{P}{Q}$ over
-$\ctxext{\Gamma}{A}$ by the family $A$, the result is the extension of (1)
-the extension of $P$ by $A$ and (2) the extension of $Q$ by $A$. The latter is
-simply $Q$ since extension acts trivially at the level of families (and terms).
-
-\subsection{The empty context and the empty families}
-\label{empty}
-We introduce an empty context and an family over $\Gamma$ for every context $\Gamma$. 
-It has been suggested by some to only include empty families and not an
-empty context because an empty context is not necessary, but we do have several
-reasons to include them. Having an empty context requires also rules asserting 
-that a context is the same thing as a family over the empty context and this
-gives the categorical structure on contexts for free once one has it for 
-families. The main ingredients that will be missing from the theory once an
-empty context is avoided are weakening by a context and identity morphisms from
-a context to itself. Including these by hand also requires to formulate all the
-compatibility rules involving weakening once more for the cases of weakening by
-a context and identity terms at contexts. We prefer to state these rules once
-and only once and including an empty context helps in this respect.
-
-We also prefer to have our scheme of compatibility rules as symmetrical as
-possible. The structure of type dependency should look exactly the same in the
-default case as in any context. In that respect, having an empty family but not
-an empty context seems a bit odd. Also, some sets of compatibility rules 
-(like the rules stating that extension is compatible with the empty families)
-will become assymetrical as a result of not including an empty context. Moreover,
-we would eventually like to include a stratification of the theory by means of
-a type judgment (asserting that a family in a context $\Gamma$ is a type) and
-study closed types (i.e.~types in the empty context). It would be possible to
-provide a notion of closed types without having an empty context, but this would
-have to be formulated separately and we would have to restate all the rules for
-types (if any) for closed types all over again.
-
-One of the main uses of the empty context and the empty families is that we
-get the property that the `action on contexts' of an operation is compatible
-with its `action on families'.
-
-The empty family over a context $\Gamma$ is introduced by the following rule
-inference rule:
-\begin{align}
-& \inference
-  { }
-  { \jctx{\emptyc}
-    }
-  \\
-& \inference
-  { \jctx{\Gamma}
-    }
-  { \jfam{\Gamma}{\emptyf[\Gamma]}
-    }
-  \\
-& \inference
-  { \jctxeq{\Gamma}{\Gamma'}
-    }
-  { \jfameq{\Gamma}{\emptyf[\Gamma]}{\emptyf[\Gamma']}
-    }
-\end{align}
-
-By regarding contexts as families of contexts over the empty context, we
-enable ourselves also to speak of terms of contexts. A term of a context
-$\Gamma$ is a term of the family $\Gamma$ over the empty context. These ideas
-are captured in the following convertibility rules:
-\begin{align}
-& \inference
-  { \jctx{\Gamma}
-    }
-  { \jfam{\emptyc}{\Gamma}
-    } 
-& & \inference
-    { \jfam{\emptyc}{\Gamma}
-      }
-    { \jctx{\Gamma}
-      }
-    \\
-& \inference
-  { \jctxeq{\Gamma}{\Delta}
-    }
-  { \jfameq{\emptyc}{\Gamma}{\Delta}
-    }
-& & \inference
-    { \jfameq{\emptyc}{\Gamma}{\Delta}
-      }
-    { \jctxeq{\Gamma}{\Delta}
-      }
-\end{align}
-
-The reader may wonder whether the empty family $\emptyf$ always has a
-term. This shall follow from the rules stating the compatibility of extension
-with the empty families in \autoref{comp-0e} below and from
-identity terms (\autoref{identityterms}).
-\begin{comment}
-When contexts are viewed as statements, the empty context is the statement that
-asserts nothing. Likewise, in any context one can assert nothing and this is
-achieved by the empty family. 
-Since somebody who states nothing asserts it all right, there is a term of the empty
-family over any context. Moreover, any two terms of the empty family are always
-judgmentally equal.
-\begin{align}
-& \inference
-  { }
-  { \jctx{\emptyc}
-    }
-  \\
-& \inference
-  { \jctx{\Gamma}
-    }
-  { \jfam{\Gamma}{\emptyf[\Gamma]}
-    }
-  \\
-& \inference
-  { \jctx{\Gamma}
-    }
-  { \jterm{\Gamma}{\emptyf[\Gamma]}{\emptytm[\Gamma]}
-    }
-  \\
-& \inference
-  { \jterm{\Gamma}{\emptyf[\Gamma]}{x}
-    }
-  { \jtermeq
-      {\Gamma}
-      {\emptyf[\Gamma]}
-      {x}
-      {\emptytm[\Gamma]}
-    }
-  \label{comp-00-t}
-\end{align}
-
-The empty context and the empty families together with the operation of
-extension will provide several shortcuts for the rest
-of the theory of families and terms. This is (in part) obtained by requiring
-that a context is exactly the same thing as a family of contexts over the
-empty context and that judgmental equality of contexts is exactly the same
-as judgmental equality of families of contexts over the empty context.
-
-\label{comp-00}
-Because every context is a family of contexts over the empty context, 
-the empty context $\emptyc$ may be considered as a family of contexts over the empty
-context. When we do this, we get $\emptyf[\emptyc]$.
-\begin{equation}\label{eq:comp-00}
-\inference
-  { }
-  { \jfameq
-      {\emptyc}
-      {\emptyc}
-      {\emptyf[\emptyc]}
-    }
-\end{equation}
-In the future, we shall denote $\emptyf[\Gamma]$ by $\emptyf$. The above rule
-guarantees that this will not cause confusion. Likewise, we shall denote
-$\emptytm[\Gamma]$ by $\emptytm$.
-\end{comment}
-
-\subsubsection{Compatibility of extension with the empty context and families}
-In the following set of inference rules we state that the empty context and
-the empty family are neutral objects for both context extension (the first two
-rules) and family extension (the last two rules).
-\label{comp-e0}\label{comp-0e}
-\begin{align}
-& \inference
-  { \jctx{\Gamma}
-    }
-  { \jctxeq{\ctxext{\emptyc}{\Gamma}}{\Gamma}
-    }
-  \label{comp-0e-c}
-  \\
-& \inference
-  { \jctx{\Gamma}
-    }
-  { \jctxeq{\ctxext{\Gamma}{\emptyf}}{\Gamma}
-    }
-  \label{comp-e0-c}\\
-& \inference
-  { \jfam{\Gamma}{A}
-    }
-  { \jfameq{\Gamma}{\ctxext{\emptyf}{A}}{A}
-    }
-  \label{comp-0e-f}
-  \\
-& \inference
-  { \jfam{\Gamma}{A}
-    }
-  { \jfameq{\Gamma}{{A}{\emptyf}}{A}
-    }
-  \label{comp-e0-f}
-\end{align}
-
-\subsubsection{Family extension restricted to the empty context}
-Strictly speaking we should have used a different notation for context extension
-as for family extension, because the following rule asserting that family extension
-in the empty context is the same thing as context extension would look tautological
-without a difference. So let us denote, only for the moment, context extension
-of $\Gamma$ by $A$ by $(\ctxext{\Gamma}{A})^c$ and family extension of $A$ by
-$P$ in context $\Gamma$ by $(\ctxext{A}{P})^\famsym$.
-
-Note that we may consider a context $\Gamma$ as a family over $\emptyc$ and
-a family $\jfam{\Gamma}{A}$ as a family $\jfam{{\emptyc}{\Gamma}}{A}$. 
-Therefore we will require the following rule:
-\begin{align}
-& \inference
-  { \jfam{\Gamma}{A}
-    }
-  { \jctxeq{(\ctxext{\Gamma}{A})^c}{(\ctxext{\Gamma}{A})^\famsym}
-    }
-\end{align}
-Note that this rule actually justifies that we have not utilized two different
-notations for context extension and family extension.
-
-\subsection{The theory of weakening}
-\label{weakening}
-When $A$ is a family in context $\Gamma$, the operation of weakening by $A$
-takes a family $B$ in context $\Gamma$ and provides a family $\ctxwk{A}{B}$
-in context $\ctxext{\Gamma}{A}$. The context family $\ctxwk{A}{B}$ can be seen
-as the constant family over $\ctxext{\Gamma}{A}$ with value $B$. This idea will
-be axiomatized in the cancellation property of weakening and substitution in
-\autoref{cancellation-ws}. In \autoref{morphisms} we will take the terms 
-$\unfold{\jhom{\Gamma}{A}{B}{f}}$ to be the morphisms of families from $A$ to 
-$B$. These will be at the heart of the categorical structure of the theory.
-
-The weakening operation acts on three levels: on contexts, on families and
-on terms. The `action on contexts' of weakening is the action we described
-above: it takes a family $B$ over $\Gamma$ to a family $\ctxwk{A}{B}$ over
-$\ctxext{\Gamma}{A}$; the `action on families' of weakening takes a family
-$Q$ over $\ctxext{\Gamma}{B}$ to a family $\ctxwk[\famsym]{A}{Q}$ over
-$\ctxext{{\Gamma}{A}}{\ctxwk{A}{B}}$; the `action on terms' of weakening takes
-a term $g$ of $Q$ to a term $\ctxwk[\tmsym]{A}{g}$ of $\ctxwk[\famsym]{A}{Q}$.
-\begin{align}
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{\Gamma}{B}
-    }
-  { \jfam{{\Gamma}{A}}{\ctxwk{A}{B}}
-    }
-& & \inference
-    { \jfameq{\Gamma}{A}{A'}
-      \jfameq{\Gamma}{B}{B'}
-      }
-    { \jfameq{{\Gamma}{A}}{\ctxwk{A}{B}}{\ctxwk{A'}{B'}}
-      }
-    \\
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{{\Gamma}{B}}{Q}
-    }
-  { \jfam{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk[\famsym]{A}{Q}}
-    }
-& & \inference
-    { \jfameq{\Gamma}{A}{A'}
-      \jfameq{{\Gamma}{B}}{Q}{Q'}
-      }
-    { \jfameq
-        {{{\Gamma}{A}}{\ctxwk{A}{B}}}
-        {\ctxwk[\famsym]{A}{Q}}
-        {\ctxwk[\famsym]{A'}{Q'}}
-      }
-    \\
-%& \inference
-%  { \jterm{\Gamma}{B}{y}
-%    }
-%  { \jterm{{\Gamma}{A}}{\ctxwk{A}{B}}{\ctxwk{A}{y}}
-%    }
-%& & \inference
-%    { \jtermeq{\Gamma}{B}{y}{y'}
-%      }
-%    { \jtermeq
-%        {{\Gamma}{A}}
-%        {\ctxwk{A}{B}}
-%        {\ctxwk{A}{y}}
-%        {\ctxwk{A'}{y'}}
-%      }
-%    \\
-& \inference
-  { \jfam{\Gamma}{A}
-    \jterm{{\Gamma}{B}}{Q}{g}
-    }
-  { \jterm{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk[\famsym]{A}{Q}}{\ctxwk[\tmsym]{A}{g}}
-    }
-& & \inference
-    { \jfameq{\Gamma}{A}{A'}
-      \jtermeq{{\Gamma}{B}}{Q}{g}{g'}
-      }
-    { \jtermeq
-        {{{\Gamma}{A}}{\ctxwk{A}{B}}}
-        {\ctxwk[\famsym]{A}{Q}}
-        {\ctxwk[\tmsym]{A}{g}}
-        {\ctxwk[\tmsym]{A'}{g'}}
-      }
-\end{align}
-
-\subsubsection{Weakening of an empty family}
-The following inference rules express that when the empty family is
-weakened, the result is the empty family.
-\label{comp-w0}
-\begin{align}
-%& \inference
-%  { \jctx{\Gamma}
-%    }
-%  { \jfameq{\Gamma}{\ctxwk{\Gamma}{\emptyc}}{\emptyf}
-%    }
-%  \\
-& \inference
-  { \jfam{\Gamma}{A}
-    }
-  { \jfameq{{\Gamma}{A}}{\ctxwk{A}{\emptyf}}{\emptyf}
-    }
-  \label{comp-w0-c}\\
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{\Gamma}{B}
-    }
-  { \jfameq
-    {{{\Gamma}{A}}{\ctxwk{A}{B}}}
-    {\ctxwk[\famsym]{A}{\emptyf}}
-    {\emptyf}
-    }
-  \label{comp-w0-f}
-\end{align}
-\begin{comment}
-Note that we do not have to axiomatize the action of weakening
-by $A$ on the term of the empty family, because the result will be in the empty 
-family and every term in the empty family is judgmentally equal to the term 
-$\emptytm$.
-\end{comment}
-
-Because a family over $\Gamma$ is the same as a family over 
-$\ctxext{\Gamma}{\emptyf}$ we can apply both the action on contexts and the
-action on families of weakening to a family $B$ over $\Gamma$. When we apply
-the action on families, we obtain a family $\ctxwk[\famsym]{A}{B}$ over the
-context $\ctxext{{\Gamma}{A}}{\ctxwk{A}{\emptyf}}$. However, since we have
-postulated the judgmental equalities $\ctxwk{A}{\emptyf}\jdeq\emptyf$ and
-$\ctxext{{\Gamma}{A}}{\emptyf}\jdeq\ctxext{\Gamma}{A}$, we see that we can
-compare $\ctxwk[\famsym]{A}{B}$ with $\ctxwk{A}{B}$. The following inference
-rule postulates that these two are judgmentally equal:
-\begin{equation}
-\inference
-{ \jfam{\Gamma}{A}
-  \jfam{\Gamma}{B}
-  }
-{ \jfameq{{\Gamma}{A}}{\ctxwk[\famsym]{A}{B}}{\ctxwk{A}{B}}
-  }
-\end{equation}
-Due to this rule, the action on contexts and the action on families of weakening
-are compatible with each other and consequently there can be no possible
-confusion when we omit the annotations $\famsym$ and $\tmsym$ alltogether. In
-the future, the weakening of a family $Q$ over $\ctxext{\Gamma}{B}$ shall
-be denoted just by $\ctxwk{A}{Q}$ and likewise the weakening of a term $g$ of
-$Q$ shall be denoted by $\ctxwk{A}{g}$.
-
-Because a family $B$ over $\Gamma$ can be treated as a family by the operation
-of weakening, weakening also acts on the terms of $B$. The weakening of a term
-$y$ of $B$ by $A$ can be seen as the constant term (or function) of the
-family $\ctxwk{A}{B}$.
-
-\subsubsection{Weakenings of extensions}
-\label{comp-we}
-The following rules assert that when an extended family is weakened, the
-weakening distributes over the extension factors.
-\begin{align}
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{{\Gamma}{B}}{Q}
-    }
-  { \jfameq
-      {\ctxext{\Gamma}{A}}
-      {\ctxwk{A}{\ctxext{B}{Q}}}
-      {\ctxext{\ctxwk{A}{B}}{\ctxwk{A}{Q}}}
-    }
-  \label{comp-we-c}
-  \\
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{{{\Gamma}{B}}{Q}}{R}
-    }
-  { \jfameq
-      {\ctxext{{\Gamma}{A}}{\ctxwk{A}{B}}}
-      {\ctxwk{A}{\ctxext{Q}{R}}}
-      {\ctxext{\ctxwk{A}{Q}}{\ctxwk{A}{R}}}
-    }
-  \label{comp-we-f}
-\end{align}
-When thinking of terms of $\ctxwk{A}{B}$ as morphisms of families from $A$ to
-$B$, this looks already like form of type theoretic choice. It is weaker in that
-it is not stated with function types, yet it is stronger in that it states a
-judgmental equality between two families.
-
-There is also a version of this property where an extended term is weakened.
-This variant is stated and proved in \autoref{comp-we-t}.
-
-\subsubsection{Weakening of weakenings}
-Weakening by a family $A$ in context $\Gamma$ brings things in context $\Gamma$
-to things in context $\ctxext{\Gamma}{A}$. Since we have all the ingredients of
-the theory of contexts, families and terms in the context $\Gamma$ as well it
-has in particular it's own weakening by families. Suppose we have a family
-$\jfam{{\Gamma}{B}}{Q}$ over $B$ in context $\Gamma$. Weakening by $Q$ brings
-things from context $\ctxext{\Gamma}{B}$ to $\ctxext{{\Gamma}{B}}{Q}$. Thus
-we can provide rules asserting what will happen when we first weaken by $Q$ and
-then (via the action on families) by $A$. We will require the following
-inference rules to be valid:
-\label{comp-ww}
-\begin{align}
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{{\Gamma}{B}}{Q}
-    \jfam{{\Gamma}{B}}{R}
-    }
-  { \jfameq
-      {{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{Q}}}
-      {\ctxwk{A}{{Q}{R}}}
-      {\ctxwk{{A}{Q}}{{A}{R}}}
-    }
-  \label{comp-ww-c}\\
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{{\Gamma}{B}}{Q}
-    \jfam{{{\Gamma}{B}}{R}}{S}
-    }
-  { \jfameq
-      {{{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{Q}}}{\ctxwk{A}{{Q}{R}}}}
-      {\ctxwk{A}{{Q}{S}}}
-      {\ctxwk{{A}{Q}}{{A}{S}}}
-    }
-  \label{comp-ww-f}\\
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{{\Gamma}{B}}{Q}
-    \jterm{{{\Gamma}{B}}{R}}{S}{k}
-    }
-  { \jtermeq
-      {{{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{Q}}}{\ctxwk{A}{{Q}{R}}}}
-      {\ctxwk{A}{{Q}{S}}}
-      {\ctxwk{A}{{Q}{k}}}
-      {\ctxwk{{A}{Q}}{{A}{k}}}
-    }
-  \label{comp-ww-t}
-\end{align}
-
-\begin{rmk}
-As an important special case of these inference rules we have the following
-valid inference rules:
-\begin{align*}
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{\Gamma}{B}
-    \jfam{\Gamma}{C}
-    }
-  { \jfameq
-      {{{\Gamma}{A}}{\ctxwk{A}{B}}}
-      {\ctxwk{A}{{B}{C}}}
-      {\ctxwk{{A}{B}}{{A}{C}}}
-    }
-  \\
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{\Gamma}{B}
-    \jfam{{\Gamma}{C}}{R}
-    }
-  { \jfameq
-      {{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{{B}{C}}}}
-      {\ctxwk{A}{{B}{R}}}
-      {\ctxwk{{A}{B}}{{A}{R}}}
-    }
-  \\
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{\Gamma}{B}
-    \jterm{{\Gamma}{C}}{R}{h}
-    }
-  { \jtermeq
-      {{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{{B}{C}}}}
-      {\ctxwk{A}{{B}{R}}}
-      {\ctxwk{A}{{B}{h}}}
-      {\ctxwk{{A}{B}}{{A}{h}}}
-    }
-\end{align*}
-Moreover, we have
-\begin{equation*}
-\inference
-  { \jfam{\Gamma}{A}
-    \jfam{\Gamma}{B}
-    \jterm{\Gamma}{C}{z}
-    }
-  { \jtermeq
-      {{{\Gamma}{A}}{\ctxwk{A}{B}}}
-      {\ctxwk{A}{{B}{C}}}
-      {\ctxwk{A}{{B}{z}}}
-      {\ctxwk{{A}{B}}{{A}{z}}}
-    }
-\end{equation*}
-\end{rmk}
-
-\subsubsection{Weakening by the empty family}
-Note that we can also weaken by the empty family over $\Gamma$.
-Weakening by the empty family $\emptyf$ over a context $\Gamma$ leaves families, 
-their terms, families over those families and terms of those unchanged:
-\label{comp-0w}
-\begin{align}
-& \inference
-  { \jfam{\Gamma}{B}
-    }
-  { \jfameq{\Gamma}{\ctxwk{\emptyf}{B}}{B}
-    }
-  \label{comp-0w-c}
-  \\
-& \inference
-  { \jfam{{\Gamma}{B}}{Q}
-    }
-  { \jfameq{{\Gamma}{B}}{\ctxwk{\emptyf}{Q}}{Q}
-    }
-  \label{comp-0w-f}
-  \\
-& \inference
-  { \jterm{{\Gamma}{B}}{Q}{g}
-    }
-  { \jtermeq{{\Gamma}{B}}{Q}{\ctxwk{\emptyf}{g}}{g}
-    }
-  \label{comp-0w-t}
-\end{align}
-
-\subsubsection{Currying for weakening}
-\label{comp-ew}
-The rules expressing that extension is compatible with weakening assert that
-weakening by an extension is the same thing as weakening twice in the
-appropriate way.
-\begin{align}
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{{\Gamma}{A}}{P}
-    \jfam{\Gamma}{B}
-    }
-  { \jfameq
-      {{{\Gamma}{A}}{P}}
-      {\ctxwk{\ctxext{A}{P}}{B}}
-      {\ctxwk{P}{{A}{B}}}
-    }
-  \label{comp-ew-c}\\
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{{\Gamma}{A}}{P}
-    \jfam{{\Gamma}{B}}{Q}
-    }
-  { \jfameq
-      {{{{\Gamma}{A}}{P}}{\ctxwk{P}{{A}{B}}}}
-      {\ctxwk{\ctxext{A}{P}}{Q}}
-      {\ctxwk{P}{{A}{Q}}}
-    }
-  \label{comp-ew-f}\\
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{{\Gamma}{A}}{P}
-    \jterm{{\Gamma}{B}}{Q}{g}
-    }
-  { \jtermeq
-      {{{{\Gamma}{A}}{P}}{\ctxwk{P}{{A}{B}}}}
-      {\ctxwk{P}{{A}{Q}}}
-      {\ctxwk{\ctxext{A}{P}}{g}}
-      {\ctxwk{P}{{A}{g}}}
-    } 
-  \label{comp-ew-t}
-\end{align}
-
-\begin{comment}
-\subsubsection{Weakening by a context}
-Since we have not required that there is an empty context with the property that
-a context is the same thing as a family in the empty context, we will have to
-explicitly introduce weakening by contexts. Since it will allways be clear
-whether something is weakened by a context or by a family, we shall use the
-same notation for weakening by contexts as for weakening by families. We shall
-have to repeat all the compatibility rules for weakening by families for
-weakening by contexts; there is no way around it. It is
-actually this sort of inconvenience that has originally led us to include an
-empty context and be done with it.
-
-In the following inference rules we assume that we have a context $\Gamma$ and
-that $\jctxeq{\Gamma}{\Gamma'}$ in the rules where $\Gamma'$ makes an appearence.
-\begin{align}
-& \inference
-  { \jctx{\Delta}
-    }
-  { \jfam{\Gamma}{\ctxwk{\Gamma}{\Delta}}
-    }
-& & \inference
-  { \jctxeq{\Delta}{\Delta'}
-    }
-  { \jfameq{\Gamma}{\ctxwk{\Gamma}{\Delta}}{\ctxwk{\Gamma'}{\Delta'}}
-    }
-  \\
-& \inference
-  { \jfam{\Delta}{B}
-    }
-  { \jfam{\ctxext{\Gamma}{\ctxwk{\Gamma}{\Delta}}}{\ctxwk{\Gamma}{B}}
-    }
-& & \inference
-  { \jfameq{\Delta}{B}{B'}
-    }
-  { \jfameq
-      {\ctxext{\Gamma}{\ctxwk{\Gamma}{\Delta}}}
-      {\ctxwk{\Gamma}{B}}
-      {\ctxwk{\Gamma'}{B'}}
-    }
-  \\
-& \inference
-  { \jterm{\Delta}{B}{y}
-    }
-  { \jterm
-      {\ctxext{\Gamma}{\ctxwk{\Gamma}{\Delta}}}
-      {\ctxwk{\Gamma}{B}}
-      {\ctxwk{\Gamma}{y}}
-    }
-& & \inference
-  { \jtermeq{\Delta}{B}{y}{y'}
-    }
-  { \jtermeq
-      {\ctxext{\Gamma}{\ctxwk{\Gamma}{\Delta}}}
-      {\ctxwk{\Gamma}{B}}
-      {\ctxwk{\Gamma}{y}}
-      {\ctxwk{\Gamma}{y'}}
-    }
-\end{align}
-Note that we do not employ three different notations for the three different
-actions of weakening by a context because there can be no confusion about which
-of the three actions is meant. If there were an empty context we should have
-distinguished the three as we did for weakening by a family.
-
-We shall have similar compatibility rules for weakening by a context as we did
-for weakening by a family.
-
-{\color{red}To do. It will look exactly the same as for weakening by families}
-\end{comment}
-
-\subsection{The theory of projections}
-\label{identityterms}
-We formulate the theory of projections on top of the theory of weakening
-described in \autoref{weakening}. In the theory of projections we introduce
-the identity terms. The identity term at a family $A$ in context $\Gamma$ is
-a term $\jterm{{\Gamma}{A}}{\ctxwk{A}{A}}{\idtm{A}}$, so weakening is necessary.
-Together with weakening, the identity terms will provide for the projections.
-The projection from $\cftext{A}{P}$ to $A$ will be the term
-$\jterm{{\Gamma}{{A}{P}}}{\ctxwk{\ctxext{A}{P}}{A}}{\ctxwk{P}{\idtm{A}}}$ and
-will be studied extensively in \autoref{extension-on-terms}.
-
-The following inference rules are required to be valid:
-\begin{align}
-& \inference
-  { \jfam{\Gamma}{A}
-    }
-  { \unfoldall{\jhom{\Gamma}{A}{A}{\idtm{A}}}
-    }
-& & \inference
-    { \jfameq{\Gamma}{A}{A'}
-      }
-    { \unfoldall{\jhomeq{\Gamma}{A}{A}{\idtm{A}}{\idtm{A'}}}
-      }
-\end{align}
-
-\begin{comment}
-We stress again
-that because we have not required there to be an empty context, we need to
-explicitly call for identity terms $\idtm{\Gamma}$ of $\ctxwk{\Gamma}{\Gamma}$
-in context $\Gamma$ for each context $\Gamma$. Thus, we will also require
-the inference rule
-\begin{align}
-& \inference
-  { \jctx{\Gamma}
-    }
-  { \jterm{\Gamma}{\ctxwk{\Gamma}{\Gamma}}{\idtm{\Gamma}}
-    }
-& &
-  \inference
-  { \jctxeq{\Gamma}{\Gamma'}
-    }
-  { \jtermeq{\Gamma}{\ctxwk{\Gamma}{\Gamma}}{\idtm{\Gamma}}{\idtm{\Gamma'}}
-    }
-\end{align}
-For the following inference rules, there should also be versions of identity
-terms at contexts.
-\end{comment}
-
-\subsubsection{The identity term of an extended family}
-The fact that identity terms are compatible with extensions is derived in
-\autoref{comp-ie}.
-
-\subsubsection{The identity term of a weakened family}
-\label{comp-wi}
-The identity term of a weakened family is the weakened identity term:
-\begin{equation}
-\inference
-  { \jfam{\Gamma}{A}
-    \jfam{{\Gamma}{B}}{Q}
-    }
-  { \unfoldall{\jhomeq
-      {{{\Gamma}{A}}{\ctxwk{A}{B}}}
-      {\ctxwk{A}{Q}}
-      {\ctxwk{A}{Q}}
-      {\ctxwk{A}{\idtm{Q}}}
-      {\idtm{\ctxwk{A}{Q}}}
-      }
-    }
-  \label{comp-wi-t}
-\end{equation}
-
-\begin{rmk}
-An important special case is the judgmental equality
-\begin{equation*}
-\unfoldall{\jhomeq
-      {{\Gamma}{A}}
-      {\ctxwk{A}{B}}
-      {\ctxwk{A}{B}}
-      {\ctxwk{A}{\idtm{B}}}
-      {\idtm{\ctxwk{A}{B}}}
-      }
-\end{equation*}
-for a family $\jfam{\Gamma}{B}$.
-\end{rmk}
-
-\subsection{The theory of substitution}
-\label{substitution}
-The theory of substitution can be formulated separately from the theory of
-weakening. Therefore, we will only require the theory of extension and the
-empty context and families. In \autoref{sec:esystem-equalities} we will
-provide the additional judgmental equality rules that will be required when
-combining the theories of substitution and of projections.
-
-Given a family $P$ over $A$ and a term $x$ of $A$, substitution gives a way to
-consider the \emph{fiber $\subst{x}{P}$ of $P$ at $x$}. As was the case with
-weakening, the substitution operation comes in three parts: the `action on
-contexts' of substitution is the operation just described; the `action on
-families' of substitution takes a family $Q$ over $\ctxext{{\Gamma}{A}}{P}$
-to a family $\subst[\famsym]{x}{Q}$ over $\ctxext{\Gamma}{\subst{x}{P}}$; the
-`action on terms' of substitution takes a term $g$ of $Q$ over
-$\ctxext{{\Gamma}{A}}{P}$ to a term $\subst[\tmsym]{x}{g}$ of 
-$\subst[\famsym]{x}{Q}$.
-\begin{align}
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jfam{{\Gamma}{A}}{P}
-    }
-  { \jfam{\Gamma}{\subst{x}{P}}
-    }
-& & \inference
-    { \jtermeq{\Gamma}{A}{x}{x'}
-      \jfameq{{\Gamma}{A}}{P}{P'}
-      }
-    { \jfameq{\Gamma}{\subst{x}{P}}{\subst{x'}{P'}}
-      }
-    \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jfam{{{\Gamma}{A}}{P}}{Q}
-    }
-  { \jfam{{\Gamma}{\subst{x}{P}}}{\subst[\famsym]{x}{Q}}
-    }
-& & \inference
-    { \jtermeq{\Gamma}{A}{x}{x'}
-      \jfameq{{{\Gamma}{A}}{P}}{Q}{Q'}
-      }
-    { \jfameq{{\Gamma}{\subst{x}{P}}}{\subst[\famsym]{x}{Q}}{\subst[\famsym]{x'}{Q'}}
-      }
-    \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jterm{{{\Gamma}{A}}{P}}{Q}{g}
-    }
-  { \jterm{{\Gamma}{\subst{x}{P}}}{\subst[\famsym]{x}{Q}}{\subst[\tmsym]{x}{g}}
-    }
-& & \inference
-    { \jtermeq{\Gamma}{A}{x}{x'}
-      \jtermeq{{{\Gamma}{A}}{P}}{Q}{g}{g'}
-      }
-    { \jtermeq
-        {{\Gamma}{\subst{x}{P}}}
-        {\subst[\famsym]{x}{Q}}
-        {\subst[\tmsym]{x}{g}}
-        {\subst[\tmsym]{x'}{g'}}
-      }
-\end{align}
-
-\subsubsection{Fibers of an empty family}
-The following inference rules establish that the fibers of the empty family are 
-the empty families:
-\label{comp-s0}
-\begin{align}
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    }
-  { \jfameq{\Gamma}{\subst{x}{\emptyf}}{\emptyf}
-    }
-  \label{comp-s0-c}
-  \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jfam{{\Gamma}{A}}{P}
-    }
-  { \jfameq
-      {{\Gamma}{\subst{x}{P}}}
-      {\subst{x}{\emptyf}}
-      {\emptyf}
-    }
-  \label{comp-s0-f}
-\end{align}
-\begin{comment}
-Note again that we do not have to axiomatize the action of substitution by $x$ on the term
-of the empty family, because the result will be in the empty family and every
-term in the empty family is judgmentally equal to the term $\emptytm$.
-\end{comment}
-
-We use the above rule to state the compatibility of the action on families of
-substitution with the action on contexts of substitution. Note that a family
-$P$ over $\ctxext{\Gamma}{A}$ may be regarded as a family over
-$\ctxext{{\Gamma}{A}}{\emptyf}$. Thus, we may consider the family
-$\subst[\famsym]{x}{P}$ over $\ctxext{\Gamma}{\subst{x}{\emptyf}}$. Since
-$\subst{x}{\emptyf}$ is judgmentally equal to the empty family, we may compare
-$\subst[\famsym]{x}{P}$ with $\subst{x}{P}$:
-\begin{equation}
-\inference
-{ \jfam{{\Gamma}{A}}{P}
-  }
-{ \jfameq{\Gamma}{\subst[\famsym]{x}{P}}{\subst{x}{P}}
-  }
-\end{equation}
-Due to this rule we need not make the annotations $\famsym$ and $\tmsym$ in
-the notation for substitution anymore and thus we shall omit them from now on.
-Note that because a family $P$ over $\ctxext{\Gamma}{A}$ is eligible for
-application of the action on families of substitution, we may also substitute
-terms of $P$. Thus, given terms $\jterm{\Gamma}{A}{x}$ and $\jterm{{\Gamma}{A}}{P}{f}$,
-we get a term $\jterm{\Gamma}{\subst{x}{P}}{\subst{x}{f}}$, the \emph{value of
-$f$ at $x$}.
-
-In \autoref{morphisms} we will use a combination of weakening and substitution
-to define composition of morphisms of families. However, we have to rely
-on the cancellation rule stated in \autoref{cancellation-ws} before we can
-meaningfully state the definition of composition.
-
-\subsubsection{Fibers of extensions}
-\label{comp-se}
-The following inference rules assert that if we take the fiber of an extended
-family at a term $x$ of $A$ in context $\Gamma$, the substitution by $x$
-distributes over the factors of the extension.
-\begin{align}
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jfam{{{\Gamma}{A}}{P}}{Q}
-    }
-  { \jfameq
-      {\Gamma}
-      {\subst{x}{\ctxext{P}{Q}}}
-      {\ctxext{\subst{x}{P}}{\subst{x}{Q}}}
-    }
-  \label{comp-se-c}
-  \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jfam{{{{\Gamma}{A}}{P}}{Q}}{R}
-    }
-  { \jfameq
-      {{\Gamma}{\subst{x}{P}}}
-      {\subst{x}{\ctxext{Q}{R}}}
-      {\ctxext{\subst{x}{Q}}{\subst{x}{R}}}
-    }
-  \label{comp-se-f}
-\end{align}
-There is also a version of this statement in which extended terms are considered.
-This variant is stated and proved in \autoref{comp-se-t}.
-
-\subsubsection{Fibers of fibers}
-\label{comp-ss}
-
-The following rules explain what happens when we first substitute by a term
-$g$ of $Q$ over $P$ in context $\ctxext{\Gamma}{A}$ and then by a term $x$ of
-$A$:
-\begin{align}
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jterm{{{\Gamma}{A}}{P}}{Q}{g}
-    \jfam{{{{\Gamma}{A}}{P}}{Q}}{R}
-    }
-  { \jfameq
-      {{{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}}
-      {\subst{x}{{g}{R}}}
-      {\subst{{x}{g}}{{x}{R}}}
-    }
-  \label{comp-ss-c}
-  \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jterm{{{\Gamma}{A}}{P}}{Q}{g}
-    \jfam{{{{{\Gamma}{A}}{P}}{Q}}{R}}{S}
-    }
-  { \jfameq
-      {{{{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}}{\subst{x}{{g}{R}}}}
-      {\subst{x}{{g}{S}}}
-      {\subst{{x}{g}}{{x}{S}}}
-    }
-  \label{comp-ss-f}
-  \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jterm{{{\Gamma}{A}}{P}}{Q}{g}
-    \jterm{{{{{\Gamma}{A}}{P}}{Q}}{R}}{S}{k}
-    }
-  { \jtermeq
-      {{{{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}}{\subst{x}{{g}{R}}}}
-      {\subst{x}{{g}{S}}}
-      {\subst{x}{{g}{k}}}
-      {\subst{{x}{g}}{{x}{k}}}
-    }
-  \label{comp-ss-t}
-\end{align}
-
-\begin{rmk}
-As an important special case of these inference rules we have the following
-valid inference rules:
-\begin{align*}
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jterm{{\Gamma}{A}}{P}{f}
-    \jfam{{{\Gamma}{A}}{P}}{Q}
-    }
-  { \jfameq
-      {\Gamma}
-      {\subst{x}{{f}{Q}}}
-      {\subst{{x}{f}}{{x}{Q}}}
-    }
-  \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jterm{{\Gamma}{A}}{P}{f}
-    \jfam{{{{\Gamma}{A}}{P}}{Q}}{R}
-    }
-  { \jfameq
-      {{\Gamma}{\subst{x}{{f}{Q}}}}
-      {\subst{x}{{f}{R}}}
-      {\subst{{x}{f}}{{x}{R}}}
-    }
-  \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jterm{{\Gamma}{A}}{P}{f}
-    \jterm{{{{\Gamma}{A}}{P}}{Q}}{R}{h}
-    }
-  { \jtermeq
-      {{\Gamma}{\subst{x}{{f}{Q}}}}
-      {\subst{x}{{f}{R}}}
-      {\subst{x}{{f}{h}}}
-      {\subst{{x}{f}}{{x}{h}}}
-    }
-\end{align*}
-Moreover, we get
-\begin{equation*}
-\inference
-  { \jterm{\Gamma}{A}{x}
-    \jterm{{\Gamma}{A}}{P}{f}
-    \jterm{{{\Gamma}{A}}{P}}{Q}{g}
-    }
-  { \jtermeq
-      {\Gamma}
-      {\subst{x}{{f}{Q}}}
-      {\subst{x}{{f}{g}}}
-      {\subst{{x}{f}}{{x}{g}}}
-    }
-\end{equation*}
-\end{rmk}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Joining the theory of projections with the theory of substitution}
-\label{sec:esystem-equalities}
-
-\begin{comment}
-We have already seen that empty families, extension, weakening, substitution
-and the identity terms are compatible with each other. There are, however, a
-few more judgmental equalities explaining what happens when
-\begin{enumerate}
-\item something is weakend by the empty family. This is used to consider a term
-      $x$ of $A$ in context $\Gamma$ as a morphism from $\emptyf$ to $A$ in
-      context $\Gamma$;
-\item something is weakened by an extended family. This is Currying for weakening;
-\item something is first weakened by a family $A$ in context $\Gamma$ and then
-      substituted by a term of $A$. This is used to show that composition is
-      compatible with itself (i.e. that composition is associative);
-\item the identity terms are joining the party. Identity terms are empty contexts
-      in certain models, so they should behave accordingly.
-\end{enumerate}
-In other words, the rules seem to not fit in the scheme that all operations are
-compatible with each other, but in fact each of these rules plays a role in
-completing the scheme (work out further).
-\end{comment}
-
-\begin{comment}
-In this section we extend the judgmental equality relation in such a way that
-the empty context, extension, weakening, substitution and the identity terms
-all become compatible with each other. Since we have five operations, there
-are going to be $25$ finite sets of rules explaining how each operation is
-compatible with each of the operations.
-
-In addition to these $25$ finite sets
-of compatibility rules, there are also two sets of rules explaining the
-cancellation properties of weakening and substitution and of the identity terms.
-
-A typical compatibility axiom captures the idea that a diagram like
-\begin{equation*}
-\begin{tikzcd}[column sep=huge]
-\jfam{\Gamma}{\blank} 
-  \ar{r}{C\mapsto\ctxwk{B}{C}} 
-  \ar{d}[swap]{C\mapsto\ctxwk{A}{C}} 
-& \jfam{{\Gamma}{B}}{\blank} 
-  \ar{d}{Q\mapsto\ctxwk{A}{Q}}
-  \\
-\jfam{{\Gamma}{A}}{\blank} 
-  \ar{r}[swap]{P\mapsto\ctxwk{{A}{B}}{P}} 
-& \jfam{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\blank}
-\end{tikzcd}
-\end{equation*}
-in the case of two consecutive weakenings, or
-\begin{equation*}
-\begin{tikzcd}[column sep=huge]
-\jfam{{\Gamma}{A}}{\blank} 
-  \ar{d}[swap]{Q\mapsto\subst{x}{Q}} 
-  \ar{r}{Q\mapsto\ctxwk{P}{Q}} 
-& \jfam{{{\Gamma}{A}}{P}}{\blank} 
-    \ar{d}{R\mapsto\subst{x}{R}}
-  \\ 
-\jfam{\Gamma}{\blank} 
-  \ar{r}[swap]{B\mapsto\ctxwk{\subst{x}{P}}{B}} 
-& \jfam{{\Gamma}{\subst{x}{P}}}{\blank}
-\end{tikzcd}
-\end{equation*}
-in the case of a combination of weakening and substitution, commutes. We won't
-actually draw any of these diagrams in this section, but readers are encouraged
-to do so if they think it can help them understanding the axioms. Another thing
-one could keep in mind with the axioms that involve weakening is that all of
-them become completely obvious when weakening is made notationally invisible.
-Of course, we have an explicit notation for weakening on purpuse, but it is the
-usual practice of type theorists to make weakening invisible.
-
-The table below indicates where
-we can find the axioms of each of the combinations of operators. The reference
-at the row of operation $X$ and the column of operation $Y$ refers to the
-axiom which explains $Y$ distributes over the arguments of $X$ when $Y$ is
-applied after $X$. The empty context has no arguments, whereas extension,
-and weakening have two arguments which are both of the family-kind, substitution
-has two arguments of which one is a term and the other is a family and
-idenitity terms have one argument, a family.
-\begin{center}
-\begin{tabular}{r|rrrr}
-& ext & wk & subst & id\\
-\hline
-extension & \autoref{comp-ee}* &  & \cmark & \cmark \\
-weakening & \autoref{comp-we}* & \autoref{comp-ww} & \autoref{comp-sw} & \autoref{comp-wi} \\
-substitution & \autoref{comp-se}* & \autoref{comp-ws} & \autoref{comp-ss} & \autoref{comp-si} \\
-identity terms & \cmark & \autoref{comp-wi} & \autoref{comp-si} & \xmark
-\end{tabular}
-\end{center}
-Apart from references to sections where axioms are given, we also see checkmarks,
-crosses, marked references and duplicated references in this table. 
-The checkmarks indicate that the desired property
-can be proved and need not be axiomatized. To clarify those:
-\begin{itemize}
-\item The identity term of the empty context is the unique term of the empty
-family. Recall that we already have axiomatized in \autoref{empty} that every
-term of the empty family is judgmentally equal to $\emptytm$.
-\item The identity term of an extension can be explained by means of the notion
-of \emph{extension on terms}. This notion is defined in \autoref{extension-on-terms}
-using only identity terms and substitution. With extension on terms we can
-combine a term $\jterm{\Gamma}{A}{x}$ and a term $\jterm{\Gamma}{\subst{x}{P}}{u}$
-into a term $\jterm{\Gamma}{{A}{P}}{\tmext{x}{u}}$, which may be seen as the
-pair consisting of $x$ and $u$. Since it is a defined notion, we will
-be able to derive lots of properties for extension of terms. One of them explains
-how we can see the identity term of an extension as the extension of two certain
-terms, see \autoref{comp-ie}.
-\item Using extension on terms we can also make sense of `fibers at an extension'
-because that would exactly be the fiber at an extended term. We will see in
-\autoref{comp-es} that, for instance, $\subst{\tmext{x}{u}}{Q}\jdeq\subst{u}{{x}{Q}}$
-for any family $\jfam{{{\Gamma}{A}}{P}}{Q}$ and any two terms $\jterm{\Gamma}{A}{x}$
-and $\jterm{\Gamma}{\subst{x}{P}}{u}$. That's a form of currying.
-\end{itemize}
-The spots with a \xmark simply make no sense 
-{\color{red}(or do they, when the term model
-is better understood (by me)?)}. The marked references point to
-axioms which \emph{do} have a variant for terms, but those require the notion
-of extension on terms, can be proved and therefore need not be included in our
-list of compatibility axioms. In each of these sections, we give a forward
-reference to the lemma in which this variant is stated and proved.
-\end{comment}
-
-\subsubsection{Fibers of weakenings}\label{comp-sw}
-The following rules assert what happens when we first weaken by a family
-$Q$ over $P$ in context $\ctxext{\Gamma}{A}$ and then substitute by a term
-$x$ of $A$:
-\begin{align}
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jfam{{{\Gamma}{A}}{P}}{Q}
-    \jfam{{{\Gamma}{A}}{P}}{R}
-    }
-  { \jfameq
-      {{{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}}
-      {\subst{x}{\ctxwk{Q}{R}}}
-      {\ctxwk{\subst{x}{Q}}{\subst{x}{R}}}
-    }
-  \label{comp-sw-c}
-  \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jfam{{{\Gamma}{A}}{P}}{Q}
-    \jfam{{{{\Gamma}{A}}{P}}{R}}{S}
-    }
-  { \jfameq
-      {{{{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}}{\subst{x}{\ctxwk{Q}{R}}}}
-      {\subst{x}{\ctxwk{Q}{S}}}
-      {\ctxwk{\subst{x}{Q}}{\subst{x}{S}}}
-    }
-  \label{comp-sw-f}
-  \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jfam{{{\Gamma}{A}}{P}}{Q}
-    \jterm{{{{\Gamma}{A}}{P}}{R}}{S}{k}
-    }
-  { \jtermeq
-      {{{{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}}{\subst{x}{\ctxwk{Q}{R}}}}
-      {\subst{x}{\ctxwk{Q}{S}}}
-      {\subst{x}{\ctxwk{Q}{k}}}
-      {\ctxwk{\subst{x}{Q}}{\subst{x}{k}}}
-    }
-  \label{comp-sw-t}
-\end{align}
-
-\begin{rmk}
-As an important special case of these inference rules we have the following
-valid inference rules:
-\begin{align*}
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jfam{{\Gamma}{A}}{P}
-    \jfam{{\Gamma}{A}}{Q}
-    }
-  { \jfameq
-      {{\Gamma}{\subst{x}{P}}}
-      {\subst{x}{\ctxwk{P}{Q}}}
-      {\ctxwk{\subst{x}{P}}{\subst{x}{Q}}}
-    }
-  \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jfam{{\Gamma}{A}}{P}
-    \jfam{{{\Gamma}{A}}{Q}}{R}
-    }
-  { \jfameq
-      {{{\Gamma}{\subst{x}{P}}}{\subst{x}{\ctxwk{P}{Q}}}}
-      {\subst{x}{\ctxwk{P}{R}}}
-      {\ctxwk{\subst{x}{P}}{\subst{x}{R}}}
-    }
-  \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jfam{{\Gamma}{A}}{P}
-    \jterm{{{\Gamma}{A}}{Q}}{R}{h}
-    }
-  { \jtermeq
-      {{{\Gamma}{\subst{x}{P}}}{\subst{x}{\ctxwk{P}{Q}}}}
-      {\subst{x}{\ctxwk{P}{R}}}
-      {\subst{x}{\ctxwk{P}{h}}}
-      {\ctxwk{\subst{x}{P}}{\subst{x}{h}}}
-    }
-\end{align*}
-Moreover, we get
-\begin{equation*}
-\inference
-  { \jterm{\Gamma}{A}{x}
-    \jfam{{\Gamma}{A}}{P}
-    \jterm{{\Gamma}{A}}{Q}{g}
-    }
-  { \jtermeq
-      {{\Gamma}{\subst{x}{P}}}
-      {\subst{x}{\ctxwk{P}{Q}}}
-      {\subst{x}{\ctxwk{P}{g}}}
-      {\ctxwk{\subst{x}{P}}{\subst{x}{g}}}
-    }
-\end{equation*}
-\end{rmk}
-
-\subsubsection{Weakenings of fibers}
-\label{comp-ws}
-The following inference rules explain what happens when we first weaken by a
-term $g$ of $Q$ in context $\ctxext{\Gamma}{B}$ and then weaken by a family
-$A$ over $\Gamma$.
-\begin{align}
-& \inference
-  { \jfam{\Gamma}{A}
-    \jterm{{\Gamma}{B}}{Q}{g}
-    \jfam{{{\Gamma}{B}}{Q}}{R}
-    }
-  { \jfameq
-      {{{\Gamma}{A}}{\ctxwk{A}{B}}}
-      {\ctxwk{A}{\subst{g}{R}}}
-      {\subst{\ctxwk{A}{g}}{\ctxwk{A}{R}}}
-    }
-  \label{comp-ws-c}
-  \\
-& \inference
-  { \jfam{\Gamma}{A}
-    \jterm{{\Gamma}{B}}{Q}{g}
-    \jfam{{{{\Gamma}{B}}{Q}}{R}}{S}
-    }
-  { \jfameq
-      {{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{\subst{g}{R}}}}
-      {\ctxwk{A}{\subst{g}{S}}}
-      {\subst{\ctxwk{A}{g}}{\ctxwk{A}{S}}}
-    }
-  \label{comp-ws-f}
-  \\
-& \inference
-  { \jfam{\Gamma}{A}
-    \jterm{{\Gamma}{B}}{Q}{g}
-    \jterm{{{{\Gamma}{B}}{Q}}{R}}{S}{k}
-    }
-  { \jtermeq
-      {{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{\subst{g}{R}}}}
-      {\ctxwk{A}{\subst{g}{S}}}
-      {\ctxwk{A}{\subst{g}{k}}}
-      {\subst{\ctxwk{A}{g}}{\ctxwk{A}{k}}}
-    }
-  \label{comp-ws-t}
-\end{align}
-
-\begin{rmk}
-As an important special case of these inference rules we have the following 
-valid inference rules:
-\begin{align*}
-& \inference
-  { \jfam{\Gamma}{A}
-    \jterm{\Gamma}{B}{y}
-    \jfam{{\Gamma}{B}}{Q}
-    }
-  { \jfameq
-      {{\Gamma}{A}}
-      {\ctxwk{A}{\subst{y}{Q}}}
-      {\subst{\ctxwk{A}{y}}{\ctxwk{A}{Q}}}
-    }
-  \\
-& \inference
-  { \jfam{\Gamma}{A}
-    \jterm{\Gamma}{B}{y}
-    \jfam{{{\Gamma}{B}}{Q}}{R}
-    }
-  { \jfameq
-      {{{\Gamma}{A}}{\ctxwk{A}{\subst{y}{Q}}}}
-      {\ctxwk{A}{\subst{y}{R}}}
-      {\subst{\ctxwk{A}{y}}{\ctxwk{A}{R}}}
-    }
-  \\
-& \inference
-  { \jfam{\Gamma}{A}
-    \jterm{\Gamma}{B}{y}
-    \jterm{{{\Gamma}{B}}{Q}}{R}{h}
-    }
-  { \jtermeq
-      {{{\Gamma}{A}}{\ctxwk{A}{\subst{y}{Q}}}}
-      {\ctxwk{A}{\subst{y}{R}}}
-      {\ctxwk{A}{\subst{y}{h}}}
-      {\subst{\ctxwk{A}{y}}{\ctxwk{A}{h}}}
-    }
-\end{align*}
-Moreover, we get
-\begin{equation*}
-\inference
-  { \jfam{\Gamma}{A}
-    \jterm{\Gamma}{B}{y}
-    \jterm{{\Gamma}{B}}{Q}{g}
-    }
-  { \jtermeq
-      {{\Gamma}{A}}
-      {\ctxwk{A}{\subst{y}{Q}}}
-      {\ctxwk{A}{\subst{y}{g}}}
-      {\subst{\ctxwk{A}{y}}{\ctxwk{A}{g}}}
-    }
-\end{equation*}
-\end{rmk}
-
-\subsubsection{The cancellation property of weakening and substitution}
-\label{cancellation-ws}
-The judgmental equalities we're about to describe assert that substituting a term
-in the weakening a thing gives you the thing back. In the case of contexts we get that each fiber
-$\subst{x}{\ctxwk{A}{B}}$ is just $B$ and in the case of terms we get 
-that $\ctxwk{A}{y}$ is the constant function
-mapping everything to $y:B$. Thus, these rules actually establish the weakening
-as the weakening. After stating the rules we will describe what it means to
-compose context morphisms (terms of weakened contexts).
-\begin{align}
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{\Gamma}{B}
-    \jterm{\Gamma}{A}{x}
-    }
-  { \jfameq
-      {\Gamma}
-      {\subst{x}{\ctxwk{A}{B}}}
-      {B}
-    }
-  \label{cancellation-ws-c}\\
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{{\Gamma}{B}}{Q}
-    \jterm{\Gamma}{A}{x}
-    }
-  { \jfameq{{\Gamma}{B}}{\subst{x}{\ctxwk{A}{Q}}}{Q}
-    }
-  \label{cancellation-ws-f}\\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jterm{{\Gamma}{B}}{Q}{g}
-    }
-  { \jtermeq{{\Gamma}{B}}{Q}{\subst{x}{\ctxwk{A}{g}}}{g}
-    }
-  \label{cancellation-ws-t}
-\end{align}
-
-\subsubsection{The identity term of a substituted family}
-\label{comp-si}
-The identity term of a substituted family is the substitution of the identity term
-\begin{equation}
-\inference
-  { \jterm{\Gamma}{A}{x}
-    \jfam{{{\Gamma}{A}}{P}}{Q}
-    }
-  { \unfoldall{\jhomeq
-      {{\Gamma}{\subst{x}{P}}}
-      {\subst{x}{Q}}
-      {\subst{x}{Q}}
-      {\subst{x}{\idtm{Q}}}
-      {\idtm{\subst{x}{Q}}}
-    }}
-  \label{comp-si-t}
-\end{equation}
-
-\begin{rmk}
-An important special case is the judgmental equality
-\begin{equation*}
-\unfoldall{\jhomeq
-      {\Gamma}
-      {\subst{x}{P}}
-      {\subst{x}{P}}
-      {\subst{x}{\idtm{P}}}
-      {\idtm{\subst{x}{P}}}
-    }
-\end{equation*}
-for a family $\jfam{{\Gamma}{A}}{P}$.
-\end{rmk}
-
-\subsubsection{The cancellation properties of identity terms}
-\label{cancellation-i}
-Identity terms are determined by their behavior with respect to substitution combined with
-weakening. The identity terms will also be subject to compatibility rules.
-\begin{align}
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    }
-  { \jtermeq{\Gamma}{A}{\subst{x}{\idtm{A}}}{x}
-    }
-  \label{cancellation-si}\\
-& \inference
-  { \jfam{{\Gamma}{A}}{P}
-    }
-  { \jfameq{{\Gamma}{A}}{\subst{\idtm{A}}{\ctxwk{A}{P}}}{P}
-    }
-  \label{precomp-idtm-c}
-  \\
-& \inference
-  { \jfam{{{\Gamma}{A}}{P}}{Q}
-    }
-  { \jfameq{{{\Gamma}{A}}{P}}{\subst{\idtm{A}}{\ctxwk{A}{Q}}}{Q}
-    }
-  \label{precomp-idtm-f}
-  \\
-& \inference
-  { \jterm{{{\Gamma}{A}}{P}}{Q}{g}
-    }
-  { \jtermeq
-      {{{\Gamma}{A}}{P}}
-      {Q}
-      {\subst{\idtm{A}}{\ctxwk{A}{g}}}
-      {g}
-    }
-  \label{precomp-idtm-t}
-\end{align}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{The possiblity of types in the theory of contexts, families and
-terms}
-\label{types}
 We have deliberately not spoken of types so far because we have taken the point
@@
 
-\begin{comment}
-We envision a
-categorical interpretation of the theory of contexts, families and terms where 
-contexts are interpreted as finite paths of arrows belonging to a predetermined
-class. The empty family would be such a path of length $0$ whereas the types
-would be such paths of length $1$.
-\end{comment}
-
 We add rules asserting that a weakened type is again a type that 

### 2014-10-28 Rules/rules_empty_context_and_families.tex
reorganization

--- 2014-09-20 article-models-tt.tex
@@
-\section{The generalized algebraic theory of contexts, families and terms (the E-system)}\label{tt}
-In this section we give a description of dependent type theory before type
-constructors. Apart from contexts, families and terms -- which provide for the
-core of the language of dependent logic -- the basic ingredients
-of this theory will be the operations of extension on context and on families, 
-and empty context and empty families, weakening, substitution
-and the identity terms. The resulting theory can be seen as a manifestation of the 
-structure underlying dependent type theory.
-
-We will
-formulate the theory of contexts, families and terms in such a way that contexts aren't defined
-to be lists of variable declarations. The variable-free (a.k.a.~\emph{name-free}) approach 
-we take here is rather different than those appearing in 
-\cite{hofmann1995extensional,TheBook} but it has appeared in the work of Coquand
-and in \cite{Dybjer1996}.
-The main reason we don't let variable declarations in is that we don't see them 
-in the internal models either. This way we also set out to a more algebraic 
-approach of type theory and higher category theory. Thirdly, we will not have to
-be burdened with superficial comments about variables being bounded or not, or 
-fresh or free or not occuring at all.
-
-In the current presentation, contexts can be seen as binary
-planar trees of which the leaves are (families of) contexts. 
-The judgmental equality relation on contexts is an equivalence relation which 
-expresses that binary planar
-trees of contexts are judgmentally equal if their leaves are, taking only
-(the isomorphism class of) 
-the order of the leaves into account \emph{and not the actual shape of the three}.
-The intuition behind this equivalence relation is indeed that unstructured
-(i.e.~unbracketed) lists such as the lists of variable declarations which
-usually appear in type theoretical syntax, may be regarded as contexts.
-
-Besides contexts, families and terms there will also be a notion of `type in
-a context', which one may assume.
-A context is eligible to be a \emph{closed type} and over a context, a family
-may be a type. We have the intuition that `being a type'
-expresses the property of `being atomic or irreducible'. With a notion of types,
-one has the means to require a stratification by which every context is a finite
-extension of types starting at a closed type, and every family is likewise a
-finite extension of types. The idea behind this is to recover the theory of
-\cite{Garner2014} {\color{red}(and possibly B-systems?)}. We stress that
-we will not include axioms asserting such a stratification in the E-system
-and we leave discussion on the possibility of a typing judgment until the
-end of this section.
-
-With families of contexts being the principal things of
-study, we have added a bit more generality to the theory, compared to the
-theory of categories with families presented in \cite{Dybjer1996}. Here are
-a couple of reasons for this extra generality:
-\begin{enumerate}
-\item We want to be very sure that we can interpret all the rules of type theory
-in a straightforward manner in any context.
-\item We do want do end up with a categorical structure on the contexts, derived
-from the theory by requiring that a morphism from $\Gamma$ to $\Delta$ is a
-term of the weakening of $\Delta$ by $\Gamma$ in context $\Gamma$. Thus, we
-must be able to lift the context $\Delta$ to become a family of contexts over
-$\Gamma$ and it is not reasonable to expect that this results in a type.
-\item It is very easy in this setting to assert that an operation has an action
-on very high levels of families, i.e.~when $F$ is an operation taking things in
-context $\Gamma$ to things in context $\Delta$, then it takes a family $A$ in
-context $\Gamma$ to $F(A)$ in context $\Delta$; a family $P$ in context
-$\ctxext{\Gamma}{A}$ to a family $F(P)$ in context $\ctxext{\Delta}{(F(A))}$,
-and so on. The reason it becomes simple is that we don't have to explain what
-`and so on' means.
-\end{enumerate}
-
-We will formulate fairly strict rules governing the judgmental equalities,
-expressing that extension, weakening and substitution are combatible with
-each other in a judgmental manner. This does not, however, diminish the role
-of isomorphisms or of homotopies could play in the theory once identity types
-are added. Indeed, types could still have non-trivial identity relations and
-the category of types in a certain context could genuinely display higher
-categorical structure, or so we conjecture.
-
-Much of the rules we state are just compatibility rules of extension, weakening
-and substitution with each other. In a way, these rules assert that our contexts
-are just structureless lists of contexts and that likewise terms are structureless
-lists of terms. They are structureless in the sense that the order in which
-they are formed by pairing up is irrelevant. We note that this causes complications
-in the traditional way that categorical sematics of type theory is implemented,
-where contexts become objects of the category which is supposed to model type
-theory. The reason for this is that context extension will not satisfy all the
-compatibility rules we're about to state. The first step to resolving this is taking
-the types in the empty context as the objects.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Judgments and valid inference rules}\label{judgments}
-The theory we describe here is a theory of contexts, families of
-contexts and terms thereof. The families of contexts are by some authors called
-dependent contexts, but they are handled a bit differently here because they
-become the primary object of study. Dependent contexts can be types; they could
-be seen as atomic or indecomposable dependent contexts.
-
-Thus we make eight kinds of judgments: ``$\Gamma$ is a context'',
-``$A$ is a family of contexts over $\Gamma$'', ``$A$ is a type in context $\Gamma$''
-and ``$x$ is a term of the family $A$ of contexts over $\Gamma$''. The other four
-judgments are for judgmental equality.
-\begin{align*}
-\jalign\jctx{\Gamma} 
-& \jalign\jctxeq{\Gamma}{\Gamma'}
-  \\
-\jalign\jfam{\Gamma}{A} 
-& \jalign\jfameq{\Gamma}{A}{B}
-  \\
-\jalign\jterm{\Gamma}{A}{x} 
-& \jalign\jtermeq{\Gamma}{A}{x}{y}.
-\end{align*}
-
-Strictly speaking, we have three different judgmental equalities in play and one
-could request for a notational difference to signify that fact. For instance,
-we could denote the judgmental equalities of contexts, families and terms by
-$\jdeq_c$, $\jdeq_f$ and $\jdeq_t$ respectively. It will, however, always be
-clear which of the three kinds of judgmental equality is meant when we assert
-a judgmental equality and therefore we shall not bother to make this notational
-distinction.
-
-We note that what we call families over contexts
-here could also have been named dependent contexts or telescopes, see
-\cite{deBruijn1991,hofmann1995extensional}. The term family is in agreement
-with the terminology scheme of \cite{TheBook}, though the reader should be
-warned that the notion of familie means something slightly different there than
-it does here.
-
-The theory of contexts, families and terms shall be described by means of a
-declaration of valid inference rules. An inference rule is an expression of the
-form
-\begin{equation*}
-\inference{\mathcal{H}_1\quad\cdots\quad\mathcal{H}_n}{\mathcal{J}}
-\end{equation*}
-where each of $\mathcal{H}_1$, ..., $\mathcal{H}_n$ and $\mathcal{J}$ is 
-judgment. The judgments $\mathcal{H}_1$, ..., $\mathcal{H}_n$ are called the
-hypotheses and the judgment $\mathcal{J}$ is the conclusion of the inference
-rule. When a judgment is the conclusion of a valid inference rule, we also say
-that the judgment is valid or well-formed. In fact, the only way to construct
-new judgments is via the valid inference rules we describe in this section.
-
-The class of valid inference rules is closed under composition and under
-rearranging the hypotheses. That
-means that if
-\begin{equation*}
-\inference{\mathcal{I}_{1k}\quad\cdots\quad\mathcal{I}_{m_kk}}{\mathcal{H}_k}
-\end{equation*}
-and
-\begin{equation*}
-\inference{\mathcal{H}_1\quad\cdots\quad\mathcal{H}_n}{\mathcal{J}}
-\end{equation*}
-are valid inference rules for $1\leq k\leq n$, then
-\begin{equation*}
-\inference{\mathcal{I}_{11}\quad\cdots\quad\mathcal{I}_{m_11}\quad\cdots\quad\mathcal{I}_{1n}\quad\cdots\quad\mathcal{I}_{m_nn}}{\mathcal{J}}
-\end{equation*}
-is also a valid inference rule. Also, in a valid inference rule 
-duplicated hypotheses may always be contracted and hypotheses may always be 
-reordered; the result is still considered a valid inference rule.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{The basic rules for judgmental equality}
-The rules for judgmental equality establish that it is an equivalence relation
-in all three cases (contexts, families and terms). Thus, the following inference
-rules shall be required to be valid:
-\bgroup\small
-\begin{align*}
-& \inference
-  { \jctx{\Gamma}
-    }
-  { \jctxeq{\Gamma}{\Gamma}
-    } 
-& & \inference
-    { \jctxeq{\Gamma}{\Delta}
-      }
-    { \jctxeq{\Delta}{\Gamma}
-      } 
-& & \inference
-    { \jctxeq{\Gamma}{\Delta}
-      \jctxeq{\Delta}{\greek{E}}
-      }
-    { \jctxeq{\Gamma}{\greek{E}}
-      }
-    \\
-& \inference
-  { \jfam{\Gamma}{A}
-    }
-  { \jfameq{\Gamma}{A}{A}
-    } 
-& & \inference
-    { \jfameq{\Gamma}{A}{B}
-      }
-    { \jfameq{\Gamma}{B}{A}
-      }
-& & \inference
-    { \jfameq{\Gamma}{A}{B}
-      \jfameq{\Gamma}{B}{C}
-      }
-    { \jfameq{\Gamma}{A}{C}
-      }
-    \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    }
-  { \jtermeq{\Gamma}{A}{x}{x}
-    }
-& & \inference
-    { \jtermeq{\Gamma}{A}{x}{y}
-      }
-    { \jtermeq{\Gamma}{A}{y}{x}
-      }
-& & \inference
-    { \jtermeq{\Gamma}{A}{x}{y}
-      \jtermeq{\Gamma}{A}{y}{z}
-      }
-    { \jtermeq{\Gamma}{A}{x}{z}
-      }
-\end{align*}
-\egroup
-
-The following convertibility rules are responsible for the strictness
-of judgmental equality, which sets it apart from equivalences or identifications:
-\begin{align*}
-& \inference
-  { \jctxeq{\Gamma}{\Delta}
-    \jfam{\Gamma}{A}
-    }
-  { \jfam{\Delta}{A}
-    }
-& & \inference
-    { \jctxeq{\Gamma}{\Delta}
-      \jfameq{\Gamma}{A}{B}
-      }
-    { \jfameq{\Delta}{A}{B}
-      }
-    \\
-& \inference
-  { \jctxeq{\Gamma}{\Delta}
-    \jterm{\Gamma}{A}{x}
-    }
-  { \jterm{\Delta}{A}{x}
-    }
-& & \inference
-    { \jctxeq{\Gamma}{\Delta}
-      \jtermeq{\Gamma}{A}{x}{y}
-      }
-    { \jtermeq{\Delta}{A}{x}{y}
-      }
-    \\
-& \inference
-  { \jfameq{\Gamma}{A}{B}
-    \jterm{\Gamma}{A}{x}
-    }
-  { \jterm{\Gamma}{B}{x}
-    }
-& & \inference
-    { \jfameq{\Gamma}{A}{B}
-      \jtermeq{\Gamma}{A}{x}{y}
-      }
-    { \jtermeq{\Gamma}{B}{x}{y}
-      }
-\end{align*}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{The operations of the theory of contexts, families and terms}
-In this subsection we introduce the notions of
-extension on context and on families, the empty context and the empty families, 
-weakening, 
-substitution and identity terms.
-
-Weakening and substitution are both
-operations of the theory of contexts, families and terms and therefore they will
-have three levels of actions: on contexts (at the appropriate level), families
-and terms.
-
-\begin{comment}
-Generally, 
-each operation should
-have an action on contexts, on families and on terms at the `appropriate level'
-and should be compatible with each of the operations we introduce here. In
-that will become more practical after the theory of contexts, families and terms
-is formulated, this means that when we will write down a type theory that
-interprets itself in \autoref{ttmodels}, each of the notions of
-extension, empty family, weakening, substitution and identity terms are going to be internalized
-as terms of the theory and all terms are going to be compatible with these.
-That vague idea will be our guiding light.
-
-So in principle, the introductions of each of the empty context, extension,
-weakening, substitution and identity terms should come in three parts. In 
-practice, they do not all come in three parts.
-
-Apart from the empty context, there will indeed be an empty family and a term
-of the empty family (which will be deemed unique). But extension will
-not have an explicit action `at the level of families' since extension
-is used to define higher families and the action would have to be trivial
-anyway. The empty context and extension provide the essentials of the rest of
-the theory in the sense that higher families are families over extensions,
-contexts are families over the empty context and the `action on families' of
-any operation applied to families over the empty context is the same as the
-original operation. Thus, when we have introduced the operations of weakening
-and substitution we will see that they also have an `action on the terms of 
-contexts'. We will state the relevant judgmental equality rules for the empty
-context in this section rather than in the next, since they are required to
-state that the action on families of an operation is compatible with the
-original operation.
-\end{comment}
-
-We note upfront that both context extension and family extension can also be
-seen as operations of contexts, families and terms, even though we will not
-introduce them as such. The reason is that extension is used to define what it
-means to be a family over a family over a context (namely, a family over an
-extended context). As such, it will have a trivial action on families and terms
-which will not need to be introduced. We will come back to this point in
-\autoref{extension}.
-
-\subsection{Extension}
-\label{extension}
-The operations of extension enable one to consider families over
-families and higher level families alike. We will need both a context extension
-operation and a family extension operation.
-
-The inference rules introducing context and family extension are as follows:
-\begin{align}
-& \inference
-  { \jfam{\Gamma}{A}
-    }
-  { \jctx{\ctxext{\Gamma}{A}}
-    }
-& & \inference
-    { \jctxeq{\Gamma}{\Delta}
-      \jfameq{\Gamma}{A}{B}
-      }
-    { \jctxeq{\ctxext{\Gamma}{A}}{\ctxext{\Delta}{B}}
-      }
-    \\
-& \inference
-  { \jfam{{\Gamma}{A}}{P}
-    }
-  { \jfam{\Gamma}{\ctxext{A}{P}}
-    }
-& & \inference
-    { \jfameq{\Gamma}{A}{B} 
-      \jfameq{{\Gamma}{A}}{P}{Q}
-      }
-    { \jfameq{\Gamma}{\ctxext{A}{P}}{\ctxext{B}{Q}}
-      }
-\end{align}
-The extension operation may also be defined to act on families over families
-and terms thereof. For instance, When $\jfam{{\Gamma}{A}}{P}$ is a family,
-we would get a family $\jfam{{\Gamma}{A}}{{\Gamma}{P}}$ and when
-$\jterm{{\Gamma}{A}}{P}{f}$ is a term, we would get a term
-$\jterm{{\Gamma}{A}}{{\Gamma}{P}}{\ctxext{\Gamma}{f}}$. Then there would be
-axioms stating the judgmental equalities $\ctxext{\Gamma}{P}\jdeq P$ and
-$\ctxext{\Gamma}{f}\jdeq f$. Thus, it is of no actual use to introduce the
-action of context extension on families and terms.
-
-In \autoref{extension-on-terms} we will also define an extension operation
-on terms, which defines a term $\jterm{\Gamma}{{A}{P}}{\tmext{x}{u}}$ for
-every $\jterm{\Gamma}{A}{x}$ and $\jterm{\Gamma}{\subst{x}{P}}{u}$. This
-operation is defined by using the substitution operation and identity terms.
-
-\subsubsection{Associativity of extension}
-\label{comp-ee}
-The inference rules asserting that extension is compatible with itself assert
-that contexts are unstructured lists of type declarations. This rule is
-unavoidable if we want that for a family $A$ in context $\Gamma$, a family over
-$A$ is the same thing as a family over $\ctxext{\Gamma}{A}$. 
-\begin{align}
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{{\Gamma}{A}}{P}
-    }
-  { \jctxeq{\ctxext{{\Gamma}{A}}{P}}{\ctxext{\Gamma}{{A}{P}}}
-    }
-  \\
-& \inference
-  { \jfam{{\Gamma}{A}}{P}
-    \jfam{{{\Gamma}{A}}{P}}{Q}
-    }
-  { \jfameq{\Gamma}{\ctxext{A}{{P}{Q}}}{\ctxext{{A}{P}}{Q}}
-    }
-  \label{comp-ee-c}
-\end{align}
-Note that we may take the point of view that extension distributes over
-extensions. If we want to extend an extended family $\ctxext{P}{Q}$ over
-$\ctxext{\Gamma}{A}$ by the family $A$, the result is the extension of (1)
-the extension of $P$ by $A$ and (2) the extension of $Q$ by $A$. The latter is
-simply $Q$ since extension acts trivially at the level of families (and terms).
-
 \subsection{The empty context and the empty families}
@@
     }
-  { \jfam{\emptyc}{\Gamma}
+  { \jfam{\emptyc}{i(\Gamma)}
     } 
-& & \inference
-    { \jfam{\emptyc}{\Gamma}
-      }
-    { \jctx{\Gamma}
-      }
-    \\
+  &
 & \inference
@@
     }
-  { \jfameq{\emptyc}{\Gamma}{\Delta}
-    }
-& & \inference
-    { \jfameq{\emptyc}{\Gamma}{\Delta}
-      }
-    { \jctxeq{\Gamma}{\Delta}
+  { \jfameq{\emptyc}{i(\Gamma)}{i(\Delta)}
     }
@@
 identity terms (\autoref{identityterms}).
-\begin{comment}
-When contexts are viewed as statements, the empty context is the statement that
-asserts nothing. Likewise, in any context one can assert nothing and this is
-achieved by the empty family. 
-Since somebody who states nothing asserts it all right, there is a term of the empty
-family over any context. Moreover, any two terms of the empty family are always
-judgmentally equal.
-\begin{align}
-& \inference
-  { }
-  { \jctx{\emptyc}
-    }
-  \\
-& \inference
-  { \jctx{\Gamma}
-    }
-  { \jfam{\Gamma}{\emptyf[\Gamma]}
-    }
-  \\
-& \inference
-  { \jctx{\Gamma}
-    }
-  { \jterm{\Gamma}{\emptyf[\Gamma]}{\emptytm[\Gamma]}
-    }
-  \\
-& \inference
-  { \jterm{\Gamma}{\emptyf[\Gamma]}{x}
-    }
-  { \jtermeq
-      {\Gamma}
-      {\emptyf[\Gamma]}
-      {x}
-      {\emptytm[\Gamma]}
-    }
-  \label{comp-00-t}
-\end{align}
-
-The empty context and the empty families together with the operation of
-extension will provide several shortcuts for the rest
-of the theory of families and terms. This is (in part) obtained by requiring
-that a context is exactly the same thing as a family of contexts over the
-empty context and that judgmental equality of contexts is exactly the same
-as judgmental equality of families of contexts over the empty context.
-
-\label{comp-00}
-Because every context is a family of contexts over the empty context, 
-the empty context $\emptyc$ may be considered as a family of contexts over the empty
-context. When we do this, we get $\emptyf[\emptyc]$.
-\begin{equation}\label{eq:comp-00}
-\inference
-  { }
-  { \jfameq
-      {\emptyc}
-      {\emptyc}
-      {\emptyf[\emptyc]}
-    }
-\end{equation}
-In the future, we shall denote $\emptyf[\Gamma]$ by $\emptyf$. The above rule
-guarantees that this will not cause confusion. Likewise, we shall denote
-$\emptytm[\Gamma]$ by $\emptytm$.
-\end{comment}
 
@@
     }
-  { \jctxeq{\ctxext{\emptyc}{\Gamma}}{\Gamma}
+  { \jctxeq{\ctxext{\emptyc}{i(\Gamma)}}{\Gamma}
     }
@@
   \label{comp-e0-f}
+  \\
+& \inference
+  { \jfam{\emptyc}{A}
+    }
+  { \jfameq{\emptyc}{i(\ctxext{\emptyc}{A})}{A}
+    }
 \end{align}
@@
     }
-  { \jctxeq{(\ctxext{\Gamma}{A})^c}{(\ctxext{\Gamma}{A})^\famsym}
+  { \jctxeq{\ctxext{\Gamma}{A}}{\ctxext{i(\Gamma)}{A}}
     }
@@
 notations for context extension and family extension.
-
-\subsection{The theory of weakening}
-\label{weakening}
-When $A$ is a family in context $\Gamma$, the operation of weakening by $A$
-takes a family $B$ in context $\Gamma$ and provides a family $\ctxwk{A}{B}$
-in context $\ctxext{\Gamma}{A}$. The context family $\ctxwk{A}{B}$ can be seen
-as the constant family over $\ctxext{\Gamma}{A}$ with value $B$. This idea will
-be axiomatized in the cancellation property of weakening and substitution in
-\autoref{cancellation-ws}. In \autoref{morphisms} we will take the terms 
-$\unfold{\jhom{\Gamma}{A}{B}{f}}$ to be the morphisms of families from $A$ to 
-$B$. These will be at the heart of the categorical structure of the theory.
-
-The weakening operation acts on three levels: on contexts, on families and
-on terms. The `action on contexts' of weakening is the action we described
-above: it takes a family $B$ over $\Gamma$ to a family $\ctxwk{A}{B}$ over
-$\ctxext{\Gamma}{A}$; the `action on families' of weakening takes a family
-$Q$ over $\ctxext{\Gamma}{B}$ to a family $\ctxwk[\famsym]{A}{Q}$ over
-$\ctxext{{\Gamma}{A}}{\ctxwk{A}{B}}$; the `action on terms' of weakening takes
-a term $g$ of $Q$ to a term $\ctxwk[\tmsym]{A}{g}$ of $\ctxwk[\famsym]{A}{Q}$.
-\begin{align}
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{\Gamma}{B}
-    }
-  { \jfam{{\Gamma}{A}}{\ctxwk{A}{B}}
-    }
-& & \inference
-    { \jfameq{\Gamma}{A}{A'}
-      \jfameq{\Gamma}{B}{B'}
-      }
-    { \jfameq{{\Gamma}{A}}{\ctxwk{A}{B}}{\ctxwk{A'}{B'}}
-      }
-    \\
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{{\Gamma}{B}}{Q}
-    }
-  { \jfam{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk[\famsym]{A}{Q}}
-    }
-& & \inference
-    { \jfameq{\Gamma}{A}{A'}
-      \jfameq{{\Gamma}{B}}{Q}{Q'}
-      }
-    { \jfameq
-        {{{\Gamma}{A}}{\ctxwk{A}{B}}}
-        {\ctxwk[\famsym]{A}{Q}}
-        {\ctxwk[\famsym]{A'}{Q'}}
-      }
-    \\
-%& \inference
-%  { \jterm{\Gamma}{B}{y}
-%    }
-%  { \jterm{{\Gamma}{A}}{\ctxwk{A}{B}}{\ctxwk{A}{y}}
-%    }
-%& & \inference
-%    { \jtermeq{\Gamma}{B}{y}{y'}
-%      }
-%    { \jtermeq
-%        {{\Gamma}{A}}
-%        {\ctxwk{A}{B}}
-%        {\ctxwk{A}{y}}
-%        {\ctxwk{A'}{y'}}
-%      }
-%    \\
-& \inference
-  { \jfam{\Gamma}{A}
-    \jterm{{\Gamma}{B}}{Q}{g}
-    }
-  { \jterm{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk[\famsym]{A}{Q}}{\ctxwk[\tmsym]{A}{g}}
-    }
-& & \inference
-    { \jfameq{\Gamma}{A}{A'}
-      \jtermeq{{\Gamma}{B}}{Q}{g}{g'}
-      }
-    { \jtermeq
-        {{{\Gamma}{A}}{\ctxwk{A}{B}}}
-        {\ctxwk[\famsym]{A}{Q}}
-        {\ctxwk[\tmsym]{A}{g}}
-        {\ctxwk[\tmsym]{A'}{g'}}
-      }
-\end{align}
-
-\subsubsection{Weakening of an empty family}
-The following inference rules express that when the empty family is
-weakened, the result is the empty family.
-\label{comp-w0}
-\begin{align}
-%& \inference
-%  { \jctx{\Gamma}
-%    }
-%  { \jfameq{\Gamma}{\ctxwk{\Gamma}{\emptyc}}{\emptyf}
-%    }
-%  \\
-& \inference
-  { \jfam{\Gamma}{A}
-    }
-  { \jfameq{{\Gamma}{A}}{\ctxwk{A}{\emptyf}}{\emptyf}
-    }
-  \label{comp-w0-c}\\
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{\Gamma}{B}
-    }
-  { \jfameq
-    {{{\Gamma}{A}}{\ctxwk{A}{B}}}
-    {\ctxwk[\famsym]{A}{\emptyf}}
-    {\emptyf}
-    }
-  \label{comp-w0-f}
-\end{align}
-\begin{comment}
-Note that we do not have to axiomatize the action of weakening
-by $A$ on the term of the empty family, because the result will be in the empty 
-family and every term in the empty family is judgmentally equal to the term 
-$\emptytm$.
-\end{comment}
-
-Because a family over $\Gamma$ is the same as a family over 
-$\ctxext{\Gamma}{\emptyf}$ we can apply both the action on contexts and the
-action on families of weakening to a family $B$ over $\Gamma$. When we apply
-the action on families, we obtain a family $\ctxwk[\famsym]{A}{B}$ over the
-context $\ctxext{{\Gamma}{A}}{\ctxwk{A}{\emptyf}}$. However, since we have
-postulated the judgmental equalities $\ctxwk{A}{\emptyf}\jdeq\emptyf$ and
-$\ctxext{{\Gamma}{A}}{\emptyf}\jdeq\ctxext{\Gamma}{A}$, we see that we can
-compare $\ctxwk[\famsym]{A}{B}$ with $\ctxwk{A}{B}$. The following inference
-rule postulates that these two are judgmentally equal:
-\begin{equation}
-\inference
-{ \jfam{\Gamma}{A}
-  \jfam{\Gamma}{B}
-  }
-{ \jfameq{{\Gamma}{A}}{\ctxwk[\famsym]{A}{B}}{\ctxwk{A}{B}}
-  }
-\end{equation}
-Due to this rule, the action on contexts and the action on families of weakening
-are compatible with each other and consequently there can be no possible
-confusion when we omit the annotations $\famsym$ and $\tmsym$ alltogether. In
-the future, the weakening of a family $Q$ over $\ctxext{\Gamma}{B}$ shall
-be denoted just by $\ctxwk{A}{Q}$ and likewise the weakening of a term $g$ of
-$Q$ shall be denoted by $\ctxwk{A}{g}$.
-
-Because a family $B$ over $\Gamma$ can be treated as a family by the operation
-of weakening, weakening also acts on the terms of $B$. The weakening of a term
-$y$ of $B$ by $A$ can be seen as the constant term (or function) of the
-family $\ctxwk{A}{B}$.
-
-\subsubsection{Weakenings of extensions}
-\label{comp-we}
-The following rules assert that when an extended family is weakened, the
-weakening distributes over the extension factors.
-\begin{align}
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{{\Gamma}{B}}{Q}
-    }
-  { \jfameq
-      {\ctxext{\Gamma}{A}}
-      {\ctxwk{A}{\ctxext{B}{Q}}}
-      {\ctxext{\ctxwk{A}{B}}{\ctxwk{A}{Q}}}
-    }
-  \label{comp-we-c}
-  \\
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{{{\Gamma}{B}}{Q}}{R}
-    }
-  { \jfameq
-      {\ctxext{{\Gamma}{A}}{\ctxwk{A}{B}}}
-      {\ctxwk{A}{\ctxext{Q}{R}}}
-      {\ctxext{\ctxwk{A}{Q}}{\ctxwk{A}{R}}}
-    }
-  \label{comp-we-f}
-\end{align}
-When thinking of terms of $\ctxwk{A}{B}$ as morphisms of families from $A$ to
-$B$, this looks already like form of type theoretic choice. It is weaker in that
-it is not stated with function types, yet it is stronger in that it states a
-judgmental equality between two families.
-
-There is also a version of this property where an extended term is weakened.
-This variant is stated and proved in \autoref{comp-we-t}.
-
-\subsubsection{Weakening of weakenings}
-Weakening by a family $A$ in context $\Gamma$ brings things in context $\Gamma$
-to things in context $\ctxext{\Gamma}{A}$. Since we have all the ingredients of
-the theory of contexts, families and terms in the context $\Gamma$ as well it
-has in particular it's own weakening by families. Suppose we have a family
-$\jfam{{\Gamma}{B}}{Q}$ over $B$ in context $\Gamma$. Weakening by $Q$ brings
-things from context $\ctxext{\Gamma}{B}$ to $\ctxext{{\Gamma}{B}}{Q}$. Thus
-we can provide rules asserting what will happen when we first weaken by $Q$ and
-then (via the action on families) by $A$. We will require the following
-inference rules to be valid:
-\label{comp-ww}
-\begin{align}
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{{\Gamma}{B}}{Q}
-    \jfam{{\Gamma}{B}}{R}
-    }
-  { \jfameq
-      {{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{Q}}}
-      {\ctxwk{A}{{Q}{R}}}
-      {\ctxwk{{A}{Q}}{{A}{R}}}
-    }
-  \label{comp-ww-c}\\
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{{\Gamma}{B}}{Q}
-    \jfam{{{\Gamma}{B}}{R}}{S}
-    }
-  { \jfameq
-      {{{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{Q}}}{\ctxwk{A}{{Q}{R}}}}
-      {\ctxwk{A}{{Q}{S}}}
-      {\ctxwk{{A}{Q}}{{A}{S}}}
-    }
-  \label{comp-ww-f}\\
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{{\Gamma}{B}}{Q}
-    \jterm{{{\Gamma}{B}}{R}}{S}{k}
-    }
-  { \jtermeq
-      {{{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{Q}}}{\ctxwk{A}{{Q}{R}}}}
-      {\ctxwk{A}{{Q}{S}}}
-      {\ctxwk{A}{{Q}{k}}}
-      {\ctxwk{{A}{Q}}{{A}{k}}}
-    }
-  \label{comp-ww-t}
-\end{align}
-
-\begin{rmk}
-As an important special case of these inference rules we have the following
-valid inference rules:
-\begin{align*}
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{\Gamma}{B}
-    \jfam{\Gamma}{C}
-    }
-  { \jfameq
-      {{{\Gamma}{A}}{\ctxwk{A}{B}}}
-      {\ctxwk{A}{{B}{C}}}
-      {\ctxwk{{A}{B}}{{A}{C}}}
-    }
-  \\
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{\Gamma}{B}
-    \jfam{{\Gamma}{C}}{R}
-    }
-  { \jfameq
-      {{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{{B}{C}}}}
-      {\ctxwk{A}{{B}{R}}}
-      {\ctxwk{{A}{B}}{{A}{R}}}
-    }
-  \\
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{\Gamma}{B}
-    \jterm{{\Gamma}{C}}{R}{h}
-    }
-  { \jtermeq
-      {{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{{B}{C}}}}
-      {\ctxwk{A}{{B}{R}}}
-      {\ctxwk{A}{{B}{h}}}
-      {\ctxwk{{A}{B}}{{A}{h}}}
-    }
-\end{align*}
-Moreover, we have
-\begin{equation*}
-\inference
-  { \jfam{\Gamma}{A}
-    \jfam{\Gamma}{B}
-    \jterm{\Gamma}{C}{z}
-    }
-  { \jtermeq
-      {{{\Gamma}{A}}{\ctxwk{A}{B}}}
-      {\ctxwk{A}{{B}{C}}}
-      {\ctxwk{A}{{B}{z}}}
-      {\ctxwk{{A}{B}}{{A}{z}}}
-    }
-\end{equation*}
-\end{rmk}
-
-\subsubsection{Weakening by the empty family}
-Note that we can also weaken by the empty family over $\Gamma$.
-Weakening by the empty family $\emptyf$ over a context $\Gamma$ leaves families, 
-their terms, families over those families and terms of those unchanged:
-\label{comp-0w}
-\begin{align}
-& \inference
-  { \jfam{\Gamma}{B}
-    }
-  { \jfameq{\Gamma}{\ctxwk{\emptyf}{B}}{B}
-    }
-  \label{comp-0w-c}
-  \\
-& \inference
-  { \jfam{{\Gamma}{B}}{Q}
-    }
-  { \jfameq{{\Gamma}{B}}{\ctxwk{\emptyf}{Q}}{Q}
-    }
-  \label{comp-0w-f}
-  \\
-& \inference
-  { \jterm{{\Gamma}{B}}{Q}{g}
-    }
-  { \jtermeq{{\Gamma}{B}}{Q}{\ctxwk{\emptyf}{g}}{g}
-    }
-  \label{comp-0w-t}
-\end{align}
-
-\subsubsection{Currying for weakening}
-\label{comp-ew}
-The rules expressing that extension is compatible with weakening assert that
-weakening by an extension is the same thing as weakening twice in the
-appropriate way.
-\begin{align}
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{{\Gamma}{A}}{P}
-    \jfam{\Gamma}{B}
-    }
-  { \jfameq
-      {{{\Gamma}{A}}{P}}
-      {\ctxwk{\ctxext{A}{P}}{B}}
-      {\ctxwk{P}{{A}{B}}}
-    }
-  \label{comp-ew-c}\\
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{{\Gamma}{A}}{P}
-    \jfam{{\Gamma}{B}}{Q}
-    }
-  { \jfameq
-      {{{{\Gamma}{A}}{P}}{\ctxwk{P}{{A}{B}}}}
-      {\ctxwk{\ctxext{A}{P}}{Q}}
-      {\ctxwk{P}{{A}{Q}}}
-    }
-  \label{comp-ew-f}\\
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{{\Gamma}{A}}{P}
-    \jterm{{\Gamma}{B}}{Q}{g}
-    }
-  { \jtermeq
-      {{{{\Gamma}{A}}{P}}{\ctxwk{P}{{A}{B}}}}
-      {\ctxwk{P}{{A}{Q}}}
-      {\ctxwk{\ctxext{A}{P}}{g}}
-      {\ctxwk{P}{{A}{g}}}
-    } 
-  \label{comp-ew-t}
-\end{align}
-
-\begin{comment}
-\subsubsection{Weakening by a context}
-Since we have not required that there is an empty context with the property that
-a context is the same thing as a family in the empty context, we will have to
-explicitly introduce weakening by contexts. Since it will allways be clear
-whether something is weakened by a context or by a family, we shall use the
-same notation for weakening by contexts as for weakening by families. We shall
-have to repeat all the compatibility rules for weakening by families for
-weakening by contexts; there is no way around it. It is
-actually this sort of inconvenience that has originally led us to include an
-empty context and be done with it.
-
-In the following inference rules we assume that we have a context $\Gamma$ and
-that $\jctxeq{\Gamma}{\Gamma'}$ in the rules where $\Gamma'$ makes an appearence.
-\begin{align}
-& \inference
-  { \jctx{\Delta}
-    }
-  { \jfam{\Gamma}{\ctxwk{\Gamma}{\Delta}}
-    }
-& & \inference
-  { \jctxeq{\Delta}{\Delta'}
-    }
-  { \jfameq{\Gamma}{\ctxwk{\Gamma}{\Delta}}{\ctxwk{\Gamma'}{\Delta'}}
-    }
-  \\
-& \inference
-  { \jfam{\Delta}{B}
-    }
-  { \jfam{\ctxext{\Gamma}{\ctxwk{\Gamma}{\Delta}}}{\ctxwk{\Gamma}{B}}
-    }
-& & \inference
-  { \jfameq{\Delta}{B}{B'}
-    }
-  { \jfameq
-      {\ctxext{\Gamma}{\ctxwk{\Gamma}{\Delta}}}
-      {\ctxwk{\Gamma}{B}}
-      {\ctxwk{\Gamma'}{B'}}
-    }
-  \\
-& \inference
-  { \jterm{\Delta}{B}{y}
-    }
-  { \jterm
-      {\ctxext{\Gamma}{\ctxwk{\Gamma}{\Delta}}}
-      {\ctxwk{\Gamma}{B}}
-      {\ctxwk{\Gamma}{y}}
-    }
-& & \inference
-  { \jtermeq{\Delta}{B}{y}{y'}
-    }
-  { \jtermeq
-      {\ctxext{\Gamma}{\ctxwk{\Gamma}{\Delta}}}
-      {\ctxwk{\Gamma}{B}}
-      {\ctxwk{\Gamma}{y}}
-      {\ctxwk{\Gamma}{y'}}
-    }
-\end{align}
-Note that we do not employ three different notations for the three different
-actions of weakening by a context because there can be no confusion about which
-of the three actions is meant. If there were an empty context we should have
-distinguished the three as we did for weakening by a family.
-
-We shall have similar compatibility rules for weakening by a context as we did
-for weakening by a family.
-
-{\color{red}To do. It will look exactly the same as for weakening by families}
-\end{comment}
-
-\subsection{The theory of projections}
-\label{identityterms}
-We formulate the theory of projections on top of the theory of weakening
-described in \autoref{weakening}. In the theory of projections we introduce
-the identity terms. The identity term at a family $A$ in context $\Gamma$ is
-a term $\jterm{{\Gamma}{A}}{\ctxwk{A}{A}}{\idtm{A}}$, so weakening is necessary.
-Together with weakening, the identity terms will provide for the projections.
-The projection from $\cftext{A}{P}$ to $A$ will be the term
-$\jterm{{\Gamma}{{A}{P}}}{\ctxwk{\ctxext{A}{P}}{A}}{\ctxwk{P}{\idtm{A}}}$ and
-will be studied extensively in \autoref{extension-on-terms}.
-
-The following inference rules are required to be valid:
-\begin{align}
-& \inference
-  { \jfam{\Gamma}{A}
-    }
-  { \unfoldall{\jhom{\Gamma}{A}{A}{\idtm{A}}}
-    }
-& & \inference
-    { \jfameq{\Gamma}{A}{A'}
-      }
-    { \unfoldall{\jhomeq{\Gamma}{A}{A}{\idtm{A}}{\idtm{A'}}}
-      }
-\end{align}
-
-\begin{comment}
-We stress again
-that because we have not required there to be an empty context, we need to
-explicitly call for identity terms $\idtm{\Gamma}$ of $\ctxwk{\Gamma}{\Gamma}$
-in context $\Gamma$ for each context $\Gamma$. Thus, we will also require
-the inference rule
-\begin{align}
-& \inference
-  { \jctx{\Gamma}
-    }
-  { \jterm{\Gamma}{\ctxwk{\Gamma}{\Gamma}}{\idtm{\Gamma}}
-    }
-& &
-  \inference
-  { \jctxeq{\Gamma}{\Gamma'}
-    }
-  { \jtermeq{\Gamma}{\ctxwk{\Gamma}{\Gamma}}{\idtm{\Gamma}}{\idtm{\Gamma'}}
-    }
-\end{align}
-For the following inference rules, there should also be versions of identity
-terms at contexts.
-\end{comment}
-
-\subsubsection{The identity term of an extended family}
-The fact that identity terms are compatible with extensions is derived in
-\autoref{comp-ie}.
-
-\subsubsection{The identity term of a weakened family}
-\label{comp-wi}
-The identity term of a weakened family is the weakened identity term:
-\begin{equation}
-\inference
-  { \jfam{\Gamma}{A}
-    \jfam{{\Gamma}{B}}{Q}
-    }
-  { \unfoldall{\jhomeq
-      {{{\Gamma}{A}}{\ctxwk{A}{B}}}
-      {\ctxwk{A}{Q}}
-      {\ctxwk{A}{Q}}
-      {\ctxwk{A}{\idtm{Q}}}
-      {\idtm{\ctxwk{A}{Q}}}
-      }
-    }
-  \label{comp-wi-t}
-\end{equation}
-
-\begin{rmk}
-An important special case is the judgmental equality
-\begin{equation*}
-\unfoldall{\jhomeq
-      {{\Gamma}{A}}
-      {\ctxwk{A}{B}}
-      {\ctxwk{A}{B}}
-      {\ctxwk{A}{\idtm{B}}}
-      {\idtm{\ctxwk{A}{B}}}
-      }
-\end{equation*}
-for a family $\jfam{\Gamma}{B}$.
-\end{rmk}
-
-\subsection{The theory of substitution}
-\label{substitution}
-The theory of substitution can be formulated separately from the theory of
-weakening. Therefore, we will only require the theory of extension and the
-empty context and families. In \autoref{sec:esystem-equalities} we will
-provide the additional judgmental equality rules that will be required when
-combining the theories of substitution and of projections.
-
-Given a family $P$ over $A$ and a term $x$ of $A$, substitution gives a way to
-consider the \emph{fiber $\subst{x}{P}$ of $P$ at $x$}. As was the case with
-weakening, the substitution operation comes in three parts: the `action on
-contexts' of substitution is the operation just described; the `action on
-families' of substitution takes a family $Q$ over $\ctxext{{\Gamma}{A}}{P}$
-to a family $\subst[\famsym]{x}{Q}$ over $\ctxext{\Gamma}{\subst{x}{P}}$; the
-`action on terms' of substitution takes a term $g$ of $Q$ over
-$\ctxext{{\Gamma}{A}}{P}$ to a term $\subst[\tmsym]{x}{g}$ of 
-$\subst[\famsym]{x}{Q}$.
-\begin{align}
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jfam{{\Gamma}{A}}{P}
-    }
-  { \jfam{\Gamma}{\subst{x}{P}}
-    }
-& & \inference
-    { \jtermeq{\Gamma}{A}{x}{x'}
-      \jfameq{{\Gamma}{A}}{P}{P'}
-      }
-    { \jfameq{\Gamma}{\subst{x}{P}}{\subst{x'}{P'}}
-      }
-    \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jfam{{{\Gamma}{A}}{P}}{Q}
-    }
-  { \jfam{{\Gamma}{\subst{x}{P}}}{\subst[\famsym]{x}{Q}}
-    }
-& & \inference
-    { \jtermeq{\Gamma}{A}{x}{x'}
-      \jfameq{{{\Gamma}{A}}{P}}{Q}{Q'}
-      }
-    { \jfameq{{\Gamma}{\subst{x}{P}}}{\subst[\famsym]{x}{Q}}{\subst[\famsym]{x'}{Q'}}
-      }
-    \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jterm{{{\Gamma}{A}}{P}}{Q}{g}
-    }
-  { \jterm{{\Gamma}{\subst{x}{P}}}{\subst[\famsym]{x}{Q}}{\subst[\tmsym]{x}{g}}
-    }
-& & \inference
-    { \jtermeq{\Gamma}{A}{x}{x'}
-      \jtermeq{{{\Gamma}{A}}{P}}{Q}{g}{g'}
-      }
-    { \jtermeq
-        {{\Gamma}{\subst{x}{P}}}
-        {\subst[\famsym]{x}{Q}}
-        {\subst[\tmsym]{x}{g}}
-        {\subst[\tmsym]{x'}{g'}}
-      }
-\end{align}
-
-\subsubsection{Fibers of an empty family}
-The following inference rules establish that the fibers of the empty family are 
-the empty families:
-\label{comp-s0}
-\begin{align}
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    }
-  { \jfameq{\Gamma}{\subst{x}{\emptyf}}{\emptyf}
-    }
-  \label{comp-s0-c}
-  \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jfam{{\Gamma}{A}}{P}
-    }
-  { \jfameq
-      {{\Gamma}{\subst{x}{P}}}
-      {\subst{x}{\emptyf}}
-      {\emptyf}
-    }
-  \label{comp-s0-f}
-\end{align}
-\begin{comment}
-Note again that we do not have to axiomatize the action of substitution by $x$ on the term
-of the empty family, because the result will be in the empty family and every
-term in the empty family is judgmentally equal to the term $\emptytm$.
-\end{comment}
-
-We use the above rule to state the compatibility of the action on families of
-substitution with the action on contexts of substitution. Note that a family
-$P$ over $\ctxext{\Gamma}{A}$ may be regarded as a family over
-$\ctxext{{\Gamma}{A}}{\emptyf}$. Thus, we may consider the family
-$\subst[\famsym]{x}{P}$ over $\ctxext{\Gamma}{\subst{x}{\emptyf}}$. Since
-$\subst{x}{\emptyf}$ is judgmentally equal to the empty family, we may compare
-$\subst[\famsym]{x}{P}$ with $\subst{x}{P}$:
-\begin{equation}
-\inference
-{ \jfam{{\Gamma}{A}}{P}
-  }
-{ \jfameq{\Gamma}{\subst[\famsym]{x}{P}}{\subst{x}{P}}
-  }
-\end{equation}
-Due to this rule we need not make the annotations $\famsym$ and $\tmsym$ in
-the notation for substitution anymore and thus we shall omit them from now on.
-Note that because a family $P$ over $\ctxext{\Gamma}{A}$ is eligible for
-application of the action on families of substitution, we may also substitute
-terms of $P$. Thus, given terms $\jterm{\Gamma}{A}{x}$ and $\jterm{{\Gamma}{A}}{P}{f}$,
-we get a term $\jterm{\Gamma}{\subst{x}{P}}{\subst{x}{f}}$, the \emph{value of
-$f$ at $x$}.
-
-In \autoref{morphisms} we will use a combination of weakening and substitution
-to define composition of morphisms of families. However, we have to rely
-on the cancellation rule stated in \autoref{cancellation-ws} before we can
-meaningfully state the definition of composition.
-
-\subsubsection{Fibers of extensions}
-\label{comp-se}
-The following inference rules assert that if we take the fiber of an extended
-family at a term $x$ of $A$ in context $\Gamma$, the substitution by $x$
-distributes over the factors of the extension.
-\begin{align}
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jfam{{{\Gamma}{A}}{P}}{Q}
-    }
-  { \jfameq
-      {\Gamma}
-      {\subst{x}{\ctxext{P}{Q}}}
-      {\ctxext{\subst{x}{P}}{\subst{x}{Q}}}
-    }
-  \label{comp-se-c}
-  \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jfam{{{{\Gamma}{A}}{P}}{Q}}{R}
-    }
-  { \jfameq
-      {{\Gamma}{\subst{x}{P}}}
-      {\subst{x}{\ctxext{Q}{R}}}
-      {\ctxext{\subst{x}{Q}}{\subst{x}{R}}}
-    }
-  \label{comp-se-f}
-\end{align}
-There is also a version of this statement in which extended terms are considered.
-This variant is stated and proved in \autoref{comp-se-t}.
-
-\subsubsection{Fibers of fibers}
-\label{comp-ss}
-
-The following rules explain what happens when we first substitute by a term
-$g$ of $Q$ over $P$ in context $\ctxext{\Gamma}{A}$ and then by a term $x$ of
-$A$:
-\begin{align}
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jterm{{{\Gamma}{A}}{P}}{Q}{g}
-    \jfam{{{{\Gamma}{A}}{P}}{Q}}{R}
-    }
-  { \jfameq
-      {{{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}}
-      {\subst{x}{{g}{R}}}
-      {\subst{{x}{g}}{{x}{R}}}
-    }
-  \label{comp-ss-c}
-  \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jterm{{{\Gamma}{A}}{P}}{Q}{g}
-    \jfam{{{{{\Gamma}{A}}{P}}{Q}}{R}}{S}
-    }
-  { \jfameq
-      {{{{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}}{\subst{x}{{g}{R}}}}
-      {\subst{x}{{g}{S}}}
-      {\subst{{x}{g}}{{x}{S}}}
-    }
-  \label{comp-ss-f}
-  \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jterm{{{\Gamma}{A}}{P}}{Q}{g}
-    \jterm{{{{{\Gamma}{A}}{P}}{Q}}{R}}{S}{k}
-    }
-  { \jtermeq
-      {{{{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}}{\subst{x}{{g}{R}}}}
-      {\subst{x}{{g}{S}}}
-      {\subst{x}{{g}{k}}}
-      {\subst{{x}{g}}{{x}{k}}}
-    }
-  \label{comp-ss-t}
-\end{align}
-
-\begin{rmk}
-As an important special case of these inference rules we have the following
-valid inference rules:
-\begin{align*}
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jterm{{\Gamma}{A}}{P}{f}
-    \jfam{{{\Gamma}{A}}{P}}{Q}
-    }
-  { \jfameq
-      {\Gamma}
-      {\subst{x}{{f}{Q}}}
-      {\subst{{x}{f}}{{x}{Q}}}
-    }
-  \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jterm{{\Gamma}{A}}{P}{f}
-    \jfam{{{{\Gamma}{A}}{P}}{Q}}{R}
-    }
-  { \jfameq
-      {{\Gamma}{\subst{x}{{f}{Q}}}}
-      {\subst{x}{{f}{R}}}
-      {\subst{{x}{f}}{{x}{R}}}
-    }
-  \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jterm{{\Gamma}{A}}{P}{f}
-    \jterm{{{{\Gamma}{A}}{P}}{Q}}{R}{h}
-    }
-  { \jtermeq
-      {{\Gamma}{\subst{x}{{f}{Q}}}}
-      {\subst{x}{{f}{R}}}
-      {\subst{x}{{f}{h}}}
-      {\subst{{x}{f}}{{x}{h}}}
-    }
-\end{align*}
-Moreover, we get
-\begin{equation*}
-\inference
-  { \jterm{\Gamma}{A}{x}
-    \jterm{{\Gamma}{A}}{P}{f}
-    \jterm{{{\Gamma}{A}}{P}}{Q}{g}
-    }
-  { \jtermeq
-      {\Gamma}
-      {\subst{x}{{f}{Q}}}
-      {\subst{x}{{f}{g}}}
-      {\subst{{x}{f}}{{x}{g}}}
-    }
-\end{equation*}
-\end{rmk}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Joining the theory of projections with the theory of substitution}
-\label{sec:esystem-equalities}
-
-\begin{comment}
-We have already seen that empty families, extension, weakening, substitution
-and the identity terms are compatible with each other. There are, however, a
-few more judgmental equalities explaining what happens when
-\begin{enumerate}
-\item something is weakend by the empty family. This is used to consider a term
-      $x$ of $A$ in context $\Gamma$ as a morphism from $\emptyf$ to $A$ in
-      context $\Gamma$;
-\item something is weakened by an extended family. This is Currying for weakening;
-\item something is first weakened by a family $A$ in context $\Gamma$ and then
-      substituted by a term of $A$. This is used to show that composition is
-      compatible with itself (i.e. that composition is associative);
-\item the identity terms are joining the party. Identity terms are empty contexts
-      in certain models, so they should behave accordingly.
-\end{enumerate}
-In other words, the rules seem to not fit in the scheme that all operations are
-compatible with each other, but in fact each of these rules plays a role in
-completing the scheme (work out further).
-\end{comment}
-
-\begin{comment}
-In this section we extend the judgmental equality relation in such a way that
-the empty context, extension, weakening, substitution and the identity terms
-all become compatible with each other. Since we have five operations, there
-are going to be $25$ finite sets of rules explaining how each operation is
-compatible with each of the operations.
-
-In addition to these $25$ finite sets
-of compatibility rules, there are also two sets of rules explaining the
-cancellation properties of weakening and substitution and of the identity terms.
-
-A typical compatibility axiom captures the idea that a diagram like
-\begin{equation*}
-\begin{tikzcd}[column sep=huge]
-\jfam{\Gamma}{\blank} 
-  \ar{r}{C\mapsto\ctxwk{B}{C}} 
-  \ar{d}[swap]{C\mapsto\ctxwk{A}{C}} 
-& \jfam{{\Gamma}{B}}{\blank} 
-  \ar{d}{Q\mapsto\ctxwk{A}{Q}}
-  \\
-\jfam{{\Gamma}{A}}{\blank} 
-  \ar{r}[swap]{P\mapsto\ctxwk{{A}{B}}{P}} 
-& \jfam{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\blank}
-\end{tikzcd}
-\end{equation*}
-in the case of two consecutive weakenings, or
-\begin{equation*}
-\begin{tikzcd}[column sep=huge]
-\jfam{{\Gamma}{A}}{\blank} 
-  \ar{d}[swap]{Q\mapsto\subst{x}{Q}} 
-  \ar{r}{Q\mapsto\ctxwk{P}{Q}} 
-& \jfam{{{\Gamma}{A}}{P}}{\blank} 
-    \ar{d}{R\mapsto\subst{x}{R}}
-  \\ 
-\jfam{\Gamma}{\blank} 
-  \ar{r}[swap]{B\mapsto\ctxwk{\subst{x}{P}}{B}} 
-& \jfam{{\Gamma}{\subst{x}{P}}}{\blank}
-\end{tikzcd}
-\end{equation*}
-in the case of a combination of weakening and substitution, commutes. We won't
-actually draw any of these diagrams in this section, but readers are encouraged
-to do so if they think it can help them understanding the axioms. Another thing
-one could keep in mind with the axioms that involve weakening is that all of
-them become completely obvious when weakening is made notationally invisible.
-Of course, we have an explicit notation for weakening on purpuse, but it is the
-usual practice of type theorists to make weakening invisible.
-
-The table below indicates where
-we can find the axioms of each of the combinations of operators. The reference
-at the row of operation $X$ and the column of operation $Y$ refers to the
-axiom which explains $Y$ distributes over the arguments of $X$ when $Y$ is
-applied after $X$. The empty context has no arguments, whereas extension,
-and weakening have two arguments which are both of the family-kind, substitution
-has two arguments of which one is a term and the other is a family and
-idenitity terms have one argument, a family.
-\begin{center}
-\begin{tabular}{r|rrrr}
-& ext & wk & subst & id\\
-\hline
-extension & \autoref{comp-ee}* &  & \cmark & \cmark \\
-weakening & \autoref{comp-we}* & \autoref{comp-ww} & \autoref{comp-sw} & \autoref{comp-wi} \\
-substitution & \autoref{comp-se}* & \autoref{comp-ws} & \autoref{comp-ss} & \autoref{comp-si} \\
-identity terms & \cmark & \autoref{comp-wi} & \autoref{comp-si} & \xmark
-\end{tabular}
-\end{center}
-Apart from references to sections where axioms are given, we also see checkmarks,
-crosses, marked references and duplicated references in this table. 
-The checkmarks indicate that the desired property
-can be proved and need not be axiomatized. To clarify those:
-\begin{itemize}
-\item The identity term of the empty context is the unique term of the empty
-family. Recall that we already have axiomatized in \autoref{empty} that every
-term of the empty family is judgmentally equal to $\emptytm$.
-\item The identity term of an extension can be explained by means of the notion
-of \emph{extension on terms}. This notion is defined in \autoref{extension-on-terms}
-using only identity terms and substitution. With extension on terms we can
-combine a term $\jterm{\Gamma}{A}{x}$ and a term $\jterm{\Gamma}{\subst{x}{P}}{u}$
-into a term $\jterm{\Gamma}{{A}{P}}{\tmext{x}{u}}$, which may be seen as the
-pair consisting of $x$ and $u$. Since it is a defined notion, we will
-be able to derive lots of properties for extension of terms. One of them explains
-how we can see the identity term of an extension as the extension of two certain
-terms, see \autoref{comp-ie}.
-\item Using extension on terms we can also make sense of `fibers at an extension'
-because that would exactly be the fiber at an extended term. We will see in
-\autoref{comp-es} that, for instance, $\subst{\tmext{x}{u}}{Q}\jdeq\subst{u}{{x}{Q}}$
-for any family $\jfam{{{\Gamma}{A}}{P}}{Q}$ and any two terms $\jterm{\Gamma}{A}{x}$
-and $\jterm{\Gamma}{\subst{x}{P}}{u}$. That's a form of currying.
-\end{itemize}
-The spots with a \xmark simply make no sense 
-{\color{red}(or do they, when the term model
-is better understood (by me)?)}. The marked references point to
-axioms which \emph{do} have a variant for terms, but those require the notion
-of extension on terms, can be proved and therefore need not be included in our
-list of compatibility axioms. In each of these sections, we give a forward
-reference to the lemma in which this variant is stated and proved.
-\end{comment}
-
-\subsubsection{Fibers of weakenings}\label{comp-sw}
-The following rules assert what happens when we first weaken by a family
-$Q$ over $P$ in context $\ctxext{\Gamma}{A}$ and then substitute by a term
-$x$ of $A$:
-\begin{align}
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jfam{{{\Gamma}{A}}{P}}{Q}
-    \jfam{{{\Gamma}{A}}{P}}{R}
-    }
-  { \jfameq
-      {{{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}}
-      {\subst{x}{\ctxwk{Q}{R}}}
-      {\ctxwk{\subst{x}{Q}}{\subst{x}{R}}}
-    }
-  \label{comp-sw-c}
-  \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jfam{{{\Gamma}{A}}{P}}{Q}
-    \jfam{{{{\Gamma}{A}}{P}}{R}}{S}
-    }
-  { \jfameq
-      {{{{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}}{\subst{x}{\ctxwk{Q}{R}}}}
-      {\subst{x}{\ctxwk{Q}{S}}}
-      {\ctxwk{\subst{x}{Q}}{\subst{x}{S}}}
-    }
-  \label{comp-sw-f}
-  \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jfam{{{\Gamma}{A}}{P}}{Q}
-    \jterm{{{{\Gamma}{A}}{P}}{R}}{S}{k}
-    }
-  { \jtermeq
-      {{{{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}}{\subst{x}{\ctxwk{Q}{R}}}}
-      {\subst{x}{\ctxwk{Q}{S}}}
-      {\subst{x}{\ctxwk{Q}{k}}}
-      {\ctxwk{\subst{x}{Q}}{\subst{x}{k}}}
-    }
-  \label{comp-sw-t}
-\end{align}
-
-\begin{rmk}
-As an important special case of these inference rules we have the following
-valid inference rules:
-\begin{align*}
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jfam{{\Gamma}{A}}{P}
-    \jfam{{\Gamma}{A}}{Q}
-    }
-  { \jfameq
-      {{\Gamma}{\subst{x}{P}}}
-      {\subst{x}{\ctxwk{P}{Q}}}
-      {\ctxwk{\subst{x}{P}}{\subst{x}{Q}}}
-    }
-  \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jfam{{\Gamma}{A}}{P}
-    \jfam{{{\Gamma}{A}}{Q}}{R}
-    }
-  { \jfameq
-      {{{\Gamma}{\subst{x}{P}}}{\subst{x}{\ctxwk{P}{Q}}}}
-      {\subst{x}{\ctxwk{P}{R}}}
-      {\ctxwk{\subst{x}{P}}{\subst{x}{R}}}
-    }
-  \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jfam{{\Gamma}{A}}{P}
-    \jterm{{{\Gamma}{A}}{Q}}{R}{h}
-    }
-  { \jtermeq
-      {{{\Gamma}{\subst{x}{P}}}{\subst{x}{\ctxwk{P}{Q}}}}
-      {\subst{x}{\ctxwk{P}{R}}}
-      {\subst{x}{\ctxwk{P}{h}}}
-      {\ctxwk{\subst{x}{P}}{\subst{x}{h}}}
-    }
-\end{align*}
-Moreover, we get
-\begin{equation*}
-\inference
-  { \jterm{\Gamma}{A}{x}
-    \jfam{{\Gamma}{A}}{P}
-    \jterm{{\Gamma}{A}}{Q}{g}
-    }
-  { \jtermeq
-      {{\Gamma}{\subst{x}{P}}}
-      {\subst{x}{\ctxwk{P}{Q}}}
-      {\subst{x}{\ctxwk{P}{g}}}
-      {\ctxwk{\subst{x}{P}}{\subst{x}{g}}}
-    }
-\end{equation*}
-\end{rmk}
-
-\subsubsection{Weakenings of fibers}
-\label{comp-ws}
-The following inference rules explain what happens when we first weaken by a
-term $g$ of $Q$ in context $\ctxext{\Gamma}{B}$ and then weaken by a family
-$A$ over $\Gamma$.
-\begin{align}
-& \inference
-  { \jfam{\Gamma}{A}
-    \jterm{{\Gamma}{B}}{Q}{g}
-    \jfam{{{\Gamma}{B}}{Q}}{R}
-    }
-  { \jfameq
-      {{{\Gamma}{A}}{\ctxwk{A}{B}}}
-      {\ctxwk{A}{\subst{g}{R}}}
-      {\subst{\ctxwk{A}{g}}{\ctxwk{A}{R}}}
-    }
-  \label{comp-ws-c}
-  \\
-& \inference
-  { \jfam{\Gamma}{A}
-    \jterm{{\Gamma}{B}}{Q}{g}
-    \jfam{{{{\Gamma}{B}}{Q}}{R}}{S}
-    }
-  { \jfameq
-      {{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{\subst{g}{R}}}}
-      {\ctxwk{A}{\subst{g}{S}}}
-      {\subst{\ctxwk{A}{g}}{\ctxwk{A}{S}}}
-    }
-  \label{comp-ws-f}
-  \\
-& \inference
-  { \jfam{\Gamma}{A}
-    \jterm{{\Gamma}{B}}{Q}{g}
-    \jterm{{{{\Gamma}{B}}{Q}}{R}}{S}{k}
-    }
-  { \jtermeq
-      {{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{\subst{g}{R}}}}
-      {\ctxwk{A}{\subst{g}{S}}}
-      {\ctxwk{A}{\subst{g}{k}}}
-      {\subst{\ctxwk{A}{g}}{\ctxwk{A}{k}}}
-    }
-  \label{comp-ws-t}
-\end{align}
-
-\begin{rmk}
-As an important special case of these inference rules we have the following 
-valid inference rules:
-\begin{align*}
-& \inference
-  { \jfam{\Gamma}{A}
-    \jterm{\Gamma}{B}{y}
-    \jfam{{\Gamma}{B}}{Q}
-    }
-  { \jfameq
-      {{\Gamma}{A}}
-      {\ctxwk{A}{\subst{y}{Q}}}
-      {\subst{\ctxwk{A}{y}}{\ctxwk{A}{Q}}}
-    }
-  \\
-& \inference
-  { \jfam{\Gamma}{A}
-    \jterm{\Gamma}{B}{y}
-    \jfam{{{\Gamma}{B}}{Q}}{R}
-    }
-  { \jfameq
-      {{{\Gamma}{A}}{\ctxwk{A}{\subst{y}{Q}}}}
-      {\ctxwk{A}{\subst{y}{R}}}
-      {\subst{\ctxwk{A}{y}}{\ctxwk{A}{R}}}
-    }
-  \\
-& \inference
-  { \jfam{\Gamma}{A}
-    \jterm{\Gamma}{B}{y}
-    \jterm{{{\Gamma}{B}}{Q}}{R}{h}
-    }
-  { \jtermeq
-      {{{\Gamma}{A}}{\ctxwk{A}{\subst{y}{Q}}}}
-      {\ctxwk{A}{\subst{y}{R}}}
-      {\ctxwk{A}{\subst{y}{h}}}
-      {\subst{\ctxwk{A}{y}}{\ctxwk{A}{h}}}
-    }
-\end{align*}
-Moreover, we get
-\begin{equation*}
-\inference
-  { \jfam{\Gamma}{A}
-    \jterm{\Gamma}{B}{y}
-    \jterm{{\Gamma}{B}}{Q}{g}
-    }
-  { \jtermeq
-      {{\Gamma}{A}}
-      {\ctxwk{A}{\subst{y}{Q}}}
-      {\ctxwk{A}{\subst{y}{g}}}
-      {\subst{\ctxwk{A}{y}}{\ctxwk{A}{g}}}
-    }
-\end{equation*}
-\end{rmk}
-
-\subsubsection{The cancellation property of weakening and substitution}
-\label{cancellation-ws}
-The judgmental equalities we're about to describe assert that substituting a term
-in the weakening a thing gives you the thing back. In the case of contexts we get that each fiber
-$\subst{x}{\ctxwk{A}{B}}$ is just $B$ and in the case of terms we get 
-that $\ctxwk{A}{y}$ is the constant function
-mapping everything to $y:B$. Thus, these rules actually establish the weakening
-as the weakening. After stating the rules we will describe what it means to
-compose context morphisms (terms of weakened contexts).
-\begin{align}
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{\Gamma}{B}
-    \jterm{\Gamma}{A}{x}
-    }
-  { \jfameq
-      {\Gamma}
-      {\subst{x}{\ctxwk{A}{B}}}
-      {B}
-    }
-  \label{cancellation-ws-c}\\
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{{\Gamma}{B}}{Q}
-    \jterm{\Gamma}{A}{x}
-    }
-  { \jfameq{{\Gamma}{B}}{\subst{x}{\ctxwk{A}{Q}}}{Q}
-    }
-  \label{cancellation-ws-f}\\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jterm{{\Gamma}{B}}{Q}{g}
-    }
-  { \jtermeq{{\Gamma}{B}}{Q}{\subst{x}{\ctxwk{A}{g}}}{g}
-    }
-  \label{cancellation-ws-t}
-\end{align}
-
-\subsubsection{The identity term of a substituted family}
-\label{comp-si}
-The identity term of a substituted family is the substitution of the identity term
-\begin{equation}
-\inference
-  { \jterm{\Gamma}{A}{x}
-    \jfam{{{\Gamma}{A}}{P}}{Q}
-    }
-  { \unfoldall{\jhomeq
-      {{\Gamma}{\subst{x}{P}}}
-      {\subst{x}{Q}}
-      {\subst{x}{Q}}
-      {\subst{x}{\idtm{Q}}}
-      {\idtm{\subst{x}{Q}}}
-    }}
-  \label{comp-si-t}
-\end{equation}
-
-\begin{rmk}
-An important special case is the judgmental equality
-\begin{equation*}
-\unfoldall{\jhomeq
-      {\Gamma}
-      {\subst{x}{P}}
-      {\subst{x}{P}}
-      {\subst{x}{\idtm{P}}}
-      {\idtm{\subst{x}{P}}}
-    }
-\end{equation*}
-for a family $\jfam{{\Gamma}{A}}{P}$.
-\end{rmk}
-
-\subsubsection{The cancellation properties of identity terms}
-\label{cancellation-i}
-Identity terms are determined by their behavior with respect to substitution combined with
-weakening. The identity terms will also be subject to compatibility rules.
-\begin{align}
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    }
-  { \jtermeq{\Gamma}{A}{\subst{x}{\idtm{A}}}{x}
-    }
-  \label{cancellation-si}\\
-& \inference
-  { \jfam{{\Gamma}{A}}{P}
-    }
-  { \jfameq{{\Gamma}{A}}{\subst{\idtm{A}}{\ctxwk{A}{P}}}{P}
-    }
-  \label{precomp-idtm-c}
-  \\
-& \inference
-  { \jfam{{{\Gamma}{A}}{P}}{Q}
-    }
-  { \jfameq{{{\Gamma}{A}}{P}}{\subst{\idtm{A}}{\ctxwk{A}{Q}}}{Q}
-    }
-  \label{precomp-idtm-f}
-  \\
-& \inference
-  { \jterm{{{\Gamma}{A}}{P}}{Q}{g}
-    }
-  { \jtermeq
-      {{{\Gamma}{A}}{P}}
-      {Q}
-      {\subst{\idtm{A}}{\ctxwk{A}{g}}}
-      {g}
-    }
-  \label{precomp-idtm-t}
-\end{align}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{The possiblity of types in the theory of contexts, families and
-terms}
-\label{types}
-We have deliberately not spoken of types so far because we have taken the point
-of view that a type in a context is nothing but a family in that context which
-belongs to the class of types. We think of types as \emph{irreducible} families,
-i.e.\ families which are neither empty nor the extension of two
-families which are both not empty (in algebraic terminology: which
-are both non-trivial). To allow ourselves to speak of types we introduce two
-new judgments: the judgment that something is a type and the judgment that two
-types are equal.
-\begin{align*}
-\jalign\jtype{\Gamma}{A} 
-& \jalign\jtypeeq{\Gamma}{A}{B}
-\end{align*}
-But only families of contexts are eligible to be types. If $A$ is a type
-in context $\Gamma$, then $A$ is also a family of contexts over $\Gamma$. 
-Moreover, two types in context $\Gamma$ are judgmentally equal precisely when they are equal
-as context families and if a family $B$ of contexts over $\Gamma$ is
-judgmentally equal to a type $A$ in context $\Gamma$, then $B$ is a type in
-context $\Gamma$. This is expressed by the following four inference rules:
-\begin{align*}
-& \inference
-  { \jtype{\Gamma}{A}
-    }
-  { \jfam{\Gamma}{A}
-    }
-& & \inference
-    { \jtypeeq{\Gamma}{A}{B}
-      }
-    { \jfameq{\Gamma}{A}{B}
-      }
-    \\
-& \inference
-  { \jtype{\Gamma}{A}
-    \jfameq{\Gamma}{A}{B}
-    }
-  { \jtype{\Gamma}{B}
-    }
-& & \inference
-    { \jtype{\Gamma}{A}
-      \jfameq{\Gamma}{A}{B}
-      }
-    { \jtypeeq{\Gamma}{A}{B}
-      }
-\end{align*}
-As pointed out at the beginning of this subsection, 
-we do not assume that the empty family is a type, that would be like
-assuming that the multiplicative unit of a ring is prime.
-
-\begin{comment}
-We envision a
-categorical interpretation of the theory of contexts, families and terms where 
-contexts are interpreted as finite paths of arrows belonging to a predetermined
-class. The empty family would be such a path of length $0$ whereas the types
-would be such paths of length $1$.
-\end{comment}
-
-We add rules asserting that a weakened type is again a type that 
-substitution preserves the property of being a type:
-\begin{align}
-& \inference
-  { \jfam{\Gamma}{B}
-    \jtype{\ctxext{\Gamma}{B}}{Q}
-    }
-  { \jtype{\ctxext{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{Q}}
-    }
-  \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jtype{\ctxext{{\Gamma}{A}}{P}}{Q}
-    }
-  { \jtype{\ctxext{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}
-    }
-\end{align}
-With only the current rules, the possibility of making the judgment that
-something is a type does not add much to the theory of contexts, families and
-terms. Nevertheless, when studying models, having an interpretation for the
-judgment that something is a type will allow for the possibility to study
-conditions such as the one asserting that every family `factorizes' uniquely
-as multiple applications of extension to types, analogous to the condition on
-unique factorization domains in ring theory. All the tradional models of type
-theory should translate to models with such a condition, simply because
-traditionally contexts are viewed as lists of variable (and type) declarations.

### 2014-10-28 Rules/rules_extension.tex
reorganization

--- 2014-09-20 article-models-tt.tex
@@
-\section{The generalized algebraic theory of contexts, families and terms (the E-system)}\label{tt}
-In this section we give a description of dependent type theory before type
-constructors. Apart from contexts, families and terms -- which provide for the
-core of the language of dependent logic -- the basic ingredients
-of this theory will be the operations of extension on context and on families, 
-and empty context and empty families, weakening, substitution
-and the identity terms. The resulting theory can be seen as a manifestation of the 
-structure underlying dependent type theory.
-
-We will
-formulate the theory of contexts, families and terms in such a way that contexts aren't defined
-to be lists of variable declarations. The variable-free (a.k.a.~\emph{name-free}) approach 
-we take here is rather different than those appearing in 
-\cite{hofmann1995extensional,TheBook} but it has appeared in the work of Coquand
-and in \cite{Dybjer1996}.
-The main reason we don't let variable declarations in is that we don't see them 
-in the internal models either. This way we also set out to a more algebraic 
-approach of type theory and higher category theory. Thirdly, we will not have to
-be burdened with superficial comments about variables being bounded or not, or 
-fresh or free or not occuring at all.
-
-In the current presentation, contexts can be seen as binary
-planar trees of which the leaves are (families of) contexts. 
-The judgmental equality relation on contexts is an equivalence relation which 
-expresses that binary planar
-trees of contexts are judgmentally equal if their leaves are, taking only
-(the isomorphism class of) 
-the order of the leaves into account \emph{and not the actual shape of the three}.
-The intuition behind this equivalence relation is indeed that unstructured
-(i.e.~unbracketed) lists such as the lists of variable declarations which
-usually appear in type theoretical syntax, may be regarded as contexts.
-
-Besides contexts, families and terms there will also be a notion of `type in
-a context', which one may assume.
-A context is eligible to be a \emph{closed type} and over a context, a family
-may be a type. We have the intuition that `being a type'
-expresses the property of `being atomic or irreducible'. With a notion of types,
-one has the means to require a stratification by which every context is a finite
-extension of types starting at a closed type, and every family is likewise a
-finite extension of types. The idea behind this is to recover the theory of
-\cite{Garner2014} {\color{red}(and possibly B-systems?)}. We stress that
-we will not include axioms asserting such a stratification in the E-system
-and we leave discussion on the possibility of a typing judgment until the
-end of this section.
-
-With families of contexts being the principal things of
-study, we have added a bit more generality to the theory, compared to the
-theory of categories with families presented in \cite{Dybjer1996}. Here are
-a couple of reasons for this extra generality:
-\begin{enumerate}
-\item We want to be very sure that we can interpret all the rules of type theory
-in a straightforward manner in any context.
-\item We do want do end up with a categorical structure on the contexts, derived
-from the theory by requiring that a morphism from $\Gamma$ to $\Delta$ is a
-term of the weakening of $\Delta$ by $\Gamma$ in context $\Gamma$. Thus, we
-must be able to lift the context $\Delta$ to become a family of contexts over
-$\Gamma$ and it is not reasonable to expect that this results in a type.
-\item It is very easy in this setting to assert that an operation has an action
-on very high levels of families, i.e.~when $F$ is an operation taking things in
-context $\Gamma$ to things in context $\Delta$, then it takes a family $A$ in
-context $\Gamma$ to $F(A)$ in context $\Delta$; a family $P$ in context
-$\ctxext{\Gamma}{A}$ to a family $F(P)$ in context $\ctxext{\Delta}{(F(A))}$,
-and so on. The reason it becomes simple is that we don't have to explain what
-`and so on' means.
-\end{enumerate}
-
-We will formulate fairly strict rules governing the judgmental equalities,
-expressing that extension, weakening and substitution are combatible with
-each other in a judgmental manner. This does not, however, diminish the role
-of isomorphisms or of homotopies could play in the theory once identity types
-are added. Indeed, types could still have non-trivial identity relations and
-the category of types in a certain context could genuinely display higher
-categorical structure, or so we conjecture.
-
-Much of the rules we state are just compatibility rules of extension, weakening
-and substitution with each other. In a way, these rules assert that our contexts
-are just structureless lists of contexts and that likewise terms are structureless
-lists of terms. They are structureless in the sense that the order in which
-they are formed by pairing up is irrelevant. We note that this causes complications
-in the traditional way that categorical sematics of type theory is implemented,
-where contexts become objects of the category which is supposed to model type
-theory. The reason for this is that context extension will not satisfy all the
-compatibility rules we're about to state. The first step to resolving this is taking
-the types in the empty context as the objects.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Judgments and valid inference rules}\label{judgments}
-The theory we describe here is a theory of contexts, families of
-contexts and terms thereof. The families of contexts are by some authors called
-dependent contexts, but they are handled a bit differently here because they
-become the primary object of study. Dependent contexts can be types; they could
-be seen as atomic or indecomposable dependent contexts.
-
-Thus we make eight kinds of judgments: ``$\Gamma$ is a context'',
-``$A$ is a family of contexts over $\Gamma$'', ``$A$ is a type in context $\Gamma$''
-and ``$x$ is a term of the family $A$ of contexts over $\Gamma$''. The other four
-judgments are for judgmental equality.
-\begin{align*}
-\jalign\jctx{\Gamma} 
-& \jalign\jctxeq{\Gamma}{\Gamma'}
-  \\
-\jalign\jfam{\Gamma}{A} 
-& \jalign\jfameq{\Gamma}{A}{B}
-  \\
-\jalign\jterm{\Gamma}{A}{x} 
-& \jalign\jtermeq{\Gamma}{A}{x}{y}.
-\end{align*}
-
-Strictly speaking, we have three different judgmental equalities in play and one
-could request for a notational difference to signify that fact. For instance,
-we could denote the judgmental equalities of contexts, families and terms by
-$\jdeq_c$, $\jdeq_f$ and $\jdeq_t$ respectively. It will, however, always be
-clear which of the three kinds of judgmental equality is meant when we assert
-a judgmental equality and therefore we shall not bother to make this notational
-distinction.
-
-We note that what we call families over contexts
-here could also have been named dependent contexts or telescopes, see
-\cite{deBruijn1991,hofmann1995extensional}. The term family is in agreement
-with the terminology scheme of \cite{TheBook}, though the reader should be
-warned that the notion of familie means something slightly different there than
-it does here.
-
-The theory of contexts, families and terms shall be described by means of a
-declaration of valid inference rules. An inference rule is an expression of the
-form
-\begin{equation*}
-\inference{\mathcal{H}_1\quad\cdots\quad\mathcal{H}_n}{\mathcal{J}}
-\end{equation*}
-where each of $\mathcal{H}_1$, ..., $\mathcal{H}_n$ and $\mathcal{J}$ is 
-judgment. The judgments $\mathcal{H}_1$, ..., $\mathcal{H}_n$ are called the
-hypotheses and the judgment $\mathcal{J}$ is the conclusion of the inference
-rule. When a judgment is the conclusion of a valid inference rule, we also say
-that the judgment is valid or well-formed. In fact, the only way to construct
-new judgments is via the valid inference rules we describe in this section.
-
-The class of valid inference rules is closed under composition and under
-rearranging the hypotheses. That
-means that if
-\begin{equation*}
-\inference{\mathcal{I}_{1k}\quad\cdots\quad\mathcal{I}_{m_kk}}{\mathcal{H}_k}
-\end{equation*}
-and
-\begin{equation*}
-\inference{\mathcal{H}_1\quad\cdots\quad\mathcal{H}_n}{\mathcal{J}}
-\end{equation*}
-are valid inference rules for $1\leq k\leq n$, then
-\begin{equation*}
-\inference{\mathcal{I}_{11}\quad\cdots\quad\mathcal{I}_{m_11}\quad\cdots\quad\mathcal{I}_{1n}\quad\cdots\quad\mathcal{I}_{m_nn}}{\mathcal{J}}
-\end{equation*}
-is also a valid inference rule. Also, in a valid inference rule 
-duplicated hypotheses may always be contracted and hypotheses may always be 
-reordered; the result is still considered a valid inference rule.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{The basic rules for judgmental equality}
-The rules for judgmental equality establish that it is an equivalence relation
-in all three cases (contexts, families and terms). Thus, the following inference
-rules shall be required to be valid:
-\bgroup\small
-\begin{align*}
-& \inference
-  { \jctx{\Gamma}
-    }
-  { \jctxeq{\Gamma}{\Gamma}
-    } 
-& & \inference
-    { \jctxeq{\Gamma}{\Delta}
-      }
-    { \jctxeq{\Delta}{\Gamma}
-      } 
-& & \inference
-    { \jctxeq{\Gamma}{\Delta}
-      \jctxeq{\Delta}{\greek{E}}
-      }
-    { \jctxeq{\Gamma}{\greek{E}}
-      }
-    \\
-& \inference
-  { \jfam{\Gamma}{A}
-    }
-  { \jfameq{\Gamma}{A}{A}
-    } 
-& & \inference
-    { \jfameq{\Gamma}{A}{B}
-      }
-    { \jfameq{\Gamma}{B}{A}
-      }
-& & \inference
-    { \jfameq{\Gamma}{A}{B}
-      \jfameq{\Gamma}{B}{C}
-      }
-    { \jfameq{\Gamma}{A}{C}
-      }
-    \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    }
-  { \jtermeq{\Gamma}{A}{x}{x}
-    }
-& & \inference
-    { \jtermeq{\Gamma}{A}{x}{y}
-      }
-    { \jtermeq{\Gamma}{A}{y}{x}
-      }
-& & \inference
-    { \jtermeq{\Gamma}{A}{x}{y}
-      \jtermeq{\Gamma}{A}{y}{z}
-      }
-    { \jtermeq{\Gamma}{A}{x}{z}
-      }
-\end{align*}
-\egroup
-
-The following convertibility rules are responsible for the strictness
-of judgmental equality, which sets it apart from equivalences or identifications:
-\begin{align*}
-& \inference
-  { \jctxeq{\Gamma}{\Delta}
-    \jfam{\Gamma}{A}
-    }
-  { \jfam{\Delta}{A}
-    }
-& & \inference
-    { \jctxeq{\Gamma}{\Delta}
-      \jfameq{\Gamma}{A}{B}
-      }
-    { \jfameq{\Delta}{A}{B}
-      }
-    \\
-& \inference
-  { \jctxeq{\Gamma}{\Delta}
-    \jterm{\Gamma}{A}{x}
-    }
-  { \jterm{\Delta}{A}{x}
-    }
-& & \inference
-    { \jctxeq{\Gamma}{\Delta}
-      \jtermeq{\Gamma}{A}{x}{y}
-      }
-    { \jtermeq{\Delta}{A}{x}{y}
-      }
-    \\
-& \inference
-  { \jfameq{\Gamma}{A}{B}
-    \jterm{\Gamma}{A}{x}
-    }
-  { \jterm{\Gamma}{B}{x}
-    }
-& & \inference
-    { \jfameq{\Gamma}{A}{B}
-      \jtermeq{\Gamma}{A}{x}{y}
-      }
-    { \jtermeq{\Gamma}{B}{x}{y}
-      }
-\end{align*}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{The operations of the theory of contexts, families and terms}
-In this subsection we introduce the notions of
-extension on context and on families, the empty context and the empty families, 
-weakening, 
-substitution and identity terms.
-
-Weakening and substitution are both
-operations of the theory of contexts, families and terms and therefore they will
-have three levels of actions: on contexts (at the appropriate level), families
-and terms.
-
-\begin{comment}
-Generally, 
-each operation should
-have an action on contexts, on families and on terms at the `appropriate level'
-and should be compatible with each of the operations we introduce here. In
-that will become more practical after the theory of contexts, families and terms
-is formulated, this means that when we will write down a type theory that
-interprets itself in \autoref{ttmodels}, each of the notions of
-extension, empty family, weakening, substitution and identity terms are going to be internalized
-as terms of the theory and all terms are going to be compatible with these.
-That vague idea will be our guiding light.
-
-So in principle, the introductions of each of the empty context, extension,
-weakening, substitution and identity terms should come in three parts. In 
-practice, they do not all come in three parts.
-
-Apart from the empty context, there will indeed be an empty family and a term
-of the empty family (which will be deemed unique). But extension will
-not have an explicit action `at the level of families' since extension
-is used to define higher families and the action would have to be trivial
-anyway. The empty context and extension provide the essentials of the rest of
-the theory in the sense that higher families are families over extensions,
-contexts are families over the empty context and the `action on families' of
-any operation applied to families over the empty context is the same as the
-original operation. Thus, when we have introduced the operations of weakening
-and substitution we will see that they also have an `action on the terms of 
-contexts'. We will state the relevant judgmental equality rules for the empty
-context in this section rather than in the next, since they are required to
-state that the action on families of an operation is compatible with the
-original operation.
-\end{comment}
-
-We note upfront that both context extension and family extension can also be
-seen as operations of contexts, families and terms, even though we will not
-introduce them as such. The reason is that extension is used to define what it
-means to be a family over a family over a context (namely, a family over an
-extended context). As such, it will have a trivial action on families and terms
-which will not need to be introduced. We will come back to this point in
-\autoref{extension}.
-
 \subsection{Extension}
 \label{extension}
+
 The operations of extension enable one to consider families over
@@
 simply $Q$ since extension acts trivially at the level of families (and terms).
-
-\subsection{The empty context and the empty families}
-\label{empty}
-We introduce an empty context and an family over $\Gamma$ for every context $\Gamma$. 
-It has been suggested by some to only include empty families and not an
-empty context because an empty context is not necessary, but we do have several
-reasons to include them. Having an empty context requires also rules asserting 
-that a context is the same thing as a family over the empty context and this
-gives the categorical structure on contexts for free once one has it for 
-families. The main ingredients that will be missing from the theory once an
-empty context is avoided are weakening by a context and identity morphisms from
-a context to itself. Including these by hand also requires to formulate all the
-compatibility rules involving weakening once more for the cases of weakening by
-a context and identity terms at contexts. We prefer to state these rules once
-and only once and including an empty context helps in this respect.
-
-We also prefer to have our scheme of compatibility rules as symmetrical as
-possible. The structure of type dependency should look exactly the same in the
-default case as in any context. In that respect, having an empty family but not
-an empty context seems a bit odd. Also, some sets of compatibility rules 
-(like the rules stating that extension is compatible with the empty families)
-will become assymetrical as a result of not including an empty context. Moreover,
-we would eventually like to include a stratification of the theory by means of
-a type judgment (asserting that a family in a context $\Gamma$ is a type) and
-study closed types (i.e.~types in the empty context). It would be possible to
-provide a notion of closed types without having an empty context, but this would
-have to be formulated separately and we would have to restate all the rules for
-types (if any) for closed types all over again.
-
-One of the main uses of the empty context and the empty families is that we
-get the property that the `action on contexts' of an operation is compatible
-with its `action on families'.
-
-The empty family over a context $\Gamma$ is introduced by the following rule
-inference rule:
-\begin{align}
-& \inference
-  { }
-  { \jctx{\emptyc}
-    }
-  \\
-& \inference
-  { \jctx{\Gamma}
-    }
-  { \jfam{\Gamma}{\emptyf[\Gamma]}
-    }
-  \\
-& \inference
-  { \jctxeq{\Gamma}{\Gamma'}
-    }
-  { \jfameq{\Gamma}{\emptyf[\Gamma]}{\emptyf[\Gamma']}
-    }
-\end{align}
-
-By regarding contexts as families of contexts over the empty context, we
-enable ourselves also to speak of terms of contexts. A term of a context
-$\Gamma$ is a term of the family $\Gamma$ over the empty context. These ideas
-are captured in the following convertibility rules:
-\begin{align}
-& \inference
-  { \jctx{\Gamma}
-    }
-  { \jfam{\emptyc}{\Gamma}
-    } 
-& & \inference
-    { \jfam{\emptyc}{\Gamma}
-      }
-    { \jctx{\Gamma}
-      }
-    \\
-& \inference
-  { \jctxeq{\Gamma}{\Delta}
-    }
-  { \jfameq{\emptyc}{\Gamma}{\Delta}
-    }
-& & \inference
-    { \jfameq{\emptyc}{\Gamma}{\Delta}
-      }
-    { \jctxeq{\Gamma}{\Delta}
-      }
-\end{align}
-
-The reader may wonder whether the empty family $\emptyf$ always has a
-term. This shall follow from the rules stating the compatibility of extension
-with the empty families in \autoref{comp-0e} below and from
-identity terms (\autoref{identityterms}).
-\begin{comment}
-When contexts are viewed as statements, the empty context is the statement that
-asserts nothing. Likewise, in any context one can assert nothing and this is
-achieved by the empty family. 
-Since somebody who states nothing asserts it all right, there is a term of the empty
-family over any context. Moreover, any two terms of the empty family are always
-judgmentally equal.
-\begin{align}
-& \inference
-  { }
-  { \jctx{\emptyc}
-    }
-  \\
-& \inference
-  { \jctx{\Gamma}
-    }
-  { \jfam{\Gamma}{\emptyf[\Gamma]}
-    }
-  \\
-& \inference
-  { \jctx{\Gamma}
-    }
-  { \jterm{\Gamma}{\emptyf[\Gamma]}{\emptytm[\Gamma]}
-    }
-  \\
-& \inference
-  { \jterm{\Gamma}{\emptyf[\Gamma]}{x}
-    }
-  { \jtermeq
-      {\Gamma}
-      {\emptyf[\Gamma]}
-      {x}
-      {\emptytm[\Gamma]}
-    }
-  \label{comp-00-t}
-\end{align}
-
-The empty context and the empty families together with the operation of
-extension will provide several shortcuts for the rest
-of the theory of families and terms. This is (in part) obtained by requiring
-that a context is exactly the same thing as a family of contexts over the
-empty context and that judgmental equality of contexts is exactly the same
-as judgmental equality of families of contexts over the empty context.
-
-\label{comp-00}
-Because every context is a family of contexts over the empty context, 
-the empty context $\emptyc$ may be considered as a family of contexts over the empty
-context. When we do this, we get $\emptyf[\emptyc]$.
-\begin{equation}\label{eq:comp-00}
-\inference
-  { }
-  { \jfameq
-      {\emptyc}
-      {\emptyc}
-      {\emptyf[\emptyc]}
-    }
-\end{equation}
-In the future, we shall denote $\emptyf[\Gamma]$ by $\emptyf$. The above rule
-guarantees that this will not cause confusion. Likewise, we shall denote
-$\emptytm[\Gamma]$ by $\emptytm$.
-\end{comment}
-
-\subsubsection{Compatibility of extension with the empty context and families}
-In the following set of inference rules we state that the empty context and
-the empty family are neutral objects for both context extension (the first two
-rules) and family extension (the last two rules).
-\label{comp-e0}\label{comp-0e}
-\begin{align}
-& \inference
-  { \jctx{\Gamma}
-    }
-  { \jctxeq{\ctxext{\emptyc}{\Gamma}}{\Gamma}
-    }
-  \label{comp-0e-c}
-  \\
-& \inference
-  { \jctx{\Gamma}
-    }
-  { \jctxeq{\ctxext{\Gamma}{\emptyf}}{\Gamma}
-    }
-  \label{comp-e0-c}\\
-& \inference
-  { \jfam{\Gamma}{A}
-    }
-  { \jfameq{\Gamma}{\ctxext{\emptyf}{A}}{A}
-    }
-  \label{comp-0e-f}
-  \\
-& \inference
-  { \jfam{\Gamma}{A}
-    }
-  { \jfameq{\Gamma}{{A}{\emptyf}}{A}
-    }
-  \label{comp-e0-f}
-\end{align}
-
-\subsubsection{Family extension restricted to the empty context}
-Strictly speaking we should have used a different notation for context extension
-as for family extension, because the following rule asserting that family extension
-in the empty context is the same thing as context extension would look tautological
-without a difference. So let us denote, only for the moment, context extension
-of $\Gamma$ by $A$ by $(\ctxext{\Gamma}{A})^c$ and family extension of $A$ by
-$P$ in context $\Gamma$ by $(\ctxext{A}{P})^\famsym$.
-
-Note that we may consider a context $\Gamma$ as a family over $\emptyc$ and
-a family $\jfam{\Gamma}{A}$ as a family $\jfam{{\emptyc}{\Gamma}}{A}$. 
-Therefore we will require the following rule:
-\begin{align}
-& \inference
-  { \jfam{\Gamma}{A}
-    }
-  { \jctxeq{(\ctxext{\Gamma}{A})^c}{(\ctxext{\Gamma}{A})^\famsym}
-    }
-\end{align}
-Note that this rule actually justifies that we have not utilized two different
-notations for context extension and family extension.
-
-\subsection{The theory of weakening}
-\label{weakening}
-When $A$ is a family in context $\Gamma$, the operation of weakening by $A$
-takes a family $B$ in context $\Gamma$ and provides a family $\ctxwk{A}{B}$
-in context $\ctxext{\Gamma}{A}$. The context family $\ctxwk{A}{B}$ can be seen
-as the constant family over $\ctxext{\Gamma}{A}$ with value $B$. This idea will
-be axiomatized in the cancellation property of weakening and substitution in
-\autoref{cancellation-ws}. In \autoref{morphisms} we will take the terms 
-$\unfold{\jhom{\Gamma}{A}{B}{f}}$ to be the morphisms of families from $A$ to 
-$B$. These will be at the heart of the categorical structure of the theory.
-
-The weakening operation acts on three levels: on contexts, on families and
-on terms. The `action on contexts' of weakening is the action we described
-above: it takes a family $B$ over $\Gamma$ to a family $\ctxwk{A}{B}$ over
-$\ctxext{\Gamma}{A}$; the `action on families' of weakening takes a family
-$Q$ over $\ctxext{\Gamma}{B}$ to a family $\ctxwk[\famsym]{A}{Q}$ over
-$\ctxext{{\Gamma}{A}}{\ctxwk{A}{B}}$; the `action on terms' of weakening takes
-a term $g$ of $Q$ to a term $\ctxwk[\tmsym]{A}{g}$ of $\ctxwk[\famsym]{A}{Q}$.
-\begin{align}
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{\Gamma}{B}
-    }
-  { \jfam{{\Gamma}{A}}{\ctxwk{A}{B}}
-    }
-& & \inference
-    { \jfameq{\Gamma}{A}{A'}
-      \jfameq{\Gamma}{B}{B'}
-      }
-    { \jfameq{{\Gamma}{A}}{\ctxwk{A}{B}}{\ctxwk{A'}{B'}}
-      }
-    \\
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{{\Gamma}{B}}{Q}
-    }
-  { \jfam{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk[\famsym]{A}{Q}}
-    }
-& & \inference
-    { \jfameq{\Gamma}{A}{A'}
-      \jfameq{{\Gamma}{B}}{Q}{Q'}
-      }
-    { \jfameq
-        {{{\Gamma}{A}}{\ctxwk{A}{B}}}
-        {\ctxwk[\famsym]{A}{Q}}
-        {\ctxwk[\famsym]{A'}{Q'}}
-      }
-    \\
-%& \inference
-%  { \jterm{\Gamma}{B}{y}
-%    }
-%  { \jterm{{\Gamma}{A}}{\ctxwk{A}{B}}{\ctxwk{A}{y}}
-%    }
-%& & \inference
-%    { \jtermeq{\Gamma}{B}{y}{y'}
-%      }
-%    { \jtermeq
-%        {{\Gamma}{A}}
-%        {\ctxwk{A}{B}}
-%        {\ctxwk{A}{y}}
-%        {\ctxwk{A'}{y'}}
-%      }
-%    \\
-& \inference
-  { \jfam{\Gamma}{A}
-    \jterm{{\Gamma}{B}}{Q}{g}
-    }
-  { \jterm{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk[\famsym]{A}{Q}}{\ctxwk[\tmsym]{A}{g}}
-    }
-& & \inference
-    { \jfameq{\Gamma}{A}{A'}
-      \jtermeq{{\Gamma}{B}}{Q}{g}{g'}
-      }
-    { \jtermeq
-        {{{\Gamma}{A}}{\ctxwk{A}{B}}}
-        {\ctxwk[\famsym]{A}{Q}}
-        {\ctxwk[\tmsym]{A}{g}}
-        {\ctxwk[\tmsym]{A'}{g'}}
-      }
-\end{align}
-
-\subsubsection{Weakening of an empty family}
-The following inference rules express that when the empty family is
-weakened, the result is the empty family.
-\label{comp-w0}
-\begin{align}
-%& \inference
-%  { \jctx{\Gamma}
-%    }
-%  { \jfameq{\Gamma}{\ctxwk{\Gamma}{\emptyc}}{\emptyf}
-%    }
-%  \\
-& \inference
-  { \jfam{\Gamma}{A}
-    }
-  { \jfameq{{\Gamma}{A}}{\ctxwk{A}{\emptyf}}{\emptyf}
-    }
-  \label{comp-w0-c}\\
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{\Gamma}{B}
-    }
-  { \jfameq
-    {{{\Gamma}{A}}{\ctxwk{A}{B}}}
-    {\ctxwk[\famsym]{A}{\emptyf}}
-    {\emptyf}
-    }
-  \label{comp-w0-f}
-\end{align}
-\begin{comment}
-Note that we do not have to axiomatize the action of weakening
-by $A$ on the term of the empty family, because the result will be in the empty 
-family and every term in the empty family is judgmentally equal to the term 
-$\emptytm$.
-\end{comment}
-
-Because a family over $\Gamma$ is the same as a family over 
-$\ctxext{\Gamma}{\emptyf}$ we can apply both the action on contexts and the
-action on families of weakening to a family $B$ over $\Gamma$. When we apply
-the action on families, we obtain a family $\ctxwk[\famsym]{A}{B}$ over the
-context $\ctxext{{\Gamma}{A}}{\ctxwk{A}{\emptyf}}$. However, since we have
-postulated the judgmental equalities $\ctxwk{A}{\emptyf}\jdeq\emptyf$ and
-$\ctxext{{\Gamma}{A}}{\emptyf}\jdeq\ctxext{\Gamma}{A}$, we see that we can
-compare $\ctxwk[\famsym]{A}{B}$ with $\ctxwk{A}{B}$. The following inference
-rule postulates that these two are judgmentally equal:
-\begin{equation}
-\inference
-{ \jfam{\Gamma}{A}
-  \jfam{\Gamma}{B}
-  }
-{ \jfameq{{\Gamma}{A}}{\ctxwk[\famsym]{A}{B}}{\ctxwk{A}{B}}
-  }
-\end{equation}
-Due to this rule, the action on contexts and the action on families of weakening
-are compatible with each other and consequently there can be no possible
-confusion when we omit the annotations $\famsym$ and $\tmsym$ alltogether. In
-the future, the weakening of a family $Q$ over $\ctxext{\Gamma}{B}$ shall
-be denoted just by $\ctxwk{A}{Q}$ and likewise the weakening of a term $g$ of
-$Q$ shall be denoted by $\ctxwk{A}{g}$.
-
-Because a family $B$ over $\Gamma$ can be treated as a family by the operation
-of weakening, weakening also acts on the terms of $B$. The weakening of a term
-$y$ of $B$ by $A$ can be seen as the constant term (or function) of the
-family $\ctxwk{A}{B}$.
-
-\subsubsection{Weakenings of extensions}
-\label{comp-we}
-The following rules assert that when an extended family is weakened, the
-weakening distributes over the extension factors.
-\begin{align}
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{{\Gamma}{B}}{Q}
-    }
-  { \jfameq
-      {\ctxext{\Gamma}{A}}
-      {\ctxwk{A}{\ctxext{B}{Q}}}
-      {\ctxext{\ctxwk{A}{B}}{\ctxwk{A}{Q}}}
-    }
-  \label{comp-we-c}
-  \\
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{{{\Gamma}{B}}{Q}}{R}
-    }
-  { \jfameq
-      {\ctxext{{\Gamma}{A}}{\ctxwk{A}{B}}}
-      {\ctxwk{A}{\ctxext{Q}{R}}}
-      {\ctxext{\ctxwk{A}{Q}}{\ctxwk{A}{R}}}
-    }
-  \label{comp-we-f}
-\end{align}
-When thinking of terms of $\ctxwk{A}{B}$ as morphisms of families from $A$ to
-$B$, this looks already like form of type theoretic choice. It is weaker in that
-it is not stated with function types, yet it is stronger in that it states a
-judgmental equality between two families.
-
-There is also a version of this property where an extended term is weakened.
-This variant is stated and proved in \autoref{comp-we-t}.
-
-\subsubsection{Weakening of weakenings}
-Weakening by a family $A$ in context $\Gamma$ brings things in context $\Gamma$
-to things in context $\ctxext{\Gamma}{A}$. Since we have all the ingredients of
-the theory of contexts, families and terms in the context $\Gamma$ as well it
-has in particular it's own weakening by families. Suppose we have a family
-$\jfam{{\Gamma}{B}}{Q}$ over $B$ in context $\Gamma$. Weakening by $Q$ brings
-things from context $\ctxext{\Gamma}{B}$ to $\ctxext{{\Gamma}{B}}{Q}$. Thus
-we can provide rules asserting what will happen when we first weaken by $Q$ and
-then (via the action on families) by $A$. We will require the following
-inference rules to be valid:
-\label{comp-ww}
-\begin{align}
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{{\Gamma}{B}}{Q}
-    \jfam{{\Gamma}{B}}{R}
-    }
-  { \jfameq
-      {{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{Q}}}
-      {\ctxwk{A}{{Q}{R}}}
-      {\ctxwk{{A}{Q}}{{A}{R}}}
-    }
-  \label{comp-ww-c}\\
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{{\Gamma}{B}}{Q}
-    \jfam{{{\Gamma}{B}}{R}}{S}
-    }
-  { \jfameq
-      {{{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{Q}}}{\ctxwk{A}{{Q}{R}}}}
-      {\ctxwk{A}{{Q}{S}}}
-      {\ctxwk{{A}{Q}}{{A}{S}}}
-    }
-  \label{comp-ww-f}\\
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{{\Gamma}{B}}{Q}
-    \jterm{{{\Gamma}{B}}{R}}{S}{k}
-    }
-  { \jtermeq
-      {{{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{Q}}}{\ctxwk{A}{{Q}{R}}}}
-      {\ctxwk{A}{{Q}{S}}}
-      {\ctxwk{A}{{Q}{k}}}
-      {\ctxwk{{A}{Q}}{{A}{k}}}
-    }
-  \label{comp-ww-t}
-\end{align}
-
-\begin{rmk}
-As an important special case of these inference rules we have the following
-valid inference rules:
-\begin{align*}
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{\Gamma}{B}
-    \jfam{\Gamma}{C}
-    }
-  { \jfameq
-      {{{\Gamma}{A}}{\ctxwk{A}{B}}}
-      {\ctxwk{A}{{B}{C}}}
-      {\ctxwk{{A}{B}}{{A}{C}}}
-    }
-  \\
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{\Gamma}{B}
-    \jfam{{\Gamma}{C}}{R}
-    }
-  { \jfameq
-      {{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{{B}{C}}}}
-      {\ctxwk{A}{{B}{R}}}
-      {\ctxwk{{A}{B}}{{A}{R}}}
-    }
-  \\
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{\Gamma}{B}
-    \jterm{{\Gamma}{C}}{R}{h}
-    }
-  { \jtermeq
-      {{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{{B}{C}}}}
-      {\ctxwk{A}{{B}{R}}}
-      {\ctxwk{A}{{B}{h}}}
-      {\ctxwk{{A}{B}}{{A}{h}}}
-    }
-\end{align*}
-Moreover, we have
-\begin{equation*}
-\inference
-  { \jfam{\Gamma}{A}
-    \jfam{\Gamma}{B}
-    \jterm{\Gamma}{C}{z}
-    }
-  { \jtermeq
-      {{{\Gamma}{A}}{\ctxwk{A}{B}}}
-      {\ctxwk{A}{{B}{C}}}
-      {\ctxwk{A}{{B}{z}}}
-      {\ctxwk{{A}{B}}{{A}{z}}}
-    }
-\end{equation*}
-\end{rmk}
-
-\subsubsection{Weakening by the empty family}
-Note that we can also weaken by the empty family over $\Gamma$.
-Weakening by the empty family $\emptyf$ over a context $\Gamma$ leaves families, 
-their terms, families over those families and terms of those unchanged:
-\label{comp-0w}
-\begin{align}
-& \inference
-  { \jfam{\Gamma}{B}
-    }
-  { \jfameq{\Gamma}{\ctxwk{\emptyf}{B}}{B}
-    }
-  \label{comp-0w-c}
-  \\
-& \inference
-  { \jfam{{\Gamma}{B}}{Q}
-    }
-  { \jfameq{{\Gamma}{B}}{\ctxwk{\emptyf}{Q}}{Q}
-    }
-  \label{comp-0w-f}
-  \\
-& \inference
-  { \jterm{{\Gamma}{B}}{Q}{g}
-    }
-  { \jtermeq{{\Gamma}{B}}{Q}{\ctxwk{\emptyf}{g}}{g}
-    }
-  \label{comp-0w-t}
-\end{align}
-
-\subsubsection{Currying for weakening}
-\label{comp-ew}
-The rules expressing that extension is compatible with weakening assert that
-weakening by an extension is the same thing as weakening twice in the
-appropriate way.
-\begin{align}
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{{\Gamma}{A}}{P}
-    \jfam{\Gamma}{B}
-    }
-  { \jfameq
-      {{{\Gamma}{A}}{P}}
-      {\ctxwk{\ctxext{A}{P}}{B}}
-      {\ctxwk{P}{{A}{B}}}
-    }
-  \label{comp-ew-c}\\
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{{\Gamma}{A}}{P}
-    \jfam{{\Gamma}{B}}{Q}
-    }
-  { \jfameq
-      {{{{\Gamma}{A}}{P}}{\ctxwk{P}{{A}{B}}}}
-      {\ctxwk{\ctxext{A}{P}}{Q}}
-      {\ctxwk{P}{{A}{Q}}}
-    }
-  \label{comp-ew-f}\\
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{{\Gamma}{A}}{P}
-    \jterm{{\Gamma}{B}}{Q}{g}
-    }
-  { \jtermeq
-      {{{{\Gamma}{A}}{P}}{\ctxwk{P}{{A}{B}}}}
-      {\ctxwk{P}{{A}{Q}}}
-      {\ctxwk{\ctxext{A}{P}}{g}}
-      {\ctxwk{P}{{A}{g}}}
-    } 
-  \label{comp-ew-t}
-\end{align}
-
-\begin{comment}
-\subsubsection{Weakening by a context}
-Since we have not required that there is an empty context with the property that
-a context is the same thing as a family in the empty context, we will have to
-explicitly introduce weakening by contexts. Since it will allways be clear
-whether something is weakened by a context or by a family, we shall use the
-same notation for weakening by contexts as for weakening by families. We shall
-have to repeat all the compatibility rules for weakening by families for
-weakening by contexts; there is no way around it. It is
-actually this sort of inconvenience that has originally led us to include an
-empty context and be done with it.
-
-In the following inference rules we assume that we have a context $\Gamma$ and
-that $\jctxeq{\Gamma}{\Gamma'}$ in the rules where $\Gamma'$ makes an appearence.
-\begin{align}
-& \inference
-  { \jctx{\Delta}
-    }
-  { \jfam{\Gamma}{\ctxwk{\Gamma}{\Delta}}
-    }
-& & \inference
-  { \jctxeq{\Delta}{\Delta'}
-    }
-  { \jfameq{\Gamma}{\ctxwk{\Gamma}{\Delta}}{\ctxwk{\Gamma'}{\Delta'}}
-    }
-  \\
-& \inference
-  { \jfam{\Delta}{B}
-    }
-  { \jfam{\ctxext{\Gamma}{\ctxwk{\Gamma}{\Delta}}}{\ctxwk{\Gamma}{B}}
-    }
-& & \inference
-  { \jfameq{\Delta}{B}{B'}
-    }
-  { \jfameq
-      {\ctxext{\Gamma}{\ctxwk{\Gamma}{\Delta}}}
-      {\ctxwk{\Gamma}{B}}
-      {\ctxwk{\Gamma'}{B'}}
-    }
-  \\
-& \inference
-  { \jterm{\Delta}{B}{y}
-    }
-  { \jterm
-      {\ctxext{\Gamma}{\ctxwk{\Gamma}{\Delta}}}
-      {\ctxwk{\Gamma}{B}}
-      {\ctxwk{\Gamma}{y}}
-    }
-& & \inference
-  { \jtermeq{\Delta}{B}{y}{y'}
-    }
-  { \jtermeq
-      {\ctxext{\Gamma}{\ctxwk{\Gamma}{\Delta}}}
-      {\ctxwk{\Gamma}{B}}
-      {\ctxwk{\Gamma}{y}}
-      {\ctxwk{\Gamma}{y'}}
-    }
-\end{align}
-Note that we do not employ three different notations for the three different
-actions of weakening by a context because there can be no confusion about which
-of the three actions is meant. If there were an empty context we should have
-distinguished the three as we did for weakening by a family.
-
-We shall have similar compatibility rules for weakening by a context as we did
-for weakening by a family.
-
-{\color{red}To do. It will look exactly the same as for weakening by families}
-\end{comment}
-
-\subsection{The theory of projections}
-\label{identityterms}
-We formulate the theory of projections on top of the theory of weakening
-described in \autoref{weakening}. In the theory of projections we introduce
-the identity terms. The identity term at a family $A$ in context $\Gamma$ is
-a term $\jterm{{\Gamma}{A}}{\ctxwk{A}{A}}{\idtm{A}}$, so weakening is necessary.
-Together with weakening, the identity terms will provide for the projections.
-The projection from $\cftext{A}{P}$ to $A$ will be the term
-$\jterm{{\Gamma}{{A}{P}}}{\ctxwk{\ctxext{A}{P}}{A}}{\ctxwk{P}{\idtm{A}}}$ and
-will be studied extensively in \autoref{extension-on-terms}.
-
-The following inference rules are required to be valid:
-\begin{align}
-& \inference
-  { \jfam{\Gamma}{A}
-    }
-  { \unfoldall{\jhom{\Gamma}{A}{A}{\idtm{A}}}
-    }
-& & \inference
-    { \jfameq{\Gamma}{A}{A'}
-      }
-    { \unfoldall{\jhomeq{\Gamma}{A}{A}{\idtm{A}}{\idtm{A'}}}
-      }
-\end{align}
-
-\begin{comment}
-We stress again
-that because we have not required there to be an empty context, we need to
-explicitly call for identity terms $\idtm{\Gamma}$ of $\ctxwk{\Gamma}{\Gamma}$
-in context $\Gamma$ for each context $\Gamma$. Thus, we will also require
-the inference rule
-\begin{align}
-& \inference
-  { \jctx{\Gamma}
-    }
-  { \jterm{\Gamma}{\ctxwk{\Gamma}{\Gamma}}{\idtm{\Gamma}}
-    }
-& &
-  \inference
-  { \jctxeq{\Gamma}{\Gamma'}
-    }
-  { \jtermeq{\Gamma}{\ctxwk{\Gamma}{\Gamma}}{\idtm{\Gamma}}{\idtm{\Gamma'}}
-    }
-\end{align}
-For the following inference rules, there should also be versions of identity
-terms at contexts.
-\end{comment}
-
-\subsubsection{The identity term of an extended family}
-The fact that identity terms are compatible with extensions is derived in
-\autoref{comp-ie}.
-
-\subsubsection{The identity term of a weakened family}
-\label{comp-wi}
-The identity term of a weakened family is the weakened identity term:
-\begin{equation}
-\inference
-  { \jfam{\Gamma}{A}
-    \jfam{{\Gamma}{B}}{Q}
-    }
-  { \unfoldall{\jhomeq
-      {{{\Gamma}{A}}{\ctxwk{A}{B}}}
-      {\ctxwk{A}{Q}}
-      {\ctxwk{A}{Q}}
-      {\ctxwk{A}{\idtm{Q}}}
-      {\idtm{\ctxwk{A}{Q}}}
-      }
-    }
-  \label{comp-wi-t}
-\end{equation}
-
-\begin{rmk}
-An important special case is the judgmental equality
-\begin{equation*}
-\unfoldall{\jhomeq
-      {{\Gamma}{A}}
-      {\ctxwk{A}{B}}
-      {\ctxwk{A}{B}}
-      {\ctxwk{A}{\idtm{B}}}
-      {\idtm{\ctxwk{A}{B}}}
-      }
-\end{equation*}
-for a family $\jfam{\Gamma}{B}$.
-\end{rmk}
-
-\subsection{The theory of substitution}
-\label{substitution}
-The theory of substitution can be formulated separately from the theory of
-weakening. Therefore, we will only require the theory of extension and the
-empty context and families. In \autoref{sec:esystem-equalities} we will
-provide the additional judgmental equality rules that will be required when
-combining the theories of substitution and of projections.
-
-Given a family $P$ over $A$ and a term $x$ of $A$, substitution gives a way to
-consider the \emph{fiber $\subst{x}{P}$ of $P$ at $x$}. As was the case with
-weakening, the substitution operation comes in three parts: the `action on
-contexts' of substitution is the operation just described; the `action on
-families' of substitution takes a family $Q$ over $\ctxext{{\Gamma}{A}}{P}$
-to a family $\subst[\famsym]{x}{Q}$ over $\ctxext{\Gamma}{\subst{x}{P}}$; the
-`action on terms' of substitution takes a term $g$ of $Q$ over
-$\ctxext{{\Gamma}{A}}{P}$ to a term $\subst[\tmsym]{x}{g}$ of 
-$\subst[\famsym]{x}{Q}$.
-\begin{align}
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jfam{{\Gamma}{A}}{P}
-    }
-  { \jfam{\Gamma}{\subst{x}{P}}
-    }
-& & \inference
-    { \jtermeq{\Gamma}{A}{x}{x'}
-      \jfameq{{\Gamma}{A}}{P}{P'}
-      }
-    { \jfameq{\Gamma}{\subst{x}{P}}{\subst{x'}{P'}}
-      }
-    \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jfam{{{\Gamma}{A}}{P}}{Q}
-    }
-  { \jfam{{\Gamma}{\subst{x}{P}}}{\subst[\famsym]{x}{Q}}
-    }
-& & \inference
-    { \jtermeq{\Gamma}{A}{x}{x'}
-      \jfameq{{{\Gamma}{A}}{P}}{Q}{Q'}
-      }
-    { \jfameq{{\Gamma}{\subst{x}{P}}}{\subst[\famsym]{x}{Q}}{\subst[\famsym]{x'}{Q'}}
-      }
-    \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jterm{{{\Gamma}{A}}{P}}{Q}{g}
-    }
-  { \jterm{{\Gamma}{\subst{x}{P}}}{\subst[\famsym]{x}{Q}}{\subst[\tmsym]{x}{g}}
-    }
-& & \inference
-    { \jtermeq{\Gamma}{A}{x}{x'}
-      \jtermeq{{{\Gamma}{A}}{P}}{Q}{g}{g'}
-      }
-    { \jtermeq
-        {{\Gamma}{\subst{x}{P}}}
-        {\subst[\famsym]{x}{Q}}
-        {\subst[\tmsym]{x}{g}}
-        {\subst[\tmsym]{x'}{g'}}
-      }
-\end{align}
-
-\subsubsection{Fibers of an empty family}
-The following inference rules establish that the fibers of the empty family are 
-the empty families:
-\label{comp-s0}
-\begin{align}
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    }
-  { \jfameq{\Gamma}{\subst{x}{\emptyf}}{\emptyf}
-    }
-  \label{comp-s0-c}
-  \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jfam{{\Gamma}{A}}{P}
-    }
-  { \jfameq
-      {{\Gamma}{\subst{x}{P}}}
-      {\subst{x}{\emptyf}}
-      {\emptyf}
-    }
-  \label{comp-s0-f}
-\end{align}
-\begin{comment}
-Note again that we do not have to axiomatize the action of substitution by $x$ on the term
-of the empty family, because the result will be in the empty family and every
-term in the empty family is judgmentally equal to the term $\emptytm$.
-\end{comment}
-
-We use the above rule to state the compatibility of the action on families of
-substitution with the action on contexts of substitution. Note that a family
-$P$ over $\ctxext{\Gamma}{A}$ may be regarded as a family over
-$\ctxext{{\Gamma}{A}}{\emptyf}$. Thus, we may consider the family
-$\subst[\famsym]{x}{P}$ over $\ctxext{\Gamma}{\subst{x}{\emptyf}}$. Since
-$\subst{x}{\emptyf}$ is judgmentally equal to the empty family, we may compare
-$\subst[\famsym]{x}{P}$ with $\subst{x}{P}$:
-\begin{equation}
-\inference
-{ \jfam{{\Gamma}{A}}{P}
-  }
-{ \jfameq{\Gamma}{\subst[\famsym]{x}{P}}{\subst{x}{P}}
-  }
-\end{equation}
-Due to this rule we need not make the annotations $\famsym$ and $\tmsym$ in
-the notation for substitution anymore and thus we shall omit them from now on.
-Note that because a family $P$ over $\ctxext{\Gamma}{A}$ is eligible for
-application of the action on families of substitution, we may also substitute
-terms of $P$. Thus, given terms $\jterm{\Gamma}{A}{x}$ and $\jterm{{\Gamma}{A}}{P}{f}$,
-we get a term $\jterm{\Gamma}{\subst{x}{P}}{\subst{x}{f}}$, the \emph{value of
-$f$ at $x$}.
-
-In \autoref{morphisms} we will use a combination of weakening and substitution
-to define composition of morphisms of families. However, we have to rely
-on the cancellation rule stated in \autoref{cancellation-ws} before we can
-meaningfully state the definition of composition.
-
-\subsubsection{Fibers of extensions}
-\label{comp-se}
-The following inference rules assert that if we take the fiber of an extended
-family at a term $x$ of $A$ in context $\Gamma$, the substitution by $x$
-distributes over the factors of the extension.
-\begin{align}
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jfam{{{\Gamma}{A}}{P}}{Q}
-    }
-  { \jfameq
-      {\Gamma}
-      {\subst{x}{\ctxext{P}{Q}}}
-      {\ctxext{\subst{x}{P}}{\subst{x}{Q}}}
-    }
-  \label{comp-se-c}
-  \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jfam{{{{\Gamma}{A}}{P}}{Q}}{R}
-    }
-  { \jfameq
-      {{\Gamma}{\subst{x}{P}}}
-      {\subst{x}{\ctxext{Q}{R}}}
-      {\ctxext{\subst{x}{Q}}{\subst{x}{R}}}
-    }
-  \label{comp-se-f}
-\end{align}
-There is also a version of this statement in which extended terms are considered.
-This variant is stated and proved in \autoref{comp-se-t}.
-
-\subsubsection{Fibers of fibers}
-\label{comp-ss}
-
-The following rules explain what happens when we first substitute by a term
-$g$ of $Q$ over $P$ in context $\ctxext{\Gamma}{A}$ and then by a term $x$ of
-$A$:
-\begin{align}
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jterm{{{\Gamma}{A}}{P}}{Q}{g}
-    \jfam{{{{\Gamma}{A}}{P}}{Q}}{R}
-    }
-  { \jfameq
-      {{{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}}
-      {\subst{x}{{g}{R}}}
-      {\subst{{x}{g}}{{x}{R}}}
-    }
-  \label{comp-ss-c}
-  \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jterm{{{\Gamma}{A}}{P}}{Q}{g}
-    \jfam{{{{{\Gamma}{A}}{P}}{Q}}{R}}{S}
-    }
-  { \jfameq
-      {{{{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}}{\subst{x}{{g}{R}}}}
-      {\subst{x}{{g}{S}}}
-      {\subst{{x}{g}}{{x}{S}}}
-    }
-  \label{comp-ss-f}
-  \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jterm{{{\Gamma}{A}}{P}}{Q}{g}
-    \jterm{{{{{\Gamma}{A}}{P}}{Q}}{R}}{S}{k}
-    }
-  { \jtermeq
-      {{{{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}}{\subst{x}{{g}{R}}}}
-      {\subst{x}{{g}{S}}}
-      {\subst{x}{{g}{k}}}
-      {\subst{{x}{g}}{{x}{k}}}
-    }
-  \label{comp-ss-t}
-\end{align}
-
-\begin{rmk}
-As an important special case of these inference rules we have the following
-valid inference rules:
-\begin{align*}
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jterm{{\Gamma}{A}}{P}{f}
-    \jfam{{{\Gamma}{A}}{P}}{Q}
-    }
-  { \jfameq
-      {\Gamma}
-      {\subst{x}{{f}{Q}}}
-      {\subst{{x}{f}}{{x}{Q}}}
-    }
-  \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jterm{{\Gamma}{A}}{P}{f}
-    \jfam{{{{\Gamma}{A}}{P}}{Q}}{R}
-    }
-  { \jfameq
-      {{\Gamma}{\subst{x}{{f}{Q}}}}
-      {\subst{x}{{f}{R}}}
-      {\subst{{x}{f}}{{x}{R}}}
-    }
-  \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jterm{{\Gamma}{A}}{P}{f}
-    \jterm{{{{\Gamma}{A}}{P}}{Q}}{R}{h}
-    }
-  { \jtermeq
-      {{\Gamma}{\subst{x}{{f}{Q}}}}
-      {\subst{x}{{f}{R}}}
-      {\subst{x}{{f}{h}}}
-      {\subst{{x}{f}}{{x}{h}}}
-    }
-\end{align*}
-Moreover, we get
-\begin{equation*}
-\inference
-  { \jterm{\Gamma}{A}{x}
-    \jterm{{\Gamma}{A}}{P}{f}
-    \jterm{{{\Gamma}{A}}{P}}{Q}{g}
-    }
-  { \jtermeq
-      {\Gamma}
-      {\subst{x}{{f}{Q}}}
-      {\subst{x}{{f}{g}}}
-      {\subst{{x}{f}}{{x}{g}}}
-    }
-\end{equation*}
-\end{rmk}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Joining the theory of projections with the theory of substitution}
-\label{sec:esystem-equalities}
-
-\begin{comment}
-We have already seen that empty families, extension, weakening, substitution
-and the identity terms are compatible with each other. There are, however, a
-few more judgmental equalities explaining what happens when
-\begin{enumerate}
-\item something is weakend by the empty family. This is used to consider a term
-      $x$ of $A$ in context $\Gamma$ as a morphism from $\emptyf$ to $A$ in
-      context $\Gamma$;
-\item something is weakened by an extended family. This is Currying for weakening;
-\item something is first weakened by a family $A$ in context $\Gamma$ and then
-      substituted by a term of $A$. This is used to show that composition is
-      compatible with itself (i.e. that composition is associative);
-\item the identity terms are joining the party. Identity terms are empty contexts
-      in certain models, so they should behave accordingly.
-\end{enumerate}
-In other words, the rules seem to not fit in the scheme that all operations are
-compatible with each other, but in fact each of these rules plays a role in
-completing the scheme (work out further).
-\end{comment}
-
-\begin{comment}
-In this section we extend the judgmental equality relation in such a way that
-the empty context, extension, weakening, substitution and the identity terms
-all become compatible with each other. Since we have five operations, there
-are going to be $25$ finite sets of rules explaining how each operation is
-compatible with each of the operations.
-
-In addition to these $25$ finite sets
-of compatibility rules, there are also two sets of rules explaining the
-cancellation properties of weakening and substitution and of the identity terms.
-
-A typical compatibility axiom captures the idea that a diagram like
-\begin{equation*}
-\begin{tikzcd}[column sep=huge]
-\jfam{\Gamma}{\blank} 
-  \ar{r}{C\mapsto\ctxwk{B}{C}} 
-  \ar{d}[swap]{C\mapsto\ctxwk{A}{C}} 
-& \jfam{{\Gamma}{B}}{\blank} 
-  \ar{d}{Q\mapsto\ctxwk{A}{Q}}
-  \\
-\jfam{{\Gamma}{A}}{\blank} 
-  \ar{r}[swap]{P\mapsto\ctxwk{{A}{B}}{P}} 
-& \jfam{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\blank}
-\end{tikzcd}
-\end{equation*}
-in the case of two consecutive weakenings, or
-\begin{equation*}
-\begin{tikzcd}[column sep=huge]
-\jfam{{\Gamma}{A}}{\blank} 
-  \ar{d}[swap]{Q\mapsto\subst{x}{Q}} 
-  \ar{r}{Q\mapsto\ctxwk{P}{Q}} 
-& \jfam{{{\Gamma}{A}}{P}}{\blank} 
-    \ar{d}{R\mapsto\subst{x}{R}}
-  \\ 
-\jfam{\Gamma}{\blank} 
-  \ar{r}[swap]{B\mapsto\ctxwk{\subst{x}{P}}{B}} 
-& \jfam{{\Gamma}{\subst{x}{P}}}{\blank}
-\end{tikzcd}
-\end{equation*}
-in the case of a combination of weakening and substitution, commutes. We won't
-actually draw any of these diagrams in this section, but readers are encouraged
-to do so if they think it can help them understanding the axioms. Another thing
-one could keep in mind with the axioms that involve weakening is that all of
-them become completely obvious when weakening is made notationally invisible.
-Of course, we have an explicit notation for weakening on purpuse, but it is the
-usual practice of type theorists to make weakening invisible.
-
-The table below indicates where
-we can find the axioms of each of the combinations of operators. The reference
-at the row of operation $X$ and the column of operation $Y$ refers to the
-axiom which explains $Y$ distributes over the arguments of $X$ when $Y$ is
-applied after $X$. The empty context has no arguments, whereas extension,
-and weakening have two arguments which are both of the family-kind, substitution
-has two arguments of which one is a term and the other is a family and
-idenitity terms have one argument, a family.
-\begin{center}
-\begin{tabular}{r|rrrr}
-& ext & wk & subst & id\\
-\hline
-extension & \autoref{comp-ee}* &  & \cmark & \cmark \\
-weakening & \autoref{comp-we}* & \autoref{comp-ww} & \autoref{comp-sw} & \autoref{comp-wi} \\
-substitution & \autoref{comp-se}* & \autoref{comp-ws} & \autoref{comp-ss} & \autoref{comp-si} \\
-identity terms & \cmark & \autoref{comp-wi} & \autoref{comp-si} & \xmark
-\end{tabular}
-\end{center}
-Apart from references to sections where axioms are given, we also see checkmarks,
-crosses, marked references and duplicated references in this table. 
-The checkmarks indicate that the desired property
-can be proved and need not be axiomatized. To clarify those:
-\begin{itemize}
-\item The identity term of the empty context is the unique term of the empty
-family. Recall that we already have axiomatized in \autoref{empty} that every
-term of the empty family is judgmentally equal to $\emptytm$.
-\item The identity term of an extension can be explained by means of the notion
-of \emph{extension on terms}. This notion is defined in \autoref{extension-on-terms}
-using only identity terms and substitution. With extension on terms we can
-combine a term $\jterm{\Gamma}{A}{x}$ and a term $\jterm{\Gamma}{\subst{x}{P}}{u}$
-into a term $\jterm{\Gamma}{{A}{P}}{\tmext{x}{u}}$, which may be seen as the
-pair consisting of $x$ and $u$. Since it is a defined notion, we will
-be able to derive lots of properties for extension of terms. One of them explains
-how we can see the identity term of an extension as the extension of two certain
-terms, see \autoref{comp-ie}.
-\item Using extension on terms we can also make sense of `fibers at an extension'
-because that would exactly be the fiber at an extended term. We will see in
-\autoref{comp-es} that, for instance, $\subst{\tmext{x}{u}}{Q}\jdeq\subst{u}{{x}{Q}}$
-for any family $\jfam{{{\Gamma}{A}}{P}}{Q}$ and any two terms $\jterm{\Gamma}{A}{x}$
-and $\jterm{\Gamma}{\subst{x}{P}}{u}$. That's a form of currying.
-\end{itemize}
-The spots with a \xmark simply make no sense 
-{\color{red}(or do they, when the term model
-is better understood (by me)?)}. The marked references point to
-axioms which \emph{do} have a variant for terms, but those require the notion
-of extension on terms, can be proved and therefore need not be included in our
-list of compatibility axioms. In each of these sections, we give a forward
-reference to the lemma in which this variant is stated and proved.
-\end{comment}
-
-\subsubsection{Fibers of weakenings}\label{comp-sw}
-The following rules assert what happens when we first weaken by a family
-$Q$ over $P$ in context $\ctxext{\Gamma}{A}$ and then substitute by a term
-$x$ of $A$:
-\begin{align}
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jfam{{{\Gamma}{A}}{P}}{Q}
-    \jfam{{{\Gamma}{A}}{P}}{R}
-    }
-  { \jfameq
-      {{{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}}
-      {\subst{x}{\ctxwk{Q}{R}}}
-      {\ctxwk{\subst{x}{Q}}{\subst{x}{R}}}
-    }
-  \label{comp-sw-c}
-  \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jfam{{{\Gamma}{A}}{P}}{Q}
-    \jfam{{{{\Gamma}{A}}{P}}{R}}{S}
-    }
-  { \jfameq
-      {{{{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}}{\subst{x}{\ctxwk{Q}{R}}}}
-      {\subst{x}{\ctxwk{Q}{S}}}
-      {\ctxwk{\subst{x}{Q}}{\subst{x}{S}}}
-    }
-  \label{comp-sw-f}
-  \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jfam{{{\Gamma}{A}}{P}}{Q}
-    \jterm{{{{\Gamma}{A}}{P}}{R}}{S}{k}
-    }
-  { \jtermeq
-      {{{{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}}{\subst{x}{\ctxwk{Q}{R}}}}
-      {\subst{x}{\ctxwk{Q}{S}}}
-      {\subst{x}{\ctxwk{Q}{k}}}
-      {\ctxwk{\subst{x}{Q}}{\subst{x}{k}}}
-    }
-  \label{comp-sw-t}
-\end{align}
-
-\begin{rmk}
-As an important special case of these inference rules we have the following
-valid inference rules:
-\begin{align*}
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jfam{{\Gamma}{A}}{P}
-    \jfam{{\Gamma}{A}}{Q}
-    }
-  { \jfameq
-      {{\Gamma}{\subst{x}{P}}}
-      {\subst{x}{\ctxwk{P}{Q}}}
-      {\ctxwk{\subst{x}{P}}{\subst{x}{Q}}}
-    }
-  \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jfam{{\Gamma}{A}}{P}
-    \jfam{{{\Gamma}{A}}{Q}}{R}
-    }
-  { \jfameq
-      {{{\Gamma}{\subst{x}{P}}}{\subst{x}{\ctxwk{P}{Q}}}}
-      {\subst{x}{\ctxwk{P}{R}}}
-      {\ctxwk{\subst{x}{P}}{\subst{x}{R}}}
-    }
-  \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jfam{{\Gamma}{A}}{P}
-    \jterm{{{\Gamma}{A}}{Q}}{R}{h}
-    }
-  { \jtermeq
-      {{{\Gamma}{\subst{x}{P}}}{\subst{x}{\ctxwk{P}{Q}}}}
-      {\subst{x}{\ctxwk{P}{R}}}
-      {\subst{x}{\ctxwk{P}{h}}}
-      {\ctxwk{\subst{x}{P}}{\subst{x}{h}}}
-    }
-\end{align*}
-Moreover, we get
-\begin{equation*}
-\inference
-  { \jterm{\Gamma}{A}{x}
-    \jfam{{\Gamma}{A}}{P}
-    \jterm{{\Gamma}{A}}{Q}{g}
-    }
-  { \jtermeq
-      {{\Gamma}{\subst{x}{P}}}
-      {\subst{x}{\ctxwk{P}{Q}}}
-      {\subst{x}{\ctxwk{P}{g}}}
-      {\ctxwk{\subst{x}{P}}{\subst{x}{g}}}
-    }
-\end{equation*}
-\end{rmk}
-
-\subsubsection{Weakenings of fibers}
-\label{comp-ws}
-The following inference rules explain what happens when we first weaken by a
-term $g$ of $Q$ in context $\ctxext{\Gamma}{B}$ and then weaken by a family
-$A$ over $\Gamma$.
-\begin{align}
-& \inference
-  { \jfam{\Gamma}{A}
-    \jterm{{\Gamma}{B}}{Q}{g}
-    \jfam{{{\Gamma}{B}}{Q}}{R}
-    }
-  { \jfameq
-      {{{\Gamma}{A}}{\ctxwk{A}{B}}}
-      {\ctxwk{A}{\subst{g}{R}}}
-      {\subst{\ctxwk{A}{g}}{\ctxwk{A}{R}}}
-    }
-  \label{comp-ws-c}
-  \\
-& \inference
-  { \jfam{\Gamma}{A}
-    \jterm{{\Gamma}{B}}{Q}{g}
-    \jfam{{{{\Gamma}{B}}{Q}}{R}}{S}
-    }
-  { \jfameq
-      {{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{\subst{g}{R}}}}
-      {\ctxwk{A}{\subst{g}{S}}}
-      {\subst{\ctxwk{A}{g}}{\ctxwk{A}{S}}}
-    }
-  \label{comp-ws-f}
-  \\
-& \inference
-  { \jfam{\Gamma}{A}
-    \jterm{{\Gamma}{B}}{Q}{g}
-    \jterm{{{{\Gamma}{B}}{Q}}{R}}{S}{k}
-    }
-  { \jtermeq
-      {{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{\subst{g}{R}}}}
-      {\ctxwk{A}{\subst{g}{S}}}
-      {\ctxwk{A}{\subst{g}{k}}}
-      {\subst{\ctxwk{A}{g}}{\ctxwk{A}{k}}}
-    }
-  \label{comp-ws-t}
-\end{align}
-
-\begin{rmk}
-As an important special case of these inference rules we have the following 
-valid inference rules:
-\begin{align*}
-& \inference
-  { \jfam{\Gamma}{A}
-    \jterm{\Gamma}{B}{y}
-    \jfam{{\Gamma}{B}}{Q}
-    }
-  { \jfameq
-      {{\Gamma}{A}}
-      {\ctxwk{A}{\subst{y}{Q}}}
-      {\subst{\ctxwk{A}{y}}{\ctxwk{A}{Q}}}
-    }
-  \\
-& \inference
-  { \jfam{\Gamma}{A}
-    \jterm{\Gamma}{B}{y}
-    \jfam{{{\Gamma}{B}}{Q}}{R}
-    }
-  { \jfameq
-      {{{\Gamma}{A}}{\ctxwk{A}{\subst{y}{Q}}}}
-      {\ctxwk{A}{\subst{y}{R}}}
-      {\subst{\ctxwk{A}{y}}{\ctxwk{A}{R}}}
-    }
-  \\
-& \inference
-  { \jfam{\Gamma}{A}
-    \jterm{\Gamma}{B}{y}
-    \jterm{{{\Gamma}{B}}{Q}}{R}{h}
-    }
-  { \jtermeq
-      {{{\Gamma}{A}}{\ctxwk{A}{\subst{y}{Q}}}}
-      {\ctxwk{A}{\subst{y}{R}}}
-      {\ctxwk{A}{\subst{y}{h}}}
-      {\subst{\ctxwk{A}{y}}{\ctxwk{A}{h}}}
-    }
-\end{align*}
-Moreover, we get
-\begin{equation*}
-\inference
-  { \jfam{\Gamma}{A}
-    \jterm{\Gamma}{B}{y}
-    \jterm{{\Gamma}{B}}{Q}{g}
-    }
-  { \jtermeq
-      {{\Gamma}{A}}
-      {\ctxwk{A}{\subst{y}{Q}}}
-      {\ctxwk{A}{\subst{y}{g}}}
-      {\subst{\ctxwk{A}{y}}{\ctxwk{A}{g}}}
-    }
-\end{equation*}
-\end{rmk}
-
-\subsubsection{The cancellation property of weakening and substitution}
-\label{cancellation-ws}
-The judgmental equalities we're about to describe assert that substituting a term
-in the weakening a thing gives you the thing back. In the case of contexts we get that each fiber
-$\subst{x}{\ctxwk{A}{B}}$ is just $B$ and in the case of terms we get 
-that $\ctxwk{A}{y}$ is the constant function
-mapping everything to $y:B$. Thus, these rules actually establish the weakening
-as the weakening. After stating the rules we will describe what it means to
-compose context morphisms (terms of weakened contexts).
-\begin{align}
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{\Gamma}{B}
-    \jterm{\Gamma}{A}{x}
-    }
-  { \jfameq
-      {\Gamma}
-      {\subst{x}{\ctxwk{A}{B}}}
-      {B}
-    }
-  \label{cancellation-ws-c}\\
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{{\Gamma}{B}}{Q}
-    \jterm{\Gamma}{A}{x}
-    }
-  { \jfameq{{\Gamma}{B}}{\subst{x}{\ctxwk{A}{Q}}}{Q}
-    }
-  \label{cancellation-ws-f}\\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jterm{{\Gamma}{B}}{Q}{g}
-    }
-  { \jtermeq{{\Gamma}{B}}{Q}{\subst{x}{\ctxwk{A}{g}}}{g}
-    }
-  \label{cancellation-ws-t}
-\end{align}
-
-\subsubsection{The identity term of a substituted family}
-\label{comp-si}
-The identity term of a substituted family is the substitution of the identity term
-\begin{equation}
-\inference
-  { \jterm{\Gamma}{A}{x}
-    \jfam{{{\Gamma}{A}}{P}}{Q}
-    }
-  { \unfoldall{\jhomeq
-      {{\Gamma}{\subst{x}{P}}}
-      {\subst{x}{Q}}
-      {\subst{x}{Q}}
-      {\subst{x}{\idtm{Q}}}
-      {\idtm{\subst{x}{Q}}}
-    }}
-  \label{comp-si-t}
-\end{equation}
-
-\begin{rmk}
-An important special case is the judgmental equality
-\begin{equation*}
-\unfoldall{\jhomeq
-      {\Gamma}
-      {\subst{x}{P}}
-      {\subst{x}{P}}
-      {\subst{x}{\idtm{P}}}
-      {\idtm{\subst{x}{P}}}
-    }
-\end{equation*}
-for a family $\jfam{{\Gamma}{A}}{P}$.
-\end{rmk}
-
-\subsubsection{The cancellation properties of identity terms}
-\label{cancellation-i}
-Identity terms are determined by their behavior with respect to substitution combined with
-weakening. The identity terms will also be subject to compatibility rules.
-\begin{align}
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    }
-  { \jtermeq{\Gamma}{A}{\subst{x}{\idtm{A}}}{x}
-    }
-  \label{cancellation-si}\\
-& \inference
-  { \jfam{{\Gamma}{A}}{P}
-    }
-  { \jfameq{{\Gamma}{A}}{\subst{\idtm{A}}{\ctxwk{A}{P}}}{P}
-    }
-  \label{precomp-idtm-c}
-  \\
-& \inference
-  { \jfam{{{\Gamma}{A}}{P}}{Q}
-    }
-  { \jfameq{{{\Gamma}{A}}{P}}{\subst{\idtm{A}}{\ctxwk{A}{Q}}}{Q}
-    }
-  \label{precomp-idtm-f}
-  \\
-& \inference
-  { \jterm{{{\Gamma}{A}}{P}}{Q}{g}
-    }
-  { \jtermeq
-      {{{\Gamma}{A}}{P}}
-      {Q}
-      {\subst{\idtm{A}}{\ctxwk{A}{g}}}
-      {g}
-    }
-  \label{precomp-idtm-t}
-\end{align}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{The possiblity of types in the theory of contexts, families and
-terms}
-\label{types}
-We have deliberately not spoken of types so far because we have taken the point
-of view that a type in a context is nothing but a family in that context which
-belongs to the class of types. We think of types as \emph{irreducible} families,
-i.e.\ families which are neither empty nor the extension of two
-families which are both not empty (in algebraic terminology: which
-are both non-trivial). To allow ourselves to speak of types we introduce two
-new judgments: the judgment that something is a type and the judgment that two
-types are equal.
-\begin{align*}
-\jalign\jtype{\Gamma}{A} 
-& \jalign\jtypeeq{\Gamma}{A}{B}
-\end{align*}
-But only families of contexts are eligible to be types. If $A$ is a type
-in context $\Gamma$, then $A$ is also a family of contexts over $\Gamma$. 
-Moreover, two types in context $\Gamma$ are judgmentally equal precisely when they are equal
-as context families and if a family $B$ of contexts over $\Gamma$ is
-judgmentally equal to a type $A$ in context $\Gamma$, then $B$ is a type in
-context $\Gamma$. This is expressed by the following four inference rules:
-\begin{align*}
-& \inference
-  { \jtype{\Gamma}{A}
-    }
-  { \jfam{\Gamma}{A}
-    }
-& & \inference
-    { \jtypeeq{\Gamma}{A}{B}
-      }
-    { \jfameq{\Gamma}{A}{B}
-      }
-    \\
-& \inference
-  { \jtype{\Gamma}{A}
-    \jfameq{\Gamma}{A}{B}
-    }
-  { \jtype{\Gamma}{B}
-    }
-& & \inference
-    { \jtype{\Gamma}{A}
-      \jfameq{\Gamma}{A}{B}
-      }
-    { \jtypeeq{\Gamma}{A}{B}
-      }
-\end{align*}
-As pointed out at the beginning of this subsection, 
-we do not assume that the empty family is a type, that would be like
-assuming that the multiplicative unit of a ring is prime.
-
-\begin{comment}
-We envision a
-categorical interpretation of the theory of contexts, families and terms where 
-contexts are interpreted as finite paths of arrows belonging to a predetermined
-class. The empty family would be such a path of length $0$ whereas the types
-would be such paths of length $1$.
-\end{comment}
-
-We add rules asserting that a weakened type is again a type that 
-substitution preserves the property of being a type:
-\begin{align}
-& \inference
-  { \jfam{\Gamma}{B}
-    \jtype{\ctxext{\Gamma}{B}}{Q}
-    }
-  { \jtype{\ctxext{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{Q}}
-    }
-  \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jtype{\ctxext{{\Gamma}{A}}{P}}{Q}
-    }
-  { \jtype{\ctxext{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}
-    }
-\end{align}
-With only the current rules, the possibility of making the judgment that
-something is a type does not add much to the theory of contexts, families and
-terms. Nevertheless, when studying models, having an interpretation for the
-judgment that something is a type will allow for the possibility to study
-conditions such as the one asserting that every family `factorizes' uniquely
-as multiple applications of extension to types, analogous to the condition on
-unique factorization domains in ring theory. All the tradional models of type
-theory should translate to models with such a condition, simply because
-traditionally contexts are viewed as lists of variable (and type) declarations.

### 2014-10-28 Rules/rules_intro.tex
reorganization

--- 2014-09-20 article-models-tt.tex
@@
-\section{The generalized algebraic theory of contexts, families and terms (the E-system)}\label{tt}
 In this section we give a description of dependent type theory before type
@@
 the types in the empty context as the objects.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Judgments and valid inference rules}\label{judgments}
-The theory we describe here is a theory of contexts, families of
-contexts and terms thereof. The families of contexts are by some authors called
-dependent contexts, but they are handled a bit differently here because they
-become the primary object of study. Dependent contexts can be types; they could
-be seen as atomic or indecomposable dependent contexts.
-
-Thus we make eight kinds of judgments: ``$\Gamma$ is a context'',
-``$A$ is a family of contexts over $\Gamma$'', ``$A$ is a type in context $\Gamma$''
-and ``$x$ is a term of the family $A$ of contexts over $\Gamma$''. The other four
-judgments are for judgmental equality.
-\begin{align*}
-\jalign\jctx{\Gamma} 
-& \jalign\jctxeq{\Gamma}{\Gamma'}
-  \\
-\jalign\jfam{\Gamma}{A} 
-& \jalign\jfameq{\Gamma}{A}{B}
-  \\
-\jalign\jterm{\Gamma}{A}{x} 
-& \jalign\jtermeq{\Gamma}{A}{x}{y}.
-\end{align*}
-
-Strictly speaking, we have three different judgmental equalities in play and one
-could request for a notational difference to signify that fact. For instance,
-we could denote the judgmental equalities of contexts, families and terms by
-$\jdeq_c$, $\jdeq_f$ and $\jdeq_t$ respectively. It will, however, always be
-clear which of the three kinds of judgmental equality is meant when we assert
-a judgmental equality and therefore we shall not bother to make this notational
-distinction.
-
-We note that what we call families over contexts
-here could also have been named dependent contexts or telescopes, see
-\cite{deBruijn1991,hofmann1995extensional}. The term family is in agreement
-with the terminology scheme of \cite{TheBook}, though the reader should be
-warned that the notion of familie means something slightly different there than
-it does here.
-
-The theory of contexts, families and terms shall be described by means of a
-declaration of valid inference rules. An inference rule is an expression of the
-form
-\begin{equation*}
-\inference{\mathcal{H}_1\quad\cdots\quad\mathcal{H}_n}{\mathcal{J}}
-\end{equation*}
-where each of $\mathcal{H}_1$, ..., $\mathcal{H}_n$ and $\mathcal{J}$ is 
-judgment. The judgments $\mathcal{H}_1$, ..., $\mathcal{H}_n$ are called the
-hypotheses and the judgment $\mathcal{J}$ is the conclusion of the inference
-rule. When a judgment is the conclusion of a valid inference rule, we also say
-that the judgment is valid or well-formed. In fact, the only way to construct
-new judgments is via the valid inference rules we describe in this section.
-
-The class of valid inference rules is closed under composition and under
-rearranging the hypotheses. That
-means that if
-\begin{equation*}
-\inference{\mathcal{I}_{1k}\quad\cdots\quad\mathcal{I}_{m_kk}}{\mathcal{H}_k}
-\end{equation*}
-and
-\begin{equation*}
-\inference{\mathcal{H}_1\quad\cdots\quad\mathcal{H}_n}{\mathcal{J}}
-\end{equation*}
-are valid inference rules for $1\leq k\leq n$, then
-\begin{equation*}
-\inference{\mathcal{I}_{11}\quad\cdots\quad\mathcal{I}_{m_11}\quad\cdots\quad\mathcal{I}_{1n}\quad\cdots\quad\mathcal{I}_{m_nn}}{\mathcal{J}}
-\end{equation*}
-is also a valid inference rule. Also, in a valid inference rule 
-duplicated hypotheses may always be contracted and hypotheses may always be 
-reordered; the result is still considered a valid inference rule.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{The basic rules for judgmental equality}
-The rules for judgmental equality establish that it is an equivalence relation
-in all three cases (contexts, families and terms). Thus, the following inference
-rules shall be required to be valid:
-\bgroup\small
-\begin{align*}
-& \inference
-  { \jctx{\Gamma}
-    }
-  { \jctxeq{\Gamma}{\Gamma}
-    } 
-& & \inference
-    { \jctxeq{\Gamma}{\Delta}
-      }
-    { \jctxeq{\Delta}{\Gamma}
-      } 
-& & \inference
-    { \jctxeq{\Gamma}{\Delta}
-      \jctxeq{\Delta}{\greek{E}}
-      }
-    { \jctxeq{\Gamma}{\greek{E}}
-      }
-    \\
-& \inference
-  { \jfam{\Gamma}{A}
-    }
-  { \jfameq{\Gamma}{A}{A}
-    } 
-& & \inference
-    { \jfameq{\Gamma}{A}{B}
-      }
-    { \jfameq{\Gamma}{B}{A}
-      }
-& & \inference
-    { \jfameq{\Gamma}{A}{B}
-      \jfameq{\Gamma}{B}{C}
-      }
-    { \jfameq{\Gamma}{A}{C}
-      }
-    \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    }
-  { \jtermeq{\Gamma}{A}{x}{x}
-    }
-& & \inference
-    { \jtermeq{\Gamma}{A}{x}{y}
-      }
-    { \jtermeq{\Gamma}{A}{y}{x}
-      }
-& & \inference
-    { \jtermeq{\Gamma}{A}{x}{y}
-      \jtermeq{\Gamma}{A}{y}{z}
-      }
-    { \jtermeq{\Gamma}{A}{x}{z}
-      }
-\end{align*}
-\egroup
-
-The following convertibility rules are responsible for the strictness
-of judgmental equality, which sets it apart from equivalences or identifications:
-\begin{align*}
-& \inference
-  { \jctxeq{\Gamma}{\Delta}
-    \jfam{\Gamma}{A}
-    }
-  { \jfam{\Delta}{A}
-    }
-& & \inference
-    { \jctxeq{\Gamma}{\Delta}
-      \jfameq{\Gamma}{A}{B}
-      }
-    { \jfameq{\Delta}{A}{B}
-      }
-    \\
-& \inference
-  { \jctxeq{\Gamma}{\Delta}
-    \jterm{\Gamma}{A}{x}
-    }
-  { \jterm{\Delta}{A}{x}
-    }
-& & \inference
-    { \jctxeq{\Gamma}{\Delta}
-      \jtermeq{\Gamma}{A}{x}{y}
-      }
-    { \jtermeq{\Delta}{A}{x}{y}
-      }
-    \\
-& \inference
-  { \jfameq{\Gamma}{A}{B}
-    \jterm{\Gamma}{A}{x}
-    }
-  { \jterm{\Gamma}{B}{x}
-    }
-& & \inference
-    { \jfameq{\Gamma}{A}{B}
-      \jtermeq{\Gamma}{A}{x}{y}
-      }
-    { \jtermeq{\Gamma}{B}{x}{y}
-      }
-\end{align*}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{The operations of the theory of contexts, families and terms}
-In this subsection we introduce the notions of
-extension on context and on families, the empty context and the empty families, 
-weakening, 
-substitution and identity terms.
-
-Weakening and substitution are both
-operations of the theory of contexts, families and terms and therefore they will
-have three levels of actions: on contexts (at the appropriate level), families
-and terms.
-
-\begin{comment}
-Generally, 
-each operation should
-have an action on contexts, on families and on terms at the `appropriate level'
-and should be compatible with each of the operations we introduce here. In
-that will become more practical after the theory of contexts, families and terms
-is formulated, this means that when we will write down a type theory that
-interprets itself in \autoref{ttmodels}, each of the notions of
-extension, empty family, weakening, substitution and identity terms are going to be internalized
-as terms of the theory and all terms are going to be compatible with these.
-That vague idea will be our guiding light.
-
-So in principle, the introductions of each of the empty context, extension,
-weakening, substitution and identity terms should come in three parts. In 
-practice, they do not all come in three parts.
-
-Apart from the empty context, there will indeed be an empty family and a term
-of the empty family (which will be deemed unique). But extension will
-not have an explicit action `at the level of families' since extension
-is used to define higher families and the action would have to be trivial
-anyway. The empty context and extension provide the essentials of the rest of
-the theory in the sense that higher families are families over extensions,
-contexts are families over the empty context and the `action on families' of
-any operation applied to families over the empty context is the same as the
-original operation. Thus, when we have introduced the operations of weakening
-and substitution we will see that they also have an `action on the terms of 
-contexts'. We will state the relevant judgmental equality rules for the empty
-context in this section rather than in the next, since they are required to
-state that the action on families of an operation is compatible with the
-original operation.
-\end{comment}
-
-We note upfront that both context extension and family extension can also be
-seen as operations of contexts, families and terms, even though we will not
-introduce them as such. The reason is that extension is used to define what it
-means to be a family over a family over a context (namely, a family over an
-extended context). As such, it will have a trivial action on families and terms
-which will not need to be introduced. We will come back to this point in
-\autoref{extension}.
-
-\subsection{Extension}
-\label{extension}
-The operations of extension enable one to consider families over
-families and higher level families alike. We will need both a context extension
-operation and a family extension operation.
-
-The inference rules introducing context and family extension are as follows:
-\begin{align}
-& \inference
-  { \jfam{\Gamma}{A}
-    }
-  { \jctx{\ctxext{\Gamma}{A}}
-    }
-& & \inference
-    { \jctxeq{\Gamma}{\Delta}
-      \jfameq{\Gamma}{A}{B}
-      }
-    { \jctxeq{\ctxext{\Gamma}{A}}{\ctxext{\Delta}{B}}
-      }
-    \\
-& \inference
-  { \jfam{{\Gamma}{A}}{P}
-    }
-  { \jfam{\Gamma}{\ctxext{A}{P}}
-    }
-& & \inference
-    { \jfameq{\Gamma}{A}{B} 
-      \jfameq{{\Gamma}{A}}{P}{Q}
-      }
-    { \jfameq{\Gamma}{\ctxext{A}{P}}{\ctxext{B}{Q}}
-      }
-\end{align}
-The extension operation may also be defined to act on families over families
-and terms thereof. For instance, When $\jfam{{\Gamma}{A}}{P}$ is a family,
-we would get a family $\jfam{{\Gamma}{A}}{{\Gamma}{P}}$ and when
-$\jterm{{\Gamma}{A}}{P}{f}$ is a term, we would get a term
-$\jterm{{\Gamma}{A}}{{\Gamma}{P}}{\ctxext{\Gamma}{f}}$. Then there would be
-axioms stating the judgmental equalities $\ctxext{\Gamma}{P}\jdeq P$ and
-$\ctxext{\Gamma}{f}\jdeq f$. Thus, it is of no actual use to introduce the
-action of context extension on families and terms.
-
-In \autoref{extension-on-terms} we will also define an extension operation
-on terms, which defines a term $\jterm{\Gamma}{{A}{P}}{\tmext{x}{u}}$ for
-every $\jterm{\Gamma}{A}{x}$ and $\jterm{\Gamma}{\subst{x}{P}}{u}$. This
-operation is defined by using the substitution operation and identity terms.
-
-\subsubsection{Associativity of extension}
-\label{comp-ee}
-The inference rules asserting that extension is compatible with itself assert
-that contexts are unstructured lists of type declarations. This rule is
-unavoidable if we want that for a family $A$ in context $\Gamma$, a family over
-$A$ is the same thing as a family over $\ctxext{\Gamma}{A}$. 
-\begin{align}
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{{\Gamma}{A}}{P}
-    }
-  { \jctxeq{\ctxext{{\Gamma}{A}}{P}}{\ctxext{\Gamma}{{A}{P}}}
-    }
-  \\
-& \inference
-  { \jfam{{\Gamma}{A}}{P}
-    \jfam{{{\Gamma}{A}}{P}}{Q}
-    }
-  { \jfameq{\Gamma}{\ctxext{A}{{P}{Q}}}{\ctxext{{A}{P}}{Q}}
-    }
-  \label{comp-ee-c}
-\end{align}
-Note that we may take the point of view that extension distributes over
-extensions. If we want to extend an extended family $\ctxext{P}{Q}$ over
-$\ctxext{\Gamma}{A}$ by the family $A$, the result is the extension of (1)
-the extension of $P$ by $A$ and (2) the extension of $Q$ by $A$. The latter is
-simply $Q$ since extension acts trivially at the level of families (and terms).
-
-\subsection{The empty context and the empty families}
-\label{empty}
-We introduce an empty context and an family over $\Gamma$ for every context $\Gamma$. 
-It has been suggested by some to only include empty families and not an
-empty context because an empty context is not necessary, but we do have several
-reasons to include them. Having an empty context requires also rules asserting 
-that a context is the same thing as a family over the empty context and this
-gives the categorical structure on contexts for free once one has it for 
-families. The main ingredients that will be missing from the theory once an
-empty context is avoided are weakening by a context and identity morphisms from
-a context to itself. Including these by hand also requires to formulate all the
-compatibility rules involving weakening once more for the cases of weakening by
-a context and identity terms at contexts. We prefer to state these rules once
-and only once and including an empty context helps in this respect.
-
-We also prefer to have our scheme of compatibility rules as symmetrical as
-possible. The structure of type dependency should look exactly the same in the
-default case as in any context. In that respect, having an empty family but not
-an empty context seems a bit odd. Also, some sets of compatibility rules 
-(like the rules stating that extension is compatible with the empty families)
-will become assymetrical as a result of not including an empty context. Moreover,
-we would eventually like to include a stratification of the theory by means of
-a type judgment (asserting that a family in a context $\Gamma$ is a type) and
-study closed types (i.e.~types in the empty context). It would be possible to
-provide a notion of closed types without having an empty context, but this would
-have to be formulated separately and we would have to restate all the rules for
-types (if any) for closed types all over again.
-
-One of the main uses of the empty context and the empty families is that we
-get the property that the `action on contexts' of an operation is compatible
-with its `action on families'.
-
-The empty family over a context $\Gamma$ is introduced by the following rule
-inference rule:
-\begin{align}
-& \inference
-  { }
-  { \jctx{\emptyc}
-    }
-  \\
-& \inference
-  { \jctx{\Gamma}
-    }
-  { \jfam{\Gamma}{\emptyf[\Gamma]}
-    }
-  \\
-& \inference
-  { \jctxeq{\Gamma}{\Gamma'}
-    }
-  { \jfameq{\Gamma}{\emptyf[\Gamma]}{\emptyf[\Gamma']}
-    }
-\end{align}
-
-By regarding contexts as families of contexts over the empty context, we
-enable ourselves also to speak of terms of contexts. A term of a context
-$\Gamma$ is a term of the family $\Gamma$ over the empty context. These ideas
-are captured in the following convertibility rules:
-\begin{align}
-& \inference
-  { \jctx{\Gamma}
-    }
-  { \jfam{\emptyc}{\Gamma}
-    } 
-& & \inference
-    { \jfam{\emptyc}{\Gamma}
-      }
-    { \jctx{\Gamma}
-      }
-    \\
-& \inference
-  { \jctxeq{\Gamma}{\Delta}
-    }
-  { \jfameq{\emptyc}{\Gamma}{\Delta}
-    }
-& & \inference
-    { \jfameq{\emptyc}{\Gamma}{\Delta}
-      }
-    { \jctxeq{\Gamma}{\Delta}
-      }
-\end{align}
-
-The reader may wonder whether the empty family $\emptyf$ always has a
-term. This shall follow from the rules stating the compatibility of extension
-with the empty families in \autoref{comp-0e} below and from
-identity terms (\autoref{identityterms}).
-\begin{comment}
-When contexts are viewed as statements, the empty context is the statement that
-asserts nothing. Likewise, in any context one can assert nothing and this is
-achieved by the empty family. 
-Since somebody who states nothing asserts it all right, there is a term of the empty
-family over any context. Moreover, any two terms of the empty family are always
-judgmentally equal.
-\begin{align}
-& \inference
-  { }
-  { \jctx{\emptyc}
-    }
-  \\
-& \inference
-  { \jctx{\Gamma}
-    }
-  { \jfam{\Gamma}{\emptyf[\Gamma]}
-    }
-  \\
-& \inference
-  { \jctx{\Gamma}
-    }
-  { \jterm{\Gamma}{\emptyf[\Gamma]}{\emptytm[\Gamma]}
-    }
-  \\
-& \inference
-  { \jterm{\Gamma}{\emptyf[\Gamma]}{x}
-    }
-  { \jtermeq
-      {\Gamma}
-      {\emptyf[\Gamma]}
-      {x}
-      {\emptytm[\Gamma]}
-    }
-  \label{comp-00-t}
-\end{align}
-
-The empty context and the empty families together with the operation of
-extension will provide several shortcuts for the rest
-of the theory of families and terms. This is (in part) obtained by requiring
-that a context is exactly the same thing as a family of contexts over the
-empty context and that judgmental equality of contexts is exactly the same
-as judgmental equality of families of contexts over the empty context.
-
-\label{comp-00}
-Because every context is a family of contexts over the empty context, 
-the empty context $\emptyc$ may be considered as a family of contexts over the empty
-context. When we do this, we get $\emptyf[\emptyc]$.
-\begin{equation}\label{eq:comp-00}
-\inference
-  { }
-  { \jfameq
-      {\emptyc}
-      {\emptyc}
-      {\emptyf[\emptyc]}
-    }
-\end{equation}
-In the future, we shall denote $\emptyf[\Gamma]$ by $\emptyf$. The above rule
-guarantees that this will not cause confusion. Likewise, we shall denote
-$\emptytm[\Gamma]$ by $\emptytm$.
-\end{comment}
-
-\subsubsection{Compatibility of extension with the empty context and families}
-In the following set of inference rules we state that the empty context and
-the empty family are neutral objects for both context extension (the first two
-rules) and family extension (the last two rules).
-\label{comp-e0}\label{comp-0e}
-\begin{align}
-& \inference
-  { \jctx{\Gamma}
-    }
-  { \jctxeq{\ctxext{\emptyc}{\Gamma}}{\Gamma}
-    }
-  \label{comp-0e-c}
-  \\
-& \inference
-  { \jctx{\Gamma}
-    }
-  { \jctxeq{\ctxext{\Gamma}{\emptyf}}{\Gamma}
-    }
-  \label{comp-e0-c}\\
-& \inference
-  { \jfam{\Gamma}{A}
-    }
-  { \jfameq{\Gamma}{\ctxext{\emptyf}{A}}{A}
-    }
-  \label{comp-0e-f}
-  \\
-& \inference
-  { \jfam{\Gamma}{A}
-    }
-  { \jfameq{\Gamma}{{A}{\emptyf}}{A}
-    }
-  \label{comp-e0-f}
-\end{align}
-
-\subsubsection{Family extension restricted to the empty context}
-Strictly speaking we should have used a different notation for context extension
-as for family extension, because the following rule asserting that family extension
-in the empty context is the same thing as context extension would look tautological
-without a difference. So let us denote, only for the moment, context extension
-of $\Gamma$ by $A$ by $(\ctxext{\Gamma}{A})^c$ and family extension of $A$ by
-$P$ in context $\Gamma$ by $(\ctxext{A}{P})^\famsym$.
-
-Note that we may consider a context $\Gamma$ as a family over $\emptyc$ and
-a family $\jfam{\Gamma}{A}$ as a family $\jfam{{\emptyc}{\Gamma}}{A}$. 
-Therefore we will require the following rule:
-\begin{align}
-& \inference
-  { \jfam{\Gamma}{A}
-    }
-  { \jctxeq{(\ctxext{\Gamma}{A})^c}{(\ctxext{\Gamma}{A})^\famsym}
-    }
-\end{align}
-Note that this rule actually justifies that we have not utilized two different
-notations for context extension and family extension.
-
-\subsection{The theory of weakening}
-\label{weakening}
-When $A$ is a family in context $\Gamma$, the operation of weakening by $A$
-takes a family $B$ in context $\Gamma$ and provides a family $\ctxwk{A}{B}$
-in context $\ctxext{\Gamma}{A}$. The context family $\ctxwk{A}{B}$ can be seen
-as the constant family over $\ctxext{\Gamma}{A}$ with value $B$. This idea will
-be axiomatized in the cancellation property of weakening and substitution in
-\autoref{cancellation-ws}. In \autoref{morphisms} we will take the terms 
-$\unfold{\jhom{\Gamma}{A}{B}{f}}$ to be the morphisms of families from $A$ to 
-$B$. These will be at the heart of the categorical structure of the theory.
-
-The weakening operation acts on three levels: on contexts, on families and
-on terms. The `action on contexts' of weakening is the action we described
-above: it takes a family $B$ over $\Gamma$ to a family $\ctxwk{A}{B}$ over
-$\ctxext{\Gamma}{A}$; the `action on families' of weakening takes a family
-$Q$ over $\ctxext{\Gamma}{B}$ to a family $\ctxwk[\famsym]{A}{Q}$ over
-$\ctxext{{\Gamma}{A}}{\ctxwk{A}{B}}$; the `action on terms' of weakening takes
-a term $g$ of $Q$ to a term $\ctxwk[\tmsym]{A}{g}$ of $\ctxwk[\famsym]{A}{Q}$.
-\begin{align}
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{\Gamma}{B}
-    }
-  { \jfam{{\Gamma}{A}}{\ctxwk{A}{B}}
-    }
-& & \inference
-    { \jfameq{\Gamma}{A}{A'}
-      \jfameq{\Gamma}{B}{B'}
-      }
-    { \jfameq{{\Gamma}{A}}{\ctxwk{A}{B}}{\ctxwk{A'}{B'}}
-      }
-    \\
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{{\Gamma}{B}}{Q}
-    }
-  { \jfam{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk[\famsym]{A}{Q}}
-    }
-& & \inference
-    { \jfameq{\Gamma}{A}{A'}
-      \jfameq{{\Gamma}{B}}{Q}{Q'}
-      }
-    { \jfameq
-        {{{\Gamma}{A}}{\ctxwk{A}{B}}}
-        {\ctxwk[\famsym]{A}{Q}}
-        {\ctxwk[\famsym]{A'}{Q'}}
-      }
-    \\
-%& \inference
-%  { \jterm{\Gamma}{B}{y}
-%    }
-%  { \jterm{{\Gamma}{A}}{\ctxwk{A}{B}}{\ctxwk{A}{y}}
-%    }
-%& & \inference
-%    { \jtermeq{\Gamma}{B}{y}{y'}
-%      }
-%    { \jtermeq
-%        {{\Gamma}{A}}
-%        {\ctxwk{A}{B}}
-%        {\ctxwk{A}{y}}
-%        {\ctxwk{A'}{y'}}
-%      }
-%    \\
-& \inference
-  { \jfam{\Gamma}{A}
-    \jterm{{\Gamma}{B}}{Q}{g}
-    }
-  { \jterm{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk[\famsym]{A}{Q}}{\ctxwk[\tmsym]{A}{g}}
-    }
-& & \inference
-    { \jfameq{\Gamma}{A}{A'}
-      \jtermeq{{\Gamma}{B}}{Q}{g}{g'}
-      }
-    { \jtermeq
-        {{{\Gamma}{A}}{\ctxwk{A}{B}}}
-        {\ctxwk[\famsym]{A}{Q}}
-        {\ctxwk[\tmsym]{A}{g}}
-        {\ctxwk[\tmsym]{A'}{g'}}
-      }
-\end{align}
-
-\subsubsection{Weakening of an empty family}
-The following inference rules express that when the empty family is
-weakened, the result is the empty family.
-\label{comp-w0}
-\begin{align}
-%& \inference
-%  { \jctx{\Gamma}
-%    }
-%  { \jfameq{\Gamma}{\ctxwk{\Gamma}{\emptyc}}{\emptyf}
-%    }
-%  \\
-& \inference
-  { \jfam{\Gamma}{A}
-    }
-  { \jfameq{{\Gamma}{A}}{\ctxwk{A}{\emptyf}}{\emptyf}
-    }
-  \label{comp-w0-c}\\
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{\Gamma}{B}
-    }
-  { \jfameq
-    {{{\Gamma}{A}}{\ctxwk{A}{B}}}
-    {\ctxwk[\famsym]{A}{\emptyf}}
-    {\emptyf}
-    }
-  \label{comp-w0-f}
-\end{align}
-\begin{comment}
-Note that we do not have to axiomatize the action of weakening
-by $A$ on the term of the empty family, because the result will be in the empty 
-family and every term in the empty family is judgmentally equal to the term 
-$\emptytm$.
-\end{comment}
-
-Because a family over $\Gamma$ is the same as a family over 
-$\ctxext{\Gamma}{\emptyf}$ we can apply both the action on contexts and the
-action on families of weakening to a family $B$ over $\Gamma$. When we apply
-the action on families, we obtain a family $\ctxwk[\famsym]{A}{B}$ over the
-context $\ctxext{{\Gamma}{A}}{\ctxwk{A}{\emptyf}}$. However, since we have
-postulated the judgmental equalities $\ctxwk{A}{\emptyf}\jdeq\emptyf$ and
-$\ctxext{{\Gamma}{A}}{\emptyf}\jdeq\ctxext{\Gamma}{A}$, we see that we can
-compare $\ctxwk[\famsym]{A}{B}$ with $\ctxwk{A}{B}$. The following inference
-rule postulates that these two are judgmentally equal:
-\begin{equation}
-\inference
-{ \jfam{\Gamma}{A}
-  \jfam{\Gamma}{B}
-  }
-{ \jfameq{{\Gamma}{A}}{\ctxwk[\famsym]{A}{B}}{\ctxwk{A}{B}}
-  }
-\end{equation}
-Due to this rule, the action on contexts and the action on families of weakening
-are compatible with each other and consequently there can be no possible
-confusion when we omit the annotations $\famsym$ and $\tmsym$ alltogether. In
-the future, the weakening of a family $Q$ over $\ctxext{\Gamma}{B}$ shall
-be denoted just by $\ctxwk{A}{Q}$ and likewise the weakening of a term $g$ of
-$Q$ shall be denoted by $\ctxwk{A}{g}$.
-
-Because a family $B$ over $\Gamma$ can be treated as a family by the operation
-of weakening, weakening also acts on the terms of $B$. The weakening of a term
-$y$ of $B$ by $A$ can be seen as the constant term (or function) of the
-family $\ctxwk{A}{B}$.
-
-\subsubsection{Weakenings of extensions}
-\label{comp-we}
-The following rules assert that when an extended family is weakened, the
-weakening distributes over the extension factors.
-\begin{align}
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{{\Gamma}{B}}{Q}
-    }
-  { \jfameq
-      {\ctxext{\Gamma}{A}}
-      {\ctxwk{A}{\ctxext{B}{Q}}}
-      {\ctxext{\ctxwk{A}{B}}{\ctxwk{A}{Q}}}
-    }
-  \label{comp-we-c}
-  \\
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{{{\Gamma}{B}}{Q}}{R}
-    }
-  { \jfameq
-      {\ctxext{{\Gamma}{A}}{\ctxwk{A}{B}}}
-      {\ctxwk{A}{\ctxext{Q}{R}}}
-      {\ctxext{\ctxwk{A}{Q}}{\ctxwk{A}{R}}}
-    }
-  \label{comp-we-f}
-\end{align}
-When thinking of terms of $\ctxwk{A}{B}$ as morphisms of families from $A$ to
-$B$, this looks already like form of type theoretic choice. It is weaker in that
-it is not stated with function types, yet it is stronger in that it states a
-judgmental equality between two families.
-
-There is also a version of this property where an extended term is weakened.
-This variant is stated and proved in \autoref{comp-we-t}.
-
-\subsubsection{Weakening of weakenings}
-Weakening by a family $A$ in context $\Gamma$ brings things in context $\Gamma$
-to things in context $\ctxext{\Gamma}{A}$. Since we have all the ingredients of
-the theory of contexts, families and terms in the context $\Gamma$ as well it
-has in particular it's own weakening by families. Suppose we have a family
-$\jfam{{\Gamma}{B}}{Q}$ over $B$ in context $\Gamma$. Weakening by $Q$ brings
-things from context $\ctxext{\Gamma}{B}$ to $\ctxext{{\Gamma}{B}}{Q}$. Thus
-we can provide rules asserting what will happen when we first weaken by $Q$ and
-then (via the action on families) by $A$. We will require the following
-inference rules to be valid:
-\label{comp-ww}
-\begin{align}
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{{\Gamma}{B}}{Q}
-    \jfam{{\Gamma}{B}}{R}
-    }
-  { \jfameq
-      {{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{Q}}}
-      {\ctxwk{A}{{Q}{R}}}
-      {\ctxwk{{A}{Q}}{{A}{R}}}
-    }
-  \label{comp-ww-c}\\
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{{\Gamma}{B}}{Q}
-    \jfam{{{\Gamma}{B}}{R}}{S}
-    }
-  { \jfameq
-      {{{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{Q}}}{\ctxwk{A}{{Q}{R}}}}
-      {\ctxwk{A}{{Q}{S}}}
-      {\ctxwk{{A}{Q}}{{A}{S}}}
-    }
-  \label{comp-ww-f}\\
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{{\Gamma}{B}}{Q}
-    \jterm{{{\Gamma}{B}}{R}}{S}{k}
-    }
-  { \jtermeq
-      {{{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{Q}}}{\ctxwk{A}{{Q}{R}}}}
-      {\ctxwk{A}{{Q}{S}}}
-      {\ctxwk{A}{{Q}{k}}}
-      {\ctxwk{{A}{Q}}{{A}{k}}}
-    }
-  \label{comp-ww-t}
-\end{align}
-
-\begin{rmk}
-As an important special case of these inference rules we have the following
-valid inference rules:
-\begin{align*}
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{\Gamma}{B}
-    \jfam{\Gamma}{C}
-    }
-  { \jfameq
-      {{{\Gamma}{A}}{\ctxwk{A}{B}}}
-      {\ctxwk{A}{{B}{C}}}
-      {\ctxwk{{A}{B}}{{A}{C}}}
-    }
-  \\
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{\Gamma}{B}
-    \jfam{{\Gamma}{C}}{R}
-    }
-  { \jfameq
-      {{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{{B}{C}}}}
-      {\ctxwk{A}{{B}{R}}}
-      {\ctxwk{{A}{B}}{{A}{R}}}
-    }
-  \\
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{\Gamma}{B}
-    \jterm{{\Gamma}{C}}{R}{h}
-    }
-  { \jtermeq
-      {{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{{B}{C}}}}
-      {\ctxwk{A}{{B}{R}}}
-      {\ctxwk{A}{{B}{h}}}
-      {\ctxwk{{A}{B}}{{A}{h}}}
-    }
-\end{align*}
-Moreover, we have
-\begin{equation*}
-\inference
-  { \jfam{\Gamma}{A}
-    \jfam{\Gamma}{B}
-    \jterm{\Gamma}{C}{z}
-    }
-  { \jtermeq
-      {{{\Gamma}{A}}{\ctxwk{A}{B}}}
-      {\ctxwk{A}{{B}{C}}}
-      {\ctxwk{A}{{B}{z}}}
-      {\ctxwk{{A}{B}}{{A}{z}}}
-    }
-\end{equation*}
-\end{rmk}
-
-\subsubsection{Weakening by the empty family}
-Note that we can also weaken by the empty family over $\Gamma$.
-Weakening by the empty family $\emptyf$ over a context $\Gamma$ leaves families, 
-their terms, families over those families and terms of those unchanged:
-\label{comp-0w}
-\begin{align}
-& \inference
-  { \jfam{\Gamma}{B}
-    }
-  { \jfameq{\Gamma}{\ctxwk{\emptyf}{B}}{B}
-    }
-  \label{comp-0w-c}
-  \\
-& \inference
-  { \jfam{{\Gamma}{B}}{Q}
-    }
-  { \jfameq{{\Gamma}{B}}{\ctxwk{\emptyf}{Q}}{Q}
-    }
-  \label{comp-0w-f}
-  \\
-& \inference
-  { \jterm{{\Gamma}{B}}{Q}{g}
-    }
-  { \jtermeq{{\Gamma}{B}}{Q}{\ctxwk{\emptyf}{g}}{g}
-    }
-  \label{comp-0w-t}
-\end{align}
-
-\subsubsection{Currying for weakening}
-\label{comp-ew}
-The rules expressing that extension is compatible with weakening assert that
-weakening by an extension is the same thing as weakening twice in the
-appropriate way.
-\begin{align}
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{{\Gamma}{A}}{P}
-    \jfam{\Gamma}{B}
-    }
-  { \jfameq
-      {{{\Gamma}{A}}{P}}
-      {\ctxwk{\ctxext{A}{P}}{B}}
-      {\ctxwk{P}{{A}{B}}}
-    }
-  \label{comp-ew-c}\\
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{{\Gamma}{A}}{P}
-    \jfam{{\Gamma}{B}}{Q}
-    }
-  { \jfameq
-      {{{{\Gamma}{A}}{P}}{\ctxwk{P}{{A}{B}}}}
-      {\ctxwk{\ctxext{A}{P}}{Q}}
-      {\ctxwk{P}{{A}{Q}}}
-    }
-  \label{comp-ew-f}\\
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{{\Gamma}{A}}{P}
-    \jterm{{\Gamma}{B}}{Q}{g}
-    }
-  { \jtermeq
-      {{{{\Gamma}{A}}{P}}{\ctxwk{P}{{A}{B}}}}
-      {\ctxwk{P}{{A}{Q}}}
-      {\ctxwk{\ctxext{A}{P}}{g}}
-      {\ctxwk{P}{{A}{g}}}
-    } 
-  \label{comp-ew-t}
-\end{align}
-
-\begin{comment}
-\subsubsection{Weakening by a context}
-Since we have not required that there is an empty context with the property that
-a context is the same thing as a family in the empty context, we will have to
-explicitly introduce weakening by contexts. Since it will allways be clear
-whether something is weakened by a context or by a family, we shall use the
-same notation for weakening by contexts as for weakening by families. We shall
-have to repeat all the compatibility rules for weakening by families for
-weakening by contexts; there is no way around it. It is
-actually this sort of inconvenience that has originally led us to include an
-empty context and be done with it.
-
-In the following inference rules we assume that we have a context $\Gamma$ and
-that $\jctxeq{\Gamma}{\Gamma'}$ in the rules where $\Gamma'$ makes an appearence.
-\begin{align}
-& \inference
-  { \jctx{\Delta}
-    }
-  { \jfam{\Gamma}{\ctxwk{\Gamma}{\Delta}}
-    }
-& & \inference
-  { \jctxeq{\Delta}{\Delta'}
-    }
-  { \jfameq{\Gamma}{\ctxwk{\Gamma}{\Delta}}{\ctxwk{\Gamma'}{\Delta'}}
-    }
-  \\
-& \inference
-  { \jfam{\Delta}{B}
-    }
-  { \jfam{\ctxext{\Gamma}{\ctxwk{\Gamma}{\Delta}}}{\ctxwk{\Gamma}{B}}
-    }
-& & \inference
-  { \jfameq{\Delta}{B}{B'}
-    }
-  { \jfameq
-      {\ctxext{\Gamma}{\ctxwk{\Gamma}{\Delta}}}
-      {\ctxwk{\Gamma}{B}}
-      {\ctxwk{\Gamma'}{B'}}
-    }
-  \\
-& \inference
-  { \jterm{\Delta}{B}{y}
-    }
-  { \jterm
-      {\ctxext{\Gamma}{\ctxwk{\Gamma}{\Delta}}}
-      {\ctxwk{\Gamma}{B}}
-      {\ctxwk{\Gamma}{y}}
-    }
-& & \inference
-  { \jtermeq{\Delta}{B}{y}{y'}
-    }
-  { \jtermeq
-      {\ctxext{\Gamma}{\ctxwk{\Gamma}{\Delta}}}
-      {\ctxwk{\Gamma}{B}}
-      {\ctxwk{\Gamma}{y}}
-      {\ctxwk{\Gamma}{y'}}
-    }
-\end{align}
-Note that we do not employ three different notations for the three different
-actions of weakening by a context because there can be no confusion about which
-of the three actions is meant. If there were an empty context we should have
-distinguished the three as we did for weakening by a family.
-
-We shall have similar compatibility rules for weakening by a context as we did
-for weakening by a family.
-
-{\color{red}To do. It will look exactly the same as for weakening by families}
-\end{comment}
-
-\subsection{The theory of projections}
-\label{identityterms}
-We formulate the theory of projections on top of the theory of weakening
-described in \autoref{weakening}. In the theory of projections we introduce
-the identity terms. The identity term at a family $A$ in context $\Gamma$ is
-a term $\jterm{{\Gamma}{A}}{\ctxwk{A}{A}}{\idtm{A}}$, so weakening is necessary.
-Together with weakening, the identity terms will provide for the projections.
-The projection from $\cftext{A}{P}$ to $A$ will be the term
-$\jterm{{\Gamma}{{A}{P}}}{\ctxwk{\ctxext{A}{P}}{A}}{\ctxwk{P}{\idtm{A}}}$ and
-will be studied extensively in \autoref{extension-on-terms}.
-
-The following inference rules are required to be valid:
-\begin{align}
-& \inference
-  { \jfam{\Gamma}{A}
-    }
-  { \unfoldall{\jhom{\Gamma}{A}{A}{\idtm{A}}}
-    }
-& & \inference
-    { \jfameq{\Gamma}{A}{A'}
-      }
-    { \unfoldall{\jhomeq{\Gamma}{A}{A}{\idtm{A}}{\idtm{A'}}}
-      }
-\end{align}
-
-\begin{comment}
-We stress again
-that because we have not required there to be an empty context, we need to
-explicitly call for identity terms $\idtm{\Gamma}$ of $\ctxwk{\Gamma}{\Gamma}$
-in context $\Gamma$ for each context $\Gamma$. Thus, we will also require
-the inference rule
-\begin{align}
-& \inference
-  { \jctx{\Gamma}
-    }
-  { \jterm{\Gamma}{\ctxwk{\Gamma}{\Gamma}}{\idtm{\Gamma}}
-    }
-& &
-  \inference
-  { \jctxeq{\Gamma}{\Gamma'}
-    }
-  { \jtermeq{\Gamma}{\ctxwk{\Gamma}{\Gamma}}{\idtm{\Gamma}}{\idtm{\Gamma'}}
-    }
-\end{align}
-For the following inference rules, there should also be versions of identity
-terms at contexts.
-\end{comment}
-
-\subsubsection{The identity term of an extended family}
-The fact that identity terms are compatible with extensions is derived in
-\autoref{comp-ie}.
-
-\subsubsection{The identity term of a weakened family}
-\label{comp-wi}
-The identity term of a weakened family is the weakened identity term:
-\begin{equation}
-\inference
-  { \jfam{\Gamma}{A}
-    \jfam{{\Gamma}{B}}{Q}
-    }
-  { \unfoldall{\jhomeq
-      {{{\Gamma}{A}}{\ctxwk{A}{B}}}
-      {\ctxwk{A}{Q}}
-      {\ctxwk{A}{Q}}
-      {\ctxwk{A}{\idtm{Q}}}
-      {\idtm{\ctxwk{A}{Q}}}
-      }
-    }
-  \label{comp-wi-t}
-\end{equation}
-
-\begin{rmk}
-An important special case is the judgmental equality
-\begin{equation*}
-\unfoldall{\jhomeq
-      {{\Gamma}{A}}
-      {\ctxwk{A}{B}}
-      {\ctxwk{A}{B}}
-      {\ctxwk{A}{\idtm{B}}}
-      {\idtm{\ctxwk{A}{B}}}
-      }
-\end{equation*}
-for a family $\jfam{\Gamma}{B}$.
-\end{rmk}
-
-\subsection{The theory of substitution}
-\label{substitution}
-The theory of substitution can be formulated separately from the theory of
-weakening. Therefore, we will only require the theory of extension and the
-empty context and families. In \autoref{sec:esystem-equalities} we will
-provide the additional judgmental equality rules that will be required when
-combining the theories of substitution and of projections.
-
-Given a family $P$ over $A$ and a term $x$ of $A$, substitution gives a way to
-consider the \emph{fiber $\subst{x}{P}$ of $P$ at $x$}. As was the case with
-weakening, the substitution operation comes in three parts: the `action on
-contexts' of substitution is the operation just described; the `action on
-families' of substitution takes a family $Q$ over $\ctxext{{\Gamma}{A}}{P}$
-to a family $\subst[\famsym]{x}{Q}$ over $\ctxext{\Gamma}{\subst{x}{P}}$; the
-`action on terms' of substitution takes a term $g$ of $Q$ over
-$\ctxext{{\Gamma}{A}}{P}$ to a term $\subst[\tmsym]{x}{g}$ of 
-$\subst[\famsym]{x}{Q}$.
-\begin{align}
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jfam{{\Gamma}{A}}{P}
-    }
-  { \jfam{\Gamma}{\subst{x}{P}}
-    }
-& & \inference
-    { \jtermeq{\Gamma}{A}{x}{x'}
-      \jfameq{{\Gamma}{A}}{P}{P'}
-      }
-    { \jfameq{\Gamma}{\subst{x}{P}}{\subst{x'}{P'}}
-      }
-    \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jfam{{{\Gamma}{A}}{P}}{Q}
-    }
-  { \jfam{{\Gamma}{\subst{x}{P}}}{\subst[\famsym]{x}{Q}}
-    }
-& & \inference
-    { \jtermeq{\Gamma}{A}{x}{x'}
-      \jfameq{{{\Gamma}{A}}{P}}{Q}{Q'}
-      }
-    { \jfameq{{\Gamma}{\subst{x}{P}}}{\subst[\famsym]{x}{Q}}{\subst[\famsym]{x'}{Q'}}
-      }
-    \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jterm{{{\Gamma}{A}}{P}}{Q}{g}
-    }
-  { \jterm{{\Gamma}{\subst{x}{P}}}{\subst[\famsym]{x}{Q}}{\subst[\tmsym]{x}{g}}
-    }
-& & \inference
-    { \jtermeq{\Gamma}{A}{x}{x'}
-      \jtermeq{{{\Gamma}{A}}{P}}{Q}{g}{g'}
-      }
-    { \jtermeq
-        {{\Gamma}{\subst{x}{P}}}
-        {\subst[\famsym]{x}{Q}}
-        {\subst[\tmsym]{x}{g}}
-        {\subst[\tmsym]{x'}{g'}}
-      }
-\end{align}
-
-\subsubsection{Fibers of an empty family}
-The following inference rules establish that the fibers of the empty family are 
-the empty families:
-\label{comp-s0}
-\begin{align}
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    }
-  { \jfameq{\Gamma}{\subst{x}{\emptyf}}{\emptyf}
-    }
-  \label{comp-s0-c}
-  \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jfam{{\Gamma}{A}}{P}
-    }
-  { \jfameq
-      {{\Gamma}{\subst{x}{P}}}
-      {\subst{x}{\emptyf}}
-      {\emptyf}
-    }
-  \label{comp-s0-f}
-\end{align}
-\begin{comment}
-Note again that we do not have to axiomatize the action of substitution by $x$ on the term
-of the empty family, because the result will be in the empty family and every
-term in the empty family is judgmentally equal to the term $\emptytm$.
-\end{comment}
-
-We use the above rule to state the compatibility of the action on families of
-substitution with the action on contexts of substitution. Note that a family
-$P$ over $\ctxext{\Gamma}{A}$ may be regarded as a family over
-$\ctxext{{\Gamma}{A}}{\emptyf}$. Thus, we may consider the family
-$\subst[\famsym]{x}{P}$ over $\ctxext{\Gamma}{\subst{x}{\emptyf}}$. Since
-$\subst{x}{\emptyf}$ is judgmentally equal to the empty family, we may compare
-$\subst[\famsym]{x}{P}$ with $\subst{x}{P}$:
-\begin{equation}
-\inference
-{ \jfam{{\Gamma}{A}}{P}
-  }
-{ \jfameq{\Gamma}{\subst[\famsym]{x}{P}}{\subst{x}{P}}
-  }
-\end{equation}
-Due to this rule we need not make the annotations $\famsym$ and $\tmsym$ in
-the notation for substitution anymore and thus we shall omit them from now on.
-Note that because a family $P$ over $\ctxext{\Gamma}{A}$ is eligible for
-application of the action on families of substitution, we may also substitute
-terms of $P$. Thus, given terms $\jterm{\Gamma}{A}{x}$ and $\jterm{{\Gamma}{A}}{P}{f}$,
-we get a term $\jterm{\Gamma}{\subst{x}{P}}{\subst{x}{f}}$, the \emph{value of
-$f$ at $x$}.
-
-In \autoref{morphisms} we will use a combination of weakening and substitution
-to define composition of morphisms of families. However, we have to rely
-on the cancellation rule stated in \autoref{cancellation-ws} before we can
-meaningfully state the definition of composition.
-
-\subsubsection{Fibers of extensions}
-\label{comp-se}
-The following inference rules assert that if we take the fiber of an extended
-family at a term $x$ of $A$ in context $\Gamma$, the substitution by $x$
-distributes over the factors of the extension.
-\begin{align}
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jfam{{{\Gamma}{A}}{P}}{Q}
-    }
-  { \jfameq
-      {\Gamma}
-      {\subst{x}{\ctxext{P}{Q}}}
-      {\ctxext{\subst{x}{P}}{\subst{x}{Q}}}
-    }
-  \label{comp-se-c}
-  \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jfam{{{{\Gamma}{A}}{P}}{Q}}{R}
-    }
-  { \jfameq
-      {{\Gamma}{\subst{x}{P}}}
-      {\subst{x}{\ctxext{Q}{R}}}
-      {\ctxext{\subst{x}{Q}}{\subst{x}{R}}}
-    }
-  \label{comp-se-f}
-\end{align}
-There is also a version of this statement in which extended terms are considered.
-This variant is stated and proved in \autoref{comp-se-t}.
-
-\subsubsection{Fibers of fibers}
-\label{comp-ss}
-
-The following rules explain what happens when we first substitute by a term
-$g$ of $Q$ over $P$ in context $\ctxext{\Gamma}{A}$ and then by a term $x$ of
-$A$:
-\begin{align}
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jterm{{{\Gamma}{A}}{P}}{Q}{g}
-    \jfam{{{{\Gamma}{A}}{P}}{Q}}{R}
-    }
-  { \jfameq
-      {{{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}}
-      {\subst{x}{{g}{R}}}
-      {\subst{{x}{g}}{{x}{R}}}
-    }
-  \label{comp-ss-c}
-  \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jterm{{{\Gamma}{A}}{P}}{Q}{g}
-    \jfam{{{{{\Gamma}{A}}{P}}{Q}}{R}}{S}
-    }
-  { \jfameq
-      {{{{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}}{\subst{x}{{g}{R}}}}
-      {\subst{x}{{g}{S}}}
-      {\subst{{x}{g}}{{x}{S}}}
-    }
-  \label{comp-ss-f}
-  \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jterm{{{\Gamma}{A}}{P}}{Q}{g}
-    \jterm{{{{{\Gamma}{A}}{P}}{Q}}{R}}{S}{k}
-    }
-  { \jtermeq
-      {{{{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}}{\subst{x}{{g}{R}}}}
-      {\subst{x}{{g}{S}}}
-      {\subst{x}{{g}{k}}}
-      {\subst{{x}{g}}{{x}{k}}}
-    }
-  \label{comp-ss-t}
-\end{align}
-
-\begin{rmk}
-As an important special case of these inference rules we have the following
-valid inference rules:
-\begin{align*}
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jterm{{\Gamma}{A}}{P}{f}
-    \jfam{{{\Gamma}{A}}{P}}{Q}
-    }
-  { \jfameq
-      {\Gamma}
-      {\subst{x}{{f}{Q}}}
-      {\subst{{x}{f}}{{x}{Q}}}
-    }
-  \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jterm{{\Gamma}{A}}{P}{f}
-    \jfam{{{{\Gamma}{A}}{P}}{Q}}{R}
-    }
-  { \jfameq
-      {{\Gamma}{\subst{x}{{f}{Q}}}}
-      {\subst{x}{{f}{R}}}
-      {\subst{{x}{f}}{{x}{R}}}
-    }
-  \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jterm{{\Gamma}{A}}{P}{f}
-    \jterm{{{{\Gamma}{A}}{P}}{Q}}{R}{h}
-    }
-  { \jtermeq
-      {{\Gamma}{\subst{x}{{f}{Q}}}}
-      {\subst{x}{{f}{R}}}
-      {\subst{x}{{f}{h}}}
-      {\subst{{x}{f}}{{x}{h}}}
-    }
-\end{align*}
-Moreover, we get
-\begin{equation*}
-\inference
-  { \jterm{\Gamma}{A}{x}
-    \jterm{{\Gamma}{A}}{P}{f}
-    \jterm{{{\Gamma}{A}}{P}}{Q}{g}
-    }
-  { \jtermeq
-      {\Gamma}
-      {\subst{x}{{f}{Q}}}
-      {\subst{x}{{f}{g}}}
-      {\subst{{x}{f}}{{x}{g}}}
-    }
-\end{equation*}
-\end{rmk}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Joining the theory of projections with the theory of substitution}
-\label{sec:esystem-equalities}
-
-\begin{comment}
-We have already seen that empty families, extension, weakening, substitution
-and the identity terms are compatible with each other. There are, however, a
-few more judgmental equalities explaining what happens when
-\begin{enumerate}
-\item something is weakend by the empty family. This is used to consider a term
-      $x$ of $A$ in context $\Gamma$ as a morphism from $\emptyf$ to $A$ in
-      context $\Gamma$;
-\item something is weakened by an extended family. This is Currying for weakening;
-\item something is first weakened by a family $A$ in context $\Gamma$ and then
-      substituted by a term of $A$. This is used to show that composition is
-      compatible with itself (i.e. that composition is associative);
-\item the identity terms are joining the party. Identity terms are empty contexts
-      in certain models, so they should behave accordingly.
-\end{enumerate}
-In other words, the rules seem to not fit in the scheme that all operations are
-compatible with each other, but in fact each of these rules plays a role in
-completing the scheme (work out further).
-\end{comment}
-
-\begin{comment}
-In this section we extend the judgmental equality relation in such a way that
-the empty context, extension, weakening, substitution and the identity terms
-all become compatible with each other. Since we have five operations, there
-are going to be $25$ finite sets of rules explaining how each operation is
-compatible with each of the operations.
-
-In addition to these $25$ finite sets
-of compatibility rules, there are also two sets of rules explaining the
-cancellation properties of weakening and substitution and of the identity terms.
-
-A typical compatibility axiom captures the idea that a diagram like
-\begin{equation*}
-\begin{tikzcd}[column sep=huge]
-\jfam{\Gamma}{\blank} 
-  \ar{r}{C\mapsto\ctxwk{B}{C}} 
-  \ar{d}[swap]{C\mapsto\ctxwk{A}{C}} 
-& \jfam{{\Gamma}{B}}{\blank} 
-  \ar{d}{Q\mapsto\ctxwk{A}{Q}}
-  \\
-\jfam{{\Gamma}{A}}{\blank} 
-  \ar{r}[swap]{P\mapsto\ctxwk{{A}{B}}{P}} 
-& \jfam{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\blank}
-\end{tikzcd}
-\end{equation*}
-in the case of two consecutive weakenings, or
-\begin{equation*}
-\begin{tikzcd}[column sep=huge]
-\jfam{{\Gamma}{A}}{\blank} 
-  \ar{d}[swap]{Q\mapsto\subst{x}{Q}} 
-  \ar{r}{Q\mapsto\ctxwk{P}{Q}} 
-& \jfam{{{\Gamma}{A}}{P}}{\blank} 
-    \ar{d}{R\mapsto\subst{x}{R}}
-  \\ 
-\jfam{\Gamma}{\blank} 
-  \ar{r}[swap]{B\mapsto\ctxwk{\subst{x}{P}}{B}} 
-& \jfam{{\Gamma}{\subst{x}{P}}}{\blank}
-\end{tikzcd}
-\end{equation*}
-in the case of a combination of weakening and substitution, commutes. We won't
-actually draw any of these diagrams in this section, but readers are encouraged
-to do so if they think it can help them understanding the axioms. Another thing
-one could keep in mind with the axioms that involve weakening is that all of
-them become completely obvious when weakening is made notationally invisible.
-Of course, we have an explicit notation for weakening on purpuse, but it is the
-usual practice of type theorists to make weakening invisible.
-
-The table below indicates where
-we can find the axioms of each of the combinations of operators. The reference
-at the row of operation $X$ and the column of operation $Y$ refers to the
-axiom which explains $Y$ distributes over the arguments of $X$ when $Y$ is
-applied after $X$. The empty context has no arguments, whereas extension,
-and weakening have two arguments which are both of the family-kind, substitution
-has two arguments of which one is a term and the other is a family and
-idenitity terms have one argument, a family.
-\begin{center}
-\begin{tabular}{r|rrrr}
-& ext & wk & subst & id\\
-\hline
-extension & \autoref{comp-ee}* &  & \cmark & \cmark \\
-weakening & \autoref{comp-we}* & \autoref{comp-ww} & \autoref{comp-sw} & \autoref{comp-wi} \\
-substitution & \autoref{comp-se}* & \autoref{comp-ws} & \autoref{comp-ss} & \autoref{comp-si} \\
-identity terms & \cmark & \autoref{comp-wi} & \autoref{comp-si} & \xmark
-\end{tabular}
-\end{center}
-Apart from references to sections where axioms are given, we also see checkmarks,
-crosses, marked references and duplicated references in this table. 
-The checkmarks indicate that the desired property
-can be proved and need not be axiomatized. To clarify those:
-\begin{itemize}
-\item The identity term of the empty context is the unique term of the empty
-family. Recall that we already have axiomatized in \autoref{empty} that every
-term of the empty family is judgmentally equal to $\emptytm$.
-\item The identity term of an extension can be explained by means of the notion
-of \emph{extension on terms}. This notion is defined in \autoref{extension-on-terms}
-using only identity terms and substitution. With extension on terms we can
-combine a term $\jterm{\Gamma}{A}{x}$ and a term $\jterm{\Gamma}{\subst{x}{P}}{u}$
-into a term $\jterm{\Gamma}{{A}{P}}{\tmext{x}{u}}$, which may be seen as the
-pair consisting of $x$ and $u$. Since it is a defined notion, we will
-be able to derive lots of properties for extension of terms. One of them explains
-how we can see the identity term of an extension as the extension of two certain
-terms, see \autoref{comp-ie}.
-\item Using extension on terms we can also make sense of `fibers at an extension'
-because that would exactly be the fiber at an extended term. We will see in
-\autoref{comp-es} that, for instance, $\subst{\tmext{x}{u}}{Q}\jdeq\subst{u}{{x}{Q}}$
-for any family $\jfam{{{\Gamma}{A}}{P}}{Q}$ and any two terms $\jterm{\Gamma}{A}{x}$
-and $\jterm{\Gamma}{\subst{x}{P}}{u}$. That's a form of currying.
-\end{itemize}
-The spots with a \xmark simply make no sense 
-{\color{red}(or do they, when the term model
-is better understood (by me)?)}. The marked references point to
-axioms which \emph{do} have a variant for terms, but those require the notion
-of extension on terms, can be proved and therefore need not be included in our
-list of compatibility axioms. In each of these sections, we give a forward
-reference to the lemma in which this variant is stated and proved.
-\end{comment}
-
-\subsubsection{Fibers of weakenings}\label{comp-sw}
-The following rules assert what happens when we first weaken by a family
-$Q$ over $P$ in context $\ctxext{\Gamma}{A}$ and then substitute by a term
-$x$ of $A$:
-\begin{align}
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jfam{{{\Gamma}{A}}{P}}{Q}
-    \jfam{{{\Gamma}{A}}{P}}{R}
-    }
-  { \jfameq
-      {{{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}}
-      {\subst{x}{\ctxwk{Q}{R}}}
-      {\ctxwk{\subst{x}{Q}}{\subst{x}{R}}}
-    }
-  \label{comp-sw-c}
-  \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jfam{{{\Gamma}{A}}{P}}{Q}
-    \jfam{{{{\Gamma}{A}}{P}}{R}}{S}
-    }
-  { \jfameq
-      {{{{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}}{\subst{x}{\ctxwk{Q}{R}}}}
-      {\subst{x}{\ctxwk{Q}{S}}}
-      {\ctxwk{\subst{x}{Q}}{\subst{x}{S}}}
-    }
-  \label{comp-sw-f}
-  \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jfam{{{\Gamma}{A}}{P}}{Q}
-    \jterm{{{{\Gamma}{A}}{P}}{R}}{S}{k}
-    }
-  { \jtermeq
-      {{{{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}}{\subst{x}{\ctxwk{Q}{R}}}}
-      {\subst{x}{\ctxwk{Q}{S}}}
-      {\subst{x}{\ctxwk{Q}{k}}}
-      {\ctxwk{\subst{x}{Q}}{\subst{x}{k}}}
-    }
-  \label{comp-sw-t}
-\end{align}
-
-\begin{rmk}
-As an important special case of these inference rules we have the following
-valid inference rules:
-\begin{align*}
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jfam{{\Gamma}{A}}{P}
-    \jfam{{\Gamma}{A}}{Q}
-    }
-  { \jfameq
-      {{\Gamma}{\subst{x}{P}}}
-      {\subst{x}{\ctxwk{P}{Q}}}
-      {\ctxwk{\subst{x}{P}}{\subst{x}{Q}}}
-    }
-  \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jfam{{\Gamma}{A}}{P}
-    \jfam{{{\Gamma}{A}}{Q}}{R}
-    }
-  { \jfameq
-      {{{\Gamma}{\subst{x}{P}}}{\subst{x}{\ctxwk{P}{Q}}}}
-      {\subst{x}{\ctxwk{P}{R}}}
-      {\ctxwk{\subst{x}{P}}{\subst{x}{R}}}
-    }
-  \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jfam{{\Gamma}{A}}{P}
-    \jterm{{{\Gamma}{A}}{Q}}{R}{h}
-    }
-  { \jtermeq
-      {{{\Gamma}{\subst{x}{P}}}{\subst{x}{\ctxwk{P}{Q}}}}
-      {\subst{x}{\ctxwk{P}{R}}}
-      {\subst{x}{\ctxwk{P}{h}}}
-      {\ctxwk{\subst{x}{P}}{\subst{x}{h}}}
-    }
-\end{align*}
-Moreover, we get
-\begin{equation*}
-\inference
-  { \jterm{\Gamma}{A}{x}
-    \jfam{{\Gamma}{A}}{P}
-    \jterm{{\Gamma}{A}}{Q}{g}
-    }
-  { \jtermeq
-      {{\Gamma}{\subst{x}{P}}}
-      {\subst{x}{\ctxwk{P}{Q}}}
-      {\subst{x}{\ctxwk{P}{g}}}
-      {\ctxwk{\subst{x}{P}}{\subst{x}{g}}}
-    }
-\end{equation*}
-\end{rmk}
-
-\subsubsection{Weakenings of fibers}
-\label{comp-ws}
-The following inference rules explain what happens when we first weaken by a
-term $g$ of $Q$ in context $\ctxext{\Gamma}{B}$ and then weaken by a family
-$A$ over $\Gamma$.
-\begin{align}
-& \inference
-  { \jfam{\Gamma}{A}
-    \jterm{{\Gamma}{B}}{Q}{g}
-    \jfam{{{\Gamma}{B}}{Q}}{R}
-    }
-  { \jfameq
-      {{{\Gamma}{A}}{\ctxwk{A}{B}}}
-      {\ctxwk{A}{\subst{g}{R}}}
-      {\subst{\ctxwk{A}{g}}{\ctxwk{A}{R}}}
-    }
-  \label{comp-ws-c}
-  \\
-& \inference
-  { \jfam{\Gamma}{A}
-    \jterm{{\Gamma}{B}}{Q}{g}
-    \jfam{{{{\Gamma}{B}}{Q}}{R}}{S}
-    }
-  { \jfameq
-      {{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{\subst{g}{R}}}}
-      {\ctxwk{A}{\subst{g}{S}}}
-      {\subst{\ctxwk{A}{g}}{\ctxwk{A}{S}}}
-    }
-  \label{comp-ws-f}
-  \\
-& \inference
-  { \jfam{\Gamma}{A}
-    \jterm{{\Gamma}{B}}{Q}{g}
-    \jterm{{{{\Gamma}{B}}{Q}}{R}}{S}{k}
-    }
-  { \jtermeq
-      {{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{\subst{g}{R}}}}
-      {\ctxwk{A}{\subst{g}{S}}}
-      {\ctxwk{A}{\subst{g}{k}}}
-      {\subst{\ctxwk{A}{g}}{\ctxwk{A}{k}}}
-    }
-  \label{comp-ws-t}
-\end{align}
-
-\begin{rmk}
-As an important special case of these inference rules we have the following 
-valid inference rules:
-\begin{align*}
-& \inference
-  { \jfam{\Gamma}{A}
-    \jterm{\Gamma}{B}{y}
-    \jfam{{\Gamma}{B}}{Q}
-    }
-  { \jfameq
-      {{\Gamma}{A}}
-      {\ctxwk{A}{\subst{y}{Q}}}
-      {\subst{\ctxwk{A}{y}}{\ctxwk{A}{Q}}}
-    }
-  \\
-& \inference
-  { \jfam{\Gamma}{A}
-    \jterm{\Gamma}{B}{y}
-    \jfam{{{\Gamma}{B}}{Q}}{R}
-    }
-  { \jfameq
-      {{{\Gamma}{A}}{\ctxwk{A}{\subst{y}{Q}}}}
-      {\ctxwk{A}{\subst{y}{R}}}
-      {\subst{\ctxwk{A}{y}}{\ctxwk{A}{R}}}
-    }
-  \\
-& \inference
-  { \jfam{\Gamma}{A}
-    \jterm{\Gamma}{B}{y}
-    \jterm{{{\Gamma}{B}}{Q}}{R}{h}
-    }
-  { \jtermeq
-      {{{\Gamma}{A}}{\ctxwk{A}{\subst{y}{Q}}}}
-      {\ctxwk{A}{\subst{y}{R}}}
-      {\ctxwk{A}{\subst{y}{h}}}
-      {\subst{\ctxwk{A}{y}}{\ctxwk{A}{h}}}
-    }
-\end{align*}
-Moreover, we get
-\begin{equation*}
-\inference
-  { \jfam{\Gamma}{A}
-    \jterm{\Gamma}{B}{y}
-    \jterm{{\Gamma}{B}}{Q}{g}
-    }
-  { \jtermeq
-      {{\Gamma}{A}}
-      {\ctxwk{A}{\subst{y}{Q}}}
-      {\ctxwk{A}{\subst{y}{g}}}
-      {\subst{\ctxwk{A}{y}}{\ctxwk{A}{g}}}
-    }
-\end{equation*}
-\end{rmk}
-
-\subsubsection{The cancellation property of weakening and substitution}
-\label{cancellation-ws}
-The judgmental equalities we're about to describe assert that substituting a term
-in the weakening a thing gives you the thing back. In the case of contexts we get that each fiber
-$\subst{x}{\ctxwk{A}{B}}$ is just $B$ and in the case of terms we get 
-that $\ctxwk{A}{y}$ is the constant function
-mapping everything to $y:B$. Thus, these rules actually establish the weakening
-as the weakening. After stating the rules we will describe what it means to
-compose context morphisms (terms of weakened contexts).
-\begin{align}
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{\Gamma}{B}
-    \jterm{\Gamma}{A}{x}
-    }
-  { \jfameq
-      {\Gamma}
-      {\subst{x}{\ctxwk{A}{B}}}
-      {B}
-    }
-  \label{cancellation-ws-c}\\
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{{\Gamma}{B}}{Q}
-    \jterm{\Gamma}{A}{x}
-    }
-  { \jfameq{{\Gamma}{B}}{\subst{x}{\ctxwk{A}{Q}}}{Q}
-    }
-  \label{cancellation-ws-f}\\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jterm{{\Gamma}{B}}{Q}{g}
-    }
-  { \jtermeq{{\Gamma}{B}}{Q}{\subst{x}{\ctxwk{A}{g}}}{g}
-    }
-  \label{cancellation-ws-t}
-\end{align}
-
-\subsubsection{The identity term of a substituted family}
-\label{comp-si}
-The identity term of a substituted family is the substitution of the identity term
-\begin{equation}
-\inference
-  { \jterm{\Gamma}{A}{x}
-    \jfam{{{\Gamma}{A}}{P}}{Q}
-    }
-  { \unfoldall{\jhomeq
-      {{\Gamma}{\subst{x}{P}}}
-      {\subst{x}{Q}}
-      {\subst{x}{Q}}
-      {\subst{x}{\idtm{Q}}}
-      {\idtm{\subst{x}{Q}}}
-    }}
-  \label{comp-si-t}
-\end{equation}
-
-\begin{rmk}
-An important special case is the judgmental equality
-\begin{equation*}
-\unfoldall{\jhomeq
-      {\Gamma}
-      {\subst{x}{P}}
-      {\subst{x}{P}}
-      {\subst{x}{\idtm{P}}}
-      {\idtm{\subst{x}{P}}}
-    }
-\end{equation*}
-for a family $\jfam{{\Gamma}{A}}{P}$.
-\end{rmk}
-
-\subsubsection{The cancellation properties of identity terms}
-\label{cancellation-i}
-Identity terms are determined by their behavior with respect to substitution combined with
-weakening. The identity terms will also be subject to compatibility rules.
-\begin{align}
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    }
-  { \jtermeq{\Gamma}{A}{\subst{x}{\idtm{A}}}{x}
-    }
-  \label{cancellation-si}\\
-& \inference
-  { \jfam{{\Gamma}{A}}{P}
-    }
-  { \jfameq{{\Gamma}{A}}{\subst{\idtm{A}}{\ctxwk{A}{P}}}{P}
-    }
-  \label{precomp-idtm-c}
-  \\
-& \inference
-  { \jfam{{{\Gamma}{A}}{P}}{Q}
-    }
-  { \jfameq{{{\Gamma}{A}}{P}}{\subst{\idtm{A}}{\ctxwk{A}{Q}}}{Q}
-    }
-  \label{precomp-idtm-f}
-  \\
-& \inference
-  { \jterm{{{\Gamma}{A}}{P}}{Q}{g}
-    }
-  { \jtermeq
-      {{{\Gamma}{A}}{P}}
-      {Q}
-      {\subst{\idtm{A}}{\ctxwk{A}{g}}}
-      {g}
-    }
-  \label{precomp-idtm-t}
-\end{align}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{The possiblity of types in the theory of contexts, families and
-terms}
-\label{types}
-We have deliberately not spoken of types so far because we have taken the point
-of view that a type in a context is nothing but a family in that context which
-belongs to the class of types. We think of types as \emph{irreducible} families,
-i.e.\ families which are neither empty nor the extension of two
-families which are both not empty (in algebraic terminology: which
-are both non-trivial). To allow ourselves to speak of types we introduce two
-new judgments: the judgment that something is a type and the judgment that two
-types are equal.
-\begin{align*}
-\jalign\jtype{\Gamma}{A} 
-& \jalign\jtypeeq{\Gamma}{A}{B}
-\end{align*}
-But only families of contexts are eligible to be types. If $A$ is a type
-in context $\Gamma$, then $A$ is also a family of contexts over $\Gamma$. 
-Moreover, two types in context $\Gamma$ are judgmentally equal precisely when they are equal
-as context families and if a family $B$ of contexts over $\Gamma$ is
-judgmentally equal to a type $A$ in context $\Gamma$, then $B$ is a type in
-context $\Gamma$. This is expressed by the following four inference rules:
-\begin{align*}
-& \inference
-  { \jtype{\Gamma}{A}
-    }
-  { \jfam{\Gamma}{A}
-    }
-& & \inference
-    { \jtypeeq{\Gamma}{A}{B}
-      }
-    { \jfameq{\Gamma}{A}{B}
-      }
-    \\
-& \inference
-  { \jtype{\Gamma}{A}
-    \jfameq{\Gamma}{A}{B}
-    }
-  { \jtype{\Gamma}{B}
-    }
-& & \inference
-    { \jtype{\Gamma}{A}
-      \jfameq{\Gamma}{A}{B}
-      }
-    { \jtypeeq{\Gamma}{A}{B}
-      }
-\end{align*}
-As pointed out at the beginning of this subsection, 
-we do not assume that the empty family is a type, that would be like
-assuming that the multiplicative unit of a ring is prime.
-
-\begin{comment}
-We envision a
-categorical interpretation of the theory of contexts, families and terms where 
-contexts are interpreted as finite paths of arrows belonging to a predetermined
-class. The empty family would be such a path of length $0$ whereas the types
-would be such paths of length $1$.
-\end{comment}
-
-We add rules asserting that a weakened type is again a type that 
-substitution preserves the property of being a type:
-\begin{align}
-& \inference
-  { \jfam{\Gamma}{B}
-    \jtype{\ctxext{\Gamma}{B}}{Q}
-    }
-  { \jtype{\ctxext{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{Q}}
-    }
-  \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jtype{\ctxext{{\Gamma}{A}}{P}}{Q}
-    }
-  { \jtype{\ctxext{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}
-    }
-\end{align}
-With only the current rules, the possibility of making the judgment that
-something is a type does not add much to the theory of contexts, families and
-terms. Nevertheless, when studying models, having an interpretation for the
-judgment that something is a type will allow for the possibility to study
-conditions such as the one asserting that every family `factorizes' uniquely
-as multiple applications of extension to types, analogous to the condition on
-unique factorization domains in ring theory. All the tradional models of type
-theory should translate to models with such a condition, simply because
-traditionally contexts are viewed as lists of variable (and type) declarations.

### 2014-10-28 Rules/rules_join.tex
reorganization

--- 2014-09-20 article-models-tt.tex
@@
-\section{The generalized algebraic theory of contexts, families and terms (the E-system)}\label{tt}
-In this section we give a description of dependent type theory before type
-constructors. Apart from contexts, families and terms -- which provide for the
-core of the language of dependent logic -- the basic ingredients
-of this theory will be the operations of extension on context and on families, 
-and empty context and empty families, weakening, substitution
-and the identity terms. The resulting theory can be seen as a manifestation of the 
-structure underlying dependent type theory.
-
-We will
-formulate the theory of contexts, families and terms in such a way that contexts aren't defined
-to be lists of variable declarations. The variable-free (a.k.a.~\emph{name-free}) approach 
-we take here is rather different than those appearing in 
-\cite{hofmann1995extensional,TheBook} but it has appeared in the work of Coquand
-and in \cite{Dybjer1996}.
-The main reason we don't let variable declarations in is that we don't see them 
-in the internal models either. This way we also set out to a more algebraic 
-approach of type theory and higher category theory. Thirdly, we will not have to
-be burdened with superficial comments about variables being bounded or not, or 
-fresh or free or not occuring at all.
-
-In the current presentation, contexts can be seen as binary
-planar trees of which the leaves are (families of) contexts. 
-The judgmental equality relation on contexts is an equivalence relation which 
-expresses that binary planar
-trees of contexts are judgmentally equal if their leaves are, taking only
-(the isomorphism class of) 
-the order of the leaves into account \emph{and not the actual shape of the three}.
-The intuition behind this equivalence relation is indeed that unstructured
-(i.e.~unbracketed) lists such as the lists of variable declarations which
-usually appear in type theoretical syntax, may be regarded as contexts.
-
-Besides contexts, families and terms there will also be a notion of `type in
-a context', which one may assume.
-A context is eligible to be a \emph{closed type} and over a context, a family
-may be a type. We have the intuition that `being a type'
-expresses the property of `being atomic or irreducible'. With a notion of types,
-one has the means to require a stratification by which every context is a finite
-extension of types starting at a closed type, and every family is likewise a
-finite extension of types. The idea behind this is to recover the theory of
-\cite{Garner2014} {\color{red}(and possibly B-systems?)}. We stress that
-we will not include axioms asserting such a stratification in the E-system
-and we leave discussion on the possibility of a typing judgment until the
-end of this section.
-
-With families of contexts being the principal things of
-study, we have added a bit more generality to the theory, compared to the
-theory of categories with families presented in \cite{Dybjer1996}. Here are
-a couple of reasons for this extra generality:
-\begin{enumerate}
-\item We want to be very sure that we can interpret all the rules of type theory
-in a straightforward manner in any context.
-\item We do want do end up with a categorical structure on the contexts, derived
-from the theory by requiring that a morphism from $\Gamma$ to $\Delta$ is a
-term of the weakening of $\Delta$ by $\Gamma$ in context $\Gamma$. Thus, we
-must be able to lift the context $\Delta$ to become a family of contexts over
-$\Gamma$ and it is not reasonable to expect that this results in a type.
-\item It is very easy in this setting to assert that an operation has an action
-on very high levels of families, i.e.~when $F$ is an operation taking things in
-context $\Gamma$ to things in context $\Delta$, then it takes a family $A$ in
-context $\Gamma$ to $F(A)$ in context $\Delta$; a family $P$ in context
-$\ctxext{\Gamma}{A}$ to a family $F(P)$ in context $\ctxext{\Delta}{(F(A))}$,
-and so on. The reason it becomes simple is that we don't have to explain what
-`and so on' means.
-\end{enumerate}
-
-We will formulate fairly strict rules governing the judgmental equalities,
-expressing that extension, weakening and substitution are combatible with
-each other in a judgmental manner. This does not, however, diminish the role
-of isomorphisms or of homotopies could play in the theory once identity types
-are added. Indeed, types could still have non-trivial identity relations and
-the category of types in a certain context could genuinely display higher
-categorical structure, or so we conjecture.
-
-Much of the rules we state are just compatibility rules of extension, weakening
-and substitution with each other. In a way, these rules assert that our contexts
-are just structureless lists of contexts and that likewise terms are structureless
-lists of terms. They are structureless in the sense that the order in which
-they are formed by pairing up is irrelevant. We note that this causes complications
-in the traditional way that categorical sematics of type theory is implemented,
-where contexts become objects of the category which is supposed to model type
-theory. The reason for this is that context extension will not satisfy all the
-compatibility rules we're about to state. The first step to resolving this is taking
-the types in the empty context as the objects.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Judgments and valid inference rules}\label{judgments}
-The theory we describe here is a theory of contexts, families of
-contexts and terms thereof. The families of contexts are by some authors called
-dependent contexts, but they are handled a bit differently here because they
-become the primary object of study. Dependent contexts can be types; they could
-be seen as atomic or indecomposable dependent contexts.
-
-Thus we make eight kinds of judgments: ``$\Gamma$ is a context'',
-``$A$ is a family of contexts over $\Gamma$'', ``$A$ is a type in context $\Gamma$''
-and ``$x$ is a term of the family $A$ of contexts over $\Gamma$''. The other four
-judgments are for judgmental equality.
-\begin{align*}
-\jalign\jctx{\Gamma} 
-& \jalign\jctxeq{\Gamma}{\Gamma'}
-  \\
-\jalign\jfam{\Gamma}{A} 
-& \jalign\jfameq{\Gamma}{A}{B}
-  \\
-\jalign\jterm{\Gamma}{A}{x} 
-& \jalign\jtermeq{\Gamma}{A}{x}{y}.
-\end{align*}
-
-Strictly speaking, we have three different judgmental equalities in play and one
-could request for a notational difference to signify that fact. For instance,
-we could denote the judgmental equalities of contexts, families and terms by
-$\jdeq_c$, $\jdeq_f$ and $\jdeq_t$ respectively. It will, however, always be
-clear which of the three kinds of judgmental equality is meant when we assert
-a judgmental equality and therefore we shall not bother to make this notational
-distinction.
-
-We note that what we call families over contexts
-here could also have been named dependent contexts or telescopes, see
-\cite{deBruijn1991,hofmann1995extensional}. The term family is in agreement
-with the terminology scheme of \cite{TheBook}, though the reader should be
-warned that the notion of familie means something slightly different there than
-it does here.
-
-The theory of contexts, families and terms shall be described by means of a
-declaration of valid inference rules. An inference rule is an expression of the
-form
-\begin{equation*}
-\inference{\mathcal{H}_1\quad\cdots\quad\mathcal{H}_n}{\mathcal{J}}
-\end{equation*}
-where each of $\mathcal{H}_1$, ..., $\mathcal{H}_n$ and $\mathcal{J}$ is 
-judgment. The judgments $\mathcal{H}_1$, ..., $\mathcal{H}_n$ are called the
-hypotheses and the judgment $\mathcal{J}$ is the conclusion of the inference
-rule. When a judgment is the conclusion of a valid inference rule, we also say
-that the judgment is valid or well-formed. In fact, the only way to construct
-new judgments is via the valid inference rules we describe in this section.
-
-The class of valid inference rules is closed under composition and under
-rearranging the hypotheses. That
-means that if
-\begin{equation*}
-\inference{\mathcal{I}_{1k}\quad\cdots\quad\mathcal{I}_{m_kk}}{\mathcal{H}_k}
-\end{equation*}
-and
-\begin{equation*}
-\inference{\mathcal{H}_1\quad\cdots\quad\mathcal{H}_n}{\mathcal{J}}
-\end{equation*}
-are valid inference rules for $1\leq k\leq n$, then
-\begin{equation*}
-\inference{\mathcal{I}_{11}\quad\cdots\quad\mathcal{I}_{m_11}\quad\cdots\quad\mathcal{I}_{1n}\quad\cdots\quad\mathcal{I}_{m_nn}}{\mathcal{J}}
-\end{equation*}
-is also a valid inference rule. Also, in a valid inference rule 
-duplicated hypotheses may always be contracted and hypotheses may always be 
-reordered; the result is still considered a valid inference rule.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{The basic rules for judgmental equality}
-The rules for judgmental equality establish that it is an equivalence relation
-in all three cases (contexts, families and terms). Thus, the following inference
-rules shall be required to be valid:
-\bgroup\small
-\begin{align*}
-& \inference
-  { \jctx{\Gamma}
-    }
-  { \jctxeq{\Gamma}{\Gamma}
-    } 
-& & \inference
-    { \jctxeq{\Gamma}{\Delta}
-      }
-    { \jctxeq{\Delta}{\Gamma}
-      } 
-& & \inference
-    { \jctxeq{\Gamma}{\Delta}
-      \jctxeq{\Delta}{\greek{E}}
-      }
-    { \jctxeq{\Gamma}{\greek{E}}
-      }
-    \\
-& \inference
-  { \jfam{\Gamma}{A}
-    }
-  { \jfameq{\Gamma}{A}{A}
-    } 
-& & \inference
-    { \jfameq{\Gamma}{A}{B}
-      }
-    { \jfameq{\Gamma}{B}{A}
-      }
-& & \inference
-    { \jfameq{\Gamma}{A}{B}
-      \jfameq{\Gamma}{B}{C}
-      }
-    { \jfameq{\Gamma}{A}{C}
-      }
-    \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    }
-  { \jtermeq{\Gamma}{A}{x}{x}
-    }
-& & \inference
-    { \jtermeq{\Gamma}{A}{x}{y}
-      }
-    { \jtermeq{\Gamma}{A}{y}{x}
-      }
-& & \inference
-    { \jtermeq{\Gamma}{A}{x}{y}
-      \jtermeq{\Gamma}{A}{y}{z}
-      }
-    { \jtermeq{\Gamma}{A}{x}{z}
-      }
-\end{align*}
-\egroup
-
-The following convertibility rules are responsible for the strictness
-of judgmental equality, which sets it apart from equivalences or identifications:
-\begin{align*}
-& \inference
-  { \jctxeq{\Gamma}{\Delta}
-    \jfam{\Gamma}{A}
-    }
-  { \jfam{\Delta}{A}
-    }
-& & \inference
-    { \jctxeq{\Gamma}{\Delta}
-      \jfameq{\Gamma}{A}{B}
-      }
-    { \jfameq{\Delta}{A}{B}
-      }
-    \\
-& \inference
-  { \jctxeq{\Gamma}{\Delta}
-    \jterm{\Gamma}{A}{x}
-    }
-  { \jterm{\Delta}{A}{x}
-    }
-& & \inference
-    { \jctxeq{\Gamma}{\Delta}
-      \jtermeq{\Gamma}{A}{x}{y}
-      }
-    { \jtermeq{\Delta}{A}{x}{y}
-      }
-    \\
-& \inference
-  { \jfameq{\Gamma}{A}{B}
-    \jterm{\Gamma}{A}{x}
-    }
-  { \jterm{\Gamma}{B}{x}
-    }
-& & \inference
-    { \jfameq{\Gamma}{A}{B}
-      \jtermeq{\Gamma}{A}{x}{y}
-      }
-    { \jtermeq{\Gamma}{B}{x}{y}
-      }
-\end{align*}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{The operations of the theory of contexts, families and terms}
-In this subsection we introduce the notions of
-extension on context and on families, the empty context and the empty families, 
-weakening, 
-substitution and identity terms.
-
-Weakening and substitution are both
-operations of the theory of contexts, families and terms and therefore they will
-have three levels of actions: on contexts (at the appropriate level), families
-and terms.
-
-\begin{comment}
-Generally, 
-each operation should
-have an action on contexts, on families and on terms at the `appropriate level'
-and should be compatible with each of the operations we introduce here. In
-that will become more practical after the theory of contexts, families and terms
-is formulated, this means that when we will write down a type theory that
-interprets itself in \autoref{ttmodels}, each of the notions of
-extension, empty family, weakening, substitution and identity terms are going to be internalized
-as terms of the theory and all terms are going to be compatible with these.
-That vague idea will be our guiding light.
-
-So in principle, the introductions of each of the empty context, extension,
-weakening, substitution and identity terms should come in three parts. In 
-practice, they do not all come in three parts.
-
-Apart from the empty context, there will indeed be an empty family and a term
-of the empty family (which will be deemed unique). But extension will
-not have an explicit action `at the level of families' since extension
-is used to define higher families and the action would have to be trivial
-anyway. The empty context and extension provide the essentials of the rest of
-the theory in the sense that higher families are families over extensions,
-contexts are families over the empty context and the `action on families' of
-any operation applied to families over the empty context is the same as the
-original operation. Thus, when we have introduced the operations of weakening
-and substitution we will see that they also have an `action on the terms of 
-contexts'. We will state the relevant judgmental equality rules for the empty
-context in this section rather than in the next, since they are required to
-state that the action on families of an operation is compatible with the
-original operation.
-\end{comment}
-
-We note upfront that both context extension and family extension can also be
-seen as operations of contexts, families and terms, even though we will not
-introduce them as such. The reason is that extension is used to define what it
-means to be a family over a family over a context (namely, a family over an
-extended context). As such, it will have a trivial action on families and terms
-which will not need to be introduced. We will come back to this point in
-\autoref{extension}.
-
-\subsection{Extension}
-\label{extension}
-The operations of extension enable one to consider families over
-families and higher level families alike. We will need both a context extension
-operation and a family extension operation.
-
-The inference rules introducing context and family extension are as follows:
-\begin{align}
-& \inference
-  { \jfam{\Gamma}{A}
-    }
-  { \jctx{\ctxext{\Gamma}{A}}
-    }
-& & \inference
-    { \jctxeq{\Gamma}{\Delta}
-      \jfameq{\Gamma}{A}{B}
-      }
-    { \jctxeq{\ctxext{\Gamma}{A}}{\ctxext{\Delta}{B}}
-      }
-    \\
-& \inference
-  { \jfam{{\Gamma}{A}}{P}
-    }
-  { \jfam{\Gamma}{\ctxext{A}{P}}
-    }
-& & \inference
-    { \jfameq{\Gamma}{A}{B} 
-      \jfameq{{\Gamma}{A}}{P}{Q}
-      }
-    { \jfameq{\Gamma}{\ctxext{A}{P}}{\ctxext{B}{Q}}
-      }
-\end{align}
-The extension operation may also be defined to act on families over families
-and terms thereof. For instance, When $\jfam{{\Gamma}{A}}{P}$ is a family,
-we would get a family $\jfam{{\Gamma}{A}}{{\Gamma}{P}}$ and when
-$\jterm{{\Gamma}{A}}{P}{f}$ is a term, we would get a term
-$\jterm{{\Gamma}{A}}{{\Gamma}{P}}{\ctxext{\Gamma}{f}}$. Then there would be
-axioms stating the judgmental equalities $\ctxext{\Gamma}{P}\jdeq P$ and
-$\ctxext{\Gamma}{f}\jdeq f$. Thus, it is of no actual use to introduce the
-action of context extension on families and terms.
-
-In \autoref{extension-on-terms} we will also define an extension operation
-on terms, which defines a term $\jterm{\Gamma}{{A}{P}}{\tmext{x}{u}}$ for
-every $\jterm{\Gamma}{A}{x}$ and $\jterm{\Gamma}{\subst{x}{P}}{u}$. This
-operation is defined by using the substitution operation and identity terms.
-
-\subsubsection{Associativity of extension}
-\label{comp-ee}
-The inference rules asserting that extension is compatible with itself assert
-that contexts are unstructured lists of type declarations. This rule is
-unavoidable if we want that for a family $A$ in context $\Gamma$, a family over
-$A$ is the same thing as a family over $\ctxext{\Gamma}{A}$. 
-\begin{align}
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{{\Gamma}{A}}{P}
-    }
-  { \jctxeq{\ctxext{{\Gamma}{A}}{P}}{\ctxext{\Gamma}{{A}{P}}}
-    }
-  \\
-& \inference
-  { \jfam{{\Gamma}{A}}{P}
-    \jfam{{{\Gamma}{A}}{P}}{Q}
-    }
-  { \jfameq{\Gamma}{\ctxext{A}{{P}{Q}}}{\ctxext{{A}{P}}{Q}}
-    }
-  \label{comp-ee-c}
-\end{align}
-Note that we may take the point of view that extension distributes over
-extensions. If we want to extend an extended family $\ctxext{P}{Q}$ over
-$\ctxext{\Gamma}{A}$ by the family $A$, the result is the extension of (1)
-the extension of $P$ by $A$ and (2) the extension of $Q$ by $A$. The latter is
-simply $Q$ since extension acts trivially at the level of families (and terms).
-
-\subsection{The empty context and the empty families}
-\label{empty}
-We introduce an empty context and an family over $\Gamma$ for every context $\Gamma$. 
-It has been suggested by some to only include empty families and not an
-empty context because an empty context is not necessary, but we do have several
-reasons to include them. Having an empty context requires also rules asserting 
-that a context is the same thing as a family over the empty context and this
-gives the categorical structure on contexts for free once one has it for 
-families. The main ingredients that will be missing from the theory once an
-empty context is avoided are weakening by a context and identity morphisms from
-a context to itself. Including these by hand also requires to formulate all the
-compatibility rules involving weakening once more for the cases of weakening by
-a context and identity terms at contexts. We prefer to state these rules once
-and only once and including an empty context helps in this respect.
-
-We also prefer to have our scheme of compatibility rules as symmetrical as
-possible. The structure of type dependency should look exactly the same in the
-default case as in any context. In that respect, having an empty family but not
-an empty context seems a bit odd. Also, some sets of compatibility rules 
-(like the rules stating that extension is compatible with the empty families)
-will become assymetrical as a result of not including an empty context. Moreover,
-we would eventually like to include a stratification of the theory by means of
-a type judgment (asserting that a family in a context $\Gamma$ is a type) and
-study closed types (i.e.~types in the empty context). It would be possible to
-provide a notion of closed types without having an empty context, but this would
-have to be formulated separately and we would have to restate all the rules for
-types (if any) for closed types all over again.
-
-One of the main uses of the empty context and the empty families is that we
-get the property that the `action on contexts' of an operation is compatible
-with its `action on families'.
-
-The empty family over a context $\Gamma$ is introduced by the following rule
-inference rule:
-\begin{align}
-& \inference
-  { }
-  { \jctx{\emptyc}
-    }
-  \\
-& \inference
-  { \jctx{\Gamma}
-    }
-  { \jfam{\Gamma}{\emptyf[\Gamma]}
-    }
-  \\
-& \inference
-  { \jctxeq{\Gamma}{\Gamma'}
-    }
-  { \jfameq{\Gamma}{\emptyf[\Gamma]}{\emptyf[\Gamma']}
-    }
-\end{align}
-
-By regarding contexts as families of contexts over the empty context, we
-enable ourselves also to speak of terms of contexts. A term of a context
-$\Gamma$ is a term of the family $\Gamma$ over the empty context. These ideas
-are captured in the following convertibility rules:
-\begin{align}
-& \inference
-  { \jctx{\Gamma}
-    }
-  { \jfam{\emptyc}{\Gamma}
-    } 
-& & \inference
-    { \jfam{\emptyc}{\Gamma}
-      }
-    { \jctx{\Gamma}
-      }
-    \\
-& \inference
-  { \jctxeq{\Gamma}{\Delta}
-    }
-  { \jfameq{\emptyc}{\Gamma}{\Delta}
-    }
-& & \inference
-    { \jfameq{\emptyc}{\Gamma}{\Delta}
-      }
-    { \jctxeq{\Gamma}{\Delta}
-      }
-\end{align}
-
-The reader may wonder whether the empty family $\emptyf$ always has a
-term. This shall follow from the rules stating the compatibility of extension
-with the empty families in \autoref{comp-0e} below and from
-identity terms (\autoref{identityterms}).
-\begin{comment}
-When contexts are viewed as statements, the empty context is the statement that
-asserts nothing. Likewise, in any context one can assert nothing and this is
-achieved by the empty family. 
-Since somebody who states nothing asserts it all right, there is a term of the empty
-family over any context. Moreover, any two terms of the empty family are always
-judgmentally equal.
-\begin{align}
-& \inference
-  { }
-  { \jctx{\emptyc}
-    }
-  \\
-& \inference
-  { \jctx{\Gamma}
-    }
-  { \jfam{\Gamma}{\emptyf[\Gamma]}
-    }
-  \\
-& \inference
-  { \jctx{\Gamma}
-    }
-  { \jterm{\Gamma}{\emptyf[\Gamma]}{\emptytm[\Gamma]}
-    }
-  \\
-& \inference
-  { \jterm{\Gamma}{\emptyf[\Gamma]}{x}
-    }
-  { \jtermeq
-      {\Gamma}
-      {\emptyf[\Gamma]}
-      {x}
-      {\emptytm[\Gamma]}
-    }
-  \label{comp-00-t}
-\end{align}
-
-The empty context and the empty families together with the operation of
-extension will provide several shortcuts for the rest
-of the theory of families and terms. This is (in part) obtained by requiring
-that a context is exactly the same thing as a family of contexts over the
-empty context and that judgmental equality of contexts is exactly the same
-as judgmental equality of families of contexts over the empty context.
-
-\label{comp-00}
-Because every context is a family of contexts over the empty context, 
-the empty context $\emptyc$ may be considered as a family of contexts over the empty
-context. When we do this, we get $\emptyf[\emptyc]$.
-\begin{equation}\label{eq:comp-00}
-\inference
-  { }
-  { \jfameq
-      {\emptyc}
-      {\emptyc}
-      {\emptyf[\emptyc]}
-    }
-\end{equation}
-In the future, we shall denote $\emptyf[\Gamma]$ by $\emptyf$. The above rule
-guarantees that this will not cause confusion. Likewise, we shall denote
-$\emptytm[\Gamma]$ by $\emptytm$.
-\end{comment}
-
-\subsubsection{Compatibility of extension with the empty context and families}
-In the following set of inference rules we state that the empty context and
-the empty family are neutral objects for both context extension (the first two
-rules) and family extension (the last two rules).
-\label{comp-e0}\label{comp-0e}
-\begin{align}
-& \inference
-  { \jctx{\Gamma}
-    }
-  { \jctxeq{\ctxext{\emptyc}{\Gamma}}{\Gamma}
-    }
-  \label{comp-0e-c}
-  \\
-& \inference
-  { \jctx{\Gamma}
-    }
-  { \jctxeq{\ctxext{\Gamma}{\emptyf}}{\Gamma}
-    }
-  \label{comp-e0-c}\\
-& \inference
-  { \jfam{\Gamma}{A}
-    }
-  { \jfameq{\Gamma}{\ctxext{\emptyf}{A}}{A}
-    }
-  \label{comp-0e-f}
-  \\
-& \inference
-  { \jfam{\Gamma}{A}
-    }
-  { \jfameq{\Gamma}{{A}{\emptyf}}{A}
-    }
-  \label{comp-e0-f}
-\end{align}
-
-\subsubsection{Family extension restricted to the empty context}
-Strictly speaking we should have used a different notation for context extension
-as for family extension, because the following rule asserting that family extension
-in the empty context is the same thing as context extension would look tautological
-without a difference. So let us denote, only for the moment, context extension
-of $\Gamma$ by $A$ by $(\ctxext{\Gamma}{A})^c$ and family extension of $A$ by
-$P$ in context $\Gamma$ by $(\ctxext{A}{P})^\famsym$.
-
-Note that we may consider a context $\Gamma$ as a family over $\emptyc$ and
-a family $\jfam{\Gamma}{A}$ as a family $\jfam{{\emptyc}{\Gamma}}{A}$. 
-Therefore we will require the following rule:
-\begin{align}
-& \inference
-  { \jfam{\Gamma}{A}
-    }
-  { \jctxeq{(\ctxext{\Gamma}{A})^c}{(\ctxext{\Gamma}{A})^\famsym}
-    }
-\end{align}
-Note that this rule actually justifies that we have not utilized two different
-notations for context extension and family extension.
-
-\subsection{The theory of weakening}
-\label{weakening}
-When $A$ is a family in context $\Gamma$, the operation of weakening by $A$
-takes a family $B$ in context $\Gamma$ and provides a family $\ctxwk{A}{B}$
-in context $\ctxext{\Gamma}{A}$. The context family $\ctxwk{A}{B}$ can be seen
-as the constant family over $\ctxext{\Gamma}{A}$ with value $B$. This idea will
-be axiomatized in the cancellation property of weakening and substitution in
-\autoref{cancellation-ws}. In \autoref{morphisms} we will take the terms 
-$\unfold{\jhom{\Gamma}{A}{B}{f}}$ to be the morphisms of families from $A$ to 
-$B$. These will be at the heart of the categorical structure of the theory.
-
-The weakening operation acts on three levels: on contexts, on families and
-on terms. The `action on contexts' of weakening is the action we described
-above: it takes a family $B$ over $\Gamma$ to a family $\ctxwk{A}{B}$ over
-$\ctxext{\Gamma}{A}$; the `action on families' of weakening takes a family
-$Q$ over $\ctxext{\Gamma}{B}$ to a family $\ctxwk[\famsym]{A}{Q}$ over
-$\ctxext{{\Gamma}{A}}{\ctxwk{A}{B}}$; the `action on terms' of weakening takes
-a term $g$ of $Q$ to a term $\ctxwk[\tmsym]{A}{g}$ of $\ctxwk[\famsym]{A}{Q}$.
-\begin{align}
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{\Gamma}{B}
-    }
-  { \jfam{{\Gamma}{A}}{\ctxwk{A}{B}}
-    }
-& & \inference
-    { \jfameq{\Gamma}{A}{A'}
-      \jfameq{\Gamma}{B}{B'}
-      }
-    { \jfameq{{\Gamma}{A}}{\ctxwk{A}{B}}{\ctxwk{A'}{B'}}
-      }
-    \\
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{{\Gamma}{B}}{Q}
-    }
-  { \jfam{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk[\famsym]{A}{Q}}
-    }
-& & \inference
-    { \jfameq{\Gamma}{A}{A'}
-      \jfameq{{\Gamma}{B}}{Q}{Q'}
-      }
-    { \jfameq
-        {{{\Gamma}{A}}{\ctxwk{A}{B}}}
-        {\ctxwk[\famsym]{A}{Q}}
-        {\ctxwk[\famsym]{A'}{Q'}}
-      }
-    \\
-%& \inference
-%  { \jterm{\Gamma}{B}{y}
-%    }
-%  { \jterm{{\Gamma}{A}}{\ctxwk{A}{B}}{\ctxwk{A}{y}}
-%    }
-%& & \inference
-%    { \jtermeq{\Gamma}{B}{y}{y'}
-%      }
-%    { \jtermeq
-%        {{\Gamma}{A}}
-%        {\ctxwk{A}{B}}
-%        {\ctxwk{A}{y}}
-%        {\ctxwk{A'}{y'}}
-%      }
-%    \\
-& \inference
-  { \jfam{\Gamma}{A}
-    \jterm{{\Gamma}{B}}{Q}{g}
-    }
-  { \jterm{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk[\famsym]{A}{Q}}{\ctxwk[\tmsym]{A}{g}}
-    }
-& & \inference
-    { \jfameq{\Gamma}{A}{A'}
-      \jtermeq{{\Gamma}{B}}{Q}{g}{g'}
-      }
-    { \jtermeq
-        {{{\Gamma}{A}}{\ctxwk{A}{B}}}
-        {\ctxwk[\famsym]{A}{Q}}
-        {\ctxwk[\tmsym]{A}{g}}
-        {\ctxwk[\tmsym]{A'}{g'}}
-      }
-\end{align}
+\subsection{Joining the theory of projections with the theory of substitution}
+\label{sec:esystem-equalities}
 
@@
 \begin{align}
-%& \inference
-%  { \jctx{\Gamma}
-%    }
-%  { \jfameq{\Gamma}{\ctxwk{\Gamma}{\emptyc}}{\emptyf}
-%    }
-%  \\
 & \inference
@@
 \end{align}
-\begin{comment}
-Note that we do not have to axiomatize the action of weakening
-by $A$ on the term of the empty family, because the result will be in the empty 
-family and every term in the empty family is judgmentally equal to the term 
-$\emptytm$.
-\end{comment}
 
@@
 
-\subsubsection{Weakenings of extensions}
-\label{comp-we}
-The following rules assert that when an extended family is weakened, the
-weakening distributes over the extension factors.
-\begin{align}
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{{\Gamma}{B}}{Q}
-    }
-  { \jfameq
-      {\ctxext{\Gamma}{A}}
-      {\ctxwk{A}{\ctxext{B}{Q}}}
-      {\ctxext{\ctxwk{A}{B}}{\ctxwk{A}{Q}}}
-    }
-  \label{comp-we-c}
-  \\
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{{{\Gamma}{B}}{Q}}{R}
-    }
-  { \jfameq
-      {\ctxext{{\Gamma}{A}}{\ctxwk{A}{B}}}
-      {\ctxwk{A}{\ctxext{Q}{R}}}
-      {\ctxext{\ctxwk{A}{Q}}{\ctxwk{A}{R}}}
-    }
-  \label{comp-we-f}
-\end{align}
-When thinking of terms of $\ctxwk{A}{B}$ as morphisms of families from $A$ to
-$B$, this looks already like form of type theoretic choice. It is weaker in that
-it is not stated with function types, yet it is stronger in that it states a
-judgmental equality between two families.
-
-There is also a version of this property where an extended term is weakened.
-This variant is stated and proved in \autoref{comp-we-t}.
-
-\subsubsection{Weakening of weakenings}
-Weakening by a family $A$ in context $\Gamma$ brings things in context $\Gamma$
-to things in context $\ctxext{\Gamma}{A}$. Since we have all the ingredients of
-the theory of contexts, families and terms in the context $\Gamma$ as well it
-has in particular it's own weakening by families. Suppose we have a family
-$\jfam{{\Gamma}{B}}{Q}$ over $B$ in context $\Gamma$. Weakening by $Q$ brings
-things from context $\ctxext{\Gamma}{B}$ to $\ctxext{{\Gamma}{B}}{Q}$. Thus
-we can provide rules asserting what will happen when we first weaken by $Q$ and
-then (via the action on families) by $A$. We will require the following
-inference rules to be valid:
-\label{comp-ww}
-\begin{align}
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{{\Gamma}{B}}{Q}
-    \jfam{{\Gamma}{B}}{R}
-    }
-  { \jfameq
-      {{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{Q}}}
-      {\ctxwk{A}{{Q}{R}}}
-      {\ctxwk{{A}{Q}}{{A}{R}}}
-    }
-  \label{comp-ww-c}\\
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{{\Gamma}{B}}{Q}
-    \jfam{{{\Gamma}{B}}{R}}{S}
-    }
-  { \jfameq
-      {{{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{Q}}}{\ctxwk{A}{{Q}{R}}}}
-      {\ctxwk{A}{{Q}{S}}}
-      {\ctxwk{{A}{Q}}{{A}{S}}}
-    }
-  \label{comp-ww-f}\\
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{{\Gamma}{B}}{Q}
-    \jterm{{{\Gamma}{B}}{R}}{S}{k}
-    }
-  { \jtermeq
-      {{{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{Q}}}{\ctxwk{A}{{Q}{R}}}}
-      {\ctxwk{A}{{Q}{S}}}
-      {\ctxwk{A}{{Q}{k}}}
-      {\ctxwk{{A}{Q}}{{A}{k}}}
-    }
-  \label{comp-ww-t}
-\end{align}
-
-\begin{rmk}
-As an important special case of these inference rules we have the following
-valid inference rules:
-\begin{align*}
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{\Gamma}{B}
-    \jfam{\Gamma}{C}
-    }
-  { \jfameq
-      {{{\Gamma}{A}}{\ctxwk{A}{B}}}
-      {\ctxwk{A}{{B}{C}}}
-      {\ctxwk{{A}{B}}{{A}{C}}}
-    }
-  \\
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{\Gamma}{B}
-    \jfam{{\Gamma}{C}}{R}
-    }
-  { \jfameq
-      {{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{{B}{C}}}}
-      {\ctxwk{A}{{B}{R}}}
-      {\ctxwk{{A}{B}}{{A}{R}}}
-    }
-  \\
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{\Gamma}{B}
-    \jterm{{\Gamma}{C}}{R}{h}
-    }
-  { \jtermeq
-      {{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{{B}{C}}}}
-      {\ctxwk{A}{{B}{R}}}
-      {\ctxwk{A}{{B}{h}}}
-      {\ctxwk{{A}{B}}{{A}{h}}}
-    }
-\end{align*}
-Moreover, we have
-\begin{equation*}
-\inference
-  { \jfam{\Gamma}{A}
-    \jfam{\Gamma}{B}
-    \jterm{\Gamma}{C}{z}
-    }
-  { \jtermeq
-      {{{\Gamma}{A}}{\ctxwk{A}{B}}}
-      {\ctxwk{A}{{B}{C}}}
-      {\ctxwk{A}{{B}{z}}}
-      {\ctxwk{{A}{B}}{{A}{z}}}
-    }
-\end{equation*}
-\end{rmk}
-
 \subsubsection{Weakening by the empty family}
@@
 
-\subsubsection{Currying for weakening}
-\label{comp-ew}
-The rules expressing that extension is compatible with weakening assert that
-weakening by an extension is the same thing as weakening twice in the
-appropriate way.
-\begin{align}
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{{\Gamma}{A}}{P}
-    \jfam{\Gamma}{B}
-    }
-  { \jfameq
-      {{{\Gamma}{A}}{P}}
-      {\ctxwk{\ctxext{A}{P}}{B}}
-      {\ctxwk{P}{{A}{B}}}
-    }
-  \label{comp-ew-c}\\
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{{\Gamma}{A}}{P}
-    \jfam{{\Gamma}{B}}{Q}
-    }
-  { \jfameq
-      {{{{\Gamma}{A}}{P}}{\ctxwk{P}{{A}{B}}}}
-      {\ctxwk{\ctxext{A}{P}}{Q}}
-      {\ctxwk{P}{{A}{Q}}}
-    }
-  \label{comp-ew-f}\\
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{{\Gamma}{A}}{P}
-    \jterm{{\Gamma}{B}}{Q}{g}
-    }
-  { \jtermeq
-      {{{{\Gamma}{A}}{P}}{\ctxwk{P}{{A}{B}}}}
-      {\ctxwk{P}{{A}{Q}}}
-      {\ctxwk{\ctxext{A}{P}}{g}}
-      {\ctxwk{P}{{A}{g}}}
-    } 
-  \label{comp-ew-t}
-\end{align}
-
-\begin{comment}
-\subsubsection{Weakening by a context}
-Since we have not required that there is an empty context with the property that
-a context is the same thing as a family in the empty context, we will have to
-explicitly introduce weakening by contexts. Since it will allways be clear
-whether something is weakened by a context or by a family, we shall use the
-same notation for weakening by contexts as for weakening by families. We shall
-have to repeat all the compatibility rules for weakening by families for
-weakening by contexts; there is no way around it. It is
-actually this sort of inconvenience that has originally led us to include an
-empty context and be done with it.
-
-In the following inference rules we assume that we have a context $\Gamma$ and
-that $\jctxeq{\Gamma}{\Gamma'}$ in the rules where $\Gamma'$ makes an appearence.
-\begin{align}
-& \inference
-  { \jctx{\Delta}
-    }
-  { \jfam{\Gamma}{\ctxwk{\Gamma}{\Delta}}
-    }
-& & \inference
-  { \jctxeq{\Delta}{\Delta'}
-    }
-  { \jfameq{\Gamma}{\ctxwk{\Gamma}{\Delta}}{\ctxwk{\Gamma'}{\Delta'}}
-    }
-  \\
-& \inference
-  { \jfam{\Delta}{B}
-    }
-  { \jfam{\ctxext{\Gamma}{\ctxwk{\Gamma}{\Delta}}}{\ctxwk{\Gamma}{B}}
-    }
-& & \inference
-  { \jfameq{\Delta}{B}{B'}
-    }
-  { \jfameq
-      {\ctxext{\Gamma}{\ctxwk{\Gamma}{\Delta}}}
-      {\ctxwk{\Gamma}{B}}
-      {\ctxwk{\Gamma'}{B'}}
-    }
-  \\
-& \inference
-  { \jterm{\Delta}{B}{y}
-    }
-  { \jterm
-      {\ctxext{\Gamma}{\ctxwk{\Gamma}{\Delta}}}
-      {\ctxwk{\Gamma}{B}}
-      {\ctxwk{\Gamma}{y}}
-    }
-& & \inference
-  { \jtermeq{\Delta}{B}{y}{y'}
-    }
-  { \jtermeq
-      {\ctxext{\Gamma}{\ctxwk{\Gamma}{\Delta}}}
-      {\ctxwk{\Gamma}{B}}
-      {\ctxwk{\Gamma}{y}}
-      {\ctxwk{\Gamma}{y'}}
-    }
-\end{align}
-Note that we do not employ three different notations for the three different
-actions of weakening by a context because there can be no confusion about which
-of the three actions is meant. If there were an empty context we should have
-distinguished the three as we did for weakening by a family.
-
-We shall have similar compatibility rules for weakening by a context as we did
-for weakening by a family.
-
-{\color{red}To do. It will look exactly the same as for weakening by families}
-\end{comment}
-
-\subsection{The theory of projections}
-\label{identityterms}
-We formulate the theory of projections on top of the theory of weakening
-described in \autoref{weakening}. In the theory of projections we introduce
-the identity terms. The identity term at a family $A$ in context $\Gamma$ is
-a term $\jterm{{\Gamma}{A}}{\ctxwk{A}{A}}{\idtm{A}}$, so weakening is necessary.
-Together with weakening, the identity terms will provide for the projections.
-The projection from $\cftext{A}{P}$ to $A$ will be the term
-$\jterm{{\Gamma}{{A}{P}}}{\ctxwk{\ctxext{A}{P}}{A}}{\ctxwk{P}{\idtm{A}}}$ and
-will be studied extensively in \autoref{extension-on-terms}.
-
-The following inference rules are required to be valid:
-\begin{align}
-& \inference
-  { \jfam{\Gamma}{A}
-    }
-  { \unfoldall{\jhom{\Gamma}{A}{A}{\idtm{A}}}
-    }
-& & \inference
-    { \jfameq{\Gamma}{A}{A'}
-      }
-    { \unfoldall{\jhomeq{\Gamma}{A}{A}{\idtm{A}}{\idtm{A'}}}
-      }
-\end{align}
-
-\begin{comment}
-We stress again
-that because we have not required there to be an empty context, we need to
-explicitly call for identity terms $\idtm{\Gamma}$ of $\ctxwk{\Gamma}{\Gamma}$
-in context $\Gamma$ for each context $\Gamma$. Thus, we will also require
-the inference rule
-\begin{align}
-& \inference
-  { \jctx{\Gamma}
-    }
-  { \jterm{\Gamma}{\ctxwk{\Gamma}{\Gamma}}{\idtm{\Gamma}}
-    }
-& &
-  \inference
-  { \jctxeq{\Gamma}{\Gamma'}
-    }
-  { \jtermeq{\Gamma}{\ctxwk{\Gamma}{\Gamma}}{\idtm{\Gamma}}{\idtm{\Gamma'}}
-    }
-\end{align}
-For the following inference rules, there should also be versions of identity
-terms at contexts.
-\end{comment}
-
-\subsubsection{The identity term of an extended family}
-The fact that identity terms are compatible with extensions is derived in
-\autoref{comp-ie}.
-
-\subsubsection{The identity term of a weakened family}
-\label{comp-wi}
-The identity term of a weakened family is the weakened identity term:
-\begin{equation}
-\inference
-  { \jfam{\Gamma}{A}
-    \jfam{{\Gamma}{B}}{Q}
-    }
-  { \unfoldall{\jhomeq
-      {{{\Gamma}{A}}{\ctxwk{A}{B}}}
-      {\ctxwk{A}{Q}}
-      {\ctxwk{A}{Q}}
-      {\ctxwk{A}{\idtm{Q}}}
-      {\idtm{\ctxwk{A}{Q}}}
-      }
-    }
-  \label{comp-wi-t}
-\end{equation}
-
-\begin{rmk}
-An important special case is the judgmental equality
-\begin{equation*}
-\unfoldall{\jhomeq
-      {{\Gamma}{A}}
-      {\ctxwk{A}{B}}
-      {\ctxwk{A}{B}}
-      {\ctxwk{A}{\idtm{B}}}
-      {\idtm{\ctxwk{A}{B}}}
-      }
-\end{equation*}
-for a family $\jfam{\Gamma}{B}$.
-\end{rmk}
-
-\subsection{The theory of substitution}
-\label{substitution}
-The theory of substitution can be formulated separately from the theory of
-weakening. Therefore, we will only require the theory of extension and the
-empty context and families. In \autoref{sec:esystem-equalities} we will
-provide the additional judgmental equality rules that will be required when
-combining the theories of substitution and of projections.
-
-Given a family $P$ over $A$ and a term $x$ of $A$, substitution gives a way to
-consider the \emph{fiber $\subst{x}{P}$ of $P$ at $x$}. As was the case with
-weakening, the substitution operation comes in three parts: the `action on
-contexts' of substitution is the operation just described; the `action on
-families' of substitution takes a family $Q$ over $\ctxext{{\Gamma}{A}}{P}$
-to a family $\subst[\famsym]{x}{Q}$ over $\ctxext{\Gamma}{\subst{x}{P}}$; the
-`action on terms' of substitution takes a term $g$ of $Q$ over
-$\ctxext{{\Gamma}{A}}{P}$ to a term $\subst[\tmsym]{x}{g}$ of 
-$\subst[\famsym]{x}{Q}$.
-\begin{align}
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jfam{{\Gamma}{A}}{P}
-    }
-  { \jfam{\Gamma}{\subst{x}{P}}
-    }
-& & \inference
-    { \jtermeq{\Gamma}{A}{x}{x'}
-      \jfameq{{\Gamma}{A}}{P}{P'}
-      }
-    { \jfameq{\Gamma}{\subst{x}{P}}{\subst{x'}{P'}}
-      }
-    \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jfam{{{\Gamma}{A}}{P}}{Q}
-    }
-  { \jfam{{\Gamma}{\subst{x}{P}}}{\subst[\famsym]{x}{Q}}
-    }
-& & \inference
-    { \jtermeq{\Gamma}{A}{x}{x'}
-      \jfameq{{{\Gamma}{A}}{P}}{Q}{Q'}
-      }
-    { \jfameq{{\Gamma}{\subst{x}{P}}}{\subst[\famsym]{x}{Q}}{\subst[\famsym]{x'}{Q'}}
-      }
-    \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jterm{{{\Gamma}{A}}{P}}{Q}{g}
-    }
-  { \jterm{{\Gamma}{\subst{x}{P}}}{\subst[\famsym]{x}{Q}}{\subst[\tmsym]{x}{g}}
-    }
-& & \inference
-    { \jtermeq{\Gamma}{A}{x}{x'}
-      \jtermeq{{{\Gamma}{A}}{P}}{Q}{g}{g'}
-      }
-    { \jtermeq
-        {{\Gamma}{\subst{x}{P}}}
-        {\subst[\famsym]{x}{Q}}
-        {\subst[\tmsym]{x}{g}}
-        {\subst[\tmsym]{x'}{g'}}
-      }
-\end{align}
-
 \subsubsection{Fibers of an empty family}
@@
 \end{align}
-\begin{comment}
-Note again that we do not have to axiomatize the action of substitution by $x$ on the term
-of the empty family, because the result will be in the empty family and every
-term in the empty family is judgmentally equal to the term $\emptytm$.
-\end{comment}
 
@@
 meaningfully state the definition of composition.
-
-\subsubsection{Fibers of extensions}
-\label{comp-se}
-The following inference rules assert that if we take the fiber of an extended
-family at a term $x$ of $A$ in context $\Gamma$, the substitution by $x$
-distributes over the factors of the extension.
-\begin{align}
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jfam{{{\Gamma}{A}}{P}}{Q}
-    }
-  { \jfameq
-      {\Gamma}
-      {\subst{x}{\ctxext{P}{Q}}}
-      {\ctxext{\subst{x}{P}}{\subst{x}{Q}}}
-    }
-  \label{comp-se-c}
-  \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jfam{{{{\Gamma}{A}}{P}}{Q}}{R}
-    }
-  { \jfameq
-      {{\Gamma}{\subst{x}{P}}}
-      {\subst{x}{\ctxext{Q}{R}}}
-      {\ctxext{\subst{x}{Q}}{\subst{x}{R}}}
-    }
-  \label{comp-se-f}
-\end{align}
-There is also a version of this statement in which extended terms are considered.
-This variant is stated and proved in \autoref{comp-se-t}.
-
-\subsubsection{Fibers of fibers}
-\label{comp-ss}
-
-The following rules explain what happens when we first substitute by a term
-$g$ of $Q$ over $P$ in context $\ctxext{\Gamma}{A}$ and then by a term $x$ of
-$A$:
-\begin{align}
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jterm{{{\Gamma}{A}}{P}}{Q}{g}
-    \jfam{{{{\Gamma}{A}}{P}}{Q}}{R}
-    }
-  { \jfameq
-      {{{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}}
-      {\subst{x}{{g}{R}}}
-      {\subst{{x}{g}}{{x}{R}}}
-    }
-  \label{comp-ss-c}
-  \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jterm{{{\Gamma}{A}}{P}}{Q}{g}
-    \jfam{{{{{\Gamma}{A}}{P}}{Q}}{R}}{S}
-    }
-  { \jfameq
-      {{{{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}}{\subst{x}{{g}{R}}}}
-      {\subst{x}{{g}{S}}}
-      {\subst{{x}{g}}{{x}{S}}}
-    }
-  \label{comp-ss-f}
-  \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jterm{{{\Gamma}{A}}{P}}{Q}{g}
-    \jterm{{{{{\Gamma}{A}}{P}}{Q}}{R}}{S}{k}
-    }
-  { \jtermeq
-      {{{{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}}{\subst{x}{{g}{R}}}}
-      {\subst{x}{{g}{S}}}
-      {\subst{x}{{g}{k}}}
-      {\subst{{x}{g}}{{x}{k}}}
-    }
-  \label{comp-ss-t}
-\end{align}
-
-\begin{rmk}
-As an important special case of these inference rules we have the following
-valid inference rules:
-\begin{align*}
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jterm{{\Gamma}{A}}{P}{f}
-    \jfam{{{\Gamma}{A}}{P}}{Q}
-    }
-  { \jfameq
-      {\Gamma}
-      {\subst{x}{{f}{Q}}}
-      {\subst{{x}{f}}{{x}{Q}}}
-    }
-  \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jterm{{\Gamma}{A}}{P}{f}
-    \jfam{{{{\Gamma}{A}}{P}}{Q}}{R}
-    }
-  { \jfameq
-      {{\Gamma}{\subst{x}{{f}{Q}}}}
-      {\subst{x}{{f}{R}}}
-      {\subst{{x}{f}}{{x}{R}}}
-    }
-  \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jterm{{\Gamma}{A}}{P}{f}
-    \jterm{{{{\Gamma}{A}}{P}}{Q}}{R}{h}
-    }
-  { \jtermeq
-      {{\Gamma}{\subst{x}{{f}{Q}}}}
-      {\subst{x}{{f}{R}}}
-      {\subst{x}{{f}{h}}}
-      {\subst{{x}{f}}{{x}{h}}}
-    }
-\end{align*}
-Moreover, we get
-\begin{equation*}
-\inference
-  { \jterm{\Gamma}{A}{x}
-    \jterm{{\Gamma}{A}}{P}{f}
-    \jterm{{{\Gamma}{A}}{P}}{Q}{g}
-    }
-  { \jtermeq
-      {\Gamma}
-      {\subst{x}{{f}{Q}}}
-      {\subst{x}{{f}{g}}}
-      {\subst{{x}{f}}{{x}{g}}}
-    }
-\end{equation*}
-\end{rmk}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Joining the theory of projections with the theory of substitution}
-\label{sec:esystem-equalities}
-
-\begin{comment}
-We have already seen that empty families, extension, weakening, substitution
-and the identity terms are compatible with each other. There are, however, a
-few more judgmental equalities explaining what happens when
-\begin{enumerate}
-\item something is weakend by the empty family. This is used to consider a term
-      $x$ of $A$ in context $\Gamma$ as a morphism from $\emptyf$ to $A$ in
-      context $\Gamma$;
-\item something is weakened by an extended family. This is Currying for weakening;
-\item something is first weakened by a family $A$ in context $\Gamma$ and then
-      substituted by a term of $A$. This is used to show that composition is
-      compatible with itself (i.e. that composition is associative);
-\item the identity terms are joining the party. Identity terms are empty contexts
-      in certain models, so they should behave accordingly.
-\end{enumerate}
-In other words, the rules seem to not fit in the scheme that all operations are
-compatible with each other, but in fact each of these rules plays a role in
-completing the scheme (work out further).
-\end{comment}
-
-\begin{comment}
-In this section we extend the judgmental equality relation in such a way that
-the empty context, extension, weakening, substitution and the identity terms
-all become compatible with each other. Since we have five operations, there
-are going to be $25$ finite sets of rules explaining how each operation is
-compatible with each of the operations.
-
-In addition to these $25$ finite sets
-of compatibility rules, there are also two sets of rules explaining the
-cancellation properties of weakening and substitution and of the identity terms.
-
-A typical compatibility axiom captures the idea that a diagram like
-\begin{equation*}
-\begin{tikzcd}[column sep=huge]
-\jfam{\Gamma}{\blank} 
-  \ar{r}{C\mapsto\ctxwk{B}{C}} 
-  \ar{d}[swap]{C\mapsto\ctxwk{A}{C}} 
-& \jfam{{\Gamma}{B}}{\blank} 
-  \ar{d}{Q\mapsto\ctxwk{A}{Q}}
-  \\
-\jfam{{\Gamma}{A}}{\blank} 
-  \ar{r}[swap]{P\mapsto\ctxwk{{A}{B}}{P}} 
-& \jfam{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\blank}
-\end{tikzcd}
-\end{equation*}
-in the case of two consecutive weakenings, or
-\begin{equation*}
-\begin{tikzcd}[column sep=huge]
-\jfam{{\Gamma}{A}}{\blank} 
-  \ar{d}[swap]{Q\mapsto\subst{x}{Q}} 
-  \ar{r}{Q\mapsto\ctxwk{P}{Q}} 
-& \jfam{{{\Gamma}{A}}{P}}{\blank} 
-    \ar{d}{R\mapsto\subst{x}{R}}
-  \\ 
-\jfam{\Gamma}{\blank} 
-  \ar{r}[swap]{B\mapsto\ctxwk{\subst{x}{P}}{B}} 
-& \jfam{{\Gamma}{\subst{x}{P}}}{\blank}
-\end{tikzcd}
-\end{equation*}
-in the case of a combination of weakening and substitution, commutes. We won't
-actually draw any of these diagrams in this section, but readers are encouraged
-to do so if they think it can help them understanding the axioms. Another thing
-one could keep in mind with the axioms that involve weakening is that all of
-them become completely obvious when weakening is made notationally invisible.
-Of course, we have an explicit notation for weakening on purpuse, but it is the
-usual practice of type theorists to make weakening invisible.
-
-The table below indicates where
-we can find the axioms of each of the combinations of operators. The reference
-at the row of operation $X$ and the column of operation $Y$ refers to the
-axiom which explains $Y$ distributes over the arguments of $X$ when $Y$ is
-applied after $X$. The empty context has no arguments, whereas extension,
-and weakening have two arguments which are both of the family-kind, substitution
-has two arguments of which one is a term and the other is a family and
-idenitity terms have one argument, a family.
-\begin{center}
-\begin{tabular}{r|rrrr}
-& ext & wk & subst & id\\
-\hline
-extension & \autoref{comp-ee}* &  & \cmark & \cmark \\
-weakening & \autoref{comp-we}* & \autoref{comp-ww} & \autoref{comp-sw} & \autoref{comp-wi} \\
-substitution & \autoref{comp-se}* & \autoref{comp-ws} & \autoref{comp-ss} & \autoref{comp-si} \\
-identity terms & \cmark & \autoref{comp-wi} & \autoref{comp-si} & \xmark
-\end{tabular}
-\end{center}
-Apart from references to sections where axioms are given, we also see checkmarks,
-crosses, marked references and duplicated references in this table. 
-The checkmarks indicate that the desired property
-can be proved and need not be axiomatized. To clarify those:
-\begin{itemize}
-\item The identity term of the empty context is the unique term of the empty
-family. Recall that we already have axiomatized in \autoref{empty} that every
-term of the empty family is judgmentally equal to $\emptytm$.
-\item The identity term of an extension can be explained by means of the notion
-of \emph{extension on terms}. This notion is defined in \autoref{extension-on-terms}
-using only identity terms and substitution. With extension on terms we can
-combine a term $\jterm{\Gamma}{A}{x}$ and a term $\jterm{\Gamma}{\subst{x}{P}}{u}$
-into a term $\jterm{\Gamma}{{A}{P}}{\tmext{x}{u}}$, which may be seen as the
-pair consisting of $x$ and $u$. Since it is a defined notion, we will
-be able to derive lots of properties for extension of terms. One of them explains
-how we can see the identity term of an extension as the extension of two certain
-terms, see \autoref{comp-ie}.
-\item Using extension on terms we can also make sense of `fibers at an extension'
-because that would exactly be the fiber at an extended term. We will see in
-\autoref{comp-es} that, for instance, $\subst{\tmext{x}{u}}{Q}\jdeq\subst{u}{{x}{Q}}$
-for any family $\jfam{{{\Gamma}{A}}{P}}{Q}$ and any two terms $\jterm{\Gamma}{A}{x}$
-and $\jterm{\Gamma}{\subst{x}{P}}{u}$. That's a form of currying.
-\end{itemize}
-The spots with a \xmark simply make no sense 
-{\color{red}(or do they, when the term model
-is better understood (by me)?)}. The marked references point to
-axioms which \emph{do} have a variant for terms, but those require the notion
-of extension on terms, can be proved and therefore need not be included in our
-list of compatibility axioms. In each of these sections, we give a forward
-reference to the lemma in which this variant is stated and proved.
-\end{comment}
 
@@
 
-\subsubsection{The cancellation properties of identity terms}
+\subsubsection{The cancellation property of identity terms}
 \label{cancellation-i}
@@
 \end{align}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{The possiblity of types in the theory of contexts, families and
-terms}
-\label{types}
-We have deliberately not spoken of types so far because we have taken the point
-of view that a type in a context is nothing but a family in that context which
-belongs to the class of types. We think of types as \emph{irreducible} families,
-i.e.\ families which are neither empty nor the extension of two
-families which are both not empty (in algebraic terminology: which
-are both non-trivial). To allow ourselves to speak of types we introduce two
-new judgments: the judgment that something is a type and the judgment that two
-types are equal.
-\begin{align*}
-\jalign\jtype{\Gamma}{A} 
-& \jalign\jtypeeq{\Gamma}{A}{B}
-\end{align*}
-But only families of contexts are eligible to be types. If $A$ is a type
-in context $\Gamma$, then $A$ is also a family of contexts over $\Gamma$. 
-Moreover, two types in context $\Gamma$ are judgmentally equal precisely when they are equal
-as context families and if a family $B$ of contexts over $\Gamma$ is
-judgmentally equal to a type $A$ in context $\Gamma$, then $B$ is a type in
-context $\Gamma$. This is expressed by the following four inference rules:
-\begin{align*}
-& \inference
-  { \jtype{\Gamma}{A}
-    }
-  { \jfam{\Gamma}{A}
-    }
-& & \inference
-    { \jtypeeq{\Gamma}{A}{B}
-      }
-    { \jfameq{\Gamma}{A}{B}
-      }
-    \\
-& \inference
-  { \jtype{\Gamma}{A}
-    \jfameq{\Gamma}{A}{B}
-    }
-  { \jtype{\Gamma}{B}
-    }
-& & \inference
-    { \jtype{\Gamma}{A}
-      \jfameq{\Gamma}{A}{B}
-      }
-    { \jtypeeq{\Gamma}{A}{B}
-      }
-\end{align*}
-As pointed out at the beginning of this subsection, 
-we do not assume that the empty family is a type, that would be like
-assuming that the multiplicative unit of a ring is prime.
-
-\begin{comment}
-We envision a
-categorical interpretation of the theory of contexts, families and terms where 
-contexts are interpreted as finite paths of arrows belonging to a predetermined
-class. The empty family would be such a path of length $0$ whereas the types
-would be such paths of length $1$.
-\end{comment}
-
-We add rules asserting that a weakened type is again a type that 
-substitution preserves the property of being a type:
-\begin{align}
-& \inference
-  { \jfam{\Gamma}{B}
-    \jtype{\ctxext{\Gamma}{B}}{Q}
-    }
-  { \jtype{\ctxext{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{Q}}
-    }
-  \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jtype{\ctxext{{\Gamma}{A}}{P}}{Q}
-    }
-  { \jtype{\ctxext{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}
-    }
-\end{align}
-With only the current rules, the possibility of making the judgment that
-something is a type does not add much to the theory of contexts, families and
-terms. Nevertheless, when studying models, having an interpretation for the
-judgment that something is a type will allow for the possibility to study
-conditions such as the one asserting that every family `factorizes' uniquely
-as multiple applications of extension to types, analogous to the condition on
-unique factorization domains in ring theory. All the tradional models of type
-theory should translate to models with such a condition, simply because
-traditionally contexts are viewed as lists of variable (and type) declarations.

### 2014-10-28 Rules/rules_projections.tex
reorganization

--- 2014-09-20 article-models-tt.tex
@@
-\section{The generalized algebraic theory of contexts, families and terms (the E-system)}\label{tt}
-In this section we give a description of dependent type theory before type
-constructors. Apart from contexts, families and terms -- which provide for the
-core of the language of dependent logic -- the basic ingredients
-of this theory will be the operations of extension on context and on families, 
-and empty context and empty families, weakening, substitution
-and the identity terms. The resulting theory can be seen as a manifestation of the 
-structure underlying dependent type theory.
-
-We will
-formulate the theory of contexts, families and terms in such a way that contexts aren't defined
-to be lists of variable declarations. The variable-free (a.k.a.~\emph{name-free}) approach 
-we take here is rather different than those appearing in 
-\cite{hofmann1995extensional,TheBook} but it has appeared in the work of Coquand
-and in \cite{Dybjer1996}.
-The main reason we don't let variable declarations in is that we don't see them 
-in the internal models either. This way we also set out to a more algebraic 
-approach of type theory and higher category theory. Thirdly, we will not have to
-be burdened with superficial comments about variables being bounded or not, or 
-fresh or free or not occuring at all.
-
-In the current presentation, contexts can be seen as binary
-planar trees of which the leaves are (families of) contexts. 
-The judgmental equality relation on contexts is an equivalence relation which 
-expresses that binary planar
-trees of contexts are judgmentally equal if their leaves are, taking only
-(the isomorphism class of) 
-the order of the leaves into account \emph{and not the actual shape of the three}.
-The intuition behind this equivalence relation is indeed that unstructured
-(i.e.~unbracketed) lists such as the lists of variable declarations which
-usually appear in type theoretical syntax, may be regarded as contexts.
-
-Besides contexts, families and terms there will also be a notion of `type in
-a context', which one may assume.
-A context is eligible to be a \emph{closed type} and over a context, a family
-may be a type. We have the intuition that `being a type'
-expresses the property of `being atomic or irreducible'. With a notion of types,
-one has the means to require a stratification by which every context is a finite
-extension of types starting at a closed type, and every family is likewise a
-finite extension of types. The idea behind this is to recover the theory of
-\cite{Garner2014} {\color{red}(and possibly B-systems?)}. We stress that
-we will not include axioms asserting such a stratification in the E-system
-and we leave discussion on the possibility of a typing judgment until the
-end of this section.
-
-With families of contexts being the principal things of
-study, we have added a bit more generality to the theory, compared to the
-theory of categories with families presented in \cite{Dybjer1996}. Here are
-a couple of reasons for this extra generality:
-\begin{enumerate}
-\item We want to be very sure that we can interpret all the rules of type theory
-in a straightforward manner in any context.
-\item We do want do end up with a categorical structure on the contexts, derived
-from the theory by requiring that a morphism from $\Gamma$ to $\Delta$ is a
-term of the weakening of $\Delta$ by $\Gamma$ in context $\Gamma$. Thus, we
-must be able to lift the context $\Delta$ to become a family of contexts over
-$\Gamma$ and it is not reasonable to expect that this results in a type.
-\item It is very easy in this setting to assert that an operation has an action
-on very high levels of families, i.e.~when $F$ is an operation taking things in
-context $\Gamma$ to things in context $\Delta$, then it takes a family $A$ in
-context $\Gamma$ to $F(A)$ in context $\Delta$; a family $P$ in context
-$\ctxext{\Gamma}{A}$ to a family $F(P)$ in context $\ctxext{\Delta}{(F(A))}$,
-and so on. The reason it becomes simple is that we don't have to explain what
-`and so on' means.
-\end{enumerate}
-
-We will formulate fairly strict rules governing the judgmental equalities,
-expressing that extension, weakening and substitution are combatible with
-each other in a judgmental manner. This does not, however, diminish the role
-of isomorphisms or of homotopies could play in the theory once identity types
-are added. Indeed, types could still have non-trivial identity relations and
-the category of types in a certain context could genuinely display higher
-categorical structure, or so we conjecture.
-
-Much of the rules we state are just compatibility rules of extension, weakening
-and substitution with each other. In a way, these rules assert that our contexts
-are just structureless lists of contexts and that likewise terms are structureless
-lists of terms. They are structureless in the sense that the order in which
-they are formed by pairing up is irrelevant. We note that this causes complications
-in the traditional way that categorical sematics of type theory is implemented,
-where contexts become objects of the category which is supposed to model type
-theory. The reason for this is that context extension will not satisfy all the
-compatibility rules we're about to state. The first step to resolving this is taking
-the types in the empty context as the objects.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Judgments and valid inference rules}\label{judgments}
-The theory we describe here is a theory of contexts, families of
-contexts and terms thereof. The families of contexts are by some authors called
-dependent contexts, but they are handled a bit differently here because they
-become the primary object of study. Dependent contexts can be types; they could
-be seen as atomic or indecomposable dependent contexts.
-
-Thus we make eight kinds of judgments: ``$\Gamma$ is a context'',
-``$A$ is a family of contexts over $\Gamma$'', ``$A$ is a type in context $\Gamma$''
-and ``$x$ is a term of the family $A$ of contexts over $\Gamma$''. The other four
-judgments are for judgmental equality.
-\begin{align*}
-\jalign\jctx{\Gamma} 
-& \jalign\jctxeq{\Gamma}{\Gamma'}
-  \\
-\jalign\jfam{\Gamma}{A} 
-& \jalign\jfameq{\Gamma}{A}{B}
-  \\
-\jalign\jterm{\Gamma}{A}{x} 
-& \jalign\jtermeq{\Gamma}{A}{x}{y}.
-\end{align*}
-
-Strictly speaking, we have three different judgmental equalities in play and one
-could request for a notational difference to signify that fact. For instance,
-we could denote the judgmental equalities of contexts, families and terms by
-$\jdeq_c$, $\jdeq_f$ and $\jdeq_t$ respectively. It will, however, always be
-clear which of the three kinds of judgmental equality is meant when we assert
-a judgmental equality and therefore we shall not bother to make this notational
-distinction.
-
-We note that what we call families over contexts
-here could also have been named dependent contexts or telescopes, see
-\cite{deBruijn1991,hofmann1995extensional}. The term family is in agreement
-with the terminology scheme of \cite{TheBook}, though the reader should be
-warned that the notion of familie means something slightly different there than
-it does here.
-
-The theory of contexts, families and terms shall be described by means of a
-declaration of valid inference rules. An inference rule is an expression of the
-form
-\begin{equation*}
-\inference{\mathcal{H}_1\quad\cdots\quad\mathcal{H}_n}{\mathcal{J}}
-\end{equation*}
-where each of $\mathcal{H}_1$, ..., $\mathcal{H}_n$ and $\mathcal{J}$ is 
-judgment. The judgments $\mathcal{H}_1$, ..., $\mathcal{H}_n$ are called the
-hypotheses and the judgment $\mathcal{J}$ is the conclusion of the inference
-rule. When a judgment is the conclusion of a valid inference rule, we also say
-that the judgment is valid or well-formed. In fact, the only way to construct
-new judgments is via the valid inference rules we describe in this section.
-
-The class of valid inference rules is closed under composition and under
-rearranging the hypotheses. That
-means that if
-\begin{equation*}
-\inference{\mathcal{I}_{1k}\quad\cdots\quad\mathcal{I}_{m_kk}}{\mathcal{H}_k}
-\end{equation*}
-and
-\begin{equation*}
-\inference{\mathcal{H}_1\quad\cdots\quad\mathcal{H}_n}{\mathcal{J}}
-\end{equation*}
-are valid inference rules for $1\leq k\leq n$, then
-\begin{equation*}
-\inference{\mathcal{I}_{11}\quad\cdots\quad\mathcal{I}_{m_11}\quad\cdots\quad\mathcal{I}_{1n}\quad\cdots\quad\mathcal{I}_{m_nn}}{\mathcal{J}}
-\end{equation*}
-is also a valid inference rule. Also, in a valid inference rule 
-duplicated hypotheses may always be contracted and hypotheses may always be 
-reordered; the result is still considered a valid inference rule.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{The basic rules for judgmental equality}
-The rules for judgmental equality establish that it is an equivalence relation
-in all three cases (contexts, families and terms). Thus, the following inference
-rules shall be required to be valid:
-\bgroup\small
-\begin{align*}
-& \inference
-  { \jctx{\Gamma}
-    }
-  { \jctxeq{\Gamma}{\Gamma}
-    } 
-& & \inference
-    { \jctxeq{\Gamma}{\Delta}
-      }
-    { \jctxeq{\Delta}{\Gamma}
-      } 
-& & \inference
-    { \jctxeq{\Gamma}{\Delta}
-      \jctxeq{\Delta}{\greek{E}}
-      }
-    { \jctxeq{\Gamma}{\greek{E}}
-      }
-    \\
-& \inference
-  { \jfam{\Gamma}{A}
-    }
-  { \jfameq{\Gamma}{A}{A}
-    } 
-& & \inference
-    { \jfameq{\Gamma}{A}{B}
-      }
-    { \jfameq{\Gamma}{B}{A}
-      }
-& & \inference
-    { \jfameq{\Gamma}{A}{B}
-      \jfameq{\Gamma}{B}{C}
-      }
-    { \jfameq{\Gamma}{A}{C}
-      }
-    \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    }
-  { \jtermeq{\Gamma}{A}{x}{x}
-    }
-& & \inference
-    { \jtermeq{\Gamma}{A}{x}{y}
-      }
-    { \jtermeq{\Gamma}{A}{y}{x}
-      }
-& & \inference
-    { \jtermeq{\Gamma}{A}{x}{y}
-      \jtermeq{\Gamma}{A}{y}{z}
-      }
-    { \jtermeq{\Gamma}{A}{x}{z}
-      }
-\end{align*}
-\egroup
-
-The following convertibility rules are responsible for the strictness
-of judgmental equality, which sets it apart from equivalences or identifications:
-\begin{align*}
-& \inference
-  { \jctxeq{\Gamma}{\Delta}
-    \jfam{\Gamma}{A}
-    }
-  { \jfam{\Delta}{A}
-    }
-& & \inference
-    { \jctxeq{\Gamma}{\Delta}
-      \jfameq{\Gamma}{A}{B}
-      }
-    { \jfameq{\Delta}{A}{B}
-      }
-    \\
-& \inference
-  { \jctxeq{\Gamma}{\Delta}
-    \jterm{\Gamma}{A}{x}
-    }
-  { \jterm{\Delta}{A}{x}
-    }
-& & \inference
-    { \jctxeq{\Gamma}{\Delta}
-      \jtermeq{\Gamma}{A}{x}{y}
-      }
-    { \jtermeq{\Delta}{A}{x}{y}
-      }
-    \\
-& \inference
-  { \jfameq{\Gamma}{A}{B}
-    \jterm{\Gamma}{A}{x}
-    }
-  { \jterm{\Gamma}{B}{x}
-    }
-& & \inference
-    { \jfameq{\Gamma}{A}{B}
-      \jtermeq{\Gamma}{A}{x}{y}
-      }
-    { \jtermeq{\Gamma}{B}{x}{y}
-      }
-\end{align*}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{The operations of the theory of contexts, families and terms}
-In this subsection we introduce the notions of
-extension on context and on families, the empty context and the empty families, 
-weakening, 
-substitution and identity terms.
-
-Weakening and substitution are both
-operations of the theory of contexts, families and terms and therefore they will
-have three levels of actions: on contexts (at the appropriate level), families
-and terms.
-
-\begin{comment}
-Generally, 
-each operation should
-have an action on contexts, on families and on terms at the `appropriate level'
-and should be compatible with each of the operations we introduce here. In
-that will become more practical after the theory of contexts, families and terms
-is formulated, this means that when we will write down a type theory that
-interprets itself in \autoref{ttmodels}, each of the notions of
-extension, empty family, weakening, substitution and identity terms are going to be internalized
-as terms of the theory and all terms are going to be compatible with these.
-That vague idea will be our guiding light.
-
-So in principle, the introductions of each of the empty context, extension,
-weakening, substitution and identity terms should come in three parts. In 
-practice, they do not all come in three parts.
-
-Apart from the empty context, there will indeed be an empty family and a term
-of the empty family (which will be deemed unique). But extension will
-not have an explicit action `at the level of families' since extension
-is used to define higher families and the action would have to be trivial
-anyway. The empty context and extension provide the essentials of the rest of
-the theory in the sense that higher families are families over extensions,
-contexts are families over the empty context and the `action on families' of
-any operation applied to families over the empty context is the same as the
-original operation. Thus, when we have introduced the operations of weakening
-and substitution we will see that they also have an `action on the terms of 
-contexts'. We will state the relevant judgmental equality rules for the empty
-context in this section rather than in the next, since they are required to
-state that the action on families of an operation is compatible with the
-original operation.
-\end{comment}
-
-We note upfront that both context extension and family extension can also be
-seen as operations of contexts, families and terms, even though we will not
-introduce them as such. The reason is that extension is used to define what it
-means to be a family over a family over a context (namely, a family over an
-extended context). As such, it will have a trivial action on families and terms
-which will not need to be introduced. We will come back to this point in
-\autoref{extension}.
-
-\subsection{Extension}
-\label{extension}
-The operations of extension enable one to consider families over
-families and higher level families alike. We will need both a context extension
-operation and a family extension operation.
-
-The inference rules introducing context and family extension are as follows:
-\begin{align}
-& \inference
-  { \jfam{\Gamma}{A}
-    }
-  { \jctx{\ctxext{\Gamma}{A}}
-    }
-& & \inference
-    { \jctxeq{\Gamma}{\Delta}
-      \jfameq{\Gamma}{A}{B}
-      }
-    { \jctxeq{\ctxext{\Gamma}{A}}{\ctxext{\Delta}{B}}
-      }
-    \\
-& \inference
-  { \jfam{{\Gamma}{A}}{P}
-    }
-  { \jfam{\Gamma}{\ctxext{A}{P}}
-    }
-& & \inference
-    { \jfameq{\Gamma}{A}{B} 
-      \jfameq{{\Gamma}{A}}{P}{Q}
-      }
-    { \jfameq{\Gamma}{\ctxext{A}{P}}{\ctxext{B}{Q}}
-      }
-\end{align}
-The extension operation may also be defined to act on families over families
-and terms thereof. For instance, When $\jfam{{\Gamma}{A}}{P}$ is a family,
-we would get a family $\jfam{{\Gamma}{A}}{{\Gamma}{P}}$ and when
-$\jterm{{\Gamma}{A}}{P}{f}$ is a term, we would get a term
-$\jterm{{\Gamma}{A}}{{\Gamma}{P}}{\ctxext{\Gamma}{f}}$. Then there would be
-axioms stating the judgmental equalities $\ctxext{\Gamma}{P}\jdeq P$ and
-$\ctxext{\Gamma}{f}\jdeq f$. Thus, it is of no actual use to introduce the
-action of context extension on families and terms.
-
-In \autoref{extension-on-terms} we will also define an extension operation
-on terms, which defines a term $\jterm{\Gamma}{{A}{P}}{\tmext{x}{u}}$ for
-every $\jterm{\Gamma}{A}{x}$ and $\jterm{\Gamma}{\subst{x}{P}}{u}$. This
-operation is defined by using the substitution operation and identity terms.
-
-\subsubsection{Associativity of extension}
-\label{comp-ee}
-The inference rules asserting that extension is compatible with itself assert
-that contexts are unstructured lists of type declarations. This rule is
-unavoidable if we want that for a family $A$ in context $\Gamma$, a family over
-$A$ is the same thing as a family over $\ctxext{\Gamma}{A}$. 
-\begin{align}
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{{\Gamma}{A}}{P}
-    }
-  { \jctxeq{\ctxext{{\Gamma}{A}}{P}}{\ctxext{\Gamma}{{A}{P}}}
-    }
-  \\
-& \inference
-  { \jfam{{\Gamma}{A}}{P}
-    \jfam{{{\Gamma}{A}}{P}}{Q}
-    }
-  { \jfameq{\Gamma}{\ctxext{A}{{P}{Q}}}{\ctxext{{A}{P}}{Q}}
-    }
-  \label{comp-ee-c}
-\end{align}
-Note that we may take the point of view that extension distributes over
-extensions. If we want to extend an extended family $\ctxext{P}{Q}$ over
-$\ctxext{\Gamma}{A}$ by the family $A$, the result is the extension of (1)
-the extension of $P$ by $A$ and (2) the extension of $Q$ by $A$. The latter is
-simply $Q$ since extension acts trivially at the level of families (and terms).
-
-\subsection{The empty context and the empty families}
-\label{empty}
-We introduce an empty context and an family over $\Gamma$ for every context $\Gamma$. 
-It has been suggested by some to only include empty families and not an
-empty context because an empty context is not necessary, but we do have several
-reasons to include them. Having an empty context requires also rules asserting 
-that a context is the same thing as a family over the empty context and this
-gives the categorical structure on contexts for free once one has it for 
-families. The main ingredients that will be missing from the theory once an
-empty context is avoided are weakening by a context and identity morphisms from
-a context to itself. Including these by hand also requires to formulate all the
-compatibility rules involving weakening once more for the cases of weakening by
-a context and identity terms at contexts. We prefer to state these rules once
-and only once and including an empty context helps in this respect.
-
-We also prefer to have our scheme of compatibility rules as symmetrical as
-possible. The structure of type dependency should look exactly the same in the
-default case as in any context. In that respect, having an empty family but not
-an empty context seems a bit odd. Also, some sets of compatibility rules 
-(like the rules stating that extension is compatible with the empty families)
-will become assymetrical as a result of not including an empty context. Moreover,
-we would eventually like to include a stratification of the theory by means of
-a type judgment (asserting that a family in a context $\Gamma$ is a type) and
-study closed types (i.e.~types in the empty context). It would be possible to
-provide a notion of closed types without having an empty context, but this would
-have to be formulated separately and we would have to restate all the rules for
-types (if any) for closed types all over again.
-
-One of the main uses of the empty context and the empty families is that we
-get the property that the `action on contexts' of an operation is compatible
-with its `action on families'.
-
-The empty family over a context $\Gamma$ is introduced by the following rule
-inference rule:
-\begin{align}
-& \inference
-  { }
-  { \jctx{\emptyc}
-    }
-  \\
-& \inference
-  { \jctx{\Gamma}
-    }
-  { \jfam{\Gamma}{\emptyf[\Gamma]}
-    }
-  \\
-& \inference
-  { \jctxeq{\Gamma}{\Gamma'}
-    }
-  { \jfameq{\Gamma}{\emptyf[\Gamma]}{\emptyf[\Gamma']}
-    }
-\end{align}
-
-By regarding contexts as families of contexts over the empty context, we
-enable ourselves also to speak of terms of contexts. A term of a context
-$\Gamma$ is a term of the family $\Gamma$ over the empty context. These ideas
-are captured in the following convertibility rules:
-\begin{align}
-& \inference
-  { \jctx{\Gamma}
-    }
-  { \jfam{\emptyc}{\Gamma}
-    } 
-& & \inference
-    { \jfam{\emptyc}{\Gamma}
-      }
-    { \jctx{\Gamma}
-      }
-    \\
-& \inference
-  { \jctxeq{\Gamma}{\Delta}
-    }
-  { \jfameq{\emptyc}{\Gamma}{\Delta}
-    }
-& & \inference
-    { \jfameq{\emptyc}{\Gamma}{\Delta}
-      }
-    { \jctxeq{\Gamma}{\Delta}
-      }
-\end{align}
-
-The reader may wonder whether the empty family $\emptyf$ always has a
-term. This shall follow from the rules stating the compatibility of extension
-with the empty families in \autoref{comp-0e} below and from
-identity terms (\autoref{identityterms}).
-\begin{comment}
-When contexts are viewed as statements, the empty context is the statement that
-asserts nothing. Likewise, in any context one can assert nothing and this is
-achieved by the empty family. 
-Since somebody who states nothing asserts it all right, there is a term of the empty
-family over any context. Moreover, any two terms of the empty family are always
-judgmentally equal.
-\begin{align}
-& \inference
-  { }
-  { \jctx{\emptyc}
-    }
-  \\
-& \inference
-  { \jctx{\Gamma}
-    }
-  { \jfam{\Gamma}{\emptyf[\Gamma]}
-    }
-  \\
-& \inference
-  { \jctx{\Gamma}
-    }
-  { \jterm{\Gamma}{\emptyf[\Gamma]}{\emptytm[\Gamma]}
-    }
-  \\
-& \inference
-  { \jterm{\Gamma}{\emptyf[\Gamma]}{x}
-    }
-  { \jtermeq
-      {\Gamma}
-      {\emptyf[\Gamma]}
-      {x}
-      {\emptytm[\Gamma]}
-    }
-  \label{comp-00-t}
-\end{align}
-
-The empty context and the empty families together with the operation of
-extension will provide several shortcuts for the rest
-of the theory of families and terms. This is (in part) obtained by requiring
-that a context is exactly the same thing as a family of contexts over the
-empty context and that judgmental equality of contexts is exactly the same
-as judgmental equality of families of contexts over the empty context.
-
-\label{comp-00}
-Because every context is a family of contexts over the empty context, 
-the empty context $\emptyc$ may be considered as a family of contexts over the empty
-context. When we do this, we get $\emptyf[\emptyc]$.
-\begin{equation}\label{eq:comp-00}
-\inference
-  { }
-  { \jfameq
-      {\emptyc}
-      {\emptyc}
-      {\emptyf[\emptyc]}
-    }
-\end{equation}
-In the future, we shall denote $\emptyf[\Gamma]$ by $\emptyf$. The above rule
-guarantees that this will not cause confusion. Likewise, we shall denote
-$\emptytm[\Gamma]$ by $\emptytm$.
-\end{comment}
-
-\subsubsection{Compatibility of extension with the empty context and families}
-In the following set of inference rules we state that the empty context and
-the empty family are neutral objects for both context extension (the first two
-rules) and family extension (the last two rules).
-\label{comp-e0}\label{comp-0e}
-\begin{align}
-& \inference
-  { \jctx{\Gamma}
-    }
-  { \jctxeq{\ctxext{\emptyc}{\Gamma}}{\Gamma}
-    }
-  \label{comp-0e-c}
-  \\
-& \inference
-  { \jctx{\Gamma}
-    }
-  { \jctxeq{\ctxext{\Gamma}{\emptyf}}{\Gamma}
-    }
-  \label{comp-e0-c}\\
-& \inference
-  { \jfam{\Gamma}{A}
-    }
-  { \jfameq{\Gamma}{\ctxext{\emptyf}{A}}{A}
-    }
-  \label{comp-0e-f}
-  \\
-& \inference
-  { \jfam{\Gamma}{A}
-    }
-  { \jfameq{\Gamma}{{A}{\emptyf}}{A}
-    }
-  \label{comp-e0-f}
-\end{align}
-
-\subsubsection{Family extension restricted to the empty context}
-Strictly speaking we should have used a different notation for context extension
-as for family extension, because the following rule asserting that family extension
-in the empty context is the same thing as context extension would look tautological
-without a difference. So let us denote, only for the moment, context extension
-of $\Gamma$ by $A$ by $(\ctxext{\Gamma}{A})^c$ and family extension of $A$ by
-$P$ in context $\Gamma$ by $(\ctxext{A}{P})^\famsym$.
-
-Note that we may consider a context $\Gamma$ as a family over $\emptyc$ and
-a family $\jfam{\Gamma}{A}$ as a family $\jfam{{\emptyc}{\Gamma}}{A}$. 
-Therefore we will require the following rule:
-\begin{align}
-& \inference
-  { \jfam{\Gamma}{A}
-    }
-  { \jctxeq{(\ctxext{\Gamma}{A})^c}{(\ctxext{\Gamma}{A})^\famsym}
-    }
-\end{align}
-Note that this rule actually justifies that we have not utilized two different
-notations for context extension and family extension.
-
-\subsection{The theory of weakening}
-\label{weakening}
-When $A$ is a family in context $\Gamma$, the operation of weakening by $A$
-takes a family $B$ in context $\Gamma$ and provides a family $\ctxwk{A}{B}$
-in context $\ctxext{\Gamma}{A}$. The context family $\ctxwk{A}{B}$ can be seen
-as the constant family over $\ctxext{\Gamma}{A}$ with value $B$. This idea will
-be axiomatized in the cancellation property of weakening and substitution in
-\autoref{cancellation-ws}. In \autoref{morphisms} we will take the terms 
-$\unfold{\jhom{\Gamma}{A}{B}{f}}$ to be the morphisms of families from $A$ to 
-$B$. These will be at the heart of the categorical structure of the theory.
-
-The weakening operation acts on three levels: on contexts, on families and
-on terms. The `action on contexts' of weakening is the action we described
-above: it takes a family $B$ over $\Gamma$ to a family $\ctxwk{A}{B}$ over
-$\ctxext{\Gamma}{A}$; the `action on families' of weakening takes a family
-$Q$ over $\ctxext{\Gamma}{B}$ to a family $\ctxwk[\famsym]{A}{Q}$ over
-$\ctxext{{\Gamma}{A}}{\ctxwk{A}{B}}$; the `action on terms' of weakening takes
-a term $g$ of $Q$ to a term $\ctxwk[\tmsym]{A}{g}$ of $\ctxwk[\famsym]{A}{Q}$.
-\begin{align}
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{\Gamma}{B}
-    }
-  { \jfam{{\Gamma}{A}}{\ctxwk{A}{B}}
-    }
-& & \inference
-    { \jfameq{\Gamma}{A}{A'}
-      \jfameq{\Gamma}{B}{B'}
-      }
-    { \jfameq{{\Gamma}{A}}{\ctxwk{A}{B}}{\ctxwk{A'}{B'}}
-      }
-    \\
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{{\Gamma}{B}}{Q}
-    }
-  { \jfam{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk[\famsym]{A}{Q}}
-    }
-& & \inference
-    { \jfameq{\Gamma}{A}{A'}
-      \jfameq{{\Gamma}{B}}{Q}{Q'}
-      }
-    { \jfameq
-        {{{\Gamma}{A}}{\ctxwk{A}{B}}}
-        {\ctxwk[\famsym]{A}{Q}}
-        {\ctxwk[\famsym]{A'}{Q'}}
-      }
-    \\
-%& \inference
-%  { \jterm{\Gamma}{B}{y}
-%    }
-%  { \jterm{{\Gamma}{A}}{\ctxwk{A}{B}}{\ctxwk{A}{y}}
-%    }
-%& & \inference
-%    { \jtermeq{\Gamma}{B}{y}{y'}
-%      }
-%    { \jtermeq
-%        {{\Gamma}{A}}
-%        {\ctxwk{A}{B}}
-%        {\ctxwk{A}{y}}
-%        {\ctxwk{A'}{y'}}
-%      }
-%    \\
-& \inference
-  { \jfam{\Gamma}{A}
-    \jterm{{\Gamma}{B}}{Q}{g}
-    }
-  { \jterm{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk[\famsym]{A}{Q}}{\ctxwk[\tmsym]{A}{g}}
-    }
-& & \inference
-    { \jfameq{\Gamma}{A}{A'}
-      \jtermeq{{\Gamma}{B}}{Q}{g}{g'}
-      }
-    { \jtermeq
-        {{{\Gamma}{A}}{\ctxwk{A}{B}}}
-        {\ctxwk[\famsym]{A}{Q}}
-        {\ctxwk[\tmsym]{A}{g}}
-        {\ctxwk[\tmsym]{A'}{g'}}
-      }
-\end{align}
-
-\subsubsection{Weakening of an empty family}
-The following inference rules express that when the empty family is
-weakened, the result is the empty family.
-\label{comp-w0}
-\begin{align}
-%& \inference
-%  { \jctx{\Gamma}
-%    }
-%  { \jfameq{\Gamma}{\ctxwk{\Gamma}{\emptyc}}{\emptyf}
-%    }
-%  \\
-& \inference
-  { \jfam{\Gamma}{A}
-    }
-  { \jfameq{{\Gamma}{A}}{\ctxwk{A}{\emptyf}}{\emptyf}
-    }
-  \label{comp-w0-c}\\
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{\Gamma}{B}
-    }
-  { \jfameq
-    {{{\Gamma}{A}}{\ctxwk{A}{B}}}
-    {\ctxwk[\famsym]{A}{\emptyf}}
-    {\emptyf}
-    }
-  \label{comp-w0-f}
-\end{align}
-\begin{comment}
-Note that we do not have to axiomatize the action of weakening
-by $A$ on the term of the empty family, because the result will be in the empty 
-family and every term in the empty family is judgmentally equal to the term 
-$\emptytm$.
-\end{comment}
-
-Because a family over $\Gamma$ is the same as a family over 
-$\ctxext{\Gamma}{\emptyf}$ we can apply both the action on contexts and the
-action on families of weakening to a family $B$ over $\Gamma$. When we apply
-the action on families, we obtain a family $\ctxwk[\famsym]{A}{B}$ over the
-context $\ctxext{{\Gamma}{A}}{\ctxwk{A}{\emptyf}}$. However, since we have
-postulated the judgmental equalities $\ctxwk{A}{\emptyf}\jdeq\emptyf$ and
-$\ctxext{{\Gamma}{A}}{\emptyf}\jdeq\ctxext{\Gamma}{A}$, we see that we can
-compare $\ctxwk[\famsym]{A}{B}$ with $\ctxwk{A}{B}$. The following inference
-rule postulates that these two are judgmentally equal:
-\begin{equation}
-\inference
-{ \jfam{\Gamma}{A}
-  \jfam{\Gamma}{B}
-  }
-{ \jfameq{{\Gamma}{A}}{\ctxwk[\famsym]{A}{B}}{\ctxwk{A}{B}}
-  }
-\end{equation}
-Due to this rule, the action on contexts and the action on families of weakening
-are compatible with each other and consequently there can be no possible
-confusion when we omit the annotations $\famsym$ and $\tmsym$ alltogether. In
-the future, the weakening of a family $Q$ over $\ctxext{\Gamma}{B}$ shall
-be denoted just by $\ctxwk{A}{Q}$ and likewise the weakening of a term $g$ of
-$Q$ shall be denoted by $\ctxwk{A}{g}$.
-
-Because a family $B$ over $\Gamma$ can be treated as a family by the operation
-of weakening, weakening also acts on the terms of $B$. The weakening of a term
-$y$ of $B$ by $A$ can be seen as the constant term (or function) of the
-family $\ctxwk{A}{B}$.
-
-\subsubsection{Weakenings of extensions}
-\label{comp-we}
-The following rules assert that when an extended family is weakened, the
-weakening distributes over the extension factors.
-\begin{align}
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{{\Gamma}{B}}{Q}
-    }
-  { \jfameq
-      {\ctxext{\Gamma}{A}}
-      {\ctxwk{A}{\ctxext{B}{Q}}}
-      {\ctxext{\ctxwk{A}{B}}{\ctxwk{A}{Q}}}
-    }
-  \label{comp-we-c}
-  \\
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{{{\Gamma}{B}}{Q}}{R}
-    }
-  { \jfameq
-      {\ctxext{{\Gamma}{A}}{\ctxwk{A}{B}}}
-      {\ctxwk{A}{\ctxext{Q}{R}}}
-      {\ctxext{\ctxwk{A}{Q}}{\ctxwk{A}{R}}}
-    }
-  \label{comp-we-f}
-\end{align}
-When thinking of terms of $\ctxwk{A}{B}$ as morphisms of families from $A$ to
-$B$, this looks already like form of type theoretic choice. It is weaker in that
-it is not stated with function types, yet it is stronger in that it states a
-judgmental equality between two families.
-
-There is also a version of this property where an extended term is weakened.
-This variant is stated and proved in \autoref{comp-we-t}.
-
-\subsubsection{Weakening of weakenings}
-Weakening by a family $A$ in context $\Gamma$ brings things in context $\Gamma$
-to things in context $\ctxext{\Gamma}{A}$. Since we have all the ingredients of
-the theory of contexts, families and terms in the context $\Gamma$ as well it
-has in particular it's own weakening by families. Suppose we have a family
-$\jfam{{\Gamma}{B}}{Q}$ over $B$ in context $\Gamma$. Weakening by $Q$ brings
-things from context $\ctxext{\Gamma}{B}$ to $\ctxext{{\Gamma}{B}}{Q}$. Thus
-we can provide rules asserting what will happen when we first weaken by $Q$ and
-then (via the action on families) by $A$. We will require the following
-inference rules to be valid:
-\label{comp-ww}
-\begin{align}
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{{\Gamma}{B}}{Q}
-    \jfam{{\Gamma}{B}}{R}
-    }
-  { \jfameq
-      {{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{Q}}}
-      {\ctxwk{A}{{Q}{R}}}
-      {\ctxwk{{A}{Q}}{{A}{R}}}
-    }
-  \label{comp-ww-c}\\
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{{\Gamma}{B}}{Q}
-    \jfam{{{\Gamma}{B}}{R}}{S}
-    }
-  { \jfameq
-      {{{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{Q}}}{\ctxwk{A}{{Q}{R}}}}
-      {\ctxwk{A}{{Q}{S}}}
-      {\ctxwk{{A}{Q}}{{A}{S}}}
-    }
-  \label{comp-ww-f}\\
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{{\Gamma}{B}}{Q}
-    \jterm{{{\Gamma}{B}}{R}}{S}{k}
-    }
-  { \jtermeq
-      {{{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{Q}}}{\ctxwk{A}{{Q}{R}}}}
-      {\ctxwk{A}{{Q}{S}}}
-      {\ctxwk{A}{{Q}{k}}}
-      {\ctxwk{{A}{Q}}{{A}{k}}}
-    }
-  \label{comp-ww-t}
-\end{align}
-
-\begin{rmk}
-As an important special case of these inference rules we have the following
-valid inference rules:
-\begin{align*}
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{\Gamma}{B}
-    \jfam{\Gamma}{C}
-    }
-  { \jfameq
-      {{{\Gamma}{A}}{\ctxwk{A}{B}}}
-      {\ctxwk{A}{{B}{C}}}
-      {\ctxwk{{A}{B}}{{A}{C}}}
-    }
-  \\
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{\Gamma}{B}
-    \jfam{{\Gamma}{C}}{R}
-    }
-  { \jfameq
-      {{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{{B}{C}}}}
-      {\ctxwk{A}{{B}{R}}}
-      {\ctxwk{{A}{B}}{{A}{R}}}
-    }
-  \\
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{\Gamma}{B}
-    \jterm{{\Gamma}{C}}{R}{h}
-    }
-  { \jtermeq
-      {{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{{B}{C}}}}
-      {\ctxwk{A}{{B}{R}}}
-      {\ctxwk{A}{{B}{h}}}
-      {\ctxwk{{A}{B}}{{A}{h}}}
-    }
-\end{align*}
-Moreover, we have
-\begin{equation*}
-\inference
-  { \jfam{\Gamma}{A}
-    \jfam{\Gamma}{B}
-    \jterm{\Gamma}{C}{z}
-    }
-  { \jtermeq
-      {{{\Gamma}{A}}{\ctxwk{A}{B}}}
-      {\ctxwk{A}{{B}{C}}}
-      {\ctxwk{A}{{B}{z}}}
-      {\ctxwk{{A}{B}}{{A}{z}}}
-    }
-\end{equation*}
-\end{rmk}
-
-\subsubsection{Weakening by the empty family}
-Note that we can also weaken by the empty family over $\Gamma$.
-Weakening by the empty family $\emptyf$ over a context $\Gamma$ leaves families, 
-their terms, families over those families and terms of those unchanged:
-\label{comp-0w}
-\begin{align}
-& \inference
-  { \jfam{\Gamma}{B}
-    }
-  { \jfameq{\Gamma}{\ctxwk{\emptyf}{B}}{B}
-    }
-  \label{comp-0w-c}
-  \\
-& \inference
-  { \jfam{{\Gamma}{B}}{Q}
-    }
-  { \jfameq{{\Gamma}{B}}{\ctxwk{\emptyf}{Q}}{Q}
-    }
-  \label{comp-0w-f}
-  \\
-& \inference
-  { \jterm{{\Gamma}{B}}{Q}{g}
-    }
-  { \jtermeq{{\Gamma}{B}}{Q}{\ctxwk{\emptyf}{g}}{g}
-    }
-  \label{comp-0w-t}
-\end{align}
-
-\subsubsection{Currying for weakening}
-\label{comp-ew}
-The rules expressing that extension is compatible with weakening assert that
-weakening by an extension is the same thing as weakening twice in the
-appropriate way.
-\begin{align}
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{{\Gamma}{A}}{P}
-    \jfam{\Gamma}{B}
-    }
-  { \jfameq
-      {{{\Gamma}{A}}{P}}
-      {\ctxwk{\ctxext{A}{P}}{B}}
-      {\ctxwk{P}{{A}{B}}}
-    }
-  \label{comp-ew-c}\\
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{{\Gamma}{A}}{P}
-    \jfam{{\Gamma}{B}}{Q}
-    }
-  { \jfameq
-      {{{{\Gamma}{A}}{P}}{\ctxwk{P}{{A}{B}}}}
-      {\ctxwk{\ctxext{A}{P}}{Q}}
-      {\ctxwk{P}{{A}{Q}}}
-    }
-  \label{comp-ew-f}\\
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{{\Gamma}{A}}{P}
-    \jterm{{\Gamma}{B}}{Q}{g}
-    }
-  { \jtermeq
-      {{{{\Gamma}{A}}{P}}{\ctxwk{P}{{A}{B}}}}
-      {\ctxwk{P}{{A}{Q}}}
-      {\ctxwk{\ctxext{A}{P}}{g}}
-      {\ctxwk{P}{{A}{g}}}
-    } 
-  \label{comp-ew-t}
-\end{align}
-
-\begin{comment}
-\subsubsection{Weakening by a context}
-Since we have not required that there is an empty context with the property that
-a context is the same thing as a family in the empty context, we will have to
-explicitly introduce weakening by contexts. Since it will allways be clear
-whether something is weakened by a context or by a family, we shall use the
-same notation for weakening by contexts as for weakening by families. We shall
-have to repeat all the compatibility rules for weakening by families for
-weakening by contexts; there is no way around it. It is
-actually this sort of inconvenience that has originally led us to include an
-empty context and be done with it.
-
-In the following inference rules we assume that we have a context $\Gamma$ and
-that $\jctxeq{\Gamma}{\Gamma'}$ in the rules where $\Gamma'$ makes an appearence.
-\begin{align}
-& \inference
-  { \jctx{\Delta}
-    }
-  { \jfam{\Gamma}{\ctxwk{\Gamma}{\Delta}}
-    }
-& & \inference
-  { \jctxeq{\Delta}{\Delta'}
-    }
-  { \jfameq{\Gamma}{\ctxwk{\Gamma}{\Delta}}{\ctxwk{\Gamma'}{\Delta'}}
-    }
-  \\
-& \inference
-  { \jfam{\Delta}{B}
-    }
-  { \jfam{\ctxext{\Gamma}{\ctxwk{\Gamma}{\Delta}}}{\ctxwk{\Gamma}{B}}
-    }
-& & \inference
-  { \jfameq{\Delta}{B}{B'}
-    }
-  { \jfameq
-      {\ctxext{\Gamma}{\ctxwk{\Gamma}{\Delta}}}
-      {\ctxwk{\Gamma}{B}}
-      {\ctxwk{\Gamma'}{B'}}
-    }
-  \\
-& \inference
-  { \jterm{\Delta}{B}{y}
-    }
-  { \jterm
-      {\ctxext{\Gamma}{\ctxwk{\Gamma}{\Delta}}}
-      {\ctxwk{\Gamma}{B}}
-      {\ctxwk{\Gamma}{y}}
-    }
-& & \inference
-  { \jtermeq{\Delta}{B}{y}{y'}
-    }
-  { \jtermeq
-      {\ctxext{\Gamma}{\ctxwk{\Gamma}{\Delta}}}
-      {\ctxwk{\Gamma}{B}}
-      {\ctxwk{\Gamma}{y}}
-      {\ctxwk{\Gamma}{y'}}
-    }
-\end{align}
-Note that we do not employ three different notations for the three different
-actions of weakening by a context because there can be no confusion about which
-of the three actions is meant. If there were an empty context we should have
-distinguished the three as we did for weakening by a family.
-
-We shall have similar compatibility rules for weakening by a context as we did
-for weakening by a family.
-
-{\color{red}To do. It will look exactly the same as for weakening by families}
-\end{comment}
-
 \subsection{The theory of projections}
 \label{identityterms}
+
 We formulate the theory of projections on top of the theory of weakening
@@
 \end{align}
-
-\begin{comment}
-We stress again
-that because we have not required there to be an empty context, we need to
-explicitly call for identity terms $\idtm{\Gamma}$ of $\ctxwk{\Gamma}{\Gamma}$
-in context $\Gamma$ for each context $\Gamma$. Thus, we will also require
-the inference rule
-\begin{align}
-& \inference
-  { \jctx{\Gamma}
-    }
-  { \jterm{\Gamma}{\ctxwk{\Gamma}{\Gamma}}{\idtm{\Gamma}}
-    }
-& &
-  \inference
-  { \jctxeq{\Gamma}{\Gamma'}
-    }
-  { \jtermeq{\Gamma}{\ctxwk{\Gamma}{\Gamma}}{\idtm{\Gamma}}{\idtm{\Gamma'}}
-    }
-\end{align}
-For the following inference rules, there should also be versions of identity
-terms at contexts.
-\end{comment}
 
@@
 \end{rmk}
-
-\subsection{The theory of substitution}
-\label{substitution}
-The theory of substitution can be formulated separately from the theory of
-weakening. Therefore, we will only require the theory of extension and the
-empty context and families. In \autoref{sec:esystem-equalities} we will
-provide the additional judgmental equality rules that will be required when
-combining the theories of substitution and of projections.
-
-Given a family $P$ over $A$ and a term $x$ of $A$, substitution gives a way to
-consider the \emph{fiber $\subst{x}{P}$ of $P$ at $x$}. As was the case with
-weakening, the substitution operation comes in three parts: the `action on
-contexts' of substitution is the operation just described; the `action on
-families' of substitution takes a family $Q$ over $\ctxext{{\Gamma}{A}}{P}$
-to a family $\subst[\famsym]{x}{Q}$ over $\ctxext{\Gamma}{\subst{x}{P}}$; the
-`action on terms' of substitution takes a term $g$ of $Q$ over
-$\ctxext{{\Gamma}{A}}{P}$ to a term $\subst[\tmsym]{x}{g}$ of 
-$\subst[\famsym]{x}{Q}$.
-\begin{align}
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jfam{{\Gamma}{A}}{P}
-    }
-  { \jfam{\Gamma}{\subst{x}{P}}
-    }
-& & \inference
-    { \jtermeq{\Gamma}{A}{x}{x'}
-      \jfameq{{\Gamma}{A}}{P}{P'}
-      }
-    { \jfameq{\Gamma}{\subst{x}{P}}{\subst{x'}{P'}}
-      }
-    \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jfam{{{\Gamma}{A}}{P}}{Q}
-    }
-  { \jfam{{\Gamma}{\subst{x}{P}}}{\subst[\famsym]{x}{Q}}
-    }
-& & \inference
-    { \jtermeq{\Gamma}{A}{x}{x'}
-      \jfameq{{{\Gamma}{A}}{P}}{Q}{Q'}
-      }
-    { \jfameq{{\Gamma}{\subst{x}{P}}}{\subst[\famsym]{x}{Q}}{\subst[\famsym]{x'}{Q'}}
-      }
-    \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jterm{{{\Gamma}{A}}{P}}{Q}{g}
-    }
-  { \jterm{{\Gamma}{\subst{x}{P}}}{\subst[\famsym]{x}{Q}}{\subst[\tmsym]{x}{g}}
-    }
-& & \inference
-    { \jtermeq{\Gamma}{A}{x}{x'}
-      \jtermeq{{{\Gamma}{A}}{P}}{Q}{g}{g'}
-      }
-    { \jtermeq
-        {{\Gamma}{\subst{x}{P}}}
-        {\subst[\famsym]{x}{Q}}
-        {\subst[\tmsym]{x}{g}}
-        {\subst[\tmsym]{x'}{g'}}
-      }
-\end{align}
-
-\subsubsection{Fibers of an empty family}
-The following inference rules establish that the fibers of the empty family are 
-the empty families:
-\label{comp-s0}
-\begin{align}
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    }
-  { \jfameq{\Gamma}{\subst{x}{\emptyf}}{\emptyf}
-    }
-  \label{comp-s0-c}
-  \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jfam{{\Gamma}{A}}{P}
-    }
-  { \jfameq
-      {{\Gamma}{\subst{x}{P}}}
-      {\subst{x}{\emptyf}}
-      {\emptyf}
-    }
-  \label{comp-s0-f}
-\end{align}
-\begin{comment}
-Note again that we do not have to axiomatize the action of substitution by $x$ on the term
-of the empty family, because the result will be in the empty family and every
-term in the empty family is judgmentally equal to the term $\emptytm$.
-\end{comment}
-
-We use the above rule to state the compatibility of the action on families of
-substitution with the action on contexts of substitution. Note that a family
-$P$ over $\ctxext{\Gamma}{A}$ may be regarded as a family over
-$\ctxext{{\Gamma}{A}}{\emptyf}$. Thus, we may consider the family
-$\subst[\famsym]{x}{P}$ over $\ctxext{\Gamma}{\subst{x}{\emptyf}}$. Since
-$\subst{x}{\emptyf}$ is judgmentally equal to the empty family, we may compare
-$\subst[\famsym]{x}{P}$ with $\subst{x}{P}$:
-\begin{equation}
-\inference
-{ \jfam{{\Gamma}{A}}{P}
-  }
-{ \jfameq{\Gamma}{\subst[\famsym]{x}{P}}{\subst{x}{P}}
-  }
-\end{equation}
-Due to this rule we need not make the annotations $\famsym$ and $\tmsym$ in
-the notation for substitution anymore and thus we shall omit them from now on.
-Note that because a family $P$ over $\ctxext{\Gamma}{A}$ is eligible for
-application of the action on families of substitution, we may also substitute
-terms of $P$. Thus, given terms $\jterm{\Gamma}{A}{x}$ and $\jterm{{\Gamma}{A}}{P}{f}$,
-we get a term $\jterm{\Gamma}{\subst{x}{P}}{\subst{x}{f}}$, the \emph{value of
-$f$ at $x$}.
-
-In \autoref{morphisms} we will use a combination of weakening and substitution
-to define composition of morphisms of families. However, we have to rely
-on the cancellation rule stated in \autoref{cancellation-ws} before we can
-meaningfully state the definition of composition.
-
-\subsubsection{Fibers of extensions}
-\label{comp-se}
-The following inference rules assert that if we take the fiber of an extended
-family at a term $x$ of $A$ in context $\Gamma$, the substitution by $x$
-distributes over the factors of the extension.
-\begin{align}
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jfam{{{\Gamma}{A}}{P}}{Q}
-    }
-  { \jfameq
-      {\Gamma}
-      {\subst{x}{\ctxext{P}{Q}}}
-      {\ctxext{\subst{x}{P}}{\subst{x}{Q}}}
-    }
-  \label{comp-se-c}
-  \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jfam{{{{\Gamma}{A}}{P}}{Q}}{R}
-    }
-  { \jfameq
-      {{\Gamma}{\subst{x}{P}}}
-      {\subst{x}{\ctxext{Q}{R}}}
-      {\ctxext{\subst{x}{Q}}{\subst{x}{R}}}
-    }
-  \label{comp-se-f}
-\end{align}
-There is also a version of this statement in which extended terms are considered.
-This variant is stated and proved in \autoref{comp-se-t}.
-
-\subsubsection{Fibers of fibers}
-\label{comp-ss}
-
-The following rules explain what happens when we first substitute by a term
-$g$ of $Q$ over $P$ in context $\ctxext{\Gamma}{A}$ and then by a term $x$ of
-$A$:
-\begin{align}
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jterm{{{\Gamma}{A}}{P}}{Q}{g}
-    \jfam{{{{\Gamma}{A}}{P}}{Q}}{R}
-    }
-  { \jfameq
-      {{{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}}
-      {\subst{x}{{g}{R}}}
-      {\subst{{x}{g}}{{x}{R}}}
-    }
-  \label{comp-ss-c}
-  \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jterm{{{\Gamma}{A}}{P}}{Q}{g}
-    \jfam{{{{{\Gamma}{A}}{P}}{Q}}{R}}{S}
-    }
-  { \jfameq
-      {{{{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}}{\subst{x}{{g}{R}}}}
-      {\subst{x}{{g}{S}}}
-      {\subst{{x}{g}}{{x}{S}}}
-    }
-  \label{comp-ss-f}
-  \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jterm{{{\Gamma}{A}}{P}}{Q}{g}
-    \jterm{{{{{\Gamma}{A}}{P}}{Q}}{R}}{S}{k}
-    }
-  { \jtermeq
-      {{{{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}}{\subst{x}{{g}{R}}}}
-      {\subst{x}{{g}{S}}}
-      {\subst{x}{{g}{k}}}
-      {\subst{{x}{g}}{{x}{k}}}
-    }
-  \label{comp-ss-t}
-\end{align}
-
-\begin{rmk}
-As an important special case of these inference rules we have the following
-valid inference rules:
-\begin{align*}
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jterm{{\Gamma}{A}}{P}{f}
-    \jfam{{{\Gamma}{A}}{P}}{Q}
-    }
-  { \jfameq
-      {\Gamma}
-      {\subst{x}{{f}{Q}}}
-      {\subst{{x}{f}}{{x}{Q}}}
-    }
-  \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jterm{{\Gamma}{A}}{P}{f}
-    \jfam{{{{\Gamma}{A}}{P}}{Q}}{R}
-    }
-  { \jfameq
-      {{\Gamma}{\subst{x}{{f}{Q}}}}
-      {\subst{x}{{f}{R}}}
-      {\subst{{x}{f}}{{x}{R}}}
-    }
-  \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jterm{{\Gamma}{A}}{P}{f}
-    \jterm{{{{\Gamma}{A}}{P}}{Q}}{R}{h}
-    }
-  { \jtermeq
-      {{\Gamma}{\subst{x}{{f}{Q}}}}
-      {\subst{x}{{f}{R}}}
-      {\subst{x}{{f}{h}}}
-      {\subst{{x}{f}}{{x}{h}}}
-    }
-\end{align*}
-Moreover, we get
-\begin{equation*}
-\inference
-  { \jterm{\Gamma}{A}{x}
-    \jterm{{\Gamma}{A}}{P}{f}
-    \jterm{{{\Gamma}{A}}{P}}{Q}{g}
-    }
-  { \jtermeq
-      {\Gamma}
-      {\subst{x}{{f}{Q}}}
-      {\subst{x}{{f}{g}}}
-      {\subst{{x}{f}}{{x}{g}}}
-    }
-\end{equation*}
-\end{rmk}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Joining the theory of projections with the theory of substitution}
-\label{sec:esystem-equalities}
-
-\begin{comment}
-We have already seen that empty families, extension, weakening, substitution
-and the identity terms are compatible with each other. There are, however, a
-few more judgmental equalities explaining what happens when
-\begin{enumerate}
-\item something is weakend by the empty family. This is used to consider a term
-      $x$ of $A$ in context $\Gamma$ as a morphism from $\emptyf$ to $A$ in
-      context $\Gamma$;
-\item something is weakened by an extended family. This is Currying for weakening;
-\item something is first weakened by a family $A$ in context $\Gamma$ and then
-      substituted by a term of $A$. This is used to show that composition is
-      compatible with itself (i.e. that composition is associative);
-\item the identity terms are joining the party. Identity terms are empty contexts
-      in certain models, so they should behave accordingly.
-\end{enumerate}
-In other words, the rules seem to not fit in the scheme that all operations are
-compatible with each other, but in fact each of these rules plays a role in
-completing the scheme (work out further).
-\end{comment}
-
-\begin{comment}
-In this section we extend the judgmental equality relation in such a way that
-the empty context, extension, weakening, substitution and the identity terms
-all become compatible with each other. Since we have five operations, there
-are going to be $25$ finite sets of rules explaining how each operation is
-compatible with each of the operations.
-
-In addition to these $25$ finite sets
-of compatibility rules, there are also two sets of rules explaining the
-cancellation properties of weakening and substitution and of the identity terms.
-
-A typical compatibility axiom captures the idea that a diagram like
-\begin{equation*}
-\begin{tikzcd}[column sep=huge]
-\jfam{\Gamma}{\blank} 
-  \ar{r}{C\mapsto\ctxwk{B}{C}} 
-  \ar{d}[swap]{C\mapsto\ctxwk{A}{C}} 
-& \jfam{{\Gamma}{B}}{\blank} 
-  \ar{d}{Q\mapsto\ctxwk{A}{Q}}
-  \\
-\jfam{{\Gamma}{A}}{\blank} 
-  \ar{r}[swap]{P\mapsto\ctxwk{{A}{B}}{P}} 
-& \jfam{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\blank}
-\end{tikzcd}
-\end{equation*}
-in the case of two consecutive weakenings, or
-\begin{equation*}
-\begin{tikzcd}[column sep=huge]
-\jfam{{\Gamma}{A}}{\blank} 
-  \ar{d}[swap]{Q\mapsto\subst{x}{Q}} 
-  \ar{r}{Q\mapsto\ctxwk{P}{Q}} 
-& \jfam{{{\Gamma}{A}}{P}}{\blank} 
-    \ar{d}{R\mapsto\subst{x}{R}}
-  \\ 
-\jfam{\Gamma}{\blank} 
-  \ar{r}[swap]{B\mapsto\ctxwk{\subst{x}{P}}{B}} 
-& \jfam{{\Gamma}{\subst{x}{P}}}{\blank}
-\end{tikzcd}
-\end{equation*}
-in the case of a combination of weakening and substitution, commutes. We won't
-actually draw any of these diagrams in this section, but readers are encouraged
-to do so if they think it can help them understanding the axioms. Another thing
-one could keep in mind with the axioms that involve weakening is that all of
-them become completely obvious when weakening is made notationally invisible.
-Of course, we have an explicit notation for weakening on purpuse, but it is the
-usual practice of type theorists to make weakening invisible.
-
-The table below indicates where
-we can find the axioms of each of the combinations of operators. The reference
-at the row of operation $X$ and the column of operation $Y$ refers to the
-axiom which explains $Y$ distributes over the arguments of $X$ when $Y$ is
-applied after $X$. The empty context has no arguments, whereas extension,
-and weakening have two arguments which are both of the family-kind, substitution
-has two arguments of which one is a term and the other is a family and
-idenitity terms have one argument, a family.
-\begin{center}
-\begin{tabular}{r|rrrr}
-& ext & wk & subst & id\\
-\hline
-extension & \autoref{comp-ee}* &  & \cmark & \cmark \\
-weakening & \autoref{comp-we}* & \autoref{comp-ww} & \autoref{comp-sw} & \autoref{comp-wi} \\
-substitution & \autoref{comp-se}* & \autoref{comp-ws} & \autoref{comp-ss} & \autoref{comp-si} \\
-identity terms & \cmark & \autoref{comp-wi} & \autoref{comp-si} & \xmark
-\end{tabular}
-\end{center}
-Apart from references to sections where axioms are given, we also see checkmarks,
-crosses, marked references and duplicated references in this table. 
-The checkmarks indicate that the desired property
-can be proved and need not be axiomatized. To clarify those:
-\begin{itemize}
-\item The identity term of the empty context is the unique term of the empty
-family. Recall that we already have axiomatized in \autoref{empty} that every
-term of the empty family is judgmentally equal to $\emptytm$.
-\item The identity term of an extension can be explained by means of the notion
-of \emph{extension on terms}. This notion is defined in \autoref{extension-on-terms}
-using only identity terms and substitution. With extension on terms we can
-combine a term $\jterm{\Gamma}{A}{x}$ and a term $\jterm{\Gamma}{\subst{x}{P}}{u}$
-into a term $\jterm{\Gamma}{{A}{P}}{\tmext{x}{u}}$, which may be seen as the
-pair consisting of $x$ and $u$. Since it is a defined notion, we will
-be able to derive lots of properties for extension of terms. One of them explains
-how we can see the identity term of an extension as the extension of two certain
-terms, see \autoref{comp-ie}.
-\item Using extension on terms we can also make sense of `fibers at an extension'
-because that would exactly be the fiber at an extended term. We will see in
-\autoref{comp-es} that, for instance, $\subst{\tmext{x}{u}}{Q}\jdeq\subst{u}{{x}{Q}}$
-for any family $\jfam{{{\Gamma}{A}}{P}}{Q}$ and any two terms $\jterm{\Gamma}{A}{x}$
-and $\jterm{\Gamma}{\subst{x}{P}}{u}$. That's a form of currying.
-\end{itemize}
-The spots with a \xmark simply make no sense 
-{\color{red}(or do they, when the term model
-is better understood (by me)?)}. The marked references point to
-axioms which \emph{do} have a variant for terms, but those require the notion
-of extension on terms, can be proved and therefore need not be included in our
-list of compatibility axioms. In each of these sections, we give a forward
-reference to the lemma in which this variant is stated and proved.
-\end{comment}
-
-\subsubsection{Fibers of weakenings}\label{comp-sw}
-The following rules assert what happens when we first weaken by a family
-$Q$ over $P$ in context $\ctxext{\Gamma}{A}$ and then substitute by a term
-$x$ of $A$:
-\begin{align}
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jfam{{{\Gamma}{A}}{P}}{Q}
-    \jfam{{{\Gamma}{A}}{P}}{R}
-    }
-  { \jfameq
-      {{{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}}
-      {\subst{x}{\ctxwk{Q}{R}}}
-      {\ctxwk{\subst{x}{Q}}{\subst{x}{R}}}
-    }
-  \label{comp-sw-c}
-  \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jfam{{{\Gamma}{A}}{P}}{Q}
-    \jfam{{{{\Gamma}{A}}{P}}{R}}{S}
-    }
-  { \jfameq
-      {{{{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}}{\subst{x}{\ctxwk{Q}{R}}}}
-      {\subst{x}{\ctxwk{Q}{S}}}
-      {\ctxwk{\subst{x}{Q}}{\subst{x}{S}}}
-    }
-  \label{comp-sw-f}
-  \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jfam{{{\Gamma}{A}}{P}}{Q}
-    \jterm{{{{\Gamma}{A}}{P}}{R}}{S}{k}
-    }
-  { \jtermeq
-      {{{{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}}{\subst{x}{\ctxwk{Q}{R}}}}
-      {\subst{x}{\ctxwk{Q}{S}}}
-      {\subst{x}{\ctxwk{Q}{k}}}
-      {\ctxwk{\subst{x}{Q}}{\subst{x}{k}}}
-    }
-  \label{comp-sw-t}
-\end{align}
-
-\begin{rmk}
-As an important special case of these inference rules we have the following
-valid inference rules:
-\begin{align*}
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jfam{{\Gamma}{A}}{P}
-    \jfam{{\Gamma}{A}}{Q}
-    }
-  { \jfameq
-      {{\Gamma}{\subst{x}{P}}}
-      {\subst{x}{\ctxwk{P}{Q}}}
-      {\ctxwk{\subst{x}{P}}{\subst{x}{Q}}}
-    }
-  \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jfam{{\Gamma}{A}}{P}
-    \jfam{{{\Gamma}{A}}{Q}}{R}
-    }
-  { \jfameq
-      {{{\Gamma}{\subst{x}{P}}}{\subst{x}{\ctxwk{P}{Q}}}}
-      {\subst{x}{\ctxwk{P}{R}}}
-      {\ctxwk{\subst{x}{P}}{\subst{x}{R}}}
-    }
-  \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jfam{{\Gamma}{A}}{P}
-    \jterm{{{\Gamma}{A}}{Q}}{R}{h}
-    }
-  { \jtermeq
-      {{{\Gamma}{\subst{x}{P}}}{\subst{x}{\ctxwk{P}{Q}}}}
-      {\subst{x}{\ctxwk{P}{R}}}
-      {\subst{x}{\ctxwk{P}{h}}}
-      {\ctxwk{\subst{x}{P}}{\subst{x}{h}}}
-    }
-\end{align*}
-Moreover, we get
-\begin{equation*}
-\inference
-  { \jterm{\Gamma}{A}{x}
-    \jfam{{\Gamma}{A}}{P}
-    \jterm{{\Gamma}{A}}{Q}{g}
-    }
-  { \jtermeq
-      {{\Gamma}{\subst{x}{P}}}
-      {\subst{x}{\ctxwk{P}{Q}}}
-      {\subst{x}{\ctxwk{P}{g}}}
-      {\ctxwk{\subst{x}{P}}{\subst{x}{g}}}
-    }
-\end{equation*}
-\end{rmk}
-
-\subsubsection{Weakenings of fibers}
-\label{comp-ws}
-The following inference rules explain what happens when we first weaken by a
-term $g$ of $Q$ in context $\ctxext{\Gamma}{B}$ and then weaken by a family
-$A$ over $\Gamma$.
-\begin{align}
-& \inference
-  { \jfam{\Gamma}{A}
-    \jterm{{\Gamma}{B}}{Q}{g}
-    \jfam{{{\Gamma}{B}}{Q}}{R}
-    }
-  { \jfameq
-      {{{\Gamma}{A}}{\ctxwk{A}{B}}}
-      {\ctxwk{A}{\subst{g}{R}}}
-      {\subst{\ctxwk{A}{g}}{\ctxwk{A}{R}}}
-    }
-  \label{comp-ws-c}
-  \\
-& \inference
-  { \jfam{\Gamma}{A}
-    \jterm{{\Gamma}{B}}{Q}{g}
-    \jfam{{{{\Gamma}{B}}{Q}}{R}}{S}
-    }
-  { \jfameq
-      {{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{\subst{g}{R}}}}
-      {\ctxwk{A}{\subst{g}{S}}}
-      {\subst{\ctxwk{A}{g}}{\ctxwk{A}{S}}}
-    }
-  \label{comp-ws-f}
-  \\
-& \inference
-  { \jfam{\Gamma}{A}
-    \jterm{{\Gamma}{B}}{Q}{g}
-    \jterm{{{{\Gamma}{B}}{Q}}{R}}{S}{k}
-    }
-  { \jtermeq
-      {{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{\subst{g}{R}}}}
-      {\ctxwk{A}{\subst{g}{S}}}
-      {\ctxwk{A}{\subst{g}{k}}}
-      {\subst{\ctxwk{A}{g}}{\ctxwk{A}{k}}}
-    }
-  \label{comp-ws-t}
-\end{align}
-
-\begin{rmk}
-As an important special case of these inference rules we have the following 
-valid inference rules:
-\begin{align*}
-& \inference
-  { \jfam{\Gamma}{A}
-    \jterm{\Gamma}{B}{y}
-    \jfam{{\Gamma}{B}}{Q}
-    }
-  { \jfameq
-      {{\Gamma}{A}}
-      {\ctxwk{A}{\subst{y}{Q}}}
-      {\subst{\ctxwk{A}{y}}{\ctxwk{A}{Q}}}
-    }
-  \\
-& \inference
-  { \jfam{\Gamma}{A}
-    \jterm{\Gamma}{B}{y}
-    \jfam{{{\Gamma}{B}}{Q}}{R}
-    }
-  { \jfameq
-      {{{\Gamma}{A}}{\ctxwk{A}{\subst{y}{Q}}}}
-      {\ctxwk{A}{\subst{y}{R}}}
-      {\subst{\ctxwk{A}{y}}{\ctxwk{A}{R}}}
-    }
-  \\
-& \inference
-  { \jfam{\Gamma}{A}
-    \jterm{\Gamma}{B}{y}
-    \jterm{{{\Gamma}{B}}{Q}}{R}{h}
-    }
-  { \jtermeq
-      {{{\Gamma}{A}}{\ctxwk{A}{\subst{y}{Q}}}}
-      {\ctxwk{A}{\subst{y}{R}}}
-      {\ctxwk{A}{\subst{y}{h}}}
-      {\subst{\ctxwk{A}{y}}{\ctxwk{A}{h}}}
-    }
-\end{align*}
-Moreover, we get
-\begin{equation*}
-\inference
-  { \jfam{\Gamma}{A}
-    \jterm{\Gamma}{B}{y}
-    \jterm{{\Gamma}{B}}{Q}{g}
-    }
-  { \jtermeq
-      {{\Gamma}{A}}
-      {\ctxwk{A}{\subst{y}{Q}}}
-      {\ctxwk{A}{\subst{y}{g}}}
-      {\subst{\ctxwk{A}{y}}{\ctxwk{A}{g}}}
-    }
-\end{equation*}
-\end{rmk}
-
-\subsubsection{The cancellation property of weakening and substitution}
-\label{cancellation-ws}
-The judgmental equalities we're about to describe assert that substituting a term
-in the weakening a thing gives you the thing back. In the case of contexts we get that each fiber
-$\subst{x}{\ctxwk{A}{B}}$ is just $B$ and in the case of terms we get 
-that $\ctxwk{A}{y}$ is the constant function
-mapping everything to $y:B$. Thus, these rules actually establish the weakening
-as the weakening. After stating the rules we will describe what it means to
-compose context morphisms (terms of weakened contexts).
-\begin{align}
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{\Gamma}{B}
-    \jterm{\Gamma}{A}{x}
-    }
-  { \jfameq
-      {\Gamma}
-      {\subst{x}{\ctxwk{A}{B}}}
-      {B}
-    }
-  \label{cancellation-ws-c}\\
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{{\Gamma}{B}}{Q}
-    \jterm{\Gamma}{A}{x}
-    }
-  { \jfameq{{\Gamma}{B}}{\subst{x}{\ctxwk{A}{Q}}}{Q}
-    }
-  \label{cancellation-ws-f}\\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jterm{{\Gamma}{B}}{Q}{g}
-    }
-  { \jtermeq{{\Gamma}{B}}{Q}{\subst{x}{\ctxwk{A}{g}}}{g}
-    }
-  \label{cancellation-ws-t}
-\end{align}
-
-\subsubsection{The identity term of a substituted family}
-\label{comp-si}
-The identity term of a substituted family is the substitution of the identity term
-\begin{equation}
-\inference
-  { \jterm{\Gamma}{A}{x}
-    \jfam{{{\Gamma}{A}}{P}}{Q}
-    }
-  { \unfoldall{\jhomeq
-      {{\Gamma}{\subst{x}{P}}}
-      {\subst{x}{Q}}
-      {\subst{x}{Q}}
-      {\subst{x}{\idtm{Q}}}
-      {\idtm{\subst{x}{Q}}}
-    }}
-  \label{comp-si-t}
-\end{equation}
-
-\begin{rmk}
-An important special case is the judgmental equality
-\begin{equation*}
-\unfoldall{\jhomeq
-      {\Gamma}
-      {\subst{x}{P}}
-      {\subst{x}{P}}
-      {\subst{x}{\idtm{P}}}
-      {\idtm{\subst{x}{P}}}
-    }
-\end{equation*}
-for a family $\jfam{{\Gamma}{A}}{P}$.
-\end{rmk}
-
-\subsubsection{The cancellation properties of identity terms}
-\label{cancellation-i}
-Identity terms are determined by their behavior with respect to substitution combined with
-weakening. The identity terms will also be subject to compatibility rules.
-\begin{align}
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    }
-  { \jtermeq{\Gamma}{A}{\subst{x}{\idtm{A}}}{x}
-    }
-  \label{cancellation-si}\\
-& \inference
-  { \jfam{{\Gamma}{A}}{P}
-    }
-  { \jfameq{{\Gamma}{A}}{\subst{\idtm{A}}{\ctxwk{A}{P}}}{P}
-    }
-  \label{precomp-idtm-c}
-  \\
-& \inference
-  { \jfam{{{\Gamma}{A}}{P}}{Q}
-    }
-  { \jfameq{{{\Gamma}{A}}{P}}{\subst{\idtm{A}}{\ctxwk{A}{Q}}}{Q}
-    }
-  \label{precomp-idtm-f}
-  \\
-& \inference
-  { \jterm{{{\Gamma}{A}}{P}}{Q}{g}
-    }
-  { \jtermeq
-      {{{\Gamma}{A}}{P}}
-      {Q}
-      {\subst{\idtm{A}}{\ctxwk{A}{g}}}
-      {g}
-    }
-  \label{precomp-idtm-t}
-\end{align}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{The possiblity of types in the theory of contexts, families and
-terms}
-\label{types}
-We have deliberately not spoken of types so far because we have taken the point
-of view that a type in a context is nothing but a family in that context which
-belongs to the class of types. We think of types as \emph{irreducible} families,
-i.e.\ families which are neither empty nor the extension of two
-families which are both not empty (in algebraic terminology: which
-are both non-trivial). To allow ourselves to speak of types we introduce two
-new judgments: the judgment that something is a type and the judgment that two
-types are equal.
-\begin{align*}
-\jalign\jtype{\Gamma}{A} 
-& \jalign\jtypeeq{\Gamma}{A}{B}
-\end{align*}
-But only families of contexts are eligible to be types. If $A$ is a type
-in context $\Gamma$, then $A$ is also a family of contexts over $\Gamma$. 
-Moreover, two types in context $\Gamma$ are judgmentally equal precisely when they are equal
-as context families and if a family $B$ of contexts over $\Gamma$ is
-judgmentally equal to a type $A$ in context $\Gamma$, then $B$ is a type in
-context $\Gamma$. This is expressed by the following four inference rules:
-\begin{align*}
-& \inference
-  { \jtype{\Gamma}{A}
-    }
-  { \jfam{\Gamma}{A}
-    }
-& & \inference
-    { \jtypeeq{\Gamma}{A}{B}
-      }
-    { \jfameq{\Gamma}{A}{B}
-      }
-    \\
-& \inference
-  { \jtype{\Gamma}{A}
-    \jfameq{\Gamma}{A}{B}
-    }
-  { \jtype{\Gamma}{B}
-    }
-& & \inference
-    { \jtype{\Gamma}{A}
-      \jfameq{\Gamma}{A}{B}
-      }
-    { \jtypeeq{\Gamma}{A}{B}
-      }
-\end{align*}
-As pointed out at the beginning of this subsection, 
-we do not assume that the empty family is a type, that would be like
-assuming that the multiplicative unit of a ring is prime.
-
-\begin{comment}
-We envision a
-categorical interpretation of the theory of contexts, families and terms where 
-contexts are interpreted as finite paths of arrows belonging to a predetermined
-class. The empty family would be such a path of length $0$ whereas the types
-would be such paths of length $1$.
-\end{comment}
-
-We add rules asserting that a weakened type is again a type that 
-substitution preserves the property of being a type:
-\begin{align}
-& \inference
-  { \jfam{\Gamma}{B}
-    \jtype{\ctxext{\Gamma}{B}}{Q}
-    }
-  { \jtype{\ctxext{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{Q}}
-    }
-  \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jtype{\ctxext{{\Gamma}{A}}{P}}{Q}
-    }
-  { \jtype{\ctxext{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}
-    }
-\end{align}
-With only the current rules, the possibility of making the judgment that
-something is a type does not add much to the theory of contexts, families and
-terms. Nevertheless, when studying models, having an interpretation for the
-judgment that something is a type will allow for the possibility to study
-conditions such as the one asserting that every family `factorizes' uniquely
-as multiple applications of extension to types, analogous to the condition on
-unique factorization domains in ring theory. All the tradional models of type
-theory should translate to models with such a condition, simply because
-traditionally contexts are viewed as lists of variable (and type) declarations.

### 2014-10-28 Rules/rules_setup.tex
reorganization

--- 2014-09-20 article-models-tt.tex
@@
-\section{The generalized algebraic theory of contexts, families and terms (the E-system)}\label{tt}
-In this section we give a description of dependent type theory before type
-constructors. Apart from contexts, families and terms -- which provide for the
-core of the language of dependent logic -- the basic ingredients
-of this theory will be the operations of extension on context and on families, 
-and empty context and empty families, weakening, substitution
-and the identity terms. The resulting theory can be seen as a manifestation of the 
-structure underlying dependent type theory.
+\subsection{Judgments and valid inference rules}
+\label{judgments}
 
-We will
-formulate the theory of contexts, families and terms in such a way that contexts aren't defined
-to be lists of variable declarations. The variable-free (a.k.a.~\emph{name-free}) approach 
-we take here is rather different than those appearing in 
-\cite{hofmann1995extensional,TheBook} but it has appeared in the work of Coquand
-and in \cite{Dybjer1996}.
-The main reason we don't let variable declarations in is that we don't see them 
-in the internal models either. This way we also set out to a more algebraic 
-approach of type theory and higher category theory. Thirdly, we will not have to
-be burdened with superficial comments about variables being bounded or not, or 
-fresh or free or not occuring at all.
-
-In the current presentation, contexts can be seen as binary
-planar trees of which the leaves are (families of) contexts. 
-The judgmental equality relation on contexts is an equivalence relation which 
-expresses that binary planar
-trees of contexts are judgmentally equal if their leaves are, taking only
-(the isomorphism class of) 
-the order of the leaves into account \emph{and not the actual shape of the three}.
-The intuition behind this equivalence relation is indeed that unstructured
-(i.e.~unbracketed) lists such as the lists of variable declarations which
-usually appear in type theoretical syntax, may be regarded as contexts.
-
-Besides contexts, families and terms there will also be a notion of `type in
-a context', which one may assume.
-A context is eligible to be a \emph{closed type} and over a context, a family
-may be a type. We have the intuition that `being a type'
-expresses the property of `being atomic or irreducible'. With a notion of types,
-one has the means to require a stratification by which every context is a finite
-extension of types starting at a closed type, and every family is likewise a
-finite extension of types. The idea behind this is to recover the theory of
-\cite{Garner2014} {\color{red}(and possibly B-systems?)}. We stress that
-we will not include axioms asserting such a stratification in the E-system
-and we leave discussion on the possibility of a typing judgment until the
-end of this section.
-
-With families of contexts being the principal things of
-study, we have added a bit more generality to the theory, compared to the
-theory of categories with families presented in \cite{Dybjer1996}. Here are
-a couple of reasons for this extra generality:
-\begin{enumerate}
-\item We want to be very sure that we can interpret all the rules of type theory
-in a straightforward manner in any context.
-\item We do want do end up with a categorical structure on the contexts, derived
-from the theory by requiring that a morphism from $\Gamma$ to $\Delta$ is a
-term of the weakening of $\Delta$ by $\Gamma$ in context $\Gamma$. Thus, we
-must be able to lift the context $\Delta$ to become a family of contexts over
-$\Gamma$ and it is not reasonable to expect that this results in a type.
-\item It is very easy in this setting to assert that an operation has an action
-on very high levels of families, i.e.~when $F$ is an operation taking things in
-context $\Gamma$ to things in context $\Delta$, then it takes a family $A$ in
-context $\Gamma$ to $F(A)$ in context $\Delta$; a family $P$ in context
-$\ctxext{\Gamma}{A}$ to a family $F(P)$ in context $\ctxext{\Delta}{(F(A))}$,
-and so on. The reason it becomes simple is that we don't have to explain what
-`and so on' means.
-\end{enumerate}
-
-We will formulate fairly strict rules governing the judgmental equalities,
-expressing that extension, weakening and substitution are combatible with
-each other in a judgmental manner. This does not, however, diminish the role
-of isomorphisms or of homotopies could play in the theory once identity types
-are added. Indeed, types could still have non-trivial identity relations and
-the category of types in a certain context could genuinely display higher
-categorical structure, or so we conjecture.
-
-Much of the rules we state are just compatibility rules of extension, weakening
-and substitution with each other. In a way, these rules assert that our contexts
-are just structureless lists of contexts and that likewise terms are structureless
-lists of terms. They are structureless in the sense that the order in which
-they are formed by pairing up is irrelevant. We note that this causes complications
-in the traditional way that categorical sematics of type theory is implemented,
-where contexts become objects of the category which is supposed to model type
-theory. The reason for this is that context extension will not satisfy all the
-compatibility rules we're about to state. The first step to resolving this is taking
-the types in the empty context as the objects.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Judgments and valid inference rules}\label{judgments}
 The theory we describe here is a theory of contexts, families of
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{The basic rules for judgmental equality}
 The rules for judgmental equality establish that it is an equivalence relation
@@
 \end{align*}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{The operations of the theory of contexts, families and terms}
-In this subsection we introduce the notions of
-extension on context and on families, the empty context and the empty families, 
-weakening, 
-substitution and identity terms.
-
-Weakening and substitution are both
-operations of the theory of contexts, families and terms and therefore they will
-have three levels of actions: on contexts (at the appropriate level), families
-and terms.
-
-\begin{comment}
-Generally, 
-each operation should
-have an action on contexts, on families and on terms at the `appropriate level'
-and should be compatible with each of the operations we introduce here. In
-that will become more practical after the theory of contexts, families and terms
-is formulated, this means that when we will write down a type theory that
-interprets itself in \autoref{ttmodels}, each of the notions of
-extension, empty family, weakening, substitution and identity terms are going to be internalized
-as terms of the theory and all terms are going to be compatible with these.
-That vague idea will be our guiding light.
-
-So in principle, the introductions of each of the empty context, extension,
-weakening, substitution and identity terms should come in three parts. In 
-practice, they do not all come in three parts.
-
-Apart from the empty context, there will indeed be an empty family and a term
-of the empty family (which will be deemed unique). But extension will
-not have an explicit action `at the level of families' since extension
-is used to define higher families and the action would have to be trivial
-anyway. The empty context and extension provide the essentials of the rest of
-the theory in the sense that higher families are families over extensions,
-contexts are families over the empty context and the `action on families' of
-any operation applied to families over the empty context is the same as the
-original operation. Thus, when we have introduced the operations of weakening
-and substitution we will see that they also have an `action on the terms of 
-contexts'. We will state the relevant judgmental equality rules for the empty
-context in this section rather than in the next, since they are required to
-state that the action on families of an operation is compatible with the
-original operation.
-\end{comment}
-
-We note upfront that both context extension and family extension can also be
-seen as operations of contexts, families and terms, even though we will not
-introduce them as such. The reason is that extension is used to define what it
-means to be a family over a family over a context (namely, a family over an
-extended context). As such, it will have a trivial action on families and terms
-which will not need to be introduced. We will come back to this point in
-\autoref{extension}.
-
-\subsection{Extension}
-\label{extension}
-The operations of extension enable one to consider families over
-families and higher level families alike. We will need both a context extension
-operation and a family extension operation.
-
-The inference rules introducing context and family extension are as follows:
-\begin{align}
-& \inference
-  { \jfam{\Gamma}{A}
-    }
-  { \jctx{\ctxext{\Gamma}{A}}
-    }
-& & \inference
-    { \jctxeq{\Gamma}{\Delta}
-      \jfameq{\Gamma}{A}{B}
-      }
-    { \jctxeq{\ctxext{\Gamma}{A}}{\ctxext{\Delta}{B}}
-      }
-    \\
-& \inference
-  { \jfam{{\Gamma}{A}}{P}
-    }
-  { \jfam{\Gamma}{\ctxext{A}{P}}
-    }
-& & \inference
-    { \jfameq{\Gamma}{A}{B} 
-      \jfameq{{\Gamma}{A}}{P}{Q}
-      }
-    { \jfameq{\Gamma}{\ctxext{A}{P}}{\ctxext{B}{Q}}
-      }
-\end{align}
-The extension operation may also be defined to act on families over families
-and terms thereof. For instance, When $\jfam{{\Gamma}{A}}{P}$ is a family,
-we would get a family $\jfam{{\Gamma}{A}}{{\Gamma}{P}}$ and when
-$\jterm{{\Gamma}{A}}{P}{f}$ is a term, we would get a term
-$\jterm{{\Gamma}{A}}{{\Gamma}{P}}{\ctxext{\Gamma}{f}}$. Then there would be
-axioms stating the judgmental equalities $\ctxext{\Gamma}{P}\jdeq P$ and
-$\ctxext{\Gamma}{f}\jdeq f$. Thus, it is of no actual use to introduce the
-action of context extension on families and terms.
-
-In \autoref{extension-on-terms} we will also define an extension operation
-on terms, which defines a term $\jterm{\Gamma}{{A}{P}}{\tmext{x}{u}}$ for
-every $\jterm{\Gamma}{A}{x}$ and $\jterm{\Gamma}{\subst{x}{P}}{u}$. This
-operation is defined by using the substitution operation and identity terms.
-
-\subsubsection{Associativity of extension}
-\label{comp-ee}
-The inference rules asserting that extension is compatible with itself assert
-that contexts are unstructured lists of type declarations. This rule is
-unavoidable if we want that for a family $A$ in context $\Gamma$, a family over
-$A$ is the same thing as a family over $\ctxext{\Gamma}{A}$. 
-\begin{align}
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{{\Gamma}{A}}{P}
-    }
-  { \jctxeq{\ctxext{{\Gamma}{A}}{P}}{\ctxext{\Gamma}{{A}{P}}}
-    }
-  \\
-& \inference
-  { \jfam{{\Gamma}{A}}{P}
-    \jfam{{{\Gamma}{A}}{P}}{Q}
-    }
-  { \jfameq{\Gamma}{\ctxext{A}{{P}{Q}}}{\ctxext{{A}{P}}{Q}}
-    }
-  \label{comp-ee-c}
-\end{align}
-Note that we may take the point of view that extension distributes over
-extensions. If we want to extend an extended family $\ctxext{P}{Q}$ over
-$\ctxext{\Gamma}{A}$ by the family $A$, the result is the extension of (1)
-the extension of $P$ by $A$ and (2) the extension of $Q$ by $A$. The latter is
-simply $Q$ since extension acts trivially at the level of families (and terms).
-
-\subsection{The empty context and the empty families}
-\label{empty}
-We introduce an empty context and an family over $\Gamma$ for every context $\Gamma$. 
-It has been suggested by some to only include empty families and not an
-empty context because an empty context is not necessary, but we do have several
-reasons to include them. Having an empty context requires also rules asserting 
-that a context is the same thing as a family over the empty context and this
-gives the categorical structure on contexts for free once one has it for 
-families. The main ingredients that will be missing from the theory once an
-empty context is avoided are weakening by a context and identity morphisms from
-a context to itself. Including these by hand also requires to formulate all the
-compatibility rules involving weakening once more for the cases of weakening by
-a context and identity terms at contexts. We prefer to state these rules once
-and only once and including an empty context helps in this respect.
-
-We also prefer to have our scheme of compatibility rules as symmetrical as
-possible. The structure of type dependency should look exactly the same in the
-default case as in any context. In that respect, having an empty family but not
-an empty context seems a bit odd. Also, some sets of compatibility rules 
-(like the rules stating that extension is compatible with the empty families)
-will become assymetrical as a result of not including an empty context. Moreover,
-we would eventually like to include a stratification of the theory by means of
-a type judgment (asserting that a family in a context $\Gamma$ is a type) and
-study closed types (i.e.~types in the empty context). It would be possible to
-provide a notion of closed types without having an empty context, but this would
-have to be formulated separately and we would have to restate all the rules for
-types (if any) for closed types all over again.
-
-One of the main uses of the empty context and the empty families is that we
-get the property that the `action on contexts' of an operation is compatible
-with its `action on families'.
-
-The empty family over a context $\Gamma$ is introduced by the following rule
-inference rule:
-\begin{align}
-& \inference
-  { }
-  { \jctx{\emptyc}
-    }
-  \\
-& \inference
-  { \jctx{\Gamma}
-    }
-  { \jfam{\Gamma}{\emptyf[\Gamma]}
-    }
-  \\
-& \inference
-  { \jctxeq{\Gamma}{\Gamma'}
-    }
-  { \jfameq{\Gamma}{\emptyf[\Gamma]}{\emptyf[\Gamma']}
-    }
-\end{align}
-
-By regarding contexts as families of contexts over the empty context, we
-enable ourselves also to speak of terms of contexts. A term of a context
-$\Gamma$ is a term of the family $\Gamma$ over the empty context. These ideas
-are captured in the following convertibility rules:
-\begin{align}
-& \inference
-  { \jctx{\Gamma}
-    }
-  { \jfam{\emptyc}{\Gamma}
-    } 
-& & \inference
-    { \jfam{\emptyc}{\Gamma}
-      }
-    { \jctx{\Gamma}
-      }
-    \\
-& \inference
-  { \jctxeq{\Gamma}{\Delta}
-    }
-  { \jfameq{\emptyc}{\Gamma}{\Delta}
-    }
-& & \inference
-    { \jfameq{\emptyc}{\Gamma}{\Delta}
-      }
-    { \jctxeq{\Gamma}{\Delta}
-      }
-\end{align}
-
-The reader may wonder whether the empty family $\emptyf$ always has a
-term. This shall follow from the rules stating the compatibility of extension
-with the empty families in \autoref{comp-0e} below and from
-identity terms (\autoref{identityterms}).
-\begin{comment}
-When contexts are viewed as statements, the empty context is the statement that
-asserts nothing. Likewise, in any context one can assert nothing and this is
-achieved by the empty family. 
-Since somebody who states nothing asserts it all right, there is a term of the empty
-family over any context. Moreover, any two terms of the empty family are always
-judgmentally equal.
-\begin{align}
-& \inference
-  { }
-  { \jctx{\emptyc}
-    }
-  \\
-& \inference
-  { \jctx{\Gamma}
-    }
-  { \jfam{\Gamma}{\emptyf[\Gamma]}
-    }
-  \\
-& \inference
-  { \jctx{\Gamma}
-    }
-  { \jterm{\Gamma}{\emptyf[\Gamma]}{\emptytm[\Gamma]}
-    }
-  \\
-& \inference
-  { \jterm{\Gamma}{\emptyf[\Gamma]}{x}
-    }
-  { \jtermeq
-      {\Gamma}
-      {\emptyf[\Gamma]}
-      {x}
-      {\emptytm[\Gamma]}
-    }
-  \label{comp-00-t}
-\end{align}
-
-The empty context and the empty families together with the operation of
-extension will provide several shortcuts for the rest
-of the theory of families and terms. This is (in part) obtained by requiring
-that a context is exactly the same thing as a family of contexts over the
-empty context and that judgmental equality of contexts is exactly the same
-as judgmental equality of families of contexts over the empty context.
-
-\label{comp-00}
-Because every context is a family of contexts over the empty context, 
-the empty context $\emptyc$ may be considered as a family of contexts over the empty
-context. When we do this, we get $\emptyf[\emptyc]$.
-\begin{equation}\label{eq:comp-00}
-\inference
-  { }
-  { \jfameq
-      {\emptyc}
-      {\emptyc}
-      {\emptyf[\emptyc]}
-    }
-\end{equation}
-In the future, we shall denote $\emptyf[\Gamma]$ by $\emptyf$. The above rule
-guarantees that this will not cause confusion. Likewise, we shall denote
-$\emptytm[\Gamma]$ by $\emptytm$.
-\end{comment}
-
-\subsubsection{Compatibility of extension with the empty context and families}
-In the following set of inference rules we state that the empty context and
-the empty family are neutral objects for both context extension (the first two
-rules) and family extension (the last two rules).
-\label{comp-e0}\label{comp-0e}
-\begin{align}
-& \inference
-  { \jctx{\Gamma}
-    }
-  { \jctxeq{\ctxext{\emptyc}{\Gamma}}{\Gamma}
-    }
-  \label{comp-0e-c}
-  \\
-& \inference
-  { \jctx{\Gamma}
-    }
-  { \jctxeq{\ctxext{\Gamma}{\emptyf}}{\Gamma}
-    }
-  \label{comp-e0-c}\\
-& \inference
-  { \jfam{\Gamma}{A}
-    }
-  { \jfameq{\Gamma}{\ctxext{\emptyf}{A}}{A}
-    }
-  \label{comp-0e-f}
-  \\
-& \inference
-  { \jfam{\Gamma}{A}
-    }
-  { \jfameq{\Gamma}{{A}{\emptyf}}{A}
-    }
-  \label{comp-e0-f}
-\end{align}
-
-\subsubsection{Family extension restricted to the empty context}
-Strictly speaking we should have used a different notation for context extension
-as for family extension, because the following rule asserting that family extension
-in the empty context is the same thing as context extension would look tautological
-without a difference. So let us denote, only for the moment, context extension
-of $\Gamma$ by $A$ by $(\ctxext{\Gamma}{A})^c$ and family extension of $A$ by
-$P$ in context $\Gamma$ by $(\ctxext{A}{P})^\famsym$.
-
-Note that we may consider a context $\Gamma$ as a family over $\emptyc$ and
-a family $\jfam{\Gamma}{A}$ as a family $\jfam{{\emptyc}{\Gamma}}{A}$. 
-Therefore we will require the following rule:
-\begin{align}
-& \inference
-  { \jfam{\Gamma}{A}
-    }
-  { \jctxeq{(\ctxext{\Gamma}{A})^c}{(\ctxext{\Gamma}{A})^\famsym}
-    }
-\end{align}
-Note that this rule actually justifies that we have not utilized two different
-notations for context extension and family extension.
-
-\subsection{The theory of weakening}
-\label{weakening}
-When $A$ is a family in context $\Gamma$, the operation of weakening by $A$
-takes a family $B$ in context $\Gamma$ and provides a family $\ctxwk{A}{B}$
-in context $\ctxext{\Gamma}{A}$. The context family $\ctxwk{A}{B}$ can be seen
-as the constant family over $\ctxext{\Gamma}{A}$ with value $B$. This idea will
-be axiomatized in the cancellation property of weakening and substitution in
-\autoref{cancellation-ws}. In \autoref{morphisms} we will take the terms 
-$\unfold{\jhom{\Gamma}{A}{B}{f}}$ to be the morphisms of families from $A$ to 
-$B$. These will be at the heart of the categorical structure of the theory.
-
-The weakening operation acts on three levels: on contexts, on families and
-on terms. The `action on contexts' of weakening is the action we described
-above: it takes a family $B$ over $\Gamma$ to a family $\ctxwk{A}{B}$ over
-$\ctxext{\Gamma}{A}$; the `action on families' of weakening takes a family
-$Q$ over $\ctxext{\Gamma}{B}$ to a family $\ctxwk[\famsym]{A}{Q}$ over
-$\ctxext{{\Gamma}{A}}{\ctxwk{A}{B}}$; the `action on terms' of weakening takes
-a term $g$ of $Q$ to a term $\ctxwk[\tmsym]{A}{g}$ of $\ctxwk[\famsym]{A}{Q}$.
-\begin{align}
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{\Gamma}{B}
-    }
-  { \jfam{{\Gamma}{A}}{\ctxwk{A}{B}}
-    }
-& & \inference
-    { \jfameq{\Gamma}{A}{A'}
-      \jfameq{\Gamma}{B}{B'}
-      }
-    { \jfameq{{\Gamma}{A}}{\ctxwk{A}{B}}{\ctxwk{A'}{B'}}
-      }
-    \\
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{{\Gamma}{B}}{Q}
-    }
-  { \jfam{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk[\famsym]{A}{Q}}
-    }
-& & \inference
-    { \jfameq{\Gamma}{A}{A'}
-      \jfameq{{\Gamma}{B}}{Q}{Q'}
-      }
-    { \jfameq
-        {{{\Gamma}{A}}{\ctxwk{A}{B}}}
-        {\ctxwk[\famsym]{A}{Q}}
-        {\ctxwk[\famsym]{A'}{Q'}}
-      }
-    \\
-%& \inference
-%  { \jterm{\Gamma}{B}{y}
-%    }
-%  { \jterm{{\Gamma}{A}}{\ctxwk{A}{B}}{\ctxwk{A}{y}}
-%    }
-%& & \inference
-%    { \jtermeq{\Gamma}{B}{y}{y'}
-%      }
-%    { \jtermeq
-%        {{\Gamma}{A}}
-%        {\ctxwk{A}{B}}
-%        {\ctxwk{A}{y}}
-%        {\ctxwk{A'}{y'}}
-%      }
-%    \\
-& \inference
-  { \jfam{\Gamma}{A}
-    \jterm{{\Gamma}{B}}{Q}{g}
-    }
-  { \jterm{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk[\famsym]{A}{Q}}{\ctxwk[\tmsym]{A}{g}}
-    }
-& & \inference
-    { \jfameq{\Gamma}{A}{A'}
-      \jtermeq{{\Gamma}{B}}{Q}{g}{g'}
-      }
-    { \jtermeq
-        {{{\Gamma}{A}}{\ctxwk{A}{B}}}
-        {\ctxwk[\famsym]{A}{Q}}
-        {\ctxwk[\tmsym]{A}{g}}
-        {\ctxwk[\tmsym]{A'}{g'}}
-      }
-\end{align}
-
-\subsubsection{Weakening of an empty family}
-The following inference rules express that when the empty family is
-weakened, the result is the empty family.
-\label{comp-w0}
-\begin{align}
-%& \inference
-%  { \jctx{\Gamma}
-%    }
-%  { \jfameq{\Gamma}{\ctxwk{\Gamma}{\emptyc}}{\emptyf}
-%    }
-%  \\
-& \inference
-  { \jfam{\Gamma}{A}
-    }
-  { \jfameq{{\Gamma}{A}}{\ctxwk{A}{\emptyf}}{\emptyf}
-    }
-  \label{comp-w0-c}\\
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{\Gamma}{B}
-    }
-  { \jfameq
-    {{{\Gamma}{A}}{\ctxwk{A}{B}}}
-    {\ctxwk[\famsym]{A}{\emptyf}}
-    {\emptyf}
-    }
-  \label{comp-w0-f}
-\end{align}
-\begin{comment}
-Note that we do not have to axiomatize the action of weakening
-by $A$ on the term of the empty family, because the result will be in the empty 
-family and every term in the empty family is judgmentally equal to the term 
-$\emptytm$.
-\end{comment}
-
-Because a family over $\Gamma$ is the same as a family over 
-$\ctxext{\Gamma}{\emptyf}$ we can apply both the action on contexts and the
-action on families of weakening to a family $B$ over $\Gamma$. When we apply
-the action on families, we obtain a family $\ctxwk[\famsym]{A}{B}$ over the
-context $\ctxext{{\Gamma}{A}}{\ctxwk{A}{\emptyf}}$. However, since we have
-postulated the judgmental equalities $\ctxwk{A}{\emptyf}\jdeq\emptyf$ and
-$\ctxext{{\Gamma}{A}}{\emptyf}\jdeq\ctxext{\Gamma}{A}$, we see that we can
-compare $\ctxwk[\famsym]{A}{B}$ with $\ctxwk{A}{B}$. The following inference
-rule postulates that these two are judgmentally equal:
-\begin{equation}
-\inference
-{ \jfam{\Gamma}{A}
-  \jfam{\Gamma}{B}
-  }
-{ \jfameq{{\Gamma}{A}}{\ctxwk[\famsym]{A}{B}}{\ctxwk{A}{B}}
-  }
-\end{equation}
-Due to this rule, the action on contexts and the action on families of weakening
-are compatible with each other and consequently there can be no possible
-confusion when we omit the annotations $\famsym$ and $\tmsym$ alltogether. In
-the future, the weakening of a family $Q$ over $\ctxext{\Gamma}{B}$ shall
-be denoted just by $\ctxwk{A}{Q}$ and likewise the weakening of a term $g$ of
-$Q$ shall be denoted by $\ctxwk{A}{g}$.
-
-Because a family $B$ over $\Gamma$ can be treated as a family by the operation
-of weakening, weakening also acts on the terms of $B$. The weakening of a term
-$y$ of $B$ by $A$ can be seen as the constant term (or function) of the
-family $\ctxwk{A}{B}$.
-
-\subsubsection{Weakenings of extensions}
-\label{comp-we}
-The following rules assert that when an extended family is weakened, the
-weakening distributes over the extension factors.
-\begin{align}
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{{\Gamma}{B}}{Q}
-    }
-  { \jfameq
-      {\ctxext{\Gamma}{A}}
-      {\ctxwk{A}{\ctxext{B}{Q}}}
-      {\ctxext{\ctxwk{A}{B}}{\ctxwk{A}{Q}}}
-    }
-  \label{comp-we-c}
-  \\
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{{{\Gamma}{B}}{Q}}{R}
-    }
-  { \jfameq
-      {\ctxext{{\Gamma}{A}}{\ctxwk{A}{B}}}
-      {\ctxwk{A}{\ctxext{Q}{R}}}
-      {\ctxext{\ctxwk{A}{Q}}{\ctxwk{A}{R}}}
-    }
-  \label{comp-we-f}
-\end{align}
-When thinking of terms of $\ctxwk{A}{B}$ as morphisms of families from $A$ to
-$B$, this looks already like form of type theoretic choice. It is weaker in that
-it is not stated with function types, yet it is stronger in that it states a
-judgmental equality between two families.
-
-There is also a version of this property where an extended term is weakened.
-This variant is stated and proved in \autoref{comp-we-t}.
-
-\subsubsection{Weakening of weakenings}
-Weakening by a family $A$ in context $\Gamma$ brings things in context $\Gamma$
-to things in context $\ctxext{\Gamma}{A}$. Since we have all the ingredients of
-the theory of contexts, families and terms in the context $\Gamma$ as well it
-has in particular it's own weakening by families. Suppose we have a family
-$\jfam{{\Gamma}{B}}{Q}$ over $B$ in context $\Gamma$. Weakening by $Q$ brings
-things from context $\ctxext{\Gamma}{B}$ to $\ctxext{{\Gamma}{B}}{Q}$. Thus
-we can provide rules asserting what will happen when we first weaken by $Q$ and
-then (via the action on families) by $A$. We will require the following
-inference rules to be valid:
-\label{comp-ww}
-\begin{align}
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{{\Gamma}{B}}{Q}
-    \jfam{{\Gamma}{B}}{R}
-    }
-  { \jfameq
-      {{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{Q}}}
-      {\ctxwk{A}{{Q}{R}}}
-      {\ctxwk{{A}{Q}}{{A}{R}}}
-    }
-  \label{comp-ww-c}\\
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{{\Gamma}{B}}{Q}
-    \jfam{{{\Gamma}{B}}{R}}{S}
-    }
-  { \jfameq
-      {{{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{Q}}}{\ctxwk{A}{{Q}{R}}}}
-      {\ctxwk{A}{{Q}{S}}}
-      {\ctxwk{{A}{Q}}{{A}{S}}}
-    }
-  \label{comp-ww-f}\\
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{{\Gamma}{B}}{Q}
-    \jterm{{{\Gamma}{B}}{R}}{S}{k}
-    }
-  { \jtermeq
-      {{{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{Q}}}{\ctxwk{A}{{Q}{R}}}}
-      {\ctxwk{A}{{Q}{S}}}
-      {\ctxwk{A}{{Q}{k}}}
-      {\ctxwk{{A}{Q}}{{A}{k}}}
-    }
-  \label{comp-ww-t}
-\end{align}
-
-\begin{rmk}
-As an important special case of these inference rules we have the following
-valid inference rules:
-\begin{align*}
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{\Gamma}{B}
-    \jfam{\Gamma}{C}
-    }
-  { \jfameq
-      {{{\Gamma}{A}}{\ctxwk{A}{B}}}
-      {\ctxwk{A}{{B}{C}}}
-      {\ctxwk{{A}{B}}{{A}{C}}}
-    }
-  \\
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{\Gamma}{B}
-    \jfam{{\Gamma}{C}}{R}
-    }
-  { \jfameq
-      {{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{{B}{C}}}}
-      {\ctxwk{A}{{B}{R}}}
-      {\ctxwk{{A}{B}}{{A}{R}}}
-    }
-  \\
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{\Gamma}{B}
-    \jterm{{\Gamma}{C}}{R}{h}
-    }
-  { \jtermeq
-      {{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{{B}{C}}}}
-      {\ctxwk{A}{{B}{R}}}
-      {\ctxwk{A}{{B}{h}}}
-      {\ctxwk{{A}{B}}{{A}{h}}}
-    }
-\end{align*}
-Moreover, we have
-\begin{equation*}
-\inference
-  { \jfam{\Gamma}{A}
-    \jfam{\Gamma}{B}
-    \jterm{\Gamma}{C}{z}
-    }
-  { \jtermeq
-      {{{\Gamma}{A}}{\ctxwk{A}{B}}}
-      {\ctxwk{A}{{B}{C}}}
-      {\ctxwk{A}{{B}{z}}}
-      {\ctxwk{{A}{B}}{{A}{z}}}
-    }
-\end{equation*}
-\end{rmk}
-
-\subsubsection{Weakening by the empty family}
-Note that we can also weaken by the empty family over $\Gamma$.
-Weakening by the empty family $\emptyf$ over a context $\Gamma$ leaves families, 
-their terms, families over those families and terms of those unchanged:
-\label{comp-0w}
-\begin{align}
-& \inference
-  { \jfam{\Gamma}{B}
-    }
-  { \jfameq{\Gamma}{\ctxwk{\emptyf}{B}}{B}
-    }
-  \label{comp-0w-c}
-  \\
-& \inference
-  { \jfam{{\Gamma}{B}}{Q}
-    }
-  { \jfameq{{\Gamma}{B}}{\ctxwk{\emptyf}{Q}}{Q}
-    }
-  \label{comp-0w-f}
-  \\
-& \inference
-  { \jterm{{\Gamma}{B}}{Q}{g}
-    }
-  { \jtermeq{{\Gamma}{B}}{Q}{\ctxwk{\emptyf}{g}}{g}
-    }
-  \label{comp-0w-t}
-\end{align}
-
-\subsubsection{Currying for weakening}
-\label{comp-ew}
-The rules expressing that extension is compatible with weakening assert that
-weakening by an extension is the same thing as weakening twice in the
-appropriate way.
-\begin{align}
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{{\Gamma}{A}}{P}
-    \jfam{\Gamma}{B}
-    }
-  { \jfameq
-      {{{\Gamma}{A}}{P}}
-      {\ctxwk{\ctxext{A}{P}}{B}}
-      {\ctxwk{P}{{A}{B}}}
-    }
-  \label{comp-ew-c}\\
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{{\Gamma}{A}}{P}
-    \jfam{{\Gamma}{B}}{Q}
-    }
-  { \jfameq
-      {{{{\Gamma}{A}}{P}}{\ctxwk{P}{{A}{B}}}}
-      {\ctxwk{\ctxext{A}{P}}{Q}}
-      {\ctxwk{P}{{A}{Q}}}
-    }
-  \label{comp-ew-f}\\
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{{\Gamma}{A}}{P}
-    \jterm{{\Gamma}{B}}{Q}{g}
-    }
-  { \jtermeq
-      {{{{\Gamma}{A}}{P}}{\ctxwk{P}{{A}{B}}}}
-      {\ctxwk{P}{{A}{Q}}}
-      {\ctxwk{\ctxext{A}{P}}{g}}
-      {\ctxwk{P}{{A}{g}}}
-    } 
-  \label{comp-ew-t}
-\end{align}
-
-\begin{comment}
-\subsubsection{Weakening by a context}
-Since we have not required that there is an empty context with the property that
-a context is the same thing as a family in the empty context, we will have to
-explicitly introduce weakening by contexts. Since it will allways be clear
-whether something is weakened by a context or by a family, we shall use the
-same notation for weakening by contexts as for weakening by families. We shall
-have to repeat all the compatibility rules for weakening by families for
-weakening by contexts; there is no way around it. It is
-actually this sort of inconvenience that has originally led us to include an
-empty context and be done with it.
-
-In the following inference rules we assume that we have a context $\Gamma$ and
-that $\jctxeq{\Gamma}{\Gamma'}$ in the rules where $\Gamma'$ makes an appearence.
-\begin{align}
-& \inference
-  { \jctx{\Delta}
-    }
-  { \jfam{\Gamma}{\ctxwk{\Gamma}{\Delta}}
-    }
-& & \inference
-  { \jctxeq{\Delta}{\Delta'}
-    }
-  { \jfameq{\Gamma}{\ctxwk{\Gamma}{\Delta}}{\ctxwk{\Gamma'}{\Delta'}}
-    }
-  \\
-& \inference
-  { \jfam{\Delta}{B}
-    }
-  { \jfam{\ctxext{\Gamma}{\ctxwk{\Gamma}{\Delta}}}{\ctxwk{\Gamma}{B}}
-    }
-& & \inference
-  { \jfameq{\Delta}{B}{B'}
-    }
-  { \jfameq
-      {\ctxext{\Gamma}{\ctxwk{\Gamma}{\Delta}}}
-      {\ctxwk{\Gamma}{B}}
-      {\ctxwk{\Gamma'}{B'}}
-    }
-  \\
-& \inference
-  { \jterm{\Delta}{B}{y}
-    }
-  { \jterm
-      {\ctxext{\Gamma}{\ctxwk{\Gamma}{\Delta}}}
-      {\ctxwk{\Gamma}{B}}
-      {\ctxwk{\Gamma}{y}}
-    }
-& & \inference
-  { \jtermeq{\Delta}{B}{y}{y'}
-    }
-  { \jtermeq
-      {\ctxext{\Gamma}{\ctxwk{\Gamma}{\Delta}}}
-      {\ctxwk{\Gamma}{B}}
-      {\ctxwk{\Gamma}{y}}
-      {\ctxwk{\Gamma}{y'}}
-    }
-\end{align}
-Note that we do not employ three different notations for the three different
-actions of weakening by a context because there can be no confusion about which
-of the three actions is meant. If there were an empty context we should have
-distinguished the three as we did for weakening by a family.
-
-We shall have similar compatibility rules for weakening by a context as we did
-for weakening by a family.
-
-{\color{red}To do. It will look exactly the same as for weakening by families}
-\end{comment}
-
-\subsection{The theory of projections}
-\label{identityterms}
-We formulate the theory of projections on top of the theory of weakening
-described in \autoref{weakening}. In the theory of projections we introduce
-the identity terms. The identity term at a family $A$ in context $\Gamma$ is
-a term $\jterm{{\Gamma}{A}}{\ctxwk{A}{A}}{\idtm{A}}$, so weakening is necessary.
-Together with weakening, the identity terms will provide for the projections.
-The projection from $\cftext{A}{P}$ to $A$ will be the term
-$\jterm{{\Gamma}{{A}{P}}}{\ctxwk{\ctxext{A}{P}}{A}}{\ctxwk{P}{\idtm{A}}}$ and
-will be studied extensively in \autoref{extension-on-terms}.
-
-The following inference rules are required to be valid:
-\begin{align}
-& \inference
-  { \jfam{\Gamma}{A}
-    }
-  { \unfoldall{\jhom{\Gamma}{A}{A}{\idtm{A}}}
-    }
-& & \inference
-    { \jfameq{\Gamma}{A}{A'}
-      }
-    { \unfoldall{\jhomeq{\Gamma}{A}{A}{\idtm{A}}{\idtm{A'}}}
-      }
-\end{align}
-
-\begin{comment}
-We stress again
-that because we have not required there to be an empty context, we need to
-explicitly call for identity terms $\idtm{\Gamma}$ of $\ctxwk{\Gamma}{\Gamma}$
-in context $\Gamma$ for each context $\Gamma$. Thus, we will also require
-the inference rule
-\begin{align}
-& \inference
-  { \jctx{\Gamma}
-    }
-  { \jterm{\Gamma}{\ctxwk{\Gamma}{\Gamma}}{\idtm{\Gamma}}
-    }
-& &
-  \inference
-  { \jctxeq{\Gamma}{\Gamma'}
-    }
-  { \jtermeq{\Gamma}{\ctxwk{\Gamma}{\Gamma}}{\idtm{\Gamma}}{\idtm{\Gamma'}}
-    }
-\end{align}
-For the following inference rules, there should also be versions of identity
-terms at contexts.
-\end{comment}
-
-\subsubsection{The identity term of an extended family}
-The fact that identity terms are compatible with extensions is derived in
-\autoref{comp-ie}.
-
-\subsubsection{The identity term of a weakened family}
-\label{comp-wi}
-The identity term of a weakened family is the weakened identity term:
-\begin{equation}
-\inference
-  { \jfam{\Gamma}{A}
-    \jfam{{\Gamma}{B}}{Q}
-    }
-  { \unfoldall{\jhomeq
-      {{{\Gamma}{A}}{\ctxwk{A}{B}}}
-      {\ctxwk{A}{Q}}
-      {\ctxwk{A}{Q}}
-      {\ctxwk{A}{\idtm{Q}}}
-      {\idtm{\ctxwk{A}{Q}}}
-      }
-    }
-  \label{comp-wi-t}
-\end{equation}
-
-\begin{rmk}
-An important special case is the judgmental equality
-\begin{equation*}
-\unfoldall{\jhomeq
-      {{\Gamma}{A}}
-      {\ctxwk{A}{B}}
-      {\ctxwk{A}{B}}
-      {\ctxwk{A}{\idtm{B}}}
-      {\idtm{\ctxwk{A}{B}}}
-      }
-\end{equation*}
-for a family $\jfam{\Gamma}{B}$.
-\end{rmk}
-
-\subsection{The theory of substitution}
-\label{substitution}
-The theory of substitution can be formulated separately from the theory of
-weakening. Therefore, we will only require the theory of extension and the
-empty context and families. In \autoref{sec:esystem-equalities} we will
-provide the additional judgmental equality rules that will be required when
-combining the theories of substitution and of projections.
-
-Given a family $P$ over $A$ and a term $x$ of $A$, substitution gives a way to
-consider the \emph{fiber $\subst{x}{P}$ of $P$ at $x$}. As was the case with
-weakening, the substitution operation comes in three parts: the `action on
-contexts' of substitution is the operation just described; the `action on
-families' of substitution takes a family $Q$ over $\ctxext{{\Gamma}{A}}{P}$
-to a family $\subst[\famsym]{x}{Q}$ over $\ctxext{\Gamma}{\subst{x}{P}}$; the
-`action on terms' of substitution takes a term $g$ of $Q$ over
-$\ctxext{{\Gamma}{A}}{P}$ to a term $\subst[\tmsym]{x}{g}$ of 
-$\subst[\famsym]{x}{Q}$.
-\begin{align}
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jfam{{\Gamma}{A}}{P}
-    }
-  { \jfam{\Gamma}{\subst{x}{P}}
-    }
-& & \inference
-    { \jtermeq{\Gamma}{A}{x}{x'}
-      \jfameq{{\Gamma}{A}}{P}{P'}
-      }
-    { \jfameq{\Gamma}{\subst{x}{P}}{\subst{x'}{P'}}
-      }
-    \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jfam{{{\Gamma}{A}}{P}}{Q}
-    }
-  { \jfam{{\Gamma}{\subst{x}{P}}}{\subst[\famsym]{x}{Q}}
-    }
-& & \inference
-    { \jtermeq{\Gamma}{A}{x}{x'}
-      \jfameq{{{\Gamma}{A}}{P}}{Q}{Q'}
-      }
-    { \jfameq{{\Gamma}{\subst{x}{P}}}{\subst[\famsym]{x}{Q}}{\subst[\famsym]{x'}{Q'}}
-      }
-    \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jterm{{{\Gamma}{A}}{P}}{Q}{g}
-    }
-  { \jterm{{\Gamma}{\subst{x}{P}}}{\subst[\famsym]{x}{Q}}{\subst[\tmsym]{x}{g}}
-    }
-& & \inference
-    { \jtermeq{\Gamma}{A}{x}{x'}
-      \jtermeq{{{\Gamma}{A}}{P}}{Q}{g}{g'}
-      }
-    { \jtermeq
-        {{\Gamma}{\subst{x}{P}}}
-        {\subst[\famsym]{x}{Q}}
-        {\subst[\tmsym]{x}{g}}
-        {\subst[\tmsym]{x'}{g'}}
-      }
-\end{align}
-
-\subsubsection{Fibers of an empty family}
-The following inference rules establish that the fibers of the empty family are 
-the empty families:
-\label{comp-s0}
-\begin{align}
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    }
-  { \jfameq{\Gamma}{\subst{x}{\emptyf}}{\emptyf}
-    }
-  \label{comp-s0-c}
-  \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jfam{{\Gamma}{A}}{P}
-    }
-  { \jfameq
-      {{\Gamma}{\subst{x}{P}}}
-      {\subst{x}{\emptyf}}
-      {\emptyf}
-    }
-  \label{comp-s0-f}
-\end{align}
-\begin{comment}
-Note again that we do not have to axiomatize the action of substitution by $x$ on the term
-of the empty family, because the result will be in the empty family and every
-term in the empty family is judgmentally equal to the term $\emptytm$.
-\end{comment}
-
-We use the above rule to state the compatibility of the action on families of
-substitution with the action on contexts of substitution. Note that a family
-$P$ over $\ctxext{\Gamma}{A}$ may be regarded as a family over
-$\ctxext{{\Gamma}{A}}{\emptyf}$. Thus, we may consider the family
-$\subst[\famsym]{x}{P}$ over $\ctxext{\Gamma}{\subst{x}{\emptyf}}$. Since
-$\subst{x}{\emptyf}$ is judgmentally equal to the empty family, we may compare
-$\subst[\famsym]{x}{P}$ with $\subst{x}{P}$:
-\begin{equation}
-\inference
-{ \jfam{{\Gamma}{A}}{P}
-  }
-{ \jfameq{\Gamma}{\subst[\famsym]{x}{P}}{\subst{x}{P}}
-  }
-\end{equation}
-Due to this rule we need not make the annotations $\famsym$ and $\tmsym$ in
-the notation for substitution anymore and thus we shall omit them from now on.
-Note that because a family $P$ over $\ctxext{\Gamma}{A}$ is eligible for
-application of the action on families of substitution, we may also substitute
-terms of $P$. Thus, given terms $\jterm{\Gamma}{A}{x}$ and $\jterm{{\Gamma}{A}}{P}{f}$,
-we get a term $\jterm{\Gamma}{\subst{x}{P}}{\subst{x}{f}}$, the \emph{value of
-$f$ at $x$}.
-
-In \autoref{morphisms} we will use a combination of weakening and substitution
-to define composition of morphisms of families. However, we have to rely
-on the cancellation rule stated in \autoref{cancellation-ws} before we can
-meaningfully state the definition of composition.
-
-\subsubsection{Fibers of extensions}
-\label{comp-se}
-The following inference rules assert that if we take the fiber of an extended
-family at a term $x$ of $A$ in context $\Gamma$, the substitution by $x$
-distributes over the factors of the extension.
-\begin{align}
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jfam{{{\Gamma}{A}}{P}}{Q}
-    }
-  { \jfameq
-      {\Gamma}
-      {\subst{x}{\ctxext{P}{Q}}}
-      {\ctxext{\subst{x}{P}}{\subst{x}{Q}}}
-    }
-  \label{comp-se-c}
-  \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jfam{{{{\Gamma}{A}}{P}}{Q}}{R}
-    }
-  { \jfameq
-      {{\Gamma}{\subst{x}{P}}}
-      {\subst{x}{\ctxext{Q}{R}}}
-      {\ctxext{\subst{x}{Q}}{\subst{x}{R}}}
-    }
-  \label{comp-se-f}
-\end{align}
-There is also a version of this statement in which extended terms are considered.
-This variant is stated and proved in \autoref{comp-se-t}.
-
-\subsubsection{Fibers of fibers}
-\label{comp-ss}
-
-The following rules explain what happens when we first substitute by a term
-$g$ of $Q$ over $P$ in context $\ctxext{\Gamma}{A}$ and then by a term $x$ of
-$A$:
-\begin{align}
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jterm{{{\Gamma}{A}}{P}}{Q}{g}
-    \jfam{{{{\Gamma}{A}}{P}}{Q}}{R}
-    }
-  { \jfameq
-      {{{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}}
-      {\subst{x}{{g}{R}}}
-      {\subst{{x}{g}}{{x}{R}}}
-    }
-  \label{comp-ss-c}
-  \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jterm{{{\Gamma}{A}}{P}}{Q}{g}
-    \jfam{{{{{\Gamma}{A}}{P}}{Q}}{R}}{S}
-    }
-  { \jfameq
-      {{{{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}}{\subst{x}{{g}{R}}}}
-      {\subst{x}{{g}{S}}}
-      {\subst{{x}{g}}{{x}{S}}}
-    }
-  \label{comp-ss-f}
-  \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jterm{{{\Gamma}{A}}{P}}{Q}{g}
-    \jterm{{{{{\Gamma}{A}}{P}}{Q}}{R}}{S}{k}
-    }
-  { \jtermeq
-      {{{{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}}{\subst{x}{{g}{R}}}}
-      {\subst{x}{{g}{S}}}
-      {\subst{x}{{g}{k}}}
-      {\subst{{x}{g}}{{x}{k}}}
-    }
-  \label{comp-ss-t}
-\end{align}
-
-\begin{rmk}
-As an important special case of these inference rules we have the following
-valid inference rules:
-\begin{align*}
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jterm{{\Gamma}{A}}{P}{f}
-    \jfam{{{\Gamma}{A}}{P}}{Q}
-    }
-  { \jfameq
-      {\Gamma}
-      {\subst{x}{{f}{Q}}}
-      {\subst{{x}{f}}{{x}{Q}}}
-    }
-  \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jterm{{\Gamma}{A}}{P}{f}
-    \jfam{{{{\Gamma}{A}}{P}}{Q}}{R}
-    }
-  { \jfameq
-      {{\Gamma}{\subst{x}{{f}{Q}}}}
-      {\subst{x}{{f}{R}}}
-      {\subst{{x}{f}}{{x}{R}}}
-    }
-  \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jterm{{\Gamma}{A}}{P}{f}
-    \jterm{{{{\Gamma}{A}}{P}}{Q}}{R}{h}
-    }
-  { \jtermeq
-      {{\Gamma}{\subst{x}{{f}{Q}}}}
-      {\subst{x}{{f}{R}}}
-      {\subst{x}{{f}{h}}}
-      {\subst{{x}{f}}{{x}{h}}}
-    }
-\end{align*}
-Moreover, we get
-\begin{equation*}
-\inference
-  { \jterm{\Gamma}{A}{x}
-    \jterm{{\Gamma}{A}}{P}{f}
-    \jterm{{{\Gamma}{A}}{P}}{Q}{g}
-    }
-  { \jtermeq
-      {\Gamma}
-      {\subst{x}{{f}{Q}}}
-      {\subst{x}{{f}{g}}}
-      {\subst{{x}{f}}{{x}{g}}}
-    }
-\end{equation*}
-\end{rmk}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Joining the theory of projections with the theory of substitution}
-\label{sec:esystem-equalities}
-
-\begin{comment}
-We have already seen that empty families, extension, weakening, substitution
-and the identity terms are compatible with each other. There are, however, a
-few more judgmental equalities explaining what happens when
-\begin{enumerate}
-\item something is weakend by the empty family. This is used to consider a term
-      $x$ of $A$ in context $\Gamma$ as a morphism from $\emptyf$ to $A$ in
-      context $\Gamma$;
-\item something is weakened by an extended family. This is Currying for weakening;
-\item something is first weakened by a family $A$ in context $\Gamma$ and then
-      substituted by a term of $A$. This is used to show that composition is
-      compatible with itself (i.e. that composition is associative);
-\item the identity terms are joining the party. Identity terms are empty contexts
-      in certain models, so they should behave accordingly.
-\end{enumerate}
-In other words, the rules seem to not fit in the scheme that all operations are
-compatible with each other, but in fact each of these rules plays a role in
-completing the scheme (work out further).
-\end{comment}
-
-\begin{comment}
-In this section we extend the judgmental equality relation in such a way that
-the empty context, extension, weakening, substitution and the identity terms
-all become compatible with each other. Since we have five operations, there
-are going to be $25$ finite sets of rules explaining how each operation is
-compatible with each of the operations.
-
-In addition to these $25$ finite sets
-of compatibility rules, there are also two sets of rules explaining the
-cancellation properties of weakening and substitution and of the identity terms.
-
-A typical compatibility axiom captures the idea that a diagram like
-\begin{equation*}
-\begin{tikzcd}[column sep=huge]
-\jfam{\Gamma}{\blank} 
-  \ar{r}{C\mapsto\ctxwk{B}{C}} 
-  \ar{d}[swap]{C\mapsto\ctxwk{A}{C}} 
-& \jfam{{\Gamma}{B}}{\blank} 
-  \ar{d}{Q\mapsto\ctxwk{A}{Q}}
-  \\
-\jfam{{\Gamma}{A}}{\blank} 
-  \ar{r}[swap]{P\mapsto\ctxwk{{A}{B}}{P}} 
-& \jfam{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\blank}
-\end{tikzcd}
-\end{equation*}
-in the case of two consecutive weakenings, or
-\begin{equation*}
-\begin{tikzcd}[column sep=huge]
-\jfam{{\Gamma}{A}}{\blank} 
-  \ar{d}[swap]{Q\mapsto\subst{x}{Q}} 
-  \ar{r}{Q\mapsto\ctxwk{P}{Q}} 
-& \jfam{{{\Gamma}{A}}{P}}{\blank} 
-    \ar{d}{R\mapsto\subst{x}{R}}
-  \\ 
-\jfam{\Gamma}{\blank} 
-  \ar{r}[swap]{B\mapsto\ctxwk{\subst{x}{P}}{B}} 
-& \jfam{{\Gamma}{\subst{x}{P}}}{\blank}
-\end{tikzcd}
-\end{equation*}
-in the case of a combination of weakening and substitution, commutes. We won't
-actually draw any of these diagrams in this section, but readers are encouraged
-to do so if they think it can help them understanding the axioms. Another thing
-one could keep in mind with the axioms that involve weakening is that all of
-them become completely obvious when weakening is made notationally invisible.
-Of course, we have an explicit notation for weakening on purpuse, but it is the
-usual practice of type theorists to make weakening invisible.
-
-The table below indicates where
-we can find the axioms of each of the combinations of operators. The reference
-at the row of operation $X$ and the column of operation $Y$ refers to the
-axiom which explains $Y$ distributes over the arguments of $X$ when $Y$ is
-applied after $X$. The empty context has no arguments, whereas extension,
-and weakening have two arguments which are both of the family-kind, substitution
-has two arguments of which one is a term and the other is a family and
-idenitity terms have one argument, a family.
-\begin{center}
-\begin{tabular}{r|rrrr}
-& ext & wk & subst & id\\
-\hline
-extension & \autoref{comp-ee}* &  & \cmark & \cmark \\
-weakening & \autoref{comp-we}* & \autoref{comp-ww} & \autoref{comp-sw} & \autoref{comp-wi} \\
-substitution & \autoref{comp-se}* & \autoref{comp-ws} & \autoref{comp-ss} & \autoref{comp-si} \\
-identity terms & \cmark & \autoref{comp-wi} & \autoref{comp-si} & \xmark
-\end{tabular}
-\end{center}
-Apart from references to sections where axioms are given, we also see checkmarks,
-crosses, marked references and duplicated references in this table. 
-The checkmarks indicate that the desired property
-can be proved and need not be axiomatized. To clarify those:
-\begin{itemize}
-\item The identity term of the empty context is the unique term of the empty
-family. Recall that we already have axiomatized in \autoref{empty} that every
-term of the empty family is judgmentally equal to $\emptytm$.
-\item The identity term of an extension can be explained by means of the notion
-of \emph{extension on terms}. This notion is defined in \autoref{extension-on-terms}
-using only identity terms and substitution. With extension on terms we can
-combine a term $\jterm{\Gamma}{A}{x}$ and a term $\jterm{\Gamma}{\subst{x}{P}}{u}$
-into a term $\jterm{\Gamma}{{A}{P}}{\tmext{x}{u}}$, which may be seen as the
-pair consisting of $x$ and $u$. Since it is a defined notion, we will
-be able to derive lots of properties for extension of terms. One of them explains
-how we can see the identity term of an extension as the extension of two certain
-terms, see \autoref{comp-ie}.
-\item Using extension on terms we can also make sense of `fibers at an extension'
-because that would exactly be the fiber at an extended term. We will see in
-\autoref{comp-es} that, for instance, $\subst{\tmext{x}{u}}{Q}\jdeq\subst{u}{{x}{Q}}$
-for any family $\jfam{{{\Gamma}{A}}{P}}{Q}$ and any two terms $\jterm{\Gamma}{A}{x}$
-and $\jterm{\Gamma}{\subst{x}{P}}{u}$. That's a form of currying.
-\end{itemize}
-The spots with a \xmark simply make no sense 
-{\color{red}(or do they, when the term model
-is better understood (by me)?)}. The marked references point to
-axioms which \emph{do} have a variant for terms, but those require the notion
-of extension on terms, can be proved and therefore need not be included in our
-list of compatibility axioms. In each of these sections, we give a forward
-reference to the lemma in which this variant is stated and proved.
-\end{comment}
-
-\subsubsection{Fibers of weakenings}\label{comp-sw}
-The following rules assert what happens when we first weaken by a family
-$Q$ over $P$ in context $\ctxext{\Gamma}{A}$ and then substitute by a term
-$x$ of $A$:
-\begin{align}
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jfam{{{\Gamma}{A}}{P}}{Q}
-    \jfam{{{\Gamma}{A}}{P}}{R}
-    }
-  { \jfameq
-      {{{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}}
-      {\subst{x}{\ctxwk{Q}{R}}}
-      {\ctxwk{\subst{x}{Q}}{\subst{x}{R}}}
-    }
-  \label{comp-sw-c}
-  \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jfam{{{\Gamma}{A}}{P}}{Q}
-    \jfam{{{{\Gamma}{A}}{P}}{R}}{S}
-    }
-  { \jfameq
-      {{{{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}}{\subst{x}{\ctxwk{Q}{R}}}}
-      {\subst{x}{\ctxwk{Q}{S}}}
-      {\ctxwk{\subst{x}{Q}}{\subst{x}{S}}}
-    }
-  \label{comp-sw-f}
-  \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jfam{{{\Gamma}{A}}{P}}{Q}
-    \jterm{{{{\Gamma}{A}}{P}}{R}}{S}{k}
-    }
-  { \jtermeq
-      {{{{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}}{\subst{x}{\ctxwk{Q}{R}}}}
-      {\subst{x}{\ctxwk{Q}{S}}}
-      {\subst{x}{\ctxwk{Q}{k}}}
-      {\ctxwk{\subst{x}{Q}}{\subst{x}{k}}}
-    }
-  \label{comp-sw-t}
-\end{align}
-
-\begin{rmk}
-As an important special case of these inference rules we have the following
-valid inference rules:
-\begin{align*}
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jfam{{\Gamma}{A}}{P}
-    \jfam{{\Gamma}{A}}{Q}
-    }
-  { \jfameq
-      {{\Gamma}{\subst{x}{P}}}
-      {\subst{x}{\ctxwk{P}{Q}}}
-      {\ctxwk{\subst{x}{P}}{\subst{x}{Q}}}
-    }
-  \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jfam{{\Gamma}{A}}{P}
-    \jfam{{{\Gamma}{A}}{Q}}{R}
-    }
-  { \jfameq
-      {{{\Gamma}{\subst{x}{P}}}{\subst{x}{\ctxwk{P}{Q}}}}
-      {\subst{x}{\ctxwk{P}{R}}}
-      {\ctxwk{\subst{x}{P}}{\subst{x}{R}}}
-    }
-  \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jfam{{\Gamma}{A}}{P}
-    \jterm{{{\Gamma}{A}}{Q}}{R}{h}
-    }
-  { \jtermeq
-      {{{\Gamma}{\subst{x}{P}}}{\subst{x}{\ctxwk{P}{Q}}}}
-      {\subst{x}{\ctxwk{P}{R}}}
-      {\subst{x}{\ctxwk{P}{h}}}
-      {\ctxwk{\subst{x}{P}}{\subst{x}{h}}}
-    }
-\end{align*}
-Moreover, we get
-\begin{equation*}
-\inference
-  { \jterm{\Gamma}{A}{x}
-    \jfam{{\Gamma}{A}}{P}
-    \jterm{{\Gamma}{A}}{Q}{g}
-    }
-  { \jtermeq
-      {{\Gamma}{\subst{x}{P}}}
-      {\subst{x}{\ctxwk{P}{Q}}}
-      {\subst{x}{\ctxwk{P}{g}}}
-      {\ctxwk{\subst{x}{P}}{\subst{x}{g}}}
-    }
-\end{equation*}
-\end{rmk}
-
-\subsubsection{Weakenings of fibers}
-\label{comp-ws}
-The following inference rules explain what happens when we first weaken by a
-term $g$ of $Q$ in context $\ctxext{\Gamma}{B}$ and then weaken by a family
-$A$ over $\Gamma$.
-\begin{align}
-& \inference
-  { \jfam{\Gamma}{A}
-    \jterm{{\Gamma}{B}}{Q}{g}
-    \jfam{{{\Gamma}{B}}{Q}}{R}
-    }
-  { \jfameq
-      {{{\Gamma}{A}}{\ctxwk{A}{B}}}
-      {\ctxwk{A}{\subst{g}{R}}}
-      {\subst{\ctxwk{A}{g}}{\ctxwk{A}{R}}}
-    }
-  \label{comp-ws-c}
-  \\
-& \inference
-  { \jfam{\Gamma}{A}
-    \jterm{{\Gamma}{B}}{Q}{g}
-    \jfam{{{{\Gamma}{B}}{Q}}{R}}{S}
-    }
-  { \jfameq
-      {{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{\subst{g}{R}}}}
-      {\ctxwk{A}{\subst{g}{S}}}
-      {\subst{\ctxwk{A}{g}}{\ctxwk{A}{S}}}
-    }
-  \label{comp-ws-f}
-  \\
-& \inference
-  { \jfam{\Gamma}{A}
-    \jterm{{\Gamma}{B}}{Q}{g}
-    \jterm{{{{\Gamma}{B}}{Q}}{R}}{S}{k}
-    }
-  { \jtermeq
-      {{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{\subst{g}{R}}}}
-      {\ctxwk{A}{\subst{g}{S}}}
-      {\ctxwk{A}{\subst{g}{k}}}
-      {\subst{\ctxwk{A}{g}}{\ctxwk{A}{k}}}
-    }
-  \label{comp-ws-t}
-\end{align}
-
-\begin{rmk}
-As an important special case of these inference rules we have the following 
-valid inference rules:
-\begin{align*}
-& \inference
-  { \jfam{\Gamma}{A}
-    \jterm{\Gamma}{B}{y}
-    \jfam{{\Gamma}{B}}{Q}
-    }
-  { \jfameq
-      {{\Gamma}{A}}
-      {\ctxwk{A}{\subst{y}{Q}}}
-      {\subst{\ctxwk{A}{y}}{\ctxwk{A}{Q}}}
-    }
-  \\
-& \inference
-  { \jfam{\Gamma}{A}
-    \jterm{\Gamma}{B}{y}
-    \jfam{{{\Gamma}{B}}{Q}}{R}
-    }
-  { \jfameq
-      {{{\Gamma}{A}}{\ctxwk{A}{\subst{y}{Q}}}}
-      {\ctxwk{A}{\subst{y}{R}}}
-      {\subst{\ctxwk{A}{y}}{\ctxwk{A}{R}}}
-    }
-  \\
-& \inference
-  { \jfam{\Gamma}{A}
-    \jterm{\Gamma}{B}{y}
-    \jterm{{{\Gamma}{B}}{Q}}{R}{h}
-    }
-  { \jtermeq
-      {{{\Gamma}{A}}{\ctxwk{A}{\subst{y}{Q}}}}
-      {\ctxwk{A}{\subst{y}{R}}}
-      {\ctxwk{A}{\subst{y}{h}}}
-      {\subst{\ctxwk{A}{y}}{\ctxwk{A}{h}}}
-    }
-\end{align*}
-Moreover, we get
-\begin{equation*}
-\inference
-  { \jfam{\Gamma}{A}
-    \jterm{\Gamma}{B}{y}
-    \jterm{{\Gamma}{B}}{Q}{g}
-    }
-  { \jtermeq
-      {{\Gamma}{A}}
-      {\ctxwk{A}{\subst{y}{Q}}}
-      {\ctxwk{A}{\subst{y}{g}}}
-      {\subst{\ctxwk{A}{y}}{\ctxwk{A}{g}}}
-    }
-\end{equation*}
-\end{rmk}
-
-\subsubsection{The cancellation property of weakening and substitution}
-\label{cancellation-ws}
-The judgmental equalities we're about to describe assert that substituting a term
-in the weakening a thing gives you the thing back. In the case of contexts we get that each fiber
-$\subst{x}{\ctxwk{A}{B}}$ is just $B$ and in the case of terms we get 
-that $\ctxwk{A}{y}$ is the constant function
-mapping everything to $y:B$. Thus, these rules actually establish the weakening
-as the weakening. After stating the rules we will describe what it means to
-compose context morphisms (terms of weakened contexts).
-\begin{align}
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{\Gamma}{B}
-    \jterm{\Gamma}{A}{x}
-    }
-  { \jfameq
-      {\Gamma}
-      {\subst{x}{\ctxwk{A}{B}}}
-      {B}
-    }
-  \label{cancellation-ws-c}\\
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{{\Gamma}{B}}{Q}
-    \jterm{\Gamma}{A}{x}
-    }
-  { \jfameq{{\Gamma}{B}}{\subst{x}{\ctxwk{A}{Q}}}{Q}
-    }
-  \label{cancellation-ws-f}\\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jterm{{\Gamma}{B}}{Q}{g}
-    }
-  { \jtermeq{{\Gamma}{B}}{Q}{\subst{x}{\ctxwk{A}{g}}}{g}
-    }
-  \label{cancellation-ws-t}
-\end{align}
-
-\subsubsection{The identity term of a substituted family}
-\label{comp-si}
-The identity term of a substituted family is the substitution of the identity term
-\begin{equation}
-\inference
-  { \jterm{\Gamma}{A}{x}
-    \jfam{{{\Gamma}{A}}{P}}{Q}
-    }
-  { \unfoldall{\jhomeq
-      {{\Gamma}{\subst{x}{P}}}
-      {\subst{x}{Q}}
-      {\subst{x}{Q}}
-      {\subst{x}{\idtm{Q}}}
-      {\idtm{\subst{x}{Q}}}
-    }}
-  \label{comp-si-t}
-\end{equation}
-
-\begin{rmk}
-An important special case is the judgmental equality
-\begin{equation*}
-\unfoldall{\jhomeq
-      {\Gamma}
-      {\subst{x}{P}}
-      {\subst{x}{P}}
-      {\subst{x}{\idtm{P}}}
-      {\idtm{\subst{x}{P}}}
-    }
-\end{equation*}
-for a family $\jfam{{\Gamma}{A}}{P}$.
-\end{rmk}
-
-\subsubsection{The cancellation properties of identity terms}
-\label{cancellation-i}
-Identity terms are determined by their behavior with respect to substitution combined with
-weakening. The identity terms will also be subject to compatibility rules.
-\begin{align}
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    }
-  { \jtermeq{\Gamma}{A}{\subst{x}{\idtm{A}}}{x}
-    }
-  \label{cancellation-si}\\
-& \inference
-  { \jfam{{\Gamma}{A}}{P}
-    }
-  { \jfameq{{\Gamma}{A}}{\subst{\idtm{A}}{\ctxwk{A}{P}}}{P}
-    }
-  \label{precomp-idtm-c}
-  \\
-& \inference
-  { \jfam{{{\Gamma}{A}}{P}}{Q}
-    }
-  { \jfameq{{{\Gamma}{A}}{P}}{\subst{\idtm{A}}{\ctxwk{A}{Q}}}{Q}
-    }
-  \label{precomp-idtm-f}
-  \\
-& \inference
-  { \jterm{{{\Gamma}{A}}{P}}{Q}{g}
-    }
-  { \jtermeq
-      {{{\Gamma}{A}}{P}}
-      {Q}
-      {\subst{\idtm{A}}{\ctxwk{A}{g}}}
-      {g}
-    }
-  \label{precomp-idtm-t}
-\end{align}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{The possiblity of types in the theory of contexts, families and
-terms}
-\label{types}
-We have deliberately not spoken of types so far because we have taken the point
-of view that a type in a context is nothing but a family in that context which
-belongs to the class of types. We think of types as \emph{irreducible} families,
-i.e.\ families which are neither empty nor the extension of two
-families which are both not empty (in algebraic terminology: which
-are both non-trivial). To allow ourselves to speak of types we introduce two
-new judgments: the judgment that something is a type and the judgment that two
-types are equal.
-\begin{align*}
-\jalign\jtype{\Gamma}{A} 
-& \jalign\jtypeeq{\Gamma}{A}{B}
-\end{align*}
-But only families of contexts are eligible to be types. If $A$ is a type
-in context $\Gamma$, then $A$ is also a family of contexts over $\Gamma$. 
-Moreover, two types in context $\Gamma$ are judgmentally equal precisely when they are equal
-as context families and if a family $B$ of contexts over $\Gamma$ is
-judgmentally equal to a type $A$ in context $\Gamma$, then $B$ is a type in
-context $\Gamma$. This is expressed by the following four inference rules:
-\begin{align*}
-& \inference
-  { \jtype{\Gamma}{A}
-    }
-  { \jfam{\Gamma}{A}
-    }
-& & \inference
-    { \jtypeeq{\Gamma}{A}{B}
-      }
-    { \jfameq{\Gamma}{A}{B}
-      }
-    \\
-& \inference
-  { \jtype{\Gamma}{A}
-    \jfameq{\Gamma}{A}{B}
-    }
-  { \jtype{\Gamma}{B}
-    }
-& & \inference
-    { \jtype{\Gamma}{A}
-      \jfameq{\Gamma}{A}{B}
-      }
-    { \jtypeeq{\Gamma}{A}{B}
-      }
-\end{align*}
-As pointed out at the beginning of this subsection, 
-we do not assume that the empty family is a type, that would be like
-assuming that the multiplicative unit of a ring is prime.
-
-\begin{comment}
-We envision a
-categorical interpretation of the theory of contexts, families and terms where 
-contexts are interpreted as finite paths of arrows belonging to a predetermined
-class. The empty family would be such a path of length $0$ whereas the types
-would be such paths of length $1$.
-\end{comment}
-
-We add rules asserting that a weakened type is again a type that 
-substitution preserves the property of being a type:
-\begin{align}
-& \inference
-  { \jfam{\Gamma}{B}
-    \jtype{\ctxext{\Gamma}{B}}{Q}
-    }
-  { \jtype{\ctxext{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{Q}}
-    }
-  \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jtype{\ctxext{{\Gamma}{A}}{P}}{Q}
-    }
-  { \jtype{\ctxext{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}
-    }
-\end{align}
-With only the current rules, the possibility of making the judgment that
-something is a type does not add much to the theory of contexts, families and
-terms. Nevertheless, when studying models, having an interpretation for the
-judgment that something is a type will allow for the possibility to study
-conditions such as the one asserting that every family `factorizes' uniquely
-as multiple applications of extension to types, analogous to the condition on
-unique factorization domains in ring theory. All the tradional models of type
-theory should translate to models with such a condition, simply because
-traditionally contexts are viewed as lists of variable (and type) declarations.

### 2014-10-28 Rules/rules_substitution.tex
reorganization

--- 2014-09-20 article-models-tt.tex
@@
-\section{The generalized algebraic theory of contexts, families and terms (the E-system)}\label{tt}
-In this section we give a description of dependent type theory before type
-constructors. Apart from contexts, families and terms -- which provide for the
-core of the language of dependent logic -- the basic ingredients
-of this theory will be the operations of extension on context and on families, 
-and empty context and empty families, weakening, substitution
-and the identity terms. The resulting theory can be seen as a manifestation of the 
-structure underlying dependent type theory.
-
-We will
-formulate the theory of contexts, families and terms in such a way that contexts aren't defined
-to be lists of variable declarations. The variable-free (a.k.a.~\emph{name-free}) approach 
-we take here is rather different than those appearing in 
-\cite{hofmann1995extensional,TheBook} but it has appeared in the work of Coquand
-and in \cite{Dybjer1996}.
-The main reason we don't let variable declarations in is that we don't see them 
-in the internal models either. This way we also set out to a more algebraic 
-approach of type theory and higher category theory. Thirdly, we will not have to
-be burdened with superficial comments about variables being bounded or not, or 
-fresh or free or not occuring at all.
-
-In the current presentation, contexts can be seen as binary
-planar trees of which the leaves are (families of) contexts. 
-The judgmental equality relation on contexts is an equivalence relation which 
-expresses that binary planar
-trees of contexts are judgmentally equal if their leaves are, taking only
-(the isomorphism class of) 
-the order of the leaves into account \emph{and not the actual shape of the three}.
-The intuition behind this equivalence relation is indeed that unstructured
-(i.e.~unbracketed) lists such as the lists of variable declarations which
-usually appear in type theoretical syntax, may be regarded as contexts.
-
-Besides contexts, families and terms there will also be a notion of `type in
-a context', which one may assume.
-A context is eligible to be a \emph{closed type} and over a context, a family
-may be a type. We have the intuition that `being a type'
-expresses the property of `being atomic or irreducible'. With a notion of types,
-one has the means to require a stratification by which every context is a finite
-extension of types starting at a closed type, and every family is likewise a
-finite extension of types. The idea behind this is to recover the theory of
-\cite{Garner2014} {\color{red}(and possibly B-systems?)}. We stress that
-we will not include axioms asserting such a stratification in the E-system
-and we leave discussion on the possibility of a typing judgment until the
-end of this section.
-
-With families of contexts being the principal things of
-study, we have added a bit more generality to the theory, compared to the
-theory of categories with families presented in \cite{Dybjer1996}. Here are
-a couple of reasons for this extra generality:
-\begin{enumerate}
-\item We want to be very sure that we can interpret all the rules of type theory
-in a straightforward manner in any context.
-\item We do want do end up with a categorical structure on the contexts, derived
-from the theory by requiring that a morphism from $\Gamma$ to $\Delta$ is a
-term of the weakening of $\Delta$ by $\Gamma$ in context $\Gamma$. Thus, we
-must be able to lift the context $\Delta$ to become a family of contexts over
-$\Gamma$ and it is not reasonable to expect that this results in a type.
-\item It is very easy in this setting to assert that an operation has an action
-on very high levels of families, i.e.~when $F$ is an operation taking things in
-context $\Gamma$ to things in context $\Delta$, then it takes a family $A$ in
-context $\Gamma$ to $F(A)$ in context $\Delta$; a family $P$ in context
-$\ctxext{\Gamma}{A}$ to a family $F(P)$ in context $\ctxext{\Delta}{(F(A))}$,
-and so on. The reason it becomes simple is that we don't have to explain what
-`and so on' means.
-\end{enumerate}
-
-We will formulate fairly strict rules governing the judgmental equalities,
-expressing that extension, weakening and substitution are combatible with
-each other in a judgmental manner. This does not, however, diminish the role
-of isomorphisms or of homotopies could play in the theory once identity types
-are added. Indeed, types could still have non-trivial identity relations and
-the category of types in a certain context could genuinely display higher
-categorical structure, or so we conjecture.
-
-Much of the rules we state are just compatibility rules of extension, weakening
-and substitution with each other. In a way, these rules assert that our contexts
-are just structureless lists of contexts and that likewise terms are structureless
-lists of terms. They are structureless in the sense that the order in which
-they are formed by pairing up is irrelevant. We note that this causes complications
-in the traditional way that categorical sematics of type theory is implemented,
-where contexts become objects of the category which is supposed to model type
-theory. The reason for this is that context extension will not satisfy all the
-compatibility rules we're about to state. The first step to resolving this is taking
-the types in the empty context as the objects.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Judgments and valid inference rules}\label{judgments}
-The theory we describe here is a theory of contexts, families of
-contexts and terms thereof. The families of contexts are by some authors called
-dependent contexts, but they are handled a bit differently here because they
-become the primary object of study. Dependent contexts can be types; they could
-be seen as atomic or indecomposable dependent contexts.
-
-Thus we make eight kinds of judgments: ``$\Gamma$ is a context'',
-``$A$ is a family of contexts over $\Gamma$'', ``$A$ is a type in context $\Gamma$''
-and ``$x$ is a term of the family $A$ of contexts over $\Gamma$''. The other four
-judgments are for judgmental equality.
-\begin{align*}
-\jalign\jctx{\Gamma} 
-& \jalign\jctxeq{\Gamma}{\Gamma'}
-  \\
-\jalign\jfam{\Gamma}{A} 
-& \jalign\jfameq{\Gamma}{A}{B}
-  \\
-\jalign\jterm{\Gamma}{A}{x} 
-& \jalign\jtermeq{\Gamma}{A}{x}{y}.
-\end{align*}
-
-Strictly speaking, we have three different judgmental equalities in play and one
-could request for a notational difference to signify that fact. For instance,
-we could denote the judgmental equalities of contexts, families and terms by
-$\jdeq_c$, $\jdeq_f$ and $\jdeq_t$ respectively. It will, however, always be
-clear which of the three kinds of judgmental equality is meant when we assert
-a judgmental equality and therefore we shall not bother to make this notational
-distinction.
-
-We note that what we call families over contexts
-here could also have been named dependent contexts or telescopes, see
-\cite{deBruijn1991,hofmann1995extensional}. The term family is in agreement
-with the terminology scheme of \cite{TheBook}, though the reader should be
-warned that the notion of familie means something slightly different there than
-it does here.
-
-The theory of contexts, families and terms shall be described by means of a
-declaration of valid inference rules. An inference rule is an expression of the
-form
-\begin{equation*}
-\inference{\mathcal{H}_1\quad\cdots\quad\mathcal{H}_n}{\mathcal{J}}
-\end{equation*}
-where each of $\mathcal{H}_1$, ..., $\mathcal{H}_n$ and $\mathcal{J}$ is 
-judgment. The judgments $\mathcal{H}_1$, ..., $\mathcal{H}_n$ are called the
-hypotheses and the judgment $\mathcal{J}$ is the conclusion of the inference
-rule. When a judgment is the conclusion of a valid inference rule, we also say
-that the judgment is valid or well-formed. In fact, the only way to construct
-new judgments is via the valid inference rules we describe in this section.
-
-The class of valid inference rules is closed under composition and under
-rearranging the hypotheses. That
-means that if
-\begin{equation*}
-\inference{\mathcal{I}_{1k}\quad\cdots\quad\mathcal{I}_{m_kk}}{\mathcal{H}_k}
-\end{equation*}
-and
-\begin{equation*}
-\inference{\mathcal{H}_1\quad\cdots\quad\mathcal{H}_n}{\mathcal{J}}
-\end{equation*}
-are valid inference rules for $1\leq k\leq n$, then
-\begin{equation*}
-\inference{\mathcal{I}_{11}\quad\cdots\quad\mathcal{I}_{m_11}\quad\cdots\quad\mathcal{I}_{1n}\quad\cdots\quad\mathcal{I}_{m_nn}}{\mathcal{J}}
-\end{equation*}
-is also a valid inference rule. Also, in a valid inference rule 
-duplicated hypotheses may always be contracted and hypotheses may always be 
-reordered; the result is still considered a valid inference rule.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{The basic rules for judgmental equality}
-The rules for judgmental equality establish that it is an equivalence relation
-in all three cases (contexts, families and terms). Thus, the following inference
-rules shall be required to be valid:
-\bgroup\small
-\begin{align*}
-& \inference
-  { \jctx{\Gamma}
-    }
-  { \jctxeq{\Gamma}{\Gamma}
-    } 
-& & \inference
-    { \jctxeq{\Gamma}{\Delta}
-      }
-    { \jctxeq{\Delta}{\Gamma}
-      } 
-& & \inference
-    { \jctxeq{\Gamma}{\Delta}
-      \jctxeq{\Delta}{\greek{E}}
-      }
-    { \jctxeq{\Gamma}{\greek{E}}
-      }
-    \\
-& \inference
-  { \jfam{\Gamma}{A}
-    }
-  { \jfameq{\Gamma}{A}{A}
-    } 
-& & \inference
-    { \jfameq{\Gamma}{A}{B}
-      }
-    { \jfameq{\Gamma}{B}{A}
-      }
-& & \inference
-    { \jfameq{\Gamma}{A}{B}
-      \jfameq{\Gamma}{B}{C}
-      }
-    { \jfameq{\Gamma}{A}{C}
-      }
-    \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    }
-  { \jtermeq{\Gamma}{A}{x}{x}
-    }
-& & \inference
-    { \jtermeq{\Gamma}{A}{x}{y}
-      }
-    { \jtermeq{\Gamma}{A}{y}{x}
-      }
-& & \inference
-    { \jtermeq{\Gamma}{A}{x}{y}
-      \jtermeq{\Gamma}{A}{y}{z}
-      }
-    { \jtermeq{\Gamma}{A}{x}{z}
-      }
-\end{align*}
-\egroup
-
-The following convertibility rules are responsible for the strictness
-of judgmental equality, which sets it apart from equivalences or identifications:
-\begin{align*}
-& \inference
-  { \jctxeq{\Gamma}{\Delta}
-    \jfam{\Gamma}{A}
-    }
-  { \jfam{\Delta}{A}
-    }
-& & \inference
-    { \jctxeq{\Gamma}{\Delta}
-      \jfameq{\Gamma}{A}{B}
-      }
-    { \jfameq{\Delta}{A}{B}
-      }
-    \\
-& \inference
-  { \jctxeq{\Gamma}{\Delta}
-    \jterm{\Gamma}{A}{x}
-    }
-  { \jterm{\Delta}{A}{x}
-    }
-& & \inference
-    { \jctxeq{\Gamma}{\Delta}
-      \jtermeq{\Gamma}{A}{x}{y}
-      }
-    { \jtermeq{\Delta}{A}{x}{y}
-      }
-    \\
-& \inference
-  { \jfameq{\Gamma}{A}{B}
-    \jterm{\Gamma}{A}{x}
-    }
-  { \jterm{\Gamma}{B}{x}
-    }
-& & \inference
-    { \jfameq{\Gamma}{A}{B}
-      \jtermeq{\Gamma}{A}{x}{y}
-      }
-    { \jtermeq{\Gamma}{B}{x}{y}
-      }
-\end{align*}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{The operations of the theory of contexts, families and terms}
-In this subsection we introduce the notions of
-extension on context and on families, the empty context and the empty families, 
-weakening, 
-substitution and identity terms.
-
-Weakening and substitution are both
-operations of the theory of contexts, families and terms and therefore they will
-have three levels of actions: on contexts (at the appropriate level), families
-and terms.
-
-\begin{comment}
-Generally, 
-each operation should
-have an action on contexts, on families and on terms at the `appropriate level'
-and should be compatible with each of the operations we introduce here. In
-that will become more practical after the theory of contexts, families and terms
-is formulated, this means that when we will write down a type theory that
-interprets itself in \autoref{ttmodels}, each of the notions of
-extension, empty family, weakening, substitution and identity terms are going to be internalized
-as terms of the theory and all terms are going to be compatible with these.
-That vague idea will be our guiding light.
-
-So in principle, the introductions of each of the empty context, extension,
-weakening, substitution and identity terms should come in three parts. In 
-practice, they do not all come in three parts.
-
-Apart from the empty context, there will indeed be an empty family and a term
-of the empty family (which will be deemed unique). But extension will
-not have an explicit action `at the level of families' since extension
-is used to define higher families and the action would have to be trivial
-anyway. The empty context and extension provide the essentials of the rest of
-the theory in the sense that higher families are families over extensions,
-contexts are families over the empty context and the `action on families' of
-any operation applied to families over the empty context is the same as the
-original operation. Thus, when we have introduced the operations of weakening
-and substitution we will see that they also have an `action on the terms of 
-contexts'. We will state the relevant judgmental equality rules for the empty
-context in this section rather than in the next, since they are required to
-state that the action on families of an operation is compatible with the
-original operation.
-\end{comment}
-
-We note upfront that both context extension and family extension can also be
-seen as operations of contexts, families and terms, even though we will not
-introduce them as such. The reason is that extension is used to define what it
-means to be a family over a family over a context (namely, a family over an
-extended context). As such, it will have a trivial action on families and terms
-which will not need to be introduced. We will come back to this point in
-\autoref{extension}.
-
-\subsection{Extension}
-\label{extension}
-The operations of extension enable one to consider families over
-families and higher level families alike. We will need both a context extension
-operation and a family extension operation.
-
-The inference rules introducing context and family extension are as follows:
-\begin{align}
-& \inference
-  { \jfam{\Gamma}{A}
-    }
-  { \jctx{\ctxext{\Gamma}{A}}
-    }
-& & \inference
-    { \jctxeq{\Gamma}{\Delta}
-      \jfameq{\Gamma}{A}{B}
-      }
-    { \jctxeq{\ctxext{\Gamma}{A}}{\ctxext{\Delta}{B}}
-      }
-    \\
-& \inference
-  { \jfam{{\Gamma}{A}}{P}
-    }
-  { \jfam{\Gamma}{\ctxext{A}{P}}
-    }
-& & \inference
-    { \jfameq{\Gamma}{A}{B} 
-      \jfameq{{\Gamma}{A}}{P}{Q}
-      }
-    { \jfameq{\Gamma}{\ctxext{A}{P}}{\ctxext{B}{Q}}
-      }
-\end{align}
-The extension operation may also be defined to act on families over families
-and terms thereof. For instance, When $\jfam{{\Gamma}{A}}{P}$ is a family,
-we would get a family $\jfam{{\Gamma}{A}}{{\Gamma}{P}}$ and when
-$\jterm{{\Gamma}{A}}{P}{f}$ is a term, we would get a term
-$\jterm{{\Gamma}{A}}{{\Gamma}{P}}{\ctxext{\Gamma}{f}}$. Then there would be
-axioms stating the judgmental equalities $\ctxext{\Gamma}{P}\jdeq P$ and
-$\ctxext{\Gamma}{f}\jdeq f$. Thus, it is of no actual use to introduce the
-action of context extension on families and terms.
-
-In \autoref{extension-on-terms} we will also define an extension operation
-on terms, which defines a term $\jterm{\Gamma}{{A}{P}}{\tmext{x}{u}}$ for
-every $\jterm{\Gamma}{A}{x}$ and $\jterm{\Gamma}{\subst{x}{P}}{u}$. This
-operation is defined by using the substitution operation and identity terms.
-
-\subsubsection{Associativity of extension}
-\label{comp-ee}
-The inference rules asserting that extension is compatible with itself assert
-that contexts are unstructured lists of type declarations. This rule is
-unavoidable if we want that for a family $A$ in context $\Gamma$, a family over
-$A$ is the same thing as a family over $\ctxext{\Gamma}{A}$. 
-\begin{align}
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{{\Gamma}{A}}{P}
-    }
-  { \jctxeq{\ctxext{{\Gamma}{A}}{P}}{\ctxext{\Gamma}{{A}{P}}}
-    }
-  \\
-& \inference
-  { \jfam{{\Gamma}{A}}{P}
-    \jfam{{{\Gamma}{A}}{P}}{Q}
-    }
-  { \jfameq{\Gamma}{\ctxext{A}{{P}{Q}}}{\ctxext{{A}{P}}{Q}}
-    }
-  \label{comp-ee-c}
-\end{align}
-Note that we may take the point of view that extension distributes over
-extensions. If we want to extend an extended family $\ctxext{P}{Q}$ over
-$\ctxext{\Gamma}{A}$ by the family $A$, the result is the extension of (1)
-the extension of $P$ by $A$ and (2) the extension of $Q$ by $A$. The latter is
-simply $Q$ since extension acts trivially at the level of families (and terms).
-
-\subsection{The empty context and the empty families}
-\label{empty}
-We introduce an empty context and an family over $\Gamma$ for every context $\Gamma$. 
-It has been suggested by some to only include empty families and not an
-empty context because an empty context is not necessary, but we do have several
-reasons to include them. Having an empty context requires also rules asserting 
-that a context is the same thing as a family over the empty context and this
-gives the categorical structure on contexts for free once one has it for 
-families. The main ingredients that will be missing from the theory once an
-empty context is avoided are weakening by a context and identity morphisms from
-a context to itself. Including these by hand also requires to formulate all the
-compatibility rules involving weakening once more for the cases of weakening by
-a context and identity terms at contexts. We prefer to state these rules once
-and only once and including an empty context helps in this respect.
-
-We also prefer to have our scheme of compatibility rules as symmetrical as
-possible. The structure of type dependency should look exactly the same in the
-default case as in any context. In that respect, having an empty family but not
-an empty context seems a bit odd. Also, some sets of compatibility rules 
-(like the rules stating that extension is compatible with the empty families)
-will become assymetrical as a result of not including an empty context. Moreover,
-we would eventually like to include a stratification of the theory by means of
-a type judgment (asserting that a family in a context $\Gamma$ is a type) and
-study closed types (i.e.~types in the empty context). It would be possible to
-provide a notion of closed types without having an empty context, but this would
-have to be formulated separately and we would have to restate all the rules for
-types (if any) for closed types all over again.
-
-One of the main uses of the empty context and the empty families is that we
-get the property that the `action on contexts' of an operation is compatible
-with its `action on families'.
-
-The empty family over a context $\Gamma$ is introduced by the following rule
-inference rule:
-\begin{align}
-& \inference
-  { }
-  { \jctx{\emptyc}
-    }
-  \\
-& \inference
-  { \jctx{\Gamma}
-    }
-  { \jfam{\Gamma}{\emptyf[\Gamma]}
-    }
-  \\
-& \inference
-  { \jctxeq{\Gamma}{\Gamma'}
-    }
-  { \jfameq{\Gamma}{\emptyf[\Gamma]}{\emptyf[\Gamma']}
-    }
-\end{align}
-
-By regarding contexts as families of contexts over the empty context, we
-enable ourselves also to speak of terms of contexts. A term of a context
-$\Gamma$ is a term of the family $\Gamma$ over the empty context. These ideas
-are captured in the following convertibility rules:
-\begin{align}
-& \inference
-  { \jctx{\Gamma}
-    }
-  { \jfam{\emptyc}{\Gamma}
-    } 
-& & \inference
-    { \jfam{\emptyc}{\Gamma}
-      }
-    { \jctx{\Gamma}
-      }
-    \\
-& \inference
-  { \jctxeq{\Gamma}{\Delta}
-    }
-  { \jfameq{\emptyc}{\Gamma}{\Delta}
-    }
-& & \inference
-    { \jfameq{\emptyc}{\Gamma}{\Delta}
-      }
-    { \jctxeq{\Gamma}{\Delta}
-      }
-\end{align}
-
-The reader may wonder whether the empty family $\emptyf$ always has a
-term. This shall follow from the rules stating the compatibility of extension
-with the empty families in \autoref{comp-0e} below and from
-identity terms (\autoref{identityterms}).
-\begin{comment}
-When contexts are viewed as statements, the empty context is the statement that
-asserts nothing. Likewise, in any context one can assert nothing and this is
-achieved by the empty family. 
-Since somebody who states nothing asserts it all right, there is a term of the empty
-family over any context. Moreover, any two terms of the empty family are always
-judgmentally equal.
-\begin{align}
-& \inference
-  { }
-  { \jctx{\emptyc}
-    }
-  \\
-& \inference
-  { \jctx{\Gamma}
-    }
-  { \jfam{\Gamma}{\emptyf[\Gamma]}
-    }
-  \\
-& \inference
-  { \jctx{\Gamma}
-    }
-  { \jterm{\Gamma}{\emptyf[\Gamma]}{\emptytm[\Gamma]}
-    }
-  \\
-& \inference
-  { \jterm{\Gamma}{\emptyf[\Gamma]}{x}
-    }
-  { \jtermeq
-      {\Gamma}
-      {\emptyf[\Gamma]}
-      {x}
-      {\emptytm[\Gamma]}
-    }
-  \label{comp-00-t}
-\end{align}
-
-The empty context and the empty families together with the operation of
-extension will provide several shortcuts for the rest
-of the theory of families and terms. This is (in part) obtained by requiring
-that a context is exactly the same thing as a family of contexts over the
-empty context and that judgmental equality of contexts is exactly the same
-as judgmental equality of families of contexts over the empty context.
-
-\label{comp-00}
-Because every context is a family of contexts over the empty context, 
-the empty context $\emptyc$ may be considered as a family of contexts over the empty
-context. When we do this, we get $\emptyf[\emptyc]$.
-\begin{equation}\label{eq:comp-00}
-\inference
-  { }
-  { \jfameq
-      {\emptyc}
-      {\emptyc}
-      {\emptyf[\emptyc]}
-    }
-\end{equation}
-In the future, we shall denote $\emptyf[\Gamma]$ by $\emptyf$. The above rule
-guarantees that this will not cause confusion. Likewise, we shall denote
-$\emptytm[\Gamma]$ by $\emptytm$.
-\end{comment}
-
-\subsubsection{Compatibility of extension with the empty context and families}
-In the following set of inference rules we state that the empty context and
-the empty family are neutral objects for both context extension (the first two
-rules) and family extension (the last two rules).
-\label{comp-e0}\label{comp-0e}
-\begin{align}
-& \inference
-  { \jctx{\Gamma}
-    }
-  { \jctxeq{\ctxext{\emptyc}{\Gamma}}{\Gamma}
-    }
-  \label{comp-0e-c}
-  \\
-& \inference
-  { \jctx{\Gamma}
-    }
-  { \jctxeq{\ctxext{\Gamma}{\emptyf}}{\Gamma}
-    }
-  \label{comp-e0-c}\\
-& \inference
-  { \jfam{\Gamma}{A}
-    }
-  { \jfameq{\Gamma}{\ctxext{\emptyf}{A}}{A}
-    }
-  \label{comp-0e-f}
-  \\
-& \inference
-  { \jfam{\Gamma}{A}
-    }
-  { \jfameq{\Gamma}{{A}{\emptyf}}{A}
-    }
-  \label{comp-e0-f}
-\end{align}
-
-\subsubsection{Family extension restricted to the empty context}
-Strictly speaking we should have used a different notation for context extension
-as for family extension, because the following rule asserting that family extension
-in the empty context is the same thing as context extension would look tautological
-without a difference. So let us denote, only for the moment, context extension
-of $\Gamma$ by $A$ by $(\ctxext{\Gamma}{A})^c$ and family extension of $A$ by
-$P$ in context $\Gamma$ by $(\ctxext{A}{P})^\famsym$.
-
-Note that we may consider a context $\Gamma$ as a family over $\emptyc$ and
-a family $\jfam{\Gamma}{A}$ as a family $\jfam{{\emptyc}{\Gamma}}{A}$. 
-Therefore we will require the following rule:
-\begin{align}
-& \inference
-  { \jfam{\Gamma}{A}
-    }
-  { \jctxeq{(\ctxext{\Gamma}{A})^c}{(\ctxext{\Gamma}{A})^\famsym}
-    }
-\end{align}
-Note that this rule actually justifies that we have not utilized two different
-notations for context extension and family extension.
-
-\subsection{The theory of weakening}
-\label{weakening}
-When $A$ is a family in context $\Gamma$, the operation of weakening by $A$
-takes a family $B$ in context $\Gamma$ and provides a family $\ctxwk{A}{B}$
-in context $\ctxext{\Gamma}{A}$. The context family $\ctxwk{A}{B}$ can be seen
-as the constant family over $\ctxext{\Gamma}{A}$ with value $B$. This idea will
-be axiomatized in the cancellation property of weakening and substitution in
-\autoref{cancellation-ws}. In \autoref{morphisms} we will take the terms 
-$\unfold{\jhom{\Gamma}{A}{B}{f}}$ to be the morphisms of families from $A$ to 
-$B$. These will be at the heart of the categorical structure of the theory.
-
-The weakening operation acts on three levels: on contexts, on families and
-on terms. The `action on contexts' of weakening is the action we described
-above: it takes a family $B$ over $\Gamma$ to a family $\ctxwk{A}{B}$ over
-$\ctxext{\Gamma}{A}$; the `action on families' of weakening takes a family
-$Q$ over $\ctxext{\Gamma}{B}$ to a family $\ctxwk[\famsym]{A}{Q}$ over
-$\ctxext{{\Gamma}{A}}{\ctxwk{A}{B}}$; the `action on terms' of weakening takes
-a term $g$ of $Q$ to a term $\ctxwk[\tmsym]{A}{g}$ of $\ctxwk[\famsym]{A}{Q}$.
-\begin{align}
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{\Gamma}{B}
-    }
-  { \jfam{{\Gamma}{A}}{\ctxwk{A}{B}}
-    }
-& & \inference
-    { \jfameq{\Gamma}{A}{A'}
-      \jfameq{\Gamma}{B}{B'}
-      }
-    { \jfameq{{\Gamma}{A}}{\ctxwk{A}{B}}{\ctxwk{A'}{B'}}
-      }
-    \\
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{{\Gamma}{B}}{Q}
-    }
-  { \jfam{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk[\famsym]{A}{Q}}
-    }
-& & \inference
-    { \jfameq{\Gamma}{A}{A'}
-      \jfameq{{\Gamma}{B}}{Q}{Q'}
-      }
-    { \jfameq
-        {{{\Gamma}{A}}{\ctxwk{A}{B}}}
-        {\ctxwk[\famsym]{A}{Q}}
-        {\ctxwk[\famsym]{A'}{Q'}}
-      }
-    \\
-%& \inference
-%  { \jterm{\Gamma}{B}{y}
-%    }
-%  { \jterm{{\Gamma}{A}}{\ctxwk{A}{B}}{\ctxwk{A}{y}}
-%    }
-%& & \inference
-%    { \jtermeq{\Gamma}{B}{y}{y'}
-%      }
-%    { \jtermeq
-%        {{\Gamma}{A}}
-%        {\ctxwk{A}{B}}
-%        {\ctxwk{A}{y}}
-%        {\ctxwk{A'}{y'}}
-%      }
-%    \\
-& \inference
-  { \jfam{\Gamma}{A}
-    \jterm{{\Gamma}{B}}{Q}{g}
-    }
-  { \jterm{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk[\famsym]{A}{Q}}{\ctxwk[\tmsym]{A}{g}}
-    }
-& & \inference
-    { \jfameq{\Gamma}{A}{A'}
-      \jtermeq{{\Gamma}{B}}{Q}{g}{g'}
-      }
-    { \jtermeq
-        {{{\Gamma}{A}}{\ctxwk{A}{B}}}
-        {\ctxwk[\famsym]{A}{Q}}
-        {\ctxwk[\tmsym]{A}{g}}
-        {\ctxwk[\tmsym]{A'}{g'}}
-      }
-\end{align}
-
-\subsubsection{Weakening of an empty family}
-The following inference rules express that when the empty family is
-weakened, the result is the empty family.
-\label{comp-w0}
-\begin{align}
-%& \inference
-%  { \jctx{\Gamma}
-%    }
-%  { \jfameq{\Gamma}{\ctxwk{\Gamma}{\emptyc}}{\emptyf}
-%    }
-%  \\
-& \inference
-  { \jfam{\Gamma}{A}
-    }
-  { \jfameq{{\Gamma}{A}}{\ctxwk{A}{\emptyf}}{\emptyf}
-    }
-  \label{comp-w0-c}\\
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{\Gamma}{B}
-    }
-  { \jfameq
-    {{{\Gamma}{A}}{\ctxwk{A}{B}}}
-    {\ctxwk[\famsym]{A}{\emptyf}}
-    {\emptyf}
-    }
-  \label{comp-w0-f}
-\end{align}
-\begin{comment}
-Note that we do not have to axiomatize the action of weakening
-by $A$ on the term of the empty family, because the result will be in the empty 
-family and every term in the empty family is judgmentally equal to the term 
-$\emptytm$.
-\end{comment}
-
-Because a family over $\Gamma$ is the same as a family over 
-$\ctxext{\Gamma}{\emptyf}$ we can apply both the action on contexts and the
-action on families of weakening to a family $B$ over $\Gamma$. When we apply
-the action on families, we obtain a family $\ctxwk[\famsym]{A}{B}$ over the
-context $\ctxext{{\Gamma}{A}}{\ctxwk{A}{\emptyf}}$. However, since we have
-postulated the judgmental equalities $\ctxwk{A}{\emptyf}\jdeq\emptyf$ and
-$\ctxext{{\Gamma}{A}}{\emptyf}\jdeq\ctxext{\Gamma}{A}$, we see that we can
-compare $\ctxwk[\famsym]{A}{B}$ with $\ctxwk{A}{B}$. The following inference
-rule postulates that these two are judgmentally equal:
-\begin{equation}
-\inference
-{ \jfam{\Gamma}{A}
-  \jfam{\Gamma}{B}
-  }
-{ \jfameq{{\Gamma}{A}}{\ctxwk[\famsym]{A}{B}}{\ctxwk{A}{B}}
-  }
-\end{equation}
-Due to this rule, the action on contexts and the action on families of weakening
-are compatible with each other and consequently there can be no possible
-confusion when we omit the annotations $\famsym$ and $\tmsym$ alltogether. In
-the future, the weakening of a family $Q$ over $\ctxext{\Gamma}{B}$ shall
-be denoted just by $\ctxwk{A}{Q}$ and likewise the weakening of a term $g$ of
-$Q$ shall be denoted by $\ctxwk{A}{g}$.
-
-Because a family $B$ over $\Gamma$ can be treated as a family by the operation
-of weakening, weakening also acts on the terms of $B$. The weakening of a term
-$y$ of $B$ by $A$ can be seen as the constant term (or function) of the
-family $\ctxwk{A}{B}$.
-
-\subsubsection{Weakenings of extensions}
-\label{comp-we}
-The following rules assert that when an extended family is weakened, the
-weakening distributes over the extension factors.
-\begin{align}
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{{\Gamma}{B}}{Q}
-    }
-  { \jfameq
-      {\ctxext{\Gamma}{A}}
-      {\ctxwk{A}{\ctxext{B}{Q}}}
-      {\ctxext{\ctxwk{A}{B}}{\ctxwk{A}{Q}}}
-    }
-  \label{comp-we-c}
-  \\
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{{{\Gamma}{B}}{Q}}{R}
-    }
-  { \jfameq
-      {\ctxext{{\Gamma}{A}}{\ctxwk{A}{B}}}
-      {\ctxwk{A}{\ctxext{Q}{R}}}
-      {\ctxext{\ctxwk{A}{Q}}{\ctxwk{A}{R}}}
-    }
-  \label{comp-we-f}
-\end{align}
-When thinking of terms of $\ctxwk{A}{B}$ as morphisms of families from $A$ to
-$B$, this looks already like form of type theoretic choice. It is weaker in that
-it is not stated with function types, yet it is stronger in that it states a
-judgmental equality between two families.
-
-There is also a version of this property where an extended term is weakened.
-This variant is stated and proved in \autoref{comp-we-t}.
-
-\subsubsection{Weakening of weakenings}
-Weakening by a family $A$ in context $\Gamma$ brings things in context $\Gamma$
-to things in context $\ctxext{\Gamma}{A}$. Since we have all the ingredients of
-the theory of contexts, families and terms in the context $\Gamma$ as well it
-has in particular it's own weakening by families. Suppose we have a family
-$\jfam{{\Gamma}{B}}{Q}$ over $B$ in context $\Gamma$. Weakening by $Q$ brings
-things from context $\ctxext{\Gamma}{B}$ to $\ctxext{{\Gamma}{B}}{Q}$. Thus
-we can provide rules asserting what will happen when we first weaken by $Q$ and
-then (via the action on families) by $A$. We will require the following
-inference rules to be valid:
-\label{comp-ww}
-\begin{align}
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{{\Gamma}{B}}{Q}
-    \jfam{{\Gamma}{B}}{R}
-    }
-  { \jfameq
-      {{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{Q}}}
-      {\ctxwk{A}{{Q}{R}}}
-      {\ctxwk{{A}{Q}}{{A}{R}}}
-    }
-  \label{comp-ww-c}\\
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{{\Gamma}{B}}{Q}
-    \jfam{{{\Gamma}{B}}{R}}{S}
-    }
-  { \jfameq
-      {{{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{Q}}}{\ctxwk{A}{{Q}{R}}}}
-      {\ctxwk{A}{{Q}{S}}}
-      {\ctxwk{{A}{Q}}{{A}{S}}}
-    }
-  \label{comp-ww-f}\\
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{{\Gamma}{B}}{Q}
-    \jterm{{{\Gamma}{B}}{R}}{S}{k}
-    }
-  { \jtermeq
-      {{{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{Q}}}{\ctxwk{A}{{Q}{R}}}}
-      {\ctxwk{A}{{Q}{S}}}
-      {\ctxwk{A}{{Q}{k}}}
-      {\ctxwk{{A}{Q}}{{A}{k}}}
-    }
-  \label{comp-ww-t}
-\end{align}
-
-\begin{rmk}
-As an important special case of these inference rules we have the following
-valid inference rules:
-\begin{align*}
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{\Gamma}{B}
-    \jfam{\Gamma}{C}
-    }
-  { \jfameq
-      {{{\Gamma}{A}}{\ctxwk{A}{B}}}
-      {\ctxwk{A}{{B}{C}}}
-      {\ctxwk{{A}{B}}{{A}{C}}}
-    }
-  \\
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{\Gamma}{B}
-    \jfam{{\Gamma}{C}}{R}
-    }
-  { \jfameq
-      {{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{{B}{C}}}}
-      {\ctxwk{A}{{B}{R}}}
-      {\ctxwk{{A}{B}}{{A}{R}}}
-    }
-  \\
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{\Gamma}{B}
-    \jterm{{\Gamma}{C}}{R}{h}
-    }
-  { \jtermeq
-      {{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{{B}{C}}}}
-      {\ctxwk{A}{{B}{R}}}
-      {\ctxwk{A}{{B}{h}}}
-      {\ctxwk{{A}{B}}{{A}{h}}}
-    }
-\end{align*}
-Moreover, we have
-\begin{equation*}
-\inference
-  { \jfam{\Gamma}{A}
-    \jfam{\Gamma}{B}
-    \jterm{\Gamma}{C}{z}
-    }
-  { \jtermeq
-      {{{\Gamma}{A}}{\ctxwk{A}{B}}}
-      {\ctxwk{A}{{B}{C}}}
-      {\ctxwk{A}{{B}{z}}}
-      {\ctxwk{{A}{B}}{{A}{z}}}
-    }
-\end{equation*}
-\end{rmk}
-
-\subsubsection{Weakening by the empty family}
-Note that we can also weaken by the empty family over $\Gamma$.
-Weakening by the empty family $\emptyf$ over a context $\Gamma$ leaves families, 
-their terms, families over those families and terms of those unchanged:
-\label{comp-0w}
-\begin{align}
-& \inference
-  { \jfam{\Gamma}{B}
-    }
-  { \jfameq{\Gamma}{\ctxwk{\emptyf}{B}}{B}
-    }
-  \label{comp-0w-c}
-  \\
-& \inference
-  { \jfam{{\Gamma}{B}}{Q}
-    }
-  { \jfameq{{\Gamma}{B}}{\ctxwk{\emptyf}{Q}}{Q}
-    }
-  \label{comp-0w-f}
-  \\
-& \inference
-  { \jterm{{\Gamma}{B}}{Q}{g}
-    }
-  { \jtermeq{{\Gamma}{B}}{Q}{\ctxwk{\emptyf}{g}}{g}
-    }
-  \label{comp-0w-t}
-\end{align}
-
-\subsubsection{Currying for weakening}
-\label{comp-ew}
-The rules expressing that extension is compatible with weakening assert that
-weakening by an extension is the same thing as weakening twice in the
-appropriate way.
-\begin{align}
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{{\Gamma}{A}}{P}
-    \jfam{\Gamma}{B}
-    }
-  { \jfameq
-      {{{\Gamma}{A}}{P}}
-      {\ctxwk{\ctxext{A}{P}}{B}}
-      {\ctxwk{P}{{A}{B}}}
-    }
-  \label{comp-ew-c}\\
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{{\Gamma}{A}}{P}
-    \jfam{{\Gamma}{B}}{Q}
-    }
-  { \jfameq
-      {{{{\Gamma}{A}}{P}}{\ctxwk{P}{{A}{B}}}}
-      {\ctxwk{\ctxext{A}{P}}{Q}}
-      {\ctxwk{P}{{A}{Q}}}
-    }
-  \label{comp-ew-f}\\
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{{\Gamma}{A}}{P}
-    \jterm{{\Gamma}{B}}{Q}{g}
-    }
-  { \jtermeq
-      {{{{\Gamma}{A}}{P}}{\ctxwk{P}{{A}{B}}}}
-      {\ctxwk{P}{{A}{Q}}}
-      {\ctxwk{\ctxext{A}{P}}{g}}
-      {\ctxwk{P}{{A}{g}}}
-    } 
-  \label{comp-ew-t}
-\end{align}
-
-\begin{comment}
-\subsubsection{Weakening by a context}
-Since we have not required that there is an empty context with the property that
-a context is the same thing as a family in the empty context, we will have to
-explicitly introduce weakening by contexts. Since it will allways be clear
-whether something is weakened by a context or by a family, we shall use the
-same notation for weakening by contexts as for weakening by families. We shall
-have to repeat all the compatibility rules for weakening by families for
-weakening by contexts; there is no way around it. It is
-actually this sort of inconvenience that has originally led us to include an
-empty context and be done with it.
-
-In the following inference rules we assume that we have a context $\Gamma$ and
-that $\jctxeq{\Gamma}{\Gamma'}$ in the rules where $\Gamma'$ makes an appearence.
-\begin{align}
-& \inference
-  { \jctx{\Delta}
-    }
-  { \jfam{\Gamma}{\ctxwk{\Gamma}{\Delta}}
-    }
-& & \inference
-  { \jctxeq{\Delta}{\Delta'}
-    }
-  { \jfameq{\Gamma}{\ctxwk{\Gamma}{\Delta}}{\ctxwk{\Gamma'}{\Delta'}}
-    }
-  \\
-& \inference
-  { \jfam{\Delta}{B}
-    }
-  { \jfam{\ctxext{\Gamma}{\ctxwk{\Gamma}{\Delta}}}{\ctxwk{\Gamma}{B}}
-    }
-& & \inference
-  { \jfameq{\Delta}{B}{B'}
-    }
-  { \jfameq
-      {\ctxext{\Gamma}{\ctxwk{\Gamma}{\Delta}}}
-      {\ctxwk{\Gamma}{B}}
-      {\ctxwk{\Gamma'}{B'}}
-    }
-  \\
-& \inference
-  { \jterm{\Delta}{B}{y}
-    }
-  { \jterm
-      {\ctxext{\Gamma}{\ctxwk{\Gamma}{\Delta}}}
-      {\ctxwk{\Gamma}{B}}
-      {\ctxwk{\Gamma}{y}}
-    }
-& & \inference
-  { \jtermeq{\Delta}{B}{y}{y'}
-    }
-  { \jtermeq
-      {\ctxext{\Gamma}{\ctxwk{\Gamma}{\Delta}}}
-      {\ctxwk{\Gamma}{B}}
-      {\ctxwk{\Gamma}{y}}
-      {\ctxwk{\Gamma}{y'}}
-    }
-\end{align}
-Note that we do not employ three different notations for the three different
-actions of weakening by a context because there can be no confusion about which
-of the three actions is meant. If there were an empty context we should have
-distinguished the three as we did for weakening by a family.
-
-We shall have similar compatibility rules for weakening by a context as we did
-for weakening by a family.
-
-{\color{red}To do. It will look exactly the same as for weakening by families}
-\end{comment}
-
-\subsection{The theory of projections}
-\label{identityterms}
-We formulate the theory of projections on top of the theory of weakening
-described in \autoref{weakening}. In the theory of projections we introduce
-the identity terms. The identity term at a family $A$ in context $\Gamma$ is
-a term $\jterm{{\Gamma}{A}}{\ctxwk{A}{A}}{\idtm{A}}$, so weakening is necessary.
-Together with weakening, the identity terms will provide for the projections.
-The projection from $\cftext{A}{P}$ to $A$ will be the term
-$\jterm{{\Gamma}{{A}{P}}}{\ctxwk{\ctxext{A}{P}}{A}}{\ctxwk{P}{\idtm{A}}}$ and
-will be studied extensively in \autoref{extension-on-terms}.
-
-The following inference rules are required to be valid:
-\begin{align}
-& \inference
-  { \jfam{\Gamma}{A}
-    }
-  { \unfoldall{\jhom{\Gamma}{A}{A}{\idtm{A}}}
-    }
-& & \inference
-    { \jfameq{\Gamma}{A}{A'}
-      }
-    { \unfoldall{\jhomeq{\Gamma}{A}{A}{\idtm{A}}{\idtm{A'}}}
-      }
-\end{align}
-
-\begin{comment}
-We stress again
-that because we have not required there to be an empty context, we need to
-explicitly call for identity terms $\idtm{\Gamma}$ of $\ctxwk{\Gamma}{\Gamma}$
-in context $\Gamma$ for each context $\Gamma$. Thus, we will also require
-the inference rule
-\begin{align}
-& \inference
-  { \jctx{\Gamma}
-    }
-  { \jterm{\Gamma}{\ctxwk{\Gamma}{\Gamma}}{\idtm{\Gamma}}
-    }
-& &
-  \inference
-  { \jctxeq{\Gamma}{\Gamma'}
-    }
-  { \jtermeq{\Gamma}{\ctxwk{\Gamma}{\Gamma}}{\idtm{\Gamma}}{\idtm{\Gamma'}}
-    }
-\end{align}
-For the following inference rules, there should also be versions of identity
-terms at contexts.
-\end{comment}
-
-\subsubsection{The identity term of an extended family}
-The fact that identity terms are compatible with extensions is derived in
-\autoref{comp-ie}.
-
-\subsubsection{The identity term of a weakened family}
-\label{comp-wi}
-The identity term of a weakened family is the weakened identity term:
-\begin{equation}
-\inference
-  { \jfam{\Gamma}{A}
-    \jfam{{\Gamma}{B}}{Q}
-    }
-  { \unfoldall{\jhomeq
-      {{{\Gamma}{A}}{\ctxwk{A}{B}}}
-      {\ctxwk{A}{Q}}
-      {\ctxwk{A}{Q}}
-      {\ctxwk{A}{\idtm{Q}}}
-      {\idtm{\ctxwk{A}{Q}}}
-      }
-    }
-  \label{comp-wi-t}
-\end{equation}
-
-\begin{rmk}
-An important special case is the judgmental equality
-\begin{equation*}
-\unfoldall{\jhomeq
-      {{\Gamma}{A}}
-      {\ctxwk{A}{B}}
-      {\ctxwk{A}{B}}
-      {\ctxwk{A}{\idtm{B}}}
-      {\idtm{\ctxwk{A}{B}}}
-      }
-\end{equation*}
-for a family $\jfam{\Gamma}{B}$.
-\end{rmk}
-
 \subsection{The theory of substitution}
 \label{substitution}
+
 The theory of substitution can be formulated separately from the theory of
@@
 \end{align}
-
-\subsubsection{Fibers of an empty family}
-The following inference rules establish that the fibers of the empty family are 
-the empty families:
-\label{comp-s0}
-\begin{align}
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    }
-  { \jfameq{\Gamma}{\subst{x}{\emptyf}}{\emptyf}
-    }
-  \label{comp-s0-c}
-  \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jfam{{\Gamma}{A}}{P}
-    }
-  { \jfameq
-      {{\Gamma}{\subst{x}{P}}}
-      {\subst{x}{\emptyf}}
-      {\emptyf}
-    }
-  \label{comp-s0-f}
-\end{align}
-\begin{comment}
-Note again that we do not have to axiomatize the action of substitution by $x$ on the term
-of the empty family, because the result will be in the empty family and every
-term in the empty family is judgmentally equal to the term $\emptytm$.
-\end{comment}
-
-We use the above rule to state the compatibility of the action on families of
-substitution with the action on contexts of substitution. Note that a family
-$P$ over $\ctxext{\Gamma}{A}$ may be regarded as a family over
-$\ctxext{{\Gamma}{A}}{\emptyf}$. Thus, we may consider the family
-$\subst[\famsym]{x}{P}$ over $\ctxext{\Gamma}{\subst{x}{\emptyf}}$. Since
-$\subst{x}{\emptyf}$ is judgmentally equal to the empty family, we may compare
-$\subst[\famsym]{x}{P}$ with $\subst{x}{P}$:
-\begin{equation}
-\inference
-{ \jfam{{\Gamma}{A}}{P}
-  }
-{ \jfameq{\Gamma}{\subst[\famsym]{x}{P}}{\subst{x}{P}}
-  }
-\end{equation}
-Due to this rule we need not make the annotations $\famsym$ and $\tmsym$ in
-the notation for substitution anymore and thus we shall omit them from now on.
-Note that because a family $P$ over $\ctxext{\Gamma}{A}$ is eligible for
-application of the action on families of substitution, we may also substitute
-terms of $P$. Thus, given terms $\jterm{\Gamma}{A}{x}$ and $\jterm{{\Gamma}{A}}{P}{f}$,
-we get a term $\jterm{\Gamma}{\subst{x}{P}}{\subst{x}{f}}$, the \emph{value of
-$f$ at $x$}.
-
-In \autoref{morphisms} we will use a combination of weakening and substitution
-to define composition of morphisms of families. However, we have to rely
-on the cancellation rule stated in \autoref{cancellation-ws} before we can
-meaningfully state the definition of composition.
 
@@
 \end{rmk}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Joining the theory of projections with the theory of substitution}
-\label{sec:esystem-equalities}
-
-\begin{comment}
-We have already seen that empty families, extension, weakening, substitution
-and the identity terms are compatible with each other. There are, however, a
-few more judgmental equalities explaining what happens when
-\begin{enumerate}
-\item something is weakend by the empty family. This is used to consider a term
-      $x$ of $A$ in context $\Gamma$ as a morphism from $\emptyf$ to $A$ in
-      context $\Gamma$;
-\item something is weakened by an extended family. This is Currying for weakening;
-\item something is first weakened by a family $A$ in context $\Gamma$ and then
-      substituted by a term of $A$. This is used to show that composition is
-      compatible with itself (i.e. that composition is associative);
-\item the identity terms are joining the party. Identity terms are empty contexts
-      in certain models, so they should behave accordingly.
-\end{enumerate}
-In other words, the rules seem to not fit in the scheme that all operations are
-compatible with each other, but in fact each of these rules plays a role in
-completing the scheme (work out further).
-\end{comment}
-
-\begin{comment}
-In this section we extend the judgmental equality relation in such a way that
-the empty context, extension, weakening, substitution and the identity terms
-all become compatible with each other. Since we have five operations, there
-are going to be $25$ finite sets of rules explaining how each operation is
-compatible with each of the operations.
-
-In addition to these $25$ finite sets
-of compatibility rules, there are also two sets of rules explaining the
-cancellation properties of weakening and substitution and of the identity terms.
-
-A typical compatibility axiom captures the idea that a diagram like
-\begin{equation*}
-\begin{tikzcd}[column sep=huge]
-\jfam{\Gamma}{\blank} 
-  \ar{r}{C\mapsto\ctxwk{B}{C}} 
-  \ar{d}[swap]{C\mapsto\ctxwk{A}{C}} 
-& \jfam{{\Gamma}{B}}{\blank} 
-  \ar{d}{Q\mapsto\ctxwk{A}{Q}}
-  \\
-\jfam{{\Gamma}{A}}{\blank} 
-  \ar{r}[swap]{P\mapsto\ctxwk{{A}{B}}{P}} 
-& \jfam{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\blank}
-\end{tikzcd}
-\end{equation*}
-in the case of two consecutive weakenings, or
-\begin{equation*}
-\begin{tikzcd}[column sep=huge]
-\jfam{{\Gamma}{A}}{\blank} 
-  \ar{d}[swap]{Q\mapsto\subst{x}{Q}} 
-  \ar{r}{Q\mapsto\ctxwk{P}{Q}} 
-& \jfam{{{\Gamma}{A}}{P}}{\blank} 
-    \ar{d}{R\mapsto\subst{x}{R}}
-  \\ 
-\jfam{\Gamma}{\blank} 
-  \ar{r}[swap]{B\mapsto\ctxwk{\subst{x}{P}}{B}} 
-& \jfam{{\Gamma}{\subst{x}{P}}}{\blank}
-\end{tikzcd}
-\end{equation*}
-in the case of a combination of weakening and substitution, commutes. We won't
-actually draw any of these diagrams in this section, but readers are encouraged
-to do so if they think it can help them understanding the axioms. Another thing
-one could keep in mind with the axioms that involve weakening is that all of
-them become completely obvious when weakening is made notationally invisible.
-Of course, we have an explicit notation for weakening on purpuse, but it is the
-usual practice of type theorists to make weakening invisible.
-
-The table below indicates where
-we can find the axioms of each of the combinations of operators. The reference
-at the row of operation $X$ and the column of operation $Y$ refers to the
-axiom which explains $Y$ distributes over the arguments of $X$ when $Y$ is
-applied after $X$. The empty context has no arguments, whereas extension,
-and weakening have two arguments which are both of the family-kind, substitution
-has two arguments of which one is a term and the other is a family and
-idenitity terms have one argument, a family.
-\begin{center}
-\begin{tabular}{r|rrrr}
-& ext & wk & subst & id\\
-\hline
-extension & \autoref{comp-ee}* &  & \cmark & \cmark \\
-weakening & \autoref{comp-we}* & \autoref{comp-ww} & \autoref{comp-sw} & \autoref{comp-wi} \\
-substitution & \autoref{comp-se}* & \autoref{comp-ws} & \autoref{comp-ss} & \autoref{comp-si} \\
-identity terms & \cmark & \autoref{comp-wi} & \autoref{comp-si} & \xmark
-\end{tabular}
-\end{center}
-Apart from references to sections where axioms are given, we also see checkmarks,
-crosses, marked references and duplicated references in this table. 
-The checkmarks indicate that the desired property
-can be proved and need not be axiomatized. To clarify those:
-\begin{itemize}
-\item The identity term of the empty context is the unique term of the empty
-family. Recall that we already have axiomatized in \autoref{empty} that every
-term of the empty family is judgmentally equal to $\emptytm$.
-\item The identity term of an extension can be explained by means of the notion
-of \emph{extension on terms}. This notion is defined in \autoref{extension-on-terms}
-using only identity terms and substitution. With extension on terms we can
-combine a term $\jterm{\Gamma}{A}{x}$ and a term $\jterm{\Gamma}{\subst{x}{P}}{u}$
-into a term $\jterm{\Gamma}{{A}{P}}{\tmext{x}{u}}$, which may be seen as the
-pair consisting of $x$ and $u$. Since it is a defined notion, we will
-be able to derive lots of properties for extension of terms. One of them explains
-how we can see the identity term of an extension as the extension of two certain
-terms, see \autoref{comp-ie}.
-\item Using extension on terms we can also make sense of `fibers at an extension'
-because that would exactly be the fiber at an extended term. We will see in
-\autoref{comp-es} that, for instance, $\subst{\tmext{x}{u}}{Q}\jdeq\subst{u}{{x}{Q}}$
-for any family $\jfam{{{\Gamma}{A}}{P}}{Q}$ and any two terms $\jterm{\Gamma}{A}{x}$
-and $\jterm{\Gamma}{\subst{x}{P}}{u}$. That's a form of currying.
-\end{itemize}
-The spots with a \xmark simply make no sense 
-{\color{red}(or do they, when the term model
-is better understood (by me)?)}. The marked references point to
-axioms which \emph{do} have a variant for terms, but those require the notion
-of extension on terms, can be proved and therefore need not be included in our
-list of compatibility axioms. In each of these sections, we give a forward
-reference to the lemma in which this variant is stated and proved.
-\end{comment}
-
-\subsubsection{Fibers of weakenings}\label{comp-sw}
-The following rules assert what happens when we first weaken by a family
-$Q$ over $P$ in context $\ctxext{\Gamma}{A}$ and then substitute by a term
-$x$ of $A$:
-\begin{align}
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jfam{{{\Gamma}{A}}{P}}{Q}
-    \jfam{{{\Gamma}{A}}{P}}{R}
-    }
-  { \jfameq
-      {{{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}}
-      {\subst{x}{\ctxwk{Q}{R}}}
-      {\ctxwk{\subst{x}{Q}}{\subst{x}{R}}}
-    }
-  \label{comp-sw-c}
-  \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jfam{{{\Gamma}{A}}{P}}{Q}
-    \jfam{{{{\Gamma}{A}}{P}}{R}}{S}
-    }
-  { \jfameq
-      {{{{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}}{\subst{x}{\ctxwk{Q}{R}}}}
-      {\subst{x}{\ctxwk{Q}{S}}}
-      {\ctxwk{\subst{x}{Q}}{\subst{x}{S}}}
-    }
-  \label{comp-sw-f}
-  \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jfam{{{\Gamma}{A}}{P}}{Q}
-    \jterm{{{{\Gamma}{A}}{P}}{R}}{S}{k}
-    }
-  { \jtermeq
-      {{{{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}}{\subst{x}{\ctxwk{Q}{R}}}}
-      {\subst{x}{\ctxwk{Q}{S}}}
-      {\subst{x}{\ctxwk{Q}{k}}}
-      {\ctxwk{\subst{x}{Q}}{\subst{x}{k}}}
-    }
-  \label{comp-sw-t}
-\end{align}
-
-\begin{rmk}
-As an important special case of these inference rules we have the following
-valid inference rules:
-\begin{align*}
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jfam{{\Gamma}{A}}{P}
-    \jfam{{\Gamma}{A}}{Q}
-    }
-  { \jfameq
-      {{\Gamma}{\subst{x}{P}}}
-      {\subst{x}{\ctxwk{P}{Q}}}
-      {\ctxwk{\subst{x}{P}}{\subst{x}{Q}}}
-    }
-  \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jfam{{\Gamma}{A}}{P}
-    \jfam{{{\Gamma}{A}}{Q}}{R}
-    }
-  { \jfameq
-      {{{\Gamma}{\subst{x}{P}}}{\subst{x}{\ctxwk{P}{Q}}}}
-      {\subst{x}{\ctxwk{P}{R}}}
-      {\ctxwk{\subst{x}{P}}{\subst{x}{R}}}
-    }
-  \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jfam{{\Gamma}{A}}{P}
-    \jterm{{{\Gamma}{A}}{Q}}{R}{h}
-    }
-  { \jtermeq
-      {{{\Gamma}{\subst{x}{P}}}{\subst{x}{\ctxwk{P}{Q}}}}
-      {\subst{x}{\ctxwk{P}{R}}}
-      {\subst{x}{\ctxwk{P}{h}}}
-      {\ctxwk{\subst{x}{P}}{\subst{x}{h}}}
-    }
-\end{align*}
-Moreover, we get
-\begin{equation*}
-\inference
-  { \jterm{\Gamma}{A}{x}
-    \jfam{{\Gamma}{A}}{P}
-    \jterm{{\Gamma}{A}}{Q}{g}
-    }
-  { \jtermeq
-      {{\Gamma}{\subst{x}{P}}}
-      {\subst{x}{\ctxwk{P}{Q}}}
-      {\subst{x}{\ctxwk{P}{g}}}
-      {\ctxwk{\subst{x}{P}}{\subst{x}{g}}}
-    }
-\end{equation*}
-\end{rmk}
-
-\subsubsection{Weakenings of fibers}
-\label{comp-ws}
-The following inference rules explain what happens when we first weaken by a
-term $g$ of $Q$ in context $\ctxext{\Gamma}{B}$ and then weaken by a family
-$A$ over $\Gamma$.
-\begin{align}
-& \inference
-  { \jfam{\Gamma}{A}
-    \jterm{{\Gamma}{B}}{Q}{g}
-    \jfam{{{\Gamma}{B}}{Q}}{R}
-    }
-  { \jfameq
-      {{{\Gamma}{A}}{\ctxwk{A}{B}}}
-      {\ctxwk{A}{\subst{g}{R}}}
-      {\subst{\ctxwk{A}{g}}{\ctxwk{A}{R}}}
-    }
-  \label{comp-ws-c}
-  \\
-& \inference
-  { \jfam{\Gamma}{A}
-    \jterm{{\Gamma}{B}}{Q}{g}
-    \jfam{{{{\Gamma}{B}}{Q}}{R}}{S}
-    }
-  { \jfameq
-      {{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{\subst{g}{R}}}}
-      {\ctxwk{A}{\subst{g}{S}}}
-      {\subst{\ctxwk{A}{g}}{\ctxwk{A}{S}}}
-    }
-  \label{comp-ws-f}
-  \\
-& \inference
-  { \jfam{\Gamma}{A}
-    \jterm{{\Gamma}{B}}{Q}{g}
-    \jterm{{{{\Gamma}{B}}{Q}}{R}}{S}{k}
-    }
-  { \jtermeq
-      {{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{\subst{g}{R}}}}
-      {\ctxwk{A}{\subst{g}{S}}}
-      {\ctxwk{A}{\subst{g}{k}}}
-      {\subst{\ctxwk{A}{g}}{\ctxwk{A}{k}}}
-    }
-  \label{comp-ws-t}
-\end{align}
-
-\begin{rmk}
-As an important special case of these inference rules we have the following 
-valid inference rules:
-\begin{align*}
-& \inference
-  { \jfam{\Gamma}{A}
-    \jterm{\Gamma}{B}{y}
-    \jfam{{\Gamma}{B}}{Q}
-    }
-  { \jfameq
-      {{\Gamma}{A}}
-      {\ctxwk{A}{\subst{y}{Q}}}
-      {\subst{\ctxwk{A}{y}}{\ctxwk{A}{Q}}}
-    }
-  \\
-& \inference
-  { \jfam{\Gamma}{A}
-    \jterm{\Gamma}{B}{y}
-    \jfam{{{\Gamma}{B}}{Q}}{R}
-    }
-  { \jfameq
-      {{{\Gamma}{A}}{\ctxwk{A}{\subst{y}{Q}}}}
-      {\ctxwk{A}{\subst{y}{R}}}
-      {\subst{\ctxwk{A}{y}}{\ctxwk{A}{R}}}
-    }
-  \\
-& \inference
-  { \jfam{\Gamma}{A}
-    \jterm{\Gamma}{B}{y}
-    \jterm{{{\Gamma}{B}}{Q}}{R}{h}
-    }
-  { \jtermeq
-      {{{\Gamma}{A}}{\ctxwk{A}{\subst{y}{Q}}}}
-      {\ctxwk{A}{\subst{y}{R}}}
-      {\ctxwk{A}{\subst{y}{h}}}
-      {\subst{\ctxwk{A}{y}}{\ctxwk{A}{h}}}
-    }
-\end{align*}
-Moreover, we get
-\begin{equation*}
-\inference
-  { \jfam{\Gamma}{A}
-    \jterm{\Gamma}{B}{y}
-    \jterm{{\Gamma}{B}}{Q}{g}
-    }
-  { \jtermeq
-      {{\Gamma}{A}}
-      {\ctxwk{A}{\subst{y}{Q}}}
-      {\ctxwk{A}{\subst{y}{g}}}
-      {\subst{\ctxwk{A}{y}}{\ctxwk{A}{g}}}
-    }
-\end{equation*}
-\end{rmk}
-
-\subsubsection{The cancellation property of weakening and substitution}
-\label{cancellation-ws}
-The judgmental equalities we're about to describe assert that substituting a term
-in the weakening a thing gives you the thing back. In the case of contexts we get that each fiber
-$\subst{x}{\ctxwk{A}{B}}$ is just $B$ and in the case of terms we get 
-that $\ctxwk{A}{y}$ is the constant function
-mapping everything to $y:B$. Thus, these rules actually establish the weakening
-as the weakening. After stating the rules we will describe what it means to
-compose context morphisms (terms of weakened contexts).
-\begin{align}
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{\Gamma}{B}
-    \jterm{\Gamma}{A}{x}
-    }
-  { \jfameq
-      {\Gamma}
-      {\subst{x}{\ctxwk{A}{B}}}
-      {B}
-    }
-  \label{cancellation-ws-c}\\
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{{\Gamma}{B}}{Q}
-    \jterm{\Gamma}{A}{x}
-    }
-  { \jfameq{{\Gamma}{B}}{\subst{x}{\ctxwk{A}{Q}}}{Q}
-    }
-  \label{cancellation-ws-f}\\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jterm{{\Gamma}{B}}{Q}{g}
-    }
-  { \jtermeq{{\Gamma}{B}}{Q}{\subst{x}{\ctxwk{A}{g}}}{g}
-    }
-  \label{cancellation-ws-t}
-\end{align}
-
-\subsubsection{The identity term of a substituted family}
-\label{comp-si}
-The identity term of a substituted family is the substitution of the identity term
-\begin{equation}
-\inference
-  { \jterm{\Gamma}{A}{x}
-    \jfam{{{\Gamma}{A}}{P}}{Q}
-    }
-  { \unfoldall{\jhomeq
-      {{\Gamma}{\subst{x}{P}}}
-      {\subst{x}{Q}}
-      {\subst{x}{Q}}
-      {\subst{x}{\idtm{Q}}}
-      {\idtm{\subst{x}{Q}}}
-    }}
-  \label{comp-si-t}
-\end{equation}
-
-\begin{rmk}
-An important special case is the judgmental equality
-\begin{equation*}
-\unfoldall{\jhomeq
-      {\Gamma}
-      {\subst{x}{P}}
-      {\subst{x}{P}}
-      {\subst{x}{\idtm{P}}}
-      {\idtm{\subst{x}{P}}}
-    }
-\end{equation*}
-for a family $\jfam{{\Gamma}{A}}{P}$.
-\end{rmk}
-
-\subsubsection{The cancellation properties of identity terms}
-\label{cancellation-i}
-Identity terms are determined by their behavior with respect to substitution combined with
-weakening. The identity terms will also be subject to compatibility rules.
-\begin{align}
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    }
-  { \jtermeq{\Gamma}{A}{\subst{x}{\idtm{A}}}{x}
-    }
-  \label{cancellation-si}\\
-& \inference
-  { \jfam{{\Gamma}{A}}{P}
-    }
-  { \jfameq{{\Gamma}{A}}{\subst{\idtm{A}}{\ctxwk{A}{P}}}{P}
-    }
-  \label{precomp-idtm-c}
-  \\
-& \inference
-  { \jfam{{{\Gamma}{A}}{P}}{Q}
-    }
-  { \jfameq{{{\Gamma}{A}}{P}}{\subst{\idtm{A}}{\ctxwk{A}{Q}}}{Q}
-    }
-  \label{precomp-idtm-f}
-  \\
-& \inference
-  { \jterm{{{\Gamma}{A}}{P}}{Q}{g}
-    }
-  { \jtermeq
-      {{{\Gamma}{A}}{P}}
-      {Q}
-      {\subst{\idtm{A}}{\ctxwk{A}{g}}}
-      {g}
-    }
-  \label{precomp-idtm-t}
-\end{align}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{The possiblity of types in the theory of contexts, families and
-terms}
-\label{types}
-We have deliberately not spoken of types so far because we have taken the point
-of view that a type in a context is nothing but a family in that context which
-belongs to the class of types. We think of types as \emph{irreducible} families,
-i.e.\ families which are neither empty nor the extension of two
-families which are both not empty (in algebraic terminology: which
-are both non-trivial). To allow ourselves to speak of types we introduce two
-new judgments: the judgment that something is a type and the judgment that two
-types are equal.
-\begin{align*}
-\jalign\jtype{\Gamma}{A} 
-& \jalign\jtypeeq{\Gamma}{A}{B}
-\end{align*}
-But only families of contexts are eligible to be types. If $A$ is a type
-in context $\Gamma$, then $A$ is also a family of contexts over $\Gamma$. 
-Moreover, two types in context $\Gamma$ are judgmentally equal precisely when they are equal
-as context families and if a family $B$ of contexts over $\Gamma$ is
-judgmentally equal to a type $A$ in context $\Gamma$, then $B$ is a type in
-context $\Gamma$. This is expressed by the following four inference rules:
-\begin{align*}
-& \inference
-  { \jtype{\Gamma}{A}
-    }
-  { \jfam{\Gamma}{A}
-    }
-& & \inference
-    { \jtypeeq{\Gamma}{A}{B}
-      }
-    { \jfameq{\Gamma}{A}{B}
-      }
-    \\
-& \inference
-  { \jtype{\Gamma}{A}
-    \jfameq{\Gamma}{A}{B}
-    }
-  { \jtype{\Gamma}{B}
-    }
-& & \inference
-    { \jtype{\Gamma}{A}
-      \jfameq{\Gamma}{A}{B}
-      }
-    { \jtypeeq{\Gamma}{A}{B}
-      }
-\end{align*}
-As pointed out at the beginning of this subsection, 
-we do not assume that the empty family is a type, that would be like
-assuming that the multiplicative unit of a ring is prime.
-
-\begin{comment}
-We envision a
-categorical interpretation of the theory of contexts, families and terms where 
-contexts are interpreted as finite paths of arrows belonging to a predetermined
-class. The empty family would be such a path of length $0$ whereas the types
-would be such paths of length $1$.
-\end{comment}
-
-We add rules asserting that a weakened type is again a type that 
-substitution preserves the property of being a type:
-\begin{align}
-& \inference
-  { \jfam{\Gamma}{B}
-    \jtype{\ctxext{\Gamma}{B}}{Q}
-    }
-  { \jtype{\ctxext{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{Q}}
-    }
-  \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jtype{\ctxext{{\Gamma}{A}}{P}}{Q}
-    }
-  { \jtype{\ctxext{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}
-    }
-\end{align}
-With only the current rules, the possibility of making the judgment that
-something is a type does not add much to the theory of contexts, families and
-terms. Nevertheless, when studying models, having an interpretation for the
-judgment that something is a type will allow for the possibility to study
-conditions such as the one asserting that every family `factorizes' uniquely
-as multiple applications of extension to types, analogous to the condition on
-unique factorization domains in ring theory. All the tradional models of type
-theory should translate to models with such a condition, simply because
-traditionally contexts are viewed as lists of variable (and type) declarations.

### 2014-10-28 Rules/rules_weakening.tex
reorganization

--- 2014-09-20 article-models-tt.tex
@@
-\section{The generalized algebraic theory of contexts, families and terms (the E-system)}\label{tt}
-In this section we give a description of dependent type theory before type
-constructors. Apart from contexts, families and terms -- which provide for the
-core of the language of dependent logic -- the basic ingredients
-of this theory will be the operations of extension on context and on families, 
-and empty context and empty families, weakening, substitution
-and the identity terms. The resulting theory can be seen as a manifestation of the 
-structure underlying dependent type theory.
-
-We will
-formulate the theory of contexts, families and terms in such a way that contexts aren't defined
-to be lists of variable declarations. The variable-free (a.k.a.~\emph{name-free}) approach 
-we take here is rather different than those appearing in 
-\cite{hofmann1995extensional,TheBook} but it has appeared in the work of Coquand
-and in \cite{Dybjer1996}.
-The main reason we don't let variable declarations in is that we don't see them 
-in the internal models either. This way we also set out to a more algebraic 
-approach of type theory and higher category theory. Thirdly, we will not have to
-be burdened with superficial comments about variables being bounded or not, or 
-fresh or free or not occuring at all.
-
-In the current presentation, contexts can be seen as binary
-planar trees of which the leaves are (families of) contexts. 
-The judgmental equality relation on contexts is an equivalence relation which 
-expresses that binary planar
-trees of contexts are judgmentally equal if their leaves are, taking only
-(the isomorphism class of) 
-the order of the leaves into account \emph{and not the actual shape of the three}.
-The intuition behind this equivalence relation is indeed that unstructured
-(i.e.~unbracketed) lists such as the lists of variable declarations which
-usually appear in type theoretical syntax, may be regarded as contexts.
-
-Besides contexts, families and terms there will also be a notion of `type in
-a context', which one may assume.
-A context is eligible to be a \emph{closed type} and over a context, a family
-may be a type. We have the intuition that `being a type'
-expresses the property of `being atomic or irreducible'. With a notion of types,
-one has the means to require a stratification by which every context is a finite
-extension of types starting at a closed type, and every family is likewise a
-finite extension of types. The idea behind this is to recover the theory of
-\cite{Garner2014} {\color{red}(and possibly B-systems?)}. We stress that
-we will not include axioms asserting such a stratification in the E-system
-and we leave discussion on the possibility of a typing judgment until the
-end of this section.
-
-With families of contexts being the principal things of
-study, we have added a bit more generality to the theory, compared to the
-theory of categories with families presented in \cite{Dybjer1996}. Here are
-a couple of reasons for this extra generality:
-\begin{enumerate}
-\item We want to be very sure that we can interpret all the rules of type theory
-in a straightforward manner in any context.
-\item We do want do end up with a categorical structure on the contexts, derived
-from the theory by requiring that a morphism from $\Gamma$ to $\Delta$ is a
-term of the weakening of $\Delta$ by $\Gamma$ in context $\Gamma$. Thus, we
-must be able to lift the context $\Delta$ to become a family of contexts over
-$\Gamma$ and it is not reasonable to expect that this results in a type.
-\item It is very easy in this setting to assert that an operation has an action
-on very high levels of families, i.e.~when $F$ is an operation taking things in
-context $\Gamma$ to things in context $\Delta$, then it takes a family $A$ in
-context $\Gamma$ to $F(A)$ in context $\Delta$; a family $P$ in context
-$\ctxext{\Gamma}{A}$ to a family $F(P)$ in context $\ctxext{\Delta}{(F(A))}$,
-and so on. The reason it becomes simple is that we don't have to explain what
-`and so on' means.
-\end{enumerate}
-
-We will formulate fairly strict rules governing the judgmental equalities,
-expressing that extension, weakening and substitution are combatible with
-each other in a judgmental manner. This does not, however, diminish the role
-of isomorphisms or of homotopies could play in the theory once identity types
-are added. Indeed, types could still have non-trivial identity relations and
-the category of types in a certain context could genuinely display higher
-categorical structure, or so we conjecture.
-
-Much of the rules we state are just compatibility rules of extension, weakening
-and substitution with each other. In a way, these rules assert that our contexts
-are just structureless lists of contexts and that likewise terms are structureless
-lists of terms. They are structureless in the sense that the order in which
-they are formed by pairing up is irrelevant. We note that this causes complications
-in the traditional way that categorical sematics of type theory is implemented,
-where contexts become objects of the category which is supposed to model type
-theory. The reason for this is that context extension will not satisfy all the
-compatibility rules we're about to state. The first step to resolving this is taking
-the types in the empty context as the objects.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Judgments and valid inference rules}\label{judgments}
-The theory we describe here is a theory of contexts, families of
-contexts and terms thereof. The families of contexts are by some authors called
-dependent contexts, but they are handled a bit differently here because they
-become the primary object of study. Dependent contexts can be types; they could
-be seen as atomic or indecomposable dependent contexts.
-
-Thus we make eight kinds of judgments: ``$\Gamma$ is a context'',
-``$A$ is a family of contexts over $\Gamma$'', ``$A$ is a type in context $\Gamma$''
-and ``$x$ is a term of the family $A$ of contexts over $\Gamma$''. The other four
-judgments are for judgmental equality.
-\begin{align*}
-\jalign\jctx{\Gamma} 
-& \jalign\jctxeq{\Gamma}{\Gamma'}
-  \\
-\jalign\jfam{\Gamma}{A} 
-& \jalign\jfameq{\Gamma}{A}{B}
-  \\
-\jalign\jterm{\Gamma}{A}{x} 
-& \jalign\jtermeq{\Gamma}{A}{x}{y}.
-\end{align*}
-
-Strictly speaking, we have three different judgmental equalities in play and one
-could request for a notational difference to signify that fact. For instance,
-we could denote the judgmental equalities of contexts, families and terms by
-$\jdeq_c$, $\jdeq_f$ and $\jdeq_t$ respectively. It will, however, always be
-clear which of the three kinds of judgmental equality is meant when we assert
-a judgmental equality and therefore we shall not bother to make this notational
-distinction.
-
-We note that what we call families over contexts
-here could also have been named dependent contexts or telescopes, see
-\cite{deBruijn1991,hofmann1995extensional}. The term family is in agreement
-with the terminology scheme of \cite{TheBook}, though the reader should be
-warned that the notion of familie means something slightly different there than
-it does here.
-
-The theory of contexts, families and terms shall be described by means of a
-declaration of valid inference rules. An inference rule is an expression of the
-form
-\begin{equation*}
-\inference{\mathcal{H}_1\quad\cdots\quad\mathcal{H}_n}{\mathcal{J}}
-\end{equation*}
-where each of $\mathcal{H}_1$, ..., $\mathcal{H}_n$ and $\mathcal{J}$ is 
-judgment. The judgments $\mathcal{H}_1$, ..., $\mathcal{H}_n$ are called the
-hypotheses and the judgment $\mathcal{J}$ is the conclusion of the inference
-rule. When a judgment is the conclusion of a valid inference rule, we also say
-that the judgment is valid or well-formed. In fact, the only way to construct
-new judgments is via the valid inference rules we describe in this section.
-
-The class of valid inference rules is closed under composition and under
-rearranging the hypotheses. That
-means that if
-\begin{equation*}
-\inference{\mathcal{I}_{1k}\quad\cdots\quad\mathcal{I}_{m_kk}}{\mathcal{H}_k}
-\end{equation*}
-and
-\begin{equation*}
-\inference{\mathcal{H}_1\quad\cdots\quad\mathcal{H}_n}{\mathcal{J}}
-\end{equation*}
-are valid inference rules for $1\leq k\leq n$, then
-\begin{equation*}
-\inference{\mathcal{I}_{11}\quad\cdots\quad\mathcal{I}_{m_11}\quad\cdots\quad\mathcal{I}_{1n}\quad\cdots\quad\mathcal{I}_{m_nn}}{\mathcal{J}}
-\end{equation*}
-is also a valid inference rule. Also, in a valid inference rule 
-duplicated hypotheses may always be contracted and hypotheses may always be 
-reordered; the result is still considered a valid inference rule.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{The basic rules for judgmental equality}
-The rules for judgmental equality establish that it is an equivalence relation
-in all three cases (contexts, families and terms). Thus, the following inference
-rules shall be required to be valid:
-\bgroup\small
-\begin{align*}
-& \inference
-  { \jctx{\Gamma}
-    }
-  { \jctxeq{\Gamma}{\Gamma}
-    } 
-& & \inference
-    { \jctxeq{\Gamma}{\Delta}
-      }
-    { \jctxeq{\Delta}{\Gamma}
-      } 
-& & \inference
-    { \jctxeq{\Gamma}{\Delta}
-      \jctxeq{\Delta}{\greek{E}}
-      }
-    { \jctxeq{\Gamma}{\greek{E}}
-      }
-    \\
-& \inference
-  { \jfam{\Gamma}{A}
-    }
-  { \jfameq{\Gamma}{A}{A}
-    } 
-& & \inference
-    { \jfameq{\Gamma}{A}{B}
-      }
-    { \jfameq{\Gamma}{B}{A}
-      }
-& & \inference
-    { \jfameq{\Gamma}{A}{B}
-      \jfameq{\Gamma}{B}{C}
-      }
-    { \jfameq{\Gamma}{A}{C}
-      }
-    \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    }
-  { \jtermeq{\Gamma}{A}{x}{x}
-    }
-& & \inference
-    { \jtermeq{\Gamma}{A}{x}{y}
-      }
-    { \jtermeq{\Gamma}{A}{y}{x}
-      }
-& & \inference
-    { \jtermeq{\Gamma}{A}{x}{y}
-      \jtermeq{\Gamma}{A}{y}{z}
-      }
-    { \jtermeq{\Gamma}{A}{x}{z}
-      }
-\end{align*}
-\egroup
-
-The following convertibility rules are responsible for the strictness
-of judgmental equality, which sets it apart from equivalences or identifications:
-\begin{align*}
-& \inference
-  { \jctxeq{\Gamma}{\Delta}
-    \jfam{\Gamma}{A}
-    }
-  { \jfam{\Delta}{A}
-    }
-& & \inference
-    { \jctxeq{\Gamma}{\Delta}
-      \jfameq{\Gamma}{A}{B}
-      }
-    { \jfameq{\Delta}{A}{B}
-      }
-    \\
-& \inference
-  { \jctxeq{\Gamma}{\Delta}
-    \jterm{\Gamma}{A}{x}
-    }
-  { \jterm{\Delta}{A}{x}
-    }
-& & \inference
-    { \jctxeq{\Gamma}{\Delta}
-      \jtermeq{\Gamma}{A}{x}{y}
-      }
-    { \jtermeq{\Delta}{A}{x}{y}
-      }
-    \\
-& \inference
-  { \jfameq{\Gamma}{A}{B}
-    \jterm{\Gamma}{A}{x}
-    }
-  { \jterm{\Gamma}{B}{x}
-    }
-& & \inference
-    { \jfameq{\Gamma}{A}{B}
-      \jtermeq{\Gamma}{A}{x}{y}
-      }
-    { \jtermeq{\Gamma}{B}{x}{y}
-      }
-\end{align*}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{The operations of the theory of contexts, families and terms}
-In this subsection we introduce the notions of
-extension on context and on families, the empty context and the empty families, 
-weakening, 
-substitution and identity terms.
-
-Weakening and substitution are both
-operations of the theory of contexts, families and terms and therefore they will
-have three levels of actions: on contexts (at the appropriate level), families
-and terms.
-
-\begin{comment}
-Generally, 
-each operation should
-have an action on contexts, on families and on terms at the `appropriate level'
-and should be compatible with each of the operations we introduce here. In
-that will become more practical after the theory of contexts, families and terms
-is formulated, this means that when we will write down a type theory that
-interprets itself in \autoref{ttmodels}, each of the notions of
-extension, empty family, weakening, substitution and identity terms are going to be internalized
-as terms of the theory and all terms are going to be compatible with these.
-That vague idea will be our guiding light.
-
-So in principle, the introductions of each of the empty context, extension,
-weakening, substitution and identity terms should come in three parts. In 
-practice, they do not all come in three parts.
-
-Apart from the empty context, there will indeed be an empty family and a term
-of the empty family (which will be deemed unique). But extension will
-not have an explicit action `at the level of families' since extension
-is used to define higher families and the action would have to be trivial
-anyway. The empty context and extension provide the essentials of the rest of
-the theory in the sense that higher families are families over extensions,
-contexts are families over the empty context and the `action on families' of
-any operation applied to families over the empty context is the same as the
-original operation. Thus, when we have introduced the operations of weakening
-and substitution we will see that they also have an `action on the terms of 
-contexts'. We will state the relevant judgmental equality rules for the empty
-context in this section rather than in the next, since they are required to
-state that the action on families of an operation is compatible with the
-original operation.
-\end{comment}
-
-We note upfront that both context extension and family extension can also be
-seen as operations of contexts, families and terms, even though we will not
-introduce them as such. The reason is that extension is used to define what it
-means to be a family over a family over a context (namely, a family over an
-extended context). As such, it will have a trivial action on families and terms
-which will not need to be introduced. We will come back to this point in
-\autoref{extension}.
-
-\subsection{Extension}
-\label{extension}
-The operations of extension enable one to consider families over
-families and higher level families alike. We will need both a context extension
-operation and a family extension operation.
-
-The inference rules introducing context and family extension are as follows:
-\begin{align}
-& \inference
-  { \jfam{\Gamma}{A}
-    }
-  { \jctx{\ctxext{\Gamma}{A}}
-    }
-& & \inference
-    { \jctxeq{\Gamma}{\Delta}
-      \jfameq{\Gamma}{A}{B}
-      }
-    { \jctxeq{\ctxext{\Gamma}{A}}{\ctxext{\Delta}{B}}
-      }
-    \\
-& \inference
-  { \jfam{{\Gamma}{A}}{P}
-    }
-  { \jfam{\Gamma}{\ctxext{A}{P}}
-    }
-& & \inference
-    { \jfameq{\Gamma}{A}{B} 
-      \jfameq{{\Gamma}{A}}{P}{Q}
-      }
-    { \jfameq{\Gamma}{\ctxext{A}{P}}{\ctxext{B}{Q}}
-      }
-\end{align}
-The extension operation may also be defined to act on families over families
-and terms thereof. For instance, When $\jfam{{\Gamma}{A}}{P}$ is a family,
-we would get a family $\jfam{{\Gamma}{A}}{{\Gamma}{P}}$ and when
-$\jterm{{\Gamma}{A}}{P}{f}$ is a term, we would get a term
-$\jterm{{\Gamma}{A}}{{\Gamma}{P}}{\ctxext{\Gamma}{f}}$. Then there would be
-axioms stating the judgmental equalities $\ctxext{\Gamma}{P}\jdeq P$ and
-$\ctxext{\Gamma}{f}\jdeq f$. Thus, it is of no actual use to introduce the
-action of context extension on families and terms.
-
-In \autoref{extension-on-terms} we will also define an extension operation
-on terms, which defines a term $\jterm{\Gamma}{{A}{P}}{\tmext{x}{u}}$ for
-every $\jterm{\Gamma}{A}{x}$ and $\jterm{\Gamma}{\subst{x}{P}}{u}$. This
-operation is defined by using the substitution operation and identity terms.
-
-\subsubsection{Associativity of extension}
-\label{comp-ee}
-The inference rules asserting that extension is compatible with itself assert
-that contexts are unstructured lists of type declarations. This rule is
-unavoidable if we want that for a family $A$ in context $\Gamma$, a family over
-$A$ is the same thing as a family over $\ctxext{\Gamma}{A}$. 
-\begin{align}
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{{\Gamma}{A}}{P}
-    }
-  { \jctxeq{\ctxext{{\Gamma}{A}}{P}}{\ctxext{\Gamma}{{A}{P}}}
-    }
-  \\
-& \inference
-  { \jfam{{\Gamma}{A}}{P}
-    \jfam{{{\Gamma}{A}}{P}}{Q}
-    }
-  { \jfameq{\Gamma}{\ctxext{A}{{P}{Q}}}{\ctxext{{A}{P}}{Q}}
-    }
-  \label{comp-ee-c}
-\end{align}
-Note that we may take the point of view that extension distributes over
-extensions. If we want to extend an extended family $\ctxext{P}{Q}$ over
-$\ctxext{\Gamma}{A}$ by the family $A$, the result is the extension of (1)
-the extension of $P$ by $A$ and (2) the extension of $Q$ by $A$. The latter is
-simply $Q$ since extension acts trivially at the level of families (and terms).
-
-\subsection{The empty context and the empty families}
-\label{empty}
-We introduce an empty context and an family over $\Gamma$ for every context $\Gamma$. 
-It has been suggested by some to only include empty families and not an
-empty context because an empty context is not necessary, but we do have several
-reasons to include them. Having an empty context requires also rules asserting 
-that a context is the same thing as a family over the empty context and this
-gives the categorical structure on contexts for free once one has it for 
-families. The main ingredients that will be missing from the theory once an
-empty context is avoided are weakening by a context and identity morphisms from
-a context to itself. Including these by hand also requires to formulate all the
-compatibility rules involving weakening once more for the cases of weakening by
-a context and identity terms at contexts. We prefer to state these rules once
-and only once and including an empty context helps in this respect.
-
-We also prefer to have our scheme of compatibility rules as symmetrical as
-possible. The structure of type dependency should look exactly the same in the
-default case as in any context. In that respect, having an empty family but not
-an empty context seems a bit odd. Also, some sets of compatibility rules 
-(like the rules stating that extension is compatible with the empty families)
-will become assymetrical as a result of not including an empty context. Moreover,
-we would eventually like to include a stratification of the theory by means of
-a type judgment (asserting that a family in a context $\Gamma$ is a type) and
-study closed types (i.e.~types in the empty context). It would be possible to
-provide a notion of closed types without having an empty context, but this would
-have to be formulated separately and we would have to restate all the rules for
-types (if any) for closed types all over again.
-
-One of the main uses of the empty context and the empty families is that we
-get the property that the `action on contexts' of an operation is compatible
-with its `action on families'.
-
-The empty family over a context $\Gamma$ is introduced by the following rule
-inference rule:
-\begin{align}
-& \inference
-  { }
-  { \jctx{\emptyc}
-    }
-  \\
-& \inference
-  { \jctx{\Gamma}
-    }
-  { \jfam{\Gamma}{\emptyf[\Gamma]}
-    }
-  \\
-& \inference
-  { \jctxeq{\Gamma}{\Gamma'}
-    }
-  { \jfameq{\Gamma}{\emptyf[\Gamma]}{\emptyf[\Gamma']}
-    }
-\end{align}
-
-By regarding contexts as families of contexts over the empty context, we
-enable ourselves also to speak of terms of contexts. A term of a context
-$\Gamma$ is a term of the family $\Gamma$ over the empty context. These ideas
-are captured in the following convertibility rules:
-\begin{align}
-& \inference
-  { \jctx{\Gamma}
-    }
-  { \jfam{\emptyc}{\Gamma}
-    } 
-& & \inference
-    { \jfam{\emptyc}{\Gamma}
-      }
-    { \jctx{\Gamma}
-      }
-    \\
-& \inference
-  { \jctxeq{\Gamma}{\Delta}
-    }
-  { \jfameq{\emptyc}{\Gamma}{\Delta}
-    }
-& & \inference
-    { \jfameq{\emptyc}{\Gamma}{\Delta}
-      }
-    { \jctxeq{\Gamma}{\Delta}
-      }
-\end{align}
-
-The reader may wonder whether the empty family $\emptyf$ always has a
-term. This shall follow from the rules stating the compatibility of extension
-with the empty families in \autoref{comp-0e} below and from
-identity terms (\autoref{identityterms}).
-\begin{comment}
-When contexts are viewed as statements, the empty context is the statement that
-asserts nothing. Likewise, in any context one can assert nothing and this is
-achieved by the empty family. 
-Since somebody who states nothing asserts it all right, there is a term of the empty
-family over any context. Moreover, any two terms of the empty family are always
-judgmentally equal.
-\begin{align}
-& \inference
-  { }
-  { \jctx{\emptyc}
-    }
-  \\
-& \inference
-  { \jctx{\Gamma}
-    }
-  { \jfam{\Gamma}{\emptyf[\Gamma]}
-    }
-  \\
-& \inference
-  { \jctx{\Gamma}
-    }
-  { \jterm{\Gamma}{\emptyf[\Gamma]}{\emptytm[\Gamma]}
-    }
-  \\
-& \inference
-  { \jterm{\Gamma}{\emptyf[\Gamma]}{x}
-    }
-  { \jtermeq
-      {\Gamma}
-      {\emptyf[\Gamma]}
-      {x}
-      {\emptytm[\Gamma]}
-    }
-  \label{comp-00-t}
-\end{align}
-
-The empty context and the empty families together with the operation of
-extension will provide several shortcuts for the rest
-of the theory of families and terms. This is (in part) obtained by requiring
-that a context is exactly the same thing as a family of contexts over the
-empty context and that judgmental equality of contexts is exactly the same
-as judgmental equality of families of contexts over the empty context.
-
-\label{comp-00}
-Because every context is a family of contexts over the empty context, 
-the empty context $\emptyc$ may be considered as a family of contexts over the empty
-context. When we do this, we get $\emptyf[\emptyc]$.
-\begin{equation}\label{eq:comp-00}
-\inference
-  { }
-  { \jfameq
-      {\emptyc}
-      {\emptyc}
-      {\emptyf[\emptyc]}
-    }
-\end{equation}
-In the future, we shall denote $\emptyf[\Gamma]$ by $\emptyf$. The above rule
-guarantees that this will not cause confusion. Likewise, we shall denote
-$\emptytm[\Gamma]$ by $\emptytm$.
-\end{comment}
-
-\subsubsection{Compatibility of extension with the empty context and families}
-In the following set of inference rules we state that the empty context and
-the empty family are neutral objects for both context extension (the first two
-rules) and family extension (the last two rules).
-\label{comp-e0}\label{comp-0e}
-\begin{align}
-& \inference
-  { \jctx{\Gamma}
-    }
-  { \jctxeq{\ctxext{\emptyc}{\Gamma}}{\Gamma}
-    }
-  \label{comp-0e-c}
-  \\
-& \inference
-  { \jctx{\Gamma}
-    }
-  { \jctxeq{\ctxext{\Gamma}{\emptyf}}{\Gamma}
-    }
-  \label{comp-e0-c}\\
-& \inference
-  { \jfam{\Gamma}{A}
-    }
-  { \jfameq{\Gamma}{\ctxext{\emptyf}{A}}{A}
-    }
-  \label{comp-0e-f}
-  \\
-& \inference
-  { \jfam{\Gamma}{A}
-    }
-  { \jfameq{\Gamma}{{A}{\emptyf}}{A}
-    }
-  \label{comp-e0-f}
-\end{align}
-
-\subsubsection{Family extension restricted to the empty context}
-Strictly speaking we should have used a different notation for context extension
-as for family extension, because the following rule asserting that family extension
-in the empty context is the same thing as context extension would look tautological
-without a difference. So let us denote, only for the moment, context extension
-of $\Gamma$ by $A$ by $(\ctxext{\Gamma}{A})^c$ and family extension of $A$ by
-$P$ in context $\Gamma$ by $(\ctxext{A}{P})^\famsym$.
-
-Note that we may consider a context $\Gamma$ as a family over $\emptyc$ and
-a family $\jfam{\Gamma}{A}$ as a family $\jfam{{\emptyc}{\Gamma}}{A}$. 
-Therefore we will require the following rule:
-\begin{align}
-& \inference
-  { \jfam{\Gamma}{A}
-    }
-  { \jctxeq{(\ctxext{\Gamma}{A})^c}{(\ctxext{\Gamma}{A})^\famsym}
-    }
-\end{align}
-Note that this rule actually justifies that we have not utilized two different
-notations for context extension and family extension.
-
 \subsection{The theory of weakening}
 \label{weakening}
+
 When $A$ is a family in context $\Gamma$, the operation of weakening by $A$
@@
 \end{align}
-
-\subsubsection{Weakening of an empty family}
-The following inference rules express that when the empty family is
-weakened, the result is the empty family.
-\label{comp-w0}
-\begin{align}
-%& \inference
-%  { \jctx{\Gamma}
-%    }
-%  { \jfameq{\Gamma}{\ctxwk{\Gamma}{\emptyc}}{\emptyf}
-%    }
-%  \\
-& \inference
-  { \jfam{\Gamma}{A}
-    }
-  { \jfameq{{\Gamma}{A}}{\ctxwk{A}{\emptyf}}{\emptyf}
-    }
-  \label{comp-w0-c}\\
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{\Gamma}{B}
-    }
-  { \jfameq
-    {{{\Gamma}{A}}{\ctxwk{A}{B}}}
-    {\ctxwk[\famsym]{A}{\emptyf}}
-    {\emptyf}
-    }
-  \label{comp-w0-f}
-\end{align}
-\begin{comment}
-Note that we do not have to axiomatize the action of weakening
-by $A$ on the term of the empty family, because the result will be in the empty 
-family and every term in the empty family is judgmentally equal to the term 
-$\emptytm$.
-\end{comment}
-
-Because a family over $\Gamma$ is the same as a family over 
-$\ctxext{\Gamma}{\emptyf}$ we can apply both the action on contexts and the
-action on families of weakening to a family $B$ over $\Gamma$. When we apply
-the action on families, we obtain a family $\ctxwk[\famsym]{A}{B}$ over the
-context $\ctxext{{\Gamma}{A}}{\ctxwk{A}{\emptyf}}$. However, since we have
-postulated the judgmental equalities $\ctxwk{A}{\emptyf}\jdeq\emptyf$ and
-$\ctxext{{\Gamma}{A}}{\emptyf}\jdeq\ctxext{\Gamma}{A}$, we see that we can
-compare $\ctxwk[\famsym]{A}{B}$ with $\ctxwk{A}{B}$. The following inference
-rule postulates that these two are judgmentally equal:
-\begin{equation}
-\inference
-{ \jfam{\Gamma}{A}
-  \jfam{\Gamma}{B}
-  }
-{ \jfameq{{\Gamma}{A}}{\ctxwk[\famsym]{A}{B}}{\ctxwk{A}{B}}
-  }
-\end{equation}
-Due to this rule, the action on contexts and the action on families of weakening
-are compatible with each other and consequently there can be no possible
-confusion when we omit the annotations $\famsym$ and $\tmsym$ alltogether. In
-the future, the weakening of a family $Q$ over $\ctxext{\Gamma}{B}$ shall
-be denoted just by $\ctxwk{A}{Q}$ and likewise the weakening of a term $g$ of
-$Q$ shall be denoted by $\ctxwk{A}{g}$.
-
-Because a family $B$ over $\Gamma$ can be treated as a family by the operation
-of weakening, weakening also acts on the terms of $B$. The weakening of a term
-$y$ of $B$ by $A$ can be seen as the constant term (or function) of the
-family $\ctxwk{A}{B}$.
 
@@
 
-\subsubsection{Weakening by the empty family}
-Note that we can also weaken by the empty family over $\Gamma$.
-Weakening by the empty family $\emptyf$ over a context $\Gamma$ leaves families, 
-their terms, families over those families and terms of those unchanged:
-\label{comp-0w}
-\begin{align}
-& \inference
-  { \jfam{\Gamma}{B}
-    }
-  { \jfameq{\Gamma}{\ctxwk{\emptyf}{B}}{B}
-    }
-  \label{comp-0w-c}
-  \\
-& \inference
-  { \jfam{{\Gamma}{B}}{Q}
-    }
-  { \jfameq{{\Gamma}{B}}{\ctxwk{\emptyf}{Q}}{Q}
-    }
-  \label{comp-0w-f}
-  \\
-& \inference
-  { \jterm{{\Gamma}{B}}{Q}{g}
-    }
-  { \jtermeq{{\Gamma}{B}}{Q}{\ctxwk{\emptyf}{g}}{g}
-    }
-  \label{comp-0w-t}
-\end{align}
-
 \subsubsection{Currying for weakening}
@@
 \end{align}
-
-\begin{comment}
-\subsubsection{Weakening by a context}
-Since we have not required that there is an empty context with the property that
-a context is the same thing as a family in the empty context, we will have to
-explicitly introduce weakening by contexts. Since it will allways be clear
-whether something is weakened by a context or by a family, we shall use the
-same notation for weakening by contexts as for weakening by families. We shall
-have to repeat all the compatibility rules for weakening by families for
-weakening by contexts; there is no way around it. It is
-actually this sort of inconvenience that has originally led us to include an
-empty context and be done with it.
-
-In the following inference rules we assume that we have a context $\Gamma$ and
-that $\jctxeq{\Gamma}{\Gamma'}$ in the rules where $\Gamma'$ makes an appearence.
-\begin{align}
-& \inference
-  { \jctx{\Delta}
-    }
-  { \jfam{\Gamma}{\ctxwk{\Gamma}{\Delta}}
-    }
-& & \inference
-  { \jctxeq{\Delta}{\Delta'}
-    }
-  { \jfameq{\Gamma}{\ctxwk{\Gamma}{\Delta}}{\ctxwk{\Gamma'}{\Delta'}}
-    }
-  \\
-& \inference
-  { \jfam{\Delta}{B}
-    }
-  { \jfam{\ctxext{\Gamma}{\ctxwk{\Gamma}{\Delta}}}{\ctxwk{\Gamma}{B}}
-    }
-& & \inference
-  { \jfameq{\Delta}{B}{B'}
-    }
-  { \jfameq
-      {\ctxext{\Gamma}{\ctxwk{\Gamma}{\Delta}}}
-      {\ctxwk{\Gamma}{B}}
-      {\ctxwk{\Gamma'}{B'}}
-    }
-  \\
-& \inference
-  { \jterm{\Delta}{B}{y}
-    }
-  { \jterm
-      {\ctxext{\Gamma}{\ctxwk{\Gamma}{\Delta}}}
-      {\ctxwk{\Gamma}{B}}
-      {\ctxwk{\Gamma}{y}}
-    }
-& & \inference
-  { \jtermeq{\Delta}{B}{y}{y'}
-    }
-  { \jtermeq
-      {\ctxext{\Gamma}{\ctxwk{\Gamma}{\Delta}}}
-      {\ctxwk{\Gamma}{B}}
-      {\ctxwk{\Gamma}{y}}
-      {\ctxwk{\Gamma}{y'}}
-    }
-\end{align}
-Note that we do not employ three different notations for the three different
-actions of weakening by a context because there can be no confusion about which
-of the three actions is meant. If there were an empty context we should have
-distinguished the three as we did for weakening by a family.
-
-We shall have similar compatibility rules for weakening by a context as we did
-for weakening by a family.
-
-{\color{red}To do. It will look exactly the same as for weakening by families}
-\end{comment}
-
-\subsection{The theory of projections}
-\label{identityterms}
-We formulate the theory of projections on top of the theory of weakening
-described in \autoref{weakening}. In the theory of projections we introduce
-the identity terms. The identity term at a family $A$ in context $\Gamma$ is
-a term $\jterm{{\Gamma}{A}}{\ctxwk{A}{A}}{\idtm{A}}$, so weakening is necessary.
-Together with weakening, the identity terms will provide for the projections.
-The projection from $\cftext{A}{P}$ to $A$ will be the term
-$\jterm{{\Gamma}{{A}{P}}}{\ctxwk{\ctxext{A}{P}}{A}}{\ctxwk{P}{\idtm{A}}}$ and
-will be studied extensively in \autoref{extension-on-terms}.
-
-The following inference rules are required to be valid:
-\begin{align}
-& \inference
-  { \jfam{\Gamma}{A}
-    }
-  { \unfoldall{\jhom{\Gamma}{A}{A}{\idtm{A}}}
-    }
-& & \inference
-    { \jfameq{\Gamma}{A}{A'}
-      }
-    { \unfoldall{\jhomeq{\Gamma}{A}{A}{\idtm{A}}{\idtm{A'}}}
-      }
-\end{align}
-
-\begin{comment}
-We stress again
-that because we have not required there to be an empty context, we need to
-explicitly call for identity terms $\idtm{\Gamma}$ of $\ctxwk{\Gamma}{\Gamma}$
-in context $\Gamma$ for each context $\Gamma$. Thus, we will also require
-the inference rule
-\begin{align}
-& \inference
-  { \jctx{\Gamma}
-    }
-  { \jterm{\Gamma}{\ctxwk{\Gamma}{\Gamma}}{\idtm{\Gamma}}
-    }
-& &
-  \inference
-  { \jctxeq{\Gamma}{\Gamma'}
-    }
-  { \jtermeq{\Gamma}{\ctxwk{\Gamma}{\Gamma}}{\idtm{\Gamma}}{\idtm{\Gamma'}}
-    }
-\end{align}
-For the following inference rules, there should also be versions of identity
-terms at contexts.
-\end{comment}
-
-\subsubsection{The identity term of an extended family}
-The fact that identity terms are compatible with extensions is derived in
-\autoref{comp-ie}.
-
-\subsubsection{The identity term of a weakened family}
-\label{comp-wi}
-The identity term of a weakened family is the weakened identity term:
-\begin{equation}
-\inference
-  { \jfam{\Gamma}{A}
-    \jfam{{\Gamma}{B}}{Q}
-    }
-  { \unfoldall{\jhomeq
-      {{{\Gamma}{A}}{\ctxwk{A}{B}}}
-      {\ctxwk{A}{Q}}
-      {\ctxwk{A}{Q}}
-      {\ctxwk{A}{\idtm{Q}}}
-      {\idtm{\ctxwk{A}{Q}}}
-      }
-    }
-  \label{comp-wi-t}
-\end{equation}
-
-\begin{rmk}
-An important special case is the judgmental equality
-\begin{equation*}
-\unfoldall{\jhomeq
-      {{\Gamma}{A}}
-      {\ctxwk{A}{B}}
-      {\ctxwk{A}{B}}
-      {\ctxwk{A}{\idtm{B}}}
-      {\idtm{\ctxwk{A}{B}}}
-      }
-\end{equation*}
-for a family $\jfam{\Gamma}{B}$.
-\end{rmk}
-
-\subsection{The theory of substitution}
-\label{substitution}
-The theory of substitution can be formulated separately from the theory of
-weakening. Therefore, we will only require the theory of extension and the
-empty context and families. In \autoref{sec:esystem-equalities} we will
-provide the additional judgmental equality rules that will be required when
-combining the theories of substitution and of projections.
-
-Given a family $P$ over $A$ and a term $x$ of $A$, substitution gives a way to
-consider the \emph{fiber $\subst{x}{P}$ of $P$ at $x$}. As was the case with
-weakening, the substitution operation comes in three parts: the `action on
-contexts' of substitution is the operation just described; the `action on
-families' of substitution takes a family $Q$ over $\ctxext{{\Gamma}{A}}{P}$
-to a family $\subst[\famsym]{x}{Q}$ over $\ctxext{\Gamma}{\subst{x}{P}}$; the
-`action on terms' of substitution takes a term $g$ of $Q$ over
-$\ctxext{{\Gamma}{A}}{P}$ to a term $\subst[\tmsym]{x}{g}$ of 
-$\subst[\famsym]{x}{Q}$.
-\begin{align}
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jfam{{\Gamma}{A}}{P}
-    }
-  { \jfam{\Gamma}{\subst{x}{P}}
-    }
-& & \inference
-    { \jtermeq{\Gamma}{A}{x}{x'}
-      \jfameq{{\Gamma}{A}}{P}{P'}
-      }
-    { \jfameq{\Gamma}{\subst{x}{P}}{\subst{x'}{P'}}
-      }
-    \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jfam{{{\Gamma}{A}}{P}}{Q}
-    }
-  { \jfam{{\Gamma}{\subst{x}{P}}}{\subst[\famsym]{x}{Q}}
-    }
-& & \inference
-    { \jtermeq{\Gamma}{A}{x}{x'}
-      \jfameq{{{\Gamma}{A}}{P}}{Q}{Q'}
-      }
-    { \jfameq{{\Gamma}{\subst{x}{P}}}{\subst[\famsym]{x}{Q}}{\subst[\famsym]{x'}{Q'}}
-      }
-    \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jterm{{{\Gamma}{A}}{P}}{Q}{g}
-    }
-  { \jterm{{\Gamma}{\subst{x}{P}}}{\subst[\famsym]{x}{Q}}{\subst[\tmsym]{x}{g}}
-    }
-& & \inference
-    { \jtermeq{\Gamma}{A}{x}{x'}
-      \jtermeq{{{\Gamma}{A}}{P}}{Q}{g}{g'}
-      }
-    { \jtermeq
-        {{\Gamma}{\subst{x}{P}}}
-        {\subst[\famsym]{x}{Q}}
-        {\subst[\tmsym]{x}{g}}
-        {\subst[\tmsym]{x'}{g'}}
-      }
-\end{align}
-
-\subsubsection{Fibers of an empty family}
-The following inference rules establish that the fibers of the empty family are 
-the empty families:
-\label{comp-s0}
-\begin{align}
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    }
-  { \jfameq{\Gamma}{\subst{x}{\emptyf}}{\emptyf}
-    }
-  \label{comp-s0-c}
-  \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jfam{{\Gamma}{A}}{P}
-    }
-  { \jfameq
-      {{\Gamma}{\subst{x}{P}}}
-      {\subst{x}{\emptyf}}
-      {\emptyf}
-    }
-  \label{comp-s0-f}
-\end{align}
-\begin{comment}
-Note again that we do not have to axiomatize the action of substitution by $x$ on the term
-of the empty family, because the result will be in the empty family and every
-term in the empty family is judgmentally equal to the term $\emptytm$.
-\end{comment}
-
-We use the above rule to state the compatibility of the action on families of
-substitution with the action on contexts of substitution. Note that a family
-$P$ over $\ctxext{\Gamma}{A}$ may be regarded as a family over
-$\ctxext{{\Gamma}{A}}{\emptyf}$. Thus, we may consider the family
-$\subst[\famsym]{x}{P}$ over $\ctxext{\Gamma}{\subst{x}{\emptyf}}$. Since
-$\subst{x}{\emptyf}$ is judgmentally equal to the empty family, we may compare
-$\subst[\famsym]{x}{P}$ with $\subst{x}{P}$:
-\begin{equation}
-\inference
-{ \jfam{{\Gamma}{A}}{P}
-  }
-{ \jfameq{\Gamma}{\subst[\famsym]{x}{P}}{\subst{x}{P}}
-  }
-\end{equation}
-Due to this rule we need not make the annotations $\famsym$ and $\tmsym$ in
-the notation for substitution anymore and thus we shall omit them from now on.
-Note that because a family $P$ over $\ctxext{\Gamma}{A}$ is eligible for
-application of the action on families of substitution, we may also substitute
-terms of $P$. Thus, given terms $\jterm{\Gamma}{A}{x}$ and $\jterm{{\Gamma}{A}}{P}{f}$,
-we get a term $\jterm{\Gamma}{\subst{x}{P}}{\subst{x}{f}}$, the \emph{value of
-$f$ at $x$}.
-
-In \autoref{morphisms} we will use a combination of weakening and substitution
-to define composition of morphisms of families. However, we have to rely
-on the cancellation rule stated in \autoref{cancellation-ws} before we can
-meaningfully state the definition of composition.
-
-\subsubsection{Fibers of extensions}
-\label{comp-se}
-The following inference rules assert that if we take the fiber of an extended
-family at a term $x$ of $A$ in context $\Gamma$, the substitution by $x$
-distributes over the factors of the extension.
-\begin{align}
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jfam{{{\Gamma}{A}}{P}}{Q}
-    }
-  { \jfameq
-      {\Gamma}
-      {\subst{x}{\ctxext{P}{Q}}}
-      {\ctxext{\subst{x}{P}}{\subst{x}{Q}}}
-    }
-  \label{comp-se-c}
-  \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jfam{{{{\Gamma}{A}}{P}}{Q}}{R}
-    }
-  { \jfameq
-      {{\Gamma}{\subst{x}{P}}}
-      {\subst{x}{\ctxext{Q}{R}}}
-      {\ctxext{\subst{x}{Q}}{\subst{x}{R}}}
-    }
-  \label{comp-se-f}
-\end{align}
-There is also a version of this statement in which extended terms are considered.
-This variant is stated and proved in \autoref{comp-se-t}.
-
-\subsubsection{Fibers of fibers}
-\label{comp-ss}
-
-The following rules explain what happens when we first substitute by a term
-$g$ of $Q$ over $P$ in context $\ctxext{\Gamma}{A}$ and then by a term $x$ of
-$A$:
-\begin{align}
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jterm{{{\Gamma}{A}}{P}}{Q}{g}
-    \jfam{{{{\Gamma}{A}}{P}}{Q}}{R}
-    }
-  { \jfameq
-      {{{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}}
-      {\subst{x}{{g}{R}}}
-      {\subst{{x}{g}}{{x}{R}}}
-    }
-  \label{comp-ss-c}
-  \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jterm{{{\Gamma}{A}}{P}}{Q}{g}
-    \jfam{{{{{\Gamma}{A}}{P}}{Q}}{R}}{S}
-    }
-  { \jfameq
-      {{{{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}}{\subst{x}{{g}{R}}}}
-      {\subst{x}{{g}{S}}}
-      {\subst{{x}{g}}{{x}{S}}}
-    }
-  \label{comp-ss-f}
-  \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jterm{{{\Gamma}{A}}{P}}{Q}{g}
-    \jterm{{{{{\Gamma}{A}}{P}}{Q}}{R}}{S}{k}
-    }
-  { \jtermeq
-      {{{{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}}{\subst{x}{{g}{R}}}}
-      {\subst{x}{{g}{S}}}
-      {\subst{x}{{g}{k}}}
-      {\subst{{x}{g}}{{x}{k}}}
-    }
-  \label{comp-ss-t}
-\end{align}
-
-\begin{rmk}
-As an important special case of these inference rules we have the following
-valid inference rules:
-\begin{align*}
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jterm{{\Gamma}{A}}{P}{f}
-    \jfam{{{\Gamma}{A}}{P}}{Q}
-    }
-  { \jfameq
-      {\Gamma}
-      {\subst{x}{{f}{Q}}}
-      {\subst{{x}{f}}{{x}{Q}}}
-    }
-  \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jterm{{\Gamma}{A}}{P}{f}
-    \jfam{{{{\Gamma}{A}}{P}}{Q}}{R}
-    }
-  { \jfameq
-      {{\Gamma}{\subst{x}{{f}{Q}}}}
-      {\subst{x}{{f}{R}}}
-      {\subst{{x}{f}}{{x}{R}}}
-    }
-  \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jterm{{\Gamma}{A}}{P}{f}
-    \jterm{{{{\Gamma}{A}}{P}}{Q}}{R}{h}
-    }
-  { \jtermeq
-      {{\Gamma}{\subst{x}{{f}{Q}}}}
-      {\subst{x}{{f}{R}}}
-      {\subst{x}{{f}{h}}}
-      {\subst{{x}{f}}{{x}{h}}}
-    }
-\end{align*}
-Moreover, we get
-\begin{equation*}
-\inference
-  { \jterm{\Gamma}{A}{x}
-    \jterm{{\Gamma}{A}}{P}{f}
-    \jterm{{{\Gamma}{A}}{P}}{Q}{g}
-    }
-  { \jtermeq
-      {\Gamma}
-      {\subst{x}{{f}{Q}}}
-      {\subst{x}{{f}{g}}}
-      {\subst{{x}{f}}{{x}{g}}}
-    }
-\end{equation*}
-\end{rmk}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Joining the theory of projections with the theory of substitution}
-\label{sec:esystem-equalities}
-
-\begin{comment}
-We have already seen that empty families, extension, weakening, substitution
-and the identity terms are compatible with each other. There are, however, a
-few more judgmental equalities explaining what happens when
-\begin{enumerate}
-\item something is weakend by the empty family. This is used to consider a term
-      $x$ of $A$ in context $\Gamma$ as a morphism from $\emptyf$ to $A$ in
-      context $\Gamma$;
-\item something is weakened by an extended family. This is Currying for weakening;
-\item something is first weakened by a family $A$ in context $\Gamma$ and then
-      substituted by a term of $A$. This is used to show that composition is
-      compatible with itself (i.e. that composition is associative);
-\item the identity terms are joining the party. Identity terms are empty contexts
-      in certain models, so they should behave accordingly.
-\end{enumerate}
-In other words, the rules seem to not fit in the scheme that all operations are
-compatible with each other, but in fact each of these rules plays a role in
-completing the scheme (work out further).
-\end{comment}
-
-\begin{comment}
-In this section we extend the judgmental equality relation in such a way that
-the empty context, extension, weakening, substitution and the identity terms
-all become compatible with each other. Since we have five operations, there
-are going to be $25$ finite sets of rules explaining how each operation is
-compatible with each of the operations.
-
-In addition to these $25$ finite sets
-of compatibility rules, there are also two sets of rules explaining the
-cancellation properties of weakening and substitution and of the identity terms.
-
-A typical compatibility axiom captures the idea that a diagram like
-\begin{equation*}
-\begin{tikzcd}[column sep=huge]
-\jfam{\Gamma}{\blank} 
-  \ar{r}{C\mapsto\ctxwk{B}{C}} 
-  \ar{d}[swap]{C\mapsto\ctxwk{A}{C}} 
-& \jfam{{\Gamma}{B}}{\blank} 
-  \ar{d}{Q\mapsto\ctxwk{A}{Q}}
-  \\
-\jfam{{\Gamma}{A}}{\blank} 
-  \ar{r}[swap]{P\mapsto\ctxwk{{A}{B}}{P}} 
-& \jfam{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\blank}
-\end{tikzcd}
-\end{equation*}
-in the case of two consecutive weakenings, or
-\begin{equation*}
-\begin{tikzcd}[column sep=huge]
-\jfam{{\Gamma}{A}}{\blank} 
-  \ar{d}[swap]{Q\mapsto\subst{x}{Q}} 
-  \ar{r}{Q\mapsto\ctxwk{P}{Q}} 
-& \jfam{{{\Gamma}{A}}{P}}{\blank} 
-    \ar{d}{R\mapsto\subst{x}{R}}
-  \\ 
-\jfam{\Gamma}{\blank} 
-  \ar{r}[swap]{B\mapsto\ctxwk{\subst{x}{P}}{B}} 
-& \jfam{{\Gamma}{\subst{x}{P}}}{\blank}
-\end{tikzcd}
-\end{equation*}
-in the case of a combination of weakening and substitution, commutes. We won't
-actually draw any of these diagrams in this section, but readers are encouraged
-to do so if they think it can help them understanding the axioms. Another thing
-one could keep in mind with the axioms that involve weakening is that all of
-them become completely obvious when weakening is made notationally invisible.
-Of course, we have an explicit notation for weakening on purpuse, but it is the
-usual practice of type theorists to make weakening invisible.
-
-The table below indicates where
-we can find the axioms of each of the combinations of operators. The reference
-at the row of operation $X$ and the column of operation $Y$ refers to the
-axiom which explains $Y$ distributes over the arguments of $X$ when $Y$ is
-applied after $X$. The empty context has no arguments, whereas extension,
-and weakening have two arguments which are both of the family-kind, substitution
-has two arguments of which one is a term and the other is a family and
-idenitity terms have one argument, a family.
-\begin{center}
-\begin{tabular}{r|rrrr}
-& ext & wk & subst & id\\
-\hline
-extension & \autoref{comp-ee}* &  & \cmark & \cmark \\
-weakening & \autoref{comp-we}* & \autoref{comp-ww} & \autoref{comp-sw} & \autoref{comp-wi} \\
-substitution & \autoref{comp-se}* & \autoref{comp-ws} & \autoref{comp-ss} & \autoref{comp-si} \\
-identity terms & \cmark & \autoref{comp-wi} & \autoref{comp-si} & \xmark
-\end{tabular}
-\end{center}
-Apart from references to sections where axioms are given, we also see checkmarks,
-crosses, marked references and duplicated references in this table. 
-The checkmarks indicate that the desired property
-can be proved and need not be axiomatized. To clarify those:
-\begin{itemize}
-\item The identity term of the empty context is the unique term of the empty
-family. Recall that we already have axiomatized in \autoref{empty} that every
-term of the empty family is judgmentally equal to $\emptytm$.
-\item The identity term of an extension can be explained by means of the notion
-of \emph{extension on terms}. This notion is defined in \autoref{extension-on-terms}
-using only identity terms and substitution. With extension on terms we can
-combine a term $\jterm{\Gamma}{A}{x}$ and a term $\jterm{\Gamma}{\subst{x}{P}}{u}$
-into a term $\jterm{\Gamma}{{A}{P}}{\tmext{x}{u}}$, which may be seen as the
-pair consisting of $x$ and $u$. Since it is a defined notion, we will
-be able to derive lots of properties for extension of terms. One of them explains
-how we can see the identity term of an extension as the extension of two certain
-terms, see \autoref{comp-ie}.
-\item Using extension on terms we can also make sense of `fibers at an extension'
-because that would exactly be the fiber at an extended term. We will see in
-\autoref{comp-es} that, for instance, $\subst{\tmext{x}{u}}{Q}\jdeq\subst{u}{{x}{Q}}$
-for any family $\jfam{{{\Gamma}{A}}{P}}{Q}$ and any two terms $\jterm{\Gamma}{A}{x}$
-and $\jterm{\Gamma}{\subst{x}{P}}{u}$. That's a form of currying.
-\end{itemize}
-The spots with a \xmark simply make no sense 
-{\color{red}(or do they, when the term model
-is better understood (by me)?)}. The marked references point to
-axioms which \emph{do} have a variant for terms, but those require the notion
-of extension on terms, can be proved and therefore need not be included in our
-list of compatibility axioms. In each of these sections, we give a forward
-reference to the lemma in which this variant is stated and proved.
-\end{comment}
-
-\subsubsection{Fibers of weakenings}\label{comp-sw}
-The following rules assert what happens when we first weaken by a family
-$Q$ over $P$ in context $\ctxext{\Gamma}{A}$ and then substitute by a term
-$x$ of $A$:
-\begin{align}
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jfam{{{\Gamma}{A}}{P}}{Q}
-    \jfam{{{\Gamma}{A}}{P}}{R}
-    }
-  { \jfameq
-      {{{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}}
-      {\subst{x}{\ctxwk{Q}{R}}}
-      {\ctxwk{\subst{x}{Q}}{\subst{x}{R}}}
-    }
-  \label{comp-sw-c}
-  \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jfam{{{\Gamma}{A}}{P}}{Q}
-    \jfam{{{{\Gamma}{A}}{P}}{R}}{S}
-    }
-  { \jfameq
-      {{{{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}}{\subst{x}{\ctxwk{Q}{R}}}}
-      {\subst{x}{\ctxwk{Q}{S}}}
-      {\ctxwk{\subst{x}{Q}}{\subst{x}{S}}}
-    }
-  \label{comp-sw-f}
-  \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jfam{{{\Gamma}{A}}{P}}{Q}
-    \jterm{{{{\Gamma}{A}}{P}}{R}}{S}{k}
-    }
-  { \jtermeq
-      {{{{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}}{\subst{x}{\ctxwk{Q}{R}}}}
-      {\subst{x}{\ctxwk{Q}{S}}}
-      {\subst{x}{\ctxwk{Q}{k}}}
-      {\ctxwk{\subst{x}{Q}}{\subst{x}{k}}}
-    }
-  \label{comp-sw-t}
-\end{align}
-
-\begin{rmk}
-As an important special case of these inference rules we have the following
-valid inference rules:
-\begin{align*}
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jfam{{\Gamma}{A}}{P}
-    \jfam{{\Gamma}{A}}{Q}
-    }
-  { \jfameq
-      {{\Gamma}{\subst{x}{P}}}
-      {\subst{x}{\ctxwk{P}{Q}}}
-      {\ctxwk{\subst{x}{P}}{\subst{x}{Q}}}
-    }
-  \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jfam{{\Gamma}{A}}{P}
-    \jfam{{{\Gamma}{A}}{Q}}{R}
-    }
-  { \jfameq
-      {{{\Gamma}{\subst{x}{P}}}{\subst{x}{\ctxwk{P}{Q}}}}
-      {\subst{x}{\ctxwk{P}{R}}}
-      {\ctxwk{\subst{x}{P}}{\subst{x}{R}}}
-    }
-  \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jfam{{\Gamma}{A}}{P}
-    \jterm{{{\Gamma}{A}}{Q}}{R}{h}
-    }
-  { \jtermeq
-      {{{\Gamma}{\subst{x}{P}}}{\subst{x}{\ctxwk{P}{Q}}}}
-      {\subst{x}{\ctxwk{P}{R}}}
-      {\subst{x}{\ctxwk{P}{h}}}
-      {\ctxwk{\subst{x}{P}}{\subst{x}{h}}}
-    }
-\end{align*}
-Moreover, we get
-\begin{equation*}
-\inference
-  { \jterm{\Gamma}{A}{x}
-    \jfam{{\Gamma}{A}}{P}
-    \jterm{{\Gamma}{A}}{Q}{g}
-    }
-  { \jtermeq
-      {{\Gamma}{\subst{x}{P}}}
-      {\subst{x}{\ctxwk{P}{Q}}}
-      {\subst{x}{\ctxwk{P}{g}}}
-      {\ctxwk{\subst{x}{P}}{\subst{x}{g}}}
-    }
-\end{equation*}
-\end{rmk}
-
-\subsubsection{Weakenings of fibers}
-\label{comp-ws}
-The following inference rules explain what happens when we first weaken by a
-term $g$ of $Q$ in context $\ctxext{\Gamma}{B}$ and then weaken by a family
-$A$ over $\Gamma$.
-\begin{align}
-& \inference
-  { \jfam{\Gamma}{A}
-    \jterm{{\Gamma}{B}}{Q}{g}
-    \jfam{{{\Gamma}{B}}{Q}}{R}
-    }
-  { \jfameq
-      {{{\Gamma}{A}}{\ctxwk{A}{B}}}
-      {\ctxwk{A}{\subst{g}{R}}}
-      {\subst{\ctxwk{A}{g}}{\ctxwk{A}{R}}}
-    }
-  \label{comp-ws-c}
-  \\
-& \inference
-  { \jfam{\Gamma}{A}
-    \jterm{{\Gamma}{B}}{Q}{g}
-    \jfam{{{{\Gamma}{B}}{Q}}{R}}{S}
-    }
-  { \jfameq
-      {{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{\subst{g}{R}}}}
-      {\ctxwk{A}{\subst{g}{S}}}
-      {\subst{\ctxwk{A}{g}}{\ctxwk{A}{S}}}
-    }
-  \label{comp-ws-f}
-  \\
-& \inference
-  { \jfam{\Gamma}{A}
-    \jterm{{\Gamma}{B}}{Q}{g}
-    \jterm{{{{\Gamma}{B}}{Q}}{R}}{S}{k}
-    }
-  { \jtermeq
-      {{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{\subst{g}{R}}}}
-      {\ctxwk{A}{\subst{g}{S}}}
-      {\ctxwk{A}{\subst{g}{k}}}
-      {\subst{\ctxwk{A}{g}}{\ctxwk{A}{k}}}
-    }
-  \label{comp-ws-t}
-\end{align}
-
-\begin{rmk}
-As an important special case of these inference rules we have the following 
-valid inference rules:
-\begin{align*}
-& \inference
-  { \jfam{\Gamma}{A}
-    \jterm{\Gamma}{B}{y}
-    \jfam{{\Gamma}{B}}{Q}
-    }
-  { \jfameq
-      {{\Gamma}{A}}
-      {\ctxwk{A}{\subst{y}{Q}}}
-      {\subst{\ctxwk{A}{y}}{\ctxwk{A}{Q}}}
-    }
-  \\
-& \inference
-  { \jfam{\Gamma}{A}
-    \jterm{\Gamma}{B}{y}
-    \jfam{{{\Gamma}{B}}{Q}}{R}
-    }
-  { \jfameq
-      {{{\Gamma}{A}}{\ctxwk{A}{\subst{y}{Q}}}}
-      {\ctxwk{A}{\subst{y}{R}}}
-      {\subst{\ctxwk{A}{y}}{\ctxwk{A}{R}}}
-    }
-  \\
-& \inference
-  { \jfam{\Gamma}{A}
-    \jterm{\Gamma}{B}{y}
-    \jterm{{{\Gamma}{B}}{Q}}{R}{h}
-    }
-  { \jtermeq
-      {{{\Gamma}{A}}{\ctxwk{A}{\subst{y}{Q}}}}
-      {\ctxwk{A}{\subst{y}{R}}}
-      {\ctxwk{A}{\subst{y}{h}}}
-      {\subst{\ctxwk{A}{y}}{\ctxwk{A}{h}}}
-    }
-\end{align*}
-Moreover, we get
-\begin{equation*}
-\inference
-  { \jfam{\Gamma}{A}
-    \jterm{\Gamma}{B}{y}
-    \jterm{{\Gamma}{B}}{Q}{g}
-    }
-  { \jtermeq
-      {{\Gamma}{A}}
-      {\ctxwk{A}{\subst{y}{Q}}}
-      {\ctxwk{A}{\subst{y}{g}}}
-      {\subst{\ctxwk{A}{y}}{\ctxwk{A}{g}}}
-    }
-\end{equation*}
-\end{rmk}
-
-\subsubsection{The cancellation property of weakening and substitution}
-\label{cancellation-ws}
-The judgmental equalities we're about to describe assert that substituting a term
-in the weakening a thing gives you the thing back. In the case of contexts we get that each fiber
-$\subst{x}{\ctxwk{A}{B}}$ is just $B$ and in the case of terms we get 
-that $\ctxwk{A}{y}$ is the constant function
-mapping everything to $y:B$. Thus, these rules actually establish the weakening
-as the weakening. After stating the rules we will describe what it means to
-compose context morphisms (terms of weakened contexts).
-\begin{align}
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{\Gamma}{B}
-    \jterm{\Gamma}{A}{x}
-    }
-  { \jfameq
-      {\Gamma}
-      {\subst{x}{\ctxwk{A}{B}}}
-      {B}
-    }
-  \label{cancellation-ws-c}\\
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{{\Gamma}{B}}{Q}
-    \jterm{\Gamma}{A}{x}
-    }
-  { \jfameq{{\Gamma}{B}}{\subst{x}{\ctxwk{A}{Q}}}{Q}
-    }
-  \label{cancellation-ws-f}\\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jterm{{\Gamma}{B}}{Q}{g}
-    }
-  { \jtermeq{{\Gamma}{B}}{Q}{\subst{x}{\ctxwk{A}{g}}}{g}
-    }
-  \label{cancellation-ws-t}
-\end{align}
-
-\subsubsection{The identity term of a substituted family}
-\label{comp-si}
-The identity term of a substituted family is the substitution of the identity term
-\begin{equation}
-\inference
-  { \jterm{\Gamma}{A}{x}
-    \jfam{{{\Gamma}{A}}{P}}{Q}
-    }
-  { \unfoldall{\jhomeq
-      {{\Gamma}{\subst{x}{P}}}
-      {\subst{x}{Q}}
-      {\subst{x}{Q}}
-      {\subst{x}{\idtm{Q}}}
-      {\idtm{\subst{x}{Q}}}
-    }}
-  \label{comp-si-t}
-\end{equation}
-
-\begin{rmk}
-An important special case is the judgmental equality
-\begin{equation*}
-\unfoldall{\jhomeq
-      {\Gamma}
-      {\subst{x}{P}}
-      {\subst{x}{P}}
-      {\subst{x}{\idtm{P}}}
-      {\idtm{\subst{x}{P}}}
-    }
-\end{equation*}
-for a family $\jfam{{\Gamma}{A}}{P}$.
-\end{rmk}
-
-\subsubsection{The cancellation properties of identity terms}
-\label{cancellation-i}
-Identity terms are determined by their behavior with respect to substitution combined with
-weakening. The identity terms will also be subject to compatibility rules.
-\begin{align}
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    }
-  { \jtermeq{\Gamma}{A}{\subst{x}{\idtm{A}}}{x}
-    }
-  \label{cancellation-si}\\
-& \inference
-  { \jfam{{\Gamma}{A}}{P}
-    }
-  { \jfameq{{\Gamma}{A}}{\subst{\idtm{A}}{\ctxwk{A}{P}}}{P}
-    }
-  \label{precomp-idtm-c}
-  \\
-& \inference
-  { \jfam{{{\Gamma}{A}}{P}}{Q}
-    }
-  { \jfameq{{{\Gamma}{A}}{P}}{\subst{\idtm{A}}{\ctxwk{A}{Q}}}{Q}
-    }
-  \label{precomp-idtm-f}
-  \\
-& \inference
-  { \jterm{{{\Gamma}{A}}{P}}{Q}{g}
-    }
-  { \jtermeq
-      {{{\Gamma}{A}}{P}}
-      {Q}
-      {\subst{\idtm{A}}{\ctxwk{A}{g}}}
-      {g}
-    }
-  \label{precomp-idtm-t}
-\end{align}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{The possiblity of types in the theory of contexts, families and
-terms}
-\label{types}
-We have deliberately not spoken of types so far because we have taken the point
-of view that a type in a context is nothing but a family in that context which
-belongs to the class of types. We think of types as \emph{irreducible} families,
-i.e.\ families which are neither empty nor the extension of two
-families which are both not empty (in algebraic terminology: which
-are both non-trivial). To allow ourselves to speak of types we introduce two
-new judgments: the judgment that something is a type and the judgment that two
-types are equal.
-\begin{align*}
-\jalign\jtype{\Gamma}{A} 
-& \jalign\jtypeeq{\Gamma}{A}{B}
-\end{align*}
-But only families of contexts are eligible to be types. If $A$ is a type
-in context $\Gamma$, then $A$ is also a family of contexts over $\Gamma$. 
-Moreover, two types in context $\Gamma$ are judgmentally equal precisely when they are equal
-as context families and if a family $B$ of contexts over $\Gamma$ is
-judgmentally equal to a type $A$ in context $\Gamma$, then $B$ is a type in
-context $\Gamma$. This is expressed by the following four inference rules:
-\begin{align*}
-& \inference
-  { \jtype{\Gamma}{A}
-    }
-  { \jfam{\Gamma}{A}
-    }
-& & \inference
-    { \jtypeeq{\Gamma}{A}{B}
-      }
-    { \jfameq{\Gamma}{A}{B}
-      }
-    \\
-& \inference
-  { \jtype{\Gamma}{A}
-    \jfameq{\Gamma}{A}{B}
-    }
-  { \jtype{\Gamma}{B}
-    }
-& & \inference
-    { \jtype{\Gamma}{A}
-      \jfameq{\Gamma}{A}{B}
-      }
-    { \jtypeeq{\Gamma}{A}{B}
-      }
-\end{align*}
-As pointed out at the beginning of this subsection, 
-we do not assume that the empty family is a type, that would be like
-assuming that the multiplicative unit of a ring is prime.
-
-\begin{comment}
-We envision a
-categorical interpretation of the theory of contexts, families and terms where 
-contexts are interpreted as finite paths of arrows belonging to a predetermined
-class. The empty family would be such a path of length $0$ whereas the types
-would be such paths of length $1$.
-\end{comment}
-
-We add rules asserting that a weakened type is again a type that 
-substitution preserves the property of being a type:
-\begin{align}
-& \inference
-  { \jfam{\Gamma}{B}
-    \jtype{\ctxext{\Gamma}{B}}{Q}
-    }
-  { \jtype{\ctxext{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{Q}}
-    }
-  \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jtype{\ctxext{{\Gamma}{A}}{P}}{Q}
-    }
-  { \jtype{\ctxext{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}
-    }
-\end{align}
-With only the current rules, the possibility of making the judgment that
-something is a type does not add much to the theory of contexts, families and
-terms. Nevertheless, when studying models, having an interpretation for the
-judgment that something is a type will allow for the possibility to study
-conditions such as the one asserting that every family `factorizes' uniquely
-as multiple applications of extension to types, analogous to the condition on
-unique factorization domains in ring theory. All the tradional models of type
-theory should translate to models with such a condition, simply because
-traditionally contexts are viewed as lists of variable (and type) declarations.

### 2014-10-28 Rules/rules_intro.tex
moved rules for empty context and families

@@
 In the current presentation, contexts can be seen as binary
-planar trees of which the leaves are (families of) contexts. 
+planar rooted trees of which the leaves are (families of) contexts. 
 The judgmental equality relation on contexts is an equivalence relation which 
-expresses that binary planar
+expresses that binary planar rooted
 trees of contexts are judgmentally equal if their leaves are, taking only

### 2014-10-28 Junk/article-models-constructors.tex
moved junk

--- 2013-11-11 article-models-constructors.tex
@@
-\section{Variable-free type theory -- other type constructors}
+\section{Introducing the type constructors}
 We will now describe the rules for the type constructors that we don't assume
@@
 
+\begin{comment}
 \section{Binary trees}
@@
 \end{proof}
+\end{comment}

### 2014-10-28 Junk/article-models-ttpretty.tex
moved junk

\section{Pretty type theory}
In this section we shall discuss conventions to algorithmically `pretty-print' judgments of
the theory of contexts, families and terms. These will include variable
declarations in contexts and omission of explicit notation of weakening. This
will render various notationally different expressions identical and thus
we shall have to show that when this happens, the expressions under consideration
were already judgmentally equal. Pretty type theory could be seen as a
semi-formal version of the theory of contexts, families and terms and it
will facilitate translation to the informal language developed in \cite{TheBook}.

\begin{comment}
\subsection{The basic judgments}
The basic judgments of pretty type theory are the same as for structural type
theory. There are judgments for: ``$\Gamma$ is a context'',
``$A(i)$ over $i:\Gamma$ is a family over $\Gamma$'', ``$A(i)$ over $i:\Gamma$ 
is a type in context $\Gamma$''
and ``$x(i)$ is a term of $A(i)$ above $i:\Gamma$''. The other four
judgments are for judgmental equality.

\begin{align*}
\jvctx*{\Gamma} & \jvctxeq*{\Gamma}{\Gamma'}\\
\jvfam*{i}{\Gamma}{A} & \jvfameq*{i}{\Gamma}{A}{B}\\
\jvtype*{i}{\Gamma}{A} & \jvtypeeq*{i}{\Gamma}{A}{B}\\
\jvterm*{i}{\Gamma}{A}{x} & \jvtermeq*{i}{\Gamma}{A}{x}{y}.
\end{align*}

We have the following basic inference rules that relate types and families:

\begin{small}
\begin{align*}
& \inference
  {\jvtype{i}{\Gamma}{A}}
  {\jvfam{i}{\Gamma}{A}}
& & \inference
    {\jvtypeeq{i}{\Gamma}{A}{B}}
    {\jvfameq{i}{\Gamma}{A}{B}}\\
& \inference
  {\jvtype{i}{\Gamma}{A}
   \jvfameq{i}{\Gamma}{A}{B}}
  {\jvtype{i}{\Gamma}{B}}
& & \inference
    {\jvtype{i}{\Gamma}{A}
     \jvfameq{i}{\Gamma}{A}{B}}
    {\jvtypeeq{i}{\Gamma}{A}{B}}
\end{align*}
\end{small}

\subsection{The basic rules for judgmental equality}
The rules for judgmental equality establish that it is an equivalence relation.
\bgroup\small
\begin{align*}
& \inference
  {\jvctx{\Gamma}}
  {\jvctxeq{\Gamma}{\Gamma}} 
& & \inference
    {\jvctxeq{\Gamma}{\Delta}}
    {\jvctxeq{\Delta}{\Gamma}} 
& & \inference
    {\jvctxeq{\Gamma}{\Delta}
     \jvctxeq{\Delta}{\greek{E}}}
    {\jvctxeq{\Gamma}{\greek{E}}}\\
& \inference
  {\jvfam{i}{\Gamma}{A}}
  {\jvfameq{i}{\Gamma}{A}{A}} 
& & \inference
    {\jvfameq{i}{\Gamma}{A}{B}}
    {\jvfameq{i}{\Gamma}{B}{A}}
& & \inference
    {\jvfameq{i}{\Gamma}{A}{B}
     \jvfameq{i}{\Gamma}{B}{C}}
    {\jvfameq{i}{\Gamma}{A}{C}}\\
& \inference
  {\jvterm{i}{\Gamma}{A}{x}}
  {\jvtermeq{i}{\Gamma}{A}{x}{x}}
& & \inference
    {\jvtermeq{i}{\Gamma}{A}{x}{y}}
    {\jvtermeq{i}{\Gamma}{A}{y}{x}}
& & \inference
    {\jvtermeq{i}{\Gamma}{A}{x}{y}
     \jvtermeq{i}{\Gamma}{A}{y}{z}}
    {\jvtermeq{i}{\Gamma}{A}{x}{z}}
\end{align*}
\egroup

The following convertibility rules are responsible for the strictness
of judgmental equality, which sets it apart from equivalences or identifications:

\begin{align*}
& \inference
  {\jvctxeq{\Gamma}{\Delta}
   \jvfam{i}{\Gamma}{A}}
  {\jvfam{i}{\Delta}{A}}
& & \inference
    {\jvctxeq{\Gamma}{\Delta}
     \jvfameq{i}{\Gamma}{A}{B}}
    {\jvfameq{i}{\Delta}{A}{B}}\\
& \inference
  {\jvctxeq{\Gamma}{\Delta}
   \jvterm{i}{\Gamma}{A}{x}}
  {\jvterm{i}{\Delta}{A}{x}}
& & \inference
    {\jvctxeq{\Gamma}{\Delta}
     \jvtermeq{i}{\Gamma}{A}{x}{y}}
    {\jvtermeq{i}{\Delta}{A}{x}{y}}\\
& \inference
  {\jvfameq{i}{\Gamma}{A}{B}
   \jvterm{i}{\Gamma}{A}{x}}
  {\jvterm{i}{\Gamma}{B}{x}}
& & \inference
    {\jvfameq{i}{\Gamma}{A}{B}
     \jvtermeq{i}{\Gamma}{A}{x}{y}}
    {\jvtermeq{i}{\Gamma}{B}{x}{y}}
\end{align*}

\subsection{The empty context}
The empty context looks a bit strange when we explicitly denote the terms. But
we will not do so anymore after this subsection.

\begin{align}
& \inference
  {}
  {\jctx{\emptyc}}\\
& \inference
  {\jctx{\Gamma}}
  {\jvfam{i}{\Gamma}{\emptyf[\Gamma]}}\\
& \inference
  {\jctx{\Gamma}}
  {\jvterm{i}{\Gamma}{\emptyf[\Gamma]}{\emptytm[\Gamma]}}\\
& \inference
  {\jvterm{i}{\Gamma}{\emptyf[\Gamma]}{x}}
  {\jvtermeq{i}{\Gamma}{\emptyf[\Gamma]}{x}{\emptytm[\Gamma]}}
\end{align}

Moreover, if $\Gamma$ is a context family over the
empty context, then $\Gamma$ is a context and every context is a context
family over the empty context. Note that this allows us to speak
of terms of contexts too.

\begin{align}
& \inference
  {\jctx{\Gamma}}
  {\jvfam{\nameless}{\emptyc}{\Gamma}} 
& & \inference
    {\jvfam{\nameless}{\emptyc}{\Gamma}}
    {\jctx{\Gamma}}\\
& \inference
  {\jctxeq{\Gamma}{\Delta}}
  {\jvfameq{\nameless}{\emptyc}{\Gamma}{\Delta}}
& & \inference
    {\jvfameq{\nameless}{\emptyc}{\Gamma}{\Delta}}
    {\jctxeq{\Gamma}{\Delta}}
\end{align}

\subsubsection{The empty context is compatible with itslef}
The empty context $\emptyc$ may be considered as a family of contexts over the empty
context. When we do this, we get $\emptyf[\emptyc]$.
\begin{equation}
\inference
  {}
  {\jvfameq{\nameless}{\emptyc}{\emptyc}{\emptyf[\emptyc]}}
\end{equation}
In the future, we shall denote $\emptyf[\Gamma]$ by $\emptyf$. The above rule
guarantees that this will not cause confusion. Likewise, we shall denote
$\emptytm[\Gamma]$ by $\emptytm$.

\subsection{Extension}
We introduce extension which not only extends a context $\Gamma$ and a family
$A$ over it to a context $\ctxext{\Gamma}{A}$, but which also extends a family $A$
in context $\Gamma$ and a family $P$ over it to a family $\ctxext{A}{P}$ over context
$\Gamma$. We do this to ensure that all of type theory can be done in a context.
For instance, we could say (1) that a context in context $\Gamma$ is the same thing
as a family over $\Gamma$; (2) When $A$ is a context in this sense, a family over
$A$ is the same thing as a family $P$ over $\ctxext{\Gamma}{A}$ and 
(3) when $P$ is a family over $A$ in this sense, a term of $P$ keeps its original meaning.

\begin{align}
& \inference
  {\jvfam{i}{\Gamma}{A}}
  {\jvfamcombi{{i}{x}}{{\Gamma}{A}}{P}}
& & \inference
    {\jctxeq{\Gamma}{\Delta}
     \jfameq{\Gamma}{A}{B}}
    {\jctxeq{\ctxext{\Gamma}{A}}{\ctxext{\Delta}{B}}}\\
& \inference
  {\jfam{{\Gamma}{A}}{P}}
  {\jfam{\Gamma}{\ctxext{A}{P}}}
& & \inference
    {\jfameq{\Gamma}{A}{B}
     \jfameq{{\Gamma}{A}}{P}{Q}}
    {\jfameq{\Gamma}{\ctxext{A}{P}}{\ctxext{B}{Q}}}
\end{align}

\subsubsection{Extension is compatible with the empty context}
The following rule asserts that extension by $\emptyc$ leaves the contexts unchanged.
\begin{align}
& \inference
  {\jctx{\Gamma}}
  {\jctxeq{\ctxext{\emptyc}{\Gamma}}{\Gamma}}\\
& \inference
  {\jctx{\Gamma}}
  {\jctxeq{\ctxext{\Gamma}{\emptyf}}{\Gamma}}\\
& \inference
  {\jfam{\Gamma}{A}}
  {\jfameq{\Gamma}{\ctxext{\emptyf}{A}}{A}}
\end{align}

\subsubsection{Extension is compatible with itself}
The inference rules asserting that extension is compatible with itself assert
that contexts are unstructured lists of type declarations. This rule is
unavoidable if we want that for a family $A$ in context $\Gamma$, a family over
$A$ is the same thing as a family over $\ctxext{\Gamma}{A}$.

\begin{align}
& \inference
  {\jfam{\Gamma}{A}
   \jfam{{\Gamma}{A}}{P}}
  {\jctxeq{\ctxext{{\Gamma}{A}}{P}}{\ctxext{\Gamma}{{A}{P}}}}\\
& \inference
  {\jfam{{\Gamma}{A}}{P}
   \jfam{{{\Gamma}{A}}{P}}{Q}}
  {\jfameq{\Gamma}{\ctxext{{A}{P}}{Q}}{\ctxext{A}{{P}{Q}}}}
\end{align}
\end{comment}

### 2014-10-28 Internal_models/article-models-ttinternal.tex
recovered those lost files

--- 2014-09-20 article-models-ttinternal.tex
@@
 pre-weakening homomorphism if the diagram
-\begin{equation*}
+\begin{equation*}\label{diag:cftwk-cftwk}
 \begin{tikzcd}[column sep=8em]
@@
 \end{align*}
-Thus we may conclude that the diagram above a pullback of the diagram in \autoref{diag:cftwk-cftwk}
+Thus we may conclude that the diagram above is a pullback of the diagram in \autoref{diag:cftwk-cftwk}
 by $\cftctxext$ and therefore it commutes as desired.
@@
 \subsection{E-adjunctions}
-
-\subsection{Pre-universes}
-Pre-universes are internal versions of the theory of contexts, families and
-terms. They interpret extension, the empty context, weakening, substitution
-and identity terms all at once in a compatible way. Besides the compatibility
-properties there will be judgmental equalities analogous to the cancellation
-properties of \autoref{cancellation-ws,cancellation-i}. Pre-universes are to
-the theory of contexts, families and terms what internal categories to a
-category.

### 2014-11-01 Algebras/article-models-cat.tex
extension objects with empty families

\section{E-objects in categories with finite limits}
In this section we assume that $\cat{C}$ is a finitely complete category and
whenever we write a pullback, we assume that it is chosen. Recall that for
any morphism $f:A\to B$ in a category $\cat{C}$ with chosen pullbacks, there
is a functor
\begin{equation*}
f^\ast : \cat{C}/B\to\cat{C}/A.
\end{equation*}
As usual, when $g:X\to B$ is a morphism, we will write $f^\ast(X)$ for the
domain of $f^\ast(g)$. When there is more than one morphism $X\to B$ involved,
as will be the case below, we will write $\pullback{A}{X}{f}{g}$. The projections
will be written as $\pullbackpr{1}{f}{g}$ and $\pullbackpr{2}{f}{g}$. So in this notation, a
typical pullback diagram has the following form:
\begin{equation*}
\begin{tikzcd}[column sep=large]
\pullback{A}{X}{f}{g}
  \ar{r}{\pullbackpr{1}{f}{g}}
  \ar{d}[swap]{\pullbackpr{2}{f}{g}}
  &
A \ar{d}{f}
  \\
X \ar{r}[swap]{g}
  &
B
\end{tikzcd}
\end{equation*}
Also, when we have a commutative diagram of the form
\begin{equation*}
\begin{tikzcd}
A \ar{r}{f}
  \ar{d}{a}
  &
X \ar{d}
  & 
B \ar{l}[swap]{g}
  \ar{d}{b}
  \\
A'
  \ar{r}[swap]{f'}
  &
X'
  &
B'
  \ar{l}{g'}
\end{tikzcd}
\end{equation*}
we will denote the unique map from $\pullback{A}{B}{f}{g}$ to $\pullback{A'}{B'}{f'}{g'}$
such that the diagram
\begin{equation*}
\begin{tikzcd}
  {}
  & 
\pullback{A'}{B'}{f'}{g'}
  \ar{dd}
  \ar{rr}
  &
  &
B'
  \ar{dd}{g'}
  \\
\pullback{A}{B}{f}{g}
  \ar{dd}
  \ar[crossing over]{rr}
  \ar[dotted]{ur}{\pullback{a}{b}{f'}{g'}}
  &
  &
B \ar{ur}{b}
  \\
  {}
  &
A'
  \ar{rr}
  &
  &
X'
  \\
A \ar{rr}[swap]{f}
  \ar{ur}{a}
  &
  &
X \ar[crossing over,leftarrow]{uu}[near end,swap]{g}
  \ar{ur}
\end{tikzcd}
\end{equation*}
commutes, by $\pullback{a}{b}{f'}{g'}$. In the current work, we shall
write $A\times B$ for the pullback of $A\rightarrow 1\leftarrow B$, and
$\pi_1$ and $\pi_2$ for its projections (thus, no separate choice of
cartesian products is made).

\subsection{Extension objects}
\begin{defn}
A \emph{pre-extension object $\stesys$ in $\cat{C}$} consists of a \emph{fundamental structure}, which is a diagram of the form
\begin{equation*}
\begin{tikzcd}
\stesyst
  \ar{d}[swap]{\ebd}
  \\
\stesysf
  \ar{d}[swap]{\eft}
  \\
\stesysc
\end{tikzcd}
\end{equation*}
in $\cat{C}$ together with the \emph{context extension} and \emph{family extension} operations
\begin{align*}
\ectxext &:\stesysf\to \stesysc\\
\efamext & :\stesysff\to \stesysf,
\end{align*}
respectively, such that the diagram
\begin{equation*}
\begin{tikzcd}
\stesysf_2 
  \ar{r}{\efamext} 
  \ar{d}[swap]{\eft[1]} 
  & 
\stesysf 
  \ar{d}{\eft}
  \\
\stesysf
  \ar{r}[swap]{\eft} 
  & 
\stesysc
\end{tikzcd}
\end{equation*}
commutes.
\end{defn}

\begin{defn}
We introduce the following notation:
\begin{align*}
\stesysf_2 
  & := \stesysff
  \\
\eft[1] 
  & := \pullbackpr{1}{\ectxext}{\eft} : \stesysf_2\to\stesysf
  \\
\stesysf_3 & := \pullback{\stesysf_2}{\stesysf_2}{\efamext}{\eft[1]}
  \\
\eft[2]
  & := \pullbackpr{1}{\efamext}{\eft[1]} : \stesysf_3\to\stesysf_2.
\end{align*}
Then it follows that the outer square in the diagram
\begin{equation*}
\begin{tikzcd}[column sep=large]
\stesysf_3
  \ar[dotted]{dr}{\eext{2}}
  \ar{rr}{\pullback{\pullbackpr{2}{\ectxext}{\eft}}{\pullbackpr{2}{\ectxext}{\eft}}{\ectxext}{\eft}}
  \ar{dd}[swap]{\eft[2]}
  & 
  &
\stesysf_2
  \ar{d}{\efamext}
  \\
  &
\stesysf_2
  \ar{d}[swap]{\eft[1]}
  \ar{r}{\pullbackpr{2}{\ectxext}{\eft}}
  &
\stesysf
  \ar{d}{\eft}
  \\
\stesysf_2
  \ar{r}[swap]{\eft[1]}
  &
\stesysf
  \ar{r}[swap]{\ectxext}
  &
\stesysc
\end{tikzcd}
\end{equation*}
commutes. We define $\eext{2}$ to be the unique morphism rendering the above diagram
commutative. Now we may continue to define
\begin{align*}
\stesysf_4 
  & := 
\pullback{\stesysf_3}{\stesysf_3}{\eext{2}}{\eft[2]}
  \\
\eft[3] 
  & := 
\pullbackpr{1}{\eext{2}}{\eft[2]}.
\end{align*}
Then we see that the outer square of the diagram
\begin{equation*}
\begin{tikzcd}[column sep=large]
\stesysf_4
  \ar[dotted]{dr}{\eext{3}}
  \ar{rr}{\pullback{\pullbackpr{2}{\efamext}{\eft[1]}}{\pullbackpr{2}{\efamext}{\eft[1]}}{\efamext}{\eft[1]}}
  \ar{dd}[swap]{\eft[3]}
  & 
  &
\stesysf_3
  \ar{d}{\eext{2}}
  \\
  &
\stesysf_3
  \ar{d}[swap]{\eft[2]}
  \ar{r}{\pullbackpr{2}{\efamext}{\eft[1]}}
  &
\stesysf_2
  \ar{d}{\eft[1]}
  \\
\stesysf_3
  \ar{r}[swap]{\eft[2]}
  &
\stesysf_2
  \ar{r}[swap]{\efamext}
  &
\stesysf
\end{tikzcd}
\end{equation*}
commutes,
so we may define $\eext{3}$ to be the unique map which renders the diagram. It
is straightforward to continue this process by induction, but we shall need not
go any further in this article.
\end{defn}

\begin{defn} An extension object is a pre-extension object $\stesys$ for which 
the diagrams
\begin{equation*}
\begin{tikzcd}
\stesysf_2 
  \ar{d}[swap]{\pullbackpr{2}{\ectxext}{\eft}} 
  \ar{r}{\efamext} 
  & 
\stesysf 
  \ar{d}{\ectxext}
  \\
\stesysf 
  \ar{r}[swap]{\ectxext} 
  & 
\stesysc
\end{tikzcd}
\qquad
\begin{tikzcd}
\stesysf_3
  \ar{d}[swap]{\pullbackpr{2}{\efamext}{\eft[1]}}
  \ar{r}{\eext{2}}
  & 
\stesysf_2 
  \ar{d}{\efamext} 
  \\
\stesysf_2 
  \ar{r}[swap]{\efamext} 
  &
\stesysf
\end{tikzcd}
\end{equation*}
commute.
\end{defn}

\begin{comment}
\begin{lem}
There exists an isomorphism $\alpha$ such that the triangle
\begin{equation*}
\begin{tikzcd}[column sep=tiny]
\pullback{\stesysf}{\stesysf_2}{\ectxext}{\eft\circ\eft[1]}
  \ar[dotted]{rr}{\alpha}
  \ar{dr}[swap]{\pullback{\catid{\stesysf}}{\efamext}{\ectxext}{\eft}}
  &
  &
\stesysf_3
  \ar{dl}{\eext{2}}
  \\
& \stesysf_2
\end{tikzcd}
\end{equation*}
commutes
\end{lem}

\begin{proof}
There is a unique morphism $\alpha:
\pullback{\stesysf}{\stesysf_2}{\ectxext}{\eft\circ\eft[1]}\to\stesysf_3$
rendering the diagram
\begin{equation*}
\begin{tikzcd}[column sep=large]
\pullback{\stesysf}{\stesysf_2}{\ectxext}{\eft\circ\eft[1]}
  \ar[bend left=10,yshift=.5ex]{drrr}{\pullbackpr{2}{\ectxext}{\eft}\circ\pullbackpr{2}{\ectxext}{\eft\circ\eft[1]}}
  \ar[bend right=10]{ddr}[swap]{\pullback{\catid{\stesysf}}{\eft[1]}{\ectxext}{\eft}}
  \ar[dotted]{dr}{\alpha}
  \\
& \stesysf_3
  \ar{r}{\pullbackpr{2}{\efamext}{\eft[1]}}
  \ar{d}{\eft[2]}
  &
\stesysf_2
  \ar{d}[swap]{\eft[1]}
  \ar{r}[swap]{\pullbackpr{2}{\ectxext}{\eft}}
  &
\stesysf
  \ar{d}{\eft}
  \\
{} & \stesysf_2
  \ar{r}[swap]{\efamext}
  &
\stesysf
  \ar{r}[swap]{\ectxext}
  &
\stesysc
\end{tikzcd}
\end{equation*}
\end{proof}
\end{comment}

\begin{defn}
Suppose $\stesys$ is a pre-extension object of $\cat{C}$. Then we define the pre-extension object
$\famesys{\stesys}$ to consist of the fundamental structure
\begin{equation*}
\begin{tikzcd}
\stesyst_2
  \ar{d}{\ebd[1]}
  \\
\stesysf_2
  \ar{d}{\eft[1]}
  \\
\stesysf
\end{tikzcd}
\end{equation*}
where
\begin{align*}
\stesyst_2 
  & := \pullback{\stesysf}{\stesyst}{\ectxext}{\eft\circ\ebd}
  \\
\ebd[1]
  & := \ectxext^\ast(\ebd),
\end{align*}
with the extension operations
\begin{align*}
\efamext 
  & 
  : \stesysf_2\to\stesysf\\
\eext{2} & : \stesysf_3\to\stesysf_2.
\end{align*}
\end{defn}

In \autoref{famextobj} below we shall show that $\famesys{\stesys}$ is an
extension algebra whenever $\stesys$ is an extension algebra. We shall need
a handful of lemmas to give the proof.

\begin{defn}
We define
\begin{align*}
\beta_1 
  & := 
\pullbackpr{2}{\ectxext}{\eft}
  & &
  : \stesysf_2\to\stesysf
  \\
\beta_2
  & :=
\pullback{\beta_1}{\beta_1}{\ectxext}{\eft}
  & &
  : \stesysf_3\to\stesysf_2
  \\
\beta_3
  & :=
\pullback{\beta_2}{\beta_2}{\efamext}{\eft[1]}
  & &
  : \stesysf_4\to\stesysf_3.
\end{align*}
\end{defn}

\begin{lem}
Let $\stesys$ be a pre-extension object. Then the square
\begin{equation*}
\begin{tikzcd}[column sep=10em]
\stesysf_4
  \ar{r}{\pullback{\pullbackpr{2}{\efamext}{\eft[1]}}{\pullbackpr{2}{\efamext}{\eft[1]}}{\efamext}{\eft[1]}}
  \ar{d}[swap]{\beta_3}
  &
\stesysf_3
  \ar{d}{\beta_2}
  \\
\stesysf_3
  \ar{r}[swap]{\beta_2}
  &
\stesysf_2
\end{tikzcd}
\end{equation*}
commutes.
\end{lem}

\begin{proof}
Left to the reader.
\end{proof}

\begin{comment}
\begin{proof}
It is straightforward to verify the equalities
\begin{align*}
\pullbackpr{1}{\ectxext}{\eft}\circ\beta_2\circ
  (\pullback{\pullbackpr{2}{\efamext}{\eft[1]}}{\pullbackpr{2}{\efamext}{\eft[1]}}{\efamext}{\eft[1]})
  & =
\pullbackpr{1}{\ectxext}{\eft}\circ\beta_2\circ\beta_3
  \\
\pullbackpr{2}{\ectxext}{\eft}\circ\beta_2\circ
  (\pullback{\pullbackpr{2}{\efamext}{\eft[1]}}{\pullbackpr{2}{\efamext}{\eft[1]}}{\efamext}{\eft[1]})
  & =
\pullbackpr{2}{\ectxext}{\eft}\circ\beta_2\circ\beta_3.\qedhere
\end{align*}
\end{proof}
\end{comment}

Note that the square
\begin{equation*}
\begin{tikzcd}
\stesysf_3
  \ar{r}{\beta_2}
  \ar{d}[swap]{\eext{2}}
  &
\stesysf_2
  \ar{d}{\efamext}
  \\
\stesysf_2
  \ar{r}[swap]{\beta_1}
  &
\stesysf
\end{tikzcd}
\end{equation*}
commutes by definition. We have a similar result relating $\eext{3}$ and
$\eext{2}$.

\begin{lem}
Let $\stesys$ be a pre-extension object. Then the square
\begin{equation*}
\begin{tikzcd}
\stesysf_4
  \ar{r}{\beta_3}
  \ar{d}[swap]{\eext{3}}
  &
\stesysf_3
  \ar{d}{\eext{2}}
  \\
\stesysf_3
  \ar{r}[swap]{\beta_2}
  &
\stesysf_2
\end{tikzcd}
\end{equation*}
commutes.
\end{lem}

\begin{proof}
Left to the reader.
\end{proof}
\begin{comment}
\begin{proof}
It is straightforward to verify the equalities
\begin{align*}
\pullbackpr{1}{\ectxext}{\eft}\circ\beta_2\circ\eext{3}
  & = \beta_1\circ\eft[2]\circ\eft[3]
  \\
\pullbackpr{1}{\ectxext}{\eft}\circ\eext{2}\circ\beta_3
  & = \beta_1\circ\eft[2]\circ\eft[3].
\end{align*}
Thus, it remains to verify that
\begin{equation*}
\pullbackpr{2}{\ectxext}{\eft}\circ\beta_2\circ\eext{3}
  = \pullbackpr{2}{\ectxext}{\eft}\circ\eext{2}\circ\beta_3.
\end{equation*}
It is straightforward to see that the diagram
\begin{equation*}
\begin{tikzcd}[column sep=large]
\stesysf_4
  \ar{dd}[swap]{\pullback{\pullbackpr{2}{\efamext}{\eft[1]}}{\pullbackpr{2}{\efamext}{\eft[1]}}{\efamext}{\eft[1]}}
  \ar{r}{\eext{3}}
  &
\stesysf_3
  \ar{r}{\beta_2}
  \ar{d}[swap]{\pullbackpr{2}{\efamext}{\eft[1]}}
  &
\stesysf_2
  \ar{dd}{\pullbackpr{2}{\ectxext}{\eft}}
  \\
  {} &
\stesysf_2
  \ar{dr}{\beta_1}
  \\
\stesysf_3
  \ar{ur}{\eext{2}}
  \ar{r}[swap]{\beta_2}
  &
\stesysf_2
  \ar{r}[swap]{\efamext}
  &
\stesysf
\end{tikzcd}
\end{equation*}
commutes. It is likewise straightforward to see that the diagram
\begin{equation*}
\begin{tikzcd}
\stesysf_4
  \ar{r}{\beta_3}
  \ar{d}[swap]{\pullback{\pullbackpr{2}{\efamext}{\eft[1]}}{\pullbackpr{2}{\efamext}{\eft[1]}}{\efamext}{\eft[1]}}
  &
\stesysf_3
  \ar{r}{\eext{2}}
  \ar{d}[swap]{\beta_2}
  &
\stesysf_2
  \ar{d}{\pullbackpr{2}{\ectxext}{\eft}}
  \\
\stesysf_3
  \ar{r}[swap]{\beta_2}
  &
\stesysf_2
  \ar{r}[swap]{\efamext}
  &
\stesysf
\end{tikzcd}
\end{equation*}
commutes, completing our goal.
\end{proof}
\end{comment}

\begin{thm}[Local extension structure]\label{famextobj}
If $\stesys$ is an extension object, then so is $\famesys{\stesys}$.
\end{thm}

\begin{proof}
Note that the diagram
\begin{equation*}
\begin{tikzcd}
\stesysf_3
  \ar{d}[swap]{\pullbackpr{2}{\efamext}{\eft[1]}}
  \ar{r}{\eext{2}}
  & 
\stesysf_2 
  \ar{d}{\efamext} 
  \\
\stesysf_2 
  \ar{r}[swap]{\efamext} 
  &
\stesysf
\end{tikzcd}
\end{equation*}
commutes by assumption. For the second condition, we have to show that the
diagram
\begin{equation*}
\begin{tikzcd}
\stesysf_4
  \ar{d}[swap]{\pullbackpr{2}{\eext{2}}{\eft[2]}}
  \ar{r}{\eext{3}}
  & 
\stesysf_3
  \ar{d}{\eext{2}} 
  \\
\stesysf_3
  \ar{r}[swap]{\eext{2}} 
  &
\stesysf_2
\end{tikzcd}
\end{equation*}
Since this is a question about two maps into a pullback, it suffices to verify
that
\begin{align*}
\pullbackpr{1}{\ectxext}{\eft}\circ\eext{2}\circ\eext{3}
  & =
\pullbackpr{1}{\ectxext}{\eft}\circ\eext{2}\circ\pullbackpr{2}{\eext{2}}{\eft[2]}
  \\
\pullbackpr{2}{\ectxext}{\eft}\circ\eext{2}\circ\eext{3}
  & =
\pullbackpr{2}{\ectxext}{\eft}\circ\eext{2}\circ\pullbackpr{2}{\eext{2}}{\eft[2]}.
\end{align*}
For the first equality, it is fairly straightforward to show that both the
equalities
\begin{equation*}
\pullbackpr{1}{\ectxext}{\eft}\circ\eext{2}\circ\eext{3}
  =
\eft[1]\circ\eft[2]\circ\eft[3]
\end{equation*}
and
\begin{equation*}
\pullbackpr{1}{\ectxext}{\eft}\circ\eext{2}\circ\pullbackpr{2}{\eext{2}}{\eft[2]}
  =
\eft[1]\circ\eft[2]\circ\eft[3].
\end{equation*}
hold. For the second subgoal (which is more tricky). Notice first that the
diagram
\begin{equation*}
\begin{tikzcd}
\stesysf_4
  \ar{r}{\eext{3}}
  \ar{d}[swap]{\beta_3}
  &
\stesysf_3
  \ar{r}{\eext{2}}
  \ar{d}[swap]{\beta_2}
  &
\stesysf_2
  \ar{d}{\pullbackpr{2}{\ectxext}{\eft}}
  \\
\stesysf_3
  \ar{r}[swap]{\eext{2}}
  &
\stesysf_2
  \ar{r}[swap]{\efamext}
  &
\stesysf
\end{tikzcd}
\end{equation*}
commutes. We also have the commutative diagram
\begin{equation*}
\begin{tikzcd}[column sep=large]
\stesysf_4
  \ar{r}{\pullbackpr{2}{\eext{2}}{\eft[2]}}
  \ar{d}[swap]{\beta_3}
  &
\stesysf_3
  \ar{r}{\eext{2}}
  \ar{d}[swap]{\beta_2}
  &
\stesysf_2
  \ar{d}{\pullbackpr{2}{\ectxext}{\eft}}
  \\
\stesysf_3
  \ar{r}{\pullbackpr{2}{\efamext}{\eft[1]}}
  \ar{dr}[swap]{\eext{2}}
  &
\stesysf_2
  \ar{r}{\efamext}
  &
\stesysf
  \\
  {} &
\stesysf_2
  \ar{ur}[swap]{\efamext}
\end{tikzcd}
\end{equation*}
completing the proof.
\end{proof}

\subsection{(Pre-)extension homomorphisms}\label{subsection:e_extension_homomorphisms}
In this subsection we start with the study of pre-extension homomorphisms, which
will include the extension homomorphisms since they will be the pre-extension
homomorphisms of which both the domain and codomain are extension objects.
Our main examples of extension homomorphisms will be the operations of weakening
and substitution. There are some basic examples of pre-extension homomorphisms
that will be useful too, which get introduced in the this section and in
\autoref{subsection:change_of_base}. In this section, we will mainly be
interested in pre-extension homomorphisms between local pre-extension objects.
We will end this section by proving that a retract of an extension object is
always an extension object.

\begin{defn}
Let $\stesys$ and $\stesys'$ be pre-extension algebras. A \emph{pre-extension 
homomorphism $f$ from $\stesys$ to $\stesys'$} is a triple $(f_0,f_1,f^t)$ 
consisting of morphisms
\begin{equation*}
\begin{tikzcd}
\stesyst 
  \ar{r}{f^t}
  \ar{d}[swap]{\ebd}
  &
\stesyst'
  \ar{d}{\ebd'}
  \\
\stesysf 
  \ar{r}{f_1}
  \ar{d}[swap]{\eft}
  &
\stesysf'
  \ar{d}{\eft'}
  \\
\stesysc 
  \ar{r}[swap]{f_0}
  &
\stesysc'
\end{tikzcd}
\end{equation*}
such that the indicated squares commute, for which furthermore the squares
\begin{equation*}
\begin{tikzcd}
\stesysf \ar{r}{f_1}
  \ar{d}[swap]{\ectxext}
  &
\stesysf'
  \ar{d}{\ectxext'}
  \\
\stesysc
  \ar{r}[swap]{f_0}
  &
\stesysc'
\end{tikzcd}
\end{equation*}
and
\begin{equation*}
\begin{tikzcd}[column sep=large]
\stesysf\times_{\ectxext,\eft} \stesysf
  \ar{r}{f_1\times_{\ectxext',\eft'} f_1}
  \ar{d}[swap]{\efamext}
  &
\stesysf'\times_{\ectxext',\eft'} \stesysf'
  \ar{d}{\efamext'}
  \\
\stesysf
  \ar{r}[swap]{f_1}
  &
\stesysf'
\end{tikzcd}
\end{equation*}
Composition and the identity homomorphism are defined in the expected way. We
define furthermore
\begin{align*}
f_2 & := \pullback{f_1}{f_1}{\ectxext'}{\eft'}
  \\
f_3 & := \pullback{f_2}{f_2}{\efamext'}{\eft[1]'}.
\end{align*}
\end{defn}

\begin{defn}
A pre-extension homomorphism between extension objects is called an extension
homomorphism.
\end{defn}

\begin{lem}
Let $\stesys$ be an extension object. Then
\begin{equation*}
\begin{tikzcd}[column sep=large]
\stesyst_2
  \ar{r}{\pullbackpr{2}{\ectxext}{\eft\circ\ebd}}
  \ar{d}[swap]{\ebd[1]}
  &
\stesyst
  \ar{d}{\ebd}
  \\
\stesysf_2
  \ar{r}{\pullbackpr{2}{\ectxext}{\eft}}
  \ar{d}[swap]{\eft[1]}
  &
\stesysf
  \ar{d}{\eft}
  \\
\stesysf
  \ar{r}[swap]{\ectxext}
  &
\stesysc
\end{tikzcd}
\end{equation*}
assembles an extension homomorphism $\mathbf{e}_0:\famesys{\stesys}\to\stesys$.
Likewise, we have an extension homomorphism
$\mathbf{e}_1:\famesys{\famesys{\stesys}}\to\famesys{\stesys}$. Thus, a
pre-extension object is an extension object if and only if $\mathbf{e}_0$
and $\mathbf{e}_1$ are pre-extension homomorphisms.
\end{lem}

\begin{proof}
Immediate from the conditions of being an extension object.
\end{proof}

\begin{defn}
Let $\stesys$ be a pre-extension object. Then
\begin{equation*}
\begin{tikzcd}
\stesyst_3
  \ar{r}{\beta^t}
  \ar{d}[swap]{\ebd[2]}
  &
\stesyst_2
  \ar{d}{\ebd[1]}
  \\
\stesysf_3
  \ar{r}{\beta_2}
  \ar{d}[swap]{\eft[2]}
  &
\stesysf_2
  \ar{d}{\eft[1]}
  \\
\stesysf_2
  \ar{r}[swap]{\beta_1}
  &
\stesysf
\end{tikzcd}
\qquad
\text{and}
\qquad
\begin{tikzcd}
\stesyst_4
  \ar{r}{\beta^t_2}
  \ar{d}[swap]{\ebd[3]}
  &
\stesyst_3
  \ar{d}{\ebd[2]}
  \\
\stesysf_4
  \ar{r}{\beta_3}
  \ar{d}[swap]{\eft[3]}
  &
\stesysf_3
  \ar{d}{\eft[2]}
  \\
\stesysf_3
  \ar{r}[swap]{\beta_2}
  &
\stesysf_2
\end{tikzcd}
\end{equation*}
assemble pre-extension homomorphisms 
\(
\boldsymbol{\beta}
  :
\famesys{\famesys{\stesys}}
  \to
\famesys{\stesys}
\) 
and
\(
\boldsymbol{\beta}_\mathbf{2}
  :
\famesys{\famesys{\famesys{\stesys}}}
  \to
\famesys{\famesys{\stesys}}
\).
\end{defn}

\begin{defn}\label{famehom}
Suppose that $f:\stesys'\to\stesys$ is a pre-extension homomorphism. Then we
define $\famehom{f}:\famesys{\stesys'}\to\famesys{\stesys}$ to consist of
\begin{equation*}
\begin{tikzcd}
\stesyst_2'
  \ar{r}{f^t_2}
  \ar{d}[swap]{\ebd[1]'}
  &
\stesyst_2
  \ar{d}{\ebd[1]}
  \\
\stesysf_2'
  \ar{r}{f_2}
  \ar{d}[swap]{\eft[1]'}
  &
\stesysf_2
  \ar{d}{\eft[1]}
  \\
\stesysf'
  \ar{r}[swap]{f_1}
  &
\stesysf
\end{tikzcd}
\end{equation*}
where we define
\begin{equation*}
f^t_2 := \pullback{f_1}{f^t}{\ectxext}{\eft\circ\ebd}.
\end{equation*}
\end{defn}

\begin{lem}
The triple $\famehom{f}$ defined in \autoref{famehom} is a pre-extension homomorphism.
\end{lem}

\begin{proof}
Note that the square
\begin{equation*}
\begin{tikzcd}
\stesysf_2'
  \ar{r}{f_2}
  \ar{d}[swap]{\efamext'}
  &
\stesysf_2
  \ar{d}{\efamext}
  \\
\stesysf'
  \ar{r}[swap]{f_1}
  &
\stesysf
\end{tikzcd}
\end{equation*}
commutes by assumption. Thus, it remains to show that the square
\begin{equation*}
\begin{tikzcd}
\stesysf_3'
  \ar{r}{f_3}
  \ar{d}[swap]{\eext{2}'}
  &
\stesysf_3
  \ar{d}{\eext{2}}
  \\
\stesysf_2'
  \ar{r}[swap]{f_2}
  &
\stesysf_2
\end{tikzcd}
\end{equation*}
commutes. It is equivalent to show that the equalities
\begin{align*}
\pullbackpr{1}{\ectxext}{\eft}\circ f_2\circ\eext{2}'
  & =
\pullbackpr{1}{\ectxext}{\eft}\circ \eext{2}\circ f_3
  \\
\pullbackpr{2}{\ectxext}{\eft}\circ f_2\circ\eext{2}'
  & =
\pullbackpr{2}{\ectxext}{\eft}\circ \eext{2}\circ f_3
\end{align*}
both hold. For the first, it is straightforward to verify that the diagram
\begin{equation*}
\begin{tikzcd}[column sep=large]
{} &
\stesysf_2'
  \ar{r}{f_2}
  \ar{dr}[near end]{\pullbackpr{1}{\ectxext'}{\eft'}}
  &
\stesysf_2
  \ar{dr}{\pullbackpr{1}{\ectxext}{\eft}}
  \\
\stesysf_3'
  \ar{ur}{\eext{2}'}
  \ar{r}[swap]{\beta_2'}
  \ar{ddr}[swap]{f_3}
  &
\stesysf_2'
  \ar{r}{\efamext'}
  \ar{dr}[swap]{f_2}
  &
\stesysf'
  \ar{r}{f_1}
  &
\stesysf
  \\
{} & {} &
\stesysf_2
  \ar{ur}[near start]{\efamext}
  \\
{} &
\stesysf_3
  \ar{r}[swap]{\eext{2}}
  \ar{ur}{\beta_2}
  &
\stesysf_2
  \ar{uur}[swap]{\pullbackpr{1}{\ectxext}{\eft}}
\end{tikzcd}
\end{equation*}
commutes. For the second, note that the diagram
\begin{equation*}
\begin{tikzcd}[column sep=large]
{} &
\stesysf_2'
  \ar{r}{f_2}
  \ar{dr}{\beta_1'}
  &
\stesysf_2
  \ar{ddr}{\pullbackpr{2}{\ectxext}{\eft}}
  \\
{} & {} &
\stesysf'
  \ar{dr}[swap,near start]{f_1}
  \\
\stesysf_3'
  \ar{uur}{\eext{2}'}
  \ar{r}{\beta_2'}
  \ar{dr}[swap]{f_3}
  &
\stesysf_2'
  \ar{r}{f_2}
  \ar{ur}{\efamext'}
  &
\stesysf_2
  \ar{r}[swap]{\efamext}
  &
\stesysf
  \\
{} &
\stesysf_3
  \ar{r}[swap]{\eext{2}}
  \ar{ur}[near start]{\beta_2}
  &
\stesysf_2
  \ar{ur}[swap]{\pullbackpr{2}{\ectxext}{\eft}}
\end{tikzcd}
\end{equation*}
commutes.
\end{proof}

\begin{lem}[Stability under retracts]\label{esys-retract}
Suppose $f:\stesys\to\stesys'$ is a pre-extension homomorphism between
pre-extension objects. If there is a pre-extension homomorphism $g:\stesys'\to
\stesys$ such that $g\circ f=\catid{\stesys}$ and $\stesys'$ is an extension
algebra, then $\stesys$ is an extension algebra.
\end{lem}

Before we start with the proof, note that we have the equalities
$g_2\circ f_2=\catid{\stesysf_2}$ and $g_3\circ f_3=\catid{\stesysf_3}$
under the hypotheses of the lemma.

\begin{proof}
Our first subgoal is to show that the square
\begin{equation*}
\begin{tikzcd}
\stesysf_2 
  \ar{r}{\efamext} 
  \ar{d}[swap]{\pullbackpr{2}{\ectxext}{\eft}} 
  & 
\stesysf 
  \ar{d}{\ectxext}
  \\
\stesysf
  \ar{r}[swap]{\ectxext} 
  & 
\stesysc
\end{tikzcd}
\end{equation*}
commutes. Note that in the diagram
\begin{equation*}
\begin{tikzcd}
  {}
  & 
\stesysf
  \ar{dd}[near start]{\ectxext}
  \ar{rr}{f_1}
  &
  &
\stesysf'
  \ar{dd}[near start]{\ectxext'}
  \ar{rr}{g_1}
  &
  &
\stesysf
  \ar{dd}{\ectxext}
  \\
\stesysf_2
  \ar{dd}[swap]{\pullbackpr{2}{\ectxext}{\eft}}
  \ar[crossing over]{rr}[swap,near start]{f_2}
  \ar{ur}{\efamext}
  &
  &
\stesysf_2'
  \ar{ur}[near start]{\efamext'}
  \ar[crossing over]{rr}[swap,near start]{g_2}
  &
  &
\stesysf_2
  \ar{ur}[swap,near start]{\efamext}
  \\
  {}
  &
\stesysc
  \ar{rr}[near start]{f_0}
  &
  &
\stesysc'
  \ar{rr}[near start]{g_0}
  &
  &
\stesysc
  \\
\stesysf 
  \ar{rr}[swap]{f_1}
  \ar{ur}{\ectxext}
  &
  &
\stesysf' 
  \ar[crossing over,leftarrow]{uu}[near end,swap]{\pullbackpr{2}{\ectxext'}{\eft'}}
  \ar{ur}[swap,near end]{\ectxext'}
  \ar{rr}[swap]{g_1}
  &
  &
\stesysf
  \ar[crossing over,leftarrow]{uu}[near end,swap]{\pullbackpr{2}{\ectxext}{\eft}}
  \ar{ur}[swap]{\ectxext}
\end{tikzcd}
\end{equation*}
all the faces minus the far left and far right face commute. Using that $g$
is a section of $f$, we can read off that also the far left face commutes,
completing our first subgoal.

For the second subgoal, note that also $\famehom{g}\circ\famehom{f}=
\catid{\famesys{\stesys}}$ and that $\famesys{\stesys'}$ is an extension object.
Thus we can apply what we have proven so far to conclude that the square
\begin{equation*}
\begin{tikzcd}
\stesysf_3 
  \ar{r}{\eext{2}} 
  \ar{d}[swap]{\pullbackpr{2}{\efamext}{\eft[1]}} 
  & 
\stesysf_2 
  \ar{d}{\efamext}
  \\
\stesysf_2
  \ar{r}[swap]{\efamext} 
  & 
\stesysf
\end{tikzcd}
\end{equation*}
commutes.
\end{proof}

\subsection{The change of base of (pre-)extension objects}
\label{subsection:change_of_base}
An important construction of (pre-)extension objects is the change of base. It
allows us to consider `parametrized homomorphisms', such as weakening and
substitution.

\begin{defn}
Suppose $f:\stesys\to\stesys'$ is a pre-extension homomorphism. We say that
a diagram
\begin{equation*}
\begin{tikzcd}
\stesys
  \ar{r}{f}
  \ar{d}[swap]{p}
  &
\stesys'
  \ar{d}{p'}
  \\
X \ar{r}[swap]{g}
  &
Y
\end{tikzcd}
\end{equation*}
commutes if the diagram
\begin{equation*}
\begin{tikzcd}
\stesysc
  \ar{r}{f_0}
  \ar{d}[swap]{p}
  &
\stesysc'
  \ar{d}{p'}
  \\
X \ar{r}[swap]{g}
  &
Y
\end{tikzcd}
\end{equation*}
commutes.
\end{defn}

The first goal in this subsection is to define for every (pre-)extension object 
$\stesys$ and every $p:\stesysc\rightarrow X\leftarrow Y:g$, a (pre-)extension
object $\cobesys{Y}{\stesys}{g}{p}$ with a homomorphism $\pullbackpr{2}{g}{p}:
\cobesys{Y}{\stesys}{g}{p}\to\stesys$ and a morphism $\pullbackpr{1}{g}{p}:
\pullback{Y}{\stesysc}{g}{p}\to Y$ such that for every diagram
\begin{equation*}
\begin{tikzcd}[column sep=large]
\stesys'
  \ar[bend right=10]{ddr}[swap]{p'}
  \ar[bend left=10]{rrd}{f}
  \ar[dotted]{dr}[near end]{[p',f]}
  \\
  {}&
\cobesys{Y}{\stesys}{g}{p}
  \ar{d}{\pullbackpr{1}{g}{p}}
  \ar{r}[swap]{\pullbackpr{2}{g}{p}}
  &
\stesys
  \ar{d}{p}
  \\
  {}&
Y \ar{r}[swap]{g}
  &
X
\end{tikzcd}
\end{equation*}
of which the outer square commutes, the (pre-)extension homomorphism $[p',f]$ exists
and is unique with the property that it renders the diagram commutative. We will
give the definition of $\cobesys{Y}{\stesys}{g}{p}$ in \autoref{cobesys}. After
proving that the change of base of a pre-extension algebra is indeed a
pre-extension algebra (\autoref{cobesys-preext}) and that the change of base
of an extension algebra is an extension algebra (\autoref{cobesys-ext}), we
will demonstrate the above unique existence in \autoref{cobesys-existence,%
cobesys-pullback}.

The second goal in this subsection is to follow the same procedure for
$\famesys{\famesys{\stesys}}$ to show that it is equivalent to
$\cobesys{\stesysf}{\famesys{\stesys}}{\ectxext}{\eft}$. We will do this by
verifying directly that it has the universal property of the change of base
described above, because we will use the ingredients in our definition of
weakening and substitution objects.

\begin{defn}[Change of base]\label{cobesys}
Suppose $\stesys$ is a pre-extension object in $\cat{C}$ and that 
$p:\stesysc\rightarrow X\leftarrow Y:g$.
Then we define the pre-extension object $\cobesys{Y}{\stesys}{g}{p}$ to consist of
\begin{equation*}
\begin{tikzcd}
\cobesys{Y}{\stesyst}{g}{p\circ\eft\circ\ebd}
  \ar{r}
  \ar{d}[swap]{g^\ast(\ebd)}
  &
\stesyst
  \ar{d}{\ebd}
  \\
\cobesys{Y}{\stesysf}{g}{p\circ\eft}
  \ar{r}
  \ar{d}[swap]{g^\ast(\eft)}
  &
\stesysf
  \ar{d}{\eft}
  \\
\cobesys{Y}{\stesysc}{g}{p}
  \ar{r}
  \ar{d}[swap]{\pullbackpr{1}{g}{p}}
  &
\stesysc
  \ar{d}{p}
  \\
Y \ar{r}[swap]{g}
  &
X
\end{tikzcd}
\end{equation*} 
and the operations
\begin{align*}
\cobesys{Y}{\ectxext}{g}{p} 
  & : \pullback{Y}{\stesysf}{g}{p\circ\eft}\to \pullback{Y}{\stesysc}{g}{p}\\
\cobesys{Y}{\efamext}{g}{p} 
  & : \pullback
    {\pullback{Y}{\stesysf}{g}{p\circ\eft}}
    {\pullback{Y}{\stesysf}{g}{p\circ\eft}}
    {\cobesys{Y}{\ectxext}{g}{p}}
    {g^\ast(\eft)}
  \to 
  \pullback{Y}{\stesysf}{g}{p\circ\eft}.
\end{align*}
defined by
\begin{equation*}
\cobesys{Y}{\ectxext}{g}{p} := \pullback{\catid{Y}}{\ectxext}{g}{p}
\end{equation*}
and where $\cobesys{Y}{\efamext}{g}{p}$ is defined by rendering the diagram
\begin{equation*}
\begin{tikzcd}[column sep=large]
(\cobesys{Y}{\stesysf}{g}{p\circ\eft})_2
  \ar{rr}{\pullback{\pullbackpr{2}{g}{p\circ\eft}}{\pullbackpr{2}{g}{p\circ\eft}}{\ectxext}{\eft}}
  \ar{dd}[swap]{\pullbackpr{1}{\cobesys{Y}{\ectxext}{g}{p}}{g^\ast(\eft)}}
  \ar[dotted]{dr}[swap]{\cobesys{Y}{\efamext}{g}{p}}
  &
  &
\stesysf_2
  \ar{d}{\efamext}
  \\
  {}&
\cobesys{Y}{\stesysf}{g}{p\circ\eft}
  \ar{r}{\pullbackpr{2}{g}{p\circ\eft}}
  \ar{d}[swap]{\pullbackpr{1}{g}{p\circ\eft}}
  &
\stesysf
  \ar{d}{p\circ\eft}
  \\
\cobesys{Y}{\stesysf}{g}{p\circ\eft}
  \ar{r}[swap]{\pullbackpr{1}{g}{p\circ\eft}}
  &
Y \ar{r}[swap]{g}
  &
X
\end{tikzcd}
\end{equation*} 
commutative. 
The process of obtaining the pre-extension object $\cobesys{Y}{\stesys}{g}{p}$ out of $\stesys$
and $g:Y\to X$ is also called the \emph{change of base}.
\end{defn}

\begin{lem}\label{cobesys-preext}
Any change of base of a pre-extension object is a pre-extension object.
\end{lem}

\begin{proof}
Let $\stesys$ be an extension algebra and consider $p:\stesysc\rightarrow X\leftarrow Y:g$.
We need to verify that the square
\begin{equation*}
\begin{tikzcd}[column sep=large]
(\pullback{Y}{\stesysf}{g}{p\circ\eft})_2
  \ar{r}{\cobesys{Y}{\efamext}{g}{p}} 
  \ar{d}[swap]{\pullbackpr{1}{\cobesys{Y}{\ectxext}{g}{p}}{g^\ast(\eft)}} 
  & 
\pullback{Y}{\stesysf}{g}{p\circ\eft}
  \ar{d}{g^\ast(\eft)}
  \\
\pullback{Y}{\stesysf}{g}{p\circ\eft}
  \ar{r}[swap]{g^\ast(\eft)} 
  & 
\pullback{Y}{\stesysc}{g}{p}
\end{tikzcd}
\end{equation*}
commutes. It is fairly obvious that
\begin{equation*}
\pullbackpr{1}{g}{p}\circ g^\ast(\eft)\circ (\cobesys{Y}{\efamext}{g}{p})
  =
\pullbackpr{1}{g}{p\circ\eft}\circ \pullbackpr{1}{\cobesys{Y}{\ectxext}{g}{p}}{g^\ast(\eft)}
\end{equation*}
and that the diagram
\begin{equation*}
\begin{tikzcd}
  {}&
  {}&
\pullback{Y}{\stesysf}{g}{p\circ\eft}
  \ar{rr}{g^\ast(\eft)}
  \ar{dr}[swap]{\pullbackpr{2}{g}{p\circ\eft}}
  &
  {}&
\pullback{Y}{\stesysc}{g}{p}
  \ar{ddr}{\pullbackpr{2}{g}{p}}
  \\
  {}&
  {}&
  {}&
\stesysf
  \ar{drr}[swap]{\eft}
  \\
(\pullback{y}{\stesysf}{g}{p\circ\eft})_2
  \ar{uurr}{\cobesys{Y}{\efamext}{g}{p}}
  \ar{rr}[swap,yshift=-.5ex]{\pullback{\pullbackpr{2}{g}{p\circ\eft}}{\pullbackpr{2}{g}{p\circ\eft}}{\ectxext}{\eft}}
  \ar{ddrr}[swap]{\pullbackpr{1}{\cobesys{Y}{\ectxext}{g}{p}}{g^\ast(\eft)}}
  &
  {}&
\stesysf_2
  \ar{ur}{\efamext}
  \ar{dr}[swap]{\eft[1]}
  &
  {}&
  {}&
\stesysc
  \\
  {}&
  {}&
  {}&
\stesysf
  \ar{urr}{\eft}
  \\
  {}&
  {}&
\pullback{Y}{\stesysf}{g}{p\circ\eft}
  \ar{rr}[swap]{g^\ast(\eft)}
  \ar{ur}{\pullbackpr{2}{g}{p\circ\eft}}
  &
  {}&
\pullback{Y}{\stesysc}{g}{p}
  \ar{uur}[swap]{\pullbackpr{2}{g}{p}}
\end{tikzcd}
\end{equation*}
commutes.
\end{proof}

\begin{thm}\label{cobesys-ext}
The change of base of an extension algebra is an extension algebra.
\end{thm}

\begin{proof}
Our first subgoal is to verify that the square
\begin{equation*}
\begin{tikzcd}[column sep=large]
(\pullback{Y}{\stesysf}{g}{p\circ\eft})_2
  \ar{r}{\cobesys{Y}{\efamext}{g}{p}} 
  \ar{d}[swap]{\pullbackpr{2}{\cobesys{Y}{\ectxext}{g}{p}}{g^\ast(\eft)}} 
  & 
\pullback{Y}{\stesysf}{g}{p\circ\eft}
  \ar{d}{\cobesys{Y}{\ectxext}{g}{p}}
  \\
\pullback{Y}{\stesysf}{g}{p\circ\eft}
  \ar{r}[swap]{\cobesys{Y}{\ectxext}{g}{p}} 
  & 
\pullback{Y}{\stesysc}{g}{p}
\end{tikzcd}
\end{equation*}
\end{proof}

The following construction is useful for defining extension homomorphisms into
`higher' extension objects

\begin{defn}\label{cobesys-existence}
Consider a commutative diagram
\begin{equation*}
\begin{tikzcd}
\stesys'
  \ar{r}{f}
  \ar{d}[swap]{p'}
  &
\stesys
  \ar{d}{p}
  \\
Y \ar{r}[swap]{g}
  &
X
\end{tikzcd}
\end{equation*}
Then we construct $[p,f]:\stesys'\to\cobesys{Y}{\stesys}{g}{p}$
\begin{itemize}
\item by defining $[p,f]_0:\stesysc'\to\pullback{Y}{\stesysc}{g}{p}$ be the uniqe
morphism rendering the diagram
\begin{equation*}
\begin{tikzcd}[column sep=large]
\stesysc'
  \ar[bend right=10]{ddr}[swap]{p'}
  \ar[bend left=10]{rrd}{f_0}
  \ar{dr}[near end]{[p,f]_0}
  \\
  {}&
\pullback{Y}{\stesysc}{g}{p}
  \ar{r}[swap]{\pullbackpr{2}{g}{p}}
  \ar{d}{\pullbackpr{1}{g}{p}}
  &
\stesysc
  \ar{d}{p}
  \\
  {}&
Y \ar{r}[swap]{g}
  &
X
\end{tikzcd}
\end{equation*}
commutative.
\item by defining $[p,f]_1:\stesysf'\to\pullback{Y}{\stesysf}{g}{p\circ\eft}$ be the uniqe
morphism rendering the diagram
\begin{equation*}
\begin{tikzcd}[column sep=huge]
\stesysf'
  \ar[bend right=10]{ddr}[swap]{p'\circ\eft'}
  \ar[bend left=10]{rrd}{f_1}
  \ar{dr}[near end]{[p,f]_1}
  \\
  {}&
\pullback{Y}{\stesysf}{g}{p\circ\eft}
  \ar{r}[swap]{\pullbackpr{2}{g}{p\circ\eft}}
  \ar{d}{\pullbackpr{1}{g}{p\circ\eft}}
  &
\stesysc
  \ar{d}{p\circ\eft}
  \\
  {}&
Y \ar{r}[swap]{g}
  &
X
\end{tikzcd}
\end{equation*}
commutative.
\item by defining $[p,f]^t:\stesyst'\to\pullback{Y}{\stesyst}{g}{p\circ\eft\circ\ebd}$ be the uniqe
morphism rendering the diagram
\begin{equation*}
\begin{tikzcd}[column sep=huge]
\stesyst'
  \ar[bend right=10]{ddr}[swap]{p'\circ\eft'\circ\ebd'}
  \ar[bend left=10]{rrd}{f^t}
  \ar{dr}[near end]{[p,f]^t}
  \\
  {}&
\pullback{Y}{\stesysf}{g}{p\circ\eft\circ\ebd}
  \ar{r}[swap]{\pullbackpr{2}{g}{p\circ\eft\circ\ebd}}
  \ar{d}{\pullbackpr{1}{g}{p\circ\eft\circ\ebd}}
  &
\stesysc
  \ar{d}{p\circ\eft\circ\ebd}
  \\
  {}&
Y \ar{r}[swap]{g}
  &
X
\end{tikzcd}
\end{equation*}
commutative.
\end{itemize}
\end{defn}

\begin{thm}\label{cobesys-pullback}
For every diagram
\begin{equation*}
\begin{tikzcd}[column sep=large]
\stesys'
  \ar[bend right=10]{ddr}[swap]{p'}
  \ar[bend left=10]{rrd}{f}
  \ar[dotted]{dr}[near end]{[p',f]}
  \\
  {}&
\cobesys{Y}{\stesys}{g}{p}
  \ar{d}{\pullbackpr{1}{g}{p}}
  \ar{r}[swap]{\pullbackpr{2}{g}{p}}
  &
\stesys
  \ar{d}{p}
  \\
  {}&
Y \ar{r}[swap]{g}
  &
X
\end{tikzcd}
\end{equation*}
of which the outer square commutes, the pre-extension homomorphism $[p',f]$
is unique with the property that it renders the whole diagram commutative.
\end{thm}

\begin{defn}\label{famfamstesys_into}
Consider a commutative square
\begin{equation*}
\begin{tikzcd}
\stesys'
  \ar{r}{f}
  \ar{d}[swap]{p}
  &
\famesys\stesys
  \ar{d}{\eft}
  \\
\stesysf \ar{r}[swap]{\ectxext}
  &
\stesysc
\end{tikzcd}
\end{equation*}
Then we construct
\begin{equation*}
|[p,f]|:\stesys'\to\famesys{\famesys{\stesys}}
\end{equation*}
as follows:
\begin{itemize}
\item let $|[p,f]|_0:\stesysc'\to\stesysf_2$ be the unique morphism rendering
the diagram
\begin{equation*}
\begin{tikzcd}[column sep=large]
\stesysc' 
  \ar[bend left=10]{rrd}{f_0}
  \ar[swap,bend right=10]{ddr}{p}
  \ar[dotted]{dr}[near end]{|[p,f]|_0}
  \\
  {}&
\stesysf_2
  \ar{r}[swap]{\pullbackpr{2}{\ectxext}{\eft}}
  \ar{d}{\eft[1]}
  &
\stesysf
  \ar{d}{\eft}
  \\
  {}&
\stesysf
  \ar{r}[swap]{\ectxext}
  &
\stesysc
\end{tikzcd}
\end{equation*}
commutative.
\item Let $|[p,f]|_1:\stesysf'\to\stesysf_3$ be the unique morphism rendering
the diagram
\begin{equation*}
\begin{tikzcd}[column sep=large]
\stesysf'
  \ar[bend left=10]{drr}{f_1}
  \ar[swap]{dd}{\eft'}
  \ar[dotted]{dr}[near end]{|[p,f]|_1}
  \\
  {}&
\stesysf_3
  \ar{r}[swap]{\pullbackpr{2}{\efamext}{\eft[1]}}
  \ar{d}{\eft[1]}
  &
\stesysf_2
  \ar{d}{\eft[1]}
  \\
\stesysc'
  \ar{r}[swap]{|[p,f]|_0}
  &
\stesysf_2
  \ar{r}[swap]{\efamext}
  &
\stesysf
\end{tikzcd}
\end{equation*}
commutative.
\item Let $|[p,f]|^t:\stesyst'\to\stesyst_3$ be the unique morphism rendering
the diagram
\begin{equation*}
\begin{tikzcd}[column sep=huge]
\stesyst'
  \ar[bend left=10]{drr}{f^t}
  \ar[swap]{dd}{\eft'\circ\ebd'}
  \ar[dotted]{dr}[near end]{|[p,f]|^t}
  \\
  {}&
\stesyst_3
  \ar{r}[swap]{\pullbackpr{2}{\efamext}{\eft[1]\circ\ebd[1]}}
  \ar{d}[swap]{\pullbackpr{1}{\efamext}{\eft[1]\circ\ebd[1]}}
  &
\stesyst_2
  \ar{d}{\eft[1]\circ\ebd[1]}
  \\
\stesysc'
  \ar{r}[swap]{|[p,f]|_0}
  &
\stesysf_2
  \ar{r}[swap]{\efamext}
  &
\stesysf
\end{tikzcd}
\end{equation*}
commutative.
\end{itemize}
\end{defn}

\begin{lem}
Under the hypotheses of \autoref{famfamstesys_into}, $|[p,f]|$ is a pre-extension
homomorphism. Moreover, it is the unique pre-extension homomorphism for which
the diagram
\begin{equation*}
\begin{tikzcd}[column sep=large]
\stesys' 
  \ar[bend left=10]{rrd}{f}
  \ar[swap,bend right=10]{ddr}{p}
  \ar[dotted]{dr}[near end]{|[p,f]|}
  \\
  {}&
\famesys{\famesys{\stesys}}
  \ar{r}[swap]{\pullbackpr{2}{\ectxext}{\eft}}
  \ar{d}{\eft[1]}
  &
\famesys{\stesys}
  \ar{d}{\eft}
  \\
  {}&
\stesysf
  \ar{r}[swap]{\ectxext}
  &
\stesysc
\end{tikzcd}
\end{equation*}
commutes.
\end{lem}

\begin{lem}
Suppose $f:\stesys\to \stesys'$ is a pre-extension homomorphism and consider a morphism
$p:\stesys'\to X$ and $g:Y\to X$. Then the change of base 
$g^\ast(f):\cobesys{Y}{\stesys}{g}{p\circ f_0}\to
\cobesys{Y}{\stesys'}{g}{p}$ is a pre-extension morphism.
\end{lem}

\begin{lem}
Let $\stesys$ be a pre-extension algebra and consider $p:\stesysc\rightarrow X\leftarrow Y:g$.
Then there is an isomorphism
\begin{equation*}
\varphi:\famesys{\cobesys{Y}{\stesys}{g}{p}}
  \simeq
\cobesys{Y}{\famesys{\stesys}}{g}{p\circ\eft}
\end{equation*}
uniquely determined by
\end{lem}

\begin{proof}
This follows from the pasting lemma for pullbacks.
\end{proof}

\subsection{Pre-weakening objects}
\begin{defn}
Let $\stesys$ be an extension object in $\cat{C}$. A pre-weakening operation
on $\stesys$ is an extension homomorphism 
$ \mathbf{w}(\stesys)
    :
  \cobesys{\stesysf}{\famesys{\stesys}}{\eft}{\eft}
    \to
  \famesys{\famesys{\stesys}}$
for which the diagram
\begin{equation*}
\begin{tikzcd}[column sep=large]
\cobesys{\stesysf}{\famesys{\stesys}}{\eft}{\eft}
  \ar{r}{\mathbf{w}(\stesys)}
  \ar{dr}[swap]{\pullbackpr{1}{\eft}{\eft}}
  &
\famesys{\famesys{\stesys}}
  \ar{d}{\eft[1]}
  \\
& \stesysf
\end{tikzcd}
\end{equation*}
commutes.
\end{defn}

\begin{defn}
Let $\stesys$ be an extension object with pre-weakening operation
$\mathbf{w}(\stesys)$. Then $\famesys{\stesys}$ has the pre-weakening operation
$\mathbf{w}(\famesys{\stesys})$ which is uniquely determined by rendering the
diagram
\begin{equation*}
\begin{tikzcd}[column sep=large]
\cobesys{\stesysf_2}{\famesys{\famesys{\stesys}}}{\eft[1]}{\eft[1]}
  \ar{rr}{%
      \pullback{\beta_1}{\boldsymbol{\beta}}{\eft}{\eft}
    }
  \ar[bend right]{ddr}[swap]{\pullbackpr{1}{\eft[1]}{\eft[1]}}
  \ar[dotted]{dr}{\mathbf{w}(\famesys{\stesys})}
  &
  {}&
\cobesys{\stesysf}{\famesys{\stesys}}{\eft}{\eft}
  \ar{r}{\mathbf{w}(\stesys)}
  &
\famesys{\famesys{\stesys}}
  \ar{d}{\boldsymbol{\beta}}
  \\
  {}&
\famesys{\famesys{\famesys{\stesys}}}
  \ar{r}{\boldsymbol{\beta}_\mathbf{2}}
  \ar{d}[swap]{\eft[2]}
  &
\famesys{\famesys{\stesys}}
  \ar{d}{\eft[1]}
  \ar{r}{\boldsymbol{\beta}}
  &
\famesys{\stesys}
  \ar{d}{\eft}
  \\
  {}&
\stesysf_2
  \ar{r}[swap]{\efamext}
  &
\stesysf
  \ar{r}[swap]{\ectxext}
  &
\stesysc
\end{tikzcd}
\end{equation*}
commutative.
\end{defn}

\begin{defn}
A pre-weakening object $\stesys$ in $\cat{C}$ is an extension object $\stesys$ 
with a pre-weakening operation 
$ \mathbf{w}(\stesys)
    :
  \cobesys{\stesysf}{\famesys{\stesys}}{\eft}{\eft}
    \to
  \famesys{\famesys{\stesys}}$
for which the diagram
\begin{equation*}
\begin{tikzcd}[column sep=15em]
\cobesys{\stesysf_2}{\famesys{\stesys}}{\eft\circ\eft[1]}{\eft}
  \ar[bend right=10]{dr}[swap]%
    { [ \pullbackpr{1}{\eft\circ\eft[1]}{\eft},%
        \mathbf{w}(\stesys)%
          \circ%
        (\pullback{\efamext}{\catid{\famesys{\stesys}}}{\eft}{\eft})%
        ]%
      }
  \ar{r}{
    [ \pullbackpr{1}{\eft\circ\eft[1]}{\eft},%
      \mathbf{w}(\stesys)%
        \circ%
      (\pullback{\eft[1]}{\catid{\famesys{\stesys}}}{\eft}{\eft})%
      ]}%
  &
\cobesys{\stesysf_2}{\famesys{\famesys{\stesys}}}{\eft[1]}{\eft[1]}
  \ar{d}{\mathbf{w}(\famesys{\stesys})}
  \\
  {}&
\famesys{\famesys{\famesys{\stesys}}}
\end{tikzcd}
\end{equation*}
commutes. This condition is called \emph{Currying for weakening}.
\end{defn}

\begin{lem}
If $\stesys$ is a pre-weakening algebra, then so is $\famesys{\stesys}$. 
\end{lem}

\begin{proof}
\end{proof}

\begin{defn}
A pre-weakening morphism between preweakening objects $\stesys$ and $\stesys'$ is an
extension homomorphism $f:\stesys\to \stesys'$ such that additionally the diagram
\begin{equation*}
\begin{tikzcd}[column sep=large]
\cobesys{\stesysf}{\famesys{\stesys}}{\eft}{\eft}
  \ar{d}[swap]{\mathbf{w}(\stesys)}
  \ar{r}{\pullback{f_1}{\famehom{f}}{\eft'}{\eft'}}
  &
\cobesys{\stesysf'}{\famesys{\stesys'}}{\eft'}{\eft'}
  \ar{d}{\mathbf{w}(\stesys')}
  \\
\famesys{\famesys{\stesys}}
  \ar{r}[swap]{\famehom{\famehom{f}}}
  &
\famesys{\famesys{\stesys'}}
\end{tikzcd}
\end{equation*}
commutes.
\end{defn}

\begin{defn}
Let $\stesys$ be a pre-weakening algebra and consider $p:\stesysc\rightarrow X\leftarrow Y:p$.
Then we define
\begin{equation*}
\mathbf{w}(\cobesys{Y}{\stesys}{g}{p})
  :
\cobesys
  { (\pullback{Y}{\stesysf}{g}{p\circ\eft})}
  { \famesys{\cobesys{Y}{\stesys}{g}{p}}}
  { g^\ast(\eft)}
  { g^\ast(\eft)}
  \to
\famesys{\famesys{\cobesys{Y}{\stesys}{g}{p}}}
\end{equation*}
to be the unique extension homomorphism rendering the diagram
\begin{equation*}
\begin{tikzcd}
\cobesys
  { \pullback{Y}{\stesysf}{g}{p\circ\eft}}
  { \famesys{\cobesys{Y}{\stesys}{g}{p}}}
  { g^\ast(\eft)}
  { g^\ast(\eft)}
  \ar[bend right]{ddr}[swap]{\pullbackpr{1}{g^\ast(\eft)}{g^\ast(\eft)}}
  \ar{rr}{%
    \pullback
      { \pullbackpr{2}{g}{p\circ\eft}}
      { \boldsymbol{\pi}_\mathbf{2}(g,p\circ\eft)}
      { \eft}
      { \eft}
    }
  \ar[dotted]{dr}{\mathbf{w}(\cobesys{Y}{\stesys}{g}{p})}
  &
  {}&
\cobesys{\stesysf}{\famesys{\stesys}}{\eft}{\eft}
  \ar{r}{\mathbf{w}(\stesys)}
  &
\famesys{\famesys{\stesys}}
  \ar{d}{\boldsymbol{\beta}}
  \\
  {}&
\famesys{\famesys{\cobesys{Y}{\stesys}{g}{p}}}
  \ar{r}{\beta}
  \ar{d}{g^\ast(\eft)_1}
  &
\famesys{\cobesys{Y}{\stesys}{g}{p}}
  \ar{r}{\boldsymbol{\pi}_2(g,p\circ\eft)}
  \ar{d}{g^\ast(\eft)}
  &
\famesys{\stesys}
  \ar{d}{\eft}
  \\
  {}&
\pullback{Y}{\stesysf}{g}{p\circ\eft}
  \ar{r}[swap]{\cobesys{Y}{\ectxext}{g}{p}}
  &
\pullback{Y}{\stesysc}{g}{p}
  \ar{r}[swap]{\pullbackpr{2}{g}{p}}
  &
\stesysc
\end{tikzcd}
\end{equation*}
commutative.
\end{defn}

\subsection{Weakening objects}
Since we have shown that the property of being a pre-weakening object is closed
under the relevant operations, we can make the following definition:

\begin{defn}
A weakening object is a pre-weakening object $\stesys$ with the property that
$\mathbf{w}(\stesys)$ is a pre-weakening morphism.
\end{defn}

\begin{defn}
A weakening homomorphism is a pre-weakening homomorphism such that the domain
and codomain are weakening objects.
\end{defn}

\begin{thm}
Suppose $\stesys$ is a weakening object, then so is $\famesys{\stesys}$
\end{thm}

\begin{thm}
The change of base of any weakening object is again a weakening object.
\end{thm}

\subsection{Projection objects}
\begin{defn}
A pre-projection object is a weakening object $\stesys$ for which there is a term
$\mathbf{i}:\stesysf\to \stesyst_2$ such that the diagram
\begin{equation*}
\begin{tikzcd}[column sep=large]
\stesysf \ar{r}{\mathbf{i}} \ar{d}[swap]{\Delta_{\eft}} & \stesyst_2 \ar{d}{\ebd[1]}\\
\pullback{\stesysf}{\stesysf}{\eft}{\eft} \ar{r}[swap]{w(\stesys)_0} & \stesysf_2
\end{tikzcd}
\end{equation*}
commutes. In this diagram $\Delta_{\eft}:\stesysf\to \pullback{\stesysf}{\stesysf}{\eft}{\eft}$ is the diagonal.
\end{defn}

\begin{defn}
A pre-projection homomorphism from $\stesys$ to $\stesys'$ is a weakening homomorphism
$f:\stesys\to \stesys'$ such that the square
\begin{equation*}
\begin{tikzcd}[column sep=large]
\stesyst_2
  \ar{r}{f^t_1}
  &
\stesyst_2'
  \\
\stesysf \ar{r}[swap]{f_1}
  \ar{u}{\mathbf{i}}
  &
\stesysf'
  \ar{u}[swap]{\mathbf{i}'}
\end{tikzcd}
\end{equation*}
commutes
\end{defn}

\begin{lem}
The change of base of a pre-projection object is again a pre-projection object.
\end{lem}

\begin{lem}
If $CFT$ is a pre-projection object, then so is $\mathbf{F}_{CFT}$, where
$\mathbf{F}_{\mathbf{i}}$ is defined to be $F\times_{e_0,c}\mathbf{i}$ is
a pre-projection algebra.
\end{lem}

\begin{defn}
A projection algebra is a pre-projection algebra for which weakening is a
pre-projection homomorphism.
\end{defn}

\begin{cor}
The change of base of a projection object is again a projection object.
\end{cor}

\begin{cor}
If $CFT$ is a projection object, then so is $\mathbf{F}_{CFT}$, where
$\mathbf{F}_{\mathbf{i}}$ is defined to be $F\times_{e_0,c}\mathbf{i}$ is
a projection algebra.
\end{cor}

\subsection{Substitution objects}

\begin{defn}
A \emph{pre-substitution} for an extension object $\stesys$ is an
extension homomorphism
\begin{equation*}
\mathbf{s}(\stesys):\cobesys{\stesyst}{\famesys{\famesys{\stesys}}}{\ebd}{\eft[1]}\to \famesys{\stesys}
\end{equation*}
for which the square
\begin{equation*}
\begin{tikzcd}[column sep=large]
\pullback{\stesyst}{\stesysf_2}{\ebd}{\eft[1]}
  \ar{r}{s(\stesys)_0}
  \ar{d}[swap]{\ebd\circ\pullbackpr{1}{\ebd}{\eft[1]}}
  &
\stesysf 
  \ar{d}{\eft}
  \\
\stesysf 
  \ar{r}[swap]{\eft}
  &
\stesysc
\end{tikzcd}
\end{equation*}
commutes. A \emph{pre-substitution object} is an extension object
together with a pre-substitution.
\end{defn}

\begin{defn}
A \emph{pre-substitution homomorphism} is an extension homomorphism $f:\stesys'\to \stesys$
for which the square
\begin{equation*}
\begin{tikzcd}[column sep=huge]
\cobesys{\stesyst'}{\famesys{\famesys{\stesys'}}}{\ebd'}{\eft[1]'}
  \ar{r}{\pullback{f^t}{\famehom{\famehom{f}}}{\ebd}{\eft[1]}}
  \ar{d}[swap]{\mathbf{s}'(\stesys')}
  &
\cobesys{\stesyst}{\famesys{\famesys{\stesys}}}{\ebd}{\eft[1]}
  \ar{d}{\mathbf{s}(\stesys)}
  \\
\famesys{\stesys'}
  \ar{r}[swap]{\famehom{f}}
  &
\famesys{\stesys}
\end{tikzcd}
\end{equation*}
commutes.
\end{defn}

\begin{lem}
The change of base of a pre-substitution object is again a pre-substitution object.
\end{lem}

\begin{lem}
If $\stesys$ is a pre-substitution object, then so is $\famesys{\stesys}$ with
$\mathbf{s}(\famesys{\stesys})$ defined to be the unique extension homomorphism
rendering the diagram
\begin{equation*}
\begin{tikzcd}
\cobesys{\stesyst_2}{\famesys{\famesys{\famesys{\stesys}}}}{\ebd[1]}{\eft[2]}
  \ar[dotted]{dr}{\mathbf{s}(\famesys{\stesys})}
  \ar{rr}{\pullback{\pullbackpr{2}{\ectxext}{\eft\circ\ebd}}{\boldsymbol{\beta}_\mathbf{2}}{\ebd}{\eft[1]}}
  \ar{dd}[swap]{\ebd[1]\circ\pullbackpr{1}{\ebd[1]}{\eft[2]}}
  &
  {}&
\cobesys{\stesyst}{\famesys{\famesys{\stesys}}}{\ebd}{\eft[1]}
  \ar{d}{\mathbf{s}(\stesys)}
  \\
  {}&
\famesys{\famesys{\stesys}}
  \ar{r}{\boldsymbol{\beta}}
  \ar{d}{\eft[1]}
  &
\famesys{\stesys}
  \ar{d}{\eft}
  \\
\stesysf_2
  \ar{r}[swap]{\eft[1]}
  &
\stesysf
  \ar{r}[swap]{\ectxext}
  &
\stesysc
\end{tikzcd}
\end{equation*}
commutative.
\end{lem}

\begin{proof}
The requirement on pre-substitutions holds by construction.
\end{proof}

It makes sense now to consider the possibility that the pre-substitution
itself is a pre-substitution homomorphism.

\begin{defn}
A \emph{substitution object} is a pre-substitution object for which substitution is
a pre-substitution homomorphism.
\end{defn}

\begin{cor}
The change of base of a substitution object is again a substitution object.
\end{cor}

\begin{cor}
If $\stesys$ is a substitution object, then so is $\famesys{\stesys}$.
\end{cor}

\subsection{Extension objects with empty context and families}

\begin{defn}
An extension object $\stesys$ is said to have \emph{empty families} if there
is a section
\begin{equation*}
\phi_1(\stesys):\stesysc\to\stesysf
\end{equation*}
of $\eft$, satisfying the following additional properties:
\begin{enumerate}
\item $\phi_1(\stesys)$ is also a section of $\ectxext$.
\item The unique morphism $\phi_2$ rendering the diagram
\begin{equation*}
\begin{tikzcd}[column sep=huge]
\stesysf
  \ar[bend left=20]{drr}{\phi_1(\stesys)\circ\ectxext}
  \ar[equals,bend right=20]{ddr}
  \ar[dotted]{dr}{\phi_2}
  \\
  {}&
\stesysf_2
  \ar{d}{\eft[1]}
  \ar{r}{\pullbackpr{2}{\ectxext}{\eft}}
  &
\stesysf
  \ar{d}{\eft}
  \\
  {}&
\stesysf
  \ar{r}[swap]{\ectxext}
  &
\stesysc
\end{tikzcd}
\end{equation*}
commutative, is a section of $\efamext$.
\item The unique morphism $\iota_1$ rendering the diagram
\begin{equation*}
\begin{tikzcd}[column sep=huge]
\stesysf
  \ar[equals,bend left=20]{drr}
  \ar[bend right=20]{ddr}[swap]{\phi_1(\stesys)\circ\eft}
  \ar[dotted]{dr}{\iota_1}
  \\
  {}&
\stesysf_2
  \ar{d}{\eft[1]}
  \ar{r}[swap]{\pullbackpr{2}{\ectxext}{\eft}}
  &
\stesysf
  \ar{d}{\eft}
  \\
  {}&
\stesysf
  \ar{r}[swap]{\ectxext}
  &
\stesysc
\end{tikzcd}
\end{equation*}
commutative, is a section of $\efamext$.
\end{enumerate}
\end{defn}

\begin{defn}
A homomorphism of extension objects with empty families is an extension
homomorphism $f:\stesys'\to\stesys$ for which the diagram
\begin{equation*}
\begin{tikzcd}
\stesysf'
  \ar{r}{f_1}
  &
\stesysf
  \\
\stesysc'
  \ar{u}{\phi_1(\stesys')}
  \ar{r}[swap]{f_0}
  &
\stesysc
  \ar{u}[swap]{\phi_1(\stesys)}
\end{tikzcd}
\end{equation*}
commutes.
\end{defn}

\begin{lem}
Suppose $\stesys$ is an extension object with empty families. Then
$\famesys{\stesys}$ is an extension object with empty families, with
$\phi_1(\famesys{\stesys}):=\phi_2$. 
\end{lem}

\begin{lem}
Suppose $\stesys$ is an extension object with empty families and consider
$p:\stesysc\rightarrow X\leftarrow Y:p$. Then
$\cobesys{Y}{\stesys}{g}{p}$ is an extension object with empty families
with $\phi_1(\cobesys{Y}{\stesys}{g}{p}):=g^\ast(\phi_1)$.
\end{lem}

\subsection{E-objects}
\begin{defn}
An \emph{E-object} is an extension object with the structure of a projection object,
the structure of a substitution object and which has an empty context and families,
such that additionally:
\begin{enumerate}
\item substitution is a projection homomorphism
\item weakening is a substitution homomorphism
\item both weakening and substitution are empty-CF homomorphisms.
\item 
\end{enumerate}
\end{defn}

### 2014-11-03 Algebras/article-models-cat.tex
worked on presentation

@@
   \ar{r}{\efamext} 
-  \ar{d}[swap]{\eft[1]} 
+  \ar{d}[swap]{\eft[2]} 
   & 
@@
   \\
-\eft[1] 
+\eft[2] 
   & := \pullbackpr{1}{\ectxext}{\eft} : \stesysf_2\to\stesysf
   \\
-\stesysf_3 & := \pullback{\stesysf_2}{\stesysf_2}{\efamext}{\eft[1]}
-  \\
-\eft[2]
-  & := \pullbackpr{1}{\efamext}{\eft[1]} : \stesysf_3\to\stesysf_2.
+\stesysf_3 & := \pullback{\stesysf_2}{\stesysf_2}{\efamext}{\eft[2]}
+  \\
+\eft[3]
+  & := \pullbackpr{1}{\efamext}{\eft[2]} : \stesysf_3\to\stesysf_2.
 \end{align*}
@@
   \ar{rr}{\pullback{\pullbackpr{2}{\ectxext}{\eft}}{\pullbackpr{2}{\ectxext}{\eft}}{\ectxext}{\eft}}
-  \ar{dd}[swap]{\eft[2]}
+  \ar{dd}[swap]{\eft[3]}
   & 
@@
 \stesysf_2
-  \ar{d}[swap]{\eft[1]}
+  \ar{d}[swap]{\eft[2]}
   \ar{r}{\pullbackpr{2}{\ectxext}{\eft}}
@@
 \stesysf_2
-  \ar{r}[swap]{\eft[1]}
+  \ar{r}[swap]{\eft[2]}
   &
@@
   & := 
-\pullback{\stesysf_3}{\stesysf_3}{\eext{2}}{\eft[2]}
-  \\
-\eft[3] 
+\pullback{\stesysf_3}{\stesysf_3}{\eext{2}}{\eft[3]}
+  \\
+\eft[4] 
   & := 
-\pullbackpr{1}{\eext{2}}{\eft[2]}.
+\pullbackpr{1}{\eext{2}}{\eft[3]}.
 \end{align*}
@@
   \ar[dotted]{dr}{\eext{3}}
-  \ar{rr}{\pullback{\pullbackpr{2}{\efamext}{\eft[1]}}{\pullbackpr{2}{\efamext}{\eft[1]}}{\efamext}{\eft[1]}}
-  \ar{dd}[swap]{\eft[3]}
+  \ar{rr}{\pullback{\pullbackpr{2}{\efamext}{\eft[2]}}{\pullbackpr{2}{\efamext}{\eft[2]}}{\efamext}{\eft[2]}}
+  \ar{dd}[swap]{\eft[4]}
   & 
@@
 \stesysf_3
-  \ar{d}[swap]{\eft[2]}
-  \ar{r}{\pullbackpr{2}{\efamext}{\eft[1]}}
-  &
-\stesysf_2
-  \ar{d}{\eft[1]}
-  \\
-\stesysf_3
-  \ar{r}[swap]{\eft[2]}
+  \ar{d}[swap]{\eft[3]}
+  \ar{r}{\pullbackpr{2}{\efamext}{\eft[2]}}
+  &
+\stesysf_2
+  \ar{d}{\eft[2]}
+  \\
+\stesysf_3
+  \ar{r}[swap]{\eft[3]}
   &
@@
 commutes,
-so we may define $\eext{3}$ to be the unique map which renders the diagram. It
+so we may define $\eext{3}$ to be the unique map which renders the diagram
+commutative. It
 is straightforward to continue this process by induction, but we shall need not
@@
 
-\begin{defn} An extension object is a pre-extension object $\stesys$ for which 
+\begin{defn} 
+An extension object is a pre-extension object $\stesys$ for which 
 the diagrams
@@
 \stesysf_3
-  \ar{d}[swap]{\pullbackpr{2}{\efamext}{\eft[1]}}
+  \ar{d}[swap]{\pullbackpr{2}{\efamext}{\eft[2]}}
   \ar{r}{\eext{2}}
@@
 \begin{tikzcd}[column sep=tiny]
-\pullback{\stesysf}{\stesysf_2}{\ectxext}{\eft\circ\eft[1]}
+\pullback{\stesysf}{\stesysf_2}{\ectxext}{\eft\circ\eft[2]}
   \ar[dotted]{rr}{\alpha}
@@
 There is a unique morphism $\alpha:
-\pullback{\stesysf}{\stesysf_2}{\ectxext}{\eft\circ\eft[1]}\to\stesysf_3$
+\pullback{\stesysf}{\stesysf_2}{\ectxext}{\eft\circ\eft[2]}\to\stesysf_3$
 rendering the diagram
@@
 \begin{tikzcd}[column sep=large]
-\pullback{\stesysf}{\stesysf_2}{\ectxext}{\eft\circ\eft[1]}
-  \ar[bend left=10,yshift=.5ex]{drrr}{\pullbackpr{2}{\ectxext}{\eft}\circ\pullbackpr{2}{\ectxext}{\eft\circ\eft[1]}}
-  \ar[bend right=10]{ddr}[swap]{\pullback{\catid{\stesysf}}{\eft[1]}{\ectxext}{\eft}}
+\pullback{\stesysf}{\stesysf_2}{\ectxext}{\eft\circ\eft[2]}
+  \ar[bend left=10,yshift=.5ex]{drrr}{\pullbackpr{2}{\ectxext}{\eft}\circ\pullbackpr{2}{\ectxext}{\eft\circ\eft[2]}}
+  \ar[bend right=10]{ddr}[swap]{\pullback{\catid{\stesysf}}{\eft[2]}{\ectxext}{\eft}}
   \ar[dotted]{dr}{\alpha}
@@
 & \stesysf_3
-  \ar{r}{\pullbackpr{2}{\efamext}{\eft[1]}}
-  \ar{d}{\eft[2]}
-  &
-\stesysf_2
-  \ar{d}[swap]{\eft[1]}
+  \ar{r}{\pullbackpr{2}{\efamext}{\eft[2]}}
+  \ar{d}{\eft[3]}
+  &
+\stesysf_2
+  \ar{d}[swap]{\eft[2]}
   \ar{r}[swap]{\pullbackpr{2}{\ectxext}{\eft}}
@@
 \stesyst_2
-  \ar{d}{\ebd[1]}
-  \\
-\stesysf_2
-  \ar{d}{\eft[1]}
+  \ar{d}{\ebd[2]}
+  \\
+\stesysf_2
+  \ar{d}{\eft[2]}
   \\
@@
   \\
-\ebd[1]
+\ebd[2]
   & := \ectxext^\ast(\ebd),
@@
   & :=
-\pullback{\beta_2}{\beta_2}{\efamext}{\eft[1]}
+\pullback{\beta_2}{\beta_2}{\efamext}{\eft[2]}
   & &
@@
 \stesysf_4
-  \ar{r}{\pullback{\pullbackpr{2}{\efamext}{\eft[1]}}{\pullbackpr{2}{\efamext}{\eft[1]}}{\efamext}{\eft[1]}}
+  \ar{r}{\pullback{\pullbackpr{2}{\efamext}{\eft[2]}}{\pullbackpr{2}{\efamext}{\eft[2]}}{\efamext}{\eft[2]}}
   \ar{d}[swap]{\beta_3}
@@
 \pullbackpr{1}{\ectxext}{\eft}\circ\beta_2\circ
-  (\pullback{\pullbackpr{2}{\efamext}{\eft[1]}}{\pullbackpr{2}{\efamext}{\eft[1]}}{\efamext}{\eft[1]})
+  (\pullback{\pullbackpr{2}{\efamext}{\eft[2]}}{\pullbackpr{2}{\efamext}{\eft[2]}}{\efamext}{\eft[2]})
   & =
@@
 \pullbackpr{2}{\ectxext}{\eft}\circ\beta_2\circ
-  (\pullback{\pullbackpr{2}{\efamext}{\eft[1]}}{\pullbackpr{2}{\efamext}{\eft[1]}}{\efamext}{\eft[1]})
+  (\pullback{\pullbackpr{2}{\efamext}{\eft[2]}}{\pullbackpr{2}{\efamext}{\eft[2]}}{\efamext}{\eft[2]})
   & =
@@
 \pullbackpr{1}{\ectxext}{\eft}\circ\beta_2\circ\eext{3}
-  & = \beta_1\circ\eft[2]\circ\eft[3]
+  & = \beta_1\circ\eft[3]\circ\eft[4]
   \\
 \pullbackpr{1}{\ectxext}{\eft}\circ\eext{2}\circ\beta_3
-  & = \beta_1\circ\eft[2]\circ\eft[3].
+  & = \beta_1\circ\eft[3]\circ\eft[4].
 \end{align*}
@@
 \stesysf_4
-  \ar{dd}[swap]{\pullback{\pullbackpr{2}{\efamext}{\eft[1]}}{\pullbackpr{2}{\efamext}{\eft[1]}}{\efamext}{\eft[1]}}
+  \ar{dd}[swap]{\pullback{\pullbackpr{2}{\efamext}{\eft[2]}}{\pullbackpr{2}{\efamext}{\eft[2]}}{\efamext}{\eft[2]}}
   \ar{r}{\eext{3}}
@@
   \ar{r}{\beta_2}
-  \ar{d}[swap]{\pullbackpr{2}{\efamext}{\eft[1]}}
+  \ar{d}[swap]{\pullbackpr{2}{\efamext}{\eft[2]}}
   &
@@
   \ar{r}{\beta_3}
-  \ar{d}[swap]{\pullback{\pullbackpr{2}{\efamext}{\eft[1]}}{\pullbackpr{2}{\efamext}{\eft[1]}}{\efamext}{\eft[1]}}
+  \ar{d}[swap]{\pullback{\pullbackpr{2}{\efamext}{\eft[2]}}{\pullbackpr{2}{\efamext}{\eft[2]}}{\efamext}{\eft[2]}}
   &
@@
 \stesysf_4
-  \ar{d}[swap]{\pullbackpr{2}{\eext{2}}{\eft[2]}}
+  \ar{d}[swap]{\pullbackpr{2}{\eext{2}}{\eft[3]}}
   \ar{r}{\eext{3}}
@@
   & =
-\pullbackpr{1}{\ectxext}{\eft}\circ\eext{2}\circ\pullbackpr{2}{\eext{2}}{\eft[2]}
+\pullbackpr{1}{\ectxext}{\eft}\circ\eext{2}\circ\pullbackpr{2}{\eext{2}}{\eft[3]}
   \\
@@
   & =
-\pullbackpr{2}{\ectxext}{\eft}\circ\eext{2}\circ\pullbackpr{2}{\eext{2}}{\eft[2]}.
+\pullbackpr{2}{\ectxext}{\eft}\circ\eext{2}\circ\pullbackpr{2}{\eext{2}}{\eft[3]}.
 \end{align*}
@@
   =
-\eft[1]\circ\eft[2]\circ\eft[3]
+\eft[2]\circ\eft[3]\circ\eft[4]
 \end{equation*}
@@
 \begin{equation*}
-\pullbackpr{1}{\ectxext}{\eft}\circ\eext{2}\circ\pullbackpr{2}{\eext{2}}{\eft[2]}
+\pullbackpr{1}{\ectxext}{\eft}\circ\eext{2}\circ\pullbackpr{2}{\eext{2}}{\eft[3]}
   =
-\eft[1]\circ\eft[2]\circ\eft[3].
+\eft[2]\circ\eft[3]\circ\eft[4].
 \end{equation*}
@@
 \stesysf_4
-  \ar{r}{\pullbackpr{2}{\eext{2}}{\eft[2]}}
+  \ar{r}{\pullbackpr{2}{\eext{2}}{\eft[3]}}
   \ar{d}[swap]{\beta_3}
@@
 \stesysf_3
-  \ar{r}{\pullbackpr{2}{\efamext}{\eft[1]}}
+  \ar{r}{\pullbackpr{2}{\efamext}{\eft[2]}}
   \ar{dr}[swap]{\eext{2}}
@@
 Let $\stesys$ and $\stesys'$ be pre-extension algebras. A \emph{pre-extension 
-homomorphism $f$ from $\stesys$ to $\stesys'$} is a triple $(f_0,f_1,f^t)$ 
+homomorphism $f$ from $\stesys'$ to $\stesys$} is a triple $(f_0,f_1,f^t)$ 
 consisting of morphisms
@@
 \begin{tikzcd}
+\stesyst' 
+  \ar{r}{f^t}
+  \ar{d}[swap]{\ebd'}
+  &
 \stesyst
-  \ar{r}{f^t}
-  \ar{d}[swap]{\ebd}
-  &
-\stesyst'
-  \ar{d}{\ebd'}
-  \\
-\stesysf 
+  \ar{d}{\ebd}
+  \\
+\stesysf'
   \ar{r}{f_1}
-  \ar{d}[swap]{\eft}
-  &
+  \ar{d}[swap]{\eft'}
+  &
+\stesysf
+  \ar{d}{\eft}
+  \\
+\stesysc' 
+  \ar{r}[swap]{f_0}
+  &
+\stesysc
+\end{tikzcd}
+\end{equation*}
+such that the indicated squares commute, for which furthermore the squares
+\begin{equation*}
+\begin{tikzcd}
 \stesysf' 
-  \ar{d}{\eft'}
-  \\
-\stesysc 
+  \ar{r}{f_1}
+  \ar{d}[swap]{\ectxext'}
+  &
+\stesysf
+  \ar{d}{\ectxext}
+  \\
+\stesysc'
   \ar{r}[swap]{f_0}
   &
-\stesysc'
-\end{tikzcd}
-\end{equation*}
-such that the indicated squares commute, for which furthermore the squares
-\begin{equation*}
-\begin{tikzcd}
-\stesysf \ar{r}{f_1}
-  \ar{d}[swap]{\ectxext}
-  &
+\stesysc
+\end{tikzcd}
+\end{equation*}
+and
+\begin{equation*}
+\begin{tikzcd}[column sep=huge]
+\stesysf'\times_{\ectxext',\eft'} \stesysf'
+  \ar{r}{f_1\times_{\ectxext,\eft} f_1}
+  \ar{d}[swap]{\efamext'}
+  &
+\stesysf\times_{\ectxext,\eft} \stesysf
+  \ar{d}{\efamext}
+  \\
 \stesysf'
-  \ar{d}{\ectxext'}
-  \\
-\stesysc
-  \ar{r}[swap]{f_0}
-  &
-\stesysc'
-\end{tikzcd}
-\end{equation*}
-and
-\begin{equation*}
-\begin{tikzcd}[column sep=large]
-\stesysf\times_{\ectxext,\eft} \stesysf
-  \ar{r}{f_1\times_{\ectxext',\eft'} f_1}
-  \ar{d}[swap]{\efamext}
-  &
-\stesysf'\times_{\ectxext',\eft'} \stesysf'
-  \ar{d}{\efamext'}
-  \\
-\stesysf
   \ar{r}[swap]{f_1}
   &
-\stesysf'
+\stesysf
 \end{tikzcd}
@@
 \begin{align*}
-f_2 & := \pullback{f_1}{f_1}{\ectxext'}{\eft'}
-  \\
-f_3 & := \pullback{f_2}{f_2}{\efamext'}{\eft[1]'}.
+f_2 & := \pullback{f_1}{f_1}{\ectxext}{\eft}
+  \\
+f_3 & := \pullback{f_2}{f_2}{\efamext}{\eft[2]}.
 \end{align*}
@@
 \begin{equation*}
-\begin{tikzcd}[column sep=large]
+\begin{tikzcd}[column sep=huge]
 \stesyst_2
   \ar{r}{\pullbackpr{2}{\ectxext}{\eft\circ\ebd}}
-  \ar{d}[swap]{\ebd[1]}
+  \ar{d}[swap]{\ebd[2]}
   &
@@
   \ar{r}{\pullbackpr{2}{\ectxext}{\eft}}
-  \ar{d}[swap]{\eft[1]}
+  \ar{d}[swap]{\eft[2]}
   &
@@
   \ar{r}{\beta^t}
-  \ar{d}[swap]{\ebd[2]}
+  \ar{d}[swap]{\ebd[3]}
   &
 \stesyst_2
-  \ar{d}{\ebd[1]}
+  \ar{d}{\ebd[2]}
   \\
@@
   \ar{r}{\beta_2}
-  \ar{d}[swap]{\eft[2]}
-  &
-\stesysf_2
-  \ar{d}{\eft[1]}
+  \ar{d}[swap]{\eft[3]}
+  &
+\stesysf_2
+  \ar{d}{\eft[2]}
   \\
@@
   \ar{r}{\beta^t_2}
-  \ar{d}[swap]{\ebd[3]}
+  \ar{d}[swap]{\ebd[4]}
   &
 \stesyst_3
-  \ar{d}{\ebd[2]}
+  \ar{d}{\ebd[3]}
   \\
@@
   \ar{r}{\beta_3}
-  \ar{d}[swap]{\eft[3]}
-  &
-\stesysf_3
-  \ar{d}{\eft[2]}
+  \ar{d}[swap]{\eft[4]}
+  &
+\stesysf_3
+  \ar{d}{\eft[3]}
   \\
@@
   \ar{r}{f^t_2}
-  \ar{d}[swap]{\ebd[1]'}
+  \ar{d}[swap]{\ebd[2]'}
   &
 \stesyst_2
-  \ar{d}{\ebd[1]}
+  \ar{d}{\ebd[2]}
   \\
@@
   \ar{r}{f_2}
-  \ar{d}[swap]{\eft[1]'}
-  &
-\stesysf_2
-  \ar{d}{\eft[1]}
+  \ar{d}[swap]{\eft[2]'}
+  &
+\stesysf_2
+  \ar{d}{\eft[2]}
   \\
@@
   \ar{r}{\eext{2}} 
-  \ar{d}[swap]{\pullbackpr{2}{\efamext}{\eft[1]}} 
+  \ar{d}[swap]{\pullbackpr{2}{\efamext}{\eft[2]}} 
   & 
@@
   \ar{ur}{\efamext}
-  \ar{dr}[swap]{\eft[1]}
+  \ar{dr}[swap]{\eft[2]}
   &
@@
 \begin{equation*}
-|[p,f]|:\stesys'\to\famesys{\famesys{\stesys}}
+[p,f]:\stesys'\to\famesys{\famesys{\stesys}}
 \end{equation*}
@@
 \begin{itemize}
-\item let $|[p,f]|_0:\stesysc'\to\stesysf_2$ be the unique morphism rendering
+\item let $[p,f]_0:\stesysc'\to\stesysf_2$ be the unique morphism rendering
 the diagram
@@
   \ar[swap,bend right=10]{ddr}{p}
-  \ar[dotted]{dr}[near end]{|[p,f]|_0}
+  \ar[dotted]{dr}[near end]{[p,f]_0}
   \\
@@
   \ar{r}[swap]{\pullbackpr{2}{\ectxext}{\eft}}
-  \ar{d}{\eft[1]}
+  \ar{d}{\eft[2]}
   &
@@
 commutative.
-\item Let $|[p,f]|_1:\stesysf'\to\stesysf_3$ be the unique morphism rendering
+\item Let $[p,f]_1:\stesysf'\to\stesysf_3$ be the unique morphism rendering
 the diagram
@@
   \ar[swap]{dd}{\eft'}
-  \ar[dotted]{dr}[near end]{|[p,f]|_1}
-  \\
-  {}&
-\stesysf_3
-  \ar{r}[swap]{\pullbackpr{2}{\efamext}{\eft[1]}}
-  \ar{d}{\eft[1]}
-  &
-\stesysf_2
-  \ar{d}{\eft[1]}
+  \ar[dotted]{dr}[near end]{[p,f]_1}
+  \\
+  {}&
+\stesysf_3
+  \ar{r}[swap]{\pullbackpr{2}{\efamext}{\eft[2]}}
+  \ar{d}{\eft[2]}
+  &
+\stesysf_2
+  \ar{d}{\eft[2]}
   \\
 \stesysc'
-  \ar{r}[swap]{|[p,f]|_0}
+  \ar{r}[swap]{[p,f]_0}
   &
@@
 commutative.
-\item Let $|[p,f]|^t:\stesyst'\to\stesyst_3$ be the unique morphism rendering
+\item Let $[p,f]^t:\stesyst'\to\stesyst_3$ be the unique morphism rendering
 the diagram
@@
   \ar[swap]{dd}{\eft'\circ\ebd'}
-  \ar[dotted]{dr}[near end]{|[p,f]|^t}
+  \ar[dotted]{dr}[near end]{[p,f]^t}
   \\
@@
 \stesyst_3
-  \ar{r}[swap]{\pullbackpr{2}{\efamext}{\eft[1]\circ\ebd[1]}}
-  \ar{d}[swap]{\pullbackpr{1}{\efamext}{\eft[1]\circ\ebd[1]}}
+  \ar{r}[swap]{\pullbackpr{2}{\efamext}{\eft[2]\circ\ebd[2]}}
+  \ar{d}[swap]{\pullbackpr{1}{\efamext}{\eft[2]\circ\ebd[2]}}
   &
 \stesyst_2
-  \ar{d}{\eft[1]\circ\ebd[1]}
+  \ar{d}{\eft[2]\circ\ebd[2]}
   \\
 \stesysc'
-  \ar{r}[swap]{|[p,f]|_0}
+  \ar{r}[swap]{[p,f]_0}
   &
@@
 \begin{lem}
-Under the hypotheses of \autoref{famfamstesys_into}, $|[p,f]|$ is a pre-extension
+Under the hypotheses of \autoref{famfamstesys_into}, $[p,f]$ is a pre-extension
 homomorphism. Moreover, it is the unique pre-extension homomorphism for which
@@
   \ar[swap,bend right=10]{ddr}{p}
-  \ar[dotted]{dr}[near end]{|[p,f]|}
+  \ar[dotted]{dr}[near end]{[p,f]}
   \\
@@
 \famesys{\famesys{\stesys}}
-  \ar{d}{\eft[1]}
+  \ar{d}{\eft[2]}
   \\
@@
 \begin{tikzcd}[column sep=large]
-\cobesys{\stesysf_2}{\famesys{\famesys{\stesys}}}{\eft[1]}{\eft[1]}
+\cobesys{\stesysf_2}{\famesys{\famesys{\stesys}}}{\eft[2]}{\eft[2]}
   \ar{rr}{%
@@
     }
-  \ar[bend right]{ddr}[swap]{\pullbackpr{1}{\eft[1]}{\eft[1]}}
+  \ar[bend right]{ddr}[swap]{\pullbackpr{1}{\eft[2]}{\eft[2]}}
   \ar[dotted]{dr}{\mathbf{w}(\famesys{\stesys})}
@@
   \ar{r}{\boldsymbol{\beta}_\mathbf{2}}
-  \ar{d}[swap]{\eft[2]}
+  \ar{d}[swap]{\eft[3]}
   &
 \famesys{\famesys{\stesys}}
-  \ar{d}{\eft[1]}
+  \ar{d}{\eft[2]}
   \ar{r}{\boldsymbol{\beta}}
@@
 \begin{tikzcd}[column sep=15em]
-\cobesys{\stesysf_2}{\famesys{\stesys}}{\eft\circ\eft[1]}{\eft}
+\cobesys{\stesysf_2}{\famesys{\stesys}}{\eft\circ\eft[2]}{\eft}
   \ar[bend right=10]{dr}[swap]%
-    { [ \pullbackpr{1}{\eft\circ\eft[1]}{\eft},%
+    { [ \pullbackpr{1}{\eft\circ\eft[2]}{\eft},%
         \mathbf{w}(\stesys)%
@@
   \ar{r}{
-    [ \pullbackpr{1}{\eft\circ\eft[1]}{\eft},%
+    [ \pullbackpr{1}{\eft\circ\eft[2]}{\eft},%
       \mathbf{w}(\stesys)%
         \circ%
-      (\pullback{\eft[1]}{\catid{\famesys{\stesys}}}{\eft}{\eft})%
+      (\pullback{\eft[2]}{\catid{\famesys{\stesys}}}{\eft}{\eft})%
       ]}%
   &
-\cobesys{\stesysf_2}{\famesys{\famesys{\stesys}}}{\eft[1]}{\eft[1]}
+\cobesys{\stesysf_2}{\famesys{\famesys{\stesys}}}{\eft[2]}{\eft[2]}
   \ar{d}{\mathbf{w}(\famesys{\stesys})}
@@
 \begin{tikzcd}[column sep=large]
-\stesysf \ar{r}{\mathbf{i}} \ar{d}[swap]{\Delta_{\eft}} & \stesyst_2 \ar{d}{\ebd[1]}\\
+\stesysf \ar{r}{\mathbf{i}} \ar{d}[swap]{\Delta_{\eft}} & \stesyst_2 \ar{d}{\ebd[2]}\\
 \pullback{\stesysf}{\stesysf}{\eft}{\eft} \ar{r}[swap]{w(\stesys)_0} & \stesysf_2
@@
 \begin{equation*}
-\mathbf{s}(\stesys):\cobesys{\stesyst}{\famesys{\famesys{\stesys}}}{\ebd}{\eft[1]}\to \famesys{\stesys}
+\mathbf{s}(\stesys):\cobesys{\stesyst}{\famesys{\famesys{\stesys}}}{\ebd}{\eft[2]}\to \famesys{\stesys}
 \end{equation*}
@@
 \begin{tikzcd}[column sep=large]
-\pullback{\stesyst}{\stesysf_2}{\ebd}{\eft[1]}
+\pullback{\stesyst}{\stesysf_2}{\ebd}{\eft[2]}
   \ar{r}{s(\stesys)_0}
-  \ar{d}[swap]{\ebd\circ\pullbackpr{1}{\ebd}{\eft[1]}}
+  \ar{d}[swap]{\ebd\circ\pullbackpr{1}{\ebd}{\eft[2]}}
   &
@@
 \begin{tikzcd}[column sep=huge]
-\cobesys{\stesyst'}{\famesys{\famesys{\stesys'}}}{\ebd'}{\eft[1]'}
-  \ar{r}{\pullback{f^t}{\famehom{\famehom{f}}}{\ebd}{\eft[1]}}
+\cobesys{\stesyst'}{\famesys{\famesys{\stesys'}}}{\ebd'}{\eft[2]'}
+  \ar{r}{\pullback{f^t}{\famehom{\famehom{f}}}{\ebd}{\eft[2]}}
   \ar{d}[swap]{\mathbf{s}'(\stesys')}
   &
-\cobesys{\stesyst}{\famesys{\famesys{\stesys}}}{\ebd}{\eft[1]}
+\cobesys{\stesyst}{\famesys{\famesys{\stesys}}}{\ebd}{\eft[2]}
   \ar{d}{\mathbf{s}(\stesys)}
@@
 \begin{tikzcd}
-\cobesys{\stesyst_2}{\famesys{\famesys{\famesys{\stesys}}}}{\ebd[1]}{\eft[2]}
+\cobesys{\stesyst_2}{\famesys{\famesys{\famesys{\stesys}}}}{\ebd[2]}{\eft[3]}
   \ar[dotted]{dr}{\mathbf{s}(\famesys{\stesys})}
-  \ar{rr}{\pullback{\pullbackpr{2}{\ectxext}{\eft\circ\ebd}}{\boldsymbol{\beta}_\mathbf{2}}{\ebd}{\eft[1]}}
-  \ar{dd}[swap]{\ebd[1]\circ\pullbackpr{1}{\ebd[1]}{\eft[2]}}
-  &
-  {}&
-\cobesys{\stesyst}{\famesys{\famesys{\stesys}}}{\ebd}{\eft[1]}
+  \ar{rr}{\pullback{\pullbackpr{2}{\ectxext}{\eft\circ\ebd}}{\boldsymbol{\beta}_\mathbf{2}}{\ebd}{\eft[2]}}
+  \ar{dd}[swap]{\ebd[2]\circ\pullbackpr{1}{\ebd[2]}{\eft[3]}}
+  &
+  {}&
+\cobesys{\stesyst}{\famesys{\famesys{\stesys}}}{\ebd}{\eft[2]}
   \ar{d}{\mathbf{s}(\stesys)}
@@
   \ar{r}{\boldsymbol{\beta}}
-  \ar{d}{\eft[1]}
+  \ar{d}{\eft[2]}
   &
@@
 \stesysf_2
-  \ar{d}{\eft[1]}
+  \ar{d}{\eft[2]}
   \ar{r}{\pullbackpr{2}{\ectxext}{\eft}}
@@
 \stesysf_2
-  \ar{d}{\eft[1]}
+  \ar{d}{\eft[2]}
   \ar{r}[swap]{\pullbackpr{2}{\ectxext}{\eft}}

### 2014-11-03 Presentation/presentation.tex
developing E-systems in categories with finite limits
worked on presentation

@@
 terms. Thus there will be 3 sorts.
+\pause
 \item The theory is algebraic and the operations on it are homomorphisms. The
@@
 require.
+\pause
 \item Finite set of rules. The theory is invariant under slicing as a consequence, rather than by
 convention. There will be no rule schemes.
+\pause
 \item The theory is capable of interpreting itself. Investigations to internal
 models of type theory originally motivated me to write type theory this way.
+\pause
+\item The meta-theory should not require anything more than rules for handling
+inferences. In particular, natural numbers should not be required.
 \end{itemize}
@@
 \begin{frame}
+\frametitle{The fundamental structure of E-systems}
+The fundamental structure of an E-system $\stesys$ in a category
+$\cat{C}$ with finite limits consists of
+\begin{equation*}
+\begin{tikzcd}
+\stesyst
+  \ar{d}{\ebd}
+  \\
+\stesysf
+  \ar{d}{\eft}
+  \\
+\stesysc
+\end{tikzcd}
+\end{equation*}
+The inferences of judgmental equality are automatically taken
+care of by the equality the ambient category
+\end{frame}
+
+\begin{frame}
 \frametitle{Extension}
@@
 \begin{frame}
-\frametitle{The empty context and family}
-\begin{small}
-Introduction rules for the empty context and family:
-\begin{align*}
-& \inference
-  { }
-  { \jctx{\emptyc}
-    }
-& & \inference
-  { \jctx{\Gamma}
-    }
-  { \jfam{\Gamma}{\emptyf[\Gamma]}
-    }
-& & \inference
-  { \jctxeq{\Gamma}{\Gamma'}
-    }
-  { \jfameq{\Gamma}{\emptyf[\Gamma]}{\emptyf[\Gamma']}
-    }
-\end{align*}
-\pause
-Compatibility rules for families and contexts:
-\begin{align*}
-& \inference
-  { \jctx{\Gamma}
-    }
-  { \jfam{\emptyc}{\Gamma}
-    } 
-& & \inference
-    { \jfam{\emptyc}{\Gamma}
-      }
-    { \jctx{\Gamma}
-      }
-    \\
-& \inference
-  { \jctxeq{\Gamma}{\Delta}
-    }
-  { \jfameq{\emptyc}{\Gamma}{\Delta}
-    }
-& & \inference
-    { \jfameq{\emptyc}{\Gamma}{\Delta}
-      }
-    { \jctxeq{\Gamma}{\Delta}
-      }
-\end{align*}
-\pause
-Compatibility rules for extension and the empty families:
-\begin{align*}
-& \inference
-  { \jctx{\Gamma}
-    }
-  { \jctxeq{\ctxext{\emptyc}{\Gamma}}{\Gamma}
-    }
-& & \inference
-  { \jctx{\Gamma}
-    }
-  { \jctxeq{\ctxext{\Gamma}{\emptyf}}{\Gamma}
-    }
-  \\
-& \inference
-  { \jfam{\Gamma}{A}
-    }
-  { \jfameq{\Gamma}{\ctxext{\emptyf}{A}}{A}
-    }
-& & \inference
-  { \jfam{\Gamma}{A}
-    }
-  { \jfameq{\Gamma}{{A}{\emptyf}}{A}
-    }
-\end{align*}
-Compatibility of context extension and family extension:
-\begin{align*}
-& \inference
-  { \jfam{\Gamma}{A}
-    }
-  { \jctxeq{(\ctxext{\Gamma}{A})^c}{(\ctxext{\Gamma}{A})^\famsym}
-    }
-\end{align*}
-\end{small}
+\frametitle{Pre-extension algebras}
+A \emph{pre-extension object $\stesys$ in $\cat{C}$} consists of a \emph{fundamental structure}
+in $\cat{C}$ together with the \emph{context extension} and \emph{family extension} operations
+\begin{align*}
+\ectxext% 
+  &%
+: \stesysf \to \stesysc%
+  \\
+\efamext% 
+  &%
+: \stesysff \to \stesysf,%
+\end{align*}
+respectively, such that the diagram
+\begin{equation*}
+\begin{tikzcd}[ampersand replacement=\&]
+\stesysff 
+  \ar{r}{\efamext} 
+  \ar{d}[swap]{\pullbackpr{1}{\ectxext}{\eft}} 
+  \& 
+\stesysf 
+  \ar{d}{\eft}
+  \\
+\stesysf
+  \ar{r}[swap]{\eft} 
+  \& 
+\stesysc
+\end{tikzcd}
+\end{equation*}
+commutes.
+\end{frame}
+
+\begin{frame}
+\frametitle{Notation for pre-extension algebras}
+We introduce the following notation:
+\begin{align*}
+\stesysf_2 
+  & := \stesysff
+  \\
+\eft[2] 
+  & := \pullbackpr{1}{\ectxext}{\eft} : \stesysf_2\to\stesysf
+  \\
+\stesysf_3 & := \pullback{\stesysf_2}{\stesysf_2}{\efamext}{\eft[2]}
+  \\
+\eft[3]
+  & := \pullbackpr{1}{\efamext}{\eft[2]} : \stesysf_3\to\stesysf_2.
+\end{align*}
+Then it follows that the outer square in the diagram
+\begin{equation*}
+\begin{tikzcd}[ampersand replacement=\&,column sep=large]
+\stesysf_3
+  \ar[dotted]{dr}{\eext{2}}
+  \ar{rr}{\pullback{\pullbackpr{2}{\ectxext}{\eft}}{\pullbackpr{2}{\ectxext}{\eft}}{\ectxext}{\eft}}
+  \ar{dd}[swap]{\eft[3]}
+  \& 
+  \&
+\stesysf_2
+  \ar{d}{\efamext}
+  \\
+  \&
+\stesysf_2
+  \ar{d}[swap]{\eft[2]}
+  \ar{r}{\pullbackpr{2}{\ectxext}{\eft}}
+  \&
+\stesysf
+  \ar{d}{\eft}
+  \\
+\stesysf_2
+  \ar{r}[swap]{\eft[2]}
+  \&
+\stesysf
+  \ar{r}[swap]{\ectxext}
+  \&
+\stesysc
+\end{tikzcd}
+\end{equation*}
+commutes. We define $\eext{2}$ to be the unique morphism rendering the above diagram
+commutative. 
+\end{frame}
+
+\begin{frame}
+\frametitle{Notation for extension algebras (continued)}
+Now we may continue to define
+\begin{align*}
+\stesysf_4 
+  & := 
+\pullback{\stesysf_3}{\stesysf_3}{\eext{2}}{\eft[3]}
+  \\
+\eft[4] 
+  & := 
+\pullbackpr{1}{\eext{2}}{\eft[3]}.
+\end{align*}
+Then we see that the outer square of the diagram
+\begin{equation*}
+\begin{tikzcd}[ampersand replacement=\&,column sep=large]
+\stesysf_4
+  \ar[dotted]{dr}{\eext{3}}
+  \ar{rr}{\pullback{\pullbackpr{2}{\efamext}{\eft[2]}}{\pullbackpr{2}{\efamext}{\eft[2]}}{\efamext}{\eft[2]}}
+  \ar{dd}[swap]{\eft[4]}
+  \& 
+  \&
+\stesysf_3
+  \ar{d}{\eext{2}}
+  \\
+  \&
+\stesysf_3
+  \ar{d}[swap]{\eft[3]}
+  \ar{r}{\pullbackpr{2}{\efamext}{\eft[2]}}
+  \&
+\stesysf_2
+  \ar{d}{\eft[2]}
+  \\
+\stesysf_3
+  \ar{r}[swap]{\eft[3]}
+  \&
+\stesysf_2
+  \ar{r}[swap]{\efamext}
+  \&
+\stesysf
+\end{tikzcd}
+\end{equation*}
+commutes,
+so we may define $\eext{3}$ to be the unique map which renders the diagram commutative.
+\end{frame}
+
+\begin{frame}
+\frametitle{Extension algebras}
+An extension object is a pre-extension object $\stesys$ for which 
+the diagrams
+\begin{equation*}
+\begin{tikzcd}[ampersand replacement=\&]
+\stesysf_2 
+  \ar{d}[swap]{\pullbackpr{2}{\ectxext}{\eft}} 
+  \ar{r}{\efamext} 
+  \& 
+\stesysf 
+  \ar{d}{\ectxext}
+  \\
+\stesysf 
+  \ar{r}[swap]{\ectxext} 
+  \& 
+\stesysc
+\end{tikzcd}
+\qquad
+\begin{tikzcd}[ampersand replacement=\&]
+\stesysf_3
+  \ar{d}[swap]{\pullbackpr{2}{\efamext}{\eft[2]}}
+  \ar{r}{\eext{2}}
+  \& 
+\stesysf_2 
+  \ar{d}{\efamext} 
+  \\
+\stesysf_2 
+  \ar{r}[swap]{\efamext} 
+  \&
+\stesysf
+\end{tikzcd}
+\end{equation*}
+commute.
+\end{frame}
+
+\begin{frame}
+\frametitle{Local pre-extension algebras}
+Suppose that $\stesys$ is a pre-extension algebra. Then we define
+the pre-extension algebra $\famesys{\stesys}$ to consist of the
+fundamental structure
+\begin{equation*}
+\begin{tikzcd}[ampersand replacement=\&]
+\stesyst_2
+  \ar{d}{\ebd[2]}
+  \\
+\stesysf_2
+  \ar{d}{\eft[2]}
+  \\
+\stesysf
+\end{tikzcd}
+\end{equation*}
+where
+\begin{align*}
+\stesyst_2 
+  & := \pullback{\stesysf}{\stesyst}{\ectxext}{\eft\circ\ebd},
+  &
+\ebd[2]
+  & := \ectxext^\ast(\ebd),
+\end{align*}
+with the extension operations
+\begin{align*}
+\efamext 
+  & 
+: \stesysf_2\to\stesysf,
+  &
+\eext{2} 
+  & 
+: \stesysf_3\to\stesysf_2.
+\end{align*}
+\end{frame}
+
+\begin{frame}
+\frametitle{Local extension algebras}
+\begin{theorem}
+If $\stesys$ is an extension algebra, then so is $\famesys{\stesys}$.
+\end{theorem}
+\end{frame}
+
+\begin{frame}
+\frametitle{Pre-extension homomorphisms}
+\begin{footnotesize}
+Let $\stesys$ and $\stesys'$ be pre-extension algebras. A \emph{pre-extension 
+homomorphism $f$ from $\stesys'$ to $\stesys$} is a triple $(f_0,f_1,f^t)$ 
+consisting of morphisms
+\begin{equation*}
+\begin{tikzcd}[ampersand replacement=\&]
+\stesyst' 
+  \ar{r}{f^t}
+  \ar{d}[swap]{\ebd'}
+  \&
+\stesyst
+  \ar{d}{\ebd}
+  \\
+\stesysf'
+  \ar{r}{f_1}
+  \ar{d}[swap]{\eft'}
+  \&
+\stesysf
+  \ar{d}{\eft}
+  \\
+\stesysc' 
+  \ar{r}[swap]{f_0}
+  \&
+\stesysc
+\end{tikzcd}
+\end{equation*}
+such that the indicated squares commute, for which furthermore the squares
+\begin{equation*}
+\begin{tikzcd}[ampersand replacement=\&]
+\stesysf' 
+  \ar{r}{f_1}
+  \ar{d}[swap]{\ectxext'}
+  \&
+\stesysf
+  \ar{d}{\ectxext}
+  \\
+\stesysc'
+  \ar{r}[swap]{f_0}
+  \&
+\stesysc
+\end{tikzcd}
+\quad
+\text{and}
+\quad
+\begin{tikzcd}[ampersand replacement=\&,column sep=huge]
+\stesysf_2'
+  \ar{r}{f_1\times_{\ectxext,\eft} f_1}
+  \ar{d}[swap]{\efamext'}
+  \&
+\stesysf_2
+  \ar{d}{\efamext}
+  \\
+\stesysf'
+  \ar{r}[swap]{f_1}
+  \&
+\stesysf
+\end{tikzcd}
+\end{equation*}
+commute. Composition and the identity homomorphism are defined in the expected way.
+\end{footnotesize}
+\end{frame}
+
+\begin{frame}
+\begin{definition}
+A pre-extension homomorphism between extension objects is called an extension
+homomorphism.
+\end{definition}
+
+\begin{lemma}
+Let $\stesys$ be a pre-extension algebra. Then $\stesys$ is an extension
+algebra if and only if we have the extension homomorphisms
+$\mathbf{e}_0:\famesys{\stesys}\to\stesys$
+and
+$\mathbf{e}_1:\famesys{\famesys{\stesys}}\to\famesys{\stesys}$
+given by
+\begin{equation*}
+\begin{tikzcd}[ampersand replacement=\&,column sep=huge]
+\stesyst_2
+  \ar{r}{\pullbackpr{2}{\ectxext}{\eft\circ\ebd}}
+  \ar{d}[swap]{\ebd[2]}
+  \&
+\stesyst
+  \ar{d}{\ebd}
+  \\
+\stesysf_2
+  \ar{r}{\pullbackpr{2}{\ectxext}{\eft}}
+  \ar{d}[swap]{\eft[2]}
+  \&
+\stesysf
+  \ar{d}{\eft}
+  \\
+\stesysf
+  \ar{r}[swap]{\ectxext}
+  \&
+\stesysc
+\end{tikzcd}
+\qquad\text{and}\qquad
+\begin{tikzcd}[ampersand replacement=\&,column sep=huge]
+\stesyst_3
+  \ar{r}{\pullbackpr{2}{\efamext}{\eft[2]\circ\ebd[2]}}
+  \ar{d}[swap]{\ebd[3]}
+  \&
+\stesyst_2
+  \ar{d}{\ebd[2]}
+  \\
+\stesysf_3
+  \ar{r}{\pullbackpr{2}{\efamext}{\eft[2]}}
+  \ar{d}[swap]{\eft[3]}
+  \&
+\stesysf_2
+  \ar{d}{\eft[2]}
+  \\
+\stesysf_2
+  \ar{r}[swap]{\efamext}
+  \&
+\stesysf
+\end{tikzcd}
+\end{equation*}
+\end{lemma}
+\end{frame}
+
+\begin{frame}
+\frametitle{Mixed commutative diagrams}
+Suppose $f:\stesys\to\stesys'$ is a pre-extension homomorphism. We say that
+a diagram
+\begin{equation*}
+\begin{tikzcd}[ampersand replacement=\&]
+\stesys
+  \ar{r}{f}
+  \ar{d}[swap]{p}
+  \&
+\stesys'
+  \ar{d}{p'}
+  \\
+X \ar{r}[swap]{g}
+  \&
+Y
+\end{tikzcd}
+\end{equation*}
+commutes if the diagram
+\begin{equation*}
+\begin{tikzcd}[ampersand replacement=\&]
+\stesysc
+  \ar{r}{f_0}
+  \ar{d}[swap]{p}
+  \&
+\stesysc'
+  \ar{d}{p'}
+  \\
+X \ar{r}[swap]{g}
+  \&
+Y
+\end{tikzcd}
+\end{equation*}
+commutes.
+\end{frame}
+
+\begin{frame}
+\frametitle{Change of base of pre-extension algebras}
+Let $\stesys$ be a pre-extension algebra and consider $p:\stesysc\rightarrow X\leftarrow Y:g$.
+Then there is a pre-extension algebra $\cobesys{Y}{\stesys}{g}{p}$ with projections, such that for every diagram
+\begin{equation*}
+\begin{tikzcd}[ampersand replacement=\&,column sep=large]
+\stesys'
+  \ar[bend right=10]{ddr}[swap]{p'}
+  \ar[bend left=10]{rrd}{f}
+  \ar[dotted]{dr}[near end]{[p',f]}
+  \\
+  {}\&
+\cobesys{Y}{\stesys}{g}{p}
+  \ar{d}{\pullbackpr{1}{g}{p}}
+  \ar{r}[swap]{\pullbackpr{2}{g}{p}}
+  \&
+\stesys
+  \ar{d}{p}
+  \\
+  {}\&
+Y \ar{r}[swap]{g}
+  \&
+X
+\end{tikzcd}
+\end{equation*}
+of which the outer square commutes, the pre-extension homomorphism $[p',f]$ exists
+and is unique with the property that it renders the diagram commutative.
+\end{frame}
+
+\begin{frame}
+\frametitle{Change of base of extension algebras}
+\begin{theorem}
+The change of base of an extension algebra is again an extension algebra
+\end{theorem}
+
+\begin{theorem}
+Let $\stesys$ be a pre-extension algebra. Then
+\begin{equation*}
+\famesys{\famesys{\stesys}} \simeq \cobesys{\stesysf}{\famesys{\stesys}}{\ectxext}{\eft}.
+\end{equation*}
+\end{theorem}
+
+\begin{theorem}
+Let $\stesys$ be a pre-extension algebra and consider $p:\stesysc\rightarrow
+X\leftarrow Y:g$. Then
+\begin{equation*}
+\famesys{\cobesys{Y}{\stesys}{g}{p}}\simeq
+\cobesys{Y}{\famesys{\stesys}}{g}{p\circ\eft}.
+\end{equation*}
+\end{theorem}
+\end{frame}
+
+\begin{frame}
+\frametitle{Introduction to the theory of weakening}
+\begin{itemize}
+\item We will introduce an operation for weakening which acts at the three
+levels of fundamental structures (contexts, families and terms).
+  \pause
+\item It will preserve extension, so it will be implemented as an extension
+homomorphism.
+  \pause
+\item It will preserve itself, so weakening will also be a weakening
+homomorphism.
+  \pause
+\item There will be a notion of `Currying for weakening', which explains
+what happens when something is weakened by an extended object.
+\end{itemize}
 \end{frame}
@@
 \begin{frame}
-\frametitle{The theory of weakening (continued)}
+\frametitle{Pre-weakening}
+Let $\stesys$ be an extension algebra in $\cat{C}$. A pre-weakening operation
+on $\stesys$ is an extension homomorphism 
+$ \mathbf{w}(\stesys)
+    :
+  \cobesys{\stesysf}{\famesys{\stesys}}{\eft}{\eft}
+    \to
+  \famesys{\famesys{\stesys}}$
+for which the diagram
+\begin{equation*}
+\begin{tikzcd}[ampersand replacement=\&,column sep=large]
+\cobesys{\stesysf}{\famesys{\stesys}}{\eft}{\eft}
+  \ar{r}{\mathbf{w}(\stesys)}
+  \ar{dr}[swap]{\pullbackpr{1}{\eft}{\eft}}
+  \&
+\famesys{\famesys{\stesys}}
+  \ar{d}{\eft[2]}
+  \\
+  {}\& 
+\stesysf
+\end{tikzcd}
+\end{equation*}
+commutes.
+\end{frame}
+
+\begin{frame}
+Let $\stesys$ be an extension algebra with pre-weakening operation
+$\mathbf{w}(\stesys)$. Then $\famesys{\stesys}$ has the pre-weakening operation
+$\mathbf{w}(\famesys{\stesys})$ which is uniquely determined by rendering the
+diagram
 \begin{small}
-Weakening preserves itself:
+\begin{equation*}
+\begin{tikzcd}[ampersand replacement=\&,column sep=large]
+\cobesys{\stesysf_2}{\famesys{\famesys{\stesys}}}{\eft[2]}{\eft[2]}
+  \ar{rr}{%
+      \pullback{\beta_1}{\boldsymbol{\beta}}{\eft}{\eft}
+    }
+  \ar[bend right]{ddr}[swap]{\pullbackpr{1}{\eft[2]}{\eft[2]}}
+  \ar[dotted]{dr}{\mathbf{w}(\famesys{\stesys})}
+  \&
+  {}\&
+\cobesys{\stesysf}{\famesys{\stesys}}{\eft}{\eft}
+  \ar{r}{\mathbf{w}(\stesys)}
+  \&
+\famesys{\famesys{\stesys}}
+  \ar{d}{\boldsymbol{\beta}}
+  \\
+  {}\&
+\famesys{\famesys{\famesys{\stesys}}}
+  \ar{r}{\boldsymbol{\beta}_\mathbf{2}}
+  \ar{d}[swap]{\eft[3]}
+  \&
+\famesys{\famesys{\stesys}}
+  \ar{d}{\eft[2]}
+  \ar{r}{\boldsymbol{\beta}}
+  \&
+\famesys{\stesys}
+  \ar{d}{\eft}
+  \\
+  {}\&
+\stesysf_2
+  \ar{r}[swap]{\efamext}
+  \&
+\stesysf
+  \ar{r}[swap]{\ectxext}
+  \&
+\stesysc
+\end{tikzcd}
+\end{equation*}
+\end{small}
+commutative.
+\end{frame}
+
+\begin{frame}
+\frametitle{The theory of weakening (Currying)}
+\begin{align*}
+& \inference
+  { \jfam{\Gamma}{A}
+    \jfam{{\Gamma}{A}}{P}
+    \jfam{\Gamma}{B}
+    }
+  { \jfameq
+      {{{\Gamma}{A}}{P}}
+      {\ctxwk{\ctxext{A}{P}}{B}}
+      {\ctxwk{P}{{A}{B}}}
+    }
+\\
+& \inference
+  { \jfam{\Gamma}{A}
+    \jfam{{\Gamma}{A}}{P}
+    \jfam{{\Gamma}{B}}{Q}
+    }
+  { \jfameq
+      {{{{\Gamma}{A}}{P}}{\ctxwk{P}{{A}{B}}}}
+      {\ctxwk{\ctxext{A}{P}}{Q}}
+      {\ctxwk{P}{{A}{Q}}}
+    }
+\\
+& \inference
+  { \jfam{\Gamma}{A}
+    \jfam{{\Gamma}{A}}{P}
+    \jterm{{\Gamma}{B}}{Q}{g}
+    }
+  { \jtermeq
+      {{{{\Gamma}{A}}{P}}{\ctxwk{P}{{A}{B}}}}
+      {\ctxwk{P}{{A}{Q}}}
+      {\ctxwk{\ctxext{A}{P}}{g}}
+      {\ctxwk{P}{{A}{g}}}
+    } 
+\end{align*}
+\end{frame}
+
+\begin{frame}
+\frametitle{Pre-weakening algebras}
+A pre-weakening algebra $\stesys$ in $\cat{C}$ is an extension algebra $\stesys$ 
+with a pre-weakening operation 
+$ \mathbf{w}(\stesys)
+    :
+  \cobesys{\stesysf}{\famesys{\stesys}}{\eft}{\eft}
+    \to
+  \famesys{\famesys{\stesys}}$
+for which the diagram
+\begin{equation*}
+\begin{tikzcd}[ampersand replacement=\&,column sep=15em]
+\cobesys{\stesysf_2}{\famesys{\stesys}}{\eft\circ\eft[2]}{\eft}
+  \ar[bend right=10]{dr}[swap]%
+    { [ \pullbackpr{1}{\eft\circ\eft[2]}{\eft},%
+        \mathbf{w}(\stesys)%
+          \circ%
+        (\pullback{\efamext}{\catid{\famesys{\stesys}}}{\eft}{\eft})%
+        ]%
+      }
+  \ar{r}{
+    [ \pullbackpr{1}{\eft\circ\eft[2]}{\eft},%
+      \mathbf{w}(\stesys)%
+        \circ%
+      (\pullback{\eft[2]}{\catid{\famesys{\stesys}}}{\eft}{\eft})%
+      ]}%
+  \&
+\cobesys{\stesysf_2}{\famesys{\famesys{\stesys}}}{\eft[2]}{\eft[2]}
+  \ar{d}{\mathbf{w}(\famesys{\stesys})}
+  \\
+  {}\&
+\famesys{\famesys{\famesys{\stesys}}}
+\end{tikzcd}
+\end{equation*}
+commutes.
+\end{frame}
+
+\begin{frame}
+\frametitle{The theory of weakening (Weakening preserves itself)}
+\begin{small}
 \begin{align*}
@@
       {\ctxwk{{A}{Q}}{{A}{k}}}
-    }
-\end{align*}
-\pause
-Currying for weakening:
-\begin{align*}
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{{\Gamma}{A}}{P}
-    \jfam{\Gamma}{B}
-    }
-  { \jfameq
-      {{{\Gamma}{A}}{P}}
-      {\ctxwk{\ctxext{A}{P}}{B}}
-      {\ctxwk{P}{{A}{B}}}
-    }
-\\
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{{\Gamma}{A}}{P}
-    \jfam{{\Gamma}{B}}{Q}
-    }
-  { \jfameq
-      {{{{\Gamma}{A}}{P}}{\ctxwk{P}{{A}{B}}}}
-      {\ctxwk{\ctxext{A}{P}}{Q}}
-      {\ctxwk{P}{{A}{Q}}}
-    }
-\\
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{{\Gamma}{A}}{P}
-    \jterm{{\Gamma}{B}}{Q}{g}
-    }
-  { \jtermeq
-      {{{{\Gamma}{A}}{P}}{\ctxwk{P}{{A}{B}}}}
-      {\ctxwk{P}{{A}{Q}}}
-      {\ctxwk{\ctxext{A}{P}}{g}}
-      {\ctxwk{P}{{A}{g}}}
     }
@@
 \end{align*}
+\end{frame}
+
+\begin{frame}
+\frametitle{The empty context and family}
+\begin{small}
+Introduction rules for the empty context and family:
+\begin{align*}
+& \inference
+  { }
+  { \jctx{\emptyc}
+    }
+& & \inference
+  { \jctx{\Gamma}
+    }
+  { \jfam{\Gamma}{\emptyf[\Gamma]}
+    }
+& & \inference
+  { \jctxeq{\Gamma}{\Gamma'}
+    }
+  { \jfameq{\Gamma}{\emptyf[\Gamma]}{\emptyf[\Gamma']}
+    }
+\end{align*}
+\pause
+Compatibility rules for families and contexts:
+\begin{align*}
+& \inference
+  { \jctx{\Gamma}
+    }
+  { \jfam{\emptyc}{\Gamma}
+    } 
+& & \inference
+    { \jfam{\emptyc}{\Gamma}
+      }
+    { \jctx{\Gamma}
+      }
+    \\
+& \inference
+  { \jctxeq{\Gamma}{\Delta}
+    }
+  { \jfameq{\emptyc}{\Gamma}{\Delta}
+    }
+& & \inference
+    { \jfameq{\emptyc}{\Gamma}{\Delta}
+      }
+    { \jctxeq{\Gamma}{\Delta}
+      }
+\end{align*}
+\pause
+Compatibility rules for extension and the empty families:
+\begin{align*}
+& \inference
+  { \jctx{\Gamma}
+    }
+  { \jctxeq{\ctxext{\emptyc}{\Gamma}}{\Gamma}
+    }
+& & \inference
+  { \jctx{\Gamma}
+    }
+  { \jctxeq{\ctxext{\Gamma}{\emptyf}}{\Gamma}
+    }
+  \\
+& \inference
+  { \jfam{\Gamma}{A}
+    }
+  { \jfameq{\Gamma}{\ctxext{\emptyf}{A}}{A}
+    }
+& & \inference
+  { \jfam{\Gamma}{A}
+    }
+  { \jfameq{\Gamma}{{A}{\emptyf}}{A}
+    }
+\end{align*}
+Compatibility of context extension and family extension:
+\begin{align*}
+& \inference
+  { \jfam{\Gamma}{A}
+    }
+  { \jctxeq{(\ctxext{\Gamma}{A})^c}{(\ctxext{\Gamma}{A})^\famsym}
+    }
+\end{align*}
+\end{small}
 \end{frame}

### 2014-11-03 main.tex
reorganization and slight change in the rules
small change
moved rules for empty context and families
moved junk

--- 2014-06-18 article-models-main.tex
@@
 
-\input{preamble-articles}
+\input{Preamble/preamble-articles}
+
+\includeonly{Algebras/article-models-cat}
 
@@
 
-\begin{comment}
-\begin{abstract}
-A project is proposed where we investigate notions of internal models of type
-theory and internal higher categories. We take the point of view that an
-internal model is an internal higher category with extra structure interpreting
-the type constructors. We propose to (1) explore possible definitions of both
-notions, (2) find a zoo of examples and (3) find extensions of the underlying
-theory and models thereof to include higher inductive types.
-\end{abstract}
-\end{comment}
-
 \tableofcontents
 
-\begin{comment}
-\part{Introduction}
-The project proposal is described in \autoref{stage1} and \autoref{stage2}. In
-the appendices we give elaborations on the topics discussed in the proposal itself.
-By including these we intend to clarify the proposal and present some of the 
-(unfinished) work that is already done.
-
-\input{article-models-stage1}
-
-\input{article-models-stage2}
-
-%\input{article-models-eg}
-\end{comment}
+\input{Introduction/article-models-intro}
 
@@
 
-\input{article-models-tt}
+\include{Rules/rules}
 
-\input{article-models-ttderived}
-
-%\input{article-models-ttpretty}
+\include{Derivations/article-models-ttderived}
 
@@
 
-\input{article-models-ttinternal}
+\include{Internal_models/article-models-ttinternal}
 
-\begin{comment}
-\input{article-models-ttmodels}
+\part{Type constructors}
 
-\input{article-models-ttuniverses}
+\include{Dependent_function_types/article-models-ttdp}
 
-\part{Inductive constructions}
+\include{Universes/article-models-ttuniverses}
 
-\input{article-models-ttind}
+\include{Junk/article-models-ttind}
 
-\input{article-models-ttconstr}
-\end{comment}
+\part{Categorical semantics}
 
-%\input{article-models-idtypes}
-
-%\input{article-models-constructors}
-
-%\input{article-models-defn}
+\include{Algebras/article-models-cat}
 
@@
 %\phantomsection\addcontentsline{toc}{section}{References}
-\bibliography{refs}
+\bibliography{References/refs}
 

### 2014-11-05 Presentation/presentation.tex
nearly finished presentation

@@
 \documentclass[handout]{beamer}
+\setbeamercolor{frametitle}{fg=teal!80!black}
+\setbeamercolor{item}{fg=red!90!black}
+\setbeamercolor{title}{fg=teal!80!black}
+\setbeamercolor{block title}{fg=teal!80!black}
 
@@
 
-\title{An algebraic formulation of dependent type theory}
+\title{\bf An algebraic formulation of dependent type theory}
 \author{Egbert Rijke}
@@
 
+\newcommand\goal[1]{\begin{huge}\begin{center}\textbf{Goal: {\color{teal}#1}}\end{center}\end{huge}}
+\newcommand\important[1]{\textbf{\color{red!90!black}#1}}
+
 \begin{document}
@@
 \begin{frame}
-\frametitle{Desired properties of the algebraic theory}
+\frametitle{\bf Goal and acknowledgements}
+We will give an essentially algebraic theory of type dependency. Thus, we will
+do two things:
+\begin{itemize}
+\item Give a syntactic description of the theory in the usual style of
+name-free type theory.
+\item Define the corresponding algebraic objects in a category with finite limits.
+\end{itemize}
+\pause
+People unfamiliar with type theory should get out of this presentation 
+a sense of what dependent type theory is.
+\\[\baselineskip]
+\pause
+For many insights in the current presentation I am grateful to
+\begin{itemize}
+\item Vladimir Voevodsky
+\item Richard Garner
+\item Steve Awodey
+\end{itemize}
+\end{frame}
+
+\begin{frame}
+\frametitle{\bf Desired properties of the algebraic theory}
 \begin{itemize}
@@
 \begin{frame}
-\frametitle{Overview of the theories}
-\begin{equation*}
-\begin{tikzcd}[ampersand replacement=\&,column sep=tiny]
-{} \& \text{Extension} \ar{dr} \ar{d} \ar{dl}
-  \\
-\text{Weakening} \ar{d} \& \text{Substitution} \ar{dd} \& \text{Empty context} \ar{ddl}
-\\
-\text{Projection} \ar{dr}
-\\
-{} \& \text{Dependent Type Theory}
-\end{tikzcd}
-\end{equation*}
-\end{frame}
-
-\begin{frame}
-\frametitle{The basic judgments}
+\frametitle{\bf Overview of the theories}
+\begin{equation*}
+\begin{tikzcd}[ampersand replacement=\&,column sep=0em]
+{} \& \textbf{Extension} \ar[very thick,-stealth]{dr} \ar[very thick,-stealth]{d} \ar[very thick,-stealth]{dl}
+  \\
+\textbf{Weakening} \ar[very thick,-stealth]{d} \& \textbf{Substitution} \ar[very thick,-stealth]{dd} \& \makebox[3cm]{\textbf{Empty context and families}} \ar[very thick,-stealth]{ddl}
+\\
+\textbf{Projection} \ar[very thick,-stealth]{dr}
+\\
+{} \& \textbf{Dependent Type Theory}
+\end{tikzcd}
+\end{equation*}
+\\[2em]
+\pause
+The theories of weakening (and projections), substitution and the empty context and families are formulated as three independent theories on top of the theory of extension.
+\end{frame}
+
+\begin{frame}
+\frametitle{\bf The basic judgments}
 \begin{align*}
@@
 \end{align*}
-\end{frame}
-
-\begin{frame}
-\frametitle{Judgmental equality is an equivalence relation}
+\\[\baselineskip]
+There are rules expressing that all three kinds of judgmental equality are
+convertible equivalence relations.
+
+\pause
+An example of a conversion rule:
+\begin{align*}
+& \inference
+  { \jctxeq{\Gamma}{\Delta}
+    \jfam{\Gamma}{A}
+    }
+  { \jfam{\Delta}{A}
+    }
+& & \inference
+    { \jctxeq{\Gamma}{\Delta}
+      \jfameq{\Gamma}{A}{B}
+      }
+    { \jfameq{\Delta}{A}{B}
+      }
+\end{align*}
+\end{frame}
+
+\begin{comment}
+\begin{frame}
+\frametitle{\bf Rules: judgmental equality is an equivalence relation}
 \begin{footnotesize}
@@
 \end{frame}
-
-\begin{frame}
-\frametitle{The fundamental structure of E-systems}
-The fundamental structure of an E-system $\stesys$ in a category
-$\cat{C}$ with finite limits consists of
+\end{comment}
+
+\begin{frame}
+\frametitle{\bf The fundamental structure of E-systems}
+The \important{fundamental structure of an E-system $\stesys$} in a category with finite limits consists of
 \begin{equation*}
@@
 \end{equation*}
-The inferences of judgmental equality are automatically taken
-care of by the equality the ambient category
-\end{frame}
-
-\begin{frame}
-\frametitle{Extension}
-Introduction rules for extension:
+\pause
+\begin{itemize}
+\item The object $\stesysc$ represents the sort of contexts.
+\pause
+\item The object $\stesysf$ represents the sort of families.
+\pause
+\item Every family has a context, this is represented by $\eft:\stesysf\to\stesysc$.
+\pause
+\item The object $\stesyst$ represents the sort of terms.
+\pause
+\item Every term has a family associated to it, this is represented by
+$\partial:\stesyst\to\stesysf$.
+\end{itemize}
+\end{frame}
+
+\begin{frame}
+\frametitle{\bf Rules for extension}
+Introduction rules for \important{context extension}:
 \begin{align*}
@@
       }
-    \\
+\pause
+\intertext{%
+Introduction rules for \important{family extension}:}
 & \inference
@@
 \pause
-Extension is associative:
+\important{Extension is associative}:
 \begin{align*}
@@
 \begin{frame}
-\frametitle{Pre-extension algebras}
-A \emph{pre-extension object $\stesys$ in $\cat{C}$} consists of a \emph{fundamental structure}
-in $\cat{C}$ together with the \emph{context extension} and \emph{family extension} operations
+\frametitle{\bf Current goal}
+\goal{Define extension algebras}
+\end{frame}
+
+\begin{frame}
+\frametitle{\bf Pre-extension algebras}
+A \important{pre-extension algebra $\stesys$} in a category with finite limits consists
+\begin{itemize}
+\item a fundamental structure $\stesys$, and
+\item context extension and family extension operations
+\end{itemize}
 \begin{align*}
@@
 \end{align*}
-respectively, such that the diagram
+implementing the introduction rules
+\begin{align*}
+& \inference
+  { \jfam{\Gamma}{A}
+    }
+  { \jctx{\ctxext{\Gamma}{A}}
+    }
+& & \inference
+  { \jfam{{\Gamma}{A}}{P}
+    }
+  { \jfam{\Gamma}{\ctxext{A}{P}}
+    }
+\end{align*}
+Thus, we will additionally require that we have a commuting diagram
 \begin{equation*}
@@
 \end{equation*}
-commutes.
-\end{frame}
-
-\begin{frame}
-\frametitle{Notation for pre-extension algebras}
+\end{frame}
+
+\begin{frame}
+\frametitle{\bf Notation for pre-extension algebras}
 We introduce the following notation:
@@
 \begin{frame}
-\frametitle{Notation for extension algebras (continued)}
-Now we may continue to define
-\begin{align*}
-\stesysf_4 
-  & := 
-\pullback{\stesysf_3}{\stesysf_3}{\eext{2}}{\eft[3]}
-  \\
-\eft[4] 
-  & := 
-\pullbackpr{1}{\eext{2}}{\eft[3]}.
-\end{align*}
-Then we see that the outer square of the diagram
-\begin{equation*}
-\begin{tikzcd}[ampersand replacement=\&,column sep=large]
-\stesysf_4
-  \ar[dotted]{dr}{\eext{3}}
-  \ar{rr}{\pullback{\pullbackpr{2}{\efamext}{\eft[2]}}{\pullbackpr{2}{\efamext}{\eft[2]}}{\efamext}{\eft[2]}}
-  \ar{dd}[swap]{\eft[4]}
-  \& 
-  \&
-\stesysf_3
-  \ar{d}{\eext{2}}
-  \\
-  \&
-\stesysf_3
-  \ar{d}[swap]{\eft[3]}
-  \ar{r}{\pullbackpr{2}{\efamext}{\eft[2]}}
-  \&
-\stesysf_2
-  \ar{d}{\eft[2]}
-  \\
-\stesysf_3
-  \ar{r}[swap]{\eft[3]}
-  \&
-\stesysf_2
-  \ar{r}[swap]{\efamext}
-  \&
-\stesysf
-\end{tikzcd}
-\end{equation*}
-commutes,
-so we may define $\eext{3}$ to be the unique map which renders the diagram commutative.
-\end{frame}
-
-\begin{frame}
-\frametitle{Extension algebras}
-An extension object is a pre-extension object $\stesys$ for which 
+\frametitle{\bf Extension algebras}
+An \important{extension algebra} is a pre-extension algebra $\stesys$ for which 
 the diagrams
@@
 commute.
-\end{frame}
-
-\begin{frame}
-\frametitle{Local pre-extension algebras}
+\\[\baselineskip]
+\begin{itemize}
+\item These diagrams implement associativity of extension.
+\end{itemize}
+\end{frame}
+
+\begin{frame}
+\frametitle{\bf Current goal}
+\goal{Develop properties of extension algebras}
+
+We need to know:
+\begin{itemize}
+\item What homomorphisms of (pre-)extension algebras are
+\item That 
+\end{itemize}
+\end{frame}
+
+\begin{frame}
+\frametitle{\bf Slicing of pre-extension algebras}
 Suppose that $\stesys$ is a pre-extension algebra. Then we define
@@
 \begin{frame}
-\frametitle{Local extension algebras}
+\frametitle{\bf Local extension algebras}
 \begin{theorem}
@@
 \begin{frame}
-\frametitle{Pre-extension homomorphisms}
+\frametitle{\bf Pre-extension homomorphisms}
 \begin{footnotesize}
@@
 \begin{definition}
-A pre-extension homomorphism between extension objects is called an extension
+A pre-extension homomorphism between extension algebras is called an extension
 homomorphism.
@@
 \begin{frame}
-\frametitle{Mixed commutative diagrams}
+\frametitle{\bf Mixed commutative diagrams}
 Suppose $f:\stesys\to\stesys'$ is a pre-extension homomorphism. We say that
@@
 \begin{frame}
-\frametitle{Change of base of pre-extension algebras}
+\frametitle{\bf Change of base of pre-extension algebras}
 Let $\stesys$ be a pre-extension algebra and consider $p:\stesysc\rightarrow X\leftarrow Y:g$.
@@
 \begin{frame}
-\frametitle{Change of base of extension algebras}
+\frametitle{\bf Change of base of extension algebras}
 \begin{theorem}
@@
 \begin{frame}
-\frametitle{Introduction to the theory of weakening}
+\frametitle{\bf Introduction to the theory of weakening}
 \begin{itemize}
@@
 \begin{frame}
-\frametitle{The theory of weakening}
+\frametitle{\bf Rules for the theory of weakening}
 \begin{footnotesize}
@@
     }
-  { \jfam{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk[\famsym]{A}{Q}}
+  { \jfam{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{Q}}
     }
@@
         {{{\Gamma}{A}}{\ctxwk{A}{B}}}
-        {\ctxwk[\famsym]{A}{Q}}
-        {\ctxwk[\famsym]{A'}{Q'}}
+        {\ctxwk{A}{Q}}
+        {\ctxwk{A'}{Q'}}
       }
@@
     }
-  { \jterm{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk[\famsym]{A}{Q}}{\ctxwk[\tmsym]{A}{g}}
+  { \jterm{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{Q}}{\ctxwk{A}{g}}
     }
@@
         {{{\Gamma}{A}}{\ctxwk{A}{B}}}
-        {\ctxwk[\famsym]{A}{Q}}
-        {\ctxwk[\tmsym]{A}{g}}
-        {\ctxwk[\tmsym]{A'}{g'}}
+        {\ctxwk{A}{Q}}
+        {\ctxwk{A}{g}}
+        {\ctxwk{A'}{g'}}
       }
@@
 \begin{frame}
-\frametitle{Pre-weakening}
-Let $\stesys$ be an extension algebra in $\cat{C}$. A pre-weakening operation
+\frametitle{\bf Pre-weakening}
+Let $\stesys$ be an extension algebra. A pre-weakening operation
 on $\stesys$ is an extension homomorphism 
-$ \mathbf{w}(\stesys)
+\begin{equation*}
+\mathbf{w}(\stesys)
     :
@@
     \to
-  \famesys{\famesys{\stesys}}$
+  \famesys{\famesys{\stesys}}
+\end{equation*}
 for which the diagram
@@
 \begin{frame}
+\frametitle{\bf Weakening for the families algebra}
 Let $\stesys$ be an extension algebra with pre-weakening operation
@@
 \begin{frame}
-\frametitle{The theory of weakening (Currying)}
+\frametitle{\bf Rules for weakening: Currying}
 \begin{align*}
@@
 \begin{frame}
-\frametitle{Pre-weakening algebras}
-A pre-weakening algebra $\stesys$ in $\cat{C}$ is an extension algebra $\stesys$ 
+\frametitle{\bf Pre-weakening algebras}
+A pre-weakening algebra $\stesys$ is an extension algebra $\stesys$ 
 with a pre-weakening operation 
@@
 \begin{frame}
-\frametitle{The theory of weakening (Weakening preserves itself)}
+\frametitle{\bf Pre-weakening homomorphisms}
+A pre-weakening homomorphism between pre-weakening algebras $\stesys'$ and $\stesys$ is an
+extension homomorphism $f:\stesys'\to \stesys$ such that additionally the diagram
+\begin{equation*}
+\begin{tikzcd}[ampersand replacement=\&,column sep=large]
+\cobesys{\stesysf'}{\famesys{\stesys'}}{\eft'}{\eft'}
+  \ar{d}[swap]{\mathbf{w}(\stesys')}
+  \ar{r}{\pullback{f_1}{\famehom{f}}{\eft}{\eft}}
+  \&
+\cobesys{\stesysf}{\famesys{\stesys}}{\eft}{\eft}
+  \ar{d}{\mathbf{w}(\stesys)}
+  \\
+\famesys{\famesys{\stesys'}}
+  \ar{r}[swap]{\famehom{\famehom{f}}}
+  \&
+\famesys{\famesys{\stesys}}
+\end{tikzcd}
+\end{equation*}
+commutes.
+\end{frame}
+
+\begin{frame}
+\frametitle{\bf Change of base of pre-weakening algebras}
+Let $\stesys$ be a pre-weakening algebra and consider $p:\stesysc\rightarrow X\leftarrow Y:g$.
+Then we define
+\begin{equation*}
+\mathbf{w}(\cobesys{Y}{\stesys}{g}{p})
+  :
+\cobesys
+  { (\pullback{Y}{\stesysf}{g}{p\circ\eft})}
+  { \famesys{\cobesys{Y}{\stesys}{g}{p}}}
+  { g^\ast(\eft)}
+  { g^\ast(\eft)}
+  \to
+\famesys{\famesys{\cobesys{Y}{\stesys}{g}{p}}}
+\end{equation*}
+to be the unique extension homomorphism rendering the diagram
+\begin{scriptsize}
+\begin{equation*}
+\begin{tikzcd}[ampersand replacement=\&]
+\cobesys
+  { \pullback{Y}{\stesysf}{g}{p\circ\eft}}
+  { \famesys{\cobesys{Y}{\stesys}{g}{p}}}
+  { g^\ast(\eft)}
+  { g^\ast(\eft)}
+  \ar[bend right]{ddr}[swap]{\pullbackpr{1}{g^\ast(\eft)}{g^\ast(\eft)}}
+  \ar{rr}{%
+    \pullback
+      { \pullbackpr{2}{g}{p\circ\eft}}
+      { \boldsymbol{\pi}_\mathbf{2}(g,p\circ\eft)}
+      { \eft}
+      { \eft}
+    }
+  \ar[dotted]{dr}{\mathbf{w}(\cobesys{Y}{\stesys}{g}{p})}
+  \&
+  {}\&
+\cobesys{\stesysf}{\famesys{\stesys}}{\eft}{\eft}
+  \ar{r}{\mathbf{w}(\stesys)}
+  \&
+\famesys{\famesys{\stesys}}
+  \ar{d}{\boldsymbol{\beta}}
+  \\
+  {}\&
+\famesys{\famesys{\cobesys{Y}{\stesys}{g}{p}}}
+  \ar{r}{\beta}
+  \ar{d}{g^\ast(\eft)_1}
+  \&
+\famesys{\cobesys{Y}{\stesys}{g}{p}}
+  \ar{r}{\boldsymbol{\pi}_2(g,p\circ\eft)}
+  \ar{d}{g^\ast(\eft)}
+  \&
+\famesys{\stesys}
+  \ar{d}{\eft}
+  \\
+  {}\&
+\pullback{Y}{\stesysf}{g}{p\circ\eft}
+  \ar{r}[swap]{\cobesys{Y}{\ectxext}{g}{p}}
+  \&
+\pullback{Y}{\stesysc}{g}{p}
+  \ar{r}[swap]{\pullbackpr{2}{g}{p}}
+  \&
+\stesysc
+\end{tikzcd}
+\end{equation*}
+\end{scriptsize}
+commutative.
+\end{frame}
+
+\begin{frame}
+\frametitle{\bf Properties of pre-weakening algebras}
+\begin{theorem}
+If $\stesys$ is a pre-weakening algebra and $p:\stesysc\rightarrow X\leftarrow Y:g$,
+then $\cobesys{Y}{\stesys}{g}{p}$ is also a pre-weakening algebra.
+\end{theorem}
+\pause
+\begin{theorem}
+If $\stesys$ is a pre-weakening algebra, then so is $\famesys{\stesys}$. 
+\end{theorem}
+\end{frame}
+
+\begin{frame}
+\frametitle{\bf Weakening algebras}
+Thus, it makes sense to require that the pre-weakening operation itself is a
+pre-weakening homomorphism.
+\begin{definition}
+A \emph{weakening algebra} is a pre-weakening algebra $\stesys$ with the property that
+$\mathbf{w}(\stesys)$ is a pre-weakening morphism.
+\end{definition}
+\pause
+\begin{theorem}
+If $\stesys$ is a weakening algebra and $p:\stesysc\rightarrow X\leftarrow Y:g$,
+then $\cobesys{Y}{\stesys}{g}{p}$ is also a weakening algebra.
+\end{theorem}
+\pause
+\begin{theorem}
+If $\stesys$ is a weakening algebra, then so is $\famesys{\stesys}$. 
+\end{theorem}
+\end{frame}
+
+\begin{frame}
+\frametitle{\bf Rules for weakening: weakening preserves itself}
+The requirement that weakening is itself a pre-weakening homomorphism
+is represented by the following inference rules which we impose on
+weakening:
 \begin{small}
@@
 \\
+\pause
 & \inference
@@
 \begin{frame}
-\frametitle{The theory of projections}
-Introduction rules for identity morphisms:
+\frametitle{\bf Rules for the theory of projections}
+Introduction rules for units:
 \begin{align*}
@@
 \end{align*}
-Weakening preserves the identity morphism:
+Weakening preserves the unit:
 \begin{equation*}
@@
 \begin{frame}
-\frametitle{The theory of substitution}
+\frametitle{\bf Pre-projection algebras}
+A pre-projection algebra is a weakening algebra $\stesys$ for which there is a term
+$\mathbf{i}(\stesys):\stesysf\to \stesyst_2$ such that the diagram
+\begin{equation*}
+\begin{tikzcd}[ampersand replacement=\&,column sep=large]
+\stesysf 
+  \ar{r}{\mathbf{i}(\stesys)} 
+  \ar{d}[swap]{\Delta_{\eft}} 
+  \& 
+\stesyst_2 
+  \ar{d}{\ebd[2]}\\
+\pullback{\stesysf}{\stesysf}{\eft}{\eft} 
+  \ar{r}[swap]{w(\stesys)_0} 
+  \& 
+\stesysf_2
+\end{tikzcd}
+\end{equation*}
+commutes. In this diagram, $\Delta_{\eft}:\stesysf\to \pullback{\stesysf}{\stesysf}{\eft}{\eft}$ is the diagonal.
+\end{frame}
+
+\begin{frame}
+\frametitle{\bf Pre-projection homomorphisms}
+A pre-projection homomorphism from $\stesys$ to $\stesys'$ is a weakening homomorphism
+$f:\stesys'\to \stesys$ such that the square
+\begin{equation*}
+\begin{tikzcd}[ampersand replacement=\&,column sep=large]
+\stesyst_2'
+  \ar{r}{f^t_2}
+  \&
+\stesyst_2
+  \\
+\stesysf' 
+  \ar{r}[swap]{f_1}
+  \ar{u}{\mathbf{i}(\stesys')}
+  \&
+\stesysf
+  \ar{u}[swap]{\mathbf{i}(\stesys)}
+\end{tikzcd}
+\end{equation*}
+commutes.
+\end{frame}
+
+\begin{frame}
+\frametitle{\bf Projection algebras}
+\begin{definition}
+A projection algebra is a pre-projection algebra for which the weakening operation
+is a pre-projection homomorphism.
+\end{definition}
+\pause
+\begin{theorem}
+If $\stesys$ is a projection algebra and $p:\stesysc\rightarrow X\leftarrow Y:g$,
+then $\cobesys{Y}{\stesys}{g}{p}$ is also a projection algebra.
+\end{theorem}
+\pause
+\begin{theorem}
+If $\stesys$ is a projection algebra, then so is $\famesys{\stesys}$. 
+\end{theorem}
+\end{frame}
+
+\begin{frame}
+\frametitle{\bf Rules for the theory of substitution}
 \begin{footnotesize}
@@
     }
-  { \jfam{{\Gamma}{\subst{x}{P}}}{\subst[\famsym]{x}{Q}}
+  { \jfam{{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}
     }
@@
       }
-    { \jfameq{{\Gamma}{\subst{x}{P}}}{\subst[\famsym]{x}{Q}}{\subst[\famsym]{x'}{Q'}}
+    { \jfameq{{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}{\subst{x'}{Q'}}
       }
@@
     }
-  { \jterm{{\Gamma}{\subst{x}{P}}}{\subst[\famsym]{x}{Q}}{\subst[\tmsym]{x}{g}}
+  { \jterm{{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}{\subst{x}{g}}
     }
@@
         {{\Gamma}{\subst{x}{P}}}
-        {\subst[\famsym]{x}{Q}}
-        {\subst[\tmsym]{x}{g}}
-        {\subst[\tmsym]{x'}{g'}}
-      }
-\end{align*}
-\end{footnotesize}
-\end{frame}
-
-\begin{frame}
-\frametitle{The theory of substitution (continued)}
+        {\subst{x}{Q}}
+        {\subst{x}{g}}
+        {\subst{x'}{g'}}
+      }
+\end{align*}
+\pause
 Substitution is compatible with extension:
@@
 \end{align*}
-\pause
-Substitution is compatible with itself:
+\end{footnotesize}
+\end{frame}
+
+\begin{frame}
+\frametitle{\bf Pre-substitution algebras}
+A \emph{pre-substitution} for an extension algebra $\stesys$ is an
+extension homomorphism
+\begin{equation*}
+\mathbf{s}(\stesys):\cobesys{\stesyst}{\famesys{\famesys{\stesys}}}{\ebd}{\eft[2]}\to \famesys{\stesys}
+\end{equation*}
+for which the square
+\begin{equation*}
+\begin{tikzcd}[ampersand replacement=\&,column sep=large]
+\pullback{\stesyst}{\stesysf_2}{\ebd}{\eft[2]}
+  \ar{r}{s(\stesys)_0}
+  \ar{d}[swap]{\ebd\circ\pullbackpr{1}{\ebd}{\eft[2]}}
+  \&
+\stesysf 
+  \ar{d}{\eft}
+  \\
+\stesysf 
+  \ar{r}[swap]{\eft}
+  \&
+\stesysc
+\end{tikzcd}
+\end{equation*}
+commutes.\\
+\pause
+A \emph{pre-substitution algebra} is an extension algebra
+together with a pre-substitution.
+\end{frame}
+
+\begin{frame}
+\frametitle{\bf Pre-substitution homomorphisms}
+A \emph{pre-substitution homomorphism} is an extension homomorphism $f:\stesys'\to \stesys$
+for which the square
+\begin{equation*}
+\begin{tikzcd}[ampersand replacement=\&,column sep=huge]
+\cobesys{\stesyst'}{\famesys{\famesys{\stesys'}}}{\ebd'}{\eft[2]'}
+  \ar{r}{\pullback{f^t}{\famehom{\famehom{f}}}{\ebd}{\eft[2]}}
+  \ar{d}[swap]{\mathbf{s}(\stesys')}
+  \&
+\cobesys{\stesyst}{\famesys{\famesys{\stesys}}}{\ebd}{\eft[2]}
+  \ar{d}{\mathbf{s}(\stesys)}
+  \\
+\famesys{\stesys'}
+  \ar{r}[swap]{\famehom{f}}
+  \&
+\famesys{\stesys}
+\end{tikzcd}
+\end{equation*}
+commutes.
+\end{frame}
+
+\begin{frame}
+\frametitle{\bf The family pre-substitution algebra}
+\begin{theorem}
+If $\stesys$ is a pre-substitution algebra, then so is $\famesys{\stesys}$\pause~with
+$\mathbf{s}(\famesys{\stesys})$ defined to be the unique extension homomorphism
+rendering the diagram
+\begin{equation*}
+\begin{tikzcd}[ampersand replacement=\&]
+\cobesys{\stesyst_2}{\famesys{\famesys{\famesys{\stesys}}}}{\ebd[2]}{\eft[3]}
+  \ar[dotted]{dr}{\mathbf{s}(\famesys{\stesys})}
+  \ar{rr}{\pullback{\pullbackpr{2}{\ectxext}{\eft\circ\ebd}}{\boldsymbol{\beta}_\mathbf{2}}{\ebd}{\eft[2]}}
+  \ar{dd}[swap]{\ebd[2]\circ\pullbackpr{1}{\ebd[2]}{\eft[3]}}
+  \&
+  {}\&
+\cobesys{\stesyst}{\famesys{\famesys{\stesys}}}{\ebd}{\eft[2]}
+  \ar{d}{\mathbf{s}(\stesys)}
+  \\
+  {}\&
+\famesys{\famesys{\stesys}}
+  \ar{r}{\boldsymbol{\beta}}
+  \ar{d}{\eft[2]}
+  \&
+\famesys{\stesys}
+  \ar{d}{\eft}
+  \\
+\stesysf_2
+  \ar{r}[swap]{\eft[2]}
+  \&
+\stesysf
+  \ar{r}[swap]{\ectxext}
+  \&
+\stesysc
+\end{tikzcd}
+\end{equation*}
+commutative.
+\end{theorem}
+\end{frame}
+
+\begin{frame}
+\frametitle{\bf Change of base of pre-substitution algebras}
+\begin{theorem}
+Let $\stesys$ be a pre-substitution algebra and consider
+$p:\stesysc\rightarrow X\leftarrow Y:g$. Then
+$\cobesys{Y}{\stesys}{g}{p}$ is a pre-substitution algebra
+\pause~with $\mathbf{s}(\cobesys{Y}{\stesys}{g}{p})$ defined
+to be the unique extension homomorphism rendering the diagram
+\begin{footnotesize}
+\begin{equation*}
+\begin{tikzcd}[ampersand replacement=\&]
+\cobesys
+  {\pullback{Y}{\stesyst}{g}{p\circ\eft\circ\ebd}}
+  {\famesys{\famesys{\cobesys{Y}{\stesys}{g}{p}}}}
+  {g^\ast(\ebd)}{g^\ast(\eft[2])}
+  \ar[dotted]{dr}{\mathbf{s}(\cobesys{Y}{\stesys}{g}{p})}
+  \ar{rr}
+  \ar[bend right]{ddr}[swap]%
+    { \pullbackpr{1}{g}{p\circ\eft\circ\ebd}%
+        \circ%
+      \pullbackpr{1}{g^\ast(\ebd)}{g^\ast(\eft[2])}%
+      }
+  \&
+  {}\&
+\cobesys{\stesyst}{\famesys{\famesys{\stesys}}}{\ebd}{\eft[2]}
+  \ar{d}{\mathbf{s}(\stesys)}
+  \\
+  {}\&
+\famesys{\cobesys{Y}{\stesys}{g}{p}}
+  \ar{r}
+  \ar{d}{\pullbackpr{1}{g}{p\circ\eft}}
+  \&
+\famesys{\stesys}
+  \ar{d}{p\circ\eft}
+  \\
+  {}\&
+Y
+  \ar{r}[swap]{g}
+  \&
+X
+\end{tikzcd}
+\end{equation*}
+\end{footnotesize}
+commutative. 
+\end{theorem}
+\end{frame}
+
+\begin{frame}
+\frametitle{\bf Substitution algebras}
+Thus, it makes sense to require that the pre-substitution operation itself is a
+pre-substitution homomorphism.
+\begin{definition}
+A \emph{substitution algebra} is a pre-substitution algebra $\stesys$ with the property that
+$\mathbf{S}(\stesys)$ is a pre-substitution morphism.
+\end{definition}
+\pause
+\begin{theorem}
+If $\stesys$ is a substitution algebra and $p:\stesysc\rightarrow X\leftarrow Y:g$,
+then $\cobesys{Y}{\stesys}{g}{p}$ is also a substitution algebra.
+\end{theorem}
+\pause
+\begin{theorem}
+If $\stesys$ is a substitution algebra, then so is $\famesys{\stesys}$. 
+\end{theorem}
+\end{frame}
+
+\begin{frame}
+\frametitle{\bf Rules for substitution: substitution is compatible with itself}
 \begin{align*}
@@
 \begin{frame}
-\frametitle{The empty context and family}
-\begin{small}
-Introduction rules for the empty context and family:
-\begin{align*}
-& \inference
-  { }
-  { \jctx{\emptyc}
-    }
-& & \inference
-  { \jctx{\Gamma}
-    }
-  { \jfam{\Gamma}{\emptyf[\Gamma]}
-    }
-& & \inference
-  { \jctxeq{\Gamma}{\Gamma'}
-    }
-  { \jfameq{\Gamma}{\emptyf[\Gamma]}{\emptyf[\Gamma']}
-    }
-\end{align*}
-\pause
-Compatibility rules for families and contexts:
-\begin{align*}
-& \inference
-  { \jctx{\Gamma}
-    }
-  { \jfam{\emptyc}{\Gamma}
-    } 
-& & \inference
-    { \jfam{\emptyc}{\Gamma}
-      }
-    { \jctx{\Gamma}
-      }
-    \\
-& \inference
-  { \jctxeq{\Gamma}{\Delta}
-    }
-  { \jfameq{\emptyc}{\Gamma}{\Delta}
-    }
-& & \inference
-    { \jfameq{\emptyc}{\Gamma}{\Delta}
-      }
-    { \jctxeq{\Gamma}{\Delta}
-      }
-\end{align*}
-\pause
-Compatibility rules for extension and the empty families:
-\begin{align*}
-& \inference
-  { \jctx{\Gamma}
-    }
-  { \jctxeq{\ctxext{\emptyc}{\Gamma}}{\Gamma}
-    }
-& & \inference
-  { \jctx{\Gamma}
-    }
-  { \jctxeq{\ctxext{\Gamma}{\emptyf}}{\Gamma}
-    }
-  \\
-& \inference
-  { \jfam{\Gamma}{A}
-    }
-  { \jfameq{\Gamma}{\ctxext{\emptyf}{A}}{A}
-    }
-& & \inference
-  { \jfam{\Gamma}{A}
-    }
-  { \jfameq{\Gamma}{{A}{\emptyf}}{A}
-    }
-\end{align*}
-Compatibility of context extension and family extension:
-\begin{align*}
-& \inference
-  { \jfam{\Gamma}{A}
-    }
-  { \jctxeq{(\ctxext{\Gamma}{A})^c}{(\ctxext{\Gamma}{A})^\famsym}
-    }
-\end{align*}
-\end{small}
-\end{frame}
-
-\begin{frame}
-\frametitle{Joining the theories of projections, substitution and the empty context}
-To join the three theories we have formulated on top of the theory of extension,
+\frametitle{\bf Joining the theories of projections and substitution}
+To join the two theories we have formulated on top of the theory of extension,
 we need to provide rules describing:
 \begin{itemize}
-\item what happens when something is weakened by the empty family;\pause
-\item that weakening preserves the empty context and family;\pause
-\item that substitution preserves the empty context and family;\pause
-\item that weakening preserves substitution;\pause
-\item that substitution preserves weakening;\pause
+\item that weakening preserves substitution (weakening is a substitution homomorphism);\pause
+\item that substitution preserves weakening (substitution is a weakening homomorphism);\pause
+\item that substitution preserves units (substitution is a projection homomorphism);\pause
 \item that weakenings are constant families;\pause
-\item that substitution preserves the identity morphisms;\pause
-\item that identity morphisms act as identity functions;\pause
-\item that pullback along the identity morphism has no effect.
+\item that weakening by a family $A$ followed by substitution of the unit of $A$ has no effect; this is precomposing with the unit.
 \end{itemize}
@@
 \begin{frame}
-\frametitle{Compatibility rules}
-Weakening by the empty family:
-\begin{align*}
-& \inference
-  { \jfam{\Gamma}{B}
-    }
-  { \jfameq{\Gamma}{\ctxwk{\emptyf}{B}}{B}
-    }
-  \\
-& \inference
-  { \jfam{{\Gamma}{B}}{Q}
-    }
-  { \jfameq{{\Gamma}{B}}{\ctxwk{\emptyf}{Q}}{Q}
-    }
-  \\
-& \inference
-  { \jterm{{\Gamma}{B}}{Q}{g}
-    }
-  { \jtermeq{{\Gamma}{B}}{Q}{\ctxwk{\emptyf}{g}}{g}
-    }
-\end{align*}
-\pause
-Weakening of the empty family:
-\begin{align*}
-& \inference
-  { \jfam{\Gamma}{A}
-    }
-  { \jfameq{{\Gamma}{A}}{\ctxwk{A}{\emptyf}}{\emptyf}
-    }
-  \\
+\frametitle{\bf Pre-E-systems}
+A \emph{pre-E-system} is an extension algebra $\stesys$ with a
+weakening operation $\mathbf{w}(\stesys)$, units
+$\mathbf{i}(\stesys)$ and a substitution operation
+$\mathbf{s}(\stesys)$ giving it both the structure of
+a projection algebra and a substitution algebra, such
+that in addition
+\begin{itemize}
+\item Weakening is a substitution homomorphism;
+\item Substitution is a projection homomorphism.
+\end{itemize}
+Pre-E-homomorphisms are extension homomorphisms which
+are both projection homomorphisms and substitution
+homomorphisms.
+\end{frame}
+
+\begin{frame}
+\frametitle{\bf Rules: Weakened families are constant families}
+\begin{align*}
 & \inference
@@
     \jfam{\Gamma}{B}
+    \jterm{\Gamma}{A}{x}
     }
   { \jfameq
-    {{{\Gamma}{A}}{\ctxwk{A}{B}}}
-    {\ctxwk[\famsym]{A}{\emptyf}}
-    {\emptyf}
-    }
-\end{align*}
-\pause
-Compatibility of weakening of a family with weakening of a context:
-\begin{equation*}
-\inference
-{ \jfam{\Gamma}{A}
-  \jfam{\Gamma}{B}
-  }
-{ \jfameq{{\Gamma}{A}}{\ctxwk[\famsym]{A}{B}}{\ctxwk{A}{B}}
-  }
-\end{equation*}
-\end{frame}
-
-\begin{frame}
-\frametitle{Compatibility rules (continued)}
-Substitution of an empty family
-\begin{align*}
+      {\Gamma}
+      {\subst{x}{\ctxwk{A}{B}}}
+      {B}
+    }
+  \\
+& \inference
+  { \jfam{\Gamma}{A}
+    \jfam{{\Gamma}{B}}{Q}
+    \jterm{\Gamma}{A}{x}
+    }
+  { \jfameq{{\Gamma}{B}}{\subst{x}{\ctxwk{A}{Q}}}{Q}
+    }
+  \\
 & \inference
   { \jterm{\Gamma}{A}{x}
-    }
-  { \jfameq{\Gamma}{\subst{x}{\emptyf}}{\emptyf}
-    }
-  \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jfam{{\Gamma}{A}}{P}
-    }
-  { \jfameq
-      {{\Gamma}{\subst{x}{P}}}
-      {\subst{x}{\emptyf}}
-      {\emptyf}
-    }
-\end{align*}
-Compatibility of substitution of a family with substitution of a context:
-\begin{equation*}
-\inference
+    \jterm{{\Gamma}{B}}{Q}{g}
+    }
+  { \jtermeq{{\Gamma}{B}}{Q}{\subst{x}{\ctxwk{A}{g}}}{g}
+    }
+\end{align*}
+\end{frame}
+
+\begin{frame}
+\frametitle{\bf Weakened families are constant families}
+In a pre-E-system, we say that \emph{weakened families are 
+constant families} if the diagram
+\begin{equation*}
+\begin{tikzcd}[ampersand replacement=\&,column sep=12em]
+\cobesys{\stesyst}{\famesys{\stesys}}{\eft\circ\ebd}{\eft}
+  \ar{r}{ [ \catid{\stesyst},
+            \mathbf{w}(\stesys)
+              \circ
+            (\pullback{\ebd}{\catid{\famesys{\stesys}}}{\eft}{\eft})
+            ]
+          }
+  \ar{dr}[swap]{\pullbackpr{2}{\eft\circ\ebd}{\eft}}
+  \&
+\cobesys
+  { \stesyst}
+  { \famesys{\famesys{\stesys}}
+    }
+  { \ebd}
+  { \eft[2]}
+  \ar{d}{\mathbf{s}(\stesys)}
+  \\
+  {}\&
+\famesys{\stesys}
+\end{tikzcd}
+\end{equation*}
+commutes.
+\end{frame}
+
+\begin{frame}
+\frametitle{\bf Rules: precomposition with a unit has no effect}
+\begin{align*}
+& \inference
   { \jfam{{\Gamma}{A}}{P}
     }
-{ \jfameq{\Gamma}{\subst[\famsym]{x}{P}}{\subst{x}{P}}
-  }
-\end{equation*}
-\end{frame}
-
-\begin{frame}
-\frametitle{Compatibility rules (continued)}
+  { \jfameq{{\Gamma}{A}}{\subst{\idtm{A}}{\ctxwk{A}{P}}}{P}
+    }
+  \\
+& \inference
+  { \jfam{{{\Gamma}{A}}{P}}{Q}
+    }
+  { \jfameq{{{\Gamma}{A}}{P}}{\subst{\idtm{A}}{\ctxwk{A}{Q}}}{Q}
+    }
+  \\
+& \inference
+  { \jterm{{{\Gamma}{A}}{P}}{Q}{g}
+    }
+  { \jtermeq
+      {{{\Gamma}{A}}{P}}
+      {Q}
+      {\subst{\idtm{A}}{\ctxwk{A}{g}}}
+      {g}
+    }
+\end{align*}
+\end{frame}
+
+\begin{frame}
+\frametitle{\bf Invariance with respect to precomposition with units}
+In a pre-E-system $\stesys$ we say that \emph{everything is invariant
+with respect to precomposition with units} if the diagram
+\begin{equation*}
+\begin{tikzcd}[ampersand replacement=\&,column sep=large]
+\famesys{\famesys{\stesys}}
+  \ar{r}
+  \ar[equals]{ddrr}
+  \&
+\famesys{\cobesys{\stesysf}{\famesys{\stesys}}{\eft}{\eft}}
+  \ar{r}[yshift=1ex]{[\pullbackpr{1}{\eft}{\eft}\circ\eft^\ast(\eft[2]),\famehom{\mathbf{w}(\stesys)}]}
+  \&
+\cobesys
+  {\stesysf}
+  {\famesys{\famesys{\famesys{\stesys}}}}
+  {w(\stesys)_0\circ\Delta_{\eft}}
+  {\eft[3]}
+  \ar{d}{\pullback{\mathbf{i}(\stesys)}{\catid{\famesys{\famesys{\stesys}}}]}{\ebd[2]}{\eft[3]}}
+  \\
+  {}\&
+  {}\&
+\cobesys{\stesyst_2}{\famesys{\famesys{\famesys{\stesys}}}}{\ebd[2]}{\eft[3]}
+  \ar{d}{\mathbf{s}(\famesys{\stesys})}
+  \\
+  {}\&
+  {}\&
+\famesys{\famesys{\stesys}}
+\end{tikzcd}
+\end{equation*}
+commutes.
+\end{frame}
+
+\begin{frame}
+\frametitle{\bf E-systems}
+An \emph{E-system} is an extension algebra with
+\begin{itemize}
+\item a weakening operation and units, giving it the structure of a projection algebra;
+\item a substitution operation, giving it the structure of a substitution algebra
+\item (an empty context and empty families);
+\end{itemize}
+such that
+\begin{itemize}
+\item Weakening is a substitution homomorphism;
+\item Substitution is a projection homomorphism;
+\item Weakened families are constant families;
+\item Precomposition with identity functions leaves everything invariant.
+\item (The requirements regarding the empty context and families according to their rules)
+\end{itemize}
+\end{frame}
+
+\begin{frame}
+\frametitle{\bf The end}
+\begin{center}
+Thank you for your attention
+\end{center}
+\end{frame}
+
+\begin{frame}
+\frametitle{\bf Appendix: weakening and substitution preserve each other}
 \begin{footnotesize}
@@
 \begin{frame}
-\frametitle{Compatibility rules (continued)}
-Weakened families are constant:
-\begin{align*}
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{\Gamma}{B}
-    \jterm{\Gamma}{A}{x}
-    }
-  { \jfameq
-      {\Gamma}
-      {\subst{x}{\ctxwk{A}{B}}}
-      {B}
-    }
-  \\
-& \inference
-  { \jfam{\Gamma}{A}
-    \jfam{{\Gamma}{B}}{Q}
-    \jterm{\Gamma}{A}{x}
-    }
-  { \jfameq{{\Gamma}{B}}{\subst{x}{\ctxwk{A}{Q}}}{Q}
-    }
-  \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
-    \jterm{{\Gamma}{B}}{Q}{g}
-    }
-  { \jtermeq{{\Gamma}{B}}{Q}{\subst{x}{\ctxwk{A}{g}}}{g}
-    }
-\end{align*}
-\pause
-Substitution preserves the identity morphisms
+\frametitle{\bf Appendix: substitution preserves units}
 \begin{equation*}
@@
 \end{equation*}
-\pause
-Identity morphisms act as identity functions:
+\end{frame}
+
+\begin{frame}
+\frametitle{\bf Appendix: the empty context and family}
+\begin{small}
+Introduction rules for the empty context and family:
+\begin{equation*}
+\inference
+  { }
+  { \jctx{\emptyc}
+    }
+\qquad
+\inference
+  { \jctx{\Gamma}
+    }
+  { \jfam{\Gamma}{\emptyf[\Gamma]}
+    }
+\qquad
+\inference
+  { \jctxeq{\Gamma}{\Gamma'}
+    }
+  { \jfameq{\Gamma}{\emptyf[\Gamma]}{\emptyf[\Gamma']}
+    }
+\end{equation*}
+\pause
+Every context induces a family in the empty context:
+\begin{align*}
+& \inference
+  { \jctx{\Gamma}
+    }
+  { \jfam{\emptyc}{i(\Gamma)}
+    } 
+  &
+& \inference
+  { \jctxeq{\Gamma}{\Delta}
+    }
+  { \jfameq{\emptyc}{i(\Gamma)}{i(\Delta)}
+    }
+\end{align*}
+\pause
+Compatibility rules for empty context and family with extension:
+\begin{align*}
+& \inference
+  { \jctx{\Gamma}
+    }
+  { \jctxeq{\ctxext{\emptyc}{i(\Gamma)}}{\Gamma}
+    }
+& 
+& \inference
+  { \jctx{\Gamma}
+    }
+  { \jctxeq{\ctxext{\Gamma}{\emptyf}}{\Gamma}
+    }
+  \\
+& \inference
+  { \jfam{\Gamma}{A}
+    }
+  { \jfameq{\Gamma}{\ctxext{\emptyf}{A}}{A}
+    }
+&
+& \inference
+  { \jfam{\Gamma}{A}
+    }
+  { \jfameq{\Gamma}{{A}{\emptyf}}{A}
+    }
+  \\
+& \inference
+  { \jfam{\emptyc}{A}
+    }
+  { \jfameq{\emptyc}{i(\ctxext{\emptyc}{A})}{A}
+    }
+\end{align*}
+\pause
+and finally
+\begin{align*}
+& \inference
+  { \jfam{\Gamma}{A}
+    }
+  { \jctxeq{\ctxext{\Gamma}{A}}{\ctxext{i(\Gamma)}{A}}
+    }
+& &
+\inference
+  { }
+  { \jfameq{\emptyc}{i(\emptyc)}{\emptyf[\emptyc]}}
+\end{align*}
+\end{small}
+\end{frame}
+
+\begin{frame}
+\frametitle{\bf Appendix: weakening and the empty families}
+Weakening by the empty family:
+\begin{align*}
+& \inference
+  { \jfam{\Gamma}{B}
+    }
+  { \jfameq{\Gamma}{\ctxwk{\emptyf}{B}}{B}
+    }
+  \\
+& \inference
+  { \jfam{{\Gamma}{B}}{Q}
+    }
+  { \jfameq{{\Gamma}{B}}{\ctxwk{\emptyf}{Q}}{Q}
+    }
+  \\
+& \inference
+  { \jterm{{\Gamma}{B}}{Q}{g}
+    }
+  { \jtermeq{{\Gamma}{B}}{Q}{\ctxwk{\emptyf}{g}}{g}
+    }
+\end{align*}
+\pause
+Weakening of the empty family:
+\begin{align*}
+& \inference
+  { \jfam{\Gamma}{A}
+    }
+  { \jfameq{{\Gamma}{A}}{\ctxwk{A}{\emptyf}}{\emptyf}
+    }
+  \\
+& \inference
+  { \jfam{\Gamma}{A}
+    \jfam{\Gamma}{B}
+    }
+  { \jfameq
+    {{{\Gamma}{A}}{\ctxwk{A}{B}}}
+    {\ctxwk{A}{\emptyf}}
+    {\emptyf}
+    }
+\end{align*}
+\pause
+Compatibility of weakening of a family with weakening of a context:
+\begin{equation*}
+\inference
+{ \jfam{\Gamma}{A}
+  \jfam{\Gamma}{B}
+  }
+{ \jfameq{{\Gamma}{A}}{\ctxwk{A}{B}}{\ctxwk{A}{B}}
+  }
+\end{equation*}
+\end{frame}
+
+\begin{frame}
+\frametitle{\bf Appendix: substitution of an empty family}
+\begin{align*}
+& \inference
+  { \jterm{\Gamma}{A}{x}
+    }
+  { \jfameq{\Gamma}{\subst{x}{\emptyf}}{\emptyf}
+    }
+  \\
+& \inference
+  { \jterm{\Gamma}{A}{x}
+    \jfam{{\Gamma}{A}}{P}
+    }
+  { \jfameq
+      {{\Gamma}{\subst{x}{P}}}
+      {\subst{x}{\emptyf}}
+      {\emptyf}
+    }
+\end{align*}
+Compatibility of substitution of a family with substitution of a context:
+\begin{equation*}
+\inference
+{ \jfam{{\Gamma}{A}}{P}
+  }
+{ \jfameq{\Gamma}{\subst{x}{P}}{\subst{x}{P}}
+  }
+\end{equation*}
+\end{frame}
+
+\begin{frame}
+\frametitle{\bf Appendix: units act as identity functions}
 \begin{equation*}
@@
 \end{equation*}
-\end{frame}
-
-\begin{frame}
-\frametitle{Compatibility rules (continued)}
-Pullback along identity morphisms has no effect:
-\begin{align*}
-& \inference
-  { \jfam{{\Gamma}{A}}{P}
-    }
-  { \jfameq{{\Gamma}{A}}{\subst{\idtm{A}}{\ctxwk{A}{P}}}{P}
-    }
-  \\
-& \inference
-  { \jfam{{{\Gamma}{A}}{P}}{Q}
-    }
-  { \jfameq{{{\Gamma}{A}}{P}}{\subst{\idtm{A}}{\ctxwk{A}{Q}}}{Q}
-    }
-  \\
-& \inference
-  { \jterm{{{\Gamma}{A}}{P}}{Q}{g}
-    }
-  { \jtermeq
-      {{{\Gamma}{A}}{P}}
-      {Q}
-      {\subst{\idtm{A}}{\ctxwk{A}{g}}}
-      {g}
-    }
-\end{align*}
+\\[\baselineskip]
+Note that the empty family is needed to have this rule.
 \end{frame}

### 2014-11-05 Presentation/presentation.tex
nearly finished presentation

@@
 \documentclass[handout]{beamer}
-\setbeamercolor{frametitle}{fg=teal!80!black}
+\setbeamercolor{frametitle}{fg=teal!90!black}
 \setbeamercolor{item}{fg=red!90!black}
-\setbeamercolor{title}{fg=teal!80!black}
-\setbeamercolor{block title}{fg=teal!80!black}
+\setbeamercolor{title}{fg=teal!90!black}
+\setbeamercolor{block title}{fg=teal!90!black}
 
@@
 
-\newcommand\goal[1]{\begin{huge}\begin{center}\textbf{Goal: {\color{teal}#1}}\end{center}\end{huge}}
+\newcommand\subgoal[1]{\begin{huge}\begin{center}\textbf{Subgoal: {\color{teal!90!black}#1}}\end{center}\end{huge}}
 \newcommand\important[1]{\textbf{\color{red!90!black}#1}}
@@
   \\
-\jalign\jterm{\Gamma}{A}{x} 
-& \jalign\jtermeq{\Gamma}{A}{x}{y}.
+\jalign\jterm{\Gamma}{A}{a} 
+& \jalign\jtermeq{\Gamma}{A}{a}{y}.
 \end{align*}
@@
 & \inference
-  { \jterm{\Gamma}{A}{x}
-    }
-  { \jtermeq{\Gamma}{A}{x}{x}
+  { \jterm{\Gamma}{A}{a}
+    }
+  { \jtermeq{\Gamma}{A}{a}{a}
     }
 & & \inference
-    { \jtermeq{\Gamma}{A}{x}{y}
-      }
-    { \jtermeq{\Gamma}{A}{y}{x}
+    { \jtermeq{\Gamma}{A}{a}{y}
+      }
+    { \jtermeq{\Gamma}{A}{y}{a}
       }
 & & \inference
-    { \jtermeq{\Gamma}{A}{x}{y}
+    { \jtermeq{\Gamma}{A}{a}{y}
       \jtermeq{\Gamma}{A}{y}{z}
       }
-    { \jtermeq{\Gamma}{A}{x}{z}
+    { \jtermeq{\Gamma}{A}{a}{z}
       }
@@
   { \jctxeq{\Gamma}{\Delta}
-    \jterm{\Gamma}{A}{x}
-    }
-  { \jterm{\Delta}{A}{x}
+    \jterm{\Gamma}{A}{a}
+    }
+  { \jterm{\Delta}{A}{a}
     }
@@
     { \jctxeq{\Gamma}{\Delta}
-      \jtermeq{\Gamma}{A}{x}{y}
-      }
-    { \jtermeq{\Delta}{A}{x}{y}
+      \jtermeq{\Gamma}{A}{a}{y}
+      }
+    { \jtermeq{\Delta}{A}{a}{y}
       }
@@
   { \jfameq{\Gamma}{A}{B}
-    \jterm{\Gamma}{A}{x}
-    }
-  { \jterm{\Gamma}{B}{x}
+    \jterm{\Gamma}{A}{a}
+    }
+  { \jterm{\Gamma}{B}{a}
     }
@@
     { \jfameq{\Gamma}{A}{B}
-      \jtermeq{\Gamma}{A}{x}{y}
-      }
-    { \jtermeq{\Gamma}{B}{x}{y}
+      \jtermeq{\Gamma}{A}{a}{y}
+      }
+    { \jtermeq{\Gamma}{B}{a}{y}
       }
@@
 \end{align*}
-\end{frame}
-
-\begin{frame}
-\frametitle{\bf Current goal}
-\goal{Define extension algebras}
 \end{frame}
@@
   & := \stesysff
+  \pause
   \\
@@
   & := \pullbackpr{1}{\ectxext}{\eft} : \stesysf_2\to\stesysf
+  \pause
   \\
 \stesysf_3 & := \pullback{\stesysf_2}{\stesysf_2}{\efamext}{\eft[2]}
+  \pause
   \\
@@
 \end{align*}
+\pause
 Then it follows that the outer square in the diagram
@@
 \end{equation*}
-commutes. We define $\eext{2}$ to be the unique morphism rendering the above diagram
+commutes. \pause
+We define \important{$\eext{2}$} to be the unique morphism rendering the above diagram
 commutative. 
@@
 \begin{frame}
-\frametitle{\bf Current goal}
-\goal{Develop properties of extension algebras}
-
+\frametitle{Subgoal: develop properties of extension algebras}
 We need to know:
 \begin{itemize}
-\item What homomorphisms of (pre-)extension algebras are
-\item That 
+\item What homomorphisms of (pre-)extension algebras are.
+\pause
+\item That each extension algebra gives locally an extension algebra (stable under slicing).
+\pause
+\item That the change of base of an extension is an extension algebra. Change of base allows
+for \important{`parametrized extension homomorphisms'} of which weakening and substitution
+are going to be examples.
 \end{itemize}
+\end{frame}
+
+\begin{frame}
+\frametitle{\bf Pre-extension homomorphisms}
+\begin{footnotesize}
+Let $\stesys$ and $\stesys'$ be pre-extension algebras. A \important{pre-extension 
+homomorphism $f$ from $\stesys'$ to $\stesys$} is a triple $(f_0,f_1,f^t)$ 
+consisting of morphisms
+\begin{equation*}
+\begin{tikzcd}[ampersand replacement=\&]
+\stesyst' 
+  \ar{r}{f^t}
+  \ar{d}[swap]{\ebd'}
+  \&
+\stesyst
+  \ar{d}{\ebd}
+  \\
+\stesysf'
+  \ar{r}{f_1}
+  \ar{d}[swap]{\eft'}
+  \&
+\stesysf
+  \ar{d}{\eft}
+  \\
+\stesysc' 
+  \ar{r}[swap]{f_0}
+  \&
+\stesysc
+\end{tikzcd}
+\end{equation*}
+such that the indicated squares commute, for which furthermore the squares
+\begin{equation*}
+\begin{tikzcd}[ampersand replacement=\&]
+\stesysf' 
+  \ar{r}{f_1}
+  \ar{d}[swap]{\ectxext'}
+  \&
+\stesysf
+  \ar{d}{\ectxext}
+  \\
+\stesysc'
+  \ar{r}[swap]{f_0}
+  \&
+\stesysc
+\end{tikzcd}
+\quad
+\text{and}
+\quad
+\begin{tikzcd}[ampersand replacement=\&,column sep=huge]
+\stesysf_2'
+  \ar{r}{f_1\times_{\ectxext,\eft} f_1}
+  \ar{d}[swap]{\efamext'}
+  \&
+\stesysf_2
+  \ar{d}{\efamext}
+  \\
+\stesysf'
+  \ar{r}[swap]{f_1}
+  \&
+\stesysf
+\end{tikzcd}
+\end{equation*}
+commute.
+\end{footnotesize}
+\end{frame}
+
+\begin{frame}
+\frametitle{Extension homomorphisms}
+\begin{definition}
+A pre-extension homomorphism between extension algebras is called an \important{extension
+homomorphism}.
+\end{definition}
 \end{frame}
@@
 Suppose that $\stesys$ is a pre-extension algebra. Then we define
-the pre-extension algebra $\famesys{\stesys}$ to consist of the
+the pre-extension algebra \important{$\famesys{\stesys}$} to consist of the
 fundamental structure
@@
 \end{theorem}
-\end{frame}
-
-\begin{frame}
-\frametitle{\bf Pre-extension homomorphisms}
-\begin{footnotesize}
-Let $\stesys$ and $\stesys'$ be pre-extension algebras. A \emph{pre-extension 
-homomorphism $f$ from $\stesys'$ to $\stesys$} is a triple $(f_0,f_1,f^t)$ 
-consisting of morphisms
-\begin{equation*}
-\begin{tikzcd}[ampersand replacement=\&]
-\stesyst' 
-  \ar{r}{f^t}
-  \ar{d}[swap]{\ebd'}
-  \&
-\stesyst
-  \ar{d}{\ebd}
-  \\
-\stesysf'
-  \ar{r}{f_1}
-  \ar{d}[swap]{\eft'}
-  \&
-\stesysf
-  \ar{d}{\eft}
-  \\
-\stesysc' 
-  \ar{r}[swap]{f_0}
-  \&
-\stesysc
-\end{tikzcd}
-\end{equation*}
-such that the indicated squares commute, for which furthermore the squares
-\begin{equation*}
-\begin{tikzcd}[ampersand replacement=\&]
-\stesysf' 
-  \ar{r}{f_1}
-  \ar{d}[swap]{\ectxext'}
-  \&
-\stesysf
-  \ar{d}{\ectxext}
-  \\
-\stesysc'
-  \ar{r}[swap]{f_0}
-  \&
-\stesysc
-\end{tikzcd}
-\quad
-\text{and}
-\quad
-\begin{tikzcd}[ampersand replacement=\&,column sep=huge]
-\stesysf_2'
-  \ar{r}{f_1\times_{\ectxext,\eft} f_1}
-  \ar{d}[swap]{\efamext'}
-  \&
-\stesysf_2
-  \ar{d}{\efamext}
-  \\
-\stesysf'
-  \ar{r}[swap]{f_1}
-  \&
-\stesysf
-\end{tikzcd}
-\end{equation*}
-commute. Composition and the identity homomorphism are defined in the expected way.
-\end{footnotesize}
-\end{frame}
-
-\begin{frame}
-\begin{definition}
-A pre-extension homomorphism between extension algebras is called an extension
-homomorphism.
-\end{definition}
 
@@
 Let $\stesys$ be a pre-extension algebra and consider $p:\stesysc\rightarrow X\leftarrow Y:g$.
-Then there is a pre-extension algebra $\cobesys{Y}{\stesys}{g}{p}$ with projections, such that for every diagram
+Then there is a pre-extension algebra \important{$\cobesys{Y}{\stesys}{g}{p}$} with projections, such that for every diagram
 \begin{equation*}
@@
 \end{equation*}
-of which the outer square commutes, the pre-extension homomorphism $[p',f]$ exists
+of which the outer square commutes, the pre-extension homomorphism \important{$[p',f]$} exists
 and is unique with the property that it renders the diagram commutative.
@@
 \begin{theorem}
-The change of base of an extension algebra is again an extension algebra
+The change of base of an extension algebra is again an extension algebra.
 \end{theorem}
@@
 \begin{frame}
-\frametitle{\bf Introduction to the theory of weakening}
+\frametitle{\bf Overview of the theory of weakening}
 \begin{itemize}
-\item We will introduce an operation for weakening which acts at the three
-levels of fundamental structures (contexts, families and terms).
-  \pause
-\item It will preserve extension, so it will be implemented as an extension
+\item In the syntax, we will introduce an operation for weakening which acts at the three
+levels of fundamental structures (contexts, families and terms). \pause
+\important{When $B$ is weakened by $A$, we denote this by $\ctxwk{A}{B}$.}
+  \pause
+\item Weakening will preserve extension, so it will be implemented as an extension
 homomorphism.
   \pause
-\item It will preserve itself, so weakening will also be a weakening
-homomorphism.
-  \pause
-\item There will be a notion of `Currying for weakening', which explains
+\item There will be a notion of \important{`Currying for weakening'}, which explains
 what happens when something is weakened by an extended object.
+  \pause
+\item Pre-weakening algebras will be extension algebras with a weakening operation
+satisfying an implementation of the currying condition. \pause Pre-weakening homomorphisms
+will preserve this structure.
+  \pause
+\item Weakening will preserve itself, so weakening we will require that weakening is
+a pre-weakening homomorphism.
 \end{itemize}
@@
       }
+  \pause
   \\
@@
 \begin{frame}
-\frametitle{\bf Pre-weakening}
-Let $\stesys$ be an extension algebra. A pre-weakening operation
-on $\stesys$ is an extension homomorphism 
+\frametitle{\bf The weakening operation}
+Let $\stesys$ be an extension algebra. A \important{weakening operation
+on $\stesys$} is an extension homomorphism 
 \begin{equation*}
@@
 \begin{frame}
-\frametitle{\bf Weakening for the families algebra}
-Let $\stesys$ be an extension algebra with pre-weakening operation
-$\mathbf{w}(\stesys)$. Then $\famesys{\stesys}$ has the pre-weakening operation
-$\mathbf{w}(\famesys{\stesys})$ which is uniquely determined by rendering the
-diagram
-\begin{small}
-\begin{equation*}
-\begin{tikzcd}[ampersand replacement=\&,column sep=large]
-\cobesys{\stesysf_2}{\famesys{\famesys{\stesys}}}{\eft[2]}{\eft[2]}
-  \ar{rr}{%
-      \pullback{\beta_1}{\boldsymbol{\beta}}{\eft}{\eft}
-    }
-  \ar[bend right]{ddr}[swap]{\pullbackpr{1}{\eft[2]}{\eft[2]}}
-  \ar[dotted]{dr}{\mathbf{w}(\famesys{\stesys})}
-  \&
-  {}\&
-\cobesys{\stesysf}{\famesys{\stesys}}{\eft}{\eft}
-  \ar{r}{\mathbf{w}(\stesys)}
-  \&
-\famesys{\famesys{\stesys}}
-  \ar{d}{\boldsymbol{\beta}}
-  \\
-  {}\&
-\famesys{\famesys{\famesys{\stesys}}}
-  \ar{r}{\boldsymbol{\beta}_\mathbf{2}}
-  \ar{d}[swap]{\eft[3]}
-  \&
-\famesys{\famesys{\stesys}}
-  \ar{d}{\eft[2]}
-  \ar{r}{\boldsymbol{\beta}}
-  \&
-\famesys{\stesys}
-  \ar{d}{\eft}
-  \\
-  {}\&
-\stesysf_2
-  \ar{r}[swap]{\efamext}
-  \&
-\stesysf
-  \ar{r}[swap]{\ectxext}
-  \&
-\stesysc
-\end{tikzcd}
-\end{equation*}
-\end{small}
-commutative.
-\end{frame}
-
-\begin{frame}
 \frametitle{\bf Rules for weakening: Currying}
@@
     }
+  \pause
 \\
@@
 \end{align*}
+\\[\baselineskip]
+\pause
+To express these rules algebraically, we need to define a weakening operation on
+$\famesys{\stesys}$, provided we have one on $\stesys$.
+\end{frame}
+
+\begin{frame}
+\frametitle{\bf Weakening for the families algebra}
+Let $\stesys$ be an extension algebra with weakening operation
+$\mathbf{w}(\stesys)$. Then $\famesys{\stesys}$ has the weakening operation
+$\mathbf{w}(\famesys{\stesys})$ which is uniquely determined by rendering the
+diagram
+\begin{small}
+\begin{equation*}
+\begin{tikzcd}[ampersand replacement=\&,column sep=large]
+\cobesys{\stesysf_2}{\famesys{\famesys{\stesys}}}{\eft[2]}{\eft[2]}
+  \ar{rr}{%
+      \pullback{\beta_1}{\boldsymbol{\beta}}{\eft}{\eft}
+    }
+  \ar[bend right]{ddr}[swap]{\pullbackpr{1}{\eft[2]}{\eft[2]}}
+  \ar[dotted]{dr}{\important{$\mathbf{w}(\famesys{\stesys})$}}
+  \&
+  {}\&
+\cobesys{\stesysf}{\famesys{\stesys}}{\eft}{\eft}
+  \ar{r}{\important{$\mathbf{w}(\stesys)$}}
+  \&
+\famesys{\famesys{\stesys}}
+  \ar{d}{\boldsymbol{\beta}}
+  \\
+  {}\&
+\famesys{\famesys{\famesys{\stesys}}}
+  \ar{r}{\boldsymbol{\beta}_\mathbf{2}}
+  \ar{d}[swap]{\eft[3]}
+  \&
+\famesys{\famesys{\stesys}}
+  \ar{d}{\eft[2]}
+  \ar{r}{\boldsymbol{\beta}}
+  \&
+\famesys{\stesys}
+  \ar{d}{\eft}
+  \\
+  {}\&
+\stesysf_2
+  \ar{r}[swap]{\efamext}
+  \&
+\stesysf
+  \ar{r}[swap]{\ectxext}
+  \&
+\stesysc
+\end{tikzcd}
+\end{equation*}
+\end{small}
+commutative.
 \end{frame}
@@
 \frametitle{\bf Pre-weakening algebras}
-A pre-weakening algebra $\stesys$ is an extension algebra $\stesys$ 
-with a pre-weakening operation 
+A \important{pre-weakening algebra} is an extension algebra $\stesys$ 
+with a weakening operation 
 $ \mathbf{w}(\stesys)
@@
     { [ \pullbackpr{1}{\eft\circ\eft[2]}{\eft},%
+        \important{$
         \mathbf{w}(\stesys)%
           \circ%
-        (\pullback{\efamext}{\catid{\famesys{\stesys}}}{\eft}{\eft})%
+        (\pullback{\efamext}{\catid{\famesys{\stesys}}}{\eft}{\eft})$}%
         ]%
@@
     [ \pullbackpr{1}{\eft\circ\eft[2]}{\eft},%
-      \mathbf{w}(\stesys)%
+      \important{$\mathbf{w}(\stesys)$}%
         \circ%
@@
 \cobesys{\stesysf_2}{\famesys{\famesys{\stesys}}}{\eft[2]}{\eft[2]}
-  \ar{d}{\mathbf{w}(\famesys{\stesys})}
+  \ar{d}{\important{$\mathbf{w}(\famesys{\stesys})$}}
   \\
@@
 \end{equation*}
-commutes.
+commutes (implementing currying for weakening).
 \end{frame}
@@
 \frametitle{\bf Pre-weakening homomorphisms}
-A pre-weakening homomorphism between pre-weakening algebras $\stesys'$ and $\stesys$ is an
+A \important{pre-weakening homomorphism} between pre-weakening algebras $\stesys'$ and $\stesys$ is an
 extension homomorphism $f:\stesys'\to \stesys$ such that additionally the diagram
@@
   \ar{r}{\beta}
-  \ar{d}{g^\ast(\eft)_1}
+  \ar{d}{g^\ast(\eft)_2}
   \&
@@
 \frametitle{\bf Weakening algebras}
-Thus, it makes sense to require that the pre-weakening operation itself is a
+Thus, it makes sense to require that the weakening operation itself is a
 pre-weakening homomorphism.
 \begin{definition}
-A \emph{weakening algebra} is a pre-weakening algebra $\stesys$ with the property that
-$\mathbf{w}(\stesys)$ is a pre-weakening morphism.
+A \important{weakening algebra} is a pre-weakening algebra $\stesys$ with the property that
+$\mathbf{w}(\stesys)$ is a pre-weakening homomorphism.
 \end{definition}
@@
 \end{small}
+\end{frame}
+
+\begin{frame}
+\frametitle{\bf Overview of the theory of projections}
+\begin{itemize}
+\item In the theory of projections, we will introduce units, which will eventually behave
+like \important{fiberwise identity morphisms of families}.
+  \pause
+\item Together with weakening, units will induce all the \important{projections}.
+  \pause
+\item \important{Pre-projection algebras} will be weakening algebras with units.
+  \pause
+\item \important{Projection algebras} will be pre-projection algebras for which
+weakening is a pre-projection homomorphism.
+\end{itemize}
 \end{frame}
@@
 \frametitle{\bf Pre-projection algebras}
-A pre-projection algebra is a weakening algebra $\stesys$ for which there is a term
+A \important{pre-projection algebra} is a weakening algebra $\stesys$ for which there is a term
 $\mathbf{i}(\stesys):\stesysf\to \stesyst_2$ such that the diagram
@@
 \frametitle{\bf Pre-projection homomorphisms}
-A pre-projection homomorphism from $\stesys$ to $\stesys'$ is a weakening homomorphism
+A \important{pre-projection homomorphism from $\stesys$ to $\stesys'$} is a weakening homomorphism
 $f:\stesys'\to \stesys$ such that the square
@@
 \begin{definition}
-A projection algebra is a pre-projection algebra for which the weakening operation
+A \important{projection algebra} is a pre-projection algebra for which the weakening operation
 is a pre-projection homomorphism.
+\end{definition}
+\pause
+\begin{definition}
+A \important{projection homomorphism} is a pre-projection homomorphism between projection
+algebras.
 \end{definition}
@@
 \begin{frame}
+\frametitle{\bf Overview of the theory of substitution}
+\begin{itemize}
+\item We will introduce an operation for substitution by a term, this allows us
+to consider \important{fibers of families}. The substitution of $P$ by $a$ will
+be denoted by $\subst{a}{P}$.
+  \pause
+\item Substitution will be compatible with extension, thus it will be implemented
+as an extension homomorphism.
+  \pause
+\item \important{Pre-substitution algebras} will be extension algebras with a substitution
+operation. \important{Pre-substitution homomorphisms} will preserve this structure.
+  \pause
+\item \important{Substitution algebras} will be pre-substitution algebras for which
+the substitution operation itself is a pre-substitution homomorphism.
+\end{itemize}
+\end{frame}
+
+\begin{frame}
 \frametitle{\bf Rules for the theory of substitution}
@@
 & \inference
-  { \jterm{\Gamma}{A}{x}
+  { \jterm{\Gamma}{A}{a}
     \jfam{{\Gamma}{A}}{P}
     }
-  { \jfam{\Gamma}{\subst{x}{P}}
+  { \jfam{\Gamma}{\subst{a}{P}}
     }
 & & \inference
-    { \jtermeq{\Gamma}{A}{x}{x'}
+    { \jtermeq{\Gamma}{A}{a}{a'}
       \jfameq{{\Gamma}{A}}{P}{P'}
       }
-    { \jfameq{\Gamma}{\subst{x}{P}}{\subst{x'}{P'}}
-      }
-    \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
+    { \jfameq{\Gamma}{\subst{a}{P}}{\subst{a'}{P'}}
+      }
+    \\
+& \inference
+  { \jterm{\Gamma}{A}{a}
     \jfam{{{\Gamma}{A}}{P}}{Q}
     }
-  { \jfam{{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}
+  { \jfam{{\Gamma}{\subst{a}{P}}}{\subst{a}{Q}}
     }
 & & \inference
-    { \jtermeq{\Gamma}{A}{x}{x'}
+    { \jtermeq{\Gamma}{A}{a}{a'}
       \jfameq{{{\Gamma}{A}}{P}}{Q}{Q'}
       }
-    { \jfameq{{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}{\subst{x'}{Q'}}
-      }
-    \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
+    { \jfameq{{\Gamma}{\subst{a}{P}}}{\subst{a}{Q}}{\subst{a'}{Q'}}
+      }
+    \\
+& \inference
+  { \jterm{\Gamma}{A}{a}
     \jterm{{{\Gamma}{A}}{P}}{Q}{g}
     }
-  { \jterm{{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}{\subst{x}{g}}
+  { \jterm{{\Gamma}{\subst{a}{P}}}{\subst{a}{Q}}{\subst{a}{g}}
     }
 & & \inference
-    { \jtermeq{\Gamma}{A}{x}{x'}
+    { \jtermeq{\Gamma}{A}{a}{a'}
       \jtermeq{{{\Gamma}{A}}{P}}{Q}{g}{g'}
@@
     { \jtermeq
-        {{\Gamma}{\subst{x}{P}}}
-        {\subst{x}{Q}}
-        {\subst{x}{g}}
-        {\subst{x'}{g'}}
+        {{\Gamma}{\subst{a}{P}}}
+        {\subst{a}{Q}}
+        {\subst{a}{g}}
+        {\subst{a'}{g'}}
       }
@@
 & \inference
-  { \jterm{\Gamma}{A}{x}
+  { \jterm{\Gamma}{A}{a}
     \jfam{{{\Gamma}{A}}{P}}{Q}
@@
       {\Gamma}
-      {\subst{x}{\ctxext{P}{Q}}}
-      {\ctxext{\subst{x}{P}}{\subst{x}{Q}}}
-    }
-  \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
+      {\subst{a}{\ctxext{P}{Q}}}
+      {\ctxext{\subst{a}{P}}{\subst{a}{Q}}}
+    }
+  \\
+& \inference
+  { \jterm{\Gamma}{A}{a}
     \jfam{{{{\Gamma}{A}}{P}}{Q}}{R}
@@
   { \jfameq
-      {{\Gamma}{\subst{x}{P}}}
-      {\subst{x}{\ctxext{Q}{R}}}
-      {\ctxext{\subst{x}{Q}}{\subst{x}{R}}}
+      {{\Gamma}{\subst{a}{P}}}
+      {\subst{a}{\ctxext{Q}{R}}}
+      {\ctxext{\subst{a}{Q}}{\subst{a}{R}}}
     }
@@
 \frametitle{\bf Pre-substitution algebras}
-A \emph{pre-substitution} for an extension algebra $\stesys$ is an
+A \important{pre-substitution} for an extension algebra $\stesys$ is an
 extension homomorphism
@@
 \pause
-A \emph{pre-substitution algebra} is an extension algebra
+A \important{pre-substitution algebra} is an extension algebra
 together with a pre-substitution.
@@
 \frametitle{\bf Pre-substitution homomorphisms}
-A \emph{pre-substitution homomorphism} is an extension homomorphism $f:\stesys'\to \stesys$
+A \important{pre-substitution homomorphism} is an extension homomorphism $f:\stesys'\to \stesys$
 for which the square
@@
 \begin{definition}
-A \emph{substitution algebra} is a pre-substitution algebra $\stesys$ with the property that
+A \important{substitution algebra} is a pre-substitution algebra $\stesys$ with the property that
 $\mathbf{S}(\stesys)$ is a pre-substitution morphism.
+\end{definition}
+\pause
+\begin{definition}
+A \important{substitution homomorphism} is a pre-substitution homomorphism between substitution
+algebras.
 \end{definition}
@@
 & \inference
-  { \jterm{\Gamma}{A}{x}
+  { \jterm{\Gamma}{A}{a}
     \jterm{{{\Gamma}{A}}{P}}{Q}{g}
@@
   { \jfameq
-      {{{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}}
-      {\subst{x}{{g}{R}}}
-      {\subst{{x}{g}}{{x}{R}}}
-    }
-  \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
+      {{{\Gamma}{\subst{a}{P}}}{\subst{a}{Q}}}
+      {\subst{a}{{g}{R}}}
+      {\subst{{a}{g}}{{a}{R}}}
+    }
+  \\
+& \inference
+  { \jterm{\Gamma}{A}{a}
     \jterm{{{\Gamma}{A}}{P}}{Q}{g}
@@
   { \jfameq
-      {{{{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}}{\subst{x}{{g}{R}}}}
-      {\subst{x}{{g}{S}}}
-      {\subst{{x}{g}}{{x}{S}}}
-    }
-  \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
+      {{{{\Gamma}{\subst{a}{P}}}{\subst{a}{Q}}}{\subst{a}{{g}{R}}}}
+      {\subst{a}{{g}{S}}}
+      {\subst{{a}{g}}{{a}{S}}}
+    }
+  \\
+& \inference
+  { \jterm{\Gamma}{A}{a}
     \jterm{{{\Gamma}{A}}{P}}{Q}{g}
@@
   { \jtermeq
-      {{{{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}}{\subst{x}{{g}{R}}}}
-      {\subst{x}{{g}{S}}}
-      {\subst{x}{{g}{k}}}
-      {\subst{{x}{g}}{{x}{k}}}
+      {{{{\Gamma}{\subst{a}{P}}}{\subst{a}{Q}}}{\subst{a}{{g}{R}}}}
+      {\subst{a}{{g}{S}}}
+      {\subst{a}{{g}{k}}}
+      {\subst{{a}{g}}{{a}{k}}}
     }
@@
 \item that weakenings are constant families;\pause
-\item that weakening by a family $A$ followed by substitution of the unit of $A$ has no effect; this is precomposing with the unit.
+\item everything is invariant with respect to precomposition with units.
 \end{itemize}
@@
 \frametitle{\bf Pre-E-systems}
-A \emph{pre-E-system} is an extension algebra $\stesys$ with a
+A \important{pre-E-system} is an extension algebra $\stesys$ with a
 weakening operation $\mathbf{w}(\stesys)$, units
@@
 \end{itemize}
-Pre-E-homomorphisms are extension homomorphisms which
+\important{Pre-E-homomorphisms} are extension homomorphisms which
 are both projection homomorphisms and substitution
@@
     \jfam{\Gamma}{B}
-    \jterm{\Gamma}{A}{x}
+    \jterm{\Gamma}{A}{a}
     }
@@
       {\Gamma}
-      {\subst{x}{\ctxwk{A}{B}}}
+      {\subst{a}{\ctxwk{A}{B}}}
       {B}
@@
     \jfam{{\Gamma}{B}}{Q}
-    \jterm{\Gamma}{A}{x}
-    }
-  { \jfameq{{\Gamma}{B}}{\subst{x}{\ctxwk{A}{Q}}}{Q}
-    }
-  \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
+    \jterm{\Gamma}{A}{a}
+    }
+  { \jfameq{{\Gamma}{B}}{\subst{a}{\ctxwk{A}{Q}}}{Q}
+    }
+  \\
+& \inference
+  { \jterm{\Gamma}{A}{a}
     \jterm{{\Gamma}{B}}{Q}{g}
     }
-  { \jtermeq{{\Gamma}{B}}{Q}{\subst{x}{\ctxwk{A}{g}}}{g}
+  { \jtermeq{{\Gamma}{B}}{Q}{\subst{a}{\ctxwk{A}{g}}}{g}
     }
@@
 \frametitle{\bf Weakened families are constant families}
-In a pre-E-system, we say that \emph{weakened families are 
+In a pre-E-system, we say that \important{weakened families are 
 constant families} if the diagram
@@
 \frametitle{\bf Invariance with respect to precomposition with units}
-In a pre-E-system $\stesys$ we say that \emph{everything is invariant
+In a pre-E-system $\stesys$ we say that \important{everything is invariant
 with respect to precomposition with units} if the diagram
@@
   { \jfameq
-      {{{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}}
-      {\subst{x}{\ctxwk{Q}{R}}}
-      {\ctxwk{\subst{x}{Q}}{\subst{x}{R}}}
-    }
-  \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
+      {{{\Gamma}{\subst{a}{P}}}{\subst{a}{Q}}}
+      {\subst{a}{\ctxwk{Q}{R}}}
+      {\ctxwk{\subst{a}{Q}}{\subst{a}{R}}}
+    }
+  \\
+& \inference
+  { \jterm{\Gamma}{A}{a}
     \jfam{{{\Gamma}{A}}{P}}{Q}
@@
   { \jfameq
-      {{{{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}}{\subst{x}{\ctxwk{Q}{R}}}}
-      {\subst{x}{\ctxwk{Q}{S}}}
-      {\ctxwk{\subst{x}{Q}}{\subst{x}{S}}}
-    }
-  \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
+      {{{{\Gamma}{\subst{a}{P}}}{\subst{a}{Q}}}{\subst{a}{\ctxwk{Q}{R}}}}
+      {\subst{a}{\ctxwk{Q}{S}}}
+      {\ctxwk{\subst{a}{Q}}{\subst{a}{S}}}
+    }
+  \\
+& \inference
+  { \jterm{\Gamma}{A}{a}
     \jfam{{{\Gamma}{A}}{P}}{Q}
@@
   { \jtermeq
-      {{{{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}}{\subst{x}{\ctxwk{Q}{R}}}}
-      {\subst{x}{\ctxwk{Q}{S}}}
-      {\subst{x}{\ctxwk{Q}{k}}}
-      {\ctxwk{\subst{x}{Q}}{\subst{x}{k}}}
+      {{{{\Gamma}{\subst{a}{P}}}{\subst{a}{Q}}}{\subst{a}{\ctxwk{Q}{R}}}}
+      {\subst{a}{\ctxwk{Q}{S}}}
+      {\subst{a}{\ctxwk{Q}{k}}}
+      {\ctxwk{\subst{a}{Q}}{\subst{a}{k}}}
     }
@@
 \inference
-  { \jterm{\Gamma}{A}{x}
+  { \jterm{\Gamma}{A}{a}
     \jfam{{{\Gamma}{A}}{P}}{Q}
@@
   { \unfoldall{\jhomeq
-      {{\Gamma}{\subst{x}{P}}}
-      {\subst{x}{Q}}
-      {\subst{x}{Q}}
-      {\subst{x}{\idtm{Q}}}
-      {\idtm{\subst{x}{Q}}}
+      {{\Gamma}{\subst{a}{P}}}
+      {\subst{a}{Q}}
+      {\subst{a}{Q}}
+      {\subst{a}{\idtm{Q}}}
+      {\idtm{\subst{a}{Q}}}
     }}
@@
 & \inference
-  { \jterm{\Gamma}{A}{x}
-    }
-  { \jfameq{\Gamma}{\subst{x}{\emptyf}}{\emptyf}
-    }
-  \\
-& \inference
-  { \jterm{\Gamma}{A}{x}
+  { \jterm{\Gamma}{A}{a}
+    }
+  { \jfameq{\Gamma}{\subst{a}{\emptyf}}{\emptyf}
+    }
+  \\
+& \inference
+  { \jterm{\Gamma}{A}{a}
     \jfam{{\Gamma}{A}}{P}
@@
   { \jfameq
-      {{\Gamma}{\subst{x}{P}}}
-      {\subst{x}{\emptyf}}
+      {{\Gamma}{\subst{a}{P}}}
+      {\subst{a}{\emptyf}}
       {\emptyf}
@@
   }
-{ \jfameq{\Gamma}{\subst{x}{P}}{\subst{x}{P}}
+{ \jfameq{\Gamma}{\subst{a}{P}}{\subst{a}{P}}
   }
@@
 \inference
-  { \jterm{\Gamma}{A}{x}
-    }
-  { \jtermeq{\Gamma}{A}{\subst{x}{\idtm{A}}}{x}
+  { \jterm{\Gamma}{A}{a}
+    }
+  { \jtermeq{\Gamma}{A}{\subst{a}{\idtm{A}}}{a}
     }

### 2014-11-06 Presentation/presentation.tex
worked on presentation

@@
-\documentclass[handout]{beamer}
+\documentclass{beamer}
 \setbeamercolor{frametitle}{fg=teal!90!black}
@@
 \author{Egbert Rijke}
-\institute{Carnegie Mellon University}
-\date{November 7th -- 10th, 2014}
+\institute{Carnegie Mellon University\\\texttt{erijke@andrew.cmu.edu}}
+\date{November 7th -- 10th 2014}
 
@@
       }
-\pause
-\intertext{%
-Introduction rules for \important{family extension}:}
+\end{align*}
+\pause
+Introduction rules for \important{family extension}:
+\begin{align*}
 & \inference
@@
 \end{align*}
+\pause
 implementing the introduction rules
@@
 \end{align*}
+\pause
 Thus, we will additionally require that we have a commuting diagram
@@
   & := \stesysff
-  \pause
   \\
@@
   & := \pullbackpr{1}{\ectxext}{\eft} : \stesysf_2\to\stesysf
-  \pause
   \\
 \stesysf_3 & := \pullback{\stesysf_2}{\stesysf_2}{\efamext}{\eft[2]}
-  \pause
   \\
@@
 \begin{frame}
-\frametitle{Subgoal: develop properties of extension algebras}
+\frametitle{\bf Subgoal: develop properties of extension algebras}
 We need to know:
@@
 \end{equation*}
+\pause
 such that the indicated squares commute, for which furthermore the squares
@@
 \end{equation*}
+\pause
 where
@@
 \end{theorem}
-
+\pause
 \begin{lemma}
@@
 \end{theorem}
-
+\pause
 \begin{theorem}
@@
       }
-  \pause
-  \\
+\end{align*}
+  \pause
+\begin{align*}
 & \inference
@@
     }
-  \pause
-\\
+\end{align*}
+\pause
+\begin{align*}
 & \inference
@@
 \end{equation*}
+\pause
 to be the unique extension homomorphism rendering the diagram
@@
     }
-\\
-\pause
+\end{align*}
+\pause
+\begin{align*}
 & \inference
@@
 \\
-\pause
 & \inference
@@
       }
-    \\
+\end{align*}
+\pause
+\begin{align*}
 & \inference
@@
 \begin{tikzcd}[ampersand replacement=\&,column sep=large]
-\pullback{\stesyst}{\stesysf_2}{\ebd}{\eft[2]}
-  \ar{r}{s(\stesys)_0}
+\pullback{\stesyst}{\famesys{\famesys{\stesysf}}}{\ebd}{\eft[2]}
+  \ar{r}{\mathbf{s}(\stesys)}
   \ar{d}[swap]{\ebd\circ\pullbackpr{1}{\ebd}{\eft[2]}}
   \&
-\stesysf 
+\famesys{\stesys} 
   \ar{d}{\eft}
@@
     }
-  \\
+\end{align*}
+\pause
+\begin{align*}
 & \inference
@@
 \end{itemize}
+\pause
 \important{Pre-E-homomorphisms} are extension homomorphisms which
@@
 \frametitle{\bf E-systems}
-An \emph{E-system} is an extension algebra with
+An \important{E-system} is an extension algebra with\pause
 \begin{itemize}
-\item a weakening operation and units, giving it the structure of a projection algebra;
-\item a substitution operation, giving it the structure of a substitution algebra
+\item a weakening operation and units, giving it the structure of a projection algebra;\pause
+\item a substitution operation, giving it the structure of a substitution algebra;\pause
 \item (an empty context and empty families);
 \end{itemize}
-such that
+such that\pause
 \begin{itemize}
-\item Weakening is a substitution homomorphism;
-\item Substitution is a projection homomorphism;
-\item Weakened families are constant families;
-\item Precomposition with identity functions leaves everything invariant.
+\item Weakening is a substitution homomorphism;\pause
+\item Substitution is a projection homomorphism;\pause
+\item Weakened families are constant families;\pause
+\item Precomposition with identity functions leaves everything invariant.\pause
 \item (The requirements regarding the empty context and families according to their rules)

### 2014-11-06 Presentation/presentation.tex
presentation finished

@@
 \frametitle{\bf Goal and acknowledgements}
-We will give an essentially algebraic theory of type dependency. Thus, we will
+In this presentation, an essentially algebraic theory of type dependency will be given. Thus, we will
 do two things:
@@
 name-free type theory.
-\item Define the corresponding algebraic objects in a category with finite limits.
+\item Define the corresponding algebraic objects, called \important{E-systems}, in a category with finite limits.
 \end{itemize}
-\pause
-People unfamiliar with type theory should get out of this presentation 
-a sense of what dependent type theory is.
-\\[\baselineskip]
 \pause
@@
 \pause
-\item The theory is capable of interpreting itself. Investigations to internal
-models of type theory originally motivated me to write type theory this way.
-\pause
 \item The meta-theory should not require anything more than rules for handling
@@
 \begin{frame}
+\frametitle{\bf Consequences of being algebraic}
+The requirement that the theory is essentially algebraic gives immediate access to the standard tools of algebra, such as
+\begin{itemize}
+\item free algebras;
+\item quotients algebras.
+\end{itemize}
+\pause
+The major source of examples of E-systems are C-systems.
+\\[\baselineskip]
+\pause
+In a way similar to the definition of E-systems, it is possible to define
+\important{internal E-systems}. One of the aims of this research was to have
+a systematic treatment of internal models.
+\begin{itemize}
+\item Nothing more than plain dependent type theory is needed to express what
+an internal E-system is.
+\end{itemize}
+\end{frame}
+
+\begin{frame}
 \frametitle{\bf Overview of the theories}
@@
   \\
-\textbf{Weakening} \ar[very thick,-stealth]{d} \& \textbf{Substitution} \ar[very thick,-stealth]{dd} \& \makebox[3cm]{\textbf{Empty context and families}} \ar[very thick,-stealth]{ddl}
-\\
-\textbf{Projection} \ar[very thick,-stealth]{dr}
+\makebox[.7cm]{\textbf{Weakening}} \ar[very thick,-stealth]{d} \& \textbf{Substitution} \ar[very thick,-stealth]{dd} \& \makebox[3cm]{\textbf{Empty context and families}} \ar[very thick,-stealth]{ddl}
+\\
+\makebox[.7cm]{\textbf{Projection}} \ar[very thick,-stealth]{dr}
 \\
@@
 \begin{frame}
-\frametitle{\bf The end}
-\begin{center}
-Thank you for your attention
-\end{center}
+\frametitle{\bf Towards algebraic Homotopy Type Theory}
+\begin{footnotesize}
+\begin{equation*}
+\begin{tikzcd}[ampersand replacement=\&,column sep=0em,row sep=1.3em]
+ {}\& 
+\textbf{Extension} 
+  \ar[very thick,-stealth]{dr} 
+  \ar[very thick,-stealth]{d} 
+  \ar[very thick,-stealth]{dl}
+  \\
+\makebox[2cm]{\textbf{Weakening}} 
+  \ar[very thick,-stealth]{d} 
+  \& 
+\textbf{Substitution} 
+  \ar[very thick,-stealth]{dd} 
+  \& 
+\makebox[3cm]{\textbf{Empty context and families}} 
+  \ar[very thick,-stealth]{ddl}
+  \\
+\makebox[2cm]{\textbf{Projection}} 
+  \ar[very thick,-stealth]{dr}
+  \\
+  {}\& 
+\textbf{Dependent Type Theory}
+  \ar[very thick,-stealth]{dr}
+  \ar[very thick,-stealth]{dl}
+  \\
+\makebox[2cm]{\bf Inductive types}
+  \ar[very thick,-stealth]{ddr}
+  \&\&
+\makebox[2cm]{\bf $\Pi$-types}
+  \ar[very thick,-stealth]{d}
+  \\
+  {}\&
+  {}\&
+\makebox[2cm]{\bf Universes}
+  \ar[very thick,-stealth]{dl}
+  \\
+  {}\&
+\mbox{\bf Univalent mathematics(?)}
+  \ar[very thick,-stealth]{d}
+  \\
+  {}\&
+\mbox{\bf Higher inductive types}
+\end{tikzcd}
+\end{equation*}
+\end{footnotesize}
 \end{frame}

### 2014-11-07 Presentation/presentation.tex
comparison with B-systems added

@@
 \newcommand\important[1]{\textbf{\color{red!90!black}#1}}
+\newcommand{\Btilde}{\skew{4}{\tilde}{B}}
 
@@
 \frametitle{\bf Goal and acknowledgements}
-In this presentation, an essentially algebraic theory of type dependency will be given. Thus, we will
-do two things:
+We will do two things:
 \begin{itemize}
-\item Give a syntactic description of the theory in the usual style of
+\item Give a syntactic reformulation of type theory in the usual style of
 name-free type theory.
@@
 \begin{frame}
-\frametitle{\bf Consequences of being algebraic}
+\frametitle{\bf The scope of the theory}
 The requirement that the theory is essentially algebraic gives immediate access to the standard tools of algebra, such as
@@
 \pause
-The major source of examples of E-systems are C-systems.
-\\[\baselineskip]
+The major source of examples of E-systems are \important{B-systems}.
+\begin{itemize}
+\item B-systems are stratified E-systems.
+\item There is a full and faithful functor from C-systems to pre-B-systems with the B-systems as its image.
+\end{itemize}
 \pause
@@
 \jalign\jctx{\Gamma} 
-& \jalign\jctxeq{\Gamma}{\Gamma'}
+& \jalign\jctxeq{\Gamma}{\Delta}
   \\
@@
 \jalign\jterm{\Gamma}{A}{a} 
-& \jalign\jtermeq{\Gamma}{A}{a}{y}.
+& \jalign\jtermeq{\Gamma}{A}{a}{b}.
 \end{align*}
@@
 
-\begin{comment}
-\begin{frame}
-\frametitle{\bf Rules: judgmental equality is an equivalence relation}
-\begin{footnotesize}
-\begin{align*}
-& \inference
-  { \jctx{\Gamma}
-    }
-  { \jctxeq{\Gamma}{\Gamma}
-    } 
-& & \inference
-    { \jctxeq{\Gamma}{\Delta}
-      }
-    { \jctxeq{\Delta}{\Gamma}
-      } 
-& & \inference
-    { \jctxeq{\Gamma}{\Delta}
-      \jctxeq{\Delta}{\greek{E}}
-      }
-    { \jctxeq{\Gamma}{\greek{E}}
-      }
-    \\
-& \inference
-  { \jfam{\Gamma}{A}
-    }
-  { \jfameq{\Gamma}{A}{A}
-    } 
-& & \inference
-    { \jfameq{\Gamma}{A}{B}
-      }
-    { \jfameq{\Gamma}{B}{A}
-      }
-& & \inference
-    { \jfameq{\Gamma}{A}{B}
-      \jfameq{\Gamma}{B}{C}
-      }
-    { \jfameq{\Gamma}{A}{C}
-      }
-    \\
-& \inference
-  { \jterm{\Gamma}{A}{a}
-    }
-  { \jtermeq{\Gamma}{A}{a}{a}
-    }
-& & \inference
-    { \jtermeq{\Gamma}{A}{a}{y}
-      }
-    { \jtermeq{\Gamma}{A}{y}{a}
-      }
-& & \inference
-    { \jtermeq{\Gamma}{A}{a}{y}
-      \jtermeq{\Gamma}{A}{y}{z}
-      }
-    { \jtermeq{\Gamma}{A}{a}{z}
-      }
-\end{align*}
-\begin{align*}
-& \inference
-  { \jctxeq{\Gamma}{\Delta}
-    \jfam{\Gamma}{A}
-    }
-  { \jfam{\Delta}{A}
-    }
-& & \inference
-    { \jctxeq{\Gamma}{\Delta}
-      \jfameq{\Gamma}{A}{B}
-      }
-    { \jfameq{\Delta}{A}{B}
-      }
-    \\
-& \inference
-  { \jctxeq{\Gamma}{\Delta}
-    \jterm{\Gamma}{A}{a}
-    }
-  { \jterm{\Delta}{A}{a}
-    }
-& & \inference
-    { \jctxeq{\Gamma}{\Delta}
-      \jtermeq{\Gamma}{A}{a}{y}
-      }
-    { \jtermeq{\Delta}{A}{a}{y}
-      }
-    \\
-& \inference
-  { \jfameq{\Gamma}{A}{B}
-    \jterm{\Gamma}{A}{a}
-    }
-  { \jterm{\Gamma}{B}{a}
-    }
-& & \inference
-    { \jfameq{\Gamma}{A}{B}
-      \jtermeq{\Gamma}{A}{a}{y}
-      }
-    { \jtermeq{\Gamma}{B}{a}{y}
-      }
-\end{align*}
-\end{footnotesize}
-\end{frame}
-\end{comment}
-
 \begin{frame}
@@
 \item Every term has a family associated to it, this is represented by
-$\partial:\stesyst\to\stesysf$.
+$\ebd:\stesyst\to\stesysf$.
 \end{itemize}
+\end{frame}
+
+\begin{frame}
+\frametitle{\bf Comparison with B-systems}
+\begin{minipage}[t]{.42\textwidth}
+The fundamental structure of E-systems:
+\\[1.8cm]
+\begin{equation*}
+\begin{tikzcd}[ampersand replacement=\&]
+\stesyst
+  \ar{dr}[swap]{\ebd}
+  \\
+  {}\&
+\stesysf
+  \ar{d}{\eft}
+  \\
+  {}\&
+\stesysc
+\end{tikzcd}
+\end{equation*}
+\end{minipage}
+\hfill
+\begin{minipage}[t]{.42\textwidth}
+The underlying structure of B-systems:
+\vfill
+\begin{equation*}
+\begin{tikzcd}[ampersand replacement=\&]
+\Btilde_2
+  \ar{dr}[swap]{\ebd}
+  \&
+\vdots
+  \ar{d}{\eft}
+  \\
+\Btilde_1
+  \ar{dr}[swap]{\ebd}
+  \&
+B_2
+  \ar{d}{\eft}
+  \\
+\Btilde_0
+  \ar{dr}[swap]{\ebd}
+  \&
+B_1
+  \ar{d}{\eft}
+  \\
+  {}\&
+B_0
+\end{tikzcd}
+\end{equation*}
+\end{minipage}
 \end{frame}
@@
 \begin{frame}
-\frametitle{\bf Mixed commutative diagrams}
+\frametitle{\bf Condensed commutative diagrams}
 Suppose $f:\stesys\to\stesys'$ is a pre-extension homomorphism. We say that
@@
 \end{align*}
+\pause
 Weakening preserves the unit:

### 2015-01-19 Rules/rules_extension.tex
fixed broken forward reference to term extension

@@
 
-In \autoref{extension-on-terms} we will also define an extension operation
-on terms, which defines a term $\jterm{\Gamma}{{A}{P}}{\tmext{x}{u}}$ for
-every $\jterm{\Gamma}{A}{x}$ and $\jterm{\Gamma}{\subst{x}{P}}{u}$. This
-operation is defined by using the substitution operation and identity terms.
+Using identity terms and substitution, it is also possible to define an 
+extension operation on terms, which defines a term 
+$\jterm{\Gamma}{{A}{P}}{\tmext{x}{u}}$ for every $\jterm{\Gamma}{A}{x}$ and 
+$\jterm{\Gamma}{\subst{x}{P}}{u}$. Thus, there is a formal definition of the
+\emph{pair} $\tmext{x}{u}$. We shall not go into this, in the current paper,
+but it will be very useful to have when one further develops the theory of
+E-systems.
 

### 2015-01-19 main.tex
fixed broken forward reference to term extension

@@
 
-\includeonly{Algebras/article-models-cat}
+%\includeonly{Algebras/article-models-cat}
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\title{Internal models of type theory and internal higher categories for the
-Univalent Foundations}
+\title{An essentially algebraic formulation of dependent type theory}
 \author{Egbert Rijke}
@@
 
-\tableofcontents
+%\tableofcontents
 
-\input{Introduction/article-models-intro}
+%\input{Introduction/article-models-intro}
 
-\part{Type theory}
+%\part{Type theory}
 
+\begin{comment}
 In this part we develop type theory from the ground up. We start with a type
@@
 models for E-systems, presented in \autoref{part:models}.
+\end{comment}
 
-\include{Rules/rules}
+\input{Rules/rules}
 
-\include{Derivations/article-models-ttderived}
+%\include{Derivations/article-models-ttderived}
 
-\part{Internal Models}\label{part:models}
+%\part{Internal Models}\label{part:models}
 
-\include{Internal_models/article-models-ttinternal}
+%\include{Internal_models/article-models-ttinternal}
 
-\part{Type constructors}
+%\part{Type constructors}
 
-\include{Dependent_function_types/article-models-ttdp}
+%\include{Dependent_function_types/article-models-ttdp}
 
-\include{Universes/article-models-ttuniverses}
+%\include{Universes/article-models-ttuniverses}
 
-\include{Junk/article-models-ttind}
+%\include{Junk/article-models-ttind}
 
-\part{Categorical semantics}
+%\part{Categorical semantics}
 
-\include{Algebras/article-models-cat}
+\input{Algebras/article-models-cat}
 

### 2015-02-10 Algebras/empty.tex
large reorganisation

--- 2014-11-03 Algebras/article-models-cat.tex
@@
-\section{E-objects in categories with finite limits}
-In this section we assume that $\cat{C}$ is a finitely complete category and
-whenever we write a pullback, we assume that it is chosen. Recall that for
-any morphism $f:A\to B$ in a category $\cat{C}$ with chosen pullbacks, there
-is a functor
-\begin{equation*}
-f^\ast : \cat{C}/B\to\cat{C}/A.
-\end{equation*}
-As usual, when $g:X\to B$ is a morphism, we will write $f^\ast(X)$ for the
-domain of $f^\ast(g)$. When there is more than one morphism $X\to B$ involved,
-as will be the case below, we will write $\pullback{A}{X}{f}{g}$. The projections
-will be written as $\pullbackpr{1}{f}{g}$ and $\pullbackpr{2}{f}{g}$. So in this notation, a
-typical pullback diagram has the following form:
-\begin{equation*}
-\begin{tikzcd}[column sep=large]
-\pullback{A}{X}{f}{g}
-  \ar{r}{\pullbackpr{1}{f}{g}}
-  \ar{d}[swap]{\pullbackpr{2}{f}{g}}
-  &
-A \ar{d}{f}
-  \\
-X \ar{r}[swap]{g}
-  &
-B
-\end{tikzcd}
-\end{equation*}
-Also, when we have a commutative diagram of the form
-\begin{equation*}
-\begin{tikzcd}
-A \ar{r}{f}
-  \ar{d}{a}
-  &
-X \ar{d}
-  & 
-B \ar{l}[swap]{g}
-  \ar{d}{b}
-  \\
-A'
-  \ar{r}[swap]{f'}
-  &
-X'
-  &
-B'
-  \ar{l}{g'}
-\end{tikzcd}
-\end{equation*}
-we will denote the unique map from $\pullback{A}{B}{f}{g}$ to $\pullback{A'}{B'}{f'}{g'}$
-such that the diagram
-\begin{equation*}
-\begin{tikzcd}
-  {}
-  & 
-\pullback{A'}{B'}{f'}{g'}
-  \ar{dd}
-  \ar{rr}
-  &
-  &
-B'
-  \ar{dd}{g'}
-  \\
-\pullback{A}{B}{f}{g}
-  \ar{dd}
-  \ar[crossing over]{rr}
-  \ar[dotted]{ur}{\pullback{a}{b}{f'}{g'}}
-  &
-  &
-B \ar{ur}{b}
-  \\
-  {}
-  &
-A'
-  \ar{rr}
-  &
-  &
-X'
-  \\
-A \ar{rr}[swap]{f}
-  \ar{ur}{a}
-  &
-  &
-X \ar[crossing over,leftarrow]{uu}[near end,swap]{g}
-  \ar{ur}
-\end{tikzcd}
-\end{equation*}
-commutes, by $\pullback{a}{b}{f'}{g'}$. In the current work, we shall
-write $A\times B$ for the pullback of $A\rightarrow 1\leftarrow B$, and
-$\pi_1$ and $\pi_2$ for its projections (thus, no separate choice of
-cartesian products is made).
-
-\subsection{Extension objects}
-\begin{defn}
-A \emph{pre-extension object $\stesys$ in $\cat{C}$} consists of a \emph{fundamental structure}, which is a diagram of the form
-\begin{equation*}
-\begin{tikzcd}
-\stesyst
-  \ar{d}[swap]{\ebd}
-  \\
-\stesysf
-  \ar{d}[swap]{\eft}
-  \\
-\stesysc
-\end{tikzcd}
-\end{equation*}
-in $\cat{C}$ together with the \emph{context extension} and \emph{family extension} operations
-\begin{align*}
-\ectxext &:\stesysf\to \stesysc\\
-\efamext & :\stesysff\to \stesysf,
-\end{align*}
-respectively, such that the diagram
-\begin{equation*}
-\begin{tikzcd}
-\stesysf_2 
-  \ar{r}{\efamext} 
-  \ar{d}[swap]{\eft[2]} 
-  & 
-\stesysf 
-  \ar{d}{\eft}
-  \\
-\stesysf
-  \ar{r}[swap]{\eft} 
-  & 
-\stesysc
-\end{tikzcd}
-\end{equation*}
-commutes.
-\end{defn}
+\subsection{Extension algebras with empty context and families}
 
 \begin{defn}
-We introduce the following notation:
-\begin{align*}
-\stesysf_2 
-  & := \stesysff
-  \\
-\eft[2] 
-  & := \pullbackpr{1}{\ectxext}{\eft} : \stesysf_2\to\stesysf
-  \\
-\stesysf_3 & := \pullback{\stesysf_2}{\stesysf_2}{\efamext}{\eft[2]}
-  \\
-\eft[3]
-  & := \pullbackpr{1}{\efamext}{\eft[2]} : \stesysf_3\to\stesysf_2.
-\end{align*}
-Then it follows that the outer square in the diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=large]
-\stesysf_3
-  \ar[dotted]{dr}{\eext{2}}
-  \ar{rr}{\pullback{\pullbackpr{2}{\ectxext}{\eft}}{\pullbackpr{2}{\ectxext}{\eft}}{\ectxext}{\eft}}
-  \ar{dd}[swap]{\eft[3]}
-  & 
-  &
-\stesysf_2
-  \ar{d}{\efamext}
-  \\
-  &
-\stesysf_2
-  \ar{d}[swap]{\eft[2]}
-  \ar{r}{\pullbackpr{2}{\ectxext}{\eft}}
-  &
-\stesysf
-  \ar{d}{\eft}
-  \\
-\stesysf_2
-  \ar{r}[swap]{\eft[2]}
-  &
-\stesysf
-  \ar{r}[swap]{\ectxext}
-  &
-\stesysc
-\end{tikzcd}
-\end{equation*}
-commutes. We define $\eext{2}$ to be the unique morphism rendering the above diagram
-commutative. Now we may continue to define
-\begin{align*}
-\stesysf_4 
-  & := 
-\pullback{\stesysf_3}{\stesysf_3}{\eext{2}}{\eft[3]}
-  \\
-\eft[4] 
-  & := 
-\pullbackpr{1}{\eext{2}}{\eft[3]}.
-\end{align*}
-Then we see that the outer square of the diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=large]
-\stesysf_4
-  \ar[dotted]{dr}{\eext{3}}
-  \ar{rr}{\pullback{\pullbackpr{2}{\efamext}{\eft[2]}}{\pullbackpr{2}{\efamext}{\eft[2]}}{\efamext}{\eft[2]}}
-  \ar{dd}[swap]{\eft[4]}
-  & 
-  &
-\stesysf_3
-  \ar{d}{\eext{2}}
-  \\
-  &
-\stesysf_3
-  \ar{d}[swap]{\eft[3]}
-  \ar{r}{\pullbackpr{2}{\efamext}{\eft[2]}}
-  &
-\stesysf_2
-  \ar{d}{\eft[2]}
-  \\
-\stesysf_3
-  \ar{r}[swap]{\eft[3]}
-  &
-\stesysf_2
-  \ar{r}[swap]{\efamext}
-  &
-\stesysf
-\end{tikzcd}
-\end{equation*}
-commutes,
-so we may define $\eext{3}$ to be the unique map which renders the diagram
-commutative. It
-is straightforward to continue this process by induction, but we shall need not
-go any further in this article.
-\end{defn}
-
-\begin{defn} 
-An extension object is a pre-extension object $\stesys$ for which 
-the diagrams
-\begin{equation*}
-\begin{tikzcd}
-\stesysf_2 
-  \ar{d}[swap]{\pullbackpr{2}{\ectxext}{\eft}} 
-  \ar{r}{\efamext} 
-  & 
-\stesysf 
-  \ar{d}{\ectxext}
-  \\
-\stesysf 
-  \ar{r}[swap]{\ectxext} 
-  & 
-\stesysc
-\end{tikzcd}
-\qquad
-\begin{tikzcd}
-\stesysf_3
-  \ar{d}[swap]{\pullbackpr{2}{\efamext}{\eft[2]}}
-  \ar{r}{\eext{2}}
-  & 
-\stesysf_2 
-  \ar{d}{\efamext} 
-  \\
-\stesysf_2 
-  \ar{r}[swap]{\efamext} 
-  &
-\stesysf
-\end{tikzcd}
-\end{equation*}
-commute.
-\end{defn}
-
-\begin{comment}
-\begin{lem}
-There exists an isomorphism $\alpha$ such that the triangle
-\begin{equation*}
-\begin{tikzcd}[column sep=tiny]
-\pullback{\stesysf}{\stesysf_2}{\ectxext}{\eft\circ\eft[2]}
-  \ar[dotted]{rr}{\alpha}
-  \ar{dr}[swap]{\pullback{\catid{\stesysf}}{\efamext}{\ectxext}{\eft}}
-  &
-  &
-\stesysf_3
-  \ar{dl}{\eext{2}}
-  \\
-& \stesysf_2
-\end{tikzcd}
-\end{equation*}
-commutes
-\end{lem}
-
-\begin{proof}
-There is a unique morphism $\alpha:
-\pullback{\stesysf}{\stesysf_2}{\ectxext}{\eft\circ\eft[2]}\to\stesysf_3$
-rendering the diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=large]
-\pullback{\stesysf}{\stesysf_2}{\ectxext}{\eft\circ\eft[2]}
-  \ar[bend left=10,yshift=.5ex]{drrr}{\pullbackpr{2}{\ectxext}{\eft}\circ\pullbackpr{2}{\ectxext}{\eft\circ\eft[2]}}
-  \ar[bend right=10]{ddr}[swap]{\pullback{\catid{\stesysf}}{\eft[2]}{\ectxext}{\eft}}
-  \ar[dotted]{dr}{\alpha}
-  \\
-& \stesysf_3
-  \ar{r}{\pullbackpr{2}{\efamext}{\eft[2]}}
-  \ar{d}{\eft[3]}
-  &
-\stesysf_2
-  \ar{d}[swap]{\eft[2]}
-  \ar{r}[swap]{\pullbackpr{2}{\ectxext}{\eft}}
-  &
-\stesysf
-  \ar{d}{\eft}
-  \\
-{} & \stesysf_2
-  \ar{r}[swap]{\efamext}
-  &
-\stesysf
-  \ar{r}[swap]{\ectxext}
-  &
-\stesysc
-\end{tikzcd}
-\end{equation*}
-\end{proof}
-\end{comment}
-
-\begin{defn}
-Suppose $\stesys$ is a pre-extension object of $\cat{C}$. Then we define the pre-extension object
-$\famesys{\stesys}$ to consist of the fundamental structure
-\begin{equation*}
-\begin{tikzcd}
-\stesyst_2
-  \ar{d}{\ebd[2]}
-  \\
-\stesysf_2
-  \ar{d}{\eft[2]}
-  \\
-\stesysf
-\end{tikzcd}
-\end{equation*}
-where
-\begin{align*}
-\stesyst_2 
-  & := \pullback{\stesysf}{\stesyst}{\ectxext}{\eft\circ\ebd}
-  \\
-\ebd[2]
-  & := \ectxext^\ast(\ebd),
-\end{align*}
-with the extension operations
-\begin{align*}
-\efamext 
-  & 
-  : \stesysf_2\to\stesysf\\
-\eext{2} & : \stesysf_3\to\stesysf_2.
-\end{align*}
-\end{defn}
-
-In \autoref{famextobj} below we shall show that $\famesys{\stesys}$ is an
-extension algebra whenever $\stesys$ is an extension algebra. We shall need
-a handful of lemmas to give the proof.
-
-\begin{defn}
-We define
-\begin{align*}
-\beta_1 
-  & := 
-\pullbackpr{2}{\ectxext}{\eft}
-  & &
-  : \stesysf_2\to\stesysf
-  \\
-\beta_2
-  & :=
-\pullback{\beta_1}{\beta_1}{\ectxext}{\eft}
-  & &
-  : \stesysf_3\to\stesysf_2
-  \\
-\beta_3
-  & :=
-\pullback{\beta_2}{\beta_2}{\efamext}{\eft[2]}
-  & &
-  : \stesysf_4\to\stesysf_3.
-\end{align*}
-\end{defn}
-
-\begin{lem}
-Let $\stesys$ be a pre-extension object. Then the square
-\begin{equation*}
-\begin{tikzcd}[column sep=10em]
-\stesysf_4
-  \ar{r}{\pullback{\pullbackpr{2}{\efamext}{\eft[2]}}{\pullbackpr{2}{\efamext}{\eft[2]}}{\efamext}{\eft[2]}}
-  \ar{d}[swap]{\beta_3}
-  &
-\stesysf_3
-  \ar{d}{\beta_2}
-  \\
-\stesysf_3
-  \ar{r}[swap]{\beta_2}
-  &
-\stesysf_2
-\end{tikzcd}
-\end{equation*}
-commutes.
-\end{lem}
-
-\begin{proof}
-Left to the reader.
-\end{proof}
-
-\begin{comment}
-\begin{proof}
-It is straightforward to verify the equalities
-\begin{align*}
-\pullbackpr{1}{\ectxext}{\eft}\circ\beta_2\circ
-  (\pullback{\pullbackpr{2}{\efamext}{\eft[2]}}{\pullbackpr{2}{\efamext}{\eft[2]}}{\efamext}{\eft[2]})
-  & =
-\pullbackpr{1}{\ectxext}{\eft}\circ\beta_2\circ\beta_3
-  \\
-\pullbackpr{2}{\ectxext}{\eft}\circ\beta_2\circ
-  (\pullback{\pullbackpr{2}{\efamext}{\eft[2]}}{\pullbackpr{2}{\efamext}{\eft[2]}}{\efamext}{\eft[2]})
-  & =
-\pullbackpr{2}{\ectxext}{\eft}\circ\beta_2\circ\beta_3.\qedhere
-\end{align*}
-\end{proof}
-\end{comment}
-
-Note that the square
-\begin{equation*}
-\begin{tikzcd}
-\stesysf_3
-  \ar{r}{\beta_2}
-  \ar{d}[swap]{\eext{2}}
-  &
-\stesysf_2
-  \ar{d}{\efamext}
-  \\
-\stesysf_2
-  \ar{r}[swap]{\beta_1}
-  &
-\stesysf
-\end{tikzcd}
-\end{equation*}
-commutes by definition. We have a similar result relating $\eext{3}$ and
-$\eext{2}$.
-
-\begin{lem}
-Let $\stesys$ be a pre-extension object. Then the square
-\begin{equation*}
-\begin{tikzcd}
-\stesysf_4
-  \ar{r}{\beta_3}
-  \ar{d}[swap]{\eext{3}}
-  &
-\stesysf_3
-  \ar{d}{\eext{2}}
-  \\
-\stesysf_3
-  \ar{r}[swap]{\beta_2}
-  &
-\stesysf_2
-\end{tikzcd}
-\end{equation*}
-commutes.
-\end{lem}
-
-\begin{proof}
-Left to the reader.
-\end{proof}
-\begin{comment}
-\begin{proof}
-It is straightforward to verify the equalities
-\begin{align*}
-\pullbackpr{1}{\ectxext}{\eft}\circ\beta_2\circ\eext{3}
-  & = \beta_1\circ\eft[3]\circ\eft[4]
-  \\
-\pullbackpr{1}{\ectxext}{\eft}\circ\eext{2}\circ\beta_3
-  & = \beta_1\circ\eft[3]\circ\eft[4].
-\end{align*}
-Thus, it remains to verify that
-\begin{equation*}
-\pullbackpr{2}{\ectxext}{\eft}\circ\beta_2\circ\eext{3}
-  = \pullbackpr{2}{\ectxext}{\eft}\circ\eext{2}\circ\beta_3.
-\end{equation*}
-It is straightforward to see that the diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=large]
-\stesysf_4
-  \ar{dd}[swap]{\pullback{\pullbackpr{2}{\efamext}{\eft[2]}}{\pullbackpr{2}{\efamext}{\eft[2]}}{\efamext}{\eft[2]}}
-  \ar{r}{\eext{3}}
-  &
-\stesysf_3
-  \ar{r}{\beta_2}
-  \ar{d}[swap]{\pullbackpr{2}{\efamext}{\eft[2]}}
-  &
-\stesysf_2
-  \ar{dd}{\pullbackpr{2}{\ectxext}{\eft}}
-  \\
-  {} &
-\stesysf_2
-  \ar{dr}{\beta_1}
-  \\
-\stesysf_3
-  \ar{ur}{\eext{2}}
-  \ar{r}[swap]{\beta_2}
-  &
-\stesysf_2
-  \ar{r}[swap]{\efamext}
-  &
-\stesysf
-\end{tikzcd}
-\end{equation*}
-commutes. It is likewise straightforward to see that the diagram
-\begin{equation*}
-\begin{tikzcd}
-\stesysf_4
-  \ar{r}{\beta_3}
-  \ar{d}[swap]{\pullback{\pullbackpr{2}{\efamext}{\eft[2]}}{\pullbackpr{2}{\efamext}{\eft[2]}}{\efamext}{\eft[2]}}
-  &
-\stesysf_3
-  \ar{r}{\eext{2}}
-  \ar{d}[swap]{\beta_2}
-  &
-\stesysf_2
-  \ar{d}{\pullbackpr{2}{\ectxext}{\eft}}
-  \\
-\stesysf_3
-  \ar{r}[swap]{\beta_2}
-  &
-\stesysf_2
-  \ar{r}[swap]{\efamext}
-  &
-\stesysf
-\end{tikzcd}
-\end{equation*}
-commutes, completing our goal.
-\end{proof}
-\end{comment}
-
-\begin{thm}[Local extension structure]\label{famextobj}
-If $\stesys$ is an extension object, then so is $\famesys{\stesys}$.
-\end{thm}
-
-\begin{proof}
-Note that the diagram
-\begin{equation*}
-\begin{tikzcd}
-\stesysf_3
-  \ar{d}[swap]{\pullbackpr{2}{\efamext}{\eft[2]}}
-  \ar{r}{\eext{2}}
-  & 
-\stesysf_2 
-  \ar{d}{\efamext} 
-  \\
-\stesysf_2 
-  \ar{r}[swap]{\efamext} 
-  &
-\stesysf
-\end{tikzcd}
-\end{equation*}
-commutes by assumption. For the second condition, we have to show that the
-diagram
-\begin{equation*}
-\begin{tikzcd}
-\stesysf_4
-  \ar{d}[swap]{\pullbackpr{2}{\eext{2}}{\eft[3]}}
-  \ar{r}{\eext{3}}
-  & 
-\stesysf_3
-  \ar{d}{\eext{2}} 
-  \\
-\stesysf_3
-  \ar{r}[swap]{\eext{2}} 
-  &
-\stesysf_2
-\end{tikzcd}
-\end{equation*}
-Since this is a question about two maps into a pullback, it suffices to verify
-that
-\begin{align*}
-\pullbackpr{1}{\ectxext}{\eft}\circ\eext{2}\circ\eext{3}
-  & =
-\pullbackpr{1}{\ectxext}{\eft}\circ\eext{2}\circ\pullbackpr{2}{\eext{2}}{\eft[3]}
-  \\
-\pullbackpr{2}{\ectxext}{\eft}\circ\eext{2}\circ\eext{3}
-  & =
-\pullbackpr{2}{\ectxext}{\eft}\circ\eext{2}\circ\pullbackpr{2}{\eext{2}}{\eft[3]}.
-\end{align*}
-For the first equality, it is fairly straightforward to show that both the
-equalities
-\begin{equation*}
-\pullbackpr{1}{\ectxext}{\eft}\circ\eext{2}\circ\eext{3}
-  =
-\eft[2]\circ\eft[3]\circ\eft[4]
-\end{equation*}
-and
-\begin{equation*}
-\pullbackpr{1}{\ectxext}{\eft}\circ\eext{2}\circ\pullbackpr{2}{\eext{2}}{\eft[3]}
-  =
-\eft[2]\circ\eft[3]\circ\eft[4].
-\end{equation*}
-hold. For the second subgoal (which is more tricky). Notice first that the
-diagram
-\begin{equation*}
-\begin{tikzcd}
-\stesysf_4
-  \ar{r}{\eext{3}}
-  \ar{d}[swap]{\beta_3}
-  &
-\stesysf_3
-  \ar{r}{\eext{2}}
-  \ar{d}[swap]{\beta_2}
-  &
-\stesysf_2
-  \ar{d}{\pullbackpr{2}{\ectxext}{\eft}}
-  \\
-\stesysf_3
-  \ar{r}[swap]{\eext{2}}
-  &
-\stesysf_2
-  \ar{r}[swap]{\efamext}
-  &
-\stesysf
-\end{tikzcd}
-\end{equation*}
-commutes. We also have the commutative diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=large]
-\stesysf_4
-  \ar{r}{\pullbackpr{2}{\eext{2}}{\eft[3]}}
-  \ar{d}[swap]{\beta_3}
-  &
-\stesysf_3
-  \ar{r}{\eext{2}}
-  \ar{d}[swap]{\beta_2}
-  &
-\stesysf_2
-  \ar{d}{\pullbackpr{2}{\ectxext}{\eft}}
-  \\
-\stesysf_3
-  \ar{r}{\pullbackpr{2}{\efamext}{\eft[2]}}
-  \ar{dr}[swap]{\eext{2}}
-  &
-\stesysf_2
-  \ar{r}{\efamext}
-  &
-\stesysf
-  \\
-  {} &
-\stesysf_2
-  \ar{ur}[swap]{\efamext}
-\end{tikzcd}
-\end{equation*}
-completing the proof.
-\end{proof}
-
-\subsection{(Pre-)extension homomorphisms}\label{subsection:e_extension_homomorphisms}
-In this subsection we start with the study of pre-extension homomorphisms, which
-will include the extension homomorphisms since they will be the pre-extension
-homomorphisms of which both the domain and codomain are extension objects.
-Our main examples of extension homomorphisms will be the operations of weakening
-and substitution. There are some basic examples of pre-extension homomorphisms
-that will be useful too, which get introduced in the this section and in
-\autoref{subsection:change_of_base}. In this section, we will mainly be
-interested in pre-extension homomorphisms between local pre-extension objects.
-We will end this section by proving that a retract of an extension object is
-always an extension object.
-
-\begin{defn}
-Let $\stesys$ and $\stesys'$ be pre-extension algebras. A \emph{pre-extension 
-homomorphism $f$ from $\stesys'$ to $\stesys$} is a triple $(f_0,f_1,f^t)$ 
-consisting of morphisms
-\begin{equation*}
-\begin{tikzcd}
-\stesyst' 
-  \ar{r}{f^t}
-  \ar{d}[swap]{\ebd'}
-  &
-\stesyst
-  \ar{d}{\ebd}
-  \\
-\stesysf'
-  \ar{r}{f_1}
-  \ar{d}[swap]{\eft'}
-  &
-\stesysf
-  \ar{d}{\eft}
-  \\
-\stesysc' 
-  \ar{r}[swap]{f_0}
-  &
-\stesysc
-\end{tikzcd}
-\end{equation*}
-such that the indicated squares commute, for which furthermore the squares
-\begin{equation*}
-\begin{tikzcd}
-\stesysf' 
-  \ar{r}{f_1}
-  \ar{d}[swap]{\ectxext'}
-  &
-\stesysf
-  \ar{d}{\ectxext}
-  \\
-\stesysc'
-  \ar{r}[swap]{f_0}
-  &
-\stesysc
-\end{tikzcd}
-\end{equation*}
-and
-\begin{equation*}
-\begin{tikzcd}[column sep=huge]
-\stesysf'\times_{\ectxext',\eft'} \stesysf'
-  \ar{r}{f_1\times_{\ectxext,\eft} f_1}
-  \ar{d}[swap]{\efamext'}
-  &
-\stesysf\times_{\ectxext,\eft} \stesysf
-  \ar{d}{\efamext}
-  \\
-\stesysf'
-  \ar{r}[swap]{f_1}
-  &
-\stesysf
-\end{tikzcd}
-\end{equation*}
-Composition and the identity homomorphism are defined in the expected way. We
-define furthermore
-\begin{align*}
-f_2 & := \pullback{f_1}{f_1}{\ectxext}{\eft}
-  \\
-f_3 & := \pullback{f_2}{f_2}{\efamext}{\eft[2]}.
-\end{align*}
-\end{defn}
-
-\begin{defn}
-A pre-extension homomorphism between extension objects is called an extension
-homomorphism.
-\end{defn}
-
-\begin{lem}
-Let $\stesys$ be an extension object. Then
-\begin{equation*}
-\begin{tikzcd}[column sep=huge]
-\stesyst_2
-  \ar{r}{\pullbackpr{2}{\ectxext}{\eft\circ\ebd}}
-  \ar{d}[swap]{\ebd[2]}
-  &
-\stesyst
-  \ar{d}{\ebd}
-  \\
-\stesysf_2
-  \ar{r}{\pullbackpr{2}{\ectxext}{\eft}}
-  \ar{d}[swap]{\eft[2]}
-  &
-\stesysf
-  \ar{d}{\eft}
-  \\
-\stesysf
-  \ar{r}[swap]{\ectxext}
-  &
-\stesysc
-\end{tikzcd}
-\end{equation*}
-assembles an extension homomorphism $\mathbf{e}_0:\famesys{\stesys}\to\stesys$.
-Likewise, we have an extension homomorphism
-$\mathbf{e}_1:\famesys{\famesys{\stesys}}\to\famesys{\stesys}$. Thus, a
-pre-extension object is an extension object if and only if $\mathbf{e}_0$
-and $\mathbf{e}_1$ are pre-extension homomorphisms.
-\end{lem}
-
-\begin{proof}
-Immediate from the conditions of being an extension object.
-\end{proof}
-
-\begin{defn}
-Let $\stesys$ be a pre-extension object. Then
-\begin{equation*}
-\begin{tikzcd}
-\stesyst_3
-  \ar{r}{\beta^t}
-  \ar{d}[swap]{\ebd[3]}
-  &
-\stesyst_2
-  \ar{d}{\ebd[2]}
-  \\
-\stesysf_3
-  \ar{r}{\beta_2}
-  \ar{d}[swap]{\eft[3]}
-  &
-\stesysf_2
-  \ar{d}{\eft[2]}
-  \\
-\stesysf_2
-  \ar{r}[swap]{\beta_1}
-  &
-\stesysf
-\end{tikzcd}
-\qquad
-\text{and}
-\qquad
-\begin{tikzcd}
-\stesyst_4
-  \ar{r}{\beta^t_2}
-  \ar{d}[swap]{\ebd[4]}
-  &
-\stesyst_3
-  \ar{d}{\ebd[3]}
-  \\
-\stesysf_4
-  \ar{r}{\beta_3}
-  \ar{d}[swap]{\eft[4]}
-  &
-\stesysf_3
-  \ar{d}{\eft[3]}
-  \\
-\stesysf_3
-  \ar{r}[swap]{\beta_2}
-  &
-\stesysf_2
-\end{tikzcd}
-\end{equation*}
-assemble pre-extension homomorphisms 
-\(
-\boldsymbol{\beta}
-  :
-\famesys{\famesys{\stesys}}
-  \to
-\famesys{\stesys}
-\) 
-and
-\(
-\boldsymbol{\beta}_\mathbf{2}
-  :
-\famesys{\famesys{\famesys{\stesys}}}
-  \to
-\famesys{\famesys{\stesys}}
-\).
-\end{defn}
-
-\begin{defn}\label{famehom}
-Suppose that $f:\stesys'\to\stesys$ is a pre-extension homomorphism. Then we
-define $\famehom{f}:\famesys{\stesys'}\to\famesys{\stesys}$ to consist of
-\begin{equation*}
-\begin{tikzcd}
-\stesyst_2'
-  \ar{r}{f^t_2}
-  \ar{d}[swap]{\ebd[2]'}
-  &
-\stesyst_2
-  \ar{d}{\ebd[2]}
-  \\
-\stesysf_2'
-  \ar{r}{f_2}
-  \ar{d}[swap]{\eft[2]'}
-  &
-\stesysf_2
-  \ar{d}{\eft[2]}
-  \\
-\stesysf'
-  \ar{r}[swap]{f_1}
-  &
-\stesysf
-\end{tikzcd}
-\end{equation*}
-where we define
-\begin{equation*}
-f^t_2 := \pullback{f_1}{f^t}{\ectxext}{\eft\circ\ebd}.
-\end{equation*}
-\end{defn}
-
-\begin{lem}
-The triple $\famehom{f}$ defined in \autoref{famehom} is a pre-extension homomorphism.
-\end{lem}
-
-\begin{proof}
-Note that the square
-\begin{equation*}
-\begin{tikzcd}
-\stesysf_2'
-  \ar{r}{f_2}
-  \ar{d}[swap]{\efamext'}
-  &
-\stesysf_2
-  \ar{d}{\efamext}
-  \\
-\stesysf'
-  \ar{r}[swap]{f_1}
-  &
-\stesysf
-\end{tikzcd}
-\end{equation*}
-commutes by assumption. Thus, it remains to show that the square
-\begin{equation*}
-\begin{tikzcd}
-\stesysf_3'
-  \ar{r}{f_3}
-  \ar{d}[swap]{\eext{2}'}
-  &
-\stesysf_3
-  \ar{d}{\eext{2}}
-  \\
-\stesysf_2'
-  \ar{r}[swap]{f_2}
-  &
-\stesysf_2
-\end{tikzcd}
-\end{equation*}
-commutes. It is equivalent to show that the equalities
-\begin{align*}
-\pullbackpr{1}{\ectxext}{\eft}\circ f_2\circ\eext{2}'
-  & =
-\pullbackpr{1}{\ectxext}{\eft}\circ \eext{2}\circ f_3
-  \\
-\pullbackpr{2}{\ectxext}{\eft}\circ f_2\circ\eext{2}'
-  & =
-\pullbackpr{2}{\ectxext}{\eft}\circ \eext{2}\circ f_3
-\end{align*}
-both hold. For the first, it is straightforward to verify that the diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=large]
-{} &
-\stesysf_2'
-  \ar{r}{f_2}
-  \ar{dr}[near end]{\pullbackpr{1}{\ectxext'}{\eft'}}
-  &
-\stesysf_2
-  \ar{dr}{\pullbackpr{1}{\ectxext}{\eft}}
-  \\
-\stesysf_3'
-  \ar{ur}{\eext{2}'}
-  \ar{r}[swap]{\beta_2'}
-  \ar{ddr}[swap]{f_3}
-  &
-\stesysf_2'
-  \ar{r}{\efamext'}
-  \ar{dr}[swap]{f_2}
-  &
-\stesysf'
-  \ar{r}{f_1}
-  &
-\stesysf
-  \\
-{} & {} &
-\stesysf_2
-  \ar{ur}[near start]{\efamext}
-  \\
-{} &
-\stesysf_3
-  \ar{r}[swap]{\eext{2}}
-  \ar{ur}{\beta_2}
-  &
-\stesysf_2
-  \ar{uur}[swap]{\pullbackpr{1}{\ectxext}{\eft}}
-\end{tikzcd}
-\end{equation*}
-commutes. For the second, note that the diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=large]
-{} &
-\stesysf_2'
-  \ar{r}{f_2}
-  \ar{dr}{\beta_1'}
-  &
-\stesysf_2
-  \ar{ddr}{\pullbackpr{2}{\ectxext}{\eft}}
-  \\
-{} & {} &
-\stesysf'
-  \ar{dr}[swap,near start]{f_1}
-  \\
-\stesysf_3'
-  \ar{uur}{\eext{2}'}
-  \ar{r}{\beta_2'}
-  \ar{dr}[swap]{f_3}
-  &
-\stesysf_2'
-  \ar{r}{f_2}
-  \ar{ur}{\efamext'}
-  &
-\stesysf_2
-  \ar{r}[swap]{\efamext}
-  &
-\stesysf
-  \\
-{} &
-\stesysf_3
-  \ar{r}[swap]{\eext{2}}
-  \ar{ur}[near start]{\beta_2}
-  &
-\stesysf_2
-  \ar{ur}[swap]{\pullbackpr{2}{\ectxext}{\eft}}
-\end{tikzcd}
-\end{equation*}
-commutes.
-\end{proof}
-
-\begin{lem}[Stability under retracts]\label{esys-retract}
-Suppose $f:\stesys\to\stesys'$ is a pre-extension homomorphism between
-pre-extension objects. If there is a pre-extension homomorphism $g:\stesys'\to
-\stesys$ such that $g\circ f=\catid{\stesys}$ and $\stesys'$ is an extension
-algebra, then $\stesys$ is an extension algebra.
-\end{lem}
-
-Before we start with the proof, note that we have the equalities
-$g_2\circ f_2=\catid{\stesysf_2}$ and $g_3\circ f_3=\catid{\stesysf_3}$
-under the hypotheses of the lemma.
-
-\begin{proof}
-Our first subgoal is to show that the square
-\begin{equation*}
-\begin{tikzcd}
-\stesysf_2 
-  \ar{r}{\efamext} 
-  \ar{d}[swap]{\pullbackpr{2}{\ectxext}{\eft}} 
-  & 
-\stesysf 
-  \ar{d}{\ectxext}
-  \\
-\stesysf
-  \ar{r}[swap]{\ectxext} 
-  & 
-\stesysc
-\end{tikzcd}
-\end{equation*}
-commutes. Note that in the diagram
-\begin{equation*}
-\begin{tikzcd}
-  {}
-  & 
-\stesysf
-  \ar{dd}[near start]{\ectxext}
-  \ar{rr}{f_1}
-  &
-  &
-\stesysf'
-  \ar{dd}[near start]{\ectxext'}
-  \ar{rr}{g_1}
-  &
-  &
-\stesysf
-  \ar{dd}{\ectxext}
-  \\
-\stesysf_2
-  \ar{dd}[swap]{\pullbackpr{2}{\ectxext}{\eft}}
-  \ar[crossing over]{rr}[swap,near start]{f_2}
-  \ar{ur}{\efamext}
-  &
-  &
-\stesysf_2'
-  \ar{ur}[near start]{\efamext'}
-  \ar[crossing over]{rr}[swap,near start]{g_2}
-  &
-  &
-\stesysf_2
-  \ar{ur}[swap,near start]{\efamext}
-  \\
-  {}
-  &
-\stesysc
-  \ar{rr}[near start]{f_0}
-  &
-  &
-\stesysc'
-  \ar{rr}[near start]{g_0}
-  &
-  &
-\stesysc
-  \\
-\stesysf 
-  \ar{rr}[swap]{f_1}
-  \ar{ur}{\ectxext}
-  &
-  &
-\stesysf' 
-  \ar[crossing over,leftarrow]{uu}[near end,swap]{\pullbackpr{2}{\ectxext'}{\eft'}}
-  \ar{ur}[swap,near end]{\ectxext'}
-  \ar{rr}[swap]{g_1}
-  &
-  &
-\stesysf
-  \ar[crossing over,leftarrow]{uu}[near end,swap]{\pullbackpr{2}{\ectxext}{\eft}}
-  \ar{ur}[swap]{\ectxext}
-\end{tikzcd}
-\end{equation*}
-all the faces minus the far left and far right face commute. Using that $g$
-is a section of $f$, we can read off that also the far left face commutes,
-completing our first subgoal.
-
-For the second subgoal, note that also $\famehom{g}\circ\famehom{f}=
-\catid{\famesys{\stesys}}$ and that $\famesys{\stesys'}$ is an extension object.
-Thus we can apply what we have proven so far to conclude that the square
-\begin{equation*}
-\begin{tikzcd}
-\stesysf_3 
-  \ar{r}{\eext{2}} 
-  \ar{d}[swap]{\pullbackpr{2}{\efamext}{\eft[2]}} 
-  & 
-\stesysf_2 
-  \ar{d}{\efamext}
-  \\
-\stesysf_2
-  \ar{r}[swap]{\efamext} 
-  & 
-\stesysf
-\end{tikzcd}
-\end{equation*}
-commutes.
-\end{proof}
-
-\subsection{The change of base of (pre-)extension objects}
-\label{subsection:change_of_base}
-An important construction of (pre-)extension objects is the change of base. It
-allows us to consider `parametrized homomorphisms', such as weakening and
-substitution.
-
-\begin{defn}
-Suppose $f:\stesys\to\stesys'$ is a pre-extension homomorphism. We say that
-a diagram
-\begin{equation*}
-\begin{tikzcd}
-\stesys
-  \ar{r}{f}
-  \ar{d}[swap]{p}
-  &
-\stesys'
-  \ar{d}{p'}
-  \\
-X \ar{r}[swap]{g}
-  &
-Y
-\end{tikzcd}
-\end{equation*}
-commutes if the diagram
-\begin{equation*}
-\begin{tikzcd}
-\stesysc
-  \ar{r}{f_0}
-  \ar{d}[swap]{p}
-  &
-\stesysc'
-  \ar{d}{p'}
-  \\
-X \ar{r}[swap]{g}
-  &
-Y
-\end{tikzcd}
-\end{equation*}
-commutes.
-\end{defn}
-
-The first goal in this subsection is to define for every (pre-)extension object 
-$\stesys$ and every $p:\stesysc\rightarrow X\leftarrow Y:g$, a (pre-)extension
-object $\cobesys{Y}{\stesys}{g}{p}$ with a homomorphism $\pullbackpr{2}{g}{p}:
-\cobesys{Y}{\stesys}{g}{p}\to\stesys$ and a morphism $\pullbackpr{1}{g}{p}:
-\pullback{Y}{\stesysc}{g}{p}\to Y$ such that for every diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=large]
-\stesys'
-  \ar[bend right=10]{ddr}[swap]{p'}
-  \ar[bend left=10]{rrd}{f}
-  \ar[dotted]{dr}[near end]{[p',f]}
-  \\
-  {}&
-\cobesys{Y}{\stesys}{g}{p}
-  \ar{d}{\pullbackpr{1}{g}{p}}
-  \ar{r}[swap]{\pullbackpr{2}{g}{p}}
-  &
-\stesys
-  \ar{d}{p}
-  \\
-  {}&
-Y \ar{r}[swap]{g}
-  &
-X
-\end{tikzcd}
-\end{equation*}
-of which the outer square commutes, the (pre-)extension homomorphism $[p',f]$ exists
-and is unique with the property that it renders the diagram commutative. We will
-give the definition of $\cobesys{Y}{\stesys}{g}{p}$ in \autoref{cobesys}. After
-proving that the change of base of a pre-extension algebra is indeed a
-pre-extension algebra (\autoref{cobesys-preext}) and that the change of base
-of an extension algebra is an extension algebra (\autoref{cobesys-ext}), we
-will demonstrate the above unique existence in \autoref{cobesys-existence,%
-cobesys-pullback}.
-
-The second goal in this subsection is to follow the same procedure for
-$\famesys{\famesys{\stesys}}$ to show that it is equivalent to
-$\cobesys{\stesysf}{\famesys{\stesys}}{\ectxext}{\eft}$. We will do this by
-verifying directly that it has the universal property of the change of base
-described above, because we will use the ingredients in our definition of
-weakening and substitution objects.
-
-\begin{defn}[Change of base]\label{cobesys}
-Suppose $\stesys$ is a pre-extension object in $\cat{C}$ and that 
-$p:\stesysc\rightarrow X\leftarrow Y:g$.
-Then we define the pre-extension object $\cobesys{Y}{\stesys}{g}{p}$ to consist of
-\begin{equation*}
-\begin{tikzcd}
-\cobesys{Y}{\stesyst}{g}{p\circ\eft\circ\ebd}
-  \ar{r}
-  \ar{d}[swap]{g^\ast(\ebd)}
-  &
-\stesyst
-  \ar{d}{\ebd}
-  \\
-\cobesys{Y}{\stesysf}{g}{p\circ\eft}
-  \ar{r}
-  \ar{d}[swap]{g^\ast(\eft)}
-  &
-\stesysf
-  \ar{d}{\eft}
-  \\
-\cobesys{Y}{\stesysc}{g}{p}
-  \ar{r}
-  \ar{d}[swap]{\pullbackpr{1}{g}{p}}
-  &
-\stesysc
-  \ar{d}{p}
-  \\
-Y \ar{r}[swap]{g}
-  &
-X
-\end{tikzcd}
-\end{equation*} 
-and the operations
-\begin{align*}
-\cobesys{Y}{\ectxext}{g}{p} 
-  & : \pullback{Y}{\stesysf}{g}{p\circ\eft}\to \pullback{Y}{\stesysc}{g}{p}\\
-\cobesys{Y}{\efamext}{g}{p} 
-  & : \pullback
-    {\pullback{Y}{\stesysf}{g}{p\circ\eft}}
-    {\pullback{Y}{\stesysf}{g}{p\circ\eft}}
-    {\cobesys{Y}{\ectxext}{g}{p}}
-    {g^\ast(\eft)}
-  \to 
-  \pullback{Y}{\stesysf}{g}{p\circ\eft}.
-\end{align*}
-defined by
-\begin{equation*}
-\cobesys{Y}{\ectxext}{g}{p} := \pullback{\catid{Y}}{\ectxext}{g}{p}
-\end{equation*}
-and where $\cobesys{Y}{\efamext}{g}{p}$ is defined by rendering the diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=large]
-(\cobesys{Y}{\stesysf}{g}{p\circ\eft})_2
-  \ar{rr}{\pullback{\pullbackpr{2}{g}{p\circ\eft}}{\pullbackpr{2}{g}{p\circ\eft}}{\ectxext}{\eft}}
-  \ar{dd}[swap]{\pullbackpr{1}{\cobesys{Y}{\ectxext}{g}{p}}{g^\ast(\eft)}}
-  \ar[dotted]{dr}[swap]{\cobesys{Y}{\efamext}{g}{p}}
-  &
-  &
-\stesysf_2
-  \ar{d}{\efamext}
-  \\
-  {}&
-\cobesys{Y}{\stesysf}{g}{p\circ\eft}
-  \ar{r}{\pullbackpr{2}{g}{p\circ\eft}}
-  \ar{d}[swap]{\pullbackpr{1}{g}{p\circ\eft}}
-  &
-\stesysf
-  \ar{d}{p\circ\eft}
-  \\
-\cobesys{Y}{\stesysf}{g}{p\circ\eft}
-  \ar{r}[swap]{\pullbackpr{1}{g}{p\circ\eft}}
-  &
-Y \ar{r}[swap]{g}
-  &
-X
-\end{tikzcd}
-\end{equation*} 
-commutative. 
-The process of obtaining the pre-extension object $\cobesys{Y}{\stesys}{g}{p}$ out of $\stesys$
-and $g:Y\to X$ is also called the \emph{change of base}.
-\end{defn}
-
-\begin{lem}\label{cobesys-preext}
-Any change of base of a pre-extension object is a pre-extension object.
-\end{lem}
-
-\begin{proof}
-Let $\stesys$ be an extension algebra and consider $p:\stesysc\rightarrow X\leftarrow Y:g$.
-We need to verify that the square
-\begin{equation*}
-\begin{tikzcd}[column sep=large]
-(\pullback{Y}{\stesysf}{g}{p\circ\eft})_2
-  \ar{r}{\cobesys{Y}{\efamext}{g}{p}} 
-  \ar{d}[swap]{\pullbackpr{1}{\cobesys{Y}{\ectxext}{g}{p}}{g^\ast(\eft)}} 
-  & 
-\pullback{Y}{\stesysf}{g}{p\circ\eft}
-  \ar{d}{g^\ast(\eft)}
-  \\
-\pullback{Y}{\stesysf}{g}{p\circ\eft}
-  \ar{r}[swap]{g^\ast(\eft)} 
-  & 
-\pullback{Y}{\stesysc}{g}{p}
-\end{tikzcd}
-\end{equation*}
-commutes. It is fairly obvious that
-\begin{equation*}
-\pullbackpr{1}{g}{p}\circ g^\ast(\eft)\circ (\cobesys{Y}{\efamext}{g}{p})
-  =
-\pullbackpr{1}{g}{p\circ\eft}\circ \pullbackpr{1}{\cobesys{Y}{\ectxext}{g}{p}}{g^\ast(\eft)}
-\end{equation*}
-and that the diagram
-\begin{equation*}
-\begin{tikzcd}
-  {}&
-  {}&
-\pullback{Y}{\stesysf}{g}{p\circ\eft}
-  \ar{rr}{g^\ast(\eft)}
-  \ar{dr}[swap]{\pullbackpr{2}{g}{p\circ\eft}}
-  &
-  {}&
-\pullback{Y}{\stesysc}{g}{p}
-  \ar{ddr}{\pullbackpr{2}{g}{p}}
-  \\
-  {}&
-  {}&
-  {}&
-\stesysf
-  \ar{drr}[swap]{\eft}
-  \\
-(\pullback{y}{\stesysf}{g}{p\circ\eft})_2
-  \ar{uurr}{\cobesys{Y}{\efamext}{g}{p}}
-  \ar{rr}[swap,yshift=-.5ex]{\pullback{\pullbackpr{2}{g}{p\circ\eft}}{\pullbackpr{2}{g}{p\circ\eft}}{\ectxext}{\eft}}
-  \ar{ddrr}[swap]{\pullbackpr{1}{\cobesys{Y}{\ectxext}{g}{p}}{g^\ast(\eft)}}
-  &
-  {}&
-\stesysf_2
-  \ar{ur}{\efamext}
-  \ar{dr}[swap]{\eft[2]}
-  &
-  {}&
-  {}&
-\stesysc
-  \\
-  {}&
-  {}&
-  {}&
-\stesysf
-  \ar{urr}{\eft}
-  \\
-  {}&
-  {}&
-\pullback{Y}{\stesysf}{g}{p\circ\eft}
-  \ar{rr}[swap]{g^\ast(\eft)}
-  \ar{ur}{\pullbackpr{2}{g}{p\circ\eft}}
-  &
-  {}&
-\pullback{Y}{\stesysc}{g}{p}
-  \ar{uur}[swap]{\pullbackpr{2}{g}{p}}
-\end{tikzcd}
-\end{equation*}
-commutes.
-\end{proof}
-
-\begin{thm}\label{cobesys-ext}
-The change of base of an extension algebra is an extension algebra.
-\end{thm}
-
-\begin{proof}
-Our first subgoal is to verify that the square
-\begin{equation*}
-\begin{tikzcd}[column sep=large]
-(\pullback{Y}{\stesysf}{g}{p\circ\eft})_2
-  \ar{r}{\cobesys{Y}{\efamext}{g}{p}} 
-  \ar{d}[swap]{\pullbackpr{2}{\cobesys{Y}{\ectxext}{g}{p}}{g^\ast(\eft)}} 
-  & 
-\pullback{Y}{\stesysf}{g}{p\circ\eft}
-  \ar{d}{\cobesys{Y}{\ectxext}{g}{p}}
-  \\
-\pullback{Y}{\stesysf}{g}{p\circ\eft}
-  \ar{r}[swap]{\cobesys{Y}{\ectxext}{g}{p}} 
-  & 
-\pullback{Y}{\stesysc}{g}{p}
-\end{tikzcd}
-\end{equation*}
-\end{proof}
-
-The following construction is useful for defining extension homomorphisms into
-`higher' extension objects
-
-\begin{defn}\label{cobesys-existence}
-Consider a commutative diagram
-\begin{equation*}
-\begin{tikzcd}
-\stesys'
-  \ar{r}{f}
-  \ar{d}[swap]{p'}
-  &
-\stesys
-  \ar{d}{p}
-  \\
-Y \ar{r}[swap]{g}
-  &
-X
-\end{tikzcd}
-\end{equation*}
-Then we construct $[p,f]:\stesys'\to\cobesys{Y}{\stesys}{g}{p}$
-\begin{itemize}
-\item by defining $[p,f]_0:\stesysc'\to\pullback{Y}{\stesysc}{g}{p}$ be the uniqe
-morphism rendering the diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=large]
-\stesysc'
-  \ar[bend right=10]{ddr}[swap]{p'}
-  \ar[bend left=10]{rrd}{f_0}
-  \ar{dr}[near end]{[p,f]_0}
-  \\
-  {}&
-\pullback{Y}{\stesysc}{g}{p}
-  \ar{r}[swap]{\pullbackpr{2}{g}{p}}
-  \ar{d}{\pullbackpr{1}{g}{p}}
-  &
-\stesysc
-  \ar{d}{p}
-  \\
-  {}&
-Y \ar{r}[swap]{g}
-  &
-X
-\end{tikzcd}
-\end{equation*}
-commutative.
-\item by defining $[p,f]_1:\stesysf'\to\pullback{Y}{\stesysf}{g}{p\circ\eft}$ be the uniqe
-morphism rendering the diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=huge]
-\stesysf'
-  \ar[bend right=10]{ddr}[swap]{p'\circ\eft'}
-  \ar[bend left=10]{rrd}{f_1}
-  \ar{dr}[near end]{[p,f]_1}
-  \\
-  {}&
-\pullback{Y}{\stesysf}{g}{p\circ\eft}
-  \ar{r}[swap]{\pullbackpr{2}{g}{p\circ\eft}}
-  \ar{d}{\pullbackpr{1}{g}{p\circ\eft}}
-  &
-\stesysc
-  \ar{d}{p\circ\eft}
-  \\
-  {}&
-Y \ar{r}[swap]{g}
-  &
-X
-\end{tikzcd}
-\end{equation*}
-commutative.
-\item by defining $[p,f]^t:\stesyst'\to\pullback{Y}{\stesyst}{g}{p\circ\eft\circ\ebd}$ be the uniqe
-morphism rendering the diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=huge]
-\stesyst'
-  \ar[bend right=10]{ddr}[swap]{p'\circ\eft'\circ\ebd'}
-  \ar[bend left=10]{rrd}{f^t}
-  \ar{dr}[near end]{[p,f]^t}
-  \\
-  {}&
-\pullback{Y}{\stesysf}{g}{p\circ\eft\circ\ebd}
-  \ar{r}[swap]{\pullbackpr{2}{g}{p\circ\eft\circ\ebd}}
-  \ar{d}{\pullbackpr{1}{g}{p\circ\eft\circ\ebd}}
-  &
-\stesysc
-  \ar{d}{p\circ\eft\circ\ebd}
-  \\
-  {}&
-Y \ar{r}[swap]{g}
-  &
-X
-\end{tikzcd}
-\end{equation*}
-commutative.
-\end{itemize}
-\end{defn}
-
-\begin{thm}\label{cobesys-pullback}
-For every diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=large]
-\stesys'
-  \ar[bend right=10]{ddr}[swap]{p'}
-  \ar[bend left=10]{rrd}{f}
-  \ar[dotted]{dr}[near end]{[p',f]}
-  \\
-  {}&
-\cobesys{Y}{\stesys}{g}{p}
-  \ar{d}{\pullbackpr{1}{g}{p}}
-  \ar{r}[swap]{\pullbackpr{2}{g}{p}}
-  &
-\stesys
-  \ar{d}{p}
-  \\
-  {}&
-Y \ar{r}[swap]{g}
-  &
-X
-\end{tikzcd}
-\end{equation*}
-of which the outer square commutes, the pre-extension homomorphism $[p',f]$
-is unique with the property that it renders the whole diagram commutative.
-\end{thm}
-
-\begin{defn}\label{famfamstesys_into}
-Consider a commutative square
-\begin{equation*}
-\begin{tikzcd}
-\stesys'
-  \ar{r}{f}
-  \ar{d}[swap]{p}
-  &
-\famesys\stesys
-  \ar{d}{\eft}
-  \\
-\stesysf \ar{r}[swap]{\ectxext}
-  &
-\stesysc
-\end{tikzcd}
-\end{equation*}
-Then we construct
-\begin{equation*}
-[p,f]:\stesys'\to\famesys{\famesys{\stesys}}
-\end{equation*}
-as follows:
-\begin{itemize}
-\item let $[p,f]_0:\stesysc'\to\stesysf_2$ be the unique morphism rendering
-the diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=large]
-\stesysc' 
-  \ar[bend left=10]{rrd}{f_0}
-  \ar[swap,bend right=10]{ddr}{p}
-  \ar[dotted]{dr}[near end]{[p,f]_0}
-  \\
-  {}&
-\stesysf_2
-  \ar{r}[swap]{\pullbackpr{2}{\ectxext}{\eft}}
-  \ar{d}{\eft[2]}
-  &
-\stesysf
-  \ar{d}{\eft}
-  \\
-  {}&
-\stesysf
-  \ar{r}[swap]{\ectxext}
-  &
-\stesysc
-\end{tikzcd}
-\end{equation*}
-commutative.
-\item Let $[p,f]_1:\stesysf'\to\stesysf_3$ be the unique morphism rendering
-the diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=large]
-\stesysf'
-  \ar[bend left=10]{drr}{f_1}
-  \ar[swap]{dd}{\eft'}
-  \ar[dotted]{dr}[near end]{[p,f]_1}
-  \\
-  {}&
-\stesysf_3
-  \ar{r}[swap]{\pullbackpr{2}{\efamext}{\eft[2]}}
-  \ar{d}{\eft[2]}
-  &
-\stesysf_2
-  \ar{d}{\eft[2]}
-  \\
-\stesysc'
-  \ar{r}[swap]{[p,f]_0}
-  &
-\stesysf_2
-  \ar{r}[swap]{\efamext}
-  &
-\stesysf
-\end{tikzcd}
-\end{equation*}
-commutative.
-\item Let $[p,f]^t:\stesyst'\to\stesyst_3$ be the unique morphism rendering
-the diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=huge]
-\stesyst'
-  \ar[bend left=10]{drr}{f^t}
-  \ar[swap]{dd}{\eft'\circ\ebd'}
-  \ar[dotted]{dr}[near end]{[p,f]^t}
-  \\
-  {}&
-\stesyst_3
-  \ar{r}[swap]{\pullbackpr{2}{\efamext}{\eft[2]\circ\ebd[2]}}
-  \ar{d}[swap]{\pullbackpr{1}{\efamext}{\eft[2]\circ\ebd[2]}}
-  &
-\stesyst_2
-  \ar{d}{\eft[2]\circ\ebd[2]}
-  \\
-\stesysc'
-  \ar{r}[swap]{[p,f]_0}
-  &
-\stesysf_2
-  \ar{r}[swap]{\efamext}
-  &
-\stesysf
-\end{tikzcd}
-\end{equation*}
-commutative.
-\end{itemize}
-\end{defn}
-
-\begin{lem}
-Under the hypotheses of \autoref{famfamstesys_into}, $[p,f]$ is a pre-extension
-homomorphism. Moreover, it is the unique pre-extension homomorphism for which
-the diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=large]
-\stesys' 
-  \ar[bend left=10]{rrd}{f}
-  \ar[swap,bend right=10]{ddr}{p}
-  \ar[dotted]{dr}[near end]{[p,f]}
-  \\
-  {}&
-\famesys{\famesys{\stesys}}
-  \ar{r}[swap]{\pullbackpr{2}{\ectxext}{\eft}}
-  \ar{d}{\eft[2]}
-  &
-\famesys{\stesys}
-  \ar{d}{\eft}
-  \\
-  {}&
-\stesysf
-  \ar{r}[swap]{\ectxext}
-  &
-\stesysc
-\end{tikzcd}
-\end{equation*}
-commutes.
-\end{lem}
-
-\begin{lem}
-Suppose $f:\stesys\to \stesys'$ is a pre-extension homomorphism and consider a morphism
-$p:\stesys'\to X$ and $g:Y\to X$. Then the change of base 
-$g^\ast(f):\cobesys{Y}{\stesys}{g}{p\circ f_0}\to
-\cobesys{Y}{\stesys'}{g}{p}$ is a pre-extension morphism.
-\end{lem}
-
-\begin{lem}
-Let $\stesys$ be a pre-extension algebra and consider $p:\stesysc\rightarrow X\leftarrow Y:g$.
-Then there is an isomorphism
-\begin{equation*}
-\varphi:\famesys{\cobesys{Y}{\stesys}{g}{p}}
-  \simeq
-\cobesys{Y}{\famesys{\stesys}}{g}{p\circ\eft}
-\end{equation*}
-uniquely determined by
-\end{lem}
-
-\begin{proof}
-This follows from the pasting lemma for pullbacks.
-\end{proof}
-
-\subsection{Pre-weakening objects}
-\begin{defn}
-Let $\stesys$ be an extension object in $\cat{C}$. A pre-weakening operation
-on $\stesys$ is an extension homomorphism 
-$ \mathbf{w}(\stesys)
-    :
-  \cobesys{\stesysf}{\famesys{\stesys}}{\eft}{\eft}
-    \to
-  \famesys{\famesys{\stesys}}$
-for which the diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=large]
-\cobesys{\stesysf}{\famesys{\stesys}}{\eft}{\eft}
-  \ar{r}{\mathbf{w}(\stesys)}
-  \ar{dr}[swap]{\pullbackpr{1}{\eft}{\eft}}
-  &
-\famesys{\famesys{\stesys}}
-  \ar{d}{\eft[2]}
-  \\
-& \stesysf
-\end{tikzcd}
-\end{equation*}
-commutes.
-\end{defn}
-
-\begin{defn}
-Let $\stesys$ be an extension object with pre-weakening operation
-$\mathbf{w}(\stesys)$. Then $\famesys{\stesys}$ has the pre-weakening operation
-$\mathbf{w}(\famesys{\stesys})$ which is uniquely determined by rendering the
-diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=large]
-\cobesys{\stesysf_2}{\famesys{\famesys{\stesys}}}{\eft[2]}{\eft[2]}
-  \ar{rr}{%
-      \pullback{\beta_1}{\boldsymbol{\beta}}{\eft}{\eft}
-    }
-  \ar[bend right]{ddr}[swap]{\pullbackpr{1}{\eft[2]}{\eft[2]}}
-  \ar[dotted]{dr}{\mathbf{w}(\famesys{\stesys})}
-  &
-  {}&
-\cobesys{\stesysf}{\famesys{\stesys}}{\eft}{\eft}
-  \ar{r}{\mathbf{w}(\stesys)}
-  &
-\famesys{\famesys{\stesys}}
-  \ar{d}{\boldsymbol{\beta}}
-  \\
-  {}&
-\famesys{\famesys{\famesys{\stesys}}}
-  \ar{r}{\boldsymbol{\beta}_\mathbf{2}}
-  \ar{d}[swap]{\eft[3]}
-  &
-\famesys{\famesys{\stesys}}
-  \ar{d}{\eft[2]}
-  \ar{r}{\boldsymbol{\beta}}
-  &
-\famesys{\stesys}
-  \ar{d}{\eft}
-  \\
-  {}&
-\stesysf_2
-  \ar{r}[swap]{\efamext}
-  &
-\stesysf
-  \ar{r}[swap]{\ectxext}
-  &
-\stesysc
-\end{tikzcd}
-\end{equation*}
-commutative.
-\end{defn}
-
-\begin{defn}
-A pre-weakening object $\stesys$ in $\cat{C}$ is an extension object $\stesys$ 
-with a pre-weakening operation 
-$ \mathbf{w}(\stesys)
-    :
-  \cobesys{\stesysf}{\famesys{\stesys}}{\eft}{\eft}
-    \to
-  \famesys{\famesys{\stesys}}$
-for which the diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=15em]
-\cobesys{\stesysf_2}{\famesys{\stesys}}{\eft\circ\eft[2]}{\eft}
-  \ar[bend right=10]{dr}[swap]%
-    { [ \pullbackpr{1}{\eft\circ\eft[2]}{\eft},%
-        \mathbf{w}(\stesys)%
-          \circ%
-        (\pullback{\efamext}{\catid{\famesys{\stesys}}}{\eft}{\eft})%
-        ]%
-      }
-  \ar{r}{
-    [ \pullbackpr{1}{\eft\circ\eft[2]}{\eft},%
-      \mathbf{w}(\stesys)%
-        \circ%
-      (\pullback{\eft[2]}{\catid{\famesys{\stesys}}}{\eft}{\eft})%
-      ]}%
-  &
-\cobesys{\stesysf_2}{\famesys{\famesys{\stesys}}}{\eft[2]}{\eft[2]}
-  \ar{d}{\mathbf{w}(\famesys{\stesys})}
-  \\
-  {}&
-\famesys{\famesys{\famesys{\stesys}}}
-\end{tikzcd}
-\end{equation*}
-commutes. This condition is called \emph{Currying for weakening}.
-\end{defn}
-
-\begin{lem}
-If $\stesys$ is a pre-weakening algebra, then so is $\famesys{\stesys}$. 
-\end{lem}
-
-\begin{proof}
-\end{proof}
-
-\begin{defn}
-A pre-weakening morphism between preweakening objects $\stesys$ and $\stesys'$ is an
-extension homomorphism $f:\stesys\to \stesys'$ such that additionally the diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=large]
-\cobesys{\stesysf}{\famesys{\stesys}}{\eft}{\eft}
-  \ar{d}[swap]{\mathbf{w}(\stesys)}
-  \ar{r}{\pullback{f_1}{\famehom{f}}{\eft'}{\eft'}}
-  &
-\cobesys{\stesysf'}{\famesys{\stesys'}}{\eft'}{\eft'}
-  \ar{d}{\mathbf{w}(\stesys')}
-  \\
-\famesys{\famesys{\stesys}}
-  \ar{r}[swap]{\famehom{\famehom{f}}}
-  &
-\famesys{\famesys{\stesys'}}
-\end{tikzcd}
-\end{equation*}
-commutes.
-\end{defn}
-
-\begin{defn}
-Let $\stesys$ be a pre-weakening algebra and consider $p:\stesysc\rightarrow X\leftarrow Y:p$.
-Then we define
-\begin{equation*}
-\mathbf{w}(\cobesys{Y}{\stesys}{g}{p})
-  :
-\cobesys
-  { (\pullback{Y}{\stesysf}{g}{p\circ\eft})}
-  { \famesys{\cobesys{Y}{\stesys}{g}{p}}}
-  { g^\ast(\eft)}
-  { g^\ast(\eft)}
-  \to
-\famesys{\famesys{\cobesys{Y}{\stesys}{g}{p}}}
-\end{equation*}
-to be the unique extension homomorphism rendering the diagram
-\begin{equation*}
-\begin{tikzcd}
-\cobesys
-  { \pullback{Y}{\stesysf}{g}{p\circ\eft}}
-  { \famesys{\cobesys{Y}{\stesys}{g}{p}}}
-  { g^\ast(\eft)}
-  { g^\ast(\eft)}
-  \ar[bend right]{ddr}[swap]{\pullbackpr{1}{g^\ast(\eft)}{g^\ast(\eft)}}
-  \ar{rr}{%
-    \pullback
-      { \pullbackpr{2}{g}{p\circ\eft}}
-      { \boldsymbol{\pi}_\mathbf{2}(g,p\circ\eft)}
-      { \eft}
-      { \eft}
-    }
-  \ar[dotted]{dr}{\mathbf{w}(\cobesys{Y}{\stesys}{g}{p})}
-  &
-  {}&
-\cobesys{\stesysf}{\famesys{\stesys}}{\eft}{\eft}
-  \ar{r}{\mathbf{w}(\stesys)}
-  &
-\famesys{\famesys{\stesys}}
-  \ar{d}{\boldsymbol{\beta}}
-  \\
-  {}&
-\famesys{\famesys{\cobesys{Y}{\stesys}{g}{p}}}
-  \ar{r}{\beta}
-  \ar{d}{g^\ast(\eft)_1}
-  &
-\famesys{\cobesys{Y}{\stesys}{g}{p}}
-  \ar{r}{\boldsymbol{\pi}_2(g,p\circ\eft)}
-  \ar{d}{g^\ast(\eft)}
-  &
-\famesys{\stesys}
-  \ar{d}{\eft}
-  \\
-  {}&
-\pullback{Y}{\stesysf}{g}{p\circ\eft}
-  \ar{r}[swap]{\cobesys{Y}{\ectxext}{g}{p}}
-  &
-\pullback{Y}{\stesysc}{g}{p}
-  \ar{r}[swap]{\pullbackpr{2}{g}{p}}
-  &
-\stesysc
-\end{tikzcd}
-\end{equation*}
-commutative.
-\end{defn}
-
-\subsection{Weakening objects}
-Since we have shown that the property of being a pre-weakening object is closed
-under the relevant operations, we can make the following definition:
-
-\begin{defn}
-A weakening object is a pre-weakening object $\stesys$ with the property that
-$\mathbf{w}(\stesys)$ is a pre-weakening morphism.
-\end{defn}
-
-\begin{defn}
-A weakening homomorphism is a pre-weakening homomorphism such that the domain
-and codomain are weakening objects.
-\end{defn}
-
-\begin{thm}
-Suppose $\stesys$ is a weakening object, then so is $\famesys{\stesys}$
-\end{thm}
-
-\begin{thm}
-The change of base of any weakening object is again a weakening object.
-\end{thm}
-
-\subsection{Projection objects}
-\begin{defn}
-A pre-projection object is a weakening object $\stesys$ for which there is a term
-$\mathbf{i}:\stesysf\to \stesyst_2$ such that the diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=large]
-\stesysf \ar{r}{\mathbf{i}} \ar{d}[swap]{\Delta_{\eft}} & \stesyst_2 \ar{d}{\ebd[2]}\\
-\pullback{\stesysf}{\stesysf}{\eft}{\eft} \ar{r}[swap]{w(\stesys)_0} & \stesysf_2
-\end{tikzcd}
-\end{equation*}
-commutes. In this diagram $\Delta_{\eft}:\stesysf\to \pullback{\stesysf}{\stesysf}{\eft}{\eft}$ is the diagonal.
-\end{defn}
-
-\begin{defn}
-A pre-projection homomorphism from $\stesys$ to $\stesys'$ is a weakening homomorphism
-$f:\stesys\to \stesys'$ such that the square
-\begin{equation*}
-\begin{tikzcd}[column sep=large]
-\stesyst_2
-  \ar{r}{f^t_1}
-  &
-\stesyst_2'
-  \\
-\stesysf \ar{r}[swap]{f_1}
-  \ar{u}{\mathbf{i}}
-  &
-\stesysf'
-  \ar{u}[swap]{\mathbf{i}'}
-\end{tikzcd}
-\end{equation*}
-commutes
-\end{defn}
-
-\begin{lem}
-The change of base of a pre-projection object is again a pre-projection object.
-\end{lem}
-
-\begin{lem}
-If $CFT$ is a pre-projection object, then so is $\mathbf{F}_{CFT}$, where
-$\mathbf{F}_{\mathbf{i}}$ is defined to be $F\times_{e_0,c}\mathbf{i}$ is
-a pre-projection algebra.
-\end{lem}
-
-\begin{defn}
-A projection algebra is a pre-projection algebra for which weakening is a
-pre-projection homomorphism.
-\end{defn}
-
-\begin{cor}
-The change of base of a projection object is again a projection object.
-\end{cor}
-
-\begin{cor}
-If $CFT$ is a projection object, then so is $\mathbf{F}_{CFT}$, where
-$\mathbf{F}_{\mathbf{i}}$ is defined to be $F\times_{e_0,c}\mathbf{i}$ is
-a projection algebra.
-\end{cor}
-
-\subsection{Substitution objects}
-
-\begin{defn}
-A \emph{pre-substitution} for an extension object $\stesys$ is an
-extension homomorphism
-\begin{equation*}
-\mathbf{s}(\stesys):\cobesys{\stesyst}{\famesys{\famesys{\stesys}}}{\ebd}{\eft[2]}\to \famesys{\stesys}
-\end{equation*}
-for which the square
-\begin{equation*}
-\begin{tikzcd}[column sep=large]
-\pullback{\stesyst}{\stesysf_2}{\ebd}{\eft[2]}
-  \ar{r}{s(\stesys)_0}
-  \ar{d}[swap]{\ebd\circ\pullbackpr{1}{\ebd}{\eft[2]}}
-  &
-\stesysf 
-  \ar{d}{\eft}
-  \\
-\stesysf 
-  \ar{r}[swap]{\eft}
-  &
-\stesysc
-\end{tikzcd}
-\end{equation*}
-commutes. A \emph{pre-substitution object} is an extension object
-together with a pre-substitution.
-\end{defn}
-
-\begin{defn}
-A \emph{pre-substitution homomorphism} is an extension homomorphism $f:\stesys'\to \stesys$
-for which the square
-\begin{equation*}
-\begin{tikzcd}[column sep=huge]
-\cobesys{\stesyst'}{\famesys{\famesys{\stesys'}}}{\ebd'}{\eft[2]'}
-  \ar{r}{\pullback{f^t}{\famehom{\famehom{f}}}{\ebd}{\eft[2]}}
-  \ar{d}[swap]{\mathbf{s}'(\stesys')}
-  &
-\cobesys{\stesyst}{\famesys{\famesys{\stesys}}}{\ebd}{\eft[2]}
-  \ar{d}{\mathbf{s}(\stesys)}
-  \\
-\famesys{\stesys'}
-  \ar{r}[swap]{\famehom{f}}
-  &
-\famesys{\stesys}
-\end{tikzcd}
-\end{equation*}
-commutes.
-\end{defn}
-
-\begin{lem}
-The change of base of a pre-substitution object is again a pre-substitution object.
-\end{lem}
-
-\begin{lem}
-If $\stesys$ is a pre-substitution object, then so is $\famesys{\stesys}$ with
-$\mathbf{s}(\famesys{\stesys})$ defined to be the unique extension homomorphism
-rendering the diagram
-\begin{equation*}
-\begin{tikzcd}
-\cobesys{\stesyst_2}{\famesys{\famesys{\famesys{\stesys}}}}{\ebd[2]}{\eft[3]}
-  \ar[dotted]{dr}{\mathbf{s}(\famesys{\stesys})}
-  \ar{rr}{\pullback{\pullbackpr{2}{\ectxext}{\eft\circ\ebd}}{\boldsymbol{\beta}_\mathbf{2}}{\ebd}{\eft[2]}}
-  \ar{dd}[swap]{\ebd[2]\circ\pullbackpr{1}{\ebd[2]}{\eft[3]}}
-  &
-  {}&
-\cobesys{\stesyst}{\famesys{\famesys{\stesys}}}{\ebd}{\eft[2]}
-  \ar{d}{\mathbf{s}(\stesys)}
-  \\
-  {}&
-\famesys{\famesys{\stesys}}
-  \ar{r}{\boldsymbol{\beta}}
-  \ar{d}{\eft[2]}
-  &
-\famesys{\stesys}
-  \ar{d}{\eft}
-  \\
-\stesysf_2
-  \ar{r}[swap]{\eft[2]}
-  &
-\stesysf
-  \ar{r}[swap]{\ectxext}
-  &
-\stesysc
-\end{tikzcd}
-\end{equation*}
-commutative.
-\end{lem}
-
-\begin{proof}
-The requirement on pre-substitutions holds by construction.
-\end{proof}
-
-It makes sense now to consider the possibility that the pre-substitution
-itself is a pre-substitution homomorphism.
-
-\begin{defn}
-A \emph{substitution object} is a pre-substitution object for which substitution is
-a pre-substitution homomorphism.
-\end{defn}
-
-\begin{cor}
-The change of base of a substitution object is again a substitution object.
-\end{cor}
-
-\begin{cor}
-If $\stesys$ is a substitution object, then so is $\famesys{\stesys}$.
-\end{cor}
-
-\subsection{Extension objects with empty context and families}
-
-\begin{defn}
-An extension object $\stesys$ is said to have \emph{empty families} if there
+An extension algebra $\stesys$ is said to have \emph{empty families} if there
 is a section
@@
 \begin{defn}
-A homomorphism of extension objects with empty families is an extension
+A homomorphism of extension algebras with empty families is an extension
 homomorphism $f:\stesys'\to\stesys$ for which the diagram
@@
 \begin{lem}
-Suppose $\stesys$ is an extension object with empty families. Then
-$\famesys{\stesys}$ is an extension object with empty families, with
+Suppose $\stesys$ is an extension algebra with empty families. Then
+$\famesys{\stesys}$ is an extension algebra with empty families, with
 $\phi_1(\famesys{\stesys}):=\phi_2$. 
@@
 \begin{lem}
-Suppose $\stesys$ is an extension object with empty families and consider
+Suppose $\stesys$ is an extension algebra with empty families and consider
 $p:\stesysc\rightarrow X\leftarrow Y:p$. Then
-$\cobesys{Y}{\stesys}{g}{p}$ is an extension object with empty families
+$\cobesys{Y}{\stesys}{g}{p}$ is an extension algebra with empty families
 with $\phi_1(\cobesys{Y}{\stesys}{g}{p}):=g^\ast(\phi_1)$.
 \end{lem}
-
-\subsection{E-objects}
-\begin{defn}
-An \emph{E-object} is an extension object with the structure of a projection object,
-the structure of a substitution object and which has an empty context and families,
-such that additionally:
-\begin{enumerate}
-\item substitution is a projection homomorphism
-\item weakening is a substitution homomorphism
-\item both weakening and substitution are empty-CF homomorphisms.
-\item 
-\end{enumerate}
-\end{defn}

### 2015-02-10 Algebras/fundamental.tex
large reorganisation

--- 2014-11-03 Algebras/article-models-cat.tex
@@
-\section{E-objects in categories with finite limits}
-In this section we assume that $\cat{C}$ is a finitely complete category and
+\subsection{Semantics of the fundamental structure of dependent type theory}
+For the semantics of the theory of E-systems, we will assume that $\cat{C}$ is a 
+category with finite limits and
 whenever we write a pullback, we assume that it is chosen. Recall that for
@@
 
-\subsection{Extension objects}
 \begin{defn}
-A \emph{pre-extension object $\stesys$ in $\cat{C}$} consists of a \emph{fundamental structure}, which is a diagram of the form
+A \emph{fundamental structure} $\stesys$ in $\cat{C}$ consists of a diagram of the form
 \begin{equation*}
@@
 \end{equation*}
-in $\cat{C}$ together with the \emph{context extension} and \emph{family extension} operations
-\begin{align*}
-\ectxext &:\stesysf\to \stesysc\\
-\efamext & :\stesysff\to \stesysf,
-\end{align*}
-respectively, such that the diagram
-\begin{equation*}
-\begin{tikzcd}
-\stesysf_2 
-  \ar{r}{\efamext} 
-  \ar{d}[swap]{\eft[2]} 
-  & 
-\stesysf 
-  \ar{d}{\eft}
-  \\
-\stesysf
-  \ar{r}[swap]{\eft} 
-  & 
-\stesysc
-\end{tikzcd}
-\end{equation*}
-commutes.
+in $\cat{C}$. In this diagram, $C$ is the object of \emph{contexts}, $F$ is
+the object of \emph{families}, and $T$ is the object of \emph{terms}.
 \end{defn}
-
-\begin{defn}
-We introduce the following notation:
-\begin{align*}
-\stesysf_2 
-  & := \stesysff
-  \\
-\eft[2] 
-  & := \pullbackpr{1}{\ectxext}{\eft} : \stesysf_2\to\stesysf
-  \\
-\stesysf_3 & := \pullback{\stesysf_2}{\stesysf_2}{\efamext}{\eft[2]}
-  \\
-\eft[3]
-  & := \pullbackpr{1}{\efamext}{\eft[2]} : \stesysf_3\to\stesysf_2.
-\end{align*}
-Then it follows that the outer square in the diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=large]
-\stesysf_3
-  \ar[dotted]{dr}{\eext{2}}
-  \ar{rr}{\pullback{\pullbackpr{2}{\ectxext}{\eft}}{\pullbackpr{2}{\ectxext}{\eft}}{\ectxext}{\eft}}
-  \ar{dd}[swap]{\eft[3]}
-  & 
-  &
-\stesysf_2
-  \ar{d}{\efamext}
-  \\
-  &
-\stesysf_2
-  \ar{d}[swap]{\eft[2]}
-  \ar{r}{\pullbackpr{2}{\ectxext}{\eft}}
-  &
-\stesysf
-  \ar{d}{\eft}
-  \\
-\stesysf_2
-  \ar{r}[swap]{\eft[2]}
-  &
-\stesysf
-  \ar{r}[swap]{\ectxext}
-  &
-\stesysc
-\end{tikzcd}
-\end{equation*}
-commutes. We define $\eext{2}$ to be the unique morphism rendering the above diagram
-commutative. Now we may continue to define
-\begin{align*}
-\stesysf_4 
-  & := 
-\pullback{\stesysf_3}{\stesysf_3}{\eext{2}}{\eft[3]}
-  \\
-\eft[4] 
-  & := 
-\pullbackpr{1}{\eext{2}}{\eft[3]}.
-\end{align*}
-Then we see that the outer square of the diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=large]
-\stesysf_4
-  \ar[dotted]{dr}{\eext{3}}
-  \ar{rr}{\pullback{\pullbackpr{2}{\efamext}{\eft[2]}}{\pullbackpr{2}{\efamext}{\eft[2]}}{\efamext}{\eft[2]}}
-  \ar{dd}[swap]{\eft[4]}
-  & 
-  &
-\stesysf_3
-  \ar{d}{\eext{2}}
-  \\
-  &
-\stesysf_3
-  \ar{d}[swap]{\eft[3]}
-  \ar{r}{\pullbackpr{2}{\efamext}{\eft[2]}}
-  &
-\stesysf_2
-  \ar{d}{\eft[2]}
-  \\
-\stesysf_3
-  \ar{r}[swap]{\eft[3]}
-  &
-\stesysf_2
-  \ar{r}[swap]{\efamext}
-  &
-\stesysf
-\end{tikzcd}
-\end{equation*}
-commutes,
-so we may define $\eext{3}$ to be the unique map which renders the diagram
-commutative. It
-is straightforward to continue this process by induction, but we shall need not
-go any further in this article.
-\end{defn}
-
-\begin{defn} 
-An extension object is a pre-extension object $\stesys$ for which 
-the diagrams
-\begin{equation*}
-\begin{tikzcd}
-\stesysf_2 
-  \ar{d}[swap]{\pullbackpr{2}{\ectxext}{\eft}} 
-  \ar{r}{\efamext} 
-  & 
-\stesysf 
-  \ar{d}{\ectxext}
-  \\
-\stesysf 
-  \ar{r}[swap]{\ectxext} 
-  & 
-\stesysc
-\end{tikzcd}
-\qquad
-\begin{tikzcd}
-\stesysf_3
-  \ar{d}[swap]{\pullbackpr{2}{\efamext}{\eft[2]}}
-  \ar{r}{\eext{2}}
-  & 
-\stesysf_2 
-  \ar{d}{\efamext} 
-  \\
-\stesysf_2 
-  \ar{r}[swap]{\efamext} 
-  &
-\stesysf
-\end{tikzcd}
-\end{equation*}
-commute.
-\end{defn}
-
-\begin{comment}
-\begin{lem}
-There exists an isomorphism $\alpha$ such that the triangle
-\begin{equation*}
-\begin{tikzcd}[column sep=tiny]
-\pullback{\stesysf}{\stesysf_2}{\ectxext}{\eft\circ\eft[2]}
-  \ar[dotted]{rr}{\alpha}
-  \ar{dr}[swap]{\pullback{\catid{\stesysf}}{\efamext}{\ectxext}{\eft}}
-  &
-  &
-\stesysf_3
-  \ar{dl}{\eext{2}}
-  \\
-& \stesysf_2
-\end{tikzcd}
-\end{equation*}
-commutes
-\end{lem}
-
-\begin{proof}
-There is a unique morphism $\alpha:
-\pullback{\stesysf}{\stesysf_2}{\ectxext}{\eft\circ\eft[2]}\to\stesysf_3$
-rendering the diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=large]
-\pullback{\stesysf}{\stesysf_2}{\ectxext}{\eft\circ\eft[2]}
-  \ar[bend left=10,yshift=.5ex]{drrr}{\pullbackpr{2}{\ectxext}{\eft}\circ\pullbackpr{2}{\ectxext}{\eft\circ\eft[2]}}
-  \ar[bend right=10]{ddr}[swap]{\pullback{\catid{\stesysf}}{\eft[2]}{\ectxext}{\eft}}
-  \ar[dotted]{dr}{\alpha}
-  \\
-& \stesysf_3
-  \ar{r}{\pullbackpr{2}{\efamext}{\eft[2]}}
-  \ar{d}{\eft[3]}
-  &
-\stesysf_2
-  \ar{d}[swap]{\eft[2]}
-  \ar{r}[swap]{\pullbackpr{2}{\ectxext}{\eft}}
-  &
-\stesysf
-  \ar{d}{\eft}
-  \\
-{} & \stesysf_2
-  \ar{r}[swap]{\efamext}
-  &
-\stesysf
-  \ar{r}[swap]{\ectxext}
-  &
-\stesysc
-\end{tikzcd}
-\end{equation*}
-\end{proof}
-\end{comment}
-
-\begin{defn}
-Suppose $\stesys$ is a pre-extension object of $\cat{C}$. Then we define the pre-extension object
-$\famesys{\stesys}$ to consist of the fundamental structure
-\begin{equation*}
-\begin{tikzcd}
-\stesyst_2
-  \ar{d}{\ebd[2]}
-  \\
-\stesysf_2
-  \ar{d}{\eft[2]}
-  \\
-\stesysf
-\end{tikzcd}
-\end{equation*}
-where
-\begin{align*}
-\stesyst_2 
-  & := \pullback{\stesysf}{\stesyst}{\ectxext}{\eft\circ\ebd}
-  \\
-\ebd[2]
-  & := \ectxext^\ast(\ebd),
-\end{align*}
-with the extension operations
-\begin{align*}
-\efamext 
-  & 
-  : \stesysf_2\to\stesysf\\
-\eext{2} & : \stesysf_3\to\stesysf_2.
-\end{align*}
-\end{defn}
-
-In \autoref{famextobj} below we shall show that $\famesys{\stesys}$ is an
-extension algebra whenever $\stesys$ is an extension algebra. We shall need
-a handful of lemmas to give the proof.
-
-\begin{defn}
-We define
-\begin{align*}
-\beta_1 
-  & := 
-\pullbackpr{2}{\ectxext}{\eft}
-  & &
-  : \stesysf_2\to\stesysf
-  \\
-\beta_2
-  & :=
-\pullback{\beta_1}{\beta_1}{\ectxext}{\eft}
-  & &
-  : \stesysf_3\to\stesysf_2
-  \\
-\beta_3
-  & :=
-\pullback{\beta_2}{\beta_2}{\efamext}{\eft[2]}
-  & &
-  : \stesysf_4\to\stesysf_3.
-\end{align*}
-\end{defn}
-
-\begin{lem}
-Let $\stesys$ be a pre-extension object. Then the square
-\begin{equation*}
-\begin{tikzcd}[column sep=10em]
-\stesysf_4
-  \ar{r}{\pullback{\pullbackpr{2}{\efamext}{\eft[2]}}{\pullbackpr{2}{\efamext}{\eft[2]}}{\efamext}{\eft[2]}}
-  \ar{d}[swap]{\beta_3}
-  &
-\stesysf_3
-  \ar{d}{\beta_2}
-  \\
-\stesysf_3
-  \ar{r}[swap]{\beta_2}
-  &
-\stesysf_2
-\end{tikzcd}
-\end{equation*}
-commutes.
-\end{lem}
-
-\begin{proof}
-Left to the reader.
-\end{proof}
-
-\begin{comment}
-\begin{proof}
-It is straightforward to verify the equalities
-\begin{align*}
-\pullbackpr{1}{\ectxext}{\eft}\circ\beta_2\circ
-  (\pullback{\pullbackpr{2}{\efamext}{\eft[2]}}{\pullbackpr{2}{\efamext}{\eft[2]}}{\efamext}{\eft[2]})
-  & =
-\pullbackpr{1}{\ectxext}{\eft}\circ\beta_2\circ\beta_3
-  \\
-\pullbackpr{2}{\ectxext}{\eft}\circ\beta_2\circ
-  (\pullback{\pullbackpr{2}{\efamext}{\eft[2]}}{\pullbackpr{2}{\efamext}{\eft[2]}}{\efamext}{\eft[2]})
-  & =
-\pullbackpr{2}{\ectxext}{\eft}\circ\beta_2\circ\beta_3.\qedhere
-\end{align*}
-\end{proof}
-\end{comment}
-
-Note that the square
-\begin{equation*}
-\begin{tikzcd}
-\stesysf_3
-  \ar{r}{\beta_2}
-  \ar{d}[swap]{\eext{2}}
-  &
-\stesysf_2
-  \ar{d}{\efamext}
-  \\
-\stesysf_2
-  \ar{r}[swap]{\beta_1}
-  &
-\stesysf
-\end{tikzcd}
-\end{equation*}
-commutes by definition. We have a similar result relating $\eext{3}$ and
-$\eext{2}$.
-
-\begin{lem}
-Let $\stesys$ be a pre-extension object. Then the square
-\begin{equation*}
-\begin{tikzcd}
-\stesysf_4
-  \ar{r}{\beta_3}
-  \ar{d}[swap]{\eext{3}}
-  &
-\stesysf_3
-  \ar{d}{\eext{2}}
-  \\
-\stesysf_3
-  \ar{r}[swap]{\beta_2}
-  &
-\stesysf_2
-\end{tikzcd}
-\end{equation*}
-commutes.
-\end{lem}
-
-\begin{proof}
-Left to the reader.
-\end{proof}
-\begin{comment}
-\begin{proof}
-It is straightforward to verify the equalities
-\begin{align*}
-\pullbackpr{1}{\ectxext}{\eft}\circ\beta_2\circ\eext{3}
-  & = \beta_1\circ\eft[3]\circ\eft[4]
-  \\
-\pullbackpr{1}{\ectxext}{\eft}\circ\eext{2}\circ\beta_3
-  & = \beta_1\circ\eft[3]\circ\eft[4].
-\end{align*}
-Thus, it remains to verify that
-\begin{equation*}
-\pullbackpr{2}{\ectxext}{\eft}\circ\beta_2\circ\eext{3}
-  = \pullbackpr{2}{\ectxext}{\eft}\circ\eext{2}\circ\beta_3.
-\end{equation*}
-It is straightforward to see that the diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=large]
-\stesysf_4
-  \ar{dd}[swap]{\pullback{\pullbackpr{2}{\efamext}{\eft[2]}}{\pullbackpr{2}{\efamext}{\eft[2]}}{\efamext}{\eft[2]}}
-  \ar{r}{\eext{3}}
-  &
-\stesysf_3
-  \ar{r}{\beta_2}
-  \ar{d}[swap]{\pullbackpr{2}{\efamext}{\eft[2]}}
-  &
-\stesysf_2
-  \ar{dd}{\pullbackpr{2}{\ectxext}{\eft}}
-  \\
-  {} &
-\stesysf_2
-  \ar{dr}{\beta_1}
-  \\
-\stesysf_3
-  \ar{ur}{\eext{2}}
-  \ar{r}[swap]{\beta_2}
-  &
-\stesysf_2
-  \ar{r}[swap]{\efamext}
-  &
-\stesysf
-\end{tikzcd}
-\end{equation*}
-commutes. It is likewise straightforward to see that the diagram
-\begin{equation*}
-\begin{tikzcd}
-\stesysf_4
-  \ar{r}{\beta_3}
-  \ar{d}[swap]{\pullback{\pullbackpr{2}{\efamext}{\eft[2]}}{\pullbackpr{2}{\efamext}{\eft[2]}}{\efamext}{\eft[2]}}
-  &
-\stesysf_3
-  \ar{r}{\eext{2}}
-  \ar{d}[swap]{\beta_2}
-  &
-\stesysf_2
-  \ar{d}{\pullbackpr{2}{\ectxext}{\eft}}
-  \\
-\stesysf_3
-  \ar{r}[swap]{\beta_2}
-  &
-\stesysf_2
-  \ar{r}[swap]{\efamext}
-  &
-\stesysf
-\end{tikzcd}
-\end{equation*}
-commutes, completing our goal.
-\end{proof}
-\end{comment}
-
-\begin{thm}[Local extension structure]\label{famextobj}
-If $\stesys$ is an extension object, then so is $\famesys{\stesys}$.
-\end{thm}
-
-\begin{proof}
-Note that the diagram
-\begin{equation*}
-\begin{tikzcd}
-\stesysf_3
-  \ar{d}[swap]{\pullbackpr{2}{\efamext}{\eft[2]}}
-  \ar{r}{\eext{2}}
-  & 
-\stesysf_2 
-  \ar{d}{\efamext} 
-  \\
-\stesysf_2 
-  \ar{r}[swap]{\efamext} 
-  &
-\stesysf
-\end{tikzcd}
-\end{equation*}
-commutes by assumption. For the second condition, we have to show that the
-diagram
-\begin{equation*}
-\begin{tikzcd}
-\stesysf_4
-  \ar{d}[swap]{\pullbackpr{2}{\eext{2}}{\eft[3]}}
-  \ar{r}{\eext{3}}
-  & 
-\stesysf_3
-  \ar{d}{\eext{2}} 
-  \\
-\stesysf_3
-  \ar{r}[swap]{\eext{2}} 
-  &
-\stesysf_2
-\end{tikzcd}
-\end{equation*}
-Since this is a question about two maps into a pullback, it suffices to verify
-that
-\begin{align*}
-\pullbackpr{1}{\ectxext}{\eft}\circ\eext{2}\circ\eext{3}
-  & =
-\pullbackpr{1}{\ectxext}{\eft}\circ\eext{2}\circ\pullbackpr{2}{\eext{2}}{\eft[3]}
-  \\
-\pullbackpr{2}{\ectxext}{\eft}\circ\eext{2}\circ\eext{3}
-  & =
-\pullbackpr{2}{\ectxext}{\eft}\circ\eext{2}\circ\pullbackpr{2}{\eext{2}}{\eft[3]}.
-\end{align*}
-For the first equality, it is fairly straightforward to show that both the
-equalities
-\begin{equation*}
-\pullbackpr{1}{\ectxext}{\eft}\circ\eext{2}\circ\eext{3}
-  =
-\eft[2]\circ\eft[3]\circ\eft[4]
-\end{equation*}
-and
-\begin{equation*}
-\pullbackpr{1}{\ectxext}{\eft}\circ\eext{2}\circ\pullbackpr{2}{\eext{2}}{\eft[3]}
-  =
-\eft[2]\circ\eft[3]\circ\eft[4].
-\end{equation*}
-hold. For the second subgoal (which is more tricky). Notice first that the
-diagram
-\begin{equation*}
-\begin{tikzcd}
-\stesysf_4
-  \ar{r}{\eext{3}}
-  \ar{d}[swap]{\beta_3}
-  &
-\stesysf_3
-  \ar{r}{\eext{2}}
-  \ar{d}[swap]{\beta_2}
-  &
-\stesysf_2
-  \ar{d}{\pullbackpr{2}{\ectxext}{\eft}}
-  \\
-\stesysf_3
-  \ar{r}[swap]{\eext{2}}
-  &
-\stesysf_2
-  \ar{r}[swap]{\efamext}
-  &
-\stesysf
-\end{tikzcd}
-\end{equation*}
-commutes. We also have the commutative diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=large]
-\stesysf_4
-  \ar{r}{\pullbackpr{2}{\eext{2}}{\eft[3]}}
-  \ar{d}[swap]{\beta_3}
-  &
-\stesysf_3
-  \ar{r}{\eext{2}}
-  \ar{d}[swap]{\beta_2}
-  &
-\stesysf_2
-  \ar{d}{\pullbackpr{2}{\ectxext}{\eft}}
-  \\
-\stesysf_3
-  \ar{r}{\pullbackpr{2}{\efamext}{\eft[2]}}
-  \ar{dr}[swap]{\eext{2}}
-  &
-\stesysf_2
-  \ar{r}{\efamext}
-  &
-\stesysf
-  \\
-  {} &
-\stesysf_2
-  \ar{ur}[swap]{\efamext}
-\end{tikzcd}
-\end{equation*}
-completing the proof.
-\end{proof}
-
-\subsection{(Pre-)extension homomorphisms}\label{subsection:e_extension_homomorphisms}
-In this subsection we start with the study of pre-extension homomorphisms, which
-will include the extension homomorphisms since they will be the pre-extension
-homomorphisms of which both the domain and codomain are extension objects.
-Our main examples of extension homomorphisms will be the operations of weakening
-and substitution. There are some basic examples of pre-extension homomorphisms
-that will be useful too, which get introduced in the this section and in
-\autoref{subsection:change_of_base}. In this section, we will mainly be
-interested in pre-extension homomorphisms between local pre-extension objects.
-We will end this section by proving that a retract of an extension object is
-always an extension object.
-
-\begin{defn}
-Let $\stesys$ and $\stesys'$ be pre-extension algebras. A \emph{pre-extension 
-homomorphism $f$ from $\stesys'$ to $\stesys$} is a triple $(f_0,f_1,f^t)$ 
-consisting of morphisms
-\begin{equation*}
-\begin{tikzcd}
-\stesyst' 
-  \ar{r}{f^t}
-  \ar{d}[swap]{\ebd'}
-  &
-\stesyst
-  \ar{d}{\ebd}
-  \\
-\stesysf'
-  \ar{r}{f_1}
-  \ar{d}[swap]{\eft'}
-  &
-\stesysf
-  \ar{d}{\eft}
-  \\
-\stesysc' 
-  \ar{r}[swap]{f_0}
-  &
-\stesysc
-\end{tikzcd}
-\end{equation*}
-such that the indicated squares commute, for which furthermore the squares
-\begin{equation*}
-\begin{tikzcd}
-\stesysf' 
-  \ar{r}{f_1}
-  \ar{d}[swap]{\ectxext'}
-  &
-\stesysf
-  \ar{d}{\ectxext}
-  \\
-\stesysc'
-  \ar{r}[swap]{f_0}
-  &
-\stesysc
-\end{tikzcd}
-\end{equation*}
-and
-\begin{equation*}
-\begin{tikzcd}[column sep=huge]
-\stesysf'\times_{\ectxext',\eft'} \stesysf'
-  \ar{r}{f_1\times_{\ectxext,\eft} f_1}
-  \ar{d}[swap]{\efamext'}
-  &
-\stesysf\times_{\ectxext,\eft} \stesysf
-  \ar{d}{\efamext}
-  \\
-\stesysf'
-  \ar{r}[swap]{f_1}
-  &
-\stesysf
-\end{tikzcd}
-\end{equation*}
-Composition and the identity homomorphism are defined in the expected way. We
-define furthermore
-\begin{align*}
-f_2 & := \pullback{f_1}{f_1}{\ectxext}{\eft}
-  \\
-f_3 & := \pullback{f_2}{f_2}{\efamext}{\eft[2]}.
-\end{align*}
-\end{defn}
-
-\begin{defn}
-A pre-extension homomorphism between extension objects is called an extension
-homomorphism.
-\end{defn}
-
-\begin{lem}
-Let $\stesys$ be an extension object. Then
-\begin{equation*}
-\begin{tikzcd}[column sep=huge]
-\stesyst_2
-  \ar{r}{\pullbackpr{2}{\ectxext}{\eft\circ\ebd}}
-  \ar{d}[swap]{\ebd[2]}
-  &
-\stesyst
-  \ar{d}{\ebd}
-  \\
-\stesysf_2
-  \ar{r}{\pullbackpr{2}{\ectxext}{\eft}}
-  \ar{d}[swap]{\eft[2]}
-  &
-\stesysf
-  \ar{d}{\eft}
-  \\
-\stesysf
-  \ar{r}[swap]{\ectxext}
-  &
-\stesysc
-\end{tikzcd}
-\end{equation*}
-assembles an extension homomorphism $\mathbf{e}_0:\famesys{\stesys}\to\stesys$.
-Likewise, we have an extension homomorphism
-$\mathbf{e}_1:\famesys{\famesys{\stesys}}\to\famesys{\stesys}$. Thus, a
-pre-extension object is an extension object if and only if $\mathbf{e}_0$
-and $\mathbf{e}_1$ are pre-extension homomorphisms.
-\end{lem}
-
-\begin{proof}
-Immediate from the conditions of being an extension object.
-\end{proof}
-
-\begin{defn}
-Let $\stesys$ be a pre-extension object. Then
-\begin{equation*}
-\begin{tikzcd}
-\stesyst_3
-  \ar{r}{\beta^t}
-  \ar{d}[swap]{\ebd[3]}
-  &
-\stesyst_2
-  \ar{d}{\ebd[2]}
-  \\
-\stesysf_3
-  \ar{r}{\beta_2}
-  \ar{d}[swap]{\eft[3]}
-  &
-\stesysf_2
-  \ar{d}{\eft[2]}
-  \\
-\stesysf_2
-  \ar{r}[swap]{\beta_1}
-  &
-\stesysf
-\end{tikzcd}
-\qquad
-\text{and}
-\qquad
-\begin{tikzcd}
-\stesyst_4
-  \ar{r}{\beta^t_2}
-  \ar{d}[swap]{\ebd[4]}
-  &
-\stesyst_3
-  \ar{d}{\ebd[3]}
-  \\
-\stesysf_4
-  \ar{r}{\beta_3}
-  \ar{d}[swap]{\eft[4]}
-  &
-\stesysf_3
-  \ar{d}{\eft[3]}
-  \\
-\stesysf_3
-  \ar{r}[swap]{\beta_2}
-  &
-\stesysf_2
-\end{tikzcd}
-\end{equation*}
-assemble pre-extension homomorphisms 
-\(
-\boldsymbol{\beta}
-  :
-\famesys{\famesys{\stesys}}
-  \to
-\famesys{\stesys}
-\) 
-and
-\(
-\boldsymbol{\beta}_\mathbf{2}
-  :
-\famesys{\famesys{\famesys{\stesys}}}
-  \to
-\famesys{\famesys{\stesys}}
-\).
-\end{defn}
-
-\begin{defn}\label{famehom}
-Suppose that $f:\stesys'\to\stesys$ is a pre-extension homomorphism. Then we
-define $\famehom{f}:\famesys{\stesys'}\to\famesys{\stesys}$ to consist of
-\begin{equation*}
-\begin{tikzcd}
-\stesyst_2'
-  \ar{r}{f^t_2}
-  \ar{d}[swap]{\ebd[2]'}
-  &
-\stesyst_2
-  \ar{d}{\ebd[2]}
-  \\
-\stesysf_2'
-  \ar{r}{f_2}
-  \ar{d}[swap]{\eft[2]'}
-  &
-\stesysf_2
-  \ar{d}{\eft[2]}
-  \\
-\stesysf'
-  \ar{r}[swap]{f_1}
-  &
-\stesysf
-\end{tikzcd}
-\end{equation*}
-where we define
-\begin{equation*}
-f^t_2 := \pullback{f_1}{f^t}{\ectxext}{\eft\circ\ebd}.
-\end{equation*}
-\end{defn}
-
-\begin{lem}
-The triple $\famehom{f}$ defined in \autoref{famehom} is a pre-extension homomorphism.
-\end{lem}
-
-\begin{proof}
-Note that the square
-\begin{equation*}
-\begin{tikzcd}
-\stesysf_2'
-  \ar{r}{f_2}
-  \ar{d}[swap]{\efamext'}
-  &
-\stesysf_2
-  \ar{d}{\efamext}
-  \\
-\stesysf'
-  \ar{r}[swap]{f_1}
-  &
-\stesysf
-\end{tikzcd}
-\end{equation*}
-commutes by assumption. Thus, it remains to show that the square
-\begin{equation*}
-\begin{tikzcd}
-\stesysf_3'
-  \ar{r}{f_3}
-  \ar{d}[swap]{\eext{2}'}
-  &
-\stesysf_3
-  \ar{d}{\eext{2}}
-  \\
-\stesysf_2'
-  \ar{r}[swap]{f_2}
-  &
-\stesysf_2
-\end{tikzcd}
-\end{equation*}
-commutes. It is equivalent to show that the equalities
-\begin{align*}
-\pullbackpr{1}{\ectxext}{\eft}\circ f_2\circ\eext{2}'
-  & =
-\pullbackpr{1}{\ectxext}{\eft}\circ \eext{2}\circ f_3
-  \\
-\pullbackpr{2}{\ectxext}{\eft}\circ f_2\circ\eext{2}'
-  & =
-\pullbackpr{2}{\ectxext}{\eft}\circ \eext{2}\circ f_3
-\end{align*}
-both hold. For the first, it is straightforward to verify that the diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=large]
-{} &
-\stesysf_2'
-  \ar{r}{f_2}
-  \ar{dr}[near end]{\pullbackpr{1}{\ectxext'}{\eft'}}
-  &
-\stesysf_2
-  \ar{dr}{\pullbackpr{1}{\ectxext}{\eft}}
-  \\
-\stesysf_3'
-  \ar{ur}{\eext{2}'}
-  \ar{r}[swap]{\beta_2'}
-  \ar{ddr}[swap]{f_3}
-  &
-\stesysf_2'
-  \ar{r}{\efamext'}
-  \ar{dr}[swap]{f_2}
-  &
-\stesysf'
-  \ar{r}{f_1}
-  &
-\stesysf
-  \\
-{} & {} &
-\stesysf_2
-  \ar{ur}[near start]{\efamext}
-  \\
-{} &
-\stesysf_3
-  \ar{r}[swap]{\eext{2}}
-  \ar{ur}{\beta_2}
-  &
-\stesysf_2
-  \ar{uur}[swap]{\pullbackpr{1}{\ectxext}{\eft}}
-\end{tikzcd}
-\end{equation*}
-commutes. For the second, note that the diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=large]
-{} &
-\stesysf_2'
-  \ar{r}{f_2}
-  \ar{dr}{\beta_1'}
-  &
-\stesysf_2
-  \ar{ddr}{\pullbackpr{2}{\ectxext}{\eft}}
-  \\
-{} & {} &
-\stesysf'
-  \ar{dr}[swap,near start]{f_1}
-  \\
-\stesysf_3'
-  \ar{uur}{\eext{2}'}
-  \ar{r}{\beta_2'}
-  \ar{dr}[swap]{f_3}
-  &
-\stesysf_2'
-  \ar{r}{f_2}
-  \ar{ur}{\efamext'}
-  &
-\stesysf_2
-  \ar{r}[swap]{\efamext}
-  &
-\stesysf
-  \\
-{} &
-\stesysf_3
-  \ar{r}[swap]{\eext{2}}
-  \ar{ur}[near start]{\beta_2}
-  &
-\stesysf_2
-  \ar{ur}[swap]{\pullbackpr{2}{\ectxext}{\eft}}
-\end{tikzcd}
-\end{equation*}
-commutes.
-\end{proof}
-
-\begin{lem}[Stability under retracts]\label{esys-retract}
-Suppose $f:\stesys\to\stesys'$ is a pre-extension homomorphism between
-pre-extension objects. If there is a pre-extension homomorphism $g:\stesys'\to
-\stesys$ such that $g\circ f=\catid{\stesys}$ and $\stesys'$ is an extension
-algebra, then $\stesys$ is an extension algebra.
-\end{lem}
-
-Before we start with the proof, note that we have the equalities
-$g_2\circ f_2=\catid{\stesysf_2}$ and $g_3\circ f_3=\catid{\stesysf_3}$
-under the hypotheses of the lemma.
-
-\begin{proof}
-Our first subgoal is to show that the square
-\begin{equation*}
-\begin{tikzcd}
-\stesysf_2 
-  \ar{r}{\efamext} 
-  \ar{d}[swap]{\pullbackpr{2}{\ectxext}{\eft}} 
-  & 
-\stesysf 
-  \ar{d}{\ectxext}
-  \\
-\stesysf
-  \ar{r}[swap]{\ectxext} 
-  & 
-\stesysc
-\end{tikzcd}
-\end{equation*}
-commutes. Note that in the diagram
-\begin{equation*}
-\begin{tikzcd}
-  {}
-  & 
-\stesysf
-  \ar{dd}[near start]{\ectxext}
-  \ar{rr}{f_1}
-  &
-  &
-\stesysf'
-  \ar{dd}[near start]{\ectxext'}
-  \ar{rr}{g_1}
-  &
-  &
-\stesysf
-  \ar{dd}{\ectxext}
-  \\
-\stesysf_2
-  \ar{dd}[swap]{\pullbackpr{2}{\ectxext}{\eft}}
-  \ar[crossing over]{rr}[swap,near start]{f_2}
-  \ar{ur}{\efamext}
-  &
-  &
-\stesysf_2'
-  \ar{ur}[near start]{\efamext'}
-  \ar[crossing over]{rr}[swap,near start]{g_2}
-  &
-  &
-\stesysf_2
-  \ar{ur}[swap,near start]{\efamext}
-  \\
-  {}
-  &
-\stesysc
-  \ar{rr}[near start]{f_0}
-  &
-  &
-\stesysc'
-  \ar{rr}[near start]{g_0}
-  &
-  &
-\stesysc
-  \\
-\stesysf 
-  \ar{rr}[swap]{f_1}
-  \ar{ur}{\ectxext}
-  &
-  &
-\stesysf' 
-  \ar[crossing over,leftarrow]{uu}[near end,swap]{\pullbackpr{2}{\ectxext'}{\eft'}}
-  \ar{ur}[swap,near end]{\ectxext'}
-  \ar{rr}[swap]{g_1}
-  &
-  &
-\stesysf
-  \ar[crossing over,leftarrow]{uu}[near end,swap]{\pullbackpr{2}{\ectxext}{\eft}}
-  \ar{ur}[swap]{\ectxext}
-\end{tikzcd}
-\end{equation*}
-all the faces minus the far left and far right face commute. Using that $g$
-is a section of $f$, we can read off that also the far left face commutes,
-completing our first subgoal.
-
-For the second subgoal, note that also $\famehom{g}\circ\famehom{f}=
-\catid{\famesys{\stesys}}$ and that $\famesys{\stesys'}$ is an extension object.
-Thus we can apply what we have proven so far to conclude that the square
-\begin{equation*}
-\begin{tikzcd}
-\stesysf_3 
-  \ar{r}{\eext{2}} 
-  \ar{d}[swap]{\pullbackpr{2}{\efamext}{\eft[2]}} 
-  & 
-\stesysf_2 
-  \ar{d}{\efamext}
-  \\
-\stesysf_2
-  \ar{r}[swap]{\efamext} 
-  & 
-\stesysf
-\end{tikzcd}
-\end{equation*}
-commutes.
-\end{proof}
-
-\subsection{The change of base of (pre-)extension objects}
-\label{subsection:change_of_base}
-An important construction of (pre-)extension objects is the change of base. It
-allows us to consider `parametrized homomorphisms', such as weakening and
-substitution.
-
-\begin{defn}
-Suppose $f:\stesys\to\stesys'$ is a pre-extension homomorphism. We say that
-a diagram
-\begin{equation*}
-\begin{tikzcd}
-\stesys
-  \ar{r}{f}
-  \ar{d}[swap]{p}
-  &
-\stesys'
-  \ar{d}{p'}
-  \\
-X \ar{r}[swap]{g}
-  &
-Y
-\end{tikzcd}
-\end{equation*}
-commutes if the diagram
-\begin{equation*}
-\begin{tikzcd}
-\stesysc
-  \ar{r}{f_0}
-  \ar{d}[swap]{p}
-  &
-\stesysc'
-  \ar{d}{p'}
-  \\
-X \ar{r}[swap]{g}
-  &
-Y
-\end{tikzcd}
-\end{equation*}
-commutes.
-\end{defn}
-
-The first goal in this subsection is to define for every (pre-)extension object 
-$\stesys$ and every $p:\stesysc\rightarrow X\leftarrow Y:g$, a (pre-)extension
-object $\cobesys{Y}{\stesys}{g}{p}$ with a homomorphism $\pullbackpr{2}{g}{p}:
-\cobesys{Y}{\stesys}{g}{p}\to\stesys$ and a morphism $\pullbackpr{1}{g}{p}:
-\pullback{Y}{\stesysc}{g}{p}\to Y$ such that for every diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=large]
-\stesys'
-  \ar[bend right=10]{ddr}[swap]{p'}
-  \ar[bend left=10]{rrd}{f}
-  \ar[dotted]{dr}[near end]{[p',f]}
-  \\
-  {}&
-\cobesys{Y}{\stesys}{g}{p}
-  \ar{d}{\pullbackpr{1}{g}{p}}
-  \ar{r}[swap]{\pullbackpr{2}{g}{p}}
-  &
-\stesys
-  \ar{d}{p}
-  \\
-  {}&
-Y \ar{r}[swap]{g}
-  &
-X
-\end{tikzcd}
-\end{equation*}
-of which the outer square commutes, the (pre-)extension homomorphism $[p',f]$ exists
-and is unique with the property that it renders the diagram commutative. We will
-give the definition of $\cobesys{Y}{\stesys}{g}{p}$ in \autoref{cobesys}. After
-proving that the change of base of a pre-extension algebra is indeed a
-pre-extension algebra (\autoref{cobesys-preext}) and that the change of base
-of an extension algebra is an extension algebra (\autoref{cobesys-ext}), we
-will demonstrate the above unique existence in \autoref{cobesys-existence,%
-cobesys-pullback}.
-
-The second goal in this subsection is to follow the same procedure for
-$\famesys{\famesys{\stesys}}$ to show that it is equivalent to
-$\cobesys{\stesysf}{\famesys{\stesys}}{\ectxext}{\eft}$. We will do this by
-verifying directly that it has the universal property of the change of base
-described above, because we will use the ingredients in our definition of
-weakening and substitution objects.
-
-\begin{defn}[Change of base]\label{cobesys}
-Suppose $\stesys$ is a pre-extension object in $\cat{C}$ and that 
-$p:\stesysc\rightarrow X\leftarrow Y:g$.
-Then we define the pre-extension object $\cobesys{Y}{\stesys}{g}{p}$ to consist of
-\begin{equation*}
-\begin{tikzcd}
-\cobesys{Y}{\stesyst}{g}{p\circ\eft\circ\ebd}
-  \ar{r}
-  \ar{d}[swap]{g^\ast(\ebd)}
-  &
-\stesyst
-  \ar{d}{\ebd}
-  \\
-\cobesys{Y}{\stesysf}{g}{p\circ\eft}
-  \ar{r}
-  \ar{d}[swap]{g^\ast(\eft)}
-  &
-\stesysf
-  \ar{d}{\eft}
-  \\
-\cobesys{Y}{\stesysc}{g}{p}
-  \ar{r}
-  \ar{d}[swap]{\pullbackpr{1}{g}{p}}
-  &
-\stesysc
-  \ar{d}{p}
-  \\
-Y \ar{r}[swap]{g}
-  &
-X
-\end{tikzcd}
-\end{equation*} 
-and the operations
-\begin{align*}
-\cobesys{Y}{\ectxext}{g}{p} 
-  & : \pullback{Y}{\stesysf}{g}{p\circ\eft}\to \pullback{Y}{\stesysc}{g}{p}\\
-\cobesys{Y}{\efamext}{g}{p} 
-  & : \pullback
-    {\pullback{Y}{\stesysf}{g}{p\circ\eft}}
-    {\pullback{Y}{\stesysf}{g}{p\circ\eft}}
-    {\cobesys{Y}{\ectxext}{g}{p}}
-    {g^\ast(\eft)}
-  \to 
-  \pullback{Y}{\stesysf}{g}{p\circ\eft}.
-\end{align*}
-defined by
-\begin{equation*}
-\cobesys{Y}{\ectxext}{g}{p} := \pullback{\catid{Y}}{\ectxext}{g}{p}
-\end{equation*}
-and where $\cobesys{Y}{\efamext}{g}{p}$ is defined by rendering the diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=large]
-(\cobesys{Y}{\stesysf}{g}{p\circ\eft})_2
-  \ar{rr}{\pullback{\pullbackpr{2}{g}{p\circ\eft}}{\pullbackpr{2}{g}{p\circ\eft}}{\ectxext}{\eft}}
-  \ar{dd}[swap]{\pullbackpr{1}{\cobesys{Y}{\ectxext}{g}{p}}{g^\ast(\eft)}}
-  \ar[dotted]{dr}[swap]{\cobesys{Y}{\efamext}{g}{p}}
-  &
-  &
-\stesysf_2
-  \ar{d}{\efamext}
-  \\
-  {}&
-\cobesys{Y}{\stesysf}{g}{p\circ\eft}
-  \ar{r}{\pullbackpr{2}{g}{p\circ\eft}}
-  \ar{d}[swap]{\pullbackpr{1}{g}{p\circ\eft}}
-  &
-\stesysf
-  \ar{d}{p\circ\eft}
-  \\
-\cobesys{Y}{\stesysf}{g}{p\circ\eft}
-  \ar{r}[swap]{\pullbackpr{1}{g}{p\circ\eft}}
-  &
-Y \ar{r}[swap]{g}
-  &
-X
-\end{tikzcd}
-\end{equation*} 
-commutative. 
-The process of obtaining the pre-extension object $\cobesys{Y}{\stesys}{g}{p}$ out of $\stesys$
-and $g:Y\to X$ is also called the \emph{change of base}.
-\end{defn}
-
-\begin{lem}\label{cobesys-preext}
-Any change of base of a pre-extension object is a pre-extension object.
-\end{lem}
-
-\begin{proof}
-Let $\stesys$ be an extension algebra and consider $p:\stesysc\rightarrow X\leftarrow Y:g$.
-We need to verify that the square
-\begin{equation*}
-\begin{tikzcd}[column sep=large]
-(\pullback{Y}{\stesysf}{g}{p\circ\eft})_2
-  \ar{r}{\cobesys{Y}{\efamext}{g}{p}} 
-  \ar{d}[swap]{\pullbackpr{1}{\cobesys{Y}{\ectxext}{g}{p}}{g^\ast(\eft)}} 
-  & 
-\pullback{Y}{\stesysf}{g}{p\circ\eft}
-  \ar{d}{g^\ast(\eft)}
-  \\
-\pullback{Y}{\stesysf}{g}{p\circ\eft}
-  \ar{r}[swap]{g^\ast(\eft)} 
-  & 
-\pullback{Y}{\stesysc}{g}{p}
-\end{tikzcd}
-\end{equation*}
-commutes. It is fairly obvious that
-\begin{equation*}
-\pullbackpr{1}{g}{p}\circ g^\ast(\eft)\circ (\cobesys{Y}{\efamext}{g}{p})
-  =
-\pullbackpr{1}{g}{p\circ\eft}\circ \pullbackpr{1}{\cobesys{Y}{\ectxext}{g}{p}}{g^\ast(\eft)}
-\end{equation*}
-and that the diagram
-\begin{equation*}
-\begin{tikzcd}
-  {}&
-  {}&
-\pullback{Y}{\stesysf}{g}{p\circ\eft}
-  \ar{rr}{g^\ast(\eft)}
-  \ar{dr}[swap]{\pullbackpr{2}{g}{p\circ\eft}}
-  &
-  {}&
-\pullback{Y}{\stesysc}{g}{p}
-  \ar{ddr}{\pullbackpr{2}{g}{p}}
-  \\
-  {}&
-  {}&
-  {}&
-\stesysf
-  \ar{drr}[swap]{\eft}
-  \\
-(\pullback{y}{\stesysf}{g}{p\circ\eft})_2
-  \ar{uurr}{\cobesys{Y}{\efamext}{g}{p}}
-  \ar{rr}[swap,yshift=-.5ex]{\pullback{\pullbackpr{2}{g}{p\circ\eft}}{\pullbackpr{2}{g}{p\circ\eft}}{\ectxext}{\eft}}
-  \ar{ddrr}[swap]{\pullbackpr{1}{\cobesys{Y}{\ectxext}{g}{p}}{g^\ast(\eft)}}
-  &
-  {}&
-\stesysf_2
-  \ar{ur}{\efamext}
-  \ar{dr}[swap]{\eft[2]}
-  &
-  {}&
-  {}&
-\stesysc
-  \\
-  {}&
-  {}&
-  {}&
-\stesysf
-  \ar{urr}{\eft}
-  \\
-  {}&
-  {}&
-\pullback{Y}{\stesysf}{g}{p\circ\eft}
-  \ar{rr}[swap]{g^\ast(\eft)}
-  \ar{ur}{\pullbackpr{2}{g}{p\circ\eft}}
-  &
-  {}&
-\pullback{Y}{\stesysc}{g}{p}
-  \ar{uur}[swap]{\pullbackpr{2}{g}{p}}
-\end{tikzcd}
-\end{equation*}
-commutes.
-\end{proof}
-
-\begin{thm}\label{cobesys-ext}
-The change of base of an extension algebra is an extension algebra.
-\end{thm}
-
-\begin{proof}
-Our first subgoal is to verify that the square
-\begin{equation*}
-\begin{tikzcd}[column sep=large]
-(\pullback{Y}{\stesysf}{g}{p\circ\eft})_2
-  \ar{r}{\cobesys{Y}{\efamext}{g}{p}} 
-  \ar{d}[swap]{\pullbackpr{2}{\cobesys{Y}{\ectxext}{g}{p}}{g^\ast(\eft)}} 
-  & 
-\pullback{Y}{\stesysf}{g}{p\circ\eft}
-  \ar{d}{\cobesys{Y}{\ectxext}{g}{p}}
-  \\
-\pullback{Y}{\stesysf}{g}{p\circ\eft}
-  \ar{r}[swap]{\cobesys{Y}{\ectxext}{g}{p}} 
-  & 
-\pullback{Y}{\stesysc}{g}{p}
-\end{tikzcd}
-\end{equation*}
-\end{proof}
-
-The following construction is useful for defining extension homomorphisms into
-`higher' extension objects
-
-\begin{defn}\label{cobesys-existence}
-Consider a commutative diagram
-\begin{equation*}
-\begin{tikzcd}
-\stesys'
-  \ar{r}{f}
-  \ar{d}[swap]{p'}
-  &
-\stesys
-  \ar{d}{p}
-  \\
-Y \ar{r}[swap]{g}
-  &
-X
-\end{tikzcd}
-\end{equation*}
-Then we construct $[p,f]:\stesys'\to\cobesys{Y}{\stesys}{g}{p}$
-\begin{itemize}
-\item by defining $[p,f]_0:\stesysc'\to\pullback{Y}{\stesysc}{g}{p}$ be the uniqe
-morphism rendering the diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=large]
-\stesysc'
-  \ar[bend right=10]{ddr}[swap]{p'}
-  \ar[bend left=10]{rrd}{f_0}
-  \ar{dr}[near end]{[p,f]_0}
-  \\
-  {}&
-\pullback{Y}{\stesysc}{g}{p}
-  \ar{r}[swap]{\pullbackpr{2}{g}{p}}
-  \ar{d}{\pullbackpr{1}{g}{p}}
-  &
-\stesysc
-  \ar{d}{p}
-  \\
-  {}&
-Y \ar{r}[swap]{g}
-  &
-X
-\end{tikzcd}
-\end{equation*}
-commutative.
-\item by defining $[p,f]_1:\stesysf'\to\pullback{Y}{\stesysf}{g}{p\circ\eft}$ be the uniqe
-morphism rendering the diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=huge]
-\stesysf'
-  \ar[bend right=10]{ddr}[swap]{p'\circ\eft'}
-  \ar[bend left=10]{rrd}{f_1}
-  \ar{dr}[near end]{[p,f]_1}
-  \\
-  {}&
-\pullback{Y}{\stesysf}{g}{p\circ\eft}
-  \ar{r}[swap]{\pullbackpr{2}{g}{p\circ\eft}}
-  \ar{d}{\pullbackpr{1}{g}{p\circ\eft}}
-  &
-\stesysc
-  \ar{d}{p\circ\eft}
-  \\
-  {}&
-Y \ar{r}[swap]{g}
-  &
-X
-\end{tikzcd}
-\end{equation*}
-commutative.
-\item by defining $[p,f]^t:\stesyst'\to\pullback{Y}{\stesyst}{g}{p\circ\eft\circ\ebd}$ be the uniqe
-morphism rendering the diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=huge]
-\stesyst'
-  \ar[bend right=10]{ddr}[swap]{p'\circ\eft'\circ\ebd'}
-  \ar[bend left=10]{rrd}{f^t}
-  \ar{dr}[near end]{[p,f]^t}
-  \\
-  {}&
-\pullback{Y}{\stesysf}{g}{p\circ\eft\circ\ebd}
-  \ar{r}[swap]{\pullbackpr{2}{g}{p\circ\eft\circ\ebd}}
-  \ar{d}{\pullbackpr{1}{g}{p\circ\eft\circ\ebd}}
-  &
-\stesysc
-  \ar{d}{p\circ\eft\circ\ebd}
-  \\
-  {}&
-Y \ar{r}[swap]{g}
-  &
-X
-\end{tikzcd}
-\end{equation*}
-commutative.
-\end{itemize}
-\end{defn}
-
-\begin{thm}\label{cobesys-pullback}
-For every diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=large]
-\stesys'
-  \ar[bend right=10]{ddr}[swap]{p'}
-  \ar[bend left=10]{rrd}{f}
-  \ar[dotted]{dr}[near end]{[p',f]}
-  \\
-  {}&
-\cobesys{Y}{\stesys}{g}{p}
-  \ar{d}{\pullbackpr{1}{g}{p}}
-  \ar{r}[swap]{\pullbackpr{2}{g}{p}}
-  &
-\stesys
-  \ar{d}{p}
-  \\
-  {}&
-Y \ar{r}[swap]{g}
-  &
-X
-\end{tikzcd}
-\end{equation*}
-of which the outer square commutes, the pre-extension homomorphism $[p',f]$
-is unique with the property that it renders the whole diagram commutative.
-\end{thm}
-
-\begin{defn}\label{famfamstesys_into}
-Consider a commutative square
-\begin{equation*}
-\begin{tikzcd}
-\stesys'
-  \ar{r}{f}
-  \ar{d}[swap]{p}
-  &
-\famesys\stesys
-  \ar{d}{\eft}
-  \\
-\stesysf \ar{r}[swap]{\ectxext}
-  &
-\stesysc
-\end{tikzcd}
-\end{equation*}
-Then we construct
-\begin{equation*}
-[p,f]:\stesys'\to\famesys{\famesys{\stesys}}
-\end{equation*}
-as follows:
-\begin{itemize}
-\item let $[p,f]_0:\stesysc'\to\stesysf_2$ be the unique morphism rendering
-the diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=large]
-\stesysc' 
-  \ar[bend left=10]{rrd}{f_0}
-  \ar[swap,bend right=10]{ddr}{p}
-  \ar[dotted]{dr}[near end]{[p,f]_0}
-  \\
-  {}&
-\stesysf_2
-  \ar{r}[swap]{\pullbackpr{2}{\ectxext}{\eft}}
-  \ar{d}{\eft[2]}
-  &
-\stesysf
-  \ar{d}{\eft}
-  \\
-  {}&
-\stesysf
-  \ar{r}[swap]{\ectxext}
-  &
-\stesysc
-\end{tikzcd}
-\end{equation*}
-commutative.
-\item Let $[p,f]_1:\stesysf'\to\stesysf_3$ be the unique morphism rendering
-the diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=large]
-\stesysf'
-  \ar[bend left=10]{drr}{f_1}
-  \ar[swap]{dd}{\eft'}
-  \ar[dotted]{dr}[near end]{[p,f]_1}
-  \\
-  {}&
-\stesysf_3
-  \ar{r}[swap]{\pullbackpr{2}{\efamext}{\eft[2]}}
-  \ar{d}{\eft[2]}
-  &
-\stesysf_2
-  \ar{d}{\eft[2]}
-  \\
-\stesysc'
-  \ar{r}[swap]{[p,f]_0}
-  &
-\stesysf_2
-  \ar{r}[swap]{\efamext}
-  &
-\stesysf
-\end{tikzcd}
-\end{equation*}
-commutative.
-\item Let $[p,f]^t:\stesyst'\to\stesyst_3$ be the unique morphism rendering
-the diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=huge]
-\stesyst'
-  \ar[bend left=10]{drr}{f^t}
-  \ar[swap]{dd}{\eft'\circ\ebd'}
-  \ar[dotted]{dr}[near end]{[p,f]^t}
-  \\
-  {}&
-\stesyst_3
-  \ar{r}[swap]{\pullbackpr{2}{\efamext}{\eft[2]\circ\ebd[2]}}
-  \ar{d}[swap]{\pullbackpr{1}{\efamext}{\eft[2]\circ\ebd[2]}}
-  &
-\stesyst_2
-  \ar{d}{\eft[2]\circ\ebd[2]}
-  \\
-\stesysc'
-  \ar{r}[swap]{[p,f]_0}
-  &
-\stesysf_2
-  \ar{r}[swap]{\efamext}
-  &
-\stesysf
-\end{tikzcd}
-\end{equation*}
-commutative.
-\end{itemize}
-\end{defn}
-
-\begin{lem}
-Under the hypotheses of \autoref{famfamstesys_into}, $[p,f]$ is a pre-extension
-homomorphism. Moreover, it is the unique pre-extension homomorphism for which
-the diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=large]
-\stesys' 
-  \ar[bend left=10]{rrd}{f}
-  \ar[swap,bend right=10]{ddr}{p}
-  \ar[dotted]{dr}[near end]{[p,f]}
-  \\
-  {}&
-\famesys{\famesys{\stesys}}
-  \ar{r}[swap]{\pullbackpr{2}{\ectxext}{\eft}}
-  \ar{d}{\eft[2]}
-  &
-\famesys{\stesys}
-  \ar{d}{\eft}
-  \\
-  {}&
-\stesysf
-  \ar{r}[swap]{\ectxext}
-  &
-\stesysc
-\end{tikzcd}
-\end{equation*}
-commutes.
-\end{lem}
-
-\begin{lem}
-Suppose $f:\stesys\to \stesys'$ is a pre-extension homomorphism and consider a morphism
-$p:\stesys'\to X$ and $g:Y\to X$. Then the change of base 
-$g^\ast(f):\cobesys{Y}{\stesys}{g}{p\circ f_0}\to
-\cobesys{Y}{\stesys'}{g}{p}$ is a pre-extension morphism.
-\end{lem}
-
-\begin{lem}
-Let $\stesys$ be a pre-extension algebra and consider $p:\stesysc\rightarrow X\leftarrow Y:g$.
-Then there is an isomorphism
-\begin{equation*}
-\varphi:\famesys{\cobesys{Y}{\stesys}{g}{p}}
-  \simeq
-\cobesys{Y}{\famesys{\stesys}}{g}{p\circ\eft}
-\end{equation*}
-uniquely determined by
-\end{lem}
-
-\begin{proof}
-This follows from the pasting lemma for pullbacks.
-\end{proof}
-
-\subsection{Pre-weakening objects}
-\begin{defn}
-Let $\stesys$ be an extension object in $\cat{C}$. A pre-weakening operation
-on $\stesys$ is an extension homomorphism 
-$ \mathbf{w}(\stesys)
-    :
-  \cobesys{\stesysf}{\famesys{\stesys}}{\eft}{\eft}
-    \to
-  \famesys{\famesys{\stesys}}$
-for which the diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=large]
-\cobesys{\stesysf}{\famesys{\stesys}}{\eft}{\eft}
-  \ar{r}{\mathbf{w}(\stesys)}
-  \ar{dr}[swap]{\pullbackpr{1}{\eft}{\eft}}
-  &
-\famesys{\famesys{\stesys}}
-  \ar{d}{\eft[2]}
-  \\
-& \stesysf
-\end{tikzcd}
-\end{equation*}
-commutes.
-\end{defn}
-
-\begin{defn}
-Let $\stesys$ be an extension object with pre-weakening operation
-$\mathbf{w}(\stesys)$. Then $\famesys{\stesys}$ has the pre-weakening operation
-$\mathbf{w}(\famesys{\stesys})$ which is uniquely determined by rendering the
-diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=large]
-\cobesys{\stesysf_2}{\famesys{\famesys{\stesys}}}{\eft[2]}{\eft[2]}
-  \ar{rr}{%
-      \pullback{\beta_1}{\boldsymbol{\beta}}{\eft}{\eft}
-    }
-  \ar[bend right]{ddr}[swap]{\pullbackpr{1}{\eft[2]}{\eft[2]}}
-  \ar[dotted]{dr}{\mathbf{w}(\famesys{\stesys})}
-  &
-  {}&
-\cobesys{\stesysf}{\famesys{\stesys}}{\eft}{\eft}
-  \ar{r}{\mathbf{w}(\stesys)}
-  &
-\famesys{\famesys{\stesys}}
-  \ar{d}{\boldsymbol{\beta}}
-  \\
-  {}&
-\famesys{\famesys{\famesys{\stesys}}}
-  \ar{r}{\boldsymbol{\beta}_\mathbf{2}}
-  \ar{d}[swap]{\eft[3]}
-  &
-\famesys{\famesys{\stesys}}
-  \ar{d}{\eft[2]}
-  \ar{r}{\boldsymbol{\beta}}
-  &
-\famesys{\stesys}
-  \ar{d}{\eft}
-  \\
-  {}&
-\stesysf_2
-  \ar{r}[swap]{\efamext}
-  &
-\stesysf
-  \ar{r}[swap]{\ectxext}
-  &
-\stesysc
-\end{tikzcd}
-\end{equation*}
-commutative.
-\end{defn}
-
-\begin{defn}
-A pre-weakening object $\stesys$ in $\cat{C}$ is an extension object $\stesys$ 
-with a pre-weakening operation 
-$ \mathbf{w}(\stesys)
-    :
-  \cobesys{\stesysf}{\famesys{\stesys}}{\eft}{\eft}
-    \to
-  \famesys{\famesys{\stesys}}$
-for which the diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=15em]
-\cobesys{\stesysf_2}{\famesys{\stesys}}{\eft\circ\eft[2]}{\eft}
-  \ar[bend right=10]{dr}[swap]%
-    { [ \pullbackpr{1}{\eft\circ\eft[2]}{\eft},%
-        \mathbf{w}(\stesys)%
-          \circ%
-        (\pullback{\efamext}{\catid{\famesys{\stesys}}}{\eft}{\eft})%
-        ]%
-      }
-  \ar{r}{
-    [ \pullbackpr{1}{\eft\circ\eft[2]}{\eft},%
-      \mathbf{w}(\stesys)%
-        \circ%
-      (\pullback{\eft[2]}{\catid{\famesys{\stesys}}}{\eft}{\eft})%
-      ]}%
-  &
-\cobesys{\stesysf_2}{\famesys{\famesys{\stesys}}}{\eft[2]}{\eft[2]}
-  \ar{d}{\mathbf{w}(\famesys{\stesys})}
-  \\
-  {}&
-\famesys{\famesys{\famesys{\stesys}}}
-\end{tikzcd}
-\end{equation*}
-commutes. This condition is called \emph{Currying for weakening}.
-\end{defn}
-
-\begin{lem}
-If $\stesys$ is a pre-weakening algebra, then so is $\famesys{\stesys}$. 
-\end{lem}
-
-\begin{proof}
-\end{proof}
-
-\begin{defn}
-A pre-weakening morphism between preweakening objects $\stesys$ and $\stesys'$ is an
-extension homomorphism $f:\stesys\to \stesys'$ such that additionally the diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=large]
-\cobesys{\stesysf}{\famesys{\stesys}}{\eft}{\eft}
-  \ar{d}[swap]{\mathbf{w}(\stesys)}
-  \ar{r}{\pullback{f_1}{\famehom{f}}{\eft'}{\eft'}}
-  &
-\cobesys{\stesysf'}{\famesys{\stesys'}}{\eft'}{\eft'}
-  \ar{d}{\mathbf{w}(\stesys')}
-  \\
-\famesys{\famesys{\stesys}}
-  \ar{r}[swap]{\famehom{\famehom{f}}}
-  &
-\famesys{\famesys{\stesys'}}
-\end{tikzcd}
-\end{equation*}
-commutes.
-\end{defn}
-
-\begin{defn}
-Let $\stesys$ be a pre-weakening algebra and consider $p:\stesysc\rightarrow X\leftarrow Y:p$.
-Then we define
-\begin{equation*}
-\mathbf{w}(\cobesys{Y}{\stesys}{g}{p})
-  :
-\cobesys
-  { (\pullback{Y}{\stesysf}{g}{p\circ\eft})}
-  { \famesys{\cobesys{Y}{\stesys}{g}{p}}}
-  { g^\ast(\eft)}
-  { g^\ast(\eft)}
-  \to
-\famesys{\famesys{\cobesys{Y}{\stesys}{g}{p}}}
-\end{equation*}
-to be the unique extension homomorphism rendering the diagram
-\begin{equation*}
-\begin{tikzcd}
-\cobesys
-  { \pullback{Y}{\stesysf}{g}{p\circ\eft}}
-  { \famesys{\cobesys{Y}{\stesys}{g}{p}}}
-  { g^\ast(\eft)}
-  { g^\ast(\eft)}
-  \ar[bend right]{ddr}[swap]{\pullbackpr{1}{g^\ast(\eft)}{g^\ast(\eft)}}
-  \ar{rr}{%
-    \pullback
-      { \pullbackpr{2}{g}{p\circ\eft}}
-      { \boldsymbol{\pi}_\mathbf{2}(g,p\circ\eft)}
-      { \eft}
-      { \eft}
-    }
-  \ar[dotted]{dr}{\mathbf{w}(\cobesys{Y}{\stesys}{g}{p})}
-  &
-  {}&
-\cobesys{\stesysf}{\famesys{\stesys}}{\eft}{\eft}
-  \ar{r}{\mathbf{w}(\stesys)}
-  &
-\famesys{\famesys{\stesys}}
-  \ar{d}{\boldsymbol{\beta}}
-  \\
-  {}&
-\famesys{\famesys{\cobesys{Y}{\stesys}{g}{p}}}
-  \ar{r}{\beta}
-  \ar{d}{g^\ast(\eft)_1}
-  &
-\famesys{\cobesys{Y}{\stesys}{g}{p}}
-  \ar{r}{\boldsymbol{\pi}_2(g,p\circ\eft)}
-  \ar{d}{g^\ast(\eft)}
-  &
-\famesys{\stesys}
-  \ar{d}{\eft}
-  \\
-  {}&
-\pullback{Y}{\stesysf}{g}{p\circ\eft}
-  \ar{r}[swap]{\cobesys{Y}{\ectxext}{g}{p}}
-  &
-\pullback{Y}{\stesysc}{g}{p}
-  \ar{r}[swap]{\pullbackpr{2}{g}{p}}
-  &
-\stesysc
-\end{tikzcd}
-\end{equation*}
-commutative.
-\end{defn}
-
-\subsection{Weakening objects}
-Since we have shown that the property of being a pre-weakening object is closed
-under the relevant operations, we can make the following definition:
-
-\begin{defn}
-A weakening object is a pre-weakening object $\stesys$ with the property that
-$\mathbf{w}(\stesys)$ is a pre-weakening morphism.
-\end{defn}
-
-\begin{defn}
-A weakening homomorphism is a pre-weakening homomorphism such that the domain
-and codomain are weakening objects.
-\end{defn}
-
-\begin{thm}
-Suppose $\stesys$ is a weakening object, then so is $\famesys{\stesys}$
-\end{thm}
-
-\begin{thm}
-The change of base of any weakening object is again a weakening object.
-\end{thm}
-
-\subsection{Projection objects}
-\begin{defn}
-A pre-projection object is a weakening object $\stesys$ for which there is a term
-$\mathbf{i}:\stesysf\to \stesyst_2$ such that the diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=large]
-\stesysf \ar{r}{\mathbf{i}} \ar{d}[swap]{\Delta_{\eft}} & \stesyst_2 \ar{d}{\ebd[2]}\\
-\pullback{\stesysf}{\stesysf}{\eft}{\eft} \ar{r}[swap]{w(\stesys)_0} & \stesysf_2
-\end{tikzcd}
-\end{equation*}
-commutes. In this diagram $\Delta_{\eft}:\stesysf\to \pullback{\stesysf}{\stesysf}{\eft}{\eft}$ is the diagonal.
-\end{defn}
-
-\begin{defn}
-A pre-projection homomorphism from $\stesys$ to $\stesys'$ is a weakening homomorphism
-$f:\stesys\to \stesys'$ such that the square
-\begin{equation*}
-\begin{tikzcd}[column sep=large]
-\stesyst_2
-  \ar{r}{f^t_1}
-  &
-\stesyst_2'
-  \\
-\stesysf \ar{r}[swap]{f_1}
-  \ar{u}{\mathbf{i}}
-  &
-\stesysf'
-  \ar{u}[swap]{\mathbf{i}'}
-\end{tikzcd}
-\end{equation*}
-commutes
-\end{defn}
-
-\begin{lem}
-The change of base of a pre-projection object is again a pre-projection object.
-\end{lem}
-
-\begin{lem}
-If $CFT$ is a pre-projection object, then so is $\mathbf{F}_{CFT}$, where
-$\mathbf{F}_{\mathbf{i}}$ is defined to be $F\times_{e_0,c}\mathbf{i}$ is
-a pre-projection algebra.
-\end{lem}
-
-\begin{defn}
-A projection algebra is a pre-projection algebra for which weakening is a
-pre-projection homomorphism.
-\end{defn}
-
-\begin{cor}
-The change of base of a projection object is again a projection object.
-\end{cor}
-
-\begin{cor}
-If $CFT$ is a projection object, then so is $\mathbf{F}_{CFT}$, where
-$\mathbf{F}_{\mathbf{i}}$ is defined to be $F\times_{e_0,c}\mathbf{i}$ is
-a projection algebra.
-\end{cor}
-
-\subsection{Substitution objects}
-
-\begin{defn}
-A \emph{pre-substitution} for an extension object $\stesys$ is an
-extension homomorphism
-\begin{equation*}
-\mathbf{s}(\stesys):\cobesys{\stesyst}{\famesys{\famesys{\stesys}}}{\ebd}{\eft[2]}\to \famesys{\stesys}
-\end{equation*}
-for which the square
-\begin{equation*}
-\begin{tikzcd}[column sep=large]
-\pullback{\stesyst}{\stesysf_2}{\ebd}{\eft[2]}
-  \ar{r}{s(\stesys)_0}
-  \ar{d}[swap]{\ebd\circ\pullbackpr{1}{\ebd}{\eft[2]}}
-  &
-\stesysf 
-  \ar{d}{\eft}
-  \\
-\stesysf 
-  \ar{r}[swap]{\eft}
-  &
-\stesysc
-\end{tikzcd}
-\end{equation*}
-commutes. A \emph{pre-substitution object} is an extension object
-together with a pre-substitution.
-\end{defn}
-
-\begin{defn}
-A \emph{pre-substitution homomorphism} is an extension homomorphism $f:\stesys'\to \stesys$
-for which the square
-\begin{equation*}
-\begin{tikzcd}[column sep=huge]
-\cobesys{\stesyst'}{\famesys{\famesys{\stesys'}}}{\ebd'}{\eft[2]'}
-  \ar{r}{\pullback{f^t}{\famehom{\famehom{f}}}{\ebd}{\eft[2]}}
-  \ar{d}[swap]{\mathbf{s}'(\stesys')}
-  &
-\cobesys{\stesyst}{\famesys{\famesys{\stesys}}}{\ebd}{\eft[2]}
-  \ar{d}{\mathbf{s}(\stesys)}
-  \\
-\famesys{\stesys'}
-  \ar{r}[swap]{\famehom{f}}
-  &
-\famesys{\stesys}
-\end{tikzcd}
-\end{equation*}
-commutes.
-\end{defn}
-
-\begin{lem}
-The change of base of a pre-substitution object is again a pre-substitution object.
-\end{lem}
-
-\begin{lem}
-If $\stesys$ is a pre-substitution object, then so is $\famesys{\stesys}$ with
-$\mathbf{s}(\famesys{\stesys})$ defined to be the unique extension homomorphism
-rendering the diagram
-\begin{equation*}
-\begin{tikzcd}
-\cobesys{\stesyst_2}{\famesys{\famesys{\famesys{\stesys}}}}{\ebd[2]}{\eft[3]}
-  \ar[dotted]{dr}{\mathbf{s}(\famesys{\stesys})}
-  \ar{rr}{\pullback{\pullbackpr{2}{\ectxext}{\eft\circ\ebd}}{\boldsymbol{\beta}_\mathbf{2}}{\ebd}{\eft[2]}}
-  \ar{dd}[swap]{\ebd[2]\circ\pullbackpr{1}{\ebd[2]}{\eft[3]}}
-  &
-  {}&
-\cobesys{\stesyst}{\famesys{\famesys{\stesys}}}{\ebd}{\eft[2]}
-  \ar{d}{\mathbf{s}(\stesys)}
-  \\
-  {}&
-\famesys{\famesys{\stesys}}
-  \ar{r}{\boldsymbol{\beta}}
-  \ar{d}{\eft[2]}
-  &
-\famesys{\stesys}
-  \ar{d}{\eft}
-  \\
-\stesysf_2
-  \ar{r}[swap]{\eft[2]}
-  &
-\stesysf
-  \ar{r}[swap]{\ectxext}
-  &
-\stesysc
-\end{tikzcd}
-\end{equation*}
-commutative.
-\end{lem}
-
-\begin{proof}
-The requirement on pre-substitutions holds by construction.
-\end{proof}
-
-It makes sense now to consider the possibility that the pre-substitution
-itself is a pre-substitution homomorphism.
-
-\begin{defn}
-A \emph{substitution object} is a pre-substitution object for which substitution is
-a pre-substitution homomorphism.
-\end{defn}
-
-\begin{cor}
-The change of base of a substitution object is again a substitution object.
-\end{cor}
-
-\begin{cor}
-If $\stesys$ is a substitution object, then so is $\famesys{\stesys}$.
-\end{cor}
-
-\subsection{Extension objects with empty context and families}
-
-\begin{defn}
-An extension object $\stesys$ is said to have \emph{empty families} if there
-is a section
-\begin{equation*}
-\phi_1(\stesys):\stesysc\to\stesysf
-\end{equation*}
-of $\eft$, satisfying the following additional properties:
-\begin{enumerate}
-\item $\phi_1(\stesys)$ is also a section of $\ectxext$.
-\item The unique morphism $\phi_2$ rendering the diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=huge]
-\stesysf
-  \ar[bend left=20]{drr}{\phi_1(\stesys)\circ\ectxext}
-  \ar[equals,bend right=20]{ddr}
-  \ar[dotted]{dr}{\phi_2}
-  \\
-  {}&
-\stesysf_2
-  \ar{d}{\eft[2]}
-  \ar{r}{\pullbackpr{2}{\ectxext}{\eft}}
-  &
-\stesysf
-  \ar{d}{\eft}
-  \\
-  {}&
-\stesysf
-  \ar{r}[swap]{\ectxext}
-  &
-\stesysc
-\end{tikzcd}
-\end{equation*}
-commutative, is a section of $\efamext$.
-\item The unique morphism $\iota_1$ rendering the diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=huge]
-\stesysf
-  \ar[equals,bend left=20]{drr}
-  \ar[bend right=20]{ddr}[swap]{\phi_1(\stesys)\circ\eft}
-  \ar[dotted]{dr}{\iota_1}
-  \\
-  {}&
-\stesysf_2
-  \ar{d}{\eft[2]}
-  \ar{r}[swap]{\pullbackpr{2}{\ectxext}{\eft}}
-  &
-\stesysf
-  \ar{d}{\eft}
-  \\
-  {}&
-\stesysf
-  \ar{r}[swap]{\ectxext}
-  &
-\stesysc
-\end{tikzcd}
-\end{equation*}
-commutative, is a section of $\efamext$.
-\end{enumerate}
-\end{defn}
-
-\begin{defn}
-A homomorphism of extension objects with empty families is an extension
-homomorphism $f:\stesys'\to\stesys$ for which the diagram
-\begin{equation*}
-\begin{tikzcd}
-\stesysf'
-  \ar{r}{f_1}
-  &
-\stesysf
-  \\
-\stesysc'
-  \ar{u}{\phi_1(\stesys')}
-  \ar{r}[swap]{f_0}
-  &
-\stesysc
-  \ar{u}[swap]{\phi_1(\stesys)}
-\end{tikzcd}
-\end{equation*}
-commutes.
-\end{defn}
-
-\begin{lem}
-Suppose $\stesys$ is an extension object with empty families. Then
-$\famesys{\stesys}$ is an extension object with empty families, with
-$\phi_1(\famesys{\stesys}):=\phi_2$. 
-\end{lem}
-
-\begin{lem}
-Suppose $\stesys$ is an extension object with empty families and consider
-$p:\stesysc\rightarrow X\leftarrow Y:p$. Then
-$\cobesys{Y}{\stesys}{g}{p}$ is an extension object with empty families
-with $\phi_1(\cobesys{Y}{\stesys}{g}{p}):=g^\ast(\phi_1)$.
-\end{lem}
-
-\subsection{E-objects}
-\begin{defn}
-An \emph{E-object} is an extension object with the structure of a projection object,
-the structure of a substitution object and which has an empty context and families,
-such that additionally:
-\begin{enumerate}
-\item substitution is a projection homomorphism
-\item weakening is a substitution homomorphism
-\item both weakening and substitution are empty-CF homomorphisms.
-\item 
-\end{enumerate}
-\end{defn}

### 2015-02-10 Algebras/projections.tex
large reorganisation

--- 2014-11-03 Algebras/article-models-cat.tex
@@
-\section{E-objects in categories with finite limits}
-In this section we assume that $\cat{C}$ is a finitely complete category and
-whenever we write a pullback, we assume that it is chosen. Recall that for
-any morphism $f:A\to B$ in a category $\cat{C}$ with chosen pullbacks, there
-is a functor
-\begin{equation*}
-f^\ast : \cat{C}/B\to\cat{C}/A.
-\end{equation*}
-As usual, when $g:X\to B$ is a morphism, we will write $f^\ast(X)$ for the
-domain of $f^\ast(g)$. When there is more than one morphism $X\to B$ involved,
-as will be the case below, we will write $\pullback{A}{X}{f}{g}$. The projections
-will be written as $\pullbackpr{1}{f}{g}$ and $\pullbackpr{2}{f}{g}$. So in this notation, a
-typical pullback diagram has the following form:
-\begin{equation*}
-\begin{tikzcd}[column sep=large]
-\pullback{A}{X}{f}{g}
-  \ar{r}{\pullbackpr{1}{f}{g}}
-  \ar{d}[swap]{\pullbackpr{2}{f}{g}}
-  &
-A \ar{d}{f}
-  \\
-X \ar{r}[swap]{g}
-  &
-B
-\end{tikzcd}
-\end{equation*}
-Also, when we have a commutative diagram of the form
-\begin{equation*}
-\begin{tikzcd}
-A \ar{r}{f}
-  \ar{d}{a}
-  &
-X \ar{d}
-  & 
-B \ar{l}[swap]{g}
-  \ar{d}{b}
-  \\
-A'
-  \ar{r}[swap]{f'}
-  &
-X'
-  &
-B'
-  \ar{l}{g'}
-\end{tikzcd}
-\end{equation*}
-we will denote the unique map from $\pullback{A}{B}{f}{g}$ to $\pullback{A'}{B'}{f'}{g'}$
-such that the diagram
-\begin{equation*}
-\begin{tikzcd}
-  {}
-  & 
-\pullback{A'}{B'}{f'}{g'}
-  \ar{dd}
-  \ar{rr}
-  &
-  &
-B'
-  \ar{dd}{g'}
-  \\
-\pullback{A}{B}{f}{g}
-  \ar{dd}
-  \ar[crossing over]{rr}
-  \ar[dotted]{ur}{\pullback{a}{b}{f'}{g'}}
-  &
-  &
-B \ar{ur}{b}
-  \\
-  {}
-  &
-A'
-  \ar{rr}
-  &
-  &
-X'
-  \\
-A \ar{rr}[swap]{f}
-  \ar{ur}{a}
-  &
-  &
-X \ar[crossing over,leftarrow]{uu}[near end,swap]{g}
-  \ar{ur}
-\end{tikzcd}
-\end{equation*}
-commutes, by $\pullback{a}{b}{f'}{g'}$. In the current work, we shall
-write $A\times B$ for the pullback of $A\rightarrow 1\leftarrow B$, and
-$\pi_1$ and $\pi_2$ for its projections (thus, no separate choice of
-cartesian products is made).
-
-\subsection{Extension objects}
+\subsection{Projection algebras}
 \begin{defn}
-A \emph{pre-extension object $\stesys$ in $\cat{C}$} consists of a \emph{fundamental structure}, which is a diagram of the form
-\begin{equation*}
-\begin{tikzcd}
-\stesyst
-  \ar{d}[swap]{\ebd}
-  \\
-\stesysf
-  \ar{d}[swap]{\eft}
-  \\
-\stesysc
-\end{tikzcd}
-\end{equation*}
-in $\cat{C}$ together with the \emph{context extension} and \emph{family extension} operations
-\begin{align*}
-\ectxext &:\stesysf\to \stesysc\\
-\efamext & :\stesysff\to \stesysf,
-\end{align*}
-respectively, such that the diagram
-\begin{equation*}
-\begin{tikzcd}
-\stesysf_2 
-  \ar{r}{\efamext} 
-  \ar{d}[swap]{\eft[2]} 
-  & 
-\stesysf 
-  \ar{d}{\eft}
-  \\
-\stesysf
-  \ar{r}[swap]{\eft} 
-  & 
-\stesysc
-\end{tikzcd}
-\end{equation*}
-commutes.
-\end{defn}
-
-\begin{defn}
-We introduce the following notation:
-\begin{align*}
-\stesysf_2 
-  & := \stesysff
-  \\
-\eft[2] 
-  & := \pullbackpr{1}{\ectxext}{\eft} : \stesysf_2\to\stesysf
-  \\
-\stesysf_3 & := \pullback{\stesysf_2}{\stesysf_2}{\efamext}{\eft[2]}
-  \\
-\eft[3]
-  & := \pullbackpr{1}{\efamext}{\eft[2]} : \stesysf_3\to\stesysf_2.
-\end{align*}
-Then it follows that the outer square in the diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=large]
-\stesysf_3
-  \ar[dotted]{dr}{\eext{2}}
-  \ar{rr}{\pullback{\pullbackpr{2}{\ectxext}{\eft}}{\pullbackpr{2}{\ectxext}{\eft}}{\ectxext}{\eft}}
-  \ar{dd}[swap]{\eft[3]}
-  & 
-  &
-\stesysf_2
-  \ar{d}{\efamext}
-  \\
-  &
-\stesysf_2
-  \ar{d}[swap]{\eft[2]}
-  \ar{r}{\pullbackpr{2}{\ectxext}{\eft}}
-  &
-\stesysf
-  \ar{d}{\eft}
-  \\
-\stesysf_2
-  \ar{r}[swap]{\eft[2]}
-  &
-\stesysf
-  \ar{r}[swap]{\ectxext}
-  &
-\stesysc
-\end{tikzcd}
-\end{equation*}
-commutes. We define $\eext{2}$ to be the unique morphism rendering the above diagram
-commutative. Now we may continue to define
-\begin{align*}
-\stesysf_4 
-  & := 
-\pullback{\stesysf_3}{\stesysf_3}{\eext{2}}{\eft[3]}
-  \\
-\eft[4] 
-  & := 
-\pullbackpr{1}{\eext{2}}{\eft[3]}.
-\end{align*}
-Then we see that the outer square of the diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=large]
-\stesysf_4
-  \ar[dotted]{dr}{\eext{3}}
-  \ar{rr}{\pullback{\pullbackpr{2}{\efamext}{\eft[2]}}{\pullbackpr{2}{\efamext}{\eft[2]}}{\efamext}{\eft[2]}}
-  \ar{dd}[swap]{\eft[4]}
-  & 
-  &
-\stesysf_3
-  \ar{d}{\eext{2}}
-  \\
-  &
-\stesysf_3
-  \ar{d}[swap]{\eft[3]}
-  \ar{r}{\pullbackpr{2}{\efamext}{\eft[2]}}
-  &
-\stesysf_2
-  \ar{d}{\eft[2]}
-  \\
-\stesysf_3
-  \ar{r}[swap]{\eft[3]}
-  &
-\stesysf_2
-  \ar{r}[swap]{\efamext}
-  &
-\stesysf
-\end{tikzcd}
-\end{equation*}
-commutes,
-so we may define $\eext{3}$ to be the unique map which renders the diagram
-commutative. It
-is straightforward to continue this process by induction, but we shall need not
-go any further in this article.
-\end{defn}
-
-\begin{defn} 
-An extension object is a pre-extension object $\stesys$ for which 
-the diagrams
-\begin{equation*}
-\begin{tikzcd}
-\stesysf_2 
-  \ar{d}[swap]{\pullbackpr{2}{\ectxext}{\eft}} 
-  \ar{r}{\efamext} 
-  & 
-\stesysf 
-  \ar{d}{\ectxext}
-  \\
-\stesysf 
-  \ar{r}[swap]{\ectxext} 
-  & 
-\stesysc
-\end{tikzcd}
-\qquad
-\begin{tikzcd}
-\stesysf_3
-  \ar{d}[swap]{\pullbackpr{2}{\efamext}{\eft[2]}}
-  \ar{r}{\eext{2}}
-  & 
-\stesysf_2 
-  \ar{d}{\efamext} 
-  \\
-\stesysf_2 
-  \ar{r}[swap]{\efamext} 
-  &
-\stesysf
-\end{tikzcd}
-\end{equation*}
-commute.
-\end{defn}
-
-\begin{comment}
-\begin{lem}
-There exists an isomorphism $\alpha$ such that the triangle
-\begin{equation*}
-\begin{tikzcd}[column sep=tiny]
-\pullback{\stesysf}{\stesysf_2}{\ectxext}{\eft\circ\eft[2]}
-  \ar[dotted]{rr}{\alpha}
-  \ar{dr}[swap]{\pullback{\catid{\stesysf}}{\efamext}{\ectxext}{\eft}}
-  &
-  &
-\stesysf_3
-  \ar{dl}{\eext{2}}
-  \\
-& \stesysf_2
-\end{tikzcd}
-\end{equation*}
-commutes
-\end{lem}
-
-\begin{proof}
-There is a unique morphism $\alpha:
-\pullback{\stesysf}{\stesysf_2}{\ectxext}{\eft\circ\eft[2]}\to\stesysf_3$
-rendering the diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=large]
-\pullback{\stesysf}{\stesysf_2}{\ectxext}{\eft\circ\eft[2]}
-  \ar[bend left=10,yshift=.5ex]{drrr}{\pullbackpr{2}{\ectxext}{\eft}\circ\pullbackpr{2}{\ectxext}{\eft\circ\eft[2]}}
-  \ar[bend right=10]{ddr}[swap]{\pullback{\catid{\stesysf}}{\eft[2]}{\ectxext}{\eft}}
-  \ar[dotted]{dr}{\alpha}
-  \\
-& \stesysf_3
-  \ar{r}{\pullbackpr{2}{\efamext}{\eft[2]}}
-  \ar{d}{\eft[3]}
-  &
-\stesysf_2
-  \ar{d}[swap]{\eft[2]}
-  \ar{r}[swap]{\pullbackpr{2}{\ectxext}{\eft}}
-  &
-\stesysf
-  \ar{d}{\eft}
-  \\
-{} & \stesysf_2
-  \ar{r}[swap]{\efamext}
-  &
-\stesysf
-  \ar{r}[swap]{\ectxext}
-  &
-\stesysc
-\end{tikzcd}
-\end{equation*}
-\end{proof}
-\end{comment}
-
-\begin{defn}
-Suppose $\stesys$ is a pre-extension object of $\cat{C}$. Then we define the pre-extension object
-$\famesys{\stesys}$ to consist of the fundamental structure
-\begin{equation*}
-\begin{tikzcd}
-\stesyst_2
-  \ar{d}{\ebd[2]}
-  \\
-\stesysf_2
-  \ar{d}{\eft[2]}
-  \\
-\stesysf
-\end{tikzcd}
-\end{equation*}
-where
-\begin{align*}
-\stesyst_2 
-  & := \pullback{\stesysf}{\stesyst}{\ectxext}{\eft\circ\ebd}
-  \\
-\ebd[2]
-  & := \ectxext^\ast(\ebd),
-\end{align*}
-with the extension operations
-\begin{align*}
-\efamext 
-  & 
-  : \stesysf_2\to\stesysf\\
-\eext{2} & : \stesysf_3\to\stesysf_2.
-\end{align*}
-\end{defn}
-
-In \autoref{famextobj} below we shall show that $\famesys{\stesys}$ is an
-extension algebra whenever $\stesys$ is an extension algebra. We shall need
-a handful of lemmas to give the proof.
-
-\begin{defn}
-We define
-\begin{align*}
-\beta_1 
-  & := 
-\pullbackpr{2}{\ectxext}{\eft}
-  & &
-  : \stesysf_2\to\stesysf
-  \\
-\beta_2
-  & :=
-\pullback{\beta_1}{\beta_1}{\ectxext}{\eft}
-  & &
-  : \stesysf_3\to\stesysf_2
-  \\
-\beta_3
-  & :=
-\pullback{\beta_2}{\beta_2}{\efamext}{\eft[2]}
-  & &
-  : \stesysf_4\to\stesysf_3.
-\end{align*}
-\end{defn}
-
-\begin{lem}
-Let $\stesys$ be a pre-extension object. Then the square
-\begin{equation*}
-\begin{tikzcd}[column sep=10em]
-\stesysf_4
-  \ar{r}{\pullback{\pullbackpr{2}{\efamext}{\eft[2]}}{\pullbackpr{2}{\efamext}{\eft[2]}}{\efamext}{\eft[2]}}
-  \ar{d}[swap]{\beta_3}
-  &
-\stesysf_3
-  \ar{d}{\beta_2}
-  \\
-\stesysf_3
-  \ar{r}[swap]{\beta_2}
-  &
-\stesysf_2
-\end{tikzcd}
-\end{equation*}
-commutes.
-\end{lem}
-
-\begin{proof}
-Left to the reader.
-\end{proof}
-
-\begin{comment}
-\begin{proof}
-It is straightforward to verify the equalities
-\begin{align*}
-\pullbackpr{1}{\ectxext}{\eft}\circ\beta_2\circ
-  (\pullback{\pullbackpr{2}{\efamext}{\eft[2]}}{\pullbackpr{2}{\efamext}{\eft[2]}}{\efamext}{\eft[2]})
-  & =
-\pullbackpr{1}{\ectxext}{\eft}\circ\beta_2\circ\beta_3
-  \\
-\pullbackpr{2}{\ectxext}{\eft}\circ\beta_2\circ
-  (\pullback{\pullbackpr{2}{\efamext}{\eft[2]}}{\pullbackpr{2}{\efamext}{\eft[2]}}{\efamext}{\eft[2]})
-  & =
-\pullbackpr{2}{\ectxext}{\eft}\circ\beta_2\circ\beta_3.\qedhere
-\end{align*}
-\end{proof}
-\end{comment}
-
-Note that the square
-\begin{equation*}
-\begin{tikzcd}
-\stesysf_3
-  \ar{r}{\beta_2}
-  \ar{d}[swap]{\eext{2}}
-  &
-\stesysf_2
-  \ar{d}{\efamext}
-  \\
-\stesysf_2
-  \ar{r}[swap]{\beta_1}
-  &
-\stesysf
-\end{tikzcd}
-\end{equation*}
-commutes by definition. We have a similar result relating $\eext{3}$ and
-$\eext{2}$.
-
-\begin{lem}
-Let $\stesys$ be a pre-extension object. Then the square
-\begin{equation*}
-\begin{tikzcd}
-\stesysf_4
-  \ar{r}{\beta_3}
-  \ar{d}[swap]{\eext{3}}
-  &
-\stesysf_3
-  \ar{d}{\eext{2}}
-  \\
-\stesysf_3
-  \ar{r}[swap]{\beta_2}
-  &
-\stesysf_2
-\end{tikzcd}
-\end{equation*}
-commutes.
-\end{lem}
-
-\begin{proof}
-Left to the reader.
-\end{proof}
-\begin{comment}
-\begin{proof}
-It is straightforward to verify the equalities
-\begin{align*}
-\pullbackpr{1}{\ectxext}{\eft}\circ\beta_2\circ\eext{3}
-  & = \beta_1\circ\eft[3]\circ\eft[4]
-  \\
-\pullbackpr{1}{\ectxext}{\eft}\circ\eext{2}\circ\beta_3
-  & = \beta_1\circ\eft[3]\circ\eft[4].
-\end{align*}
-Thus, it remains to verify that
-\begin{equation*}
-\pullbackpr{2}{\ectxext}{\eft}\circ\beta_2\circ\eext{3}
-  = \pullbackpr{2}{\ectxext}{\eft}\circ\eext{2}\circ\beta_3.
-\end{equation*}
-It is straightforward to see that the diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=large]
-\stesysf_4
-  \ar{dd}[swap]{\pullback{\pullbackpr{2}{\efamext}{\eft[2]}}{\pullbackpr{2}{\efamext}{\eft[2]}}{\efamext}{\eft[2]}}
-  \ar{r}{\eext{3}}
-  &
-\stesysf_3
-  \ar{r}{\beta_2}
-  \ar{d}[swap]{\pullbackpr{2}{\efamext}{\eft[2]}}
-  &
-\stesysf_2
-  \ar{dd}{\pullbackpr{2}{\ectxext}{\eft}}
-  \\
-  {} &
-\stesysf_2
-  \ar{dr}{\beta_1}
-  \\
-\stesysf_3
-  \ar{ur}{\eext{2}}
-  \ar{r}[swap]{\beta_2}
-  &
-\stesysf_2
-  \ar{r}[swap]{\efamext}
-  &
-\stesysf
-\end{tikzcd}
-\end{equation*}
-commutes. It is likewise straightforward to see that the diagram
-\begin{equation*}
-\begin{tikzcd}
-\stesysf_4
-  \ar{r}{\beta_3}
-  \ar{d}[swap]{\pullback{\pullbackpr{2}{\efamext}{\eft[2]}}{\pullbackpr{2}{\efamext}{\eft[2]}}{\efamext}{\eft[2]}}
-  &
-\stesysf_3
-  \ar{r}{\eext{2}}
-  \ar{d}[swap]{\beta_2}
-  &
-\stesysf_2
-  \ar{d}{\pullbackpr{2}{\ectxext}{\eft}}
-  \\
-\stesysf_3
-  \ar{r}[swap]{\beta_2}
-  &
-\stesysf_2
-  \ar{r}[swap]{\efamext}
-  &
-\stesysf
-\end{tikzcd}
-\end{equation*}
-commutes, completing our goal.
-\end{proof}
-\end{comment}
-
-\begin{thm}[Local extension structure]\label{famextobj}
-If $\stesys$ is an extension object, then so is $\famesys{\stesys}$.
-\end{thm}
-
-\begin{proof}
-Note that the diagram
-\begin{equation*}
-\begin{tikzcd}
-\stesysf_3
-  \ar{d}[swap]{\pullbackpr{2}{\efamext}{\eft[2]}}
-  \ar{r}{\eext{2}}
-  & 
-\stesysf_2 
-  \ar{d}{\efamext} 
-  \\
-\stesysf_2 
-  \ar{r}[swap]{\efamext} 
-  &
-\stesysf
-\end{tikzcd}
-\end{equation*}
-commutes by assumption. For the second condition, we have to show that the
-diagram
-\begin{equation*}
-\begin{tikzcd}
-\stesysf_4
-  \ar{d}[swap]{\pullbackpr{2}{\eext{2}}{\eft[3]}}
-  \ar{r}{\eext{3}}
-  & 
-\stesysf_3
-  \ar{d}{\eext{2}} 
-  \\
-\stesysf_3
-  \ar{r}[swap]{\eext{2}} 
-  &
-\stesysf_2
-\end{tikzcd}
-\end{equation*}
-Since this is a question about two maps into a pullback, it suffices to verify
-that
-\begin{align*}
-\pullbackpr{1}{\ectxext}{\eft}\circ\eext{2}\circ\eext{3}
-  & =
-\pullbackpr{1}{\ectxext}{\eft}\circ\eext{2}\circ\pullbackpr{2}{\eext{2}}{\eft[3]}
-  \\
-\pullbackpr{2}{\ectxext}{\eft}\circ\eext{2}\circ\eext{3}
-  & =
-\pullbackpr{2}{\ectxext}{\eft}\circ\eext{2}\circ\pullbackpr{2}{\eext{2}}{\eft[3]}.
-\end{align*}
-For the first equality, it is fairly straightforward to show that both the
-equalities
-\begin{equation*}
-\pullbackpr{1}{\ectxext}{\eft}\circ\eext{2}\circ\eext{3}
-  =
-\eft[2]\circ\eft[3]\circ\eft[4]
-\end{equation*}
-and
-\begin{equation*}
-\pullbackpr{1}{\ectxext}{\eft}\circ\eext{2}\circ\pullbackpr{2}{\eext{2}}{\eft[3]}
-  =
-\eft[2]\circ\eft[3]\circ\eft[4].
-\end{equation*}
-hold. For the second subgoal (which is more tricky). Notice first that the
-diagram
-\begin{equation*}
-\begin{tikzcd}
-\stesysf_4
-  \ar{r}{\eext{3}}
-  \ar{d}[swap]{\beta_3}
-  &
-\stesysf_3
-  \ar{r}{\eext{2}}
-  \ar{d}[swap]{\beta_2}
-  &
-\stesysf_2
-  \ar{d}{\pullbackpr{2}{\ectxext}{\eft}}
-  \\
-\stesysf_3
-  \ar{r}[swap]{\eext{2}}
-  &
-\stesysf_2
-  \ar{r}[swap]{\efamext}
-  &
-\stesysf
-\end{tikzcd}
-\end{equation*}
-commutes. We also have the commutative diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=large]
-\stesysf_4
-  \ar{r}{\pullbackpr{2}{\eext{2}}{\eft[3]}}
-  \ar{d}[swap]{\beta_3}
-  &
-\stesysf_3
-  \ar{r}{\eext{2}}
-  \ar{d}[swap]{\beta_2}
-  &
-\stesysf_2
-  \ar{d}{\pullbackpr{2}{\ectxext}{\eft}}
-  \\
-\stesysf_3
-  \ar{r}{\pullbackpr{2}{\efamext}{\eft[2]}}
-  \ar{dr}[swap]{\eext{2}}
-  &
-\stesysf_2
-  \ar{r}{\efamext}
-  &
-\stesysf
-  \\
-  {} &
-\stesysf_2
-  \ar{ur}[swap]{\efamext}
-\end{tikzcd}
-\end{equation*}
-completing the proof.
-\end{proof}
-
-\subsection{(Pre-)extension homomorphisms}\label{subsection:e_extension_homomorphisms}
-In this subsection we start with the study of pre-extension homomorphisms, which
-will include the extension homomorphisms since they will be the pre-extension
-homomorphisms of which both the domain and codomain are extension objects.
-Our main examples of extension homomorphisms will be the operations of weakening
-and substitution. There are some basic examples of pre-extension homomorphisms
-that will be useful too, which get introduced in the this section and in
-\autoref{subsection:change_of_base}. In this section, we will mainly be
-interested in pre-extension homomorphisms between local pre-extension objects.
-We will end this section by proving that a retract of an extension object is
-always an extension object.
-
-\begin{defn}
-Let $\stesys$ and $\stesys'$ be pre-extension algebras. A \emph{pre-extension 
-homomorphism $f$ from $\stesys'$ to $\stesys$} is a triple $(f_0,f_1,f^t)$ 
-consisting of morphisms
-\begin{equation*}
-\begin{tikzcd}
-\stesyst' 
-  \ar{r}{f^t}
-  \ar{d}[swap]{\ebd'}
-  &
-\stesyst
-  \ar{d}{\ebd}
-  \\
-\stesysf'
-  \ar{r}{f_1}
-  \ar{d}[swap]{\eft'}
-  &
-\stesysf
-  \ar{d}{\eft}
-  \\
-\stesysc' 
-  \ar{r}[swap]{f_0}
-  &
-\stesysc
-\end{tikzcd}
-\end{equation*}
-such that the indicated squares commute, for which furthermore the squares
-\begin{equation*}
-\begin{tikzcd}
-\stesysf' 
-  \ar{r}{f_1}
-  \ar{d}[swap]{\ectxext'}
-  &
-\stesysf
-  \ar{d}{\ectxext}
-  \\
-\stesysc'
-  \ar{r}[swap]{f_0}
-  &
-\stesysc
-\end{tikzcd}
-\end{equation*}
-and
-\begin{equation*}
-\begin{tikzcd}[column sep=huge]
-\stesysf'\times_{\ectxext',\eft'} \stesysf'
-  \ar{r}{f_1\times_{\ectxext,\eft} f_1}
-  \ar{d}[swap]{\efamext'}
-  &
-\stesysf\times_{\ectxext,\eft} \stesysf
-  \ar{d}{\efamext}
-  \\
-\stesysf'
-  \ar{r}[swap]{f_1}
-  &
-\stesysf
-\end{tikzcd}
-\end{equation*}
-Composition and the identity homomorphism are defined in the expected way. We
-define furthermore
-\begin{align*}
-f_2 & := \pullback{f_1}{f_1}{\ectxext}{\eft}
-  \\
-f_3 & := \pullback{f_2}{f_2}{\efamext}{\eft[2]}.
-\end{align*}
-\end{defn}
-
-\begin{defn}
-A pre-extension homomorphism between extension objects is called an extension
-homomorphism.
-\end{defn}
-
-\begin{lem}
-Let $\stesys$ be an extension object. Then
-\begin{equation*}
-\begin{tikzcd}[column sep=huge]
-\stesyst_2
-  \ar{r}{\pullbackpr{2}{\ectxext}{\eft\circ\ebd}}
-  \ar{d}[swap]{\ebd[2]}
-  &
-\stesyst
-  \ar{d}{\ebd}
-  \\
-\stesysf_2
-  \ar{r}{\pullbackpr{2}{\ectxext}{\eft}}
-  \ar{d}[swap]{\eft[2]}
-  &
-\stesysf
-  \ar{d}{\eft}
-  \\
-\stesysf
-  \ar{r}[swap]{\ectxext}
-  &
-\stesysc
-\end{tikzcd}
-\end{equation*}
-assembles an extension homomorphism $\mathbf{e}_0:\famesys{\stesys}\to\stesys$.
-Likewise, we have an extension homomorphism
-$\mathbf{e}_1:\famesys{\famesys{\stesys}}\to\famesys{\stesys}$. Thus, a
-pre-extension object is an extension object if and only if $\mathbf{e}_0$
-and $\mathbf{e}_1$ are pre-extension homomorphisms.
-\end{lem}
-
-\begin{proof}
-Immediate from the conditions of being an extension object.
-\end{proof}
-
-\begin{defn}
-Let $\stesys$ be a pre-extension object. Then
-\begin{equation*}
-\begin{tikzcd}
-\stesyst_3
-  \ar{r}{\beta^t}
-  \ar{d}[swap]{\ebd[3]}
-  &
-\stesyst_2
-  \ar{d}{\ebd[2]}
-  \\
-\stesysf_3
-  \ar{r}{\beta_2}
-  \ar{d}[swap]{\eft[3]}
-  &
-\stesysf_2
-  \ar{d}{\eft[2]}
-  \\
-\stesysf_2
-  \ar{r}[swap]{\beta_1}
-  &
-\stesysf
-\end{tikzcd}
-\qquad
-\text{and}
-\qquad
-\begin{tikzcd}
-\stesyst_4
-  \ar{r}{\beta^t_2}
-  \ar{d}[swap]{\ebd[4]}
-  &
-\stesyst_3
-  \ar{d}{\ebd[3]}
-  \\
-\stesysf_4
-  \ar{r}{\beta_3}
-  \ar{d}[swap]{\eft[4]}
-  &
-\stesysf_3
-  \ar{d}{\eft[3]}
-  \\
-\stesysf_3
-  \ar{r}[swap]{\beta_2}
-  &
-\stesysf_2
-\end{tikzcd}
-\end{equation*}
-assemble pre-extension homomorphisms 
-\(
-\boldsymbol{\beta}
-  :
-\famesys{\famesys{\stesys}}
-  \to
-\famesys{\stesys}
-\) 
-and
-\(
-\boldsymbol{\beta}_\mathbf{2}
-  :
-\famesys{\famesys{\famesys{\stesys}}}
-  \to
-\famesys{\famesys{\stesys}}
-\).
-\end{defn}
-
-\begin{defn}\label{famehom}
-Suppose that $f:\stesys'\to\stesys$ is a pre-extension homomorphism. Then we
-define $\famehom{f}:\famesys{\stesys'}\to\famesys{\stesys}$ to consist of
-\begin{equation*}
-\begin{tikzcd}
-\stesyst_2'
-  \ar{r}{f^t_2}
-  \ar{d}[swap]{\ebd[2]'}
-  &
-\stesyst_2
-  \ar{d}{\ebd[2]}
-  \\
-\stesysf_2'
-  \ar{r}{f_2}
-  \ar{d}[swap]{\eft[2]'}
-  &
-\stesysf_2
-  \ar{d}{\eft[2]}
-  \\
-\stesysf'
-  \ar{r}[swap]{f_1}
-  &
-\stesysf
-\end{tikzcd}
-\end{equation*}
-where we define
-\begin{equation*}
-f^t_2 := \pullback{f_1}{f^t}{\ectxext}{\eft\circ\ebd}.
-\end{equation*}
-\end{defn}
-
-\begin{lem}
-The triple $\famehom{f}$ defined in \autoref{famehom} is a pre-extension homomorphism.
-\end{lem}
-
-\begin{proof}
-Note that the square
-\begin{equation*}
-\begin{tikzcd}
-\stesysf_2'
-  \ar{r}{f_2}
-  \ar{d}[swap]{\efamext'}
-  &
-\stesysf_2
-  \ar{d}{\efamext}
-  \\
-\stesysf'
-  \ar{r}[swap]{f_1}
-  &
-\stesysf
-\end{tikzcd}
-\end{equation*}
-commutes by assumption. Thus, it remains to show that the square
-\begin{equation*}
-\begin{tikzcd}
-\stesysf_3'
-  \ar{r}{f_3}
-  \ar{d}[swap]{\eext{2}'}
-  &
-\stesysf_3
-  \ar{d}{\eext{2}}
-  \\
-\stesysf_2'
-  \ar{r}[swap]{f_2}
-  &
-\stesysf_2
-\end{tikzcd}
-\end{equation*}
-commutes. It is equivalent to show that the equalities
-\begin{align*}
-\pullbackpr{1}{\ectxext}{\eft}\circ f_2\circ\eext{2}'
-  & =
-\pullbackpr{1}{\ectxext}{\eft}\circ \eext{2}\circ f_3
-  \\
-\pullbackpr{2}{\ectxext}{\eft}\circ f_2\circ\eext{2}'
-  & =
-\pullbackpr{2}{\ectxext}{\eft}\circ \eext{2}\circ f_3
-\end{align*}
-both hold. For the first, it is straightforward to verify that the diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=large]
-{} &
-\stesysf_2'
-  \ar{r}{f_2}
-  \ar{dr}[near end]{\pullbackpr{1}{\ectxext'}{\eft'}}
-  &
-\stesysf_2
-  \ar{dr}{\pullbackpr{1}{\ectxext}{\eft}}
-  \\
-\stesysf_3'
-  \ar{ur}{\eext{2}'}
-  \ar{r}[swap]{\beta_2'}
-  \ar{ddr}[swap]{f_3}
-  &
-\stesysf_2'
-  \ar{r}{\efamext'}
-  \ar{dr}[swap]{f_2}
-  &
-\stesysf'
-  \ar{r}{f_1}
-  &
-\stesysf
-  \\
-{} & {} &
-\stesysf_2
-  \ar{ur}[near start]{\efamext}
-  \\
-{} &
-\stesysf_3
-  \ar{r}[swap]{\eext{2}}
-  \ar{ur}{\beta_2}
-  &
-\stesysf_2
-  \ar{uur}[swap]{\pullbackpr{1}{\ectxext}{\eft}}
-\end{tikzcd}
-\end{equation*}
-commutes. For the second, note that the diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=large]
-{} &
-\stesysf_2'
-  \ar{r}{f_2}
-  \ar{dr}{\beta_1'}
-  &
-\stesysf_2
-  \ar{ddr}{\pullbackpr{2}{\ectxext}{\eft}}
-  \\
-{} & {} &
-\stesysf'
-  \ar{dr}[swap,near start]{f_1}
-  \\
-\stesysf_3'
-  \ar{uur}{\eext{2}'}
-  \ar{r}{\beta_2'}
-  \ar{dr}[swap]{f_3}
-  &
-\stesysf_2'
-  \ar{r}{f_2}
-  \ar{ur}{\efamext'}
-  &
-\stesysf_2
-  \ar{r}[swap]{\efamext}
-  &
-\stesysf
-  \\
-{} &
-\stesysf_3
-  \ar{r}[swap]{\eext{2}}
-  \ar{ur}[near start]{\beta_2}
-  &
-\stesysf_2
-  \ar{ur}[swap]{\pullbackpr{2}{\ectxext}{\eft}}
-\end{tikzcd}
-\end{equation*}
-commutes.
-\end{proof}
-
-\begin{lem}[Stability under retracts]\label{esys-retract}
-Suppose $f:\stesys\to\stesys'$ is a pre-extension homomorphism between
-pre-extension objects. If there is a pre-extension homomorphism $g:\stesys'\to
-\stesys$ such that $g\circ f=\catid{\stesys}$ and $\stesys'$ is an extension
-algebra, then $\stesys$ is an extension algebra.
-\end{lem}
-
-Before we start with the proof, note that we have the equalities
-$g_2\circ f_2=\catid{\stesysf_2}$ and $g_3\circ f_3=\catid{\stesysf_3}$
-under the hypotheses of the lemma.
-
-\begin{proof}
-Our first subgoal is to show that the square
-\begin{equation*}
-\begin{tikzcd}
-\stesysf_2 
-  \ar{r}{\efamext} 
-  \ar{d}[swap]{\pullbackpr{2}{\ectxext}{\eft}} 
-  & 
-\stesysf 
-  \ar{d}{\ectxext}
-  \\
-\stesysf
-  \ar{r}[swap]{\ectxext} 
-  & 
-\stesysc
-\end{tikzcd}
-\end{equation*}
-commutes. Note that in the diagram
-\begin{equation*}
-\begin{tikzcd}
-  {}
-  & 
-\stesysf
-  \ar{dd}[near start]{\ectxext}
-  \ar{rr}{f_1}
-  &
-  &
-\stesysf'
-  \ar{dd}[near start]{\ectxext'}
-  \ar{rr}{g_1}
-  &
-  &
-\stesysf
-  \ar{dd}{\ectxext}
-  \\
-\stesysf_2
-  \ar{dd}[swap]{\pullbackpr{2}{\ectxext}{\eft}}
-  \ar[crossing over]{rr}[swap,near start]{f_2}
-  \ar{ur}{\efamext}
-  &
-  &
-\stesysf_2'
-  \ar{ur}[near start]{\efamext'}
-  \ar[crossing over]{rr}[swap,near start]{g_2}
-  &
-  &
-\stesysf_2
-  \ar{ur}[swap,near start]{\efamext}
-  \\
-  {}
-  &
-\stesysc
-  \ar{rr}[near start]{f_0}
-  &
-  &
-\stesysc'
-  \ar{rr}[near start]{g_0}
-  &
-  &
-\stesysc
-  \\
-\stesysf 
-  \ar{rr}[swap]{f_1}
-  \ar{ur}{\ectxext}
-  &
-  &
-\stesysf' 
-  \ar[crossing over,leftarrow]{uu}[near end,swap]{\pullbackpr{2}{\ectxext'}{\eft'}}
-  \ar{ur}[swap,near end]{\ectxext'}
-  \ar{rr}[swap]{g_1}
-  &
-  &
-\stesysf
-  \ar[crossing over,leftarrow]{uu}[near end,swap]{\pullbackpr{2}{\ectxext}{\eft}}
-  \ar{ur}[swap]{\ectxext}
-\end{tikzcd}
-\end{equation*}
-all the faces minus the far left and far right face commute. Using that $g$
-is a section of $f$, we can read off that also the far left face commutes,
-completing our first subgoal.
-
-For the second subgoal, note that also $\famehom{g}\circ\famehom{f}=
-\catid{\famesys{\stesys}}$ and that $\famesys{\stesys'}$ is an extension object.
-Thus we can apply what we have proven so far to conclude that the square
-\begin{equation*}
-\begin{tikzcd}
-\stesysf_3 
-  \ar{r}{\eext{2}} 
-  \ar{d}[swap]{\pullbackpr{2}{\efamext}{\eft[2]}} 
-  & 
-\stesysf_2 
-  \ar{d}{\efamext}
-  \\
-\stesysf_2
-  \ar{r}[swap]{\efamext} 
-  & 
-\stesysf
-\end{tikzcd}
-\end{equation*}
-commutes.
-\end{proof}
-
-\subsection{The change of base of (pre-)extension objects}
-\label{subsection:change_of_base}
-An important construction of (pre-)extension objects is the change of base. It
-allows us to consider `parametrized homomorphisms', such as weakening and
-substitution.
-
-\begin{defn}
-Suppose $f:\stesys\to\stesys'$ is a pre-extension homomorphism. We say that
-a diagram
-\begin{equation*}
-\begin{tikzcd}
-\stesys
-  \ar{r}{f}
-  \ar{d}[swap]{p}
-  &
-\stesys'
-  \ar{d}{p'}
-  \\
-X \ar{r}[swap]{g}
-  &
-Y
-\end{tikzcd}
-\end{equation*}
-commutes if the diagram
-\begin{equation*}
-\begin{tikzcd}
-\stesysc
-  \ar{r}{f_0}
-  \ar{d}[swap]{p}
-  &
-\stesysc'
-  \ar{d}{p'}
-  \\
-X \ar{r}[swap]{g}
-  &
-Y
-\end{tikzcd}
-\end{equation*}
-commutes.
-\end{defn}
-
-The first goal in this subsection is to define for every (pre-)extension object 
-$\stesys$ and every $p:\stesysc\rightarrow X\leftarrow Y:g$, a (pre-)extension
-object $\cobesys{Y}{\stesys}{g}{p}$ with a homomorphism $\pullbackpr{2}{g}{p}:
-\cobesys{Y}{\stesys}{g}{p}\to\stesys$ and a morphism $\pullbackpr{1}{g}{p}:
-\pullback{Y}{\stesysc}{g}{p}\to Y$ such that for every diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=large]
-\stesys'
-  \ar[bend right=10]{ddr}[swap]{p'}
-  \ar[bend left=10]{rrd}{f}
-  \ar[dotted]{dr}[near end]{[p',f]}
-  \\
-  {}&
-\cobesys{Y}{\stesys}{g}{p}
-  \ar{d}{\pullbackpr{1}{g}{p}}
-  \ar{r}[swap]{\pullbackpr{2}{g}{p}}
-  &
-\stesys
-  \ar{d}{p}
-  \\
-  {}&
-Y \ar{r}[swap]{g}
-  &
-X
-\end{tikzcd}
-\end{equation*}
-of which the outer square commutes, the (pre-)extension homomorphism $[p',f]$ exists
-and is unique with the property that it renders the diagram commutative. We will
-give the definition of $\cobesys{Y}{\stesys}{g}{p}$ in \autoref{cobesys}. After
-proving that the change of base of a pre-extension algebra is indeed a
-pre-extension algebra (\autoref{cobesys-preext}) and that the change of base
-of an extension algebra is an extension algebra (\autoref{cobesys-ext}), we
-will demonstrate the above unique existence in \autoref{cobesys-existence,%
-cobesys-pullback}.
-
-The second goal in this subsection is to follow the same procedure for
-$\famesys{\famesys{\stesys}}$ to show that it is equivalent to
-$\cobesys{\stesysf}{\famesys{\stesys}}{\ectxext}{\eft}$. We will do this by
-verifying directly that it has the universal property of the change of base
-described above, because we will use the ingredients in our definition of
-weakening and substitution objects.
-
-\begin{defn}[Change of base]\label{cobesys}
-Suppose $\stesys$ is a pre-extension object in $\cat{C}$ and that 
-$p:\stesysc\rightarrow X\leftarrow Y:g$.
-Then we define the pre-extension object $\cobesys{Y}{\stesys}{g}{p}$ to consist of
-\begin{equation*}
-\begin{tikzcd}
-\cobesys{Y}{\stesyst}{g}{p\circ\eft\circ\ebd}
-  \ar{r}
-  \ar{d}[swap]{g^\ast(\ebd)}
-  &
-\stesyst
-  \ar{d}{\ebd}
-  \\
-\cobesys{Y}{\stesysf}{g}{p\circ\eft}
-  \ar{r}
-  \ar{d}[swap]{g^\ast(\eft)}
-  &
-\stesysf
-  \ar{d}{\eft}
-  \\
-\cobesys{Y}{\stesysc}{g}{p}
-  \ar{r}
-  \ar{d}[swap]{\pullbackpr{1}{g}{p}}
-  &
-\stesysc
-  \ar{d}{p}
-  \\
-Y \ar{r}[swap]{g}
-  &
-X
-\end{tikzcd}
-\end{equation*} 
-and the operations
-\begin{align*}
-\cobesys{Y}{\ectxext}{g}{p} 
-  & : \pullback{Y}{\stesysf}{g}{p\circ\eft}\to \pullback{Y}{\stesysc}{g}{p}\\
-\cobesys{Y}{\efamext}{g}{p} 
-  & : \pullback
-    {\pullback{Y}{\stesysf}{g}{p\circ\eft}}
-    {\pullback{Y}{\stesysf}{g}{p\circ\eft}}
-    {\cobesys{Y}{\ectxext}{g}{p}}
-    {g^\ast(\eft)}
-  \to 
-  \pullback{Y}{\stesysf}{g}{p\circ\eft}.
-\end{align*}
-defined by
-\begin{equation*}
-\cobesys{Y}{\ectxext}{g}{p} := \pullback{\catid{Y}}{\ectxext}{g}{p}
-\end{equation*}
-and where $\cobesys{Y}{\efamext}{g}{p}$ is defined by rendering the diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=large]
-(\cobesys{Y}{\stesysf}{g}{p\circ\eft})_2
-  \ar{rr}{\pullback{\pullbackpr{2}{g}{p\circ\eft}}{\pullbackpr{2}{g}{p\circ\eft}}{\ectxext}{\eft}}
-  \ar{dd}[swap]{\pullbackpr{1}{\cobesys{Y}{\ectxext}{g}{p}}{g^\ast(\eft)}}
-  \ar[dotted]{dr}[swap]{\cobesys{Y}{\efamext}{g}{p}}
-  &
-  &
-\stesysf_2
-  \ar{d}{\efamext}
-  \\
-  {}&
-\cobesys{Y}{\stesysf}{g}{p\circ\eft}
-  \ar{r}{\pullbackpr{2}{g}{p\circ\eft}}
-  \ar{d}[swap]{\pullbackpr{1}{g}{p\circ\eft}}
-  &
-\stesysf
-  \ar{d}{p\circ\eft}
-  \\
-\cobesys{Y}{\stesysf}{g}{p\circ\eft}
-  \ar{r}[swap]{\pullbackpr{1}{g}{p\circ\eft}}
-  &
-Y \ar{r}[swap]{g}
-  &
-X
-\end{tikzcd}
-\end{equation*} 
-commutative. 
-The process of obtaining the pre-extension object $\cobesys{Y}{\stesys}{g}{p}$ out of $\stesys$
-and $g:Y\to X$ is also called the \emph{change of base}.
-\end{defn}
-
-\begin{lem}\label{cobesys-preext}
-Any change of base of a pre-extension object is a pre-extension object.
-\end{lem}
-
-\begin{proof}
-Let $\stesys$ be an extension algebra and consider $p:\stesysc\rightarrow X\leftarrow Y:g$.
-We need to verify that the square
-\begin{equation*}
-\begin{tikzcd}[column sep=large]
-(\pullback{Y}{\stesysf}{g}{p\circ\eft})_2
-  \ar{r}{\cobesys{Y}{\efamext}{g}{p}} 
-  \ar{d}[swap]{\pullbackpr{1}{\cobesys{Y}{\ectxext}{g}{p}}{g^\ast(\eft)}} 
-  & 
-\pullback{Y}{\stesysf}{g}{p\circ\eft}
-  \ar{d}{g^\ast(\eft)}
-  \\
-\pullback{Y}{\stesysf}{g}{p\circ\eft}
-  \ar{r}[swap]{g^\ast(\eft)} 
-  & 
-\pullback{Y}{\stesysc}{g}{p}
-\end{tikzcd}
-\end{equation*}
-commutes. It is fairly obvious that
-\begin{equation*}
-\pullbackpr{1}{g}{p}\circ g^\ast(\eft)\circ (\cobesys{Y}{\efamext}{g}{p})
-  =
-\pullbackpr{1}{g}{p\circ\eft}\circ \pullbackpr{1}{\cobesys{Y}{\ectxext}{g}{p}}{g^\ast(\eft)}
-\end{equation*}
-and that the diagram
-\begin{equation*}
-\begin{tikzcd}
-  {}&
-  {}&
-\pullback{Y}{\stesysf}{g}{p\circ\eft}
-  \ar{rr}{g^\ast(\eft)}
-  \ar{dr}[swap]{\pullbackpr{2}{g}{p\circ\eft}}
-  &
-  {}&
-\pullback{Y}{\stesysc}{g}{p}
-  \ar{ddr}{\pullbackpr{2}{g}{p}}
-  \\
-  {}&
-  {}&
-  {}&
-\stesysf
-  \ar{drr}[swap]{\eft}
-  \\
-(\pullback{y}{\stesysf}{g}{p\circ\eft})_2
-  \ar{uurr}{\cobesys{Y}{\efamext}{g}{p}}
-  \ar{rr}[swap,yshift=-.5ex]{\pullback{\pullbackpr{2}{g}{p\circ\eft}}{\pullbackpr{2}{g}{p\circ\eft}}{\ectxext}{\eft}}
-  \ar{ddrr}[swap]{\pullbackpr{1}{\cobesys{Y}{\ectxext}{g}{p}}{g^\ast(\eft)}}
-  &
-  {}&
-\stesysf_2
-  \ar{ur}{\efamext}
-  \ar{dr}[swap]{\eft[2]}
-  &
-  {}&
-  {}&
-\stesysc
-  \\
-  {}&
-  {}&
-  {}&
-\stesysf
-  \ar{urr}{\eft}
-  \\
-  {}&
-  {}&
-\pullback{Y}{\stesysf}{g}{p\circ\eft}
-  \ar{rr}[swap]{g^\ast(\eft)}
-  \ar{ur}{\pullbackpr{2}{g}{p\circ\eft}}
-  &
-  {}&
-\pullback{Y}{\stesysc}{g}{p}
-  \ar{uur}[swap]{\pullbackpr{2}{g}{p}}
-\end{tikzcd}
-\end{equation*}
-commutes.
-\end{proof}
-
-\begin{thm}\label{cobesys-ext}
-The change of base of an extension algebra is an extension algebra.
-\end{thm}
-
-\begin{proof}
-Our first subgoal is to verify that the square
-\begin{equation*}
-\begin{tikzcd}[column sep=large]
-(\pullback{Y}{\stesysf}{g}{p\circ\eft})_2
-  \ar{r}{\cobesys{Y}{\efamext}{g}{p}} 
-  \ar{d}[swap]{\pullbackpr{2}{\cobesys{Y}{\ectxext}{g}{p}}{g^\ast(\eft)}} 
-  & 
-\pullback{Y}{\stesysf}{g}{p\circ\eft}
-  \ar{d}{\cobesys{Y}{\ectxext}{g}{p}}
-  \\
-\pullback{Y}{\stesysf}{g}{p\circ\eft}
-  \ar{r}[swap]{\cobesys{Y}{\ectxext}{g}{p}} 
-  & 
-\pullback{Y}{\stesysc}{g}{p}
-\end{tikzcd}
-\end{equation*}
-\end{proof}
-
-The following construction is useful for defining extension homomorphisms into
-`higher' extension objects
-
-\begin{defn}\label{cobesys-existence}
-Consider a commutative diagram
-\begin{equation*}
-\begin{tikzcd}
-\stesys'
-  \ar{r}{f}
-  \ar{d}[swap]{p'}
-  &
-\stesys
-  \ar{d}{p}
-  \\
-Y \ar{r}[swap]{g}
-  &
-X
-\end{tikzcd}
-\end{equation*}
-Then we construct $[p,f]:\stesys'\to\cobesys{Y}{\stesys}{g}{p}$
-\begin{itemize}
-\item by defining $[p,f]_0:\stesysc'\to\pullback{Y}{\stesysc}{g}{p}$ be the uniqe
-morphism rendering the diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=large]
-\stesysc'
-  \ar[bend right=10]{ddr}[swap]{p'}
-  \ar[bend left=10]{rrd}{f_0}
-  \ar{dr}[near end]{[p,f]_0}
-  \\
-  {}&
-\pullback{Y}{\stesysc}{g}{p}
-  \ar{r}[swap]{\pullbackpr{2}{g}{p}}
-  \ar{d}{\pullbackpr{1}{g}{p}}
-  &
-\stesysc
-  \ar{d}{p}
-  \\
-  {}&
-Y \ar{r}[swap]{g}
-  &
-X
-\end{tikzcd}
-\end{equation*}
-commutative.
-\item by defining $[p,f]_1:\stesysf'\to\pullback{Y}{\stesysf}{g}{p\circ\eft}$ be the uniqe
-morphism rendering the diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=huge]
-\stesysf'
-  \ar[bend right=10]{ddr}[swap]{p'\circ\eft'}
-  \ar[bend left=10]{rrd}{f_1}
-  \ar{dr}[near end]{[p,f]_1}
-  \\
-  {}&
-\pullback{Y}{\stesysf}{g}{p\circ\eft}
-  \ar{r}[swap]{\pullbackpr{2}{g}{p\circ\eft}}
-  \ar{d}{\pullbackpr{1}{g}{p\circ\eft}}
-  &
-\stesysc
-  \ar{d}{p\circ\eft}
-  \\
-  {}&
-Y \ar{r}[swap]{g}
-  &
-X
-\end{tikzcd}
-\end{equation*}
-commutative.
-\item by defining $[p,f]^t:\stesyst'\to\pullback{Y}{\stesyst}{g}{p\circ\eft\circ\ebd}$ be the uniqe
-morphism rendering the diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=huge]
-\stesyst'
-  \ar[bend right=10]{ddr}[swap]{p'\circ\eft'\circ\ebd'}
-  \ar[bend left=10]{rrd}{f^t}
-  \ar{dr}[near end]{[p,f]^t}
-  \\
-  {}&
-\pullback{Y}{\stesysf}{g}{p\circ\eft\circ\ebd}
-  \ar{r}[swap]{\pullbackpr{2}{g}{p\circ\eft\circ\ebd}}
-  \ar{d}{\pullbackpr{1}{g}{p\circ\eft\circ\ebd}}
-  &
-\stesysc
-  \ar{d}{p\circ\eft\circ\ebd}
-  \\
-  {}&
-Y \ar{r}[swap]{g}
-  &
-X
-\end{tikzcd}
-\end{equation*}
-commutative.
-\end{itemize}
-\end{defn}
-
-\begin{thm}\label{cobesys-pullback}
-For every diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=large]
-\stesys'
-  \ar[bend right=10]{ddr}[swap]{p'}
-  \ar[bend left=10]{rrd}{f}
-  \ar[dotted]{dr}[near end]{[p',f]}
-  \\
-  {}&
-\cobesys{Y}{\stesys}{g}{p}
-  \ar{d}{\pullbackpr{1}{g}{p}}
-  \ar{r}[swap]{\pullbackpr{2}{g}{p}}
-  &
-\stesys
-  \ar{d}{p}
-  \\
-  {}&
-Y \ar{r}[swap]{g}
-  &
-X
-\end{tikzcd}
-\end{equation*}
-of which the outer square commutes, the pre-extension homomorphism $[p',f]$
-is unique with the property that it renders the whole diagram commutative.
-\end{thm}
-
-\begin{defn}\label{famfamstesys_into}
-Consider a commutative square
-\begin{equation*}
-\begin{tikzcd}
-\stesys'
-  \ar{r}{f}
-  \ar{d}[swap]{p}
-  &
-\famesys\stesys
-  \ar{d}{\eft}
-  \\
-\stesysf \ar{r}[swap]{\ectxext}
-  &
-\stesysc
-\end{tikzcd}
-\end{equation*}
-Then we construct
-\begin{equation*}
-[p,f]:\stesys'\to\famesys{\famesys{\stesys}}
-\end{equation*}
-as follows:
-\begin{itemize}
-\item let $[p,f]_0:\stesysc'\to\stesysf_2$ be the unique morphism rendering
-the diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=large]
-\stesysc' 
-  \ar[bend left=10]{rrd}{f_0}
-  \ar[swap,bend right=10]{ddr}{p}
-  \ar[dotted]{dr}[near end]{[p,f]_0}
-  \\
-  {}&
-\stesysf_2
-  \ar{r}[swap]{\pullbackpr{2}{\ectxext}{\eft}}
-  \ar{d}{\eft[2]}
-  &
-\stesysf
-  \ar{d}{\eft}
-  \\
-  {}&
-\stesysf
-  \ar{r}[swap]{\ectxext}
-  &
-\stesysc
-\end{tikzcd}
-\end{equation*}
-commutative.
-\item Let $[p,f]_1:\stesysf'\to\stesysf_3$ be the unique morphism rendering
-the diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=large]
-\stesysf'
-  \ar[bend left=10]{drr}{f_1}
-  \ar[swap]{dd}{\eft'}
-  \ar[dotted]{dr}[near end]{[p,f]_1}
-  \\
-  {}&
-\stesysf_3
-  \ar{r}[swap]{\pullbackpr{2}{\efamext}{\eft[2]}}
-  \ar{d}{\eft[2]}
-  &
-\stesysf_2
-  \ar{d}{\eft[2]}
-  \\
-\stesysc'
-  \ar{r}[swap]{[p,f]_0}
-  &
-\stesysf_2
-  \ar{r}[swap]{\efamext}
-  &
-\stesysf
-\end{tikzcd}
-\end{equation*}
-commutative.
-\item Let $[p,f]^t:\stesyst'\to\stesyst_3$ be the unique morphism rendering
-the diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=huge]
-\stesyst'
-  \ar[bend left=10]{drr}{f^t}
-  \ar[swap]{dd}{\eft'\circ\ebd'}
-  \ar[dotted]{dr}[near end]{[p,f]^t}
-  \\
-  {}&
-\stesyst_3
-  \ar{r}[swap]{\pullbackpr{2}{\efamext}{\eft[2]\circ\ebd[2]}}
-  \ar{d}[swap]{\pullbackpr{1}{\efamext}{\eft[2]\circ\ebd[2]}}
-  &
-\stesyst_2
-  \ar{d}{\eft[2]\circ\ebd[2]}
-  \\
-\stesysc'
-  \ar{r}[swap]{[p,f]_0}
-  &
-\stesysf_2
-  \ar{r}[swap]{\efamext}
-  &
-\stesysf
-\end{tikzcd}
-\end{equation*}
-commutative.
-\end{itemize}
-\end{defn}
-
-\begin{lem}
-Under the hypotheses of \autoref{famfamstesys_into}, $[p,f]$ is a pre-extension
-homomorphism. Moreover, it is the unique pre-extension homomorphism for which
-the diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=large]
-\stesys' 
-  \ar[bend left=10]{rrd}{f}
-  \ar[swap,bend right=10]{ddr}{p}
-  \ar[dotted]{dr}[near end]{[p,f]}
-  \\
-  {}&
-\famesys{\famesys{\stesys}}
-  \ar{r}[swap]{\pullbackpr{2}{\ectxext}{\eft}}
-  \ar{d}{\eft[2]}
-  &
-\famesys{\stesys}
-  \ar{d}{\eft}
-  \\
-  {}&
-\stesysf
-  \ar{r}[swap]{\ectxext}
-  &
-\stesysc
-\end{tikzcd}
-\end{equation*}
-commutes.
-\end{lem}
-
-\begin{lem}
-Suppose $f:\stesys\to \stesys'$ is a pre-extension homomorphism and consider a morphism
-$p:\stesys'\to X$ and $g:Y\to X$. Then the change of base 
-$g^\ast(f):\cobesys{Y}{\stesys}{g}{p\circ f_0}\to
-\cobesys{Y}{\stesys'}{g}{p}$ is a pre-extension morphism.
-\end{lem}
-
-\begin{lem}
-Let $\stesys$ be a pre-extension algebra and consider $p:\stesysc\rightarrow X\leftarrow Y:g$.
-Then there is an isomorphism
-\begin{equation*}
-\varphi:\famesys{\cobesys{Y}{\stesys}{g}{p}}
-  \simeq
-\cobesys{Y}{\famesys{\stesys}}{g}{p\circ\eft}
-\end{equation*}
-uniquely determined by
-\end{lem}
-
-\begin{proof}
-This follows from the pasting lemma for pullbacks.
-\end{proof}
-
-\subsection{Pre-weakening objects}
-\begin{defn}
-Let $\stesys$ be an extension object in $\cat{C}$. A pre-weakening operation
-on $\stesys$ is an extension homomorphism 
-$ \mathbf{w}(\stesys)
-    :
-  \cobesys{\stesysf}{\famesys{\stesys}}{\eft}{\eft}
-    \to
-  \famesys{\famesys{\stesys}}$
-for which the diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=large]
-\cobesys{\stesysf}{\famesys{\stesys}}{\eft}{\eft}
-  \ar{r}{\mathbf{w}(\stesys)}
-  \ar{dr}[swap]{\pullbackpr{1}{\eft}{\eft}}
-  &
-\famesys{\famesys{\stesys}}
-  \ar{d}{\eft[2]}
-  \\
-& \stesysf
-\end{tikzcd}
-\end{equation*}
-commutes.
-\end{defn}
-
-\begin{defn}
-Let $\stesys$ be an extension object with pre-weakening operation
-$\mathbf{w}(\stesys)$. Then $\famesys{\stesys}$ has the pre-weakening operation
-$\mathbf{w}(\famesys{\stesys})$ which is uniquely determined by rendering the
-diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=large]
-\cobesys{\stesysf_2}{\famesys{\famesys{\stesys}}}{\eft[2]}{\eft[2]}
-  \ar{rr}{%
-      \pullback{\beta_1}{\boldsymbol{\beta}}{\eft}{\eft}
-    }
-  \ar[bend right]{ddr}[swap]{\pullbackpr{1}{\eft[2]}{\eft[2]}}
-  \ar[dotted]{dr}{\mathbf{w}(\famesys{\stesys})}
-  &
-  {}&
-\cobesys{\stesysf}{\famesys{\stesys}}{\eft}{\eft}
-  \ar{r}{\mathbf{w}(\stesys)}
-  &
-\famesys{\famesys{\stesys}}
-  \ar{d}{\boldsymbol{\beta}}
-  \\
-  {}&
-\famesys{\famesys{\famesys{\stesys}}}
-  \ar{r}{\boldsymbol{\beta}_\mathbf{2}}
-  \ar{d}[swap]{\eft[3]}
-  &
-\famesys{\famesys{\stesys}}
-  \ar{d}{\eft[2]}
-  \ar{r}{\boldsymbol{\beta}}
-  &
-\famesys{\stesys}
-  \ar{d}{\eft}
-  \\
-  {}&
-\stesysf_2
-  \ar{r}[swap]{\efamext}
-  &
-\stesysf
-  \ar{r}[swap]{\ectxext}
-  &
-\stesysc
-\end{tikzcd}
-\end{equation*}
-commutative.
-\end{defn}
-
-\begin{defn}
-A pre-weakening object $\stesys$ in $\cat{C}$ is an extension object $\stesys$ 
-with a pre-weakening operation 
-$ \mathbf{w}(\stesys)
-    :
-  \cobesys{\stesysf}{\famesys{\stesys}}{\eft}{\eft}
-    \to
-  \famesys{\famesys{\stesys}}$
-for which the diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=15em]
-\cobesys{\stesysf_2}{\famesys{\stesys}}{\eft\circ\eft[2]}{\eft}
-  \ar[bend right=10]{dr}[swap]%
-    { [ \pullbackpr{1}{\eft\circ\eft[2]}{\eft},%
-        \mathbf{w}(\stesys)%
-          \circ%
-        (\pullback{\efamext}{\catid{\famesys{\stesys}}}{\eft}{\eft})%
-        ]%
-      }
-  \ar{r}{
-    [ \pullbackpr{1}{\eft\circ\eft[2]}{\eft},%
-      \mathbf{w}(\stesys)%
-        \circ%
-      (\pullback{\eft[2]}{\catid{\famesys{\stesys}}}{\eft}{\eft})%
-      ]}%
-  &
-\cobesys{\stesysf_2}{\famesys{\famesys{\stesys}}}{\eft[2]}{\eft[2]}
-  \ar{d}{\mathbf{w}(\famesys{\stesys})}
-  \\
-  {}&
-\famesys{\famesys{\famesys{\stesys}}}
-\end{tikzcd}
-\end{equation*}
-commutes. This condition is called \emph{Currying for weakening}.
-\end{defn}
-
-\begin{lem}
-If $\stesys$ is a pre-weakening algebra, then so is $\famesys{\stesys}$. 
-\end{lem}
-
-\begin{proof}
-\end{proof}
-
-\begin{defn}
-A pre-weakening morphism between preweakening objects $\stesys$ and $\stesys'$ is an
-extension homomorphism $f:\stesys\to \stesys'$ such that additionally the diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=large]
-\cobesys{\stesysf}{\famesys{\stesys}}{\eft}{\eft}
-  \ar{d}[swap]{\mathbf{w}(\stesys)}
-  \ar{r}{\pullback{f_1}{\famehom{f}}{\eft'}{\eft'}}
-  &
-\cobesys{\stesysf'}{\famesys{\stesys'}}{\eft'}{\eft'}
-  \ar{d}{\mathbf{w}(\stesys')}
-  \\
-\famesys{\famesys{\stesys}}
-  \ar{r}[swap]{\famehom{\famehom{f}}}
-  &
-\famesys{\famesys{\stesys'}}
-\end{tikzcd}
-\end{equation*}
-commutes.
-\end{defn}
-
-\begin{defn}
-Let $\stesys$ be a pre-weakening algebra and consider $p:\stesysc\rightarrow X\leftarrow Y:p$.
-Then we define
-\begin{equation*}
-\mathbf{w}(\cobesys{Y}{\stesys}{g}{p})
-  :
-\cobesys
-  { (\pullback{Y}{\stesysf}{g}{p\circ\eft})}
-  { \famesys{\cobesys{Y}{\stesys}{g}{p}}}
-  { g^\ast(\eft)}
-  { g^\ast(\eft)}
-  \to
-\famesys{\famesys{\cobesys{Y}{\stesys}{g}{p}}}
-\end{equation*}
-to be the unique extension homomorphism rendering the diagram
-\begin{equation*}
-\begin{tikzcd}
-\cobesys
-  { \pullback{Y}{\stesysf}{g}{p\circ\eft}}
-  { \famesys{\cobesys{Y}{\stesys}{g}{p}}}
-  { g^\ast(\eft)}
-  { g^\ast(\eft)}
-  \ar[bend right]{ddr}[swap]{\pullbackpr{1}{g^\ast(\eft)}{g^\ast(\eft)}}
-  \ar{rr}{%
-    \pullback
-      { \pullbackpr{2}{g}{p\circ\eft}}
-      { \boldsymbol{\pi}_\mathbf{2}(g,p\circ\eft)}
-      { \eft}
-      { \eft}
-    }
-  \ar[dotted]{dr}{\mathbf{w}(\cobesys{Y}{\stesys}{g}{p})}
-  &
-  {}&
-\cobesys{\stesysf}{\famesys{\stesys}}{\eft}{\eft}
-  \ar{r}{\mathbf{w}(\stesys)}
-  &
-\famesys{\famesys{\stesys}}
-  \ar{d}{\boldsymbol{\beta}}
-  \\
-  {}&
-\famesys{\famesys{\cobesys{Y}{\stesys}{g}{p}}}
-  \ar{r}{\beta}
-  \ar{d}{g^\ast(\eft)_1}
-  &
-\famesys{\cobesys{Y}{\stesys}{g}{p}}
-  \ar{r}{\boldsymbol{\pi}_2(g,p\circ\eft)}
-  \ar{d}{g^\ast(\eft)}
-  &
-\famesys{\stesys}
-  \ar{d}{\eft}
-  \\
-  {}&
-\pullback{Y}{\stesysf}{g}{p\circ\eft}
-  \ar{r}[swap]{\cobesys{Y}{\ectxext}{g}{p}}
-  &
-\pullback{Y}{\stesysc}{g}{p}
-  \ar{r}[swap]{\pullbackpr{2}{g}{p}}
-  &
-\stesysc
-\end{tikzcd}
-\end{equation*}
-commutative.
-\end{defn}
-
-\subsection{Weakening objects}
-Since we have shown that the property of being a pre-weakening object is closed
-under the relevant operations, we can make the following definition:
-
-\begin{defn}
-A weakening object is a pre-weakening object $\stesys$ with the property that
-$\mathbf{w}(\stesys)$ is a pre-weakening morphism.
-\end{defn}
-
-\begin{defn}
-A weakening homomorphism is a pre-weakening homomorphism such that the domain
-and codomain are weakening objects.
-\end{defn}
-
-\begin{thm}
-Suppose $\stesys$ is a weakening object, then so is $\famesys{\stesys}$
-\end{thm}
-
-\begin{thm}
-The change of base of any weakening object is again a weakening object.
-\end{thm}
-
-\subsection{Projection objects}
-\begin{defn}
-A pre-projection object is a weakening object $\stesys$ for which there is a term
+A pre-projection algebra is a weakening algebra $\stesys$ for which there is a term
 $\mathbf{i}:\stesysf\to \stesyst_2$ such that the diagram
@@
 \end{equation*}
-commutes
+commutes.
 \end{defn}
@@
 \begin{lem}
-The change of base of a pre-projection object is again a pre-projection object.
+The change of base of a pre-projection algebra is again a pre-projection algebra.
 \end{lem}
@@
 \begin{lem}
-If $CFT$ is a pre-projection object, then so is $\mathbf{F}_{CFT}$, where
+If $CFT$ is a pre-projection algebra, then so is $\mathbf{F}_{CFT}$, where
 $\mathbf{F}_{\mathbf{i}}$ is defined to be $F\times_{e_0,c}\mathbf{i}$ is
@@
 \begin{cor}
-The change of base of a projection object is again a projection object.
+The change of base of a projection algebra is again a projection algebra.
 \end{cor}
@@
 \begin{cor}
-If $CFT$ is a projection object, then so is $\mathbf{F}_{CFT}$, where
+If $CFT$ is a projection algebra, then so is $\mathbf{F}_{CFT}$, where
 $\mathbf{F}_{\mathbf{i}}$ is defined to be $F\times_{e_0,c}\mathbf{i}$ is
@@
 \end{cor}
-
-\subsection{Substitution objects}
-
-\begin{defn}
-A \emph{pre-substitution} for an extension object $\stesys$ is an
-extension homomorphism
-\begin{equation*}
-\mathbf{s}(\stesys):\cobesys{\stesyst}{\famesys{\famesys{\stesys}}}{\ebd}{\eft[2]}\to \famesys{\stesys}
-\end{equation*}
-for which the square
-\begin{equation*}
-\begin{tikzcd}[column sep=large]
-\pullback{\stesyst}{\stesysf_2}{\ebd}{\eft[2]}
-  \ar{r}{s(\stesys)_0}
-  \ar{d}[swap]{\ebd\circ\pullbackpr{1}{\ebd}{\eft[2]}}
-  &
-\stesysf 
-  \ar{d}{\eft}
-  \\
-\stesysf 
-  \ar{r}[swap]{\eft}
-  &
-\stesysc
-\end{tikzcd}
-\end{equation*}
-commutes. A \emph{pre-substitution object} is an extension object
-together with a pre-substitution.
-\end{defn}
-
-\begin{defn}
-A \emph{pre-substitution homomorphism} is an extension homomorphism $f:\stesys'\to \stesys$
-for which the square
-\begin{equation*}
-\begin{tikzcd}[column sep=huge]
-\cobesys{\stesyst'}{\famesys{\famesys{\stesys'}}}{\ebd'}{\eft[2]'}
-  \ar{r}{\pullback{f^t}{\famehom{\famehom{f}}}{\ebd}{\eft[2]}}
-  \ar{d}[swap]{\mathbf{s}'(\stesys')}
-  &
-\cobesys{\stesyst}{\famesys{\famesys{\stesys}}}{\ebd}{\eft[2]}
-  \ar{d}{\mathbf{s}(\stesys)}
-  \\
-\famesys{\stesys'}
-  \ar{r}[swap]{\famehom{f}}
-  &
-\famesys{\stesys}
-\end{tikzcd}
-\end{equation*}
-commutes.
-\end{defn}
-
-\begin{lem}
-The change of base of a pre-substitution object is again a pre-substitution object.
-\end{lem}
-
-\begin{lem}
-If $\stesys$ is a pre-substitution object, then so is $\famesys{\stesys}$ with
-$\mathbf{s}(\famesys{\stesys})$ defined to be the unique extension homomorphism
-rendering the diagram
-\begin{equation*}
-\begin{tikzcd}
-\cobesys{\stesyst_2}{\famesys{\famesys{\famesys{\stesys}}}}{\ebd[2]}{\eft[3]}
-  \ar[dotted]{dr}{\mathbf{s}(\famesys{\stesys})}
-  \ar{rr}{\pullback{\pullbackpr{2}{\ectxext}{\eft\circ\ebd}}{\boldsymbol{\beta}_\mathbf{2}}{\ebd}{\eft[2]}}
-  \ar{dd}[swap]{\ebd[2]\circ\pullbackpr{1}{\ebd[2]}{\eft[3]}}
-  &
-  {}&
-\cobesys{\stesyst}{\famesys{\famesys{\stesys}}}{\ebd}{\eft[2]}
-  \ar{d}{\mathbf{s}(\stesys)}
-  \\
-  {}&
-\famesys{\famesys{\stesys}}
-  \ar{r}{\boldsymbol{\beta}}
-  \ar{d}{\eft[2]}
-  &
-\famesys{\stesys}
-  \ar{d}{\eft}
-  \\
-\stesysf_2
-  \ar{r}[swap]{\eft[2]}
-  &
-\stesysf
-  \ar{r}[swap]{\ectxext}
-  &
-\stesysc
-\end{tikzcd}
-\end{equation*}
-commutative.
-\end{lem}
-
-\begin{proof}
-The requirement on pre-substitutions holds by construction.
-\end{proof}
-
-It makes sense now to consider the possibility that the pre-substitution
-itself is a pre-substitution homomorphism.
-
-\begin{defn}
-A \emph{substitution object} is a pre-substitution object for which substitution is
-a pre-substitution homomorphism.
-\end{defn}
-
-\begin{cor}
-The change of base of a substitution object is again a substitution object.
-\end{cor}
-
-\begin{cor}
-If $\stesys$ is a substitution object, then so is $\famesys{\stesys}$.
-\end{cor}
-
-\subsection{Extension objects with empty context and families}
-
-\begin{defn}
-An extension object $\stesys$ is said to have \emph{empty families} if there
-is a section
-\begin{equation*}
-\phi_1(\stesys):\stesysc\to\stesysf
-\end{equation*}
-of $\eft$, satisfying the following additional properties:
-\begin{enumerate}
-\item $\phi_1(\stesys)$ is also a section of $\ectxext$.
-\item The unique morphism $\phi_2$ rendering the diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=huge]
-\stesysf
-  \ar[bend left=20]{drr}{\phi_1(\stesys)\circ\ectxext}
-  \ar[equals,bend right=20]{ddr}
-  \ar[dotted]{dr}{\phi_2}
-  \\
-  {}&
-\stesysf_2
-  \ar{d}{\eft[2]}
-  \ar{r}{\pullbackpr{2}{\ectxext}{\eft}}
-  &
-\stesysf
-  \ar{d}{\eft}
-  \\
-  {}&
-\stesysf
-  \ar{r}[swap]{\ectxext}
-  &
-\stesysc
-\end{tikzcd}
-\end{equation*}
-commutative, is a section of $\efamext$.
-\item The unique morphism $\iota_1$ rendering the diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=huge]
-\stesysf
-  \ar[equals,bend left=20]{drr}
-  \ar[bend right=20]{ddr}[swap]{\phi_1(\stesys)\circ\eft}
-  \ar[dotted]{dr}{\iota_1}
-  \\
-  {}&
-\stesysf_2
-  \ar{d}{\eft[2]}
-  \ar{r}[swap]{\pullbackpr{2}{\ectxext}{\eft}}
-  &
-\stesysf
-  \ar{d}{\eft}
-  \\
-  {}&
-\stesysf
-  \ar{r}[swap]{\ectxext}
-  &
-\stesysc
-\end{tikzcd}
-\end{equation*}
-commutative, is a section of $\efamext$.
-\end{enumerate}
-\end{defn}
-
-\begin{defn}
-A homomorphism of extension objects with empty families is an extension
-homomorphism $f:\stesys'\to\stesys$ for which the diagram
-\begin{equation*}
-\begin{tikzcd}
-\stesysf'
-  \ar{r}{f_1}
-  &
-\stesysf
-  \\
-\stesysc'
-  \ar{u}{\phi_1(\stesys')}
-  \ar{r}[swap]{f_0}
-  &
-\stesysc
-  \ar{u}[swap]{\phi_1(\stesys)}
-\end{tikzcd}
-\end{equation*}
-commutes.
-\end{defn}
-
-\begin{lem}
-Suppose $\stesys$ is an extension object with empty families. Then
-$\famesys{\stesys}$ is an extension object with empty families, with
-$\phi_1(\famesys{\stesys}):=\phi_2$. 
-\end{lem}
-
-\begin{lem}
-Suppose $\stesys$ is an extension object with empty families and consider
-$p:\stesysc\rightarrow X\leftarrow Y:p$. Then
-$\cobesys{Y}{\stesys}{g}{p}$ is an extension object with empty families
-with $\phi_1(\cobesys{Y}{\stesys}{g}{p}):=g^\ast(\phi_1)$.
-\end{lem}
-
-\subsection{E-objects}
-\begin{defn}
-An \emph{E-object} is an extension object with the structure of a projection object,
-the structure of a substitution object and which has an empty context and families,
-such that additionally:
-\begin{enumerate}
-\item substitution is a projection homomorphism
-\item weakening is a substitution homomorphism
-\item both weakening and substitution are empty-CF homomorphisms.
-\item 
-\end{enumerate}
-\end{defn}

### 2015-02-10 Algebras/substitution.tex
large reorganisation

--- 2014-11-03 Algebras/article-models-cat.tex
@@
-\section{E-objects in categories with finite limits}
-In this section we assume that $\cat{C}$ is a finitely complete category and
-whenever we write a pullback, we assume that it is chosen. Recall that for
-any morphism $f:A\to B$ in a category $\cat{C}$ with chosen pullbacks, there
-is a functor
-\begin{equation*}
-f^\ast : \cat{C}/B\to\cat{C}/A.
-\end{equation*}
-As usual, when $g:X\to B$ is a morphism, we will write $f^\ast(X)$ for the
-domain of $f^\ast(g)$. When there is more than one morphism $X\to B$ involved,
-as will be the case below, we will write $\pullback{A}{X}{f}{g}$. The projections
-will be written as $\pullbackpr{1}{f}{g}$ and $\pullbackpr{2}{f}{g}$. So in this notation, a
-typical pullback diagram has the following form:
-\begin{equation*}
-\begin{tikzcd}[column sep=large]
-\pullback{A}{X}{f}{g}
-  \ar{r}{\pullbackpr{1}{f}{g}}
-  \ar{d}[swap]{\pullbackpr{2}{f}{g}}
-  &
-A \ar{d}{f}
-  \\
-X \ar{r}[swap]{g}
-  &
-B
-\end{tikzcd}
-\end{equation*}
-Also, when we have a commutative diagram of the form
-\begin{equation*}
-\begin{tikzcd}
-A \ar{r}{f}
-  \ar{d}{a}
-  &
-X \ar{d}
-  & 
-B \ar{l}[swap]{g}
-  \ar{d}{b}
-  \\
-A'
-  \ar{r}[swap]{f'}
-  &
-X'
-  &
-B'
-  \ar{l}{g'}
-\end{tikzcd}
-\end{equation*}
-we will denote the unique map from $\pullback{A}{B}{f}{g}$ to $\pullback{A'}{B'}{f'}{g'}$
-such that the diagram
-\begin{equation*}
-\begin{tikzcd}
-  {}
-  & 
-\pullback{A'}{B'}{f'}{g'}
-  \ar{dd}
-  \ar{rr}
-  &
-  &
-B'
-  \ar{dd}{g'}
-  \\
-\pullback{A}{B}{f}{g}
-  \ar{dd}
-  \ar[crossing over]{rr}
-  \ar[dotted]{ur}{\pullback{a}{b}{f'}{g'}}
-  &
-  &
-B \ar{ur}{b}
-  \\
-  {}
-  &
-A'
-  \ar{rr}
-  &
-  &
-X'
-  \\
-A \ar{rr}[swap]{f}
-  \ar{ur}{a}
-  &
-  &
-X \ar[crossing over,leftarrow]{uu}[near end,swap]{g}
-  \ar{ur}
-\end{tikzcd}
-\end{equation*}
-commutes, by $\pullback{a}{b}{f'}{g'}$. In the current work, we shall
-write $A\times B$ for the pullback of $A\rightarrow 1\leftarrow B$, and
-$\pi_1$ and $\pi_2$ for its projections (thus, no separate choice of
-cartesian products is made).
-
-\subsection{Extension objects}
-\begin{defn}
-A \emph{pre-extension object $\stesys$ in $\cat{C}$} consists of a \emph{fundamental structure}, which is a diagram of the form
-\begin{equation*}
-\begin{tikzcd}
-\stesyst
-  \ar{d}[swap]{\ebd}
-  \\
-\stesysf
-  \ar{d}[swap]{\eft}
-  \\
-\stesysc
-\end{tikzcd}
-\end{equation*}
-in $\cat{C}$ together with the \emph{context extension} and \emph{family extension} operations
-\begin{align*}
-\ectxext &:\stesysf\to \stesysc\\
-\efamext & :\stesysff\to \stesysf,
-\end{align*}
-respectively, such that the diagram
-\begin{equation*}
-\begin{tikzcd}
-\stesysf_2 
-  \ar{r}{\efamext} 
-  \ar{d}[swap]{\eft[2]} 
-  & 
-\stesysf 
-  \ar{d}{\eft}
-  \\
-\stesysf
-  \ar{r}[swap]{\eft} 
-  & 
-\stesysc
-\end{tikzcd}
-\end{equation*}
-commutes.
-\end{defn}
+\subsection{Substitution algebras}
 
 \begin{defn}
-We introduce the following notation:
-\begin{align*}
-\stesysf_2 
-  & := \stesysff
-  \\
-\eft[2] 
-  & := \pullbackpr{1}{\ectxext}{\eft} : \stesysf_2\to\stesysf
-  \\
-\stesysf_3 & := \pullback{\stesysf_2}{\stesysf_2}{\efamext}{\eft[2]}
-  \\
-\eft[3]
-  & := \pullbackpr{1}{\efamext}{\eft[2]} : \stesysf_3\to\stesysf_2.
-\end{align*}
-Then it follows that the outer square in the diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=large]
-\stesysf_3
-  \ar[dotted]{dr}{\eext{2}}
-  \ar{rr}{\pullback{\pullbackpr{2}{\ectxext}{\eft}}{\pullbackpr{2}{\ectxext}{\eft}}{\ectxext}{\eft}}
-  \ar{dd}[swap]{\eft[3]}
-  & 
-  &
-\stesysf_2
-  \ar{d}{\efamext}
-  \\
-  &
-\stesysf_2
-  \ar{d}[swap]{\eft[2]}
-  \ar{r}{\pullbackpr{2}{\ectxext}{\eft}}
-  &
-\stesysf
-  \ar{d}{\eft}
-  \\
-\stesysf_2
-  \ar{r}[swap]{\eft[2]}
-  &
-\stesysf
-  \ar{r}[swap]{\ectxext}
-  &
-\stesysc
-\end{tikzcd}
-\end{equation*}
-commutes. We define $\eext{2}$ to be the unique morphism rendering the above diagram
-commutative. Now we may continue to define
-\begin{align*}
-\stesysf_4 
-  & := 
-\pullback{\stesysf_3}{\stesysf_3}{\eext{2}}{\eft[3]}
-  \\
-\eft[4] 
-  & := 
-\pullbackpr{1}{\eext{2}}{\eft[3]}.
-\end{align*}
-Then we see that the outer square of the diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=large]
-\stesysf_4
-  \ar[dotted]{dr}{\eext{3}}
-  \ar{rr}{\pullback{\pullbackpr{2}{\efamext}{\eft[2]}}{\pullbackpr{2}{\efamext}{\eft[2]}}{\efamext}{\eft[2]}}
-  \ar{dd}[swap]{\eft[4]}
-  & 
-  &
-\stesysf_3
-  \ar{d}{\eext{2}}
-  \\
-  &
-\stesysf_3
-  \ar{d}[swap]{\eft[3]}
-  \ar{r}{\pullbackpr{2}{\efamext}{\eft[2]}}
-  &
-\stesysf_2
-  \ar{d}{\eft[2]}
-  \\
-\stesysf_3
-  \ar{r}[swap]{\eft[3]}
-  &
-\stesysf_2
-  \ar{r}[swap]{\efamext}
-  &
-\stesysf
-\end{tikzcd}
-\end{equation*}
-commutes,
-so we may define $\eext{3}$ to be the unique map which renders the diagram
-commutative. It
-is straightforward to continue this process by induction, but we shall need not
-go any further in this article.
-\end{defn}
-
-\begin{defn} 
-An extension object is a pre-extension object $\stesys$ for which 
-the diagrams
-\begin{equation*}
-\begin{tikzcd}
-\stesysf_2 
-  \ar{d}[swap]{\pullbackpr{2}{\ectxext}{\eft}} 
-  \ar{r}{\efamext} 
-  & 
-\stesysf 
-  \ar{d}{\ectxext}
-  \\
-\stesysf 
-  \ar{r}[swap]{\ectxext} 
-  & 
-\stesysc
-\end{tikzcd}
-\qquad
-\begin{tikzcd}
-\stesysf_3
-  \ar{d}[swap]{\pullbackpr{2}{\efamext}{\eft[2]}}
-  \ar{r}{\eext{2}}
-  & 
-\stesysf_2 
-  \ar{d}{\efamext} 
-  \\
-\stesysf_2 
-  \ar{r}[swap]{\efamext} 
-  &
-\stesysf
-\end{tikzcd}
-\end{equation*}
-commute.
-\end{defn}
-
-\begin{comment}
-\begin{lem}
-There exists an isomorphism $\alpha$ such that the triangle
-\begin{equation*}
-\begin{tikzcd}[column sep=tiny]
-\pullback{\stesysf}{\stesysf_2}{\ectxext}{\eft\circ\eft[2]}
-  \ar[dotted]{rr}{\alpha}
-  \ar{dr}[swap]{\pullback{\catid{\stesysf}}{\efamext}{\ectxext}{\eft}}
-  &
-  &
-\stesysf_3
-  \ar{dl}{\eext{2}}
-  \\
-& \stesysf_2
-\end{tikzcd}
-\end{equation*}
-commutes
-\end{lem}
-
-\begin{proof}
-There is a unique morphism $\alpha:
-\pullback{\stesysf}{\stesysf_2}{\ectxext}{\eft\circ\eft[2]}\to\stesysf_3$
-rendering the diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=large]
-\pullback{\stesysf}{\stesysf_2}{\ectxext}{\eft\circ\eft[2]}
-  \ar[bend left=10,yshift=.5ex]{drrr}{\pullbackpr{2}{\ectxext}{\eft}\circ\pullbackpr{2}{\ectxext}{\eft\circ\eft[2]}}
-  \ar[bend right=10]{ddr}[swap]{\pullback{\catid{\stesysf}}{\eft[2]}{\ectxext}{\eft}}
-  \ar[dotted]{dr}{\alpha}
-  \\
-& \stesysf_3
-  \ar{r}{\pullbackpr{2}{\efamext}{\eft[2]}}
-  \ar{d}{\eft[3]}
-  &
-\stesysf_2
-  \ar{d}[swap]{\eft[2]}
-  \ar{r}[swap]{\pullbackpr{2}{\ectxext}{\eft}}
-  &
-\stesysf
-  \ar{d}{\eft}
-  \\
-{} & \stesysf_2
-  \ar{r}[swap]{\efamext}
-  &
-\stesysf
-  \ar{r}[swap]{\ectxext}
-  &
-\stesysc
-\end{tikzcd}
-\end{equation*}
-\end{proof}
-\end{comment}
-
-\begin{defn}
-Suppose $\stesys$ is a pre-extension object of $\cat{C}$. Then we define the pre-extension object
-$\famesys{\stesys}$ to consist of the fundamental structure
-\begin{equation*}
-\begin{tikzcd}
-\stesyst_2
-  \ar{d}{\ebd[2]}
-  \\
-\stesysf_2
-  \ar{d}{\eft[2]}
-  \\
-\stesysf
-\end{tikzcd}
-\end{equation*}
-where
-\begin{align*}
-\stesyst_2 
-  & := \pullback{\stesysf}{\stesyst}{\ectxext}{\eft\circ\ebd}
-  \\
-\ebd[2]
-  & := \ectxext^\ast(\ebd),
-\end{align*}
-with the extension operations
-\begin{align*}
-\efamext 
-  & 
-  : \stesysf_2\to\stesysf\\
-\eext{2} & : \stesysf_3\to\stesysf_2.
-\end{align*}
-\end{defn}
-
-In \autoref{famextobj} below we shall show that $\famesys{\stesys}$ is an
-extension algebra whenever $\stesys$ is an extension algebra. We shall need
-a handful of lemmas to give the proof.
-
-\begin{defn}
-We define
-\begin{align*}
-\beta_1 
-  & := 
-\pullbackpr{2}{\ectxext}{\eft}
-  & &
-  : \stesysf_2\to\stesysf
-  \\
-\beta_2
-  & :=
-\pullback{\beta_1}{\beta_1}{\ectxext}{\eft}
-  & &
-  : \stesysf_3\to\stesysf_2
-  \\
-\beta_3
-  & :=
-\pullback{\beta_2}{\beta_2}{\efamext}{\eft[2]}
-  & &
-  : \stesysf_4\to\stesysf_3.
-\end{align*}
-\end{defn}
-
-\begin{lem}
-Let $\stesys$ be a pre-extension object. Then the square
-\begin{equation*}
-\begin{tikzcd}[column sep=10em]
-\stesysf_4
-  \ar{r}{\pullback{\pullbackpr{2}{\efamext}{\eft[2]}}{\pullbackpr{2}{\efamext}{\eft[2]}}{\efamext}{\eft[2]}}
-  \ar{d}[swap]{\beta_3}
-  &
-\stesysf_3
-  \ar{d}{\beta_2}
-  \\
-\stesysf_3
-  \ar{r}[swap]{\beta_2}
-  &
-\stesysf_2
-\end{tikzcd}
-\end{equation*}
-commutes.
-\end{lem}
-
-\begin{proof}
-Left to the reader.
-\end{proof}
-
-\begin{comment}
-\begin{proof}
-It is straightforward to verify the equalities
-\begin{align*}
-\pullbackpr{1}{\ectxext}{\eft}\circ\beta_2\circ
-  (\pullback{\pullbackpr{2}{\efamext}{\eft[2]}}{\pullbackpr{2}{\efamext}{\eft[2]}}{\efamext}{\eft[2]})
-  & =
-\pullbackpr{1}{\ectxext}{\eft}\circ\beta_2\circ\beta_3
-  \\
-\pullbackpr{2}{\ectxext}{\eft}\circ\beta_2\circ
-  (\pullback{\pullbackpr{2}{\efamext}{\eft[2]}}{\pullbackpr{2}{\efamext}{\eft[2]}}{\efamext}{\eft[2]})
-  & =
-\pullbackpr{2}{\ectxext}{\eft}\circ\beta_2\circ\beta_3.\qedhere
-\end{align*}
-\end{proof}
-\end{comment}
-
-Note that the square
-\begin{equation*}
-\begin{tikzcd}
-\stesysf_3
-  \ar{r}{\beta_2}
-  \ar{d}[swap]{\eext{2}}
-  &
-\stesysf_2
-  \ar{d}{\efamext}
-  \\
-\stesysf_2
-  \ar{r}[swap]{\beta_1}
-  &
-\stesysf
-\end{tikzcd}
-\end{equation*}
-commutes by definition. We have a similar result relating $\eext{3}$ and
-$\eext{2}$.
-
-\begin{lem}
-Let $\stesys$ be a pre-extension object. Then the square
-\begin{equation*}
-\begin{tikzcd}
-\stesysf_4
-  \ar{r}{\beta_3}
-  \ar{d}[swap]{\eext{3}}
-  &
-\stesysf_3
-  \ar{d}{\eext{2}}
-  \\
-\stesysf_3
-  \ar{r}[swap]{\beta_2}
-  &
-\stesysf_2
-\end{tikzcd}
-\end{equation*}
-commutes.
-\end{lem}
-
-\begin{proof}
-Left to the reader.
-\end{proof}
-\begin{comment}
-\begin{proof}
-It is straightforward to verify the equalities
-\begin{align*}
-\pullbackpr{1}{\ectxext}{\eft}\circ\beta_2\circ\eext{3}
-  & = \beta_1\circ\eft[3]\circ\eft[4]
-  \\
-\pullbackpr{1}{\ectxext}{\eft}\circ\eext{2}\circ\beta_3
-  & = \beta_1\circ\eft[3]\circ\eft[4].
-\end{align*}
-Thus, it remains to verify that
-\begin{equation*}
-\pullbackpr{2}{\ectxext}{\eft}\circ\beta_2\circ\eext{3}
-  = \pullbackpr{2}{\ectxext}{\eft}\circ\eext{2}\circ\beta_3.
-\end{equation*}
-It is straightforward to see that the diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=large]
-\stesysf_4
-  \ar{dd}[swap]{\pullback{\pullbackpr{2}{\efamext}{\eft[2]}}{\pullbackpr{2}{\efamext}{\eft[2]}}{\efamext}{\eft[2]}}
-  \ar{r}{\eext{3}}
-  &
-\stesysf_3
-  \ar{r}{\beta_2}
-  \ar{d}[swap]{\pullbackpr{2}{\efamext}{\eft[2]}}
-  &
-\stesysf_2
-  \ar{dd}{\pullbackpr{2}{\ectxext}{\eft}}
-  \\
-  {} &
-\stesysf_2
-  \ar{dr}{\beta_1}
-  \\
-\stesysf_3
-  \ar{ur}{\eext{2}}
-  \ar{r}[swap]{\beta_2}
-  &
-\stesysf_2
-  \ar{r}[swap]{\efamext}
-  &
-\stesysf
-\end{tikzcd}
-\end{equation*}
-commutes. It is likewise straightforward to see that the diagram
-\begin{equation*}
-\begin{tikzcd}
-\stesysf_4
-  \ar{r}{\beta_3}
-  \ar{d}[swap]{\pullback{\pullbackpr{2}{\efamext}{\eft[2]}}{\pullbackpr{2}{\efamext}{\eft[2]}}{\efamext}{\eft[2]}}
-  &
-\stesysf_3
-  \ar{r}{\eext{2}}
-  \ar{d}[swap]{\beta_2}
-  &
-\stesysf_2
-  \ar{d}{\pullbackpr{2}{\ectxext}{\eft}}
-  \\
-\stesysf_3
-  \ar{r}[swap]{\beta_2}
-  &
-\stesysf_2
-  \ar{r}[swap]{\efamext}
-  &
-\stesysf
-\end{tikzcd}
-\end{equation*}
-commutes, completing our goal.
-\end{proof}
-\end{comment}
-
-\begin{thm}[Local extension structure]\label{famextobj}
-If $\stesys$ is an extension object, then so is $\famesys{\stesys}$.
-\end{thm}
-
-\begin{proof}
-Note that the diagram
-\begin{equation*}
-\begin{tikzcd}
-\stesysf_3
-  \ar{d}[swap]{\pullbackpr{2}{\efamext}{\eft[2]}}
-  \ar{r}{\eext{2}}
-  & 
-\stesysf_2 
-  \ar{d}{\efamext} 
-  \\
-\stesysf_2 
-  \ar{r}[swap]{\efamext} 
-  &
-\stesysf
-\end{tikzcd}
-\end{equation*}
-commutes by assumption. For the second condition, we have to show that the
-diagram
-\begin{equation*}
-\begin{tikzcd}
-\stesysf_4
-  \ar{d}[swap]{\pullbackpr{2}{\eext{2}}{\eft[3]}}
-  \ar{r}{\eext{3}}
-  & 
-\stesysf_3
-  \ar{d}{\eext{2}} 
-  \\
-\stesysf_3
-  \ar{r}[swap]{\eext{2}} 
-  &
-\stesysf_2
-\end{tikzcd}
-\end{equation*}
-Since this is a question about two maps into a pullback, it suffices to verify
-that
-\begin{align*}
-\pullbackpr{1}{\ectxext}{\eft}\circ\eext{2}\circ\eext{3}
-  & =
-\pullbackpr{1}{\ectxext}{\eft}\circ\eext{2}\circ\pullbackpr{2}{\eext{2}}{\eft[3]}
-  \\
-\pullbackpr{2}{\ectxext}{\eft}\circ\eext{2}\circ\eext{3}
-  & =
-\pullbackpr{2}{\ectxext}{\eft}\circ\eext{2}\circ\pullbackpr{2}{\eext{2}}{\eft[3]}.
-\end{align*}
-For the first equality, it is fairly straightforward to show that both the
-equalities
-\begin{equation*}
-\pullbackpr{1}{\ectxext}{\eft}\circ\eext{2}\circ\eext{3}
-  =
-\eft[2]\circ\eft[3]\circ\eft[4]
-\end{equation*}
-and
-\begin{equation*}
-\pullbackpr{1}{\ectxext}{\eft}\circ\eext{2}\circ\pullbackpr{2}{\eext{2}}{\eft[3]}
-  =
-\eft[2]\circ\eft[3]\circ\eft[4].
-\end{equation*}
-hold. For the second subgoal (which is more tricky). Notice first that the
-diagram
-\begin{equation*}
-\begin{tikzcd}
-\stesysf_4
-  \ar{r}{\eext{3}}
-  \ar{d}[swap]{\beta_3}
-  &
-\stesysf_3
-  \ar{r}{\eext{2}}
-  \ar{d}[swap]{\beta_2}
-  &
-\stesysf_2
-  \ar{d}{\pullbackpr{2}{\ectxext}{\eft}}
-  \\
-\stesysf_3
-  \ar{r}[swap]{\eext{2}}
-  &
-\stesysf_2
-  \ar{r}[swap]{\efamext}
-  &
-\stesysf
-\end{tikzcd}
-\end{equation*}
-commutes. We also have the commutative diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=large]
-\stesysf_4
-  \ar{r}{\pullbackpr{2}{\eext{2}}{\eft[3]}}
-  \ar{d}[swap]{\beta_3}
-  &
-\stesysf_3
-  \ar{r}{\eext{2}}
-  \ar{d}[swap]{\beta_2}
-  &
-\stesysf_2
-  \ar{d}{\pullbackpr{2}{\ectxext}{\eft}}
-  \\
-\stesysf_3
-  \ar{r}{\pullbackpr{2}{\efamext}{\eft[2]}}
-  \ar{dr}[swap]{\eext{2}}
-  &
-\stesysf_2
-  \ar{r}{\efamext}
-  &
-\stesysf
-  \\
-  {} &
-\stesysf_2
-  \ar{ur}[swap]{\efamext}
-\end{tikzcd}
-\end{equation*}
-completing the proof.
-\end{proof}
-
-\subsection{(Pre-)extension homomorphisms}\label{subsection:e_extension_homomorphisms}
-In this subsection we start with the study of pre-extension homomorphisms, which
-will include the extension homomorphisms since they will be the pre-extension
-homomorphisms of which both the domain and codomain are extension objects.
-Our main examples of extension homomorphisms will be the operations of weakening
-and substitution. There are some basic examples of pre-extension homomorphisms
-that will be useful too, which get introduced in the this section and in
-\autoref{subsection:change_of_base}. In this section, we will mainly be
-interested in pre-extension homomorphisms between local pre-extension objects.
-We will end this section by proving that a retract of an extension object is
-always an extension object.
-
-\begin{defn}
-Let $\stesys$ and $\stesys'$ be pre-extension algebras. A \emph{pre-extension 
-homomorphism $f$ from $\stesys'$ to $\stesys$} is a triple $(f_0,f_1,f^t)$ 
-consisting of morphisms
-\begin{equation*}
-\begin{tikzcd}
-\stesyst' 
-  \ar{r}{f^t}
-  \ar{d}[swap]{\ebd'}
-  &
-\stesyst
-  \ar{d}{\ebd}
-  \\
-\stesysf'
-  \ar{r}{f_1}
-  \ar{d}[swap]{\eft'}
-  &
-\stesysf
-  \ar{d}{\eft}
-  \\
-\stesysc' 
-  \ar{r}[swap]{f_0}
-  &
-\stesysc
-\end{tikzcd}
-\end{equation*}
-such that the indicated squares commute, for which furthermore the squares
-\begin{equation*}
-\begin{tikzcd}
-\stesysf' 
-  \ar{r}{f_1}
-  \ar{d}[swap]{\ectxext'}
-  &
-\stesysf
-  \ar{d}{\ectxext}
-  \\
-\stesysc'
-  \ar{r}[swap]{f_0}
-  &
-\stesysc
-\end{tikzcd}
-\end{equation*}
-and
-\begin{equation*}
-\begin{tikzcd}[column sep=huge]
-\stesysf'\times_{\ectxext',\eft'} \stesysf'
-  \ar{r}{f_1\times_{\ectxext,\eft} f_1}
-  \ar{d}[swap]{\efamext'}
-  &
-\stesysf\times_{\ectxext,\eft} \stesysf
-  \ar{d}{\efamext}
-  \\
-\stesysf'
-  \ar{r}[swap]{f_1}
-  &
-\stesysf
-\end{tikzcd}
-\end{equation*}
-Composition and the identity homomorphism are defined in the expected way. We
-define furthermore
-\begin{align*}
-f_2 & := \pullback{f_1}{f_1}{\ectxext}{\eft}
-  \\
-f_3 & := \pullback{f_2}{f_2}{\efamext}{\eft[2]}.
-\end{align*}
-\end{defn}
-
-\begin{defn}
-A pre-extension homomorphism between extension objects is called an extension
-homomorphism.
-\end{defn}
-
-\begin{lem}
-Let $\stesys$ be an extension object. Then
-\begin{equation*}
-\begin{tikzcd}[column sep=huge]
-\stesyst_2
-  \ar{r}{\pullbackpr{2}{\ectxext}{\eft\circ\ebd}}
-  \ar{d}[swap]{\ebd[2]}
-  &
-\stesyst
-  \ar{d}{\ebd}
-  \\
-\stesysf_2
-  \ar{r}{\pullbackpr{2}{\ectxext}{\eft}}
-  \ar{d}[swap]{\eft[2]}
-  &
-\stesysf
-  \ar{d}{\eft}
-  \\
-\stesysf
-  \ar{r}[swap]{\ectxext}
-  &
-\stesysc
-\end{tikzcd}
-\end{equation*}
-assembles an extension homomorphism $\mathbf{e}_0:\famesys{\stesys}\to\stesys$.
-Likewise, we have an extension homomorphism
-$\mathbf{e}_1:\famesys{\famesys{\stesys}}\to\famesys{\stesys}$. Thus, a
-pre-extension object is an extension object if and only if $\mathbf{e}_0$
-and $\mathbf{e}_1$ are pre-extension homomorphisms.
-\end{lem}
-
-\begin{proof}
-Immediate from the conditions of being an extension object.
-\end{proof}
-
-\begin{defn}
-Let $\stesys$ be a pre-extension object. Then
-\begin{equation*}
-\begin{tikzcd}
-\stesyst_3
-  \ar{r}{\beta^t}
-  \ar{d}[swap]{\ebd[3]}
-  &
-\stesyst_2
-  \ar{d}{\ebd[2]}
-  \\
-\stesysf_3
-  \ar{r}{\beta_2}
-  \ar{d}[swap]{\eft[3]}
-  &
-\stesysf_2
-  \ar{d}{\eft[2]}
-  \\
-\stesysf_2
-  \ar{r}[swap]{\beta_1}
-  &
-\stesysf
-\end{tikzcd}
-\qquad
-\text{and}
-\qquad
-\begin{tikzcd}
-\stesyst_4
-  \ar{r}{\beta^t_2}
-  \ar{d}[swap]{\ebd[4]}
-  &
-\stesyst_3
-  \ar{d}{\ebd[3]}
-  \\
-\stesysf_4
-  \ar{r}{\beta_3}
-  \ar{d}[swap]{\eft[4]}
-  &
-\stesysf_3
-  \ar{d}{\eft[3]}
-  \\
-\stesysf_3
-  \ar{r}[swap]{\beta_2}
-  &
-\stesysf_2
-\end{tikzcd}
-\end{equation*}
-assemble pre-extension homomorphisms 
-\(
-\boldsymbol{\beta}
-  :
-\famesys{\famesys{\stesys}}
-  \to
-\famesys{\stesys}
-\) 
-and
-\(
-\boldsymbol{\beta}_\mathbf{2}
-  :
-\famesys{\famesys{\famesys{\stesys}}}
-  \to
-\famesys{\famesys{\stesys}}
-\).
-\end{defn}
-
-\begin{defn}\label{famehom}
-Suppose that $f:\stesys'\to\stesys$ is a pre-extension homomorphism. Then we
-define $\famehom{f}:\famesys{\stesys'}\to\famesys{\stesys}$ to consist of
-\begin{equation*}
-\begin{tikzcd}
-\stesyst_2'
-  \ar{r}{f^t_2}
-  \ar{d}[swap]{\ebd[2]'}
-  &
-\stesyst_2
-  \ar{d}{\ebd[2]}
-  \\
-\stesysf_2'
-  \ar{r}{f_2}
-  \ar{d}[swap]{\eft[2]'}
-  &
-\stesysf_2
-  \ar{d}{\eft[2]}
-  \\
-\stesysf'
-  \ar{r}[swap]{f_1}
-  &
-\stesysf
-\end{tikzcd}
-\end{equation*}
-where we define
-\begin{equation*}
-f^t_2 := \pullback{f_1}{f^t}{\ectxext}{\eft\circ\ebd}.
-\end{equation*}
-\end{defn}
-
-\begin{lem}
-The triple $\famehom{f}$ defined in \autoref{famehom} is a pre-extension homomorphism.
-\end{lem}
-
-\begin{proof}
-Note that the square
-\begin{equation*}
-\begin{tikzcd}
-\stesysf_2'
-  \ar{r}{f_2}
-  \ar{d}[swap]{\efamext'}
-  &
-\stesysf_2
-  \ar{d}{\efamext}
-  \\
-\stesysf'
-  \ar{r}[swap]{f_1}
-  &
-\stesysf
-\end{tikzcd}
-\end{equation*}
-commutes by assumption. Thus, it remains to show that the square
-\begin{equation*}
-\begin{tikzcd}
-\stesysf_3'
-  \ar{r}{f_3}
-  \ar{d}[swap]{\eext{2}'}
-  &
-\stesysf_3
-  \ar{d}{\eext{2}}
-  \\
-\stesysf_2'
-  \ar{r}[swap]{f_2}
-  &
-\stesysf_2
-\end{tikzcd}
-\end{equation*}
-commutes. It is equivalent to show that the equalities
-\begin{align*}
-\pullbackpr{1}{\ectxext}{\eft}\circ f_2\circ\eext{2}'
-  & =
-\pullbackpr{1}{\ectxext}{\eft}\circ \eext{2}\circ f_3
-  \\
-\pullbackpr{2}{\ectxext}{\eft}\circ f_2\circ\eext{2}'
-  & =
-\pullbackpr{2}{\ectxext}{\eft}\circ \eext{2}\circ f_3
-\end{align*}
-both hold. For the first, it is straightforward to verify that the diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=large]
-{} &
-\stesysf_2'
-  \ar{r}{f_2}
-  \ar{dr}[near end]{\pullbackpr{1}{\ectxext'}{\eft'}}
-  &
-\stesysf_2
-  \ar{dr}{\pullbackpr{1}{\ectxext}{\eft}}
-  \\
-\stesysf_3'
-  \ar{ur}{\eext{2}'}
-  \ar{r}[swap]{\beta_2'}
-  \ar{ddr}[swap]{f_3}
-  &
-\stesysf_2'
-  \ar{r}{\efamext'}
-  \ar{dr}[swap]{f_2}
-  &
-\stesysf'
-  \ar{r}{f_1}
-  &
-\stesysf
-  \\
-{} & {} &
-\stesysf_2
-  \ar{ur}[near start]{\efamext}
-  \\
-{} &
-\stesysf_3
-  \ar{r}[swap]{\eext{2}}
-  \ar{ur}{\beta_2}
-  &
-\stesysf_2
-  \ar{uur}[swap]{\pullbackpr{1}{\ectxext}{\eft}}
-\end{tikzcd}
-\end{equation*}
-commutes. For the second, note that the diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=large]
-{} &
-\stesysf_2'
-  \ar{r}{f_2}
-  \ar{dr}{\beta_1'}
-  &
-\stesysf_2
-  \ar{ddr}{\pullbackpr{2}{\ectxext}{\eft}}
-  \\
-{} & {} &
-\stesysf'
-  \ar{dr}[swap,near start]{f_1}
-  \\
-\stesysf_3'
-  \ar{uur}{\eext{2}'}
-  \ar{r}{\beta_2'}
-  \ar{dr}[swap]{f_3}
-  &
-\stesysf_2'
-  \ar{r}{f_2}
-  \ar{ur}{\efamext'}
-  &
-\stesysf_2
-  \ar{r}[swap]{\efamext}
-  &
-\stesysf
-  \\
-{} &
-\stesysf_3
-  \ar{r}[swap]{\eext{2}}
-  \ar{ur}[near start]{\beta_2}
-  &
-\stesysf_2
-  \ar{ur}[swap]{\pullbackpr{2}{\ectxext}{\eft}}
-\end{tikzcd}
-\end{equation*}
-commutes.
-\end{proof}
-
-\begin{lem}[Stability under retracts]\label{esys-retract}
-Suppose $f:\stesys\to\stesys'$ is a pre-extension homomorphism between
-pre-extension objects. If there is a pre-extension homomorphism $g:\stesys'\to
-\stesys$ such that $g\circ f=\catid{\stesys}$ and $\stesys'$ is an extension
-algebra, then $\stesys$ is an extension algebra.
-\end{lem}
-
-Before we start with the proof, note that we have the equalities
-$g_2\circ f_2=\catid{\stesysf_2}$ and $g_3\circ f_3=\catid{\stesysf_3}$
-under the hypotheses of the lemma.
-
-\begin{proof}
-Our first subgoal is to show that the square
-\begin{equation*}
-\begin{tikzcd}
-\stesysf_2 
-  \ar{r}{\efamext} 
-  \ar{d}[swap]{\pullbackpr{2}{\ectxext}{\eft}} 
-  & 
-\stesysf 
-  \ar{d}{\ectxext}
-  \\
-\stesysf
-  \ar{r}[swap]{\ectxext} 
-  & 
-\stesysc
-\end{tikzcd}
-\end{equation*}
-commutes. Note that in the diagram
-\begin{equation*}
-\begin{tikzcd}
-  {}
-  & 
-\stesysf
-  \ar{dd}[near start]{\ectxext}
-  \ar{rr}{f_1}
-  &
-  &
-\stesysf'
-  \ar{dd}[near start]{\ectxext'}
-  \ar{rr}{g_1}
-  &
-  &
-\stesysf
-  \ar{dd}{\ectxext}
-  \\
-\stesysf_2
-  \ar{dd}[swap]{\pullbackpr{2}{\ectxext}{\eft}}
-  \ar[crossing over]{rr}[swap,near start]{f_2}
-  \ar{ur}{\efamext}
-  &
-  &
-\stesysf_2'
-  \ar{ur}[near start]{\efamext'}
-  \ar[crossing over]{rr}[swap,near start]{g_2}
-  &
-  &
-\stesysf_2
-  \ar{ur}[swap,near start]{\efamext}
-  \\
-  {}
-  &
-\stesysc
-  \ar{rr}[near start]{f_0}
-  &
-  &
-\stesysc'
-  \ar{rr}[near start]{g_0}
-  &
-  &
-\stesysc
-  \\
-\stesysf 
-  \ar{rr}[swap]{f_1}
-  \ar{ur}{\ectxext}
-  &
-  &
-\stesysf' 
-  \ar[crossing over,leftarrow]{uu}[near end,swap]{\pullbackpr{2}{\ectxext'}{\eft'}}
-  \ar{ur}[swap,near end]{\ectxext'}
-  \ar{rr}[swap]{g_1}
-  &
-  &
-\stesysf
-  \ar[crossing over,leftarrow]{uu}[near end,swap]{\pullbackpr{2}{\ectxext}{\eft}}
-  \ar{ur}[swap]{\ectxext}
-\end{tikzcd}
-\end{equation*}
-all the faces minus the far left and far right face commute. Using that $g$
-is a section of $f$, we can read off that also the far left face commutes,
-completing our first subgoal.
-
-For the second subgoal, note that also $\famehom{g}\circ\famehom{f}=
-\catid{\famesys{\stesys}}$ and that $\famesys{\stesys'}$ is an extension object.
-Thus we can apply what we have proven so far to conclude that the square
-\begin{equation*}
-\begin{tikzcd}
-\stesysf_3 
-  \ar{r}{\eext{2}} 
-  \ar{d}[swap]{\pullbackpr{2}{\efamext}{\eft[2]}} 
-  & 
-\stesysf_2 
-  \ar{d}{\efamext}
-  \\
-\stesysf_2
-  \ar{r}[swap]{\efamext} 
-  & 
-\stesysf
-\end{tikzcd}
-\end{equation*}
-commutes.
-\end{proof}
-
-\subsection{The change of base of (pre-)extension objects}
-\label{subsection:change_of_base}
-An important construction of (pre-)extension objects is the change of base. It
-allows us to consider `parametrized homomorphisms', such as weakening and
-substitution.
-
-\begin{defn}
-Suppose $f:\stesys\to\stesys'$ is a pre-extension homomorphism. We say that
-a diagram
-\begin{equation*}
-\begin{tikzcd}
-\stesys
-  \ar{r}{f}
-  \ar{d}[swap]{p}
-  &
-\stesys'
-  \ar{d}{p'}
-  \\
-X \ar{r}[swap]{g}
-  &
-Y
-\end{tikzcd}
-\end{equation*}
-commutes if the diagram
-\begin{equation*}
-\begin{tikzcd}
-\stesysc
-  \ar{r}{f_0}
-  \ar{d}[swap]{p}
-  &
-\stesysc'
-  \ar{d}{p'}
-  \\
-X \ar{r}[swap]{g}
-  &
-Y
-\end{tikzcd}
-\end{equation*}
-commutes.
-\end{defn}
-
-The first goal in this subsection is to define for every (pre-)extension object 
-$\stesys$ and every $p:\stesysc\rightarrow X\leftarrow Y:g$, a (pre-)extension
-object $\cobesys{Y}{\stesys}{g}{p}$ with a homomorphism $\pullbackpr{2}{g}{p}:
-\cobesys{Y}{\stesys}{g}{p}\to\stesys$ and a morphism $\pullbackpr{1}{g}{p}:
-\pullback{Y}{\stesysc}{g}{p}\to Y$ such that for every diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=large]
-\stesys'
-  \ar[bend right=10]{ddr}[swap]{p'}
-  \ar[bend left=10]{rrd}{f}
-  \ar[dotted]{dr}[near end]{[p',f]}
-  \\
-  {}&
-\cobesys{Y}{\stesys}{g}{p}
-  \ar{d}{\pullbackpr{1}{g}{p}}
-  \ar{r}[swap]{\pullbackpr{2}{g}{p}}
-  &
-\stesys
-  \ar{d}{p}
-  \\
-  {}&
-Y \ar{r}[swap]{g}
-  &
-X
-\end{tikzcd}
-\end{equation*}
-of which the outer square commutes, the (pre-)extension homomorphism $[p',f]$ exists
-and is unique with the property that it renders the diagram commutative. We will
-give the definition of $\cobesys{Y}{\stesys}{g}{p}$ in \autoref{cobesys}. After
-proving that the change of base of a pre-extension algebra is indeed a
-pre-extension algebra (\autoref{cobesys-preext}) and that the change of base
-of an extension algebra is an extension algebra (\autoref{cobesys-ext}), we
-will demonstrate the above unique existence in \autoref{cobesys-existence,%
-cobesys-pullback}.
-
-The second goal in this subsection is to follow the same procedure for
-$\famesys{\famesys{\stesys}}$ to show that it is equivalent to
-$\cobesys{\stesysf}{\famesys{\stesys}}{\ectxext}{\eft}$. We will do this by
-verifying directly that it has the universal property of the change of base
-described above, because we will use the ingredients in our definition of
-weakening and substitution objects.
-
-\begin{defn}[Change of base]\label{cobesys}
-Suppose $\stesys$ is a pre-extension object in $\cat{C}$ and that 
-$p:\stesysc\rightarrow X\leftarrow Y:g$.
-Then we define the pre-extension object $\cobesys{Y}{\stesys}{g}{p}$ to consist of
-\begin{equation*}
-\begin{tikzcd}
-\cobesys{Y}{\stesyst}{g}{p\circ\eft\circ\ebd}
-  \ar{r}
-  \ar{d}[swap]{g^\ast(\ebd)}
-  &
-\stesyst
-  \ar{d}{\ebd}
-  \\
-\cobesys{Y}{\stesysf}{g}{p\circ\eft}
-  \ar{r}
-  \ar{d}[swap]{g^\ast(\eft)}
-  &
-\stesysf
-  \ar{d}{\eft}
-  \\
-\cobesys{Y}{\stesysc}{g}{p}
-  \ar{r}
-  \ar{d}[swap]{\pullbackpr{1}{g}{p}}
-  &
-\stesysc
-  \ar{d}{p}
-  \\
-Y \ar{r}[swap]{g}
-  &
-X
-\end{tikzcd}
-\end{equation*} 
-and the operations
-\begin{align*}
-\cobesys{Y}{\ectxext}{g}{p} 
-  & : \pullback{Y}{\stesysf}{g}{p\circ\eft}\to \pullback{Y}{\stesysc}{g}{p}\\
-\cobesys{Y}{\efamext}{g}{p} 
-  & : \pullback
-    {\pullback{Y}{\stesysf}{g}{p\circ\eft}}
-    {\pullback{Y}{\stesysf}{g}{p\circ\eft}}
-    {\cobesys{Y}{\ectxext}{g}{p}}
-    {g^\ast(\eft)}
-  \to 
-  \pullback{Y}{\stesysf}{g}{p\circ\eft}.
-\end{align*}
-defined by
-\begin{equation*}
-\cobesys{Y}{\ectxext}{g}{p} := \pullback{\catid{Y}}{\ectxext}{g}{p}
-\end{equation*}
-and where $\cobesys{Y}{\efamext}{g}{p}$ is defined by rendering the diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=large]
-(\cobesys{Y}{\stesysf}{g}{p\circ\eft})_2
-  \ar{rr}{\pullback{\pullbackpr{2}{g}{p\circ\eft}}{\pullbackpr{2}{g}{p\circ\eft}}{\ectxext}{\eft}}
-  \ar{dd}[swap]{\pullbackpr{1}{\cobesys{Y}{\ectxext}{g}{p}}{g^\ast(\eft)}}
-  \ar[dotted]{dr}[swap]{\cobesys{Y}{\efamext}{g}{p}}
-  &
-  &
-\stesysf_2
-  \ar{d}{\efamext}
-  \\
-  {}&
-\cobesys{Y}{\stesysf}{g}{p\circ\eft}
-  \ar{r}{\pullbackpr{2}{g}{p\circ\eft}}
-  \ar{d}[swap]{\pullbackpr{1}{g}{p\circ\eft}}
-  &
-\stesysf
-  \ar{d}{p\circ\eft}
-  \\
-\cobesys{Y}{\stesysf}{g}{p\circ\eft}
-  \ar{r}[swap]{\pullbackpr{1}{g}{p\circ\eft}}
-  &
-Y \ar{r}[swap]{g}
-  &
-X
-\end{tikzcd}
-\end{equation*} 
-commutative. 
-The process of obtaining the pre-extension object $\cobesys{Y}{\stesys}{g}{p}$ out of $\stesys$
-and $g:Y\to X$ is also called the \emph{change of base}.
-\end{defn}
-
-\begin{lem}\label{cobesys-preext}
-Any change of base of a pre-extension object is a pre-extension object.
-\end{lem}
-
-\begin{proof}
-Let $\stesys$ be an extension algebra and consider $p:\stesysc\rightarrow X\leftarrow Y:g$.
-We need to verify that the square
-\begin{equation*}
-\begin{tikzcd}[column sep=large]
-(\pullback{Y}{\stesysf}{g}{p\circ\eft})_2
-  \ar{r}{\cobesys{Y}{\efamext}{g}{p}} 
-  \ar{d}[swap]{\pullbackpr{1}{\cobesys{Y}{\ectxext}{g}{p}}{g^\ast(\eft)}} 
-  & 
-\pullback{Y}{\stesysf}{g}{p\circ\eft}
-  \ar{d}{g^\ast(\eft)}
-  \\
-\pullback{Y}{\stesysf}{g}{p\circ\eft}
-  \ar{r}[swap]{g^\ast(\eft)} 
-  & 
-\pullback{Y}{\stesysc}{g}{p}
-\end{tikzcd}
-\end{equation*}
-commutes. It is fairly obvious that
-\begin{equation*}
-\pullbackpr{1}{g}{p}\circ g^\ast(\eft)\circ (\cobesys{Y}{\efamext}{g}{p})
-  =
-\pullbackpr{1}{g}{p\circ\eft}\circ \pullbackpr{1}{\cobesys{Y}{\ectxext}{g}{p}}{g^\ast(\eft)}
-\end{equation*}
-and that the diagram
-\begin{equation*}
-\begin{tikzcd}
-  {}&
-  {}&
-\pullback{Y}{\stesysf}{g}{p\circ\eft}
-  \ar{rr}{g^\ast(\eft)}
-  \ar{dr}[swap]{\pullbackpr{2}{g}{p\circ\eft}}
-  &
-  {}&
-\pullback{Y}{\stesysc}{g}{p}
-  \ar{ddr}{\pullbackpr{2}{g}{p}}
-  \\
-  {}&
-  {}&
-  {}&
-\stesysf
-  \ar{drr}[swap]{\eft}
-  \\
-(\pullback{y}{\stesysf}{g}{p\circ\eft})_2
-  \ar{uurr}{\cobesys{Y}{\efamext}{g}{p}}
-  \ar{rr}[swap,yshift=-.5ex]{\pullback{\pullbackpr{2}{g}{p\circ\eft}}{\pullbackpr{2}{g}{p\circ\eft}}{\ectxext}{\eft}}
-  \ar{ddrr}[swap]{\pullbackpr{1}{\cobesys{Y}{\ectxext}{g}{p}}{g^\ast(\eft)}}
-  &
-  {}&
-\stesysf_2
-  \ar{ur}{\efamext}
-  \ar{dr}[swap]{\eft[2]}
-  &
-  {}&
-  {}&
-\stesysc
-  \\
-  {}&
-  {}&
-  {}&
-\stesysf
-  \ar{urr}{\eft}
-  \\
-  {}&
-  {}&
-\pullback{Y}{\stesysf}{g}{p\circ\eft}
-  \ar{rr}[swap]{g^\ast(\eft)}
-  \ar{ur}{\pullbackpr{2}{g}{p\circ\eft}}
-  &
-  {}&
-\pullback{Y}{\stesysc}{g}{p}
-  \ar{uur}[swap]{\pullbackpr{2}{g}{p}}
-\end{tikzcd}
-\end{equation*}
-commutes.
-\end{proof}
-
-\begin{thm}\label{cobesys-ext}
-The change of base of an extension algebra is an extension algebra.
-\end{thm}
-
-\begin{proof}
-Our first subgoal is to verify that the square
-\begin{equation*}
-\begin{tikzcd}[column sep=large]
-(\pullback{Y}{\stesysf}{g}{p\circ\eft})_2
-  \ar{r}{\cobesys{Y}{\efamext}{g}{p}} 
-  \ar{d}[swap]{\pullbackpr{2}{\cobesys{Y}{\ectxext}{g}{p}}{g^\ast(\eft)}} 
-  & 
-\pullback{Y}{\stesysf}{g}{p\circ\eft}
-  \ar{d}{\cobesys{Y}{\ectxext}{g}{p}}
-  \\
-\pullback{Y}{\stesysf}{g}{p\circ\eft}
-  \ar{r}[swap]{\cobesys{Y}{\ectxext}{g}{p}} 
-  & 
-\pullback{Y}{\stesysc}{g}{p}
-\end{tikzcd}
-\end{equation*}
-\end{proof}
-
-The following construction is useful for defining extension homomorphisms into
-`higher' extension objects
-
-\begin{defn}\label{cobesys-existence}
-Consider a commutative diagram
-\begin{equation*}
-\begin{tikzcd}
-\stesys'
-  \ar{r}{f}
-  \ar{d}[swap]{p'}
-  &
-\stesys
-  \ar{d}{p}
-  \\
-Y \ar{r}[swap]{g}
-  &
-X
-\end{tikzcd}
-\end{equation*}
-Then we construct $[p,f]:\stesys'\to\cobesys{Y}{\stesys}{g}{p}$
-\begin{itemize}
-\item by defining $[p,f]_0:\stesysc'\to\pullback{Y}{\stesysc}{g}{p}$ be the uniqe
-morphism rendering the diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=large]
-\stesysc'
-  \ar[bend right=10]{ddr}[swap]{p'}
-  \ar[bend left=10]{rrd}{f_0}
-  \ar{dr}[near end]{[p,f]_0}
-  \\
-  {}&
-\pullback{Y}{\stesysc}{g}{p}
-  \ar{r}[swap]{\pullbackpr{2}{g}{p}}
-  \ar{d}{\pullbackpr{1}{g}{p}}
-  &
-\stesysc
-  \ar{d}{p}
-  \\
-  {}&
-Y \ar{r}[swap]{g}
-  &
-X
-\end{tikzcd}
-\end{equation*}
-commutative.
-\item by defining $[p,f]_1:\stesysf'\to\pullback{Y}{\stesysf}{g}{p\circ\eft}$ be the uniqe
-morphism rendering the diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=huge]
-\stesysf'
-  \ar[bend right=10]{ddr}[swap]{p'\circ\eft'}
-  \ar[bend left=10]{rrd}{f_1}
-  \ar{dr}[near end]{[p,f]_1}
-  \\
-  {}&
-\pullback{Y}{\stesysf}{g}{p\circ\eft}
-  \ar{r}[swap]{\pullbackpr{2}{g}{p\circ\eft}}
-  \ar{d}{\pullbackpr{1}{g}{p\circ\eft}}
-  &
-\stesysc
-  \ar{d}{p\circ\eft}
-  \\
-  {}&
-Y \ar{r}[swap]{g}
-  &
-X
-\end{tikzcd}
-\end{equation*}
-commutative.
-\item by defining $[p,f]^t:\stesyst'\to\pullback{Y}{\stesyst}{g}{p\circ\eft\circ\ebd}$ be the uniqe
-morphism rendering the diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=huge]
-\stesyst'
-  \ar[bend right=10]{ddr}[swap]{p'\circ\eft'\circ\ebd'}
-  \ar[bend left=10]{rrd}{f^t}
-  \ar{dr}[near end]{[p,f]^t}
-  \\
-  {}&
-\pullback{Y}{\stesysf}{g}{p\circ\eft\circ\ebd}
-  \ar{r}[swap]{\pullbackpr{2}{g}{p\circ\eft\circ\ebd}}
-  \ar{d}{\pullbackpr{1}{g}{p\circ\eft\circ\ebd}}
-  &
-\stesysc
-  \ar{d}{p\circ\eft\circ\ebd}
-  \\
-  {}&
-Y \ar{r}[swap]{g}
-  &
-X
-\end{tikzcd}
-\end{equation*}
-commutative.
-\end{itemize}
-\end{defn}
-
-\begin{thm}\label{cobesys-pullback}
-For every diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=large]
-\stesys'
-  \ar[bend right=10]{ddr}[swap]{p'}
-  \ar[bend left=10]{rrd}{f}
-  \ar[dotted]{dr}[near end]{[p',f]}
-  \\
-  {}&
-\cobesys{Y}{\stesys}{g}{p}
-  \ar{d}{\pullbackpr{1}{g}{p}}
-  \ar{r}[swap]{\pullbackpr{2}{g}{p}}
-  &
-\stesys
-  \ar{d}{p}
-  \\
-  {}&
-Y \ar{r}[swap]{g}
-  &
-X
-\end{tikzcd}
-\end{equation*}
-of which the outer square commutes, the pre-extension homomorphism $[p',f]$
-is unique with the property that it renders the whole diagram commutative.
-\end{thm}
-
-\begin{defn}\label{famfamstesys_into}
-Consider a commutative square
-\begin{equation*}
-\begin{tikzcd}
-\stesys'
-  \ar{r}{f}
-  \ar{d}[swap]{p}
-  &
-\famesys\stesys
-  \ar{d}{\eft}
-  \\
-\stesysf \ar{r}[swap]{\ectxext}
-  &
-\stesysc
-\end{tikzcd}
-\end{equation*}
-Then we construct
-\begin{equation*}
-[p,f]:\stesys'\to\famesys{\famesys{\stesys}}
-\end{equation*}
-as follows:
-\begin{itemize}
-\item let $[p,f]_0:\stesysc'\to\stesysf_2$ be the unique morphism rendering
-the diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=large]
-\stesysc' 
-  \ar[bend left=10]{rrd}{f_0}
-  \ar[swap,bend right=10]{ddr}{p}
-  \ar[dotted]{dr}[near end]{[p,f]_0}
-  \\
-  {}&
-\stesysf_2
-  \ar{r}[swap]{\pullbackpr{2}{\ectxext}{\eft}}
-  \ar{d}{\eft[2]}
-  &
-\stesysf
-  \ar{d}{\eft}
-  \\
-  {}&
-\stesysf
-  \ar{r}[swap]{\ectxext}
-  &
-\stesysc
-\end{tikzcd}
-\end{equation*}
-commutative.
-\item Let $[p,f]_1:\stesysf'\to\stesysf_3$ be the unique morphism rendering
-the diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=large]
-\stesysf'
-  \ar[bend left=10]{drr}{f_1}
-  \ar[swap]{dd}{\eft'}
-  \ar[dotted]{dr}[near end]{[p,f]_1}
-  \\
-  {}&
-\stesysf_3
-  \ar{r}[swap]{\pullbackpr{2}{\efamext}{\eft[2]}}
-  \ar{d}{\eft[2]}
-  &
-\stesysf_2
-  \ar{d}{\eft[2]}
-  \\
-\stesysc'
-  \ar{r}[swap]{[p,f]_0}
-  &
-\stesysf_2
-  \ar{r}[swap]{\efamext}
-  &
-\stesysf
-\end{tikzcd}
-\end{equation*}
-commutative.
-\item Let $[p,f]^t:\stesyst'\to\stesyst_3$ be the unique morphism rendering
-the diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=huge]
-\stesyst'
-  \ar[bend left=10]{drr}{f^t}
-  \ar[swap]{dd}{\eft'\circ\ebd'}
-  \ar[dotted]{dr}[near end]{[p,f]^t}
-  \\
-  {}&
-\stesyst_3
-  \ar{r}[swap]{\pullbackpr{2}{\efamext}{\eft[2]\circ\ebd[2]}}
-  \ar{d}[swap]{\pullbackpr{1}{\efamext}{\eft[2]\circ\ebd[2]}}
-  &
-\stesyst_2
-  \ar{d}{\eft[2]\circ\ebd[2]}
-  \\
-\stesysc'
-  \ar{r}[swap]{[p,f]_0}
-  &
-\stesysf_2
-  \ar{r}[swap]{\efamext}
-  &
-\stesysf
-\end{tikzcd}
-\end{equation*}
-commutative.
-\end{itemize}
-\end{defn}
-
-\begin{lem}
-Under the hypotheses of \autoref{famfamstesys_into}, $[p,f]$ is a pre-extension
-homomorphism. Moreover, it is the unique pre-extension homomorphism for which
-the diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=large]
-\stesys' 
-  \ar[bend left=10]{rrd}{f}
-  \ar[swap,bend right=10]{ddr}{p}
-  \ar[dotted]{dr}[near end]{[p,f]}
-  \\
-  {}&
-\famesys{\famesys{\stesys}}
-  \ar{r}[swap]{\pullbackpr{2}{\ectxext}{\eft}}
-  \ar{d}{\eft[2]}
-  &
-\famesys{\stesys}
-  \ar{d}{\eft}
-  \\
-  {}&
-\stesysf
-  \ar{r}[swap]{\ectxext}
-  &
-\stesysc
-\end{tikzcd}
-\end{equation*}
-commutes.
-\end{lem}
-
-\begin{lem}
-Suppose $f:\stesys\to \stesys'$ is a pre-extension homomorphism and consider a morphism
-$p:\stesys'\to X$ and $g:Y\to X$. Then the change of base 
-$g^\ast(f):\cobesys{Y}{\stesys}{g}{p\circ f_0}\to
-\cobesys{Y}{\stesys'}{g}{p}$ is a pre-extension morphism.
-\end{lem}
-
-\begin{lem}
-Let $\stesys$ be a pre-extension algebra and consider $p:\stesysc\rightarrow X\leftarrow Y:g$.
-Then there is an isomorphism
-\begin{equation*}
-\varphi:\famesys{\cobesys{Y}{\stesys}{g}{p}}
-  \simeq
-\cobesys{Y}{\famesys{\stesys}}{g}{p\circ\eft}
-\end{equation*}
-uniquely determined by
-\end{lem}
-
-\begin{proof}
-This follows from the pasting lemma for pullbacks.
-\end{proof}
-
-\subsection{Pre-weakening objects}
-\begin{defn}
-Let $\stesys$ be an extension object in $\cat{C}$. A pre-weakening operation
-on $\stesys$ is an extension homomorphism 
-$ \mathbf{w}(\stesys)
-    :
-  \cobesys{\stesysf}{\famesys{\stesys}}{\eft}{\eft}
-    \to
-  \famesys{\famesys{\stesys}}$
-for which the diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=large]
-\cobesys{\stesysf}{\famesys{\stesys}}{\eft}{\eft}
-  \ar{r}{\mathbf{w}(\stesys)}
-  \ar{dr}[swap]{\pullbackpr{1}{\eft}{\eft}}
-  &
-\famesys{\famesys{\stesys}}
-  \ar{d}{\eft[2]}
-  \\
-& \stesysf
-\end{tikzcd}
-\end{equation*}
-commutes.
-\end{defn}
-
-\begin{defn}
-Let $\stesys$ be an extension object with pre-weakening operation
-$\mathbf{w}(\stesys)$. Then $\famesys{\stesys}$ has the pre-weakening operation
-$\mathbf{w}(\famesys{\stesys})$ which is uniquely determined by rendering the
-diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=large]
-\cobesys{\stesysf_2}{\famesys{\famesys{\stesys}}}{\eft[2]}{\eft[2]}
-  \ar{rr}{%
-      \pullback{\beta_1}{\boldsymbol{\beta}}{\eft}{\eft}
-    }
-  \ar[bend right]{ddr}[swap]{\pullbackpr{1}{\eft[2]}{\eft[2]}}
-  \ar[dotted]{dr}{\mathbf{w}(\famesys{\stesys})}
-  &
-  {}&
-\cobesys{\stesysf}{\famesys{\stesys}}{\eft}{\eft}
-  \ar{r}{\mathbf{w}(\stesys)}
-  &
-\famesys{\famesys{\stesys}}
-  \ar{d}{\boldsymbol{\beta}}
-  \\
-  {}&
-\famesys{\famesys{\famesys{\stesys}}}
-  \ar{r}{\boldsymbol{\beta}_\mathbf{2}}
-  \ar{d}[swap]{\eft[3]}
-  &
-\famesys{\famesys{\stesys}}
-  \ar{d}{\eft[2]}
-  \ar{r}{\boldsymbol{\beta}}
-  &
-\famesys{\stesys}
-  \ar{d}{\eft}
-  \\
-  {}&
-\stesysf_2
-  \ar{r}[swap]{\efamext}
-  &
-\stesysf
-  \ar{r}[swap]{\ectxext}
-  &
-\stesysc
-\end{tikzcd}
-\end{equation*}
-commutative.
-\end{defn}
-
-\begin{defn}
-A pre-weakening object $\stesys$ in $\cat{C}$ is an extension object $\stesys$ 
-with a pre-weakening operation 
-$ \mathbf{w}(\stesys)
-    :
-  \cobesys{\stesysf}{\famesys{\stesys}}{\eft}{\eft}
-    \to
-  \famesys{\famesys{\stesys}}$
-for which the diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=15em]
-\cobesys{\stesysf_2}{\famesys{\stesys}}{\eft\circ\eft[2]}{\eft}
-  \ar[bend right=10]{dr}[swap]%
-    { [ \pullbackpr{1}{\eft\circ\eft[2]}{\eft},%
-        \mathbf{w}(\stesys)%
-          \circ%
-        (\pullback{\efamext}{\catid{\famesys{\stesys}}}{\eft}{\eft})%
-        ]%
-      }
-  \ar{r}{
-    [ \pullbackpr{1}{\eft\circ\eft[2]}{\eft},%
-      \mathbf{w}(\stesys)%
-        \circ%
-      (\pullback{\eft[2]}{\catid{\famesys{\stesys}}}{\eft}{\eft})%
-      ]}%
-  &
-\cobesys{\stesysf_2}{\famesys{\famesys{\stesys}}}{\eft[2]}{\eft[2]}
-  \ar{d}{\mathbf{w}(\famesys{\stesys})}
-  \\
-  {}&
-\famesys{\famesys{\famesys{\stesys}}}
-\end{tikzcd}
-\end{equation*}
-commutes. This condition is called \emph{Currying for weakening}.
-\end{defn}
-
-\begin{lem}
-If $\stesys$ is a pre-weakening algebra, then so is $\famesys{\stesys}$. 
-\end{lem}
-
-\begin{proof}
-\end{proof}
-
-\begin{defn}
-A pre-weakening morphism between preweakening objects $\stesys$ and $\stesys'$ is an
-extension homomorphism $f:\stesys\to \stesys'$ such that additionally the diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=large]
-\cobesys{\stesysf}{\famesys{\stesys}}{\eft}{\eft}
-  \ar{d}[swap]{\mathbf{w}(\stesys)}
-  \ar{r}{\pullback{f_1}{\famehom{f}}{\eft'}{\eft'}}
-  &
-\cobesys{\stesysf'}{\famesys{\stesys'}}{\eft'}{\eft'}
-  \ar{d}{\mathbf{w}(\stesys')}
-  \\
-\famesys{\famesys{\stesys}}
-  \ar{r}[swap]{\famehom{\famehom{f}}}
-  &
-\famesys{\famesys{\stesys'}}
-\end{tikzcd}
-\end{equation*}
-commutes.
-\end{defn}
-
-\begin{defn}
-Let $\stesys$ be a pre-weakening algebra and consider $p:\stesysc\rightarrow X\leftarrow Y:p$.
-Then we define
-\begin{equation*}
-\mathbf{w}(\cobesys{Y}{\stesys}{g}{p})
-  :
-\cobesys
-  { (\pullback{Y}{\stesysf}{g}{p\circ\eft})}
-  { \famesys{\cobesys{Y}{\stesys}{g}{p}}}
-  { g^\ast(\eft)}
-  { g^\ast(\eft)}
-  \to
-\famesys{\famesys{\cobesys{Y}{\stesys}{g}{p}}}
-\end{equation*}
-to be the unique extension homomorphism rendering the diagram
-\begin{equation*}
-\begin{tikzcd}
-\cobesys
-  { \pullback{Y}{\stesysf}{g}{p\circ\eft}}
-  { \famesys{\cobesys{Y}{\stesys}{g}{p}}}
-  { g^\ast(\eft)}
-  { g^\ast(\eft)}
-  \ar[bend right]{ddr}[swap]{\pullbackpr{1}{g^\ast(\eft)}{g^\ast(\eft)}}
-  \ar{rr}{%
-    \pullback
-      { \pullbackpr{2}{g}{p\circ\eft}}
-      { \boldsymbol{\pi}_\mathbf{2}(g,p\circ\eft)}
-      { \eft}
-      { \eft}
-    }
-  \ar[dotted]{dr}{\mathbf{w}(\cobesys{Y}{\stesys}{g}{p})}
-  &
-  {}&
-\cobesys{\stesysf}{\famesys{\stesys}}{\eft}{\eft}
-  \ar{r}{\mathbf{w}(\stesys)}
-  &
-\famesys{\famesys{\stesys}}
-  \ar{d}{\boldsymbol{\beta}}
-  \\
-  {}&
-\famesys{\famesys{\cobesys{Y}{\stesys}{g}{p}}}
-  \ar{r}{\beta}
-  \ar{d}{g^\ast(\eft)_1}
-  &
-\famesys{\cobesys{Y}{\stesys}{g}{p}}
-  \ar{r}{\boldsymbol{\pi}_2(g,p\circ\eft)}
-  \ar{d}{g^\ast(\eft)}
-  &
-\famesys{\stesys}
-  \ar{d}{\eft}
-  \\
-  {}&
-\pullback{Y}{\stesysf}{g}{p\circ\eft}
-  \ar{r}[swap]{\cobesys{Y}{\ectxext}{g}{p}}
-  &
-\pullback{Y}{\stesysc}{g}{p}
-  \ar{r}[swap]{\pullbackpr{2}{g}{p}}
-  &
-\stesysc
-\end{tikzcd}
-\end{equation*}
-commutative.
-\end{defn}
-
-\subsection{Weakening objects}
-Since we have shown that the property of being a pre-weakening object is closed
-under the relevant operations, we can make the following definition:
-
-\begin{defn}
-A weakening object is a pre-weakening object $\stesys$ with the property that
-$\mathbf{w}(\stesys)$ is a pre-weakening morphism.
-\end{defn}
-
-\begin{defn}
-A weakening homomorphism is a pre-weakening homomorphism such that the domain
-and codomain are weakening objects.
-\end{defn}
-
-\begin{thm}
-Suppose $\stesys$ is a weakening object, then so is $\famesys{\stesys}$
-\end{thm}
-
-\begin{thm}
-The change of base of any weakening object is again a weakening object.
-\end{thm}
-
-\subsection{Projection objects}
-\begin{defn}
-A pre-projection object is a weakening object $\stesys$ for which there is a term
-$\mathbf{i}:\stesysf\to \stesyst_2$ such that the diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=large]
-\stesysf \ar{r}{\mathbf{i}} \ar{d}[swap]{\Delta_{\eft}} & \stesyst_2 \ar{d}{\ebd[2]}\\
-\pullback{\stesysf}{\stesysf}{\eft}{\eft} \ar{r}[swap]{w(\stesys)_0} & \stesysf_2
-\end{tikzcd}
-\end{equation*}
-commutes. In this diagram $\Delta_{\eft}:\stesysf\to \pullback{\stesysf}{\stesysf}{\eft}{\eft}$ is the diagonal.
-\end{defn}
-
-\begin{defn}
-A pre-projection homomorphism from $\stesys$ to $\stesys'$ is a weakening homomorphism
-$f:\stesys\to \stesys'$ such that the square
-\begin{equation*}
-\begin{tikzcd}[column sep=large]
-\stesyst_2
-  \ar{r}{f^t_1}
-  &
-\stesyst_2'
-  \\
-\stesysf \ar{r}[swap]{f_1}
-  \ar{u}{\mathbf{i}}
-  &
-\stesysf'
-  \ar{u}[swap]{\mathbf{i}'}
-\end{tikzcd}
-\end{equation*}
-commutes
-\end{defn}
-
-\begin{lem}
-The change of base of a pre-projection object is again a pre-projection object.
-\end{lem}
-
-\begin{lem}
-If $CFT$ is a pre-projection object, then so is $\mathbf{F}_{CFT}$, where
-$\mathbf{F}_{\mathbf{i}}$ is defined to be $F\times_{e_0,c}\mathbf{i}$ is
-a pre-projection algebra.
-\end{lem}
-
-\begin{defn}
-A projection algebra is a pre-projection algebra for which weakening is a
-pre-projection homomorphism.
-\end{defn}
-
-\begin{cor}
-The change of base of a projection object is again a projection object.
-\end{cor}
-
-\begin{cor}
-If $CFT$ is a projection object, then so is $\mathbf{F}_{CFT}$, where
-$\mathbf{F}_{\mathbf{i}}$ is defined to be $F\times_{e_0,c}\mathbf{i}$ is
-a projection algebra.
-\end{cor}
-
-\subsection{Substitution objects}
-
-\begin{defn}
-A \emph{pre-substitution} for an extension object $\stesys$ is an
+A \emph{pre-substitution} for an extension algebra $\stesys$ is an
 extension homomorphism
@@
 \end{equation*}
-commutes. A \emph{pre-substitution object} is an extension object
+commutes. A \emph{pre-substitution algebra} is an extension algebra
 together with a pre-substitution.
@@
 \begin{lem}
-The change of base of a pre-substitution object is again a pre-substitution object.
+The change of base of a pre-substitution algebra is again a pre-substitution algebra.
 \end{lem}
@@
 \begin{lem}
-If $\stesys$ is a pre-substitution object, then so is $\famesys{\stesys}$ with
+If $\stesys$ is a pre-substitution algebra, then so is $\famesys{\stesys}$ with
 $\mathbf{s}(\famesys{\stesys})$ defined to be the unique extension homomorphism
@@
 \begin{defn}
-A \emph{substitution object} is a pre-substitution object for which substitution is
+A \emph{substitution algebra} is a pre-substitution algebra for which substitution is
 a pre-substitution homomorphism.
@@
 \begin{cor}
-The change of base of a substitution object is again a substitution object.
+The change of base of a substitution algebra is again a substitution algebra.
 \end{cor}
@@
 \begin{cor}
-If $\stesys$ is a substitution object, then so is $\famesys{\stesys}$.
+If $\stesys$ is a substitution algebra, then so is $\famesys{\stesys}$.
 \end{cor}
-
-\subsection{Extension objects with empty context and families}
-
-\begin{defn}
-An extension object $\stesys$ is said to have \emph{empty families} if there
-is a section
-\begin{equation*}
-\phi_1(\stesys):\stesysc\to\stesysf
-\end{equation*}
-of $\eft$, satisfying the following additional properties:
-\begin{enumerate}
-\item $\phi_1(\stesys)$ is also a section of $\ectxext$.
-\item The unique morphism $\phi_2$ rendering the diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=huge]
-\stesysf
-  \ar[bend left=20]{drr}{\phi_1(\stesys)\circ\ectxext}
-  \ar[equals,bend right=20]{ddr}
-  \ar[dotted]{dr}{\phi_2}
-  \\
-  {}&
-\stesysf_2
-  \ar{d}{\eft[2]}
-  \ar{r}{\pullbackpr{2}{\ectxext}{\eft}}
-  &
-\stesysf
-  \ar{d}{\eft}
-  \\
-  {}&
-\stesysf
-  \ar{r}[swap]{\ectxext}
-  &
-\stesysc
-\end{tikzcd}
-\end{equation*}
-commutative, is a section of $\efamext$.
-\item The unique morphism $\iota_1$ rendering the diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=huge]
-\stesysf
-  \ar[equals,bend left=20]{drr}
-  \ar[bend right=20]{ddr}[swap]{\phi_1(\stesys)\circ\eft}
-  \ar[dotted]{dr}{\iota_1}
-  \\
-  {}&
-\stesysf_2
-  \ar{d}{\eft[2]}
-  \ar{r}[swap]{\pullbackpr{2}{\ectxext}{\eft}}
-  &
-\stesysf
-  \ar{d}{\eft}
-  \\
-  {}&
-\stesysf
-  \ar{r}[swap]{\ectxext}
-  &
-\stesysc
-\end{tikzcd}
-\end{equation*}
-commutative, is a section of $\efamext$.
-\end{enumerate}
-\end{defn}
-
-\begin{defn}
-A homomorphism of extension objects with empty families is an extension
-homomorphism $f:\stesys'\to\stesys$ for which the diagram
-\begin{equation*}
-\begin{tikzcd}
-\stesysf'
-  \ar{r}{f_1}
-  &
-\stesysf
-  \\
-\stesysc'
-  \ar{u}{\phi_1(\stesys')}
-  \ar{r}[swap]{f_0}
-  &
-\stesysc
-  \ar{u}[swap]{\phi_1(\stesys)}
-\end{tikzcd}
-\end{equation*}
-commutes.
-\end{defn}
-
-\begin{lem}
-Suppose $\stesys$ is an extension object with empty families. Then
-$\famesys{\stesys}$ is an extension object with empty families, with
-$\phi_1(\famesys{\stesys}):=\phi_2$. 
-\end{lem}
-
-\begin{lem}
-Suppose $\stesys$ is an extension object with empty families and consider
-$p:\stesysc\rightarrow X\leftarrow Y:p$. Then
-$\cobesys{Y}{\stesys}{g}{p}$ is an extension object with empty families
-with $\phi_1(\cobesys{Y}{\stesys}{g}{p}):=g^\ast(\phi_1)$.
-\end{lem}
-
-\subsection{E-objects}
-\begin{defn}
-An \emph{E-object} is an extension object with the structure of a projection object,
-the structure of a substitution object and which has an empty context and families,
-such that additionally:
-\begin{enumerate}
-\item substitution is a projection homomorphism
-\item weakening is a substitution homomorphism
-\item both weakening and substitution are empty-CF homomorphisms.
-\item 
-\end{enumerate}
-\end{defn}

### 2015-02-10 Rules/rules_extension.tex
removed comma
large reorganisation
removed a comment

@@
 $\jterm{\Gamma}{\subst{x}{P}}{u}$. Thus, there is a formal definition of the
-\emph{pair} $\tmext{x}{u}$. We shall not go into this, in the current paper,
+\emph{pair} $\tmext{x}{u}$. We shall not go into this in the current paper,
 but it will be very useful to have when one further develops the theory of
@@
 \end{align}
-Note that we may take the point of view that extension distributes over
-extensions. If we want to extend an extended family $\ctxext{P}{Q}$ over
-$\ctxext{\Gamma}{A}$ by the family $A$, the result is the extension of (1)
-the extension of $P$ by $A$ and (2) the extension of $Q$ by $A$. The latter is
-simply $Q$ since extension acts trivially at the level of families (and terms).

### 2015-02-10 Rules/rules_setup.tex
rewrite definition of valid inference rules

@@
 \label{judgments}
+
+The theory of contexts, families and terms shall be described by means of a
+declaration of valid inference rules. An inference rule is a finite list
+$\mathcal{H}_1,\ldots,\mathcal{H}_n;\mathcal{J}$ of expressions called
+judgments. Usually we represent an inference rule in the following
+form
+\begin{equation*}
+\inference{\mathcal{H}_1\quad\cdots\quad\mathcal{H}_n}{\mathcal{J}}
+\end{equation*}
+The judgments $\mathcal{H}_1$, \ldots, $\mathcal{H}_n$ are called the
+hypotheses and the judgment $\mathcal{J}$ is the conclusion of the inference
+rule.
+
+\begin{defn}
+An inference rule is called \emph{valid} in either of the following cases:
+\begin{itemize}
+\item It is declared valid as an axiom;
+\item The inference rule
+\begin{equation*}
+\inference{\mathcal{H}_1\quad\cdots\quad\mathcal{H}_n}{\mathcal{H}_n}
+\end{equation*}
+is a valid inference rule;
+\item If
+\begin{equation*}
+\inference{\mathcal{I}_{1k}\quad\cdots\quad\mathcal{I}_{m_kk}}{\mathcal{H}_k}
+\end{equation*}
+and
+\begin{equation*}
+\inference{\mathcal{H}_1\quad\cdots\quad\mathcal{H}_n}{\mathcal{J}}
+\end{equation*}
+are valid inference rules for $1\leq k\leq n$, then
+\begin{equation*}
+\inference{\mathcal{I}_{11}\quad\cdots\quad\mathcal{I}_{m_11}\quad\cdots\quad\mathcal{I}_{1n}\quad\cdots\quad\mathcal{I}_{m_nn}}{\mathcal{J}}
+\end{equation*}
+is also a valid inference rule.
+\item If
+\begin{equation*}
+\inference{\mathcal{H}_1\quad\cdots\quad\mathcal{H}_n}{\mathcal{H}_{n+1}}
+\qquad
+\text{and}
+\qquad
+\inference{\mathcal{H}_1\quad\cdots\quad\mathcal{H}_{n+m+1}}{\mathcal{J}}
+\end{equation*}
+are valid inference rules, then
+\begin{equation*}
+\inference{\mathcal{H}_1\quad\cdots\quad\hat{\mathcal{H}}_{n+1}\quad\cdots\quad\mathcal{H}_{n+m+1}}{\mathcal{J}}
+\end{equation*}
+where the hypothesis $\mathcal{H}_{n+1}$ is omitted, is a valid inference rule.
+\end{itemize}
+\end{defn}
 
@@
 it does here.
-
-The theory of contexts, families and terms shall be described by means of a
-declaration of valid inference rules. An inference rule is an expression of the
-form
-\begin{equation*}
-\inference{\mathcal{H}_1\quad\cdots\quad\mathcal{H}_n}{\mathcal{J}}
-\end{equation*}
-where each of $\mathcal{H}_1$, ..., $\mathcal{H}_n$ and $\mathcal{J}$ is 
-judgment. The judgments $\mathcal{H}_1$, ..., $\mathcal{H}_n$ are called the
-hypotheses and the judgment $\mathcal{J}$ is the conclusion of the inference
-rule. When a judgment is the conclusion of a valid inference rule, we also say
-that the judgment is valid or well-formed. In fact, the only way to construct
-new judgments is via the valid inference rules we describe in this section.
-
-The class of valid inference rules is closed under composition and under
-rearranging the hypotheses. That
-means that if
-\begin{equation*}
-\inference{\mathcal{I}_{1k}\quad\cdots\quad\mathcal{I}_{m_kk}}{\mathcal{H}_k}
-\end{equation*}
-and
-\begin{equation*}
-\inference{\mathcal{H}_1\quad\cdots\quad\mathcal{H}_n}{\mathcal{J}}
-\end{equation*}
-are valid inference rules for $1\leq k\leq n$, then
-\begin{equation*}
-\inference{\mathcal{I}_{11}\quad\cdots\quad\mathcal{I}_{m_11}\quad\cdots\quad\mathcal{I}_{1n}\quad\cdots\quad\mathcal{I}_{m_nn}}{\mathcal{J}}
-\end{equation*}
-is also a valid inference rule. Also, in a valid inference rule 
-duplicated hypotheses may always be contracted and hypotheses may always be 
-reordered; the result is still considered a valid inference rule.
 

### 2015-02-26 Algebras/fundamental.tex
added definitions to fundamental structure; renamed section titles

@@
 \subsection{Semantics of the fundamental structure of dependent type theory}
-For the semantics of the theory of E-systems, we will assume that $\cat{C}$ is a 
-category with finite limits and
-whenever we write a pullback, we assume that it is chosen. Recall that for
-any morphism $f:A\to B$ in a category $\cat{C}$ with chosen pullbacks, there
-is a functor
-\begin{equation*}
-f^\ast : \cat{C}/B\to\cat{C}/A.
-\end{equation*}
-As usual, when $g:X\to B$ is a morphism, we will write $f^\ast(X)$ for the
-domain of $f^\ast(g)$. When there is more than one morphism $X\to B$ involved,
-as will be the case below, we will write $\pullback{A}{X}{f}{g}$. The projections
-will be written as $\pullbackpr{1}{f}{g}$ and $\pullbackpr{2}{f}{g}$. So in this notation, a
-typical pullback diagram has the following form:
-\begin{equation*}
-\begin{tikzcd}[column sep=large]
-\pullback{A}{X}{f}{g}
-  \ar{r}{\pullbackpr{1}{f}{g}}
-  \ar{d}[swap]{\pullbackpr{2}{f}{g}}
-  &
-A \ar{d}{f}
-  \\
-X \ar{r}[swap]{g}
-  &
-B
-\end{tikzcd}
-\end{equation*}
-Also, when we have a commutative diagram of the form
-\begin{equation*}
-\begin{tikzcd}
-A \ar{r}{f}
-  \ar{d}{a}
-  &
-X \ar{d}
-  & 
-B \ar{l}[swap]{g}
-  \ar{d}{b}
-  \\
-A'
-  \ar{r}[swap]{f'}
-  &
-X'
-  &
-B'
-  \ar{l}{g'}
-\end{tikzcd}
-\end{equation*}
-we will denote the unique map from $\pullback{A}{B}{f}{g}$ to $\pullback{A'}{B'}{f'}{g'}$
-such that the diagram
-\begin{equation*}
-\begin{tikzcd}
-  {}
-  & 
-\pullback{A'}{B'}{f'}{g'}
-  \ar{dd}
-  \ar{rr}
-  &
-  &
-B'
-  \ar{dd}{g'}
-  \\
-\pullback{A}{B}{f}{g}
-  \ar{dd}
-  \ar[crossing over]{rr}
-  \ar[dotted]{ur}{\pullback{a}{b}{f'}{g'}}
-  &
-  &
-B \ar{ur}{b}
-  \\
-  {}
-  &
-A'
-  \ar{rr}
-  &
-  &
-X'
-  \\
-A \ar{rr}[swap]{f}
-  \ar{ur}{a}
-  &
-  &
-X \ar[crossing over,leftarrow]{uu}[near end,swap]{g}
-  \ar{ur}
-\end{tikzcd}
-\end{equation*}
-commutes, by $\pullback{a}{b}{f'}{g'}$. In the current work, we shall
-write $A\times B$ for the pullback of $A\rightarrow 1\leftarrow B$, and
-$\pi_1$ and $\pi_2$ for its projections (thus, no separate choice of
-cartesian products is made).
 
@@
 \end{equation*}
-in $\cat{C}$. In this diagram, $C$ is the object of \emph{contexts}, $F$ is
-the object of \emph{families}, and $T$ is the object of \emph{terms}.
+in $\cat{C}$. In this diagram, $\stesysc$ is the object of \emph{contexts}, $\stesysf$ is
+the object of \emph{families}, and $\stesyst$ is the object of \emph{terms}.
 \end{defn}
+
+\begin{defn}
+A \emph{context of $\stesys$ under an object $X$ of $\cat{C}$} is a morphism $\Gamma:X\to \stesysc$ in
+$\cat{C}$. A \emph{family over $\Gamma$ under $X$} is a morphism
+$A:X\to (\pullback{X}{\stesysf}{\Gamma}{\eft})$. A 
+\emph{term of the family $A$ over $\Gamma$ under $X$}
+is a morphism $x:X\to(\pullback{X}{\stesyst}{\Gamma}{\ebd\circ\eft})$.
+
+Contexts, families and terms of $\stesys$ under the terminal object $\mathbf{1}$
+are called \emph{contexts, families and terms of $\stesys$}, respectively.
+\end{defn}
+
+\begin{lem}
+For any object $X$ of $\stesys$, contexts, families and terms of $\stesys$ under
+$X$ satisfy all the judgmental equality rules stated in \autoref{judgments}.
+\end{lem}
+
+\begin{proof}
+This follows immediately from the fact that equality on the hom-sets are
+equivalence relations.
+\end{proof}
+
+\begin{defn}
+Let $\stesys'$ and $\stesys$ be fundamental structures in $\cat{C}$. A
+homomorphism $f:\stesys'\to\stesys$ is a triple $(f_0,f_1,f^t)$ 
+consisting of morphisms
+\begin{equation*}
+\begin{tikzcd}
+\stesyst' 
+  \ar{r}{f^t}
+  \ar{d}[swap]{\ebd'}
+  &
+\stesyst
+  \ar{d}{\ebd}
+  \\
+\stesysf'
+  \ar{r}{f_1}
+  \ar{d}[swap]{\eft'}
+  &
+\stesysf
+  \ar{d}{\eft}
+  \\
+\stesysc' 
+  \ar{r}[swap]{f_0}
+  &
+\stesysc
+\end{tikzcd}
+\end{equation*}
+such that the indicated squares commute.
+\end{defn}
+
+\begin{defn}
+Suppose $\stesys$ is a fundamental structure in $\cat{C}$ and that 
+$p:\stesysc\rightarrow X\leftarrow Y:g$.
+Then we define the fundamental structure $\cobesys{Y}{\stesys}{g}{p}$ to consist
+of the part of the left column above $Y$ in the diagram
+\begin{equation*}
+\begin{tikzcd}
+\cobesys{Y}{\stesyst}{g}{p\circ\eft\circ\ebd}
+  \ar{r}
+  \ar{d}[swap]{g^\ast(\ebd)}
+  &
+\stesyst
+  \ar{d}{\ebd}
+  \\
+\cobesys{Y}{\stesysf}{g}{p\circ\eft}
+  \ar{r}
+  \ar{d}[swap]{g^\ast(\eft)}
+  &
+\stesysf
+  \ar{d}{\eft}
+  \\
+\cobesys{Y}{\stesysc}{g}{p}
+  \ar{r}
+  \ar{d}[swap]{\pullbackpr{1}{g}{p}}
+  &
+\stesysc
+  \ar{d}{p}
+  \\
+Y \ar{r}[swap]{g}
+  &
+X
+\end{tikzcd}
+\end{equation*}
+The horizontal arrows form a homomorphism
+of fundamental structures.
+\end{defn}

### 2015-02-26 Rules/rules_setup.tex
added definitions to fundamental structure; renamed section titles

--- 2014-10-28 Rules/rules_setup.tex
@@
-\subsection{Judgments and valid inference rules}
+\subsection{The judgments of dependent type theory}
 \label{judgments}

### 2015-02-26 main.tex
large reorganisation
added definitions to fundamental structure; renamed section titles

@@
 
-\input{Rules/rules}
+\section{Introduction}
+In this article we define the essentially algebraic objects which correspond to dependent
+type theories, which are closely related to Voevodsky's B-systems. This provides
+a new axiomatization of dependent type theory, with the following properties:
+\begin{itemize}
+\item The meta-theory should not require anything more than rules for handling
+inferences. In particular, natural numbers, or an infinite set of variables 
+should not be required.
+\item Finitely many sorts. The basic ingredients are contexts, families and
+terms. Thus there will be 3 sorts.
+\item The theory is algebraic and the operations on it are homomorphisms. From
+the requirement that type theoretic operations must be homomorphisms, we can
+read off which judgmental equalities to require require.
+\item Finite set of rules.
+\item The theory is invariant under slicing in a canonical way.
+\end{itemize}
+
+We build
+dependent type theory by imposing several structures of increasing complexities
+on top of each other. We begin by the theory of extension, which incorporates
+a context extension and a family extension operation to the theory which merely
+has three sorts. On top of the theory of extension we can formulate the theories
+of weakening, substitution and extension-units (i.e.~the empty context and families).
+The theory of weakening can be enriched with units playing the role of identity
+morphisms, which gives a theory of projections. The three resulting theories
+can then be combined into dependent type theory. Thus, the theory of dependent
+types is outlined in the following diagram.
+
+\begin{equation*}
+\begin{tikzcd}[ampersand replacement=\&,column sep=0em]
+{} \& \textbf{Extension} \ar[very thick,-stealth]{dr} \ar[very thick,-stealth]{d} \ar[very thick,-stealth]{dl}
+  \\
+\makebox[.7cm]{\textbf{Weakening}} \ar[very thick,-stealth]{d} \& \textbf{Substitution} \ar[very thick,-stealth]{dd} \& \makebox[3cm]{\textbf{Empty context and families}} \ar[very thick,-stealth]{ddl}
+\\
+\makebox[.7cm]{\textbf{Projection}} \ar[very thick,-stealth]{dr}
+\\
+{} \& \textbf{Dependent Type Theory}
+\end{tikzcd}
+\end{equation*}
+
+We give two presentations of the theory simultaneously. First, we directly
+define the algebras for the corresponding theories in an arbitrary category
+with finite limits and second, we describe the theory in terms of inference
+rules.
+
+For the semantics of the theory of E-systems, we will assume that $\cat{C}$ is a 
+category with finite limits and
+whenever we write a pullback, we assume that it is chosen. Recall that for
+any morphism $f:A\to B$ in a category $\cat{C}$ with chosen pullbacks, there
+is a functor
+\begin{equation*}
+f^\ast : \cat{C}/B\to\cat{C}/A.
+\end{equation*}
+As usual, when $g:X\to B$ is a morphism, we will write $f^\ast(X)$ for the
+domain of $f^\ast(g)$. When there is more than one morphism $X\to B$ involved,
+as will be the case below, we will write $\pullback{A}{X}{f}{g}$. The projections
+will be written as $\pullbackpr{1}{f}{g}$ and $\pullbackpr{2}{f}{g}$. So in this notation, a
+typical pullback diagram has the following form:
+\begin{equation*}
+\begin{tikzcd}[column sep=large]
+\pullback{A}{X}{f}{g}
+  \ar{r}{\pullbackpr{1}{f}{g}}
+  \ar{d}[swap]{\pullbackpr{2}{f}{g}}
+  &
+A \ar{d}{f}
+  \\
+X \ar{r}[swap]{g}
+  &
+B
+\end{tikzcd}
+\end{equation*}
+Also, when we have a commutative diagram of the form
+\begin{equation*}
+\begin{tikzcd}
+A \ar{r}{f}
+  \ar{d}{a}
+  &
+X \ar{d}
+  & 
+B \ar{l}[swap]{g}
+  \ar{d}{b}
+  \\
+A'
+  \ar{r}[swap]{f'}
+  &
+X'
+  &
+B'
+  \ar{l}{g'}
+\end{tikzcd}
+\end{equation*}
+we will denote the unique map from $\pullback{A}{B}{f}{g}$ to $\pullback{A'}{B'}{f'}{g'}$
+such that the diagram
+\begin{equation*}
+\begin{tikzcd}
+  {}
+  & 
+\pullback{A'}{B'}{f'}{g'}
+  \ar{dd}
+  \ar{rr}
+  &
+  &
+B'
+  \ar{dd}{g'}
+  \\
+\pullback{A}{B}{f}{g}
+  \ar{dd}
+  \ar[crossing over]{rr}
+  \ar[dotted]{ur}{\pullback{a}{b}{f'}{g'}}
+  &
+  &
+B \ar{ur}{b}
+  \\
+  {}
+  &
+A'
+  \ar{rr}
+  &
+  &
+X'
+  \\
+A \ar{rr}[swap]{f}
+  \ar{ur}{a}
+  &
+  &
+X \ar[crossing over,leftarrow]{uu}[near end,swap]{g}
+  \ar{ur}
+\end{tikzcd}
+\end{equation*}
+commutes, by $\pullback{a}{b}{f'}{g'}$. In the current work, we shall
+write $A\times B$ for the pullback of $A\rightarrow 1\leftarrow B$, and
+$\pi_1$ and $\pi_2$ for its projections (thus, no separate choice of
+cartesian products is made).
+
+The formal theory of contexts, families and terms shall be described by means of a
+declaration of valid inference rules. Informally, an inference rule is a finite list
+$\mathcal{H}_1,\ldots,\mathcal{H}_n;\mathcal{J}$ of expressions called
+judgments. Usually we represent an inference rule in the following
+form
+\begin{equation*}
+\inference{\mathcal{H}_1\quad\cdots\quad\mathcal{H}_n}{\mathcal{J}}
+\end{equation*}
+The judgments $\mathcal{H}_1$, \ldots, $\mathcal{H}_n$ are called the
+hypotheses and the judgment $\mathcal{J}$ is the conclusion of the inference
+rule.
+
+An inference rule is called \emph{valid} in either of the following cases:
+\begin{description}
+\item[(Axiom)] It is declared valid as an axiom;
+\item[(Projection)] The inference rule
+\begin{equation*}
+\inference{\mathcal{H}_1\quad\cdots\quad\mathcal{H}_n}{\mathcal{H}_i}
+\end{equation*}
+is a valid inference rule for any $1\leq i\leq n$;
+\item[(Composition)] If
+\begin{equation*}
+\inference{\mathcal{I}_{1k}\quad\cdots\quad\mathcal{I}_{m_kk}}{\mathcal{H}_k}
+\qquad\text{and}\qquad
+\inference{\mathcal{H}_1\quad\cdots\quad\mathcal{H}_n}{\mathcal{J}}
+\end{equation*}
+are valid inference rules for $1\leq k\leq n$, then
+\begin{equation*}
+\inference{\mathcal{I}_{11}\quad\cdots\quad\mathcal{I}_{m_11}\quad\cdots\quad\mathcal{I}_{1n}\quad\cdots\quad\mathcal{I}_{m_nn}}{\mathcal{J}}
+\end{equation*}
+is also a valid inference rule;
+\item[(Cut elimination)] If
+\begin{equation*}
+\inference{\mathcal{H}_1\quad\cdots\quad\mathcal{H}_n}{\mathcal{H}_{n+1}}
+\qquad
+\text{and}
+\qquad
+\inference{\mathcal{H}_1\quad\cdots\quad\mathcal{H}_{n+m+1}}{\mathcal{J}}
+\end{equation*}
+are valid inference rules, then
+\begin{equation*}
+\inference{\mathcal{H}_1\quad\cdots\quad\hat{\mathcal{H}}_{n+1}\quad\cdots\quad\mathcal{H}_{n+m+1}}{\mathcal{J}}
+\end{equation*}
+where the hypothesis $\mathcal{H}_{n+1}$ is omitted, is a valid inference rule.
+\end{description}
+In this article, we shall restrict ourselves to formulating the axioms for
+dependent type theory. Derivations shall be made in a follow-up.
+
+\input{fundamental_structure}
+
+\input{extension}
+
+\input{weakening}
+
+\input{projections}
+
+\input{substitution}
+
+\input{empty}
+
+\input{dependent_type_theory}
+
+\section{The relation between E-systems and B-systems}
+
+In this section we answer the following questions posed by Voevodsky in private
+correspondence.
+\begin{itemize}
+\item Is the construction that takes a (unital) B-system to an E-system a functor?
+\item Is this functor fully faithful?
+\item Is there a simple way to characterize the E-systems that arise from B-systems (the image of this functor on objects)?
+\item Construct an example of an E-system that does not arise from a B-system.
+\end{itemize}
+
+\subsection{The functor from B-systems to E-systems}
+Consider a B-system $\mathbb{B}$. We will construct an E-system $E(\mathbb{B})$
+in the category of sets. This E-systems consists of sets $C(\mathbb{B})$,
+$F(\mathbb{B})$ and $T(\mathbb{B})$ for the sorts of contexts, families and
+terms respectively, with functions $\eft:F(\mathbb{B})\to C(\mathbb{B})$ and
+$\ebd:T(\mathbb{B})\to T(\mathbb{B})$ and the operations of extension,
+weakening, units, substitution and empty contexts and families.
+
+The set $C(\mathbb{B})$ consists of pairs $(n,X)$, where $n\in\mathbb{N}$ and
+$X\in B_n$. The set $F(\mathbb{B})$ consists of quadruples $(n,X,m,Y)$, where
+$n,m\in\mathbb{N}$ such that $n<m$, $X\in B_n$ and $Y\in B_{m}$ such that
+$ft^{m-n}:=ft^{n-1}\circ\cdots\circ ft^m(Y)=X$.
 
@@
 
-\input{Algebras/article-models-cat}
+%\input{Algebras/article-models-cat}
 

### 2015-02-27 Algebras/fundamental.tex
fixed the proof of Lemma 3.18

@@
 \begin{lem}
-For any object $X$ of $\stesys$, contexts, families and terms of $\stesys$ under
-$X$ satisfy all the judgmental equality rules stated in \autoref{judgments}.
+For any object $X$ of $\cat{C}$ and any $\stesys$ in $\cat{C}$, contexts,
+families and terms of $\stesys$ under $X$ satisfy all the judgmental equality 
+rules stated in \autoref{judgments}.
 \end{lem}
@@
 Let $\stesys'$ and $\stesys$ be fundamental structures in $\cat{C}$. A
-homomorphism $f:\stesys'\to\stesys$ is a triple $(f_0,f_1,f^t)$ 
-consisting of morphisms
+\emph{fundamental homomorphism} $f:\stesys'\to\stesys$ is a triple 
+$(f_0,f_1,f_t)$ consisting of morphisms
 \begin{equation*}
@@
 \stesyst' 
-  \ar{r}{f^t}
+  \ar{r}{f_t}
   \ar{d}[swap]{\ebd'}
@@
 
+In other words, the category $\mathbf{Fund}(\cat{C})$ of fundamental structures
+is the functor category $(0\leftarrow 1\leftarrow t)\to\cat{C}$.
+
+The definitions of weakening algebra, projection algebra and substitution
+algebra (and ultimately the definition of e-systems) will require the
+change of base construction. Proving that the definitions are stable under
+change of base is therefore a crucial aspect of the development of the theory.
+The change of base of fundamental structures is the simplest case.
+
 \begin{defn}
@@
 $p:\stesysc\rightarrow X\leftarrow Y:g$.
-Then we define the fundamental structure $\cobesys{Y}{\stesys}{g}{p}$ to consist
-of the part of the left column above $Y$ in the diagram
+Then we define the \emph{change of base $\cobesys{Y}{\stesys}{g}{p}$ of 
+$\stesys$ along $g:Y\to X$} to be the fundamental structure above $Y$ in the 
+diagram
 \begin{equation*}

### 2015-02-27 main.tex
fixed the proof of Lemma 3.18

@@
 
+\subsection{Categories with finite limits}
 For the semantics of the theory of E-systems, we will assume that $\cat{C}$ is a 
@@
 
-The formal theory of contexts, families and terms shall be described by means of a
-declaration of valid inference rules. Informally, an inference rule is a finite list
-$\mathcal{H}_1,\ldots,\mathcal{H}_n;\mathcal{J}$ of expressions called
-judgments. Usually we represent an inference rule in the following
-form
+\subsection{Inference rules}
+The syntactic theory of contexts, families and terms shall be described by means
+of a declaration of valid inference rules. Although their appearance suggests 
+otherwise, the notion of inference rule is ultimately informal, in the sense
+that we state the explanations and possibilities with inference rules in
+natural language. One may attempt to study systems of inference rules within
+other systems to study the applications of the various concepts involved, but
+this study will not make the basic notion of inference rules more formal. We do
+not see this as a problem, as systems of inference rules have been implemented
+in proof assistant software in which one essentially only manipulates
+judgments. In this paper we use inference rules as an alternative presentation
+of the theory, which some readers may prefer.
+
+An inference rule is a finite list $\mathcal{H}_1,\ldots,\mathcal{H}_n;
+\mathcal{J}$ of expressions called judgments. Usually we represent an inference
+rule in the following form
 \begin{equation*}
@@
 \end{description}
-In this article, we shall restrict ourselves to formulating the axioms for
-dependent type theory. Derivations shall be made in a follow-up.
+In categories with finite limits, valid inference rules will always be 
+interpreted by  arrows.
 

### 2015-03-03 Algebras/fundamental.tex
split in the section on extension algebras

@@
 
+The following definition provides us with a useful abbreviation to display
+diagrams in which both fundamental structures and arbitrary objects are 
+involved.
+
+\begin{defn}
+Suppose $f:\stesys\to\stesys'$ is a fundamental homomorphism. We say that
+a diagram
+\begin{equation*}
+\begin{tikzcd}
+\stesys
+  \ar{r}{f}
+  \ar{d}[swap]{p}
+  &
+\stesys'
+  \ar{d}{p'}
+  \\
+X \ar{r}[swap]{g}
+  &
+Y
+\end{tikzcd}
+\end{equation*}
+commutes if the diagram
+\begin{equation*}
+\begin{tikzcd}
+\stesysc
+  \ar{r}{f_0}
+  \ar{d}[swap]{p}
+  &
+\stesysc'
+  \ar{d}{p'}
+  \\
+X \ar{r}[swap]{g}
+  &
+Y
+\end{tikzcd}
+\end{equation*}
+commutes.
+\end{defn}
+
+Thus, our goal of defining the change of base for a fundamental structure
+can be restated in the following way: given a fundamental structure 
+$\stesys$ and arrows $p:\stesysc\rightarrow X\leftarrow Y:g$, we want to
+define a fundamental structure $\cobesys{Y}{\stesys}{g}{p}$ with a 
+homomorphism $\pullbackpr{2}{g}{p}: \cobesys{Y}{\stesys}{g}{p}\to\stesys$ and a 
+morphism $\pullbackpr{1}{g}{p}: \pullback{Y}{\stesysc}{g}{p}\to Y$ such that for
+every diagram
+\begin{equation*}
+\begin{tikzcd}[column sep=large]
+\stesys'
+  \ar[bend right=10]{ddr}[swap]{p'}
+  \ar[bend left=10]{rrd}{f}
+  \ar[dotted]{dr}[near end]{[p',f]}
+  \\
+  {}&
+\cobesys{Y}{\stesys}{g}{p}
+  \ar{d}{\pullbackpr{1}{g}{p}}
+  \ar{r}[swap]{\pullbackpr{2}{g}{p}}
+  &
+\stesys
+  \ar{d}{p}
+  \\
+  {}&
+Y \ar{r}[swap]{g}
+  &
+X
+\end{tikzcd}
+\end{equation*}
+of which the outer square commutes, the fundamental homomorphism $[p',f]$ 
+exists and is unique with the property that it renders the diagram commutative.
+
 \begin{defn}
@@
 \begin{equation*}
-\begin{tikzcd}
+\begin{tikzcd}[column sep=7em]
 \cobesys{Y}{\stesyst}{g}{p\circ\eft\circ\ebd}
-  \ar{r}
+  \ar{r}{\pullbackpr{2}{g}{p\circ\eft\circ\ebd}}
   \ar{d}[swap]{g^\ast(\ebd)}
@@
 \cobesys{Y}{\stesysf}{g}{p\circ\eft}
-  \ar{r}
+  \ar{r}{\pullbackpr{2}{g}{p\circ\eft}}
   \ar{d}[swap]{g^\ast(\eft)}
@@
 \cobesys{Y}{\stesysc}{g}{p}
-  \ar{r}
+  \ar{r}{\pullbackpr{2}{g}{p}}
   \ar{d}[swap]{\pullbackpr{1}{g}{p}}
@@
 \end{equation*}
-The horizontal arrows form a homomorphism
-of fundamental structures.
+The horizontal arrows form a homomorphism of fundamental structures, we will
+denote it by $\pullbackpr{2}{g}{p}:\cobesys{Y}{\stesys}{g}{p}\to\stesys$.
 \end{defn}

### 2015-03-03 main.tex
split in the section on extension algebras

@@
 \end{equation*}
-Also, when we have a commutative diagram of the form
+Furthermore, given $h:Y\to A$ and $k:Y\to B$ such that $h\circ f=g\circ k$, we
+denote the arrow $Y\to\pullback{A}{X}{f}{g}$ given by the universal property of
+pullbacks, by $[h,k]$. Also, when we have a commutative diagram of the form
 \begin{equation*}

### 2015-03-12 weakeningcats.tex
The new E-systems

\section{Weakening systems}

\begin{defn}
A \emph{pre-weakening structure on a category $\cat{C}$} consists of a functor
$W_f:\cat{C}/Y\to \cat{C}/X$ for every $f:X\to Y$ in $\cat{C}$ such 
that
\begin{enumerate}
\item $W_{\catid{X}}=\catid{\cat{C}/X}$ for every object $X\in\cat{C}$.
\item $W_{f\circ f'}=W_{f'}\circ W_f$ for every $f':X'\to X$ and $f:X\to Y$ in
$\cat{C}$.
\item $W_f$ preserves the final object, i.e.~$W_f(\catid{Y})=\catid{X}$.
%\item $W^1_f\circ W^0_g= W^0_{W^1_f(g)}\circ W^1_f$.
\end{enumerate}
A \emph{pre-weakening system} is a category together with a pre-weakening structure.
\end{defn}

\begin{defn}
A \emph{pre-weakening homomorphism} $F:\cat{C}\to\cat{D}$ between pre-weakening systems
is a functor $F:\cat{C}\to\cat{D}$ such that the square
\begin{equation*}
\begin{tikzcd}
\cat{C}/X
  \arrow[r,"F/X"]
  &
\cat{D}/F(X)
  \\
\cat{C}/Y
  \arrow[u,"W_f"]
  \arrow[r,swap,"F/Y"]
  &
\cat{D}/F(Y)
  \arrow[u,swap,"W_{F(f)}"]
\end{tikzcd}
\end{equation*}
commutes for any $f:X\to Y$ in $\cat{C}$.
\end{defn}

\begin{lem}
Let $\cat{C}$ be a pre-weakening system. Then for any object $X$ of $\cat{C}$,
the slice category $\cat{C}/X$ is also a pre-weakening system.
\end{lem}

\begin{defn}
A \emph{weakening system} is a pre-weakening system for which every $W_f$ is a
pre-weakening homomorphism. A \emph{weakening homomorphism} is a pre-weakening 
homomorphism between weakening systems.
\end{defn}

\begin{cor}
For any object $X$ of a weakening system $\cat{C}$, the slice category
$\cat{C}/X$ is a weakening system.
\end{cor}

\begin{rmk}
If the underlying category is a group, the axioms of a weakening system
obtain the following form: There is a group homomorphism $W_x:G\to G$ for
every $x\in G$, in such a way that $W_e=\catid{G}$ and $W_{xy}=W_y\circ W_x$
for every $x,y\in G$, and for any $x,y\in G$ the diagram
\begin{equation*}
\begin{tikzcd}
G \arrow[r,"W_y"]
  &
G 
  \\
G \arrow[u,"W_x"]
  \arrow[r,swap,"W_y"]
  &
G \arrow[u,swap,"W_{W_y(x)}"]
\end{tikzcd}
\end{equation*}
is required to commute.
\end{rmk}

\begin{lem}
For any group, conjugation is a weakening structure. Thus, any group is
canonically a weakening system.
\end{lem}

\begin{proof}
Note that $e^{-1}xe=x$ and $(xy)^{-1}z(xy)=y^{-1}(x^{-1}zx)y$ for any $x,y,z\in G$.
Also, the diagram
\begin{equation*}
\begin{tikzcd}
G \arrow[r,"\mathrm{conj}_y"]
  &
G 
  \\
G \arrow[u,"\mathrm{conj}_x"]
  \arrow[r,swap,"\mathrm{conj}_y"]
  &
G \arrow[u,swap,"\mathrm{conj}_{y^{-1}xy}"]
\end{tikzcd}
\end{equation*}
commutes.
\end{proof}

\begin{rmk}
Similarly, every groupoid has a canonical weakening structure given by
conjugation.
\end{rmk}

\section{Weakening systems with term structure}
\begin{defn}
A \emph{category with a term structure} is a category $\cat{C}$ with a set $T(f)$ for every
morphism $f$ of $\cat{C}$. A \emph{functor with term structure} from $\cat{C}$ to $\cat{D}$
is a functor $F:\cat{C}\to\cat{D}$ with a function $T(f)\to T(F(f))$ for
every $f:X\to Y$ in $\cat{C}$.
\end{defn}

\begin{defn}
Suppose $\cat{C}$ is a category with term structure, and $X$ is an object of $\cat{C}$. 
Then $\cat{C}/X$ is a category with term structure, where $T_{\cat{C}/X}(g)=T_\cat{C}(g)$.
\end{defn}

\begin{defn}
A \emph{weakening system with term term structure} is the same thing as a weakening system,
except that the underlying category and the functors $W_f$ are required to
be functors with term term structure. The equalities are also required to hold on the term structure.
\end{defn}

\begin{cor}
For any object $X$ of a weakening system with term structure, the slice category
$\cat{C}/X$ is a weakening system with term structure.
\end{cor}

\section{Projection systems}
\begin{defn}
A \emph{pre-projection system} is a weakening system with term structure with an element
$\tfid{f}\in T(W_f(f))$ for every $f:X\to Y$ in $\cat{C}$. 
\end{defn}

\begin{defn}
A \emph{pre-projection homomorphism} $F:\cat{C}\to\cat{D}$ is a weakening homomorphism for which
\begin{equation*}
F(\tfid{f})=\tfid{F(f)}
\end{equation*}
for every $f:X\to Y$ in $\cat{C}$.
\end{defn}

\begin{lem}
For any object $X$ in a pre-projection system $\cat{C}$, the slice category
$\cat{C}/X$ is a pre-projection system.
\end{lem}

\begin{defn}
A \emph{projection system} is a pre-projection system for which every $W_f$ is
a pre-projection homomorphism. A \emph{projection homomorphism} is a pre-projection homomorphism
between projection systems.
\end{defn}

\begin{cor}
For any object $X$ of a projection system $\cat{C}$, the slice category $\cat{C}/X$
is a projection system.
\end{cor}

\section{Substitution systems}

\begin{defn}
A \emph{pre-substitution structure} on a category with term structure $\cat{C}$ consists
of a functor with term structure $S_t:\cat{C}/X\to\cat{C}/Y$ for every $t\in T(f)$ and
$f:X\to Y$, such that $S_t(\catid{X})=\catid{Y}$.

A \emph{pre-substitution system} is a category with term structure together with a
pre-substitution structure. 
\end{defn}

\begin{defn}
A \emph{pre-substitution homomorphism} $F:\cat{C}\to\cat{D}$ is a functor with term structure for
which the diagram
\begin{equation*}
\begin{tikzcd}[column sep=large]
\cat{C}/X
  \arrow[r,"F/X"]
  \arrow[d,swap,"S_t"]
  &
\cat{D}/F(X)
  \arrow[d,"S_{F(t)}"]
  \\
\cat{C}/Y
  \arrow[r,swap,"F/Y"]
  &
\cat{D}/F(Y)
\end{tikzcd}
\end{equation*}
commutes.
\end{defn}

\begin{lem}
For any object $X$ of a pre-substitution system $\cat{C}$, the slice category
$\cat{C}/X$ is a pre-substitution system.
\end{lem}

\begin{defn}
A \emph{substitution system} is a pre-substitution system for which each
$S_t$ is a pre-substitution homomorphism. A \emph{substitution homomorphism} is a
pre-substitution homomorphism between substitution systems.
\end{defn}

\begin{cor}
For any object $X$ of a substitution system, the slice category $\cat{C}/X$
is a substitution system.
\end{cor}

\section{E-systems}
\begin{defn}
An \emph{E-system} is a category with term structure with
\begin{itemize}
\item a weakening structure $W$,
\item the structure $\tfid{}$ of a projection system,
\item a substitution structure $S$,
\end{itemize}
such that
\begin{enumerate}
\item each $W_f$ is a substitution homomorphism;
\item each $S_t$ is a projection homomorphism;
\item $S_t(\tfid{f})=t$ for any $t\in T(f)$ and $f:X\to Y$;
\item $S_t\circ W_f=\catid{\cat{C}/Y}$ for any $t\in T(f)$ and $f:X\to Y$;
\item $S_{\tfid{f}}\circ W_{W_f(f)}=\catid{\cat{C}/X}$ for any $f:X\to Y$.
\end{enumerate}
\end{defn}

\begin{cor}
For any object $X$ of an E-system $\cat{C}$, the slice category $\cat{C}/X$
is an E-system.
\end{cor}

\begin{defn}
An \emph{E-homomorphism} $F:\cat{C}\to\cat{D}$ is a functor which is both a projection
homomorphism and a substitution homomorphism.
\end{defn}

\begin{comment}
\begin{lem}
Suppose $\cat{D}$ is a retract of a category $\cat{C}$, i.e. there are functors
$S:\cat{D}\to\cat{C}$ and $R:\cat{C}\to\cat{D}$ with a natural isomorphism
$\tau:R\circ S \simeq \catid{\cat{D}}$, and suppose that $\cat{C}$ is a weakening
category. Then $\cat{D}$ is a weakening system too.
\end{lem}

\begin{proof}
Let $f:X\to Y$ be a morphism in $\cat{D}$. Then we define $W_f$ to be the
composition
\begin{equation*}
\begin{tikzcd}[column sep=3.2em]
\cat{D}/Y
  \arrow[r,"S/Y"]
  &
\cat{C}/S(Y)
  \arrow[r,"W_{S(f)}"]
  &
\cat{C}/S(X)
  \arrow[r,"R/S(X)"]
  &
\cat{D}/R(S(X))
  \arrow[r,"\tau_X\circ{-}"]
  &
\cat{D}/X
\end{tikzcd}
\end{equation*}
We first verify whether this assembles a pre-weakening structure on $\cat{D}$.
\begin{enumerate}
\item Let $X$ be an object of $\cat{D}$. Then $W_{\catid{X}}$ is the composition
\begin{equation*}
\begin{tikzcd}[column sep=3.2em]
\cat{D}/X
  \arrow[r,"S/X"]
  &
\cat{C}/S(X)
  \arrow[r,"R/S(X)"]
  &
\cat{D}/R(S(X))
  \arrow[r,"{-}\circ\tau_X"]
  &
\cat{D}/X
\end{tikzcd}
\end{equation*}
This functor maps $g:X'\to X$ consecutively to $S(g)$, to $R(S(g))$, to
$\tau_X\circ R(S(g))$. Note that $\tau_X\circ R(S(g))=g\circ \tau_{X'}$ by
naturality.
\end{enumerate}
\end{proof}
\end{comment}

### 2015-04-24 weakeningcats.tex
Theorems about E-systems

@@
+\begin{rmk}
+Recall that the slice pre-category $\cat{C}/\Gamma/(\Delta,f)$ is isomorphic to the slice pre-category
+$\cat{C}/\Delta$, in the essentially algebraic theory of pre-categories. In the following,
+we shall usually identify these two categories. Recall also that for every functor
+$F:\cat{C}\to\cat{D}$ and every object $\Delta$ of $\cat{C}$, there is a functor
+$F/\Delta:\cat{C}/\Delta\to \cat{D}/F(\Delta)$ given by the action on morphisms of $F$. 
+\end{rmk}
+
+\begin{rmk}
+It will be convenient to denote the objects of the category $\cat{C}$ with capital
+greek letters $\Gamma$, $\Delta$, \ldots, and the morphisms by capital latin letters
+$A$, $B$, $P$, $Q$, etcetera. When $A$ is a morphism into $\Gamma$, we will denote
+its codomain by $\ctxext{\Gamma}{A}$. When $A$ is a morphism into $\Gamma$ and
+$P$ is a morphism into $\ctxext{\Gamma}{A}$, it will be convenient to denote the
+composition of $P$ with $A$ by $\ctxext{A}{P}$ as well. Note that $\ctxext{A}{P}$
+is indeed the domain of $P$ in the slice category $\cat{C}/\Gamma$. Also, objects
+of $\cat{C}$ will always have a capital greek letter in their notation, e.g.~as
+in $\ctxext{{\Gamma}{A}}{P}$. 
+\end{rmk}
+
 \section{Weakening systems}
@@
 \begin{defn}
-A \emph{pre-weakening structure on a category $\cat{C}$} consists of a functor
-$W_f:\cat{C}/Y\to \cat{C}/X$ for every $f:X\to Y$ in $\cat{C}$ such 
+A \emph{pre-weakening structure on a pre-category $\cat{C}$} consists of a functor
+$W_A:\cat{C}/\Gamma\to \cat{C}/\ctxext{\Gamma}{A}$ for every $A:\ctxext{\Gamma}{A}\to \Gamma$ in $\cat{C}$ such 
 that
 \begin{enumerate}
-\item $W_{\catid{X}}=\catid{\cat{C}/X}$ for every object $X\in\cat{C}$.
-\item $W_{f\circ f'}=W_{f'}\circ W_f$ for every $f':X'\to X$ and $f:X\to Y$ in
-$\cat{C}$.
-\item $W_f$ preserves the final object, i.e.~$W_f(\catid{Y})=\catid{X}$.
-%\item $W^1_f\circ W^0_g= W^0_{W^1_f(g)}\circ W^1_f$.
+\item $W_{\catid{\Gamma}}=\catid{\cat{C}/\Gamma}$ for every object $\Gamma\in\cat{C}$.
+\item $W_{\ctxext{A}{P}}=W_P\circ W_A$ for every $P\in\cat{C}/\ctxext{\Gamma}{A}$ and $A\in\cat{C}/\Gamma$.
+\item $W_A$ preserves the final object, i.e.~$W_A(\catid{\Gamma})=\catid{\ctxext{\Gamma}{A}}$.
 \end{enumerate}
-A \emph{pre-weakening system} is a category together with a pre-weakening structure.
+A \emph{pre-weakening system} is a pre-category together with a pre-weakening structure.
 \end{defn}
@@
 \begin{equation*}
-\begin{tikzcd}
-\cat{C}/X
-  \arrow[r,"F/X"]
+\begin{tikzcd}[column sep=large]
+\cat{C}/\ctxext{\Gamma}{A}
+  \arrow[r,"F/\ctxext{\Gamma}{A}"]
   &
-\cat{D}/F(X)
+\cat{D}/F(\ctxext{\Gamma}{A})
   \\
-\cat{C}/Y
-  \arrow[u,"W_f"]
-  \arrow[r,swap,"F/Y"]
+\cat{C}/\Gamma
+  \arrow[u,"W_A"]
+  \arrow[r,swap,"F/\Gamma"]
   &
-\cat{D}/F(Y)
-  \arrow[u,swap,"W_{F(f)}"]
+\cat{D}/F(\Gamma)
+  \arrow[u,swap,"W_{F(A)}"]
 \end{tikzcd}
 \end{equation*}
-commutes for any $f:X\to Y$ in $\cat{C}$.
+commutes for any $A\in\cat{C}/\Gamma$.
 \end{defn}
@@
 \begin{lem}
-Let $\cat{C}$ be a pre-weakening system. Then for any object $X$ of $\cat{C}$,
-the slice category $\cat{C}/X$ is also a pre-weakening system.
+Let $\cat{C}$ be a pre-weakening system. Then for any object $\Gamma$ of $\cat{C}$,
+the slice pre-category $\cat{C}/\Gamma$ is also a pre-weakening system, with the weakening
+structure given by $W(\cat{C}/\Gamma)_P=W(\cat{C})_P$ for each $P\in\cat{C}/\ctxext{\Gamma}{A}$
+and each $A\in\cat{C}/\Gamma$.
 \end{lem}
@@
 
+\begin{rmk}
+In other words, a weakening system is a pre-weakening system with a weakening
+structure $W$ for which the square
+\begin{equation*}
+\begin{tikzcd}[column sep=huge]
+\cat{C}/\ctxext{{\Gamma}{B}}{Q}
+  \arrow[r,"W_A/\ctxext{B}{Q}"]
+  &
+\cat{C}/\ctxext{{\Gamma}{A}}{W_A(\ctxext{B}{Q})}
+  \\
+\cat{C}/\ctxext{\Gamma}{B}
+  \arrow[u,"W_Q"]
+  \arrow[r,swap,"W_A/B"]
+  &
+\cat{C}/\ctxext{{\Gamma}{A}}{W_A(B)}
+  \arrow[u,swap,"W_{W_A(Q)}"]
+\end{tikzcd}
+\end{equation*}
+for each $A,B\in\cat{C}/\Gamma$ and $Q:\cat{C}/\ctxext{\Gamma}{B}$. The objects part of this 
+property asserts that for any $k\in \cat{C}/\greek{E}$, the dotted arrows in the diagram
+\begin{equation*}
+\begin{tikzcd}[column sep=tiny]
+\bullet \arrow[dr,densely dotted,swap,near start,"(W_A/\ctxext{B}{Q})(W_Q(R))"] & & \bullet \arrow[dl,densely dotted,near start,"W_{(W_A/B)(Q)}((W_A/B)(R))"] & & & & & & & & & & \bullet \arrow[dl,"W_Q(R)"] \\
+& \bullet \arrow[dr,swap,"(W_A/B)(Q)"] & & \bullet \arrow[dl,"(W_A/B)(R)"] & & & & & & & & \bullet \arrow[dr,swap,"Q"] & & \bullet \arrow[dl,"R"] \\
+& & \bullet \arrow[d,swap,"W_A(B)"] & & & & & & & & & & \ctxext{\Gamma}{B} \arrow[d,"B"] \\
+& & \ctxext{\Gamma}{A} \arrow[rrrrrrrrrr,swap,"A"] & & & & & & & & & & \Gamma
+\end{tikzcd}
+\end{equation*}
+are equal.
+
+A useful special case of this property is where $B=\catid{\Gamma}$. Thus, if $W$ is
+a weakening system, then the diagram
+\begin{equation*}
+\begin{tikzcd}[column sep=huge]
+\cat{C}/\ctxext{\Gamma}{Q}
+  \arrow[r,"W_A/Q"]
+  &
+\cat{C}/\ctxext{{\Gamma}{A}}{W_A(Q)}
+  \\
+\cat{C}/\Gamma
+  \arrow[u,"W_Q"]
+  \arrow[r,swap,"W_A"]
+  &
+\cat{C}/\ctxext{\Gamma}{A}
+  \arrow[u,swap,"W_{W_A(Q)}"]
+\end{tikzcd}
+\end{equation*}
+commutes for every $A\in\cat{C}/\Gamma$ and $Q\in\cat{C}/\Gamma$. 
+\end{rmk}
+
 \begin{cor}
-For any object $X$ of a weakening system $\cat{C}$, the slice category
-$\cat{C}/X$ is a weakening system.
+For any object $\Gamma$ of a weakening system $\cat{C}$, the slice pre-category
+$\cat{C}/\Gamma$ is a weakening system.
 \end{cor}
 
-\begin{rmk}
-If the underlying category is a group, the axioms of a weakening system
+\begin{eg}
+If the underlying pre-category is a group, the axioms of a weakening system
 obtain the following form: There is a group homomorphism $W_x:G\to G$ for
@@
 \end{equation*}
-is required to commute.
-\end{rmk}
+is required to commute. Thus, a weakening system on a group $G$ consists of
+a group homomorphism $W:\op G\to\mathrm{Aut}(G)$ such that $W_a\circ W_b=
+W_{W_a(b)}\circ W_a$. The latter requirement is satisfied if $b\cdot a=a\cdot W_a(b)$
+for every $a,b\in G$, or equivalently if $[b,a]=b^{-1}\cdot W_a(b)$.
+\end{eg}
 
@@
 
+\begin{comment}
 \begin{rmk}
@@
 \end{rmk}
+\end{comment}
+
+\begin{lem}
+Let $G$ be a graph, and let $\cat{C}_G$ be the free pre-category generated by
+$G$. A pre-weakening structure on $\cat{C}_G$ is determined by functors
+$W_e:\cat{C}_G/\Gamma\to\cat{C}_G/\Delta$ for every edge $e:\Delta\to \Gamma$ in $G$, with the
+property that $W_e(\catid{\Gamma})=\catid{\Delta}$.
+
+A weakening structure on $\cat{C}_G$ is
+a pre-weakening structure on $\cat{C}_G$ such that the diagram
+\begin{equation*}
+\begin{tikzcd}[column sep=huge]
+\cat{C}_G/\greek{Z} \arrow[r,"W_e/(g\circ h)"] & \cat{C}_G/\mathrm{dom}(W_e(g\circ h)) \\
+\cat{C}_G/\greek{E} \arrow[u,"W_h"] \arrow[r,swap,"W_e/g"] & \cat{C}_G/\mathrm{dom}(W_e(g)) \arrow[u,swap,"W_{W_e(h)}"]
+\end{tikzcd}
+\end{equation*}
+commutes for every arrow $g:\greek{E}\to \Gamma$ in $\cat{C}_G$ and every two edges $e:\Delta\to \Gamma$ and
+$h:\greek{Z}\to \greek{E}$ in $G$.
+\end{lem}
+
+\begin{eg}
+Let $G$ be the graph consisting of one vertex $\ast$ and one edge $e:\ast\to\ast$.
+Then $\cat{C}_G/\ast=\N$. Then the only weakening structure on $\cat{C}/G$ is
+given by $W_e=\catid{\N}$. The reason is that 
+\end{eg}
+
+\begin{lem}
+Let $P$ be a pre-order. A pre-weakening structure on $P$ is determined by order
+preserving maps $W_{a,b}:{\downarrow}(b)\to{\downarrow}(a)$ satisfying
+$W_{a,b}(b)=a$, for every $a\leq b$, where
+${\downarrow}(x)$ is the set $\{y\in P\mid y\leq x\}$.
+
+A weakening structure on $P$ is a pre-weakening structure on $P$ such that the
+diagram
+\begin{equation*}
+\begin{tikzcd}[column sep=huge]
+{\downarrow}(c') \arrow[r,"W_{a,b}|_{{\downarrow}(c')}"] & {\downarrow}(W_{a,b}(c')) \\
+{\downarrow}(c) \arrow[u,"W_{c',c}"] \arrow[r,swap,"W_{a,b}|_{{\downarrow}(c)}"] & {\downarrow}(W_{a,b}(c)) \arrow[u,swap,"W_{W_{a,b}(c'),W_{a,b}(c)}"]
+\end{tikzcd}
+\end{equation*}
+for any $a\leq b$ and $c'\leq c\leq b$.
+\end{lem}
+
+\begin{eg}
+In a poset with meets, the functions $W_{a,b}:{\downarrow}(b)\to{\downarrow}(a)$
+given by $W_{a,b}(x):= a\land x$ assemble a weakening structure, because for
+any $a\leq b$ and $x,c'\leq c\leq b$ we have $a\land(c'\land x)=(a\land c')\land
+(a\land x)$.  
+\end{eg}
+
+\begin{eg}
+Consider the poset $(\N,\geq)$, which has a morphism from $a$ to $b$ if $a\geq b$.
+On this poset, we may consider the weakening structure defined by $W_{a,b}(x)=
+a+(x-b)$.
+\end{eg}
 
@@
 \begin{defn}
-A \emph{category with a term structure} is a category $\cat{C}$ with a set $T(f)$ for every
-morphism $f$ of $\cat{C}$. A \emph{functor with term structure} from $\cat{C}$ to $\cat{D}$
-is a functor $F:\cat{C}\to\cat{D}$ with a function $T(f)\to T(F(f))$ for
-every $f:X\to Y$ in $\cat{C}$.
+A \emph{category with a term structure} is a pre-category $\cat{C}$ with a set $T(A)$ for every
+$A\in\cat{C}/\Gamma$ for every $\Gamma\in\cat{C}$. A \emph{functor with term structure} from $\cat{C}$ to $\cat{D}$
+is a functor $F:\cat{C}\to\cat{D}$ with a function $T(A)\to T(F(A))$ for
+every $A\in\cat{C}/\Gamma$, for every $\Gamma\in\cat{C}$.
 \end{defn}
@@
 \begin{defn}
-Suppose $\cat{C}$ is a category with term structure, and $X$ is an object of $\cat{C}$. 
-Then $\cat{C}/X$ is a category with term structure, where $T_{\cat{C}/X}(g)=T_\cat{C}(g)$.
+Suppose $\cat{C}$ is a pre-category with term structure, and $\Gamma$ is an object of $\cat{C}$. 
+Then $\cat{C}/\Gamma$ is a pre-category with term structure, where $T_{\cat{C}/\Gamma}(A)=T_\cat{C}(A)$.
 \end{defn}
@@
 A \emph{weakening system with term term structure} is the same thing as a weakening system,
-except that the underlying category and the functors $W_f$ are required to
+except that the underlying pre-category and the functors $W_A$ are required to
 be functors with term term structure. The equalities are also required to hold on the term structure.
@@
 \begin{cor}
-For any object $X$ of a weakening system with term structure, the slice category
-$\cat{C}/X$ is a weakening system with term structure.
+For any object $\Gamma$ of a weakening system with term structure, the slice pre-category
+$\cat{C}/\Gamma$ is a weakening system with term structure.
 \end{cor}
+
+\begin{eg}
+%For any group $G$ with weakening structure $W:\op{G}\to\mathrm{Aut}(G)$, a term 
+%structure is determined by a presheaf on $G/\ker(W)$.
+
+Let $G$ be a group, with the weakening structure given by conjugation. A term
+structure on $G$ is given by a set $X$ with a right $G$-action $\mu:\op G\times X\to X$.
+We take $T(a):= X$ for every $a\in G$.
+\end{eg}
 
@@
 A \emph{pre-projection system} is a weakening system with term structure with an element
-$\tfid{f}\in T(W_f(f))$ for every $f:X\to Y$ in $\cat{C}$. 
+$\tfid{A}\in T(W_A(A))$ for every $A\in\cat{C}/\Gamma$ and $\Gamma\in\cat{C}$. 
 \end{defn}
@@
 \begin{equation*}
-F(\tfid{f})=\tfid{F(f)}
+F(\tfid{A})=\tfid{F(A)}
 \end{equation*}
-for every $f:X\to Y$ in $\cat{C}$.
+for every $A\in\cat{C}/\Gamma$ and $\Gamma\in\cat{C}$.
 \end{defn}
@@
 \begin{lem}
-For any object $X$ in a pre-projection system $\cat{C}$, the slice category
-$\cat{C}/X$ is a pre-projection system.
+For any object $\Gamma$ in a pre-projection system $\cat{C}$, the slice pre-category
+$\cat{C}/\Gamma$ is a pre-projection system.
 \end{lem}
@@
 \begin{defn}
-A \emph{projection system} is a pre-projection system for which every $W_f$ is
+A \emph{projection system} is a pre-projection system for which every $W_A$ is
 a pre-projection homomorphism. A \emph{projection homomorphism} is a pre-projection homomorphism
@@
 \begin{cor}
-For any object $X$ of a projection system $\cat{C}$, the slice category $\cat{C}/X$
+For any object $\Gamma$ of a projection system $\cat{C}$, the slice pre-category $\cat{C}/\Gamma$
 is a projection system.
 \end{cor}
+
+\begin{eg}
+Consider a group $G$ with a right $G$-set $X$. A projection structure on $(G,X)$
+consists of an element $\tfid{a}\in X$ for every $a\in G$, so that $\tfid{a}\cdot b=
+\tfid{b^{-1}ab}$ for every $a,b\in G$.
+
+%For instance, this is the case if we take $X$ to be a pointed set with $\ast\in X$,
+%and take $\tfid{a}:=\ast$ for all $a\in G$.
+\end{eg}
 
@@
 \begin{defn}
-A \emph{pre-substitution structure} on a category with term structure $\cat{C}$ consists
-of a functor with term structure $S_t:\cat{C}/X\to\cat{C}/Y$ for every $t\in T(f)$ and
-$f:X\to Y$, such that $S_t(\catid{X})=\catid{Y}$.
+A \emph{pre-substitution structure} on a pre-category with term structure $\cat{C}$ consists
+of a functor with term structure $S_x:\cat{C}/\ctxext{\Gamma}{A}\to\cat{C}/\Gamma$ for every $x\in T(A)$ and
+$A\in\cat{C}/\Gamma$, such that $S_x(\catid{\ctxext{\Gamma}{A}})=\catid{\Gamma}$.
 
-A \emph{pre-substitution system} is a category with term structure together with a
+A \emph{pre-substitution system} is a pre-category with term structure together with a
 pre-substitution structure. 
@@
 \begin{tikzcd}[column sep=large]
-\cat{C}/X
-  \arrow[r,"F/X"]
-  \arrow[d,swap,"S_t"]
+\cat{C}/\ctxext{\Gamma}{A}
+  \arrow[r,"F/\ctxext{\Gamma}{A}"]
+  \arrow[d,swap,"S_x"]
   &
-\cat{D}/F(X)
-  \arrow[d,"S_{F(t)}"]
+\cat{D}/F(\ctxext{\Gamma}{A})
+  \arrow[d,"S_{F(x)}"]
   \\
-\cat{C}/Y
-  \arrow[r,swap,"F/Y"]
+\cat{C}/\Gamma
+  \arrow[r,swap,"F/\Gamma"]
   &
-\cat{D}/F(Y)
+\cat{D}/F(\Gamma)
 \end{tikzcd}
 \end{equation*}
-commutes.
+commutes for every $x\in T(A)$ and $A\in\cat{C}/\Gamma$.
 \end{defn}
@@
 \begin{lem}
-For any object $X$ of a pre-substitution system $\cat{C}$, the slice category
-$\cat{C}/X$ is a pre-substitution system.
+For any object $\Delta$ of a pre-substitution system $\cat{C}$, the slice pre-category
+$\cat{C}/\Delta$ is a pre-substitution system.
 \end{lem}
@@
 \begin{cor}
-For any object $X$ of a substitution system, the slice category $\cat{C}/X$
+For any object $\Delta$ of a substitution system, the slice pre-category $\cat{C}/\Delta$
 is a substitution system.
@@
 
+\begin{rmk}
+The condition that every $S_x$ is a substitution homomorphism, asserts that
+the diagram
+\begin{equation*}
+\begin{tikzcd}[column sep=large]
+\cat{C}/\ctxext{{{\Gamma}{A}}{P}}{Q}
+  \arrow[r,"S_x/\ctxext{P}{Q}"]
+  \arrow[d,swap,"S_y"]
+  &
+\cat{C}/\ctxext{{\Gamma}{S_x(P)}}{S_x(Q)}
+  \arrow[d,"S_{S_x(y)}"]
+  \\
+\cat{C}/\ctxext{{\Gamma}{A}}{P}
+  \arrow[r,swap,"S_x/P"]
+  &
+\cat{C}/\ctxext{\Gamma}{S_x(P)}
+\end{tikzcd}
+\end{equation*}
+commutes for every $y\in T(Q)$.
+\end{rmk}
+
+\begin{eg}
+Let $G$ be a group and let $H$ be a set, forming a category with term 
+structure of which $T(a)=H$ for every $a\in G$. A substitution
+system on $(G,H)$ consists of a function $\sigma_G : H\times G\to G$ and
+a function $\sigma_H:H\times H\to H$, for which the diagrams
+\begin{equation*}
+\begin{tikzcd}[column sep=large]
+G \arrow[r,"\sigma_G(h)"]
+  \arrow[d,swap,"\sigma_G(k)"]
+  &
+G \arrow[d,"{\sigma_G(\sigma_H(h,k))}"]
+  \\
+G \arrow[r,swap,"\sigma_G(h)"] 
+& G
+\end{tikzcd}
+\end{equation*}
+and
+\begin{equation*}
+\begin{tikzcd}[column sep=large]
+H \arrow[r,"\sigma_H(h)"]
+  \arrow[d,swap,"\sigma_H(k)"]
+  &
+H \arrow[d,"{\sigma_H(\sigma_H(h,k))}"]
+  \\
+H \arrow[r,swap,"\sigma_H(h)"]
+  &
+H
+\end{tikzcd}
+\end{equation*}
+commute for every $h,k\in H$. 
+\end{eg}
+
 \section{E-systems}
 \begin{defn}
-An \emph{E-system} is a category with term structure with
+\label{defn:esystems}
+An \emph{E-system} is a pre-category $\cat{C}$ with term structure with
 \begin{itemize}
@@
 \begin{enumerate}
-\item each $W_f$ is a substitution homomorphism;
-\item each $S_t$ is a projection homomorphism;
-\item $S_t(\tfid{f})=t$ for any $t\in T(f)$ and $f:X\to Y$;
-\item $S_t\circ W_f=\catid{\cat{C}/Y}$ for any $t\in T(f)$ and $f:X\to Y$;
-\item $S_{\tfid{f}}\circ W_{W_f(f)}=\catid{\cat{C}/X}$ for any $f:X\to Y$.
+\item each $W_A$ is a substitution homomorphism, i.e.~the diagram
+\begin{equation*}
+\begin{tikzcd}[column sep=huge]
+\cat{C}/\ctxext{{\Gamma}{B}}{Q}
+  \arrow[r,"W_A/\ctxext{B}{Q}"]
+  \arrow[d,swap,"S_y"]
+  &
+\cat{C}/\ctxext{{{\Gamma}{A}}{W_A(B)}}{W_A(Q)}
+  \arrow[d,"S_{W_A(y)}"]
+  \\
+\cat{C}/\ctxext{\Gamma}{B}
+  \arrow[r,swap,"W_A/B"]
+  &
+\cat{C}/\ctxext{{\Gamma}{A}}{W_A(B)}
+\end{tikzcd}
+\end{equation*}
+commutes for every $Q\in\cat{C}/\ctxext{\Gamma}{B}$, $B\in\cat{C}/\Gamma$ and each 
+$y\in T(Q)$; 
+\item each $S_x$ is a projection homomorphism, i.e.~the diagram
+\begin{equation*}
+\begin{tikzcd}[column sep=huge]
+\cat{C}/\ctxext{{\Gamma}{A}}{P}
+  \arrow[r,"S_x/P"]
+  \arrow[d,swap,"W_Q"]
+  &
+\cat{C}/\ctxext{\Gamma}{S_x(P)}
+  \arrow[d,"W_{S_x(Q)}"]
+  \\
+\cat{C}/\ctxext{{{\Gamma}{A}}{P}}{Q}
+  \arrow[r,swap,"S_x/\ctxext{P}{Q}"]
+  &
+\cat{C}/\ctxext{{\Gamma}{S_x(P)}}{S_x(Q)}
+\end{tikzcd}
+\end{equation*}
+commutes for every $Q\in\cat{C}/\ctxext{{\Gamma}{A}}{P}$, and we have
+\begin{equation*}
+S_x(\tfid{Q})=\tfid{S_x(Q)}
+\end{equation*}
+for every $Q\in\cat{C}/\ctxext{{\Gamma}{A}}{P}$;
+\item $S_x(\tfid{A})=x$ for any $x\in T(A)$ and $A\in\cat{C}/\Gamma$;
+\item \label{tTf:StWf_id} $S_x\circ W_A=\catid{\cat{C}/\Gamma}$ for any $x\in T(A)$ and $A\in\cat{C}/\Gamma$;
+\item \label{f:SidfWff_id} $S_{\tfid{A}}\circ W_A/A=\catid{\cat{C}/\ctxext{\Gamma}{A}}$ for any $A\in\cat{C}/\Gamma$.
 \end{enumerate}
-\end{defn}
-
-\begin{cor}
-For any object $X$ of an E-system $\cat{C}$, the slice category $\cat{C}/X$
-is an E-system.
-\end{cor}
-
-\begin{defn}
 An \emph{E-homomorphism} $F:\cat{C}\to\cat{D}$ is a functor which is both a projection
@@
 
+\begin{cor}
+For any object $\Delta$ of an E-system $\cat{C}$, the slice pre-category $\cat{C}/\Delta$
+is an E-system.
+\end{cor}
+
+\begin{rmk}[Needs more precision]
+Consider the category $E$ generated by the graph consisting of the objects of 
+the underlying category of an $E$-system $\mathcal{E}$ as edges, and the morphisms
+$S_t:X\to Y$ and $W_A:Y\to X$ as edges, subject to the axioms of an $E$-systems.
+
+Any morphism of $E$ can be rewritten in a unique way
+as a composition $\tilde S\circ \tilde W$, where $\tilde S$ is a morphism in
+substitution-normal form, $\tilde W$ is in weakening normal form, every term
+expression is in normal form, and where
+each occurence of $S_x\circ W_A$ and $S_{\tfid{A}}\circ W_A/A$ is eliminated.
+\end{rmk}
+
+\begin{defn}
+An E-system of which the underlying category has a terminal object $\emptyc$, is said to
+be \emph{E-system with empty context}. An E-homomorphism is said to preserve the
+empty context when it preserves the terminal object $\emptyc$ on the nose.
+\end{defn}
+
+\begin{rmk}
+When there is a terminal object, we get a weakening functor $W_\Gamma:\cat{C}\to
+\cat{C}/\Gamma$ for every object $\Gamma$ of $\cat{C}$; we get a set $T(\Gamma)$;
+and for $i\in T(\Gamma)$ we get substitutions $S_i:\cat{C}/\Gamma\to\cat{C}$. 
+\end{rmk}
+
+\section{Properties of E-systems}
+
+\begin{defn}
+For every $A,B\in\cat{C}/\Gamma$ we define the set 
+\begin{equation*}
+\thom{A}{B}:=T(W_A(B)).
+\end{equation*} 
+\end{defn}
+
+\begin{rmk}
+Note that $\thom{\catid{\Gamma}}{A}=T(A)$ for any $A\in\cat{C}/\Gamma$, 
+because we have $W_{\catid{\ctxext{\Gamma}{A}}}=\catid{\cat{C}/\ctxext{\Gamma}{A}}$.
+
+Note also that $\thom{\ctxext{A}{P}}{B}=\thom{P}{W_A(B)}$ for any $P\in\cat{C}/\ctxext{\Gamma}{A}$
+and $B\in\cat{C}/\Gamma$, because $W_{\ctxext{A}{P}}=
+W_P\circ W_A$.
+\end{rmk}
+
+\begin{defn}
+Let $A,B\in\cat{C}/\Gamma$.
+For any $f\in\thom{A}{B}$ we define the pre-composition E-homomorphism
+\begin{equation*}
+f_\ast := S_f\circ W_A/B : \cat{C}/\ctxext{\Gamma}{B}\to\cat{C}/\ctxext{\Gamma}{A}.
+\end{equation*}
+Furthermore, for $g\in\thom{B}{C}$ we may consider $W_B(C)\in\cat{C}/\ctxext{\Gamma}{B}$ as a morphism from $W_B(C)$ to $\catid{\ctxext{\Gamma}{B}}$. Therefore, $f_\ast$ acts on the terms of
+$W_B(C)$, so it makes sense to define $g\circ f:=f_\ast(g)$. 
+\end{defn}
+
+Note that $g\circ f\in T(S_f(W_A/B(W_B(C))))$, whereas we would like that
+$g\circ f\in\thom{A}{C}$. The following lemma confirms this:
+
+\begin{lem}
+Suppose $A$ and $B$ are morphisms of the underlying pre-category of $\mathcal{E}$
+with the same codomain, and let $f\in\thom{A}{B}$. Then we have 
+\begin{equation*}
+S_f\circ (W_A/B)\circ W_B=W_A.
+\end{equation*}
+\end{lem}
+
+\begin{proof}
+Since weakening is a weakening homomorphism, we have
+\begin{equation*}
+S_f\circ (W_A/B)\circ W_B=S_f\circ W_{W_A(B)}\circ W_A
+\end{equation*}
+By \autoref{tTf:StWf_id} of \autoref{defn:esystems} it we get that
+\begin{equation*}
+S_f\circ W_{W_A(B)}\circ W_A=W_A.\qedhere
+\end{equation*}
+\end{proof}
+
+\begin{rmk}
+Note that \autoref{f:SidfWff_id} of \autoref{defn:esystems} asserts precisely
+that ${(\tfid{A})}_\ast=\catid{\cat{C}/\Delta}$ for any $A:\Delta\to \Gamma$ in $\cat{C}$. In
+particular, it follows that $g\circ\tfid{A}=g$ for any $g\in\thom{A}{B}$
+\end{rmk}
+
+\begin{lem}
+For any $f\in\thom{A}{B}$ we have $\tfid{B}\circ f=f$.
+\end{lem}
+
+\begin{proof}
+We have
+\begin{equation*}
+\tfid{B}\circ f=S_f(W_A(\tfid{B}))=S_f(\tfid{W_A(B)})=f.\qedhere
+\end{equation*}
+\end{proof}
+
+\begin{lem}
+For any $f\in\thom{A}{B}$ and $g\in\thom{B}{C}$ we have $f_\ast\circ g_\ast=(g\circ f)_\ast$. 
+\end{lem}
+
+\begin{proof}
+\begin{align*}
+f_\ast\circ g_\ast & = S_f\circ (W_A/B) \circ S_g \circ (W_B/C)\\
+& = S_f\circ S_{W_A(g)}\circ(W_A/\ctxext{B}{W_B(C)})\circ (W_B/C) \\
+& = S_{S_f(W_A(g))}\circ (S_f/W_A(W_B(C)))\circ (W_A/\ctxext{B}{W_B(C)}) \circ W_B/C \\
+& = S_{S_f(W_A(g))}\circ ((S_f\circ (W_A/B) \circ W_B)/C) \\
+& = S_{S_f(W_A(g))}\circ ((S_f\circ W_{W_A(B)}\circ W_A)/C) \\
+& = S_{S_f(W_A(g))}\circ W_A/C \\
+& = (g\circ f)_\ast.\qedhere
+\end{align*}
+\end{proof}
+
+\begin{cor}
+Composition is associative. Thus, we have constructed a pre-category for each
+object $\Gamma$, with the same objects as $\cat{C}/\Gamma$, of which the hom-sets
+are the sets $\thom{A}{B}$. 
+\end{cor}
+
+\begin{rmk}
+Since any E-homomorphism is both a substitution homomorphism and a weakening
+homomorphism, it follows immediately that any E-homomorphism preserves any
+$f_\ast$, i.e.~for any $f\in\thom{A}{B}$ in $\mathcal{E}$ and any E-homomorphism
+$F:\mathcal{E}\to\mathcal{D}$, the diagram
+\begin{equation*}
+\begin{tikzcd}[column sep=large]
+\mathcal{E}/\ctxext{\Gamma}{A} \arrow[r,"F/\ctxext{\Gamma}{A}"] & \mathcal{D}/F(\ctxext{\Gamma}{A})
+  \\
+\mathcal{E}/\ctxext{\Gamma}{B} \arrow[u,"f_\ast"] \arrow[r,swap,"F/\ctxext{\Gamma}{B}"] & \mathcal{D}/F(\ctxext{\Gamma}{B}) \arrow[u,swap,"F(f)_\ast"]
+\end{tikzcd}
+\end{equation*}
+commutes. This applies to weakening, substitution and pre-composition.
+\end{rmk}
+
+\begin{cor}
+Suppose $F:\mathcal{E}\to\mathcal{D}$ is an E-homomorphism. Then $F$ induces
+a functor $\mathbb{T}_F:\mathbb{T}_\mathcal{E}\to \mathbb{T}_\mathcal{D}$. 
+\end{cor}
+
+\subsection{Pairing and the projections}
+
+\begin{defn}
+Let $x\in T(A)$ and $u\in T(S_x(P))$ for $A\in\cat{C}/\Gamma$ and $P\in\cat{C}/\ctxext{\Gamma}{A}$. 
+We define
+\begin{equation*}
+\tmext{x}{u}\defeq S_u(S_x(\tfid{\ctxext{A}{P}}))\in T(\ctxext{A}{P}).
+\end{equation*}
+\end{defn}
+
+\begin{thm}
+Let $x\in T(A)$ and $u\in T(S_x(P))$ for $A\in\cat{C}/\Gamma$ and $P\in\cat{C}/\ctxext{\Gamma}{A}$.
+Then we have
+\begin{equation*}
+S_{\tmext{x}{u}}=S_u\circ (S_x/P):\cat{C}/\ctxext{\Gamma}{{A}{P}}\to\cat{C}/\Gamma
+\end{equation*}
+\end{thm}
+
+\begin{defn}
+Let $A\in\cat{C}/\Gamma$ and $P\in\cat{C}/\ctxext{\Gamma}{A}$. We define
+\begin{align*}
+\cprojfstf{A}{P} & \defeq W_P(\tfid{A})\in\thom{\ctxext{A}{P}}{A}\\
+\cprojsndf{A}{P} & \defeq \tfid{P}\in\thomd{\cprojfstf{A}{P}}{\ctxext{A}{P}}{P}
+\end{align*}
+\end{defn}
+
+\begin{lem}
+Let $x\in T(A)$ and $u\in T(S_x(P))$. then we have
+\begin{align*}
+S_{\tmext{x}{u}}(\cprojfstf{A}{P}) & \jdeq x\\
+S_{\tmext{x}{u}}(\cprojsndf{A}{P}) & \jdeq u.
+\end{align*}
+Also, for $w\in T(\ctxext{A}{P})$ we have
+\begin{equation*}
+\tmext{S_w(\cprojfstf{A}{P})}{S_w(\cprojsndf{A}{P})} \jdeq w.
+\end{equation*}
+\end{lem}
+
+\begin{lem}
+For any $x\in T(A)$, $u\in T(S_x(P))$ and $v\in T(S_{\tmext{x}{u}}(Q))$ we have
+\begin{equation*}
+\tmext{{x}{u}}{v}\jdeq \tmext{x}{{u}{v}}\in T(\ctxext{{A}{P}}{Q}).
+\end{equation*}
+Also, we have
+\begin{align*}
+\cprojfstf{A}{P}\circ\cprojfstf{\ctxext{A}{P}}{Q} & \jdeq\cprojfstf{A}{\ctxext{P}{Q}} \\
+\cprojsndf{A}{P}\circ\cprojfstf{\ctxext{A}{P}}{Q} & \jdeq \cprojfstf{P}{Q}\circ\cprojsndf{A}{\ctxext{P}{Q}} \\
+\cprojsndf{\ctxext{A}{P}}{Q} & \jdeq \cprojsndf{P}{Q}\circ\cprojsndf{A}{\ctxext{P}{Q}}.
+\end{align*}
+\end{lem}
+
+\begin{thm}
+Let $F:\mathcal{E}\to\mathcal{D}$ be an E-homomorphism and let $x\in T(A)$ and
+$u\in T(S_x(P))$. Then we have
+\begin{equation*}
+F(\tmext{x}{u})\jdeq \tmext{F(x)}{F(u)}.
+\end{equation*}
+Also, we have
+\begin{align*}
+F(\cprojfstf{A}{P}) & \jdeq \cprojfstf{F(A)}{F(P)} \\
+F(\cprojsndf{A}{P}) & \jdeq \cprojsndf{F(A)}{F(P)}.
+\end{align*}
+\end{thm}
+
+\begin{lem}
+Let $f\in\thom{A}{\ctxext{B}{Q}}$ Then we have
+\begin{equation*}
+f \jdeq \tmext{\jcomp{A}{f}{\cprojfstf{B}{Q}}}{\jcomp{A}{f}{\cprojsndf{B}{Q}}}.
+\end{equation*}
+Alternatively, when $f_0\in\thom{A}{B}$ and $f_1\in\thom{A}{\jcomp{A}{f}{Q}}$,
+then we have $\tmext{f_0}{f_1}\in\thom{A}{\ctxext{B}{Q}}$ and
+\begin{align*}
+\jcomp{A}{\tmext{f_0}{f_1}}{\cprojfstf{B}{Q}} & \jdeq f_0 \\
+\jcomp{A}{\tmext{f_0}{f_1}}{\cprojsndf{B}{Q}} & \jdeq f_1.
+\end{align*}
+\end{lem}
+
+\begin{thm}
+Precomposition with $\cprojfstf{A}{P}$ is weakening by $P$, i.e.
+\begin{equation*}
+S_{\cprojfstf{A}{P}}\circ W_{\ctxext{A}{P}}\jdeq W_P.
+\end{equation*}
+\end{thm}
+
+\section{Natural transformations of E-homomorphisms}
+\begin{rmk}
+When $\mathcal{E}$ is an E-system with empty context, then we have the category
+structure on the contexts given by $\thom{\Gamma}{\Delta}= T(W_\Gamma(\Delta))$.
+\end{rmk}
+
+\begin{defn}
+A \emph{natural transformation} of E-homomorphisms from $F:\mathcal{E}\to\mathcal{D}$
+to $G:\mathcal{E}\to\mathcal{D}$ preserving the empty context consists of
+\begin{enumerate}
+\item A t-morphism $\tau0_A\in\thom{F(\Gamma)}{G(\Gamma)}$ for every $\Gamma\in\mathcal{E}$.
+\item A commuting square square
+\begin{equation*}
+\begin{tikzcd}
+F(A) \arrow[r,"\tau1_A"] \arrow[d,twoheadrightarrow] & G(A) \arrow[d,twoheadrightarrow] \\
+F(\Gamma) \arrow[r,swap,"\tau0_\Gamma"] & G(\Gamma)
+\end{tikzcd}
+\end{equation*}
+for every family $P$ over $A$ of solid arrows, such that the diagram
+\begin{equation*}
+\begin{tikzcd}
+F(A) \arrow[r,"\tau1_A"] & G(A) \\
+F(\Gamma) \arrow[r,swap,"\tau0_\Gamma"] \arrow[densely dotted,u,"F(x)"] & G(\Gamma) \arrow[densely dotted,u,swap,"G(x)"]
+\end{tikzcd}
+\end{equation*}
+commutes for every $x\in T(A)$.
+\end{enumerate}
+such that
+\begin{enumerate}
+\item the square
+\begin{equation*}
+\begin{tikzcd}[column sep=huge]
+W_{F(A)}(F(B)) \arrow[r,"W_{F(A)}(\tau1_B)"] \arrow[d,->>] & W_{G(A)}(G(B)) \arrow[d,->>] \\
+F(A) \arrow[r,"\tau1_A"] & G(A)
+\end{tikzcd}
+\end{equation*}
+commutes for every $A,B\in\mathcal{E}/\Gamma$. Note that 
+\begin{align*}
+W_{F(A)}(\tau1_B)
+  & \in 
+T(W_{F(A)}(W_{F(B)}(G(B))))
+  \\
+  & =
+T(W_{W_{F(A)}(F(B))}(W_{F(A)}(G(B))))
+\end{align*}
+\end{enumerate}
+\end{defn}
+
+\section{Indexed E-systems}
+\begin{defn}
+Let $\cat{A}$ be a category. An $\cat{A}$-indexed E-system is an internal
+E-system in the category $Psh(\cat{A})$.
+\end{defn}
+
+\section{Examples of adjunctions}
+
+\begin{lem}
+For each $A\in\mathcal{E}/\Gamma$, $W_A$ is left adjoint to $\Pi_A$. 
+\end{lem}
+
+\begin{proof}
+The unit $\eta:1_{\cat{C}/\ctxext{\Gamma}{A}}\Rightarrow W_A\circ\Pi_A$ is given by
+\begin{equation*}
+\eta_P:\lam{f}{x} ev(\lambda(f),x)
+\end{equation*}
+$\eta_P\in\thom{P}{W_A(\Pi_A(P))}$
+\end{proof}
+
 \begin{comment}
 \begin{lem}
-Suppose $\cat{D}$ is a retract of a category $\cat{C}$, i.e. there are functors
+Suppose $\cat{D}$ is a retract of a pre-category $\cat{C}$, i.e. there are functors
 $S:\cat{D}\to\cat{C}$ and $R:\cat{C}\to\cat{D}$ with a natural isomorphism
@@
 \begin{proof}
-Let $f:X\to Y$ be a morphism in $\cat{D}$. Then we define $W_f$ to be the
+Let $f:\Delta\to \Gamma$ be a morphism in $\cat{D}$. Then we define $W_f$ to be the
 composition
@@
 \begin{tikzcd}[column sep=3.2em]
-\cat{D}/Y
-  \arrow[r,"S/Y"]
+\cat{D}/\Gamma
+  \arrow[r,"S/\Gamma"]
   &
-\cat{C}/S(Y)
+\cat{C}/S(\Gamma)
   \arrow[r,"W_{S(f)}"]
   &
-\cat{C}/S(X)
-  \arrow[r,"R/S(X)"]
+\cat{C}/S(\Delta)
+  \arrow[r,"R/S(\Delta)"]
   &
-\cat{D}/R(S(X))
-  \arrow[r,"\tau_X\circ{-}"]
+\cat{D}/R(S(\Delta))
+  \arrow[r,"\tau_\Delta\circ{-}"]
   &
-\cat{D}/X
+\cat{D}/\Delta
 \end{tikzcd}
@@
 \begin{enumerate}
-\item Let $X$ be an object of $\cat{D}$. Then $W_{\catid{X}}$ is the composition
+\item Let $\Delta$ be an object of $\cat{D}$. Then $W_{\catid{\Delta}}$ is the composition
 \begin{equation*}
 \begin{tikzcd}[column sep=3.2em]
-\cat{D}/X
-  \arrow[r,"S/X"]
+\cat{D}/\Delta
+  \arrow[r,"S/\Delta"]
   &
-\cat{C}/S(X)
-  \arrow[r,"R/S(X)"]
+\cat{C}/S(\Delta)
+  \arrow[r,"R/S(\Delta)"]
   &
-\cat{D}/R(S(X))
-  \arrow[r,"{-}\circ\tau_X"]
+\cat{D}/R(S(\Delta))
+  \arrow[r,"{-}\circ\tau_\Delta"]
   &
-\cat{D}/X
+\cat{D}/\Delta
 \end{tikzcd}
 \end{equation*}
-This functor maps $g:X'\to X$ consecutively to $S(g)$, to $R(S(g))$, to
-$\tau_X\circ R(S(g))$. Note that $\tau_X\circ R(S(g))=g\circ \tau_{X'}$ by
+This functor maps $g:\Delta'\to \Delta$ consecutively to $S(g)$, to $R(S(g))$, to
+$\tau_\Delta\circ R(S(g))$. Note that $\tau_\Delta\circ R(S(g))=g\circ \tau_{\Delta'}$ by
 naturality.

### 2015-04-24 weakeningcats.tex
small reorg

@@
 
+\subsection{The category of internal functions}
+
 \begin{defn}
@@
 
+The composition $\ctxext{A}{P}$ of $A\in\cat{C}/\Gamma$ and $P\in\cat{C}/\ctxext{\Gamma}{A}$
+behaves like a strict $\Sigma$-type. We may define the pairing term
+$\typefont{pair}^{A,P}\defeq\tfid{\ctxext{A}{P}}\in T(W_P(W_A(\ctxext{A}{P})))$
+and the projections and prove several useful properties about them. The strictness
+is found, among other things, in the fact that we can prove judgmental $\eta$-equality,
+and that pairing is strictly associative.
+
 \begin{defn}
@@
 \end{defn}
+
+\begin{cor}
+Let $F:\mathcal{E}\to\mathcal{D}$ be an E-homomorphism and let $x\in T(A)$ and
+$u\in T(S_x(P))$. Then we have
+\begin{equation*}
+F(\tmext{x}{u})\jdeq \tmext{F(x)}{F(u)}.
+\end{equation*}
+Also, we have
+\begin{align*}
+F(\cprojfstf{A}{P}) & \jdeq \cprojfstf{F(A)}{F(P)} \\
+F(\cprojsndf{A}{P}) & \jdeq \cprojsndf{F(A)}{F(P)}.
+\end{align*}
+\end{cor}
 
@@
 \end{lem}
-
-\begin{thm}
-Let $F:\mathcal{E}\to\mathcal{D}$ be an E-homomorphism and let $x\in T(A)$ and
-$u\in T(S_x(P))$. Then we have
-\begin{equation*}
-F(\tmext{x}{u})\jdeq \tmext{F(x)}{F(u)}.
-\end{equation*}
-Also, we have
-\begin{align*}
-F(\cprojfstf{A}{P}) & \jdeq \cprojfstf{F(A)}{F(P)} \\
-F(\cprojsndf{A}{P}) & \jdeq \cprojsndf{F(A)}{F(P)}.
-\end{align*}
-\end{thm}
 

### 2015-06-12 main.tex
The new E-systems
Theorems about E-systems
huge revision

@@
 \title{An essentially algebraic formulation of dependent type theory}
-\author{Egbert Rijke}
+\author{Egbert Rijke\\ {\sc Carnegie Mellon University}}
 \date{\today}
@@
 \maketitle
+
+\begin{abstract}
+We present an algebraic formulation of dependent type theory with
+three sorts for contexts, families and terms, with finitely many operations
+between them and finitely many equality axioms. Algebras for this theory are
+called E-systems. We then show that there is a full and faithful functor
+from Voevodsky's B-systems into E-systems, of which the image is the category
+of stratified E-systems. 
+\end{abstract}
+
+\section{Introduction}
+
+When type theory is regarded as a foundational system for doing mathematics,
+we would first of all like it to have finitely many axioms which do not depend
+on any presumed mathematical structures. In particular, natural numbers should
+not be needed to state what type dependency is. Nevertheless, many modern
+accounts on what type theory is, do rely on the existence of the natural
+numbers in the meta-theory. The type theory of \cite{TheBook} uses an infinte
+hierarchy $\UU_0$, $\UU_1$, $\UU_2$, \ldots of universes, as do the type
+type theories of \Coq and \Agda. Likewise, models of type theory considered
+by Cartmell, Voevodsky and Garner do rely on the natural numbers.
+
+That leaves Dybjer's categories with families as
+
+\subsection*{Existing models of type theory}
+
+\subsection*{Essentially algebraic theories}
+An essentially algebraic theory is an equational theory of partial operations
+which may be ordered, and of which the domain of each operation is declared by 
+equations on the preceeding operations. The concept of essentially algebraic
+theories is due to Freyd, who writes that a critical feature of essentially
+algebraic theories is that their models are closed under 
+colimits \cite{Aspects_of_topoi}.
+
+In \cite{Adamek_Locally}, Ad\'amek and Rosicky
+define an essentially algebraic theory to be a list 
+\begin{equation*}
+(S,\Sigma,a,d,c,E,\Sigma_t,\mathrm{Def})
+\end{equation*} 
+consisting of a set $S$ of sorts, a set $\Sigma$ of operation symbols,
+an arity function $a:\Sigma\to\mathbf{Set}$ specifying the arity of each operation
+symbol, a map $d_\sigma:a(\sigma)\to S$
+for each $\sigma\in\Sigma$ specifying the sort of the domain of $\sigma$,
+a map $c:\Sigma\to S$ specifying the sort of the codomain of each operation
+symbol,
+a set $E$ of $\Sigma$-equations, a set $\Sigma_t\subseteq\Sigma$ of total
+operation symbols, and a function $\mathrm{Def}$ assigning to each operation
+symbol $\sigma:\Pi_{i\in a(\sigma)} s_i\to s$ in $\Sigma-\Sigma_t$ a set
+$\mathrm{Def}(\sigma)$ of $\Sigma_t$-equations in the standard variables
+$x_i\in V_s$, $i\in a(\sigma)$. Subsequently, they show that a category of models of an
+essentially algebraic theory is the same thing as a locally presentable 
+category.
+
+Every essentially algebraic theory $\mathbb{T}$ in the sense of Ad\'amek and Rosicky
+determines a syntactic category $\mathcal{C}_\mathbb{T}$, which has all finite
+limits, and $\mathbb{T}$ can be seen as a presentation of the category
+$\mathcal{C}_\mathbb{T}$. Then there is a natural equivalence between the
+category of models of $\mathbb{T}$ in a category $\cat{C}$ with finite limits,
+and the category of finite limit preserving functors $\mathcal{C}_\mathbb{T}\to
+\cat{C}$. Thus, in this sense an essentially algebraic theory \emph{is} just
+a category with finite limits.
+
+Because essentially algebraic theories have a pleasant theory of models, 
+phrasing a theory as an essentially algebraic one gives immediate access to
+a well-developed toolbox which can be used in its study. Therefore, recognizing
+that the theory of type dependency is essentially algebraic is a critical step
+in the development of its theory of models.
+
+\subsection*{Logical frameworks}
+
+\subsection*{The essentially algebraic theory of E-systems}
+
+\subsection*{Overview and conventions}
+In \autoref{sec:esys_defn}, we give the categorical definition of E-system in 
+several stages. First we define weakening systems, then projection systems and
+substitution systems, and finally we bundle them together into the definition
+of E-systems. The structure of the definition of weakening systems is 
+prototypcial: we first define the notion of pre-weakening systems, which are
+categories that come with a pre-weakening structure; then we define the notion
+of pre-weakening homomorphism, which is a functor preserving the weakening
+structure and we note that it makes sense to ask for the property that the
+functors of which the weakening system consists, are also pre-weakening
+homomorphisms; and finally, adding this requirement will be give the definition
+of weakening systems. We go through this process for each of the definitions
+of weakening systems, projection systems, substitution systems and E-systems.
+
+Throughout this section, our focus is on (set-valued) models of the essentially algebraic
+theory of type dependency, and in fact we shall not give the syntactic
+presentation of the theory. A syntactic presentation of type theory in the
+usual type theoretic style and models in arbitrary categories with finite limits
+have been presented by the author during the Homotopy Type Theory workshop in
+Oxford, November 2014%
+\footnote{See \url{http://www.qmac.ox.ac.uk/events/Talk\%20slides/Egbert\%20Rijke.pdf}}.
+
+In \autoref{sec:esys_props}, we derive several properties from the axioms of
+E-systems. This section is meant to give some insight in what can generally be
+done with E-systems, and provide some of the intuition behind the constructions
+in \autoref{sec:esys_compared}.
+
+Although we have introduced a notion of E-homomorphism, a better behaved category
+of E-systems is the one where E-homomorphisms lift factorizations. The condition
+that a functor lifts factorizations is also algebraic, so we still have an
+essentially algebraic theory of E-systems if we require this.
+In \autoref{sec:esys_compared}, where we introduce the category of E-systems
+with E-homomorphisms lifting factorizations, we construct an embedding from Voevodsky's
+B-systems into E-systems and prove that its image is the (full) subcategory
+of stratified E-systems.
+
+\begin{rmk}
+In the present article we use $\jdeq$ for equality, because the equality we
+consider here will interpret the judgmental equality of type theory, 
+and we wish to reserve
+the more convenient symbol $=$ for the internal equality of E-systems, i.e.~for
+the identification types, as in \cite{TheBook}. 
+\end{rmk}
+
+\begin{rmk}
+Recall that the slice pre-category $\cat{C}/\Gamma/(\Delta,f)$ is isomorphic to the slice pre-category
+$\cat{C}/\Delta$, in the essentially algebraic theory of pre-categories. In the following,
+we shall usually identify these two categories. Recall also that for every functor
+$F:\cat{C}\to\cat{D}$ and every object $\Delta$ of $\cat{C}$, there is a functor
+$F/\Delta:\cat{C}/\Delta\to \cat{D}/F(\Delta)$ given by the action on morphisms of $F$. 
+\end{rmk}
+
+\begin{rmk}
+It will be convenient to denote the objects of the category $\cat{C}$ with capital
+greek letters $\Gamma$, $\Delta$, \ldots, and the morphisms by capital latin letters
+$A$, $B$, $P$, $Q$, etcetera. When $A$ is a morphism into $\Gamma$, we will denote
+its codomain by $\ctxext{\Gamma}{A}$. When $A$ is a morphism into $\Gamma$ and
+$P$ is a morphism into $\ctxext{\Gamma}{A}$, it will be convenient to denote the
+composition of $P$ with $A$ by $\ctxext{A}{P}$ as well. Note that $\ctxext{A}{P}$
+is indeed the domain of $P$ in the slice category $\cat{C}/\Gamma$. Also, objects
+of $\cat{C}$ will always have a capital greek letter in their notation, e.g.~as
+in $\ctxext{{\Gamma}{A}}{P}$. 
+\end{rmk}
+
+\subsection*{Acknowledgements}
+Several people have been indispensible throughout the research process.
+I would like to thank 
+Richard Garner for the early discussions on the inference rules of type 
+dependency we had in Barcelona and in Paris; 
+Joachim Kock for the many fruitful discussions and for hosting me at the Universitat
+Aut\`onoma de Barcelona during the year 2013-2014;
+Steve Awodey for sharing many insights and useful advice which lead to the
+definition of an E-system in arbitrary categories with finite limits;
+and Vladimir Voevodsky for inviting me to present E-systems at the Homotopy Type
+Theory workshop in Oxford (November 2014) and to visit the Institute for Advanced 
+Study in Princeton (March 2015), where the definition presented here was
+discovered.
 
@@
 
+\begin{comment}
 \section{Introduction}
@@
 
+\part{The new E-systems}
+\end{comment}
+
+\input{weakeningcats}
+
 \bibliographystyle{plain}

### 2015-06-12 weakeningcats.tex
huge revision

@@
-\begin{rmk}
-Recall that the slice pre-category $\cat{C}/\Gamma/(\Delta,f)$ is isomorphic to the slice pre-category
-$\cat{C}/\Delta$, in the essentially algebraic theory of pre-categories. In the following,
-we shall usually identify these two categories. Recall also that for every functor
-$F:\cat{C}\to\cat{D}$ and every object $\Delta$ of $\cat{C}$, there is a functor
-$F/\Delta:\cat{C}/\Delta\to \cat{D}/F(\Delta)$ given by the action on morphisms of $F$. 
-\end{rmk}
-
-\begin{rmk}
-It will be convenient to denote the objects of the category $\cat{C}$ with capital
-greek letters $\Gamma$, $\Delta$, \ldots, and the morphisms by capital latin letters
-$A$, $B$, $P$, $Q$, etcetera. When $A$ is a morphism into $\Gamma$, we will denote
-its codomain by $\ctxext{\Gamma}{A}$. When $A$ is a morphism into $\Gamma$ and
-$P$ is a morphism into $\ctxext{\Gamma}{A}$, it will be convenient to denote the
-composition of $P$ with $A$ by $\ctxext{A}{P}$ as well. Note that $\ctxext{A}{P}$
-is indeed the domain of $P$ in the slice category $\cat{C}/\Gamma$. Also, objects
-of $\cat{C}$ will always have a capital greek letter in their notation, e.g.~as
-in $\ctxext{{\Gamma}{A}}{P}$. 
-\end{rmk}
-
-\section{Weakening systems}
+\section{The definition of E-systems}\label{sec:esys_defn}
+\subsection{Weakening systems}
 
@@
 \begin{enumerate}
-\item $W_{\catid{\Gamma}}=\catid{\cat{C}/\Gamma}$ for every object $\Gamma\in\cat{C}$.
-\item $W_{\ctxext{A}{P}}=W_P\circ W_A$ for every $P\in\cat{C}/\ctxext{\Gamma}{A}$ and $A\in\cat{C}/\Gamma$.
-\item $W_A$ preserves the final object, i.e.~$W_A(\catid{\Gamma})=\catid{\ctxext{\Gamma}{A}}$.
+\item $W_{\catid{\Gamma}}\jdeq \catid{\cat{C}/\Gamma}$ for every object $\Gamma\in\cat{C}$.
+\item $W_{\ctxext{A}{P}}\jdeq W_P\circ W_A$ for every $P\in\cat{C}/\ctxext{\Gamma}{A}$ and $A\in\cat{C}/\Gamma$.
+\item $W_A$ preserves the final object, i.e.~$W_A(\catid{\Gamma})\jdeq \catid{\ctxext{\Gamma}{A}}$.
 \end{enumerate}
@@
 the slice pre-category $\cat{C}/\Gamma$ is also a pre-weakening system, with the weakening
-structure given by $W(\cat{C}/\Gamma)_P=W(\cat{C})_P$ for each $P\in\cat{C}/\ctxext{\Gamma}{A}$
+structure given by $W(\cat{C}/\Gamma)_P\jdeq W(\cat{C})_P$ for each $P\in\cat{C}/\ctxext{\Gamma}{A}$
 and each $A\in\cat{C}/\Gamma$.
@@
 
-A useful special case of this property is where $B=\catid{\Gamma}$. Thus, if $W$ is
+A useful special case of this property is where $B\jdeq \catid{\Gamma}$. Thus, if $W$ is
 a weakening system, then the diagram
@@
 \begin{tikzcd}[column sep=huge]
-\cat{C}/\ctxext{\Gamma}{Q}
-  \arrow[r,"W_A/Q"]
-  &
-\cat{C}/\ctxext{{\Gamma}{A}}{W_A(Q)}
+\cat{C}/\ctxext{\Gamma}{C}
+  \arrow[r,"W_A/C"]
+  &
+\cat{C}/\ctxext{{\Gamma}{A}}{W_A(C)}
   \\
 \cat{C}/\Gamma
-  \arrow[u,"W_Q"]
+  \arrow[u,"W_C"]
   \arrow[r,swap,"W_A"]
@@
 \cat{C}/\ctxext{\Gamma}{A}
-  \arrow[u,swap,"W_{W_A(Q)}"]
-\end{tikzcd}
-\end{equation*}
-commutes for every $A\in\cat{C}/\Gamma$ and $Q\in\cat{C}/\Gamma$. 
+  \arrow[u,swap,"W_{W_A(C)}"]
+\end{tikzcd}
+\end{equation*}
+commutes for every $A,C\in\cat{C}/\Gamma$. In particular, we see
+that $W_A(W_C(D))\jdeq W_{W_A(C)}(W_A(D))$ for any $D\in\cat{C}/\Gamma$,
+i.e.~that weakening is a self-distributive operation. 
 \end{rmk}
@@
 obtain the following form: There is a group homomorphism $W_x:G\to G$ for
-every $x\in G$, in such a way that $W_e=\catid{G}$ and $W_{xy}=W_y\circ W_x$
+every $x\in G$, in such a way that $W_e\jdeq \catid{G}$ and $W_{xy}\jdeq W_y\circ W_x$
 for every $x,y\in G$, and for any $x,y\in G$ the diagram
@@
 is required to commute. Thus, a weakening system on a group $G$ consists of
-a group homomorphism $W:\op G\to\mathrm{Aut}(G)$ such that $W_a\circ W_b=
-W_{W_a(b)}\circ W_a$. The latter requirement is satisfied if $b\cdot a=a\cdot W_a(b)$
-for every $a,b\in G$, or equivalently if $[b,a]=b^{-1}\cdot W_a(b)$.
+a group homomorphism $W:\op G\to\mathrm{Aut}(G)$ such that $W_a\circ W_b\jdeq 
+W_{W_a(b)}\circ W_a$. The latter requirement is satisfied if $b\cdot a\jdeq a\cdot W_a(b)$
+for every $a,b\in G$, or equivalently if $[b,a]\jdeq b^{-1}\cdot W_a(b)$.
 \end{eg}
@@
 \begin{proof}
-Note that $e^{-1}xe=x$ and $(xy)^{-1}z(xy)=y^{-1}(x^{-1}zx)y$ for any $x,y,z\in G$.
+Note that $e^{-1}xe\jdeq x$ and $(xy)^{-1}z(xy)\jdeq y^{-1}(x^{-1}zx)y$ for any $x,y,z\in G$.
 Also, the diagram
@@
 
+\begin{comment}
 \begin{lem}
@@
 $W_e:\cat{C}_G/\Gamma\to\cat{C}_G/\Delta$ for every edge $e:\Delta\to \Gamma$ in $G$, with the
-property that $W_e(\catid{\Gamma})=\catid{\Delta}$.
+property that $W_e(\catid{\Gamma})\jdeq \catid{\Delta}$.
 
@@
 Let $G$ be the graph consisting of one vertex $\ast$ and one edge $e:\ast\to\ast$.
-Then $\cat{C}_G/\ast=\N$. Then the only weakening structure on $\cat{C}/G$ is
-given by $W_e=\catid{\N}$. The reason is that 
+Then $\cat{C}_G/\ast\jdeq \N$. Then the only weakening structure on $\cat{C}/G$ is
+given by $W_e\jdeq \catid{\N}$. The reason is that 
 \end{eg}
+\end{comment}
 
@@
 preserving maps $W_{a,b}:{\downarrow}(b)\to{\downarrow}(a)$ satisfying
-$W_{a,b}(b)=a$, for every $a\leq b$, where
+$W_{a,b}(b)\jdeq a$, for every $a\leq b$, where
 ${\downarrow}(x)$ is the set $\{y\in P\mid y\leq x\}$.
@@
 In a poset with meets, the functions $W_{a,b}:{\downarrow}(b)\to{\downarrow}(a)$
-given by $W_{a,b}(x):= a\land x$ assemble a weakening structure, because for
-any $a\leq b$ and $x,c'\leq c\leq b$ we have $a\land(c'\land x)=(a\land c')\land
+given by $W_{a,b}(x)\defeq a\land x$ assemble a weakening structure, because for
+any $a\leq b$ and $x,c'\leq c\leq b$ we have $a\land(c'\land x)\jdeq (a\land c')\land
 (a\land x)$.  
@@
 Consider the poset $(\N,\geq)$, which has a morphism from $a$ to $b$ if $a\geq b$.
-On this poset, we may consider the weakening structure defined by $W_{a,b}(x)=
+On this poset, we may consider the weakening structure defined by $W_{a,b}(x)\jdeq
 a+(x-b)$.
@@
 
-\section{Weakening systems with term structure}
+\subsection{Weakening systems with term structure}
 \begin{defn}
@@
 Suppose $\cat{C}$ is a pre-category with term structure, and $\Gamma$ is an object of $\cat{C}$. 
-Then $\cat{C}/\Gamma$ is a pre-category with term structure, where $T_{\cat{C}/\Gamma}(A)=T_\cat{C}(A)$.
+Then $\cat{C}/\Gamma$ is a pre-category with term structure, where $T_{\cat{C}/\Gamma}(A)\jdeq T_\cat{C}(A)$.
 \end{defn}
@@
 structure on $G$ is given by a set $X$ with a right $G$-action $\mu:\op G\times X\to X$.
-We take $T(a):= X$ for every $a\in G$.
+We take $T(a)\defeq X$ for every $a\in G$.
 \end{eg}
 
-\section{Projection systems}
-\begin{defn}
-A \emph{pre-projection system} is a weakening system with term structure with an element
+\subsection{Projection systems}
+\begin{defn}
+A \emph{pre-projection system} is a pre-weakening system with term structure with an element
 $\tfid{A}\in T(W_A(A))$ for every $A\in\cat{C}/\Gamma$ and $\Gamma\in\cat{C}$. 
@@
 \begin{defn}
-A \emph{pre-projection homomorphism} $F:\cat{C}\to\cat{D}$ is a weakening homomorphism for which
-\begin{equation*}
-F(\tfid{A})=\tfid{F(A)}
+A \emph{pre-projection homomorphism} $F:\cat{C}\to\cat{D}$ is a pre-weakening homomorphism for which
+\begin{equation*}
+F(\tfid{A})\jdeq \tfid{F(A)}
 \end{equation*}
@@
 
+\begin{proof}
+Straightforward.
+\end{proof}
+
 \begin{defn}
@@
 
-\begin{eg}
-Consider a group $G$ with a right $G$-set $X$. A projection structure on $(G,X)$
-consists of an element $\tfid{a}\in X$ for every $a\in G$, so that $\tfid{a}\cdot b=
-\tfid{b^{-1}ab}$ for every $a,b\in G$.
-
+%\begin{eg}
+%Consider a group $G$ with a right $G$-set $X$. A projection structure on $(G,X)$
+%consists of an element $\tfid{a}\in X$ for every $a\in G$, so that $\tfid{a}\cdot b=
+%\tfid{b^{-1}ab}$ for every $a,b\in G$. 
 %For instance, this is the case if we take $X$ to be a pointed set with $\ast\in X$,
 %and take $\tfid{a}:=\ast$ for all $a\in G$.
-\end{eg}
-
-\section{Substitution systems}
+%\end{eg}
+
+\subsection{Substitution systems}
 
@@
 of a functor with term structure $S_x:\cat{C}/\ctxext{\Gamma}{A}\to\cat{C}/\Gamma$ for every $x\in T(A)$ and
-$A\in\cat{C}/\Gamma$, such that $S_x(\catid{\ctxext{\Gamma}{A}})=\catid{\Gamma}$.
+$A\in\cat{C}/\Gamma$, such that $S_x(\catid{\ctxext{\Gamma}{A}})\jdeq\catid{\Gamma}$.
 
@@
 \begin{lem}
-For any object $\Delta$ of a pre-substitution system $\cat{C}$, the slice pre-category
-$\cat{C}/\Delta$ is a pre-substitution system.
+For any object $\Gamma$ of a pre-substitution system $\cat{C}$, the slice pre-category
+$\cat{C}/\Gamma	$ is a pre-substitution system.
 \end{lem}
@@
 \begin{cor}
-For any object $\Delta$ of a substitution system, the slice pre-category $\cat{C}/\Delta$
+For any object $\Gamma$ of a substitution system, the slice pre-category $\cat{C}/\Gamma$
 is a substitution system.
@@
 Let $G$ be a group and let $H$ be a set, forming a category with term 
-structure of which $T(a)=H$ for every $a\in G$. A substitution
+structure of which $T(a)\jdeq H$ for every $a\in G$. A substitution
 system on $(G,H)$ consists of a function $\sigma_G : H\times G\to G$ and
@@
 
-\section{E-systems}
+\subsection{E-systems}
+
+Although it is now possible to state the definition of an E-system without
+any difficulty, it shall be convenient later to also have a category of
+pre-E-systems at our disposal. Pre-E-systems shall by default be non-unital,
+i.e.~they do not come with a pre-projection system, while E-systems shall
+always be assumed to be unital. In the present context, we shall not need
+unital pre-E-systems or non-unital E-systems.
+
+\begin{defn}
+A \emph{pre-E-system} is a pre-category $\cat{C}$ with term structure
+with a pre-weakening structure $W$, a pre-substitution structure $S$ and a terminal
+object. A \emph{unital pre-E-system}, or simply \emph{pre-E-system}, is a non-unital pre-E-system
+which also has the structure of a pre-projection system. 
+\end{defn}
+
+\begin{defn}
+A \emph{pre-E-homomorphism} from $\mathbb{E}\to\mathbb{D}$ is a functor
+$H:\mathbb{C}_\mathbb{E}\to\mathbb{C}_\mathbb{D}$ between the underlying categories
+with term structure, which is a pre-weakening homomorphism, a pre-substitution
+homomorphism, and which preserves the terminal object.
+
+A \emph{unital} pre-E-homomorphism is a non-unital pre-E-homomorphism which is 
+also a pre-projection homomorphism. 
+\end{defn}
+
 \begin{defn}
 \label{defn:esystems}
-An \emph{E-system} is a pre-category $\cat{C}$ with term structure with
-\begin{itemize}
-\item a weakening structure $W$,
-\item the structure $\tfid{}$ of a projection system,
-\item a substitution structure $S$,
-\end{itemize}
-such that
+An \emph{E-system} is a unital pre-E-system in which 
 \begin{enumerate}
-\item each $W_A$ is a substitution homomorphism, i.e.~the diagram
+\item each $W_A$ is a unital pre-E-homomorphism,
+\item each $S_x$ is a unital pre-E-homomorphism,
+\item \label{tTf:StWf_id} weakened families are constant, i.e.~$S_x\circ W_A\jdeq \catid{\cat{C}/\Gamma}$ for any $x\in T(A)$ and $A\in\cat{C}/\Gamma$,
+\item $S_x(\tfid{A})\jdeq x$ for any $x\in T(A)$ and $A\in\cat{C}/\Gamma$, and
+\item \label{f:SidfWff_id} $S_{\tfid{A}}\circ W_A/A\jdeq \catid{\cat{C}/\ctxext{\Gamma}{A}}$ for any $A\in\cat{C}/\Gamma$.
+\end{enumerate}
+An \emph{E-homomorphism} $H:\mathbb{E}\to\mathbb{D}$ is a functor from
+$\cat{C}_{\mathbb{E}}$ to $\cat{C}_{\mathbb{D}}$ which is both a projection
+homomorphism and a substitution homomorphism.
+\end{defn}
+
+\begin{rmk}
+The condition that each $W_A$ is a substitution homomorphism asserts that
+the diagram
 \begin{equation*}
@@
 \end{equation*}
-commutes for every $Q\in\cat{C}/\ctxext{\Gamma}{B}$, $B\in\cat{C}/\Gamma$ and each 
-$y\in T(Q)$; 
-\item each $S_x$ is a projection homomorphism, i.e.~the diagram
+of functors with term structure commutes for every 
+$Q\in\cat{C}/\ctxext{\Gamma}{B}$, $B\in\cat{C}/\Gamma$ and each $y\in T(Q)$.
+
+Likewise, the condition that each $S_x$ is a weakening homomorphism
+asserts that the diagram
 \begin{equation*}
@@
 \end{equation*}
-commutes for every $Q\in\cat{C}/\ctxext{{\Gamma}{A}}{P}$, and we have
-\begin{equation*}
-S_x(\tfid{Q})=\tfid{S_x(Q)}
-\end{equation*}
-for every $Q\in\cat{C}/\ctxext{{\Gamma}{A}}{P}$;
-\item $S_x(\tfid{A})=x$ for any $x\in T(A)$ and $A\in\cat{C}/\Gamma$;
-\item \label{tTf:StWf_id} $S_x\circ W_A=\catid{\cat{C}/\Gamma}$ for any $x\in T(A)$ and $A\in\cat{C}/\Gamma$;
-\item \label{f:SidfWff_id} $S_{\tfid{A}}\circ W_A/A=\catid{\cat{C}/\ctxext{\Gamma}{A}}$ for any $A\in\cat{C}/\Gamma$.
-\end{enumerate}
-An \emph{E-homomorphism} $F:\cat{C}\to\cat{D}$ is a functor which is both a projection
-homomorphism and a substitution homomorphism.
-\end{defn}
+of functors with term structure commutes for every 
+$Q\in\cat{C}/\ctxext{{\Gamma}{A}}{P}$.
+\end{rmk}
 
 \begin{cor}
-For any object $\Delta$ of an E-system $\cat{C}$, the slice pre-category $\cat{C}/\Delta$
+For any object $\Gamma$ of an E-system $\cat{C}$, the slice pre-category $\cat{C}/\Gamma$
 is an E-system.
@@
 
+\begin{comment}
 \begin{rmk}[Needs more precision]
@@
 \end{rmk}
-
-\begin{defn}
-An E-system of which the underlying category has a terminal object $\emptyc$, is said to
-be \emph{E-system with empty context}. An E-homomorphism is said to preserve the
-empty context when it preserves the terminal object $\emptyc$ on the nose.
+\end{comment}
+
+\section{Properties of E-systems}\label{sec:esys_props}
+
+The theory of E-systems could be seen as a sort of category theory with only
+projections and their sections. We elaborate on this point of view by defining
+the category of internal homomorphisms, which Lawvere might have called the
+category of terms. 
+In the category of internal morphisms over a context $\Gamma$,
+a morphism $f:A\to B$ is a term of the constant family $W_A(B)$ over 
+$\ctxext{\Gamma}{A}$. Rather than directly defining composition, we will define
+for every $f:A\to B$ in context $\Gamma$, a pre-composition operation
+$f_\ast$, which is going to be an E-homomorphism from 
+$\cat{C}/\ctxext{\Gamma}{B}$ to $\cat{C}/\ctxext{\Gamma}{A}$. This allows us
+to consider diagrams of the following form:
+\begin{equation*}
+\begin{tikzcd}
+P \arrow[d,->>] \arrow[r,"g"] & Q \arrow[d,->>] \\
+A \arrow[r,swap,"f"] & B
+\end{tikzcd}
+\end{equation*}
+where we consider the morphism $g$ over $f$ to be an internal morphism from 
+$P$ to $f_\ast(Q)$ in context $\ctxext{\Gamma}{A}$, capturing the
+essence of being a morphism from the family $P$ over $A$ to the family
+$Q$ over $B$, reindexed by $f$. 
+Whenever we draw such diagrams, a double headed arrow will always
+refer to a family, and an ordinary arrow will refer to an internal morphism,
+possibly over another internal morphism
+
+A morphism $g$ over $f$ can be vertically composed with $f$ so
+that we obtain a morphism $\jvcomp{P}{f}{g}:\ctxext{A}{P}\to\ctxext{B}{Q}$ in 
+context $\Gamma$, and it can be horizontally composed with a morphism $k:Q\to R$
+over $h:B\to C$ so that we obtain a morphism $\jfcomp{A}{f}{P}{g}{k}:P\to R$ 
+over $\jcomp{A}{f}{h}:A\to C$. We will show interchange laws between these
+sorts of composition. The interchange laws have as a consequence that if we
+assert the commutativity of a large composed diagram.
+
+We then demonstrate how to define a category internal to an E-system. The main
+point of this lies mainly in demonstrating how the operations of E-systems can
+be used to define a category, because it is no surprise that a category can be
+defined in type theory. Nevertheless, it is noteworthy that we need no further 
+conditions on E-systems.
+This is unlike the situation in category theory, where we need
+to assume that the ambient category has finite limits in order to say what an
+internal category is. In fact, it is possible to define an E-system internal
+to an E-system, just using the algebraic structure of an E-system, and actually
+defining an internal category was the hardest part of that definition.
+
+Before we begin this section, we introduce some handy shorthands.
+
+\begin{defn}
+Let $A\in\cat{C}/\Gamma$. Recall that $W_A:\cat{C}/\Gamma\to\cat{C}/\ctxext{\Gamma}{A}$ acts on objects,
+morphisms and terms. We introduce the infix form of weakening by $A\in\cat{C}/\Gamma$ to be
+$\ctxwk{A}{\blank}$. Thus, we will write
+\begin{align*}
+\ctxwk{A}{B} & \defeq W_A(B) & & \text{for $B\in\cat{C}/\Gamma$} \\
+\ctxwk{A}{Q} & \defeq W_A(Q) & & \text{for $B\in\cat{C}/\Gamma$ and $Q\in\cat{C}/\ctxext{\Gamma}{B}$}\\
+\ctxwk{A}{g} & \defeq W_A(g) & & \text{for $B\in\cat{C}/\Gamma$, $Q\in\cat{C}/\ctxext{\Gamma}{B}$ and $g\in T(Q)$}
+\end{align*}
+\end{defn}
+
+\begin{defn}
+Let $x\in T(A)$ for a family $A\in\cat{C}/\Gamma$. The infix form of substitution
+by $x$ is taken to be $\subst{x}{\blank}$. Thus, we will write
+\begin{align*}
+\subst{x}{P} & \defeq S_x(P) & & \text{for $P\in\cat{C}/\ctxext{\Gamma}{A}$} \\
+\subst{x}{Q} & \defeq S_x(Q) & & \text{for $P\in\cat{C}/\ctxext{\Gamma}{A}$ and $Q\in\cat{C}/\ctxext{{\Gamma}{A}}{P}$} \\
+\subst{x}{g} & \defeq S_x(g) & & \text{for $P\in\cat{C}/\ctxext{\Gamma}{A}$, $Q\in\cat{C}/\ctxext{{\Gamma}{A}}{P}$ and $g\in T(Q)$}
+\end{align*}
+\end{defn}
+
+\subsection{The category of internal morphisms}\label{subsec:im_cat}
+
+\begin{defn}
+For every $A,B\in\cat{C}/\Gamma$ we define the set 
+\begin{equation*}
+\thom{A}{B}\defeq T(\ctxwk{A}{B}).
+\end{equation*}
+An element $f\in \thom{A}{B}$ is called an \emph{internal morphism in context
+$\Gamma$}. We sometimes write $\jhom{\Gamma}{A}{B}{f}$ to indicate that $f$ is
+an internal morphism over $\Gamma$, or we may draw a diagram of the form
+\begin{equation*}
+\begin{tikzcd}[column sep=tiny]
+A \arrow[rr,"f"] \arrow[dr,fib] & & B \arrow[dl,fib] \\
+& \Gamma
+\end{tikzcd}
+\end{equation*}
+but we shall generally omit the arrows down to $\Gamma$ and say instead that we have
+a diagram in context $\Gamma$. 
 \end{defn}
@@
 \begin{rmk}
-When there is a terminal object, we get a weakening functor $W_\Gamma:\cat{C}\to
-\cat{C}/\Gamma$ for every object $\Gamma$ of $\cat{C}$; we get a set $T(\Gamma)$;
-and for $i\in T(\Gamma)$ we get substitutions $S_i:\cat{C}/\Gamma\to\cat{C}$. 
-\end{rmk}
-
-\section{Properties of E-systems}
-
-\subsection{The category of internal functions}
-
-\begin{defn}
-For every $A,B\in\cat{C}/\Gamma$ we define the set 
-\begin{equation*}
-\thom{A}{B}:=T(W_A(B)).
-\end{equation*} 
-\end{defn}
-
-\begin{rmk}
-Note that $\thom{\catid{\Gamma}}{A}=T(A)$ for any $A\in\cat{C}/\Gamma$, 
-because we have $W_{\catid{\ctxext{\Gamma}{A}}}=\catid{\cat{C}/\ctxext{\Gamma}{A}}$.
-
-Note also that $\thom{\ctxext{A}{P}}{B}=\thom{P}{W_A(B)}$ for any $P\in\cat{C}/\ctxext{\Gamma}{A}$
-and $B\in\cat{C}/\Gamma$, because $W_{\ctxext{A}{P}}=
-W_P\circ W_A$.
+Note that $\thom{\catid{\Gamma}}{A}\jdeq T(A)$ for any $A\in\cat{C}/\Gamma$, 
+because we have 
+$W_{\catid{\ctxext{\Gamma}{A}}}\jdeq \catid{\cat{C}/\ctxext{\Gamma}{A}}$.
+
+Note also that $\thom{\ctxext{A}{P}}{B}\jdeq \thom{P}{\ctxwk{A}{B}}$ 
+for any $P\in\cat{C}/\ctxext{\Gamma}{A}$ and $B\in\cat{C}/\Gamma$,
+because $W_{\ctxext{A}{P}}\jdeq W_P\circ W_A$.
+Once we have established a category of which the morphisms are given by 
+$\thom{\blank}{\blank}$, we therefore get that 
+\begin{equation*}
+\ctxext{A}{(\blank)}\dashv W_A.
+\end{equation*}
+The right adjoint to weakening by $A$, if it exists, will be the dependent
+product $\Pi_A$. 
 \end{rmk}
@@
 \begin{equation*}
-f_\ast := S_f\circ W_A/B : \cat{C}/\ctxext{\Gamma}{B}\to\cat{C}/\ctxext{\Gamma}{A}.
+f_\ast \defeq S_f\circ W_A/B : \cat{C}/\ctxext{\Gamma}{B}\to\cat{C}/\ctxext{\Gamma}{A}.
 \end{equation*}
 Furthermore, for $g\in\thom{B}{C}$ we may consider $W_B(C)\in\cat{C}/\ctxext{\Gamma}{B}$ as a morphism from $W_B(C)$ to $\catid{\ctxext{\Gamma}{B}}$. Therefore, $f_\ast$ acts on the terms of
-$W_B(C)$, so it makes sense to define $g\circ f:=f_\ast(g)$. 
-\end{defn}
-
-Note that $g\circ f\in T(S_f(W_A/B(W_B(C))))$, whereas we would like that
-$g\circ f\in\thom{A}{C}$. The following lemma confirms this:
-
-\begin{lem}
-Suppose $A$ and $B$ are morphisms of the underlying pre-category of $\mathcal{E}$
-with the same codomain, and let $f\in\thom{A}{B}$. Then we have 
-\begin{equation*}
-S_f\circ (W_A/B)\circ W_B=W_A.
-\end{equation*}
-\end{lem}
-
-\begin{proof}
+$W_B(C)$, so it makes sense to define $g\circ f\defeq f_\ast(g)$. In fact, it makes
+sense to let $(\blank)\circ f$ be the infix notation for $f_\ast$. 
+\end{defn}
+
+\begin{rmk}
+Note that for $\jhom{\Gamma}{B}{C}{g}$, 
+we have $g\circ f\in T(S_f(W_A/B(W_B(C))))$, whereas we would like that
+$g\circ f\in\thom{A}{C}$. More generally, we can show that
+\begin{equation*}
+S_f\circ (W_A/B)\circ W_B\jdeq W_A.
+\end{equation*}
 Since weakening is a weakening homomorphism, we have
 \begin{equation*}
-S_f\circ (W_A/B)\circ W_B=S_f\circ W_{W_A(B)}\circ W_A
+S_f\circ (W_A/B)\circ W_B\jdeq S_f\circ W_{W_A(B)}\circ W_A
 \end{equation*}
@@
 \begin{equation*}
-S_f\circ W_{W_A(B)}\circ W_A=W_A.\qedhere
-\end{equation*}
-\end{proof}
+S_f\circ W_{W_A(B)}\circ W_A\jdeq W_A.
+\end{equation*}
+\end{rmk}
+
+\begin{rmk}
+To see how this extended notion of precomposition works, suppose we have a diagram
+\begin{equation*}
+\begin{tikzcd}[column sep=tiny]
+Q \arrow[rr,"g"] \arrow[dr,fib] & & R \arrow[dl,fib] \\
+& B
+\end{tikzcd}
+\end{equation*}
+in context $\Gamma$, i.e.~$\jhom{\ctxext{\Gamma}{B}}{Q}{R}{g}$, and let $\jhom{\Gamma}{A}{B}{f}$. Precomposing by $f$ gives
+us a diagram
+\begin{equation*}
+\begin{tikzcd}[column sep=tiny]
+\jcomp{A}{f}{Q} \arrow[rr,"\jcomp{A}{f}{g}"] \arrow[dr,fib] & & \jcomp{A}{f}{R} \arrow[dl,fib] \\
+& A
+\end{tikzcd}
+\end{equation*}
+in context $\Gamma$.
+\end{rmk}
 
@@
 Note that \autoref{f:SidfWff_id} of \autoref{defn:esystems} asserts precisely
-that ${(\tfid{A})}_\ast=\catid{\cat{C}/\Delta}$ for any $A:\Delta\to \Gamma$ in $\cat{C}$. In
-particular, it follows that $g\circ\tfid{A}=g$ for any $g\in\thom{A}{B}$
+that ${(\tfid{A})}_\ast\jdeq \catid{\cat{C}/\ctxext{\Gamma}{A}}$ for any $A\in\cat{C}/\Gamma$. In
+particular, it follows that $g\circ\tfid{A}\jdeq g$ for any $g\in\thom{A}{B}$
 \end{rmk}
 
-\begin{lem}
-For any $f\in\thom{A}{B}$ we have $\tfid{B}\circ f=f$.
+\begin{lem}\label{lem:compcomp}
+For any $f\in\thom{A}{B}$ and $g\in\thom{B}{C}$ we have $f_\ast\circ g_\ast\jdeq (g\circ f)_\ast$. 
 \end{lem}
@@
 \begin{proof}
-We have
-\begin{equation*}
-\tfid{B}\circ f=S_f(W_A(\tfid{B}))=S_f(\tfid{W_A(B)})=f.\qedhere
-\end{equation*}
-\end{proof}
-
-\begin{lem}
-For any $f\in\thom{A}{B}$ and $g\in\thom{B}{C}$ we have $f_\ast\circ g_\ast=(g\circ f)_\ast$. 
-\end{lem}
-
-\begin{proof}
-\begin{align*}
-f_\ast\circ g_\ast & = S_f\circ (W_A/B) \circ S_g \circ (W_B/C)\\
-& = S_f\circ S_{W_A(g)}\circ(W_A/\ctxext{B}{W_B(C)})\circ (W_B/C) \\
-& = S_{S_f(W_A(g))}\circ (S_f/W_A(W_B(C)))\circ (W_A/\ctxext{B}{W_B(C)}) \circ W_B/C \\
-& = S_{S_f(W_A(g))}\circ ((S_f\circ (W_A/B) \circ W_B)/C) \\
-& = S_{S_f(W_A(g))}\circ ((S_f\circ W_{W_A(B)}\circ W_A)/C) \\
-& = S_{S_f(W_A(g))}\circ W_A/C \\
-& = (g\circ f)_\ast.\qedhere
-\end{align*}
-\end{proof}
-
-\begin{cor}
-Composition is associative. Thus, we have constructed a pre-category for each
-object $\Gamma$, with the same objects as $\cat{C}/\Gamma$, of which the hom-sets
-are the sets $\thom{A}{B}$. 
-\end{cor}
+\begin{align*}
+f_\ast\circ g_\ast & \jdeq S_f\circ (W_A/B) \circ S_g \circ (W_B/C)\\
+& \jdeq S_f\circ S_{W_A(g)}\circ(W_A/\ctxext{B}{W_B(C)})\circ (W_B/C) \\
+& \jdeq S_{S_f(W_A(g))}\circ (S_f/W_A(W_B(C)))\circ (W_A/\ctxext{B}{W_B(C)}) \circ W_B/C \\
+& \jdeq S_{S_f(W_A(g))}\circ ((S_f\circ (W_A/B) \circ W_B)/C) \\
+& \jdeq S_{S_f(W_A(g))}\circ ((S_f\circ W_{W_A(B)}\circ W_A)/C) \\
+& \jdeq S_{S_f(W_A(g))}\circ W_A/C \\
+& \jdeq (g\circ f)_\ast.\qedhere
+\end{align*}
+\end{proof}
+
+\begin{thm}
+For each object $\Gamma$ of an E-system $\mathbb{E}$, we have a category
+$\mathbb{C}_I(\mathbb{E},\Gamma)$ of internal morphisms of $\mathbb{E}$,
+consisting of the same objects as $\cat{C}/\Gamma$ and the hom-sets 
+$\thom{A}{B}$. 
+\end{thm}
+
+\begin{proof}
+The fact that composition is associative is a direct corollary of
+\autoref{lem:compcomp}. The axiom $(\tfid{A})_\ast\jdeq\catid{\ctxext{\Gamma}{A}}$
+implies that the identity morphisms satisfy the right identity law. It remains
+to show that $\tfid{B}\circ f\jdeq f$. This is a simple calculation:
+\begin{equation*}
+\tfid{B}\circ f 
+  \jdeq
+\subst{f}{\ctxwk{A}{\tfid{B}}}
+  \jdeq
+\subst{f}{\tfid{\ctxwk{A}{B}}}
+  \jdeq
+f.\qedhere
+\end{equation*}
+\end{proof}
 
@@
 \end{equation*}
-commutes. This applies to weakening, substitution and pre-composition.
+commutes. This applies to weakening, substitution and pre-composition, and it
+follows that for any E-homomorphism $F:\mathcal{E}\to\mathcal{D}$ and any
+$\Gamma\in\mathcal{E}$, we have a functor $\mathbb{C}_I(F,\Gamma):
+\mathbb{C}_I(\mathcal{E},\Gamma)\to\mathbb{C}_I(\mathcal{D},F(\Gamma))$.
 \end{rmk}
 
-\begin{cor}
-Suppose $F:\mathcal{E}\to\mathcal{D}$ is an E-homomorphism. Then $F$ induces
-a functor $\mathbb{T}_F:\mathbb{T}_\mathcal{E}\to \mathbb{T}_\mathcal{D}$. 
-\end{cor}
+The following fact will be useful:
+
+\begin{lem}\label{lem:compW_W}
+Let $f\in\thom{A}{B}$ be an internal morphism in context $\Gamma$. Then one has
+\begin{equation*}
+f_\ast\circ W_B\jdeq W_A.
+\end{equation*}
+\end{lem}
+
+\begin{proof}
+The proof is a simple calculation:
+\begin{equation*}
+f_\ast\circ W_B\jdeq S_f\circ W_A/B\circ W_B\jdeq S_f\circ W_{W_A(B)}\circ W_A
+\jdeq W_A.\qedhere
+\end{equation*}
+\end{proof}
 
@@
 
+In this section we will make more extensive use of the infix form of the
+weakening and substitution operations.
+
 \begin{defn}
 Let $x\in T(A)$ and $u\in T(S_x(P))$ for $A\in\cat{C}/\Gamma$ and $P\in\cat{C}/\ctxext{\Gamma}{A}$. 
-We define
-\begin{equation*}
-\tmext{x}{u}\defeq S_u(S_x(\tfid{\ctxext{A}{P}}))\in T(\ctxext{A}{P}).
-\end{equation*}
-\end{defn}
-
-\begin{thm}
+We define the \emph{term extension of $x$ and $u$} to be
+\begin{equation*}
+\tmext{x}{u}\defeq \subst{u}{{x}{\tfid{\ctxext{A}{P}}}}\in T(\ctxext{A}{P}).
+\end{equation*}
+\end{defn}
+
+To prove anything about the term $\tmext{x}{u}$, we need the following property.
+
+\begin{thm}\label{subst_by_tmext}
 Let $x\in T(A)$ and $u\in T(S_x(P))$ for $A\in\cat{C}/\Gamma$ and $P\in\cat{C}/\ctxext{\Gamma}{A}$.
@@
 \begin{equation*}
-S_{\tmext{x}{u}}=S_u\circ (S_x/P):\cat{C}/\ctxext{\Gamma}{{A}{P}}\to\cat{C}/\Gamma
+S_{\tmext{x}{u}}\jdeq S_u\circ (S_x/P):\cat{C}/\ctxext{\Gamma}{{A}{P}}\to\cat{C}/\Gamma
 \end{equation*}
@@
 
+\begin{proof}
+\begin{align*}
+S_{\tmext{x}{u}} & \jdeq S_{S_u(S_x(\tfid{\ctxext{A}{P}}))} \\
+& \jdeq S_{S_u(S_x(\tfid{\ctxext{A}{P}}))}\circ (S_u\circ W_{S_x(P)})\circ (S_x\circ W_{A}) \\
+& \jdeq S_u \circ S_{S_x(\tfid{\ctxext{A}{P}})} \circ W_{S_x(P)}\circ S_x\circ W_A \\
+& \jdeq S_u \circ (S_x/P) \circ S_{\tfid{\ctxext{A}{P}}} \circ W_P\circ W_A \\
+& \jdeq S_u \circ (S_x/P) \circ S_{\tfid{\ctxext{A}{P}}} \circ W_{\ctxext{A}{P}}\\
+& \jdeq S_u \circ (S_x/P).\qedhere
+\end{align*}
+\end{proof}
+
 \begin{defn}
@@
 \begin{align*}
-\cprojfstf{A}{P} & \defeq W_P(\tfid{A})\in\thom{\ctxext{A}{P}}{A}\\
+\cprojfstf{A}{P} & \defeq \ctxwk{P}{\tfid{A}}\in\thom{\ctxext{A}{P}}{A}\\
 \cprojsndf{A}{P} & \defeq \tfid{P}\in\thomd{\cprojfstf{A}{P}}{\ctxext{A}{P}}{P}
@@
 
-\begin{cor}
+\begin{cor}\label{cor:pairing_ehom}
 Let $F:\mathcal{E}\to\mathcal{D}$ be an E-homomorphism and let $x\in T(A)$ and
@@
 
-\begin{lem}
-Let $x\in T(A)$ and $u\in T(S_x(P))$. then we have
-\begin{align*}
-S_{\tmext{x}{u}}(\cprojfstf{A}{P}) & \jdeq x\\
-S_{\tmext{x}{u}}(\cprojsndf{A}{P}) & \jdeq u.
-\end{align*}
-Also, for $w\in T(\ctxext{A}{P})$ we have
-\begin{equation*}
-\tmext{S_w(\cprojfstf{A}{P})}{S_w(\cprojsndf{A}{P})} \jdeq w.
-\end{equation*}
-\end{lem}
-
-\begin{lem}
+\begin{thm}\label{thm:famext_up}
+For any $A\in\cat{C}/\Gamma$ and $P\in\cat{C}/\ctxext{\Gamma}{A}$,
+the map
+\begin{equation*}
+(x,u)\mapsto \tmext{x}{u}:\big(\bigsqcup\nolimits_{x\in T(A)} T(\subst{x}{P})\big)\to T(\ctxext{A}{P}).
+\end{equation*}
+is an isomorphism.  
+\end{thm}
+
+\begin{proof}
+Lets write $\varphi$ for the indicated map. The inverse $\psi$ of $\varphi$ is defined
+by $w\mapsto (\subst{w}{\cprojfstf{A}{P}},\subst{w}{\cprojsndf{A}{P}})$. We have
+to show that for any $x\in T(A)$ and $u\in T(\subst{x}{P})$, one has
+\begin{align*}
+\subst{\tmext{x}{u}}{\cprojfstf{A}{P}} & \jdeq x\\
+\subst{\tmext{x}{u}}{\cprojsndf{A}{P}} & \jdeq u,
+\end{align*}
+and that for any $w\in T(\ctxext{A}{P})$, one has
+\begin{equation*}
+\tmext{\subst{w}{\cprojfstf{A}{P}}}{\subst{w}{\cprojsndf{A}{P}}} \jdeq w.
+\end{equation*}
+To show that $\subst{\tmext{x}{u}}{\cprojfstf{A}{P}} \jdeq x$, we use that
+$S_{\tmext{x}{u}}\jdeq S_u\circ S_x/P$ to show that
+\begin{align*}
+\subst{\tmext{x}{u}}{\cprojfstf{A}{P}}
+  & \jdeq 
+\subst{\tmext{x}{u}}{\ctxwk{P}{\tfid{A}}} \\
+  & \jdeq
+\subst{u}{{x}{\ctxwk{P}{\tfid{A}}}} \tag{By \autoref{subst_by_tmext}}\\
+  & \jdeq
+\subst{u}{\ctxwk{\subst{x}{P}}{\subst{x}{\tfid{A}}}} \\
+  & \jdeq
+\subst{x}{\tfid{A}} \\
+  & \jdeq
+x
+\end{align*}
+To show that $\subst{\tmext{x}{u}}{\cprojsndf{A}{P}}\jdeq u$, note that
+\begin{equation*}
+\subst{\tmext{x}{u}}{\cprojsndf{A}{P}}
+  \jdeq
+\subst{u}{{x}{\tfid{P}}}
+  \jdeq
+\subst{u}{\tfid{\subst{x}{P}}}
+  \jdeq
+u
+\end{equation*}
+Therefore, $\psi$ is a left-inverse of $\varphi$. To show that it is also a
+right inverse, note that \autoref{cor:pairing_ehom} gives us that
+\begin{equation*}
+\tmext{\cprojfst{A}{P}{w}}{\cprojsnd{A}{P}{w}}
+  \jdeq
+\subst{w}{(\tmext{\cprojfstf{A}{P}}{\cprojsndf{A}{P}})}.
+\end{equation*}
+Thus, it suffices to show that $\tmext{\cprojfstf{A}{P}}{\cprojsndf{A}{P}}\jdeq
+\tfid{\ctxext{A}{P}}$, which is straightforward.
+\end{proof}
+
+\begin{lem}\label{lem:tmext_assoc}
 For any $x\in T(A)$, $u\in T(S_x(P))$ and $v\in T(S_{\tmext{x}{u}}(Q))$ we have
@@
 \end{equation*}
+\begin{comment}
 Also, we have
@@
 \end{align*}
+\end{comment}
 \end{lem}
 
+\begin{proof}
+By \autoref{subst_by_tmext}, we have $S_v\circ S_{\tmext{x}{u}}\jdeq S_v\circ S_u\circ S_x \jdeq S_{\tmext{u}{v}}\circ S_x$,
+so associativity of term extension follows.
+\end{proof}
+
+\begin{comment}
 \begin{lem}
@@
 
+\begin{proof}
+Straightforward
+\end{proof}
+\end{comment}
+
+We end this section by showing that weakening is the same operation as pre-composition
+with a first projection.
+
+\begin{thm}\label{precomp_by_proj}
+Precomposition with $\cprojfstf{A}{P}$ is weakening by $P$, i.e.
+\begin{equation*}
+(\cprojfstf{A}{P})_\ast \jdeq W_P.
+\end{equation*}
+\end{thm}
+
+\begin{proof}
+\begin{align*}
+(\cprojfstf{A}{P})_\ast & \jdeq S_{\cprojfstf{A}{P}}\circ W_{\ctxext{A}{P}} \\
+& \jdeq S_{W_P(\tfid{A})}\circ W_P\circ W_A \\
+& \jdeq W_P\circ S_{\tfid{A}}\circ W_A \\
+& \jdeq W_P.\qedhere 
+\end{align*}
+\end{proof}
+
+\subsection{The interchange laws}
+
+We are now in the position to define vertical and horizontal composition, and
+prove properties of them.
+
+\begin{defn}
+Let $\jhom{\Gamma}{A}{B}{f}$ and $\jfhom{\Gamma}{A}{B}{f}{P}{Q}{F}$. Then we
+define 
+\begin{equation*}
+\jhomdefn{\Gamma}{{A}{P}}{{B}{Q}}{\jvcomp{P}{f}{F}}{\unfold{\jvcomp{P}{f}{F}}}.
+\end{equation*}
+Whenever we say that we have a diagram of the form
+\begin{equation*}
+\begin{tikzcd}
+R \arrow[r,"f_2"] \arrow[d,fib] &
+S \arrow[d,fib] \\
+P \arrow[r,"f_1"] \arrow[d,fib] &
+Q \arrow[d,fib] \\
+A \arrow[r,"f_0"] &
+B
+\end{tikzcd}
+\end{equation*}
+we mean that we have
+$f_0\in\thom{A}{B}$, $f_1\in\thomd{f_0}{P}{Q}$ and 
+$f_2\in\thomd{\jvcomp{P}{f_0}{f_1}}{R}{S}$.
+\end{defn}
+
+\begin{cor}
+Let $\jhom{\Gamma}{A}{B}{f}$ and $\jfhom{\Gamma}{A}{B}{f}{P}{Q}{F}$. Then we
+have
+\begin{equation*}
+H(\jvcomp{P}{f}{F})\jdeq \jvcomp{H(P)}{H(f)}{H(F)}
+\end{equation*}
+for any E-homomorphism $H:\mathbb{E}\to\mathbb{D}$.
+\end{cor}
+
+\begin{lem}
+Vertical composition is associative.
+\end{lem}
+
+\begin{proof}
+Consider the diagram
+\begin{equation*}
+\begin{tikzcd}
+R \arrow[r,"f_2"] \arrow[d,fib] &
+S \arrow[d,fib] \\
+P \arrow[r,"f_1"] \arrow[d,fib] &
+Q \arrow[d,fib] \\
+A \arrow[r,"f_0"] &
+B
+\end{tikzcd}
+\end{equation*}
+in context $\Gamma$.
+Because weakening distributes over term extension, and term extension is
+associative, we have
+\begin{align*}
+\jvcomp{R}{\jvcomp{P}{f_0}{f_1}}{f_2}
+  & \jdeq
+\tmext{\ctxwk{R}{\tmext{\ctxwk{P}{f_0}}{f_1}}}{f_2}
+  \\
+  & \jdeq
+\tmext{\tmext{\ctxwk{R}{{P}{f_0}}}{\ctxwk{R}{f_1}}}{f_2}
+  \\
+  & \jdeq
+\tmext{\ctxwk{\ctxext{P}{R}}{f_0}}{\tmext{\ctxwk{R}{f_1}}{f_2}}
+  \tag{By \autoref{lem:tmext_assoc}}
+  \\
+  & \jdeq
+\jvcomp{{P}{R}}{f_0}{\jvcomp{R}{f_1}{f_2}}.\qedhere
+\end{align*}
+\end{proof}
+
+\begin{defn}
+Let $\jhom{\Gamma}{A}{B}{f}$ and $\jfhom{\Gamma}{A}{B}{f}{P}{Q}{F}$. Then we
+define the E-homomorphism 
+\begin{equation*}
+F_\circledast\defeq F_\ast\circ (f_\ast/Q):
+\mathbb{E}/\ctxext{{\Gamma}{B}}{Q}\to\mathbb{E}/\ctxext{{\Gamma}{A}}{P}.
+\end{equation*}
+The infix notation of $F_\circledast$ is taken to be $\jfcomp{\blank}{\blank}{\blank}{F}{\blank}$.
+\end{defn}
+
+\begin{lem}\label{lem:three-composition}
+Let $\jhom{\Gamma}{A}{B}{f}$ and $\jfhom{\Gamma}{A}{B}{f}{P}{Q}{F}$. Then we
+have the equality
+\begin{equation*}
+F_\circledast\jdeq (\jvcomp{P}{f}{F})_\ast.
+\end{equation*}
+\end{lem}
+
+\begin{proof}
+\begin{align*}
+F_\ast\circ (f_\ast/Q)
+  & \jdeq
+S_F\circ W_P\circ S_f/(W_A(Q))\circ W_A/\ctxext{B}{Q}
+  \\
+  & \jdeq
+S_F\circ S_{W_P(f)}/W_P(W_A(Q))\circ W_P/W_A(\ctxext{B}{Q})\circ W_A/\ctxext{B}{Q}
+  \\
+  & \jdeq
+S_F\circ S_{W_P(f)}/W_P(W_A(Q))\circ W_{\ctxext{A}{P}}/\ctxext{B}{Q}
+  \\
+  & \jdeq
+S_{\tmext{W_P(f)}{F}}\circ W_{\ctxext{A}{P}}/\ctxext{B}{Q}
+  \tag{By \autoref{subst_by_tmext}}
+  \\
+  & \jdeq
+(\jvcomp{P}{f}{F})_\ast.\qedhere
+\end{align*}
+\end{proof}
+
+We are now in position to state and prove the interchange law of horizontal
+and vertical composition.
+
+\begin{thm}\label{thm:interchange}
+Consider the diagram
+\begin{equation*}
+\begin{tikzcd}
+P \arrow[r,"F"] \arrow[d,fib] &
+Q \arrow[r,"G"] \arrow[d,fib] &
+R \arrow[d,fib] \\
+A \arrow[r,"f"] &
+B \arrow[r,"g"] &
+C
+\end{tikzcd}
+\end{equation*}
+in context $\Gamma$. Then the equality
+\begin{equation*}
+{\jcomp{{A}{P}}{\jvcomp{P}{f}{F}}{\jvcomp{Q}{g}{G}}}
+  \jdeq
+{\jvcomp{P}{\jcomp{A}{f}{g}}{\jfcomp{A}{f}{P}{F}{G}}}
+\end{equation*}
+of morphisms from $\ctxext{A}{P}$ to $\ctxext{C}{R}$ in context $\Gamma$ holds.
+\end{thm}
+
+\begin{proof}
+By \autoref{lem:three-composition}, we have
+\begin{align*}
+\jcomp{{A}{P}}{\jvcomp{P}{f}{F}}{\jvcomp{Q}{g}{G}}
+  & \jdeq
+\jcomp{P}{F}{\jcomp{A}{f}{\tmext{\ctxwk{Q}{g}}{G}}}
+  \tag{By \autoref{lem:three-composition}}
+  \\
+  & \jdeq
+\jcomp{P}{F}{\tmext{\ctxwk{\jcomp{A}{f}{Q}}{\jcomp{A}{f}{g}}}{\jcomp{A}{f}{G}}}
+  \\
+  & \jdeq
+\tmext{\jcomp{P}{F}{\ctxwk{\jcomp{A}{f}{Q}}{\jcomp{A}{f}{g}}}}{\jcomp{P}{F}{{A}{f}{G}}}
+  \\
+  & \jdeq
+\tmext{\jcomp{P}{F}{\ctxwk{\jcomp{A}{f}{Q}}{\jcomp{A}{f}{g}}}}{\jfcomp{A}{f}{P}{F}{G}}
+  \\
+  & \jdeq
+\tmext{\ctxwk{P}{\jcomp{A}{f}{g}}}{\jfcomp{A}{f}{P}{F}{G}}
+  \tag{By \autoref{lem:compW_W}}
+  \\
+  & \jdeq
+\jvcomp{P}{\jcomp{A}{f}{g}}{\jfcomp{A}{f}{P}{F}{G}}.\qedhere
+\end{align*}
+\end{proof}
+
 \begin{thm}
-Precomposition with $\cprojfstf{A}{P}$ is weakening by $P$, i.e.
-\begin{equation*}
-S_{\cprojfstf{A}{P}}\circ W_{\ctxext{A}{P}}\jdeq W_P.
-\end{equation*}
+Consider the diagram
+\begin{equation*}
+\begin{tikzcd}
+P \arrow[r,"F"] \arrow[d,fib] &
+Q \arrow[r,"G"] \arrow[d,fib] &
+R \arrow[d,fib] \\
+A \arrow[r,"f"] &
+B \arrow[r,"g"] &
+C
+\end{tikzcd}
+\end{equation*}
+in context $\Gamma$. 
+Then $F_\circledast\circ G_\circledast\jdeq (\jfcomp{A}{f}{P}{F}{G})_\circledast$.
+In other words the composition $\jfcomp{}{}{}{\blank}{\blank}$ is associative.
 \end{thm}
+
+\begin{proof}
+\begin{align*}
+F_\circledast\circ G_\circledast
+  & \jdeq
+(\jvcomp{P}{f}{F})_\ast\circ(\jvcomp{Q}{g}{G})_\ast 
+  \tag{By \autoref{lem:three-composition}}\\
+  & \jdeq
+(\jcomp{\ctxext{A}{P}}{\jvcomp{P}{f}{F}}{\jvcomp{Q}{g}{G}})_\ast 
+  \tag{By \autoref{lem:compcomp}}\\
+  & \jdeq
+(\jvcomp{}{\jcomp{A}{f}{g}}{\jfcomp{A}{f}{P}{F}{G}})_\ast 
+  \tag{By \autoref{thm:interchange}} \\
+  & \jdeq
+(\jfcomp{A}{f}{P}{F}{G})_\circledast.
+  \tag{By \autoref{lem:three-composition}}
+\end{align*}
+\end{proof}
+
+\begin{lem}
+Let $\jhom{\Gamma}{A}{B}{f}$ and $\jfhom{\Gamma}{A}{B}{f}{P}{Q}{F}$. Then
+$\jvcomp{P}{f}{F}$ is the unique morphism from $\ctxext{A}{P}$ to $\ctxext{B}{Q}$
+with the property that both the diagram
+\begin{equation*}
+\begin{tikzcd}
+\ctxext{A}{P}
+  \ar{r}{\jvcomp{P}{f}{F}}
+  \ar{d}[swap]{\cprojfstf{A}{P}}
+& \ctxext{B}{Q}
+  \ar{d}{\cprojfstf{B}{Q}}
+  \\
+A \ar{r}[swap]{f}
+& B
+\end{tikzcd}
+\end{equation*}
+commutes, and $\jcomp{}{\jvcomp{P}{f}{F}}{\cprojsndf{B}{Q}}\jdeq F$.
+\end{lem}
+
+\begin{proof}
+We first note that
+\begin{align*}
+\jcomp{}{\jvcomp{P}{f}{F}}{\cprojfstf{B}{Q}}
+  & \jdeq
+\jcomp{P}{F}{\jcomp{A}{f}{\cprojfstf{B}{Q}}} 
+  \tag{By \autoref{lem:three-composition}} \\
+  & \jdeq
+\jcomp{P}{F}{\ctxwk{\jcomp{A}{f}{Q}}{\jcomp{A}{f}{\idtm{B}}}} \\
+  & \jdeq
+\jcomp{P}{F}{\ctxwk{\jcomp{A}{f}{Q}}{f}} \\
+  & \jdeq
+\ctxwk{P} f \tag{By \autoref{lem:compW_W}} \\
+  & \jdeq
+\jcomp{\ctxext{A}{P}}{\cprojfstf{A}{P}}{f}.
+  \tag{By \autoref{precomp_by_proj}}
+\end{align*}
+Also, we have
+\begin{align*}
+\jcomp{}{\jvcomp{P}{f}{F}}{\cprojsndf{B}{Q}}
+  & \jdeq
+\jcomp{P}{F}{\jcomp{A}{f}{\idtm{Q}}}
+  \tag{By \autoref{lem:three-composition}}
+  \\
+  & \jdeq
+\jcomp{P}{F}{\idtm{\jcomp{A}{f}{Q}}}
+  \\
+  & \jdeq
+F.
+\end{align*}
+Thus, we conclude that $\jvcomp{P}{f}{F}$ has indeed the stated property. For
+the uniqueness, let $G:\ctxext{A}{P}\to\ctxext{B}{Q}$ be a morphism such that
+$\jcomp{}{G}{\cprojfstf{B}{Q}}\jdeq\jcomp{}{\cprojfstf{A}{P}}{f}$ and
+$\jcomp{}{G}{\cprojsndf{B}{Q}}\jdeq F$. Then it follows that
+\begin{equation*}
+G \jdeq \tmext{\jcomp{}{\cprojfstf{A}{P}}{f}}{F}
+  \jdeq \tmext{\ctxwk{P}{f}}{F}
+  \jdeq \jvcomp{P}{f}{F}.\qedhere
+\end{equation*}
+\end{proof}
+
+\subsection{Categories internal to E-systems}
+The definition of internal category we're about to give, has a type of objects
+and a type of morphisms, such that the codomain function is a projection. It might
+be more appropriate to call such structures `right categories', and the corresponding
+structures where the domain function is a projection might be called `left
+categories'. In particular, the opposite of a right category will be a left
+category. There is also a symmetric notion of a category.
+
+\begin{defn}
+A category-framework $\cftalg{A}$ in context $\Gamma$ in an E-system $\mathbb{E}$ consists of
+\begin{enumerate}
+\item A family $\cftalgc{\cftalg{A}}\in\cat{C}/\Gamma$, representing the type of
+objects of the category framework;
+\item A family $\cftalgf{\cftalg{A}}\in\cat{C}/\ctxext{\Gamma}{\cftalgc{\cftalg{A}}}$,
+representing the type of morphisms of the category framework, where a term of
+$\subst{X}{\cftalgf{\cftalg{A}}}$ is considered to be a morphism into $X\in T(\cftalgc{\cftalg{A}})$;
+\item A morphism $\jhom{\Gamma}{\ctxext{\cftalgc{\cftalg{A}}}{\cftalgf{\cftalg{A}}}}{\cftalgc{\cftalg{A}}}{\cftctxext}$, representing the domain function;
+\item A morphism $\jhom{{\Gamma}{\cftalgc{\cftalg{A}}}}{{\cftalgf{\cftalg{A}}}{\jcomp{}{\cftctxext}{\cftalgf{\cftalg{A}}}}}{\cftalgf{\cftalg{A}}}{\cftfamext}$, representing composition;
+\item A term $\jterm{{\Gamma}{\cftalgc{\cftalg{A}}}}{\cftalgf{\cftalg{A}}}{\cftempf{\cftalg{A}}}$,
+representing the identity morphisms.
+\end{enumerate}
+\end{defn}
+
+\begin{rmk}
+Instead of looking at $\cftfamext$ as a morphism from 
+$\ctxext{\cftalgf{\cftalg{A}}}{\jcomp{}{\cftctxext}{\cftalgf{\cftalg{A}}}}$
+to $\cftalgf{\cftalg{A}}$ in context $\ctxext{\Gamma}{\cftalgc{\cftalg{A}}}$, 
+one could also look at $\cftfamext$ as a 
+morphism \emph{over $\cprojfstf{\cftalgc{\cftalg{A}}}{\cftalgf{\cftalg{A}}}$} 
+in context $\Gamma$, as indicated in the following diagram:
+\begin{equation*}
+\begin{tikzcd}[column sep=large]
+\cftalgf{\cftalg{A}}
+  \ar[fib]{d}
+& \jcomp{}{\cftctxext}{\cftalgf{\cftalg{A}}}
+  \ar[fib]{d}
+  \ar{l}[swap]{\cftfamext}
+  \ar{r}
+& \cftalgf{\cftalg{A}}
+  \ar[fib]{d}
+  \\
+\cftalgc{\cftalg{A}}
+& \ctxext{A}{\cftalgf{\cftalg{A}}}
+  \ar{l}{\cprojfstf{A}{\cftalgf{\cftalg{A}}}}
+  \ar{r}[swap]{\cftctxext}
+& \cftalgc{\cftalg{A}}
+\end{tikzcd}
+\end{equation*}
+From this point of view, we see that $\cftfamext$ has an object $X$ and a 
+morphism $f$ into $X$ as parameters, 
+then takes a morphism into the domain of $f$ as an argument, 
+and outputs a morphism into $X$. The projection maps $(X,f)$ to $X$, so we see
+that composition is indeed an operation of this kind.
+\end{rmk}
+
+A category-framework is not a category unless several axioms of the theory of
+categories are satisfied. The domain of a composed arrow $g\circ f$ must be the
+domain of $f$, composition must be associative, the domains of the identity
+morphisms must be the same as their codomains, and identities must satisfy
+the left and right identity laws. However, the axioms expressing these 
+conditions are not very insightful when they're written down directly. Thus, we
+take the usual way and define the notion of functor before we describe the
+actual conditions of being a category.
+
+\begin{defn}
+Let $\cftalg{A}$ and $\cftalg{B}$ be category-frameworks in context $\Gamma$,
+in an E-system $\mathbb{E}$.
+An \emph{internal functor $\cfthom{F}$ from $\extalg{A}$ to
+$\extalg{B}$ in context $\Gamma$} is a pair 
+$(\cfthomc{\cfthom{F}},\cfthomf{\cfthom{F}})$ consisting of
+\begin{align*}
+\jalign\jhom
+  {\Gamma}
+  {\cftalgc{\cftalg{A}}}
+  {\cftalgc{\cftalg{B}}}
+  {\cfthomc{\cfthom{F}}}
+  \\
+\jalign\jfhom
+  {\Gamma}
+  {\cftalgc{\cftalg{A}}}
+  {\cftalgc{\cftalg{B}}}
+  {\cfthomc{\cfthom{F}}}
+  {\cftalgf{\cftalg{A}}}
+  {\cftalgf{\cftalg{B}}}
+  {\cfthomf{\cfthom{F}}}
+\end{align*}
+for which the diagrams
+\begin{equation}\label{eq:exthom1}
+\begin{tikzcd}
+\ctxext{\cftalgc{\cftalg{A}}}{\cftalgf{\cftalg{A}}}
+  \ar{r}{\jvcomp{}{\cfthomc{\cfthom{F}}}{\cfthomf{\cfthom{F}}}}
+  \ar{d}[swap]{\cftctxext[\cftalg{A}]}
+& \ctxext{\cftalgc{\cftalg{B}}}{\cftalgf{\cftalg{B}}}
+  \ar{d}{\cftctxext[\cftalg{B}]}
+  \\
+\cftalgc{\cftalg{A}}
+  \ar{r}[swap]{\cfthomc{\cfthom{F}}}
+& \cftalgc{\cftalg{B}}
+\end{tikzcd}
+\end{equation}
+and
+\begin{equation}\label{eq:exthom2}
+\begin{tikzcd}[column sep=huge]
+\ctxext{\cftalgf{\cftalg{A}}}{\jcomp{}{\cftctxext[\cftalg{A}]}{\cftalgf{\cftalg{A}}}}
+  \ar{r}{\jvcomp{}{\cfthomf{\cfthom{F}}}{\jcomp{}{\cftctxext[\cftalg{A}]}{\cfthomf{\cfthom{F}}}}}
+  \ar{d}[swap]{\cftfamext[\cftalg{A}]}
+& \jcomp{}{\cfthomc{\cfthom{F}}}{\ctxext{\cftalgf{\cftalg{B}}}{\jcomp{}{\cftctxext[\cftalg{B}]}{\cftalgf{\cftalg{B}}}}}
+  \ar{d}{\jcomp{}{\cfthomc{\cfthom{F}}}{\cftfamext[\cftalg{B}]}}
+  \\
+\cftalgf{\cftalg{A}}
+  \ar{r}[swap]{\cfthomf{\cfthom{F}}}
+& \jcomp{}{\cfthomc{\cfthom{F}}}{\cftalgf{\cftalg{B}}}
+\end{tikzcd}
+\end{equation}
+commute, and for which $\subst{\cftempf{\cftalg{A}}}{\cfthomf{\cfthom{F}}}
+\jdeq\jcomp{}{\cfthomc{\cfthom{F}}}{\cftempf{\cftalg{B}}}$.
+\end{defn}
+
+\begin{defn}
+Let $\cftalg{A}$ be a category-framework. We define the category-framework
+\begin{equation*}
+\cftfamalg{\cftalg{A}}\defeq(\cftalgf{\cftalg{A}},\jcomp{}{\cftctxext}%
+{\cftalgf{\cftalg{A}}},\cftfamext,\jcomp{}{\cftctxext}{\cftfamext},
+\jcomp{}{\cftctxext}{\cftempf{\cftalg{A}}}),
+\end{equation*}
+When $\cftalg{P}$ is a category-framework in context $\ctxext{\Gamma}{A}$, then
+and we define the category-framework
+\begin{equation*}
+\ctxext{\cftalgc{\cftalg{A}}}{\cftalg{P}}
+  \defeq
+( \ctxext{A}{\cftalgc{\cftalg{P}}},
+  \cftalgf{\cftalg{P}},
+  \jvcomp{}{\idtm{A}}{\cftctxext[\cftalg{P}]},
+  \cftfamext[\cftalg{P}],
+  \cftempf{\cftalg{P}}).
+\end{equation*}
+\end{defn}
+
+\begin{defn}
+An internal category $\cftalg{A}$ in an E-system $\mathbb{E}$ consists of a
+category-framework $\cftalg{A}$ satisfying the following conditions:
+\begin{enumerate}
+\item \emph{The domain condition.} The pair 
+$(\cftctxext,\idtm{\jcomp{}{\cftctxext}{\cftalgf{\cftalg{A}}}})$ is a
+functor from $\ctxext{\cftalgc{\cftalg{A}}}{\cftfamalg{\cftalg{A}}}$ to
+$\cftalg{A}$.
+\item \emph{The associativity law.} The pair
+$(\cftfamext,\idtm{\jcomp{}{\cftfamext}{\cftalgf{\cftfamalg{\cftalg{A}}}}})$
+is a functor from $\ctxext{\cftalgf{\cftalg{A}}}{\cftfamalg{\cftfamalg{\cftalg{A}}}}$ to
+$\cftfamalg{\cftalg{A}}$.
+\item \emph{The identity domain condition.} The domain of the identity morphism
+is the same as its codomain:
+\begin{equation*}
+\jtermeq
+  {{\Gamma}{\cftalgc{\cftalg{A}}}}
+  {\ctxwk{\cftalgc{\cftalg{A}}}{\cftalgc{\cftalg{A}}}}
+  {\subst{\cftempf{\cftalg{A}}}{\cftctxext[\cftalg{A}]}}
+  {\idtm{\cftalgc{\cftalg{A}}}}.
+\end{equation*}
+\item \emph{The left identity law.} A morphism precomposed by the identity morphism
+is equal to the original morphism:
+\begin{equation*}
+\jtermeq
+  {{{\Gamma}{\cftalgc{\cftalg{A}}}}{\cftalgf{\cftalg{A}}}}
+  {\ctxwk{\cftalgf{\cftalg{A}}}{\cftalgf{\cftalg{A}}}}
+  {\subst{\cftempf{\cftalg{A}}}{\cftfamext}}
+  {\idtm{\cftalgf{\cftalg{A}}}}.
+\end{equation*}
+\item \emph{The right identity law.} A morphism postcomposed by the identity morphism
+is equal to the original morphism:
+\begin{equation*}
+\jtermeq
+  {{{\Gamma}{\cftalgc{\cftalg{A}}}}{\cftalgf{\cftalg{A}}}}
+  {\ctxwk{\cftalgf{\cftalg{A}}}{\cftalgf{\cftalg{A}}}}
+  {\subst{\jcomp{}{\cftctxext}{\cftempf{\cftalg{A}}}}{\cftfamext}}
+  {\idtm{\cftalgf{\cftalg{A}}}}.
+\end{equation*}
+\end{enumerate}
+\end{defn}
+
+\begin{rmk}
+Since we have only used operations of E-systems to define internal categories,
+it immediately follows that if $\cftalg{A}$ is an internal category in
+$\mathbb{E}$ and if $F:\mathbb{E}\to\mathbb{D}$ is an E-homomorphism, then
+$F(\cftalg{A})$ is an internal category in $\mathbb{D}$. In particular, if
+$\cftalg{Q}$ is an internal category in context $\ctxext{\Gamma}{B}$ and
+$f:A\to B$, then the change of base $f_\ast(\cftalg{Q})$ is an internal category
+in context $\ctxext{\Gamma}{A}$. Thus, it is immediate that the notion of
+internal category is stable under change of base.
+\end{rmk}
+
+\subsection{Internal E-systems}
+
+\subsubsection{Internal weakening systems}
+
+\subsubsection{Internal projection systems}
+
+\subsubsection{Internal substitution systems}
+
+\subsubsection{Internal E-systems}
+
+\begin{comment}
+\section{Dependent morphisms}
+
+\begin{defn}
+Let $f_0\in\thom{A}{B}$ be an internal morphism and let $P\in\cat{C}/\ctxext{\Gamma}{A}$ and
+$Q\in\cat{C}/\ctxext{\Gamma}{B}$. We define
+\begin{equation*}
+\thomd{f_0}{P}{Q} \defeq \thom{P}{{f_0}_\ast(Q)}
+\end{equation*}
+We indicate that $f_1\in\thomd{f_0}{P}{Q}$ by drawing the diagram
+\begin{equation*}
+\begin{tikzcd}
+P \arrow[r,"f_1"] \arrow[d,->>] & Q \arrow[d,->>] \\
+A \arrow[r,swap,"f_0"] & B.
+\end{tikzcd}
+\end{equation*}
+\end{defn}
+
+\begin{defn}
+Consider the diagram
+\begin{equation*}
+\begin{tikzcd}
+P \arrow[r,"f_1"] \arrow[d,->>] & Q \arrow[d,->>] \arrow[r,"g_1"] & R \arrow[d,->>] \\
+A \arrow[r,swap,"f_0"] & B \arrow[r,"g_0"] & C.
+\end{tikzcd}
+\end{equation*}
+We define
+\begin{align*}
+\jvcomp{P}{f_0}{f_1} & \defeq\unfold\jvcomp{P}{f_0}{f_1} & & \in \thom{\ctxext{A}{P}}{\ctxext{B}{Q}} \\
+\jfcomp{A}{f_0}{P}{f_1}{g_1} & \defeq \unfold\jfcomp{A}{f_0}{P}{f_1}{g_1} & & \in \thomd{\jcomp{A}{f_0}{g_0}}{P}{R}.
+\end{align*}
+\end{defn}
+
+\begin{rmk}
+Consider
+\begin{equation*}
+\begin{tikzcd}
+P \arrow[r,"f_1"] \arrow[d,->>] & Q \arrow[d,->>] \\
+A \arrow[r,swap,"f_0"] & B.
+\end{tikzcd}
+\end{equation*}
+Then the diagram
+\begin{equation*}
+\begin{tikzcd}[column sep=large]
+\ctxext{A}{P} \arrow[r,"\jvcomp{P}{f_0}{f_1}"] \arrow[d,swap,"\cprojfstf{A}{P}"] & \ctxext{B}{Q} \arrow[d,"\cprojfstf{B}{Q}"] \\
+A \arrow[r,swap,"f_0"] & B
+\end{tikzcd}
+\end{equation*}
+commutes. Also, we have
+\begin{equation*}
+S_{\tmext{x}{u}}(\jvcomp{P}{f_0}{f_1}) \jdeq \tmext{S_x(f_0)}{S_{\tmext{x}{u}}(f_1)}
+\end{equation*}
+for each $x\in T(A)$ and $u\in T(S_x(P))$.
+\end{rmk}
 
@@
 
-\begin{comment}
 \begin{lem}
@@
 \end{proof}
+
+\section{Weak E-systems}
+
+\begin{rmk}
+Note that $W$ is a strict $2$-functor frm $\cat{C}$ to $\mathbb{Cat}$ of which
+the value at an object $\Gamma$ is specified to be $\cat{C}/\Gamma$, and which
+is required to preserve the final object. We could have considered $W$ to be a 
+pseudo-functor from $\cat{C}$ to $\mathbb{Cat}$ instead. A coherence theorem
+by Power \cite{Power89} then allows for the strictification of this pseudo-functor
+into a (strict) weakening structure. (Does this map $\Gamma$ to $\cat{C}/\Gamma$
+and final to final?). Thus, we know what a weak pre-weakening system has to be.
+The following questions need to be answered:
+\begin{itemize}
+\item What is a weak weakening/projection/substitution system.
+\item What is a weak E-system? I.e.~what are the coherence rules for something
+to be a weakening/projection/substitution homomorphism?
+\item Can they all be strictified? Are there more examples of weak E-systems?
+\end{itemize}
+\end{rmk}
 \end{comment}
+
+\section{E-systems compared to other systems of type dependency}\label{sec:esys_compared}
+In an effort to formalize the simplicial set model of Martin-L\"of type theory
+with the univalence axiom, Voevodsky has studied various notions of models of
+type theory, taking Cartmell's \emph{contextual categories} as a starting
+point. Voevodsky's definition of a \emph{C-system} is equivalent to that of Cartmell
+\cite{VV_C-systems_quotients}. There are several techniques of defining 
+C-systems: in \cite{VV_Csys_univ}, C-systems are constructed out of a universe
+in a category; in \cite{VV_C-systems_quotients}, sub-C-systems and regular
+quotients of C-systems are defined; in \cite{VV_C-systems_monad}, a C-system
+is constructed out of a monad $R$ on the category of sets and a left $R$-module
+with values in $\mathbf{Set}$.
+
+In \cite{VV_B-systems}, Voevodsky introduced his notion of \emph{B-systems}, which is
+an essentially algebraic theory of type dependency with infinitely many sorts 
+and operations, and subsequently he showed that there is a full and faithful
+functor from the category of C-systems to the category of B0-systems, of which
+the image is contained in the subcategory of B-systems. This fact is important at least
+for a theoretical reason: it tells us that the conditions of being a B-system
+are necessary to axiomatize type dependency. In other words,
+this is a theorem that tells us that we need no further axioms. It is
+unknown whether the categories of C-systems and B-systems are actually equivalent.
+This question would settle the sufficiency of the conditions of being a B-system.
+
+The category of B-systems will be shown to be equivalent to the category of
+stratified E-systems. However, we do not arrive at the fact that the functor
+$\mathcal{E}:\mathbf{Bsys}\to\mathbf{Esys}$ is full without imposing an additional
+condition on the E-homomorphisms. 
+To ensure that the category of stratified E-systems with stratified 
+E-homomorphisms between them is a full subcategory of the category 
+$\mathbf{Esys}$, we will require that the underlying functor of each E-homomorphism
+lifts factorizations. It is
+important to note that this condition is also algebraic, so that the category
+of E-systems remains algebraic.
+
+\subsection{Lifting factorizations}
+Recall
+that for any morphism $f:X\to Y$ in a category $\cat{C}$ there is a category
+$\mathbf{fact}(f)$ of factorizations of $f$. Also, any functor $F:\cat{C}\to\cat{D}$
+determines a functor $F_{\mathbf{fact}(f)}:\mathbf{fact}(f)\to\mathbf{fact}(H(f))$
+for any morphism $f$ of $\cat{C}$.
+
+\begin{defn}
+A functor $F:\cat{C}\to\cat{D}$ is said to
+\emph{lift factorizations} if for any $f\in\cat{C}/X$, the functor
+$F_{\mathbf{fact}(f)}:\mathbf{fact}(f)\to\mathbf{fact}(F(f))$ is an isomorphism
+of categories. 
+\end{defn}
+
+\begin{comment}
+\begin{rmk}
+We may choose the property of lifting factorizations
+to involve either an equivalence or an isomorphism of categories. The
+isomorphism-version, would say that for every factorization $h'\circ g'$
+of $F(f)$ there is a unique factorization $f\jdeq h\circ g$ in $\cat{C}$ such
+that $F(g)\jdeq g'$ and $F(h)\jdeq h'$. In the version with equivalences,
+the uniqueness is replaced by uniqueness up to isomorphism, and the equalites
+are replaced by an isomorphism in $\mathbf{fact}(F(f))$.
+
+It is easier to state the isomorphism version of lifting factorizations with
+inference rules and it might be easier to explain this condition on type theoretical
+grounds, although the equivalence version has the advantage of being
+categorical (i.e.~invariant under equivalence of categories).
+
+In the present
+context, it doesn't matter very much which one we pick: the categories in which
+we're interested are all posets.
+\end{rmk}
+\end{comment}
+
+\begin{eg}
+For any slice category $\cat{C}/X$, the forgetful functor $\cat{C}/X\to\cat{C}$
+lifts factorizations.
+\end{eg}
+
+\begin{defn}
+We write $\mathbf{Esys}$ for the category of E-systems where the underlying
+functor of each E-homomorphism lifts factorizations.
+\end{defn}
+
+\subsection{Stratified E-systems}
+
+In this subsection we define the notion of stratified E-system. Then we show
+that the category $\mathbf{Esys_s}$ of stratified E-systems with homomorphisms preserving the
+stratification, is a full subcategory of $\mathbf{Esys}$.
+
+\begin{defn}
+A category $\cat{C}$ with terminal object is said to be \emph{stratified} if there exists a 
+\emph{stratification functor}
+\begin{equation*}
+L : \cat{C}\to (\mathbb{N},\geq)
+\end{equation*}
+such that
+\begin{enumerate}
+\item $L(X)\jdeq 0$ if and only if $X$ is terminal, and for any $f:X\to Y$ we have
+$L(X)\jdeq L(Y)$ if and only if $X\jdeq Y$ and $f\jdeq\catid{X}$. 
+\item every morphism $f:X\to Y$ in $\cat{C}_\mathbb{E}$, where $L(X)\jdeq
+n+m+1$ and $L(Y)\jdeq n$, has a unique factorization 
+\begin{equation*}
+\begin{tikzcd}
+X \arrow[r,"f_m"] & X_m \arrow[r,"f_{m-1}"] & \cdots \arrow[r,"f_1"] & X_1 \arrow[r,"f_0"] & Y
+\end{tikzcd}
+\end{equation*}
+where $L(X_i)\jdeq n+i$.
+\end{enumerate}
+A functor $F:\cat{C}\to\cat{D}$ between stratified categories is said to be stratified 
+if $L_{\cat{C}}\jdeq L_{\cat{D}}\circ F$. 
+\end{defn}
+
+%\begin{rmk}
+%The categorical version of being stratified, is that there exists a factofibration
+%$L:\cat{C}\to (N,\geq)$ which preserves the terminal object.
+%\end{rmk}
+
+\begin{lem}
+Any stratified category is a rooted tree.
+\end{lem}
+
+\begin{proof}
+Let $\cat{C}$ be a stratified category. The root of the tree is going to be the terminal
+object of $\cat{C}$. Note that objects of $\cat{C}$ have no automorphisms other
+than the identity morphisms. Therefore, to show that $\cat{C}$ is a tree,
+it suffices to show that for every object $X$ with $L(X)>0$ there is
+a unique object $\eft(X)$ with $L(\eft(X))\jdeq L(X)-1$, with a morphism $X\to \eft(X)$.
+The existence follows from the existence of the factorization of $X\to 1$. The 
+uniqueness follows from the observation that for any arrow
+$X\to Y$ with $L(Y)\jdeq L(X)-1$, we can factorize $X\to 1$ through $Y$. Since
+factorizations are unique, we will have $Y\jdeq \eft(X)$. 
+\end{proof}
+
+\begin{lem}
+The category of small stratified categories is a full subcategory of the category
+of small categories with terminal objects and between them functors which lift 
+factorizations and preserve the terminal object.
+\end{lem}
+
+\begin{proof}
+Let $F:\cat{C}\to\cat{D}$ be a functor between stratified categories which lifts
+factorizations and 
+which preserves terminal objects. We show by induction that $F$ is stratified.
+Since $F$ preserves the terminal object, $L(Y)\jdeq 0$ implies $L(F(Y))\jdeq 0$.
+Now suppose that for $n\in\mathbb{N}$, we have that $L(\Gamma)\jdeq n$ implies
+$L(F(Y))\jdeq n$, and let $X$ be such that $L(X)\jdeq n+1$ with
+$p_X:X\to\eft(X)$. Then the poset $\mathbf{fact}(p_X)$ is isomorphic to the
+poset $\mathbf{2}$ with two objects, one of which is smaller than the other. By the
+assumption that $F$ lifts factorizations, the poset $\mathbf{fact}(F(p_X))$
+is also isomorphic to $\mathbf{2}$. Thus, the only factorization
+of $F(X)\to\unit$ starts with $F(p_X):H(X)\to F(\eft(X))$, which implies that
+$L(F(X))\jdeq n+1$. 
+\end{proof}
+
+\begin{rmk}
+It is easy to show that a category can only be stratified in at most one way,
+and that if a category is stratified, then so are any of its slices. Therefore,
+it makes sense to ask for the property that the weakening and substitution
+functors in a pre-weakening or pre-substitution system are stratified.
+\end{rmk}
+
+\begin{defn}
+A pre-E-system is said to be \emph{stratified} if its underlying category is
+stratified and if each $W_A$ and $S_x$ is a stratified functor. The category of
+stratified pre-E-systems with stratified E-homomorphisms between them is denoted by
+$\mathbf{Esys_s}$.
+\end{defn}
+
+\begin{cor}\label{lem:strat_full}
+The category of stratified pre-E-systems is a full subcategory of the category of
+pre-E-systems.
+\end{cor}
+
+\subsection{The embedding of B0-systems into pre-E-systems}
+
+Voevodsky distinguishes between pre-B-systems, which is an essentially algebraic theory with
+only operations and no equalities; B0-systems, which is an extension of the
+essentially algebraic theory of pre-B-systems which ensures that all the operations behave
+well with two sets of the operations ($\eft$ and $\ebd$); and B-systems, which extends the 
+theory of B0-systems with equalities ensuring that the rest of the operations 
+are compatible with each other. Voevodsky, furthermore distinguishes
+non-unital and unital versions of these, and noted that not all B-homomorphisms
+between unital B-systems are unital. The same distinguishment could be made for 
+E-systems by forgetting the projection structure, but we have not been able to construct a
+functor from unital B0-systems to unital pre-E-systems. Therefore, we shall only
+consider non-unital B0-systems and unital B-systems. In this section we will
+recall the definition of non-unital B0-systems and construct a full and faithful
+functor from them to the category of pre-E-systems.
+
+One reason for introducing B-systems by first introducing pre-B-systems is that
+it is clear in the definition of pre-B-systems that there are only countably
+many operations. Since we are interested in embedding the B-systems into E-systems,
+we shall organize the definition of B-systems a bit differently, but of course
+the resulting definition of B0-systems and B-systems will be (trivially) 
+equivalent.
+
+\begin{defn}
+A \emph{B-framework} is a collection of data of the following form:
+\begin{enumerate}
+\item for all $n\in\mathbb{N}$ two sets $B_n$ and $\tilde{B}_n$. 
+\item for all $n\in\mathbb{N}$ maps of the form
+\begin{align*}
+\eft[n] & : B_{n+1}\to B_n \\
+\ebd[n] & : \tilde{B}_n\to B_n.
+\end{align*}
+For $m,n\in\mathbb{N}$, we denote the composition $\eft[n]\circ\cdots\circ\eft[n+m]:B_{n+m+1}\to B_n$ by $\eft[n]^m$. 
+\item $B_0$ is a singleton $\{\pt\}$.
+\end{enumerate} 
+\end{defn}
+
+\begin{rmk}
+Since the indices $n,m\in\mathbb{N}$ can be infered, we will usually omit them
+in the infix form of the operators.
+\end{rmk}
+
+\begin{defn}
+A homomorphism $H:\mathbb{B}\to\mathbb{A}$ of B-frameworks consists of maps
+$H_n:B_n\to A_n$ and $\tilde{H}_n:\tilde{B}_n\to\tilde{A}_n$ such that
+\begin{align*}
+\eft(H(X)) & \jdeq H(\eft(X)) \\
+\ebd(\tilde{H}(x)) & \jdeq H(\ebd(x))
+\end{align*}
+for any $X\in B_n$ and $x\in\tilde{B}_n$. The category of B-frameworks is
+denoted by $\mathbf{Bfr}$. 
+\end{defn}
+
+\begin{defn}
+For every B-framework $\mathbb{B}$ and any $X\in B_n$, there is a B-framework
+$\mathbb{B}/X$ given by
+\begin{align*}
+(B/X)_{m} & \jdeq \{Y\in B_{n+m}\mid\eft^{m}(Y)\jdeq X\}\\
+(\tilde{B}/X)_m & \jdeq \{y\in \tilde{B}_{n+m}\mid\eft^m(\ebd(y))\jdeq X\}.
+\end{align*}
+Also, for any homomorphism $H:\mathbb{B}\to\mathbb{A}$ of B-frameworks and any
+$X\in B_n$, there is a homomorphism $H/X:\mathbb{B}/X\to\mathbb{A}/H(X)$
+defined in the obvious way.
+\end{defn}
+
+\begin{rmk}
+Note that for $X\in B_n$ and $Y\in B_{n+m}$ such that $\eft^m(Y)\jdeq X$, 
+we have $(\mathbb{B}/X)/Y\cong B/Y$.
+\end{rmk}
+
+\begin{defn}
+A B0-system $\mathbb{B}$ consists of a B-framework $\mathbb{B}$ and homomorphisms
+\begin{align*}
+W_{X} & : \mathbb{B}/\eft(X)\to\mathbb{B}/X\\
+S_{x} & : \mathbb{B}/\ebd(x)\to\mathbb{B}/\eft(\ebd(x))
+\end{align*}
+of B-frameworks, for any $X\in B_{n+1}$ and any $x\in\tilde{B}_{n+1}$. 
+\end{defn}
+
+\begin{lem}
+When $\mathbb{B}$ is a B0-system, then so is each $\mathbb{B}/X$.
+\end{lem}
+
+\begin{proof}
+Straightforward.
+\end{proof}
+
+\begin{defn}
+A B0-homomorphism $H:\mathbb{B}\to\mathbb{A}$ is a homomorphism of B-frameworks
+for which the diagrams
+\begin{equation*}
+\begin{tikzcd}[column sep=huge]
+\mathbb{B}/X \arrow[r,"H/X"] & \mathbb{A}/H(X) \\
+\mathbb{B}/\eft(X) \arrow[u,"W_X"] \arrow[r,swap,"H/\eft(X)"] & \mathbb{A}/\eft(H(X)) \arrow[u,swap,"W_{H(X)}"]
+\end{tikzcd}
+\end{equation*}
+and
+\begin{equation*}
+\begin{tikzcd}[column sep=huge]
+\mathbb{B}/\ebd(x) \arrow[r,"H/\ebd(x)"] \arrow[d,swap,"S_x"] & \mathbb{A}/\ebd(H(X)) \arrow[d,"S_{\tilde{H}(x)}"] \\
+\mathbb{B}/\eft(\ebd(x)) \arrow[r,swap,"H/\eft(\ebd(x))"] & \mathbb{A}/\eft(\ebd(H(X)))
+\end{tikzcd}
+\end{equation*}
+of homomorphisms of B-frameworks commute for every 
+$X\in B_n$ and $x\in\tilde{B}_n$. The category of B0-systems is denoted
+by $\mathbf{B0sys}$.
+\end{defn}
+
+Now that we have defined the category of B0systems, we can start defining the
+embedding into pre-E-systems.
+
+\begin{defn}
+Let $\mathbb{B}$ be a B-framework. Then $\mathbb{B}$ determines
+a stratified category $U(\mathbb{B})$. Also, any homomorphism of B-frameworks
+determines a stratified functor.
+\end{defn}
+
+\begin{constr}
+The set of objects of $U(\mathbb{B})$ is taken to be the set
+$\bigsqcup_{(n\in\mathbb{N})}B_n$. Now we define
+$(k,X)\leq (n,Y)$ if and only if $n\leq k$ and $\eft[n]^{k-n}(X)\jdeq Y$.
+In other words, we only have $(n+m,X)\leq (n,\eft^m(X))$. It is immediate from
+this definition that for any $X\in B_{n+1}$ there is a unique $\Gamma\in
+B_n$ such that $(n+1,X)\leq (n,\Gamma)$, so $U(\mathbb{B})$ is stratified.
+
+To show that the functor $U(H):U(\mathbb{B})\to U(\mathbb{A})$ is stratified for
+any homomorphism $H:\mathbb{B}\to\mathbb{A}$ of B-frameworks, note that
+by \autoref{lem:strat_full} it suffices to show that each $U(H)$ 
+lifts factorizations.
+Let $H:\mathbb{B}\to\mathbb{A}$ be a homomorphism of B-frameworks, and let
+$(m,X):(n+m,X)\leq (n,\Gamma)$. Then the category $\mathbf{fact}_{(m,X)}$ of
+factorizations of $(m,X)$ is a finite linear order with
+$m+1$ elements, and so is the category $\mathbf{fact}_{(m,H(X))}$. 
+\end{constr}
+
+\begin{rmk} 
+It is in fact the case that $U(\mathbb{B})$ is the free category generated by the
+graph with $\bigsqcup_{(n\in\mathbb{N})}B_n$ as the set of vertices, and
+for each object of the form $(n+1,X)$ an edge to $(n,\eft[n](X))$. 
+Essentially all $U$ does, is forgetting about the term structure 
+$\tilde{B}$ of $\mathbb{B}$.
+
+Furthermore, it is useful to note that for any $X\in B_n$, we get an 
+isomorphism $U(\mathbb{B}/X)\cong  U(\mathbb{B})/(n,X)$ of rooted trees,
+natural in $X$. Therefore we will usually not distinguish between $U(\mathbb{B})/X$
+and $U(\mathbb{B})/(n,X)$. 
+\end{rmk}
+
+\begin{defn}
+Let $\mathbb{B}$ be a B0-system. Then $U(\mathbb{B})$ can be given the structure
+of a pre-E-system. We write $\mathcal{E}(\mathbb{B})$ for the resulting
+pre-E-system. Thus we get a functor $\mathcal{E}:\mathbf{B0sys}\to\mathbf{Esys_s}$.
+\end{defn}
+
+\begin{constr}
+We begin by simultaneously defining the term structure and the pre-substitution
+structure on $U(\mathbb{B})$ by induction on the natural numbers. More precisely,
+for any $m\in\mathbb{N}$ we will define, for any $(n+m,X)\leq (n,\Gamma)$ a set
+$T(m,X)$ and for any $x\in T(m,X)$ a homomorphism 
+$S_x:\mathbb{B}/X\to\mathbb{B}/\Gamma$ of B-frameworks. Then it follows
+immediately that we get stratified functors $U(S_x):U(\mathbb{B})/X\to U(\mathbb{B})/\Gamma$
+defining the pre-substitution structure of $\mathcal{E}(\mathbb{B})$.
+
+For $(n,X)\in U(\mathbb{B})$, we define $T(0,X)\defeq
+\{\pt[X]\}$ and $S_{\pt[X]}\defeq \catid{\mathbb{B}/X}$. We also define for
+$(n+1,X)\leq (n,\Gamma)$ the set $T(1,X)\jdeq\{x\in B_{n+1}\mid \ebd(x)\jdeq X\}$.
+Note that for $x\in T(1,X)$, the homomorphism $S_x:\mathbb{B}/X \to\mathbb{B}/\Gamma$
+is already assumed to exist, because $\mathbb{B}$ is a B0-system.
+
+For the inductive step, suppose that for $m\in\mathbb{N}$ we have sets $T(m,X)$
+for any $(n+m,X)\leq (n,\Gamma)$ and homomorphisms $S_x:\mathbb{B}/X \to \mathbb{B}/\Gamma$
+of B-frameworks, for any $x\in T(m,X)$. Then we define
+\begin{equation*}
+T(m+1,X)\defeq\bigsqcup_{x\in T(m,\eft(X))} T(1,S_x(X)).
+\end{equation*}
+For $(x,u)\in T(m+1,X)$ we define the homomorphism $S_{(x,u)}:\mathbb{B}/X\to
+\mathbb{B}/\Gamma$ as
+\begin{equation*}
+S_{(x,u)}\defeq S_u\circ S_x/X.
+\end{equation*}
+It remains to define a pre-weakening structure on $U(\mathbb{B})$. To do this,
+it suffices to define homomorphisms $W_{(m,X)}:\mathbb{B}/\Gamma \to\mathbb{B}/X$
+of B-frameworks, for any $(n+m,X)\leq (n,\Gamma)$. We define
+\begin{equation*}
+W_{(m,X)}\defeq W_X\circ\cdots\circ W_{\eft^{m-1}(X)}.\qedhere
+\end{equation*}
+\end{constr}
+
+\begin{thm}
+The functor $\mathcal{E}$ is full and faithful.
+\end{thm}
+
+\begin{proof}
+We have to show that the map $\mathrm{hom}(\mathbb{B},\mathbb{A})\to
+\mathrm{hom}(\mathcal{E}(\mathbb{B}),\mathcal{E}(\mathbb{A}))$ is a bijection
+for any two B0-systems $\mathbb{B}$ and $\mathbb{A}$.
+
+To show that this map is injective, suppose $F,G:\mathbb{B}\to\mathbb{A}$ are 
+B0-homomorphisms such that $\mathcal{E}(F)\jdeq\mathcal{E}(G)$. Then we have
+$F_n(X)\jdeq\mathcal{E}(F)(n,X)\jdeq \mathcal{E}(G)(n,X)\jdeq G_n(X)$ for each
+object $x\in B_n$. Also, we have $\tilde{F}_n(x)\jdeq T_{\mathcal{E}(F)}(n,x)
+\jdeq T_{\mathcal{E}(G)}(n,x)\jdeq \tilde{G}_n(x)$ for each
+$x\in\tilde{B}_n$. This shows that $F\jdeq G$, so we conclude that $\mathcal{E}$
+is faithful.
+
+To show that the map $\mathrm{hom}(\mathbb{B},\mathbb{A})\to
+\mathrm{hom}(\mathcal{E}(\mathbb{B}),\mathcal{E}(\mathbb{A}))$ 
+is surjective, let $H:\mathcal{E}(\mathbb{B})\to
+\mathcal{E}(\mathbb{A})$. Since both $\mathcal{E}(\mathbb{B})$ and
+$\mathcal{E}(\mathbb{A})$ are stratified, we get from
+\autoref{lem:strat_full} that $H$ is a stratified pre-E-homomorphism. 
+Because $H$ is stratified, we get maps $H_n:B_n\to A_n$ for any $n\in\mathbb{N}$.
+The term structure of $H$ gives maps $\tilde{H}_n:
+\tilde{B}_n\to\tilde{A}_n$. This defines a B0-homomorphism since $H$ 
+preserves the pre-weakening and pre-substitution structures of $\mathcal{E}(\mathbb{B})$.
+It is immediate from the definition of $H_n$ and
+$\tilde{H}_n$ that this B0-homomorphism is mapped back to $H$, so we conclude
+that $\mathcal{E}$ is full.
+\end{proof}
+
+\subsection{B-systems are stratified E-systems}
+There are two properties which hold for pre-E-systems of the form $\mathcal{E}
+(\mathbb{B})$ for some B0-system $\mathbb{B}$, 
+which stand in the way of proving that the category of stratified
+pre-E-systems is equivalent to the category of B0-systems. The first is that
+for any $A\in U(\mathbb{B})/\Gamma$ and $P\in U(\mathbb{B})/\ctxext{\Gamma}{A}$,
+we have an isomorphism $T(\ctxext{A}{P})\cong\bigsqcup_{x\in T(A)}T(S_x(P))$.
+The second, which depends on the first, is that for any $(x,u)\in T(\ctxext{A}{P})$
+we have $S(x,u)\jdeq S_u\circ S_x/P$. In other words, both the term structure
+and the pre-substitution structure of $\mathcal{E}(\mathbb{B})$ are compatible
+with the categorical structure of the underlying category $U(\mathbb{B})$.
+However, to prove the corresponding facts for E-systems, we needed the full
+set of conditions on unital E-systems: to prove that the term structure of
+$U(\mathbb{B})$ is compatible with composition we needed units, and to prove 
+that the operations $S_x$ are compatible with the pairing function we needed
+that they are E-homomorphisms.
+
+Presumably, we can assume these compatibility conditions on pre-E-systems in order
+to prove the desired equivalence of B0-systems and stratified pre-E-systems. However,
+that seems somewhat unnatural, especially given that they automatically hold
+for E-systems. Thus, for the remainder of this section we set out to show that
+the category of unital B-systems is equivalent to the category of stratified 
+unital E-systems.
+
+\begin{defn}
+A unital B0-system is a B0-system with maps
+\begin{equation*}
+\delta_n : B_{n+1}\to \tilde{B}_{n+2}.
+\end{equation*}
+such that $\ebd(\delta_n(X))\jdeq W_{X}(X)$ for any $X\in B_n$. 
+A B0-homomorphism $H$ is said to be unital if one has
+\begin{equation*}
+H(\delta(X))\jdeq\delta(H(X))
+\end{equation*}
+for any $X\in B_n$.  
+\end{defn}
+
+\begin{rmk}
+Voevodsky showed in \cite{VV_B-systems} that the category of unital B0-systems
+with unital B0-homomorphisms between them is not a full
+subcategory of $\mathbf{B0sys}$. 
+\end{rmk}
+
+\begin{defn}
+A \emph{(unital) B-system} is a B0-system for which the following conditions
+hold:
+\begin{enumerate}
+\item Every $W_X$ and $S_x$ is a unital B0-homomorphism.
+\item For every $x\in \tilde{B}_{n+1}$ one has $S_x\circ W_{\ebd(x)}\jdeq
+\catid{\mathbb{B}/\eft(\ebd(x))}$. 
+\item For every $x\in\tilde{B}_{n+1}$ one has $S_x(\delta(\ebd(x)))\jdeq x$.
+\item For every $X\in B_{n+1}$ one has $S_{\delta(X)}\circ W_X/X\jdeq
+\catid{\mathbb{B}/X}$. 
+\end{enumerate}
+\emph{B-homomorphisms} are unital B0-homomorphisms between B-systems.
+We denote the category of B-systems by $\mathbf{Bsys}$. 
+\end{defn}
+
+\begin{lem}
+For any non-unital B-system $\mathbb{B}$, $\mathcal{E}(\mathbb{B})$ is a 
+stratified non-unital E-system. 
+\end{lem}
+
+\begin{proof}
+Straightforward.
+\end{proof}
+
+\begin{defn}
+For any unital B-system $\mathbb{B}$, $\mathcal{E}(\mathbb{B})$ has a
+projection structure. We thus obtain a functor 
+$\mathcal{E}:\mathbf{Bsys}\to\mathbf{Esys}$ from unital B-systems to E-systems.
+\end{defn}
+
+\begin{constr}
+We first need to extend the functor $\mathcal{E}:\mathbf{B0sys}\to\mathbf{Esys_p}$
+from non-unital B0-systems to non-unital pre-E-systems, to a functor from
+unital B-systems to E-systems.
+
+We need to give $\mathcal{E}(\mathbb{B})$ the structure of a unital
+pre-E-system. We construct by induction on $m\in \mathbb{N}$, a term of type 
+$T(m,W_{(m,X)}(X))$ with the property that that 
+\begin{equation*}
+S_{\tfid{(m,X)}}\circ W_{(m,X)}/X\jdeq\catid{U(\mathbb{B})/X},
+\end{equation*} 
+for every $(n+m,X)\leq (n,\Gamma)$ by induction on $m$. The extra
+condition is part of the inductive construction, because we need it on the way.
+
+For $m\jdeq 0$ we have but one
+choice, and for $m\jdeq 1$ we can directly use the unital structure of
+$\mathbb{B}$. For the inductive step, recall that
+\begin{equation*}
+T(m+1,W_{(m+1,X)}(X))\jdeq\bigsqcup_{x\in T(m,W_{(m+1,X)}(\eft(X)))}T(1,S_x(W_{(m+1,X)}(X))).
+\end{equation*}
+We have the term $x\defeq \tilde{W}_X(\tfid{(m,\eft(X))})\in T(m,W_{(m+1,X)}(\eft(X)))$.
+Now we may observe that 
+\begin{align*}
+S_x(W_{(m+1,X)}(X))
+  & \jdeq 
+S_{\tilde{W}_X(\tfid{(m,\eft(X))})}(W_{(m+1,X)}(X))\\
+  & \jdeq 
+S_{\tilde{W}_X(\tfid{(m,\eft(X))})}(W_X(W_{(m,\eft(X))}(X))) \\
+  & \jdeq
+W_{X}(S_{\tfid{(m,\eft(X))}}(W_{(m,\eft(X))}(X))) \\
+  & \jdeq 
+W_{X}(X).
+\end{align*}
+Therefore, we see that $T(1,S_x(W_{(m+1,X)}(X)))\jdeq T(1,W_X(X))$, where we can
+choose the term $\delta(X)$. Thus, we define
+\begin{equation*}
+\tfid{(m+1,X)}\defeq (\tilde{W}_X(\tfid{(m,\eft(X))}),\delta(X))
+\end{equation*} 
+for $m\geq 1$. Now we need to check that $S_{\tfid{(m+1,X)}}\circ 
+W_{(m+1,X)}/X\jdeq \catid{U(\mathbb{B})/X}$. This is indeed the case:
+\begin{align*}
+S_{\tfid{(m+1,X)}}\circ W_{(m+1,X)}/X
+  & \jdeq
+S_{\delta(X)}\circ S_{\tilde{W}_X(\tfid{(m,\eft(X))})}\circ W_{(m+1,X)}/X \\
+  & \jdeq
+S_{\delta(X)}\circ W_X/X \\
+  & \jdeq
+\catid{U(\mathbb{B})/X}.
+\end{align*}
+This finishes the definition of the pre-projection structure on $\mathcal{E}(\mathbb{B})$.
+Straightforward proves by induction give that each $W_{(m,X)}$ and $S_x$ is a 
+pre-projection homomorphism and that $S_x(\tfid{(m,X)})\jdeq x$, for any 
+$x\in T(m,X)$.
+\end{constr}
+
+\begin{thm}
+The functor $\mathcal{E}:\mathbf{Bsys}\to\mathbf{Esys}$ lifts to an equivalence
+to the subcategory $\mathbf{Esys_s}$ of stratified E-systems: i.e.~we have a
+commuting diagram
+\begin{equation*}
+\begin{tikzcd}
+& \mathbf{Esys_s} \arrow[dr,"\subseteq"]\\
+\mathbf{Bsys} \arrow[ur,"\simeq"] \arrow[rr,swap,"\mathcal{E}"] & & \mathbf{Esys}
+\end{tikzcd}
+\end{equation*}
+\end{thm}
+
+\begin{proof}
+We already have shown that $\mathcal{E}$ factors through the stratified
+E-systems and that it is full and faithful, so it remains
+to show that every stratified E-system is in the image of $\mathcal{E}$. In
+other words, for every stratified E-system $\mathbb{E}$ we need to construct
+a B0-system $\mathbb{B}$ such that $\mathcal{E}(\mathbb{B})\cong\mathbb{E}$.
+
+Let $\mathbb{E}$ be a stratified E-system. From the stratification functor
+$L:\mathbb{C}_{\mathbb{E}}\to (\mathbb{N},\geq)$ we obtain the pre-image
+presheaf $B(\mathbb{E}):\op{(\mathbb{N},\leq)}\to\mathbf{Set}$ given by
+\begin{align*}
+B(\mathbb{E})_n & \defeq \{X\in\mathrm{Ob}(\mathbb{C}_{\mathbb{E}})\mid L(X)\jdeq n\}
+\end{align*}
+and $\eft(X)\defeq\mathrm{cod}(p_X)$. We define the term structure of $B(\mathbb{E})_n$
+to be given by
+\begin{equation*}
+\tilde{B}(\mathbb{E})_n \defeq \bigsqcup_{X\in B(\mathbb{E})_n} T(p_X)
+\end{equation*}
+with the projection map $\ebd[n]\defeq\pi_1:\tilde{B}(\mathbb{E})_n\to B(\mathbb{E})_n$.
+This defines the B-framework $B(\mathbb{E})$, and we obtain a functor
+$B:\mathbf{Esys_s}\to\mathbf{B0sys}$. It is immediate that we get isomorphisms
+$U(B(\mathbb{E}))\cong\cat{C}_{\mathbb{E}}$, natural in $\mathbb{E}$. Also, we
+get natural isomorphisms $B(\mathbb{E}/\Gamma)\cong B(\mathbb{E})/\Gamma$.
+
+The weakening and substitution structures on
+$B(\mathbb{E})$ are given by $W_X\defeq B(W_{p_X})$ and $S_{(X,x)}\defeq B(S_x)$,
+respectively. Since these functors are assumed to be stratified, they define
+homomorphisms of B-frameworks. Because a weakening system of an E-system is
+required to be functorial, i.e.~$W_{\ctxext{A}{P}}\jdeq W_P\circ W_A$ and
+$W_{\catid{\Gamma}}\jdeq\catid{\cat{C}_{\mathbb{E}}/\Gamma}$, it follows
+immediately that the weakening structure on $U(B(\mathbb{E}))$ is isomorphic to
+that of $\mathbb{E}$. 
+\end{proof}
+
+\begin{comment}
+In the following definition $\Delta(X)$ denotes the discrete category on the set
+$X$ and $\mathbf{2}$ is the category $s\to t$.
+\begin{defn}
+Define $\mathbb{S}$ to be the pushout
+\begin{equation*}
+\begin{tikzcd}[column sep=large]
+\Delta(\mathbb{N}) \arrow[r,"{n\mapsto(n,s)}"] \arrow[d,swap,"i"] & \Delta(\mathbb{N})\times\mathbf{2} \arrow[d] \\
+(\mathbb{N},\leq) \arrow[r] & \mathbb{S}
+\end{tikzcd}
+\end{equation*}
+in $\mathbf{Cat}$.
+\end{defn}
+
+Note that for any $n\in\mathbb{N}$, we have the map $S^n := (m\mapsto m+n)$ which
+is monotone. By the universal property of pushouts, we obtain a functor $\sigma^n:\mathbb{S}\to\mathbb{S}$
+given by
+\begin{equation*}
+\begin{tikzcd}[column sep=huge]
+\Delta(\mathbb{N}) 
+  \arrow[r,"{n\mapsto(n,s)}"] 
+  \arrow[d,swap,"i"] 
+  & 
+\Delta(\mathbb{N})\times\mathbf{2} 
+  \arrow[d]
+  \arrow[r,"{\Delta(S^n)\times\catid{\mathbf{2}}}"]
+  &
+\Delta(\mathbb{N})\times\mathbf{2}
+  \arrow[dd]
+  \\
+(\mathbb{N},\leq) 
+  \arrow[r] 
+  \arrow[d,swap,"{(S^n,\leq)}"]
+  & 
+\mathbb{S}
+  \arrow[dr,densely dotted,yshift=.5ex,"\sigma^n"]
+  \\
+(\mathbb{N},\leq)
+  \arrow[rr] & &
+\mathbb{S}
+\end{tikzcd}
+\end{equation*}
+
+\begin{defn}
+A non-unital pre-B0-system is a presheaf $B$ on $\mathbb{S}$ such that $B(0)$ is
+a singleton, and the following additional structure:
+\begin{enumerate}
+\item For any $n\in\mathbb{N}$ and $X\in B(n)$, write $B/X$ for the sub-presheaf
+of $B\circ\sigma^n$ given by $B/X(0)=\{X\}$ and $B/X(n+1)=B_{n\leq n+1}^{-1}(B/X(n))$. 
+We require that for any $n\in\mathbb{N}$ and any $A\in B(n+1)$, there is a 
+natural transformation $W_A:B/(B_{n\leq n+1}(A))\Rightarrow B/A$.
+\end{enumerate}
+\end{defn}
+\end{comment}
+
+\subsection{From D-systems to E-systems}
+
+We use `D-system' to refer to Dybjer's categories with families. We recall
+Dybjer's definition. Note that $\mathbf{Fam}$ is the category of families of
+sets, of which the objects consist of a set $A$ and a family $\{B(x)\}_{x\in A}$
+over $A$. Although a terminal object is assumed to exist, not every object 
+of a D-system is `fibrant', i.e.~it is
+not the case that every $X\to 1$ is a composition of projections.
+We could only hope that there is a functor from D-sytems to E-systems without
+terminal objects. We write $\mathbf{Esys}_{\circ}$ for the category of E-systems
+without terminal objects.
+
+\begin{defn}
+A category is said to be locally stratified if each of its slice categories
+are stratified.
+\end{defn}
+
+\begin{defn}
+A \emph{setup for categories with families} consists of
+\begin{enumerate}
+\item A category $\cat{C}$ with a terminal object $1$.
+\item A presheaf $\mathrm{Ty}:\op{\cat{C}}\to\mathbf{Set}$. We will write
+$A[\gamma]$ for $\mathrm{Ty}(\gamma,A)$. 
+\item A commuting triangle
+\begin{equation*}
+\begin{tikzcd}
+\int_\cat{C}\mathrm{Ty} \arrow[dr,swap,"\pi"] \arrow[rr,"p"] & & \cat{C}^{\to} \arrow[dl,"\mathrm{cod}"] \\
+& \cat{C}
+\end{tikzcd}
+\end{equation*} 
+We will write $p_A:\ctxext{\Gamma}{A}\to\Gamma$
+for the morphism $U(\Gamma,A)$ into $\Gamma$, for each object $\Gamma\in\cat{C}$ and
+$A\in\mathrm{Ty}(\Gamma)$.
+\end{enumerate}
+\end{defn}
+
+\begin{defn}
+A \emph{category with families} consists of a setup with
+\begin{enumerate}
+\item[(iv)] a preasheaf $\mathrm{Tm}:\op{(\int_\cat{C}\mathrm{Ty})}\to\mathbf{Set}$,
+\item[(v)] a term $q_A\in\mathrm{Tm}(\ctxext{\Gamma}{A},A[p_A])$ for every
+$A\in\mathrm{Ty}(\Gamma)$, 
+\end{enumerate}
+such that for every $\gamma:\Delta\to\Gamma$ in $\cat{C}$ and 
+$a\in \mathrm{Tm}(\Delta,A[\gamma])$ there is a unique morphism 
+$\theta:\Delta\to \ctxext{\Gamma}{A}$ such that $p_A\circ\theta\jdeq\gamma$ and 
+$q_A[\theta]\jdeq a$.
+\end{defn}
+
+\begin{lem}
+Let $\mathbb{D}$ be a category with families. Then there are bijections
+\begin{equation*}
+\mathrm{Tm}(\Delta,A[\gamma])\cong\{\theta:\Delta\to\ctxext{\Gamma}{A}\mid p_A\circ\theta\jdeq \gamma\}
+\end{equation*}
+natural in $\gamma$ and $A$.
+\end{lem}
+
+\begin{proof}
+Let $\theta:\Delta\to\ctxext{\Gamma}{A}$ be such that $p_A\circ\theta\jdeq\gamma$. 
+Then $\subst{\theta}{q_A}$ is a term
+of $\subst{\theta}{{p_A}{A}}$. By the functoriality of $\mathrm{Ty}$, we have
+$\subst{\theta}{{p_A}{A}}\jdeq \subst{p_A\circ\theta}{A}\jdeq\subst{\gamma}{A}$. 
+So we obtain $\subst{\theta}{q_A}\in\mathrm{Tm}(\Delta,\subst{\gamma}{A})$. Now
+the uniqueness of $\theta$ implies that the map $\theta\mapsto\subst{\theta}{q_A}$
+is a bijection. 
+\end{proof}
+
+The previous lemma renders the presheaf $\mathrm{Tm}$ superfluous. Therefore,
+we define D-systems as follows.
+
+\begin{defn}
+A \emph{D-system} consists of a setup for categories with families
+such that for every $\gamma:\Delta\to\Gamma$ and $A\in\mathrm{Ty}(\Gamma)$, there
+is a bijection
+\begin{equation*}
+\{\eta:\Delta\to\ctxext{\Delta}{\subst{\gamma}{A}}\mid p_{\subst{\gamma}{A}}\circ\eta\jdeq\catid{\Delta}\}
+  \cong
+\{\theta:\Delta\to\ctxext{\Gamma}{A}\mid p_A\circ\theta\jdeq\gamma\}
+\end{equation*}
+natural in $\gamma$ and $A$.
+\end{defn}
+
+\begin{thm}
+A category with families is the same thing as a D-system.
+\end{thm}
+
+\begin{defn}
+There is a functor $U$ from D-systems to locally stratified categories.
+\end{defn}
+
+\begin{constr}
+Let $\mathbb{D}$ be a D-system with underlying category $\cat{C}$. We define
+$U(\mathbb{D})$ to be the category with the same objects as $\cat{C}$, where the
+morphisms are generated by the projections $p_A:\ctxext{\Gamma}{A}\to\Gamma$.
+
+We have to prove that this category is locally stratified. Let $\Gamma\in\cat{C}$. 
+\end{constr}
+
+\begin{defn}
+For any D-system $\mathbb{D}$, $U(\mathbb{D})$ can be given the structure of
+an E-system without terminal object. Thus, we get a functor
+$\mathbf{Dsys}\to\mathbf{Esys}_{\circ}$
+\end{defn}
+
+\begin{constr}
+We first need to define a term structure on $U(\mathbb{D})$. 
+Next, we define a weakening structure on $\cat{C}_p$. Since the morphisms of
+$\cat{C}_p$ are composites of projections, we will define
+\begin{equation*}
+W_{p_{A_n}\circ\cdots\circ p_{A_1}}\defeq W_{p_{A_n}}\circ\cdots W_{p_{A_1}}.
+\end{equation*}
+Therefore, we only have to define functors 
+$W_A:\cat{C}_p/\Gamma\to\cat{C}_p/\ctxext{\Gamma}{A}$ with term structure for every 
+$A\in\mathrm{Ty}(\Gamma)$. Since morphisms are composites of projections, we
+can define this functor by induction. Let $W_A(\catid{\Gamma})\defeq
+\catid{\ctxext{\Gamma}{A}}$ and $W_A(p_Q\circ f)\defeq$
+\end{constr}
+
+\begin{thm}
+There is a full and faithful functor from the category of D-systems to the
+category of E-systems.
+\end{thm}
+
+\section{An alternative characterization of E-systems}
+
+\begin{defn}
+An E'-system consists of a category $\cat{C}$ with a subcategory $\cat{F}$ with the
+same objects as $\cat{C}$, with
+the following structure
+\begin{enumerate}
+\item $\cat{F}$ has a terminal object.
+\item For any $f:\Delta\to\Gamma$ in $\cat{C}$, a functor $f_\ast:\cat{F}/\Gamma\to
+\cat{F}/\Delta$ %and
+%for any $Q\in\cat{F}/\ctxext{\Gamma}{B}$ and any section $g$ of $Q$ a section
+%$f_\ast(g)$ of $f_\ast(Q)$, 
+so that $f_\ast$ is contravariant in $f$ and preserves the terminal object. 
+\item For any $A\in\cat{F}/\Gamma$ and $f:\Delta\to\Gamma$, factorizations of
+$f$ through $A$ are naturally the same thing as sections of
+$f_\ast(A)$, i.e. there is an isomorphism
+\begin{equation*}
+\zeta_{A,f}%\defeq\theta\mapsto f_\ast(\theta)
+:\{\theta:\Delta\to\ctxext{\Gamma}{A}\mid A\circ\theta\jdeq f\}
+  \cong
+\{\eta:\Delta\to\ctxext{\Delta}{(f_\ast(A))}\mid f_\ast(A)\circ\eta\jdeq\catid{\Delta}\}
+\end{equation*}
+in the presheaf category 
+$\mathrm{Psh}(\op{(\cat{F}/\Gamma)}\times(\cat{C}/\Gamma))$. 
+\end{enumerate}
+\end{defn}
+
+\begin{rmk}
+Explicitly, the
+naturality condition states that for any $P\in\cat{F}/\ctxext{\Gamma}{A}$,
+$g:\greek{E}\to\Delta$ and $\theta:\Delta\to\ctxext{{\Gamma}{A}}{P}$ such that
+$P\circ\theta\jdeq f$, we have the equality
+\begin{equation*}
+\zeta_{A,f\circ g}(P\circ\theta\circ g)\jdeq f^\ast(P)\circ\zeta_{P,f}(\theta)\circ g.
+\end{equation*}
+\end{rmk}
+
+We can give the category $\cat{F}$ a term structure as follows:
+
+\begin{defn}
+For any E'-system $\mathbb{E}$, the category $\cat{F}$ has a term structure
+where the terms of $A$ are sections of $A$, i.e.
+\begin{equation*}
+T(A)\defeq\{\theta:\Gamma\to\ctxext{\Gamma}{A}\mid A\circ\theta\jdeq\catid{\Gamma}\}.
+\end{equation*}
+\end{defn}
+
+\begin{lem}
+Let $\mathbb{E}$ be an E'-system. Then there is a function $T(A)\to T(f_\ast(A))$
+for any $f:\Delta\to\Gamma$ in $\cat{C}$ and $A\in\cat{F}/\Gamma$, contravariant
+in $f$. 
+\end{lem}
+
+\begin{proof}
+Using the natural isomorphism $\zeta_{A,f}$, we see that
+\begin{align*}
+T(f_\ast(A)) & \jdeq \{\eta : \Delta\to\ctxext{\Delta}{f_\ast(A)}\mid f_\ast(A)\circ\eta\jdeq\catid{\Delta}\}
+  \\
+  & \cong
+\{\theta : \Delta\to \ctxext{\Gamma}{A}\mid A\circ\theta\jdeq f\}.
+\end{align*}
+Therefore, we have the function $x\mapsto \zeta_{A,f}(x\circ f)$ from $T(A)$ to
+$T(f_\ast(A))$. 
+\end{proof}
+
+\begin{defn}
+Consider $P\in\cat{F}/\ctxext{\Gamma}{A}$ and $f:\Delta\to\Gamma$,
+$\theta_1:\Delta\to\ctxext{\Gamma}{A}$ such that $A\circ\theta_1\jdeq f$, 
+and $\theta_2:\Delta\to\ctxext{{\Gamma}{A}}{P}$ such that $P\circ\theta_2\jdeq\theta_1$.  
+\end{defn}
+
+\begin{defn}
+For any $P\in\cat{F}/\ctxext{\Gamma}{A}$ and $f:\Delta\to\Gamma$ 
+there is a morphism $\tilde{f}:\ctxext{\Delta}{f_\ast(A)}\to\ctxext{\Gamma}{A}$
+with the property that $A\circ\tilde{f}\jdeq f\circ f_\ast(A)$, and
+we have isomorphisms
+\begin{align*}
+\{\theta:\ctxext{\Delta}{f_\ast(A)}\to\ctxext{{\Gamma}{A}}{P}\mid P\circ\theta\jdeq \tilde{f}\}
+&  \cong
+\{\eta:\ctxext{\Delta}{f_\ast(A)}\to\ctxext{{\Delta}{(f_\ast(A))}}{f_\ast(P)}\mid f_\ast(P)\circ\eta\jdeq\catid{\ctxext{\Delta}{f_\ast(A)}}\}
+\end{align*}
+\end{defn}
+
+\begin{constr}
+\begin{align*}
+& \{ \theta : \ctxext{\Delta}{f_\ast(A)}\to\ctxext{\Gamma}{A}\mid A\circ\theta\jdeq f\circ f_\ast(A) \} \\
+  & \cong
+\{ \eta : \ctxext{\Delta}{f_\ast(A)}\to\ctxext{{\Delta}{f_\ast(A)}}{(f\circ f_\ast(A))_\ast(A)}
+  \mid (f\circ f_\ast(A))_\ast(A)\circ\eta\jdeq\catid{\ctxext{\Delta}{f_\ast(A)}}\} \\
+  & \jdeq
+\{ \eta : \ctxext{\Delta}{f_\ast(A)}\to\ctxext{{\Delta}{f_\ast(A)}}{(f_\ast(A))_\ast(f_\ast(A))}
+  \mid (f\circ f_\ast(A))_\ast(A)\circ\eta\jdeq\catid{\ctxext{\Delta}{f_\ast(A)}}\} \\
+  & \jdeq
+T(W_{f_\ast(A)}(f_\ast(A))).
+\end{align*}
+Since we have $\tfid{f_\ast(A)}\in T(W_{f_\ast(A)}(f_\ast(A)))$, we obtain 
+\begin{equation*}
+\tilde{f}\defeq \zeta_{A,f\circ f_\ast(A)}^{-1}(\tfid{f_\ast(A)}):\ctxext{\Delta}{f_\ast(A)}\to\ctxext{\Gamma}{A}.
+\end{equation*}
+It remains to prove the asserted isomorphism. 
+\begin{align*}
+\{\theta:\ctxext{\Delta}{f_\ast(A)}\to\ctxext{{\Gamma}{A}}{P}\mid P\circ\theta\jdeq \tilde{f}\}
+&  \cong
+\{\eta : \ctxext{\Delta}{f_\ast(A)}\to\ctxext{{\Delta}{f_\ast(A)}}{\tilde{f}_\ast(P)} \mid
+  \tilde{f}_\ast(P)\circ\eta\jdeq\catid{}\}
+  \\
+  & \cong
+\{\eta:\ctxext{\Delta}{f_\ast(A)}\to\ctxext{{\Delta}{(f_\ast(A))}}{f_\ast(P)}\mid f_\ast(P)\circ\eta\jdeq\catid{\ctxext{\Delta}{f_\ast(A)}}\}
+\end{align*}
+\end{constr}
+
+\begin{lem}
+Let $\mathbb{E}$ be an E'-system. Then there is a function $T(P)\to T(f_\ast(P))$
+for any $f:\Delta\to\Gamma$ in $\cat{C}$ and $P\in\cat{F}/\ctxext{\Gamma}{A}$, contravariant
+in $f$. 
+\end{lem}
+
+\begin{proof}
+Using the natural isomorphism $\zeta_{A,f}$, we see that
+\begin{align*}
+T(f_\ast(P)) & \jdeq \{\eta : \ctxext{\Delta}{f_\ast(A)}\to\ctxext{{\Delta}{f_\ast(A)}}{f_\ast(P)}\mid f_\ast(P)\circ\eta\jdeq\catid{\ctxext{\Delta}{f_\ast(A)}}\}.
+\intertext{By all means, this should be isomorphic to}
+  & \cong
+\{\theta : \ctxext{\Delta}{f_\ast(A)}\to \ctxext{{\Gamma}{A}}{P}\mid P\circ\theta\jdeq \tilde{f}\}.
+\end{align*}
+Therefore, we have the function $x\mapsto \zeta_{A,f}(x\circ f)$ from $T(A)$ to
+$T(f_\ast(A))$. 
+\end{proof}
+
+\begin{thm}
+E-systems are the same things as E'-systems.
+\end{thm}
+
+\begin{proof}
+Given an E-system $\mathbb{E}$, we have the category $\cat{C}$ of internal
+morphisms in the empty context, with the subcategory $\cat{F}$ of projections $\cprojfstf{A}{P}$. 
+The functors $f_\ast:\cat{F}/\Gamma\to\cat{F}/\Delta$ are contravariant in
+$f$, so we have to verify whether we have the stated isomorphisms.
+
+Note first that for any $A\in\cat{F}/\Gamma$, we have
+$W_A\defeq A_\ast:\cat{F}/\Gamma\to\cat{F}/\ctxext{\Gamma}{A}$. Then $W_A$ is
+contravariant in $A$ and that it preserves the terminal object. We give $W_A$
+an action on terms by defining
+\begin{equation*}
+W_A(\theta)\defeq
+\end{equation*}
+for any $\theta\in T(Q)$ and $Q\in\cat{F}/\ctxext{\Gamma}{B}$.
+
+Thus, the $W_A$
+assemble a pre-weakening structure.
+
+Note that for any $A\in\cat{F}/\Gamma$, we have
+\begin{align*}
+T(W_A(A)) & \jdeq \{\eta:\ctxext{\Gamma}{A}\to\ctxext{{\Gamma}{A}}{W_A(A)}\mid W_A(A)\circ\eta\jdeq\catid{\ctxext{\Gamma}{A}}\} \\
+  & \cong
+\{\theta:\ctxext{\Gamma}{A}\to\ctxext{\Gamma}{A}\mid A\circ\theta\jdeq A\},
+\end{align*}
+so we find $\idtm{A}\defeq\zeta_{A,A}(\catid{\ctxext{\Gamma}{A}})\in T(W_A(A))$.
+Now let $\theta:\Delta\to\ctxext{\Gamma}{A}$ so that $A\circ\theta\jdeq f$
+for some $f:\Delta\to\Gamma$. Then we have $\zeta_{A,f}(\theta)\in T(f_\ast(A))$.  
+We wish to define $\theta_\ast(\idtm{A}):\Delta\to\ctxext{\Delta}{f_\ast(A)}$
+
+for any $A\in\cat{F}/\Gamma$ and $Q\in\cat{F}/\ctxext{\Gamma}{B}$, we have
+\begin{align*}
+T(B) 
+  & \jdeq 
+\{ \eta : \Gamma \to\ctxext{\Gamma}{B}
+        \mid B\circ\eta\jdeq\catid{\Gamma}\}
+  \\
+T(W_A(B))
+  & \jdeq
+\{ \eta : \ctxext{\Gamma}{A}\to\ctxext{{\Gamma}{A}}{W_A(B)}
+        \mid W_A(B)\circ\eta\jdeq\catid{\ctxext{\Gamma}{A}}\}
+  \\
+  & \cong
+\{ \theta : \ctxext{\Gamma}{A}\to\ctxext{\Gamma}{B}\mid B\circ\theta\jdeq A\}
+\end{align*}
+Therefore, $\eta\mapsto \zeta_{B,A}(\eta\circ A)$ defines a function from
+$T(B)$ to $T(W_A(B))$. 
+\end{proof}
+
+\begin{comment}
+\section{Acknowledgments}
+The research that eventually became the study of E-systems began while analyzing
+the internel models I encountered during investigations of the type theoretical
+descent theorem, in the summer of 2013. In a private conversation during the Conference on Type Theory, 
+Homotopy Theory and Univalent Foundations at the Universitat Aut\`onoma de 
+Barcelona, I presented the first version of the axioms to Richard Garner, who
+told me he found a similar structure  while analyzing the combinatorics of type
+dependency, see \cite{Garner2014}. After that conference, Joachim
+Kock generously offered me to stay for a few months at the UAB. He was intrigued
+by the idea of doing category theory using only sections, and during the sessions
+we had in this period I finalized the inference rules that correspond to the
+axioms of E-systems.
+
+In Paris, in the spring of 2014, Richard Garner introduced me to Vladimir 
+Voevodsky, who had been working on similar problems and wanted to know exactly
+what axioms should hold on B0-systems so that it is in the image of the functor
+from C-systems to B0-systems. In other words, he wanted to know what the axioms
+of B-systems were, and be confident that we had them all.
+
+In August 2015, I started a grad student of Steve Awodey at the Carnegie 
+Mellon University. With Steve we elaborated much more on the fact that the theory
+of E-systems is essentially algebraic, and I wrote down what an E-system is in
+an arbitrary category with finite limits.
+
+The discussions that followed both with Awodey and with Voevodsky, 
+at the CMU, at the Homotopy Type Theory workshop in Oxford (November 2014), and
+while visiting Voevodsky at the Institute for Advanced Study in Princeton (March 2015), 
+essentially shaped E-systems into their current presentation.
+
+I am grateful to all the people mentioned for sharing many valuable insights 
+with me. 
+\end{comment}

### 2015-06-16 weakeningcats.tex
E'-systems

@@
-\section{The definition of E-systems}\label{sec:esys_defn}
+\section{E-systems}\label{sec:esys_defn}
+
 \subsection{Weakening systems}
@@
 
-\subsection{E-systems}
+\subsection{The definition of E-systems}
 
@@
 \end{rmk}
-
-\subsection{Internal E-systems}
-
-\subsubsection{Internal weakening systems}
-
-\subsubsection{Internal projection systems}
-
-\subsubsection{Internal substitution systems}
-
-\subsubsection{Internal E-systems}
 
@@
 
-\subsection{From D-systems to E-systems}
+\subsection{From D-systems to locally stratified E'-systems}
 
@@
 \item A presheaf $\mathrm{Ty}:\op{\cat{C}}\to\mathbf{Set}$. We will write
-$A[\gamma]$ for $\mathrm{Ty}(\gamma,A)$. 
+$f_\ast(A)$ for $\mathrm{Ty}(f,A)$. 
 \item A commuting triangle
@@
 \begin{enumerate}
-\item[(iv)] a preasheaf $\mathrm{Tm}:\op{(\int_\cat{C}\mathrm{Ty})}\to\mathbf{Set}$,
-\item[(v)] a term $q_A\in\mathrm{Tm}(\ctxext{\Gamma}{A},A[p_A])$ for every
+\item[(iv)] a preasheaf $\mathrm{Tm}:\op{(\int_\cat{C}\mathrm{Ty})}\to\mathbf{Set}$.
+The action of a morphism $f:(\Delta,f_\ast(A))\to(\Gamma,A)$ of $\int_\cat{C}\mathrm{Ty}$
+on a term $x\in \mathrm{Tm}(A)$ is also written by $f_\ast(x)$.
+\item[(v)] a term $q_A\in\mathrm{Tm}(\ctxext{\Gamma}{A},(p_A)_\ast(A))$ for every
 $A\in\mathrm{Ty}(\Gamma)$, 
 \end{enumerate}
-such that for every $\gamma:\Delta\to\Gamma$ in $\cat{C}$ and 
-$a\in \mathrm{Tm}(\Delta,A[\gamma])$ there is a unique morphism 
-$\theta:\Delta\to \ctxext{\Gamma}{A}$ such that $p_A\circ\theta\jdeq\gamma$ and 
-$q_A[\theta]\jdeq a$.
-\end{defn}
-
-\begin{lem}
+such that for every $f:\Delta\to\Gamma$ in $\cat{C}$ and 
+$x\in \mathrm{Tm}(\Delta,f_\ast(A))$ there is a unique morphism 
+$\tilde{x}:\Delta\to \ctxext{\Gamma}{A}$ such that $p_A\circ\tilde{x}\jdeq f$ and 
+$\tilde{x}_\ast(q_A)\jdeq x$.
+\end{defn}
+
+\begin{lem}\label{lem:cwf_to_dsys}
 Let $\mathbb{D}$ be a category with families. Then there are bijections
 \begin{equation*}
-\mathrm{Tm}(\Delta,A[\gamma])\cong\{\theta:\Delta\to\ctxext{\Gamma}{A}\mid p_A\circ\theta\jdeq \gamma\}
-\end{equation*}
-natural in $\gamma$ and $A$.
+\zeta_{A,f}:\{\theta:\Delta\to\ctxext{\Gamma}{A}\mid p_A\circ\theta\jdeq f\}\cong \mathrm{Tm}(\Delta,f_\ast(A))
+\end{equation*}
+coherent in $f$ in the sense that
+\begin{align*}
+\zeta_{A,\catid{\Gamma}}(\theta) & \jdeq \theta \\
+\zeta_{A,f\circ g}(\theta) & \jdeq \zeta_{f_\ast(A),g}(\zeta_{A,f}(\theta)\circ g)
+\end{align*}
+for any $g:X\to\Delta$. 
 \end{lem}
@@
 \begin{proof}
-Let $\theta:\Delta\to\ctxext{\Gamma}{A}$ be such that $p_A\circ\theta\jdeq\gamma$. 
-Then $\subst{\theta}{q_A}$ is a term
-of $\subst{\theta}{{p_A}{A}}$. By the functoriality of $\mathrm{Ty}$, we have
-$\subst{\theta}{{p_A}{A}}\jdeq \subst{p_A\circ\theta}{A}\jdeq\subst{\gamma}{A}$. 
-So we obtain $\subst{\theta}{q_A}\in\mathrm{Tm}(\Delta,\subst{\gamma}{A})$. Now
-the uniqueness of $\theta$ implies that the map $\theta\mapsto\subst{\theta}{q_A}$
+Let $\theta:\Delta\to\ctxext{\Gamma}{A}$ be such that $p_A\circ\theta\jdeq f$. 
+Then $\theta_\ast(q_A)$ is an element
+of $\mathrm{Tm}(\theta_\ast((p_A)_\ast(A)))$. By the functoriality of $\mathrm{Ty}$, we have
+$\theta_\ast((p_A)_\ast(A))\jdeq (p_A\circ\theta)_\ast(A)\jdeq f_\ast(A)$. 
+So we obtain $\theta_\ast(q_A)\in\mathrm{Tm}(\Delta,f_\ast(A))$. Now
+the uniqueness of $\theta$ implies that the map $\theta\mapsto \theta_\ast(q_A)$
 is a bijection.
+
+It remains to verify the coherence identities. Let $\theta : \Gamma\to\ctxext{\Gamma}{A}$
+be such that $p_A\circ\theta\jdeq\catid{\Gamma}$. We wish to show that
+$\theta_\ast(q_A)\jdeq\theta$.  
+\begin{align*}
+\theta_\ast(q_A)
+& \jdeq \theta_\ast(\zeta_{A,p_A}(\catid{\ctxext{\Gamma}{A}})) \\
+& \jdeq \cdots \\
+& \jdeq \theta
+\end{align*}
 \end{proof}
@@
 A \emph{D-system} consists of a setup for categories with families
-such that for every $\gamma:\Delta\to\Gamma$ and $A\in\mathrm{Ty}(\Gamma)$, there
+such that for every $f:\Delta\to\Gamma$ and $A\in\mathrm{Ty}(\Gamma)$, there
 is a bijection
 \begin{equation*}
-\{\eta:\Delta\to\ctxext{\Delta}{\subst{\gamma}{A}}\mid p_{\subst{\gamma}{A}}\circ\eta\jdeq\catid{\Delta}\}
+\zeta_{A,f}:\{\theta:\Delta\to\ctxext{\Gamma}{A}\mid p_A\circ\theta\jdeq f\}
   \cong
-\{\theta:\Delta\to\ctxext{\Gamma}{A}\mid p_A\circ\theta\jdeq\gamma\}
-\end{equation*}
-natural in $\gamma$ and $A$.
+\{\eta:\Delta\to\ctxext{\Delta}{f_\ast(A)}\mid p_{f_\ast(A)}\circ\eta\jdeq\catid{\Delta}\}
+\end{equation*}
+natural in $f$ in the sense that the equalities
+\begin{align*}
+\zeta_{A,\catid{\Gamma}}(\theta) & \jdeq \theta \\
+\zeta_{A,f\circ g}(\theta\circ g) & \jdeq \zeta_{f_\ast(A),g}(\zeta_{A,f}(\theta)\circ g)
+\end{align*}
+hold for any $f:\Delta\to\Gamma$ and $g:X\to\Delta$, and any $\theta:\Delta\to\ctxext{\Gamma}{A}$
+such that $p_A\circ\theta\jdeq f$. 
 \end{defn}
@@
 
-\begin{defn}
-There is a functor $U$ from D-systems to locally stratified categories.
+\begin{proof}
+In \autoref{lem:cwf_to_dsys} we have shown that every category with families is
+a D-system. Thus, it suffices to show the converse.
+
+Let $\mathbb{D}$ be a D-system. For any object $\Gamma$ of $\mathbb{D}$ and any
+type $A\in\mathrm{Ty}(\Gamma)$, we define 
+\begin{equation*}
+\mathrm{Tm}(A)\defeq \{ x :\Gamma\to\ctxext{\Gamma}{A}\mid p_A\circ x\jdeq\catid{\Gamma}\}.
+\end{equation*}
+For any morphism $f:\Delta\to\Gamma$ and $A\in\mathrm{Ty}(\Gamma)$, we define
+the function $\mathrm{Tm}(f):\mathrm{Tm}(A)\to\mathrm{Tm}(f_\ast(A))$ by
+\begin{equation*}
+\mathrm{Tm}(f,x)\defeq \zeta_{A,f}(x\circ f).
+\end{equation*}
+We have to verify that this satisfies the functorial identities. For $f:\Delta\to\Gamma$
+and $g:X\to\Delta$ and any $x\in\mathrm{Tm}(A)$, we have
+\begin{align*}
+\mathrm{Tm}(f\circ g,x) & \jdeq \zeta_{A,f\circ g}(x\circ f\circ g) \\
+& \jdeq \zeta_{f_\ast(A),g}(\zeta_{A,f}(x\circ f)\circ g) \\
+& \jdeq \mathrm{Tm}(g,\mathrm{Tm}(f,x)).
+\end{align*}
+and
+\begin{equation*}
+\mathrm{Tm}(\catid{\Gamma},x) \jdeq \zeta_{A,\catid{\Gamma}}(x) \jdeq x.
+\end{equation*}
+This finishes the construction of the presheaf $\mathrm{Tm}$ on $\int_{\cat{C}}\mathrm{Ty}$.
+
+Next, we have to find $q_A\in\mathrm{Tm}((p_A)_\ast(A))$. Note that we have the
+isomorphism
+\begin{align*}
+\mathrm{Tm}((p_A)_\ast(A))
+& \jdeq \{\eta :\ctxext{\Gamma}{A}\to\ctxext{{\Gamma}{A}}{(p_A)_\ast(A)}\mid p_{(p_A)_\ast(A)}\circ\theta\jdeq\catid{\ctxext{\Gamma}{A}}\} \\
+& \cong \{\theta : \ctxext{\Gamma}{A}\to\ctxext{\Gamma}{A} \mid p_A\circ \theta\jdeq p_A\}
+\end{align*}
+The latter set contains the identity morphism $\catid{\ctxext{\Gamma}{A}}$. Therefore,
+we find
+\begin{equation*}
+q_A \defeq \zeta_{A,p_A}(\catid{\ctxext{\Gamma}{A}})\in\mathrm{Tm}((p_A)_\ast(A)).
+\end{equation*}
+To verify that we have constructed a category with families, let $f:\Delta\to\Gamma$
+and $x\in T(f_\ast(A))$. Note that we have the morphism 
+$\tilde{x}\defeq \zeta_{A,f}^{-1}(x):\Delta\to\ctxext{\Gamma}{A}$. 
+We need to check that $\tilde{x}$ is unique with the property that 
+$\mathrm{Tm}(\tilde{x},q_A)\jdeq x$. The coherence identities can be used to see
+that this is indeed the case:
+\begin{align*}
+\mathrm{Tm}(\tilde{x},q_A)
+& \jdeq \zeta_{(p_A)_\ast(A),\tilde{x}}(q_A\circ\tilde{x}) \\
+& \jdeq \zeta_{(p_A)_\ast(A),\tilde{x}}(\zeta_{A,p_A}(\catid{\ctxext{\Gamma}{A}})\circ\tilde{x}) \\
+& \jdeq \zeta_{A,p_A\circ\tilde{x}}(\tilde{x}) \\
+& \jdeq \zeta_{A,f}(\tilde{x}) \\
+& \jdeq x.
+\end{align*}
+For the uniqueness, let $\theta:\Delta\to\ctxext{\Gamma}{A}$ be any morphism such that
+$p_A\circ\theta\jdeq f$ and $\mathrm{Tm}(\theta,q_A)\jdeq x$. Then we have
+\begin{align*}
+x & \jdeq \mathrm{Tm}(\theta,q_A) \\
+& \jdeq \zeta_{(p_A)_\ast(A),\theta}(q_A\circ\theta) \\
+& \jdeq \zeta_{(p_A)_\ast(A),\theta}(\zeta_{A,p_A}(\catid{\ctxext{\Gamma}{A}})\circ\theta \\
+& \jdeq \zeta_{A,p_A\circ\theta}(\theta) \\
+& \jdeq \zeta_{A,f}(\theta).
+\end{align*}
+Therefore we see that $\tilde{x}\jdeq\zeta_{A,f}^{-1}(x)\jdeq\theta$.
+\end{proof}
+
+\begin{defn}
+An E'-system is said to be locally stratified, if the category $\cat{F}$ is
+locally stratified.
+\end{defn}
+
+\begin{comment}
+\begin{rmk}
+In the following definition, one could take the underlying category of the
+E-system $\mathcal{E}(\mathbb{D})$ associated to a D-system $\mathbb{D}$ to
+be the subcategory generated by the projections $p_A$, and construct the
+rest of the structure of E-systems on top of that. However, it is not so easy
+to characterize the image of the functor $\mathcal{E}$ in that case. Intuitively,
+D-systems should correspond to locally stratified E-systems. To be able to prove
+such a theorem, we've gone with a slightly modified definition of the underlying
+category.
+\end{rmk}
+\end{comment}
+
+\begin{defn}
+For any D-system $\mathbb{D}$ there is 
+a locally stratified E'-system $U(\mathbb{D})$ without terminal object in 
+$\cat{F}$. Thus, we get a functor $\mathbf{Dsys}\to\mathbf{E'sys}_{\circ}$
 \end{defn}
@@
 Let $\mathbb{D}$ be a D-system with underlying category $\cat{C}$. We define
-$U(\mathbb{D})$ to be the category with the same objects as $\cat{C}$, where the
-morphisms are generated by the projections $p_A:\ctxext{\Gamma}{A}\to\Gamma$.
-
-We have to prove that this category is locally stratified. Let $\Gamma\in\cat{C}$. 
-\end{constr}
-
-\begin{defn}
-For any D-system $\mathbb{D}$, $U(\mathbb{D})$ can be given the structure of
-an E-system without terminal object. Thus, we get a functor
-$\mathbf{Dsys}\to\mathbf{Esys}_{\circ}$
-\end{defn}
-
-\begin{constr}
-We first need to define a term structure on $U(\mathbb{D})$. 
+$\cat{F}\subseteq\cat{C}$ to be the
+\begin{comment}
+category consisting of 
+\begin{enumerate}
+\item equivalence classes of finite lists of the form $(\Gamma,A_1,\ldots A_n)$, where $\Gamma$ is
+an object of $\cat{C}$ and $A_{i+1}\in\mathrm{Ty}(\ctxext{{{\Gamma}{A_1}}{\cdots}}{A_i})$, where
+the equivalence relation is generated by
+\begin{equation*}
+(\ctxext{{{\Gamma}{A_1}}{\cdots}}{A_n},A_{n+1},\ldots,A_{n+m})\sim (\Gamma,A_1,\ldots,A_{n+m}).
+\end{equation*}
+\item
+\end{enumerate}
+\end{comment}
+subcategory with the same objects as
+$\cat{C}$, where the morphisms are generated by the projections $p_A$.
+
+Let $f:\Delta\to\Gamma$. We define $f_\ast:\cat{F}/\Gamma\to\cat{F}/\Delta$ by
+induction on the length of the arrow $A$ into $\Gamma$, simultaneously with a
+morphism $\pi_2(f,A)$ from $\ctxext{\Delta}{f_\ast(A)}$ into $\ctxext{\Gamma}{A}$
+for which the square
+\begin{equation*}
+\begin{tikzcd}
+\ctxext{\Delta}{f_\ast(A)} \arrow[r,"{\pi_2(f,A)}"] \arrow[d,fib] & \ctxext{\Gamma}{A} \arrow[d,fib] \\
+\Delta \arrow[r,swap,"f"] & \Gamma
+\end{tikzcd}
+\end{equation*}
+is a pullback square. We define
+$f_\ast(\catid{\Gamma})\defeq\catid{\Delta}$, and $\pi_2(f,\catid{\Gamma})\defeq f:\ctxext{\Delta}{f_\ast(\catid{\Gamma})}\to\ctxext{\Gamma}{\catid{\Gamma}}$. The resulting square is trivially
+a pullback square. Now suppose we have a pullback square as above.
+and let $B\in\mathrm{Ty}(\ctxext{\Gamma}{A})$. Then we define
+\begin{equation*}
+f_\ast(A\circ p_B)\defeq \pi_2(f,A)_\ast(B).
+\end{equation*}
+By \autoref{lem:EE_pb}, we get the top pullback square in the diagram
+\begin{equation*}
+\begin{tikzcd}[column sep=6em]
+\ctxext{{\Delta}{f_\ast(A)}}{f_\ast(B)} \arrow[d,fib] \arrow[r,"{\pi_2(\pi_2(f,A),B)}"] & \ctxext{{\Gamma}{A}}{B} \arrow[d,fib] \\
+\ctxext{\Delta}{f_\ast(A)} \arrow[d,fib] \arrow[r,"{\pi_2(f,A)}"] & \ctxext{\Gamma}{A} \arrow[d,fib] \\
+\Delta \arrow[r,"f"] & \Gamma
+\end{tikzcd}
+\end{equation*}
+Since the bottom square is assumed to be a pullback, we get that the outer rectangle
+is a pullback, as desired.
+
 Next, we define a weakening structure on $\cat{C}_p$. Since the morphisms of
@@
 
-\section{An alternative characterization of E-systems}
-
-\begin{defn}
-An E'-system consists of a category $\cat{C}$ with a subcategory $\cat{F}$ with the
+\subsection{The equivalence of E-systems and E'-systems}
+
+The definition of D-system suggests a more general structure related to E-systems.
+
+\begin{defn}
+An \emph{E'-system} consists of a category $\cat{C}$ with a subcategory $\cat{F}$ with the
 same objects as $\cat{C}$, with
@@
 so that $f_\ast$ is contravariant in $f$ and preserves the terminal object. 
-\item For any $A\in\cat{F}/\Gamma$ and $f:\Delta\to\Gamma$, factorizations of
-$f$ through $A$ are naturally the same thing as sections of
+\item For any $A\in\cat{F}/\Gamma$ and $f:\Delta\to\Gamma$, lifts of
+$f$ to $A$ are naturally the same thing as sections of
 $f_\ast(A)$, i.e. there is an isomorphism
@@
 \end{equation*}
-in the presheaf category 
-$\mathrm{Psh}(\op{(\cat{F}/\Gamma)}\times(\cat{C}/\Gamma))$. 
+which are natural in $A$ and $f$, in the sense that
+\begin{enumerate}
+\item For any $P\in\cat{F}/\ctxext{\Gamma}{A}$, $f:\Delta\to\Gamma$ and
+$\theta:\Delta\to\ctxext{{\Gamma}{A}}{P}$ such that $(\ctxext{A}{P})\circ\theta\jdeq f$,
+one hase
+\begin{equation*}
+\zeta_{A,f}(P\circ\theta)\jdeq f_\ast(P)\circ \zeta_{\ctxext{A}{P},f}(\theta).
+\end{equation*}
+\item For any $A\in\cat{F}/\Gamma$, $f:\Delta\to\Gamma$, $g:X\to\Delta$
+and $\theta:\Delta\to\ctxext{\Gamma}{A}$ such that $A\circ\theta\jdeq f$, one
+has
+\begin{align*}
+\zeta_{A,\catid{\Gamma}}(\theta) & \jdeq \theta \\
+\zeta_{A,f\circ g}(\theta\circ g) & \jdeq\zeta_{f_\ast(A),g}(\zeta_{A,f}(\theta)\circ g).
+\end{align*}
 \end{enumerate}
-\end{defn}
-
-\begin{rmk}
-Explicitly, the
-naturality condition states that for any $P\in\cat{F}/\ctxext{\Gamma}{A}$,
-$g:\greek{E}\to\Delta$ and $\theta:\Delta\to\ctxext{{\Gamma}{A}}{P}$ such that
-$P\circ\theta\jdeq f$, we have the equality
-\begin{equation*}
-\zeta_{A,f\circ g}(P\circ\theta\circ g)\jdeq f^\ast(P)\circ\zeta_{P,f}(\theta)\circ g.
-\end{equation*}
-\end{rmk}
+\end{enumerate}
+\end{defn}
 
@@
 
-\begin{lem}
+\begin{lem}\label{lem:ptm}
 Let $\mathbb{E}$ be an E'-system. Then there is a function $T(A)\to T(f_\ast(A))$
@@
 \begin{defn}
-Consider $P\in\cat{F}/\ctxext{\Gamma}{A}$ and $f:\Delta\to\Gamma$,
-$\theta_1:\Delta\to\ctxext{\Gamma}{A}$ such that $A\circ\theta_1\jdeq f$, 
-and $\theta_2:\Delta\to\ctxext{{\Gamma}{A}}{P}$ such that $P\circ\theta_2\jdeq\theta_1$.  
-\end{defn}
+For any $A\in\cat{F}/\Gamma$, we define
+$W_A\defeq A_\ast:\cat{F}/\Gamma\to\cat{F}/\ctxext{\Gamma}{A}$. Then we get
+for any $A,B\in\cat{F}/\Gamma$ an isomorphism
+\begin{equation*}
+\mathrm{hom}_{\cat{C}/\Gamma}(A,B)
+  \cong
+T(W_A(B)).
+\end{equation*}
+In particular, we get a term $\idtm{A}\in T(W_A(A))$. 
+\end{defn}
+
+\begin{constr}
+Let $A,B\in\cat{F}/\Gamma$. Then we have the isomorphism
+\begin{align*}
+T(W_A(B)) & \jdeq \{\eta:\ctxext{\Gamma}{A}\to\ctxext{{\Gamma}{A}}{W_A(B)}\mid W_A(B)\circ\eta\jdeq\catid{\ctxext{\Gamma}{A}}\} \\
+  & \cong
+\{\theta:\ctxext{\Gamma}{A}\to\ctxext{\Gamma}{B}\mid B\circ\theta\jdeq A\},
+\end{align*}
+so we find $\idtm{A}\defeq\zeta_{A,A}(\catid{\ctxext{\Gamma}{A}})\in T(W_A(A))$.
+\end{constr}
 
@@
 For any $P\in\cat{F}/\ctxext{\Gamma}{A}$ and $f:\Delta\to\Gamma$ 
-there is a morphism $\tilde{f}:\ctxext{\Delta}{f_\ast(A)}\to\ctxext{\Gamma}{A}$
-with the property that $A\circ\tilde{f}\jdeq f\circ f_\ast(A)$, and
-we have isomorphisms
-\begin{align*}
-\{\theta:\ctxext{\Delta}{f_\ast(A)}\to\ctxext{{\Gamma}{A}}{P}\mid P\circ\theta\jdeq \tilde{f}\}
-&  \cong
-\{\eta:\ctxext{\Delta}{f_\ast(A)}\to\ctxext{{\Delta}{(f_\ast(A))}}{f_\ast(P)}\mid f_\ast(P)\circ\eta\jdeq\catid{\ctxext{\Delta}{f_\ast(A)}}\}
-\end{align*}
+there is a morphism $\pi_2(f,A):\ctxext{\Delta}{f_\ast(A)}\to\ctxext{\Gamma}{A}$
+such that $A\circ \pi_2(f,A)\jdeq f\circ f_\ast(A)$.
 \end{defn}
@@
 \begin{equation*}
-\tilde{f}\defeq \zeta_{A,f\circ f_\ast(A)}^{-1}(\tfid{f_\ast(A)}):\ctxext{\Delta}{f_\ast(A)}\to\ctxext{\Gamma}{A}.
-\end{equation*}
-It remains to prove the asserted isomorphism. 
-\begin{align*}
-\{\theta:\ctxext{\Delta}{f_\ast(A)}\to\ctxext{{\Gamma}{A}}{P}\mid P\circ\theta\jdeq \tilde{f}\}
-&  \cong
-\{\eta : \ctxext{\Delta}{f_\ast(A)}\to\ctxext{{\Delta}{f_\ast(A)}}{\tilde{f}_\ast(P)} \mid
-  \tilde{f}_\ast(P)\circ\eta\jdeq\catid{}\}
-  \\
+\pi_2(f,A)\defeq \zeta_{A,f\circ f_\ast(A)}^{-1}(\tfid{f_\ast(A)}):\ctxext{\Delta}{f_\ast(A)}\to\ctxext{\Gamma}{A}
+\end{equation*}
+satisfying $A\circ\pi_2(f,A)\jdeq f\circ f_\ast(A)$.
+\end{constr}
+
+\begin{lem}
+For any $f:\Delta\to\Gamma$, $g:X\to\Delta$ and $A\in\cat{F}/\Gamma$, one has
+\begin{equation*}
+\pi_2(f,A)\circ\pi_2(g,f_\ast(A))\jdeq\pi_2(f\circ g,A).
+\end{equation*}
+\end{lem}
+
+\begin{proof}
+It suffices to show that
+\begin{equation*}
+\zeta_{A,f\circ g\circ g_\ast(f_\ast(A))}(\pi_2(f,A)\circ\pi_2(g,f_\ast(A)))\jdeq\tfid{g_\ast(f_\ast(A))}.
+\end{equation*}
+Because $g\circ g_\ast(f_\ast(A))\jdeq f_\ast(A)\circ\pi_2(g,f_\ast(A))$, we get
+to use the naturality of $\zeta$:
+\begin{align*}
+& \zeta_{A,f\circ g\circ g_\ast(f_\ast(A))}(\pi_2(f,A)\circ\pi_2(g,f_\ast(A))) \\
+& \jdeq
+\zeta_{A,f\circ f_\ast(A)\circ\pi_2(g,f_\ast(A))}(\pi_2(f,A)\circ\pi_2(g,f_\ast(A))) \\
+& \jdeq
+\zeta_{f_\ast(A)_\ast(f_\ast(A)),\pi_2(g,f_\ast(A))}(\zeta_{A,f\circ f_\ast(A)}(\pi_2(f,a))\circ\pi_2(g,f_\ast(A))) \\
+& \jdeq
+\zeta_{f_\ast(A)_\ast(f_\ast(A)),\pi_2(g,f_\ast(A))}(\tfid{f_\ast(A)}\circ\pi_2(g,f_\ast(A))) \\
+& \jdeq
+\zeta_{f_\ast(A)_\ast(f_\ast(A)),\pi_2(g,f_\ast(A))}(\zeta_{f_\ast(A),f_\ast(A)}(\catid{\ctxext{\Delta}{f_\ast(A)}})\circ\pi_2(g,f_\ast(A))) \\
+& \jdeq
+\zeta_{f_\ast(A),f_\ast(A)\circ\pi_2(g,f_\ast(A))}(\catid{\ctxext{\Delta}{f_\ast(A)}}\circ\pi_2(g,f_\ast(A))) \\
+& \jdeq
+\zeta_{f_\ast(A),f_\ast(A)\circ\pi_2(g,f_\ast(A))}(\pi_2(g,f_\ast(A))) \\
+& \jdeq
+\zeta_{f_\ast(A),g\circ g_\ast(f_\ast(A))}(\pi_2(g,f_\ast(A)))\\
+& \jdeq
+\tfid{g_\ast(f_\ast(A))}.\qedhere
+\end{align*}
+\end{proof}
+
+\begin{lem}
+Let $f:\Delta\to\Gamma$ and $A\in\cat{F}/\Gamma$. Then one has 
+\begin{equation*}
+\pi_2(f,A)\circ\zeta_{A,f}(\theta)\jdeq\theta
+\end{equation*}
+for any $\theta:\Delta\to\ctxext{\Gamma}{A}$ satisfying $A\circ\theta\jdeq f$.
+\end{lem}
+
+\begin{proof}
+We will write $\tilde\theta\defeq\zeta_{A,f}(\theta)$. Note that
+\begin{equation*}
+A\circ\pi_2(f,A)\circ\tilde\theta\jdeq f\circ f_\ast(A)\circ\tilde\theta\jdeq f.
+\end{equation*} 
+Therefore, we can prove that $\pi_2(f,A)\circ\tilde\theta\jdeq\theta$ by showing
+that $\zeta_{A,f}(\pi_2(f,A)\circ\tilde\theta)\jdeq\tilde\theta$. This is
+shown using the naturality of $\zeta$:
+\begin{align*}
+\zeta_{A,f}(\pi_2(f,A)\circ\tilde\theta)
+  & \jdeq
+\zeta_{A,f\circ f_\ast(A)\circ\tilde\theta}(\pi_2(f,A)\circ\tilde\theta)
+  \\
+  & \jdeq
+\zeta_{(f_\ast(A))_\ast(f_\ast(A)),\tilde{\theta}}(\zeta_{A,f\circ f_\ast(A)}(\pi_2(f,A))\circ\tilde\theta)
+  \\
+  & \jdeq
+\zeta_{(f_\ast(A))_\ast(f_\ast(A)),\tilde{\theta}}(\idtm{f_\ast(A)}\circ\tilde\theta)
+  \\
+  & \jdeq
+\zeta_{(f_\ast(A))_\ast(f_\ast(A)),\tilde{\theta}}(\zeta_{f_\ast(A),f_\ast(A)}(\catid{\ctxext{\Delta}{f_\ast(A)}})\circ\tilde\theta)
+  \\
+  & \jdeq
+\zeta_{f_\ast(A),f_\ast(A)\circ\tilde\theta}(\catid{\ctxext{\Delta}{f_\ast(A)}}\circ\tilde\theta) \\
+  & \jdeq
+\zeta_{f_\ast(A),\catid{\Delta}}(\tilde\theta) \\
+  & \jdeq
+\theta.\qedhere
+\end{align*}
+\end{proof}
+
+\begin{lem}\label{lem:EE_pb}
+The diagram
+\begin{equation*}
+\begin{tikzcd}[column sep=large]
+\ctxext{\Delta}{f_\ast(A)} \arrow[d,fib] \arrow[r,"{\pi_2(f,A)}"] & \ctxext{\Gamma}{A} \arrow[d,fib] \\
+\Delta \arrow[r,"f"] & \Gamma
+\end{tikzcd}
+\end{equation*}
+is a pullback square.
+\end{lem}
+
+\begin{proof}
+To verify the universal mapping property of the pullback square, consider a
+commuting diagram
+\begin{equation*}
+\begin{tikzcd}
+X \arrow[d,swap,"g"] \arrow[r,"h"] & \ctxext{\Gamma}{A} \arrow[d,fib] \\
+\Delta \arrow[r,"f"] & \Gamma
+\end{tikzcd}
+\end{equation*}
+Then $h$ corresponds uniquely to a section $k$ of $(f\circ g)_\ast(A)$. Since
+$(f\circ g)_\ast(A)\jdeq g_\ast(f_\ast(A))$, the section $k$ fits in the
+commuting diagram
+\begin{equation*}
+\begin{tikzcd}[column sep=huge]
+\ctxext{X}{g_\ast(f_\ast(A))} \arrow[d,xshift=.7ex,fib] \arrow[r,"{\pi_2(g,f_\ast(A))}"]
+& \ctxext{\Delta}{f_\ast(A)} \arrow[d,fib] \arrow[r,"{\pi_2(f,A)}"]
+& \ctxext{\Gamma}{A} \arrow[d,fib] \\
+X \arrow[r,swap,"g"] \arrow[u,xshift=-.7ex,"k"] & \Delta \arrow[r,swap,"f"] & \Gamma
+\end{tikzcd}
+\end{equation*}
+so we obtain $u\defeq\pi_2(g,f_\ast(A))\circ k:X\to\ctxext{\Delta}{f_\ast(A)}$. 
+Then we have $f_\ast(A)\circ u\jdeq g\circ g_\ast(f_\ast(A))\circ k\jdeq g$ and
+$\pi_2(f,A)\circ u\jdeq \pi_2(f\circ g,A)\circ k\jdeq h$. Note that
+$u$ corresponds uniquely to $k$, which corresponds uniquely to $h$, so the
+uniqueness of $u$ is automatic.
+\end{proof}
+
+\begin{lem}
+For any $f:\Delta\to\Gamma$ and $P\in\cat{F}/\ctxext{\Gamma}{A}$, there is a unique isomorphism
+$\varphi_{f,P}:\pi_2(f,A)_\ast(P)\cong (f_\ast/A)(P)$ in 
+$\mathbb{C}/\ctxext{\Delta}{f_\ast(A)}$ such that
+\begin{equation*}
+\pi_2(f,\ctxext{A}{P})\jdeq \pi_2(\pi_2(f,A),P)\circ\varphi_{f,P}.
+\end{equation*}
+Consequently, we get isomorphisms
+\begin{align*}
+\{\theta:\ctxext{\Delta}{f_\ast(A)}\to\ctxext{{\Gamma}{A}}{P}\mid P\circ\theta\jdeq \pi_2(f,A)\}
 &  \cong
@@
 \end{align*}
-\end{constr}
+natural in $P$ and $f$.
+\end{lem}
+
+\begin{proof}
+Note that both the lower square and the outer rectangle in the diagram
+\begin{equation*}
+\begin{tikzcd}[column sep=huge]
+\ctxext{{\Delta}{f_\ast(A)}}{f_\ast(P)} \arrow[d,fib] \arrow[r,"{\pi_2(f,\ctxext{A}{P})}"]
+& \ctxext{{\Gamma}{A}}{P} \arrow[d,fib] \\
+\ctxext{\Delta}{f_\ast(A)} \arrow[d,fib] \arrow[r,"{\pi_2(f,A)}"] 
+& \ctxext{\Gamma}{A} \arrow[d,fib] \\
+\Delta \arrow[r,swap,"f"] & \Gamma
+\end{tikzcd}
+\end{equation*}
+are pullback squares. By the pasting lemma for pullbacks, it follows that the
+top square must also be a pullback. On the other hand, we have the pullback
+square
+\begin{equation*}
+\begin{tikzcd}
+\ctxext{{\Delta}{f_\ast(A)}}{\pi_2(f,A)_\ast(P)} \arrow[d,fib] \arrow[r]
+& \ctxext{{\Gamma}{A}}{P} \arrow[d,fib] \\
+\ctxext{\Delta}{f_\ast(A)} \arrow[r,swap,"{\pi_2(f,A)}"]
+& \ctxext{\Gamma}{A}
+\end{tikzcd}
+\end{equation*}
+Since pullbacks are uniquely unique, we get the desired isomorphism $\varphi_{f,P}$.
+
+It remains to prove the asserted isomorphism. 
+\begin{align*}
+& \{\theta:\ctxext{\Delta}{f_\ast(A)}\to\ctxext{{\Gamma}{A}}{P}\mid P\circ\theta\jdeq \pi_2(f,A)\} \\
+&  \cong
+\{\eta : \ctxext{\Delta}{f_\ast(A)}\to\ctxext{{\Delta}{f_\ast(A)}}{\pi_2(f,A)_\ast(P)} \mid
+  \pi_2(f,A)_\ast(P)\circ\eta\jdeq\catid{}\}
+  \\
+  & \cong
+\{\eta:\ctxext{\Delta}{f_\ast(A)}\to\ctxext{{\Delta}{(f_\ast(A))}}{f_\ast(P)}\mid f_\ast(P)\circ\eta\jdeq\catid{\ctxext{\Delta}{f_\ast(A)}}\}
+\end{align*}
+\end{proof}
 
@@
 \begin{proof}
-Using the natural isomorphism $\zeta_{A,f}$, we see that
-\begin{align*}
-T(f_\ast(P)) & \jdeq \{\eta : \ctxext{\Delta}{f_\ast(A)}\to\ctxext{{\Delta}{f_\ast(A)}}{f_\ast(P)}\mid f_\ast(P)\circ\eta\jdeq\catid{\ctxext{\Delta}{f_\ast(A)}}\}.
-\intertext{By all means, this should be isomorphic to}
-  & \cong
-\{\theta : \ctxext{\Delta}{f_\ast(A)}\to \ctxext{{\Gamma}{A}}{P}\mid P\circ\theta\jdeq \tilde{f}\}.
-\end{align*}
-Therefore, we have the function $x\mapsto \zeta_{A,f}(x\circ f)$ from $T(A)$ to
-$T(f_\ast(A))$. 
+Since $f_\ast(P)$ is isomorphic to $\pi_2(f,A)_\ast(P)$ in $\cat{C}/\ctxext{\Gamma}{A}$,
+it follows that the sets $T(f_\ast(P))$ and $T(\pi_2(f,A)_\ast(P))$ are isomorphic
+via composition with $\varphi_{f,P}$. By \autoref{lem:ptm}, we have a function
+from $T(P)$ to $T(\pi_2(f,A)_\ast(P))$.
+\end{proof}
+
+\begin{defn}
+For any $\Gamma\in\cat{C}$, define $\cat{C}_\cat{F}/\Gamma$ to be the full
+subcategory of $\cat{C}/\Gamma$ with the same objects as $\cat{F}/\Gamma$. 
+\end{defn}
+
+\begin{thm}
+Let $\mathbb{E}$ be an E'-system. Then $f_\ast$ can be extended to a functor
+$\cat{C}_\cat{F}/\Gamma\to\cat{C}_\cat{F}/\Delta$, for each $f:\Delta\to\Gamma$, and moreover the diagram
+\begin{equation*}
+\begin{tikzcd}[column sep=large]
+\cat{C}_\cat{F}/\ctxext{\Gamma}{B} \arrow[r,"{f_\ast/B}"] \arrow[d,swap,"{g_\ast}"]
+& \cat{C}_\cat{F}/\ctxext{\Delta}{f_\ast(B)} \arrow[d,"{(f_\ast(g))_\ast}"] \\
+\cat{C}_\cat{F}/\ctxext{\Gamma}{A} \arrow[r,swap,"{f_\ast/A}"]
+& \cat{C}_\cat{F}/\ctxext{\Delta}{f_\ast(A)}
+\end{tikzcd}
+\end{equation*}
+commutes for every $g:A\to B$ in $\cat{C}_\cat{F}/\Gamma$.
+\end{thm}
+
+\begin{proof}
+For any $A,B\in\cat{F}/\Gamma$, we have an isomorphism
+$\mathrm{hom}_{\cat{C}/\Gamma}(A,B)\cong T(W_A(B))$. We also get a function
+$T(W_A(B))\to T(f_\ast(W_A(B)))$. Therefore, it suffices to show that
+$f_\ast(W_A(B))\cong W_{f_\ast(A)}(f_\ast(B))$. 
+\begin{align*}
+f_\ast(W_A(B))
+& \jdeq f_\ast(A_\ast(B)) \\
+& \cong \pi_2(f,A)_\ast(A_\ast(B)) \\
+& \jdeq (A\circ\pi_2(f,A))_\ast(B) \\
+& \jdeq (f\circ f_\ast(A))_\ast(B) \\
+& \jdeq f_\ast(A)_\ast(f_\ast(B)) \\
+& \jdeq W_{f_\ast(A)}(f_\ast(B)). \qedhere
+\end{align*}
 \end{proof}
@@
 
-Note first that for any $A\in\cat{F}/\Gamma$, we have
-$W_A\defeq A_\ast:\cat{F}/\Gamma\to\cat{F}/\ctxext{\Gamma}{A}$. Then $W_A$ is
-contravariant in $A$ and that it preserves the terminal object. We give $W_A$
-an action on terms by defining
-\begin{equation*}
-W_A(\theta)\defeq
-\end{equation*}
-for any $\theta\in T(Q)$ and $Q\in\cat{F}/\ctxext{\Gamma}{B}$.
-
-Thus, the $W_A$
-assemble a pre-weakening structure.
-
-Note that for any $A\in\cat{F}/\Gamma$, we have
-\begin{align*}
-T(W_A(A)) & \jdeq \{\eta:\ctxext{\Gamma}{A}\to\ctxext{{\Gamma}{A}}{W_A(A)}\mid W_A(A)\circ\eta\jdeq\catid{\ctxext{\Gamma}{A}}\} \\
-  & \cong
-\{\theta:\ctxext{\Gamma}{A}\to\ctxext{\Gamma}{A}\mid A\circ\theta\jdeq A\},
-\end{align*}
-so we find $\idtm{A}\defeq\zeta_{A,A}(\catid{\ctxext{\Gamma}{A}})\in T(W_A(A))$.
+Note first that for any $A\in\cat{F}/\Gamma$, $W_A$ is
+contravariant in $A$ and that it preserves the terminal object.
+
 Now let $\theta:\Delta\to\ctxext{\Gamma}{A}$ so that $A\circ\theta\jdeq f$

### 2015-06-30 main.tex
E'-systems
comprehension categories, cwfs

@@
 
+%\tableofcontents
+
 \section{Introduction}
-
-When type theory is regarded as a foundational system for doing mathematics,
-we would first of all like it to have finitely many axioms which do not depend
-on any presumed mathematical structures. In particular, natural numbers should
-not be needed to state what type dependency is. Nevertheless, many modern
-accounts on what type theory is, do rely on the existence of the natural
-numbers in the meta-theory. The type theory of \cite{TheBook} uses an infinte
-hierarchy $\UU_0$, $\UU_1$, $\UU_2$, \ldots of universes, as do the type
-type theories of \Coq and \Agda. Likewise, models of type theory considered
-by Cartmell, Voevodsky and Garner do rely on the natural numbers.
-
-That leaves Dybjer's categories with families as
-
-\subsection*{Existing models of type theory}
 
@@
 
-\subsection*{The essentially algebraic theory of E-systems}
+\subsection*{Setting up the essentially algebraic theory of E-systems}
+The theory of type dependency as we present it here, consists of three sorts
+$C$, $F$ and $T$ for contexts, families and terms. There is a map 
+$\eft:F\to C$ which determines for each family, in which context it is a
+family; and there is a map $\ebd:T\to F$ which determines for each term, of
+which family it is a term. Thus, the backbones of an E-system can be represented
+as
+\begin{equation*}
+\begin{tikzcd}
+T \arrow[d,"\ebd"] \\ F \arrow[d,"\eft"] \\ C
+\end{tikzcd}
+\end{equation*}
+When $A$ is a family in context $\Gamma$, there is also a context $\ctxext{\Gamma}{A}$,
+called the \emph{context extension of $A$}. Context extension allows us to consider
+`families over families', where a family $P$ over $A$ in context $\Gamma$, would
+simply be a family $P$ over $\ctxext{\Gamma}{A}$. Context extension is represented
+by another function $\cftctxext : F\to C$.
+
+Since we want the rules of
+type theory to be valid in any context $\Gamma$, there is also a notion of family
+extension. Thus, for any family $P$ over $A$ in context $\Gamma$, there is a
+family $\ctxext{A}{P}$ in context $\Gamma$. In other words, we have a function
+$\cftfamext : F\times_{\cftctxext,\eft} F\to F$.
+
+Two associativity laws will require
+that context and family extension are compatible with each other: we have
+the equality $\ctxext{{\Gamma}{A}}{P}\jdeq\ctxext{\Gamma}{{A}{P}}$ of contexts,
+and the equality $\ctxext{{A}{P}}{Q}\jdeq\ctxext{A}{{P}{Q}}$ of families.
+
+Finally, there is an empty context $\emptyc\in C$, and an empty family $\emptyf[\Gamma]$
+in context $\Gamma$, for every $\Gamma\in C$. The empty context is such that context extension
+restricted to families over the empty context is an isomorphism, and the empty
+families are neutral with respect to context and family extension in the sense
+that $\ctxext{\Gamma}{\emptyf[\Gamma]}\jdeq\Gamma$, $\ctxext{\emptyf[\Gamma]}{A}
+\jdeq A$ and $\ctxext{A}{\emptyf[\ctxext{\Gamma}{A}]}\jdeq A$.
+
+One may recognize in these ingredients of E-systems a category with a
+terminal object, and a set $T(f)$ of terms for every morphism $f$ (satisfying
+no further axioms, so far). Ignoring the set of terms for the moment, $C$ is
+the set of objects and $F$ is the set of morphisms; $\eft$ assigns the codomain
+to each morphism and $\cftctxext$ assigns the domain to each  morphism. The
+operation $\cftfamext$ composes two morphisms. The first associativity rule
+asserts that the domain of the composition $\ctxext{A}{P}$ is the domain of
+$P$, and the second associativity rule asserts that composition is associative
+in the usual sense. The empty context satisfies exactly the properties of the
+terminal object, and the empty family $\emptyf[\Gamma]$ satisfies the properties
+of the identity morphisms.
+
+Thus, an E-system $\mathbf{E}$ will have an underlying category, which may be called the
+category of families of $\mathbf{E}$. The structure of type dependency will by
+built on top of this category. This structure will consist of an E-homomorphism
+$W_A$ called \emph{weakening by $A$}, for every family $A$ in context $\Gamma$,
+an E-homomorphism $S_x$ called \emph{substitution by $x$}, for every term
+$x$ of $A$ in context $\Gamma$, and units $\idtm{A}\in T(W_A(A))$ for every
+family $A$ in context $\Gamma$, satisfying a handful of further axioms. 
+The weakening and substitution homomorphisms
+will preserve the extension structure we have discussed above; in other words,
+they will be functors preserving the terminal object.
+
+\subsection*{The pre-categorical nature of E-systems}
+Although we start our definition of E-systems with a category with term structure,
+we do not set out to have a categorical notion of E-systems, in the sense that the structure of
+being an E-system can be transported along equivalences of categories. The goal
+in this paper is to describe the algebraic structures belonging to an essentially
+algebraic theory of type dependency, and this structure turns out to be pre-categorical
+rather than categorical. The fact that we've encountered a category at the outset
+of our definition of E-systems is in a sense coincidence, rather than born out
+of a desire to do type theory categorically and we would consider E-systems to
+be pre-categorical models of type theory rather than categorical models of type
+theory. The difference between pre-categories and categories is that equality
+of objects in a pre-category is considered to be (strict) equality, whereas
+equality of objects in a category is considered to be isomorphism of objects,
+and we have chosen to use the former notion of equality.
+
+\subsection*{Existing notions of categorical models of dependent type theory}
+
+When introducing a new notion of model of dependent type theory, one necessarily
+faces the task of comparing the new notion with existing notions, and there are
+quite a few.
+
+\begin{itemize}
+\item Contextual categories were introduced by Cartmell \cite{Cartmell1986} and were studied
+extensively by Voevodsky in \cite{VV_Csys_univ,VV_C-systems_monad,VV_C-systems_quotients}
+under the name C-systems.
+\item Categories with families, introduced by Dybjer in \cite{Dybjer1996}, model a
+theory of type dependency with sorts for contexts, context morphisms, types
+and terms. Among the virtues of categories with families are that its theory
+is essentially algebraic, and they are relatively simple to work with.
+
+Categories with families are models of a dependent type theory with a
+category of contexts, for each context $\Gamma$ a sort $\mathrm{Ty}(\Gamma)$ of
+types in context $\Gamma$, and for each type $A\in\mathrm{Ty}(\Gamma)$ a sort
+$\mathrm{Tm}(A)$ of terms of type $A$ in context $\Gamma$. We will see that
+the sorts $\mathrm{Tm}(A)$ are superfluous, because it will always be 
+naturally isomorphic to the set of sections of the projection morphism
+$p_A:\ctxext{\Gamma}{A}\to\Gamma$ in the category of contexts. The evident
+functor from categories with families to E-systems will forget that the contexts
+form a category, and therefore it will not be faithful.
+\item Comprehension categories were introduced by Jacobs in \cite{Jacobs1993}, as
+a minimal notion of categorical model in type theory, without really making
+precise what that means. 
+Comprehension categories are models of a dependent type theory with a
+category of context, and for each context $\Gamma$ a category of types. 
+\end{itemize}
+
+There is a simple test to quickly figure out some basic differences between different
+dependent type theories, asking what kind of structure remains if there are no
+types at all. In a contextual category, only the terminal object remains.
+However, when the presheaf $\mathrm{Ty}$ of a category with families is assumed
+to be empty, or when a category $\cat{E}$ fibered over some category is assumed
+to be empty, the underlying category of context remains. We shall see that
+E-systems are like contextual categories in this respect: only a singleton can
+be an E-system with no types. This complicates a thorough comparison between
+categories with families and comprehension categories on one side, and E-systems
+on the other. We shall construct functors from the categories of categories with
+families and comprehension categories directly preserving type dependency, but
+these functors will not be faithful.
+
+Thus, it seems that both categories with families and comprehension categories
+have more structure than needed to model dependent type theory.
 
@@
 
-In \autoref{sec:esys_props}, we derive several properties from the axioms of
-E-systems. This section is meant to give some insight in what can generally be
-done with E-systems, and provide some of the intuition behind the constructions
-in \autoref{sec:esys_compared}.
+In \autoref{sec:esys_props}, we introduce the auxiliary notion of E'-system.
+An E'-system consists of a pre-category $\cat{C}$ with a sub-pre-category
+$\cat{F}$ with the same objects as $\cat{C}$, and a functorial choice of
+pullbacks for any morphism $f:\Delta\to\Gamma$ in $\cat{C}$ and any
+$A\in\cat{F}/\Gamma$. The goal of this section is to show that 
+E'-systems and E-systems are different axiomatizations of the same structures,
+and apart from that the goal is to show some equational reasoning in E-systems.
+The properties proven in this section also serve to provide some of the 
+intuition behind the constructions in \autoref{sec:esys_compared}.
 
@@
 \begin{rmk}
-It will be convenient to denote the objects of the category $\cat{C}$ with capital
+When we think of a category $\cat{F}$ as a category of fibrations or families
+over a context -- which will be the outset of the definition of E-systems --
+it will be convenient to denote the objects of the $\cat{F}$ with capital
 greek letters $\Gamma$, $\Delta$, \ldots, and the morphisms by capital latin letters
-$A$, $B$, $P$, $Q$, etcetera. When $A$ is a morphism into $\Gamma$, we will denote
-its codomain by $\ctxext{\Gamma}{A}$. When $A$ is a morphism into $\Gamma$ and
-$P$ is a morphism into $\ctxext{\Gamma}{A}$, it will be convenient to denote the
-composition of $P$ with $A$ by $\ctxext{A}{P}$ as well. Note that $\ctxext{A}{P}$
-is indeed the domain of $P$ in the slice category $\cat{C}/\Gamma$. Also, objects
-of $\cat{C}$ will always have a capital greek letter in their notation, e.g.~as
-in $\ctxext{{\Gamma}{A}}{P}$. 
+$A$, $B$, $P$, $Q$, \ldots. When $A$ is a morphism into $\Gamma$, we will denote
+its domain by $\ctxext{\Gamma}{A}$, so we don't always have to introduce a name
+for the domain. When $A$ is a morphism into $\Gamma$ and
+$P$ is a morphism into $\ctxext{\Gamma}{A}$, it will also be convenient to denote the
+composition of $P$ with $A$ by $\ctxext{A}{P}$. Note that $\ctxext{A}{P}$
+is indeed the domain of $P$ in the slice category $\cat{F}/\Gamma$.
+\end{rmk}
+
+\begin{rmk}
+We have chosen not to use common type theoretical notation involving special
+expressions for judgments and for inference rules. One reason is that it has been
+proven to be hard for the uninitiated to read and comprehend such expressions.
+Another is that set theoretic
+notation serves very well for getting the algebraic structure accross, which
+is what we want to focus on.
+
+We would also like to remark that although it appears that we introduce possibly
+large families of operations (a weakening operation $W_A$ for each morphism $A$,
+for instance), E-systems can actually be defined with finitely many
+operations. This becomes clearer when one defines an E-system in an arbitrary
+category with finite limits.
 \end{rmk}

### 2015-06-30 weakeningcats.tex
comprehension categories, cwfs

@@
 \begin{defn}
-A \emph{pre-weakening structure on a pre-category $\cat{C}$} consists of a functor
-$W_A:\cat{C}/\Gamma\to \cat{C}/\ctxext{\Gamma}{A}$ for every $A:\ctxext{\Gamma}{A}\to \Gamma$ in $\cat{C}$ such 
+A \emph{pre-weakening structure on a pre-category $\cat{F}$} consists of a functor
+$W_A:\cat{F}/\Gamma\to \cat{F}/\ctxext{\Gamma}{A}$ for every $A:\ctxext{\Gamma}{A}\to \Gamma$ in $\cat{F}$ such 
 that
 \begin{enumerate}
-\item $W_{\catid{\Gamma}}\jdeq \catid{\cat{C}/\Gamma}$ for every object $\Gamma\in\cat{C}$.
-\item $W_{\ctxext{A}{P}}\jdeq W_P\circ W_A$ for every $P\in\cat{C}/\ctxext{\Gamma}{A}$ and $A\in\cat{C}/\Gamma$.
+\item $W_{\catid{\Gamma}}\jdeq \catid{\cat{F}/\Gamma}$ for every object $\Gamma\in\cat{F}$.
+\item $W_{\ctxext{A}{P}}\jdeq W_P\circ W_A$ for every $P\in\cat{F}/\ctxext{\Gamma}{A}$ and $A\in\cat{F}/\Gamma$.
 \item $W_A$ preserves the final object, i.e.~$W_A(\catid{\Gamma})\jdeq \catid{\ctxext{\Gamma}{A}}$.
@@
 \begin{defn}
-A \emph{pre-weakening homomorphism} $F:\cat{C}\to\cat{D}$ between pre-weakening systems
-is a functor $F:\cat{C}\to\cat{D}$ such that the square
+A \emph{pre-weakening homomorphism} $F:\cat{F}\to\cat{D}$ between pre-weakening systems
+is a functor $F:\cat{F}\to\cat{D}$ such that the square
 \begin{equation*}
 \begin{tikzcd}[column sep=large]
-\cat{C}/\ctxext{\Gamma}{A}
+\cat{F}/\ctxext{\Gamma}{A}
   \arrow[r,"F/\ctxext{\Gamma}{A}"]
@@
   \\
-\cat{C}/\Gamma
+\cat{F}/\Gamma
   \arrow[u,"W_A"]
@@
 \end{equation*}
-commutes for any $A\in\cat{C}/\Gamma$.
+commutes for any $A\in\cat{F}/\Gamma$.
 \end{defn}
@@
 \begin{lem}
-Let $\cat{C}$ be a pre-weakening system. Then for any object $\Gamma$ of $\cat{C}$,
-the slice pre-category $\cat{C}/\Gamma$ is also a pre-weakening system, with the weakening
-structure given by $W(\cat{C}/\Gamma)_P\jdeq W(\cat{C})_P$ for each $P\in\cat{C}/\ctxext{\Gamma}{A}$
-and each $A\in\cat{C}/\Gamma$.
+Let $\cat{F}$ be a pre-weakening system. Then for any object $\Gamma$ of $\cat{F}$,
+the slice pre-category $\cat{F}/\Gamma$ is also a pre-weakening system, with the weakening
+structure given by $W(\cat{F}/\Gamma)_P\jdeq W(\cat{F})_P$ for each $P\in\cat{F}/\ctxext{\Gamma}{A}$
+and each $A\in\cat{F}/\Gamma$.
 \end{lem}
@@
 \begin{tikzcd}[column sep=huge]
-\cat{C}/\ctxext{{\Gamma}{B}}{Q}
+\cat{F}/\ctxext{{\Gamma}{B}}{Q}
   \arrow[r,"W_A/\ctxext{B}{Q}"]
   &
-\cat{C}/\ctxext{{\Gamma}{A}}{W_A(\ctxext{B}{Q})}
-  \\
-\cat{C}/\ctxext{\Gamma}{B}
+\cat{F}/\ctxext{{\Gamma}{A}}{W_A(\ctxext{B}{Q})}
+  \\
+\cat{F}/\ctxext{\Gamma}{B}
   \arrow[u,"W_Q"]
@@
   &
-\cat{C}/\ctxext{{\Gamma}{A}}{W_A(B)}
+\cat{F}/\ctxext{{\Gamma}{A}}{W_A(B)}
   \arrow[u,swap,"W_{W_A(Q)}"]
@@
 \end{equation*}
-for each $A,B\in\cat{C}/\Gamma$ and $Q:\cat{C}/\ctxext{\Gamma}{B}$. The objects part of this 
-property asserts that for any $k\in \cat{C}/\greek{E}$, the dotted arrows in the diagram
+for each $A,B\in\cat{F}/\Gamma$ and $Q:\cat{F}/\ctxext{\Gamma}{B}$. The objects part of this 
+property asserts that for any $k\in \cat{F}/\greek{E}$, the dotted arrows in the diagram
 \begin{equation*}
@@
 \begin{tikzcd}[column sep=huge]
-\cat{C}/\ctxext{\Gamma}{C}
+\cat{F}/\ctxext{\Gamma}{C}
   \arrow[r,"W_A/C"]
   &
-\cat{C}/\ctxext{{\Gamma}{A}}{W_A(C)}
-  \\
-\cat{C}/\Gamma
+\cat{F}/\ctxext{{\Gamma}{A}}{W_A(C)}
+  \\
+\cat{F}/\Gamma
   \arrow[u,"W_C"]
@@
   &
-\cat{C}/\ctxext{\Gamma}{A}
+\cat{F}/\ctxext{\Gamma}{A}
   \arrow[u,swap,"W_{W_A(C)}"]
@@
 \end{equation*}
-commutes for every $A,C\in\cat{C}/\Gamma$. In particular, we see
-that $W_A(W_C(D))\jdeq W_{W_A(C)}(W_A(D))$ for any $D\in\cat{C}/\Gamma$,
+commutes for every $A,C\in\cat{F}/\Gamma$. In particular, we see
+that $W_A(W_C(D))\jdeq W_{W_A(C)}(W_A(D))$ for any $D\in\cat{F}/\Gamma$,
 i.e.~that weakening is a self-distributive operation. 
@@
 \begin{cor}
-For any object $\Gamma$ of a weakening system $\cat{C}$, the slice pre-category
-$\cat{C}/\Gamma$ is a weakening system.
+For any object $\Gamma$ of a weakening system $\cat{F}$, the slice pre-category
+$\cat{F}/\Gamma$ is a weakening system.
 \end{cor}
@@
 \begin{lem}
-Let $G$ be a graph, and let $\cat{C}_G$ be the free pre-category generated by
-$G$. A pre-weakening structure on $\cat{C}_G$ is determined by functors
-$W_e:\cat{C}_G/\Gamma\to\cat{C}_G/\Delta$ for every edge $e:\Delta\to \Gamma$ in $G$, with the
+Let $G$ be a graph, and let $\cat{F}_G$ be the free pre-category generated by
+$G$. A pre-weakening structure on $\cat{F}_G$ is determined by functors
+$W_e:\cat{F}_G/\Gamma\to\cat{F}_G/\Delta$ for every edge $e:\Delta\to \Gamma$ in $G$, with the
 property that $W_e(\catid{\Gamma})\jdeq \catid{\Delta}$.
 
-A weakening structure on $\cat{C}_G$ is
-a pre-weakening structure on $\cat{C}_G$ such that the diagram
+A weakening structure on $\cat{F}_G$ is
+a pre-weakening structure on $\cat{F}_G$ such that the diagram
 \begin{equation*}
 \begin{tikzcd}[column sep=huge]
-\cat{C}_G/\greek{Z} \arrow[r,"W_e/(g\circ h)"] & \cat{C}_G/\mathrm{dom}(W_e(g\circ h)) \\
-\cat{C}_G/\greek{E} \arrow[u,"W_h"] \arrow[r,swap,"W_e/g"] & \cat{C}_G/\mathrm{dom}(W_e(g)) \arrow[u,swap,"W_{W_e(h)}"]
-\end{tikzcd}
-\end{equation*}
-commutes for every arrow $g:\greek{E}\to \Gamma$ in $\cat{C}_G$ and every two edges $e:\Delta\to \Gamma$ and
+\cat{F}_G/\greek{Z} \arrow[r,"W_e/(g\circ h)"] & \cat{F}_G/\mathrm{dom}(W_e(g\circ h)) \\
+\cat{F}_G/\greek{E} \arrow[u,"W_h"] \arrow[r,swap,"W_e/g"] & \cat{F}_G/\mathrm{dom}(W_e(g)) \arrow[u,swap,"W_{W_e(h)}"]
+\end{tikzcd}
+\end{equation*}
+commutes for every arrow $g:\greek{E}\to \Gamma$ in $\cat{F}_G$ and every two edges $e:\Delta\to \Gamma$ and
 $h:\greek{Z}\to \greek{E}$ in $G$.
@@
 Let $G$ be the graph consisting of one vertex $\ast$ and one edge $e:\ast\to\ast$.
-Then $\cat{C}_G/\ast\jdeq \N$. Then the only weakening structure on $\cat{C}/G$ is
+Then $\cat{F}_G/\ast\jdeq \N$. Then the only weakening structure on $\cat{F}/G$ is
 given by $W_e\jdeq \catid{\N}$. The reason is that 
@@
 \begin{defn}
-A \emph{category with a term structure} is a pre-category $\cat{C}$ with a set $T(A)$ for every
-$A\in\cat{C}/\Gamma$ for every $\Gamma\in\cat{C}$. A \emph{functor with term structure} from $\cat{C}$ to $\cat{D}$
-is a functor $F:\cat{C}\to\cat{D}$ with a function $T(A)\to T(F(A))$ for
-every $A\in\cat{C}/\Gamma$, for every $\Gamma\in\cat{C}$.
-\end{defn}
-
-\begin{defn}
-Suppose $\cat{C}$ is a pre-category with term structure, and $\Gamma$ is an object of $\cat{C}$. 
-Then $\cat{C}/\Gamma$ is a pre-category with term structure, where $T_{\cat{C}/\Gamma}(A)\jdeq T_\cat{C}(A)$.
+A \emph{category with a term structure} is a pre-category $\cat{F}$ with a set $T(A)$ for every
+$A\in\cat{F}/\Gamma$ for every $\Gamma\in\cat{F}$. A \emph{functor with term structure} from $\cat{F}$ to $\cat{D}$
+is a functor $F:\cat{F}\to\cat{D}$ with a function $T(A)\to T(F(A))$ for
+every $A\in\cat{F}/\Gamma$, for every $\Gamma\in\cat{F}$.
+\end{defn}
+
+\begin{defn}
+Suppose $\cat{F}$ is a pre-category with term structure, and $\Gamma$ is an object of $\cat{F}$. 
+Then $\cat{F}/\Gamma$ is a pre-category with term structure, where $T_{\cat{F}/\Gamma}(A)\jdeq T_\cat{F}(A)$.
 \end{defn}
@@
 For any object $\Gamma$ of a weakening system with term structure, the slice pre-category
-$\cat{C}/\Gamma$ is a weakening system with term structure.
+$\cat{F}/\Gamma$ is a weakening system with term structure.
 \end{cor}
@@
 A \emph{pre-projection system} is a pre-weakening system with term structure with an element
-$\tfid{A}\in T(W_A(A))$ for every $A\in\cat{C}/\Gamma$ and $\Gamma\in\cat{C}$. 
-\end{defn}
-
-\begin{defn}
-A \emph{pre-projection homomorphism} $F:\cat{C}\to\cat{D}$ is a pre-weakening homomorphism for which
+$\tfid{A}\in T(W_A(A))$ for every $A\in\cat{F}/\Gamma$ and $\Gamma\in\cat{F}$. 
+\end{defn}
+
+\begin{defn}
+A \emph{pre-projection homomorphism} $F:\cat{F}\to\cat{D}$ is a pre-weakening homomorphism for which
 \begin{equation*}
@@
 \end{equation*}
-for every $A\in\cat{C}/\Gamma$ and $\Gamma\in\cat{C}$.
+for every $A\in\cat{F}/\Gamma$ and $\Gamma\in\cat{F}$.
 \end{defn}
@@
 \begin{lem}
-For any object $\Gamma$ in a pre-projection system $\cat{C}$, the slice pre-category
-$\cat{C}/\Gamma$ is a pre-projection system.
+For any object $\Gamma$ in a pre-projection system $\cat{F}$, the slice pre-category
+$\cat{F}/\Gamma$ is a pre-projection system.
 \end{lem}
@@
 \begin{cor}
-For any object $\Gamma$ of a projection system $\cat{C}$, the slice pre-category $\cat{C}/\Gamma$
+For any object $\Gamma$ of a projection system $\cat{F}$, the slice pre-category $\cat{F}/\Gamma$
 is a projection system.
@@
 \begin{defn}
-A \emph{pre-substitution structure} on a pre-category with term structure $\cat{C}$ consists
-of a functor with term structure $S_x:\cat{C}/\ctxext{\Gamma}{A}\to\cat{C}/\Gamma$ for every $x\in T(A)$ and
-$A\in\cat{C}/\Gamma$, such that $S_x(\catid{\ctxext{\Gamma}{A}})\jdeq\catid{\Gamma}$.
+A \emph{pre-substitution structure} on a pre-category with term structure $\cat{F}$ consists
+of a functor with term structure $S_x:\cat{F}/\ctxext{\Gamma}{A}\to\cat{F}/\Gamma$ for every $x\in T(A)$ and
+$A\in\cat{F}/\Gamma$, such that $S_x(\catid{\ctxext{\Gamma}{A}})\jdeq\catid{\Gamma}$.
 
@@
 \begin{defn}
-A \emph{pre-substitution homomorphism} $F:\cat{C}\to\cat{D}$ is a functor with term structure for
+A \emph{pre-substitution homomorphism} $F:\cat{F}\to\cat{D}$ is a functor with term structure for
 which the diagram
@@
 \begin{tikzcd}[column sep=large]
-\cat{C}/\ctxext{\Gamma}{A}
+\cat{F}/\ctxext{\Gamma}{A}
   \arrow[r,"F/\ctxext{\Gamma}{A}"]
@@
   \\
-\cat{C}/\Gamma
+\cat{F}/\Gamma
   \arrow[r,swap,"F/\Gamma"]
@@
 \end{equation*}
-commutes for every $x\in T(A)$ and $A\in\cat{C}/\Gamma$.
+commutes for every $x\in T(A)$ and $A\in\cat{F}/\Gamma$.
 \end{defn}
@@
 \begin{lem}
-For any object $\Gamma$ of a pre-substitution system $\cat{C}$, the slice pre-category
-$\cat{C}/\Gamma	$ is a pre-substitution system.
+For any object $\Gamma$ of a pre-substitution system $\cat{F}$, the slice pre-category
+$\cat{F}/\Gamma	$ is a pre-substitution system.
 \end{lem}
@@
 \begin{cor}
-For any object $\Gamma$ of a substitution system, the slice pre-category $\cat{C}/\Gamma$
+For any object $\Gamma$ of a substitution system, the slice pre-category $\cat{F}/\Gamma$
 is a substitution system.
@@
 \begin{tikzcd}[column sep=large]
-\cat{C}/\ctxext{{{\Gamma}{A}}{P}}{Q}
+\cat{F}/\ctxext{{{\Gamma}{A}}{P}}{Q}
   \arrow[r,"S_x/\ctxext{P}{Q}"]
@@
   &
-\cat{C}/\ctxext{{\Gamma}{S_x(P)}}{S_x(Q)}
+\cat{F}/\ctxext{{\Gamma}{S_x(P)}}{S_x(Q)}
   \arrow[d,"S_{S_x(y)}"]
   \\
-\cat{C}/\ctxext{{\Gamma}{A}}{P}
+\cat{F}/\ctxext{{\Gamma}{A}}{P}
   \arrow[r,swap,"S_x/P"]
   &
-\cat{C}/\ctxext{\Gamma}{S_x(P)}
+\cat{F}/\ctxext{\Gamma}{S_x(P)}
 \end{tikzcd}
@@
 \begin{defn}
-A \emph{pre-E-system} is a pre-category $\cat{C}$ with term structure
+A \emph{pre-E-system} is a pre-category $\cat{F}$ with term structure
 with a pre-weakening structure $W$, a pre-substitution structure $S$ and a terminal
@@
 \item each $S_x$ is a unital pre-E-homomorphism,
-\item \label{tTf:StWf_id} weakened families are constant, i.e.~$S_x\circ W_A\jdeq \catid{\cat{C}/\Gamma}$ for any $x\in T(A)$ and $A\in\cat{C}/\Gamma$,
-\item $S_x(\tfid{A})\jdeq x$ for any $x\in T(A)$ and $A\in\cat{C}/\Gamma$, and
-\item \label{f:SidfWff_id} $S_{\tfid{A}}\circ W_A/A\jdeq \catid{\cat{C}/\ctxext{\Gamma}{A}}$ for any $A\in\cat{C}/\Gamma$.
+\item \label{tTf:StWf_id} weakened families are constant, i.e.~$S_x\circ W_A\jdeq \catid{\cat{F}/\Gamma}$ for any $x\in T(A)$ and $A\in\cat{F}/\Gamma$,
+\item $S_x(\tfid{A})\jdeq x$ for any $x\in T(A)$ and $A\in\cat{F}/\Gamma$, and
+\item \label{f:SidfWff_id} $S_{\tfid{A}}\circ W_A/A\jdeq \catid{\cat{F}/\ctxext{\Gamma}{A}}$ for any $A\in\cat{F}/\Gamma$.
 \end{enumerate}
 An \emph{E-homomorphism} $H:\mathbb{E}\to\mathbb{D}$ is a functor from
-$\cat{C}_{\mathbb{E}}$ to $\cat{C}_{\mathbb{D}}$ which is both a projection
+$\cat{F}_{\mathbb{E}}$ to $\cat{F}_{\mathbb{D}}$ which is both a projection
 homomorphism and a substitution homomorphism.
@@
 \begin{tikzcd}[column sep=huge]
-\cat{C}/\ctxext{{\Gamma}{B}}{Q}
+\cat{F}/\ctxext{{\Gamma}{B}}{Q}
   \arrow[r,"W_A/\ctxext{B}{Q}"]
@@
   &
-\cat{C}/\ctxext{{{\Gamma}{A}}{W_A(B)}}{W_A(Q)}
+\cat{F}/\ctxext{{{\Gamma}{A}}{W_A(B)}}{W_A(Q)}
   \arrow[d,"S_{W_A(y)}"]
   \\
-\cat{C}/\ctxext{\Gamma}{B}
+\cat{F}/\ctxext{\Gamma}{B}
   \arrow[r,swap,"W_A/B"]
   &
-\cat{C}/\ctxext{{\Gamma}{A}}{W_A(B)}
+\cat{F}/\ctxext{{\Gamma}{A}}{W_A(B)}
 \end{tikzcd}
@@
 of functors with term structure commutes for every 
-$Q\in\cat{C}/\ctxext{\Gamma}{B}$, $B\in\cat{C}/\Gamma$ and each $y\in T(Q)$.
+$Q\in\cat{F}/\ctxext{\Gamma}{B}$, $B\in\cat{F}/\Gamma$ and each $y\in T(Q)$.
 
@@
 \begin{tikzcd}[column sep=huge]
-\cat{C}/\ctxext{{\Gamma}{A}}{P}
+\cat{F}/\ctxext{{\Gamma}{A}}{P}
   \arrow[r,"S_x/P"]
@@
   &
-\cat{C}/\ctxext{\Gamma}{S_x(P)}
+\cat{F}/\ctxext{\Gamma}{S_x(P)}
   \arrow[d,"W_{S_x(Q)}"]
   \\
-\cat{C}/\ctxext{{{\Gamma}{A}}{P}}{Q}
+\cat{F}/\ctxext{{{\Gamma}{A}}{P}}{Q}
   \arrow[r,swap,"S_x/\ctxext{P}{Q}"]
   &
-\cat{C}/\ctxext{{\Gamma}{S_x(P)}}{S_x(Q)}
+\cat{F}/\ctxext{{\Gamma}{S_x(P)}}{S_x(Q)}
 \end{tikzcd}
@@
 of functors with term structure commutes for every 
-$Q\in\cat{C}/\ctxext{{\Gamma}{A}}{P}$.
+$Q\in\cat{F}/\ctxext{{\Gamma}{A}}{P}$.
 \end{rmk}
@@
 \begin{cor}
-For any object $\Gamma$ of an E-system $\cat{C}$, the slice pre-category $\cat{C}/\Gamma$
+For any object $\Gamma$ of an E-system $\cat{F}$, the slice pre-category $\cat{F}/\Gamma$
 is an E-system.
@@
 
-\section{Properties of E-systems}\label{sec:esys_props}
+\section{The category of internal morphisms of an E-system}\label{sec:esys_props}
 
@@
 projections and their sections. We elaborate on this point of view by defining
-the category of internal homomorphisms, which Lawvere might have called the
+the category of internal morphisms, which Lawvere might have called the
 category of terms. 
 In the category of internal morphisms over a context $\Gamma$,
+which we define in \autoref{subsec:im_cat},
 a morphism $f:A\to B$ is a term of the constant family $W_A(B)$ over 
@@
 $f_\ast$, which is going to be an E-homomorphism from 
-$\cat{C}/\ctxext{\Gamma}{B}$ to $\cat{C}/\ctxext{\Gamma}{A}$. This allows us
+$\cat{F}/\ctxext{\Gamma}{B}$ to $\cat{F}/\ctxext{\Gamma}{A}$. This allows us
 to consider diagrams of the following form:
@@
 refer to a family, and an ordinary arrow will refer to an internal morphism,
-possibly over another internal morphism
+possibly over another internal morphism.
 
@@
 over $h:B\to C$ so that we obtain a morphism $\jfcomp{A}{f}{P}{g}{k}:P\to R$ 
-over $\jcomp{A}{f}{h}:A\to C$. We will show interchange laws between these
+over $\jcomp{A}{f}{h}:A\to C$. 
+In \autoref{sec:interchange}, we will show interchange laws between these
 sorts of composition. The interchange laws have as a consequence that if we
@@
 
+The category of internal morphisms forms a category of which the families form
+a subcategory. This, together with the pre-composition operation described
+earlier, gives rise to the auxiliary notion of E'-systems, which we define
+in \autoref{sec:E'}. The thread running through this whole section is proving
+that the notions of E-system and E'-system are the same. In \autoref{sec:E'}
+we show directly that any E'-system is an E-system. We end this section by
+showing that the converse also holds.
+
+Before we begin this section, we introduce some more convenient notation.
+
+\begin{defn}
+Let $A\in\cat{F}/\Gamma$. Recall that $W_A:\cat{F}/\Gamma\to\cat{F}/\ctxext{\Gamma}{A}$ acts on objects,
+morphisms and terms. We introduce the infix form of weakening by $A\in\cat{F}/\Gamma$ to be
+$\ctxwk{A}{\blank}$. Thus, we will write
+\begin{align*}
+\ctxwk{A}{B} & \defeq W_A(B) & & \text{for $B\in\cat{F}/\Gamma$} \\
+\ctxwk{A}{Q} & \defeq W_A(Q) & & \text{for $B\in\cat{F}/\Gamma$ and $Q\in\cat{F}/\ctxext{\Gamma}{B}$}\\
+\ctxwk{A}{g} & \defeq W_A(g) & & \text{for $B\in\cat{F}/\Gamma$, $Q\in\cat{F}/\ctxext{\Gamma}{B}$ and $g\in T(Q)$}
+\end{align*}
+\end{defn}
+
+\begin{defn}
+Let $x\in T(A)$ for a family $A\in\cat{F}/\Gamma$. The infix form of substitution
+by $x$ is taken to be $\subst{x}{\blank}$. Thus, we will write
+\begin{align*}
+\subst{x}{P} & \defeq S_x(P) & & \text{for $P\in\cat{F}/\ctxext{\Gamma}{A}$} \\
+\subst{x}{Q} & \defeq S_x(Q) & & \text{for $P\in\cat{F}/\ctxext{\Gamma}{A}$ and $Q\in\cat{F}/\ctxext{{\Gamma}{A}}{P}$} \\
+\subst{x}{g} & \defeq S_x(g) & & \text{for $P\in\cat{F}/\ctxext{\Gamma}{A}$, $Q\in\cat{F}/\ctxext{{\Gamma}{A}}{P}$ and $g\in T(Q)$}
+\end{align*}
+\end{defn}
+
+\subsection{E'-systems}\label{sec:E'}
+
+E'-systems can be thought of as strict versions of tribes%
+\footnote{Tribes are an invention of Joyal, see \url{http://ncatlab.org/homotopytypetheory/files/Joyal.pdf}}.
+
+\begin{defn}\label{defn:E'sys}
+An \emph{E'-system} consists of a category $\cat{C}$ and a subcategory $\cat{F}$ with
+the same objects as $\cat{C}$, such that
+\begin{enumerate}
+\item $\cat{F}$ has a terminal object.
+\item For any $f:\Delta\to\Gamma$ in $\cat{C}$ and any $A\in\cat{F}/\Gamma$, a
+functorial choice of a pullback square 
+\begin{equation*}
+\begin{tikzcd}[column sep=large]
+\ctxext{\Delta}{f_\ast(A)} \arrow[r,"{\pi_2(f,A)}"] \arrow[d,fib,swap,"{f_\ast(A)}"] & \ctxext{\Gamma}{A} \arrow[d,fib,"A"] \\
+\Delta \arrow[r,"f"] & \Gamma
+\end{tikzcd}
+\end{equation*}
+such that $f_\ast(A)\in\cat{F}/\Delta$. Explicitly, the functoriality
+requirement is that
+\begin{enumerate}
+\item For any $f:\Delta\to\Gamma$, one has 
+$f_\ast(\catid{\Gamma})\jdeq\catid{\Delta}$ and
+\begin{equation*}
+\pi_2(f,\catid{\Gamma})\jdeq f.
+\end{equation*}
+\item \label{defn:E'sys:b} For any $A\in\cat{F}/\Gamma$, one has $(\catid{\Gamma})_\ast(A)\jdeq A$ and
+\begin{equation*}
+\pi_2(\catid{\Gamma},A)\jdeq\catid{\ctxext{\Gamma}{A}}
+\end{equation*}
+\item \label{defn:E'sys:c} For any $f:\Delta\to\Gamma$, $g:X\to\Delta$ and $A\in\cat{F}/\Gamma$, one
+has $(f\circ g)_\ast(A)\jdeq g_\ast(f_\ast(A))$ and 
+\begin{equation*}
+\pi_2(f\circ g,A)\jdeq \pi_2(f,A)\circ\pi_2(g,f_\ast(A))
+\end{equation*}
+\item For any $P\in\cat{F}/\ctxext{\Gamma}{A}$ and $f:\Delta\to\Gamma$, one has
+$f_\ast(P)\jdeq\pi_2(f,A)_\ast(P)$ and 
+\begin{equation*}
+\pi_2(f,\ctxext{A}{P}) \jdeq \pi_2(\pi_2(f,a),P)
+\end{equation*}
+\end{enumerate}
+\end{enumerate}
+For any $f:\Delta\to\Gamma$ we shall write $f_\ast$ for the induced functor
+$\cat{F}/\Gamma\to\cat{F}/\Delta$. 
+\end{defn}
+
+\begin{rmk}
+Without a terminal object in $\cat{F}$, the category $\cat{C}$ does not correspond
+directly to anything in an E-system.
+\end{rmk}
+
+\begin{rmk}
+For E'-systems, we do not require that every pullback of $A\in\cat{F}/\Gamma$ along $f:\Delta\to\Gamma$
+is in $\cat{F}/\Delta$, just that there is a choice of such a pullback which is
+in $\cat{F}/\Delta$. Thus, the definition of E'-system is slightly different from
+Taylor's notion of display map category. Furthermore, the class of display maps in
+a display map category is not required to be closed under composition, and neither
+is there assumed to exist an object $T$ such that for every object $X$ there is a 
+unique display map $X\to T$.
+\end{rmk}
+
+\begin{rmk}
+Also, we do not require that $\cat{C}$ has a terminal object, and neither do
+we require that $\cat{F}$ contains all the isomorphisms of $\cat{C}$. These are
+essential differences with Joyal's notion of tribes.   
+\end{rmk}
+
+\begin{defn}
+For any $\Gamma\in\cat{C}$, define $\cat{C}_\cat{F}/\Gamma$ to be the full
+subcategory of $\cat{C}/\Gamma$ with the same objects as $\cat{F}/\Gamma$. 
+\end{defn}
+
+\begin{lem}\label{lem:pb_selfdistributive}
+Let $\mathbb{E}$ be an E'-system. Then $f_\ast$ can be extended to a functor from
+$\cat{C}_\cat{F}/\Gamma$ to $\cat{C}_\cat{F}/\Delta$, for each $f:\Delta\to\Gamma$, and moreover the diagram
+\begin{equation*}
+\begin{tikzcd}[column sep=large]
+\cat{C}_\cat{F}/\ctxext{\Gamma}{B} \arrow[r,"{f_\ast/B}"] \arrow[d,swap,"{g_\ast}"]
+& \cat{C}_\cat{F}/\ctxext{\Delta}{f_\ast(B)} \arrow[d,"{(f_\ast(g))_\ast}"] \\
+\cat{C}_\cat{F}/\ctxext{\Gamma}{A} \arrow[r,swap,"{f_\ast/A}"]
+& \cat{C}_\cat{F}/\ctxext{\Delta}{f_\ast(A)}
+\end{tikzcd}
+\end{equation*}
+commutes for every $g:A\to B$ in $\cat{C}_\cat{F}/\Gamma$.
+\end{lem}
+
+\begin{proof}
+It is straightforward to use the universal property of the pullback to extend
+$f_\ast$ to a functor from $\cat{C}_\cat{F}/\Gamma$ to $\cat{C}_\cat{F}/\Delta$.
+To see that the diagram commutes, we calculate
+\begin{align*}
+f_\ast/A\circ g_\ast
+& \jdeq
+\pi_2(f,A)_\ast\circ g_\ast \\
+& \jdeq
+(g\circ\pi_2(f,A))_\ast \\
+& \jdeq
+(\pi_2(f,B)\circ f_\ast(g))_\ast \\
+& \jdeq
+(f_\ast(g))_\ast\circ\pi_2(f,B)_\ast \\
+& \jdeq
+(f_\ast(g))_\ast\circ f_\ast/B.\qedhere
+\end{align*}
+\end{proof}
+
+\begin{thm}
+An E-system is the same thing as an E'-system.
+\end{thm}
+
+\begin{proof}
+We define for any $A\in\cat{F}/\Gamma$ the set
+\begin{equation*}
+T(A)\defeq\{x:\Gamma\to\ctxext{\Gamma}{A}\mid A\circ x\jdeq\catid{\Gamma}\}.
+\end{equation*}
+Since each functor $f_\ast$ is extended to all of $\cat{C}_{\cat{F}}/\Gamma$,
+it follows immediately that those are functors with term structure.
+
+We define for any $A\in\cat{F}/\Gamma$, the functor $W_A\defeq A_\ast:\cat{F}/\Gamma\to\cat{F}/\ctxext{\Gamma}{A}$.
+Likewise, we define for any $x\in T(A)$, the functor $S_x\defeq x_\ast:\cat{F}/\ctxext{\Gamma}{A}\to\cat{F}/\Gamma$.
+We also define $\tfid{A}:T(W_A(A))$ by the universal property of pullbacks:
+\begin{equation*}
+\begin{tikzcd}[column sep=large]
+\ctxext{\Gamma}{A} \arrow[drr,bend left=15,equals] \arrow[ddr,bend right=15,equals] \arrow[dr,densely dotted,near end,"{\tfid{A}}"] \\
+& \ctxext{{\Gamma}{A}}{W_A(A)} \arrow[r,swap,"{\pi_2(A,A)}"] \arrow[d,fib] & \ctxext{\Gamma}{A} \arrow[d,fib] \\
+& \ctxext{\Gamma}{A} \arrow[r,swap,"A"] & \Gamma
+\end{tikzcd}
+\end{equation*}
+As an immediate consequence of \autoref{lem:pb_selfdistributive}, we get that
+each functor $W_A$ and $S_x$ is both a weakening functor and a substitution functor.
+Since each $W_A$ and $S_x$ preserves identity morphisms, it also preserves the
+terms $\tfid{A}$.
+
+It remains to verify the rest of the conditions of E-systems.
+\begin{itemize}
+\item To show that weakened families are constant, note that
+\begin{equation*}
+S_x\circ W_A\jdeq x_\ast\circ A_\ast\jdeq (A\circ x)_\ast\jdeq (\catid{\Gamma})_\ast\jdeq \catid{\cat{C}_{\cat{F}}/\Gamma}.
+\end{equation*}
+\item The identity terms are neutral for pre-composition:
+\begin{equation*}
+S_{\tfid{A}}\circ W_A/A \jdeq S_{\tfid{A}}\circ \pi_2(A,A)_\ast \jdeq (\pi_2(A,A)\circ\tfid{A})_\ast \jdeq
+(\catid{\ctxext{\Gamma}{A}})_\ast \jdeq \catid{\cat{C}_{\cat{F}}/\ctxext{\Gamma}{A}} 
+\end{equation*}
+\item The identity terms behave like identity functions: by the universal property,
+$S_x(\tfid{A})$ is the unique section of $A$ such that the square
+\begin{equation*}
+\begin{tikzcd}[column sep=6em]
+\Gamma \arrow[d,swap,"{S_x(\tfid{A})}"] \arrow[r,"{\pi_2(x,\catid{\ctxext{\Gamma}{A}})}"] & \ctxext{\Gamma}{A} \arrow[d,"{\tfid{A}}"] \\
+\ctxext{\Gamma}{A} \arrow[r,swap,"{\pi_2(x,W_A(A))}"] & \ctxext{{\Gamma}{A}}{W_A(A)}
+\end{tikzcd}
+\end{equation*}
+commutes. Thus, it suffices to show that this square also commutes with $x$ in the place of
+$S_x(\tfid{A})$. Note that $\pi_2(x,\catid{\ctxext{\Gamma}{A}})\jdeq x$. 
+Since $\ctxext{{\Gamma}{A}}{W_A(A)}$ is itself a pullback, it suffices
+and it is straightforward to verify the equalities
+\begin{align*}
+W_A(A)\circ\pi_2(x,W_A(A))\circ x & \jdeq W_A(A)\circ\idtm{A}\circ x \\
+\pi_2(A,A)\circ\pi_2(x,W_A(A))\circ x & \jdeq \pi_2(A,A)\circ\idtm{A}\circ x.\qedhere
+\end{align*}
+\end{itemize}
+\end{proof}
+
+\begin{comment}
 We then demonstrate how to define a category internal to an E-system. The main
@@
 defining an internal category was the hardest part of that definition.
-
-Before we begin this section, we introduce some handy shorthands.
-
-\begin{defn}
-Let $A\in\cat{C}/\Gamma$. Recall that $W_A:\cat{C}/\Gamma\to\cat{C}/\ctxext{\Gamma}{A}$ acts on objects,
-morphisms and terms. We introduce the infix form of weakening by $A\in\cat{C}/\Gamma$ to be
-$\ctxwk{A}{\blank}$. Thus, we will write
-\begin{align*}
-\ctxwk{A}{B} & \defeq W_A(B) & & \text{for $B\in\cat{C}/\Gamma$} \\
-\ctxwk{A}{Q} & \defeq W_A(Q) & & \text{for $B\in\cat{C}/\Gamma$ and $Q\in\cat{C}/\ctxext{\Gamma}{B}$}\\
-\ctxwk{A}{g} & \defeq W_A(g) & & \text{for $B\in\cat{C}/\Gamma$, $Q\in\cat{C}/\ctxext{\Gamma}{B}$ and $g\in T(Q)$}
-\end{align*}
-\end{defn}
-
-\begin{defn}
-Let $x\in T(A)$ for a family $A\in\cat{C}/\Gamma$. The infix form of substitution
-by $x$ is taken to be $\subst{x}{\blank}$. Thus, we will write
-\begin{align*}
-\subst{x}{P} & \defeq S_x(P) & & \text{for $P\in\cat{C}/\ctxext{\Gamma}{A}$} \\
-\subst{x}{Q} & \defeq S_x(Q) & & \text{for $P\in\cat{C}/\ctxext{\Gamma}{A}$ and $Q\in\cat{C}/\ctxext{{\Gamma}{A}}{P}$} \\
-\subst{x}{g} & \defeq S_x(g) & & \text{for $P\in\cat{C}/\ctxext{\Gamma}{A}$, $Q\in\cat{C}/\ctxext{{\Gamma}{A}}{P}$ and $g\in T(Q)$}
-\end{align*}
-\end{defn}
+\end{comment}
 
@@
 \begin{defn}
-For every $A,B\in\cat{C}/\Gamma$ we define the set 
+For every $A,B\in\cat{F}/\Gamma$ we define the set 
 \begin{equation*}
@@
 \begin{rmk}
-Note that $\thom{\catid{\Gamma}}{A}\jdeq T(A)$ for any $A\in\cat{C}/\Gamma$, 
+Note that $\thom{\catid{\Gamma}}{A}\jdeq T(A)$ for any $A\in\cat{F}/\Gamma$, 
 because we have 
-$W_{\catid{\ctxext{\Gamma}{A}}}\jdeq \catid{\cat{C}/\ctxext{\Gamma}{A}}$.
+$W_{\catid{\ctxext{\Gamma}{A}}}\jdeq \catid{\cat{F}/\ctxext{\Gamma}{A}}$.
 
 Note also that $\thom{\ctxext{A}{P}}{B}\jdeq \thom{P}{\ctxwk{A}{B}}$ 
-for any $P\in\cat{C}/\ctxext{\Gamma}{A}$ and $B\in\cat{C}/\Gamma$,
+for any $P\in\cat{F}/\ctxext{\Gamma}{A}$ and $B\in\cat{F}/\Gamma$,
 because $W_{\ctxext{A}{P}}\jdeq W_P\circ W_A$.
@@
 \begin{defn}
-Let $A,B\in\cat{C}/\Gamma$.
+Let $A,B\in\cat{F}/\Gamma$.
 For any $f\in\thom{A}{B}$ we define the pre-composition E-homomorphism
 \begin{equation*}
-f_\ast \defeq S_f\circ W_A/B : \cat{C}/\ctxext{\Gamma}{B}\to\cat{C}/\ctxext{\Gamma}{A}.
-\end{equation*}
-Furthermore, for $g\in\thom{B}{C}$ we may consider $W_B(C)\in\cat{C}/\ctxext{\Gamma}{B}$ as a morphism from $W_B(C)$ to $\catid{\ctxext{\Gamma}{B}}$. Therefore, $f_\ast$ acts on the terms of
+f_\ast \defeq S_f\circ W_A/B : \cat{F}/\ctxext{\Gamma}{B}\to\cat{F}/\ctxext{\Gamma}{A}.
+\end{equation*}
+Furthermore, for $g\in\thom{B}{C}$ we may consider $W_B(C)\in\cat{F}/\ctxext{\Gamma}{B}$ as a morphism from $W_B(C)$ to $\catid{\ctxext{\Gamma}{B}}$. Therefore, $f_\ast$ acts on the terms of
 $W_B(C)$, so it makes sense to define $g\circ f\defeq f_\ast(g)$. In fact, it makes
@@
 Note that \autoref{f:SidfWff_id} of \autoref{defn:esystems} asserts precisely
-that ${(\tfid{A})}_\ast\jdeq \catid{\cat{C}/\ctxext{\Gamma}{A}}$ for any $A\in\cat{C}/\Gamma$. In
+that ${(\tfid{A})}_\ast\jdeq \catid{\cat{F}/\ctxext{\Gamma}{A}}$ for any $A\in\cat{F}/\Gamma$. In
 particular, it follows that $g\circ\tfid{A}\jdeq g$ for any $g\in\thom{A}{B}$
@@
 $\mathbb{C}_I(\mathbb{E},\Gamma)$ of internal morphisms of $\mathbb{E}$,
-consisting of the same objects as $\cat{C}/\Gamma$ and the hom-sets 
+consisting of the same objects as $\cat{F}/\Gamma$ and the hom-sets 
 $\thom{A}{B}$. 
@@
 
-The composition $\ctxext{A}{P}$ of $A\in\cat{C}/\Gamma$ and $P\in\cat{C}/\ctxext{\Gamma}{A}$
+The composition $\ctxext{A}{P}$ of $A\in\cat{F}/\Gamma$ and $P\in\cat{F}/\ctxext{\Gamma}{A}$
 behaves like a strict $\Sigma$-type. We may define the pairing term
@@
 \begin{defn}
-Let $x\in T(A)$ and $u\in T(S_x(P))$ for $A\in\cat{C}/\Gamma$ and $P\in\cat{C}/\ctxext{\Gamma}{A}$. 
+Let $x\in T(A)$ and $u\in T(S_x(P))$ for $A\in\cat{F}/\Gamma$ and $P\in\cat{F}/\ctxext{\Gamma}{A}$. 
 We define the \emph{term extension of $x$ and $u$} to be
@@
 \begin{thm}\label{subst_by_tmext}
-Let $x\in T(A)$ and $u\in T(S_x(P))$ for $A\in\cat{C}/\Gamma$ and $P\in\cat{C}/\ctxext{\Gamma}{A}$.
+Let $x\in T(A)$ and $u\in T(S_x(P))$ for $A\in\cat{F}/\Gamma$ and $P\in\cat{F}/\ctxext{\Gamma}{A}$.
 Then we have
 \begin{equation*}
-S_{\tmext{x}{u}}\jdeq S_u\circ (S_x/P):\cat{C}/\ctxext{\Gamma}{{A}{P}}\to\cat{C}/\Gamma
+S_{\tmext{x}{u}}\jdeq S_u\circ (S_x/P):\cat{F}/\ctxext{\Gamma}{{A}{P}}\to\cat{F}/\Gamma
 \end{equation*}
@@
 \begin{defn}
-Let $A\in\cat{C}/\Gamma$ and $P\in\cat{C}/\ctxext{\Gamma}{A}$. We define
+Let $A\in\cat{F}/\Gamma$ and $P\in\cat{F}/\ctxext{\Gamma}{A}$. We define
 \begin{align*}
@@
 \begin{thm}\label{thm:famext_up}
-For any $A\in\cat{C}/\Gamma$ and $P\in\cat{C}/\ctxext{\Gamma}{A}$,
+For any $A\in\cat{F}/\Gamma$ and $P\in\cat{F}/\ctxext{\Gamma}{A}$,
 the map
@@
 
-\subsection{The interchange laws}
+\subsection{The interchange laws}\label{sec:interchange}
 
@@
 
+\subsection{E-systems are E'-systems}
+
+\begin{lem}
+Consider $A\in\cat{F}/\Gamma$ and $Q\in\cat{F}/\ctxext{\Gamma}{B}$. Then there
+is an isomorphism
+\begin{equation*}
+\{h\in\thom{A}{\ctxext{B}{Q}}\mid \jcomp{A}{h}{\cprojfstf{B}{Q}}\jdeq f\}\cong T(\jcomp{A}{f}{Q}).
+\end{equation*}
+for any $f\in\thom{A}{B}$.
+\end{lem}
+
+\begin{proof}
+We have the isomorphism
+\begin{align*}
+\thom{A}{\ctxext{B}{Q}} 
+& \jdeq T(\ctxwk{A}{\ctxext{B}{Q}}) \\
+& \jdeq T(\ctxext{\ctxwk{A}{B}}{\ctxwk{A}{Q}}) \\
+& \cong \bigsqcup\nolimits_{(f\in T(\ctxwk{A}{B}))}T(\subst{f}{\ctxwk{A}{Q}})
+\tag{By \autoref{thm:famext_up}} \\
+& \jdeq \bigsqcup\nolimits_{(f\in\thom{A}{B})}T(\jcomp{A}{f}{Q}).
+\end{align*}
+Also, we find $\cprojfst{\ctxwk{A}{B}}{\ctxwk{A}{Q}}{h}\jdeq\subst{h}{\ctxwk{A}{\cprojfstf{B}{Q}}}\jdeq
+\jcomp{A}{h}{\cprojfstf{B}{Q}}$.
+\end{proof}
+
+\begin{thm}
+E-systems are the same things as E'-systems.
+\end{thm}
+
+\begin{proof}
+For any $f:A\to B$ in context $\Gamma$ and any $Q\in\cat{F}/B$, we define
+\begin{equation*}
+\pi_2(f,Q)\defeq \jvcomp{\jcomp{A}{f}{Q}}{f}{\tfid{\jcomp{A}{f}{Q}}}
+  : \ctxext{A}{\jcomp{A}{f}{Q}}\to\ctxext{B}{Q}.
+\end{equation*}
+Then we have the commuting diagram
+\begin{equation*}
+\begin{tikzcd}[column sep=huge]
+\ctxext{A}{\jcomp{A}{f}{Q}} \arrow[r,"{\pi_2(f,Q)}"] \arrow[d,swap,"{\cprojfstf{A}{\jcomp{A}{f}{Q}}}"] & \ctxext{B}{Q} \arrow[d,"{\cprojfstf{B}{Q}}"] \\
+A \arrow[r,"f"] & B
+\end{tikzcd}
+\end{equation*}
+in context $\Gamma$. To show that this is a pullback diagram, consider a
+morphism $g:X\to A$ in context $\Gamma$. Then we have the isomorphisms
+\begin{equation*}
+\{h\in\thom{X}{\ctxext{B}{Q}}\mid \jcomp{X}{h}{\cprojfstf{B}{Q}}\jdeq\jcomp{X}{g}{f}\}
+  \cong
+T((\jcomp{X}{g}{f})_\ast(Q))
+  \cong
+\{u\in\thom{X}{\ctxext{A}{\jcomp{A}{f}{Q}}}\mid\jcomp{X}{u}{\cprojfstf{A}{\jcomp{A}{f}{Q}}}\jdeq g\}
+\end{equation*}
+Thus, we find for every $h:X\to\ctxext{B}{Q}$ satisfying
+$\jcomp{X}{h}{\cprojfstf{B}{Q}}\jdeq\jcomp{X}{g}{f}$, a unique morphism
+$u:X\to\ctxext{A}{\jcomp{A}{f}{Q}}$ satisfying
+$\jcomp{X}{u}{\cprojfstf{A}{\jcomp{A}{f}{Q}}}\jdeq g$. It is easy to verify
+that $\jcomp{X}{u}{\pi_2(f,Q)}\jdeq h$, so the universal property of pullbacks
+holds. The functoriality conditions follow immediately from the interchange
+laws proven in \autoref{sec:interchange}.
+
+We have described functors from E'-systems to E-systems and vice versa. It is
+a straightforward check that these are mutual inverses.
+\end{proof}
+
+\begin{comment}
 \subsection{Categories internal to E-systems}
@@
 \begin{enumerate}
-\item A family $\cftalgc{\cftalg{A}}\in\cat{C}/\Gamma$, representing the type of
+\item A family $\cftalgc{\cftalg{A}}\in\cat{F}/\Gamma$, representing the type of
 objects of the category framework;
-\item A family $\cftalgf{\cftalg{A}}\in\cat{C}/\ctxext{\Gamma}{\cftalgc{\cftalg{A}}}$,
+\item A family $\cftalgf{\cftalg{A}}\in\cat{F}/\ctxext{\Gamma}{\cftalgc{\cftalg{A}}}$,
 representing the type of morphisms of the category framework, where a term of
@@
 \begin{defn}
-Let $f_0\in\thom{A}{B}$ be an internal morphism and let $P\in\cat{C}/\ctxext{\Gamma}{A}$ and
-$Q\in\cat{C}/\ctxext{\Gamma}{B}$. We define
+Let $f_0\in\thom{A}{B}$ be an internal morphism and let $P\in\cat{F}/\ctxext{\Gamma}{A}$ and
+$Q\in\cat{F}/\ctxext{\Gamma}{B}$. We define
 \begin{equation*}
@@
 \begin{proof}
-The unit $\eta:1_{\cat{C}/\ctxext{\Gamma}{A}}\Rightarrow W_A\circ\Pi_A$ is given by
+The unit $\eta:1_{\cat{F}/\ctxext{\Gamma}{A}}\Rightarrow W_A\circ\Pi_A$ is given by
 \begin{equation*}
@@
 \begin{lem}
-Suppose $\cat{D}$ is a retract of a pre-category $\cat{C}$, i.e. there are functors
-$S:\cat{D}\to\cat{C}$ and $R:\cat{C}\to\cat{D}$ with a natural isomorphism
-$\tau:R\circ S \simeq \catid{\cat{D}}$, and suppose that $\cat{C}$ is a weakening
+Suppose $\cat{D}$ is a retract of a pre-category $\cat{F}$, i.e. there are functors
+$S:\cat{D}\to\cat{F}$ and $R:\cat{F}\to\cat{D}$ with a natural isomorphism
+$\tau:R\circ S \simeq \catid{\cat{D}}$, and suppose that $\cat{F}$ is a weakening
 category. Then $\cat{D}$ is a weakening system too.
@@
   &
-\cat{C}/S(\Gamma)
+\cat{F}/S(\Gamma)
   \arrow[r,"W_{S(f)}"]
   &
-\cat{C}/S(\Delta)
+\cat{F}/S(\Delta)
   \arrow[r,"R/S(\Delta)"]
@@
   &
-\cat{C}/S(\Delta)
+\cat{F}/S(\Delta)
   \arrow[r,"R/S(\Delta)"]
@@
 \begin{rmk}
-Note that $W$ is a strict $2$-functor frm $\cat{C}$ to $\mathbb{Cat}$ of which
-the value at an object $\Gamma$ is specified to be $\cat{C}/\Gamma$, and which
+Note that $W$ is a strict $2$-functor frm $\cat{F}$ to $\mathbb{Cat}$ of which
+the value at an object $\Gamma$ is specified to be $\cat{F}/\Gamma$, and which
 is required to preserve the final object. We could have considered $W$ to be a 
-pseudo-functor from $\cat{C}$ to $\mathbb{Cat}$ instead. A coherence theorem
+pseudo-functor from $\cat{F}$ to $\mathbb{Cat}$ instead. A coherence theorem
 by Power \cite{Power89} then allows for the strictification of this pseudo-functor
-into a (strict) weakening structure. (Does this map $\Gamma$ to $\cat{C}/\Gamma$
+into a (strict) weakening structure. (Does this map $\Gamma$ to $\cat{F}/\Gamma$
 and final to final?). Thus, we know what a weak pre-weakening system has to be.
@@
 $B:\mathbf{Esys_s}\to\mathbf{B0sys}$. It is immediate that we get isomorphisms
-$U(B(\mathbb{E}))\cong\cat{C}_{\mathbb{E}}$, natural in $\mathbb{E}$. Also, we
+$U(B(\mathbb{E}))\cong\cat{F}_{\mathbb{E}}$, natural in $\mathbb{E}$. Also, we
 get natural isomorphisms $B(\mathbb{E}/\Gamma)\cong B(\mathbb{E})/\Gamma$.
@@
 required to be functorial, i.e.~$W_{\ctxext{A}{P}}\jdeq W_P\circ W_A$ and
-$W_{\catid{\Gamma}}\jdeq\catid{\cat{C}_{\mathbb{E}}/\Gamma}$, it follows
+$W_{\catid{\Gamma}}\jdeq\catid{\cat{F}_{\mathbb{E}}/\Gamma}$, it follows
 immediately that the weakening structure on $U(B(\mathbb{E}))$ is isomorphic to
@@
 
-\subsection{From D-systems to locally stratified E'-systems}
-
-We use `D-system' to refer to Dybjer's categories with families. We recall
-Dybjer's definition. Note that $\mathbf{Fam}$ is the category of families of
-sets, of which the objects consist of a set $A$ and a family $\{B(x)\}_{x\in A}$
-over $A$. Although a terminal object is assumed to exist, not every object 
-of a D-system is `fibrant', i.e.~it is
-not the case that every $X\to 1$ is a composition of projections.
-We could only hope that there is a functor from D-sytems to E-systems without
-terminal objects. We write $\mathbf{Esys}_{\circ}$ for the category of E-systems
-without terminal objects.
-
-\begin{defn}
-A category is said to be locally stratified if each of its slice categories
-are stratified.
-\end{defn}
-
-\begin{defn}
-A \emph{setup for categories with families} consists of
+\subsection{From categories with families and conceptual categories to E-systems}
+
+In this section we recall Dybjer's notion of categories with families and we
+introduce the equivalent notion of D-systems. After having proved that categories with families
+are the same things as D-systems, we will focus on constructing a functor from
+D-systems to E-systems without empty context. However, due to the existence of pathological examples
+of categories with families, this functor is not going to be faithful. We then
+introduce the notion of well-founded D-systems, and show that the
+functor from D-systems to E'-systems without empty context restricts to an
+embedding from the well-founded D-systems to E'-systems.
+
+\begin{defn}
+A \emph{category with families} consists of
 \begin{enumerate}
@@
 $f_\ast(A)$ for $\mathrm{Ty}(f,A)$.
-\item A commuting triangle
-\begin{equation*}
-\begin{tikzcd}
-\int_\cat{C}\mathrm{Ty} \arrow[dr,swap,"\pi"] \arrow[rr,"p"] & & \cat{C}^{\to} \arrow[dl,"\mathrm{cod}"] \\
-& \cat{C}
-\end{tikzcd}
-\end{equation*} 
-We will write $p_A:\ctxext{\Gamma}{A}\to\Gamma$
-for the morphism $U(\Gamma,A)$ into $\Gamma$, for each object $\Gamma\in\cat{C}$ and
-$A\in\mathrm{Ty}(\Gamma)$.
-\end{enumerate}
-\end{defn}
-
-\begin{defn}
-A \emph{category with families} consists of a setup with
-\begin{enumerate}
-\item[(iv)] a preasheaf $\mathrm{Tm}:\op{(\int_\cat{C}\mathrm{Ty})}\to\mathbf{Set}$.
+\item For any $A\in\mathrm{Ty}(\Gamma)$, a morphism
+\begin{equation*}
+p_A : \ctxext{\Gamma}{A}\twoheadrightarrow \Gamma
+\end{equation*}
+in $\cat{C}$.
+\item a preasheaf $\mathrm{Tm}:\op{(\int_\cat{C}\mathrm{Ty})}\to\mathbf{Set}$.
 The action of a morphism $f:(\Delta,f_\ast(A))\to(\Gamma,A)$ of $\int_\cat{C}\mathrm{Ty}$
 on a term $x\in \mathrm{Tm}(A)$ is also written by $f_\ast(x)$.
-\item[(v)] a term $q_A\in\mathrm{Tm}(\ctxext{\Gamma}{A},(p_A)_\ast(A))$ for every
+\item a term $\delta_A\in\mathrm{Tm}(\ctxext{\Gamma}{A},(p_A)_\ast(A))$ for every
 $A\in\mathrm{Ty}(\Gamma)$, 
 \end{enumerate}
-such that for every $f:\Delta\to\Gamma$ in $\cat{C}$ and 
-$x\in \mathrm{Tm}(\Delta,f_\ast(A))$ there is a unique morphism 
-$\tilde{x}:\Delta\to \ctxext{\Gamma}{A}$ such that $p_A\circ\tilde{x}\jdeq f$ and 
-$\tilde{x}_\ast(q_A)\jdeq x$.
-\end{defn}
-
-\begin{lem}\label{lem:cwf_to_dsys}
-Let $\mathbb{D}$ be a category with families. Then there are bijections
-\begin{equation*}
-\zeta_{A,f}:\{\theta:\Delta\to\ctxext{\Gamma}{A}\mid p_A\circ\theta\jdeq f\}\cong \mathrm{Tm}(\Delta,f_\ast(A))
-\end{equation*}
-coherent in $f$ in the sense that
-\begin{align*}
-\zeta_{A,\catid{\Gamma}}(\theta) & \jdeq \theta \\
-\zeta_{A,f\circ g}(\theta) & \jdeq \zeta_{f_\ast(A),g}(\zeta_{A,f}(\theta)\circ g)
-\end{align*}
-for any $g:X\to\Delta$. 
+with the universal property that for every $f:\Delta\to\Gamma$ in $\cat{C}$, 
+the function
+\begin{equation*}
+\theta\mapsto\theta_\ast(\delta_A)
+  : \{\theta:\Delta\to\ctxext{\Gamma}{A}\mid p_A\circ\theta\jdeq f\}\to
+    \mathrm{Tm}(f_\ast(A))
+\end{equation*}
+is a bijection. We denote the inverse of this map by $\mu_{A,f}$. 
+\end{defn}
+
+There are of course many more sensible examples of categories with families,
+but the following class of examples causes us to doubt whether it is possible
+at all to have a faithful functor from the category of categories
+with families to the category of E-systems.
+
+\begin{eg}
+Every category $\cat{C}$ with a terminal object and the empty presheaf 
+$\mathrm{Ty}$, is a category with families.
+\end{eg}
+
+\begin{rmk}
+Although a terminal object is usually assumed to exist, not every object 
+of a D-system is assumed to be `fibrant', i.e.~it is
+not the case that every $X\to 1$ is a composition of projections $p_A$.
+We could only hope that there is a functor from D-sytems to E-systems without
+terminal objects.
+\end{rmk}
+
+There is essentially only one choice for the presheaf $\mathrm{Tm}$ and the
+term $\delta_A$, so the extra structure on a setup for categories with families
+is determined essentially uniquely once it exists. In other words, being a
+category with families is a property on setups for categories with families.
+
+\begin{defn}\label{lem:cwf_to_dsys}
+Let $\mathbb{D}$ be a category with families. Then there is a natural isomorphism
+\begin{equation*}
+\varphi_A:\mathrm{Tm}(A)\cong\{\theta:\Gamma\to\ctxext{\Gamma}{A}\mid p_A\circ\theta\jdeq\catid{\Gamma}\}
+\end{equation*}
+in the presheaf category $\mathrm{Psh}(\int_{\cat{C}}\mathrm{Ty})$. 
+\begin{comment}
+The
+natural isomorphism $\varphi$ is such that the term $\delta_A\in\mathrm{Tm}((p_A)_\ast(A))$
+corresponds to the unique arrow $\tilde{q}_A$ which fits in the diagram
+\begin{equation*}
+\begin{tikzcd}
+\ctxext{\Gamma}{A} \arrow[ddr,bend right=15,equals] \arrow[drr,bend left=15,equals] \arrow[dr,"{\tilde \delta_A}"] \\
+& \ctxext{{\Gamma}{A}}{(p_A)_\ast(A)} \arrow[d,fib] & \ctxext{\Gamma}{A} \arrow[d,fib] \\
+& \ctxext{\Gamma}{A} \arrow[r,"{p_A}"] & \Gamma
+\end{tikzcd}
+\end{equation*}
+\end{comment}
+\end{defn}
+
+\begin{constr}
+We define the inverse of $\varphi_A$ by $\theta\mapsto\theta_\ast(\delta_A)$. Verifying
+that these form a natural isomorphism is straightforward.
+\end{constr}
+
+\begin{defn}
+A \emph{D-system} consists of
+\begin{enumerate}
+\item A category $\cat{C}$ with a terminal object $1$.
+\item A presheaf $\mathrm{Ty}:\op{\cat{C}}\to\mathbf{Set}$. 
+\item A commuting triangle
+\begin{equation*}
+\begin{tikzcd}
+\int_{\cat{C}}\mathrm{Ty} \arrow[rr,"p"] \arrow[dr,"\pi"] & & \cat{C}^\square \arrow[dl,"\mathrm{cod}"] \\
+& \cat{C}
+\end{tikzcd}
+\end{equation*}
+of functors, where $\cat{C}^\square$ is the sub-pre-category of $\cat{C}^{\to}$ of
+which the morphisms are pullback squares. For any $A\in\mathrm{Ty}(\Gamma)$,
+the morphism $p(\Gamma,A)$ will be denoted by $p_A$; for any
+$(f,A):(\Delta,f_\ast(A))\to(\Gamma,A)$ in $\int_{\cat{C}}\mathrm{Ty}$, the morphism
+$p(f,A)$ will be denoted by $\pi_2(f,A)$.
+\end{enumerate}
+\end{defn}
+
+\begin{rmk}
+For any $f:\Delta\to\Gamma$, $g:X\to\Delta$ and $A\in\mathrm{Ty}(\Gamma)$ in a
+D-system $\mathbb{D}$, one has 
+$\pi_2(\catid{\Gamma},A)\jdeq\catid{\ctxext{\Gamma}{A}}$ and
+$\pi_2(f\circ g,A)\jdeq \pi_2(f,A)\circ\pi_2(g,f_\ast(A))$.
+\end{rmk}
+
+\begin{defn}
+A \emph{D-homomorphism from $\mathbb{D}\to\mathbb{D}'$} consists of a functor 
+$\mathcal{F}:\cat{C}\to\cat{C}'$ and a natural transformation
+$\mathrm{Ty}\Rightarrow\mathrm{Ty}'\circ\mathcal{F}$, such that 
+$\mathcal{F}(p_A)\jdeq p_{\mathcal{F}(A)}$ for any
+$A\in\mathrm{Ty}(\Gamma)$.
+\end{defn}
+
+\begin{lem}
+In the definition of D-system, the assumed structure beyond a choice of
+projections $p_A:\ctxext{\Gamma}{A}\twoheadrightarrow\Gamma$ is property-like.
 \end{lem}
 
-\begin{proof}
-Let $\theta:\Delta\to\ctxext{\Gamma}{A}$ be such that $p_A\circ\theta\jdeq f$. 
-Then $\theta_\ast(q_A)$ is an element
-of $\mathrm{Tm}(\theta_\ast((p_A)_\ast(A)))$. By the functoriality of $\mathrm{Ty}$, we have
-$\theta_\ast((p_A)_\ast(A))\jdeq (p_A\circ\theta)_\ast(A)\jdeq f_\ast(A)$. 
-So we obtain $\theta_\ast(q_A)\in\mathrm{Tm}(\Delta,f_\ast(A))$. Now
-the uniqueness of $\theta$ implies that the map $\theta\mapsto \theta_\ast(q_A)$
-is a bijection.
-
-It remains to verify the coherence identities. Let $\theta : \Gamma\to\ctxext{\Gamma}{A}$
-be such that $p_A\circ\theta\jdeq\catid{\Gamma}$. We wish to show that
-$\theta_\ast(q_A)\jdeq\theta$.  
-\begin{align*}
-\theta_\ast(q_A)
-& \jdeq \theta_\ast(\zeta_{A,p_A}(\catid{\ctxext{\Gamma}{A}})) \\
-& \jdeq \cdots \\
-& \jdeq \theta
-\end{align*}
-\end{proof}
-
-The previous lemma renders the presheaf $\mathrm{Tm}$ superfluous. Therefore,
-we define D-systems as follows.
-
-\begin{defn}
-A \emph{D-system} consists of a setup for categories with families
-such that for every $f:\Delta\to\Gamma$ and $A\in\mathrm{Ty}(\Gamma)$, there
-is a bijection
+\begin{thm}\label{thm:DD}
+A category with families is the same thing as a D-system.
+\end{thm}
+
+\begin{proof}
+Let $\mathbf{CWF}$ be a category with families. We show that $\mathbf{CWF}$ is
+a D-system in five steps.
+\begin{enumerate}
+\item We first show that
+\begingroup\it for every $f:\Delta\to\Gamma$ and $A\in\mathrm{Ty}(\Gamma)$, there is an
+isomorphism
 \begin{equation*}
@@
 such that $p_A\circ\theta\jdeq f$.
-\end{defn}
-
-\begin{thm}
-A category with families is the same thing as a D-system.
-\end{thm}
-
-\begin{proof}
-In \autoref{lem:cwf_to_dsys} we have shown that every category with families is
-a D-system. Thus, it suffices to show the converse.
-
-Let $\mathbb{D}$ be a D-system. For any object $\Gamma$ of $\mathbb{D}$ and any
-type $A\in\mathrm{Ty}(\Gamma)$, we define 
-\begin{equation*}
-\mathrm{Tm}(A)\defeq \{ x :\Gamma\to\ctxext{\Gamma}{A}\mid p_A\circ x\jdeq\catid{\Gamma}\}.
-\end{equation*}
-For any morphism $f:\Delta\to\Gamma$ and $A\in\mathrm{Ty}(\Gamma)$, we define
-the function $\mathrm{Tm}(f):\mathrm{Tm}(A)\to\mathrm{Tm}(f_\ast(A))$ by
-\begin{equation*}
-\mathrm{Tm}(f,x)\defeq \zeta_{A,f}(x\circ f).
-\end{equation*}
-We have to verify that this satisfies the functorial identities. For $f:\Delta\to\Gamma$
-and $g:X\to\Delta$ and any $x\in\mathrm{Tm}(A)$, we have
-\begin{align*}
-\mathrm{Tm}(f\circ g,x) & \jdeq \zeta_{A,f\circ g}(x\circ f\circ g) \\
-& \jdeq \zeta_{f_\ast(A),g}(\zeta_{A,f}(x\circ f)\circ g) \\
-& \jdeq \mathrm{Tm}(g,\mathrm{Tm}(f,x)).
-\end{align*}
-and
-\begin{equation*}
-\mathrm{Tm}(\catid{\Gamma},x) \jdeq \zeta_{A,\catid{\Gamma}}(x) \jdeq x.
-\end{equation*}
-This finishes the construction of the presheaf $\mathrm{Tm}$ on $\int_{\cat{C}}\mathrm{Ty}$.
-
-Next, we have to find $q_A\in\mathrm{Tm}((p_A)_\ast(A))$. Note that we have the
-isomorphism
-\begin{align*}
-\mathrm{Tm}((p_A)_\ast(A))
-& \jdeq \{\eta :\ctxext{\Gamma}{A}\to\ctxext{{\Gamma}{A}}{(p_A)_\ast(A)}\mid p_{(p_A)_\ast(A)}\circ\theta\jdeq\catid{\ctxext{\Gamma}{A}}\} \\
-& \cong \{\theta : \ctxext{\Gamma}{A}\to\ctxext{\Gamma}{A} \mid p_A\circ \theta\jdeq p_A\}
-\end{align*}
-The latter set contains the identity morphism $\catid{\ctxext{\Gamma}{A}}$. Therefore,
-we find
-\begin{equation*}
-q_A \defeq \zeta_{A,p_A}(\catid{\ctxext{\Gamma}{A}})\in\mathrm{Tm}((p_A)_\ast(A)).
-\end{equation*}
-To verify that we have constructed a category with families, let $f:\Delta\to\Gamma$
-and $x\in T(f_\ast(A))$. Note that we have the morphism 
-$\tilde{x}\defeq \zeta_{A,f}^{-1}(x):\Delta\to\ctxext{\Gamma}{A}$. 
-We need to check that $\tilde{x}$ is unique with the property that 
-$\mathrm{Tm}(\tilde{x},q_A)\jdeq x$. The coherence identities can be used to see
-that this is indeed the case:
-\begin{align*}
-\mathrm{Tm}(\tilde{x},q_A)
-& \jdeq \zeta_{(p_A)_\ast(A),\tilde{x}}(q_A\circ\tilde{x}) \\
-& \jdeq \zeta_{(p_A)_\ast(A),\tilde{x}}(\zeta_{A,p_A}(\catid{\ctxext{\Gamma}{A}})\circ\tilde{x}) \\
-& \jdeq \zeta_{A,p_A\circ\tilde{x}}(\tilde{x}) \\
-& \jdeq \zeta_{A,f}(\tilde{x}) \\
-& \jdeq x.
-\end{align*}
-For the uniqueness, let $\theta:\Delta\to\ctxext{\Gamma}{A}$ be any morphism such that
-$p_A\circ\theta\jdeq f$ and $\mathrm{Tm}(\theta,q_A)\jdeq x$. Then we have
-\begin{align*}
-x & \jdeq \mathrm{Tm}(\theta,q_A) \\
-& \jdeq \zeta_{(p_A)_\ast(A),\theta}(q_A\circ\theta) \\
-& \jdeq \zeta_{(p_A)_\ast(A),\theta}(\zeta_{A,p_A}(\catid{\ctxext{\Gamma}{A}})\circ\theta \\
-& \jdeq \zeta_{A,p_A\circ\theta}(\theta) \\
-& \jdeq \zeta_{A,f}(\theta).
-\end{align*}
-Therefore we see that $\tilde{x}\jdeq\zeta_{A,f}^{-1}(x)\jdeq\theta$.
-\end{proof}
-
-\begin{defn}
-An E'-system is said to be locally stratified, if the category $\cat{F}$ is
-locally stratified.
-\end{defn}
-
+
+Furthermore, we get $\zeta_{A,p_A}(\catid{\ctxext{\Gamma}{A}})\jdeq\varphi_{(p_A)_\ast(A)}(\delta_A)$%
+\endgroup
+\medskip
+
+\renewcommand\qedsymbol{\(\blacksquare\)}
+By \autoref{lem:cwf_to_dsys} we may define
+\begin{equation*}
+\zeta_{A,f}(\theta)\defeq \varphi_{f_\ast(A)}(\theta_\ast(\delta_A))
+\end{equation*}
+for any $\theta\in\{\theta:\Delta\to\ctxext{\Gamma}{A}\mid p_A\circ\theta\jdeq f\}$. 
+Since $\zeta_{A,f}$ is a composition of isomorphism, it is itself an isomorphism.
+Also, we get immediately that $\zeta_{A,\catid{\Gamma}}(\theta)\jdeq\theta$. To
+show the remaining identity:
+\begin{align*}
+\zeta_{A,f\circ g}(\theta\circ g)
+& \jdeq \varphi_{g_\ast(f_\ast(A))}((\theta\circ g)_\ast(\delta_A)) \\
+& \jdeq \varphi_{g_\ast(f_\ast(A))}(g_\ast(\theta_\ast(\delta_A))) \\
+& \jdeq \varphi_{g_\ast(f_\ast(A))}(g_\ast(\varphi_{f_\ast(A)}(\theta_\ast(\delta_A))_\ast(\delta_A))) \\
+& \jdeq \varphi_{g_\ast(f_\ast(A))}(g_\ast(\zeta_{A,f}(\theta)_\ast(\delta_A))) \\
+& \jdeq \varphi_{g_\ast(f_\ast(A))}((\zeta_{A,f}(\theta)\circ g)_\ast(\delta_A)) \\
+& \jdeq \zeta_{f_\ast(A),g}(\zeta_{A,f}(\theta)\circ g).
+\end{align*}
+This completes the proof that the isomorphisms $\zeta_{A,f}$ satisfy the required
+equalities.
+
+\item
+For any $A\in\mathrm{Ty}(\Gamma)$, we define
+$W_A\defeq A_\ast:\cat{F}/\Gamma\to\cat{F}/\ctxext{\Gamma}{A}$. Then we get
+for any $A,B\in\cat{F}/\Gamma$ an isomorphism
+\begin{equation*}
+\mathrm{hom}_{\cat{C}/\Gamma}(A,B)
+  \cong
+T(W_A(B)).
+\end{equation*}
+In particular, we get a term $\idtm{A}\in T((p_A)_\ast(A))$. 
+\medskip
+
+Let $A,B\in\cat{F}/\Gamma$. Then we have the isomorphism
+\begin{align*}
+T(W_A(B)) & \jdeq \{\eta:\ctxext{\Gamma}{A}\to\ctxext{{\Gamma}{A}}{W_A(B)}\mid W_A(B)\circ\eta\jdeq\catid{\ctxext{\Gamma}{A}}\} \\
+  & \cong
+\{\theta:\ctxext{\Gamma}{A}\to\ctxext{\Gamma}{B}\mid B\circ\theta\jdeq A\},
+\end{align*}
+so we find $\idtm{A}\defeq\zeta_{A,A}(\catid{\ctxext{\Gamma}{A}})\in T(W_A(A))$.
+
+\item Now we show that 
+\begingroup\it for any $A\in\mathrm{Ty}(\Gamma)$ and $f:\Delta\to\Gamma$ 
+there is a morphism $\pi_2(f,A):\ctxext{\Delta}{f_\ast(A)}\to\ctxext{\Gamma}{A}$
+such that $p_A\circ \pi_2(f,A)\jdeq f\circ p_{f_\ast(A)}$.
+\endgroup
+\medskip
+
+Note that we have the isomorphism
+\begin{align*}
+& \{ \theta : \ctxext{\Delta}{f_\ast(A)}\to\ctxext{\Gamma}{A}\mid p_A\circ\theta\jdeq f\circ p_{f_\ast(A)} \} \\
+  & \cong
+\{ \eta : \ctxext{\Delta}{f_\ast(A)}\to\ctxext{{\Delta}{f_\ast(A)}}{(f\circ p_{f_\ast(A)})_\ast(A)}
+  \mid (f\circ p_{f_\ast(A)})_\ast(A)\circ\eta\jdeq\catid{\ctxext{\Delta}{f_\ast(A)}}\} \\
+  & \jdeq
+\{ \eta : \ctxext{\Delta}{f_\ast(A)}\to\ctxext{{\Delta}{f_\ast(A)}}{(p_{f_\ast(A)})_\ast(f_\ast(A))}
+  \mid (f\circ p_{f_\ast(A)})_\ast(A)\circ\eta\jdeq\catid{\ctxext{\Delta}{f_\ast(A)}}\} \\
+  & \jdeq
+T((p_{f_\ast(A)})_\ast(f_\ast(A))).
+\end{align*}
+Since we have $\delta_{f_\ast(A)}\in T((p_{f_\ast(A)})_\ast(f_\ast(A)))$, we obtain 
+\begin{equation*}
+\pi_2(f,A)\defeq \zeta_{A,f\circ p_{f_\ast(A)}}^{-1}(\delta_{f_\ast(A)}):\ctxext{\Delta}{f_\ast(A)}\to\ctxext{\Gamma}{A}
+\end{equation*}
+satisfying $p_A\circ\pi_2(f,A)\jdeq f\circ p_{f_\ast(A)}$.
+\item Next, we show that
+\begingroup\it
+for any $f:\Delta\to\Gamma$, $g:X\to\Delta$ and $A\in\mathrm{Ty}(\Gamma)$, one has
+\begin{equation*}
+\pi_2(f,A)\circ\pi_2(g,f_\ast(A))\jdeq\pi_2(f\circ g,A).
+\end{equation*}
+\endgroup
+\medskip
+
+It suffices to show that
+\begin{equation*}
+\zeta_{A,f\circ g\circ p_{g_\ast(f_\ast(A))}}(\pi_2(f,A)\circ\pi_2(g,f_\ast(A)))\jdeq\delta_{g_\ast(f_\ast(A))}.
+\end{equation*}
+Because $g\circ p_{g_\ast(f_\ast(A))}\jdeq p_{f_\ast(A)}\circ\pi_2(g,f_\ast(A))$, we get
+to use the naturality of $\zeta$:
+\begin{align*}
+& \zeta_{A,f\circ g\circ p_{g_\ast(f_\ast(A))}}(\pi_2(f,A)\circ\pi_2(g,f_\ast(A))) \\
+& \jdeq
+\zeta_{A,f\circ p_{f_\ast(A)}\circ\pi_2(g,f_\ast(A))}(\pi_2(f,A)\circ\pi_2(g,f_\ast(A))) \\
+& \jdeq
+\zeta_{(p_{f_\ast(A)})_\ast(f_\ast(A)),\pi_2(g,f_\ast(A))}(\zeta_{A,f\circ p_{f_\ast(A)}}(\pi_2(f,a))\circ\pi_2(g,f_\ast(A))) \\
+& \jdeq
+\zeta_{(p_{f_\ast(A)})_\ast(f_\ast(A)),\pi_2(g,f_\ast(A))}(\delta_{f_\ast(A)}\circ\pi_2(g,f_\ast(A))) \\
+& \jdeq
+\zeta_{(p_{f_\ast(A)})_\ast(f_\ast(A)),\pi_2(g,f_\ast(A))}(\zeta_{f_\ast(A),p_{f_\ast(A)}}(\catid{\ctxext{\Delta}{f_\ast(A)}})\circ\pi_2(g,f_\ast(A))) \\
+& \jdeq
+\zeta_{f_\ast(A),p_{f_\ast(A)}\circ\pi_2(g,f_\ast(A))}(\catid{\ctxext{\Delta}{f_\ast(A)}}\circ\pi_2(g,f_\ast(A))) \\
+& \jdeq
+\zeta_{f_\ast(A),p_{f_\ast(A)}\circ\pi_2(g,f_\ast(A))}(\pi_2(g,f_\ast(A))) \\
+& \jdeq
+\zeta_{f_\ast(A),g\circ p_{g_\ast(f_\ast(A))}}(\pi_2(g,f_\ast(A)))\\
+& \jdeq
+\delta_{g_\ast(f_\ast(A))}.
+\end{align*}
+\item
+\begingroup\it
+Let $f:\Delta\to\Gamma$ and $A\in\mathrm{Ty}(\Gamma)$. Then one has 
+\begin{equation*}
+\pi_2(f,A)\circ\zeta_{A,f}(\theta)\jdeq\theta
+\end{equation*}
+for any $\theta:\Delta\to\ctxext{\Gamma}{A}$ satisfying $A\circ\theta\jdeq f$.
+\endgroup
+\medskip
+
+We will write $\tilde\theta\defeq\zeta_{A,f}(\theta)$. Note that
+\begin{equation*}
+p_A\circ\pi_2(f,A)\circ\tilde\theta\jdeq f\circ p_{f_\ast(A)}\circ\tilde\theta\jdeq f.
+\end{equation*} 
+Therefore, we can prove that $\pi_2(f,A)\circ\tilde\theta\jdeq\theta$ by showing
+that $\zeta_{A,f}(\pi_2(f,A)\circ\tilde\theta)\jdeq\tilde\theta$. This is
+shown using the naturality of $\zeta$:
+\begin{align*}
+\zeta_{A,f}(\pi_2(f,A)\circ\tilde\theta)
+  & \jdeq
+\zeta_{A,f\circ p_{f_\ast(A)}\circ\tilde\theta}(\pi_2(f,A)\circ\tilde\theta)
+  \\
+  & \jdeq
+\zeta_{(p_{f_\ast(A)})_\ast(f_\ast(A)),\tilde{\theta}}(\zeta_{A,f\circ p_{f_\ast(A)}}(\pi_2(f,A))\circ\tilde\theta)
+  \\
+  & \jdeq
+\zeta_{(p_{f_\ast(A)})_\ast(f_\ast(A)),\tilde{\theta}}(\idtm{f_\ast(A)}\circ\tilde\theta)
+  \\
+  & \jdeq
+\zeta_{(p_{f_\ast(A)})_\ast(f_\ast(A)),\tilde{\theta}}(\zeta_{f_\ast(A),p_{f_\ast(A)}}(\catid{\ctxext{\Delta}{f_\ast(A)}})\circ\tilde\theta)
+  \\
+  & \jdeq
+\zeta_{f_\ast(A),p_{f_\ast(A)}\circ\tilde\theta}(\catid{\ctxext{\Delta}{f_\ast(A)}}\circ\tilde\theta) \\
+  & \jdeq
+\zeta_{f_\ast(A),\catid{\Delta}}(\tilde\theta) \\
+  & \jdeq 
+\theta.
+\end{align*}
+\item 
+\begingroup\it
+The diagram
+\begin{equation*}
+\begin{tikzcd}[column sep=large]
+\ctxext{\Delta}{f_\ast(A)} \arrow[d,fib] \arrow[r,"{\pi_2(f,A)}"] & \ctxext{\Gamma}{A} \arrow[d,fib] \\
+\Delta \arrow[r,"f"] & \Gamma
+\end{tikzcd}
+\end{equation*}
+is a pullback square.
+\endgroup
+\medskip
+
+To verify the universal mapping property of the pullback square, consider a
+commuting diagram
+\begin{equation*}
+\begin{tikzcd}
+X \arrow[d,swap,"g"] \arrow[r,"h"] & \ctxext{\Gamma}{A} \arrow[d,fib] \\
+\Delta \arrow[r,"f"] & \Gamma
+\end{tikzcd}
+\end{equation*}
+Then $h$ corresponds uniquely to a section $k$ of $(f\circ g)_\ast(A)$. Since
+$(f\circ g)_\ast(A)\jdeq g_\ast(f_\ast(A))$, the section $k$ fits in the
+commuting diagram
+\begin{equation*}
+\begin{tikzcd}[column sep=huge]
+\ctxext{X}{g_\ast(f_\ast(A))} \arrow[d,xshift=.7ex,fib] \arrow[r,"{\pi_2(g,f_\ast(A))}"]
+& \ctxext{\Delta}{f_\ast(A)} \arrow[d,fib] \arrow[r,"{\pi_2(f,A)}"]
+& \ctxext{\Gamma}{A} \arrow[d,fib] \\
+X \arrow[r,swap,"g"] \arrow[u,xshift=-.7ex,"k"] & \Delta \arrow[r,swap,"f"] & \Gamma
+\end{tikzcd}
+\end{equation*}
+so we obtain $u\defeq\pi_2(g,f_\ast(A))\circ k:X\to\ctxext{\Delta}{f_\ast(A)}$. 
+Then we have $p_{f_\ast(A)}\circ u\jdeq g\circ p_{g_\ast(f_\ast(A))}\circ k\jdeq g$ and
+$\pi_2(f,A)\circ u\jdeq \pi_2(f\circ g,A)\circ k\jdeq h$. Note that
+$u$ corresponds uniquely to $k$, which corresponds uniquely to $h$, so the
+uniqueness of $u$ is automatic.
+\end{enumerate}
+Now let $\mathbb{D}$ be a D-system. We define the presheaf $\mathrm{Tm}:
+\op{(\int_{\cat{C}}\mathrm{Ty})}\to\mathbf{Set}$ by
+\begin{equation*}
+\mathrm{Tm}(A) \defeq \{x:\Gamma\to\ctxext{\Gamma}{A}\mid p_A\circ x\jdeq \catid{\Gamma}\}.
+\end{equation*}
+For $f:\Delta\to\Gamma$ and $A\in\mathrm{Ty}(\Gamma)$, we get a function
+$x\mapsto f_\ast(x):\mathrm{Tm}(A)\to\mathrm{Tm}(f_\ast(A))$ by defining $f_\ast(x)$
+to be the unique morphism such that the diagram
+\begin{equation*}
+\begin{tikzcd}[column sep=huge] 
+\Delta \arrow[ddr,bend right=15,equals] \arrow[drr,bend left=15,"x\circ f"] \arrow[dr,densely dotted,near end,"{f_\ast(x)}"] \\
+& \ctxext{\Delta}{f_\ast(A)} \arrow[d,fib] \arrow[r,swap,"{\pi_2(f,A)}"] & \ctxext{\Gamma}{A} \arrow[d,fib] \\
+& \Delta \arrow[r,swap,"f"] & \Gamma
+\end{tikzcd}
+\end{equation*}
+commutes. A combination of the assumed uniqueness and the assumed equalities on
+$\pi_2(\blank,\blank)$ gives the functoriality of $\mathrm{Tm}$.
+
+The term $\delta_A\in\mathrm{Tm}((p_A)_\ast(A))$ is defined to be the unique morphism
+such that the diagram
+\begin{equation*}
+\begin{tikzcd}[column sep=large]
+\ctxext{\Gamma}{A} \arrow[drr,bend left=15,equals] \arrow[ddr,bend right=15,equals] \arrow[dr,densely dotted,near end,"{\delta_{A}}"] \\
+& \ctxext{{\Gamma}{A}}{(p_A)_\ast(A)} \arrow[r,swap,"{\pi_2(A,A)}"] \arrow[d,fib] & \ctxext{\Gamma}{A} \arrow[d,fib] \\
+& \ctxext{\Gamma}{A} \arrow[r,swap,"A"] & \Gamma
+\end{tikzcd}
+\end{equation*}
+commutes.
+
+The universality assumption of categories with families is a consequence of
+the universal property of pullbacks: the map $x\mapsto\pi_2(f,X)\circ x$
+from $\{x:\Delta\to\ctxext{\Delta}{f_\ast(A)}\mid p_{f_\ast(A)}\circ x\jdeq\catid{\Delta}\}$
+to $\{\theta:\Delta\to\ctxext{\Gamma}{A}\mid p_A\circ\theta\jdeq f\}$ is a bijection.
+\end{proof}
+
+\begin{defn}
+Let $\mathbb{D}$ be a D-system with underlying category $\cat{C}$. We define
+$\cat{F}\subseteq\cat{C}$ to be the subcategory with the same objects as
+$\cat{C}$, where the morphisms are generated by the projections $p_A$. Note that
+there is no reason why $\cat{F}$ should possess a terminal object.
+\end{defn}
+
+To be able to even construct an E'-system out of a D-system, we need to impose
+a rather restricting conditon on the D-system.
+
+\begin{defn}
+Let $\mathbb{D}$ be a D-system. Then we define for any $\Gamma\in\mathbb{D}$ 
+the category $\Lambda(\Gamma)$ of which the objects are finite sequences
+of the form $(A_1,\ldots,A_n)$ where
+\begin{equation*}
+A_1\in\mathrm{Ty}(\Gamma),\ A_2\in\mathrm{Ty}(\ctxext{\Gamma}{A_1}),\ \ldots,\ 
+A_n\in\mathrm{Ty}(\ctxext{{{\Gamma}{A_1}}{\cdots}}{A_{n-1}})
+\end{equation*}
+where the morphisms are generated by $(A_1,\ldots,A_{n+1})\leq (A_1,\ldots,A_{n})$.
+\end{defn}
+
+\begin{defn}
+Let $\mathbb{D}$ be a D-system and let $f:\Delta\to\Gamma$ in $\mathbb{D}$.
+Then there is a functor $\Lambda(f):\Lambda(\Gamma)\to\Lambda(\Delta)$, with
+morphisms $\pi_2(f,A):\ctxext{\Delta}{p_{\Lambda(f,A)}}\to
+\ctxext{\Gamma}{p_A}$ in $\mathbb{D}$, such that the square
+\begin{equation*}
+\begin{tikzcd}[column sep=large]
+\ctxext{\Delta}{\Lambda(f,A)} \arrow[r,"{\pi_2(f,A)}"] \arrow[d,fib] & \ctxext{\Gamma}{A} \arrow[d,fib] \\
+\Delta \arrow[r,swap,"f"] & \Gamma
+\end{tikzcd}
+\end{equation*}
+is a pullback square.
+\end{defn}
+
+\begin{constr}
+We construct the asserted structure by
+induction on the length of the arrow $A$ into $\Gamma$. We define
+$\Lambda(f,\emptyf[\Gamma])\defeq \emptyf[\Delta]$, and $\pi_2(f,\emptyf[\Gamma])\defeq f:\ctxext{\Delta}{\catid{\Delta}}\to\ctxext{\Gamma}{\catid{\Gamma}}$. The resulting square is trivially
+a pullback square. Now suppose we have a pullback square as above.
+and let $B\in\mathrm{Ty}(\ctxext{\Gamma}{A})$. Then we define
+\begin{equation*}
+\Lambda(f,[A,B])\defeq [\Lambda(f,A),\pi_2(f,A)_\ast(B)].
+\end{equation*}
+By \autoref{thm:DD}, the top square in the diagram
+\begin{equation*}
+\begin{tikzcd}[column sep=6em]
+\ctxext{{\Delta}{\Lambda(f,A)}}{(\pi_2(f,A)_\ast(B))} \arrow[d,fib] \arrow[r,"{\pi_2(\pi_2(f,A),B)}"] & \ctxext{{\Gamma}{A}}{B} \arrow[d,fib] \\
+\ctxext{\Delta}{\Lambda(f,A)} \arrow[d,fib] \arrow[r,"{\pi_2(f,A)}"] & \ctxext{\Gamma}{A} \arrow[d,fib] \\
+\Delta \arrow[r,"f"] & \Gamma
+\end{tikzcd}
+\end{equation*}
+is a pullback square.
+Since the bottom square is assumed to be a pullback, we get that the outer rectangle
+is a pullback, as desired. 
+\end{constr}
+
+\begin{rmk}
+The construction above does not help us define a functor $f_\ast:\cat{F}/\Gamma
+\to\cat{F}/\Delta$ without further assumptions on the D-system $\mathbb{D}$.
+The functor $f_\ast$ would have the property that the square
+\begin{equation*}
+\begin{tikzcd}
+\Lambda(\Gamma) \arrow[d,swap,"p"] \arrow[r,"\Lambda(f)"] & \Lambda(\Delta) \arrow[d,"p"] \\
+\cat{F}/\Gamma \arrow[r,densely dotted,"{f_\ast}"] & \cat{F}/\Delta
+\end{tikzcd}
+\end{equation*}
+commutes. The above square can only commute, if one has
+\begin{equation*}
+\Lambda(f)/A\jdeq\Lambda(f)/A'
+\end{equation*} 
+for any $A,A'\in\Lambda(\Gamma)$ with $p_A\jdeq p_{A'}$. This would entail that
+$\pi_2(f,A)\jdeq\pi_2(f,A')$, a property that seems to not hold generally.
+\end{rmk}
+
+\begin{defn}
+A D-system is said to be coherent if for any $f:\Delta\to\Gamma$ and any
+$A,A'\in\Lambda(\Gamma)$ satisfying $p_A\jdeq p_{A'}$, one has
+\begin{equation*}
+\pi_2(f,A)\jdeq \pi_2(f,A').
+\end{equation*}
+\end{defn}
+
+\begin{defn}
+We say that a D-system $\cat{D}$ is \emph{locally well-founded} if the obvious functor
+$p:\Lambda(\mathbb{D})\to\cat{F}/\Gamma$ is an isomorphism of pre-categories.
+\end{defn}
+
+\begin{rmk}
+In other words, a D-system is locally well-founded if and only if
+for any $A\in\cat{F}$ there is a unique sequence
+$A_1\in\mathrm{Ty}(\Gamma)$, $A_2\in\mathrm{Ty}(\ctxext{\Gamma}{A_1})$, \ldots,
+$A_n\in\mathrm{Ty}(\ctxext{{{\Gamma}{A_1}}{\cdots}}{A_{n-1}})$ such that
+$A\jdeq p_{A_1}\circ\cdots\circ p_{A_n}$. Locally well-founded D-systems 
+are examples of coherent D-systems.
+\end{rmk}
+
+\begin{defn}
+For any locally well-founded D-system $\mathbb{D}$ there is 
+an E'-system $U(\mathbb{D})$ without terminal object in 
+$\cat{F}$. Thus, we get a functor $\mathbf{Dsys}\to\mathbf{E'sys}_{\circ}$
+\end{defn}
+
+\begin{constr}
+Let $f:\Delta\to\Gamma$. By the assumption that $\mathbb{D}$ is locally well-founded,
+
+To verify that this gives indeed a functor from $\cat{F}/\Gamma$ to
+$\cat{F}/\Delta$, we need to show that for any two objects or $A,A'\in\Lambda(\Gamma)$
+which get mapped to the same object of $\cat{F}/\Gamma$, get mapped to the
+same object of $\cat{F}/\Delta$. Let $A,A'\in\Lambda(\Gamma)$ be such that
+$p_A\jdeq p_{A'}$. It suffices to show that $\pi_2(f,A)\jdeq\pi_2(f,A')$
+
+The inductive proofs to verify properties
+\autoref{defn:E'sys:b} and \autoref{defn:E'sys:c} of \autoref{defn:E'sys} are
+completely straightforward, so $U(\mathbb{D})$ is an E'-system without terminal
+object in $\cat{F}$.
 \begin{comment}
-\begin{rmk}
-In the following definition, one could take the underlying category of the
-E-system $\mathcal{E}(\mathbb{D})$ associated to a D-system $\mathbb{D}$ to
-be the subcategory generated by the projections $p_A$, and construct the
-rest of the structure of E-systems on top of that. However, it is not so easy
-to characterize the image of the functor $\mathcal{E}$ in that case. Intuitively,
-D-systems should correspond to locally stratified E-systems. To be able to prove
-such a theorem, we've gone with a slightly modified definition of the underlying
-category.
-\end{rmk}
-\end{comment}
-
-\begin{defn}
-For any D-system $\mathbb{D}$ there is 
-a locally stratified E'-system $U(\mathbb{D})$ without terminal object in 
-$\cat{F}$. Thus, we get a functor $\mathbf{Dsys}\to\mathbf{E'sys}_{\circ}$
-\end{defn}
-
-\begin{constr}
-Let $\mathbb{D}$ be a D-system with underlying category $\cat{C}$. We define
-$\cat{F}\subseteq\cat{C}$ to be the
-\begin{comment}
-category consisting of 
-\begin{enumerate}
-\item equivalence classes of finite lists of the form $(\Gamma,A_1,\ldots A_n)$, where $\Gamma$ is
-an object of $\cat{C}$ and $A_{i+1}\in\mathrm{Ty}(\ctxext{{{\Gamma}{A_1}}{\cdots}}{A_i})$, where
-the equivalence relation is generated by
-\begin{equation*}
-(\ctxext{{{\Gamma}{A_1}}{\cdots}}{A_n},A_{n+1},\ldots,A_{n+m})\sim (\Gamma,A_1,\ldots,A_{n+m}).
-\end{equation*}
-\item
+Finally, we need to verify that the structure we've defined satisfies the properties
+of being an E'-system without terminal object. We go through them systematically,
+following the same order as in \autoref{defn:E'sys}:
+\begin{enumerate}[label=(\alph*)]
+\item This property holds by definition.
+\item Let $\Gamma\in\cat{C}$. We will prove by induction on $A\in\cat{F}/\Gamma$
+that $(\catid{\Gamma})_\ast(A)\jdeq A$ and that $\pi_2(\catid{\Gamma},A)\jdeq \catid{\ctxext{\Gamma}{A}}$.
+Note that these properties hold by definition for $A\jdeq\catid{\Gamma}$. For the
+inductive step, assume that $A\in\cat{F}/\Gamma$ is such that
+$(\catid{\Gamma})_\ast(A)\jdeq A$ and $\pi_2(\catid{\Gamma},A)\jdeq
+\catid{\ctxext{\Gamma}{A}}$ and let $B\in\mathrm{Ty}(\ctxext{\Gamma}{A})$. Then
+we have
+\begin{align*}
+(\catid{\Gamma})_\ast(A\circ p_B) & \jdeq (\catid{\Gamma})_\ast(A)\circ \pi_2(\catid{\Gamma},A)_\ast(p_B) \\
+& \jdeq A\circ p_B
+\end{align*}
+Also, we have
+\begin{align*}
+\pi_2(\catid{\Gamma},A\circ p_B) & \jdeq \pi_2(\pi_2(\catid{\Gamma},A),p_B) \\
+& \jdeq \pi_2(\catid{\ctxext{\Gamma}{A}},p_B) \\
+& \jdeq \catid{\ctxext{{\Gamma}{A}}{B}}
+\end{align*}
+\item Let $f:\Delta\to \Gamma$ and $g:X\to \Delta$. We will prove by induction
+on $A\in\cat{F}/\Gamma$ that $(f\circ g)_\ast(A)\jdeq g_\ast(f_\ast(A))$ and
+that $\pi_2(f\circ g,A)\jdeq \pi_2(f,A)\circ \pi_2(g,f_\ast(A))$. Note that
+these properties hold by definition for $A\jdeq\catid{\Gamma}$. For the inductive
+step, assume that $A\in\cat{F}/\Gamma$ is such that $(f\circ g)_\ast(A)\jdeq g_\ast(f_\ast(A))$ and
+$\pi_2(f\circ g,A)\jdeq \pi_2(f,A)\circ \pi_2(g,f_\ast(A))$, and let
+$B\in\mathrm{Ty}(\ctxext{\Gamma}{A})$. Then we have
+\begin{align*}
+(f\circ g)_\ast(A\circ p_B) & \jdeq \pi_2(f\circ g,A)_\ast(p_B) \\
+& \jdeq (\pi_2(f,A)\circ\pi_2(g,f_\ast(A)))_\ast(p_B) \\
+& \jdeq \pi_2(g,f_\ast)_\ast(\pi_2(f,A)_\ast(p_B)) \\
+& \jdeq g_\ast(f_\ast(A\circ p_B))
+\end{align*}
+
+\item This property holds by definition.
 \end{enumerate}
 \end{comment}
-subcategory with the same objects as
-$\cat{C}$, where the morphisms are generated by the projections $p_A$.
-
-Let $f:\Delta\to\Gamma$. We define $f_\ast:\cat{F}/\Gamma\to\cat{F}/\Delta$ by
-induction on the length of the arrow $A$ into $\Gamma$, simultaneously with a
-morphism $\pi_2(f,A)$ from $\ctxext{\Delta}{f_\ast(A)}$ into $\ctxext{\Gamma}{A}$
-for which the square
+\end{constr}
+
+\begin{thm}
+The functor $U:\mathbf{Dsys}\to\mathbf{E'sys_{\circ}}$ is full and faithful.
+\end{thm}
+
+\begin{proof}
+The functor $U$ is faithful, because for any morphism 
+$\mathcal{F}:\mathbb{D}\to\mathbb{D}'$ of locally well-founded D-systems, 
+$U(\mathcal{F})$ and $\mathcal{F}$ have the same underlying functor
+from $\cat{C}$ to $\cat{C}'$.
+
+It remains to show that $U$ is full. Let
+$\mathbb{D}$ and $\mathbb{D}$ be D-systems and let $\mathcal{G}:U(\mathbb{D})
+\to U(\mathbb{D}')$. Then $\mathcal{G}$ restricted to $\cat{F}$ lifts factorizations.
+We  first have to define a natural transformation
+$\mathrm{Ty}\Rightarrow \mathrm{Ty}'\circ\mathcal{G}$. Let
+$\Gamma\in\cat{C}$ and $A\in\mathrm{Ty}(\Gamma)$. Then we have
+$\mathcal{G}(p_A):\mathcal{G}(\ctxext{\Gamma}{A})\to \mathcal{G}(\Gamma)$.
+The category of factorizations of $p_A$ in $\cat{F}$ is a isomorphic to $\mathbf{2}$,
+so the category of factorizations of $\mathcal{G}(p_A)$ is also isomorphic to
+$\mathbf{2}$. This implies that $\mathcal{G}(p_A)$ is of the form $p_B$ for
+some $B\in\mathrm{Ty}'(\mathcal{G}(\Gamma))$. This $B$ is called $\mathcal{G}(A)$.
+
+To verify the naturality, let $f:\Delta\to\Gamma$. We need to show that the
+square
 \begin{equation*}
 \begin{tikzcd}
+\mathrm{Ty}(\Delta) \arrow[r,"{\mathcal{G}_\Delta}"] & \mathrm{Ty}'(\mathcal{G}(\Delta)) \\
+\mathrm{Ty}(\Gamma) \arrow[u,"\mathrm{Ty}(f)"] \arrow[r,"{\mathcal{G}_\Gamma}"] &
+\mathrm{Ty}'(\mathcal{G}(\Gamma)) \arrow[u,swap,"{\mathrm{Ty}'(\mathcal{G}(f))}"]
+\end{tikzcd}
+\end{equation*}
+commutes. Let $A\in\mathrm{Ty}(\Gamma)$. Then we need to show that
+$\mathcal{G}(f)_\ast(\mathcal{G}(A))\jdeq \mathcal{G}(f_\ast(A))$. By the uniqueness
+assumption of well-founded D-systems, it suffices to show that
+$p_{\mathcal{G}(f)_\ast(\mathcal{G}(A))}\jdeq p_{\mathcal{G}(f_\ast(A))}$. To
+see this, we do a simple calculation
+\begin{align*}
+p_{\mathcal{G}(f)_\ast(\mathcal{G}(A))}
+& \jdeq \mathcal{G}(f)_\ast(p_{\mathcal{G}(A)}) \\
+& \jdeq \mathcal{G}(f)_\ast(\mathcal{G}(p_A)) \\
+& \jdeq \mathcal{G}(f_\ast(p_A)) \\
+& \jdeq \mathcal{G}(p_{f_\ast(A)}) \\
+& \jdeq p_{\mathcal{G}(f_\ast(A))} \qedhere
+\end{align*}
+\end{proof}
+
+\begin{thm}
+The image of $U$ is the subcategory of locally stratified E'-systems without
+empty context. 
+\end{thm}
+
+\begin{proof}
+For any $\Gamma\in\mathbb{D}$, there is an obvious stratification functor
+$\Lambda(\Gamma)\to (\N,\geq)$, so $U(\mathbb{D})$ is a locally stratified
+E'-system.
+
+Now let $\mathbb{E}$ be a locally stratified E'-system without empty context.
+For $\Gamma\in\cat{C}$, we define $\mathrm{Ty}(\Gamma)$ to be the set of
+families of length one. Since any $f_\ast$ is stratified, it follows that
+$\mathrm{Ty}$ is a presheaf on $\cat{C}$. The choice of pullback squares
+\begin{equation*}
+\begin{tikzcd}[column sep=large]
 \ctxext{\Delta}{f_\ast(A)} \arrow[r,"{\pi_2(f,A)}"] \arrow[d,fib] & \ctxext{\Gamma}{A} \arrow[d,fib] \\
@@
 \end{equation*}
-is a pullback square. We define
-$f_\ast(\catid{\Gamma})\defeq\catid{\Delta}$, and $\pi_2(f,\catid{\Gamma})\defeq f:\ctxext{\Delta}{f_\ast(\catid{\Gamma})}\to\ctxext{\Gamma}{\catid{\Gamma}}$. The resulting square is trivially
-a pullback square. Now suppose we have a pullback square as above.
-and let $B\in\mathrm{Ty}(\ctxext{\Gamma}{A})$. Then we define
-\begin{equation*}
-f_\ast(A\circ p_B)\defeq \pi_2(f,A)_\ast(B).
-\end{equation*}
-By \autoref{lem:EE_pb}, we get the top pullback square in the diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=6em]
-\ctxext{{\Delta}{f_\ast(A)}}{f_\ast(B)} \arrow[d,fib] \arrow[r,"{\pi_2(\pi_2(f,A),B)}"] & \ctxext{{\Gamma}{A}}{B} \arrow[d,fib] \\
-\ctxext{\Delta}{f_\ast(A)} \arrow[d,fib] \arrow[r,"{\pi_2(f,A)}"] & \ctxext{\Gamma}{A} \arrow[d,fib] \\
-\Delta \arrow[r,"f"] & \Gamma
-\end{tikzcd}
-\end{equation*}
-Since the bottom square is assumed to be a pullback, we get that the outer rectangle
-is a pullback, as desired.
-
-Next, we define a weakening structure on $\cat{C}_p$. Since the morphisms of
-$\cat{C}_p$ are composites of projections, we will define
-\begin{equation*}
-W_{p_{A_n}\circ\cdots\circ p_{A_1}}\defeq W_{p_{A_n}}\circ\cdots W_{p_{A_1}}.
-\end{equation*}
-Therefore, we only have to define functors 
-$W_A:\cat{C}_p/\Gamma\to\cat{C}_p/\ctxext{\Gamma}{A}$ with term structure for every 
-$A\in\mathrm{Ty}(\Gamma)$. Since morphisms are composites of projections, we
-can define this functor by induction. Let $W_A(\catid{\Gamma})\defeq
-\catid{\ctxext{\Gamma}{A}}$ and $W_A(p_Q\circ f)\defeq$
-\end{constr}
-
-\begin{thm}
-There is a full and faithful functor from the category of D-systems to the
-category of E-systems.
-\end{thm}
-
+is functorial by assumption, so we have defined a category with families. It
+is straigtforward to verify that this category with families is locally well-founded,
+and that its construction is inverse to $U$. 
+\end{proof}
+
+\begin{comment}
 \subsection{The equivalence of E-systems and E'-systems}
@@
 \end{proof}
+\end{comment}
+
+\subsection{Comprehension categories}
+
+\begin{defn}
+Let $p:\cat{E}\to\cat{B}$ be a functor and let $g:A\to B$ be a morphism in $\cat{B}$.
+A morphism $f:X\to Y$ in $\cat{E}$ is said to be \emph{cartesian over $g$}, if
+\begin{enumerate}
+\item $p(f)\jdeq g$
+\item For all $f':X'\to Y$ with $p(f')\jdeq g$, there is a unique $u:X'\to X$
+such that $f'=f\circ u$ and $p(u)\jdeq\catid{A}$. 
+\end{enumerate}
+A cartesian morphism over $g$ is also called a \emph{cartesian lift of $g$}.
+\end{defn}
+
+\begin{defn}
+A functor $p:\cat{E}\to\cat{B}$ is called a \emph{Grothendieck fibration} if
+every morphism into an object of the form $p(X)$ has a cartesian lift, 
+and cartesian morphisms are closed under composition.
+\end{defn}
+
+\begin{defn}
+A Grothendieck fibration $p:\cat{E}\to\cat{B}$ is said to be \emph{split}, if
+there is a functorial choice of cartesian lifts, i.e.~for every $g:A\to B$ in $\cat{B}$
+there is a cartesian lift $j(g):j(A)\to j(B)$, such that $j(\catid{A})\jdeq\catid{j(A)}$
+and $j(h\circ g)\jdeq j(h)\circ j(g)$.
+\end{defn}
+
+\begin{defn}
+A comprehension category $\mathcal{P}$ consists of categories $\cat{E}$, $\cat{C}$ and a
+commuting triangle
+\begin{equation*}
+\begin{tikzcd}
+\cat{E} \arrow[rr,"\mathcal{P}"] \arrow[dr,swap,"p"] & & \cat{C}^{\to} \arrow[dl,"\mathrm{cod}"] \\
+& \cat{C}
+\end{tikzcd}
+\end{equation*}
+of functors, such that $p$ is a Grothendieck fibration, and $\mathcal{P}$ maps
+cartesian morphisms in $\cat{E}$ to pullback squares in $\cat{C}$. We say that
+$\mathcal{P}$ is a full comprehension category if the functor $\mathcal{P}:
+\cat{E}\to\cat{C}^{\to}$ is full and faithful; and we say that
+$\mathcal{P}$ is split if $p$ is split.
+\end{defn}
+
+\begin{defn}
+There is a functor $\mathcal{V}:\mathbf{Comp}\to\mathbf{E'sys_{\circ}}$, from
+the category $\mathbf{Comp}$ of split full comprehension categories to
+E'-systems without empty context.
+\end{defn}
+
+\begin{constr}
+Let $\mathcal{P}$ be a comprehension category with underlying category
+$\cat{C}$. We define $\cat{F}$ to be the sub-pre-category of $\cat{C}$ generated
+by the morphisms $\mathcal{P}(A):\ctxext{\Gamma}{A}\to\Gamma$ for $A\in\cat{E}$.
+
+For any $f:\Delta\to\Gamma$, we define a functor $f_\ast:\cat{F}/\Gamma\to
+\cat{F}/\Delta$ by induction on the length of the morphism
+Let $f:\Delta\to\Gamma$ and $p(A)\jdeq\Gamma$. Then we get a pullback square
+\begin{equation*}
+\begin{tikzcd}[column sep=large]
+\ctxext{\Delta}{f_\ast(A)} \arrow[d,fib] \arrow[r,"{\pi_2(f,A)}"] & \ctxext{\Gamma}{A} \arrow[d,fib] \\
+\Delta \arrow[r,"f"] & \Gamma
+\end{tikzcd}
+\end{equation*}
+There seems to be no way to show that
+$\pi_2(f,\ctxext{A}{P})\jdeq \pi_2(\pi_2(f,A),P)$.
+\end{constr}
+
+\begin{conj}
+An E'-system is the same thing as a split full comprehension category in which
+$\cat{E}$ has a terminal object.
+\end{conj}
+
+\begin{proof}
+Let $\mathcal{E}$ be an E'-system. Define $\cat{E}$ to be the full subcategory
+of $\cat{C}^\to$ whose objects are the morphisms of $\cat{F}$. Thus, for any
+$\Gamma\in\cat{C}$, the category $p^{-1}(\Gamma)$ is the category of internal
+morphisms the E-system $\mathcal{E}/\Gamma$.
+
+We first show that $p$ is a Grothendieck fibration. Let $f:\Delta\to\Gamma$ be
+a morphism in $\cat{C}$ and let $A\in\cat{F}/\Gamma$. Then we have $\catid{\Delta},\catid{\Gamma}\in\cat{E}$,
+and $(f,\pi_2(f,A)):f_\ast(A)\to A$. It is routine to verify that $(f,\pi_2(f,A))$
+is a cartesian lift of $f$, and that this choice of catesian lifts is functorial.
+
+Now let $g:X\to\Delta$ and $f:\Delta\to \Gamma$, and let $\tilde{f}:\tilde{\Delta}\to\tilde{\Gamma}$ and
+$\tilde{g}:\tilde{X}\to\tilde{\Delta}$ be cartesian lifts of $f$ and $g$, respectively. We have to show
+that $\tilde{f}\circ \tilde{g}$ is a cartesian lift of $f\circ g$. Let $C\in\cat{F}/X$ and
+let $h : C\to \tilde{\Gamma}$ be a morphism such that $\mathrm{cod}(h)\jdeq f\circ g$. 
+\end{proof}
 
 \begin{comment}
-\section{Acknowledgments}
-The research that eventually became the study of E-systems began while analyzing
-the internel models I encountered during investigations of the type theoretical
-descent theorem, in the summer of 2013. In a private conversation during the Conference on Type Theory, 
-Homotopy Theory and Univalent Foundations at the Universitat Aut\`onoma de 
-Barcelona, I presented the first version of the axioms to Richard Garner, who
-told me he found a similar structure  while analyzing the combinatorics of type
-dependency, see \cite{Garner2014}. After that conference, Joachim
-Kock generously offered me to stay for a few months at the UAB. He was intrigued
-by the idea of doing category theory using only sections, and during the sessions
-we had in this period I finalized the inference rules that correspond to the
-axioms of E-systems.
-
-In Paris, in the spring of 2014, Richard Garner introduced me to Vladimir 
-Voevodsky, who had been working on similar problems and wanted to know exactly
-what axioms should hold on B0-systems so that it is in the image of the functor
-from C-systems to B0-systems. In other words, he wanted to know what the axioms
-of B-systems were, and be confident that we had them all.
-
-In August 2015, I started a grad student of Steve Awodey at the Carnegie 
-Mellon University. With Steve we elaborated much more on the fact that the theory
-of E-systems is essentially algebraic, and I wrote down what an E-system is in
-an arbitrary category with finite limits.
-
-The discussions that followed both with Awodey and with Voevodsky, 
-at the CMU, at the Homotopy Type Theory workshop in Oxford (November 2014), and
-while visiting Voevodsky at the Institute for Advanced Study in Princeton (March 2015), 
-essentially shaped E-systems into their current presentation.
-
-I am grateful to all the people mentioned for sharing many valuable insights 
-with me. 
+\subsection{Quotients of E-systems}
+
+\begin{defn}
+A regular congruence relation $\sim$ on a category $\cat{F}$ with term structure,
+is a triple $(\sim_c,\sim_f,\sim_t)$ consisting of
+\begin{enumerate}
+\item an equivalence relation $\sim_c$ on the set of objects of $\cat{F}$,
+\item an equivalence relation $\sim_f$ on the set of morphisms of $\cat{F}$,
+\item an equivalence relation $\sim_t$ on the set of terms of $\cat{F}$,
+\end{enumerate}
+which are compatible with $\mathrm{cod}$, $\mathrm{dom}$, $\catid{}$, $\circ$
+and $\ebd$
+\end{defn}
 \end{comment}

### 2015-07-29 e-systems.tex
separation of sections into different files

--- 2015-06-12 weakeningcats.tex
@@
 \end{equation*}
-commutes for any $A\in\cat{C}/\Gamma$.
-\end{defn}
-
-\begin{lem}
-Let $\cat{C}$ be a pre-weakening system. Then for any object $\Gamma$ of $\cat{C}$,
-the slice pre-category $\cat{C}/\Gamma$ is also a pre-weakening system, with the weakening
-structure given by $W(\cat{C}/\Gamma)_P\jdeq W(\cat{C})_P$ for each $P\in\cat{C}/\ctxext{\Gamma}{A}$
-and each $A\in\cat{C}/\Gamma$.
+commutes for any $A\in\cat{F}/\Gamma$.
+\end{defn}
+
+\begin{lem}
+Let $\cat{F}$ be a pre-weakening system. Then for any object $\Gamma$ of $\cat{F}$,
+the slice pre-category $\cat{F}/\Gamma$ is also a pre-weakening system, with the weakening
+structure given by $W(\cat{F}/\Gamma)_P\jdeq W(\cat{F})_P$ for each $P\in\cat{F}/\ctxext{\Gamma}{A}$
+and each $A\in\cat{F}/\Gamma$.
 \end{lem}
@@
 \begin{eg}
-If the underlying pre-category is a group, the axioms of a weakening system
-obtain the following form: There is a group homomorphism $W_x:G\to G$ for
-every $x\in G$, in such a way that $W_e\jdeq \catid{G}$ and $W_{xy}\jdeq W_y\circ W_x$
-for every $x,y\in G$, and for any $x,y\in G$ the diagram
+If the underlying pre-category is a monoid $M$, the axioms of a weakening system
+obtain the following form: There is a group homomorphism $W_x:M\to M$ for
+every $x\in M$, in such a way that $W_e\jdeq \catid{M}$ and $W_{xy}\jdeq W_y\circ W_x$
+for every $x,y\in M$, and for any $x,y\in M$ the diagram
 \begin{equation*}
 \begin{tikzcd}
-G \arrow[r,"W_y"]
-  &
-G 
-  \\
-G \arrow[u,"W_x"]
+M \arrow[r,"W_y"]
+  &
+M 
+  \\
+M \arrow[u,"W_x"]
   \arrow[r,swap,"W_y"]
   &
-G \arrow[u,swap,"W_{W_y(x)}"]
-\end{tikzcd}
-\end{equation*}
-is required to commute. Thus, a weakening system on a group $G$ consists of
-a group homomorphism $W:\op G\to\mathrm{Aut}(G)$ such that $W_a\circ W_b\jdeq 
-W_{W_a(b)}\circ W_a$. The latter requirement is satisfied if $b\cdot a\jdeq a\cdot W_a(b)$
-for every $a,b\in G$, or equivalently if $[b,a]\jdeq b^{-1}\cdot W_a(b)$.
-\end{eg}
-
+M \arrow[u,swap,"W_{W_y(x)}"]
+\end{tikzcd}
+\end{equation*}
+is required to commute. Thus, a weakening system on a monoid $M$ consists of
+a homomorphism $W:\op M\to\mathrm{Aut}(M)$ such that $W_x\circ W_y\jdeq 
+W_{W_x(y)}\circ W_x$. The latter requirement is satisfied if $y\cdot x\jdeq x\cdot W_x(y)$
+for every $x,y\in M$. Group conjugation forms a weakening structure on any group.
+\end{eg}
+
+\begin{comment}
 \begin{lem}
@@
 \begin{lem}
-Let $G$ be a graph, and let $\cat{C}_G$ be the free pre-category generated by
-$G$. A pre-weakening structure on $\cat{C}_G$ is determined by functors
-$W_e:\cat{C}_G/\Gamma\to\cat{C}_G/\Delta$ for every edge $e:\Delta\to \Gamma$ in $G$, with the
+Let $G$ be a graph, and let $\cat{F}_G$ be the free pre-category generated by
+$G$. A pre-weakening structure on $\cat{F}_G$ is determined by functors
+$W_e:\cat{F}_G/\Gamma\to\cat{F}_G/\Delta$ for every edge $e:\Delta\to \Gamma$ in $G$, with the
 property that $W_e(\catid{\Gamma})\jdeq \catid{\Delta}$.
 
-A weakening structure on $\cat{C}_G$ is
-a pre-weakening structure on $\cat{C}_G$ such that the diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=huge]
-\cat{C}_G/\greek{Z} \arrow[r,"W_e/(g\circ h)"] & \cat{C}_G/\mathrm{dom}(W_e(g\circ h)) \\
-\cat{C}_G/\greek{E} \arrow[u,"W_h"] \arrow[r,swap,"W_e/g"] & \cat{C}_G/\mathrm{dom}(W_e(g)) \arrow[u,swap,"W_{W_e(h)}"]
-\end{tikzcd}
-\end{equation*}
-commutes for every arrow $g:\greek{E}\to \Gamma$ in $\cat{C}_G$ and every two edges $e:\Delta\to \Gamma$ and
+A weakening structure on $\cat{F}_G$ is
+a pre-weakening structure on $\cat{F}_G$ such that the diagram
+\begin{equation*}
+\begin{tikzcd}[column sep=huge]
+\cat{F}_G/\greek{Z} \arrow[r,"W_e/(g\circ h)"] & \cat{F}_G/\mathrm{dom}(W_e(g\circ h)) \\
+\cat{F}_G/\greek{E} \arrow[u,"W_h"] \arrow[r,swap,"W_e/g"] & \cat{F}_G/\mathrm{dom}(W_e(g)) \arrow[u,swap,"W_{W_e(h)}"]
+\end{tikzcd}
+\end{equation*}
+commutes for every arrow $g:\greek{E}\to \Gamma$ in $\cat{F}_G$ and every two edges $e:\Delta\to \Gamma$ and
 $h:\greek{Z}\to \greek{E}$ in $G$.
@@
 Let $G$ be the graph consisting of one vertex $\ast$ and one edge $e:\ast\to\ast$.
-Then $\cat{C}_G/\ast\jdeq \N$. Then the only weakening structure on $\cat{C}/G$ is
+Then $\cat{F}_G/\ast\jdeq \N$. Then the only weakening structure on $\cat{F}/G$ is
 given by $W_e\jdeq \catid{\N}$. The reason is that 
 \end{eg}
-\end{comment}
 
@@
 \end{eg}
+\end{comment}
 
@@
 \end{equation*}
-for every $A\in\cat{C}/\Gamma$ and $\Gamma\in\cat{C}$.
-\end{defn}
-
-\begin{lem}
-For any object $\Gamma$ in a pre-projection system $\cat{C}$, the slice pre-category
-$\cat{C}/\Gamma$ is a pre-projection system.
+for every $A\in\cat{F}/\Gamma$ and $\Gamma\in\cat{F}$.
+\end{defn}
+
+\begin{lem}
+For any object $\Gamma$ in a pre-projection system $\cat{F}$, the slice pre-category
+$\cat{F}/\Gamma$ is a pre-projection system.
 \end{lem}
@@
 \end{equation*}
-commutes for every $x\in T(A)$ and $A\in\cat{C}/\Gamma$.
-\end{defn}
-
-\begin{lem}
-For any object $\Gamma$ of a pre-substitution system $\cat{C}$, the slice pre-category
-$\cat{C}/\Gamma	$ is a pre-substitution system.
+commutes for every $x\in T(A)$ and $A\in\cat{F}/\Gamma$.
+\end{defn}
+
+\begin{lem}
+For any object $\Gamma$ of a pre-substitution system $\cat{F}$, the slice pre-category
+$\cat{F}/\Gamma	$ is a pre-substitution system.
 \end{lem}
@@
 \end{comment}
-
-\section{Properties of E-systems}\label{sec:esys_props}
-
-The theory of E-systems could be seen as a sort of category theory with only
-projections and their sections. We elaborate on this point of view by defining
-the category of internal homomorphisms, which Lawvere might have called the
-category of terms. 
-In the category of internal morphisms over a context $\Gamma$,
-a morphism $f:A\to B$ is a term of the constant family $W_A(B)$ over 
-$\ctxext{\Gamma}{A}$. Rather than directly defining composition, we will define
-for every $f:A\to B$ in context $\Gamma$, a pre-composition operation
-$f_\ast$, which is going to be an E-homomorphism from 
-$\cat{C}/\ctxext{\Gamma}{B}$ to $\cat{C}/\ctxext{\Gamma}{A}$. This allows us
-to consider diagrams of the following form:
-\begin{equation*}
-\begin{tikzcd}
-P \arrow[d,->>] \arrow[r,"g"] & Q \arrow[d,->>] \\
-A \arrow[r,swap,"f"] & B
-\end{tikzcd}
-\end{equation*}
-where we consider the morphism $g$ over $f$ to be an internal morphism from 
-$P$ to $f_\ast(Q)$ in context $\ctxext{\Gamma}{A}$, capturing the
-essence of being a morphism from the family $P$ over $A$ to the family
-$Q$ over $B$, reindexed by $f$. 
-Whenever we draw such diagrams, a double headed arrow will always
-refer to a family, and an ordinary arrow will refer to an internal morphism,
-possibly over another internal morphism
-
-A morphism $g$ over $f$ can be vertically composed with $f$ so
-that we obtain a morphism $\jvcomp{P}{f}{g}:\ctxext{A}{P}\to\ctxext{B}{Q}$ in 
-context $\Gamma$, and it can be horizontally composed with a morphism $k:Q\to R$
-over $h:B\to C$ so that we obtain a morphism $\jfcomp{A}{f}{P}{g}{k}:P\to R$ 
-over $\jcomp{A}{f}{h}:A\to C$. We will show interchange laws between these
-sorts of composition. The interchange laws have as a consequence that if we
-assert the commutativity of a large composed diagram.
-
-We then demonstrate how to define a category internal to an E-system. The main
-point of this lies mainly in demonstrating how the operations of E-systems can
-be used to define a category, because it is no surprise that a category can be
-defined in type theory. Nevertheless, it is noteworthy that we need no further 
-conditions on E-systems.
-This is unlike the situation in category theory, where we need
-to assume that the ambient category has finite limits in order to say what an
-internal category is. In fact, it is possible to define an E-system internal
-to an E-system, just using the algebraic structure of an E-system, and actually
-defining an internal category was the hardest part of that definition.
-
-Before we begin this section, we introduce some handy shorthands.
-
-\begin{defn}
-Let $A\in\cat{C}/\Gamma$. Recall that $W_A:\cat{C}/\Gamma\to\cat{C}/\ctxext{\Gamma}{A}$ acts on objects,
-morphisms and terms. We introduce the infix form of weakening by $A\in\cat{C}/\Gamma$ to be
-$\ctxwk{A}{\blank}$. Thus, we will write
-\begin{align*}
-\ctxwk{A}{B} & \defeq W_A(B) & & \text{for $B\in\cat{C}/\Gamma$} \\
-\ctxwk{A}{Q} & \defeq W_A(Q) & & \text{for $B\in\cat{C}/\Gamma$ and $Q\in\cat{C}/\ctxext{\Gamma}{B}$}\\
-\ctxwk{A}{g} & \defeq W_A(g) & & \text{for $B\in\cat{C}/\Gamma$, $Q\in\cat{C}/\ctxext{\Gamma}{B}$ and $g\in T(Q)$}
-\end{align*}
-\end{defn}
-
-\begin{defn}
-Let $x\in T(A)$ for a family $A\in\cat{C}/\Gamma$. The infix form of substitution
-by $x$ is taken to be $\subst{x}{\blank}$. Thus, we will write
-\begin{align*}
-\subst{x}{P} & \defeq S_x(P) & & \text{for $P\in\cat{C}/\ctxext{\Gamma}{A}$} \\
-\subst{x}{Q} & \defeq S_x(Q) & & \text{for $P\in\cat{C}/\ctxext{\Gamma}{A}$ and $Q\in\cat{C}/\ctxext{{\Gamma}{A}}{P}$} \\
-\subst{x}{g} & \defeq S_x(g) & & \text{for $P\in\cat{C}/\ctxext{\Gamma}{A}$, $Q\in\cat{C}/\ctxext{{\Gamma}{A}}{P}$ and $g\in T(Q)$}
-\end{align*}
-\end{defn}
-
-\subsection{The category of internal morphisms}\label{subsec:im_cat}
-
-\begin{defn}
-For every $A,B\in\cat{C}/\Gamma$ we define the set 
-\begin{equation*}
-\thom{A}{B}\defeq T(\ctxwk{A}{B}).
-\end{equation*}
-An element $f\in \thom{A}{B}$ is called an \emph{internal morphism in context
-$\Gamma$}. We sometimes write $\jhom{\Gamma}{A}{B}{f}$ to indicate that $f$ is
-an internal morphism over $\Gamma$, or we may draw a diagram of the form
-\begin{equation*}
-\begin{tikzcd}[column sep=tiny]
-A \arrow[rr,"f"] \arrow[dr,fib] & & B \arrow[dl,fib] \\
-& \Gamma
-\end{tikzcd}
-\end{equation*}
-but we shall generally omit the arrows down to $\Gamma$ and say instead that we have
-a diagram in context $\Gamma$. 
-\end{defn}
-
-\begin{rmk}
-Note that $\thom{\catid{\Gamma}}{A}\jdeq T(A)$ for any $A\in\cat{C}/\Gamma$, 
-because we have 
-$W_{\catid{\ctxext{\Gamma}{A}}}\jdeq \catid{\cat{C}/\ctxext{\Gamma}{A}}$.
-
-Note also that $\thom{\ctxext{A}{P}}{B}\jdeq \thom{P}{\ctxwk{A}{B}}$ 
-for any $P\in\cat{C}/\ctxext{\Gamma}{A}$ and $B\in\cat{C}/\Gamma$,
-because $W_{\ctxext{A}{P}}\jdeq W_P\circ W_A$.
-Once we have established a category of which the morphisms are given by 
-$\thom{\blank}{\blank}$, we therefore get that 
-\begin{equation*}
-\ctxext{A}{(\blank)}\dashv W_A.
-\end{equation*}
-The right adjoint to weakening by $A$, if it exists, will be the dependent
-product $\Pi_A$. 
-\end{rmk}
-
-\begin{defn}
-Let $A,B\in\cat{C}/\Gamma$.
-For any $f\in\thom{A}{B}$ we define the pre-composition E-homomorphism
-\begin{equation*}
-f_\ast \defeq S_f\circ W_A/B : \cat{C}/\ctxext{\Gamma}{B}\to\cat{C}/\ctxext{\Gamma}{A}.
-\end{equation*}
-Furthermore, for $g\in\thom{B}{C}$ we may consider $W_B(C)\in\cat{C}/\ctxext{\Gamma}{B}$ as a morphism from $W_B(C)$ to $\catid{\ctxext{\Gamma}{B}}$. Therefore, $f_\ast$ acts on the terms of
-$W_B(C)$, so it makes sense to define $g\circ f\defeq f_\ast(g)$. In fact, it makes
-sense to let $(\blank)\circ f$ be the infix notation for $f_\ast$. 
-\end{defn}
-
-\begin{rmk}
-Note that for $\jhom{\Gamma}{B}{C}{g}$, 
-we have $g\circ f\in T(S_f(W_A/B(W_B(C))))$, whereas we would like that
-$g\circ f\in\thom{A}{C}$. More generally, we can show that
-\begin{equation*}
-S_f\circ (W_A/B)\circ W_B\jdeq W_A.
-\end{equation*}
-Since weakening is a weakening homomorphism, we have
-\begin{equation*}
-S_f\circ (W_A/B)\circ W_B\jdeq S_f\circ W_{W_A(B)}\circ W_A
-\end{equation*}
-By \autoref{tTf:StWf_id} of \autoref{defn:esystems} it we get that
-\begin{equation*}
-S_f\circ W_{W_A(B)}\circ W_A\jdeq W_A.
-\end{equation*}
-\end{rmk}
-
-\begin{rmk}
-To see how this extended notion of precomposition works, suppose we have a diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=tiny]
-Q \arrow[rr,"g"] \arrow[dr,fib] & & R \arrow[dl,fib] \\
-& B
-\end{tikzcd}
-\end{equation*}
-in context $\Gamma$, i.e.~$\jhom{\ctxext{\Gamma}{B}}{Q}{R}{g}$, and let $\jhom{\Gamma}{A}{B}{f}$. Precomposing by $f$ gives
-us a diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=tiny]
-\jcomp{A}{f}{Q} \arrow[rr,"\jcomp{A}{f}{g}"] \arrow[dr,fib] & & \jcomp{A}{f}{R} \arrow[dl,fib] \\
-& A
-\end{tikzcd}
-\end{equation*}
-in context $\Gamma$.
-\end{rmk}
-
-\begin{rmk}
-Note that \autoref{f:SidfWff_id} of \autoref{defn:esystems} asserts precisely
-that ${(\tfid{A})}_\ast\jdeq \catid{\cat{C}/\ctxext{\Gamma}{A}}$ for any $A\in\cat{C}/\Gamma$. In
-particular, it follows that $g\circ\tfid{A}\jdeq g$ for any $g\in\thom{A}{B}$
-\end{rmk}
-
-\begin{lem}\label{lem:compcomp}
-For any $f\in\thom{A}{B}$ and $g\in\thom{B}{C}$ we have $f_\ast\circ g_\ast\jdeq (g\circ f)_\ast$. 
-\end{lem}
-
-\begin{proof}
-\begin{align*}
-f_\ast\circ g_\ast & \jdeq S_f\circ (W_A/B) \circ S_g \circ (W_B/C)\\
-& \jdeq S_f\circ S_{W_A(g)}\circ(W_A/\ctxext{B}{W_B(C)})\circ (W_B/C) \\
-& \jdeq S_{S_f(W_A(g))}\circ (S_f/W_A(W_B(C)))\circ (W_A/\ctxext{B}{W_B(C)}) \circ W_B/C \\
-& \jdeq S_{S_f(W_A(g))}\circ ((S_f\circ (W_A/B) \circ W_B)/C) \\
-& \jdeq S_{S_f(W_A(g))}\circ ((S_f\circ W_{W_A(B)}\circ W_A)/C) \\
-& \jdeq S_{S_f(W_A(g))}\circ W_A/C \\
-& \jdeq (g\circ f)_\ast.\qedhere
-\end{align*}
-\end{proof}
-
-\begin{thm}
-For each object $\Gamma$ of an E-system $\mathbb{E}$, we have a category
-$\mathbb{C}_I(\mathbb{E},\Gamma)$ of internal morphisms of $\mathbb{E}$,
-consisting of the same objects as $\cat{C}/\Gamma$ and the hom-sets 
-$\thom{A}{B}$. 
-\end{thm}
-
-\begin{proof}
-The fact that composition is associative is a direct corollary of
-\autoref{lem:compcomp}. The axiom $(\tfid{A})_\ast\jdeq\catid{\ctxext{\Gamma}{A}}$
-implies that the identity morphisms satisfy the right identity law. It remains
-to show that $\tfid{B}\circ f\jdeq f$. This is a simple calculation:
-\begin{equation*}
-\tfid{B}\circ f 
-  \jdeq
-\subst{f}{\ctxwk{A}{\tfid{B}}}
-  \jdeq
-\subst{f}{\tfid{\ctxwk{A}{B}}}
-  \jdeq
-f.\qedhere
-\end{equation*}
-\end{proof}
-
-\begin{rmk}
-Since any E-homomorphism is both a substitution homomorphism and a weakening
-homomorphism, it follows immediately that any E-homomorphism preserves any
-$f_\ast$, i.e.~for any $f\in\thom{A}{B}$ in $\mathcal{E}$ and any E-homomorphism
-$F:\mathcal{E}\to\mathcal{D}$, the diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=large]
-\mathcal{E}/\ctxext{\Gamma}{A} \arrow[r,"F/\ctxext{\Gamma}{A}"] & \mathcal{D}/F(\ctxext{\Gamma}{A})
-  \\
-\mathcal{E}/\ctxext{\Gamma}{B} \arrow[u,"f_\ast"] \arrow[r,swap,"F/\ctxext{\Gamma}{B}"] & \mathcal{D}/F(\ctxext{\Gamma}{B}) \arrow[u,swap,"F(f)_\ast"]
-\end{tikzcd}
-\end{equation*}
-commutes. This applies to weakening, substitution and pre-composition, and it
-follows that for any E-homomorphism $F:\mathcal{E}\to\mathcal{D}$ and any
-$\Gamma\in\mathcal{E}$, we have a functor $\mathbb{C}_I(F,\Gamma):
-\mathbb{C}_I(\mathcal{E},\Gamma)\to\mathbb{C}_I(\mathcal{D},F(\Gamma))$.
-\end{rmk}
-
-The following fact will be useful:
-
-\begin{lem}\label{lem:compW_W}
-Let $f\in\thom{A}{B}$ be an internal morphism in context $\Gamma$. Then one has
-\begin{equation*}
-f_\ast\circ W_B\jdeq W_A.
-\end{equation*}
-\end{lem}
-
-\begin{proof}
-The proof is a simple calculation:
-\begin{equation*}
-f_\ast\circ W_B\jdeq S_f\circ W_A/B\circ W_B\jdeq S_f\circ W_{W_A(B)}\circ W_A
-\jdeq W_A.\qedhere
-\end{equation*}
-\end{proof}
-
-\subsection{Pairing and the projections}
-
-The composition $\ctxext{A}{P}$ of $A\in\cat{C}/\Gamma$ and $P\in\cat{C}/\ctxext{\Gamma}{A}$
-behaves like a strict $\Sigma$-type. We may define the pairing term
-$\typefont{pair}^{A,P}\defeq\tfid{\ctxext{A}{P}}\in T(W_P(W_A(\ctxext{A}{P})))$
-and the projections and prove several useful properties about them. The strictness
-is found, among other things, in the fact that we can prove judgmental $\eta$-equality,
-and that pairing is strictly associative.
-
-In this section we will make more extensive use of the infix form of the
-weakening and substitution operations.
-
-\begin{defn}
-Let $x\in T(A)$ and $u\in T(S_x(P))$ for $A\in\cat{C}/\Gamma$ and $P\in\cat{C}/\ctxext{\Gamma}{A}$. 
-We define the \emph{term extension of $x$ and $u$} to be
-\begin{equation*}
-\tmext{x}{u}\defeq \subst{u}{{x}{\tfid{\ctxext{A}{P}}}}\in T(\ctxext{A}{P}).
-\end{equation*}
-\end{defn}
-
-To prove anything about the term $\tmext{x}{u}$, we need the following property.
-
-\begin{thm}\label{subst_by_tmext}
-Let $x\in T(A)$ and $u\in T(S_x(P))$ for $A\in\cat{C}/\Gamma$ and $P\in\cat{C}/\ctxext{\Gamma}{A}$.
-Then we have
-\begin{equation*}
-S_{\tmext{x}{u}}\jdeq S_u\circ (S_x/P):\cat{C}/\ctxext{\Gamma}{{A}{P}}\to\cat{C}/\Gamma
-\end{equation*}
-\end{thm}
-
-\begin{proof}
-\begin{align*}
-S_{\tmext{x}{u}} & \jdeq S_{S_u(S_x(\tfid{\ctxext{A}{P}}))} \\
-& \jdeq S_{S_u(S_x(\tfid{\ctxext{A}{P}}))}\circ (S_u\circ W_{S_x(P)})\circ (S_x\circ W_{A}) \\
-& \jdeq S_u \circ S_{S_x(\tfid{\ctxext{A}{P}})} \circ W_{S_x(P)}\circ S_x\circ W_A \\
-& \jdeq S_u \circ (S_x/P) \circ S_{\tfid{\ctxext{A}{P}}} \circ W_P\circ W_A \\
-& \jdeq S_u \circ (S_x/P) \circ S_{\tfid{\ctxext{A}{P}}} \circ W_{\ctxext{A}{P}}\\
-& \jdeq S_u \circ (S_x/P).\qedhere
-\end{align*}
-\end{proof}
-
-\begin{defn}
-Let $A\in\cat{C}/\Gamma$ and $P\in\cat{C}/\ctxext{\Gamma}{A}$. We define
-\begin{align*}
-\cprojfstf{A}{P} & \defeq \ctxwk{P}{\tfid{A}}\in\thom{\ctxext{A}{P}}{A}\\
-\cprojsndf{A}{P} & \defeq \tfid{P}\in\thomd{\cprojfstf{A}{P}}{\ctxext{A}{P}}{P}
-\end{align*}
-\end{defn}
-
-\begin{cor}\label{cor:pairing_ehom}
-Let $F:\mathcal{E}\to\mathcal{D}$ be an E-homomorphism and let $x\in T(A)$ and
-$u\in T(S_x(P))$. Then we have
-\begin{equation*}
-F(\tmext{x}{u})\jdeq \tmext{F(x)}{F(u)}.
-\end{equation*}
-Also, we have
-\begin{align*}
-F(\cprojfstf{A}{P}) & \jdeq \cprojfstf{F(A)}{F(P)} \\
-F(\cprojsndf{A}{P}) & \jdeq \cprojsndf{F(A)}{F(P)}.
-\end{align*}
-\end{cor}
-
-\begin{thm}\label{thm:famext_up}
-For any $A\in\cat{C}/\Gamma$ and $P\in\cat{C}/\ctxext{\Gamma}{A}$,
-the map
-\begin{equation*}
-(x,u)\mapsto \tmext{x}{u}:\big(\bigsqcup\nolimits_{x\in T(A)} T(\subst{x}{P})\big)\to T(\ctxext{A}{P}).
-\end{equation*}
-is an isomorphism.  
-\end{thm}
-
-\begin{proof}
-Lets write $\varphi$ for the indicated map. The inverse $\psi$ of $\varphi$ is defined
-by $w\mapsto (\subst{w}{\cprojfstf{A}{P}},\subst{w}{\cprojsndf{A}{P}})$. We have
-to show that for any $x\in T(A)$ and $u\in T(\subst{x}{P})$, one has
-\begin{align*}
-\subst{\tmext{x}{u}}{\cprojfstf{A}{P}} & \jdeq x\\
-\subst{\tmext{x}{u}}{\cprojsndf{A}{P}} & \jdeq u,
-\end{align*}
-and that for any $w\in T(\ctxext{A}{P})$, one has
-\begin{equation*}
-\tmext{\subst{w}{\cprojfstf{A}{P}}}{\subst{w}{\cprojsndf{A}{P}}} \jdeq w.
-\end{equation*}
-To show that $\subst{\tmext{x}{u}}{\cprojfstf{A}{P}} \jdeq x$, we use that
-$S_{\tmext{x}{u}}\jdeq S_u\circ S_x/P$ to show that
-\begin{align*}
-\subst{\tmext{x}{u}}{\cprojfstf{A}{P}}
-  & \jdeq 
-\subst{\tmext{x}{u}}{\ctxwk{P}{\tfid{A}}} \\
-  & \jdeq
-\subst{u}{{x}{\ctxwk{P}{\tfid{A}}}} \tag{By \autoref{subst_by_tmext}}\\
-  & \jdeq
-\subst{u}{\ctxwk{\subst{x}{P}}{\subst{x}{\tfid{A}}}} \\
-  & \jdeq
-\subst{x}{\tfid{A}} \\
-  & \jdeq
-x
-\end{align*}
-To show that $\subst{\tmext{x}{u}}{\cprojsndf{A}{P}}\jdeq u$, note that
-\begin{equation*}
-\subst{\tmext{x}{u}}{\cprojsndf{A}{P}}
-  \jdeq
-\subst{u}{{x}{\tfid{P}}}
-  \jdeq
-\subst{u}{\tfid{\subst{x}{P}}}
-  \jdeq
-u
-\end{equation*}
-Therefore, $\psi$ is a left-inverse of $\varphi$. To show that it is also a
-right inverse, note that \autoref{cor:pairing_ehom} gives us that
-\begin{equation*}
-\tmext{\cprojfst{A}{P}{w}}{\cprojsnd{A}{P}{w}}
-  \jdeq
-\subst{w}{(\tmext{\cprojfstf{A}{P}}{\cprojsndf{A}{P}})}.
-\end{equation*}
-Thus, it suffices to show that $\tmext{\cprojfstf{A}{P}}{\cprojsndf{A}{P}}\jdeq
-\tfid{\ctxext{A}{P}}$, which is straightforward.
-\end{proof}
-
-\begin{lem}\label{lem:tmext_assoc}
-For any $x\in T(A)$, $u\in T(S_x(P))$ and $v\in T(S_{\tmext{x}{u}}(Q))$ we have
-\begin{equation*}
-\tmext{{x}{u}}{v}\jdeq \tmext{x}{{u}{v}}\in T(\ctxext{{A}{P}}{Q}).
-\end{equation*}
-\begin{comment}
-Also, we have
-\begin{align*}
-\cprojfstf{A}{P}\circ\cprojfstf{\ctxext{A}{P}}{Q} & \jdeq\cprojfstf{A}{\ctxext{P}{Q}} \\
-\cprojsndf{A}{P}\circ\cprojfstf{\ctxext{A}{P}}{Q} & \jdeq \cprojfstf{P}{Q}\circ\cprojsndf{A}{\ctxext{P}{Q}} \\
-\cprojsndf{\ctxext{A}{P}}{Q} & \jdeq \cprojsndf{P}{Q}\circ\cprojsndf{A}{\ctxext{P}{Q}}.
-\end{align*}
-\end{comment}
-\end{lem}
-
-\begin{proof}
-By \autoref{subst_by_tmext}, we have $S_v\circ S_{\tmext{x}{u}}\jdeq S_v\circ S_u\circ S_x \jdeq S_{\tmext{u}{v}}\circ S_x$,
-so associativity of term extension follows.
-\end{proof}
-
-\begin{comment}
-\begin{lem}
-Let $f\in\thom{A}{\ctxext{B}{Q}}$ Then we have
-\begin{equation*}
-f \jdeq \tmext{\jcomp{A}{f}{\cprojfstf{B}{Q}}}{\jcomp{A}{f}{\cprojsndf{B}{Q}}}.
-\end{equation*}
-Alternatively, when $f_0\in\thom{A}{B}$ and $f_1\in\thom{A}{\jcomp{A}{f}{Q}}$,
-then we have $\tmext{f_0}{f_1}\in\thom{A}{\ctxext{B}{Q}}$ and
-\begin{align*}
-\jcomp{A}{\tmext{f_0}{f_1}}{\cprojfstf{B}{Q}} & \jdeq f_0 \\
-\jcomp{A}{\tmext{f_0}{f_1}}{\cprojsndf{B}{Q}} & \jdeq f_1.
-\end{align*}
-\end{lem}
-
-\begin{proof}
-Straightforward
-\end{proof}
-\end{comment}
-
-We end this section by showing that weakening is the same operation as pre-composition
-with a first projection.
-
-\begin{thm}\label{precomp_by_proj}
-Precomposition with $\cprojfstf{A}{P}$ is weakening by $P$, i.e.
-\begin{equation*}
-(\cprojfstf{A}{P})_\ast \jdeq W_P.
-\end{equation*}
-\end{thm}
-
-\begin{proof}
-\begin{align*}
-(\cprojfstf{A}{P})_\ast & \jdeq S_{\cprojfstf{A}{P}}\circ W_{\ctxext{A}{P}} \\
-& \jdeq S_{W_P(\tfid{A})}\circ W_P\circ W_A \\
-& \jdeq W_P\circ S_{\tfid{A}}\circ W_A \\
-& \jdeq W_P.\qedhere 
-\end{align*}
-\end{proof}
-
-\subsection{The interchange laws}
-
-We are now in the position to define vertical and horizontal composition, and
-prove properties of them.
-
-\begin{defn}
-Let $\jhom{\Gamma}{A}{B}{f}$ and $\jfhom{\Gamma}{A}{B}{f}{P}{Q}{F}$. Then we
-define 
-\begin{equation*}
-\jhomdefn{\Gamma}{{A}{P}}{{B}{Q}}{\jvcomp{P}{f}{F}}{\unfold{\jvcomp{P}{f}{F}}}.
-\end{equation*}
-Whenever we say that we have a diagram of the form
-\begin{equation*}
-\begin{tikzcd}
-R \arrow[r,"f_2"] \arrow[d,fib] &
-S \arrow[d,fib] \\
-P \arrow[r,"f_1"] \arrow[d,fib] &
-Q \arrow[d,fib] \\
-A \arrow[r,"f_0"] &
-B
-\end{tikzcd}
-\end{equation*}
-we mean that we have
-$f_0\in\thom{A}{B}$, $f_1\in\thomd{f_0}{P}{Q}$ and 
-$f_2\in\thomd{\jvcomp{P}{f_0}{f_1}}{R}{S}$.
-\end{defn}
-
-\begin{cor}
-Let $\jhom{\Gamma}{A}{B}{f}$ and $\jfhom{\Gamma}{A}{B}{f}{P}{Q}{F}$. Then we
-have
-\begin{equation*}
-H(\jvcomp{P}{f}{F})\jdeq \jvcomp{H(P)}{H(f)}{H(F)}
-\end{equation*}
-for any E-homomorphism $H:\mathbb{E}\to\mathbb{D}$.
-\end{cor}
-
-\begin{lem}
-Vertical composition is associative.
-\end{lem}
-
-\begin{proof}
-Consider the diagram
-\begin{equation*}
-\begin{tikzcd}
-R \arrow[r,"f_2"] \arrow[d,fib] &
-S \arrow[d,fib] \\
-P \arrow[r,"f_1"] \arrow[d,fib] &
-Q \arrow[d,fib] \\
-A \arrow[r,"f_0"] &
-B
-\end{tikzcd}
-\end{equation*}
-in context $\Gamma$.
-Because weakening distributes over term extension, and term extension is
-associative, we have
-\begin{align*}
-\jvcomp{R}{\jvcomp{P}{f_0}{f_1}}{f_2}
-  & \jdeq
-\tmext{\ctxwk{R}{\tmext{\ctxwk{P}{f_0}}{f_1}}}{f_2}
-  \\
-  & \jdeq
-\tmext{\tmext{\ctxwk{R}{{P}{f_0}}}{\ctxwk{R}{f_1}}}{f_2}
-  \\
-  & \jdeq
-\tmext{\ctxwk{\ctxext{P}{R}}{f_0}}{\tmext{\ctxwk{R}{f_1}}{f_2}}
-  \tag{By \autoref{lem:tmext_assoc}}
-  \\
-  & \jdeq
-\jvcomp{{P}{R}}{f_0}{\jvcomp{R}{f_1}{f_2}}.\qedhere
-\end{align*}
-\end{proof}
-
-\begin{defn}
-Let $\jhom{\Gamma}{A}{B}{f}$ and $\jfhom{\Gamma}{A}{B}{f}{P}{Q}{F}$. Then we
-define the E-homomorphism 
-\begin{equation*}
-F_\circledast\defeq F_\ast\circ (f_\ast/Q):
-\mathbb{E}/\ctxext{{\Gamma}{B}}{Q}\to\mathbb{E}/\ctxext{{\Gamma}{A}}{P}.
-\end{equation*}
-The infix notation of $F_\circledast$ is taken to be $\jfcomp{\blank}{\blank}{\blank}{F}{\blank}$.
-\end{defn}
-
-\begin{lem}\label{lem:three-composition}
-Let $\jhom{\Gamma}{A}{B}{f}$ and $\jfhom{\Gamma}{A}{B}{f}{P}{Q}{F}$. Then we
-have the equality
-\begin{equation*}
-F_\circledast\jdeq (\jvcomp{P}{f}{F})_\ast.
-\end{equation*}
-\end{lem}
-
-\begin{proof}
-\begin{align*}
-F_\ast\circ (f_\ast/Q)
-  & \jdeq
-S_F\circ W_P\circ S_f/(W_A(Q))\circ W_A/\ctxext{B}{Q}
-  \\
-  & \jdeq
-S_F\circ S_{W_P(f)}/W_P(W_A(Q))\circ W_P/W_A(\ctxext{B}{Q})\circ W_A/\ctxext{B}{Q}
-  \\
-  & \jdeq
-S_F\circ S_{W_P(f)}/W_P(W_A(Q))\circ W_{\ctxext{A}{P}}/\ctxext{B}{Q}
-  \\
-  & \jdeq
-S_{\tmext{W_P(f)}{F}}\circ W_{\ctxext{A}{P}}/\ctxext{B}{Q}
-  \tag{By \autoref{subst_by_tmext}}
-  \\
-  & \jdeq
-(\jvcomp{P}{f}{F})_\ast.\qedhere
-\end{align*}
-\end{proof}
-
-We are now in position to state and prove the interchange law of horizontal
-and vertical composition.
-
-\begin{thm}\label{thm:interchange}
-Consider the diagram
-\begin{equation*}
-\begin{tikzcd}
-P \arrow[r,"F"] \arrow[d,fib] &
-Q \arrow[r,"G"] \arrow[d,fib] &
-R \arrow[d,fib] \\
-A \arrow[r,"f"] &
-B \arrow[r,"g"] &
-C
-\end{tikzcd}
-\end{equation*}
-in context $\Gamma$. Then the equality
-\begin{equation*}
-{\jcomp{{A}{P}}{\jvcomp{P}{f}{F}}{\jvcomp{Q}{g}{G}}}
-  \jdeq
-{\jvcomp{P}{\jcomp{A}{f}{g}}{\jfcomp{A}{f}{P}{F}{G}}}
-\end{equation*}
-of morphisms from $\ctxext{A}{P}$ to $\ctxext{C}{R}$ in context $\Gamma$ holds.
-\end{thm}
-
-\begin{proof}
-By \autoref{lem:three-composition}, we have
-\begin{align*}
-\jcomp{{A}{P}}{\jvcomp{P}{f}{F}}{\jvcomp{Q}{g}{G}}
-  & \jdeq
-\jcomp{P}{F}{\jcomp{A}{f}{\tmext{\ctxwk{Q}{g}}{G}}}
-  \tag{By \autoref{lem:three-composition}}
-  \\
-  & \jdeq
-\jcomp{P}{F}{\tmext{\ctxwk{\jcomp{A}{f}{Q}}{\jcomp{A}{f}{g}}}{\jcomp{A}{f}{G}}}
-  \\
-  & \jdeq
-\tmext{\jcomp{P}{F}{\ctxwk{\jcomp{A}{f}{Q}}{\jcomp{A}{f}{g}}}}{\jcomp{P}{F}{{A}{f}{G}}}
-  \\
-  & \jdeq
-\tmext{\jcomp{P}{F}{\ctxwk{\jcomp{A}{f}{Q}}{\jcomp{A}{f}{g}}}}{\jfcomp{A}{f}{P}{F}{G}}
-  \\
-  & \jdeq
-\tmext{\ctxwk{P}{\jcomp{A}{f}{g}}}{\jfcomp{A}{f}{P}{F}{G}}
-  \tag{By \autoref{lem:compW_W}}
-  \\
-  & \jdeq
-\jvcomp{P}{\jcomp{A}{f}{g}}{\jfcomp{A}{f}{P}{F}{G}}.\qedhere
-\end{align*}
-\end{proof}
-
-\begin{thm}
-Consider the diagram
-\begin{equation*}
-\begin{tikzcd}
-P \arrow[r,"F"] \arrow[d,fib] &
-Q \arrow[r,"G"] \arrow[d,fib] &
-R \arrow[d,fib] \\
-A \arrow[r,"f"] &
-B \arrow[r,"g"] &
-C
-\end{tikzcd}
-\end{equation*}
-in context $\Gamma$. 
-Then $F_\circledast\circ G_\circledast\jdeq (\jfcomp{A}{f}{P}{F}{G})_\circledast$.
-In other words the composition $\jfcomp{}{}{}{\blank}{\blank}$ is associative.
-\end{thm}
-
-\begin{proof}
-\begin{align*}
-F_\circledast\circ G_\circledast
-  & \jdeq
-(\jvcomp{P}{f}{F})_\ast\circ(\jvcomp{Q}{g}{G})_\ast 
-  \tag{By \autoref{lem:three-composition}}\\
-  & \jdeq
-(\jcomp{\ctxext{A}{P}}{\jvcomp{P}{f}{F}}{\jvcomp{Q}{g}{G}})_\ast 
-  \tag{By \autoref{lem:compcomp}}\\
-  & \jdeq
-(\jvcomp{}{\jcomp{A}{f}{g}}{\jfcomp{A}{f}{P}{F}{G}})_\ast 
-  \tag{By \autoref{thm:interchange}} \\
-  & \jdeq
-(\jfcomp{A}{f}{P}{F}{G})_\circledast.
-  \tag{By \autoref{lem:three-composition}}
-\end{align*}
-\end{proof}
-
-\begin{lem}
-Let $\jhom{\Gamma}{A}{B}{f}$ and $\jfhom{\Gamma}{A}{B}{f}{P}{Q}{F}$. Then
-$\jvcomp{P}{f}{F}$ is the unique morphism from $\ctxext{A}{P}$ to $\ctxext{B}{Q}$
-with the property that both the diagram
-\begin{equation*}
-\begin{tikzcd}
-\ctxext{A}{P}
-  \ar{r}{\jvcomp{P}{f}{F}}
-  \ar{d}[swap]{\cprojfstf{A}{P}}
-& \ctxext{B}{Q}
-  \ar{d}{\cprojfstf{B}{Q}}
-  \\
-A \ar{r}[swap]{f}
-& B
-\end{tikzcd}
-\end{equation*}
-commutes, and $\jcomp{}{\jvcomp{P}{f}{F}}{\cprojsndf{B}{Q}}\jdeq F$.
-\end{lem}
-
-\begin{proof}
-We first note that
-\begin{align*}
-\jcomp{}{\jvcomp{P}{f}{F}}{\cprojfstf{B}{Q}}
-  & \jdeq
-\jcomp{P}{F}{\jcomp{A}{f}{\cprojfstf{B}{Q}}} 
-  \tag{By \autoref{lem:three-composition}} \\
-  & \jdeq
-\jcomp{P}{F}{\ctxwk{\jcomp{A}{f}{Q}}{\jcomp{A}{f}{\idtm{B}}}} \\
-  & \jdeq
-\jcomp{P}{F}{\ctxwk{\jcomp{A}{f}{Q}}{f}} \\
-  & \jdeq
-\ctxwk{P} f \tag{By \autoref{lem:compW_W}} \\
-  & \jdeq
-\jcomp{\ctxext{A}{P}}{\cprojfstf{A}{P}}{f}.
-  \tag{By \autoref{precomp_by_proj}}
-\end{align*}
-Also, we have
-\begin{align*}
-\jcomp{}{\jvcomp{P}{f}{F}}{\cprojsndf{B}{Q}}
-  & \jdeq
-\jcomp{P}{F}{\jcomp{A}{f}{\idtm{Q}}}
-  \tag{By \autoref{lem:three-composition}}
-  \\
-  & \jdeq
-\jcomp{P}{F}{\idtm{\jcomp{A}{f}{Q}}}
-  \\
-  & \jdeq
-F.
-\end{align*}
-Thus, we conclude that $\jvcomp{P}{f}{F}$ has indeed the stated property. For
-the uniqueness, let $G:\ctxext{A}{P}\to\ctxext{B}{Q}$ be a morphism such that
-$\jcomp{}{G}{\cprojfstf{B}{Q}}\jdeq\jcomp{}{\cprojfstf{A}{P}}{f}$ and
-$\jcomp{}{G}{\cprojsndf{B}{Q}}\jdeq F$. Then it follows that
-\begin{equation*}
-G \jdeq \tmext{\jcomp{}{\cprojfstf{A}{P}}{f}}{F}
-  \jdeq \tmext{\ctxwk{P}{f}}{F}
-  \jdeq \jvcomp{P}{f}{F}.\qedhere
-\end{equation*}
-\end{proof}
-
-\subsection{Categories internal to E-systems}
-The definition of internal category we're about to give, has a type of objects
-and a type of morphisms, such that the codomain function is a projection. It might
-be more appropriate to call such structures `right categories', and the corresponding
-structures where the domain function is a projection might be called `left
-categories'. In particular, the opposite of a right category will be a left
-category. There is also a symmetric notion of a category.
-
-\begin{defn}
-A category-framework $\cftalg{A}$ in context $\Gamma$ in an E-system $\mathbb{E}$ consists of
-\begin{enumerate}
-\item A family $\cftalgc{\cftalg{A}}\in\cat{C}/\Gamma$, representing the type of
-objects of the category framework;
-\item A family $\cftalgf{\cftalg{A}}\in\cat{C}/\ctxext{\Gamma}{\cftalgc{\cftalg{A}}}$,
-representing the type of morphisms of the category framework, where a term of
-$\subst{X}{\cftalgf{\cftalg{A}}}$ is considered to be a morphism into $X\in T(\cftalgc{\cftalg{A}})$;
-\item A morphism $\jhom{\Gamma}{\ctxext{\cftalgc{\cftalg{A}}}{\cftalgf{\cftalg{A}}}}{\cftalgc{\cftalg{A}}}{\cftctxext}$, representing the domain function;
-\item A morphism $\jhom{{\Gamma}{\cftalgc{\cftalg{A}}}}{{\cftalgf{\cftalg{A}}}{\jcomp{}{\cftctxext}{\cftalgf{\cftalg{A}}}}}{\cftalgf{\cftalg{A}}}{\cftfamext}$, representing composition;
-\item A term $\jterm{{\Gamma}{\cftalgc{\cftalg{A}}}}{\cftalgf{\cftalg{A}}}{\cftempf{\cftalg{A}}}$,
-representing the identity morphisms.
-\end{enumerate}
-\end{defn}
-
-\begin{rmk}
-Instead of looking at $\cftfamext$ as a morphism from 
-$\ctxext{\cftalgf{\cftalg{A}}}{\jcomp{}{\cftctxext}{\cftalgf{\cftalg{A}}}}$
-to $\cftalgf{\cftalg{A}}$ in context $\ctxext{\Gamma}{\cftalgc{\cftalg{A}}}$, 
-one could also look at $\cftfamext$ as a 
-morphism \emph{over $\cprojfstf{\cftalgc{\cftalg{A}}}{\cftalgf{\cftalg{A}}}$} 
-in context $\Gamma$, as indicated in the following diagram:
-\begin{equation*}
-\begin{tikzcd}[column sep=large]
-\cftalgf{\cftalg{A}}
-  \ar[fib]{d}
-& \jcomp{}{\cftctxext}{\cftalgf{\cftalg{A}}}
-  \ar[fib]{d}
-  \ar{l}[swap]{\cftfamext}
-  \ar{r}
-& \cftalgf{\cftalg{A}}
-  \ar[fib]{d}
-  \\
-\cftalgc{\cftalg{A}}
-& \ctxext{A}{\cftalgf{\cftalg{A}}}
-  \ar{l}{\cprojfstf{A}{\cftalgf{\cftalg{A}}}}
-  \ar{r}[swap]{\cftctxext}
-& \cftalgc{\cftalg{A}}
-\end{tikzcd}
-\end{equation*}
-From this point of view, we see that $\cftfamext$ has an object $X$ and a 
-morphism $f$ into $X$ as parameters, 
-then takes a morphism into the domain of $f$ as an argument, 
-and outputs a morphism into $X$. The projection maps $(X,f)$ to $X$, so we see
-that composition is indeed an operation of this kind.
-\end{rmk}
-
-A category-framework is not a category unless several axioms of the theory of
-categories are satisfied. The domain of a composed arrow $g\circ f$ must be the
-domain of $f$, composition must be associative, the domains of the identity
-morphisms must be the same as their codomains, and identities must satisfy
-the left and right identity laws. However, the axioms expressing these 
-conditions are not very insightful when they're written down directly. Thus, we
-take the usual way and define the notion of functor before we describe the
-actual conditions of being a category.
-
-\begin{defn}
-Let $\cftalg{A}$ and $\cftalg{B}$ be category-frameworks in context $\Gamma$,
-in an E-system $\mathbb{E}$.
-An \emph{internal functor $\cfthom{F}$ from $\extalg{A}$ to
-$\extalg{B}$ in context $\Gamma$} is a pair 
-$(\cfthomc{\cfthom{F}},\cfthomf{\cfthom{F}})$ consisting of
-\begin{align*}
-\jalign\jhom
-  {\Gamma}
-  {\cftalgc{\cftalg{A}}}
-  {\cftalgc{\cftalg{B}}}
-  {\cfthomc{\cfthom{F}}}
-  \\
-\jalign\jfhom
-  {\Gamma}
-  {\cftalgc{\cftalg{A}}}
-  {\cftalgc{\cftalg{B}}}
-  {\cfthomc{\cfthom{F}}}
-  {\cftalgf{\cftalg{A}}}
-  {\cftalgf{\cftalg{B}}}
-  {\cfthomf{\cfthom{F}}}
-\end{align*}
-for which the diagrams
-\begin{equation}\label{eq:exthom1}
-\begin{tikzcd}
-\ctxext{\cftalgc{\cftalg{A}}}{\cftalgf{\cftalg{A}}}
-  \ar{r}{\jvcomp{}{\cfthomc{\cfthom{F}}}{\cfthomf{\cfthom{F}}}}
-  \ar{d}[swap]{\cftctxext[\cftalg{A}]}
-& \ctxext{\cftalgc{\cftalg{B}}}{\cftalgf{\cftalg{B}}}
-  \ar{d}{\cftctxext[\cftalg{B}]}
-  \\
-\cftalgc{\cftalg{A}}
-  \ar{r}[swap]{\cfthomc{\cfthom{F}}}
-& \cftalgc{\cftalg{B}}
-\end{tikzcd}
-\end{equation}
-and
-\begin{equation}\label{eq:exthom2}
-\begin{tikzcd}[column sep=huge]
-\ctxext{\cftalgf{\cftalg{A}}}{\jcomp{}{\cftctxext[\cftalg{A}]}{\cftalgf{\cftalg{A}}}}
-  \ar{r}{\jvcomp{}{\cfthomf{\cfthom{F}}}{\jcomp{}{\cftctxext[\cftalg{A}]}{\cfthomf{\cfthom{F}}}}}
-  \ar{d}[swap]{\cftfamext[\cftalg{A}]}
-& \jcomp{}{\cfthomc{\cfthom{F}}}{\ctxext{\cftalgf{\cftalg{B}}}{\jcomp{}{\cftctxext[\cftalg{B}]}{\cftalgf{\cftalg{B}}}}}
-  \ar{d}{\jcomp{}{\cfthomc{\cfthom{F}}}{\cftfamext[\cftalg{B}]}}
-  \\
-\cftalgf{\cftalg{A}}
-  \ar{r}[swap]{\cfthomf{\cfthom{F}}}
-& \jcomp{}{\cfthomc{\cfthom{F}}}{\cftalgf{\cftalg{B}}}
-\end{tikzcd}
-\end{equation}
-commute, and for which $\subst{\cftempf{\cftalg{A}}}{\cfthomf{\cfthom{F}}}
-\jdeq\jcomp{}{\cfthomc{\cfthom{F}}}{\cftempf{\cftalg{B}}}$.
-\end{defn}
-
-\begin{defn}
-Let $\cftalg{A}$ be a category-framework. We define the category-framework
-\begin{equation*}
-\cftfamalg{\cftalg{A}}\defeq(\cftalgf{\cftalg{A}},\jcomp{}{\cftctxext}%
-{\cftalgf{\cftalg{A}}},\cftfamext,\jcomp{}{\cftctxext}{\cftfamext},
-\jcomp{}{\cftctxext}{\cftempf{\cftalg{A}}}),
-\end{equation*}
-When $\cftalg{P}$ is a category-framework in context $\ctxext{\Gamma}{A}$, then
-and we define the category-framework
-\begin{equation*}
-\ctxext{\cftalgc{\cftalg{A}}}{\cftalg{P}}
-  \defeq
-( \ctxext{A}{\cftalgc{\cftalg{P}}},
-  \cftalgf{\cftalg{P}},
-  \jvcomp{}{\idtm{A}}{\cftctxext[\cftalg{P}]},
-  \cftfamext[\cftalg{P}],
-  \cftempf{\cftalg{P}}).
-\end{equation*}
-\end{defn}
-
-\begin{defn}
-An internal category $\cftalg{A}$ in an E-system $\mathbb{E}$ consists of a
-category-framework $\cftalg{A}$ satisfying the following conditions:
-\begin{enumerate}
-\item \emph{The domain condition.} The pair 
-$(\cftctxext,\idtm{\jcomp{}{\cftctxext}{\cftalgf{\cftalg{A}}}})$ is a
-functor from $\ctxext{\cftalgc{\cftalg{A}}}{\cftfamalg{\cftalg{A}}}$ to
-$\cftalg{A}$.
-\item \emph{The associativity law.} The pair
-$(\cftfamext,\idtm{\jcomp{}{\cftfamext}{\cftalgf{\cftfamalg{\cftalg{A}}}}})$
-is a functor from $\ctxext{\cftalgf{\cftalg{A}}}{\cftfamalg{\cftfamalg{\cftalg{A}}}}$ to
-$\cftfamalg{\cftalg{A}}$.
-\item \emph{The identity domain condition.} The domain of the identity morphism
-is the same as its codomain:
-\begin{equation*}
-\jtermeq
-  {{\Gamma}{\cftalgc{\cftalg{A}}}}
-  {\ctxwk{\cftalgc{\cftalg{A}}}{\cftalgc{\cftalg{A}}}}
-  {\subst{\cftempf{\cftalg{A}}}{\cftctxext[\cftalg{A}]}}
-  {\idtm{\cftalgc{\cftalg{A}}}}.
-\end{equation*}
-\item \emph{The left identity law.} A morphism precomposed by the identity morphism
-is equal to the original morphism:
-\begin{equation*}
-\jtermeq
-  {{{\Gamma}{\cftalgc{\cftalg{A}}}}{\cftalgf{\cftalg{A}}}}
-  {\ctxwk{\cftalgf{\cftalg{A}}}{\cftalgf{\cftalg{A}}}}
-  {\subst{\cftempf{\cftalg{A}}}{\cftfamext}}
-  {\idtm{\cftalgf{\cftalg{A}}}}.
-\end{equation*}
-\item \emph{The right identity law.} A morphism postcomposed by the identity morphism
-is equal to the original morphism:
-\begin{equation*}
-\jtermeq
-  {{{\Gamma}{\cftalgc{\cftalg{A}}}}{\cftalgf{\cftalg{A}}}}
-  {\ctxwk{\cftalgf{\cftalg{A}}}{\cftalgf{\cftalg{A}}}}
-  {\subst{\jcomp{}{\cftctxext}{\cftempf{\cftalg{A}}}}{\cftfamext}}
-  {\idtm{\cftalgf{\cftalg{A}}}}.
-\end{equation*}
-\end{enumerate}
-\end{defn}
-
-\begin{rmk}
-Since we have only used operations of E-systems to define internal categories,
-it immediately follows that if $\cftalg{A}$ is an internal category in
-$\mathbb{E}$ and if $F:\mathbb{E}\to\mathbb{D}$ is an E-homomorphism, then
-$F(\cftalg{A})$ is an internal category in $\mathbb{D}$. In particular, if
-$\cftalg{Q}$ is an internal category in context $\ctxext{\Gamma}{B}$ and
-$f:A\to B$, then the change of base $f_\ast(\cftalg{Q})$ is an internal category
-in context $\ctxext{\Gamma}{A}$. Thus, it is immediate that the notion of
-internal category is stable under change of base.
-\end{rmk}
-
-\subsection{Internal E-systems}
-
-\subsubsection{Internal weakening systems}
-
-\subsubsection{Internal projection systems}
-
-\subsubsection{Internal substitution systems}
-
-\subsubsection{Internal E-systems}
-
-\begin{comment}
-\section{Dependent morphisms}
-
-\begin{defn}
-Let $f_0\in\thom{A}{B}$ be an internal morphism and let $P\in\cat{C}/\ctxext{\Gamma}{A}$ and
-$Q\in\cat{C}/\ctxext{\Gamma}{B}$. We define
-\begin{equation*}
-\thomd{f_0}{P}{Q} \defeq \thom{P}{{f_0}_\ast(Q)}
-\end{equation*}
-We indicate that $f_1\in\thomd{f_0}{P}{Q}$ by drawing the diagram
-\begin{equation*}
-\begin{tikzcd}
-P \arrow[r,"f_1"] \arrow[d,->>] & Q \arrow[d,->>] \\
-A \arrow[r,swap,"f_0"] & B.
-\end{tikzcd}
-\end{equation*}
-\end{defn}
-
-\begin{defn}
-Consider the diagram
-\begin{equation*}
-\begin{tikzcd}
-P \arrow[r,"f_1"] \arrow[d,->>] & Q \arrow[d,->>] \arrow[r,"g_1"] & R \arrow[d,->>] \\
-A \arrow[r,swap,"f_0"] & B \arrow[r,"g_0"] & C.
-\end{tikzcd}
-\end{equation*}
-We define
-\begin{align*}
-\jvcomp{P}{f_0}{f_1} & \defeq\unfold\jvcomp{P}{f_0}{f_1} & & \in \thom{\ctxext{A}{P}}{\ctxext{B}{Q}} \\
-\jfcomp{A}{f_0}{P}{f_1}{g_1} & \defeq \unfold\jfcomp{A}{f_0}{P}{f_1}{g_1} & & \in \thomd{\jcomp{A}{f_0}{g_0}}{P}{R}.
-\end{align*}
-\end{defn}
-
-\begin{rmk}
-Consider
-\begin{equation*}
-\begin{tikzcd}
-P \arrow[r,"f_1"] \arrow[d,->>] & Q \arrow[d,->>] \\
-A \arrow[r,swap,"f_0"] & B.
-\end{tikzcd}
-\end{equation*}
-Then the diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=large]
-\ctxext{A}{P} \arrow[r,"\jvcomp{P}{f_0}{f_1}"] \arrow[d,swap,"\cprojfstf{A}{P}"] & \ctxext{B}{Q} \arrow[d,"\cprojfstf{B}{Q}"] \\
-A \arrow[r,swap,"f_0"] & B
-\end{tikzcd}
-\end{equation*}
-commutes. Also, we have
-\begin{equation*}
-S_{\tmext{x}{u}}(\jvcomp{P}{f_0}{f_1}) \jdeq \tmext{S_x(f_0)}{S_{\tmext{x}{u}}(f_1)}
-\end{equation*}
-for each $x\in T(A)$ and $u\in T(S_x(P))$.
-\end{rmk}
-
-\section{Natural transformations of E-homomorphisms}
-\begin{rmk}
-When $\mathcal{E}$ is an E-system with empty context, then we have the category
-structure on the contexts given by $\thom{\Gamma}{\Delta}= T(W_\Gamma(\Delta))$.
-\end{rmk}
-
-\begin{defn}
-A \emph{natural transformation} of E-homomorphisms from $F:\mathcal{E}\to\mathcal{D}$
-to $G:\mathcal{E}\to\mathcal{D}$ preserving the empty context consists of
-\begin{enumerate}
-\item A t-morphism $\tau0_A\in\thom{F(\Gamma)}{G(\Gamma)}$ for every $\Gamma\in\mathcal{E}$.
-\item A commuting square square
-\begin{equation*}
-\begin{tikzcd}
-F(A) \arrow[r,"\tau1_A"] \arrow[d,twoheadrightarrow] & G(A) \arrow[d,twoheadrightarrow] \\
-F(\Gamma) \arrow[r,swap,"\tau0_\Gamma"] & G(\Gamma)
-\end{tikzcd}
-\end{equation*}
-for every family $P$ over $A$ of solid arrows, such that the diagram
-\begin{equation*}
-\begin{tikzcd}
-F(A) \arrow[r,"\tau1_A"] & G(A) \\
-F(\Gamma) \arrow[r,swap,"\tau0_\Gamma"] \arrow[densely dotted,u,"F(x)"] & G(\Gamma) \arrow[densely dotted,u,swap,"G(x)"]
-\end{tikzcd}
-\end{equation*}
-commutes for every $x\in T(A)$.
-\end{enumerate}
-such that
-\begin{enumerate}
-\item the square
-\begin{equation*}
-\begin{tikzcd}[column sep=huge]
-W_{F(A)}(F(B)) \arrow[r,"W_{F(A)}(\tau1_B)"] \arrow[d,->>] & W_{G(A)}(G(B)) \arrow[d,->>] \\
-F(A) \arrow[r,"\tau1_A"] & G(A)
-\end{tikzcd}
-\end{equation*}
-commutes for every $A,B\in\mathcal{E}/\Gamma$. Note that 
-\begin{align*}
-W_{F(A)}(\tau1_B)
-  & \in 
-T(W_{F(A)}(W_{F(B)}(G(B))))
-  \\
-  & =
-T(W_{W_{F(A)}(F(B))}(W_{F(A)}(G(B))))
-\end{align*}
-\end{enumerate}
-\end{defn}
-
-\section{Indexed E-systems}
-\begin{defn}
-Let $\cat{A}$ be a category. An $\cat{A}$-indexed E-system is an internal
-E-system in the category $Psh(\cat{A})$.
-\end{defn}
-
-\section{Examples of adjunctions}
-
-\begin{lem}
-For each $A\in\mathcal{E}/\Gamma$, $W_A$ is left adjoint to $\Pi_A$. 
-\end{lem}
-
-\begin{proof}
-The unit $\eta:1_{\cat{C}/\ctxext{\Gamma}{A}}\Rightarrow W_A\circ\Pi_A$ is given by
-\begin{equation*}
-\eta_P:\lam{f}{x} ev(\lambda(f),x)
-\end{equation*}
-$\eta_P\in\thom{P}{W_A(\Pi_A(P))}$
-\end{proof}
-
-\begin{lem}
-Suppose $\cat{D}$ is a retract of a pre-category $\cat{C}$, i.e. there are functors
-$S:\cat{D}\to\cat{C}$ and $R:\cat{C}\to\cat{D}$ with a natural isomorphism
-$\tau:R\circ S \simeq \catid{\cat{D}}$, and suppose that $\cat{C}$ is a weakening
-category. Then $\cat{D}$ is a weakening system too.
-\end{lem}
-
-\begin{proof}
-Let $f:\Delta\to \Gamma$ be a morphism in $\cat{D}$. Then we define $W_f$ to be the
-composition
-\begin{equation*}
-\begin{tikzcd}[column sep=3.2em]
-\cat{D}/\Gamma
-  \arrow[r,"S/\Gamma"]
-  &
-\cat{C}/S(\Gamma)
-  \arrow[r,"W_{S(f)}"]
-  &
-\cat{C}/S(\Delta)
-  \arrow[r,"R/S(\Delta)"]
-  &
-\cat{D}/R(S(\Delta))
-  \arrow[r,"\tau_\Delta\circ{-}"]
-  &
-\cat{D}/\Delta
-\end{tikzcd}
-\end{equation*}
-We first verify whether this assembles a pre-weakening structure on $\cat{D}$.
-\begin{enumerate}
-\item Let $\Delta$ be an object of $\cat{D}$. Then $W_{\catid{\Delta}}$ is the composition
-\begin{equation*}
-\begin{tikzcd}[column sep=3.2em]
-\cat{D}/\Delta
-  \arrow[r,"S/\Delta"]
-  &
-\cat{C}/S(\Delta)
-  \arrow[r,"R/S(\Delta)"]
-  &
-\cat{D}/R(S(\Delta))
-  \arrow[r,"{-}\circ\tau_\Delta"]
-  &
-\cat{D}/\Delta
-\end{tikzcd}
-\end{equation*}
-This functor maps $g:\Delta'\to \Delta$ consecutively to $S(g)$, to $R(S(g))$, to
-$\tau_\Delta\circ R(S(g))$. Note that $\tau_\Delta\circ R(S(g))=g\circ \tau_{\Delta'}$ by
-naturality.
-\end{enumerate}
-\end{proof}
-
-\section{Weak E-systems}
-
-\begin{rmk}
-Note that $W$ is a strict $2$-functor frm $\cat{C}$ to $\mathbb{Cat}$ of which
-the value at an object $\Gamma$ is specified to be $\cat{C}/\Gamma$, and which
-is required to preserve the final object. We could have considered $W$ to be a 
-pseudo-functor from $\cat{C}$ to $\mathbb{Cat}$ instead. A coherence theorem
-by Power \cite{Power89} then allows for the strictification of this pseudo-functor
-into a (strict) weakening structure. (Does this map $\Gamma$ to $\cat{C}/\Gamma$
-and final to final?). Thus, we know what a weak pre-weakening system has to be.
-The following questions need to be answered:
-\begin{itemize}
-\item What is a weak weakening/projection/substitution system.
-\item What is a weak E-system? I.e.~what are the coherence rules for something
-to be a weakening/projection/substitution homomorphism?
-\item Can they all be strictified? Are there more examples of weak E-systems?
-\end{itemize}
-\end{rmk}
-\end{comment}
-
-\section{E-systems compared to other systems of type dependency}\label{sec:esys_compared}
-In an effort to formalize the simplicial set model of Martin-L\"of type theory
-with the univalence axiom, Voevodsky has studied various notions of models of
-type theory, taking Cartmell's \emph{contextual categories} as a starting
-point. Voevodsky's definition of a \emph{C-system} is equivalent to that of Cartmell
-\cite{VV_C-systems_quotients}. There are several techniques of defining 
-C-systems: in \cite{VV_Csys_univ}, C-systems are constructed out of a universe
-in a category; in \cite{VV_C-systems_quotients}, sub-C-systems and regular
-quotients of C-systems are defined; in \cite{VV_C-systems_monad}, a C-system
-is constructed out of a monad $R$ on the category of sets and a left $R$-module
-with values in $\mathbf{Set}$.
-
-In \cite{VV_B-systems}, Voevodsky introduced his notion of \emph{B-systems}, which is
-an essentially algebraic theory of type dependency with infinitely many sorts 
-and operations, and subsequently he showed that there is a full and faithful
-functor from the category of C-systems to the category of B0-systems, of which
-the image is contained in the subcategory of B-systems. This fact is important at least
-for a theoretical reason: it tells us that the conditions of being a B-system
-are necessary to axiomatize type dependency. In other words,
-this is a theorem that tells us that we need no further axioms. It is
-unknown whether the categories of C-systems and B-systems are actually equivalent.
-This question would settle the sufficiency of the conditions of being a B-system.
-
-The category of B-systems will be shown to be equivalent to the category of
-stratified E-systems. However, we do not arrive at the fact that the functor
-$\mathcal{E}:\mathbf{Bsys}\to\mathbf{Esys}$ is full without imposing an additional
-condition on the E-homomorphisms. 
-To ensure that the category of stratified E-systems with stratified 
-E-homomorphisms between them is a full subcategory of the category 
-$\mathbf{Esys}$, we will require that the underlying functor of each E-homomorphism
-lifts factorizations. It is
-important to note that this condition is also algebraic, so that the category
-of E-systems remains algebraic.
-
-\subsection{Lifting factorizations}
-Recall
-that for any morphism $f:X\to Y$ in a category $\cat{C}$ there is a category
-$\mathbf{fact}(f)$ of factorizations of $f$. Also, any functor $F:\cat{C}\to\cat{D}$
-determines a functor $F_{\mathbf{fact}(f)}:\mathbf{fact}(f)\to\mathbf{fact}(H(f))$
-for any morphism $f$ of $\cat{C}$.
-
-\begin{defn}
-A functor $F:\cat{C}\to\cat{D}$ is said to
-\emph{lift factorizations} if for any $f\in\cat{C}/X$, the functor
-$F_{\mathbf{fact}(f)}:\mathbf{fact}(f)\to\mathbf{fact}(F(f))$ is an isomorphism
-of categories. 
-\end{defn}
-
-\begin{comment}
-\begin{rmk}
-We may choose the property of lifting factorizations
-to involve either an equivalence or an isomorphism of categories. The
-isomorphism-version, would say that for every factorization $h'\circ g'$
-of $F(f)$ there is a unique factorization $f\jdeq h\circ g$ in $\cat{C}$ such
-that $F(g)\jdeq g'$ and $F(h)\jdeq h'$. In the version with equivalences,
-the uniqueness is replaced by uniqueness up to isomorphism, and the equalites
-are replaced by an isomorphism in $\mathbf{fact}(F(f))$.
-
-It is easier to state the isomorphism version of lifting factorizations with
-inference rules and it might be easier to explain this condition on type theoretical
-grounds, although the equivalence version has the advantage of being
-categorical (i.e.~invariant under equivalence of categories).
-
-In the present
-context, it doesn't matter very much which one we pick: the categories in which
-we're interested are all posets.
-\end{rmk}
-\end{comment}
-
-\begin{eg}
-For any slice category $\cat{C}/X$, the forgetful functor $\cat{C}/X\to\cat{C}$
-lifts factorizations.
-\end{eg}
-
-\begin{defn}
-We write $\mathbf{Esys}$ for the category of E-systems where the underlying
-functor of each E-homomorphism lifts factorizations.
-\end{defn}
-
-\subsection{Stratified E-systems}
-
-In this subsection we define the notion of stratified E-system. Then we show
-that the category $\mathbf{Esys_s}$ of stratified E-systems with homomorphisms preserving the
-stratification, is a full subcategory of $\mathbf{Esys}$.
-
-\begin{defn}
-A category $\cat{C}$ with terminal object is said to be \emph{stratified} if there exists a 
-\emph{stratification functor}
-\begin{equation*}
-L : \cat{C}\to (\mathbb{N},\geq)
-\end{equation*}
-such that
-\begin{enumerate}
-\item $L(X)\jdeq 0$ if and only if $X$ is terminal, and for any $f:X\to Y$ we have
-$L(X)\jdeq L(Y)$ if and only if $X\jdeq Y$ and $f\jdeq\catid{X}$. 
-\item every morphism $f:X\to Y$ in $\cat{C}_\mathbb{E}$, where $L(X)\jdeq
-n+m+1$ and $L(Y)\jdeq n$, has a unique factorization 
-\begin{equation*}
-\begin{tikzcd}
-X \arrow[r,"f_m"] & X_m \arrow[r,"f_{m-1}"] & \cdots \arrow[r,"f_1"] & X_1 \arrow[r,"f_0"] & Y
-\end{tikzcd}
-\end{equation*}
-where $L(X_i)\jdeq n+i$.
-\end{enumerate}
-A functor $F:\cat{C}\to\cat{D}$ between stratified categories is said to be stratified 
-if $L_{\cat{C}}\jdeq L_{\cat{D}}\circ F$. 
-\end{defn}
-
-%\begin{rmk}
-%The categorical version of being stratified, is that there exists a factofibration
-%$L:\cat{C}\to (N,\geq)$ which preserves the terminal object.
-%\end{rmk}
-
-\begin{lem}
-Any stratified category is a rooted tree.
-\end{lem}
-
-\begin{proof}
-Let $\cat{C}$ be a stratified category. The root of the tree is going to be the terminal
-object of $\cat{C}$. Note that objects of $\cat{C}$ have no automorphisms other
-than the identity morphisms. Therefore, to show that $\cat{C}$ is a tree,
-it suffices to show that for every object $X$ with $L(X)>0$ there is
-a unique object $\eft(X)$ with $L(\eft(X))\jdeq L(X)-1$, with a morphism $X\to \eft(X)$.
-The existence follows from the existence of the factorization of $X\to 1$. The 
-uniqueness follows from the observation that for any arrow
-$X\to Y$ with $L(Y)\jdeq L(X)-1$, we can factorize $X\to 1$ through $Y$. Since
-factorizations are unique, we will have $Y\jdeq \eft(X)$. 
-\end{proof}
-
-\begin{lem}
-The category of small stratified categories is a full subcategory of the category
-of small categories with terminal objects and between them functors which lift 
-factorizations and preserve the terminal object.
-\end{lem}
-
-\begin{proof}
-Let $F:\cat{C}\to\cat{D}$ be a functor between stratified categories which lifts
-factorizations and 
-which preserves terminal objects. We show by induction that $F$ is stratified.
-Since $F$ preserves the terminal object, $L(Y)\jdeq 0$ implies $L(F(Y))\jdeq 0$.
-Now suppose that for $n\in\mathbb{N}$, we have that $L(\Gamma)\jdeq n$ implies
-$L(F(Y))\jdeq n$, and let $X$ be such that $L(X)\jdeq n+1$ with
-$p_X:X\to\eft(X)$. Then the poset $\mathbf{fact}(p_X)$ is isomorphic to the
-poset $\mathbf{2}$ with two objects, one of which is smaller than the other. By the
-assumption that $F$ lifts factorizations, the poset $\mathbf{fact}(F(p_X))$
-is also isomorphic to $\mathbf{2}$. Thus, the only factorization
-of $F(X)\to\unit$ starts with $F(p_X):H(X)\to F(\eft(X))$, which implies that
-$L(F(X))\jdeq n+1$. 
-\end{proof}
-
-\begin{rmk}
-It is easy to show that a category can only be stratified in at most one way,
-and that if a category is stratified, then so are any of its slices. Therefore,
-it makes sense to ask for the property that the weakening and substitution
-functors in a pre-weakening or pre-substitution system are stratified.
-\end{rmk}
-
-\begin{defn}
-A pre-E-system is said to be \emph{stratified} if its underlying category is
-stratified and if each $W_A$ and $S_x$ is a stratified functor. The category of
-stratified pre-E-systems with stratified E-homomorphisms between them is denoted by
-$\mathbf{Esys_s}$.
-\end{defn}
-
-\begin{cor}\label{lem:strat_full}
-The category of stratified pre-E-systems is a full subcategory of the category of
-pre-E-systems.
-\end{cor}
-
-\subsection{The embedding of B0-systems into pre-E-systems}
-
-Voevodsky distinguishes between pre-B-systems, which is an essentially algebraic theory with
-only operations and no equalities; B0-systems, which is an extension of the
-essentially algebraic theory of pre-B-systems which ensures that all the operations behave
-well with two sets of the operations ($\eft$ and $\ebd$); and B-systems, which extends the 
-theory of B0-systems with equalities ensuring that the rest of the operations 
-are compatible with each other. Voevodsky, furthermore distinguishes
-non-unital and unital versions of these, and noted that not all B-homomorphisms
-between unital B-systems are unital. The same distinguishment could be made for 
-E-systems by forgetting the projection structure, but we have not been able to construct a
-functor from unital B0-systems to unital pre-E-systems. Therefore, we shall only
-consider non-unital B0-systems and unital B-systems. In this section we will
-recall the definition of non-unital B0-systems and construct a full and faithful
-functor from them to the category of pre-E-systems.
-
-One reason for introducing B-systems by first introducing pre-B-systems is that
-it is clear in the definition of pre-B-systems that there are only countably
-many operations. Since we are interested in embedding the B-systems into E-systems,
-we shall organize the definition of B-systems a bit differently, but of course
-the resulting definition of B0-systems and B-systems will be (trivially) 
-equivalent.
-
-\begin{defn}
-A \emph{B-framework} is a collection of data of the following form:
-\begin{enumerate}
-\item for all $n\in\mathbb{N}$ two sets $B_n$ and $\tilde{B}_n$. 
-\item for all $n\in\mathbb{N}$ maps of the form
-\begin{align*}
-\eft[n] & : B_{n+1}\to B_n \\
-\ebd[n] & : \tilde{B}_n\to B_n.
-\end{align*}
-For $m,n\in\mathbb{N}$, we denote the composition $\eft[n]\circ\cdots\circ\eft[n+m]:B_{n+m+1}\to B_n$ by $\eft[n]^m$. 
-\item $B_0$ is a singleton $\{\pt\}$.
-\end{enumerate} 
-\end{defn}
-
-\begin{rmk}
-Since the indices $n,m\in\mathbb{N}$ can be infered, we will usually omit them
-in the infix form of the operators.
-\end{rmk}
-
-\begin{defn}
-A homomorphism $H:\mathbb{B}\to\mathbb{A}$ of B-frameworks consists of maps
-$H_n:B_n\to A_n$ and $\tilde{H}_n:\tilde{B}_n\to\tilde{A}_n$ such that
-\begin{align*}
-\eft(H(X)) & \jdeq H(\eft(X)) \\
-\ebd(\tilde{H}(x)) & \jdeq H(\ebd(x))
-\end{align*}
-for any $X\in B_n$ and $x\in\tilde{B}_n$. The category of B-frameworks is
-denoted by $\mathbf{Bfr}$. 
-\end{defn}
-
-\begin{defn}
-For every B-framework $\mathbb{B}$ and any $X\in B_n$, there is a B-framework
-$\mathbb{B}/X$ given by
-\begin{align*}
-(B/X)_{m} & \jdeq \{Y\in B_{n+m}\mid\eft^{m}(Y)\jdeq X\}\\
-(\tilde{B}/X)_m & \jdeq \{y\in \tilde{B}_{n+m}\mid\eft^m(\ebd(y))\jdeq X\}.
-\end{align*}
-Also, for any homomorphism $H:\mathbb{B}\to\mathbb{A}$ of B-frameworks and any
-$X\in B_n$, there is a homomorphism $H/X:\mathbb{B}/X\to\mathbb{A}/H(X)$
-defined in the obvious way.
-\end{defn}
-
-\begin{rmk}
-Note that for $X\in B_n$ and $Y\in B_{n+m}$ such that $\eft^m(Y)\jdeq X$, 
-we have $(\mathbb{B}/X)/Y\cong B/Y$.
-\end{rmk}
-
-\begin{defn}
-A B0-system $\mathbb{B}$ consists of a B-framework $\mathbb{B}$ and homomorphisms
-\begin{align*}
-W_{X} & : \mathbb{B}/\eft(X)\to\mathbb{B}/X\\
-S_{x} & : \mathbb{B}/\ebd(x)\to\mathbb{B}/\eft(\ebd(x))
-\end{align*}
-of B-frameworks, for any $X\in B_{n+1}$ and any $x\in\tilde{B}_{n+1}$. 
-\end{defn}
-
-\begin{lem}
-When $\mathbb{B}$ is a B0-system, then so is each $\mathbb{B}/X$.
-\end{lem}
-
-\begin{proof}
-Straightforward.
-\end{proof}
-
-\begin{defn}
-A B0-homomorphism $H:\mathbb{B}\to\mathbb{A}$ is a homomorphism of B-frameworks
-for which the diagrams
-\begin{equation*}
-\begin{tikzcd}[column sep=huge]
-\mathbb{B}/X \arrow[r,"H/X"] & \mathbb{A}/H(X) \\
-\mathbb{B}/\eft(X) \arrow[u,"W_X"] \arrow[r,swap,"H/\eft(X)"] & \mathbb{A}/\eft(H(X)) \arrow[u,swap,"W_{H(X)}"]
-\end{tikzcd}
-\end{equation*}
-and
-\begin{equation*}
-\begin{tikzcd}[column sep=huge]
-\mathbb{B}/\ebd(x) \arrow[r,"H/\ebd(x)"] \arrow[d,swap,"S_x"] & \mathbb{A}/\ebd(H(X)) \arrow[d,"S_{\tilde{H}(x)}"] \\
-\mathbb{B}/\eft(\ebd(x)) \arrow[r,swap,"H/\eft(\ebd(x))"] & \mathbb{A}/\eft(\ebd(H(X)))
-\end{tikzcd}
-\end{equation*}
-of homomorphisms of B-frameworks commute for every 
-$X\in B_n$ and $x\in\tilde{B}_n$. The category of B0-systems is denoted
-by $\mathbf{B0sys}$.
-\end{defn}
-
-Now that we have defined the category of B0systems, we can start defining the
-embedding into pre-E-systems.
-
-\begin{defn}
-Let $\mathbb{B}$ be a B-framework. Then $\mathbb{B}$ determines
-a stratified category $U(\mathbb{B})$. Also, any homomorphism of B-frameworks
-determines a stratified functor.
-\end{defn}
-
-\begin{constr}
-The set of objects of $U(\mathbb{B})$ is taken to be the set
-$\bigsqcup_{(n\in\mathbb{N})}B_n$. Now we define
-$(k,X)\leq (n,Y)$ if and only if $n\leq k$ and $\eft[n]^{k-n}(X)\jdeq Y$.
-In other words, we only have $(n+m,X)\leq (n,\eft^m(X))$. It is immediate from
-this definition that for any $X\in B_{n+1}$ there is a unique $\Gamma\in
-B_n$ such that $(n+1,X)\leq (n,\Gamma)$, so $U(\mathbb{B})$ is stratified.
-
-To show that the functor $U(H):U(\mathbb{B})\to U(\mathbb{A})$ is stratified for
-any homomorphism $H:\mathbb{B}\to\mathbb{A}$ of B-frameworks, note that
-by \autoref{lem:strat_full} it suffices to show that each $U(H)$ 
-lifts factorizations.
-Let $H:\mathbb{B}\to\mathbb{A}$ be a homomorphism of B-frameworks, and let
-$(m,X):(n+m,X)\leq (n,\Gamma)$. Then the category $\mathbf{fact}_{(m,X)}$ of
-factorizations of $(m,X)$ is a finite linear order with
-$m+1$ elements, and so is the category $\mathbf{fact}_{(m,H(X))}$. 
-\end{constr}
-
-\begin{rmk} 
-It is in fact the case that $U(\mathbb{B})$ is the free category generated by the
-graph with $\bigsqcup_{(n\in\mathbb{N})}B_n$ as the set of vertices, and
-for each object of the form $(n+1,X)$ an edge to $(n,\eft[n](X))$. 
-Essentially all $U$ does, is forgetting about the term structure 
-$\tilde{B}$ of $\mathbb{B}$.
-
-Furthermore, it is useful to note that for any $X\in B_n$, we get an 
-isomorphism $U(\mathbb{B}/X)\cong  U(\mathbb{B})/(n,X)$ of rooted trees,
-natural in $X$. Therefore we will usually not distinguish between $U(\mathbb{B})/X$
-and $U(\mathbb{B})/(n,X)$. 
-\end{rmk}
-
-\begin{defn}
-Let $\mathbb{B}$ be a B0-system. Then $U(\mathbb{B})$ can be given the structure
-of a pre-E-system. We write $\mathcal{E}(\mathbb{B})$ for the resulting
-pre-E-system. Thus we get a functor $\mathcal{E}:\mathbf{B0sys}\to\mathbf{Esys_s}$.
-\end{defn}
-
-\begin{constr}
-We begin by simultaneously defining the term structure and the pre-substitution
-structure on $U(\mathbb{B})$ by induction on the natural numbers. More precisely,
-for any $m\in\mathbb{N}$ we will define, for any $(n+m,X)\leq (n,\Gamma)$ a set
-$T(m,X)$ and for any $x\in T(m,X)$ a homomorphism 
-$S_x:\mathbb{B}/X\to\mathbb{B}/\Gamma$ of B-frameworks. Then it follows
-immediately that we get stratified functors $U(S_x):U(\mathbb{B})/X\to U(\mathbb{B})/\Gamma$
-defining the pre-substitution structure of $\mathcal{E}(\mathbb{B})$.
-
-For $(n,X)\in U(\mathbb{B})$, we define $T(0,X)\defeq
-\{\pt[X]\}$ and $S_{\pt[X]}\defeq \catid{\mathbb{B}/X}$. We also define for
-$(n+1,X)\leq (n,\Gamma)$ the set $T(1,X)\jdeq\{x\in B_{n+1}\mid \ebd(x)\jdeq X\}$.
-Note that for $x\in T(1,X)$, the homomorphism $S_x:\mathbb{B}/X \to\mathbb{B}/\Gamma$
-is already assumed to exist, because $\mathbb{B}$ is a B0-system.
-
-For the inductive step, suppose that for $m\in\mathbb{N}$ we have sets $T(m,X)$
-for any $(n+m,X)\leq (n,\Gamma)$ and homomorphisms $S_x:\mathbb{B}/X \to \mathbb{B}/\Gamma$
-of B-frameworks, for any $x\in T(m,X)$. Then we define
-\begin{equation*}
-T(m+1,X)\defeq\bigsqcup_{x\in T(m,\eft(X))} T(1,S_x(X)).
-\end{equation*}
-For $(x,u)\in T(m+1,X)$ we define the homomorphism $S_{(x,u)}:\mathbb{B}/X\to
-\mathbb{B}/\Gamma$ as
-\begin{equation*}
-S_{(x,u)}\defeq S_u\circ S_x/X.
-\end{equation*}
-It remains to define a pre-weakening structure on $U(\mathbb{B})$. To do this,
-it suffices to define homomorphisms $W_{(m,X)}:\mathbb{B}/\Gamma \to\mathbb{B}/X$
-of B-frameworks, for any $(n+m,X)\leq (n,\Gamma)$. We define
-\begin{equation*}
-W_{(m,X)}\defeq W_X\circ\cdots\circ W_{\eft^{m-1}(X)}.\qedhere
-\end{equation*}
-\end{constr}
-
-\begin{thm}
-The functor $\mathcal{E}$ is full and faithful.
-\end{thm}
-
-\begin{proof}
-We have to show that the map $\mathrm{hom}(\mathbb{B},\mathbb{A})\to
-\mathrm{hom}(\mathcal{E}(\mathbb{B}),\mathcal{E}(\mathbb{A}))$ is a bijection
-for any two B0-systems $\mathbb{B}$ and $\mathbb{A}$.
-
-To show that this map is injective, suppose $F,G:\mathbb{B}\to\mathbb{A}$ are 
-B0-homomorphisms such that $\mathcal{E}(F)\jdeq\mathcal{E}(G)$. Then we have
-$F_n(X)\jdeq\mathcal{E}(F)(n,X)\jdeq \mathcal{E}(G)(n,X)\jdeq G_n(X)$ for each
-object $x\in B_n$. Also, we have $\tilde{F}_n(x)\jdeq T_{\mathcal{E}(F)}(n,x)
-\jdeq T_{\mathcal{E}(G)}(n,x)\jdeq \tilde{G}_n(x)$ for each
-$x\in\tilde{B}_n$. This shows that $F\jdeq G$, so we conclude that $\mathcal{E}$
-is faithful.
-
-To show that the map $\mathrm{hom}(\mathbb{B},\mathbb{A})\to
-\mathrm{hom}(\mathcal{E}(\mathbb{B}),\mathcal{E}(\mathbb{A}))$ 
-is surjective, let $H:\mathcal{E}(\mathbb{B})\to
-\mathcal{E}(\mathbb{A})$. Since both $\mathcal{E}(\mathbb{B})$ and
-$\mathcal{E}(\mathbb{A})$ are stratified, we get from
-\autoref{lem:strat_full} that $H$ is a stratified pre-E-homomorphism. 
-Because $H$ is stratified, we get maps $H_n:B_n\to A_n$ for any $n\in\mathbb{N}$.
-The term structure of $H$ gives maps $\tilde{H}_n:
-\tilde{B}_n\to\tilde{A}_n$. This defines a B0-homomorphism since $H$ 
-preserves the pre-weakening and pre-substitution structures of $\mathcal{E}(\mathbb{B})$.
-It is immediate from the definition of $H_n$ and
-$\tilde{H}_n$ that this B0-homomorphism is mapped back to $H$, so we conclude
-that $\mathcal{E}$ is full.
-\end{proof}
-
-\subsection{B-systems are stratified E-systems}
-There are two properties which hold for pre-E-systems of the form $\mathcal{E}
-(\mathbb{B})$ for some B0-system $\mathbb{B}$, 
-which stand in the way of proving that the category of stratified
-pre-E-systems is equivalent to the category of B0-systems. The first is that
-for any $A\in U(\mathbb{B})/\Gamma$ and $P\in U(\mathbb{B})/\ctxext{\Gamma}{A}$,
-we have an isomorphism $T(\ctxext{A}{P})\cong\bigsqcup_{x\in T(A)}T(S_x(P))$.
-The second, which depends on the first, is that for any $(x,u)\in T(\ctxext{A}{P})$
-we have $S(x,u)\jdeq S_u\circ S_x/P$. In other words, both the term structure
-and the pre-substitution structure of $\mathcal{E}(\mathbb{B})$ are compatible
-with the categorical structure of the underlying category $U(\mathbb{B})$.
-However, to prove the corresponding facts for E-systems, we needed the full
-set of conditions on unital E-systems: to prove that the term structure of
-$U(\mathbb{B})$ is compatible with composition we needed units, and to prove 
-that the operations $S_x$ are compatible with the pairing function we needed
-that they are E-homomorphisms.
-
-Presumably, we can assume these compatibility conditions on pre-E-systems in order
-to prove the desired equivalence of B0-systems and stratified pre-E-systems. However,
-that seems somewhat unnatural, especially given that they automatically hold
-for E-systems. Thus, for the remainder of this section we set out to show that
-the category of unital B-systems is equivalent to the category of stratified 
-unital E-systems.
-
-\begin{defn}
-A unital B0-system is a B0-system with maps
-\begin{equation*}
-\delta_n : B_{n+1}\to \tilde{B}_{n+2}.
-\end{equation*}
-such that $\ebd(\delta_n(X))\jdeq W_{X}(X)$ for any $X\in B_n$. 
-A B0-homomorphism $H$ is said to be unital if one has
-\begin{equation*}
-H(\delta(X))\jdeq\delta(H(X))
-\end{equation*}
-for any $X\in B_n$.  
-\end{defn}
-
-\begin{rmk}
-Voevodsky showed in \cite{VV_B-systems} that the category of unital B0-systems
-with unital B0-homomorphisms between them is not a full
-subcategory of $\mathbf{B0sys}$. 
-\end{rmk}
-
-\begin{defn}
-A \emph{(unital) B-system} is a B0-system for which the following conditions
-hold:
-\begin{enumerate}
-\item Every $W_X$ and $S_x$ is a unital B0-homomorphism.
-\item For every $x\in \tilde{B}_{n+1}$ one has $S_x\circ W_{\ebd(x)}\jdeq
-\catid{\mathbb{B}/\eft(\ebd(x))}$. 
-\item For every $x\in\tilde{B}_{n+1}$ one has $S_x(\delta(\ebd(x)))\jdeq x$.
-\item For every $X\in B_{n+1}$ one has $S_{\delta(X)}\circ W_X/X\jdeq
-\catid{\mathbb{B}/X}$. 
-\end{enumerate}
-\emph{B-homomorphisms} are unital B0-homomorphisms between B-systems.
-We denote the category of B-systems by $\mathbf{Bsys}$. 
-\end{defn}
-
-\begin{lem}
-For any non-unital B-system $\mathbb{B}$, $\mathcal{E}(\mathbb{B})$ is a 
-stratified non-unital E-system. 
-\end{lem}
-
-\begin{proof}
-Straightforward.
-\end{proof}
-
-\begin{defn}
-For any unital B-system $\mathbb{B}$, $\mathcal{E}(\mathbb{B})$ has a
-projection structure. We thus obtain a functor 
-$\mathcal{E}:\mathbf{Bsys}\to\mathbf{Esys}$ from unital B-systems to E-systems.
-\end{defn}
-
-\begin{constr}
-We first need to extend the functor $\mathcal{E}:\mathbf{B0sys}\to\mathbf{Esys_p}$
-from non-unital B0-systems to non-unital pre-E-systems, to a functor from
-unital B-systems to E-systems.
-
-We need to give $\mathcal{E}(\mathbb{B})$ the structure of a unital
-pre-E-system. We construct by induction on $m\in \mathbb{N}$, a term of type 
-$T(m,W_{(m,X)}(X))$ with the property that that 
-\begin{equation*}
-S_{\tfid{(m,X)}}\circ W_{(m,X)}/X\jdeq\catid{U(\mathbb{B})/X},
-\end{equation*} 
-for every $(n+m,X)\leq (n,\Gamma)$ by induction on $m$. The extra
-condition is part of the inductive construction, because we need it on the way.
-
-For $m\jdeq 0$ we have but one
-choice, and for $m\jdeq 1$ we can directly use the unital structure of
-$\mathbb{B}$. For the inductive step, recall that
-\begin{equation*}
-T(m+1,W_{(m+1,X)}(X))\jdeq\bigsqcup_{x\in T(m,W_{(m+1,X)}(\eft(X)))}T(1,S_x(W_{(m+1,X)}(X))).
-\end{equation*}
-We have the term $x\defeq \tilde{W}_X(\tfid{(m,\eft(X))})\in T(m,W_{(m+1,X)}(\eft(X)))$.
-Now we may observe that 
-\begin{align*}
-S_x(W_{(m+1,X)}(X))
-  & \jdeq 
-S_{\tilde{W}_X(\tfid{(m,\eft(X))})}(W_{(m+1,X)}(X))\\
-  & \jdeq 
-S_{\tilde{W}_X(\tfid{(m,\eft(X))})}(W_X(W_{(m,\eft(X))}(X))) \\
-  & \jdeq
-W_{X}(S_{\tfid{(m,\eft(X))}}(W_{(m,\eft(X))}(X))) \\
-  & \jdeq 
-W_{X}(X).
-\end{align*}
-Therefore, we see that $T(1,S_x(W_{(m+1,X)}(X)))\jdeq T(1,W_X(X))$, where we can
-choose the term $\delta(X)$. Thus, we define
-\begin{equation*}
-\tfid{(m+1,X)}\defeq (\tilde{W}_X(\tfid{(m,\eft(X))}),\delta(X))
-\end{equation*} 
-for $m\geq 1$. Now we need to check that $S_{\tfid{(m+1,X)}}\circ 
-W_{(m+1,X)}/X\jdeq \catid{U(\mathbb{B})/X}$. This is indeed the case:
-\begin{align*}
-S_{\tfid{(m+1,X)}}\circ W_{(m+1,X)}/X
-  & \jdeq
-S_{\delta(X)}\circ S_{\tilde{W}_X(\tfid{(m,\eft(X))})}\circ W_{(m+1,X)}/X \\
-  & \jdeq
-S_{\delta(X)}\circ W_X/X \\
-  & \jdeq
-\catid{U(\mathbb{B})/X}.
-\end{align*}
-This finishes the definition of the pre-projection structure on $\mathcal{E}(\mathbb{B})$.
-Straightforward proves by induction give that each $W_{(m,X)}$ and $S_x$ is a 
-pre-projection homomorphism and that $S_x(\tfid{(m,X)})\jdeq x$, for any 
-$x\in T(m,X)$.
-\end{constr}
-
-\begin{thm}
-The functor $\mathcal{E}:\mathbf{Bsys}\to\mathbf{Esys}$ lifts to an equivalence
-to the subcategory $\mathbf{Esys_s}$ of stratified E-systems: i.e.~we have a
-commuting diagram
-\begin{equation*}
-\begin{tikzcd}
-& \mathbf{Esys_s} \arrow[dr,"\subseteq"]\\
-\mathbf{Bsys} \arrow[ur,"\simeq"] \arrow[rr,swap,"\mathcal{E}"] & & \mathbf{Esys}
-\end{tikzcd}
-\end{equation*}
-\end{thm}
-
-\begin{proof}
-We already have shown that $\mathcal{E}$ factors through the stratified
-E-systems and that it is full and faithful, so it remains
-to show that every stratified E-system is in the image of $\mathcal{E}$. In
-other words, for every stratified E-system $\mathbb{E}$ we need to construct
-a B0-system $\mathbb{B}$ such that $\mathcal{E}(\mathbb{B})\cong\mathbb{E}$.
-
-Let $\mathbb{E}$ be a stratified E-system. From the stratification functor
-$L:\mathbb{C}_{\mathbb{E}}\to (\mathbb{N},\geq)$ we obtain the pre-image
-presheaf $B(\mathbb{E}):\op{(\mathbb{N},\leq)}\to\mathbf{Set}$ given by
-\begin{align*}
-B(\mathbb{E})_n & \defeq \{X\in\mathrm{Ob}(\mathbb{C}_{\mathbb{E}})\mid L(X)\jdeq n\}
-\end{align*}
-and $\eft(X)\defeq\mathrm{cod}(p_X)$. We define the term structure of $B(\mathbb{E})_n$
-to be given by
-\begin{equation*}
-\tilde{B}(\mathbb{E})_n \defeq \bigsqcup_{X\in B(\mathbb{E})_n} T(p_X)
-\end{equation*}
-with the projection map $\ebd[n]\defeq\pi_1:\tilde{B}(\mathbb{E})_n\to B(\mathbb{E})_n$.
-This defines the B-framework $B(\mathbb{E})$, and we obtain a functor
-$B:\mathbf{Esys_s}\to\mathbf{B0sys}$. It is immediate that we get isomorphisms
-$U(B(\mathbb{E}))\cong\cat{C}_{\mathbb{E}}$, natural in $\mathbb{E}$. Also, we
-get natural isomorphisms $B(\mathbb{E}/\Gamma)\cong B(\mathbb{E})/\Gamma$.
-
-The weakening and substitution structures on
-$B(\mathbb{E})$ are given by $W_X\defeq B(W_{p_X})$ and $S_{(X,x)}\defeq B(S_x)$,
-respectively. Since these functors are assumed to be stratified, they define
-homomorphisms of B-frameworks. Because a weakening system of an E-system is
-required to be functorial, i.e.~$W_{\ctxext{A}{P}}\jdeq W_P\circ W_A$ and
-$W_{\catid{\Gamma}}\jdeq\catid{\cat{C}_{\mathbb{E}}/\Gamma}$, it follows
-immediately that the weakening structure on $U(B(\mathbb{E}))$ is isomorphic to
-that of $\mathbb{E}$. 
-\end{proof}
-
-\begin{comment}
-In the following definition $\Delta(X)$ denotes the discrete category on the set
-$X$ and $\mathbf{2}$ is the category $s\to t$.
-\begin{defn}
-Define $\mathbb{S}$ to be the pushout
-\begin{equation*}
-\begin{tikzcd}[column sep=large]
-\Delta(\mathbb{N}) \arrow[r,"{n\mapsto(n,s)}"] \arrow[d,swap,"i"] & \Delta(\mathbb{N})\times\mathbf{2} \arrow[d] \\
-(\mathbb{N},\leq) \arrow[r] & \mathbb{S}
-\end{tikzcd}
-\end{equation*}
-in $\mathbf{Cat}$.
-\end{defn}
-
-Note that for any $n\in\mathbb{N}$, we have the map $S^n := (m\mapsto m+n)$ which
-is monotone. By the universal property of pushouts, we obtain a functor $\sigma^n:\mathbb{S}\to\mathbb{S}$
-given by
-\begin{equation*}
-\begin{tikzcd}[column sep=huge]
-\Delta(\mathbb{N}) 
-  \arrow[r,"{n\mapsto(n,s)}"] 
-  \arrow[d,swap,"i"] 
-  & 
-\Delta(\mathbb{N})\times\mathbf{2} 
-  \arrow[d]
-  \arrow[r,"{\Delta(S^n)\times\catid{\mathbf{2}}}"]
-  &
-\Delta(\mathbb{N})\times\mathbf{2}
-  \arrow[dd]
-  \\
-(\mathbb{N},\leq) 
-  \arrow[r] 
-  \arrow[d,swap,"{(S^n,\leq)}"]
-  & 
-\mathbb{S}
-  \arrow[dr,densely dotted,yshift=.5ex,"\sigma^n"]
-  \\
-(\mathbb{N},\leq)
-  \arrow[rr] & &
-\mathbb{S}
-\end{tikzcd}
-\end{equation*}
-
-\begin{defn}
-A non-unital pre-B0-system is a presheaf $B$ on $\mathbb{S}$ such that $B(0)$ is
-a singleton, and the following additional structure:
-\begin{enumerate}
-\item For any $n\in\mathbb{N}$ and $X\in B(n)$, write $B/X$ for the sub-presheaf
-of $B\circ\sigma^n$ given by $B/X(0)=\{X\}$ and $B/X(n+1)=B_{n\leq n+1}^{-1}(B/X(n))$. 
-We require that for any $n\in\mathbb{N}$ and any $A\in B(n+1)$, there is a 
-natural transformation $W_A:B/(B_{n\leq n+1}(A))\Rightarrow B/A$.
-\end{enumerate}
-\end{defn}
-\end{comment}
-
-\subsection{From D-systems to E-systems}
-
-We use `D-system' to refer to Dybjer's categories with families. We recall
-Dybjer's definition. Note that $\mathbf{Fam}$ is the category of families of
-sets, of which the objects consist of a set $A$ and a family $\{B(x)\}_{x\in A}$
-over $A$. Although a terminal object is assumed to exist, not every object 
-of a D-system is `fibrant', i.e.~it is
-not the case that every $X\to 1$ is a composition of projections.
-We could only hope that there is a functor from D-sytems to E-systems without
-terminal objects. We write $\mathbf{Esys}_{\circ}$ for the category of E-systems
-without terminal objects.
-
-\begin{defn}
-A category is said to be locally stratified if each of its slice categories
-are stratified.
-\end{defn}
-
-\begin{defn}
-A \emph{setup for categories with families} consists of
-\begin{enumerate}
-\item A category $\cat{C}$ with a terminal object $1$.
-\item A presheaf $\mathrm{Ty}:\op{\cat{C}}\to\mathbf{Set}$. We will write
-$A[\gamma]$ for $\mathrm{Ty}(\gamma,A)$. 
-\item A commuting triangle
-\begin{equation*}
-\begin{tikzcd}
-\int_\cat{C}\mathrm{Ty} \arrow[dr,swap,"\pi"] \arrow[rr,"p"] & & \cat{C}^{\to} \arrow[dl,"\mathrm{cod}"] \\
-& \cat{C}
-\end{tikzcd}
-\end{equation*} 
-We will write $p_A:\ctxext{\Gamma}{A}\to\Gamma$
-for the morphism $U(\Gamma,A)$ into $\Gamma$, for each object $\Gamma\in\cat{C}$ and
-$A\in\mathrm{Ty}(\Gamma)$.
-\end{enumerate}
-\end{defn}
-
-\begin{defn}
-A \emph{category with families} consists of a setup with
-\begin{enumerate}
-\item[(iv)] a preasheaf $\mathrm{Tm}:\op{(\int_\cat{C}\mathrm{Ty})}\to\mathbf{Set}$,
-\item[(v)] a term $q_A\in\mathrm{Tm}(\ctxext{\Gamma}{A},A[p_A])$ for every
-$A\in\mathrm{Ty}(\Gamma)$, 
-\end{enumerate}
-such that for every $\gamma:\Delta\to\Gamma$ in $\cat{C}$ and 
-$a\in \mathrm{Tm}(\Delta,A[\gamma])$ there is a unique morphism 
-$\theta:\Delta\to \ctxext{\Gamma}{A}$ such that $p_A\circ\theta\jdeq\gamma$ and 
-$q_A[\theta]\jdeq a$.
-\end{defn}
-
-\begin{lem}
-Let $\mathbb{D}$ be a category with families. Then there are bijections
-\begin{equation*}
-\mathrm{Tm}(\Delta,A[\gamma])\cong\{\theta:\Delta\to\ctxext{\Gamma}{A}\mid p_A\circ\theta\jdeq \gamma\}
-\end{equation*}
-natural in $\gamma$ and $A$.
-\end{lem}
-
-\begin{proof}
-Let $\theta:\Delta\to\ctxext{\Gamma}{A}$ be such that $p_A\circ\theta\jdeq\gamma$. 
-Then $\subst{\theta}{q_A}$ is a term
-of $\subst{\theta}{{p_A}{A}}$. By the functoriality of $\mathrm{Ty}$, we have
-$\subst{\theta}{{p_A}{A}}\jdeq \subst{p_A\circ\theta}{A}\jdeq\subst{\gamma}{A}$. 
-So we obtain $\subst{\theta}{q_A}\in\mathrm{Tm}(\Delta,\subst{\gamma}{A})$. Now
-the uniqueness of $\theta$ implies that the map $\theta\mapsto\subst{\theta}{q_A}$
-is a bijection. 
-\end{proof}
-
-The previous lemma renders the presheaf $\mathrm{Tm}$ superfluous. Therefore,
-we define D-systems as follows.
-
-\begin{defn}
-A \emph{D-system} consists of a setup for categories with families
-such that for every $\gamma:\Delta\to\Gamma$ and $A\in\mathrm{Ty}(\Gamma)$, there
-is a bijection
-\begin{equation*}
-\{\eta:\Delta\to\ctxext{\Delta}{\subst{\gamma}{A}}\mid p_{\subst{\gamma}{A}}\circ\eta\jdeq\catid{\Delta}\}
-  \cong
-\{\theta:\Delta\to\ctxext{\Gamma}{A}\mid p_A\circ\theta\jdeq\gamma\}
-\end{equation*}
-natural in $\gamma$ and $A$.
-\end{defn}
-
-\begin{thm}
-A category with families is the same thing as a D-system.
-\end{thm}
-
-\begin{defn}
-There is a functor $U$ from D-systems to locally stratified categories.
-\end{defn}
-
-\begin{constr}
-Let $\mathbb{D}$ be a D-system with underlying category $\cat{C}$. We define
-$U(\mathbb{D})$ to be the category with the same objects as $\cat{C}$, where the
-morphisms are generated by the projections $p_A:\ctxext{\Gamma}{A}\to\Gamma$.
-
-We have to prove that this category is locally stratified. Let $\Gamma\in\cat{C}$. 
-\end{constr}
-
-\begin{defn}
-For any D-system $\mathbb{D}$, $U(\mathbb{D})$ can be given the structure of
-an E-system without terminal object. Thus, we get a functor
-$\mathbf{Dsys}\to\mathbf{Esys}_{\circ}$
-\end{defn}
-
-\begin{constr}
-We first need to define a term structure on $U(\mathbb{D})$. 
-Next, we define a weakening structure on $\cat{C}_p$. Since the morphisms of
-$\cat{C}_p$ are composites of projections, we will define
-\begin{equation*}
-W_{p_{A_n}\circ\cdots\circ p_{A_1}}\defeq W_{p_{A_n}}\circ\cdots W_{p_{A_1}}.
-\end{equation*}
-Therefore, we only have to define functors 
-$W_A:\cat{C}_p/\Gamma\to\cat{C}_p/\ctxext{\Gamma}{A}$ with term structure for every 
-$A\in\mathrm{Ty}(\Gamma)$. Since morphisms are composites of projections, we
-can define this functor by induction. Let $W_A(\catid{\Gamma})\defeq
-\catid{\ctxext{\Gamma}{A}}$ and $W_A(p_Q\circ f)\defeq$
-\end{constr}
-
-\begin{thm}
-There is a full and faithful functor from the category of D-systems to the
-category of E-systems.
-\end{thm}
-
-\section{An alternative characterization of E-systems}
-
-\begin{defn}
-An E'-system consists of a category $\cat{C}$ with a subcategory $\cat{F}$ with the
-same objects as $\cat{C}$, with
-the following structure
-\begin{enumerate}
-\item $\cat{F}$ has a terminal object.
-\item For any $f:\Delta\to\Gamma$ in $\cat{C}$, a functor $f_\ast:\cat{F}/\Gamma\to
-\cat{F}/\Delta$ %and
-%for any $Q\in\cat{F}/\ctxext{\Gamma}{B}$ and any section $g$ of $Q$ a section
-%$f_\ast(g)$ of $f_\ast(Q)$, 
-so that $f_\ast$ is contravariant in $f$ and preserves the terminal object. 
-\item For any $A\in\cat{F}/\Gamma$ and $f:\Delta\to\Gamma$, factorizations of
-$f$ through $A$ are naturally the same thing as sections of
-$f_\ast(A)$, i.e. there is an isomorphism
-\begin{equation*}
-\zeta_{A,f}%\defeq\theta\mapsto f_\ast(\theta)
-:\{\theta:\Delta\to\ctxext{\Gamma}{A}\mid A\circ\theta\jdeq f\}
-  \cong
-\{\eta:\Delta\to\ctxext{\Delta}{(f_\ast(A))}\mid f_\ast(A)\circ\eta\jdeq\catid{\Delta}\}
-\end{equation*}
-in the presheaf category 
-$\mathrm{Psh}(\op{(\cat{F}/\Gamma)}\times(\cat{C}/\Gamma))$. 
-\end{enumerate}
-\end{defn}
-
-\begin{rmk}
-Explicitly, the
-naturality condition states that for any $P\in\cat{F}/\ctxext{\Gamma}{A}$,
-$g:\greek{E}\to\Delta$ and $\theta:\Delta\to\ctxext{{\Gamma}{A}}{P}$ such that
-$P\circ\theta\jdeq f$, we have the equality
-\begin{equation*}
-\zeta_{A,f\circ g}(P\circ\theta\circ g)\jdeq f^\ast(P)\circ\zeta_{P,f}(\theta)\circ g.
-\end{equation*}
-\end{rmk}
-
-We can give the category $\cat{F}$ a term structure as follows:
-
-\begin{defn}
-For any E'-system $\mathbb{E}$, the category $\cat{F}$ has a term structure
-where the terms of $A$ are sections of $A$, i.e.
-\begin{equation*}
-T(A)\defeq\{\theta:\Gamma\to\ctxext{\Gamma}{A}\mid A\circ\theta\jdeq\catid{\Gamma}\}.
-\end{equation*}
-\end{defn}
-
-\begin{lem}
-Let $\mathbb{E}$ be an E'-system. Then there is a function $T(A)\to T(f_\ast(A))$
-for any $f:\Delta\to\Gamma$ in $\cat{C}$ and $A\in\cat{F}/\Gamma$, contravariant
-in $f$. 
-\end{lem}
-
-\begin{proof}
-Using the natural isomorphism $\zeta_{A,f}$, we see that
-\begin{align*}
-T(f_\ast(A)) & \jdeq \{\eta : \Delta\to\ctxext{\Delta}{f_\ast(A)}\mid f_\ast(A)\circ\eta\jdeq\catid{\Delta}\}
-  \\
-  & \cong
-\{\theta : \Delta\to \ctxext{\Gamma}{A}\mid A\circ\theta\jdeq f\}.
-\end{align*}
-Therefore, we have the function $x\mapsto \zeta_{A,f}(x\circ f)$ from $T(A)$ to
-$T(f_\ast(A))$. 
-\end{proof}
-
-\begin{defn}
-Consider $P\in\cat{F}/\ctxext{\Gamma}{A}$ and $f:\Delta\to\Gamma$,
-$\theta_1:\Delta\to\ctxext{\Gamma}{A}$ such that $A\circ\theta_1\jdeq f$, 
-and $\theta_2:\Delta\to\ctxext{{\Gamma}{A}}{P}$ such that $P\circ\theta_2\jdeq\theta_1$.  
-\end{defn}
-
-\begin{defn}
-For any $P\in\cat{F}/\ctxext{\Gamma}{A}$ and $f:\Delta\to\Gamma$ 
-there is a morphism $\tilde{f}:\ctxext{\Delta}{f_\ast(A)}\to\ctxext{\Gamma}{A}$
-with the property that $A\circ\tilde{f}\jdeq f\circ f_\ast(A)$, and
-we have isomorphisms
-\begin{align*}
-\{\theta:\ctxext{\Delta}{f_\ast(A)}\to\ctxext{{\Gamma}{A}}{P}\mid P\circ\theta\jdeq \tilde{f}\}
-&  \cong
-\{\eta:\ctxext{\Delta}{f_\ast(A)}\to\ctxext{{\Delta}{(f_\ast(A))}}{f_\ast(P)}\mid f_\ast(P)\circ\eta\jdeq\catid{\ctxext{\Delta}{f_\ast(A)}}\}
-\end{align*}
-\end{defn}
-
-\begin{constr}
-\begin{align*}
-& \{ \theta : \ctxext{\Delta}{f_\ast(A)}\to\ctxext{\Gamma}{A}\mid A\circ\theta\jdeq f\circ f_\ast(A) \} \\
-  & \cong
-\{ \eta : \ctxext{\Delta}{f_\ast(A)}\to\ctxext{{\Delta}{f_\ast(A)}}{(f\circ f_\ast(A))_\ast(A)}
-  \mid (f\circ f_\ast(A))_\ast(A)\circ\eta\jdeq\catid{\ctxext{\Delta}{f_\ast(A)}}\} \\
-  & \jdeq
-\{ \eta : \ctxext{\Delta}{f_\ast(A)}\to\ctxext{{\Delta}{f_\ast(A)}}{(f_\ast(A))_\ast(f_\ast(A))}
-  \mid (f\circ f_\ast(A))_\ast(A)\circ\eta\jdeq\catid{\ctxext{\Delta}{f_\ast(A)}}\} \\
-  & \jdeq
-T(W_{f_\ast(A)}(f_\ast(A))).
-\end{align*}
-Since we have $\tfid{f_\ast(A)}\in T(W_{f_\ast(A)}(f_\ast(A)))$, we obtain 
-\begin{equation*}
-\tilde{f}\defeq \zeta_{A,f\circ f_\ast(A)}^{-1}(\tfid{f_\ast(A)}):\ctxext{\Delta}{f_\ast(A)}\to\ctxext{\Gamma}{A}.
-\end{equation*}
-It remains to prove the asserted isomorphism. 
-\begin{align*}
-\{\theta:\ctxext{\Delta}{f_\ast(A)}\to\ctxext{{\Gamma}{A}}{P}\mid P\circ\theta\jdeq \tilde{f}\}
-&  \cong
-\{\eta : \ctxext{\Delta}{f_\ast(A)}\to\ctxext{{\Delta}{f_\ast(A)}}{\tilde{f}_\ast(P)} \mid
-  \tilde{f}_\ast(P)\circ\eta\jdeq\catid{}\}
-  \\
-  & \cong
-\{\eta:\ctxext{\Delta}{f_\ast(A)}\to\ctxext{{\Delta}{(f_\ast(A))}}{f_\ast(P)}\mid f_\ast(P)\circ\eta\jdeq\catid{\ctxext{\Delta}{f_\ast(A)}}\}
-\end{align*}
-\end{constr}
-
-\begin{lem}
-Let $\mathbb{E}$ be an E'-system. Then there is a function $T(P)\to T(f_\ast(P))$
-for any $f:\Delta\to\Gamma$ in $\cat{C}$ and $P\in\cat{F}/\ctxext{\Gamma}{A}$, contravariant
-in $f$. 
-\end{lem}
-
-\begin{proof}
-Using the natural isomorphism $\zeta_{A,f}$, we see that
-\begin{align*}
-T(f_\ast(P)) & \jdeq \{\eta : \ctxext{\Delta}{f_\ast(A)}\to\ctxext{{\Delta}{f_\ast(A)}}{f_\ast(P)}\mid f_\ast(P)\circ\eta\jdeq\catid{\ctxext{\Delta}{f_\ast(A)}}\}.
-\intertext{By all means, this should be isomorphic to}
-  & \cong
-\{\theta : \ctxext{\Delta}{f_\ast(A)}\to \ctxext{{\Gamma}{A}}{P}\mid P\circ\theta\jdeq \tilde{f}\}.
-\end{align*}
-Therefore, we have the function $x\mapsto \zeta_{A,f}(x\circ f)$ from $T(A)$ to
-$T(f_\ast(A))$. 
-\end{proof}
-
-\begin{thm}
-E-systems are the same things as E'-systems.
-\end{thm}
-
-\begin{proof}
-Given an E-system $\mathbb{E}$, we have the category $\cat{C}$ of internal
-morphisms in the empty context, with the subcategory $\cat{F}$ of projections $\cprojfstf{A}{P}$. 
-The functors $f_\ast:\cat{F}/\Gamma\to\cat{F}/\Delta$ are contravariant in
-$f$, so we have to verify whether we have the stated isomorphisms.
-
-Note first that for any $A\in\cat{F}/\Gamma$, we have
-$W_A\defeq A_\ast:\cat{F}/\Gamma\to\cat{F}/\ctxext{\Gamma}{A}$. Then $W_A$ is
-contravariant in $A$ and that it preserves the terminal object. We give $W_A$
-an action on terms by defining
-\begin{equation*}
-W_A(\theta)\defeq
-\end{equation*}
-for any $\theta\in T(Q)$ and $Q\in\cat{F}/\ctxext{\Gamma}{B}$.
-
-Thus, the $W_A$
-assemble a pre-weakening structure.
-
-Note that for any $A\in\cat{F}/\Gamma$, we have
-\begin{align*}
-T(W_A(A)) & \jdeq \{\eta:\ctxext{\Gamma}{A}\to\ctxext{{\Gamma}{A}}{W_A(A)}\mid W_A(A)\circ\eta\jdeq\catid{\ctxext{\Gamma}{A}}\} \\
-  & \cong
-\{\theta:\ctxext{\Gamma}{A}\to\ctxext{\Gamma}{A}\mid A\circ\theta\jdeq A\},
-\end{align*}
-so we find $\idtm{A}\defeq\zeta_{A,A}(\catid{\ctxext{\Gamma}{A}})\in T(W_A(A))$.
-Now let $\theta:\Delta\to\ctxext{\Gamma}{A}$ so that $A\circ\theta\jdeq f$
-for some $f:\Delta\to\Gamma$. Then we have $\zeta_{A,f}(\theta)\in T(f_\ast(A))$.  
-We wish to define $\theta_\ast(\idtm{A}):\Delta\to\ctxext{\Delta}{f_\ast(A)}$
-
-for any $A\in\cat{F}/\Gamma$ and $Q\in\cat{F}/\ctxext{\Gamma}{B}$, we have
-\begin{align*}
-T(B) 
-  & \jdeq 
-\{ \eta : \Gamma \to\ctxext{\Gamma}{B}
-        \mid B\circ\eta\jdeq\catid{\Gamma}\}
-  \\
-T(W_A(B))
-  & \jdeq
-\{ \eta : \ctxext{\Gamma}{A}\to\ctxext{{\Gamma}{A}}{W_A(B)}
-        \mid W_A(B)\circ\eta\jdeq\catid{\ctxext{\Gamma}{A}}\}
-  \\
-  & \cong
-\{ \theta : \ctxext{\Gamma}{A}\to\ctxext{\Gamma}{B}\mid B\circ\theta\jdeq A\}
-\end{align*}
-Therefore, $\eta\mapsto \zeta_{B,A}(\eta\circ A)$ defines a function from
-$T(B)$ to $T(W_A(B))$. 
-\end{proof}
-
-\begin{comment}
-\section{Acknowledgments}
-The research that eventually became the study of E-systems began while analyzing
-the internel models I encountered during investigations of the type theoretical
-descent theorem, in the summer of 2013. In a private conversation during the Conference on Type Theory, 
-Homotopy Theory and Univalent Foundations at the Universitat Aut\`onoma de 
-Barcelona, I presented the first version of the axioms to Richard Garner, who
-told me he found a similar structure  while analyzing the combinatorics of type
-dependency, see \cite{Garner2014}. After that conference, Joachim
-Kock generously offered me to stay for a few months at the UAB. He was intrigued
-by the idea of doing category theory using only sections, and during the sessions
-we had in this period I finalized the inference rules that correspond to the
-axioms of E-systems.
-
-In Paris, in the spring of 2014, Richard Garner introduced me to Vladimir 
-Voevodsky, who had been working on similar problems and wanted to know exactly
-what axioms should hold on B0-systems so that it is in the image of the functor
-from C-systems to B0-systems. In other words, he wanted to know what the axioms
-of B-systems were, and be confident that we had them all.
-
-In August 2015, I started a grad student of Steve Awodey at the Carnegie 
-Mellon University. With Steve we elaborated much more on the fact that the theory
-of E-systems is essentially algebraic, and I wrote down what an E-system is in
-an arbitrary category with finite limits.
-
-The discussions that followed both with Awodey and with Voevodsky, 
-at the CMU, at the Homotopy Type Theory workshop in Oxford (November 2014), and
-while visiting Voevodsky at the Institute for Advanced Study in Princeton (March 2015), 
-essentially shaped E-systems into their current presentation.
-
-I am grateful to all the people mentioned for sharing many valuable insights 
-with me. 
-\end{comment}

### 2015-07-29 internal_homs.tex
separation of sections into different files

--- 2015-06-12 weakeningcats.tex
@@
-\section{The definition of E-systems}\label{sec:esys_defn}
-\subsection{Weakening systems}
-
-\begin{defn}
-A \emph{pre-weakening structure on a pre-category $\cat{C}$} consists of a functor
-$W_A:\cat{C}/\Gamma\to \cat{C}/\ctxext{\Gamma}{A}$ for every $A:\ctxext{\Gamma}{A}\to \Gamma$ in $\cat{C}$ such 
-that
-\begin{enumerate}
-\item $W_{\catid{\Gamma}}\jdeq \catid{\cat{C}/\Gamma}$ for every object $\Gamma\in\cat{C}$.
-\item $W_{\ctxext{A}{P}}\jdeq W_P\circ W_A$ for every $P\in\cat{C}/\ctxext{\Gamma}{A}$ and $A\in\cat{C}/\Gamma$.
-\item $W_A$ preserves the final object, i.e.~$W_A(\catid{\Gamma})\jdeq \catid{\ctxext{\Gamma}{A}}$.
-\end{enumerate}
-A \emph{pre-weakening system} is a pre-category together with a pre-weakening structure.
-\end{defn}
-
-\begin{defn}
-A \emph{pre-weakening homomorphism} $F:\cat{C}\to\cat{D}$ between pre-weakening systems
-is a functor $F:\cat{C}\to\cat{D}$ such that the square
-\begin{equation*}
-\begin{tikzcd}[column sep=large]
-\cat{C}/\ctxext{\Gamma}{A}
-  \arrow[r,"F/\ctxext{\Gamma}{A}"]
-  &
-\cat{D}/F(\ctxext{\Gamma}{A})
-  \\
-\cat{C}/\Gamma
-  \arrow[u,"W_A"]
-  \arrow[r,swap,"F/\Gamma"]
-  &
-\cat{D}/F(\Gamma)
-  \arrow[u,swap,"W_{F(A)}"]
-\end{tikzcd}
-\end{equation*}
-commutes for any $A\in\cat{C}/\Gamma$.
-\end{defn}
-
-\begin{lem}
-Let $\cat{C}$ be a pre-weakening system. Then for any object $\Gamma$ of $\cat{C}$,
-the slice pre-category $\cat{C}/\Gamma$ is also a pre-weakening system, with the weakening
-structure given by $W(\cat{C}/\Gamma)_P\jdeq W(\cat{C})_P$ for each $P\in\cat{C}/\ctxext{\Gamma}{A}$
-and each $A\in\cat{C}/\Gamma$.
-\end{lem}
-
-\begin{defn}
-A \emph{weakening system} is a pre-weakening system for which every $W_f$ is a
-pre-weakening homomorphism. A \emph{weakening homomorphism} is a pre-weakening 
-homomorphism between weakening systems.
-\end{defn}
-
-\begin{rmk}
-In other words, a weakening system is a pre-weakening system with a weakening
-structure $W$ for which the square
-\begin{equation*}
-\begin{tikzcd}[column sep=huge]
-\cat{C}/\ctxext{{\Gamma}{B}}{Q}
-  \arrow[r,"W_A/\ctxext{B}{Q}"]
-  &
-\cat{C}/\ctxext{{\Gamma}{A}}{W_A(\ctxext{B}{Q})}
-  \\
-\cat{C}/\ctxext{\Gamma}{B}
-  \arrow[u,"W_Q"]
-  \arrow[r,swap,"W_A/B"]
-  &
-\cat{C}/\ctxext{{\Gamma}{A}}{W_A(B)}
-  \arrow[u,swap,"W_{W_A(Q)}"]
-\end{tikzcd}
-\end{equation*}
-for each $A,B\in\cat{C}/\Gamma$ and $Q:\cat{C}/\ctxext{\Gamma}{B}$. The objects part of this 
-property asserts that for any $k\in \cat{C}/\greek{E}$, the dotted arrows in the diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=tiny]
-\bullet \arrow[dr,densely dotted,swap,near start,"(W_A/\ctxext{B}{Q})(W_Q(R))"] & & \bullet \arrow[dl,densely dotted,near start,"W_{(W_A/B)(Q)}((W_A/B)(R))"] & & & & & & & & & & \bullet \arrow[dl,"W_Q(R)"] \\
-& \bullet \arrow[dr,swap,"(W_A/B)(Q)"] & & \bullet \arrow[dl,"(W_A/B)(R)"] & & & & & & & & \bullet \arrow[dr,swap,"Q"] & & \bullet \arrow[dl,"R"] \\
-& & \bullet \arrow[d,swap,"W_A(B)"] & & & & & & & & & & \ctxext{\Gamma}{B} \arrow[d,"B"] \\
-& & \ctxext{\Gamma}{A} \arrow[rrrrrrrrrr,swap,"A"] & & & & & & & & & & \Gamma
-\end{tikzcd}
-\end{equation*}
-are equal.
-
-A useful special case of this property is where $B\jdeq \catid{\Gamma}$. Thus, if $W$ is
-a weakening system, then the diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=huge]
-\cat{C}/\ctxext{\Gamma}{C}
-  \arrow[r,"W_A/C"]
-  &
-\cat{C}/\ctxext{{\Gamma}{A}}{W_A(C)}
-  \\
-\cat{C}/\Gamma
-  \arrow[u,"W_C"]
-  \arrow[r,swap,"W_A"]
-  &
-\cat{C}/\ctxext{\Gamma}{A}
-  \arrow[u,swap,"W_{W_A(C)}"]
-\end{tikzcd}
-\end{equation*}
-commutes for every $A,C\in\cat{C}/\Gamma$. In particular, we see
-that $W_A(W_C(D))\jdeq W_{W_A(C)}(W_A(D))$ for any $D\in\cat{C}/\Gamma$,
-i.e.~that weakening is a self-distributive operation. 
-\end{rmk}
-
-\begin{cor}
-For any object $\Gamma$ of a weakening system $\cat{C}$, the slice pre-category
-$\cat{C}/\Gamma$ is a weakening system.
-\end{cor}
-
-\begin{eg}
-If the underlying pre-category is a group, the axioms of a weakening system
-obtain the following form: There is a group homomorphism $W_x:G\to G$ for
-every $x\in G$, in such a way that $W_e\jdeq \catid{G}$ and $W_{xy}\jdeq W_y\circ W_x$
-for every $x,y\in G$, and for any $x,y\in G$ the diagram
-\begin{equation*}
-\begin{tikzcd}
-G \arrow[r,"W_y"]
-  &
-G 
-  \\
-G \arrow[u,"W_x"]
-  \arrow[r,swap,"W_y"]
-  &
-G \arrow[u,swap,"W_{W_y(x)}"]
-\end{tikzcd}
-\end{equation*}
-is required to commute. Thus, a weakening system on a group $G$ consists of
-a group homomorphism $W:\op G\to\mathrm{Aut}(G)$ such that $W_a\circ W_b\jdeq 
-W_{W_a(b)}\circ W_a$. The latter requirement is satisfied if $b\cdot a\jdeq a\cdot W_a(b)$
-for every $a,b\in G$, or equivalently if $[b,a]\jdeq b^{-1}\cdot W_a(b)$.
-\end{eg}
-
-\begin{lem}
-For any group, conjugation is a weakening structure. Thus, any group is
-canonically a weakening system.
-\end{lem}
-
-\begin{proof}
-Note that $e^{-1}xe\jdeq x$ and $(xy)^{-1}z(xy)\jdeq y^{-1}(x^{-1}zx)y$ for any $x,y,z\in G$.
-Also, the diagram
-\begin{equation*}
-\begin{tikzcd}
-G \arrow[r,"\mathrm{conj}_y"]
-  &
-G 
-  \\
-G \arrow[u,"\mathrm{conj}_x"]
-  \arrow[r,swap,"\mathrm{conj}_y"]
-  &
-G \arrow[u,swap,"\mathrm{conj}_{y^{-1}xy}"]
-\end{tikzcd}
-\end{equation*}
-commutes.
-\end{proof}
-
-\begin{comment}
-\begin{rmk}
-Similarly, every groupoid has a canonical weakening structure given by
-conjugation.
-\end{rmk}
-\end{comment}
-
-\begin{comment}
-\begin{lem}
-Let $G$ be a graph, and let $\cat{C}_G$ be the free pre-category generated by
-$G$. A pre-weakening structure on $\cat{C}_G$ is determined by functors
-$W_e:\cat{C}_G/\Gamma\to\cat{C}_G/\Delta$ for every edge $e:\Delta\to \Gamma$ in $G$, with the
-property that $W_e(\catid{\Gamma})\jdeq \catid{\Delta}$.
-
-A weakening structure on $\cat{C}_G$ is
-a pre-weakening structure on $\cat{C}_G$ such that the diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=huge]
-\cat{C}_G/\greek{Z} \arrow[r,"W_e/(g\circ h)"] & \cat{C}_G/\mathrm{dom}(W_e(g\circ h)) \\
-\cat{C}_G/\greek{E} \arrow[u,"W_h"] \arrow[r,swap,"W_e/g"] & \cat{C}_G/\mathrm{dom}(W_e(g)) \arrow[u,swap,"W_{W_e(h)}"]
-\end{tikzcd}
-\end{equation*}
-commutes for every arrow $g:\greek{E}\to \Gamma$ in $\cat{C}_G$ and every two edges $e:\Delta\to \Gamma$ and
-$h:\greek{Z}\to \greek{E}$ in $G$.
-\end{lem}
-
-\begin{eg}
-Let $G$ be the graph consisting of one vertex $\ast$ and one edge $e:\ast\to\ast$.
-Then $\cat{C}_G/\ast\jdeq \N$. Then the only weakening structure on $\cat{C}/G$ is
-given by $W_e\jdeq \catid{\N}$. The reason is that 
-\end{eg}
-\end{comment}
-
-\begin{lem}
-Let $P$ be a pre-order. A pre-weakening structure on $P$ is determined by order
-preserving maps $W_{a,b}:{\downarrow}(b)\to{\downarrow}(a)$ satisfying
-$W_{a,b}(b)\jdeq a$, for every $a\leq b$, where
-${\downarrow}(x)$ is the set $\{y\in P\mid y\leq x\}$.
-
-A weakening structure on $P$ is a pre-weakening structure on $P$ such that the
-diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=huge]
-{\downarrow}(c') \arrow[r,"W_{a,b}|_{{\downarrow}(c')}"] & {\downarrow}(W_{a,b}(c')) \\
-{\downarrow}(c) \arrow[u,"W_{c',c}"] \arrow[r,swap,"W_{a,b}|_{{\downarrow}(c)}"] & {\downarrow}(W_{a,b}(c)) \arrow[u,swap,"W_{W_{a,b}(c'),W_{a,b}(c)}"]
-\end{tikzcd}
-\end{equation*}
-for any $a\leq b$ and $c'\leq c\leq b$.
-\end{lem}
-
-\begin{eg}
-In a poset with meets, the functions $W_{a,b}:{\downarrow}(b)\to{\downarrow}(a)$
-given by $W_{a,b}(x)\defeq a\land x$ assemble a weakening structure, because for
-any $a\leq b$ and $x,c'\leq c\leq b$ we have $a\land(c'\land x)\jdeq (a\land c')\land
-(a\land x)$.  
-\end{eg}
-
-\begin{eg}
-Consider the poset $(\N,\geq)$, which has a morphism from $a$ to $b$ if $a\geq b$.
-On this poset, we may consider the weakening structure defined by $W_{a,b}(x)\jdeq
-a+(x-b)$.
-\end{eg}
-
-\subsection{Weakening systems with term structure}
-\begin{defn}
-A \emph{category with a term structure} is a pre-category $\cat{C}$ with a set $T(A)$ for every
-$A\in\cat{C}/\Gamma$ for every $\Gamma\in\cat{C}$. A \emph{functor with term structure} from $\cat{C}$ to $\cat{D}$
-is a functor $F:\cat{C}\to\cat{D}$ with a function $T(A)\to T(F(A))$ for
-every $A\in\cat{C}/\Gamma$, for every $\Gamma\in\cat{C}$.
-\end{defn}
-
-\begin{defn}
-Suppose $\cat{C}$ is a pre-category with term structure, and $\Gamma$ is an object of $\cat{C}$. 
-Then $\cat{C}/\Gamma$ is a pre-category with term structure, where $T_{\cat{C}/\Gamma}(A)\jdeq T_\cat{C}(A)$.
-\end{defn}
-
-\begin{defn}
-A \emph{weakening system with term term structure} is the same thing as a weakening system,
-except that the underlying pre-category and the functors $W_A$ are required to
-be functors with term term structure. The equalities are also required to hold on the term structure.
-\end{defn}
-
-\begin{cor}
-For any object $\Gamma$ of a weakening system with term structure, the slice pre-category
-$\cat{C}/\Gamma$ is a weakening system with term structure.
-\end{cor}
-
-\begin{eg}
-%For any group $G$ with weakening structure $W:\op{G}\to\mathrm{Aut}(G)$, a term 
-%structure is determined by a presheaf on $G/\ker(W)$.
-
-Let $G$ be a group, with the weakening structure given by conjugation. A term
-structure on $G$ is given by a set $X$ with a right $G$-action $\mu:\op G\times X\to X$.
-We take $T(a)\defeq X$ for every $a\in G$.
-\end{eg}
-
-\subsection{Projection systems}
-\begin{defn}
-A \emph{pre-projection system} is a pre-weakening system with term structure with an element
-$\tfid{A}\in T(W_A(A))$ for every $A\in\cat{C}/\Gamma$ and $\Gamma\in\cat{C}$. 
-\end{defn}
-
-\begin{defn}
-A \emph{pre-projection homomorphism} $F:\cat{C}\to\cat{D}$ is a pre-weakening homomorphism for which
-\begin{equation*}
-F(\tfid{A})\jdeq \tfid{F(A)}
-\end{equation*}
-for every $A\in\cat{C}/\Gamma$ and $\Gamma\in\cat{C}$.
-\end{defn}
-
-\begin{lem}
-For any object $\Gamma$ in a pre-projection system $\cat{C}$, the slice pre-category
-$\cat{C}/\Gamma$ is a pre-projection system.
-\end{lem}
-
-\begin{proof}
-Straightforward.
-\end{proof}
-
-\begin{defn}
-A \emph{projection system} is a pre-projection system for which every $W_A$ is
-a pre-projection homomorphism. A \emph{projection homomorphism} is a pre-projection homomorphism
-between projection systems.
-\end{defn}
-
-\begin{cor}
-For any object $\Gamma$ of a projection system $\cat{C}$, the slice pre-category $\cat{C}/\Gamma$
-is a projection system.
-\end{cor}
-
-%\begin{eg}
-%Consider a group $G$ with a right $G$-set $X$. A projection structure on $(G,X)$
-%consists of an element $\tfid{a}\in X$ for every $a\in G$, so that $\tfid{a}\cdot b=
-%\tfid{b^{-1}ab}$ for every $a,b\in G$. 
-%For instance, this is the case if we take $X$ to be a pointed set with $\ast\in X$,
-%and take $\tfid{a}:=\ast$ for all $a\in G$.
-%\end{eg}
-
-\subsection{Substitution systems}
-
-\begin{defn}
-A \emph{pre-substitution structure} on a pre-category with term structure $\cat{C}$ consists
-of a functor with term structure $S_x:\cat{C}/\ctxext{\Gamma}{A}\to\cat{C}/\Gamma$ for every $x\in T(A)$ and
-$A\in\cat{C}/\Gamma$, such that $S_x(\catid{\ctxext{\Gamma}{A}})\jdeq\catid{\Gamma}$.
-
-A \emph{pre-substitution system} is a pre-category with term structure together with a
-pre-substitution structure. 
-\end{defn}
-
-\begin{defn}
-A \emph{pre-substitution homomorphism} $F:\cat{C}\to\cat{D}$ is a functor with term structure for
-which the diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=large]
-\cat{C}/\ctxext{\Gamma}{A}
-  \arrow[r,"F/\ctxext{\Gamma}{A}"]
-  \arrow[d,swap,"S_x"]
-  &
-\cat{D}/F(\ctxext{\Gamma}{A})
-  \arrow[d,"S_{F(x)}"]
-  \\
-\cat{C}/\Gamma
-  \arrow[r,swap,"F/\Gamma"]
-  &
-\cat{D}/F(\Gamma)
-\end{tikzcd}
-\end{equation*}
-commutes for every $x\in T(A)$ and $A\in\cat{C}/\Gamma$.
-\end{defn}
-
-\begin{lem}
-For any object $\Gamma$ of a pre-substitution system $\cat{C}$, the slice pre-category
-$\cat{C}/\Gamma	$ is a pre-substitution system.
-\end{lem}
-
-\begin{defn}
-A \emph{substitution system} is a pre-substitution system for which each
-$S_t$ is a pre-substitution homomorphism. A \emph{substitution homomorphism} is a
-pre-substitution homomorphism between substitution systems.
-\end{defn}
-
-\begin{cor}
-For any object $\Gamma$ of a substitution system, the slice pre-category $\cat{C}/\Gamma$
-is a substitution system.
-\end{cor}
-
-\begin{rmk}
-The condition that every $S_x$ is a substitution homomorphism, asserts that
-the diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=large]
-\cat{C}/\ctxext{{{\Gamma}{A}}{P}}{Q}
-  \arrow[r,"S_x/\ctxext{P}{Q}"]
-  \arrow[d,swap,"S_y"]
-  &
-\cat{C}/\ctxext{{\Gamma}{S_x(P)}}{S_x(Q)}
-  \arrow[d,"S_{S_x(y)}"]
-  \\
-\cat{C}/\ctxext{{\Gamma}{A}}{P}
-  \arrow[r,swap,"S_x/P"]
-  &
-\cat{C}/\ctxext{\Gamma}{S_x(P)}
-\end{tikzcd}
-\end{equation*}
-commutes for every $y\in T(Q)$.
-\end{rmk}
-
-\begin{eg}
-Let $G$ be a group and let $H$ be a set, forming a category with term 
-structure of which $T(a)\jdeq H$ for every $a\in G$. A substitution
-system on $(G,H)$ consists of a function $\sigma_G : H\times G\to G$ and
-a function $\sigma_H:H\times H\to H$, for which the diagrams
-\begin{equation*}
-\begin{tikzcd}[column sep=large]
-G \arrow[r,"\sigma_G(h)"]
-  \arrow[d,swap,"\sigma_G(k)"]
-  &
-G \arrow[d,"{\sigma_G(\sigma_H(h,k))}"]
-  \\
-G \arrow[r,swap,"\sigma_G(h)"] 
-& G
-\end{tikzcd}
-\end{equation*}
-and
-\begin{equation*}
-\begin{tikzcd}[column sep=large]
-H \arrow[r,"\sigma_H(h)"]
-  \arrow[d,swap,"\sigma_H(k)"]
-  &
-H \arrow[d,"{\sigma_H(\sigma_H(h,k))}"]
-  \\
-H \arrow[r,swap,"\sigma_H(h)"]
-  &
-H
-\end{tikzcd}
-\end{equation*}
-commute for every $h,k\in H$. 
-\end{eg}
-
-\subsection{E-systems}
-
-Although it is now possible to state the definition of an E-system without
-any difficulty, it shall be convenient later to also have a category of
-pre-E-systems at our disposal. Pre-E-systems shall by default be non-unital,
-i.e.~they do not come with a pre-projection system, while E-systems shall
-always be assumed to be unital. In the present context, we shall not need
-unital pre-E-systems or non-unital E-systems.
-
-\begin{defn}
-A \emph{pre-E-system} is a pre-category $\cat{C}$ with term structure
-with a pre-weakening structure $W$, a pre-substitution structure $S$ and a terminal
-object. A \emph{unital pre-E-system}, or simply \emph{pre-E-system}, is a non-unital pre-E-system
-which also has the structure of a pre-projection system. 
-\end{defn}
-
-\begin{defn}
-A \emph{pre-E-homomorphism} from $\mathbb{E}\to\mathbb{D}$ is a functor
-$H:\mathbb{C}_\mathbb{E}\to\mathbb{C}_\mathbb{D}$ between the underlying categories
-with term structure, which is a pre-weakening homomorphism, a pre-substitution
-homomorphism, and which preserves the terminal object.
-
-A \emph{unital} pre-E-homomorphism is a non-unital pre-E-homomorphism which is 
-also a pre-projection homomorphism. 
-\end{defn}
-
-\begin{defn}
-\label{defn:esystems}
-An \emph{E-system} is a unital pre-E-system in which 
-\begin{enumerate}
-\item each $W_A$ is a unital pre-E-homomorphism,
-\item each $S_x$ is a unital pre-E-homomorphism,
-\item \label{tTf:StWf_id} weakened families are constant, i.e.~$S_x\circ W_A\jdeq \catid{\cat{C}/\Gamma}$ for any $x\in T(A)$ and $A\in\cat{C}/\Gamma$,
-\item $S_x(\tfid{A})\jdeq x$ for any $x\in T(A)$ and $A\in\cat{C}/\Gamma$, and
-\item \label{f:SidfWff_id} $S_{\tfid{A}}\circ W_A/A\jdeq \catid{\cat{C}/\ctxext{\Gamma}{A}}$ for any $A\in\cat{C}/\Gamma$.
-\end{enumerate}
-An \emph{E-homomorphism} $H:\mathbb{E}\to\mathbb{D}$ is a functor from
-$\cat{C}_{\mathbb{E}}$ to $\cat{C}_{\mathbb{D}}$ which is both a projection
-homomorphism and a substitution homomorphism.
-\end{defn}
-
-\begin{rmk}
-The condition that each $W_A$ is a substitution homomorphism asserts that
-the diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=huge]
-\cat{C}/\ctxext{{\Gamma}{B}}{Q}
-  \arrow[r,"W_A/\ctxext{B}{Q}"]
-  \arrow[d,swap,"S_y"]
-  &
-\cat{C}/\ctxext{{{\Gamma}{A}}{W_A(B)}}{W_A(Q)}
-  \arrow[d,"S_{W_A(y)}"]
-  \\
-\cat{C}/\ctxext{\Gamma}{B}
-  \arrow[r,swap,"W_A/B"]
-  &
-\cat{C}/\ctxext{{\Gamma}{A}}{W_A(B)}
-\end{tikzcd}
-\end{equation*}
-of functors with term structure commutes for every 
-$Q\in\cat{C}/\ctxext{\Gamma}{B}$, $B\in\cat{C}/\Gamma$ and each $y\in T(Q)$.
-
-Likewise, the condition that each $S_x$ is a weakening homomorphism
-asserts that the diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=huge]
-\cat{C}/\ctxext{{\Gamma}{A}}{P}
-  \arrow[r,"S_x/P"]
-  \arrow[d,swap,"W_Q"]
-  &
-\cat{C}/\ctxext{\Gamma}{S_x(P)}
-  \arrow[d,"W_{S_x(Q)}"]
-  \\
-\cat{C}/\ctxext{{{\Gamma}{A}}{P}}{Q}
-  \arrow[r,swap,"S_x/\ctxext{P}{Q}"]
-  &
-\cat{C}/\ctxext{{\Gamma}{S_x(P)}}{S_x(Q)}
-\end{tikzcd}
-\end{equation*}
-of functors with term structure commutes for every 
-$Q\in\cat{C}/\ctxext{{\Gamma}{A}}{P}$.
-\end{rmk}
-
-\begin{cor}
-For any object $\Gamma$ of an E-system $\cat{C}$, the slice pre-category $\cat{C}/\Gamma$
-is an E-system.
-\end{cor}
-
-\begin{comment}
-\begin{rmk}[Needs more precision]
-Consider the category $E$ generated by the graph consisting of the objects of 
-the underlying category of an $E$-system $\mathcal{E}$ as edges, and the morphisms
-$S_t:X\to Y$ and $W_A:Y\to X$ as edges, subject to the axioms of an $E$-systems.
-
-Any morphism of $E$ can be rewritten in a unique way
-as a composition $\tilde S\circ \tilde W$, where $\tilde S$ is a morphism in
-substitution-normal form, $\tilde W$ is in weakening normal form, every term
-expression is in normal form, and where
-each occurence of $S_x\circ W_A$ and $S_{\tfid{A}}\circ W_A/A$ is eliminated.
-\end{rmk}
-\end{comment}
-
-\section{Properties of E-systems}\label{sec:esys_props}
+\section{The category of internal morphisms of an E-system}\label{sec:esys_props}
 
@@
 \end{rmk}
-
-\subsection{Internal E-systems}
-
-\subsubsection{Internal weakening systems}
-
-\subsubsection{Internal projection systems}
-
-\subsubsection{Internal substitution systems}
-
-\subsubsection{Internal E-systems}
+\end{comment}
 
@@
 \end{comment}
-
-\section{E-systems compared to other systems of type dependency}\label{sec:esys_compared}
-In an effort to formalize the simplicial set model of Martin-L\"of type theory
-with the univalence axiom, Voevodsky has studied various notions of models of
-type theory, taking Cartmell's \emph{contextual categories} as a starting
-point. Voevodsky's definition of a \emph{C-system} is equivalent to that of Cartmell
-\cite{VV_C-systems_quotients}. There are several techniques of defining 
-C-systems: in \cite{VV_Csys_univ}, C-systems are constructed out of a universe
-in a category; in \cite{VV_C-systems_quotients}, sub-C-systems and regular
-quotients of C-systems are defined; in \cite{VV_C-systems_monad}, a C-system
-is constructed out of a monad $R$ on the category of sets and a left $R$-module
-with values in $\mathbf{Set}$.
-
-In \cite{VV_B-systems}, Voevodsky introduced his notion of \emph{B-systems}, which is
-an essentially algebraic theory of type dependency with infinitely many sorts 
-and operations, and subsequently he showed that there is a full and faithful
-functor from the category of C-systems to the category of B0-systems, of which
-the image is contained in the subcategory of B-systems. This fact is important at least
-for a theoretical reason: it tells us that the conditions of being a B-system
-are necessary to axiomatize type dependency. In other words,
-this is a theorem that tells us that we need no further axioms. It is
-unknown whether the categories of C-systems and B-systems are actually equivalent.
-This question would settle the sufficiency of the conditions of being a B-system.
-
-The category of B-systems will be shown to be equivalent to the category of
-stratified E-systems. However, we do not arrive at the fact that the functor
-$\mathcal{E}:\mathbf{Bsys}\to\mathbf{Esys}$ is full without imposing an additional
-condition on the E-homomorphisms. 
-To ensure that the category of stratified E-systems with stratified 
-E-homomorphisms between them is a full subcategory of the category 
-$\mathbf{Esys}$, we will require that the underlying functor of each E-homomorphism
-lifts factorizations. It is
-important to note that this condition is also algebraic, so that the category
-of E-systems remains algebraic.
-
-\subsection{Lifting factorizations}
-Recall
-that for any morphism $f:X\to Y$ in a category $\cat{C}$ there is a category
-$\mathbf{fact}(f)$ of factorizations of $f$. Also, any functor $F:\cat{C}\to\cat{D}$
-determines a functor $F_{\mathbf{fact}(f)}:\mathbf{fact}(f)\to\mathbf{fact}(H(f))$
-for any morphism $f$ of $\cat{C}$.
-
-\begin{defn}
-A functor $F:\cat{C}\to\cat{D}$ is said to
-\emph{lift factorizations} if for any $f\in\cat{C}/X$, the functor
-$F_{\mathbf{fact}(f)}:\mathbf{fact}(f)\to\mathbf{fact}(F(f))$ is an isomorphism
-of categories. 
-\end{defn}
-
-\begin{comment}
-\begin{rmk}
-We may choose the property of lifting factorizations
-to involve either an equivalence or an isomorphism of categories. The
-isomorphism-version, would say that for every factorization $h'\circ g'$
-of $F(f)$ there is a unique factorization $f\jdeq h\circ g$ in $\cat{C}$ such
-that $F(g)\jdeq g'$ and $F(h)\jdeq h'$. In the version with equivalences,
-the uniqueness is replaced by uniqueness up to isomorphism, and the equalites
-are replaced by an isomorphism in $\mathbf{fact}(F(f))$.
-
-It is easier to state the isomorphism version of lifting factorizations with
-inference rules and it might be easier to explain this condition on type theoretical
-grounds, although the equivalence version has the advantage of being
-categorical (i.e.~invariant under equivalence of categories).
-
-In the present
-context, it doesn't matter very much which one we pick: the categories in which
-we're interested are all posets.
-\end{rmk}
-\end{comment}
-
-\begin{eg}
-For any slice category $\cat{C}/X$, the forgetful functor $\cat{C}/X\to\cat{C}$
-lifts factorizations.
-\end{eg}
-
-\begin{defn}
-We write $\mathbf{Esys}$ for the category of E-systems where the underlying
-functor of each E-homomorphism lifts factorizations.
-\end{defn}
-
-\subsection{Stratified E-systems}
-
-In this subsection we define the notion of stratified E-system. Then we show
-that the category $\mathbf{Esys_s}$ of stratified E-systems with homomorphisms preserving the
-stratification, is a full subcategory of $\mathbf{Esys}$.
-
-\begin{defn}
-A category $\cat{C}$ with terminal object is said to be \emph{stratified} if there exists a 
-\emph{stratification functor}
-\begin{equation*}
-L : \cat{C}\to (\mathbb{N},\geq)
-\end{equation*}
-such that
-\begin{enumerate}
-\item $L(X)\jdeq 0$ if and only if $X$ is terminal, and for any $f:X\to Y$ we have
-$L(X)\jdeq L(Y)$ if and only if $X\jdeq Y$ and $f\jdeq\catid{X}$. 
-\item every morphism $f:X\to Y$ in $\cat{C}_\mathbb{E}$, where $L(X)\jdeq
-n+m+1$ and $L(Y)\jdeq n$, has a unique factorization 
-\begin{equation*}
-\begin{tikzcd}
-X \arrow[r,"f_m"] & X_m \arrow[r,"f_{m-1}"] & \cdots \arrow[r,"f_1"] & X_1 \arrow[r,"f_0"] & Y
-\end{tikzcd}
-\end{equation*}
-where $L(X_i)\jdeq n+i$.
-\end{enumerate}
-A functor $F:\cat{C}\to\cat{D}$ between stratified categories is said to be stratified 
-if $L_{\cat{C}}\jdeq L_{\cat{D}}\circ F$. 
-\end{defn}
-
-%\begin{rmk}
-%The categorical version of being stratified, is that there exists a factofibration
-%$L:\cat{C}\to (N,\geq)$ which preserves the terminal object.
-%\end{rmk}
-
-\begin{lem}
-Any stratified category is a rooted tree.
-\end{lem}
-
-\begin{proof}
-Let $\cat{C}$ be a stratified category. The root of the tree is going to be the terminal
-object of $\cat{C}$. Note that objects of $\cat{C}$ have no automorphisms other
-than the identity morphisms. Therefore, to show that $\cat{C}$ is a tree,
-it suffices to show that for every object $X$ with $L(X)>0$ there is
-a unique object $\eft(X)$ with $L(\eft(X))\jdeq L(X)-1$, with a morphism $X\to \eft(X)$.
-The existence follows from the existence of the factorization of $X\to 1$. The 
-uniqueness follows from the observation that for any arrow
-$X\to Y$ with $L(Y)\jdeq L(X)-1$, we can factorize $X\to 1$ through $Y$. Since
-factorizations are unique, we will have $Y\jdeq \eft(X)$. 
-\end{proof}
-
-\begin{lem}
-The category of small stratified categories is a full subcategory of the category
-of small categories with terminal objects and between them functors which lift 
-factorizations and preserve the terminal object.
-\end{lem}
-
-\begin{proof}
-Let $F:\cat{C}\to\cat{D}$ be a functor between stratified categories which lifts
-factorizations and 
-which preserves terminal objects. We show by induction that $F$ is stratified.
-Since $F$ preserves the terminal object, $L(Y)\jdeq 0$ implies $L(F(Y))\jdeq 0$.
-Now suppose that for $n\in\mathbb{N}$, we have that $L(\Gamma)\jdeq n$ implies
-$L(F(Y))\jdeq n$, and let $X$ be such that $L(X)\jdeq n+1$ with
-$p_X:X\to\eft(X)$. Then the poset $\mathbf{fact}(p_X)$ is isomorphic to the
-poset $\mathbf{2}$ with two objects, one of which is smaller than the other. By the
-assumption that $F$ lifts factorizations, the poset $\mathbf{fact}(F(p_X))$
-is also isomorphic to $\mathbf{2}$. Thus, the only factorization
-of $F(X)\to\unit$ starts with $F(p_X):H(X)\to F(\eft(X))$, which implies that
-$L(F(X))\jdeq n+1$. 
-\end{proof}
-
-\begin{rmk}
-It is easy to show that a category can only be stratified in at most one way,
-and that if a category is stratified, then so are any of its slices. Therefore,
-it makes sense to ask for the property that the weakening and substitution
-functors in a pre-weakening or pre-substitution system are stratified.
-\end{rmk}
-
-\begin{defn}
-A pre-E-system is said to be \emph{stratified} if its underlying category is
-stratified and if each $W_A$ and $S_x$ is a stratified functor. The category of
-stratified pre-E-systems with stratified E-homomorphisms between them is denoted by
-$\mathbf{Esys_s}$.
-\end{defn}
-
-\begin{cor}\label{lem:strat_full}
-The category of stratified pre-E-systems is a full subcategory of the category of
-pre-E-systems.
-\end{cor}
-
-\subsection{The embedding of B0-systems into pre-E-systems}
-
-Voevodsky distinguishes between pre-B-systems, which is an essentially algebraic theory with
-only operations and no equalities; B0-systems, which is an extension of the
-essentially algebraic theory of pre-B-systems which ensures that all the operations behave
-well with two sets of the operations ($\eft$ and $\ebd$); and B-systems, which extends the 
-theory of B0-systems with equalities ensuring that the rest of the operations 
-are compatible with each other. Voevodsky, furthermore distinguishes
-non-unital and unital versions of these, and noted that not all B-homomorphisms
-between unital B-systems are unital. The same distinguishment could be made for 
-E-systems by forgetting the projection structure, but we have not been able to construct a
-functor from unital B0-systems to unital pre-E-systems. Therefore, we shall only
-consider non-unital B0-systems and unital B-systems. In this section we will
-recall the definition of non-unital B0-systems and construct a full and faithful
-functor from them to the category of pre-E-systems.
-
-One reason for introducing B-systems by first introducing pre-B-systems is that
-it is clear in the definition of pre-B-systems that there are only countably
-many operations. Since we are interested in embedding the B-systems into E-systems,
-we shall organize the definition of B-systems a bit differently, but of course
-the resulting definition of B0-systems and B-systems will be (trivially) 
-equivalent.
-
-\begin{defn}
-A \emph{B-framework} is a collection of data of the following form:
-\begin{enumerate}
-\item for all $n\in\mathbb{N}$ two sets $B_n$ and $\tilde{B}_n$. 
-\item for all $n\in\mathbb{N}$ maps of the form
-\begin{align*}
-\eft[n] & : B_{n+1}\to B_n \\
-\ebd[n] & : \tilde{B}_n\to B_n.
-\end{align*}
-For $m,n\in\mathbb{N}$, we denote the composition $\eft[n]\circ\cdots\circ\eft[n+m]:B_{n+m+1}\to B_n$ by $\eft[n]^m$. 
-\item $B_0$ is a singleton $\{\pt\}$.
-\end{enumerate} 
-\end{defn}
-
-\begin{rmk}
-Since the indices $n,m\in\mathbb{N}$ can be infered, we will usually omit them
-in the infix form of the operators.
-\end{rmk}
-
-\begin{defn}
-A homomorphism $H:\mathbb{B}\to\mathbb{A}$ of B-frameworks consists of maps
-$H_n:B_n\to A_n$ and $\tilde{H}_n:\tilde{B}_n\to\tilde{A}_n$ such that
-\begin{align*}
-\eft(H(X)) & \jdeq H(\eft(X)) \\
-\ebd(\tilde{H}(x)) & \jdeq H(\ebd(x))
-\end{align*}
-for any $X\in B_n$ and $x\in\tilde{B}_n$. The category of B-frameworks is
-denoted by $\mathbf{Bfr}$. 
-\end{defn}
-
-\begin{defn}
-For every B-framework $\mathbb{B}$ and any $X\in B_n$, there is a B-framework
-$\mathbb{B}/X$ given by
-\begin{align*}
-(B/X)_{m} & \jdeq \{Y\in B_{n+m}\mid\eft^{m}(Y)\jdeq X\}\\
-(\tilde{B}/X)_m & \jdeq \{y\in \tilde{B}_{n+m}\mid\eft^m(\ebd(y))\jdeq X\}.
-\end{align*}
-Also, for any homomorphism $H:\mathbb{B}\to\mathbb{A}$ of B-frameworks and any
-$X\in B_n$, there is a homomorphism $H/X:\mathbb{B}/X\to\mathbb{A}/H(X)$
-defined in the obvious way.
-\end{defn}
-
-\begin{rmk}
-Note that for $X\in B_n$ and $Y\in B_{n+m}$ such that $\eft^m(Y)\jdeq X$, 
-we have $(\mathbb{B}/X)/Y\cong B/Y$.
-\end{rmk}
-
-\begin{defn}
-A B0-system $\mathbb{B}$ consists of a B-framework $\mathbb{B}$ and homomorphisms
-\begin{align*}
-W_{X} & : \mathbb{B}/\eft(X)\to\mathbb{B}/X\\
-S_{x} & : \mathbb{B}/\ebd(x)\to\mathbb{B}/\eft(\ebd(x))
-\end{align*}
-of B-frameworks, for any $X\in B_{n+1}$ and any $x\in\tilde{B}_{n+1}$. 
-\end{defn}
-
-\begin{lem}
-When $\mathbb{B}$ is a B0-system, then so is each $\mathbb{B}/X$.
-\end{lem}
-
-\begin{proof}
-Straightforward.
-\end{proof}
-
-\begin{defn}
-A B0-homomorphism $H:\mathbb{B}\to\mathbb{A}$ is a homomorphism of B-frameworks
-for which the diagrams
-\begin{equation*}
-\begin{tikzcd}[column sep=huge]
-\mathbb{B}/X \arrow[r,"H/X"] & \mathbb{A}/H(X) \\
-\mathbb{B}/\eft(X) \arrow[u,"W_X"] \arrow[r,swap,"H/\eft(X)"] & \mathbb{A}/\eft(H(X)) \arrow[u,swap,"W_{H(X)}"]
-\end{tikzcd}
-\end{equation*}
-and
-\begin{equation*}
-\begin{tikzcd}[column sep=huge]
-\mathbb{B}/\ebd(x) \arrow[r,"H/\ebd(x)"] \arrow[d,swap,"S_x"] & \mathbb{A}/\ebd(H(X)) \arrow[d,"S_{\tilde{H}(x)}"] \\
-\mathbb{B}/\eft(\ebd(x)) \arrow[r,swap,"H/\eft(\ebd(x))"] & \mathbb{A}/\eft(\ebd(H(X)))
-\end{tikzcd}
-\end{equation*}
-of homomorphisms of B-frameworks commute for every 
-$X\in B_n$ and $x\in\tilde{B}_n$. The category of B0-systems is denoted
-by $\mathbf{B0sys}$.
-\end{defn}
-
-Now that we have defined the category of B0systems, we can start defining the
-embedding into pre-E-systems.
-
-\begin{defn}
-Let $\mathbb{B}$ be a B-framework. Then $\mathbb{B}$ determines
-a stratified category $U(\mathbb{B})$. Also, any homomorphism of B-frameworks
-determines a stratified functor.
-\end{defn}
-
-\begin{constr}
-The set of objects of $U(\mathbb{B})$ is taken to be the set
-$\bigsqcup_{(n\in\mathbb{N})}B_n$. Now we define
-$(k,X)\leq (n,Y)$ if and only if $n\leq k$ and $\eft[n]^{k-n}(X)\jdeq Y$.
-In other words, we only have $(n+m,X)\leq (n,\eft^m(X))$. It is immediate from
-this definition that for any $X\in B_{n+1}$ there is a unique $\Gamma\in
-B_n$ such that $(n+1,X)\leq (n,\Gamma)$, so $U(\mathbb{B})$ is stratified.
-
-To show that the functor $U(H):U(\mathbb{B})\to U(\mathbb{A})$ is stratified for
-any homomorphism $H:\mathbb{B}\to\mathbb{A}$ of B-frameworks, note that
-by \autoref{lem:strat_full} it suffices to show that each $U(H)$ 
-lifts factorizations.
-Let $H:\mathbb{B}\to\mathbb{A}$ be a homomorphism of B-frameworks, and let
-$(m,X):(n+m,X)\leq (n,\Gamma)$. Then the category $\mathbf{fact}_{(m,X)}$ of
-factorizations of $(m,X)$ is a finite linear order with
-$m+1$ elements, and so is the category $\mathbf{fact}_{(m,H(X))}$. 
-\end{constr}
-
-\begin{rmk} 
-It is in fact the case that $U(\mathbb{B})$ is the free category generated by the
-graph with $\bigsqcup_{(n\in\mathbb{N})}B_n$ as the set of vertices, and
-for each object of the form $(n+1,X)$ an edge to $(n,\eft[n](X))$. 
-Essentially all $U$ does, is forgetting about the term structure 
-$\tilde{B}$ of $\mathbb{B}$.
-
-Furthermore, it is useful to note that for any $X\in B_n$, we get an 
-isomorphism $U(\mathbb{B}/X)\cong  U(\mathbb{B})/(n,X)$ of rooted trees,
-natural in $X$. Therefore we will usually not distinguish between $U(\mathbb{B})/X$
-and $U(\mathbb{B})/(n,X)$. 
-\end{rmk}
-
-\begin{defn}
-Let $\mathbb{B}$ be a B0-system. Then $U(\mathbb{B})$ can be given the structure
-of a pre-E-system. We write $\mathcal{E}(\mathbb{B})$ for the resulting
-pre-E-system. Thus we get a functor $\mathcal{E}:\mathbf{B0sys}\to\mathbf{Esys_s}$.
-\end{defn}
-
-\begin{constr}
-We begin by simultaneously defining the term structure and the pre-substitution
-structure on $U(\mathbb{B})$ by induction on the natural numbers. More precisely,
-for any $m\in\mathbb{N}$ we will define, for any $(n+m,X)\leq (n,\Gamma)$ a set
-$T(m,X)$ and for any $x\in T(m,X)$ a homomorphism 
-$S_x:\mathbb{B}/X\to\mathbb{B}/\Gamma$ of B-frameworks. Then it follows
-immediately that we get stratified functors $U(S_x):U(\mathbb{B})/X\to U(\mathbb{B})/\Gamma$
-defining the pre-substitution structure of $\mathcal{E}(\mathbb{B})$.
-
-For $(n,X)\in U(\mathbb{B})$, we define $T(0,X)\defeq
-\{\pt[X]\}$ and $S_{\pt[X]}\defeq \catid{\mathbb{B}/X}$. We also define for
-$(n+1,X)\leq (n,\Gamma)$ the set $T(1,X)\jdeq\{x\in B_{n+1}\mid \ebd(x)\jdeq X\}$.
-Note that for $x\in T(1,X)$, the homomorphism $S_x:\mathbb{B}/X \to\mathbb{B}/\Gamma$
-is already assumed to exist, because $\mathbb{B}$ is a B0-system.
-
-For the inductive step, suppose that for $m\in\mathbb{N}$ we have sets $T(m,X)$
-for any $(n+m,X)\leq (n,\Gamma)$ and homomorphisms $S_x:\mathbb{B}/X \to \mathbb{B}/\Gamma$
-of B-frameworks, for any $x\in T(m,X)$. Then we define
-\begin{equation*}
-T(m+1,X)\defeq\bigsqcup_{x\in T(m,\eft(X))} T(1,S_x(X)).
-\end{equation*}
-For $(x,u)\in T(m+1,X)$ we define the homomorphism $S_{(x,u)}:\mathbb{B}/X\to
-\mathbb{B}/\Gamma$ as
-\begin{equation*}
-S_{(x,u)}\defeq S_u\circ S_x/X.
-\end{equation*}
-It remains to define a pre-weakening structure on $U(\mathbb{B})$. To do this,
-it suffices to define homomorphisms $W_{(m,X)}:\mathbb{B}/\Gamma \to\mathbb{B}/X$
-of B-frameworks, for any $(n+m,X)\leq (n,\Gamma)$. We define
-\begin{equation*}
-W_{(m,X)}\defeq W_X\circ\cdots\circ W_{\eft^{m-1}(X)}.\qedhere
-\end{equation*}
-\end{constr}
-
-\begin{thm}
-The functor $\mathcal{E}$ is full and faithful.
-\end{thm}
-
-\begin{proof}
-We have to show that the map $\mathrm{hom}(\mathbb{B},\mathbb{A})\to
-\mathrm{hom}(\mathcal{E}(\mathbb{B}),\mathcal{E}(\mathbb{A}))$ is a bijection
-for any two B0-systems $\mathbb{B}$ and $\mathbb{A}$.
-
-To show that this map is injective, suppose $F,G:\mathbb{B}\to\mathbb{A}$ are 
-B0-homomorphisms such that $\mathcal{E}(F)\jdeq\mathcal{E}(G)$. Then we have
-$F_n(X)\jdeq\mathcal{E}(F)(n,X)\jdeq \mathcal{E}(G)(n,X)\jdeq G_n(X)$ for each
-object $x\in B_n$. Also, we have $\tilde{F}_n(x)\jdeq T_{\mathcal{E}(F)}(n,x)
-\jdeq T_{\mathcal{E}(G)}(n,x)\jdeq \tilde{G}_n(x)$ for each
-$x\in\tilde{B}_n$. This shows that $F\jdeq G$, so we conclude that $\mathcal{E}$
-is faithful.
-
-To show that the map $\mathrm{hom}(\mathbb{B},\mathbb{A})\to
-\mathrm{hom}(\mathcal{E}(\mathbb{B}),\mathcal{E}(\mathbb{A}))$ 
-is surjective, let $H:\mathcal{E}(\mathbb{B})\to
-\mathcal{E}(\mathbb{A})$. Since both $\mathcal{E}(\mathbb{B})$ and
-$\mathcal{E}(\mathbb{A})$ are stratified, we get from
-\autoref{lem:strat_full} that $H$ is a stratified pre-E-homomorphism. 
-Because $H$ is stratified, we get maps $H_n:B_n\to A_n$ for any $n\in\mathbb{N}$.
-The term structure of $H$ gives maps $\tilde{H}_n:
-\tilde{B}_n\to\tilde{A}_n$. This defines a B0-homomorphism since $H$ 
-preserves the pre-weakening and pre-substitution structures of $\mathcal{E}(\mathbb{B})$.
-It is immediate from the definition of $H_n$ and
-$\tilde{H}_n$ that this B0-homomorphism is mapped back to $H$, so we conclude
-that $\mathcal{E}$ is full.
-\end{proof}
-
-\subsection{B-systems are stratified E-systems}
-There are two properties which hold for pre-E-systems of the form $\mathcal{E}
-(\mathbb{B})$ for some B0-system $\mathbb{B}$, 
-which stand in the way of proving that the category of stratified
-pre-E-systems is equivalent to the category of B0-systems. The first is that
-for any $A\in U(\mathbb{B})/\Gamma$ and $P\in U(\mathbb{B})/\ctxext{\Gamma}{A}$,
-we have an isomorphism $T(\ctxext{A}{P})\cong\bigsqcup_{x\in T(A)}T(S_x(P))$.
-The second, which depends on the first, is that for any $(x,u)\in T(\ctxext{A}{P})$
-we have $S(x,u)\jdeq S_u\circ S_x/P$. In other words, both the term structure
-and the pre-substitution structure of $\mathcal{E}(\mathbb{B})$ are compatible
-with the categorical structure of the underlying category $U(\mathbb{B})$.
-However, to prove the corresponding facts for E-systems, we needed the full
-set of conditions on unital E-systems: to prove that the term structure of
-$U(\mathbb{B})$ is compatible with composition we needed units, and to prove 
-that the operations $S_x$ are compatible with the pairing function we needed
-that they are E-homomorphisms.
-
-Presumably, we can assume these compatibility conditions on pre-E-systems in order
-to prove the desired equivalence of B0-systems and stratified pre-E-systems. However,
-that seems somewhat unnatural, especially given that they automatically hold
-for E-systems. Thus, for the remainder of this section we set out to show that
-the category of unital B-systems is equivalent to the category of stratified 
-unital E-systems.
-
-\begin{defn}
-A unital B0-system is a B0-system with maps
-\begin{equation*}
-\delta_n : B_{n+1}\to \tilde{B}_{n+2}.
-\end{equation*}
-such that $\ebd(\delta_n(X))\jdeq W_{X}(X)$ for any $X\in B_n$. 
-A B0-homomorphism $H$ is said to be unital if one has
-\begin{equation*}
-H(\delta(X))\jdeq\delta(H(X))
-\end{equation*}
-for any $X\in B_n$.  
-\end{defn}
-
-\begin{rmk}
-Voevodsky showed in \cite{VV_B-systems} that the category of unital B0-systems
-with unital B0-homomorphisms between them is not a full
-subcategory of $\mathbf{B0sys}$. 
-\end{rmk}
-
-\begin{defn}
-A \emph{(unital) B-system} is a B0-system for which the following conditions
-hold:
-\begin{enumerate}
-\item Every $W_X$ and $S_x$ is a unital B0-homomorphism.
-\item For every $x\in \tilde{B}_{n+1}$ one has $S_x\circ W_{\ebd(x)}\jdeq
-\catid{\mathbb{B}/\eft(\ebd(x))}$. 
-\item For every $x\in\tilde{B}_{n+1}$ one has $S_x(\delta(\ebd(x)))\jdeq x$.
-\item For every $X\in B_{n+1}$ one has $S_{\delta(X)}\circ W_X/X\jdeq
-\catid{\mathbb{B}/X}$. 
-\end{enumerate}
-\emph{B-homomorphisms} are unital B0-homomorphisms between B-systems.
-We denote the category of B-systems by $\mathbf{Bsys}$. 
-\end{defn}
-
-\begin{lem}
-For any non-unital B-system $\mathbb{B}$, $\mathcal{E}(\mathbb{B})$ is a 
-stratified non-unital E-system. 
-\end{lem}
-
-\begin{proof}
-Straightforward.
-\end{proof}
-
-\begin{defn}
-For any unital B-system $\mathbb{B}$, $\mathcal{E}(\mathbb{B})$ has a
-projection structure. We thus obtain a functor 
-$\mathcal{E}:\mathbf{Bsys}\to\mathbf{Esys}$ from unital B-systems to E-systems.
-\end{defn}
-
-\begin{constr}
-We first need to extend the functor $\mathcal{E}:\mathbf{B0sys}\to\mathbf{Esys_p}$
-from non-unital B0-systems to non-unital pre-E-systems, to a functor from
-unital B-systems to E-systems.
-
-We need to give $\mathcal{E}(\mathbb{B})$ the structure of a unital
-pre-E-system. We construct by induction on $m\in \mathbb{N}$, a term of type 
-$T(m,W_{(m,X)}(X))$ with the property that that 
-\begin{equation*}
-S_{\tfid{(m,X)}}\circ W_{(m,X)}/X\jdeq\catid{U(\mathbb{B})/X},
-\end{equation*} 
-for every $(n+m,X)\leq (n,\Gamma)$ by induction on $m$. The extra
-condition is part of the inductive construction, because we need it on the way.
-
-For $m\jdeq 0$ we have but one
-choice, and for $m\jdeq 1$ we can directly use the unital structure of
-$\mathbb{B}$. For the inductive step, recall that
-\begin{equation*}
-T(m+1,W_{(m+1,X)}(X))\jdeq\bigsqcup_{x\in T(m,W_{(m+1,X)}(\eft(X)))}T(1,S_x(W_{(m+1,X)}(X))).
-\end{equation*}
-We have the term $x\defeq \tilde{W}_X(\tfid{(m,\eft(X))})\in T(m,W_{(m+1,X)}(\eft(X)))$.
-Now we may observe that 
-\begin{align*}
-S_x(W_{(m+1,X)}(X))
-  & \jdeq 
-S_{\tilde{W}_X(\tfid{(m,\eft(X))})}(W_{(m+1,X)}(X))\\
-  & \jdeq 
-S_{\tilde{W}_X(\tfid{(m,\eft(X))})}(W_X(W_{(m,\eft(X))}(X))) \\
-  & \jdeq
-W_{X}(S_{\tfid{(m,\eft(X))}}(W_{(m,\eft(X))}(X))) \\
-  & \jdeq 
-W_{X}(X).
-\end{align*}
-Therefore, we see that $T(1,S_x(W_{(m+1,X)}(X)))\jdeq T(1,W_X(X))$, where we can
-choose the term $\delta(X)$. Thus, we define
-\begin{equation*}
-\tfid{(m+1,X)}\defeq (\tilde{W}_X(\tfid{(m,\eft(X))}),\delta(X))
-\end{equation*} 
-for $m\geq 1$. Now we need to check that $S_{\tfid{(m+1,X)}}\circ 
-W_{(m+1,X)}/X\jdeq \catid{U(\mathbb{B})/X}$. This is indeed the case:
-\begin{align*}
-S_{\tfid{(m+1,X)}}\circ W_{(m+1,X)}/X
-  & \jdeq
-S_{\delta(X)}\circ S_{\tilde{W}_X(\tfid{(m,\eft(X))})}\circ W_{(m+1,X)}/X \\
-  & \jdeq
-S_{\delta(X)}\circ W_X/X \\
-  & \jdeq
-\catid{U(\mathbb{B})/X}.
-\end{align*}
-This finishes the definition of the pre-projection structure on $\mathcal{E}(\mathbb{B})$.
-Straightforward proves by induction give that each $W_{(m,X)}$ and $S_x$ is a 
-pre-projection homomorphism and that $S_x(\tfid{(m,X)})\jdeq x$, for any 
-$x\in T(m,X)$.
-\end{constr}
-
-\begin{thm}
-The functor $\mathcal{E}:\mathbf{Bsys}\to\mathbf{Esys}$ lifts to an equivalence
-to the subcategory $\mathbf{Esys_s}$ of stratified E-systems: i.e.~we have a
-commuting diagram
-\begin{equation*}
-\begin{tikzcd}
-& \mathbf{Esys_s} \arrow[dr,"\subseteq"]\\
-\mathbf{Bsys} \arrow[ur,"\simeq"] \arrow[rr,swap,"\mathcal{E}"] & & \mathbf{Esys}
-\end{tikzcd}
-\end{equation*}
-\end{thm}
-
-\begin{proof}
-We already have shown that $\mathcal{E}$ factors through the stratified
-E-systems and that it is full and faithful, so it remains
-to show that every stratified E-system is in the image of $\mathcal{E}$. In
-other words, for every stratified E-system $\mathbb{E}$ we need to construct
-a B0-system $\mathbb{B}$ such that $\mathcal{E}(\mathbb{B})\cong\mathbb{E}$.
-
-Let $\mathbb{E}$ be a stratified E-system. From the stratification functor
-$L:\mathbb{C}_{\mathbb{E}}\to (\mathbb{N},\geq)$ we obtain the pre-image
-presheaf $B(\mathbb{E}):\op{(\mathbb{N},\leq)}\to\mathbf{Set}$ given by
-\begin{align*}
-B(\mathbb{E})_n & \defeq \{X\in\mathrm{Ob}(\mathbb{C}_{\mathbb{E}})\mid L(X)\jdeq n\}
-\end{align*}
-and $\eft(X)\defeq\mathrm{cod}(p_X)$. We define the term structure of $B(\mathbb{E})_n$
-to be given by
-\begin{equation*}
-\tilde{B}(\mathbb{E})_n \defeq \bigsqcup_{X\in B(\mathbb{E})_n} T(p_X)
-\end{equation*}
-with the projection map $\ebd[n]\defeq\pi_1:\tilde{B}(\mathbb{E})_n\to B(\mathbb{E})_n$.
-This defines the B-framework $B(\mathbb{E})$, and we obtain a functor
-$B:\mathbf{Esys_s}\to\mathbf{B0sys}$. It is immediate that we get isomorphisms
-$U(B(\mathbb{E}))\cong\cat{C}_{\mathbb{E}}$, natural in $\mathbb{E}$. Also, we
-get natural isomorphisms $B(\mathbb{E}/\Gamma)\cong B(\mathbb{E})/\Gamma$.
-
-The weakening and substitution structures on
-$B(\mathbb{E})$ are given by $W_X\defeq B(W_{p_X})$ and $S_{(X,x)}\defeq B(S_x)$,
-respectively. Since these functors are assumed to be stratified, they define
-homomorphisms of B-frameworks. Because a weakening system of an E-system is
-required to be functorial, i.e.~$W_{\ctxext{A}{P}}\jdeq W_P\circ W_A$ and
-$W_{\catid{\Gamma}}\jdeq\catid{\cat{C}_{\mathbb{E}}/\Gamma}$, it follows
-immediately that the weakening structure on $U(B(\mathbb{E}))$ is isomorphic to
-that of $\mathbb{E}$. 
-\end{proof}
-
-\begin{comment}
-In the following definition $\Delta(X)$ denotes the discrete category on the set
-$X$ and $\mathbf{2}$ is the category $s\to t$.
-\begin{defn}
-Define $\mathbb{S}$ to be the pushout
-\begin{equation*}
-\begin{tikzcd}[column sep=large]
-\Delta(\mathbb{N}) \arrow[r,"{n\mapsto(n,s)}"] \arrow[d,swap,"i"] & \Delta(\mathbb{N})\times\mathbf{2} \arrow[d] \\
-(\mathbb{N},\leq) \arrow[r] & \mathbb{S}
-\end{tikzcd}
-\end{equation*}
-in $\mathbf{Cat}$.
-\end{defn}
-
-Note that for any $n\in\mathbb{N}$, we have the map $S^n := (m\mapsto m+n)$ which
-is monotone. By the universal property of pushouts, we obtain a functor $\sigma^n:\mathbb{S}\to\mathbb{S}$
-given by
-\begin{equation*}
-\begin{tikzcd}[column sep=huge]
-\Delta(\mathbb{N}) 
-  \arrow[r,"{n\mapsto(n,s)}"] 
-  \arrow[d,swap,"i"] 
-  & 
-\Delta(\mathbb{N})\times\mathbf{2} 
-  \arrow[d]
-  \arrow[r,"{\Delta(S^n)\times\catid{\mathbf{2}}}"]
-  &
-\Delta(\mathbb{N})\times\mathbf{2}
-  \arrow[dd]
-  \\
-(\mathbb{N},\leq) 
-  \arrow[r] 
-  \arrow[d,swap,"{(S^n,\leq)}"]
-  & 
-\mathbb{S}
-  \arrow[dr,densely dotted,yshift=.5ex,"\sigma^n"]
-  \\
-(\mathbb{N},\leq)
-  \arrow[rr] & &
-\mathbb{S}
-\end{tikzcd}
-\end{equation*}
-
-\begin{defn}
-A non-unital pre-B0-system is a presheaf $B$ on $\mathbb{S}$ such that $B(0)$ is
-a singleton, and the following additional structure:
-\begin{enumerate}
-\item For any $n\in\mathbb{N}$ and $X\in B(n)$, write $B/X$ for the sub-presheaf
-of $B\circ\sigma^n$ given by $B/X(0)=\{X\}$ and $B/X(n+1)=B_{n\leq n+1}^{-1}(B/X(n))$. 
-We require that for any $n\in\mathbb{N}$ and any $A\in B(n+1)$, there is a 
-natural transformation $W_A:B/(B_{n\leq n+1}(A))\Rightarrow B/A$.
-\end{enumerate}
-\end{defn}
-\end{comment}
-
-\subsection{From D-systems to E-systems}
-
-We use `D-system' to refer to Dybjer's categories with families. We recall
-Dybjer's definition. Note that $\mathbf{Fam}$ is the category of families of
-sets, of which the objects consist of a set $A$ and a family $\{B(x)\}_{x\in A}$
-over $A$. Although a terminal object is assumed to exist, not every object 
-of a D-system is `fibrant', i.e.~it is
-not the case that every $X\to 1$ is a composition of projections.
-We could only hope that there is a functor from D-sytems to E-systems without
-terminal objects. We write $\mathbf{Esys}_{\circ}$ for the category of E-systems
-without terminal objects.
-
-\begin{defn}
-A category is said to be locally stratified if each of its slice categories
-are stratified.
-\end{defn}
-
-\begin{defn}
-A \emph{setup for categories with families} consists of
-\begin{enumerate}
-\item A category $\cat{C}$ with a terminal object $1$.
-\item A presheaf $\mathrm{Ty}:\op{\cat{C}}\to\mathbf{Set}$. We will write
-$A[\gamma]$ for $\mathrm{Ty}(\gamma,A)$. 
-\item A commuting triangle
-\begin{equation*}
-\begin{tikzcd}
-\int_\cat{C}\mathrm{Ty} \arrow[dr,swap,"\pi"] \arrow[rr,"p"] & & \cat{C}^{\to} \arrow[dl,"\mathrm{cod}"] \\
-& \cat{C}
-\end{tikzcd}
-\end{equation*} 
-We will write $p_A:\ctxext{\Gamma}{A}\to\Gamma$
-for the morphism $U(\Gamma,A)$ into $\Gamma$, for each object $\Gamma\in\cat{C}$ and
-$A\in\mathrm{Ty}(\Gamma)$.
-\end{enumerate}
-\end{defn}
-
-\begin{defn}
-A \emph{category with families} consists of a setup with
-\begin{enumerate}
-\item[(iv)] a preasheaf $\mathrm{Tm}:\op{(\int_\cat{C}\mathrm{Ty})}\to\mathbf{Set}$,
-\item[(v)] a term $q_A\in\mathrm{Tm}(\ctxext{\Gamma}{A},A[p_A])$ for every
-$A\in\mathrm{Ty}(\Gamma)$, 
-\end{enumerate}
-such that for every $\gamma:\Delta\to\Gamma$ in $\cat{C}$ and 
-$a\in \mathrm{Tm}(\Delta,A[\gamma])$ there is a unique morphism 
-$\theta:\Delta\to \ctxext{\Gamma}{A}$ such that $p_A\circ\theta\jdeq\gamma$ and 
-$q_A[\theta]\jdeq a$.
-\end{defn}
-
-\begin{lem}
-Let $\mathbb{D}$ be a category with families. Then there are bijections
-\begin{equation*}
-\mathrm{Tm}(\Delta,A[\gamma])\cong\{\theta:\Delta\to\ctxext{\Gamma}{A}\mid p_A\circ\theta\jdeq \gamma\}
-\end{equation*}
-natural in $\gamma$ and $A$.
-\end{lem}
-
-\begin{proof}
-Let $\theta:\Delta\to\ctxext{\Gamma}{A}$ be such that $p_A\circ\theta\jdeq\gamma$. 
-Then $\subst{\theta}{q_A}$ is a term
-of $\subst{\theta}{{p_A}{A}}$. By the functoriality of $\mathrm{Ty}$, we have
-$\subst{\theta}{{p_A}{A}}\jdeq \subst{p_A\circ\theta}{A}\jdeq\subst{\gamma}{A}$. 
-So we obtain $\subst{\theta}{q_A}\in\mathrm{Tm}(\Delta,\subst{\gamma}{A})$. Now
-the uniqueness of $\theta$ implies that the map $\theta\mapsto\subst{\theta}{q_A}$
-is a bijection. 
-\end{proof}
-
-The previous lemma renders the presheaf $\mathrm{Tm}$ superfluous. Therefore,
-we define D-systems as follows.
-
-\begin{defn}
-A \emph{D-system} consists of a setup for categories with families
-such that for every $\gamma:\Delta\to\Gamma$ and $A\in\mathrm{Ty}(\Gamma)$, there
-is a bijection
-\begin{equation*}
-\{\eta:\Delta\to\ctxext{\Delta}{\subst{\gamma}{A}}\mid p_{\subst{\gamma}{A}}\circ\eta\jdeq\catid{\Delta}\}
-  \cong
-\{\theta:\Delta\to\ctxext{\Gamma}{A}\mid p_A\circ\theta\jdeq\gamma\}
-\end{equation*}
-natural in $\gamma$ and $A$.
-\end{defn}
-
-\begin{thm}
-A category with families is the same thing as a D-system.
-\end{thm}
-
-\begin{defn}
-There is a functor $U$ from D-systems to locally stratified categories.
-\end{defn}
-
-\begin{constr}
-Let $\mathbb{D}$ be a D-system with underlying category $\cat{C}$. We define
-$U(\mathbb{D})$ to be the category with the same objects as $\cat{C}$, where the
-morphisms are generated by the projections $p_A:\ctxext{\Gamma}{A}\to\Gamma$.
-
-We have to prove that this category is locally stratified. Let $\Gamma\in\cat{C}$. 
-\end{constr}
-
-\begin{defn}
-For any D-system $\mathbb{D}$, $U(\mathbb{D})$ can be given the structure of
-an E-system without terminal object. Thus, we get a functor
-$\mathbf{Dsys}\to\mathbf{Esys}_{\circ}$
-\end{defn}
-
-\begin{constr}
-We first need to define a term structure on $U(\mathbb{D})$. 
-Next, we define a weakening structure on $\cat{C}_p$. Since the morphisms of
-$\cat{C}_p$ are composites of projections, we will define
-\begin{equation*}
-W_{p_{A_n}\circ\cdots\circ p_{A_1}}\defeq W_{p_{A_n}}\circ\cdots W_{p_{A_1}}.
-\end{equation*}
-Therefore, we only have to define functors 
-$W_A:\cat{C}_p/\Gamma\to\cat{C}_p/\ctxext{\Gamma}{A}$ with term structure for every 
-$A\in\mathrm{Ty}(\Gamma)$. Since morphisms are composites of projections, we
-can define this functor by induction. Let $W_A(\catid{\Gamma})\defeq
-\catid{\ctxext{\Gamma}{A}}$ and $W_A(p_Q\circ f)\defeq$
-\end{constr}
-
-\begin{thm}
-There is a full and faithful functor from the category of D-systems to the
-category of E-systems.
-\end{thm}
-
-\section{An alternative characterization of E-systems}
-
-\begin{defn}
-An E'-system consists of a category $\cat{C}$ with a subcategory $\cat{F}$ with the
-same objects as $\cat{C}$, with
-the following structure
-\begin{enumerate}
-\item $\cat{F}$ has a terminal object.
-\item For any $f:\Delta\to\Gamma$ in $\cat{C}$, a functor $f_\ast:\cat{F}/\Gamma\to
-\cat{F}/\Delta$ %and
-%for any $Q\in\cat{F}/\ctxext{\Gamma}{B}$ and any section $g$ of $Q$ a section
-%$f_\ast(g)$ of $f_\ast(Q)$, 
-so that $f_\ast$ is contravariant in $f$ and preserves the terminal object. 
-\item For any $A\in\cat{F}/\Gamma$ and $f:\Delta\to\Gamma$, factorizations of
-$f$ through $A$ are naturally the same thing as sections of
-$f_\ast(A)$, i.e. there is an isomorphism
-\begin{equation*}
-\zeta_{A,f}%\defeq\theta\mapsto f_\ast(\theta)
-:\{\theta:\Delta\to\ctxext{\Gamma}{A}\mid A\circ\theta\jdeq f\}
-  \cong
-\{\eta:\Delta\to\ctxext{\Delta}{(f_\ast(A))}\mid f_\ast(A)\circ\eta\jdeq\catid{\Delta}\}
-\end{equation*}
-in the presheaf category 
-$\mathrm{Psh}(\op{(\cat{F}/\Gamma)}\times(\cat{C}/\Gamma))$. 
-\end{enumerate}
-\end{defn}
-
-\begin{rmk}
-Explicitly, the
-naturality condition states that for any $P\in\cat{F}/\ctxext{\Gamma}{A}$,
-$g:\greek{E}\to\Delta$ and $\theta:\Delta\to\ctxext{{\Gamma}{A}}{P}$ such that
-$P\circ\theta\jdeq f$, we have the equality
-\begin{equation*}
-\zeta_{A,f\circ g}(P\circ\theta\circ g)\jdeq f^\ast(P)\circ\zeta_{P,f}(\theta)\circ g.
-\end{equation*}
-\end{rmk}
-
-We can give the category $\cat{F}$ a term structure as follows:
-
-\begin{defn}
-For any E'-system $\mathbb{E}$, the category $\cat{F}$ has a term structure
-where the terms of $A$ are sections of $A$, i.e.
-\begin{equation*}
-T(A)\defeq\{\theta:\Gamma\to\ctxext{\Gamma}{A}\mid A\circ\theta\jdeq\catid{\Gamma}\}.
-\end{equation*}
-\end{defn}
-
-\begin{lem}
-Let $\mathbb{E}$ be an E'-system. Then there is a function $T(A)\to T(f_\ast(A))$
-for any $f:\Delta\to\Gamma$ in $\cat{C}$ and $A\in\cat{F}/\Gamma$, contravariant
-in $f$. 
-\end{lem}
-
-\begin{proof}
-Using the natural isomorphism $\zeta_{A,f}$, we see that
-\begin{align*}
-T(f_\ast(A)) & \jdeq \{\eta : \Delta\to\ctxext{\Delta}{f_\ast(A)}\mid f_\ast(A)\circ\eta\jdeq\catid{\Delta}\}
-  \\
-  & \cong
-\{\theta : \Delta\to \ctxext{\Gamma}{A}\mid A\circ\theta\jdeq f\}.
-\end{align*}
-Therefore, we have the function $x\mapsto \zeta_{A,f}(x\circ f)$ from $T(A)$ to
-$T(f_\ast(A))$. 
-\end{proof}
-
-\begin{defn}
-Consider $P\in\cat{F}/\ctxext{\Gamma}{A}$ and $f:\Delta\to\Gamma$,
-$\theta_1:\Delta\to\ctxext{\Gamma}{A}$ such that $A\circ\theta_1\jdeq f$, 
-and $\theta_2:\Delta\to\ctxext{{\Gamma}{A}}{P}$ such that $P\circ\theta_2\jdeq\theta_1$.  
-\end{defn}
-
-\begin{defn}
-For any $P\in\cat{F}/\ctxext{\Gamma}{A}$ and $f:\Delta\to\Gamma$ 
-there is a morphism $\tilde{f}:\ctxext{\Delta}{f_\ast(A)}\to\ctxext{\Gamma}{A}$
-with the property that $A\circ\tilde{f}\jdeq f\circ f_\ast(A)$, and
-we have isomorphisms
-\begin{align*}
-\{\theta:\ctxext{\Delta}{f_\ast(A)}\to\ctxext{{\Gamma}{A}}{P}\mid P\circ\theta\jdeq \tilde{f}\}
-&  \cong
-\{\eta:\ctxext{\Delta}{f_\ast(A)}\to\ctxext{{\Delta}{(f_\ast(A))}}{f_\ast(P)}\mid f_\ast(P)\circ\eta\jdeq\catid{\ctxext{\Delta}{f_\ast(A)}}\}
-\end{align*}
-\end{defn}
-
-\begin{constr}
-\begin{align*}
-& \{ \theta : \ctxext{\Delta}{f_\ast(A)}\to\ctxext{\Gamma}{A}\mid A\circ\theta\jdeq f\circ f_\ast(A) \} \\
-  & \cong
-\{ \eta : \ctxext{\Delta}{f_\ast(A)}\to\ctxext{{\Delta}{f_\ast(A)}}{(f\circ f_\ast(A))_\ast(A)}
-  \mid (f\circ f_\ast(A))_\ast(A)\circ\eta\jdeq\catid{\ctxext{\Delta}{f_\ast(A)}}\} \\
-  & \jdeq
-\{ \eta : \ctxext{\Delta}{f_\ast(A)}\to\ctxext{{\Delta}{f_\ast(A)}}{(f_\ast(A))_\ast(f_\ast(A))}
-  \mid (f\circ f_\ast(A))_\ast(A)\circ\eta\jdeq\catid{\ctxext{\Delta}{f_\ast(A)}}\} \\
-  & \jdeq
-T(W_{f_\ast(A)}(f_\ast(A))).
-\end{align*}
-Since we have $\tfid{f_\ast(A)}\in T(W_{f_\ast(A)}(f_\ast(A)))$, we obtain 
-\begin{equation*}
-\tilde{f}\defeq \zeta_{A,f\circ f_\ast(A)}^{-1}(\tfid{f_\ast(A)}):\ctxext{\Delta}{f_\ast(A)}\to\ctxext{\Gamma}{A}.
-\end{equation*}
-It remains to prove the asserted isomorphism. 
-\begin{align*}
-\{\theta:\ctxext{\Delta}{f_\ast(A)}\to\ctxext{{\Gamma}{A}}{P}\mid P\circ\theta\jdeq \tilde{f}\}
-&  \cong
-\{\eta : \ctxext{\Delta}{f_\ast(A)}\to\ctxext{{\Delta}{f_\ast(A)}}{\tilde{f}_\ast(P)} \mid
-  \tilde{f}_\ast(P)\circ\eta\jdeq\catid{}\}
-  \\
-  & \cong
-\{\eta:\ctxext{\Delta}{f_\ast(A)}\to\ctxext{{\Delta}{(f_\ast(A))}}{f_\ast(P)}\mid f_\ast(P)\circ\eta\jdeq\catid{\ctxext{\Delta}{f_\ast(A)}}\}
-\end{align*}
-\end{constr}
-
-\begin{lem}
-Let $\mathbb{E}$ be an E'-system. Then there is a function $T(P)\to T(f_\ast(P))$
-for any $f:\Delta\to\Gamma$ in $\cat{C}$ and $P\in\cat{F}/\ctxext{\Gamma}{A}$, contravariant
-in $f$. 
-\end{lem}
-
-\begin{proof}
-Using the natural isomorphism $\zeta_{A,f}$, we see that
-\begin{align*}
-T(f_\ast(P)) & \jdeq \{\eta : \ctxext{\Delta}{f_\ast(A)}\to\ctxext{{\Delta}{f_\ast(A)}}{f_\ast(P)}\mid f_\ast(P)\circ\eta\jdeq\catid{\ctxext{\Delta}{f_\ast(A)}}\}.
-\intertext{By all means, this should be isomorphic to}
-  & \cong
-\{\theta : \ctxext{\Delta}{f_\ast(A)}\to \ctxext{{\Gamma}{A}}{P}\mid P\circ\theta\jdeq \tilde{f}\}.
-\end{align*}
-Therefore, we have the function $x\mapsto \zeta_{A,f}(x\circ f)$ from $T(A)$ to
-$T(f_\ast(A))$. 
-\end{proof}
-
-\begin{thm}
-E-systems are the same things as E'-systems.
-\end{thm}
-
-\begin{proof}
-Given an E-system $\mathbb{E}$, we have the category $\cat{C}$ of internal
-morphisms in the empty context, with the subcategory $\cat{F}$ of projections $\cprojfstf{A}{P}$. 
-The functors $f_\ast:\cat{F}/\Gamma\to\cat{F}/\Delta$ are contravariant in
-$f$, so we have to verify whether we have the stated isomorphisms.
-
-Note first that for any $A\in\cat{F}/\Gamma$, we have
-$W_A\defeq A_\ast:\cat{F}/\Gamma\to\cat{F}/\ctxext{\Gamma}{A}$. Then $W_A$ is
-contravariant in $A$ and that it preserves the terminal object. We give $W_A$
-an action on terms by defining
-\begin{equation*}
-W_A(\theta)\defeq
-\end{equation*}
-for any $\theta\in T(Q)$ and $Q\in\cat{F}/\ctxext{\Gamma}{B}$.
-
-Thus, the $W_A$
-assemble a pre-weakening structure.
-
-Note that for any $A\in\cat{F}/\Gamma$, we have
-\begin{align*}
-T(W_A(A)) & \jdeq \{\eta:\ctxext{\Gamma}{A}\to\ctxext{{\Gamma}{A}}{W_A(A)}\mid W_A(A)\circ\eta\jdeq\catid{\ctxext{\Gamma}{A}}\} \\
-  & \cong
-\{\theta:\ctxext{\Gamma}{A}\to\ctxext{\Gamma}{A}\mid A\circ\theta\jdeq A\},
-\end{align*}
-so we find $\idtm{A}\defeq\zeta_{A,A}(\catid{\ctxext{\Gamma}{A}})\in T(W_A(A))$.
-Now let $\theta:\Delta\to\ctxext{\Gamma}{A}$ so that $A\circ\theta\jdeq f$
-for some $f:\Delta\to\Gamma$. Then we have $\zeta_{A,f}(\theta)\in T(f_\ast(A))$.  
-We wish to define $\theta_\ast(\idtm{A}):\Delta\to\ctxext{\Delta}{f_\ast(A)}$
-
-for any $A\in\cat{F}/\Gamma$ and $Q\in\cat{F}/\ctxext{\Gamma}{B}$, we have
-\begin{align*}
-T(B) 
-  & \jdeq 
-\{ \eta : \Gamma \to\ctxext{\Gamma}{B}
-        \mid B\circ\eta\jdeq\catid{\Gamma}\}
-  \\
-T(W_A(B))
-  & \jdeq
-\{ \eta : \ctxext{\Gamma}{A}\to\ctxext{{\Gamma}{A}}{W_A(B)}
-        \mid W_A(B)\circ\eta\jdeq\catid{\ctxext{\Gamma}{A}}\}
-  \\
-  & \cong
-\{ \theta : \ctxext{\Gamma}{A}\to\ctxext{\Gamma}{B}\mid B\circ\theta\jdeq A\}
-\end{align*}
-Therefore, $\eta\mapsto \zeta_{B,A}(\eta\circ A)$ defines a function from
-$T(B)$ to $T(W_A(B))$. 
-\end{proof}
-
-\begin{comment}
-\section{Acknowledgments}
-The research that eventually became the study of E-systems began while analyzing
-the internel models I encountered during investigations of the type theoretical
-descent theorem, in the summer of 2013. In a private conversation during the Conference on Type Theory, 
-Homotopy Theory and Univalent Foundations at the Universitat Aut\`onoma de 
-Barcelona, I presented the first version of the axioms to Richard Garner, who
-told me he found a similar structure  while analyzing the combinatorics of type
-dependency, see \cite{Garner2014}. After that conference, Joachim
-Kock generously offered me to stay for a few months at the UAB. He was intrigued
-by the idea of doing category theory using only sections, and during the sessions
-we had in this period I finalized the inference rules that correspond to the
-axioms of E-systems.
-
-In Paris, in the spring of 2014, Richard Garner introduced me to Vladimir 
-Voevodsky, who had been working on similar problems and wanted to know exactly
-what axioms should hold on B0-systems so that it is in the image of the functor
-from C-systems to B0-systems. In other words, he wanted to know what the axioms
-of B-systems were, and be confident that we had them all.
-
-In August 2015, I started a grad student of Steve Awodey at the Carnegie 
-Mellon University. With Steve we elaborated much more on the fact that the theory
-of E-systems is essentially algebraic, and I wrote down what an E-system is in
-an arbitrary category with finite limits.
-
-The discussions that followed both with Awodey and with Voevodsky, 
-at the CMU, at the Homotopy Type Theory workshop in Oxford (November 2014), and
-while visiting Voevodsky at the Institute for Advanced Study in Princeton (March 2015), 
-essentially shaped E-systems into their current presentation.
-
-I am grateful to all the people mentioned for sharing many valuable insights 
-with me. 
-\end{comment}

### 2015-07-29 main.tex
separation of sections into different files

@@
 between them and finitely many equality axioms. Algebras for this theory are
-called E-systems. We then show that there is a full and faithful functor
-from Voevodsky's B-systems into E-systems, of which the image is the category
-of stratified E-systems. 
+called E-systems. A distinctive feature of the dependent type theory of
+E-systems is that it takes `telescopic families' as the primitive notion of
+families.
+
+We then compare Voevodsky's B-systems, Dybjer's categories
+with families and Jacob's comprehension categories with E-systems. There is
+a full and faithful functor from B-systems to E-systems, of which the image is
+the category of stratified E-systems. To have a functor from categories with
+families to E-systems, we need to impose an additonal coherence condition, which
+makes sure that the substitution operation of a category with families extends 
+to telescopic families. 
 \end{abstract}
@@
 \section{Introduction}
+
+\subsection*{Principles of mathematical fuseki}
+Mathematics is the study of abstract concepts such as space or quantity, and
+when we design a language for such a study, we want to be able to efficiently
+express things about the objects under investigation. The subject matter of
+this paper is a logical framework of dependent type theory.
+
+Pfenning \cite{PfenningLFintro} defines logical frameworks
+to be meta-languages for the formalization of deductive system. Examples of
+logical framework therefore include propositional logic, first-order predicate 
+logic, Russell's ramified type theory, the lambda-calculus, dependent type 
+theory, and so on. By considering models of a logical framework, it becomes
+itself the subject of mathematical study.
+
+It may actually be useful to see logical frameworks themselves
+as objects subject to mathematical investigation.
+
+Logical frameworks for mathematics are not set in stone.
+
+ and neither to they
+remain the same once they are declared.
+
+When designing and maintaining a logical framework
 
@@
 We have chosen not to use common type theoretical notation involving special
-expressions for judgments and for inference rules. One reason is that it has been
-proven to be hard for the uninitiated to read and comprehend such expressions.
-Another is that set theoretic
-notation serves very well for getting the algebraic structure accross, which
-is what we want to focus on.
+expressions for judgments and for inference rules, because we want to focus
+on the algebraic structure of type dependency rather than on its syntax.
 
@@
 
-\input{weakeningcats}
+\input{e-systems}
+
+\input{internal_homs}
+
+\input{compared}
+
+\section{Type constructors in E-systems}
+
+\begin{defn}
+An E-system $\mathbb{E}$ is said to have $\Pi$-types, if 
+\begin{enumerate}
+\item for any 
+$A\in\cat{F}/\Gamma$, there is an E-homomorphism
+\begin{equation*}
+\Pi_A : \mathbb{E}/\ctxext{\Gamma}{A}\to\mathbb{E}/\Gamma
+\end{equation*}
+For terms $g:Q$ in context $\ctxext{{\Gamma}{A}}{P}$, we will usually write
+$\lambda_{A}(g)$ for the term $\Pi_A(g)$ of $\Pi_A(Q)$ in context
+$\ctxext{\Gamma}{\Pi_A(P)}$. Note that when $P$ is empty, then we retrieve the
+usual lambda-abstraction.
+\item for any $A\in\cat{F}/\Gamma$ and $Q\in\cat{F}/\ctxext{{\Gamma}{A}}{P}$, 
+there is a term 
+\begin{equation*}
+\tfev^{A,Q} : \ctxwk{\Pi_A(Q)}{{\Pi_A(P)}{Q}}
+\end{equation*}
+in context $\ctxext{{{{\Gamma}{\Pi_A(P)}}{\Pi_A(Q)}}{\ctxwk{\Pi_A(\ctxext{P}{Q})}{A}}}{\ctxwk{\Pi_A(\ctxext{P}{Q})}{P}}$.
+Note that for each $g:\Pi_A(Q)$, we obtain a term $\subst{g}{\tfev^{A,Q}}: \ctxwk{\Pi_A(P)}{Q}$ in
+context $\ctxext{{\Gamma}{\Pi_A(P)}}{\ctxwk{\Pi_A(P)}{A}}$.
+\item for any $A\in\cat{F}/\Gamma$ there is a term
+\begin{equation*}
+{\tfev^{A,P}} : {\ctxwk{\Pi_A(P)}{P}}
+\end{equation*}
+in context ${\ctxext{{\Gamma}{\Pi_A(P)}}{\ctxwk{\Pi_A(P)}{A}}}$
+\end{enumerate}
+satisfying
+\begin{enumerate}
+\item 
+\begin{enumerate}
+\item For any $P\in\cat{F}/\ctxext{\Gamma}{A}$ one has
+\begin{equation*}
+\Pi_{\ctxext{A}{P}} \jdeq \Pi_A\circ \Pi_P
+\end{equation*}
+\item For any $Q\in\cat{F}/\ctxext{{\Gamma}{A}}{P}$, one has
+\begin{equation*}
+{\evtm{\ctxext{A}{P}}{Q}}
+  \jdeq
+{\subst{\evtm{A}{\mprd{P}{Q}}}{\ctxwk{\mprd{A}{\mprd{P}{Q}}}{\evtm{P}{Q}}}}
+\end{equation*}
+\end{enumerate}
+\item 
+\begin{enumerate}
+\item For any $\Gamma\in\mathbb{E}$, one has
+\begin{equation*}
+\Pi_{\catid{\Gamma}}\jdeq \catid{\mathbb{E}/\Gamma}.
+\end{equation*}
+\item For any $A\in\mathbb{F}/\Gamma$, one has
+\begin{equation*}
+\tfev^{\catid{\Gamma},A} \jdeq \tfid{A}
+\end{equation*}
+\end{enumerate}
+\item for any term $f:P$ in context $\ctxext{\Gamma}{A}$, one has
+\begin{equation*}
+\subst{\lambda_A(f)}{\tfev^{A,P}}\jdeq f.
+\end{equation*}
+\end{enumerate}
+\end{defn}
+
+\begin{rmk}
+The evaluation term determines a term
+\begin{equation*}
+\tilde\tfev^{A,P} : \thom{\ctxwk{A}{\Pi_A(P)}}{P}
+\end{equation*}
+\end{rmk}
+
+\begin{defn}
+A $\Pi$E-homomorphism from $\mathbb{E}$ to $\mathbb{E}'$ is an E-homomorphism
+$H:\mathbb{E}\to\mathbb{E}'$ such that
+\begin{enumerate}
+\item for any $A\in\cat{F}/\Gamma$, one has
+\begin{equation*}
+\begin{tikzcd}[column sep=large]
+\cat{F}/\ctxext{\Gamma}{A} \arrow[d,swap,"\Pi_A"] \arrow[r,"H/\ctxext{\Gamma}{A}"] & \cat{F}'/\ctxext{H(\Gamma)}{H(A)} \arrow[d,"\Pi_{H(A)}"] \\
+\cat{F}/\Gamma \arrow[r,swap,"H/\Gamma"] & \cat{F}'/F(\Gamma)
+\end{tikzcd}
+\end{equation*}
+\item For any $P\in\cat{F}/\ctxext{\Gamma}{A}$, one has
+\begin{equation*}
+H(\tfev^{A,P})\jdeq \tfev^{H(A),H(P)}
+\end{equation*}
+\end{enumerate}
+\end{defn}
 

### 2015-07-29 weakeningcats.tex
separation of sections into different files

@@
 \begin{eg}
-If the underlying pre-category is a group, the axioms of a weakening system
-obtain the following form: There is a group homomorphism $W_x:G\to G$ for
-every $x\in G$, in such a way that $W_e\jdeq \catid{G}$ and $W_{xy}\jdeq W_y\circ W_x$
-for every $x,y\in G$, and for any $x,y\in G$ the diagram
+If the underlying pre-category is a monoid $M$, the axioms of a weakening system
+obtain the following form: There is a group homomorphism $W_x:M\to M$ for
+every $x\in M$, in such a way that $W_e\jdeq \catid{M}$ and $W_{xy}\jdeq W_y\circ W_x$
+for every $x,y\in M$, and for any $x,y\in M$ the diagram
 \begin{equation*}
 \begin{tikzcd}
-G \arrow[r,"W_y"]
+M \arrow[r,"W_y"]
   &
-G 
-  \\
-G \arrow[u,"W_x"]
+M 
+  \\
+M \arrow[u,"W_x"]
   \arrow[r,swap,"W_y"]
   &
-G \arrow[u,swap,"W_{W_y(x)}"]
-\end{tikzcd}
-\end{equation*}
-is required to commute. Thus, a weakening system on a group $G$ consists of
-a group homomorphism $W:\op G\to\mathrm{Aut}(G)$ such that $W_a\circ W_b\jdeq 
-W_{W_a(b)}\circ W_a$. The latter requirement is satisfied if $b\cdot a\jdeq a\cdot W_a(b)$
-for every $a,b\in G$, or equivalently if $[b,a]\jdeq b^{-1}\cdot W_a(b)$.
+M \arrow[u,swap,"W_{W_y(x)}"]
+\end{tikzcd}
+\end{equation*}
+is required to commute. Thus, a weakening system on a monoid $M$ consists of
+a homomorphism $W:\op M\to\mathrm{Aut}(M)$ such that $W_x\circ W_y\jdeq 
+W_{W_x(y)}\circ W_x$. The latter requirement is satisfied if $y\cdot x\jdeq x\cdot W_x(y)$
+for every $x,y\in M$. Group conjugation forms a weakening structure on any group.
 \end{eg}
 
+\begin{comment}
 \begin{lem}
@@
 \end{eg}
-\end{comment}
 
@@
 introduce the equivalent notion of D-systems. After having proved that categories with families
-are the same things as D-systems, we will focus on constructing a functor from
-D-systems to E-systems without empty context. However, due to the existence of pathological examples
-of categories with families, this functor is not going to be faithful. We then
-introduce the notion of well-founded D-systems, and show that the
-functor from D-systems to E'-systems without empty context restricts to an
-embedding from the well-founded D-systems to E'-systems.
+are the same things as D-systems, we will focus on constructing a canonical functor from
+D-systems to E'-systems, in which the category $\cat{F}$ of families has contexts
+as objects and compositions of projections as morphisms. There is no reason why
+this category should possess a terminal object, so we naturally have to retreat
+to E'-systems without empty contex. However, the construction of a
+canonical functor in which families are represented by compositions of projections
+is still not possible without a further coherence condition. To make sure that
+the pullback operations of a D-system extend to the category
+$\cat{F}$, we introduce \emph{coherent} D-systems. So we only have a functor $U$
+from the category $\mathbf{Dsys}$ of coherent D-systems to the category
+$\mathbf{E'sys_{\circ}}$ of E'-systems without empty context.
+We see that it is not evident at all to turn categories with families into
+E-systems, even if we ignore the issue that categories with families map
+naturally into E'-systems without empty context.
+Thus it seems that the type theories of categories with families and of E-systems
+are not directly comparable.
+
+We end this section by showing that the functor 
+$U:\mathbf{Dsys}\to\mathbf{E'sys_{\circ}}$ restricts to a full and faithful 
+functor on the coherent D-systems which satisfy the condition that every 
+composition of projections is a composition of projections in precisely one way.
+We call coherent D-systems with this property
+locally well-founded. Among the locally well-founded D-systems are the B-systems.
+As a by-product, we get that the full and faithful functor from B-systems to
+E-systems factors through D-systems.
 
@@
 induction on the length of the arrow $A$ into $\Gamma$. We define
-$\Lambda(f,\emptyf[\Gamma])\defeq \emptyf[\Delta]$, and $\pi_2(f,\emptyf[\Gamma])\defeq f:\ctxext{\Delta}{\catid{\Delta}}\to\ctxext{\Gamma}{\catid{\Gamma}}$. The resulting square is trivially
+$\Lambda(f)$ on the empty sequence as
+$\Lambda(f,\emptyf[\Gamma])\defeq \emptyf[\Delta]$, and we define
+$\pi_2(f,\emptyf[\Gamma])\defeq f:\ctxext{\Delta}{\catid{\Delta}}\to
+\ctxext{\Gamma}{\catid{\Gamma}}$. The resulting square is trivially
 a pullback square. Now suppose we have a pullback square as above.
@@
 \begin{rmk}
-The construction above does not help us define a functor $f_\ast:\cat{F}/\Gamma
-\to\cat{F}/\Delta$ without further assumptions on the D-system $\mathbb{D}$.
-The functor $f_\ast$ would have the property that the square
+We can't directly use the functors $\Lambda(f)$ to give a D-system the structure
+of an E'-system, because there is a coherence condition missing.
+\end{rmk}
+
+\begin{thm}\label{thm:Dsys_coh}
+The following are equivalent:
+\begin{enumerate}
+\item The functors $\Lambda(f):\Lambda(\Gamma)\to\Lambda(\Delta)$ extend to 
+the structure of an E'-system without empty context.
+\item For any $f:\Delta\to\Gamma$ and all $A,A'\in\Lambda(\Gamma)$ satisfying 
+$p_A\jdeq p_{A'}$, one has
+\begin{equation*}
+p_{\Lambda(f,A)}\jdeq p_{\Lambda(f,A')}\qquad\text{and}\qquad\pi_2(f,A)\jdeq \pi_2(f,A').
+\end{equation*}
+\end{enumerate}
+\end{thm}
+
+\begin{proof}
+Suppose we can extend $\Lambda(f)$ for all $f:\Delta\to\Gamma$, to form commuting
+diagrams
 \begin{equation*}
@@
 \end{equation*}
-commutes. The above square can only commute, if one has
-\begin{equation*}
-\Lambda(f)/A\jdeq\Lambda(f)/A'
-\end{equation*} 
-for any $A,A'\in\Lambda(\Gamma)$ with $p_A\jdeq p_{A'}$. This would entail that
-$\pi_2(f,A)\jdeq\pi_2(f,A')$, a property that seems to not hold generally.
-\end{rmk}
-
-\begin{defn}
-A D-system is said to be coherent if for any $f:\Delta\to\Gamma$ and any
-$A,A'\in\Lambda(\Gamma)$ satisfying $p_A\jdeq p_{A'}$, one has
-\begin{equation*}
-\pi_2(f,A)\jdeq \pi_2(f,A').
-\end{equation*}
+of functors, forming an E'-system without empty context. Let 
+$f:\Delta\to\Gamma$, and let $A,A':\Lambda(\Gamma)$ be such that 
+$p_A\jdeq p_{A'}$. Then it follows
+that $\pi_2(f,A)_\ast\jdeq f_\ast/A\jdeq f_\ast/A'\jdeq\pi_2(f,A')_\ast$, so
+we get the desired equality $\pi_2(f,A)\jdeq\pi_2(f,A')$.
+
+For the converse, suppose that for all $f:\Delta\to\Gamma$ and $A,A'\in
+\Lambda(\Gamma)$ such that $p_A\jdeq p_{A'}$, one has 
+$p_{\Lambda(f,A)}\jdeq p_{\Lambda(f,A')}$ and $\pi_2(f,A)\jdeq\pi_2(f,A')$.
+Let $f:\Delta\to\Gamma$. By the assumption that $p_{\Lambda(f,A)}\jdeq p_{\Lambda(f,A')}$,
+it follows that the objects part of $p^\Delta\circ\Lambda(f):\Lambda(\Gamma)\to\cat{F}/\Delta$ 
+factors through $p^\Gamma:\Lambda(\Gamma)\to\cat{F}/\Gamma$. Moreover, by the
+assumption that $\pi_2(f,A)\jdeq \pi_2(f,A')$ it follows that the pullback
+squares
+\begin{equation*}
+\begin{tikzcd}[column sep=large]
+\ctxext{\Delta}{f_\ast(A)} \arrow[d,fib] \arrow[r,"{\pi_2(f,A)}"] & \ctxext{\Gamma}{A} \arrow[d,fib] \\
+\Delta \arrow[r,"f"] & \Gamma
+\end{tikzcd}
+\qquad\text{and}\qquad
+\begin{tikzcd}[column sep=large]
+\ctxext{\Delta}{f_\ast(A')} \arrow[d,fib] \arrow[r,"{\pi_2(f,A')}"] & \ctxext{\Gamma}{A} \arrow[d,fib] \\
+\Delta \arrow[r,"f"] & \Gamma
+\end{tikzcd}
+\end{equation*}
+are the same. Since we have this property
+for any object $\Gamma$, it also follows that the morphisms part of
+$p^\Delta\circ\Lambda(f)$ factors through $p^\Gamma$. This defines the functor
+$f_\ast$ in the commuting square
+\begin{equation*}
+\begin{tikzcd}
+\Lambda(\Gamma) \arrow[d,swap,"p"] \arrow[r,"\Lambda(f)"] & \Lambda(\Delta) \arrow[d,"p"] \\
+\cat{F}/\Gamma \arrow[r,densely dotted,"{f_\ast}"] & \cat{F}/\Delta
+\end{tikzcd}
+\end{equation*}
+We need to verify that $f_\ast$ is a functorial choice of pullbacks, as in the
+definition of \autoref{defn:E'sys}. Since $f_\ast$ is defined as an extension
+of $\Lambda(f)$, it is enough to verify the analoguous functoriality properties
+for $\Lambda(f)$. We do this by induction on the objects of $\Lambda(\Gamma)$,
+in the same order as in \autoref{defn:E'sys}.
+\begin{enumerate}[label=(\alph*)]
+\item This property holds by definition.
+\item Let $\Gamma\in\cat{C}$. We will prove by induction on $A\in\Lambda(\Gamma)$
+that $(\catid{\Gamma})_\ast(A)\jdeq A$ and that $\pi_2(\catid{\Gamma},A)\jdeq \catid{\ctxext{\Gamma}{A}}$.
+Note that these properties hold by definition for $A\jdeq\catid{\Gamma}$. For the
+inductive step, assume that $A\in\Lambda(\Gamma)$ is such that
+$(\catid{\Gamma})_\ast(A)\jdeq A$ and $\pi_2(\catid{\Gamma},A)\jdeq
+\catid{\ctxext{\Gamma}{A}}$ and let $B\in\mathrm{Ty}(\ctxext{\Gamma}{A})$. Then
+we have
+\begin{align*}
+(\catid{\Gamma})_\ast(A\circ p_B) & \jdeq (\catid{\Gamma})_\ast(A)\circ \pi_2(\catid{\Gamma},A)_\ast(p_B) \\
+& \jdeq A\circ p_B
+\end{align*}
+Also, we have
+\begin{align*}
+\pi_2(\catid{\Gamma},A\circ p_B) & \jdeq \pi_2(\pi_2(\catid{\Gamma},A),p_B) \\
+& \jdeq \pi_2(\catid{\ctxext{\Gamma}{A}},p_B) \\
+& \jdeq \catid{\ctxext{{\Gamma}{A}}{B}}
+\end{align*}
+\item Let $f:\Delta\to \Gamma$ and $g:X\to \Delta$. We will prove by induction
+on $A\in\Lambda(\Gamma)$ that $(f\circ g)_\ast(A)\jdeq g_\ast(f_\ast(A))$ and
+that $\pi_2(f\circ g,A)\jdeq \pi_2(f,A)\circ \pi_2(g,f_\ast(A))$. Note that
+these properties hold by definition for $A\jdeq\catid{\Gamma}$. For the inductive
+step, assume that $A\in\Lambda(\Gamma)$ is such that $(f\circ g)_\ast(A)\jdeq g_\ast(f_\ast(A))$ and
+$\pi_2(f\circ g,A)\jdeq \pi_2(f,A)\circ \pi_2(g,f_\ast(A))$, and let
+$B\in\mathrm{Ty}(\ctxext{\Gamma}{A})$. Then we have
+\begin{align*}
+(f\circ g)_\ast(A\circ p_B) & \jdeq \pi_2(f\circ g,A)_\ast(p_B) \\
+& \jdeq (\pi_2(f,A)\circ\pi_2(g,f_\ast(A)))_\ast(p_B) \\
+& \jdeq \pi_2(g,f_\ast)_\ast(\pi_2(f,A)_\ast(p_B)) \\
+& \jdeq g_\ast(f_\ast(A\circ p_B))
+\end{align*}
+\item This property holds by definition.\qedhere
+\end{enumerate}
+\end{proof}
+
+\begin{defn}
+If either of the equivalent conditions of \autoref{thm:Dsys_coh} hold, we say 
+that the D-system is \emph{coherent}. Thus, we have a functor
+$U:\mathbf{Dsys}\to\mathbf{E'sys_{\circ}}$ from coherent D-systems to E'-systems
 \end{defn}
@@
 
-\begin{defn}
-For any locally well-founded D-system $\mathbb{D}$ there is 
-an E'-system $U(\mathbb{D})$ without terminal object in 
-$\cat{F}$. Thus, we get a functor $\mathbf{Dsys}\to\mathbf{E'sys}_{\circ}$
-\end{defn}
-
-\begin{constr}
-Let $f:\Delta\to\Gamma$. By the assumption that $\mathbb{D}$ is locally well-founded,
-
-To verify that this gives indeed a functor from $\cat{F}/\Gamma$ to
-$\cat{F}/\Delta$, we need to show that for any two objects or $A,A'\in\Lambda(\Gamma)$
-which get mapped to the same object of $\cat{F}/\Gamma$, get mapped to the
-same object of $\cat{F}/\Delta$. Let $A,A'\in\Lambda(\Gamma)$ be such that
-$p_A\jdeq p_{A'}$. It suffices to show that $\pi_2(f,A)\jdeq\pi_2(f,A')$
-
-The inductive proofs to verify properties
-\autoref{defn:E'sys:b} and \autoref{defn:E'sys:c} of \autoref{defn:E'sys} are
-completely straightforward, so $U(\mathbb{D})$ is an E'-system without terminal
-object in $\cat{F}$.
-\begin{comment}
-Finally, we need to verify that the structure we've defined satisfies the properties
-of being an E'-system without terminal object. We go through them systematically,
-following the same order as in \autoref{defn:E'sys}:
-\begin{enumerate}[label=(\alph*)]
-\item This property holds by definition.
-\item Let $\Gamma\in\cat{C}$. We will prove by induction on $A\in\cat{F}/\Gamma$
-that $(\catid{\Gamma})_\ast(A)\jdeq A$ and that $\pi_2(\catid{\Gamma},A)\jdeq \catid{\ctxext{\Gamma}{A}}$.
-Note that these properties hold by definition for $A\jdeq\catid{\Gamma}$. For the
-inductive step, assume that $A\in\cat{F}/\Gamma$ is such that
-$(\catid{\Gamma})_\ast(A)\jdeq A$ and $\pi_2(\catid{\Gamma},A)\jdeq
-\catid{\ctxext{\Gamma}{A}}$ and let $B\in\mathrm{Ty}(\ctxext{\Gamma}{A})$. Then
-we have
-\begin{align*}
-(\catid{\Gamma})_\ast(A\circ p_B) & \jdeq (\catid{\Gamma})_\ast(A)\circ \pi_2(\catid{\Gamma},A)_\ast(p_B) \\
-& \jdeq A\circ p_B
-\end{align*}
-Also, we have
-\begin{align*}
-\pi_2(\catid{\Gamma},A\circ p_B) & \jdeq \pi_2(\pi_2(\catid{\Gamma},A),p_B) \\
-& \jdeq \pi_2(\catid{\ctxext{\Gamma}{A}},p_B) \\
-& \jdeq \catid{\ctxext{{\Gamma}{A}}{B}}
-\end{align*}
-\item Let $f:\Delta\to \Gamma$ and $g:X\to \Delta$. We will prove by induction
-on $A\in\cat{F}/\Gamma$ that $(f\circ g)_\ast(A)\jdeq g_\ast(f_\ast(A))$ and
-that $\pi_2(f\circ g,A)\jdeq \pi_2(f,A)\circ \pi_2(g,f_\ast(A))$. Note that
-these properties hold by definition for $A\jdeq\catid{\Gamma}$. For the inductive
-step, assume that $A\in\cat{F}/\Gamma$ is such that $(f\circ g)_\ast(A)\jdeq g_\ast(f_\ast(A))$ and
-$\pi_2(f\circ g,A)\jdeq \pi_2(f,A)\circ \pi_2(g,f_\ast(A))$, and let
-$B\in\mathrm{Ty}(\ctxext{\Gamma}{A})$. Then we have
-\begin{align*}
-(f\circ g)_\ast(A\circ p_B) & \jdeq \pi_2(f\circ g,A)_\ast(p_B) \\
-& \jdeq (\pi_2(f,A)\circ\pi_2(g,f_\ast(A)))_\ast(p_B) \\
-& \jdeq \pi_2(g,f_\ast)_\ast(\pi_2(f,A)_\ast(p_B)) \\
-& \jdeq g_\ast(f_\ast(A\circ p_B))
-\end{align*}
-
-\item This property holds by definition.
-\end{enumerate}
-\end{comment}
-\end{constr}
-
 \begin{thm}
-The functor $U:\mathbf{Dsys}\to\mathbf{E'sys_{\circ}}$ is full and faithful.
+The restriction of the functor $U:\mathbf{Dsys}\to\mathbf{E'sys_{\circ}}$ to the
+locally well-founded D-systems is full and faithful.
 \end{thm}
@@
 \begin{proof}
-The functor $U$ is faithful, because for any morphism 
+The functor $U$ is faithful on the locally well-founded D-systems, because for any morphism 
 $\mathcal{F}:\mathbb{D}\to\mathbb{D}'$ of locally well-founded D-systems, 
@@
 
-It remains to show that $U$ is full. Let
-$\mathbb{D}$ and $\mathbb{D}$ be D-systems and let $\mathcal{G}:U(\mathbb{D})
+It remains to show that $U$ is full on the locally well-founded D-systems. Let
+$\mathbb{D}$ and $\mathbb{D}$ be locally well-founded D-systems and let $\mathcal{G}:U(\mathbb{D})
 \to U(\mathbb{D}')$. Then $\mathcal{G}$ restricted to $\cat{F}$ lifts factorizations.

### 2015-08-16 compared.tex
didn't keep track of the work I did

--- 2015-07-29 weakeningcats.tex
@@
-\section{E-systems}\label{sec:esys_defn}
-
-\subsection{Weakening systems}
-
-\begin{defn}
-A \emph{pre-weakening structure on a pre-category $\cat{F}$} consists of a functor
-$W_A:\cat{F}/\Gamma\to \cat{F}/\ctxext{\Gamma}{A}$ for every $A:\ctxext{\Gamma}{A}\to \Gamma$ in $\cat{F}$ such 
-that
-\begin{enumerate}
-\item $W_{\catid{\Gamma}}\jdeq \catid{\cat{F}/\Gamma}$ for every object $\Gamma\in\cat{F}$.
-\item $W_{\ctxext{A}{P}}\jdeq W_P\circ W_A$ for every $P\in\cat{F}/\ctxext{\Gamma}{A}$ and $A\in\cat{F}/\Gamma$.
-\item $W_A$ preserves the final object, i.e.~$W_A(\catid{\Gamma})\jdeq \catid{\ctxext{\Gamma}{A}}$.
-\end{enumerate}
-A \emph{pre-weakening system} is a pre-category together with a pre-weakening structure.
-\end{defn}
-
-\begin{defn}
-A \emph{pre-weakening homomorphism} $F:\cat{F}\to\cat{D}$ between pre-weakening systems
-is a functor $F:\cat{F}\to\cat{D}$ such that the square
-\begin{equation*}
-\begin{tikzcd}[column sep=large]
-\cat{F}/\ctxext{\Gamma}{A}
-  \arrow[r,"F/\ctxext{\Gamma}{A}"]
-  &
-\cat{D}/F(\ctxext{\Gamma}{A})
-  \\
-\cat{F}/\Gamma
-  \arrow[u,"W_A"]
-  \arrow[r,swap,"F/\Gamma"]
-  &
-\cat{D}/F(\Gamma)
-  \arrow[u,swap,"W_{F(A)}"]
-\end{tikzcd}
-\end{equation*}
-commutes for any $A\in\cat{F}/\Gamma$.
-\end{defn}
-
-\begin{lem}
-Let $\cat{F}$ be a pre-weakening system. Then for any object $\Gamma$ of $\cat{F}$,
-the slice pre-category $\cat{F}/\Gamma$ is also a pre-weakening system, with the weakening
-structure given by $W(\cat{F}/\Gamma)_P\jdeq W(\cat{F})_P$ for each $P\in\cat{F}/\ctxext{\Gamma}{A}$
-and each $A\in\cat{F}/\Gamma$.
-\end{lem}
-
-\begin{defn}
-A \emph{weakening system} is a pre-weakening system for which every $W_f$ is a
-pre-weakening homomorphism. A \emph{weakening homomorphism} is a pre-weakening 
-homomorphism between weakening systems.
-\end{defn}
-
-\begin{rmk}
-In other words, a weakening system is a pre-weakening system with a weakening
-structure $W$ for which the square
-\begin{equation*}
-\begin{tikzcd}[column sep=huge]
-\cat{F}/\ctxext{{\Gamma}{B}}{Q}
-  \arrow[r,"W_A/\ctxext{B}{Q}"]
-  &
-\cat{F}/\ctxext{{\Gamma}{A}}{W_A(\ctxext{B}{Q})}
-  \\
-\cat{F}/\ctxext{\Gamma}{B}
-  \arrow[u,"W_Q"]
-  \arrow[r,swap,"W_A/B"]
-  &
-\cat{F}/\ctxext{{\Gamma}{A}}{W_A(B)}
-  \arrow[u,swap,"W_{W_A(Q)}"]
-\end{tikzcd}
-\end{equation*}
-for each $A,B\in\cat{F}/\Gamma$ and $Q:\cat{F}/\ctxext{\Gamma}{B}$. The objects part of this 
-property asserts that for any $k\in \cat{F}/\greek{E}$, the dotted arrows in the diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=tiny]
-\bullet \arrow[dr,densely dotted,swap,near start,"(W_A/\ctxext{B}{Q})(W_Q(R))"] & & \bullet \arrow[dl,densely dotted,near start,"W_{(W_A/B)(Q)}((W_A/B)(R))"] & & & & & & & & & & \bullet \arrow[dl,"W_Q(R)"] \\
-& \bullet \arrow[dr,swap,"(W_A/B)(Q)"] & & \bullet \arrow[dl,"(W_A/B)(R)"] & & & & & & & & \bullet \arrow[dr,swap,"Q"] & & \bullet \arrow[dl,"R"] \\
-& & \bullet \arrow[d,swap,"W_A(B)"] & & & & & & & & & & \ctxext{\Gamma}{B} \arrow[d,"B"] \\
-& & \ctxext{\Gamma}{A} \arrow[rrrrrrrrrr,swap,"A"] & & & & & & & & & & \Gamma
-\end{tikzcd}
-\end{equation*}
-are equal.
-
-A useful special case of this property is where $B\jdeq \catid{\Gamma}$. Thus, if $W$ is
-a weakening system, then the diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=huge]
-\cat{F}/\ctxext{\Gamma}{C}
-  \arrow[r,"W_A/C"]
-  &
-\cat{F}/\ctxext{{\Gamma}{A}}{W_A(C)}
-  \\
-\cat{F}/\Gamma
-  \arrow[u,"W_C"]
-  \arrow[r,swap,"W_A"]
-  &
-\cat{F}/\ctxext{\Gamma}{A}
-  \arrow[u,swap,"W_{W_A(C)}"]
-\end{tikzcd}
-\end{equation*}
-commutes for every $A,C\in\cat{F}/\Gamma$. In particular, we see
-that $W_A(W_C(D))\jdeq W_{W_A(C)}(W_A(D))$ for any $D\in\cat{F}/\Gamma$,
-i.e.~that weakening is a self-distributive operation. 
-\end{rmk}
-
-\begin{cor}
-For any object $\Gamma$ of a weakening system $\cat{F}$, the slice pre-category
-$\cat{F}/\Gamma$ is a weakening system.
-\end{cor}
-
-\begin{eg}
-If the underlying pre-category is a monoid $M$, the axioms of a weakening system
-obtain the following form: There is a group homomorphism $W_x:M\to M$ for
-every $x\in M$, in such a way that $W_e\jdeq \catid{M}$ and $W_{xy}\jdeq W_y\circ W_x$
-for every $x,y\in M$, and for any $x,y\in M$ the diagram
-\begin{equation*}
-\begin{tikzcd}
-M \arrow[r,"W_y"]
-  &
-M 
-  \\
-M \arrow[u,"W_x"]
-  \arrow[r,swap,"W_y"]
-  &
-M \arrow[u,swap,"W_{W_y(x)}"]
-\end{tikzcd}
-\end{equation*}
-is required to commute. Thus, a weakening system on a monoid $M$ consists of
-a homomorphism $W:\op M\to\mathrm{Aut}(M)$ such that $W_x\circ W_y\jdeq 
-W_{W_x(y)}\circ W_x$. The latter requirement is satisfied if $y\cdot x\jdeq x\cdot W_x(y)$
-for every $x,y\in M$. Group conjugation forms a weakening structure on any group.
-\end{eg}
-
-\begin{comment}
-\begin{lem}
-For any group, conjugation is a weakening structure. Thus, any group is
-canonically a weakening system.
-\end{lem}
-
-\begin{proof}
-Note that $e^{-1}xe\jdeq x$ and $(xy)^{-1}z(xy)\jdeq y^{-1}(x^{-1}zx)y$ for any $x,y,z\in G$.
-Also, the diagram
-\begin{equation*}
-\begin{tikzcd}
-G \arrow[r,"\mathrm{conj}_y"]
-  &
-G 
-  \\
-G \arrow[u,"\mathrm{conj}_x"]
-  \arrow[r,swap,"\mathrm{conj}_y"]
-  &
-G \arrow[u,swap,"\mathrm{conj}_{y^{-1}xy}"]
-\end{tikzcd}
-\end{equation*}
-commutes.
-\end{proof}
-\end{comment}
-
-\begin{comment}
-\begin{rmk}
-Similarly, every groupoid has a canonical weakening structure given by
-conjugation.
-\end{rmk}
-\end{comment}
-
-\begin{comment}
-\begin{lem}
-Let $G$ be a graph, and let $\cat{F}_G$ be the free pre-category generated by
-$G$. A pre-weakening structure on $\cat{F}_G$ is determined by functors
-$W_e:\cat{F}_G/\Gamma\to\cat{F}_G/\Delta$ for every edge $e:\Delta\to \Gamma$ in $G$, with the
-property that $W_e(\catid{\Gamma})\jdeq \catid{\Delta}$.
-
-A weakening structure on $\cat{F}_G$ is
-a pre-weakening structure on $\cat{F}_G$ such that the diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=huge]
-\cat{F}_G/\greek{Z} \arrow[r,"W_e/(g\circ h)"] & \cat{F}_G/\mathrm{dom}(W_e(g\circ h)) \\
-\cat{F}_G/\greek{E} \arrow[u,"W_h"] \arrow[r,swap,"W_e/g"] & \cat{F}_G/\mathrm{dom}(W_e(g)) \arrow[u,swap,"W_{W_e(h)}"]
-\end{tikzcd}
-\end{equation*}
-commutes for every arrow $g:\greek{E}\to \Gamma$ in $\cat{F}_G$ and every two edges $e:\Delta\to \Gamma$ and
-$h:\greek{Z}\to \greek{E}$ in $G$.
-\end{lem}
-
-\begin{eg}
-Let $G$ be the graph consisting of one vertex $\ast$ and one edge $e:\ast\to\ast$.
-Then $\cat{F}_G/\ast\jdeq \N$. Then the only weakening structure on $\cat{F}/G$ is
-given by $W_e\jdeq \catid{\N}$. The reason is that 
-\end{eg}
-
-\begin{lem}
-Let $P$ be a pre-order. A pre-weakening structure on $P$ is determined by order
-preserving maps $W_{a,b}:{\downarrow}(b)\to{\downarrow}(a)$ satisfying
-$W_{a,b}(b)\jdeq a$, for every $a\leq b$, where
-${\downarrow}(x)$ is the set $\{y\in P\mid y\leq x\}$.
-
-A weakening structure on $P$ is a pre-weakening structure on $P$ such that the
-diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=huge]
-{\downarrow}(c') \arrow[r,"W_{a,b}|_{{\downarrow}(c')}"] & {\downarrow}(W_{a,b}(c')) \\
-{\downarrow}(c) \arrow[u,"W_{c',c}"] \arrow[r,swap,"W_{a,b}|_{{\downarrow}(c)}"] & {\downarrow}(W_{a,b}(c)) \arrow[u,swap,"W_{W_{a,b}(c'),W_{a,b}(c)}"]
-\end{tikzcd}
-\end{equation*}
-for any $a\leq b$ and $c'\leq c\leq b$.
-\end{lem}
-
-\begin{eg}
-In a poset with meets, the functions $W_{a,b}:{\downarrow}(b)\to{\downarrow}(a)$
-given by $W_{a,b}(x)\defeq a\land x$ assemble a weakening structure, because for
-any $a\leq b$ and $x,c'\leq c\leq b$ we have $a\land(c'\land x)\jdeq (a\land c')\land
-(a\land x)$.  
-\end{eg}
-
-\begin{eg}
-Consider the poset $(\N,\geq)$, which has a morphism from $a$ to $b$ if $a\geq b$.
-On this poset, we may consider the weakening structure defined by $W_{a,b}(x)\jdeq
-a+(x-b)$.
-\end{eg}
-\end{comment}
-
-\subsection{Weakening systems with term structure}
-\begin{defn}
-A \emph{category with a term structure} is a pre-category $\cat{F}$ with a set $T(A)$ for every
-$A\in\cat{F}/\Gamma$ for every $\Gamma\in\cat{F}$. A \emph{functor with term structure} from $\cat{F}$ to $\cat{D}$
-is a functor $F:\cat{F}\to\cat{D}$ with a function $T(A)\to T(F(A))$ for
-every $A\in\cat{F}/\Gamma$, for every $\Gamma\in\cat{F}$.
-\end{defn}
-
-\begin{defn}
-Suppose $\cat{F}$ is a pre-category with term structure, and $\Gamma$ is an object of $\cat{F}$. 
-Then $\cat{F}/\Gamma$ is a pre-category with term structure, where $T_{\cat{F}/\Gamma}(A)\jdeq T_\cat{F}(A)$.
-\end{defn}
-
-\begin{defn}
-A \emph{weakening system with term term structure} is the same thing as a weakening system,
-except that the underlying pre-category and the functors $W_A$ are required to
-be functors with term term structure. The equalities are also required to hold on the term structure.
-\end{defn}
-
-\begin{cor}
-For any object $\Gamma$ of a weakening system with term structure, the slice pre-category
-$\cat{F}/\Gamma$ is a weakening system with term structure.
-\end{cor}
-
-\begin{eg}
-%For any group $G$ with weakening structure $W:\op{G}\to\mathrm{Aut}(G)$, a term 
-%structure is determined by a presheaf on $G/\ker(W)$.
-
-Let $G$ be a group, with the weakening structure given by conjugation. A term
-structure on $G$ is given by a set $X$ with a right $G$-action $\mu:\op G\times X\to X$.
-We take $T(a)\defeq X$ for every $a\in G$.
-\end{eg}
-
-\subsection{Projection systems}
-\begin{defn}
-A \emph{pre-projection system} is a pre-weakening system with term structure with an element
-$\tfid{A}\in T(W_A(A))$ for every $A\in\cat{F}/\Gamma$ and $\Gamma\in\cat{F}$. 
-\end{defn}
-
-\begin{defn}
-A \emph{pre-projection homomorphism} $F:\cat{F}\to\cat{D}$ is a pre-weakening homomorphism for which
-\begin{equation*}
-F(\tfid{A})\jdeq \tfid{F(A)}
-\end{equation*}
-for every $A\in\cat{F}/\Gamma$ and $\Gamma\in\cat{F}$.
-\end{defn}
-
-\begin{lem}
-For any object $\Gamma$ in a pre-projection system $\cat{F}$, the slice pre-category
-$\cat{F}/\Gamma$ is a pre-projection system.
-\end{lem}
-
-\begin{proof}
-Straightforward.
-\end{proof}
-
-\begin{defn}
-A \emph{projection system} is a pre-projection system for which every $W_A$ is
-a pre-projection homomorphism. A \emph{projection homomorphism} is a pre-projection homomorphism
-between projection systems.
-\end{defn}
-
-\begin{cor}
-For any object $\Gamma$ of a projection system $\cat{F}$, the slice pre-category $\cat{F}/\Gamma$
-is a projection system.
-\end{cor}
-
-%\begin{eg}
-%Consider a group $G$ with a right $G$-set $X$. A projection structure on $(G,X)$
-%consists of an element $\tfid{a}\in X$ for every $a\in G$, so that $\tfid{a}\cdot b=
-%\tfid{b^{-1}ab}$ for every $a,b\in G$. 
-%For instance, this is the case if we take $X$ to be a pointed set with $\ast\in X$,
-%and take $\tfid{a}:=\ast$ for all $a\in G$.
-%\end{eg}
-
-\subsection{Substitution systems}
-
-\begin{defn}
-A \emph{pre-substitution structure} on a pre-category with term structure $\cat{F}$ consists
-of a functor with term structure $S_x:\cat{F}/\ctxext{\Gamma}{A}\to\cat{F}/\Gamma$ for every $x\in T(A)$ and
-$A\in\cat{F}/\Gamma$, such that $S_x(\catid{\ctxext{\Gamma}{A}})\jdeq\catid{\Gamma}$.
-
-A \emph{pre-substitution system} is a pre-category with term structure together with a
-pre-substitution structure. 
-\end{defn}
-
-\begin{defn}
-A \emph{pre-substitution homomorphism} $F:\cat{F}\to\cat{D}$ is a functor with term structure for
-which the diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=large]
-\cat{F}/\ctxext{\Gamma}{A}
-  \arrow[r,"F/\ctxext{\Gamma}{A}"]
-  \arrow[d,swap,"S_x"]
-  &
-\cat{D}/F(\ctxext{\Gamma}{A})
-  \arrow[d,"S_{F(x)}"]
-  \\
-\cat{F}/\Gamma
-  \arrow[r,swap,"F/\Gamma"]
-  &
-\cat{D}/F(\Gamma)
-\end{tikzcd}
-\end{equation*}
-commutes for every $x\in T(A)$ and $A\in\cat{F}/\Gamma$.
-\end{defn}
-
-\begin{lem}
-For any object $\Gamma$ of a pre-substitution system $\cat{F}$, the slice pre-category
-$\cat{F}/\Gamma	$ is a pre-substitution system.
-\end{lem}
-
-\begin{defn}
-A \emph{substitution system} is a pre-substitution system for which each
-$S_t$ is a pre-substitution homomorphism. A \emph{substitution homomorphism} is a
-pre-substitution homomorphism between substitution systems.
-\end{defn}
-
-\begin{cor}
-For any object $\Gamma$ of a substitution system, the slice pre-category $\cat{F}/\Gamma$
-is a substitution system.
-\end{cor}
-
-\begin{rmk}
-The condition that every $S_x$ is a substitution homomorphism, asserts that
-the diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=large]
-\cat{F}/\ctxext{{{\Gamma}{A}}{P}}{Q}
-  \arrow[r,"S_x/\ctxext{P}{Q}"]
-  \arrow[d,swap,"S_y"]
-  &
-\cat{F}/\ctxext{{\Gamma}{S_x(P)}}{S_x(Q)}
-  \arrow[d,"S_{S_x(y)}"]
-  \\
-\cat{F}/\ctxext{{\Gamma}{A}}{P}
-  \arrow[r,swap,"S_x/P"]
-  &
-\cat{F}/\ctxext{\Gamma}{S_x(P)}
-\end{tikzcd}
-\end{equation*}
-commutes for every $y\in T(Q)$.
-\end{rmk}
-
-\begin{eg}
-Let $G$ be a group and let $H$ be a set, forming a category with term 
-structure of which $T(a)\jdeq H$ for every $a\in G$. A substitution
-system on $(G,H)$ consists of a function $\sigma_G : H\times G\to G$ and
-a function $\sigma_H:H\times H\to H$, for which the diagrams
-\begin{equation*}
-\begin{tikzcd}[column sep=large]
-G \arrow[r,"\sigma_G(h)"]
-  \arrow[d,swap,"\sigma_G(k)"]
-  &
-G \arrow[d,"{\sigma_G(\sigma_H(h,k))}"]
-  \\
-G \arrow[r,swap,"\sigma_G(h)"] 
-& G
-\end{tikzcd}
-\end{equation*}
-and
-\begin{equation*}
-\begin{tikzcd}[column sep=large]
-H \arrow[r,"\sigma_H(h)"]
-  \arrow[d,swap,"\sigma_H(k)"]
-  &
-H \arrow[d,"{\sigma_H(\sigma_H(h,k))}"]
-  \\
-H \arrow[r,swap,"\sigma_H(h)"]
-  &
-H
-\end{tikzcd}
-\end{equation*}
-commute for every $h,k\in H$. 
-\end{eg}
-
-\subsection{The definition of E-systems}
-
-Although it is now possible to state the definition of an E-system without
-any difficulty, it shall be convenient later to also have a category of
-pre-E-systems at our disposal. Pre-E-systems shall by default be non-unital,
-i.e.~they do not come with a pre-projection system, while E-systems shall
-always be assumed to be unital. In the present context, we shall not need
-unital pre-E-systems or non-unital E-systems.
-
-\begin{defn}
-A \emph{pre-E-system} is a pre-category $\cat{F}$ with term structure
-with a pre-weakening structure $W$, a pre-substitution structure $S$ and a terminal
-object. A \emph{unital pre-E-system}, or simply \emph{pre-E-system}, is a non-unital pre-E-system
-which also has the structure of a pre-projection system. 
-\end{defn}
-
-\begin{defn}
-A \emph{pre-E-homomorphism} from $\mathbb{E}\to\mathbb{D}$ is a functor
-$H:\mathbb{C}_\mathbb{E}\to\mathbb{C}_\mathbb{D}$ between the underlying categories
-with term structure, which is a pre-weakening homomorphism, a pre-substitution
-homomorphism, and which preserves the terminal object.
-
-A \emph{unital} pre-E-homomorphism is a non-unital pre-E-homomorphism which is 
-also a pre-projection homomorphism. 
-\end{defn}
-
-\begin{defn}
-\label{defn:esystems}
-An \emph{E-system} is a unital pre-E-system in which 
-\begin{enumerate}
-\item each $W_A$ is a unital pre-E-homomorphism,
-\item each $S_x$ is a unital pre-E-homomorphism,
-\item \label{tTf:StWf_id} weakened families are constant, i.e.~$S_x\circ W_A\jdeq \catid{\cat{F}/\Gamma}$ for any $x\in T(A)$ and $A\in\cat{F}/\Gamma$,
-\item $S_x(\tfid{A})\jdeq x$ for any $x\in T(A)$ and $A\in\cat{F}/\Gamma$, and
-\item \label{f:SidfWff_id} $S_{\tfid{A}}\circ W_A/A\jdeq \catid{\cat{F}/\ctxext{\Gamma}{A}}$ for any $A\in\cat{F}/\Gamma$.
-\end{enumerate}
-An \emph{E-homomorphism} $H:\mathbb{E}\to\mathbb{D}$ is a functor from
-$\cat{F}_{\mathbb{E}}$ to $\cat{F}_{\mathbb{D}}$ which is both a projection
-homomorphism and a substitution homomorphism.
-\end{defn}
-
-\begin{rmk}
-The condition that each $W_A$ is a substitution homomorphism asserts that
-the diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=huge]
-\cat{F}/\ctxext{{\Gamma}{B}}{Q}
-  \arrow[r,"W_A/\ctxext{B}{Q}"]
-  \arrow[d,swap,"S_y"]
-  &
-\cat{F}/\ctxext{{{\Gamma}{A}}{W_A(B)}}{W_A(Q)}
-  \arrow[d,"S_{W_A(y)}"]
-  \\
-\cat{F}/\ctxext{\Gamma}{B}
-  \arrow[r,swap,"W_A/B"]
-  &
-\cat{F}/\ctxext{{\Gamma}{A}}{W_A(B)}
-\end{tikzcd}
-\end{equation*}
-of functors with term structure commutes for every 
-$Q\in\cat{F}/\ctxext{\Gamma}{B}$, $B\in\cat{F}/\Gamma$ and each $y\in T(Q)$.
-
-Likewise, the condition that each $S_x$ is a weakening homomorphism
-asserts that the diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=huge]
-\cat{F}/\ctxext{{\Gamma}{A}}{P}
-  \arrow[r,"S_x/P"]
-  \arrow[d,swap,"W_Q"]
-  &
-\cat{F}/\ctxext{\Gamma}{S_x(P)}
-  \arrow[d,"W_{S_x(Q)}"]
-  \\
-\cat{F}/\ctxext{{{\Gamma}{A}}{P}}{Q}
-  \arrow[r,swap,"S_x/\ctxext{P}{Q}"]
-  &
-\cat{F}/\ctxext{{\Gamma}{S_x(P)}}{S_x(Q)}
-\end{tikzcd}
-\end{equation*}
-of functors with term structure commutes for every 
-$Q\in\cat{F}/\ctxext{{\Gamma}{A}}{P}$.
-\end{rmk}
-
-\begin{cor}
-For any object $\Gamma$ of an E-system $\cat{F}$, the slice pre-category $\cat{F}/\Gamma$
-is an E-system.
-\end{cor}
-
-\begin{comment}
-\begin{rmk}[Needs more precision]
-Consider the category $E$ generated by the graph consisting of the objects of 
-the underlying category of an $E$-system $\mathcal{E}$ as edges, and the morphisms
-$S_t:X\to Y$ and $W_A:Y\to X$ as edges, subject to the axioms of an $E$-systems.
-
-Any morphism of $E$ can be rewritten in a unique way
-as a composition $\tilde S\circ \tilde W$, where $\tilde S$ is a morphism in
-substitution-normal form, $\tilde W$ is in weakening normal form, every term
-expression is in normal form, and where
-each occurence of $S_x\circ W_A$ and $S_{\tfid{A}}\circ W_A/A$ is eliminated.
-\end{rmk}
-\end{comment}
-
-\section{The category of internal morphisms of an E-system}\label{sec:esys_props}
-
-The theory of E-systems could be seen as a sort of category theory with only
-projections and their sections. We elaborate on this point of view by defining
-the category of internal morphisms, which Lawvere might have called the
-category of terms. 
-In the category of internal morphisms over a context $\Gamma$,
-which we define in \autoref{subsec:im_cat},
-a morphism $f:A\to B$ is a term of the constant family $W_A(B)$ over 
-$\ctxext{\Gamma}{A}$. Rather than directly defining composition, we will define
-for every $f:A\to B$ in context $\Gamma$, a pre-composition operation
-$f_\ast$, which is going to be an E-homomorphism from 
-$\cat{F}/\ctxext{\Gamma}{B}$ to $\cat{F}/\ctxext{\Gamma}{A}$. This allows us
-to consider diagrams of the following form:
-\begin{equation*}
-\begin{tikzcd}
-P \arrow[d,->>] \arrow[r,"g"] & Q \arrow[d,->>] \\
-A \arrow[r,swap,"f"] & B
-\end{tikzcd}
-\end{equation*}
-where we consider the morphism $g$ over $f$ to be an internal morphism from 
-$P$ to $f_\ast(Q)$ in context $\ctxext{\Gamma}{A}$, capturing the
-essence of being a morphism from the family $P$ over $A$ to the family
-$Q$ over $B$, reindexed by $f$. 
-Whenever we draw such diagrams, a double headed arrow will always
-refer to a family, and an ordinary arrow will refer to an internal morphism,
-possibly over another internal morphism.
-
-A morphism $g$ over $f$ can be vertically composed with $f$ so
-that we obtain a morphism $\jvcomp{P}{f}{g}:\ctxext{A}{P}\to\ctxext{B}{Q}$ in 
-context $\Gamma$, and it can be horizontally composed with a morphism $k:Q\to R$
-over $h:B\to C$ so that we obtain a morphism $\jfcomp{A}{f}{P}{g}{k}:P\to R$ 
-over $\jcomp{A}{f}{h}:A\to C$. 
-In \autoref{sec:interchange}, we will show interchange laws between these
-sorts of composition. The interchange laws have as a consequence that if we
-assert the commutativity of a large composed diagram.
-
-The category of internal morphisms forms a category of which the families form
-a subcategory. This, together with the pre-composition operation described
-earlier, gives rise to the auxiliary notion of E'-systems, which we define
-in \autoref{sec:E'}. The thread running through this whole section is proving
-that the notions of E-system and E'-system are the same. In \autoref{sec:E'}
-we show directly that any E'-system is an E-system. We end this section by
-showing that the converse also holds.
-
-Before we begin this section, we introduce some more convenient notation.
-
-\begin{defn}
-Let $A\in\cat{F}/\Gamma$. Recall that $W_A:\cat{F}/\Gamma\to\cat{F}/\ctxext{\Gamma}{A}$ acts on objects,
-morphisms and terms. We introduce the infix form of weakening by $A\in\cat{F}/\Gamma$ to be
-$\ctxwk{A}{\blank}$. Thus, we will write
-\begin{align*}
-\ctxwk{A}{B} & \defeq W_A(B) & & \text{for $B\in\cat{F}/\Gamma$} \\
-\ctxwk{A}{Q} & \defeq W_A(Q) & & \text{for $B\in\cat{F}/\Gamma$ and $Q\in\cat{F}/\ctxext{\Gamma}{B}$}\\
-\ctxwk{A}{g} & \defeq W_A(g) & & \text{for $B\in\cat{F}/\Gamma$, $Q\in\cat{F}/\ctxext{\Gamma}{B}$ and $g\in T(Q)$}
-\end{align*}
-\end{defn}
-
-\begin{defn}
-Let $x\in T(A)$ for a family $A\in\cat{F}/\Gamma$. The infix form of substitution
-by $x$ is taken to be $\subst{x}{\blank}$. Thus, we will write
-\begin{align*}
-\subst{x}{P} & \defeq S_x(P) & & \text{for $P\in\cat{F}/\ctxext{\Gamma}{A}$} \\
-\subst{x}{Q} & \defeq S_x(Q) & & \text{for $P\in\cat{F}/\ctxext{\Gamma}{A}$ and $Q\in\cat{F}/\ctxext{{\Gamma}{A}}{P}$} \\
-\subst{x}{g} & \defeq S_x(g) & & \text{for $P\in\cat{F}/\ctxext{\Gamma}{A}$, $Q\in\cat{F}/\ctxext{{\Gamma}{A}}{P}$ and $g\in T(Q)$}
-\end{align*}
-\end{defn}
-
-\subsection{E'-systems}\label{sec:E'}
-
-E'-systems can be thought of as strict versions of tribes%
-\footnote{Tribes are an invention of Joyal, see \url{http://ncatlab.org/homotopytypetheory/files/Joyal.pdf}}.
-
-\begin{defn}\label{defn:E'sys}
-An \emph{E'-system} consists of a category $\cat{C}$ and a subcategory $\cat{F}$ with
-the same objects as $\cat{C}$, such that
-\begin{enumerate}
-\item $\cat{F}$ has a terminal object.
-\item For any $f:\Delta\to\Gamma$ in $\cat{C}$ and any $A\in\cat{F}/\Gamma$, a
-functorial choice of a pullback square 
-\begin{equation*}
-\begin{tikzcd}[column sep=large]
-\ctxext{\Delta}{f_\ast(A)} \arrow[r,"{\pi_2(f,A)}"] \arrow[d,fib,swap,"{f_\ast(A)}"] & \ctxext{\Gamma}{A} \arrow[d,fib,"A"] \\
-\Delta \arrow[r,"f"] & \Gamma
-\end{tikzcd}
-\end{equation*}
-such that $f_\ast(A)\in\cat{F}/\Delta$. Explicitly, the functoriality
-requirement is that
-\begin{enumerate}
-\item For any $f:\Delta\to\Gamma$, one has 
-$f_\ast(\catid{\Gamma})\jdeq\catid{\Delta}$ and
-\begin{equation*}
-\pi_2(f,\catid{\Gamma})\jdeq f.
-\end{equation*}
-\item \label{defn:E'sys:b} For any $A\in\cat{F}/\Gamma$, one has $(\catid{\Gamma})_\ast(A)\jdeq A$ and
-\begin{equation*}
-\pi_2(\catid{\Gamma},A)\jdeq\catid{\ctxext{\Gamma}{A}}
-\end{equation*}
-\item \label{defn:E'sys:c} For any $f:\Delta\to\Gamma$, $g:X\to\Delta$ and $A\in\cat{F}/\Gamma$, one
-has $(f\circ g)_\ast(A)\jdeq g_\ast(f_\ast(A))$ and 
-\begin{equation*}
-\pi_2(f\circ g,A)\jdeq \pi_2(f,A)\circ\pi_2(g,f_\ast(A))
-\end{equation*}
-\item For any $P\in\cat{F}/\ctxext{\Gamma}{A}$ and $f:\Delta\to\Gamma$, one has
-$f_\ast(P)\jdeq\pi_2(f,A)_\ast(P)$ and 
-\begin{equation*}
-\pi_2(f,\ctxext{A}{P}) \jdeq \pi_2(\pi_2(f,a),P)
-\end{equation*}
-\end{enumerate}
-\end{enumerate}
-For any $f:\Delta\to\Gamma$ we shall write $f_\ast$ for the induced functor
-$\cat{F}/\Gamma\to\cat{F}/\Delta$. 
-\end{defn}
-
-\begin{rmk}
-Without a terminal object in $\cat{F}$, the category $\cat{C}$ does not correspond
-directly to anything in an E-system.
-\end{rmk}
-
-\begin{rmk}
-For E'-systems, we do not require that every pullback of $A\in\cat{F}/\Gamma$ along $f:\Delta\to\Gamma$
-is in $\cat{F}/\Delta$, just that there is a choice of such a pullback which is
-in $\cat{F}/\Delta$. Thus, the definition of E'-system is slightly different from
-Taylor's notion of display map category. Furthermore, the class of display maps in
-a display map category is not required to be closed under composition, and neither
-is there assumed to exist an object $T$ such that for every object $X$ there is a 
-unique display map $X\to T$.
-\end{rmk}
-
-\begin{rmk}
-Also, we do not require that $\cat{C}$ has a terminal object, and neither do
-we require that $\cat{F}$ contains all the isomorphisms of $\cat{C}$. These are
-essential differences with Joyal's notion of tribes.   
-\end{rmk}
-
-\begin{defn}
-For any $\Gamma\in\cat{C}$, define $\cat{C}_\cat{F}/\Gamma$ to be the full
-subcategory of $\cat{C}/\Gamma$ with the same objects as $\cat{F}/\Gamma$. 
-\end{defn}
-
-\begin{lem}\label{lem:pb_selfdistributive}
-Let $\mathbb{E}$ be an E'-system. Then $f_\ast$ can be extended to a functor from
-$\cat{C}_\cat{F}/\Gamma$ to $\cat{C}_\cat{F}/\Delta$, for each $f:\Delta\to\Gamma$, and moreover the diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=large]
-\cat{C}_\cat{F}/\ctxext{\Gamma}{B} \arrow[r,"{f_\ast/B}"] \arrow[d,swap,"{g_\ast}"]
-& \cat{C}_\cat{F}/\ctxext{\Delta}{f_\ast(B)} \arrow[d,"{(f_\ast(g))_\ast}"] \\
-\cat{C}_\cat{F}/\ctxext{\Gamma}{A} \arrow[r,swap,"{f_\ast/A}"]
-& \cat{C}_\cat{F}/\ctxext{\Delta}{f_\ast(A)}
-\end{tikzcd}
-\end{equation*}
-commutes for every $g:A\to B$ in $\cat{C}_\cat{F}/\Gamma$.
-\end{lem}
-
-\begin{proof}
-It is straightforward to use the universal property of the pullback to extend
-$f_\ast$ to a functor from $\cat{C}_\cat{F}/\Gamma$ to $\cat{C}_\cat{F}/\Delta$.
-To see that the diagram commutes, we calculate
-\begin{align*}
-f_\ast/A\circ g_\ast
-& \jdeq
-\pi_2(f,A)_\ast\circ g_\ast \\
-& \jdeq
-(g\circ\pi_2(f,A))_\ast \\
-& \jdeq
-(\pi_2(f,B)\circ f_\ast(g))_\ast \\
-& \jdeq
-(f_\ast(g))_\ast\circ\pi_2(f,B)_\ast \\
-& \jdeq
-(f_\ast(g))_\ast\circ f_\ast/B.\qedhere
-\end{align*}
-\end{proof}
-
-\begin{thm}
-An E-system is the same thing as an E'-system.
-\end{thm}
-
-\begin{proof}
-We define for any $A\in\cat{F}/\Gamma$ the set
-\begin{equation*}
-T(A)\defeq\{x:\Gamma\to\ctxext{\Gamma}{A}\mid A\circ x\jdeq\catid{\Gamma}\}.
-\end{equation*}
-Since each functor $f_\ast$ is extended to all of $\cat{C}_{\cat{F}}/\Gamma$,
-it follows immediately that those are functors with term structure.
-
-We define for any $A\in\cat{F}/\Gamma$, the functor $W_A\defeq A_\ast:\cat{F}/\Gamma\to\cat{F}/\ctxext{\Gamma}{A}$.
-Likewise, we define for any $x\in T(A)$, the functor $S_x\defeq x_\ast:\cat{F}/\ctxext{\Gamma}{A}\to\cat{F}/\Gamma$.
-We also define $\tfid{A}:T(W_A(A))$ by the universal property of pullbacks:
-\begin{equation*}
-\begin{tikzcd}[column sep=large]
-\ctxext{\Gamma}{A} \arrow[drr,bend left=15,equals] \arrow[ddr,bend right=15,equals] \arrow[dr,densely dotted,near end,"{\tfid{A}}"] \\
-& \ctxext{{\Gamma}{A}}{W_A(A)} \arrow[r,swap,"{\pi_2(A,A)}"] \arrow[d,fib] & \ctxext{\Gamma}{A} \arrow[d,fib] \\
-& \ctxext{\Gamma}{A} \arrow[r,swap,"A"] & \Gamma
-\end{tikzcd}
-\end{equation*}
-As an immediate consequence of \autoref{lem:pb_selfdistributive}, we get that
-each functor $W_A$ and $S_x$ is both a weakening functor and a substitution functor.
-Since each $W_A$ and $S_x$ preserves identity morphisms, it also preserves the
-terms $\tfid{A}$.
-
-It remains to verify the rest of the conditions of E-systems.
-\begin{itemize}
-\item To show that weakened families are constant, note that
-\begin{equation*}
-S_x\circ W_A\jdeq x_\ast\circ A_\ast\jdeq (A\circ x)_\ast\jdeq (\catid{\Gamma})_\ast\jdeq \catid{\cat{C}_{\cat{F}}/\Gamma}.
-\end{equation*}
-\item The identity terms are neutral for pre-composition:
-\begin{equation*}
-S_{\tfid{A}}\circ W_A/A \jdeq S_{\tfid{A}}\circ \pi_2(A,A)_\ast \jdeq (\pi_2(A,A)\circ\tfid{A})_\ast \jdeq
-(\catid{\ctxext{\Gamma}{A}})_\ast \jdeq \catid{\cat{C}_{\cat{F}}/\ctxext{\Gamma}{A}} 
-\end{equation*}
-\item The identity terms behave like identity functions: by the universal property,
-$S_x(\tfid{A})$ is the unique section of $A$ such that the square
-\begin{equation*}
-\begin{tikzcd}[column sep=6em]
-\Gamma \arrow[d,swap,"{S_x(\tfid{A})}"] \arrow[r,"{\pi_2(x,\catid{\ctxext{\Gamma}{A}})}"] & \ctxext{\Gamma}{A} \arrow[d,"{\tfid{A}}"] \\
-\ctxext{\Gamma}{A} \arrow[r,swap,"{\pi_2(x,W_A(A))}"] & \ctxext{{\Gamma}{A}}{W_A(A)}
-\end{tikzcd}
-\end{equation*}
-commutes. Thus, it suffices to show that this square also commutes with $x$ in the place of
-$S_x(\tfid{A})$. Note that $\pi_2(x,\catid{\ctxext{\Gamma}{A}})\jdeq x$. 
-Since $\ctxext{{\Gamma}{A}}{W_A(A)}$ is itself a pullback, it suffices
-and it is straightforward to verify the equalities
-\begin{align*}
-W_A(A)\circ\pi_2(x,W_A(A))\circ x & \jdeq W_A(A)\circ\idtm{A}\circ x \\
-\pi_2(A,A)\circ\pi_2(x,W_A(A))\circ x & \jdeq \pi_2(A,A)\circ\idtm{A}\circ x.\qedhere
-\end{align*}
-\end{itemize}
-\end{proof}
-
-\begin{comment}
-We then demonstrate how to define a category internal to an E-system. The main
-point of this lies mainly in demonstrating how the operations of E-systems can
-be used to define a category, because it is no surprise that a category can be
-defined in type theory. Nevertheless, it is noteworthy that we need no further 
-conditions on E-systems.
-This is unlike the situation in category theory, where we need
-to assume that the ambient category has finite limits in order to say what an
-internal category is. In fact, it is possible to define an E-system internal
-to an E-system, just using the algebraic structure of an E-system, and actually
-defining an internal category was the hardest part of that definition.
-\end{comment}
-
-\subsection{The category of internal morphisms}\label{subsec:im_cat}
-
-\begin{defn}
-For every $A,B\in\cat{F}/\Gamma$ we define the set 
-\begin{equation*}
-\thom{A}{B}\defeq T(\ctxwk{A}{B}).
-\end{equation*}
-An element $f\in \thom{A}{B}$ is called an \emph{internal morphism in context
-$\Gamma$}. We sometimes write $\jhom{\Gamma}{A}{B}{f}$ to indicate that $f$ is
-an internal morphism over $\Gamma$, or we may draw a diagram of the form
-\begin{equation*}
-\begin{tikzcd}[column sep=tiny]
-A \arrow[rr,"f"] \arrow[dr,fib] & & B \arrow[dl,fib] \\
-& \Gamma
-\end{tikzcd}
-\end{equation*}
-but we shall generally omit the arrows down to $\Gamma$ and say instead that we have
-a diagram in context $\Gamma$. 
-\end{defn}
-
-\begin{rmk}
-Note that $\thom{\catid{\Gamma}}{A}\jdeq T(A)$ for any $A\in\cat{F}/\Gamma$, 
-because we have 
-$W_{\catid{\ctxext{\Gamma}{A}}}\jdeq \catid{\cat{F}/\ctxext{\Gamma}{A}}$.
-
-Note also that $\thom{\ctxext{A}{P}}{B}\jdeq \thom{P}{\ctxwk{A}{B}}$ 
-for any $P\in\cat{F}/\ctxext{\Gamma}{A}$ and $B\in\cat{F}/\Gamma$,
-because $W_{\ctxext{A}{P}}\jdeq W_P\circ W_A$.
-Once we have established a category of which the morphisms are given by 
-$\thom{\blank}{\blank}$, we therefore get that 
-\begin{equation*}
-\ctxext{A}{(\blank)}\dashv W_A.
-\end{equation*}
-The right adjoint to weakening by $A$, if it exists, will be the dependent
-product $\Pi_A$. 
-\end{rmk}
-
-\begin{defn}
-Let $A,B\in\cat{F}/\Gamma$.
-For any $f\in\thom{A}{B}$ we define the pre-composition E-homomorphism
-\begin{equation*}
-f_\ast \defeq S_f\circ W_A/B : \cat{F}/\ctxext{\Gamma}{B}\to\cat{F}/\ctxext{\Gamma}{A}.
-\end{equation*}
-Furthermore, for $g\in\thom{B}{C}$ we may consider $W_B(C)\in\cat{F}/\ctxext{\Gamma}{B}$ as a morphism from $W_B(C)$ to $\catid{\ctxext{\Gamma}{B}}$. Therefore, $f_\ast$ acts on the terms of
-$W_B(C)$, so it makes sense to define $g\circ f\defeq f_\ast(g)$. In fact, it makes
-sense to let $(\blank)\circ f$ be the infix notation for $f_\ast$. 
-\end{defn}
-
-\begin{rmk}
-Note that for $\jhom{\Gamma}{B}{C}{g}$, 
-we have $g\circ f\in T(S_f(W_A/B(W_B(C))))$, whereas we would like that
-$g\circ f\in\thom{A}{C}$. More generally, we can show that
-\begin{equation*}
-S_f\circ (W_A/B)\circ W_B\jdeq W_A.
-\end{equation*}
-Since weakening is a weakening homomorphism, we have
-\begin{equation*}
-S_f\circ (W_A/B)\circ W_B\jdeq S_f\circ W_{W_A(B)}\circ W_A
-\end{equation*}
-By \autoref{tTf:StWf_id} of \autoref{defn:esystems} it we get that
-\begin{equation*}
-S_f\circ W_{W_A(B)}\circ W_A\jdeq W_A.
-\end{equation*}
-\end{rmk}
-
-\begin{rmk}
-To see how this extended notion of precomposition works, suppose we have a diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=tiny]
-Q \arrow[rr,"g"] \arrow[dr,fib] & & R \arrow[dl,fib] \\
-& B
-\end{tikzcd}
-\end{equation*}
-in context $\Gamma$, i.e.~$\jhom{\ctxext{\Gamma}{B}}{Q}{R}{g}$, and let $\jhom{\Gamma}{A}{B}{f}$. Precomposing by $f$ gives
-us a diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=tiny]
-\jcomp{A}{f}{Q} \arrow[rr,"\jcomp{A}{f}{g}"] \arrow[dr,fib] & & \jcomp{A}{f}{R} \arrow[dl,fib] \\
-& A
-\end{tikzcd}
-\end{equation*}
-in context $\Gamma$.
-\end{rmk}
-
-\begin{rmk}
-Note that \autoref{f:SidfWff_id} of \autoref{defn:esystems} asserts precisely
-that ${(\tfid{A})}_\ast\jdeq \catid{\cat{F}/\ctxext{\Gamma}{A}}$ for any $A\in\cat{F}/\Gamma$. In
-particular, it follows that $g\circ\tfid{A}\jdeq g$ for any $g\in\thom{A}{B}$
-\end{rmk}
-
-\begin{lem}\label{lem:compcomp}
-For any $f\in\thom{A}{B}$ and $g\in\thom{B}{C}$ we have $f_\ast\circ g_\ast\jdeq (g\circ f)_\ast$. 
-\end{lem}
-
-\begin{proof}
-\begin{align*}
-f_\ast\circ g_\ast & \jdeq S_f\circ (W_A/B) \circ S_g \circ (W_B/C)\\
-& \jdeq S_f\circ S_{W_A(g)}\circ(W_A/\ctxext{B}{W_B(C)})\circ (W_B/C) \\
-& \jdeq S_{S_f(W_A(g))}\circ (S_f/W_A(W_B(C)))\circ (W_A/\ctxext{B}{W_B(C)}) \circ W_B/C \\
-& \jdeq S_{S_f(W_A(g))}\circ ((S_f\circ (W_A/B) \circ W_B)/C) \\
-& \jdeq S_{S_f(W_A(g))}\circ ((S_f\circ W_{W_A(B)}\circ W_A)/C) \\
-& \jdeq S_{S_f(W_A(g))}\circ W_A/C \\
-& \jdeq (g\circ f)_\ast.\qedhere
-\end{align*}
-\end{proof}
-
-\begin{thm}
-For each object $\Gamma$ of an E-system $\mathbb{E}$, we have a category
-$\mathbb{C}_I(\mathbb{E},\Gamma)$ of internal morphisms of $\mathbb{E}$,
-consisting of the same objects as $\cat{F}/\Gamma$ and the hom-sets 
-$\thom{A}{B}$. 
-\end{thm}
-
-\begin{proof}
-The fact that composition is associative is a direct corollary of
-\autoref{lem:compcomp}. The axiom $(\tfid{A})_\ast\jdeq\catid{\ctxext{\Gamma}{A}}$
-implies that the identity morphisms satisfy the right identity law. It remains
-to show that $\tfid{B}\circ f\jdeq f$. This is a simple calculation:
-\begin{equation*}
-\tfid{B}\circ f 
-  \jdeq
-\subst{f}{\ctxwk{A}{\tfid{B}}}
-  \jdeq
-\subst{f}{\tfid{\ctxwk{A}{B}}}
-  \jdeq
-f.\qedhere
-\end{equation*}
-\end{proof}
-
-\begin{rmk}
-Since any E-homomorphism is both a substitution homomorphism and a weakening
-homomorphism, it follows immediately that any E-homomorphism preserves any
-$f_\ast$, i.e.~for any $f\in\thom{A}{B}$ in $\mathcal{E}$ and any E-homomorphism
-$F:\mathcal{E}\to\mathcal{D}$, the diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=large]
-\mathcal{E}/\ctxext{\Gamma}{A} \arrow[r,"F/\ctxext{\Gamma}{A}"] & \mathcal{D}/F(\ctxext{\Gamma}{A})
-  \\
-\mathcal{E}/\ctxext{\Gamma}{B} \arrow[u,"f_\ast"] \arrow[r,swap,"F/\ctxext{\Gamma}{B}"] & \mathcal{D}/F(\ctxext{\Gamma}{B}) \arrow[u,swap,"F(f)_\ast"]
-\end{tikzcd}
-\end{equation*}
-commutes. This applies to weakening, substitution and pre-composition, and it
-follows that for any E-homomorphism $F:\mathcal{E}\to\mathcal{D}$ and any
-$\Gamma\in\mathcal{E}$, we have a functor $\mathbb{C}_I(F,\Gamma):
-\mathbb{C}_I(\mathcal{E},\Gamma)\to\mathbb{C}_I(\mathcal{D},F(\Gamma))$.
-\end{rmk}
-
-The following fact will be useful:
-
-\begin{lem}\label{lem:compW_W}
-Let $f\in\thom{A}{B}$ be an internal morphism in context $\Gamma$. Then one has
-\begin{equation*}
-f_\ast\circ W_B\jdeq W_A.
-\end{equation*}
-\end{lem}
-
-\begin{proof}
-The proof is a simple calculation:
-\begin{equation*}
-f_\ast\circ W_B\jdeq S_f\circ W_A/B\circ W_B\jdeq S_f\circ W_{W_A(B)}\circ W_A
-\jdeq W_A.\qedhere
-\end{equation*}
-\end{proof}
-
-\subsection{Pairing and the projections}
-
-The composition $\ctxext{A}{P}$ of $A\in\cat{F}/\Gamma$ and $P\in\cat{F}/\ctxext{\Gamma}{A}$
-behaves like a strict $\Sigma$-type. We may define the pairing term
-$\typefont{pair}^{A,P}\defeq\tfid{\ctxext{A}{P}}\in T(W_P(W_A(\ctxext{A}{P})))$
-and the projections and prove several useful properties about them. The strictness
-is found, among other things, in the fact that we can prove judgmental $\eta$-equality,
-and that pairing is strictly associative.
-
-In this section we will make more extensive use of the infix form of the
-weakening and substitution operations.
-
-\begin{defn}
-Let $x\in T(A)$ and $u\in T(S_x(P))$ for $A\in\cat{F}/\Gamma$ and $P\in\cat{F}/\ctxext{\Gamma}{A}$. 
-We define the \emph{term extension of $x$ and $u$} to be
-\begin{equation*}
-\tmext{x}{u}\defeq \subst{u}{{x}{\tfid{\ctxext{A}{P}}}}\in T(\ctxext{A}{P}).
-\end{equation*}
-\end{defn}
-
-To prove anything about the term $\tmext{x}{u}$, we need the following property.
-
-\begin{thm}\label{subst_by_tmext}
-Let $x\in T(A)$ and $u\in T(S_x(P))$ for $A\in\cat{F}/\Gamma$ and $P\in\cat{F}/\ctxext{\Gamma}{A}$.
-Then we have
-\begin{equation*}
-S_{\tmext{x}{u}}\jdeq S_u\circ (S_x/P):\cat{F}/\ctxext{\Gamma}{{A}{P}}\to\cat{F}/\Gamma
-\end{equation*}
-\end{thm}
-
-\begin{proof}
-\begin{align*}
-S_{\tmext{x}{u}} & \jdeq S_{S_u(S_x(\tfid{\ctxext{A}{P}}))} \\
-& \jdeq S_{S_u(S_x(\tfid{\ctxext{A}{P}}))}\circ (S_u\circ W_{S_x(P)})\circ (S_x\circ W_{A}) \\
-& \jdeq S_u \circ S_{S_x(\tfid{\ctxext{A}{P}})} \circ W_{S_x(P)}\circ S_x\circ W_A \\
-& \jdeq S_u \circ (S_x/P) \circ S_{\tfid{\ctxext{A}{P}}} \circ W_P\circ W_A \\
-& \jdeq S_u \circ (S_x/P) \circ S_{\tfid{\ctxext{A}{P}}} \circ W_{\ctxext{A}{P}}\\
-& \jdeq S_u \circ (S_x/P).\qedhere
-\end{align*}
-\end{proof}
-
-\begin{defn}
-Let $A\in\cat{F}/\Gamma$ and $P\in\cat{F}/\ctxext{\Gamma}{A}$. We define
-\begin{align*}
-\cprojfstf{A}{P} & \defeq \ctxwk{P}{\tfid{A}}\in\thom{\ctxext{A}{P}}{A}\\
-\cprojsndf{A}{P} & \defeq \tfid{P}\in\thomd{\cprojfstf{A}{P}}{\ctxext{A}{P}}{P}
-\end{align*}
-\end{defn}
-
-\begin{cor}\label{cor:pairing_ehom}
-Let $F:\mathcal{E}\to\mathcal{D}$ be an E-homomorphism and let $x\in T(A)$ and
-$u\in T(S_x(P))$. Then we have
-\begin{equation*}
-F(\tmext{x}{u})\jdeq \tmext{F(x)}{F(u)}.
-\end{equation*}
-Also, we have
-\begin{align*}
-F(\cprojfstf{A}{P}) & \jdeq \cprojfstf{F(A)}{F(P)} \\
-F(\cprojsndf{A}{P}) & \jdeq \cprojsndf{F(A)}{F(P)}.
-\end{align*}
-\end{cor}
-
-\begin{thm}\label{thm:famext_up}
-For any $A\in\cat{F}/\Gamma$ and $P\in\cat{F}/\ctxext{\Gamma}{A}$,
-the map
-\begin{equation*}
-(x,u)\mapsto \tmext{x}{u}:\big(\bigsqcup\nolimits_{x\in T(A)} T(\subst{x}{P})\big)\to T(\ctxext{A}{P}).
-\end{equation*}
-is an isomorphism.  
-\end{thm}
-
-\begin{proof}
-Lets write $\varphi$ for the indicated map. The inverse $\psi$ of $\varphi$ is defined
-by $w\mapsto (\subst{w}{\cprojfstf{A}{P}},\subst{w}{\cprojsndf{A}{P}})$. We have
-to show that for any $x\in T(A)$ and $u\in T(\subst{x}{P})$, one has
-\begin{align*}
-\subst{\tmext{x}{u}}{\cprojfstf{A}{P}} & \jdeq x\\
-\subst{\tmext{x}{u}}{\cprojsndf{A}{P}} & \jdeq u,
-\end{align*}
-and that for any $w\in T(\ctxext{A}{P})$, one has
-\begin{equation*}
-\tmext{\subst{w}{\cprojfstf{A}{P}}}{\subst{w}{\cprojsndf{A}{P}}} \jdeq w.
-\end{equation*}
-To show that $\subst{\tmext{x}{u}}{\cprojfstf{A}{P}} \jdeq x$, we use that
-$S_{\tmext{x}{u}}\jdeq S_u\circ S_x/P$ to show that
-\begin{align*}
-\subst{\tmext{x}{u}}{\cprojfstf{A}{P}}
-  & \jdeq 
-\subst{\tmext{x}{u}}{\ctxwk{P}{\tfid{A}}} \\
-  & \jdeq
-\subst{u}{{x}{\ctxwk{P}{\tfid{A}}}} \tag{By \autoref{subst_by_tmext}}\\
-  & \jdeq
-\subst{u}{\ctxwk{\subst{x}{P}}{\subst{x}{\tfid{A}}}} \\
-  & \jdeq
-\subst{x}{\tfid{A}} \\
-  & \jdeq
-x
-\end{align*}
-To show that $\subst{\tmext{x}{u}}{\cprojsndf{A}{P}}\jdeq u$, note that
-\begin{equation*}
-\subst{\tmext{x}{u}}{\cprojsndf{A}{P}}
-  \jdeq
-\subst{u}{{x}{\tfid{P}}}
-  \jdeq
-\subst{u}{\tfid{\subst{x}{P}}}
-  \jdeq
-u
-\end{equation*}
-Therefore, $\psi$ is a left-inverse of $\varphi$. To show that it is also a
-right inverse, note that \autoref{cor:pairing_ehom} gives us that
-\begin{equation*}
-\tmext{\cprojfst{A}{P}{w}}{\cprojsnd{A}{P}{w}}
-  \jdeq
-\subst{w}{(\tmext{\cprojfstf{A}{P}}{\cprojsndf{A}{P}})}.
-\end{equation*}
-Thus, it suffices to show that $\tmext{\cprojfstf{A}{P}}{\cprojsndf{A}{P}}\jdeq
-\tfid{\ctxext{A}{P}}$, which is straightforward.
-\end{proof}
-
-\begin{lem}\label{lem:tmext_assoc}
-For any $x\in T(A)$, $u\in T(S_x(P))$ and $v\in T(S_{\tmext{x}{u}}(Q))$ we have
-\begin{equation*}
-\tmext{{x}{u}}{v}\jdeq \tmext{x}{{u}{v}}\in T(\ctxext{{A}{P}}{Q}).
-\end{equation*}
-\begin{comment}
-Also, we have
-\begin{align*}
-\cprojfstf{A}{P}\circ\cprojfstf{\ctxext{A}{P}}{Q} & \jdeq\cprojfstf{A}{\ctxext{P}{Q}} \\
-\cprojsndf{A}{P}\circ\cprojfstf{\ctxext{A}{P}}{Q} & \jdeq \cprojfstf{P}{Q}\circ\cprojsndf{A}{\ctxext{P}{Q}} \\
-\cprojsndf{\ctxext{A}{P}}{Q} & \jdeq \cprojsndf{P}{Q}\circ\cprojsndf{A}{\ctxext{P}{Q}}.
-\end{align*}
-\end{comment}
-\end{lem}
-
-\begin{proof}
-By \autoref{subst_by_tmext}, we have $S_v\circ S_{\tmext{x}{u}}\jdeq S_v\circ S_u\circ S_x \jdeq S_{\tmext{u}{v}}\circ S_x$,
-so associativity of term extension follows.
-\end{proof}
-
-\begin{comment}
-\begin{lem}
-Let $f\in\thom{A}{\ctxext{B}{Q}}$ Then we have
-\begin{equation*}
-f \jdeq \tmext{\jcomp{A}{f}{\cprojfstf{B}{Q}}}{\jcomp{A}{f}{\cprojsndf{B}{Q}}}.
-\end{equation*}
-Alternatively, when $f_0\in\thom{A}{B}$ and $f_1\in\thom{A}{\jcomp{A}{f}{Q}}$,
-then we have $\tmext{f_0}{f_1}\in\thom{A}{\ctxext{B}{Q}}$ and
-\begin{align*}
-\jcomp{A}{\tmext{f_0}{f_1}}{\cprojfstf{B}{Q}} & \jdeq f_0 \\
-\jcomp{A}{\tmext{f_0}{f_1}}{\cprojsndf{B}{Q}} & \jdeq f_1.
-\end{align*}
-\end{lem}
-
-\begin{proof}
-Straightforward
-\end{proof}
-\end{comment}
-
-We end this section by showing that weakening is the same operation as pre-composition
-with a first projection.
-
-\begin{thm}\label{precomp_by_proj}
-Precomposition with $\cprojfstf{A}{P}$ is weakening by $P$, i.e.
-\begin{equation*}
-(\cprojfstf{A}{P})_\ast \jdeq W_P.
-\end{equation*}
-\end{thm}
-
-\begin{proof}
-\begin{align*}
-(\cprojfstf{A}{P})_\ast & \jdeq S_{\cprojfstf{A}{P}}\circ W_{\ctxext{A}{P}} \\
-& \jdeq S_{W_P(\tfid{A})}\circ W_P\circ W_A \\
-& \jdeq W_P\circ S_{\tfid{A}}\circ W_A \\
-& \jdeq W_P.\qedhere 
-\end{align*}
-\end{proof}
-
-\subsection{The interchange laws}\label{sec:interchange}
-
-We are now in the position to define vertical and horizontal composition, and
-prove properties of them.
-
-\begin{defn}
-Let $\jhom{\Gamma}{A}{B}{f}$ and $\jfhom{\Gamma}{A}{B}{f}{P}{Q}{F}$. Then we
-define 
-\begin{equation*}
-\jhomdefn{\Gamma}{{A}{P}}{{B}{Q}}{\jvcomp{P}{f}{F}}{\unfold{\jvcomp{P}{f}{F}}}.
-\end{equation*}
-Whenever we say that we have a diagram of the form
-\begin{equation*}
-\begin{tikzcd}
-R \arrow[r,"f_2"] \arrow[d,fib] &
-S \arrow[d,fib] \\
-P \arrow[r,"f_1"] \arrow[d,fib] &
-Q \arrow[d,fib] \\
-A \arrow[r,"f_0"] &
-B
-\end{tikzcd}
-\end{equation*}
-we mean that we have
-$f_0\in\thom{A}{B}$, $f_1\in\thomd{f_0}{P}{Q}$ and 
-$f_2\in\thomd{\jvcomp{P}{f_0}{f_1}}{R}{S}$.
-\end{defn}
-
-\begin{cor}
-Let $\jhom{\Gamma}{A}{B}{f}$ and $\jfhom{\Gamma}{A}{B}{f}{P}{Q}{F}$. Then we
-have
-\begin{equation*}
-H(\jvcomp{P}{f}{F})\jdeq \jvcomp{H(P)}{H(f)}{H(F)}
-\end{equation*}
-for any E-homomorphism $H:\mathbb{E}\to\mathbb{D}$.
-\end{cor}
-
-\begin{lem}
-Vertical composition is associative.
-\end{lem}
-
-\begin{proof}
-Consider the diagram
-\begin{equation*}
-\begin{tikzcd}
-R \arrow[r,"f_2"] \arrow[d,fib] &
-S \arrow[d,fib] \\
-P \arrow[r,"f_1"] \arrow[d,fib] &
-Q \arrow[d,fib] \\
-A \arrow[r,"f_0"] &
-B
-\end{tikzcd}
-\end{equation*}
-in context $\Gamma$.
-Because weakening distributes over term extension, and term extension is
-associative, we have
-\begin{align*}
-\jvcomp{R}{\jvcomp{P}{f_0}{f_1}}{f_2}
-  & \jdeq
-\tmext{\ctxwk{R}{\tmext{\ctxwk{P}{f_0}}{f_1}}}{f_2}
-  \\
-  & \jdeq
-\tmext{\tmext{\ctxwk{R}{{P}{f_0}}}{\ctxwk{R}{f_1}}}{f_2}
-  \\
-  & \jdeq
-\tmext{\ctxwk{\ctxext{P}{R}}{f_0}}{\tmext{\ctxwk{R}{f_1}}{f_2}}
-  \tag{By \autoref{lem:tmext_assoc}}
-  \\
-  & \jdeq
-\jvcomp{{P}{R}}{f_0}{\jvcomp{R}{f_1}{f_2}}.\qedhere
-\end{align*}
-\end{proof}
-
-\begin{defn}
-Let $\jhom{\Gamma}{A}{B}{f}$ and $\jfhom{\Gamma}{A}{B}{f}{P}{Q}{F}$. Then we
-define the E-homomorphism 
-\begin{equation*}
-F_\circledast\defeq F_\ast\circ (f_\ast/Q):
-\mathbb{E}/\ctxext{{\Gamma}{B}}{Q}\to\mathbb{E}/\ctxext{{\Gamma}{A}}{P}.
-\end{equation*}
-The infix notation of $F_\circledast$ is taken to be $\jfcomp{\blank}{\blank}{\blank}{F}{\blank}$.
-\end{defn}
-
-\begin{lem}\label{lem:three-composition}
-Let $\jhom{\Gamma}{A}{B}{f}$ and $\jfhom{\Gamma}{A}{B}{f}{P}{Q}{F}$. Then we
-have the equality
-\begin{equation*}
-F_\circledast\jdeq (\jvcomp{P}{f}{F})_\ast.
-\end{equation*}
-\end{lem}
-
-\begin{proof}
-\begin{align*}
-F_\ast\circ (f_\ast/Q)
-  & \jdeq
-S_F\circ W_P\circ S_f/(W_A(Q))\circ W_A/\ctxext{B}{Q}
-  \\
-  & \jdeq
-S_F\circ S_{W_P(f)}/W_P(W_A(Q))\circ W_P/W_A(\ctxext{B}{Q})\circ W_A/\ctxext{B}{Q}
-  \\
-  & \jdeq
-S_F\circ S_{W_P(f)}/W_P(W_A(Q))\circ W_{\ctxext{A}{P}}/\ctxext{B}{Q}
-  \\
-  & \jdeq
-S_{\tmext{W_P(f)}{F}}\circ W_{\ctxext{A}{P}}/\ctxext{B}{Q}
-  \tag{By \autoref{subst_by_tmext}}
-  \\
-  & \jdeq
-(\jvcomp{P}{f}{F})_\ast.\qedhere
-\end{align*}
-\end{proof}
-
-We are now in position to state and prove the interchange law of horizontal
-and vertical composition.
-
-\begin{thm}\label{thm:interchange}
-Consider the diagram
-\begin{equation*}
-\begin{tikzcd}
-P \arrow[r,"F"] \arrow[d,fib] &
-Q \arrow[r,"G"] \arrow[d,fib] &
-R \arrow[d,fib] \\
-A \arrow[r,"f"] &
-B \arrow[r,"g"] &
-C
-\end{tikzcd}
-\end{equation*}
-in context $\Gamma$. Then the equality
-\begin{equation*}
-{\jcomp{{A}{P}}{\jvcomp{P}{f}{F}}{\jvcomp{Q}{g}{G}}}
-  \jdeq
-{\jvcomp{P}{\jcomp{A}{f}{g}}{\jfcomp{A}{f}{P}{F}{G}}}
-\end{equation*}
-of morphisms from $\ctxext{A}{P}$ to $\ctxext{C}{R}$ in context $\Gamma$ holds.
-\end{thm}
-
-\begin{proof}
-By \autoref{lem:three-composition}, we have
-\begin{align*}
-\jcomp{{A}{P}}{\jvcomp{P}{f}{F}}{\jvcomp{Q}{g}{G}}
-  & \jdeq
-\jcomp{P}{F}{\jcomp{A}{f}{\tmext{\ctxwk{Q}{g}}{G}}}
-  \tag{By \autoref{lem:three-composition}}
-  \\
-  & \jdeq
-\jcomp{P}{F}{\tmext{\ctxwk{\jcomp{A}{f}{Q}}{\jcomp{A}{f}{g}}}{\jcomp{A}{f}{G}}}
-  \\
-  & \jdeq
-\tmext{\jcomp{P}{F}{\ctxwk{\jcomp{A}{f}{Q}}{\jcomp{A}{f}{g}}}}{\jcomp{P}{F}{{A}{f}{G}}}
-  \\
-  & \jdeq
-\tmext{\jcomp{P}{F}{\ctxwk{\jcomp{A}{f}{Q}}{\jcomp{A}{f}{g}}}}{\jfcomp{A}{f}{P}{F}{G}}
-  \\
-  & \jdeq
-\tmext{\ctxwk{P}{\jcomp{A}{f}{g}}}{\jfcomp{A}{f}{P}{F}{G}}
-  \tag{By \autoref{lem:compW_W}}
-  \\
-  & \jdeq
-\jvcomp{P}{\jcomp{A}{f}{g}}{\jfcomp{A}{f}{P}{F}{G}}.\qedhere
-\end{align*}
-\end{proof}
-
-\begin{thm}
-Consider the diagram
-\begin{equation*}
-\begin{tikzcd}
-P \arrow[r,"F"] \arrow[d,fib] &
-Q \arrow[r,"G"] \arrow[d,fib] &
-R \arrow[d,fib] \\
-A \arrow[r,"f"] &
-B \arrow[r,"g"] &
-C
-\end{tikzcd}
-\end{equation*}
-in context $\Gamma$. 
-Then $F_\circledast\circ G_\circledast\jdeq (\jfcomp{A}{f}{P}{F}{G})_\circledast$.
-In other words the composition $\jfcomp{}{}{}{\blank}{\blank}$ is associative.
-\end{thm}
-
-\begin{proof}
-\begin{align*}
-F_\circledast\circ G_\circledast
-  & \jdeq
-(\jvcomp{P}{f}{F})_\ast\circ(\jvcomp{Q}{g}{G})_\ast 
-  \tag{By \autoref{lem:three-composition}}\\
-  & \jdeq
-(\jcomp{\ctxext{A}{P}}{\jvcomp{P}{f}{F}}{\jvcomp{Q}{g}{G}})_\ast 
-  \tag{By \autoref{lem:compcomp}}\\
-  & \jdeq
-(\jvcomp{}{\jcomp{A}{f}{g}}{\jfcomp{A}{f}{P}{F}{G}})_\ast 
-  \tag{By \autoref{thm:interchange}} \\
-  & \jdeq
-(\jfcomp{A}{f}{P}{F}{G})_\circledast.
-  \tag{By \autoref{lem:three-composition}}
-\end{align*}
-\end{proof}
-
-\begin{lem}
-Let $\jhom{\Gamma}{A}{B}{f}$ and $\jfhom{\Gamma}{A}{B}{f}{P}{Q}{F}$. Then
-$\jvcomp{P}{f}{F}$ is the unique morphism from $\ctxext{A}{P}$ to $\ctxext{B}{Q}$
-with the property that both the diagram
-\begin{equation*}
-\begin{tikzcd}
-\ctxext{A}{P}
-  \ar{r}{\jvcomp{P}{f}{F}}
-  \ar{d}[swap]{\cprojfstf{A}{P}}
-& \ctxext{B}{Q}
-  \ar{d}{\cprojfstf{B}{Q}}
-  \\
-A \ar{r}[swap]{f}
-& B
-\end{tikzcd}
-\end{equation*}
-commutes, and $\jcomp{}{\jvcomp{P}{f}{F}}{\cprojsndf{B}{Q}}\jdeq F$.
-\end{lem}
-
-\begin{proof}
-We first note that
-\begin{align*}
-\jcomp{}{\jvcomp{P}{f}{F}}{\cprojfstf{B}{Q}}
-  & \jdeq
-\jcomp{P}{F}{\jcomp{A}{f}{\cprojfstf{B}{Q}}} 
-  \tag{By \autoref{lem:three-composition}} \\
-  & \jdeq
-\jcomp{P}{F}{\ctxwk{\jcomp{A}{f}{Q}}{\jcomp{A}{f}{\idtm{B}}}} \\
-  & \jdeq
-\jcomp{P}{F}{\ctxwk{\jcomp{A}{f}{Q}}{f}} \\
-  & \jdeq
-\ctxwk{P} f \tag{By \autoref{lem:compW_W}} \\
-  & \jdeq
-\jcomp{\ctxext{A}{P}}{\cprojfstf{A}{P}}{f}.
-  \tag{By \autoref{precomp_by_proj}}
-\end{align*}
-Also, we have
-\begin{align*}
-\jcomp{}{\jvcomp{P}{f}{F}}{\cprojsndf{B}{Q}}
-  & \jdeq
-\jcomp{P}{F}{\jcomp{A}{f}{\idtm{Q}}}
-  \tag{By \autoref{lem:three-composition}}
-  \\
-  & \jdeq
-\jcomp{P}{F}{\idtm{\jcomp{A}{f}{Q}}}
-  \\
-  & \jdeq
-F.
-\end{align*}
-Thus, we conclude that $\jvcomp{P}{f}{F}$ has indeed the stated property. For
-the uniqueness, let $G:\ctxext{A}{P}\to\ctxext{B}{Q}$ be a morphism such that
-$\jcomp{}{G}{\cprojfstf{B}{Q}}\jdeq\jcomp{}{\cprojfstf{A}{P}}{f}$ and
-$\jcomp{}{G}{\cprojsndf{B}{Q}}\jdeq F$. Then it follows that
-\begin{equation*}
-G \jdeq \tmext{\jcomp{}{\cprojfstf{A}{P}}{f}}{F}
-  \jdeq \tmext{\ctxwk{P}{f}}{F}
-  \jdeq \jvcomp{P}{f}{F}.\qedhere
-\end{equation*}
-\end{proof}
-
-\subsection{E-systems are E'-systems}
-
-\begin{lem}
-Consider $A\in\cat{F}/\Gamma$ and $Q\in\cat{F}/\ctxext{\Gamma}{B}$. Then there
-is an isomorphism
-\begin{equation*}
-\{h\in\thom{A}{\ctxext{B}{Q}}\mid \jcomp{A}{h}{\cprojfstf{B}{Q}}\jdeq f\}\cong T(\jcomp{A}{f}{Q}).
-\end{equation*}
-for any $f\in\thom{A}{B}$.
-\end{lem}
-
-\begin{proof}
-We have the isomorphism
-\begin{align*}
-\thom{A}{\ctxext{B}{Q}} 
-& \jdeq T(\ctxwk{A}{\ctxext{B}{Q}}) \\
-& \jdeq T(\ctxext{\ctxwk{A}{B}}{\ctxwk{A}{Q}}) \\
-& \cong \bigsqcup\nolimits_{(f\in T(\ctxwk{A}{B}))}T(\subst{f}{\ctxwk{A}{Q}})
-\tag{By \autoref{thm:famext_up}} \\
-& \jdeq \bigsqcup\nolimits_{(f\in\thom{A}{B})}T(\jcomp{A}{f}{Q}).
-\end{align*}
-Also, we find $\cprojfst{\ctxwk{A}{B}}{\ctxwk{A}{Q}}{h}\jdeq\subst{h}{\ctxwk{A}{\cprojfstf{B}{Q}}}\jdeq
-\jcomp{A}{h}{\cprojfstf{B}{Q}}$.
-\end{proof}
-
-\begin{thm}
-E-systems are the same things as E'-systems.
-\end{thm}
-
-\begin{proof}
-For any $f:A\to B$ in context $\Gamma$ and any $Q\in\cat{F}/B$, we define
-\begin{equation*}
-\pi_2(f,Q)\defeq \jvcomp{\jcomp{A}{f}{Q}}{f}{\tfid{\jcomp{A}{f}{Q}}}
-  : \ctxext{A}{\jcomp{A}{f}{Q}}\to\ctxext{B}{Q}.
-\end{equation*}
-Then we have the commuting diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=huge]
-\ctxext{A}{\jcomp{A}{f}{Q}} \arrow[r,"{\pi_2(f,Q)}"] \arrow[d,swap,"{\cprojfstf{A}{\jcomp{A}{f}{Q}}}"] & \ctxext{B}{Q} \arrow[d,"{\cprojfstf{B}{Q}}"] \\
-A \arrow[r,"f"] & B
-\end{tikzcd}
-\end{equation*}
-in context $\Gamma$. To show that this is a pullback diagram, consider a
-morphism $g:X\to A$ in context $\Gamma$. Then we have the isomorphisms
-\begin{equation*}
-\{h\in\thom{X}{\ctxext{B}{Q}}\mid \jcomp{X}{h}{\cprojfstf{B}{Q}}\jdeq\jcomp{X}{g}{f}\}
-  \cong
-T((\jcomp{X}{g}{f})_\ast(Q))
-  \cong
-\{u\in\thom{X}{\ctxext{A}{\jcomp{A}{f}{Q}}}\mid\jcomp{X}{u}{\cprojfstf{A}{\jcomp{A}{f}{Q}}}\jdeq g\}
-\end{equation*}
-Thus, we find for every $h:X\to\ctxext{B}{Q}$ satisfying
-$\jcomp{X}{h}{\cprojfstf{B}{Q}}\jdeq\jcomp{X}{g}{f}$, a unique morphism
-$u:X\to\ctxext{A}{\jcomp{A}{f}{Q}}$ satisfying
-$\jcomp{X}{u}{\cprojfstf{A}{\jcomp{A}{f}{Q}}}\jdeq g$. It is easy to verify
-that $\jcomp{X}{u}{\pi_2(f,Q)}\jdeq h$, so the universal property of pullbacks
-holds. The functoriality conditions follow immediately from the interchange
-laws proven in \autoref{sec:interchange}.
-
-We have described functors from E'-systems to E-systems and vice versa. It is
-a straightforward check that these are mutual inverses.
-\end{proof}
-
-\begin{comment}
-\subsection{Categories internal to E-systems}
-The definition of internal category we're about to give, has a type of objects
-and a type of morphisms, such that the codomain function is a projection. It might
-be more appropriate to call such structures `right categories', and the corresponding
-structures where the domain function is a projection might be called `left
-categories'. In particular, the opposite of a right category will be a left
-category. There is also a symmetric notion of a category.
-
-\begin{defn}
-A category-framework $\cftalg{A}$ in context $\Gamma$ in an E-system $\mathbb{E}$ consists of
-\begin{enumerate}
-\item A family $\cftalgc{\cftalg{A}}\in\cat{F}/\Gamma$, representing the type of
-objects of the category framework;
-\item A family $\cftalgf{\cftalg{A}}\in\cat{F}/\ctxext{\Gamma}{\cftalgc{\cftalg{A}}}$,
-representing the type of morphisms of the category framework, where a term of
-$\subst{X}{\cftalgf{\cftalg{A}}}$ is considered to be a morphism into $X\in T(\cftalgc{\cftalg{A}})$;
-\item A morphism $\jhom{\Gamma}{\ctxext{\cftalgc{\cftalg{A}}}{\cftalgf{\cftalg{A}}}}{\cftalgc{\cftalg{A}}}{\cftctxext}$, representing the domain function;
-\item A morphism $\jhom{{\Gamma}{\cftalgc{\cftalg{A}}}}{{\cftalgf{\cftalg{A}}}{\jcomp{}{\cftctxext}{\cftalgf{\cftalg{A}}}}}{\cftalgf{\cftalg{A}}}{\cftfamext}$, representing composition;
-\item A term $\jterm{{\Gamma}{\cftalgc{\cftalg{A}}}}{\cftalgf{\cftalg{A}}}{\cftempf{\cftalg{A}}}$,
-representing the identity morphisms.
-\end{enumerate}
-\end{defn}
-
-\begin{rmk}
-Instead of looking at $\cftfamext$ as a morphism from 
-$\ctxext{\cftalgf{\cftalg{A}}}{\jcomp{}{\cftctxext}{\cftalgf{\cftalg{A}}}}$
-to $\cftalgf{\cftalg{A}}$ in context $\ctxext{\Gamma}{\cftalgc{\cftalg{A}}}$, 
-one could also look at $\cftfamext$ as a 
-morphism \emph{over $\cprojfstf{\cftalgc{\cftalg{A}}}{\cftalgf{\cftalg{A}}}$} 
-in context $\Gamma$, as indicated in the following diagram:
-\begin{equation*}
-\begin{tikzcd}[column sep=large]
-\cftalgf{\cftalg{A}}
-  \ar[fib]{d}
-& \jcomp{}{\cftctxext}{\cftalgf{\cftalg{A}}}
-  \ar[fib]{d}
-  \ar{l}[swap]{\cftfamext}
-  \ar{r}
-& \cftalgf{\cftalg{A}}
-  \ar[fib]{d}
-  \\
-\cftalgc{\cftalg{A}}
-& \ctxext{A}{\cftalgf{\cftalg{A}}}
-  \ar{l}{\cprojfstf{A}{\cftalgf{\cftalg{A}}}}
-  \ar{r}[swap]{\cftctxext}
-& \cftalgc{\cftalg{A}}
-\end{tikzcd}
-\end{equation*}
-From this point of view, we see that $\cftfamext$ has an object $X$ and a 
-morphism $f$ into $X$ as parameters, 
-then takes a morphism into the domain of $f$ as an argument, 
-and outputs a morphism into $X$. The projection maps $(X,f)$ to $X$, so we see
-that composition is indeed an operation of this kind.
-\end{rmk}
-
-A category-framework is not a category unless several axioms of the theory of
-categories are satisfied. The domain of a composed arrow $g\circ f$ must be the
-domain of $f$, composition must be associative, the domains of the identity
-morphisms must be the same as their codomains, and identities must satisfy
-the left and right identity laws. However, the axioms expressing these 
-conditions are not very insightful when they're written down directly. Thus, we
-take the usual way and define the notion of functor before we describe the
-actual conditions of being a category.
-
-\begin{defn}
-Let $\cftalg{A}$ and $\cftalg{B}$ be category-frameworks in context $\Gamma$,
-in an E-system $\mathbb{E}$.
-An \emph{internal functor $\cfthom{F}$ from $\extalg{A}$ to
-$\extalg{B}$ in context $\Gamma$} is a pair 
-$(\cfthomc{\cfthom{F}},\cfthomf{\cfthom{F}})$ consisting of
-\begin{align*}
-\jalign\jhom
-  {\Gamma}
-  {\cftalgc{\cftalg{A}}}
-  {\cftalgc{\cftalg{B}}}
-  {\cfthomc{\cfthom{F}}}
-  \\
-\jalign\jfhom
-  {\Gamma}
-  {\cftalgc{\cftalg{A}}}
-  {\cftalgc{\cftalg{B}}}
-  {\cfthomc{\cfthom{F}}}
-  {\cftalgf{\cftalg{A}}}
-  {\cftalgf{\cftalg{B}}}
-  {\cfthomf{\cfthom{F}}}
-\end{align*}
-for which the diagrams
-\begin{equation}\label{eq:exthom1}
-\begin{tikzcd}
-\ctxext{\cftalgc{\cftalg{A}}}{\cftalgf{\cftalg{A}}}
-  \ar{r}{\jvcomp{}{\cfthomc{\cfthom{F}}}{\cfthomf{\cfthom{F}}}}
-  \ar{d}[swap]{\cftctxext[\cftalg{A}]}
-& \ctxext{\cftalgc{\cftalg{B}}}{\cftalgf{\cftalg{B}}}
-  \ar{d}{\cftctxext[\cftalg{B}]}
-  \\
-\cftalgc{\cftalg{A}}
-  \ar{r}[swap]{\cfthomc{\cfthom{F}}}
-& \cftalgc{\cftalg{B}}
-\end{tikzcd}
-\end{equation}
-and
-\begin{equation}\label{eq:exthom2}
-\begin{tikzcd}[column sep=huge]
-\ctxext{\cftalgf{\cftalg{A}}}{\jcomp{}{\cftctxext[\cftalg{A}]}{\cftalgf{\cftalg{A}}}}
-  \ar{r}{\jvcomp{}{\cfthomf{\cfthom{F}}}{\jcomp{}{\cftctxext[\cftalg{A}]}{\cfthomf{\cfthom{F}}}}}
-  \ar{d}[swap]{\cftfamext[\cftalg{A}]}
-& \jcomp{}{\cfthomc{\cfthom{F}}}{\ctxext{\cftalgf{\cftalg{B}}}{\jcomp{}{\cftctxext[\cftalg{B}]}{\cftalgf{\cftalg{B}}}}}
-  \ar{d}{\jcomp{}{\cfthomc{\cfthom{F}}}{\cftfamext[\cftalg{B}]}}
-  \\
-\cftalgf{\cftalg{A}}
-  \ar{r}[swap]{\cfthomf{\cfthom{F}}}
-& \jcomp{}{\cfthomc{\cfthom{F}}}{\cftalgf{\cftalg{B}}}
-\end{tikzcd}
-\end{equation}
-commute, and for which $\subst{\cftempf{\cftalg{A}}}{\cfthomf{\cfthom{F}}}
-\jdeq\jcomp{}{\cfthomc{\cfthom{F}}}{\cftempf{\cftalg{B}}}$.
-\end{defn}
-
-\begin{defn}
-Let $\cftalg{A}$ be a category-framework. We define the category-framework
-\begin{equation*}
-\cftfamalg{\cftalg{A}}\defeq(\cftalgf{\cftalg{A}},\jcomp{}{\cftctxext}%
-{\cftalgf{\cftalg{A}}},\cftfamext,\jcomp{}{\cftctxext}{\cftfamext},
-\jcomp{}{\cftctxext}{\cftempf{\cftalg{A}}}),
-\end{equation*}
-When $\cftalg{P}$ is a category-framework in context $\ctxext{\Gamma}{A}$, then
-and we define the category-framework
-\begin{equation*}
-\ctxext{\cftalgc{\cftalg{A}}}{\cftalg{P}}
-  \defeq
-( \ctxext{A}{\cftalgc{\cftalg{P}}},
-  \cftalgf{\cftalg{P}},
-  \jvcomp{}{\idtm{A}}{\cftctxext[\cftalg{P}]},
-  \cftfamext[\cftalg{P}],
-  \cftempf{\cftalg{P}}).
-\end{equation*}
-\end{defn}
-
-\begin{defn}
-An internal category $\cftalg{A}$ in an E-system $\mathbb{E}$ consists of a
-category-framework $\cftalg{A}$ satisfying the following conditions:
-\begin{enumerate}
-\item \emph{The domain condition.} The pair 
-$(\cftctxext,\idtm{\jcomp{}{\cftctxext}{\cftalgf{\cftalg{A}}}})$ is a
-functor from $\ctxext{\cftalgc{\cftalg{A}}}{\cftfamalg{\cftalg{A}}}$ to
-$\cftalg{A}$.
-\item \emph{The associativity law.} The pair
-$(\cftfamext,\idtm{\jcomp{}{\cftfamext}{\cftalgf{\cftfamalg{\cftalg{A}}}}})$
-is a functor from $\ctxext{\cftalgf{\cftalg{A}}}{\cftfamalg{\cftfamalg{\cftalg{A}}}}$ to
-$\cftfamalg{\cftalg{A}}$.
-\item \emph{The identity domain condition.} The domain of the identity morphism
-is the same as its codomain:
-\begin{equation*}
-\jtermeq
-  {{\Gamma}{\cftalgc{\cftalg{A}}}}
-  {\ctxwk{\cftalgc{\cftalg{A}}}{\cftalgc{\cftalg{A}}}}
-  {\subst{\cftempf{\cftalg{A}}}{\cftctxext[\cftalg{A}]}}
-  {\idtm{\cftalgc{\cftalg{A}}}}.
-\end{equation*}
-\item \emph{The left identity law.} A morphism precomposed by the identity morphism
-is equal to the original morphism:
-\begin{equation*}
-\jtermeq
-  {{{\Gamma}{\cftalgc{\cftalg{A}}}}{\cftalgf{\cftalg{A}}}}
-  {\ctxwk{\cftalgf{\cftalg{A}}}{\cftalgf{\cftalg{A}}}}
-  {\subst{\cftempf{\cftalg{A}}}{\cftfamext}}
-  {\idtm{\cftalgf{\cftalg{A}}}}.
-\end{equation*}
-\item \emph{The right identity law.} A morphism postcomposed by the identity morphism
-is equal to the original morphism:
-\begin{equation*}
-\jtermeq
-  {{{\Gamma}{\cftalgc{\cftalg{A}}}}{\cftalgf{\cftalg{A}}}}
-  {\ctxwk{\cftalgf{\cftalg{A}}}{\cftalgf{\cftalg{A}}}}
-  {\subst{\jcomp{}{\cftctxext}{\cftempf{\cftalg{A}}}}{\cftfamext}}
-  {\idtm{\cftalgf{\cftalg{A}}}}.
-\end{equation*}
-\end{enumerate}
-\end{defn}
-
-\begin{rmk}
-Since we have only used operations of E-systems to define internal categories,
-it immediately follows that if $\cftalg{A}$ is an internal category in
-$\mathbb{E}$ and if $F:\mathbb{E}\to\mathbb{D}$ is an E-homomorphism, then
-$F(\cftalg{A})$ is an internal category in $\mathbb{D}$. In particular, if
-$\cftalg{Q}$ is an internal category in context $\ctxext{\Gamma}{B}$ and
-$f:A\to B$, then the change of base $f_\ast(\cftalg{Q})$ is an internal category
-in context $\ctxext{\Gamma}{A}$. Thus, it is immediate that the notion of
-internal category is stable under change of base.
-\end{rmk}
-\end{comment}
-
-\begin{comment}
-\section{Dependent morphisms}
-
-\begin{defn}
-Let $f_0\in\thom{A}{B}$ be an internal morphism and let $P\in\cat{F}/\ctxext{\Gamma}{A}$ and
-$Q\in\cat{F}/\ctxext{\Gamma}{B}$. We define
-\begin{equation*}
-\thomd{f_0}{P}{Q} \defeq \thom{P}{{f_0}_\ast(Q)}
-\end{equation*}
-We indicate that $f_1\in\thomd{f_0}{P}{Q}$ by drawing the diagram
-\begin{equation*}
-\begin{tikzcd}
-P \arrow[r,"f_1"] \arrow[d,->>] & Q \arrow[d,->>] \\
-A \arrow[r,swap,"f_0"] & B.
-\end{tikzcd}
-\end{equation*}
-\end{defn}
-
-\begin{defn}
-Consider the diagram
-\begin{equation*}
-\begin{tikzcd}
-P \arrow[r,"f_1"] \arrow[d,->>] & Q \arrow[d,->>] \arrow[r,"g_1"] & R \arrow[d,->>] \\
-A \arrow[r,swap,"f_0"] & B \arrow[r,"g_0"] & C.
-\end{tikzcd}
-\end{equation*}
-We define
-\begin{align*}
-\jvcomp{P}{f_0}{f_1} & \defeq\unfold\jvcomp{P}{f_0}{f_1} & & \in \thom{\ctxext{A}{P}}{\ctxext{B}{Q}} \\
-\jfcomp{A}{f_0}{P}{f_1}{g_1} & \defeq \unfold\jfcomp{A}{f_0}{P}{f_1}{g_1} & & \in \thomd{\jcomp{A}{f_0}{g_0}}{P}{R}.
-\end{align*}
-\end{defn}
-
-\begin{rmk}
-Consider
-\begin{equation*}
-\begin{tikzcd}
-P \arrow[r,"f_1"] \arrow[d,->>] & Q \arrow[d,->>] \\
-A \arrow[r,swap,"f_0"] & B.
-\end{tikzcd}
-\end{equation*}
-Then the diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=large]
-\ctxext{A}{P} \arrow[r,"\jvcomp{P}{f_0}{f_1}"] \arrow[d,swap,"\cprojfstf{A}{P}"] & \ctxext{B}{Q} \arrow[d,"\cprojfstf{B}{Q}"] \\
-A \arrow[r,swap,"f_0"] & B
-\end{tikzcd}
-\end{equation*}
-commutes. Also, we have
-\begin{equation*}
-S_{\tmext{x}{u}}(\jvcomp{P}{f_0}{f_1}) \jdeq \tmext{S_x(f_0)}{S_{\tmext{x}{u}}(f_1)}
-\end{equation*}
-for each $x\in T(A)$ and $u\in T(S_x(P))$.
-\end{rmk}
-
-\section{Natural transformations of E-homomorphisms}
-\begin{rmk}
-When $\mathcal{E}$ is an E-system with empty context, then we have the category
-structure on the contexts given by $\thom{\Gamma}{\Delta}= T(W_\Gamma(\Delta))$.
-\end{rmk}
-
-\begin{defn}
-A \emph{natural transformation} of E-homomorphisms from $F:\mathcal{E}\to\mathcal{D}$
-to $G:\mathcal{E}\to\mathcal{D}$ preserving the empty context consists of
-\begin{enumerate}
-\item A t-morphism $\tau0_A\in\thom{F(\Gamma)}{G(\Gamma)}$ for every $\Gamma\in\mathcal{E}$.
-\item A commuting square square
-\begin{equation*}
-\begin{tikzcd}
-F(A) \arrow[r,"\tau1_A"] \arrow[d,twoheadrightarrow] & G(A) \arrow[d,twoheadrightarrow] \\
-F(\Gamma) \arrow[r,swap,"\tau0_\Gamma"] & G(\Gamma)
-\end{tikzcd}
-\end{equation*}
-for every family $P$ over $A$ of solid arrows, such that the diagram
-\begin{equation*}
-\begin{tikzcd}
-F(A) \arrow[r,"\tau1_A"] & G(A) \\
-F(\Gamma) \arrow[r,swap,"\tau0_\Gamma"] \arrow[densely dotted,u,"F(x)"] & G(\Gamma) \arrow[densely dotted,u,swap,"G(x)"]
-\end{tikzcd}
-\end{equation*}
-commutes for every $x\in T(A)$.
-\end{enumerate}
-such that
-\begin{enumerate}
-\item the square
-\begin{equation*}
-\begin{tikzcd}[column sep=huge]
-W_{F(A)}(F(B)) \arrow[r,"W_{F(A)}(\tau1_B)"] \arrow[d,->>] & W_{G(A)}(G(B)) \arrow[d,->>] \\
-F(A) \arrow[r,"\tau1_A"] & G(A)
-\end{tikzcd}
-\end{equation*}
-commutes for every $A,B\in\mathcal{E}/\Gamma$. Note that 
-\begin{align*}
-W_{F(A)}(\tau1_B)
-  & \in 
-T(W_{F(A)}(W_{F(B)}(G(B))))
-  \\
-  & =
-T(W_{W_{F(A)}(F(B))}(W_{F(A)}(G(B))))
-\end{align*}
-\end{enumerate}
-\end{defn}
-
-\section{Indexed E-systems}
-\begin{defn}
-Let $\cat{A}$ be a category. An $\cat{A}$-indexed E-system is an internal
-E-system in the category $Psh(\cat{A})$.
-\end{defn}
-
-\section{Examples of adjunctions}
-
-\begin{lem}
-For each $A\in\mathcal{E}/\Gamma$, $W_A$ is left adjoint to $\Pi_A$. 
-\end{lem}
-
-\begin{proof}
-The unit $\eta:1_{\cat{F}/\ctxext{\Gamma}{A}}\Rightarrow W_A\circ\Pi_A$ is given by
-\begin{equation*}
-\eta_P:\lam{f}{x} ev(\lambda(f),x)
-\end{equation*}
-$\eta_P\in\thom{P}{W_A(\Pi_A(P))}$
-\end{proof}
-
-\begin{lem}
-Suppose $\cat{D}$ is a retract of a pre-category $\cat{F}$, i.e. there are functors
-$S:\cat{D}\to\cat{F}$ and $R:\cat{F}\to\cat{D}$ with a natural isomorphism
-$\tau:R\circ S \simeq \catid{\cat{D}}$, and suppose that $\cat{F}$ is a weakening
-category. Then $\cat{D}$ is a weakening system too.
-\end{lem}
-
-\begin{proof}
-Let $f:\Delta\to \Gamma$ be a morphism in $\cat{D}$. Then we define $W_f$ to be the
-composition
-\begin{equation*}
-\begin{tikzcd}[column sep=3.2em]
-\cat{D}/\Gamma
-  \arrow[r,"S/\Gamma"]
-  &
-\cat{F}/S(\Gamma)
-  \arrow[r,"W_{S(f)}"]
-  &
-\cat{F}/S(\Delta)
-  \arrow[r,"R/S(\Delta)"]
-  &
-\cat{D}/R(S(\Delta))
-  \arrow[r,"\tau_\Delta\circ{-}"]
-  &
-\cat{D}/\Delta
-\end{tikzcd}
-\end{equation*}
-We first verify whether this assembles a pre-weakening structure on $\cat{D}$.
-\begin{enumerate}
-\item Let $\Delta$ be an object of $\cat{D}$. Then $W_{\catid{\Delta}}$ is the composition
-\begin{equation*}
-\begin{tikzcd}[column sep=3.2em]
-\cat{D}/\Delta
-  \arrow[r,"S/\Delta"]
-  &
-\cat{F}/S(\Delta)
-  \arrow[r,"R/S(\Delta)"]
-  &
-\cat{D}/R(S(\Delta))
-  \arrow[r,"{-}\circ\tau_\Delta"]
-  &
-\cat{D}/\Delta
-\end{tikzcd}
-\end{equation*}
-This functor maps $g:\Delta'\to \Delta$ consecutively to $S(g)$, to $R(S(g))$, to
-$\tau_\Delta\circ R(S(g))$. Note that $\tau_\Delta\circ R(S(g))=g\circ \tau_{\Delta'}$ by
-naturality.
-\end{enumerate}
-\end{proof}
-
-\section{Weak E-systems}
-
-\begin{rmk}
-Note that $W$ is a strict $2$-functor frm $\cat{F}$ to $\mathbb{Cat}$ of which
-the value at an object $\Gamma$ is specified to be $\cat{F}/\Gamma$, and which
-is required to preserve the final object. We could have considered $W$ to be a 
-pseudo-functor from $\cat{F}$ to $\mathbb{Cat}$ instead. A coherence theorem
-by Power \cite{Power89} then allows for the strictification of this pseudo-functor
-into a (strict) weakening structure. (Does this map $\Gamma$ to $\cat{F}/\Gamma$
-and final to final?). Thus, we know what a weak pre-weakening system has to be.
-The following questions need to be answered:
-\begin{itemize}
-\item What is a weak weakening/projection/substitution system.
-\item What is a weak E-system? I.e.~what are the coherence rules for something
-to be a weakening/projection/substitution homomorphism?
-\item Can they all be strictified? Are there more examples of weak E-systems?
-\end{itemize}
-\end{rmk}
-\end{comment}
-
 \section{E-systems compared to other systems of type dependency}\label{sec:esys_compared}
@@
 
-\subsection{From categories with families and conceptual categories to E-systems}
+\subsection{From categories with families and comprehension categories to E-systems}
 
@@
 
-\begin{defn}
-A \emph{D-system} consists of
+The definition of D-system we give below, is practically the same as the definition
+of category with attributes in \cite[Definition 3.10]{Hofmann_syntax_semantics}.
+Following Voevodsky, we wish to reserve the word `category' for usage in mathematical terms
+for notions which are `categorical' in the sense of being invariant under 
+equivalence of categories. The definition of categories with attributes involves
+an equality of functors (which would have to be replaced by a natural 
+isomorphism), so it is at least not obviously categorical.
+
+\begin{defn}
+A \emph{D-system} (a category with attributes, see \cite{Hofmann_syntax_semantics}) consists of
 \begin{enumerate}
@@
 the morphism $p(\Gamma,A)$ will be denoted by $p_A$; for any
-$(f,A):(\Delta,f_\ast(A))\to(\Gamma,A)$ in $\int_{\cat{C}}\mathrm{Ty}$, the morphism
-$p(f,A)$ will be denoted by $\pi_2(f,A)$.
+$(f,A):(\Delta,f_\ast(A))\to(\Gamma,A)$ in $\int_{\cat{C}}\mathrm{Ty}$, the top
+morphism of the pullback square $p(f,A)$ will be denoted by $\pi_2(f,A)$.
 \end{enumerate}
@@
 A \emph{D-homomorphism from $\mathbb{D}\to\mathbb{D}'$} consists of a functor 
+$\mathcal{F}:\cat{C}\to\cat{C}'$ which maps the terminal object of
+$\cat{C}$ to the terminal object of $\cat{C}'$, and a natural transformation
+$\tau:\mathrm{Ty}\Rightarrow\mathrm{Ty}'\circ\mathcal{F}$, such that 
+$\mathcal{F}(p_A)\jdeq p_{\mathcal{F}(A)}$ for any
+$A\in\mathrm{Ty}(\Gamma)$, and $\mathcal{F}(\pi_2(f,A))\jdeq
+\pi_2(\mathcal{F}(f),\tau_\Gamma(A))$ for any $f:\Delta\to\Gamma$ in $\cat{C}$ and
+any $A\in\mathrm{Ty}(\Gamma)$. 
+\end{defn}
+
+The following lemma restates the last condition under the assumption that
+$\mathcal{F}$ preserves \emph{all} finite limits. Note that
+in the definition of D-homomorphism, the underlying functor is only
+required to preserve the pullbacks of the projections, of which there may be
+none.
+
+\begin{lem}
+Suppose $\mathbb{D}$ and $\mathbb{D}'$ are D-systems with underlying categories
+$\cat{C}$ and $\cat{C}'$, respectively. Then a pair consisting of a finite
+limit preserving functor
 $\mathcal{F}:\cat{C}\to\cat{C}'$ and a natural transformation
-$\mathrm{Ty}\Rightarrow\mathrm{Ty}'\circ\mathcal{F}$, such that 
-$\mathcal{F}(p_A)\jdeq p_{\mathcal{F}(A)}$ for any
-$A\in\mathrm{Ty}(\Gamma)$.
-\end{defn}
-
-\begin{lem}
-In the definition of D-system, the assumed structure beyond a choice of
-projections $p_A:\ctxext{\Gamma}{A}\twoheadrightarrow\Gamma$ is property-like.
+$\tau:\mathrm{Ty}\to \mathrm{Ty'}\circ\mathcal{F}$ is a D-homomorphism if and
+only if $\mathcal{F}$ preserves the terminal object on the nose and the induced
+square
+\begin{equation*}
+\begin{tikzcd}
+\int_{\cat{C}}\mathrm{Ty} \arrow[r,"p"] \arrow[d,swap,"{\int_{\mathcal{F}}\tau}"] & \cat{C}^{\square} \arrow[d,"\mathcal{F}^\square"] \\
+\int_{\cat{C}'}\mathrm{Ty}' \arrow[r,"{p'}"] & {\cat{C}'}^\square
+\end{tikzcd}
+\end{equation*}
+commutes.
 \end{lem}
 
+\begin{proof}
+Straightforward.
+\end{proof}
+
+%\begin{lem}
+%In the definition of D-system, the assumed structure beyond a choice of
+%projections $p_A:\ctxext{\Gamma}{A}\twoheadrightarrow\Gamma$ is property-like.
+%\end{lem}
+
+The following theorem is due to Hofmann, \cite[section 3.2]{Hofmann_syntax_semantics}. 
+%In the proof that every category with families is a D-system, we
+%outline the successive steps. The omitted details are marked as omitted, and
+%they are straightforward to fill in.
+
 \begin{thm}\label{thm:DD}
-A category with families is the same thing as a D-system.
+The category of categories with families is equivalent to the category of 
+D-systems.\hfill$\square$
 \end{thm}
 
-\begin{proof}
-Let $\mathbf{CWF}$ be a category with families. We show that $\mathbf{CWF}$ is
-a D-system in five steps.
+\begin{comment}
+Since this theorem basically asserts that a category with families is a
+comprehension category
+
+\begin{proof}[Outline of the proof]
+Let $\mathbf{CWF}$ be a category with families. The fact that $\mathbf{CWF}$ is
+a D-system can be shown by proving each of the assertions in the following six 
+steps.
 \begin{enumerate}
-\item We first show that
-\begingroup\it for every $f:\Delta\to\Gamma$ and $A\in\mathrm{Ty}(\Gamma)$, there is an
+\item For every $f:\Delta\to\Gamma$ and $A\in\mathrm{Ty}(\Gamma)$, there is an
 isomorphism
@@
 such that $p_A\circ\theta\jdeq f$.
-
-Furthermore, we get $\zeta_{A,p_A}(\catid{\ctxext{\Gamma}{A}})\jdeq\varphi_{(p_A)_\ast(A)}(\delta_A)$%
-\endgroup
+Furthermore, we get $\zeta_{A,p_A}(\catid{\ctxext{\Gamma}{A}})\jdeq\varphi_{(p_A)_\ast(A)}(\delta_A)$.
+
 \medskip
-
-\renewcommand\qedsymbol{\(\blacksquare\)}
+(Proof omitted).
+\begin{comment}
 By \autoref{lem:cwf_to_dsys} we may define
@@
 equalities.
-
+\end{comment}
+
+\begin{comment}
 \item
 For any $A\in\mathrm{Ty}(\Gamma)$, we define
-$W_A\defeq A_\ast:\cat{F}/\Gamma\to\cat{F}/\ctxext{\Gamma}{A}$. Then we get
+$W_A\defeq (p_A)_\ast:\cat{F}/\Gamma\to\cat{F}/\ctxext{\Gamma}{A}$. Then we get
 for any $A,B\in\cat{F}/\Gamma$ an isomorphism
@@
 
-\item Now we show that 
-\begingroup\it for any $A\in\mathrm{Ty}(\Gamma)$ and $f:\Delta\to\Gamma$ 
+\item For any $A\in\mathrm{Ty}(\Gamma)$ and $f:\Delta\to\Gamma$ 
 there is a morphism $\pi_2(f,A):\ctxext{\Delta}{f_\ast(A)}\to\ctxext{\Gamma}{A}$
 such that $p_A\circ \pi_2(f,A)\jdeq f\circ p_{f_\ast(A)}$.
-\endgroup
+
 \medskip
-
+(Proof omitted).
+\begin{comment}
 Note that we have the isomorphism
@@
 satisfying $p_A\circ\pi_2(f,A)\jdeq f\circ p_{f_\ast(A)}$.
-\item Next, we show that
-\begingroup\it
-for any $f:\Delta\to\Gamma$, $g:X\to\Delta$ and $A\in\mathrm{Ty}(\Gamma)$, one has
+\end{comment}
+\begin{comment}
+\item For any $f:\Delta\to\Gamma$, $g:X\to\Delta$ and $A\in\mathrm{Ty}(\Gamma)$, one has
 \begin{equation*}
@@
 \end{equation*}
-\endgroup
+
 \medskip
-
+(Proof omitted).
+\begin{comment}
 It suffices to show that
@@
 \end{align*}
+\end{comment}\begin{comment}
 \item
-\begingroup\it
 Let $f:\Delta\to\Gamma$ and $A\in\mathrm{Ty}(\Gamma)$. Then one has 
@@
 for any $\theta:\Delta\to\ctxext{\Gamma}{A}$ satisfying $A\circ\theta\jdeq f$.
-\endgroup
+
 \medskip
-
+(Proof omitted).
+\begin{comment}
 We will write $\tilde\theta\defeq\zeta_{A,f}(\theta)$. Note that
@@
 \end{align*}
+\end{comment}\begin{comment}
 \item 
-\begingroup\it
 The diagram
@@
 is a pullback square.
-\endgroup
-\medskip
-
-To verify the universal mapping property of the pullback square, consider a
+
+Proof of (vi): To verify the universal mapping property of the pullback square, consider a
 commuting diagram
@@
 \end{enumerate}
+This completes the proof outline that $\mathbf{CWF}$ is a D-system.
+
 Now let $\mathbb{D}$ be a D-system. We define the presheaf $\mathrm{Tm}:
@@
 \end{defn}
-
-To be able to even construct an E'-system out of a D-system, we need to impose
-a rather restricting conditon on the D-system.
 
@@
 \begin{defn}
-Let $p:\cat{E}\to\cat{B}$ be a functor and let $g:A\to B$ be a morphism in $\cat{B}$.
-A morphism $f:X\to Y$ in $\cat{E}$ is said to be \emph{cartesian over $g$}, if
+Let $p:\cat{E}\to\cat{B}$ be a functor and let $f:\Delta\to \Gamma$ be a morphism in $\cat{B}$.
+A morphism $g:A\to B$ in $\cat{E}$ is said to be \emph{cartesian over $f$}, if
 \begin{enumerate}
-\item $p(f)\jdeq g$
-\item For all $f':X'\to Y$ with $p(f')\jdeq g$, there is a unique $u:X'\to X$
-such that $f'=f\circ u$ and $p(u)\jdeq\catid{A}$. 
+\item $p(g)\jdeq f$
+\item For all $g':A'\to B$ with $p(g')\jdeq f$, there is a unique $u:A'\to A$
+such that $g'=g\circ u$ and $p(u)\jdeq\catid{\Delta}$. 
 \end{enumerate}
-A cartesian morphism over $g$ is also called a \emph{cartesian lift of $g$}.
+A cartesian morphism over $f$ is also called a \emph{cartesian lift of $f$}.
 \end{defn}
@@
 A functor $p:\cat{E}\to\cat{B}$ is called a \emph{Grothendieck fibration} if
-every morphism into an object of the form $p(X)$ has a cartesian lift, 
-and cartesian morphisms are closed under composition.
+every morphism into an object of the form $p(A)$ for some object $A$ of $\cat{E}$ 
+has a cartesian lift, and cartesian morphisms are closed under composition.
 \end{defn}
@@
 A Grothendieck fibration $p:\cat{E}\to\cat{B}$ is said to be \emph{split}, if
-there is a functorial choice of cartesian lifts, i.e.~for every $g:A\to B$ in $\cat{B}$
-there is a cartesian lift $j(g):j(A)\to j(B)$, such that $j(\catid{A})\jdeq\catid{j(A)}$
-and $j(h\circ g)\jdeq j(h)\circ j(g)$.
-\end{defn}
+there is a functorial choice of cartesian lifts, i.e.~for every $f:\Delta\to \Gamma$ in 
+$\cat{B}$ and any $B\in \cat{E}$ such that $p(B)\jdeq \Delta$,
+there is a cartesian lift $j(B,g):A \to B$ for some object $A$ of $\cat{E}$, such that 
+$j(B,p(B))\jdeq\catid{B}$ and $j(B, g\circ h)\jdeq j(B,g)\circ j(\mathrm{dom}(j(B,g)),h)$.
+\end{defn}
+
+\begin{eg}
+For any presheaf $P:\cat{C}\to\mathbb{Set}$, the projection functor
+$\int_\cat{C} P\to\cat{C}$ is a Grothendieck fibration.
+\end{eg}
 
@@
 
+\begin{eg}
+Since any morphism of $\int_{\cat{C}}$ is cartesian, it follows that D-systems
+are comprehension categories. As D-systems are built up from a presheaf of types
+on the underlying category, they are in fact discrete comprehension categories.
+\end{eg}
+
+\begin{defn}
+Let $\mathcal{P}$ be a comprehension category. Then we define the pre-category
+$\cat{F}$ to be the category with the same objects as $\cat{C}$, of which the
+morphisms are generated by the morphisms of the form $P(A)$, for objects
+$A\in\cat{E}$. 
+\end{defn}
+
+\begin{defn}
+Let $\mathcal{P}$ be a comprehension category. Then we define for any 
+$\Gamma\in\cat{C}$, the poset $\Lambda(\Gamma)$ to be the category of finite 
+sequences  $A_1,\ldots,A_n$ of objects of $\cat{E}$, such that 
+$p(A_1)\jdeq\Gamma$ and $p(A_{i+1})\jdeq\mathrm{dom}(\mathcal{P}(A_i))$ 
+for any $1\leq i < n$. The partial order on 
+$\Lambda(\Gamma)$ is defined by $(A_1,\ldots,A_n)\leq (B_1,\ldots,B_m)$ if and 
+only if $n\leq m$ and $A_i\jdeq B_i$ for all $1\leq i\leq n$.
+
+The functor $\mathcal{P}$ induces a functor $\mathcal{P}:\Lambda(\Gamma)\to
+\cat{F}/\Gamma$ by $\mathcal{P}(A_1,\ldots,A_n)\jdeq\mathcal{P}(A_1)\circ
+\cdots\circ\mathcal{P}(A_n)$.
+\end{defn}
+
+\begin{defn}
+Let $f_0\defeq f:\Delta\to\Gamma$ be a morphism of $\cat{C}$, the base category of a
+split comprehension category $\mathcal{P}$ wich a choice $j$ of cartesian lifts.
+By induction we define an element $\Lambda(f,A)\in\Lambda(\Delta)$ together with
+a morphism $\pi_2(f,A):\mathrm{dom}(\mathcal{P}(\Lambda(f,A)))\to 
+\mathrm{dom}(\mathcal{P}(A))$, for any $A\jdeq (A_1,\ldots,A_n) 
+\in\Lambda(\Gamma)$.
+
+For $n\jdeq 0$, i.e.~if $A$ is the empty sequence, we let $\Lambda(f,A)$ be the 
+empty sequence, and $\pi_2(f,A) \jdeq f$. For the inductive step, consider a 
+sequence $A_1,\ldots,A_{n+1}\in \Lambda(\Gamma)$. Then we define
+\begin{align*}
+\Lambda(f,A) & \jdeq \mathrm{dom}(j(A_{n+1},\pi_2(f,(A_1,\ldots,A_n))))
+\intertext{and}
+\pi_2(f,A) & \jdeq \mathrm{dom}(\mathcal{P}(j(A_{n+1},\pi_2(f,(A_1,\ldots,A_n)))))
+\end{align*}
+where we consider in the definition of $\pi_2(f,A)$, $\mathrm{dom}$ to be the 
+domain functor from $\cat{C}^{\to}\to\cat{C}$. 
+\end{defn}
+
+\begin{rmk}
+It follows immediately from the definition of comprehension categories and the
+pasting lemma of pullbacks, that the square
+\begin{equation*}
+\begin{tikzcd}[column sep=large]
+\mathrm{dom}(\mathcal{P}(\Lambda(f,A))) \arrow[r,"{\pi_2(f,A)}"] \arrow[d,swap,"{\mathcal{P}(\Lambda(f,A))}"] & \mathrm{dom}(\mathcal{P}(A)) \arrow[d,"{\mathcal{P}(A)}"] \\
+\Delta \arrow[r,"f"] & \Gamma
+\end{tikzcd}
+\end{equation*}
+is a pullback square.
+\end{rmk}
+
+\begin{thm}
+Let $\mathcal{P}$ be a comprehension category. Then the following are equivalent:
+\begin{enumerate}
+\item The functors $\Lambda(f)$ can be extended to the structure of an E'-system
+without empty context.
+\item For any
+$A,A'\in\Lambda(\Gamma)$ such that $\mathcal{P}(A)\jdeq\mathcal{P}(A')$
+and for any $f:\Delta\to\Gamma$ in $\cat{C}$, the equalities
+$\mathcal{P}(\Lambda(f,A))\jdeq\mathcal{P}(\Lambda(f,A'))$ and
+$\pi_2(f,A)\jdeq\pi_2(f,A')$ hold. 
+\end{enumerate}
+\end{thm}
+
+\begin{proof}
+Suppose we can extend $\Lambda(f)$ for all $f:\Delta\to\Gamma$, to form commuting
+diagrams
+\begin{equation*}
+\begin{tikzcd}
+\Lambda(\Gamma) \arrow[d,swap,"\mathcal{P}"] \arrow[r,"\Lambda(f)"] & \Lambda(\Delta) \arrow[d,"\mathcal{P}"] \\
+\cat{F}/\Gamma \arrow[r,densely dotted,"{f_\ast}"] & \cat{F}/\Delta
+\end{tikzcd}
+\end{equation*}
+of functors, forming an E'-system without empty context, where $\mathcal{P}(A)
+\defeq \mathcal{P}(A_1)\circ\cdots\circ\mathcal{P}(A_n)$ for any
+$A\jdeq (A_1,\ldots,A_n)\in\Lambda(\Gamma)$. Then it must hold that for any
+$A,A'\in\Lambda(\Gamma)$ such that $\mathcal{P}(A)\jdeq\mathcal{P}(A')$, we have
+$\mathcal{P}(\Lambda(f,A))\jdeq\mathcal{P}(\Lambda(f,A'))$. Also, for such
+$A$ and $A'$ it follows from the way $f_\ast$ is defined, 
+that $\pi_2(f,A)_\ast\jdeq f_\ast/\mathcal{P}(A)\jdeq
+f_\ast/\mathcal{P}(A')\jdeq\pi_2(f,A')_\ast$, so the desired equality
+$\pi_2(f,A)\jdeq\pi_2(f,A')$ holds.
+
+For the converse, suppose that for all $f:\Delta\to\Gamma$ and $A,A'\in
+\Lambda(\Gamma)$ such that $p_A\jdeq p_{A'}$, one has 
+$p_{\Lambda(f,A)}\jdeq p_{\Lambda(f,A')}$ and $\pi_2(f,A)\jdeq\pi_2(f,A')$.
+Let $f:\Delta\to\Gamma$. By the assumption that $p_{\Lambda(f,A)}\jdeq p_{\Lambda(f,A')}$,
+it follows that the objects part of $p^\Delta\circ\Lambda(f):\Lambda(\Gamma)\to\cat{F}/\Delta$ 
+factors through $p^\Gamma:\Lambda(\Gamma)\to\cat{F}/\Gamma$. Moreover, by the
+assumption that $\pi_2(f,A)\jdeq \pi_2(f,A')$ it follows that the pullback
+squares
+\begin{equation*}
+\begin{tikzcd}[column sep=large]
+\ctxext{\Delta}{f_\ast(A)} \arrow[d,fib] \arrow[r,"{\pi_2(f,A)}"] & \ctxext{\Gamma}{A} \arrow[d,fib] \\
+\Delta \arrow[r,"f"] & \Gamma
+\end{tikzcd}
+\qquad\text{and}\qquad
+\begin{tikzcd}[column sep=large]
+\ctxext{\Delta}{f_\ast(A')} \arrow[d,fib] \arrow[r,"{\pi_2(f,A')}"] & \ctxext{\Gamma}{A} \arrow[d,fib] \\
+\Delta \arrow[r,"f"] & \Gamma
+\end{tikzcd}
+\end{equation*}
+are the same. Since we have this property
+for any object $\Gamma$, it also follows that the morphisms part of
+$p^\Delta\circ\Lambda(f)$ factors through $p^\Gamma$. This defines the functor
+$f_\ast$ in the commuting square
+\begin{equation*}
+\begin{tikzcd}
+\Lambda(\Gamma) \arrow[d,swap,"p"] \arrow[r,"\Lambda(f)"] & \Lambda(\Delta) \arrow[d,"p"] \\
+\cat{F}/\Gamma \arrow[r,densely dotted,"{f_\ast}"] & \cat{F}/\Delta
+\end{tikzcd}
+\end{equation*}
+We need to verify that $f_\ast$ is a functorial choice of pullbacks, as in the
+definition of \autoref{defn:E'sys}. Since $f_\ast$ is defined as an extension
+of $\Lambda(f)$, it is enough to verify the analoguous functoriality properties
+for $\Lambda(f)$. We do this by induction on the objects of $\Lambda(\Gamma)$,
+in the same order as in \autoref{defn:E'sys}.
+\begin{enumerate}[label=(\alph*)]
+\item This property holds by definition.
+\item Let $\Gamma\in\cat{C}$. We will prove by induction on $A\in\Lambda(\Gamma)$
+that $(\catid{\Gamma})_\ast(A)\jdeq A$ and that $\pi_2(\catid{\Gamma},A)\jdeq \catid{\ctxext{\Gamma}{A}}$.
+Note that these properties hold by definition for $A\jdeq\catid{\Gamma}$. For the
+inductive step, assume that $A\in\Lambda(\Gamma)$ is such that
+$(\catid{\Gamma})_\ast(A)\jdeq A$ and $\pi_2(\catid{\Gamma},A)\jdeq
+\catid{\ctxext{\Gamma}{A}}$ and let $B\in\mathrm{Ty}(\ctxext{\Gamma}{A})$. Then
+we have
+\begin{align*}
+(\catid{\Gamma})_\ast(A\circ p_B) & \jdeq (\catid{\Gamma})_\ast(A)\circ \pi_2(\catid{\Gamma},A)_\ast(p_B) \\
+& \jdeq A\circ p_B
+\end{align*}
+Also, we have
+\begin{align*}
+\pi_2(\catid{\Gamma},A\circ p_B) & \jdeq \pi_2(\pi_2(\catid{\Gamma},A),p_B) \\
+& \jdeq \pi_2(\catid{\ctxext{\Gamma}{A}},p_B) \\
+& \jdeq \catid{\ctxext{{\Gamma}{A}}{B}}
+\end{align*}
+\item Let $f:\Delta\to \Gamma$ and $g:X\to \Delta$. We will prove by induction
+on $A\in\Lambda(\Gamma)$ that $(f\circ g)_\ast(A)\jdeq g_\ast(f_\ast(A))$ and
+that $\pi_2(f\circ g,A)\jdeq \pi_2(f,A)\circ \pi_2(g,f_\ast(A))$. Note that
+these properties hold by definition for $A\jdeq\catid{\Gamma}$. For the inductive
+step, assume that $A\in\Lambda(\Gamma)$ is such that $(f\circ g)_\ast(A)\jdeq g_\ast(f_\ast(A))$ and
+$\pi_2(f\circ g,A)\jdeq \pi_2(f,A)\circ \pi_2(g,f_\ast(A))$, and let
+$B\in\mathrm{Ty}(\ctxext{\Gamma}{A})$. Then we have
+\begin{align*}
+(f\circ g)_\ast(A\circ p_B) & \jdeq \pi_2(f\circ g,A)_\ast(p_B) \\
+& \jdeq (\pi_2(f,A)\circ\pi_2(g,f_\ast(A)))_\ast(p_B) \\
+& \jdeq \pi_2(g,f_\ast)_\ast(\pi_2(f,A)_\ast(p_B)) \\
+& \jdeq g_\ast(f_\ast(A\circ p_B))
+\end{align*}
+\item This property holds by definition.\qedhere
+\end{enumerate}
+\end{proof}
+
 \begin{defn}
@@
 \end{proof}
-
-\begin{comment}
-\subsection{Quotients of E-systems}
-
-\begin{defn}
-A regular congruence relation $\sim$ on a category $\cat{F}$ with term structure,
-is a triple $(\sim_c,\sim_f,\sim_t)$ consisting of
-\begin{enumerate}
-\item an equivalence relation $\sim_c$ on the set of objects of $\cat{F}$,
-\item an equivalence relation $\sim_f$ on the set of morphisms of $\cat{F}$,
-\item an equivalence relation $\sim_t$ on the set of terms of $\cat{F}$,
-\end{enumerate}
-which are compatible with $\mathrm{cod}$, $\mathrm{dom}$, $\catid{}$, $\circ$
-and $\ebd$
-\end{defn}
-\end{comment}

### 2015-08-16 e-systems.tex
didn't keep track of the work I did

@@
 \end{comment}
+
+\begin{comment}
+\section{An example with finite sets}
+
+\begin{defn}
+Let $\cat{C}$ be pre-category with chosen finite limits in
+which the terminal object is also the initial object (e.g.~the category of
+non-empty sets or groups), 
+and write $\cat{C}'$ for the category with the
+same objects as $\cat{C}$, and with the monomorphisms of $\cat{C}$ as morphisms. 
+We define the E-system $\mathbf{E}_{\cat{C}}$ to have underlying category
+$\cat{C}'$, where a term of a monomorphism $m$ is a section of $m$.
+
+\end{defn}
+
+\begin{defn}
+We define the E-system $\mathbf{E_{fin}}$ of finite sets. A context of
+$\mathbf{E_{fin}}$ is a finite set. A family $A$ over $\Gamma$
+\end{defn}
+\end{comment}

### 2015-08-16 internal_homs.tex
didn't keep track of the work I did

@@
-\section{The category of internal morphisms of an E-system}\label{sec:esys_props}
+\section{The pre-category of internal morphisms of an E-system}\label{sec:esys_props}
 
@@
 projections and their sections. We elaborate on this point of view by defining
-the category of internal morphisms, which Lawvere might have called the
+the pre-category of internal morphisms, which Lawvere might have called the
 category of terms. 
-In the category of internal morphisms over a context $\Gamma$,
+In the pre-category of internal morphisms over a context $\Gamma$,
 which we define in \autoref{subsec:im_cat},
@@
 
-The category of internal morphisms forms a category of which the families form
-a subcategory. This, together with the pre-composition operation described
+The pre-category of internal morphisms forms a pre-category of which the families form
+a subpre-category. This, together with the pre-composition operation described
 earlier, gives rise to the auxiliary notion of E'-systems, which we define
@@
 \begin{defn}\label{defn:E'sys}
-An \emph{E'-system} consists of a category $\cat{C}$ and a subcategory $\cat{F}$ with
+An \emph{E'-system} consists of a pre-category $\cat{C}$ and a subpre-category $\cat{F}$ with
 the same objects as $\cat{C}$, such that
@@
 \begin{rmk}
-Without a terminal object in $\cat{F}$, the category $\cat{C}$ does not correspond
+Without a terminal object in $\cat{F}$, the pre-category $\cat{C}$ does not correspond
 directly to anything in an E-system.
@@
 For any $\Gamma\in\cat{C}$, define $\cat{C}_\cat{F}/\Gamma$ to be the full
-subcategory of $\cat{C}/\Gamma$ with the same objects as $\cat{F}/\Gamma$. 
+subpre-category of $\cat{C}/\Gamma$ with the same objects as $\cat{F}/\Gamma$. 
 \end{defn}
@@
 
-\subsection{The category of internal morphisms}\label{subsec:im_cat}
+\subsection{The pre-category of internal morphisms}\label{subsec:im_cat}
 
@@
 because $W_{\ctxext{A}{P}}\jdeq W_P\circ W_A$.
-Once we have established a category of which the morphisms are given by 
+Once we have established a pre-category of which the morphisms are given by 
 $\thom{\blank}{\blank}$, we therefore get that 
@@
 \begin{thm}
-For each object $\Gamma$ of an E-system $\mathbb{E}$, we have a category
+For each object $\Gamma$ of an E-system $\mathbb{E}$, we have a pre-category
 $\mathbb{C}_I(\mathbb{E},\Gamma)$ of internal morphisms of $\mathbb{E}$,

### 2015-08-16 main.tex
didn't keep track of the work I did

@@
 \title{An essentially algebraic formulation of dependent type theory}
-\author{Egbert Rijke\\ {\sc Carnegie Mellon University}}
+\author{Egbert Rijke\footnote{{\sc Carnegie Mellon University}, department of
+philosophy; \url{erijke@andrew.cmu.edu}}}
 \date{\today}
@@
 
-\section{Introduction}
+\section{Introduction --- Mathematical languages as designed and maintained languages}
+Mathematical research is aided by the development of languages that help 
+researchers to investigate their subject efficiently. Such a development is an
+ongoing process, both for informal and formal languages, that may have both
+mathematical and meta-mathematical aspects.
 
-\subsection*{Principles of mathematical fuseki}
-Mathematics is the study of abstract concepts such as space or quantity, and
-when we design a language for such a study, we want to be able to efficiently
-express things about the objects under investigation. The subject matter of
-this paper is a logical framework of dependent type theory.
+There is an amount of attraction and repulsion that works simultaneously
+between formal and informal languages for mathematics, which is a driving force
+for the development of mathematical languages. The use of
+a natural language in mathematical arguments facilitates people to engage
+with mathematical constructs at a conceptual level, while a formal mathematical
+language aims to encode the intuitive development. At the same time and 
+especially nowadays, formal languages aim to possess a high level of rigor and
+precision, so that mathematical arguments and constructions can be executed and
+verified algorithmically, and informal mathematics is usually not considered
+acceptable if it cannot in principle be carried out in a formal system.
 
+It may be the case that formal languages are developed in order to be more
+precise and certain about bodies of existing mathematical work, but there are
+also cases where the conceptual thinking
+
+The book on Homotopy Type Theory, \cite{TheBook}, was written in informal
+language (maybe `semi-formal' would be a more accurate description), 
+partly because it was conceived that the underlying type theory might
+not always remain the same, while principles and concepts developed starting
+from type theory exist separately from any particular version of type 
+theory, though perhaps not completely independent.
+
+Today, formal languages for mathematics, logic and computer science, are studied
+under the name logical frameworks.
 Pfenning \cite{PfenningLFintro} defines logical frameworks
@@
 theory, and so on. By considering models of a logical framework, it becomes
-itself the subject of mathematical study.
+itself the subject of mathematical study. Thus, logical frameworks typically
+contain laws for handling logical inferences, the specification of the sorts
+and operations of the language or theory, and axioms the operations are required to
+satisfy.
 
-It may actually be useful to see logical frameworks themselves
-as objects subject to mathematical investigation.
+While there are aspects of logical frameworks that are of meta-mathematical 
+nature, such as descriptions of handling inference rules, it may be useful to
+consider a mathematical language as a mathematical object, especially if one
+is interested in models of the theory. For practically any theory there is a
+wish to do so, so it may be useful to keep this in mind in the specification
+of a logical framework. One way to do this, is to make sure that the theory
+belongs to a class of theories of which the model theory is reasonably well-%
+understood. Among such classes of theories are algebraic theories and
+essentially algebraic theories.
 
-Logical frameworks for mathematics are not set in stone.
-
- and neither to they
-remain the same once they are declared.
-
-When designing and maintaining a logical framework
+The choices made in the design principles and specification of a logical 
+framework, will determine both the subject of study and how effective we can be
+in its study. Investigations of logical framework can therefore be seen as
+`opening theories for mathematics', in analogy of opening theories for games.
+Logical frameworks for mathematics are not set in stone and 
+neither to they remain the same once they are declared.
 
@@
 algebraic theories is that their models are closed under 
-colimits \cite{Aspects_of_topoi}.
+colimits \cite{Aspects_of_topoi}. Many classes of theories have been described
+and shown to be equivalent to essentially algebraic theories. Among them are
+Partial Horn Logic and quasi-equational theories \cite{Palmgren_Vickers},
+cartesian theories \cite{Freyd2002}, finite limit sketches 
+\cite{Toposes_triples_theories}, left exact theories
+\cite{McLarty1986}, and they can be equivalently represented as locally
+presentable categories \cite{Adamek_Locally}.
 
@@
 
-\subsection*{Logical frameworks}
+\subsection*{Setting up the essentially algebraic theory of E-systems}
+The theory of E-systems we present in the present article is an iteration in the
+development of theory of type
+dependency. The goal has been to describe an essentially algebraic theory of
+type dependency that is in close correspondence to the type theory of
+\cite{TheBook}, while at the same time it is as economic as possible in the
+demand of external resources (such as the natural numbers). Instead of explaning
+what a dependent type is via an infinite hierarchy of universes $\UU_0$,
+$\UU_1$, \ldots, we set out to describe type families using only the three sorts
+that typically occur in type theoretical expressions: contexts, families and 
+terms. Although an infinite hierarchy of universes allows one to consider everything
+to be a term (because every type is a term of some specified universe), and 
+therefore use the literally same mechanisms for handling type expressions as for
+term expressions in the process of compilation and verification of a 
+construction, this specific setup requires the natural numbers to exist in the 
+meta-theory.
 
-\subsection*{Setting up the essentially algebraic theory of E-systems}
-The theory of type dependency as we present it here, consists of three sorts
-$C$, $F$ and $T$ for contexts, families and terms. There is a map 
+An E-system is therefore an algebra consisting of three sorts: $C$ for contexts,
+$F$ for families, and $T$ for terms. There is a map 
 $\eft:F\to C$ which determines for each family, in which context it is a

### 2015-08-16 weakeningcats.tex
didn't keep track of the work I did

--- 2015-06-16 weakeningcats.tex
@@
-\section{E-systems}\label{sec:esys_defn}
-
-\subsection{Weakening systems}
+\begin{comment}
+\subsection{Quotients of E-systems}
 
 \begin{defn}
-A \emph{pre-weakening structure on a pre-category $\cat{C}$} consists of a functor
-$W_A:\cat{C}/\Gamma\to \cat{C}/\ctxext{\Gamma}{A}$ for every $A:\ctxext{\Gamma}{A}\to \Gamma$ in $\cat{C}$ such 
-that
+A regular congruence relation $\sim$ on a category $\cat{F}$ with term structure,
+is a triple $(\sim_c,\sim_f,\sim_t)$ consisting of
 \begin{enumerate}
-\item $W_{\catid{\Gamma}}\jdeq \catid{\cat{C}/\Gamma}$ for every object $\Gamma\in\cat{C}$.
-\item $W_{\ctxext{A}{P}}\jdeq W_P\circ W_A$ for every $P\in\cat{C}/\ctxext{\Gamma}{A}$ and $A\in\cat{C}/\Gamma$.
-\item $W_A$ preserves the final object, i.e.~$W_A(\catid{\Gamma})\jdeq \catid{\ctxext{\Gamma}{A}}$.
+\item an equivalence relation $\sim_c$ on the set of objects of $\cat{F}$,
+\item an equivalence relation $\sim_f$ on the set of morphisms of $\cat{F}$,
+\item an equivalence relation $\sim_t$ on the set of terms of $\cat{F}$,
 \end{enumerate}
-A \emph{pre-weakening system} is a pre-category together with a pre-weakening structure.
-\end{defn}
-
-\begin{defn}
-A \emph{pre-weakening homomorphism} $F:\cat{C}\to\cat{D}$ between pre-weakening systems
-is a functor $F:\cat{C}\to\cat{D}$ such that the square
-\begin{equation*}
-\begin{tikzcd}[column sep=large]
-\cat{C}/\ctxext{\Gamma}{A}
-  \arrow[r,"F/\ctxext{\Gamma}{A}"]
-  &
-\cat{D}/F(\ctxext{\Gamma}{A})
-  \\
-\cat{C}/\Gamma
-  \arrow[u,"W_A"]
-  \arrow[r,swap,"F/\Gamma"]
-  &
-\cat{D}/F(\Gamma)
-  \arrow[u,swap,"W_{F(A)}"]
-\end{tikzcd}
-\end{equation*}
-commutes for any $A\in\cat{C}/\Gamma$.
-\end{defn}
-
-\begin{lem}
-Let $\cat{C}$ be a pre-weakening system. Then for any object $\Gamma$ of $\cat{C}$,
-the slice pre-category $\cat{C}/\Gamma$ is also a pre-weakening system, with the weakening
-structure given by $W(\cat{C}/\Gamma)_P\jdeq W(\cat{C})_P$ for each $P\in\cat{C}/\ctxext{\Gamma}{A}$
-and each $A\in\cat{C}/\Gamma$.
-\end{lem}
-
-\begin{defn}
-A \emph{weakening system} is a pre-weakening system for which every $W_f$ is a
-pre-weakening homomorphism. A \emph{weakening homomorphism} is a pre-weakening 
-homomorphism between weakening systems.
-\end{defn}
-
-\begin{rmk}
-In other words, a weakening system is a pre-weakening system with a weakening
-structure $W$ for which the square
-\begin{equation*}
-\begin{tikzcd}[column sep=huge]
-\cat{C}/\ctxext{{\Gamma}{B}}{Q}
-  \arrow[r,"W_A/\ctxext{B}{Q}"]
-  &
-\cat{C}/\ctxext{{\Gamma}{A}}{W_A(\ctxext{B}{Q})}
-  \\
-\cat{C}/\ctxext{\Gamma}{B}
-  \arrow[u,"W_Q"]
-  \arrow[r,swap,"W_A/B"]
-  &
-\cat{C}/\ctxext{{\Gamma}{A}}{W_A(B)}
-  \arrow[u,swap,"W_{W_A(Q)}"]
-\end{tikzcd}
-\end{equation*}
-for each $A,B\in\cat{C}/\Gamma$ and $Q:\cat{C}/\ctxext{\Gamma}{B}$. The objects part of this 
-property asserts that for any $k\in \cat{C}/\greek{E}$, the dotted arrows in the diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=tiny]
-\bullet \arrow[dr,densely dotted,swap,near start,"(W_A/\ctxext{B}{Q})(W_Q(R))"] & & \bullet \arrow[dl,densely dotted,near start,"W_{(W_A/B)(Q)}((W_A/B)(R))"] & & & & & & & & & & \bullet \arrow[dl,"W_Q(R)"] \\
-& \bullet \arrow[dr,swap,"(W_A/B)(Q)"] & & \bullet \arrow[dl,"(W_A/B)(R)"] & & & & & & & & \bullet \arrow[dr,swap,"Q"] & & \bullet \arrow[dl,"R"] \\
-& & \bullet \arrow[d,swap,"W_A(B)"] & & & & & & & & & & \ctxext{\Gamma}{B} \arrow[d,"B"] \\
-& & \ctxext{\Gamma}{A} \arrow[rrrrrrrrrr,swap,"A"] & & & & & & & & & & \Gamma
-\end{tikzcd}
-\end{equation*}
-are equal.
-
-A useful special case of this property is where $B\jdeq \catid{\Gamma}$. Thus, if $W$ is
-a weakening system, then the diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=huge]
-\cat{C}/\ctxext{\Gamma}{C}
-  \arrow[r,"W_A/C"]
-  &
-\cat{C}/\ctxext{{\Gamma}{A}}{W_A(C)}
-  \\
-\cat{C}/\Gamma
-  \arrow[u,"W_C"]
-  \arrow[r,swap,"W_A"]
-  &
-\cat{C}/\ctxext{\Gamma}{A}
-  \arrow[u,swap,"W_{W_A(C)}"]
-\end{tikzcd}
-\end{equation*}
-commutes for every $A,C\in\cat{C}/\Gamma$. In particular, we see
-that $W_A(W_C(D))\jdeq W_{W_A(C)}(W_A(D))$ for any $D\in\cat{C}/\Gamma$,
-i.e.~that weakening is a self-distributive operation. 
-\end{rmk}
-
-\begin{cor}
-For any object $\Gamma$ of a weakening system $\cat{C}$, the slice pre-category
-$\cat{C}/\Gamma$ is a weakening system.
-\end{cor}
-
-\begin{eg}
-If the underlying pre-category is a group, the axioms of a weakening system
-obtain the following form: There is a group homomorphism $W_x:G\to G$ for
-every $x\in G$, in such a way that $W_e\jdeq \catid{G}$ and $W_{xy}\jdeq W_y\circ W_x$
-for every $x,y\in G$, and for any $x,y\in G$ the diagram
-\begin{equation*}
-\begin{tikzcd}
-G \arrow[r,"W_y"]
-  &
-G 
-  \\
-G \arrow[u,"W_x"]
-  \arrow[r,swap,"W_y"]
-  &
-G \arrow[u,swap,"W_{W_y(x)}"]
-\end{tikzcd}
-\end{equation*}
-is required to commute. Thus, a weakening system on a group $G$ consists of
-a group homomorphism $W:\op G\to\mathrm{Aut}(G)$ such that $W_a\circ W_b\jdeq 
-W_{W_a(b)}\circ W_a$. The latter requirement is satisfied if $b\cdot a\jdeq a\cdot W_a(b)$
-for every $a,b\in G$, or equivalently if $[b,a]\jdeq b^{-1}\cdot W_a(b)$.
-\end{eg}
-
-\begin{lem}
-For any group, conjugation is a weakening structure. Thus, any group is
-canonically a weakening system.
-\end{lem}
-
-\begin{proof}
-Note that $e^{-1}xe\jdeq x$ and $(xy)^{-1}z(xy)\jdeq y^{-1}(x^{-1}zx)y$ for any $x,y,z\in G$.
-Also, the diagram
-\begin{equation*}
-\begin{tikzcd}
-G \arrow[r,"\mathrm{conj}_y"]
-  &
-G 
-  \\
-G \arrow[u,"\mathrm{conj}_x"]
-  \arrow[r,swap,"\mathrm{conj}_y"]
-  &
-G \arrow[u,swap,"\mathrm{conj}_{y^{-1}xy}"]
-\end{tikzcd}
-\end{equation*}
-commutes.
-\end{proof}
-
-\begin{comment}
-\begin{rmk}
-Similarly, every groupoid has a canonical weakening structure given by
-conjugation.
-\end{rmk}
-\end{comment}
-
-\begin{comment}
-\begin{lem}
-Let $G$ be a graph, and let $\cat{C}_G$ be the free pre-category generated by
-$G$. A pre-weakening structure on $\cat{C}_G$ is determined by functors
-$W_e:\cat{C}_G/\Gamma\to\cat{C}_G/\Delta$ for every edge $e:\Delta\to \Gamma$ in $G$, with the
-property that $W_e(\catid{\Gamma})\jdeq \catid{\Delta}$.
-
-A weakening structure on $\cat{C}_G$ is
-a pre-weakening structure on $\cat{C}_G$ such that the diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=huge]
-\cat{C}_G/\greek{Z} \arrow[r,"W_e/(g\circ h)"] & \cat{C}_G/\mathrm{dom}(W_e(g\circ h)) \\
-\cat{C}_G/\greek{E} \arrow[u,"W_h"] \arrow[r,swap,"W_e/g"] & \cat{C}_G/\mathrm{dom}(W_e(g)) \arrow[u,swap,"W_{W_e(h)}"]
-\end{tikzcd}
-\end{equation*}
-commutes for every arrow $g:\greek{E}\to \Gamma$ in $\cat{C}_G$ and every two edges $e:\Delta\to \Gamma$ and
-$h:\greek{Z}\to \greek{E}$ in $G$.
-\end{lem}
-
-\begin{eg}
-Let $G$ be the graph consisting of one vertex $\ast$ and one edge $e:\ast\to\ast$.
-Then $\cat{C}_G/\ast\jdeq \N$. Then the only weakening structure on $\cat{C}/G$ is
-given by $W_e\jdeq \catid{\N}$. The reason is that 
-\end{eg}
-\end{comment}
-
-\begin{lem}
-Let $P$ be a pre-order. A pre-weakening structure on $P$ is determined by order
-preserving maps $W_{a,b}:{\downarrow}(b)\to{\downarrow}(a)$ satisfying
-$W_{a,b}(b)\jdeq a$, for every $a\leq b$, where
-${\downarrow}(x)$ is the set $\{y\in P\mid y\leq x\}$.
-
-A weakening structure on $P$ is a pre-weakening structure on $P$ such that the
-diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=huge]
-{\downarrow}(c') \arrow[r,"W_{a,b}|_{{\downarrow}(c')}"] & {\downarrow}(W_{a,b}(c')) \\
-{\downarrow}(c) \arrow[u,"W_{c',c}"] \arrow[r,swap,"W_{a,b}|_{{\downarrow}(c)}"] & {\downarrow}(W_{a,b}(c)) \arrow[u,swap,"W_{W_{a,b}(c'),W_{a,b}(c)}"]
-\end{tikzcd}
-\end{equation*}
-for any $a\leq b$ and $c'\leq c\leq b$.
-\end{lem}
-
-\begin{eg}
-In a poset with meets, the functions $W_{a,b}:{\downarrow}(b)\to{\downarrow}(a)$
-given by $W_{a,b}(x)\defeq a\land x$ assemble a weakening structure, because for
-any $a\leq b$ and $x,c'\leq c\leq b$ we have $a\land(c'\land x)\jdeq (a\land c')\land
-(a\land x)$.  
-\end{eg}
-
-\begin{eg}
-Consider the poset $(\N,\geq)$, which has a morphism from $a$ to $b$ if $a\geq b$.
-On this poset, we may consider the weakening structure defined by $W_{a,b}(x)\jdeq
-a+(x-b)$.
-\end{eg}
-
-\subsection{Weakening systems with term structure}
-\begin{defn}
-A \emph{category with a term structure} is a pre-category $\cat{C}$ with a set $T(A)$ for every
-$A\in\cat{C}/\Gamma$ for every $\Gamma\in\cat{C}$. A \emph{functor with term structure} from $\cat{C}$ to $\cat{D}$
-is a functor $F:\cat{C}\to\cat{D}$ with a function $T(A)\to T(F(A))$ for
-every $A\in\cat{C}/\Gamma$, for every $\Gamma\in\cat{C}$.
-\end{defn}
-
-\begin{defn}
-Suppose $\cat{C}$ is a pre-category with term structure, and $\Gamma$ is an object of $\cat{C}$. 
-Then $\cat{C}/\Gamma$ is a pre-category with term structure, where $T_{\cat{C}/\Gamma}(A)\jdeq T_\cat{C}(A)$.
-\end{defn}
-
-\begin{defn}
-A \emph{weakening system with term term structure} is the same thing as a weakening system,
-except that the underlying pre-category and the functors $W_A$ are required to
-be functors with term term structure. The equalities are also required to hold on the term structure.
-\end{defn}
-
-\begin{cor}
-For any object $\Gamma$ of a weakening system with term structure, the slice pre-category
-$\cat{C}/\Gamma$ is a weakening system with term structure.
-\end{cor}
-
-\begin{eg}
-%For any group $G$ with weakening structure $W:\op{G}\to\mathrm{Aut}(G)$, a term 
-%structure is determined by a presheaf on $G/\ker(W)$.
-
-Let $G$ be a group, with the weakening structure given by conjugation. A term
-structure on $G$ is given by a set $X$ with a right $G$-action $\mu:\op G\times X\to X$.
-We take $T(a)\defeq X$ for every $a\in G$.
-\end{eg}
-
-\subsection{Projection systems}
-\begin{defn}
-A \emph{pre-projection system} is a pre-weakening system with term structure with an element
-$\tfid{A}\in T(W_A(A))$ for every $A\in\cat{C}/\Gamma$ and $\Gamma\in\cat{C}$. 
-\end{defn}
-
-\begin{defn}
-A \emph{pre-projection homomorphism} $F:\cat{C}\to\cat{D}$ is a pre-weakening homomorphism for which
-\begin{equation*}
-F(\tfid{A})\jdeq \tfid{F(A)}
-\end{equation*}
-for every $A\in\cat{C}/\Gamma$ and $\Gamma\in\cat{C}$.
-\end{defn}
-
-\begin{lem}
-For any object $\Gamma$ in a pre-projection system $\cat{C}$, the slice pre-category
-$\cat{C}/\Gamma$ is a pre-projection system.
-\end{lem}
-
-\begin{proof}
-Straightforward.
-\end{proof}
-
-\begin{defn}
-A \emph{projection system} is a pre-projection system for which every $W_A$ is
-a pre-projection homomorphism. A \emph{projection homomorphism} is a pre-projection homomorphism
-between projection systems.
-\end{defn}
-
-\begin{cor}
-For any object $\Gamma$ of a projection system $\cat{C}$, the slice pre-category $\cat{C}/\Gamma$
-is a projection system.
-\end{cor}
-
-%\begin{eg}
-%Consider a group $G$ with a right $G$-set $X$. A projection structure on $(G,X)$
-%consists of an element $\tfid{a}\in X$ for every $a\in G$, so that $\tfid{a}\cdot b=
-%\tfid{b^{-1}ab}$ for every $a,b\in G$. 
-%For instance, this is the case if we take $X$ to be a pointed set with $\ast\in X$,
-%and take $\tfid{a}:=\ast$ for all $a\in G$.
-%\end{eg}
-
-\subsection{Substitution systems}
-
-\begin{defn}
-A \emph{pre-substitution structure} on a pre-category with term structure $\cat{C}$ consists
-of a functor with term structure $S_x:\cat{C}/\ctxext{\Gamma}{A}\to\cat{C}/\Gamma$ for every $x\in T(A)$ and
-$A\in\cat{C}/\Gamma$, such that $S_x(\catid{\ctxext{\Gamma}{A}})\jdeq\catid{\Gamma}$.
-
-A \emph{pre-substitution system} is a pre-category with term structure together with a
-pre-substitution structure. 
-\end{defn}
-
-\begin{defn}
-A \emph{pre-substitution homomorphism} $F:\cat{C}\to\cat{D}$ is a functor with term structure for
-which the diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=large]
-\cat{C}/\ctxext{\Gamma}{A}
-  \arrow[r,"F/\ctxext{\Gamma}{A}"]
-  \arrow[d,swap,"S_x"]
-  &
-\cat{D}/F(\ctxext{\Gamma}{A})
-  \arrow[d,"S_{F(x)}"]
-  \\
-\cat{C}/\Gamma
-  \arrow[r,swap,"F/\Gamma"]
-  &
-\cat{D}/F(\Gamma)
-\end{tikzcd}
-\end{equation*}
-commutes for every $x\in T(A)$ and $A\in\cat{C}/\Gamma$.
-\end{defn}
-
-\begin{lem}
-For any object $\Gamma$ of a pre-substitution system $\cat{C}$, the slice pre-category
-$\cat{C}/\Gamma	$ is a pre-substitution system.
-\end{lem}
-
-\begin{defn}
-A \emph{substitution system} is a pre-substitution system for which each
-$S_t$ is a pre-substitution homomorphism. A \emph{substitution homomorphism} is a
-pre-substitution homomorphism between substitution systems.
-\end{defn}
-
-\begin{cor}
-For any object $\Gamma$ of a substitution system, the slice pre-category $\cat{C}/\Gamma$
-is a substitution system.
-\end{cor}
-
-\begin{rmk}
-The condition that every $S_x$ is a substitution homomorphism, asserts that
-the diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=large]
-\cat{C}/\ctxext{{{\Gamma}{A}}{P}}{Q}
-  \arrow[r,"S_x/\ctxext{P}{Q}"]
-  \arrow[d,swap,"S_y"]
-  &
-\cat{C}/\ctxext{{\Gamma}{S_x(P)}}{S_x(Q)}
-  \arrow[d,"S_{S_x(y)}"]
-  \\
-\cat{C}/\ctxext{{\Gamma}{A}}{P}
-  \arrow[r,swap,"S_x/P"]
-  &
-\cat{C}/\ctxext{\Gamma}{S_x(P)}
-\end{tikzcd}
-\end{equation*}
-commutes for every $y\in T(Q)$.
-\end{rmk}
-
-\begin{eg}
-Let $G$ be a group and let $H$ be a set, forming a category with term 
-structure of which $T(a)\jdeq H$ for every $a\in G$. A substitution
-system on $(G,H)$ consists of a function $\sigma_G : H\times G\to G$ and
-a function $\sigma_H:H\times H\to H$, for which the diagrams
-\begin{equation*}
-\begin{tikzcd}[column sep=large]
-G \arrow[r,"\sigma_G(h)"]
-  \arrow[d,swap,"\sigma_G(k)"]
-  &
-G \arrow[d,"{\sigma_G(\sigma_H(h,k))}"]
-  \\
-G \arrow[r,swap,"\sigma_G(h)"] 
-& G
-\end{tikzcd}
-\end{equation*}
-and
-\begin{equation*}
-\begin{tikzcd}[column sep=large]
-H \arrow[r,"\sigma_H(h)"]
-  \arrow[d,swap,"\sigma_H(k)"]
-  &
-H \arrow[d,"{\sigma_H(\sigma_H(h,k))}"]
-  \\
-H \arrow[r,swap,"\sigma_H(h)"]
-  &
-H
-\end{tikzcd}
-\end{equation*}
-commute for every $h,k\in H$. 
-\end{eg}
-
-\subsection{The definition of E-systems}
-
-Although it is now possible to state the definition of an E-system without
-any difficulty, it shall be convenient later to also have a category of
-pre-E-systems at our disposal. Pre-E-systems shall by default be non-unital,
-i.e.~they do not come with a pre-projection system, while E-systems shall
-always be assumed to be unital. In the present context, we shall not need
-unital pre-E-systems or non-unital E-systems.
-
-\begin{defn}
-A \emph{pre-E-system} is a pre-category $\cat{C}$ with term structure
-with a pre-weakening structure $W$, a pre-substitution structure $S$ and a terminal
-object. A \emph{unital pre-E-system}, or simply \emph{pre-E-system}, is a non-unital pre-E-system
-which also has the structure of a pre-projection system. 
-\end{defn}
-
-\begin{defn}
-A \emph{pre-E-homomorphism} from $\mathbb{E}\to\mathbb{D}$ is a functor
-$H:\mathbb{C}_\mathbb{E}\to\mathbb{C}_\mathbb{D}$ between the underlying categories
-with term structure, which is a pre-weakening homomorphism, a pre-substitution
-homomorphism, and which preserves the terminal object.
-
-A \emph{unital} pre-E-homomorphism is a non-unital pre-E-homomorphism which is 
-also a pre-projection homomorphism. 
-\end{defn}
-
-\begin{defn}
-\label{defn:esystems}
-An \emph{E-system} is a unital pre-E-system in which 
-\begin{enumerate}
-\item each $W_A$ is a unital pre-E-homomorphism,
-\item each $S_x$ is a unital pre-E-homomorphism,
-\item \label{tTf:StWf_id} weakened families are constant, i.e.~$S_x\circ W_A\jdeq \catid{\cat{C}/\Gamma}$ for any $x\in T(A)$ and $A\in\cat{C}/\Gamma$,
-\item $S_x(\tfid{A})\jdeq x$ for any $x\in T(A)$ and $A\in\cat{C}/\Gamma$, and
-\item \label{f:SidfWff_id} $S_{\tfid{A}}\circ W_A/A\jdeq \catid{\cat{C}/\ctxext{\Gamma}{A}}$ for any $A\in\cat{C}/\Gamma$.
-\end{enumerate}
-An \emph{E-homomorphism} $H:\mathbb{E}\to\mathbb{D}$ is a functor from
-$\cat{C}_{\mathbb{E}}$ to $\cat{C}_{\mathbb{D}}$ which is both a projection
-homomorphism and a substitution homomorphism.
-\end{defn}
-
-\begin{rmk}
-The condition that each $W_A$ is a substitution homomorphism asserts that
-the diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=huge]
-\cat{C}/\ctxext{{\Gamma}{B}}{Q}
-  \arrow[r,"W_A/\ctxext{B}{Q}"]
-  \arrow[d,swap,"S_y"]
-  &
-\cat{C}/\ctxext{{{\Gamma}{A}}{W_A(B)}}{W_A(Q)}
-  \arrow[d,"S_{W_A(y)}"]
-  \\
-\cat{C}/\ctxext{\Gamma}{B}
-  \arrow[r,swap,"W_A/B"]
-  &
-\cat{C}/\ctxext{{\Gamma}{A}}{W_A(B)}
-\end{tikzcd}
-\end{equation*}
-of functors with term structure commutes for every 
-$Q\in\cat{C}/\ctxext{\Gamma}{B}$, $B\in\cat{C}/\Gamma$ and each $y\in T(Q)$.
-
-Likewise, the condition that each $S_x$ is a weakening homomorphism
-asserts that the diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=huge]
-\cat{C}/\ctxext{{\Gamma}{A}}{P}
-  \arrow[r,"S_x/P"]
-  \arrow[d,swap,"W_Q"]
-  &
-\cat{C}/\ctxext{\Gamma}{S_x(P)}
-  \arrow[d,"W_{S_x(Q)}"]
-  \\
-\cat{C}/\ctxext{{{\Gamma}{A}}{P}}{Q}
-  \arrow[r,swap,"S_x/\ctxext{P}{Q}"]
-  &
-\cat{C}/\ctxext{{\Gamma}{S_x(P)}}{S_x(Q)}
-\end{tikzcd}
-\end{equation*}
-of functors with term structure commutes for every 
-$Q\in\cat{C}/\ctxext{{\Gamma}{A}}{P}$.
-\end{rmk}
-
-\begin{cor}
-For any object $\Gamma$ of an E-system $\cat{C}$, the slice pre-category $\cat{C}/\Gamma$
-is an E-system.
-\end{cor}
-
-\begin{comment}
-\begin{rmk}[Needs more precision]
-Consider the category $E$ generated by the graph consisting of the objects of 
-the underlying category of an $E$-system $\mathcal{E}$ as edges, and the morphisms
-$S_t:X\to Y$ and $W_A:Y\to X$ as edges, subject to the axioms of an $E$-systems.
-
-Any morphism of $E$ can be rewritten in a unique way
-as a composition $\tilde S\circ \tilde W$, where $\tilde S$ is a morphism in
-substitution-normal form, $\tilde W$ is in weakening normal form, every term
-expression is in normal form, and where
-each occurence of $S_x\circ W_A$ and $S_{\tfid{A}}\circ W_A/A$ is eliminated.
-\end{rmk}
-\end{comment}
-
-\section{Properties of E-systems}\label{sec:esys_props}
-
-The theory of E-systems could be seen as a sort of category theory with only
-projections and their sections. We elaborate on this point of view by defining
-the category of internal homomorphisms, which Lawvere might have called the
-category of terms. 
-In the category of internal morphisms over a context $\Gamma$,
-a morphism $f:A\to B$ is a term of the constant family $W_A(B)$ over 
-$\ctxext{\Gamma}{A}$. Rather than directly defining composition, we will define
-for every $f:A\to B$ in context $\Gamma$, a pre-composition operation
-$f_\ast$, which is going to be an E-homomorphism from 
-$\cat{C}/\ctxext{\Gamma}{B}$ to $\cat{C}/\ctxext{\Gamma}{A}$. This allows us
-to consider diagrams of the following form:
-\begin{equation*}
-\begin{tikzcd}
-P \arrow[d,->>] \arrow[r,"g"] & Q \arrow[d,->>] \\
-A \arrow[r,swap,"f"] & B
-\end{tikzcd}
-\end{equation*}
-where we consider the morphism $g$ over $f$ to be an internal morphism from 
-$P$ to $f_\ast(Q)$ in context $\ctxext{\Gamma}{A}$, capturing the
-essence of being a morphism from the family $P$ over $A$ to the family
-$Q$ over $B$, reindexed by $f$. 
-Whenever we draw such diagrams, a double headed arrow will always
-refer to a family, and an ordinary arrow will refer to an internal morphism,
-possibly over another internal morphism
-
-A morphism $g$ over $f$ can be vertically composed with $f$ so
-that we obtain a morphism $\jvcomp{P}{f}{g}:\ctxext{A}{P}\to\ctxext{B}{Q}$ in 
-context $\Gamma$, and it can be horizontally composed with a morphism $k:Q\to R$
-over $h:B\to C$ so that we obtain a morphism $\jfcomp{A}{f}{P}{g}{k}:P\to R$ 
-over $\jcomp{A}{f}{h}:A\to C$. We will show interchange laws between these
-sorts of composition. The interchange laws have as a consequence that if we
-assert the commutativity of a large composed diagram.
-
-We then demonstrate how to define a category internal to an E-system. The main
-point of this lies mainly in demonstrating how the operations of E-systems can
-be used to define a category, because it is no surprise that a category can be
-defined in type theory. Nevertheless, it is noteworthy that we need no further 
-conditions on E-systems.
-This is unlike the situation in category theory, where we need
-to assume that the ambient category has finite limits in order to say what an
-internal category is. In fact, it is possible to define an E-system internal
-to an E-system, just using the algebraic structure of an E-system, and actually
-defining an internal category was the hardest part of that definition.
-
-Before we begin this section, we introduce some handy shorthands.
-
-\begin{defn}
-Let $A\in\cat{C}/\Gamma$. Recall that $W_A:\cat{C}/\Gamma\to\cat{C}/\ctxext{\Gamma}{A}$ acts on objects,
-morphisms and terms. We introduce the infix form of weakening by $A\in\cat{C}/\Gamma$ to be
-$\ctxwk{A}{\blank}$. Thus, we will write
-\begin{align*}
-\ctxwk{A}{B} & \defeq W_A(B) & & \text{for $B\in\cat{C}/\Gamma$} \\
-\ctxwk{A}{Q} & \defeq W_A(Q) & & \text{for $B\in\cat{C}/\Gamma$ and $Q\in\cat{C}/\ctxext{\Gamma}{B}$}\\
-\ctxwk{A}{g} & \defeq W_A(g) & & \text{for $B\in\cat{C}/\Gamma$, $Q\in\cat{C}/\ctxext{\Gamma}{B}$ and $g\in T(Q)$}
-\end{align*}
-\end{defn}
-
-\begin{defn}
-Let $x\in T(A)$ for a family $A\in\cat{C}/\Gamma$. The infix form of substitution
-by $x$ is taken to be $\subst{x}{\blank}$. Thus, we will write
-\begin{align*}
-\subst{x}{P} & \defeq S_x(P) & & \text{for $P\in\cat{C}/\ctxext{\Gamma}{A}$} \\
-\subst{x}{Q} & \defeq S_x(Q) & & \text{for $P\in\cat{C}/\ctxext{\Gamma}{A}$ and $Q\in\cat{C}/\ctxext{{\Gamma}{A}}{P}$} \\
-\subst{x}{g} & \defeq S_x(g) & & \text{for $P\in\cat{C}/\ctxext{\Gamma}{A}$, $Q\in\cat{C}/\ctxext{{\Gamma}{A}}{P}$ and $g\in T(Q)$}
-\end{align*}
-\end{defn}
-
-\subsection{The category of internal morphisms}\label{subsec:im_cat}
-
-\begin{defn}
-For every $A,B\in\cat{C}/\Gamma$ we define the set 
-\begin{equation*}
-\thom{A}{B}\defeq T(\ctxwk{A}{B}).
-\end{equation*}
-An element $f\in \thom{A}{B}$ is called an \emph{internal morphism in context
-$\Gamma$}. We sometimes write $\jhom{\Gamma}{A}{B}{f}$ to indicate that $f$ is
-an internal morphism over $\Gamma$, or we may draw a diagram of the form
-\begin{equation*}
-\begin{tikzcd}[column sep=tiny]
-A \arrow[rr,"f"] \arrow[dr,fib] & & B \arrow[dl,fib] \\
-& \Gamma
-\end{tikzcd}
-\end{equation*}
-but we shall generally omit the arrows down to $\Gamma$ and say instead that we have
-a diagram in context $\Gamma$. 
-\end{defn}
-
-\begin{rmk}
-Note that $\thom{\catid{\Gamma}}{A}\jdeq T(A)$ for any $A\in\cat{C}/\Gamma$, 
-because we have 
-$W_{\catid{\ctxext{\Gamma}{A}}}\jdeq \catid{\cat{C}/\ctxext{\Gamma}{A}}$.
-
-Note also that $\thom{\ctxext{A}{P}}{B}\jdeq \thom{P}{\ctxwk{A}{B}}$ 
-for any $P\in\cat{C}/\ctxext{\Gamma}{A}$ and $B\in\cat{C}/\Gamma$,
-because $W_{\ctxext{A}{P}}\jdeq W_P\circ W_A$.
-Once we have established a category of which the morphisms are given by 
-$\thom{\blank}{\blank}$, we therefore get that 
-\begin{equation*}
-\ctxext{A}{(\blank)}\dashv W_A.
-\end{equation*}
-The right adjoint to weakening by $A$, if it exists, will be the dependent
-product $\Pi_A$. 
-\end{rmk}
-
-\begin{defn}
-Let $A,B\in\cat{C}/\Gamma$.
-For any $f\in\thom{A}{B}$ we define the pre-composition E-homomorphism
-\begin{equation*}
-f_\ast \defeq S_f\circ W_A/B : \cat{C}/\ctxext{\Gamma}{B}\to\cat{C}/\ctxext{\Gamma}{A}.
-\end{equation*}
-Furthermore, for $g\in\thom{B}{C}$ we may consider $W_B(C)\in\cat{C}/\ctxext{\Gamma}{B}$ as a morphism from $W_B(C)$ to $\catid{\ctxext{\Gamma}{B}}$. Therefore, $f_\ast$ acts on the terms of
-$W_B(C)$, so it makes sense to define $g\circ f\defeq f_\ast(g)$. In fact, it makes
-sense to let $(\blank)\circ f$ be the infix notation for $f_\ast$. 
-\end{defn}
-
-\begin{rmk}
-Note that for $\jhom{\Gamma}{B}{C}{g}$, 
-we have $g\circ f\in T(S_f(W_A/B(W_B(C))))$, whereas we would like that
-$g\circ f\in\thom{A}{C}$. More generally, we can show that
-\begin{equation*}
-S_f\circ (W_A/B)\circ W_B\jdeq W_A.
-\end{equation*}
-Since weakening is a weakening homomorphism, we have
-\begin{equation*}
-S_f\circ (W_A/B)\circ W_B\jdeq S_f\circ W_{W_A(B)}\circ W_A
-\end{equation*}
-By \autoref{tTf:StWf_id} of \autoref{defn:esystems} it we get that
-\begin{equation*}
-S_f\circ W_{W_A(B)}\circ W_A\jdeq W_A.
-\end{equation*}
-\end{rmk}
-
-\begin{rmk}
-To see how this extended notion of precomposition works, suppose we have a diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=tiny]
-Q \arrow[rr,"g"] \arrow[dr,fib] & & R \arrow[dl,fib] \\
-& B
-\end{tikzcd}
-\end{equation*}
-in context $\Gamma$, i.e.~$\jhom{\ctxext{\Gamma}{B}}{Q}{R}{g}$, and let $\jhom{\Gamma}{A}{B}{f}$. Precomposing by $f$ gives
-us a diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=tiny]
-\jcomp{A}{f}{Q} \arrow[rr,"\jcomp{A}{f}{g}"] \arrow[dr,fib] & & \jcomp{A}{f}{R} \arrow[dl,fib] \\
-& A
-\end{tikzcd}
-\end{equation*}
-in context $\Gamma$.
-\end{rmk}
-
-\begin{rmk}
-Note that \autoref{f:SidfWff_id} of \autoref{defn:esystems} asserts precisely
-that ${(\tfid{A})}_\ast\jdeq \catid{\cat{C}/\ctxext{\Gamma}{A}}$ for any $A\in\cat{C}/\Gamma$. In
-particular, it follows that $g\circ\tfid{A}\jdeq g$ for any $g\in\thom{A}{B}$
-\end{rmk}
-
-\begin{lem}\label{lem:compcomp}
-For any $f\in\thom{A}{B}$ and $g\in\thom{B}{C}$ we have $f_\ast\circ g_\ast\jdeq (g\circ f)_\ast$. 
-\end{lem}
-
-\begin{proof}
-\begin{align*}
-f_\ast\circ g_\ast & \jdeq S_f\circ (W_A/B) \circ S_g \circ (W_B/C)\\
-& \jdeq S_f\circ S_{W_A(g)}\circ(W_A/\ctxext{B}{W_B(C)})\circ (W_B/C) \\
-& \jdeq S_{S_f(W_A(g))}\circ (S_f/W_A(W_B(C)))\circ (W_A/\ctxext{B}{W_B(C)}) \circ W_B/C \\
-& \jdeq S_{S_f(W_A(g))}\circ ((S_f\circ (W_A/B) \circ W_B)/C) \\
-& \jdeq S_{S_f(W_A(g))}\circ ((S_f\circ W_{W_A(B)}\circ W_A)/C) \\
-& \jdeq S_{S_f(W_A(g))}\circ W_A/C \\
-& \jdeq (g\circ f)_\ast.\qedhere
-\end{align*}
-\end{proof}
-
-\begin{thm}
-For each object $\Gamma$ of an E-system $\mathbb{E}$, we have a category
-$\mathbb{C}_I(\mathbb{E},\Gamma)$ of internal morphisms of $\mathbb{E}$,
-consisting of the same objects as $\cat{C}/\Gamma$ and the hom-sets 
-$\thom{A}{B}$. 
-\end{thm}
-
-\begin{proof}
-The fact that composition is associative is a direct corollary of
-\autoref{lem:compcomp}. The axiom $(\tfid{A})_\ast\jdeq\catid{\ctxext{\Gamma}{A}}$
-implies that the identity morphisms satisfy the right identity law. It remains
-to show that $\tfid{B}\circ f\jdeq f$. This is a simple calculation:
-\begin{equation*}
-\tfid{B}\circ f 
-  \jdeq
-\subst{f}{\ctxwk{A}{\tfid{B}}}
-  \jdeq
-\subst{f}{\tfid{\ctxwk{A}{B}}}
-  \jdeq
-f.\qedhere
-\end{equation*}
-\end{proof}
-
-\begin{rmk}
-Since any E-homomorphism is both a substitution homomorphism and a weakening
-homomorphism, it follows immediately that any E-homomorphism preserves any
-$f_\ast$, i.e.~for any $f\in\thom{A}{B}$ in $\mathcal{E}$ and any E-homomorphism
-$F:\mathcal{E}\to\mathcal{D}$, the diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=large]
-\mathcal{E}/\ctxext{\Gamma}{A} \arrow[r,"F/\ctxext{\Gamma}{A}"] & \mathcal{D}/F(\ctxext{\Gamma}{A})
-  \\
-\mathcal{E}/\ctxext{\Gamma}{B} \arrow[u,"f_\ast"] \arrow[r,swap,"F/\ctxext{\Gamma}{B}"] & \mathcal{D}/F(\ctxext{\Gamma}{B}) \arrow[u,swap,"F(f)_\ast"]
-\end{tikzcd}
-\end{equation*}
-commutes. This applies to weakening, substitution and pre-composition, and it
-follows that for any E-homomorphism $F:\mathcal{E}\to\mathcal{D}$ and any
-$\Gamma\in\mathcal{E}$, we have a functor $\mathbb{C}_I(F,\Gamma):
-\mathbb{C}_I(\mathcal{E},\Gamma)\to\mathbb{C}_I(\mathcal{D},F(\Gamma))$.
-\end{rmk}
-
-The following fact will be useful:
-
-\begin{lem}\label{lem:compW_W}
-Let $f\in\thom{A}{B}$ be an internal morphism in context $\Gamma$. Then one has
-\begin{equation*}
-f_\ast\circ W_B\jdeq W_A.
-\end{equation*}
-\end{lem}
-
-\begin{proof}
-The proof is a simple calculation:
-\begin{equation*}
-f_\ast\circ W_B\jdeq S_f\circ W_A/B\circ W_B\jdeq S_f\circ W_{W_A(B)}\circ W_A
-\jdeq W_A.\qedhere
-\end{equation*}
-\end{proof}
-
-\subsection{Pairing and the projections}
-
-The composition $\ctxext{A}{P}$ of $A\in\cat{C}/\Gamma$ and $P\in\cat{C}/\ctxext{\Gamma}{A}$
-behaves like a strict $\Sigma$-type. We may define the pairing term
-$\typefont{pair}^{A,P}\defeq\tfid{\ctxext{A}{P}}\in T(W_P(W_A(\ctxext{A}{P})))$
-and the projections and prove several useful properties about them. The strictness
-is found, among other things, in the fact that we can prove judgmental $\eta$-equality,
-and that pairing is strictly associative.
-
-In this section we will make more extensive use of the infix form of the
-weakening and substitution operations.
-
-\begin{defn}
-Let $x\in T(A)$ and $u\in T(S_x(P))$ for $A\in\cat{C}/\Gamma$ and $P\in\cat{C}/\ctxext{\Gamma}{A}$. 
-We define the \emph{term extension of $x$ and $u$} to be
-\begin{equation*}
-\tmext{x}{u}\defeq \subst{u}{{x}{\tfid{\ctxext{A}{P}}}}\in T(\ctxext{A}{P}).
-\end{equation*}
-\end{defn}
-
-To prove anything about the term $\tmext{x}{u}$, we need the following property.
-
-\begin{thm}\label{subst_by_tmext}
-Let $x\in T(A)$ and $u\in T(S_x(P))$ for $A\in\cat{C}/\Gamma$ and $P\in\cat{C}/\ctxext{\Gamma}{A}$.
-Then we have
-\begin{equation*}
-S_{\tmext{x}{u}}\jdeq S_u\circ (S_x/P):\cat{C}/\ctxext{\Gamma}{{A}{P}}\to\cat{C}/\Gamma
-\end{equation*}
-\end{thm}
-
-\begin{proof}
-\begin{align*}
-S_{\tmext{x}{u}} & \jdeq S_{S_u(S_x(\tfid{\ctxext{A}{P}}))} \\
-& \jdeq S_{S_u(S_x(\tfid{\ctxext{A}{P}}))}\circ (S_u\circ W_{S_x(P)})\circ (S_x\circ W_{A}) \\
-& \jdeq S_u \circ S_{S_x(\tfid{\ctxext{A}{P}})} \circ W_{S_x(P)}\circ S_x\circ W_A \\
-& \jdeq S_u \circ (S_x/P) \circ S_{\tfid{\ctxext{A}{P}}} \circ W_P\circ W_A \\
-& \jdeq S_u \circ (S_x/P) \circ S_{\tfid{\ctxext{A}{P}}} \circ W_{\ctxext{A}{P}}\\
-& \jdeq S_u \circ (S_x/P).\qedhere
-\end{align*}
-\end{proof}
-
-\begin{defn}
-Let $A\in\cat{C}/\Gamma$ and $P\in\cat{C}/\ctxext{\Gamma}{A}$. We define
-\begin{align*}
-\cprojfstf{A}{P} & \defeq \ctxwk{P}{\tfid{A}}\in\thom{\ctxext{A}{P}}{A}\\
-\cprojsndf{A}{P} & \defeq \tfid{P}\in\thomd{\cprojfstf{A}{P}}{\ctxext{A}{P}}{P}
-\end{align*}
-\end{defn}
-
-\begin{cor}\label{cor:pairing_ehom}
-Let $F:\mathcal{E}\to\mathcal{D}$ be an E-homomorphism and let $x\in T(A)$ and
-$u\in T(S_x(P))$. Then we have
-\begin{equation*}
-F(\tmext{x}{u})\jdeq \tmext{F(x)}{F(u)}.
-\end{equation*}
-Also, we have
-\begin{align*}
-F(\cprojfstf{A}{P}) & \jdeq \cprojfstf{F(A)}{F(P)} \\
-F(\cprojsndf{A}{P}) & \jdeq \cprojsndf{F(A)}{F(P)}.
-\end{align*}
-\end{cor}
-
-\begin{thm}\label{thm:famext_up}
-For any $A\in\cat{C}/\Gamma$ and $P\in\cat{C}/\ctxext{\Gamma}{A}$,
-the map
-\begin{equation*}
-(x,u)\mapsto \tmext{x}{u}:\big(\bigsqcup\nolimits_{x\in T(A)} T(\subst{x}{P})\big)\to T(\ctxext{A}{P}).
-\end{equation*}
-is an isomorphism.  
-\end{thm}
-
-\begin{proof}
-Lets write $\varphi$ for the indicated map. The inverse $\psi$ of $\varphi$ is defined
-by $w\mapsto (\subst{w}{\cprojfstf{A}{P}},\subst{w}{\cprojsndf{A}{P}})$. We have
-to show that for any $x\in T(A)$ and $u\in T(\subst{x}{P})$, one has
-\begin{align*}
-\subst{\tmext{x}{u}}{\cprojfstf{A}{P}} & \jdeq x\\
-\subst{\tmext{x}{u}}{\cprojsndf{A}{P}} & \jdeq u,
-\end{align*}
-and that for any $w\in T(\ctxext{A}{P})$, one has
-\begin{equation*}
-\tmext{\subst{w}{\cprojfstf{A}{P}}}{\subst{w}{\cprojsndf{A}{P}}} \jdeq w.
-\end{equation*}
-To show that $\subst{\tmext{x}{u}}{\cprojfstf{A}{P}} \jdeq x$, we use that
-$S_{\tmext{x}{u}}\jdeq S_u\circ S_x/P$ to show that
-\begin{align*}
-\subst{\tmext{x}{u}}{\cprojfstf{A}{P}}
-  & \jdeq 
-\subst{\tmext{x}{u}}{\ctxwk{P}{\tfid{A}}} \\
-  & \jdeq
-\subst{u}{{x}{\ctxwk{P}{\tfid{A}}}} \tag{By \autoref{subst_by_tmext}}\\
-  & \jdeq
-\subst{u}{\ctxwk{\subst{x}{P}}{\subst{x}{\tfid{A}}}} \\
-  & \jdeq
-\subst{x}{\tfid{A}} \\
-  & \jdeq
-x
-\end{align*}
-To show that $\subst{\tmext{x}{u}}{\cprojsndf{A}{P}}\jdeq u$, note that
-\begin{equation*}
-\subst{\tmext{x}{u}}{\cprojsndf{A}{P}}
-  \jdeq
-\subst{u}{{x}{\tfid{P}}}
-  \jdeq
-\subst{u}{\tfid{\subst{x}{P}}}
-  \jdeq
-u
-\end{equation*}
-Therefore, $\psi$ is a left-inverse of $\varphi$. To show that it is also a
-right inverse, note that \autoref{cor:pairing_ehom} gives us that
-\begin{equation*}
-\tmext{\cprojfst{A}{P}{w}}{\cprojsnd{A}{P}{w}}
-  \jdeq
-\subst{w}{(\tmext{\cprojfstf{A}{P}}{\cprojsndf{A}{P}})}.
-\end{equation*}
-Thus, it suffices to show that $\tmext{\cprojfstf{A}{P}}{\cprojsndf{A}{P}}\jdeq
-\tfid{\ctxext{A}{P}}$, which is straightforward.
-\end{proof}
-
-\begin{lem}\label{lem:tmext_assoc}
-For any $x\in T(A)$, $u\in T(S_x(P))$ and $v\in T(S_{\tmext{x}{u}}(Q))$ we have
-\begin{equation*}
-\tmext{{x}{u}}{v}\jdeq \tmext{x}{{u}{v}}\in T(\ctxext{{A}{P}}{Q}).
-\end{equation*}
-\begin{comment}
-Also, we have
-\begin{align*}
-\cprojfstf{A}{P}\circ\cprojfstf{\ctxext{A}{P}}{Q} & \jdeq\cprojfstf{A}{\ctxext{P}{Q}} \\
-\cprojsndf{A}{P}\circ\cprojfstf{\ctxext{A}{P}}{Q} & \jdeq \cprojfstf{P}{Q}\circ\cprojsndf{A}{\ctxext{P}{Q}} \\
-\cprojsndf{\ctxext{A}{P}}{Q} & \jdeq \cprojsndf{P}{Q}\circ\cprojsndf{A}{\ctxext{P}{Q}}.
-\end{align*}
-\end{comment}
-\end{lem}
-
-\begin{proof}
-By \autoref{subst_by_tmext}, we have $S_v\circ S_{\tmext{x}{u}}\jdeq S_v\circ S_u\circ S_x \jdeq S_{\tmext{u}{v}}\circ S_x$,
-so associativity of term extension follows.
-\end{proof}
-
-\begin{comment}
-\begin{lem}
-Let $f\in\thom{A}{\ctxext{B}{Q}}$ Then we have
-\begin{equation*}
-f \jdeq \tmext{\jcomp{A}{f}{\cprojfstf{B}{Q}}}{\jcomp{A}{f}{\cprojsndf{B}{Q}}}.
-\end{equation*}
-Alternatively, when $f_0\in\thom{A}{B}$ and $f_1\in\thom{A}{\jcomp{A}{f}{Q}}$,
-then we have $\tmext{f_0}{f_1}\in\thom{A}{\ctxext{B}{Q}}$ and
-\begin{align*}
-\jcomp{A}{\tmext{f_0}{f_1}}{\cprojfstf{B}{Q}} & \jdeq f_0 \\
-\jcomp{A}{\tmext{f_0}{f_1}}{\cprojsndf{B}{Q}} & \jdeq f_1.
-\end{align*}
-\end{lem}
-
-\begin{proof}
-Straightforward
-\end{proof}
-\end{comment}
-
-We end this section by showing that weakening is the same operation as pre-composition
-with a first projection.
-
-\begin{thm}\label{precomp_by_proj}
-Precomposition with $\cprojfstf{A}{P}$ is weakening by $P$, i.e.
-\begin{equation*}
-(\cprojfstf{A}{P})_\ast \jdeq W_P.
-\end{equation*}
-\end{thm}
-
-\begin{proof}
-\begin{align*}
-(\cprojfstf{A}{P})_\ast & \jdeq S_{\cprojfstf{A}{P}}\circ W_{\ctxext{A}{P}} \\
-& \jdeq S_{W_P(\tfid{A})}\circ W_P\circ W_A \\
-& \jdeq W_P\circ S_{\tfid{A}}\circ W_A \\
-& \jdeq W_P.\qedhere 
-\end{align*}
-\end{proof}
-
-\subsection{The interchange laws}
-
-We are now in the position to define vertical and horizontal composition, and
-prove properties of them.
-
-\begin{defn}
-Let $\jhom{\Gamma}{A}{B}{f}$ and $\jfhom{\Gamma}{A}{B}{f}{P}{Q}{F}$. Then we
-define 
-\begin{equation*}
-\jhomdefn{\Gamma}{{A}{P}}{{B}{Q}}{\jvcomp{P}{f}{F}}{\unfold{\jvcomp{P}{f}{F}}}.
-\end{equation*}
-Whenever we say that we have a diagram of the form
-\begin{equation*}
-\begin{tikzcd}
-R \arrow[r,"f_2"] \arrow[d,fib] &
-S \arrow[d,fib] \\
-P \arrow[r,"f_1"] \arrow[d,fib] &
-Q \arrow[d,fib] \\
-A \arrow[r,"f_0"] &
-B
-\end{tikzcd}
-\end{equation*}
-we mean that we have
-$f_0\in\thom{A}{B}$, $f_1\in\thomd{f_0}{P}{Q}$ and 
-$f_2\in\thomd{\jvcomp{P}{f_0}{f_1}}{R}{S}$.
-\end{defn}
-
-\begin{cor}
-Let $\jhom{\Gamma}{A}{B}{f}$ and $\jfhom{\Gamma}{A}{B}{f}{P}{Q}{F}$. Then we
-have
-\begin{equation*}
-H(\jvcomp{P}{f}{F})\jdeq \jvcomp{H(P)}{H(f)}{H(F)}
-\end{equation*}
-for any E-homomorphism $H:\mathbb{E}\to\mathbb{D}$.
-\end{cor}
-
-\begin{lem}
-Vertical composition is associative.
-\end{lem}
-
-\begin{proof}
-Consider the diagram
-\begin{equation*}
-\begin{tikzcd}
-R \arrow[r,"f_2"] \arrow[d,fib] &
-S \arrow[d,fib] \\
-P \arrow[r,"f_1"] \arrow[d,fib] &
-Q \arrow[d,fib] \\
-A \arrow[r,"f_0"] &
-B
-\end{tikzcd}
-\end{equation*}
-in context $\Gamma$.
-Because weakening distributes over term extension, and term extension is
-associative, we have
-\begin{align*}
-\jvcomp{R}{\jvcomp{P}{f_0}{f_1}}{f_2}
-  & \jdeq
-\tmext{\ctxwk{R}{\tmext{\ctxwk{P}{f_0}}{f_1}}}{f_2}
-  \\
-  & \jdeq
-\tmext{\tmext{\ctxwk{R}{{P}{f_0}}}{\ctxwk{R}{f_1}}}{f_2}
-  \\
-  & \jdeq
-\tmext{\ctxwk{\ctxext{P}{R}}{f_0}}{\tmext{\ctxwk{R}{f_1}}{f_2}}
-  \tag{By \autoref{lem:tmext_assoc}}
-  \\
-  & \jdeq
-\jvcomp{{P}{R}}{f_0}{\jvcomp{R}{f_1}{f_2}}.\qedhere
-\end{align*}
-\end{proof}
-
-\begin{defn}
-Let $\jhom{\Gamma}{A}{B}{f}$ and $\jfhom{\Gamma}{A}{B}{f}{P}{Q}{F}$. Then we
-define the E-homomorphism 
-\begin{equation*}
-F_\circledast\defeq F_\ast\circ (f_\ast/Q):
-\mathbb{E}/\ctxext{{\Gamma}{B}}{Q}\to\mathbb{E}/\ctxext{{\Gamma}{A}}{P}.
-\end{equation*}
-The infix notation of $F_\circledast$ is taken to be $\jfcomp{\blank}{\blank}{\blank}{F}{\blank}$.
-\end{defn}
-
-\begin{lem}\label{lem:three-composition}
-Let $\jhom{\Gamma}{A}{B}{f}$ and $\jfhom{\Gamma}{A}{B}{f}{P}{Q}{F}$. Then we
-have the equality
-\begin{equation*}
-F_\circledast\jdeq (\jvcomp{P}{f}{F})_\ast.
-\end{equation*}
-\end{lem}
-
-\begin{proof}
-\begin{align*}
-F_\ast\circ (f_\ast/Q)
-  & \jdeq
-S_F\circ W_P\circ S_f/(W_A(Q))\circ W_A/\ctxext{B}{Q}
-  \\
-  & \jdeq
-S_F\circ S_{W_P(f)}/W_P(W_A(Q))\circ W_P/W_A(\ctxext{B}{Q})\circ W_A/\ctxext{B}{Q}
-  \\
-  & \jdeq
-S_F\circ S_{W_P(f)}/W_P(W_A(Q))\circ W_{\ctxext{A}{P}}/\ctxext{B}{Q}
-  \\
-  & \jdeq
-S_{\tmext{W_P(f)}{F}}\circ W_{\ctxext{A}{P}}/\ctxext{B}{Q}
-  \tag{By \autoref{subst_by_tmext}}
-  \\
-  & \jdeq
-(\jvcomp{P}{f}{F})_\ast.\qedhere
-\end{align*}
-\end{proof}
-
-We are now in position to state and prove the interchange law of horizontal
-and vertical composition.
-
-\begin{thm}\label{thm:interchange}
-Consider the diagram
-\begin{equation*}
-\begin{tikzcd}
-P \arrow[r,"F"] \arrow[d,fib] &
-Q \arrow[r,"G"] \arrow[d,fib] &
-R \arrow[d,fib] \\
-A \arrow[r,"f"] &
-B \arrow[r,"g"] &
-C
-\end{tikzcd}
-\end{equation*}
-in context $\Gamma$. Then the equality
-\begin{equation*}
-{\jcomp{{A}{P}}{\jvcomp{P}{f}{F}}{\jvcomp{Q}{g}{G}}}
-  \jdeq
-{\jvcomp{P}{\jcomp{A}{f}{g}}{\jfcomp{A}{f}{P}{F}{G}}}
-\end{equation*}
-of morphisms from $\ctxext{A}{P}$ to $\ctxext{C}{R}$ in context $\Gamma$ holds.
-\end{thm}
-
-\begin{proof}
-By \autoref{lem:three-composition}, we have
-\begin{align*}
-\jcomp{{A}{P}}{\jvcomp{P}{f}{F}}{\jvcomp{Q}{g}{G}}
-  & \jdeq
-\jcomp{P}{F}{\jcomp{A}{f}{\tmext{\ctxwk{Q}{g}}{G}}}
-  \tag{By \autoref{lem:three-composition}}
-  \\
-  & \jdeq
-\jcomp{P}{F}{\tmext{\ctxwk{\jcomp{A}{f}{Q}}{\jcomp{A}{f}{g}}}{\jcomp{A}{f}{G}}}
-  \\
-  & \jdeq
-\tmext{\jcomp{P}{F}{\ctxwk{\jcomp{A}{f}{Q}}{\jcomp{A}{f}{g}}}}{\jcomp{P}{F}{{A}{f}{G}}}
-  \\
-  & \jdeq
-\tmext{\jcomp{P}{F}{\ctxwk{\jcomp{A}{f}{Q}}{\jcomp{A}{f}{g}}}}{\jfcomp{A}{f}{P}{F}{G}}
-  \\
-  & \jdeq
-\tmext{\ctxwk{P}{\jcomp{A}{f}{g}}}{\jfcomp{A}{f}{P}{F}{G}}
-  \tag{By \autoref{lem:compW_W}}
-  \\
-  & \jdeq
-\jvcomp{P}{\jcomp{A}{f}{g}}{\jfcomp{A}{f}{P}{F}{G}}.\qedhere
-\end{align*}
-\end{proof}
-
-\begin{thm}
-Consider the diagram
-\begin{equation*}
-\begin{tikzcd}
-P \arrow[r,"F"] \arrow[d,fib] &
-Q \arrow[r,"G"] \arrow[d,fib] &
-R \arrow[d,fib] \\
-A \arrow[r,"f"] &
-B \arrow[r,"g"] &
-C
-\end{tikzcd}
-\end{equation*}
-in context $\Gamma$. 
-Then $F_\circledast\circ G_\circledast\jdeq (\jfcomp{A}{f}{P}{F}{G})_\circledast$.
-In other words the composition $\jfcomp{}{}{}{\blank}{\blank}$ is associative.
-\end{thm}
-
-\begin{proof}
-\begin{align*}
-F_\circledast\circ G_\circledast
-  & \jdeq
-(\jvcomp{P}{f}{F})_\ast\circ(\jvcomp{Q}{g}{G})_\ast 
-  \tag{By \autoref{lem:three-composition}}\\
-  & \jdeq
-(\jcomp{\ctxext{A}{P}}{\jvcomp{P}{f}{F}}{\jvcomp{Q}{g}{G}})_\ast 
-  \tag{By \autoref{lem:compcomp}}\\
-  & \jdeq
-(\jvcomp{}{\jcomp{A}{f}{g}}{\jfcomp{A}{f}{P}{F}{G}})_\ast 
-  \tag{By \autoref{thm:interchange}} \\
-  & \jdeq
-(\jfcomp{A}{f}{P}{F}{G})_\circledast.
-  \tag{By \autoref{lem:three-composition}}
-\end{align*}
-\end{proof}
-
-\begin{lem}
-Let $\jhom{\Gamma}{A}{B}{f}$ and $\jfhom{\Gamma}{A}{B}{f}{P}{Q}{F}$. Then
-$\jvcomp{P}{f}{F}$ is the unique morphism from $\ctxext{A}{P}$ to $\ctxext{B}{Q}$
-with the property that both the diagram
-\begin{equation*}
-\begin{tikzcd}
-\ctxext{A}{P}
-  \ar{r}{\jvcomp{P}{f}{F}}
-  \ar{d}[swap]{\cprojfstf{A}{P}}
-& \ctxext{B}{Q}
-  \ar{d}{\cprojfstf{B}{Q}}
-  \\
-A \ar{r}[swap]{f}
-& B
-\end{tikzcd}
-\end{equation*}
-commutes, and $\jcomp{}{\jvcomp{P}{f}{F}}{\cprojsndf{B}{Q}}\jdeq F$.
-\end{lem}
-
-\begin{proof}
-We first note that
-\begin{align*}
-\jcomp{}{\jvcomp{P}{f}{F}}{\cprojfstf{B}{Q}}
-  & \jdeq
-\jcomp{P}{F}{\jcomp{A}{f}{\cprojfstf{B}{Q}}} 
-  \tag{By \autoref{lem:three-composition}} \\
-  & \jdeq
-\jcomp{P}{F}{\ctxwk{\jcomp{A}{f}{Q}}{\jcomp{A}{f}{\idtm{B}}}} \\
-  & \jdeq
-\jcomp{P}{F}{\ctxwk{\jcomp{A}{f}{Q}}{f}} \\
-  & \jdeq
-\ctxwk{P} f \tag{By \autoref{lem:compW_W}} \\
-  & \jdeq
-\jcomp{\ctxext{A}{P}}{\cprojfstf{A}{P}}{f}.
-  \tag{By \autoref{precomp_by_proj}}
-\end{align*}
-Also, we have
-\begin{align*}
-\jcomp{}{\jvcomp{P}{f}{F}}{\cprojsndf{B}{Q}}
-  & \jdeq
-\jcomp{P}{F}{\jcomp{A}{f}{\idtm{Q}}}
-  \tag{By \autoref{lem:three-composition}}
-  \\
-  & \jdeq
-\jcomp{P}{F}{\idtm{\jcomp{A}{f}{Q}}}
-  \\
-  & \jdeq
-F.
-\end{align*}
-Thus, we conclude that $\jvcomp{P}{f}{F}$ has indeed the stated property. For
-the uniqueness, let $G:\ctxext{A}{P}\to\ctxext{B}{Q}$ be a morphism such that
-$\jcomp{}{G}{\cprojfstf{B}{Q}}\jdeq\jcomp{}{\cprojfstf{A}{P}}{f}$ and
-$\jcomp{}{G}{\cprojsndf{B}{Q}}\jdeq F$. Then it follows that
-\begin{equation*}
-G \jdeq \tmext{\jcomp{}{\cprojfstf{A}{P}}{f}}{F}
-  \jdeq \tmext{\ctxwk{P}{f}}{F}
-  \jdeq \jvcomp{P}{f}{F}.\qedhere
-\end{equation*}
-\end{proof}
-
-\subsection{Categories internal to E-systems}
-The definition of internal category we're about to give, has a type of objects
-and a type of morphisms, such that the codomain function is a projection. It might
-be more appropriate to call such structures `right categories', and the corresponding
-structures where the domain function is a projection might be called `left
-categories'. In particular, the opposite of a right category will be a left
-category. There is also a symmetric notion of a category.
-
-\begin{defn}
-A category-framework $\cftalg{A}$ in context $\Gamma$ in an E-system $\mathbb{E}$ consists of
-\begin{enumerate}
-\item A family $\cftalgc{\cftalg{A}}\in\cat{C}/\Gamma$, representing the type of
-objects of the category framework;
-\item A family $\cftalgf{\cftalg{A}}\in\cat{C}/\ctxext{\Gamma}{\cftalgc{\cftalg{A}}}$,
-representing the type of morphisms of the category framework, where a term of
-$\subst{X}{\cftalgf{\cftalg{A}}}$ is considered to be a morphism into $X\in T(\cftalgc{\cftalg{A}})$;
-\item A morphism $\jhom{\Gamma}{\ctxext{\cftalgc{\cftalg{A}}}{\cftalgf{\cftalg{A}}}}{\cftalgc{\cftalg{A}}}{\cftctxext}$, representing the domain function;
-\item A morphism $\jhom{{\Gamma}{\cftalgc{\cftalg{A}}}}{{\cftalgf{\cftalg{A}}}{\jcomp{}{\cftctxext}{\cftalgf{\cftalg{A}}}}}{\cftalgf{\cftalg{A}}}{\cftfamext}$, representing composition;
-\item A term $\jterm{{\Gamma}{\cftalgc{\cftalg{A}}}}{\cftalgf{\cftalg{A}}}{\cftempf{\cftalg{A}}}$,
-representing the identity morphisms.
-\end{enumerate}
-\end{defn}
-
-\begin{rmk}
-Instead of looking at $\cftfamext$ as a morphism from 
-$\ctxext{\cftalgf{\cftalg{A}}}{\jcomp{}{\cftctxext}{\cftalgf{\cftalg{A}}}}$
-to $\cftalgf{\cftalg{A}}$ in context $\ctxext{\Gamma}{\cftalgc{\cftalg{A}}}$, 
-one could also look at $\cftfamext$ as a 
-morphism \emph{over $\cprojfstf{\cftalgc{\cftalg{A}}}{\cftalgf{\cftalg{A}}}$} 
-in context $\Gamma$, as indicated in the following diagram:
-\begin{equation*}
-\begin{tikzcd}[column sep=large]
-\cftalgf{\cftalg{A}}
-  \ar[fib]{d}
-& \jcomp{}{\cftctxext}{\cftalgf{\cftalg{A}}}
-  \ar[fib]{d}
-  \ar{l}[swap]{\cftfamext}
-  \ar{r}
-& \cftalgf{\cftalg{A}}
-  \ar[fib]{d}
-  \\
-\cftalgc{\cftalg{A}}
-& \ctxext{A}{\cftalgf{\cftalg{A}}}
-  \ar{l}{\cprojfstf{A}{\cftalgf{\cftalg{A}}}}
-  \ar{r}[swap]{\cftctxext}
-& \cftalgc{\cftalg{A}}
-\end{tikzcd}
-\end{equation*}
-From this point of view, we see that $\cftfamext$ has an object $X$ and a 
-morphism $f$ into $X$ as parameters, 
-then takes a morphism into the domain of $f$ as an argument, 
-and outputs a morphism into $X$. The projection maps $(X,f)$ to $X$, so we see
-that composition is indeed an operation of this kind.
-\end{rmk}
-
-A category-framework is not a category unless several axioms of the theory of
-categories are satisfied. The domain of a composed arrow $g\circ f$ must be the
-domain of $f$, composition must be associative, the domains of the identity
-morphisms must be the same as their codomains, and identities must satisfy
-the left and right identity laws. However, the axioms expressing these 
-conditions are not very insightful when they're written down directly. Thus, we
-take the usual way and define the notion of functor before we describe the
-actual conditions of being a category.
-
-\begin{defn}
-Let $\cftalg{A}$ and $\cftalg{B}$ be category-frameworks in context $\Gamma$,
-in an E-system $\mathbb{E}$.
-An \emph{internal functor $\cfthom{F}$ from $\extalg{A}$ to
-$\extalg{B}$ in context $\Gamma$} is a pair 
-$(\cfthomc{\cfthom{F}},\cfthomf{\cfthom{F}})$ consisting of
-\begin{align*}
-\jalign\jhom
-  {\Gamma}
-  {\cftalgc{\cftalg{A}}}
-  {\cftalgc{\cftalg{B}}}
-  {\cfthomc{\cfthom{F}}}
-  \\
-\jalign\jfhom
-  {\Gamma}
-  {\cftalgc{\cftalg{A}}}
-  {\cftalgc{\cftalg{B}}}
-  {\cfthomc{\cfthom{F}}}
-  {\cftalgf{\cftalg{A}}}
-  {\cftalgf{\cftalg{B}}}
-  {\cfthomf{\cfthom{F}}}
-\end{align*}
-for which the diagrams
-\begin{equation}\label{eq:exthom1}
-\begin{tikzcd}
-\ctxext{\cftalgc{\cftalg{A}}}{\cftalgf{\cftalg{A}}}
-  \ar{r}{\jvcomp{}{\cfthomc{\cfthom{F}}}{\cfthomf{\cfthom{F}}}}
-  \ar{d}[swap]{\cftctxext[\cftalg{A}]}
-& \ctxext{\cftalgc{\cftalg{B}}}{\cftalgf{\cftalg{B}}}
-  \ar{d}{\cftctxext[\cftalg{B}]}
-  \\
-\cftalgc{\cftalg{A}}
-  \ar{r}[swap]{\cfthomc{\cfthom{F}}}
-& \cftalgc{\cftalg{B}}
-\end{tikzcd}
-\end{equation}
-and
-\begin{equation}\label{eq:exthom2}
-\begin{tikzcd}[column sep=huge]
-\ctxext{\cftalgf{\cftalg{A}}}{\jcomp{}{\cftctxext[\cftalg{A}]}{\cftalgf{\cftalg{A}}}}
-  \ar{r}{\jvcomp{}{\cfthomf{\cfthom{F}}}{\jcomp{}{\cftctxext[\cftalg{A}]}{\cfthomf{\cfthom{F}}}}}
-  \ar{d}[swap]{\cftfamext[\cftalg{A}]}
-& \jcomp{}{\cfthomc{\cfthom{F}}}{\ctxext{\cftalgf{\cftalg{B}}}{\jcomp{}{\cftctxext[\cftalg{B}]}{\cftalgf{\cftalg{B}}}}}
-  \ar{d}{\jcomp{}{\cfthomc{\cfthom{F}}}{\cftfamext[\cftalg{B}]}}
-  \\
-\cftalgf{\cftalg{A}}
-  \ar{r}[swap]{\cfthomf{\cfthom{F}}}
-& \jcomp{}{\cfthomc{\cfthom{F}}}{\cftalgf{\cftalg{B}}}
-\end{tikzcd}
-\end{equation}
-commute, and for which $\subst{\cftempf{\cftalg{A}}}{\cfthomf{\cfthom{F}}}
-\jdeq\jcomp{}{\cfthomc{\cfthom{F}}}{\cftempf{\cftalg{B}}}$.
-\end{defn}
-
-\begin{defn}
-Let $\cftalg{A}$ be a category-framework. We define the category-framework
-\begin{equation*}
-\cftfamalg{\cftalg{A}}\defeq(\cftalgf{\cftalg{A}},\jcomp{}{\cftctxext}%
-{\cftalgf{\cftalg{A}}},\cftfamext,\jcomp{}{\cftctxext}{\cftfamext},
-\jcomp{}{\cftctxext}{\cftempf{\cftalg{A}}}),
-\end{equation*}
-When $\cftalg{P}$ is a category-framework in context $\ctxext{\Gamma}{A}$, then
-and we define the category-framework
-\begin{equation*}
-\ctxext{\cftalgc{\cftalg{A}}}{\cftalg{P}}
-  \defeq
-( \ctxext{A}{\cftalgc{\cftalg{P}}},
-  \cftalgf{\cftalg{P}},
-  \jvcomp{}{\idtm{A}}{\cftctxext[\cftalg{P}]},
-  \cftfamext[\cftalg{P}],
-  \cftempf{\cftalg{P}}).
-\end{equation*}
-\end{defn}
-
-\begin{defn}
-An internal category $\cftalg{A}$ in an E-system $\mathbb{E}$ consists of a
-category-framework $\cftalg{A}$ satisfying the following conditions:
-\begin{enumerate}
-\item \emph{The domain condition.} The pair 
-$(\cftctxext,\idtm{\jcomp{}{\cftctxext}{\cftalgf{\cftalg{A}}}})$ is a
-functor from $\ctxext{\cftalgc{\cftalg{A}}}{\cftfamalg{\cftalg{A}}}$ to
-$\cftalg{A}$.
-\item \emph{The associativity law.} The pair
-$(\cftfamext,\idtm{\jcomp{}{\cftfamext}{\cftalgf{\cftfamalg{\cftalg{A}}}}})$
-is a functor from $\ctxext{\cftalgf{\cftalg{A}}}{\cftfamalg{\cftfamalg{\cftalg{A}}}}$ to
-$\cftfamalg{\cftalg{A}}$.
-\item \emph{The identity domain condition.} The domain of the identity morphism
-is the same as its codomain:
-\begin{equation*}
-\jtermeq
-  {{\Gamma}{\cftalgc{\cftalg{A}}}}
-  {\ctxwk{\cftalgc{\cftalg{A}}}{\cftalgc{\cftalg{A}}}}
-  {\subst{\cftempf{\cftalg{A}}}{\cftctxext[\cftalg{A}]}}
-  {\idtm{\cftalgc{\cftalg{A}}}}.
-\end{equation*}
-\item \emph{The left identity law.} A morphism precomposed by the identity morphism
-is equal to the original morphism:
-\begin{equation*}
-\jtermeq
-  {{{\Gamma}{\cftalgc{\cftalg{A}}}}{\cftalgf{\cftalg{A}}}}
-  {\ctxwk{\cftalgf{\cftalg{A}}}{\cftalgf{\cftalg{A}}}}
-  {\subst{\cftempf{\cftalg{A}}}{\cftfamext}}
-  {\idtm{\cftalgf{\cftalg{A}}}}.
-\end{equation*}
-\item \emph{The right identity law.} A morphism postcomposed by the identity morphism
-is equal to the original morphism:
-\begin{equation*}
-\jtermeq
-  {{{\Gamma}{\cftalgc{\cftalg{A}}}}{\cftalgf{\cftalg{A}}}}
-  {\ctxwk{\cftalgf{\cftalg{A}}}{\cftalgf{\cftalg{A}}}}
-  {\subst{\jcomp{}{\cftctxext}{\cftempf{\cftalg{A}}}}{\cftfamext}}
-  {\idtm{\cftalgf{\cftalg{A}}}}.
-\end{equation*}
-\end{enumerate}
-\end{defn}
-
-\begin{rmk}
-Since we have only used operations of E-systems to define internal categories,
-it immediately follows that if $\cftalg{A}$ is an internal category in
-$\mathbb{E}$ and if $F:\mathbb{E}\to\mathbb{D}$ is an E-homomorphism, then
-$F(\cftalg{A})$ is an internal category in $\mathbb{D}$. In particular, if
-$\cftalg{Q}$ is an internal category in context $\ctxext{\Gamma}{B}$ and
-$f:A\to B$, then the change of base $f_\ast(\cftalg{Q})$ is an internal category
-in context $\ctxext{\Gamma}{A}$. Thus, it is immediate that the notion of
-internal category is stable under change of base.
-\end{rmk}
-
-\begin{comment}
-\section{Dependent morphisms}
-
-\begin{defn}
-Let $f_0\in\thom{A}{B}$ be an internal morphism and let $P\in\cat{C}/\ctxext{\Gamma}{A}$ and
-$Q\in\cat{C}/\ctxext{\Gamma}{B}$. We define
-\begin{equation*}
-\thomd{f_0}{P}{Q} \defeq \thom{P}{{f_0}_\ast(Q)}
-\end{equation*}
-We indicate that $f_1\in\thomd{f_0}{P}{Q}$ by drawing the diagram
-\begin{equation*}
-\begin{tikzcd}
-P \arrow[r,"f_1"] \arrow[d,->>] & Q \arrow[d,->>] \\
-A \arrow[r,swap,"f_0"] & B.
-\end{tikzcd}
-\end{equation*}
-\end{defn}
-
-\begin{defn}
-Consider the diagram
-\begin{equation*}
-\begin{tikzcd}
-P \arrow[r,"f_1"] \arrow[d,->>] & Q \arrow[d,->>] \arrow[r,"g_1"] & R \arrow[d,->>] \\
-A \arrow[r,swap,"f_0"] & B \arrow[r,"g_0"] & C.
-\end{tikzcd}
-\end{equation*}
-We define
-\begin{align*}
-\jvcomp{P}{f_0}{f_1} & \defeq\unfold\jvcomp{P}{f_0}{f_1} & & \in \thom{\ctxext{A}{P}}{\ctxext{B}{Q}} \\
-\jfcomp{A}{f_0}{P}{f_1}{g_1} & \defeq \unfold\jfcomp{A}{f_0}{P}{f_1}{g_1} & & \in \thomd{\jcomp{A}{f_0}{g_0}}{P}{R}.
-\end{align*}
-\end{defn}
-
-\begin{rmk}
-Consider
-\begin{equation*}
-\begin{tikzcd}
-P \arrow[r,"f_1"] \arrow[d,->>] & Q \arrow[d,->>] \\
-A \arrow[r,swap,"f_0"] & B.
-\end{tikzcd}
-\end{equation*}
-Then the diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=large]
-\ctxext{A}{P} \arrow[r,"\jvcomp{P}{f_0}{f_1}"] \arrow[d,swap,"\cprojfstf{A}{P}"] & \ctxext{B}{Q} \arrow[d,"\cprojfstf{B}{Q}"] \\
-A \arrow[r,swap,"f_0"] & B
-\end{tikzcd}
-\end{equation*}
-commutes. Also, we have
-\begin{equation*}
-S_{\tmext{x}{u}}(\jvcomp{P}{f_0}{f_1}) \jdeq \tmext{S_x(f_0)}{S_{\tmext{x}{u}}(f_1)}
-\end{equation*}
-for each $x\in T(A)$ and $u\in T(S_x(P))$.
-\end{rmk}
-
-\section{Natural transformations of E-homomorphisms}
-\begin{rmk}
-When $\mathcal{E}$ is an E-system with empty context, then we have the category
-structure on the contexts given by $\thom{\Gamma}{\Delta}= T(W_\Gamma(\Delta))$.
-\end{rmk}
-
-\begin{defn}
-A \emph{natural transformation} of E-homomorphisms from $F:\mathcal{E}\to\mathcal{D}$
-to $G:\mathcal{E}\to\mathcal{D}$ preserving the empty context consists of
-\begin{enumerate}
-\item A t-morphism $\tau0_A\in\thom{F(\Gamma)}{G(\Gamma)}$ for every $\Gamma\in\mathcal{E}$.
-\item A commuting square square
-\begin{equation*}
-\begin{tikzcd}
-F(A) \arrow[r,"\tau1_A"] \arrow[d,twoheadrightarrow] & G(A) \arrow[d,twoheadrightarrow] \\
-F(\Gamma) \arrow[r,swap,"\tau0_\Gamma"] & G(\Gamma)
-\end{tikzcd}
-\end{equation*}
-for every family $P$ over $A$ of solid arrows, such that the diagram
-\begin{equation*}
-\begin{tikzcd}
-F(A) \arrow[r,"\tau1_A"] & G(A) \\
-F(\Gamma) \arrow[r,swap,"\tau0_\Gamma"] \arrow[densely dotted,u,"F(x)"] & G(\Gamma) \arrow[densely dotted,u,swap,"G(x)"]
-\end{tikzcd}
-\end{equation*}
-commutes for every $x\in T(A)$.
-\end{enumerate}
-such that
-\begin{enumerate}
-\item the square
-\begin{equation*}
-\begin{tikzcd}[column sep=huge]
-W_{F(A)}(F(B)) \arrow[r,"W_{F(A)}(\tau1_B)"] \arrow[d,->>] & W_{G(A)}(G(B)) \arrow[d,->>] \\
-F(A) \arrow[r,"\tau1_A"] & G(A)
-\end{tikzcd}
-\end{equation*}
-commutes for every $A,B\in\mathcal{E}/\Gamma$. Note that 
-\begin{align*}
-W_{F(A)}(\tau1_B)
-  & \in 
-T(W_{F(A)}(W_{F(B)}(G(B))))
-  \\
-  & =
-T(W_{W_{F(A)}(F(B))}(W_{F(A)}(G(B))))
-\end{align*}
-\end{enumerate}
-\end{defn}
-
-\section{Indexed E-systems}
-\begin{defn}
-Let $\cat{A}$ be a category. An $\cat{A}$-indexed E-system is an internal
-E-system in the category $Psh(\cat{A})$.
-\end{defn}
-
-\section{Examples of adjunctions}
-
-\begin{lem}
-For each $A\in\mathcal{E}/\Gamma$, $W_A$ is left adjoint to $\Pi_A$. 
-\end{lem}
-
-\begin{proof}
-The unit $\eta:1_{\cat{C}/\ctxext{\Gamma}{A}}\Rightarrow W_A\circ\Pi_A$ is given by
-\begin{equation*}
-\eta_P:\lam{f}{x} ev(\lambda(f),x)
-\end{equation*}
-$\eta_P\in\thom{P}{W_A(\Pi_A(P))}$
-\end{proof}
-
-\begin{lem}
-Suppose $\cat{D}$ is a retract of a pre-category $\cat{C}$, i.e. there are functors
-$S:\cat{D}\to\cat{C}$ and $R:\cat{C}\to\cat{D}$ with a natural isomorphism
-$\tau:R\circ S \simeq \catid{\cat{D}}$, and suppose that $\cat{C}$ is a weakening
-category. Then $\cat{D}$ is a weakening system too.
-\end{lem}
-
-\begin{proof}
-Let $f:\Delta\to \Gamma$ be a morphism in $\cat{D}$. Then we define $W_f$ to be the
-composition
-\begin{equation*}
-\begin{tikzcd}[column sep=3.2em]
-\cat{D}/\Gamma
-  \arrow[r,"S/\Gamma"]
-  &
-\cat{C}/S(\Gamma)
-  \arrow[r,"W_{S(f)}"]
-  &
-\cat{C}/S(\Delta)
-  \arrow[r,"R/S(\Delta)"]
-  &
-\cat{D}/R(S(\Delta))
-  \arrow[r,"\tau_\Delta\circ{-}"]
-  &
-\cat{D}/\Delta
-\end{tikzcd}
-\end{equation*}
-We first verify whether this assembles a pre-weakening structure on $\cat{D}$.
-\begin{enumerate}
-\item Let $\Delta$ be an object of $\cat{D}$. Then $W_{\catid{\Delta}}$ is the composition
-\begin{equation*}
-\begin{tikzcd}[column sep=3.2em]
-\cat{D}/\Delta
-  \arrow[r,"S/\Delta"]
-  &
-\cat{C}/S(\Delta)
-  \arrow[r,"R/S(\Delta)"]
-  &
-\cat{D}/R(S(\Delta))
-  \arrow[r,"{-}\circ\tau_\Delta"]
-  &
-\cat{D}/\Delta
-\end{tikzcd}
-\end{equation*}
-This functor maps $g:\Delta'\to \Delta$ consecutively to $S(g)$, to $R(S(g))$, to
-$\tau_\Delta\circ R(S(g))$. Note that $\tau_\Delta\circ R(S(g))=g\circ \tau_{\Delta'}$ by
-naturality.
-\end{enumerate}
-\end{proof}
-
-\section{Weak E-systems}
-
-\begin{rmk}
-Note that $W$ is a strict $2$-functor frm $\cat{C}$ to $\mathbb{Cat}$ of which
-the value at an object $\Gamma$ is specified to be $\cat{C}/\Gamma$, and which
-is required to preserve the final object. We could have considered $W$ to be a 
-pseudo-functor from $\cat{C}$ to $\mathbb{Cat}$ instead. A coherence theorem
-by Power \cite{Power89} then allows for the strictification of this pseudo-functor
-into a (strict) weakening structure. (Does this map $\Gamma$ to $\cat{C}/\Gamma$
-and final to final?). Thus, we know what a weak pre-weakening system has to be.
-The following questions need to be answered:
-\begin{itemize}
-\item What is a weak weakening/projection/substitution system.
-\item What is a weak E-system? I.e.~what are the coherence rules for something
-to be a weakening/projection/substitution homomorphism?
-\item Can they all be strictified? Are there more examples of weak E-systems?
-\end{itemize}
-\end{rmk}
-\end{comment}
-
-\section{E-systems compared to other systems of type dependency}\label{sec:esys_compared}
-In an effort to formalize the simplicial set model of Martin-L\"of type theory
-with the univalence axiom, Voevodsky has studied various notions of models of
-type theory, taking Cartmell's \emph{contextual categories} as a starting
-point. Voevodsky's definition of a \emph{C-system} is equivalent to that of Cartmell
-\cite{VV_C-systems_quotients}. There are several techniques of defining 
-C-systems: in \cite{VV_Csys_univ}, C-systems are constructed out of a universe
-in a category; in \cite{VV_C-systems_quotients}, sub-C-systems and regular
-quotients of C-systems are defined; in \cite{VV_C-systems_monad}, a C-system
-is constructed out of a monad $R$ on the category of sets and a left $R$-module
-with values in $\mathbf{Set}$.
-
-In \cite{VV_B-systems}, Voevodsky introduced his notion of \emph{B-systems}, which is
-an essentially algebraic theory of type dependency with infinitely many sorts 
-and operations, and subsequently he showed that there is a full and faithful
-functor from the category of C-systems to the category of B0-systems, of which
-the image is contained in the subcategory of B-systems. This fact is important at least
-for a theoretical reason: it tells us that the conditions of being a B-system
-are necessary to axiomatize type dependency. In other words,
-this is a theorem that tells us that we need no further axioms. It is
-unknown whether the categories of C-systems and B-systems are actually equivalent.
-This question would settle the sufficiency of the conditions of being a B-system.
-
-The category of B-systems will be shown to be equivalent to the category of
-stratified E-systems. However, we do not arrive at the fact that the functor
-$\mathcal{E}:\mathbf{Bsys}\to\mathbf{Esys}$ is full without imposing an additional
-condition on the E-homomorphisms. 
-To ensure that the category of stratified E-systems with stratified 
-E-homomorphisms between them is a full subcategory of the category 
-$\mathbf{Esys}$, we will require that the underlying functor of each E-homomorphism
-lifts factorizations. It is
-important to note that this condition is also algebraic, so that the category
-of E-systems remains algebraic.
-
-\subsection{Lifting factorizations}
-Recall
-that for any morphism $f:X\to Y$ in a category $\cat{C}$ there is a category
-$\mathbf{fact}(f)$ of factorizations of $f$. Also, any functor $F:\cat{C}\to\cat{D}$
-determines a functor $F_{\mathbf{fact}(f)}:\mathbf{fact}(f)\to\mathbf{fact}(H(f))$
-for any morphism $f$ of $\cat{C}$.
-
-\begin{defn}
-A functor $F:\cat{C}\to\cat{D}$ is said to
-\emph{lift factorizations} if for any $f\in\cat{C}/X$, the functor
-$F_{\mathbf{fact}(f)}:\mathbf{fact}(f)\to\mathbf{fact}(F(f))$ is an isomorphism
-of categories. 
-\end{defn}
-
-\begin{comment}
-\begin{rmk}
-We may choose the property of lifting factorizations
-to involve either an equivalence or an isomorphism of categories. The
-isomorphism-version, would say that for every factorization $h'\circ g'$
-of $F(f)$ there is a unique factorization $f\jdeq h\circ g$ in $\cat{C}$ such
-that $F(g)\jdeq g'$ and $F(h)\jdeq h'$. In the version with equivalences,
-the uniqueness is replaced by uniqueness up to isomorphism, and the equalites
-are replaced by an isomorphism in $\mathbf{fact}(F(f))$.
-
-It is easier to state the isomorphism version of lifting factorizations with
-inference rules and it might be easier to explain this condition on type theoretical
-grounds, although the equivalence version has the advantage of being
-categorical (i.e.~invariant under equivalence of categories).
-
-In the present
-context, it doesn't matter very much which one we pick: the categories in which
-we're interested are all posets.
-\end{rmk}
-\end{comment}
-
-\begin{eg}
-For any slice category $\cat{C}/X$, the forgetful functor $\cat{C}/X\to\cat{C}$
-lifts factorizations.
-\end{eg}
-
-\begin{defn}
-We write $\mathbf{Esys}$ for the category of E-systems where the underlying
-functor of each E-homomorphism lifts factorizations.
-\end{defn}
-
-\subsection{Stratified E-systems}
-
-In this subsection we define the notion of stratified E-system. Then we show
-that the category $\mathbf{Esys_s}$ of stratified E-systems with homomorphisms preserving the
-stratification, is a full subcategory of $\mathbf{Esys}$.
-
-\begin{defn}
-A category $\cat{C}$ with terminal object is said to be \emph{stratified} if there exists a 
-\emph{stratification functor}
-\begin{equation*}
-L : \cat{C}\to (\mathbb{N},\geq)
-\end{equation*}
-such that
-\begin{enumerate}
-\item $L(X)\jdeq 0$ if and only if $X$ is terminal, and for any $f:X\to Y$ we have
-$L(X)\jdeq L(Y)$ if and only if $X\jdeq Y$ and $f\jdeq\catid{X}$. 
-\item every morphism $f:X\to Y$ in $\cat{C}_\mathbb{E}$, where $L(X)\jdeq
-n+m+1$ and $L(Y)\jdeq n$, has a unique factorization 
-\begin{equation*}
-\begin{tikzcd}
-X \arrow[r,"f_m"] & X_m \arrow[r,"f_{m-1}"] & \cdots \arrow[r,"f_1"] & X_1 \arrow[r,"f_0"] & Y
-\end{tikzcd}
-\end{equation*}
-where $L(X_i)\jdeq n+i$.
-\end{enumerate}
-A functor $F:\cat{C}\to\cat{D}$ between stratified categories is said to be stratified 
-if $L_{\cat{C}}\jdeq L_{\cat{D}}\circ F$. 
-\end{defn}
-
-%\begin{rmk}
-%The categorical version of being stratified, is that there exists a factofibration
-%$L:\cat{C}\to (N,\geq)$ which preserves the terminal object.
-%\end{rmk}
-
-\begin{lem}
-Any stratified category is a rooted tree.
-\end{lem}
-
-\begin{proof}
-Let $\cat{C}$ be a stratified category. The root of the tree is going to be the terminal
-object of $\cat{C}$. Note that objects of $\cat{C}$ have no automorphisms other
-than the identity morphisms. Therefore, to show that $\cat{C}$ is a tree,
-it suffices to show that for every object $X$ with $L(X)>0$ there is
-a unique object $\eft(X)$ with $L(\eft(X))\jdeq L(X)-1$, with a morphism $X\to \eft(X)$.
-The existence follows from the existence of the factorization of $X\to 1$. The 
-uniqueness follows from the observation that for any arrow
-$X\to Y$ with $L(Y)\jdeq L(X)-1$, we can factorize $X\to 1$ through $Y$. Since
-factorizations are unique, we will have $Y\jdeq \eft(X)$. 
-\end{proof}
-
-\begin{lem}
-The category of small stratified categories is a full subcategory of the category
-of small categories with terminal objects and between them functors which lift 
-factorizations and preserve the terminal object.
-\end{lem}
-
-\begin{proof}
-Let $F:\cat{C}\to\cat{D}$ be a functor between stratified categories which lifts
-factorizations and 
-which preserves terminal objects. We show by induction that $F$ is stratified.
-Since $F$ preserves the terminal object, $L(Y)\jdeq 0$ implies $L(F(Y))\jdeq 0$.
-Now suppose that for $n\in\mathbb{N}$, we have that $L(\Gamma)\jdeq n$ implies
-$L(F(Y))\jdeq n$, and let $X$ be such that $L(X)\jdeq n+1$ with
-$p_X:X\to\eft(X)$. Then the poset $\mathbf{fact}(p_X)$ is isomorphic to the
-poset $\mathbf{2}$ with two objects, one of which is smaller than the other. By the
-assumption that $F$ lifts factorizations, the poset $\mathbf{fact}(F(p_X))$
-is also isomorphic to $\mathbf{2}$. Thus, the only factorization
-of $F(X)\to\unit$ starts with $F(p_X):H(X)\to F(\eft(X))$, which implies that
-$L(F(X))\jdeq n+1$. 
-\end{proof}
-
-\begin{rmk}
-It is easy to show that a category can only be stratified in at most one way,
-and that if a category is stratified, then so are any of its slices. Therefore,
-it makes sense to ask for the property that the weakening and substitution
-functors in a pre-weakening or pre-substitution system are stratified.
-\end{rmk}
-
-\begin{defn}
-A pre-E-system is said to be \emph{stratified} if its underlying category is
-stratified and if each $W_A$ and $S_x$ is a stratified functor. The category of
-stratified pre-E-systems with stratified E-homomorphisms between them is denoted by
-$\mathbf{Esys_s}$.
-\end{defn}
-
-\begin{cor}\label{lem:strat_full}
-The category of stratified pre-E-systems is a full subcategory of the category of
-pre-E-systems.
-\end{cor}
-
-\subsection{The embedding of B0-systems into pre-E-systems}
-
-Voevodsky distinguishes between pre-B-systems, which is an essentially algebraic theory with
-only operations and no equalities; B0-systems, which is an extension of the
-essentially algebraic theory of pre-B-systems which ensures that all the operations behave
-well with two sets of the operations ($\eft$ and $\ebd$); and B-systems, which extends the 
-theory of B0-systems with equalities ensuring that the rest of the operations 
-are compatible with each other. Voevodsky, furthermore distinguishes
-non-unital and unital versions of these, and noted that not all B-homomorphisms
-between unital B-systems are unital. The same distinguishment could be made for 
-E-systems by forgetting the projection structure, but we have not been able to construct a
-functor from unital B0-systems to unital pre-E-systems. Therefore, we shall only
-consider non-unital B0-systems and unital B-systems. In this section we will
-recall the definition of non-unital B0-systems and construct a full and faithful
-functor from them to the category of pre-E-systems.
-
-One reason for introducing B-systems by first introducing pre-B-systems is that
-it is clear in the definition of pre-B-systems that there are only countably
-many operations. Since we are interested in embedding the B-systems into E-systems,
-we shall organize the definition of B-systems a bit differently, but of course
-the resulting definition of B0-systems and B-systems will be (trivially) 
-equivalent.
-
-\begin{defn}
-A \emph{B-framework} is a collection of data of the following form:
-\begin{enumerate}
-\item for all $n\in\mathbb{N}$ two sets $B_n$ and $\tilde{B}_n$. 
-\item for all $n\in\mathbb{N}$ maps of the form
-\begin{align*}
-\eft[n] & : B_{n+1}\to B_n \\
-\ebd[n] & : \tilde{B}_n\to B_n.
-\end{align*}
-For $m,n\in\mathbb{N}$, we denote the composition $\eft[n]\circ\cdots\circ\eft[n+m]:B_{n+m+1}\to B_n$ by $\eft[n]^m$. 
-\item $B_0$ is a singleton $\{\pt\}$.
-\end{enumerate} 
-\end{defn}
-
-\begin{rmk}
-Since the indices $n,m\in\mathbb{N}$ can be infered, we will usually omit them
-in the infix form of the operators.
-\end{rmk}
-
-\begin{defn}
-A homomorphism $H:\mathbb{B}\to\mathbb{A}$ of B-frameworks consists of maps
-$H_n:B_n\to A_n$ and $\tilde{H}_n:\tilde{B}_n\to\tilde{A}_n$ such that
-\begin{align*}
-\eft(H(X)) & \jdeq H(\eft(X)) \\
-\ebd(\tilde{H}(x)) & \jdeq H(\ebd(x))
-\end{align*}
-for any $X\in B_n$ and $x\in\tilde{B}_n$. The category of B-frameworks is
-denoted by $\mathbf{Bfr}$. 
-\end{defn}
-
-\begin{defn}
-For every B-framework $\mathbb{B}$ and any $X\in B_n$, there is a B-framework
-$\mathbb{B}/X$ given by
-\begin{align*}
-(B/X)_{m} & \jdeq \{Y\in B_{n+m}\mid\eft^{m}(Y)\jdeq X\}\\
-(\tilde{B}/X)_m & \jdeq \{y\in \tilde{B}_{n+m}\mid\eft^m(\ebd(y))\jdeq X\}.
-\end{align*}
-Also, for any homomorphism $H:\mathbb{B}\to\mathbb{A}$ of B-frameworks and any
-$X\in B_n$, there is a homomorphism $H/X:\mathbb{B}/X\to\mathbb{A}/H(X)$
-defined in the obvious way.
-\end{defn}
-
-\begin{rmk}
-Note that for $X\in B_n$ and $Y\in B_{n+m}$ such that $\eft^m(Y)\jdeq X$, 
-we have $(\mathbb{B}/X)/Y\cong B/Y$.
-\end{rmk}
-
-\begin{defn}
-A B0-system $\mathbb{B}$ consists of a B-framework $\mathbb{B}$ and homomorphisms
-\begin{align*}
-W_{X} & : \mathbb{B}/\eft(X)\to\mathbb{B}/X\\
-S_{x} & : \mathbb{B}/\ebd(x)\to\mathbb{B}/\eft(\ebd(x))
-\end{align*}
-of B-frameworks, for any $X\in B_{n+1}$ and any $x\in\tilde{B}_{n+1}$. 
-\end{defn}
-
-\begin{lem}
-When $\mathbb{B}$ is a B0-system, then so is each $\mathbb{B}/X$.
-\end{lem}
-
-\begin{proof}
-Straightforward.
-\end{proof}
-
-\begin{defn}
-A B0-homomorphism $H:\mathbb{B}\to\mathbb{A}$ is a homomorphism of B-frameworks
-for which the diagrams
-\begin{equation*}
-\begin{tikzcd}[column sep=huge]
-\mathbb{B}/X \arrow[r,"H/X"] & \mathbb{A}/H(X) \\
-\mathbb{B}/\eft(X) \arrow[u,"W_X"] \arrow[r,swap,"H/\eft(X)"] & \mathbb{A}/\eft(H(X)) \arrow[u,swap,"W_{H(X)}"]
-\end{tikzcd}
-\end{equation*}
-and
-\begin{equation*}
-\begin{tikzcd}[column sep=huge]
-\mathbb{B}/\ebd(x) \arrow[r,"H/\ebd(x)"] \arrow[d,swap,"S_x"] & \mathbb{A}/\ebd(H(X)) \arrow[d,"S_{\tilde{H}(x)}"] \\
-\mathbb{B}/\eft(\ebd(x)) \arrow[r,swap,"H/\eft(\ebd(x))"] & \mathbb{A}/\eft(\ebd(H(X)))
-\end{tikzcd}
-\end{equation*}
-of homomorphisms of B-frameworks commute for every 
-$X\in B_n$ and $x\in\tilde{B}_n$. The category of B0-systems is denoted
-by $\mathbf{B0sys}$.
-\end{defn}
-
-Now that we have defined the category of B0systems, we can start defining the
-embedding into pre-E-systems.
-
-\begin{defn}
-Let $\mathbb{B}$ be a B-framework. Then $\mathbb{B}$ determines
-a stratified category $U(\mathbb{B})$. Also, any homomorphism of B-frameworks
-determines a stratified functor.
-\end{defn}
-
-\begin{constr}
-The set of objects of $U(\mathbb{B})$ is taken to be the set
-$\bigsqcup_{(n\in\mathbb{N})}B_n$. Now we define
-$(k,X)\leq (n,Y)$ if and only if $n\leq k$ and $\eft[n]^{k-n}(X)\jdeq Y$.
-In other words, we only have $(n+m,X)\leq (n,\eft^m(X))$. It is immediate from
-this definition that for any $X\in B_{n+1}$ there is a unique $\Gamma\in
-B_n$ such that $(n+1,X)\leq (n,\Gamma)$, so $U(\mathbb{B})$ is stratified.
-
-To show that the functor $U(H):U(\mathbb{B})\to U(\mathbb{A})$ is stratified for
-any homomorphism $H:\mathbb{B}\to\mathbb{A}$ of B-frameworks, note that
-by \autoref{lem:strat_full} it suffices to show that each $U(H)$ 
-lifts factorizations.
-Let $H:\mathbb{B}\to\mathbb{A}$ be a homomorphism of B-frameworks, and let
-$(m,X):(n+m,X)\leq (n,\Gamma)$. Then the category $\mathbf{fact}_{(m,X)}$ of
-factorizations of $(m,X)$ is a finite linear order with
-$m+1$ elements, and so is the category $\mathbf{fact}_{(m,H(X))}$. 
-\end{constr}
-
-\begin{rmk} 
-It is in fact the case that $U(\mathbb{B})$ is the free category generated by the
-graph with $\bigsqcup_{(n\in\mathbb{N})}B_n$ as the set of vertices, and
-for each object of the form $(n+1,X)$ an edge to $(n,\eft[n](X))$. 
-Essentially all $U$ does, is forgetting about the term structure 
-$\tilde{B}$ of $\mathbb{B}$.
-
-Furthermore, it is useful to note that for any $X\in B_n$, we get an 
-isomorphism $U(\mathbb{B}/X)\cong  U(\mathbb{B})/(n,X)$ of rooted trees,
-natural in $X$. Therefore we will usually not distinguish between $U(\mathbb{B})/X$
-and $U(\mathbb{B})/(n,X)$. 
-\end{rmk}
-
-\begin{defn}
-Let $\mathbb{B}$ be a B0-system. Then $U(\mathbb{B})$ can be given the structure
-of a pre-E-system. We write $\mathcal{E}(\mathbb{B})$ for the resulting
-pre-E-system. Thus we get a functor $\mathcal{E}:\mathbf{B0sys}\to\mathbf{Esys_s}$.
-\end{defn}
-
-\begin{constr}
-We begin by simultaneously defining the term structure and the pre-substitution
-structure on $U(\mathbb{B})$ by induction on the natural numbers. More precisely,
-for any $m\in\mathbb{N}$ we will define, for any $(n+m,X)\leq (n,\Gamma)$ a set
-$T(m,X)$ and for any $x\in T(m,X)$ a homomorphism 
-$S_x:\mathbb{B}/X\to\mathbb{B}/\Gamma$ of B-frameworks. Then it follows
-immediately that we get stratified functors $U(S_x):U(\mathbb{B})/X\to U(\mathbb{B})/\Gamma$
-defining the pre-substitution structure of $\mathcal{E}(\mathbb{B})$.
-
-For $(n,X)\in U(\mathbb{B})$, we define $T(0,X)\defeq
-\{\pt[X]\}$ and $S_{\pt[X]}\defeq \catid{\mathbb{B}/X}$. We also define for
-$(n+1,X)\leq (n,\Gamma)$ the set $T(1,X)\jdeq\{x\in B_{n+1}\mid \ebd(x)\jdeq X\}$.
-Note that for $x\in T(1,X)$, the homomorphism $S_x:\mathbb{B}/X \to\mathbb{B}/\Gamma$
-is already assumed to exist, because $\mathbb{B}$ is a B0-system.
-
-For the inductive step, suppose that for $m\in\mathbb{N}$ we have sets $T(m,X)$
-for any $(n+m,X)\leq (n,\Gamma)$ and homomorphisms $S_x:\mathbb{B}/X \to \mathbb{B}/\Gamma$
-of B-frameworks, for any $x\in T(m,X)$. Then we define
-\begin{equation*}
-T(m+1,X)\defeq\bigsqcup_{x\in T(m,\eft(X))} T(1,S_x(X)).
-\end{equation*}
-For $(x,u)\in T(m+1,X)$ we define the homomorphism $S_{(x,u)}:\mathbb{B}/X\to
-\mathbb{B}/\Gamma$ as
-\begin{equation*}
-S_{(x,u)}\defeq S_u\circ S_x/X.
-\end{equation*}
-It remains to define a pre-weakening structure on $U(\mathbb{B})$. To do this,
-it suffices to define homomorphisms $W_{(m,X)}:\mathbb{B}/\Gamma \to\mathbb{B}/X$
-of B-frameworks, for any $(n+m,X)\leq (n,\Gamma)$. We define
-\begin{equation*}
-W_{(m,X)}\defeq W_X\circ\cdots\circ W_{\eft^{m-1}(X)}.\qedhere
-\end{equation*}
-\end{constr}
-
-\begin{thm}
-The functor $\mathcal{E}$ is full and faithful.
-\end{thm}
-
-\begin{proof}
-We have to show that the map $\mathrm{hom}(\mathbb{B},\mathbb{A})\to
-\mathrm{hom}(\mathcal{E}(\mathbb{B}),\mathcal{E}(\mathbb{A}))$ is a bijection
-for any two B0-systems $\mathbb{B}$ and $\mathbb{A}$.
-
-To show that this map is injective, suppose $F,G:\mathbb{B}\to\mathbb{A}$ are 
-B0-homomorphisms such that $\mathcal{E}(F)\jdeq\mathcal{E}(G)$. Then we have
-$F_n(X)\jdeq\mathcal{E}(F)(n,X)\jdeq \mathcal{E}(G)(n,X)\jdeq G_n(X)$ for each
-object $x\in B_n$. Also, we have $\tilde{F}_n(x)\jdeq T_{\mathcal{E}(F)}(n,x)
-\jdeq T_{\mathcal{E}(G)}(n,x)\jdeq \tilde{G}_n(x)$ for each
-$x\in\tilde{B}_n$. This shows that $F\jdeq G$, so we conclude that $\mathcal{E}$
-is faithful.
-
-To show that the map $\mathrm{hom}(\mathbb{B},\mathbb{A})\to
-\mathrm{hom}(\mathcal{E}(\mathbb{B}),\mathcal{E}(\mathbb{A}))$ 
-is surjective, let $H:\mathcal{E}(\mathbb{B})\to
-\mathcal{E}(\mathbb{A})$. Since both $\mathcal{E}(\mathbb{B})$ and
-$\mathcal{E}(\mathbb{A})$ are stratified, we get from
-\autoref{lem:strat_full} that $H$ is a stratified pre-E-homomorphism. 
-Because $H$ is stratified, we get maps $H_n:B_n\to A_n$ for any $n\in\mathbb{N}$.
-The term structure of $H$ gives maps $\tilde{H}_n:
-\tilde{B}_n\to\tilde{A}_n$. This defines a B0-homomorphism since $H$ 
-preserves the pre-weakening and pre-substitution structures of $\mathcal{E}(\mathbb{B})$.
-It is immediate from the definition of $H_n$ and
-$\tilde{H}_n$ that this B0-homomorphism is mapped back to $H$, so we conclude
-that $\mathcal{E}$ is full.
-\end{proof}
-
-\subsection{B-systems are stratified E-systems}
-There are two properties which hold for pre-E-systems of the form $\mathcal{E}
-(\mathbb{B})$ for some B0-system $\mathbb{B}$, 
-which stand in the way of proving that the category of stratified
-pre-E-systems is equivalent to the category of B0-systems. The first is that
-for any $A\in U(\mathbb{B})/\Gamma$ and $P\in U(\mathbb{B})/\ctxext{\Gamma}{A}$,
-we have an isomorphism $T(\ctxext{A}{P})\cong\bigsqcup_{x\in T(A)}T(S_x(P))$.
-The second, which depends on the first, is that for any $(x,u)\in T(\ctxext{A}{P})$
-we have $S(x,u)\jdeq S_u\circ S_x/P$. In other words, both the term structure
-and the pre-substitution structure of $\mathcal{E}(\mathbb{B})$ are compatible
-with the categorical structure of the underlying category $U(\mathbb{B})$.
-However, to prove the corresponding facts for E-systems, we needed the full
-set of conditions on unital E-systems: to prove that the term structure of
-$U(\mathbb{B})$ is compatible with composition we needed units, and to prove 
-that the operations $S_x$ are compatible with the pairing function we needed
-that they are E-homomorphisms.
-
-Presumably, we can assume these compatibility conditions on pre-E-systems in order
-to prove the desired equivalence of B0-systems and stratified pre-E-systems. However,
-that seems somewhat unnatural, especially given that they automatically hold
-for E-systems. Thus, for the remainder of this section we set out to show that
-the category of unital B-systems is equivalent to the category of stratified 
-unital E-systems.
-
-\begin{defn}
-A unital B0-system is a B0-system with maps
-\begin{equation*}
-\delta_n : B_{n+1}\to \tilde{B}_{n+2}.
-\end{equation*}
-such that $\ebd(\delta_n(X))\jdeq W_{X}(X)$ for any $X\in B_n$. 
-A B0-homomorphism $H$ is said to be unital if one has
-\begin{equation*}
-H(\delta(X))\jdeq\delta(H(X))
-\end{equation*}
-for any $X\in B_n$.  
-\end{defn}
-
-\begin{rmk}
-Voevodsky showed in \cite{VV_B-systems} that the category of unital B0-systems
-with unital B0-homomorphisms between them is not a full
-subcategory of $\mathbf{B0sys}$. 
-\end{rmk}
-
-\begin{defn}
-A \emph{(unital) B-system} is a B0-system for which the following conditions
-hold:
-\begin{enumerate}
-\item Every $W_X$ and $S_x$ is a unital B0-homomorphism.
-\item For every $x\in \tilde{B}_{n+1}$ one has $S_x\circ W_{\ebd(x)}\jdeq
-\catid{\mathbb{B}/\eft(\ebd(x))}$. 
-\item For every $x\in\tilde{B}_{n+1}$ one has $S_x(\delta(\ebd(x)))\jdeq x$.
-\item For every $X\in B_{n+1}$ one has $S_{\delta(X)}\circ W_X/X\jdeq
-\catid{\mathbb{B}/X}$. 
-\end{enumerate}
-\emph{B-homomorphisms} are unital B0-homomorphisms between B-systems.
-We denote the category of B-systems by $\mathbf{Bsys}$. 
-\end{defn}
-
-\begin{lem}
-For any non-unital B-system $\mathbb{B}$, $\mathcal{E}(\mathbb{B})$ is a 
-stratified non-unital E-system. 
-\end{lem}
-
-\begin{proof}
-Straightforward.
-\end{proof}
-
-\begin{defn}
-For any unital B-system $\mathbb{B}$, $\mathcal{E}(\mathbb{B})$ has a
-projection structure. We thus obtain a functor 
-$\mathcal{E}:\mathbf{Bsys}\to\mathbf{Esys}$ from unital B-systems to E-systems.
-\end{defn}
-
-\begin{constr}
-We first need to extend the functor $\mathcal{E}:\mathbf{B0sys}\to\mathbf{Esys_p}$
-from non-unital B0-systems to non-unital pre-E-systems, to a functor from
-unital B-systems to E-systems.
-
-We need to give $\mathcal{E}(\mathbb{B})$ the structure of a unital
-pre-E-system. We construct by induction on $m\in \mathbb{N}$, a term of type 
-$T(m,W_{(m,X)}(X))$ with the property that that 
-\begin{equation*}
-S_{\tfid{(m,X)}}\circ W_{(m,X)}/X\jdeq\catid{U(\mathbb{B})/X},
-\end{equation*} 
-for every $(n+m,X)\leq (n,\Gamma)$ by induction on $m$. The extra
-condition is part of the inductive construction, because we need it on the way.
-
-For $m\jdeq 0$ we have but one
-choice, and for $m\jdeq 1$ we can directly use the unital structure of
-$\mathbb{B}$. For the inductive step, recall that
-\begin{equation*}
-T(m+1,W_{(m+1,X)}(X))\jdeq\bigsqcup_{x\in T(m,W_{(m+1,X)}(\eft(X)))}T(1,S_x(W_{(m+1,X)}(X))).
-\end{equation*}
-We have the term $x\defeq \tilde{W}_X(\tfid{(m,\eft(X))})\in T(m,W_{(m+1,X)}(\eft(X)))$.
-Now we may observe that 
-\begin{align*}
-S_x(W_{(m+1,X)}(X))
-  & \jdeq 
-S_{\tilde{W}_X(\tfid{(m,\eft(X))})}(W_{(m+1,X)}(X))\\
-  & \jdeq 
-S_{\tilde{W}_X(\tfid{(m,\eft(X))})}(W_X(W_{(m,\eft(X))}(X))) \\
-  & \jdeq
-W_{X}(S_{\tfid{(m,\eft(X))}}(W_{(m,\eft(X))}(X))) \\
-  & \jdeq 
-W_{X}(X).
-\end{align*}
-Therefore, we see that $T(1,S_x(W_{(m+1,X)}(X)))\jdeq T(1,W_X(X))$, where we can
-choose the term $\delta(X)$. Thus, we define
-\begin{equation*}
-\tfid{(m+1,X)}\defeq (\tilde{W}_X(\tfid{(m,\eft(X))}),\delta(X))
-\end{equation*} 
-for $m\geq 1$. Now we need to check that $S_{\tfid{(m+1,X)}}\circ 
-W_{(m+1,X)}/X\jdeq \catid{U(\mathbb{B})/X}$. This is indeed the case:
-\begin{align*}
-S_{\tfid{(m+1,X)}}\circ W_{(m+1,X)}/X
-  & \jdeq
-S_{\delta(X)}\circ S_{\tilde{W}_X(\tfid{(m,\eft(X))})}\circ W_{(m+1,X)}/X \\
-  & \jdeq
-S_{\delta(X)}\circ W_X/X \\
-  & \jdeq
-\catid{U(\mathbb{B})/X}.
-\end{align*}
-This finishes the definition of the pre-projection structure on $\mathcal{E}(\mathbb{B})$.
-Straightforward proves by induction give that each $W_{(m,X)}$ and $S_x$ is a 
-pre-projection homomorphism and that $S_x(\tfid{(m,X)})\jdeq x$, for any 
-$x\in T(m,X)$.
-\end{constr}
-
-\begin{thm}
-The functor $\mathcal{E}:\mathbf{Bsys}\to\mathbf{Esys}$ lifts to an equivalence
-to the subcategory $\mathbf{Esys_s}$ of stratified E-systems: i.e.~we have a
-commuting diagram
-\begin{equation*}
-\begin{tikzcd}
-& \mathbf{Esys_s} \arrow[dr,"\subseteq"]\\
-\mathbf{Bsys} \arrow[ur,"\simeq"] \arrow[rr,swap,"\mathcal{E}"] & & \mathbf{Esys}
-\end{tikzcd}
-\end{equation*}
-\end{thm}
-
-\begin{proof}
-We already have shown that $\mathcal{E}$ factors through the stratified
-E-systems and that it is full and faithful, so it remains
-to show that every stratified E-system is in the image of $\mathcal{E}$. In
-other words, for every stratified E-system $\mathbb{E}$ we need to construct
-a B0-system $\mathbb{B}$ such that $\mathcal{E}(\mathbb{B})\cong\mathbb{E}$.
-
-Let $\mathbb{E}$ be a stratified E-system. From the stratification functor
-$L:\mathbb{C}_{\mathbb{E}}\to (\mathbb{N},\geq)$ we obtain the pre-image
-presheaf $B(\mathbb{E}):\op{(\mathbb{N},\leq)}\to\mathbf{Set}$ given by
-\begin{align*}
-B(\mathbb{E})_n & \defeq \{X\in\mathrm{Ob}(\mathbb{C}_{\mathbb{E}})\mid L(X)\jdeq n\}
-\end{align*}
-and $\eft(X)\defeq\mathrm{cod}(p_X)$. We define the term structure of $B(\mathbb{E})_n$
-to be given by
-\begin{equation*}
-\tilde{B}(\mathbb{E})_n \defeq \bigsqcup_{X\in B(\mathbb{E})_n} T(p_X)
-\end{equation*}
-with the projection map $\ebd[n]\defeq\pi_1:\tilde{B}(\mathbb{E})_n\to B(\mathbb{E})_n$.
-This defines the B-framework $B(\mathbb{E})$, and we obtain a functor
-$B:\mathbf{Esys_s}\to\mathbf{B0sys}$. It is immediate that we get isomorphisms
-$U(B(\mathbb{E}))\cong\cat{C}_{\mathbb{E}}$, natural in $\mathbb{E}$. Also, we
-get natural isomorphisms $B(\mathbb{E}/\Gamma)\cong B(\mathbb{E})/\Gamma$.
-
-The weakening and substitution structures on
-$B(\mathbb{E})$ are given by $W_X\defeq B(W_{p_X})$ and $S_{(X,x)}\defeq B(S_x)$,
-respectively. Since these functors are assumed to be stratified, they define
-homomorphisms of B-frameworks. Because a weakening system of an E-system is
-required to be functorial, i.e.~$W_{\ctxext{A}{P}}\jdeq W_P\circ W_A$ and
-$W_{\catid{\Gamma}}\jdeq\catid{\cat{C}_{\mathbb{E}}/\Gamma}$, it follows
-immediately that the weakening structure on $U(B(\mathbb{E}))$ is isomorphic to
-that of $\mathbb{E}$. 
-\end{proof}
-
-\begin{comment}
-In the following definition $\Delta(X)$ denotes the discrete category on the set
-$X$ and $\mathbf{2}$ is the category $s\to t$.
-\begin{defn}
-Define $\mathbb{S}$ to be the pushout
-\begin{equation*}
-\begin{tikzcd}[column sep=large]
-\Delta(\mathbb{N}) \arrow[r,"{n\mapsto(n,s)}"] \arrow[d,swap,"i"] & \Delta(\mathbb{N})\times\mathbf{2} \arrow[d] \\
-(\mathbb{N},\leq) \arrow[r] & \mathbb{S}
-\end{tikzcd}
-\end{equation*}
-in $\mathbf{Cat}$.
-\end{defn}
-
-Note that for any $n\in\mathbb{N}$, we have the map $S^n := (m\mapsto m+n)$ which
-is monotone. By the universal property of pushouts, we obtain a functor $\sigma^n:\mathbb{S}\to\mathbb{S}$
-given by
-\begin{equation*}
-\begin{tikzcd}[column sep=huge]
-\Delta(\mathbb{N}) 
-  \arrow[r,"{n\mapsto(n,s)}"] 
-  \arrow[d,swap,"i"] 
-  & 
-\Delta(\mathbb{N})\times\mathbf{2} 
-  \arrow[d]
-  \arrow[r,"{\Delta(S^n)\times\catid{\mathbf{2}}}"]
-  &
-\Delta(\mathbb{N})\times\mathbf{2}
-  \arrow[dd]
-  \\
-(\mathbb{N},\leq) 
-  \arrow[r] 
-  \arrow[d,swap,"{(S^n,\leq)}"]
-  & 
-\mathbb{S}
-  \arrow[dr,densely dotted,yshift=.5ex,"\sigma^n"]
-  \\
-(\mathbb{N},\leq)
-  \arrow[rr] & &
-\mathbb{S}
-\end{tikzcd}
-\end{equation*}
-
-\begin{defn}
-A non-unital pre-B0-system is a presheaf $B$ on $\mathbb{S}$ such that $B(0)$ is
-a singleton, and the following additional structure:
-\begin{enumerate}
-\item For any $n\in\mathbb{N}$ and $X\in B(n)$, write $B/X$ for the sub-presheaf
-of $B\circ\sigma^n$ given by $B/X(0)=\{X\}$ and $B/X(n+1)=B_{n\leq n+1}^{-1}(B/X(n))$. 
-We require that for any $n\in\mathbb{N}$ and any $A\in B(n+1)$, there is a 
-natural transformation $W_A:B/(B_{n\leq n+1}(A))\Rightarrow B/A$.
-\end{enumerate}
+which are compatible with $\mathrm{cod}$, $\mathrm{dom}$, $\catid{}$, $\circ$
+and $\ebd$
 \end{defn}
 \end{comment}
-
-\subsection{From D-systems to locally stratified E'-systems}
-
-We use `D-system' to refer to Dybjer's categories with families. We recall
-Dybjer's definition. Note that $\mathbf{Fam}$ is the category of families of
-sets, of which the objects consist of a set $A$ and a family $\{B(x)\}_{x\in A}$
-over $A$. Although a terminal object is assumed to exist, not every object 
-of a D-system is `fibrant', i.e.~it is
-not the case that every $X\to 1$ is a composition of projections.
-We could only hope that there is a functor from D-sytems to E-systems without
-terminal objects. We write $\mathbf{Esys}_{\circ}$ for the category of E-systems
-without terminal objects.
-
-\begin{defn}
-A category is said to be locally stratified if each of its slice categories
-are stratified.
-\end{defn}
-
-\begin{defn}
-A \emph{setup for categories with families} consists of
-\begin{enumerate}
-\item A category $\cat{C}$ with a terminal object $1$.
-\item A presheaf $\mathrm{Ty}:\op{\cat{C}}\to\mathbf{Set}$. We will write
-$f_\ast(A)$ for $\mathrm{Ty}(f,A)$. 
-\item A commuting triangle
-\begin{equation*}
-\begin{tikzcd}
-\int_\cat{C}\mathrm{Ty} \arrow[dr,swap,"\pi"] \arrow[rr,"p"] & & \cat{C}^{\to} \arrow[dl,"\mathrm{cod}"] \\
-& \cat{C}
-\end{tikzcd}
-\end{equation*} 
-We will write $p_A:\ctxext{\Gamma}{A}\to\Gamma$
-for the morphism $U(\Gamma,A)$ into $\Gamma$, for each object $\Gamma\in\cat{C}$ and
-$A\in\mathrm{Ty}(\Gamma)$.
-\end{enumerate}
-\end{defn}
-
-\begin{defn}
-A \emph{category with families} consists of a setup with
-\begin{enumerate}
-\item[(iv)] a preasheaf $\mathrm{Tm}:\op{(\int_\cat{C}\mathrm{Ty})}\to\mathbf{Set}$.
-The action of a morphism $f:(\Delta,f_\ast(A))\to(\Gamma,A)$ of $\int_\cat{C}\mathrm{Ty}$
-on a term $x\in \mathrm{Tm}(A)$ is also written by $f_\ast(x)$.
-\item[(v)] a term $q_A\in\mathrm{Tm}(\ctxext{\Gamma}{A},(p_A)_\ast(A))$ for every
-$A\in\mathrm{Ty}(\Gamma)$, 
-\end{enumerate}
-such that for every $f:\Delta\to\Gamma$ in $\cat{C}$ and 
-$x\in \mathrm{Tm}(\Delta,f_\ast(A))$ there is a unique morphism 
-$\tilde{x}:\Delta\to \ctxext{\Gamma}{A}$ such that $p_A\circ\tilde{x}\jdeq f$ and 
-$\tilde{x}_\ast(q_A)\jdeq x$.
-\end{defn}
-
-\begin{lem}\label{lem:cwf_to_dsys}
-Let $\mathbb{D}$ be a category with families. Then there are bijections
-\begin{equation*}
-\zeta_{A,f}:\{\theta:\Delta\to\ctxext{\Gamma}{A}\mid p_A\circ\theta\jdeq f\}\cong \mathrm{Tm}(\Delta,f_\ast(A))
-\end{equation*}
-coherent in $f$ in the sense that
-\begin{align*}
-\zeta_{A,\catid{\Gamma}}(\theta) & \jdeq \theta \\
-\zeta_{A,f\circ g}(\theta) & \jdeq \zeta_{f_\ast(A),g}(\zeta_{A,f}(\theta)\circ g)
-\end{align*}
-for any $g:X\to\Delta$. 
-\end{lem}
-
-\begin{proof}
-Let $\theta:\Delta\to\ctxext{\Gamma}{A}$ be such that $p_A\circ\theta\jdeq f$. 
-Then $\theta_\ast(q_A)$ is an element
-of $\mathrm{Tm}(\theta_\ast((p_A)_\ast(A)))$. By the functoriality of $\mathrm{Ty}$, we have
-$\theta_\ast((p_A)_\ast(A))\jdeq (p_A\circ\theta)_\ast(A)\jdeq f_\ast(A)$. 
-So we obtain $\theta_\ast(q_A)\in\mathrm{Tm}(\Delta,f_\ast(A))$. Now
-the uniqueness of $\theta$ implies that the map $\theta\mapsto \theta_\ast(q_A)$
-is a bijection.
-
-It remains to verify the coherence identities. Let $\theta : \Gamma\to\ctxext{\Gamma}{A}$
-be such that $p_A\circ\theta\jdeq\catid{\Gamma}$. We wish to show that
-$\theta_\ast(q_A)\jdeq\theta$.  
-\begin{align*}
-\theta_\ast(q_A)
-& \jdeq \theta_\ast(\zeta_{A,p_A}(\catid{\ctxext{\Gamma}{A}})) \\
-& \jdeq \cdots \\
-& \jdeq \theta
-\end{align*}
-\end{proof}
-
-The previous lemma renders the presheaf $\mathrm{Tm}$ superfluous. Therefore,
-we define D-systems as follows.
-
-\begin{defn}
-A \emph{D-system} consists of a setup for categories with families
-such that for every $f:\Delta\to\Gamma$ and $A\in\mathrm{Ty}(\Gamma)$, there
-is a bijection
-\begin{equation*}
-\zeta_{A,f}:\{\theta:\Delta\to\ctxext{\Gamma}{A}\mid p_A\circ\theta\jdeq f\}
-  \cong
-\{\eta:\Delta\to\ctxext{\Delta}{f_\ast(A)}\mid p_{f_\ast(A)}\circ\eta\jdeq\catid{\Delta}\}
-\end{equation*}
-natural in $f$ in the sense that the equalities
-\begin{align*}
-\zeta_{A,\catid{\Gamma}}(\theta) & \jdeq \theta \\
-\zeta_{A,f\circ g}(\theta\circ g) & \jdeq \zeta_{f_\ast(A),g}(\zeta_{A,f}(\theta)\circ g)
-\end{align*}
-hold for any $f:\Delta\to\Gamma$ and $g:X\to\Delta$, and any $\theta:\Delta\to\ctxext{\Gamma}{A}$
-such that $p_A\circ\theta\jdeq f$. 
-\end{defn}
-
-\begin{thm}
-A category with families is the same thing as a D-system.
-\end{thm}
-
-\begin{proof}
-In \autoref{lem:cwf_to_dsys} we have shown that every category with families is
-a D-system. Thus, it suffices to show the converse.
-
-Let $\mathbb{D}$ be a D-system. For any object $\Gamma$ of $\mathbb{D}$ and any
-type $A\in\mathrm{Ty}(\Gamma)$, we define 
-\begin{equation*}
-\mathrm{Tm}(A)\defeq \{ x :\Gamma\to\ctxext{\Gamma}{A}\mid p_A\circ x\jdeq\catid{\Gamma}\}.
-\end{equation*}
-For any morphism $f:\Delta\to\Gamma$ and $A\in\mathrm{Ty}(\Gamma)$, we define
-the function $\mathrm{Tm}(f):\mathrm{Tm}(A)\to\mathrm{Tm}(f_\ast(A))$ by
-\begin{equation*}
-\mathrm{Tm}(f,x)\defeq \zeta_{A,f}(x\circ f).
-\end{equation*}
-We have to verify that this satisfies the functorial identities. For $f:\Delta\to\Gamma$
-and $g:X\to\Delta$ and any $x\in\mathrm{Tm}(A)$, we have
-\begin{align*}
-\mathrm{Tm}(f\circ g,x) & \jdeq \zeta_{A,f\circ g}(x\circ f\circ g) \\
-& \jdeq \zeta_{f_\ast(A),g}(\zeta_{A,f}(x\circ f)\circ g) \\
-& \jdeq \mathrm{Tm}(g,\mathrm{Tm}(f,x)).
-\end{align*}
-and
-\begin{equation*}
-\mathrm{Tm}(\catid{\Gamma},x) \jdeq \zeta_{A,\catid{\Gamma}}(x) \jdeq x.
-\end{equation*}
-This finishes the construction of the presheaf $\mathrm{Tm}$ on $\int_{\cat{C}}\mathrm{Ty}$.
-
-Next, we have to find $q_A\in\mathrm{Tm}((p_A)_\ast(A))$. Note that we have the
-isomorphism
-\begin{align*}
-\mathrm{Tm}((p_A)_\ast(A))
-& \jdeq \{\eta :\ctxext{\Gamma}{A}\to\ctxext{{\Gamma}{A}}{(p_A)_\ast(A)}\mid p_{(p_A)_\ast(A)}\circ\theta\jdeq\catid{\ctxext{\Gamma}{A}}\} \\
-& \cong \{\theta : \ctxext{\Gamma}{A}\to\ctxext{\Gamma}{A} \mid p_A\circ \theta\jdeq p_A\}
-\end{align*}
-The latter set contains the identity morphism $\catid{\ctxext{\Gamma}{A}}$. Therefore,
-we find
-\begin{equation*}
-q_A \defeq \zeta_{A,p_A}(\catid{\ctxext{\Gamma}{A}})\in\mathrm{Tm}((p_A)_\ast(A)).
-\end{equation*}
-To verify that we have constructed a category with families, let $f:\Delta\to\Gamma$
-and $x\in T(f_\ast(A))$. Note that we have the morphism 
-$\tilde{x}\defeq \zeta_{A,f}^{-1}(x):\Delta\to\ctxext{\Gamma}{A}$. 
-We need to check that $\tilde{x}$ is unique with the property that 
-$\mathrm{Tm}(\tilde{x},q_A)\jdeq x$. The coherence identities can be used to see
-that this is indeed the case:
-\begin{align*}
-\mathrm{Tm}(\tilde{x},q_A)
-& \jdeq \zeta_{(p_A)_\ast(A),\tilde{x}}(q_A\circ\tilde{x}) \\
-& \jdeq \zeta_{(p_A)_\ast(A),\tilde{x}}(\zeta_{A,p_A}(\catid{\ctxext{\Gamma}{A}})\circ\tilde{x}) \\
-& \jdeq \zeta_{A,p_A\circ\tilde{x}}(\tilde{x}) \\
-& \jdeq \zeta_{A,f}(\tilde{x}) \\
-& \jdeq x.
-\end{align*}
-For the uniqueness, let $\theta:\Delta\to\ctxext{\Gamma}{A}$ be any morphism such that
-$p_A\circ\theta\jdeq f$ and $\mathrm{Tm}(\theta,q_A)\jdeq x$. Then we have
-\begin{align*}
-x & \jdeq \mathrm{Tm}(\theta,q_A) \\
-& \jdeq \zeta_{(p_A)_\ast(A),\theta}(q_A\circ\theta) \\
-& \jdeq \zeta_{(p_A)_\ast(A),\theta}(\zeta_{A,p_A}(\catid{\ctxext{\Gamma}{A}})\circ\theta \\
-& \jdeq \zeta_{A,p_A\circ\theta}(\theta) \\
-& \jdeq \zeta_{A,f}(\theta).
-\end{align*}
-Therefore we see that $\tilde{x}\jdeq\zeta_{A,f}^{-1}(x)\jdeq\theta$.
-\end{proof}
-
-\begin{defn}
-An E'-system is said to be locally stratified, if the category $\cat{F}$ is
-locally stratified.
-\end{defn}
-
-\begin{comment}
-\begin{rmk}
-In the following definition, one could take the underlying category of the
-E-system $\mathcal{E}(\mathbb{D})$ associated to a D-system $\mathbb{D}$ to
-be the subcategory generated by the projections $p_A$, and construct the
-rest of the structure of E-systems on top of that. However, it is not so easy
-to characterize the image of the functor $\mathcal{E}$ in that case. Intuitively,
-D-systems should correspond to locally stratified E-systems. To be able to prove
-such a theorem, we've gone with a slightly modified definition of the underlying
-category.
-\end{rmk}
-\end{comment}
-
-\begin{defn}
-For any D-system $\mathbb{D}$ there is 
-a locally stratified E'-system $U(\mathbb{D})$ without terminal object in 
-$\cat{F}$. Thus, we get a functor $\mathbf{Dsys}\to\mathbf{E'sys}_{\circ}$
-\end{defn}
-
-\begin{constr}
-Let $\mathbb{D}$ be a D-system with underlying category $\cat{C}$. We define
-$\cat{F}\subseteq\cat{C}$ to be the
-\begin{comment}
-category consisting of 
-\begin{enumerate}
-\item equivalence classes of finite lists of the form $(\Gamma,A_1,\ldots A_n)$, where $\Gamma$ is
-an object of $\cat{C}$ and $A_{i+1}\in\mathrm{Ty}(\ctxext{{{\Gamma}{A_1}}{\cdots}}{A_i})$, where
-the equivalence relation is generated by
-\begin{equation*}
-(\ctxext{{{\Gamma}{A_1}}{\cdots}}{A_n},A_{n+1},\ldots,A_{n+m})\sim (\Gamma,A_1,\ldots,A_{n+m}).
-\end{equation*}
-\item
-\end{enumerate}
-\end{comment}
-subcategory with the same objects as
-$\cat{C}$, where the morphisms are generated by the projections $p_A$.
-
-Let $f:\Delta\to\Gamma$. We define $f_\ast:\cat{F}/\Gamma\to\cat{F}/\Delta$ by
-induction on the length of the arrow $A$ into $\Gamma$, simultaneously with a
-morphism $\pi_2(f,A)$ from $\ctxext{\Delta}{f_\ast(A)}$ into $\ctxext{\Gamma}{A}$
-for which the square
-\begin{equation*}
-\begin{tikzcd}
-\ctxext{\Delta}{f_\ast(A)} \arrow[r,"{\pi_2(f,A)}"] \arrow[d,fib] & \ctxext{\Gamma}{A} \arrow[d,fib] \\
-\Delta \arrow[r,swap,"f"] & \Gamma
-\end{tikzcd}
-\end{equation*}
-is a pullback square. We define
-$f_\ast(\catid{\Gamma})\defeq\catid{\Delta}$, and $\pi_2(f,\catid{\Gamma})\defeq f:\ctxext{\Delta}{f_\ast(\catid{\Gamma})}\to\ctxext{\Gamma}{\catid{\Gamma}}$. The resulting square is trivially
-a pullback square. Now suppose we have a pullback square as above.
-and let $B\in\mathrm{Ty}(\ctxext{\Gamma}{A})$. Then we define
-\begin{equation*}
-f_\ast(A\circ p_B)\defeq \pi_2(f,A)_\ast(B).
-\end{equation*}
-By \autoref{lem:EE_pb}, we get the top pullback square in the diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=6em]
-\ctxext{{\Delta}{f_\ast(A)}}{f_\ast(B)} \arrow[d,fib] \arrow[r,"{\pi_2(\pi_2(f,A),B)}"] & \ctxext{{\Gamma}{A}}{B} \arrow[d,fib] \\
-\ctxext{\Delta}{f_\ast(A)} \arrow[d,fib] \arrow[r,"{\pi_2(f,A)}"] & \ctxext{\Gamma}{A} \arrow[d,fib] \\
-\Delta \arrow[r,"f"] & \Gamma
-\end{tikzcd}
-\end{equation*}
-Since the bottom square is assumed to be a pullback, we get that the outer rectangle
-is a pullback, as desired.
-
-Next, we define a weakening structure on $\cat{C}_p$. Since the morphisms of
-$\cat{C}_p$ are composites of projections, we will define
-\begin{equation*}
-W_{p_{A_n}\circ\cdots\circ p_{A_1}}\defeq W_{p_{A_n}}\circ\cdots W_{p_{A_1}}.
-\end{equation*}
-Therefore, we only have to define functors 
-$W_A:\cat{C}_p/\Gamma\to\cat{C}_p/\ctxext{\Gamma}{A}$ with term structure for every 
-$A\in\mathrm{Ty}(\Gamma)$. Since morphisms are composites of projections, we
-can define this functor by induction. Let $W_A(\catid{\Gamma})\defeq
-\catid{\ctxext{\Gamma}{A}}$ and $W_A(p_Q\circ f)\defeq$
-\end{constr}
-
-\begin{thm}
-There is a full and faithful functor from the category of D-systems to the
-category of E-systems.
-\end{thm}
-
-\subsection{The equivalence of E-systems and E'-systems}
-
-The definition of D-system suggests a more general structure related to E-systems.
-
-\begin{defn}
-An \emph{E'-system} consists of a category $\cat{C}$ with a subcategory $\cat{F}$ with the
-same objects as $\cat{C}$, with
-the following structure
-\begin{enumerate}
-\item $\cat{F}$ has a terminal object.
-\item For any $f:\Delta\to\Gamma$ in $\cat{C}$, a functor $f_\ast:\cat{F}/\Gamma\to
-\cat{F}/\Delta$ %and
-%for any $Q\in\cat{F}/\ctxext{\Gamma}{B}$ and any section $g$ of $Q$ a section
-%$f_\ast(g)$ of $f_\ast(Q)$, 
-so that $f_\ast$ is contravariant in $f$ and preserves the terminal object. 
-\item For any $A\in\cat{F}/\Gamma$ and $f:\Delta\to\Gamma$, lifts of
-$f$ to $A$ are naturally the same thing as sections of
-$f_\ast(A)$, i.e. there is an isomorphism
-\begin{equation*}
-\zeta_{A,f}%\defeq\theta\mapsto f_\ast(\theta)
-:\{\theta:\Delta\to\ctxext{\Gamma}{A}\mid A\circ\theta\jdeq f\}
-  \cong
-\{\eta:\Delta\to\ctxext{\Delta}{(f_\ast(A))}\mid f_\ast(A)\circ\eta\jdeq\catid{\Delta}\}
-\end{equation*}
-which are natural in $A$ and $f$, in the sense that
-\begin{enumerate}
-\item For any $P\in\cat{F}/\ctxext{\Gamma}{A}$, $f:\Delta\to\Gamma$ and
-$\theta:\Delta\to\ctxext{{\Gamma}{A}}{P}$ such that $(\ctxext{A}{P})\circ\theta\jdeq f$,
-one hase
-\begin{equation*}
-\zeta_{A,f}(P\circ\theta)\jdeq f_\ast(P)\circ \zeta_{\ctxext{A}{P},f}(\theta).
-\end{equation*}
-\item For any $A\in\cat{F}/\Gamma$, $f:\Delta\to\Gamma$, $g:X\to\Delta$
-and $\theta:\Delta\to\ctxext{\Gamma}{A}$ such that $A\circ\theta\jdeq f$, one
-has
-\begin{align*}
-\zeta_{A,\catid{\Gamma}}(\theta) & \jdeq \theta \\
-\zeta_{A,f\circ g}(\theta\circ g) & \jdeq\zeta_{f_\ast(A),g}(\zeta_{A,f}(\theta)\circ g).
-\end{align*}
-\end{enumerate}
-\end{enumerate}
-\end{defn}
-
-We can give the category $\cat{F}$ a term structure as follows:
-
-\begin{defn}
-For any E'-system $\mathbb{E}$, the category $\cat{F}$ has a term structure
-where the terms of $A$ are sections of $A$, i.e.
-\begin{equation*}
-T(A)\defeq\{\theta:\Gamma\to\ctxext{\Gamma}{A}\mid A\circ\theta\jdeq\catid{\Gamma}\}.
-\end{equation*}
-\end{defn}
-
-\begin{lem}\label{lem:ptm}
-Let $\mathbb{E}$ be an E'-system. Then there is a function $T(A)\to T(f_\ast(A))$
-for any $f:\Delta\to\Gamma$ in $\cat{C}$ and $A\in\cat{F}/\Gamma$, contravariant
-in $f$. 
-\end{lem}
-
-\begin{proof}
-Using the natural isomorphism $\zeta_{A,f}$, we see that
-\begin{align*}
-T(f_\ast(A)) & \jdeq \{\eta : \Delta\to\ctxext{\Delta}{f_\ast(A)}\mid f_\ast(A)\circ\eta\jdeq\catid{\Delta}\}
-  \\
-  & \cong
-\{\theta : \Delta\to \ctxext{\Gamma}{A}\mid A\circ\theta\jdeq f\}.
-\end{align*}
-Therefore, we have the function $x\mapsto \zeta_{A,f}(x\circ f)$ from $T(A)$ to
-$T(f_\ast(A))$. 
-\end{proof}
-
-\begin{defn}
-For any $A\in\cat{F}/\Gamma$, we define
-$W_A\defeq A_\ast:\cat{F}/\Gamma\to\cat{F}/\ctxext{\Gamma}{A}$. Then we get
-for any $A,B\in\cat{F}/\Gamma$ an isomorphism
-\begin{equation*}
-\mathrm{hom}_{\cat{C}/\Gamma}(A,B)
-  \cong
-T(W_A(B)).
-\end{equation*}
-In particular, we get a term $\idtm{A}\in T(W_A(A))$. 
-\end{defn}
-
-\begin{constr}
-Let $A,B\in\cat{F}/\Gamma$. Then we have the isomorphism
-\begin{align*}
-T(W_A(B)) & \jdeq \{\eta:\ctxext{\Gamma}{A}\to\ctxext{{\Gamma}{A}}{W_A(B)}\mid W_A(B)\circ\eta\jdeq\catid{\ctxext{\Gamma}{A}}\} \\
-  & \cong
-\{\theta:\ctxext{\Gamma}{A}\to\ctxext{\Gamma}{B}\mid B\circ\theta\jdeq A\},
-\end{align*}
-so we find $\idtm{A}\defeq\zeta_{A,A}(\catid{\ctxext{\Gamma}{A}})\in T(W_A(A))$.
-\end{constr}
-
-\begin{defn}
-For any $P\in\cat{F}/\ctxext{\Gamma}{A}$ and $f:\Delta\to\Gamma$ 
-there is a morphism $\pi_2(f,A):\ctxext{\Delta}{f_\ast(A)}\to\ctxext{\Gamma}{A}$
-such that $A\circ \pi_2(f,A)\jdeq f\circ f_\ast(A)$.
-\end{defn}
-
-\begin{constr}
-\begin{align*}
-& \{ \theta : \ctxext{\Delta}{f_\ast(A)}\to\ctxext{\Gamma}{A}\mid A\circ\theta\jdeq f\circ f_\ast(A) \} \\
-  & \cong
-\{ \eta : \ctxext{\Delta}{f_\ast(A)}\to\ctxext{{\Delta}{f_\ast(A)}}{(f\circ f_\ast(A))_\ast(A)}
-  \mid (f\circ f_\ast(A))_\ast(A)\circ\eta\jdeq\catid{\ctxext{\Delta}{f_\ast(A)}}\} \\
-  & \jdeq
-\{ \eta : \ctxext{\Delta}{f_\ast(A)}\to\ctxext{{\Delta}{f_\ast(A)}}{(f_\ast(A))_\ast(f_\ast(A))}
-  \mid (f\circ f_\ast(A))_\ast(A)\circ\eta\jdeq\catid{\ctxext{\Delta}{f_\ast(A)}}\} \\
-  & \jdeq
-T(W_{f_\ast(A)}(f_\ast(A))).
-\end{align*}
-Since we have $\tfid{f_\ast(A)}\in T(W_{f_\ast(A)}(f_\ast(A)))$, we obtain 
-\begin{equation*}
-\pi_2(f,A)\defeq \zeta_{A,f\circ f_\ast(A)}^{-1}(\tfid{f_\ast(A)}):\ctxext{\Delta}{f_\ast(A)}\to\ctxext{\Gamma}{A}
-\end{equation*}
-satisfying $A\circ\pi_2(f,A)\jdeq f\circ f_\ast(A)$.
-\end{constr}
-
-\begin{lem}
-For any $f:\Delta\to\Gamma$, $g:X\to\Delta$ and $A\in\cat{F}/\Gamma$, one has
-\begin{equation*}
-\pi_2(f,A)\circ\pi_2(g,f_\ast(A))\jdeq\pi_2(f\circ g,A).
-\end{equation*}
-\end{lem}
-
-\begin{proof}
-It suffices to show that
-\begin{equation*}
-\zeta_{A,f\circ g\circ g_\ast(f_\ast(A))}(\pi_2(f,A)\circ\pi_2(g,f_\ast(A)))\jdeq\tfid{g_\ast(f_\ast(A))}.
-\end{equation*}
-Because $g\circ g_\ast(f_\ast(A))\jdeq f_\ast(A)\circ\pi_2(g,f_\ast(A))$, we get
-to use the naturality of $\zeta$:
-\begin{align*}
-& \zeta_{A,f\circ g\circ g_\ast(f_\ast(A))}(\pi_2(f,A)\circ\pi_2(g,f_\ast(A))) \\
-& \jdeq
-\zeta_{A,f\circ f_\ast(A)\circ\pi_2(g,f_\ast(A))}(\pi_2(f,A)\circ\pi_2(g,f_\ast(A))) \\
-& \jdeq
-\zeta_{f_\ast(A)_\ast(f_\ast(A)),\pi_2(g,f_\ast(A))}(\zeta_{A,f\circ f_\ast(A)}(\pi_2(f,a))\circ\pi_2(g,f_\ast(A))) \\
-& \jdeq
-\zeta_{f_\ast(A)_\ast(f_\ast(A)),\pi_2(g,f_\ast(A))}(\tfid{f_\ast(A)}\circ\pi_2(g,f_\ast(A))) \\
-& \jdeq
-\zeta_{f_\ast(A)_\ast(f_\ast(A)),\pi_2(g,f_\ast(A))}(\zeta_{f_\ast(A),f_\ast(A)}(\catid{\ctxext{\Delta}{f_\ast(A)}})\circ\pi_2(g,f_\ast(A))) \\
-& \jdeq
-\zeta_{f_\ast(A),f_\ast(A)\circ\pi_2(g,f_\ast(A))}(\catid{\ctxext{\Delta}{f_\ast(A)}}\circ\pi_2(g,f_\ast(A))) \\
-& \jdeq
-\zeta_{f_\ast(A),f_\ast(A)\circ\pi_2(g,f_\ast(A))}(\pi_2(g,f_\ast(A))) \\
-& \jdeq
-\zeta_{f_\ast(A),g\circ g_\ast(f_\ast(A))}(\pi_2(g,f_\ast(A)))\\
-& \jdeq
-\tfid{g_\ast(f_\ast(A))}.\qedhere
-\end{align*}
-\end{proof}
-
-\begin{lem}
-Let $f:\Delta\to\Gamma$ and $A\in\cat{F}/\Gamma$. Then one has 
-\begin{equation*}
-\pi_2(f,A)\circ\zeta_{A,f}(\theta)\jdeq\theta
-\end{equation*}
-for any $\theta:\Delta\to\ctxext{\Gamma}{A}$ satisfying $A\circ\theta\jdeq f$.
-\end{lem}
-
-\begin{proof}
-We will write $\tilde\theta\defeq\zeta_{A,f}(\theta)$. Note that
-\begin{equation*}
-A\circ\pi_2(f,A)\circ\tilde\theta\jdeq f\circ f_\ast(A)\circ\tilde\theta\jdeq f.
-\end{equation*} 
-Therefore, we can prove that $\pi_2(f,A)\circ\tilde\theta\jdeq\theta$ by showing
-that $\zeta_{A,f}(\pi_2(f,A)\circ\tilde\theta)\jdeq\tilde\theta$. This is
-shown using the naturality of $\zeta$:
-\begin{align*}
-\zeta_{A,f}(\pi_2(f,A)\circ\tilde\theta)
-  & \jdeq
-\zeta_{A,f\circ f_\ast(A)\circ\tilde\theta}(\pi_2(f,A)\circ\tilde\theta)
-  \\
-  & \jdeq
-\zeta_{(f_\ast(A))_\ast(f_\ast(A)),\tilde{\theta}}(\zeta_{A,f\circ f_\ast(A)}(\pi_2(f,A))\circ\tilde\theta)
-  \\
-  & \jdeq
-\zeta_{(f_\ast(A))_\ast(f_\ast(A)),\tilde{\theta}}(\idtm{f_\ast(A)}\circ\tilde\theta)
-  \\
-  & \jdeq
-\zeta_{(f_\ast(A))_\ast(f_\ast(A)),\tilde{\theta}}(\zeta_{f_\ast(A),f_\ast(A)}(\catid{\ctxext{\Delta}{f_\ast(A)}})\circ\tilde\theta)
-  \\
-  & \jdeq
-\zeta_{f_\ast(A),f_\ast(A)\circ\tilde\theta}(\catid{\ctxext{\Delta}{f_\ast(A)}}\circ\tilde\theta) \\
-  & \jdeq
-\zeta_{f_\ast(A),\catid{\Delta}}(\tilde\theta) \\
-  & \jdeq
-\theta.\qedhere
-\end{align*}
-\end{proof}
-
-\begin{lem}\label{lem:EE_pb}
-The diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=large]
-\ctxext{\Delta}{f_\ast(A)} \arrow[d,fib] \arrow[r,"{\pi_2(f,A)}"] & \ctxext{\Gamma}{A} \arrow[d,fib] \\
-\Delta \arrow[r,"f"] & \Gamma
-\end{tikzcd}
-\end{equation*}
-is a pullback square.
-\end{lem}
-
-\begin{proof}
-To verify the universal mapping property of the pullback square, consider a
-commuting diagram
-\begin{equation*}
-\begin{tikzcd}
-X \arrow[d,swap,"g"] \arrow[r,"h"] & \ctxext{\Gamma}{A} \arrow[d,fib] \\
-\Delta \arrow[r,"f"] & \Gamma
-\end{tikzcd}
-\end{equation*}
-Then $h$ corresponds uniquely to a section $k$ of $(f\circ g)_\ast(A)$. Since
-$(f\circ g)_\ast(A)\jdeq g_\ast(f_\ast(A))$, the section $k$ fits in the
-commuting diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=huge]
-\ctxext{X}{g_\ast(f_\ast(A))} \arrow[d,xshift=.7ex,fib] \arrow[r,"{\pi_2(g,f_\ast(A))}"]
-& \ctxext{\Delta}{f_\ast(A)} \arrow[d,fib] \arrow[r,"{\pi_2(f,A)}"]
-& \ctxext{\Gamma}{A} \arrow[d,fib] \\
-X \arrow[r,swap,"g"] \arrow[u,xshift=-.7ex,"k"] & \Delta \arrow[r,swap,"f"] & \Gamma
-\end{tikzcd}
-\end{equation*}
-so we obtain $u\defeq\pi_2(g,f_\ast(A))\circ k:X\to\ctxext{\Delta}{f_\ast(A)}$. 
-Then we have $f_\ast(A)\circ u\jdeq g\circ g_\ast(f_\ast(A))\circ k\jdeq g$ and
-$\pi_2(f,A)\circ u\jdeq \pi_2(f\circ g,A)\circ k\jdeq h$. Note that
-$u$ corresponds uniquely to $k$, which corresponds uniquely to $h$, so the
-uniqueness of $u$ is automatic.
-\end{proof}
-
-\begin{lem}
-For any $f:\Delta\to\Gamma$ and $P\in\cat{F}/\ctxext{\Gamma}{A}$, there is a unique isomorphism
-$\varphi_{f,P}:\pi_2(f,A)_\ast(P)\cong (f_\ast/A)(P)$ in 
-$\mathbb{C}/\ctxext{\Delta}{f_\ast(A)}$ such that
-\begin{equation*}
-\pi_2(f,\ctxext{A}{P})\jdeq \pi_2(\pi_2(f,A),P)\circ\varphi_{f,P}.
-\end{equation*}
-Consequently, we get isomorphisms
-\begin{align*}
-\{\theta:\ctxext{\Delta}{f_\ast(A)}\to\ctxext{{\Gamma}{A}}{P}\mid P\circ\theta\jdeq \pi_2(f,A)\}
-&  \cong
-\{\eta:\ctxext{\Delta}{f_\ast(A)}\to\ctxext{{\Delta}{(f_\ast(A))}}{f_\ast(P)}\mid f_\ast(P)\circ\eta\jdeq\catid{\ctxext{\Delta}{f_\ast(A)}}\}
-\end{align*}
-natural in $P$ and $f$.
-\end{lem}
-
-\begin{proof}
-Note that both the lower square and the outer rectangle in the diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=huge]
-\ctxext{{\Delta}{f_\ast(A)}}{f_\ast(P)} \arrow[d,fib] \arrow[r,"{\pi_2(f,\ctxext{A}{P})}"]
-& \ctxext{{\Gamma}{A}}{P} \arrow[d,fib] \\
-\ctxext{\Delta}{f_\ast(A)} \arrow[d,fib] \arrow[r,"{\pi_2(f,A)}"] 
-& \ctxext{\Gamma}{A} \arrow[d,fib] \\
-\Delta \arrow[r,swap,"f"] & \Gamma
-\end{tikzcd}
-\end{equation*}
-are pullback squares. By the pasting lemma for pullbacks, it follows that the
-top square must also be a pullback. On the other hand, we have the pullback
-square
-\begin{equation*}
-\begin{tikzcd}
-\ctxext{{\Delta}{f_\ast(A)}}{\pi_2(f,A)_\ast(P)} \arrow[d,fib] \arrow[r]
-& \ctxext{{\Gamma}{A}}{P} \arrow[d,fib] \\
-\ctxext{\Delta}{f_\ast(A)} \arrow[r,swap,"{\pi_2(f,A)}"]
-& \ctxext{\Gamma}{A}
-\end{tikzcd}
-\end{equation*}
-Since pullbacks are uniquely unique, we get the desired isomorphism $\varphi_{f,P}$.
-
-It remains to prove the asserted isomorphism. 
-\begin{align*}
-& \{\theta:\ctxext{\Delta}{f_\ast(A)}\to\ctxext{{\Gamma}{A}}{P}\mid P\circ\theta\jdeq \pi_2(f,A)\} \\
-&  \cong
-\{\eta : \ctxext{\Delta}{f_\ast(A)}\to\ctxext{{\Delta}{f_\ast(A)}}{\pi_2(f,A)_\ast(P)} \mid
-  \pi_2(f,A)_\ast(P)\circ\eta\jdeq\catid{}\}
-  \\
-  & \cong
-\{\eta:\ctxext{\Delta}{f_\ast(A)}\to\ctxext{{\Delta}{(f_\ast(A))}}{f_\ast(P)}\mid f_\ast(P)\circ\eta\jdeq\catid{\ctxext{\Delta}{f_\ast(A)}}\}
-\end{align*}
-\end{proof}
-
-\begin{lem}
-Let $\mathbb{E}$ be an E'-system. Then there is a function $T(P)\to T(f_\ast(P))$
-for any $f:\Delta\to\Gamma$ in $\cat{C}$ and $P\in\cat{F}/\ctxext{\Gamma}{A}$, contravariant
-in $f$. 
-\end{lem}
-
-\begin{proof}
-Since $f_\ast(P)$ is isomorphic to $\pi_2(f,A)_\ast(P)$ in $\cat{C}/\ctxext{\Gamma}{A}$,
-it follows that the sets $T(f_\ast(P))$ and $T(\pi_2(f,A)_\ast(P))$ are isomorphic
-via composition with $\varphi_{f,P}$. By \autoref{lem:ptm}, we have a function
-from $T(P)$ to $T(\pi_2(f,A)_\ast(P))$.
-\end{proof}
-
-\begin{defn}
-For any $\Gamma\in\cat{C}$, define $\cat{C}_\cat{F}/\Gamma$ to be the full
-subcategory of $\cat{C}/\Gamma$ with the same objects as $\cat{F}/\Gamma$. 
-\end{defn}
-
-\begin{thm}
-Let $\mathbb{E}$ be an E'-system. Then $f_\ast$ can be extended to a functor
-$\cat{C}_\cat{F}/\Gamma\to\cat{C}_\cat{F}/\Delta$, for each $f:\Delta\to\Gamma$, and moreover the diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=large]
-\cat{C}_\cat{F}/\ctxext{\Gamma}{B} \arrow[r,"{f_\ast/B}"] \arrow[d,swap,"{g_\ast}"]
-& \cat{C}_\cat{F}/\ctxext{\Delta}{f_\ast(B)} \arrow[d,"{(f_\ast(g))_\ast}"] \\
-\cat{C}_\cat{F}/\ctxext{\Gamma}{A} \arrow[r,swap,"{f_\ast/A}"]
-& \cat{C}_\cat{F}/\ctxext{\Delta}{f_\ast(A)}
-\end{tikzcd}
-\end{equation*}
-commutes for every $g:A\to B$ in $\cat{C}_\cat{F}/\Gamma$.
-\end{thm}
-
-\begin{proof}
-For any $A,B\in\cat{F}/\Gamma$, we have an isomorphism
-$\mathrm{hom}_{\cat{C}/\Gamma}(A,B)\cong T(W_A(B))$. We also get a function
-$T(W_A(B))\to T(f_\ast(W_A(B)))$. Therefore, it suffices to show that
-$f_\ast(W_A(B))\cong W_{f_\ast(A)}(f_\ast(B))$. 
-\begin{align*}
-f_\ast(W_A(B))
-& \jdeq f_\ast(A_\ast(B)) \\
-& \cong \pi_2(f,A)_\ast(A_\ast(B)) \\
-& \jdeq (A\circ\pi_2(f,A))_\ast(B) \\
-& \jdeq (f\circ f_\ast(A))_\ast(B) \\
-& \jdeq f_\ast(A)_\ast(f_\ast(B)) \\
-& \jdeq W_{f_\ast(A)}(f_\ast(B)). \qedhere
-\end{align*}
-\end{proof}
-
-\begin{thm}
-E-systems are the same things as E'-systems.
-\end{thm}
-
-\begin{proof}
-Given an E-system $\mathbb{E}$, we have the category $\cat{C}$ of internal
-morphisms in the empty context, with the subcategory $\cat{F}$ of projections $\cprojfstf{A}{P}$. 
-The functors $f_\ast:\cat{F}/\Gamma\to\cat{F}/\Delta$ are contravariant in
-$f$, so we have to verify whether we have the stated isomorphisms.
-
-Note first that for any $A\in\cat{F}/\Gamma$, $W_A$ is
-contravariant in $A$ and that it preserves the terminal object.
-
-Now let $\theta:\Delta\to\ctxext{\Gamma}{A}$ so that $A\circ\theta\jdeq f$
-for some $f:\Delta\to\Gamma$. Then we have $\zeta_{A,f}(\theta)\in T(f_\ast(A))$.  
-We wish to define $\theta_\ast(\idtm{A}):\Delta\to\ctxext{\Delta}{f_\ast(A)}$
-
-for any $A\in\cat{F}/\Gamma$ and $Q\in\cat{F}/\ctxext{\Gamma}{B}$, we have
-\begin{align*}
-T(B) 
-  & \jdeq 
-\{ \eta : \Gamma \to\ctxext{\Gamma}{B}
-        \mid B\circ\eta\jdeq\catid{\Gamma}\}
-  \\
-T(W_A(B))
-  & \jdeq
-\{ \eta : \ctxext{\Gamma}{A}\to\ctxext{{\Gamma}{A}}{W_A(B)}
-        \mid W_A(B)\circ\eta\jdeq\catid{\ctxext{\Gamma}{A}}\}
-  \\
-  & \cong
-\{ \theta : \ctxext{\Gamma}{A}\to\ctxext{\Gamma}{B}\mid B\circ\theta\jdeq A\}
-\end{align*}
-Therefore, $\eta\mapsto \zeta_{B,A}(\eta\circ A)$ defines a function from
-$T(B)$ to $T(W_A(B))$. 
-\end{proof}
-
-\begin{comment}
-\section{Acknowledgments}
-The research that eventually became the study of E-systems began while analyzing
-the internel models I encountered during investigations of the type theoretical
-descent theorem, in the summer of 2013. In a private conversation during the Conference on Type Theory, 
-Homotopy Theory and Univalent Foundations at the Universitat Aut\`onoma de 
-Barcelona, I presented the first version of the axioms to Richard Garner, who
-told me he found a similar structure  while analyzing the combinatorics of type
-dependency, see \cite{Garner2014}. After that conference, Joachim
-Kock generously offered me to stay for a few months at the UAB. He was intrigued
-by the idea of doing category theory using only sections, and during the sessions
-we had in this period I finalized the inference rules that correspond to the
-axioms of E-systems.
-
-In Paris, in the spring of 2014, Richard Garner introduced me to Vladimir 
-Voevodsky, who had been working on similar problems and wanted to know exactly
-what axioms should hold on B0-systems so that it is in the image of the functor
-from C-systems to B0-systems. In other words, he wanted to know what the axioms
-of B-systems were, and be confident that we had them all.
-
-In August 2015, I started a grad student of Steve Awodey at the Carnegie 
-Mellon University. With Steve we elaborated much more on the fact that the theory
-of E-systems is essentially algebraic, and I wrote down what an E-system is in
-an arbitrary category with finite limits.
-
-The discussions that followed both with Awodey and with Voevodsky, 
-at the CMU, at the Homotopy Type Theory workshop in Oxford (November 2014), and
-while visiting Voevodsky at the Institute for Advanced Study in Princeton (March 2015), 
-essentially shaped E-systems into their current presentation.
-
-I am grateful to all the people mentioned for sharing many valuable insights 
-with me. 
-\end{comment}

### 2015-10-01 compared.tex
moved intro to new file

@@
 
-\subsection{From categories with families and comprehension categories to E-systems}
-
+\subsection{Comprehension categories and categories with families}
 In this section we recall Dybjer's notion of categories with families and we
@@
 \begin{defn}
+Let $p:\cat{E}\to\cat{B}$ be a functor and let $f:\Delta\to \Gamma$ be a morphism in $\cat{B}$.
+A morphism $g:A\to B$ in $\cat{E}$ is said to be \emph{cartesian over $f$}, if
+\begin{enumerate}
+\item $p(g)\jdeq f$, and 
+\item For all $g':A'\to B$ with $p(g')\jdeq f$, there is a unique $u:A'\to A$
+such that $g'=g\circ u$ and $p(u)\jdeq\catid{\Delta}$. 
+\end{enumerate}
+Dually, a morphism $g:A\to B$ in $\cat{E}$ is said to be \emph{cocartesian over
+$f$}, if
+\begin{enumerate}
+\item $p(g)\jdeq f$, and
+\item For all $g':A\to B'$ with $p(g')\jdeq f$, there is a unique $v:B\to B'$
+such that $g'=v\circ g$ and $p(v)\jdeq\catid{\Gamma}$. 
+\end{enumerate}
+A (co)cartesian morphism over $f$ is also called a \emph{(co)cartesian lift of $f$}.
+\end{defn}
+
+\begin{defn}
+A functor $p:\cat{E}\to\cat{B}$ is called a \emph{Grothendieck fibration} if
+every morphism into an object of the form $p(A)$ for some object $A$ of $\cat{E}$ 
+has a cartesian lift, and cartesian morphisms are closed under composition.
+\end{defn}
+
+\begin{defn}
+A Grothendieck fibration $p:\cat{E}\to\cat{B}$ is said to be \emph{split}, if
+there is a functorial choice of cartesian lifts, i.e.~for every $f:\Delta\to \Gamma$ in 
+$\cat{B}$ and any $B\in \cat{E}$ such that $p(B)\jdeq \Delta$,
+there is a cartesian lift $j(B,g):A \to B$ for some object $A$ of $\cat{E}$, such that 
+$j(B,p(B))\jdeq\catid{B}$ and $j(B, g\circ h)\jdeq j(B,g)\circ j(\mathrm{dom}(j(B,g)),h)$.
+\end{defn}
+
+\begin{eg}
+For any presheaf $P:\cat{C}\to\mathbb{Set}$, the projection functor
+$\int_\cat{C} P\to\cat{C}$ is a Grothendieck fibration.
+\end{eg}
+
+\begin{defn}
+A comprehension category $\mathcal{P}$ consists of categories $\cat{E}$, $\cat{C}$ and a
+commuting triangle
+\begin{equation*}
+\begin{tikzcd}
+\cat{E} \arrow[rr,"\mathcal{P}"] \arrow[dr,swap,"p"] & & \cat{C}^{\to} \arrow[dl,"\mathrm{cod}"] \\
+& \cat{C}
+\end{tikzcd}
+\end{equation*}
+of functors, such that $p$ is a Grothendieck fibration, and $\mathcal{P}$ maps
+cartesian morphisms in $\cat{E}$ to pullback squares in $\cat{C}$. We say that
+$\mathcal{P}$ is a full comprehension category if the functor $\mathcal{P}:
+\cat{E}\to\cat{C}^{\to}$ is full and faithful; and we say that
+$\mathcal{P}$ is split if $p$ is split.
+\end{defn}
+
+We now recall the notion of categories with families, which forms a class of
+examples of comprehension categories.
+
+\begin{defn}
 A \emph{category with families} consists of
@@
 
-\begin{defn}
-Let $\mathbb{D}$ be a D-system with underlying category $\cat{C}$. We define
-$\cat{F}\subseteq\cat{C}$ to be the subcategory with the same objects as
-$\cat{C}$, where the morphisms are generated by the projections $p_A$. Note that
-there is no reason why $\cat{F}$ should possess a terminal object.
-\end{defn}
-
-\begin{defn}
-Let $\mathbb{D}$ be a D-system. Then we define for any $\Gamma\in\mathbb{D}$ 
-the category $\Lambda(\Gamma)$ of which the objects are finite sequences
-of the form $(A_1,\ldots,A_n)$ where
-\begin{equation*}
-A_1\in\mathrm{Ty}(\Gamma),\ A_2\in\mathrm{Ty}(\ctxext{\Gamma}{A_1}),\ \ldots,\ 
-A_n\in\mathrm{Ty}(\ctxext{{{\Gamma}{A_1}}{\cdots}}{A_{n-1}})
-\end{equation*}
-where the morphisms are generated by $(A_1,\ldots,A_{n+1})\leq (A_1,\ldots,A_{n})$.
-\end{defn}
-
-\begin{defn}
-Let $\mathbb{D}$ be a D-system and let $f:\Delta\to\Gamma$ in $\mathbb{D}$.
-Then there is a functor $\Lambda(f):\Lambda(\Gamma)\to\Lambda(\Delta)$, with
-morphisms $\pi_2(f,A):\ctxext{\Delta}{p_{\Lambda(f,A)}}\to
-\ctxext{\Gamma}{p_A}$ in $\mathbb{D}$, such that the square
+\begin{eg}
+Since any morphism of $\int_{\cat{C}}$ is cartesian, it follows that D-systems
+are comprehension categories. As D-systems are built up from a presheaf of types
+on the underlying category, they are in fact discrete comprehension categories.
+\end{eg}
+
+The following two definitions define subclasses of closed D-systems and 
+coherent closed D-systems, and the latter structure is the same as an E'-system
+(without empty context). In the next section we will show how to close a comprehension
+category, so that it becomes a closed D-system which satisfies an additional coherence condition.
+
+\begin{defn}
+Let $\cat{C}$ be a D-system. In this definition, we will write
+$\mathrm{Fam}$ for the presheaf of types. The D-system $\cat{C}$
+is said to be \emph{closed}, if we have elements $\emptyf[\Gamma]\in
+\mathrm{Fam}(\Gamma)$ for each $\Gamma\in\cat{C}$, and maps 
+$P\mapsto\ctxext{A}{P}:\mathrm{Fam}(\ctxext{\Gamma}{A})\to \mathrm{Fam}(\Gamma)$
+such that
+\begin{enumerate}
+\item 
+\begin{enumerate}
+\item for any $A\in\mathrm{Fam}(\Gamma)$, $P\in\mathrm{Fam}(\ctxext{\Gamma}{A})$
+and $Q\in\mathrm{Fam}(\ctxext{{\Gamma}{A}}{P})$, one has
+\begin{equation*}
+\ctxext{{A}{P}}{Q}\jdeq\ctxext{A}{{P}{Q}},
+\end{equation*}
+\item and for any $A\in\mathrm{Fam}(\Gamma)$, one has
+\begin{equation*}
+\ctxext{\emptyf[\Gamma]}{A}\jdeq\ctxext{A}{\emptyf[\ctxext{\Gamma}{A}]}\jdeq A,
+\end{equation*}
+\end{enumerate}
+\item 
+\begin{enumerate}
+\item for any $A\in\mathrm{Fam}(\Gamma)$ and $P\in\mathrm{Fam}(\ctxext{\Gamma}{A})$,
+one has
+\begin{equation*}
+p_{\ctxext{A}{P}}\jdeq p_A\circ p_P,
+\end{equation*}
+\item and for any $\Gamma\in\cat{C}$ one has
+\begin{equation*}
+p_{\emptyf[\Gamma]}\jdeq\catid{\Gamma}
+\end{equation*}
+\end{enumerate}
+\item 
+\begin{enumerate}
+\item for any $A\in\mathrm{Fam}(\Gamma)$, $P\in\mathrm{Fam}(\ctxext{\Gamma}{A})$
+and any $f:\Delta\to\Gamma$ in $\cat{C}$, one has
+\begin{equation*}
+f_\ast(\ctxext{A}{P})\jdeq f_\ast(A)\circ \pi_2(f,A)_\ast(P),
+\end{equation*}
+\item and for any $f:\Delta\to\Gamma$ in $\cat{C}$, one has
+\begin{equation*}
+f_\ast(\emptyf[\Gamma])\jdeq\emptyf[\Delta].
+\end{equation*}
+\end{enumerate}
+\end{enumerate} 
+\end{defn}
+
+\begin{lem}
+A D-system is closed if and only if  the following two conditions hold:
+\begin{enumerate}
+\item There is a section $\phi:\cat{C}\to \int_{\cat{C}}\mathrm{Ty}$ of $\pi$.
+\item For every $P\in\mathrm{Ty}(\ctxext{\Gamma}{A})$,
+the morphism $p_A:\ctxext{\Gamma}{A}\to\Gamma$ has a unique cocartesian lift
+at $P$.
+\end{enumerate}
+\end{lem}
+
+\begin{proof}
+Let $\mathbb{D}$ be a closed D-system. To show that $\mathbb{D}$ has the two
+stated properties:
+\begin{enumerate}
+\item
+\item We have to show that each for $P\in\mathrm{Ty}(\ctxext{\Gamma}{A})$,
+the projection $p_A:\ctxext{\Gamma}{A}\to\Gamma$ has a unique cocartesian lift
+at $P$. Note that the morphism $(\ctxext{\Gamma}{A},P)\to (\Gamma,\ctxext{A}{P})$
+in $\int_{\cat{C}}\mathrm{Ty}$ 
+\end{enumerate}
+
+Let $\mathbb{D}$ be a D-system for which the stated two conditions hold. To show
+that $\mathbb{D}$ is a closed D-system, we systematically go through each of the
+properties:
+\begin{enumerate}
+\item 
+\end{enumerate} 
+\end{proof}
+
+\begin{defn}
+A closed D-system is said to be \emph{coherent}, if for any $A,A'\in\mathrm{Fam}(\Gamma)$
+such that $p_A\jdeq p_{A'}$, and any $f:\Delta\to\Gamma$, we have
+$p_{f_\ast(A)}\jdeq p_{f_\ast(A')}$ and $\pi_2(f,A)\jdeq \pi_2(f,A')$.
+\end{defn}
+
+\begin{thm}
+An E'-system without empty context is the same thing as a coherent closed D-system.
+\end{thm}
+
+\begin{proof}
+Let $\mathbb{E}$ be an E'-system without empty context. Then we define the
+presheaf $\mathrm{Fam}$ on $\cat{C}$ on objects by
+\begin{equation*}
+\mathrm{Fam}(\Gamma)\defeq \mathrm{Obj}(\cat{F}/\Gamma)
+\end{equation*}
+and on morphisms by the action on objects of $f_\ast$, for each $f:\Delta\to
+\Gamma$ in $\cat{C}$. By condition (b) and (c) in the definition of E'-systems, it
+follows that this defines a contravariant functor.
+
+We define a functor $\mathcal{P}:\int_{\cat{C}}\to\cat{C}^\square$ by sending
+the pair $(\Gamma,A)$ to the morphism $A:\ctxext{\Gamma}{A}\to\Gamma$ in
+$\cat{C}$, and by sending $f:(\Delta,f_\ast(A))\to (\Gamma,A)$ to the pullback
+square
 \begin{equation*}
 \begin{tikzcd}[column sep=large]
-\ctxext{\Delta}{\Lambda(f,A)} \arrow[r,"{\pi_2(f,A)}"] \arrow[d,fib] & \ctxext{\Gamma}{A} \arrow[d,fib] \\
-\Delta \arrow[r,swap,"f"] & \Gamma
+\ctxext{\Delta}{f_\ast(A)} \arrow[d,swap,"{f_\ast(A)}"] \arrow[r,"{\pi_2(f,A)}"] &
+\ctxext{\Gamma}{A} \arrow[d,"A"] \\ \Delta \arrow[r,"f"] & \Gamma
+\end{tikzcd}
+\end{equation*}
+This is also functorial because of conditions (b) and (c) in the definition of
+E'-systems. Thus, we have defined a D-system
+\begin{equation*}
+\begin{tikzcd}
+\int_{\cat{C}}\mathrm{Fam} \arrow[rr,"{\mathcal{P}}"] \arrow[dr,"{\pi}"] & & \cat{C}^\square \arrow[dl,"{\mathrm{cod}}"] \\ & \cat{C}
+\end{tikzcd}
+\end{equation*}
+and it is straightforward to verify that this D-system is closed and coherent.
+
+Conversely, suppose $\mathbb{D}$ is a closed D-system with underlying category
+$\cat{C}$. Define $\cat{F}$ to be the category with the same objects as
+$\cat{C}$, and morphisms of the form $p_A$. To check that this forms a category,
+note that $p_{\emptyf[\Gamma]}\jdeq\catid{\Gamma}$ and $p_A\circ p_P\jdeq
+p_{\ctxext{A}{P}}$.
+
+Now let $f:\Delta\to\Gamma$ be a morphism in $\cat{C}$. For any $A\in\cat{F}/\Gamma$,
+we have an element $\tilde{A}\in\mathrm{Fam}(\Gamma)$ such that $p_{\tilde{A}}
+\jdeq A$, and we can form a pullback square
+\begin{equation*}
+\begin{tikzcd}
+\ctxext{\Delta}{f_\ast(\tilde{A})} \arrow[r,"\tilde{A}"] \arrow[d,fib,swap,"p_{f_\ast(\tilde{A})}"]
+& \ctxext{\Gamma}{\tilde{A}} \arrow[d,fib,"{p_{\tilde{A}}}"] \\
+\Delta \arrow[r,"f"] & \Gamma
+\end{tikzcd}
+\end{equation*}
+By the coherence condition, this choice of pullback square does not depend on the
+choice of $\tilde{A}$ such that $p_{\tilde{A}}\jdeq A$ (having this fact 
+available is precisely what the coherence condition is for). The verification of conditions
+(a) to (d) of \autoref{defn:E'sys} is straightforward.
+\end{proof}
+
+\subsection{From comprehension categories to E-systems}
+To get from comprehension categories to E'-systems without empty contexts, we
+need to turn comprehension categories into closed D-systems.
+
+\begin{defn}
+Let $\mathcal{P}$ be a comprehension category. Then we define the pre-category
+$\cat{F}$ to be the category with the same objects as $\cat{C}$, of which the
+morphisms are generated by the morphisms of the form $P(A)$, for objects
+$A\in\cat{E}$. 
+\end{defn}
+
+\begin{defn}
+Let $\mathcal{P}$ be a comprehension category. Then we define for any 
+$\Gamma\in\cat{C}$, the poset $\Lambda(\Gamma)$ to be the category of finite 
+sequences  $A_1,\ldots,A_n$ of objects of $\cat{E}$, such that 
+$p(A_1)\jdeq\Gamma$ and $p(A_{i+1})\jdeq\mathrm{dom}(\mathcal{P}(A_i))$ 
+for any $1\leq i < n$. The partial order on 
+$\Lambda(\Gamma)$ is defined by $(A_1,\ldots,A_n)\leq (B_1,\ldots,B_m)$ if and 
+only if $n\leq m$ and $A_i\jdeq B_i$ for all $1\leq i\leq n$.
+
+The functor $\mathcal{P}$ induces a functor $\mathcal{P}:\Lambda(\Gamma)\to
+\cat{F}/\Gamma$ by $\mathcal{P}(A_1,\ldots,A_n)\jdeq\mathcal{P}(A_1)\circ
+\cdots\circ\mathcal{P}(A_n)$.
+\end{defn}
+
+\begin{defn}
+Let $\mathcal{P}$ be a split comprehension category with choice $j$ of cartesian
+lifts. We define a closed D-system
+\begin{equation*}
+\begin{tikzcd}
+\int_{\cat{C}}\Lambda \arrow[rr,"\bar{\mathcal{P}}"] \arrow[dr,swap,"\pi"] & &
+\cat{C}^\square \arrow[dl,"\mathrm{cod}"] \\ & \cat{C}
+\end{tikzcd}
+\end{equation*}
+\end{defn}
+
+\begin{constr}
+We first define, for any morphism $f_0\defeq f:\Delta\to\Gamma$ of $\cat{C}$,
+and for any $A\jdeq (A_1,\ldots,A_n)\in\Lambda(\Gamma)$,
+an element $\Lambda(f,A)\in\Lambda(\Delta)$ together with
+a morphism $q(f,A):\mathrm{dom}(\mathcal{P}(\Lambda(f,A)))\to 
+\mathrm{dom}(\mathcal{P}(A))$. We do this by induction on the length of $A$.
+
+For $n\jdeq 0$, i.e.~if $A$ is the empty sequence, we let $\Lambda(f,A)$ be the 
+empty sequence, and $q(f,A) \jdeq f$. For the inductive step, consider a 
+sequence $A_1,\ldots,A_{n+1}\in \Lambda(\Gamma)$. Then we define
+\begin{align*}
+\Lambda(f,A) & \jdeq \mathrm{dom}(j(A_{n+1},q(f,(A_1,\ldots,A_n))))
+\intertext{and}
+q(f,A) & \jdeq \mathrm{dom}(\mathcal{P}(j(A_{n+1},q(f,(A_1,\ldots,A_n)))))
+\end{align*}
+where we consider in the definition of $q(f,A)$, $\mathrm{dom}$ to be the 
+domain functor from $\cat{C}^{\to}\to\cat{C}$.
+
+It follows immediately from the definition of comprehension categories and the
+pasting lemma of pullbacks, that the square
+\begin{equation*}
+\begin{tikzcd}[column sep=large]
+\mathrm{dom}(\mathcal{P}(\Lambda(f,A))) \arrow[r,"{q(f,A)}"] \arrow[d,swap,"{\mathcal{P}(\Lambda(f,A))}"] & \mathrm{dom}(\mathcal{P}(A)) \arrow[d,"{\mathcal{P}(A)}"] \\
+\Delta \arrow[r,"f"] & \Gamma
 \end{tikzcd}
@@
 is a pullback square.
-\end{defn}
-
-\begin{constr}
-We construct the asserted structure by
-induction on the length of the arrow $A$ into $\Gamma$. We define
-$\Lambda(f)$ on the empty sequence as
-$\Lambda(f,\emptyf[\Gamma])\defeq \emptyf[\Delta]$, and we define
-$\pi_2(f,\emptyf[\Gamma])\defeq f:\ctxext{\Delta}{\catid{\Delta}}\to
-\ctxext{\Gamma}{\catid{\Gamma}}$. The resulting square is trivially
-a pullback square. Now suppose we have a pullback square as above.
-and let $B\in\mathrm{Ty}(\ctxext{\Gamma}{A})$. Then we define
-\begin{equation*}
-\Lambda(f,[A,B])\defeq [\Lambda(f,A),\pi_2(f,A)_\ast(B)].
-\end{equation*}
-By \autoref{thm:DD}, the top square in the diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=6em]
-\ctxext{{\Delta}{\Lambda(f,A)}}{(\pi_2(f,A)_\ast(B))} \arrow[d,fib] \arrow[r,"{\pi_2(\pi_2(f,A),B)}"] & \ctxext{{\Gamma}{A}}{B} \arrow[d,fib] \\
-\ctxext{\Delta}{\Lambda(f,A)} \arrow[d,fib] \arrow[r,"{\pi_2(f,A)}"] & \ctxext{\Gamma}{A} \arrow[d,fib] \\
+
+From the above construction we obtain a poset-valued presheaf $\Lambda:\op{\cat{C}}
+\to\mathbf{Pos}$, to which we can apply the Grothendieck construction to obtain
+the Grothendieck fibration $\pi:\int_{\cat{C}}\Lambda\to\cat{C}$. 
+\end{constr}
+
+\begin{comment}
+\begin{defn}
+Let $f_0\defeq f:\Delta\to\Gamma$ be a morphism of $\cat{C}$, the base category of a
+split comprehension category $\mathcal{P}$ wich a choice $j$ of cartesian lifts.
+By induction we define an element $\Lambda(f,A)\in\Lambda(\Delta)$ together with
+a morphism $q(f,A):\mathrm{dom}(\mathcal{P}(\Lambda(f,A)))\to 
+\mathrm{dom}(\mathcal{P}(A))$, for any $A\jdeq (A_1,\ldots,A_n) 
+\in\Lambda(\Gamma)$.
+
+For $n\jdeq 0$, i.e.~if $A$ is the empty sequence, we let $\Lambda(f,A)$ be the 
+empty sequence, and $q(f,A) \jdeq f$. For the inductive step, consider a 
+sequence $A_1,\ldots,A_{n+1}\in \Lambda(\Gamma)$. Then we define
+\begin{align*}
+\Lambda(f,A) & \jdeq \mathrm{dom}(j(A_{n+1},q(f,(A_1,\ldots,A_n))))
+\intertext{and}
+q(f,A) & \jdeq \mathrm{dom}(\mathcal{P}(j(A_{n+1},q(f,(A_1,\ldots,A_n)))))
+\end{align*}
+where we consider in the definition of $q(f,A)$, $\mathrm{dom}$ to be the 
+domain functor from $\cat{C}^{\to}\to\cat{C}$. 
+\end{defn}
+
+\begin{rmk}
+It follows immediately from the definition of comprehension categories and the
+pasting lemma of pullbacks, that the square
+\begin{equation*}
+\begin{tikzcd}[column sep=large]
+\mathrm{dom}(\mathcal{P}(\Lambda(f,A))) \arrow[r,"{q(f,A)}"] \arrow[d,swap,"{\mathcal{P}(\Lambda(f,A))}"] & \mathrm{dom}(\mathcal{P}(A)) \arrow[d,"{\mathcal{P}(A)}"] \\
 \Delta \arrow[r,"f"] & \Gamma
@@
 is a pullback square.
-Since the bottom square is assumed to be a pullback, we get that the outer rectangle
-is a pullback, as desired. 
-\end{constr}
-
-\begin{rmk}
-We can't directly use the functors $\Lambda(f)$ to give a D-system the structure
-of an E'-system, because there is a coherence condition missing.
 \end{rmk}
-
-\begin{thm}\label{thm:Dsys_coh}
-The following are equivalent:
+\end{comment}
+
+\begin{thm}
+Let $\mathcal{P}$ be a comprehension category. Then the following are equivalent:
 \begin{enumerate}
-\item The functors $\Lambda(f):\Lambda(\Gamma)\to\Lambda(\Delta)$ extend to 
-the structure of an E'-system without empty context.
-\item For any $f:\Delta\to\Gamma$ and all $A,A'\in\Lambda(\Gamma)$ satisfying 
-$p_A\jdeq p_{A'}$, one has
-\begin{equation*}
-p_{\Lambda(f,A)}\jdeq p_{\Lambda(f,A')}\qquad\text{and}\qquad\pi_2(f,A)\jdeq \pi_2(f,A').
-\end{equation*}
+\item The choice functors $\Lambda(f)$ and lifts $q(f)$ can be extended to the 
+structure of an E'-system without empty context.
+\item For any
+$A,A'\in\Lambda(\Gamma)$ such that $\mathcal{P}(A)\jdeq\mathcal{P}(A')$
+and for any $f:\Delta\to\Gamma$ in $\cat{C}$, the equalities
+$\mathcal{P}(\Lambda(f,A))\jdeq\mathcal{P}(\Lambda(f,A'))$ and
+$\pi_2(f,A)\jdeq\pi_2(f,A')$ hold. 
 \end{enumerate}
@@
 \begin{tikzcd}
-\Lambda(\Gamma) \arrow[d,swap,"p"] \arrow[r,"\Lambda(f)"] & \Lambda(\Delta) \arrow[d,"p"] \\
+\Lambda(\Gamma) \arrow[d,swap,"\mathcal{P}"] \arrow[r,"\Lambda(f)"] & \Lambda(\Delta) \arrow[d,"\mathcal{P}"] \\
 \cat{F}/\Gamma \arrow[r,densely dotted,"{f_\ast}"] & \cat{F}/\Delta
@@
 \end{equation*}
-of functors, forming an E'-system without empty context. Let 
-$f:\Delta\to\Gamma$, and let $A,A':\Lambda(\Gamma)$ be such that 
-$p_A\jdeq p_{A'}$. Then it follows
-that $\pi_2(f,A)_\ast\jdeq f_\ast/A\jdeq f_\ast/A'\jdeq\pi_2(f,A')_\ast$, so
-we get the desired equality $\pi_2(f,A)\jdeq\pi_2(f,A')$.
+of functors, forming an E'-system without empty context, where $\mathcal{P}(A)
+\defeq \mathcal{P}(A_1)\circ\cdots\circ\mathcal{P}(A_n)$ for any
+$A\jdeq (A_1,\ldots,A_n)\in\Lambda(\Gamma)$. Then it must hold that for any
+$A,A'\in\Lambda(\Gamma)$ such that $\mathcal{P}(A)\jdeq\mathcal{P}(A')$, we have
+$\mathcal{P}(\Lambda(f,A))\jdeq\mathcal{P}(\Lambda(f,A'))$. Also, for such
+$A$ and $A'$ it follows from the way $\Lambda(f)$ is defined, 
+that $\Lambda(\pi_2(f,A))\jdeq \Lambda(f)/A\jdeq
+\Lambda(f)/A' \jdeq \Lambda(\pi_2(f,A'))$, and therefore we also get the equalities
+\begin{equation*}
+\pi_2(f,A)\jdeq \pi_2(\pi_2(f,A),())\jdeq \pi_2(\pi_2(f,A'),()) \jdeq \pi_2(f,A'),
+\end{equation*}
+so the desired equality holds.
 
 For the converse, suppose that for all $f:\Delta\to\Gamma$ and $A,A'\in
-\Lambda(\Gamma)$ such that $p_A\jdeq p_{A'}$, one has 
-$p_{\Lambda(f,A)}\jdeq p_{\Lambda(f,A')}$ and $\pi_2(f,A)\jdeq\pi_2(f,A')$.
-Let $f:\Delta\to\Gamma$. By the assumption that $p_{\Lambda(f,A)}\jdeq p_{\Lambda(f,A')}$,
-it follows that the objects part of $p^\Delta\circ\Lambda(f):\Lambda(\Gamma)\to\cat{F}/\Delta$ 
-factors through $p^\Gamma:\Lambda(\Gamma)\to\cat{F}/\Gamma$. Moreover, by the
+\Lambda(\Gamma)$ such that $\mathcal{P}(A)\jdeq \mathcal{P}(A')$, one has 
+$\mathcal{P}(\Lambda(f,A))\jdeq \mathcal{P}(\Lambda(f,A'))$ and $\pi_2(f,A)\jdeq\pi_2(f,A')$.
+Let $f:\Delta\to\Gamma$. By the assumption that $\mathcal{P}(\Lambda(f,A))\jdeq \mathcal{P}(\Lambda(f,A'))$,
+it follows that the objects part of $\mathcal{P}\circ\Lambda(f):\Lambda(\Gamma)\to\cat{F}/\Delta$ 
+factors through $\mathcal{P}:\Lambda(\Gamma)\to\cat{F}/\Gamma$. Moreover, by the
 assumption that $\pi_2(f,A)\jdeq \pi_2(f,A')$ it follows that the pullback
@@
 \begin{defn}
-If either of the equivalent conditions of \autoref{thm:Dsys_coh} hold, we say 
-that the D-system is \emph{coherent}. Thus, we have a functor
-$U:\mathbf{Dsys}\to\mathbf{E'sys_{\circ}}$ from coherent D-systems to E'-systems
-\end{defn}
-
-\begin{defn}
-We say that a D-system $\cat{D}$ is \emph{locally well-founded} if the obvious functor
-$p:\Lambda(\mathbb{D})\to\cat{F}/\Gamma$ is an isomorphism of pre-categories.
-\end{defn}
-
-\begin{rmk}
-In other words, a D-system is locally well-founded if and only if
-for any $A\in\cat{F}$ there is a unique sequence
-$A_1\in\mathrm{Ty}(\Gamma)$, $A_2\in\mathrm{Ty}(\ctxext{\Gamma}{A_1})$, \ldots,
-$A_n\in\mathrm{Ty}(\ctxext{{{\Gamma}{A_1}}{\cdots}}{A_{n-1}})$ such that
-$A\jdeq p_{A_1}\circ\cdots\circ p_{A_n}$. Locally well-founded D-systems 
-are examples of coherent D-systems.
-\end{rmk}
-
-\begin{thm}
-The restriction of the functor $U:\mathbf{Dsys}\to\mathbf{E'sys_{\circ}}$ to the
-locally well-founded D-systems is full and faithful.
-\end{thm}
-
-\begin{proof}
-The functor $U$ is faithful on the locally well-founded D-systems, because for any morphism 
-$\mathcal{F}:\mathbb{D}\to\mathbb{D}'$ of locally well-founded D-systems, 
-$U(\mathcal{F})$ and $\mathcal{F}$ have the same underlying functor
-from $\cat{C}$ to $\cat{C}'$.
-
-It remains to show that $U$ is full on the locally well-founded D-systems. Let
-$\mathbb{D}$ and $\mathbb{D}$ be locally well-founded D-systems and let $\mathcal{G}:U(\mathbb{D})
-\to U(\mathbb{D}')$. Then $\mathcal{G}$ restricted to $\cat{F}$ lifts factorizations.
-We  first have to define a natural transformation
-$\mathrm{Ty}\Rightarrow \mathrm{Ty}'\circ\mathcal{G}$. Let
-$\Gamma\in\cat{C}$ and $A\in\mathrm{Ty}(\Gamma)$. Then we have
-$\mathcal{G}(p_A):\mathcal{G}(\ctxext{\Gamma}{A})\to \mathcal{G}(\Gamma)$.
-The category of factorizations of $p_A$ in $\cat{F}$ is a isomorphic to $\mathbf{2}$,
-so the category of factorizations of $\mathcal{G}(p_A)$ is also isomorphic to
-$\mathbf{2}$. This implies that $\mathcal{G}(p_A)$ is of the form $p_B$ for
-some $B\in\mathrm{Ty}'(\mathcal{G}(\Gamma))$. This $B$ is called $\mathcal{G}(A)$.
-
-To verify the naturality, let $f:\Delta\to\Gamma$. We need to show that the
-square
-\begin{equation*}
-\begin{tikzcd}
-\mathrm{Ty}(\Delta) \arrow[r,"{\mathcal{G}_\Delta}"] & \mathrm{Ty}'(\mathcal{G}(\Delta)) \\
-\mathrm{Ty}(\Gamma) \arrow[u,"\mathrm{Ty}(f)"] \arrow[r,"{\mathcal{G}_\Gamma}"] &
-\mathrm{Ty}'(\mathcal{G}(\Gamma)) \arrow[u,swap,"{\mathrm{Ty}'(\mathcal{G}(f))}"]
-\end{tikzcd}
-\end{equation*}
-commutes. Let $A\in\mathrm{Ty}(\Gamma)$. Then we need to show that
-$\mathcal{G}(f)_\ast(\mathcal{G}(A))\jdeq \mathcal{G}(f_\ast(A))$. By the uniqueness
-assumption of well-founded D-systems, it suffices to show that
-$p_{\mathcal{G}(f)_\ast(\mathcal{G}(A))}\jdeq p_{\mathcal{G}(f_\ast(A))}$. To
-see this, we do a simple calculation
-\begin{align*}
-p_{\mathcal{G}(f)_\ast(\mathcal{G}(A))}
-& \jdeq \mathcal{G}(f)_\ast(p_{\mathcal{G}(A)}) \\
-& \jdeq \mathcal{G}(f)_\ast(\mathcal{G}(p_A)) \\
-& \jdeq \mathcal{G}(f_\ast(p_A)) \\
-& \jdeq \mathcal{G}(p_{f_\ast(A)}) \\
-& \jdeq p_{\mathcal{G}(f_\ast(A))} \qedhere
-\end{align*}
-\end{proof}
-
-\begin{thm}
-The image of $U$ is the subcategory of locally stratified E'-systems without
-empty context. 
-\end{thm}
-
-\begin{proof}
-For any $\Gamma\in\mathbb{D}$, there is an obvious stratification functor
-$\Lambda(\Gamma)\to (\N,\geq)$, so $U(\mathbb{D})$ is a locally stratified
-E'-system.
-
-Now let $\mathbb{E}$ be a locally stratified E'-system without empty context.
-For $\Gamma\in\cat{C}$, we define $\mathrm{Ty}(\Gamma)$ to be the set of
-families of length one. Since any $f_\ast$ is stratified, it follows that
-$\mathrm{Ty}$ is a presheaf on $\cat{C}$. The choice of pullback squares
-\begin{equation*}
-\begin{tikzcd}[column sep=large]
-\ctxext{\Delta}{f_\ast(A)} \arrow[r,"{\pi_2(f,A)}"] \arrow[d,fib] & \ctxext{\Gamma}{A} \arrow[d,fib] \\
-\Delta \arrow[r,swap,"f"] & \Gamma
-\end{tikzcd}
-\end{equation*}
-is functorial by assumption, so we have defined a category with families. It
-is straigtforward to verify that this category with families is locally well-founded,
-and that its construction is inverse to $U$. 
-\end{proof}
-
-\begin{comment}
-\subsection{The equivalence of E-systems and E'-systems}
-
-The definition of D-system suggests a more general structure related to E-systems.
-
-\begin{defn}
-An \emph{E'-system} consists of a category $\cat{C}$ with a subcategory $\cat{F}$ with the
-same objects as $\cat{C}$, with
-the following structure
-\begin{enumerate}
-\item $\cat{F}$ has a terminal object.
-\item For any $f:\Delta\to\Gamma$ in $\cat{C}$, a functor $f_\ast:\cat{F}/\Gamma\to
-\cat{F}/\Delta$ %and
-%for any $Q\in\cat{F}/\ctxext{\Gamma}{B}$ and any section $g$ of $Q$ a section
-%$f_\ast(g)$ of $f_\ast(Q)$, 
-so that $f_\ast$ is contravariant in $f$ and preserves the terminal object. 
-\item For any $A\in\cat{F}/\Gamma$ and $f:\Delta\to\Gamma$, lifts of
-$f$ to $A$ are naturally the same thing as sections of
-$f_\ast(A)$, i.e. there is an isomorphism
-\begin{equation*}
-\zeta_{A,f}%\defeq\theta\mapsto f_\ast(\theta)
-:\{\theta:\Delta\to\ctxext{\Gamma}{A}\mid A\circ\theta\jdeq f\}
-  \cong
-\{\eta:\Delta\to\ctxext{\Delta}{(f_\ast(A))}\mid f_\ast(A)\circ\eta\jdeq\catid{\Delta}\}
-\end{equation*}
-which are natural in $A$ and $f$, in the sense that
-\begin{enumerate}
-\item For any $P\in\cat{F}/\ctxext{\Gamma}{A}$, $f:\Delta\to\Gamma$ and
-$\theta:\Delta\to\ctxext{{\Gamma}{A}}{P}$ such that $(\ctxext{A}{P})\circ\theta\jdeq f$,
-one hase
-\begin{equation*}
-\zeta_{A,f}(P\circ\theta)\jdeq f_\ast(P)\circ \zeta_{\ctxext{A}{P},f}(\theta).
-\end{equation*}
-\item For any $A\in\cat{F}/\Gamma$, $f:\Delta\to\Gamma$, $g:X\to\Delta$
-and $\theta:\Delta\to\ctxext{\Gamma}{A}$ such that $A\circ\theta\jdeq f$, one
-has
-\begin{align*}
-\zeta_{A,\catid{\Gamma}}(\theta) & \jdeq \theta \\
-\zeta_{A,f\circ g}(\theta\circ g) & \jdeq\zeta_{f_\ast(A),g}(\zeta_{A,f}(\theta)\circ g).
-\end{align*}
-\end{enumerate}
-\end{enumerate}
-\end{defn}
-
-We can give the category $\cat{F}$ a term structure as follows:
-
-\begin{defn}
-For any E'-system $\mathbb{E}$, the category $\cat{F}$ has a term structure
-where the terms of $A$ are sections of $A$, i.e.
-\begin{equation*}
-T(A)\defeq\{\theta:\Gamma\to\ctxext{\Gamma}{A}\mid A\circ\theta\jdeq\catid{\Gamma}\}.
-\end{equation*}
-\end{defn}
-
-\begin{lem}\label{lem:ptm}
-Let $\mathbb{E}$ be an E'-system. Then there is a function $T(A)\to T(f_\ast(A))$
-for any $f:\Delta\to\Gamma$ in $\cat{C}$ and $A\in\cat{F}/\Gamma$, contravariant
-in $f$. 
-\end{lem}
-
-\begin{proof}
-Using the natural isomorphism $\zeta_{A,f}$, we see that
-\begin{align*}
-T(f_\ast(A)) & \jdeq \{\eta : \Delta\to\ctxext{\Delta}{f_\ast(A)}\mid f_\ast(A)\circ\eta\jdeq\catid{\Delta}\}
-  \\
-  & \cong
-\{\theta : \Delta\to \ctxext{\Gamma}{A}\mid A\circ\theta\jdeq f\}.
-\end{align*}
-Therefore, we have the function $x\mapsto \zeta_{A,f}(x\circ f)$ from $T(A)$ to
-$T(f_\ast(A))$. 
-\end{proof}
-
-\begin{defn}
-For any $A\in\cat{F}/\Gamma$, we define
-$W_A\defeq A_\ast:\cat{F}/\Gamma\to\cat{F}/\ctxext{\Gamma}{A}$. Then we get
-for any $A,B\in\cat{F}/\Gamma$ an isomorphism
-\begin{equation*}
-\mathrm{hom}_{\cat{C}/\Gamma}(A,B)
-  \cong
-T(W_A(B)).
-\end{equation*}
-In particular, we get a term $\idtm{A}\in T(W_A(A))$. 
+There is a functor $\mathcal{V}:\mathbf{Comp}\to\mathbf{E'sys_{\circ}}$, from
+the category $\mathbf{Comp}$ of split full comprehension categories to
+E'-systems without empty context.
 \end{defn}
@@
 \begin{constr}
-Let $A,B\in\cat{F}/\Gamma$. Then we have the isomorphism
-\begin{align*}
-T(W_A(B)) & \jdeq \{\eta:\ctxext{\Gamma}{A}\to\ctxext{{\Gamma}{A}}{W_A(B)}\mid W_A(B)\circ\eta\jdeq\catid{\ctxext{\Gamma}{A}}\} \\
-  & \cong
-\{\theta:\ctxext{\Gamma}{A}\to\ctxext{\Gamma}{B}\mid B\circ\theta\jdeq A\},
-\end{align*}
-so we find $\idtm{A}\defeq\zeta_{A,A}(\catid{\ctxext{\Gamma}{A}})\in T(W_A(A))$.
-\end{constr}
-
-\begin{defn}
-For any $P\in\cat{F}/\ctxext{\Gamma}{A}$ and $f:\Delta\to\Gamma$ 
-there is a morphism $\pi_2(f,A):\ctxext{\Delta}{f_\ast(A)}\to\ctxext{\Gamma}{A}$
-such that $A\circ \pi_2(f,A)\jdeq f\circ f_\ast(A)$.
-\end{defn}
-
-\begin{constr}
-\begin{align*}
-& \{ \theta : \ctxext{\Delta}{f_\ast(A)}\to\ctxext{\Gamma}{A}\mid A\circ\theta\jdeq f\circ f_\ast(A) \} \\
-  & \cong
-\{ \eta : \ctxext{\Delta}{f_\ast(A)}\to\ctxext{{\Delta}{f_\ast(A)}}{(f\circ f_\ast(A))_\ast(A)}
-  \mid (f\circ f_\ast(A))_\ast(A)\circ\eta\jdeq\catid{\ctxext{\Delta}{f_\ast(A)}}\} \\
-  & \jdeq
-\{ \eta : \ctxext{\Delta}{f_\ast(A)}\to\ctxext{{\Delta}{f_\ast(A)}}{(f_\ast(A))_\ast(f_\ast(A))}
-  \mid (f\circ f_\ast(A))_\ast(A)\circ\eta\jdeq\catid{\ctxext{\Delta}{f_\ast(A)}}\} \\
-  & \jdeq
-T(W_{f_\ast(A)}(f_\ast(A))).
-\end{align*}
-Since we have $\tfid{f_\ast(A)}\in T(W_{f_\ast(A)}(f_\ast(A)))$, we obtain 
-\begin{equation*}
-\pi_2(f,A)\defeq \zeta_{A,f\circ f_\ast(A)}^{-1}(\tfid{f_\ast(A)}):\ctxext{\Delta}{f_\ast(A)}\to\ctxext{\Gamma}{A}
-\end{equation*}
-satisfying $A\circ\pi_2(f,A)\jdeq f\circ f_\ast(A)$.
-\end{constr}
-
-\begin{lem}
-For any $f:\Delta\to\Gamma$, $g:X\to\Delta$ and $A\in\cat{F}/\Gamma$, one has
-\begin{equation*}
-\pi_2(f,A)\circ\pi_2(g,f_\ast(A))\jdeq\pi_2(f\circ g,A).
-\end{equation*}
-\end{lem}
-
-\begin{proof}
-It suffices to show that
-\begin{equation*}
-\zeta_{A,f\circ g\circ g_\ast(f_\ast(A))}(\pi_2(f,A)\circ\pi_2(g,f_\ast(A)))\jdeq\tfid{g_\ast(f_\ast(A))}.
-\end{equation*}
-Because $g\circ g_\ast(f_\ast(A))\jdeq f_\ast(A)\circ\pi_2(g,f_\ast(A))$, we get
-to use the naturality of $\zeta$:
-\begin{align*}
-& \zeta_{A,f\circ g\circ g_\ast(f_\ast(A))}(\pi_2(f,A)\circ\pi_2(g,f_\ast(A))) \\
-& \jdeq
-\zeta_{A,f\circ f_\ast(A)\circ\pi_2(g,f_\ast(A))}(\pi_2(f,A)\circ\pi_2(g,f_\ast(A))) \\
-& \jdeq
-\zeta_{f_\ast(A)_\ast(f_\ast(A)),\pi_2(g,f_\ast(A))}(\zeta_{A,f\circ f_\ast(A)}(\pi_2(f,a))\circ\pi_2(g,f_\ast(A))) \\
-& \jdeq
-\zeta_{f_\ast(A)_\ast(f_\ast(A)),\pi_2(g,f_\ast(A))}(\tfid{f_\ast(A)}\circ\pi_2(g,f_\ast(A))) \\
-& \jdeq
-\zeta_{f_\ast(A)_\ast(f_\ast(A)),\pi_2(g,f_\ast(A))}(\zeta_{f_\ast(A),f_\ast(A)}(\catid{\ctxext{\Delta}{f_\ast(A)}})\circ\pi_2(g,f_\ast(A))) \\
-& \jdeq
-\zeta_{f_\ast(A),f_\ast(A)\circ\pi_2(g,f_\ast(A))}(\catid{\ctxext{\Delta}{f_\ast(A)}}\circ\pi_2(g,f_\ast(A))) \\
-& \jdeq
-\zeta_{f_\ast(A),f_\ast(A)\circ\pi_2(g,f_\ast(A))}(\pi_2(g,f_\ast(A))) \\
-& \jdeq
-\zeta_{f_\ast(A),g\circ g_\ast(f_\ast(A))}(\pi_2(g,f_\ast(A)))\\
-& \jdeq
-\tfid{g_\ast(f_\ast(A))}.\qedhere
-\end{align*}
-\end{proof}
-
-\begin{lem}
-Let $f:\Delta\to\Gamma$ and $A\in\cat{F}/\Gamma$. Then one has 
-\begin{equation*}
-\pi_2(f,A)\circ\zeta_{A,f}(\theta)\jdeq\theta
-\end{equation*}
-for any $\theta:\Delta\to\ctxext{\Gamma}{A}$ satisfying $A\circ\theta\jdeq f$.
-\end{lem}
-
-\begin{proof}
-We will write $\tilde\theta\defeq\zeta_{A,f}(\theta)$. Note that
-\begin{equation*}
-A\circ\pi_2(f,A)\circ\tilde\theta\jdeq f\circ f_\ast(A)\circ\tilde\theta\jdeq f.
-\end{equation*} 
-Therefore, we can prove that $\pi_2(f,A)\circ\tilde\theta\jdeq\theta$ by showing
-that $\zeta_{A,f}(\pi_2(f,A)\circ\tilde\theta)\jdeq\tilde\theta$. This is
-shown using the naturality of $\zeta$:
-\begin{align*}
-\zeta_{A,f}(\pi_2(f,A)\circ\tilde\theta)
-  & \jdeq
-\zeta_{A,f\circ f_\ast(A)\circ\tilde\theta}(\pi_2(f,A)\circ\tilde\theta)
-  \\
-  & \jdeq
-\zeta_{(f_\ast(A))_\ast(f_\ast(A)),\tilde{\theta}}(\zeta_{A,f\circ f_\ast(A)}(\pi_2(f,A))\circ\tilde\theta)
-  \\
-  & \jdeq
-\zeta_{(f_\ast(A))_\ast(f_\ast(A)),\tilde{\theta}}(\idtm{f_\ast(A)}\circ\tilde\theta)
-  \\
-  & \jdeq
-\zeta_{(f_\ast(A))_\ast(f_\ast(A)),\tilde{\theta}}(\zeta_{f_\ast(A),f_\ast(A)}(\catid{\ctxext{\Delta}{f_\ast(A)}})\circ\tilde\theta)
-  \\
-  & \jdeq
-\zeta_{f_\ast(A),f_\ast(A)\circ\tilde\theta}(\catid{\ctxext{\Delta}{f_\ast(A)}}\circ\tilde\theta) \\
-  & \jdeq
-\zeta_{f_\ast(A),\catid{\Delta}}(\tilde\theta) \\
-  & \jdeq
-\theta.\qedhere
-\end{align*}
-\end{proof}
-
-\begin{lem}\label{lem:EE_pb}
-The diagram
+Let $\mathcal{P}$ be a comprehension category with underlying category
+$\cat{C}$. We define $\cat{F}$ to be the sub-pre-category of $\cat{C}$ generated
+by the morphisms $\mathcal{P}(A):\ctxext{\Gamma}{A}\to\Gamma$ for $A\in\cat{E}$.
+
+For any $f:\Delta\to\Gamma$, we define a functor $f_\ast:\cat{F}/\Gamma\to
+\cat{F}/\Delta$ by induction on the length of the morphism
+Let $f:\Delta\to\Gamma$ and $p(A)\jdeq\Gamma$. Then we get a pullback square
 \begin{equation*}
@@
 \end{equation*}
+There seems to be no way to show that
+$\pi_2(f,\ctxext{A}{P})\jdeq \pi_2(\pi_2(f,A),P)$.
+\end{constr}
+
+\begin{conj}
+An E'-system is the same thing as a split full comprehension category in which
+$\cat{E}$ has a terminal object.
+\end{conj}
+
+\begin{proof}
+Let $\mathcal{E}$ be an E'-system. Define $\cat{E}$ to be the full subcategory
+of $\cat{C}^\to$ whose objects are the morphisms of $\cat{F}$. Thus, for any
+$\Gamma\in\cat{C}$, the category $p^{-1}(\Gamma)$ is the category of internal
+morphisms the E-system $\mathcal{E}/\Gamma$.
+
+We first show that $p$ is a Grothendieck fibration. Let $f:\Delta\to\Gamma$ be
+a morphism in $\cat{C}$ and let $A\in\cat{F}/\Gamma$. Then we have $\catid{\Delta},\catid{\Gamma}\in\cat{E}$,
+and $(f,\pi_2(f,A)):f_\ast(A)\to A$. It is routine to verify that $(f,\pi_2(f,A))$
+is a cartesian lift of $f$, and that this choice of catesian lifts is functorial.
+
+Now let $g:X\to\Delta$ and $f:\Delta\to \Gamma$, and let $\tilde{f}:\tilde{\Delta}\to\tilde{\Gamma}$ and
+$\tilde{g}:\tilde{X}\to\tilde{\Delta}$ be cartesian lifts of $f$ and $g$, respectively. We have to show
+that $\tilde{f}\circ \tilde{g}$ is a cartesian lift of $f\circ g$. Let $C\in\cat{F}/X$ and
+let $h : C\to \tilde{\Gamma}$ be a morphism such that $\mathrm{cod}(h)\jdeq f\circ g$. 
+\end{proof}
+
+\subsection{From categories with families and comprehension categories to E-systems}
+
+\begin{defn}
+Let $\mathbb{D}$ be a D-system with underlying category $\cat{C}$. We define
+$\cat{F}\subseteq\cat{C}$ to be the subcategory with the same objects as
+$\cat{C}$, where the morphisms are generated by the projections $p_A$. Note that
+there is no reason why $\cat{F}$ should possess a terminal object.
+\end{defn}
+
+\begin{defn}
+Let $\mathbb{D}$ be a D-system. Then we define for any $\Gamma\in\mathbb{D}$ 
+the category $\Lambda(\Gamma)$ of which the objects are finite sequences
+of the form $(A_1,\ldots,A_n)$ where
+\begin{equation*}
+A_1\in\mathrm{Ty}(\Gamma),\ A_2\in\mathrm{Ty}(\ctxext{\Gamma}{A_1}),\ \ldots,\ 
+A_n\in\mathrm{Ty}(\ctxext{{{\Gamma}{A_1}}{\cdots}}{A_{n-1}})
+\end{equation*}
+where the morphisms are generated by $(A_1,\ldots,A_{n+1})\leq (A_1,\ldots,A_{n})$.
+\end{defn}
+
+\begin{defn}
+Let $\mathbb{D}$ be a D-system and let $f:\Delta\to\Gamma$ in $\mathbb{D}$.
+Then there is a functor $\Lambda(f):\Lambda(\Gamma)\to\Lambda(\Delta)$, with
+morphisms $\pi_2(f,A):\ctxext{\Delta}{p_{\Lambda(f,A)}}\to
+\ctxext{\Gamma}{p_A}$ in $\mathbb{D}$, such that the square
+\begin{equation*}
+\begin{tikzcd}[column sep=large]
+\ctxext{\Delta}{\Lambda(f,A)} \arrow[r,"{\pi_2(f,A)}"] \arrow[d,fib] & \ctxext{\Gamma}{A} \arrow[d,fib] \\
+\Delta \arrow[r,swap,"f"] & \Gamma
+\end{tikzcd}
+\end{equation*}
 is a pullback square.
-\end{lem}
-
-\begin{proof}
-To verify the universal mapping property of the pullback square, consider a
-commuting diagram
-\begin{equation*}
-\begin{tikzcd}
-X \arrow[d,swap,"g"] \arrow[r,"h"] & \ctxext{\Gamma}{A} \arrow[d,fib] \\
+\end{defn}
+
+\begin{constr}
+We construct the asserted structure by
+induction on the length of the arrow $A$ into $\Gamma$. We define
+$\Lambda(f)$ on the empty sequence as
+$\Lambda(f,\emptyf[\Gamma])\defeq \emptyf[\Delta]$, and we define
+$\pi_2(f,\emptyf[\Gamma])\defeq f:\ctxext{\Delta}{\catid{\Delta}}\to
+\ctxext{\Gamma}{\catid{\Gamma}}$. The resulting square is trivially
+a pullback square. Now suppose we have a pullback square as above.
+and let $B\in\mathrm{Ty}(\ctxext{\Gamma}{A})$. Then we define
+\begin{equation*}
+\Lambda(f,[A,B])\defeq [\Lambda(f,A),\pi_2(f,A)_\ast(B)].
+\end{equation*}
+By \autoref{thm:DD}, the top square in the diagram
+\begin{equation*}
+\begin{tikzcd}[column sep=6em]
+\ctxext{{\Delta}{\Lambda(f,A)}}{(\pi_2(f,A)_\ast(B))} \arrow[d,fib] \arrow[r,"{\pi_2(\pi_2(f,A),B)}"] & \ctxext{{\Gamma}{A}}{B} \arrow[d,fib] \\
+\ctxext{\Delta}{\Lambda(f,A)} \arrow[d,fib] \arrow[r,"{\pi_2(f,A)}"] & \ctxext{\Gamma}{A} \arrow[d,fib] \\
 \Delta \arrow[r,"f"] & \Gamma
@@
 \end{equation*}
-Then $h$ corresponds uniquely to a section $k$ of $(f\circ g)_\ast(A)$. Since
-$(f\circ g)_\ast(A)\jdeq g_\ast(f_\ast(A))$, the section $k$ fits in the
-commuting diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=huge]
-\ctxext{X}{g_\ast(f_\ast(A))} \arrow[d,xshift=.7ex,fib] \arrow[r,"{\pi_2(g,f_\ast(A))}"]
-& \ctxext{\Delta}{f_\ast(A)} \arrow[d,fib] \arrow[r,"{\pi_2(f,A)}"]
-& \ctxext{\Gamma}{A} \arrow[d,fib] \\
-X \arrow[r,swap,"g"] \arrow[u,xshift=-.7ex,"k"] & \Delta \arrow[r,swap,"f"] & \Gamma
-\end{tikzcd}
-\end{equation*}
-so we obtain $u\defeq\pi_2(g,f_\ast(A))\circ k:X\to\ctxext{\Delta}{f_\ast(A)}$. 
-Then we have $f_\ast(A)\circ u\jdeq g\circ g_\ast(f_\ast(A))\circ k\jdeq g$ and
-$\pi_2(f,A)\circ u\jdeq \pi_2(f\circ g,A)\circ k\jdeq h$. Note that
-$u$ corresponds uniquely to $k$, which corresponds uniquely to $h$, so the
-uniqueness of $u$ is automatic.
-\end{proof}
-
-\begin{lem}
-For any $f:\Delta\to\Gamma$ and $P\in\cat{F}/\ctxext{\Gamma}{A}$, there is a unique isomorphism
-$\varphi_{f,P}:\pi_2(f,A)_\ast(P)\cong (f_\ast/A)(P)$ in 
-$\mathbb{C}/\ctxext{\Delta}{f_\ast(A)}$ such that
-\begin{equation*}
-\pi_2(f,\ctxext{A}{P})\jdeq \pi_2(\pi_2(f,A),P)\circ\varphi_{f,P}.
-\end{equation*}
-Consequently, we get isomorphisms
-\begin{align*}
-\{\theta:\ctxext{\Delta}{f_\ast(A)}\to\ctxext{{\Gamma}{A}}{P}\mid P\circ\theta\jdeq \pi_2(f,A)\}
-&  \cong
-\{\eta:\ctxext{\Delta}{f_\ast(A)}\to\ctxext{{\Delta}{(f_\ast(A))}}{f_\ast(P)}\mid f_\ast(P)\circ\eta\jdeq\catid{\ctxext{\Delta}{f_\ast(A)}}\}
-\end{align*}
-natural in $P$ and $f$.
-\end{lem}
-
-\begin{proof}
-Note that both the lower square and the outer rectangle in the diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=huge]
-\ctxext{{\Delta}{f_\ast(A)}}{f_\ast(P)} \arrow[d,fib] \arrow[r,"{\pi_2(f,\ctxext{A}{P})}"]
-& \ctxext{{\Gamma}{A}}{P} \arrow[d,fib] \\
-\ctxext{\Delta}{f_\ast(A)} \arrow[d,fib] \arrow[r,"{\pi_2(f,A)}"] 
-& \ctxext{\Gamma}{A} \arrow[d,fib] \\
-\Delta \arrow[r,swap,"f"] & \Gamma
-\end{tikzcd}
-\end{equation*}
-are pullback squares. By the pasting lemma for pullbacks, it follows that the
-top square must also be a pullback. On the other hand, we have the pullback
-square
-\begin{equation*}
-\begin{tikzcd}
-\ctxext{{\Delta}{f_\ast(A)}}{\pi_2(f,A)_\ast(P)} \arrow[d,fib] \arrow[r]
-& \ctxext{{\Gamma}{A}}{P} \arrow[d,fib] \\
-\ctxext{\Delta}{f_\ast(A)} \arrow[r,swap,"{\pi_2(f,A)}"]
-& \ctxext{\Gamma}{A}
-\end{tikzcd}
-\end{equation*}
-Since pullbacks are uniquely unique, we get the desired isomorphism $\varphi_{f,P}$.
-
-It remains to prove the asserted isomorphism. 
-\begin{align*}
-& \{\theta:\ctxext{\Delta}{f_\ast(A)}\to\ctxext{{\Gamma}{A}}{P}\mid P\circ\theta\jdeq \pi_2(f,A)\} \\
-&  \cong
-\{\eta : \ctxext{\Delta}{f_\ast(A)}\to\ctxext{{\Delta}{f_\ast(A)}}{\pi_2(f,A)_\ast(P)} \mid
-  \pi_2(f,A)_\ast(P)\circ\eta\jdeq\catid{}\}
-  \\
-  & \cong
-\{\eta:\ctxext{\Delta}{f_\ast(A)}\to\ctxext{{\Delta}{(f_\ast(A))}}{f_\ast(P)}\mid f_\ast(P)\circ\eta\jdeq\catid{\ctxext{\Delta}{f_\ast(A)}}\}
-\end{align*}
-\end{proof}
-
-\begin{lem}
-Let $\mathbb{E}$ be an E'-system. Then there is a function $T(P)\to T(f_\ast(P))$
-for any $f:\Delta\to\Gamma$ in $\cat{C}$ and $P\in\cat{F}/\ctxext{\Gamma}{A}$, contravariant
-in $f$. 
-\end{lem}
-
-\begin{proof}
-Since $f_\ast(P)$ is isomorphic to $\pi_2(f,A)_\ast(P)$ in $\cat{C}/\ctxext{\Gamma}{A}$,
-it follows that the sets $T(f_\ast(P))$ and $T(\pi_2(f,A)_\ast(P))$ are isomorphic
-via composition with $\varphi_{f,P}$. By \autoref{lem:ptm}, we have a function
-from $T(P)$ to $T(\pi_2(f,A)_\ast(P))$.
-\end{proof}
-
-\begin{defn}
-For any $\Gamma\in\cat{C}$, define $\cat{C}_\cat{F}/\Gamma$ to be the full
-subcategory of $\cat{C}/\Gamma$ with the same objects as $\cat{F}/\Gamma$. 
-\end{defn}
-
-\begin{thm}
-Let $\mathbb{E}$ be an E'-system. Then $f_\ast$ can be extended to a functor
-$\cat{C}_\cat{F}/\Gamma\to\cat{C}_\cat{F}/\Delta$, for each $f:\Delta\to\Gamma$, and moreover the diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=large]
-\cat{C}_\cat{F}/\ctxext{\Gamma}{B} \arrow[r,"{f_\ast/B}"] \arrow[d,swap,"{g_\ast}"]
-& \cat{C}_\cat{F}/\ctxext{\Delta}{f_\ast(B)} \arrow[d,"{(f_\ast(g))_\ast}"] \\
-\cat{C}_\cat{F}/\ctxext{\Gamma}{A} \arrow[r,swap,"{f_\ast/A}"]
-& \cat{C}_\cat{F}/\ctxext{\Delta}{f_\ast(A)}
-\end{tikzcd}
-\end{equation*}
-commutes for every $g:A\to B$ in $\cat{C}_\cat{F}/\Gamma$.
-\end{thm}
-
-\begin{proof}
-For any $A,B\in\cat{F}/\Gamma$, we have an isomorphism
-$\mathrm{hom}_{\cat{C}/\Gamma}(A,B)\cong T(W_A(B))$. We also get a function
-$T(W_A(B))\to T(f_\ast(W_A(B)))$. Therefore, it suffices to show that
-$f_\ast(W_A(B))\cong W_{f_\ast(A)}(f_\ast(B))$. 
-\begin{align*}
-f_\ast(W_A(B))
-& \jdeq f_\ast(A_\ast(B)) \\
-& \cong \pi_2(f,A)_\ast(A_\ast(B)) \\
-& \jdeq (A\circ\pi_2(f,A))_\ast(B) \\
-& \jdeq (f\circ f_\ast(A))_\ast(B) \\
-& \jdeq f_\ast(A)_\ast(f_\ast(B)) \\
-& \jdeq W_{f_\ast(A)}(f_\ast(B)). \qedhere
-\end{align*}
-\end{proof}
-
-\begin{thm}
-E-systems are the same things as E'-systems.
-\end{thm}
-
-\begin{proof}
-Given an E-system $\mathbb{E}$, we have the category $\cat{C}$ of internal
-morphisms in the empty context, with the subcategory $\cat{F}$ of projections $\cprojfstf{A}{P}$. 
-The functors $f_\ast:\cat{F}/\Gamma\to\cat{F}/\Delta$ are contravariant in
-$f$, so we have to verify whether we have the stated isomorphisms.
-
-Note first that for any $A\in\cat{F}/\Gamma$, $W_A$ is
-contravariant in $A$ and that it preserves the terminal object.
-
-Now let $\theta:\Delta\to\ctxext{\Gamma}{A}$ so that $A\circ\theta\jdeq f$
-for some $f:\Delta\to\Gamma$. Then we have $\zeta_{A,f}(\theta)\in T(f_\ast(A))$.  
-We wish to define $\theta_\ast(\idtm{A}):\Delta\to\ctxext{\Delta}{f_\ast(A)}$
-
-for any $A\in\cat{F}/\Gamma$ and $Q\in\cat{F}/\ctxext{\Gamma}{B}$, we have
-\begin{align*}
-T(B) 
-  & \jdeq 
-\{ \eta : \Gamma \to\ctxext{\Gamma}{B}
-        \mid B\circ\eta\jdeq\catid{\Gamma}\}
-  \\
-T(W_A(B))
-  & \jdeq
-\{ \eta : \ctxext{\Gamma}{A}\to\ctxext{{\Gamma}{A}}{W_A(B)}
-        \mid W_A(B)\circ\eta\jdeq\catid{\ctxext{\Gamma}{A}}\}
-  \\
-  & \cong
-\{ \theta : \ctxext{\Gamma}{A}\to\ctxext{\Gamma}{B}\mid B\circ\theta\jdeq A\}
-\end{align*}
-Therefore, $\eta\mapsto \zeta_{B,A}(\eta\circ A)$ defines a function from
-$T(B)$ to $T(W_A(B))$. 
-\end{proof}
-\end{comment}
-
-\subsection{Comprehension categories}
-
-\begin{defn}
-Let $p:\cat{E}\to\cat{B}$ be a functor and let $f:\Delta\to \Gamma$ be a morphism in $\cat{B}$.
-A morphism $g:A\to B$ in $\cat{E}$ is said to be \emph{cartesian over $f$}, if
+is a pullback square.
+Since the bottom square is assumed to be a pullback, we get that the outer rectangle
+is a pullback, as desired. 
+\end{constr}
+
+\begin{rmk}
+We can't directly use the functors $\Lambda(f)$ to give a D-system the structure
+of an E'-system, because there is a coherence condition missing.
+\end{rmk}
+
+\begin{thm}\label{thm:Dsys_coh}
+The following are equivalent:
 \begin{enumerate}
-\item $p(g)\jdeq f$
-\item For all $g':A'\to B$ with $p(g')\jdeq f$, there is a unique $u:A'\to A$
-such that $g'=g\circ u$ and $p(u)\jdeq\catid{\Delta}$. 
-\end{enumerate}
-A cartesian morphism over $f$ is also called a \emph{cartesian lift of $f$}.
-\end{defn}
-
-\begin{defn}
-A functor $p:\cat{E}\to\cat{B}$ is called a \emph{Grothendieck fibration} if
-every morphism into an object of the form $p(A)$ for some object $A$ of $\cat{E}$ 
-has a cartesian lift, and cartesian morphisms are closed under composition.
-\end{defn}
-
-\begin{defn}
-A Grothendieck fibration $p:\cat{E}\to\cat{B}$ is said to be \emph{split}, if
-there is a functorial choice of cartesian lifts, i.e.~for every $f:\Delta\to \Gamma$ in 
-$\cat{B}$ and any $B\in \cat{E}$ such that $p(B)\jdeq \Delta$,
-there is a cartesian lift $j(B,g):A \to B$ for some object $A$ of $\cat{E}$, such that 
-$j(B,p(B))\jdeq\catid{B}$ and $j(B, g\circ h)\jdeq j(B,g)\circ j(\mathrm{dom}(j(B,g)),h)$.
-\end{defn}
-
-\begin{eg}
-For any presheaf $P:\cat{C}\to\mathbb{Set}$, the projection functor
-$\int_\cat{C} P\to\cat{C}$ is a Grothendieck fibration.
-\end{eg}
-
-\begin{defn}
-A comprehension category $\mathcal{P}$ consists of categories $\cat{E}$, $\cat{C}$ and a
-commuting triangle
-\begin{equation*}
-\begin{tikzcd}
-\cat{E} \arrow[rr,"\mathcal{P}"] \arrow[dr,swap,"p"] & & \cat{C}^{\to} \arrow[dl,"\mathrm{cod}"] \\
-& \cat{C}
-\end{tikzcd}
-\end{equation*}
-of functors, such that $p$ is a Grothendieck fibration, and $\mathcal{P}$ maps
-cartesian morphisms in $\cat{E}$ to pullback squares in $\cat{C}$. We say that
-$\mathcal{P}$ is a full comprehension category if the functor $\mathcal{P}:
-\cat{E}\to\cat{C}^{\to}$ is full and faithful; and we say that
-$\mathcal{P}$ is split if $p$ is split.
-\end{defn}
-
-\begin{eg}
-Since any morphism of $\int_{\cat{C}}$ is cartesian, it follows that D-systems
-are comprehension categories. As D-systems are built up from a presheaf of types
-on the underlying category, they are in fact discrete comprehension categories.
-\end{eg}
-
-\begin{defn}
-Let $\mathcal{P}$ be a comprehension category. Then we define the pre-category
-$\cat{F}$ to be the category with the same objects as $\cat{C}$, of which the
-morphisms are generated by the morphisms of the form $P(A)$, for objects
-$A\in\cat{E}$. 
-\end{defn}
-
-\begin{defn}
-Let $\mathcal{P}$ be a comprehension category. Then we define for any 
-$\Gamma\in\cat{C}$, the poset $\Lambda(\Gamma)$ to be the category of finite 
-sequences  $A_1,\ldots,A_n$ of objects of $\cat{E}$, such that 
-$p(A_1)\jdeq\Gamma$ and $p(A_{i+1})\jdeq\mathrm{dom}(\mathcal{P}(A_i))$ 
-for any $1\leq i < n$. The partial order on 
-$\Lambda(\Gamma)$ is defined by $(A_1,\ldots,A_n)\leq (B_1,\ldots,B_m)$ if and 
-only if $n\leq m$ and $A_i\jdeq B_i$ for all $1\leq i\leq n$.
-
-The functor $\mathcal{P}$ induces a functor $\mathcal{P}:\Lambda(\Gamma)\to
-\cat{F}/\Gamma$ by $\mathcal{P}(A_1,\ldots,A_n)\jdeq\mathcal{P}(A_1)\circ
-\cdots\circ\mathcal{P}(A_n)$.
-\end{defn}
-
-\begin{defn}
-Let $f_0\defeq f:\Delta\to\Gamma$ be a morphism of $\cat{C}$, the base category of a
-split comprehension category $\mathcal{P}$ wich a choice $j$ of cartesian lifts.
-By induction we define an element $\Lambda(f,A)\in\Lambda(\Delta)$ together with
-a morphism $\pi_2(f,A):\mathrm{dom}(\mathcal{P}(\Lambda(f,A)))\to 
-\mathrm{dom}(\mathcal{P}(A))$, for any $A\jdeq (A_1,\ldots,A_n) 
-\in\Lambda(\Gamma)$.
-
-For $n\jdeq 0$, i.e.~if $A$ is the empty sequence, we let $\Lambda(f,A)$ be the 
-empty sequence, and $\pi_2(f,A) \jdeq f$. For the inductive step, consider a 
-sequence $A_1,\ldots,A_{n+1}\in \Lambda(\Gamma)$. Then we define
-\begin{align*}
-\Lambda(f,A) & \jdeq \mathrm{dom}(j(A_{n+1},\pi_2(f,(A_1,\ldots,A_n))))
-\intertext{and}
-\pi_2(f,A) & \jdeq \mathrm{dom}(\mathcal{P}(j(A_{n+1},\pi_2(f,(A_1,\ldots,A_n)))))
-\end{align*}
-where we consider in the definition of $\pi_2(f,A)$, $\mathrm{dom}$ to be the 
-domain functor from $\cat{C}^{\to}\to\cat{C}$. 
-\end{defn}
-
-\begin{rmk}
-It follows immediately from the definition of comprehension categories and the
-pasting lemma of pullbacks, that the square
-\begin{equation*}
-\begin{tikzcd}[column sep=large]
-\mathrm{dom}(\mathcal{P}(\Lambda(f,A))) \arrow[r,"{\pi_2(f,A)}"] \arrow[d,swap,"{\mathcal{P}(\Lambda(f,A))}"] & \mathrm{dom}(\mathcal{P}(A)) \arrow[d,"{\mathcal{P}(A)}"] \\
-\Delta \arrow[r,"f"] & \Gamma
-\end{tikzcd}
-\end{equation*}
-is a pullback square.
-\end{rmk}
-
-\begin{thm}
-Let $\mathcal{P}$ be a comprehension category. Then the following are equivalent:
-\begin{enumerate}
-\item The functors $\Lambda(f)$ can be extended to the structure of an E'-system
-without empty context.
-\item For any
-$A,A'\in\Lambda(\Gamma)$ such that $\mathcal{P}(A)\jdeq\mathcal{P}(A')$
-and for any $f:\Delta\to\Gamma$ in $\cat{C}$, the equalities
-$\mathcal{P}(\Lambda(f,A))\jdeq\mathcal{P}(\Lambda(f,A'))$ and
-$\pi_2(f,A)\jdeq\pi_2(f,A')$ hold. 
+\item The functors $\Lambda(f):\Lambda(\Gamma)\to\Lambda(\Delta)$ extend to 
+the structure of an E'-system without empty context.
+\item For any $f:\Delta\to\Gamma$ and all $A,A'\in\Lambda(\Gamma)$ satisfying 
+$p_A\jdeq p_{A'}$, one has
+\begin{equation*}
+p_{\Lambda(f,A)}\jdeq p_{\Lambda(f,A')}\qquad\text{and}\qquad\pi_2(f,A)\jdeq \pi_2(f,A').
+\end{equation*}
 \end{enumerate}
@@
 \begin{tikzcd}
-\Lambda(\Gamma) \arrow[d,swap,"\mathcal{P}"] \arrow[r,"\Lambda(f)"] & \Lambda(\Delta) \arrow[d,"\mathcal{P}"] \\
+\Lambda(\Gamma) \arrow[d,swap,"p"] \arrow[r,"\Lambda(f)"] & \Lambda(\Delta) \arrow[d,"p"] \\
 \cat{F}/\Gamma \arrow[r,densely dotted,"{f_\ast}"] & \cat{F}/\Delta
@@
 \end{equation*}
-of functors, forming an E'-system without empty context, where $\mathcal{P}(A)
-\defeq \mathcal{P}(A_1)\circ\cdots\circ\mathcal{P}(A_n)$ for any
-$A\jdeq (A_1,\ldots,A_n)\in\Lambda(\Gamma)$. Then it must hold that for any
-$A,A'\in\Lambda(\Gamma)$ such that $\mathcal{P}(A)\jdeq\mathcal{P}(A')$, we have
-$\mathcal{P}(\Lambda(f,A))\jdeq\mathcal{P}(\Lambda(f,A'))$. Also, for such
-$A$ and $A'$ it follows from the way $f_\ast$ is defined, 
-that $\pi_2(f,A)_\ast\jdeq f_\ast/\mathcal{P}(A)\jdeq
-f_\ast/\mathcal{P}(A')\jdeq\pi_2(f,A')_\ast$, so the desired equality
-$\pi_2(f,A)\jdeq\pi_2(f,A')$ holds.
+of functors, forming an E'-system without empty context. Let 
+$f:\Delta\to\Gamma$, and let $A,A':\Lambda(\Gamma)$ be such that 
+$p_A\jdeq p_{A'}$. Then it follows
+that $\pi_2(f,A)_\ast\jdeq f_\ast/A\jdeq f_\ast/A'\jdeq\pi_2(f,A')_\ast$, so
+we get the desired equality $\pi_2(f,A)\jdeq\pi_2(f,A')$.
 
@@
 \begin{defn}
-There is a functor $\mathcal{V}:\mathbf{Comp}\to\mathbf{E'sys_{\circ}}$, from
-the category $\mathbf{Comp}$ of split full comprehension categories to
-E'-systems without empty context.
-\end{defn}
-
-\begin{constr}
-Let $\mathcal{P}$ be a comprehension category with underlying category
-$\cat{C}$. We define $\cat{F}$ to be the sub-pre-category of $\cat{C}$ generated
-by the morphisms $\mathcal{P}(A):\ctxext{\Gamma}{A}\to\Gamma$ for $A\in\cat{E}$.
-
-For any $f:\Delta\to\Gamma$, we define a functor $f_\ast:\cat{F}/\Gamma\to
-\cat{F}/\Delta$ by induction on the length of the morphism
-Let $f:\Delta\to\Gamma$ and $p(A)\jdeq\Gamma$. Then we get a pullback square
+If either of the equivalent conditions of \autoref{thm:Dsys_coh} hold, we say 
+that the D-system is \emph{coherent}. Thus, we have a functor
+$U:\mathbf{Dsys}\to\mathbf{E'sys_{\circ}}$ from coherent D-systems to E'-systems
+\end{defn}
+
+\begin{defn}
+We say that a D-system $\cat{D}$ is \emph{locally well-founded} if the obvious functor
+$p:\Lambda(\mathbb{D})\to\cat{F}/\Gamma$ is an isomorphism of pre-categories.
+\end{defn}
+
+\begin{rmk}
+In other words, a D-system is locally well-founded if and only if
+for any $A\in\cat{F}$ there is a unique sequence
+$A_1\in\mathrm{Ty}(\Gamma)$, $A_2\in\mathrm{Ty}(\ctxext{\Gamma}{A_1})$, \ldots,
+$A_n\in\mathrm{Ty}(\ctxext{{{\Gamma}{A_1}}{\cdots}}{A_{n-1}})$ such that
+$A\jdeq p_{A_1}\circ\cdots\circ p_{A_n}$. Locally well-founded D-systems 
+are examples of coherent D-systems.
+\end{rmk}
+
+\begin{thm}
+The restriction of the functor $U:\mathbf{Dsys}\to\mathbf{E'sys_{\circ}}$ to the
+locally well-founded D-systems is full and faithful.
+\end{thm}
+
+\begin{proof}
+The functor $U$ is faithful on the locally well-founded D-systems, because for any morphism 
+$\mathcal{F}:\mathbb{D}\to\mathbb{D}'$ of locally well-founded D-systems, 
+$U(\mathcal{F})$ and $\mathcal{F}$ have the same underlying functor
+from $\cat{C}$ to $\cat{C}'$.
+
+It remains to show that $U$ is full on the locally well-founded D-systems. Let
+$\mathbb{D}$ and $\mathbb{D}$ be locally well-founded D-systems and let $\mathcal{G}:U(\mathbb{D})
+\to U(\mathbb{D}')$. Then $\mathcal{G}$ restricted to $\cat{F}$ lifts factorizations.
+We  first have to define a natural transformation
+$\mathrm{Ty}\Rightarrow \mathrm{Ty}'\circ\mathcal{G}$. Let
+$\Gamma\in\cat{C}$ and $A\in\mathrm{Ty}(\Gamma)$. Then we have
+$\mathcal{G}(p_A):\mathcal{G}(\ctxext{\Gamma}{A})\to \mathcal{G}(\Gamma)$.
+The category of factorizations of $p_A$ in $\cat{F}$ is a isomorphic to $\mathbf{2}$,
+so the category of factorizations of $\mathcal{G}(p_A)$ is also isomorphic to
+$\mathbf{2}$. This implies that $\mathcal{G}(p_A)$ is of the form $p_B$ for
+some $B\in\mathrm{Ty}'(\mathcal{G}(\Gamma))$. This $B$ is called $\mathcal{G}(A)$.
+
+To verify the naturality, let $f:\Delta\to\Gamma$. We need to show that the
+square
+\begin{equation*}
+\begin{tikzcd}
+\mathrm{Ty}(\Delta) \arrow[r,"{\mathcal{G}_\Delta}"] & \mathrm{Ty}'(\mathcal{G}(\Delta)) \\
+\mathrm{Ty}(\Gamma) \arrow[u,"\mathrm{Ty}(f)"] \arrow[r,"{\mathcal{G}_\Gamma}"] &
+\mathrm{Ty}'(\mathcal{G}(\Gamma)) \arrow[u,swap,"{\mathrm{Ty}'(\mathcal{G}(f))}"]
+\end{tikzcd}
+\end{equation*}
+commutes. Let $A\in\mathrm{Ty}(\Gamma)$. Then we need to show that
+$\mathcal{G}(f)_\ast(\mathcal{G}(A))\jdeq \mathcal{G}(f_\ast(A))$. By the uniqueness
+assumption of well-founded D-systems, it suffices to show that
+$p_{\mathcal{G}(f)_\ast(\mathcal{G}(A))}\jdeq p_{\mathcal{G}(f_\ast(A))}$. To
+see this, we do a simple calculation
+\begin{align*}
+p_{\mathcal{G}(f)_\ast(\mathcal{G}(A))}
+& \jdeq \mathcal{G}(f)_\ast(p_{\mathcal{G}(A)}) \\
+& \jdeq \mathcal{G}(f)_\ast(\mathcal{G}(p_A)) \\
+& \jdeq \mathcal{G}(f_\ast(p_A)) \\
+& \jdeq \mathcal{G}(p_{f_\ast(A)}) \\
+& \jdeq p_{\mathcal{G}(f_\ast(A))} \qedhere
+\end{align*}
+\end{proof}
+
+\begin{thm}
+The image of $U$ is the subcategory of locally stratified E'-systems without
+empty context. 
+\end{thm}
+
+\begin{proof}
+For any $\Gamma\in\mathbb{D}$, there is an obvious stratification functor
+$\Lambda(\Gamma)\to (\N,\geq)$, so $U(\mathbb{D})$ is a locally stratified
+E'-system.
+
+Now let $\mathbb{E}$ be a locally stratified E'-system without empty context.
+For $\Gamma\in\cat{C}$, we define $\mathrm{Ty}(\Gamma)$ to be the set of
+families of length one. Since any $f_\ast$ is stratified, it follows that
+$\mathrm{Ty}$ is a presheaf on $\cat{C}$. The choice of pullback squares
 \begin{equation*}
 \begin{tikzcd}[column sep=large]
-\ctxext{\Delta}{f_\ast(A)} \arrow[d,fib] \arrow[r,"{\pi_2(f,A)}"] & \ctxext{\Gamma}{A} \arrow[d,fib] \\
-\Delta \arrow[r,"f"] & \Gamma
-\end{tikzcd}
-\end{equation*}
-There seems to be no way to show that
-$\pi_2(f,\ctxext{A}{P})\jdeq \pi_2(\pi_2(f,A),P)$.
-\end{constr}
-
-\begin{conj}
-An E'-system is the same thing as a split full comprehension category in which
-$\cat{E}$ has a terminal object.
-\end{conj}
-
-\begin{proof}
-Let $\mathcal{E}$ be an E'-system. Define $\cat{E}$ to be the full subcategory
-of $\cat{C}^\to$ whose objects are the morphisms of $\cat{F}$. Thus, for any
-$\Gamma\in\cat{C}$, the category $p^{-1}(\Gamma)$ is the category of internal
-morphisms the E-system $\mathcal{E}/\Gamma$.
-
-We first show that $p$ is a Grothendieck fibration. Let $f:\Delta\to\Gamma$ be
-a morphism in $\cat{C}$ and let $A\in\cat{F}/\Gamma$. Then we have $\catid{\Delta},\catid{\Gamma}\in\cat{E}$,
-and $(f,\pi_2(f,A)):f_\ast(A)\to A$. It is routine to verify that $(f,\pi_2(f,A))$
-is a cartesian lift of $f$, and that this choice of catesian lifts is functorial.
-
-Now let $g:X\to\Delta$ and $f:\Delta\to \Gamma$, and let $\tilde{f}:\tilde{\Delta}\to\tilde{\Gamma}$ and
-$\tilde{g}:\tilde{X}\to\tilde{\Delta}$ be cartesian lifts of $f$ and $g$, respectively. We have to show
-that $\tilde{f}\circ \tilde{g}$ is a cartesian lift of $f\circ g$. Let $C\in\cat{F}/X$ and
-let $h : C\to \tilde{\Gamma}$ be a morphism such that $\mathrm{cod}(h)\jdeq f\circ g$. 
+\ctxext{\Delta}{f_\ast(A)} \arrow[r,"{\pi_2(f,A)}"] \arrow[d,fib] & \ctxext{\Gamma}{A} \arrow[d,fib] \\
+\Delta \arrow[r,swap,"f"] & \Gamma
+\end{tikzcd}
+\end{equation*}
+is functorial by assumption, so we have defined a category with families. It
+is straigtforward to verify that this category with families is locally well-founded,
+and that its construction is inverse to $U$. 
 \end{proof}

### 2015-10-01 internal_homs.tex
moved intro to new file

@@
-\section{The pre-category of internal morphisms of an E-system}\label{sec:esys_props}
+\section{Elementary properties of E-systems}\label{sec:esys_props}
 
@@
 \begin{equation*}
-\pi_2(f,\ctxext{A}{P}) \jdeq \pi_2(\pi_2(f,a),P)
+\pi_2(f,\ctxext{A}{P}) \jdeq \pi_2(\pi_2(f,A),P)
 \end{equation*}
@@
 \end{proof}
-
-\begin{comment}
-\subsection{Categories internal to E-systems}
-The definition of internal category we're about to give, has a type of objects
-and a type of morphisms, such that the codomain function is a projection. It might
-be more appropriate to call such structures `right categories', and the corresponding
-structures where the domain function is a projection might be called `left
-categories'. In particular, the opposite of a right category will be a left
-category. There is also a symmetric notion of a category.
-
-\begin{defn}
-A category-framework $\cftalg{A}$ in context $\Gamma$ in an E-system $\mathbb{E}$ consists of
-\begin{enumerate}
-\item A family $\cftalgc{\cftalg{A}}\in\cat{F}/\Gamma$, representing the type of
-objects of the category framework;
-\item A family $\cftalgf{\cftalg{A}}\in\cat{F}/\ctxext{\Gamma}{\cftalgc{\cftalg{A}}}$,
-representing the type of morphisms of the category framework, where a term of
-$\subst{X}{\cftalgf{\cftalg{A}}}$ is considered to be a morphism into $X\in T(\cftalgc{\cftalg{A}})$;
-\item A morphism $\jhom{\Gamma}{\ctxext{\cftalgc{\cftalg{A}}}{\cftalgf{\cftalg{A}}}}{\cftalgc{\cftalg{A}}}{\cftctxext}$, representing the domain function;
-\item A morphism $\jhom{{\Gamma}{\cftalgc{\cftalg{A}}}}{{\cftalgf{\cftalg{A}}}{\jcomp{}{\cftctxext}{\cftalgf{\cftalg{A}}}}}{\cftalgf{\cftalg{A}}}{\cftfamext}$, representing composition;
-\item A term $\jterm{{\Gamma}{\cftalgc{\cftalg{A}}}}{\cftalgf{\cftalg{A}}}{\cftempf{\cftalg{A}}}$,
-representing the identity morphisms.
-\end{enumerate}
-\end{defn}
-
-\begin{rmk}
-Instead of looking at $\cftfamext$ as a morphism from 
-$\ctxext{\cftalgf{\cftalg{A}}}{\jcomp{}{\cftctxext}{\cftalgf{\cftalg{A}}}}$
-to $\cftalgf{\cftalg{A}}$ in context $\ctxext{\Gamma}{\cftalgc{\cftalg{A}}}$, 
-one could also look at $\cftfamext$ as a 
-morphism \emph{over $\cprojfstf{\cftalgc{\cftalg{A}}}{\cftalgf{\cftalg{A}}}$} 
-in context $\Gamma$, as indicated in the following diagram:
-\begin{equation*}
-\begin{tikzcd}[column sep=large]
-\cftalgf{\cftalg{A}}
-  \ar[fib]{d}
-& \jcomp{}{\cftctxext}{\cftalgf{\cftalg{A}}}
-  \ar[fib]{d}
-  \ar{l}[swap]{\cftfamext}
-  \ar{r}
-& \cftalgf{\cftalg{A}}
-  \ar[fib]{d}
-  \\
-\cftalgc{\cftalg{A}}
-& \ctxext{A}{\cftalgf{\cftalg{A}}}
-  \ar{l}{\cprojfstf{A}{\cftalgf{\cftalg{A}}}}
-  \ar{r}[swap]{\cftctxext}
-& \cftalgc{\cftalg{A}}
-\end{tikzcd}
-\end{equation*}
-From this point of view, we see that $\cftfamext$ has an object $X$ and a 
-morphism $f$ into $X$ as parameters, 
-then takes a morphism into the domain of $f$ as an argument, 
-and outputs a morphism into $X$. The projection maps $(X,f)$ to $X$, so we see
-that composition is indeed an operation of this kind.
-\end{rmk}
-
-A category-framework is not a category unless several axioms of the theory of
-categories are satisfied. The domain of a composed arrow $g\circ f$ must be the
-domain of $f$, composition must be associative, the domains of the identity
-morphisms must be the same as their codomains, and identities must satisfy
-the left and right identity laws. However, the axioms expressing these 
-conditions are not very insightful when they're written down directly. Thus, we
-take the usual way and define the notion of functor before we describe the
-actual conditions of being a category.
-
-\begin{defn}
-Let $\cftalg{A}$ and $\cftalg{B}$ be category-frameworks in context $\Gamma$,
-in an E-system $\mathbb{E}$.
-An \emph{internal functor $\cfthom{F}$ from $\extalg{A}$ to
-$\extalg{B}$ in context $\Gamma$} is a pair 
-$(\cfthomc{\cfthom{F}},\cfthomf{\cfthom{F}})$ consisting of
-\begin{align*}
-\jalign\jhom
-  {\Gamma}
-  {\cftalgc{\cftalg{A}}}
-  {\cftalgc{\cftalg{B}}}
-  {\cfthomc{\cfthom{F}}}
-  \\
-\jalign\jfhom
-  {\Gamma}
-  {\cftalgc{\cftalg{A}}}
-  {\cftalgc{\cftalg{B}}}
-  {\cfthomc{\cfthom{F}}}
-  {\cftalgf{\cftalg{A}}}
-  {\cftalgf{\cftalg{B}}}
-  {\cfthomf{\cfthom{F}}}
-\end{align*}
-for which the diagrams
-\begin{equation}\label{eq:exthom1}
-\begin{tikzcd}
-\ctxext{\cftalgc{\cftalg{A}}}{\cftalgf{\cftalg{A}}}
-  \ar{r}{\jvcomp{}{\cfthomc{\cfthom{F}}}{\cfthomf{\cfthom{F}}}}
-  \ar{d}[swap]{\cftctxext[\cftalg{A}]}
-& \ctxext{\cftalgc{\cftalg{B}}}{\cftalgf{\cftalg{B}}}
-  \ar{d}{\cftctxext[\cftalg{B}]}
-  \\
-\cftalgc{\cftalg{A}}
-  \ar{r}[swap]{\cfthomc{\cfthom{F}}}
-& \cftalgc{\cftalg{B}}
-\end{tikzcd}
-\end{equation}
-and
-\begin{equation}\label{eq:exthom2}
-\begin{tikzcd}[column sep=huge]
-\ctxext{\cftalgf{\cftalg{A}}}{\jcomp{}{\cftctxext[\cftalg{A}]}{\cftalgf{\cftalg{A}}}}
-  \ar{r}{\jvcomp{}{\cfthomf{\cfthom{F}}}{\jcomp{}{\cftctxext[\cftalg{A}]}{\cfthomf{\cfthom{F}}}}}
-  \ar{d}[swap]{\cftfamext[\cftalg{A}]}
-& \jcomp{}{\cfthomc{\cfthom{F}}}{\ctxext{\cftalgf{\cftalg{B}}}{\jcomp{}{\cftctxext[\cftalg{B}]}{\cftalgf{\cftalg{B}}}}}
-  \ar{d}{\jcomp{}{\cfthomc{\cfthom{F}}}{\cftfamext[\cftalg{B}]}}
-  \\
-\cftalgf{\cftalg{A}}
-  \ar{r}[swap]{\cfthomf{\cfthom{F}}}
-& \jcomp{}{\cfthomc{\cfthom{F}}}{\cftalgf{\cftalg{B}}}
-\end{tikzcd}
-\end{equation}
-commute, and for which $\subst{\cftempf{\cftalg{A}}}{\cfthomf{\cfthom{F}}}
-\jdeq\jcomp{}{\cfthomc{\cfthom{F}}}{\cftempf{\cftalg{B}}}$.
-\end{defn}
-
-\begin{defn}
-Let $\cftalg{A}$ be a category-framework. We define the category-framework
-\begin{equation*}
-\cftfamalg{\cftalg{A}}\defeq(\cftalgf{\cftalg{A}},\jcomp{}{\cftctxext}%
-{\cftalgf{\cftalg{A}}},\cftfamext,\jcomp{}{\cftctxext}{\cftfamext},
-\jcomp{}{\cftctxext}{\cftempf{\cftalg{A}}}),
-\end{equation*}
-When $\cftalg{P}$ is a category-framework in context $\ctxext{\Gamma}{A}$, then
-and we define the category-framework
-\begin{equation*}
-\ctxext{\cftalgc{\cftalg{A}}}{\cftalg{P}}
-  \defeq
-( \ctxext{A}{\cftalgc{\cftalg{P}}},
-  \cftalgf{\cftalg{P}},
-  \jvcomp{}{\idtm{A}}{\cftctxext[\cftalg{P}]},
-  \cftfamext[\cftalg{P}],
-  \cftempf{\cftalg{P}}).
-\end{equation*}
-\end{defn}
-
-\begin{defn}
-An internal category $\cftalg{A}$ in an E-system $\mathbb{E}$ consists of a
-category-framework $\cftalg{A}$ satisfying the following conditions:
-\begin{enumerate}
-\item \emph{The domain condition.} The pair 
-$(\cftctxext,\idtm{\jcomp{}{\cftctxext}{\cftalgf{\cftalg{A}}}})$ is a
-functor from $\ctxext{\cftalgc{\cftalg{A}}}{\cftfamalg{\cftalg{A}}}$ to
-$\cftalg{A}$.
-\item \emph{The associativity law.} The pair
-$(\cftfamext,\idtm{\jcomp{}{\cftfamext}{\cftalgf{\cftfamalg{\cftalg{A}}}}})$
-is a functor from $\ctxext{\cftalgf{\cftalg{A}}}{\cftfamalg{\cftfamalg{\cftalg{A}}}}$ to
-$\cftfamalg{\cftalg{A}}$.
-\item \emph{The identity domain condition.} The domain of the identity morphism
-is the same as its codomain:
-\begin{equation*}
-\jtermeq
-  {{\Gamma}{\cftalgc{\cftalg{A}}}}
-  {\ctxwk{\cftalgc{\cftalg{A}}}{\cftalgc{\cftalg{A}}}}
-  {\subst{\cftempf{\cftalg{A}}}{\cftctxext[\cftalg{A}]}}
-  {\idtm{\cftalgc{\cftalg{A}}}}.
-\end{equation*}
-\item \emph{The left identity law.} A morphism precomposed by the identity morphism
-is equal to the original morphism:
-\begin{equation*}
-\jtermeq
-  {{{\Gamma}{\cftalgc{\cftalg{A}}}}{\cftalgf{\cftalg{A}}}}
-  {\ctxwk{\cftalgf{\cftalg{A}}}{\cftalgf{\cftalg{A}}}}
-  {\subst{\cftempf{\cftalg{A}}}{\cftfamext}}
-  {\idtm{\cftalgf{\cftalg{A}}}}.
-\end{equation*}
-\item \emph{The right identity law.} A morphism postcomposed by the identity morphism
-is equal to the original morphism:
-\begin{equation*}
-\jtermeq
-  {{{\Gamma}{\cftalgc{\cftalg{A}}}}{\cftalgf{\cftalg{A}}}}
-  {\ctxwk{\cftalgf{\cftalg{A}}}{\cftalgf{\cftalg{A}}}}
-  {\subst{\jcomp{}{\cftctxext}{\cftempf{\cftalg{A}}}}{\cftfamext}}
-  {\idtm{\cftalgf{\cftalg{A}}}}.
-\end{equation*}
-\end{enumerate}
-\end{defn}
-
-\begin{rmk}
-Since we have only used operations of E-systems to define internal categories,
-it immediately follows that if $\cftalg{A}$ is an internal category in
-$\mathbb{E}$ and if $F:\mathbb{E}\to\mathbb{D}$ is an E-homomorphism, then
-$F(\cftalg{A})$ is an internal category in $\mathbb{D}$. In particular, if
-$\cftalg{Q}$ is an internal category in context $\ctxext{\Gamma}{B}$ and
-$f:A\to B$, then the change of base $f_\ast(\cftalg{Q})$ is an internal category
-in context $\ctxext{\Gamma}{A}$. Thus, it is immediate that the notion of
-internal category is stable under change of base.
-\end{rmk}
-\end{comment}
 

### 2015-10-01 introduction.tex
moved intro to new file

--- 2015-08-16 main.tex
@@
-\documentclass{article}
-
-\input{Preamble/preamble-articles}
-
-%\includeonly{Algebras/article-models-cat}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\title{An essentially algebraic formulation of dependent type theory}
-\author{Egbert Rijke\footnote{{\sc Carnegie Mellon University}, department of
-philosophy; \url{erijke@andrew.cmu.edu}}}
-\date{\today}
-
-\begin{document}
-
-\maketitle
-
-\begin{abstract}
-We present an algebraic formulation of dependent type theory with
-three sorts for contexts, families and terms, with finitely many operations
-between them and finitely many equality axioms. Algebras for this theory are
-called E-systems. A distinctive feature of the dependent type theory of
-E-systems is that it takes `telescopic families' as the primitive notion of
-families.
-
-We then compare Voevodsky's B-systems, Dybjer's categories
-with families and Jacob's comprehension categories with E-systems. There is
-a full and faithful functor from B-systems to E-systems, of which the image is
-the category of stratified E-systems. To have a functor from categories with
-families to E-systems, we need to impose an additonal coherence condition, which
-makes sure that the substitution operation of a category with families extends 
-to telescopic families. 
-\end{abstract}
-
-%\tableofcontents
-
 \section{Introduction --- Mathematical languages as designed and maintained languages}
@@
 discovered.
-
-%\tableofcontents
-
-%\input{Introduction/article-models-intro}
-
-%\part{Type theory}
-
-\begin{comment}
-In this part we develop type theory from the ground up. We start with a type
-theory without any of the basic constructors. This is the theory of contexts
-families and terms which has the basic operations of extension, weakening,
-substitution and identity terms. Type theory before type constructors has not
-been studied very much. Dependent product types or even universes tend to make
-an early appearance in just about any presentation of type theory.
-A noteable exception is the theory of categories with families of Dybjer
-in \cite{Dybjer1996}, which has been elaborated on further by Dybjer and
-Clairambault in \cite{DybjerClairambault2011} and in unpublished work by
-Awodey \cite{Awodey2013} on natural models of type theory,
-which makes the connection between categories with families and representable
-transformations of presheaves. In the way type theory is presented by the
-Univalent Foundations Project in \cite{TheBook}, which seems to have won the race of introducing universes
-as early as possible hands down, it seems entirely unfeasible to study type
-theory without type constructors. Also, the 
-proof-assistant \Coq\ {\color{red}(and \Agda too?)} has universes and dependent product and pair types 
-built-in, making it impossible to study type theory without type constructors in 
-that environment.
-
-Nevertheless, type theory without type constructors has received some attention
-contributors to the Univalent Foundations Program recently. We name two
-further investigations on this topic, other than the mentioned work by Awodey.
-In \cite{Garner2014}, Garner describes the combinatorial structure
-of the type operations of the weakening, substitution and projection monads
-(their projections are our identity terms) and suggests lots of further research
-that can be done on type theory without constructors. Also, Joyal has presented
-his theory of tribes, which is a categorical explanation of type theory
-without type constructors.
-
-After we have described the E-system, which is the flavor of type theory without
-type constructors presented in \autoref{tt}, we will demonstrate in \autoref{ttderived}
-that the theory gives rise to a rich categorical structure. The notions and
-properties we derive here will be essential for the further work on internal
-models for E-systems, presented in \autoref{part:models}.
-\end{comment}
-
-\begin{comment}
-\section{Introduction}
-In this article we define the essentially algebraic objects which correspond to dependent
-type theories, which are closely related to Voevodsky's B-systems. This provides
-a new axiomatization of dependent type theory, with the following properties:
-\begin{itemize}
-\item The meta-theory should not require anything more than rules for handling
-inferences. In particular, natural numbers, or an infinite set of variables 
-should not be required.
-\item Finitely many sorts. The basic ingredients are contexts, families and
-terms. Thus there will be 3 sorts.
-\item The theory is algebraic and the operations on it are homomorphisms. From
-the requirement that type theoretic operations must be homomorphisms, we can
-read off which judgmental equalities to require require.
-\item Finite set of rules.
-\item The theory is invariant under slicing in a canonical way.
-\end{itemize}
-
-We build
-dependent type theory by imposing several structures of increasing complexities
-on top of each other. We begin by the theory of extension, which incorporates
-a context extension and a family extension operation to the theory which merely
-has three sorts. On top of the theory of extension we can formulate the theories
-of weakening, substitution and extension-units (i.e.~the empty context and families).
-The theory of weakening can be enriched with units playing the role of identity
-morphisms, which gives a theory of projections. The three resulting theories
-can then be combined into dependent type theory. Thus, the theory of dependent
-types is outlined in the following diagram.
-
-\begin{equation*}
-\begin{tikzcd}[ampersand replacement=\&,column sep=0em]
-{} \& \textbf{Extension} \ar[very thick,-stealth]{dr} \ar[very thick,-stealth]{d} \ar[very thick,-stealth]{dl}
-  \\
-\makebox[.7cm]{\textbf{Weakening}} \ar[very thick,-stealth]{d} \& \textbf{Substitution} \ar[very thick,-stealth]{dd} \& \makebox[3cm]{\textbf{Empty context and families}} \ar[very thick,-stealth]{ddl}
-\\
-\makebox[.7cm]{\textbf{Projection}} \ar[very thick,-stealth]{dr}
-\\
-{} \& \textbf{Dependent Type Theory}
-\end{tikzcd}
-\end{equation*}
-
-We give two presentations of the theory simultaneously. First, we directly
-define the algebras for the corresponding theories in an arbitrary category
-with finite limits and second, we describe the theory in terms of inference
-rules.
-
-\subsection{Categories with finite limits}
-For the semantics of the theory of E-systems, we will assume that $\cat{C}$ is a 
-category with finite limits and
-whenever we write a pullback, we assume that it is chosen. Recall that for
-any morphism $f:A\to B$ in a category $\cat{C}$ with chosen pullbacks, there
-is a functor
-\begin{equation*}
-f^\ast : \cat{C}/B\to\cat{C}/A.
-\end{equation*}
-As usual, when $g:X\to B$ is a morphism, we will write $f^\ast(X)$ for the
-domain of $f^\ast(g)$. When there is more than one morphism $X\to B$ involved,
-as will be the case below, we will write $\pullback{A}{X}{f}{g}$. The projections
-will be written as $\pullbackpr{1}{f}{g}$ and $\pullbackpr{2}{f}{g}$. So in this notation, a
-typical pullback diagram has the following form:
-\begin{equation*}
-\begin{tikzcd}[column sep=large]
-\pullback{A}{X}{f}{g}
-  \ar{r}{\pullbackpr{1}{f}{g}}
-  \ar{d}[swap]{\pullbackpr{2}{f}{g}}
-  &
-A \ar{d}{f}
-  \\
-X \ar{r}[swap]{g}
-  &
-B
-\end{tikzcd}
-\end{equation*}
-Furthermore, given $h:Y\to A$ and $k:Y\to B$ such that $h\circ f=g\circ k$, we
-denote the arrow $Y\to\pullback{A}{X}{f}{g}$ given by the universal property of
-pullbacks, by $[h,k]$. Also, when we have a commutative diagram of the form
-\begin{equation*}
-\begin{tikzcd}
-A \ar{r}{f}
-  \ar{d}{a}
-  &
-X \ar{d}
-  & 
-B \ar{l}[swap]{g}
-  \ar{d}{b}
-  \\
-A'
-  \ar{r}[swap]{f'}
-  &
-X'
-  &
-B'
-  \ar{l}{g'}
-\end{tikzcd}
-\end{equation*}
-we will denote the unique map from $\pullback{A}{B}{f}{g}$ to $\pullback{A'}{B'}{f'}{g'}$
-such that the diagram
-\begin{equation*}
-\begin{tikzcd}
-  {}
-  & 
-\pullback{A'}{B'}{f'}{g'}
-  \ar{dd}
-  \ar{rr}
-  &
-  &
-B'
-  \ar{dd}{g'}
-  \\
-\pullback{A}{B}{f}{g}
-  \ar{dd}
-  \ar[crossing over]{rr}
-  \ar[dotted]{ur}{\pullback{a}{b}{f'}{g'}}
-  &
-  &
-B \ar{ur}{b}
-  \\
-  {}
-  &
-A'
-  \ar{rr}
-  &
-  &
-X'
-  \\
-A \ar{rr}[swap]{f}
-  \ar{ur}{a}
-  &
-  &
-X \ar[crossing over,leftarrow]{uu}[near end,swap]{g}
-  \ar{ur}
-\end{tikzcd}
-\end{equation*}
-commutes, by $\pullback{a}{b}{f'}{g'}$. In the current work, we shall
-write $A\times B$ for the pullback of $A\rightarrow 1\leftarrow B$, and
-$\pi_1$ and $\pi_2$ for its projections (thus, no separate choice of
-cartesian products is made).
-
-\subsection{Inference rules}
-The syntactic theory of contexts, families and terms shall be described by means
-of a declaration of valid inference rules. Although their appearance suggests 
-otherwise, the notion of inference rule is ultimately informal, in the sense
-that we state the explanations and possibilities with inference rules in
-natural language. One may attempt to study systems of inference rules within
-other systems to study the applications of the various concepts involved, but
-this study will not make the basic notion of inference rules more formal. We do
-not see this as a problem, as systems of inference rules have been implemented
-in proof assistant software in which one essentially only manipulates
-judgments. In this paper we use inference rules as an alternative presentation
-of the theory, which some readers may prefer.
-
-An inference rule is a finite list $\mathcal{H}_1,\ldots,\mathcal{H}_n;
-\mathcal{J}$ of expressions called judgments. Usually we represent an inference
-rule in the following form
-\begin{equation*}
-\inference{\mathcal{H}_1\quad\cdots\quad\mathcal{H}_n}{\mathcal{J}}
-\end{equation*}
-The judgments $\mathcal{H}_1$, \ldots, $\mathcal{H}_n$ are called the
-hypotheses and the judgment $\mathcal{J}$ is the conclusion of the inference
-rule.
-
-An inference rule is called \emph{valid} in either of the following cases:
-\begin{description}
-\item[(Axiom)] It is declared valid as an axiom;
-\item[(Projection)] The inference rule
-\begin{equation*}
-\inference{\mathcal{H}_1\quad\cdots\quad\mathcal{H}_n}{\mathcal{H}_i}
-\end{equation*}
-is a valid inference rule for any $1\leq i\leq n$;
-\item[(Composition)] If
-\begin{equation*}
-\inference{\mathcal{I}_{1k}\quad\cdots\quad\mathcal{I}_{m_kk}}{\mathcal{H}_k}
-\qquad\text{and}\qquad
-\inference{\mathcal{H}_1\quad\cdots\quad\mathcal{H}_n}{\mathcal{J}}
-\end{equation*}
-are valid inference rules for $1\leq k\leq n$, then
-\begin{equation*}
-\inference{\mathcal{I}_{11}\quad\cdots\quad\mathcal{I}_{m_11}\quad\cdots\quad\mathcal{I}_{1n}\quad\cdots\quad\mathcal{I}_{m_nn}}{\mathcal{J}}
-\end{equation*}
-is also a valid inference rule;
-\item[(Cut elimination)] If
-\begin{equation*}
-\inference{\mathcal{H}_1\quad\cdots\quad\mathcal{H}_n}{\mathcal{H}_{n+1}}
-\qquad
-\text{and}
-\qquad
-\inference{\mathcal{H}_1\quad\cdots\quad\mathcal{H}_{n+m+1}}{\mathcal{J}}
-\end{equation*}
-are valid inference rules, then
-\begin{equation*}
-\inference{\mathcal{H}_1\quad\cdots\quad\hat{\mathcal{H}}_{n+1}\quad\cdots\quad\mathcal{H}_{n+m+1}}{\mathcal{J}}
-\end{equation*}
-where the hypothesis $\mathcal{H}_{n+1}$ is omitted, is a valid inference rule.
-\end{description}
-In categories with finite limits, valid inference rules will always be 
-interpreted by  arrows.
-
-\input{fundamental_structure}
-
-\input{extension}
-
-\input{weakening}
-
-\input{projections}
-
-\input{substitution}
-
-\input{empty}
-
-\input{dependent_type_theory}
-
-\section{The relation between E-systems and B-systems}
-
-In this section we answer the following questions posed by Voevodsky in private
-correspondence.
-\begin{itemize}
-\item Is the construction that takes a (unital) B-system to an E-system a functor?
-\item Is this functor fully faithful?
-\item Is there a simple way to characterize the E-systems that arise from B-systems (the image of this functor on objects)?
-\item Construct an example of an E-system that does not arise from a B-system.
-\end{itemize}
-
-\subsection{The functor from B-systems to E-systems}
-Consider a B-system $\mathbb{B}$. We will construct an E-system $E(\mathbb{B})$
-in the category of sets. This E-systems consists of sets $C(\mathbb{B})$,
-$F(\mathbb{B})$ and $T(\mathbb{B})$ for the sorts of contexts, families and
-terms respectively, with functions $\eft:F(\mathbb{B})\to C(\mathbb{B})$ and
-$\ebd:T(\mathbb{B})\to T(\mathbb{B})$ and the operations of extension,
-weakening, units, substitution and empty contexts and families.
-
-The set $C(\mathbb{B})$ consists of pairs $(n,X)$, where $n\in\mathbb{N}$ and
-$X\in B_n$. The set $F(\mathbb{B})$ consists of quadruples $(n,X,m,Y)$, where
-$n,m\in\mathbb{N}$ such that $n<m$, $X\in B_n$ and $Y\in B_{m}$ such that
-$ft^{m-n}:=ft^{n-1}\circ\cdots\circ ft^m(Y)=X$.
-
-%\include{Derivations/article-models-ttderived}
-
-%\part{Internal Models}\label{part:models}
-
-%\include{Internal_models/article-models-ttinternal}
-
-%\part{Type constructors}
-
-%\include{Dependent_function_types/article-models-ttdp}
-
-%\include{Universes/article-models-ttuniverses}
-
-%\include{Junk/article-models-ttind}
-
-%\part{Categorical semantics}
-
-%\input{Algebras/article-models-cat}
-
-\part{The new E-systems}
-\end{comment}
-
-\input{e-systems}
-
-\input{internal_homs}
-
-\input{compared}
-
-\section{Type constructors in E-systems}
-
-\begin{defn}
-An E-system $\mathbb{E}$ is said to have $\Pi$-types, if 
-\begin{enumerate}
-\item for any 
-$A\in\cat{F}/\Gamma$, there is an E-homomorphism
-\begin{equation*}
-\Pi_A : \mathbb{E}/\ctxext{\Gamma}{A}\to\mathbb{E}/\Gamma
-\end{equation*}
-For terms $g:Q$ in context $\ctxext{{\Gamma}{A}}{P}$, we will usually write
-$\lambda_{A}(g)$ for the term $\Pi_A(g)$ of $\Pi_A(Q)$ in context
-$\ctxext{\Gamma}{\Pi_A(P)}$. Note that when $P$ is empty, then we retrieve the
-usual lambda-abstraction.
-\item for any $A\in\cat{F}/\Gamma$ and $Q\in\cat{F}/\ctxext{{\Gamma}{A}}{P}$, 
-there is a term 
-\begin{equation*}
-\tfev^{A,Q} : \ctxwk{\Pi_A(Q)}{{\Pi_A(P)}{Q}}
-\end{equation*}
-in context $\ctxext{{{{\Gamma}{\Pi_A(P)}}{\Pi_A(Q)}}{\ctxwk{\Pi_A(\ctxext{P}{Q})}{A}}}{\ctxwk{\Pi_A(\ctxext{P}{Q})}{P}}$.
-Note that for each $g:\Pi_A(Q)$, we obtain a term $\subst{g}{\tfev^{A,Q}}: \ctxwk{\Pi_A(P)}{Q}$ in
-context $\ctxext{{\Gamma}{\Pi_A(P)}}{\ctxwk{\Pi_A(P)}{A}}$.
-\item for any $A\in\cat{F}/\Gamma$ there is a term
-\begin{equation*}
-{\tfev^{A,P}} : {\ctxwk{\Pi_A(P)}{P}}
-\end{equation*}
-in context ${\ctxext{{\Gamma}{\Pi_A(P)}}{\ctxwk{\Pi_A(P)}{A}}}$
-\end{enumerate}
-satisfying
-\begin{enumerate}
-\item 
-\begin{enumerate}
-\item For any $P\in\cat{F}/\ctxext{\Gamma}{A}$ one has
-\begin{equation*}
-\Pi_{\ctxext{A}{P}} \jdeq \Pi_A\circ \Pi_P
-\end{equation*}
-\item For any $Q\in\cat{F}/\ctxext{{\Gamma}{A}}{P}$, one has
-\begin{equation*}
-{\evtm{\ctxext{A}{P}}{Q}}
-  \jdeq
-{\subst{\evtm{A}{\mprd{P}{Q}}}{\ctxwk{\mprd{A}{\mprd{P}{Q}}}{\evtm{P}{Q}}}}
-\end{equation*}
-\end{enumerate}
-\item 
-\begin{enumerate}
-\item For any $\Gamma\in\mathbb{E}$, one has
-\begin{equation*}
-\Pi_{\catid{\Gamma}}\jdeq \catid{\mathbb{E}/\Gamma}.
-\end{equation*}
-\item For any $A\in\mathbb{F}/\Gamma$, one has
-\begin{equation*}
-\tfev^{\catid{\Gamma},A} \jdeq \tfid{A}
-\end{equation*}
-\end{enumerate}
-\item for any term $f:P$ in context $\ctxext{\Gamma}{A}$, one has
-\begin{equation*}
-\subst{\lambda_A(f)}{\tfev^{A,P}}\jdeq f.
-\end{equation*}
-\end{enumerate}
-\end{defn}
-
-\begin{rmk}
-The evaluation term determines a term
-\begin{equation*}
-\tilde\tfev^{A,P} : \thom{\ctxwk{A}{\Pi_A(P)}}{P}
-\end{equation*}
-\end{rmk}
-
-\begin{defn}
-A $\Pi$E-homomorphism from $\mathbb{E}$ to $\mathbb{E}'$ is an E-homomorphism
-$H:\mathbb{E}\to\mathbb{E}'$ such that
-\begin{enumerate}
-\item for any $A\in\cat{F}/\Gamma$, one has
-\begin{equation*}
-\begin{tikzcd}[column sep=large]
-\cat{F}/\ctxext{\Gamma}{A} \arrow[d,swap,"\Pi_A"] \arrow[r,"H/\ctxext{\Gamma}{A}"] & \cat{F}'/\ctxext{H(\Gamma)}{H(A)} \arrow[d,"\Pi_{H(A)}"] \\
-\cat{F}/\Gamma \arrow[r,swap,"H/\Gamma"] & \cat{F}'/F(\Gamma)
-\end{tikzcd}
-\end{equation*}
-\item For any $P\in\cat{F}/\ctxext{\Gamma}{A}$, one has
-\begin{equation*}
-H(\tfev^{A,P})\jdeq \tfev^{H(A),H(P)}
-\end{equation*}
-\end{enumerate}
-\end{defn}
-
-\bibliographystyle{plain}
-%\phantomsection\addcontentsline{toc}{section}{References}
-\bibliography{References/refs}
-
-\end{document}

### 2015-10-01 main.tex
typo Jacob*s -> Jacobs*
moved intro to new file

--- 2015-07-29 main.tex
@@
 
+\newcommand{\thetitle}{An essentially algebraic formulation of dependent type theory}
+\newcommand{\theauthor}{Egbert Rijke}
+\hypersetup{pdftitle={\thetitle},pdfauthor={\theauthor},pdfhighlight={/I}}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\title{An essentially algebraic formulation of dependent type theory}
-\author{Egbert Rijke\\ {\sc Carnegie Mellon University}}
+\title{\thetitle}
+\author{\theauthor\footnote{{\sc Carnegie Mellon University}, department of
+philosophy; \url{erijke@andrew.cmu.edu}}}
 \date{\today}
@@
 We then compare Voevodsky's B-systems, Dybjer's categories
-with families and Jacob's comprehension categories with E-systems. There is
+with families and Jacobs' comprehension categories with E-systems. There is
 a full and faithful functor from B-systems to E-systems, of which the image is
@@
 
-%\tableofcontents
+\tableofcontents
 
-\section{Introduction}
-
-\subsection*{Principles of mathematical fuseki}
-Mathematics is the study of abstract concepts such as space or quantity, and
-when we design a language for such a study, we want to be able to efficiently
-express things about the objects under investigation. The subject matter of
-this paper is a logical framework of dependent type theory.
-
-Pfenning \cite{PfenningLFintro} defines logical frameworks
-to be meta-languages for the formalization of deductive system. Examples of
-logical framework therefore include propositional logic, first-order predicate 
-logic, Russell's ramified type theory, the lambda-calculus, dependent type 
-theory, and so on. By considering models of a logical framework, it becomes
-itself the subject of mathematical study.
-
-It may actually be useful to see logical frameworks themselves
-as objects subject to mathematical investigation.
-
-Logical frameworks for mathematics are not set in stone.
-
- and neither to they
-remain the same once they are declared.
-
-When designing and maintaining a logical framework
-
-\subsection*{Essentially algebraic theories}
-An essentially algebraic theory is an equational theory of partial operations
-which may be ordered, and of which the domain of each operation is declared by 
-equations on the preceeding operations. The concept of essentially algebraic
-theories is due to Freyd, who writes that a critical feature of essentially
-algebraic theories is that their models are closed under 
-colimits \cite{Aspects_of_topoi}.
-
-In \cite{Adamek_Locally}, Ad\'amek and Rosicky
-define an essentially algebraic theory to be a list 
-\begin{equation*}
-(S,\Sigma,a,d,c,E,\Sigma_t,\mathrm{Def})
-\end{equation*} 
-consisting of a set $S$ of sorts, a set $\Sigma$ of operation symbols,
-an arity function $a:\Sigma\to\mathbf{Set}$ specifying the arity of each operation
-symbol, a map $d_\sigma:a(\sigma)\to S$
-for each $\sigma\in\Sigma$ specifying the sort of the domain of $\sigma$,
-a map $c:\Sigma\to S$ specifying the sort of the codomain of each operation
-symbol,
-a set $E$ of $\Sigma$-equations, a set $\Sigma_t\subseteq\Sigma$ of total
-operation symbols, and a function $\mathrm{Def}$ assigning to each operation
-symbol $\sigma:\Pi_{i\in a(\sigma)} s_i\to s$ in $\Sigma-\Sigma_t$ a set
-$\mathrm{Def}(\sigma)$ of $\Sigma_t$-equations in the standard variables
-$x_i\in V_s$, $i\in a(\sigma)$. Subsequently, they show that a category of models of an
-essentially algebraic theory is the same thing as a locally presentable 
-category.
-
-Every essentially algebraic theory $\mathbb{T}$ in the sense of Ad\'amek and Rosicky
-determines a syntactic category $\mathcal{C}_\mathbb{T}$, which has all finite
-limits, and $\mathbb{T}$ can be seen as a presentation of the category
-$\mathcal{C}_\mathbb{T}$. Then there is a natural equivalence between the
-category of models of $\mathbb{T}$ in a category $\cat{C}$ with finite limits,
-and the category of finite limit preserving functors $\mathcal{C}_\mathbb{T}\to
-\cat{C}$. Thus, in this sense an essentially algebraic theory \emph{is} just
-a category with finite limits.
-
-Because essentially algebraic theories have a pleasant theory of models, 
-phrasing a theory as an essentially algebraic one gives immediate access to
-a well-developed toolbox which can be used in its study. Therefore, recognizing
-that the theory of type dependency is essentially algebraic is a critical step
-in the development of its theory of models.
-
-\subsection*{Logical frameworks}
-
-\subsection*{Setting up the essentially algebraic theory of E-systems}
-The theory of type dependency as we present it here, consists of three sorts
-$C$, $F$ and $T$ for contexts, families and terms. There is a map 
-$\eft:F\to C$ which determines for each family, in which context it is a
-family; and there is a map $\ebd:T\to F$ which determines for each term, of
-which family it is a term. Thus, the backbones of an E-system can be represented
-as
-\begin{equation*}
-\begin{tikzcd}
-T \arrow[d,"\ebd"] \\ F \arrow[d,"\eft"] \\ C
-\end{tikzcd}
-\end{equation*}
-When $A$ is a family in context $\Gamma$, there is also a context $\ctxext{\Gamma}{A}$,
-called the \emph{context extension of $A$}. Context extension allows us to consider
-`families over families', where a family $P$ over $A$ in context $\Gamma$, would
-simply be a family $P$ over $\ctxext{\Gamma}{A}$. Context extension is represented
-by another function $\cftctxext : F\to C$.
-
-Since we want the rules of
-type theory to be valid in any context $\Gamma$, there is also a notion of family
-extension. Thus, for any family $P$ over $A$ in context $\Gamma$, there is a
-family $\ctxext{A}{P}$ in context $\Gamma$. In other words, we have a function
-$\cftfamext : F\times_{\cftctxext,\eft} F\to F$.
-
-Two associativity laws will require
-that context and family extension are compatible with each other: we have
-the equality $\ctxext{{\Gamma}{A}}{P}\jdeq\ctxext{\Gamma}{{A}{P}}$ of contexts,
-and the equality $\ctxext{{A}{P}}{Q}\jdeq\ctxext{A}{{P}{Q}}$ of families.
-
-Finally, there is an empty context $\emptyc\in C$, and an empty family $\emptyf[\Gamma]$
-in context $\Gamma$, for every $\Gamma\in C$. The empty context is such that context extension
-restricted to families over the empty context is an isomorphism, and the empty
-families are neutral with respect to context and family extension in the sense
-that $\ctxext{\Gamma}{\emptyf[\Gamma]}\jdeq\Gamma$, $\ctxext{\emptyf[\Gamma]}{A}
-\jdeq A$ and $\ctxext{A}{\emptyf[\ctxext{\Gamma}{A}]}\jdeq A$.
-
-One may recognize in these ingredients of E-systems a category with a
-terminal object, and a set $T(f)$ of terms for every morphism $f$ (satisfying
-no further axioms, so far). Ignoring the set of terms for the moment, $C$ is
-the set of objects and $F$ is the set of morphisms; $\eft$ assigns the codomain
-to each morphism and $\cftctxext$ assigns the domain to each  morphism. The
-operation $\cftfamext$ composes two morphisms. The first associativity rule
-asserts that the domain of the composition $\ctxext{A}{P}$ is the domain of
-$P$, and the second associativity rule asserts that composition is associative
-in the usual sense. The empty context satisfies exactly the properties of the
-terminal object, and the empty family $\emptyf[\Gamma]$ satisfies the properties
-of the identity morphisms.
-
-Thus, an E-system $\mathbf{E}$ will have an underlying category, which may be called the
-category of families of $\mathbf{E}$. The structure of type dependency will by
-built on top of this category. This structure will consist of an E-homomorphism
-$W_A$ called \emph{weakening by $A$}, for every family $A$ in context $\Gamma$,
-an E-homomorphism $S_x$ called \emph{substitution by $x$}, for every term
-$x$ of $A$ in context $\Gamma$, and units $\idtm{A}\in T(W_A(A))$ for every
-family $A$ in context $\Gamma$, satisfying a handful of further axioms. 
-The weakening and substitution homomorphisms
-will preserve the extension structure we have discussed above; in other words,
-they will be functors preserving the terminal object.
-
-\subsection*{The pre-categorical nature of E-systems}
-Although we start our definition of E-systems with a category with term structure,
-we do not set out to have a categorical notion of E-systems, in the sense that the structure of
-being an E-system can be transported along equivalences of categories. The goal
-in this paper is to describe the algebraic structures belonging to an essentially
-algebraic theory of type dependency, and this structure turns out to be pre-categorical
-rather than categorical. The fact that we've encountered a category at the outset
-of our definition of E-systems is in a sense coincidence, rather than born out
-of a desire to do type theory categorically and we would consider E-systems to
-be pre-categorical models of type theory rather than categorical models of type
-theory. The difference between pre-categories and categories is that equality
-of objects in a pre-category is considered to be (strict) equality, whereas
-equality of objects in a category is considered to be isomorphism of objects,
-and we have chosen to use the former notion of equality.
-
-\subsection*{Existing notions of categorical models of dependent type theory}
-
-When introducing a new notion of model of dependent type theory, one necessarily
-faces the task of comparing the new notion with existing notions, and there are
-quite a few.
-
-\begin{itemize}
-\item Contextual categories were introduced by Cartmell \cite{Cartmell1986} and were studied
-extensively by Voevodsky in \cite{VV_Csys_univ,VV_C-systems_monad,VV_C-systems_quotients}
-under the name C-systems.
-\item Categories with families, introduced by Dybjer in \cite{Dybjer1996}, model a
-theory of type dependency with sorts for contexts, context morphisms, types
-and terms. Among the virtues of categories with families are that its theory
-is essentially algebraic, and they are relatively simple to work with.
-
-Categories with families are models of a dependent type theory with a
-category of contexts, for each context $\Gamma$ a sort $\mathrm{Ty}(\Gamma)$ of
-types in context $\Gamma$, and for each type $A\in\mathrm{Ty}(\Gamma)$ a sort
-$\mathrm{Tm}(A)$ of terms of type $A$ in context $\Gamma$. We will see that
-the sorts $\mathrm{Tm}(A)$ are superfluous, because it will always be 
-naturally isomorphic to the set of sections of the projection morphism
-$p_A:\ctxext{\Gamma}{A}\to\Gamma$ in the category of contexts. The evident
-functor from categories with families to E-systems will forget that the contexts
-form a category, and therefore it will not be faithful.
-\item Comprehension categories were introduced by Jacobs in \cite{Jacobs1993}, as
-a minimal notion of categorical model in type theory, without really making
-precise what that means. 
-Comprehension categories are models of a dependent type theory with a
-category of context, and for each context $\Gamma$ a category of types. 
-\end{itemize}
-
-There is a simple test to quickly figure out some basic differences between different
-dependent type theories, asking what kind of structure remains if there are no
-types at all. In a contextual category, only the terminal object remains.
-However, when the presheaf $\mathrm{Ty}$ of a category with families is assumed
-to be empty, or when a category $\cat{E}$ fibered over some category is assumed
-to be empty, the underlying category of context remains. We shall see that
-E-systems are like contextual categories in this respect: only a singleton can
-be an E-system with no types. This complicates a thorough comparison between
-categories with families and comprehension categories on one side, and E-systems
-on the other. We shall construct functors from the categories of categories with
-families and comprehension categories directly preserving type dependency, but
-these functors will not be faithful.
-
-Thus, it seems that both categories with families and comprehension categories
-have more structure than needed to model dependent type theory.
-
-\subsection*{Overview and conventions}
-In \autoref{sec:esys_defn}, we give the categorical definition of E-system in 
-several stages. First we define weakening systems, then projection systems and
-substitution systems, and finally we bundle them together into the definition
-of E-systems. The structure of the definition of weakening systems is 
-prototypcial: we first define the notion of pre-weakening systems, which are
-categories that come with a pre-weakening structure; then we define the notion
-of pre-weakening homomorphism, which is a functor preserving the weakening
-structure and we note that it makes sense to ask for the property that the
-functors of which the weakening system consists, are also pre-weakening
-homomorphisms; and finally, adding this requirement will be give the definition
-of weakening systems. We go through this process for each of the definitions
-of weakening systems, projection systems, substitution systems and E-systems.
-
-Throughout this section, our focus is on (set-valued) models of the essentially algebraic
-theory of type dependency, and in fact we shall not give the syntactic
-presentation of the theory. A syntactic presentation of type theory in the
-usual type theoretic style and models in arbitrary categories with finite limits
-have been presented by the author during the Homotopy Type Theory workshop in
-Oxford, November 2014%
-\footnote{See \url{http://www.qmac.ox.ac.uk/events/Talk\%20slides/Egbert\%20Rijke.pdf}}.
-
-In \autoref{sec:esys_props}, we introduce the auxiliary notion of E'-system.
-An E'-system consists of a pre-category $\cat{C}$ with a sub-pre-category
-$\cat{F}$ with the same objects as $\cat{C}$, and a functorial choice of
-pullbacks for any morphism $f:\Delta\to\Gamma$ in $\cat{C}$ and any
-$A\in\cat{F}/\Gamma$. The goal of this section is to show that 
-E'-systems and E-systems are different axiomatizations of the same structures,
-and apart from that the goal is to show some equational reasoning in E-systems.
-The properties proven in this section also serve to provide some of the 
-intuition behind the constructions in \autoref{sec:esys_compared}.
-
-Although we have introduced a notion of E-homomorphism, a better behaved category
-of E-systems is the one where E-homomorphisms lift factorizations. The condition
-that a functor lifts factorizations is also algebraic, so we still have an
-essentially algebraic theory of E-systems if we require this.
-In \autoref{sec:esys_compared}, where we introduce the category of E-systems
-with E-homomorphisms lifting factorizations, we construct an embedding from Voevodsky's
-B-systems into E-systems and prove that its image is the (full) subcategory
-of stratified E-systems.
-
-\begin{rmk}
-In the present article we use $\jdeq$ for equality, because the equality we
-consider here will interpret the judgmental equality of type theory, 
-and we wish to reserve
-the more convenient symbol $=$ for the internal equality of E-systems, i.e.~for
-the identification types, as in \cite{TheBook}. 
-\end{rmk}
-
-\begin{rmk}
-Recall that the slice pre-category $\cat{C}/\Gamma/(\Delta,f)$ is isomorphic to the slice pre-category
-$\cat{C}/\Delta$, in the essentially algebraic theory of pre-categories. In the following,
-we shall usually identify these two categories. Recall also that for every functor
-$F:\cat{C}\to\cat{D}$ and every object $\Delta$ of $\cat{C}$, there is a functor
-$F/\Delta:\cat{C}/\Delta\to \cat{D}/F(\Delta)$ given by the action on morphisms of $F$. 
-\end{rmk}
-
-\begin{rmk}
-When we think of a category $\cat{F}$ as a category of fibrations or families
-over a context -- which will be the outset of the definition of E-systems --
-it will be convenient to denote the objects of the $\cat{F}$ with capital
-greek letters $\Gamma$, $\Delta$, \ldots, and the morphisms by capital latin letters
-$A$, $B$, $P$, $Q$, \ldots. When $A$ is a morphism into $\Gamma$, we will denote
-its domain by $\ctxext{\Gamma}{A}$, so we don't always have to introduce a name
-for the domain. When $A$ is a morphism into $\Gamma$ and
-$P$ is a morphism into $\ctxext{\Gamma}{A}$, it will also be convenient to denote the
-composition of $P$ with $A$ by $\ctxext{A}{P}$. Note that $\ctxext{A}{P}$
-is indeed the domain of $P$ in the slice category $\cat{F}/\Gamma$.
-\end{rmk}
-
-\begin{rmk}
-We have chosen not to use common type theoretical notation involving special
-expressions for judgments and for inference rules, because we want to focus
-on the algebraic structure of type dependency rather than on its syntax.
-
-We would also like to remark that although it appears that we introduce possibly
-large families of operations (a weakening operation $W_A$ for each morphism $A$,
-for instance), E-systems can actually be defined with finitely many
-operations. This becomes clearer when one defines an E-system in an arbitrary
-category with finite limits.
-\end{rmk}
-
-\subsection*{Acknowledgements}
-Several people have been indispensible throughout the research process.
-I would like to thank 
-Richard Garner for the early discussions on the inference rules of type 
-dependency we had in Barcelona and in Paris; 
-Joachim Kock for the many fruitful discussions and for hosting me at the Universitat
-Aut\`onoma de Barcelona during the year 2013-2014;
-Steve Awodey for sharing many insights and useful advice which lead to the
-definition of an E-system in arbitrary categories with finite limits;
-and Vladimir Voevodsky for inviting me to present E-systems at the Homotopy Type
-Theory workshop in Oxford (November 2014) and to visit the Institute for Advanced 
-Study in Princeton (March 2015), where the definition presented here was
-discovered.
+\input{introduction}
 
@@
 
+\input{theories}
+
+\begin{comment}
 \section{Type constructors in E-systems}
@@
 \end{defn}
+\end{comment}
 

### 2015-11-24 introduction.tex
added basic rules for equality

@@
 
+An essentially algebraic theory can be built step by step in the following way.
+First, one fixes the set $S$ of sorts. Next, one fixes a set $\Sigma_0$ of total
+operations.
+
 \subsection*{Setting up the essentially algebraic theory of E-systems}
@@
 
+In the specification of a formal language of type theory, we would therefore
+have three pairs of kinds of judgments that can be made:
+\begin{align*}
+\jalign\jctx{\Gamma} 
+& \jalign\jctxeq{\Gamma}{\Gamma'}
+  \\
+\jalign\jfam{\Gamma}{A} 
+& \jalign\jfameq{\Gamma}{A}{B}
+  \\
+\jalign\jterm{\Gamma}{A}{x} 
+& \jalign\jtermeq{\Gamma}{A}{x}{y}.
+\end{align*}
+These are asserting that $\Gamma$ is a context; that $\Gamma$ and $\Gamma'$ are
+judgementally equal contexts; that $A$ is a family in context $\Gamma$; that
+$A$ and $B$ are judgmentally equal families in context $\Gamma$; that $x$ is
+a term of $A$ in context $\Gamma$; and that $x$ and $y$ are judgmentally equal
+terms of $A$ in context $\Gamma$.
+
 An E-system is therefore an algebra consisting of three sorts: $C$ for contexts,
@@
 \end{equation*}
+This structure is represented by the following set of rules governing judgmental
+equality. First, judgmental equality is postulated to be an equivalence relation
+in all three cases (contexts, families and terms), so we require
+%\bgroup\small
+\begin{align*}
+& \inference
+  { \jctx{\Gamma}
+    }
+  { \jctxeq{\Gamma}{\Gamma}
+    } 
+& & \inference
+    { \jctxeq{\Gamma}{\Delta}
+      }
+    { \jctxeq{\Delta}{\Gamma}
+      } 
+& & \inference
+    { \jctxeq{\Gamma}{\Delta}
+      \jctxeq{\Delta}{\greek{E}}
+      }
+    { \jctxeq{\Gamma}{\greek{E}}
+      }
+    \\
+& \inference
+  { \jfam{\Gamma}{A}
+    }
+  { \jfameq{\Gamma}{A}{A}
+    } 
+& & \inference
+    { \jfameq{\Gamma}{A}{B}
+      }
+    { \jfameq{\Gamma}{B}{A}
+      }
+& & \inference
+    { \jfameq{\Gamma}{A}{B}
+      \jfameq{\Gamma}{B}{C}
+      }
+    { \jfameq{\Gamma}{A}{C}
+      }
+    \\
+& \inference
+  { \jterm{\Gamma}{A}{x}
+    }
+  { \jtermeq{\Gamma}{A}{x}{x}
+    }
+& & \inference
+    { \jtermeq{\Gamma}{A}{x}{y}
+      }
+    { \jtermeq{\Gamma}{A}{y}{x}
+      }
+& & \inference
+    { \jtermeq{\Gamma}{A}{x}{y}
+      \jtermeq{\Gamma}{A}{y}{z}
+      }
+    { \jtermeq{\Gamma}{A}{x}{z}
+      }
+\end{align*}
+%\egroup
+Second, there are convertibility rules assuring that the notion of being a
+family in a certain context, or a term of a family, is compatible with judgmental
+equality.
+\begin{align*}
+& \inference
+  { \jctxeq{\Gamma}{\Delta}
+    \jfam{\Gamma}{A}
+    }
+  { \jfam{\Delta}{A}
+    }
+& & \inference
+    { \jctxeq{\Gamma}{\Delta}
+      \jfameq{\Gamma}{A}{B}
+      }
+    { \jfameq{\Delta}{A}{B}
+      }
+    \\
+& \inference
+  { \jctxeq{\Gamma}{\Delta}
+    \jterm{\Gamma}{A}{x}
+    }
+  { \jterm{\Delta}{A}{x}
+    }
+& & \inference
+    { \jctxeq{\Gamma}{\Delta}
+      \jtermeq{\Gamma}{A}{x}{y}
+      }
+    { \jtermeq{\Delta}{A}{x}{y}
+      }
+    \\
+& \inference
+  { \jfameq{\Gamma}{A}{B}
+    \jterm{\Gamma}{A}{x}
+    }
+  { \jterm{\Gamma}{B}{x}
+    }
+& & \inference
+    { \jfameq{\Gamma}{A}{B}
+      \jtermeq{\Gamma}{A}{x}{y}
+      }
+    { \jtermeq{\Gamma}{B}{x}{y}
+      }
+\end{align*}
 When $A$ is a family in context $\Gamma$, there is also a context $\ctxext{\Gamma}{A}$,

### 2015-11-26 compared.tex
telescopic CWFs

@@
 A functor $F:\cat{C}\to\cat{D}$ is said to
-\emph{lift factorizations} if for any $f\in\cat{C}/X$, the functor
+\define{lift factorizations} if for any $f\in\cat{C}/X$, the functor
 $F_{\mathbf{fact}(f)}:\mathbf{fact}(f)\to\mathbf{fact}(F(f))$ is an isomorphism
@@
 \begin{defn}
-A category $\cat{C}$ with terminal object is said to be \emph{stratified} if there exists a 
+A category $\cat{C}$ with terminal object is said to be \define{stratified} if there exists a 
 \emph{stratification functor}
@@
 \begin{defn}
-A pre-E-system is said to be \emph{stratified} if its underlying category is
+A pre-E-system is said to be \define{stratified} if its underlying category is
 stratified and if each $W_A$ and $S_x$ is a stratified functor. The category of
@@
 \begin{defn}
-A \emph{B-framework} is a collection of data of the following form:
+A \define{B-framework} is a collection of data of the following form:
 \begin{enumerate}
@@
 \begin{defn}
-A homomorphism $H:\mathbb{B}\to\mathbb{A}$ of B-frameworks consists of maps
+A \define{homomorphism $H:\mathbb{B}\to\mathbb{A}$ of B-frameworks} consists of maps
 $H_n:B_n\to A_n$ and $\tilde{H}_n:\tilde{B}_n\to\tilde{A}_n$ such that
@@
 \begin{defn}
-A B0-system $\mathbb{B}$ consists of a B-framework $\mathbb{B}$ and homomorphisms
+A \define{B0-system} $\mathbb{B}$ consists of a B-framework $\mathbb{B}$ and homomorphisms
 \begin{align*}
@@
 \begin{defn}
-A B0-homomorphism $H:\mathbb{B}\to\mathbb{A}$ is a homomorphism of B-frameworks
+A \define{B0-homomorphism} $H:\mathbb{B}\to\mathbb{A}$ is a homomorphism of B-frameworks
 for which the diagrams
@@
 \begin{defn}
-A unital B0-system is a B0-system with maps
+A \define{unital B0-system} is a B0-system with maps
 \begin{equation*}
@@
 such that $\ebd(\delta_n(X))\jdeq W_{X}(X)$ for any $X\in B_n$. 
-A B0-homomorphism $H$ is said to be unital if one has
+A B0-homomorphism $H$ is said to be \define{unital} if one has
 \begin{equation*}
@@
 
-\subsection{Comprehension categories and categories with families}
-In this section we recall Dybjer's notion of categories with families and we
-introduce the equivalent notion of D-systems. After having proved that categories with families
-are the same things as D-systems, we will focus on constructing a canonical functor from
-D-systems to E'-systems, in which the category $\cat{F}$ of families has contexts
-as objects and compositions of projections as morphisms. There is no reason why
+\subsection{Telescopic categories with families}
+
+\begin{defn}
+A \define{category with families} consists of
+\begin{enumerate}
+\item A category $\cat{C}$ with a terminal object $1$.
+\item A presheaf $\mathrm{Ty}:\op{\cat{C}}\to\mathbf{Set}$. We will write
+$f_\ast(A)$ for $\mathrm{Ty}(f,A)$.
+\item For any $A\in\mathrm{Ty}(\Gamma)$, a morphism
+\begin{equation*}
+p_A : \ctxext{\Gamma}{A}\twoheadrightarrow \Gamma
+\end{equation*}
+in $\cat{C}$.
+\item a preasheaf $\mathrm{Tm}:\op{(\int_\cat{C}\mathrm{Ty})}\to\mathbf{Set}$.
+The action of a morphism $f:(\Delta,f_\ast(A))\to(\Gamma,A)$ of $\int_\cat{C}\mathrm{Ty}$
+on a term $x\in \mathrm{Tm}(A)$ is also written by $f_\ast(x)$.
+\item a term $\delta_A\in\mathrm{Tm}(\ctxext{\Gamma}{A},(p_A)_\ast(A))$ for every
+$A\in\mathrm{Ty}(\Gamma)$, 
+\end{enumerate}
+with the universal property that for every $f:\Delta\to\Gamma$ in $\cat{C}$, 
+the function
+\begin{equation*}
+\theta\mapsto\theta_\ast(\delta_A)
+  : \{\theta:\Delta\to\ctxext{\Gamma}{A}\mid p_A\circ\theta\jdeq f\}\to
+    \mathrm{Tm}(f_\ast(A))
+\end{equation*}
+is a bijection. We denote the inverse of this map by $\mu_{A,f}$. 
+\end{defn}
+
+There are of course many more sensible examples of categories with families,
+but the following class of examples causes us to doubt whether it is possible
+at all to have a faithful functor from the category of categories
+with families to the category of E-systems.
+
+\begin{eg}
+Every category $\cat{C}$ with a terminal object and the empty presheaf 
+$\mathrm{Ty}$, is a category with families.
+\end{eg}
+
+\begin{rmk}
+Although a terminal object is usually assumed to exist, not every object 
+of a D-system is assumed to be `fibrant', i.e.~it is
+not the case that every $X\to 1$ is a composition of projections $p_A$.
+We could only hope that there is a functor from D-sytems to E-systems without
+terminal objects.
+\end{rmk}
+
+There is essentially only one choice for the presheaf $\mathrm{Tm}$ and the
+term $\delta_A$, so the extra structure on a setup for categories with families
+is determined essentially uniquely once it exists.
+
+\begin{defn}\label{lem:cwf_to_dsys}
+Let $\mathbb{D}$ be a category with families. Then there is a natural isomorphism
+\begin{equation*}
+\varphi_A:\mathrm{Tm}(A)\cong\{\theta:\Gamma\to\ctxext{\Gamma}{A}\mid p_A\circ\theta\jdeq\catid{\Gamma}\}
+\end{equation*}
+in the presheaf category $\mathrm{Psh}(\int_{\cat{C}}\mathrm{Ty})$. 
+\begin{comment}
+The
+natural isomorphism $\varphi$ is such that the term $\delta_A\in\mathrm{Tm}((p_A)_\ast(A))$
+corresponds to the unique arrow $\tilde{q}_A$ which fits in the diagram
+\begin{equation*}
+\begin{tikzcd}
+\ctxext{\Gamma}{A} \arrow[ddr,bend right=15,equals] \arrow[drr,bend left=15,equals] \arrow[dr,"{\tilde \delta_A}"] \\
+& \ctxext{{\Gamma}{A}}{(p_A)_\ast(A)} \arrow[d,fib] & \ctxext{\Gamma}{A} \arrow[d,fib] \\
+& \ctxext{\Gamma}{A} \arrow[r,"{p_A}"] & \Gamma
+\end{tikzcd}
+\end{equation*}
+\end{comment}
+\end{defn}
+
+\begin{constr}
+We define the inverse of $\varphi_A$ by $\theta\mapsto\theta_\ast(\delta_A)$. Verifying
+that these form a natural isomorphism is straightforward.
+\end{constr}
+
+The definition of D-system we give below, is practically the same as the definition
+of category with attributes in \cite[Definition 3.10]{Hofmann_syntax_semantics}.
+Following Voevodsky, we wish to reserve the word `category' for usage in mathematical terms
+for notions which are `categorical' in the sense of being invariant under 
+equivalence of categories. The definition of categories with attributes involves
+an equality of functors (which would have to be replaced by a natural 
+isomorphism), so it is at least not obviously categorical.
+
+\begin{defn}
+A \define{D-system} (a category with attributes, see \cite{Hofmann_syntax_semantics}) consists of
+\begin{enumerate}
+\item A category $\cat{C}$ with a terminal object $1$.
+\item A presheaf $\mathrm{Ty}:\op{\cat{C}}\to\mathbf{Set}$. 
+\item A commuting triangle
+\begin{equation*}
+\begin{tikzcd}
+\int_{\cat{C}}\mathrm{Ty} \arrow[rr,"p"] \arrow[dr,"\pi"] & & \cat{C}^\square \arrow[dl,"\mathrm{cod}"] \\
+& \cat{C}
+\end{tikzcd}
+\end{equation*}
+of functors, where $\cat{C}^\square$ is the sub-pre-category of $\cat{C}^{\to}$ of
+which the morphisms are pullback squares. For any $A\in\mathrm{Ty}(\Gamma)$,
+the morphism $p(\Gamma,A)$ will be denoted by $p_A$; for any
+$(f,A):(\Delta,f_\ast(A))\to(\Gamma,A)$ in $\int_{\cat{C}}\mathrm{Ty}$, the top
+morphism of the pullback square $p(f,A)$ will be denoted by $\pi_2(f,A)$.
+\end{enumerate}
+\end{defn}
+
+\begin{rmk}
+For any $f:\Delta\to\Gamma$, $g:X\to\Delta$ and $A\in\mathrm{Ty}(\Gamma)$ in a
+D-system $\mathbb{D}$, one has 
+$\pi_2(\catid{\Gamma},A)\jdeq\catid{\ctxext{\Gamma}{A}}$ and
+$\pi_2(f\circ g,A)\jdeq \pi_2(f,A)\circ\pi_2(g,f_\ast(A))$.
+\end{rmk}
+
+\begin{defn}
+A \define{D-homomorphism} from $\mathbb{D}$ to $\mathbb{D}'$ consists of a functor 
+$\mathcal{F}:\cat{C}\to\cat{C}'$ which maps the terminal object of
+$\cat{C}$ to the terminal object of $\cat{C}'$, and a natural transformation
+$\tau:\mathrm{Ty}\Rightarrow\mathrm{Ty}'\circ\mathcal{F}$, such that 
+$\mathcal{F}(p_A)\jdeq p_{\mathcal{F}(A)}$ for any
+$A\in\mathrm{Ty}(\Gamma)$, and $\mathcal{F}(\pi_2(f,A))\jdeq
+\pi_2(\mathcal{F}(f),\tau_\Gamma(A))$ for any $f:\Delta\to\Gamma$ in $\cat{C}$ and
+any $A\in\mathrm{Ty}(\Gamma)$. 
+\end{defn}
+
+The following lemma restates the last condition under the assumption that
+$\mathcal{F}$ preserves \emph{all} finite limits. Note that
+in the definition of D-homomorphism, the underlying functor is only
+required to preserve the pullbacks of the projections, of which there may be
+none.
+
+\begin{lem}
+Suppose $\mathbb{D}$ and $\mathbb{D}'$ are D-systems with underlying categories
+$\cat{C}$ and $\cat{C}'$, respectively. Then a pair consisting of a finite
+limit preserving functor
+$\mathcal{F}:\cat{C}\to\cat{C}'$ and a natural transformation
+$\tau:\mathrm{Ty}\to \mathrm{Ty'}\circ\mathcal{F}$ is a D-homomorphism if and
+only if $\mathcal{F}$ preserves the terminal object on the nose and the induced
+square
+\begin{equation*}
+\begin{tikzcd}
+\int_{\cat{C}}\mathrm{Ty} \arrow[r,"p"] \arrow[d,swap,"{\int_{\mathcal{F}}\tau}"] & \cat{C}^{\square} \arrow[d,"\mathcal{F}^\square"] \\
+\int_{\cat{C}'}\mathrm{Ty}' \arrow[r,"{p'}"] & {\cat{C}'}^\square
+\end{tikzcd}
+\end{equation*}
+commutes.
+\end{lem}
+
+\begin{proof}
+Straightforward.
+\end{proof}
+
+%\begin{lem}
+%In the definition of D-system, the assumed structure beyond a choice of
+%projections $p_A:\ctxext{\Gamma}{A}\twoheadrightarrow\Gamma$ is property-like.
+%\end{lem}
+
+The following theorem is due to Hofmann, \cite[section 3.2]{Hofmann_syntax_semantics}. 
+%In the proof that every category with families is a D-system, we
+%outline the successive steps. The omitted details are marked as omitted, and
+%they are straightforward to fill in.
+
+\begin{thm}\label{thm:DD}
+The category of categories with families is equivalent to the category of 
+D-systems.%\hfill$\square$
+\end{thm}
+
+\begin{comment}
+Since this theorem basically asserts that a category with families is a
+comprehension category
+\end{comment}
+
+\begin{proof}%[Outline of the proof]
+Let $\mathbf{CWF}$ be a category with families. The fact that $\mathbf{CWF}$ is
+a D-system can be shown by proving each of the assertions in the following six 
+steps.
+\begin{enumerate}
+\item For every $f:\Delta\to\Gamma$ and $A\in\mathrm{Ty}(\Gamma)$, there is an
+isomorphism
+\begin{equation*}
+\zeta_{A,f}:\{\theta:\Delta\to\ctxext{\Gamma}{A}\mid p_A\circ\theta\jdeq f\}
+  \cong
+\{\eta:\Delta\to\ctxext{\Delta}{f_\ast(A)}\mid p_{f_\ast(A)}\circ\eta\jdeq\catid{\Delta}\}
+\end{equation*}
+natural in $f$ in the sense that the equalities
+\begin{align*}
+\zeta_{A,\catid{\Gamma}}(\theta) & \jdeq \theta \\
+\zeta_{A,f\circ g}(\theta\circ g) & \jdeq \zeta_{f_\ast(A),g}(\zeta_{A,f}(\theta)\circ g)
+\end{align*}
+hold for any $f:\Delta\to\Gamma$ and $g:X\to\Delta$, and any $\theta:\Delta\to\ctxext{\Gamma}{A}$
+such that $p_A\circ\theta\jdeq f$.
+Furthermore, we get $\zeta_{A,p_A}(\catid{\ctxext{\Gamma}{A}})\jdeq\varphi_{(p_A)_\ast(A)}(\delta_A)$.
+
+\medskip
+%(Proof omitted).
+%\begin{comment}
+By \autoref{lem:cwf_to_dsys} we may define
+\begin{equation*}
+\zeta_{A,f}(\theta)\defeq \varphi_{f_\ast(A)}(\theta_\ast(\delta_A))
+\end{equation*}
+for any $\theta\in\{\theta:\Delta\to\ctxext{\Gamma}{A}\mid p_A\circ\theta\jdeq f\}$. 
+Since $\zeta_{A,f}$ is a composition of isomorphism, it is itself an isomorphism.
+Also, we get immediately that $\zeta_{A,\catid{\Gamma}}(\theta)\jdeq\theta$. To
+show the remaining identity:
+\begin{align*}
+\zeta_{A,f\circ g}(\theta\circ g)
+& \jdeq \varphi_{g_\ast(f_\ast(A))}((\theta\circ g)_\ast(\delta_A)) \\
+& \jdeq \varphi_{g_\ast(f_\ast(A))}(g_\ast(\theta_\ast(\delta_A))) \\
+& \jdeq \varphi_{g_\ast(f_\ast(A))}(g_\ast(\varphi_{f_\ast(A)}(\theta_\ast(\delta_A))_\ast(\delta_A))) \\
+& \jdeq \varphi_{g_\ast(f_\ast(A))}(g_\ast(\zeta_{A,f}(\theta)_\ast(\delta_A))) \\
+& \jdeq \varphi_{g_\ast(f_\ast(A))}((\zeta_{A,f}(\theta)\circ g)_\ast(\delta_A)) \\
+& \jdeq \zeta_{f_\ast(A),g}(\zeta_{A,f}(\theta)\circ g).
+\end{align*}
+This completes the proof that the isomorphisms $\zeta_{A,f}$ satisfy the required
+equalities.
+%\end{comment}
+
+%\begin{comment}
+\item
+For any $A\in\mathrm{Ty}(\Gamma)$, we define
+$W_A\defeq (p_A)_\ast:\cat{F}/\Gamma\to\cat{F}/\ctxext{\Gamma}{A}$. Then we get
+for any $A,B\in\cat{F}/\Gamma$ an isomorphism
+\begin{equation*}
+\mathrm{hom}_{\cat{C}/\Gamma}(A,B)
+  \cong
+T(W_A(B)).
+\end{equation*}
+In particular, we get a term $\idtm{A}\in T((p_A)_\ast(A))$. 
+\medskip
+
+Let $A,B\in\cat{F}/\Gamma$. Then we have the isomorphism
+\begin{align*}
+T(W_A(B)) & \jdeq \{\eta:\ctxext{\Gamma}{A}\to\ctxext{{\Gamma}{A}}{W_A(B)}\mid W_A(B)\circ\eta\jdeq\catid{\ctxext{\Gamma}{A}}\} \\
+  & \cong
+\{\theta:\ctxext{\Gamma}{A}\to\ctxext{\Gamma}{B}\mid B\circ\theta\jdeq A\},
+\end{align*}
+so we find $\idtm{A}\defeq\zeta_{A,A}(\catid{\ctxext{\Gamma}{A}})\in T(W_A(A))$.
+
+\item For any $A\in\mathrm{Ty}(\Gamma)$ and $f:\Delta\to\Gamma$ 
+there is a morphism $\pi_2(f,A):\ctxext{\Delta}{f_\ast(A)}\to\ctxext{\Gamma}{A}$
+such that $p_A\circ \pi_2(f,A)\jdeq f\circ p_{f_\ast(A)}$.
+
+\medskip
+%(Proof omitted).
+%\begin{comment}
+Note that we have the isomorphism
+\begin{align*}
+& \{ \theta : \ctxext{\Delta}{f_\ast(A)}\to\ctxext{\Gamma}{A}\mid p_A\circ\theta\jdeq f\circ p_{f_\ast(A)} \} \\
+  & \cong
+\{ \eta : \ctxext{\Delta}{f_\ast(A)}\to\ctxext{{\Delta}{f_\ast(A)}}{(f\circ p_{f_\ast(A)})_\ast(A)}
+  \mid (f\circ p_{f_\ast(A)})_\ast(A)\circ\eta\jdeq\catid{\ctxext{\Delta}{f_\ast(A)}}\} \\
+  & \jdeq
+\{ \eta : \ctxext{\Delta}{f_\ast(A)}\to\ctxext{{\Delta}{f_\ast(A)}}{(p_{f_\ast(A)})_\ast(f_\ast(A))}
+  \mid (f\circ p_{f_\ast(A)})_\ast(A)\circ\eta\jdeq\catid{\ctxext{\Delta}{f_\ast(A)}}\} \\
+  & \jdeq
+T((p_{f_\ast(A)})_\ast(f_\ast(A))).
+\end{align*}
+Since we have $\delta_{f_\ast(A)}\in T((p_{f_\ast(A)})_\ast(f_\ast(A)))$, we obtain 
+\begin{equation*}
+\pi_2(f,A)\defeq \zeta_{A,f\circ p_{f_\ast(A)}}^{-1}(\delta_{f_\ast(A)}):\ctxext{\Delta}{f_\ast(A)}\to\ctxext{\Gamma}{A}
+\end{equation*}
+satisfying $p_A\circ\pi_2(f,A)\jdeq f\circ p_{f_\ast(A)}$.
+%\end{comment}
+%\begin{comment}
+\item For any $f:\Delta\to\Gamma$, $g:X\to\Delta$ and $A\in\mathrm{Ty}(\Gamma)$, one has
+\begin{equation*}
+\pi_2(f,A)\circ\pi_2(g,f_\ast(A))\jdeq\pi_2(f\circ g,A).
+\end{equation*}
+
+\medskip
+%(Proof omitted).
+%\begin{comment}
+It suffices to show that
+\begin{equation*}
+\zeta_{A,f\circ g\circ p_{g_\ast(f_\ast(A))}}(\pi_2(f,A)\circ\pi_2(g,f_\ast(A)))\jdeq\delta_{g_\ast(f_\ast(A))}.
+\end{equation*}
+Because $g\circ p_{g_\ast(f_\ast(A))}\jdeq p_{f_\ast(A)}\circ\pi_2(g,f_\ast(A))$, we get
+to use the naturality of $\zeta$:
+\begin{align*}
+& \zeta_{A,f\circ g\circ p_{g_\ast(f_\ast(A))}}(\pi_2(f,A)\circ\pi_2(g,f_\ast(A))) \\
+& \jdeq
+\zeta_{A,f\circ p_{f_\ast(A)}\circ\pi_2(g,f_\ast(A))}(\pi_2(f,A)\circ\pi_2(g,f_\ast(A))) \\
+& \jdeq
+\zeta_{(p_{f_\ast(A)})_\ast(f_\ast(A)),\pi_2(g,f_\ast(A))}(\zeta_{A,f\circ p_{f_\ast(A)}}(\pi_2(f,a))\circ\pi_2(g,f_\ast(A))) \\
+& \jdeq
+\zeta_{(p_{f_\ast(A)})_\ast(f_\ast(A)),\pi_2(g,f_\ast(A))}(\delta_{f_\ast(A)}\circ\pi_2(g,f_\ast(A))) \\
+& \jdeq
+\zeta_{(p_{f_\ast(A)})_\ast(f_\ast(A)),\pi_2(g,f_\ast(A))}(\zeta_{f_\ast(A),p_{f_\ast(A)}}(\catid{\ctxext{\Delta}{f_\ast(A)}})\circ\pi_2(g,f_\ast(A))) \\
+& \jdeq
+\zeta_{f_\ast(A),p_{f_\ast(A)}\circ\pi_2(g,f_\ast(A))}(\catid{\ctxext{\Delta}{f_\ast(A)}}\circ\pi_2(g,f_\ast(A))) \\
+& \jdeq
+\zeta_{f_\ast(A),p_{f_\ast(A)}\circ\pi_2(g,f_\ast(A))}(\pi_2(g,f_\ast(A))) \\
+& \jdeq
+\zeta_{f_\ast(A),g\circ p_{g_\ast(f_\ast(A))}}(\pi_2(g,f_\ast(A)))\\
+& \jdeq
+\delta_{g_\ast(f_\ast(A))}.
+\end{align*}
+%\end{comment}\begin{comment}
+\item
+Let $f:\Delta\to\Gamma$ and $A\in\mathrm{Ty}(\Gamma)$. Then one has 
+\begin{equation*}
+\pi_2(f,A)\circ\zeta_{A,f}(\theta)\jdeq\theta
+\end{equation*}
+for any $\theta:\Delta\to\ctxext{\Gamma}{A}$ satisfying $A\circ\theta\jdeq f$.
+
+\medskip
+%(Proof omitted).
+%\begin{comment}
+We will write $\tilde\theta\defeq\zeta_{A,f}(\theta)$. Note that
+\begin{equation*}
+p_A\circ\pi_2(f,A)\circ\tilde\theta\jdeq f\circ p_{f_\ast(A)}\circ\tilde\theta\jdeq f.
+\end{equation*} 
+Therefore, we can prove that $\pi_2(f,A)\circ\tilde\theta\jdeq\theta$ by showing
+that $\zeta_{A,f}(\pi_2(f,A)\circ\tilde\theta)\jdeq\tilde\theta$. This is
+shown using the naturality of $\zeta$:
+\begin{align*}
+\zeta_{A,f}(\pi_2(f,A)\circ\tilde\theta)
+  & \jdeq
+\zeta_{A,f\circ p_{f_\ast(A)}\circ\tilde\theta}(\pi_2(f,A)\circ\tilde\theta)
+  \\
+  & \jdeq
+\zeta_{(p_{f_\ast(A)})_\ast(f_\ast(A)),\tilde{\theta}}(\zeta_{A,f\circ p_{f_\ast(A)}}(\pi_2(f,A))\circ\tilde\theta)
+  \\
+  & \jdeq
+\zeta_{(p_{f_\ast(A)})_\ast(f_\ast(A)),\tilde{\theta}}(\idtm{f_\ast(A)}\circ\tilde\theta)
+  \\
+  & \jdeq
+\zeta_{(p_{f_\ast(A)})_\ast(f_\ast(A)),\tilde{\theta}}(\zeta_{f_\ast(A),p_{f_\ast(A)}}(\catid{\ctxext{\Delta}{f_\ast(A)}})\circ\tilde\theta)
+  \\
+  & \jdeq
+\zeta_{f_\ast(A),p_{f_\ast(A)}\circ\tilde\theta}(\catid{\ctxext{\Delta}{f_\ast(A)}}\circ\tilde\theta) \\
+  & \jdeq
+\zeta_{f_\ast(A),\catid{\Delta}}(\tilde\theta) \\
+  & \jdeq 
+\theta.
+\end{align*}
+%\end{comment}\begin{comment}
+\item 
+The diagram
+\begin{equation*}
+\begin{tikzcd}[column sep=large]
+\ctxext{\Delta}{f_\ast(A)} \arrow[d,fib] \arrow[r,"{\pi_2(f,A)}"] & \ctxext{\Gamma}{A} \arrow[d,fib] \\
+\Delta \arrow[r,"f"] & \Gamma
+\end{tikzcd}
+\end{equation*}
+is a pullback square.
+
+Proof of (vi): To verify the universal mapping property of the pullback square, consider a
+commuting diagram
+\begin{equation*}
+\begin{tikzcd}
+X \arrow[d,swap,"g"] \arrow[r,"h"] & \ctxext{\Gamma}{A} \arrow[d,fib] \\
+\Delta \arrow[r,"f"] & \Gamma
+\end{tikzcd}
+\end{equation*}
+Then $h$ corresponds uniquely to a section $k$ of $(f\circ g)_\ast(A)$. Since
+$(f\circ g)_\ast(A)\jdeq g_\ast(f_\ast(A))$, the section $k$ fits in the
+commuting diagram
+\begin{equation*}
+\begin{tikzcd}[column sep=huge]
+\ctxext{X}{g_\ast(f_\ast(A))} \arrow[d,xshift=.7ex,fib] \arrow[r,"{\pi_2(g,f_\ast(A))}"]
+& \ctxext{\Delta}{f_\ast(A)} \arrow[d,fib] \arrow[r,"{\pi_2(f,A)}"]
+& \ctxext{\Gamma}{A} \arrow[d,fib] \\
+X \arrow[r,swap,"g"] \arrow[u,xshift=-.7ex,"k"] & \Delta \arrow[r,swap,"f"] & \Gamma
+\end{tikzcd}
+\end{equation*}
+so we obtain $u\defeq\pi_2(g,f_\ast(A))\circ k:X\to\ctxext{\Delta}{f_\ast(A)}$. 
+Then we have $p_{f_\ast(A)}\circ u\jdeq g\circ p_{g_\ast(f_\ast(A))}\circ k\jdeq g$ and
+$\pi_2(f,A)\circ u\jdeq \pi_2(f\circ g,A)\circ k\jdeq h$. Note that
+$u$ corresponds uniquely to $k$, which corresponds uniquely to $h$, so the
+uniqueness of $u$ is automatic.
+\end{enumerate}
+This completes the proof %outline 
+that $\mathbf{CWF}$ is a D-system.
+
+Now let $\mathbb{D}$ be a D-system. We define the presheaf $\mathrm{Tm}:
+\op{(\int_{\cat{C}}\mathrm{Ty})}\to\mathbf{Set}$ by
+\begin{equation*}
+\mathrm{Tm}(A) \defeq \{x:\Gamma\to\ctxext{\Gamma}{A}\mid p_A\circ x\jdeq \catid{\Gamma}\}.
+\end{equation*}
+For $f:\Delta\to\Gamma$ and $A\in\mathrm{Ty}(\Gamma)$, we get a function
+$x\mapsto f_\ast(x):\mathrm{Tm}(A)\to\mathrm{Tm}(f_\ast(A))$ by defining $f_\ast(x)$
+to be the unique morphism such that the diagram
+\begin{equation*}
+\begin{tikzcd}[column sep=huge] 
+\Delta \arrow[ddr,bend right=15,equals] \arrow[drr,bend left=15,"x\circ f"] \arrow[dr,densely dotted,near end,"{f_\ast(x)}"] \\
+& \ctxext{\Delta}{f_\ast(A)} \arrow[d,fib] \arrow[r,swap,"{\pi_2(f,A)}"] & \ctxext{\Gamma}{A} \arrow[d,fib] \\
+& \Delta \arrow[r,swap,"f"] & \Gamma
+\end{tikzcd}
+\end{equation*}
+commutes. A combination of the assumed uniqueness and the assumed equalities on
+$\pi_2(\blank,\blank)$ gives the functoriality of $\mathrm{Tm}$.
+
+The term $\delta_A\in\mathrm{Tm}((p_A)_\ast(A))$ is defined to be the unique morphism
+such that the diagram
+\begin{equation*}
+\begin{tikzcd}[column sep=large]
+\ctxext{\Gamma}{A} \arrow[drr,bend left=15,equals] \arrow[ddr,bend right=15,equals] \arrow[dr,densely dotted,near end,"{\delta_{A}}"] \\
+& \ctxext{{\Gamma}{A}}{(p_A)_\ast(A)} \arrow[r,swap,"{\pi_2(A,A)}"] \arrow[d,fib] & \ctxext{\Gamma}{A} \arrow[d,fib] \\
+& \ctxext{\Gamma}{A} \arrow[r,swap,"A"] & \Gamma
+\end{tikzcd}
+\end{equation*}
+commutes.
+
+The universality assumption of categories with families is a consequence of
+the universal property of pullbacks: the map $x\mapsto\pi_2(f,X)\circ x$
+from $\{x:\Delta\to\ctxext{\Delta}{f_\ast(A)}\mid p_{f_\ast(A)}\circ x\jdeq\catid{\Delta}\}$
+to $\{\theta:\Delta\to\ctxext{\Gamma}{A}\mid p_A\circ\theta\jdeq f\}$ is a bijection.
+\end{proof}
+%\end{comment}
+
+The following two definitions define subclasses of telescopic D-systems and 
+coherent telescopic D-systems, and the latter structure is the same as an E'-system
+(without empty context). In \autoref{defn:comp_to_telescopic}, we will show how
+to obtain a telescopic D-system from any comprehension category. This construction
+will also generate a telescopic D-system out of any D-system, because D-systems
+are comprehension categories.
+
+\begin{defn}
+Let $\cat{C}$ be a D-system. In this definition, we will write
+$\mathrm{Fam}$ for the presheaf of types. The D-system $\cat{C}$
+is said to be \define{telescopic}, if we have elements $\emptyf[\Gamma]\in
+\mathrm{Fam}(\Gamma)$ for each $\Gamma\in\cat{C}$, and maps 
+$P\mapsto\ctxext{A}{P}:\mathrm{Fam}(\ctxext{\Gamma}{A})\to \mathrm{Fam}(\Gamma)$
+such that
+\begin{enumerate}
+\item 
+\begin{enumerate}
+\item for any $A\in\mathrm{Fam}(\Gamma)$ and $P\in\mathrm{Fam}(\ctxext{\Gamma}{A})$,
+one has $\ctxext{\Gamma}{{A}{P}}\jdeq\ctxext{{\Gamma}{A}}{P}$, and
+\begin{equation*}
+p_{\ctxext{A}{P}}\jdeq p_A\circ p_P,
+\end{equation*}
+\item and for any $\Gamma\in\cat{C}$ one has $\ctxext{\Gamma}{\emptyf[\Gamma]}\jdeq\Gamma$, and
+\begin{equation*}
+p_{\emptyf[\Gamma]}\jdeq\catid{\Gamma}.
+\end{equation*}
+\end{enumerate}
+\item 
+\begin{enumerate}
+\item for any $A\in\mathrm{Fam}(\Gamma)$, $P\in\mathrm{Fam}(\ctxext{\Gamma}{A})$
+and $Q\in\mathrm{Fam}(\ctxext{{\Gamma}{A}}{P})$, one has
+\begin{equation*}
+\ctxext{{A}{P}}{Q}\jdeq\ctxext{A}{{P}{Q}},
+\end{equation*}
+\item and for any $A\in\mathrm{Fam}(\Gamma)$, one has
+\begin{equation*}
+\ctxext{\emptyf[\Gamma]}{A}\jdeq\ctxext{A}{\emptyf[\ctxext{\Gamma}{A}]}\jdeq A.
+\end{equation*}
+\end{enumerate}
+\item 
+\begin{enumerate}
+\item for any $A\in\mathrm{Fam}(\Gamma)$, $P\in\mathrm{Fam}(\ctxext{\Gamma}{A})$
+and any $f:\Delta\to\Gamma$ in $\cat{C}$, one has
+\begin{equation*}
+f_\ast(\ctxext{A}{P})\jdeq \ctxext{(f_\ast(A))}{(\pi_2(f,A)_\ast(P))},
+\end{equation*}
+\item and for any $f:\Delta\to\Gamma$ in $\cat{C}$, one has
+\begin{equation*}
+f_\ast(\emptyf[\Gamma])\jdeq\emptyf[\Delta].
+\end{equation*}
+\end{enumerate}
+\end{enumerate} 
+\end{defn}
+
+\begin{lem}
+A D-system is telescopic if and only if  the following two conditions hold:
+\begin{enumerate}
+\item There is a section $\phi:\cat{C}\to \int_{\cat{C}}\mathrm{Ty}$ of $\pi$.
+\item For every $P\in\mathrm{Ty}(\ctxext{\Gamma}{A})$,
+the morphism $p_A:\ctxext{\Gamma}{A}\to\Gamma$ has a unique cocartesian lift
+at $P$.
+\end{enumerate}
+\end{lem}
+
+\begin{proof}
+Let $\mathbb{D}$ be a telescopic D-system. To show that $\mathbb{D}$ has the two
+stated properties:
+\begin{enumerate}
+\item
+\item We have to show that each for $P\in\mathrm{Ty}(\ctxext{\Gamma}{A})$,
+the projection $p_A:\ctxext{\Gamma}{A}\to\Gamma$ has a unique cocartesian lift
+at $P$. Note that the morphism $(\ctxext{\Gamma}{A},P)\to (\Gamma,\ctxext{A}{P})$
+in $\int_{\cat{C}}\mathrm{Ty}$ 
+\end{enumerate}
+
+Let $\mathbb{D}$ be a D-system for which the stated two conditions hold. To show
+that $\mathbb{D}$ is a telescopic D-system, we systematically go through each of the
+properties:
+\begin{enumerate}
+\item 
+\end{enumerate} 
+\end{proof}
+
+\begin{conj}
+The notion of being telescopic is a property.
+\end{conj}
+
+\begin{defn}
+A telescopic D-system is said to be \define{coherent}, if for any $A,A'\in\mathrm{Fam}(\Gamma)$
+such that $p_A\jdeq p_{A'}$, and any $f:\Delta\to\Gamma$, we have
+$p_{f_\ast(A)}\jdeq p_{f_\ast(A')}$ and $\pi_2(f,A)\jdeq \pi_2(f,A')$.
+\end{defn}
+
+\begin{thm}
+An E'-system without empty context is the same thing as a coherent telescopic D-system.
+\end{thm}
+
+\begin{proof}
+Let $\mathbb{E}$ be an E'-system without empty context. Then we define the
+presheaf $\mathrm{Fam}$ on $\cat{C}$ on objects by
+\begin{equation*}
+\mathrm{Fam}(\Gamma)\defeq \mathrm{Obj}(\cat{F}/\Gamma)
+\end{equation*}
+and on morphisms by the action on objects of $f_\ast$, for each $f:\Delta\to
+\Gamma$ in $\cat{C}$. By condition (b) and (c) in the definition of E'-systems, it
+follows that this defines a contravariant functor.
+
+We define a functor $\mathcal{P}:\int_{\cat{C}}\to\cat{C}^\square$ by sending
+the pair $(\Gamma,A)$ to the morphism $A:\ctxext{\Gamma}{A}\to\Gamma$ in
+$\cat{C}$, and by sending $f:(\Delta,f_\ast(A))\to (\Gamma,A)$ to the pullback
+square
+\begin{equation*}
+\begin{tikzcd}[column sep=large]
+\ctxext{\Delta}{f_\ast(A)} \arrow[d,swap,"{f_\ast(A)}"] \arrow[r,"{\pi_2(f,A)}"] &
+\ctxext{\Gamma}{A} \arrow[d,"A"] \\ \Delta \arrow[r,"f"] & \Gamma
+\end{tikzcd}
+\end{equation*}
+This is also functorial because of conditions (b) and (c) in the definition of
+E'-systems. Thus, we have defined a D-system
+\begin{equation*}
+\begin{tikzcd}
+\int_{\cat{C}}\mathrm{Fam} \arrow[rr,"{\mathcal{P}}"] \arrow[dr,"{\pi}"] & & \cat{C}^\square \arrow[dl,"{\mathrm{cod}}"] \\ & \cat{C}
+\end{tikzcd}
+\end{equation*}
+and it is straightforward to verify that this D-system is telescopic and coherent.
+
+Conversely, suppose $\mathbb{D}$ is a telescopic D-system with underlying category
+$\cat{C}$. Define $\cat{F}$ to be the category with the same objects as
+$\cat{C}$, and morphisms of the form $p_A$. To check that this forms a category,
+note that $p_{\emptyf[\Gamma]}\jdeq\catid{\Gamma}$ and $p_A\circ p_P\jdeq
+p_{\ctxext{A}{P}}$.
+
+Now let $f:\Delta\to\Gamma$ be a morphism in $\cat{C}$. For any $A\in\cat{F}/\Gamma$,
+we have an element $\tilde{A}\in\mathrm{Fam}(\Gamma)$ such that $p_{\tilde{A}}
+\jdeq A$, and we can form a pullback square
+\begin{equation*}
+\begin{tikzcd}[column sep=large]
+\ctxext{\Delta}{f_\ast(\tilde{A})} \arrow[r,"{\pi_2(f,\tilde{A})}"] \arrow[d,fib,swap,"p_{f_\ast(\tilde{A})}"]
+& \ctxext{\Gamma}{\tilde{A}} \arrow[d,fib,"{p_{\tilde{A}}}"] \\
+\Delta \arrow[r,"f"] & \Gamma
+\end{tikzcd}
+\end{equation*}
+By the coherence condition, this choice of pullback square does not depend on the
+choice of $\tilde{A}$ such that $p_{\tilde{A}}\jdeq A$ (having this fact 
+available is precisely what the coherence condition is for). The verification of conditions
+(a) to (d) of \autoref{defn:E'sys} is straightforward.
+\end{proof}
+
+\subsection{Coherent comprehension categories}
+In this section we recall Jacobs' notion of comprehension categories, and 
+Dybjer's notion of categories with families. It is shown in 
+\cite{Hofmann_syntax_semantics} that the category of categoriew with families
+is equivalent to the category of categories with attributes, which forms a
+subcategory of comprehension categories. Thus, for our purpose of comparing
+E-systems with various existing notions of models of type dependency, it suffice
+to make the comparison with comprehension categories.
+
+In this section we do the preparateions for this comparison. After having
+introduced categories with families, we prove that categories with families
+with the extra structure of `being closed under family extension' in a coherent
+way, are E'-systems without empty context.
+
+\begin{defn}
+Let $p:\cat{E}\to\cat{B}$ be a functor and let $f:\Delta\to \Gamma$ be a morphism in $\cat{B}$.
+A morphism $g:A\to B$ in $\cat{E}$ is said to be \define{cartesian over $f$}, if
+\begin{enumerate}
+\item $p(g)\jdeq f$, and 
+\item For all $g':A'\to B$ with $p(g')\jdeq f$, there is a unique $u:A'\to A$
+such that $g'=g\circ u$ and $p(u)\jdeq\catid{\Delta}$. 
+\end{enumerate}
+Dually, a morphism $g:A\to B$ in $\cat{E}$ is said to be \define{cocartesian over
+$f$}, if
+\begin{enumerate}
+\item $p(g)\jdeq f$, and
+\item For all $g':A\to B'$ with $p(g')\jdeq f$, there is a unique $v:B\to B'$
+such that $g'=v\circ g$ and $p(v)\jdeq\catid{\Gamma}$. 
+\end{enumerate}
+A (co)cartesian morphism over $f$ is also called a \define{(co)cartesian lift of $f$}.
+\end{defn}
+
+\begin{defn}
+A functor $p:\cat{E}\to\cat{B}$ is called a \define{Grothendieck fibration} if
+every morphism into an object of the form $p(A)$ for some object $A$ of $\cat{E}$ 
+has a cartesian lift, and cartesian morphisms are closed under composition.
+\end{defn}
+
+\begin{defn}
+A Grothendieck fibration $p:\cat{E}\to\cat{B}$ is said to be \define{split}, if
+there is a functorial choice of cartesian lifts, i.e.~for every $f:\Delta\to \Gamma$ in 
+$\cat{B}$ and any $B\in \cat{E}$ such that $p(B)\jdeq \Gamma$,
+there is a cartesian lift $j(B,f):A \to B$ for some object $A$ of $\cat{E}$, such that 
+$j(B,\idfunc[\Gamma])\jdeq\catid{B}$ and 
+$j(B, f\circ g)\jdeq j(B,f)\circ j(\mathrm{dom}(j(B,f)),g)$ for each
+$f:\Delta\to\Gamma$ and each $g:\greek{E}\to\Delta$. 
+\end{defn}
+
+\begin{eg}
+For any presheaf $P:\cat{C}\to\mathbb{Set}$, the projection functor
+$\int_\cat{C} P\to\cat{C}$ is a Grothendieck fibration.
+\end{eg}
+
+\begin{defn}
+A \define{comprehension category} $\mathcal{P}$ consists of categories $\cat{E}$, $\cat{C}$ and a
+commuting triangle
+\begin{equation*}
+\begin{tikzcd}
+\cat{E} \arrow[rr,"\mathcal{P}"] \arrow[dr,swap,"p"] & & \cat{C}^{\to} \arrow[dl,"\mathrm{cod}"] \\
+& \cat{C}
+\end{tikzcd}
+\end{equation*}
+of functors, such that $p$ is a Grothendieck fibration, and $\mathcal{P}$ maps
+cartesian morphisms in $\cat{E}$ to pullback squares in $\cat{C}$. We say that
+$\mathcal{P}$ is a \define{full} comprehension category if the functor $\mathcal{P}:
+\cat{E}\to\cat{C}^{\to}$ is full and faithful; and we say that
+$\mathcal{P}$ is \define{split} if $p$ is split.
+\end{defn}
+
+\begin{eg}
+Since any morphism of $\int_{\cat{C}}\mathrm{Ty}$ is cartesian, it follows that D-systems
+are comprehension categories. As D-systems are built up from a presheaf of types
+on the underlying category, they are in fact discrete comprehension categories.
+\end{eg}
+
+Each comprehension category determines a subcategory $\cat{F}$ generated by
+the projections, as we shall see soon. However, there is no reason why
 this category should possess a terminal object, so we naturally have to retreat
-to E'-systems without empty contex. However, the construction of a
-canonical functor in which families are represented by compositions of projections
-is still not possible without a further coherence condition. To make sure that
-the pullback operations of a D-system extend to the category
-$\cat{F}$, we introduce \emph{coherent} D-systems. So we only have a functor $U$
+to E'-systems without empty contex. Even when we make this generalization, 
+the construction of a canonical functor in which families are represented by 
+compositions of projections is still not possible without a further coherence 
+condition. To make sure that
+the pullback operations of a comprehension category extend to the category
+$\cat{F}$, we introduce \emph{coherent} comprehension categories. So we only have a functor $U$
 from the category $\mathbf{Dsys}$ of coherent D-systems to the category
@@
 
-\begin{defn}
-Let $p:\cat{E}\to\cat{B}$ be a functor and let $f:\Delta\to \Gamma$ be a morphism in $\cat{B}$.
-A morphism $g:A\to B$ in $\cat{E}$ is said to be \emph{cartesian over $f$}, if
-\begin{enumerate}
-\item $p(g)\jdeq f$, and 
-\item For all $g':A'\to B$ with $p(g')\jdeq f$, there is a unique $u:A'\to A$
-such that $g'=g\circ u$ and $p(u)\jdeq\catid{\Delta}$. 
-\end{enumerate}
-Dually, a morphism $g:A\to B$ in $\cat{E}$ is said to be \emph{cocartesian over
-$f$}, if
-\begin{enumerate}
-\item $p(g)\jdeq f$, and
-\item For all $g':A\to B'$ with $p(g')\jdeq f$, there is a unique $v:B\to B'$
-such that $g'=v\circ g$ and $p(v)\jdeq\catid{\Gamma}$. 
-\end{enumerate}
-A (co)cartesian morphism over $f$ is also called a \emph{(co)cartesian lift of $f$}.
-\end{defn}
-
-\begin{defn}
-A functor $p:\cat{E}\to\cat{B}$ is called a \emph{Grothendieck fibration} if
-every morphism into an object of the form $p(A)$ for some object $A$ of $\cat{E}$ 
-has a cartesian lift, and cartesian morphisms are closed under composition.
-\end{defn}
-
-\begin{defn}
-A Grothendieck fibration $p:\cat{E}\to\cat{B}$ is said to be \emph{split}, if
-there is a functorial choice of cartesian lifts, i.e.~for every $f:\Delta\to \Gamma$ in 
-$\cat{B}$ and any $B\in \cat{E}$ such that $p(B)\jdeq \Delta$,
-there is a cartesian lift $j(B,g):A \to B$ for some object $A$ of $\cat{E}$, such that 
-$j(B,p(B))\jdeq\catid{B}$ and $j(B, g\circ h)\jdeq j(B,g)\circ j(\mathrm{dom}(j(B,g)),h)$.
-\end{defn}
-
-\begin{eg}
-For any presheaf $P:\cat{C}\to\mathbb{Set}$, the projection functor
-$\int_\cat{C} P\to\cat{C}$ is a Grothendieck fibration.
-\end{eg}
-
-\begin{defn}
-A comprehension category $\mathcal{P}$ consists of categories $\cat{E}$, $\cat{C}$ and a
-commuting triangle
-\begin{equation*}
-\begin{tikzcd}
-\cat{E} \arrow[rr,"\mathcal{P}"] \arrow[dr,swap,"p"] & & \cat{C}^{\to} \arrow[dl,"\mathrm{cod}"] \\
-& \cat{C}
-\end{tikzcd}
-\end{equation*}
-of functors, such that $p$ is a Grothendieck fibration, and $\mathcal{P}$ maps
-cartesian morphisms in $\cat{E}$ to pullback squares in $\cat{C}$. We say that
-$\mathcal{P}$ is a full comprehension category if the functor $\mathcal{P}:
-\cat{E}\to\cat{C}^{\to}$ is full and faithful; and we say that
-$\mathcal{P}$ is split if $p$ is split.
-\end{defn}
-
-We now recall the notion of categories with families, which forms a class of
-examples of comprehension categories.
-
-\begin{defn}
-A \emph{category with families} consists of
-\begin{enumerate}
-\item A category $\cat{C}$ with a terminal object $1$.
-\item A presheaf $\mathrm{Ty}:\op{\cat{C}}\to\mathbf{Set}$. We will write
-$f_\ast(A)$ for $\mathrm{Ty}(f,A)$.
-\item For any $A\in\mathrm{Ty}(\Gamma)$, a morphism
-\begin{equation*}
-p_A : \ctxext{\Gamma}{A}\twoheadrightarrow \Gamma
-\end{equation*}
-in $\cat{C}$.
-\item a preasheaf $\mathrm{Tm}:\op{(\int_\cat{C}\mathrm{Ty})}\to\mathbf{Set}$.
-The action of a morphism $f:(\Delta,f_\ast(A))\to(\Gamma,A)$ of $\int_\cat{C}\mathrm{Ty}$
-on a term $x\in \mathrm{Tm}(A)$ is also written by $f_\ast(x)$.
-\item a term $\delta_A\in\mathrm{Tm}(\ctxext{\Gamma}{A},(p_A)_\ast(A))$ for every
-$A\in\mathrm{Ty}(\Gamma)$, 
-\end{enumerate}
-with the universal property that for every $f:\Delta\to\Gamma$ in $\cat{C}$, 
-the function
-\begin{equation*}
-\theta\mapsto\theta_\ast(\delta_A)
-  : \{\theta:\Delta\to\ctxext{\Gamma}{A}\mid p_A\circ\theta\jdeq f\}\to
-    \mathrm{Tm}(f_\ast(A))
-\end{equation*}
-is a bijection. We denote the inverse of this map by $\mu_{A,f}$. 
-\end{defn}
-
-There are of course many more sensible examples of categories with families,
-but the following class of examples causes us to doubt whether it is possible
-at all to have a faithful functor from the category of categories
-with families to the category of E-systems.
-
-\begin{eg}
-Every category $\cat{C}$ with a terminal object and the empty presheaf 
-$\mathrm{Ty}$, is a category with families.
-\end{eg}
-
-\begin{rmk}
-Although a terminal object is usually assumed to exist, not every object 
-of a D-system is assumed to be `fibrant', i.e.~it is
-not the case that every $X\to 1$ is a composition of projections $p_A$.
-We could only hope that there is a functor from D-sytems to E-systems without
-terminal objects.
-\end{rmk}
-
-There is essentially only one choice for the presheaf $\mathrm{Tm}$ and the
-term $\delta_A$, so the extra structure on a setup for categories with families
-is determined essentially uniquely once it exists. In other words, being a
-category with families is a property on setups for categories with families.
-
-\begin{defn}\label{lem:cwf_to_dsys}
-Let $\mathbb{D}$ be a category with families. Then there is a natural isomorphism
-\begin{equation*}
-\varphi_A:\mathrm{Tm}(A)\cong\{\theta:\Gamma\to\ctxext{\Gamma}{A}\mid p_A\circ\theta\jdeq\catid{\Gamma}\}
-\end{equation*}
-in the presheaf category $\mathrm{Psh}(\int_{\cat{C}}\mathrm{Ty})$. 
-\begin{comment}
-The
-natural isomorphism $\varphi$ is such that the term $\delta_A\in\mathrm{Tm}((p_A)_\ast(A))$
-corresponds to the unique arrow $\tilde{q}_A$ which fits in the diagram
-\begin{equation*}
-\begin{tikzcd}
-\ctxext{\Gamma}{A} \arrow[ddr,bend right=15,equals] \arrow[drr,bend left=15,equals] \arrow[dr,"{\tilde \delta_A}"] \\
-& \ctxext{{\Gamma}{A}}{(p_A)_\ast(A)} \arrow[d,fib] & \ctxext{\Gamma}{A} \arrow[d,fib] \\
-& \ctxext{\Gamma}{A} \arrow[r,"{p_A}"] & \Gamma
-\end{tikzcd}
-\end{equation*}
-\end{comment}
-\end{defn}
-
-\begin{constr}
-We define the inverse of $\varphi_A$ by $\theta\mapsto\theta_\ast(\delta_A)$. Verifying
-that these form a natural isomorphism is straightforward.
-\end{constr}
-
-The definition of D-system we give below, is practically the same as the definition
-of category with attributes in \cite[Definition 3.10]{Hofmann_syntax_semantics}.
-Following Voevodsky, we wish to reserve the word `category' for usage in mathematical terms
-for notions which are `categorical' in the sense of being invariant under 
-equivalence of categories. The definition of categories with attributes involves
-an equality of functors (which would have to be replaced by a natural 
-isomorphism), so it is at least not obviously categorical.
-
-\begin{defn}
-A \emph{D-system} (a category with attributes, see \cite{Hofmann_syntax_semantics}) consists of
-\begin{enumerate}
-\item A category $\cat{C}$ with a terminal object $1$.
-\item A presheaf $\mathrm{Ty}:\op{\cat{C}}\to\mathbf{Set}$. 
-\item A commuting triangle
-\begin{equation*}
-\begin{tikzcd}
-\int_{\cat{C}}\mathrm{Ty} \arrow[rr,"p"] \arrow[dr,"\pi"] & & \cat{C}^\square \arrow[dl,"\mathrm{cod}"] \\
-& \cat{C}
-\end{tikzcd}
-\end{equation*}
-of functors, where $\cat{C}^\square$ is the sub-pre-category of $\cat{C}^{\to}$ of
-which the morphisms are pullback squares. For any $A\in\mathrm{Ty}(\Gamma)$,
-the morphism $p(\Gamma,A)$ will be denoted by $p_A$; for any
-$(f,A):(\Delta,f_\ast(A))\to(\Gamma,A)$ in $\int_{\cat{C}}\mathrm{Ty}$, the top
-morphism of the pullback square $p(f,A)$ will be denoted by $\pi_2(f,A)$.
-\end{enumerate}
-\end{defn}
-
-\begin{rmk}
-For any $f:\Delta\to\Gamma$, $g:X\to\Delta$ and $A\in\mathrm{Ty}(\Gamma)$ in a
-D-system $\mathbb{D}$, one has 
-$\pi_2(\catid{\Gamma},A)\jdeq\catid{\ctxext{\Gamma}{A}}$ and
-$\pi_2(f\circ g,A)\jdeq \pi_2(f,A)\circ\pi_2(g,f_\ast(A))$.
-\end{rmk}
-
-\begin{defn}
-A \emph{D-homomorphism from $\mathbb{D}\to\mathbb{D}'$} consists of a functor 
-$\mathcal{F}:\cat{C}\to\cat{C}'$ which maps the terminal object of
-$\cat{C}$ to the terminal object of $\cat{C}'$, and a natural transformation
-$\tau:\mathrm{Ty}\Rightarrow\mathrm{Ty}'\circ\mathcal{F}$, such that 
-$\mathcal{F}(p_A)\jdeq p_{\mathcal{F}(A)}$ for any
-$A\in\mathrm{Ty}(\Gamma)$, and $\mathcal{F}(\pi_2(f,A))\jdeq
-\pi_2(\mathcal{F}(f),\tau_\Gamma(A))$ for any $f:\Delta\to\Gamma$ in $\cat{C}$ and
-any $A\in\mathrm{Ty}(\Gamma)$. 
-\end{defn}
-
-The following lemma restates the last condition under the assumption that
-$\mathcal{F}$ preserves \emph{all} finite limits. Note that
-in the definition of D-homomorphism, the underlying functor is only
-required to preserve the pullbacks of the projections, of which there may be
-none.
-
-\begin{lem}
-Suppose $\mathbb{D}$ and $\mathbb{D}'$ are D-systems with underlying categories
-$\cat{C}$ and $\cat{C}'$, respectively. Then a pair consisting of a finite
-limit preserving functor
-$\mathcal{F}:\cat{C}\to\cat{C}'$ and a natural transformation
-$\tau:\mathrm{Ty}\to \mathrm{Ty'}\circ\mathcal{F}$ is a D-homomorphism if and
-only if $\mathcal{F}$ preserves the terminal object on the nose and the induced
-square
-\begin{equation*}
-\begin{tikzcd}
-\int_{\cat{C}}\mathrm{Ty} \arrow[r,"p"] \arrow[d,swap,"{\int_{\mathcal{F}}\tau}"] & \cat{C}^{\square} \arrow[d,"\mathcal{F}^\square"] \\
-\int_{\cat{C}'}\mathrm{Ty}' \arrow[r,"{p'}"] & {\cat{C}'}^\square
-\end{tikzcd}
-\end{equation*}
-commutes.
-\end{lem}
-
-\begin{proof}
-Straightforward.
-\end{proof}
-
-%\begin{lem}
-%In the definition of D-system, the assumed structure beyond a choice of
-%projections $p_A:\ctxext{\Gamma}{A}\twoheadrightarrow\Gamma$ is property-like.
-%\end{lem}
-
-The following theorem is due to Hofmann, \cite[section 3.2]{Hofmann_syntax_semantics}. 
-%In the proof that every category with families is a D-system, we
-%outline the successive steps. The omitted details are marked as omitted, and
-%they are straightforward to fill in.
-
-\begin{thm}\label{thm:DD}
-The category of categories with families is equivalent to the category of 
-D-systems.\hfill$\square$
-\end{thm}
-
-\begin{comment}
-Since this theorem basically asserts that a category with families is a
-comprehension category
-
-\begin{proof}[Outline of the proof]
-Let $\mathbf{CWF}$ be a category with families. The fact that $\mathbf{CWF}$ is
-a D-system can be shown by proving each of the assertions in the following six 
-steps.
-\begin{enumerate}
-\item For every $f:\Delta\to\Gamma$ and $A\in\mathrm{Ty}(\Gamma)$, there is an
-isomorphism
-\begin{equation*}
-\zeta_{A,f}:\{\theta:\Delta\to\ctxext{\Gamma}{A}\mid p_A\circ\theta\jdeq f\}
-  \cong
-\{\eta:\Delta\to\ctxext{\Delta}{f_\ast(A)}\mid p_{f_\ast(A)}\circ\eta\jdeq\catid{\Delta}\}
-\end{equation*}
-natural in $f$ in the sense that the equalities
-\begin{align*}
-\zeta_{A,\catid{\Gamma}}(\theta) & \jdeq \theta \\
-\zeta_{A,f\circ g}(\theta\circ g) & \jdeq \zeta_{f_\ast(A),g}(\zeta_{A,f}(\theta)\circ g)
-\end{align*}
-hold for any $f:\Delta\to\Gamma$ and $g:X\to\Delta$, and any $\theta:\Delta\to\ctxext{\Gamma}{A}$
-such that $p_A\circ\theta\jdeq f$.
-Furthermore, we get $\zeta_{A,p_A}(\catid{\ctxext{\Gamma}{A}})\jdeq\varphi_{(p_A)_\ast(A)}(\delta_A)$.
-
-\medskip
-(Proof omitted).
-\begin{comment}
-By \autoref{lem:cwf_to_dsys} we may define
-\begin{equation*}
-\zeta_{A,f}(\theta)\defeq \varphi_{f_\ast(A)}(\theta_\ast(\delta_A))
-\end{equation*}
-for any $\theta\in\{\theta:\Delta\to\ctxext{\Gamma}{A}\mid p_A\circ\theta\jdeq f\}$. 
-Since $\zeta_{A,f}$ is a composition of isomorphism, it is itself an isomorphism.
-Also, we get immediately that $\zeta_{A,\catid{\Gamma}}(\theta)\jdeq\theta$. To
-show the remaining identity:
-\begin{align*}
-\zeta_{A,f\circ g}(\theta\circ g)
-& \jdeq \varphi_{g_\ast(f_\ast(A))}((\theta\circ g)_\ast(\delta_A)) \\
-& \jdeq \varphi_{g_\ast(f_\ast(A))}(g_\ast(\theta_\ast(\delta_A))) \\
-& \jdeq \varphi_{g_\ast(f_\ast(A))}(g_\ast(\varphi_{f_\ast(A)}(\theta_\ast(\delta_A))_\ast(\delta_A))) \\
-& \jdeq \varphi_{g_\ast(f_\ast(A))}(g_\ast(\zeta_{A,f}(\theta)_\ast(\delta_A))) \\
-& \jdeq \varphi_{g_\ast(f_\ast(A))}((\zeta_{A,f}(\theta)\circ g)_\ast(\delta_A)) \\
-& \jdeq \zeta_{f_\ast(A),g}(\zeta_{A,f}(\theta)\circ g).
-\end{align*}
-This completes the proof that the isomorphisms $\zeta_{A,f}$ satisfy the required
-equalities.
-\end{comment}
-
-\begin{comment}
-\item
-For any $A\in\mathrm{Ty}(\Gamma)$, we define
-$W_A\defeq (p_A)_\ast:\cat{F}/\Gamma\to\cat{F}/\ctxext{\Gamma}{A}$. Then we get
-for any $A,B\in\cat{F}/\Gamma$ an isomorphism
-\begin{equation*}
-\mathrm{hom}_{\cat{C}/\Gamma}(A,B)
-  \cong
-T(W_A(B)).
-\end{equation*}
-In particular, we get a term $\idtm{A}\in T((p_A)_\ast(A))$. 
-\medskip
-
-Let $A,B\in\cat{F}/\Gamma$. Then we have the isomorphism
-\begin{align*}
-T(W_A(B)) & \jdeq \{\eta:\ctxext{\Gamma}{A}\to\ctxext{{\Gamma}{A}}{W_A(B)}\mid W_A(B)\circ\eta\jdeq\catid{\ctxext{\Gamma}{A}}\} \\
-  & \cong
-\{\theta:\ctxext{\Gamma}{A}\to\ctxext{\Gamma}{B}\mid B\circ\theta\jdeq A\},
-\end{align*}
-so we find $\idtm{A}\defeq\zeta_{A,A}(\catid{\ctxext{\Gamma}{A}})\in T(W_A(A))$.
-
-\item For any $A\in\mathrm{Ty}(\Gamma)$ and $f:\Delta\to\Gamma$ 
-there is a morphism $\pi_2(f,A):\ctxext{\Delta}{f_\ast(A)}\to\ctxext{\Gamma}{A}$
-such that $p_A\circ \pi_2(f,A)\jdeq f\circ p_{f_\ast(A)}$.
-
-\medskip
-(Proof omitted).
-\begin{comment}
-Note that we have the isomorphism
-\begin{align*}
-& \{ \theta : \ctxext{\Delta}{f_\ast(A)}\to\ctxext{\Gamma}{A}\mid p_A\circ\theta\jdeq f\circ p_{f_\ast(A)} \} \\
-  & \cong
-\{ \eta : \ctxext{\Delta}{f_\ast(A)}\to\ctxext{{\Delta}{f_\ast(A)}}{(f\circ p_{f_\ast(A)})_\ast(A)}
-  \mid (f\circ p_{f_\ast(A)})_\ast(A)\circ\eta\jdeq\catid{\ctxext{\Delta}{f_\ast(A)}}\} \\
-  & \jdeq
-\{ \eta : \ctxext{\Delta}{f_\ast(A)}\to\ctxext{{\Delta}{f_\ast(A)}}{(p_{f_\ast(A)})_\ast(f_\ast(A))}
-  \mid (f\circ p_{f_\ast(A)})_\ast(A)\circ\eta\jdeq\catid{\ctxext{\Delta}{f_\ast(A)}}\} \\
-  & \jdeq
-T((p_{f_\ast(A)})_\ast(f_\ast(A))).
-\end{align*}
-Since we have $\delta_{f_\ast(A)}\in T((p_{f_\ast(A)})_\ast(f_\ast(A)))$, we obtain 
-\begin{equation*}
-\pi_2(f,A)\defeq \zeta_{A,f\circ p_{f_\ast(A)}}^{-1}(\delta_{f_\ast(A)}):\ctxext{\Delta}{f_\ast(A)}\to\ctxext{\Gamma}{A}
-\end{equation*}
-satisfying $p_A\circ\pi_2(f,A)\jdeq f\circ p_{f_\ast(A)}$.
-\end{comment}
-\begin{comment}
-\item For any $f:\Delta\to\Gamma$, $g:X\to\Delta$ and $A\in\mathrm{Ty}(\Gamma)$, one has
-\begin{equation*}
-\pi_2(f,A)\circ\pi_2(g,f_\ast(A))\jdeq\pi_2(f\circ g,A).
-\end{equation*}
-
-\medskip
-(Proof omitted).
-\begin{comment}
-It suffices to show that
-\begin{equation*}
-\zeta_{A,f\circ g\circ p_{g_\ast(f_\ast(A))}}(\pi_2(f,A)\circ\pi_2(g,f_\ast(A)))\jdeq\delta_{g_\ast(f_\ast(A))}.
-\end{equation*}
-Because $g\circ p_{g_\ast(f_\ast(A))}\jdeq p_{f_\ast(A)}\circ\pi_2(g,f_\ast(A))$, we get
-to use the naturality of $\zeta$:
-\begin{align*}
-& \zeta_{A,f\circ g\circ p_{g_\ast(f_\ast(A))}}(\pi_2(f,A)\circ\pi_2(g,f_\ast(A))) \\
-& \jdeq
-\zeta_{A,f\circ p_{f_\ast(A)}\circ\pi_2(g,f_\ast(A))}(\pi_2(f,A)\circ\pi_2(g,f_\ast(A))) \\
-& \jdeq
-\zeta_{(p_{f_\ast(A)})_\ast(f_\ast(A)),\pi_2(g,f_\ast(A))}(\zeta_{A,f\circ p_{f_\ast(A)}}(\pi_2(f,a))\circ\pi_2(g,f_\ast(A))) \\
-& \jdeq
-\zeta_{(p_{f_\ast(A)})_\ast(f_\ast(A)),\pi_2(g,f_\ast(A))}(\delta_{f_\ast(A)}\circ\pi_2(g,f_\ast(A))) \\
-& \jdeq
-\zeta_{(p_{f_\ast(A)})_\ast(f_\ast(A)),\pi_2(g,f_\ast(A))}(\zeta_{f_\ast(A),p_{f_\ast(A)}}(\catid{\ctxext{\Delta}{f_\ast(A)}})\circ\pi_2(g,f_\ast(A))) \\
-& \jdeq
-\zeta_{f_\ast(A),p_{f_\ast(A)}\circ\pi_2(g,f_\ast(A))}(\catid{\ctxext{\Delta}{f_\ast(A)}}\circ\pi_2(g,f_\ast(A))) \\
-& \jdeq
-\zeta_{f_\ast(A),p_{f_\ast(A)}\circ\pi_2(g,f_\ast(A))}(\pi_2(g,f_\ast(A))) \\
-& \jdeq
-\zeta_{f_\ast(A),g\circ p_{g_\ast(f_\ast(A))}}(\pi_2(g,f_\ast(A)))\\
-& \jdeq
-\delta_{g_\ast(f_\ast(A))}.
-\end{align*}
-\end{comment}\begin{comment}
-\item
-Let $f:\Delta\to\Gamma$ and $A\in\mathrm{Ty}(\Gamma)$. Then one has 
-\begin{equation*}
-\pi_2(f,A)\circ\zeta_{A,f}(\theta)\jdeq\theta
-\end{equation*}
-for any $\theta:\Delta\to\ctxext{\Gamma}{A}$ satisfying $A\circ\theta\jdeq f$.
-
-\medskip
-(Proof omitted).
-\begin{comment}
-We will write $\tilde\theta\defeq\zeta_{A,f}(\theta)$. Note that
-\begin{equation*}
-p_A\circ\pi_2(f,A)\circ\tilde\theta\jdeq f\circ p_{f_\ast(A)}\circ\tilde\theta\jdeq f.
-\end{equation*} 
-Therefore, we can prove that $\pi_2(f,A)\circ\tilde\theta\jdeq\theta$ by showing
-that $\zeta_{A,f}(\pi_2(f,A)\circ\tilde\theta)\jdeq\tilde\theta$. This is
-shown using the naturality of $\zeta$:
-\begin{align*}
-\zeta_{A,f}(\pi_2(f,A)\circ\tilde\theta)
-  & \jdeq
-\zeta_{A,f\circ p_{f_\ast(A)}\circ\tilde\theta}(\pi_2(f,A)\circ\tilde\theta)
-  \\
-  & \jdeq
-\zeta_{(p_{f_\ast(A)})_\ast(f_\ast(A)),\tilde{\theta}}(\zeta_{A,f\circ p_{f_\ast(A)}}(\pi_2(f,A))\circ\tilde\theta)
-  \\
-  & \jdeq
-\zeta_{(p_{f_\ast(A)})_\ast(f_\ast(A)),\tilde{\theta}}(\idtm{f_\ast(A)}\circ\tilde\theta)
-  \\
-  & \jdeq
-\zeta_{(p_{f_\ast(A)})_\ast(f_\ast(A)),\tilde{\theta}}(\zeta_{f_\ast(A),p_{f_\ast(A)}}(\catid{\ctxext{\Delta}{f_\ast(A)}})\circ\tilde\theta)
-  \\
-  & \jdeq
-\zeta_{f_\ast(A),p_{f_\ast(A)}\circ\tilde\theta}(\catid{\ctxext{\Delta}{f_\ast(A)}}\circ\tilde\theta) \\
-  & \jdeq
-\zeta_{f_\ast(A),\catid{\Delta}}(\tilde\theta) \\
-  & \jdeq 
-\theta.
-\end{align*}
-\end{comment}\begin{comment}
-\item 
-The diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=large]
-\ctxext{\Delta}{f_\ast(A)} \arrow[d,fib] \arrow[r,"{\pi_2(f,A)}"] & \ctxext{\Gamma}{A} \arrow[d,fib] \\
-\Delta \arrow[r,"f"] & \Gamma
-\end{tikzcd}
-\end{equation*}
-is a pullback square.
-
-Proof of (vi): To verify the universal mapping property of the pullback square, consider a
-commuting diagram
-\begin{equation*}
-\begin{tikzcd}
-X \arrow[d,swap,"g"] \arrow[r,"h"] & \ctxext{\Gamma}{A} \arrow[d,fib] \\
-\Delta \arrow[r,"f"] & \Gamma
-\end{tikzcd}
-\end{equation*}
-Then $h$ corresponds uniquely to a section $k$ of $(f\circ g)_\ast(A)$. Since
-$(f\circ g)_\ast(A)\jdeq g_\ast(f_\ast(A))$, the section $k$ fits in the
-commuting diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=huge]
-\ctxext{X}{g_\ast(f_\ast(A))} \arrow[d,xshift=.7ex,fib] \arrow[r,"{\pi_2(g,f_\ast(A))}"]
-& \ctxext{\Delta}{f_\ast(A)} \arrow[d,fib] \arrow[r,"{\pi_2(f,A)}"]
-& \ctxext{\Gamma}{A} \arrow[d,fib] \\
-X \arrow[r,swap,"g"] \arrow[u,xshift=-.7ex,"k"] & \Delta \arrow[r,swap,"f"] & \Gamma
-\end{tikzcd}
-\end{equation*}
-so we obtain $u\defeq\pi_2(g,f_\ast(A))\circ k:X\to\ctxext{\Delta}{f_\ast(A)}$. 
-Then we have $p_{f_\ast(A)}\circ u\jdeq g\circ p_{g_\ast(f_\ast(A))}\circ k\jdeq g$ and
-$\pi_2(f,A)\circ u\jdeq \pi_2(f\circ g,A)\circ k\jdeq h$. Note that
-$u$ corresponds uniquely to $k$, which corresponds uniquely to $h$, so the
-uniqueness of $u$ is automatic.
-\end{enumerate}
-This completes the proof outline that $\mathbf{CWF}$ is a D-system.
-
-Now let $\mathbb{D}$ be a D-system. We define the presheaf $\mathrm{Tm}:
-\op{(\int_{\cat{C}}\mathrm{Ty})}\to\mathbf{Set}$ by
-\begin{equation*}
-\mathrm{Tm}(A) \defeq \{x:\Gamma\to\ctxext{\Gamma}{A}\mid p_A\circ x\jdeq \catid{\Gamma}\}.
-\end{equation*}
-For $f:\Delta\to\Gamma$ and $A\in\mathrm{Ty}(\Gamma)$, we get a function
-$x\mapsto f_\ast(x):\mathrm{Tm}(A)\to\mathrm{Tm}(f_\ast(A))$ by defining $f_\ast(x)$
-to be the unique morphism such that the diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=huge] 
-\Delta \arrow[ddr,bend right=15,equals] \arrow[drr,bend left=15,"x\circ f"] \arrow[dr,densely dotted,near end,"{f_\ast(x)}"] \\
-& \ctxext{\Delta}{f_\ast(A)} \arrow[d,fib] \arrow[r,swap,"{\pi_2(f,A)}"] & \ctxext{\Gamma}{A} \arrow[d,fib] \\
-& \Delta \arrow[r,swap,"f"] & \Gamma
-\end{tikzcd}
-\end{equation*}
-commutes. A combination of the assumed uniqueness and the assumed equalities on
-$\pi_2(\blank,\blank)$ gives the functoriality of $\mathrm{Tm}$.
-
-The term $\delta_A\in\mathrm{Tm}((p_A)_\ast(A))$ is defined to be the unique morphism
-such that the diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=large]
-\ctxext{\Gamma}{A} \arrow[drr,bend left=15,equals] \arrow[ddr,bend right=15,equals] \arrow[dr,densely dotted,near end,"{\delta_{A}}"] \\
-& \ctxext{{\Gamma}{A}}{(p_A)_\ast(A)} \arrow[r,swap,"{\pi_2(A,A)}"] \arrow[d,fib] & \ctxext{\Gamma}{A} \arrow[d,fib] \\
-& \ctxext{\Gamma}{A} \arrow[r,swap,"A"] & \Gamma
-\end{tikzcd}
-\end{equation*}
-commutes.
-
-The universality assumption of categories with families is a consequence of
-the universal property of pullbacks: the map $x\mapsto\pi_2(f,X)\circ x$
-from $\{x:\Delta\to\ctxext{\Delta}{f_\ast(A)}\mid p_{f_\ast(A)}\circ x\jdeq\catid{\Delta}\}$
-to $\{\theta:\Delta\to\ctxext{\Gamma}{A}\mid p_A\circ\theta\jdeq f\}$ is a bijection.
-\end{proof}
-\end{comment}
-
-\begin{eg}
-Since any morphism of $\int_{\cat{C}}$ is cartesian, it follows that D-systems
-are comprehension categories. As D-systems are built up from a presheaf of types
-on the underlying category, they are in fact discrete comprehension categories.
-\end{eg}
-
-The following two definitions define subclasses of closed D-systems and 
-coherent closed D-systems, and the latter structure is the same as an E'-system
-(without empty context). In the next section we will show how to close a comprehension
-category, so that it becomes a closed D-system which satisfies an additional coherence condition.
-
-\begin{defn}
-Let $\cat{C}$ be a D-system. In this definition, we will write
-$\mathrm{Fam}$ for the presheaf of types. The D-system $\cat{C}$
-is said to be \emph{closed}, if we have elements $\emptyf[\Gamma]\in
-\mathrm{Fam}(\Gamma)$ for each $\Gamma\in\cat{C}$, and maps 
-$P\mapsto\ctxext{A}{P}:\mathrm{Fam}(\ctxext{\Gamma}{A})\to \mathrm{Fam}(\Gamma)$
-such that
-\begin{enumerate}
-\item 
-\begin{enumerate}
-\item for any $A\in\mathrm{Fam}(\Gamma)$, $P\in\mathrm{Fam}(\ctxext{\Gamma}{A})$
-and $Q\in\mathrm{Fam}(\ctxext{{\Gamma}{A}}{P})$, one has
-\begin{equation*}
-\ctxext{{A}{P}}{Q}\jdeq\ctxext{A}{{P}{Q}},
-\end{equation*}
-\item and for any $A\in\mathrm{Fam}(\Gamma)$, one has
-\begin{equation*}
-\ctxext{\emptyf[\Gamma]}{A}\jdeq\ctxext{A}{\emptyf[\ctxext{\Gamma}{A}]}\jdeq A,
-\end{equation*}
-\end{enumerate}
-\item 
-\begin{enumerate}
-\item for any $A\in\mathrm{Fam}(\Gamma)$ and $P\in\mathrm{Fam}(\ctxext{\Gamma}{A})$,
-one has
-\begin{equation*}
-p_{\ctxext{A}{P}}\jdeq p_A\circ p_P,
-\end{equation*}
-\item and for any $\Gamma\in\cat{C}$ one has
-\begin{equation*}
-p_{\emptyf[\Gamma]}\jdeq\catid{\Gamma}
-\end{equation*}
-\end{enumerate}
-\item 
-\begin{enumerate}
-\item for any $A\in\mathrm{Fam}(\Gamma)$, $P\in\mathrm{Fam}(\ctxext{\Gamma}{A})$
-and any $f:\Delta\to\Gamma$ in $\cat{C}$, one has
-\begin{equation*}
-f_\ast(\ctxext{A}{P})\jdeq f_\ast(A)\circ \pi_2(f,A)_\ast(P),
-\end{equation*}
-\item and for any $f:\Delta\to\Gamma$ in $\cat{C}$, one has
-\begin{equation*}
-f_\ast(\emptyf[\Gamma])\jdeq\emptyf[\Delta].
-\end{equation*}
-\end{enumerate}
-\end{enumerate} 
-\end{defn}
-
-\begin{lem}
-A D-system is closed if and only if  the following two conditions hold:
-\begin{enumerate}
-\item There is a section $\phi:\cat{C}\to \int_{\cat{C}}\mathrm{Ty}$ of $\pi$.
-\item For every $P\in\mathrm{Ty}(\ctxext{\Gamma}{A})$,
-the morphism $p_A:\ctxext{\Gamma}{A}\to\Gamma$ has a unique cocartesian lift
-at $P$.
-\end{enumerate}
-\end{lem}
-
-\begin{proof}
-Let $\mathbb{D}$ be a closed D-system. To show that $\mathbb{D}$ has the two
-stated properties:
-\begin{enumerate}
-\item
-\item We have to show that each for $P\in\mathrm{Ty}(\ctxext{\Gamma}{A})$,
-the projection $p_A:\ctxext{\Gamma}{A}\to\Gamma$ has a unique cocartesian lift
-at $P$. Note that the morphism $(\ctxext{\Gamma}{A},P)\to (\Gamma,\ctxext{A}{P})$
-in $\int_{\cat{C}}\mathrm{Ty}$ 
-\end{enumerate}
-
-Let $\mathbb{D}$ be a D-system for which the stated two conditions hold. To show
-that $\mathbb{D}$ is a closed D-system, we systematically go through each of the
-properties:
-\begin{enumerate}
-\item 
-\end{enumerate} 
-\end{proof}
-
-\begin{defn}
-A closed D-system is said to be \emph{coherent}, if for any $A,A'\in\mathrm{Fam}(\Gamma)$
-such that $p_A\jdeq p_{A'}$, and any $f:\Delta\to\Gamma$, we have
-$p_{f_\ast(A)}\jdeq p_{f_\ast(A')}$ and $\pi_2(f,A)\jdeq \pi_2(f,A')$.
-\end{defn}
-
-\begin{thm}
-An E'-system without empty context is the same thing as a coherent closed D-system.
-\end{thm}
-
-\begin{proof}
-Let $\mathbb{E}$ be an E'-system without empty context. Then we define the
-presheaf $\mathrm{Fam}$ on $\cat{C}$ on objects by
-\begin{equation*}
-\mathrm{Fam}(\Gamma)\defeq \mathrm{Obj}(\cat{F}/\Gamma)
-\end{equation*}
-and on morphisms by the action on objects of $f_\ast$, for each $f:\Delta\to
-\Gamma$ in $\cat{C}$. By condition (b) and (c) in the definition of E'-systems, it
-follows that this defines a contravariant functor.
-
-We define a functor $\mathcal{P}:\int_{\cat{C}}\to\cat{C}^\square$ by sending
-the pair $(\Gamma,A)$ to the morphism $A:\ctxext{\Gamma}{A}\to\Gamma$ in
-$\cat{C}$, and by sending $f:(\Delta,f_\ast(A))\to (\Gamma,A)$ to the pullback
-square
-\begin{equation*}
-\begin{tikzcd}[column sep=large]
-\ctxext{\Delta}{f_\ast(A)} \arrow[d,swap,"{f_\ast(A)}"] \arrow[r,"{\pi_2(f,A)}"] &
-\ctxext{\Gamma}{A} \arrow[d,"A"] \\ \Delta \arrow[r,"f"] & \Gamma
-\end{tikzcd}
-\end{equation*}
-This is also functorial because of conditions (b) and (c) in the definition of
-E'-systems. Thus, we have defined a D-system
-\begin{equation*}
-\begin{tikzcd}
-\int_{\cat{C}}\mathrm{Fam} \arrow[rr,"{\mathcal{P}}"] \arrow[dr,"{\pi}"] & & \cat{C}^\square \arrow[dl,"{\mathrm{cod}}"] \\ & \cat{C}
-\end{tikzcd}
-\end{equation*}
-and it is straightforward to verify that this D-system is closed and coherent.
-
-Conversely, suppose $\mathbb{D}$ is a closed D-system with underlying category
-$\cat{C}$. Define $\cat{F}$ to be the category with the same objects as
-$\cat{C}$, and morphisms of the form $p_A$. To check that this forms a category,
-note that $p_{\emptyf[\Gamma]}\jdeq\catid{\Gamma}$ and $p_A\circ p_P\jdeq
-p_{\ctxext{A}{P}}$.
-
-Now let $f:\Delta\to\Gamma$ be a morphism in $\cat{C}$. For any $A\in\cat{F}/\Gamma$,
-we have an element $\tilde{A}\in\mathrm{Fam}(\Gamma)$ such that $p_{\tilde{A}}
-\jdeq A$, and we can form a pullback square
-\begin{equation*}
-\begin{tikzcd}
-\ctxext{\Delta}{f_\ast(\tilde{A})} \arrow[r,"\tilde{A}"] \arrow[d,fib,swap,"p_{f_\ast(\tilde{A})}"]
-& \ctxext{\Gamma}{\tilde{A}} \arrow[d,fib,"{p_{\tilde{A}}}"] \\
-\Delta \arrow[r,"f"] & \Gamma
-\end{tikzcd}
-\end{equation*}
-By the coherence condition, this choice of pullback square does not depend on the
-choice of $\tilde{A}$ such that $p_{\tilde{A}}\jdeq A$ (having this fact 
-available is precisely what the coherence condition is for). The verification of conditions
-(a) to (d) of \autoref{defn:E'sys} is straightforward.
-\end{proof}
-
-\subsection{From comprehension categories to E-systems}
 To get from comprehension categories to E'-systems without empty contexts, we
-need to turn comprehension categories into closed D-systems.
+need to turn comprehension categories into telescopic D-systems.
 
@@
 $\cat{F}$ to be the category with the same objects as $\cat{C}$, of which the
-morphisms are generated by the morphisms of the form $P(A)$, for objects
+morphisms are generated by the morphisms of the form $\mathcal{P}(A)$, for objects
 $A\in\cat{E}$. 
@@
 
-\begin{defn}
+\begin{defn}\label{defn:comp_to_telescopic}
 Let $\mathcal{P}$ be a split comprehension category with choice $j$ of cartesian
-lifts. We define a closed D-system
+lifts. We define a telescopic D-system
 \begin{equation*}
@@
 \begin{enumerate}
-\item The choice functors $\Lambda(f)$ and lifts $q(f)$ can be extended to the 
+\item The choice of functors $\Lambda(f)$ and lifts $q(f,A)$ can be extended to the 
 structure of an E'-system without empty context.
@@
 
-\begin{defn}
-There is a functor $\mathcal{V}:\mathbf{Comp}\to\mathbf{E'sys_{\circ}}$, from
-the category $\mathbf{Comp}$ of split full comprehension categories to
-E'-systems without empty context.
-\end{defn}
-
-\begin{constr}
-Let $\mathcal{P}$ be a comprehension category with underlying category
-$\cat{C}$. We define $\cat{F}$ to be the sub-pre-category of $\cat{C}$ generated
-by the morphisms $\mathcal{P}(A):\ctxext{\Gamma}{A}\to\Gamma$ for $A\in\cat{E}$.
-
-For any $f:\Delta\to\Gamma$, we define a functor $f_\ast:\cat{F}/\Gamma\to
-\cat{F}/\Delta$ by induction on the length of the morphism
-Let $f:\Delta\to\Gamma$ and $p(A)\jdeq\Gamma$. Then we get a pullback square
-\begin{equation*}
-\begin{tikzcd}[column sep=large]
-\ctxext{\Delta}{f_\ast(A)} \arrow[d,fib] \arrow[r,"{\pi_2(f,A)}"] & \ctxext{\Gamma}{A} \arrow[d,fib] \\
-\Delta \arrow[r,"f"] & \Gamma
-\end{tikzcd}
-\end{equation*}
-There seems to be no way to show that
-$\pi_2(f,\ctxext{A}{P})\jdeq \pi_2(\pi_2(f,A),P)$.
-\end{constr}
-
 \begin{conj}
-An E'-system is the same thing as a split full comprehension category in which
-$\cat{E}$ has a terminal object.
+The functor from coherent split comprehension categories to E'-systems without
+empty context is full and faithful.
 \end{conj}
@@
 
-\subsection{From categories with families and comprehension categories to E-systems}
-
-\begin{defn}
-Let $\mathbb{D}$ be a D-system with underlying category $\cat{C}$. We define
-$\cat{F}\subseteq\cat{C}$ to be the subcategory with the same objects as
-$\cat{C}$, where the morphisms are generated by the projections $p_A$. Note that
-there is no reason why $\cat{F}$ should possess a terminal object.
-\end{defn}
-
-\begin{defn}
-Let $\mathbb{D}$ be a D-system. Then we define for any $\Gamma\in\mathbb{D}$ 
-the category $\Lambda(\Gamma)$ of which the objects are finite sequences
-of the form $(A_1,\ldots,A_n)$ where
-\begin{equation*}
-A_1\in\mathrm{Ty}(\Gamma),\ A_2\in\mathrm{Ty}(\ctxext{\Gamma}{A_1}),\ \ldots,\ 
-A_n\in\mathrm{Ty}(\ctxext{{{\Gamma}{A_1}}{\cdots}}{A_{n-1}})
-\end{equation*}
-where the morphisms are generated by $(A_1,\ldots,A_{n+1})\leq (A_1,\ldots,A_{n})$.
-\end{defn}
-
-\begin{defn}
-Let $\mathbb{D}$ be a D-system and let $f:\Delta\to\Gamma$ in $\mathbb{D}$.
-Then there is a functor $\Lambda(f):\Lambda(\Gamma)\to\Lambda(\Delta)$, with
-morphisms $\pi_2(f,A):\ctxext{\Delta}{p_{\Lambda(f,A)}}\to
-\ctxext{\Gamma}{p_A}$ in $\mathbb{D}$, such that the square
-\begin{equation*}
-\begin{tikzcd}[column sep=large]
-\ctxext{\Delta}{\Lambda(f,A)} \arrow[r,"{\pi_2(f,A)}"] \arrow[d,fib] & \ctxext{\Gamma}{A} \arrow[d,fib] \\
-\Delta \arrow[r,swap,"f"] & \Gamma
-\end{tikzcd}
-\end{equation*}
-is a pullback square.
-\end{defn}
-
-\begin{constr}
-We construct the asserted structure by
-induction on the length of the arrow $A$ into $\Gamma$. We define
-$\Lambda(f)$ on the empty sequence as
-$\Lambda(f,\emptyf[\Gamma])\defeq \emptyf[\Delta]$, and we define
-$\pi_2(f,\emptyf[\Gamma])\defeq f:\ctxext{\Delta}{\catid{\Delta}}\to
-\ctxext{\Gamma}{\catid{\Gamma}}$. The resulting square is trivially
-a pullback square. Now suppose we have a pullback square as above.
-and let $B\in\mathrm{Ty}(\ctxext{\Gamma}{A})$. Then we define
-\begin{equation*}
-\Lambda(f,[A,B])\defeq [\Lambda(f,A),\pi_2(f,A)_\ast(B)].
-\end{equation*}
-By \autoref{thm:DD}, the top square in the diagram
-\begin{equation*}
-\begin{tikzcd}[column sep=6em]
-\ctxext{{\Delta}{\Lambda(f,A)}}{(\pi_2(f,A)_\ast(B))} \arrow[d,fib] \arrow[r,"{\pi_2(\pi_2(f,A),B)}"] & \ctxext{{\Gamma}{A}}{B} \arrow[d,fib] \\
-\ctxext{\Delta}{\Lambda(f,A)} \arrow[d,fib] \arrow[r,"{\pi_2(f,A)}"] & \ctxext{\Gamma}{A} \arrow[d,fib] \\
-\Delta \arrow[r,"f"] & \Gamma
-\end{tikzcd}
-\end{equation*}
-is a pullback square.
-Since the bottom square is assumed to be a pullback, we get that the outer rectangle
-is a pullback, as desired. 
-\end{constr}
-
-\begin{rmk}
-We can't directly use the functors $\Lambda(f)$ to give a D-system the structure
-of an E'-system, because there is a coherence condition missing.
-\end{rmk}
-
-\begin{thm}\label{thm:Dsys_coh}
-The following are equivalent:
-\begin{enumerate}
-\item The functors $\Lambda(f):\Lambda(\Gamma)\to\Lambda(\Delta)$ extend to 
-the structure of an E'-system without empty context.
-\item For any $f:\Delta\to\Gamma$ and all $A,A'\in\Lambda(\Gamma)$ satisfying 
-$p_A\jdeq p_{A'}$, one has
-\begin{equation*}
-p_{\Lambda(f,A)}\jdeq p_{\Lambda(f,A')}\qquad\text{and}\qquad\pi_2(f,A)\jdeq \pi_2(f,A').
-\end{equation*}
-\end{enumerate}
-\end{thm}
-
-\begin{proof}
-Suppose we can extend $\Lambda(f)$ for all $f:\Delta\to\Gamma$, to form commuting
-diagrams
-\begin{equation*}
-\begin{tikzcd}
-\Lambda(\Gamma) \arrow[d,swap,"p"] \arrow[r,"\Lambda(f)"] & \Lambda(\Delta) \arrow[d,"p"] \\
-\cat{F}/\Gamma \arrow[r,densely dotted,"{f_\ast}"] & \cat{F}/\Delta
-\end{tikzcd}
-\end{equation*}
-of functors, forming an E'-system without empty context. Let 
-$f:\Delta\to\Gamma$, and let $A,A':\Lambda(\Gamma)$ be such that 
-$p_A\jdeq p_{A'}$. Then it follows
-that $\pi_2(f,A)_\ast\jdeq f_\ast/A\jdeq f_\ast/A'\jdeq\pi_2(f,A')_\ast$, so
-we get the desired equality $\pi_2(f,A)\jdeq\pi_2(f,A')$.
-
-For the converse, suppose that for all $f:\Delta\to\Gamma$ and $A,A'\in
-\Lambda(\Gamma)$ such that $p_A\jdeq p_{A'}$, one has 
-$p_{\Lambda(f,A)}\jdeq p_{\Lambda(f,A')}$ and $\pi_2(f,A)\jdeq\pi_2(f,A')$.
-Let $f:\Delta\to\Gamma$. By the assumption that $p_{\Lambda(f,A)}\jdeq p_{\Lambda(f,A')}$,
-it follows that the objects part of $p^\Delta\circ\Lambda(f):\Lambda(\Gamma)\to\cat{F}/\Delta$ 
-factors through $p^\Gamma:\Lambda(\Gamma)\to\cat{F}/\Gamma$. Moreover, by the
-assumption that $\pi_2(f,A)\jdeq \pi_2(f,A')$ it follows that the pullback
-squares
-\begin{equation*}
-\begin{tikzcd}[column sep=large]
-\ctxext{\Delta}{f_\ast(A)} \arrow[d,fib] \arrow[r,"{\pi_2(f,A)}"] & \ctxext{\Gamma}{A} \arrow[d,fib] \\
-\Delta \arrow[r,"f"] & \Gamma
-\end{tikzcd}
-\qquad\text{and}\qquad
-\begin{tikzcd}[column sep=large]
-\ctxext{\Delta}{f_\ast(A')} \arrow[d,fib] \arrow[r,"{\pi_2(f,A')}"] & \ctxext{\Gamma}{A} \arrow[d,fib] \\
-\Delta \arrow[r,"f"] & \Gamma
-\end{tikzcd}
-\end{equation*}
-are the same. Since we have this property
-for any object $\Gamma$, it also follows that the morphisms part of
-$p^\Delta\circ\Lambda(f)$ factors through $p^\Gamma$. This defines the functor
-$f_\ast$ in the commuting square
-\begin{equation*}
-\begin{tikzcd}
-\Lambda(\Gamma) \arrow[d,swap,"p"] \arrow[r,"\Lambda(f)"] & \Lambda(\Delta) \arrow[d,"p"] \\
-\cat{F}/\Gamma \arrow[r,densely dotted,"{f_\ast}"] & \cat{F}/\Delta
-\end{tikzcd}
-\end{equation*}
-We need to verify that $f_\ast$ is a functorial choice of pullbacks, as in the
-definition of \autoref{defn:E'sys}. Since $f_\ast$ is defined as an extension
-of $\Lambda(f)$, it is enough to verify the analoguous functoriality properties
-for $\Lambda(f)$. We do this by induction on the objects of $\Lambda(\Gamma)$,
-in the same order as in \autoref{defn:E'sys}.
-\begin{enumerate}[label=(\alph*)]
-\item This property holds by definition.
-\item Let $\Gamma\in\cat{C}$. We will prove by induction on $A\in\Lambda(\Gamma)$
-that $(\catid{\Gamma})_\ast(A)\jdeq A$ and that $\pi_2(\catid{\Gamma},A)\jdeq \catid{\ctxext{\Gamma}{A}}$.
-Note that these properties hold by definition for $A\jdeq\catid{\Gamma}$. For the
-inductive step, assume that $A\in\Lambda(\Gamma)$ is such that
-$(\catid{\Gamma})_\ast(A)\jdeq A$ and $\pi_2(\catid{\Gamma},A)\jdeq
-\catid{\ctxext{\Gamma}{A}}$ and let $B\in\mathrm{Ty}(\ctxext{\Gamma}{A})$. Then
-we have
-\begin{align*}
-(\catid{\Gamma})_\ast(A\circ p_B) & \jdeq (\catid{\Gamma})_\ast(A)\circ \pi_2(\catid{\Gamma},A)_\ast(p_B) \\
-& \jdeq A\circ p_B
-\end{align*}
-Also, we have
-\begin{align*}
-\pi_2(\catid{\Gamma},A\circ p_B) & \jdeq \pi_2(\pi_2(\catid{\Gamma},A),p_B) \\
-& \jdeq \pi_2(\catid{\ctxext{\Gamma}{A}},p_B) \\
-& \jdeq \catid{\ctxext{{\Gamma}{A}}{B}}
-\end{align*}
-\item Let $f:\Delta\to \Gamma$ and $g:X\to \Delta$. We will prove by induction
-on $A\in\Lambda(\Gamma)$ that $(f\circ g)_\ast(A)\jdeq g_\ast(f_\ast(A))$ and
-that $\pi_2(f\circ g,A)\jdeq \pi_2(f,A)\circ \pi_2(g,f_\ast(A))$. Note that
-these properties hold by definition for $A\jdeq\catid{\Gamma}$. For the inductive
-step, assume that $A\in\Lambda(\Gamma)$ is such that $(f\circ g)_\ast(A)\jdeq g_\ast(f_\ast(A))$ and
-$\pi_2(f\circ g,A)\jdeq \pi_2(f,A)\circ \pi_2(g,f_\ast(A))$, and let
-$B\in\mathrm{Ty}(\ctxext{\Gamma}{A})$. Then we have
-\begin{align*}
-(f\circ g)_\ast(A\circ p_B) & \jdeq \pi_2(f\circ g,A)_\ast(p_B) \\
-& \jdeq (\pi_2(f,A)\circ\pi_2(g,f_\ast(A)))_\ast(p_B) \\
-& \jdeq \pi_2(g,f_\ast)_\ast(\pi_2(f,A)_\ast(p_B)) \\
-& \jdeq g_\ast(f_\ast(A\circ p_B))
-\end{align*}
-\item This property holds by definition.\qedhere
-\end{enumerate}
-\end{proof}
-
-\begin{defn}
-If either of the equivalent conditions of \autoref{thm:Dsys_coh} hold, we say 
-that the D-system is \emph{coherent}. Thus, we have a functor
-$U:\mathbf{Dsys}\to\mathbf{E'sys_{\circ}}$ from coherent D-systems to E'-systems
-\end{defn}
-
-\begin{defn}
-We say that a D-system $\cat{D}$ is \emph{locally well-founded} if the obvious functor
+Merge the previous theorem with the following... (work in progress).
+
+\begin{defn}
+We say that a D-system $\cat{D}$ is \define{locally well-founded} if the obvious functor
 $p:\Lambda(\mathbb{D})\to\cat{F}/\Gamma$ is an isomorphism of pre-categories.

### 2015-11-26 e-systems.tex
telescopic CWFs

@@
 \begin{defn}
-A \emph{pre-weakening structure on a pre-category $\cat{F}$} consists of a functor
+A \define{pre-weakening structure on a pre-category $\cat{F}$} consists of a functor
 $W_A:\cat{F}/\Gamma\to \cat{F}/\ctxext{\Gamma}{A}$ for every $A:\ctxext{\Gamma}{A}\to \Gamma$ in $\cat{F}$ such 
@@
 \end{enumerate}
-A \emph{pre-weakening system} is a pre-category together with a pre-weakening structure.
+A \define{pre-weakening system} is a pre-category together with a pre-weakening structure.
 \end{defn}
@@
 \begin{defn}
-A \emph{pre-weakening homomorphism} $F:\cat{F}\to\cat{D}$ between pre-weakening systems
+A \define{pre-weakening homomorphism} $F:\cat{F}\to\cat{D}$ between pre-weakening systems
 is a functor $F:\cat{F}\to\cat{D}$ such that the square
@@
 \begin{defn}
-A \emph{weakening system} is a pre-weakening system for which every $W_f$ is a
-pre-weakening homomorphism. A \emph{weakening homomorphism} is a pre-weakening 
+A \define{weakening system} is a pre-weakening system for which every $W_f$ is a
+pre-weakening homomorphism. A \define{weakening homomorphism} is a pre-weakening 
 homomorphism between weakening systems.
@@
 \begin{defn}
-A \emph{category with a term structure} is a pre-category $\cat{F}$ with a set $T(A)$ for every
-$A\in\cat{F}/\Gamma$ for every $\Gamma\in\cat{F}$. A \emph{functor with term structure} from $\cat{F}$ to $\cat{D}$
+A \define{category with a term structure} is a pre-category $\cat{F}$ with a set $T(A)$ for every
+$A\in\cat{F}/\Gamma$ for every $\Gamma\in\cat{F}$. A \define{functor with term structure} from $\cat{F}$ to $\cat{D}$
 is a functor $F:\cat{F}\to\cat{D}$ with a function $T(A)\to T(F(A))$ for
@@
 \begin{defn}
-A \emph{weakening system with term term structure} is the same thing as a weakening system,
+A \define{weakening system with term term structure} is the same thing as a weakening system,
 except that the underlying pre-category and the functors $W_A$ are required to
@@
 \begin{defn}
-A \emph{pre-projection system} is a pre-weakening system with term structure with an element
+A \define{pre-projection system} is a pre-weakening system with term structure with an element
 $\tfid{A}\in T(W_A(A))$ for every $A\in\cat{F}/\Gamma$ and $\Gamma\in\cat{F}$. 
@@
 \begin{defn}
-A \emph{pre-projection homomorphism} $F:\cat{F}\to\cat{D}$ is a pre-weakening homomorphism for which
+A \define{pre-projection homomorphism} $F:\cat{F}\to\cat{D}$ is a pre-weakening homomorphism for which
 \begin{equation*}
@@
 \begin{defn}
-A \emph{projection system} is a pre-projection system for which every $W_A$ is
-a pre-projection homomorphism. A \emph{projection homomorphism} is a pre-projection homomorphism
+A \define{projection system} is a pre-projection system for which every $W_A$ is
+a pre-projection homomorphism. A \define{projection homomorphism} is a pre-projection homomorphism
 between projection systems.
@@
 \begin{defn}
-A \emph{pre-substitution structure} on a pre-category with term structure $\cat{F}$ consists
+A \define{pre-substitution structure} on a pre-category with term structure $\cat{F}$ consists
 of a functor with term structure $S_x:\cat{F}/\ctxext{\Gamma}{A}\to\cat{F}/\Gamma$ for every $x\in T(A)$ and
@@
 
-A \emph{pre-substitution system} is a pre-category with term structure together with a
+A \define{pre-substitution system} is a pre-category with term structure together with a
 pre-substitution structure. 
@@
 \begin{defn}
-A \emph{pre-substitution homomorphism} $F:\cat{F}\to\cat{D}$ is a functor with term structure for
+A \define{pre-substitution homomorphism} $F:\cat{F}\to\cat{D}$ is a functor with term structure for
 which the diagram
@@
 \begin{defn}
-A \emph{substitution system} is a pre-substitution system for which each
-$S_t$ is a pre-substitution homomorphism. A \emph{substitution homomorphism} is a
+A \define{substitution system} is a pre-substitution system for which each
+$S_t$ is a pre-substitution homomorphism. A \define{substitution homomorphism} is a
 pre-substitution homomorphism between substitution systems.
@@
 \begin{defn}
-A \emph{pre-E-system} is a pre-category $\cat{F}$ with term structure
+A \define{pre-E-system} is a pre-category $\cat{F}$ with term structure
 with a pre-weakening structure $W$, a pre-substitution structure $S$ and a terminal
-object. A \emph{unital pre-E-system}, or simply \emph{pre-E-system}, is a non-unital pre-E-system
+object. A \define{unital pre-E-system}, or simply \define{pre-E-system}, is a non-unital pre-E-system
 which also has the structure of a pre-projection system. 
@@
 \begin{defn}
-A \emph{pre-E-homomorphism} from $\mathbb{E}\to\mathbb{D}$ is a functor
+A \define{pre-E-homomorphism} from $\mathbb{E}\to\mathbb{D}$ is a functor
 $H:\mathbb{C}_\mathbb{E}\to\mathbb{C}_\mathbb{D}$ between the underlying categories
@@
 
-A \emph{unital} pre-E-homomorphism is a non-unital pre-E-homomorphism which is 
+A \define{unital} pre-E-homomorphism is a non-unital pre-E-homomorphism which is 
 also a pre-projection homomorphism. 
@@
 \label{defn:esystems}
-An \emph{E-system} is a unital pre-E-system in which 
+An \define{E-system} is a unital pre-E-system in which 
 \begin{enumerate}
@@
 \end{enumerate}
-An \emph{E-homomorphism} $H:\mathbb{E}\to\mathbb{D}$ is a functor from
+An \define{E-homomorphism} $H:\mathbb{E}\to\mathbb{D}$ is a functor from
 $\cat{F}_{\mathbb{E}}$ to $\cat{F}_{\mathbb{D}}$ which is both a projection

### 2015-11-26 internal_homs.tex
telescopic CWFs

@@
 \begin{defn}\label{defn:E'sys}
-An \emph{E'-system} consists of a pre-category $\cat{C}$ and a subpre-category $\cat{F}$ with
+An \define{E'-system} consists of a pre-category $\cat{C}$ and a subpre-category $\cat{F}$ with
 the same objects as $\cat{C}$, such that
@@
 \end{equation*}
-\item \label{defn:E'sys:c} For any $f:\Delta\to\Gamma$, $g:X\to\Delta$ and $A\in\cat{F}/\Gamma$, one
+\item \label{defn:E'sys:c} For any $f:\Delta\to\Gamma$, $g:\greek{E}\to\Delta$ and $A\in\cat{F}/\Gamma$, one
 has $(f\circ g)_\ast(A)\jdeq g_\ast(f_\ast(A))$ and 
@@
 \end{equation*}
-An element $f\in \thom{A}{B}$ is called an \emph{internal morphism in context
+An element $f\in \thom{A}{B}$ is called an \define{internal morphism in context
 $\Gamma$}. We sometimes write $\jhom{\Gamma}{A}{B}{f}$ to indicate that $f$ is
@@
 Let $A,B\in\cat{F}/\Gamma$.
-For any $f\in\thom{A}{B}$ we define the pre-composition E-homomorphism
+For any $f\in\thom{A}{B}$ we define the \define{pre-composition} E-homomorphism
 \begin{equation*}
@@
 Let $x\in T(A)$ and $u\in T(S_x(P))$ for $A\in\cat{F}/\Gamma$ and $P\in\cat{F}/\ctxext{\Gamma}{A}$. 
-We define the \emph{term extension of $x$ and $u$} to be
+We define the \define{term extension of $x$ and $u$} to be
 \begin{equation*}

### 2016-02-12 compared.tex
asterisk up

@@
-\section{E-systems compared to other systems of type dependency}\label{sec:esys_compared}
+\section{Telescopic type systems compared to Voevodsky's B-systems}\label{sec:esys_compared}
 In an effort to formalize the simplicial set model of Martin-L\"of type theory
@@
 
-The category of B-systems will be shown to be equivalent to the category of
-stratified E-systems. However, we do not arrive at the fact that the functor
-$\mathcal{E}:\mathbf{Bsys}\to\mathbf{Esys}$ is full without imposing an additional
-condition on the E-homomorphisms. 
-To ensure that the category of stratified E-systems with stratified 
-E-homomorphisms between them is a full subcategory of the category 
-$\mathbf{Esys}$, we will require that the underlying functor of each E-homomorphism
-lifts factorizations. It is
-important to note that this condition is also algebraic, so that the category
-of E-systems remains algebraic.
-
-\subsection{Lifting factorizations}
-Recall
-that for any morphism $f:X\to Y$ in a category $\cat{C}$ there is a category
-$\mathbf{fact}(f)$ of factorizations of $f$. Also, any functor $F:\cat{C}\to\cat{D}$
-determines a functor $F_{\mathbf{fact}(f)}:\mathbf{fact}(f)\to\mathbf{fact}(H(f))$
-for any morphism $f$ of $\cat{C}$.
-
-\begin{defn}
-A functor $F:\cat{C}\to\cat{D}$ is said to
-\define{lift factorizations} if for any $f\in\cat{C}/X$, the functor
-$F_{\mathbf{fact}(f)}:\mathbf{fact}(f)\to\mathbf{fact}(F(f))$ is an isomorphism
-of categories. 
-\end{defn}
-
-\begin{comment}
-\begin{rmk}
-We may choose the property of lifting factorizations
-to involve either an equivalence or an isomorphism of categories. The
-isomorphism-version, would say that for every factorization $h'\circ g'$
-of $F(f)$ there is a unique factorization $f\jdeq h\circ g$ in $\cat{C}$ such
-that $F(g)\jdeq g'$ and $F(h)\jdeq h'$. In the version with equivalences,
-the uniqueness is replaced by uniqueness up to isomorphism, and the equalites
-are replaced by an isomorphism in $\mathbf{fact}(F(f))$.
-
-It is easier to state the isomorphism version of lifting factorizations with
-inference rules and it might be easier to explain this condition on type theoretical
-grounds, although the equivalence version has the advantage of being
-categorical (i.e.~invariant under equivalence of categories).
-
-In the present
-context, it doesn't matter very much which one we pick: the categories in which
-we're interested are all posets.
-\end{rmk}
-\end{comment}
-
-\begin{eg}
-For any slice category $\cat{C}/X$, the forgetful functor $\cat{C}/X\to\cat{C}$
-lifts factorizations.
-\end{eg}
-
-\begin{defn}
-We write $\mathbf{Esys}$ for the category of E-systems where the underlying
-functor of each E-homomorphism lifts factorizations.
-\end{defn}
-
-\subsection{Stratified E-systems}
-
-In this subsection we define the notion of stratified E-system. Then we show
-that the category $\mathbf{Esys_s}$ of stratified E-systems with homomorphisms preserving the
-stratification, is a full subcategory of $\mathbf{Esys}$.
+\subsection{Stratified telescopic type systems}
+
+In this subsection we define the notion of stratified telescopic type system.
+Then we show that the category $\mathbf{Tel}_s$ of stratified telescopic type systems with homomorphisms preserving the stratification, is a full subcategory of $\mathbf{Tel}$.
 
@@
 \begin{defn}
-A pre-E-system is said to be \define{stratified} if its underlying category is
+A pre-telescopic type system is said to be \define{stratified} if its underlying category is
 stratified and if each $W_A$ and $S_x$ is a stratified functor. The category of
-stratified pre-E-systems with stratified E-homomorphisms between them is denoted by
-$\mathbf{Esys_s}$.
+stratified pre-telescopic type systems with stratified E-homomorphisms between them is denoted by
+$\mathbf{Tel}_s$.
 \end{defn}
@@
 \begin{cor}\label{lem:strat_full}
-The category of stratified pre-E-systems is a full subcategory of the category of
-pre-E-systems.
+The category of stratified pre-telescopic type systems is a full subcategory of the category of
+pre-telescopic type systems.
 \end{cor}
 
-\subsection{The embedding of B0-systems into pre-E-systems}
+\subsection{The embedding of B0-systems into pre-telescopic type systems}
 
@@
 between unital B-systems are unital. The same distinguishment could be made for 
-E-systems by forgetting the projection structure, but we have not been able to construct a
-functor from unital B0-systems to unital pre-E-systems. Therefore, we shall only
+telescopic type systems by forgetting the projection structure, but we have not been able to construct a
+functor from unital B0-systems to unital pre-telescopic type systems. Therefore, we shall only
 consider non-unital B0-systems and unital B-systems. In this section we will
 recall the definition of non-unital B0-systems and construct a full and faithful
-functor from them to the category of pre-E-systems.
+functor from them to the category of pre-telescopic type systems.
 
@@
 it is clear in the definition of pre-B-systems that there are only countably
-many operations. Since we are interested in embedding the B-systems into E-systems,
+many operations. Since we are interested in embedding the B-systems into telescopic type systems,
 we shall organize the definition of B-systems a bit differently, but of course
@@
 Now that we have defined the category of B0systems, we can start defining the
-embedding into pre-E-systems.
+embedding into pre-telescopic type systems.
 
@@
 Let $\mathbb{B}$ be a B0-system. Then $U(\mathbb{B})$ can be given the structure
-of a pre-E-system. We write $\mathcal{E}(\mathbb{B})$ for the resulting
-pre-E-system. Thus we get a functor $\mathcal{E}:\mathbf{B0sys}\to\mathbf{Esys_s}$.
+of a pre-telescopic type system. We write $\mathcal{E}(\mathbb{B})$ for the resulting
+pre-telescopic type system. Thus we get a functor $\mathcal{E}:\mathbf{B0sys}\to\mathbf{Tel}_s$.
 \end{defn}
@@
 
-\subsection{B-systems are stratified E-systems}
-There are two properties which hold for pre-E-systems of the form $\mathcal{E}
+\subsection{B-systems are stratified telescopic type systems}
+There are two properties which hold for pre-telescopic type systems of the form $\mathcal{E}
 (\mathbb{B})$ for some B0-system $\mathbb{B}$, 
 which stand in the way of proving that the category of stratified
-pre-E-systems is equivalent to the category of B0-systems. The first is that
+pre-telescopic type systems is equivalent to the category of B0-systems. The first is that
 for any $A\in U(\mathbb{B})/\Gamma$ and $P\in U(\mathbb{B})/\ctxext{\Gamma}{A}$,
@@
 with the categorical structure of the underlying category $U(\mathbb{B})$.
-However, to prove the corresponding facts for E-systems, we needed the full
-set of conditions on unital E-systems: to prove that the term structure of
+However, to prove the corresponding facts for telescopic type systems, we needed the full
+set of conditions on unital telescopic type systems: to prove that the term structure of
 $U(\mathbb{B})$ is compatible with composition we needed units, and to prove 
@@
 
-Presumably, we can assume these compatibility conditions on pre-E-systems in order
-to prove the desired equivalence of B0-systems and stratified pre-E-systems. However,
+Presumably, we can assume these compatibility conditions on pre-telescopic type systems in order
+to prove the desired equivalence of B0-systems and stratified pre-telescopic type systems. However,
 that seems somewhat unnatural, especially given that they automatically hold
-for E-systems. Thus, for the remainder of this section we set out to show that
+for telescopic type systems. Thus, for the remainder of this section we set out to show that
 the category of unital B-systems is equivalent to the category of stratified 
-unital E-systems.
+unital telescopic type systems.
 
@@
 For any non-unital B-system $\mathbb{B}$, $\mathcal{E}(\mathbb{B})$ is a 
-stratified non-unital E-system. 
+stratified non-unital telescopic type system. 
 \end{lem}
@@
 projection structure. We thus obtain a functor 
-$\mathcal{E}:\mathbf{Bsys}\to\mathbf{Esys}$ from unital B-systems to E-systems.
+$\mathcal{E}:\mathbf{Bsys}\to\mathbf{Tel}$ from unital B-systems to telescopic type systems.
 \end{defn}
@@
 We first need to extend the functor $\mathcal{E}:\mathbf{B0sys}\to\mathbf{Esys_p}$
-from non-unital B0-systems to non-unital pre-E-systems, to a functor from
-unital B-systems to E-systems.
+from non-unital B0-systems to non-unital pre-telescopic type systems, to a functor from
+unital B-systems to telescopic type systems.
 
 We need to give $\mathcal{E}(\mathbb{B})$ the structure of a unital
-pre-E-system. We construct by induction on $m\in \mathbb{N}$, a term of type 
+pre-telescopic type system. We construct by induction on $m\in \mathbb{N}$, a term of type 
 $T(m,W_{(m,X)}(X))$ with the property that that 
@@
 \begin{thm}
-The functor $\mathcal{E}:\mathbf{Bsys}\to\mathbf{Esys}$ lifts to an equivalence
-to the subcategory $\mathbf{Esys_s}$ of stratified E-systems: i.e.~we have a
+The functor $\mathcal{E}:\mathbf{Bsys}\to\mathbf{Tel}$ lifts to an equivalence
+to the subcategory $\mathbf{Tel}_s$ of stratified telescopic type systems: i.e.~we have a
 commuting diagram
@@
 \begin{tikzcd}
-& \mathbf{Esys_s} \arrow[dr,"\subseteq"]\\
-\mathbf{Bsys} \arrow[ur,"\simeq"] \arrow[rr,swap,"\mathcal{E}"] & & \mathbf{Esys}
+& \mathbf{Tel}_s \arrow[dr,"\subseteq"]\\
+\mathbf{Bsys} \arrow[ur,"\simeq"] \arrow[rr,swap,"\mathcal{E}"] & & \mathbf{Tel}
 \end{tikzcd}
@@
 We already have shown that $\mathcal{E}$ factors through the stratified
-E-systems and that it is full and faithful, so it remains
-to show that every stratified E-system is in the image of $\mathcal{E}$. In
-other words, for every stratified E-system $\mathbb{E}$ we need to construct
+telescopic type systems and that it is full and faithful, so it remains
+to show that every stratified telescopic type system is in the image of $\mathcal{E}$. In
+other words, for every stratified telescopic type system $\mathbb{E}$ we need to construct
 a B0-system $\mathbb{B}$ such that $\mathcal{E}(\mathbb{B})\cong\mathbb{E}$.
 
-Let $\mathbb{E}$ be a stratified E-system. From the stratification functor
+Let $\mathbb{E}$ be a stratified telescopi type system. From the stratification functor
 $L:\mathbb{C}_{\mathbb{E}}\to (\mathbb{N},\geq)$ we obtain the pre-image
@@
 This defines the B-framework $B(\mathbb{E})$, and we obtain a functor
-$B:\mathbf{Esys_s}\to\mathbf{B0sys}$. It is immediate that we get isomorphisms
+$B:\mathbf{Tel}_s\to\mathbf{B0sys}$. It is immediate that we get isomorphisms
 $U(B(\mathbb{E}))\cong\cat{F}_{\mathbb{E}}$, natural in $\mathbb{E}$. Also, we
@@
 respectively. Since these functors are assumed to be stratified, they define
-homomorphisms of B-frameworks. Because a weakening system of an E-system is
+homomorphisms of B-frameworks. Because a weakening system of a telescopic type system is
 required to be functorial, i.e.~$W_{\ctxext{A}{P}}\jdeq W_P\circ W_A$ and
@@
 
+\section{Telescopic type systems compared to comprehension categories and to categories with families}
 \subsection{Telescopic categories with families}
@@
 \item A presheaf $\mathrm{Ty}:\op{\cat{C}}\to\mathbf{Set}$. We will write
-$f_\ast(A)$ for $\mathrm{Ty}(f,A)$.
+$f^\ast(A)$ for $\mathrm{Ty}(f,A)$.
 \item For any $A\in\mathrm{Ty}(\Gamma)$, a morphism
@@
 \item a preasheaf $\mathrm{Tm}:\op{(\int_\cat{C}\mathrm{Ty})}\to\mathbf{Set}$.
-The action of a morphism $f:(\Delta,f_\ast(A))\to(\Gamma,A)$ of $\int_\cat{C}\mathrm{Ty}$
-on a term $x\in \mathrm{Tm}(A)$ is also written by $f_\ast(x)$.
-\item a term $\delta_A\in\mathrm{Tm}(\ctxext{\Gamma}{A},(p_A)_\ast(A))$ for every
+The action of a morphism $f:(\Delta,f^\ast(A))\to(\Gamma,A)$ of $\int_\cat{C}\mathrm{Ty}$
+on a term $x\in \mathrm{Tm}(A)$ is also written by $f^\ast(x)$.
+\item a term $\delta_A\in\mathrm{Tm}(\ctxext{\Gamma}{A},(p_A)^\ast(A))$ for every
 $A\in\mathrm{Ty}(\Gamma)$, 
@@
 \begin{equation*}
-\theta\mapsto\theta_\ast(\delta_A)
+\theta\mapsto\theta^\ast(\delta_A)
   : \{\theta:\Delta\to\ctxext{\Gamma}{A}\mid p_A\circ\theta\jdeq f\}\to
-    \mathrm{Tm}(f_\ast(A))
+    \mathrm{Tm}(f^\ast(A))
 \end{equation*}
@@
 at all to have a faithful functor from the category of categories
-with families to the category of E-systems.
+with families to the category of telescopic type systems.
 
@@
 Although a terminal object is usually assumed to exist, not every object 
-of a D-system is assumed to be `fibrant', i.e.~it is
+of a category with families is assumed to be `fibrant', i.e.~it is
 not the case that every $X\to 1$ is a composition of projections $p_A$.
-We could only hope that there is a functor from D-sytems to E-systems without
+We could only hope that there is a functor from D-sytems to telescopic type systems without
 terminal objects.
@@
 The
-natural isomorphism $\varphi$ is such that the term $\delta_A\in\mathrm{Tm}((p_A)_\ast(A))$
+natural isomorphism $\varphi$ is such that the term $\delta_A\in\mathrm{Tm}((p_A)^\ast(A))$
 corresponds to the unique arrow $\tilde{q}_A$ which fits in the diagram
@@
 \ctxext{\Gamma}{A} \arrow[ddr,bend right=15,equals] \arrow[drr,bend left=15,equals] \arrow[dr,"{\tilde \delta_A}"] \\
-& \ctxext{{\Gamma}{A}}{(p_A)_\ast(A)} \arrow[d,fib] & \ctxext{\Gamma}{A} \arrow[d,fib] \\
+& \ctxext{{\Gamma}{A}}{(p_A)^\ast(A)} \arrow[d,fib] & \ctxext{\Gamma}{A} \arrow[d,fib] \\
 & \ctxext{\Gamma}{A} \arrow[r,"{p_A}"] & \Gamma
@@
 \begin{constr}
-We define the inverse of $\varphi_A$ by $\theta\mapsto\theta_\ast(\delta_A)$. Verifying
+We define the inverse of $\varphi_A$ by $\theta\mapsto\theta^\ast(\delta_A)$. Verifying
 that these form a natural isomorphism is straightforward.
@@
 
+The following two definitions define subclasses of telescopic D-systems and 
+coherent telescopic D-systems, and the latter structure is the same as an E'-system
+(without empty context). In \autoref{defn:comp_to_telescopic}, we will show how
+to obtain a telescopic D-system from any comprehension category. This construction
+will also generate a telescopic D-system out of any D-system, because D-systems
+are comprehension categories.
+
+\begin{defn}
+Let $\cat{C}$ be a D-system. In this definition, we will write
+$\mathrm{Fam}$ for the presheaf of types. The D-system $\cat{C}$
+is said to be \define{telescopic}, if we have elements $\emptyf[\Gamma]\in
+\mathrm{Fam}(\Gamma)$ for each $\Gamma\in\cat{C}$, and maps 
+$P\mapsto\ctxext{A}{P}:\mathrm{Fam}(\ctxext{\Gamma}{A})\to \mathrm{Fam}(\Gamma)$
+such that
+\begin{enumerate}
+\item 
+\begin{enumerate}
+\item for any $A\in\mathrm{Fam}(\Gamma)$ and $P\in\mathrm{Fam}(\ctxext{\Gamma}{A})$,
+one has $\ctxext{\Gamma}{{A}{P}}\jdeq\ctxext{{\Gamma}{A}}{P}$, and
+\begin{equation*}
+p_{\ctxext{A}{P}}\jdeq p_A\circ p_P,
+\end{equation*}
+\item and for any $\Gamma\in\cat{C}$ one has $\ctxext{\Gamma}{\emptyf[\Gamma]}\jdeq\Gamma$, and
+\begin{equation*}
+p_{\emptyf[\Gamma]}\jdeq\catid{\Gamma}.
+\end{equation*}
+\end{enumerate}
+\item 
+\begin{enumerate}
+\item for any $A\in\mathrm{Fam}(\Gamma)$, $P\in\mathrm{Fam}(\ctxext{\Gamma}{A})$
+and $Q\in\mathrm{Fam}(\ctxext{{\Gamma}{A}}{P})$, one has
+\begin{equation*}
+\ctxext{{A}{P}}{Q}\jdeq\ctxext{A}{{P}{Q}},
+\end{equation*}
+\item and for any $A\in\mathrm{Fam}(\Gamma)$, one has
+\begin{equation*}
+\ctxext{\emptyf[\Gamma]}{A}\jdeq\ctxext{A}{\emptyf[\ctxext{\Gamma}{A}]}\jdeq A.
+\end{equation*}
+\end{enumerate}
+\item 
+\begin{enumerate}
+\item for any $A\in\mathrm{Fam}(\Gamma)$, $P\in\mathrm{Fam}(\ctxext{\Gamma}{A})$
+and any $f:\Delta\to\Gamma$ in $\cat{C}$, one has
+\begin{equation*}
+f^\ast(\ctxext{A}{P})\jdeq \ctxext{(f^\ast(A))}{(\pi_2(f,A)^\ast(P))},
+\end{equation*}
+\item and for any $f:\Delta\to\Gamma$ in $\cat{C}$, one has
+\begin{equation*}
+f^\ast(\emptyf[\Gamma])\jdeq\emptyf[\Delta].
+\end{equation*}
+\end{enumerate}
+\end{enumerate} 
+\end{defn}
+
+\begin{lem}
+A D-system is telescopic if and only if  the following two conditions hold:
+\begin{enumerate}
+\item There is a global section $\phi:\cat{C}\to \int_{\cat{C}}\mathrm{Ty}$ of $\pi$.
+\item For every $P\in\mathrm{Ty}(\ctxext{\Gamma}{A})$,
+the morphism $p_A:\ctxext{\Gamma}{A}\to\Gamma$ has a unique cocartesian lift
+at $P$.
+\end{enumerate}
+\end{lem}
+
+\begin{proof}
+Let $\mathbb{D}$ be a telescopic D-system. To show that $\mathbb{D}$ has the two
+stated properties:
+\begin{enumerate}
+\item The choice $\emptyf[\Gamma]\in\mathrm{Ty}(\Gamma)$ together with the fact
+that $f^\ast(\emptyf[\Gamma])\jdeq\emptyf[\Delta]$ for any $f:\Delta\to\Gamma$,
+is a global section of $\mathrm{Ty}$. 
+\item We have to show that each for $P\in\mathrm{Ty}(\ctxext{\Gamma}{A})$,
+the projection $p_A:\ctxext{\Gamma}{A}\to\Gamma$ has a unique cocartesian lift
+at $P$. We first define a morphism $g:(\ctxext{\Gamma}{A},P)\to(\Gamma,\ctxext{A}{P})$
+in $\int_{\cat{C}}\mathrm{Ty}$. Thus, we have to find a morphism
+$g:\ctxext{\Gamma}{A}\to\Gamma$ with the property that $g^\ast(\ctxext{A}{P})
+\jdeq P$.
+
+In other words, we have to show that there is a unique 
+$Q\in\mathrm{Ty}(\Gamma)$, and a unique morphism $g:(\ctxext{\Gamma}{A},P)\to
+(\Gamma,Q)$ in $\int_C\mathrm{Ty}$, such that $\pi(g)\jdeq p_A$ and such that
+for every $g':(\ctxext{\Gamma}{A},P)\to(\Gamma,Q')$ with the property that
+$\pi(g')\jdeq p_A$, there is a unique $u:Q\to Q'$ such that $g'=u\circ g$ and
+$\pi(u)\jdeq\idfunc[\Gamma]$.
+
+Note that the morphism $(\ctxext{\Gamma}{A},P)\to (\Gamma,\ctxext{A}{P})$
+in $\int_{\cat{C}}\mathrm{Ty}$.  
+\end{enumerate}
+
+Let $\mathbb{D}$ be a D-system for which the stated two conditions hold. To show
+that $\mathbb{D}$ is a telescopic D-system, we systematically go through each of the
+properties:
+\begin{enumerate}
+\item 
+\end{enumerate} 
+\end{proof}
+
+\begin{conj}
+The notion of being telescopic is a property.
+\end{conj}
+
+\begin{defn}
+A telescopic D-system is said to be \define{coherent}, if for any $A,A'\in\mathrm{Fam}(\Gamma)$
+such that $p_A\jdeq p_{A'}$, and any $f:\Delta\to\Gamma$, we have
+$p_{f^\ast(A)}\jdeq p_{f^\ast(A')}$ and $\pi_2(f,A)\jdeq \pi_2(f,A')$.
+\end{defn}
+
+\begin{thm}
+An E'-system without empty context is the same thing as a coherent telescopic D-system.
+\end{thm}
+
+\begin{proof}
+Let $\mathbb{E}$ be an E'-system without empty context. Then we define the
+presheaf $\mathrm{Fam}$ on $\cat{C}$ on objects by
+\begin{equation*}
+\mathrm{Fam}(\Gamma)\defeq \mathrm{Obj}(\cat{F}/\Gamma)
+\end{equation*}
+and on morphisms by the action on objects of $f^\ast$, for each $f:\Delta\to
+\Gamma$ in $\cat{C}$. By condition (b) and (c) in the definition of E'-systems, it
+follows that this defines a contravariant functor.
+
+We define a functor $\mathcal{P}:\int_{\cat{C}}\to\cat{C}^\square$ by sending
+the pair $(\Gamma,A)$ to the morphism $A:\ctxext{\Gamma}{A}\to\Gamma$ in
+$\cat{C}$, and by sending $f:(\Delta,f^\ast(A))\to (\Gamma,A)$ to the pullback
+square
+\begin{equation*}
+\begin{tikzcd}[column sep=large]
+\ctxext{\Delta}{f^\ast(A)} \arrow[d,swap,"{f^\ast(A)}"] \arrow[r,"{\pi_2(f,A)}"] &
+\ctxext{\Gamma}{A} \arrow[d,"A"] \\ \Delta \arrow[r,"f"] & \Gamma
+\end{tikzcd}
+\end{equation*}
+This is also functorial because of conditions (b) and (c) in the definition of
+E'-systems. Thus, we have defined a D-system
+\begin{equation*}
+\begin{tikzcd}
+\int_{\cat{C}}\mathrm{Fam} \arrow[rr,"{\mathcal{P}}"] \arrow[dr,"{\pi}"] & & \cat{C}^\square \arrow[dl,"{\mathrm{cod}}"] \\ & \cat{C}
+\end{tikzcd}
+\end{equation*}
+and it is straightforward to verify that this D-system is telescopic and coherent.
+
+Conversely, suppose $\mathbb{D}$ is a telescopic D-system with underlying category
+$\cat{C}$. Define $\cat{F}$ to be the category with the same objects as
+$\cat{C}$, and morphisms of the form $p_A$. To check that this forms a category,
+note that $p_{\emptyf[\Gamma]}\jdeq\catid{\Gamma}$ and $p_A\circ p_P\jdeq
+p_{\ctxext{A}{P}}$.
+
+Now let $f:\Delta\to\Gamma$ be a morphism in $\cat{C}$. For any $A\in\cat{F}/\Gamma$,
+we have an element $\tilde{A}\in\mathrm{Fam}(\Gamma)$ such that $p_{\tilde{A}}
+\jdeq A$, and we can form a pullback square
+\begin{equation*}
+\begin{tikzcd}[column sep=large]
+\ctxext{\Delta}{f^\ast(\tilde{A})} \arrow[r,"{\pi_2(f,\tilde{A})}"] \arrow[d,fib,swap,"p_{f^\ast(\tilde{A})}"]
+& \ctxext{\Gamma}{\tilde{A}} \arrow[d,fib,"{p_{\tilde{A}}}"] \\
+\Delta \arrow[r,"f"] & \Gamma
+\end{tikzcd}
+\end{equation*}
+By the coherence condition, this choice of pullback square does not depend on the
+choice of $\tilde{A}$ such that $p_{\tilde{A}}\jdeq A$ (having this fact 
+available is precisely what the coherence condition is for). The verification of conditions
+(a) to (d) of \autoref{defn:E'sys} is straightforward.
+\end{proof}
+
+\subsection{Coherent comprehension categories}
+In this section we recall Jacobs' notion of comprehension categories, and 
+Dybjer's notion of categories with families. It is shown in 
+\cite{Hofmann_syntax_semantics} that the category of categoriew with families
+is equivalent to the category of categories with attributes, which forms a
+subcategory of comprehension categories. Thus, for our purpose of comparing
+telescopic type systems with various existing notions of models of type dependency, it suffice
+to make the comparison with comprehension categories.
+
+In this section we do the preparateions for this comparison. After having
+introduced categories with families, we prove that categories with families
+with the extra structure of `being closed under family extension' in a coherent
+way, are E'-systems without empty context.
+
+\begin{defn}
+Let $p:\cat{E}\to\cat{B}$ be a functor and let $f:\Delta\to \Gamma$ be a morphism in $\cat{B}$.
+A morphism $g:A\to B$ in $\cat{E}$ is said to be \define{cartesian over $f$}, if
+\begin{enumerate}
+\item $p(g)\jdeq f$, and 
+\item For all $g':A'\to B$ with $p(g')\jdeq f$, there is a unique $u:A'\to A$
+such that $g'=g\circ u$ and $p(u)\jdeq\catid{\Delta}$. 
+\end{enumerate}
+Dually, a morphism $g:A\to B$ in $\cat{E}$ is said to be \define{cocartesian over
+$f$}, if
+\begin{enumerate}
+\item $p(g)\jdeq f$, and
+\item For all $g':A\to B'$ with $p(g')\jdeq f$, there is a unique $v:B\to B'$
+such that $g'=v\circ g$ and $p(v)\jdeq\catid{\Gamma}$. 
+\end{enumerate}
+A (co)cartesian morphism over $f$ is also called a \define{(co)cartesian lift of $f$}.
+\end{defn}
+
+\begin{defn}
+A functor $p:\cat{E}\to\cat{B}$ is called a \define{Grothendieck fibration} if
+every morphism into an object of the form $p(A)$ for some object $A$ of $\cat{E}$ 
+has a cartesian lift, and cartesian morphisms are closed under composition.
+\end{defn}
+
+\begin{defn}
+A Grothendieck fibration $p:\cat{E}\to\cat{B}$ is said to be \define{split}, if
+there is a functorial choice of cartesian lifts, i.e.~for every $f:\Delta\to \Gamma$ in 
+$\cat{B}$ and any $B\in \cat{E}$ such that $p(B)\jdeq \Gamma$,
+there is a cartesian lift $j(B,f):A \to B$ for some object $A$ of $\cat{E}$, such that 
+$j(B,\idfunc[\Gamma])\jdeq\catid{B}$ and 
+$j(B, f\circ g)\jdeq j(B,f)\circ j(\mathrm{dom}(j(B,f)),g)$ for each
+$f:\Delta\to\Gamma$ and each $g:\greek{E}\to\Delta$. 
+\end{defn}
+
+\begin{eg}
+For any presheaf $P:\cat{C}\to\mathbb{Set}$, the projection functor
+$\int_\cat{C} P\to\cat{C}$ is a Grothendieck fibration.
+\end{eg}
+
+\begin{defn}
+A \define{comprehension category} $\mathcal{P}$ consists of categories $\cat{E}$, $\cat{C}$ and a
+commuting triangle
+\begin{equation*}
+\begin{tikzcd}
+\cat{E} \arrow[rr,"\mathcal{P}"] \arrow[dr,swap,"p"] & & \cat{C}^{\to} \arrow[dl,"\mathrm{cod}"] \\
+& \cat{C}
+\end{tikzcd}
+\end{equation*}
+of functors, such that $p$ is a Grothendieck fibration, and $\mathcal{P}$ maps
+cartesian morphisms in $\cat{E}$ to pullback squares in $\cat{C}$. We say that
+$\mathcal{P}$ is a \define{full} comprehension category if the functor $\mathcal{P}:
+\cat{E}\to\cat{C}^{\to}$ is full and faithful; and we say that
+$\mathcal{P}$ is \define{split} if $p$ is split.
+\end{defn}
+
 The definition of D-system we give below, is practically the same as the definition
 of category with attributes in \cite[Definition 3.10]{Hofmann_syntax_semantics}.
-Following Voevodsky, we wish to reserve the word `category' for usage in mathematical terms
-for notions which are `categorical' in the sense of being invariant under 
-equivalence of categories. The definition of categories with attributes involves
-an equality of functors (which would have to be replaced by a natural 
-isomorphism), so it is at least not obviously categorical.
+
+\begin{rmk}
+Since any morphism of $\int_{\cat{C}}\mathrm{Ty}$ is cartesian, it follows that D-systems
+are comprehension categories. As D-systems are built up from a presheaf of types
+on the underlying category, they are in fact discrete comprehension categories.
+\end{rmk}
 
@@
 the morphism $p(\Gamma,A)$ will be denoted by $p_A$; for any
-$(f,A):(\Delta,f_\ast(A))\to(\Gamma,A)$ in $\int_{\cat{C}}\mathrm{Ty}$, the top
+$(f,A):(\Delta,f^\ast(A))\to(\Gamma,A)$ in $\int_{\cat{C}}\mathrm{Ty}$, the top
 morphism of the pullback square $p(f,A)$ will be denoted by $\pi_2(f,A)$.
@@
 $\pi_2(\catid{\Gamma},A)\jdeq\catid{\ctxext{\Gamma}{A}}$ and
-$\pi_2(f\circ g,A)\jdeq \pi_2(f,A)\circ\pi_2(g,f_\ast(A))$.
+$\pi_2(f\circ g,A)\jdeq \pi_2(f,A)\circ\pi_2(g,f^\ast(A))$.
 \end{rmk}
@@
   \cong
-\{\eta:\Delta\to\ctxext{\Delta}{f_\ast(A)}\mid p_{f_\ast(A)}\circ\eta\jdeq\catid{\Delta}\}
+\{\eta:\Delta\to\ctxext{\Delta}{f^\ast(A)}\mid p_{f^\ast(A)}\circ\eta\jdeq\catid{\Delta}\}
 \end{equation*}
@@
 \zeta_{A,\catid{\Gamma}}(\theta) & \jdeq \theta \\
-\zeta_{A,f\circ g}(\theta\circ g) & \jdeq \zeta_{f_\ast(A),g}(\zeta_{A,f}(\theta)\circ g)
+\zeta_{A,f\circ g}(\theta\circ g) & \jdeq \zeta_{f^\ast(A),g}(\zeta_{A,f}(\theta)\circ g)
 \end{align*}
@@
 such that $p_A\circ\theta\jdeq f$.
-Furthermore, we get $\zeta_{A,p_A}(\catid{\ctxext{\Gamma}{A}})\jdeq\varphi_{(p_A)_\ast(A)}(\delta_A)$.
+Furthermore, we get $\zeta_{A,p_A}(\catid{\ctxext{\Gamma}{A}})\jdeq\varphi_{(p_A)^\ast(A)}(\delta_A)$.
 
@@
 \begin{equation*}
-\zeta_{A,f}(\theta)\defeq \varphi_{f_\ast(A)}(\theta_\ast(\delta_A))
+\zeta_{A,f}(\theta)\defeq \varphi_{f^\ast(A)}(\theta^\ast(\delta_A))
 \end{equation*}
@@
 \zeta_{A,f\circ g}(\theta\circ g)
-& \jdeq \varphi_{g_\ast(f_\ast(A))}((\theta\circ g)_\ast(\delta_A)) \\
-& \jdeq \varphi_{g_\ast(f_\ast(A))}(g_\ast(\theta_\ast(\delta_A))) \\
-& \jdeq \varphi_{g_\ast(f_\ast(A))}(g_\ast(\varphi_{f_\ast(A)}(\theta_\ast(\delta_A))_\ast(\delta_A))) \\
-& \jdeq \varphi_{g_\ast(f_\ast(A))}(g_\ast(\zeta_{A,f}(\theta)_\ast(\delta_A))) \\
-& \jdeq \varphi_{g_\ast(f_\ast(A))}((\zeta_{A,f}(\theta)\circ g)_\ast(\delta_A)) \\
-& \jdeq \zeta_{f_\ast(A),g}(\zeta_{A,f}(\theta)\circ g).
+& \jdeq \varphi_{g^\ast(f^\ast(A))}((\theta\circ g)^\ast(\delta_A)) \\
+& \jdeq \varphi_{g^\ast(f^\ast(A))}(g^\ast(\theta^\ast(\delta_A))) \\
+& \jdeq \varphi_{g^\ast(f^\ast(A))}(g^\ast(\varphi_{f^\ast(A)}(\theta^\ast(\delta_A))^\ast(\delta_A))) \\
+& \jdeq \varphi_{g^\ast(f^\ast(A))}(g^\ast(\zeta_{A,f}(\theta)^\ast(\delta_A))) \\
+& \jdeq \varphi_{g^\ast(f^\ast(A))}((\zeta_{A,f}(\theta)\circ g)^\ast(\delta_A)) \\
+& \jdeq \zeta_{f^\ast(A),g}(\zeta_{A,f}(\theta)\circ g).
 \end{align*}
@@
 For any $A\in\mathrm{Ty}(\Gamma)$, we define
-$W_A\defeq (p_A)_\ast:\cat{F}/\Gamma\to\cat{F}/\ctxext{\Gamma}{A}$. Then we get
+$W_A\defeq (p_A)^\ast:\cat{F}/\Gamma\to\cat{F}/\ctxext{\Gamma}{A}$. Then we get
 for any $A,B\in\cat{F}/\Gamma$ an isomorphism
@@
 \end{equation*}
-In particular, we get a term $\idtm{A}\in T((p_A)_\ast(A))$. 
+In particular, we get a term $\idtm{A}\in T((p_A)^\ast(A))$. 
 \medskip
@@
 \item For any $A\in\mathrm{Ty}(\Gamma)$ and $f:\Delta\to\Gamma$ 
-there is a morphism $\pi_2(f,A):\ctxext{\Delta}{f_\ast(A)}\to\ctxext{\Gamma}{A}$
-such that $p_A\circ \pi_2(f,A)\jdeq f\circ p_{f_\ast(A)}$.
+there is a morphism $\pi_2(f,A):\ctxext{\Delta}{f^\ast(A)}\to\ctxext{\Gamma}{A}$
+such that $p_A\circ \pi_2(f,A)\jdeq f\circ p_{f^\ast(A)}$.
 
@@
 \begin{align*}
-& \{ \theta : \ctxext{\Delta}{f_\ast(A)}\to\ctxext{\Gamma}{A}\mid p_A\circ\theta\jdeq f\circ p_{f_\ast(A)} \} \\
+& \{ \theta : \ctxext{\Delta}{f^\ast(A)}\to\ctxext{\Gamma}{A}\mid p_A\circ\theta\jdeq f\circ p_{f^\ast(A)} \} \\
   & \cong
-\{ \eta : \ctxext{\Delta}{f_\ast(A)}\to\ctxext{{\Delta}{f_\ast(A)}}{(f\circ p_{f_\ast(A)})_\ast(A)}
-  \mid (f\circ p_{f_\ast(A)})_\ast(A)\circ\eta\jdeq\catid{\ctxext{\Delta}{f_\ast(A)}}\} \\
-  & \jdeq
-\{ \eta : \ctxext{\Delta}{f_\ast(A)}\to\ctxext{{\Delta}{f_\ast(A)}}{(p_{f_\ast(A)})_\ast(f_\ast(A))}
-  \mid (f\circ p_{f_\ast(A)})_\ast(A)\circ\eta\jdeq\catid{\ctxext{\Delta}{f_\ast(A)}}\} \\
-  & \jdeq
-T((p_{f_\ast(A)})_\ast(f_\ast(A))).
-\end{align*}
-Since we have $\delta_{f_\ast(A)}\in T((p_{f_\ast(A)})_\ast(f_\ast(A)))$, we obtain 
-\begin{equation*}
-\pi_2(f,A)\defeq \zeta_{A,f\circ p_{f_\ast(A)}}^{-1}(\delta_{f_\ast(A)}):\ctxext{\Delta}{f_\ast(A)}\to\ctxext{\Gamma}{A}
-\end{equation*}
-satisfying $p_A\circ\pi_2(f,A)\jdeq f\circ p_{f_\ast(A)}$.
+\{ \eta : \ctxext{\Delta}{f^\ast(A)}\to\ctxext{{\Delta}{f^\ast(A)}}{(f\circ p_{f^\ast(A)})^\ast(A)}
+  \mid (f\circ p_{f^\ast(A)})^\ast(A)\circ\eta\jdeq\catid{\ctxext{\Delta}{f^\ast(A)}}\} \\
+  & \jdeq
+\{ \eta : \ctxext{\Delta}{f^\ast(A)}\to\ctxext{{\Delta}{f^\ast(A)}}{(p_{f^\ast(A)})^\ast(f^\ast(A))}
+  \mid (f\circ p_{f^\ast(A)})^\ast(A)\circ\eta\jdeq\catid{\ctxext{\Delta}{f^\ast(A)}}\} \\
+  & \jdeq
+T((p_{f^\ast(A)})^\ast(f^\ast(A))).
+\end{align*}
+Since we have $\delta_{f^\ast(A)}\in T((p_{f^\ast(A)})^\ast(f^\ast(A)))$, we obtain 
+\begin{equation*}
+\pi_2(f,A)\defeq \zeta_{A,f\circ p_{f^\ast(A)}}^{-1}(\delta_{f^\ast(A)}):\ctxext{\Delta}{f^\ast(A)}\to\ctxext{\Gamma}{A}
+\end{equation*}
+satisfying $p_A\circ\pi_2(f,A)\jdeq f\circ p_{f^\ast(A)}$.
 %\end{comment}
@@
 \begin{equation*}
-\pi_2(f,A)\circ\pi_2(g,f_\ast(A))\jdeq\pi_2(f\circ g,A).
+\pi_2(f,A)\circ\pi_2(g,f^\ast(A))\jdeq\pi_2(f\circ g,A).
 \end{equation*}
@@
 \begin{equation*}
-\zeta_{A,f\circ g\circ p_{g_\ast(f_\ast(A))}}(\pi_2(f,A)\circ\pi_2(g,f_\ast(A)))\jdeq\delta_{g_\ast(f_\ast(A))}.
-\end{equation*}
-Because $g\circ p_{g_\ast(f_\ast(A))}\jdeq p_{f_\ast(A)}\circ\pi_2(g,f_\ast(A))$, we get
+\zeta_{A,f\circ g\circ p_{g^\ast(f^\ast(A))}}(\pi_2(f,A)\circ\pi_2(g,f^\ast(A)))\jdeq\delta_{g^\ast(f^\ast(A))}.
+\end{equation*}
+Because $g\circ p_{g^\ast(f^\ast(A))}\jdeq p_{f^\ast(A)}\circ\pi_2(g,f^\ast(A))$, we get
 to use the naturality of $\zeta$:
 \begin{align*}
-& \zeta_{A,f\circ g\circ p_{g_\ast(f_\ast(A))}}(\pi_2(f,A)\circ\pi_2(g,f_\ast(A))) \\
-& \jdeq
-\zeta_{A,f\circ p_{f_\ast(A)}\circ\pi_2(g,f_\ast(A))}(\pi_2(f,A)\circ\pi_2(g,f_\ast(A))) \\
-& \jdeq
-\zeta_{(p_{f_\ast(A)})_\ast(f_\ast(A)),\pi_2(g,f_\ast(A))}(\zeta_{A,f\circ p_{f_\ast(A)}}(\pi_2(f,a))\circ\pi_2(g,f_\ast(A))) \\
-& \jdeq
-\zeta_{(p_{f_\ast(A)})_\ast(f_\ast(A)),\pi_2(g,f_\ast(A))}(\delta_{f_\ast(A)}\circ\pi_2(g,f_\ast(A))) \\
-& \jdeq
-\zeta_{(p_{f_\ast(A)})_\ast(f_\ast(A)),\pi_2(g,f_\ast(A))}(\zeta_{f_\ast(A),p_{f_\ast(A)}}(\catid{\ctxext{\Delta}{f_\ast(A)}})\circ\pi_2(g,f_\ast(A))) \\
-& \jdeq
-\zeta_{f_\ast(A),p_{f_\ast(A)}\circ\pi_2(g,f_\ast(A))}(\catid{\ctxext{\Delta}{f_\ast(A)}}\circ\pi_2(g,f_\ast(A))) \\
-& \jdeq
-\zeta_{f_\ast(A),p_{f_\ast(A)}\circ\pi_2(g,f_\ast(A))}(\pi_2(g,f_\ast(A))) \\
-& \jdeq
-\zeta_{f_\ast(A),g\circ p_{g_\ast(f_\ast(A))}}(\pi_2(g,f_\ast(A)))\\
-& \jdeq
-\delta_{g_\ast(f_\ast(A))}.
+& \zeta_{A,f\circ g\circ p_{g^\ast(f^\ast(A))}}(\pi_2(f,A)\circ\pi_2(g,f^\ast(A))) \\
+& \jdeq
+\zeta_{A,f\circ p_{f^\ast(A)}\circ\pi_2(g,f^\ast(A))}(\pi_2(f,A)\circ\pi_2(g,f^\ast(A))) \\
+& \jdeq
+\zeta_{(p_{f^\ast(A)})^\ast(f^\ast(A)),\pi_2(g,f^\ast(A))}(\zeta_{A,f\circ p_{f^\ast(A)}}(\pi_2(f,a))\circ\pi_2(g,f^\ast(A))) \\
+& \jdeq
+\zeta_{(p_{f^\ast(A)})^\ast(f^\ast(A)),\pi_2(g,f^\ast(A))}(\delta_{f^\ast(A)}\circ\pi_2(g,f^\ast(A))) \\
+& \jdeq
+\zeta_{(p_{f^\ast(A)})^\ast(f^\ast(A)),\pi_2(g,f^\ast(A))}(\zeta_{f^\ast(A),p_{f^\ast(A)}}(\catid{\ctxext{\Delta}{f^\ast(A)}})\circ\pi_2(g,f^\ast(A))) \\
+& \jdeq
+\zeta_{f^\ast(A),p_{f^\ast(A)}\circ\pi_2(g,f^\ast(A))}(\catid{\ctxext{\Delta}{f^\ast(A)}}\circ\pi_2(g,f^\ast(A))) \\
+& \jdeq
+\zeta_{f^\ast(A),p_{f^\ast(A)}\circ\pi_2(g,f^\ast(A))}(\pi_2(g,f^\ast(A))) \\
+& \jdeq
+\zeta_{f^\ast(A),g\circ p_{g^\ast(f^\ast(A))}}(\pi_2(g,f^\ast(A)))\\
+& \jdeq
+\delta_{g^\ast(f^\ast(A))}.
 \end{align*}
@@
 \begin{equation*}
-p_A\circ\pi_2(f,A)\circ\tilde\theta\jdeq f\circ p_{f_\ast(A)}\circ\tilde\theta\jdeq f.
+p_A\circ\pi_2(f,A)\circ\tilde\theta\jdeq f\circ p_{f^\ast(A)}\circ\tilde\theta\jdeq f.
 \end{equation*} 
@@
   & \jdeq
-\zeta_{A,f\circ p_{f_\ast(A)}\circ\tilde\theta}(\pi_2(f,A)\circ\tilde\theta)
+\zeta_{A,f\circ p_{f^\ast(A)}\circ\tilde\theta}(\pi_2(f,A)\circ\tilde\theta)
   \\
   & \jdeq
-\zeta_{(p_{f_\ast(A)})_\ast(f_\ast(A)),\tilde{\theta}}(\zeta_{A,f\circ p_{f_\ast(A)}}(\pi_2(f,A))\circ\tilde\theta)
+\zeta_{(p_{f^\ast(A)})^\ast(f^\ast(A)),\tilde{\theta}}(\zeta_{A,f\circ p_{f^\ast(A)}}(\pi_2(f,A))\circ\tilde\theta)
   \\
   & \jdeq
-\zeta_{(p_{f_\ast(A)})_\ast(f_\ast(A)),\tilde{\theta}}(\idtm{f_\ast(A)}\circ\tilde\theta)
+\zeta_{(p_{f^\ast(A)})^\ast(f^\ast(A)),\tilde{\theta}}(\idtm{f^\ast(A)}\circ\tilde\theta)
   \\
   & \jdeq
-\zeta_{(p_{f_\ast(A)})_\ast(f_\ast(A)),\tilde{\theta}}(\zeta_{f_\ast(A),p_{f_\ast(A)}}(\catid{\ctxext{\Delta}{f_\ast(A)}})\circ\tilde\theta)
+\zeta_{(p_{f^\ast(A)})^\ast(f^\ast(A)),\tilde{\theta}}(\zeta_{f^\ast(A),p_{f^\ast(A)}}(\catid{\ctxext{\Delta}{f^\ast(A)}})\circ\tilde\theta)
   \\
   & \jdeq
-\zeta_{f_\ast(A),p_{f_\ast(A)}\circ\tilde\theta}(\catid{\ctxext{\Delta}{f_\ast(A)}}\circ\tilde\theta) \\
-  & \jdeq
-\zeta_{f_\ast(A),\catid{\Delta}}(\tilde\theta) \\
+\zeta_{f^\ast(A),p_{f^\ast(A)}\circ\tilde\theta}(\catid{\ctxext{\Delta}{f^\ast(A)}}\circ\tilde\theta) \\
+  & \jdeq
+\zeta_{f^\ast(A),\catid{\Delta}}(\tilde\theta) \\
   & \jdeq 
@@
 \begin{tikzcd}[column sep=large]
-\ctxext{\Delta}{f_\ast(A)} \arrow[d,fib] \arrow[r,"{\pi_2(f,A)}"] & \ctxext{\Gamma}{A} \arrow[d,fib] \\
+\ctxext{\Delta}{f^\ast(A)} \arrow[d,fib] \arrow[r,"{\pi_2(f,A)}"] & \ctxext{\Gamma}{A} \arrow[d,fib] \\
 \Delta \arrow[r,"f"] & \Gamma
@@
 \end{equation*}
-Then $h$ corresponds uniquely to a section $k$ of $(f\circ g)_\ast(A)$. Since
-$(f\circ g)_\ast(A)\jdeq g_\ast(f_\ast(A))$, the section $k$ fits in the
+Then $h$ corresponds uniquely to a section $k$ of $(f\circ g)^\ast(A)$. Since
+$(f\circ g)^\ast(A)\jdeq g^\ast(f^\ast(A))$, the section $k$ fits in the
 commuting diagram
@@
 \begin{tikzcd}[column sep=huge]
-\ctxext{X}{g_\ast(f_\ast(A))} \arrow[d,xshift=.7ex,fib] \arrow[r,"{\pi_2(g,f_\ast(A))}"]
-& \ctxext{\Delta}{f_\ast(A)} \arrow[d,fib] \arrow[r,"{\pi_2(f,A)}"]
+\ctxext{X}{g^\ast(f^\ast(A))} \arrow[d,xshift=.7ex,fib] \arrow[r,"{\pi_2(g,f^\ast(A))}"]
+& \ctxext{\Delta}{f^\ast(A)} \arrow[d,fib] \arrow[r,"{\pi_2(f,A)}"]
 & \ctxext{\Gamma}{A} \arrow[d,fib] \\
@@
 \end{equation*}
-so we obtain $u\defeq\pi_2(g,f_\ast(A))\circ k:X\to\ctxext{\Delta}{f_\ast(A)}$. 
-Then we have $p_{f_\ast(A)}\circ u\jdeq g\circ p_{g_\ast(f_\ast(A))}\circ k\jdeq g$ and
+so we obtain $u\defeq\pi_2(g,f^\ast(A))\circ k:X\to\ctxext{\Delta}{f^\ast(A)}$. 
+Then we have $p_{f^\ast(A)}\circ u\jdeq g\circ p_{g^\ast(f^\ast(A))}\circ k\jdeq g$ and
 $\pi_2(f,A)\circ u\jdeq \pi_2(f\circ g,A)\circ k\jdeq h$. Note that
@@
 For $f:\Delta\to\Gamma$ and $A\in\mathrm{Ty}(\Gamma)$, we get a function
-$x\mapsto f_\ast(x):\mathrm{Tm}(A)\to\mathrm{Tm}(f_\ast(A))$ by defining $f_\ast(x)$
+$x\mapsto f^\ast(x):\mathrm{Tm}(A)\to\mathrm{Tm}(f^\ast(A))$ by defining $f^\ast(x)$
 to be the unique morphism such that the diagram
@@
 \begin{tikzcd}[column sep=huge] 
-\Delta \arrow[ddr,bend right=15,equals] \arrow[drr,bend left=15,"x\circ f"] \arrow[dr,densely dotted,near end,"{f_\ast(x)}"] \\
-& \ctxext{\Delta}{f_\ast(A)} \arrow[d,fib] \arrow[r,swap,"{\pi_2(f,A)}"] & \ctxext{\Gamma}{A} \arrow[d,fib] \\
+\Delta \arrow[ddr,bend right=15,equals] \arrow[drr,bend left=15,"x\circ f"] \arrow[dr,densely dotted,near end,"{f^\ast(x)}"] \\
+& \ctxext{\Delta}{f^\ast(A)} \arrow[d,fib] \arrow[r,swap,"{\pi_2(f,A)}"] & \ctxext{\Gamma}{A} \arrow[d,fib] \\
 & \Delta \arrow[r,swap,"f"] & \Gamma
@@
 
-The term $\delta_A\in\mathrm{Tm}((p_A)_\ast(A))$ is defined to be the unique morphism
+The term $\delta_A\in\mathrm{Tm}((p_A)^\ast(A))$ is defined to be the unique morphism
 such that the diagram
@@
 \ctxext{\Gamma}{A} \arrow[drr,bend left=15,equals] \arrow[ddr,bend right=15,equals] \arrow[dr,densely dotted,near end,"{\delta_{A}}"] \\
-& \ctxext{{\Gamma}{A}}{(p_A)_\ast(A)} \arrow[r,swap,"{\pi_2(A,A)}"] \arrow[d,fib] & \ctxext{\Gamma}{A} \arrow[d,fib] \\
+& \ctxext{{\Gamma}{A}}{(p_A)^\ast(A)} \arrow[r,swap,"{\pi_2(A,A)}"] \arrow[d,fib] & \ctxext{\Gamma}{A} \arrow[d,fib] \\
 & \ctxext{\Gamma}{A} \arrow[r,swap,"A"] & \Gamma
@@
 the universal property of pullbacks: the map $x\mapsto\pi_2(f,X)\circ x$
-from $\{x:\Delta\to\ctxext{\Delta}{f_\ast(A)}\mid p_{f_\ast(A)}\circ x\jdeq\catid{\Delta}\}$
+from $\{x:\Delta\to\ctxext{\Delta}{f^\ast(A)}\mid p_{f^\ast(A)}\circ x\jdeq\catid{\Delta}\}$
 to $\{\theta:\Delta\to\ctxext{\Gamma}{A}\mid p_A\circ\theta\jdeq f\}$ is a bijection.
@@
 %\end{comment}
-
-The following two definitions define subclasses of telescopic D-systems and 
-coherent telescopic D-systems, and the latter structure is the same as an E'-system
-(without empty context). In \autoref{defn:comp_to_telescopic}, we will show how
-to obtain a telescopic D-system from any comprehension category. This construction
-will also generate a telescopic D-system out of any D-system, because D-systems
-are comprehension categories.
-
-\begin{defn}
-Let $\cat{C}$ be a D-system. In this definition, we will write
-$\mathrm{Fam}$ for the presheaf of types. The D-system $\cat{C}$
-is said to be \define{telescopic}, if we have elements $\emptyf[\Gamma]\in
-\mathrm{Fam}(\Gamma)$ for each $\Gamma\in\cat{C}$, and maps 
-$P\mapsto\ctxext{A}{P}:\mathrm{Fam}(\ctxext{\Gamma}{A})\to \mathrm{Fam}(\Gamma)$
-such that
-\begin{enumerate}
-\item 
-\begin{enumerate}
-\item for any $A\in\mathrm{Fam}(\Gamma)$ and $P\in\mathrm{Fam}(\ctxext{\Gamma}{A})$,
-one has $\ctxext{\Gamma}{{A}{P}}\jdeq\ctxext{{\Gamma}{A}}{P}$, and
-\begin{equation*}
-p_{\ctxext{A}{P}}\jdeq p_A\circ p_P,
-\end{equation*}
-\item and for any $\Gamma\in\cat{C}$ one has $\ctxext{\Gamma}{\emptyf[\Gamma]}\jdeq\Gamma$, and
-\begin{equation*}
-p_{\emptyf[\Gamma]}\jdeq\catid{\Gamma}.
-\end{equation*}
-\end{enumerate}
-\item 
-\begin{enumerate}
-\item for any $A\in\mathrm{Fam}(\Gamma)$, $P\in\mathrm{Fam}(\ctxext{\Gamma}{A})$
-and $Q\in\mathrm{Fam}(\ctxext{{\Gamma}{A}}{P})$, one has
-\begin{equation*}
-\ctxext{{A}{P}}{Q}\jdeq\ctxext{A}{{P}{Q}},
-\end{equation*}
-\item and for any $A\in\mathrm{Fam}(\Gamma)$, one has
-\begin{equation*}
-\ctxext{\emptyf[\Gamma]}{A}\jdeq\ctxext{A}{\emptyf[\ctxext{\Gamma}{A}]}\jdeq A.
-\end{equation*}
-\end{enumerate}
-\item 
-\begin{enumerate}
-\item for any $A\in\mathrm{Fam}(\Gamma)$, $P\in\mathrm{Fam}(\ctxext{\Gamma}{A})$
-and any $f:\Delta\to\Gamma$ in $\cat{C}$, one has
-\begin{equation*}
-f_\ast(\ctxext{A}{P})\jdeq \ctxext{(f_\ast(A))}{(\pi_2(f,A)_\ast(P))},
-\end{equation*}
-\item and for any $f:\Delta\to\Gamma$ in $\cat{C}$, one has
-\begin{equation*}
-f_\ast(\emptyf[\Gamma])\jdeq\emptyf[\Delta].
-\end{equation*}
-\end{enumerate}
-\end{enumerate} 
-\end{defn}
-
-\begin{lem}
-A D-system is telescopic if and only if  the following two conditions hold:
-\begin{enumerate}
-\item There is a section $\phi:\cat{C}\to \int_{\cat{C}}\mathrm{Ty}$ of $\pi$.
-\item For every $P\in\mathrm{Ty}(\ctxext{\Gamma}{A})$,
-the morphism $p_A:\ctxext{\Gamma}{A}\to\Gamma$ has a unique cocartesian lift
-at $P$.
-\end{enumerate}
-\end{lem}
-
-\begin{proof}
-Let $\mathbb{D}$ be a telescopic D-system. To show that $\mathbb{D}$ has the two
-stated properties:
-\begin{enumerate}
-\item
-\item We have to show that each for $P\in\mathrm{Ty}(\ctxext{\Gamma}{A})$,
-the projection $p_A:\ctxext{\Gamma}{A}\to\Gamma$ has a unique cocartesian lift
-at $P$. Note that the morphism $(\ctxext{\Gamma}{A},P)\to (\Gamma,\ctxext{A}{P})$
-in $\int_{\cat{C}}\mathrm{Ty}$ 
-\end{enumerate}
-
-Let $\mathbb{D}$ be a D-system for which the stated two conditions hold. To show
-that $\mathbb{D}$ is a telescopic D-system, we systematically go through each of the
-properties:
-\begin{enumerate}
-\item 
-\end{enumerate} 
-\end{proof}
-
-\begin{conj}
-The notion of being telescopic is a property.
-\end{conj}
-
-\begin{defn}
-A telescopic D-system is said to be \define{coherent}, if for any $A,A'\in\mathrm{Fam}(\Gamma)$
-such that $p_A\jdeq p_{A'}$, and any $f:\Delta\to\Gamma$, we have
-$p_{f_\ast(A)}\jdeq p_{f_\ast(A')}$ and $\pi_2(f,A)\jdeq \pi_2(f,A')$.
-\end{defn}
-
-\begin{thm}
-An E'-system without empty context is the same thing as a coherent telescopic D-system.
-\end{thm}
-
-\begin{proof}
-Let $\mathbb{E}$ be an E'-system without empty context. Then we define the
-presheaf $\mathrm{Fam}$ on $\cat{C}$ on objects by
-\begin{equation*}
-\mathrm{Fam}(\Gamma)\defeq \mathrm{Obj}(\cat{F}/\Gamma)
-\end{equation*}
-and on morphisms by the action on objects of $f_\ast$, for each $f:\Delta\to
-\Gamma$ in $\cat{C}$. By condition (b) and (c) in the definition of E'-systems, it
-follows that this defines a contravariant functor.
-
-We define a functor $\mathcal{P}:\int_{\cat{C}}\to\cat{C}^\square$ by sending
-the pair $(\Gamma,A)$ to the morphism $A:\ctxext{\Gamma}{A}\to\Gamma$ in
-$\cat{C}$, and by sending $f:(\Delta,f_\ast(A))\to (\Gamma,A)$ to the pullback
-square
-\begin{equation*}
-\begin{tikzcd}[column sep=large]
-\ctxext{\Delta}{f_\ast(A)} \arrow[d,swap,"{f_\ast(A)}"] \arrow[r,"{\pi_2(f,A)}"] &
-\ctxext{\Gamma}{A} \arrow[d,"A"] \\ \Delta \arrow[r,"f"] & \Gamma
-\end{tikzcd}
-\end{equation*}
-This is also functorial because of conditions (b) and (c) in the definition of
-E'-systems. Thus, we have defined a D-system
-\begin{equation*}
-\begin{tikzcd}
-\int_{\cat{C}}\mathrm{Fam} \arrow[rr,"{\mathcal{P}}"] \arrow[dr,"{\pi}"] & & \cat{C}^\square \arrow[dl,"{\mathrm{cod}}"] \\ & \cat{C}
-\end{tikzcd}
-\end{equation*}
-and it is straightforward to verify that this D-system is telescopic and coherent.
-
-Conversely, suppose $\mathbb{D}$ is a telescopic D-system with underlying category
-$\cat{C}$. Define $\cat{F}$ to be the category with the same objects as
-$\cat{C}$, and morphisms of the form $p_A$. To check that this forms a category,
-note that $p_{\emptyf[\Gamma]}\jdeq\catid{\Gamma}$ and $p_A\circ p_P\jdeq
-p_{\ctxext{A}{P}}$.
-
-Now let $f:\Delta\to\Gamma$ be a morphism in $\cat{C}$. For any $A\in\cat{F}/\Gamma$,
-we have an element $\tilde{A}\in\mathrm{Fam}(\Gamma)$ such that $p_{\tilde{A}}
-\jdeq A$, and we can form a pullback square
-\begin{equation*}
-\begin{tikzcd}[column sep=large]
-\ctxext{\Delta}{f_\ast(\tilde{A})} \arrow[r,"{\pi_2(f,\tilde{A})}"] \arrow[d,fib,swap,"p_{f_\ast(\tilde{A})}"]
-& \ctxext{\Gamma}{\tilde{A}} \arrow[d,fib,"{p_{\tilde{A}}}"] \\
-\Delta \arrow[r,"f"] & \Gamma
-\end{tikzcd}
-\end{equation*}
-By the coherence condition, this choice of pullback square does not depend on the
-choice of $\tilde{A}$ such that $p_{\tilde{A}}\jdeq A$ (having this fact 
-available is precisely what the coherence condition is for). The verification of conditions
-(a) to (d) of \autoref{defn:E'sys} is straightforward.
-\end{proof}
-
-\subsection{Coherent comprehension categories}
-In this section we recall Jacobs' notion of comprehension categories, and 
-Dybjer's notion of categories with families. It is shown in 
-\cite{Hofmann_syntax_semantics} that the category of categoriew with families
-is equivalent to the category of categories with attributes, which forms a
-subcategory of comprehension categories. Thus, for our purpose of comparing
-E-systems with various existing notions of models of type dependency, it suffice
-to make the comparison with comprehension categories.
-
-In this section we do the preparateions for this comparison. After having
-introduced categories with families, we prove that categories with families
-with the extra structure of `being closed under family extension' in a coherent
-way, are E'-systems without empty context.
-
-\begin{defn}
-Let $p:\cat{E}\to\cat{B}$ be a functor and let $f:\Delta\to \Gamma$ be a morphism in $\cat{B}$.
-A morphism $g:A\to B$ in $\cat{E}$ is said to be \define{cartesian over $f$}, if
-\begin{enumerate}
-\item $p(g)\jdeq f$, and 
-\item For all $g':A'\to B$ with $p(g')\jdeq f$, there is a unique $u:A'\to A$
-such that $g'=g\circ u$ and $p(u)\jdeq\catid{\Delta}$. 
-\end{enumerate}
-Dually, a morphism $g:A\to B$ in $\cat{E}$ is said to be \define{cocartesian over
-$f$}, if
-\begin{enumerate}
-\item $p(g)\jdeq f$, and
-\item For all $g':A\to B'$ with $p(g')\jdeq f$, there is a unique $v:B\to B'$
-such that $g'=v\circ g$ and $p(v)\jdeq\catid{\Gamma}$. 
-\end{enumerate}
-A (co)cartesian morphism over $f$ is also called a \define{(co)cartesian lift of $f$}.
-\end{defn}
-
-\begin{defn}
-A functor $p:\cat{E}\to\cat{B}$ is called a \define{Grothendieck fibration} if
-every morphism into an object of the form $p(A)$ for some object $A$ of $\cat{E}$ 
-has a cartesian lift, and cartesian morphisms are closed under composition.
-\end{defn}
-
-\begin{defn}
-A Grothendieck fibration $p:\cat{E}\to\cat{B}$ is said to be \define{split}, if
-there is a functorial choice of cartesian lifts, i.e.~for every $f:\Delta\to \Gamma$ in 
-$\cat{B}$ and any $B\in \cat{E}$ such that $p(B)\jdeq \Gamma$,
-there is a cartesian lift $j(B,f):A \to B$ for some object $A$ of $\cat{E}$, such that 
-$j(B,\idfunc[\Gamma])\jdeq\catid{B}$ and 
-$j(B, f\circ g)\jdeq j(B,f)\circ j(\mathrm{dom}(j(B,f)),g)$ for each
-$f:\Delta\to\Gamma$ and each $g:\greek{E}\to\Delta$. 
-\end{defn}
-
-\begin{eg}
-For any presheaf $P:\cat{C}\to\mathbb{Set}$, the projection functor
-$\int_\cat{C} P\to\cat{C}$ is a Grothendieck fibration.
-\end{eg}
-
-\begin{defn}
-A \define{comprehension category} $\mathcal{P}$ consists of categories $\cat{E}$, $\cat{C}$ and a
-commuting triangle
-\begin{equation*}
-\begin{tikzcd}
-\cat{E} \arrow[rr,"\mathcal{P}"] \arrow[dr,swap,"p"] & & \cat{C}^{\to} \arrow[dl,"\mathrm{cod}"] \\
-& \cat{C}
-\end{tikzcd}
-\end{equation*}
-of functors, such that $p$ is a Grothendieck fibration, and $\mathcal{P}$ maps
-cartesian morphisms in $\cat{E}$ to pullback squares in $\cat{C}$. We say that
-$\mathcal{P}$ is a \define{full} comprehension category if the functor $\mathcal{P}:
-\cat{E}\to\cat{C}^{\to}$ is full and faithful; and we say that
-$\mathcal{P}$ is \define{split} if $p$ is split.
-\end{defn}
-
-\begin{eg}
-Since any morphism of $\int_{\cat{C}}\mathrm{Ty}$ is cartesian, it follows that D-systems
-are comprehension categories. As D-systems are built up from a presheaf of types
-on the underlying category, they are in fact discrete comprehension categories.
-\end{eg}
 
@@
 We see that it is not evident at all to turn categories with families into
-E-systems, even if we ignore the issue that categories with families map
+telescopic type systems, even if we ignore the issue that categories with families map
 naturally into E'-systems without empty context.
-Thus it seems that the type theories of categories with families and of E-systems
+Thus it seems that the type theories of categories with families and of telescopic type systems
 are not directly comparable.
@@
 As a by-product, we get that the full and faithful functor from B-systems to
-E-systems factors through D-systems.
+telescopic type systems factors through D-systems.
 
@@
 \Lambda(\Gamma) \arrow[d,swap,"\mathcal{P}"] \arrow[r,"\Lambda(f)"] & \Lambda(\Delta) \arrow[d,"\mathcal{P}"] \\
-\cat{F}/\Gamma \arrow[r,densely dotted,"{f_\ast}"] & \cat{F}/\Delta
+\cat{F}/\Gamma \arrow[r,densely dotted,"{f^\ast}"] & \cat{F}/\Delta
 \end{tikzcd}
@@
 \begin{tikzcd}[column sep=large]
-\ctxext{\Delta}{f_\ast(A')} \arrow[d,fib] \arrow[r,"{\pi_2(f,A')}"] & \ctxext{\Gamma}{A} \arrow[d,fib] \\
+\ctxext{\Delta}{f^\ast(A')} \arrow[d,fib] \arrow[r,"{\pi_2(f,A')}"] & \ctxext{\Gamma}{A} \arrow[d,fib] \\
 \Delta \arrow[r,"f"] & \Gamma
@@
 $p^\Delta\circ\Lambda(f)$ factors through $p^\Gamma$. This defines the functor
-$f_\ast$ in the commuting square
+$f^\ast$ in the commuting square
 \begin{equation*}
@@
 \Lambda(\Gamma) \arrow[d,swap,"p"] \arrow[r,"\Lambda(f)"] & \Lambda(\Delta) \arrow[d,"p"] \\
-\cat{F}/\Gamma \arrow[r,densely dotted,"{f_\ast}"] & \cat{F}/\Delta
-\end{tikzcd}
-\end{equation*}
-We need to verify that $f_\ast$ is a functorial choice of pullbacks, as in the
-definition of \autoref{defn:E'sys}. Since $f_\ast$ is defined as an extension
+\cat{F}/\Gamma \arrow[r,densely dotted,"{f^\ast}"] & \cat{F}/\Delta
+\end{tikzcd}
+\end{equation*}
+We need to verify that $f^\ast$ is a functorial choice of pullbacks, as in the
+definition of \autoref{defn:E'sys}. Since $f^\ast$ is defined as an extension
 of $\Lambda(f)$, it is enough to verify the analoguous functoriality properties
@@
 \item Let $\Gamma\in\cat{C}$. We will prove by induction on $A\in\Lambda(\Gamma)$
-that $(\catid{\Gamma})_\ast(A)\jdeq A$ and that $\pi_2(\catid{\Gamma},A)\jdeq \catid{\ctxext{\Gamma}{A}}$.
+that $(\catid{\Gamma})^\ast(A)\jdeq A$ and that $\pi_2(\catid{\Gamma},A)\jdeq \catid{\ctxext{\Gamma}{A}}$.
 Note that these properties hold by definition for $A\jdeq\catid{\Gamma}$. For the
 inductive step, assume that $A\in\Lambda(\Gamma)$ is such that
-$(\catid{\Gamma})_\ast(A)\jdeq A$ and $\pi_2(\catid{\Gamma},A)\jdeq
+$(\catid{\Gamma})^\ast(A)\jdeq A$ and $\pi_2(\catid{\Gamma},A)\jdeq
 \catid{\ctxext{\Gamma}{A}}$ and let $B\in\mathrm{Ty}(\ctxext{\Gamma}{A})$. Then
@@
 \begin{align*}
-(\catid{\Gamma})_\ast(A\circ p_B) & \jdeq (\catid{\Gamma})_\ast(A)\circ \pi_2(\catid{\Gamma},A)_\ast(p_B) \\
+(\catid{\Gamma})^\ast(A\circ p_B) & \jdeq (\catid{\Gamma})^\ast(A)\circ \pi_2(\catid{\Gamma},A)^\ast(p_B) \\
 & \jdeq A\circ p_B
@@
 \item Let $f:\Delta\to \Gamma$ and $g:X\to \Delta$. We will prove by induction
-on $A\in\Lambda(\Gamma)$ that $(f\circ g)_\ast(A)\jdeq g_\ast(f_\ast(A))$ and
-that $\pi_2(f\circ g,A)\jdeq \pi_2(f,A)\circ \pi_2(g,f_\ast(A))$. Note that
+on $A\in\Lambda(\Gamma)$ that $(f\circ g)^\ast(A)\jdeq g^\ast(f^\ast(A))$ and
+that $\pi_2(f\circ g,A)\jdeq \pi_2(f,A)\circ \pi_2(g,f^\ast(A))$. Note that
 these properties hold by definition for $A\jdeq\catid{\Gamma}$. For the inductive
-step, assume that $A\in\Lambda(\Gamma)$ is such that $(f\circ g)_\ast(A)\jdeq g_\ast(f_\ast(A))$ and
-$\pi_2(f\circ g,A)\jdeq \pi_2(f,A)\circ \pi_2(g,f_\ast(A))$, and let
+step, assume that $A\in\Lambda(\Gamma)$ is such that $(f\circ g)^\ast(A)\jdeq g^\ast(f^\ast(A))$ and
+$\pi_2(f\circ g,A)\jdeq \pi_2(f,A)\circ \pi_2(g,f^\ast(A))$, and let
 $B\in\mathrm{Ty}(\ctxext{\Gamma}{A})$. Then we have
 \begin{align*}
-(f\circ g)_\ast(A\circ p_B) & \jdeq \pi_2(f\circ g,A)_\ast(p_B) \\
-& \jdeq (\pi_2(f,A)\circ\pi_2(g,f_\ast(A)))_\ast(p_B) \\
-& \jdeq \pi_2(g,f_\ast)_\ast(\pi_2(f,A)_\ast(p_B)) \\
-& \jdeq g_\ast(f_\ast(A\circ p_B))
+(f\circ g)^\ast(A\circ p_B) & \jdeq \pi_2(f\circ g,A)^\ast(p_B) \\
+& \jdeq (\pi_2(f,A)\circ\pi_2(g,f^\ast(A)))^\ast(p_B) \\
+& \jdeq \pi_2(g,f^\ast)^\ast(\pi_2(f,A)^\ast(p_B)) \\
+& \jdeq g^\ast(f^\ast(A\circ p_B))
 \end{align*}
@@
 $\Gamma\in\cat{C}$, the category $p^{-1}(\Gamma)$ is the category of internal
-morphisms the E-system $\mathcal{E}/\Gamma$.
+morphisms the telescopic type system $\mathcal{E}/\Gamma$.
 
@@
 a morphism in $\cat{C}$ and let $A\in\cat{F}/\Gamma$. Then we have $\catid{\Delta},\catid{\Gamma}\in\cat{E}$,
-and $(f,\pi_2(f,A)):f_\ast(A)\to A$. It is routine to verify that $(f,\pi_2(f,A))$
+and $(f,\pi_2(f,A)):f^\ast(A)\to A$. It is routine to verify that $(f,\pi_2(f,A))$
 is a cartesian lift of $f$, and that this choice of catesian lifts is functorial.
@@
 commutes. Let $A\in\mathrm{Ty}(\Gamma)$. Then we need to show that
-$\mathcal{G}(f)_\ast(\mathcal{G}(A))\jdeq \mathcal{G}(f_\ast(A))$. By the uniqueness
+$\mathcal{G}(f)^\ast(\mathcal{G}(A))\jdeq \mathcal{G}(f^\ast(A))$. By the uniqueness
 assumption of well-founded D-systems, it suffices to show that
-$p_{\mathcal{G}(f)_\ast(\mathcal{G}(A))}\jdeq p_{\mathcal{G}(f_\ast(A))}$. To
+$p_{\mathcal{G}(f)^\ast(\mathcal{G}(A))}\jdeq p_{\mathcal{G}(f^\ast(A))}$. To
 see this, we do a simple calculation
 \begin{align*}
-p_{\mathcal{G}(f)_\ast(\mathcal{G}(A))}
-& \jdeq \mathcal{G}(f)_\ast(p_{\mathcal{G}(A)}) \\
-& \jdeq \mathcal{G}(f)_\ast(\mathcal{G}(p_A)) \\
-& \jdeq \mathcal{G}(f_\ast(p_A)) \\
-& \jdeq \mathcal{G}(p_{f_\ast(A)}) \\
-& \jdeq p_{\mathcal{G}(f_\ast(A))} \qedhere
+p_{\mathcal{G}(f)^\ast(\mathcal{G}(A))}
+& \jdeq \mathcal{G}(f)^\ast(p_{\mathcal{G}(A)}) \\
+& \jdeq \mathcal{G}(f)^\ast(\mathcal{G}(p_A)) \\
+& \jdeq \mathcal{G}(f^\ast(p_A)) \\
+& \jdeq \mathcal{G}(p_{f^\ast(A)}) \\
+& \jdeq p_{\mathcal{G}(f^\ast(A))} \qedhere
 \end{align*}
@@
 For $\Gamma\in\cat{C}$, we define $\mathrm{Ty}(\Gamma)$ to be the set of
-families of length one. Since any $f_\ast$ is stratified, it follows that
+families of length one. Since any $f^\ast$ is stratified, it follows that
 $\mathrm{Ty}$ is a presheaf on $\cat{C}$. The choice of pullback squares
@@
 \begin{tikzcd}[column sep=large]
-\ctxext{\Delta}{f_\ast(A)} \arrow[r,"{\pi_2(f,A)}"] \arrow[d,fib] & \ctxext{\Gamma}{A} \arrow[d,fib] \\
+\ctxext{\Delta}{f^\ast(A)} \arrow[r,"{\pi_2(f,A)}"] \arrow[d,fib] & \ctxext{\Gamma}{A} \arrow[d,fib] \\
 \Delta \arrow[r,swap,"f"] & \Gamma

### 2016-02-12 e-systems.tex
asterisk up

@@
-\section{E-systems}\label{sec:esys_defn}
+\section{Telescopic type systems}\label{sec:esys_defn}
 
@@
 
-\subsection{The definition of E-systems}
+\subsection{The definition of telescopic type systems}
 
-Although it is now possible to state the definition of an E-system without
+Although it is now possible to state the definition of a telescopic type system without
 any difficulty, it shall be convenient later to also have a category of
-pre-E-systems at our disposal. Pre-E-systems shall by default be non-unital,
-i.e.~they do not come with a pre-projection system, while E-systems shall
+pre-telescopic type systems at our disposal. Pre-telescopic type systems shall by default be non-unital,
+i.e.~they do not come with a pre-projection system, while telescopic type systems shall
 always be assumed to be unital. In the present context, we shall not need
-unital pre-E-systems or non-unital E-systems.
+unital pre-telescopic type systems or non-unital telescopic type systems.
 
 \begin{defn}
-A \define{pre-E-system} is a pre-category $\cat{F}$ with term structure
+A \define{pre-telescopic type system} is a pre-category $\cat{F}$ with term structure
 with a pre-weakening structure $W$, a pre-substitution structure $S$ and a terminal
-object. A \define{unital pre-E-system}, or simply \define{pre-E-system}, is a non-unital pre-E-system
-which also has the structure of a pre-projection system. 
+object.
+A \define{unital pre-telescopic type system}, or simply \define{pre-telescopic type system}, is a non-unital pre-telescopic type system which also has the structure of a pre-projection system.
 \end{defn}
@@
 \begin{defn}
-A \define{pre-E-homomorphism} from $\mathbb{E}\to\mathbb{D}$ is a functor
+A \define{pre-telescopic homomorphism} from $\mathbb{E}\to\mathbb{D}$ is a functor
 $H:\mathbb{C}_\mathbb{E}\to\mathbb{C}_\mathbb{D}$ between the underlying categories
@@
 \label{defn:esystems}
-An \define{E-system} is a unital pre-E-system in which 
+A \define{telescopic type system} is a unital pre-telescopic type system in which 
 \begin{enumerate}
@@
 \end{enumerate}
-An \define{E-homomorphism} $H:\mathbb{E}\to\mathbb{D}$ is a functor from
+An \define{telescopic homomorphism} $H:\mathbb{E}\to\mathbb{D}$ is a functor from
 $\cat{F}_{\mathbb{E}}$ to $\cat{F}_{\mathbb{D}}$ which is both a projection
@@
 \begin{cor}
-For any object $\Gamma$ of an E-system $\cat{F}$, the slice pre-category $\cat{F}/\Gamma$
-is an E-system.
+For any object $\Gamma$ of a telescopic type system $\cat{F}$, the slice pre-category $\cat{F}/\Gamma$
+is a telescopic type system.
 \end{cor}
+
+\begin{defn}
+We write $\mathbf{Tel}$ for the category of telescopic type systems where the underlying
+functor of each E-homomorphism lifts factorizations.
+\end{defn}
 

### 2016-02-12 internal_homs.tex
asterisk up

@@
-\section{Elementary properties of E-systems}\label{sec:esys_props}
-
-The theory of E-systems could be seen as a sort of category theory with only
+\section{Elementary properties of telescopic type systems}\label{sec:esys_props}
+
+The theory of telescopic type systems could be seen as a sort of category theory with only
 projections and their sections. We elaborate on this point of view by defining
@@
 for every $f:A\to B$ in context $\Gamma$, a pre-composition operation
-$f_\ast$, which is going to be an E-homomorphism from 
+$f^\ast$, which is going to be an E-homomorphism from 
 $\cat{F}/\ctxext{\Gamma}{B}$ to $\cat{F}/\ctxext{\Gamma}{A}$. This allows us
@@
 where we consider the morphism $g$ over $f$ to be an internal morphism from 
-$P$ to $f_\ast(Q)$ in context $\ctxext{\Gamma}{A}$, capturing the
+$P$ to $f^\ast(Q)$ in context $\ctxext{\Gamma}{A}$, capturing the
 essence of being a morphism from the family $P$ over $A$ to the family
@@
 in \autoref{sec:E'}. The thread running through this whole section is proving
-that the notions of E-system and E'-system are the same. In \autoref{sec:E'}
-we show directly that any E'-system is an E-system. We end this section by
+that the notions of telescopic type system and E'-system are the same. In \autoref{sec:E'}
+we show directly that any E'-system is a telescopic type system. We end this section by
 showing that the converse also holds.
@@
 \begin{tikzcd}[column sep=large]
-\ctxext{\Delta}{f_\ast(A)} \arrow[r,"{\pi_2(f,A)}"] \arrow[d,fib,swap,"{f_\ast(A)}"] & \ctxext{\Gamma}{A} \arrow[d,fib,"A"] \\
+\ctxext{\Delta}{f^\ast(A)} \arrow[r,"{\pi_2(f,A)}"] \arrow[d,fib,swap,"{f^\ast(A)}"] & \ctxext{\Gamma}{A} \arrow[d,fib,"A"] \\
 \Delta \arrow[r,"f"] & \Gamma
@@
 \end{equation*}
-such that $f_\ast(A)\in\cat{F}/\Delta$. Explicitly, the functoriality
+such that $f^\ast(A)\in\cat{F}/\Delta$. Explicitly, the functoriality
 requirement is that
@@
 \item For any $f:\Delta\to\Gamma$, one has 
-$f_\ast(\catid{\Gamma})\jdeq\catid{\Delta}$ and
+$f^\ast(\catid{\Gamma})\jdeq\catid{\Delta}$ and
 \begin{equation*}
@@
 \end{equation*}
-\item \label{defn:E'sys:b} For any $A\in\cat{F}/\Gamma$, one has $(\catid{\Gamma})_\ast(A)\jdeq A$ and
+\item \label{defn:E'sys:b} For any $A\in\cat{F}/\Gamma$, one has $(\catid{\Gamma})^\ast(A)\jdeq A$ and
 \begin{equation*}
@@
 \item \label{defn:E'sys:c} For any $f:\Delta\to\Gamma$, $g:\greek{E}\to\Delta$ and $A\in\cat{F}/\Gamma$, one
-has $(f\circ g)_\ast(A)\jdeq g_\ast(f_\ast(A))$ and 
-\begin{equation*}
-\pi_2(f\circ g,A)\jdeq \pi_2(f,A)\circ\pi_2(g,f_\ast(A))
+has $(f\circ g)^\ast(A)\jdeq g^\ast(f^\ast(A))$ and 
+\begin{equation*}
+\pi_2(f\circ g,A)\jdeq \pi_2(f,A)\circ\pi_2(g,f^\ast(A))
 \end{equation*}
 \item For any $P\in\cat{F}/\ctxext{\Gamma}{A}$ and $f:\Delta\to\Gamma$, one has
-$f_\ast(P)\jdeq\pi_2(f,A)_\ast(P)$ and 
+$f^\ast(P)\jdeq\pi_2(f,A)^\ast(P)$ and 
 \begin{equation*}
@@
 \end{enumerate}
-For any $f:\Delta\to\Gamma$ we shall write $f_\ast$ for the induced functor
+For any $f:\Delta\to\Gamma$ we shall write $f^\ast$ for the induced functor
 $\cat{F}/\Gamma\to\cat{F}/\Delta$. 
@@
 Without a terminal object in $\cat{F}$, the pre-category $\cat{C}$ does not correspond
-directly to anything in an E-system.
+directly to anything in a telescopic type system.
 \end{rmk}
@@
 \begin{lem}\label{lem:pb_selfdistributive}
-Let $\mathbb{E}$ be an E'-system. Then $f_\ast$ can be extended to a functor from
+Let $\mathbb{E}$ be an E'-system. Then $f^\ast$ can be extended to a functor from
 $\cat{C}_\cat{F}/\Gamma$ to $\cat{C}_\cat{F}/\Delta$, for each $f:\Delta\to\Gamma$, and moreover the diagram
@@
 \begin{tikzcd}[column sep=large]
-\cat{C}_\cat{F}/\ctxext{\Gamma}{B} \arrow[r,"{f_\ast/B}"] \arrow[d,swap,"{g_\ast}"]
-& \cat{C}_\cat{F}/\ctxext{\Delta}{f_\ast(B)} \arrow[d,"{(f_\ast(g))_\ast}"] \\
-\cat{C}_\cat{F}/\ctxext{\Gamma}{A} \arrow[r,swap,"{f_\ast/A}"]
-& \cat{C}_\cat{F}/\ctxext{\Delta}{f_\ast(A)}
+\cat{C}_\cat{F}/\ctxext{\Gamma}{B} \arrow[r,"{f^\ast/B}"] \arrow[d,swap,"{g^\ast}"]
+& \cat{C}_\cat{F}/\ctxext{\Delta}{f^\ast(B)} \arrow[d,"{(f^\ast(g))^\ast}"] \\
+\cat{C}_\cat{F}/\ctxext{\Gamma}{A} \arrow[r,swap,"{f^\ast/A}"]
+& \cat{C}_\cat{F}/\ctxext{\Delta}{f^\ast(A)}
 \end{tikzcd}
@@
 It is straightforward to use the universal property of the pullback to extend
-$f_\ast$ to a functor from $\cat{C}_\cat{F}/\Gamma$ to $\cat{C}_\cat{F}/\Delta$.
+$f^\ast$ to a functor from $\cat{C}_\cat{F}/\Gamma$ to $\cat{C}_\cat{F}/\Delta$.
 To see that the diagram commutes, we calculate
 \begin{align*}
-f_\ast/A\circ g_\ast
-& \jdeq
-\pi_2(f,A)_\ast\circ g_\ast \\
-& \jdeq
-(g\circ\pi_2(f,A))_\ast \\
-& \jdeq
-(\pi_2(f,B)\circ f_\ast(g))_\ast \\
-& \jdeq
-(f_\ast(g))_\ast\circ\pi_2(f,B)_\ast \\
-& \jdeq
-(f_\ast(g))_\ast\circ f_\ast/B.\qedhere
+f^\ast/A\circ g^\ast
+& \jdeq
+\pi_2(f,A)^\ast\circ g^\ast \\
+& \jdeq
+(g\circ\pi_2(f,A))^\ast \\
+& \jdeq
+(\pi_2(f,B)\circ f^\ast(g))^\ast \\
+& \jdeq
+(f^\ast(g))^\ast\circ\pi_2(f,B)^\ast \\
+& \jdeq
+(f^\ast(g))^\ast\circ f^\ast/B.\qedhere
 \end{align*}
@@
 \begin{thm}
-An E-system is the same thing as an E'-system.
+A telescopic type system is the same thing as an E'-system.
 \end{thm}
@@
 \end{equation*}
-Since each functor $f_\ast$ is extended to all of $\cat{C}_{\cat{F}}/\Gamma$,
+Since each functor $f^\ast$ is extended to all of $\cat{C}_{\cat{F}}/\Gamma$,
 it follows immediately that those are functors with term structure.
 
-We define for any $A\in\cat{F}/\Gamma$, the functor $W_A\defeq A_\ast:\cat{F}/\Gamma\to\cat{F}/\ctxext{\Gamma}{A}$.
-Likewise, we define for any $x\in T(A)$, the functor $S_x\defeq x_\ast:\cat{F}/\ctxext{\Gamma}{A}\to\cat{F}/\Gamma$.
+We define for any $A\in\cat{F}/\Gamma$, the functor $W_A\defeq A^\ast:\cat{F}/\Gamma\to\cat{F}/\ctxext{\Gamma}{A}$.
+Likewise, we define for any $x\in T(A)$, the functor $S_x\defeq x^\ast:\cat{F}/\ctxext{\Gamma}{A}\to\cat{F}/\Gamma$.
 We also define $\tfid{A}:T(W_A(A))$ by the universal property of pullbacks:
@@
 
-It remains to verify the rest of the conditions of E-systems.
+It remains to verify the rest of the conditions of telescopic type systems.
 \begin{itemize}
@@
 \begin{equation*}
-S_x\circ W_A\jdeq x_\ast\circ A_\ast\jdeq (A\circ x)_\ast\jdeq (\catid{\Gamma})_\ast\jdeq \catid{\cat{C}_{\cat{F}}/\Gamma}.
+S_x\circ W_A\jdeq x^\ast\circ A^\ast\jdeq (A\circ x)^\ast\jdeq (\catid{\Gamma})^\ast\jdeq \catid{\cat{C}_{\cat{F}}/\Gamma}.
 \end{equation*}
@@
 \begin{equation*}
-S_{\tfid{A}}\circ W_A/A \jdeq S_{\tfid{A}}\circ \pi_2(A,A)_\ast \jdeq (\pi_2(A,A)\circ\tfid{A})_\ast \jdeq
-(\catid{\ctxext{\Gamma}{A}})_\ast \jdeq \catid{\cat{C}_{\cat{F}}/\ctxext{\Gamma}{A}} 
+S_{\tfid{A}}\circ W_A/A \jdeq S_{\tfid{A}}\circ \pi_2(A,A)^\ast \jdeq (\pi_2(A,A)\circ\tfid{A})^\ast \jdeq
+(\catid{\ctxext{\Gamma}{A}})^\ast \jdeq \catid{\cat{C}_{\cat{F}}/\ctxext{\Gamma}{A}} 
 \end{equation*}
@@
 \begin{comment}
-We then demonstrate how to define a category internal to an E-system. The main
-point of this lies mainly in demonstrating how the operations of E-systems can
+We then demonstrate how to define a category internal to a telescopic type system. The main
+point of this lies mainly in demonstrating how the operations of telescopic type systems can
 be used to define a category, because it is no surprise that a category can be
 defined in type theory. Nevertheless, it is noteworthy that we need no further 
-conditions on E-systems.
+conditions on telescopic type systems.
 This is unlike the situation in category theory, where we need
 to assume that the ambient category has finite limits in order to say what an
-internal category is. In fact, it is possible to define an E-system internal
-to an E-system, just using the algebraic structure of an E-system, and actually
+internal category is. In fact, it is possible to define a telescopic type system internal
+to a telescopic type system, just using the algebraic structure of a telescopic type system, and actually
 defining an internal category was the hardest part of that definition.
@@
 \begin{equation*}
-f_\ast \defeq S_f\circ W_A/B : \cat{F}/\ctxext{\Gamma}{B}\to\cat{F}/\ctxext{\Gamma}{A}.
-\end{equation*}
-Furthermore, for $g\in\thom{B}{C}$ we may consider $W_B(C)\in\cat{F}/\ctxext{\Gamma}{B}$ as a morphism from $W_B(C)$ to $\catid{\ctxext{\Gamma}{B}}$. Therefore, $f_\ast$ acts on the terms of
-$W_B(C)$, so it makes sense to define $g\circ f\defeq f_\ast(g)$. In fact, it makes
-sense to let $(\blank)\circ f$ be the infix notation for $f_\ast$. 
+f^\ast \defeq S_f\circ W_A/B : \cat{F}/\ctxext{\Gamma}{B}\to\cat{F}/\ctxext{\Gamma}{A}.
+\end{equation*}
+Furthermore, for $g\in\thom{B}{C}$ we may consider $W_B(C)\in\cat{F}/\ctxext{\Gamma}{B}$ as a morphism from $W_B(C)$ to $\catid{\ctxext{\Gamma}{B}}$. Therefore, $f^\ast$ acts on the terms of
+$W_B(C)$, so it makes sense to define $g\circ f\defeq f^\ast(g)$. In fact, it makes
+sense to let $(\blank)\circ f$ be the infix notation for $f^\ast$. 
 \end{defn}
@@
 Note that \autoref{f:SidfWff_id} of \autoref{defn:esystems} asserts precisely
-that ${(\tfid{A})}_\ast\jdeq \catid{\cat{F}/\ctxext{\Gamma}{A}}$ for any $A\in\cat{F}/\Gamma$. In
+that ${(\tfid{A})}^\ast\jdeq \catid{\cat{F}/\ctxext{\Gamma}{A}}$ for any $A\in\cat{F}/\Gamma$. In
 particular, it follows that $g\circ\tfid{A}\jdeq g$ for any $g\in\thom{A}{B}$
@@
 \begin{lem}\label{lem:compcomp}
-For any $f\in\thom{A}{B}$ and $g\in\thom{B}{C}$ we have $f_\ast\circ g_\ast\jdeq (g\circ f)_\ast$. 
+For any $f\in\thom{A}{B}$ and $g\in\thom{B}{C}$ we have $f^\ast\circ g^\ast\jdeq (g\circ f)^\ast$. 
 \end{lem}
@@
 \begin{align*}
-f_\ast\circ g_\ast & \jdeq S_f\circ (W_A/B) \circ S_g \circ (W_B/C)\\
+f^\ast\circ g^\ast & \jdeq S_f\circ (W_A/B) \circ S_g \circ (W_B/C)\\
 & \jdeq S_f\circ S_{W_A(g)}\circ(W_A/\ctxext{B}{W_B(C)})\circ (W_B/C) \\
@@
 & \jdeq S_{S_f(W_A(g))}\circ W_A/C \\
-& \jdeq (g\circ f)_\ast.\qedhere
+& \jdeq (g\circ f)^\ast.\qedhere
 \end{align*}
@@
 \begin{thm}
-For each object $\Gamma$ of an E-system $\mathbb{E}$, we have a pre-category
+For each object $\Gamma$ of a telescopic type system $\mathbb{E}$, we have a pre-category
 $\mathbb{C}_I(\mathbb{E},\Gamma)$ of internal morphisms of $\mathbb{E}$,
@@
 The fact that composition is associative is a direct corollary of
-\autoref{lem:compcomp}. The axiom $(\tfid{A})_\ast\jdeq\catid{\ctxext{\Gamma}{A}}$
+\autoref{lem:compcomp}. The axiom $(\tfid{A})^\ast\jdeq\catid{\ctxext{\Gamma}{A}}$
 implies that the identity morphisms satisfy the right identity law. It remains
@@
 homomorphism, it follows immediately that any E-homomorphism preserves any
-$f_\ast$, i.e.~for any $f\in\thom{A}{B}$ in $\mathcal{E}$ and any E-homomorphism
+$f^\ast$, i.e.~for any $f\in\thom{A}{B}$ in $\mathcal{E}$ and any E-homomorphism
 $F:\mathcal{E}\to\mathcal{D}$, the diagram
@@
   \\
-\mathcal{E}/\ctxext{\Gamma}{B} \arrow[u,"f_\ast"] \arrow[r,swap,"F/\ctxext{\Gamma}{B}"] & \mathcal{D}/F(\ctxext{\Gamma}{B}) \arrow[u,swap,"F(f)_\ast"]
+\mathcal{E}/\ctxext{\Gamma}{B} \arrow[u,"f^\ast"] \arrow[r,swap,"F/\ctxext{\Gamma}{B}"] & \mathcal{D}/F(\ctxext{\Gamma}{B}) \arrow[u,swap,"F(f)^\ast"]
 \end{tikzcd}
@@
 \begin{equation*}
-f_\ast\circ W_B\jdeq W_A.
+f^\ast\circ W_B\jdeq W_A.
 \end{equation*}
@@
 \begin{equation*}
-f_\ast\circ W_B\jdeq S_f\circ W_A/B\circ W_B\jdeq S_f\circ W_{W_A(B)}\circ W_A
+f^\ast\circ W_B\jdeq S_f\circ W_A/B\circ W_B\jdeq S_f\circ W_{W_A(B)}\circ W_A
 \jdeq W_A.\qedhere
@@
 \begin{equation*}
-(\cprojfstf{A}{P})_\ast \jdeq W_P.
+(\cprojfstf{A}{P})^\ast \jdeq W_P.
 \end{equation*}
@@
 \begin{align*}
-(\cprojfstf{A}{P})_\ast & \jdeq S_{\cprojfstf{A}{P}}\circ W_{\ctxext{A}{P}} \\
+(\cprojfstf{A}{P})^\ast & \jdeq S_{\cprojfstf{A}{P}}\circ W_{\ctxext{A}{P}} \\
 & \jdeq S_{W_P(\tfid{A})}\circ W_P\circ W_A \\
@@
 \begin{equation*}
-F_\circledast\defeq F_\ast\circ (f_\ast/Q):
+F^\circledast\defeq F^\ast\circ (f^\ast/Q):
 \mathbb{E}/\ctxext{{\Gamma}{B}}{Q}\to\mathbb{E}/\ctxext{{\Gamma}{A}}{P}.
 \end{equation*}
-The infix notation of $F_\circledast$ is taken to be $\jfcomp{\blank}{\blank}{\blank}{F}{\blank}$.
+The infix notation of $F^\circledast$ is taken to be $\jfcomp{\blank}{\blank}{\blank}{F}{\blank}$.
 \end{defn}
@@
 \begin{equation*}
-F_\circledast\jdeq (\jvcomp{P}{f}{F})_\ast.
+F^\circledast\jdeq (\jvcomp{P}{f}{F})^\ast.
 \end{equation*}
@@
 \begin{align*}
-F_\ast\circ (f_\ast/Q)
+F^\ast\circ (f^\ast/Q)
   & \jdeq
@@
   & \jdeq
-(\jvcomp{P}{f}{F})_\ast.\qedhere
+(\jvcomp{P}{f}{F})^\ast.\qedhere
 \end{align*}
@@
 in context $\Gamma$. 
-Then $F_\circledast\circ G_\circledast\jdeq (\jfcomp{A}{f}{P}{F}{G})_\circledast$.
+Then $F^\circledast\circ G^\circledast\jdeq (\jfcomp{A}{f}{P}{F}{G})^\circledast$.
 In other words the composition $\jfcomp{}{}{}{\blank}{\blank}$ is associative.
@@
 \begin{align*}
-F_\circledast\circ G_\circledast
-  & \jdeq
-(\jvcomp{P}{f}{F})_\ast\circ(\jvcomp{Q}{g}{G})_\ast 
+F^\circledast\circ G^\circledast
+  & \jdeq
+(\jvcomp{P}{f}{F})^\ast\circ(\jvcomp{Q}{g}{G})^\ast 
   \tag{By \autoref{lem:three-composition}}\\
   & \jdeq
-(\jcomp{\ctxext{A}{P}}{\jvcomp{P}{f}{F}}{\jvcomp{Q}{g}{G}})_\ast 
+(\jcomp{\ctxext{A}{P}}{\jvcomp{P}{f}{F}}{\jvcomp{Q}{g}{G}})^\ast 
   \tag{By \autoref{lem:compcomp}}\\
   & \jdeq
-(\jvcomp{}{\jcomp{A}{f}{g}}{\jfcomp{A}{f}{P}{F}{G}})_\ast 
+(\jvcomp{}{\jcomp{A}{f}{g}}{\jfcomp{A}{f}{P}{F}{G}})^\ast 
   \tag{By \autoref{thm:interchange}} \\
   & \jdeq
-(\jfcomp{A}{f}{P}{F}{G})_\circledast.
+(\jfcomp{A}{f}{P}{F}{G})^\circledast.
   \tag{By \autoref{lem:three-composition}}
@@
 
-\subsection{E-systems are E'-systems}
+\subsection{Telescopic type systems are E'-systems}
 
@@
 \begin{thm}
-E-systems are the same things as E'-systems.
+Telescopic type systems are the same things as E'-systems.
 \end{thm}
@@
   \cong
-T((\jcomp{X}{g}{f})_\ast(Q))
+T((\jcomp{X}{g}{f})^\ast(Q))
   \cong
@@
 
-We have described functors from E'-systems to E-systems and vice versa. It is
+We have described functors from E'-systems to telescopic type systems and vice versa. It is
 a straightforward check that these are mutual inverses.
@@
 \begin{equation*}
-\thomd{f_0}{P}{Q} \defeq \thom{P}{{f_0}_\ast(Q)}
+\thomd{f_0}{P}{Q} \defeq \thom{P}{{f_0}^\ast(Q)}
 \end{equation*}
@@
 \begin{rmk}
-When $\mathcal{E}$ is an E-system with empty context, then we have the category
+When $\mathcal{E}$ is a telescopic type system with empty context, then we have the category
 structure on the contexts given by $\thom{\Gamma}{\Delta}= T(W_\Gamma(\Delta))$.
@@
 
-\section{Indexed E-systems}
+\section{Indexed telescopic type systems}
 \begin{defn}
@@
 
-\section{Weak E-systems}
+\section{Weak telescopic type systems}
 
@@
 to be a weakening/projection/substitution homomorphism?
-\item Can they all be strictified? Are there more examples of weak E-systems?
+\item Can they all be strictified? Are there more examples of weak telescopic type systems?
 \end{itemize}

### 2016-02-12 introduction.tex
telescopic CWFs
asterisk up

@@
 \section{Introduction --- Mathematical languages as designed and maintained languages}
+
+\begin{comment}
 Mathematical research is aided by the development of languages that help 
@@
 An essentially algebraic theory is an equational theory of partial operations
-which may be ordered, and of which the domain of each operation is declared by 
+which may be given the structure of a presheaf over usual ordered set of the
+natural numbers, and of which the domain of each operation is declared by 
 equations on the preceeding operations. The concept of essentially algebraic
@@
 operations. 
+\end{comment}
 
-\subsection*{Setting up the essentially algebraic theory of E-systems}
-The theory of E-systems we present in the present article is an iteration in the
+The theory of telescopic type systems is a theory of type dependency in which we have a notion of family extension.
+They are called such because if we try to interpret a category with families as a telescopic type systems, one of the things we have to do is to consider lists of dependent types as our notion of families.
+These are the objects De Bruijn called telescopes.
+On the other hand, the theory of telescopic type systems is an (essentially) algebraic theory of the notion of structure.
+A family over a context $\Gamma$ in a telescopic type system is a structure in context $\Gamma$.
+Therefore I am also considering the name `structure algebras'.
+Types are a defined notion in telescopic type systems.
+They are the atomic families (or structures), i.e.~the ones which do not decompose as the extension of two non-trivial structures.
+
+\subsection*{Setting up the essentially algebraic theory of telescopic type systems}
+The theory of telescopic type systems we present in the present article is an iteration in the
 development of theory of type
@@
 
-An E-system is therefore an algebra consisting of three sorts: $C$ for contexts,
+a telescopic type system is therefore an algebra consisting of three sorts: $C$ for contexts,
 $F$ for families, and $T$ for terms. There is a map 
@@
 family; and there is a map $\ebd:T\to F$ which determines for each term, of
-which family it is a term. Thus, the backbones of an E-system can be represented
+which family it is a term. Thus, the backbones of a telescopic type system can be represented
 as
@@
 
-One may recognize in these ingredients of E-systems a category with a
+One may recognize in these ingredients of telescopic type systems a category with a
 terminal object, and a set $T(f)$ of terms for every morphism $f$ (satisfying
@@
 
-Thus, an E-system $\mathbf{E}$ will have an underlying category, which may be called the
+Thus, a telescopic type system $\mathbf{E}$ will have an underlying category, which may be called the
 category of families of $\mathbf{E}$. The structure of type dependency will by
-built on top of this category. This structure will consist of an E-homomorphism
+built on top of this category. This structure will consist of a telescopic homomorphism
 $W_A$ called \emph{weakening by $A$}, for every family $A$ in context $\Gamma$,
-an E-homomorphism $S_x$ called \emph{substitution by $x$}, for every term
+a telescopic homomorphism $S_x$ called \emph{substitution by $x$}, for every term
 $x$ of $A$ in context $\Gamma$, and units $\idtm{A}\in T(W_A(A))$ for every
@@
 
-\subsection*{The pre-categorical nature of E-systems}
-Although we start our definition of E-systems with a category with term structure,
-we do not set out to have a categorical notion of E-systems, in the sense that the structure of
-being an E-system can be transported along equivalences of categories. The goal
+The category of B-systems will be shown to be equivalent to the category of
+stratified telescopic type systems. However, we do not arrive at the fact that the functor
+$\mathcal{E}:\mathbf{Bsys}\to\mathbf{Tel}$ is full without imposing an additional
+condition on the E-homomorphisms. 
+To ensure that the category of stratified telescopic type systems with stratified 
+E-homomorphisms between them is a full subcategory of the category 
+$\mathbf{Tel}$, we will require that the underlying functor of each E-homomorphism
+lifts factorizations. It is
+important to note that this condition is also algebraic, so that the category
+of telescopic type systems remains algebraic.
+
+Recall that for any morphism $f:X\to Y$ in a category $\cat{C}$ there is a category
+$\mathbf{fact}(f)$ of factorizations of $f$. Also, any functor $F:\cat{C}\to\cat{D}$
+determines a functor $F_{\mathbf{fact}(f)}:\mathbf{fact}(f)\to\mathbf{fact}(H(f))$
+for any morphism $f$ of $\cat{C}$.
+
+\begin{defn}
+A functor $F:\cat{C}\to\cat{D}$ is said to
+\define{lift factorizations} if for any $f\in\cat{C}/X$, the functor
+$F_{\mathbf{fact}(f)}:\mathbf{fact}(f)\to\mathbf{fact}(F(f))$ is an isomorphism
+of categories. 
+\end{defn}
+
+\begin{comment}
+\begin{rmk}
+We may choose the property of lifting factorizations
+to involve either an equivalence or an isomorphism of categories. The
+isomorphism-version, would say that for every factorization $h'\circ g'$
+of $F(f)$ there is a unique factorization $f\jdeq h\circ g$ in $\cat{C}$ such
+that $F(g)\jdeq g'$ and $F(h)\jdeq h'$. In the version with equivalences,
+the uniqueness is replaced by uniqueness up to isomorphism, and the equalites
+are replaced by an isomorphism in $\mathbf{fact}(F(f))$.
+
+It is easier to state the isomorphism version of lifting factorizations with
+inference rules and it might be easier to explain this condition on type theoretical
+grounds, although the equivalence version has the advantage of being
+categorical (i.e.~invariant under equivalence of categories).
+
+In the present
+context, it doesn't matter very much which one we pick: the categories in which
+we're interested are all posets.
+\end{rmk}
+\end{comment}
+
+\begin{eg}
+For any slice category $\cat{C}/X$, the forgetful functor $\cat{C}/X\to\cat{C}$
+lifts factorizations.
+\end{eg}
+
+Convention: A functor in this paper is a functor which lifts factorizations.
+
+\subsection*{The pre-categorical nature of telescopic type systems}
+Although we start our definition of telescopic type systems with a category with term structure,
+we do not set out to have a categorical notion of telescopic type systems, in the sense that the structure of
+being a telescopic type system can be transported along equivalences of categories. The goal
 in this paper is to describe the algebraic structures belonging to an essentially
@@
 rather than categorical. The fact that we've encountered a category at the outset
-of our definition of E-systems is in a sense coincidence, rather than born out
-of a desire to do type theory categorically and we would consider E-systems to
+of our definition of telescopic type systems is in a sense coincidence, rather than born out
+of a desire to do type theory categorically and we would consider telescopic type systems to
 be pre-categorical models of type theory rather than categorical models of type
@@
 $p_A:\ctxext{\Gamma}{A}\to\Gamma$ in the category of contexts. The evident
-functor from categories with families to E-systems will forget that the contexts
+functor from categories with families to telescopic type systems will forget that the contexts
 form a category, and therefore it will not be faithful.
@@
 category of context, and for each context $\Gamma$ a category of types.
+However, it is not quite clear what kind of theory the theory of comprehension
+categories is, which compromises the study of the notion itself.
 \end{itemize}
@@
 to be empty, the underlying category of context remains. We shall see that
-E-systems are like contextual categories in this respect: only a singleton can
-be an E-system with no types. This complicates a thorough comparison between
-categories with families and comprehension categories on one side, and E-systems
+telescopic type systems are like contextual categories in this respect: only a singleton can
+be a telescopic type system with no types. This complicates a thorough comparison between
+categories with families and comprehension categories on one side, and telescopic type systems
 on the other. We shall construct functors from the categories of categories with
@@
 \subsection*{Overview and conventions}
-In \autoref{sec:esys_defn}, we give the categorical definition of E-system in 
+In \autoref{sec:esys_defn}, we give the categorical definition of telescopic type systems in 
 several stages. First we define weakening systems, then projection systems and
 substitution systems, and finally we bundle them together into the definition
-of E-systems. The structure of the definition of weakening systems is 
+of telescopic type systems. The structure of the definition of weakening systems is 
 prototypcial: we first define the notion of pre-weakening systems, which are
@@
 of weakening systems. We go through this process for each of the definitions
-of weakening systems, projection systems, substitution systems and E-systems.
+of weakening systems, projection systems, substitution systems and telescopic type systems.
 
@@
 $A\in\cat{F}/\Gamma$. The goal of this section is to show that 
-E'-systems and E-systems are different axiomatizations of the same structures,
-and apart from that the goal is to show some equational reasoning in E-systems.
+E'-systems and telescopic type systems are different axiomatizations of the same structures,
+and apart from that the goal is to show some equational reasoning in telescopic type systems.
 The properties proven in this section also serve to provide some of the 
@@
 Although we have introduced a notion of E-homomorphism, a better behaved category
-of E-systems is the one where E-homomorphisms lift factorizations. The condition
+of telescopic type systems is the one where E-homomorphisms lift factorizations. The condition
 that a functor lifts factorizations is also algebraic, so we still have an
-essentially algebraic theory of E-systems if we require this.
-In \autoref{sec:esys_compared}, where we introduce the category of E-systems
+essentially algebraic theory of telescopic type systems if we require this.
+In \autoref{sec:esys_compared}, where we introduce the category of telescopic type systems
 with E-homomorphisms lifting factorizations, we construct an embedding from Voevodsky's
-B-systems into E-systems and prove that its image is the (full) subcategory
-of stratified E-systems.
+B-systems into telescopic type systems and prove that its image is the (full) subcategory
+of stratified telescopic type systems.
 
@@
 and we wish to reserve
-the more convenient symbol $=$ for the internal equality of E-systems, i.e.~for
+the more convenient symbol $=$ for the internal equality of telescopic type systems, i.e.~for
 the identification types, as in \cite{TheBook}. 
@@
 When we think of a category $\cat{F}$ as a category of fibrations or families
-over a context -- which will be the outset of the definition of E-systems --
+over a context -- which will be the outset of the definition of telescopic type systems --
 it will be convenient to denote the objects of the $\cat{F}$ with capital
@@
 large families of operations (a weakening operation $W_A$ for each morphism $A$,
-for instance), E-systems can actually be defined with finitely many
-operations. This becomes clearer when one defines an E-system in an arbitrary
+for instance), telescopic type systems can actually be defined with finitely many
+operations. This becomes clearer when one defines a telescopic type system in an arbitrary
 category with finite limits.
@@
 Steve Awodey for sharing many insights and useful advice which lead to the
-definition of an E-system in arbitrary categories with finite limits;
-and Vladimir Voevodsky for inviting me to present E-systems at the Homotopy Type
+definition of a telescopic type system in arbitrary categories with finite limits;
+and Vladimir Voevodsky for inviting me to present telescopic type systems at the Homotopy Type
 Theory workshop in Oxford (November 2014) and to visit the Institute for Advanced 

### 2016-05-05 e-systems.tex
changing telescopic type system to context algebra

--- 2015-11-26 e-systems.tex
@@
-\section{E-systems}\label{sec:esys_defn}
+\section{Context algebras}\label{sec:esys_defn}
 
@@
 
-\subsection{The definition of E-systems}
+\subsection{The definition of context algebras}
 
-Although it is now possible to state the definition of an E-system without
-any difficulty, it shall be convenient later to also have a category of
-pre-E-systems at our disposal. Pre-E-systems shall by default be non-unital,
-i.e.~they do not come with a pre-projection system, while E-systems shall
-always be assumed to be unital. In the present context, we shall not need
-unital pre-E-systems or non-unital E-systems.
+We can now give the definition of context algebras.
 
 \begin{defn}
-A \define{pre-E-system} is a pre-category $\cat{F}$ with term structure
+A \define{pre-context algebra} is a pre-category $\cat{F}$ with term structure
 with a pre-weakening structure $W$, a pre-substitution structure $S$ and a terminal
-object. A \define{unital pre-E-system}, or simply \define{pre-E-system}, is a non-unital pre-E-system
-which also has the structure of a pre-projection system. 
+object. A \define{unital pre-context algebra}, or simply \define{pre-context algebra}, is a non-unital pre-context algebra which also has the structure of a pre-projection system. 
 \end{defn}
@@
 \begin{defn}
-A \define{pre-E-homomorphism} from $\mathbb{E}\to\mathbb{D}$ is a functor
+A \define{pre-context homomorphism} from $\mathbb{E}\to\mathbb{D}$ is a functor
 $H:\mathbb{C}_\mathbb{E}\to\mathbb{C}_\mathbb{D}$ between the underlying categories
@@
 \label{defn:esystems}
-An \define{E-system} is a unital pre-E-system in which 
+A \define{context algebra} is a unital pre-context algebra in which 
 \begin{enumerate}
@@
 \end{enumerate}
-An \define{E-homomorphism} $H:\mathbb{E}\to\mathbb{D}$ is a functor from
+An \define{context homomorphism} $H:\mathbb{E}\to\mathbb{D}$ is a functor from
 $\cat{F}_{\mathbb{E}}$ to $\cat{F}_{\mathbb{D}}$ which is both a projection
@@
 \begin{cor}
-For any object $\Gamma$ of an E-system $\cat{F}$, the slice pre-category $\cat{F}/\Gamma$
-is an E-system.
+For any object $\Gamma$ of a context algebra $\cat{F}$, the slice pre-category $\cat{F}/\Gamma$
+is a context algebra.
 \end{cor}
+
+\begin{defn}
+We write $\mathbf{Ctx}$ for the category of context algebras where the underlying
+functor of each E-homomorphism lifts factorizations.
+\end{defn}
 

### 2016-05-05 introduction.tex
changing telescopic type system to context algebra

@@
-\section{Introduction --- Mathematical languages as designed and maintained languages}
+\section{Introduction}
 
@@
 
-The theory of telescopic type systems is a theory of type dependency in which we have a notion of family extension.
-They are called such because if we try to interpret a category with families as a telescopic type systems, one of the things we have to do is to consider lists of dependent types as our notion of families.
+\begin{comment}
+The theory of context algebras is a theory of type dependency in which we have a notion of family extension.
+They are called such because if we try to interpret a category with families as a context algebras, one of the things we have to do is to consider lists of dependent types as our notion of families.
 These are the objects De Bruijn called telescopes.
-On the other hand, the theory of telescopic type systems is an (essentially) algebraic theory of the notion of structure.
-A family over a context $\Gamma$ in a telescopic type system is a structure in context $\Gamma$.
+On the other hand, the theory of context algebras is an (essentially) algebraic theory of the notion of structure.
+A family over a context $\Gamma$ in a context algebra is a structure in context $\Gamma$.
 Therefore I am also considering the name `structure algebras'.
-Types are a defined notion in telescopic type systems.
+Types are a defined notion in context algebras.
 They are the atomic families (or structures), i.e.~the ones which do not decompose as the extension of two non-trivial structures.
+\end{comment}
 
-\subsection*{Setting up the essentially algebraic theory of telescopic type systems}
-The theory of telescopic type systems we present in the present article is an iteration in the
-development of theory of type
-dependency. The goal has been to describe an essentially algebraic theory of
+\subsection*{Setting up the essentially algebraic theory of context algebras}
+The first goal of this paper is to describe an essentially algebraic theory of
 type dependency that is in close correspondence to the type theory of
 \cite{TheBook}, while at the same time it is as economic as possible in the
-demand of external resources (such as the natural numbers). Instead of explaning
+demand of requirements of the meta-theory (such as the availability of a type of natural numbers). Instead of explaning
 what a dependent type is via an infinite hierarchy of universes $\UU_0$,
-$\UU_1$, \ldots, we set out to describe type families using only the three sorts
-that typically occur in type theoretical expressions: contexts, families and 
-terms. Although an infinite hierarchy of universes allows one to consider everything
-to be a term (because every type is a term of some specified universe), and 
+$\UU_1$, \ldots, we describe type dependency using only the three sorts
+that typically occur in type theoretical expressions: contexts, dependent contexts and 
+terms. Although an infinite hierarchy of universes allows one to reduce everything
+to a term (because every type is a term of some specified universe), and 
 therefore use the literally same mechanisms for handling type expressions as for
@@
 
-a telescopic type system is therefore an algebra consisting of three sorts: $C$ for contexts,
+a context algebra is therefore an algebra consisting of three sorts: $C$ for contexts,
 $F$ for families, and $T$ for terms. There is a map 
@@
 family; and there is a map $\ebd:T\to F$ which determines for each term, of
-which family it is a term. Thus, the backbones of a telescopic type system can be represented
+which family it is a term. Thus, the backbones of a context algebra can be represented
 as
@@
 
-One may recognize in these ingredients of telescopic type systems a category with a
+One may recognize in these ingredients of context algebras a category with a
 terminal object, and a set $T(f)$ of terms for every morphism $f$ (satisfying
@@
 
-Thus, a telescopic type system $\mathbf{E}$ will have an underlying category, which may be called the
+Thus, a context algebra $\mathbf{E}$ will have an underlying category, which may be called the
 category of families of $\mathbf{E}$. The structure of type dependency will by
@@
 The category of B-systems will be shown to be equivalent to the category of
-stratified telescopic type systems. However, we do not arrive at the fact that the functor
-$\mathcal{E}:\mathbf{Bsys}\to\mathbf{Tel}$ is full without imposing an additional
+stratified context algebras. However, we do not arrive at the fact that the functor
+$\mathcal{E}:\mathbf{Bsys}\to\mathbf{Ctx}$ is full without imposing an additional
 condition on the E-homomorphisms. 
-To ensure that the category of stratified telescopic type systems with stratified 
+To ensure that the category of stratified context algebras with stratified 
 E-homomorphisms between them is a full subcategory of the category 
-$\mathbf{Tel}$, we will require that the underlying functor of each E-homomorphism
+$\mathbf{Ctx}$, we will require that the underlying functor of each E-homomorphism
 lifts factorizations. It is
 important to note that this condition is also algebraic, so that the category
-of telescopic type systems remains algebraic.
+of context algebras remains algebraic.
 
@@
 
-\subsection*{The pre-categorical nature of telescopic type systems}
-Although we start our definition of telescopic type systems with a category with term structure,
-we do not set out to have a categorical notion of telescopic type systems, in the sense that the structure of
-being a telescopic type system can be transported along equivalences of categories. The goal
+\subsection*{The pre-categorical nature of context algebras}
+Although we start our definition of context algebras with a category with term structure,
+we do not set out to have a categorical notion of context algebras, in the sense that the structure of
+being a context algebra can be transported along equivalences of categories. The goal
 in this paper is to describe the algebraic structures belonging to an essentially
@@
 rather than categorical. The fact that we've encountered a category at the outset
-of our definition of telescopic type systems is in a sense coincidence, rather than born out
-of a desire to do type theory categorically and we would consider telescopic type systems to
+of our definition of context algebras is in a sense coincidence, rather than born out
+of a desire to do type theory categorically and we would consider context algebras to
 be pre-categorical models of type theory rather than categorical models of type
@@
 $p_A:\ctxext{\Gamma}{A}\to\Gamma$ in the category of contexts. The evident
-functor from categories with families to telescopic type systems will forget that the contexts
+functor from categories with families to context algebras will forget that the contexts
 form a category, and therefore it will not be faithful.
@@
 to be empty, the underlying category of context remains. We shall see that
-telescopic type systems are like contextual categories in this respect: only a singleton can
-be a telescopic type system with no types. This complicates a thorough comparison between
-categories with families and comprehension categories on one side, and telescopic type systems
+context algebras are like contextual categories in this respect: only a singleton can
+be a context algebra with no types. This complicates a thorough comparison between
+categories with families and comprehension categories on one side, and context algebras
 on the other. We shall construct functors from the categories of categories with
@@
 \subsection*{Overview and conventions}
-In \autoref{sec:esys_defn}, we give the categorical definition of telescopic type systems in 
+In \autoref{sec:esys_defn}, we give the categorical definition of context algebras in 
 several stages. First we define weakening systems, then projection systems and
 substitution systems, and finally we bundle them together into the definition
-of telescopic type systems. The structure of the definition of weakening systems is 
+of context algebras. The structure of the definition of weakening systems is 
 prototypcial: we first define the notion of pre-weakening systems, which are
@@
 of weakening systems. We go through this process for each of the definitions
-of weakening systems, projection systems, substitution systems and telescopic type systems.
+of weakening systems, projection systems, substitution systems and context algebras.
 
@@
 $A\in\cat{F}/\Gamma$. The goal of this section is to show that 
-E'-systems and telescopic type systems are different axiomatizations of the same structures,
-and apart from that the goal is to show some equational reasoning in telescopic type systems.
+E'-systems and context algebras are different axiomatizations of the same structures,
+and apart from that the goal is to show some equational reasoning in context algebras.
 The properties proven in this section also serve to provide some of the 
@@
 Although we have introduced a notion of E-homomorphism, a better behaved category
-of telescopic type systems is the one where E-homomorphisms lift factorizations. The condition
+of context algebras is the one where E-homomorphisms lift factorizations. The condition
 that a functor lifts factorizations is also algebraic, so we still have an
-essentially algebraic theory of telescopic type systems if we require this.
-In \autoref{sec:esys_compared}, where we introduce the category of telescopic type systems
+essentially algebraic theory of context algebras if we require this.
+In \autoref{sec:esys_compared}, where we introduce the category of context algebras
 with E-homomorphisms lifting factorizations, we construct an embedding from Voevodsky's
-B-systems into telescopic type systems and prove that its image is the (full) subcategory
-of stratified telescopic type systems.
+B-systems into context algebras and prove that its image is the (full) subcategory
+of stratified context algebras.
 
@@
 and we wish to reserve
-the more convenient symbol $=$ for the internal equality of telescopic type systems, i.e.~for
+the more convenient symbol $=$ for the internal equality of context algebras, i.e.~for
 the identification types, as in \cite{TheBook}. 
@@
 When we think of a category $\cat{F}$ as a category of fibrations or families
-over a context -- which will be the outset of the definition of telescopic type systems --
+over a context -- which will be the outset of the definition of context algebras --
 it will be convenient to denote the objects of the $\cat{F}$ with capital
@@
 large families of operations (a weakening operation $W_A$ for each morphism $A$,
-for instance), telescopic type systems can actually be defined with finitely many
-operations. This becomes clearer when one defines a telescopic type system in an arbitrary
+for instance), context algebras can actually be defined with finitely many
+operations. This becomes clearer when one defines a context algebra in an arbitrary
 category with finite limits.
@@
 Steve Awodey for sharing many insights and useful advice which lead to the
-definition of a telescopic type system in arbitrary categories with finite limits;
-and Vladimir Voevodsky for inviting me to present telescopic type systems at the Homotopy Type
+definition of a context algebra in arbitrary categories with finite limits;
+and Vladimir Voevodsky for inviting me to present context algebras at the Homotopy Type
 Theory workshop in Oxford (November 2014) and to visit the Institute for Advanced 

### 2016-05-05 main.tex
asterisk up
changing telescopic type system to context algebra

@@
 
-\newcommand{\thetitle}{An essentially algebraic formulation of dependent type theory}
+\newcommand{\thetitle}{An essentially algebraic theory of dependent contexts}
 \newcommand{\theauthor}{Egbert Rijke}
@@
 \begin{abstract}
-We present an algebraic formulation of dependent type theory with
-three sorts for contexts, families and terms, with finitely many operations
+We present an algebraic theory of dependent contexts with
+three sorts for contexts, dependent contexts and terms, with finitely many operations
 between them and finitely many equality axioms. Algebras for this theory are
-called E-systems. A distinctive feature of the dependent type theory of
-E-systems is that it takes `telescopic families' as the primitive notion of
-families.
+called context algebras. A distinctive feature of the dependent type theory of
+E-systems is that it takes dependent contexts as the notion of `type dependency'
 
 We then compare Voevodsky's B-systems, Dybjer's categories
-with families and Jacobs' comprehension categories with E-systems. There is
-a full and faithful functor from B-systems to E-systems, of which the image is
-the category of stratified E-systems. To have a functor from categories with
-families to E-systems, we need to impose an additonal coherence condition, which
+with families and Jacobs' comprehension categories with context algebras. There is
+a full and faithful functor from B-systems to context algebras, of which the image is
+the category of stratified context algebras. To have a functor from categories with
+families to context algebras, we are using an additonal coherence condition, which
 makes sure that the substitution operation of a category with families extends 
@@
 
-\input{theories}
+%\input{theories}
 
@@
 
+\appendix
+
+\input{syntax}
+
 \bibliographystyle{plain}

### 2016-05-05 compared.tex
change to context algebras

@@
-\section{Telescopic type systems compared to Voevodsky's B-systems}\label{sec:esys_compared}
+\section{Context algebras compared to Voevodsky's B-systems}\label{sec:esys_compared}
 In an effort to formalize the simplicial set model of Martin-L\"of type theory
@@
 
-\subsection{Stratified telescopic type systems}
-
-In this subsection we define the notion of stratified telescopic type system.
-Then we show that the category $\mathbf{Tel}_s$ of stratified telescopic type systems with homomorphisms preserving the stratification, is a full subcategory of $\mathbf{Tel}$.
+\subsection{Stratified context algebras}
+
+In this subsection we define the notion of stratified context algebras.
+Then we show that the category $\mathbf{Tel}_s$ of stratified context algebras with homomorphisms preserving the stratification, is a full subcategory of $\mathbf{Tel}$.
 
@@
 \begin{defn}
-A pre-telescopic type system is said to be \define{stratified} if its underlying category is
+A pre-Context algebras is said to be \define{stratified} if its underlying category is
 stratified and if each $W_A$ and $S_x$ is a stratified functor. The category of
-stratified pre-telescopic type systems with stratified E-homomorphisms between them is denoted by
+stratified pre-Context algebras with stratified E-homomorphisms between them is denoted by
 $\mathbf{Tel}_s$.
@@
 \begin{cor}\label{lem:strat_full}
-The category of stratified pre-telescopic type systems is a full subcategory of the category of
-pre-telescopic type systems.
+The category of stratified pre-Context algebras is a full subcategory of the category of
+pre-Context algebras.
 \end{cor}
 
-\subsection{The embedding of B0-systems into pre-telescopic type systems}
+\subsection{The embedding of B0-systems into pre-Context algebras}
 
@@
 between unital B-systems are unital. The same distinguishment could be made for 
-telescopic type systems by forgetting the projection structure, but we have not been able to construct a
-functor from unital B0-systems to unital pre-telescopic type systems. Therefore, we shall only
+Context algebras by forgetting the projection structure, but we have not been able to construct a
+functor from unital B0-systems to unital pre-Context algebras. Therefore, we shall only
 consider non-unital B0-systems and unital B-systems. In this section we will
 recall the definition of non-unital B0-systems and construct a full and faithful
-functor from them to the category of pre-telescopic type systems.
+functor from them to the category of pre-Context algebras.
 
@@
 it is clear in the definition of pre-B-systems that there are only countably
-many operations. Since we are interested in embedding the B-systems into telescopic type systems,
+many operations. Since we are interested in embedding the B-systems into context algebras,
 we shall organize the definition of B-systems a bit differently, but of course
@@
 Now that we have defined the category of B0systems, we can start defining the
-embedding into pre-telescopic type systems.
+embedding into pre-Context algebras.
 
@@
 Let $\mathbb{B}$ be a B0-system. Then $U(\mathbb{B})$ can be given the structure
-of a pre-telescopic type system. We write $\mathcal{E}(\mathbb{B})$ for the resulting
-pre-telescopic type system. Thus we get a functor $\mathcal{E}:\mathbf{B0sys}\to\mathbf{Tel}_s$.
+of a pre-Context algebras. We write $\mathcal{E}(\mathbb{B})$ for the resulting
+pre-Context algebras. Thus we get a functor $\mathcal{E}:\mathbf{B0sys}\to\mathbf{Tel}_s$.
 \end{defn}
@@
 
-\subsection{B-systems are stratified telescopic type systems}
-There are two properties which hold for pre-telescopic type systems of the form $\mathcal{E}
+\subsection{B-systems are stratified context algebras}
+There are two properties which hold for pre-Context algebras of the form $\mathcal{E}
 (\mathbb{B})$ for some B0-system $\mathbb{B}$, 
 which stand in the way of proving that the category of stratified
-pre-telescopic type systems is equivalent to the category of B0-systems. The first is that
+pre-Context algebras is equivalent to the category of B0-systems. The first is that
 for any $A\in U(\mathbb{B})/\Gamma$ and $P\in U(\mathbb{B})/\ctxext{\Gamma}{A}$,
@@
 with the categorical structure of the underlying category $U(\mathbb{B})$.
-However, to prove the corresponding facts for telescopic type systems, we needed the full
-set of conditions on unital telescopic type systems: to prove that the term structure of
+However, to prove the corresponding facts for context algebras, we needed the full
+set of conditions on unital context algebras: to prove that the term structure of
 $U(\mathbb{B})$ is compatible with composition we needed units, and to prove 
@@
 
-Presumably, we can assume these compatibility conditions on pre-telescopic type systems in order
-to prove the desired equivalence of B0-systems and stratified pre-telescopic type systems. However,
+Presumably, we can assume these compatibility conditions on pre-Context algebras in order
+to prove the desired equivalence of B0-systems and stratified pre-Context algebras. However,
 that seems somewhat unnatural, especially given that they automatically hold
-for telescopic type systems. Thus, for the remainder of this section we set out to show that
+for context algebras. Thus, for the remainder of this section we set out to show that
 the category of unital B-systems is equivalent to the category of stratified 
-unital telescopic type systems.
+unital context algebras.
 
@@
 For any non-unital B-system $\mathbb{B}$, $\mathcal{E}(\mathbb{B})$ is a 
-stratified non-unital telescopic type system. 
+stratified non-unital context algebras. 
 \end{lem}
@@
 projection structure. We thus obtain a functor 
-$\mathcal{E}:\mathbf{Bsys}\to\mathbf{Tel}$ from unital B-systems to telescopic type systems.
+$\mathcal{E}:\mathbf{Bsys}\to\mathbf{Tel}$ from unital B-systems to context algebras.
 \end{defn}
@@
 We first need to extend the functor $\mathcal{E}:\mathbf{B0sys}\to\mathbf{Esys_p}$
-from non-unital B0-systems to non-unital pre-telescopic type systems, to a functor from
-unital B-systems to telescopic type systems.
+from non-unital B0-systems to non-unital pre-Context algebras, to a functor from
+unital B-systems to context algebras.
 
 We need to give $\mathcal{E}(\mathbb{B})$ the structure of a unital
-pre-telescopic type system. We construct by induction on $m\in \mathbb{N}$, a term of type 
+pre-Context algebras. We construct by induction on $m\in \mathbb{N}$, a term of type 
 $T(m,W_{(m,X)}(X))$ with the property that that 
@@
 The functor $\mathcal{E}:\mathbf{Bsys}\to\mathbf{Tel}$ lifts to an equivalence
-to the subcategory $\mathbf{Tel}_s$ of stratified telescopic type systems: i.e.~we have a
+to the subcategory $\mathbf{Tel}_s$ of stratified context algebras: i.e.~we have a
 commuting diagram
@@
 We already have shown that $\mathcal{E}$ factors through the stratified
-telescopic type systems and that it is full and faithful, so it remains
-to show that every stratified telescopic type system is in the image of $\mathcal{E}$. In
-other words, for every stratified telescopic type system $\mathbb{E}$ we need to construct
+Context algebras and that it is full and faithful, so it remains
+to show that every stratified context algebras is in the image of $\mathcal{E}$. In
+other words, for every stratified context algebras $\mathbb{E}$ we need to construct
 a B0-system $\mathbb{B}$ such that $\mathcal{E}(\mathbb{B})\cong\mathbb{E}$.
@@
 respectively. Since these functors are assumed to be stratified, they define
-homomorphisms of B-frameworks. Because a weakening system of a telescopic type system is
+homomorphisms of B-frameworks. Because a weakening system of a context algebras is
 required to be functorial, i.e.~$W_{\ctxext{A}{P}}\jdeq W_P\circ W_A$ and
@@
 
-\section{Telescopic type systems compared to comprehension categories and to categories with families}
+\section{Context algebras compared to comprehension categories and to categories with families}
 \subsection{Telescopic categories with families}
@@
 at all to have a faithful functor from the category of categories
-with families to the category of telescopic type systems.
+with families to the category of context algebras.
 
@@
 not the case that every $X\to 1$ is a composition of projections $p_A$.
-We could only hope that there is a functor from D-sytems to telescopic type systems without
+We could only hope that there is a functor from D-sytems to context algebras without
 terminal objects.
@@
 subcategory of comprehension categories. Thus, for our purpose of comparing
-telescopic type systems with various existing notions of models of type dependency, it suffice
+Context algebras with various existing notions of models of type dependency, it suffice
 to make the comparison with comprehension categories.
@@
 We see that it is not evident at all to turn categories with families into
-telescopic type systems, even if we ignore the issue that categories with families map
+Context algebras, even if we ignore the issue that categories with families map
 naturally into E'-systems without empty context.
-Thus it seems that the type theories of categories with families and of telescopic type systems
+Thus it seems that the type theories of categories with families and of context algebras
 are not directly comparable.
@@
 As a by-product, we get that the full and faithful functor from B-systems to
-telescopic type systems factors through D-systems.
+Context algebras factors through D-systems.
 
@@
 $\Gamma\in\cat{C}$, the category $p^{-1}(\Gamma)$ is the category of internal
-morphisms the telescopic type system $\mathcal{E}/\Gamma$.
+morphisms the context algebras $\mathcal{E}/\Gamma$.
 

### 2016-05-05 internal_homs.tex
change to context algebras

@@
-\section{Elementary properties of telescopic type systems}\label{sec:esys_props}
-
-The theory of telescopic type systems could be seen as a sort of category theory with only
+\section{Elementary properties of context algebras}\label{sec:esys_props}
+
+The theory of context algebras could be seen as a sort of category theory with only
 projections and their sections. We elaborate on this point of view by defining
@@
 in \autoref{sec:E'}. The thread running through this whole section is proving
-that the notions of telescopic type system and E'-system are the same. In \autoref{sec:E'}
-we show directly that any E'-system is a telescopic type system. We end this section by
+that the notions of context algebra and E'-system are the same. In \autoref{sec:E'}
+we show directly that any E'-system is a context algebra. We end this section by
 showing that the converse also holds.
@@
 Without a terminal object in $\cat{F}$, the pre-category $\cat{C}$ does not correspond
-directly to anything in a telescopic type system.
+directly to anything in a context algebra.
 \end{rmk}
@@
 \begin{thm}
-A telescopic type system is the same thing as an E'-system.
+A context algebra is the same thing as an E'-system.
 \end{thm}
@@
 
-It remains to verify the rest of the conditions of telescopic type systems.
+It remains to verify the rest of the conditions of context algebras.
 \begin{itemize}
@@
 \begin{comment}
-We then demonstrate how to define a category internal to a telescopic type system. The main
-point of this lies mainly in demonstrating how the operations of telescopic type systems can
+We then demonstrate how to define a category internal to a context algebra. The main
+point of this lies mainly in demonstrating how the operations of context algebras can
 be used to define a category, because it is no surprise that a category can be
 defined in type theory. Nevertheless, it is noteworthy that we need no further 
-conditions on telescopic type systems.
+conditions on context algebras.
 This is unlike the situation in category theory, where we need
 to assume that the ambient category has finite limits in order to say what an
-internal category is. In fact, it is possible to define a telescopic type system internal
-to a telescopic type system, just using the algebraic structure of a telescopic type system, and actually
+internal category is. In fact, it is possible to define a context algebra internal
+to a context algebra, just using the algebraic structure of a context algebra, and actually
 defining an internal category was the hardest part of that definition.
@@
 \begin{thm}
-For each object $\Gamma$ of a telescopic type system $\mathbb{E}$, we have a pre-category
+For each object $\Gamma$ of a context algebra $\mathbb{E}$, we have a pre-category
 $\mathbb{C}_I(\mathbb{E},\Gamma)$ of internal morphisms of $\mathbb{E}$,
@@
 
-\subsection{Telescopic type systems are E'-systems}
+\subsection{Context algebras are E'-systems}
 
@@
 \begin{thm}
-Telescopic type systems are the same things as E'-systems.
+Context algebras are the same things as E'-systems.
 \end{thm}
@@
 
-We have described functors from E'-systems to telescopic type systems and vice versa. It is
+We have described functors from E'-systems to context algebras and vice versa. It is
 a straightforward check that these are mutual inverses.
@@
 \begin{rmk}
-When $\mathcal{E}$ is a telescopic type system with empty context, then we have the category
+When $\mathcal{E}$ is a context algebra with empty context, then we have the category
 structure on the contexts given by $\thom{\Gamma}{\Delta}= T(W_\Gamma(\Delta))$.
@@
 
-\section{Indexed telescopic type systems}
+\section{Indexed context algebras}
 \begin{defn}
@@
 
-\section{Weak telescopic type systems}
+\section{Weak context algebras}
 
@@
 to be a weakening/projection/substitution homomorphism?
-\item Can they all be strictified? Are there more examples of weak telescopic type systems?
+\item Can they all be strictified? Are there more examples of weak context algebras?
 \end{itemize}

### 2017-07-03 compared.tex
convert terminology

@@
-\section{Context algebras compared to Voevodsky's B-systems}\label{sec:esys_compared}
+\section{E-systems compared to Voevodsky's B-systems}\label{sec:esys_compared}
 In an effort to formalize the simplicial set model of Martin-L\"of type theory
@@
 
-\subsection{Stratified context algebras}
-
-In this subsection we define the notion of stratified context algebras.
-Then we show that the category $\mathbf{Tel}_s$ of stratified context algebras with homomorphisms preserving the stratification, is a full subcategory of $\mathbf{Tel}$.
+\subsection{Stratified E-systems}
+
+In this subsection we define the notion of stratified E-systems.
+Then we show that the category $\mathbf{Esys}_s$ of stratified E-systems with homomorphisms preserving the stratification, is a full subcategory of $\mathbf{Esys}$.
 
@@
 \begin{defn}
-A pre-Context algebras is said to be \define{stratified} if its underlying category is
+A pre-E-systems is said to be \define{stratified} if its underlying category is
 stratified and if each $W_A$ and $S_x$ is a stratified functor. The category of
-stratified pre-Context algebras with stratified E-homomorphisms between them is denoted by
-$\mathbf{Tel}_s$.
+stratified pre-E-systems with stratified E-homomorphisms between them is denoted by
+$\mathbf{Esys}_s$.
 \end{defn}
@@
 \begin{cor}\label{lem:strat_full}
-The category of stratified pre-Context algebras is a full subcategory of the category of
-pre-Context algebras.
+The category of stratified pre-E-systems is a full subcategory of the category of
+pre-E-systems.
 \end{cor}
 
-\subsection{The embedding of B0-systems into pre-Context algebras}
+\subsection{The embedding of B0-systems into pre-E-systems}
 
@@
 between unital B-systems are unital. The same distinguishment could be made for 
-Context algebras by forgetting the projection structure, but we have not been able to construct a
-functor from unital B0-systems to unital pre-Context algebras. Therefore, we shall only
+E-systems by forgetting the projection structure, but we have not been able to construct a
+functor from unital B0-systems to unital pre-E-systems. Therefore, we shall only
 consider non-unital B0-systems and unital B-systems. In this section we will
 recall the definition of non-unital B0-systems and construct a full and faithful
-functor from them to the category of pre-Context algebras.
+functor from them to the category of pre-E-systems.
 
@@
 it is clear in the definition of pre-B-systems that there are only countably
-many operations. Since we are interested in embedding the B-systems into context algebras,
+many operations. Since we are interested in embedding the B-systems into E-systems,
 we shall organize the definition of B-systems a bit differently, but of course
@@
 Now that we have defined the category of B0systems, we can start defining the
-embedding into pre-Context algebras.
+embedding into pre-E-systems.
 
@@
 Let $\mathbb{B}$ be a B0-system. Then $U(\mathbb{B})$ can be given the structure
-of a pre-Context algebras. We write $\mathcal{E}(\mathbb{B})$ for the resulting
-pre-Context algebras. Thus we get a functor $\mathcal{E}:\mathbf{B0sys}\to\mathbf{Tel}_s$.
+of a pre-E-systems. We write $\mathcal{E}(\mathbb{B})$ for the resulting
+pre-E-systems. Thus we get a functor $\mathcal{E}:\mathbf{B0sys}\to\mathbf{Esys}_s$.
 \end{defn}
@@
 
-\subsection{B-systems are stratified context algebras}
-There are two properties which hold for pre-Context algebras of the form $\mathcal{E}
+\subsection{B-systems are stratified E-systems}
+There are two properties which hold for pre-E-systems of the form $\mathcal{E}
 (\mathbb{B})$ for some B0-system $\mathbb{B}$, 
 which stand in the way of proving that the category of stratified
-pre-Context algebras is equivalent to the category of B0-systems. The first is that
+pre-E-systems is equivalent to the category of B0-systems. The first is that
 for any $A\in U(\mathbb{B})/\Gamma$ and $P\in U(\mathbb{B})/\ctxext{\Gamma}{A}$,
@@
 with the categorical structure of the underlying category $U(\mathbb{B})$.
-However, to prove the corresponding facts for context algebras, we needed the full
-set of conditions on unital context algebras: to prove that the term structure of
+However, to prove the corresponding facts for E-systems, we needed the full
+set of conditions on unital E-systems: to prove that the term structure of
 $U(\mathbb{B})$ is compatible with composition we needed units, and to prove 
@@
 
-Presumably, we can assume these compatibility conditions on pre-Context algebras in order
-to prove the desired equivalence of B0-systems and stratified pre-Context algebras. However,
+Presumably, we can assume these compatibility conditions on pre-E-systems in order
+to prove the desired equivalence of B0-systems and stratified pre-E-systems. However,
 that seems somewhat unnatural, especially given that they automatically hold
-for context algebras. Thus, for the remainder of this section we set out to show that
+for E-systems. Thus, for the remainder of this section we set out to show that
 the category of unital B-systems is equivalent to the category of stratified 
-unital context algebras.
+unital E-systems.
 
@@
 For any non-unital B-system $\mathbb{B}$, $\mathcal{E}(\mathbb{B})$ is a 
-stratified non-unital context algebras. 
+stratified non-unital E-systems. 
 \end{lem}
@@
 projection structure. We thus obtain a functor 
-$\mathcal{E}:\mathbf{Bsys}\to\mathbf{Tel}$ from unital B-systems to context algebras.
+$\mathcal{E}:\mathbf{Bsys}\to\mathbf{Esys}$ from unital B-systems to E-systems.
 \end{defn}
@@
 We first need to extend the functor $\mathcal{E}:\mathbf{B0sys}\to\mathbf{Esys_p}$
-from non-unital B0-systems to non-unital pre-Context algebras, to a functor from
-unital B-systems to context algebras.
+from non-unital B0-systems to non-unital pre-E-systems, to a functor from
+unital B-systems to E-systems.
 
 We need to give $\mathcal{E}(\mathbb{B})$ the structure of a unital
-pre-Context algebras. We construct by induction on $m\in \mathbb{N}$, a term of type 
+pre-E-systems. We construct by induction on $m\in \mathbb{N}$, a term of type 
 $T(m,W_{(m,X)}(X))$ with the property that that 
@@
 \begin{thm}
-The functor $\mathcal{E}:\mathbf{Bsys}\to\mathbf{Tel}$ lifts to an equivalence
-to the subcategory $\mathbf{Tel}_s$ of stratified context algebras: i.e.~we have a
+The functor $\mathcal{E}:\mathbf{Bsys}\to\mathbf{Esys}$ lifts to an equivalence
+to the subcategory $\mathbf{Esys}_s$ of stratified E-systems: i.e.~we have a
 commuting diagram
@@
 \begin{tikzcd}
-& \mathbf{Tel}_s \arrow[dr,"\subseteq"]\\
-\mathbf{Bsys} \arrow[ur,"\simeq"] \arrow[rr,swap,"\mathcal{E}"] & & \mathbf{Tel}
+& \mathbf{Esys}_s \arrow[dr,"\subseteq"]\\
+\mathbf{Bsys} \arrow[ur,"\simeq"] \arrow[rr,swap,"\mathcal{E}"] & & \mathbf{Esys}
 \end{tikzcd}
@@
 We already have shown that $\mathcal{E}$ factors through the stratified
-Context algebras and that it is full and faithful, so it remains
-to show that every stratified context algebras is in the image of $\mathcal{E}$. In
-other words, for every stratified context algebras $\mathbb{E}$ we need to construct
+E-systems and that it is full and faithful, so it remains
+to show that every stratified E-systems is in the image of $\mathcal{E}$. In
+other words, for every stratified E-systems $\mathbb{E}$ we need to construct
 a B0-system $\mathbb{B}$ such that $\mathcal{E}(\mathbb{B})\cong\mathbb{E}$.
@@
 This defines the B-framework $B(\mathbb{E})$, and we obtain a functor
-$B:\mathbf{Tel}_s\to\mathbf{B0sys}$. It is immediate that we get isomorphisms
+$B:\mathbf{Esys}_s\to\mathbf{B0sys}$. It is immediate that we get isomorphisms
 $U(B(\mathbb{E}))\cong\cat{F}_{\mathbb{E}}$, natural in $\mathbb{E}$. Also, we
@@
 respectively. Since these functors are assumed to be stratified, they define
-homomorphisms of B-frameworks. Because a weakening system of a context algebras is
+homomorphisms of B-frameworks. Because a weakening system of a E-systems is
 required to be functorial, i.e.~$W_{\ctxext{A}{P}}\jdeq W_P\circ W_A$ and
@@
 
-\section{Context algebras compared to comprehension categories and to categories with families}
+\section{E-systems compared to comprehension categories and to categories with families}
 \subsection{Telescopic categories with families}
@@
 at all to have a faithful functor from the category of categories
-with families to the category of context algebras.
+with families to the category of E-systems.
 
@@
 not the case that every $X\to 1$ is a composition of projections $p_A$.
-We could only hope that there is a functor from D-sytems to context algebras without
+We could only hope that there is a functor from D-sytems to E-systems without
 terminal objects.
@@
 The following two definitions define subclasses of telescopic D-systems and 
-coherent telescopic D-systems, and the latter structure is the same as an E'-system
+coherent telescopic D-systems, and the latter structure is the same as an CE-system
 (without empty context). In \autoref{defn:comp_to_telescopic}, we will show how
@@
 \begin{thm}
-An E'-system without empty context is the same thing as a coherent telescopic D-system.
+An CE-system without empty context is the same thing as a coherent telescopic D-system.
 \end{thm}
@@
 \begin{proof}
-Let $\mathbb{E}$ be an E'-system without empty context. Then we define the
+Let $\mathbb{E}$ be an CE-system without empty context. Then we define the
 presheaf $\mathrm{Fam}$ on $\cat{C}$ on objects by
@@
 and on morphisms by the action on objects of $f^\ast$, for each $f:\Delta\to
-\Gamma$ in $\cat{C}$. By condition (b) and (c) in the definition of E'-systems, it
+\Gamma$ in $\cat{C}$. By condition (b) and (c) in the definition of CE-systems, it
 follows that this defines a contravariant functor.
@@
 This is also functorial because of conditions (b) and (c) in the definition of
-E'-systems. Thus, we have defined a D-system
+CE-systems. Thus, we have defined a D-system
 \begin{equation*}
@@
 subcategory of comprehension categories. Thus, for our purpose of comparing
-Context algebras with various existing notions of models of type dependency, it suffice
+E-systems with various existing notions of models of type dependency, it suffice
 to make the comparison with comprehension categories.
@@
 with the extra structure of `being closed under family extension' in a coherent
-way, are E'-systems without empty context.
+way, are CE-systems without empty context.
 
@@
 this category should possess a terminal object, so we naturally have to retreat
-to E'-systems without empty contex. Even when we make this generalization, 
+to CE-systems without empty contex. Even when we make this generalization, 
 the construction of a canonical functor in which families are represented by 
@@
 from the category $\mathbf{Dsys}$ of coherent D-systems to the category
-$\mathbf{E'sys_{\circ}}$ of E'-systems without empty context.
+$\mathbf{CEsys_{\circ}}$ of CE-systems without empty context.
 We see that it is not evident at all to turn categories with families into
-Context algebras, even if we ignore the issue that categories with families map
-naturally into E'-systems without empty context.
-Thus it seems that the type theories of categories with families and of context algebras
+E-systems, even if we ignore the issue that categories with families map
+naturally into CE-systems without empty context.
+Thus it seems that the type theories of categories with families and of E-systems
 are not directly comparable.
@@
 We end this section by showing that the functor 
-$U:\mathbf{Dsys}\to\mathbf{E'sys_{\circ}}$ restricts to a full and faithful 
+$U:\mathbf{Dsys}\to\mathbf{CEsys_{\circ}}$ restricts to a full and faithful 
 functor on the coherent D-systems which satisfy the condition that every 
@@
 As a by-product, we get that the full and faithful functor from B-systems to
-Context algebras factors through D-systems.
-
-To get from comprehension categories to E'-systems without empty contexts, we
+E-systems factors through D-systems.
+
+To get from comprehension categories to CE-systems without empty contexts, we
 need to turn comprehension categories into telescopic D-systems.
@@
 \item The choice of functors $\Lambda(f)$ and lifts $q(f,A)$ can be extended to the 
-structure of an E'-system without empty context.
+structure of an CE-system without empty context.
 \item For any
@@
 \end{equation*}
-of functors, forming an E'-system without empty context, where $\mathcal{P}(A)
+of functors, forming an CE-system without empty context, where $\mathcal{P}(A)
 \defeq \mathcal{P}(A_1)\circ\cdots\circ\mathcal{P}(A_n)$ for any
@@
 \begin{conj}
-The functor from coherent split comprehension categories to E'-systems without
+The functor from coherent split comprehension categories to CE-systems without
 empty context is full and faithful.
@@
 \begin{proof}
-Let $\mathcal{E}$ be an E'-system. Define $\cat{E}$ to be the full subcategory
+Let $\mathcal{E}$ be an CE-system. Define $\cat{E}$ to be the full subcategory
 of $\cat{C}^\to$ whose objects are the morphisms of $\cat{F}$. Thus, for any
 $\Gamma\in\cat{C}$, the category $p^{-1}(\Gamma)$ is the category of internal
-morphisms the context algebras $\mathcal{E}/\Gamma$.
+morphisms the E-systems $\mathcal{E}/\Gamma$.
 
@@
 \begin{thm}
-The restriction of the functor $U:\mathbf{Dsys}\to\mathbf{E'sys_{\circ}}$ to the
+The restriction of the functor $U:\mathbf{Dsys}\to\mathbf{CEsys_{\circ}}$ to the
 locally well-founded D-systems is full and faithful.
@@
 \begin{thm}
-The image of $U$ is the subcategory of locally stratified E'-systems without
+The image of $U$ is the subcategory of locally stratified CE-systems without
 empty context. 
@@
 $\Lambda(\Gamma)\to (\N,\geq)$, so $U(\mathbb{D})$ is a locally stratified
-E'-system.
-
-Now let $\mathbb{E}$ be a locally stratified E'-system without empty context.
+CE-system.
+
+Now let $\mathbb{E}$ be a locally stratified CE-system without empty context.
 For $\Gamma\in\cat{C}$, we define $\mathrm{Ty}(\Gamma)$ to be the set of

### 2017-07-03 e-systems.tex
convert terminology

--- 2015-11-26 e-systems.tex
@@
 
-Although it is now possible to state the definition of an E-system without
-any difficulty, it shall be convenient later to also have a category of
-pre-E-systems at our disposal. Pre-E-systems shall by default be non-unital,
-i.e.~they do not come with a pre-projection system, while E-systems shall
-always be assumed to be unital. In the present context, we shall not need
-unital pre-E-systems or non-unital E-systems.
+We can now give the definition of E-systems.
 
@@
 with a pre-weakening structure $W$, a pre-substitution structure $S$ and a terminal
-object. A \define{unital pre-E-system}, or simply \define{pre-E-system}, is a non-unital pre-E-system
-which also has the structure of a pre-projection system. 
+object. A \define{unital pre-E-system}, or simply \define{pre-E-system}, is a non-unital pre-E-system which also has the structure of a pre-projection system. 
 \end{defn}
@@
 \label{defn:esystems}
-An \define{E-system} is a unital pre-E-system in which 
+A \define{E-system} is a unital pre-E-system in which 
 \begin{enumerate}
@@
 \begin{cor}
-For any object $\Gamma$ of an E-system $\cat{F}$, the slice pre-category $\cat{F}/\Gamma$
-is an E-system.
+For any object $\Gamma$ of a E-system $\cat{F}$, the slice pre-category $\cat{F}/\Gamma$
+is a E-system.
 \end{cor}
+
+\begin{defn}
+We write $\mathbf{Ctx}$ for the category of E-systems where the underlying
+functor of each E-homomorphism lifts factorizations.
+\end{defn}
 

### 2017-07-03 internal_homs.tex
convert terminology

@@
-\section{Elementary properties of context algebras}\label{sec:esys_props}
-
-The theory of context algebras could be seen as a sort of category theory with only
+\section{Elementary properties of E-systems}\label{sec:esys_props}
+
+The theory of E-systems could be seen as a sort of category theory with only
 projections and their sections. We elaborate on this point of view by defining
@@
 a subpre-category. This, together with the pre-composition operation described
-earlier, gives rise to the auxiliary notion of E'-systems, which we define
-in \autoref{sec:E'}. The thread running through this whole section is proving
-that the notions of context algebra and E'-system are the same. In \autoref{sec:E'}
-we show directly that any E'-system is a context algebra. We end this section by
+earlier, gives rise to the auxiliary notion of CE-systems, which we define
+in \autoref{sec:CE}. The thread running through this whole section is proving
+that the notions of E-system and CE-system are the same. In \autoref{sec:CE}
+we show directly that any CE-system is a E-system. We end this section by
 showing that the converse also holds.
@@
 
-\subsection{E'-systems}\label{sec:E'}
-
-E'-systems can be thought of as strict versions of tribes%
+\subsection{CE-systems}\label{sec:CE}
+
+CE-systems can be thought of as strict versions of tribes%
 \footnote{Tribes are an invention of Joyal, see \url{http://ncatlab.org/homotopytypetheory/files/Joyal.pdf}}.
 
-\begin{defn}\label{defn:E'sys}
-An \define{E'-system} consists of a pre-category $\cat{C}$ and a subpre-category $\cat{F}$ with
+\begin{defn}\label{defn:CEsys}
+An \define{CE-system} consists of a pre-category $\cat{C}$ and a subpre-category $\cat{F}$ with
 the same objects as $\cat{C}$, such that
@@
 \end{equation*}
-\item \label{defn:E'sys:b} For any $A\in\cat{F}/\Gamma$, one has $(\catid{\Gamma})^\ast(A)\jdeq A$ and
+\item \label{defn:CEsys:b} For any $A\in\cat{F}/\Gamma$, one has $(\catid{\Gamma})^\ast(A)\jdeq A$ and
 \begin{equation*}
@@
 \end{equation*}
-\item \label{defn:E'sys:c} For any $f:\Delta\to\Gamma$, $g:\greek{E}\to\Delta$ and $A\in\cat{F}/\Gamma$, one
+\item \label{defn:CEsys:c} For any $f:\Delta\to\Gamma$, $g:\greek{E}\to\Delta$ and $A\in\cat{F}/\Gamma$, one
 has $(f\circ g)^\ast(A)\jdeq g^\ast(f^\ast(A))$ and 
@@
 Without a terminal object in $\cat{F}$, the pre-category $\cat{C}$ does not correspond
-directly to anything in a context algebra.
+directly to anything in a E-system.
 \end{rmk}
@@
 \begin{rmk}
-For E'-systems, we do not require that every pullback of $A\in\cat{F}/\Gamma$ along $f:\Delta\to\Gamma$
+For CE-systems, we do not require that every pullback of $A\in\cat{F}/\Gamma$ along $f:\Delta\to\Gamma$
 is in $\cat{F}/\Delta$, just that there is a choice of such a pullback which is
-in $\cat{F}/\Delta$. Thus, the definition of E'-system is slightly different from
+in $\cat{F}/\Delta$. Thus, the definition of CE-system is slightly different from
 Taylor's notion of display map category. Furthermore, the class of display maps in
@@
 \begin{lem}\label{lem:pb_selfdistributive}
-Let $\mathbb{E}$ be an E'-system. Then $f^\ast$ can be extended to a functor from
+Let $\mathbb{E}$ be an CE-system. Then $f^\ast$ can be extended to a functor from
 $\cat{C}_\cat{F}/\Gamma$ to $\cat{C}_\cat{F}/\Delta$, for each $f:\Delta\to\Gamma$, and moreover the diagram
@@
 \begin{thm}
-A context algebra is the same thing as an E'-system.
+A E-system is the same thing as an CE-system.
 \end{thm}
@@
 
-It remains to verify the rest of the conditions of context algebras.
+It remains to verify the rest of the conditions of E-systems.
 \begin{itemize}
@@
 \begin{comment}
-We then demonstrate how to define a category internal to a context algebra. The main
-point of this lies mainly in demonstrating how the operations of context algebras can
+We then demonstrate how to define a category internal to a E-system. The main
+point of this lies mainly in demonstrating how the operations of E-systems can
 be used to define a category, because it is no surprise that a category can be
 defined in type theory. Nevertheless, it is noteworthy that we need no further 
-conditions on context algebras.
+conditions on E-systems.
 This is unlike the situation in category theory, where we need
 to assume that the ambient category has finite limits in order to say what an
-internal category is. In fact, it is possible to define a context algebra internal
-to a context algebra, just using the algebraic structure of a context algebra, and actually
+internal category is. In fact, it is possible to define a E-system internal
+to a E-system, just using the algebraic structure of a E-system, and actually
 defining an internal category was the hardest part of that definition.
@@
 \begin{thm}
-For each object $\Gamma$ of a context algebra $\mathbb{E}$, we have a pre-category
+For each object $\Gamma$ of a E-system $\mathbb{E}$, we have a pre-category
 $\mathbb{C}_I(\mathbb{E},\Gamma)$ of internal morphisms of $\mathbb{E}$,
@@
 
-\subsection{Context algebras are E'-systems}
+\subsection{E-systems are CE-systems}
 
@@
 \begin{thm}
-Context algebras are the same things as E'-systems.
+E-systems are the same things as CE-systems.
 \end{thm}
@@
 
-We have described functors from E'-systems to context algebras and vice versa. It is
+We have described functors from CE-systems to E-systems and vice versa. It is
 a straightforward check that these are mutual inverses.
@@
 \begin{rmk}
-When $\mathcal{E}$ is a context algebra with empty context, then we have the category
+When $\mathcal{E}$ is a E-system with empty context, then we have the category
 structure on the contexts given by $\thom{\Gamma}{\Delta}= T(W_\Gamma(\Delta))$.
@@
 
-\section{Indexed context algebras}
+\section{Indexed E-systems}
 \begin{defn}
@@
 
-\section{Weak context algebras}
+\section{Weak E-systems}
 
@@
 to be a weakening/projection/substitution homomorphism?
-\item Can they all be strictified? Are there more examples of weak context algebras?
+\item Can they all be strictified? Are there more examples of weak E-systems?
 \end{itemize}

### 2017-07-03 introduction.tex
convert terminology

--- 2015-11-24 introduction.tex
@@
-\section{Introduction --- Mathematical languages as designed and maintained languages}
+\section{Introduction}
+
+\begin{comment}
 Mathematical research is aided by the development of languages that help 
@@
 operations. 
+\end{comment}
+
+\begin{comment}
+The theory of E-systems is a theory of type dependency in which we have a notion of family extension.
+They are called such because if we try to interpret a category with families as a E-systems, one of the things we have to do is to consider lists of dependent types as our notion of families.
+These are the objects De Bruijn called telescopes.
+On the other hand, the theory of E-systems is an (essentially) algebraic theory of the notion of structure.
+A family over a context $\Gamma$ in a E-system is a structure in context $\Gamma$.
+Therefore I am also considering the name `structure algebras'.
+Types are a defined notion in E-systems.
+They are the atomic families (or structures), i.e.~the ones which do not decompose as the extension of two non-trivial structures.
+\end{comment}
 
 \subsection*{Setting up the essentially algebraic theory of E-systems}
-The theory of E-systems we present in the present article is an iteration in the
-development of theory of type
-dependency. The goal has been to describe an essentially algebraic theory of
+The first goal of this paper is to describe an essentially algebraic theory of
 type dependency that is in close correspondence to the type theory of
 \cite{TheBook}, while at the same time it is as economic as possible in the
-demand of external resources (such as the natural numbers). Instead of explaning
+demand of requirements of the meta-theory (such as the availability of a type of natural numbers). Instead of explaning
 what a dependent type is via an infinite hierarchy of universes $\UU_0$,
-$\UU_1$, \ldots, we set out to describe type families using only the three sorts
-that typically occur in type theoretical expressions: contexts, families and 
-terms. Although an infinite hierarchy of universes allows one to consider everything
-to be a term (because every type is a term of some specified universe), and 
+$\UU_1$, \ldots, we describe type dependency using only the three sorts
+that typically occur in type theoretical expressions: contexts, dependent contexts and 
+terms. Although an infinite hierarchy of universes allows one to reduce everything
+to a term (because every type is a term of some specified universe), and 
 therefore use the literally same mechanisms for handling type expressions as for
@@
 
-An E-system is therefore an algebra consisting of three sorts: $C$ for contexts,
+a E-system is therefore an algebra consisting of three sorts: $C$ for contexts,
 $F$ for families, and $T$ for terms. There is a map 
@@
 family; and there is a map $\ebd:T\to F$ which determines for each term, of
-which family it is a term. Thus, the backbones of an E-system can be represented
+which family it is a term. Thus, the backbones of a E-system can be represented
 as
@@
 
-Thus, an E-system $\mathbf{E}$ will have an underlying category, which may be called the
+Thus, a E-system $\mathbf{E}$ will have an underlying category, which may be called the
 category of families of $\mathbf{E}$. The structure of type dependency will by
-built on top of this category. This structure will consist of an E-homomorphism
+built on top of this category. This structure will consist of a telescopic homomorphism
 $W_A$ called \emph{weakening by $A$}, for every family $A$ in context $\Gamma$,
-an E-homomorphism $S_x$ called \emph{substitution by $x$}, for every term
+a telescopic homomorphism $S_x$ called \emph{substitution by $x$}, for every term
 $x$ of $A$ in context $\Gamma$, and units $\idtm{A}\in T(W_A(A))$ for every
@@
 
+The category of B-systems will be shown to be equivalent to the category of
+stratified E-systems. However, we do not arrive at the fact that the functor
+$\mathcal{E}:\mathbf{Bsys}\to\mathbf{Ctx}$ is full without imposing an additional
+condition on the E-homomorphisms. 
+To ensure that the category of stratified E-systems with stratified 
+E-homomorphisms between them is a full subcategory of the category 
+$\mathbf{Ctx}$, we will require that the underlying functor of each E-homomorphism
+lifts factorizations. It is
+important to note that this condition is also algebraic, so that the category
+of E-systems remains algebraic.
+
+Recall that for any morphism $f:X\to Y$ in a category $\cat{C}$ there is a category
+$\mathbf{fact}(f)$ of factorizations of $f$. Also, any functor $F:\cat{C}\to\cat{D}$
+determines a functor $F_{\mathbf{fact}(f)}:\mathbf{fact}(f)\to\mathbf{fact}(H(f))$
+for any morphism $f$ of $\cat{C}$.
+
+\begin{defn}
+A functor $F:\cat{C}\to\cat{D}$ is said to
+\define{lift factorizations} if for any $f\in\cat{C}/X$, the functor
+$F_{\mathbf{fact}(f)}:\mathbf{fact}(f)\to\mathbf{fact}(F(f))$ is an isomorphism
+of categories. 
+\end{defn}
+
+\begin{comment}
+\begin{rmk}
+We may choose the property of lifting factorizations
+to involve either an equivalence or an isomorphism of categories. The
+isomorphism-version, would say that for every factorization $h'\circ g'$
+of $F(f)$ there is a unique factorization $f\jdeq h\circ g$ in $\cat{C}$ such
+that $F(g)\jdeq g'$ and $F(h)\jdeq h'$. In the version with equivalences,
+the uniqueness is replaced by uniqueness up to isomorphism, and the equalites
+are replaced by an isomorphism in $\mathbf{fact}(F(f))$.
+
+It is easier to state the isomorphism version of lifting factorizations with
+inference rules and it might be easier to explain this condition on type theoretical
+grounds, although the equivalence version has the advantage of being
+categorical (i.e.~invariant under equivalence of categories).
+
+In the present
+context, it doesn't matter very much which one we pick: the categories in which
+we're interested are all posets.
+\end{rmk}
+\end{comment}
+
+\begin{eg}
+For any slice category $\cat{C}/X$, the forgetful functor $\cat{C}/X\to\cat{C}$
+lifts factorizations.
+\end{eg}
+
+Convention: A functor in this paper is a functor which lifts factorizations.
+
 \subsection*{The pre-categorical nature of E-systems}
@@
 we do not set out to have a categorical notion of E-systems, in the sense that the structure of
-being an E-system can be transported along equivalences of categories. The goal
+being a E-system can be transported along equivalences of categories. The goal
 in this paper is to describe the algebraic structures belonging to an essentially
@@
 E-systems are like contextual categories in this respect: only a singleton can
-be an E-system with no types. This complicates a thorough comparison between
+be a E-system with no types. This complicates a thorough comparison between
 categories with families and comprehension categories on one side, and E-systems
@@
 \subsection*{Overview and conventions}
-In \autoref{sec:esys_defn}, we give the categorical definition of E-system in 
+In \autoref{sec:esys_defn}, we give the categorical definition of E-systems in 
 several stages. First we define weakening systems, then projection systems and
@@
 for instance), E-systems can actually be defined with finitely many
-operations. This becomes clearer when one defines an E-system in an arbitrary
+operations. This becomes clearer when one defines a E-system in an arbitrary
 category with finite limits.
@@
 Steve Awodey for sharing many insights and useful advice which lead to the
-definition of an E-system in arbitrary categories with finite limits;
+definition of a E-system in arbitrary categories with finite limits;
 and Vladimir Voevodsky for inviting me to present E-systems at the Homotopy Type

### 2017-07-03 main.tex
convert terminology

@@
 between them and finitely many equality axioms. Algebras for this theory are
-called context algebras. A distinctive feature of the dependent type theory of
+called E-systems. A distinctive feature of the dependent type theory of
 E-systems is that it takes dependent contexts as the notion of `type dependency'
@@
 We then compare Voevodsky's B-systems, Dybjer's categories
-with families and Jacobs' comprehension categories with context algebras. There is
-a full and faithful functor from B-systems to context algebras, of which the image is
-the category of stratified context algebras. To have a functor from categories with
-families to context algebras, we are using an additonal coherence condition, which
+with families and Jacobs' comprehension categories with E-systems. There is
+a full and faithful functor from B-systems to E-systems, of which the image is
+the category of stratified E-systems. To have a functor from categories with
+families to E-systems, we are using an additonal coherence condition, which
 makes sure that the substitution operation of a category with families extends 

### 2017-07-18 b-systems.tex
starting to restructure the text

\section{B-systems}

\subsection{B-frames and stratified categories with term structure}

\begin{defn}
A \define{B-frame} is a collection of data of the following form:
\begin{enumerate}
\item for all $n\in\mathbb{N}$ two sets $B_n$ and $\tilde{B}_n$. 
\item for all $n\in\mathbb{N}$ maps of the form
\begin{align*}
\eft[n] & : B_{n+1}\to B_n \\
\ebd[n] & : \tilde{B}_n\to B_n.
\end{align*}
For $m,n\in\mathbb{N}$, we denote the composition $\eft[n]\circ\cdots\circ\eft[n+m]:B_{n+m+1}\to B_n$ by $\eft[n]^m$. 
\item $B_0$ is a singleton $\{\pt\}$.
\end{enumerate} 
A \define{homomorphism $H:\mathbb{B}\to\mathbb{A}$ of B-frames} consists of maps
$H_n:B_n\to A_n$ and $\tilde{H}_n:\tilde{B}_n\to\tilde{A}_n$ such that
\begin{align*}
\eft(H(X)) & \jdeq H(\eft(X)) \\
\ebd(\tilde{H}(x)) & \jdeq H(\ebd(x))
\end{align*}
for any $X\in B_n$ and $x\in\tilde{B}_n$. The category of B-frames is
denoted by $\mathbf{Bfr}$. 
\end{defn}

There is a forgetful functor from B-frames to rooted trees, which is established in the following definition.

\begin{defn}
We define a functor $\mathcal{E}:\mathbf{Bfr}\to\mathbf{Cat}$.
\end{defn}

\begin{constr}
Let $\mathbb{B}$ be a B-frame.
The set of objects of $\mathcal{E}(\mathbb{B})$ is taken to be the set $\bigsqcup_{(n\in\mathbb{N})}B_n$. 
The partial order on $\bigsqcup_{(n\in\mathbb{N})}B_n$ as an inductive family of propositions, by the clauses
\begin{equation*}
(n,X)\leq (n,X) \qquad\text{and}\qquad
((m,\eft(Y))\leq (n,X))\to((m+1,Y)\leq (n,X)).
\end{equation*}
This defines a partial order, hence we obtain a category.

For any homomorphism $H:\mathbb{B}\to\mathbb{A}$, an order-preserving map $\mathcal{E}(H):\mathcal{E}(\mathbb{B})\to\mathcal{E}(\mathbb{A})$ is defined in the obvious way. 
\end{constr}

We introduce \emph{stratified} pre-categories to characterize the image of this forgetful functor.

\begin{defn}
A pre-category $\cat{C}$ with terminal object is said to be \define{stratified} if there exists a 
\emph{stratification functor}
\begin{equation*}
L : \cat{C}\to (\mathbb{N},\geq)
\end{equation*}
such that
\begin{enumerate}
\item $L(X)\jdeq 0$ if and only if $X$ is terminal, and for any $f:X\to Y$ we have
$L(X)\jdeq L(Y)$ if and only if $X\jdeq Y$ and $f\jdeq\catid{X}$. 
\item every morphism $f:X\to Y$ in $\cat{C}$, where $L(X)\jdeq
n+m+1$ and $L(Y)\jdeq n$, has a unique factorization 
\begin{equation*}
\begin{tikzcd}
X \arrow[r,"f_m"] & X_m \arrow[r,"f_{m-1}"] & \cdots \arrow[r,"f_1"] & X_1 \arrow[r,"f_0"] & Y
\end{tikzcd}
\end{equation*}
where $L(X_i)\jdeq n+i$.
\end{enumerate}
A functor $F:\cat{C}\to\cat{D}$ between stratified pre-categories is said to be stratified 
if $L_{\cat{C}}\jdeq L_{\cat{D}}\circ F$.

We define $\mathbf{Cat_s}$ to be the category of stratified pre-categories, and stratified functors between them.
\end{defn}

%\begin{rmk}
%The categorical version of being stratified, is that there exists a factofibration
%$L:\cat{C}\to (N,\geq)$ which preserves the terminal object.
%\end{rmk}

\begin{lem}
Any stratified pre-category is a rooted tree.
\end{lem}

\begin{proof}
Let $\cat{C}$ be a stratified pre-category. The root of the tree is going to be the terminal
object of $\cat{C}$. Note that objects of $\cat{C}$ have no endomorphisms other
than the identity morphisms. Therefore, to show that $\cat{C}$ is a tree,
it suffices to show that for every object $X$ with $L(X)>0$ there is
a unique object $\eft(X)$ with $L(\eft(X))\jdeq L(X)-1$, with a morphism $p_X:X\to \eft(X)$.
The existence follows from the existence of the factorization of $X\to 1$. The 
uniqueness follows from the observation that for any arrow
$f:X\to Y$ with $L(Y)\jdeq L(X)-1$, we can factorize $X\to 1$ through $Y$. Since
factorizations are unique, we will have $Y\jdeq \eft(X)$, and $p_X\jdeq f$. 
\end{proof}

\begin{lem}
The category of small stratified categories is a full subcategory of the category
of small pre-categories with terminal objects and between them functors which lift 
factorizations in the sense of \autoref{defn:lift_factorizations}, and preserve the terminal object.
\end{lem}

\begin{proof}
Let $F:\cat{C}\to\cat{D}$ be a functor between stratified pre-categories which lifts
factorizations and 
which preserves terminal objects. We show by induction that $F$ is stratified, i.e.~that $L_{\mathcal{C}}(X)\jdeq L_{\mathcal{D}}(F(X))$, for any object $X$ of $\mathcal{C}$.

For the base case, we observe that since $F$ preserves the terminal object, $L(X)\jdeq 0$ implies $L(F(X))\jdeq 0$.

For the inductive step, suppose that for $n\in\mathbb{N}$, we have that $L(X)\jdeq n$ implies
$L(F(X))\jdeq n$, for any object $X$ of $\mathcal{C}$, and let $Y$ be such that $L(Y)\jdeq n+1$ with
$p_Y:Y\to\eft(Y)$. Then the poset $\mathbf{fact}(p_Y)$ is isomorphic to the
poset $\mathbf{2}$ with two objects, one of which is smaller than the other. By the
assumption that $F$ lifts factorizations, the poset $\mathbf{fact}(F(p_Y))$
is also isomorphic to $\mathbf{2}$. Thus, the only factorization
of $F(Y)\to\unit$ starts with $F(p_Y):H(Y)\to F(\eft(Y))$, which implies that
$L(F(Y))\jdeq n+1$. 
\end{proof}

\begin{cor}
A pre-category can only be stratified in at most one way. That is, being stratified is a property of pre-categories.
\end{cor}

\begin{thm}
The functor $\mathcal{E}:\mathbf{Bfr}\to\mathbf{Cat}$ factors through the forgetful functor $\mathbf{Cat_s}\to\mathbf{Cat}$.
\end{thm}

\begin{constr}
Let $\mathbb{B}$ be a B-frame. Note that for any $X\in B_{n+1}$ there is a unique $\Gamma\in
B_n$ such that $(n+1,X)\leq (n,\Gamma)$, so $\mathcal{E}(\mathbb{B})$ is stratified.

To show that the functor $\mathcal{E}(H):\mathcal{E}(\mathbb{B})\to \mathcal{E}(\mathbb{A})$ is stratified for
any homomorphism $H:\mathbb{B}\to\mathbb{A}$ of B-frames, note that
by \autoref{lem:strat_full} it suffices to show that each $\mathcal{E}(H)$ 
lifts factorizations.
Let $H:\mathbb{B}\to\mathbb{A}$ be a homomorphism of B-frames, and let
$(m,X):(n+m,X)\leq (n,\Gamma)$. Then the pre-category $\mathbf{fact}_{(m,X)}$ of
factorizations of $(m,X)$ is a finite linear order with
$m+1$ elements, and so is the category $\mathbf{fact}_{(m,H(X))}$. 
\end{constr}

\begin{rmk} 
It is in fact the case that $\mathcal{E}(\mathbb{B})$ is the free pre-category generated by the
graph with $\bigsqcup_{(n\in\mathbb{N})}B_n$ as the set of vertices, and
for each object of the form $(n+1,X)$ an edge to $(n,\eft[n](X))$. 
Essentially all $\mathcal{E}$ does, is forgetting about the term structure 
$\tilde{B}$ of $\mathbb{B}$.
%
%Furthermore, it is useful to note that for any $X\in B_n$, we get an 
%isomorphism $\mathcal{E}(\mathbb{B}/X)\cong  \mathcal{E}(\mathbb{B})/(n,X)$ of rooted trees,
%natural in $X$. Therefore we will usually not distinguish between $\mathcal{E}(\mathbb{B})/X$
%and $\mathcal{E}(\mathbb{B})/(n,X)$. 
\end{rmk}

\subsection{Non-unital pre-B-systems}

For every B-frame $\mathbb{B}$ and any $X\in B_n$, there is a B-frame
$\mathbb{B}/X$ given by
\begin{align*}
(B/X)_{m} & \jdeq \{Y\in B_{n+m}\mid\eft^{m}(Y)\jdeq X\}\\
(\tilde{B}/X)_m & \jdeq \{y\in \tilde{B}_{n+m}\mid\eft^m(\ebd(y))\jdeq X\}.
\end{align*}
Also, for any homomorphism $H:\mathbb{B}\to\mathbb{A}$ of B-frames and any
$X\in B_n$, there is a homomorphism $H/X:\mathbb{B}/X\to\mathbb{A}/H(X)$
defined in the obvious way.

Note that for $X\in B_n$ and $Y\in B_{n+m}$ such that $\eft^m(Y)\jdeq X$, 
we have an isomorphism $(\mathbb{B}/X)/Y\cong B/Y$ of B-frames, constructed in the obvious way, which is natural in the sense that for any homomorphism $H:\mathbb{B}\to\mathbb{A}$ of B-frames, the square
\begin{equation*}
\begin{tikzcd}
(\mathbb{B}/X)/Y \arrow[r,"\cong"] \arrow[d,swap,"(H/X)/Y"] &  \mathbb{B}/Y \arrow[d,"H/Y"] \\
(\mathbb{A}/H(X))/H(Y) \arrow[r,"\cong"] & \mathbb{A}/H(Y)
\end{tikzcd}
\end{equation*}
commutes.

\begin{defn}
A \define{non-unital pre-B-system} $\mathbb{B}$ consists of a B-frame $\mathbb{B}$ and homomorphisms
\begin{align*}
W_{X} & : \mathbb{B}/\eft(X)\to\mathbb{B}/X\\
S_{x} & : \mathbb{B}/\ebd(x)\to\mathbb{B}/\eft(\ebd(x))
\end{align*}
of B-frames, for any $X\in B_{n+1}$ and any $x\in\tilde{B}_{n+1}$.

A \define{non-unital pre-B-homomorphism} $H:\mathbb{B}\to\mathbb{A}$ is a homomorphism of B-frames
for which the diagrams
\begin{equation*}
\begin{tikzcd}[column sep=huge]
\mathbb{B}/X \arrow[r,"H/X"] & \mathbb{A}/H(X) \\
\mathbb{B}/\eft(X) \arrow[u,"W_X"] \arrow[r,swap,"H/\eft(X)"] & \mathbb{A}/\eft(H(X)) \arrow[u,swap,"W_{H(X)}"]
\end{tikzcd}
\end{equation*}
and
\begin{equation*}
\begin{tikzcd}[column sep=huge]
\mathbb{B}/\ebd(x) \arrow[r,"H/\ebd(x)"] \arrow[d,swap,"S_x"] & \mathbb{A}/\ebd(H(X)) \arrow[d,"S_{\tilde{H}(x)}"] \\
\mathbb{B}/\eft(\ebd(x)) \arrow[r,swap,"H/\eft(\ebd(x))"] & \mathbb{A}/\eft(\ebd(H(X)))
\end{tikzcd}
\end{equation*}
of homomorphisms of B-frames commute for every 
$X\in B_n$ and $x\in\tilde{B}_n$. The category of non-unital pre-B-systems is denoted
by $\mathbf{B0sys}$.
\end{defn}

\begin{comment}
\begin{lem}
When $\mathbb{B}$ has the structure of a pre-B-system, then so does each $\mathbb{B}/X$.
\end{lem}

\begin{proof}
Straightforward.
\end{proof}
\end{comment}

\begin{defn}
Let $\mathbb{B}$ be a pre-B-system. Then $U(\mathbb{B})$ can be given the structure
of a pre-E-system. We write $\mathcal{E}(\mathbb{B})$ for the resulting
pre-E-system. Thus we get a functor $\mathcal{E}:\mathbf{B0sys}\to\mathbf{E0sys}_s$.
\end{defn}

\begin{constr}
We begin by \emph{simultaneously} defining the term structure and the pre-substitution
structure on $U(\mathbb{B})$ by induction on the natural numbers. More precisely,
for any $m\in\mathbb{N}$ we will define, for any $(n+m,X)\leq (n,\Gamma)$ a set
$T(m,X)$ and for any $x\in T(m,X)$ a homomorphism 
$S_x:\mathbb{B}/X\to\mathbb{B}/\Gamma$ of B-frames. Then it follows
immediately that we get stratified functors $U(S_x):U(\mathbb{B})/X\to U(\mathbb{B})/\Gamma$
defining the pre-substitution structure of $\mathcal{E}(\mathbb{B})$.

In the base case, we define for $(n,X)\in U(\mathbb{B})$, the set
$T(0,X)\defeq \{\pt[X]\}$ and $S_{\pt[X]}\defeq \catid{\mathbb{B}/X}$.

%We also define for
%$(n+1,X)\leq (n,\Gamma)$ the set $T(1,X)\jdeq\{x\in B_{n+1}\mid \ebd(x)\jdeq X\}$.
%Note that for $x\in T(1,X)$, the homomorphism $S_x:\mathbb{B}/X \to\mathbb{B}/\Gamma$
%is already assumed to exist, because $\mathbb{B}$ is a pre-B-system.

For the inductive step, suppose that for $m\in\mathbb{N}$ we have sets $T(m,X)$
for any $(n+m,X)\leq (n,\Gamma)$ and homomorphisms $S_x:\mathbb{B}/X \to \mathbb{B}/\Gamma$
of B-frames, for any $x\in T(m,X)$. Then we define
\begin{equation*}
T(m+1,X)\defeq\bigsqcup_{x\in T(m,\eft(X))} \{u\in \tilde{B}_{n+1}\mid \ebd(u)\jdeq S_x(X)\}.
\end{equation*}
For $(x,u)\in T(m+1,X)$ we define the homomorphism $S_{(x,u)}:\mathbb{B}/X\to
\mathbb{B}/\Gamma$ as
\begin{equation*}
S_{(x,u)}\defeq S_u\circ S_x/X,
\end{equation*}
where $S_u$ is taken from the structure of $\mathbb{B}$ of being a pre-B-system.
It remains to define a pre-weakening structure on $U(\mathbb{B})$. To do this,
it suffices to define homomorphisms $W_{(m,X)}:\mathbb{B}/\Gamma \to\mathbb{B}/X$
of B-frames, for any $(n+m,X)\leq (n,\Gamma)$. We define
\begin{equation*}
W_{(m,X)}\defeq W_X\circ\cdots\circ W_{\eft^{m-1}(X)}.\qedhere
\end{equation*}
\end{constr}

\begin{thm}
The functor $\mathcal{E}$ is full and faithful.
\end{thm}

\begin{proof}
We have to show that the map $\mathrm{hom}(\mathbb{B},\mathbb{A})\to
\mathrm{hom}(\mathcal{E}(\mathbb{B}),\mathcal{E}(\mathbb{A}))$ is a bijection
for any two pre-B-systems $\mathbb{B}$ and $\mathbb{A}$.

To show that this map is injective, suppose $F,G:\mathbb{B}\to\mathbb{A}$ are 
pre-B-homomorphisms such that $\mathcal{E}(F)\jdeq\mathcal{E}(G)$. Then we have
$F_n(X)\jdeq\mathcal{E}(F)(n,X)\jdeq \mathcal{E}(G)(n,X)\jdeq G_n(X)$ for each
object $x\in B_n$. Also, we have $\tilde{F}_n(x)\jdeq T_{\mathcal{E}(F)}(n,x)
\jdeq T_{\mathcal{E}(G)}(n,x)\jdeq \tilde{G}_n(x)$ for each
$x\in\tilde{B}_n$. This shows that $F\jdeq G$, so we conclude that $\mathcal{E}$
is faithful.

To show that the map $\mathrm{hom}(\mathbb{B},\mathbb{A})\to
\mathrm{hom}(\mathcal{E}(\mathbb{B}),\mathcal{E}(\mathbb{A}))$ 
is surjective, let $H:\mathcal{E}(\mathbb{B})\to
\mathcal{E}(\mathbb{A})$. Since both $\mathcal{E}(\mathbb{B})$ and
$\mathcal{E}(\mathbb{A})$ are stratified, we get from
\autoref{lem:strat_full} that $H$ is a stratified pre-E-homomorphism. 
Because $H$ is stratified, we get maps $H_n:B_n\to A_n$ for any $n\in\mathbb{N}$.
The term structure of $H$ gives maps $\tilde{H}_n:
\tilde{B}_n\to\tilde{A}_n$. This defines a pre-B-homomorphism since $H$ 
preserves the pre-weakening and pre-substitution structures of $\mathcal{E}(\mathbb{B})$.
It is immediate from the definition of $H_n$ and
$\tilde{H}_n$ that this pre-B-homomorphism is mapped back to $H$, so we conclude
that $\mathcal{E}$ is full.
\end{proof}

### 2017-07-18 compared.tex
starting to restructure the text

@@
 Then we show that the category $\mathbf{Esys}_s$ of stratified E-systems with homomorphisms preserving the stratification, is a full subcategory of $\mathbf{Esys}$.
+Recall from \autoref{defn:esys_cat}, that we assume that morphisms between E-systems are functors that lift factorizations in the sense of \autoref{defn:lift_factorizations}.
 
@@
 $L(X)\jdeq L(Y)$ if and only if $X\jdeq Y$ and $f\jdeq\catid{X}$. 
-\item every morphism $f:X\to Y$ in $\cat{C}_\mathbb{E}$, where $L(X)\jdeq
+\item every morphism $f:X\to Y$ in $\cat{C}$, where $L(X)\jdeq
 n+m+1$ and $L(Y)\jdeq n$, has a unique factorization 
@@
 if $L_{\cat{C}}\jdeq L_{\cat{D}}\circ F$.
+
+The category of stratified categories is denoted by $\mathbf{Str}$.
 \end{defn}
@@
 Let $\cat{C}$ be a stratified category. The root of the tree is going to be the terminal
-object of $\cat{C}$. Note that objects of $\cat{C}$ have no automorphisms other
+object of $\cat{C}$. Note that objects of $\cat{C}$ have no endomorphisms other
 than the identity morphisms. Therefore, to show that $\cat{C}$ is a tree,
 it suffices to show that for every object $X$ with $L(X)>0$ there is
-a unique object $\eft(X)$ with $L(\eft(X))\jdeq L(X)-1$, with a morphism $X\to \eft(X)$.
+a unique object $\eft(X)$ with $L(\eft(X))\jdeq L(X)-1$, with a morphism $p_X:X\to \eft(X)$.
 The existence follows from the existence of the factorization of $X\to 1$. The 
 uniqueness follows from the observation that for any arrow
-$X\to Y$ with $L(Y)\jdeq L(X)-1$, we can factorize $X\to 1$ through $Y$. Since
-factorizations are unique, we will have $Y\jdeq \eft(X)$. 
+$f:X\to Y$ with $L(Y)\jdeq L(X)-1$, we can factorize $X\to 1$ through $Y$. Since
+factorizations are unique, we will have $Y\jdeq \eft(X)$, and $p_X\jdeq f$. 
 \end{proof}
@@
 of small categories with terminal objects and between them functors which lift 
-factorizations and preserve the terminal object.
+factorizations in the sense of \autoref{defn:lift_factorizations}, and preserve the terminal object.
 \end{lem}
@@
 factorizations and 
-which preserves terminal objects. We show by induction that $F$ is stratified.
-Since $F$ preserves the terminal object, $L(Y)\jdeq 0$ implies $L(F(Y))\jdeq 0$.
-Now suppose that for $n\in\mathbb{N}$, we have that $L(\Gamma)\jdeq n$ implies
-$L(F(Y))\jdeq n$, and let $X$ be such that $L(X)\jdeq n+1$ with
-$p_X:X\to\eft(X)$. Then the poset $\mathbf{fact}(p_X)$ is isomorphic to the
+which preserves terminal objects. We show by induction that $F$ is stratified, i.e.~that $L_{\mathcal{C}}(X)\jdeq L_{\mathcal{D}}(F(X))$, for any object $X$ of $\mathcal{C}$.
+
+For the base case, we observe that since $F$ preserves the terminal object, $L(X)\jdeq 0$ implies $L(F(X))\jdeq 0$.
+
+For the inductive step, suppose that for $n\in\mathbb{N}$, we have that $L(X)\jdeq n$ implies
+$L(F(X))\jdeq n$, for any object $X$ of $\mathcal{C}$, and let $Y$ be such that $L(Y)\jdeq n+1$ with
+$p_Y:Y\to\eft(Y)$. Then the poset $\mathbf{fact}(p_Y)$ is isomorphic to the
 poset $\mathbf{2}$ with two objects, one of which is smaller than the other. By the
-assumption that $F$ lifts factorizations, the poset $\mathbf{fact}(F(p_X))$
+assumption that $F$ lifts factorizations, the poset $\mathbf{fact}(F(p_Y))$
 is also isomorphic to $\mathbf{2}$. Thus, the only factorization
-of $F(X)\to\unit$ starts with $F(p_X):H(X)\to F(\eft(X))$, which implies that
-$L(F(X))\jdeq n+1$. 
+of $F(Y)\to\unit$ starts with $F(p_Y):H(Y)\to F(\eft(Y))$, which implies that
+$L(F(Y))\jdeq n+1$. 
 \end{proof}
 
+\begin{cor}
+A category can only be stratified in at most one way. That is, being stratified is a property of categories.
+\end{cor}
+
 \begin{rmk}
-It is easy to show that a category can only be stratified in at most one way,
-and that if a category is stratified, then so are any of its slices. Therefore,
+It is easy to show that if a category is stratified, then so are any of its slices. Therefore,
 it makes sense to ask for the property that the weakening and substitution
@@
 \begin{defn}
-A pre-E-systems is said to be \define{stratified} if its underlying category is
+A pre-E-system is said to be \define{stratified} if its underlying category is
 stratified and if each $W_A$ and $S_x$ is a stratified functor. The category of
@@
 theory of B0-systems with equalities ensuring that the rest of the operations 
-are compatible with each other. Voevodsky, furthermore distinguishes
+are compatible with each other. 
+Voevodsky furthermore distinguishes
 non-unital and unital versions of these, and noted that not all B-homomorphisms
@@
 \begin{rmk}
-Since the indices $n,m\in\mathbb{N}$ can be infered, we will usually omit them
+Since the indices $n,m\in\mathbb{N}$ can be inferred, we will usually omit them
 in the infix form of the operators.
@@
 \begin{defn}
-For every B-framework $\mathbb{B}$ and any $X\in B_n$, there is a B-framework
-$\mathbb{B}/X$ given by
-\begin{align*}
-(B/X)_{m} & \jdeq \{Y\in B_{n+m}\mid\eft^{m}(Y)\jdeq X\}\\
-(\tilde{B}/X)_m & \jdeq \{y\in \tilde{B}_{n+m}\mid\eft^m(\ebd(y))\jdeq X\}.
-\end{align*}
-Also, for any homomorphism $H:\mathbb{B}\to\mathbb{A}$ of B-frameworks and any
-$X\in B_n$, there is a homomorphism $H/X:\mathbb{B}/X\to\mathbb{A}/H(X)$
-defined in the obvious way.
-\end{defn}
-
-\begin{rmk}
-Note that for $X\in B_n$ and $Y\in B_{n+m}$ such that $\eft^m(Y)\jdeq X$, 
-we have $(\mathbb{B}/X)/Y\cong B/Y$.
-\end{rmk}
-
-\begin{defn}
-A \define{B0-system} $\mathbb{B}$ consists of a B-framework $\mathbb{B}$ and homomorphisms
-\begin{align*}
-W_{X} & : \mathbb{B}/\eft(X)\to\mathbb{B}/X\\
-S_{x} & : \mathbb{B}/\ebd(x)\to\mathbb{B}/\eft(\ebd(x))
-\end{align*}
-of B-frameworks, for any $X\in B_{n+1}$ and any $x\in\tilde{B}_{n+1}$. 
-\end{defn}
-
-\begin{lem}
-When $\mathbb{B}$ is a B0-system, then so is each $\mathbb{B}/X$.
-\end{lem}
-
-\begin{proof}
-Straightforward.
-\end{proof}
-
-\begin{defn}
-A \define{B0-homomorphism} $H:\mathbb{B}\to\mathbb{A}$ is a homomorphism of B-frameworks
-for which the diagrams
-\begin{equation*}
-\begin{tikzcd}[column sep=huge]
-\mathbb{B}/X \arrow[r,"H/X"] & \mathbb{A}/H(X) \\
-\mathbb{B}/\eft(X) \arrow[u,"W_X"] \arrow[r,swap,"H/\eft(X)"] & \mathbb{A}/\eft(H(X)) \arrow[u,swap,"W_{H(X)}"]
-\end{tikzcd}
-\end{equation*}
-and
-\begin{equation*}
-\begin{tikzcd}[column sep=huge]
-\mathbb{B}/\ebd(x) \arrow[r,"H/\ebd(x)"] \arrow[d,swap,"S_x"] & \mathbb{A}/\ebd(H(X)) \arrow[d,"S_{\tilde{H}(x)}"] \\
-\mathbb{B}/\eft(\ebd(x)) \arrow[r,swap,"H/\eft(\ebd(x))"] & \mathbb{A}/\eft(\ebd(H(X)))
-\end{tikzcd}
-\end{equation*}
-of homomorphisms of B-frameworks commute for every 
-$X\in B_n$ and $x\in\tilde{B}_n$. The category of B0-systems is denoted
-by $\mathbf{B0sys}$.
-\end{defn}
-
-Now that we have defined the category of B0systems, we can start defining the
-embedding into pre-E-systems.
-
-\begin{defn}
-Let $\mathbb{B}$ be a B-framework. Then $\mathbb{B}$ determines
-a stratified category $U(\mathbb{B})$. Also, any homomorphism of B-frameworks
-determines a stratified functor.
+We construct a functor $U:\mathbf{Bfr}\to\mathbf{tCat_s}$ from the category
+of B-frameworks to the category of stratified categories with term structure.
 \end{defn}
@@
 \begin{constr}
+Let $\mathbb{B}$ be a B-framework.
 The set of objects of $U(\mathbb{B})$ is taken to be the set
@@
 isomorphism $U(\mathbb{B}/X)\cong  U(\mathbb{B})/(n,X)$ of rooted trees,
-natural in $X$. Therefore we will usually not distinguish between $U(\mathbb{B})/X$
+natural in $X$. Therefore we will usually not distinguish between $U(\mathbb{B}/X)$
 and $U(\mathbb{B})/(n,X)$. 
@@
 \begin{defn}
+For every B-framework $\mathbb{B}$ and any $X\in B_n$, there is a B-framework
+$\mathbb{B}/X$ given by
+\begin{align*}
+(B/X)_{m} & \jdeq \{Y\in B_{n+m}\mid\eft^{m}(Y)\jdeq X\}\\
+(\tilde{B}/X)_m & \jdeq \{y\in \tilde{B}_{n+m}\mid\eft^m(\ebd(y))\jdeq X\}.
+\end{align*}
+Also, for any homomorphism $H:\mathbb{B}\to\mathbb{A}$ of B-frameworks and any
+$X\in B_n$, there is a homomorphism $H/X:\mathbb{B}/X\to\mathbb{A}/H(X)$
+defined in the obvious way.
+\end{defn}
+
+\begin{rmk}
+Note that for $X\in B_n$ and $Y\in B_{n+m}$ such that $\eft^m(Y)\jdeq X$, 
+we have an isomorphism $(\mathbb{B}/X)/Y\cong B/Y$ of B-frameworks, constructed in the obvious way, which is natural in the sense that for any homomorphism $H:\mathbb{B}\to\mathbb{A}$ of B-frameworks, the square
+\begin{equation*}
+\begin{tikzcd}
+(\mathbb{B}/X)/Y \arrow[r,"\cong"] \arrow[d,swap,"(H/X)/Y"] &  \mathbb{B}/Y \arrow[d,"H/Y"] \\
+(\mathbb{A}/H(X))/H(Y) \arrow[r,"\cong"] & \mathbb{A}/H(Y)
+\end{tikzcd}
+\end{equation*}
+commutes.
+\end{rmk}
+
+\begin{defn}
+A \define{non-unital B0-system} $\mathbb{B}$ consists of a B-framework $\mathbb{B}$ and homomorphisms
+\begin{align*}
+W_{X} & : \mathbb{B}/\eft(X)\to\mathbb{B}/X\\
+S_{x} & : \mathbb{B}/\ebd(x)\to\mathbb{B}/\eft(\ebd(x))
+\end{align*}
+of B-frameworks, for any $X\in B_{n+1}$ and any $x\in\tilde{B}_{n+1}$. 
+\end{defn}
+
+\begin{comment}
+\begin{lem}
+When $\mathbb{B}$ has the structure of a B0-system, then so does each $\mathbb{B}/X$.
+\end{lem}
+
+\begin{proof}
+Straightforward.
+\end{proof}
+\end{comment}
+
+\begin{defn}
+A \define{non-unital B0-homomorphism} $H:\mathbb{B}\to\mathbb{A}$ is a homomorphism of B-frameworks
+for which the diagrams
+\begin{equation*}
+\begin{tikzcd}[column sep=huge]
+\mathbb{B}/X \arrow[r,"H/X"] & \mathbb{A}/H(X) \\
+\mathbb{B}/\eft(X) \arrow[u,"W_X"] \arrow[r,swap,"H/\eft(X)"] & \mathbb{A}/\eft(H(X)) \arrow[u,swap,"W_{H(X)}"]
+\end{tikzcd}
+\end{equation*}
+and
+\begin{equation*}
+\begin{tikzcd}[column sep=huge]
+\mathbb{B}/\ebd(x) \arrow[r,"H/\ebd(x)"] \arrow[d,swap,"S_x"] & \mathbb{A}/\ebd(H(X)) \arrow[d,"S_{\tilde{H}(x)}"] \\
+\mathbb{B}/\eft(\ebd(x)) \arrow[r,swap,"H/\eft(\ebd(x))"] & \mathbb{A}/\eft(\ebd(H(X)))
+\end{tikzcd}
+\end{equation*}
+of homomorphisms of B-frameworks commute for every 
+$X\in B_n$ and $x\in\tilde{B}_n$. The category of non-unital B0-systems is denoted
+by $\mathbf{B0sys}^\circ$.
+\end{defn}
+
+\begin{defn}
 Let $\mathbb{B}$ be a B0-system. Then $U(\mathbb{B})$ can be given the structure
-of a pre-E-systems. We write $\mathcal{E}(\mathbb{B})$ for the resulting
-pre-E-systems. Thus we get a functor $\mathcal{E}:\mathbf{B0sys}\to\mathbf{Esys}_s$.
+of a pre-E-system. We write $\mathcal{E}(\mathbb{B})$ for the resulting
+pre-E-system. Thus we get a functor $\mathcal{E}:\mathbf{B0sys}\to\mathbf{E0sys}_s$.
 \end{defn}
@@
 \begin{constr}
-We begin by simultaneously defining the term structure and the pre-substitution
+We begin by \emph{simultaneously} defining the term structure and the pre-substitution
 structure on $U(\mathbb{B})$ by induction on the natural numbers. More precisely,
@@
 
-For $(n,X)\in U(\mathbb{B})$, we define $T(0,X)\defeq
-\{\pt[X]\}$ and $S_{\pt[X]}\defeq \catid{\mathbb{B}/X}$. We also define for
-$(n+1,X)\leq (n,\Gamma)$ the set $T(1,X)\jdeq\{x\in B_{n+1}\mid \ebd(x)\jdeq X\}$.
-Note that for $x\in T(1,X)$, the homomorphism $S_x:\mathbb{B}/X \to\mathbb{B}/\Gamma$
-is already assumed to exist, because $\mathbb{B}$ is a B0-system.
+In the base case, we define for $(n,X)\in U(\mathbb{B})$, the set
+$T(0,X)\defeq \{\pt[X]\}$ and $S_{\pt[X]}\defeq \catid{\mathbb{B}/X}$.
+
+%We also define for
+%$(n+1,X)\leq (n,\Gamma)$ the set $T(1,X)\jdeq\{x\in B_{n+1}\mid \ebd(x)\jdeq X\}$.
+%Note that for $x\in T(1,X)$, the homomorphism $S_x:\mathbb{B}/X \to\mathbb{B}/\Gamma$
+%is already assumed to exist, because $\mathbb{B}$ is a B0-system.
 
@@
 \begin{equation*}
-T(m+1,X)\defeq\bigsqcup_{x\in T(m,\eft(X))} T(1,S_x(X)).
+T(m+1,X)\defeq\bigsqcup_{x\in T(m,\eft(X))} \{u\in \tilde{B}_{n+1}\mid \ebd(u)\jdeq S_x(X)\}.
 \end{equation*}
@@
 \begin{equation*}
-S_{(x,u)}\defeq S_u\circ S_x/X.
-\end{equation*}
+S_{(x,u)}\defeq S_u\circ S_x/X,
+\end{equation*}
+where $S_u$ is taken from the structure of $\mathbb{B}$ of being a B0-system.
 It remains to define a pre-weakening structure on $U(\mathbb{B})$. To do this,
@@
 \begin{defn}
-A \emph{(unital) B-system} is a B0-system for which the following conditions
+A \define{(unital) B-system} is a B0-system for which the following conditions
 hold:
@@
 \end{enumerate}
-\emph{B-homomorphisms} are unital B0-homomorphisms between B-systems.
+\define{B-homomorphisms} are unital B0-homomorphisms between B-systems.
 We denote the category of B-systems by $\mathbf{Bsys}$. 
@@
 For any non-unital B-system $\mathbb{B}$, $\mathcal{E}(\mathbb{B})$ is a 
-stratified non-unital E-systems. 
+stratified non-unital E-system. 
 \end{lem}
@@
 \begin{defn}
-For any unital B-system $\mathbb{B}$, $\mathcal{E}(\mathbb{B})$ has a
-projection structure. We thus obtain a functor 
-$\mathcal{E}:\mathbf{Bsys}\to\mathbf{Esys}$ from unital B-systems to E-systems.
+We construct a functor $\mathcal{E}:\mathbf{Bsys}\to\mathbf{Esys_s}$ from (unital) B-systems to stratified E-systems.
 \end{defn}
@@
 \begin{constr}
-We first need to extend the functor $\mathcal{E}:\mathbf{B0sys}\to\mathbf{Esys_p}$
-from non-unital B0-systems to non-unital pre-E-systems, to a functor from
-unital B-systems to E-systems.
+We begin by defining a functor $\mathcal{E'}:\mathbf{Bsys}\to\mathbf{E0sys}$
+from (unital) B-systems to (unital) pre-E-systems.
+In the construction of the pre-projection structure on $\mathcal{E}(\mathbb{B})$ we use equations that are satisfied for B-systems, but not necessarily for B0-systems.
 
 We need to give $\mathcal{E}(\mathbb{B})$ the structure of a unital
-pre-E-systems. We construct by induction on $m\in \mathbb{N}$, a term of type 
-$T(m,W_{(m,X)}(X))$ with the property that that 
-\begin{equation*}
-S_{\tfid{(m,X)}}\circ W_{(m,X)}/X\jdeq\catid{U(\mathbb{B})/X},
-\end{equation*} 
-for every $(n+m,X)\leq (n,\Gamma)$ by induction on $m$. The extra
-condition is part of the inductive construction, because we need it on the way.
+pre-E-system. We construct by induction on $m\in \mathbb{N}$, 
+for every $(n+m,X)\leq (n,\Gamma)$, an element $\tfid{(m,X)}\in T(m,W_{(m,X)}(X))$ with the property that that 
+\begin{equation*}
+S_{\tfid{(m,X)}}\circ W_{(m,X)}/X\jdeq\catid{U(\mathbb{B})/X}.
+\end{equation*}
+(This extra condition is needed in the inductive construction.)
 
@@
 We have the term $x\defeq \tilde{W}_X(\tfid{(m,\eft(X))})\in T(m,W_{(m+1,X)}(\eft(X)))$.
-Now we may observe that 
+Now we observe that 
 \begin{align*}
@@
 This finishes the definition of the pre-projection structure on $\mathcal{E}(\mathbb{B})$.
-Straightforward proves by induction give that each $W_{(m,X)}$ and $S_x$ is a 
+Straightforward proofs by induction give that each $W_{(m,X)}$ and $S_x$ is a 
 pre-projection homomorphism and that $S_x(\tfid{(m,X)})\jdeq x$, for any 
-$x\in T(m,X)$.
+$x\in T(m,X)$, to show that $\mathcal{E}'(\mathbb{B})$ is in fact an E-system.
 \end{constr}
@@
 \begin{thm}
-The functor $\mathcal{E}:\mathbf{Bsys}\to\mathbf{Esys}$ lifts to an equivalence
+The functor $\mathcal{E}:\mathbf{Bsys}\to\mathbf{Esys_s}$ is an equivalence.
+\begin{comment}
 to the subcategory $\mathbf{Esys}_s$ of stratified E-systems: i.e.~we have a
@@
 \end{equation*}
+\end{comment}
 \end{thm}
@@
 E-systems and that it is full and faithful, so it remains
-to show that every stratified E-systems is in the image of $\mathcal{E}$. In
+to show that every stratified E-system is in the image of $\mathcal{E}$. In
 other words, for every stratified E-systems $\mathbb{E}$ we need to construct
@@
 
-Let $\mathbb{E}$ be a stratified telescopi type system. From the stratification functor
+Let $\mathbb{E}$ be a stratified E-system. From the stratification functor
 $L:\mathbb{C}_{\mathbb{E}}\to (\mathbb{N},\geq)$ we obtain the pre-image
@@
 $f^\ast(A)$ for $\mathrm{Ty}(f,A)$.
-\item For any $A\in\mathrm{Ty}(\Gamma)$, a morphism
+\item For any $A\in\mathrm{Ty}(\Gamma)$, an object $\ctxext{\Gamma}{A}$ of $\cat{C}$, and a morphism
 \begin{equation*}
@@
 in $\cat{C}$.
-\item a preasheaf $\mathrm{Tm}:\op{(\int_\cat{C}\mathrm{Ty})}\to\mathbf{Set}$.
+\item a presheaf $\mathrm{Tm}:\op{(\int_\cat{C}\mathrm{Ty})}\to\mathbf{Set}$.
 The action of a morphism $f:(\Delta,f^\ast(A))\to(\Gamma,A)$ of $\int_\cat{C}\mathrm{Ty}$
@@
 Although a terminal object is usually assumed to exist, not every object 
-of a category with families is assumed to be `fibrant', i.e.~it is
+of a category with families is assumed to be `fib rant', i.e.~it is
 not the case that every $X\to 1$ is a composition of projections $p_A$.
-We could only hope that there is a functor from D-sytems to E-systems without
+We could only hope that there is a functor from categories with families to E-systems without
 terminal objects.
@@
 
-There is essentially only one choice for the presheaf $\mathrm{Tm}$ and the
-term $\delta_A$, so the extra structure on a setup for categories with families
-is determined essentially uniquely once it exists.
+There is essentially only one choice for the presheaf $\mathrm{Tm}$ and the term $\delta_A$, satisfying the universal property described in the definition of categories with families.
 
@@
 
-The following two definitions define subclasses of telescopic D-systems and 
-coherent telescopic D-systems, and the latter structure is the same as an CE-system
+\begin{defn}
+A \define{category with dependent contexts} consists of
+\begin{enumerate}
+\item A category $\cat{C}$.
+\item A presheaf $\mathcal{F}:\op{\cat{C}}\to\mathbf{Set}$.
+\item For each $A\in\mathcal{F}(\Gamma)$, an object $\ctxext{\Gamma}{A}$, and a morphism $p_A:\ctxext{\Gamma}{A}\to\Gamma$.
+\item For each $\Gamma\in\cat{C}$, an element $\emptyf[\Gamma]\in\mathcal{F}(\Gamma)$ and for each $A\in\mathcal{F}(\Gamma)$ and each $P\in\mathcal{F}(\ctxext{\Gamma}{A})$, an element $\ctxext{A}{P}\in\mathcal{F}(\Gamma)$, such that
+\begin{enumerate}
+\item 
+\begin{itemize}
+\item for any $A\in\mathrm{Fam}(\Gamma)$ and $P\in\mathrm{Fam}(\ctxext{\Gamma}{A})$,
+one has $\ctxext{\Gamma}{{A}{P}}\jdeq\ctxext{{\Gamma}{A}}{P}$, and
+\begin{equation*}
+p_{\ctxext{A}{P}}\jdeq p_A\circ p_P,
+\end{equation*}
+\item and for any $\Gamma\in\cat{C}$ one has $\ctxext{\Gamma}{\emptyf[\Gamma]}\jdeq\Gamma$, and
+\begin{equation*}
+p_{\emptyf[\Gamma]}\jdeq\catid{\Gamma}.
+\end{equation*}
+\end{itemize}
+\item 
+\begin{itemize}
+\item for any $A\in\mathrm{Fam}(\Gamma)$, $P\in\mathrm{Fam}(\ctxext{\Gamma}{A})$
+and $Q\in\mathrm{Fam}(\ctxext{{\Gamma}{A}}{P})$, one has
+\begin{equation*}
+\ctxext{{A}{P}}{Q}\jdeq\ctxext{A}{{P}{Q}},
+\end{equation*}
+\item and for any $A\in\mathrm{Fam}(\Gamma)$, one has
+\begin{equation*}
+\ctxext{\emptyf[\Gamma]}{A}\jdeq\ctxext{A}{\emptyf[\ctxext{\Gamma}{A}]}\jdeq A.
+\end{equation*}
+\end{itemize}
+\item 
+\begin{itemize}
+\item for any $A\in\mathrm{Fam}(\Gamma)$, $P\in\mathrm{Fam}(\ctxext{\Gamma}{A})$
+and any $f:\Delta\to\Gamma$ in $\cat{C}$, one has
+\begin{equation*}
+f^\ast(\ctxext{A}{P})\jdeq \ctxext{(f^\ast(A))}{(\pi_2(f,A)^\ast(P))},
+\end{equation*}
+\item and for any $f:\Delta\to\Gamma$ in $\cat{C}$, one has
+\begin{equation*}
+f^\ast(\emptyf[\Gamma])\jdeq\emptyf[\Delta].
+\end{equation*}
+\end{itemize}
+\end{enumerate} 
+\end{enumerate}
+\end{defn}
+
+The following two definitions define subclasses of \emph{telescopic} categories with families and 
+\emph{coherent} telescopic categories with families, and the latter structure is the same as an CE-system
 (without empty context). In \autoref{defn:comp_to_telescopic}, we will show how
-to obtain a telescopic D-system from any comprehension category. This construction
-will also generate a telescopic D-system out of any D-system, because D-systems
+to obtain a telescopic category with families from any comprehension category. This construction
+will also generate a telescopic categories with families out of any category with families, because categories with families
 are comprehension categories.
@@
 \begin{defn}
-Let $\cat{C}$ be a D-system. In this definition, we will write
-$\mathrm{Fam}$ for the presheaf of types. The D-system $\cat{C}$
+Let $\mathbb{D}$ be a category with families with underlying category $\cat{C}$. In this definition, we will write
+$\mathrm{Fam}$ for the presheaf of types. The category with families $\mathbb{D}$
 is said to be \define{telescopic}, if we have elements $\emptyf[\Gamma]\in
@@
 
-\begin{lem}
-A D-system is telescopic if and only if  the following two conditions hold:
+\begin{comment}
+\begin{defn}
+Consider a functor $p:\cat{E}\to\cat{B}$, and let $\phi:e\to e'$ be a morphism in $\cat{E}$.
+We say that $\phi$ is \define{cocartesian} if for any morphism $\psi:e\to e''$ and any $g:p(e')\to p(e'')$ such that $g\circ p(\phi)=p(\psi)$, there is a unique morphism $h:e'\to e''$ such that $h\circ\phi=\psi$.
+
+A \define{cocartesian lift} at $e$ of a morphism $\pi:p(e)\to b'$ in $\cat{B}$ is a cocartesian morphism $\phi:e\to e'$ for which $p(\phi)=\pi$.
+\end{defn}
+\end{comment}
+
+\begin{comment}
+\begin{lem}[Work in progress]
+A category with families is telescopic if and only if  the following two conditions hold:
 \begin{enumerate}
@@
 
-\begin{proof}
-Let $\mathbb{D}$ be a telescopic D-system. To show that $\mathbb{D}$ has the two
+\begin{proof}[Work in progress]
+Let $\mathbb{D}$ be a telescopic category with families. To show that $\mathbb{D}$ has the two
 stated properties:
@@
 
-Let $\mathbb{D}$ be a D-system for which the stated two conditions hold. To show
-that $\mathbb{D}$ is a telescopic D-system, we systematically go through each of the
+Let $\mathbb{D}$ be a category with families for which the stated two conditions hold. To show
+that $\mathbb{D}$ is a telescopic category with families, we systematically go through each of the
 properties:
@@
 \end{proof}
-
-\begin{conj}
+\end{comment}
+
+\begin{conj}[Work in progress]
 The notion of being telescopic is a property.
@@
 \begin{defn}
-A telescopic D-system is said to be \define{coherent}, if for any $A,A'\in\mathrm{Fam}(\Gamma)$
+A category with dependent contexts is said to be \define{coherent}, if for any $A,A'\in\mathrm{Fam}(\Gamma)$
 such that $p_A\jdeq p_{A'}$, and any $f:\Delta\to\Gamma$, we have
@@
 
+\begin{defn}
+A category with dependent contexts is said to be \define{separated}, if for any $A,A'\in\mathrm{Fam}(\Gamma)$, we have
+\begin{equation*}
+((\ctxext{\Gamma}{A},p_A) \jdeq (\ctxext{\Gamma}{A'},p_{A'}))\to (A\jdeq A').
+\end{equation*}
+\end{defn}
+
+\begin{lem}
+Any separated category with dependent contexts is coherent.
+\end{lem}
+
+\begin{proof}
+Straightforward.
+\end{proof}
+
 \begin{thm}
-An CE-system without empty context is the same thing as a coherent telescopic D-system.
+Every CE-system without empty context is a coherent telescopic D-system.
 \end{thm}
@@
 
-We define a functor $\mathcal{P}:\int_{\cat{C}}\to\cat{C}^\square$ by sending
+We define a functor $\mathcal{P}:\int_{\cat{C}}\mathrm{Fam}\to\cat{C}^\square$ by sending
 the pair $(\Gamma,A)$ to the morphism $A:\ctxext{\Gamma}{A}\to\Gamma$ in
@@
 available is precisely what the coherence condition is for). The verification of conditions
-(a) to (d) of \autoref{defn:E'sys} is straightforward.
+(a) to (d) of \autoref{defn:CEsys} is straightforward.
 \end{proof}
@@
 Dybjer's notion of categories with families. It is shown in 
-\cite{Hofmann_syntax_semantics} that the category of categoriew with families
+\cite{Hofmann_syntax_semantics} that the category of categories with families
 is equivalent to the category of categories with attributes, which forms a
@@
 
-In this section we do the preparateions for this comparison. After having
+In this section we do the preparations for this comparison. After having
 introduced categories with families, we prove that categories with families
@@
 this category should possess a terminal object, so we naturally have to retreat
-to CE-systems without empty contex. Even when we make this generalization, 
+to CE-systems without empty context. Even when we make this generalization, 
 the construction of a canonical functor in which families are represented by 
@@
 
-\begin{defn}\label{defn:comp_to_telescopic}
+\begin{defn}[To do: transport the notion of telescopic from CWFs to D-systems (probably by reading the proof below)]\label{defn:comp_to_telescopic}
 Let $\mathcal{P}$ be a split comprehension category with choice $j$ of cartesian
@@
 Let $f_0\defeq f:\Delta\to\Gamma$ be a morphism of $\cat{C}$, the base category of a
-split comprehension category $\mathcal{P}$ wich a choice $j$ of cartesian lifts.
+split comprehension category $\mathcal{P}$ which a choice $j$ of cartesian lifts.
 By induction we define an element $\Lambda(f,A)\in\Lambda(\Delta)$ together with
@@
 \begin{thm}
-Let $\mathcal{P}$ be a comprehension category. Then the following are equivalent:
+Let $\mathcal{P}$ be a (extra assumption: split) comprehension category. Then the following are equivalent:
 \begin{enumerate}
@@
 definition of \autoref{defn:E'sys}. Since $f^\ast$ is defined as an extension
-of $\Lambda(f)$, it is enough to verify the analoguous functoriality properties
+of $\Lambda(f)$, it is enough to verify the analogous functoriality properties
 for $\Lambda(f)$. We do this by induction on the objects of $\Lambda(\Gamma)$,
@@
 and $(f,\pi_2(f,A)):f^\ast(A)\to A$. It is routine to verify that $(f,\pi_2(f,A))$
-is a cartesian lift of $f$, and that this choice of catesian lifts is functorial.
+is a cartesian lift of $f$, and that this choice of cartesian lifts is functorial.
 
@@
 is functorial by assumption, so we have defined a category with families. It
-is straigtforward to verify that this category with families is locally well-founded,
+is straightforward to verify that this category with families is locally well-founded,
 and that its construction is inverse to $U$. 

### 2017-07-18 e-systems.tex
starting to restructure the text

@@
 \begin{defn}
-A \define{pre-E-system} is a pre-category $\cat{F}$ with term structure
+A \define{non-unital pre-E-system} is a pre-category $\cat{F}$ with term structure
 with a pre-weakening structure $W$, a pre-substitution structure $S$ and a terminal
@@
 
-\begin{defn}
+\begin{defn}\label{defn:E0sys_cat}
 A \define{pre-E-homomorphism} from $\mathbb{E}\to\mathbb{D}$ is a functor
@@
 also a pre-projection homomorphism.
+
+We write $\mathbf{E0sys}$ for the category of pre-E-systems. 
 \end{defn}
@@
 \label{defn:esystems}
-A \define{E-system} is a unital pre-E-system in which 
+An \define{E-system} is a unital pre-E-system in which 
 \begin{enumerate}
@@
 
-\begin{defn}
-We write $\mathbf{Ctx}$ for the category of E-systems where the underlying
+\begin{defn}\label{defn:esys_cat}
+We write $\mathbf{Esys}$ for the category of E-systems where the underlying
 functor of each E-homomorphism lifts factorizations.
@@
 expression is in normal form, and where
-each occurence of $S_x\circ W_A$ and $S_{\tfid{A}}\circ W_A/A$ is eliminated.
+each occurrence of $S_x\circ W_A$ and $S_{\tfid{A}}\circ W_A/A$ is eliminated.
 \end{rmk}

### 2017-07-18 internal_homs.tex
starting to restructure the text

@@
 essence of being a morphism from the family $P$ over $A$ to the family
-$Q$ over $B$, reindexed by $f$. 
+$Q$ over $B$, re-indexed by $f$. 
 Whenever we draw such diagrams, a double headed arrow will always
@@
 \begin{thm}
-A E-system is the same thing as an CE-system.
+Every CE-system is an E-system.
 \end{thm}
@@
 
+In this section we define for every E-system $\mathbb{E}$, and every context $\Gamma$ in $\mathbb{E}$, a category $\mathcal{C}_\mathbb{E}(\Gamma)$.
+This goal is accomplished in \autoref{thm:internal_hom}.
+Note that we do not have a category structure on the set of contexts of $\mathbb{E}$, unless $\mathbb{E}$ possesses an empty context $[]$.
+In this case, the category structure is inherited from the category $\mathcal{C}_{\mathbb{E}}([])$.
+
 \begin{defn}
@@
 
-\begin{thm}
+\begin{thm}\label{thm:internal_hom}
 For each object $\Gamma$ of a E-system $\mathbb{E}$, we have a pre-category
-$\mathbb{C}_I(\mathbb{E},\Gamma)$ of internal morphisms of $\mathbb{E}$,
+$\mathcal{C}_{\mathbb{E}}(\Gamma)$ of internal morphisms of $\mathbb{E}$,
 consisting of the same objects as $\cat{F}/\Gamma$ and the hom-sets 
@@
 homomorphism, it follows immediately that any E-homomorphism preserves any
-$f^\ast$, i.e.~for any $f\in\thom{A}{B}$ in $\mathcal{E}$ and any E-homomorphism
-$F:\mathcal{E}\to\mathcal{D}$, the diagram
+$f^\ast$, i.e.~for any $f\in\thom{A}{B}$ in $\mathbb{E}$ and any E-homomorphism
+$F:\mathbb{E}\to\mathbb{D}$, the diagram
 \begin{equation*}
 \begin{tikzcd}[column sep=large]
-\mathcal{E}/\ctxext{\Gamma}{A} \arrow[r,"F/\ctxext{\Gamma}{A}"] & \mathcal{D}/F(\ctxext{\Gamma}{A})
-  \\
-\mathcal{E}/\ctxext{\Gamma}{B} \arrow[u,"f^\ast"] \arrow[r,swap,"F/\ctxext{\Gamma}{B}"] & \mathcal{D}/F(\ctxext{\Gamma}{B}) \arrow[u,swap,"F(f)^\ast"]
+\mathbb{E}/\ctxext{\Gamma}{A} \arrow[r,"F/\ctxext{\Gamma}{A}"] & \mathbb{D}/F(\ctxext{\Gamma}{A})
+  \\
+\mathbb{E}/\ctxext{\Gamma}{B} \arrow[u,"f^\ast"] \arrow[r,swap,"F/\ctxext{\Gamma}{B}"] & \mathbb{D}/F(\ctxext{\Gamma}{B}) \arrow[u,swap,"F(f)^\ast"]
 \end{tikzcd}
@@
 commutes. This applies to weakening, substitution and pre-composition, and it
-follows that for any E-homomorphism $F:\mathcal{E}\to\mathcal{D}$ and any
-$\Gamma\in\mathcal{E}$, we have a functor $\mathbb{C}_I(F,\Gamma):
-\mathbb{C}_I(\mathcal{E},\Gamma)\to\mathbb{C}_I(\mathcal{D},F(\Gamma))$.
+follows that for any E-homomorphism $F:\mathbb{E}\to\mathbb{D}$ and any
+$\Gamma\in\mathbb{E}$, we have a functor $\mathcal{C}_F(\Gamma):
+\mathcal{C}_{\mathbb{E}}(\Gamma)\to\mathcal{C}_{\mathbb{D}}(F(\Gamma))$.
 \end{rmk}
-
-The following fact will be useful:
-
-\begin{lem}\label{lem:compW_W}
-Let $f\in\thom{A}{B}$ be an internal morphism in context $\Gamma$. Then one has
-\begin{equation*}
-f^\ast\circ W_B\jdeq W_A.
-\end{equation*}
-\end{lem}
-
-\begin{proof}
-The proof is a simple calculation:
-\begin{equation*}
-f^\ast\circ W_B\jdeq S_f\circ W_A/B\circ W_B\jdeq S_f\circ W_{W_A(B)}\circ W_A
-\jdeq W_A.\qedhere
-\end{equation*}
-\end{proof}
 
@@
 \begin{cor}\label{cor:pairing_ehom}
-Let $F:\mathcal{E}\to\mathcal{D}$ be an E-homomorphism and let $x\in T(A)$ and
+Let $F:\mathbb{E}\to\mathbb{D}$ be an E-homomorphism and let $x\in T(A)$ and
 $u\in T(S_x(P))$. Then we have
@@
 \begin{equation*}
-F^\circledast\defeq F^\ast\circ (f^\ast/Q):
+F^\bullet\defeq F^\ast\circ (f^\ast/Q):
 \mathbb{E}/\ctxext{{\Gamma}{B}}{Q}\to\mathbb{E}/\ctxext{{\Gamma}{A}}{P}.
 \end{equation*}
-The infix notation of $F^\circledast$ is taken to be $\jfcomp{\blank}{\blank}{\blank}{F}{\blank}$.
+The infix notation of $F^\bullet$ is taken to be $\jfcomp{\blank}{\blank}{\blank}{F}{\blank}$.
 \end{defn}
@@
 \begin{equation*}
-F^\circledast\jdeq (\jvcomp{P}{f}{F})^\ast.
+F^\bullet\jdeq (\jvcomp{P}{f}{F})^\ast.
 \end{equation*}
@@
 
-We are now in position to state and prove the interchange law of horizontal
-and vertical composition.
+In the next theorem we prove the interchange law of horizontal and vertical composition.
+Its proof uses the following fact.
+
+\begin{lem}\label{lem:compW_W}
+Let $f\in\thom{A}{B}$ be an internal morphism in context $\Gamma$. Then one has
+\begin{equation*}
+f^\ast\circ W_B\jdeq W_A.
+\end{equation*}
+\end{lem}
+
+\begin{proof}
+The proof is a simple calculation:
+\begin{equation*}
+f^\ast\circ W_B\jdeq S_f\circ W_A/B\circ W_B\jdeq S_f\circ W_{W_A(B)}\circ W_A
+\jdeq W_A.\qedhere
+\end{equation*}
+\end{proof}
 
@@
 in context $\Gamma$. 
-Then $F^\circledast\circ G^\circledast\jdeq (\jfcomp{A}{f}{P}{F}{G})^\circledast$.
+Then $F^\bullet\circ G^\bullet\jdeq (\jfcomp{A}{f}{P}{F}{G})^\bullet$.
 In other words the composition $\jfcomp{}{}{}{\blank}{\blank}$ is associative.
@@
 \begin{align*}
-F^\circledast\circ G^\circledast
+F^\bullet\circ G^\bullet
   & \jdeq
@@
   & \jdeq
-(\jfcomp{A}{f}{P}{F}{G})^\circledast.
+(\jfcomp{A}{f}{P}{F}{G})^\bullet.
   \tag{By \autoref{lem:three-composition}}
@@
 
-\begin{lem}
+\begin{thm}
 Let $\jhom{\Gamma}{A}{B}{f}$ and $\jfhom{\Gamma}{A}{B}{f}{P}{Q}{F}$. Then
@@
 commutes, and $\jcomp{}{\jvcomp{P}{f}{F}}{\cprojsndf{B}{Q}}\jdeq F$.
-\end{lem}
+\end{thm}
 
@@
 \begin{thm}
-E-systems are the same things as CE-systems.
+Every E-system is a CE-system.
 \end{thm}
@@
 \begin{rmk}
-When $\mathcal{E}$ is a E-system with empty context, then we have the category
+When $\mathbb{E}$ is a E-system with empty context, then we have the category
 structure on the contexts given by $\thom{\Gamma}{\Delta}= T(W_\Gamma(\Delta))$.
@@
 \begin{defn}
-A \emph{natural transformation} of E-homomorphisms from $F:\mathcal{E}\to\mathcal{D}$
-to $G:\mathcal{E}\to\mathcal{D}$ preserving the empty context consists of
+A \emph{natural transformation} of E-homomorphisms from $F:\mathbb{E}\to\mathbb{D}$
+to $G:\mathbb{E}\to\mathbb{D}$ preserving the empty context consists of
 \begin{enumerate}
-\item A t-morphism $\tau0_A\in\thom{F(\Gamma)}{G(\Gamma)}$ for every $\Gamma\in\mathcal{E}$.
+\item A t-morphism $\tau0_A\in\thom{F(\Gamma)}{G(\Gamma)}$ for every $\Gamma\in\mathbb{E}$.
 \item A commuting square square
@@
 \end{equation*}
-commutes for every $A,B\in\mathcal{E}/\Gamma$. Note that 
+commutes for every $A,B\in\mathbb{E}/\Gamma$. Note that 
 \begin{align*}
@@
 \begin{lem}
-For each $A\in\mathcal{E}/\Gamma$, $W_A$ is left adjoint to $\Pi_A$. 
+For each $A\in\mathbb{E}/\Gamma$, $W_A$ is left adjoint to $\Pi_A$. 
 \end{lem}

### 2017-07-18 introduction.tex
starting to restructure the text

@@
 natural numbers, and of which the domain of each operation is declared by 
-equations on the preceeding operations. The concept of essentially algebraic
+equations on the preceding operations. The concept of essentially algebraic
 theories is due to Freyd, who writes that a critical feature of essentially
@@
 \cite{TheBook}, while at the same time it is as economic as possible in the
-demand of requirements of the meta-theory (such as the availability of a type of natural numbers). Instead of explaning
+demand of requirements of the meta-theory (such as the availability of a type of natural numbers). Instead of explaining
 what a dependent type is via an infinite hierarchy of universes $\UU_0$,
@@
 These are asserting that $\Gamma$ is a context; that $\Gamma$ and $\Gamma'$ are
-judgementally equal contexts; that $A$ is a family in context $\Gamma$; that
+judgmentally equal contexts; that $A$ is a family in context $\Gamma$; that
 $A$ and $B$ are judgmentally equal families in context $\Gamma$; that $x$ is
@@
 
-a E-system is therefore an algebra consisting of three sorts: $C$ for contexts,
+An E-system is therefore going to be defined as an algebra consisting of three sorts: $C$ for contexts,
 $F$ for families, and $T$ for terms. There is a map 
@@
 family; and there is a map $\ebd:T\to F$ which determines for each term, of
-which family it is a term. Thus, the backbones of a E-system can be represented
-as
+which family it is a term. Thus, an E-system is a structure on top of the following simpler structure:
 \begin{equation*}
@@
 stratified E-systems. However, we do not arrive at the fact that the functor
-$\mathcal{E}:\mathbf{Bsys}\to\mathbf{Ctx}$ is full without imposing an additional
+$\mathcal{E}:\mathbf{Bsys}\to\mathbf{Esys}$ is full without imposing an additional
 condition on the E-homomorphisms. 
@@
 E-homomorphisms between them is a full subcategory of the category 
-$\mathbf{Ctx}$, we will require that the underlying functor of each E-homomorphism
+$\mathbf{Esys}$, we will require that the underlying functor of each E-homomorphism
 lifts factorizations. It is
@@
 
-Recall that for any morphism $f:X\to Y$ in a category $\cat{C}$ there is a category
+Recall that for any morphism $f:X\to Y$ in a pre-category $\cat{C}$ there is a pre-category
 $\mathbf{fact}(f)$ of factorizations of $f$. Also, any functor $F:\cat{C}\to\cat{D}$
@@
 
-\begin{defn}
+\begin{defn}\label{defn:lift_factorizations}
 A functor $F:\cat{C}\to\cat{D}$ is said to
@@
 $F_{\mathbf{fact}(f)}:\mathbf{fact}(f)\to\mathbf{fact}(F(f))$ is an isomorphism
-of categories. 
+of pre-categories. 
 \end{defn}
@@
 that $F(g)\jdeq g'$ and $F(h)\jdeq h'$. In the version with equivalences,
-the uniqueness is replaced by uniqueness up to isomorphism, and the equalites
+the uniqueness is replaced by uniqueness up to isomorphism, and the equalities
 are replaced by an isomorphism in $\mathbf{fact}(F(f))$.
@@
 of E-systems. The structure of the definition of weakening systems is 
-prototypcial: we first define the notion of pre-weakening systems, which are
+prototypical: we first define the notion of pre-weakening systems, which are
 categories that come with a pre-weakening structure; then we define the notion
@@
 it will be convenient to denote the objects of the $\cat{F}$ with capital
-greek letters $\Gamma$, $\Delta$, \ldots, and the morphisms by capital latin letters
+Greek letters $\Gamma$, $\Delta$, \ldots, and the morphisms by capital Latin letters
 $A$, $B$, $P$, $Q$, \ldots. When $A$ is a morphism into $\Gamma$, we will denote
@@
 \subsection*{Acknowledgements}
-Several people have been indispensible throughout the research process.
+Several people have been indispensable throughout the research process.
 I would like to thank 
